### 2010-10-21 doc/USER_MANUAL/manual_SPECFEM3D.tex
important modifications : modified the tree structure of the software package. source code files are located in src/ and binaries are created in bin/. To compile type './configure' in the main folder and then 'make' or 'make generate_databases' for instance.

%% LyX 1.5.1 created this file.  For more info, see http://www.lyx.org/.
%% Do not edit unless you really know what you are doing.
\documentclass[oneside,english,onecolumn]{book}
\usepackage[T1]{fontenc}
\usepackage[latin1]{inputenc}
\usepackage{geometry}
\geometry{verbose,letterpaper,tmargin=1in,bmargin=1in,lmargin=1in,rmargin=1in}
\setcounter{secnumdepth}{3}
\setcounter{tocdepth}{3}
\usepackage{float}
\usepackage{textcomp}
\usepackage{amsmath}

% figures
\usepackage[pdftex]{graphicx}

% we are running pdflatex, so convert .eps files to .pdf
\usepackage{epstopdf}

\usepackage{amssymb}
\IfFileExists{url.sty}{\usepackage{url}}
                      {\newcommand{\url}{\texttt}}
\usepackage[authoryear]{natbib}

\makeatletter

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% LyX specific LaTeX commands.
\newcommand{\noun}[1]{\textsc{#1}}
%% A simple dot to overcome graphicx limitations
\newcommand{\lyxdot}{.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Textclass specific LaTeX commands.
\newenvironment{lyxcode}
{\begin{list}{}{
\setlength{\rightmargin}{\leftmargin}
\setlength{\listparindent}{0pt}% needed for AMS classes
\raggedright
\setlength{\itemsep}{0pt}
\setlength{\parsep}{0pt}
\normalfont\ttfamily}%
 \item[]}
{\end{list}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% User specified LaTeX commands.
%\renewcommand{\baselinestretch}{1.5}

% figures
\usepackage[dvips]{epsfig}

\usepackage{wrapfig}

% fonts
\usepackage{times}

% hyperlinks to sections and references
\usepackage[pdftex,bookmarks=true,bookmarksnumbered=true,pdfpagemode=None,pdfstartview=FitH,pdfpagelayout=SinglePage,pdfborder={0 0 0}]{hyperref}

\let\myUrl\url
\renewcommand{\url}[1]{(\myUrl{#1})}

% biblio GJI
\bibliographystyle{abbrvnat}

\newcommand{\toall}[1]{\textbf{*** All: #1 ***}}
\newcommand{\tojeroen}[1]{\textbf{*** Jeroen: #1 ***}}
\newcommand{\tobrian}[1]{\textbf{*** Brian: #1 ***}}
\newcommand{\tovala}[1]{\textbf{*** Vala: #1 ***}}
\newcommand{\tovalabrian}[1]{\textbf{*** Vala \& Brian: #1 ***}}
\newcommand{\tovalaqinya}[1]{\textbf{*** Vala \& Qinya: #1 ***}}
\newcommand{\toqinya}[1]{\textbf{*** Qinya: #1 ***}}
\newcommand{\tomin}[1]{\textbf{*** Min: #1 ***}}
\newcommand{\toalessia}[1]{\textbf{*** Alessia: #1 ***}}
\newcommand{\todimitri}[1]{\textbf{*** Dimitri: #1 ***}}

\newcommand{\nexxi}{\mbox{\texttt{NEX\_XI\/}}}
\newcommand{\nexeta}{\mbox{\texttt{NEX\_ETA\/}}}
\newcommand{\nprocxi}{\mbox{\texttt{NPROC\_XI\/}}}
\newcommand{\nproceta}{\mbox{\texttt{NPROC\_ETA\/}}}
\newcommand{\nchunks}{\mbox{\texttt{NCHUNKS\/}}}

\usepackage{babel}
\makeatother

\begin{document}
\thispagestyle{empty}\textbf{}%
\begin{figure}[H]
\noindent \includegraphics[width=0.75\paperwidth]{figures/specfem_3d-cover.pdf}
\end{figure}

\title{\textbf{SPECFEM3D}\\
\textbf{User Manual}}

\author{$\copyright$ California Institute of Technology (USA) and\\
University of Pau / CNRS / INRIA (France)\\
Version 2.0.0}

\maketitle
\newpage{}

\tableofcontents{}

\chapter{Introduction}

The software package SPECFEM3D simulates seismic
wave propagation at the local or regional scale
based upon the spectral-element method (SEM).
The SEM is a continuous Galerkin technique, which can easily be made discontinous \citep{BeMaPa94,KoWoHu02,ChCaVi03,Kop06}; it is then a particular case of the discontinuous Galerkin technique \citep{ReHi73,FaRi99,HuHuRa99,CoKaSh00,GiHeWa02,RiWh03,MoRi05,GrScSc06,BeLaPi06,DuKa06,DeSeWh08,WiStBuGh10,DeSe10}, with optimized efficiency because of its tensorized basis functions.
Effects due to lateral variations in compressional-wave speed, shear-wave
speed, density, a 3D crustal model, topography and bathymetry are
included. For a detailed introduction to the SEM as applied to regional
seismic wave propagation, please consult \citet{KoVi98,KoTr99,ChKoViCaVaFe07,TrKoLi08} and
in particular \citet{KoLiTrSuStSh04}. If you use the 3D southern
California model, please cite \citet{SuSh03} (LA), \citet{lovelyetal06}
(Salton Trough), and \citet{hauksson2000} (southern California).
The Moho map was determined by \citet{zhu&kanamori2000}. The 1D SoCal
model was developed by \citet{DrHe90}. The package can accommodate
full 21-parameter anisotropy (see~\citet{ChTr07}) as well as lateral
variations in attenuation. Adjoint capabilities and finite-frequency
kernel simulations are included \citep{LiTr06,TrKoLi08}.

All SPECFEM3D\_GLOBE software is written in Fortran90 with full portability
in mind, and conforms strictly to the Fortran95 standard. It uses
no obsolete or obsolescent features of Fortran77. The package uses
parallel programming based upon the Message Passing Interface (MPI)
\citep{GrLuSk94,Pac97}.

SPECFEM3D won the Gordon Bell award for best performance at the SuperComputing~2003
conference in Phoenix, Arizona (USA) by running at 5 teraflops (sustained)
on 1944 processors of the Japanese Earth Simulator using 14.6 billion
degrees of freedom stored in 2.5 terabytes of memory; see \cite{KoTsChTr03}
and the Gordon Bell Awards News Release \url{www.sc-conference.org/sc2003/nr_finalaward.html }
for details. It was a finalist again in 2008 for a run at 0.16 petaflops (sustained) on 149,784 processors of the `Jaguar' Cray XT5 system at Oak Ridge National Laboratories (USA) \citep{CaKoLaTiMiLeSnTr08}.

The next release of the code will include support for GPU graphics card acceleration \citep{KoMiEr09,KoErGoMi10,MiKo10}
as well as Convolutional or Auxiliary Differential Equation Perfectly Matched absorbing Layers (C-PML or ADE-PML) \citep{KoMa07,MaKoEz08,MaKoGe08,MaKo09,MaKoGeBr10}.
The next release will use the PT-Scotch and ParMetis parallel libraries for mesh partitioning.

\section{Citation}

If you use SPECFEM3D for your own research, please cite at least one
of the following articles: \cite{TrKoLi08,VaCaSaKoVi99,LeChLiKoHuTr08,LeChKoHuTr09,LeKoHuTr09,KoMiEr09,KoErGoMi10,KoGoErMi10,WiKoScTr04,KoLiTrSuStSh04,ChKoViCaVaFe07,MaKoDi09,KoViCh10,CaKoLaTiMiLeSnTr08,TrKoHjLiZhPeBoMcFrTrHu10,KoRiTr02,KoTr02a,KoTr02b,KoTr99} or \cite{KoVi98}.
If you work on geophysical applications, you may be interested in citing some of these application articles as well, among others: \cite{WiKoScTr04,JiTsKoTr05,KrJiKoTr06a,KrJiKoTr06b,LeChLiKoHuTr08,LeChKoHuTr09,LeKoHuTr09,ChFaKo04,FaChKo04,RiRiKoTrHe02,GoAmTaCaSmSaMaKo09,TrKo00,SaKoTr10}.
The corresponding Bib\TeX{} entries may be found
in file \texttt{USER\_MANUAL/bibliography.bib} or in comments at the
beginning of file \texttt{specfem3D.f90}.

\section{Support}

This material is based upon work supported by the USA National Science
Foundation under Grants No. EAR-0406751 and EAR-0711177, by the French
CNRS, French INRIA Sud-Ouest MAGIQUE-3D, French ANR NUMASIS under
Grant No. ANR-05-CIGC-002, and European FP6 Marie Curie International
Reintegration Grant No. MIRG-CT-2005-017461. 
Older versions of the code were initially developed by Dimitri Komatitsch at Institut de Physique du Globe (France)
and then by Dimitri Komatitsch and Jeroen Tromp at Harvard University (USA).
Any opinions, findings,
and conclusions or recommendations expressed in this material are
those of the authors and do not necessarily reflect the views of the
USA National Science Foundation, CNRS, INRIA, ANR or the European
Marie Curie program.

\chapter{\label{cha:Getting-Started}Getting Started}

The SPECFEM3D software package comes in a gzipped tar ball. In the
directory in which you want to install the package, type

\begin{lyxcode}
{\small tar~-zxvf~SPECFEM3D\_V2.0.0.tar.gz}{\small \par}
\end{lyxcode}
The directory \texttt{\small SPECFEM3D\_V2.0.0} will then contain
the source code. To configure the software for your system, run the
\texttt{configure} shell script. This script will attempt to guess
the appropriate configuration values for your system. However, at
a minimum, it is recommended that you explicitly specify the appropriate
command names for your Fortran90 compiler and MPI package:

\begin{lyxcode}
./configure~FC=ifort~MPIFC=mpif90
\end{lyxcode}

The SPECFEM3D software package relies on the Scotch library to partition the meshes created with CUBIT. Because it is not standard, the path of this library needs to be set with the option \texttt{-{}-with-scotch-lib}. Just as an example:
\begin{lyxcode}
./configure~FC=ifort~MPIFC=mpif90~-{}-with-scotch-lib=/opt/scotch/gcc64/lib
\end{lyxcode}

To compile a serial version of the code for small meshes that fit
on one compute node and can therefore be run serially, run \texttt{configure}
with the \texttt{-{}-without-mpi} option to suppress all calls to
MPI.

A summary of the most important configuration variables follows.

\begin{description}
\item [{\texttt{F90}}] Path to the Fortran90 compiler.
\item [{\texttt{MPIF90}}] Path to MPI Fortran90.
\item [{\texttt{MPI\_FLAGS}}] Some systems require this flag to link to
MPI libraries.
\item [{\texttt{FLAGS\_CHECK}}] Compiler flag for non-critical subroutines.
\item [{\texttt{FLAGS\_NO\_CHECK}}] Compiler flag for creating fast, production-run
code for critical subroutines.
\end{description}
The \texttt{Makefile} contains a number of suggested entries for various
compilers, e.g., Portland, Intel, Absoft, NAG, and Lahey. The software
has run on a wide variety of compute platforms, e.g., various PC clusters
and machines from Sun, SGI, IBM, Compaq, and NEC. Select the compiler
you wish to use on your system and choose the related optimization
flags. Note that the default flags in the \texttt{Makefile} are undoubtedly
not optimal for your system, so we encourage you to experiment with
these flags and to solicit advice from your systems administrator.
Selecting the right compiler and optimization flags can make a tremendous
difference in terms of performance. We welcome feedback on your experience
with various compilers and flags.

Now that you have set the compiler information, you need to select
a number of flags in the \texttt{constants.h} file depending on your
system:

\begin{description}
\item [{\texttt{LOCAL\_PATH\_IS\_ALSO\_GLOBAL}}] Set to \texttt{.false.}
on most cluster applications. For reasons of speed, the (parallel)
mesher typically writes a (parallel) database for the solver on the
local disks of the compute nodes. Some systems have no local disks,
e.g., BlueGene or the Earth Simulator, and other systems have a fast
parallel file system, in which case this flag should be set to \texttt{.true.}.
Note that this flag is not used by the mesher or the solver; it is
only used for some of the post-processing.
\end{description}
The package can run either in single or in double precision. The default
is single precision mode because this requires exactly half as much
memory. Select your preference by selecting the appropriate setting
in the \texttt{constants.h} file:

\begin{description}
\item [{\texttt{CUSTOM\_REAL}}] Set to \texttt{SIZE\_REAL} for single precision
and \texttt{SIZE\_DOUBLE} for double precision.
\end{description}
In the \texttt{precision.h} file:

\begin{description}
\item [{\texttt{CUSTOM\_MPI\_TYPE}}] Set to \texttt{MPI\_REAL} for single
precision and \texttt{MPI\_DOUBLE\_PRECISION} for double precision.
\end{description}
On a new system, it is definitely worth experimenting with single
versus double precision simulations to determine which is faster.
Note that on many current processors (e.g., Intel, AMD, IBM Power),
single precision calculations are often significantly faster; the
difference can typically be 10\% to 25\%. It is therefore often worth
using single precision if you can. We recommend running the same calculation
once in single precision and in double precision on your system and
comparing the seismograms. If they are identical, you should probably
select single precision for your future runs.

When running on an SGI add ``\texttt{setenv TRAP\_FPE OFF}'' to your
.cshrc file \textit{before} compiling in order to turn underflow trapping
off.

Finally, before compiling make sure that the subdirectories \texttt{obj}
and \texttt{OUTPUT\_FILES} exist within the directory with the source
code (\texttt{SPECFEM3D\_V2.0.0}). The \texttt{go\_mesher} script
discussed in Chapter~\ref{cha:Running-the-Mesher-Meshfem3D} automatically
takes care of creating the \texttt{OUTPUT\_FILES} directory.

Note that if you run very large meshes on a relatively small number
of processors, the memory size needed on each processor might become
greater than 2 gigabytes, which is the upper limit for 32-bit addressing;
in this case, on some compilers you may need to add \char`\"{}\texttt{-mcmodel=medium}\char`\"{}
to the compiler options otherwise the compiler will display an error
message.

\chapter{\label{cha:Mesh-Generation}Mesh Generation}

\section{\label{cha:Running-the-Mesher-CUBIT}Meshing with \texttt{CUBIT}}
\subsection{Exporting the Mesh}
\subsection{Partitioning the Mesh with \texttt{xdecompose\_mesh\_SCOTCH}}

The SPECFEM3D software package performs large scale simulations in a parallel 'Single Process Multiple Data' way. 
The spectral element mesh created with CUBIT needs to be distributed on the processors. This partitioning is executed once and for all prior to
the execution of the solver so it is referred to as a static mapping.

An efficient partitioning is important because it leverages the overall running time of the application.
It amounts to balance the number of elements in each slice while minimizing the communication coasts
resulting from the placement of adjacent elements to different processors. \texttt{decompose\_mesh\_SCOTCH} depends on the Scotch library
which provides efficent static mapping, graph and mesh partitioning routines. Scotch is a free software package downloadable from the web page \texttt{https://gforge.inria.fr/projects/scotch/}.
\\

Prior to compiling \texttt{decompose\_mesh\_SCOTCH}, make sure you have correctly specified the path 
of the Scotch library with the option \texttt{-{}-with-scotch-lib} of the \texttt{configure} script. Then you are ready to compile, 
in the directory \texttt{decompose\_mesh\_SCOTCH} type '\texttt{make decompose\_mesh\_SCOTCH}'.
If all paths and flags have been set correctly, the executable \texttt{xdecompose\_mesh\_SCOTCH} should be produced.

The partitioning is done in serial, 
the synopsis is :
\begin{lyxcode}
./decompose\_mesh\_SCOTCH  nparts  input\_directory output\_directory
\end{lyxcode}

All the files generated by the Python scripts must be placed in the \texttt{input\_directory} folder before running the program.

\section{\label{cha:Running-the-Mesher-Meshfem3D}Meshing with \texttt{xmeshfem3D}}

\noindent \begin{flushleft}
%
\begin{figure}[t]
\noindent \begin{centering}
\includegraphics[scale=0.5]{figures/socal_map_mpi}
\par\end{centering}

\caption{\label{fig:For-parallel-computing}For parallel computing purposes,
the model block is subdivided in $\nprocxi\times\nproceta$ slices
of elements. In this example we use $5^{2}=25$ processors. }

\end{figure}

\par\end{flushleft}

You are now ready to compile the internal mesher. This is an alternative to CUBIT 
for the mesh generation of relatively simple geological models. The mesher is no longer 
dedicated to Southern California and more flexiblity is provided in this version of the package.

In the directory \texttt{meshfem3D} type `\texttt{make meshfem3D}'. If all paths and flags
have been set correctly, the mesher should now compile and produce
the executable \texttt{xmeshfem3D}.

Input for the mesh generation program is provided through the parameter
file \texttt{Par\_file}, which resides in the subdirectory \texttt{meshfem3D/DATA}.
Before running the mesher, a number of parameters need to be set in
the \texttt{Par\_file}. This requires a basic understanding of how
the SEM is implemented, and we encourage you to read \citet{KoVi98,KoTr99}
and \citet{KoLiTrSuStSh04}.

The mesher and the solver use UTM coordinates internally, therefore
you need to define the zone number for the UTM projection (e.g., zone
11 for Los Angeles). Use decimal values for latitude and longitude
(no minutes/seconds). These values are approximate; the mesher will
round them off to define a square mesh in UTM coordinates. When running
benchmarks on rectangular models, turn the UTM projection off by using
the flag \texttt{\small SUPPRESS\_UTM\_PROJECTION}, in which case
all `longitude' parameters simply refer to the $x$~axis, and all
`latitude' parameters simply refer to the $y$~axis. To run the mesher
for a global simulation, the following parameters need to be set in
the \texttt{Par\_file}:

\begin{description}
\item [{\texttt{LATITUDE\_MIN}}] Minimum latitude in the block (negative
for South).
\item [{\texttt{LATITUDE\_MAX}}] Maximum latitude in the block.
\item [{\texttt{LONGITUDE\_MIN}}] Minimum longitude in the block (negative
for West).
\item [{\texttt{LONGITUDE\_MAX}}] Maximum longitude in the block.
\item [{\texttt{DEPTH\_BLOCK\_KM}}] Depth of bottom of mesh in kilometers.
\item [{\texttt{\noun{UTM\_PROJECTION\_ZONE}}}] UTM projection zone in
which your model resides, only valid when \texttt{SUPPRESS\_UTM\_}~\\
\texttt{PROJECTION} is \texttt{.false.}.
\item [{\texttt{SUPPRESS\_UTM\_PROJECTION}}] set to be \texttt{.false.}
when your model range is specified in the geographical coordinates,
and needs to be \texttt{.true.} when your model is specified in a
cartesian coordinates. \noun{UTM projection zone in which your simulation
region resides.}
\item [{\texttt{INTERFACES\_FILE }}] File in which contains the description of the topography and 
of the interfaces between the different layers of the model, if any. 
The number of spectral elements in the vertical direction within each layer is also defined in this file.

\item [{$\nexxi$}] The number of spectral elements along one side of the
block. This number \textit{must} be 8~$\times$~a multiple of $\nprocxi$
defined below. Based upon benchmarks against semi-analytical discrete
wavenumber synthetic seismograms \citep{KoLiTrSuStSh04}, determined
that a $\nexxi=288$ run is accurate to a shortest period of roughly
2~s. Therefore, since accuracy is determined by the number of grid
points per shortest wavelength, for any particular value of $\nexxi$
the simulation will be accurate to a shortest period determined by
\begin{equation}
\mbox{shortest period (s)}=(288/\nexxi)\times2.\label{eq:shortest_period}\end{equation}
The number of grid points in each orthogonal direction of the reference
element, i.e., the number of Gauss-Lobatto-Legendre points, is determined
by \texttt{NGLLX} in the \texttt{constants.h} file. We generally use
$\mbox{\texttt{NGLLX\/}}=5$, for a total of $5^{3}=125$ points per
elements. We suggest not to change this value.
\item [{$\nexeta$}] The number of spectral elements along the other side
of the block. This number \textit{must} be 8~$\times$~a multiple
of $\nproceta$ defined below.
\item [{$\nprocxi$}] The number of processors or slices along one side
of the block (see Figure~\ref{fig:For-parallel-computing}); we must
have $\nexxi=8\times c\times\nprocxi$, where $c\ge1$ is a positive
integer.
\item [{$\nproceta$}] The number of processors or slices along the other
side of the block; we must have $\nexeta=8\times c\times\nproceta$,
where $c\ge1$ is a positive integer.
\item [{\texttt{USE\_REGULAR\_MESH}}] set to be \texttt{.true.} if you want a perfectly regular
mesh or \texttt{.false.} if you want to add doubling horizontal layers to coarsen the mesh. In this case, you
also need to provide additional information by setting up the next three parameters.  
\item [{\texttt{NDOUBLINGS}}] The number of horizontal doubling layers. Must be set to \texttt{1} or \texttt{2} 
if \texttt{USE\_REGULAR\_MESH} is set to \texttt{.true.}.
\item [{\texttt{NZ\_DOUBLING\_1}}] The position of the first doubling layer (only interpreted
if \texttt{USE\_REGULAR\_MESH} is set to \texttt{.true.}).
\item [{\texttt{NZ\_DOUBLING\_2}}] The position of the second doubling layer (only interpreted
if \texttt{USE\_REGULAR\_MESH} is set to \texttt{.true.} and if \texttt{NDOUBLINGS} is set to \texttt{2}). 
\item [{\texttt{CREATE\_ABAQUS\_FILES}}] Set this flag to \texttt{.true.} to save Abaqus FEA \url{www.simulia.com}
mesh files for subsequent viewing. Turning the flag on generates files in the \texttt{LOCAL\_PATH} directory. 
See Section~\ref{sec:Mesh-graphics} for a discussion of mesh viewing features.
\item [{\texttt{CREATE\_DX\_FILES}}] Set this flag to \texttt{.true.} to save OpenDX \url{www.opendx.org}
mesh files for subsequent viewing.
\item [{\texttt{LOCAL\_PATH}}] Directory in which the databases generated
by the mesher will be written. Generally one uses a directory on the
local disk of the compute nodes, although on some machines these databases
are written on a parallel (global) file system (see also the earlier
discussion of the \texttt{LOCAL\_PATH\_IS\_ALSO\_GLOBAL} flag in Chapter~\ref{cha:Getting-Started}).
The mesher generates the necessary databases in parallel, one set
for each of the $\nprocxi\times\nproceta$ slices that constitutes
the mesh (see Figure~\ref{fig:For-parallel-computing}). After the
mesher finishes, you can log in to one of the compute nodes and view
the contents of the \texttt{LOCAL\_PATH} directory to see the (many)
files generated by the mesher.
\item [{\texttt{NMATERIALS}}] The number of different materials in your model. In the following lines, each material needs to be defined as :
\begin{lyxcode}
\texttt{material\_ID}  \texttt{rho}  \texttt{vp}  \texttt{vs}  \texttt{Q\_flag}  \texttt{anisotropy\_flag} \texttt{domain\_ID}
\end{lyxcode}
where
\begin{itemize}
     \item \texttt{Q\_flag}           : 0=no attenuation / standard \texttt{Q} attenutation value 
     \item \texttt{anisotropy\_flag}  : 0=no anisotropy / 1,2,.. check with implementation in \texttt{aniso\_model.f90}
     \item \texttt{domain\_id}        : 1=acoustic / 2=elastic / 3=poroelastic
\end{itemize}
\item [{\texttt{NMATERIALS}}] The number of regions in the mesh. 
In the following lines, because the mesh is regular or 'almost regular', each region is defined as :
\begin{lyxcode}
  \texttt{XI\_begin}  \texttt{XI\_end}  \texttt{ETA\_begin}  \texttt{ETA\_end}  \texttt{material\_ID}
\end{lyxcode}
\end{description}

The topography of the model is defined as a set of elevation
 values on a regular 2D grid. It is also posible to
define interfaces between the layers of the model in the same way.

The file defined in \texttt{INTERFACES\_FILE} contains the settings of the topography grid
and of the interfaces grids. The number 
of interfaces, including the topography, needs to be set at the first line. Then, from the bottom 
to the top of the model, you need to define the grids with several parameters :
 number of points along $x$ and $y$, minimal $x$ and $y$ coordinates, spacing between points and the file 
in which the elevation values are stored. At the end of this file, you simply need to set the
number of spectral elements in the vertical direction for each layer.\\

Finally, depending on your system, you might need to provide a file that tells MPI what compute nodes
to use for the simulations. The file must have a number of entries
(one entry per line) at least equal to the number of processors needed
for the run. A sample file is provided in the file \texttt{mymachines}.
This file is not used by the mesher or solver, but is required by
the \texttt{go\_mesher} and \texttt{go\_solver} default job submission
scripts. See Chapter~\ref{cha:Scheduler} for information about running
the code on a system with a scheduler, e.g., LSF.

Now that you have set the appropriate parameters in the \texttt{Par\_file}
and have compiled the mesher, you are ready to launch it! This is
most easily accomplished based upon the \texttt{go\_mesher} script.
When you run on a PC cluster, the script assumes that the nodes are
named n001, n002, etc. If this is not the case, change the \texttt{tr
-d `n'} line in the script. You may also need to edit the last command
at the end of the script that invokes the \texttt{mpirun} command.
See Chapter~\ref{cha:Scheduler} for information about running the
code on a system with a scheduler, e.g., LSF.

Mesher output is provided in the \texttt{OUTPUT\_FILES} directory
in \texttt{output\_mesher.txt}; this file provides lots of details
about the mesh that was generated. Alternatively, output can be directed
to the screen instead by uncommenting a line in \texttt{constants.h}:

\begin{lyxcode}
!~uncomment~this~to~write~messages~to~the~screen~

!~integer,~parameter~::~IMAIN~=~ISTANDARD\_OUTPUT
\end{lyxcode}

%% \section{Checking the MPI Buffers (Optional)}

%% The mesher writes MPI communication tables in the \texttt{OUTPUT\_FILES}
%% subdirectory in the files \texttt{addressing.txt}, \texttt{list\_messages\_corners.txt}
%% and \texttt{list\_messages\_faces.txt}, and MPI communication buffers
%% to the local disks. Use the four serial codes

%% \begin{lyxcode}
%% check\_buffers\_2D.f90

%% check\_buffers\_1D.f90
%% \end{lyxcode}
%% to check that all the MPI buffers created by the mesher have been
%% generated correctly. For example, typing `\texttt{make check\_buffers\_2D}'
%% and then `\texttt{xcheck\_buffers\_2D}' checks the communication buffers
%% between faces common to the mesh slices.

%% \begin{quote}
%% Please note that running these codes is optional because no information
%% needed by the solver is generated.
%% \end{quote}

%%\section{\label{sec:quality}Checking the Mesh Quality (Optional)}

The quality of the mesh may be inspected more precisely based upon the serial code
\texttt{check\_mesh\_quality\_CUBIT\_Abaqus.f90}. In the directory \texttt{check\_mesh\_quality\_CUBIT\_Abaqus}, type

\begin{lyxcode}
make~check\_mesh\_quality\_CUBIT\_Abaqus
\end{lyxcode}
and then use

\begin{lyxcode}
xcheck\_mesh\_quality\_CUBIT\_Abaqus
\end{lyxcode}
to generate an AVS output file (\texttt{\small AVS\_meshquality.inp}
in AVS UCD format) or OpenDX output file (\texttt{\small DX\_meshquality.}~\\
\texttt{\small dx}) that can be used to investigate mesh quality,
e.g, skewness of elements and a Gnuplot histogram (\texttt{\small mesh\_quality\_}~\\
\texttt{\small histogram.txt}) that can be plotted with gnuplot (type
`\texttt{\small gnuplot plot\_mesh\_quality\_histogram.gnu}'). The
histogram is also printed to the screen. If you want to start designing
your own meshes, this tool is useful for viewing your creations. You
are striving for meshes with elements with `cube-like' dimensions,
e.g., the mesh should contain no very elongated or skewed elements.

Running this code is optional because no information needed by the
solver is generated.

\chapter{\label{cha:Creating-Distributed-Databases}Creating the Distributed Databases}

Either you previously used  \texttt{xmeshfem3D} or \texttt{xdecompose\_mesh\_SCOTCH}, the next step in the workflow is to compile \texttt{xgenerate\_databases}. This program is going to create all the missing information needed by the SEM solver. In the main directory type '\texttt{make generate\_databases}'. Input for the program is provided through the main parameter file \texttt{Par\_file}, which resides in the subdirectory \texttt{DATA}. Before running \texttt{xgenerate\_databases}, a number of parameters need to be set in the \texttt{Par\_file}:

\begin{description}
\item [{\texttt{SIMULATION\_TYPE}}] is set to 1 for forward simulations,
2 for adjoint simulations (see Section \ref{sec:Adjoint-simulation-finite})
and 3 for kernel simulations (see Section \ref{sec:Finite-Frequency-Kernels}).
\item [{\texttt{SAVE\_FORWARD}}] is only set to \texttt{.true.} for a forward
simulation with the last frame of the simulation saved, as part of
the finite-frequency kernel calculations (see Section \ref{sec:Finite-Frequency-Kernels}).
For a regular forward simulation, leave \texttt{SIMULATION\_TYPE}
and \texttt{SAVE\_FORWARD} at their default values.
\item [{\texttt{\noun{UTM\_PROJECTION\_ZONE}}}] UTM projection zone in
which your model resides, only valid when \texttt{SUPPRESS\_UTM\_}~\\
\texttt{PROJECTION} is \texttt{.false.}.
\item [{\texttt{SUPPRESS\_UTM\_PROJECTION}}] set to be \texttt{.false.}
when your model range is specified in the geographical coordinates,
and needs to be \texttt{.true.} when your model is specified in a
cartesian coordinates. \noun{UTM projection zone in which your simulation
region resides.}
\item [{\texttt{NPROC}}] The number of MPI processors, each one is assigned one slice of the whole mesh. 
\item [{\texttt{NSTEP}}] The number of time steps of the simulation. This controls the
length of the numerical simulation, i.e., twice the number of time steps
requires twice as much CPU time. This feature is not used at the time
of meshing but is required for the solver, i.e., you may change this
parameter after running  \texttt{xgenerate\_databases}.
\item [{\texttt{DT}}] The length of each time step in seconds. 
If you used xmeshfem3D to generate your mesh, you should set the value suggested in \texttt{meshfem3D/OUTPUT\_FILES/output\_mesher.txt}.

\item [{\texttt{OCEANS}}] Set to \texttt{.true.} if the effect of the oceans
on seismic wave propagation should be incorporated based upon the
approximate treatment discussed in \citet{KoTr02b}. This feature
is inexpensive from a numerical perspective, both in terms of memory
requirements and CPU time. This approximation is accurate at periods
of roughly 20~s and longer. At shorter periods the effect of water
phases/reverberations is not taken into account, even when the flag
is on.
\item [{\texttt{TOPOGRAPHY}}] Set to \texttt{.true.} if topography and
bathymetry should be incorporated based upon model ETOPO5 \citep{Etopo5}.
This feature adds no cost to the simulation.
\item [{\texttt{ATTENUATION}}] Set to \texttt{.true.} if attenuation should
be incorporated. Turning this feature on increases the memory requirements
significantly (roughly by a factor of~1.5), and is numerically fairly
expensive. See \citet{KoTr99,KoTr02a} for a discussion on the implementation
of attenuation based upon standard linear solids.
\item [{\texttt{USE\_OLSEN\_ATTENUATION}}] Set to \texttt{.true.} if you
want to use the attenuation model that scaled from the velocity model
using Olsen's empirical relation (reference).
\item [{\texttt{ABSORBING\_CONDITIONS}}] Set to \texttt{.true.} to turn
on Clayton-Enquist absorbing boundary conditions (see \citet{KoTr99}).
\item [{\texttt{RECORD\_LENGTH\_IN\_MINUTES}}] Choose the desired record
length of the synthetic seismograms (in minutes). This controls the
length of the numerical simulation, i.e., twice the record length
requires twice as much CPU time. This feature is not used at the time
of meshing but is required for the solver, i.e., you may change this
parameter after running xgenerate\_databases.
\item [{\texttt{MOVIE\_SURFACE}}] Set to \texttt{.false.}, unless you want
to create a movie of seismic wave propagation on the Earth's surface.
Turning this option on generates large output files. See Section \ref{sec:Movies}
for a discussion on the generation of movies. This feature is only relevant for the solver.
\item [{\texttt{MOVIE\_VOLUME}}] Set to \texttt{.false.}, unless you want
to create a movie of seismic wave propagation in the Earth's interior.
Turning this option on generates huge output files. See Section \ref{sec:Movies}
for a discussion on the generation of movies. This feature is only relevant for the solver.
\item [{\texttt{NTSTEP\_BETWEEN\_FRAMES}}] Determines the number of timesteps
between movie frames. Typically you want to save a snapshot every
100 timesteps. The smaller you make this number the more output will
be generated! See Section \ref{sec:Movies} for a discussion on the
generation of movies. This feature is only relevant for the solver.
\item [{\texttt{CREATE\_SHAKEMAP}}] Set this flag to \texttt{.true.} to
create a ShakeMap\textregistered, i.e., a peak ground velocity map
of the maximum absolute value of the two horizontal components of the velocity vector.
\item [{\texttt{SAVE\_DISPLACEMENT}}] Set this flag to \texttt{.true.}
if you want to save the displacement instead of velocity for the movie
frames.
\item [{\texttt{USE\_HIGHRES\_FOR\_MOVIES}}] Set this flag to \texttt{.true.}
if you want to save the values at all the NGLL grid points for the
movie frames.
\item [{\texttt{SAVE\_MESH\_FILES}}] Set this flag to \texttt{.true.} to
save AVS \url{www.avs.com}, OpenDX \url{www.opendx.org}, or ParaView \url{www.paraview.org}
mesh files for subsequent viewing. Turning the flag on generates large
(distributed) files in the \texttt{LOCAL\_PATH} directory. See Section~\ref{sec:Mesh-graphics}
for a discussion of mesh viewing features.
\item [{\texttt{LOCAL\_PATH}}] Directory in which the databases and the synthetics 
will be written. Generally one uses a directory on the
local disk of the compute nodes, although on some machines these databases
are written on a parallel (global) file system (see also the earlier
discussion of the \texttt{LOCAL\_PATH\_IS\_ALSO\_GLOBAL} flag in Chapter~\ref{cha:Getting-Started}).
xgenerate\_databases generates the necessary databases in parallel, one set
for each of the $\nprocxi\times\nproceta$ slices that constitutes
the mesh (see Figure~\ref{fig:For-parallel-computing}). After the
mesher finishes, you can log in to one of the compute nodes and view
the contents of the \texttt{LOCAL\_PATH} directory to see the (many)
files generated by xgenerate\_databases.
\item [{\texttt{NTSTEP\_BETWEEN\_OUTPUT\_INFO}}] This parameter specifies
the interval at which basic information about a run is written to
the file system (\texttt{timestamp{*}} files in the \texttt{OUTPUT\_FILES}
directory). If you have access to a fast machine, set \texttt{NTSTEP\_BETWEEN\_OUTPUT\_INFO}
to a relatively high value (e.g., at least 100, or even 1000 or more)
to avoid writing output text files too often. This feature is not
used at the time of meshing. One can set this parameter to a larger
value than the number of time steps to avoid writing output during
the run.
\item [{\texttt{NTSTEP\_BETWEEN\_OUTPUT\_SEISMOS}}] This parameter specifies
the interval at which synthetic seismograms are written in the \texttt{LOCAL\_PATH}
directory. If a run crashes, you may still find usable (but shorter
than requested) seismograms in this directory. On a fast machine set
\texttt{NTSTEP\_BETWEEN\_OUTPUT\_SEISMOS} to a relatively high value
to avoid writing to the seismograms too often. This feature is only relevant for the solver.
\item [{\texttt{PRINT\_SOURCE\_TIME\_FUNCTION}}] Turn this flag on to print
information about the source time function in the file \texttt{OUTPUT\_FILES/plot\_source\_time\_function.txt}.
This feature is only relevant for the solver.
\end{description}

%% Another file generated by \texttt{xgenerate\_databases} is the header file \texttt{OUTPUT\_FILES/values\_from\_mesher.h}.
%% This file specifies a number of constants and flags needed by the
%% solver. These values are passed statically to the solver for reasons
%% of speed. Some useful statistics about the mesh are also provided
%% in this file.

%% For a given model, set of nodes, and set of parameters in \texttt{Par\_file},
%% one only needs to run the mesher once and for all, even if one wants
%% to run several simulations with different sources and/or receivers
%% (the source and receiver information is used in the solver only).

\chapter{\label{cha:Running-the-Solver}Running the Solver \texttt{xspecfem3D}}

Now that you have successfully generated the databases, you are ready to compile
the solver. For reasons of speed, the solver uses static memory allocation.
Therefore it needs to be recompiled (type `\texttt{make clean}' and
`\texttt{make specfem3D}') every time one reruns \texttt{xgenerate\_databases}. To compile
the solver one needs a file created by \texttt{xgenerate\_databases} in the directory
\texttt{OUTPUT\_FILES} called \texttt{values\_from\_mesher.h}, which
contains parameters describing the static size of the arrays as well
as the setting of certain flags.

The solver needs three input files in the \texttt{DATA} directory
to run: the \texttt{Par\_file} which was discussed in detail in Chapters~\ref{cha:Running-the-Mesher-CUBIT} and~\ref{cha:Running-the-Mesher-Meshfem3D},
the earthquake source parameter file \texttt{CMTSOLUTION}, and the
stations file \texttt{STATIONS}. Most parameters in the \texttt{Par\_file}
should be set prior to running the mesher. Only the following parameters
may be changed after running \texttt{xgenerate\_databases}:

\begin{itemize}
\item the simulation type control parameters: \texttt{SIMULATION\_TYPE}
and \texttt{SAVE\_FORWARD}
\item the time step parameters \texttt{NSTEP} and \texttt{DT}
\item the movie control parameters \texttt{MOVIE\_SURFACE}, \texttt{MOVIE\_VOLUME},
and \texttt{NTSTEPS\_BETWEEN\_FRAMES}
\item the ShakeMap\textregistered option \texttt{CREATE\_SHAKEMAP}
\item the output information parameters \texttt{NTSTEP\_BETWEEN\_OUTPUT\_INFO}
and \texttt{NTSTEP\_BETWEEN\_OUTPUT\_}~\\
\texttt{SEISMOS}
\item the \texttt{PRINT\_SOURCE\_TIME\_FUNCTION} flags
\end{itemize}
Any other change to the \texttt{Par\_file} implies rerunning both
the mesher and the solver.

For any particular earthquake, the \texttt{CMTSOLUTION} file that
represents the point source may be obtained directly from the Harvard Centroid-Moment Tensor (CMT) web page \url{www.seismology.harvard.edu}.
It looks like this:

\begin{lyxcode}
{\small }%
\begin{figure}[H]
\noindent \begin{centering}
{\small \includegraphics[width=1\textwidth]{figures/Hollywood_CMT} }
\par\end{centering}{\small \par}

\caption{\label{fig:CMTSOLUTION-file}\texttt{CMTSOLUTION} file based on the
format from the Harvard CMT catalog. \textbf{M} is the moment tensor,
$M_{0}${\small{} }is the seismic moment, and $M_{w}$ is the moment
magnitude.}

\end{figure}
{\small \par}
\end{lyxcode}
The \texttt{CMTSOLUTION} should be edited in the following way:

\begin{itemize}
\item Set the \texttt{time shift} parameter equal to $0.0$ (the solver
will not run otherwise.) The time shift parameter would simply apply
an overall time shift to the synthetics, something that can be done
in the post-processing (see Section \ref{sec:Process-data-and-syn}).
\item For point-source simulations (see finite sources, page \pageref{To-simulate-a})
we recommend setting the source half-duration parameter \texttt{half
duration} equal to zero, which corresponds to simulating a step source-time
function, i.e., a moment-rate function that is a delta function. If
\texttt{half duration} is not set to zero, the code will use a Gaussian
(i.e., a signal with a shape similar to a `smoothed triangle', as
explained in \citet{KoTr02a} and shown in Fig~\ref{fig:gauss.vs.triangle})
source-time function with half-width \texttt{half duration}. We prefer
to run the solver with \texttt{half duration} set to zero and convolve
the resulting synthetic seismograms in post-processing after the run,
because this way it is easy to use a variety of source-time functions
(see Section \ref{sec:Process-data-and-syn}). \citet{KoTr02a} determined
that the noise generated in the simulation by using a step source
time function may be safely filtered out afterward based upon a convolution
with the desired source time function and/or low-pass filtering. Use
the serial code \texttt{convolve\_source\_timefunction.f90} and the
script \texttt{convolve\_source\_timefunction.csh} for this purpose,
or alternatively use signal-processing software packages such as SAC \url{www.llnl.gov/sac}.
Type

\begin{lyxcode}
make~convolve\_source\_timefunction
\end{lyxcode}
to compile the code and then set the parameter \texttt{hdur} in \texttt{convolve\_source\_timefunction.csh}
to the desired half-duration.

\item The zero time of the simulation corresponds to the center of the triangle/Gaussian,
or the centroid time of the earthquake. The start time of the simulation
is $t=-1.5*\texttt{half duration}$ (the 1.5 is to make sure the moment
rate function is very close to zero when starting the simulation).
To convert to absolute time $t_{\mathrm{abs}}$, set

\begin{lyxcode}
$t_{\mathrm{abs}}=t_{\mathrm{pde}}+\texttt{time shift}+t_{\mathrm{synthetic}}$
\end{lyxcode}
where $t_{\mathrm{pde}}$ is the time given in the first line of the
\texttt{CMTSOLUTION}, \texttt{time shift} is the corresponding value
from the original \texttt{CMTSOLUTION} file and $t_{\mathrm{synthetic}}$
is the time in the first column of the output seismogram.

\end{itemize}
%
\begin{figure}
\noindent \begin{centering}
\includegraphics[width=3in]{figures/gauss_vs_triangle_mod.eps}
\par\end{centering}

\caption{Comparison of the shape of a triangle and the Gaussian function actually
used.}

\label{fig:gauss.vs.triangle}
\end{figure}

Centroid latitude and longitude should be provided in geographical
coordinates. The code converts these coordinates to geocentric coordinates~\citep{DaTr98}.
Of course you may provide your own source representations by designing
your own \texttt{CMTSOLUTION} file. Just make sure that the resulting
file adheres to the Harvard CMT conventions (see Appendix~\ref{cha:Coordinates}).
Note that the first line in the \texttt{CMTSOLUTION} file is the Preliminary Determination of Earthquakes (PDE) solution performed by the USGS NEIC, which is used as a seed for the Harvard CMT inversion. The PDE solution is based upon P waves and often gives the hypocenter of the earthquake, i.e., the rupture initiation point, whereas the CMT solution gives the `centroid location', which is the location with dominant moment release. The PDE solution is not used by our software package but must be present anyway in the first line of the file.

\label{To-simulate-a}To simulate a kinematic rupture, i.e., a finite-source
event, represented in terms of $N_{\mathrm{sources}}$ point sources,
provide a \texttt{CMTSOLUTION} file that has $N_{\mathrm{sources}}$
entries, one for each subevent (i.e., concatenate $N_{\mathrm{sources}}$
\texttt{CMTSOLUTION} files to a single \texttt{CMTSOLUTION} file).
At least one entry (not necessarily the first) must have a zero \texttt{time
shift}, and all the other entries must have non-negative \texttt{time
shift}. Each subevent can have its own half duration, latitude, longitude,
depth, and moment tensor (effectively, the local moment-density tensor).

Note that the zero in the synthetics does NOT represent the hypocentral
time or centroid time in general, but the timing of the \textit{center}
of the source triangle with zero \texttt{time shift} (Fig~\ref{fig:source_timing}).

Although it is convenient to think of each source as a triangle, in
the simulation they are actually Gaussians (as they have better frequency
characteristics). The relationship between the triangle and the gaussian
used is shown in Fig~\ref{fig:gauss.vs.triangle}. For finite fault
simulations it is usually not advisable to use a zero half duration
and convolve afterwards, since the half duration is generally fixed
by the finite fault model.

{\small }%
\begin{figure}[H]
\noindent \begin{centering}
{\small \includegraphics[width=5in]{figures/source_timing.eps} }
\par\end{centering}{\small \par}

\caption{Example of timing for three sources. The center of the first source
triangle is defined to be time zero. Note that this is NOT in general
the hypocentral time, or the start time of the source (marked as tstart).
The parameter \texttt{time shift} in the \texttt{CMTSOLUTION} file
would be t1(=0), t2, t3 in this case, and the parameter \texttt{half
duration} would be hdur1, hdur2, hdur3 for the sources 1, 2, 3 respectively.}

{\small \label{fig:source_timing} }
\end{figure}
{\small \par}

The solver can calculate seismograms at any number of stations for
basically the same numerical cost, so the user is encouraged to include
as many stations as conceivably useful in the \texttt{STATIONS} file,
which looks like this:

{\small }%
\begin{figure}[H]
\noindent \begin{centering}
{\small \includegraphics{figures/STATIONS_basin_explained} }
\par\end{centering}{\small \par}

\caption{Sample \texttt{STATIONS} file. Station latitude and longitude should
be provided in geographical coordinates. The width of the station
label should be no more than 32 characters (see \texttt{MAX\_LENGTH\_STATION\_NAME}
in the \texttt{constants.h} file), and the network label should be
no more than 8 characters (see \texttt{MAX\_LENGTH\_NETWORK\_NAME}
in the \texttt{constants.h} file).}

\end{figure}
{\small \par}

Each line represents one station in the following format:

\begin{lyxcode}
{\small Station~Network~Latitude~(degrees)~Longitude~(degrees)~Elevation~(m)~burial~(m)~}{\small \par}
\end{lyxcode}
The mesher filters the list of stations in file \texttt{DATA/STATIONS}
to exclude stations that are not located within the region given in
the \texttt{Par\_file} (between \texttt{LATITUDE\_MIN} and \texttt{LATITUDE\_MAX}
and between \texttt{LONGITUDE\_MIN} and \texttt{LONGITUDE\_MAX}).
The filtered file is called \texttt{DATA/STATIONS\_FILTERED}.

Solver output is provided in the \texttt{OUTPUT\_FILES} directory
in the \texttt{output\_solver.txt} file. Output can be directed to
the screen instead by uncommenting a line in \texttt{constants.h}:

\begin{lyxcode}
!~uncomment~this~to~write~messages~to~the~screen~

!~integer,~parameter~::~IMAIN~=~ISTANDARD\_OUTPUT~
\end{lyxcode}
On PC clusters the seismogram files are generally written to the local
disks (the path \texttt{LOCAL\_PATH} in the \texttt{Par\_file}) and
need to be gathered at the end of the simulation.

While the solver is running, its progress may be tracked by monitoring
the `\texttt{\small timestamp{*}}' files in the \texttt{\small OUTPUT\_FILES}
directory. These tiny files look something like this:

\begin{lyxcode}
Time~step~\#~~~~~~~~~~10000~

Time:~~~~~108.4890~~~~~~seconds~

Elapsed~time~in~seconds~=~~~~1153.28696703911~

Elapsed~time~in~hh:mm:ss~=~~~~~0~h~19~m~13~s~

Mean~elapsed~time~per~time~step~in~seconds~=~~~~~0.115328696703911~

Max~norm~displacement~vector~U~in~all~slices~(m)~=~~~~1.0789589E-02~
\end{lyxcode}
The \texttt{\small timestamp{*}} files provide the \texttt{Mean elapsed
time per time step in seconds}, which may be used to assess performance
on various machines (assuming you are the only user on a node), as
well as the \texttt{Max norm displacement vector U in all slices~(m)}.
If something is wrong with the model, the mesh, or the source, you
will see the code become unstable through exponentially growing values
of the displacement and fluid potential with time, and ultimately
the run will be terminated by the program. You can control the rate
at which the timestamp files are written based upon the parameter
\texttt{NTSTEP\_BETWEEN\_OUTPUT\_INFO} in the \texttt{Par\_file}.

Having set the \texttt{Par\_file} parameters, and having provided
the \texttt{CMTSOLUTION} and \texttt{STATIONS} files, you are now
ready to launch the solver! This is most easily accomplished based
upon the \texttt{go\_solver} script (See Chapter \ref{cha:Scheduler}
for information about running through a scheduler, e.g., LSF). You
may need to edit the last command at the end of the script that invokes
the \texttt{mpirun} command. The \texttt{runall} script compiles and
runs both mesher and solver in sequence. This is a safe approach that
ensures using the correct combination of mesher output and solver
input.

It is important to realize that the CPU and memory requirements of
the solver are closely tied to choices about attenuation (\texttt{ATTENUATION})
and the nature of the model (i.e., isotropic models are cheaper than
anisotropic models). We encourage you to run a variety of simulations
with various flags turned on or off to develop a sense for what is
involved.

For the same model, one can rerun the solver for different events
by simply changing the \texttt{CMTSOLUTION} file, or for different
stations by changing the \texttt{STATIONS} file. There is no need
to rerun the mesher. Of course it is best to include as many stations
as possible, since this does not add to the cost of the simulation.

\chapter{\label{cha:Adjoint-Simulations}Adjoint Simulations}

Adjoint simulations are generally performed for two distinct applications.
First, they can be used for earthquake source inversions, especially
earthquakes with large ruptures such as the Lander's earthquake \citep{WaHe94}.
Second, they can be used to generate finite-frequency sensitivity
kernels that are a critical part of tomographic inversions based upon
3D reference models \citep{TrTaLi05,LiTr06,TrKoLi08,LiTr08}. In either
case, source parameter or velocity structure updates are sought to
minimize a specific misfit function (e.g., waveform or traveltime
differences), and the adjoint simulation provides a means of computing
the gradient of the misfit function and further reducing it in successive
iterations. Applications and procedures pertaining to source studies
and finite-frequency kernels are discussed in Sections~\ref{sec:Adjoint-simulation-sources}
and \ref{sec:Adjoint-simulation-finite}, respectively. The two related
parameters in the \texttt{Par\_file} are \texttt{SIMULATION\_TYPE}
(1 or 2) and the \texttt{SAVE\_FORWARD} (boolean).

\section{\label{sec:Adjoint-simulation-sources}Adjoint Simulations for Sources}

In the case where a specific misfit function is minimized to invert
for the earthquake source parameters, the gradient of the misfit function
with respect to these source parameters can be computed by placing
time-reversed seismograms at the receivers and using them as sources
in an adjoint simulation, and then the value of the gradient is obtained
from the adjoint seismograms recorded at the original earthquake location.

\begin{enumerate}
\item \textbf{Prepare the adjoint sources} \label{enu:Prepare-the-adjoint}

\begin{enumerate}
\item First, run a regular forward simlation (\texttt{SIMULATION\_TYPE =
1} and \texttt{SAVE\_FORWARD = .false.}). You can automatically set
these two variables using the \texttt{\small UTILS/change\_simulation\_type.pl}
script:

\begin{lyxcode}
UTILS/change\_simulation\_type.pl~-f~
\end{lyxcode}
and then collect the recorded seismograms at all the stations given
in \texttt{DATA/STATIONS}.

\item Then select the stations for which you want to compute the time-reversed
adjoint sources and run the adjoint simulation, and compile them into
the \texttt{DATA/STATIONS\_ADJOINT} file, which has the same format
as the regular \texttt{DATA/STATIONS} file.

\begin{itemize}
\item Depending on what type of misfit function is used for the source inversion,
adjoint sources need to be computed from the original recorded seismograms
for the selected stations and saved in the \texttt{SEM/} directory
with the format \texttt{STA.NT.BH?.adj}, where \texttt{STA}, \texttt{NT}
are the station name and network code given in the \texttt{DATA/STATIONS\_ADJOINT}
file, and \texttt{BH?} represents the component name of a particular
adjoint seismogram.
\item The adjoint seismograms are in the same format as the original seismogram
(\texttt{STA.NT.BH?.sem?}), with the same start time, time interval
and record length.
\end{itemize}
\item Notice that even if you choose to time reverse only one component
from one specific station, you still need to supply all three components
because the code is expecting them (you can set the other two components
to be zero).
\item Also note that since time-reversal is done in the code itself, no
explicit time-reversing is needed for the preparation of the adjoint
sources, i.e., the adjoint sources are in the same forward time sense
as the original recorded seismograms.
\end{enumerate}
\item \textbf{Set the related parameters and run the adjoint simulation}\\
In the \texttt{DATA/Par\_file}, set the two related parameters to
be \texttt{SIMULATION\_TYPE = 2} and \texttt{SAVE\_FORWARD = .false.}.
More conveniently, use the scripts \texttt{UTILS/change\_simulation\_type.pl}
to modify the \texttt{Par\_file} automatically (\texttt{change\_simulation\_type.pl
-a}). Then run the solver to launch the adjoint simulation.
\item \textbf{Collect the seismograms at the original source location}

After the adjoint simulation has completed successfully, collect the
seismograms from \texttt{LOCAL\_PATH}.

\begin{itemize}
\item These adjoint seismograms are recorded at the locations of the original
earthquake sources given by the \texttt{DATA/CMTSOLUTION} file, and
have names of the form \texttt{S?????.NT.S??.sem} for the six-component
strain tensor (\texttt{SNN,SEE,SZZ,SNE,SNZ,SEZ}) at these locations,
and \texttt{S?????.NT.BH?.sem} for the three-component displacements
(\texttt{BHN,BHE,BHZ}) recorded at these locations.
\item \texttt{S?????} denotes the source number; for example, if the original
\texttt{CMTSOLUTION} provides only a point source, then the seismograms
collected will start with \texttt{S00001}.
\item These adjoint seismograms provide critical information for the computation
of the gradient of the misfit function.
\end{itemize}
\end{enumerate}

\section{\label{sec:Adjoint-simulation-finite}Adjoint Simulations for Finite-Frequency
Kernels (Kernel Simulation)}

Finite-frequency sensitivity kernels are computed in two successive
simulations (please refer to \citet{LiTr06} and \citet{TrKoLi08} for details).

\begin{enumerate}
\item \textbf{Run a forward simulation with the state variables saved at
the end of the simulation}

Prepare the \texttt{\small CMTSOLUTION} and \texttt{\small STATIONS}
files, set the parameters \texttt{\small SIMULATION\_TYPE}{\small{}
}\texttt{\small =}{\small{} }\texttt{\small 1} and \texttt{\small SAVE\_FORWARD
=}{\small{} }\texttt{\small .true.} in the \texttt{Par\_file} (\texttt{change\_simulation\_type
-F}), and run the solver.

\begin{itemize}
\item Notice that attenuation is not implemented yet for the computation
of finite-frequency kernels; therefore set \texttt{ATTENUATION = .false.}
in the \texttt{Par\_file}.
\item We also suggest you modify the half duration of the \texttt{CMTSOLUTION}
to be similar to the accuracy of the simulation (see Equation \ref{eq:shortest_period})
to avoid too much high-frequency noise in the forward wavefield, although
theoretically the high-frequency noise should be eliminated when convolved
with an adjoint wavefield with the proper frequency content.
\item This forward simulation differs from the regular simulations (\texttt{\small SIMULATION\_TYPE}{\small{}
}\texttt{\small =}{\small{} }\texttt{\small 1} and \texttt{\small SAVE\_FORWARD}{\small{}
}\texttt{\small =}{\small{} }\texttt{\small .false.}) described in
the previous chapters in that the state variables for the last time
step of the simulation, including wavefields of the displacement,
velocity, acceleration, etc., are saved to the \texttt{LOCAL\_PATH}
to be used for the subsequent simulation.
\item For regional simulations, the files recording the absorbing boundary
contribution are also written to the \texttt{LOCAL\_PATH} when \texttt{SAVE\_FORWARD
= .true.}.
\end{itemize}
\item \textbf{Prepare the adjoint sources}

The adjoint sources need to be prepared the same way as described
in the Section \ref{enu:Prepare-the-adjoint}.

\begin{itemize}
\item In the case of travel-time finite-frequency kernel for one source-receiver
pair, i.e., point source from the \texttt{CMTSOLUTION}, and one station
in the \texttt{STATIONS\_ADJOINT} list, we supply a sample program
in \texttt{UTILS/xcut\_velocity} to cut a certain portion of the original
displacement seismograms and convert it into the proper adjoint source
to compute the finite-frequency kernel.

\begin{lyxcode}
xcut\_velocity~t1~t2~ifile{[}0-5]~E/N/Z-ascii-files~{[}baz]
\end{lyxcode}
where \texttt{t1} and \texttt{t2} are the start and end time of the
portion you are interested in, \texttt{ifile} denotes the component
of the seismograms to be used (0 for all three components, 1 for East,
2 for North, and 3 for vertical, 4 for transverse, and 5 for radial
component), \texttt{E/N/Z-ascii-files} indicate the three-component
displacement seismograms in the right order, and \texttt{baz} is the
back-azimuth of the station. Note that \texttt{baz} is only supplied
when \texttt{ifile} = 4 or 5.

\end{itemize}
\item \textbf{Run the kernel simulation}

With the successful forward simulation and the adjoint source ready
in \texttt{SEM/}, set \texttt{SIMULATION\_TYPE = 3} and \texttt{SAVE\_FORWARD
= .false.} in the \texttt{Par\_file(change\_simulation\_type.pl -b),}
and rerun the solver.

\begin{itemize}
\item The adjoint simulation is launched together with the back reconstruction
of the original forward wavefield from the state variables saved from
the previous forward simulation, and the finite-frequency kernels
are computed by the interaction of the reconstructed forward wavefield
and the adjoint wavefield.
\item The back-reconstructed seismograms at the original station locations
are saved to the \texttt{LOCAL\_PATH} at the end of the kernel simulations,
and can be collected to the local disk.
\item These back-constructed seismograms can be compared with the time-reversed
original seismograms to assess the accuracy of the backward reconstruction,
and they should match very well.
\item The arrays for density, P-wave speed and S-wave speed kernels are
also saved in the \texttt{LOCAL\_PATH} with the names \texttt{proc??????\_rho(alpha,beta)\_kernel.bin},
where \texttt{proc??????} represents the processor number, \texttt{rho(alpha,beta)}
are the different types of kernels.
\end{itemize}
\end{enumerate}
In general, the three steps need to be run sequentially to assure
proper access to the necessary files. If the simulations are run through
some cluster scheduling system (e.g., LSF), and the forward simulation
and the subsequent kernel simulations cannot be assigned to the same
set of computer nodes, the kernel simulation will not be able to access
the database files saved by the forward simulation. Solutions for
this dilemma are provided in Chapter~\ref{cha:Scheduler}. Visualization
of the finite-frequency kernels is discussed in Section~\ref{sec:Finite-Frequency-Kernels}.

\chapter{Graphics}

\section{\label{sec:Mesh-graphics}Meshes}

Use the serial code \texttt{combine\_AVS\_DX.f90} (type `\texttt{make
combine\_AVS\_DX}' and then `\texttt{xcombine\_AVS\_DX}') to generate
AVS \url{www.avs.com} output files (in AVS UCD format) or OpenDX \url{www.opendx.org}
output files showing the mesh, the MPI partition (slices), the $\nchunks$
chunks, the source and receiver location, etc. Use the AVS UCD files
\texttt{AVS\_continent\_boundaries.inp} and \texttt{AVS\_plate\_boundaries.inp}
or the OpenDX files \texttt{DX\_continent\_boundaries.dx} and \texttt{DX\_plate\_boundaries.dx}
for reference.

\section{\label{sec:Movies}Movies}

To make a surface or volume movie of the simulation, set parameters
\texttt{MOVIE\_SURFACE}, \texttt{MOVIE\_VOLUME}, and \texttt{NTSTEP\_BETWEEN\_FRAMES}
in the \texttt{Par\_file}. Turning on the movie flags, in particular
\texttt{MOVIE\_VOLUME}, produces large output files. \texttt{MOVIE\_VOLUME}
files are saved in the \texttt{LOCAL\_PATH} directory, whereas \texttt{MOVIE\_SURFACE}
output files are saved in the \texttt{OUTPUT\_FILES} directory. We
save the velocity field. The look of a movie is determined by the
half-duration of the source. The half-duration should be large enough
so that the movie does not contain frequencies that are not resolved
by the mesh, i.e., it should not contain numerical noise. This can
be accomplished by selecting a CMT \texttt{HALF\_DURATION} > 1.1 $\times$
smallest period (see figure \ref{fig:CMTSOLUTION-file}). When \texttt{MOVIE\_SURFACE}
= .\texttt{true.}, the half duration of each source in the \texttt{CMTSOLUTION}
file is replaced by

\begin{quote}
\[
\sqrt{(}\mathrm{\mathtt{HALF\_DURATIO}\mathtt{N}^{2}}+\mathrm{\mathtt{HDUR\_MOVI}\mathtt{E}^{2}})\]
\textbf{NOTE:} If \texttt{HDUR\_MOVIE} is set to 0.0, the code will
select the appropriate value of 1.1 $\times$ smallest period. As
usual, for a point source one can set \texttt{HALF\_DURATION} in the
\texttt{Par\_file} to be 0.0 and \texttt{HDUR\_MOVIE} = 0.0 to get
the highest frequencies resolved by the simulation, but for a finite
source one would keep all the \texttt{HALF\_DURATIONs} as prescribed
by the finite source model and set \texttt{HDUR\_MOVIE} = 0.0.
\end{quote}

\subsection{Movie Surface}

When running \texttt{xspecfem3D} with the \texttt{MOVIE\_SURFACE}
flag turned on, the code outputs \texttt{moviedata??????} files in
the \texttt{OUTPUT\_FILES} directory. The files are in a fairly complicated
binary format, but there are two programs provided to convert the
output into more user friendly formats. The first one, \texttt{create\_movie\_AVS\_DX.f90},
outputs data in ASCII, OpenDX, AVS, or ParaView format. Run the code
from the source directory (type `\texttt{make} \texttt{create\_movie\_AVS\_DX}'
first) to create an input file in your format of choice. The code
will prompt the user for input parameters. The second program \texttt{create\_movie\_GMT.f90}
outputs ascii xyz files, convenient for use with GMT. This code uses
significantly less memory than \texttt{create\_movie\_AVS\_DX.f90}
and is therefore useful for high resolution runs.

The \texttt{SPECFEM3D} code is running in near real-time to produce
animations of southern California earthquakes via the web; see Southern California ShakeMovie\textregistered \url{www.shakemovie.caltech.edu}.

\section{\label{sec:Finite-Frequency-Kernels}Finite-Frequency Kernels}

The finite-frequency kernels computed as explained in Section \ref{sec:Adjoint-simulation-finite}
are saved in the \texttt{LOCAL\_PATH} at the end of the simulation.
Therefore, we first need to collect these files on the front end,
combine them into one mesh file, and visualize them with some auxilliary
programs.

\begin{enumerate}
\item \textbf{Create slice files}

We will only discuss the case of one source-receiver pair, i.e., the
so-called banana-doughnut kernels. Although it is possible to collect
the kernel files from all slices on the front end, it usually takes
up too much storage space (at least tens of gigabytes). Since the
sensitivity kernels are the strongest along the source-receiver great
circle path, it is sufficient to collect only the slices that are
along or close to the great circle path.

A Perl script \texttt{UTILS/slice\_number.pl} that calls MATLAB can
help to figure out the slice numbers that lie along the great circle
path (both the minor and major arcs), as well as the slice numbers
required to produce a full picture of the inner core if your kernel
also illuminates the inner core.

\begin{enumerate}
\item On machines where you have MATLAB access, copy the \texttt{CMTSOLUTION}
file, \texttt{STATIONS\_ADJOINT}, and \texttt{Par\_file}, and run:

\begin{lyxcode}
{\small UTILS/slice\_number.pl~Par\_file~output\_solver.txt~slice\_file}{\small \par}
\end{lyxcode}
which will generate a \texttt{slices\_file}.

\item For cases with multiple sources and multiple receivers, you need to
provide a slice file before proceeding to the next step.
\end{enumerate}
\item \textbf{Collect the kernel files}

After obtaining the slice files, you can collect the corresponding
kernel files from the given slices.

\begin{enumerate}
\item You can use or modify the script \texttt{UTILS/copy\_databases.pl}
to accomplish this:

\begin{lyxcode}
{\small UTILS/copy\_database.pl~slice\_file~lsf\_machine\_file~filename~{[}jobid]}{\small \par}
\end{lyxcode}
where \texttt{\small lsf\_machine\_file} is the machine file generated
by the LSF scheduler, \texttt{\small filename} is the kernel name
(e.g., \texttt{\small rho\_kernel}, \texttt{\small alpha\_kernel}
and \texttt{\small beta\_kernel}), and the optional \texttt{\small jobid}
is the name of the subdirectory under \texttt{\small LOCAL\_PATH}
where all the kernel files are stored.

\item After executing this script, all the necessary mesh topology files
as well as the kernel array files are collected to the local directory
of the front end.
\end{enumerate}
\item \textbf{Combine kernel files into one mesh file}

We use an auxilliary program \texttt{combine\_paraview\_data.f90}
to combine the kernel files from all slices into one mesh file.

\begin{enumerate}
\item Compile it in the global code directory:

\begin{lyxcode}
{\footnotesize make~combine\_paraview\_data~}{\footnotesize \par}

{\footnotesize xcombine\_paraview\_data~slice\_list~filename~input\_dir~output\_dir~high/low-resolution~}{\footnotesize \par}
\end{lyxcode}
where \texttt{input\_dir} is the directory where all the individual
kernel files are stored, and \texttt{output\_dir} is where the mesh
file will be written.

\item Use 1 for a high-resolution mesh, outputting all the GLL points to
the mesh file, or use 0 for low resolution, outputting only the corner
points of the elements to the mesh file.
\item The output mesh file will have the name \texttt{filename\_rho(alpha,beta).mesh}
\end{enumerate}
\item \textbf{Convert mesh files into .vtu files}

\begin{enumerate}
\item We next convert the \texttt{.mesh} file into the VTU (Unstructured
grid file) format which can be viewed in ParaView, for example:

\begin{lyxcode}
UTILS/mesh2vtu.pl~-i~file.mesh~-o~file.vtu
\end{lyxcode}
\item Notice that this Perl script uses a program \texttt{mesh2vtu} in the
\texttt{UTILS/mesh2vtu} directory, which further uses the VTK \url{http://www.vtk.org/}
run-time library for its execution. Therefore, make sure you have
them properly set in the script according to your system.
\end{enumerate}
\item \textbf{Copy over the source and receiver .vtk file}

In the case of a single source and a single receiver, the simulation
also generates the \texttt{OUTPUT\_FILES/sr.vtk} file to describe
the source and receiver locations, which can also be viewed in Paraview
in the next step.

\item \textbf{View the mesh in ParaView}

Finally, we can view the mesh in ParaView \url{www.paraview.org}.

\begin{enumerate}
\item Open ParaView.
\item From the top menu, \textsf{File} $\rightarrow$\textsf{Open data},
select \texttt{file.vtu}, and click the \textsf{Accept} button.

\begin{itemize}
\item If the mesh file is of moderate size, it shows up on the screen; otherwise,
only the bounding box is shown.
\end{itemize}
\item Click \textsf{Display Tab} $\rightarrow$ \textsf{Display Style} $\rightarrow$
\textsf{Representation} and select \textsf{wireframe of surface} to
display it.
\item To create a cross-section of the volumetric mesh, choose \textsf{Filter}
$\rightarrow$ \textsf{cut}, and under \textsf{Parameters Tab}, choose
\textsf{Cut Function} $\rightarrow$ \textsf{plane}.
\item Fill in center and normal information given by the \texttt{global\_slice\_number.pl}
script (either from the standard output or from \texttt{normal\_plane.txt}
file).
\item To change the color scale, go to \textsf{Display Tab} $\rightarrow$
\textsf{Color} $\rightarrow$ \textsf{Edit Color Map} and reselect
lower and upper limits, or change the color scheme.
\item Now load in the source and receiver location file by \textsf{File}
$\rightarrow$ \textsf{Open data}, select \texttt{sr.vt}k, and click
the \textsf{Accept} button. Choose \textsf{Filter} $\rightarrow$
\textsf{Glyph}, and represent the points by `\textsf{spheres}'.
\item For more information about ParaView, see the ParaView Users Guide \url{www.paraview.org/files/v1.6/ParaViewUsersGuide.PDF}.
\end{enumerate}
\end{enumerate}
%
\begin{figure}[H]
\noindent \begin{centering}
\includegraphics[scale=0.6]{figures/3D-S-Kernel}
\par\end{centering}

\caption{(a) Top Panel: Vertical source-receiver cross-section of the S-wave
finite-frequency sensitivity kernel $K_{\beta}$ for station GSC at
an epicentral distance of 176 km from the September 3, 2002, Yorba
Linda earthquake. Lower Panel: Vertical source-receiver cross-section
of the 3D S-wave velocity model used for the spectral-element simulations
\citep{KoLiTrSuStSh04}. (b) The same as (a) but for station HEC at
an epicentral distance of 165 km \citep{LiTr06}.}

\label{figure:P-wave-speed-finite-frequency}
\end{figure}

\chapter{\label{cha:Scheduler}Running through a Scheduler}

The code is usually run on large parallel machines, often PC clusters,
most of which use schedulers, i.e., queuing or batch management systems
to manage the running of jobs from a large number of users. The following
considerations need to be taken into account when running on a system
that uses a scheduler:

\begin{itemize}
\item The processors/nodes to be used for each run are assigned dynamically
by the scheduler, based on availability. Therefore, in order for the
mesher and the solver (or between successive runs of the solver) to
have access to the same database files (if they are stored on hard
drives local to the nodes on which the code is run), they must be
launched in sequence as a single job.
\item On some systems, the nodes to which running jobs are assigned are
not configured for compilation. It may therefore be necessary to pre-compile
both the mesher and the solver. A small program provided in the distribution
called \texttt{\small create\_header\_file.f90} can be used to directly
create \texttt{\small OUTPUT\_FILES/values\_from\_mesher.h} using
the information in the \texttt{\small DATA/Par\_file} without having
to run the mesher (type \texttt{\small `make}{\small{} }\texttt{\small create\_header\_}~\\
\texttt{\small file}' to compile it and `\texttt{\small xcreate\_header\_file}'
to run it; refer to the sample scripts below). The solver can now
be compiled as explained above.
\item One feature of schedulers/queuing systems is that they allow submission
of multiple jobs in a ``launch and forget'' mode. In order to take
advantage of this property, care needs to be taken that output and
intermediate files from separate jobs do not overwrite each other,
or otherwise interfere with other running jobs.
\end{itemize}
We describe here in some detail a job submission procedure for the
Caltech 1024-node cluster, CITerra, under the LSF scheduling system.
We consider the submission of a regular forward simulation. The two
main scripts are \texttt{\small run\_lsf.bash}, which compiles the
Fortran code and submits the job to the scheduler, and \texttt{\small go\_mesher\_solver\_lsf}~\\
\texttt{\small .bash}, which contains the instructions that make up
the job itself. These scripts can be found in \texttt{\small UTILS/}
directory and can straightforwardly be modified and adapted to meet
more specific running needs.

\section{\texttt{run\_lsf.bash}}

This script first sets the job queue to be `normal'. It then compiles
the mesher and solver together, figures out the number of processors
required for this simulation from the \texttt{DATA/Par\_file}, and
submits the LSF job.

\begin{lyxcode}
\#!/bin/bash

\#~use~the~normal~queue~unless~otherwise~directed~queue=\char`\"{}-q~normal\char`\"{}~

if~{[}~\$\#~-eq~1~];~then

~~~~~~~~echo~\char`\"{}Setting~the~queue~to~\$1\char`\"{}

~~~~~~~~queue=\char`\"{}-q~\$1\char`\"{}~

fi~\\
~\\
\#~compile~the~mesher~and~the~solver~

d=`date'~echo~\char`\"{}Starting~compilation~\$d\char`\"{}~

make~clean~

make~meshfem3D~

make~create\_header\_file~

xcreate\_header\_file~

make~specfem3D~

d=`date'~

echo~\char`\"{}Finished~compilation~\$d\char`\"{}~\\
~\\
\#~compute~total~number~of~nodes~needed~

NPROC\_XI=`grep~NPROC\_XI~DATA/Par\_file~|~cut~-c~34-~'~

NPROC\_ETA=`grep~NPROC\_ETA~DATA/Par\_file~|~cut~-c~34-~'~~\\
~\\
\#~total~number~of~nodes~is~the~product~of~the~values~read~

numnodes=\$((~\$NPROC\_XI~{*}~\$NPROC\_ETA~))~\\
~\\
echo~\char`\"{}Submitting~job\char`\"{}~

bsub~\$queue~-n~\$numnodes~-W~60~-K~<go\_mesher\_solver\_lsf.bash~
\end{lyxcode}

\section{\texttt{go\_mesher\_solver\_lsf.bash}}

This script describes the job itself, including setup steps that can
only be done once the scheduler has assigned a job-ID and a set of
compute nodes to the job, the \texttt{run\_lsf.bash} commands used
to run the mesher and the solver, and calls to scripts that collect
the output seismograms from the compute nodes and perform clean-up
operations.

\begin{enumerate}
\item First the script directs the scheduler to save its own output and
output from \texttt{stdout} into \texttt{\small OUTPUT\_FILES/\%J.o},
where \texttt{\%J} is short-hand for the job-ID; it also tells the
scheduler what version of \texttt{mpich} to use (\texttt{mpich\_gm})
and how to name this job (\texttt{go\_mesher\_solver\_lsf}).
\item The script then creates a list of the nodes allocated to this job
by echoing the value of a dynamically set environment variable \texttt{LSB\_MCPU\_HOSTS}
and parsing the output into a one-column list using the Perl script
\texttt{UTILS/remap\_lsf\_machines.pl}. It then creates a set of scratch
directories on these nodes (\texttt{\small /scratch/}~\\
\texttt{\small \$USER/DATABASES\_MPI}) to be used as the \texttt{LOCAL\_PATH}
for temporary storage of the database files. The scratch directories
are created using \texttt{shmux}, a shell multiplexor that can execute
the same commands on many hosts in parallel. \texttt{shmux} is available
from Shmux \url{web.taranis.org/shmux/}. Make sure that the \texttt{LOCAL\_PATH}
parameter in \texttt{DATA/Par\_file} is also set properly.
\item The next portion of the script launches the mesher and then the solver
using \texttt{run\_lsf.bash}.
\item The final portion of the script collects the seismograms and performs
clean up on the nodes, using the Perl scripts \texttt{collect\_seismo\_lsf\_multi.pl}
and \texttt{cleanmulti.pl}.
\end{enumerate}
\begin{lyxcode}
\#!/bin/bash~-v~

\#BSUB~-o~OUTPUT\_FILES/\%J.o~

\#BSUB~-a~mpich\_gm~

\#BSUB~-J~go\_mesher\_solver\_lsf~\\
~\\
\#~set~up~local~scratch~directories

BASEMPIDIR=/scratch/\$USER/DATABASES\_MPI

mkdir~-p~OUTPUT\_FILE

echo~\char`\"{}\$LSB\_MCPU\_HOSTS\char`\"{}~>~OUTPUT\_FILES/lsf\_machines~

echo~\char`\"{}\$LSB\_JOBID\char`\"{}~>~OUTPUT\_FILES/jobid

remap\_lsf\_machines.pl~OUTPUT\_FILES/lsf\_machines~>OUTPUT\_FILES/machines

shmux~-M50~-Sall~-c~\char`\"{}rm~-r~-f~/scratch/\$USER;~\textbackslash{}

~~~~~~mkdir~-p~/scratch/\$USER;~mkdir~-p~\$BASEMPIDIR\char`\"{}~\textbackslash{}

~~~~~~-~<~OUTPUT\_FILES/machines~>/dev/null~\\
~\\
\#~run~the~specfem~program

current\_pwd=\$PWD

run\_lsf.bash~-{}-gm-no-shmem~-{}-gm-copy-env~\$current\_pwd/xmeshfem3D

run\_lsf.bash~-{}-gm-no-shmem~-{}-gm-copy-env~\$current\_pwd/xspecfem3D~\\
~\\
\#~collect~seismograms~and~clean~up

mkdir~-p~SEM

cd~SEM

collect\_seismo.pl~../OUTPUT\_FILES/lsf\_machines

cleanbase.pl~../OUTPUT\_FILES/machines
\end{lyxcode}

\chapter{Post-Processing Scripts}

Several post-processing scripts/programs are provided in the \texttt{UTILS/}
directory, and most of them need to be adjusted when used on different
systems, for example, the path of the executable programs. Here we
only list the available scripts and provide a brief description, and
you can either refer to the related sections for detailed usage or,
in a lot of cases, type the script/program name without arguments
for its usage.

\section{Collect Synthetic Seismograms}

The forward and adjoint simulations generate synthetic seismograms
in the \texttt{LOCAL\_PATH}. For the forward simulation, the files
are named \texttt{STA.NT.BH?.semd} for two-column time series, or
\texttt{STA.NT.BH?.semd.sac} for ASCII SAC format, where STA and NT
are the station name and network code, and \texttt{BH?} stands for
the component name. The adjont simulations generate synthetic seismograms
with the name \texttt{S?????.NT.S??.sem} (refer to Section \ref{sec:Adjoint-simulation-sources}
for details). The kernel simulations output the back-reconstructed
synthetic seismogram in the name \texttt{STA.NT.BH?.semd}, mainly
for the purpose of checking the accuracy of the reconstruction. Refer
to Section \ref{sec:Adjoint-simulation-finite} for further details.

To collect the synthetics onto the frontend, you can use the \texttt{UTILS/collect\_seismo.pl}
machines script:

\begin{lyxcode}
collect\_seismo.pl~machines
\end{lyxcode}

\section{Clean Local Database}

After all the simulations are done, the seismograms are collected,
and the useful database files are copied to the frontend, you may
need to clean the local scratch disk for the next simulation. This
is especially important in the case of 1- or 2-chunk kernel simulation,
where very large files are generated for the absorbing boundaries
to help with the reconstruction of the regular forward wavefield.
A sample script is provided in \texttt{UTILS/}:

\begin{lyxcode}
cleanbase.pl~machines
\end{lyxcode}

\section{Process Data and Synthetics\label{sec:Process-data-and-syn}}

In many cases, the SEM synthetics are calculated and compared to data
seismograms recorded at seismic stations. Since the SEM synthetics
are accurate for a certain frequency range, both the original data
and the synthetics need to be processed before a comparison can be
made. We generally use the following scripts:

\subsection{\texttt{process\_trinet\_data.pl}}

This script cuts a given portion of the original data, filters it,
transfers the data into a displacement record, and picks the first
P and S arrivals. For more functionality, type `\texttt{process\_trinet\_data.pl}'
without any argument. An example of the usage of the script:

\begin{lyxcode}
{\footnotesize process\_trinet\_data.pl~-m~CMTSOLUTION~-l~0/180~-t~2/40~-i~dir~-p~-x~bp~~9703873{*}.BH?.SAC}{\footnotesize \par}
\end{lyxcode}
which has cut all the sac files between 0 and 180 seconds, filtered
them between 2 and 40 seconds, transfered them into displacement records
using the polezero files in \texttt{dir} directory, picked the first
P and S arrivals, and added suffix `\texttt{bp}' to the file names.

Note that all of the scripts in this section actually use the SAC
and/or IASP91 to do the core operations; therefore make sure that
the SAC and IASP91 packages are installed properly on your system,
and that all the environment variables are set properly before running
these scripts.

\subsection{\texttt{process\_trinet\_syn.pl}}

This script converts the synthetic output from the SEM code from ASCII
to SAC format, and performs similar operations as `\texttt{process\_trinet\_data.pl}'.
An example of the usage of the script:

\begin{lyxcode}
{\footnotesize process\_trinet\_syn.pl~-m~CMTSOLUTION~-a~STATIONS~-l~0/180~-t~2/40~-p~-x~bp~syn/{*}.BH?.semd}{\footnotesize \par}
\end{lyxcode}
which will convert the synthetics into SAC format, add event and station
information into the SAC headers, cut the SAC files between 0 and
180 seconds, filter them between 2 and 40 seconds, pick the first
P and S arrivals, and add the suffix `\texttt{bp}' to the file names.

More options are available for this script, such as adding time shift
to the origin time of the synthetics, convolving the synthetics with
a triangular source time function with a given half duration, etc.
Type \texttt{process\_trinet\_syn.pl} without any argument for a detailed
usage.

\subsection{\texttt{rotate.pl}}

The original data and synthetics have three components: vertical (BHZ),
north (BHN) and east (BHE). However, for most seismology applications,
transverse and radial components are also desirable. Therefore, we
need to rotate the horizontal components of both the data and the
synthetics to the transverse and radial direction, and \texttt{\small rotate.pl}
can be used to accomplish this:

\begin{lyxcode}
rotate.pl~-l~0~-L~180~-d~DATA/{*}.BHE.SAC.bp~

rotate.pl~-l~0~-L~180~SEM/{*}.BHE.semd.sac.bp~
\end{lyxcode}
where the first command performs rotation on the SAC data obtained
through Seismogram Transfer Program (STP) \url{http://www.data.scec.org/STP/stp.html},
while the second command rotates the processed SEM synthetics.

\section{Plot Movie Snapshots and Synthetic Shakemaps}

\subsection{\texttt{movie2gif.pl}}

With the movie data saved in \texttt{OUTPUT\_FILES/} at the end of
a movie simulation (\texttt{\small MOVIE\_SURFACE=.true.}), you can
run the \texttt{`create\_movie\_GMT}' code to convert these binary
movie data into GMT xyz files for futher processing. A sample script
\texttt{movie2gif.pl} is provided to do this conversion, and then
plot the movie snapshots in GMT, for example:

\begin{lyxcode}
movie2gif.pl~-m~CMTSOLUTION~-g~-f~1/40~-n~-2~-p~
\end{lyxcode}
which for the first through the 40th movie frame, converts the \texttt{moviedata}
files into GMT xyz files, interpolates them using the 'nearneighbor'
command in GMT, and plots them on a 2D topography map. Note that `\texttt{-2}'
and `\texttt{-p}' are both optional.

\subsection{\texttt{plot\_shakemap.pl}}

With the shakemap data saved in \texttt{OUTPUT\_FILES/} at the end
of a shakemap simulation (\texttt{CREATE\_SHAKEMAP=.true.}), you can
also run \texttt{`create\_movie\_GMT}' code to convert the binary
shakemap data into GMT xyz files. A sample script \texttt{plot\_shakemap.pl}
is provided to do this conversion, and then plot the shakemaps in
GMT, for example:

\begin{lyxcode}
plot\_shakemap.pl~data\_dir~type(1,2,3)~CMTSOLUTION~
\end{lyxcode}
where \texttt{type=1} for a displacement shakemap, \texttt{2} for
velocity, and \texttt{3} for acceleration.

\section{Map Local Database}

A sample program \texttt{remap\_database} is provided to map the local
database from a set of machines to another set of machines. This is
especially useful when you want to run mesher and solver, or different
types of solvers separately through a scheduler (refer to Chapter~\ref{cha:Scheduler}).

\begin{lyxcode}
run\_lsf.bash~-{}-gm-no-shmem~-{}-gm-copy-env~remap\_database~old\_machines~150
\end{lyxcode}
where \texttt{old\_machines} is the LSF machine file used in the previous
simulation, and \texttt{150} is the number of processors in total.

\chapter*{\label{cha:Bug-Reports-and}Bug Reports and Suggestions for Improvements}

To report bugs or suggest improvements to the code, please send an
e-mail to the CIG Computational Seismology Mailing List \url{cig-seismo@geodynamics.org}
or Jeroen Tromp \url{jtromp-AT-princeton.edu}, and/or use our online
bug tracking system Roundup \url{www.geodynamics.org/roundup}.

\chapter*{Notes \& Acknowledgments}

In order to keep the software package thread-safe in case a multithreaded
implementation of MPI is used, developers should not add modules or
common blocks to the source code but rather use regular subroutine
arguments (which can be grouped in ``derived types'' if needed for
clarity).

The Gauss-Lobatto-Legendre subroutines in \texttt{gll\_library.f90}
are based in part on software libraries from the Massachusetts Institute
of Technology, Department of Mechanical Engineering (Cambridge, Massachusetts, USA).
The non-structured global numbering software was provided by Paul
F. Fischer (Brown University, Providence, Rhode Island, USA, now at Argonne National Laboratory, USA).

OpenDX \url{http://www.opendx.org} is open-source based on IBM Data
Explorer, AVS \url{http://www.avs.com} is a trademark of Advanced
Visualization Systems, and ParaView \url{http://www.paraview.com}
is an open-source visualization platform.

The main developers of the \texttt{SPECFEM3D} source code are Dimitri
Komatitsch, Jeroen Tromp, and Qinya Liu. The following individuals
(listed in alphabetical order) have also contributed to the development
or improvement of the source code: Min Chen, Vala Hj\"orleifsd\'ottir,
Jes\'us Labarta, and Leif Strand. The following individuals (listed in alphabetical
order) contributed to this manual: Min Chen, Vala Hj\"orleifsd\'ottir,
Sue Kientz, Dimitri Komatitsch, Qinya Liu, Alessia Maggi,
Carl Tape, and Jeroen Tromp. The manual's cover graphic was created
by Santiago Lombeyda from Caltech's Center for Advanced Computing Research (CACR) \url{http://www.cacr.caltech.edu}.
Older versions of the code were initially developed by Dimitri Komatitsch at Institut de Physique du Globe (France)
and then by Dimitri Komatitsch and Jeroen Tromp at Harvard University (USA).

Please e-mail your feedback, questions, comments, and suggestions
to Jeroen Tromp \url{jtromp-AT-princeton.edu} or to the CIG Computational Seismology Mailing List \url{cig-seismo@geodynamics.org}.

\chapter*{Copyright}

Main authors: Dimitri Komatitsch and Jeroen Tromp

Seismological Laboratory, California Institute of Technology, USA,
and University of Pau / CNRS / INRIA, France

$\copyright$ California Institute of Technology and University of Pau / CNRS / INRIA, February 2010

This program is free software; you can redistribute it and/or modify
it under the terms of the GNU General Public License as published
by the Free Software Foundation (see Appendix \ref{cha:License}).

\bibliography{bibliography}

\appendix

\chapter{\label{cha:Coordinates}Reference Frame Convention}

The code uses the following convention for the Cartesian reference
frame:

\begin{itemize}
\item the $x$ axis points East
\item the $y$ axis points North
\item the $z$ axis points up
\end{itemize}
Note that this convention is different from both the \citet{AkRi80}
convention and the Harvard Centroid-Moment Tensor (CMT) convention.
The Aki \& Richards convention is

\begin{itemize}
\item the $x$ axis points North
\item the $y$ axis points East
\item the $z$ axis points down
\end{itemize}
and the Harvard CMT convention is

\begin{itemize}
\item the $x$ axis points South
\item the $y$ axis points East
\item the $z$ axis points up
\end{itemize}

\chapter{\label{cha:License}License }

\textbf{GNU GENERAL PUBLIC LICENSE Version 2, June 1991. Copyright
(C) 1989, 1991 Free Software Foundation, Inc. 59 Temple Place, Suite
330, Boston, MA 02111-1307 USA} \\
Everyone is permitted to copy and distribute verbatim copies of this
license document, but changing it is not allowed.

\section*{Preamble}

The licenses for most software are designed to take away your freedom
to share and change it. By contrast, the GNU General Public License
is intended to guarantee your freedom to share and change free software
-- to make sure the software is free for all its users. This General
Public License applies to most of the Free Software Foundation's software
and to any other program whose authors commit to using it. (Some other
Free Software Foundation software is covered by the GNU Library General
Public License instead.) You can apply it to your programs, too.

When we speak of free software, we are referring to freedom, not price.
Our General Public Licenses are designed to make sure that you have
the freedom to distribute copies of free software (and charge for
this service if you wish), that you receive source code or can get
it if you want it, that you can change the software or use pieces
of it in new free programs; and that you know you can do these things.

To protect your rights, we need to make restrictions that forbid anyone
to deny you these rights or to ask you to surrender the rights. These
restrictions translate to certain responsibilities for you if you
distribute copies of the software, or if you modify it.

For example, if you distribute copies of such a program, whether gratis
or for a fee, you must give the recipients all the rights that you
have. You must make sure that they, too, receive or can get the source
code. And you must show them these terms so they know their rights.

We protect your rights with two steps:

\begin{enumerate}
\item Copyright the software, and
\item Offer you this license which gives you legal permission to copy, distribute
and/or modify the software.
\end{enumerate}
Also, for each author's protection and ours, we want to make certain
that everyone understands that there is no warranty for this free
software. If the software is modified by someone else and passed on,
we want its recipients to know that what they have is not the original,
so that any problems introduced by others will not reflect on the
original authors' reputations.

Finally, any free program is threatened constantly by software patents.
We wish to avoid the danger that redistributors of a free program
will individually obtain patent licenses, in effect making the program
proprietary. To prevent this, we have made it clear that any patent
must be licensed for everyone's free use or not licensed at all.

The precise terms and conditions for copying, distribution and modification
follow.

\section*{GNU GENERAL PUBLIC LICENSE TERMS AND CONDITIONS FOR COPYING, DISTRIBUTION
AND MODIFICATION }

\begin{itemize}

\item[0.]This License applies to any program or other work which
contains a notice placed by the copyright holder saying it may be
distributed under the terms of this General Public License. The ``Program''
below refers to any such program or work, and a ``work based on the
Program'' means either the Program or any derivative work under copyright
law: that is to say, a work containing the Program or a portion of
it, either verbatim or with modifications and/or translated into another
language. (Hereinafter, translation is included without limitation
in the term ``modification.'') Each licensee is addressed as ``you.''\\
\\
Activities other than copying, distribution and modification are not
covered by this License; they are outside its scope. The act of running
the Program is not restricted, and the output from the Program is
covered only if its contents constitute a work based on the Program
(independent of having been made by running the Program). Whether
that is true depends on what the Program does.

\end{itemize}

\begin{enumerate}
\item You may copy and distribute verbatim copies of the Program's source
code as you receive it, in any medium, provided that you conspicuously
and appropriately publish on each copy an appropriate copyright notice
and disclaimer of warranty; keep intact all the notices that refer
to this License and to the absence of any warranty; and give any other
recipients of the Program a copy of this License along with the Program.

You may charge a fee for the physical act of transferring a copy,
and you may at your option offer warranty protection in exchange for
a fee.

\item You may modify your copy or copies of the Program or any portion of
it, thus forming a work based on the Program, and copy and distribute
such modifications or work under the terms of Section 1 above, provided
that you also meet all of these conditions:

\begin{enumerate}
\item You must cause the modified files to carry prominent notices stating
that you changed the files and the date of any change.
\item You must cause any work that you distribute or publish, that in whole
or in part contains or is derived from the Program or any part thereof,
to be licensed as a whole at no charge to all third parties under
the terms of this License.
\item If the modified program normally reads commands interactively when
run, you must cause it, when started running for such interactive
use in the most ordinary way, to print or display an announcement
including an appropriate copyright notice and a notice that there
is no warranty (or else, saying that you provide a warranty) and that
users may redistribute the program under these conditions, and telling
the user how to view a copy of this License. (Exception: if the Program
itself is interactive but does not normally print such an announcement,
your work based on the Program is not required to print an announcement.)
\end{enumerate}
These requirements apply to the modified work as a whole. If identifiable
sections of that work are not derived from the Program, and can be
reasonably considered independent and separate works in themselves,
then this License, and its terms, do not apply to those sections when
you distribute them as separate works. But when you distribute the
same sections as part of a whole which is a work based on the Program,
the distribution of the whole must be on the terms of this License,
whose permissions for other licensees extend to the entire whole,
and thus to each and every part regardless of who wrote it.

Thus, it is not the intent of this section to claim rights or contest
your rights to work written entirely by you; rather, the intent is
to exercise the right to control the distribution of derivative or
collective works based on the Program.

In addition, mere aggregation of another work not based on the Program
with the Program (or with a work based on the Program) on a volume
of a storage or distribution medium does not bring the other work
under the scope of this License.

\item You may copy and distribute the Program (or a work based on it, under
Section 2) in object code or executable form under the terms of Sections
1 and 2 above provided that you also do one of the following:

\begin{enumerate}
\item Accompany it with the complete corresponding machine-readable source
code, which must be distributed under the terms of Sections 1 and
2 above on a medium customarily used for software interchange; or,
\item Accompany it with a written offer, valid for at least three years,
to give any third party, for a charge no more than your cost of physically
performing source distribution, a complete machine-readable copy of
the corresponding source code, to be distributed under the terms of
Sections 1 and 2 above on a medium customarily used for software interchange;
or,
\item Accompany it with the information you received as to the offer to
distribute corresponding source code. (This alternative is allowed
only for noncommercial distribution and only if you received the program
in object code or executable form with such an offer, in accord with
Subsection b above.)
\end{enumerate}
The source code for a work means the preferred form of the work for
making modifications to it. For an executable work, complete source
code means all the source code for all modules it contains, plus any
associated interface definition files, plus the scripts used to control
compilation and installation of the executable. However, as a special
exception, the source code distributed need not include anything that
is normally distributed (in either source or binary form) with the
major components (compiler, kernel, and so on) of the operating system
on which the executable runs, unless that component itself accompanies
the executable.

If distribution of executable or object code is made by offering access
to copy from a designated place, then offering equivalent access to
copy the source code from the same place counts as distribution of
the source code, even though third parties are not compelled to copy
the source along with the object code.

\item You may not copy, modify, sublicense, or distribute the Program except
as expressly provided under this License. Any attempt otherwise to
copy, modify, sublicense or distribute the Program is void, and will
automatically terminate your rights under this License. However, parties
who have received copies, or rights, from you under this License will
not have their licenses terminated so long as such parties remain
in full compliance.
\item You are not required to accept this License, since you have not signed
it. However, nothing else grants you permission to modify or distribute
the Program or its derivative works. These actions are prohibited
by law if you do not accept this License. Therefore, by modifying
or distributing the Program (or any work based on the Program), you
indicate your acceptance of this License to do so, and all its terms
and conditions for copying, distributing or modifying the Program
or works based on it.
\item Each time you redistribute the Program (or any work based on the Program),
the recipient automatically receives a license from the original licensor
to copy, distribute or modify the Program subject to these terms and
conditions. You may not impose any further restrictions on the recipients'
exercise of the rights granted herein. You are not responsible for
enforcing compliance by third parties to this License.
\item If, as a consequence of a court judgment or allegation of patent infringement
or for any other reason (not limited to patent issues), conditions
are imposed on you (whether by court order, agreement or otherwise)
that contradict the conditions of this License, they do not excuse
you from the conditions of this License. If you cannot distribute
so as to satisfy simultaneously your obligations under this License
and any other pertinent obligations, then as a consequence you may
not distribute the Program at all. For example, if a patent license
would not permit royalty-free redistribution of the Program by all
those who receive copies directly or indirectly through you, then
the only way you could satisfy both it and this License would be to
refrain entirely from distribution of the Program.

If any portion of this section is held invalid or unenforceable under
any particular circumstance, the balance of the section is intended
to apply and the section as a whole is intended to apply in other
circumstances.

It is not the purpose of this section to induce you to infringe any
patents or other property right claims or to contest validity of any
such claims; this section has the sole purpose of protecting the integrity
of the free software distribution system, which is implemented by
public license practices. Many people have made generous contributions
to the wide range of software distributed through that system in reliance
on consistent application of that system; it is up to the author/donor
to decide if he or she is willing to distribute software through any
other system and a licensee cannot impose that choice.

This section is intended to make thoroughly clear what is believed
to be a consequence of the rest of this License.

\item If the distribution and/or use of the Program is restricted in certain
countries either by patents or by copyrighted interfaces, the original
copyright holder who places the Program under this License may add
an explicit geographical distribution limitation excluding those countries,
so that distribution is permitted only in or among countries not thus
excluded. In such case, this License incorporates the limitation as
if written in the body of this License.
\item The Free Software Foundation may publish revised and/or new versions
of the General Public License from time to time. Such new versions
will be similar in spirit to the present version, but may differ in
detail to address new problems or concerns.

Each version is given a distinguishing version number. If the Program
specifies a version number of this License which applies to it and
``any later version,'' you have the option of following the terms
and conditions either of that version or of any later version published
by the Free Software Foundation. If the Program does not specify a
version number of this License, you may choose any version ever published
by the Free Software Foundation.

\item If you wish to incorporate parts of the Program into other free programs
whose distribution conditions are different, write to the author to
ask for permission. For software which is copyrighted by the Free
Software Foundation, write to the Free Software Foundation; we sometimes
make exceptions for this. Our decision will be guided by the two goals
of preserving the free status of all derivatives of our free software
and of promoting the sharing and reuse of software generally.
\end{enumerate}

\subsection*{NO WARRANTY }

\begin{itemize}

\item[11.]BECAUSE THE PROGRAM IS LICENSED FREE OF CHARGE, THERE IS
NO WARRANTY FOR THE PROGRAM, TO THE EXTENT PERMITTED BY APPLICABLE
LAW. EXCEPT WHEN OTHERWISE STATED IN WRITING THE COPYRIGHT HOLDERS
AND/OR OTHER PARTIES PROVIDE THE PROGRAM ``AS IS'' WITHOUT WARRANTY
OF ANY KIND, EITHER EXPRESSED OR IMPLIED, INCLUDING, BUT NOT LIMITED
TO, THE IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR
PURPOSE. THE ENTIRE RISK AS TO THE QUALITY AND PERFORMANCE OF THE
PROGRAM IS WITH YOU. SHOULD THE PROGRAM PROVE DEFECTIVE, YOU ASSUME
THE COST OF ALL NECESSARY SERVICING, REPAIR OR CORRECTION.

\item[12.]IN NO EVENT UNLESS REQUIRED BY APPLICABLE LAW OR AGREED
TO IN WRITING WILL ANY COPYRIGHT HOLDER, OR ANY OTHER PARTY WHO MAY
MODIFY AND/OR REDISTRIBUTE THE PROGRAM AS PERMITTED ABOVE, BE LIABLE
TO YOU FOR DAMAGES, INCLUDING ANY GENERAL, SPECIAL, INCIDENTAL OR
CONSEQUENTIAL DAMAGES ARISING OUT OF THE USE OR INABILITY TO USE THE
PROGRAM (INCLUDING BUT NOT LIMITED TO LOSS OF DATA OR DATA BEING RENDERED
INACCURATE OR LOSSES SUSTAINED BY YOU OR THIRD PARTIES OR A FAILURE
OF THE PROGRAM TO OPERATE WITH ANY OTHER PROGRAMS), EVEN IF SUCH HOLDER
OR OTHER PARTY HAS BEEN ADVISED OF THE POSSIBILITY OF SUCH DAMAGES.

\end{itemize}

\section*{END OF TERMS AND CONDITIONS }

\subsection*{How to Apply These Terms to Your New Programs}

If you develop a new program, and you want it to be of the greatest
possible use to the public, the best way to achieve this is to make
it free software which everyone can redistribute and change under
these terms.

To do so, attach the following notices to the program. It is safest
to attach them to the start of each source file to most effectively
convey the exclusion of warranty; and each file should have at least
the ``copyright'' line and a pointer to where the full notice is found.
For example:

\begin{quote}
One line to give the program's name and a brief idea of what it does.
Copyright {\footnotesize $\copyright$ (}year) (name of author)

This program is free software; you can redistribute it and/or modify
it under the terms of the GNU General Public License as published
by the Free Software Foundation; either version 2 of the License,
or (at your option) any later version.

This program is distributed in the hope that it will be useful, but
WITHOUT ANY WARRANTY; without even the implied warranty of MERCHANTABILITY
or FITNESS FOR A PARTICULAR PURPOSE. See the GNU General Public License
for more details.

You should have received a copy of the GNU General Public License
along with this program; if not, write to the Free Software Foundation,
Inc., 59 Temple Place, Suite 330, Boston, MA 02111-1307 USA
\end{quote}
Also add information on how to contact you by electronic and paper
mail.

If the program is interactive, make it output a short notice like
this when it starts in an interactive mode:

\begin{quote}
Gnomovision version 69, Copyright $\copyright$ year name of author Gnomovision
comes with ABSOLUTELY NO WARRANTY; for details type `show w'. This
is free software, and you are welcome to redistribute it under certain
conditions; type `show c' for details.
\end{quote}
The hypothetical commands `show w' and `show c' should show the appropriate
parts of the General Public License. Of course, the commands you use
may be called something other than `show w' and `show c'; they could
even be mouse-clicks or menu items -- whatever suits your program.

You should also get your employer (if you work as a programmer) or
your school, if any, to sign a ``copyright disclaimer'' for the program,
if necessary. Here is a sample; alter the names:

\begin{quote}
Yoyodyne, Inc., hereby disclaims all copyright interest in the program
`Gnomovision' (which makes passes at compilers) written by James Hacker.

(signature of Ty Coon)\\
1 April 1989 \\
Ty Coon, President of Vice
\end{quote}
This General Public License does not permit incorporating your program
into proprietary programs. If your program is a subroutine library,
you may consider it more useful to permit linking proprietary applications
with the library. If this is what you want to do, use the GNU Library
General Public License instead of this License.
\end{document}

### 2010-10-21 doc/USER_MANUAL/compile_LaTeX_manual.sh

@@ -1,0 +1,39 @@
+#!/bin/sh
+
+/bin/rm -rf *.dvi >  /dev/null
+/bin/rm -rf *.log >  /dev/null
+/bin/rm -rf *.out >  /dev/null
+/bin/rm -rf *.aux >  /dev/null
+/bin/rm -rf *.toc >  /dev/null
+/bin/rm -rf *.blg >  /dev/null
+/bin/rm -rf *.bbl >  /dev/null
+/bin/rm -rf *.lof >  /dev/null
+/bin/rm -rf *.lot >  /dev/null
+/bin/rm -rf *.plt >  /dev/null
+/bin/rm -rf *.fff >  /dev/null
+/bin/rm -rf *.ttt >  /dev/null
+/bin/rm -rf *.tit >  /dev/null
+/bin/rm -rf *.spl >  /dev/null
+
+	pdflatex manual_SPECFEM3D
+	bibtex manual_SPECFEM3D
+	pdflatex manual_SPECFEM3D
+	pdflatex manual_SPECFEM3D
+	pdflatex manual_SPECFEM3D
+	pdflatex manual_SPECFEM3D
+
+/bin/rm -rf *.dvi >  /dev/null
+/bin/rm -rf *.log >  /dev/null
+/bin/rm -rf *.out >  /dev/null
+/bin/rm -rf *.aux >  /dev/null
+/bin/rm -rf *.toc >  /dev/null
+/bin/rm -rf *.blg >  /dev/null
+/bin/rm -rf *.bbl >  /dev/null
+/bin/rm -rf *.lof >  /dev/null
+/bin/rm -rf *.lot >  /dev/null
+/bin/rm -rf *.plt >  /dev/null
+/bin/rm -rf *.fff >  /dev/null
+/bin/rm -rf *.ttt >  /dev/null
+/bin/rm -rf *.tit >  /dev/null
+/bin/rm -rf *.spl >  /dev/null
+

### 2010-10-21 src/shared/exit_mpi.f90

@@ -1,0 +1,82 @@
+!=====================================================================
+!
+!               S p e c f e m 3 D  V e r s i o n  1 . 4
+!               ---------------------------------------
+!
+!                 Dimitri Komatitsch and Jeroen Tromp
+!    Seismological Laboratory - California Institute of Technology
+!         (c) California Institute of Technology September 2006
+!
+! This program is free software; you can redistribute it and/or modify
+! it under the terms of the GNU General Public License as published by
+! the Free Software Foundation; either version 2 of the License, or
+! (at your option) any later version.
+!
+! This program is distributed in the hope that it will be useful,
+! but WITHOUT ANY WARRANTY; without even the implied warranty of
+! MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+! GNU General Public License for more details.
+!
+! You should have received a copy of the GNU General Public License along
+! with this program; if not, write to the Free Software Foundation, Inc.,
+! 51 Franklin Street, Fifth Floor, Boston, MA 02110-1301 USA.
+!
+!=====================================================================
+
+! end the simulation and exit MPI
+
+  subroutine exit_MPI(myrank,error_msg)
+
+  implicit none
+
+  include "constants.h"
+
+! identifier for error message file
+  integer, parameter :: IERROR = 30
+
+  integer myrank
+  character(len=*) error_msg
+
+  character(len=80) outputname
+  character(len=256) OUTPUT_FILES
+
+! write error message to screen
+  write(*,*) error_msg(1:len(error_msg))
+  write(*,*) 'Error detected, aborting MPI... proc ',myrank
+
+! write error message to file
+  call get_value_string(OUTPUT_FILES, 'OUTPUT_FILES', OUTPUT_FILES_PATH(1:len_trim(OUTPUT_FILES_PATH)))
+  write(outputname,"('/error_message',i6.6,'.txt')") myrank
+  open(unit=IERROR,file=trim(OUTPUT_FILES)//outputname,status='unknown')
+  write(IERROR,*) error_msg(1:len(error_msg))
+  write(IERROR,*) 'Error detected, aborting MPI... proc ',myrank
+  close(IERROR)
+
+! close output file
+  if(myrank == 0 .and. IMAIN /= ISTANDARD_OUTPUT) close(IMAIN)
+
+  call stop_all()
+
+  end subroutine exit_MPI
+
+!
+!----
+!
+
[…]

### 2010-10-21 utils/Visualization/opendx_AVS/evaluate_colorbar.f90

@@ -1,0 +1,40 @@
+
+ program rescalenonlinear
+
+! determine GMT colorbar mapping
+
+ implicit none
+
+ double precision, parameter :: POWERVAL = 0.25d0
+
+ double precision a
+
+ a = 0.1d0
+ write(*,*) a,' gives ',a**POWERVAL
+
+ a = 0.2d0
+ write(*,*) a,' gives ',a**POWERVAL
+
+ a = 0.3d0
+ write(*,*) a,' gives ',a**POWERVAL
+
+ a = 0.4d0
+ write(*,*) a,' gives ',a**POWERVAL
+
+ a = 0.5d0
+ write(*,*) a,' gives ',a**POWERVAL
+
+ a = 0.6d0
+ write(*,*) a,' gives ',a**POWERVAL
+
+ a = 0.7d0
+ write(*,*) a,' gives ',a**POWERVAL
+
+ a = 0.8d0
+ write(*,*) a,' gives ',a**POWERVAL
+
+ a = 0.9d0
+ write(*,*) a,' gives ',a**POWERVAL
+
+ end
+

### 2010-10-21 utils/cmt_frechet/cmt.h

@@ -1,0 +1,68 @@
+!=====================================================================
+!
+!          S p e c f e m 3 D  B a s i n  V e r s i o n  1 . 2
+!          --------------------------------------------------
+!
+!                 Dimitri Komatitsch and Jeroen Tromp
+!    Seismological Laboratory - California Institute of Technology
+!         (c) California Institute of Technology July 2004
+!
+! This program is free software; you can redistribute it and/or modify
+! it under the terms of the GNU General Public License as published by
+! the Free Software Foundation; either version 2 of the License, or
+! (at your option) any later version.
+!
+! This program is distributed in the hope that it will be useful,
+! but WITHOUT ANY WARRANTY; without even the implied warranty of
+! MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+! GNU General Public License for more details.
+!
+! You should have received a copy of the GNU General Public License along
+! with this program; if not, write to the Free Software Foundation, Inc.,
+! 51 Franklin Street, Fifth Floor, Boston, MA 02110-1301 USA.
+!
+!=====================================================================
+
+  double precision, parameter :: PI = 3.141592653589793d0
+
+! small perturbation in distance (degrees)
+! (0.0025 degrees corresponds to roughly 250 meters)
+  double precision, parameter :: DDELTA = 0.0025
+
+! small perturbation in depth (km)
+  double precision, parameter :: DDEPTH = 1.0
+
+! basic scalar moment
+  double precision, parameter :: MOMENT = 1.0e+22
+
+! maximum record length
+  integer, parameter :: NDATAMAX = 25000
+
+! maximum number of records
+  integer, parameter :: NRECORDSMAX = 250
+
+! number of inversion parameters:
+! 6 for moment tensor only,
+! 7 for moment tensor plus centroid time,
+! 9 for moment tensor plus location,
+! 10 for moment tensor plus location and centroid time,
+  integer, parameter :: NPAR = 7
+
+! weigh the data in the inversion using a NOISE_WINDOW before P
+  logical, parameter :: WEIGHTING = .false.
+
+! length of noise window before P in seconds if WEIGHTING is set to true
+  double precision, parameter :: NOISE_WINDOW = 10.
+
+! select data
+  logical, parameter :: SELECT_DATA = .false.
+
+! show the fit to the data before inversion (after offset removal)
+  logical, parameter :: GRAPHICS = .true.
+
[…]

### 2010-10-21 utils/obsolete_routines_from_SPECFEM3D_v1.4.4/get_MPI_cutplanes_xi.f90

@@ -1,0 +1,178 @@
+!=====================================================================
+!
+!               S p e c f e m 3 D  V e r s i o n  1 . 4
+!               ---------------------------------------
+!
+!                 Dimitri Komatitsch and Jeroen Tromp
+!    Seismological Laboratory - California Institute of Technology
+!         (c) California Institute of Technology September 2006
+!
+! This program is free software; you can redistribute it and/or modify
+! it under the terms of the GNU General Public License as published by
+! the Free Software Foundation; either version 2 of the License, or
+! (at your option) any later version.
+!
+! This program is distributed in the hope that it will be useful,
+! but WITHOUT ANY WARRANTY; without even the implied warranty of
+! MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+! GNU General Public License for more details.
+!
+! You should have received a copy of the GNU General Public License along
+! with this program; if not, write to the Free Software Foundation, Inc.,
+! 51 Franklin Street, Fifth Floor, Boston, MA 02110-1301 USA.
+!
+!=====================================================================
+
+  subroutine get_MPI_cutplanes_xi(myrank,prname,nspec,iMPIcut_xi,ibool, &
+                        xstore,ystore,zstore,mask_ibool,npointot, &
+                        NSPEC2D_A_ETA,NSPEC2D_B_ETA)
+
+! this routine detects cut planes along xi
+! In principle the left cut plane of the first slice
+! and the right cut plane of the last slice are not used
+! in the solver except if we want to have periodic conditions
+
+  implicit none
+
+  include "constants.h"
+
+  integer nspec,myrank
+  integer NSPEC2D_A_ETA,NSPEC2D_B_ETA
+
+  logical iMPIcut_xi(2,nspec)
+
+  integer ibool(NGLLX,NGLLY,NGLLZ,nspec)
+
+  double precision xstore(NGLLX,NGLLY,NGLLZ,nspec)
+  double precision ystore(NGLLX,NGLLY,NGLLZ,nspec)
+  double precision zstore(NGLLX,NGLLY,NGLLZ,nspec)
+
+! logical mask used to create arrays iboolleft_xi and iboolright_xi
[…]

### 2010-10-21 utils/oldstuff/convert_finite_sources/convert_ASCII_tsurf_to_CMTSOLUTION.f90

@@ -1,0 +1,940 @@
+
+! convert ASCII data based on tsurf to set of CMTSOLUTION files
+
+  program convert_tsurf_to_CMTSOLUTION
+
+  implicit none
+
+  include "constants.h"
+
+! constants provided by the user
+
+! average slip length at each center (in meters)
+!! DK DK mean value for SAF 1857 from Sieh (1978) is 4.90 m
+  double precision, parameter :: AVERAGE_SLIP_LENGTH = 4.90d0
+
+! fictitious constant value of mu = 1
+! varies along the fault plane, therefore will be added later in the solver
+  double precision, parameter :: mu = 1.d0
+
+! rupture velocity in m/s
+  double precision, parameter :: rupture_velocity = 2800.d0
+
+! depth above which we suppress source elements that reach the surface
+  double precision, parameter :: DEPTH_REMOVED_TOPO = 20.d0
+
+! length of normal vectors for DX display
+  double precision, parameter :: length_normal_display_DX = +20000.d0
+
+ integer npoin    ! number of VRTX in GoCad file
+ integer nspec    ! number of TRGL in GoCad file
+
+ integer, dimension(:), allocatable :: iglob1,iglob2,iglob3
+ integer, dimension(:), allocatable :: iglob1_copy,iglob2_copy,iglob3_copy
+ double precision, dimension(:), allocatable :: x,y,z
+
+ double precision horiz_dist_fault,time_shift,time_shift_min,time_shift_max
+
+ integer ipoin,ispec,isource,NSOURCES,isourceshiftmin,isource_current
+ integer iglob1_store,iglob2_store,iglob3_store,iglob_dummy
+
+ logical exclude_source
+
+ double precision x1,y1,z1,x2,y2,z2,x3,y3,z3,horizdistval,TOLERANCE
+ double precision area_current,area_min,area_max,area_sum
+ double precision nx,ny,nz,norm
+ double precision x_center_triangle,y_center_triangle,z_center_triangle
+ double precision Mxx,Myy,Mzz,Mxy,Mxz,Myz,long,lat
+ double precision moment_tensor(6)
+
+! fault edges
+ double precision x_begin,x_end,y_begin,y_end
[…]

### 2010-10-21 utils/seis_process/process_trinet_data.pl

@@ -1,0 +1,225 @@
+#!/usr/bin/perl
+
+use Time::Local;
+use Math::Trig;
+use Getopt::Std;
+use POSIX;
+use lib '/opt/seismo-util/lib/perl/';
+use CMT_TOOLS;
+
+sub Usage{
+
+  print STDERR <<END;
+
+  process_trinet_data_new.pl  -m CMTFILE -l Start/End -t Ts/Tl 
+           -[i|I] Dir -p -x Ext -d OutDir -s Sps -P n
+           -c -y t1/s1/...  -v num_of_zeros names_of_files
+  with
+
+       -m -- Using CMTSOLUTION file to set event info
+       -l Start/End of trace to be cut
+       -t Ts/Tl -- shortest and longest period
+       -[i|I] Dir-- transfer recordings to displacement with pz files in Dir
+        (use none to set to defaulted Jascha''s dir)
+        (use -I instead to only perform transfer no bp)
+       -p -- pick event p and s arrival in sac (use socal model)
+       -x Ext -- extension of new file name
+       -d Dir -- directory to put the output files
+       -s Sps -- sampling rate per second. (default 20)
+       -P n -- number of poles for the butterworth filter (default 4)
+       -c  -- check sac output on the screen
+       -y t0/slope/t0/slope... (t3/t4/...)(for SC, try -20/0.15/11-/0.2)
+       -v # -- convert the data to either velocity or displacement by specifying
+               the number of zeros [default no change from pz files]
+
+       name_of_files --- name of sac files to be processed
+
+Notice: 1) main difference from process_data_new.pl is to specify 
+        the pz files directory and the default sampling rate is for BH component.
+        2) We always use the polezeros files corresponding to the correct
+        station, component in the pz directory. Polezero files are in the
+        format of StationName_CompName_StartTime_EndTime and recommended
+        pz directory is /opt/seismo-util/data/TriNet_PZs/, in other
[…]

### 2010-10-23 doc/USER_MANUAL/manual_SPECFEM3D.tex
added a DG reference and a comment about discontinuous meshes to the manual

@@
 based upon the spectral-element method (SEM).
-The SEM is a continuous Galerkin technique, which can easily be made discontinous \citep{BeMaPa94,KoWoHu02,ChCaVi03,Kop06}; it is then a particular case of the discontinuous Galerkin technique \citep{ReHi73,FaRi99,HuHuRa99,CoKaSh00,GiHeWa02,RiWh03,MoRi05,GrScSc06,BeLaPi06,DuKa06,DeSeWh08,WiStBuGh10,DeSe10}, with optimized efficiency because of its tensorized basis functions.
+The SEM is a continuous Galerkin technique, which can easily be made discontinous \citep{BeMaPa94,KoWoHu02,ChCaVi03,Kop06}; it is then a particular case of the discontinuous Galerkin technique \citep{ReHi73,FaRi99,HuHuRa99,CoKaSh00,GiHeWa02,RiWh03,MoRi05,GrScSc06,BeLaPi06,DuKa06,DeSeWh08,WiStBuGh10,DeSe10,EtChViGl10}, with optimized efficiency because of its tensorized basis functions.
+Note that in many (most?) geological models in the context of seismic wave propagation studies
+a discontinous mesh is not needed because material property contrasts are not drastic and thus a continuous formulation is sufficient.
 Effects due to lateral variations in compressional-wave speed, shear-wave

### 2010-11-04 doc/USER_MANUAL/manual_SPECFEM3D.tex
added the source code of SCOTCH; updated the manual (history, citations, SCOTCH).

@@
 Note that in many (most?) geological models in the context of seismic wave propagation studies
+(except for fault dynamic rupture studies)
 a discontinous mesh is not needed because material property contrasts are not drastic and thus a continuous formulation is sufficient.
@@
 seismic wave propagation, please consult \citet{KoVi98,KoTr99,ChKoViCaVaFe07,TrKoLi08} and
-in particular \citet{KoLiTrSuStSh04}. If you use the 3D southern
-California model, please cite \citet{SuSh03} (LA), \citet{lovelyetal06}
-(Salton Trough), and \citet{hauksson2000} (southern California).
-The Moho map was determined by \citet{zhu&kanamori2000}. The 1D SoCal
-model was developed by \citet{DrHe90}. The package can accommodate
+in particular \citet{KoLiTrSuStSh04}. The package can accommodate
 full 21-parameter anisotropy (see~\citet{ChTr07}) as well as lateral
@@
 kernel simulations are included \citep{LiTr06,TrKoLi08}.
+
+The SPECFEM3D package was first developed by Dimitri Komatitsch and Jeroen Tromp at Harvard University, then Caltech and Princeton, USA, starting in 1998, based in part on earlier work by Dimitri Komatitsch and Jean-Pierre Vilotte at IPG in Paris, France from 1994 to 1997.
+Since then it has been developed and maintained by a development team: in alphabetical order, C\'eline Blitz, Emiljana Jorgji, Dimitri Komatitsch, Nicolas Le Goff, Pieyre Le Loher, Roland Martin, David Mich\'ea, Jeroen Tromp \toall{add full list of developers here}.
 
@@
 It was a finalist again in 2008 for a run at 0.16 petaflops (sustained) on 149,784 processors of the `Jaguar' Cray XT5 system at Oak Ridge National Laboratories (USA) \citep{CaKoLaTiMiLeSnTr08}.
+It also won the BULL Joseph Fourier supercomputing award in 2010.
 
@@
 
+If you use the 3D southern
+California model, please cite \citet{SuSh03} (Los Angeles model), \citet{lovelyetal06}
+(Salton Trough), and \citet{hauksson2000} (southern California).
+The Moho map was determined by \citet{zhu&kanamori2000}. The 1D SoCal
+model was developed by \citet{DrHe90}.
+
 \section{Support}
@@
 Reintegration Grant No. MIRG-CT-2005-017461.
-Older versions of the code were initially developed by Dimitri Komatitsch at Institut de Physique du Globe (France)
-and then by Dimitri Komatitsch and Jeroen Tromp at Harvard University (USA).
 Any opinions, findings,
@@
 resulting from the placement of adjacent elements to different processors. \texttt{decompose\_mesh\_SCOTCH} depends on the Scotch library
-which provides efficent static mapping, graph and mesh partitioning routines. Scotch is a free software package downloadable from the web page \texttt{https://gforge.inria.fr/projects/scotch/}.
+which provides efficent static mapping, graph and mesh partitioning routines. Scotch is a free software package developed by
+Fran\c {c}ois Pellegrini et al.
+from LaBRI and INRIA in Bordeaux, France, downloadable from the web page \texttt{https://gforge.inria.fr/projects/scotch/}.
+It is more recent than METIS, actively maintained and performs better in many cases.
 \\
@@
 Prior to compiling \texttt{decompose\_mesh\_SCOTCH}, make sure you have correctly specified the path
-of the Scotch library with the option \texttt{-{}-with-scotch-lib} of the \texttt{configure} script. Then you are ready to compile, 
+of the Scotch library with the option \texttt{-{}-with-scotch-lib} of the \texttt{configure} script.
+For convenience we provide the source code of SCOTCH, which is released open source under the French CeCILL-C version 1 license,
+in directory src/decompose\_mesh\_SCOTCH/scotch\_5.1.10b. Please refer to file INSTALL.txt in that directory to see how to compile it.
+In the future you should be able to find for recent versions at http://www.labri.fr/perso/pelegrin/scotch/scotch\_en.html .
+\\
+
+Then you are ready to compile,
 in the directory \texttt{decompose\_mesh\_SCOTCH} type '\texttt{make decompose\_mesh\_SCOTCH}'.
@@
 
-The partitioning is done in serial, 
-the synopsis is :
+The partitioning is done in serial for now (in the next release we will provide a parallel version of that code),
+the synopsis is:
 \begin{lyxcode}

### 2010-11-04 src/decompose_mesh_SCOTCH/scotch_5.1.10b/src/libscotch/arch_mesh.h

@@ -1,0 +1,133 @@
+/* Copyright 2004,2007,2008 ENSEIRB, INRIA & CNRS
+**
+** This file is part of the Scotch software package for static mapping,
+** graph partitioning and sparse matrix ordering.
+**
+** This software is governed by the CeCILL-C license under French law
+** and abiding by the rules of distribution of free software. You can
+** use, modify and/or redistribute the software under the terms of the
+** CeCILL-C license as circulated by CEA, CNRS and INRIA at the following
+** URL: "http://www.cecill.info".
+** 
+** As a counterpart to the access to the source code and rights to copy,
+** modify and redistribute granted by the license, users are provided
+** only with a limited warranty and the software's author, the holder of
+** the economic rights, and the successive licensors have only limited
+** liability.
+** 
+** In this respect, the user's attention is drawn to the risks associated
+** with loading, using, modifying and/or developing or reproducing the
+** software by the user in light of its specific status of free software,
+** that may mean that it is complicated to manipulate, and that also
+** therefore means that it is reserved for developers and experienced
+** professionals having in-depth computer knowledge. Users are therefore
+** encouraged to load and test the software's suitability as regards
+** their requirements in conditions enabling the security of their
+** systems and/or data to be ensured and, more generally, to use and
+** operate it in the same conditions as regards security.
+** 
+** The fact that you are presently reading this means that you have had
+** knowledge of the CeCILL-C license and that you accept its terms.
+*/
+/************************************************************/
+/**                                                        **/
+/**   NAME       : arch_mesh.h                             **/
+/**                                                        **/
+/**   AUTHOR     : Francois PELLEGRINI                     **/
+/**                                                        **/
+/**   FUNCTION   : These lines are the data declaration    **/
+/**                for the mesh graph target architecture  **/
+/**                functions.                              **/
+/**                                                        **/
+/**   DATES      : # Version 0.0  : from : 01 dec 1992     **/
+/**                                 to   : 24 mar 1993     **/
[…]

### 2010-11-04 src/decompose_mesh_SCOTCH/scotch_5.1.10b/src/libscotch/arch_vcmplt.h

@@ -1,0 +1,97 @@
+/* Copyright 2004,2007,2008 ENSEIRB, INRIA & CNRS
+**
+** This file is part of the Scotch software package for static mapping,
+** graph partitioning and sparse matrix ordering.
+**
+** This software is governed by the CeCILL-C license under French law
+** and abiding by the rules of distribution of free software. You can
+** use, modify and/or redistribute the software under the terms of the
+** CeCILL-C license as circulated by CEA, CNRS and INRIA at the following
+** URL: "http://www.cecill.info".
+** 
+** As a counterpart to the access to the source code and rights to copy,
+** modify and redistribute granted by the license, users are provided
+** only with a limited warranty and the software's author, the holder of
+** the economic rights, and the successive licensors have only limited
+** liability.
+** 
+** In this respect, the user's attention is drawn to the risks associated
+** with loading, using, modifying and/or developing or reproducing the
+** software by the user in light of its specific status of free software,
+** that may mean that it is complicated to manipulate, and that also
+** therefore means that it is reserved for developers and experienced
+** professionals having in-depth computer knowledge. Users are therefore
+** encouraged to load and test the software's suitability as regards
+** their requirements in conditions enabling the security of their
+** systems and/or data to be ensured and, more generally, to use and
+** operate it in the same conditions as regards security.
+** 
+** The fact that you are presently reading this means that you have had
+** knowledge of the CeCILL-C license and that you accept its terms.
+*/
+/************************************************************/
+/**                                                        **/
+/**   NAME       : arch_vcmplt.h                           **/
+/**                                                        **/
+/**   AUTHOR     : Francois PELLEGRINI                     **/
+/**                                                        **/
+/**   FUNCTION   : These lines are the data declaration    **/
+/**                for the variable-sized complete graph   **/
+/**                target architecture functions.          **/
+/**                                                        **/
+/**   DATES      : # Version 3.0  : from : 01 jul 1995     **/
+/**                                 to     09 aug 1995     **/
+/**                # Version 3.1  : from : 20 jul 1996     **/
+/**                                 to     20 jul 1996     **/
[…]

### 2010-11-04 src/decompose_mesh_SCOTCH/scotch_5.1.10b/src/libscotch/dgraph_coarsen_edge.c

@@ -1,0 +1,147 @@
+/* Copyright 2007,2008 ENSEIRB, INRIA & CNRS
+**
+** This file is part of the Scotch software package for static mapping,
+** graph partitioning and sparse matrix ordering.
+**
+** This software is governed by the CeCILL-C license under French law
+** and abiding by the rules of distribution of free software. You can
+** use, modify and/or redistribute the software under the terms of the
+** CeCILL-C license as circulated by CEA, CNRS and INRIA at the following
+** URL: "http://www.cecill.info".
+** 
+** As a counterpart to the access to the source code and rights to copy,
+** modify and redistribute granted by the license, users are provided
+** only with a limited warranty and the software's author, the holder of
+** the economic rights, and the successive licensors have only limited
+** liability.
+** 
+** In this respect, the user's attention is drawn to the risks associated
+** with loading, using, modifying and/or developing or reproducing the
+** software by the user in light of its specific status of free software,
+** that may mean that it is complicated to manipulate, and that also
+** therefore means that it is reserved for developers and experienced
+** professionals having in-depth computer knowledge. Users are therefore
+** encouraged to load and test the software's suitability as regards
+** their requirements in conditions enabling the security of their
+** systems and/or data to be ensured and, more generally, to use and
+** operate it in the same conditions as regards security.
+** 
+** The fact that you are presently reading this means that you have had
+** knowledge of the CeCILL-C license and that you accept its terms.
+*/
+/************************************************************/
+/**                                                        **/
+/**   NAME       : dgraph_coarsen_edge.c                   **/
+/**                                                        **/
+/**   AUTHOR     : Francois PELLEGRINI                     **/
+/**                                                        **/
+/**   FUNCTION   : This commodity file contains the edge   **/
+/**                arrays building subroutine which is     **/
+/**                duplicated, with minor modifications,   **/
+/**                into dgraph_coarsen.c.                  **/
+/**                                                        **/
+/**   DATES      : # Version 5.2  : from : 11 dec 2008     **/
[…]

### 2010-11-04 src/decompose_mesh_SCOTCH/scotch_5.1.10b/src/libscotch/graph_induce.c

@@ -1,0 +1,368 @@
+/* Copyright 2004,2007-2009 ENSEIRB, INRIA & CNRS
+**
+** This file is part of the Scotch software package for static mapping,
+** graph partitioning and sparse matrix ordering.
+**
+** This software is governed by the CeCILL-C license under French law
+** and abiding by the rules of distribution of free software. You can
+** use, modify and/or redistribute the software under the terms of the
+** CeCILL-C license as circulated by CEA, CNRS and INRIA at the following
+** URL: "http://www.cecill.info".
+** 
+** As a counterpart to the access to the source code and rights to copy,
+** modify and redistribute granted by the license, users are provided
+** only with a limited warranty and the software's author, the holder of
+** the economic rights, and the successive licensors have only limited
+** liability.
+** 
+** In this respect, the user's attention is drawn to the risks associated
+** with loading, using, modifying and/or developing or reproducing the
+** software by the user in light of its specific status of free software,
+** that may mean that it is complicated to manipulate, and that also
+** therefore means that it is reserved for developers and experienced
+** professionals having in-depth computer knowledge. Users are therefore
+** encouraged to load and test the software's suitability as regards
+** their requirements in conditions enabling the security of their
+** systems and/or data to be ensured and, more generally, to use and
+** operate it in the same conditions as regards security.
+** 
+** The fact that you are presently reading this means that you have had
+** knowledge of the CeCILL-C license and that you accept its terms.
+*/
+/************************************************************/
+/**                                                        **/
+/**   NAME       : graph_induce.c                          **/
+/**                                                        **/
+/**   AUTHOR     : Francois PELLEGRINI                     **/
+/**                                                        **/
+/**   FUNCTION   : This module handles the source graph    **/
+/**                subgraph-making functions.              **/
+/**                                                        **/
+/**   DATES      : # Version 0.0  : from : 01 dec 1992     **/
[…]

### 2010-11-04 src/decompose_mesh_SCOTCH/scotch_5.1.10b/src/libscotch/library_order.h

@@ -1,0 +1,62 @@
+/* Copyright 2004,2007 ENSEIRB, INRIA & CNRS
+**
+** This file is part of the Scotch software package for static mapping,
+** graph partitioning and sparse matrix ordering.
+**
+** This software is governed by the CeCILL-C license under French law
+** and abiding by the rules of distribution of free software. You can
+** use, modify and/or redistribute the software under the terms of the
+** CeCILL-C license as circulated by CEA, CNRS and INRIA at the following
+** URL: "http://www.cecill.info".
+** 
+** As a counterpart to the access to the source code and rights to copy,
+** modify and redistribute granted by the license, users are provided
+** only with a limited warranty and the software's author, the holder of
+** the economic rights, and the successive licensors have only limited
+** liability.
+** 
+** In this respect, the user's attention is drawn to the risks associated
+** with loading, using, modifying and/or developing or reproducing the
+** software by the user in light of its specific status of free software,
+** that may mean that it is complicated to manipulate, and that also
+** therefore means that it is reserved for developers and experienced
+** professionals having in-depth computer knowledge. Users are therefore
+** encouraged to load and test the software's suitability as regards
+** their requirements in conditions enabling the security of their
+** systems and/or data to be ensured and, more generally, to use and
+** operate it in the same conditions as regards security.
+** 
+** The fact that you are presently reading this means that you have had
+** knowledge of the CeCILL-C license and that you accept its terms.
+*/
+/************************************************************/
+/**                                                        **/
+/**   NAME       : library_order.h                         **/
+/**                                                        **/
+/**   AUTHOR     : Francois PELLEGRINI                     **/
+/**                                                        **/
+/**   FUNCTION   : This module contains the data declara-  **/
+/**                tions for the library ordering          **/
+/**                structure.                              **/
+/**                                                        **/
+/**   DATES      : # Version 3.3  : from : 08 oct 1998     **/
+/**                                 to   : 08 oct 1998     **/
+/**                # Version 4.0  : from : 11 dec 2001     **/
+/**                                 to     25 dec 2004     **/
+/**                                                        **/
+/************************************************************/
+
+/*
+**  The type and structure definitions.
+*/
+
+/*+ Ordering. +*/
+
+typedef struct LibOrder_ {
+  Order                     o;                    /*+ Ordering data                      +*/
+  Gnum * restrict           permtab;              /*+ Direct permutation array           +*/
[…]

### 2010-11-04 src/decompose_mesh_SCOTCH/scotch_5.1.10b/src/libscotchmetis/metis_graph_order.c

@@ -1,0 +1,144 @@
+/* Copyright 2007,2008,2010 ENSEIRB, INRIA & CNRS
+**
+** This file is part of the Scotch software package for static mapping,
+** graph partitioning and sparse matrix ordering.
+**
+** This software is governed by the CeCILL-C license under French law
+** and abiding by the rules of distribution of free software. You can
+** use, modify and/or redistribute the software under the terms of the
+** CeCILL-C license as circulated by CEA, CNRS and INRIA at the following
+** URL: "http://www.cecill.info".
+** 
+** As a counterpart to the access to the source code and rights to copy,
+** modify and redistribute granted by the license, users are provided
+** only with a limited warranty and the software's author, the holder of
+** the economic rights, and the successive licensors have only limited
+** liability.
+** 
+** In this respect, the user's attention is drawn to the risks associated
+** with loading, using, modifying and/or developing or reproducing the
+** software by the user in light of its specific status of free software,
+** that may mean that it is complicated to manipulate, and that also
+** therefore means that it is reserved for developers and experienced
+** professionals having in-depth computer knowledge. Users are therefore
+** encouraged to load and test the software's suitability as regards
+** their requirements in conditions enabling the security of their
+** systems and/or data to be ensured and, more generally, to use and
+** operate it in the same conditions as regards security.
+** 
+** The fact that you are presently reading this means that you have had
+** knowledge of the CeCILL-C license and that you accept its terms.
+*/
+/************************************************************/
+/**                                                        **/
+/**   NAME       : metis_graph_order.c                     **/
+/**                                                        **/
+/**   AUTHOR     : Francois PELLEGRINI                     **/
+/**                                                        **/
+/**   FUNCTION   : This module is the compatibility        **/
+/**                library for the MeTiS ordering          **/
+/**                routines.                               **/
+/**                                                        **/
+/**   DATES      : # Version 5.0  : from : 08 sep 2006     **/
+/**                                 to     07 jun 2007     **/
+/**                # Version 5.1  : from : 30 jun 2010     **/
[…]

### 2010-11-04 src/decompose_mesh_SCOTCH/scotch_5.1.10b/src/libscotch/bgraph_bipart_bd.c
added the source code of SCOTCH; updated the manual (history, citations, SCOTCH).

@@ -1,0 +1,625 @@
+/* Copyright 2004,2007,2008,2010 ENSEIRB, INRIA & CNRS
+**
+** This file is part of the Scotch software package for static mapping,
+** graph partitioning and sparse matrix ordering.
+**
+** This software is governed by the CeCILL-C license under French law
+** and abiding by the rules of distribution of free software. You can
+** use, modify and/or redistribute the software under the terms of the
+** CeCILL-C license as circulated by CEA, CNRS and INRIA at the following
+** URL: "http://www.cecill.info".
+** 
+** As a counterpart to the access to the source code and rights to copy,
+** modify and redistribute granted by the license, users are provided
+** only with a limited warranty and the software's author, the holder of
+** the economic rights, and the successive licensors have only limited
+** liability.
+** 
+** In this respect, the user's attention is drawn to the risks associated
+** with loading, using, modifying and/or developing or reproducing the
+** software by the user in light of its specific status of free software,
+** that may mean that it is complicated to manipulate, and that also
+** therefore means that it is reserved for developers and experienced
+** professionals having in-depth computer knowledge. Users are therefore
+** encouraged to load and test the software's suitability as regards
+** their requirements in conditions enabling the security of their
+** systems and/or data to be ensured and, more generally, to use and
+** operate it in the same conditions as regards security.
+** 
+** The fact that you are presently reading this means that you have had
+** knowledge of the CeCILL-C license and that you accept its terms.
+*/
+/************************************************************/
+/**                                                        **/
+/**   NAME       : bgraph_bipart_bd.c                      **/
+/**                                                        **/
+/**   AUTHOR     : Francois PELLEGRINI                     **/
+/**                                                        **/
+/**   FUNCTION   : This module builds a band graph around  **/
+/**                the frontier in order to decrease       **/
+/**                problem size for the strategy to be     **/
+/**                applied.                                **/
+/**                                                        **/
[…]

### 2010-11-04 src/decompose_mesh_SCOTCH/scotch_5.1.10b/src/libscotch/dgraph_band.c

@@ -1,0 +1,956 @@
+/* Copyright 2007-2010 ENSEIRB, INRIA & CNRS
+**
+** This file is part of the Scotch software package for static mapping,
+** graph partitioning and sparse matrix ordering.
+**
+** This software is governed by the CeCILL-C license under French law
+** and abiding by the rules of distribution of free software. You can
+** use, modify and/or redistribute the software under the terms of the
+** CeCILL-C license as circulated by CEA, CNRS and INRIA at the following
+** URL: "http://www.cecill.info".
+** 
+** As a counterpart to the access to the source code and rights to copy,
+** modify and redistribute granted by the license, users are provided
+** only with a limited warranty and the software's author, the holder of
+** the economic rights, and the successive licensors have only limited
+** liability.
+** 
+** In this respect, the user's attention is drawn to the risks associated
+** with loading, using, modifying and/or developing or reproducing the
+** software by the user in light of its specific status of free software,
+** that may mean that it is complicated to manipulate, and that also
+** therefore means that it is reserved for developers and experienced
+** professionals having in-depth computer knowledge. Users are therefore
+** encouraged to load and test the software's suitability as regards
+** their requirements in conditions enabling the security of their
+** systems and/or data to be ensured and, more generally, to use and
+** operate it in the same conditions as regards security.
+** 
+** The fact that you are presently reading this means that you have had
+** knowledge of the CeCILL-C license and that you accept its terms.
+*/
+/************************************************************/
+/**                                                        **/
+/**   NAME       : dgraph_band.c                           **/
+/**                                                        **/
+/**   AUTHOR     : Francois PELLEGRINI                     **/
+/**                                                        **/
+/**   FUNCTION   : This module computes a distributed band **/
+/**                graph from the given frontier array.    **/
+/**                                                        **/
+/**   DATES      : # Version 5.1  : from : 11 nov 2007     **/
[…]

### 2010-11-04 src/decompose_mesh_SCOTCH/scotch_5.1.10b/src/libscotch/dmapping.c

@@ -1,0 +1,302 @@
+/* Copyright 2008 ENSEIRB, INRIA & CNRS
+**
+** This file is part of the Scotch software package for static mapping,
+** graph partitioning and sparse matrix ordering.
+**
+** This software is governed by the CeCILL-C license under French law
+** and abiding by the rules of distribution of free software. You can
+** use, modify and/or redistribute the software under the terms of the
+** CeCILL-C license as circulated by CEA, CNRS and INRIA at the following
+** URL: "http://www.cecill.info".
+** 
+** As a counterpart to the access to the source code and rights to copy,
+** modify and redistribute granted by the license, users are provided
+** only with a limited warranty and the software's author, the holder of
+** the economic rights, and the successive licensors have only limited
+** liability.
+** 
+** In this respect, the user's attention is drawn to the risks associated
+** with loading, using, modifying and/or developing or reproducing the
+** software by the user in light of its specific status of free software,
+** that may mean that it is complicated to manipulate, and that also
+** therefore means that it is reserved for developers and experienced
+** professionals having in-depth computer knowledge. Users are therefore
+** encouraged to load and test the software's suitability as regards
+** their requirements in conditions enabling the security of their
+** systems and/or data to be ensured and, more generally, to use and
+** operate it in the same conditions as regards security.
+** 
+** The fact that you are presently reading this means that you have had
+** knowledge of the CeCILL-C license and that you accept its terms.
+*/
+/************************************************************/
+/**                                                        **/
+/**   NAME       : dmapping.c                              **/
+/**                                                        **/
+/**   AUTHOR     : Francois PELLEGRINI                     **/
[…]

### 2010-11-04 src/decompose_mesh_SCOTCH/scotch_5.1.10b/src/libscotch/vdgraph_separate_ml.c

@@ -1,0 +1,810 @@
+/* Copyright 2007-2010 ENSEIRB, INRIA & CNRS
+**
+** This file is part of the Scotch software package for static mapping,
+** graph partitioning and sparse matrix ordering.
+**
+** This software is governed by the CeCILL-C license under French law
+** and abiding by the rules of distribution of free software. You can
+** use, modify and/or redistribute the software under the terms of the
+** CeCILL-C license as circulated by CEA, CNRS and INRIA at the following
+** URL: "http://www.cecill.info".
+** 
+** As a counterpart to the access to the source code and rights to copy,
+** modify and redistribute granted by the license, users are provided
+** only with a limited warranty and the software's author, the holder of
+** the economic rights, and the successive licensors have only limited
+** liability.
+** 
+** In this respect, the user's attention is drawn to the risks associated
+** with loading, using, modifying and/or developing or reproducing the
+** software by the user in light of its specific status of free software,
+** that may mean that it is complicated to manipulate, and that also
+** therefore means that it is reserved for developers and experienced
+** professionals having in-depth computer knowledge. Users are therefore
+** encouraged to load and test the software's suitability as regards
+** their requirements in conditions enabling the security of their
+** systems and/or data to be ensured and, more generally, to use and
+** operate it in the same conditions as regards security.
+** 
+** The fact that you are presently reading this means that you have had
+** knowledge of the CeCILL-C license and that you accept its terms.
+*/
+/************************************************************/
+/**                                                        **/
+/**   NAME       : vdgraph_separate_ml.c                   **/
+/**                                                        **/
+/**   AUTHOR     : Francois PELLEGRINI                     **/
+/**                Cedric CHEVALIER (v5.0)                 **/
+/**                                                        **/
+/**   FUNCTION   : This module contains the multi-level    **/
+/**                separation strategy.                    **/
[…]

### 2010-11-04 src/decompose_mesh_SCOTCH/scotch_5.1.10b/src/libscotchmetis/metis_graph_order_f.c

@@ -1,0 +1,123 @@
+/* Copyright 2007,2010 ENSEIRB, INRIA & CNRS
+**
+** This file is part of the Scotch software package for static mapping,
+** graph partitioning and sparse matrix ordering.
+**
+** This software is governed by the CeCILL-C license under French law
+** and abiding by the rules of distribution of free software. You can
+** use, modify and/or redistribute the software under the terms of the
+** CeCILL-C license as circulated by CEA, CNRS and INRIA at the following
+** URL: "http://www.cecill.info".
+** 
+** As a counterpart to the access to the source code and rights to copy,
+** modify and redistribute granted by the license, users are provided
+** only with a limited warranty and the software's author, the holder of
+** the economic rights, and the successive licensors have only limited
+** liability.
+** 
+** In this respect, the user's attention is drawn to the risks associated
+** with loading, using, modifying and/or developing or reproducing the
+** software by the user in light of its specific status of free software,
+** that may mean that it is complicated to manipulate, and that also
+** therefore means that it is reserved for developers and experienced
+** professionals having in-depth computer knowledge. Users are therefore
+** encouraged to load and test the software's suitability as regards
+** their requirements in conditions enabling the security of their
+** systems and/or data to be ensured and, more generally, to use and
+** operate it in the same conditions as regards security.
+** 
+** The fact that you are presently reading this means that you have had
+** knowledge of the CeCILL-C license and that you accept its terms.
+*/
+/************************************************************/
+/**                                                        **/
+/**   NAME       : metis_graph_order_f.c                   **/
+/**                                                        **/
+/**   AUTHOR     : Francois PELLEGRINI                     **/
+/**                                                        **/
+/**   FUNCTION   : This file contains the Fortran API of   **/
+/**                the compatibility library for the       **/
+/**                MeTiS ordering routines.                **/
+/**                                                        **/
+/**   DATES      : # Version 5.0  : from : 10 sep 2006     **/
+/**                                 to     10 sep 2006     **/
+/**                # Version 5.1  : from : 30 jun 2010     **/
+/**                                 to     30 jun 2010     **/
+/**                                                        **/
+/************************************************************/
+
+/*
+**  The defines and includes.
+*/
+
+#define LIBRARY
+
+#include "common.h"
+#include "scotch.h"
+#include "metis.h"                                /* Our "metis.h" file */
+
+/**************************************/
[…]

### 2010-11-04 src/decompose_mesh_SCOTCH/scotch_5.1.10b/src/scotch/amk_p2.c

@@ -1,0 +1,150 @@
+/* Copyright 2004,2007,2008,2010 ENSEIRB, INRIA & CNRS
+**
+** This file is part of the Scotch software package for static mapping,
+** graph partitioning and sparse matrix ordering.
+**
+** This software is governed by the CeCILL-C license under French law
+** and abiding by the rules of distribution of free software. You can
+** use, modify and/or redistribute the software under the terms of the
+** CeCILL-C license as circulated by CEA, CNRS and INRIA at the following
+** URL: "http://www.cecill.info".
+** 
+** As a counterpart to the access to the source code and rights to copy,
+** modify and redistribute granted by the license, users are provided
+** only with a limited warranty and the software's author, the holder of
+** the economic rights, and the successive licensors have only limited
+** liability.
+** 
+** In this respect, the user's attention is drawn to the risks associated
+** with loading, using, modifying and/or developing or reproducing the
+** software by the user in light of its specific status of free software,
+** that may mean that it is complicated to manipulate, and that also
+** therefore means that it is reserved for developers and experienced
+** professionals having in-depth computer knowledge. Users are therefore
+** encouraged to load and test the software's suitability as regards
+** their requirements in conditions enabling the security of their
+** systems and/or data to be ensured and, more generally, to use and
+** operate it in the same conditions as regards security.
+** 
+** The fact that you are presently reading this means that you have had
+** knowledge of the CeCILL-C license and that you accept its terms.
+*/
+/************************************************************/
+/**                                                        **/
+/**   NAME       : amk_p2.c                                **/
+/**                                                        **/
+/**   AUTHOR     : Francois PELLEGRINI                     **/
+/**                                                        **/
+/**   FUNCTION   : Creates the target architecture file    **/
+/**                for a weighted path with two vertices   **/
+/**                used to bipartition graphs in parts of  **/
+/**                different sizes.                        **/
+/**                                                        **/
+/**   DATES      : # Version 3.0  : from : 17 jul 1995     **/
+/**                                 to   : 17 jul 1995     **/
+/**                # Version 3.2  : from : 02 jun 1997     **/
[…]

### 2010-11-04 doc/USER_MANUAL/manual_SPECFEM3D.tex
added comments about other meshers: GiD, Gmsh, and TetGen.

@@
 The SPECFEM3D software package relies on the Scotch library to partition the meshes created with CUBIT.
-Because it is not standard, the path of this library needs to be set with the option \texttt{-{}-with-scotch-lib}.
-Just as an example:
+Note that we use CUBIT to create meshes of hexahedra, but other packages can be used as well, for instance GiD from http://gid.cimne.upc.es, Gmsh from http://geuz.org/gmsh or even TetGen from http://tetgen.berlios.de (TetGen generates tetrahedra, but each of them can easily be decomposed into four hexahedra as shown in the picture of the TetGen logo at http://tetgen.berlios.de/figs/Delaunay-Voronoi-3D.gif ).
+
+The path to the SCOTCH library needs to be set with the option \texttt{-{}-with-scotch-lib}. Just as an example:
 \begin{lyxcode}

### 2010-11-06 doc/USER_MANUAL/manual_SPECFEM3D.tex
committed a new version of param_reader.c by Dennis McRitchie that works on Mac OS. Also updated the citation section of the manual.

@@
 (except for fault dynamic rupture studies)
-a discontinous mesh is not needed because material property contrasts are not drastic and thus a continuous formulation is sufficient.
+a discontinous mesh is not needed because material property contrasts are not drastic and thus a continuous formulation is sufficient. For a detailed introduction to the SEM as applied to regional
+seismic wave propagation, please consult \citet{KoVi98,KoTr99,ChKoViCaVaFe07,TrKoLi08} and
+in particular \citet{KoLiTrSuStSh04}.
+
 Effects due to lateral variations in compressional-wave speed, shear-wave
 speed, density, a 3D crustal model, topography and bathymetry are
-included. For a detailed introduction to the SEM as applied to regional
-seismic wave propagation, please consult \citet{KoVi98,KoTr99,ChKoViCaVaFe07,TrKoLi08} and
-in particular \citet{KoLiTrSuStSh04}. The package can accommodate
+included. The package can accommodate
 full 21-parameter anisotropy (see~\citet{ChTr07}) as well as lateral
@@
 SPECFEM3D won the Gordon Bell award for best performance at the SuperComputing~2003
-conference in Phoenix, Arizona (USA) by running at 5 teraflops (sustained)
-on 1944 processors of the Japanese Earth Simulator using 14.6 billion
-degrees of freedom stored in 2.5 terabytes of memory; see \cite{KoTsChTr03}
-and the Gordon Bell Awards News Release \url{www.sc-conference.org/sc2003/nr_finalaward.html }
-for details.
+conference in Phoenix, Arizona (USA) (see \cite{KoTsChTr03}
+and \url{www.sc-conference.org/sc2003/nr_finalaward.html}).
 It was a finalist again in 2008 for a run at 0.16 petaflops (sustained) on 149,784 processors of the `Jaguar' Cray XT5 system at Oak Ridge National Laboratories (USA) \citep{CaKoLaTiMiLeSnTr08}.
@@
 as well as Convolutional or Auxiliary Differential Equation Perfectly Matched absorbing Layers (C-PML or ADE-PML) \citep{KoMa07,MaKoEz08,MaKoGe08,MaKo09,MaKoGeBr10}.
-The next release will use the PT-Scotch and ParMetis parallel libraries for mesh partitioning.
+The next release will use the PT-Scotch parallel library for mesh partitioning.
 

### 2010-11-07 doc/USER_MANUAL/manual_SPECFEM3D.tex
updated both manuals (added reference to Oliveira and Seriani, also to Fichtner et al.)

@@
 The SEM is a continuous Galerkin technique, which can easily be made discontinous \citep{BeMaPa94,KoWoHu02,ChCaVi03,Kop06}; it is then a particular case of the discontinuous Galerkin technique \citep{ReHi73,FaRi99,HuHuRa99,CoKaSh00,GiHeWa02,RiWh03,MoRi05,GrScSc06,BeLaPi06,DuKa06,DeSeWh08,WiStBuGh10,DeSe10,EtChViGl10}, with optimized efficiency because of its tensorized basis functions.
+In particular, it can accurately handle very distorted mesh elements \citep{OlSe11}.
 Note that in many (most?) geological models in the context of seismic wave propagation studies
@@
 variations in attenuation. Adjoint capabilities and finite-frequency
-kernel simulations are included \citep{LiTr06,TrKoLi08}.
+kernel simulations are included \citep{LiTr06,TrKoLi08,FiIgBuKe09}.
 
@@
 
-The next release of the code will include support for GPU graphics card acceleration \citep{KoMiEr09,KoErGoMi10,MiKo10}
+The next release of the code will include support for GPU graphics card acceleration \citep{KoMiEr09,KoErGoMi10,MiKo10,Kom11}
 as well as Convolutional or Auxiliary Differential Equation Perfectly Matched absorbing Layers (C-PML or ADE-PML) \citep{KoMa07,MaKoEz08,MaKoGe08,MaKo09,MaKoGeBr10}.
@@
 The SPECFEM3D software package relies on the Scotch library to partition the meshes created with CUBIT.
-Note that we use CUBIT to create meshes of hexahedra, but other packages can be used as well, for instance GiD from http://gid.cimne.upc.es, Gmsh from http://geuz.org/gmsh or even TetGen from http://tetgen.berlios.de (TetGen generates tetrahedra, but each of them can easily be decomposed into four hexahedra as shown in the picture of the TetGen logo at http://tetgen.berlios.de/figs/Delaunay-Voronoi-3D.gif ).
+Note that we use CUBIT to create meshes of hexahedra, but other packages can be used as well, for instance GiD from http://gid.cimne.upc.es or Gmsh from http://geuz.org/gmsh.
+Even mesh creation packages that generate tetrahedra, for instance TetGen from http://tetgen.berlios.de, can be used because each tetrahedron can then easily be decomposed into four hexahedra as shown in the picture of the TetGen logo at http://tetgen.berlios.de/figs/Delaunay-Voronoi-3D.gif; while this approach does not generate hexahedra of optimal quality, it can ease mesh creation in some situations and it has been shown that the spectral-element method can very accurately handle distorted mesh elements \citep{OlSe11}.
 

### 2010-11-08 doc/USER_MANUAL/manual_SPECFEM3D.tex
added a reference to Wilcox et al.

@@
 based upon the spectral-element method (SEM).
-The SEM is a continuous Galerkin technique, which can easily be made discontinous \citep{BeMaPa94,KoWoHu02,ChCaVi03,Kop06}; it is then a particular case of the discontinuous Galerkin technique \citep{ReHi73,FaRi99,HuHuRa99,CoKaSh00,GiHeWa02,RiWh03,MoRi05,GrScSc06,BeLaPi06,DuKa06,DeSeWh08,WiStBuGh10,DeSe10,EtChViGl10}, with optimized efficiency because of its tensorized basis functions.
+The SEM is a continuous Galerkin technique, which can easily be made discontinous \citep{BeMaPa94,KoWoHu02,ChCaVi03,Kop06,WiStBuGh10}; it is then a particular case of the discontinuous Galerkin technique \citep{ReHi73,FaRi99,HuHuRa99,CoKaSh00,GiHeWa02,RiWh03,MoRi05,GrScSc06,BeLaPi06,DuKa06,DeSeWh08,WiStBuGh10,DeSe10,EtChViGl10}, with optimized efficiency because of its tensorized basis functions.
 In particular, it can accurately handle very distorted mesh elements \citep{OlSe11}.

### 2010-11-09 doc/USER_MANUAL/manual_SPECFEM3D.tex
updated the SPECFEM3D manual

@@
 The SPECFEM3D software package relies on the Scotch library to partition the meshes created with CUBIT.
-Note that we use CUBIT to create meshes of hexahedra, but other packages can be used as well, for instance GiD from http://gid.cimne.upc.es or Gmsh from http://geuz.org/gmsh.
-Even mesh creation packages that generate tetrahedra, for instance TetGen from http://tetgen.berlios.de, can be used because each tetrahedron can then easily be decomposed into four hexahedra as shown in the picture of the TetGen logo at http://tetgen.berlios.de/figs/Delaunay-Voronoi-3D.gif; while this approach does not generate hexahedra of optimal quality, it can ease mesh creation in some situations and it has been shown that the spectral-element method can very accurately handle distorted mesh elements \citep{OlSe11}.
-
-The path to the SCOTCH library needs to be set with the option \texttt{-{}-with-scotch-lib}. Just as an example:
-\begin{lyxcode}
-./configure~FC=ifort~MPIFC=mpif90~-{}-with-scotch-lib=/opt/scotch/gcc64/lib
+Note that we use CUBIT to create meshes of hexahedra, but other packages can be used as well, for instance GiD from \url{http://gid.cimne.upc.es} or Gmsh from \url{http://geuz.org/gmsh}.
+Even mesh creation packages that generate tetrahedra, for instance TetGen from \url{http://tetgen.berlios.de}, can be used because each tetrahedron can then easily be decomposed into four hexahedra as shown in the picture of the TetGen logo at \url{http://tetgen.berlios.de/figs/Delaunay-Voronoi-3D.gif}; while this approach does not generate hexahedra of optimal quality, it can ease mesh creation in some situations and it has been shown that the spectral-element method can very accurately handle distorted mesh elements \citep{OlSe11}.
+
+The path to the SCOTCH installation needs to be set with the option \texttt{-{}-with-scotch-path}. Just as an example:
+\begin{lyxcode}
+./configure~FC=ifort~MPIFC=mpif90~-{}-with-scotch-path=/opt/scotch
 \end{lyxcode}
@@
 \end{description}
+The configuration scripts automatically creates for each executable a corresponding
+\texttt{Makefile} in the \texttt{src/} subdirectoy.
 The \texttt{Makefile} contains a number of suggested entries for various
@@
 
-Finally, before compiling make sure that the subdirectories \texttt{obj}
-and \texttt{OUTPUT\_FILES} exist within the directory with the source
-code (\texttt{SPECFEM3D\_V2.0.0}). The \texttt{go\_mesher} script
-discussed in Chapter~\ref{cha:Running-the-Mesher-Meshfem3D} automatically
-takes care of creating the \texttt{OUTPUT\_FILES} directory.
+% obsolete
+%Finally, before compiling make sure that the subdirectories \texttt{obj}
+%and \texttt{OUTPUT\_FILES} exist within the directory with the source
+%code (\texttt{SPECFEM3D\_V2.0.0}). The \texttt{go\_mesher} script
+%discussed in Chapter~\ref{cha:Running-the-Mesher-Meshfem3D} automatically
+%takes care of creating the \texttt{OUTPUT\_FILES} directory.
 
@@
 
+The first step in running a spectral-element method computation consists of constructing a high-quality mesh for the region under consideration.
+We provide two possibilities to do so: (1) relying on an external, hexahedral mesher CUBIT or (2) using the provided, internal mesher \texttt{xmeshfem3D}.
+In the following, we explain these two approaches.
+
 \section{\label{cha:Running-the-Mesher-CUBIT}Meshing with \texttt{CUBIT}}
-\subsection{Exporting the Mesh}
+
+CUBIT is a meshing tool suite for the creation of finite-element meshes for arbitrarily shaped models.
+It has been developed and maintained at Sandia National Laboratories and can be purchased for a small academic institutional fee at \url{http://cubit.sandia.gov}.
+Our experience showed that using CUBIT greatly facilitates and speeds up the generation and preparation of hexahedral, conforming meshes for a variety of geophysical models with increasing complexity.
+
+\begin{figure}[htbp]
+\noindent \begin{centering}
+\includegraphics[width=.6\textwidth]{figures/mount-cubit.jpg}
+\par\end{centering}
+
+\caption{Example of the graphical user interface of CUBIT. The hexahedral mesh shown in the main display consists of a hexahedral discretization of a single volume with topography.}
+
+\label{fig:mount.cubit}
+\end{figure}
+
+The basic steps in creating a load-balanced, partitioned mesh with CUBIT are: 
+\begin{description}
+\item [1.] setting up a hexahedral mesh with CUBIT, 
+\item [2.] exporting the CUBIT mesh into a specfem3d file format and 
+\item [3.] partitioning the specfem3d mesh files for a chosen number of cores. 
+\end{description}
+
+Examples are provided in the SPECFEM3D package in the subdirectory \texttt{examples/}.
+Please feel free to contribute your own example to this package by contacting the CIG Computational Seismology Mailing List  \url{cig-seismo@geodynamics.org}.
+
+\subsection{Creating the Mesh with CUBIT}
+
+For the installation and handling of the CUBIT meshing tool suite, please refer to the CUBIT user manual and documentation. 
+In order to give you a basic understanding of how to use CUBIT for our purposes, examples are provided in the SPECFEM3D package in the subdirectory \texttt{examples/}.
+For your convenience, the following examples are provided in the subdirectory \texttt{examples/}:
+\begin{description}
+\item [{\texttt{homogeneous\_halfspace}}] Basic example creates a single block model and assigns elastic material parameters.
+\item [{\texttt{layered\_halfspace}}]  Combines two different, elastic material volumes and creates a refinement layer between the two.
+This example can be compared for validation against the solutions provided in subdirectory ~\\
+\texttt{VALIDATION\_3D\_SEM\_SIMPLER\_LAYER\_SOURCE\_DEPTH/}.
+\item [{\texttt{waterlayered\_halfspace}}] Combines an acoustic and elastic material volume as in a schematic marine survey example.
+\item [{\texttt{tomographic\_model}}] Creates a single block model whose material properties will have to be read in from a tomographic model file during the databases creation by \texttt{xgenerate\_databases}.
+\end{description}
+
+\begin{figure}[htbp]
+\noindent \begin{centering}
+\includegraphics[width=.45\textwidth]{figures/example-homogeneous.jpg}
+\includegraphics[width=.45\textwidth]{figures/example-2layers.jpg} \\
+\includegraphics[width=.45\textwidth]{figures/example-water.jpg}
+\includegraphics[width=.45\textwidth]{figures/example-tomo.jpg}
+\par\end{centering}
+
+\caption{Screenshots of the CUBIT examples provided in subdirectory  \texttt{examples/}: homogeneous halfspace, layered halfspace, waterlayered halfspace and tomographic model.}
+
+\label{fig:examples.cubit}
+\end{figure}
+
+In each example subdirectory you will find a \texttt{README} file, which explains in a step-by-step tutorial the workflow for the example.
+Please feel free to contribute your own example to this package by contacting the CIG Computational Seismology Mailing List  \url{cig-seismo@geodynamics.org}.
+
+\subsection{Exporting the Mesh with \texttt{cubit2specfem3d.py} }
+
+Once the geometric model volumes in CUBIT are meshed, you prepare the model for the exportation with the definition of the material blocks and of the boundary surfaces.
+Thus, prior to exporting the mesh, you need to define blocks specifying the materials and absorbing boundaries in CUBIT.
+This process could be done automatically using the script \texttt{boundary\_definition.py} if the mesh meets some conditions or manually, following the block convention:
+
+\begin{description}
+\item [material\_name] Each material should have a specific block defined by a unique name.
+The name convention of the material is to start with either 'elastic' or 'acoustic'.
+It must be then followed by a unique identifier, e.g.
+'elastic 1', 'elastic 2', etc.
+The additional attributes to the block define the material description.
+
+For an elastic material:
+\begin{description}
+\item [material\_id] An integer value which is unique for this material. 
+\item [Vp] P-Velocity of the material (given in m/s).
+\item [Vs] S-Velocity of the material (given in m/s).
+\item [rho] density of the material (given in kg/$m^3$).
+\item [Q] quality factor to use in case of a simulation with attenuation turned on.
+It should be between 1 and 9000.
+In case no attenuation information is available, it can be set to zero.
+Please note, that the Vp- and Vs-velocities are given for a reference frequency.
+To change this reference frequency, you change the value of
+\texttt{ATTENUATION\_f0\_REFERENCE} in the main constants file \texttt{constants.h} found in subdirectory \texttt{src/shared/}.
+\item [anisotropic\_flag] Flag describing the anisotropic model to use in case an anisotropic simulation should be conducted.
+See the file \texttt{model\_aniso.f90} in subdirectory \texttt{src/generate\_databases/} for an implementation of the anisotropic models.
+In case no anisotropy is available, it can be set to zero.
+\end{description}
+Note that this material block has to be defined using all the volumes which belong to this elastic material.
+For volumes belonging to another, different material, you will need to define a new material block.
+
+For an acoustic material:
+\begin{description}
+\item [material\_id] An integer value which is unique for this material. 
+\item [Vp] P-Velocity of the material (given in m/s).
+\item [0] S-Velocity of the material is ignored.
+\item [rho] density of the material (given in kg/$m^3$).
+\end{description}
+
+\item [face\_topo] Block definition for the surface which defines the free surface (which can have topography).
+The name of this block must be 'face\_topo', the block has to be defined using all the surfaces which constitute the complete free surface of the model.
+
+\item [face\_abs\_xmin] Block definition for the faces on the absorbing boundaries, one block for each surface with x=Xmin. 
+\item [face\_abs\_xmax] Block definition for the faces on the absorbing boundaries, one block for each surface with x=Xmax.
+\item [face\_abs\_ymin] Block definition for the faces on the absorbing boundaries, one block for each surface with y=Ymin.
+\item [face\_abs\_ymax] Block definition for the faces on the absorbing boundaries, one block for each surface with y=Ymax.
+\item [face\_abs\_bottom] Block definition for the faces on the absorbing boundaries, one block for each surface with z=bottom.
+\end{description}
+
+\begin{figure}[htbp]
+\noindent \begin{centering}
+\includegraphics[width=.45\textwidth]{figures/mount-surface.jpg}
+\includegraphics[width=.45\textwidth]{figures/mount-abs.jpg}
+\par\end{centering}
+
+\caption{Example of the block definitions for the free surface 'face\_topo' (left) and the absorbing boundaries, defined in a single block 'face\_abs\_xmin' (right) in CUBIT.}
+
+\label{fig:mount.abs}
+\end{figure}
+
+Optionally, instead of specifying for each surface at the boundaries a single block like mentioned above, you can also specify a single block for all boundary surfaces and name it as one of the absorbing blocks above, e.g.
+'face\_abs\_xmin'.
+
+After the block definitions are done, you export the mesh using the script \texttt{cubit2specfem3d.py} provided in each of the example directories.
+If the export was successful, you should find the following files a the subdirectory \texttt{MESH/}:
+\begin{description}
+\item [nummaterial\_velocity\_file] Defines the material properties. For fully defined materials, the formats is:
+\begin{lyxcode}
+\texttt{domain\_ID} \texttt{material\_ID}  \texttt{rho}  \texttt{vp}  \texttt{vs}  \texttt{Q}  \texttt{anisotropy\_flag} 
+\end{lyxcode}
+where 
+\texttt{domain\_ID} is 1 for acoustic or 2 for elastic materials, \texttt{material\_ID} a unique identifier,  \texttt{rho} the density in $kg \, m^{-3}$,  \texttt{vp} the P-velocity in $m \, s^{-1}$, \texttt{Vs}  the S-velocity in $m \, s^{-1}$, \texttt{Q} the quality factor and \texttt{anisotropy\_flag} an identifier for anisotropic models.
+For a tomographic model, the material definition format is: 
+\begin{lyxcode}
+\texttt{domain\_ID} \texttt{material\_ID}  \texttt{tomography}  \texttt{name}  
+\end{lyxcode}
+where
+\texttt{domain\_ID} is 1 for acoustic or 2 for elastic materials, \texttt{material\_ID} a negative, unique identifier, \texttt{tomography} keyword for tomographic material definition and \texttt{name} the name of the tomography file.
+The name is not used so far, rather change the filename defined in the file ~\\
+\texttt{src/generate\_databases/model\_tomography.f90}.
+\item [materials\_file] Contains the material associations for each element. 
+\item [nodes\_coords\_file] Contains the point locations in Cartesian coordinates of the mesh element corners.
+\item [mesh\_file] Contains the mesh element connectivity.
+\item [free\_surface\_file] Contains the free surface connectivity.
+\item [absorbing\_surface\_file\_xmax] Contains the surface connectivity of the absorbing boundary surface at the Xmax.
+\item [absorbing\_surface\_file\_xmin] Contains the surface connectivity of the absorbing boundary surface at the Xmin.
+\item [absorbing\_surface\_file\_ymax] Contains the surface connectivity of the absorbing boundary surface at the Ymax.
+\item [absorbing\_surface\_file\_ymin] Contains the surface connectivity of the absorbing boundary surface at the Ymin.
+\item [absorbing\_surface\_file\_bottom] Contains the surface connectivity of the absorbing boundary surface at the bottom.
+\end{description}
+
 \subsection{Partitioning the Mesh with \texttt{xdecompose\_mesh\_SCOTCH}}
@@
 Fran\c {c}ois Pellegrini et al.
-from LaBRI and INRIA in Bordeaux, France, downloadable from the web page \texttt{https://gforge.inria.fr/projects/scotch/}.
+from LaBRI and INRIA in Bordeaux, France, downloadable from the web page \url{https://gforge.inria.fr/projects/scotch/}.
 It is more recent than METIS, actively maintained and performs better in many cases.
@@
 Prior to compiling \texttt{decompose\_mesh\_SCOTCH}, make sure you have correctly specified the path
-of the Scotch library with the option \texttt{-{}-with-scotch-lib} of the \texttt{configure} script.
+of the Scotch library with the option \texttt{-{}-with-scotch-path} of the \texttt{configure} script.
 For convenience we provide the source code of SCOTCH, which is released open source under the French CeCILL-C version 1 license,
-in directory src/decompose\_mesh\_SCOTCH/scotch\_5.1.10b. Please refer to file INSTALL.txt in that directory to see how to compile it.
-In the future you should be able to find for recent versions at http://www.labri.fr/perso/pelegrin/scotch/scotch\_en.html .
+in directory ~\\
+\texttt{src/decompose\_mesh\_SCOTCH/scotch\_5.1.10b}. Please refer to file INSTALL.txt in that directory to see how to compile it.
+In the future you should be able to find for recent versions at  \url{http://www.labri.fr/perso/pelegrin/scotch/scotch\_en.html} .
 \\
 
+\begin{figure}[htbp]
+\noindent \begin{centering}
+\includegraphics[width=.45\textwidth]{figures/mount-partitions.jpg}
+\includegraphics[width=.45\textwidth]{figures/mount-partitions2.jpg}
+\par\end{centering}
+
+\caption{Example of a mesh partitioning onto four cores.
+Each single core partition is colored differently.
+The executable \texttt{xdecompose\_mesh\_SCOTCH} can equally distribute the mesh on any arbitrary number of cores.}
+
+\label{fig:mount.partitions}
+\end{figure}
+
 Then you are ready to compile,
-in the directory \texttt{decompose\_mesh\_SCOTCH} type '\texttt{make decompose\_mesh\_SCOTCH}'.
-If all paths and flags have been set correctly, the executable \texttt{xdecompose\_mesh\_SCOTCH} should be produced.
+in the main directory type '\texttt{make decompose\_mesh\_SCOTCH}'.
+If all paths and flags have been set correctly, the executable \texttt{bin/xdecompose\_mesh\_SCOTCH} should be produced.
 
@@
 \begin{lyxcode}
-./decompose\_mesh\_SCOTCH  nparts  input\_directory output\_directory
+./bin/xdecompose\_mesh\_SCOTCH  nparts  input\_directory output\_directory
 \end{lyxcode}
@@
 %
-\begin{figure}[t]
+\begin{figure}[htbp]
 \noindent \begin{centering}
@@
 
-You are now ready to compile the internal mesher. This is an alternative to CUBIT
+In case you successfully ran the configuration script, you are also ready to compile the internal mesher. 
+This is an alternative to CUBIT
 for the mesh generation of relatively simple geological models. The mesher is no longer
@@
 
-In the directory \texttt{meshfem3D} type `\texttt{make meshfem3D}'. If all paths and flags
+In the main directory type `\texttt{make meshfem3D}'. If all paths and flags
 have been set correctly, the mesher should now compile and produce
-the executable \texttt{xmeshfem3D}.
+the executable \texttt{bin/xmeshfem3D}.
 
 Input for the mesh generation program is provided through the parameter
-file \texttt{Par\_file}, which resides in the subdirectory \texttt{meshfem3D/DATA}.
+file \texttt{Mesh\_Par\_file}, which resides in the subdirectory \texttt{in\_data\_files/meshfem3D\_files/}.
 Before running the mesher, a number of parameters need to be set in
-the \texttt{Par\_file}. This requires a basic understanding of how
+the \texttt{Mesh\_Par\_file}. This requires a basic understanding of how
 the SEM is implemented, and we encourage you to read \citet{KoVi98,KoTr99}
@@
 for a global simulation, the following parameters need to be set in
-the \texttt{Par\_file}:
+the \texttt{Mesh\_Par\_file}:
 
@@
 \begin{lyxcode}
-\texttt{material\_ID}  \texttt{rho}  \texttt{vp}  \texttt{vs}  \texttt{Q\_flag}  \texttt{anisotropy\_flag} \texttt{domain\_ID}
+\texttt{material\_ID}  \texttt{rho}  \texttt{vp}  \texttt{vs}  \texttt{Q}  \texttt{anisotropy\_flag} \texttt{domain\_ID}
 \end{lyxcode}
@@
 \begin{itemize}
-     \item \texttt{Q\_flag}           : 0=no attenuation / standard \texttt{Q} attenutation value
+     \item \texttt{Q}           : 0=no attenuation / standard \texttt{Q} attenutation value
      \item \texttt{anisotropy\_flag}  : 0=no anisotropy / 1,2,.. check with implementation in \texttt{aniso\_model.f90}
@@
 
-Mesher output is provided in the \texttt{OUTPUT\_FILES} directory
+Mesher output is provided in the \texttt{in\_out\_files/OUTPUT\_FILES} directory
 in \texttt{output\_mesher.txt}; this file provides lots of details
@@
 The quality of the mesh may be inspected more precisely based upon the serial code
+in the file  ~\\
 \texttt{check\_mesh\_quality\_CUBIT\_Abaqus.f90}. In the directory \texttt{check\_mesh\_quality\_CUBIT\_Abaqus}, type
@@
 This program is going to create all the missing information needed by the SEM solver.
+
+\begin{figure}[htbp]
+\noindent \begin{centering}
+\includegraphics[width=.6\textwidth]{figures/workflow.jpg}
+\par\end{centering}
+
+\caption{Schematic workflow for a SPECFEM3D simulation. The exectable \texttt{xgenerate\_databases} creates the GLL mesh points and assigns the specific model parameters.}
+
+\label{fig:workflow.databases}
+\end{figure}
+
 In the main directory type '\texttt{make generate\_databases}'.
-Input for the program is provided through the main parameter file \texttt{Par\_file}, which resides in the subdirectory \texttt{DATA}.
+Input for the program is provided through the main parameter file \texttt{Par\_file}, which resides in the subdirectory \texttt{in\_data\_files/}.
 Before running \texttt{xgenerate\_databases}, a number of parameters need to be set in the \texttt{Par\_file}:
@@
 \item [{\texttt{DT}}] The length of each time step in seconds.
-If you used xmeshfem3D to generate your mesh, you should set the value suggested in \texttt{meshfem3D/OUTPUT\_FILES/output\_mesher.txt}.
+If you used xmeshfem3D to generate your mesh, you should set the value suggested in \texttt{in\_out\_files/OUTPUT\_FILES/output\_mesher.txt}.
 
@@
 of meshing but is required for the solver, i.e., you may change this
-parameter after running xgenerate\_databases.
+parameter after running \texttt{xgenerate\_databases}.
 \item [{\texttt{MOVIE\_SURFACE}}] Set to \texttt{.false.}, unless you want
@@
 discussion of the \texttt{LOCAL\_PATH\_IS\_ALSO\_GLOBAL} flag in Chapter~\ref{cha:Getting-Started}).
-xgenerate\_databases generates the necessary databases in parallel, one set
+\texttt{xgenerate\_databases} generates the necessary databases in parallel, one set
 for each of the $\nprocxi\times\nproceta$ slices that constitutes
@@
 the contents of the \texttt{LOCAL\_PATH} directory to see the (many)
-files generated by xgenerate\_databases.
+files generated by \texttt{xgenerate\_databases}.
 \item [{\texttt{NTSTEP\_BETWEEN\_OUTPUT\_INFO}}] This parameter specifies
 the interval at which basic information about a run is written to
-the file system (\texttt{timestamp{*}} files in the \texttt{OUTPUT\_FILES}
+the file system (\texttt{timestamp{*}} files in the \texttt{in\_out\_files/OUTPUT\_FILES}
 directory). If you have access to a fast machine, set \texttt{NTSTEP\_BETWEEN\_OUTPUT\_INFO}
@@
 \item [{\texttt{PRINT\_SOURCE\_TIME\_FUNCTION}}] Turn this flag on to print
-information about the source time function in the file \texttt{OUTPUT\_FILES/plot\_source\_time\_function.txt}.
+information about the source time function in the file \texttt{in\_out\_files/OUTPUT\_FILES/plot\_source\_time\_function.txt}.
 This feature is only relevant for the solver.
@@
 Now that you have successfully generated the databases, you are ready to compile
-the solver. For reasons of speed, the solver uses static memory allocation.
-Therefore it needs to be recompiled (type `\texttt{make clean}' and
-`\texttt{make specfem3D}') every time one reruns \texttt{xgenerate\_databases}. To compile
-the solver one needs a file created by \texttt{xgenerate\_databases} in the directory
-\texttt{OUTPUT\_FILES} called \texttt{values\_from\_mesher.h}, which
-contains parameters describing the static size of the arrays as well
-as the setting of certain flags.
-
-The solver needs three input files in the \texttt{DATA} directory
-to run: the \texttt{Par\_file} which was discussed in detail in Chapters~\ref{cha:Running-the-Mesher-CUBIT} and~\ref{cha:Running-the-Mesher-Meshfem3D},
-the earthquake source parameter file \texttt{CMTSOLUTION}, and the
-stations file \texttt{STATIONS}. Most parameters in the \texttt{Par\_file}
-should be set prior to running the mesher. Only the following parameters
+the solver. In the main directory type `\texttt{make specfem3D}'.
+%obsolete
+%For reasons of speed, the solver uses static memory allocation.
+%Therefore it needs to be recompiled (type `\texttt{make clean}' and
+%`\texttt{make specfem3D}') every time one reruns \texttt{xgenerate\_databases}. To compile
+%the solver one needs a file created by \texttt{xgenerate\_databases} in the directory
+%\texttt{OUTPUT\_FILES} called \texttt{values\_from\_mesher.h}, which
+%contains parameters describing the static size of the arrays as well
+%as the setting of certain flags.
+
+The solver needs three input files in the \texttt{in\_data\_files/} directory
+to run: 
+\begin{description}
+\item [{\texttt{Par\_file}}] the main parameter file which was discussed in detail in the previous 
+Chapter~\ref{cha:Creating-Distributed-Databases},
+\item [{\texttt{CMTSOLUTION}}] the earthquake source parameter file, and 
+\item [{\texttt{STATIONS}}] the stations file. 
+\end{description}
+
+Most parameters in the \texttt{Par\_file}
+should be set prior to running the databases generation. Only the following parameters
 may be changed after running \texttt{xgenerate\_databases}:
@@
 Any other change to the \texttt{Par\_file} implies rerunning both
-the mesher and the solver.
+the generate\_databases and the solver.
 
@@
 \end{lyxcode}
-The mesher filters the list of stations in file \texttt{DATA/STATIONS}
+The mesher filters the list of stations in file \texttt{in\_data\_files/STATIONS}
 to exclude stations that are not located within the region given in
@@
 and between \texttt{LONGITUDE\_MIN} and \texttt{LONGITUDE\_MAX}).
-The filtered file is called \texttt{DATA/STATIONS\_FILTERED}.
-
-Solver output is provided in the \texttt{OUTPUT\_FILES} directory
+The filtered file is called \texttt{in\_data\_files/STATIONS\_FILTERED}.
+
+Solver output is provided in the \texttt{in\_out\_files/OUTPUT\_FILES} directory
 in the \texttt{output\_solver.txt} file. Output can be directed to
@@
 While the solver is running, its progress may be tracked by monitoring
-the `\texttt{\small timestamp{*}}' files in the \texttt{\small OUTPUT\_FILES}
-directory. These tiny files look something like this:
+the `\texttt{\small timestamp{*}}' files in the ~\\
+\texttt{\small in\_out\_files/OUTPUT\_FILES} directory. 
+These tiny files look something like this:
 
@@
 and then collect the recorded seismograms at all the stations given
-in \texttt{DATA/STATIONS}.
+in \texttt{in\_data\_files/STATIONS}.
 
@@
 adjoint sources and run the adjoint simulation, and compile them into
-the \texttt{DATA/STATIONS\_ADJOINT} file, which has the same format
-as the regular \texttt{DATA/STATIONS} file.
+the \texttt{in\_data\_files/STATIONS\_ADJOINT} file, which has the same format
+as the regular \texttt{in\_data\_files/STATIONS} file.
 
@@
 with the format \texttt{STA.NT.BH?.adj}, where \texttt{STA}, \texttt{NT}
-are the station name and network code given in the \texttt{DATA/STATIONS\_ADJOINT}
+are the station name and network code given in the \texttt{in\_data\_files/STATIONS\_ADJOINT}
 file, and \texttt{BH?} represents the component name of a particular
@@
 \item \textbf{Set the related parameters and run the adjoint simulation}\\
-In the \texttt{DATA/Par\_file}, set the two related parameters to
+In the \texttt{in\_data\_files/Par\_file}, set the two related parameters to
 be \texttt{SIMULATION\_TYPE = 2} and \texttt{SAVE\_FORWARD = .false.}.
@@
 \item These adjoint seismograms are recorded at the locations of the original
-earthquake sources given by the \texttt{DATA/CMTSOLUTION} file, and
+earthquake sources given by the \texttt{in\_data\_files/CMTSOLUTION} file, and
 have names of the form \texttt{S?????.NT.S??.sem} for the six-component
 strain tensor (\texttt{SNN,SEE,SZZ,SNE,SNZ,SEZ}) at these locations, 
-and \texttt{S?????.NT.BH?.sem} for the three-component displacements
+and ~\\
+\texttt{S?????.NT.BH?.sem} for the three-component displacements
 (\texttt{BHN,BHE,BHZ}) recorded at these locations.
@@
 files are saved in the \texttt{LOCAL\_PATH} directory, whereas \texttt{MOVIE\_SURFACE}
-output files are saved in the \texttt{OUTPUT\_FILES} directory. We
+output files are saved in the \texttt{in\_out\_files/OUTPUT\_FILES} directory. We
 save the velocity field. The look of a movie is determined by the
@@
 flag turned on, the code outputs \texttt{moviedata??????} files in
-the \texttt{OUTPUT\_FILES} directory. The files are in a fairly complicated
+the \texttt{in\_out\_files/OUTPUT\_FILES} directory. The files are in a fairly complicated
 binary format, but there are two programs provided to convert the
@@
 In the case of a single source and a single receiver, the simulation
-also generates the \texttt{OUTPUT\_FILES/sr.vtk} file to describe
+also generates the file \texttt{sr.vtk} located in the 
+\texttt{in\_out\_files/OUTPUT\_FILES/} directory to describe
 the source and receiver locations, which can also be viewed in Paraview
@@
 not configured for compilation. It may therefore be necessary to pre-compile
-both the mesher and the solver. A small program provided in the distribution
-called \texttt{\small create\_header\_file.f90} can be used to directly
-create \texttt{\small OUTPUT\_FILES/values\_from\_mesher.h} using
-the information in the \texttt{\small DATA/Par\_file} without having
-to run the mesher (type \texttt{\small `make}{\small{} }\texttt{\small create\_header\_}~\\
-\texttt{\small file}' to compile it and `\texttt{\small xcreate\_header\_file}'
-to run it; refer to the sample scripts below). The solver can now
-be compiled as explained above.
+both the mesher and the solver. 
+%obsolete
+%A small program provided in the distribution
+%called \texttt{\small create\_header\_file.f90} can be used to directly
+%create \texttt{\small in\_out\_files/OUTPUT\_FILES/values\_from\_mesher.h} using
+%the information in the \texttt{\small in\_data\_files/Par\_file} without having
+%to run the mesher (type \texttt{\small `make}{\small{} }\texttt{\small create\_header\_}~\\
+%\texttt{\small file}' to compile it and `\texttt{\small xcreate\_header\_file}'
+%to run it; refer to the sample scripts below). The solver can now
+%be compiled as explained above.
 \item One feature of schedulers/queuing systems is that they allow submission
@@
 the mesher and solver together, figures out the number of processors
-required for this simulation from the \texttt{DATA/Par\_file}, and
+required for this simulation from the \texttt{in\_data\_files/Par\_file}, and
 submits the LSF job.
@@
 
-make~meshfem3D~
-
-make~create\_header\_file~
-
-xcreate\_header\_file~
+make~generate\_databases~
 
@@
 
-NPROC\_XI=`grep~NPROC\_XI~DATA/Par\_file~|~cut~-c~34-~'~
-
-NPROC\_ETA=`grep~NPROC\_ETA~DATA/Par\_file~|~cut~-c~34-~'~~\\
+NPROC=`grep~NPROC~in\_data\_files/Par\_file~|~cut~-c~34-~'~
+
 ~\\
@@
 
-numnodes=\$((~\$NPROC\_XI~{*}~\$NPROC\_ETA~))~\\
+numnodes=\$((~\$NPROC))~\\
 ~\\
@@
 \item First the script directs the scheduler to save its own output and
-output from \texttt{stdout} into \texttt{\small OUTPUT\_FILES/\%J.o},
+output from \texttt{stdout} into ~\\
+\texttt{\small in\_out\_files/OUTPUT\_FILES/\%J.o},
 where \texttt{\%J} is short-hand for the job-ID; it also tells the
@@
 from Shmux \url{web.taranis.org/shmux/}. Make sure that the \texttt{LOCAL\_PATH}
-parameter in \texttt{DATA/Par\_file} is also set properly.
+parameter in \texttt{in\_data\_files/Par\_file} is also set properly.
 \item The next portion of the script launches the mesher and then the solver
@@
 
-\#BSUB~-o~OUTPUT\_FILES/\%J.o~
+\#BSUB~-o~in\_out\_files/OUTPUT\_FILES/\%J.o~
 
@@
 
-mkdir~-p~OUTPUT\_FILE
-
-echo~\char`\"{}\$LSB\_MCPU\_HOSTS\char`\"{}~>~OUTPUT\_FILES/lsf\_machines~
-
-echo~\char`\"{}\$LSB\_JOBID\char`\"{}~>~OUTPUT\_FILES/jobid
-
-remap\_lsf\_machines.pl~OUTPUT\_FILES/lsf\_machines~>OUTPUT\_FILES/machines
+mkdir~-p~in\_out\_files/OUTPUT\_FILE
+
+echo~\char`\"{}\$LSB\_MCPU\_HOSTS\char`\"{}~>~in\_out\_files/OUTPUT\_FILES/lsf\_machines~
+
+echo~\char`\"{}\$LSB\_JOBID\char`\"{}~>~in\_out\_files/OUTPUT\_FILES/jobid
+
+remap\_lsf\_machines.pl~in\_out\_files/OUTPUT\_FILES/lsf\_machines~> in\_out\_files/OUTPUT\_FILES/machines
 
@@
 
-~~~~~~-~<~OUTPUT\_FILES/machines~>/dev/null~\\
+~~~~~~-~<~in\_out\_files/OUTPUT\_FILES/machines~>/dev/null~\\
 ~\\
@@
 
-current\_pwd=\$PWD
+current\_pwd=\$PWD \\
+
+cd bin/ \\
 
@@
 
-mkdir~-p~SEM
-
-cd~SEM
+cd current\_pwd \\
+
+mkdir~-p~in\_out\_files/SEM
+
+cd~in\_out\_files/SEM
 
@@
 
-With the movie data saved in \texttt{OUTPUT\_FILES/} at the end of
+With the movie data saved in \texttt{in\_out\_files/OUTPUT\_FILES/} at the end of
 a movie simulation (\texttt{\small MOVIE\_SURFACE=.true.}), you can
@@
 
-With the shakemap data saved in \texttt{OUTPUT\_FILES/} at the end
-of a shakemap simulation (\texttt{CREATE\_SHAKEMAP=.true.}), you can
+With the shakemap data saved in \texttt{in\_out\_files/OUTPUT\_FILES/} at the end
+of a shakemap simulation ~\\
+(\texttt{CREATE\_SHAKEMAP=.true.}), you can
 also run \texttt{`create\_movie\_GMT}' code to convert the binary

### 2010-11-09 doc/USER_MANUAL/manual_SPECFEM3D.tex
updated the list of authors and the manual

@@
 The SPECFEM3D package was first developed by Dimitri Komatitsch and Jeroen Tromp at Harvard University, then Caltech and Princeton, USA, starting in 1998, based in part on earlier work by Dimitri Komatitsch and Jean-Pierre Vilotte at IPG in Paris, France from 1994 to 1997.
-Since then it has been developed and maintained by a development team: in alphabetical order, C\'eline Blitz, Emiljana Jorgji, Dimitri Komatitsch, Nicolas Le Goff, Pieyre Le Loher, Roland Martin, David Mich\'ea, Jeroen Tromp \toall{add full list of developers here}.
+Since then it has been developed and maintained by a development team: in alphabetical order, C\'eline Blitz, Emiljana Jorgji, Dimitri Komatitsch, Nicolas Le Goff, Pieyre Le Loher, Roland Martin, David Mich\'ea, Jeroen Tromp \toall{add full list of developers here and update the AUTHORS file accordingly}.
 
@@
 have been set correctly, the mesher should now compile and produce
-the executable \texttt{bin/xmeshfem3D}.
+the executable \texttt{bin/xmeshfem3D}. Please note that  
+\texttt{xmeshfem3D} must be called directly from the \texttt{bin/} directory, 
+as most of the binaries of the package.
 
@@
 
-%% \section{Checking the MPI Buffers (Optional)}
-
-%% The mesher writes MPI communication tables in the \texttt{OUTPUT\_FILES}
-%% subdirectory in the files \texttt{addressing.txt}, \texttt{list\_messages\_corners.txt}
-%% and \texttt{list\_messages\_faces.txt}, and MPI communication buffers
-%% to the local disks. Use the four serial codes
-
-%% \begin{lyxcode}
-%% check\_buffers\_2D.f90
-
-%% check\_buffers\_1D.f90
-%% \end{lyxcode}
-%% to check that all the MPI buffers created by the mesher have been
-%% generated correctly. For example, typing `\texttt{make check\_buffers\_2D}'
-%% and then `\texttt{xcheck\_buffers\_2D}' checks the communication buffers
-%% between faces common to the mesh slices.
-
-%% \begin{quote}
-%% Please note that running these codes is optional because no information
-%% needed by the solver is generated.
-%% \end{quote}
-
-%%\section{\label{sec:quality}Checking the Mesh Quality (Optional)}
-
 The quality of the mesh may be inspected more precisely based upon the serial code
@@
 
-In the main directory type '\texttt{make generate\_databases}'.
-Input for the program is provided through the main parameter file \texttt{Par\_file}, which resides in the subdirectory \texttt{in\_data\_files/}.
+In the main directory type '\texttt{make generate\_databases}'. Input for the program is provided through 
+the main parameter file \texttt{Par\_file}, which resides in the subdirectory \texttt{in\_data\_files/}. Please note that  
+\texttt{xgenerate\_databases} must be called directly from the \texttt{bin/} directory, as most of the binaries of the package.
 Before running \texttt{xgenerate\_databases}, a number of parameters need to be set in the \texttt{Par\_file}:
@@
 the solver. In the main directory type `\texttt{make specfem3D}'.
+Please note that \texttt{xspecfem3D} must be called directly from 
+the \texttt{bin/} directory, as most of the binaries of the package.
+
 %obsolete

### 2010-11-11 doc/USER_MANUAL/manual_SPECFEM3D.tex
updates SPECFEM3D manual; modified output path for adjoint seismograms to have same behavior as for forward seismograms; updates some utils scripts which referred to older version

@@
 on most cluster applications. For reasons of speed, the (parallel)
-mesher typically writes a (parallel) database for the solver on the
+distributed database generator typically writes a (parallel) database for the solver on the
 local disks of the compute nodes. Some systems have no local disks,
@@
 parallel file system, in which case this flag should be set to \texttt{.true.}.
-Note that this flag is not used by the mesher or the solver; it is
+Note that this flag is not used by the database generator or the solver; it is
 only used for some of the post-processing.
@@
 
-\caption{Screenshots of the CUBIT examples provided in subdirectory  \texttt{examples/}: homogeneous halfspace, layered halfspace, waterlayered halfspace and tomographic model.}
+\caption{Screenshots of the CUBIT examples provided in subdirectory  \texttt{examples/}: homogeneous halfspace (top-left), layered halfspace (top-right), waterlayered halfspace (bottom-left) and tomographic model (bottom-right).}
 
@@
 \subsection{Exporting the Mesh with \texttt{cubit2specfem3d.py} }
+\label{subsec:Exporting-the-Mesh}
 
@@
 In case no attenuation information is available, it can be set to zero.
-Please note, that the Vp- and Vs-velocities are given for a reference frequency.
+Please note that your Vp- and Vs-velocities are given for a reference frequency.
 To change this reference frequency, you change the value of
@@
 After the block definitions are done, you export the mesh using the script \texttt{cubit2specfem3d.py} provided in each of the example directories.
-If the export was successful, you should find the following files a the subdirectory \texttt{MESH/}:
+If the export was successful, you should find the following files in a subdirectory \texttt{MESH/}:
+
 \begin{description}
@@
 where 
-\texttt{domain\_ID} is 1 for acoustic or 2 for elastic materials, \texttt{material\_ID} a unique identifier,  \texttt{rho} the density in $kg \, m^{-3}$,  \texttt{vp} the P-velocity in $m \, s^{-1}$, \texttt{Vs}  the S-velocity in $m \, s^{-1}$, \texttt{Q} the quality factor and \texttt{anisotropy\_flag} an identifier for anisotropic models.
+\texttt{domain\_ID} is 1 for acoustic or 2 for elastic materials, \texttt{material\_ID} a unique identifier,  \texttt{rho} the density in $kg \, m^{-3}$,  \texttt{vp} the P-velocity in $m \, s^{-1}$, \texttt{vs}  the S-velocity in $m \, s^{-1}$, \texttt{Q} the quality factor and \texttt{anisotropy\_flag} an identifier for anisotropic models.
 For a tomographic model, the material definition format is: 
@@
 \texttt{domain\_ID} is 1 for acoustic or 2 for elastic materials, \texttt{material\_ID} a negative, unique identifier, \texttt{tomography} keyword for tomographic material definition and \texttt{name} the name of the tomography file.
-The name is not used so far, rather change the filename defined in the file ~\\
-\texttt{src/generate\_databases/model\_tomography.f90}.
+The name is not used so far, rather change the filename defined in the file \texttt{model\_tomography.f90} located in the
+\texttt{src/generate\_databases/} directory.
 \item [materials\_file] Contains the material associations for each element. 
@@
 \end{lyxcode}
-
-All the files generated by the Python scripts must be placed in the \texttt{input\_directory} folder before running the program.
+where
+\begin{itemize}
+     \item \texttt{nparts} is the number of partitions, i.e. number of cores for the parallel simulations,
+     \item \texttt{input\_directory} is the directory which holds all the files generated by the Python script \texttt{cubit2specfem3d.py} explained in the previous Section~\ref{subsec:Exporting-the-Mesh}, e.g.
+\texttt{MESH/}, and
+     \item \texttt{output\_directory} is the directory for the output of this partitioner which stores ACII-format files named like \texttt{proc******\_Database} for each partition.
+These files will be needed for creating the distributed databases, and have to reside in the directory  \texttt{LOCAL\_PATH} specified in the main \texttt{Par\_file}, e.g.
+in directory \texttt{in\_out\_files/DATABASES\_MPI}.
+Please see Chapter~\ref{cha:Creating-Distributed-Databases} for further details.
+\end{itemize}
+Note that all the files generated by the Python script  \texttt{cubit2specfem3d.py} must be placed in the \texttt{input\_directory} folder before running the program.
 
@@
 
-\noindent \begin{flushleft}
+In case you successfully ran the configuration script, you are also ready to compile the internal mesher. 
+This is an alternative to CUBIT
+for the mesh generation of relatively simple geological models. The mesher is no longer
+dedicated to Southern California and more flexiblity is provided in this version of the package.
+
+%\noindent \begin{flushleft}
 %
@@
 
-\par\end{flushleft}
-
-In case you successfully ran the configuration script, you are also ready to compile the internal mesher. 
-This is an alternative to CUBIT
-for the mesh generation of relatively simple geological models. The mesher is no longer
-dedicated to Southern California and more flexiblity is provided in this version of the package.
+%\par\end{flushleft}
 
@@
 mesh files for subsequent viewing.
-\item [{\texttt{LOCAL\_PATH}}] Directory in which the databases generated
+\item [{\texttt{LOCAL\_PATH}}] Directory in which the partitions generated
 by the mesher will be written. Generally one uses a directory on the
-local disk of the compute nodes, although on some machines these databases
+local disk of the compute nodes, although on some machines these partitions
 are written on a parallel (global) file system (see also the earlier
 discussion of the \texttt{LOCAL\_PATH\_IS\_ALSO\_GLOBAL} flag in Chapter~\ref{cha:Getting-Started}).
-The mesher generates the necessary databases in parallel, one set
+The mesher generates the necessary partitions in parallel, one set
 for each of the $\nprocxi\times\nproceta$ slices that constitutes
@@
 mesher finishes, you can log in to one of the compute nodes and view
-the contents of the \texttt{LOCAL\_PATH} directory to see the (many)
+the contents of the \texttt{LOCAL\_PATH} directory to see the
 files generated by the mesher.
+These files will be needed for creating the distributed databases, and have to reside in the directory  \texttt{LOCAL\_PATH} specified in the main \texttt{Par\_file}, e.g.
+in directory \texttt{in\_out\_files/DATABASES\_MPI}.
+Please see Chapter~\ref{cha:Creating-Distributed-Databases} for further details.
 \item [{\texttt{NMATERIALS}}] The number of different materials in your model. In the following lines, each material needs to be defined as :
@@
 \begin{itemize}
-     \item \texttt{Q}           : 0=no attenuation / standard \texttt{Q} attenutation value
+     \item \texttt{Q}           : 0=no attenuation / standard \texttt{Q} attenuation value
      \item \texttt{anisotropy\_flag}  : 0=no anisotropy / 1,2,.. check with implementation in \texttt{aniso\_model.f90}
@@
 \texttt{xgenerate\_databases} must be called directly from the \texttt{bin/} directory, as most of the binaries of the package.
-Before running \texttt{xgenerate\_databases}, a number of parameters need to be set in the \texttt{Par\_file}:
+
+\section{\label{cha:Main-Parameter}Main parameter file \texttt{Par\_file}}
+
+Before running \texttt{xgenerate\_databases}, a number of parameters need to be set in the main parameter \texttt{Par\_file}
+located in the subdirectory \texttt{in\_data\_files/}:
 
@@
 requires twice as much CPU time. This feature is not used at the time
-of meshing but is required for the solver, i.e., you may change this
+of generating the distributed databases but is required for the solver, i.e., you may change this
 parameter after running  \texttt{xgenerate\_databases}.
 \item [{\texttt{DT}}] The length of each time step in seconds. 
-If you used xmeshfem3D to generate your mesh, you should set the value suggested in \texttt{in\_out\_files/OUTPUT\_FILES/output\_mesher.txt}.
-
+This feature is not used at the time of generating the distributed databases but is required for the solver.
+Please see also Section~\ref{cha:Choosing-the-Time-Step} for further details.
 \item [{\texttt{OCEANS}}] Set to \texttt{.true.} if the effect of the oceans
@@
 is on.
-\item [{\texttt{TOPOGRAPHY}}] Set to \texttt{.true.} if topography and
-bathymetry should be incorporated based upon model ETOPO5 \citep{Etopo5}.
-This feature adds no cost to the simulation.
+%obsolete...
+%\item [{\texttt{TOPOGRAPHY}}] Set to \texttt{.true.} if topography and
+%bathymetry should be incorporated based upon model ETOPO5 \citep{Etopo5}.
+%This feature adds no cost to the simulation.
 \item [{\texttt{ATTENUATION}}] Set to \texttt{.true.} if attenuation should
@@
 of attenuation based upon standard linear solids.
+Please note that the Vp- and Vs-velocities of your model are given for a reference frequency. 
+To change this reference frequency, you change the value of 
+\texttt{ATTENUATION\_f0\_REFERENCE} in the main constants file \texttt{constants.h} found in subdirectory \texttt{src/shared/}.
 \item [{\texttt{USE\_OLSEN\_ATTENUATION}}] Set to \texttt{.true.} if you
@@
 using Olsen's empirical relation (reference).
+\item [{\texttt{ANISOTROPY}}] Set to \texttt{.true.} if you
+want to use an anisotropy model. Please see the file \texttt{model\_aniso.f90} in subdirectory \texttt{src/generate\_databases/} for the current implementation of anisotropic models.
 \item [{\texttt{ABSORBING\_CONDITIONS}}] Set to \texttt{.true.} to turn
 on Clayton-Enquist absorbing boundary conditions (see \citet{KoTr99}).
-\item [{\texttt{RECORD\_LENGTH\_IN\_MINUTES}}] Choose the desired record
-length of the synthetic seismograms (in minutes). This controls the
-length of the numerical simulation, i.e., twice the record length
-requires twice as much CPU time. This feature is not used at the time
-of meshing but is required for the solver, i.e., you may change this
-parameter after running \texttt{xgenerate\_databases}.
+%obsolete...
+%\item [{\texttt{RECORD\_LENGTH\_IN\_MINUTES}}] Choose the desired record
+%length of the synthetic seismograms (in minutes). This controls the
+%length of the numerical simulation, i.e., twice the record length
+%requires twice as much CPU time. This feature is not used at the time
+%of meshing but is required for the solver, i.e., you may change this
+%parameter after running \texttt{xgenerate\_databases}.
 \item [{\texttt{MOVIE\_SURFACE}}] Set to \texttt{.false.}, unless you want
@@
 movie frames.
+\item [{\texttt{HDUR\_MOVIE}}] determines the half duration of the source time function for the movie simulations. 
+When this parameter is set to be 0, a default half duration that corresponds to the accuracy of the simulation is provided. 
+Otherwise, it adds this half duration to the half duration specified in the source file \texttt{CMTSOLUTION},
+thus simulates longer periods to make the movie images look smoother.
 \item [{\texttt{SAVE\_MESH\_FILES}}] Set this flag to \texttt{.true.} to
@@
 for a discussion of mesh viewing features.
-\item [{\texttt{LOCAL\_PATH}}] Directory in which the databases and the synthetics
+\item [{\texttt{LOCAL\_PATH}}] Directory in which the distributed databases
 will be written. Generally one uses a directory on the
@@
 \texttt{xgenerate\_databases} generates the necessary databases in parallel, one set
-for each of the $\nprocxi\times\nproceta$ slices that constitutes
-the mesh (see Figure~\ref{fig:For-parallel-computing}). After the
-mesher finishes, you can log in to one of the compute nodes and view
+for each of the \texttt{NPROC} slices that constitutes
+the mesh (see Figure~\ref{fig:mount.partitions} and Figure~\ref{fig:For-parallel-computing}). After
+the executable finishes, you can log in to one of the compute nodes and view
 the contents of the \texttt{LOCAL\_PATH} directory to see the (many)
 files generated by \texttt{xgenerate\_databases}. 
+Please note that the \texttt{LOCAL\_PATH} directory should already contain the output files of 
+the partitioner, i.e. from \texttt{xdecompose\_mesh\_SCOTCH} or \texttt{xmeshfem3D}.
 \item [{\texttt{NTSTEP\_BETWEEN\_OUTPUT\_INFO}}] This parameter specifies
@@
 
+\section{\label{cha:Choosing-the-Time-Step}Choosing the time step \texttt{DT}}
+
+The parameter \texttt{DT} sets the length of each time step in seconds. The value of this parameter is crucial for the stability of 
+the spectral-element simulation.
+Your time step \texttt{DT} will depend on the minimum ratio between the distance $h$ of neighboring mesh points and the wave speeds $v$ defined in your model.
+The condition for the time step $\Delta t$ is:
+
+\begin{lyxcode}
+$\Delta t < C ~\mathrm{min}_\Omega (~ h / v ~)$
+\end{lyxcode}
+where $C$ is the so-called Courant number and $\Omega$ denotes the model volume. 
+The distance $h$ depends on the mesh element size and the number of
+GLL points \texttt{NGLL} specified in the main constants file \texttt{constants.h} located in the \texttt{src/shared/} subdirectory.
+The wave speed $v$ is determined based on your model's P- (or S-) velocity values.
+
+The database generator \texttt{xgenerate\_databases}, as well as the internal mesher \texttt{xmeshfem3D}, are trying
+to evaluate the value of $\Delta t$ for empirically chosen Courant numbers $C \sim 0.3$.
+If you used the mesher \texttt{xmeshfem3D} to generate your mesh, you should set the value suggested in \texttt{in\_out\_files/OUTPUT\_FILES/output\_mesher.txt} file, which is created after the mesher completed.
+In case you used CUBIT to create the mesh, you might use an arbitrary value when running \texttt{xgenerate\_databases} and then use the value suggested in the ~\\
+\texttt{in\_out\_files/OUTPUT\_FILES/output\_mesher.txt} file after the database generation completed. 
+Note that the implemented Newmark time scheme uses this time step globally, thus your simulations become more expensive for very small mesh elements in high wave-speed regions.
+Please be aware of this restriction when constructing your mesh in Chapter~\ref{cha:Mesh-Generation}.
+
 \chapter{\label{cha:Running-the-Solver}Running the Solver \texttt{xspecfem3D}}
@@
 \item the time step parameters \texttt{NSTEP} and \texttt{DT}
+\item the attenuation control parameters \texttt{ATTENUATION} and \texttt{USE\_OLSEN\_ATTENUATION}
+\item the absorbing boundary control parameter \texttt{ABSORBING\_CONDITIONS} 
 \item the movie control parameters \texttt{MOVIE\_SURFACE}, \texttt{MOVIE\_VOLUME},
@@
 Any other change to the \texttt{Par\_file} implies rerunning both
-the generate\_databases and the solver.
+the database generator \texttt{xgenerate\_databases} and the solver \texttt{xspecfem3D}.
 
@@
 \end{lyxcode}
-The mesher filters the list of stations in file \texttt{in\_data\_files/STATIONS}
+The solver \texttt{xspecfem3D} filters the list of stations in file \texttt{in\_data\_files/STATIONS}
 to exclude stations that are not located within the region given in
@@
 the \texttt{mpirun} command. The \texttt{runall} script compiles and
-runs both mesher and solver in sequence. This is a safe approach that
-ensures using the correct combination of mesher output and solver
+runs both \texttt{xgenerate\_databases} and \texttt{xspecfem3D} in sequence. This is a safe approach that
+ensures using the correct combination of distributed database output and solver
 input.
@@
 stations by changing the \texttt{STATIONS} file. There is no need
-to rerun the mesher. Of course it is best to include as many stations
+to rerun the \texttt{xgenerate\_databases} executable. Of course it is best to include as many stations
 as possible, since this does not add to the cost of the simulation.
@@
 by the scheduler, based on availability. Therefore, in order for the
-mesher and the solver (or between successive runs of the solver) to
+\texttt{xgenerate\_databases} and the \texttt{xspecfem3D} executables (or between successive runs of the solver) to
 have access to the same database files (if they are stored on hard
@@
 not configured for compilation. It may therefore be necessary to pre-compile
-both the mesher and the solver. 
+both the \texttt{xgenerate\_databases} and the \texttt{xspecfem3D} executables. 
 %obsolete
@@
 \end{itemize}
+Examples of job scripts can be found in \texttt{\small UTILS/Cluster/}
+directory and can straightforwardly be modified and adapted to meet
+more specific running needs.
+
 We describe here in some detail a job submission procedure for the
 Caltech 1024-node cluster, CITerra, under the LSF scheduling system.
-We consider the submission of a regular forward simulation. The two
+We consider the submission of a regular forward simulation using the internal mesher
+to create mesh partitions. The two
 main scripts are \texttt{\small run\_lsf.bash}, which compiles the
-Fortran code and submits the job to the scheduler, and \texttt{\small go\_mesher\_solver\_lsf}~\\
-\texttt{\small .bash}, which contains the instructions that make up
-the job itself. These scripts can be found in \texttt{\small UTILS/}
-directory and can straightforwardly be modified and adapted to meet
-more specific running needs.
-
-\section{\texttt{run\_lsf.bash}}
+Fortran code and submits the job to the scheduler, and \texttt{\small go\_mesher\_solver\_lsf\_basin.forward}, 
+which contains the instructions that make up
+the job itself. These scripts can be found in \texttt{\small UTILS/Cluster/lsf/} directory
+
+\section{Job submission \texttt{run\_lsf.bash}}
 
 This script first sets the job queue to be `normal'. It then compiles
-the mesher and solver together, figures out the number of processors
+the mesher, database generator and solver together, figures out the number of processors
 required for this simulation from the \texttt{in\_data\_files/Par\_file}, and
@@
 
+make~meshfem3D~
+
 make~generate\_databases~
@@
 ~\\
-\#~compute~total~number~of~nodes~needed~
-
-NPROC=`grep~NPROC~in\_data\_files/Par\_file~|~cut~-c~34-~'~
-
+\#~get~total~number~of~nodes~needed~for~solver~
+
+NPROC=`grep~NPROC~in\_data\_files/Par\_file~|~cut~-c~34-~'~\\
+~\\
+
+\#~compute~total~number~of~nodes~needed~for~mesher~
+
+NPROC\_XI=`grep~NPROC\_XI~in\_data\_files/meshfem3D\_files/Mesh\_Par\_file~|~cut~-c~34-~'~
+~\\
+
+NPROC\_ETA=`grep~NPROC\_ETA~in\_data\_files/meshfem3D\_files/Mesh\_Par\_file~|~cut~-c~34-~'~
 ~\\
@@
 
-numnodes=\$((~\$NPROC))~\\
+numnodes=\$((~\$NPROC\_XI~*~\$NPROC\_ETA~))~\\
+~\\
+
+\#~checks~total~number~of~nodes~
+
+if~{[}~\$numnodes~-neq~\$NPROC~];~then
+
+~~~~~~~~echo~\char`\"{}error number of procs mismatch\char`\"{}
+
+~~~~~~~~exit~
+
+fi~\\
+
 ~\\
@@
 
-bsub~\$queue~-n~\$numnodes~-W~60~-K~<go\_mesher\_solver\_lsf.bash~
-\end{lyxcode}
-
-\section{\texttt{go\_mesher\_solver\_lsf.bash}}
+bsub~\$queue~-n~\$numnodes~-W~60~-K~<go\_mesher\_solver\_lsf.forward~
+\end{lyxcode}
+
+\section{Job script \texttt{go\_mesher\_solver\_lsf.forward}}
 
@@
 compute nodes to the job, the \texttt{run\_lsf.bash} commands used
-to run the mesher and the solver, and calls to scripts that collect
+to run the mesher, database generator and the solver, and calls to scripts that collect
 the output seismograms from the compute nodes and perform clean-up
@@
 parameter in \texttt{in\_data\_files/Par\_file} is also set properly.
-\item The next portion of the script launches the mesher and then the solver
+\item The next portion of the script launches the mesher, database generator and then the solver
 using \texttt{run\_lsf.bash}.
@@
 
-mkdir~-p~in\_out\_files/OUTPUT\_FILE
+mkdir~-p~in\_out\_files/OUTPUT\_FILES~
 
@@
 
-run\_lsf.bash~-{}-gm-no-shmem~-{}-gm-copy-env~\$current\_pwd/xmeshfem3D
+run\_lsf.bash~-{}-gm-no-shmem~-{}-gm-copy-env~\$current\_pwd/xmeshfem3D~\\
+
+run\_lsf.bash~-{}-gm-no-shmem~-{}-gm-copy-env~\$current\_pwd/xgenerate\_databases~\\
 
@@
 
-cd current\_pwd \\
+cd current\_pwd/~ \\
 
@@
 
-cd~in\_out\_files/SEM
+cd~in\_out\_files/SEM/~
 
@@
 systems, for example, the path of the executable programs. Here we
-only list the available scripts and provide a brief description, and
+only list a few of the available scripts and provide a brief description, and
 you can either refer to the related sections for detailed usage or,
@@
 
+\section{Process Data and Synthetics\label{sec:Process-data-and-syn}}
+
+In many cases, the SEM synthetics are calculated and compared to data
+seismograms recorded at seismic stations. Since the SEM synthetics
+are accurate for a certain frequency range, both the original data
+and the synthetics need to be processed before a comparison can be
+made. We generally use the following scripts provided in the ~\\
+\texttt{UTILS/seis\_process/} directory:
+
+\subsection{Data processing script \texttt{process\_trinet\_data.pl}}
+
+This script cuts a given portion of the original data, filters it,
+transfers the data into a displacement record, and picks the first
+P and S arrivals. For more functionality, type `\texttt{process\_trinet\_data.pl}'
+without any argument. An example of the usage of the script:
+
+\begin{lyxcode}
+{\footnotesize process\_trinet\_data.pl~-m~CMTSOLUTION~-l~0/180~-t~2/40~-i~dir~-p~-x~bp~~9703873{*}.BH?.SAC}{\footnotesize \par}
+\end{lyxcode}
+which has cut all the sac files between 0 and 180 seconds, filtered
+them between 2 and 40 seconds, transfered them into displacement records
+using the polezero files in \texttt{dir} directory, picked the first
+P and S arrivals, and added suffix `\texttt{bp}' to the file names.
+
+Note that all of the scripts in this section actually use the SAC
+and/or IASP91 to do the core operations; therefore make sure that
+the SAC and IASP91 packages are installed properly on your system,
+and that all the environment variables are set properly before running
+these scripts.
+
+\subsection{Synthetics processing script \texttt{process\_trinet\_syn.pl}}
+
+This script converts the synthetic output from the SEM code from ASCII
+to SAC format, and performs similar operations as `\texttt{process\_trinet\_data.pl}'.
+An example of the usage of the script:
+
+\begin{lyxcode}
+{\footnotesize process\_trinet\_syn.pl~-m~CMTSOLUTION~-a~STATIONS~-l~0/180~-t~2/40~-p~-x~bp~syn/{*}.BH?.semd}{\footnotesize \par}
+\end{lyxcode}
+which will convert the synthetics into SAC format, add event and station
+information into the SAC headers, cut the SAC files between 0 and
+180 seconds, filter them between 2 and 40 seconds, pick the first
+P and S arrivals, and add the suffix `\texttt{bp}' to the file names.
+
+More options are available for this script, such as adding time shift
+to the origin time of the synthetics, convolving the synthetics with
+a triangular source time function with a given half duration, etc.
+Type \texttt{process\_trinet\_syn.pl} without any argument for a detailed
+usage.
+
+\subsection{Script \texttt{rotate.pl}}
+
+The original data and synthetics have three components: vertical (BHZ),
+north (BHN) and east (BHE). However, for most seismology applications,
+transverse and radial components are also desirable. Therefore, we
+need to rotate the horizontal components of both the data and the
+synthetics to the transverse and radial direction, and \texttt{\small rotate.pl}
+can be used to accomplish this:
+
+\begin{lyxcode}
+rotate.pl~-l~0~-L~180~-d~DATA/{*}.BHE.SAC.bp~
+
+rotate.pl~-l~0~-L~180~SEM/{*}.BHE.semd.sac.bp~
+\end{lyxcode}
+where the first command performs rotation on the SAC data obtained
+through Seismogram Transfer Program (STP) \url{http://www.data.scec.org/STP/stp.html},
+while the second command rotates the processed SEM synthetics.
+
 \section{Collect Synthetic Seismograms}
@@
 The forward and adjoint simulations generate synthetic seismograms
-in the \texttt{LOCAL\_PATH}. For the forward simulation, the files
+in the \texttt{in\_out\_files/OUTPUT\_FILES/} directory by default. For the forward simulation, the files
 are named \texttt{STA.NT.BH?.semd} for two-column time series, or
@@
 
-To collect the synthetics onto the frontend, you can use the \texttt{UTILS/collect\_seismo.pl}
-machines script:
-
-\begin{lyxcode}
-collect\_seismo.pl~machines
-\end{lyxcode}
+You do have two options to change this default output behavior, given in the main constants file
+\texttt{constants.h} located in \texttt{src/shared/} directory: 
+\begin{description}
+\item [{\texttt{WRITE\_SEISMOGRAMS\_BY\_MASTER}}] Set to \texttt{.true.} to have only the master process writing out seismograms.
+This can be useful on a cluster, where only the master process node has access to the output directory.
+\item [{\texttt{USE\_OUTPUT\_FILES\_PATH}}] Set to \texttt{.false.} to have the seismograms output to \texttt{LOCAL\_PATH} directory specified in the main parameter file \texttt{in\_data\_files/Par\_file}.
+In this case, you could collect the synthetics onto the frontend using the \texttt{collect\_seismo\_lsf\_multi.pl} script
+located in the \texttt{UTILS/Cluster/lsf/} directory. The usage of the script would be e.g.:
+
+\begin{lyxcode}
+collect\_seismo.pl~machines~in\_data\_files/Par\_file
+\end{lyxcode}
+
+\end{description}
+where  \texttt{machines} is a file containing the node names and  \texttt{in\_data\_files/Par\_file} the parameter file used to extract the \texttt{LOCAL\_PATH} directory used for the simulation.
 
@@
 need to clean the local scratch disk for the next simulation. This
-is especially important in the case of 1- or 2-chunk kernel simulation,
+is especially important in the case of kernel simulation,
 where very large files are generated for the absorbing boundaries
@@
 \end{lyxcode}
-
-\section{Process Data and Synthetics\label{sec:Process-data-and-syn}}
-
-In many cases, the SEM synthetics are calculated and compared to data
-seismograms recorded at seismic stations. Since the SEM synthetics
-are accurate for a certain frequency range, both the original data
-and the synthetics need to be processed before a comparison can be
-made. We generally use the following scripts:
-
-\subsection{\texttt{process\_trinet\_data.pl}}
-
-This script cuts a given portion of the original data, filters it,
-transfers the data into a displacement record, and picks the first
-P and S arrivals. For more functionality, type `\texttt{process\_trinet\_data.pl}'
-without any argument. An example of the usage of the script:
-
-\begin{lyxcode}
-{\footnotesize process\_trinet\_data.pl~-m~CMTSOLUTION~-l~0/180~-t~2/40~-i~dir~-p~-x~bp~~9703873{*}.BH?.SAC}{\footnotesize \par}
-\end{lyxcode}
-which has cut all the sac files between 0 and 180 seconds, filtered
-them between 2 and 40 seconds, transfered them into displacement records
-using the polezero files in \texttt{dir} directory, picked the first
-P and S arrivals, and added suffix `\texttt{bp}' to the file names.
-
-Note that all of the scripts in this section actually use the SAC
-and/or IASP91 to do the core operations; therefore make sure that
-the SAC and IASP91 packages are installed properly on your system,
-and that all the environment variables are set properly before running
-these scripts.
-
-\subsection{\texttt{process\_trinet\_syn.pl}}
-
-This script converts the synthetic output from the SEM code from ASCII
-to SAC format, and performs similar operations as `\texttt{process\_trinet\_data.pl}'.
-An example of the usage of the script:
-
-\begin{lyxcode}
-{\footnotesize process\_trinet\_syn.pl~-m~CMTSOLUTION~-a~STATIONS~-l~0/180~-t~2/40~-p~-x~bp~syn/{*}.BH?.semd}{\footnotesize \par}
-\end{lyxcode}
-which will convert the synthetics into SAC format, add event and station
-information into the SAC headers, cut the SAC files between 0 and
-180 seconds, filter them between 2 and 40 seconds, pick the first
-P and S arrivals, and add the suffix `\texttt{bp}' to the file names.
-
-More options are available for this script, such as adding time shift
-to the origin time of the synthetics, convolving the synthetics with
-a triangular source time function with a given half duration, etc.
-Type \texttt{process\_trinet\_syn.pl} without any argument for a detailed
-usage.
-
-\subsection{\texttt{rotate.pl}}
-
-The original data and synthetics have three components: vertical (BHZ),
-north (BHN) and east (BHE). However, for most seismology applications,
-transverse and radial components are also desirable. Therefore, we
-need to rotate the horizontal components of both the data and the
-synthetics to the transverse and radial direction, and \texttt{\small rotate.pl}
-can be used to accomplish this:
-
-\begin{lyxcode}
-rotate.pl~-l~0~-L~180~-d~DATA/{*}.BHE.SAC.bp~
-
-rotate.pl~-l~0~-L~180~SEM/{*}.BHE.semd.sac.bp~
-\end{lyxcode}
-where the first command performs rotation on the SAC data obtained
-through Seismogram Transfer Program (STP) \url{http://www.data.scec.org/STP/stp.html},
-while the second command rotates the processed SEM synthetics.
+where  \texttt{machines} is a file containing the node names.
 

### 2010-11-12 doc/USER_MANUAL/manual_SPECFEM3D.tex
partially updated the list of authors

@@
 The SPECFEM3D package was first developed by Dimitri Komatitsch and Jeroen Tromp at Harvard University, then Caltech and Princeton, USA, starting in 1998, based in part on earlier work by Dimitri Komatitsch and Jean-Pierre Vilotte at IPG in Paris, France from 1994 to 1997.
-Since then it has been developed and maintained by a development team: in alphabetical order, C\'eline Blitz, Emiljana Jorgji, Dimitri Komatitsch, Nicolas Le Goff, Pieyre Le Loher, Roland Martin, David Mich\'ea, Jeroen Tromp \toall{add full list of developers here and update the AUTHORS file accordingly}.
+Since then it has been developed and maintained by a development team: in alphabetical order, Emiljana Jorgji, Dimitri Komatitsch, Nicolas Le Goff, Pieyre Le Loher, Roland Martin, David Mich\'ea, Jeroen Tromp \toall{add full list of developers here and update the AUTHORS file accordingly}.
 

### 2010-11-12 doc/USER_MANUAL/manual_SPECFEM3D.tex
second partial update of list of authors

@@
 The SPECFEM3D package was first developed by Dimitri Komatitsch and Jeroen Tromp at Harvard University, then Caltech and Princeton, USA, starting in 1998, based in part on earlier work by Dimitri Komatitsch and Jean-Pierre Vilotte at IPG in Paris, France from 1994 to 1997.
-Since then it has been developed and maintained by a development team: in alphabetical order, Emiljana Jorgji, Dimitri Komatitsch, Nicolas Le Goff, Pieyre Le Loher, Roland Martin, David Mich\'ea, Jeroen Tromp \toall{add full list of developers here and update the AUTHORS file accordingly}.
+Since then it has been developed and maintained by a development team: in alphabetical order,
+Emanuele Casarotti,
+Min Chen,
+Vala Hj\"orleifsd\'ottir,
+Emiljana Jorgji,
+Sue Kientz,
+Dimitri Komatitsch,
+Qinya Liu,
+Yang Luo,
+Nicolas Le Goff,
+Pieyre Le Loher,
+Alessia Maggi,
+Roland Martin,
+Dennis McRitchie,
+David Mich\'ea,
+Daniel Peter,
+Brian Savage,
+Leif Strand,
+Carl Tape,
+Jeroen Tromp
+\toall{add full list of developers here and update the AUTHORS file accordingly}.
 

### 2010-11-16 doc/USER_MANUAL/manual_SPECFEM3D.tex
added a note about the cover page, which will need to be updated
modifies SPECFEM3D attenuation to precompute needed arrays in xgenerate_databases; minor updates in manual

@@
 \toall{add full list of developers here and update the AUTHORS file accordingly}.
+
+\toall{We should also update the Adobe Illustrator / PDF cover page as well at some point}
 
@@
 \item the time step parameters \texttt{NSTEP} and \texttt{DT}
-\item the attenuation control parameters \texttt{ATTENUATION} and \texttt{USE\_OLSEN\_ATTENUATION}
+%\item the attenuation control parameters \texttt{ATTENUATION} and \texttt{USE\_OLSEN\_ATTENUATION}
 \item the absorbing boundary control parameter \texttt{ABSORBING\_CONDITIONS}
@@
 1} and \texttt{SAVE\_FORWARD = .false.}). You can automatically set
-these two variables using the \texttt{\small UTILS/change\_simulation\_type.pl}
+these two variables using the \texttt{\small utils/change\_simulation\_type.pl}
 script:
@@
 \begin{lyxcode}
-UTILS/change\_simulation\_type.pl~-f~
+utils/change\_simulation\_type.pl~-f~
 \end{lyxcode}
@@
 adjoint sources need to be computed from the original recorded seismograms
-for the selected stations and saved in the \texttt{SEM/} directory
+for the selected stations and saved in the \texttt{in\_out\_files/SEM/} directory
 with the format \texttt{STA.NT.BH?.adj}, where \texttt{STA}, \texttt{NT}
@@
 be \texttt{SIMULATION\_TYPE = 2} and \texttt{SAVE\_FORWARD = .false.}.
-More conveniently, use the scripts \texttt{UTILS/change\_simulation\_type.pl}
+More conveniently, use the scripts \texttt{utils/change\_simulation\_type.pl}
 to modify the \texttt{Par\_file} automatically (\texttt{change\_simulation\_type.pl
@@
 in the \texttt{STATIONS\_ADJOINT} list, we supply a sample program
-in \texttt{UTILS/xcut\_velocity} to cut a certain portion of the original
+in \texttt{utils/cut\_velocity/xcut\_velocity} to cut a certain portion of the original
 displacement seismograms and convert it into the proper adjoint source
@@
 With the successful forward simulation and the adjoint source ready
-in \texttt{SEM/}, set \texttt{SIMULATION\_TYPE = 3} and \texttt{SAVE\_FORWARD
+in the \texttt{in\_out\_files/SEM/} directory, set \texttt{SIMULATION\_TYPE = 3} and \texttt{SAVE\_FORWARD
 = .false.} in the \texttt{Par\_file(change\_simulation\_type.pl -b),}
@@
 
-A Perl script \texttt{UTILS/slice\_number.pl} that calls MATLAB can
+A Perl script \texttt{slice\_number.pl} located in directory \texttt{utils/Visualization/Paraview/} can
 help to figure out the slice numbers that lie along the great circle
-path (both the minor and major arcs), as well as the slice numbers
-required to produce a full picture of the inner core if your kernel
-also illuminates the inner core.
+path. It applies to meshes created with the internal mesher \texttt{xmeshfem3D}.
 
 \begin{enumerate}
-\item On machines where you have MATLAB access, copy the \texttt{CMTSOLUTION}
-file, \texttt{STATIONS\_ADJOINT}, and \texttt{Par\_file}, and run:
-
-\begin{lyxcode}
-{\small UTILS/slice\_number.pl~Par\_file~output\_solver.txt~slice\_file}{\small \par}
+\item On machines where you have access to the script, copy the \texttt{Mesh\_Par\_file}, and
+\texttt{output\_solver} files, and run:
+
+\begin{lyxcode}
+{\small slice\_number.pl~Mesh\_Par\_file~output\_solver.txt~slice\_file}{\small \par}
 \end{lyxcode}
@@
 \begin{enumerate}
-\item You can use or modify the script \texttt{UTILS/copy\_databases.pl}
+\item You can use or modify the script \texttt{utils/copy\_basin\_database.pl}
 to accomplish this:
@@
 \begin{lyxcode}
-{\small UTILS/copy\_database.pl~slice\_file~lsf\_machine\_file~filename~{[}jobid]}{\small \par}
+{\small utils/copy\_database.pl~slice\_file~lsf\_machine\_file~filename~{[}jobid]}{\small \par}
 \end{lyxcode}
@@
 
-We use an auxilliary program \texttt{combine\_paraview\_data.f90}
+We use an auxilliary program \texttt{combine\_vol\_data.f90}
 to combine the kernel files from all slices into one mesh file.
@@
 \begin{enumerate}
-\item Compile it in the global code directory:
-
-\begin{lyxcode}
-{\footnotesize make~combine\_paraview\_data~}{\footnotesize \par}
-
-{\footnotesize xcombine\_paraview\_data~slice\_list~filename~input\_dir~output\_dir~high/low-resolution~}{\footnotesize \par}
+\item Compile it in the root directory:
+
+\begin{lyxcode}
+{\footnotesize make~combine\_vol\_data~}{\footnotesize \par}
+
+{\footnotesize ./bin/xcombine\_vol\_data~slice\_list~filename~input\_dir~output\_dir~high/low-resolution~}{\footnotesize \par}
 \end{lyxcode}
@@
 \item We next convert the \texttt{.mesh} file into the VTU (Unstructured
-grid file) format which can be viewed in ParaView, for example:
-
-\begin{lyxcode}
-UTILS/mesh2vtu.pl~-i~file.mesh~-o~file.vtu
+grid file) format which can be viewed in ParaView. For this task, you can use and modify the script \texttt{mesh2vtu.pl}
+located in directory \texttt{utils/Visualization/Paraview/}, for example:
+
+\begin{lyxcode}
+mesh2vtu.pl~-i~file.mesh~-o~file.vtu
 \end{lyxcode}
 \item Notice that this Perl script uses a program \texttt{mesh2vtu} in the
-\texttt{UTILS/mesh2vtu} directory, which further uses the VTK \url{http://www.vtk.org/}
+\texttt{utils/Visualization/Paraview/mesh2vtu} directory, which further uses the VTK \url{http://www.vtk.org/}
 run-time library for its execution. Therefore, make sure you have
@@
 \end{itemize}
-Examples of job scripts can be found in \texttt{\small UTILS/Cluster/}
+Examples of job scripts can be found in the \texttt{\small utils/Cluster/}
 directory and can straightforwardly be modified and adapted to meet
@@
 which contains the instructions that make up
-the job itself. These scripts can be found in \texttt{\small UTILS/Cluster/lsf/} directory
+the job itself. These scripts can be found in \texttt{\small utils/Cluster/lsf/} directory
 
@@
 and parsing the output into a one-column list using the Perl script
-\texttt{UTILS/remap\_lsf\_machines.pl}. It then creates a set of scratch
+\texttt{utils/Cluster/lsf/remap\_lsf\_machines.pl}. It then creates a set of scratch
 directories on these nodes (\texttt{\small /scratch/}~\\
@@
 
-Several post-processing scripts/programs are provided in the \texttt{UTILS/}
+Several post-processing scripts/programs are provided in the \texttt{utils/}
 directory, and most of them need to be adjusted when used on different
@@
 made. We generally use the following scripts provided in the ~\\
-\texttt{UTILS/seis\_process/} directory:
+\texttt{utils/seis\_process/} directory:
 
@@
 In this case, you could collect the synthetics onto the frontend using the \texttt{collect\_seismo\_lsf\_multi.pl} script
-located in the \texttt{UTILS/Cluster/lsf/} directory. The usage of the script would be e.g.:
+located in the \texttt{utils/Cluster/lsf/} directory. The usage of the script would be e.g.:
 
@@
 to help with the reconstruction of the regular forward wavefield.
-A sample script is provided in \texttt{UTILS/}:
+A sample script is provided in \texttt{utils/}:
 

### 2010-11-17 doc/USER_MANUAL/manual_SPECFEM3D.tex
added examples for meshfem3D and updated the manual

@@
 \begin{itemize}
-     \item \texttt{Q}           : 0=no attenuation / standard \texttt{Q} attenuation value
+     \item \texttt{Q}           : quality factor
      \item \texttt{anisotropy\_flag}  : 0=no anisotropy / 1,2,.. check with implementation in \texttt{aniso\_model.f90}
-     \item \texttt{domain\_id}        : 1=acoustic / 2=elastic / 3=poroelastic
+     \item \texttt{domain\_id}        : 1=acoustic / 2=elastic 
 \end{itemize}
@@
 
-The file defined in \texttt{INTERFACES\_FILE} contains the settings of the topography grid
-and of the interfaces grids. The number
-of interfaces, including the topography, needs to be set at the first line. Then, from the bottom
-to the top of the model, you need to define the grids with several parameters :
- number of points along $x$ and $y$, minimal $x$ and $y$ coordinates, spacing between points and the file
-in which the elevation values are stored. At the end of this file, you simply need to set the
-number of spectral elements in the vertical direction for each layer.\\
+The file defined in the \texttt{INTERFACES\_FILE} parameter of \texttt{Mesh\_Par\_File} contains the settings of the topography grid
+and of the interfaces grids. The number of interfaces, including the topography, needs to be set at the first line. Then, from the bottom
+to the top of the model, you need to define the grids with several parameters : {\texttt{SUPPRESS\_UTM\_PROJECTION}
+as described previously, number of points along $x$ and $y$, minimal $x$ and $y$ coordinates, spacing between points and the name of the file 
+which contains the elevation values (in $y$.$x$ increasing order).  
+At the end of this file, you simply need to set the
+number of spectral elements in the vertical direction for each layer. We provide a few models in the {\texttt{examples/}} directory.  \\
 

### 2010-11-18 doc/USER_MANUAL/manual_SPECFEM3D.tex
updates SPECFEM3D manual and example readme files; bug fixes coupling surface detection for single layer cases

@@
 
-\author{$\copyright$ California Institute of Technology (USA) and\\
+\author{$\copyright$ Princeton University (USA) and\\
 University of Pau / CNRS / INRIA (France)\\
-Version 2.0.0}
+Version 2.0.0 \\
+"Sesame" 
+}
 
 \maketitle
-\newpage{}
-
-\tableofcontents{}
-
-\chapter{Introduction}
-
-The software package SPECFEM3D simulates seismic
-wave propagation at the local or regional scale
-based upon the spectral-element method (SEM).
-The SEM is a continuous Galerkin technique, which can easily be made discontinous \citep{BeMaPa94,KoWoHu02,ChCaVi03,Kop06,WiStBuGh10}; it is then a particular case of the discontinuous Galerkin technique \citep{ReHi73,FaRi99,HuHuRa99,CoKaSh00,GiHeWa02,RiWh03,MoRi05,GrScSc06,BeLaPi06,DuKa06,DeSeWh08,WiStBuGh10,DeSe10,EtChViGl10}, with optimized efficiency because of its tensorized basis functions.
-In particular, it can accurately handle very distorted mesh elements \citep{OlSe11}.
-Note that in many (most?) geological models in the context of seismic wave propagation studies
-(except for fault dynamic rupture studies)
-a discontinous mesh is not needed because material property contrasts are not drastic and thus a continuous formulation is sufficient. For a detailed introduction to the SEM as applied to regional
-seismic wave propagation, please consult \citet{KoVi98,KoTr99,ChKoViCaVaFe07,TrKoLi08} and
-in particular \citet{KoLiTrSuStSh04}.
-
-Effects due to lateral variations in compressional-wave speed, shear-wave
-speed, density, a 3D crustal model, topography and bathymetry are
-included. The package can accommodate
-full 21-parameter anisotropy (see~\citet{ChTr07}) as well as lateral
-variations in attenuation. Adjoint capabilities and finite-frequency
-kernel simulations are included \citep{LiTr06,TrKoLi08,FiIgBuKe09}.
-
+
+\section*{Authors}
 The SPECFEM3D package was first developed by Dimitri Komatitsch and Jeroen Tromp at Harvard University, then Caltech and Princeton, USA, starting in 1998, based in part on earlier work by Dimitri Komatitsch and Jean-Pierre Vilotte at IPG in Paris, France from 1994 to 1997.
+\\
+
+%Older versions of the code were initially developed by Dimitri Komatitsch at Institut de Physique du Globe (France)
+%and then by Dimitri Komatitsch and Jeroen Tromp at Harvard University (USA).
+
 Since then it has been developed and maintained by a development team: in alphabetical order,
@@
 Emiljana Jorgji,
+Jes\'us Labarta,
 Sue Kientz,
@@
 Carl Tape,
-Jeroen Tromp
-\toall{add full list of developers here and update the AUTHORS file accordingly}.
-
-\toall{We should also update the Adobe Illustrator / PDF cover page as well at some point}
-
-All SPECFEM3D\_GLOBE software is written in Fortran90 with full portability
+Jeroen Tromp.\\
+
+\toall{add full list of developers here and update the AUTHORS file accordingly} \\
+
+%The main developers of the \texttt{SPECFEM3D} source code are Dimitri
+%Komatitsch, Jeroen Tromp, and Qinya Liu. The following individuals
+%(listed in alphabetical order) have also contributed to the development
+%or improvement of the source code: Min Chen, Vala Hj\"orleifsd\'ottir,
+%Jes\'us Labarta, and Leif Strand. The following individuals (listed in alphabetical
+%order) contributed to this manual: Min Chen, Vala Hj\"orleifsd\'ottir,
+%Sue Kientz, Dimitri Komatitsch, Qinya Liu, Alessia Maggi,
+%Carl Tape, and Jeroen Tromp.
+
+The manual's cover graphic was created
+by Santiago Lombeyda from Caltech's Center for Advanced Computing Research (CACR) \url{http://www.cacr.caltech.edu}. \\
+
+\toall{We should also update the Adobe Illustrator / PDF cover page as well at some point} \\
+
+\newpage{}
+
+\tableofcontents{}
+
+\chapter{Introduction}
+
+The software package SPECFEM3D simulates seismic
+wave propagation at the local or regional scale
+based upon the spectral-element method (SEM).
+The SEM is a continuous Galerkin technique, which can easily be made discontinous \citep{BeMaPa94,KoWoHu02,ChCaVi03,Kop06,WiStBuGh10}; it is then a particular case of the discontinuous Galerkin technique \citep{ReHi73,FaRi99,HuHuRa99,CoKaSh00,GiHeWa02,RiWh03,MoRi05,GrScSc06,BeLaPi06,DuKa06,DeSeWh08,WiStBuGh10,DeSe10,EtChViGl10}, with optimized efficiency because of its tensorized basis functions.
+In particular, it can accurately handle very distorted mesh elements \citep{OlSe11}.
+Note that in most geological models in the context of seismic wave propagation studies
+(except for fault dynamic rupture studies)
+a discontinous mesh is not needed because material property contrasts are not drastic and thus a continuous formulation is sufficient. For a detailed introduction to the SEM as applied to regional
+seismic wave propagation, please consult \citet{KoVi98,KoTr99,ChKoViCaVaFe07,TrKoLi08} and
+in particular \citet{KoLiTrSuStSh04}.
+
+Effects due to lateral variations in compressional-wave speed, shear-wave
+speed, density, a 3D crustal model, topography and bathymetry are
+included. The package can accommodate
+full 21-parameter anisotropy (see~\citet{ChTr07}) as well as lateral
+variations in attenuation \citep{SaKoTr10}. Adjoint capabilities and finite-frequency
+kernel simulations are included \citep{LiTr06,TrKoLi08,FiIgBuKe09}.
+
+%The SPECFEM3D package was first developed by Dimitri Komatitsch and Jeroen Tromp at Harvard University, then Caltech and Princeton, USA, starting in 1998, based in part on earlier work by Dimitri Komatitsch and Jean-Pierre Vilotte at IPG in Paris, France from 1994 to 1997.
+Since then it has been developed and maintained by a development team: in alphabetical order,
+%Emanuele Casarotti,
+%Min Chen,
+%Vala Hj\"orleifsd\'ottir,
+%Emiljana Jorgji,
+%Sue Kientz,
+%Dimitri Komatitsch,
+%Qinya Liu,
+%Yang Luo,
+%Nicolas Le Goff,
+%Pieyre Le Loher,
+%Alessia Maggi,
+%Roland Martin,
+%Dennis McRitchie,
+%David Mich\'ea,
+%Daniel Peter,
+%Brian Savage,
+%Leif Strand,
+%Carl Tape,
+%Jeroen Tromp
+%\toall{add full list of developers here and update the AUTHORS file accordingly}.
+
+%\toall{We should also update the Adobe Illustrator / PDF cover page as well at some point}
+
+All SPECFEM3D software is written in Fortran90 with full portability
 in mind, and conforms strictly to the Fortran95 standard. It uses
@@
 If you use SPECFEM3D for your own research, please cite at least one
-of the following articles: \cite{TrKoLi08,VaCaSaKoVi99,LeChLiKoHuTr08,LeChKoHuTr09,LeKoHuTr09,KoMiEr09,KoErGoMi10,KoGoErMi10,WiKoScTr04,KoLiTrSuStSh04,ChKoViCaVaFe07,MaKoDi09,KoViCh10,CaKoLaTiMiLeSnTr08,TrKoHjLiZhPeBoMcFrTrHu10,KoRiTr02,KoTr02a,KoTr02b,KoTr99} or \cite{KoVi98}.
-If you work on geophysical applications, you may be interested in citing some of these application articles as well, among others: \cite{WiKoScTr04,JiTsKoTr05,KrJiKoTr06a,KrJiKoTr06b,LeChLiKoHuTr08,LeChKoHuTr09,LeKoHuTr09,ChFaKo04,FaChKo04,RiRiKoTrHe02,GoAmTaCaSmSaMaKo09,TrKo00,SaKoTr10}.
+of the following articles:
+
+%\cite{TrKoLi08,VaCaSaKoVi99,LeChLiKoHuTr08,LeChKoHuTr09,LeKoHuTr09,KoMiEr09,KoErGoMi10,KoGoErMi10,WiKoScTr04,KoLiTrSuStSh04,ChKoViCaVaFe07,MaKoDi09,KoViCh10,CaKoLaTiMiLeSnTr08,TrKoHjLiZhPeBoMcFrTrHu10,KoRiTr02,KoTr02a,KoTr02b,KoTr99} or \cite{KoVi98}.
+
+%If you work on geophysical applications, you may be interested in citing some of these application articles as well, among others: \cite{WiKoScTr04,JiTsKoTr05,KrJiKoTr06a,KrJiKoTr06b,LeChLiKoHuTr08,LeChKoHuTr09,LeKoHuTr09,ChFaKo04,FaChKo04,RiRiKoTrHe02,GoAmTaCaSmSaMaKo09,TrKo00,SaKoTr10}.
+
+\begin{description}
+
+\item [Forward simulations] ~\\
+\cite{TrKoLi08,VaCaSaKoVi99,KoMiEr09,KoErGoMi10,KoGoErMi10,ChKoViCaVaFe07,MaKoDi09,KoViCh10,CaKoLaTiMiLeSnTr08,TrKoHjLiZhPeBoMcFrTrHu10,KoRiTr02,KoTr02a,KoTr02b,KoTr99} or \cite{KoVi98}
+
+\item [Adjoint simulations] ~\\
+\cite{TrTaLi05,LiTr06,TrKoLi08,LiTr08,TrKoLi08,TrKoHjLiZhPeBoMcFrTrHu10}
+
+\item [Southern California simulations] ~\\
+\cite{KoLiTrSuStSh04,KrJiKoTr06a,KrJiKoTr06b}.
+
+If you use the 3D southern California model, please cite 
+\citet{SuSh03} (Los Angeles model), 
+\citet{lovelyetal06} (Salton Trough), and
+\citet{hauksson2000} (southern California).
+The Moho map was determined by 
+\citet{zhu&kanamori2000}. 
+The 1D SoCal model was developed by \citet{DrHe90}.
+
+\end{description}
+
+\noindent
+If you work on geophysical applications, you may be interested in citing some of these application articles as well, among others:
+
+\begin{description}
+
+\item [Anisotropy] ~\\
+\cite{ChTr07,JiTsKoTr05,ChFaKo04,FaChKo04,RiRiKoTrHe02,TrKo00}.
+
+\item [Attenuation] ~\\
+\cite{SaKoTr10,KoTr02a,KoTr99}.
+
+\item [Topography] ~\\
+\cite{LeKoHuTr09,LeChKoHuTr09,LeChLiKoHuTr08,GoAmTaCaSmSaMaKo09,WiKoScTr04}.
+
+\end{description}
+
 The corresponding Bib\TeX{} entries may be found
@@
 beginning of file \texttt{specfem3D.f90}.
-
-If you use the 3D southern
-California model, please cite \citet{SuSh03} (Los Angeles model), \citet{lovelyetal06}
-(Salton Trough), and \citet{hauksson2000} (southern California).
-The Moho map was determined by \citet{zhu&kanamori2000}. The 1D SoCal
-model was developed by \citet{DrHe90}.
 
@@
 
-The SPECFEM3D software package relies on the Scotch library to partition the meshes created with CUBIT.
+The SPECFEM3D software package relies on the Scotch library to partition meshes created with CUBIT.
 Note that we use CUBIT to create meshes of hexahedra, but other packages can be used as well, for instance GiD from \url{http://gid.cimne.upc.es} or Gmsh from \url{http://geuz.org/gmsh}.
@@
 \end{description}
-The configuration scripts automatically creates for each executable a corresponding
+The configuration script automatically creates for each executable a corresponding
 \texttt{Makefile} in the \texttt{src/} subdirectoy.
@@
 once in single precision and in double precision on your system and
-comparing the seismograms. If they are identical, you should probably
+then comparing the seismograms. If they are identical, you should probably
 select single precision for your future runs.
@@
 
-The first step in running a spectral-element method computation consists of constructing a high-quality mesh for the region under consideration.
-We provide two possibilities to do so: (1) relying on an external, hexahedral mesher CUBIT or (2) using the provided, internal mesher \texttt{xmeshfem3D}.
+The first step in running a spectral-element simulation consists of constructing a high-quality mesh for the region under consideration.
+We provide two possibilities to do so: (1) relying on the external, hexahedral mesher CUBIT, or (2) using the provided, internal mesher \texttt{xmeshfem3D}.
 In the following, we explain these two approaches.
@@
 \item [1.] setting up a hexahedral mesh with CUBIT,
-\item [2.] exporting the CUBIT mesh into a specfem3d file format and
-\item [3.] partitioning the specfem3d mesh files for a chosen number of cores.
+\item [2.] exporting the CUBIT mesh into a SPECFEM3D file format and
+\item [3.] partitioning the SPECFEM3D mesh files for a chosen number of cores.
 \end{description}
@@
 Examples are provided in the SPECFEM3D package in the subdirectory \texttt{examples/}.
-Please feel free to contribute your own example to this package by contacting the CIG Computational Seismology Mailing List  \url{cig-seismo@geodynamics.org}.
+We strongly encourage you to contribute your own example to this package by contacting the CIG Computational Seismology Mailing List  \url{cig-seismo@geodynamics.org}.
 
@@
 For the installation and handling of the CUBIT meshing tool suite, please refer to the CUBIT user manual and documentation.
-In order to give you a basic understanding of how to use CUBIT for our purposes, examples are provided in the SPECFEM3D package in the subdirectory \texttt{examples/}.
-For your convenience, the following examples are provided in the subdirectory \texttt{examples/}:
+In order to give you a basic understanding of how to use CUBIT for our purposes, examples are provided in the SPECFEM3D package in the subdirectory \texttt{examples/}:
 \begin{description}
-\item [{\texttt{homogeneous\_halfspace}}] Basic example creates a single block model and assigns elastic material parameters.
+\item [{\texttt{homogeneous\_halfspace}}] Creates a single block model and assigns elastic material parameters.
 \item [{\texttt{layered\_halfspace}}]  Combines two different, elastic material volumes and creates a refinement layer between the two.
@@
 
-\caption{Screenshots of the CUBIT examples provided in subdirectory  \texttt{examples/}: homogeneous halfspace (top-left), layered halfspace (top-right), waterlayered halfspace (bottom-left) and tomographic model (bottom-right).}
+\caption{Screenshots of the CUBIT examples provided in subdirectory  \texttt{examples/}: homogeneous halfspace (top-left), layered halfspace (top-right), water layered halfspace (bottom-left) and tomographic model (bottom-right).}
 
@@
 
-Once the geometric model volumes in CUBIT are meshed, you prepare the model for the exportation with the definition of the material blocks and of the boundary surfaces.
+Once the geometric model volumes in CUBIT are meshed, you prepare the model for exportation with the definition of material blocks and boundary surfaces.
 Thus, prior to exporting the mesh, you need to define blocks specifying the materials and absorbing boundaries in CUBIT.
@@
 \item [material\_id] An integer value which is unique for this material.
-\item [Vp] P-Velocity of the material (given in m/s).
-\item [Vs] S-Velocity of the material (given in m/s).
-\item [rho] density of the material (given in kg/$m^3$).
+\item [Vp] P-wave speed of the material (given in m/s).
+\item [Vs] S-wave speed of the material (given in m/s).
+\item [rho] density of the material (given in kg/m$^3$).
 \item [Q] quality factor to use in case of a simulation with attenuation turned on.
@@
 In case no attenuation information is available, it can be set to zero.
-Please note that your Vp- and Vs-velocities are given for a reference frequency.
+Please note that your Vp- and Vs-speeds are given for a reference frequency.
 To change this reference frequency, you change the value of
@@
 \item [material\_id] An integer value which is unique for this material.
-\item [Vp] P-Velocity of the material (given in m/s).
-\item [0] S-Velocity of the material is ignored.
-\item [rho] density of the material (given in kg/$m^3$).
+\item [Vp] P-wave speed of the material (given in m/s).
+\item [0] S-wave speed of the material is ignored.
+\item [rho] density of the material (given in kg/m$^3$).
 \end{description}
@@
 where
-\texttt{domain\_ID} is 1 for acoustic or 2 for elastic materials, \texttt{material\_ID} a unique identifier,  \texttt{rho} the density in $kg \, m^{-3}$,  \texttt{vp} the P-velocity in $m \, s^{-1}$, \texttt{vs}  the S-velocity in $m \, s^{-1}$, \texttt{Q} the quality factor and \texttt{anisotropy\_flag} an identifier for anisotropic models.
+\texttt{domain\_ID} is 1 for acoustic or 2 for elastic materials, \texttt{material\_ID} a unique identifier,  \texttt{rho} the density in $kg \, m^{-3}$,  \texttt{vp} the P-wave speed in $m \, s^{-1}$, \texttt{vs}  the S-wave speed in $m \, s^{-1}$, \texttt{Q} the quality factor and \texttt{anisotropy\_flag} an identifier for anisotropic models.
 For a tomographic model, the material definition format is:
@@
 
+These mesh files are needed as input files for the partitioner  \texttt{xdecompose\_mesh\_SCOTCH}
+to load-balance the mesh. Please see the next section for further details.
+
+\subsubsection*{Checking the mesh quality}
+The quality of the mesh may be inspected more precisely based upon the serial code
+in the file  ~\\
+\texttt{check\_mesh\_quality\_CUBIT\_Abaqus.f90}. 
+Running this code is optional because no information needed by the
+solver is generated.
+
+Prior to running and compiling this code, you have to export your mesh in CUBIT to an ABAQUS (.inp) format.
+You also have to determine a number of parameters of your mesh, such as the number of nodes and number of elements
+and modify the header of the \texttt{check\_mesh\_quality\_CUBIT\_Abaqus.f90} source file. Then, 
+in the directory \texttt{src/check\_mesh\_quality\_CUBIT\_Abaqus}, type
+\begin{lyxcode}
+{\small ./compile\_all.csh}{\small \par}
+\end{lyxcode}
+and use
+\begin{lyxcode}
+{\small xcheck\_mesh\_quality\_CUBIT\_Abaqus}{\small \par}
+\end{lyxcode}
+to generate an AVS output file (\texttt{\small AVS\_meshquality.inp}
+in AVS UCD format) or OpenDX output file (\texttt{\small DX\_meshquality.}~\\
+\texttt{\small dx}) that can be used to investigate mesh quality,
+e.g, skewness of elements and a Gnuplot histogram (\texttt{\small mesh\_quality\_}~\\
+\texttt{\small histogram.txt}) that can be plotted with gnuplot (type
+`\texttt{\small gnuplot plot\_mesh\_quality\_histogram.gnu}'). The
+histogram is also printed to the screen. If you want to start designing
+your own meshes, this tool is useful for viewing your creations. You
+are striving for meshes with elements with `cube-like' dimensions,
+e.g., the mesh should contain no very elongated or skewed elements.
+
 \subsection{Partitioning the Mesh with \texttt{xdecompose\_mesh\_SCOTCH}}
@@
 The SPECFEM3D software package performs large scale simulations in a parallel 'Single Process Multiple Data' way.
-The spectral element mesh created with CUBIT needs to be distributed on the processors. This partitioning is executed once and for all prior to
+The spectral-element mesh created with CUBIT needs to be distributed on the processors. This partitioning is executed once and for all prior to
 the execution of the solver so it is referred to as a static mapping.
@@
 An efficient partitioning is important because it leverages the overall running time of the application.
-It amounts to balance the number of elements in each slice while minimizing the communication coasts
-resulting from the placement of adjacent elements to different processors. \texttt{decompose\_mesh\_SCOTCH} depends on the Scotch library
+It amounts to balance the number of elements in each slice while minimizing the communication costs
+resulting from the placement of adjacent elements on different processors. \texttt{decompose\_mesh\_SCOTCH} depends on the Scotch library
 which provides efficent static mapping, graph and mesh partitioning routines. Scotch is a free software package developed by
@@
 \texttt{src/decompose\_mesh\_SCOTCH/scotch\_5.1.10b}. Please refer to file INSTALL.txt in that directory to see how to compile it.
-In the future you should be able to find for recent versions at  \url{http://www.labri.fr/perso/pelegrin/scotch/scotch\_en.html} .
+In the future you should be able to find more recent versions at  \url{http://www.labri.fr/perso/pelegrin/scotch/scotch\_en.html} .
 \\
@@
 
-Then you are ready to compile,
+When you are ready to compile,
 in the main directory type '\texttt{make decompose\_mesh\_SCOTCH}'.
@@
 \begin{itemize}
-     \item \texttt{nparts} is the number of partitions, i.e. number of cores for the parallel simulations,
+     \item \texttt{nparts} is the number of partitions, i.e., the number of cores for the parallel simulations,
      \item \texttt{input\_directory} is the directory which holds all the files generated by the Python script \texttt{cubit2specfem3d.py} explained in the previous Section~\ref{subsec:Exporting-the-Mesh}, e.g.
@@
 \item [{\texttt{SUPPRESS\_UTM\_PROJECTION}}] set to be \texttt{.false.}
-when your model range is specified in the geographical coordinates,
-and needs to be \texttt{.true.} when your model is specified in a
-cartesian coordinates. \noun{UTM projection zone in which your simulation
+when your model range is specified in geographical coordinates,
+and needs to be \texttt{.true.} when your model is specified in
+Cartesian coordinates. \noun{UTM projection zone in which your simulation
 region resides.}
-\item [{\texttt{INTERFACES\_FILE }}] File in which contains the description of the topography and
+\item [{\texttt{INTERFACES\_FILE }}] File which contains the description of the topography and
 of the interfaces between the different layers of the model, if any.
@@
 \begin{itemize}
-     \item \texttt{Q}           : quality factor
+     \item \texttt{Q}           : quality factor (0=no attenuation)
      \item \texttt{anisotropy\_flag}  : 0=no anisotropy / 1,2,.. check with implementation in \texttt{aniso\_model.f90}
@@
 
-The topography of the model is defined as a set of elevation
- values on a regular 2D grid. It is also posible to
+The \texttt{INTERFACES\_FILE} parameter of \texttt{Mesh\_Par\_File} defines the file which contains the settings of the topography grid and of the interfaces grids.
+Topography is defined as a set of elevation values on a regular 2D grid.
+It is also possible to
 define interfaces between the layers of the model in the same way.
-
-The file defined in the \texttt{INTERFACES\_FILE} parameter of \texttt{Mesh\_Par\_File} contains the settings of the topography grid
-and of the interfaces grids. The number of interfaces, including the topography, needs to be set at the first line. Then, from the bottom
-to the top of the model, you need to define the grids with several parameters : {\texttt{SUPPRESS\_UTM\_PROJECTION}
-as described previously, number of points along $x$ and $y$, minimal $x$ and $y$ coordinates, spacing between points and the name of the file 
-which contains the elevation values (in $y$.$x$ increasing order).  
-At the end of this file, you simply need to set the
-number of spectral elements in the vertical direction for each layer. We provide a few models in the {\texttt{examples/}} directory.  \\
+The file needs to define several parameters:
+\begin{itemize} 
+\item The number of interfaces, including the topography. This needs to be set at the first line. Then, from the bottom
+to the top of the model, you need to define the grids with: 
+\item \texttt{SUPPRESS\_UTM\_PROJECTION} flag as described previously, 
+\item number of points along $x$ and $y$ direction (NXI and NETA), 
+\item minimal $x$ and $y$ coordinates (LONG\_MIN and LAT\_MIN), 
+\item spacing between points along $x$ and $y$ (SPACING\_XI and SPACING\_ETA) and 
+\item the name of the file which contains the elevation values (in $y$.$x$ increasing order). 
+\end{itemize}
+At the end of this file, you simply need to set the number of spectral elements in the vertical direction for each layer. We provide a few models in the {\texttt{examples/}} directory.  \\
 
@@
 
-Now that you have set the appropriate parameters in the \texttt{Par\_file}
+Now that you have set the appropriate parameters in the \texttt{Mesh\_Par\_file}
 and have compiled the mesher, you are ready to launch it! This is
@@
 in \texttt{output\_mesher.txt}; this file provides lots of details
-about the mesh that was generated. Alternatively, output can be directed
+about the mesh that was generated. Please note that the mesher suggests a time step \texttt{DT} to run
+the solver with. The mesher output file also contains a table about the quality of the mesh to indicate
+possible problems with the distortions of elements. 
+Alternatively, output can be directed
 to the screen instead by uncommenting a line in \texttt{constants.h}:
-
 \begin{lyxcode}
@@
 
-The quality of the mesh may be inspected more precisely based upon the serial code
-in the file  ~\\
-\texttt{check\_mesh\_quality\_CUBIT\_Abaqus.f90}. In the directory \texttt{check\_mesh\_quality\_CUBIT\_Abaqus}, type
-
-\begin{lyxcode}
-make~check\_mesh\_quality\_CUBIT\_Abaqus
-\end{lyxcode}
-and then use
-
-\begin{lyxcode}
-xcheck\_mesh\_quality\_CUBIT\_Abaqus
-\end{lyxcode}
-to generate an AVS output file (\texttt{\small AVS\_meshquality.inp}
-in AVS UCD format) or OpenDX output file (\texttt{\small DX\_meshquality.}~\\
-\texttt{\small dx}) that can be used to investigate mesh quality,
-e.g, skewness of elements and a Gnuplot histogram (\texttt{\small mesh\_quality\_}~\\
-\texttt{\small histogram.txt}) that can be plotted with gnuplot (type
-`\texttt{\small gnuplot plot\_mesh\_quality\_histogram.gnu}'). The
-histogram is also printed to the screen. If you want to start designing
-your own meshes, this tool is useful for viewing your creations. You
-are striving for meshes with elements with `cube-like' dimensions,
-e.g., the mesh should contain no very elongated or skewed elements.
-
-Running this code is optional because no information needed by the
-solver is generated.
-
 \chapter{\label{cha:Creating-Distributed-Databases}Creating the Distributed Databases}
 
-Either you previously used  \texttt{xmeshfem3D} or \texttt{xdecompose\_mesh\_SCOTCH}, the next step in the workflow is to compile \texttt{xgenerate\_databases}.
+After using \texttt{xmeshfem3D} or \texttt{xdecompose\_mesh\_SCOTCH}, the next step in the workflow is to compile \texttt{xgenerate\_databases}.
 This program is going to create all the missing information needed by the SEM solver.
@@
 
-\caption{Schematic workflow for a SPECFEM3D simulation. The exectable \texttt{xgenerate\_databases} creates the GLL mesh points and assigns the specific model parameters.}
+\caption{Schematic workflow for a SPECFEM3D simulation. The exectable \texttt{xgenerate\_databases} creates the GLL mesh points and assigns specific model parameters.}
 
@@
 \item [{\texttt{USE\_OLSEN\_ATTENUATION}}] Set to \texttt{.true.} if you
-want to use the attenuation model that scaled from the velocity model
-using Olsen's empirical relation (reference).
+want to use the attenuation model that scaled from the S-wave speed model
+using Olsen's empirical relation (see \citet{OlDaBr03}).
 \item [{\texttt{ANISOTROPY}}] Set to \texttt{.true.} if you
@@
 \item [{\texttt{SAVE\_MESH\_FILES}}] Set this flag to \texttt{.true.} to
-save AVS \url{www.avs.com}, OpenDX \url{www.opendx.org}, or ParaView \url{www.paraview.org}
+save ParaView \url{www.paraview.org}
 mesh files for subsequent viewing. Turning the flag on generates large
@@
 GLL points \texttt{NGLL} specified in the main constants file \texttt{constants.h} located in the \texttt{src/shared/} subdirectory.
-The wave speed $v$ is determined based on your model's P- (or S-) velocity values.
+The wave speed $v$ is determined based on your model's P- (or S-) wave speed values.
 
@@
 in the \texttt{in\_out\_files/SEM/} directory, set \texttt{SIMULATION\_TYPE = 3} and \texttt{SAVE\_FORWARD
-= .false.} in the \texttt{Par\_file(change\_simulation\_type.pl -b),}
+= .false.} in the \texttt{Par\_file} (you can use \texttt{change\_simulation\_type.pl -b}),
 and rerun the solver.
@@
 
-Use the serial code \texttt{combine\_AVS\_DX.f90} (type `\texttt{make
-combine\_AVS\_DX}' and then `\texttt{xcombine\_AVS\_DX}') to generate
-AVS \url{www.avs.com} output files (in AVS UCD format) or OpenDX \url{www.opendx.org}
-output files showing the mesh, the MPI partition (slices), the $\nchunks$
-chunks, the source and receiver location, etc. Use the AVS UCD files
-\texttt{AVS\_continent\_boundaries.inp} and \texttt{AVS\_plate\_boundaries.inp}
-or the OpenDX files \texttt{DX\_continent\_boundaries.dx} and \texttt{DX\_plate\_boundaries.dx}
-for reference.
+In case you used the internal mesher \texttt{xmeshfem3D} to create and partition your mesh,
+you can output mesh files in ABAQUS (.INP) and DX (.dx) format to visualize them. For this, you must set either the
+flag \texttt{CREATE\_DX\_FILES} or \texttt{CREATE\_ABAQUS\_FILES} to \texttt{.true.}  
+in the mesher's parameter file \texttt{Mesh\_Par\_file}
+prior to running the mesher (see Chapter~\ref{cha:Running-the-Mesher-Meshfem3D} for details).
+You can then use AVS \url{www.avs.com} or OpenDX \url{www.opendx.org}
+to visualize the mesh and MPI partition (slices).
+
+% obsolete: only for global package...
+%Use the serial code \texttt{combine\_AVS\_DX.f90} (type `\texttt{make
+%combine\_AVS\_DX}' and then `\texttt{xcombine\_AVS\_DX}') to generate
+%AVS \url{www.avs.com} output files (in AVS UCD format) or OpenDX \url{www.opendx.org}
+%output files showing the mesh, the MPI partition (slices), the $\nchunks$
+%chunks, the source and receiver location, etc. 
+%Use the AVS UCD files \texttt{AVS\_continent\_boundaries.inp} and \texttt{AVS\_plate\_boundaries.inp}
+%or the OpenDX files \texttt{DX\_continent\_boundaries.dx} and \texttt{DX\_plate\_boundaries.dx}
+%for reference.
+
+\begin{figure}[htbp]
+\noindent \begin{centering}
+\includegraphics[width=.49\textwidth]{figures/vtk_mesh_vp.jpg}
+\includegraphics[width=.49\textwidth]{figures/vtk_mesh_vs.jpg}
+\par\end{centering}
+
+\caption{Visualization using Paraview of VTK files created by \texttt{xgenerate\_databases} showing P- and S-wave velocities assigned to the mesh points.
+The mesh was created by \texttt{xmeshfem3D} for 4 processors.}
+
+\label{fig:vtk.mesh}
+\end{figure}
+
+You have also the option to visualize the distributed databases produced by \texttt{xgenerate\_databases}
+using Paraview \url{www.paraview.org}. For this, you must set the flag \texttt{SAVE\_MESH\_FILES} to \texttt{.true.}
+in the main parameter file \texttt{Par\_file} (see Chapter~\ref{cha:Main-Parameter} for details).
+This will create VTK files for each single partition. You can then use Paraview \url{www.paraview.org} to visualized these partitions.
 
@@
 the \texttt{in\_out\_files/OUTPUT\_FILES} directory. The files are in a fairly complicated
-binary format, but there are two programs provided to convert the
-output into more user friendly formats. The first one, \texttt{create\_movie\_AVS\_DX.f90},
-outputs data in ASCII, OpenDX, AVS, or ParaView format. Run the code
-from the source directory (type `\texttt{make} \texttt{create\_movie\_AVS\_DX}'
-first) to create an input file in your format of choice. The code
-will prompt the user for input parameters. The second program \texttt{create\_movie\_GMT.f90}
-outputs ascii xyz files, convenient for use with GMT. This code uses
-significantly less memory than \texttt{create\_movie\_AVS\_DX.f90}
-and is therefore useful for high resolution runs.
+binary format, but there is a program provided to convert the
+output into more user friendly formats:
+
+\begin{description}
+\item [\texttt{xcreate\_movie\_shakemap\_AVS\_DX\_GMT}] 
+From \texttt{create\_movie\_shakemap\_AVS\_DX\_GMT.f90}, it
+outputs data in ASCII, OpenDX, or AVS format (also readable in ParaView). 
+Before compiling the code, make sure you have the file \texttt{surface\_from\_mesher.h}
+in the  \texttt{in\_out\_files/OUTPUT\_FILES/} directory. This file will be created by the solver run.
+Then type `\texttt{make} \texttt{create\_movie\_shakemap\_AVS\_DX\_GMT}'
+and run the executable \texttt{xcreate\_movie\_shakemap\_AVS\_DX\_GMT}
+in the \texttt{bin/} subdirectory. It will create visualization files in your format of choice. The code
+will prompt the user for input parameters.
+
+\begin{figure}[htbp]
+\noindent \begin{centering}
+\includegraphics[width=.32\textwidth]{figures/movie_surf_1.jpg}
+\includegraphics[width=.32\textwidth]{figures/movie_surf_2.jpg}
+\includegraphics[width=.32\textwidth]{figures/movie_surf_3.jpg}
+\par\end{centering}
+
+\caption{Visualization using AVS files created by \texttt{xcreate\_movie\_shakemap\_AVS\_DX\_GMT} showing 
+movie snapshots of vertical velocity components at different times.}
+
+\label{fig:movie.surf}
+\end{figure}
+
+% uses obsolete routines...
+%\item [\texttt{xcreate\_movie\_GMT}]
+%From \texttt{create\_movie\_GMT.f90},
+%outputs ascii xyz files, convenient for use with GMT. This code uses
+%significantly less memory than \texttt{create\_movie\_shakemap\_AVS\_DX\_GMT.f90}
+%and is therefore useful for high resolution runs.
+
+\end{description}
 
@@
 Linda earthquake. Lower Panel: Vertical source-receiver cross-section
-of the 3D S-wave velocity model used for the spectral-element simulations
+of the 3D S-wave speed model used for the spectral-element simulations
 \citep{KoLiTrSuStSh04}. (b) The same as (a) but for station HEC at
@@
 (\texttt{CREATE\_SHAKEMAP=.true.}), you can
-also run \texttt{`create\_movie\_GMT}' code to convert the binary
+also run \texttt{`create\_movie\_shakemap\_AVS\_DX\_GMT}' code to convert the binary
 shakemap data into GMT xyz files. A sample script \texttt{plot\_shakemap.pl}
@@
 
-The main developers of the \texttt{SPECFEM3D} source code are Dimitri
-Komatitsch, Jeroen Tromp, and Qinya Liu. The following individuals
-(listed in alphabetical order) have also contributed to the development
-or improvement of the source code: Min Chen, Vala Hj\"orleifsd\'ottir,
-Jes\'us Labarta, and Leif Strand. The following individuals (listed in alphabetical
-order) contributed to this manual: Min Chen, Vala Hj\"orleifsd\'ottir,
-Sue Kientz, Dimitri Komatitsch, Qinya Liu, Alessia Maggi,
-Carl Tape, and Jeroen Tromp. The manual's cover graphic was created
-by Santiago Lombeyda from Caltech's Center for Advanced Computing Research (CACR) \url{http://www.cacr.caltech.edu}.
-Older versions of the code were initially developed by Dimitri Komatitsch at Institut de Physique du Globe (France)
-and then by Dimitri Komatitsch and Jeroen Tromp at Harvard University (USA).
+%The main developers of the \texttt{SPECFEM3D} source code are Dimitri
+%Komatitsch, Jeroen Tromp, and Qinya Liu. The following individuals
+%(listed in alphabetical order) have also contributed to the development
+%or improvement of the source code: Min Chen, Vala Hj\"orleifsd\'ottir,
+%Jes\'us Labarta, and Leif Strand. The following individuals (listed in alphabetical
+%order) contributed to this manual: Min Chen, Vala Hj\"orleifsd\'ottir,
+%Sue Kientz, Dimitri Komatitsch, Qinya Liu, Alessia Maggi,
+%Carl Tape, and Jeroen Tromp. The manual's cover graphic was created
+%by Santiago Lombeyda from Caltech's Center for Advanced Computing Research (CACR) \url{http://www.cacr.caltech.edu}.
+%Older versions of the code were initially developed by Dimitri Komatitsch at Institut de Physique du Globe (France)
+%and then by Dimitri Komatitsch and Jeroen Tromp at Harvard University (USA).
 
@@
 
-Seismological Laboratory, California Institute of Technology, USA,
+Princeton University, USA,
 and University of Pau / CNRS / INRIA, France
 
-$\copyright$ California Institute of Technology and University of Pau / CNRS / INRIA, February 2010
+$\copyright$ Princeton University and University of Pau / CNRS / INRIA, February 2010
 

### 2010-11-19 doc/USER_MANUAL/manual_SPECFEM3D.tex
updated both manuals

@@
 
-%Older versions of the code were initially developed by Dimitri Komatitsch at Institut de Physique du Globe (France)
-%and then by Dimitri Komatitsch and Jeroen Tromp at Harvard University (USA).
-
 Since then it has been developed and maintained by a development team: in alphabetical order,
@@
 Jes\'us Labarta,
-Sue Kientz,
 Dimitri Komatitsch,
@@
 
-%The main developers of the \texttt{SPECFEM3D} source code are Dimitri
-%Komatitsch, Jeroen Tromp, and Qinya Liu. The following individuals
-%(listed in alphabetical order) have also contributed to the development
-%or improvement of the source code: Min Chen, Vala Hj\"orleifsd\'ottir,
-%Jes\'us Labarta, and Leif Strand. The following individuals (listed in alphabetical
-%order) contributed to this manual: Min Chen, Vala Hj\"orleifsd\'ottir,
-%Sue Kientz, Dimitri Komatitsch, Qinya Liu, Alessia Maggi,
-%Carl Tape, and Jeroen Tromp.
-
 The manual's cover graphic was created
-by Santiago Lombeyda from Caltech's Center for Advanced Computing Research (CACR) \url{http://www.cacr.caltech.edu}. \\
+by Santiago Lombeyda from Caltech's Center for Advanced Computing Research (CACR) \url{http://www.cacr.caltech.edu}.
+Sue Kientz assembled the manual.\\
 

### 2010-11-19 doc/USER_MANUAL/manual_SPECFEM3D.tex
Manual updated for noise simulations

@@
 
+%<YANGL
+\chapter{Noise Cross-correlation Simulations}
+
+Besides earthquake simulations, SPECFEM3D includes functionality for seismic noise tomography as well.
+In order to proceed successfully in this chapter, it is critical that you have already familiarized yourself
+with procedures for meshing (Chapter \ref{cha:Mesh-Generation}), 
+creating distributed databases (Chapter \ref{cha:Creating-Distributed-Databases}),
+running earthquake simulations (Chapters \ref{cha:Running-the-Solver}) 
+and adjoint simulations (Chapter \ref{cha:Adjoint-Simulations}).
+Also, make sure you read the paper `Noise cross-correlation sensitivity kernels' \citep{trompetal2010},
+in order to understand noise simulations from a theoretical perspective.
+
+\section{Input Parameter Files}
+
+As usual, the three main input files are crucial: \texttt{Par\_file}, \texttt{CMTSOLUTION} and \texttt{STATIONS}.
+Unless otherwise specified, those input files should be located in directory \texttt{in\_data\_files/}.\\
+
+\texttt{CMTSOLUTION} is required for all simulations.
+At a first glance, it may seem unexpected to have it here,
+since the noise simulations should have nothing to do with the earthquake -- \texttt{CMTSOLUTION}.
+However, for noise simulations, it is critical to have no earthquakes. 
+In other words, the moment tensor specified in \texttt{CMTSOLUTION} must be set to zero manually!\\
+
+\texttt{STATIONS} remains the same as in previous earthquake simulations,
+except that the order of receivers listed in \texttt{STATIONS} is now important.
+The order will be used to determine the `master' receiver,
+i.e., the one that simultaneously cross correlates with the others.\\
+
+\texttt{Par\_file} also requires careful attention.
+A parameter called \texttt{NOISE\_TOMOGRAPHY} has been added which specifies the type of simulation to be run.
+\texttt{NOISE\_TOMOGRAPHY} is an integer with possible values 0, 1, 2 and 3.
+For example, when \texttt{NOISE\_TOMOGRAPHY} equals 0, a regular earthquake simulation will be run.
+When it is 1/2/3, you are about to run
+step 1/2/3 of the noise simulations respectively.
+Should you be confused by the three steps, refer to \citet{trompetal2010} for details.\\
+
+Another change to \texttt{Par\_file} involves the parameter \texttt{NSTEP}.
+While for regular earthquake simulations this parameter specifies the length of synthetic seismograms generated,
+for noise simulations it specifies the length of the seismograms used to compute cross correlations.
+The actual cross correlations are thus twice this length, i.e., $2~\mathrm{NSTEP}-1$.
+The code automatically makes the modification accordingly, if \texttt{NOISE\_TOMOGRAPHY} is not zero.\\
+
+There are other parameters in \texttt{Par\_file} which should be given specific values.
+For instance, since the first two steps for calculating noise sensitivity kernels correspond to
+forward simulations, \texttt{SIMULATION\_TYPE} must be 1 when \texttt{NOISE\_TOMOGRAPHY} equals 1 or 2.
+Also, we have to reconstruct the ensemble forward wavefields in adjoint simulations,
+therefore we need to set \texttt{SAVE\_FORWARD} to \texttt{.true.} for the second step,
+i.e., when \texttt{NOISE\_TOMOGRAPHY} equals 2.
+The third step is for kernel constructions. Hence \texttt{SIMULATION\_TYPE} should be 3, whereas
+\texttt{SAVE\_FORWARD} must be \texttt{.false.}.\\
+
+Finally, for most system architectures,
+please make sure that \texttt{LOCAL\_PATH} in \texttt{Par\_file} is in fact local, not globally shared. 
+Because we have to save the wavefields at the earth's surface at every time step,
+it is quite problematic to have a globally shared \texttt{LOCAL\_PATH}, 
+in terms of both disk storage and I/O speed.
+
+\section{Noise Simulations: Step by Step}
+
+Proper parameters in those parameter files are not enough for noise simulations to run.
+We have more parameters to specify:
+for example, the ensemble-averaged noise spectrum, the noise distribution etc.
+However, since there are a few `new' files, it is better to introduce them sequentially.
+In this section, standard procedures for noise simulations are described.
+
+\subsection{Pre-simulation}
+
+\begin{itemize}
+
+\item
+As usual, we first configure the software package using:
+
+\texttt{./configure FC=ifort MPIFC=mpif90}
+
+Use the following if SCOTCH is needed:
+
+\texttt{./configure FC=ifort MPIFC=mpif90 --with-scotch-path=/opt/scotch}\\
+
+\item
+Next, we need to compile the source code using:
+
+\texttt{make generate\_databases}
+
+\texttt{make specfem3D} \\
+
+\item
+Before we can run noise simulations, we have to specify the noise statistics,
+e.g., the ensemble-averaged noise spectrum.
+Matlab scripts are provided to help you to generate the necessary file.
+
+\texttt{examples/noise\_tomography/NOISE\_TOMOGRAPHY.m}  (main program)\\
+\texttt{examples/noise\_tomography/PetersonNoiseModel.m}
+
+In Matlab, simply run:
+
+\texttt{NOISE\_TOMOGRAPHY(NSTEP, DT, Tmin, Tmax, NOISE\_MODEL)}\\
+
+\texttt{DT} is given in \texttt{Par\_file},
+but \texttt{NSTEP} is NOT the one specified in \texttt{Par\_file}.
+Instead, you have to feed $2~\mathrm{NSTEP}-1$ to account for the doubled length of cross correlations.
+\texttt{Tmin} and \texttt{Tmax} correspond to the period range you are interested in,
+whereas \texttt{NOISE\_MODEL} denotes the noise model you will be using.
+Details can be found in the Matlab script.
+
+After running the Matlab script, you will be given the following information (plus a figure in Matlab):
+
+\texttt{*************************************************************}\\
+\texttt{the source time function has been saved in:}\\
+\texttt{/data2/yangl/3D\_NOISE/S\_squared} (note this path must be different)\\
+\texttt{S\_squared should be put into directory:}\\
+\texttt{in\_out\_files/NOISE\_TOMOGRAPHY/ in the SPECFEM3D package}\\
+
+In other words, the Matlab script creates a file called \texttt{S\_squared},
+which is the first `new' input file we encounter for noise simulations.
+
+One may choose a flat noise spectrum rather than Peterson's noise model.
+This can be done easily by modifying the Matlab script a little.
+
+\item
+Create a new directory in the SPECFEM3D package, name it as \texttt{in\_out\_files/NOISE\_TOMOGRAPHY/}.
+We will add some parameter files later in this folder.
+
+\item
+Put the Matlab-generated-file \texttt{S\_squared} in \texttt{in\_out\_files/NOISE\_TOMOGRAPHY/}.
+
+That's to say, you will have a file \texttt{in\_out\_files/NOISE\_TOMOGRAPHY/S\_squared} in the SPECFEM3D package.
+
+\item
+Create a file called \texttt{in\_out\_files/NOISE\_TOMOGRAPHY/irec\_master\_noise}. Note that this file is located
+in directory \texttt{in\_out\_files/NOISE\_TOMOGRAPHY/} as well. In general, all noise simulation related parameter
+files go into that directory.
+\texttt{irec\_master\_noise} contains only one integer, which is the ID
+of the `master' receiver. For example, if this file contains 5, it means that the fifth receiver
+listed in \texttt{in\_data\_files/STATIONS} becomes the `master'. That's why we mentioned previously that
+the order of receivers in \texttt{in\_data\_files/STATIONS} is important.
+
+Note that in regional simulations, the \texttt{in\_data\_files/STATIONS}
+might contain receivers which are outside of our computational domains.
+Therefore, the integer in \texttt{irec\_master\_noise}
+is actually the ID in \texttt{in\_data\_files/STATIONS\_FILTERED} 
+(which is generated by \texttt{bin/xgenerate\_databases}).
+
+\item
+Create a file called \texttt{in\_out\_files/NOISE\_TOMOGRAPHY/nu\_master}. This file holds three numbers,
+forming a (unit) vector. It describes which component we are cross-correlating at the `master'
+receiver, i.e., ${\hat{\bf \nu}}^{\alpha}$ in \citet{trompetal2010}.
+The three numbers correspond to E/N/Z components respectively.
+Most often, the vertical component is used, and in those cases the three numbers should be 0, 0 and 1.
+
+\item
+Describe the noise direction and distributions in \texttt{src/shared/noise\_tomography.f90}.
+Search for a subroutine called \texttt{noise\_distribution\_direction}
+in \texttt{noise\_tomography.f90}. It is actually located at the very beginning
+of \texttt{noise\_tomography.f90}. The default assumes vertical noise and a uniform distribution
+across the whole free surface of the model. It should be quite self-explanatory for modifications.
+Should you modify this part, you have to re-compile the source code.
+
+\end{itemize}
+
+\subsection{Simulations}
+
+With all of the above done, we can finally launch our simulations.
+Again, please make sure that the \texttt{LOCAL\_PATH} in \texttt{in\_data\_files/Par\_file} is not globally shared.
+It is quite problematic to have a globally shared \texttt{LOCAL\_PATH}, in terms of both disk storage and
+speed of I/O (we have to save the wavefields at the earth's surface at every time step).
+
+As discussed in \citet{trompetal2010}, it takes three steps/simulations to obtain one contribution
+of the ensemble sensitivity kernels:
+
+\begin{itemize}
+\item
+Step 1: simulation for generating wavefields\\
+\texttt{SIMULATION\_TYPE=1}\\
+\texttt{NOISE\_TOMOGRAPHY=1}\\
+\texttt{SAVE\_FORWARD} not used, can be either .true. or .false.
+
+\item
+Step 2: simulation for ensemble forward wavefields\\
+\texttt{SIMULATION\_TYPE=1}\\
+\texttt{NOISE\_TOMOGRAPHY=2}\\
+\texttt{SAVE\_FORWARD=.true.}
+
+\item
+Step 3: simulation for ensemble adjoint wavefields and sensitivity kernels\\
+\texttt{SIMULATION\_TYPE=3}\\
+\texttt{NOISE\_TOMOGRAPHY=3}\\
+\texttt{SAVE\_FORWARD=.false.}
+
+Note Step 3 is an adjoint simulation, please refer to previous chapters on how to prepare adjoint sources
+and other necessary files, as well as how adjoint simulations work.
+
+\end{itemize}
+
+Note that it's better to run the three steps continuously within the same job, 
+otherwise you have to collect the saved surface movies from the old nodes/CPUs to the new nodes/CPUs.
+This process varies from cluster to cluster and thus cannot be discussed here.
+Please ask your cluster administrator for information/configuration of the cluster you are using.\\
+
+\subsection{Post-simulation}
+
+After those simulations, you have all stuff you need, either in the \texttt{in\_out\_files/OUTPUT\_FILES/} or
+in the directory specified by \texttt{LOCAL\_PATH} in \texttt{in\_data\_files/Par\_file}
+(which are most probably on local nodes).
+Collect whatever you want from the local nodes to your workstation, and then visualize them.
+This process is the same as what you may have done for regular earthquake simulations.
+Refer to other chapters if you have problems.\\
+
+Simply speaking, two outputs are the most interesting: the simulated ensemble cross correlations and
+one contribution of the ensemble sensitivity kernels.\\
+
+The simulated ensemble cross correlations are obtained after the second simulation (Step 2).
+Seismograms in \texttt{in\_out\_files/OUTPUT\_FILES/} are actually the simulated ensemble cross correlations.
+Collect them immediately after Step 2, or the Step 3 will overwrite them.
+Note that we have a `master' receiver specified by \texttt{irec\_master\_noise},
+the seismogram at one station corresponds to the cross correlation between that station and the `master'.
+Since the seismograms have three components, we may obtain cross correlations
+for different components as well, not necessarily the cross correlations between vertical components.\\
+
+One contribution of the ensemble cross-correlation sensitivity kernels are obtained after Step 3,
+stored in the \texttt{in\_data\_files/LOCAL\_PATH} on local nodes.
+The ensemble kernel files are named the same as regular earthquake kernels.\\
+
+You need to run another three simulations to get the other contribution of the ensemble kernels,
+using different forward and adjoint sources given in \citet{trompetal2010}.
+
+\section{Example}
+
+In order to illustrate noise simulations in an easy way, one example is provided in \texttt{examples/noise\_tomography/}.
+See \texttt{examples/noise\_tomography/README\_NOISE\_TOMOGRAPHY} for explanations. \\
+
+Note, however, that they are created for a specific workstation (CLOVER@PRINCETON),
+which has at least 4 cores with `mpif90' working properly. \\
+
+If your workstation is suitable, you can run the example in \texttt{examples/noise\_tomography/} using:\\
+
+\texttt{./pre-processing}\\
+
+Even if this script does not work on your workstation, the procedure it describes is universal.
+You may review the whole process described in the last section by following the commands in \texttt{pre-processing},
+which should contain enough explanations for all the commands.
+%>YANGL
+
 \chapter{Graphics}

### 2010-11-19 doc/USER_MANUAL/manual_SPECFEM3D.tex
sorted the authors alphabetically

@@
 Emiljana Jorgji,
+Dimitri Komatitsch,
 Jes\'us Labarta,
-Dimitri Komatitsch,
+Nicolas Le Goff,
+Pieyre Le Loher,
 Qinya Liu,
 Yang Luo,
-Nicolas Le Goff,
-Pieyre Le Loher,
 Alessia Maggi,

### 2010-11-19 doc/USER_MANUAL/manual_SPECFEM3D.tex
changed "manual's cover graphic" to "cover graphic of the manual"

@@
 
-The manual's cover graphic was created
+The cover graphic of the manual was created
 by Santiago Lombeyda from Caltech's Center for Advanced Computing Research (CACR) \url{http://www.cacr.caltech.edu}.

### 2010-11-21 doc/USER_MANUAL/manual_SPECFEM3D.tex
minor restructuring of examples files; adding flag USE_SOURCES_RECVS_Z in constants.h for locating source/receivers in randomly shaped meshes; updates some files to avoid compiler warnings and unifies character lengths for LOCAL_PATH; updates manual

@@
 Since then it has been developed and maintained by a development team: in alphabetical order,
+Piero Basini,
+Ebru Bozdag,
 Emanuele Casarotti,
 Min Chen,
+Hom Nath Gharti,
 Vala Hj\"orleifsd\'ottir,
 Emiljana Jorgji,
+Sue Kientz,
 Dimitri Komatitsch,
@@
 Alessia Maggi,
+Federica Magnoni,
 Roland Martin,
@@
 David Mich\'ea,
+Tarje Nissen-Meyer,
 Daniel Peter,
@@
 In order to illustrate noise simulations in an easy way, one example is provided in \texttt{examples/noise\_tomography/}.
-See \texttt{examples/noise\_tomography/README\_NOISE\_TOMOGRAPHY} for explanations. \\
+See \texttt{examples/noise\_tomography/README} for explanations. \\
 
@@
 
-\texttt{./pre-processing}\\
+\texttt{./pre-processing.sh}\\
 
 Even if this script does not work on your workstation, the procedure it describes is universal.
-You may review the whole process described in the last section by following the commands in \texttt{pre-processing},
+You may review the whole process described in the last section by following the commands in \texttt{pre-processing.sh},
 which should contain enough explanations for all the commands.
@@
 \end{itemize}
+
+The SPECFEM3D software code internally uses Cartesian coordinates.
+%This requires to convert the given locations for sources and receiver locations.
+Sources and receiver locations are read in from the
+\texttt{CMTSOLUTION} and \texttt{STATIONS} file.
+Note that e.g. the \texttt{CMTSOLUTION} file denotes the location by "longitude/latitude/depth". We read in
+longitude as $x$ coordinate, latitude as $y$ coordinate.
+In case the flag \texttt{SUPPRESS\_UTM\_PROJECTION} is set to \texttt{.false.}
+in the main parameter file (see Chapter~\ref{cha:Creating-Distributed-Databases}),
+the $x$/$y$ coordinates have to be given in degrees and are converted to Cartesian coordinates
+using a UTM conversion with the specified UTM zone.
+The value for depth (given in $km$ in \texttt{CMTSOLUTION}) or burial depth (given in $m$ in \texttt{STATIONS})
+is evaluated with respect to the surface of the mesh at the specified $x$/$y$ location to find the corresponding $z$ coordinate.
+It is possible to use this depth value directly as $z$ coordinate by
+changing the flag \texttt{USE\_SOURCES\_RECVS\_Z} to \texttt{.true.} in the file \texttt{constants.h} located in the \texttt{src/shared/} subdirectory.
+
+The seismogram output directions are given in Cartesian $x$/$y$/$z$ directions and not rotated any further.
+For the labeling of the seismograms, we name "$x$/$y$/$z$" to "BHE/BHN/BHZ" components.
+Changing flags in \texttt{constants.h} in the \texttt{src/shared/} subdirectory only rotates the seismogram outputs
+if receivers are forced to be located at the surface (\texttt{RECVS\_CAN\_BE\_BURIED\_EXT\_MESH} set to \texttt{.false.}) and the normal to the surface at the receiver location should be used (\texttt{EXT\_MESH\_RECV\_NORMAL} set to
+\texttt{.true.}) as vertical.
+In this case, the outputs are rotated to have the BHZ component normal to the surface of the mesh, BHE and BHN are somewhat arbitrary orthogonal directions along the surface.
+
+\chapter{\label{cha:Troubleshooting}Troubleshooting}
+
+\section*{FAQ}
+
+\begin{description}
+\item [configuration fails:]
+
+   Examine the log file 'config.log'. It contains detailed informations.
+   In many cases, the path's to these specific compiler commands F90,
+   CC and MPIF90 won't be correct if `./configure` fails.
+
+   Please make sure that you have a working installation of a Fortran compiler,
+   a C compiler and an MPI implementation. You should be able to compile this
+   little program code:
+
+\begin{lyxcode}
+
+      program main ~\\
+        include 'mpif.h' ~\\
+        integer, parameter :: CUSTOM\_MPI\_TYPE = MPI\_REAL ~\\
+        integer ier ~\\
+        call MPI\_INIT(ier) ~\\
+        call MPI\_BARRIER(MPI\_COMM\_WORLD,ier) ~\\
+        call MPI\_FINALIZE(ier) ~\\
+      end
+
+\end{lyxcode}
+
+\item [compilation fails stating:] ~\\
+\begin{lyxcode}
+    ...  ~\\
+    obj/program\_generate\_databases.o: In function `MAIN\_\_':  ~\\
+    program\_generate\_databases.f90:(.text+0x14): undefined reference to `\_gfortran\_set\_std'  ~\\
+    ...  ~\\
+\end{lyxcode}
+
+  Make sure you're pointing to the right 'mpif90' wrapper command.
+
+  Normally, this message will appear when you're mixing two different Fortran
+  compilers. That is, using e.g. gfortran to compile non-MPI files
+  and mpif90, wrapper provided for e.g. ifort, to compile MPI-files.
+
+  fix: e.g. specify > ./configure FC=gfortran MPIF90=/usr/local/openmpi-gfortran/bin/mpif90
+
+\end{description}
 

### 2010-11-21 examples/VALIDATION_3D_SEM_SIMPLER_LAYER_SOURCE_DEPTH/endw1.f

@@ -1,56 +1,0 @@
-c
-c-----------------------------------------------------------------------
-c
-      double precision function endw1 (n,alpha,beta)
-c
-c=======================================================================
-c
-c     E n d w 1 :
-c     ---------
-c
-c=======================================================================
-c
-      implicit double precision (a-h,o-z)
-c
-c---- remove above card for single precision operation
-c
-      parameter(zero=0.d0,one=1.d0,two=2.d0,three=3.d0,four=4.d0)
-c
-c-----------------------------------------------------------------------
-c
-      apb   = alpha+beta
-      if (n.eq.0) then
-         endw1 = zero
-         return
-      endif
-      f1   = gammaf(alpha+two)*gammaf(beta+one)/gammaf(apb+three)
-      f1   = f1*(apb+two)*two**(apb+two)/two
-      if (n.eq.1) then
-         endw1 = f1
-         return
-      endif
-      fint1 = gammaf(alpha+two)*gammaf(beta+one)/gammaf(apb+three)
-      fint1 = fint1*two**(apb+two)
-      fint2 = gammaf(alpha+two)*gammaf(beta+two)/gammaf(apb+four)
-      fint2 = fint2*two**(apb+three)
-      f2    = (-two*(beta+two)*fint1 + (apb+four)*fint2)
-     $        * (apb+three)/four
-      if (n.eq.2) then
-         endw1 = f2
-         return
-      endif
-      do 100 i=3,n
-         di   = dble(i-1)
-         abn  = alpha+beta+di
-         abnn = abn+di
-         a1   = -(two*(di+alpha)*(di+beta))/(abn*abnn*(abnn+one))
-         a2   =  (two*(alpha-beta))/(abnn*(abnn+two))
-         a3   =  (two*(abn+one))/((abnn+two)*(abnn+one))
-         f3   =  -(a2*f2+a1*f1)/a3
-         f1   = f2
-         f2   = f3
- 100  continue
-      endw1  = f3
-c
-      return
-      end

### 2010-11-21 examples/VALIDATION_3D_SEM_SIMPLER_LAYER_SOURCE_DEPTH/zwgjd.f

@@ -1,55 +1,0 @@
-      subroutine zwgjd (z,w,np,alpha,beta)
-c
-c=======================================================================
-c
-c     Z w g j d : Generate np Gauss Jacobi points and weights
-c                 associated with Jacobi polynomial of degree n = np-1
-c
-c=======================================================================
-c
-c     Note : Coefficients alpha and beta must be greater than -1.
-c     ----
-c
-c=======================================================================
-c
-      implicit double precision (a-h,o-z)
-c
-c---- remove above card for single precision operations
-c
-      parameter(one=1.d0,two=2.d0)
-c
-      dimension z(*),w(*)
-c
-c-----------------------------------------------------------------------
-c
-      n     = np-1
-      apb   = alpha+beta
-
-      if (np.le.0) stop 'Minimum number of Gauss points is 1'
-      if ((alpha.le.-one).or.(beta.le.-one))
-     .   stop 'Alpha and Beta must be greater than -1'
-
-      if (np.eq.1) then
-         z(1) = (beta-alpha)/(apb+two)
-         w(1) = gammaf(alpha+one)*gammaf(beta+one)/gammaf(apb+two)
-     .          * two**(apb+one)
-         return
-      endif
-
-      call jacg (z,np,alpha,beta)
-
-      np1   = n+1
-      np2   = n+2
-      dnp1  = dble(np1)
-      dnp2  = dble(np2)
-      fac1  = dnp1+alpha+beta+one
-      fac2  = fac1+dnp1
-      fac3  = fac2+one
-      fnorm = pnormj(np1,alpha,beta)
-      rcoef = (fnorm*fac2*fac3)/(two*fac1*dnp2)
-      do 100 i=1,np
-         call jacobf (p,pd,pm1,pdm1,pm2,pdm2,np2,alpha,beta,z(i))
-         w(i) = -rcoef/(p*pdm1)
- 100  continue
-      return
-      end

### 2010-11-21 doc/USER_MANUAL/manual_SPECFEM3D.tex
final version of the new covers (in Adobe Illustrator and in PDF formats); final version of the list of authors

@@
 
-% figures
-\usepackage[dvips]{epsfig}
-
 \usepackage{wrapfig}
@@
 \begin{document}
-\thispagestyle{empty}\textbf{}%
-\begin{figure}[H]
-\noindent \includegraphics[width=0.75\paperwidth]{figures/specfem_3d-cover.pdf}
-\end{figure}
-
+\begin{center}
+\thispagestyle{empty}%
+\vspace*{-1.8truecm}
+\noindent\makebox[\textwidth]{%
+\noindent\includegraphics[width=0.83\paperwidth]{figures/specfem_3d-cover.pdf}
+}
+\end{center}
+%
 \title{\textbf{SPECFEM3D}\\
@@
 Version 2.0.0 \\
-"Sesame"
-}
+"Sesame"}
+
+\date{\noindent \today}
 
@@
 \section*{Authors}
-The SPECFEM3D package was first developed by Dimitri Komatitsch and Jeroen Tromp at Harvard University, then Caltech and Princeton, USA, starting in 1998, based in part on earlier work by Dimitri Komatitsch and Jean-Pierre Vilotte at IPG in Paris, France from 1994 to 1997.
-\\
+The SPECFEM3D package was first developed by Dimitri Komatitsch and Jeroen Tromp at Harvard University and Caltech, USA, starting in 1998, based in part on earlier work by Dimitri Komatitsch and Jean-Pierre Vilotte at Institut de Physique du Globe (IPGP) in Paris, France from 1994 to 1997.\\
 
@@
 Piero Basini,
+C\'eline Blitz,
 Ebru Bozdag,
@@
 Min Chen,
-Hom Nath Gharti,
 Vala Hj\"orleifsd\'ottir,
@@
 David Mich\'ea,
+Hom Nath Gharti,
 Tarje Nissen-Meyer,
@@
 Jeroen Tromp.\\
-
-\toall{add full list of developers here and update the AUTHORS file accordingly} \\
+% \toall{add full list of developers here and update the AUTHORS file accordingly}
 
 The cover graphic of the manual was created
-by Santiago Lombeyda from Caltech's Center for Advanced Computing Research (CACR) \url{http://www.cacr.caltech.edu}.
-Sue Kientz assembled the manual.\\
-
-\toall{We should also update the Adobe Illustrator / PDF cover page as well at some point} \\
+by Santiago Lombeyda from Caltech's Center for Advanced Computing Research (CACR) \url{http://www.cacr.caltech.edu}.\\
 
@@
 in particular \citet{KoLiTrSuStSh04}.
+A detailed theoretical analysis of the dispersion
+and stability properties of the SEM is available in \citet{Coh02}, \citet{DeSe07} and \citet{SeOl07}.
 
@@
 variations in attenuation \citep{SaKoTr10}. Adjoint capabilities and finite-frequency
-kernel simulations are included \citep{LiTr06,TrKoLi08,FiIgBuKe09}.
-
-%The SPECFEM3D package was first developed by Dimitri Komatitsch and Jeroen Tromp at Harvard University, then Caltech and Princeton, USA, starting in 1998, based in part on earlier work by Dimitri Komatitsch and Jean-Pierre Vilotte at IPG in Paris, France from 1994 to 1997.
-Since then it has been developed and maintained by a development team: in alphabetical order,
-%Emanuele Casarotti,
-%Min Chen,
-%Vala Hj\"orleifsd\'ottir,
-%Emiljana Jorgji,
-%Sue Kientz,
-%Dimitri Komatitsch,
-%Qinya Liu,
-%Yang Luo,
-%Nicolas Le Goff,
-%Pieyre Le Loher,
-%Alessia Maggi,
-%Roland Martin,
-%Dennis McRitchie,
-%David Mich\'ea,
-%Daniel Peter,
-%Brian Savage,
-%Leif Strand,
-%Carl Tape,
-%Jeroen Tromp
-%\toall{add full list of developers here and update the AUTHORS file accordingly}.
-
-%\toall{We should also update the Adobe Illustrator / PDF cover page as well at some point}
+kernel simulations are included \citep{LiTr06,TrKoLi08,FiIgBuKe09,ViOp09}.
 
@@
 of the following articles:
-
-%\cite{TrKoLi08,VaCaSaKoVi99,LeChLiKoHuTr08,LeChKoHuTr09,LeKoHuTr09,KoMiEr09,KoErGoMi10,KoGoErMi10,WiKoScTr04,KoLiTrSuStSh04,ChKoViCaVaFe07,MaKoDi09,KoViCh10,CaKoLaTiMiLeSnTr08,TrKoHjLiZhPeBoMcFrTrHu10,KoRiTr02,KoTr02a,KoTr02b,KoTr99} or \cite{KoVi98}.
-
-%If you work on geophysical applications, you may be interested in citing some of these application articles as well, among others: \cite{WiKoScTr04,JiTsKoTr05,KrJiKoTr06a,KrJiKoTr06b,LeChLiKoHuTr08,LeChKoHuTr09,LeKoHuTr09,ChFaKo04,FaChKo04,RiRiKoTrHe02,GoAmTaCaSmSaMaKo09,TrKo00,SaKoTr10}.
 
@@
 Reintegration Grant No. MIRG-CT-2005-017461.
-Any opinions, findings,
-and conclusions or recommendations expressed in this material are
+Any opinions, findings, and conclusions or recommendations expressed in this material are
 those of the authors and do not necessarily reflect the views of the

### 2010-11-21 src/shared/combine_surf_data.f90
updated all the headers

@@ -1,11 +1,12 @@
 !=====================================================================
 !
-!               S p e c f e m 3 D  V e r s i o n  1 . 4
+!               S p e c f e m 3 D  V e r s i o n  2 . 0
 !               ---------------------------------------
 !
-!                 Dimitri Komatitsch and Jeroen Tromp
-!    Seismological Laboratory - California Institute of Technology
-!         (c) California Institute of Technology September 2006
+!          Main authors: Dimitri Komatitsch and Jeroen Tromp
+!                        Princeton University, USA
+! (c) Princeton University / California Institute of Technology and University of Pau / CNRS / INRIA
+!                            November 2010
 !
 ! This program is free software; you can redistribute it and/or modify
 ! it under the terms of the GNU General Public License as published by
@@ -50,7 +51,7 @@ program combine_surf_data
   integer, dimension(:,:,:,:),allocatable :: ibool
   logical, dimension(:),allocatable :: mask_ibool
   integer :: NSPEC_AB,NGLOB_AB
-  
+
   real x, y, z
   real, dimension(:,:,:,:), allocatable :: dat3D
   real, dimension(:,:,:), allocatable :: dat2D
@@ -60,7 +61,7 @@ program combine_surf_data
 
 !  integer :: num_ibool(NGLOB_AB)
   integer,dimension(:),allocatable :: num_ibool
-  
+
   logical :: HIGH_RESOLUTION_MESH,  FILE_ARRAY_IS_3D
   integer :: ires, nspec_surf, npoint1, npoint2, ispec_surf, inx, iny, idim
   integer,dimension(:), allocatable ::  ibelm_surf
@@ -160,15 +161,15 @@ program combine_surf_data
     open(unit=27,file=prname(1:len_trim(prname))//'external_mesh.bin',&
           status='old',action='read',form='unformatted',iostat=ios)
     read(27) NSPEC_AB
-    read(27) NGLOB_AB 
+    read(27) NGLOB_AB
     close(27)
     nspec = NSPEC_AB
-    nglob = NGLOB_AB    
-    
+    nglob = NGLOB_AB
+
     ! allocates arrays
-    allocate(ibool(NGLLX,NGLLY,NGLLZ,NSPEC_AB))    
[…]

### 2010-11-21 src/specfem3D/assemble_MPI_vector.f90

@@ -1,11 +1,12 @@
 !=====================================================================
 !
-!               S p e c f e m 3 D  V e r s i o n  1 . 4
+!               S p e c f e m 3 D  V e r s i o n  2 . 0
 !               ---------------------------------------
 !
-!                 Dimitri Komatitsch and Jeroen Tromp
-!    Seismological Laboratory - California Institute of Technology
-!         (c) California Institute of Technology September 2006
+!          Main authors: Dimitri Komatitsch and Jeroen Tromp
+!                        Princeton University, USA
+! (c) Princeton University / California Institute of Technology and University of Pau / CNRS / INRIA
+!                            November 2010
 !
 ! This program is free software; you can redistribute it and/or modify
 ! it under the terms of the GNU General Public License as published by
@@ -47,17 +48,17 @@
   integer, dimension(max_nibool_interfaces_ext_mesh,num_interfaces_ext_mesh) :: ibool_interfaces_ext_mesh
 
   ! local parameters
-  
+
   ! send/receive temporary buffers
   !real(kind=CUSTOM_REAL), dimension(NDIM,max_nibool_interfaces_ext_mesh,num_interfaces_ext_mesh) :: &
   !     buffer_send_vector_ext_mesh,buffer_recv_vector_ext_mesh
   real(kind=CUSTOM_REAL), dimension(:,:,:), allocatable :: buffer_send_vector_ext_mesh
-  real(kind=CUSTOM_REAL), dimension(:,:,:), allocatable :: buffer_recv_vector_ext_mesh  
+  real(kind=CUSTOM_REAL), dimension(:,:,:), allocatable :: buffer_recv_vector_ext_mesh
 
-  ! requests      
+  ! requests
   !integer, dimension(num_interfaces_ext_mesh) :: request_send_vector_ext_mesh,request_recv_vector_ext_mesh
   integer, dimension(:), allocatable :: request_send_vector_ext_mesh
-  integer, dimension(:), allocatable :: request_recv_vector_ext_mesh  
+  integer, dimension(:), allocatable :: request_recv_vector_ext_mesh
 
   integer ipoin,iinterface
 
@@ -71,7 +72,7 @@
     allocate(buffer_recv_vector_ext_mesh(NDIM,max_nibool_interfaces_ext_mesh,num_interfaces_ext_mesh))
     allocate(request_send_vector_ext_mesh(num_interfaces_ext_mesh))
[…]

### 2010-11-21 src/specfem3D/compute_coupling_elastic_ac.f90

@@ -1,11 +1,12 @@
 !=====================================================================
 !
-!               S p e c f e m 3 D  V e r s i o n  1 . 4
+!               S p e c f e m 3 D  V e r s i o n  2 . 0
 !               ---------------------------------------
 !
-!                 Dimitri Komatitsch and Jeroen Tromp
-!    Seismological Laboratory - California Institute of Technology
-!         (c) California Institute of Technology September 2006
+!          Main authors: Dimitri Komatitsch and Jeroen Tromp
+!                        Princeton University, USA
+! (c) Princeton University / California Institute of Technology and University of Pau / CNRS / INRIA
+!                            November 2010
 !
 ! This program is free software; you can redistribute it and/or modify
 ! it under the terms of the GNU General Public License as published by
@@ -33,7 +34,7 @@
                         coupling_ac_el_jacobian2Dw, &
                         ispec_is_inner,phase_is_inner)
 
-! returns the updated acceleration array: accel                        
+! returns the updated acceleration array: accel
 
   implicit none
   include 'constants.h'
@@ -43,16 +44,16 @@
 ! displacement and pressure
   real(kind=CUSTOM_REAL), dimension(NDIM,NGLOB_AB) :: accel
   real(kind=CUSTOM_REAL), dimension(NGLOB_AB) :: potential_dot_dot_acoustic
-  
+
 ! global indexing
   integer, dimension(NGLLX,NGLLY,NGLLZ,NSPEC_AB) :: ibool
 
 ! acoustic-elastic coupling surface
   integer :: num_coupling_ac_el_faces
-  real(kind=CUSTOM_REAL) :: coupling_ac_el_normal(NDIM,NGLLSQUARE,num_coupling_ac_el_faces) 
+  real(kind=CUSTOM_REAL) :: coupling_ac_el_normal(NDIM,NGLLSQUARE,num_coupling_ac_el_faces)
   real(kind=CUSTOM_REAL) :: coupling_ac_el_jacobian2Dw(NGLLSQUARE,num_coupling_ac_el_faces)
   integer :: coupling_ac_el_ijk(3,NGLLSQUARE,num_coupling_ac_el_faces)
-  integer :: coupling_ac_el_ispec(num_coupling_ac_el_faces)   
+  integer :: coupling_ac_el_ispec(num_coupling_ac_el_faces)
 
 ! communication overlap
[…]

### 2010-11-21 src/specfem3D/write_movie_output.f90

@@ -1,11 +1,12 @@
 !=====================================================================
 !
-!               S p e c f e m 3 D  V e r s i o n  1 . 4
+!               S p e c f e m 3 D  V e r s i o n  2 . 0
 !               ---------------------------------------
 !
-!                 Dimitri Komatitsch and Jeroen Tromp
-!    Seismological Laboratory - California Institute of Technology
-!         (c) California Institute of Technology September 2006
+!          Main authors: Dimitri Komatitsch and Jeroen Tromp
+!                        Princeton University, USA
+! (c) Princeton University / California Institute of Technology and University of Pau / CNRS / INRIA
+!                            November 2010
 !
 ! This program is free software; you can redistribute it and/or modify
 ! it under the terms of the GNU General Public License as published by
@@ -28,13 +29,13 @@
   subroutine write_movie_output()
 
   use specfem_par
-  use specfem_par_movie  
+  use specfem_par_movie
   implicit none
 
   ! shakemap creation
   if (EXTERNAL_MESH_CREATE_SHAKEMAP) then
     call wmo_create_shakemap_em()
-  endif 
+  endif
 
   ! movie file creation
   if(EXTERNAL_MESH_MOVIE_SURFACE .and. mod(it,NTSTEP_BETWEEN_FRAMES) == 0) then
@@ -62,21 +63,21 @@
   endif
 
   end subroutine write_movie_output
-  
-  
-  
+
+
+
 !================================================================
-  
+
   subroutine wmo_create_shakemap_em()
 
 ! creation of shapemap file
-  
+
   use specfem_par
   use specfem_par_elastic
   use specfem_par_acoustic
   use specfem_par_movie
   implicit none
-  
+
   real(kind=CUSTOM_REAL),dimension(NDIM,NGLLX,NGLLY,NGLLZ):: &
     displ_element,veloc_element,accel_element
   integer :: ipoin,ispec,iglob,ispec2D
@@ -103,7 +104,7 @@
           ! x,y,z coordinates
           store_val_x_external_mesh(NGNOD2D*(ispec2D-1)+ipoin) = xstore(iglob)
           store_val_y_external_mesh(NGNOD2D*(ispec2D-1)+ipoin) = ystore(iglob)
-          store_val_z_external_mesh(NGNOD2D*(ispec2D-1)+ipoin) = zstore(iglob)        
[…]

### 2010-11-21 utils/files_needed_asteroid/constants.h

@@ -1,10 +1,10 @@
 !=====================================================================
 !
-!               S p e c f e m 3 D  V e r s i o n  1 . 4
+!               S p e c f e m 3 D  V e r s i o n  2 . 0
 !               ---------------------------------------
 !
-!                 Dimitri Komatitsch and Jeroen Tromp
-!    Seismological Laboratory - California Institute of Technology
+!          Main authors: Dimitri Komatitsch and Jeroen Tromp
+!                        Princeton University, USA
 !         (c) California Institute of Technology July 2005
 !
 ! This program is free software; you can redistribute it and/or modify
@@ -126,7 +126,7 @@
   logical, parameter :: FASTER_RECEIVERS_POINTS_ONLY = .true.
   logical, parameter :: FASTER_SOURCES_POINTS_ONLY = .true.
 
-! the receivers can be located inside the model 
+! the receivers can be located inside the model
   logical, parameter :: RECVS_CAN_BE_BURIED_EXT_MESH = .false.
   logical, parameter :: SOURCES_CAN_BE_BURIED_EXT_MESH = .false.
 

### 2010-11-21 src/shared/parallel.f90
updated all the headers

@@ -1,11 +1,12 @@
 !=====================================================================
 !
-!               S p e c f e m 3 D  V e r s i o n  1 . 4
+!               S p e c f e m 3 D  V e r s i o n  2 . 0
 !               ---------------------------------------
 !
-!                 Dimitri Komatitsch and Jeroen Tromp
-!    Seismological Laboratory - California Institute of Technology
-!         (c) California Institute of Technology September 2006
+!          Main authors: Dimitri Komatitsch and Jeroen Tromp
+!                        Princeton University, USA
+! (c) Princeton University / California Institute of Technology and University of Pau / CNRS / INRIA
+!                            November 2010
 !
 ! This program is free software; you can redistribute it and/or modify
 ! it under the terms of the GNU General Public License as published by
@@ -105,8 +106,8 @@
 
 ! standard include of the MPI library
   include 'mpif.h'
-  
-  include "constants.h"  
+
+  include "constants.h"
   include "precision.h"
 
   integer count
@@ -797,14 +798,14 @@
 
 ! standard include of the MPI library
   include 'mpif.h'
-  
+
   integer dest,recvtag
   integer recvcount
   !integer recvbuf
   integer,dimension(recvcount):: recvbuf
   integer req(MPI_STATUS_SIZE)
   integer ier
-  
+
   call MPI_RECV(recvbuf,recvcount,MPI_INTEGER,dest,recvtag,MPI_COMM_WORLD,req,ier)
 
   end subroutine recv_i
@@ -819,15 +820,15 @@
 
 ! standard include of the MPI library
   include 'mpif.h'
-  
+
   include "constants.h"
   include "precision.h"
-  
+
   integer recvcount,dest,recvtag
   real(kind=CUSTOM_REAL),dimension(recvcount) :: recvbuf
   integer req(MPI_STATUS_SIZE)
   integer ier
-  
+
   call MPI_RECV(recvbuf,recvcount,CUSTOM_MPI_TYPE,dest,recvtag,MPI_COMM_WORLD,req,ier)
 
 
@@ -844,13 +845,13 @@
 
 ! standard include of the MPI library
   include 'mpif.h'
-  
+
   !integer sendbuf,sendcount,dest,sendtag
   integer dest,sendtag
   integer sendcount
   integer,dimension(sendcount):: sendbuf
   integer ier
-  
+
[…]

### 2010-11-21 src/specfem3D/initialize_simulation.f90

@@ -1,11 +1,12 @@
 !=====================================================================
 !
-!               S p e c f e m 3 D  V e r s i o n  1 . 4
+!               S p e c f e m 3 D  V e r s i o n  2 . 0
 !               ---------------------------------------
 !
-!                 Dimitri Komatitsch and Jeroen Tromp
-!    Seismological Laboratory - California Institute of Technology
-!         (c) California Institute of Technology September 2006
+!          Main authors: Dimitri Komatitsch and Jeroen Tromp
+!                        Princeton University, USA
+! (c) Princeton University / California Institute of Technology and University of Pau / CNRS / INRIA
+!                            November 2010
 !
 ! This program is free software; you can redistribute it and/or modify
 ! it under the terms of the GNU General Public License as published by
@@ -30,12 +31,12 @@
   use specfem_par
   use specfem_par_elastic
   use specfem_par_acoustic
-  use specfem_par_poroelastic  
+  use specfem_par_poroelastic
   use specfem_par_movie
   implicit none
-  
+
   integer :: ier
-  
+
   ! read the parameter file
   call read_parameter_file( NPROC,NTSTEP_BETWEEN_OUTPUT_SEISMOS,NSTEP,DT, &
                         UTM_PROJECTION_ZONE,SUPPRESS_UTM_PROJECTION, &
@@ -57,7 +58,7 @@
 
   ! checks flags
   call initialize_simulation_check()
-   
+
   ! open main output file, only written to by process 0
   if(myrank == 0 .and. IMAIN /= ISTANDARD_OUTPUT) &
     open(unit=IMAIN,file=trim(OUTPUT_FILES)//'/output_solver.txt',status='unknown')
@@ -94,7 +95,7 @@
     write(IMAIN,*)
   endif
 
-  ! reads in numbers of spectral elements and points for this process' domain  
+  ! reads in numbers of spectral elements and points for this process' domain
   call create_name_database(prname,myrank,LOCAL_PATH)
[…]

### 2010-11-22 doc/USER_MANUAL/manual_SPECFEM3D.tex
updated all the headers
final modifications to the manual; changed "Harvard" to "Princeton" on the cover; added a sentence about conforming mesh doublings

@@
 \usepackage{geometry}
-\geometry{verbose,letterpaper,tmargin=1in,bmargin=1in,lmargin=1in,rmargin=1in}
+\geometry{letterpaper,tmargin=1in,bmargin=1in,lmargin=1in,rmargin=1in}
 \setcounter{secnumdepth}{3}
@@
 (except for fault dynamic rupture studies)
-a discontinous mesh is not needed because material property contrasts are not drastic and thus a continuous formulation is sufficient. For a detailed introduction to the SEM as applied to regional
+a discontinous mesh is not needed because material property contrasts are not drastic and thus a continuous formulation is sufficient;
+conforming mesh doubling bricks can efficiently handle mesh size variations \citep{KoTr02a,KoLiTrSuStSh04}.\\
+
+For a detailed introduction to the SEM as applied to regional
 seismic wave propagation, please consult \citet{KoVi98,KoTr99,ChKoViCaVaFe07,TrKoLi08} and
@@
 A detailed theoretical analysis of the dispersion
-and stability properties of the SEM is available in \citet{Coh02}, \citet{DeSe07} and \citet{SeOl07}.
+and stability properties of the SEM is available in \citet{Coh02}, \citet{DeSe07} and \citet{SeOl07}.\\
 
@@
 variations in attenuation \citep{SaKoTr10}. Adjoint capabilities and finite-frequency
-kernel simulations are included \citep{LiTr06,TrKoLi08,FiIgBuKe09,ViOp09}.
+kernel simulations are included \citep{LiTr06,TrKoLi08,FiIgBuKe09,ViOp09}.\\
 
@@
 parallel programming based upon the Message Passing Interface (MPI)
-\citep{GrLuSk94,Pac97}.
+\citep{GrLuSk94,Pac97}.\\
 
@@
 It was a finalist again in 2008 for a run at 0.16 petaflops (sustained) on 149,784 processors of the `Jaguar' Cray XT5 system at Oak Ridge National Laboratories (USA) \citep{CaKoLaTiMiLeSnTr08}.
-It also won the BULL Joseph Fourier supercomputing award in 2010.
+It also won the BULL Joseph Fourier supercomputing award in 2010.\\
 
@@
 
-\item [Forward simulations] ~\\
-\cite{TrKoLi08,VaCaSaKoVi99,KoMiEr09,KoErGoMi10,KoGoErMi10,ChKoViCaVaFe07,MaKoDi09,KoViCh10,CaKoLaTiMiLeSnTr08,TrKoHjLiZhPeBoMcFrTrHu10,KoRiTr02,KoTr02a,KoTr02b,KoTr99} or \cite{KoVi98}
-
-\item [Adjoint simulations] ~\\
-\cite{TrTaLi05,LiTr06,TrKoLi08,LiTr08,TrKoLi08,TrKoHjLiZhPeBoMcFrTrHu10}
+\item [Numerical simulations in general] ~\\
+Forward simulations are described in detail in \cite{TrKoLi08,VaCaSaKoVi99,KoMiEr09,KoErGoMi10,KoGoErMi10,ChKoViCaVaFe07,MaKoDi09,KoViCh10,CaKoLaTiMiLeSnTr08,TrKoHjLiZhPeBoMcFrTrHu10,KoRiTr02,KoTr02a,KoTr02b,KoTr99} or \cite{KoVi98}.
+Additional aspects dealing with adjoint simulations are described in
+\cite{TrTaLi05,LiTr06,TrKoLi08,LiTr08,TrKoHjLiZhPeBoMcFrTrHu10}.
+
+\end{description}
+
+\noindent
+If you work on geophysical applications, you may be interested in citing some of these application articles as well, among others:
+
+\begin{description}
 
@@
 
-\end{description}
-
-\noindent
-If you work on geophysical applications, you may be interested in citing some of these application articles as well, among others:
-
-\begin{description}
-
 \item [Anisotropy] ~\\
@@
 The corresponding Bib\TeX{} entries may be found
-in file \texttt{USER\_MANUAL/bibliography.bib} or in comments at the
-beginning of file \texttt{specfem3D.f90}.
+in file \texttt{doc/USER\_MANUAL/bibliography.bib}.
 
@@
 
-$\copyright$ Princeton University and University of Pau / CNRS / INRIA, February 2010
+$\copyright$ Princeton University and University of Pau / CNRS / INRIA, November 2010
 

### 2010-11-22 src/specfem3D/compute_interpolated_dva.f90
adapts IRIS name convention for seismograms; updates seis_process/ scripts; updates manual

@@ -119,7 +119,8 @@ end subroutine compute_interpolated_dva
 !
 
 subroutine compute_interpolated_dva_ac(displ_element,veloc_element,&
-                        potential_dot_dot_acoustic,NGLOB_AB, &
+                        potential_dot_dot_acoustic,potential_dot_acoustic,&
+                        potential_acoustic,NGLOB_AB, &
                         ispec,NSPEC_AB,ibool, &
                         xi_r,eta_r,gamma_r, &
                         hxir,hetar,hgammar, &
@@ -138,6 +139,8 @@ subroutine compute_interpolated_dva_ac(displ_element,veloc_element,&
   integer :: NSPEC_AB,NGLOB_AB
   real(kind=CUSTOM_REAL),dimension(NDIM,NGLLX,NGLLY,NGLLZ):: displ_element,veloc_element
   real(kind=CUSTOM_REAL),dimension(NGLOB_AB) :: potential_dot_dot_acoustic
+  real(kind=CUSTOM_REAL),dimension(NGLOB_AB) :: potential_dot_acoustic
+  real(kind=CUSTOM_REAL),dimension(NGLOB_AB) :: potential_acoustic
 
   integer,dimension(NGLLX,NGLLY,NGLLZ,NSPEC_AB):: ibool
 
@@ -174,10 +177,14 @@ subroutine compute_interpolated_dva_ac(displ_element,veloc_element,&
     vyd = veloc_element(2,nint(xi_r),nint(eta_r),nint(gamma_r))
     vzd = veloc_element(3,nint(xi_r),nint(eta_r),nint(gamma_r))
 
-    ! pressure
+    ! global index
     iglob = ibool(nint(xi_r),nint(eta_r),nint(gamma_r),ispec)
-    axd = - potential_dot_dot_acoustic(iglob)
-    ayd = - potential_dot_dot_acoustic(iglob)
+
+    ! x component -> acoustic potential
+    axd = potential_acoustic(iglob)
+    ! y component -> first time derivative of potential
+    ayd = potential_dot_acoustic(iglob)
+    ! z component -> pressure
     azd = - potential_dot_dot_acoustic(iglob)
 
   else
@@ -198,9 +205,12 @@ subroutine compute_interpolated_dva_ac(displ_element,veloc_element,&
           vxd = vxd + hlagrange*veloc_element(1,i,j,k)
           vyd = vxd + hlagrange*veloc_element(2,i,j,k)
[…]

### 2010-11-22 src/specfem3D/setup_sources_receivers.f90

@@ -454,11 +454,10 @@ subroutine setup_sources_precompute_arrays()
   integer :: isource,ispec
   integer :: irec !,irec_local
   integer :: icomp,itime,nadj_files_found,nadj_files_found_tot,ier
-  character(len=3),dimension(NDIM) :: comp = (/ "BHE", "BHN", "BHZ" /)
+  character(len=3),dimension(NDIM) :: comp ! = (/ "BHE", "BHN", "BHZ" /)
   character(len=256) :: filename
 
-
-! forward simulations
+  ! forward simulations
   if (SIMULATION_TYPE == 1  .or. SIMULATION_TYPE == 3) then
     allocate(sourcearray(NDIM,NGLLX,NGLLY,NGLLZ))
     allocate(sourcearrays(NSOURCES,NDIM,NGLLX,NGLLY,NGLLZ))
@@ -508,11 +507,17 @@ subroutine setup_sources_precompute_arrays()
     enddo
   endif
 
-! ADJOINT simulations
+  ! ADJOINT simulations
   if (SIMULATION_TYPE == 2 .or. SIMULATION_TYPE == 3) then
+
+    ! gets channel names
+    do icomp=1,NDIM
+      call write_channel_name(icomp,comp(icomp))
+    enddo
 
     ! counts local receivers which become adjoint sources
     nadj_rec_local = 0
+
     ! temporary counter to check if any files are found at all
     nadj_files_found = 0
     do irec = 1,nrec

### 2010-11-22 doc/USER_MANUAL/manual_SPECFEM3D.tex
adapts IRIS name convention for seismograms; updates seis_process/ scripts; updates manual

@@
 
+% figures
+\usepackage[dvips]{epsfig}
+
 \usepackage{wrapfig}
@@
 Min Chen,
+Hom Nath Gharti,
 Vala Hj\"orleifsd\'ottir,
@@
 David Mich\'ea,
-Hom Nath Gharti,
 Tarje Nissen-Meyer,
@@
 Jeroen Tromp.\\
-% \toall{add full list of developers here and update the AUTHORS file accordingly}
+%\toall{add full list of developers here and update the AUTHORS file accordingly}
 
 The cover graphic of the manual was created
-by Santiago Lombeyda from Caltech's Center for Advanced Computing Research (CACR) \url{http://www.cacr.caltech.edu}.\\
+by Santiago Lombeyda from Caltech's Center for Advanced Computing Research (CACR) \url{http://www.cacr.caltech.edu}. \\
 
@@
 for the selected stations and saved in the \texttt{in\_out\_files/SEM/} directory
-with the format \texttt{STA.NT.BH?.adj}, where \texttt{STA}, \texttt{NT}
+with the format \texttt{STA.NT.BX?.adj}, where \texttt{STA}, \texttt{NT}
 are the station name and network code given in the \texttt{in\_data\_files/STATIONS\_ADJOINT}
-file, and \texttt{BH?} represents the component name of a particular
-adjoint seismogram.
+file, and \texttt{BX?} represents the component name of a particular
+adjoint seismogram. Please note that the band code can change depending on 
+your sampling rate (see Appendix~\ref{cha:channel-codes} for further details).
 \item The adjoint seismograms are in the same format as the original seismogram
-(\texttt{STA.NT.BH?.sem?}), with the same start time, time interval
+(\texttt{STA.NT.BX?.sem?}), with the same start time, time interval
 and record length.
@@
 and ~\\
-\texttt{S?????.NT.BH?.sem} for the three-component displacements
-(\texttt{BHN,BHE,BHZ}) recorded at these locations.
+\texttt{S?????.NT.BX?.sem} for the three-component displacements
+(\texttt{BXN,BXE,BXZ}) recorded at these locations.
 \item \texttt{S?????} denotes the source number; for example, if the original
@@
 and the synthetics need to be processed before a comparison can be
-made. We generally use the following scripts provided in the ~\\
-\texttt{utils/seis\_process/} directory:
-
-\subsection{Data processing script \texttt{process\_trinet\_data.pl}}
+made.
+
+ For such comparisons, the following steps are recommended:
+
+\begin{enumerate}
+\item Make sure that both synthetic and observed seismograms have the correct station/event and timing information.
+\item Convolve synthetic seismograms with a source time function with the half duration specified in the \texttt{CMTSOLUTION} file, provided, as recommended, you used a zero half duration in the SEM simulations.
+\item Resample both observed and synthetic seismograms to a common sampling rate.
+\item Cut the records using the same window.
+\item Remove the trend and mean from the records and taper them.
+\item Remove the instrument response from the observed seismograms (recommended) or convolve the synthetic seismograms with the instrument response.
+\item Make sure that you apply the same filters to both observed and synthetic seismograms. Preferably, avoid filtering your records more than once.
+\item Now, you are ready to compare your synthetic and observed seismograms.
+\end{enumerate}
+
+\noindent
+We generally use the following scripts provided in the \texttt{utils/seis\_process/} directory:
+
+\subsection{Data processing script \texttt{process\_data.pl}}
 
@@
 transfers the data into a displacement record, and picks the first
-P and S arrivals. For more functionality, type `\texttt{process\_trinet\_data.pl}'
+P and S arrivals. For more functionality, type `\texttt{process\_data.pl}'
 without any argument. An example of the usage of the script:
@@
 \begin{lyxcode}
-{\footnotesize process\_trinet\_data.pl~-m~CMTSOLUTION~-l~0/180~-t~2/40~-i~dir~-p~-x~bp~~9703873{*}.BH?.SAC}{\footnotesize \par}
-\end{lyxcode}
-which has cut all the sac files between 0 and 180 seconds, filtered
-them between 2 and 40 seconds, transfered them into displacement records
-using the polezero files in \texttt{dir} directory, picked the first
-P and S arrivals, and added suffix `\texttt{bp}' to the file names.
+%{\footnotesize process\_trinet\_data.pl~-m~CMTSOLUTION~-l~0/180~-t~2/40~-i~dir~-p~-x~bp~~9703873{*}.BH?.SAC}
+{\footnotesize process\_data.pl~-m~CMTSOLUTION~-s~1.0~-l~0/4000~-i~-f~-t~40/500~-p~-x~bp~DATA/1999.330{*}.BH?.SAC}
+{\footnotesize \par}
+\end{lyxcode}
+%which has cut all the sac files between 0 and 180 seconds, filtered
+%them between 2 and 40 seconds, transfered them into displacement records
+%using the polezero files in \texttt{dir} directory, picked the first
+%P and S arrivals, and added suffix `\texttt{bp}' to the file names.
+
+\noindent which has resampled the SAC files to a sampling rate of 1 seconds, cut 
+them between 0 and 4000 seconds, transfered them into displacement
+records and filtered them between 40 and 500 seconds, 
+picked the first P and S arrivals, and added suffix `\texttt{bp}' to the file names.
 
@@
 
-\subsection{Synthetics processing script \texttt{process\_trinet\_syn.pl}}
+\subsection{Synthetics processing script \texttt{process\_syn.pl}}
 
 This script converts the synthetic output from the SEM code from ASCII
-to SAC format, and performs similar operations as `\texttt{process\_trinet\_data.pl}'.
+to SAC format, and performs similar operations as `\texttt{process\_data.pl}'.
 An example of the usage of the script:
@@
 \begin{lyxcode}
-{\footnotesize process\_trinet\_syn.pl~-m~CMTSOLUTION~-a~STATIONS~-l~0/180~-t~2/40~-p~-x~bp~syn/{*}.BH?.semd}{\footnotesize \par}
-\end{lyxcode}
-which will convert the synthetics into SAC format, add event and station
-information into the SAC headers, cut the SAC files between 0 and
-180 seconds, filter them between 2 and 40 seconds, pick the first
-P and S arrivals, and add the suffix `\texttt{bp}' to the file names.
+%{\footnotesize process\_trinet\_syn.pl~-m~CMTSOLUTION~-a~STATIONS~-l~0/180~-t~2/40~-p~-x~bp~syn/{*}.BH?.semd}
+{\footnotesize process\_syn.pl~-m~CMTSOLUTION~-h~-a~STATIONS~-s~1.0~-l~0/4000~-f~-t~40/500~-p~-x~bp~SEM/{*}.BX?.semd}
+{\footnotesize \par}
+\end{lyxcode}
+%which will convert the synthetics into SAC format, add event and station
+%information into the SAC headers, cut the SAC files between 0 and
+%180 seconds, filter them between 2 and 40 seconds, pick the first
+%P and S arrivals, and add the suffix `\texttt{bp}' to the file names.
+which will convolve the synthetics with a triangular source-time function
+from the \texttt{CMTSOLUTION} file, convert the synthetics into SAC
+format, add event and station information into the SAC headers, resample the SAC files with a sampling rate of 1 seconds, cut them between 0 and 4000 seconds, filter them between 40 and
+500 seconds with the same filter used for the observed data, pick the first P and S arrivals, and add the suffix `\texttt{bp}'
+to the file names.
 
@@
 a triangular source time function with a given half duration, etc.
-Type \texttt{process\_trinet\_syn.pl} without any argument for a detailed
+Type \texttt{process\_syn.pl} without any argument for a detailed
 usage.
 
+In order to convert between SAC format and ASCII files, useful scripts
+are provided in the subdirectories ~\\ 
+\texttt{utils/sac2000\_alpha\_convert/} and \texttt{utils/seis\_process/asc2sac/}.
+
 \subsection{Script \texttt{rotate.pl}}
 
-The original data and synthetics have three components: vertical (BHZ),
-north (BHN) and east (BHE). However, for most seismology applications,
+The original data and synthetics have three components: vertical (BHZ resp. BXZ),
+north (BHN resp. BXN) and east (BHE resp. BXE). However, for most seismology applications,
 transverse and radial components are also desirable. Therefore, we
@@
 
-rotate.pl~-l~0~-L~180~SEM/{*}.BHE.semd.sac.bp~
+rotate.pl~-l~0~-L~180~SEM/{*}.BXE.semd.sac.bp~
 \end{lyxcode}
@@
 in the \texttt{in\_out\_files/OUTPUT\_FILES/} directory by default. For the forward simulation, the files
-are named \texttt{STA.NT.BH?.semd} for two-column time series, or
-\texttt{STA.NT.BH?.semd.sac} for ASCII SAC format, where STA and NT
-are the station name and network code, and \texttt{BH?} stands for
-the component name. The adjont simulations generate synthetic seismograms
+are named like \texttt{STA.NT.BX?.semd} for two-column time series, or
+\texttt{STA.NT.BX?.semd.sac} for ASCII SAC format, where STA and NT
+are the station name and network code, and \texttt{BX?} stands for
+the component name. Please see the Appendix~\ref{cha:Coordinates} and \ref{cha:channel-codes}
+for further details.
+
+The adjont simulations generate synthetic seismograms
 with the name \texttt{S?????.NT.S??.sem} (refer to Section \ref{sec:Adjoint-simulation-sources}
 for details). The kernel simulations output the back-reconstructed
-synthetic seismogram in the name \texttt{STA.NT.BH?.semd}, mainly
+synthetic seismogram in the name \texttt{STA.NT.BX?.semd}, mainly
 for the purpose of checking the accuracy of the reconstruction. Refer
@@
 
-You do have two options to change this default output behavior, given in the main constants file
+You do have further options to change this default output behavior, given in the main constants file
 \texttt{constants.h} located in \texttt{src/shared/} directory:
 \begin{description}
+\item [{\texttt{SEISMOGRAMS\_BINARY}}] set to \texttt{.true.} to have seismograms written out in binary format.
 \item [{\texttt{WRITE\_SEISMOGRAMS\_BY\_MASTER}}] Set to \texttt{.true.} to have only the master process writing out seismograms.
@@
 
-\subsection{\texttt{plot\_shakemap.pl}}
+\subsection{\texttt{plot\_shakemap.gmt.pl}}
 
@@
 also run \texttt{`create\_movie\_shakemap\_AVS\_DX\_GMT}' code to convert the binary
-shakemap data into GMT xyz files. A sample script \texttt{plot\_shakemap.pl}
+shakemap data into GMT xyz files. A sample script \texttt{plot\_shakemap.gmt.pl}
 is provided to do this conversion, and then plot the shakemaps in
@@
 \begin{lyxcode}
-plot\_shakemap.pl~data\_dir~type(1,2,3)~CMTSOLUTION~
+plot\_shakemap.gmt.pl~data\_dir~type(1,2,3)~CMTSOLUTION~
 \end{lyxcode}
@@
 
+\subsection*{Source and receiver locations}
+
 The SPECFEM3D software code internally uses Cartesian coordinates.
-%This requires to convert the given locations for sources and receiver locations.
+The given locations for sources and receiver locations thus may get converted.
 Sources and receiver locations are read in from the
@@
 longitude as $x$ coordinate, latitude as $y$ coordinate.
+
 In case the flag \texttt{SUPPRESS\_UTM\_PROJECTION} is set to \texttt{.false.}
@@
 using a UTM conversion with the specified UTM zone.
+
 The value for depth (given in $km$ in \texttt{CMTSOLUTION}) or burial depth (given in $m$ in \texttt{STATIONS})
@@
 
+\subsection*{Seismogram outputs}
+
 The seismogram output directions are given in Cartesian $x$/$y$/$z$ directions and not rotated any further.
-For the labeling of the seismograms, we name "$x$/$y$/$z$" to "BHE/BHN/BHZ" components.
 Changing flags in \texttt{constants.h} in the \texttt{src/shared/} subdirectory only rotates the seismogram outputs
@@
 \texttt{.true.}) as vertical.
-In this case, the outputs are rotated to have the BHZ component normal to the surface of the mesh, BHE and BHN are somewhat arbitrary orthogonal directions along the surface.
+In this case, the outputs are rotated to have the vertical component normal to the surface of the mesh, $x$ and $y$ directions are somewhat arbitrary orthogonal directions along the surface.
+
+For the labeling of the seismogram channels, see Appendix~\ref{cha:channel-codes}. Additionally, we add labels to the synthetics using the following convention:
+For a receiver station located in an
+\begin{description}
+\item[elastic domain:] ~\\
+\begin{itemize}
+\item \texttt{semd} for the displacement vector
+\item \texttt{semv} for the velocity vector
+\item \texttt{sema} for the acceleration vector
+\end{itemize}
+\item[acoustic domain:] ~\\
+(please note that receiver stations in acoustic domains must be buried. This is due to the free surface condition which enforces a zero pressure at the surface.)
+\begin{itemize}
+\item \texttt{semd} for the displacement vector
+\item \texttt{semv} for the velocity vector
+\item \texttt{sema} for pressure which will be stored in the vertical component \texttt{Z}.
+Note that pressure in the acoustic media is isotropic and thus the pressure record would be the same in the other two component directions.
+We therefore use the other two seismogram components to store the acoustic potential in component \texttt{X} (or \texttt{N}) and the first time derivative of the acoustic potential in component \texttt{Y} (or \texttt{E}).
+\end{itemize}
+\end{description}
+The seismograms are by default written out in ASCII-format to the \texttt{in\_out\_files/OUTPUT\_FILES/} subdirectory by each parallel process.
+You can change this behavior by changing the following flags in the \texttt{constants.h} file located in the \texttt{src/shared/} subdirectory:
+\begin{description}
+\item [{\texttt{SEISMOGRAMS\_BINARY}}] set to \texttt{.true.} to have seismograms written out in binary format.
+\item [{\texttt{WRITE\_SEISMOGRAMS\_BY\_MASTER}}] Set to \texttt{.true.} to have only the master process writing out seismograms.
+This can be useful on a cluster, where only the master process node has access to the output directory.
+\item [{\texttt{USE\_OUTPUT\_FILES\_PATH}}] Set to \texttt{.false.} to have the seismograms output to \texttt{LOCAL\_PATH} directory specified in the main parameter file \texttt{in\_data\_files/Par\_file}.
+\end{description}
+
+\chapter{\label{cha:channel-codes}Channel Codes of Seismograms}
+
+Seismic networks, such as the Global Seismographic Network (GSN), generally involve various types of instruments with different bandwidths, sampling properties and component configurations.
+There are standards to name channel codes depending on instrument properties.
+IRIS  \texttt{(www.iris.edu)} uses SEED format for channel codes, which are represented by three letters, such as \texttt{LHN}, \texttt{BHZ}, etc.
+In older versions of the SPECFEM3D package, a common format was used for the channel codes of all seismograms, which was \texttt{BHE/BHN/BHZ} for three components.
+To avoid confusion when comparison are made to observed data, we are now using the IRIS convention for SEM seismograms.
+In the following, we give a brief explanation of the IRIS convention and how it is adopted in SEM seismograms.
+Please visit \texttt{www.iris.edu/manuals/SEED\_appA.htm} for further information.\\
+
+\noindent \textbf{\texttt{Band code:}} The first letter in the channel code denotes the band code of seismograms, which depends on the response band and the sampling rate of instruments.
+The list of band codes used by IRIS is shown in Figure \ref{fig:IRIS_band_codes}.
+The sampling rate of SEM synthetics is controlled by the resolution of simulations rather than instrument properties.
+However, for consistency, we follow the IRIS convention for SEM seismograms governed by their sampling rate.
+For SEM synthetics, we consider band codes for which $dt \leq 1$ s.
+IRIS also considers the response band of instruments.
+For instance, short-period and broad-band seismograms with the same sampling rate correspond to different band codes, such as S and B, respectively.
+In such cases, we consider SEM seismograms as broad band, ignoring the corner period ($\geq 10$ s) of the response band of instruments (note that at these resolutions, the minimum period in the SEM synthetics will be less than $10$ s).
+Accordingly, when you run a simulation the band code will be chosen depending on the resolution of the synthetics, and channel codes of SEM seismograms will start with either \texttt{L}, \texttt{M}, \texttt{B}, \texttt{H}, \texttt{C} or \texttt{F}, shown by red color in the figure.\\
+
+\begin{figure}[ht]
+\noindent \begin{centering}
+\includegraphics[scale=0.6]{figures/IRIS_band_codes.eps}
+\par\end{centering}
+
+\caption{The IRIS band code convention is based on the sampling rate and the response band of instruments.
+Please visit \texttt{www.iris.edu/manuals/SEED\_appA.htm} for further information.
+Grey rows show the relative band-code range in SPECFEM3D, and the band codes used to name SEM seismograms are denoted in red.}
+
+\label{fig:IRIS_band_codes}
+\end{figure}
+
+\noindent \textbf{\texttt{Instrument code:}} The second letter in the channel code corresponds to instrument codes, which specify the family to which the sensor belongs.
+For instance, \texttt{H} and \texttt{L} are used for high-gain and low-gain seismometers, respectively.
+The instrument code of SEM seismograms will always be \texttt{X}, as assigned by IRIS for synthetic seismograms.
+\\
+
+\noindent \textbf{\texttt{Orientation code:}} The third letter in channel codes is an orientation code, which generally describes the physical configuration of the components of instrument packages.
+SPECFEM3D uses the traditional orientation code  \texttt{E/N/Z} (East-West, North-South, Vertical) for three components when a UTM projection is used.
+If the UTM conversion is suppressed, i.e.
+the flag \texttt{SUPPRESS\_UTM\_PROJECTION} is set to \texttt{.true.}, then the three components are labelled \texttt{X/Y/Z} according to the Cartesian reference frame. \\
+
+\noindent \textbf{\texttt{EXAMPLE:}} The sampling rate is given by \texttt{DT} in the main parameter file \texttt{Par\_file} located in the \texttt{in\_data\_files/} subdirectory.
+Depending on the resolution of your simulations, if you choose a sampling rate greater than $0.01$ s and less than $1$ s, a seismogram recording displacements on the vertical component of a station \texttt{ASBS} (network \texttt{AZ}) will be named \texttt{ASBS.AZ.MXZ.semd.sac}, whereas it will be \texttt{ASBS.AZ.BXZ.semd.sac}, if the sampling rate is greater than 0.0125 and less equal to 0.1 s.
 

### 2010-11-22 src/specfem3D/noise_tomography.f90

@@ -1,0 +1,695 @@
+!=====================================================================
+!
+!               S p e c f e m 3 D  V e r s i o n  2 . 0
+!               ---------------------------------------
+!
+!          Main authors: Dimitri Komatitsch and Jeroen Tromp
+!                        Princeton University, USA
+! (c) Princeton University / California Institute of Technology and University of Pau / CNRS / INRIA
+!                            November 2010
+!
+! This program is free software; you can redistribute it and/or modify
+! it under the terms of the GNU General Public License as published by
+! the Free Software Foundation; either version 2 of the License, or
+! (at your option) any later version.
+!
+! This program is distributed in the hope that it will be useful,
+! but WITHOUT ANY WARRANTY; without even the implied warranty of
+! MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+! GNU General Public License for more details.
+!
+! You should have received a copy of the GNU General Public License along
+! with this program; if not, write to the Free Software Foundation, Inc.,
+! 51 Franklin Street, Fifth Floor, Boston, MA 02110-1301 USA.
+!
+!=====================================================================
+
+
+! !!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!! ATTENTION !!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!
+!
+! This file is first implemented for SPECFEM3D_GLOBE, and therefore it follows variables in GLOBAL package.
+! Do NOT worry about strange names containing, cause they are from GLOBAL package.
+! However, please be informed that some of those varibles (even input parameters) can be dummy (not used),
+!    since they may exist only in GLOBAL package.
+! Also, slight modifications are included due to different structures in SPECFEM3D_GLOBE and SPECFEM3D.
+! In general, this file in SPECFEM3D should be better than the one in SPECFEM3D_GLOBE,
[…]

### 2010-11-22 doc/USER_MANUAL/manual_SPECFEM3D.tex
changed "it's" to "it is" and "paper" to "article"

@@
 as well as Convolutional or Auxiliary Differential Equation Perfectly Matched absorbing Layers (C-PML or ADE-PML) \citep{KoMa07,MaKoEz08,MaKoGe08,MaKo09,MaKoGeBr10}.
-The next release will use the PT-Scotch parallel library for mesh partitioning.
+It will also use the PT-Scotch parallel library for mesh partitioning.
 
@@
 and adjoint simulations (Chapter \ref{cha:Adjoint-Simulations}).
-Also, make sure you read the paper `Noise cross-correlation sensitivity kernels' \citep{trompetal2010},
+Also, make sure you read the article `Noise cross-correlation sensitivity kernels' \citep{trompetal2010},
 in order to understand noise simulations from a theoretical perspective.
@@
 
-Note that it's better to run the three steps continuously within the same job,
+Note that it is better to run the three steps continuously within the same job,
 otherwise you have to collect the saved surface movies from the old nodes/CPUs to the new nodes/CPUs.

### 2010-11-22 doc/USER_MANUAL/manual_SPECFEM3D.tex
added reference to Arnold 1982

@@
 based upon the spectral-element method (SEM).
-The SEM is a continuous Galerkin technique, which can easily be made discontinous \citep{BeMaPa94,KoWoHu02,ChCaVi03,Kop06,WiStBuGh10}; it is then a particular case of the discontinuous Galerkin technique \citep{ReHi73,FaRi99,HuHuRa99,CoKaSh00,GiHeWa02,RiWh03,MoRi05,GrScSc06,BeLaPi06,DuKa06,DeSeWh08,WiStBuGh10,DeSe10,EtChViGl10}, with optimized efficiency because of its tensorized basis functions.
+The SEM is a continuous Galerkin technique, which can easily be made discontinous \citep{BeMaPa94,KoWoHu02,ChCaVi03,Kop06,WiStBuGh10}; it is then a particular case of the discontinuous Galerkin technique \citep{ReHi73,Arn82,FaRi99,HuHuRa99,CoKaSh00,GiHeWa02,RiWh03,MoRi05,GrScSc06,BeLaPi06,DuKa06,DeSeWh08,WiStBuGh10,DeSe10,EtChViGl10}, with optimized efficiency because of its tensorized basis functions.
 In particular, it can accurately handle very distorted mesh elements \citep{OlSe11}.

### 2010-11-22 doc/USER_MANUAL/manual_SPECFEM3D.tex
added reference to Ainsworth et al. (2006)

--- 2010-11-22 doc/USER_MANUAL/manual_SPECFEM3D.tex
@@
 based upon the spectral-element method (SEM).
-The SEM is a continuous Galerkin technique, which can easily be made discontinous \citep{BeMaPa94,KoWoHu02,ChCaVi03,Kop06,WiStBuGh10}; it is then a particular case of the discontinuous Galerkin technique \citep{ReHi73,FaRi99,HuHuRa99,CoKaSh00,GiHeWa02,RiWh03,MoRi05,GrScSc06,BeLaPi06,DuKa06,DeSeWh08,WiStBuGh10,DeSe10,EtChViGl10}, with optimized efficiency because of its tensorized basis functions.
+The SEM is a continuous Galerkin technique, which can easily be made discontinous \citep{BeMaPa94,KoWoHu02,ChCaVi03,Kop06,WiStBuGh10}; it is then a particular case of the discontinuous Galerkin technique \citep{ReHi73,Arn82,FaRi99,HuHuRa99,CoKaSh00,GiHeWa02,RiWh03,MoRi05,GrScSc06,AiMoMu06,BeLaPi06,DuKa06,DeSeWh08,WiStBuGh10,DeSe10,EtChViGl10}, with optimized efficiency because of its tensorized basis functions.
 In particular, it can accurately handle very distorted mesh elements \citep{OlSe11}.

### 2010-11-22 doc/USER_MANUAL/manual_SPECFEM3D.tex
added reference to SCOTCH

@@
 as well as Convolutional or Auxiliary Differential Equation Perfectly Matched absorbing Layers (C-PML or ADE-PML) \citep{KoMa07,MaKoEz08,MaKoGe08,MaKo09,MaKoGeBr10}.
-It will also use the PT-Scotch parallel library for mesh partitioning.
+It will also use the PT-SCOTCH parallel library for mesh partitioning.
 
@@
 
-The SPECFEM3D software package relies on the Scotch library to partition meshes created with CUBIT.
+The SPECFEM3D software package relies on the SCOTCH library to partition meshes created with CUBIT.
 Note that we use CUBIT to create meshes of hexahedra, but other packages can be used as well, for instance GiD from \url{http://gid.cimne.upc.es} or Gmsh from \url{http://geuz.org/gmsh}.
@@
 
+The SCOTCH library \citep{PeRo96}
+provides efficent static mapping, graph and mesh partitioning routines. SCOTCH is a free software package developed by
+Fran\c {c}ois Pellegrini et al.
+from LaBRI and INRIA in Bordeaux, France, downloadable from the web page \url{https://gforge.inria.fr/projects/scotch/}.
+It is more recent than METIS, actively maintained and performs better in many cases.
 The path to the SCOTCH installation needs to be set with the option \texttt{-{}-with-scotch-path}. Just as an example:
@@
 It amounts to balance the number of elements in each slice while minimizing the communication costs
-resulting from the placement of adjacent elements on different processors. \texttt{decompose\_mesh\_SCOTCH} depends on the Scotch library
-which provides efficent static mapping, graph and mesh partitioning routines. Scotch is a free software package developed by
+resulting from the placement of adjacent elements on different processors. \texttt{decompose\_mesh\_SCOTCH} depends on the SCOTCH library \citep{PeRo96},
+which provides efficent static mapping, graph and mesh partitioning routines. SCOTCH is a free software package developed by
 Fran\c {c}ois Pellegrini et al.
@@
 Prior to compiling \texttt{decompose\_mesh\_SCOTCH}, make sure you have correctly specified the path
-of the Scotch library with the option \texttt{-{}-with-scotch-path} of the \texttt{configure} script.
+of the SCOTCH library with the option \texttt{-{}-with-scotch-path} of the \texttt{configure} script.
 For convenience we provide the source code of SCOTCH, which is released open source under the French CeCILL-C version 1 license,

### 2010-11-23 doc/USER_MANUAL/manual_SPECFEM3D.tex
added a sentence about the path to the SCOTCH source code
moves unused/old/deprecated files and routines into SPECFEM3D/utils/unused_routines; updates manual and Makefiles

@@
 It is more recent than METIS, actively maintained and performs better in many cases.
+A recent version of its source code is provided in directory \texttt{src/decompose\_mesh\_SCOTCH/scotch\_5.1.10b}.
 The path to the SCOTCH installation needs to be set with the option \texttt{-{}-with-scotch-path}. Just as an example:
@@
 It is more recent than METIS, actively maintained and performs better in many cases.
+A recent version of its source code is provided in directory \texttt{src/decompose\_mesh\_SCOTCH/scotch\_5.1.10b}.
 \\
@@
 
-\subsection{\texttt{movie2gif.pl}}
+\subsection{Script \texttt{movie2gif.gmt.pl}}
 
@@
 a movie simulation (\texttt{\small MOVIE\_SURFACE=.true.}), you can
-run the \texttt{`create\_movie\_GMT}' code to convert these binary
+run the \texttt{`create\_movie\_shakemap\_AVS\_DX\_GMT}' code to convert these binary
 movie data into GMT xyz files for futher processing. A sample script
-\texttt{movie2gif.pl} is provided to do this conversion, and then
+\texttt{movie2gif.gmt.pl} is provided to do this conversion, and then
 plot the movie snapshots in GMT, for example:
@@
 \begin{lyxcode}
-movie2gif.pl~-m~CMTSOLUTION~-g~-f~1/40~-n~-2~-p~
+movie2gif.gmt.pl~-m~CMTSOLUTION~-g~-f~1/40~-n~-2~-p~
 \end{lyxcode}
@@
 
-\subsection{\texttt{plot\_shakemap.gmt.pl}}
+\subsection{Script \texttt{plot\_shakemap.gmt.pl}}
 

### 2010-11-23 doc/USER_MANUAL/manual_SPECFEM3D.tex
changed www.paraview.com to www.paraview.org

@@
 Explorer, AVS \url{http://www.avs.com} is a trademark of Advanced
-Visualization Systems, and ParaView \url{http://www.paraview.com}
+Visualization Systems, and ParaView \url{http://www.paraview.org}
 is an open-source visualization platform.

### 2010-11-24 doc/USER_MANUAL/manual_SPECFEM3D.tex
fixed a bug with URLs and parentheses

@@
 
-\let\myUrl\url
-\renewcommand{\url}[1]{(\myUrl{#1})}
+\newcommand{\urlwithparentheses}[1]{(\url{#1})}
 
@@
 The cover graphic of the manual was created
-by Santiago Lombeyda from Caltech's Center for Advanced Computing Research (CACR) \url{http://www.cacr.caltech.edu}. \\
+by Santiago Lombeyda from Caltech's Center for Advanced Computing Research (CACR) \urlwithparentheses{http://www.cacr.caltech.edu}. \\
 
@@
 Examples are provided in the SPECFEM3D package in the subdirectory \texttt{examples/}.
-We strongly encourage you to contribute your own example to this package by contacting the CIG Computational Seismology Mailing List  \url{cig-seismo@geodynamics.org}.
+We strongly encourage you to contribute your own example to this package by contacting the CIG Computational Seismology Mailing List  \urlwithparentheses{cig-seismo@geodynamics.org}.
 
@@
 In each example subdirectory you will find a \texttt{README} file, which explains in a step-by-step tutorial the workflow for the example.
-Please feel free to contribute your own example to this package by contacting the CIG Computational Seismology Mailing List  \url{cig-seismo@geodynamics.org}.
+Please feel free to contribute your own example to this package by contacting the CIG Computational Seismology Mailing List  \urlwithparentheses{cig-seismo@geodynamics.org}.
 
@@
 if \texttt{USE\_REGULAR\_MESH} is set to \texttt{.true.} and if \texttt{NDOUBLINGS} is set to \texttt{2}).
-\item [{\texttt{CREATE\_ABAQUS\_FILES}}] Set this flag to \texttt{.true.} to save Abaqus FEA \url{www.simulia.com}
+\item [{\texttt{CREATE\_ABAQUS\_FILES}}] Set this flag to \texttt{.true.} to save Abaqus FEA \urlwithparentheses{www.simulia.com}
 mesh files for subsequent viewing. Turning the flag on generates files in the \texttt{LOCAL\_PATH} directory.
 See Section~\ref{sec:Mesh-graphics} for a discussion of mesh viewing features.
-\item [{\texttt{CREATE\_DX\_FILES}}] Set this flag to \texttt{.true.} to save OpenDX \url{www.opendx.org}
+\item [{\texttt{CREATE\_DX\_FILES}}] Set this flag to \texttt{.true.} to save OpenDX \urlwithparentheses{www.opendx.org}
 mesh files for subsequent viewing.
@@
 \item [{\texttt{SAVE\_MESH\_FILES}}] Set this flag to \texttt{.true.} to
-save ParaView \url{www.paraview.org}
+save ParaView \urlwithparentheses{www.paraview.org}
 mesh files for subsequent viewing. Turning the flag on generates large
@@
 For any particular earthquake, the \texttt{CMTSOLUTION} file that
-represents the point source may be obtained directly from the Harvard Centroid-Moment Tensor (CMT) web page \url{www.seismology.harvard.edu}.
+represents the point source may be obtained directly from the Harvard Centroid-Moment Tensor (CMT) web page \urlwithparentheses{www.seismology.harvard.edu}.
 It looks like this:
@@
 script \texttt{convolve\_source\_timefunction.csh} for this purpose,
-or alternatively use signal-processing software packages such as SAC \url{www.llnl.gov/sac}.
+or alternatively use signal-processing software packages such as SAC \urlwithparentheses{www.llnl.gov/sac}.
 Type
@@
 prior to running the mesher (see Chapter~\ref{cha:Running-the-Mesher-Meshfem3D} for details).
-You can then use AVS \url{www.avs.com} or OpenDX \url{www.opendx.org}
+You can then use AVS \urlwithparentheses{www.avs.com} or OpenDX \urlwithparentheses{www.opendx.org}
 to visualize the mesh and MPI partition (slices).
-
-% obsolete: only for global package...
-%Use the serial code \texttt{combine\_AVS\_DX.f90} (type `\texttt{make
-%combine\_AVS\_DX}' and then `\texttt{xcombine\_AVS\_DX}') to generate
-%AVS \url{www.avs.com} output files (in AVS UCD format) or OpenDX \url{www.opendx.org}
-%output files showing the mesh, the MPI partition (slices), the $\nchunks$
-%chunks, the source and receiver location, etc.
-%Use the AVS UCD files \texttt{AVS\_continent\_boundaries.inp} and \texttt{AVS\_plate\_boundaries.inp}
-%or the OpenDX files \texttt{DX\_continent\_boundaries.dx} and \texttt{DX\_plate\_boundaries.dx}
-%for reference.
 
@@
 You have also the option to visualize the distributed databases produced by \texttt{xgenerate\_databases}
-using Paraview \url{www.paraview.org}. For this, you must set the flag \texttt{SAVE\_MESH\_FILES} to \texttt{.true.}
+using Paraview \urlwithparentheses{www.paraview.org}. For this, you must set the flag \texttt{SAVE\_MESH\_FILES} to \texttt{.true.}
 in the main parameter file \texttt{Par\_file} (see Chapter~\ref{cha:Main-Parameter} for details).
-This will create VTK files for each single partition. You can then use Paraview \url{www.paraview.org} to visualized these partitions.
+This will create VTK files for each single partition. You can then use Paraview \urlwithparentheses{www.paraview.org} to visualized these partitions.
 
@@
 The \texttt{SPECFEM3D} code is running in near real-time to produce
-animations of southern California earthquakes via the web; see Southern California ShakeMovie\textregistered \url{www.shakemovie.caltech.edu}.
+animations of southern California earthquakes via the web; see Southern California ShakeMovie\textregistered \urlwithparentheses{www.shakemovie.caltech.edu}.
 
@@
 \item Notice that this Perl script uses a program \texttt{mesh2vtu} in the
-\texttt{utils/Visualization/Paraview/mesh2vtu} directory, which further uses the VTK \url{http://www.vtk.org/}
+\texttt{utils/Visualization/Paraview/mesh2vtu} directory, which further uses the VTK \urlwithparentheses{http://www.vtk.org/}
 run-time library for its execution. Therefore, make sure you have
@@
 
-Finally, we can view the mesh in ParaView \url{www.paraview.org}.
+Finally, we can view the mesh in ParaView \urlwithparentheses{www.paraview.org}.
 
@@
 \textsf{Glyph}, and represent the points by `\textsf{spheres}'.
-\item For more information about ParaView, see the ParaView Users Guide \url{www.paraview.org/files/v1.6/ParaViewUsersGuide.PDF}.
+\item For more information about ParaView, see the ParaView Users Guide \urlwithparentheses{www.paraview.org/files/v1.6/ParaViewUsersGuide.PDF}.
 \end{enumerate}
@@
 the same commands on many hosts in parallel. \texttt{shmux} is available
-from Shmux \url{web.taranis.org/shmux/}. Make sure that the \texttt{LOCAL\_PATH}
+from Shmux \urlwithparentheses{web.taranis.org/shmux/}. Make sure that the \texttt{LOCAL\_PATH}
 parameter in \texttt{in\_data\_files/Par\_file} is also set properly.
@@
 where the first command performs rotation on the SAC data obtained
-through Seismogram Transfer Program (STP) \url{http://www.data.scec.org/STP/stp.html},
+through Seismogram Transfer Program (STP) \urlwithparentheses{http://www.data.scec.org/STP/stp.html},
 while the second command rotates the processed SEM synthetics.
@@
 To report bugs or suggest improvements to the code, please send an
-e-mail to the CIG Computational Seismology Mailing List \url{cig-seismo@geodynamics.org}
-or Jeroen Tromp \url{jtromp-AT-princeton.edu}, and/or use our online
-bug tracking system Roundup \url{www.geodynamics.org/roundup}.
+e-mail to the CIG Computational Seismology Mailing List \urlwithparentheses{cig-seismo@geodynamics.org}
+or Jeroen Tromp \urlwithparentheses{jtromp-AT-princeton.edu}, and/or use our online
+bug tracking system Roundup \urlwithparentheses{www.geodynamics.org/roundup}.
 
@@
 
-OpenDX \url{http://www.opendx.org} is open-source based on IBM Data
-Explorer, AVS \url{http://www.avs.com} is a trademark of Advanced
-Visualization Systems, and ParaView \url{http://www.paraview.org}
+OpenDX \urlwithparentheses{http://www.opendx.org} is open-source based on IBM Data
+Explorer, AVS \urlwithparentheses{http://www.avs.com} is a trademark of Advanced
+Visualization Systems, and ParaView \urlwithparentheses{http://www.paraview.org}
 is an open-source visualization platform.
 
-%The main developers of the \texttt{SPECFEM3D} source code are Dimitri
-%Komatitsch, Jeroen Tromp, and Qinya Liu. The following individuals
-%(listed in alphabetical order) have also contributed to the development
-%or improvement of the source code: Min Chen, Vala Hj\"orleifsd\'ottir,
-%Jes\'us Labarta, and Leif Strand. The following individuals (listed in alphabetical
-%order) contributed to this manual: Min Chen, Vala Hj\"orleifsd\'ottir,
-%Sue Kientz, Dimitri Komatitsch, Qinya Liu, Alessia Maggi,
-%Carl Tape, and Jeroen Tromp. The manual's cover graphic was created
-%by Santiago Lombeyda from Caltech's Center for Advanced Computing Research (CACR) \url{http://www.cacr.caltech.edu}.
-%Older versions of the code were initially developed by Dimitri Komatitsch at Institut de Physique du Globe (France)
-%and then by Dimitri Komatitsch and Jeroen Tromp at Harvard University (USA).
-
 Please e-mail your feedback, questions, comments, and suggestions
-to Jeroen Tromp \url{jtromp-AT-princeton.edu} or to the CIG Computational Seismology Mailing List \url{cig-seismo@geodynamics.org}.
+to Jeroen Tromp \urlwithparentheses{jtromp-AT-princeton.edu} or to the CIG Computational Seismology Mailing List \urlwithparentheses{cig-seismo@geodynamics.org}.
 

### 2010-11-24 doc/USER_MANUAL/manual_SPECFEM3D.tex
fixed all the warnings detected by IBM xlf

@@
 then comparing the seismograms. If they are identical, you should probably
-select single precision for your future runs.
-
-When running on an SGI add ``\texttt{setenv TRAP\_FPE OFF}'' to your
-.cshrc file \textit{before} compiling in order to turn underflow trapping
-off.
-
-% obsolete
-%Finally, before compiling make sure that the subdirectories \texttt{obj}
-%and \texttt{OUTPUT\_FILES} exist within the directory with the source
-%code (\texttt{SPECFEM3D\_V2.0.0}). The \texttt{go\_mesher} script
-%discussed in Chapter~\ref{cha:Running-the-Mesher-Meshfem3D} automatically
-%takes care of creating the \texttt{OUTPUT\_FILES} directory.
+select single precision for your future runs.\\
+
+When compiling on an IBM machine with the \texttt{xlf} and \texttt{xlc} compilers, we suggest running the \texttt{configure} script
+with the following options:\\
+
+\noindent\texttt{./configure FC=xlf90\_r MPIFC=mpif90 CC=xlc\_r CFLAGS="-O3 -q64" FCFLAGS="-O3 -q64" -with-scotch-path=...}.\\
+
+On SGI systems, \texttt{flags.guess} automatically informs \texttt{configure}
+to insert `\texttt{`TRAP\_FPE=OFF}'' into the generated \texttt{Makefile}
+in order to turn underflow trapping off.\\
 
@@
 in this case, on some compilers you may need to add \char`\"{}\texttt{-mcmodel=medium}\char`\"{}
-to the compiler options otherwise the compiler will display an error
-message.
+to the compiler options otherwise the compiler will display an error message; on an IBM machine with the \texttt{xlf} and \texttt{xlc} compilers, using \texttt{-q64} is usually sufficient.
 
@@
 
-%\noindent \begin{flushleft}
-%
 \begin{figure}[htbp]
@@
 \end{figure}
-
-%\par\end{flushleft}
 
@@
 is on.
-%obsolete...
-%\item [{\texttt{TOPOGRAPHY}}] Set to \texttt{.true.} if topography and
-%bathymetry should be incorporated based upon model ETOPO5 \citep{Etopo5}.
-%This feature adds no cost to the simulation.
 \item [{\texttt{ATTENUATION}}] Set to \texttt{.true.} if attenuation should
@@
 on Clayton-Enquist absorbing boundary conditions (see \citet{KoTr99}).
-%obsolete...
-%\item [{\texttt{RECORD\_LENGTH\_IN\_MINUTES}}] Choose the desired record
-%length of the synthetic seismograms (in minutes). This controls the
-%length of the numerical simulation, i.e., twice the record length
-%requires twice as much CPU time. This feature is not used at the time
-%of meshing but is required for the solver, i.e., you may change this
-%parameter after running \texttt{xgenerate\_databases}.
 \item [{\texttt{MOVIE\_SURFACE}}] Set to \texttt{.false.}, unless you want
@@
 
-%% Another file generated by \texttt{xgenerate\_databases} is the header file \texttt{OUTPUT\_FILES/values\_from\_mesher.h}.
-%% This file specifies a number of constants and flags needed by the
-%% solver. These values are passed statically to the solver for reasons
-%% of speed. Some useful statistics about the mesh are also provided
-%% in this file.
-
-%% For a given model, set of nodes, and set of parameters in \texttt{Par\_file},
-%% one only needs to run the mesher once and for all, even if one wants
-%% to run several simulations with different sources and/or receivers
-%% (the source and receiver information is used in the solver only).
-
 \section{\label{cha:Choosing-the-Time-Step}Choosing the time step \texttt{DT}}
@@
 the \texttt{bin/} directory, as most of the binaries of the package.
-
-%obsolete
-%For reasons of speed, the solver uses static memory allocation.
-%Therefore it needs to be recompiled (type `\texttt{make clean}' and
-%`\texttt{make specfem3D}') every time one reruns \texttt{xgenerate\_databases}. To compile
-%the solver one needs a file created by \texttt{xgenerate\_databases} in the directory
-%\texttt{OUTPUT\_FILES} called \texttt{values\_from\_mesher.h}, which
-%contains parameters describing the static size of the arrays as well
-%as the setting of certain flags.
 
@@
 \item the time step parameters \texttt{NSTEP} and \texttt{DT}
-%\item the attenuation control parameters \texttt{ATTENUATION} and \texttt{USE\_OLSEN\_ATTENUATION}
 \item the absorbing boundary control parameter \texttt{ABSORBING\_CONDITIONS}
@@
 
-%<YANGL
 \chapter{Noise Cross-correlation Simulations}
@@
 which should contain enough explanations for all the commands.
-%>YANGL
 
@@
 \end{figure}
-
-% uses obsolete routines...
-%\item [\texttt{xcreate\_movie\_GMT}]
-%From \texttt{create\_movie\_GMT.f90},
-%outputs ascii xyz files, convenient for use with GMT. This code uses
-%significantly less memory than \texttt{create\_movie\_shakemap\_AVS\_DX\_GMT.f90}
-%and is therefore useful for high resolution runs.
 
@@
 both the \texttt{xgenerate\_databases} and the \texttt{xspecfem3D} executables.
-%obsolete
-%A small program provided in the distribution
-%called \texttt{\small create\_header\_file.f90} can be used to directly
-%create \texttt{\small in\_out\_files/OUTPUT\_FILES/values\_from\_mesher.h} using
-%the information in the \texttt{\small in\_data\_files/Par\_file} without having
-%to run the mesher (type \texttt{\small `make}{\small{} }\texttt{\small create\_header\_}~\\
-%\texttt{\small file}' to compile it and `\texttt{\small xcreate\_header\_file}'
-%to run it; refer to the sample scripts below). The solver can now
-%be compiled as explained above.
 \item One feature of schedulers/queuing systems is that they allow submission
@@
 \begin{lyxcode}
-%{\footnotesize process\_trinet\_data.pl~-m~CMTSOLUTION~-l~0/180~-t~2/40~-i~dir~-p~-x~bp~~9703873{*}.BH?.SAC}
 {\footnotesize process\_data.pl~-m~CMTSOLUTION~-s~1.0~-l~0/4000~-i~-f~-t~40/500~-p~-x~bp~DATA/1999.330{*}.BH?.SAC}
@@
 \end{lyxcode}
-%which has cut all the sac files between 0 and 180 seconds, filtered
-%them between 2 and 40 seconds, transfered them into displacement records
-%using the polezero files in \texttt{dir} directory, picked the first
-%P and S arrivals, and added suffix `\texttt{bp}' to the file names.
 
@@
 \begin{lyxcode}
-%{\footnotesize process\_trinet\_syn.pl~-m~CMTSOLUTION~-a~STATIONS~-l~0/180~-t~2/40~-p~-x~bp~syn/{*}.BH?.semd}
 {\footnotesize process\_syn.pl~-m~CMTSOLUTION~-h~-a~STATIONS~-s~1.0~-l~0/4000~-f~-t~40/500~-p~-x~bp~SEM/{*}.BX?.semd}
@@
 \end{lyxcode}
-%which will convert the synthetics into SAC format, add event and station
-%information into the SAC headers, cut the SAC files between 0 and
-%180 seconds, filter them between 2 and 40 seconds, pick the first
-%P and S arrivals, and add the suffix `\texttt{bp}' to the file names.
 which will convolve the synthetics with a triangular source-time function

### 2010-11-25 doc/USER_MANUAL/manual_SPECFEM3D.tex
final list of authors; fixed a typo in a comment

@@
 C\'eline Blitz,
-Ebru Bozdag,
+Ebru Bozda\u{g},
 Emanuele Casarotti,
@@
 Dennis McRitchie,
+Matthias Meschede,
 David Mich\'ea,
@@
 Brian Savage,
+Bernhard Schuberth,
+Anne Sieminski,
 Leif Strand,
 Carl Tape,
-Jeroen Tromp.\\
+Jeroen Tromp,
+Hejun Zhu.\\
 %\toall{add full list of developers here and update the AUTHORS file accordingly}

### 2010-11-25 doc/USER_MANUAL/manual_SPECFEM3D.tex
added references to our domain decomposition and weak scaling papers

@@
 \cite{TrTaLi05,LiTr06,TrKoLi08,LiTr08,TrKoHjLiZhPeBoMcFrTrHu10}.
+Domain decomposition is explained in detail in \cite{MaKoBlLe08}, and excellent scaling up to 150,000 processor cores in shown for instance in
+\cite{CaKoLaTiMiLeSnTr08,KoLaMi08a,MaKoBlLe08,KoErGoMi10,KoGoErMi10,Kom11}.
 
@@
 Each single core partition is colored differently.
-The executable \texttt{xdecompose\_mesh\_SCOTCH} can equally distribute the mesh on any arbitrary number of cores.}
+The executable \texttt{xdecompose\_mesh\_SCOTCH} can equally distribute the mesh on any arbitrary number of cores.
+Domain decomposition is explained in detail in \cite{MaKoBlLe08}, and excellent scaling up to 150,000 processor cores in shown for instance in
+\cite{CaKoLaTiMiLeSnTr08,KoLaMi08a,MaKoBlLe08,KoErGoMi10,KoGoErMi10,Kom11}.}
 

### 2010-11-25 doc/USER_MANUAL/manual_SPECFEM3D.tex
updates SPECFEM3D configure script and manual regarding scotch installation

@@
 A recent version of its source code is provided in directory \texttt{src/decompose\_mesh\_SCOTCH/scotch\_5.1.10b}.
-The path to the SCOTCH installation needs to be set with the option \texttt{-{}-with-scotch-path}. Just as an example:
-\begin{lyxcode}
-./configure~FC=ifort~MPIFC=mpif90~-{}-with-scotch-path=/opt/scotch
+The path to an existing SCOTCH installation can to be set with the option \texttt{-{}-with-scotch-dir}. Just as an example:
+\begin{lyxcode}
+./configure~FC=ifort~MPIFC=mpif90~-{}-with-scotch-dir=/opt/scotch
 \end{lyxcode}
@@
 
-\noindent\texttt{./configure FC=xlf90\_r MPIFC=mpif90 CC=xlc\_r CFLAGS="-O3 -q64" FCFLAGS="-O3 -q64" -with-scotch-path=...}.\\
+\noindent\texttt{./configure FC=xlf90\_r MPIFC=mpif90 CC=xlc\_r CFLAGS="-O3 -q64" FCFLAGS="-O3 -q64" -with-scotch-dir=...}.\\
 
@@
 Prior to compiling \texttt{decompose\_mesh\_SCOTCH}, make sure you have correctly specified the path
-of the SCOTCH library with the option \texttt{-{}-with-scotch-path} of the \texttt{configure} script.
+of the SCOTCH library with the option \texttt{-{}-with-scotch-dir} of the \texttt{configure} script.
 For convenience we provide the source code of SCOTCH, which is released open source under the French CeCILL-C version 1 license,
@@
 
-\texttt{./configure FC=ifort MPIFC=mpif90 --with-scotch-path=/opt/scotch}\\
+\texttt{./configure FC=ifort MPIFC=mpif90 --with-scotch-dir=/opt/scotch}\\
 

### 2010-11-29 doc/USER_MANUAL/manual_SPECFEM3D.tex
updates SPECFEM3D configuration with scotch, manual and examples; further code clean up

@@
 \cite{TrTaLi05,LiTr06,TrKoLi08,LiTr08,TrKoHjLiZhPeBoMcFrTrHu10}.
-Domain decomposition is explained in detail in \cite{MaKoBlLe08}, and excellent scaling up to 150,000 processor cores in shown for instance in
+Domain decomposition is explained in detail in \cite{MaKoBlLe08}, and excellent scaling up to 150,000 processor cores is shown for instance in
 \cite{CaKoLaTiMiLeSnTr08,KoLaMi08a,MaKoBlLe08,KoErGoMi10,KoGoErMi10,Kom11}.
@@
 \end{lyxcode}
-The directory \texttt{\small SPECFEM3D\_V2.0.0} will then contain
+The directory \texttt{\small SPECFEM3D} will then contain
 the source code. To configure the software for your system, run the
@@
 A recent version of its source code is provided in directory \texttt{src/decompose\_mesh\_SCOTCH/scotch\_5.1.10b}.
-The path to an existing SCOTCH installation can to be set with the option \texttt{-{}-with-scotch-dir}. Just as an example:
+In case no SCOTCH libraries can be found on the system, the configuration will bundle this version for compilation.
+The path to an existing SCOTCH installation can to be set explicitly with the option \texttt{-{}-with-scotch-dir}. Just as an example:
 \begin{lyxcode}
@@
 
-Prior to compiling \texttt{decompose\_mesh\_SCOTCH}, make sure you have correctly specified the path
-of the SCOTCH library with the option \texttt{-{}-with-scotch-dir} of the \texttt{configure} script.
-For convenience we provide the source code of SCOTCH, which is released open source under the French CeCILL-C version 1 license,
-in directory ~\\
-\texttt{src/decompose\_mesh\_SCOTCH/scotch\_5.1.10b}. Please refer to file INSTALL.txt in that directory to see how to compile it.
+In most cases, the configuration with \texttt{./configure~FC=ifort} should be sufficient.
+During the configuration process, the script tries to find existing SCOTCH installations. 
+In case your system has no pre-existing SCOTCH installation,
+we provide the source code of SCOTCH, which is released open source under the French CeCILL-C version 1 license,
+in directory 
+\texttt{src/decompose\_mesh\_SCOTCH/scotch\_5.1.10b}.
+This version gets bundled with the compilation of the SPECFEM3D package if no libraries could have been found.
+If this automatic compilation of the SCOTCH libraries fails,
+please refer to file INSTALL.txt in that directory to see further details how to compile it on your system.
+In case you want to use a pre-existing installation, make sure you have correctly specified the path
+of the SCOTCH library when using the option \texttt{-{}-with-scotch-dir} with the \texttt{./configure} script.
 In the future you should be able to find more recent versions at  \url{http://www.labri.fr/perso/pelegrin/scotch/scotch\_en.html} .

### 2010-11-29 src/generate_databases/save_arrays_solver.f90

@@ -402,463 +402,3 @@
 
 
   end subroutine save_arrays_solver_ext_mesh
-
-
-
-!=============================================================
-!
-!! old way
-!! regular mesh
-!
-!  subroutine save_arrays_solver(flag_sediments,not_fully_in_bedrock,rho_vp,rho_vs,prname,xixstore,xiystore,xizstore, &
-!            etaxstore,etaystore,etazstore, &
-!            gammaxstore,gammaystore,gammazstore,jacobianstore, &
-!            xstore,ystore,zstore,kappastore,mustore, &
-!            ANISOTROPY, &
-!            c11store,c12store,c13store,c14store,c15store,c16store, &
-!            c22store,c23store,c24store,c25store,c26store,c33store,c34store,c35store,c36store, &
-!            c44store,c45store,c46store,c55store,c56store,c66store, &
-!            ibool,idoubling,rmass,rmass_ocean_load,npointot_oceans, &
-!            ibelm_xmin,ibelm_xmax,ibelm_ymin,ibelm_ymax,ibelm_bottom,ibelm_top, &
-!            nspec2D_xmin,nspec2D_xmax,nspec2D_ymin,nspec2D_ymax, &
-!            normal_xmin,normal_xmax,normal_ymin,normal_ymax,normal_bottom,normal_top, &
-!            jacobian2D_xmin,jacobian2D_xmax,jacobian2D_ymin,jacobian2D_ymax, &
-!            jacobian2D_bottom,jacobian2D_top, &
-!            iMPIcut_xi,iMPIcut_eta,nspec,nglob, &
-!            NSPEC2DMAX_XMIN_XMAX,NSPEC2DMAX_YMIN_YMAX,NSPEC2D_BOTTOM,NSPEC2D_TOP,OCEANS)
-!
-!  implicit none
-!
-!  include "constants.h"
-!
-!  integer nspec,nglob
-!  integer NSPEC2DMAX_XMIN_XMAX,NSPEC2DMAX_YMIN_YMAX,NSPEC2D_BOTTOM,NSPEC2D_TOP
-!  integer npointot_oceans
-!
-!  logical OCEANS
-!  logical ANISOTROPY
-!
-!! arrays with jacobian matrix
[…]

### 2010-12-01 doc/USER_MANUAL/manual_SPECFEM3D.tex
adds paper draft to SPECFEM3D/doc

@@
 In older versions of the SPECFEM3D package, a common format was used for the channel codes of all seismograms, which was \texttt{BHE/BHN/BHZ} for three components.
-To avoid confusion when comparison are made to observed data, we are now using the IRIS convention for SEM seismograms.
-In the following, we give a brief explanation of the IRIS convention and how it is adopted in SEM seismograms.
+To avoid confusion when comparison are made to observed data, we are now using the FDSN convention (\url{http://www.fdsn.org/}) for SEM seismograms.
+In the following, we give a brief explanation of the FDSN convention used by IRIS, and how it is adopted in SEM seismograms.
 Please visit \texttt{www.iris.edu/manuals/SEED\_appA.htm} for further information.\\
@@
 The sampling rate of SEM synthetics is controlled by the resolution of simulations rather than instrument properties.
-However, for consistency, we follow the IRIS convention for SEM seismograms governed by their sampling rate.
+However, for consistency, we follow the FDSN convention for SEM seismograms governed by their sampling rate.
 For SEM synthetics, we consider band codes for which $dt \leq 1$ s.
@@
 
-\caption{The IRIS band code convention is based on the sampling rate and the response band of instruments.
+\caption{The band code convention is based on the sampling rate and the response band of instruments.
 Please visit \texttt{www.iris.edu/manuals/SEED\_appA.htm} for further information.
@@
 For instance, \texttt{H} and \texttt{L} are used for high-gain and low-gain seismometers, respectively.
-The instrument code of SEM seismograms will always be \texttt{X}, as assigned by IRIS for synthetic seismograms.
+The instrument code of SEM seismograms will always be \texttt{X}, as assigned by FDSN for synthetic seismograms.
 \\

### 2010-12-08 doc/USER_MANUAL/manual_SPECFEM3D.tex
added ref to PuAmKa09 about DG for fault rupture dynamics

@@
 based upon the spectral-element method (SEM).
-The SEM is a continuous Galerkin technique, which can easily be made discontinous \citep{BeMaPa94,KoWoHu02,ChCaVi03,Kop06,WiStBuGh10}; it is then a particular case of the discontinuous Galerkin technique \citep{ReHi73,Arn82,FaRi99,HuHuRa99,CoKaSh00,GiHeWa02,RiWh03,MoRi05,GrScSc06,AiMoMu06,BeLaPi06,DuKa06,DeSeWh08,WiStBuGh10,DeSe10,EtChViGl10}, with optimized efficiency because of its tensorized basis functions.
+The SEM is a continuous Galerkin technique, which can easily be made discontinous \citep{BeMaPa94,KoWoHu02,ChCaVi03,Kop06,WiStBuGh10}; it is then a particular case of the discontinuous Galerkin technique \citep{ReHi73,Arn82,FaRi99,HuHuRa99,CoKaSh00,GiHeWa02,RiWh03,MoRi05,GrScSc06,AiMoMu06,BeLaPi06,DuKa06,DeSeWh08,PuAmKa09,WiStBuGh10,DeSe10,EtChViGl10}, with optimized efficiency because of its tensorized basis functions.
 In particular, it can accurately handle very distorted mesh elements \citep{OlSe11}.
 Note that in most geological models in the context of seismic wave propagation studies
-(except for fault dynamic rupture studies)
+(except for fault dynamic rupture studies, in which very high frequencies need to be modeled near the fault, see e.g. \cite{PuAmKa09})
 a discontinous mesh is not needed because material property contrasts are not drastic and thus a continuous formulation is sufficient;

### 2010-12-10 doc/USER_MANUAL/manual_SPECFEM3D.tex
added ref to LaWaBe05

@@
 based upon the spectral-element method (SEM).
-The SEM is a continuous Galerkin technique, which can easily be made discontinous \citep{BeMaPa94,KoWoHu02,ChCaVi03,Kop06,WiStBuGh10}; it is then a particular case of the discontinuous Galerkin technique \citep{ReHi73,Arn82,FaRi99,HuHuRa99,CoKaSh00,GiHeWa02,RiWh03,MoRi05,GrScSc06,AiMoMu06,BeLaPi06,DuKa06,DeSeWh08,PuAmKa09,WiStBuGh10,DeSe10,EtChViGl10}, with optimized efficiency because of its tensorized basis functions.
+The SEM is a continuous Galerkin technique, which can easily be made discontinous \citep{BeMaPa94,KoWoHu02,ChCaVi03,LaWaBe05,Kop06,WiStBuGh10}; it is then a particular case of the discontinuous Galerkin technique \citep{ReHi73,Arn82,FaRi99,HuHuRa99,CoKaSh00,GiHeWa02,RiWh03,MoRi05,GrScSc06,AiMoMu06,BeLaPi06,DuKa06,DeSeWh08,PuAmKa09,WiStBuGh10,DeSe10,EtChViGl10}, with optimized efficiency because of its tensorized basis functions.
 In particular, it can accurately handle very distorted mesh elements \citep{OlSe11}.

### 2011-01-17 doc/USER_MANUAL/manual_SPECFEM3D.tex
added a reference to AcKo11

--- 2010-12-08 doc/USER_MANUAL/manual_SPECFEM3D.tex
@@
 based upon the spectral-element method (SEM).
-The SEM is a continuous Galerkin technique, which can easily be made discontinous \citep{BeMaPa94,KoWoHu02,ChCaVi03,Kop06,WiStBuGh10}; it is then a particular case of the discontinuous Galerkin technique \citep{ReHi73,Arn82,FaRi99,HuHuRa99,CoKaSh00,GiHeWa02,RiWh03,MoRi05,GrScSc06,AiMoMu06,BeLaPi06,DuKa06,DeSeWh08,PuAmKa09,WiStBuGh10,DeSe10,EtChViGl10}, with optimized efficiency because of its tensorized basis functions.
+The SEM is a continuous Galerkin technique, which can easily be made discontinous \citep{BeMaPa94,KoWoHu02,ChCaVi03,LaWaBe05,Kop06,WiStBuGh10,AcKo11}; it is then a particular case of the discontinuous Galerkin technique \citep{ReHi73,Arn82,FaRi99,HuHuRa99,CoKaSh00,GiHeWa02,RiWh03,MoRi05,GrScSc06,AiMoMu06,BeLaPi06,DuKa06,DeSeWh08,PuAmKa09,WiStBuGh10,DeSe10,EtChViGl10}, with optimized efficiency because of its tensorized basis functions \citep{WiStBuGh10,AcKo11}.
 In particular, it can accurately handle very distorted mesh elements \citep{OlSe11}.

### 2011-01-19 doc/USER_MANUAL/manual_SPECFEM3D.tex
done modifying the Sesame paper; made similar modifications in the introduction of both manuals

--- 2010-12-08 doc/USER_MANUAL/manual_SPECFEM3D.tex
@@
 University of Pau / CNRS / INRIA (France)\\
-Version 2.0.0 \\
+Version 2.0 \\
 "Sesame"}
@@
 based upon the spectral-element method (SEM).
-The SEM is a continuous Galerkin technique, which can easily be made discontinous \citep{BeMaPa94,KoWoHu02,ChCaVi03,Kop06,WiStBuGh10}; it is then a particular case of the discontinuous Galerkin technique \citep{ReHi73,Arn82,FaRi99,HuHuRa99,CoKaSh00,GiHeWa02,RiWh03,MoRi05,GrScSc06,AiMoMu06,BeLaPi06,DuKa06,DeSeWh08,PuAmKa09,WiStBuGh10,DeSe10,EtChViGl10}, with optimized efficiency because of its tensorized basis functions.
-In particular, it can accurately handle very distorted mesh elements \citep{OlSe11}.
+The SEM is a continuous Galerkin technique, which can easily be made discontinous \citep{BeMaPa94,Ch00,KoWoHu02,ChCaVi03,LaWaBe05,Kop06,WiStBuGh10,AcKo11}; it is then a particular case of the discontinuous Galerkin technique \citep{ReHi73,Arn82,FaRi99,HuHuRa99,CoKaSh00,GiHeWa02,RiWh03,MoRi05,GrScSc06,AiMoMu06,BeLaPi06,DuKa06,DeSeWh08,PuAmKa09,WiStBuGh10,DeSe10,EtChViGl10}, with optimized efficiency because of its tensorized basis functions \citep{WiStBuGh10,AcKo11}.
+In particular, it can accurately handle very distorted mesh elements \citep{OlSe11}.\\
+
+It has very good convergence and accuracy properties \citep{MaPa89,SePr94,DeFiMu02,Coh02,DeSe07,SeOl08}.
+It is also very well suited to parallel implementation on very large supercomputers \citep{KoTr02a,KoTsChTr03,TsKoChTr03,KoLaMi08a,CaKoLaTiMiLeSnTr08,KoViCh10} as well as on clusters of GPU accelerating graphics cards \citep{KoMiEr09,KoErGoMi10,Kom11}.
+Tensor products inside each element can be optimized to reach very high efficiency \citep{DeFiMu02}, and mesh point and element numbering can be optimized to reduce processor cache misses and improve cache reuse \citep{KoLaMi08a}.
+The SEM can also handle triangular (in 2D) or tetrahedral (in 3D) elements \citep{WinBoyd96,TaWi00,KoMaTrTaWi01,Coh02,MeViSa06} as well as mixed meshes, although with increased cost and reduced accuracy in these elements, as in the discontinuous Galerkin method.\\
+
 Note that in most geological models in the context of seismic wave propagation studies

### 2011-01-20 doc/USER_MANUAL/manual_SPECFEM3D.tex
made changes to the introduction of the manuals

@@
 based upon the spectral-element method (SEM).
-The SEM is a continuous Galerkin technique, which can easily be made discontinous \citep{BeMaPa94,Ch00,KoWoHu02,ChCaVi03,LaWaBe05,Kop06,WiStBuGh10,AcKo11}; it is then a particular case of the discontinuous Galerkin technique \citep{ReHi73,Arn82,FaRi99,HuHuRa99,CoKaSh00,GiHeWa02,RiWh03,MoRi05,GrScSc06,AiMoMu06,BeLaPi06,DuKa06,DeSeWh08,PuAmKa09,WiStBuGh10,DeSe10,EtChViGl10}, with optimized efficiency because of its tensorized basis functions \citep{WiStBuGh10,AcKo11}.
+The SEM is a continuous Galerkin technique, which can easily be made discontinous \citep{BeMaPa94,Ch00,KoWoHu02,ChCaVi03,LaWaBe05,Kop06,WiStBuGh10,AcKo11}; it is then close to a particular case of the discontinuous Galerkin technique \citep{ReHi73,Arn82,FaRi99,HuHuRa99,CoKaSh00,GiHeWa02,RiWh03,MoRi05,GrScSc06,AiMoMu06,BeLaPi06,DuKa06,DeSeWh08,PuAmKa09,WiStBuGh10,DeSe10,EtChViGl10}, with optimized efficiency because of its tensorized basis functions \citep{WiStBuGh10,AcKo11}.
 In particular, it can accurately handle very distorted mesh elements \citep{OlSe11}.\\
 
-It has very good convergence and accuracy properties \citep{MaPa89,SePr94,DeFiMu02,Coh02,DeSe07,SeOl08}.
+It has very good accuracy and convergence properties \citep{MaPa89,SePr94,DeFiMu02,Coh02,DeSe07,SeOl08}.
 It is also very well suited to parallel implementation on very large supercomputers \citep{KoTr02a,KoTsChTr03,TsKoChTr03,KoLaMi08a,CaKoLaTiMiLeSnTr08,KoViCh10} as well as on clusters of GPU accelerating graphics cards \citep{KoMiEr09,KoErGoMi10,Kom11}.
@@
 a discontinous mesh is not needed because material property contrasts are not drastic and thus a continuous formulation is sufficient;
-conforming mesh doubling bricks can efficiently handle mesh size variations \citep{KoTr02a,KoLiTrSuStSh04}.\\
+conforming mesh doubling bricks can efficiently handle mesh size variations \citep{KoTr02a,KoLiTrSuStSh04,LeChLiKoHuTr08,LeChKoHuTr09,LeKoHuTr09}.\\
 

### 2011-01-23 doc/USER_MANUAL/manual_SPECFEM3D.tex
added a reference to GeRe09 and also to DaNa98

@@
 The SPECFEM3D software package relies on the SCOTCH library to partition meshes created with CUBIT.
-Note that we use CUBIT to create meshes of hexahedra, but other packages can be used as well, for instance GiD from \url{http://gid.cimne.upc.es} or Gmsh from \url{http://geuz.org/gmsh}.
+Note that we use CUBIT to create meshes of hexahedra, but other packages can be used as well, for instance GiD from \url{http://gid.cimne.upc.es} or Gmsh from \url{http://geuz.org/gmsh} \citep{GeRe09}.
 Even mesh creation packages that generate tetrahedra, for instance TetGen from \url{http://tetgen.berlios.de}, can be used because each tetrahedron can then easily be decomposed into four hexahedra as shown in the picture of the TetGen logo at \url{http://tetgen.berlios.de/figs/Delaunay-Voronoi-3D.gif}; while this approach does not generate hexahedra of optimal quality, it can ease mesh creation in some situations and it has been shown that the spectral-element method can very accurately handle distorted mesh elements \citep{OlSe11}.

### 2011-01-23 doc/USER_MANUAL/manual_SPECFEM3D.tex
removed Emiljana from the current list of authors, since she will contribute to the GPU version but has not worked on the current CPU version

@@
 Vala Hj\"orleifsd\'ottir,
-Emiljana Jorgji,
+%%%%Emiljana Jorgji,
 Sue Kientz,

### 2011-01-23 doc/USER_MANUAL/manual_SPECFEM3D.tex
added a description of how to check the quality of a mesh generated using our in-house mesher

@@
 `\texttt{\small gnuplot plot\_mesh\_quality\_histogram.gnu}'). The
-histogram is also printed to the screen. If you want to start designing
+histogram is also printed to the screen.
+Analyze that skewness histogram of mesh elements to make sure no element has a skewness above approximately 0.75, otherwise the mesh is of poor quality (and if even a single element has a skewness value above 0.80, then you must definitely improve the mesh).
+If you want to start designing
 your own meshes, this tool is useful for viewing your creations. You
@@
 \end{lyxcode}
+
+To control the quality of the mesh, check the standard output (either on the screen or in the \texttt{in\_out\_files/OUTPUT\_FILES} directory in \texttt{output\_mesher.txt}) and analyze the skewness histogram of mesh elements to make sure no element has a skewness above approximately 0.75, otherwise the mesh is of poor quality (and if even a single element has a skewness value above 0.80, then you must definitely improve the mesh).
+To draw the skewness histogram on the screen, type \texttt{gnuplot plot\_mesh\_quality\_histogram.gnu}.
 

### 2011-01-27 doc/USER_MANUAL/manual_SPECFEM3D.tex
updates movie_volume info in SPECFEM3D manual

@@
 flag turned on, the code outputs \texttt{moviedata??????} files in
-the \texttt{in\_out\_files/OUTPUT\_FILES} directory. The files are in a fairly complicated
+the \texttt{in\_out\_files/OUTPUT\_FILES} directory. There are several flags
+in the main parameter file \texttt{Par\_file} that control the output of these moviedata files 
+(see section \ref{cha:Main-Parameter} for details):
+\texttt{\small NTSTEP\_BETWEEN\_FRAMES} to set the timesteps between frames,
+\texttt{\small SAVE\_DISPLACEMENT} to save displacement instead of velocity, 
+\texttt{USE\_HIGHRES\_FOR\_MOVIES} to save values at all GLL point instead of element edges.
+In order to output additionally shakemaps, you would set
+the parameter \texttt{CREATE\_SHAKEMAP} to \texttt{.true.}.
+
+The files are in a fairly complicated
 binary format, but there is a program provided to convert the
 output into more user friendly formats:
-
 \begin{description}
@@
 animations of southern California earthquakes via the web; see Southern California ShakeMovie\textregistered \urlwithparentheses{www.shakemovie.caltech.edu}.
+
+\subsection{\label{sub:Movie-Volume}Movie Volume}
+
+When running xspecfem3D with the \texttt{\small MOVIE\_VOLUME} flag
+turned on, the code outputs several files in \texttt{\small LOCAL\_PATH} specified in the main
+\texttt{Par\_file}, e.g. in directory \texttt{in\_out\_files/DATABASES\_MPI}.
+%As the files can be very large, there are several flags in the \texttt{\small Par\_file}
+%that control the region in space and time that is saved. These are:
+%\texttt{\small NTSTEP\_BETWEEN\_FRAMES}
+The output is saved by each processor
+at the time interval specified by \texttt{\small NTSTEP\_BETWEEN\_FRAMES}.
+For all domains, the velocity field is output to files: 
+\texttt{\small proc??????\_velocity\_X\_it??????.bin},
+\texttt{\small proc??????\_velocity\_Y\_it??????.bin} and
+\texttt{\small proc??????\_velocity\_Z\_it??????.bin}.
+For elastic domains, the divergence and curl taken from the velocity field, i.e. $\nabla \cdot \bf{v}$
+and $\nabla \times \bf{v}$, get stored as well: 
+\texttt{\small proc??????\_div\_it??????.bin},
+\texttt{\small proc??????\_curl\_X\_t??????.bin},
+\texttt{\small proc??????\_curl\_Y\_it??????.bin} and
+\texttt{\small proc??????\_curl\_Z\_it??????.bin}.
+The files denoted 
+\texttt{\small proc??????\_div\_}~\\
+\texttt{\small glob\_it??????.bin} and
+\texttt{\small proc??????\_curl\_glob\_it??????.bin} 
+are stored on the global points only, all the other arrays
+are stored on all GLL points. Note that the components X/Y/Z can change to E/N/Z according to 
+the \texttt{SUPPRESS\_UTM\_PROJECTION} flag (see also Appendix~\ref{cha:Coordinates} and \ref{cha:channel-codes}).
+
+\begin{figure}[htbp]
+\noindent \begin{centering}
+\includegraphics[width=.32\textwidth]{figures/movie_volume_1.jpg}
+\includegraphics[width=.32\textwidth]{figures/movie_volume_2.jpg}
+\includegraphics[width=.32\textwidth]{figures/movie_volume_3.jpg}
+\par\end{centering}
+
+\caption{Paraview visualization using movie volume files (converted by \texttt{xcombine\_vol\_data} and \texttt{mesh2vtu.pl}) and showing
+snapshots of vertical velocity components at different times.}
+
+\label{fig:movie.volume}
+\end{figure}
+
+To visualize these files, we use an auxilliary program \texttt{combine\_vol\_data.f90}
+to combine the data from all slices into one mesh file. To compile it in the root directory, type:
+\begin{lyxcode}
+{\footnotesize make~combine\_vol\_data~}{\footnotesize \par}
+\end{lyxcode}
+which will create the executable \texttt{xcombine\_vol\_data} in the directory \texttt{bin/}. To output the
+usage of this executable, type `./xcombine\_vol\_data` without arguments.
+As an example, you would use
+\begin{lyxcode}
+{\footnotesize cd~bin/~}{\footnotesize \par}
+{\footnotesize ./xcombine\_vol\_data~0~3~velocity\_Z\_it000400~../in\_out\_files/DATABASES\_MPI~../in\_out\_files/OUTPUT\_FILES~0~}{\footnotesize \par}
+\end{lyxcode}
+to create a low-resolution mesh file. The output mesh file will have the name \texttt{velocity\_Z\_it000400.mesh}.
+We next convert the \texttt{.mesh} file into the VTU (Unstructured
+grid file) format which can be viewed in ParaView. For this task, you can use and modify the script \texttt{mesh2vtu.pl}
+located in directory \texttt{utils/Visualization/Paraview/}, for example:
+\begin{lyxcode}
+{\footnotesize mesh2vtu.pl~-i~velocity\_Z\_it000400.mesh~-o~velocity\_Z\_it000400.vtu}{\footnotesize \par}
+\end{lyxcode}
+Notice that this Perl script uses a program \texttt{mesh2vtu} in the
+\texttt{utils/Visualization/Paraview/mesh2vtu} directory, which further uses the VTK \urlwithparentheses{http://www.vtk.org/}
+run-time library for its execution. Therefore, make sure you have
+them properly set in the script according to your system.
 

### 2011-01-30 doc/USER_MANUAL/manual_SPECFEM3D.tex
generated new small versions of the covers for the Web site, and removed old comments from the list of authors.

@@
 \section*{Authors}
-The SPECFEM3D package was first developed by Dimitri Komatitsch and Jeroen Tromp at Harvard University and Caltech, USA, starting in 1998, based in part on earlier work by Dimitri Komatitsch and Jean-Pierre Vilotte at Institut de Physique du Globe (IPGP) in Paris, France from 1994 to 1997.\\
+The SPECFEM3D package was first developed by Dimitri Komatitsch and Jeroen Tromp at Harvard University and Caltech, USA, starting in 1998, based in part on earlier work by Dimitri Komatitsch and Jean-Pierre Vilotte at Institut de Physique du Globe (IPGP) in Paris, France from 1995 to 1997.\\
 
@@
 Vala Hj\"orleifsd\'ottir,
-%%%%Emiljana Jorgji,
 Sue Kientz,

### 2011-01-31 doc/USER_MANUAL/manual_SPECFEM3D.tex
updates check_mesh_quality_CUBIT_Abaqus/ and manual; updates paper_draft/

@@
 directory in which you want to install the package, type
-
 \begin{lyxcode}
@@
 \end{lyxcode}
-The directory \texttt{\small SPECFEM3D} will then contain
+The directory \texttt{\small SPECFEM3D/} will then contain
 the source code. To configure the software for your system, run the
@@
 The quality of the mesh may be inspected more precisely based upon the serial code
-in the file  ~\\
-\texttt{check\_mesh\_quality\_CUBIT\_Abaqus.f90}.
+in the file  \texttt{check\_mesh\_quality\_}~\\
+\texttt{CUBIT\_Abaqus.f90} located in the directory \texttt{src/check\_mesh\_quality\_CUBIT\_Abaqus/}.
 Running this code is optional because no information needed by the
@@
 Prior to running and compiling this code, you have to export your mesh in CUBIT to an ABAQUS (.inp) format.
+For example, export mesh block IDs belonging to volumes in order to check the quality of the hexahedral elements.
 You also have to determine a number of parameters of your mesh, such as the number of nodes and number of elements
 and modify the header of the \texttt{check\_mesh\_quality\_CUBIT\_Abaqus.f90} source file. Then,
-in the directory \texttt{src/check\_mesh\_quality\_CUBIT\_Abaqus}, type
-\begin{lyxcode}
-{\small ./compile\_all.csh}{\small \par}
+in the main directory, type
+\begin{lyxcode}
+{\small make xcheck\_mesh\_quality\_CUBIT\_Abaqus}{\small \par}
 \end{lyxcode}
@@
 \begin{lyxcode}
-{\small xcheck\_mesh\_quality\_CUBIT\_Abaqus}{\small \par}
+{\small ./bin/xcheck\_mesh\_quality\_CUBIT\_Abaqus}{\small \par}
 \end{lyxcode}
@@
 When you are ready to compile,
-in the main directory type '\texttt{make decompose\_mesh\_SCOTCH}'.
+in the main directory type
+\begin{lyxcode}
+{\small make xdecompose\_mesh\_SCOTCH}{\small \par}
+\end{lyxcode}
 If all paths and flags have been set correctly, the executable \texttt{bin/xdecompose\_mesh\_SCOTCH} should be produced.
@@
 
-In the main directory type `\texttt{make meshfem3D}'. If all paths and flags
+In the main directory, type 
+\begin{lyxcode}
+{\small make xmeshfem3D}{\small \par}
+\end{lyxcode}
+If all paths and flags
 have been set correctly, the mesher should now compile and produce
@@
 
-After using \texttt{xmeshfem3D} or \texttt{xdecompose\_mesh\_SCOTCH}, the next step in the workflow is to compile \texttt{xgenerate\_databases}.
-This program is going to create all the missing information needed by the SEM solver.
+After using \texttt{xmeshfem3D} or \texttt{xdecompose\_mesh\_SCOTCH}, the next step in the workflow is to compile 
+\texttt{xgenerate\_}~\\
+\texttt{databases}. This program is going to create all the missing information needed by the SEM solver.
 
@@
 
-In the main directory type '\texttt{make generate\_databases}'. Input for the program is provided through
+In the main directory, type 
+\begin{lyxcode}
+{\small make xgenerate\_databases}{\small \par}
+\end{lyxcode}
+Input for the program is provided through
 the main parameter file \texttt{Par\_file}, which resides in the subdirectory \texttt{in\_data\_files/}. Please note that
@@
 Now that you have successfully generated the databases, you are ready to compile
-the solver. In the main directory type `\texttt{make specfem3D}'.
+the solver. In the main directory, type
+\begin{lyxcode}
+{\small make xspecfem3D}{\small \par}
+\end{lyxcode}
 Please note that \texttt{xspecfem3D} must be called directly from
@@
 Type
-
-\begin{lyxcode}
-make~convolve\_source\_timefunction
+\begin{lyxcode}
+{\small make xconvolve\_source\_timefunction}{\small \par}
 \end{lyxcode}
@@
 
-\texttt{make generate\_databases}
-
-\texttt{make specfem3D} \\
+\texttt{make xgenerate\_databases}
+
+\texttt{make xspecfem3D} \\
 
@@
 in the  \texttt{in\_out\_files/OUTPUT\_FILES/} directory. This file will be created by the solver run.
-Then type `\texttt{make} \texttt{create\_movie\_shakemap\_AVS\_DX\_GMT}'
+Then type 
+\begin{lyxcode}
+{\footnotesize make xcreate\_movie\_shakemap\_AVS\_DX\_GMT}{\footnotesize \par}
+\end{lyxcode}
 and run the executable \texttt{xcreate\_movie\_shakemap\_AVS\_DX\_GMT}
@@
 \begin{lyxcode}
-{\footnotesize make~combine\_vol\_data~}{\footnotesize \par}
+{\footnotesize make~xcombine\_vol\_data~}{\footnotesize \par}
 \end{lyxcode}
 which will create the executable \texttt{xcombine\_vol\_data} in the directory \texttt{bin/}. To output the
-usage of this executable, type `./xcombine\_vol\_data` without arguments.
-As an example, you would use
+usage of this executable, type `./bin/xcombine\_vol\_data` without arguments.
+As an example, to run the executable you would use
 \begin{lyxcode}
@@
 \begin{lyxcode}
-{\footnotesize make~combine\_vol\_data~}{\footnotesize \par}
+{\footnotesize make~xcombine\_vol\_data~}{\footnotesize \par}
 
@@
 \begin{lyxcode}
-mesh2vtu.pl~-i~file.mesh~-o~file.vtu
+{\footnotesize mesh2vtu.pl~-i~file.mesh~-o~file.vtu~}{\footnotesize \par}
 \end{lyxcode}
@@
 
-make~meshfem3D~
-
-make~generate\_databases~
-
-make~specfem3D~
+make~xmeshfem3D~
+
+make~xgenerate\_databases~
+
+make~xspecfem3D~
 
@@
 An example of the usage of the script:
-
 \begin{lyxcode}
@@
 types of solvers separately through a scheduler (refer to Chapter~\ref{cha:Scheduler}).
-
 \begin{lyxcode}

### 2011-02-01 src/shared/combine_vol_data.f90
re-adds TOPOGRAPHY flag to use with ocean load approximation; updates Makefiles; updates layered_halfspace example

@@ -71,7 +71,7 @@
   logical :: MOVIE_SURFACE,MOVIE_VOLUME,CREATE_SHAKEMAP,SAVE_DISPLACEMENT, &
             USE_HIGHRES_FOR_MOVIES,SUPPRESS_UTM_PROJECTION
   logical :: ATTENUATION,USE_OLSEN_ATTENUATION, &
-            OCEANS
+            OCEANS,TOPOGRAPHY
   logical :: ABSORBING_CONDITIONS,SAVE_FORWARD
   logical :: ANISOTROPY,SAVE_MESH_FILES,PRINT_SOURCE_TIME_FUNCTION
   character(len=256) LOCAL_PATH
@@ -146,7 +146,7 @@
   call read_parameter_file( NPROC,NTSTEP_BETWEEN_OUTPUT_SEISMOS,NSTEP,DT, &
                         UTM_PROJECTION_ZONE,SUPPRESS_UTM_PROJECTION, &
                         ATTENUATION,USE_OLSEN_ATTENUATION,LOCAL_PATH,NSOURCES, &
-                        OCEANS,ANISOTROPY,ABSORBING_CONDITIONS, &
+                        OCEANS,TOPOGRAPHY,ANISOTROPY,ABSORBING_CONDITIONS, &
                         MOVIE_SURFACE,MOVIE_VOLUME,CREATE_SHAKEMAP,SAVE_DISPLACEMENT, &
                         NTSTEP_BETWEEN_FRAMES,USE_HIGHRES_FOR_MOVIES,HDUR_MOVIE, &
                         SAVE_MESH_FILES,PRINT_SOURCE_TIME_FUNCTION, &

### 2011-02-01 layered_halfspace/process.sh
re-adds TOPOGRAPHY flag to use with ocean load approximation; updates Makefiles; updates layered_halfspace example

@@ -18,7 +18,7 @@ echo "running example: `date`"
 currentdir=`pwd`
 
 echo
-echo "(will take about 1h 10 minutes)"
+echo "(will take about 27 minutes)"
 echo
 
 
@@ -40,7 +40,9 @@ cd ../../
 # compiles with flag for a point force (with a ricker source time function)
 cd src/shared/
 cp constants.h constants.h.org
-sed -e "s:USE_FORCE_POINT_SOURCE .*:USE_FORCE_POINT_SOURCE = .true. :" constants.h.org > constants.h
+sed -e "s:USE_FORCE_POINT_SOURCE.*:USE_FORCE_POINT_SOURCE = .true. :" constants.h.org > constants.h.tmp
+sed -e "s:FACTOR_FORCE_SOURCE.*:FACTOR_FORCE_SOURCE = -1.d15 :" constants.h.tmp > constants.h
+rm -f constants.h.tmp
 cd ../../
 make > tmp.log
 

### 2011-02-01 doc/USER_MANUAL/manual_SPECFEM3D.tex
re-adds TOPOGRAPHY flag to use with ocean load approximation; updates Makefiles; updates layered_halfspace example

@@
 is on.
+\item [{\texttt{TOPOGRAPHY}}] This feature is only effective if \texttt{OCEANS} is set to \texttt{.true.}.
+Set to \texttt{.true.} if topography and
+bathymetry should be read in based upon the topography file specified in the main
+constants file \texttt{constants.h} found in subdirectory \texttt{src/shared/} to evaluate
+elevations. If not set, elevations will be read from the numerical mesh.
 \item [{\texttt{ATTENUATION}}] Set to \texttt{.true.} if attenuation should

### 2011-02-01 src/shared/read_parameter_file.f90

@@ -27,7 +27,7 @@
   subroutine read_parameter_file( NPROC,NTSTEP_BETWEEN_OUTPUT_SEISMOS,NSTEP,DT, &
                         UTM_PROJECTION_ZONE,SUPPRESS_UTM_PROJECTION, &
                         ATTENUATION,USE_OLSEN_ATTENUATION,LOCAL_PATH,NSOURCES, &
-                        OCEANS,ANISOTROPY,ABSORBING_CONDITIONS, &
+                        OCEANS,TOPOGRAPHY,ANISOTROPY,ABSORBING_CONDITIONS, &
                         MOVIE_SURFACE,MOVIE_VOLUME,CREATE_SHAKEMAP,SAVE_DISPLACEMENT, &
                         NTSTEP_BETWEEN_FRAMES,USE_HIGHRES_FOR_MOVIES,HDUR_MOVIE, &
                         SAVE_MESH_FILES,PRINT_SOURCE_TIME_FUNCTION,NTSTEP_BETWEEN_OUTPUT_INFO, &
@@ -44,7 +44,7 @@
 
   double precision DT,HDUR_MOVIE
 
-  logical ATTENUATION,USE_OLSEN_ATTENUATION,OCEANS,ABSORBING_CONDITIONS,SAVE_FORWARD
+  logical ATTENUATION,USE_OLSEN_ATTENUATION,OCEANS,TOPOGRAPHY,ABSORBING_CONDITIONS,SAVE_FORWARD
   logical MOVIE_SURFACE,MOVIE_VOLUME,CREATE_SHAKEMAP,SAVE_DISPLACEMENT,USE_HIGHRES_FOR_MOVIES
   logical ANISOTROPY,SAVE_MESH_FILES,PRINT_SOURCE_TIME_FUNCTION,SUPPRESS_UTM_PROJECTION
 
@@ -94,6 +94,8 @@
   call read_value_double_precision(DT, 'solver.DT')
   if(err_occurred() /= 0) return
   call read_value_logical(OCEANS, 'model.OCEANS')
+  if(err_occurred() /= 0) return
+  call read_value_logical(TOPOGRAPHY, 'model.TOPOGRAPHY')
   if(err_occurred() /= 0) return
   call read_value_logical(ATTENUATION, 'model.ATTENUATION')
   if(err_occurred() /= 0) return

### 2011-02-15 doc/USER_MANUAL/manual_SPECFEM3D.tex
added DG references to the manual

@@
 It has very good accuracy and convergence properties \citep{MaPa89,SePr94,DeFiMu02,Coh02,DeSe07,SeOl08}.
+The spectral element approach admits spectral rates of convergence and allows exploiting $hp$-convergence schemes.
 It is also very well suited to parallel implementation on very large supercomputers \citep{KoTr02a,KoTsChTr03,TsKoChTr03,KoLaMi08a,CaKoLaTiMiLeSnTr08,KoViCh10} as well as on clusters of GPU accelerating graphics cards \citep{KoMiEr09,KoErGoMi10,Kom11}.
@@
 Note that in most geological models in the context of seismic wave propagation studies
-(except for fault dynamic rupture studies, in which very high frequencies need to be modeled near the fault, see e.g. \cite{PuAmKa09})
+(except for fault dynamic rupture studies, in which very high frequencies or supershear rupture need to be modeled near the fault, see e.g. \cite{BeGlCrViPi07,BeGlCrVi09,PuAmKa09,TaCrEtViBeSa10})
 a discontinous mesh is not needed because material property contrasts are not drastic and thus a continuous formulation is sufficient;
-conforming mesh doubling bricks can efficiently handle mesh size variations \citep{KoTr02a,KoLiTrSuStSh04,LeChLiKoHuTr08,LeChKoHuTr09,LeKoHuTr09}.\\
+conforming mesh doubling bricks can efficiently handle mesh size variations \citep{KoTr02a,KoLiTrSuStSh04,LeChLiKoHuTr08,LeChKoHuTr09,LeKoHuTr09}.
+Also, without significant modifications in the complexity of the
+discontinuous technique \citep{CaKaBr10,SmCoObOvSch10},
+usually those methods need to assume constant material properties inside
+each element, while the SEM can handle gradients inside each element.\\
 

### 2011-02-21 doc/USER_MANUAL/manual_SPECFEM3D.tex
added references to non blocking MPI articles

@@
 \newcommand{\nchunks}{\mbox{\texttt{NCHUNKS\/}}}
+
+% colors for the text
+\usepackage[dvipsnames,usenames]{color}
+\newcommand{\red}[1]{\textcolor{Red}{#1}}
 
@@
 \end{description}
+
+\noindent
+\red{If you use this new version 2.0, which has non blocking MPI for much better performance for medium or large runs, please cite at least one of these six articles,
+in which results of non blocking MPI runs are presented: \cite{PeKoLuMaLeCaLeMaLiBlNiBaTr11,KoErGoMi10,KoViCh10,Kom11,CaKoLaTiMiLeSnTr08,MaKoBlLe08}.}\\
 

### 2011-03-27 doc/USER_MANUAL/manual_SPECFEM3D.tex
added a note about using Intel icc for Scotch

@@
 
-To compile a serial version of the code for small meshes that fit
+If you use the Intel ifort compiler to compile the code, we recommand that you use the Intel icc C compiler to compile Scotch, i.e., use:
+\begin{lyxcode}
+./configure~CC=icc~FC=ifort~MPIFC=mpif90
+\end{lyxcode}
+
+To compile a serial version of the code for small meshes that fits
 on one compute node and can therefore be run serially, run \texttt{configure}

### 2011-04-07 doc/USER_MANUAL/manual_SPECFEM3D.tex
fixed a typo
adds utils/Cubit_import folder for scripts to deal with CUBIT meshing; updates manual

@@
 \noindent
-\red{If you use this new version 2.0, which has non blocking MPI for much better performance for medium or large runs, please cite at least one of these six articles,
-in which results of non blocking MPI runs are presented: \cite{PeKoLuMaLeCaLeMaLiBlNiBaTr11,KoErGoMi10,KoViCh10,Kom11,CaKoLaTiMiLeSnTr08,MaKoBlLe08}.}\\
+If you use this new version 2.0, which has non blocking MPI for much better performance for medium or large runs, please cite at least one of these six articles,
+in which results of non blocking MPI runs are presented: \cite{PeKoLuMaLeCaLeMaLiBlNiBaTr11,KoErGoMi10,KoViCh10,Kom11,CaKoLaTiMiLeSnTr08,MaKoBlLe08}.\\
 
@@
 
-If you use the Intel ifort compiler to compile the code, we recommand that you use the Intel icc C compiler to compile Scotch, i.e., use:
+If you use the Intel ifort compiler to compile the code, we recommend that you use the Intel icc C compiler to compile Scotch, i.e., use:
 \begin{lyxcode}
@@
 in the \texttt{STATIONS\_ADJOINT} list, we supply a sample program
-in \texttt{utils/cut\_velocity/xcut\_velocity} to cut a certain portion of the original
+in \texttt{utils/adjoint\_sources/traveltime/xcreate\_adjsrc\_traveltime} to cut a certain portion of the original
 displacement seismograms and convert it into the proper adjoint source
 to compute the finite-frequency kernel.
-
-\begin{lyxcode}
-xcut\_velocity~t1~t2~ifile{[}0-5]~E/N/Z-ascii-files~{[}baz]
+\begin{lyxcode}
+xcreate\_adjsrc\_traveltime~t1~t2~ifile{[}0-5]~E/N/Z-ascii-files~{[}baz]
 \end{lyxcode}
@@
 
+\item Similarly, a sample program to compute adjoint sources for amplitude finite-frequency kernels 
+may be found in \texttt{utils/adjoint\_sources/amplitude} and used in the same way as described for 
+traveltime measurements
+\begin{lyxcode}
+xcreate\_adjsrc\_amplitude~t1~t2~ifile{[}0-5]~E/N/Z-ascii-files~{[}baz].
+\end{lyxcode}
+
 \end{itemize}
+
 \item \textbf{Run the kernel simulation}
@@
   fix: e.g. specify > ./configure FC=gfortran MPIF90=/usr/local/openmpi-gfortran/bin/mpif90
+
+\item [after executing \texttt{xmeshfem3D} I've got elements with skewness of 81\% percent, what does this mean:] 
+Look at the skewness table printed in the \texttt{output\_mesher.txt} file after executing \texttt{xmeshfem3D} for the
+example given in \texttt{examples/meshfem3D\_examples/simple\_model/}:
+\begin{lyxcode}
+...~\\
+histogram of skewness (0. good - 1. bad): ~\\
+ 0.0000000E+00  -   5.0000001E-02       27648     81.81818      \% ~\\
+ 5.0000001E-02  -   0.1000000               0    0.0000000E+00  \% ~\\
+... ~\\
+\end{lyxcode}
+The first line means that you have 27,648 elements with a skewness value between 0 and 0.05 (which means the element is basically not skewed, just plain regular hexahedral element).
+The total number of elements you have in this mesh is (see in the \texttt{output\_mesher.txt} file a bit further down):
+\begin{lyxcode}
+...~\\
+total number of elements in entire mesh:        33792 ~\\
+... ~\\
+\end{lyxcode}
+which gives you that:  27,648 / 33,792 $\sim$ 81.8 \% of all elements are not skewed, i.e. regular elements. a fantastic value :)
+
+The histogram lists for this mesh also some stronger skewed elements, for example the worst ones belong to:
+\begin{lyxcode}
+...~\\
+0.6000000      -   0.6500000            2048     6.060606      \% ~\\
+... ~\\
+\end{lyxcode}
+
+about 6 \% of all elements have distortions with a skewness value between 0.6 and 0.65.
+The skewness values give you a hint of how good your mesh is.
+In an ideal world, you would want to have no distortions, just like the 81\% from above.
+Those elements give you the best approximate values by the GLL quadrature used in the spectral-element method.
+However, having weakly distorted elements is still fine and the solutions are still accurate enough.
+So empirically, values up to around 0.7 are tolerable, above that you should consider remeshing...
+
+To give you an idea why some of the elements are distorted, see the following figure \ref{fig:mesh.vp} of the mesh you obtain in the example \texttt{examples/meshfem3D\_examples/simple\_model/}. 
+\begin{figure}[htbp]
+\noindent \begin{centering}
+\includegraphics[width=.9\textwidth]{figures/mesh_vp.jpg}
+\par\end{centering}
+\caption{Paraview visualization using the mesh vtk-files for the example given in \texttt{examples/meshfem3D\_examples/simple\_model/}.}
+\label{fig:mesh.vp}
+\end{figure}
+You will see that the mesh contains a doubling layer, where we stitch elements together such that the size of two elements will transition to the size of one element (very useful to keep the ratio of wavespeed / element\_size about constant).
+Those elements in this doubling layer have higher skewness values and make up those 6 \% in the histogram.
+
+\item [the code gives following error message "need at least one receiver":] 
+This means that no stations given in the input file \texttt{in\_data\_files/STATIONS} could be located within the dimensions of the mesh.
+This can happen for example when the mesh was created with the in-house mesher \texttt{xmeshfem3D} while using the Universal Transverse Mercator (UTM) projection but the simulation with \texttt{xspecfem3D} was suppressing this projection from latitude/longitude to x/y/z coordinates.
+
+In such cases, try to change your \texttt{in\_data\_files/Par\_file} and set e.g.:
+\begin{lyxcode}
+SUPPRESS\_UTM\_PROJECTION         = .false. ~\\
+\end{lyxcode}
+to be the same in \texttt{Mesh\_Par\_file} and \texttt{Par\_file}.
+This flag should be identical when using the in-house mesher \texttt{xmeshfem3D}, \texttt{xgenerate\_databases} and \texttt{xspecfem3D} together to run simulations.
+
+The flag determines if the coordinates you specify for your source and station locations are given as lat/lon degrees and
+must be converted to UTM coordinates.
+As an example, if you use \texttt{.false.} within \texttt{Mesh\_Par\_file} then you create a mesh with \texttt{xmeshfem3D} using the UTM projection from lat/lon as input format to UTM projected coordinates to store the mesh point positions, which is fine.
+The error then may occur if in the \texttt{Par\_file} you have this set to \texttt{.true.} so that the \texttt{xgenerate\_databases} and \texttt{xspecfem3D} suppress the UTM projection and assume that all coordinates you use now for source and receiver locations are given in meters (that is, converted) already.
+So it won't find the specified locations in the used mesh.
+As a solutions, just change the flag in \texttt{Par\_file} to be the same as in \texttt{Mesh\_Par\_file} and rerun \texttt{xgenerate\_databases} and \texttt{xspecfem3D} to make sure that your simulation works fine.
+
+\item [I get the following error message "forward simulation became unstable and blew up":] 
+In most cases this means that your time step size \texttt{DT} is chosen too big.
+Look at your files \texttt{output\_mesher.txt} or \texttt{output\_solver.txt} created in the folder \texttt{in\_out\_files/OUTPUT\_FILES/}.
+In these output files, find the section:
+\begin{lyxcode}
+... ~\\
+********************************************* ~\\
+*** Verification of simulation parameters *** ~\\
+********************************************* ~\\
+... ~\\
+*** Minimum period resolved =    4.308774 ~\\
+*** Maximum suggested time step =   6.8863556E-02 ~\\
+... ~\\
+... ~\\
+\end{lyxcode}
+
+then change \texttt{DT} in the \texttt{in\_data\_files/Par\_file} to be somewhere close to the maximum suggested time step. In the example above: \\
+\texttt{DT                              = 0.05d0} \\
+would (most probably) work fine. It could be also bigger than the 0.068 s suggested. This depends a bit on the distortions of
+your mesh elements. The more regular they are, the bigger you can choose \texttt{DT}. Just play with this value a bit and see
+when the simulation becomes stable ...
 

### 2011-04-07 utils/Cubit_import/AUTOCAD_VTK/dxf2vtk.m

@@ -1,0 +1,132 @@
+% This function converts AUTOCAD 2000 (other ?) DXF files to VTK
+% unstructured mesh files (.vtu). The function extracts only the faces
+% represented by 'AcDbFace' tokens.
+% RUN
+% >> dxf2vtk(inpfname)
+% e.g., dxf2vtk('example.dxf')
+% INPUT
+%   inpfname: DXF input file name
+% OUTPUT
+%   .vtu file which can be visualized with ParaView/VTK
+% AUTHOR
+%   Hom Nath Gharti
+%   NORSAR
+%   homnath_AT_norsar_DOT_com
+% REVISION
+%   Feb 15,2010
+function dxf2vtk(inpf_name)
+%fprintf(1,'loading input file %s...\n',inpf_name);
+if ~exist('inpf_name','var')
+    error('dxf2vtk accepts exactly 1 argument! Enter dxf file as an argument!');
+end
+[path,fheader]=fileparts(inpf_name);
+if isempty(path)
+    path='.';
+end
+
+inpf=fopen(inpf_name,'r');
+if inpf<=0
+	error('File %s cannot be opened!\n',inpf_name);
+end
+
+% count number of 3D faces
+nface=0; nline=0;
+fprintf(1,'counting faces...');
+while(~feof(inpf))    
+    if strcmp(fgetl(inpf),'AcDbFace')
+        nface=nface+1;        
+    end
+    if nface==0
+        nline=nline+1;
+    end
+end
+fclose(inpf);
+fprintf(1,'complete!\n');
+fprintf(1,'total number of faces: %d\n',nface);
+
+fprintf(1,'extracting coordinates...');
+coord=zeros(3*nface,3);
+inpf=fopen(inpf_name,'r');
+for i_line=1:nline; fgetl(inpf); end
+for i_face=1:nface
+    inode=[1 2 3]+(i_face-1)*3;
+    fgetl(inpf);
+    strblk=textscan(inpf,'%s',18);    
+    coord(inode,:)=reshape(str2double(strblk{1}(2:2:18)),[3,3])';
+    for i_line=1:22; fgetl(inpf); end
+end
+fprintf(1,'complete!\n');
+
+fprintf(1,'removing duplicate nodes...');
+% Remove duplicates and renumber
+[ucoord, m, n]=uunique(coord);
+clear coord;
+fprintf(1,'complete!\n');
+
[…]

### 2011-04-08 src/shared/utm_geo.f90
updates source code headers; updates package version number to version 2.0.1

@@ -1,5 +1,31 @@
 !=====================================================================
 !
+!               S p e c f e m 3 D  V e r s i o n  2 . 0
+!               ---------------------------------------
+!
+!          Main authors: Dimitri Komatitsch and Jeroen Tromp
+!    Princeton University, USA and University of Pau / CNRS / INRIA
+! (c) Princeton University / California Institute of Technology and University of Pau / CNRS / INRIA
+!                            April 2011
+!
+! This program is free software; you can redistribute it and/or modify
+! it under the terms of the GNU General Public License as published by
+! the Free Software Foundation; either version 2 of the License, or
+! (at your option) any later version.
+!
+! This program is distributed in the hope that it will be useful,
+! but WITHOUT ANY WARRANTY; without even the implied warranty of
+! MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+! GNU General Public License for more details.
+!
+! You should have received a copy of the GNU General Public License along
+! with this program; if not, write to the Free Software Foundation, Inc.,
+! 51 Franklin Street, Fifth Floor, Boston, MA 02110-1301 USA.
+!
+!=====================================================================
+
+!=====================================================================
+!
 !  UTM (Universal Transverse Mercator) projection from the USGS
 !
 !=====================================================================

### 2011-04-08 doc/USER_MANUAL/manual_SPECFEM3D.tex
updates source code headers; updates package version number to version 2.0.1

@@
 \begin{lyxcode}
-{\small tar~-zxvf~SPECFEM3D\_V2.0.0.tar.gz}{\small \par}
+{\small tar~-zxvf~SPECFEM3D\_V2.0.1.tar.gz}{\small \par}
 \end{lyxcode}
@@
 
-$\copyright$ Princeton University and University of Pau / CNRS / INRIA, November 2010
+$\copyright$ Princeton University and University of Pau / CNRS / INRIA, April 2011
 

### 2011-04-09 doc/USER_MANUAL/manual_SPECFEM3D.tex
added a note about removing the Unix stack size limit

@@
 The directory \texttt{\small SPECFEM3D/} will then contain
-the source code. To configure the software for your system, run the
+the source code.\\
+
+We recommend that you add {\texttt{ulimit -S -s unlimited}} to your {\texttt{.bash\_profile}} file and/or {\texttt{limit stacksize unlimited }} to your {\texttt{.cshrc}} file to suppress any potential limit to the size of the Unix stack.\\
+
+Then, to configure the software for your system, run the
 \texttt{configure} shell script. This script will attempt to guess
@@
 
-Note that if you run very large meshes on a relatively small number
+If you run very large meshes on a relatively small number
 of processors, the memory size needed on each processor might become
-greater than 2 gigabytes, which is the upper limit for 32-bit addressing;
-in this case, on some compilers you may need to add \char`\"{}\texttt{-mcmodel=medium}\char`\"{}
-to the compiler options otherwise the compiler will display an error message; on an IBM machine with the \texttt{xlf} and \texttt{xlc} compilers, using \texttt{-q64} is usually sufficient.
+greater than 2 gigabytes, which is the upper limit for 32-bit addressing.
+In this case, on some compilers you may need to add \texttt{``-mcmodel=medium}''
+to the compiler options otherwise the compiler will display an error
+message (for instance \texttt{``relocation truncated to fit: R\_X86\_64\_PC32 against .bss''} or something similar);
+on an IBM machine with the \texttt{xlf} and \texttt{xlc} compilers, using \texttt{-q64} is usually sufficient.
 

### 2011-04-09 doc/USER_MANUAL/manual_SPECFEM3D.tex
rephrased and shortened the paragraph about Discontinuous Galerkin

@@
 
-Note that in most geological models in the context of seismic wave propagation studies
-(except for fault dynamic rupture studies, in which very high frequencies or supershear rupture need to be modeled near the fault, see e.g. \cite{BeGlCrViPi07,BeGlCrVi09,PuAmKa09,TaCrEtViBeSa10})
-a discontinous mesh is not needed because material property contrasts are not drastic and thus a continuous formulation is sufficient;
-conforming mesh doubling bricks can efficiently handle mesh size variations \citep{KoTr02a,KoLiTrSuStSh04,LeChLiKoHuTr08,LeChKoHuTr09,LeKoHuTr09}.
-Also, without significant modifications in the complexity of the
-discontinuous technique \citep{CaKaBr10,SmCoObOvSch10},
-usually those methods need to assume constant material properties inside
-each element, while the SEM can handle gradients inside each element.\\
+Note that in many geological models in the context of seismic wave propagation studies
+(except for instance for fault dynamic rupture studies, in which very high frequencies or supershear rupture need to be modeled near the fault, see e.g.
+\cite{BeGlCrViPi07,BeGlCrVi09,PuAmKa09,TaCrEtViBeSa10})
+a continuous formulation is sufficient because material property contrasts are not drastic and thus
+conforming mesh doubling bricks can efficiently handle mesh size variations \citep{KoTr02a,KoLiTrSuStSh04,LeChLiKoHuTr08,LeChKoHuTr09,LeKoHuTr09}.\\
 

### 2011-04-10 doc/USER_MANUAL/manual_SPECFEM3D.tex
added evolution of the code to the manual

@@
 it under the terms of the GNU General Public License as published
-by the Free Software Foundation (see Appendix \ref{cha:License}).
+by the Free Software Foundation (see Appendix \ref{cha:License}).\\
+
+\textbf{\underline{Evolution of the code:}}\\
+
+ MPI v. 2.0 `Sesame' (Spectral ElementS on Any MEsh), November 2010:
+ Dimitri Komatitsch, Nicolas Le Goff, Roland Martin and Pieyre Le Loher, University of Pau, France,
+ Jeroen Tromp and the Princeton group of developers, Princeton University, USA,
+ and Emanuele Casarotti, INGV Roma, Italy:
+  support for CUBIT meshes decomposed by SCOTCH, METIS or ZOLTAN;
+  much faster solver using Michel Deville's inlined matrix products.\\
+
+ MPI v. 1.4 Dimitri Komatitsch, University of Pau, Qinya Liu and others, Caltech, September 2006:
+  better adjoint and kernel calculations, faster and better I/Os
+  on very large systems, many small improvements and bug fixes.\\
+
+ MPI v. 1.3 Dimitri Komatitsch, University of Pau, and Qinya Liu, Caltech, July 2005:
+  serial version, regular mesh, adjoint and kernel calculations, ParaView support.\\
+
+ MPI v. 1.2 Min Chen and Dimitri Komatitsch, Caltech, July 2004:
+  full anisotropy, volume movie.\\
+
+ MPI v. 1.1 Dimitri Komatitsch, Caltech, October 2002: Zhu's Moho map, scaling
+  of $V_s$ with depth, Hauksson's regional model, attenuation, oceans, movies.\\
+
+ MPI v. 1.0 Dimitri Komatitsch, Caltech, USA, May 2002: first MPI version based on global code.\\
+
+ Dimitri Komatitsch, IPG Paris, France, December 1996: first 3-D solver for the CM-5 Connection Machine,
+    parallelized on 128 processors using Connection Machine Fortran.\\
 

### 2011-04-10 doc/USER_MANUAL/manual_SPECFEM3D.tex
added references to Patera 1984, Seriani et al., Faccioli et al., and also to GeoElse.

@@
 
+The SEM was originally developed in computational fluid dynamics \citep{Pat84,MaPa89}
+and has been successfully adapted to address problems in seismic wave propagation.
+Early seismic wave propagation applications of the SEM, utilizing Legendre basis functions and a
+perfectly diagonal mass matrix, include \cite{CoJoTo93}, \cite{Kom97},
+\cite{FaMaPaQu97}, \cite{CaGa97}, \cite{KoVi98} and \cite{KoTr99},
+whereas applications involving Chebyshev basis functions and a nondiagonal mass matrix
+include \cite{SePr94}, \cite{PrCaSe94} and \cite{SePrPr95}.\\
+
 All SPECFEM3D software is written in Fortran90 with full portability
@@
 as well as Convolutional or Auxiliary Differential Equation Perfectly Matched absorbing Layers (C-PML or ADE-PML) \citep{KoMa07,MaKoEz08,MaKoGe08,MaKo09,MaKoGeBr10}.
-It will also use the PT-SCOTCH parallel library for mesh partitioning.
+It will also use the PT-SCOTCH parallel library for mesh partitioning.\\
+
+SPECFEM3D includes coupled fluid-solid domains and
+adjoint capabilities, which enables one to address seismological inverse problems, but for linear rheologies so far.
+To accommodate visco-plastic or non-linear rheologies, the reader can refer to 
+the \href{http://geoelse.stru.polimi.it/}{GeoELSE} software package \citep{CaGa97,StPaIg09,ChMoTsBaKrKaStKr10}.
 

### 2011-10-26 doc/USER_MANUAL/manual_SPECFEM3D.tex
removed "in part"

@@
 \section*{Authors}
-The SPECFEM3D package was first developed by Dimitri Komatitsch and Jeroen Tromp at Harvard University and Caltech, USA, starting in 1998, based in part on earlier work by Dimitri Komatitsch and Jean-Pierre Vilotte at Institut de Physique du Globe (IPGP) in Paris, France from 1995 to 1997.\\
+The SPECFEM3D package was first developed by Dimitri Komatitsch and Jeroen Tromp at Harvard University and Caltech, USA, starting in 1998, based on earlier work by Dimitri Komatitsch and Jean-Pierre Vilotte at Institut de Physique du Globe (IPGP) in Paris, France from 1995 to 1997.\\
 

### 2011-11-18 doc/USER_MANUAL/manual_SPECFEM3D.tex
manual updates: added Ch 11 on tomographic inversion with kernels; added a couple lines to Table of Contents; added one reference (Maggi2009)

@@
 
+\chapter{\label{cha:tomo}Tomographic inversion using sensitivity kernels}
+
+One of the fundamental reasons for computing sensitivity kernels (Section~\ref{sec:Adjoint-simulation-finite}) is to use them within a tomographic inversion.
+In other words, use recorded seismograms, make measurements with synthetic seismograms, and use the misfit between the two to iteratively improve the model described by (at least) $V_{\rm p}$, $V_{\rm s}$, and $\rho$ as a function of space.
+
+Whatever misfit function you use for the tomographic inversion \citep[several examples in][]{Tromp2005}, you will weight the sensitivity kernels with measurements.
+Furthermore, you will use as many measurements (stations, components, time windows) as possible per event; hence, we call these composite kernels ``event kernels,'' which are volumetric fields representing the gradient of the misfit function with respect to one of the variables (\eg $V_{\rm s}$).
+The basic features of an adjoint-based tomographic inversion were illustrated in \citet{Tape2007} using a conjugate-gradient algorithm; there are dozens of versions of gradient-based inversion algorithms that could alternatively be used.
+The tomographic inversion of \citet{Tape2009,Tape2010} used SPECFEM3D as well as several additional components which are also stored on the CIG svn server, described next.
+
+The directory containing utilities for tomographic inversion using SPECFEM3D (or other packages that evaluate misfit functions and gradients) is here on the CIG svn server:
+%
+\begin{verbatim}
+/cig/seismo/3D/ADJOINT_TOMO/
+   flexwin/     -- FLEXWIN algorithm for automated picking of time windows
+   measure_adj/ -- reads FLEXWIN output file and makes measurements, 
+                   with the option for computing adjoint sources
+   iterate_adj/ -- various tools for iterative inversion
+                   (requires pre-computed "event kernels")
+\end{verbatim}
+%
+This directory also contains a brief \verb+README+ file indicating the role of the three subdirectories, \verb+flexwin+ \citep{Maggi2009}, \verb+measure_adj+, and \verb+iterate_adj+.
+The components for making the model update are there; however, there are no explicit rules for computing the model update, just as with any optimization problem.
+There are options for computing a conjugate gradient step, as well as a source subspace projection step.
+The workflow could use substantial ``stitching together,'' and we welcome suggestions; undoubtedly, the tomographic inversion procedure will improve with time.
+{\bf This is a work in progress.}
+
+The best single file to read is probably: \verb+ADJOINT_TOMO/iterate_adj/cluster/README+.
+
+%============================================================
+
 \chapter*{\label{cha:Bug-Reports-and}Bug Reports and Suggestions for Improvements}
@@
 
+\addcontentsline{toc}{chapter}{Notes \& Acknowledgments}
 \chapter*{Notes \& Acknowledgments}
@@
 
+\addcontentsline{toc}{chapter}{Copyright}
 \chapter*{Copyright}
@@
 
+\addcontentsline{toc}{chapter}{References}
 \bibliography{bibliography}

### 2011-11-18 doc/USER_MANUAL/manual_SPECFEM3D.tex
fixed missing \eg
updated citations in manual (Ch 11)

@@
 \newcommand{\nchunks}{\mbox{\texttt{NCHUNKS\/}}}
+\newcommand{\eg}{e.g.,}
 
@@
 
-Whatever misfit function you use for the tomographic inversion \citep[several examples in][]{Tromp2005}, you will weight the sensitivity kernels with measurements.
+Whatever misfit function you use for the tomographic inversion \citep[several examples in][]{TrTaLi05}, you will weight the sensitivity kernels with measurements.
 Furthermore, you will use as many measurements (stations, components, time windows) as possible per event; hence, we call these composite kernels ``event kernels,'' which are volumetric fields representing the gradient of the misfit function with respect to one of the variables (\eg $V_{\rm s}$).
-The basic features of an adjoint-based tomographic inversion were illustrated in \citet{Tape2007} using a conjugate-gradient algorithm; there are dozens of versions of gradient-based inversion algorithms that could alternatively be used.
-The tomographic inversion of \citet{Tape2009,Tape2010} used SPECFEM3D as well as several additional components which are also stored on the CIG svn server, described next.
+The basic features of an adjoint-based tomographic inversion were illustrated in \citet{TaLiTr07} using a conjugate-gradient algorithm; there are dozens of versions of gradient-based inversion algorithms that could alternatively be used.
+The tomographic inversion of \citet{TaLiMaTr09,TaLiMaTr2010} used SPECFEM3D as well as several additional components which are also stored on the CIG svn server, described next.
 

### 2011-11-18 doc/USER_MANUAL/manual_SPECFEM3D.tex
fixed the references in the new chapter about inversions

@@
 
-Whatever misfit function you use for the tomographic inversion \citep[several examples in][]{TrTaLi05}, you will weight the sensitivity kernels with measurements.
+Whatever misfit function you use for the tomographic inversion (several examples in \cite{TrTaLi05} and \cite{TrKoLi08}), you will weight the sensitivity kernels with measurements.
 Furthermore, you will use as many measurements (stations, components, time windows) as possible per event; hence, we call these composite kernels ``event kernels,'' which are volumetric fields representing the gradient of the misfit function with respect to one of the variables (\eg $V_{\rm s}$).
-The basic features of an adjoint-based tomographic inversion were illustrated in \citet{TaLiTr07} using a conjugate-gradient algorithm; there are dozens of versions of gradient-based inversion algorithms that could alternatively be used.
+The basic features of an adjoint-based tomographic inversion were illustrated in \citet{TaLiTr07} and \citet{TrKoLi08} using a conjugate-gradient algorithm; there are dozens of versions of gradient-based inversion algorithms that could alternatively be used.
 The tomographic inversion of \citet{TaLiMaTr09,TaLiMaTr2010} used SPECFEM3D as well as several additional components which are also stored on the CIG svn server, described next.

### 2011-11-21 doc/USER_MANUAL/manual_SPECFEM3D.tex
improved the paragraph about single precision in the manual

@@
 \end{description}
-The package can run either in single or in double precision. The default
-is single precision mode because this requires exactly half as much
-memory. Select your preference by selecting the appropriate setting
-in the \texttt{constants.h} file:
+The package can run either
+in single or in double precision mode. The default is single precision
+because for almost all calculations performed using the spectral-element method
+using single precision is sufficient and gives the same results (i.e. the same seismograms);
+and the single precision code is faster and requires exactly half as much memory.
+Select your preference by selecting the appropriate setting in the \texttt{constants.h} file:
 
@@
 \end{description}
-On a new system, it is definitely worth experimenting with single
-versus double precision simulations to determine which is faster.
-Note that on many current processors (e.g., Intel, AMD, IBM Power),
-single precision calculations are often significantly faster; the
-difference can typically be 10\% to 25\%. It is therefore often worth
-using single precision if you can. We recommend running the same calculation
-once in single precision and in double precision on your system and
-then comparing the seismograms. If they are identical, you should probably
-select single precision for your future runs.\\
+On many current processors (e.g., Intel, AMD, IBM Power), single precision calculations
+are significantly faster; the difference can typically be 10\%
+to 25\%. It is therefore better to use single precision.
+What you can do once for the physical problem you want to study is run the same calculation in single precision
+and in double precision on your system and compare the seismograms.
+If they are identical (and in most cases they will), you can select single precision for your future runs.\\
 

### 2011-11-29 doc/USER_MANUAL/manual_SPECFEM3D.tex
adds normalization option to create_movie_shakemap_AVS_DX_GMT.f90; updates movie outputs for saving displacements; updates manual in movies section

@@
 output files are saved in the \texttt{in\_out\_files/OUTPUT\_FILES} directory. We
-save the velocity field. The look of a movie is determined by the
+save the displacement field if the parameter \texttt{SAVE\_DISPLACEMENT} is set, otherwise the velocity field is saved. 
+The look of a movie is determined by the
 half-duration of the source. The half-duration should be large enough
@@
 select the appropriate value of 1.1 $\times$ smallest period. As
-usual, for a point source one can set \texttt{HALF\_DURATION} in the
-\texttt{Par\_file} to be 0.0 and \texttt{HDUR\_MOVIE} = 0.0 to get
+usual, for a point source one can set \texttt{half duration} in the
+\texttt{CMTSOLUTION} file to be 0.0 and \texttt{HDUR\_MOVIE} = 0.0 in the \texttt{Par\_file} to get
 the highest frequencies resolved by the simulation, but for a finite
-source one would keep all the \texttt{HALF\_DURATIONs} as prescribed
+source one would keep all the \texttt{half durations} as prescribed
 by the finite source model and set \texttt{HDUR\_MOVIE} = 0.0.

### 2012-01-10 doc/USER_MANUAL/manual_SPECFEM3D.tex
updates example Mount_StHelens; adds mesh output for courant and minimum period; updates manual for xcheck_mesh_quality_CUBIT_Abaqus outputs

@@
 \end{lyxcode}
-to generate an AVS output file (\texttt{\small AVS\_meshquality.inp}
-in AVS UCD format) or OpenDX output file (\texttt{\small DX\_meshquality.}~\\
-\texttt{\small dx}) that can be used to investigate mesh quality,
-e.g, skewness of elements and a Gnuplot histogram (\texttt{\small mesh\_quality\_}~\\
-\texttt{\small histogram.txt}) that can be plotted with gnuplot (type
+to generate an 
+%not supported: AVS output file (\texttt{\small AVS\_meshquality.inp} in AVS UCD format) or 
+OpenDX output file (\texttt{\small DX\_mesh\_quality.dx}) 
+that can be used to investigate mesh quality, e.g. skewness of elements, 
+and a Gnuplot histogram (\texttt{\small mesh\_quality\_histogram.txt}) 
+that can be plotted with gnuplot (type
 `\texttt{\small gnuplot plot\_mesh\_quality\_histogram.gnu}'). The

### 2012-01-10 src/shared/check_mesh_resolution.f90

@@ -26,7 +26,8 @@
 
   subroutine check_mesh_resolution(myrank,NSPEC_AB,NGLOB_AB,ibool,xstore,ystore,zstore, &
                                     kappastore,mustore,rho_vp,rho_vs, &
-                                    DT, model_speed_max,min_resolved_period )
+                                    DT, model_speed_max,min_resolved_period, &
+                                    LOCAL_PATH,SAVE_MESH_FILES )
 
 ! check the mesh, stability and resolved period for acoustic / elastic domains
 !
@@ -42,6 +43,9 @@
   integer, dimension(NGLLX,NGLLY,NGLLZ,NSPEC_AB) :: ibool
   double precision :: DT
   real(kind=CUSTOM_REAL) :: model_speed_max,min_resolved_period
+
+  character(len=256):: LOCAL_PATH
+  logical :: SAVE_MESH_FILES
 
   ! local parameters
   real(kind=CUSTOM_REAL) :: vpmin,vpmax,vsmin,vsmax,vpmin_glob,vpmax_glob,vsmin_glob,vsmax_glob
@@ -68,9 +72,13 @@
   !********************************************************************************
   !real(kind=CUSTOM_REAL),parameter :: NELEM_PER_WAVELENGTH = 1.5
 
-
   logical :: has_vs_zero
   real(kind=CUSTOM_REAL),dimension(1) :: tmp_val
+
+  ! debug: for vtk output
+  real(kind=CUSTOM_REAL),dimension(:),allocatable :: tmp1,tmp2
+  integer:: ier
+  character(len=256) :: filename,prname
 
   ! initializations
   if( DT <= 0.0d0) then
@@ -97,6 +105,15 @@
   dt_suggested_glob = HUGEVAL
 
   has_vs_zero = .false.
+
+  ! debug: for vtk output
+  if( SAVE_MESH_FILES ) then
+    allocate(tmp1(NSPEC_AB),stat=ier)
+    allocate(tmp2(NSPEC_AB),stat=ier)
+    if( ier /= 0 ) stop 'error allocating array tmp'
+    tmp1(:) = 0.0
+    tmp2(:) = 0.0
+  endif
 
   ! checks courant number & minimum resolved period for each grid cell
   do ispec=1,NSPEC_AB
@@ -132,6 +149,9 @@
     if( DT_PRESENT ) then
       cmax = max( vpmax,vsmax ) * DT / distance_min
       cmax_glob = max(cmax_glob,cmax)
+
[…]

### 2012-01-30 doc/USER_MANUAL/manual_SPECFEM3D.tex
updated the manual to describe the options one may need to change in flags.guess

@@
 
+Before running the \texttt{configure} script, you should probably edit file \texttt{flags.guess} to make sure that it contains the best compiler options for your system.
+Known problems or things to check are:
+
+\begin{description}
+\item [{\texttt{Intel ifort compiler}}] See if you need to add \texttt{-assume byterecl} for your machine.
+\item [{\texttt{IBM or Cray compilers}}] See if you need to add \texttt{-qsave} or \texttt{-qnosave} for your machine.
+\item [{\texttt{Mac OS}}] You will probably need to install \texttt{XCODE}.
+In addition, the \texttt{clock\_gettime} routine, which is used by the \texttt{SCOTCH} library that we use, does not exist in Mac OS.
+You will need to replace it with \texttt{clock\_get\_time} if you want to use \texttt{SCOTCH}.
+\end{description}
+
+When compiling on an IBM machine with the \texttt{xlf} and \texttt{xlc} compilers, we suggest running the \texttt{configure} script
+with the following options:\\
+
+\noindent\texttt{./configure FC=xlf90\_r MPIFC=mpif90 CC=xlc\_r CFLAGS="-O3 -q64" FCFLAGS="-O3 -q64" -with-scotch-dir=...}.\\
+
+On SGI systems, \texttt{flags.guess} automatically informs \texttt{configure}
+to insert `\texttt{`TRAP\_FPE=OFF}'' into the generated \texttt{Makefile}
+in order to turn underflow trapping off.\\
+
+If you run very large meshes on a relatively small number
+of processors, the memory size needed on each processor might become
+greater than 2 gigabytes, which is the upper limit for 32-bit addressing.
+In this case, on some compilers you may need to add \texttt{``-mcmodel=medium}''
+to the compiler options otherwise the compiler will display an error
+message (for instance \texttt{``relocation truncated to fit: R\_X86\_64\_PC32 against .bss''} or something similar);
+on an IBM machine with the \texttt{xlf} and \texttt{xlc} compilers, using \texttt{-q64} is usually sufficient.
+
 The SPECFEM3D software package relies on the SCOTCH library to partition meshes created with CUBIT.
@@
 If they are identical (and in most cases they will), you can select single precision for your future runs.\\
-
-When compiling on an IBM machine with the \texttt{xlf} and \texttt{xlc} compilers, we suggest running the \texttt{configure} script
-with the following options:\\
-
-\noindent\texttt{./configure FC=xlf90\_r MPIFC=mpif90 CC=xlc\_r CFLAGS="-O3 -q64" FCFLAGS="-O3 -q64" -with-scotch-dir=...}.\\
-
-On SGI systems, \texttt{flags.guess} automatically informs \texttt{configure}
-to insert `\texttt{`TRAP\_FPE=OFF}'' into the generated \texttt{Makefile}
-in order to turn underflow trapping off.\\
-
-If you run very large meshes on a relatively small number
-of processors, the memory size needed on each processor might become
-greater than 2 gigabytes, which is the upper limit for 32-bit addressing.
-In this case, on some compilers you may need to add \texttt{``-mcmodel=medium}''
-to the compiler options otherwise the compiler will display an error
-message (for instance \texttt{``relocation truncated to fit: R\_X86\_64\_PC32 against .bss''} or something similar);
-on an IBM machine with the \texttt{xlf} and \texttt{xlc} compilers, using \texttt{-q64} is usually sufficient.
 

### 2012-02-02 doc/USER_MANUAL/manual_SPECFEM3D.tex
fixed minor details in the LaTeX manual

@@
 Before running the \texttt{configure} script, you should probably edit file \texttt{flags.guess} to make sure that it contains the best compiler options for your system.
-Known problems or things to check are:
+Known issues or things to check are:
 
@@
 message (for instance \texttt{``relocation truncated to fit: R\_X86\_64\_PC32 against .bss''} or something similar);
-on an IBM machine with the \texttt{xlf} and \texttt{xlc} compilers, using \texttt{-q64} is usually sufficient.
+on an IBM machine with the \texttt{xlf} and \texttt{xlc} compilers, using \texttt{-q64} is usually sufficient.\\
 

### 2012-03-02 doc/USER_MANUAL/manual_SPECFEM3D.tex
removed "Cray"
added more details about "relocation truncated to fit"

@@
 \item [{\texttt{Intel ifort compiler}}] See if you need to add \texttt{-assume byterecl} for your machine.
-\item [{\texttt{IBM or Cray compilers}}] See if you need to add \texttt{-qsave} or \texttt{-qnosave} for your machine.
+\item [{\texttt{IBM compiler}}] See if you need to add \texttt{-qsave} or \texttt{-qnosave} for your machine.
 \item [{\texttt{Mac OS}}] You will probably need to install \texttt{XCODE}.
@@
 In this case, on some compilers you may need to add \texttt{``-mcmodel=medium}''
-to the compiler options otherwise the compiler will display an error
+to the compiler options of CFLAGS, FCFLAGS and LDFLAGS otherwise the compiler will display an error
 message (for instance \texttt{``relocation truncated to fit: R\_X86\_64\_PC32 against .bss''} or something similar);

### 2012-05-12 doc/USER_MANUAL/manual_SPECFEM3D.tex
fixed NREGIONS

@@
 \end{itemize}
-\item [{\texttt{NMATERIALS}}] The number of regions in the mesh.
+\item [{\texttt{NREGIONS}}] The number of regions in the mesh.
 In the following lines, because the mesh is regular or 'almost regular', each region is defined as :
 \begin{lyxcode}
-  \texttt{XI\_begin}  \texttt{XI\_end}  \texttt{ETA\_begin}  \texttt{ETA\_end}  \texttt{material\_ID}
+  \texttt{NEX\_XI\_BEGIN}  \texttt{NEX\_XI\_END}  \texttt{NEX\_ETA\_BEGIN}  \texttt{NEX\_ETA\_END}  \texttt{NZ\_BEGIN}  \texttt{NZ\_END}  \texttt{material\_ID}
 \end{lyxcode}

### 2012-05-18 doc/GJI_2011_paper_about_the_code/paper_sesame_2.0.tex
added the coffee mug mesh from our 2011 GJI paper

%% ------------------------------------------------------------------------ %%
% GJI ARTICLE MODE
%% ------------------------------------------------------------------------ %%
\documentclass[referee,extra]{gji}
%\documentclass[extra,mreferee]{gji}
%%%%% older
%\documentclass[11pt,extra]{gji}

%\documentclass[referee,extra]{gji} % galley/ referee
\usepackage{appendix}

%%%%\usepackage{timet}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage[dvipsnames,usenames]{color}
\usepackage[colorlinks=true,citecolor=black,linkcolor=black,urlcolor=blue]{hyperref}
\usepackage{amssymb} % for mathfrak font

%% useful commands
\newcommand{\bequ}{\begin{equation} }
\newcommand{\eequ}{\end{equation} }
\newcommand{\barr}{\begin{eqnarray} }
\newcommand{\earr}{\end{eqnarray} }

\newcommand{\bsplitequ}{\begin{split} }
\newcommand{\esplitequ}{\end{split} }

\newcommand{\oneoverrho}{\frac{1}{\rho} \,  }
\newcommand{\oneoverkappa}{\frac{1}{\kappa} \, }
\newcommand{\intOmega}{\int_\Omega \, }
\newcommand{\intSurface}{\int_{\partial \Omega}}

\newcommand{\bnabla}{ \, {\bf \nabla } }
\newcommand{\bnablaparallel}{ \bnabla^{\|} }

\newcommand{\bs}{ {\bf s } }
\newcommand{\bc}{ {\bf c } }
\newcommand{\bT}{ {\bf T } }
\newcommand{\bforce}{ {\bf f } }
\newcommand{\bM}{ {\bf M } }
\newcommand{\bI}{ {\bf I } }
\newcommand{\bm}{ {\bf m } }
\newcommand{\bx}{ {\bf x } }
\newcommand{\bw}{ {\bf w } }
\newcommand{\bu}{ {\bf u } }
\newcommand{\bv}{ {\bf v } }
\newcommand{\bxi}{ {\bf \xi} }
\newcommand{\bnormal}{ \hat{ \bf{n} } \, }
\newcommand{\bD}{ {\bf D } }

\newcommand{\testw}{ {\rm w } \, }
\newcommand{\tilden}{ \tilde{n} }
\newcommand{\dndntilde}{\frac{\partial_n}{\partial_{\tilden}} }
\newcommand{\abg}{ {\alpha \beta \gamma} }
\newcommand{\aprime}{ {\alpha^{\prime}} }
\newcommand{\bprime}{ {\beta^{\prime}} }
\newcommand{\gprime}{ {\gamma^{\prime}} }

\newcommand{\phionedot }{\dot{ \widetilde{\phi_1} } }
\newcommand{\phioneddot }{\ddot{ \widetilde{\phi_1} } }
\newcommand{\phitwodot }{\dot{ \widetilde{\phi_2} } }
\newcommand{\phitwoddot }{\ddot{ \widetilde{\phi_2} } }
\newcommand{\phithreedot }{\dot{ \widetilde{\phi_3} } }
\newcommand{\phithreeddot }{\ddot{ \widetilde{\phi_3} } }
\newcommand{\phifourdot }{\dot{ \phi_4} }
\newcommand{\phifourddot }{\ddot{ \phi_4 } }

\newcommand{\etal}{ {\it et al. } }
%% biblio GJI
\bibliographystyle{gji}
\renewcommand{\cite}[1]{\citet{#1}}
%% colors to show the corrections
\newcommand{\red}[1]{\textbf{\textcolor{Red}{#1}}}
\newcommand{\blue}[1]{\textbf{\textcolor{Blue}{#1}}}
\newcommand{\cyan}[1]{\textbf{\textcolor{Cyan}{#1}}}
\newcommand{\green}[1]{\textbf{\textcolor{Green}{#1}}}
\newcommand{\magenta}[1]{\textbf{\textcolor{Magenta}{#1}}}
\newcommand{\orange}[1]{\textbf{\textcolor{Orange}{#1}}}
%\newcommand{\red}[1]{#1}
%\newcommand{\blue}[1]{#1}
%\newcommand{\cyan}[1]{#1}

%%%%%%%%%% DK DK
% comments between authors
\newcommand{\todaniel}[1]{\textbf{\red{*** To Daniel, from Dimitri: #1 ***}}}
%\newcommand{\todimi}[1]{\textbf{\magenta{*** To Dimitri, from myself: #1 ***}}}
\newcommand{\fromdimi}[1]{\textbf{\magenta{*** From Dimitri: #1 ***}}}
%%%%%%%%%% RM RM
% comments between authors
\newcommand{\rolandtodaniel}[1]{\textbf{\blue{*** To Daniel, from Roland: #1 ***}}}

\begin{document}

% Uncomment the following code (as well as \usepackage{graphicx} above)
% if you need to include images in draft mode
%%%%%\setkeys{Gin}{draft=false}

%% ------------------------------------------------------------------------ %%
%  AUTHORS
%% ------------------------------------------------------------------------ %%
% Author address will appear at end of article, may repeat
% this command for each author.

\author[Peter et al.]{
\parbox[t]{1.\textwidth}{
Daniel Peter$^1$,
Dimitri Komatitsch$^{2,3}$,
Yang Luo$^1$,
Roland Martin$^2$,
Nicolas Le Goff$^2$,
Emanuele Casarotti$^4$,
Pieyre Le Loher$^2$,
Federica Magnoni$^4$,
Qinya Liu$^5$,
C\'eline Blitz$^2$,
Tarje Nissen-Meyer$^6$,
Piero Basini$^{6}$
and Jeroen Tromp$^{1,7}$
\vspace{1cm}
\newline
{\small
$^1$ Princeton University, Department of Geosciences, 318 Guyot Hall, Princeton, NJ 08544, USA \newline
$^2$ Universit\'e de Pau et des Pays de l'Adour, CNRS \& INRIA Magique-3D,
Laboratoire de Mod\'elisation et d'Imagerie en G\'eosciences UMR 5212,
Avenue de l'Universit\'e, 64013 Pau Cedex, France \newline
$^3$ Institut universitaire de France, 103 boulevard Saint-Michel, 75005 Paris, France \newline
$^4$ Istituto Nazionale di Geofisica e Vulcanologia, Via di Vigna Murata 605, 00143, Rome, Italy \newline
$^5$ Department of Physics, University of Toronto, Ontario, Canada \newline
$^6$ Institute of Geophysics, ETH Zurich, Sonneggstr. 5, CH-8092 Zurich, Switzerland \newline
$^7$ Princeton University, Program in Applied \& Computational Mathematics, Princeton, NJ 08544, USA \newline
{\it Email: dpeter@princeton.edu}
}
}
}

%% ------------------------------------------------------------------------ %%
%  TITLE
%% ------------------------------------------------------------------------ %%
\title[SPECFEM3D Version~2.0 `Sesame']{Forward and adjoint simulations of seismic wave propagation on fully unstructured hexahedral meshes}

\date{January 2011}

\maketitle

\keywords{Computational seismology, Tomography, Interferometry, Wave propagation}

\begin{summary}
We present forward and adjoint
spectral-element simulations of coupled acoustic and (an)elastic seismic
wave propagation on fully unstructured hexahedral meshes.
Simulations benefit from recent advances in hexahedral meshing, load balancing
and software optimization.
Meshing may be accomplished using a mesh generation tool kit such as \href{http://cubit.sandia.gov}{CUBIT},
and load balancing is facilitated by graph partitioning based on the \href{http://www.labri.fr/perso/pelegrin/scotch/}{SCOTCH} library.
Coupling between fluid and solid regions is incorporated in straightforward fashion using domain decomposition.
Topography, bathymetry and Moho undulations may be readily included in the mesh,
and physical dispersion and attenuation associated with anelasticity are accounted for using a series
of standard linear solids.
Finite-frequency Fr\'echet derivatives are calculated using adjoint methods in both fluid and solid domains.
The software is benchmarked for a layercake model.
We present various examples of fully unstructured meshes, snapshots of wavefields and finite-frequency kernels
generated by Version~2.0 `Sesame' of our widely used open source spectral-element package \href{http://www.geodynamics.org/cig/software/specfem3d}{SPECFEM3D}.
 \end{summary}

%% ------------------------------------------------------------------------ %%
\section{Introduction}\label{sec:Introduction}
%% ------------------------------------------------------------------------ %%
We present a new software package, \href{http://www.geodynamics.org/cig/software/specfem3d}{SPECFEM3D} Version~2.0 `Sesame',
capable of simulating forward and adjoint seismic wave propagation on fully unstructured hexahedral meshes of
arbitrary shaped model domains.
In view of unrelenting  growth in computational power,
it has become more-and-more important to develop software capable of
harnessing powerful computers to address a broad range of seismological forward and inverse problems.
A well-established numerical technique for solving such problems in a fast and highly accurate manner
is the spectral-element method (SEM).
The SEM was originally developed in computational fluid dynamics \citep{Pat84,MaPa89}
and has been successfully adapted to address problems in seismic wave propagation.
Early seismic wave propagation applications of the SEM, utilizing Legendre basis functions and a
perfectly diagonal mass matrix, include \cite{CoJoTo93}, \cite{Kom97},
\cite{FaMaPaQu97}, \cite{CaGa97}, \cite{KoVi98} and \cite{KoTr99},
whereas applications involving Chebyshev basis functions and a nondiagonal mass matrix
include \cite{SePr94}, \cite{PrCaSe94} and \cite{SePrPr95}.

The SEM is a continuous Galerkin technique, which may be made discontinous \citep{BeMaPa94,Ch00,KoWoHu02,ChCaVi03,LaWaBe05,Kop06,WiStBuGh10,AcKo11};
it is then close to a particular case of the discontinuous Galerkin technique \citep{ReHi73,Arn82,FaRi99,HuHuRa99,CoKaSh00,GiHeWa02,RiWh03,MoRi05,GrScSc06,AiMoMu06,BeLaPi06,DuKa06,DeSeWh08,PuAmKa09,WiStBuGh10,DeSe10,EtChViGl10},
with optimized efficiency because of its tensorized basis functions \citep{WiStBuGh10,AcKo11}.

An important feature of the SEM is that it can accurately handle very distorted mesh elements \citep{OlSe11},
and thus conforming non-structured mesh doubling bricks can efficiently accommodate mesh size variations
\citep{KoTr02a,KoLiTrSuStSh04,LeChLiKoHuTr08,LeChKoHuTr09,LeKoHuTr09}.
The method has very good accuracy and convergence properties, such as a spectral rate of convergence
 \citep{CaHuQuZa88,MaPa89,SePr94,DeFiMu02,Coh02,DeSe07,SeOl08}.
In this sense the SEM is close to the family of pseudo-spectral methods
\citep[see e.g.,][]{CaHuQuZa88,CaKoKo88,CaKoBeSe92,CaWa93,KoCoMo96},
but combined with the flexibility of finite elements, in particular in terms of mesh design.
For reviews of the SEM in seismology, see e.g., \citet{KoTsTr05}, \cite{ChKoViCaVaFe07}, \cite{TrKoLi08} and \cite{Fi11}.

The SEM is well suited to parallel implementations on very large supercomputers \citep{KoTr02a,KoTsChTr03,TsKoChTr03,KoLaMi08a,CaKoLaTiMiLeSnTr08,KoViCh10}
as well as on clusters of GPU accelerating graphics cards \citep{KoMiEr09,KoErGoMi10,Kom11}.
Tensor products inside each element may be optimized to reach very high efficiency \citep{DeFiMu02},
and mesh point and element numbering may be optimized to reduce processor cache misses
and improve cache reuse \citep{KoLaMi08a}.
The SEM can handle triangular (in 2D) or tetrahedral (in 3D) elements \citep{WinBoyd96,TaWi00,KoMaTrTaWi01,Coh02,MeViSa06},
as well as mixed meshes, although with increased cost and reduced accuracy in these non-tensorized elements,
as in the discontinuous Galerkin method.

In many cases of practical seismological interest,
using a conforming mesh and a continuous formulation is sufficient,
because in most geological models material property contrasts are not too dramatic.
When this ceases to be true, requiring a discontinuous formulation, one can either turn
to a discontinuous version of the SEM \citep{BeMaPa94,Ch00,KoWoHu02,ChCaVi03,LaWaBe05,Kop06,WiStBuGh10,AcKo11}
or to a discontinuous Galerkin technique.
A discontinuous formulation is particularly suitable for dynamic rupture simulations,
because high frequencies or supershear rupture need to be accommodated near the fault,
where a significantly denser mesh and a more sophisticated (upwind) time scheme are required,
thereby suppressing the amplification of unstable modes
\citep[see e.g.,][]{BeGlCrViPi07,PuAmKa09,BeGlCrVi09,TaCrEtViBeSa10}.
Another example that may require a discontinuous formulation involves the resolution of a shallow geotechnical layer,
in which seismic shear wavespeeds may be reduced by an order of magnitude.

For seismological applications, the SEM has been successfully implemented for three-dimensional
global- and regional-scale simulations
\citep{KoVi98,PaFaMa99,Ch00,KoTr02a,KoTr02b,CaChViMo03,ChVa04,FiIgBuKe09},
as well as local-scale simulations in complex and/or densely populated regions, for example in southern California, USA \citep{KoLiTrSuStSh04,TaLiMaTr09,TaLiMaTr2010},
Taipei, Taiwan \citep{LeChLiKoHuTr08,LeChKoHuTr09,LeKoHuTr09},
Caracas, Venezuela \citep{DeCuFeVi06}
and Grenoble, France \citep{CCGCK05,StPaIg09,ChMoTsBaKrKaStKr10}.
The SEM may also be used to study elastic wave propagation on smaller scales, for instance the propagation of ultrasonic waves in crystals \citep{WiKoScTr04}.

Two complementary SEM software packages ---namely,
\href{http://www.geodynamics.org/cig/software/specfem3d-globe}{SPECFEM3D\_GLOBE} for global and regional simulations, and \href{http://www.geodynamics.org/cig/software/specfem3d}{SPECFEM3D} for local simulations---
are feature-rich, well benchmarked and documented implementations.
Data parallelism in the SEM is efficiently exploited using the
Message-Passing Interface (MPI) standard, crucial for modern high-performance computing.
These open source packages are freely available via the Computational Infrastructure for Geodynamics
(\href{http://www.geodynamics.org}{CIG}) and widely used by the seismological community.

To extend the range of local-scale applications, easing the task of mesh generation is paramount.
The two community software packages separate a simulation into two distinct steps: first, creation of a hexahedral mesh, and
second, solution of the seismic wave equation.
This separation avoids the overhead of remeshing when running multiple simulations for the same region,
e.g., repeated simulations at the same resolution.
Focussing on local-scale simulations, previous versions of \href{http://www.geodynamics.org/cig/software/specfem3d}{SPECFEM3D} used an internal mesher which was explicitly tied to the specific purposes of the package:
all geological models were based on a layercake model.
Consequently, the solver was restricted by its internal mesher.
It was impossible to run spectral-element simulations on more complex 3D models without significant recoding, nor was it possible to run such simulations in regions of interest for on- and off-shore exploration seismology,
because acoustic wave propagation in fluids was not supported by the package.

The purpose of this article is to present forward and adjoint simulations in various 3D models using the new software package, \href{http://www.geodynamics.org/cig/software/specfem3d}{SPECFEM3D} Version~2.0 `Sesame', thereby illustrating its current capabilities.
The original \href{http://www.geodynamics.org/cig/software/specfem3d}{SPECFEM3D} package for local simulations was extended, improved and optimized in various ways.
The Version~2.0 `Sesame' release includes a more flexible internal mesher and accommodates more powerful external meshers,
such as \href{http://cubit.sandia.gov}{CUBIT}
\citep{Bl94,WhMiBe95,Mi96,CaStLeKoPiTr08}.
Adding such external meshers into the workflow greatly increases flexibility for high-performance applications,
as illustrated by the \href{http://geoelse.stru.polimi.it/}{GeoELSE} software package \citep{CaGa97,StPaIg09,ChMoTsBaKrKaStKr10}.
Advantages of \href{http://geoelse.stru.polimi.it/}{GeoELSE} include the accommodation of visco-plastic and non-linear rheologies,
whereas benefits of \href{http://www.geodynamics.org/cig/software/specfem3d}{SPECFEM3D} include coupled fluid-solid domains and
adjoint capabilities; the latter enable one to address seismological inverse problems.
Load balancing parallel simulations in \href{http://www.geodynamics.org/cig/software/specfem3d}{SPECFEM3D} is accomplished based on the graph partitioning
software package \href{http://www.labri.fr/perso/pelegrin/scotch/}{SCOTCH}
\citep{PeRo96,ChPe08}.
The new package facilitates coupled forward and adjoint acoustic/(an)elastic simulations,
which are especially interesting for problems in exploration seismology, ocean acoustics and medical tomography.
The new software is freely available under the GNU GPL Version~2 license via \href{http://www.geodynamics.org}{CIG}.

%% ------------------------------------------------------------------------ %%
\section{Governing equations}\label{sec:Method}
%% ------------------------------------------------------------------------ %%

Let us briefly summarize the equations governing seismic wave propagation implemented in \href{http://www.geodynamics.org/cig/software/specfem3d}{SPECFEM3D}.
For more technical details, the reader is referred to \citet{KoTr99}.
\href{http://www.geodynamics.org/cig/software/specfem3d}{SPECFEM3D} Version~2.0 `Sesame' implements wave propagation in coupled (an)elastic and acoustic materials
on local scales.
We may thus safely neglect additional effects that would arise from self-gravitation and rotation
\citep{KoTr02b,KoTsTr05,ChKoViCaVaFe07}, which are important at longer periods.
In the following, we first discuss (an)elastic wave propagation and subsequently consider acoustic waves.

\subsection{Elastic domain}\label{subsec:elastic}
%% ------------------------------------------------------------------------ %%

For elastic materials, the displacement wavefield $\bs(\bx, t)$ is governed by
\bequ \label{equ:motions}
\rho\, \partial_t^2 \bs = \bnabla \cdot \bT + \bforce \,,
\eequ
where $\rho$ denotes mass density, $\bT$ the stress tensor and $\bforce$ the seismic source.
On free surfaces, the traction vector must vanish, i.e.,
\bequ
\bnormal \cdot \bT = {\bf 0}\,,
\eequ
where $\bnormal$ denotes the unit outward normal on the surface.
On boundaries between different elastic materials,
both traction $\bnormal \cdot \bT$ and displacement $\bs$ need to be continuous.
On boundaries between elastic and acoustic domains,
traction $\bnormal \cdot \bT$ and the normal component of displacement $\bnormal \cdot \bs$ need to be continuous.
The initial conditions are
\bequ
\bs(\bx,0) = {\bf 0}, \hspace{10mm} \partial_t \bs(\bx,0) = {\bf 0}\,.
\eequ
We thus initiate the simulation in a medium at rest.
To accommodate simulations under pre-stressed conditions,
these initial conditions may be modified in an appropriate manner.

For elastic materials, the force $\bforce$ in eq. (\ref{equ:motions}) represents the earthquake,
which for a simple point source may be written as
\bequ
\bforce = \mbox{}- \bM\cdot \bnabla \delta(\bx - \bx_s ) \, S(t) \,,
\eequ
where $\bM$ denotes the moment tensor, $\bx_s$ the source location,
$\delta(\bx - \bx_s)$ the Dirac delta distribution located at $\bx_s$ and $S(t)$ the source-time function.
The software also accommodates kinematic rupture simulations,
which may be captured by prescribing a moment-density tensor field.

The stress tensor $\bT$ is linearly related to the strain via the constitutive relationship
\bequ
\bT = \bc : \bnabla \bs\, ,
\eequ
where $\bc$ denotes the stiffness tensor that describes the elastic properties of the medium.
The implementation is general and can handle a fully anisotropic tensor with 21 independent parameters
\citep{ChTr07,SiLiTrTr07a,SiLiTrTr07b}.
Using a linear constitutive relationship is valid under the assumption that perturbations to the reference state
are small.
Note that nonlinear effects are sometimes observed, e.g., nonlinear soil amplification,
and nonlinear constitutive relationships become important for studying such effects,
e.g., for risk mitigation \citep{XuBiGhWa03,DuDeFoKoRo10}.

In an anelastic medium,
we approximate an absorption-band solid using a series of $L$ standard linear solids
\citep{LiAnKa76},
and model the time evolution of the isotropic shear modulus $\mu$ by
\bequ
\mu (t) = \mu_R \left[ 1 - \sum_{l=1}^L \left( 1 - \frac{\tau_l^\epsilon}{\tau_l^\sigma}  \right) e^{-t/\tau_l^\sigma} \right] H(t) \, ,
\eequ
where $\mu_R$ denotes the relaxed modulus, $H(t)$ the Heaviside function and $\tau_l^\sigma$ \& $\tau_l^\epsilon$
 the stress and strain relaxation times of the $l$th standard linear solid.
Experience shows that three solids generally suffice for simulating an absorption band
\citep{EmKo87}.
For further details, see \cite{CaKoKo88b}, \cite{Rob96}, \cite{DaBr01}, \cite{MoKr05}, \cite{KoTsTr05}, \cite{Car07} and \cite{SaKoTr10}.
Simulations of seismic wave propagation in laboratory-scale rock samples or in the context of medical tomography
involve very high frequencies (in the kHz or even MHz range), and strong attenuation must be taken into account.

The SEM solves the equations of motion in the weak form,
which is obtained by dotting the momentum equation (\ref{equ:motions}) with an arbitrary test vector $\bw$ and integrating by parts over the model volume $\Omega$.
We focus on elastic domains and consider coupling interfaces with acoustic domains. Thus, we obtain
\bequ
\intOmega \rho \,\bw \cdot \partial_t^2 \bs \, \mathrm{d}^3\bx =
\intSurface  \bnormal  \cdot \bT\cdot \bw \,\mathrm{d}^2\bx
\mbox{}- \intOmega \bnabla \bw : \bT \,\mathrm{d}^3\bx + \bM : \bnabla \bw( \bx_s ) \,S(t) \, .
\eequ
Note that in this formulation the traction-free surface condition is implicitly accounted for by setting the
contribution from the free surface to zero.

When and where necessary, we use Clayton-Engquist-Stacey absorbing conditions \citep{ClEn77,Sta88,QuTaZa98} to absorb outgoing waves on  fictitious boundaries of the mesh, thereby representing a semi-infinite domain.
It would be more efficient to use a Perfectly Matched Layer (PML) \citep[see e.g.,][]{KoMa07,MaKoGe08,MaKo09},
but a parallel implementation with good load-balancing properties is challenging because additional equations need to be solved.
This issue becomes important when high-order time marching is required to reduce numerical dispersion in difficult case studies that involve complex media with poroelastic or viscoelastic rheologies \citep{MaKoEz08,MaKoGeBr10} or Newtonian compressible fluids \citep{MaCo2010}. Consequently, additional computations need to be performed in PML layers, in particular in corners,
where contributions along several directions are summed \citep{KoMa07}.

At a solid-fluid boundary, the interface
integral over the coupling surface $\partial \Omega$ is used to exchange pressure from the fluid $p_\mathrm{fluid}$ to the solid:
$ \bnormal \cdot \bT \, = \mbox{}-p_\mathrm{fluid}\,\bnormal $.

\subsection{Acoustic domain}\label{subsec:acoustic}
%% ------------------------------------------------------------------------ %%

We define a scalar potential $\phi$ such that the displacement $\bs$ may be written as
\bequ \label{equ:acousticdisplacement}
\bs  =   \rho^{-1} \bnabla \phi \, .
\eequ
The equation of motion in terms of the potential $\phi$ becomes
\bequ \label{equ:acoustic}
\kappa^{-1}\, \partial_t^2 \phi = \bnabla \cdot ( \rho^{-1}\, \bnabla \phi ) + f \,,
\eequ
where $\kappa$ denotes the bulk modulus.
It follows that velocity $\bv$ and pressure $p$ may be expressed as:
\bequ
 \bv   =   \rho^{-1}\,\bnabla \partial_t \phi\, ,
 \eequ
\bequ
  p  =   \mbox{} - \kappa\, (\bnabla \cdot \bs ) =  \mbox{}- \partial_t^2 \phi\, .	\label{equ:pressure}
\eequ
The resulting formulation for pressure $p$ is the reason why we choose to define the potential $\phi$ as in equation (\ref{equ:acousticdisplacement}).
Since pressure is continuous across first-order discontinuities,
it follows that $\partial_t^2 \phi$ and thus $\phi$ must be continuous, a requirement which is
honored automatically by the basis functions of the SEM.
The source $f$ may be expressed in terms of pressure $P$ acting at location $\bx_s$:
\bequ \label{equ:pressuresource}
f = \mbox{}-\kappa^{-1}\, P(t)\, \delta(\bx - \bx_s )\, .
\eequ
Note that the source is multiplied by a factor $\kappa^{-1}$ due to the formulation used in eq.~(\ref{equ:acoustic}).

Using Gauss' theorem and a scalar test function $\testw$\,, the weak form becomes
\bequ \label{equ:acousticweak}
\intOmega \kappa^{-1}\,\testw\, \partial_t^2\phi\, \mathrm{d}^3\bx =
 \intSurface \rho^{-1}\,\testw \hat{\bf n} \cdot \bnabla \phi \, \mathrm{d}^2\bx
- \intOmega  \rho^{-1} \,\bnabla \testw\cdot\bnabla \phi \, \mathrm{d}^3\bx
-  \kappa^{-1}  \, P(t)\, \testw (\bx_s) \,.
\eequ
At the free surface $\partial \Omega$ we set the pressure $p = \mbox{}- \partial_t^2 \phi = 0$,
thereby enforcing $\phi = 0$, $\partial_t \phi = 0$ and $\partial_t^2 \phi = 0$, i.e.,
we implement a Dirichlet boundary condition along the surface.
At a fluid-solid boundary,
the interface coupling integral may be used to exchange the normal component of displacement between fluid and solid:
$\rho^{-1}\,\bnormal\cdot\bnabla \phi = \bnormal\cdot\bs_\mathrm{solid} $\,.

%% ------------------------------------------------------------------------ %%
\section{Meshing, mesh partitioning and load balancing}\label{sec:meshing}
%% ------------------------------------------------------------------------ %%

The first step in a SEM consists of constructing a high-quality mesh for the region of interest.
In this section,
we outline the key issues based on various 3D examples.
Fig.~\ref{figure:processing} draws the schematic workflow from meshing and
partitioning to finally running spectral-element simulations.
We discuss each phase separately,
focussing on the use of an external mesher, in our case \href{http://cubit.sandia.gov}{CUBIT}
\citep{Bl94}.

\subsection{Hexahedral meshing}\label{subsec:discretizing}
%% ------------------------------------------------------------------------ %%

We subdivide the model volume $\Omega$ into a set of non-overlapping, hexahedral elements.
We impose that the discretization creates a conforming mesh,
i.e., elements match on a full face or edge, and the mesh cannot be discontinuous.
Using the SEM with hexahedral elements leads to computational benefits
over tetrahedral finite elements
\citep{KoMaTrTaWi01,MeViSa06,VoShKi10}.
Especially for parallel implementations,
taking advantage of the diagonal mass matrix and optimized tensor products
is critical in terms of computational speed
\citep{KoTsChTr03,CaKoLaTiMiLeSnTr08,VoShKi10}.
Hexahedral meshing is also attractive for the SEM because it benefits from
reduced errors and generally smaller element counts compared to tetrahedral meshing
\citep{Hest99,KoMaTrTaWi01,VoShKi10}.

Unfortunately,
automatic 3D hexahedral mesh generation is more demanding than unstructured tetrahedral meshing
\citep{ShJo06,StKeOwBlStSh10}.
In order to construct hexahedral meshes, our examples make use of an external hexahedral mesher,
such as \href{http://cubit.sandia.gov}{CUBIT}
\citep{Bl94}.
We focus on this particular mesh generation tool kit because it is a well documented and feature-rich package,
on which most of our own experience is based.
One may readily use other meshing tools,
such as  \href{http://www.simulia.com/products/abaqus_fea.html}{Abaqus} \citep{ABAQUS08},
\href{http://www.ansys.com/}{ANSYS} \citep{ANSYS11},
\href{http://www.gocad.org/www/}{GOCAD} \citep{Ma92,CaCoCaSaVi09},
\href{http://gid.cimne.upc.es}{GiD} \citep{GaGaGo10,Gid11},
\href{http://geuz.org/gmsh}{Gmsh} \citep{GeRe09},
\href{http://www.truegrid.com/}{TrueGrid} \citep{TrueGrid01,NoNu04}
or  \href{http://www.salome-platform.org/}{Salome} \citep{RiCa07,BeLeRo10}.

Fig.~\ref{figure:mountsthelens} shows several examples of fully unstructured hexahedral meshes.
In the Mount St.~Helens region, the mesh employs a mesh tripling layer to increase resolution at the topographic surface.
Tripling is the default refinement in \href{http://cubit.sandia.gov}{CUBIT} for subdividing hexahedral elements in a conforming fashion.
Surface topography is imported using Shuttle Radar Topographic Mission (\href{http://srtm.csi.cgiar.org/}{SRTM})
data, converted to Universal Transverse Mercator (UTM) coordinates with an original resolution of 90~m \citep{JaReNeGue2008}.
Meshing is performed automatically by \href{http://cubit.sandia.gov}{CUBIT} using a sweep algorithm.
The resolution of the mesh enables seismic wave simulations with frequencies up to $\sim$1.5~Hz.
The Mesh for the L'Aquila region, Italy, consists of $\sim$7~M hexahedra with an element size of $\sim$90~m at the top surface.
This mesh facilitates simulations of seismic wave propagation up to $\sim$5~Hz.
For the exploration geophysics model,
the hexahedral mesh honors a salt dome body inside a 3D model capped by a water layer.
The Mesh for asteroid 433-Eros with a close-bound surface has a resolution of roughly 300~m.
Finally, the filled coffee cup model discretized into hexahedra
couples an elastic domain for the cup with an acoustic domain for the coffee inside the cup.

In order to ensure compatibility with previous versions of \href{http://www.geodynamics.org/cig/software/specfem3d}{SPECFEM3D}
\citep[see e.g.,][]{KoLiTrSuStSh04,LiPoKoTr04},
the in-house mesher based on analytical linear interpolation from the top to the bottom of the mesh has been
adapted to the new code structure.
It facilitates the design of simpler, alternative meshes for layercake models.

\subsection{Partitioning and load balancing}\label{subsec:partitioning}
%% ------------------------------------------------------------------------ %%

Balancing the computational load and distributing the mesh on a large number of cores is crucial for
optimized high-performance simulations \citep{MaKoBlLe08}.
In order to do so, we make use of an external partitioner, namely
\href{http://www.labri.fr/perso/pelegrin/scotch/}{SCOTCH} \citep{PeRo96,ChPe08}, which we use to
balance spectral-element computations on an arbitrary number of cores.
An alternative partitioner able to fulfill these tasks is
\href{http://glaros.dtc.umn.edu/gkhome/views/metis}{METIS} \citep{KaKu98c}, but \href{http://www.labri.fr/perso/pelegrin/scotch/}{SCOTCH} is more actively maintained
\citep{ChPe08} and performs better in many cases that we have tested.

Especially for simulations involving coupled elastic and acoustic domains,
balancing the mesh becomes paramount.
Most of the computation time is spent resolving the divergence of the stress tensor in each element.
The computational cost for an elastic element is approximately four times larger than for an acoustic element,
which may be established by running simulations for  one domain at a time.
During partitioning, we therefore weight each element according to its associated domain type and computational cost
to balance the overall numerical cost rather than simply the number of elements between partitions.
The major improvement in \href{http://www.geodynamics.org/cig/software/specfem3d}{SPECFEM3D} code performance focuses on these tensor products, using highly efficient algorithms developed by \citet{DeFiMu02} and optimizing cache usage.
Another key aspect of mesh partitioning is minimization of the number of edge cuts,
because this reduces the amount of MPI communications
between processor cores (an edge cut occurs when two contiguous elements are assigned to distinct cores).
On machines comprising a very large number of cores,
it is crucial to resort to non-blocking communications between compute nodes,
for instance using non-blocking MPI message passing, in order to obtain good performance scaling
\citep{DaNa98,MaKoBlLe08,KoLaMi08a}.

Fig.~\ref{figure:partitions}
presents a simple example of partitioning and load balancing the mesh
around Mount St.~Helens, as shown in Fig.~\ref{figure:mountsthelens}.
For illustrative purposes,
we decompose the mesh onto four cores using the \href{http://www.labri.fr/perso/pelegrin/scotch/}{SCOTCH} library.
The total number of spectral-elements is $\sim$24,000,
such that each partition contains $\sim$6,000 elements after decomposition.
Partitioning and load balancing equally distributes the elements over the different cores,
since the whole domain is purely elastic.
A partitioner such as \href{http://www.labri.fr/perso/pelegrin/scotch/}{SCOTCH}
can also load balance computationally more complex meshes,
for example containing PML elements
along absorbing boundaries of the model; this is the subject of future research.

In a final, separate step we generate mesh databases for each partition needed for the spectral-element solver.
These databases contain Gauss-Lobatto-Legendre (GLL) points for all spectral elements.
Material properties are assigned to these GLL points,
and thus sampling resolution of a geological model not only depends on element size but also on polynomial degree.
Furthermore, the generation of mesh databases automatically detects interfaces between elastic and acoustic domains,
needed for coupling seismic waves from one domain to another.
Load-balancing of the simulation persists, because we keep the polynomial degree fixed for all spectral elements.
Note that this final step of generating mesh databases provides additional freedom in assigning and changing wavespeeds,
which is important for seismic inversion procedures.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Overlapping computation and communication}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

The elements that compose the mesh slices shown in Figures~\ref{figure:mountsthelens} and~\ref{figure:partitions} are in contact
through a common face, edge or point.
To allow for overlap of communication between compute nodes with computations within each mesh slice
---thereby speeding up the simulation---
a list of all elements in contact with any other mesh
slice through a common face, edge or point is created.
Members of this list are termed `outer' elements,
and all other elements are termed `inner' elements,
as illustrated in Figure~\ref{inner_outer_elements}.

Once the outer elements have been identified following a standard procedure
\citep[see e.g.,][]{DaNa98,MaKoBlLe08,Mic09,MiKo10,KoErGoMi10,Kom11},
MPI buffers are filled and a non-blocking MPI call is issued, which initiates communication and returns immediately.
While MPI messages are traveling across the network, computations are performed on inner elements.
Achieving effective overlap requires that the ratio of the number of inner to outer elements is sufficiently
large, a condition which is satisfied for suitably large mesh slices.
Under these circumstances, MPI data transfer will generally finish before the completion of  computations on inner elements.

%% ------------------------------------------------------------------------ %%
\section{Sample simulations}\label{sec:example}
%% ------------------------------------------------------------------------ %%

In this section, we present various simulations with increasing complexity to highlight the flexibility of
our new spectral-element package.
We start with a layercake model and finish with an example
of an arbitrarily-shaped model.

\subsection{Validation example: Two-Layer model}\label{subsec:examplevalidation}
%% ------------------------------------------------------------------------ %%

The SEM has been well benchmarked against discrete wavenumber methods
for layercake models by \cite{KoTr99}.
Here we compare their two-layer model solution (Figure 8, left) against the solution obtained by the new code.
The model has a horizontal size of 134~km $\times$ 134~km,
with a depth of 60~km.
We discretize the model into 70,200 elements,
using an approximate element size of 1,000~m at the top and 4,500~m at the bottom.
A mesh tripling layer is placed below the upper layer, between 3~km and 10~km,
with the wavespeed properties of the lower layer.
We use \href{http://www.labri.fr/perso/pelegrin/scotch/}{SCOTCH} to partition the model onto six cores, each with 11,700 elements.
The final mesh is generated using GLL points for a polynomial degree $N = 5$,
which results in 9,025,941 global mesh points.

A vertical force is placed at a depth of 25.05~km in the middle of the model.
The source-time function is a Ricker wavelet with a dominant frequency of 0.4~Hz.
The simulation uses a time step of 6.5~ms and propagates for 6,000 steps.
We compare our solution with seismograms obtained by \cite{KoTr99} (Figure~9).
The mesh and seismograms are shown in Fig.~\ref{figure:validation}.
The seismograms match very closely with the reference solutions,
exhibiting almost identical displacements.
Maximum waveform differences reach $\sim$0.3\%, arising from differences in mesh geometry and source implementation.

The performance of the code is summarized in Fig.~\ref{figure:scaling},
using simulations with the optimized routines by
\cite{DeFiMu02} and a polynomial degree $N = 4$.
We are interested in how the code behaves when the number of calculations is decreased linearly with the number of CPU cores (strong scaling),
and how performance varies when the number of calculations on each core is kept constant while increasing the total number of CPU cores (weak scaling).
To assess strong scaling, we fix the total mesh size but vary the number of CPU cores used for the simulation.
We run the simulation for a duration of 4,000 time steps and show the corresponding average elapsed time
per time step in Fig.~\ref{figure:scaling}(a).
More interesting for high-performance applications, we assess weak scaling
by fixing the problem size per processor and varying the number of CPU cores.
This leads to higher mesh resolutions for an increasing number of CPU cores but should keep the average elapsed time
per time step constant.
We summarize the simulation times in Fig.~\ref{figure:scaling}(b).
The computations were performed on a high-performance cluster with compute nodes consisting of two
Intel Nehalem quad-core processors; each core has 3~GB of RAM.
The code scales linearly within $\sim$90\% up to 256~CPU cores for both strong and weak scaling,
achieving excellent performance on this parallel system.
Note that for the strong scaling examples shown here,
simulations using more than 64~CPUs see a performance decrease since
communications  no longer overlap, thus
they no longer profit from the default non-blocking MPI scheme \citep{MaKoBlLe08}.

\subsection{Mount St.~Helens example: Layercake model with surface topography}\label{subsec:examplemount}
%% ------------------------------------------------------------------------ %%

In order to include surface topography,
we import \href{http://srtm.csi.cgiar.org/}{SRTM} data
with an original resolution of 90~m \citep{JaReNeGue2008}
and convert it to UTM coordinates for the corresponding UTM zone.
We read in this dataset using \href{http://cubit.sandia.gov}{CUBIT} and create a surface honoring these data points.
A 3D volume is built manually with topography on top.

The simulation uses an explosive source at a depth of 5~km.
In Fig.~\ref{figure:mountsnapshots},
we show the vertical displacement field at the free surface at consecutive times.
Note that once the wavefield hits the model boundary, it gets absorbed by the Clayton-Engquist-Stacey absorbing boundary conditions.

\subsection{L'Aquila example: Layercake model honoring surface and Moho topography}\label{subsec:exampleaquila}
%% ------------------------------------------------------------------------ %%

The purpose of this example is to show that additional surfaces may be honored by the mesh,
for example the Moho.
We import not only surface topography,
but also create a Moho surface that is honored by the boundaries of the spectral elements.
The mesh for the L'Aquila region was built using an additional `Python' library that semi-automates the mesh creation process with \href{http://cubit.sandia.gov}{CUBIT} \citep{CaStLeKoPiTr08}.
Once these mesh files are constructed,
the default partitioning and database generation process may be used to create fully load-balanced
spectral-element simulations on an arbitrary number of parallel processors.

Fig.~\ref{figure:aquila} shows several snapshots of the
seismic wavefield at consecutive times for an anelastic material,
using a kinematic source description for the April~6, 2009, L'Aquila earthquake.
Simulations are accurate up to 5~Hz
and may be used to discriminate between different wavespeed models and/or kinematic source solutions.
These high-frequency simulations may be used to assess the response of engineered structures and may guide
the development of better seismic building codes for the L'Aquila region.

\subsection{SEG/EAGE salt dome example: Exploration model}\label{subsec:exampleSEG}
%% ------------------------------------------------------------------------ %%

Our new spectral-element package can combine acoustic and (an)elastic simulations
by coupling these distinct domains.
In this example, we generate acoustic waves in the top water layer and propagate them
down through a salt dome body included in the lower, anelastic domain.
The mesh honors the surface of the salt dome and the fluid-solid boundary, i.e., the bathymetry.

Fig.~\ref{figure:saltdome} shows the acoustic wavefield at the free surface of the water layer at different times.
The source is a pressure source, located slightly below the free surface in the water layer, with a Ricker source-time function.
The wavefield is reflected and refracted by the salt dome in the anelastic domain below the water layer.
Note how these reflected/refracted waves, which include P-to-S converted waves, are recorded in the water layer.

\subsection{Asteroid example: Arbitrarily-shaped model}\label{subsec:exampleAsteroid}
%% ------------------------------------------------------------------------ %%

This final example shows that our new software package may be used for simulating wave propagation in arbitrarily-shaped models, such as asteroid Eros, which was imaged by the NEAR spacecraft in 2000--2001.
This silicated asteroid is 34~km long with a peanut-like shape and is thought to be covered with a regolith layer,
corresponding to a blanket of loose material crushed by impacts \citep{RiMeGrBr05}.
We meshed the asteroid with 5,797,440 hexahedral elements having an approximate resolution of 70~m.
To simulate a thin, 70~m regolith layer superimposed on strong bedrock, as suggested by \cite{RoThVeMuWi02},
we assigned a low-wavespeed material to the elements touching the free surface and a high-wavespeed material to elements inside the asteroid, representing solid bedrock.

We simulated the propagation of seismic waves from a source represented by a point force normal to the surface.
The source-time function corresponds to a Dirac pulse lowpass filtered up to a cutoff frequency of 5 Hz.
Fig.~\ref{figure:asteroid} displays wavefield snapshots for the first $\sim$10 seconds of the simulation. It shows the propagation of P, S and surface waves with a refocusing effect on the opposite side.
The regolith layer strongly increases physical dispersion of  surface waves.
Peak ground accelerations are plotted in Fig.~\ref{figure:asteroidshakemap} for a simulation without a regolith layer,
showing that refocussing occurs on the asteroid.

%% ------------------------------------------------------------------------ %%
\section{Adjoint sensitivity kernels}\label{sec:adjointkernels}
%% ------------------------------------------------------------------------ %%

An important goal in seismology is to use differences between observed and simulated seismograms
to improve Earth and source models, that is, we are interested in the inverse problem.
An elegant way to address this issue is to take advantage of adjoint methods \citep{Tar84,TrTaLi05}
to calculate Fr\'echet derivatives for a predefined objective function.
These derivatives may then be used in a conjugate-gradient approach to minimize differences
between data and synthetics.
The key ingredients of such an adjoint approach are sensitivity kernels.
Following \cite{TrTaLi05}, \cite{LiTr06,LiTr08} and \cite{trompetal2010},
we show examples of sensitivity kernels for various models using our new software package.

\subsection{Elastic sensitivity kernels}\label{subsec:elastickernels}
%% ------------------------------------------------------------------------ %%

Following \cite{TrTaLi05},
we may write the variation of a misfit function $\chi$ as
\bequ \label{equ:misfitperturbation}
\delta \chi = \int_V \left(K_\rho  \, \delta \ln \rho   + K_{c_{jklm}}\,   \delta c_{jklm}     \right) \,\mathrm{d}^3 \bx\,,
\eequ
where $\delta {\mathrm{ln}} \rho = \delta \rho / \rho$ denotes relative perturbations in density
and $\delta c_{jklm}  $ denotes perturbations in the elastic tensor.
The misfit kernels are given by
\bequ
K_\rho    = \mbox{}- \rho  \, \int_0^T \bs^{\dag} (  T-t ) \cdot \partial_t^2 \bs (t) \,\mathrm{d}t \,,
\eequ
\bequ
K_{c_{jklm}}   = \mbox{}- \int_0^T \epsilon_{jk}^\dag (  T-t ) \, \epsilon_{lm}( t ) \,\mathrm{d}t \,,
\eequ
where $\epsilon_{lm} $ and $\epsilon_{jk}^\dag $ denote elements of the strain and adjoint strain tensors,
and where we have suppressed the spatial dependence to avoid clutter.

In an isotropic model, we have
$c_{jklm} = ( \kappa - 2 \mu / 3 )\, \delta_{jk} \delta_{lm} + \mu\, ( \delta_{jl} \delta_{km} + \delta_{jm} \delta_{kl} ) $,
and thus eq.~(\ref{equ:misfitperturbation}) may be rewritten as
\bequ
\delta \chi = \int_V \left( K_\rho   \,\delta {\mathrm{ln}} \rho
+ K_\mu  \, \delta {\mathrm{ln}} \mu
+ K_\kappa   \, \delta {\mathrm{ln}} \kappa     \right) \, \mathrm{d}^3 \bx \, .
\eequ
The isotropic misfit kernels $K_\mu$ and $K_\kappa$ are defined by
\bequ \label{equ:kerneliso}
K_\mu     =  \mbox{}- 2 \mu   \,\int_0^T \bD^{\dag} (  T-t ) : \bD (t) \,\mathrm{d}t \,,
\eequ
\bequ
K_\kappa    =  \mbox{}- \kappa   \,\int_0^T  \left[ \bnabla \cdot \bs^\dag (  T-t ) \right] \left[ \bnabla \cdot \bs (  t ) \right]\, \mathrm{d}t \,,
\eequ
where $\bD =  \frac{1}{2} [ \bnabla \bs + ( \bnabla \bs )^T ] - \frac{1}{3}( \bnabla \cdot \bs )\, {\bf I} $
and $\bD^\dag = \frac{1}{2} [ \bnabla \bs^\dag + ( \bnabla \bs^\dag )^T ] - \frac{1}{3}( \bnabla \cdot \bs^\dag )\, {\bf I} $
are the traceless strain deviator and its adjoint, respectively.
In terms of a parameterization involving compressional wavespeed $\alpha$, shear wavespeed $\beta$
and density $\rho$, the corresponding kernels are given by a linear combination of these primary kernels (\cite{TrTaLi05}, eq. 20):
\bequ \label{equ:kernelalpha}
K_\alpha    = 2 \left( \frac{\kappa + \frac{4}{3} \mu}{\kappa} \right) K_\kappa  \,,
\qquad
K_\beta  = 2 \left( K_\mu    - \frac{4}{3} \frac{\mu}{\kappa}\, K_\kappa \right)\,,
\qquad
K'_{\rho}   = K_\kappa   + K_\mu   + K_\rho  \,.
\eequ
Note that a suitable parameterization for isotropic inversions is to use bulk sound
wavespeed $\Phi= \sqrt{\kappa/\rho}$, shear wavespeed $\beta$ and density $\rho$  \citep{Tar87}.
Bulk sound and shear wavespeeds are independent combinations of the bulk and shear moduli $\kappa$ and $\mu$.
The corresponding kernels are given by
\bequ \label{equ:kernelbulk}
K_\Phi   = 2 K_\kappa  \,,
\qquad
K'_\beta  = 2 K_\mu  \,,
\qquad
K'_{\rho}   = K_\kappa   + K_\mu   + K_\rho  \,.
\eequ

We place an explosive source at a depth of 7~km and a horizontal distance of 16~km from the receiver
in a homogeneous model with topography around Mount St.~Helens.
The P wave at the receiver is used to construct a traveltime adjoint source for the kernel simulation.
Fig.~\ref{figure:kerneliso}(a) shows the isotropic kernels $K_\kappa$, $K_\mu$ and $K_\rho$, and
Fig.~\ref{figure:kerneliso}(b) the isotropic kernels $K_\alpha$, $K_\beta$ and $K'_{\rho}$
for the same model and source-receiver geometry.

Note that although we construct the adjoint source using the P wave,
significant non-zero S-wave sensitivity is visible in the $K_\beta$ and $K_\mu$ kernels.
We interpret these areas of high sensitivity as P-to-S scattering locations,
which affect the signal within the chosen time window.
As may be observed in Fig.~\ref{figure:kerneliso},
such scattering sensitivity is especially strong at the free surface close to the receiver.

\subsection{Acoustic sensitivity kernels}\label{subsec:acoustickernels}
%% ------------------------------------------------------------------------ %%

\cite{LiTr08} calculated global sensitivity kernels, which include sensitivity to the liquid outer core.
In this section, we present acoustic sensitivity kernels for general local- or regional-scale models.
Such kernels may be used, for example, in ocean acoustics, non-destructive testing and medical tomography.

For acoustic simulations, the kernels are given by
\bequ
K_\rho    =  \rho^{-1} \,\int_0^T [\bnabla\partial_t\phi^{\dag} (  T-t )] \cdot [\bnabla\partial_t\phi(t)] \,\mathrm{d}t \,,
\eequ
\bequ
K_\kappa   = \mbox{}-\kappa^{-1}  \, \int_0^T     \partial_t^2 \phi^\dag (  T-t ) \, \partial_t^2 \phi (  t ) \, \mathrm{d}t \,,
\eequ
where $\phi$ and $\phi^\dag$ denote the acoustic scalar potential and adjoint potential, respectively.
To illustrate these kernels,
we use a model with acoustic and elastic regions.
The model combines a water layer on top of a homogeneous elastic layer,
separated by a bathymetric surface.
The dimensions of the model volume are approximately 2~km $\times$ 2~km horizontally and 1~km in depth.
Bathymetry is taken from a location in front of Pearl Harbor (Hawaii, USA),
with a resolution of $\sim$11~m.
For the forward simulation,
we use a pressure source in the form of an explosion with a Gaussian source-time function,
and record pressure variations at the receiver.
Both source and receiver are in the acoustic domain at a depth of 10~m and $\sim$1.1~km apart from each other.
We use the simulated pressure variation within the measurement window as the pressure misfit for the adjoint source,
as explained in Appendix \ref{appendix:acoustickernels1}.

Fig.~\ref{figure:kernelwater} shows the corresponding combined acoustic and elastic kernels.
The kernels highlight how the pressure waveform in the chosen measurement window is affected by a head
wave (a Scholte wave) traveling along the sea floor.
Since the acoustic region does not support shear waves,
the kernels $K_\mu$ and $K_\beta$ are zero in this upper domain.
However, they do exhibit non-zero sensitivity in the elastic domain, due to P-to-S coupling along the sea floor.

\subsection{Noise sensitivity kernels}\label{subsec:noisekernels}
%% ------------------------------------------------------------------------ %%

\newcommand{\bfPhi}{\mbox{\boldmath $\bf \Phi$}}
\def\bfD{{\mathbf{D}}}
\def\bfI{{\mathbf{I}}}

As demonstrated by \cite{trompetal2010},
noise cross-correlation sensitivity kernels may also be calculated based on an adjoint method,
and the new package has the necessary capabilities to perform such calculations.
Consider two receivers located at $\bx^\alpha$ and $\bx^\beta$\,.
In seismic interferometry,
ensemble sensitivity kernels are calculated in terms of interactions between an
ensemble forward wavefield $\bfPhi^{\alpha}$\,,
generated at location $\bx^\alpha$\,,
and an ensemble adjoint wavefield $\bfPhi^{\dagger\,\alpha\beta}$\,,
generated at $\bx^\beta$ and
triggered by the differences between simulated and observed ensemble-averaged cross correlations
at $\bx^\alpha$ and $\bx^\beta$\,.
The isotropic ensemble sensitivity kernels are given by
\begin{equation}
\langle K_\rho\rangle=\mbox{}-\int\rho\,[\bfPhi^{\dagger\,\alpha\beta}(\mbox{}-t)\cdot\partial_t^2\bfPhi^{\alpha}(t)
+\bfPhi^{\dagger\,\beta\alpha}(\mbox{}-t)\cdot\partial_t^2\bfPhi^{\beta}(t)]\,d t
\,,
\label{eq:K_rho}
\end{equation}
\begin{equation}
\langle K_\mu\rangle=\mbox{}-\int 2\mu\,[\bfD^{\dagger\,\alpha\beta}(\mbox{}-t)\!:\!\bfD^\alpha(t)
+\bfD^{\dagger\,\beta\alpha}(\mbox{}-t)\!:\!\bfD^\beta(t)]\,d t
\,,
\label{eq:K_mu}
\end{equation}
\begin{equation}
\langle K_\kappa\rangle=\mbox{}-\int\kappa\,[\bnabla\cdot\bfPhi^{\dagger\,\alpha\beta}(\mbox{}-t)\,\bnabla\cdot\bfPhi^{\alpha}(t)
+ \bnabla\cdot\bfPhi^{\dagger\,\beta\alpha}(\mbox{}-t)\,\bnabla\cdot\bfPhi^{\beta}(t)]\,d t
\,,
\label{eq:K_kappa}
\end{equation}
where
\begin{equation}
\bfD^\alpha=\tfrac{1}{2}[\bnabla\bfPhi^\alpha+(\bnabla\bfPhi^\alpha)^T]-\tfrac{1}{3}(\bnabla\cdot\bfPhi^\alpha)\,\bfI
\,,
\end{equation}
\begin{equation}
\bfD^{\dagger\,\alpha\beta}=\tfrac{1}{2}[\bnabla\bfPhi^{\dagger\,\alpha\beta}+(\bnabla\bfPhi^{\dagger\,\alpha\beta})^T]
-\tfrac{1}{3}(\bnabla\cdot\bfPhi^{\dagger\,\alpha\beta})\,\bfI
\,,
\end{equation}
denote the traceless
ensemble strain deviator and corresponding adjoint.

Figure \ref{figure:kernelnoise} shows the isotropic kernel $\langle K_\beta \rangle$
calculated according to eq. (\ref{equ:kernelalpha}) using the primary isotropic ensemble sensitivity kernels given above.
Plotted are the two contributions from ensemble wavefields generated at the first receiver location,
$\bx^\alpha$, and the second receiver location, $\bx^\beta$,
as well as the combined ensemble sensitivity kernel $\langle K_\beta \rangle$, which is the sum of the two contributions.
The two receivers are placed at a distance of 65~km from each other on top of a homogeneous block model with dimensions of 134~km $\times$ 134~km horizontally and 60~km in depth.
We smooth the kernel contributions using a 3D Gaussian filter
with a standard deviation of 2~km in the horizontal and vertical directions.
Note that these noise sensitivity kernels exhibit strong three-dimensional variability.
Depth sensitivity is controlled by the period range (5--100~s in this example).

%% ------------------------------------------------------------------------ %%
\section{Conclusions and future work}\label{sec:Conclusions}
%% ------------------------------------------------------------------------ %%

We have taken advantage of recent advances in high-performance computing, fully unstructured hexahedral meshing, load balancing and
mesh partitioning to facilitate forward and adjoint simulations of seismic wave propagation in coupled fluid and solid domains.
Our new open source software package,
\href{http://www.geodynamics.org/cig/software/specfem3d}{SPECFEM3D} Version~2.0 `Sesame',
performs acoustic and (an)elastic simulations of seismic wave propagation in complex geological models.
Hexahedral meshes may be generated based on packages such as
\href{http://cubit.sandia.gov}{CUBIT},
\href{http://www.simulia.com/products/abaqus_fea.html}{Abaqus},
\href{http://www.ansys.com/}{ANSYS},
\href{http://www.gocad.org/www/}{GOCAD},
\href{http://gid.cimne.upc.es}{GiD},
\href{http://geuz.org/gmsh}{Gmsh},
\href{http://www.truegrid.com/}{TrueGrid}
or  \href{http://www.salome-platform.org/}{Salome},
but the simple in-house mesher used in previous versions of \href{http://www.geodynamics.org/cig/software/specfem3d}{SPECFEM3D}
remains available for back-compatibility.

Partitioning and load balancing meshes may be accomplished based on graph partitioning software,
such as \href{http://www.labri.fr/perso/pelegrin/scotch/}{SCOTCH}.
By coupling acoustic and (an)elastic wave propagation,
we are able to calculate related sensitivity kernels,
which are useful for waveform inversions in off-shore exploration seismology,
ocean acoustics, non-destructive testing and medical tomography.
Attenuation is important on all scales of seismic wave propagation and is accommodated based on a series of standard linear solids.
In particular for  simulations in medical tomography, strong attenuation and related dispersion play a dominant role.
Finally, the new package can be used to calculate finite-frequency noise cross-correlation sensitivity kernels,
which may be used for seismic interferometry.
In future work, we will add C-PML and GPU support to the package.
Visco-plastic and non-linear elastic rheologies \citep[e.g.,][]{XuBiGhWa03,Priscoetal2007} are accommodated by the
\href{http://geoelse.stru.polimi.it/}{GeoELSE} software package  \citep{StPaIg09,ChMoTsBaKrKaStKr10},
and we will consider such non-linear constitutive relationships in future releases of
\href{http://www.geodynamics.org/cig/software/specfem3d}{SPECFEM3D}.

The next grand challenge involves the development of 3D seismic imaging and inversion tools for the characterization
of earthquakes, Earth `noise', and mapping of Earth's interior on all scales,
that is, to address the seismological inverse problem.
The goal is to harness the power of forward and adjoint modeling tools,
such as \href{http://www.geodynamics.org/cig/software/specfem3d}{SPECFEM3D},
together with modern computers to enhance
the quality of images of Earth’s interior and the earthquake rupture process.
Most traditional tomographic methods utilize traveltime or phase information measured by comparing data with simulations,
and interpret such measurements based on ray theory or other approximate methods.
Because of the limitations of these approximate theories, only parts of seismograms can be used,
and initial models are generally restricted to be spherically symmetric.
With the new generation of modeling tools we can go beyond classical tomography
by using fully 3D initial models \citep[e.g.,][]{Akcelik02,Akcelik03,Askan07a,PChen07b,Fichtner09,Fi11},
and utilizing as much information contained in seismograms as possible \citep[e.g.,][]{Maggi2009,Valentine2010}.
Our approach will be to minimize frequency-dependent phase and amplitude
differences between simulated and observed seismograms based on adjoint techniques in combination with conjugate
gradient methods,
an approach we refer to as `adjoint tomography' \citep{TaLiMaTr09,TaLiMaTr2010}.
The development of such capabilities will affect the fields of exploration geophysics,
regional \& global seismology, ocean acoustics, non-destructive testing, medical tomography and helioseismology.

\begin{acknowledgments}
We thank two anonymous reviewers for comments which improved the manuscript.
All simulations were performed on a Dell cluster built and maintained by the
Princeton Institute for Computational Science \& Engineering (USA).
The open source spectral-element software package \href{http://www.geodynamics.org/cig/software/specfem3d}{SPECFEM3D} Version~2.0 `Sesame' used
for this article is available for download via the Computational Infrastructure for Geodynamics
(\href{http://geodynamics.org/}{CIG}).
3D graphics were produced with the open source visualization application \href{http://www.paraview.org/}{Paraview}.
We acknowledge support by the US National Science Foundation under grant EAR-0711177.
\end{acknowledgments}

%% ------------------------------------------------------------------------ %%
%  REFERENCE LIST
%% ------------------------------------------------------------------------ %%

\bibliography{Biblio1}

%% ------------------------------------------------------------------------ %%
%
% APPENDIX
%
% \appendix resets counters and redefines section heads
%\appendix
\begin{appendices}
%% ------------------------------------------------------------------------ %%

\section{Acoustic adjoint equations}\label{appendix:acoustickernels}
%% ------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------ %%
\subsection{Pressure waveform misfit kernels}\label{appendix:acoustickernels1}
%% ------------------ ------------------------------------------------------ %%

For acoustic tomographic studies, it is convenient to define a pressure misfit function
\bequ
\chi = \frac{1}{2} \sum_i \int || p_i^{\mathrm{syn}}(\bm) - p_i^{\mathrm{obs}} ||^2 \, \mathrm{d}t\, ,
\eequ
where $p_i^{\mathrm{syn}}$ is the numerically computed pressure and $p_i^{\mathrm{obs}}$ the observed pressure
at location~$\bx_i$\,.
The variation in pressure may be written in terms of the variation in the potential~$\phi$ as
\bequ
\delta p = \mbox{}- \partial_t^2 \delta \phi \,,
\eequ
which follows from the definition of the scalar potential $\phi$ in eq. (\ref{equ:pressure}).
The corresponding action in the acoustic case is given by
\bequ
\chi = \frac{1}{2} \sum_i \int || p_i^{\mathrm{syn}} - p_i^{\mathrm{obs}} ||^2 \, \mathrm{d}t
						- \int \int_{\Omega} \lambda\, \left[ \kappa^{-1}\, \partial_t^2 \phi - \bnabla \cdot \left( \rho^{-1} \,\bnabla \phi \right) - f  \right] \,\mathrm{d}^3\bx \, \mathrm{d}t\,,
\eequ
where $\lambda$ denotes a scalar Lagrange multiplier.
Setting $ \Delta p_i = p_i^{\mathrm{syn}}  - p_i^{\mathrm{obs}}$
and taking the variation of the action, we obtain
\barr \label{equ:acousticvariation}
\delta \chi & = & \sum_i \int \Delta p_i \, \delta p_i \, \mathrm{d}t
 - \int \int_{\Omega} \left[ \delta\kappa^{-1} \, \lambda \, \partial_t^2 \phi - \bnabla \cdot \left( \delta \rho^{-1} \, \lambda \,\bnabla \phi \right) - \lambda\, \delta f  \right] \, \mathrm{d}^3\bx \, \mathrm{d}t   \nonumber \\
& & \!\!\!\!\!\!\!\!\!\!\mbox{}- \int \int_{\Omega} \left[\kappa^{-1} \, \partial_t^2 \lambda - \bnabla \cdot \left( \rho^{-1} \bnabla \lambda \right) \right] \delta \phi \, \mathrm{d}^3\bx \, \mathrm{d}t
 - \int \int_{ \partial \Omega} \bnormal \cdot \left( \rho^{-1}\, \bnabla \lambda \right) \, \delta \phi \, \mathrm{d}^2\bx \, \mathrm{d}t \, .
\earr

Since eq. (\ref{equ:acousticvariation} ) must be stationary when no model perturbations are present,
i.e., $\delta\rho = 0$\,, $ \delta\kappa= 0$ and $ \delta f = 0$,  we obtain
\barr \label{equ:lagrangecondition}
&& \int \int_{\Omega} \left[ \kappa^{-1} \, \partial_t^2 \lambda - \bnabla \cdot \left( \rho^{-1} \bnabla \lambda \right) \right]\, \delta \phi \, \mathrm{d}^3\bx \, \mathrm{d}t  \nonumber \\
&& \mbox{} = \int \int_{\Omega}  \sum_i \Delta p_i \,   \delta( \bx - \bx_{i} ) \, \delta p \, \mathrm{d}^3\bx \, \mathrm{d}t
- \int \int_{ \partial \Omega} \bnormal \cdot \left( \rho^{-1} \bnabla \lambda \right) \,\delta \phi \, \mathrm{d}^2\bx \,\mathrm{d}t \nonumber \\
&& \mbox{} = \mbox{}-\int \int_{\Omega}  \sum_i \Delta p_i \;  \delta( \bx - \bx_{i} ) \,  \partial_t^2 \delta \phi  \, \mathrm{d}^3\bx \, \mathrm{d}t
- \int \int_{ \partial \Omega} \rho^{-1}\,\bnormal \cdot \bnabla \lambda\, \delta \phi \, \mathrm{d}^2\bx \, \mathrm{d}t \nonumber \\
&& \mbox{} = \mbox{}-\int \int_{\Omega}  \sum_i \partial_t^2\Delta p_i \;  \delta( \bx - \bx_{i} ) \,\delta \phi \, \mathrm{d}^3\bx \, \mathrm{d}t
- \int \int_{ \partial \Omega} \rho^{-1} \,\bnormal \cdot \bnabla \lambda \, \delta \phi \, \mathrm{d}^2\bx \, \mathrm{d}t\,,
\earr
where $\bx_{i}$ is the station location of the corresponding $i$th measurement.
Note that the last integration by parts of the first term
is valid under the assumption that $\Delta p_i$ and $ \partial_t \Delta p_i $ vanish at the limits of the time integration,
i.e., for a given measurement window $[0,T]$, $ \Delta p_i ( \bx, 0 ) = \Delta p_i ( \bx, T ) = 0 $ and $ \partial_t \Delta p_i ( \bx , 0 ) = \partial_t \Delta p_i ( \bx , T ) = 0 $.
This is generally true because we taper the ends of the misfit window.

Let us define the adjoint scalar potential as
\bequ
\phi^\dag ( \bx , t ) \equiv \lambda (\bx , T - t )\, .
\eequ
It follows from (\ref{equ:lagrangecondition}) that $\phi^\dag$ must satisfy the adjoint wave equation
\bequ
\kappa^{-1}  \partial_t^2 \phi^\dag - \bnabla \cdot \left( \rho^{-1} \bnabla \phi^\dag \right)
= f^\dag \,,
\eequ
where the adjoint source is given by
\bequ
f^\dag ( \bx, t) = \mbox{}- \sum_i  \partial_t^2 \Delta p_i(T-t )  \;  \delta( \bx - \bx_{i} ) \,.
\eequ
The initial conditions for the adjoint potential must satisfy $\phi^\dag (T ) = 0 $ and $\partial_t \phi^\dag (T ) = 0$\,.
The corresponding fluid-solid boundary conditions involve terms with $ \rho^{-1} \bnormal \cdot \bnabla\phi^\dag$\,.

For acoustic simulations, there is no shear contribution and we may set $K_\mu = 0$.
Using
\bequ
 \bnabla \cdot \bs = \mbox{}- \kappa^{-1} \, p  =  \kappa^{-1} \, \partial_t^2 \phi \,,
\eequ
\bequ
\bnabla \cdot \bs^\dag = \kappa^{-1} \, \partial_t^2 \phi^\dag \,,
\eequ
the kernel $K_\kappa$ given in eq.~(\ref{equ:kerneliso} ) becomes
\bequ
K_\kappa   = \mbox{} - \int_0^T  \kappa^{-1} \,  \partial_t^2 \phi^\dag (T-t )\,  \partial_t^2 \phi (t ) \, \mathrm{d}t\, .
\eequ
It is this last kernel expression that is actually implemented,
since the values for $ \partial_t^2 \phi $ and $ \partial_t^2 \phi^\dag$ are obtained at each time step
in the Newark time scheme used to propagate acoustic waves.

%% ------------------------------------------------------------------------ %%
\subsection{Pressure traveltime adjoint sources}\label{appendix:acoustickernels2}
%% ------------------ ------------------------------------------------------ %%

Instead of measuring waveform misfits, one may also define a traveltime misfit for pressure signals, i.e.,
\bequ
\chi = \frac{1}{2}  \sum_i  || T_i^{\mathrm{syn}}(\bm) - T_i^{\mathrm{obs}}||^2 \,,
\eequ
where $T_i^{\mathrm{syn}}(\bm)$ denotes the arrival time in the synthetic pressure records
computed for model $\bm$, and $T_i^{\mathrm{obs}}$ the arrival time of the observed pressure wave.
The variation in traveltime $\delta T$ may be written to first order in terms of perturbations in pressure as $\delta p$ (Hung \& Dahlen 2000)
\bequ
\delta T = \frac{1}{N} \int \partial_t p \, \delta p \, \mathrm{d}t \,,
\eequ
where $ N =  \int p\,\partial_t^2 p\, \mathrm{d}t$\,.
Using $ \delta p = \mbox{}- \partial_t^2 \delta \phi $, this leads to
\bequ
\delta T = \mbox{}- \frac{1}{N} \int \partial_t p \, \partial_t^2 \delta \phi \, \mathrm{d}t\, .
\eequ
Defining $\Delta T_i \equiv T_i^{\mathrm{syn}}(\bm) - T_i^{\mathrm{obs}} $\,,
the variation of the action becomes
\barr
\sum_i \Delta T_i\, \delta T_i
& = & \mbox{}- \sum_i \frac{1}{N} \,\Delta T_i \,\int \partial_t p \, \partial_t^2 \delta \phi \, \mathrm{d}t  \\
& = & \mbox{}- \int \int_\Omega \sum_i  \frac{1}{N}\, \Delta T_i\, \delta(\bx - \bx_{i}) \,\partial_t p \, \partial_t^2 \delta \phi \,  \mathrm{d}^3\bx \, \mathrm{d}t  \,.
\earr
Under the assumption that $\partial_t p(0) = \partial_t p(T) = 0$ and $\partial_t^2 p (0) = \partial_t^2 p(T) = 0$
(which can be achieved by carefully selecting and tapering the measurement time windows),
we find after some further manipulation that the adjoint source for a traveltime misfit between observed
and simulated pressure signals is given by
\bequ
f^\dag ( \bx, t) = \mbox{}- \sum_i \frac{1}{N} \, \Delta T_i \, \partial_t^3 p( \bx, T- t)  \,  \delta( \bx - \bx_{i} ) \,.
\eequ

\end{appendices}
%% ------------------------------------------------------------------------ %%
%%
%% FIGURES
%%
%% PLEASE NOTE: WHEN YOU SUBMIT YOUR LATEX FILE TO GEMS, COMMENT OUT ANY COMMANDS
%% THAT INCLUDE GRAPHICS (see example below).
%%
%% ---------------
\clearpage

\begin{figure}
\begin{center}
\includegraphics[width=0.8\textwidth]{./images/processing.jpg}
\end{center}
\caption{Workflow for running spectral-element simulations with \href{http://www.geodynamics.org/cig/software/specfem3d}{SPECFEM3D} Version~2.0 `Sesame'.
}
\label{figure:processing}
\end{figure}

\begin{figure}
\begin{center}
\begin{minipage}[t]{0.45\textwidth}
\begin{center}
(a)\\
\includegraphics[width=1.\textwidth]{./images/mesh_mountsthelens.jpg}
\end{center}
\end{minipage}
\begin{minipage}[t]{0.45\textwidth}
\begin{center}
(b)\\
\includegraphics[width=1.\textwidth]{./images/mesh_aquila.jpg}
\end{center}
\end{minipage}
\begin{minipage}[t]{0.45\textwidth}
\begin{center}
(c)\\
\includegraphics[width=1.\textwidth]{./images/mesh_salt.jpg}
\end{center}
\end{minipage}
\begin{minipage}[t]{0.45\textwidth}
\begin{center}
(d)\\
\includegraphics[width=1.\textwidth]{./images/mesh_asteroid.jpg}
\end{center}
\end{minipage}
\begin{minipage}[t]{0.45\textwidth}
\begin{center}
(e)\\
\includegraphics[width=1.\textwidth]{./images/mesh_coffeecup.jpg}
\end{center}
\end{minipage}
\end{center}
\caption{Mesh examples: (a) Mount St.~Helens meshed by hexahedral elements.
The mesh honors surface topography and includes a mesh tripling layer in the middle of the model.
The smallest element size is approximately 280~m.
(b) L'Aquila, Italy, region discretized for high-frequency simulations.
The mesh honors surface and Moho topography and includes two mesh tripling layers.
The yellow and blue volumes denote slower and faster than average wavespeeds, respectively.
(c) Salt dome body meshed inside an exploration model for a SEG/EAGE benchmark test.
(d) 3D hexahedral mesh of the asteroid 433-Eros.
(e) Arbitrarily-shaped mesh for coupled solid-fluid simulations involving a coffee cup.
}
\label{figure:mountsthelens}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[width=0.49\textwidth]{./images/mount-partitions.jpg}
\includegraphics[width=0.49\textwidth]{./images/mount-partitions2.jpg}
\end{center}
\caption{Mount St.~Helens mesh partitioned and load balanced to run in parallel on four cores.
The four partitions are indicated by different colors.
}
\label{figure:partitions}
\end{figure}

\begin{figure}
\includegraphics[width=0.49\textwidth]{./images/mesh_inner_outer.jpg}
\includegraphics[width=0.49\textwidth]{./images/mesh_inner_outer2.jpg}
\caption{
Outer (highlighted) and inner (transparent colors) elements for the mesh shown in
Figure~\ref{figure:partitions}. Outer elements have at least one point in common with an element
from another slice and must therefore be computed first, before initiating non-blocking MPI
communications.
While MPI messages are traveling across the computer network,
simultaneous computations are performed on inner elements.
Non-blocking MPI communication is crucial to obtain good scaling
results for simulations running on a large number of parallel cores.}
\label{inner_outer_elements}
\end{figure}

\begin{figure}
\begin{center}
\begin{minipage}[t]{0.45\textwidth}
\includegraphics[width=1.\textwidth]{./images/2lay_mesh_boundary_fig8.jpg} \\
\includegraphics[width=1.\textwidth]{./images/2lay_mesh_partitions.jpg}
\end{minipage}
\begin{minipage}[t]{0.45\textwidth}
\includegraphics[width=1.\textwidth]{./images/2lay_traces55_R.pdf} \\
\includegraphics[width=1.\textwidth]{./images/2lay_traces31_R.pdf}
\end{minipage}
\end{center}
\caption{Validation for a two-layer mesh (left, top), using six partitions (left, bottom),
and seismograms recorded at the surface at horizontal distances of 2.39~km (right, top) and 31.11~km (right, bottom).
Plotted are radial displacements (SEM, red) against reference solutions (REF, black) from \cite{KoTr99},
as well as their exaggerated differences (blue) .
}
\label{figure:validation}
\end{figure}

\begin{figure}
\begin{center}
\begin{minipage}[t]{0.49\textwidth}
\begin{center}
(a) \\
\includegraphics[width=.9\textwidth]{./images/output_elapsed_time_strong.pdf}
\end{center}
\end{minipage}
\begin{minipage}[t]{0.49\textwidth}
\begin{center}
(b) \\
\includegraphics[width=.9\textwidth]{./images/output_elapsed_time_weak.pdf}
\end{center}
\end{minipage}
\end{center}
\caption{CPU scaling results for the model shown in Fig.~\ref{figure:validation}, (a) using a fixed total problem size (strong scaling) and (b) a fixed problem size per processor (weak scaling) for up to 256 cores.
Perfect weak scaling deviates slightly from a straight line,
because a larger number of processors involves more MPI buffers and therefore more computational overhead.
}
\label{figure:scaling}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[width=0.45\textwidth]{./images/mount2.jpg}
\includegraphics[width=0.45\textwidth]{./images/mount3.jpg}
\includegraphics[width=0.45\textwidth]{./images/mount4.jpg}
\includegraphics[width=0.45\textwidth]{./images/mount5.jpg}
\end{center}
\caption{Wavefield snapshots around Mount St.~Helens. Plotted are vertical displacements (up/down colored red/blue respectively) at the free surface of the model.
}
\label{figure:mountsnapshots}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[width=0.45\textwidth]{./images/aquila_1d_7000source-1.jpg}
\includegraphics[width=0.45\textwidth]{./images/aquila_1d_7000source-2.jpg}
\includegraphics[width=0.45\textwidth]{./images/aquila_1d_7000source-3.jpg}
\includegraphics[width=0.45\textwidth]{./images/aquila_1d_7000source-4.jpg}
\end{center}
\caption{Wavefield snapshots for the April~6, 2009, L'Aquila earthquake,
taken after 6~s, 11~s, 16~s and 21~s.
Plotted are vertical displacements (up/down as red/blue).
}
\label{figure:aquila}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[width=0.45\textwidth]{./images/salt-mesh_1_0005.jpg}
\includegraphics[width=0.45\textwidth]{./images/salt-mesh_1_0010.jpg}
\includegraphics[width=0.45\textwidth]{./images/salt-mesh_1_0015.jpg}
\includegraphics[width=0.45\textwidth]{./images/salt-mesh_1_0020.jpg}
\end{center}
\caption{Wavefield snapshots for an exploration geophysics simulation taken after 5~s,
10~s, 15~s and 20~s.
Plotted are vertical velocities at the free surface of the water layer.
}
\label{figure:saltdome}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[width=0.45\textwidth]{./images/asteroid_1.jpg}
\includegraphics[width=0.45\textwidth]{./images/asteroid_2.jpg}
\includegraphics[width=0.45\textwidth]{./images/asteroid_3.jpg}
\includegraphics[width=0.45\textwidth]{./images/asteroid_4.jpg}
\end{center}
\caption{Wavefield snapshots for an asteroid simulation taken after 3~s,
4.5~s, 6.5~s and 10.5~s.
Plotted is the norm of the velocity at the free surface of the asteroid.
}
\label{figure:asteroid}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[width=1.\textwidth]{./images/asteroid_shakemap1.jpg}
\end{center}
\caption{ShakeMap views for an asteroid simulation.
Plotted are different views of the peak ground accelerations at the free surface of the asteroid.
}
\label{figure:asteroidshakemap}
\end{figure}

\begin{figure}
\begin{center}
\begin{minipage}[t]{0.49\textwidth}
\begin{center}
(a)\\
\includegraphics[width=1.\textwidth]{./images/mount_kappa_kernel.jpg} \\
\includegraphics[width=1.\textwidth]{./images/mount_mu_kernel.jpg}\\
\includegraphics[width=1.\textwidth]{./images/mount_rho_kernel.jpg}
\end{center}
\end{minipage}
\begin{minipage}[t]{0.49\textwidth}
\begin{center}
(b)\\
\includegraphics[width=1.\textwidth]{./images/mount_alpha_kernel.jpg} \\
\includegraphics[width=1.\textwidth]{./images/mount_beta_kernel.jpg} \\
\includegraphics[width=1.\textwidth]{./images/mount_rhop_kernel.jpg}
\end{center}
\end{minipage}
\end{center}
\caption{Traveltime sensitivity to elastic structure.
Fr\'echet derivatives for isotropic parameterizations (a) $K_\kappa$, $K_\mu$ \& $K_\rho$ and (b) $K_\alpha$, $K_\beta$ \& $K'_{\rho}$ are compared in a model of Mount St.~Helens using traveltime adjoint sources for the P wave.
Shown are vertical cross sections through the source-receiver line and perpendicular to this line.
}
\label{figure:kerneliso}
\end{figure}

\begin{figure}
\begin{center}
\begin{minipage}[t]{0.49\textwidth}
\begin{center}
(a)\\
\includegraphics[width=1.\textwidth]{./images/pearl_kappa_kernel.jpg}\\
\includegraphics[width=1.\textwidth]{./images/pearl_mu_kernel.jpg}\\
\includegraphics[width=1.\textwidth]{./images/pearl_rho_kernel.jpg}
\end{center}
\end{minipage}
\begin{minipage}[t]{0.49\textwidth}
\begin{center}
(b)\\
\includegraphics[width=1.\textwidth]{./images/pearl_alpha_kernel.jpg}\\
\includegraphics[width=1.\textwidth]{./images/pearl_beta_kernel.jpg}\\
\includegraphics[width=1.\textwidth]{./images/pearl_rhop_kernel.jpg}
\end{center}
\end{minipage}
\end{center}
\caption{Waveform sensitivity to acoustic and elastic structure in a coupled fluid-solid simulation.
The bathymetric surface of the Pearl Harbor model, separating the two media, is shown in gray
together with a vertical cross-section through source (right) and station (left).
Plotted are combined acoustic and elastic kernels using a parameterization (a) $K_\kappa $, $K_\mu$ \& $K_\rho$
and (b) $K_\alpha$, $K_\beta$ \& $K'_{\rho}$.   }
\label{figure:kernelwater}
\end{figure}

\begin{figure}
\begin{center}
\begin{minipage}[t]{0.49\textwidth}
\begin{center}
(a)\\
\includegraphics[width=1.\textwidth]{./images/noise_beta_1st.jpg}\\
\end{center}
\end{minipage}
\begin{minipage}[t]{0.49\textwidth}
\begin{center}
(b)\\
\includegraphics[width=1.\textwidth]{./images/noise_beta_2nd.jpg}\\
\end{center}
\end{minipage}
\begin{minipage}[t]{0.49\textwidth}
\begin{center}
(c)\\
\includegraphics[width=1.\textwidth]{./images/noise_beta_sum.jpg}\\
\end{center}
\end{minipage}
\end{center}
\caption{Noise cross-correlation sensitivity to elastic structure.
Shown are (a) first, (b) second
and (c) summed contributions to the $\langle K_\beta\rangle$ Fr\'echet derivative
in a homogeneous isotropic model.
Plotted are vertical and horizontal cross sections through the line connecting the two receivers (white dots) and perpendicular to this line.
}
\label{figure:kernelnoise}
\end{figure}

\end{document}

### 2012-05-18 doc/GJI_2011_paper_about_the_code/response-reviews.tex
added the coffee mug mesh from our 2011 GJI paper

% Cover letter using letter.sty
\documentclass{letter} % Uses 10pt
\usepackage[dvipsnames,usenames]{color}
%Use \documentstyle[newcent]{letter} for New Century Schoolbook postscript font
% the following commands control the margins:
\topmargin=-1in    % Make letterhead start about 1 inch from top of page 
\textheight=8in  % text height can be bigger for a longer letter
\oddsidemargin=0pt % leftmargin is 1 inch
\textwidth=6.5in   % textwidth of 6.5in leaves 1 inch for right margin

%% More layout: Get rid of indenting throughout entire document
\setlength{\parindent}{0pt}

\usepackage{graphicx}

\newcommand{\point}[2]{
 %\begin{flushright}
  %\fcolorbox{black}{myblue}{
    \parbox[t]{1.\textwidth}{
	%\begin{itemize}
	{\it \textbf{#1:}} {\it #2}
	%\end{itemize}
	}
   %}
 %\end{flushright}
 % \vspace{5mm} 
}

\newcommand{\response}[1]{
 %\begin{flushright}
  %\fcolorbox{black}{myblue}{
    \parbox[t]{1.\textwidth}{
	%\begin{itemize}
	{\bf Response: } #1
	%\end{itemize}
	}
   %}
 %\end{flushright}
  \vspace{5mm} 
}

%% colors to show the corrections
\newcommand{\red}[1]{\textbf{\textcolor{Red}{#1}}}
\newcommand{\blue}[1]{\textbf{\textcolor{Blue}{#1}}}
\newcommand{\cyan}[1]{\textbf{\textcolor{Cyan}{#1}}}
\newcommand{\green}[1]{\textbf{\textcolor{Green}{#1}}}
\newcommand{\magenta}[1]{\textbf{\textcolor{Magenta}{#1}}}
\newcommand{\orange}[1]{\textbf{\textcolor{Orange}{#1}}}
%\newcommand{\red}[1]{#1}
%\newcommand{\blue}[1]{#1}
%\newcommand{\cyan}[1]{#1}

%%%%%%%%%% DK DK
% comments between authors
\newcommand{\todaniel}[1]{\textbf{\red{*** To Daniel, from Dimitri: #1 ***}}}
%\newcommand{\todimi}[1]{\textbf{\magenta{*** To Dimitri, from myself: #1 ***}}}
\newcommand{\fromdimi}[1]{\textbf{\magenta{*** From Dimitri: #1 ***}}}
%%%%%%%%%% RM RM
% comments between authors
\newcommand{\rolandtodaniel}[1]{\textbf{\blue{*** To Daniel, from Roland: #1 ***}}}

\begin{document}

\signature{Daniel Peter, \\
Dimitri Komatitsch,\\ 
Yang Luo,\\ 
Roland Martin, \\
Nicolas Le Goff, \\
Emanuele Casarotti, \\
Pieyre Le Loher, \\
Federica Magnoni, \\
Qinya Liu, \\
C\'eline Blitz, \\
Tarje Nissen-Meyer, \\
Piero Basini, \\
Jeroen Tromp }           % name for signature 
\longindentation=0pt                       % needed to get closing flush left
\let\raggedleft\raggedright                % needed to get date flush left

 %%------------------------------------------------------------------------------------------------%%
\begin{letter}
{Prof. Jeannot Trampert \\
Editor, Geophysical Journal International}
 %%------------------------------------------------------------------------------------------------%%

%%------------------------------------------------------------------------------------------------%%

\opening{Dear Jeannot:}

This is our response to the reviews of GJI manuscript GJI-S-11-0077 entitled ``Forward and adjoint simulations of seismic wave propagation on unstructured hexahedral meshes''. Reviewer's comments are in italics, our response is in normal type. All equation and figure numbers refer to the resubmitted manuscript, unless noted otherwise.

\vspace{10mm}
%%------------------------------------------------------------------------------------------------%%

{\bf {\large Response to reviewer: 1}}

%%------------------------------------------------------------------------------------------------%%

\point{1.1}{GENERAL REMARK:
With the exception of the introduction, the paper is well-written and interesting to read. Nevertheless, there are substantial deficiencies. These concern (1) the scientific content that is not sufficiently clear; (2) the comparison between continuous and discontinuous Galerkin methods that seems subjective, over-simplified, insufficiently quantitative and therefore incorrect; and (3) the failure to make proper reference to other researcher’s work.
}

\textbf{Response:} Each of these three issues is addressed in subsequent responses.

\point{1.2}{
SCIENTIFIC CONTENT: It is not easy to discover the scientific content of the manuscript and the problems that have been solved. The use of unstructured meshes is standard throughout the finite-element community, and a brief look at the websites of the most common meshing software is sufficient to confirm this. In fact, for most finite-element codes – including SEM – it does not matter where the mesh is coming from, and whether it is structured or not. Examples from seismic wave propagation for mesher/solver independence and unstructured meshes may be found, for instance in Stupazzini (2004), Kaeser et al. (2007), Stupazzini et al. (2009) or at http://geoelse.stru.polimi.it/.

For this manuscript to become publishable in a scientific journal, the authors should clarify which difficulties they had to overcome. Which mathematical or physical problem has been solved?

At least from my point of view, it is not sufficient to generate a collection of unstructured meshed with commercial software and to plug them into a pre-existing finite-element solver.

In its current form, the manuscript gives the impression that it is intended to be advertisement rather than a communication of scientific results.
}

\textbf{Response:}
We believe that the development of tools and services for seismology, e.g., Sambridge's ``Neighborhood Algorithm'',
Maggi et al.'s ``Flexwin'' package, Valentine \& Woodhouse's automated data selection algorithm, or Tromp et al.'s ``ShakeMovie'' system, is of tremendous scientific value, and such papers have important ``scientific content''.
Even though our paper may not be addressing a specific seismological problem,
it presents a tool that has the potential to transform seismology at a fundamental level, and we firmly believe that such work merits publication in GJI.
As computation comes to play a more-and-more important role in geophysics in general, and seismology in particular,
students and postdocs who spend enormous amounts of time and effort on the development of open source  tools that empower the broader community must get the recognition and admiration they deserve.
A publication such as this provides a proper reference for their efforts, and serves as a good introduction to the SEM and the related software.

We find it rather upsetting that some of our efforts are trivialized by the reviewer through statements such as ``...and plug them into a pre-existing finite-element solver''. If this were true, the development of software such as ours would not involve several years worth of painstaking effort on the part of a large number of people.

Finally, let us note that Appendix~A contains entirely new material related to adjoint simulations in acoustic media, something that hopefully matches the reviewer's standard of ``scientific content''.
Additionally, the Conclusions section now contains a paragraph describing the next grand challenge, which is the inverse problem.

\point{1.3}{
CONTINUOUS AND DISCONTINUOUS METHODS: The part of the introduction that compares continuous SEM to DG is excessive and not as fair as it should be.

It is certainly a good idea to start a paper on one particular numerical method with a comparison to other methods. Therefore, I have difficulties to see why the authors are so extremely focused on DG, ignoring all the other methods that have been developed over the past couple of decades. Do not forget that FD remains to be the most widely used method in seismology. If you really want to make the point that SEM is better – in whatever sense – then you need to make a much more precise and comprehensive comparison.

Furthermore, the comparison to DG is rather biased. This will be obvious to any reader simply from the style of your expressions, even when he knows little about numerical wave propagation. Sentences like note that the SEM supersedes earlier developments of high-order methods based on discontinuous techniques are not very meaningful because they are completely over-simplified and therefore wrong (see more details below). Discontinuous Galerkin methods are a large and diverse class of numerical techniques that cannot be classified as inferior with a few imprecise statements. Also, it is well known that the suitability of a particular method depends strongly on the application. I am sure, that SEM would have trouble to compete, for instance, with FD when it comes to body wave propagation in simple media.

In summary, I think that the introduction must be improved substantially.
}

\response{
We have improved the introduction, in particular our discussion of continuous and discontinuous Galerkin methods is now more balanced.
The paper in general and the introduction in particular already contain a large number of references ($>130$!), and it is not our aim to argue that one numerical method is better than another, rather we aim to provide some basic context and perspective.
}

\point{1.4}{
REFERENCE TO OTHER RESEARCHER'S WORK: An example should be sufficient to explain what I mean: On page 2 the authors write:

``The SEM was originally developed in computational fluid dynamics (Patera 1984; Maday \& Patera 1989) and has been successfully adapted to address seismological problems. For an introduction to the SEM in seismology, see e.g., Komatitsch et al. (2005), Chaljub et al. (2007) and Tromp et al. (2008).''

While this is formally correct, the authors completely ignore that the SEM was introduced to seismology by Seriani et al. (1995), Faccioli et al. (1997) and Seriani (1998).  The reader gets the impression that references to SEM can only be found in their own papers.

The manuscript is replete with such examples related, for instance to the use of unstructured meshes for wave propagation, spectral-element codes for seismic applications, adjoint techniques, wave propagation in complex media and so on and so forth. Instead of giving a historically correct collection of references, the authors almost exclusively refer to themselves.

I think this is not only a question of completeness, but also a contradiction to common scientific fairness and good practice!
}

\response{
Again, the paper contains a large number of references, reflecting the fact that we have attempted to be as comprehensive as possible.
Still, we have reordered references to better reflect scientific history, and expanded the number of citations further to provide more balance.
Let us note that the statement ``the authors almost exclusively refer to themselves'' is rather an overreaction considering the volume of references.
Furthermore, the manuscript has 13 authors, so there is bound to be a significant number of references to previous work by this group.
}

%{\it {\large Minor remarks}}

\point{1.5}{P2L39: The statement that the SEM has been successfully adapted to seismological problems requires references to Seriani et al. (1995), Faccioli et al. (1997) and Seriani (1998) who introduced the method to seismology.
}

\response{
We reordered and added references, citing earlier work of Cohen et al. (1993), who introduced the SEM to seismic wave propagation problems.
}

\point{1.6}{P3L13: Yes, dynamic rupture simulations are an exception, but not for the reason given by the authors. The reason why DG with dynamic rupture works so well is probably the upwind scheme combined with the good dispersion properties. One can, however, show analytically that dynamic rupture with SEM does not work at all, because certain modes are amplified. See for instance, de la Puente et al. (2009). 
}

\response{
We modified the discussion, referring, among others, to Puente et al. (2009).
}

\point{1.7}{P3L19: The statement concerning constant material properties within the DG elements is not correct. In fact a nodal DG method with high-order solution representation looks quite similar to SEM! People may choose to represent their material properties as constant, but this is their own choice, and not a restriction on the method. For DG with non-constant material properties see for instance Castro et al. (2010) or Wilcox et al. (2010).
}

\response{
We removed this statement. But note that a benchmark of DG against normal-mode synthetics for a spherically symmetric Earth model, such as PREM, would fail to show acceptable results unless material properties vary across elements, thereby capturing wavespeed gradients. As far as we know, such a DG benchmark has thus far not been successfully performed.
}

\point{1.8}{P3L33: It is ``close'' in the sense that they are both Galerkin projections, just as FEM is.  However DG could in theory duplicate the SEM implementation with respect to tensorised basis functions.
}

\response{
Perhaps. We would love to see confirmation of this statement, and assess if the numerical cost is the same; we doubt it!
}

\point{1.9}{General: The text is full of superlatives such as very distorted mesh elements, very good accuracy, very high efficiency, and so on. Such expressions do not bear any scientific meaning. They must be either quantified or eliminated.
}

\response{
We have reduced the number of such statements, but we note that all such remaining statements include corroborating references for quantification and/or justification.
}

\point{1.10}{P7L4: What do you mean by average isotropic shear modulus? Does this refer to an average of elastic parameters that leads to an effective isotropic modulus? (In the sense of a Voigt average, for instance.)
}

\response{
It is a misnomer, which we removed. Equation (6) simply gives the time-evolution of the isotropic shear modulus.
}

\point{1.11}{P7L16: I think it is a bit too general when you write that experience has shown that three solids generally suffice for simulating an absorption band. In fact, this depends on the width of the absorption band. When the upper and lower cutoffs differ by a factor of around 10, then 3 solids are indeed sufficient. However, you will need more when the band is wider (see for instance Robertsson (1996)).
}

\response{
Yes, this issue is addressed in various references we give, specifically, in the context of the SEM, Savage et al. (2010).
}

\point{1.12}{Typos: Please spell-check the reference list. Very often words like Galerkin, France, Grenoble, Caracas, etc. do not start with capital letters.
}

\response{
Done, thanks.
}

\vspace{10mm}
%%------------------------------------------------------------------------------------------------%%

{\bf {\large Response to reviewer: 2}}

%%------------------------------------------------------------------------------------------------%%

\point{2.1}{
Pag. 8. Eq. 11 should hold when the source contribution is zero. If it is the case, please correct it.
}

\response{
Yes, it is valid away from the source, or after cessation of the source.
}

\point{2.2}{
Pag.8. I got missed in the discussion about the coupling between an acoustic medium and an elastic medium. Can authors explain how the exchange of pressure and normal displacement is performed at the interface? Does the code solve a linear system at each time step for the computation of the physical quantities at the interface or is the algorithm time-staggered and the exchanged contributions come from the previous time step ? 
}

\response{
For details regarding the implementation, the reviewer is referred to Komatitsch \& Tromp (2002ab).
Basically, exchanges between the fluid and solid are captured by the last sentences of sections 2.1 and 2.2.
The exchange involves the current time step, and there is no need to solve a linear system.
}

\point{2.3}{
Pag. 9. I am not convinced that hexahedral meshes have a smaller number of elements than tetrahedral meshes. It is true for regular grids, but as soon as corners are present in the model, coming from the modeling of a basin for instance, mesh refinement becomes more complex for hexahedra, and you can come out with a comparable or larger number of elements. I suggest to delete the part of the sentence “smaller element counts”. 
}

\response{
The references we provide corroborate this statement. Geological models generally contain large, simple volumes, which may be meshed with fewer hexahedral elements than tetrahedral elements.
}

\point{2.4}{
Pag. 9. Authors list several mesh generators, but only one of them comes with a reference. Can authors add references to all of the mesh generators ? 
}

\response{
All references involve active web links and we added more references to each of the meshing tools. Perhaps GJI could include the weblink as well in the References?
}

\point{2.5}{
Pag. 10 In the presented examples (Mount S. Helens, Aquila),  can authors also write the size of the smallest element in the mesh ? It will help in understanding the quality of the mesh as compared to the computational costs. In the figure of Aquila, some yellow bodies appear in the mesh. What are they referred to ? 
}

\response{
The smallest element size is approximately 280~m. The yellow and blue volumes in the L'Aquila model denote slower and faster than average wavespeeds.
}

\point{2.6}{
Pag. 13 L’Aquila case. Authors say that they honor the Moho interface. Can they add a reference to the Moho and discuss the accuracy at which they know it ? Authors state that with their mesh, they can discriminate between wave speed models. I’m a bit surprised because the mesh size itself depends on the wave speed. As compared with Mount S. Helens mesh, the two models mostly have the same discretization (90 m is the average element size), but Aquila model can tackle a larger frequency band. Why ? (Is the velocity model faster for this case ? Which one has been used ? )
}

\response{
See Casarotti et al. 2008 for a more detailed discussion of the L'Aquila model. We just mean to say that with this mesh for the L'Aquila region one can now play with various 3D wavespeed  and/or kinematic rupture models to assess their relative performance.
Per usual, element size (grid spacing) and the wavespeed model govern the highest frequencies on can resolve on a mesh.
}

\vspace{10mm}
%%------------------------------------------------------------------------------------------------%%

{\bf {\large Response to editor Prof. Jeannot Trampert:}}

%%------------------------------------------------------------------------------------------------%%

\point{4.1}{
While reviewer two has only minor comments, reviewer one suggests to put your work better into context of existing literature. I think this can all be addressed in a moderate revision. I am looking forward to a revised manuscript and a detailed response to all comments.
}

\response{
We have restructured and added references, thereby putting our work in a better historical context with existing literature.
In particular, the introduction has been significantly improved.
Note, however, that in the submitted paper we already included tens of references to papers written by others, thus reviewer \#1's statement about this is an exaggeration.
We updated Fig.~6 to present a more standard weak and strong scaling analysis.
We also added a figure explaining the benefits of non-blocking MPI communications.
}

%--------------------------------------------------------------------------------------------
% SUMMARY

We hope that with these modifications the manuscript is now acceptable for publication in GJI.

\closing{Sincerely,}

%\encl{}  				% Enclosures

\end{letter}

\end{document}

### 2012-05-19 doc/USER_MANUAL/manual_SPECFEM3D.tex
better IBM compiler options. also updated the manual

@@
 greater than 2 gigabytes, which is the upper limit for 32-bit addressing.
-In this case, on some compilers you may need to add \texttt{``-mcmodel=medium}''
+In this case, on some compilers you may need to add \texttt{``-mcmodel=medium}'' or \texttt{``-mcmodel=medium -shared-intel}''
 to the compiler options of CFLAGS, FCFLAGS and LDFLAGS otherwise the compiler will display an error
@@
 If they are identical (and in most cases they will), you can select single precision for your future runs.\\
+
+\section{Visualizing the subroutine calling tree of the source code}
+
+Packages such as \texttt{doxywizard} can be used to visualize the subroutine calling tree of the source code.
+\texttt{Doxywizard} is a GUI front-end for configuring and running \texttt{doxygen}.
 

### 2012-05-19 doc/USER_MANUAL/manual_SPECFEM3D.tex
added a description about OpenMP

@@
 
+\section{Adding OpenMP support in addition to MPI}
+
+OpenMP support can be enabled in addition to MPI. However, in many cases performance will not improve
+because our pure MPI implementation is already heavily optimized and thus the resulting code will in fact       
+be slightly slower. A possible exception could be IBM BlueGene-type architectures.
+
+To enable OpenMP, uncomment the OpenMP compiler option in two lines in file \texttt{src/specfem3D/Makefile.in}
+(before running \texttt{configure}) and also uncomment the \texttt{\#define USE\_OPENMP} statement in file\\
+\texttt{src/specfem3D/specfem3D.F90}.
+
+The DO-loop using OpenMP threads has a SCHEDULE property. The OMP\_SCHEDULE
+environment variable can set the scheduling policy of that DO-loop.
+Tests performed by Marcin Zielinski at SARA (The Netherlands) showed that often
+the best scheduling policy is DYNAMIC with the size of the chunk equal to the number of
+OpenMP threads, but most preferably being twice as the number of
+OpenMP threads (thus chunk size = 8 for 4 OpenMP threads etc).
+If OMP\_SCHEDULE is not set or is empty, the DO-loop will assume generic
+scheduling policy, which will slow down the job quite a bit.
+
 \section{Visualizing the subroutine calling tree of the source code}
@@
 \texttt{Doxywizard} is a GUI front-end for configuring and running \texttt{doxygen}.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 

### 2012-05-19 doc/USER_MANUAL/manual_SPECFEM3D.tex
new developers

@@
 Emanuele Casarotti,
+Joseph Charles,
 Min Chen,
@@
 Jeroen Tromp,
+Zhinan Xie,
 Hejun Zhu.\\

### 2012-05-20 doc/USER_MANUAL/manual_SPECFEM3D.tex
updated some references in the manual, and fixed a statement that was not fully correct in the PDF file of the time scheme we use

@@
 A detailed theoretical analysis of the dispersion
-and stability properties of the SEM is available in \citet{Coh02}, \citet{DeSe07} and \citet{SeOl07}.\\
+and stability properties of the SEM is available in \citet{Coh02}, \citet{DeSe07}, \citet{SeOl07} and \citet{SeOl08}.\\
 

### 2012-07-02 doc/USER_MANUAL/manual_SPECFEM3D.tex
updates user manual to explain new MODEL parameter in Par_file and adds chapter 10 about changing models

@@
 \texttt{domain\_ID} is 1 for acoustic or 2 for elastic materials, \texttt{material\_ID} a unique identifier,  \texttt{rho} the density in $kg \, m^{-3}$,  \texttt{vp} the P-wave speed in $m \, s^{-1}$, \texttt{vs}  the S-wave speed in $m \, s^{-1}$, \texttt{Q} the quality factor and \texttt{anisotropy\_flag} an identifier for anisotropic models.
-For a tomographic model, the material definition format is:
-\begin{lyxcode}
-\texttt{domain\_ID} \texttt{material\_ID}  \texttt{tomography}  \texttt{name}
-\end{lyxcode}
-where
-\texttt{domain\_ID} is 1 for acoustic or 2 for elastic materials, \texttt{material\_ID} a negative, unique identifier, \texttt{tomography} keyword for tomographic material definition and \texttt{name} the name of the tomography file.
+
+%%magnoni
+For tomographic velocity models, please read Chapter \ref{cha:-Changing-the} and Section \ref{sec:Using-tomographic} `Using external tomographic Earth models' for further details.
+%
+%For a tomographic model, the material definition format is:
+%\begin{lyxcode}
+%\texttt{domain\_ID} \texttt{material\_ID}  \texttt{tomography}  \texttt{name}
+%\end{lyxcode}
+%where
+%\texttt{domain\_ID} is 1 for acoustic or 2 for elastic materials, \texttt{material\_ID} a negative, unique identifier, \texttt{tomography} keyword for tomographic material definition and \texttt{name} the name of the tomography file.
 The name is not used so far, rather change the filename defined in the file \texttt{model\_tomography.f90} located in the
-\texttt{src/generate\_databases/} directory.
+%\texttt{src/generate\_databases/} directory.
+%
 \item [materials\_file] Contains the material associations for each element.
@@
 This feature is not used at the time of generating the distributed databases but is required for the solver.
-Please see also Section~\ref{cha:Choosing-the-Time-Step} for further details.
+Please see also Section~\ref{sec:Choosing-the-Time-Step} for further details.
+%
+\item [{\texttt{MODEL}}] Must be set to one of the following:
+%
+\begin{description}
+\item [{\textmd{Models defined by mesh parameters:}}]~
+\begin{description}
+\item [{\texttt{default}}] Uses model parameters as defined by meshing procedures described 
+in the previous Chapter~\ref{cha:Mesh-Generation}.
+\end{description}
+\end{description}
+%
+\begin{description}
+\item [{\textmd{1D~models~with~real~structure:}}]~
+\begin{description}
+\item [{\texttt{1D\_prem}}] Isotropic version of the spherically
+symmetric Preliminary Reference Earth Model (PREM) \citep{DzAn81}.
+\item [{\texttt{1D\_socal}}] A standard isotropic 1D model for Southern California.
+\item [{\texttt{1D\_cascadia}}] Isotropic 1D profile for the Cascadia region.
+\end{description}
+\end{description}
+%
+\begin{description}
+\item [{\textmd{Fully~3D~models:}}]~
+\begin{description}
+\item [{\texttt{aniso}}] For a user-specified fully anisotropic model.
+Parameters are set up in routines located in file \texttt{model\_aniso.f90} in directory \texttt{src/generate\_databases/}.
+See Chapter~\ref{cha:-Changing-the} for a discussion on how to specify your own 3D model.
+\item [{\texttt{external}}] For a user-specified isotropic model which uses externally defined model parameters.
+Uses external model definitions set up in routines located in file \texttt{model\_external\_values.f90} in directory \texttt{src/generate\_databases/}.
+Please modify these generic template routines to use your own model definitions.
+\item [{\texttt{gll}}] For a user-specified isotropic model which uses external binary files for $v_p$, $v_s$ and $\rho$.
+Binary files are given in the same format as when outputted by the \texttt{xgenerate\_databases} executable when using option \texttt{SAVE\_MESH\_FILES}.
+These binary files define the model parameters on all GLL points which can be used for iterative inversion procedures.
+\item [{\texttt{salton\_trough}}] A 3D $V_p$ model for Southern California.
+Users must provide the corresponding data file \texttt{regrid3\_vel\_p.bin} in directory \texttt{DATA/st\_3D\_block\_harvard/}.
+\item [{\texttt{tomo}}] For a user-specified 3D isotropic model which uses a tomographic model file \texttt{tomographic\_model.xyz} in directory \texttt{in\_data\_files/}.
+See Section \ref{sec:Using-tomographic},
+for a discussion on how to specify your own 3D tomographic model.
+\end{description}
+\end{description}
+%
+%
 \item [{\texttt{OCEANS}}] Set to \texttt{.true.} if the effect of the oceans
@@
 to create a movie of seismic wave propagation on the Earth's surface.
-Turning this option on generates large output files. See Section \ref{sec:Movies}
+Turning this option on generates large output files. See Section~\ref{sec:Movies}
 for a discussion on the generation of movies. This feature is only relevant for the solver.
@@
 to create a movie of seismic wave propagation in the Earth's interior.
-Turning this option on generates huge output files. See Section \ref{sec:Movies}
+Turning this option on generates huge output files. See Section~\ref{sec:Movies}
 for a discussion on the generation of movies. This feature is only relevant for the solver.
@@
 100 timesteps. The smaller you make this number the more output will
-be generated! See Section \ref{sec:Movies} for a discussion on the
+be generated! See Section~\ref{sec:Movies} for a discussion on the
 generation of movies. This feature is only relevant for the solver.
@@
 
-\section{\label{cha:Choosing-the-Time-Step}Choosing the time step \texttt{DT}}
+\section{\label{sec:Choosing-the-Time-Step}Choosing the time step \texttt{DT}}
 
@@
 \end{lyxcode}
+
+\chapter{{\normalsize \label{cha:-Changing-the}} Changing the Model}
+
+In this section we explain how to change the velocity model used for your simulations. 
+These changes involve contributing specific subroutines
+that replace existing subroutines in the \texttt{SPECFEM3D} package.
+Note that \texttt{SPECFEM3D} can handle Earth models with material properties that vary within each spectral element.
+
+%% magnoni 6/12
+\section{{\normalsize \label{sec:Using-tomographic}}Using external tomographic Earth models}
+
+To implement your own external tomographic model, you must provide your own tomography file
+\texttt{tomography\_model.xyz} and choose between two possible options: 
+(1) set in the \texttt{Par\_file} the model parameter \texttt{MODEL = tomo}, or
+(2) for more user control, set in the \texttt{Par\_file} the model parameter \texttt{MODEL = default} and use the following format in the file \texttt{MESH/nummaterial\_velocity\_file} when using CUBIT to construct your mesh (see also section \ref{subsec:Exporting-the-Mesh}):
+\begin{lyxcode}
+\texttt{domain\_ID} \texttt{material\_ID} \texttt{tomography} \texttt{elastic} \texttt{file\_name} 1
+\end{lyxcode} 
+where
+\texttt{domain\_ID} is 1 for acoustic or 2 for elastic materials, 
+\texttt{material\_ID} a negative, unique identifier (i.e., -1), 
+\texttt{tomography} keyword for tomographic material definition, 
+\texttt{elastic} keyword for elastic material definition, 
+\texttt{file\_name} the name of the tomography file and 
+1 a positive unique identifier. 
+The name of the tomography file is not used so far, rather change the filename (\texttt{TOMO\_FILENAME}) defined in the file \texttt{model\_tomography.f90} located in the \texttt{src/generate\_databases/} directory.
+
+The external tomographic model is represented by a grid of points with assigned material properties and homogeneous resolution along each spatial direction x, y and z.
+The ASCII file \texttt{tomography\_model.xyz} that describes the tomography should be located in the \texttt{in\_data\_files/} directory. 
+The format of the file, as read from \texttt{model\_tomography.f90}, looks like Figure \ref{fig:tomography_file},
+where
+%
+\begin{description}
+\item [{\texttt{ORIG\_X}, \texttt{END\_X}}] are, respectively, the coordinates of the initial and final tomographic grid points along the x direction (in the mesh units, e.g., $m$);
+\item [{\texttt{ORIG\_Y}, \texttt{END\_Y}}] respectively the coordinates of the initial and final tomographic grid points along the y direction (in the mesh units, e.g., $m$);
+\item [{\texttt{ORIG\_Z}, \texttt{END\_Z}}] respectively the coordinates of the initial and final tomographic grid points along the z direction (in the mesh units, e.g., $m$);
+\item [\texttt{SPACING\_X}, \texttt{SPACING\_Y}, \texttt{SPACING\_Z}] the spacing between the tomographic grid points along the x, y and z directions, respectively (in the mesh units, e.g., $m$);
+\item [\texttt{NX}, \texttt{NY}, \texttt{NZ}] the number of grid points along the spatial directions x, y and z, respectively; \texttt{NX} is given by [(\texttt{END\_X} - \texttt{ORIG\_X})/\texttt{SPACING\_X}]+1; \texttt{NY} and \texttt{NZ} are the same as \texttt{NX}, but for the y and z directions, respectively;
+\item [\texttt{VP\_MIN}, \texttt{VP\_MAX}, \texttt{VS\_MIN}, \texttt{VS\_MAX}, \texttt{RHO\_MIN}, \texttt{RHO\_MAX}] the minimum and maximum values of the wave speed \texttt{vp} and \texttt{vs} (in $m \, s^{-1}$) and of the density \texttt{rho} (in $kg \, m^{-3}$); these values could be the actual limits of the tomographic parameters in the grid or the minimum and maximum values to which we force the cut of velocity and density in the model.
+\end{description}
+%
+After the first four lines, in the file \texttt{tomography\_model.xyz} the tomographic grid points are listed with the corresponding values of \texttt{vp}, \texttt{vs} and \texttt{rho}, scanning the grid along the x coordinate (from \texttt{ORIG\_X} to \texttt{END\_X} with step of \texttt{SPACING\_X}) for each given y (from \texttt{ORIG\_Y} to \texttt{END\_Y}, with step of \texttt{SPACING\_Y}) and each given z (from \texttt{ORIG\_Z} to \texttt{END\_Z}, with step of \texttt{SPACING\_Z}).
+%fig 
+\begin{figure}[htbp]
+\noindent \begin{centering}
+\includegraphics[width=1.0\textwidth]{figures/tomo_file.jpg}
+\par\end{centering}
+\caption{Tomography file \texttt{tomography\_model.xyz} that describes an external Earth model and that is read by \texttt{model\_tomography.f90}.
+The coordinates x, y and z of the grid, their limits ORIG\_X, ORIG\_Y, ORIG\_Z, END\_X, END\_Y, END\_Z and the grid spacings SPACING\_X, SPACING\_Y, SPACING\_Z should be in the units of the constructed mesh (e.g., UTM coordinates in meters).}
+\label{fig:tomography_file}
+\end{figure}
+
+The user can implement his own interpolation algorithm for the tomography model by changing the routine \texttt{model\_tomography.f90} located in the \texttt{src/generate\_databases/} directory. 
+Moreover, for models that involve both fully defined materials and a tomography description, the \texttt{nummaterial\_velocity\_file} has multiple lines each with the corresponding suitable format described above.
+
+\section{{\normalsize \label{sec:Anelastic-Models}}External (an)elastic Models}
+
+To use your own external model, you can set in the \texttt{Par\_file} the model parameter \texttt{MODEL = external}.
+Three-dimensional acoustic and/or (an)elastic (attenuation) models may then be superimposed
+onto the mesh based upon your subroutine in
+\texttt{model\_external\_values.f90} located in directory \texttt{src/generate\_databases}.
+The call to this routine would be as follows
+
+\begin{lyxcode}
+call~model\_external\_values(xmesh,~ymesh,~zmesh, ~\& \\
+~~~~~~~~~~rho,~vp,~vs,~qmu\_atten,~iflag\_aniso,~idomain\_id)
+\end{lyxcode}
+%
+Input to this routine consists of:
+\begin{description}
+\item [{\texttt{xmesh,ymesh,zmesh}}] location of mesh point 
+\end{description}
+%
+Output to this routine consists of:
+\begin{description}
+\item [{\texttt{rho,vp,vs}}] isotropic model parameters for density $\rho$ ($kg/m^3$), $v_p$ ($m/s$) and $v_s$ ($m/s$)
+\item [{\texttt{qmu\_atten}}] Shear wave quality factor: $0<Q_{\mu}<9000$
+\item [{\texttt{iflag\_aniso}}] anisotropic model flag, $0$ indicating no anisotropy or $1$ using anisotropic model parameters as defined in routine file \texttt{model\_aniso.f90}
+\item [{\texttt{idomain\_id}}] domain identifier, $1$ for acoustic, $2$ for elastic, $3$ for poroelastic materials.
+\end{description}
+
+Note that the resolution and maximum value of anelastic models are
+truncated. This speeds the construction of the standard linear solids
+during the meshing stage. To change the resolution, currently at one
+significant figure following the decimal, or the maximum value (9000),
+consult \texttt{shared/constants.h}.
+
+\section{{\normalsize \label{sec:Anisotropic-Models}}Anisotropic Models}
+
+To use your anisotropic models, you can either set in the \texttt{Par\_file} the model parameter \texttt{ANISOTROPY} to \texttt{.true.} or set the model parameter \texttt{MODEL = aniso}.
+Three-dimensional anisotropic models may be superimposed on
+the mesh based upon the subroutine in file \texttt{model\_aniso.f90} 
+located in directory \texttt{src/generate\_databases}.
+The call to this subroutine is of the form
+
+\begin{lyxcode}
+call~model\_aniso(iflag\_aniso,~rho,~vp,~vs,~\& \\
+~~~~~~~~~~c11,c12,c13,c14,c15,c16,c22,c23,c24,c25,c26,~\& \\
+~~~~~~~~~~c33,c34,c35,c36,c44,c45,c46,c55,c56,c66)~
+\end{lyxcode}
+%
+Input to this routine consists of:
+\begin{description}
+\item [{\texttt{iflag\_aniso}}] flag indicating the type of the anisotropic model, i.e. $0$ for no superimposed anisotropy, $1$ or $2$ for generic pre-defined anisotropic models.
+\item [{\texttt{rho,vp,vs}}] reference isotropic model parameters for density $\rho$, $v_p$ and $v_s$.
+\end{description}
+%
+Output from the routine consists of the following non-dimensional
+model parameters:
+\begin{description}
+\item [{\texttt{c11},}] \textbf{$\cdots$,} \texttt{\textbf{c66}} 21~dimensionalized
+anisotropic elastic parameters.
+\end{description}
+
+You can replace the \texttt{model\_aniso.f90} file by
+your own version \textit{provided you do not change the call structure
+of the routine}, i.e., the new routine should take exactly the same
+input and produce the required relative output.
 

### 2012-07-06 src/shared/sum_kernels.f90
updates code file version numbers to version 2.1

@@ -1,3 +1,30 @@
+!=====================================================================
+!
+!               S p e c f e m 3 D  V e r s i o n  2 . 1
+!               ---------------------------------------
+!
+!          Main authors: Dimitri Komatitsch and Jeroen Tromp
+!    Princeton University, USA and University of Pau / CNRS / INRIA
+! (c) Princeton University / California Institute of Technology and University of Pau / CNRS / INRIA
+!                            April 2011
+!
+! This program is free software; you can redistribute it and/or modify
+! it under the terms of the GNU General Public License as published by
+! the Free Software Foundation; either version 2 of the License, or
+! (at your option) any later version.
+!
+! This program is distributed in the hope that it will be useful,
+! but WITHOUT ANY WARRANTY; without even the implied warranty of
+! MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+! GNU General Public License for more details.
+!
+! You should have received a copy of the GNU General Public License along
+! with this program; if not, write to the Free Software Foundation, Inc.,
+! 51 Franklin Street, Fifth Floor, Boston, MA 02110-1301 USA.
+!
+!=====================================================================
+
+
 ! sum_preconditioned_kernels
 !
 ! this program can be used for event kernel summation,

### 2012-07-06 doc/USER_MANUAL/manual_SPECFEM3D.tex
updates version tag to v2.0.2 in manual
updates code file version numbers to version 2.1

@@
 University of Pau / CNRS / INRIA (France)\\
-Version 2.0 \\
+Version 2.1 \\
 "Sesame"}
@@
 \noindent
-If you use this new version 2.0, which has non blocking MPI for much better performance for medium or large runs, please cite at least one of these six articles,
+If you use this new version 2.1, which has non blocking MPI for much better performance for medium or large runs, please cite at least one of these six articles,
 in which results of non blocking MPI runs are presented: \cite{PeKoLuMaLeCaLeMaLiBlNiBaTr11,KoErGoMi10,KoViCh10,Kom11,CaKoLaTiMiLeSnTr08,MaKoBlLe08}.\\
@@
 \begin{lyxcode}
-{\small tar~-zxvf~SPECFEM3D\_V2.0.1.tar.gz}{\small \par}
+{\small tar~-zxvf~SPECFEM3D\_V2.1.0.tar.gz}{\small \par}
 \end{lyxcode}
@@
 
+ MPI v 2.1 `Sesame' (Spectral ElementS on Any MEsh), July 2012: 
+ Max Rietmann, Peter Messmer, Daniel Peter: support for CUDA GPUs. \\
+
  MPI v. 2.0 `Sesame' (Spectral ElementS on Any MEsh), November 2010:

### 2012-07-07 src/shared/force_ftz.c
updated the copyright information and a few things in the manual, to prepare for the release of version 2.1

@@ -5,9 +5,9 @@
 !               ---------------------------------------
 !
 !          Main authors: Dimitri Komatitsch and Jeroen Tromp
-!    Princeton University, USA and University of Pau / CNRS / INRIA
-! (c) Princeton University / California Institute of Technology and University of Pau / CNRS / INRIA
-!                            April 2011
+!    Princeton University, USA and CNRS / INRIA / University of Pau
+! (c) Princeton University / California Institute of Technology and CNRS / INRIA / University of Pau
+!                             July 2012
 !
 ! This program is free software; you can redistribute it and/or modify
 ! it under the terms of the GNU General Public License as published by

### 2012-07-07 doc/USER_MANUAL/manual_SPECFEM3D.tex
updated the copyright information and a few things in the manual, to prepare for the release of version 2.1

@@
 \author{$\copyright$ Princeton University (USA) and\\
-University of Pau / CNRS / INRIA (France)\\
-Version 2.1 \\
-"Sesame"}
+CNRS / INRIA / University of Pau (France)\\
+Version 2.1}
 
@@
 Matthias Meschede,
+Peter Messmer,
 David Mich\'ea,
@@
 Daniel Peter,
+Max Rietmann,
 Brian Savage,
@@
 based upon the spectral-element method (SEM).
-The SEM is a continuous Galerkin technique, which can easily be made discontinous \citep{BeMaPa94,Ch00,KoWoHu02,ChCaVi03,LaWaBe05,Kop06,WiStBuGh10,AcKo11}; it is then close to a particular case of the discontinuous Galerkin technique \citep{ReHi73,Arn82,FaRi99,HuHuRa99,CoKaSh00,GiHeWa02,RiWh03,MoRi05,GrScSc06,AiMoMu06,BeLaPi06,DuKa06,DeSeWh08,PuAmKa09,WiStBuGh10,DeSe10,EtChViGl10}, with optimized efficiency because of its tensorized basis functions \citep{WiStBuGh10,AcKo11}.
+The SEM is a continuous Galerkin technique \citep{TrKoLi08,PeKoLuMaLeCaLeMaLiBlNiBaTr11},
+which can easily be made discontinous \citep{BeMaPa94,Ch00,KoWoHu02,ChCaVi03,LaWaBe05,Kop06,WiStBuGh10,AcKo11};
+it is then close to a particular case of the discontinuous Galerkin technique \citep{ReHi73,Arn82,FaRi99,HuHuRa99,CoKaSh00,GiHeWa02,RiWh03,MoRi05,GrScSc06,AiMoMu06,BeLaPi06,DuKa06,DeSeWh08,PuAmKa09,WiStBuGh10,DeSe10,EtChViGl10}, with optimized efficiency because of its tensorized basis functions \citep{WiStBuGh10,AcKo11}.
 In particular, it can accurately handle very distorted mesh elements \citep{OlSe11}.\\
@@
 For a detailed introduction to the SEM as applied to regional
-seismic wave propagation, please consult \citet{KoVi98,KoTr99,ChKoViCaVaFe07,TrKoLi08} and
-in particular \citet{KoLiTrSuStSh04}.
+seismic wave propagation, please consult \citet{KoVi98,KoTr99,ChKoViCaVaFe07,TrKoLi08,PeKoLuMaLeCaLeMaLiBlNiBaTr11} and
+in particular \citet{KoLiTrSuStSh04,LeKoHuTr09,LeChKoHuTr09,LeChLiKoHuTr08,GoAmTaCaSmSaMaKo09,WiKoScTr04}.
 A detailed theoretical analysis of the dispersion
@@
 variations in attenuation \citep{SaKoTr10}. Adjoint capabilities and finite-frequency
-kernel simulations are included \citep{LiTr06,TrKoLi08,FiIgBuKe09,ViOp09}.\\
+kernel simulations are included \citep{TrKoLi08,PeKoLuMaLeCaLeMaLiBlNiBaTr11,LiTr06,FiIgBuKe09,ViOp09}.\\
 
@@
 
-The next release of the code will include support for GPU graphics card acceleration \citep{KoMiEr09,KoErGoMi10,MiKo10,Kom11}
-as well as Convolutional or Auxiliary Differential Equation Perfectly Matched absorbing Layers (C-PML or ADE-PML) \citep{KoMa07,MaKoEz08,MaKoGe08,MaKo09,MaKoGeBr10}.
+This new release of the code (version 2.1) includes support for GPU graphics card acceleration \citep{KoMiEr09,KoGoErMi10,KoErGoMi10,MiKo10,Kom11}.\\
+
+The next release of the code will include support Convolution or Auxiliary Differential Equation Perfectly Matched absorbing Layers
+(C-PML or ADE-PML) \citep{KoMa07,MaKoEz08,MaKoGe08,MaKo09,MaKoGeBr10}.
 It will also use the PT-SCOTCH parallel library for mesh partitioning.\\
@@
 SPECFEM3D includes coupled fluid-solid domains and
-adjoint capabilities, which enables one to address seismological inverse problems, but for linear rheologies so far.
+adjoint capabilities, which enables one to address seismological inverse problems, but for linear rheologies only so far.
 To accommodate visco-plastic or non-linear rheologies, the reader can refer to
@@
 \item [Numerical simulations in general] ~\\
-Forward simulations are described in detail in \cite{TrKoLi08,VaCaSaKoVi99,KoMiEr09,KoErGoMi10,KoGoErMi10,ChKoViCaVaFe07,MaKoDi09,KoViCh10,CaKoLaTiMiLeSnTr08,TrKoHjLiZhPeBoMcFrTrHu10,KoRiTr02,KoTr02a,KoTr02b,KoTr99} or \cite{KoVi98}.
+Forward and adjoint simulations are described in detail in \cite{TrKoLi08,PeKoLuMaLeCaLeMaLiBlNiBaTr11,VaCaSaKoVi99,KoMiEr09,KoErGoMi10,KoGoErMi10,ChKoViCaVaFe07,MaKoDi09,KoViCh10,CaKoLaTiMiLeSnTr08,TrKoHjLiZhPeBoMcFrTrHu10,KoRiTr02,KoTr02a,KoTr02b,KoTr99} or \cite{KoVi98}.
 Additional aspects dealing with adjoint simulations are described in
-\cite{TrTaLi05,LiTr06,TrKoLi08,LiTr08,TrKoHjLiZhPeBoMcFrTrHu10}.
+\cite{TrTaLi05,LiTr06,TrKoLi08,LiTr08,TrKoHjLiZhPeBoMcFrTrHu10,PeKoLuMaLeCaLeMaLiBlNiBaTr11}.
 Domain decomposition is explained in detail in \cite{MaKoBlLe08}, and excellent scaling up to 150,000 processor cores is shown for instance in
-\cite{CaKoLaTiMiLeSnTr08,KoLaMi08a,MaKoBlLe08,KoErGoMi10,KoGoErMi10,Kom11}.
+\cite{CaKoLaTiMiLeSnTr08,KoLaMi08a,MaKoBlLe08,KoErGoMi10,KoGoErMi10,Kom11},
+
+\item [GPU computing] ~\\
+Computing on GPU graphics cards for acoustic or seismic wave propagation applications
+is described in detail in \cite{KoMiEr09,KoGoErMi10,KoErGoMi10,MiKo10,Kom11}.
 
@@
 
-Princeton University, USA,
-and University of Pau / CNRS / INRIA, France
-
-$\copyright$ Princeton University and University of Pau / CNRS / INRIA, April 2011
+Princeton University, USA, and CNRS / INRIA / University of Pau, France
+
+$\copyright$ Princeton University and CNRS / INRIA / University of Pau, July 2012
 
@@
 
- MPI v 2.1 `Sesame' (Spectral ElementS on Any MEsh), July 2012: 
- Max Rietmann, Peter Messmer, Daniel Peter: support for CUDA GPUs. \\
-
- MPI v. 2.0 `Sesame' (Spectral ElementS on Any MEsh), November 2010:
+ MPI v 2.1, July 2012:
+ Max Rietmann, Peter Messmer, Daniel Peter, Dimitri Komatitsch, Joseph Charles, Zhinan Xie: support for CUDA GPUs,
+better CFL stability for the Stacey absorbing conditions. \\
+
+ MPI v. 2.0, November 2010:
  Dimitri Komatitsch, Nicolas Le Goff, Roland Martin and Pieyre Le Loher, University of Pau, France,

### 2012-07-07 doc/USER_MANUAL/manual_SPECFEM3D.tex
added a reference

@@
 If you use this new version 2.1, which has non blocking MPI for much better performance for medium or large runs, please cite at least one of these six articles,
-in which results of non blocking MPI runs are presented: \cite{PeKoLuMaLeCaLeMaLiBlNiBaTr11,KoErGoMi10,KoViCh10,Kom11,CaKoLaTiMiLeSnTr08,MaKoBlLe08}.\\
+in which results of non blocking MPI runs are presented: \cite{PeKoLuMaLeCaLeMaLiBlNiBaTr11,KoGoErMi10,KoErGoMi10,KoViCh10,Kom11,CaKoLaTiMiLeSnTr08,MaKoBlLe08}.\\
 

### 2012-07-07 doc/USER_MANUAL/manual_SPECFEM3D.tex
updated the manual

@@
 For a detailed introduction to the SEM as applied to regional
-seismic wave propagation, please consult \citet{KoVi98,KoTr99,ChKoViCaVaFe07,TrKoLi08,PeKoLuMaLeCaLeMaLiBlNiBaTr11} and
-in particular \citet{KoLiTrSuStSh04,LeKoHuTr09,LeChKoHuTr09,LeChLiKoHuTr08,GoAmTaCaSmSaMaKo09,WiKoScTr04}.
+seismic wave propagation, please consult \citet{PeKoLuMaLeCaLeMaLiBlNiBaTr11,TrKoLi08,KoVi98,KoTr99,ChKoViCaVaFe07} and
+in particular \citet{LeKoHuTr09,LeChKoHuTr09,LeChLiKoHuTr08,GoAmTaCaSmSaMaKo09,WiKoScTr04,KoLiTrSuStSh04}.
 A detailed theoretical analysis of the dispersion

### 2012-07-07 doc/USER_MANUAL/manual_SPECFEM3D.tex
removed an unneeded reference

@@
 The spectral element approach admits spectral rates of convergence and allows exploiting $hp$-convergence schemes.
-It is also very well suited to parallel implementation on very large supercomputers \citep{KoTr02a,KoTsChTr03,TsKoChTr03,KoLaMi08a,CaKoLaTiMiLeSnTr08,KoViCh10} as well as on clusters of GPU accelerating graphics cards \citep{KoMiEr09,KoErGoMi10,Kom11}.
+It is also very well suited to parallel implementation on very large supercomputers \citep{KoTsChTr03,TsKoChTr03,KoLaMi08a,CaKoLaTiMiLeSnTr08,KoViCh10} as well as on clusters of GPU accelerating graphics cards \citep{KoMiEr09,KoErGoMi10,Kom11}.
 Tensor products inside each element can be optimized to reach very high efficiency \citep{DeFiMu02}, and mesh point and element numbering can be optimized to reduce processor cache misses and improve cache reuse \citep{KoLaMi08a}.

### 2012-08-20 doc/USER_MANUAL/manual_SPECFEM3D.tex
improved the sentence about mcmodel=medium

@@
 In this case, on some compilers you may need to add \texttt{``-mcmodel=medium}'' or \texttt{``-mcmodel=medium -shared-intel}''
-to the compiler options of CFLAGS, FCFLAGS and LDFLAGS otherwise the compiler will display an error
+to the configure options of CFLAGS, FCFLAGS and LDFLAGS otherwise the compiler will display an error
 message (for instance \texttt{``relocation truncated to fit: R\_X86\_64\_PC32 against .bss''} or something similar);
@@
 
-\caption{Schematic workflow for a SPECFEM3D simulation. The exectable \texttt{xgenerate\_databases} creates the GLL mesh points and assigns specific model parameters.}
+\caption{Schematic workflow for a SPECFEM3D simulation. The executable \texttt{xgenerate\_databases} creates the GLL mesh points and assigns specific model parameters.}
 

### 2012-09-19 src/specfem3D/compute_add_sources_poroelastic.f90
adds point source flags in Par_file related to the use of an inclined force

@@ -32,7 +32,7 @@
                         rhoarraystore,phistore,tortstore,&
                         ibool,ispec_is_inner,phase_is_inner, &
                         NSOURCES,myrank,it,islice_selected_source,ispec_selected_source,&
-                        xi_source,eta_source,gamma_source,nu_source, &
+                        xi_source,eta_source,gamma_source, &
                         hdur,hdur_gaussian,tshift_cmt,dt,t0,sourcearrays, &
                         ispec_is_poroelastic,SIMULATION_TYPE,NSTEP,NGLOB_ADJOINT, &
                         nrec,islice_selected_rec,ispec_selected_rec, &
@@ -42,7 +42,8 @@
   use specfem_par,only: PRINT_SOURCE_TIME_FUNCTION,stf_used_total, &
                         xigll,yigll,zigll,xi_receiver,eta_receiver,gamma_receiver,&
                         station_name,network_name,adj_source_file, &
-                        USE_FORCE_POINT_SOURCE,FACTOR_FORCE_SOURCE,COMPONENT_FORCE_SOURCE
+                        USE_FORCE_POINT_SOURCE,FACTOR_FORCE_SOURCE
+
   implicit none
 
   include "constants.h"
@@ -66,7 +67,6 @@
   integer :: NSOURCES,myrank,it
   integer, dimension(NSOURCES) :: islice_selected_source,ispec_selected_source
   double precision, dimension(NSOURCES) :: xi_source,eta_source,gamma_source
-  double precision, dimension(3,3,NSOURCES) :: nu_source
   double precision, dimension(NSOURCES) :: hdur,hdur_gaussian,tshift_cmt
   double precision :: dt,t0
   real(kind=CUSTOM_REAL), dimension(NSOURCES,NDIM,NGLLX,NGLLY,NGLLZ) :: sourcearrays
@@ -154,19 +154,32 @@
               endif
 
               ! This is the expression of a Ricker; should be changed according maybe to the Par_file.
-              stf_used = FACTOR_FORCE_SOURCE * comp_source_time_function_rickr(dble(it-1)*DT-t0-tshift_cmt(isource),f0)
+              stf = comp_source_time_function_rickr(dble(it-1)*DT-t0-tshift_cmt(isource),f0)
               !stf_used = comp_source_time_function_gauss(dble(it-1)*DT-t0-tshift_cmt(isource),f0)
 
-              ! we use a force in a single direction along one of the components:
[…]

### 2012-09-19 doc/USER_MANUAL/manual_SPECFEM3D.tex
adds point source flags in Par_file related to the use of an inclined force

@@
 to avoid writing to the seismograms too often. This feature is only relevant for the solver.
+\item [{\texttt{USE\_FORCE\_POINT\_SOURCE}}] Turn this flag on to use a force source
+instead of a \texttt{CMTSOLUTION} moment-tensor source. This can be useful e.g. for oil industry 
+foothills simulations in which the source is a vertical force, normal force, 
+tilted force, or an impact etc. You will need to give the East, North and vertical components of the force vector;
+thus refer to Appendix~\ref{cha:Coordinates} for the orientation of the reference frame.
+\item [{\texttt{FACTOR\_FORCE\_SOURCE}}] This parameter specifies the magnitude of the force source.
+\item [{\texttt{COMPONENT\_DIR\_VECT\_SOURCE\_E}}] This parameter specifies the East component of 
+an arbitrary (non-unitary) direction vector of the force source.
+\item [{\texttt{COMPONENT\_DIR\_VECT\_SOURCE\_N}}] This parameter specifies the North component of 
+an arbitrary (non-unitary) direction vector of the force source.
+\item [{\texttt{COMPONENT\_DIR\_VECT\_SOURCE\_ZUP}}] This parameter specifies the vertical component of 
+an arbitrary (non-unitary) direction vector of the force source.
 \item [{\texttt{PRINT\_SOURCE\_TIME\_FUNCTION}}] Turn this flag on to print
@@
 \noindent \begin{centering}
-\includegraphics[width=3in]{figures/gauss_vs_triangle_mod.eps}
+%%%\includegraphics[width=3in]{figures/gauss_vs_triangle_mod.eps}
+\includegraphics[width=3in]{figures/gauss_vs_triangle_mod.pdf}
 \par\end{centering}
@@
 \noindent \begin{centering}
-{\small \includegraphics[width=5in]{figures/source_timing.eps} }
+{\small \includegraphics[width=5in]{figures/source_timing.pdf} }
 \par\end{centering}{\small \par}
@@
 \noindent \begin{centering}
-\includegraphics[scale=0.6]{figures/IRIS_band_codes.eps}
+\includegraphics[scale=0.6]{figures/IRIS_band_codes.pdf}
 \par\end{centering}

### 2012-09-19 src/specfem3D/compute_forces_elastic.F90

@@ -235,7 +235,7 @@ subroutine compute_forces_elastic()
     call compute_add_sources_elastic( NSPEC_AB,NGLOB_AB,accel, &
                         ibool,ispec_is_inner,phase_is_inner, &
                         NSOURCES,myrank,it,islice_selected_source,ispec_selected_source,&
-                        xi_source,eta_source,gamma_source,nu_source, &
+                        xi_source,eta_source,gamma_source, &
                         hdur,hdur_gaussian,tshift_cmt,dt,t0,sourcearrays, &
                         ispec_is_elastic,SIMULATION_TYPE,NSTEP,NGLOB_ADJOINT, &
                         nrec,islice_selected_rec,ispec_selected_rec, &

### 2012-09-19 doc/USER_MANUAL/manual_SPECFEM3D.tex
updates the user manual for the definition of an inclined force vector

@@
 foothills simulations in which the source is a vertical force, normal force, 
-tilted force, or an impact etc. You will need to give the East, North and vertical components of the force vector;
-thus refer to Appendix~\ref{cha:Coordinates} for the orientation of the reference frame.
+tilted force, or an impact etc. You will need to give the East, North and vertical components of an 
+arbitrary (non-unitary) direction vector of the force vector; thus refer to Appendix~\ref{cha:Coordinates} 
+for the orientation of the reference frame.
 \item [{\texttt{FACTOR\_FORCE\_SOURCE}}] This parameter specifies the magnitude of the force source.

### 2012-09-21 src/cuda/specfem3D_gpu_cuda_method_stubs.c
moves ABSORB_FREE_SURFACE flag from constants.h.in into ABSORB_INSTEAD_OF_FREE_SURFACE flag in Par_file

@@ -212,7 +212,7 @@ void FC_FUNC_(compute_forces_acoustic_cuda,
 
 void FC_FUNC_(acoustic_enforce_free_surf_cuda,
               ACOUSTIC_ENFORCE_FREE_SURF_CUDA)(long* Mesh_pointer_f,
-                                               int* ABSORB_FREE_SURFACE) {}
+                                               int* ABSORB_INSTEAD_OF_FREE_SURFACE) {}
 
 
 //

### 2012-09-21 src/generate_databases/generate_databases_par.f90

@@ -64,7 +64,8 @@
 
   logical :: ATTENUATION,USE_OLSEN_ATTENUATION, &
           OCEANS,TOPOGRAPHY,SAVE_FORWARD,USE_FORCE_POINT_SOURCE
-  logical :: ANISOTROPY,ABSORBING_CONDITIONS,SAVE_MESH_FILES,PRINT_SOURCE_TIME_FUNCTION
+  logical :: ANISOTROPY,ABSORBING_CONDITIONS,SAVE_MESH_FILES,PRINT_SOURCE_TIME_FUNCTION, &
+          ABSORB_INSTEAD_OF_FREE_SURFACE
 
   logical :: MOVIE_SURFACE,MOVIE_VOLUME,CREATE_SHAKEMAP,SAVE_DISPLACEMENT, &
           USE_HIGHRES_FOR_MOVIES

### 2012-09-21 doc/USER_MANUAL/manual_SPECFEM3D.tex
made the description of the force source more precise
moves ABSORB_FREE_SURFACE flag from constants.h.in into ABSORB_INSTEAD_OF_FREE_SURFACE flag in Par_file

@@
 on Clayton-Enquist absorbing boundary conditions (see \citet{KoTr99}).
+\item [{\texttt{ABSORB\_INSTEAD\_OF\_FREE\_SURFACE}}] Set to \texttt{.true.} to turn on 
+absorbing boundary conditions on the top surface which by default constitutes a free surface of the model.
 \item [{\texttt{MOVIE\_SURFACE}}] Set to \texttt{.false.}, unless you want
@@
 for the orientation of the reference frame.
+This vector is made unitary internally in the code and thus only its direction matters here;
+its norm is ignored and the norm of the force used is \texttt{FACTOR\_FORCE\_SOURCE} times the source time function.
 \item [{\texttt{FACTOR\_FORCE\_SOURCE}}] This parameter specifies the magnitude of the force source.
@@
 an arbitrary (non-unitary) direction vector of the force source.
-\item [{\texttt{COMPONENT\_DIR\_VECT\_SOURCE\_ZUP}}] This parameter specifies the vertical component of 
+\item [{\texttt{COMPONENT\_DIR\_VECT\_SOURCE\_Z\_UP}}] This parameter specifies the vertical component of 
 an arbitrary (non-unitary) direction vector of the force source.

### 2012-09-25 src/generate_databases/finalize_databases.f90
adds a check on the use of the ABSORB_INSTEAD_OF_FREE_SURFACE flag

@@ -101,7 +101,7 @@
 ! copy number of elements and points in an include file for the solver
   if( myrank == 0 ) then
     call save_header_file(NSPEC_AB,NGLOB_AB,NPROC, &
-               ATTENUATION,ANISOTROPY,NSTEP,DT, &
+               ATTENUATION,ANISOTROPY,NSTEP,DT,ABSORB_INSTEAD_OF_FREE_SURFACE, &
                SIMULATION_TYPE,max_memory_size,nfaces_surface_glob_ext_mesh)
   endif
 

### 2012-09-28 doc/USER_MANUAL/manual_SPECFEM3D.tex
adds a check on the use of the ABSORB_INSTEAD_OF_FREE_SURFACE flag
adds a set of FORCESOLUTION files to set force source parameters, used instead of CMTSOLUTION files if "USE_FORCE_POINT_SOURCE = .true".

@@
 to avoid writing to the seismograms too often. This feature is only relevant for the solver.
-\item [{\texttt{USE\_FORCE\_POINT\_SOURCE}}] Turn this flag on to use a force source
+\item [{\texttt{USE\_FORCE\_POINT\_SOURCE}}] Turn this flag on to use a \texttt{FORCESOLUTION} force source
 instead of a \texttt{CMTSOLUTION} moment-tensor source. This can be useful e.g. for oil industry 
@@
 This vector is made unitary internally in the code and thus only its direction matters here;
-its norm is ignored and the norm of the force used is \texttt{FACTOR\_FORCE\_SOURCE} times the source time function.
+its norm is ignored and the norm of the force used is the factor force source times the source time function.
 \item [{\texttt{FACTOR\_FORCE\_SOURCE}}] This parameter specifies the magnitude of the force source.
@@
 Chapter~\ref{cha:Creating-Distributed-Databases},
-\item [{\texttt{CMTSOLUTION}}] the earthquake source parameter file, and
+\item [{\texttt{CMTSOLUTION}} or {\texttt{FORCESOLUTION}}] the earthquake source parameter file or the force source parameter file or, and
 \item [{\texttt{STATIONS}}] the stations file.
@@
 \item the time step parameters \texttt{NSTEP} and \texttt{DT}
-\item the absorbing boundary control parameter \texttt{ABSORBING\_CONDITIONS}
+\item the absorbing boundary control parameter \texttt{ABSORBING\_CONDITIONS} on condition that the\\ 
+  \texttt{ABSORB\_INSTEAD\_OF\_FREE\_SURFACE} flag remains unmodified after running the databases generation. 
 \item the movie control parameters \texttt{MOVIE\_SURFACE}, \texttt{MOVIE\_VOLUME},
@@
 
+\vspace{1cm}
+
+\noindent The \texttt{FORCESOLUTION} should be edited in the following way:
+
+\begin{itemize}
+\item Set the \texttt{time shift} parameter equal to $0.0$ (the solver
+will not run otherwise.) The time shift parameter would simply apply
+an overall time shift to the synthetics, something that can be done
+in the post-processing (see Section \ref{sec:Process-data-and-syn}).
+\item Set the latitude, longitude, depth of the source in geographical
+  coordinates.
+\item Set the magnitude of the force source.
+\item Set the components of a (non-unitary) direction vector for the
+  force source in the East/North/Vertical basis (see Appendix A for
+  the orientation of the reference frame).
+\end{itemize}
+
 {\small }%
@@
 
-Having set the \texttt{Par\_file} parameters, and having provided
-the \texttt{CMTSOLUTION} and \texttt{STATIONS} files, you are now
-ready to launch the solver! This is most easily accomplished based
-upon the \texttt{go\_solver} script (See Chapter \ref{cha:Scheduler}
-for information about running through a scheduler, e.g., LSF). You
-may need to edit the last command at the end of the script that invokes
-the \texttt{mpirun} command. The \texttt{runall} script compiles and
-runs both \texttt{xgenerate\_databases} and \texttt{xspecfem3D} in sequence. This is a safe approach that
-ensures using the correct combination of distributed database output and solver
-input.
+Having set the \texttt{Par\_file} parameters, and having provided the
+\texttt{CMTSOLUTION} (or the \texttt{FORCESOLUTION}) and
+\texttt{STATIONS} files, you are now ready to launch the solver! This
+is most easily accomplished based upon the \texttt{go\_solver} script
+(See Chapter \ref{cha:Scheduler} for information about running through
+a scheduler, e.g., LSF). You may need to edit the last command at the
+end of the script that invokes the \texttt{mpirun} command. The
+\texttt{runall} script compiles and runs both
+\texttt{xgenerate\_databases} and \texttt{xspecfem3D} in
+sequence. This is a safe approach that ensures using the correct
+combination of distributed database output and solver input.
 
@@
 
-For the same model, one can rerun the solver for different events
-by simply changing the \texttt{CMTSOLUTION} file, or for different
-stations by changing the \texttt{STATIONS} file. There is no need
-to rerun the \texttt{xgenerate\_databases} executable. Of course it is best to include as many stations
-as possible, since this does not add to the cost of the simulation.
+For the same model, one can rerun the solver for different events by
+simply changing the \texttt{CMTSOLUTION} or \texttt{FORCESOLUTION}
+file, or for different stations by changing the \texttt{STATIONS}
+file. There is no need to rerun the \texttt{xgenerate\_databases}
+executable. Of course it is best to include as many stations as
+possible, since this does not add to the cost of the simulation.
 
@@
 The SPECFEM3D software code internally uses Cartesian coordinates.
-The given locations for sources and receiver locations thus may get converted.
-Sources and receiver locations are read in from the
-\texttt{CMTSOLUTION} and \texttt{STATIONS} file.
-Note that e.g. the \texttt{CMTSOLUTION} file denotes the location by "longitude/latitude/depth". We read in
-longitude as $x$ coordinate, latitude as $y$ coordinate.
+The given locations for sources and receiver locations thus may get
+converted.  Sources and receiver locations are read in from the
+\texttt{CMTSOLUTION} (or \texttt{FORCESOLUTION}) and \texttt{STATIONS}
+files.  Note that e.g. the \texttt{CMTSOLUTION} and
+\texttt{FORCESOLUTION} files denotes the location by
+"longitude/latitude/depth". We read in longitude as $x$ coordinate,
+latitude as $y$ coordinate.
 
@@
 
-The value for depth (given in $km$ in \texttt{CMTSOLUTION}) or burial depth (given in $m$ in \texttt{STATIONS})
-is evaluated with respect to the surface of the mesh at the specified $x$/$y$ location to find the corresponding $z$ coordinate.
-It is possible to use this depth value directly as $z$ coordinate by
-changing the flag \texttt{USE\_SOURCES\_RECVS\_Z} to \texttt{.true.} in the file \texttt{constants.h} located in the \texttt{src/shared/} subdirectory.
+The value for depth (given in $km$ in \texttt{CMTSOLUTION} and
+\texttt{FORCESOLUTION}) or burial depth (given in $m$ in
+\texttt{STATIONS}) is evaluated with respect to the surface of the
+mesh at the specified $x$/$y$ location to find the corresponding $z$
+coordinate.  It is possible to use this depth value directly as $z$
+coordinate by changing the flag \texttt{USE\_SOURCES\_RECVS\_Z} to
+\texttt{.true.} in the file \texttt{constants.h} located in the
+\texttt{src/shared/} subdirectory.
 

### 2012-09-28 src/specfem3D/compute_add_sources_acoustic.f90

@@ -42,7 +42,7 @@
                         xigll,yigll,zigll,xi_receiver,eta_receiver,gamma_receiver,&
                         station_name,network_name,adj_source_file,nrec_local,number_receiver_global, &
                         pm1_source_encoding,nsources_local,USE_FORCE_POINT_SOURCE, &
-                        FACTOR_FORCE_SOURCE
+                        factor_force_source
   implicit none
 
   include "constants.h"
@@ -123,7 +123,7 @@
         do isource = 1,NSOURCES
           if(USE_FORCE_POINT_SOURCE) then
             ! precomputes source time function factor
-            stf_pre_compute(isource) = FACTOR_FORCE_SOURCE * comp_source_time_function_rickr( &
+            stf_pre_compute(isource) = factor_force_source(isource) * comp_source_time_function_rickr( &
                   dble(it-1)*DT-t0-tshift_cmt(isource),hdur(isource))
           else
             if( USE_RICKER_IPATI ) then
@@ -175,7 +175,7 @@
 
                 ! we use nu_source(:,3) here because we want a source normal to the surface.
                 ! This is the expression of a Ricker; should be changed according maybe to the Par_file.
-                stf_used = FACTOR_FORCE_SOURCE * &
+                stf_used = factor_force_source(isource) * &
                            comp_source_time_function_rickr(dble(it-1)*DT-t0-tshift_cmt(isource),f0)
 
 
@@ -419,7 +419,7 @@
         do isource = 1,NSOURCES
           if(USE_FORCE_POINT_SOURCE) then
             ! precomputes source time function factors
-            stf_pre_compute(isource) = FACTOR_FORCE_SOURCE * comp_source_time_function_rickr( &
+            stf_pre_compute(isource) = factor_force_source(isource) * comp_source_time_function_rickr( &
                   dble(NSTEP-it)*DT-t0-tshift_cmt(isource),hdur(isource))
           else
             if( USE_RICKER_IPATI ) then
@@ -472,7 +472,7 @@
 
                 ! we use nu_source(:,3) here because we want a source normal to the surface.
                 ! This is the expression of a Ricker; should be changed according maybe to the Par_file.
-                stf_used = FACTOR_FORCE_SOURCE * comp_source_time_function_rickr( &
+                stf_used = factor_force_source(isource) * comp_source_time_function_rickr( &
                                         dble(NSTEP-it)*DT-t0-tshift_cmt(isource),f0)
 
                 ! source encoding

### 2012-09-28 doc/USER_MANUAL/manual_SPECFEM3D.tex
updates the user guide

@@
 its norm is ignored and the norm of the force used is the factor force source times the source time function.
-\item [{\texttt{FACTOR\_FORCE\_SOURCE}}] This parameter specifies the magnitude of the force source.
-\item [{\texttt{COMPONENT\_DIR\_VECT\_SOURCE\_E}}] This parameter specifies the East component of 
-an arbitrary (non-unitary) direction vector of the force source.
-\item [{\texttt{COMPONENT\_DIR\_VECT\_SOURCE\_N}}] This parameter specifies the North component of 
-an arbitrary (non-unitary) direction vector of the force source.
-\item [{\texttt{COMPONENT\_DIR\_VECT\_SOURCE\_Z\_UP}}] This parameter specifies the vertical component of 
-an arbitrary (non-unitary) direction vector of the force source.
 \item [{\texttt{PRINT\_SOURCE\_TIME\_FUNCTION}}] Turn this flag on to print

### 2012-10-01 doc/USER_MANUAL/manual_SPECFEM3D.tex
small update of the user manual

@@
 %%%\includegraphics[width=3in]{figures/gauss_vs_triangle_mod.eps}
-\includegraphics[width=3in]{figures/gauss_vs_triangle_mod.pdf}
+\includegraphics[width=2.5in]{figures/gauss_vs_triangle_mod.pdf}
 \par\end{centering}
@@
 \end{itemize}
+
+\noindent Where  necessary, set  a \texttt{FORCESOLUTION} file  in the
+same   way   you    configure   a   \texttt{CMTSOLUTION}   file   with
+$N_{\mathrm{sources}}$   entries,  one   for   each  subevent   (i.e.,
+concatenate  $N_{\mathrm{sources}}$ \texttt{FORCESOLUTION} files  to a
+single   \texttt{FORCESOLUTION}  file).  At   least  one   entry  (not
+necessarily the first) must have a  zero time shift, and all the other
+entries must have non-negative time  shift. Each subevent can have its
+own half latitude, longitude, depth, and force parameters.
 

### 2012-10-17 doc/USER_MANUAL/manual_SPECFEM3D.tex
adds the hdur parameter in FORCESOLUTION files and the use of a quasi-Heaviside source time function for runs involving a FORCESOLUTION file

@@
 foothills simulations in which the source is a vertical force, normal force, 
-tilted force, or an impact etc. You will need to give the East, North and vertical components of an 
+tilted force, or an impact etc. By default a quasi-Heaviside time function is used to represent the force point source but a Ricker 
+source time function can be computed instead by changing the value of \texttt{USE\_RICKER\_IPATI} in the main constants file 
+\texttt{constants.h} located in the \texttt{src/shared/} subdirectory.
+Note that in the \texttt{FORCESOLUTION} file, you will need to edit the East, North and vertical components of an 
 arbitrary (non-unitary) direction vector of the force vector; thus refer to Appendix~\ref{cha:Coordinates} 
-for the orientation of the reference frame.
-This vector is made unitary internally in the code and thus only its direction matters here;
+for the orientation of the reference frame. This vector is made unitary internally in the code and thus only its direction matters here;
 its norm is ignored and the norm of the force used is the factor force source times the source time function.
@@
 Chapter~\ref{cha:Creating-Distributed-Databases},
-\item [{\texttt{CMTSOLUTION}} or {\texttt{FORCESOLUTION}}] the earthquake source parameter file or the force source parameter file or, and
+\item [{\texttt{CMTSOLUTION}} or {\texttt{FORCESOLUTION}}] the earthquake source parameter file or the force source parameter file, and
 \item [{\texttt{STATIONS}}] the stations file.
@@
 \end{lyxcode}
-The \texttt{CMTSOLUTION} should be edited in the following way:
+The \texttt{CMTSOLUTION} file should be edited in the following way:
 
@@
 
-\noindent The \texttt{FORCESOLUTION} should be edited in the following way:
+\noindent The \texttt{FORCESOLUTION} file should be edited in the following way:
 
@@
 in the post-processing (see Section \ref{sec:Process-data-and-syn}).
+\item Set the \texttt{half duration} parameter of the source time function.
 \item Set the latitude, longitude, depth of the source in geographical
@@
 single   \texttt{FORCESOLUTION}  file).  At   least  one   entry  (not
-necessarily the first) must have a  zero time shift, and all the other
-entries must have non-negative time  shift. Each subevent can have its
-own half latitude, longitude, depth, and force parameters.
+necessarily the first) must have a  zero \texttt{time shift}, and all the other
+entries must have non-negative \texttt{time  shift}. Each subevent can have its
+own half latitude, longitude, depth, \texttt{half duration} and force parameters.
 

### 2012-10-17 src/specfem3D/setup_sources_receivers.f90

@@ -74,7 +74,7 @@
   use specfem_par_movie
   implicit none
 
-  double precision :: t0_acoustic,min_tshift_cmt_original
+  double precision :: t0_acoustic,min_tshift_src_original
   integer :: yr,jda,ho,mi
   integer :: isource,ispec,ier
 
@@ -90,7 +90,7 @@
           xi_source(NSOURCES), &
           eta_source(NSOURCES), &
           gamma_source(NSOURCES), &
-          tshift_cmt(NSOURCES), &
+          tshift_src(NSOURCES), &
           hdur(NSOURCES), &
           hdur_gaussian(NSOURCES), &
           utm_x_source(NSOURCES), &
@@ -104,6 +104,12 @@
           comp_dir_vect_source_N(NSOURCES), &
           comp_dir_vect_source_Z_UP(NSOURCES),stat=ier)
      if( ier /= 0 ) stop 'error allocating arrays for force point sources'
+  else
+     allocate(factor_force_source(1), &
+          comp_dir_vect_source_E(1), &
+          comp_dir_vect_source_N(1), &
+          comp_dir_vect_source_Z_UP(1),stat=ier)
+     if( ier /= 0 ) stop 'error allocating dummy FORCESOLUTION arrays'
   endif
 
   ! for source encoding (acoustic sources so far only)
@@ -116,7 +122,7 @@
 !                xi_source, eta_source & gamma_source
   call locate_source(ibool,NSOURCES,myrank,NSPEC_AB,NGLOB_AB, &
           xstore,ystore,zstore,xigll,yigll,zigll,NPROC, &
-          tshift_cmt,min_tshift_cmt_original,yr,jda,ho,mi,utm_x_source,utm_y_source, &
+          tshift_src,min_tshift_src_original,yr,jda,ho,mi,utm_x_source,utm_y_source, &
           DT,hdur,Mxx,Myy,Mzz,Mxy,Mxz,Myz, &
           islice_selected_source,ispec_selected_source, &
           xi_source,eta_source,gamma_source, &
@@ -128,7 +134,7 @@
           USE_FORCE_POINT_SOURCE,factor_force_source,comp_dir_vect_source_E, &
           comp_dir_vect_source_N,comp_dir_vect_source_Z_UP)
 
-  if(abs(minval(tshift_cmt)) > TINYVAL) call exit_MPI(myrank,'one tshift_cmt must be zero, others must be positive')
[…]

### 2012-10-22 doc/USER_MANUAL/manual_SPECFEM3D_Cartesian.tex
added "Cartesian" to the name of the package and to the manual

--- 2012-10-17 doc/USER_MANUAL/manual_SPECFEM3D.tex
@@
 \noindent\makebox[\textwidth]{%
-\noindent\includegraphics[width=0.83\paperwidth]{figures/specfem_3d-cover.pdf}
+\noindent\includegraphics[width=0.83\paperwidth]{figures/specfem_3d_Cartesian-cover.pdf}
 }
@@
 %
-\title{\textbf{SPECFEM3D}\\
+\title{\textbf{SPECFEM3D Cartesian}\\
 \textbf{User Manual}}
@@
 \section*{Authors}
-The SPECFEM3D package was first developed by Dimitri Komatitsch and Jeroen Tromp at Harvard University and Caltech, USA, starting in 1998, based on earlier work by Dimitri Komatitsch and Jean-Pierre Vilotte at Institut de Physique du Globe (IPGP) in Paris, France from 1995 to 1997.\\
+The SPECFEM3D Cartesian package was first developed by Dimitri Komatitsch and Jeroen Tromp at Harvard University and Caltech, USA, starting in 1998, based on earlier work by Dimitri Komatitsch and Jean-Pierre Vilotte at Institut de Physique du Globe (IPGP) in Paris, France from 1995 to 1997.\\
 
@@
 
-The software package SPECFEM3D simulates seismic
+The software package SPECFEM3D Cartesian simulates seismic
 wave propagation at the local or regional scale
@@
 
-All SPECFEM3D software is written in Fortran90 with full portability
+All SPECFEM3D Cartesian software is written in Fortran90 with full portability
 in mind, and conforms strictly to the Fortran95 standard. It uses
@@
 
-SPECFEM3D includes coupled fluid-solid domains and
+SPECFEM3D Cartesian includes coupled fluid-solid domains and
 adjoint capabilities, which enables one to address seismological inverse problems, but for linear rheologies only so far.
@@
 
-If you use SPECFEM3D for your own research, please cite at least one
+If you use SPECFEM3D Cartesian for your own research, please cite at least one
 of the following articles:
@@
 
-The SPECFEM3D software package comes in a gzipped tar ball. In the
+The SPECFEM3D Cartesian software package comes in a gzipped tar ball. In the
 directory in which you want to install the package, type
@@
 
-The SPECFEM3D software package relies on the SCOTCH library to partition meshes created with CUBIT.
+The SPECFEM3D Cartesian software package relies on the SCOTCH library to partition meshes created with CUBIT.
 Note that we use CUBIT to create meshes of hexahedra, but other packages can be used as well, for instance GiD from \url{http://gid.cimne.upc.es} or Gmsh from \url{http://geuz.org/gmsh} \citep{GeRe09}.
@@
 \item [1.] setting up a hexahedral mesh with CUBIT,
-\item [2.] exporting the CUBIT mesh into a SPECFEM3D file format and
-\item [3.] partitioning the SPECFEM3D mesh files for a chosen number of cores.
+\item [2.] exporting the CUBIT mesh into a SPECFEM3D Cartesian file format and
+\item [3.] partitioning the SPECFEM3D Cartesian mesh files for a chosen number of cores.
 \end{description}
 
-Examples are provided in the SPECFEM3D package in the subdirectory \texttt{examples/}.
+Examples are provided in the SPECFEM3D Cartesian package in the subdirectory \texttt{examples/}.
 We strongly encourage you to contribute your own example to this package by contacting the CIG Computational Seismology Mailing List  \urlwithparentheses{cig-seismo@geodynamics.org}.
@@
 For the installation and handling of the CUBIT meshing tool suite, please refer to the CUBIT user manual and documentation.
-In order to give you a basic understanding of how to use CUBIT for our purposes, examples are provided in the SPECFEM3D package in the subdirectory \texttt{examples/}:
+In order to give you a basic understanding of how to use CUBIT for our purposes, examples are provided in the SPECFEM3D Cartesian package in the subdirectory \texttt{examples/}:
 \begin{description}
@@
 \item [mesh\_file] Contains the mesh element connectivity.
-\item [free\_surface\_file] Contains the free surface connectivity.
+\item [free\_surface\_file] Contains the free surface connectivity. You should put both the surface of acoustic regions and of elastic regions in that file;
+that is, list all the element faces that constitute the surface of the model in that file.
 \item [absorbing\_surface\_file\_xmax] Contains the surface connectivity of the absorbing boundary surface at the Xmax.
@@
 
-The SPECFEM3D software package performs large scale simulations in a parallel 'Single Process Multiple Data' way.
+The SPECFEM3D Cartesian software package performs large scale simulations in a parallel 'Single Process Multiple Data' way.
 The spectral-element mesh created with CUBIT needs to be distributed on the processors. This partitioning is executed once and for all prior to
@@
 \texttt{src/decompose\_mesh\_SCOTCH/scotch\_5.1.10b}.
-This version gets bundled with the compilation of the SPECFEM3D package if no libraries could have been found.
+This version gets bundled with the compilation of the SPECFEM3D Cartesian package if no libraries could have been found.
 If this automatic compilation of the SCOTCH libraries fails,
@@
 These files will be needed for creating the distributed databases, and have to reside in the directory  \texttt{LOCAL\_PATH} specified in the main \texttt{Par\_file}, e.g.
-in directory \texttt{in\_out\_files/DATABASES\_MPI}.
+in directory \texttt{OUTPUT\_FILES/DATABASES\_MPI}.
 Please see Chapter~\ref{cha:Creating-Distributed-Databases} for further details.
@@
 Input for the mesh generation program is provided through the parameter
-file \texttt{Mesh\_Par\_file}, which resides in the subdirectory \texttt{in\_data\_files/meshfem3D\_files/}.
+file \texttt{Mesh\_Par\_file}, which resides in the subdirectory \texttt{DATA/meshfem3D\_files/}.
 Before running the mesher, a number of parameters need to be set in
@@
 
-Mesher output is provided in the \texttt{in\_out\_files/OUTPUT\_FILES} directory
+Mesher output is provided in the \texttt{OUTPUT\_FILES} directory
 in \texttt{output\_mesher.txt}; this file provides lots of details
@@
 
-To control the quality of the mesh, check the standard output (either on the screen or in the \texttt{in\_out\_files/OUTPUT\_FILES} directory in \texttt{output\_mesher.txt}) and analyze the skewness histogram of mesh elements to make sure no element has a skewness above approximately 0.75, otherwise the mesh is of poor quality (and if even a single element has a skewness value above 0.80, then you must definitely improve the mesh).
+To control the quality of the mesh, check the standard output (either on the screen or in the \texttt{OUTPUT\_FILES} directory in \texttt{output\_mesher.txt}) and analyze the skewness histogram of mesh elements to make sure no element has a skewness above approximately 0.75, otherwise the mesh is of poor quality (and if even a single element has a skewness value above 0.80, then you must definitely improve the mesh).
 To draw the skewness histogram on the screen, type \texttt{gnuplot plot\_mesh\_quality\_histogram.gnu}.
@@
 
-\caption{Schematic workflow for a SPECFEM3D simulation. The executable \texttt{xgenerate\_databases} creates the GLL mesh points and assigns specific model parameters.}
+\caption{Schematic workflow for a SPECFEM3D Cartesian simulation. The executable \texttt{xgenerate\_databases} creates the GLL mesh points and assigns specific model parameters.}
 
@@
 Input for the program is provided through
-the main parameter file \texttt{Par\_file}, which resides in the subdirectory \texttt{in\_data\_files/}. Please note that
+the main parameter file \texttt{Par\_file}, which resides in the subdirectory \texttt{DATA}. Please note that
 \texttt{xgenerate\_databases} must be called directly from the \texttt{bin/} directory, as most of the binaries of the package.
@@
 Before running \texttt{xgenerate\_databases}, a number of parameters need to be set in the main parameter \texttt{Par\_file}
-located in the subdirectory \texttt{in\_data\_files/}:
+located in the subdirectory \texttt{DATA}:
 
@@
 Users must provide the corresponding data file \texttt{regrid3\_vel\_p.bin} in directory \texttt{DATA/st\_3D\_block\_harvard/}.
-\item [{\texttt{tomo}}] For a user-specified 3D isotropic model which uses a tomographic model file \texttt{tomographic\_model.xyz} in directory \texttt{in\_data\_files/}.
-See Section \ref{sec:Using-tomographic},
+\item [{\texttt{tomo}}] For a user-specified 3D isotropic model which uses a tomographic model file \texttt{tomographic\_model.xyz}
+in directory \texttt{DATA}. See Section \ref{sec:Using-tomographic},
 for a discussion on how to specify your own 3D tomographic model.
@@
 the interval at which basic information about a run is written to
-the file system (\texttt{timestamp{*}} files in the \texttt{in\_out\_files/OUTPUT\_FILES}
+the file system (\texttt{timestamp{*}} files in the \texttt{OUTPUT\_FILES}
 directory). If you have access to a fast machine, set \texttt{NTSTEP\_BETWEEN\_OUTPUT\_INFO}
@@
 \item [{\texttt{PRINT\_SOURCE\_TIME\_FUNCTION}}] Turn this flag on to print
-information about the source time function in the file \texttt{in\_out\_files/OUTPUT\_FILES/plot\_source\_time\_function.txt}.
+information about the source time function in the file \texttt{OUTPUT\_FILES/plot\_source\_time\_function.txt}.
 This feature is only relevant for the solver.
@@
 to evaluate the value of $\Delta t$ for empirically chosen Courant numbers $C \sim 0.3$.
-If you used the mesher \texttt{xmeshfem3D} to generate your mesh, you should set the value suggested in \texttt{in\_out\_files/OUTPUT\_FILES/output\_mesher.txt} file, which is created after the mesher completed.
+If you used the mesher \texttt{xmeshfem3D} to generate your mesh, you should set the value suggested in \texttt{OUTPUT\_FILES/output\_mesher.txt} file, which is created after the mesher completed.
 In case you used CUBIT to create the mesh, you might use an arbitrary value when running \texttt{xgenerate\_databases} and then use the value suggested in the ~\\
-\texttt{in\_out\_files/OUTPUT\_FILES/output\_mesher.txt} file after the database generation completed.
+\texttt{OUTPUT\_FILES/output\_mesher.txt} file after the database generation completed.
 Note that the implemented Newmark time scheme uses this time step globally, thus your simulations become more expensive for very small mesh elements in high wave-speed regions.
@@
 
-The solver needs three input files in the \texttt{in\_data\_files/} directory
+The solver needs three input files in the \texttt{DATA} directory
 to run:
@@
 \end{lyxcode}
-The solver \texttt{xspecfem3D} filters the list of stations in file \texttt{in\_data\_files/STATIONS}
+The solver \texttt{xspecfem3D} filters the list of stations in file \texttt{DATA/STATIONS}
 to exclude stations that are not located within the region given in
@@
 and between \texttt{LONGITUDE\_MIN} and \texttt{LONGITUDE\_MAX}).
-The filtered file is called \texttt{in\_data\_files/STATIONS\_FILTERED}.
-
-Solver output is provided in the \texttt{in\_out\_files/OUTPUT\_FILES} directory
+The filtered file is called \texttt{DATA/STATIONS\_FILTERED}.
+
+Solver output is provided in the \texttt{OUTPUT\_FILES} directory
 in the \texttt{output\_solver.txt} file. Output can be directed to
@@
 the `\texttt{\small timestamp{*}}' files in the ~\\
-\texttt{\small in\_out\_files/OUTPUT\_FILES} directory.
+\texttt{\small OUTPUT\_FILES} directory.
 These tiny files look something like this:
@@
 and then collect the recorded seismograms at all the stations given
-in \texttt{in\_data\_files/STATIONS}.
+in \texttt{DATA/STATIONS}.
 
@@
 adjoint sources and run the adjoint simulation, and compile them into
-the \texttt{in\_data\_files/STATIONS\_ADJOINT} file, which has the same format
-as the regular \texttt{in\_data\_files/STATIONS} file.
+the \texttt{DATA/STATIONS\_ADJOINT} file, which has the same format
+as the regular \texttt{DATA/STATIONS} file.
 
@@
 adjoint sources need to be computed from the original recorded seismograms
-for the selected stations and saved in the \texttt{in\_out\_files/SEM/} directory
+for the selected stations and saved in the \texttt{OUTPUT\_FILES/SEM/} directory
 with the format \texttt{STA.NT.BX?.adj}, where \texttt{STA}, \texttt{NT}
-are the station name and network code given in the \texttt{in\_data\_files/STATIONS\_ADJOINT}
+are the station name and network code given in the \texttt{DATA/STATIONS\_ADJOINT}
 file, and \texttt{BX?} represents the component name of a particular
@@
 \item \textbf{Set the related parameters and run the adjoint simulation}\\
-In the \texttt{in\_data\_files/Par\_file}, set the two related parameters to
+In the \texttt{DATA/Par\_file}, set the two related parameters to
 be \texttt{SIMULATION\_TYPE = 2} and \texttt{SAVE\_FORWARD = .false.}.
@@
 \item These adjoint seismograms are recorded at the locations of the original
-earthquake sources given by the \texttt{in\_data\_files/CMTSOLUTION} file, and
+earthquake sources given by the \texttt{DATA/CMTSOLUTION} file, and
 have names of the form \texttt{S?????.NT.S??.sem} for the six-component
@@
 With the successful forward simulation and the adjoint source ready
-in the \texttt{in\_out\_files/SEM/} directory, set \texttt{SIMULATION\_TYPE = 3} and \texttt{SAVE\_FORWARD
+in the \texttt{OUTPUT\_FILES/SEM/} directory, set \texttt{SIMULATION\_TYPE = 3} and \texttt{SAVE\_FORWARD
 = .false.} in the \texttt{Par\_file} (you can use \texttt{change\_simulation\_type.pl -b}),
@@
 
-Besides earthquake simulations, SPECFEM3D includes functionality for seismic noise tomography as well.
+Besides earthquake simulations, SPECFEM3D Cartesian includes functionality for seismic noise tomography as well.
 In order to proceed successfully in this chapter, it is critical that you have already familiarized yourself
@@
 As usual, the three main input files are crucial: \texttt{Par\_file}, \texttt{CMTSOLUTION} and \texttt{STATIONS}.
-Unless otherwise specified, those input files should be located in directory \texttt{in\_data\_files/}.\\
+Unless otherwise specified, those input files should be located in directory \texttt{DATA/}.\\
 
@@
 \texttt{S\_squared should be put into directory:}\\
-\texttt{in\_out\_files/NOISE\_TOMOGRAPHY/ in the SPECFEM3D package}\\
+\texttt{OUTPUT\_FILES/NOISE\_TOMOGRAPHY/ in the SPECFEM3D Cartesian package}\\
 
@@
 \item
-Create a new directory in the SPECFEM3D package, name it as \texttt{in\_out\_files/NOISE\_TOMOGRAPHY/}.
+Create a new directory in the SPECFEM3D Cartesian package, name it as \texttt{OUTPUT\_FILES/NOISE\_TOMOGRAPHY/}.
 We will add some parameter files later in this folder.
@@
 \item
-Put the Matlab-generated-file \texttt{S\_squared} in \texttt{in\_out\_files/NOISE\_TOMOGRAPHY/}.
-
-That's to say, you will have a file \texttt{in\_out\_files/NOISE\_TOMOGRAPHY/S\_squared} in the SPECFEM3D package.
+Put the Matlab-generated-file \texttt{S\_squared} in \texttt{OUTPUT\_FILES/NOISE\_TOMOGRAPHY/}.
+
+That's to say, you will have a file \texttt{OUTPUT\_FILES/NOISE\_TOMOGRAPHY/S\_squared} in the SPECFEM3D Cartesian package.
 
 \item
-Create a file called \texttt{in\_out\_files/NOISE\_TOMOGRAPHY/irec\_master\_noise}. Note that this file is located
-in directory \texttt{in\_out\_files/NOISE\_TOMOGRAPHY/} as well. In general, all noise simulation related parameter
+Create a file called \texttt{OUTPUT\_FILES/NOISE\_TOMOGRAPHY/irec\_master\_noise}. Note that this file is located
+in directory \texttt{OUTPUT\_FILES/NOISE\_TOMOGRAPHY/} as well. In general, all noise simulation related parameter
 files go into that directory.
@@
 of the `master' receiver. For example, if this file contains 5, it means that the fifth receiver
-listed in \texttt{in\_data\_files/STATIONS} becomes the `master'. That's why we mentioned previously that
-the order of receivers in \texttt{in\_data\_files/STATIONS} is important.
-
-Note that in regional simulations, the \texttt{in\_data\_files/STATIONS}
+listed in \texttt{DATA/STATIONS} becomes the `master'. That's why we mentioned previously that
+the order of receivers in \texttt{DATA/STATIONS} is important.
+
+Note that in regional simulations, the \texttt{DATA/STATIONS}
 might contain receivers which are outside of our computational domains.
 Therefore, the integer in \texttt{irec\_master\_noise}
-is actually the ID in \texttt{in\_data\_files/STATIONS\_FILTERED}
+is actually the ID in \texttt{DATA/STATIONS\_FILTERED}
 (which is generated by \texttt{bin/xgenerate\_databases}).
@@
 \item
-Create a file called \texttt{in\_out\_files/NOISE\_TOMOGRAPHY/nu\_master}. This file holds three numbers,
+Create a file called \texttt{OUTPUT\_FILES/NOISE\_TOMOGRAPHY/nu\_master}. This file holds three numbers,
 forming a (unit) vector. It describes which component we are cross-correlating at the `master'
@@
 With all of the above done, we can finally launch our simulations.
-Again, please make sure that the \texttt{LOCAL\_PATH} in \texttt{in\_data\_files/Par\_file} is not globally shared.
+Again, please make sure that the \texttt{LOCAL\_PATH} in \texttt{DATA/Par\_file} is not globally shared.
 It is quite problematic to have a globally shared \texttt{LOCAL\_PATH}, in terms of both disk storage and
@@
 
-After those simulations, you have all stuff you need, either in the \texttt{in\_out\_files/OUTPUT\_FILES/} or
-in the directory specified by \texttt{LOCAL\_PATH} in \texttt{in\_data\_files/Par\_file}
+After those simulations, you have all stuff you need, either in the \texttt{OUTPUT\_FILES/} or
+in the directory specified by \texttt{LOCAL\_PATH} in \texttt{DATA/Par\_file}
 (which are most probably on local nodes).
@@
 The simulated ensemble cross correlations are obtained after the second simulation (Step 2).
-Seismograms in \texttt{in\_out\_files/OUTPUT\_FILES/} are actually the simulated ensemble cross correlations.
+Seismograms in \texttt{OUTPUT\_FILES/} are actually the simulated ensemble cross correlations.
 Collect them immediately after Step 2, or the Step 3 will overwrite them.
@@
 One contribution of the ensemble cross-correlation sensitivity kernels are obtained after Step 3,
-stored in the \texttt{in\_data\_files/LOCAL\_PATH} on local nodes.
+stored in the \texttt{DATA/LOCAL\_PATH} on local nodes.
 The ensemble kernel files are named the same as regular earthquake kernels.\\
@@
 files are saved in the \texttt{LOCAL\_PATH} directory, whereas \texttt{MOVIE\_SURFACE}
-output files are saved in the \texttt{in\_out\_files/OUTPUT\_FILES} directory. We
+output files are saved in the \texttt{OUTPUT\_FILES} directory. We
 save the displacement field if the parameter \texttt{SAVE\_DISPLACEMENT} is set, otherwise the velocity field is saved.
@@
 flag turned on, the code outputs \texttt{moviedata??????} files in
-the \texttt{in\_out\_files/OUTPUT\_FILES} directory. There are several flags
+the \texttt{OUTPUT\_FILES} directory. There are several flags
 in the main parameter file \texttt{Par\_file} that control the output of these moviedata files
@@
 Before compiling the code, make sure you have the file \texttt{surface\_from\_mesher.h}
-in the  \texttt{in\_out\_files/OUTPUT\_FILES/} directory. This file will be created by the solver run.
+in the  \texttt{OUTPUT\_FILES/} directory. This file will be created by the solver run.
 Then type
@@
 
-The \texttt{SPECFEM3D} code is running in near real-time to produce
+The \texttt{SPECFEM3D Cartesian} code is running in near real-time to produce
 animations of southern California earthquakes via the web; see Southern California ShakeMovie\textregistered \urlwithparentheses{www.shakemovie.caltech.edu}.
@@
 turned on, the code outputs several files in \texttt{\small LOCAL\_PATH} specified in the main
-\texttt{Par\_file}, e.g. in directory \texttt{in\_out\_files/DATABASES\_MPI}.
+\texttt{Par\_file}, e.g. in directory \texttt{OUTPUT\_FILES/DATABASES\_MPI}.
 %As the files can be very large, there are several flags in the \texttt{\small Par\_file}
@@
 {\footnotesize cd~bin/~}{\footnotesize \par}
-{\footnotesize ./xcombine\_vol\_data~0~3~velocity\_Z\_it000400~../in\_out\_files/DATABASES\_MPI~../in\_out\_files/OUTPUT\_FILES~0~}{\footnotesize \par}
+{\footnotesize ./xcombine\_vol\_data~0~3~velocity\_Z\_it000400~../OUTPUT\_FILES/DATABASES\_MPI~../OUTPUT\_FILES~0~}{\footnotesize \par}
 \end{lyxcode}
@@
 also generates the file \texttt{sr.vtk} located in the
-\texttt{in\_out\_files/OUTPUT\_FILES/} directory to describe
+\texttt{OUTPUT\_FILES/} directory to describe
 the source and receiver locations, which can also be viewed in Paraview
@@
 the mesher, database generator and solver together, figures out the number of processors
-required for this simulation from the \texttt{in\_data\_files/Par\_file}, and
+required for this simulation from the \texttt{DATA/Par\_file}, and
 submits the LSF job.
@@
 
-NPROC=`grep~NPROC~in\_data\_files/Par\_file~|~cut~-c~34-~'~\\
+NPROC=`grep~NPROC~DATA/Par\_file~|~cut~-c~34-~'~\\
 ~\\
@@
 
-NPROC\_XI=`grep~NPROC\_XI~in\_data\_files/meshfem3D\_files/Mesh\_Par\_file~|~cut~-c~34-~'~
+NPROC\_XI=`grep~NPROC\_XI~DATA/meshfem3D\_files/Mesh\_Par\_file~|~cut~-c~34-~'~
 ~\\
 
-NPROC\_ETA=`grep~NPROC\_ETA~in\_data\_files/meshfem3D\_files/Mesh\_Par\_file~|~cut~-c~34-~'~
+NPROC\_ETA=`grep~NPROC\_ETA~DATA/meshfem3D\_files/Mesh\_Par\_file~|~cut~-c~34-~'~
 ~\\
@@
 output from \texttt{stdout} into ~\\
-\texttt{\small in\_out\_files/OUTPUT\_FILES/\%J.o},
+\texttt{\small OUTPUT\_FILES/\%J.o},
 where \texttt{\%J} is short-hand for the job-ID; it also tells the
@@
 from Shmux \urlwithparentheses{web.taranis.org/shmux/}. Make sure that the \texttt{LOCAL\_PATH}
-parameter in \texttt{in\_data\_files/Par\_file} is also set properly.
+parameter in \texttt{DATA/Par\_file} is also set properly.
 \item The next portion of the script launches the mesher, database generator and then the solver
@@
 
-\#BSUB~-o~in\_out\_files/OUTPUT\_FILES/\%J.o~
+\#BSUB~-o~OUTPUT\_FILES/\%J.o~
 
@@
 
-mkdir~-p~in\_out\_files/OUTPUT\_FILES~
-
-echo~\char`\"{}\$LSB\_MCPU\_HOSTS\char`\"{}~>~in\_out\_files/OUTPUT\_FILES/lsf\_machines~
-
-echo~\char`\"{}\$LSB\_JOBID\char`\"{}~>~in\_out\_files/OUTPUT\_FILES/jobid
-
-remap\_lsf\_machines.pl~in\_out\_files/OUTPUT\_FILES/lsf\_machines~> in\_out\_files/OUTPUT\_FILES/machines
+mkdir~-p~OUTPUT\_FILES~
+
+echo~\char`\"{}\$LSB\_MCPU\_HOSTS\char`\"{}~>~OUTPUT\_FILES/lsf\_machines~
+
+echo~\char`\"{}\$LSB\_JOBID\char`\"{}~>~OUTPUT\_FILES/jobid
+
+remap\_lsf\_machines.pl~OUTPUT\_FILES/lsf\_machines~> OUTPUT\_FILES/machines
 
@@
 
-~~~~~~-~<~in\_out\_files/OUTPUT\_FILES/machines~>/dev/null~\\
+~~~~~~-~<~OUTPUT\_FILES/machines~>/dev/null~\\
 ~\\
@@
 
-mkdir~-p~in\_out\_files/SEM
-
-cd~in\_out\_files/SEM/~
+mkdir~-p~OUTPUT\_FILES/SEM
+
+cd~OUTPUT\_FILES/SEM/~
 
@@
 These changes involve contributing specific subroutines
-that replace existing subroutines in the \texttt{SPECFEM3D} package.
-Note that \texttt{SPECFEM3D} can handle Earth models with material properties that vary within each spectral element.
+that replace existing subroutines in the \texttt{SPECFEM3D Cartesian} package.
+Note that \texttt{SPECFEM3D Cartesian} can handle Earth models with material properties that vary within each spectral element.
 
@@
 The external tomographic model is represented by a grid of points with assigned material properties and homogeneous resolution along each spatial direction x, y and z.
-The ASCII file \texttt{tomography\_model.xyz} that describes the tomography should be located in the \texttt{in\_data\_files/} directory.
+The ASCII file \texttt{tomography\_model.xyz} that describes the tomography should be located in the \texttt{DATA/} directory.
 The format of the file, as read from \texttt{model\_tomography.f90}, looks like Figure \ref{fig:tomography_file},
@@
 The forward and adjoint simulations generate synthetic seismograms
-in the \texttt{in\_out\_files/OUTPUT\_FILES/} directory by default. For the forward simulation, the files
+in the \texttt{OUTPUT\_FILES/} directory by default. For the forward simulation, the files
 are named like \texttt{STA.NT.BX?.semd} for two-column time series, or
@@
 This can be useful on a cluster, where only the master process node has access to the output directory.
-\item [{\texttt{USE\_OUTPUT\_FILES\_PATH}}] Set to \texttt{.false.} to have the seismograms output to \texttt{LOCAL\_PATH} directory specified in the main parameter file \texttt{in\_data\_files/Par\_file}.
+\item [{\texttt{USE\_OUTPUT\_FILES\_PATH}}] Set to \texttt{.false.} to have the seismograms output to \texttt{LOCAL\_PATH} directory specified in the main parameter file \texttt{DATA/Par\_file}.
 In this case, you could collect the synthetics onto the frontend using the \texttt{collect\_seismo\_lsf\_multi.pl} script
@@
 \begin{lyxcode}
-collect\_seismo.pl~machines~in\_data\_files/Par\_file
+collect\_seismo.pl~machines~DATA/Par\_file
 \end{lyxcode}
@@
 \end{description}
-where  \texttt{machines} is a file containing the node names and  \texttt{in\_data\_files/Par\_file} the parameter file used to extract the \texttt{LOCAL\_PATH} directory used for the simulation.
+where  \texttt{machines} is a file containing the node names and  \texttt{DATA/Par\_file} the parameter file used to extract the \texttt{LOCAL\_PATH} directory used for the simulation.
 
@@
 
-With the movie data saved in \texttt{in\_out\_files/OUTPUT\_FILES/} at the end of
+With the movie data saved in \texttt{OUTPUT\_FILES/} at the end of
 a movie simulation (\texttt{\small MOVIE\_SURFACE=.true.}), you can
@@
 
-With the shakemap data saved in \texttt{in\_out\_files/OUTPUT\_FILES/} at the end
+With the shakemap data saved in \texttt{OUTPUT\_FILES/} at the end
 of a shakemap simulation ~\\
@@
 The basic features of an adjoint-based tomographic inversion were illustrated in \citet{TaLiTr07} and \citet{TrKoLi08} using a conjugate-gradient algorithm; there are dozens of versions of gradient-based inversion algorithms that could alternatively be used.
-The tomographic inversion of \citet{TaLiMaTr09,TaLiMaTr2010} used SPECFEM3D as well as several additional components which are also stored on the CIG svn server, described next.
-
-The directory containing utilities for tomographic inversion using SPECFEM3D (or other packages that evaluate misfit functions and gradients) is here on the CIG svn server:
+The tomographic inversion of \citet{TaLiMaTr09,TaLiMaTr2010} used SPECFEM3D Cartesian as well as several additional components which are also stored on the CIG svn server, described next.
+
+The directory containing utilities for tomographic inversion using SPECFEM3D Cartesian (or other packages that evaluate misfit functions and gradients) is here on the CIG svn server:
 %
@@
 
-The SPECFEM3D software code internally uses Cartesian coordinates.
+The SPECFEM3D Cartesian software code internally uses Cartesian coordinates.
 The given locations for sources and receiver locations thus may get
@@
 \end{description}
-The seismograms are by default written out in ASCII-format to the \texttt{in\_out\_files/OUTPUT\_FILES/} subdirectory by each parallel process.
+The seismograms are by default written out in ASCII-format to the \texttt{OUTPUT\_FILES/} subdirectory by each parallel process.
 You can change this behavior by changing the following flags in the \texttt{constants.h} file located in the \texttt{src/shared/} subdirectory:
@@
 This can be useful on a cluster, where only the master process node has access to the output directory.
-\item [{\texttt{USE\_OUTPUT\_FILES\_PATH}}] Set to \texttt{.false.} to have the seismograms output to \texttt{LOCAL\_PATH} directory specified in the main parameter file \texttt{in\_data\_files/Par\_file}.
+\item [{\texttt{USE\_OUTPUT\_FILES\_PATH}}] Set to \texttt{.false.} to have the seismograms output to \texttt{LOCAL\_PATH} directory specified in the main parameter file \texttt{DATA/Par\_file}.
 \end{description}
@@
 IRIS  \texttt{(www.iris.edu)} uses SEED format for channel codes, which are represented by three letters, such as \texttt{LHN}, \texttt{BHZ}, etc.
-In older versions of the SPECFEM3D package, a common format was used for the channel codes of all seismograms, which was \texttt{BHE/BHN/BHZ} for three components.
+In older versions of the SPECFEM3D Cartesian package, a common format was used for the channel codes of all seismograms, which was \texttt{BHE/BHN/BHZ} for three components.
 To avoid confusion when comparison are made to observed data, we are now using the FDSN convention (\url{http://www.fdsn.org/}) for SEM seismograms.
@@
 Please visit \texttt{www.iris.edu/manuals/SEED\_appA.htm} for further information.
-Grey rows show the relative band-code range in SPECFEM3D, and the band codes used to name SEM seismograms are denoted in red.}
+Grey rows show the relative band-code range in SPECFEM3D Cartesian, and the band codes used to name SEM seismograms are denoted in red.}
 
@@
 \noindent \textbf{\texttt{Orientation code:}} The third letter in channel codes is an orientation code, which generally describes the physical configuration of the components of instrument packages.
-SPECFEM3D uses the traditional orientation code  \texttt{E/N/Z} (East-West, North-South, Vertical) for three components when a UTM projection is used.
+SPECFEM3D Cartesian uses the traditional orientation code  \texttt{E/N/Z} (East-West, North-South, Vertical) for three components when a UTM projection is used.
 If the UTM conversion is suppressed, i.e.
@@
 
-\noindent \textbf{\texttt{EXAMPLE:}} The sampling rate is given by \texttt{DT} in the main parameter file \texttt{Par\_file} located in the \texttt{in\_data\_files/} subdirectory.
+\noindent \textbf{\texttt{EXAMPLE:}} The sampling rate is given by \texttt{DT} in the main parameter file \texttt{Par\_file} located in the \texttt{DATA/} subdirectory.
 Depending on the resolution of your simulations, if you choose a sampling rate greater than $0.01$ s and less than $1$ s, a seismogram recording displacements on the vertical component of a station \texttt{ASBS} (network \texttt{AZ}) will be named \texttt{ASBS.AZ.MXZ.semd.sac}, whereas it will be \texttt{ASBS.AZ.BXZ.semd.sac}, if the sampling rate is greater than 0.0125 and less equal to 0.1 s.
@@
 \item [the code gives following error message "need at least one receiver":]
-This means that no stations given in the input file \texttt{in\_data\_files/STATIONS} could be located within the dimensions of the mesh.
+This means that no stations given in the input file \texttt{DATA/STATIONS} could be located within the dimensions of the mesh.
 This can happen for example when the mesh was created with the in-house mesher \texttt{xmeshfem3D} while using the Universal Transverse Mercator (UTM) projection but the simulation with \texttt{xspecfem3D} was suppressing this projection from latitude/longitude to x/y/z coordinates.
 
-In such cases, try to change your \texttt{in\_data\_files/Par\_file} and set e.g.:
+In such cases, try to change your \texttt{DATA/Par\_file} and set e.g.:
 \begin{lyxcode}
@@
 In most cases this means that your time step size \texttt{DT} is chosen too big.
-Look at your files \texttt{output\_mesher.txt} or \texttt{output\_solver.txt} created in the folder \texttt{in\_out\_files/OUTPUT\_FILES/}.
+Look at your files \texttt{output\_mesher.txt} or \texttt{output\_solver.txt} created in the folder \texttt{OUTPUT\_FILES}.
 In these output files, find the section:
@@
 
-then change \texttt{DT} in the \texttt{in\_data\_files/Par\_file} to be somewhere close to the maximum suggested time step. In the example above: \\
+then change \texttt{DT} in the \texttt{DATA/Par\_file} to be somewhere close to the maximum suggested time step. In the example above: \\
 \texttt{DT                              = 0.05d0} \\

### 2012-10-22 doc/USER_MANUAL/manual_SPECFEM3D_Cartesian.tex
added IBM BlueGene compilation instructions

@@
 Roland Martin,
+Ren\'e Matzen,
 Dennis McRitchie,
@@
 scheduling policy, which will slow down the job quite a bit.
+
+\section{Compiling on an IBM BlueGene}
+
+\underline{More recent installation instruction for IBM BlueGene, from October 2012:}\\
+
+\noindent
+Edit file \texttt{flags.guess} and put this for \texttt{FLAGS\_CHECK} and \texttt{FLAGS\_NO\_CHECK}:\\
+
+\noindent
+\texttt{-g -qfullpath -O2 -qsave -qstrict -qtune=qp -qarch=qp -qcache=auto -qhalt=w}\\
+\noindent
+\texttt{-qfree=f90 -qsuffix=f=f90 -qlanglvl=95pure -Wl,-relax -Q -Q+rank,swap\_all}.\\
+
+\noindent
+The most relevant are the -qarch and -qtune flags, otherwise if these flags are set to ``auto'' then they are wrongly assigned to
+the architecture of the frond-end node, which is different from that on the compute nodes.
+You will need to set these flags to the right architecture for your BlueGene compute nodes, which is not necessarily ``qp'';
+ask your system administrator.
+On some machines if is necessary to use -O2 in these flags instead of -O3 due to a compiler bug of the XLF version installed.
+We thus suggest to first try -O3, and then if the code does not compile or does not run fine then switch back to -O2.
+The debug flags (-g, -qfullpath) do not influence performance but are useful to get at least some insights in case of problems.\\
+
+\noindent
+Before running \texttt{configure}, select the XL Fortran compiler by typing \texttt{module load bgq-xl/1.0}
+or \texttt{module load bgq-xl} (another, less efficient option is to load the GNU compilers using \texttt{module load bgq-gnu/4.4.6} or similar).\\
+
+\noindent
+Then, to configure the code, for the SPECFEM3D\_GLOBE version type this:\\
+
+\noindent
+\texttt{./configure FC=bgxlf90\_r MPIFC=mpixlf90\_r CC=bgxlc\_r LOCAL\_PATH\_IS\_ALSO\_GLOBAL=true}\\
+
+\noindent
+and for SPECFEM3D\_Cartesian type this:\\
+
+\noindent
+\texttt{./configure FC=bgxlf90\_r MPIFC=mpixlf90\_r CC=bgxlc\_r LOCAL\_PATH\_IS\_ALSO\_GLOBAL=true}.\\
+
+\noindent
+\underline{Older installation instruction for IBM BlueGene, from 2011:}\\
+
+\noindent
+To compile the code on an IBM BlueGene, Laurent L\'eger from IDRIS, France, suggests the following: compile the code with
+
+\noindent
+\texttt{   FLAGS\_NO\_CHECK="-O3 -qsave -qstrict -qtune=auto -qarch=450d -qcache=auto \\}
+\texttt{   -qfree=f90 -qsuffix=f=f90 -g -qlanglvl=95pure -qhalt=w -Q -Q+rank,swap\_all -Wl,-relax"}
+
+\noindent
+Option "-Wl,-relax" must be added on many (but not all) BlueGene systems to be able to link the binaries \texttt{xmeshfem3D}
+and \texttt{xspecfem3D} because the final link step is done by the GNU \texttt{ld} linker even if
+one uses \texttt{FC=bgxlf90\_r, MPIFC=mpixlf90\_r} and \texttt{CC=bgxlc\_r} to create all the object files.
+On the contrary, on some BlueGene systems that use the native AIX linker option "-Wl,-relax" can lead to problems and must be suppressed from \texttt{flags.guess}.
+
+\noindent
+Also, \texttt{AR=ar, ARFLAGS=cru} and \texttt{RANLIB=ranlib} are hardwired in all \texttt{Makefile.in} files by default, but to cross-compile on BlueGene/P one needs to change
+these values to \texttt{AR=bgar, ARFLAGS=cru} and \texttt{RANLIB=bgranlib}. Thus the easiest thing to do is to modify
+all \texttt{Makefile.in} files and the \texttt{configure} script to set them automatically by \texttt{configure}.
+One then just needs to pass the right commands to the \texttt{configure} script:
+
+\noindent
+\texttt{   ./configure --prefix=/path/to/SPECFEM3DG\_SP --host=Babel --build=BGP \\}
+\texttt{      FC=bgxlf90\_r MPIFC=mpixlf90\_r CC=bgxlc\_r AR=bgar ARFLAGS=cru \\}
+\texttt{      RANLIB=bgranlib LOCAL\_PATH\_IS\_ALSO\_GLOBAL=false}
+
+\noindent
+This trick can be useful for all hosts on which one needs to cross-compile.
+
+\noindent
+On BlueGene, one also needs to run the \texttt{xcreate\_header\_file} binary file manually rather than in the Makefile:
+
+\noindent
+   \texttt{bgrun -np 1 -mode VN -exe ./bin/xcreate\_header\_file}
 

### 2012-10-24 doc/USER_MANUAL/manual_SPECFEM3D_Cartesian.tex
fixed a file name in the manual
simplified the IBM BlueGene section

@@
 To accommodate visco-plastic or non-linear rheologies, the reader can refer to
-the \href{http://geoelse.stru.polimi.it/}{GeoELSE} software package \citep{CaGa97,StPaIg09,ChMoTsBaKrKaStKr10}.
+the \href{http://geoelse.stru.polimi.it/}{GeoELSE} software package \citep{CaGa97,StPaIg09}.
 
@@
 \begin{lyxcode}
-{\small tar~-zxvf~SPECFEM3D\_V2.1.0.tar.gz}{\small \par}
+{\small tar~-zxvf~SPECFEM3D\_V2.1.0\_Cartesian.tar.gz}{\small \par}
 \end{lyxcode}
@@
 \noindent
-Then, to configure the code, for the SPECFEM3D\_GLOBE version type this:\\
+Then, to configure the code, type this:\\
 
@@
 \texttt{./configure FC=bgxlf90\_r MPIFC=mpixlf90\_r CC=bgxlc\_r LOCAL\_PATH\_IS\_ALSO\_GLOBAL=true}\\
-
-\noindent
-and for SPECFEM3D\_Cartesian type this:\\
-
-\noindent
-\texttt{./configure FC=bgxlf90\_r MPIFC=mpixlf90\_r CC=bgxlc\_r LOCAL\_PATH\_IS\_ALSO\_GLOBAL=true}.\\
 
@@
 There are options for computing a conjugate gradient step, as well as a source subspace projection step.
-The workflow could use substantial ``stitching together,'' and we welcome suggestions; undoubtedly, the tomographic inversion procedure will improve with time.
+The workflow could use substantial ``stitching together,'' and we welcome suggestions;
+undoubtedly, the tomographic inversion procedure will improve with time.
 {\bf This is a work in progress.}

### 2012-10-25 doc/USER_MANUAL/manual_SPECFEM3D_Cartesian.tex
updated the manual

@@
 \section*{Authors}
-The SPECFEM3D Cartesian package was first developed by Dimitri Komatitsch and Jeroen Tromp at Harvard University and Caltech, USA, starting in 1998, based on earlier work by Dimitri Komatitsch and Jean-Pierre Vilotte at Institut de Physique du Globe (IPGP) in Paris, France from 1995 to 1997.\\
+The SPECFEM3D Cartesian package was first developed by Dimitri Komatitsch and Jean-Pierre Vilotte at Institut de Physique du Globe (IPGP) in Paris, France 
+from 1995 to 1997 and then by Dimitri Komatitsch and Jeroen Tromp at Harvard University and Caltech, USA, starting in 1998.
 
@@
 Min Chen,
+Dominik G\"oddeke,
 Vala Hj\"orleifsd\'ottir,
@@
 Jeroen Tromp,
+Jean-Pierre Vilotte,
 Zhinan Xie,
 Hejun Zhu.\\
-%\toall{add full list of developers here and update the AUTHORS file accordingly}
 
@@
 Forward and adjoint simulations are described in detail in \cite{TrKoLi08,PeKoLuMaLeCaLeMaLiBlNiBaTr11,VaCaSaKoVi99,KoMiEr09,KoErGoMi10,KoGoErMi10,ChKoViCaVaFe07,MaKoDi09,KoViCh10,CaKoLaTiMiLeSnTr08,TrKoHjLiZhPeBoMcFrTrHu10,KoRiTr02,KoTr02a,KoTr02b,KoTr99} or \cite{KoVi98}.
-Additional aspects dealing with adjoint simulations are described in
+Additional aspects of adjoint simulations are described in
 \cite{TrTaLi05,LiTr06,TrKoLi08,LiTr08,TrKoHjLiZhPeBoMcFrTrHu10,PeKoLuMaLeCaLeMaLiBlNiBaTr11}.
@@
 provides efficent static mapping, graph and mesh partitioning routines. SCOTCH is a free software package developed by
-Fran\c {c}ois Pellegrini et al.
-from LaBRI and INRIA in Bordeaux, France, downloadable from the web page \url{https://gforge.inria.fr/projects/scotch/}.
-It is more recent than METIS, actively maintained and performs better in many cases.
-A recent version of its source code is provided in directory \texttt{src/decompose\_mesh\_SCOTCH/scotch\_5.1.10b}.
-In case no SCOTCH libraries can be found on the system, the configuration will bundle this version for compilation.
+Fran\c {c}ois Pellegrini et al. from LaBRI and INRIA in Bordeaux, France, downloadable from the web page \url{https://gforge.inria.fr/projects/scotch/}.
+In case no SCOTCH libraries can be found on the system, the configuration will bundle the version provided with the source code for compilation.
 The path to an existing SCOTCH installation can to be set explicitly with the option \texttt{-{}-with-scotch-dir}. Just as an example:
@@
 \noindent
-\texttt{-qfree=f90 -qsuffix=f=f90 -qlanglvl=95pure -Wl,-relax -Q -Q+rank,swap\_all}.\\
+\texttt{-qfree=f90 -qsuffix=f=f90 -qlanglvl=95pure -Q -Q+rank,swap\_all -Wl,-relax}\\
 
@@
 which provides efficent static mapping, graph and mesh partitioning routines. SCOTCH is a free software package developed by
-Fran\c {c}ois Pellegrini et al.
-from LaBRI and INRIA in Bordeaux, France, downloadable from the web page \url{https://gforge.inria.fr/projects/scotch/}.
-It is more recent than METIS, actively maintained and performs better in many cases.
-A recent version of its source code is provided in directory \texttt{src/decompose\_mesh\_SCOTCH/scotch\_5.1.10b}.
+Fran\c {c}ois Pellegrini et al. from LaBRI and INRIA in Bordeaux, France, downloadable from the web page \url{https://gforge.inria.fr/projects/scotch/}.
 \\
@@
 
-Main authors: Dimitri Komatitsch and Jeroen Tromp
+Main historical authors: Dimitri Komatitsch and Jeroen Tromp
 
@@
  Dimitri Komatitsch, Nicolas Le Goff, Roland Martin and Pieyre Le Loher, University of Pau, France,
- Jeroen Tromp and the Princeton group of developers, Princeton University, USA,
+ Daniel Peter, Jeroen Tromp and the Princeton group of developers, Princeton University, USA,
  and Emanuele Casarotti, INGV Roma, Italy:
-  support for CUBIT meshes decomposed by SCOTCH, METIS or ZOLTAN;
+  support for CUBIT meshes decomposed by SCOTCH;
   much faster solver using Michel Deville's inlined matrix products.\\

### 2012-10-26 doc/USER_MANUAL/manual_SPECFEM3D_Cartesian.tex
updates the user manual; modifies all parameter files

@@
 \texttt{ATTENUATION\_f0\_REFERENCE} in the main constants file \texttt{constants.h} found in subdirectory \texttt{src/shared/}.
+\item [{\texttt{ANISOTROPY}}] Set to \texttt{.true.} if you
+want to use an anisotropy model. Please see the file \texttt{model\_aniso.f90} in subdirectory \texttt{src/generate\_databases/} for the current implementation of anisotropic models.
 \item [{\texttt{USE\_OLSEN\_ATTENUATION}}] Set to \texttt{.true.} if you
@@
 using Olsen's empirical relation (see \citet{OlDaBr03}).
-\item [{\texttt{ANISOTROPY}}] Set to \texttt{.true.} if you
-want to use an anisotropy model. Please see the file \texttt{model\_aniso.f90} in subdirectory \texttt{src/generate\_databases/} for the current implementation of anisotropic models.
+\item [{\texttt{OLSEN\_ATTENUATION\_RATIO}}] Determines the Olsen's constant in Olsen's empirical relation (see \citet{OlDaBr03}).
 \item [{\texttt{ABSORBING\_CONDITIONS}}] Set to \texttt{.true.} to turn
@@
 absorbing boundary conditions on the top surface which by default constitutes a free surface of the model.
+\item [{\texttt{CREATE\_SHAKEMAP}}] Set this flag to \texttt{.true.} to
+create a ShakeMap\textregistered, i.e., a peak ground velocity map
+of the maximum absolute value of the two horizontal components of the velocity vector.
 \item [{\texttt{MOVIE\_SURFACE}}] Set to \texttt{.false.}, unless you want
@@
 for a discussion on the generation of movies. This feature is only relevant for the solver.
+\item [{\texttt{MOVIE\_TYPE}}] Set this flag to 1 to show the top
+  surface (tomography + oceans) only, to 2 to show all external faces
+  of the mesh (i.e. topography + vertical edges + bottom) in shakemaps and surface movies.
 \item [{\texttt{MOVIE\_VOLUME}}] Set to \texttt{.false.}, unless you want
@@
 for a discussion on the generation of movies. This feature is only relevant for the solver.
-\item [{\texttt{NTSTEP\_BETWEEN\_FRAMES}}] Determines the number of timesteps
-between movie frames. Typically you want to save a snapshot every
-100 timesteps. The smaller you make this number the more output will
-be generated! See Section~\ref{sec:Movies} for a discussion on the
-generation of movies. This feature is only relevant for the solver.
-\item [{\texttt{CREATE\_SHAKEMAP}}] Set this flag to \texttt{.true.} to
-create a ShakeMap\textregistered, i.e., a peak ground velocity map
-of the maximum absolute value of the two horizontal components of the velocity vector.
 \item [{\texttt{SAVE\_DISPLACEMENT}}] Set this flag to \texttt{.true.}
@@
 movie frames.
-\item [{\texttt{HDUR\_MOVIE}}] determines the half duration of the source time function for the movie simulations.
+\item [{\texttt{NTSTEP\_BETWEEN\_FRAMES}}] Determines the number of timesteps
+between movie frames. Typically you want to save a snapshot every
+100 timesteps. The smaller you make this number the more output will
+be generated! See Section~\ref{sec:Movies} for a discussion on the
+generation of movies. This feature is only relevant for the solver.
+\item [{\texttt{HDUR\_MOVIE}}] Determines the half duration of the source time function for the movie simulations.
 When this parameter is set to be 0, a default half duration that corresponds to the accuracy of the simulation is provided.
@@
 \item the ShakeMap\textregistered option \texttt{CREATE\_SHAKEMAP}
-\item the output information parameters \texttt{NTSTEP\_BETWEEN\_OUTPUT\_INFO}
+\item the output information parameters \texttt{MOVIE\_TYPE}, \texttt{NTSTEP\_BETWEEN\_OUTPUT\_INFO}
 and \texttt{NTSTEP\_BETWEEN\_OUTPUT\_}~\\
@@
 To make a surface or volume movie of the simulation, set parameters
-\texttt{MOVIE\_SURFACE}, \texttt{MOVIE\_VOLUME}, and \texttt{NTSTEP\_BETWEEN\_FRAMES}
+\texttt{MOVIE\_SURFACE}, \texttt{MOVIE\_VOLUME}, \texttt{MOVIE\_TYPE}, and \texttt{NTSTEP\_BETWEEN\_FRAMES}
 in the \texttt{Par\_file}. Turning on the movie flags, in particular
@@
 the \texttt{OUTPUT\_FILES} directory. There are several flags
-in the main parameter file \texttt{Par\_file} that control the output of these moviedata files
+in the main parameter file \texttt{Par\_file} that control the output of these movie data files
 (see section \ref{cha:Main-Parameter} for details):

### 2012-10-27 doc/USER_MANUAL/manual_SPECFEM3D_Cartesian.tex
renamed decompose_mesh_SCOTCH to decompose_mesh because METIS is now supported as well, in addition to SCOTCH (and ZOLTAN may also be added in the future).

@@
 (C-PML or ADE-PML) \citep{KoMa07,MaKoEz08,MaKoGe08,MaKo09,MaKoGeBr10}.
-It will also use the PT-SCOTCH parallel library for mesh partitioning.\\
+It will also use the PT-SCOTCH parallel and threaded version of SCOTCH for mesh partitioning instead of the serial version.\\
 
@@
 
-The SPECFEM3D Cartesian software package relies on the SCOTCH library to partition meshes created with CUBIT.
 Note that we use CUBIT to create meshes of hexahedra, but other packages can be used as well, for instance GiD from \url{http://gid.cimne.upc.es} or Gmsh from \url{http://geuz.org/gmsh} \citep{GeRe09}.
 Even mesh creation packages that generate tetrahedra, for instance TetGen from \url{http://tetgen.berlios.de}, can be used because each tetrahedron can then easily be decomposed into four hexahedra as shown in the picture of the TetGen logo at \url{http://tetgen.berlios.de/figs/Delaunay-Voronoi-3D.gif}; while this approach does not generate hexahedra of optimal quality, it can ease mesh creation in some situations and it has been shown that the spectral-element method can very accurately handle distorted mesh elements \citep{OlSe11}.
+
+The SPECFEM3D Cartesian software package relies on the SCOTCH library to partition meshes created with CUBIT.
+METIS \citep{KaKu98a,KaKu98c,KaKu98b} can also be used instead of SCOTCH if you prefer, by editing file src/decompose\_mesh/decompose\_mesh.F90 and uncommenting the flag USE\_METIS\_INSTEAD\_OF\_SCOTCH.
+You will also then need to install and compile Metis version 4.0 (do *NOT* install Metis version 5.0, which has incompatible function calls) and edit \texttt{src/decompose\_mesh/Makefile.in} and uncomment the METIS link flag in that file before running \texttt{configure}.\\
 
@@
 
-These mesh files are needed as input files for the partitioner  \texttt{xdecompose\_mesh\_SCOTCH}
+These mesh files are needed as input files for the partitioner  \texttt{xdecompose\_mesh}
 to load-balance the mesh. Please see the next section for further details.
@@
 
-\subsection{Partitioning the Mesh with \texttt{xdecompose\_mesh\_SCOTCH}}
+\subsection{Partitioning the Mesh with \texttt{xdecompose\_mesh}}
 
@@
 It amounts to balance the number of elements in each slice while minimizing the communication costs
-resulting from the placement of adjacent elements on different processors. \texttt{decompose\_mesh\_SCOTCH} depends on the SCOTCH library \citep{PeRo96},
+resulting from the placement of adjacent elements on different processors. \texttt{decompose\_mesh} depends on the SCOTCH library \citep{PeRo96},
 which provides efficent static mapping, graph and mesh partitioning routines. SCOTCH is a free software package developed by
@@
 in directory
-\texttt{src/decompose\_mesh\_SCOTCH/scotch\_5.1.10b}.
+\texttt{src/decompose\_mesh/scotch\_5.1.12b}.
 This version gets bundled with the compilation of the SPECFEM3D Cartesian package if no libraries could have been found.
@@
 Each single core partition is colored differently.
-The executable \texttt{xdecompose\_mesh\_SCOTCH} can equally distribute the mesh on any arbitrary number of cores.
+The executable \texttt{xdecompose\_mesh} can equally distribute the mesh on any arbitrary number of cores.
 Domain decomposition is explained in detail in \cite{MaKoBlLe08}, and excellent scaling up to 150,000 processor cores in shown for instance in
@@
 \begin{lyxcode}
-{\small make xdecompose\_mesh\_SCOTCH}{\small \par}
-\end{lyxcode}
-If all paths and flags have been set correctly, the executable \texttt{bin/xdecompose\_mesh\_SCOTCH} should be produced.
+{\small make xdecompose\_mesh}{\small \par}
+\end{lyxcode}
+If all paths and flags have been set correctly, the executable \texttt{bin/xdecompose\_mesh} should be produced.
 
@@
 \begin{lyxcode}
-./bin/xdecompose\_mesh\_SCOTCH  nparts  input\_directory output\_directory
+./bin/xdecompose\_mesh  nparts  input\_directory output\_directory
 \end{lyxcode}
@@
 
-After using \texttt{xmeshfem3D} or \texttt{xdecompose\_mesh\_SCOTCH}, the next step in the workflow is to compile
+After using \texttt{xmeshfem3D} or \texttt{xdecompose\_mesh}, the next step in the workflow is to compile
 \texttt{xgenerate\_}~\\
@@
 Please note that the \texttt{LOCAL\_PATH} directory should already contain the output files of
-the partitioner, i.e. from \texttt{xdecompose\_mesh\_SCOTCH} or \texttt{xmeshfem3D}.
+the partitioner, i.e. from \texttt{xdecompose\_mesh} or \texttt{xmeshfem3D}.
 \item [{\texttt{NTSTEP\_BETWEEN\_OUTPUT\_INFO}}] This parameter specifies

### 2012-10-31 doc/USER_MANUAL/manual_SPECFEM3D_Cartesian.tex
adds a (pseudo) Dirac delta source time function set by default when FORCESOLUTION is used; adds calculation of source arrays for interpolated locations of force point sources; corrects source time function factors used in GPU code; updates the user manual in Par_file and FORCESOLUTION related pages.

@@
 to avoid writing to the seismograms too often. This feature is only relevant for the solver.
-\item [{\texttt{USE\_FORCE\_POINT\_SOURCE}}] Turn this flag on to use a \texttt{FORCESOLUTION} force source
-instead of a \texttt{CMTSOLUTION} moment-tensor source. This can be useful e.g. for oil industry
-foothills simulations in which the source is a vertical force, normal force,
-tilted force, or an impact etc. By default a quasi-Heaviside time function is used to represent the force point source but a Ricker
-source time function can be computed instead by changing the value of \texttt{USE\_RICKER\_IPATI} in the main constants file
-\texttt{constants.h} located in the \texttt{src/shared/} subdirectory.
-Note that in the \texttt{FORCESOLUTION} file, you will need to edit the East, North and vertical components of an
-arbitrary (non-unitary) direction vector of the force vector; thus refer to Appendix~\ref{cha:Coordinates}
-for the orientation of the reference frame. This vector is made unitary internally in the code and thus only its direction matters here;
-its norm is ignored and the norm of the force used is the factor force source times the source time function.
+\item [{\texttt{USE\_FORCE\_POINT\_SOURCE}}] Turn this flag on to use a (tilted) \texttt{FORCESOLUTION} force 
+point source located exactly at a grid point instead of a \texttt{CMTSOLUTION} moment-tensor source. 
+When the force source does not fall exactly at a \texttt{GLL} point, the solver interpolates the force between grid points 
+using Lagrange interpolants. This can be useful e.g. for oil industry foothills simulations 
+in which the source is a vertical force, normal force, tilted force, or an impact etc. 
+Note that in the \texttt{FORCESOLUTION} file, you will need to edit
+the East, North and vertical components of an arbitrary (non-unitary)
+direction vector of the force vector; thus refer to
+Appendix~\ref{cha:Coordinates} for the orientation of the reference
+frame. This vector is made unitary internally in the solver and thus
+only its direction matters here; its norm is ignored and the norm of
+the force used is the factor force source times the source time
+function.
+\item [{\texttt{USE\_RICKER\_TIME\_FUNCTION}}] Turn this flag on to 
+use a Ricker source time function instead of the source time functions set by default 
+to represent a (tilted) \texttt{FORCESOLUTION} force point source or a \texttt{CMTSOLUTION} moment-tensor source. 
+Normally, if a \texttt{CMTSOLUTION} moment-tensor source is used, a (pseudo) Heaviside step 
+function with a very short half duration is used for elastic cases to represent the permanent slip on the fault while 
+in the acoustic case a Gaussian source time function with a similarly short half duration is used to physically describe 
+actions within the fluid. Similarly, if a \texttt{FORCESOLUTION} force source is used, we need a source time function 
+with finite support, so a (pseudo) Dirac delta source time function (which is actually a Gaussian of very
+short half duration, i.e. hdur = 5 * DT) is used by default.
+Any other source-time function may then be obtained by convolution.
 \item [{\texttt{PRINT\_SOURCE\_TIME\_FUNCTION}}] Turn this flag on to print
@@
 This feature is only relevant for the solver.
+\item [{\texttt{GPU\_MODE}}] Turn this flag on to use GPUs.
 \end{description}
@@
 in the post-processing (see Section \ref{sec:Process-data-and-syn}).
-\item Set the \texttt{half duration} parameter of the source time function.
+\item Set the \texttt{half duration} parameter of the Ricker source
+  time function when {\texttt{USE\_RICKER\_TIME\_FUNCTION}} is turned
+  on in the main parameter file {\texttt{Par\_file}}. In the case the
+  solver uses a (pseudo) Dirac delta source time function to represent
+  a force point source, a very short \texttt{half duration} parameter
+  (hdur = 5 * DT) is automatically set by default.
 \item Set the latitude, longitude, depth of the source in geographical

### 2012-10-31 doc/USER_MANUAL/manual_SPECFEM3D_Cartesian.tex
updates the user manual

@@
 \item [{\texttt{USE\_FORCE\_POINT\_SOURCE}}] Turn this flag on to use a (tilted) \texttt{FORCESOLUTION} force 
-point source located exactly at a grid point instead of a \texttt{CMTSOLUTION} moment-tensor source. 
-When the force source does not fall exactly at a \texttt{GLL} point, the solver interpolates the force between grid points 
+point source instead of a \texttt{CMTSOLUTION} moment-tensor source. 
+When the force source does not fall exactly at a grid point, the solver interpolates the force between grid points 
 using Lagrange interpolants. This can be useful e.g. for oil industry foothills simulations 
@@
 to represent a (tilted) \texttt{FORCESOLUTION} force point source or a \texttt{CMTSOLUTION} moment-tensor source. 
-Normally, if a \texttt{CMTSOLUTION} moment-tensor source is used, a (pseudo) Heaviside step 
-function with a very short half duration is used for elastic cases to represent the permanent slip on the fault while 
-in the acoustic case a Gaussian source time function with a similarly short half duration is used to physically describe 
-actions within the fluid. Similarly, if a \texttt{FORCESOLUTION} force source is used, we need a source time function 
-with finite support, so a (pseudo) Dirac delta source time function (which is actually a Gaussian of very
-short half duration, i.e. hdur = 5 * DT) is used by default.
-Any other source-time function may then be obtained by convolution.
+Originally, if a \texttt{CMTSOLUTION} moment-tensor source is used, a (pseudo) Heaviside step 
+function with a very short half duration is defined for elastic cases to represent the permanent slip on the fault while 
+in the acoustic case a Gaussian source time function with a similarly short half duration is defined to physically describe 
+actions within the fluid. Otherwise, if a \texttt{FORCESOLUTION} force source is used, a (pseudo) Dirac delta source time 
+function is defined by default. Any other source-time function may then be obtained by convolution.
 \item [{\texttt{PRINT\_SOURCE\_TIME\_FUNCTION}}] Turn this flag on to print
@@
   time function when {\texttt{USE\_RICKER\_TIME\_FUNCTION}} is turned
-  on in the main parameter file {\texttt{Par\_file}}. In the case the
+  on in the main parameter file {\texttt{Par\_file}}. In case that the
   solver uses a (pseudo) Dirac delta source time function to represent

### 2012-11-12 doc/USER_MANUAL/manual_SPECFEM3D_Cartesian.tex
changed a few things in the manual
allows the use of multiple user-defined tomographic files instead of having the same and only "tomography_model.xyz" file, moves the TOMOGRAPHY_PATH flag in Par_file, organises the "done" items of the todo list

@@
 %
-\item [materials\_file] Contains the material associations for each element.
+\item [materials\_file] Contains the material associations for each element. The format is: 
+\begin{lyxcode}
+\texttt{element\_ID} \texttt{material\_ID}
+\end{lyxcode}
+where \texttt{element\_ID} is the element identifier and \texttt{material\_ID} is a unique identifier, positive (for materials) or negative (for tomographic models).
 \item [nodes\_coords\_file] Contains the point locations in Cartesian coordinates of the mesh element corners.
@@
 \begin{lyxcode}
-./bin/xdecompose\_mesh  nparts  input\_directory output\_directory
+./xdecompose\_mesh  nparts  input\_directory output\_directory
 \end{lyxcode}
@@
 want to use an anisotropy model. Please see the file \texttt{model\_aniso.f90} in subdirectory \texttt{src/generate\_databases/} for the current implementation of anisotropic models.
+\item [{\texttt{TOMOGRAPHY\_PATH}}] Directory in which the tomography files are stored for using external 
+tomographic Earth models (please read Chapter \ref{cha:-Changing-the} and Section \ref{sec:Using-tomographic} 
+`Using external tomographic Earth models' for further details.).
 \item [{\texttt{USE\_OLSEN\_ATTENUATION}}] Set to \texttt{.true.} if you
@@
 
-To implement your own external tomographic model, you must provide your own tomography file
-\texttt{tomography\_model.xyz} and choose between two possible options:
-(1) set in the \texttt{Par\_file} the model parameter \texttt{MODEL = tomo}, or
-(2) for more user control, set in the \texttt{Par\_file} the model parameter \texttt{MODEL = default} and use the following format in the file \texttt{MESH/nummaterial\_velocity\_file} when using CUBIT to construct your mesh (see also section \ref{subsec:Exporting-the-Mesh}):
+To implement your own external tomographic model(s), you must provide your own external tomography file(s), 
+and choose between two possible options:\\
+\indent (1) set in the \texttt{Par\_file} the model parameter \texttt{MODEL = tomo}, \\ 
+or for more user control: \\
+\indent (2) set in the \texttt{Par\_file} the model parameter \texttt{MODEL = default}, define the negative \texttt{material\_ID} identifier for each element in the file \texttt{MESH/materials\_file} and use the following format in the file \texttt{MESH/nummaterial\_velocity\_file} when using CUBIT to construct your mesh (see also section \ref{subsec:Exporting-the-Mesh}):
 \begin{lyxcode}
@@
 \end{lyxcode}
-where
-\texttt{domain\_ID} is 1 for acoustic or 2 for elastic materials,
-\texttt{material\_ID} a negative, unique identifier (i.e., -1),
-\texttt{tomography} keyword for tomographic material definition,
-\texttt{elastic} keyword for elastic material definition,
-\texttt{file\_name} the name of the tomography file and
-1 a positive unique identifier.
-The name of the tomography file is not used so far, rather change the filename (\texttt{TOMO\_FILENAME}) defined in the file \texttt{model\_tomography.f90} located in the \texttt{src/generate\_databases/} directory.
+where: \\
+\indent \texttt{domain\_ID} is 1 for acoustic or 2 for elastic materials, \\
+\indent \texttt{material\_ID} a negative, unique identifier (i.e., -1,-2,...), \\
+\indent \texttt{tomography} keyword for tomographic material definition, \\
+\indent \texttt{elastic} keyword for elastic material definition, \\
+\indent \texttt{file\_name} the name of the tomography file and 1 a positive unique identifier.\\
 
 The external tomographic model is represented by a grid of points with assigned material properties and homogeneous resolution along each spatial direction x, y and z.
-The ASCII file \texttt{tomography\_model.xyz} that describes the tomography should be located in the \texttt{DATA/} directory.
-The format of the file, as read from \texttt{model\_tomography.f90}, looks like Figure \ref{fig:tomography_file},
+The ASCII file \texttt{file\_name} that describe the tomography should be located in the \texttt{TOMOGRAPHY\_PATH} directory, set in the \texttt{Par\_file}.
+The format of the file, as read from \texttt{model\_tomography.f90} located in the \texttt{src/generate\_databases} directory, looks like Figure \ref{fig:tomography_file},
 where
@@
 %
-After the first four lines, in the file \texttt{tomography\_model.xyz} the tomographic grid points are listed with the corresponding values of \texttt{vp}, \texttt{vs} and \texttt{rho}, scanning the grid along the x coordinate (from \texttt{ORIG\_X} to \texttt{END\_X} with step of \texttt{SPACING\_X}) for each given y (from \texttt{ORIG\_Y} to \texttt{END\_Y}, with step of \texttt{SPACING\_Y}) and each given z (from \texttt{ORIG\_Z} to \texttt{END\_Z}, with step of \texttt{SPACING\_Z}).
+After the first four lines, in the file \texttt{file\_name} the tomographic grid points are listed with the corresponding values of \texttt{vp}, \texttt{vs} and \texttt{rho}, scanning the grid along the x coordinate (from \texttt{ORIG\_X} to \texttt{END\_X} with step of \texttt{SPACING\_X}) for each given y (from \texttt{ORIG\_Y} to \texttt{END\_Y}, with step of \texttt{SPACING\_Y}) and each given z (from \texttt{ORIG\_Z} to \texttt{END\_Z}, with step of \texttt{SPACING\_Z}).
 %fig
@@
 \par\end{centering}
-\caption{Tomography file \texttt{tomography\_model.xyz} that describes an external Earth model and that is read by \texttt{model\_tomography.f90}.
+\caption{Tomography file \texttt{file\_name} that describes an external Earth model.
 The coordinates x, y and z of the grid, their limits ORIG\_X, ORIG\_Y, ORIG\_Z, END\_X, END\_Y, END\_Z and the grid spacings SPACING\_X, SPACING\_Y, SPACING\_Z should be in the units of the constructed mesh (e.g., UTM coordinates in meters).}

### 2012-11-12 src/decompose_mesh/decompose_mesh.F90

@@ -144,7 +144,7 @@ module decompose_mesh
             OCEANS,TOPOGRAPHY,USE_FORCE_POINT_SOURCE
   logical :: ABSORBING_CONDITIONS,SAVE_FORWARD,ABSORB_INSTEAD_OF_FREE_SURFACE
   logical :: ANISOTROPY,SAVE_MESH_FILES,USE_RICKER_TIME_FUNCTION,PRINT_SOURCE_TIME_FUNCTION
-  character(len=256) LOCAL_PATH
+  character(len=256) LOCAL_PATH,TOMOGRAPHY_PATH
   integer :: IMODEL
 
   contains
@@ -243,10 +243,6 @@ module decompose_mesh
     enddo
     close(98)
 
-  ! TODO:
-  ! must be changed, if  mat(1,i) < 0  1 == interface , 2 == tomography
-    mat(2,:) = 1
-
   ! reads material definitions
   !
   ! note: format of nummaterial_velocity_file must be
@@ -255,12 +251,12 @@ module decompose_mesh
   !
   ! where
   !     material_domain_id : 1=acoustic / 2=elastic / 3=poroelastic
-  !     material_id               : number of material/volume
-  !     rho                           : density
-  !     vp                             : P-velocity
-  !     vs                             : S-velocity
-  !     Q_mu                      : 0=no attenuation
-  !     anisotropy_flag        : 0=no anisotropy/ 1,2,.. check with implementation in aniso_model.f90
+  !     material_id        : number of material/volume
+  !     rho                : density
+  !     vp                 : P-velocity
+  !     vs                 : S-velocity
+  !     Q_mu               : 0=no attenuation
+  !     anisotropy_flag    : 0=no anisotropy/ 1,2,.. check with implementation in aniso_model.f90
   ! Note that when poroelastic material, this file is a dummy except for material_domain_id & material_id,
   ! and that poroelastic materials are actually read from nummaterial_poroelastic_file, because CUBIT
   ! cannot support more than 10 attributes
@@ -340,7 +336,7 @@ module decompose_mesh
        ! format: note that we save the arguments in a slightly different order in mat_prop(:,:)
        !              #(6) material_domain_id #(0) material_id  #(1) rho #(2) vp #(3) vs #(4) Q_mu #(5) anisotropy_flag
        !
-       ! reads lines unti it reaches a defined material
+       ! reads lines until it reaches a defined material
        num_mat = -1
        do while( num_mat < 0 .and. ier == 0)
[…]

### 2012-11-15 doc/USER_MANUAL/manual_SPECFEM3D_Cartesian.tex
adds the description of NGNOD in the user manuel; updates Makefile.in

@@
 Please see also Section~\ref{sec:Choosing-the-Time-Step} for further details.
-%
+\item [{\texttt{NGNOD}}] The number of nodes for 2D and 3D shape functions for hexahedra. We use either 
+8-node mesh elements (bricks) or 27-node elements. If you use the internal mesher, the only option is 8-node bricks 
+(27-node elements are not supported). \texttt{CUBIT} does not support HEX27 elements either (it can generate them, 
+but they are flat, i.e. identical to HEX8). To generate HEX27 elements with curvature properly taken into account, 
+you can use \texttt{Gmsh} \url{http://geuz.org/gmsh/}
 \item [{\texttt{MODEL}}] Must be set to one of the following:

### 2013-01-16 doc/USER_MANUAL/manual_SPECFEM3D_Cartesian.tex
added source version of the workflow figure
adds a brief description of C-PML mesh layers created by CUBIT and the related configuration flags in Par_file

@@
 \begin{description}
-\item [nummaterial\_velocity\_file] Defines the material properties. For fully defined materials, the formats is:
+\item [absorbing\_cpml\_file] {\bf (only in case of C-PML absorbing conditions)} Contains at the first line respectively the "number of C-PML spectral elements"
+and the "thickness of C-PML layers" (currently, a common thickness for all faces is defined but this can be easily adapted in the future if needed by the setting of CPML\_width\_x, CPML\_width\_y, CPML\_width\_z parameters).
+On all following lines are defined the "C-PML spectral elements global indexing" and "C-PML regions" for each
+C-PML element according to the following convention:
+
+\begin{itemize}
+\item \#id\_cpml\_regions = 1 : X\_surface C-PML,
+\item \#id\_cpml\_regions = 2 : Y\_surface C-PML,
+\item \#id\_cpml\_regions = 3 : Z\_surface C-PML,
+\item \#id\_cpml\_regions = 4 : XY\_edge C-PML,
+\item \#id\_cpml\_regions = 5 : XZ\_edge C-PML,
+\item \#id\_cpml\_regions = 6 : YZ\_edge C-PML,
+\item \#id\_cpml\_regions = 7 : XYZ\_corner C-PML.
+\end{itemize}
+
+\item [materials\_file] Contains the material associations for each element. The format is: 
+\begin{lyxcode}
+\texttt{element\_ID} \texttt{material\_ID}
+\end{lyxcode}
+where \texttt{element\_ID} is the element identifier and \texttt{material\_ID} is a unique identifier, positive (for materials) or negative (for tomographic models or C-PML elements).
+For C-PML absorbing conditions, \texttt{material\_ID} is a unique negative identifier between -2001 and -2007 in correspondance to the convention seen above in the description of the 
+absorbing\_cpml\_file: -2001 <-> 1, -2002 <-> 2, etc\ldots
+
+\item [nummaterial\_velocity\_file] Defines the material properties. 
+\begin{itemize}
+\item For fully defined materials, the format is:
 \begin{lyxcode}
@@
 \end{lyxcode}
-where
-\texttt{domain\_ID} is 1 for acoustic or 2 for elastic materials, \texttt{material\_ID} a unique identifier,  \texttt{rho} the density in $kg \, m^{-3}$,  \texttt{vp} the P-wave speed in $m \, s^{-1}$, \texttt{vs}  the S-wave speed in $m \, s^{-1}$, \texttt{Q} the quality factor and \texttt{anisotropy\_flag} an identifier for anisotropic models.
-
+where \texttt{domain\_ID} is 1 for acoustic or 2 for elastic materials, \texttt{material\_ID} a unique identifier,  \texttt{rho} the density in $kg \, m^{-3}$,  \texttt{vp} the P-wave speed in $m \, s^{-1}$, \texttt{vs}  the S-wave speed in $m \, s^{-1}$, \texttt{Q} the quality factor and \texttt{anisotropy\_flag} an identifier for anisotropic models.
+
+\item For C-PML absorbing conditions, , the format is:
+\begin{lyxcode}
+\texttt{domain\_ID} \texttt{material\_ID}  \texttt{rho}  \texttt{vp}  \texttt{vs}
+\end{lyxcode}
+where \texttt{domain\_ID} is 1 for acoustic or 2 for elastic materials, \texttt{material\_ID} a unique identifier between -2001 and -2007 to correspond with the definition given in materials\_file,  \texttt{rho} the density in $kg \, m^{-3}$,  \texttt{vp} the P-wave speed in $m \, s^{-1}$, \texttt{vs}  the S-wave speed in $m \, s^{-1}$
 %%magnoni
-For tomographic velocity models, please read Chapter \ref{cha:-Changing-the} and Section \ref{sec:Using-tomographic} `Using external tomographic Earth models' for further details.
+\item For tomographic velocity models, please read Chapter \ref{cha:-Changing-the} and Section \ref{sec:Using-tomographic} `Using external tomographic Earth models' for further details. 
 %
@@
 %
-\item [materials\_file] Contains the material associations for each element. The format is: 
-\begin{lyxcode}
-\texttt{element\_ID} \texttt{material\_ID}
-\end{lyxcode}
-where \texttt{element\_ID} is the element identifier and \texttt{material\_ID} is a unique identifier, positive (for materials) or negative (for tomographic models).
+\end{itemize}
+
 \item [nodes\_coords\_file] Contains the point locations in Cartesian coordinates of the mesh element corners.
@@
 that is, list all the element faces that constitute the surface of the model in that file.
-\item [absorbing\_surface\_file\_xmax] Contains the surface connectivity of the absorbing boundary surface at the Xmax.
-\item [absorbing\_surface\_file\_xmin] Contains the surface connectivity of the absorbing boundary surface at the Xmin.
-\item [absorbing\_surface\_file\_ymax] Contains the surface connectivity of the absorbing boundary surface at the Ymax.
-\item [absorbing\_surface\_file\_ymin] Contains the surface connectivity of the absorbing boundary surface at the Ymin.
-\item [absorbing\_surface\_file\_bottom] Contains the surface connectivity of the absorbing boundary surface at the bottom.
+\item [absorbing\_surface\_file\_xmax] Contains the surface connectivity of the absorbing boundary surface at the Xmax \\(irrelevant in case of C-PML absorbing conditions).
+\item [absorbing\_surface\_file\_xmin] Contains the surface connectivity of the absorbing boundary surface at the Xmin \\(irrelevant in case of C-PML absorbing conditions).
+\item [absorbing\_surface\_file\_ymax] Contains the surface connectivity of the absorbing boundary surface at the Ymax \\(irrelevant in case of C-PML absorbing conditions).
+\item [absorbing\_surface\_file\_ymin] Contains the surface connectivity of the absorbing boundary surface at the Ymin \\(irrelevant in case of C-PML absorbing conditions).
+\item [absorbing\_surface\_file\_bottom] Contains the surface connectivity of the absorbing boundary surface at the bottom \\(irrelevant in case of C-PML absorbing conditions).
 \end{description}
@@
 \noindent \begin{centering}
-\includegraphics[width=.6\textwidth]{figures/workflow.jpg}
+\includegraphics[width=.6\textwidth]{figures/workflow.pdf}
 \par\end{centering}
@@
 absorbing boundary conditions on the top surface which by default constitutes a free surface of the model.
+\item [{\texttt{PML\_CONDITIONS}}] Set to \texttt{.true.} to turn on C-PML boundary conditions for a regional simulation. Both fluids and elastic solids are supported.
+\item [{\texttt{PML\_INSTEAD\_OF\_FREE\_SURFACE}}] Set to \texttt{.true.} to turn on C-PML boundary conditions on the top surface which by default constitutes a free surface of the model.
+\item [{\texttt{PML\_WIDTH\_MIN}}] Gives a min. authorized bound to the solver to target an optimized thickness of C-PML layers. 
+\item [{\texttt{PML\_WIDTH\_MAX}}] Gives a max. authorized bound to the solver to target an optimized thickness of C-PML layers. 
+\item [{\texttt{f0\_FOR\_PML}}] Determines the dominant frequency that will be used in the calculation of: \\ ALPHA\_MAX\_PML = PI*f0\_FOR\_PML.
 \item [{\texttt{CREATE\_SHAKEMAP}}] Set this flag to \texttt{.true.} to

### 2013-01-17 doc/USER_MANUAL/manual_SPECFEM3D_Cartesian.tex
removed FLAGS_NO_CHECK from the manual

@@
 MPI libraries.
-\item [{\texttt{FLAGS\_CHECK}}] Compiler flag for non-critical subroutines.
-\item [{\texttt{FLAGS\_NO\_CHECK}}] Compiler flag for creating fast, production-run
-code for critical subroutines.
+\item [{\texttt{FLAGS\_CHECK}}] Compiler flags.
 \end{description}
@@
 \noindent
-Edit file \texttt{flags.guess} and put this for \texttt{FLAGS\_CHECK} and \texttt{FLAGS\_NO\_CHECK}:\\
+Edit file \texttt{flags.guess} and put this for \texttt{FLAGS\_CHECK}:\\
 
@@
 \noindent
-\texttt{   FLAGS\_NO\_CHECK="-O3 -qsave -qstrict -qtune=auto -qarch=450d -qcache=auto \\}
+\texttt{   FLAGS\_CHECK="-O3 -qsave -qstrict -qtune=auto -qarch=450d -qcache=auto \\}
 \texttt{   -qfree=f90 -qsuffix=f=f90 -g -qlanglvl=95pure -qhalt=w -Q -Q+rank,swap\_all -Wl,-relax"}

### 2013-01-18 doc/USER_MANUAL/manual_SPECFEM3D_Cartesian.tex
switched to Fortran2003 in the manual

@@
 
-All SPECFEM3D Cartesian software is written in Fortran90 with full portability
-in mind, and conforms strictly to the Fortran95 standard. It uses
-no obsolete or obsolescent features of Fortran77. The package uses
+All SPECFEM2D software is written in Fortran2003 with full portability
+in mind, and conforms strictly to the Fortran2003 standard. It uses
+no obsolete or obsolescent features of Fortran. The package uses
 parallel programming based upon the Message Passing Interface (MPI)
@@
 a minimum, it is recommended that you explicitly specify the appropriate
-command names for your Fortran90 compiler and MPI package:
+command names for your Fortran compiler and MPI package:
 
@@
 \begin{description}
-\item [{\texttt{F90}}] Path to the Fortran90 compiler.
-\item [{\texttt{MPIF90}}] Path to MPI Fortran90.
+\item [{\texttt{F90}}] Path to the Fortran compiler.
+\item [{\texttt{MPIF90}}] Path to MPI Fortran.
 \item [{\texttt{MPI\_FLAGS}}] Some systems require this flag to link to
@@
    Examine the log file 'config.log'. It contains detailed informations.
-   In many cases, the path's to these specific compiler commands F90,
-   CC and MPIF90 won't be correct if `./configure` fails.
+   In many cases, the paths to these specific compiler commands F90,
+   CC and MPIF90 will not be correct if `./configure` fails.
 
@@
 
-  Normally, this message will appear when you're mixing two different Fortran
+  Normally, this message will appear when you are mixing two different Fortran
   compilers. That is, using e.g. gfortran to compile non-MPI files

### 2013-02-11 doc/USER_MANUAL/manual_SPECFEM3D_Cartesian.tex
fixed a typo
added new contributors to the list of authors

@@
 Since then it has been developed and maintained by a development team: in alphabetical order,
+Jean-Paul (Pablo) Ampuero,
 Piero Basini,
@@
 Min Chen,
+Percy Galvez,
 Dominik G\"oddeke,
@@
 David Mich\'ea,
+Surendra Nadh Somala,
 Tarje Nissen-Meyer,
@@
 Max Rietmann,
+Elliott Sales de Andrade,
 Brian Savage,
@@
 
-All SPECFEM2D software is written in Fortran2003 with full portability
+All SPECFEM3D software is written in Fortran2003 with full portability
 in mind, and conforms strictly to the Fortran2003 standard. It uses
@@
 In such cases, we consider SEM seismograms as broad band, ignoring the corner period ($\geq 10$ s) of the response band of instruments (note that at these resolutions, the minimum period in the SEM synthetics will be less than $10$ s).
-Accordingly, when you run a simulation the band code will be chosen depending on the resolution of the synthetics, and channel codes of SEM seismograms will start with either \texttt{L}, \texttt{M}, \texttt{B}, \texttt{H}, \texttt{C} or \texttt{F}, shown by red color in the figure.\\
+Accordingly, when
+you run a simulation the band code will be chosen depending on the resolution of the synthetics, and channel codes of SEM seismograms will start with either \texttt{L}, \texttt{M}, \texttt{B}, \texttt{H}, \texttt{C} or \texttt{F}, shown by red color in the figure.\\
 

### 2013-03-04 doc/USER_MANUAL/manual_SPECFEM3D_Cartesian.tex
more explanations about how to use PML with an external tomographic velocity and density model

@@
 
+If you use PML and an external tomographic velocity and density model,
+you should be careful because mathematically a PML cannot handle heterogeneities along the
+normal to the PML edge inside the PML layer. This comes from the fact that the damping profile
+that is defined assumes a constant velocity and density model along the normal
+direction.
+
+Thus, you need to modify your velocity and density model in order for it to be 1D inside
+the PML, as shown in Figure~\ref{fig:modify_external_velocity_model_to_use_PML}.
+
+This applies to the bottom layer as well; there you should make sure
+that your model is 1D and thus constant along the vertical direction.
+
+To summarize, only use a 3D velocity and density model inside the physical region, and in
+all the PML layers extend it by continuity from its values along the
+inner PML edge.
+
+%%
+\begin{figure}[htbp]
+\noindent \begin{centering}
+\includegraphics[width=3in]{figures/how_to_use_PML_when_a_tomographic_velocity_model_is_used.png}
+\par\end{centering}
+\caption{How to modify your external 3D velocity and density model in order to use PML.
+Such a modification is not needed when using Stacey absorbing boundary conditions (but such conditions
+are significantly less efficient).}
+\label{fig:modify_external_velocity_model_to_use_PML}
+\end{figure}
+%%
+
 \section{\label{sec:Choosing-the-Time-Step}Choosing the time step \texttt{DT}}

### 2013-03-11 doc/Version_Control/Git-SVN_Quick_Reference.tex
Explanations on how to use Git in addition to SVN. Dirty and elementary. 	new file:   doc/Version_Control/Git-SVN_Quick_Reference.tex

\documentclass[11pt,twoside]{article}

\begin{document}

\title{GIT-SVN basic commands}
\maketitle

\subsection*{Important remark}

It might be useful to issue some git-svn commands beforehand with the \verb|-n| option (or with its longer name \verb|--dry-run|).
This option shows what will be impacted by the \texttt{git-svn} command before issuing the actual command.
This can be used with the \texttt{dcommit}, \texttt{rebase}, \texttt{branch} and \texttt{tag} commands.

\subsection*{Git-svn -- Initial checkout}
\begin{verbatim}
$ git svn clone -s \
  svn+ssh://svn@geodynamics.org/cig/seismo/3D/SPECFEM3D_GLOBE
\end{verbatim}
The previous command checkouts the full svn history and converts it into a git history. It takes a huge amount of time. If the full history is not needed, one can checkout a partial history.
For instance, checking out revisions from $20\:000$ to the last one (\texttt{HEAD}) is done with:
\begin{verbatim}
$ git svn clone -s -r 20000:HEAD \
  svn+ssh://svn@geodynamics.org/cig/seismo/3D/SPECFEM3D\_GLOBE
\end{verbatim}

\subsection*{Git-svn -- SVN branches}

Taking a look at the different branches available:
\begin{verbatim}
$ git branch -a
* master
  remotes/NOISE_TOMOGRAPHY
  remotes/ORIGINAL
  remotes/SPECFEM3D_GLOBE_SUNFLOWER
  remotes/adjoint
  remotes/pluggable
  remotes/tags/v5.1.5
  remotes/trunk
\end{verbatim}

Creating a new branch \texttt{SPECFEM3D\_GLOBE\_ADIOS} on the svn server:
\begin{verbatim}
$ git svn branch SPECFEM3D_GLOBE_ADIOS
\end{verbatim}

Switching to a remote branch. It is mandatory, else the commit will be done in the master branch
\begin{verbatim}
$ git checkout -b local/SPECFEM3D_GLOBE_ADIOS \
  SPECFEM3D_GLOBE_ADIOS
\end{verbatim}

\subsection*{Git -- Checking the status of the file.}
Before doing any modification it is mandatory to know which files are tracked, have been modified or deleted.

\begin{verbatim}
$ git status
\end{verbatim}

If some files appear to be untracked they can be added to the local git repository with:

\begin{verbatim}
$ git add filename
\end{verbatim}

\begin{verbatim}
$ git status 
# should now return the added file
# as 'new file' unstead of 'untracked'.
\end{verbatim}

If some files were modified, before committing the modifications, it may be useful to see what has changed :

\begin{verbatim}
$ git diff filename
\end{verbatim}

\subsection*{Git -- Committing changes into the local repository}

\begin{verbatim}
$ git commit
\end{verbatim}

This command will open your favorite text editor (it is likely to be vi without any particular configuration) and ask you for a comment about the commit. If nothing is written, the commit will abort.

\subsection*{Git -- Removing files}

\begin{verbatim}
$ rm filename # remove the file from your workspace.
\end{verbatim}
\begin{verbatim}
$ git rm filename 
# prepare to remove the file from the git local repository.
\end{verbatim}
\begin{verbatim}
$ git commit #... 
\end{verbatim}

\subsection*{Git -- Moving files}

\begin{verbatim}
$ git mv old_file new_file
\end{verbatim}
\begin{verbatim}
$ git commit
\end{verbatim}

OR

\begin{verbatim}
$ mv old_file new_file
\end{verbatim}
\begin{verbatim}
$ git rm old_file
\end{verbatim}
\begin{verbatim}
$ git add new_file
\end{verbatim}
\begin{verbatim}
$ git commit
\end{verbatim}

\subsection*{Git -- Commit history}

\begin{verbatim}
$ git log
\end{verbatim}

\subsection*{Git -- Changing the last commit}

\begin{verbatim}
$ git commit -m 'initial commit'
\end{verbatim}
\begin{verbatim}
$ git add forgotten_file
\end{verbatim}
\begin{verbatim}
$ git commit --amend
\end{verbatim}
\subsection*{Git -- Unstaging a staged file (i.e. Removing a file from the list of files to be 
committed)}

\begin{verbatim}
$ git status 
# the file has to be in the 'changes to be committed' section
\end{verbatim}

\begin{verbatim}
$ git reset HEAD filename
filename: locally modified
\end{verbatim}
\begin{verbatim}
$ git status  
# the file should be in the 'Changes not staged for commit' section
\end{verbatim}

\subsection*{Git -- Unmodifying a Modified File}

\begin{verbatim}
$ git checkout -- filename
\end{verbatim}
\begin{verbatim}
$ git status
\end{verbatim}

\subsection*{Git-svn -- Back to the svn server}

Push modifications up to the svn server:
\begin{verbatim}
$ git svn dcommit
\end{verbatim}
That will also bring your local tree up to date. To bring your tree up to date in general, run: 
\begin{verbatim}
$ git svn rebase
\end{verbatim}
This will update your local checkout and then re-apply your local un-submitted commits on top of the new trunk. 
If you want to get the commits for all branches that exist in your clone: 
\begin{verbatim}
$ git svn fetch
\end{verbatim}

\subsection*{References and usefull readings}

\verb|http://git-scm.com/book|
\\
\verb|http://git.or.cz/course/svn.html|
\\
\verb|http://trac.parrot.org/parrot/wiki/git-svn-tutorial|

\end{document}

### 2013-03-13 src/shared/combine_vol_data.f90
replaced the OCEANS flag with APPROXIMATE_OCEAN_LOAD to avoid confusion

@@ -97,7 +97,7 @@
   logical :: MOVIE_SURFACE,MOVIE_VOLUME,CREATE_SHAKEMAP,SAVE_DISPLACEMENT, &
             USE_HIGHRES_FOR_MOVIES,SUPPRESS_UTM_PROJECTION
   logical :: ATTENUATION,USE_OLSEN_ATTENUATION, &
-            OCEANS,TOPOGRAPHY,USE_FORCE_POINT_SOURCE
+            APPROXIMATE_OCEAN_LOAD,TOPOGRAPHY,USE_FORCE_POINT_SOURCE
   logical :: ABSORBING_CONDITIONS,SAVE_FORWARD,ABSORB_INSTEAD_OF_FREE_SURFACE
   logical :: ANISOTROPY,SAVE_MESH_FILES,USE_RICKER_TIME_FUNCTION,PRINT_SOURCE_TIME_FUNCTION
   logical :: PML_CONDITIONS,PML_INSTEAD_OF_FREE_SURFACE
@@ -177,7 +177,7 @@
   call read_parameter_file( NPROC,NTSTEP_BETWEEN_OUTPUT_SEISMOS,NSTEP,DT,NGNOD,NGNOD2D, &
                         UTM_PROJECTION_ZONE,SUPPRESS_UTM_PROJECTION,TOMOGRAPHY_PATH, &
                         ATTENUATION,USE_OLSEN_ATTENUATION,LOCAL_PATH,NSOURCES, &
-                        OCEANS,TOPOGRAPHY,ANISOTROPY,ABSORBING_CONDITIONS,MOVIE_TYPE, &
+                        APPROXIMATE_OCEAN_LOAD,TOPOGRAPHY,ANISOTROPY,ABSORBING_CONDITIONS,MOVIE_TYPE, &
                         MOVIE_SURFACE,MOVIE_VOLUME,CREATE_SHAKEMAP,SAVE_DISPLACEMENT, &
                         NTSTEP_BETWEEN_FRAMES,USE_HIGHRES_FOR_MOVIES,HDUR_MOVIE, &
                         SAVE_MESH_FILES,PRINT_SOURCE_TIME_FUNCTION, &

### 2013-03-13 src/specfem3D/iterate_time.f90

@@ -817,7 +817,7 @@
                               SAVE_FORWARD, &
                               ACOUSTIC_SIMULATION,ELASTIC_SIMULATION, &
                               ABSORBING_CONDITIONS,NOISE_TOMOGRAPHY,COMPUTE_AND_STORE_STRAIN, &
-                              ATTENUATION,ANISOTROPY,OCEANS, &
+                              ATTENUATION,ANISOTROPY,APPROXIMATE_OCEAN_LOAD, &
                               APPROXIMATE_HESS_KL)
 
   end subroutine it_transfer_from_GPU

### 2013-03-13 doc/USER_MANUAL/manual_SPECFEM3D_Cartesian.tex
replaced the OCEANS flag with APPROXIMATE_OCEAN_LOAD to avoid confusion

@@
 %
-\item [{\texttt{OCEANS}}] Set to \texttt{.true.} if the effect of the oceans
+\item [{\texttt{APPROXIMATE\_OCEAN\_LOAD}}] Set to \texttt{.true.} if the effect of the oceans
 on seismic wave propagation should be incorporated based upon the
-approximate treatment discussed in \citet{KoTr02b}. This feature
+(rough) approximate treatment discussed in \citet{KoTr02b}. This feature
 is inexpensive from a numerical perspective, both in terms of memory
@@
 of roughly 20~s and longer. At shorter periods the effect of water
-phases/reverberations is not taken into account, even when the flag
-is on.
-\item [{\texttt{TOPOGRAPHY}}] This feature is only effective if \texttt{OCEANS} is set to \texttt{.true.}.
+phases/reverberations is not taken into account, even when the flag is on.
+\item [{\texttt{TOPOGRAPHY}}] This feature is only effective if \texttt{APPROXIMATE\_OCEAN\_LOAD} is set to \texttt{.true.}.
 Set to \texttt{.true.} if topography and

### 2013-03-13 src/specfem3D/initialize_simulation.f90

@@ -42,7 +42,7 @@
   call read_parameter_file( NPROC,NTSTEP_BETWEEN_OUTPUT_SEISMOS,NSTEP,DT,NGNOD,NGNOD2D, &
                         UTM_PROJECTION_ZONE,SUPPRESS_UTM_PROJECTION,TOMOGRAPHY_PATH, &
                         ATTENUATION,USE_OLSEN_ATTENUATION,LOCAL_PATH,NSOURCES, &
-                        OCEANS,TOPOGRAPHY,ANISOTROPY,ABSORBING_CONDITIONS,MOVIE_TYPE, &
+                        APPROXIMATE_OCEAN_LOAD,TOPOGRAPHY,ANISOTROPY,ABSORBING_CONDITIONS,MOVIE_TYPE, &
                         MOVIE_SURFACE,MOVIE_VOLUME,CREATE_SHAKEMAP,SAVE_DISPLACEMENT, &
                         NTSTEP_BETWEEN_FRAMES,USE_HIGHRES_FOR_MOVIES,HDUR_MOVIE, &
                         SAVE_MESH_FILES,PRINT_SOURCE_TIME_FUNCTION, &

### 2013-03-13 doc/USER_MANUAL/manual_SPECFEM3D_Cartesian.tex
added more details about SIMULATION_TYPE == 2

@@
 
-This new release of the code (version 2.1) includes support for GPU graphics card acceleration \citep{KoMiEr09,KoGoErMi10,KoErGoMi10,MiKo10,Kom11}.\\
+This new release of the code (version 2.1) includes support for GPU graphics card acceleration \citep{KoMiEr09,KoErGoMi10,MiKo10,Kom11}.\\
 
@@
 \item [Numerical simulations in general] ~\\
-Forward and adjoint simulations are described in detail in \cite{TrKoLi08,PeKoLuMaLeCaLeMaLiBlNiBaTr11,VaCaSaKoVi99,KoMiEr09,KoErGoMi10,KoGoErMi10,ChKoViCaVaFe07,MaKoDi09,KoViCh10,CaKoLaTiMiLeSnTr08,TrKoHjLiZhPeBoMcFrTrHu10,KoRiTr02,KoTr02a,KoTr02b,KoTr99} or \cite{KoVi98}.
+Forward and adjoint simulations are described in detail in \cite{TrKoLi08,PeKoLuMaLeCaLeMaLiBlNiBaTr11,VaCaSaKoVi99,KoMiEr09,KoErGoMi10,ChKoViCaVaFe07,MaKoDi09,KoViCh10,CaKoLaTiMiLeSnTr08,TrKoHjLiZhPeBoMcFrTrHu10,KoRiTr02,KoTr02a,KoTr02b,KoTr99} or \cite{KoVi98}.
 Additional aspects of adjoint simulations are described in
@@
 Domain decomposition is explained in detail in \cite{MaKoBlLe08}, and excellent scaling up to 150,000 processor cores is shown for instance in
-\cite{CaKoLaTiMiLeSnTr08,KoLaMi08a,MaKoBlLe08,KoErGoMi10,KoGoErMi10,Kom11},
+\cite{CaKoLaTiMiLeSnTr08,KoLaMi08a,MaKoBlLe08,KoErGoMi10,Kom11},
 
@@
 Computing on GPU graphics cards for acoustic or seismic wave propagation applications
-is described in detail in \cite{KoMiEr09,KoGoErMi10,KoErGoMi10,MiKo10,Kom11}.
+is described in detail in \cite{KoMiEr09,KoErGoMi10,MiKo10,Kom11}.
 
@@
 If you use this new version 2.1, which has non blocking MPI for much better performance for medium or large runs, please cite at least one of these six articles,
-in which results of non blocking MPI runs are presented: \cite{PeKoLuMaLeCaLeMaLiBlNiBaTr11,KoGoErMi10,KoErGoMi10,KoViCh10,Kom11,CaKoLaTiMiLeSnTr08,MaKoBlLe08}.\\
+in which results of non blocking MPI runs are presented: \cite{PeKoLuMaLeCaLeMaLiBlNiBaTr11,KoErGoMi10,KoViCh10,Kom11,CaKoLaTiMiLeSnTr08,MaKoBlLe08}.\\
 
@@
 Domain decomposition is explained in detail in \cite{MaKoBlLe08}, and excellent scaling up to 150,000 processor cores in shown for instance in
-\cite{CaKoLaTiMiLeSnTr08,KoLaMi08a,MaKoBlLe08,KoErGoMi10,KoGoErMi10,Kom11}.}
+\cite{CaKoLaTiMiLeSnTr08,KoLaMi08a,MaKoBlLe08,KoErGoMi10,Kom11}.}
 
@@
 parameters in the \texttt{Par\_file} are \texttt{SIMULATION\_TYPE}
-(1 or 2) and the \texttt{SAVE\_FORWARD} (boolean).
-
-\section{\label{sec:Adjoint-simulation-sources}Adjoint Simulations for Sources}
+(1, 2 or 3) and the \texttt{SAVE\_FORWARD} (boolean).
+
+\section{\label{sec:Adjoint-simulation-sources}Adjoint Simulations for Sources Only (not for the Model)}
 

### 2013-03-13 doc/USER_MANUAL/manual_SPECFEM3D_Cartesian.tex
explained what is done if PML elements are viscoelastic or poroelastic

@@
 The next release of the code will include support Convolution or Auxiliary Differential Equation Perfectly Matched absorbing Layers
-(C-PML or ADE-PML) \citep{KoMa07,MaKoEz08,MaKoGe08,MaKo09,MaKoGeBr10}.
+(C-PML or ADE-PML) \citep{MaKoEz08,MaKoGe08,MaKo09,MaKoGeBr10,KoMa07}.
 It will also use the PT-SCOTCH parallel and threaded version of SCOTCH for mesh partitioning instead of the serial version.\\
@@
 
+If you use PML, the mesh elements that belong to the PML layers can be acoustic or elastic, but not viscoelastic nor poroelastic.
+Then, when defining your model, you should define these absorbing elements as either acoustic or elastic.
+In you forget to do that, the code will fix the problem by automatically converting the viscoelastic or poroelastic PML
+elements to elastic. This means that strictly speaking the PML layer will not be perfectly matched any more, since the physical
+model will change from viscoelastic or poroelastic to elastic at the entrance of the PML, but in practice this is sufficient and
+produces only tiny / negligible spurious reflections.
+
 If you use PML and an external tomographic velocity and density model,

### 2013-03-14 doc/USER_MANUAL/manual_SPECFEM3D_Cartesian.tex
Added fault kinematics & dynamics as separate chapter to manual

@@
 
+\chapter{\label{cha:Fault-Kinematics-Dynamics}Fault Kinematics \& Dynamics}
+
+Fault is a discontinuity in the continuum that needs special treatment
+with two nodes at the same locations. These are called split nodes
+that allow to capture deformation on the fault plane termed slip.
+Fault motion determines the deformation of the rest of the bulk. This
+fault motion can be of two types either physics based spontaneous
+rupture propagation or a prescribed rupture description. The former
+is known as fault dynamics while the later is fault kinematics. We
+describe here first the common features to both fault dynamics and
+kinematics and then give detailed description of parameters required
+for each of them separately.
+
+Input files : Two additional input files are required in \texttt{DATA}
+directory for faults with split nodes\\
+
+\texttt{\textbf{Par\_file\_faults}}\texttt{ - }specialized parameters
+required by fault modules (\emph{if this file doesn't exist, code
+assumes that there are no faults})
+
+\texttt{\textbf{FAULT\_STATIONS}}\texttt{ - }stations on the fault
+where timeseries output of slip, sliprate and traction is required
+\texttt{}~\\
+
+Currently faults can only be run with \texttt{xdecompose\_mesh }and\texttt{
+CUBIT. xmeshfem3D }isn't yet ready to handle faults.\\
+
+Important module related to fault :
+
+Decomposing mesh stage : $\mathtt{src/decompose\_mesh\_SCOTCH/fault\_scotch.f90}$
+
+Generating databases stage : $\mathtt{src/generate\_databases/fault\_generate\_databases.f90}$
+
+Solver stage : $\mathtt{src/specfem3D/fault\_solver\_common.f90}$
+
+{\small \qquad{}}$\mathtt{src/specfem3D/fault\_solver\_kinematic.f90}$
+
+{\small \qquad{}}$\mathtt{src/specfem3D/fault\_solver\_dynamic.f90}$\\
+
+The features and the format of inputs and outputs of fault routines
+are subject to change.\\
+
+\section{Mesh Generation with Split Nodes}
+
+Faults need to be handled in a special way during mesh generation.
+A fault surface must lie at the interface between elements (the mesh
+must honor the fault surfaces). Moreover, a fault is made of two surfaces
+in contact. Each of these two surfaces needs a separate set of nodes.
+This approach is known as \textquotedbl{}split nodes\textquotedbl{}.
+To facilitate the mesh generation with split nodes in Cubit, we need
+to separate the two fault surfaces by a small distance, effectively
+creating a tiny opening of the fault. Note that only the interior
+of the fault must be opened, its edges must remain closed (except
+the edge on the free surface). The fault is automatically closed later
+by SPECFEM3D.\\
+
+Here is an example Cubit script to generate a mesh with split nodes
+for a buried vertical strike-slip fault:\\
+reset \\
+brick x 10 y 10 z 10 \\
+webcut volume all with plane xplane \\
+webcut volume all with plane yplane \\
+webcut volume all with plane xplane offset 3 \\
+webcut volume all with plane zplane offset 3 \\
+webcut volume all with plane zplane offset -3 \\
+imprint all \\
+merge all \\
+unmerge surf 160 \\
+mesh vol all \\
+set node constraint off \\
+node in surf 168 move X 0 Y 0.01 Z 0 \\
+node in surf 160 move X 0 Y -0.01 Z 0\\
+
+\begin{figure}[htbp]
+\noindent \begin{centering}
+\includegraphics[scale=0.3]{\lyxdot \lyxdot /src/MAIN_TRUNK/SPECFEM3D_Cartesian/doc/USER_MANUAL/figures/faultmesh}
+\\
+\includegraphics[scale=0.3]{\lyxdot \lyxdot /src/MAIN_TRUNK/SPECFEM3D_Cartesian/doc/USER_MANUAL/figures/surf168}
+\includegraphics[scale=0.3]{\lyxdot \lyxdot /src/MAIN_TRUNK/SPECFEM3D_Cartesian/doc/USER_MANUAL/figures/splitnodes} 
+\par\end{centering}
+
+\caption{Screenshots of the CUBIT example for embedded fault with split nodes:
+entire mesh decomposed into several volumes bounding the fault (top),
+fault surface 168 is selected (bottom-left) and split nodes of the
+fault illustrating only interior nodes of fault are split while those
+on the edges of fault surface are touching each other (bottom-right).}
+
+\label{fig:examples.splitnodes} 
+\end{figure}
+
+The Cubit scripts ({*}.jou and {*}.py) in the directory $\mathtt{EXAMPLES}$
+generate more complicated meshes. The {*}.py files are Python scripts
+that execute Cubit commands and use the Cubit-python interface for
+SPECFEM3D (see next section). The Python language allows to define
+and manipulate variables to parameterize the mesh. Alternatively,
+the Python script can call a Cubit journal file ({*}.jou), which looks
+like the example above. Variables can be defined and manipulated there
+using the $\mathtt{APREPRO}$ language built in Cubit.\\
+
+\texttt{Note}: you should avoid gaps in the list of indices of mesh
+objects with the following Cubit command:
+
+compress ids hex face edge node')
+
+(otherwise you will get a segmentation fault during domain decomposition)\\
+
+\section{Cubit-Python Scripts for Faults}
+
+The mesh generated in Cubit needs to be processed and exported in
+a format compatible with SPECFEM3D. This is achieved in the Python
+scripts by calling the Python-Cubit interface functions defined in
+the CUBIT directory: 
+\begin{enumerate}
+\item Function \textquotedbl{}$\mathtt{define\_bc}$\textquotedbl{} (or
+similar ones) must be called to set up the absorbing boundaries database 
+\item Function \textquotedbl{}$\mathtt{fault\_input}$\textquotedbl{} must
+be called once for each fault to set up the fault database 
+\item Function \textquotedbl{}$\mathtt{cubit2specfem3d.export2SESAME}$\textquotedbl{}
+must be called at the very end of the script to export the mesh in
+a SPECFEM3D format. 
+\end{enumerate}
+The functions in \#1 and \#3 are for bulk and are not documented here.
+We focus here on \#2:\\
+
+\textbf{Function}: $\mathtt{fault\_input}$
+
+\textbf{Purpose}: export a fault mesh database from Cubit to a SPECFEM3D-compliant
+file
+
+\textbf{Syntax}: fault\_input(id\_fault, ids\_surf\_1, ids\_surf\_2)\\
+
+\textbf{Inputs}: id\_fault integer index assigned to the fault.
+
+(The user must number all the faults, starting at 1, with unit increments) 
+\begin{lyxlist}{00.00.0000}
+\item [{ids\_surf\_1}] list of Cubit ids of all surfaces that form side
+1 of the fault
+\item [{ids\_surf\_2}] list of Cubit ids of all surfaces that form side
+2 of the fault 
+\end{lyxlist}
+(The user must decide which side of the fault is side 1. This choice
+affects the sign conventions of fault quantities as explained in a
+later section)\\
+
+\textbf{Outputs}: file \textquotedbl{}fault\_file\_X.dat\textquotedbl{},
+where X is the fault id (id\_fault).\\
+
+\textbf{Example}: For the example in the previous section:\\
+
+$\hphantom{}$A1 = {[}168{]}
+
+$\hphantom{}$A2 = {[}160{]}
+
+$\hphantom{}$faultA = fault\_input(1,A1,A2)\\
+
+\section{Running a Simulation }
+\begin{enumerate}
+\item Copy the input files to the work directory:
+
+>\textcompwordmark{}> cd \textasciitilde{}/$\mathtt{SPECFEM3D}$$\mathtt{}$
+
+>\textcompwordmark{}> cp -r $\mathtt{EXAMPLES/splay\_faults/DATA}$
+./
+
+\item Create a mesh with CUBIT: (If you have not installed CUBIT yet skip
+this step and use the mesh files included with the examples.)
+
+See the previous section \textquotedbl{}Mesh Generation with Split
+Nodes\textquotedbl{} or move to the directory CUBIT, open the cubit
+GUI, run a CUBIT script: >\textcompwordmark{}> cd \textasciitilde{}$\mathtt{/SPECFEM3D/CUBIT}$
+>\textcompwordmark{}> cubit In CUBIT's menu \textquotedbl{}Tools\textquotedbl{},
+select \textquotedbl{}Play Journal File\textquotedbl{} and select
+a script file, for instance: $\mathtt{EXAMPLES/splay\_faults/splay\_faults.py}$
+
+This creates several mesh files in directory $\mathtt{CUBIT/MESH/}$:
+absorbing\_surface\_file\_{*} (5 files) free\_surface\_file materials\_file
+mesh\_file nodes\_coords\_file nummaterial\_velocity\_file
+
+The CUBIT graphics window should show the mesh (e.g. $\mathtt{EXAMPLES/splay_{f}aults/splay_{f}aults.jpg}$)
+You can zoom, pan or rotate using mouse gestures. To look at the block
+of your interest, set \textquotedbl{}visibility off\textquotedbl{}
+in the other blocks.
+
+To run CUBIT from different working directories it is convenient to
+include the path to $\mathtt{SPECFEM3D/CUBIT}$ in an environment
+variable $\mathtt{PYTHONPATH}$
+
+\item Partition the mesh with the domain decomposition software SCOTCH.
+
+>\textcompwordmark{}> cd \textasciitilde{}/$\mathtt{SPECFEM3D/decompose\_mesh\_SCOTH}$
+
+>\textcompwordmark{}> ./$\mathtt{decompose\_mesh\_SCOTCH}$ 'nproc'
+../$\mathtt{CUBIT/MESH/}$ ../$\mathtt{DATABASES\_MPI}$ /
+
+where 'nproc' is the number of processors that you will use in the
+simulation. The second argument (../$\mathtt{CUBIT/MESH/}$ in this
+example) is the path to the directory containing the mesh files that
+were generated by CUBIT. This creates one mesh file proc000{*}{*}{*}\_Database
+per processor in directory $\mathtt{DATABASES\_MPI}$ . These files
+can be large, so you might want to place $\mathtt{DATABASES\_MPI}$
+in a scratch disk and specify the path accordingly when executing
+$x\mathtt{decompose\_mesh\_SCOTCH}$.
+
+\item Edit the file $\mathtt{DATA/Par_{f}ile}$: (copy an example here,
+see III.2)
+
+$\mathtt{LOCAL\_PATH}$= should be the path to $\mathtt{DATABASES\_MPI}$
+
+$\mathtt{NPROC}$= number of processors. The same as the number of
+partitions in SCOTCH (step 3).
+
+\item Generate databases:
+
+>\textcompwordmark{}> cd \textasciitilde{}/$\mathtt{SPECFEM3D}$
+
+>\textcompwordmark{}> mpirun -np nproc ./$\mathtt{xgenerate_{d}atabases}$
+
+or
+
+>\textcompwordmark{}> $\mathtt{./run/run.xdatabases}$
+
+or
+
+>\textcompwordmark{}> qsub -l nodes=... -l walltime=... go\_mesher
+(the script go\_mesher is in $\mathtt{utils/}$)
+
+This creates several binary mesh files for each processor (proc000{*}{*}{*}.bin)
+in directory $\mathtt{DATABASES\_MPI}$.
+
+\item If this is the first time you run a simulation with this mesh, set
+the parameter DT (the time step) DATA/Par\_file to a value smaller
+than the \textquotedbl{}Maximum suggested time step\textquotedbl{}
+in $\mathtt{OUTPUT\_FILES/output\_mesher.txt}$. Consider the note
+in section IV about the effect of Kelvin-Voigt damping on the critical
+timestep. Edit the variables $\mathtt{NSTEP}$ and $\mathtt{NTSTEP\_*}$
+in $\mathtt{DATA/Par\_file}$ accordingly.
+\item Run the solver:
+
+>\textcompwordmark{}> mpirun -np nproc ./$\mathtt{xspecfem3D}$
+
+or
+
+>\textcompwordmark{}> $\mathtt{./run/run.xspecfem3d}$
+
+or
+
+>\textcompwordmark{}> qsub -l nodes=... -l walltime=... go\_solver
+(the script go\_solver is in $\mathtt{utils/}$)
+
+\end{enumerate}
+
+\section{Examples}
+
+The package includes examples, the SCEC benchmark problems: 
+\begin{itemize}
+\item TPV5, a planar vertical strike-slip fault 
+\item TPV14 and TPV15, vertical strike-slip fault system with a fault branch 
+\item TPV102 and TPV103, rate-and-state benchmarks
+\item TPV16, heterogenous initial stress
+\item Splay fault models from Wendt et al. (2009)\\
+
+\end{itemize}
+To run the examples:
+\begin{enumerate}
+\item Read the documents in the directory $\mathtt{EXAMPLES/*/description}$.
+They contain a description of the example and additional instructions
+to run it. 
+\item Replace the contents of directory $\mathtt{DATA/*}$ by one of the
+$\mathtt{EXAMPLES/tpv*/DATA}$, for instance:
+
+>\textcompwordmark{}> cp -r $\mathtt{EXAMPLES/tpv5/DATA}$ ./
+
+\item Follow all the steps in section II above.
+\item Visualize the results with the matlab scripts in the directory $\mathtt{EXAMPLES/*/post}$
+\end{enumerate}
+
+\section{Sign Convention for Fault Quantities}
+
+During mesh generation, the fault is defined by two surfaces in contact.
+Let's denote as \textquotedbl{}\emph{side 1}\textquotedbl{} the SECOND
+surface declared by the user in the call to the python function \textquotedbl{}\emph{fault\_input}\textquotedbl{},
+and the FIRST surface as \textquotedbl{}\emph{side 2}\textquotedbl{}.
+The local coordinate system on the fault is defined as the right-handed
+coordinate system defined by (strike, dip, normal), where \textquotedbl{}normal\textquotedbl{}
+is the normal vector outgoing from side 1, \textquotedbl{}\emph{dip}\textquotedbl{}
+is parallel to the along-dip direction pointing downwards, and \textquotedbl{}\emph{strike}\textquotedbl{}
+is the horizontal along-strike vector such that the system is right-handed.\\
+
+Slip is defined as displacement on side 2 minus displacement on side
+1. In the local coordinate system on the fault, positive along-strike
+slip is right-lateral and positive along-dip slip is thrust if side
+1 is on the hanging wall (normal faulting if side 1 is on the foot
+wall).\\
+
+Traction is defined as the stress induced on side 1 by side 2, which
+is the stress tensor times the normal vector outgoing from side 1.
+In the local coordinate system on the fault, the normal traction is
+negative in compression, positive along-strike traction generates
+right-lateral slip and positive along-dip traction generates thrust
+slip if side 1 is on the hanging wall (normal faulting if side 1 is
+on the foot wall).
+
+\section{Input Files}
+
+\textbf{DATA/Par\_file} 
+\begin{lyxlist}{00.00.0000}
+\item [{\textbf{DATA/STATIONS}}] List of stations outside the fault 
+\item [{\textbf{DATA/Par\_file\_faults}}] contains parameters of the fault.
+The first part of this file has a strict format:
+
+\begin{description}
+\item [{Line}] 1: Number of faults (NF) 
+\item [{Lines}] 2 to NF+1: Kelvin Voigt damping (in seconds) for each fault.
+(See below how to set this parameter) 
+\item [{Line}] NF+2: Type of simulation (1=dynamic , 2 = kinematic) 
+\item [{Line}] NF+3: Number of time steps between updates of the time series
+outputs at selected fault points (see DATA/FAULT\_STATIONS), usually
+a large number (100s or 1000s). Note that the sampling rate of the
+time series is usually much higher.
+\item [{Line}] NF+4: Number of time steps between fault snapshot outputs
+(quantities at every fault point exported at regular times), usually
+a large number (100s or 1000s). 
+\item [{Line}] NF+5: Slip velocity threshold below which frictional healing
+is set (friction coefficient is reset to its static value). If this
+value is negative healing is disabled. 
+\item [{Line}] NF+6: Slip velocity threshold to define the rupture front.
+Only used for outputs.
+\end{description}
+
+The rest of this file is made of namelist input blocks (see \textquotedbl{}namelist\textquotedbl{}
+in a Fortran 9x manual). The input for each fault has the following
+sequence (arguments in {[}brackets{]} are optional):
+
+\end{lyxlist}
+\&\textbf{BEGIN\_FAULT} /
+
+\&\textbf{INIT\_STRESS} S1, S2, S3 {[},n1, n2, n3{]} /
+
+followed by (n1+n2+n3) \&\textbf{DIST2D} blocks
+
+\&\textbf{SWF} mus, mud, dc {[}, nmus, nmud, ndc{]} /
+
+followed by (nmus+nmud+ndc) \&\textbf{DIST2D} blocks\\
+
+\begin{lyxlist}{00.00.0000}
+\item [{\&\textbf{INIT\_STRESS}}] input block sets the initial fault stresses
+relative to the foot-wall side of the fault. Initial stresses are
+composed of a constant background value possibly overwritten in prescribed
+regions by heterogeneous distributions (see \&\textbf{DIST2D} blocks
+below):
+
+S1 = constant background value of along-strike shear stress (positive
+in the usual strike direction)
+
+S2 = constant background value of along-dip shear (positive is down-dip,
+normal faulting)
+
+S3 = constant background value of normal stress (negative in compresion)
+
+n1 = number of heterogeneous items for along-strike shear stress {[}default
+is 0{]}
+
+n2 = number of heterogeneous items for along-dip shear stress {[}default
+is 0{]}
+
+n3 = number of heterogeneous items for normal stress {[}default is
+0{]}\\
+
+\item [{\&\textbf{SWF}}] input block sets the slip-weakening friction parameters
+of the fault:
+
+mus = constant background value of static friction coefficient
+
+mud = constant background value of dynamic friction coefficient
+
+dc = constant background value of critical slip-weakening distance
+
+nmus = number of heterogeneous items for static friction coefficient
+{[}default is 0{]}
+
+nmud = number of heterogeneous items for dynamic friction coefficient
+{[}default is 0{]}
+
+ndc = number of heterogeneous items for critical slip-weakening distance
+{[}default is 0{]}\\
+
+\item [{\&\textbf{DIST2D}}] input blocks modify (overwrite) the value of
+a fault parameter by a heterogeneous spatial distribution:
+
+\&\textbf{DIST2D} shape='$\mathtt{square}$', val, xc, yc, zc, l /
+
+$\;$sets a constant value (val) within a cube with center (xc,yc,zc)
+and edge size l.
+
+\&\textbf{DIST2D} shape='$\mathtt{rectangle}$', val, xc, yc, zc,
+lx, ly, lz /
+
+$\;$sets a constant value (val) within a rectangular prism with center
+(xc,yc,zc) and edge sizes (lx,ly,lz).
+
+\&\textbf{DIST2D} shape='$\mathtt{rectangle-taper}$', val, valh,
+xc, yc, zc, lx, ly, lz /
+
+$\;$sets a vertical linear gradient within a rectangular prism with
+center (xc,yc,zc) and edge sizes (lx,ly,lz). Values vary linearly
+as a function of vertical position z between value val at z = zc-lz/2
+and value valh at z = zc+lz/2 .
+
+\&\textbf{DIST2D} shape='$\mathtt{circular}$', val, xc, yc, zc, r
+/
+
+$\;$sets a constant value (val) within a sphere with center (xc,yc,zc)
+and radius r. \\
+
+\item [{\textbf{DATA/FAULT\_STATIONS}}] Stations in the fault plane.
+
+Line 1: number of stations.
+
+Line 2 to end: 5 columns: X, Y, Z (-depth), station name, fault-id
+
+The fault-id identifies the fault that contains the station. It is
+the index of appearance in the faults list after line 2 of Par\_file\_faults
+
+\item [{\textbf{DATA/input\_file.txt}}] Heterogeneous stresses and friction
+parameters Documented in page 10 of EXAMPLES/tpv16.crack/description/TPV16\_17\_Description\_v03.pdf
+
+To activate this feature, in fault\_solver.f90 set TPV16 = .true.
+then re-compile the code: cd SPECFEM3D make
+
+\item [{Heterogeneous}] velocity models can be given in a regular grid.
+
+In the mesh generation Python file (.py) or in the CUBIT journal file
+(.jou), set the material 1st attribute to -1:
+
+\qquad{}block {*} attribute index 1 -1
+
+In module model\_tomography.f90:
+
+\qquad{}Set the parameter TOMO\_FILENAME to the name of the file
+containing the velocity model.
+
+Re-compile the code:
+
+\qquad{}cd SPECFEM3D
+
+\qquad{}make \\
+
+The format of the velocity file is: 
+\begin{description}
+\item [{Line}] 1: ORIG\_X, ORIG\_Y, ORIG\_Z, END\_X, END\_Y, END\_Z (coordinates
+of the two extreme corners of the box) 
+\item [{Line}] 2: SPACING\_X, SPACING\_Y, SPACING\_Z (regular grid spacing
+in each direction) 
+\item [{Line}] 3: NX, NY, NZ (number of grid point in each direction) 
+\item [{Line}] 4: VP\_MIN, VP\_MAX, VS\_MIN, VS\_MAX, RHO\_MIN, RHO\_MAX
+(min and max values of P wave speed, S wave speed and density) 
+\item [{Line}] 5 to 4+NX{*}NY{*}NZ: x\_tomo,y\_tomo,z\_tomo,vp\_tomo,vs\_tomo,rho\_tomo
+(position, P wave speed, S wave speed, density)\\
+
+\end{description}
+\end{lyxlist}
+Several files are generated automatically by xgenerate\_databases
+in directory DATABASES\_MPI and do not need to be modified by the
+user.
+
+\section{How to set the Kelvin-Voigt Damping Parameter }
+
+The purpose of the Kelvin-Voigt viscosity in the dynamic fault solver
+is to damp spurious oscillations generated by the fault slip at frequencies
+that are too high to be resolved by the mesh. The viscosity \textquotedbl{}\emph{eta}\textquotedbl{}
+(in seconds) depends on the size of the elements on the fault. Here
+is how to set it:
+\begin{enumerate}
+\item Determine the average linear size of the elements on the fault plane,
+\textquotedbl{}\emph{h\_fault}\textquotedbl{}. Usually this value
+is prescribed by the user during mesh generation. Otherwise it can
+be found by inspection of the mesh inside the Cubit GUI. 
+\item Use the matlab function $\mathtt{utils/critical_{t}imestep.m}$ to
+compute dtc\_fault = critical\_timestep(cp,h\_fault,ngll) This is
+the critical time step in an elastic medium for a hypothetical element
+of cubic shape with size equal to h\_fault. 
+\item Set eta in Par\_file\_faults to (0.1 to 0.3){*}dtc\_fault. A larger
+eta damps high-frequencies more aggresively but it might also affect
+lower frequencies and rupture speed.
+
+Viscosity reduces numerical stability: the critical timestep in a
+simulation with Kelvin-Voigt damping needs to be smaller than that
+in a purely elastic simulation. Here is how to set the time step accordingly:
+
+\item Run a test simulation without viscosity (eta=0 and only a few time
+steps) 
+\item Look for the \textquotedbl{}maximum suggested time step\textquotedbl{}
+in $\mathtt{OUTPUT_{F}ILES/output_{m}esher.txt}$ This is the critical
+timestep of a purely elastic simulation, \textquotedbl{}dtc\_bulk\textquotedbl{}. 
+\item Reset the timestep of the simulation with a Kelvin-Voigt material
+to a value smaller than dtc\_kv = eta{*}( sqrt(1+dtc\_bulk\textasciicircum{}2/eta\textasciicircum{}2)-1
+)
+\end{enumerate}
+Note that in general dtc\_bulk is smaller than dtc\_fault, because
+elements off the fault might be smaller or more distorted than element
+faces on the fault.
+
+\section{Output Files }
+
+Several output files are saved in \textasciitilde{}/$\mathtt{SPECFEM3D/OUTPUT_{F}ILES}$:\\
+
+\begin{enumerate}
+\item Seismograms for each station on the fault plane given in $\mathtt{DATA/FAULT_{S}TATIONS}$.
+One output file is generated for each station, named after the station.
+The files are ascii and start with a header (22 lines long) followed
+by a data block with the following format, one line per time sample:\\
+
+\# Column \#1 = Time (s)
+
+\# Column \#2 = horizontal right-lateral slip (m)
+
+\# Column \#3 = horizontal right-lateral slip rate (m/s)
+
+\# Column \#4 = horizontal right-lateral shear stress (MPa)
+
+\# Column \#5 = vertical up-dip slip (m)
+
+\# Column \#6 = vertical up-dip slip rate (m/s)
+
+\# Column \#7 = vertical up-dip shear stress (MPa)
+
+\# Column \#8 = normal stress (MPa)\\
+
+The stresses are relative to the footwall side of the fault (this
+convention controls their sign, but not their amplitude). Slip is
+defined as displacement of the hanging wall relative to the footwall.
+
+\item Seismograms at stations in the bulk (out of the fault plane) given
+in $\mathtt{DATA/STATIONS}$. 
+\item Rupture time files are named Rupture\_time\_FAULT-id. One file is
+generated for each fault. The files are ascii and start with a header
+(12 lines long) followed by a data block with the following format,
+one line per fault node:
+
+\# Column \#1 = horizontal coordinate, distance along strike (m)
+
+\# Column \#2 = vertical coordinate, distance down-dip (m)
+
+\# Column \#3 = rupture time (s)
+
+\item Fault quantities (slip, slip rate, stresses, etc) at regular times
+are stored in binary data files called Snapshot\#it\#.bin, where \#it\#
+is the timestep number. These can be read in Matlab with the function
+$\mathtt{Post-processing/FSEM3D\_snapshot.m}$
+\end{enumerate}
+
+\section{Post-processing and Visualization }
+
+Some Matlab functions for post-processing and visualization are included
+in directory Post-processing. The functions are internally documented
+(see their matlab help).\\
+
+\texttt{FSEM3D\_snapshot} reads a fault data snapshot\\
+
+The directories$\mathtt{EXAMPLES/*/post}$ contain additional Matlab
+scripts to generate figures specific to each example.
+
 \chapter{\label{cha:Adjoint-Simulations}Adjoint Simulations}

### 2013-03-16 doc/USER_MANUAL/manual_SPECFEM3D_Cartesian.tex
Made some corrections to manual and updated the pdf
added a comment about APPROXIMATE_OCEAN_LOAD; also fixed problems introduced by Lyx in the new fault dynamics section.

@@
  \item[]}
+{\end{list}}
+
+\newenvironment{lyxlist}[1]
+{\begin{list}{}
+{\settowidth{\labelwidth}{#1}
+\setlength{\leftmargin}{\labelwidth}
+\addtolength{\leftmargin}{\labelsep}
+\renewcommand{\makelabel}[1]{#1\hfil}}}
 {\end{list}}
@@
 phases/reverberations is not taken into account, even when the flag is on.
+If you want to model the effect of a fluid-solid model at short periods,
+then set this flag to \texttt{.false.} and mesh the fluid layer explicitly in your mesher,
+so that it is computed accurately and without this approximation.
 \item [{\texttt{TOPOGRAPHY}}] This feature is only effective if \texttt{APPROXIMATE\_OCEAN\_LOAD} is set to \texttt{.true.}.
@@
 \noindent \begin{centering}
-\includegraphics[scale=0.3]{\lyxdot \lyxdot /src/MAIN_TRUNK/SPECFEM3D_Cartesian/doc/USER_MANUAL/figures/faultmesh}
+\includegraphics[scale=0.3]{figures/faultmesh}
 \\
-\includegraphics[scale=0.3]{\lyxdot \lyxdot /src/MAIN_TRUNK/SPECFEM3D_Cartesian/doc/USER_MANUAL/figures/surf168}
-\includegraphics[scale=0.3]{\lyxdot \lyxdot /src/MAIN_TRUNK/SPECFEM3D_Cartesian/doc/USER_MANUAL/figures/splitnodes} 
+\includegraphics[scale=0.3]{figures/surf168}
+\includegraphics[scale=0.3]{figures/splitnodes}
 \par\end{centering}
-
 \caption{Screenshots of the CUBIT example for embedded fault with split nodes:
@@
 on the edges of fault surface are touching each other (bottom-right).}
-
 \label{fig:examples.splitnodes}
@@
 (The user must number all the faults, starting at 1, with unit increments)
-\begin{lyxlist}{00.00.0000}
+\begin{lyxlist}{-}
 \item [{ids\_surf\_1}] list of Cubit ids of all surfaces that form side
@@
 \item TPV102 and TPV103, rate-and-state benchmarks
+\item TPV22 and TPV23, step-over benchmarks
 \item TPV16, heterogenous initial stress
@@
 \textbf{DATA/Par\_file}
-\begin{lyxlist}{00.00.0000}
+\begin{lyxlist}{-}
 \item [{\textbf{DATA/STATIONS}}] List of stations outside the fault
@@
 
-\begin{lyxlist}{00.00.0000}
+\begin{lyxlist}{-}
 \item [{\&\textbf{INIT\_STRESS}}] input block sets the initial fault stresses

### 2013-03-27 doc/Version_Control/Git-SVN_Quick_Reference.tex
added doc/Version_Control/compile_LaTeX_manual.sh; fixed a typo
Some minor tweaks to the git-svn instructions.

@@
 \documentclass[11pt,twoside]{article}
+
+\usepackage[top=3cm,bottom=2.5cm,left=3cm,right=2cm]{geometry}
+\setlength{\parindent}{0pt}
 
@@
 \begin{verbatim}
-$ git svn clone -s \
-  svn+ssh://svn@geodynamics.org/cig/seismo/3D/SPECFEM3D_GLOBE
+  $ git svn clone -s svn+ssh://svn@geodynamics.org/cig/seismo/3D/SPECFEM3D_GLOBE
 \end{verbatim}
-The previous command checkouts the full svn history and converts it into a git history. It takes a huge amount of time. If the full history is not needed, one can checkout a partial history.
+The previous command checks out the full svn history and converts it into a git history. It takes a huge amount of time. If the full history is not needed, one can check out a partial history.
 For instance, checking out revisions from $20\:000$ to the last one (\texttt{HEAD}) is done with:
@@
   $ git svn clone -s -r 20000:HEAD \
-  svn+ssh://svn@geodynamics.org/cig/seismo/3D/SPECFEM3D\_GLOBE
+      svn+ssh://svn@geodynamics.org/cig/seismo/3D/SPECFEM3D_GLOBE
 \end{verbatim}
@@
 
-Creating a new branch \texttt{SPECFEM3D\_GLOBE\_ADIOS} on the svn server:
+Creating a new branch \texttt{SPECFEM3D\_GLOBE\_ADIOS} on the SVN server:
 \begin{verbatim}
@@
 \begin{verbatim}
-$ git checkout -b local/SPECFEM3D_GLOBE_ADIOS \
-  SPECFEM3D_GLOBE_ADIOS
+  $ git checkout -b local/SPECFEM3D_GLOBE_ADIOS SPECFEM3D_GLOBE_ADIOS
 \end{verbatim}
 
-\subsection*{Git -- Checking the status of the file.}
+\subsection*{Git -- Checking the status of the file}
 Before doing any modification it is mandatory to know which files are tracked, have been modified or deleted.
@@
   $ git add filename
-\end{verbatim}
 
-\begin{verbatim}
   $ git status
   # should now return the added file
-# as 'new file' unstead of 'untracked'.
+  # as `new file' instead of `untracked'.
 \end{verbatim}
@@
 \begin{verbatim}
-$ rm filename # remove the file from your workspace.
-\end{verbatim}
-\begin{verbatim}
   $ git rm filename
-# prepare to remove the file from the git local repository.
-\end{verbatim}
-\begin{verbatim}
+  # prepare to remove the file from the git local repository
+  # and remove the file from your workspace.
+
   $ git commit #...
@@
   $ git mv old_file new_file
-\end{verbatim}
-\begin{verbatim}
+
   $ git commit
@@
   $ mv old_file new_file
-\end{verbatim}
-\begin{verbatim}
+
   $ git rm old_file
-\end{verbatim}
-\begin{verbatim}
+
   $ git add new_file
-\end{verbatim}
-\begin{verbatim}
+
   $ git commit
@@
 
+Do not do this on commits that have already been sent to the SVN server.
+
 \begin{verbatim}
   $ git commit -m 'initial commit'
-\end{verbatim}
-\begin{verbatim}
+
   $ git add forgotten_file
-\end{verbatim}
-\begin{verbatim}
+
   $ git commit --amend
 \end{verbatim}
+
 \subsection*{Git -- Unstaging a staged file (i.e. Removing a file from the list of files to be
@@
   $ git status
-# the file has to be in the 'changes to be committed' section
-\end{verbatim}
+  # the file has to be in the `changes to be committed' section
 
-\begin{verbatim}
   $ git reset HEAD filename
   filename: locally modified
-\end{verbatim}
-\begin{verbatim}
+
   $ git status
-# the file should be in the 'Changes not staged for commit' section
+  # the file should be in the `Changes not staged for commit' section
 \end{verbatim}
@@
   $ git checkout -- filename
-\end{verbatim}
-\begin{verbatim}
+
   $ git status
@@
 
-\subsection*{Git-svn -- Back to the svn server}
+\subsection*{Git-svn -- Back to the SVN server}
 
-Push modifications up to the svn server:
+Push modifications up to the SVN server:
 \begin{verbatim}
@@
 \end{verbatim}
-This will update your local checkout and then re-apply your local un-submitted commits on top of the new trunk. 
-If you want to get the commits for all branches that exist in your clone: 
+This will update your local checkout and then re-apply your local unsubmitted commits on top of the new trunk so that the history remains linear.
+If you want to get the commits for all branches that exist in your clone (without modifying local changes):
 \begin{verbatim}
@@
 
-\subsection*{References and usefull readings}
+\subsection*{Git -- Stashing changes temporarily}
 
-\verb|http://git-scm.com/book|
-\\
-\verb|http://git.or.cz/course/svn.html|
-\\
-\verb|http://trac.parrot.org/parrot/wiki/git-svn-tutorial|
+If you have changes that have not been committed (locally), you will run into errors when you try to rebase from the SVN server.
+You can temporarily stash your changes before updating the tree and then restore them afterwards.
+
+\begin{verbatim}
+  $ git stash
+  # Changes to workspace from previous commit will be stored in a temporary area.
+
+  $ git svn rebase
+  # Do remote-to-local tree update.
+
+  $ git stash pop
+  # Previous changes are restored to workspace and deleted from temporary area.
+\end{verbatim}
+
+\subsection*{References and useful readings}
+
+\begin{itemize}
+\item \verb|http://git-scm.com/book|
+\item \verb|http://git.or.cz/course/svn.html|
+\item \verb|http://trac.parrot.org/parrot/wiki/git-svn-tutorial|
+\end{itemize}
 

### 2013-03-31 doc/USER_MANUAL/manual_SPECFEM3D_Cartesian.tex
changed & to and
updated the name of the absorbing condition flags in the manual

@@
 \item [{\texttt{OLSEN\_ATTENUATION\_RATIO}}] Determines the Olsen's constant in Olsen's empirical relation (see \citet{OlDaBr03}).
-\item [{\texttt{ABSORBING\_CONDITIONS}}] Set to \texttt{.true.} to turn
-on Clayton-Enquist absorbing boundary conditions (see \citet{KoTr99}).
-\item [{\texttt{ABSORB\_INSTEAD\_OF\_FREE\_SURFACE}}] Set to \texttt{.true.} to turn on
+\item [{\texttt{STACEY\_ABSORBING\_CONDITIONS}}] Set to \texttt{.true.} to turn
+on Clayton-Enquist absorbing boundary conditions (see \citet{KoTr99}). In almost all cases it is much better to use CPML absorbing
+layers (see the options below) and leave this flag to \texttt{.false.}.
+\item [{\texttt{STACEY\_INSTEAD\_OF\_FREE\_SURFACE}}] Set to \texttt{.true.} to turn on
 absorbing boundary conditions on the top surface which by default constitutes a free surface of the model.
@@
 \item [{\texttt{PML\_INSTEAD\_OF\_FREE\_SURFACE}}] Set to \texttt{.true.} to turn on C-PML boundary conditions on the top surface which by default constitutes a free surface of the model.
-\item [{\texttt{PML\_WIDTH\_MIN}}] Gives a min. authorized bound to the solver to target an optimized thickness of C-PML layers.
-\item [{\texttt{PML\_WIDTH\_MAX}}] Gives a max. authorized bound to the solver to target an optimized thickness of C-PML layers.
 \item [{\texttt{f0\_FOR\_PML}}] Determines the dominant frequency that will be used in the calculation of: \\ ALPHA\_MAX\_PML = PI*f0\_FOR\_PML.
@@
 \item the time step parameters \texttt{NSTEP} and \texttt{DT}
-\item the absorbing boundary control parameter \texttt{ABSORBING\_CONDITIONS} on condition that the\\
-  \texttt{ABSORB\_INSTEAD\_OF\_FREE\_SURFACE} flag remains unmodified after running the databases generation.
+\item the absorbing boundary control parameter \texttt{PML\_CONDITIONS} on condition that the\\
+  \texttt{PML\_INSTEAD\_OF\_FREE\_SURFACE} flag remains unmodified after running the databases generation.
 \item the movie control parameters \texttt{MOVIE\_SURFACE}, \texttt{MOVIE\_VOLUME},
@@
 
-\chapter{\label{cha:Fault-Kinematics-Dynamics}Fault Kinematics \& Dynamics}
+\chapter{\label{cha:Fault-Kinematics-Dynamics}Fault Kinematics and Dynamics}
 

### 2013-04-04 doc/USER_MANUAL/manual_SPECFEM3D_Cartesian.tex
acknowledge Yvon Maday in the manual

@@
 from 1995 to 1997 and then by Dimitri Komatitsch and Jeroen Tromp at Harvard University and Caltech, USA, starting in 1998.
+The story started on March 28, 1995, when Prof. Yvon Maday from CNRS and University of Paris, France, gave a lecture to
+Dimitri Komatitsch and Jean-Pierre Vilotte at IPG about the nice properties of the spectral-element method that he had used for
+other equations. We are deeply indebted and thankful to him for that.
 

### 2013-04-08 doc/USER_MANUAL/manual_SPECFEM3D_Cartesian.tex
updated the users manual regarding CPML

@@
 \begin{description}
-\item [absorbing\_cpml\_file] {\bf (only in case of C-PML absorbing conditions)} Contains at the first line respectively the "number of C-PML spectral elements"
-and the "thickness of C-PML layers" (currently, a common thickness for all faces is defined but this can be easily adapted in the future if needed by the setting of CPML\_width\_x, CPML\_width\_y, CPML\_width\_z parameters).
-On all following lines are defined the "C-PML spectral elements global indexing" and "C-PML regions" for each
-C-PML element according to the following convention:
+\item [absorbing\_cpml\_file] {\bf (only needed in case of C-PML absorbing conditions)}
+Contains on the first line the total number of C-PML spectral elements in the mesh,
+and then on the following line the list of all these C-PML elements with two numbers per line:
+first the spectral element number, and then a C-PML flag indicating to which C-PML layer(s)
+that element belongs, according to the following convention:
 
 \begin{itemize}
-\item \#id\_cpml\_regions = 1 : X\_surface C-PML,
-\item \#id\_cpml\_regions = 2 : Y\_surface C-PML,
-\item \#id\_cpml\_regions = 3 : Z\_surface C-PML,
-\item \#id\_cpml\_regions = 4 : XY\_edge C-PML,
-\item \#id\_cpml\_regions = 5 : XZ\_edge C-PML,
-\item \#id\_cpml\_regions = 6 : YZ\_edge C-PML,
-\item \#id\_cpml\_regions = 7 : XYZ\_corner C-PML.
+\item Flag = 1 : element belongs to a X CPML layer only (either in Xmin or in Xmax),
+\item Flag = 2 : element belongs to a Y CPML layer only (either in Ymin or in Ymax),
+\item Flag = 3 : element belongs to a Z CPML layer only (either in Zmin or in Zmax),
+\item Flag = 4 : element belongs to a X CPML layer and also to a Y CPML layer,
+\item Flag = 5 : element belongs to a X CPML layer and also to a Z CPML layer,
+\item Flag = 6 : element belongs to a Y CPML layer and also to a Z CPML layer,
+\item Flag = 7 : element belongs to a X, to a Y and to a Z CPML layer, i.e., it belongs to a CPML corner.
 \end{itemize}
+
+Note that it does not matter whether an element belongs to a Xmin or to a Xmax CPML, the flag
+is the same in both cases; the same is true for Ymin and Ymax, and also Zmin and Zmax.
+
+When you have an existing CUBIT (or similar) mesh stored in SPECFEM3D format,
+i.e., if you have existing "nodes\_coords\_file" and "mesh\_file" files but do not know how to assign
+CPML flags to them, we have created a small serial Fortran program that will do that automatically
+for you, i.e., which will create the "absorbing\_cpml\_file" for you.
+That program is
+utils/CPML/convert\_external\_layers\_of\_a\_given\_mesh\_to\_CPML\_layers.f90, and a small Makefile
+is provided in that directory (utils/CPML).
+
+IMPORTANT: it is your responsibility to make sure that in the input CUBIT (or similar) mesh
+that this code will read in SPECFEM3D format from files "nodes\_coords\_file" and "mesh\_file"
+you have created layers of elements that constitute a layer of constant thickness aligned
+with the coordinate grid axes (X, Y and/or Z), so that this code can assign CPML flags to them.
+This code does NOT check that (because it cannot, in any easy way).
+The mesh inside these CPML layers does not need to be structured nor regular, any non-structured
+mesh is fine as long as it has flat PML inner and outer faces, parallel to the axes, and thus
+of a constant thickness.
+The thickness can be different between the X, Y and Z sides. But for X it must not vary,
+for Y it must not vary, and for Z it must not vary.
+If you do not know the exact thickness, or if you think the thickness may have tiny variations,
+for instance due to roundoff issues when exporting the mesh, you can use a slightly LARGER value
+in this code (say 2\% to 5\% more), and this code will fix that and will adjust it; never use
+a SMALLER value otherwise this code will miss some CPML elements.
+
+Note: in a future release we will remove the constraint of having CPML layers aligned with the
+coordinate axes; we will allow for meshes that are titled by any constant angle in the horizontal plane.
+However this is not implemented yet.
 

### 2013-04-08 doc/USER_MANUAL/manual_SPECFEM3D_Cartesian.tex
removed negative -2001 .. -2007 fictitious CPML materials, which are now unused.

@@
 \end{lyxcode}
-where \texttt{element\_ID} is the element identifier and \texttt{material\_ID} is a unique identifier, positive (for materials) or negative (for tomographic models or C-PML elements).
-For C-PML absorbing conditions, \texttt{material\_ID} is a unique negative identifier between -2001 and -2007 in correspondance to the convention seen above in the description of the
-absorbing\_cpml\_file: -2001 <-> 1, -2002 <-> 2, etc\ldots
+where \texttt{element\_ID} is the element identifier and \texttt{material\_ID} is a unique identifier,
+positive (for materials taken from this list of materials, i.e. for which each spectral
+element has constant material properties taken from this list) or negative (for tomographic models,
+i.e. for spectral element whose real velocities and density will be assigned later by calling an external function
+to define model variations, for instance in the case of tomographic models; in such a case, material properties
+can vary inside each spectral element, i.e. be different at each of its Gauss-Lobatto-Legendre grid points).
 
@@
 \begin{itemize}
-\item For fully defined materials, the format is:
+\item For classical materials (i.e., spectral elements for which the velocity and density
+model will not be assigned by calling an external function to define for instance a tomographic model), the format is:
 \begin{lyxcode}
@@
 where \texttt{domain\_ID} is 1 for acoustic or 2 for elastic materials, \texttt{material\_ID} a unique identifier,  \texttt{rho} the density in $kg \, m^{-3}$,  \texttt{vp} the P-wave speed in $m \, s^{-1}$, \texttt{vs}  the S-wave speed in $m \, s^{-1}$, \texttt{Q} the quality factor and \texttt{anisotropy\_flag} an identifier for anisotropic models.
-
-\item For C-PML absorbing conditions, , the format is:
-\begin{lyxcode}
-\texttt{domain\_ID} \texttt{material\_ID}  \texttt{rho}  \texttt{vp}  \texttt{vs}
-\end{lyxcode}
-where \texttt{domain\_ID} is 1 for acoustic or 2 for elastic materials, \texttt{material\_ID} a unique identifier between -2001 and -2007 to correspond with the definition given in materials\_file,  \texttt{rho} the density in $kg \, m^{-3}$,  \texttt{vp} the P-wave speed in $m \, s^{-1}$, \texttt{vs}  the S-wave speed in $m \, s^{-1}$
 %%magnoni
 \item For tomographic velocity models, please read Chapter \ref{cha:-Changing-the} and Section \ref{sec:Using-tomographic} `Using external tomographic Earth models' for further details.
-%
-%For a tomographic model, the material definition format is:
-%\begin{lyxcode}
-%\texttt{domain\_ID} \texttt{material\_ID}  \texttt{tomography}  \texttt{name}
-%\end{lyxcode}
-%where
-%\texttt{domain\_ID} is 1 for acoustic or 2 for elastic materials, \texttt{material\_ID} a negative, unique identifier, \texttt{tomography} keyword for tomographic material definition and \texttt{name} the name of the tomography file.
-The name is not used so far, rather change the filename defined in the file \texttt{model\_tomography.f90} located in the
-%\texttt{src/generate\_databases/} directory.
 %

### 2013-04-09 doc/USER_MANUAL/manual_SPECFEM3D_Cartesian.tex
improved two comments

@@
 of a constant thickness.
-The thickness can be different between the X, Y and Z sides. But for X it must not vary,
+The thickness can be different for the X, Y and Z sides. But for X it must not vary,
 for Y it must not vary, and for Z it must not vary.
-If you do not know the exact thickness, or if you think the thickness may have tiny variations,
-for instance due to roundoff issues when exporting the mesh, you can use a slightly LARGER value
-in this code (say 2\% to 5\% more), and this code will fix that and will adjust it; never use
+If you do not know the exact thickness,
+you can use a slightly LARGER value
+in this code (say 2\% to 5\% more) and this code will fix that and will adjust it; never use
 a SMALLER value otherwise this code will miss some CPML elements.
@@
 \end{lyxcode}
-where \texttt{domain\_ID} is 1 for acoustic or 2 for elastic materials, \texttt{material\_ID} a unique identifier,  \texttt{rho} the density in $kg \, m^{-3}$,  \texttt{vp} the P-wave speed in $m \, s^{-1}$, \texttt{vs}  the S-wave speed in $m \, s^{-1}$, \texttt{Q} the quality factor and \texttt{anisotropy\_flag} an identifier for anisotropic models.
+where \textbf{\texttt{domain\_ID} is 1 for acoustic and 2 for elastic or viscoelastic materials,}
+\texttt{material\_ID} a unique identifier,  \texttt{rho} the density in $kg \, m^{-3}$,  \texttt{vp} the P-wave speed in $m \, s^{-1}$, \texttt{vs}  the S-wave speed in $m \, s^{-1}$, \texttt{Q} the quality factor and \texttt{anisotropy\_flag} an identifier for anisotropic models.
 %%magnoni

### 2013-04-11 doc/USER_MANUAL/manual_SPECFEM3D_Cartesian.tex
moved the Stacey parameters to after the PML parameters in the Par_file, since using CPML is much better

@@
 \item [{\texttt{OLSEN\_ATTENUATION\_RATIO}}] Determines the Olsen's constant in Olsen's empirical relation (see \citet{OlDaBr03}).
+\item [{\texttt{PML\_CONDITIONS}}] Set to \texttt{.true.} to turn on C-PML boundary conditions for a regional simulation. Both fluids and elastic solids are supported.
+\item [{\texttt{PML\_INSTEAD\_OF\_FREE\_SURFACE}}] Set to \texttt{.true.} to turn on C-PML boundary conditions on the top surface
+instead of the usual free surface.
+\item [{\texttt{f0\_FOR\_PML}}] Determines the dominant frequency that will be used in the calculation of PML damping profiles; this should be
+set to the same (or similar) dominant frequency as that of the source that you will use in your simulation. If you plan to use a Dirac source,
+then use the dominant frequency of the source wavelet with which you plan to convolve your seismograms later on in post-processing.
 \item [{\texttt{STACEY\_ABSORBING\_CONDITIONS}}] Set to \texttt{.true.} to turn
 on Clayton-Enquist absorbing boundary conditions (see \citet{KoTr99}). In almost all cases it is much better to use CPML absorbing
-layers (see the options below) and leave this flag to \texttt{.false.}.
+layers (see the options above) and leave this flag to \texttt{.false.}.
 \item [{\texttt{STACEY\_INSTEAD\_OF\_FREE\_SURFACE}}] Set to \texttt{.true.} to turn on
 absorbing boundary conditions on the top surface which by default constitutes a free surface of the model.
-\item [{\texttt{PML\_CONDITIONS}}] Set to \texttt{.true.} to turn on C-PML boundary conditions for a regional simulation. Both fluids and elastic solids are supported.
-\item [{\texttt{PML\_INSTEAD\_OF\_FREE\_SURFACE}}] Set to \texttt{.true.} to turn on C-PML boundary conditions on the top surface which by default constitutes a free surface of the model.
-\item [{\texttt{f0\_FOR\_PML}}] Determines the dominant frequency that will be used in the calculation of: \\ ALPHA\_MAX\_PML = PI*f0\_FOR\_PML.
 \item [{\texttt{CREATE\_SHAKEMAP}}] Set this flag to \texttt{.true.} to

### 2013-04-11 doc/USER_MANUAL/manual_SPECFEM3D_Cartesian.tex
added FULL_ATTENUATION_SOLID description to the manual

@@
 \texttt{ATTENUATION\_f0\_REFERENCE} in the main constants file \texttt{constants.h} found in subdirectory \texttt{src/shared/}.
+\item [{\texttt{FULL\_ATTENUATION\_SOLID}}] Set to \texttt{.true.} to add $Q_\kappa$ attenuation to the simulations in addition to $Q_\mu$
+attenuation when the \texttt{ATTENUATION} flag above is on.
+Should be off in almost all cases for seismic wave propagation, since $Q_\kappa$ is almost always negligible in Earth models.
+However, this parameter can be useful for instance for ocean acoustics simulations, when $Q_\kappa$ is not negligible in the solid
+layer at the sea bottom.
 \item [{\texttt{ANISOTROPY}}] Set to \texttt{.true.} if you

### 2013-04-11 doc/USER_MANUAL/manual_SPECFEM3D_Cartesian.tex
renamed free_surface_file to free_or_absorbing_surface_file_zmax for clarity

@@
 \item [nodes\_coords\_file] Contains the point locations in Cartesian coordinates of the mesh element corners.
-\item [mesh\_file] Contains the mesh element connectivity.
-\item [free\_surface\_file] Contains the free surface connectivity. You should put both the surface of acoustic regions and of elastic regions in that file;
+\item [mesh\_file] Contains the mesh element connectivity. The hexahedral elements can have 8 or 27 nodes.\\
+See picture doc/mesh\_numbering\_convention/numbering\_convention\_27\_nodes.jpg to see\\
+in which (standard) order the points must be cited. In the case of 8 nodes, just include the first 8 points.
+\item [free\_or\_absorbing\_surface\_file\_zmax] Contains the free surface connectivity or \\
+the surface connectivity of the absorbing boundary surface at the top (Zmax), \\
+depending on whether the top surface is defined as free or absorbing in DATA/Par\_file. \\
+You should put both the surface of acoustic regions and of elastic regions in that file;
 that is, list all the element faces that constitute the surface of the model in that file.
-\item [absorbing\_surface\_file\_xmax] Contains the surface connectivity of the absorbing boundary surface at the Xmax \\(irrelevant in case of C-PML absorbing conditions).
-\item [absorbing\_surface\_file\_xmin] Contains the surface connectivity of the absorbing boundary surface at the Xmin \\(irrelevant in case of C-PML absorbing conditions).
-\item [absorbing\_surface\_file\_ymax] Contains the surface connectivity of the absorbing boundary surface at the Ymax \\(irrelevant in case of C-PML absorbing conditions).
-\item [absorbing\_surface\_file\_ymin] Contains the surface connectivity of the absorbing boundary surface at the Ymin \\(irrelevant in case of C-PML absorbing conditions).
-\item [absorbing\_surface\_file\_bottom] Contains the surface connectivity of the absorbing boundary surface at the bottom \\(irrelevant in case of C-PML absorbing conditions).
+\item [absorbing\_surface\_file\_xmax] Contains the surface connectivity of the absorbing boundary surface at Xmax \\(also needed in the case of C-PML absorbing conditions,
+in order for the code to be able to impose Dirichlet conditions on their outer edge).
+\item [absorbing\_surface\_file\_xmin] Contains the surface connectivity of the absorbing boundary surface at Xmin \\(also needed in the case of C-PML absorbing conditions,
+in order for the code to be able to impose Dirichlet conditions on their outer edge).
+\item [absorbing\_surface\_file\_ymax] Contains the surface connectivity of the absorbing boundary surface at Ymax \\(also needed in the case of C-PML absorbing conditions,
+in order for the code to be able to impose Dirichlet conditions on their outer edge).
+\item [absorbing\_surface\_file\_ymin] Contains the surface connectivity of the absorbing boundary surface at Ymin \\(also needed in the case of C-PML absorbing conditions,
+in order for the code to be able to impose Dirichlet conditions on their outer edge).
+\item [absorbing\_surface\_file\_bottom] Contains the surface connectivity of the absorbing boundary surface at the bottom (Zmin) \\(also needed in the case of C-PML absorbing conditions,
+in order for the code to be able to impose Dirichlet conditions on their outer edge).
 \end{description}
@@
 This creates several mesh files in directory $\mathtt{CUBIT/MESH/}$:
-absorbing\_surface\_file\_{*} (5 files) free\_surface\_file materials\_file
+absorbing\_surface\_file\_{*} (5 files) free\_or\_absorbing\_surface\_file\_zmax materials\_file
 mesh\_file nodes\_coords\_file nummaterial\_velocity\_file

### 2013-04-12 doc/USER_MANUAL/manual_SPECFEM3D_Cartesian.tex

@@
 \begin{lyxcode}
-\texttt{domain\_ID} \texttt{material\_ID}  \texttt{rho}  \texttt{vp}  \texttt{vs}  \texttt{Q}  \texttt{anisotropy\_flag}
+\texttt{domain\_ID} \texttt{material\_ID}  \texttt{rho}  \texttt{vp}  \texttt{vs}  \texttt{Qkappa}  \texttt{Qmu} \texttt{anisotropy\_flag}
 \end{lyxcode}
 where \textbf{\texttt{domain\_ID} is 1 for acoustic and 2 for elastic or viscoelastic materials,}
-\texttt{material\_ID} a unique identifier,  \texttt{rho} the density in $kg \, m^{-3}$,  \texttt{vp} the P-wave speed in $m \, s^{-1}$, \texttt{vs}  the S-wave speed in $m \, s^{-1}$, \texttt{Q} the quality factor and \texttt{anisotropy\_flag} an identifier for anisotropic models.
+\texttt{material\_ID} a unique identifier,  \texttt{rho} the density in $kg \, m^{-3}$,  \texttt{vp} the P-wave speed in $m \, s^{-1}$,
+\texttt{vs}  the S-wave speed in $m \, s^{-1}$, \texttt{Q} the quality factor and \texttt{anisotropy\_flag} an identifier for anisotropic models.
+Note that the \texttt{Qkappa} value is ignored by the code unless \texttt{FULL\_ATTENUATION\_SOLID} is set.
+Note also that both \texttt{Qkappa} and \texttt{Qmu} are ignored by the code unless \texttt{ATTENUATION} is set.
+If you want a model with no \texttt{Qmu} attenuation, both set \texttt{ATTENUATION} to \texttt{.false.} in the \texttt{Par\_file} and set \texttt{Qmu} to 9999 here.
+If you want a model with no \texttt{Qkappa} attenuation, both set \texttt{FULL\_ATTENUATION\_SOLID} to \texttt{.false.} in the \texttt{Par\_file} and set \texttt{Qkappa} to 9999 here.
 %%magnoni
@@
 or for more user control: \\
-\indent (2) set in the \texttt{Par\_file} the model parameter \texttt{MODEL = default}, define the negative \texttt{material\_ID} identifier for each element in the file \texttt{MESH/materials\_file} and use the following format in the file \texttt{MESH/nummaterial\_velocity\_file} when using CUBIT to construct your mesh (see also section \ref{subsec:Exporting-the-Mesh}):
+\indent (2) set in the \texttt{Par\_file} the model parameter \texttt{MODEL = default}, define the negative \texttt{material\_ID} identifier for each element in the file \texttt{MESH/materials\_file}
+and use the following format in the file \texttt{MESH/nummaterial\_velocity\_file}
+when using CUBIT to construct your mesh (see also section \ref{subsec:Exporting-the-Mesh}):
 \begin{lyxcode}
@@
 The user can implement his own interpolation algorithm for the tomography model by changing the routine \texttt{model\_tomography.f90} located in the \texttt{src/generate\_databases/} directory.
-Moreover, for models that involve both fully defined materials and a tomography description, the \texttt{nummaterial\_velocity\_file} has multiple lines each with the corresponding suitable format described above.
+Moreover, for models that involve both fully defined materials and a tomography description, the \texttt{nummaterial\_velocity\_file}
+has multiple lines each with the corresponding suitable format described above.
 

### 2013-04-12 doc/USER_MANUAL/manual_SPECFEM3D_Cartesian.tex
added IBM Blue Gene compilation instructions for SCOTCH

@@
 
-\underline{More recent installation instruction for IBM BlueGene, from October 2012:}\\
+\underline{More recent installation instruction for IBM BlueGene, from April 2013:}\\
 
@@
 \texttt{./configure FC=bgxlf90\_r MPIFC=mpixlf90\_r CC=bgxlc\_r LOCAL\_PATH\_IS\_ALSO\_GLOBAL=true}\\
+
+In order for the SCOTCH domain decomposer to compile, on some (but not all) Blue Gene systems after
+running \texttt{configure} you may need to edit file src/decompose\_mesh/scotch/src/Makefile.inc and replace\\
+
+\noindent
+CCS    = bgxlc\_r \\
+\noindent
+CCD    = bgxlc\_r (or similar)\\
+
+with:
+
+\noindent
+CCS    = gcc \\
+\noindent
+CCD    = gcc \\
+
+and then type "make clean" and "make all".
 

### 2013-04-13 doc/USER_MANUAL/manual_SPECFEM3D_Cartesian.tex
more precise instructions for IBM Blue Gene

--- 2013-04-12 doc/USER_MANUAL/manual_SPECFEM3D_Cartesian.tex
@@
 \texttt{./configure FC=bgxlf90\_r MPIFC=mpixlf90\_r CC=bgxlc\_r LOCAL\_PATH\_IS\_ALSO\_GLOBAL=true}\\
+
+In order for the SCOTCH domain decomposer to compile, on some (but not all) Blue Gene systems you may need to
+run \texttt{configure} with \texttt{CC=gcc} instead of \texttt{CC=bgxlc\_r}.
 

### 2013-05-06 doc/USER_MANUAL/manual_SPECFEM3D_Cartesian.tex
Final Merge from branch 'full_aniso_kernels_matharu' to master. Anisotropy arrays are allocated only if the corresponding flag is set to true.

@@
 \end{itemize}
+
+\item \textbf{Run the anisotropic kernel simulation}
+
+Instead of the kernels for the isotropic wave speeds, you can also
+compute the kernels for the 21 independent components $C_{IJ},\, I,J=1,...,6$
+(using Voigt's notation) of the elastic tensor in the cartesian coordinate system. 
+This is done by setting \texttt{ANISOTROPIC\_KL} \texttt{=} \texttt{.true.} 
+in \texttt{constants.h} before compiling the package. The definition of the parameters 
+$C_{IJ}$ in terms of the corresponding components $c_{ijkl},ijkl,i,j,k,l=1,2,3$ of 
+the elastic tensor in cartesian coordinates follows \citet{ChTr07}. The 21 anisotropic 
+kernels are saved in the \texttt{LOCAL\_PATH} in one file with the name of 
+\texttt{proc??????\_cijkl\_kernel.bin} (with \texttt{proc??????} 
+the processor number). The output kernels correspond to absolute perturbations 
+$\delta C_{IJ}$ of the elastic parameters and their unit is in $s/GPa/km^{3}$. 
+For consistency, the output density kernels with this option turned on 
+are for a perturbation $\delta\rho$ (and not $\frac{\delta\rho}{\rho}$) and their unit
+is in s / (kg/m$^{3}$) / km$^{3}$. These `primary' anisotropic kernels
+can then be combined to obtain the kernels for different parameterizations of anisotropy. 
+This can be done, for example, when combining the kernel files from slices into one mesh file 
+(see Section~\ref{sec:Finite-Frequency-Kernels}).
+
+If \texttt{ANISOTROPIC\_KL} \texttt{=} \texttt{.true.} by additionally setting \texttt{ANISOTROPIC\_KL} 
+\texttt{=} \texttt{.true.} in \texttt{constants.h} the package will save anisotropic kernels paramterized
+as velocities related to transverse isotropy based on the the Chen and Tromp parameters \citet{ChTr07}.
+The kernels are saved as relative perturbations for horizontal and vertical P and S velocities, $\alpha_{v},
+\alpha_{h}, \beta_{v}, \beta_{h}$. Explicit relations can be found in appendix B. of \citet{SiLiTrTr07b}
+
+The anisotropic kernels are only currently available for CPU mode.
+
 \end{enumerate}
+
 In general, the three steps need to be run sequentially to assure

### 2013-05-06 doc/Version_Control/Git-SVN_Quick_Reference.tex
Additional directions for git-svn. How to merge in a clean way. 	modified:   Git-SVN_Quick_Reference.tex

@@
 \end{verbatim}
+
+Merging a development branch into the trunk. The master branch should be merged
+beforehand into your development branch. This way the conflicts will only
+occure in the developement branch. This also helps to contribute with
+up-to-date modifications. Therefore, the following steps should be followed:
+\begin{enumerate}
+  \item \verb|git checkout master| -- Switch to the \texttt{master}.
+  \item \verb|git svn rebase| -- The \texttt{master} is now up-to-date with the repository.
+  \item \verb|git checkout mybranch| -- Switch to the branch you want to merge.
+  \item \verb|git merge master| -- Update \texttt{mybranch} with the
+\texttt{master}. When using git only \verb|--squash| is an option to bundle
+multiple modifications in a single one. When using git-svn, it is important to
+use the \verb|--no-ff| option, otherwise git will loose branch information.
+  \item \verb|git mergetool| -- If there is any conflict to solve.
+  \item \verb|git commit| -- Commit the merge.
+  \item \verb|git checkout master| -- Switch to the \texttt{master}.
+  \item \verb|git merge [--no-ff] branch| -- Update the \texttt{master} with \texttt{mybranch}.
+  \item \verb|git commit| -- Commit the merge.
+  \item \verb|git svn dcommit| -- Send the merge to the svn repository.
+\end{enumerate}
 

### 2013-05-06 doc/USER_MANUAL/manual_SPECFEM3D_Cartesian.tex
added a comment about fluid-solid matching inside the CPML; also restored doc/USER_MANUAL/manual_SPECFEM3D_Cartesian.pdf, which needs to be there because it is used on the CIG web site (the link to the manual points to this SVN file; thus please do NOT remove it)

@@
 
-This new release of the code (version 2.1) includes support for GPU graphics card acceleration \citep{KoMiEr09,KoErGoMi10,MiKo10,Kom11}.\\
-
-The next release of the code will include support Convolution or Auxiliary Differential Equation Perfectly Matched absorbing Layers
+This new release of the code (version 2.1)
+includes Convolution or Auxiliary Differential Equation Perfectly Matched absorbing Layers
 (C-PML or ADE-PML) \citep{MaKoEz08,MaKoGe08,MaKo09,MaKoGeBr10,KoMa07}.
-It will also use the PT-SCOTCH parallel and threaded version of SCOTCH for mesh partitioning instead of the serial version.\\
+It also includes support for GPU graphics card acceleration \citep{KoMiEr09,KoErGoMi10,MiKo10,Kom11}.\\
+
+The next release of the code will use the PT-SCOTCH parallel and threaded version of SCOTCH for mesh partitioning instead of the serial version.\\
 
@@
 However this is not implemented yet.
+
+Note: in the case of fluid-solid layers in contact, the fluid-solid interface currently needs to be flat and horizontal
+inside the CPML layer (i.e., bathymetry should become flat and horizontal when entering the CPML); this (small) constraint
+will probably remain in the code for a while because it makes fluid-solid matching inside the CPML much easier.
 

### 2013-05-08 doc/USER_MANUAL/manual_SPECFEM3D_Cartesian.tex
added description of utils/strike_dip_rake_to_CMTSOLUTION.c

@@
 \noindent \begin{centering}
-%%%\includegraphics[width=3in]{figures/gauss_vs_triangle_mod.eps}
 \includegraphics[width=2.5in]{figures/gauss_vs_triangle_mod.pdf}
 \par\end{centering}
-
-\caption{Comparison of the shape of a triangle and the Gaussian function actually
-used.}
-
+\caption{Comparison of the shape of a triangle and the Gaussian function actually used.}
 \label{fig:gauss.vs.triangle}
 \end{figure}
+
+If you know the earthquake source in strike/dip/rake format rather than in \texttt{CMTSOLUTION} format,
+use the C code \texttt{SPECFEM3D\_GLOBE/utils/strike\_dip\_rake\_to\_CMTSOLUTION.c} to convert it.
+The conversion formulas are given for instance in \cite{AkRi80}.
+Note that the \cite{AkRi80} convention is slightly different from the Harvard \texttt{CMTSOLUTION} convention
+(the sign of some components is different). The C code outputs both.
 

### 2013-05-10 doc/USER_MANUAL/manual_SPECFEM3D_Cartesian.tex
added a sentence about defining FC, CC and MPIF90 in .bashrc or .cshrc

@@
 a minimum, it is recommended that you explicitly specify the appropriate
-command names for your Fortran compiler and MPI package:
+command names for your Fortran compiler and MPI package (another option is to define FC, CC and MPIF90 in your .bash\_profile 
+or your .cshrc file):
 

### 2013-06-05 doc/USER_MANUAL/manual_SPECFEM3D_Cartesian.tex
Improved the Chapter on fault kinematic/dynamics in the manual

@@
-%% LyX 1.5.1 created this file.  For more info, see http://www.lyx.org/.
+%% LyX 2.0.2 created this file.  For more info, see http://www.lyx.org/.
 %% Do not edit unless you really know what you are doing.
-\documentclass[oneside,english,onecolumn]{book}
+\documentclass[oneside,english]{book}
 \usepackage[T1]{fontenc}
 \usepackage[latin1]{inputenc}
-\usepackage{geometry}
-\geometry{letterpaper,tmargin=1in,bmargin=1in,lmargin=1in,rmargin=1in}
+\usepackage[letterpaper]{geometry}
+\geometry{verbose,tmargin=1in,bmargin=1in,lmargin=1in,rmargin=1in}
 \setcounter{secnumdepth}{3}
 \setcounter{tocdepth}{3}
+\usepackage{babel}
 \usepackage{float}
 \usepackage{textcomp}
+\usepackage{url}
 \usepackage{amsmath}
-
-% figures
-\usepackage[pdftex]{graphicx}
-
-% we are running pdflatex, so convert .eps files to .pdf
-\usepackage{epstopdf}
-
 \usepackage{amssymb}
-\IfFileExists{url.sty}{\usepackage{url}}
-                      {\newcommand{\url}{\texttt}}
+\usepackage{graphicx}
 \usepackage[authoryear]{natbib}
+\usepackage[unicode=true,pdfusetitle,
+ bookmarks=true,bookmarksnumbered=true,bookmarksopen=false,
+ breaklinks=false,pdfborder={0 0 0},backref=section,colorlinks=false,pdfpagemode=None]
+ {hyperref}
+\hypersetup{
+ pdftex,pdfstartview=FitH,pdfpagelayout=SinglePage}
+\usepackage{breakurl}
 
@@
 \newcommand{\noun}[1]{\textsc{#1}}
-%% A simple dot to overcome graphicx limitations
-\newcommand{\lyxdot}{.}
 
@@
 \newenvironment{lyxcode}
-{\begin{list}{}{
+{\par\begin{list}{}{
 \setlength{\rightmargin}{\leftmargin}
@@
 {\end{list}}
-
 \newenvironment{lyxlist}[1]
@@
  \addtolength{\leftmargin}{\labelsep}
-\renewcommand{\makelabel}[1]{#1\hfil}}}
+ \renewcommand{\makelabel}[1]{##1\hfil}}}
 {\end{list}}
@@
 
-\usepackage{wrapfig}
-
 % fonts
@@
 % hyperlinks to sections and references
-\usepackage[pdftex,bookmarks=true,bookmarksnumbered=true,pdfpagemode=None,pdfstartview=FitH,pdfpagelayout=SinglePage,pdfborder={0 0 0}]{hyperref}
 
@@
 % colors for the text
-\usepackage[dvipsnames,usenames]{color}
+\@ifundefined{definecolor}
+ {\usepackage{color}}{}
 \newcommand{\red}[1]{\textcolor{Red}{#1}}
@@
 \usepackage{babel}
+
 \makeatother
@@
 \begin{center}
-\thispagestyle{empty}%
-\vspace*{-1.8truecm}
-\noindent\makebox[\textwidth]{%
-\noindent\includegraphics[width=0.83\paperwidth]{figures/specfem_3d_Cartesian-cover.pdf}
+\thispagestyle{empty}\vspace*{-1.8truecm} %
+\makebox[1\textwidth]{%
+\includegraphics[width=0.83\paperwidth]{figures/specfem_3d_Cartesian-cover} %
 } 
-\end{center}
-%
+\par\end{center}
+
 \title{\textbf{SPECFEM3D Cartesian}\\
@@
 
-\date{\noindent \today}
+\date{\today}
 
@@
 \section*{Authors}
-The SPECFEM3D Cartesian package was first developed by Dimitri Komatitsch and Jean-Pierre Vilotte at Institut de Physique du Globe (IPGP) in Paris, France
-from 1995 to 1997 and then by Dimitri Komatitsch and Jeroen Tromp at Harvard University and Caltech, USA, starting in 1998.
-The story started on March 28, 1995, when Prof. Yvon Maday from CNRS and University of Paris, France, gave a lecture to
-Dimitri Komatitsch and Jean-Pierre Vilotte at IPG about the nice properties of the spectral-element method that he had used for
-other equations. We are deeply indebted and thankful to him for that.
-
-Since then it has been developed and maintained by a development team: in alphabetical order,
-Jean-Paul (Pablo) Ampuero,
-Piero Basini,
-C\'eline Blitz,
-Ebru Bozda\u{g},
-Emanuele Casarotti,
-Joseph Charles,
-Min Chen,
-Percy Galvez,
-Dominik G\"oddeke,
-Vala Hj\"orleifsd\'ottir,
-Sue Kientz,
-Dimitri Komatitsch,
-Jes\'us Labarta,
-Nicolas Le Goff,
-Pieyre Le Loher,
-Qinya Liu,
-Yang Luo,
-Alessia Maggi,
-Federica Magnoni,
-Roland Martin,
-Ren\'e Matzen,
-Dennis McRitchie,
-Matthias Meschede,
-Peter Messmer,
-David Mich\'ea,
-Surendra Nadh Somala,
-Tarje Nissen-Meyer,
-Daniel Peter,
-Max Rietmann,
-Elliott Sales de Andrade,
-Brian Savage,
-Bernhard Schuberth,
-Anne Sieminski,
-Leif Strand,
-Carl Tape,
-Jeroen Tromp,
-Jean-Pierre Vilotte,
-Zhinan Xie,
-Hejun Zhu.\\
-
-The cover graphic of the manual was created
-by Santiago Lombeyda from Caltech's Center for Advanced Computing Research (CACR) \urlwithparentheses{http://www.cacr.caltech.edu}. \\
+
+\noindent The SPECFEM3D Cartesian package was first developed by Dimitri
+Komatitsch and Jean-Pierre Vilotte at Institut de Physique du Globe
+(IPGP) in Paris, France from 1995 to 1997 and then by Dimitri Komatitsch
+and Jeroen Tromp at Harvard University and Caltech, USA, starting
+in 1998. The story started on March 28, 1995, when Prof. Yvon Maday
+from CNRS and University of Paris, France, gave a lecture to Dimitri
+Komatitsch and Jean-Pierre Vilotte at IPG about the nice properties
+of the spectral-element method that he had used for other equations.
+We are deeply indebted and thankful to him for that.
+
+Since then it has been developed and maintained by a development team:
+in alphabetical order, Jean-Paul (Pablo) Ampuero, Piero Basini, Céline
+Blitz, Ebru Bozda\u{g}, Emanuele Casarotti, Joseph Charles, Min Chen,
+Percy Galvez, Dominik Göddeke, Vala Hjörleifsdóttir, Sue Kientz, Dimitri
+Komatitsch, Jesús Labarta, Nicolas Le Goff, Pieyre Le Loher, Qinya
+Liu, Yang Luo, Alessia Maggi, Federica Magnoni, Roland Martin, René
+Matzen, Dennis McRitchie, Matthias Meschede, Peter Messmer, David
+Michéa, Surendra Nadh Somala, Tarje Nissen-Meyer, Daniel Peter, Max
+Rietmann, Elliott Sales de Andrade, Brian Savage, Bernhard Schuberth,
+Anne Sieminski, Leif Strand, Carl Tape, Jeroen Tromp, Jean-Pierre
+Vilotte, Zhinan Xie, Hejun Zhu.\\
+
+The cover graphic of the manual was created by Santiago Lombeyda from
+Caltech's Center for Advanced Computing Research (CACR) \urlwithparentheses{http://www.cacr.caltech.edu}.
+\\
 
@@
 
-The software package SPECFEM3D Cartesian simulates seismic
-wave propagation at the local or regional scale
-based upon the spectral-element method (SEM).
-The SEM is a continuous Galerkin technique \citep{TrKoLi08,PeKoLuMaLeCaLeMaLiBlNiBaTr11},
+The software package SPECFEM3D Cartesian simulates seismic wave propagation
+at the local or regional scale based upon the spectral-element method
+(SEM). The SEM is a continuous Galerkin technique \citep{TrKoLi08,PeKoLuMaLeCaLeMaLiBlNiBaTr11},
 which can easily be made discontinous \citep{BeMaPa94,Ch00,KoWoHu02,ChCaVi03,LaWaBe05,Kop06,WiStBuGh10,AcKo11};
-it is then close to a particular case of the discontinuous Galerkin technique \citep{ReHi73,Arn82,FaRi99,HuHuRa99,CoKaSh00,GiHeWa02,RiWh03,MoRi05,GrScSc06,AiMoMu06,BeLaPi06,DuKa06,DeSeWh08,PuAmKa09,WiStBuGh10,DeSe10,EtChViGl10}, with optimized efficiency because of its tensorized basis functions \citep{WiStBuGh10,AcKo11}.
-In particular, it can accurately handle very distorted mesh elements \citep{OlSe11}.\\
+it is then close to a particular case of the discontinuous Galerkin
+technique \citep{ReHi73,Arn82,FaRi99,HuHuRa99,CoKaSh00,GiHeWa02,RiWh03,MoRi05,GrScSc06,AiMoMu06,BeLaPi06,DuKa06,DeSeWh08,PuAmKa09,WiStBuGh10,DeSe10,EtChViGl10},
+with optimized efficiency because of its tensorized basis functions
+\citep{WiStBuGh10,AcKo11}. In particular, it can accurately handle
+very distorted mesh elements \citep{OlSe11}.\\
 
 It has very good accuracy and convergence properties \citep{MaPa89,SePr94,DeFiMu02,Coh02,DeSe07,SeOl08}.
-The spectral element approach admits spectral rates of convergence and allows exploiting $hp$-convergence schemes.
-It is also very well suited to parallel implementation on very large supercomputers \citep{KoTsChTr03,TsKoChTr03,KoLaMi08a,CaKoLaTiMiLeSnTr08,KoViCh10} as well as on clusters of GPU accelerating graphics cards \citep{KoMiEr09,KoErGoMi10,Kom11}.
-Tensor products inside each element can be optimized to reach very high efficiency \citep{DeFiMu02}, and mesh point and element numbering can be optimized to reduce processor cache misses and improve cache reuse \citep{KoLaMi08a}.
-The SEM can also handle triangular (in 2D) or tetrahedral (in 3D) elements \citep{WinBoyd96,TaWi00,KoMaTrTaWi01,Coh02,MeViSa06} as well as mixed meshes, although with increased cost and reduced accuracy in these elements, as in the discontinuous Galerkin method.\\
-
-Note that in many geological models in the context of seismic wave propagation studies
-(except for instance for fault dynamic rupture studies, in which very high frequencies or supershear rupture need to be modeled near the fault, see e.g.
-\cite{BeGlCrViPi07,BeGlCrVi09,PuAmKa09,TaCrEtViBeSa10})
-a continuous formulation is sufficient because material property contrasts are not drastic and thus
-conforming mesh doubling bricks can efficiently handle mesh size variations \citep{KoTr02a,KoLiTrSuStSh04,LeChLiKoHuTr08,LeChKoHuTr09,LeKoHuTr09}.\\
-
-For a detailed introduction to the SEM as applied to regional
-seismic wave propagation, please consult \citet{PeKoLuMaLeCaLeMaLiBlNiBaTr11,TrKoLi08,KoVi98,KoTr99,ChKoViCaVaFe07} and
-in particular \citet{LeKoHuTr09,LeChKoHuTr09,LeChLiKoHuTr08,GoAmTaCaSmSaMaKo09,WiKoScTr04,KoLiTrSuStSh04}.
-A detailed theoretical analysis of the dispersion
-and stability properties of the SEM is available in \citet{Coh02}, \citet{DeSe07}, \citet{SeOl07} and \citet{SeOl08}.\\
+The spectral element approach admits spectral rates of convergence
+and allows exploiting $hp$-convergence schemes. It is also very well
+suited to parallel implementation on very large supercomputers \citep{KoTsChTr03,TsKoChTr03,KoLaMi08a,CaKoLaTiMiLeSnTr08,KoViCh10}
+as well as on clusters of GPU accelerating graphics cards \citep{KoMiEr09,KoErGoMi10,Kom11}.
+Tensor products inside each element can be optimized to reach very
+high efficiency \citep{DeFiMu02}, and mesh point and element numbering
+can be optimized to reduce processor cache misses and improve cache
+reuse \citep{KoLaMi08a}. The SEM can also handle triangular (in 2D)
+or tetrahedral (in 3D) elements \citep{WinBoyd96,TaWi00,KoMaTrTaWi01,Coh02,MeViSa06}
+as well as mixed meshes, although with increased cost and reduced
+accuracy in these elements, as in the discontinuous Galerkin method.\\
+
+Note that in many geological models in the context of seismic wave
+propagation studies (except for instance for fault dynamic rupture
+studies, in which very high frequencies or supershear rupture need
+to be modeled near the fault, see e.g. \citet{BeGlCrViPi07,BeGlCrVi09,PuAmKa09,TaCrEtViBeSa10})
+a continuous formulation is sufficient because material property contrasts
+are not drastic and thus conforming mesh doubling bricks can efficiently
+handle mesh size variations \citep{KoTr02a,KoLiTrSuStSh04,LeChLiKoHuTr08,LeChKoHuTr09,LeKoHuTr09}.\\
+
+For a detailed introduction to the SEM as applied to regional seismic
+wave propagation, please consult \citet{PeKoLuMaLeCaLeMaLiBlNiBaTr11,TrKoLi08,KoVi98,KoTr99,ChKoViCaVaFe07}
+and in particular \citet{LeKoHuTr09,LeChKoHuTr09,LeChLiKoHuTr08,GoAmTaCaSmSaMaKo09,WiKoScTr04,KoLiTrSuStSh04}.
+A detailed theoretical analysis of the dispersion and stability properties
+of the SEM is available in \citet{Coh02}, \citet{DeSe07}, \citet{SeOl07}
+and \citet{SeOl08}.\\
 
@@
 speed, density, a 3D crustal model, topography and bathymetry are
-included. The package can accommodate
-full 21-parameter anisotropy (see~\citet{ChTr07}) as well as lateral
-variations in attenuation \citep{SaKoTr10}. Adjoint capabilities and finite-frequency
-kernel simulations are included \citep{TrKoLi08,PeKoLuMaLeCaLeMaLiBlNiBaTr11,LiTr06,FiIgBuKe09,ViOp09}.\\
+included. The package can accommodate full 21-parameter anisotropy
+(see~\citet{ChTr07}) as well as lateral variations in attenuation
+\citep{SaKoTr10}. Adjoint capabilities and finite-frequency kernel
+simulations are included \citep{TrKoLi08,PeKoLuMaLeCaLeMaLiBlNiBaTr11,LiTr06,FiIgBuKe09,ViOp09}.\\
 
 The SEM was originally developed in computational fluid dynamics \citep{Pat84,MaPa89}
-and has been successfully adapted to address problems in seismic wave propagation.
-Early seismic wave propagation applications of the SEM, utilizing Legendre basis functions and a
-perfectly diagonal mass matrix, include \cite{CoJoTo93}, \cite{Kom97},
-\cite{FaMaPaQu97}, \cite{CaGa97}, \cite{KoVi98} and \cite{KoTr99},
-whereas applications involving Chebyshev basis functions and a nondiagonal mass matrix
-include \cite{SePr94}, \cite{PrCaSe94} and \cite{SePrPr95}.\\
+and has been successfully adapted to address problems in seismic wave
+propagation. Early seismic wave propagation applications of the SEM,
+utilizing Legendre basis functions and a perfectly diagonal mass matrix,
+include \citet{CoJoTo93}, \citet{Kom97}, \citet{FaMaPaQu97}, \citet{CaGa97},
+\citet{KoVi98} and \citet{KoTr99}, whereas applications involving
+Chebyshev basis functions and a nondiagonal mass matrix include \citet{SePr94},
+\citet{PrCaSe94} and \citet{SePrPr95}.\\
 
@@
 in mind, and conforms strictly to the Fortran2003 standard. It uses
-no obsolete or obsolescent features of Fortran. The package uses
-parallel programming based upon the Message Passing Interface (MPI)
-\citep{GrLuSk94,Pac97}.\\
+no obsolete or obsolescent features of Fortran. The package uses parallel
+programming based upon the Message Passing Interface (MPI) \citep{GrLuSk94,Pac97}.\\
 
 SPECFEM3D won the Gordon Bell award for best performance at the SuperComputing~2003
-conference in Phoenix, Arizona (USA) (see \cite{KoTsChTr03}
-and \url{www.sc-conference.org/sc2003/nr_finalaward.html}).
-It was a finalist again in 2008 for a run at 0.16 petaflops (sustained) on 149,784 processors of the `Jaguar' Cray XT5 system at Oak Ridge National Laboratories (USA) \citep{CaKoLaTiMiLeSnTr08}.
-It also won the BULL Joseph Fourier supercomputing award in 2010.\\
-
-This new release of the code (version 2.1)
-includes Convolution or Auxiliary Differential Equation Perfectly Matched absorbing Layers
+conference in Phoenix, Arizona (USA) (see \citet{KoTsChTr03} and
+\url{www.sc-conference.org/sc2003/nrfinalaward.html}). It was a finalist
+again in 2008 for a run at 0.16 petaflops (sustained) on 149,784 processors
+of the `Jaguar' Cray XT5 system at Oak Ridge National Laboratories
+(USA) \citep{CaKoLaTiMiLeSnTr08}. It also won the BULL Joseph Fourier
+supercomputing award in 2010.\\
+
+This new release of the code (version 2.1) includes Convolution or
+Auxiliary Differential Equation Perfectly Matched absorbing Layers
 (C-PML or ADE-PML) \citep{MaKoEz08,MaKoGe08,MaKo09,MaKoGeBr10,KoMa07}.
@@
 
-The next release of the code will use the PT-SCOTCH parallel and threaded version of SCOTCH for mesh partitioning instead of the serial version.\\
-
-SPECFEM3D Cartesian includes coupled fluid-solid domains and
-adjoint capabilities, which enables one to address seismological inverse problems, but for linear rheologies only so far.
-To accommodate visco-plastic or non-linear rheologies, the reader can refer to
-the \href{http://geoelse.stru.polimi.it/}{GeoELSE} software package \citep{CaGa97,StPaIg09}.
+The next release of the code will use the PT-SCOTCH parallel and threaded
+version of SCOTCH for mesh partitioning instead of the serial version.\\
+
+SPECFEM3D Cartesian includes coupled fluid-solid domains and adjoint
+capabilities, which enables one to address seismological inverse problems,
+but for linear rheologies only so far. To accommodate visco-plastic
+or non-linear rheologies, the reader can refer to the \href{http://geoelse.stru.polimi.it/}{GeoELSE}software
+package \citep{CaGa97,StPaIg09}.
 
@@
 
-If you use SPECFEM3D Cartesian for your own research, please cite at least one
-of the following articles:
-
+If you use SPECFEM3D Cartesian for your own research, please cite
+at least one of the following articles:
 \begin{description}
-
-\item [Numerical simulations in general] ~\\
-Forward and adjoint simulations are described in detail in \cite{TrKoLi08,PeKoLuMaLeCaLeMaLiBlNiBaTr11,VaCaSaKoVi99,KoMiEr09,KoErGoMi10,ChKoViCaVaFe07,MaKoDi09,KoViCh10,CaKoLaTiMiLeSnTr08,TrKoHjLiZhPeBoMcFrTrHu10,KoRiTr02,KoTr02a,KoTr02b,KoTr99} or \cite{KoVi98}.
-Additional aspects of adjoint simulations are described in
-\cite{TrTaLi05,LiTr06,TrKoLi08,LiTr08,TrKoHjLiZhPeBoMcFrTrHu10,PeKoLuMaLeCaLeMaLiBlNiBaTr11}.
-Domain decomposition is explained in detail in \cite{MaKoBlLe08}, and excellent scaling up to 150,000 processor cores is shown for instance in
-\cite{CaKoLaTiMiLeSnTr08,KoLaMi08a,MaKoBlLe08,KoErGoMi10,Kom11},
-
-\item [GPU computing] ~\\
-Computing on GPU graphics cards for acoustic or seismic wave propagation applications
-is described in detail in \cite{KoMiEr09,KoErGoMi10,MiKo10,Kom11}.
-
+\item [{Numerical simulations in general}] ~\\
+ Forward and adjoint simulations are described in detail in \citet{TrKoLi08,PeKoLuMaLeCaLeMaLiBlNiBaTr11,VaCaSaKoVi99,KoMiEr09,KoErGoMi10,ChKoViCaVaFe07,MaKoDi09,KoViCh10,CaKoLaTiMiLeSnTr08,TrKoHjLiZhPeBoMcFrTrHu10,KoRiTr02,KoTr02a,KoTr02b,KoTr99}
+or \citet{KoVi98}. Additional aspects of adjoint simulations are
+described in \citet{TrTaLi05,LiTr06,TrKoLi08,LiTr08,TrKoHjLiZhPeBoMcFrTrHu10,PeKoLuMaLeCaLeMaLiBlNiBaTr11}.
+Domain decomposition is explained in detail in \citet{MaKoBlLe08},
+and excellent scaling up to 150,000 processor cores is shown for instance
+in \citet{CaKoLaTiMiLeSnTr08,KoLaMi08a,MaKoBlLe08,KoErGoMi10,Kom11},
+\item [{GPU computing}] ~\\
+ Computing on GPU graphics cards for acoustic or seismic wave propagation
+applications is described in detail in \citet{KoMiEr09,KoErGoMi10,MiKo10,Kom11}.
 \end{description}
-
-\noindent
-If you use this new version 2.1, which has non blocking MPI for much better performance for medium or large runs, please cite at least one of these six articles,
-in which results of non blocking MPI runs are presented: \cite{PeKoLuMaLeCaLeMaLiBlNiBaTr11,KoErGoMi10,KoViCh10,Kom11,CaKoLaTiMiLeSnTr08,MaKoBlLe08}.\\
-
-\noindent
-If you work on geophysical applications, you may be interested in citing some of these application articles as well, among others:
-
+\noindent If you use this new version 2.1, which has non blocking
+MPI for much better performance for medium or large runs, please cite
+at least one of these six articles, in which results of non blocking
+MPI runs are presented: \citet{PeKoLuMaLeCaLeMaLiBlNiBaTr11,KoErGoMi10,KoViCh10,Kom11,CaKoLaTiMiLeSnTr08,MaKoBlLe08}.\\
+
+\noindent If you work on geophysical applications, you may be interested
+in citing some of these application articles as well, among others:
 \begin{description}
-
-\item [Southern California simulations] ~\\
-\cite{KoLiTrSuStSh04,KrJiKoTr06a,KrJiKoTr06b}.
-
-If you use the 3D southern California model, please cite
-\citet{SuSh03} (Los Angeles model),
-\citet{lovelyetal06} (Salton Trough), and
-\citet{hauksson2000} (southern California).
-The Moho map was determined by
-\citet{zhu&kanamori2000}.
+\item [{Southern California simulations}] ~\\
+ \citet{KoLiTrSuStSh04,KrJiKoTr06a,KrJiKoTr06b}.
+
+If you use the 3D southern California model, please cite \citet{SuSh03}
+(Los Angeles model), \citet{lovelyetal06} (Salton Trough), and \citet{hauksson2000}
+(southern California). The Moho map was determined by \citet{zhu&kanamori2000}.
 The 1D SoCal model was developed by \citet{DrHe90}.
 
-\item [Anisotropy] ~\\
-\cite{ChTr07,JiTsKoTr05,ChFaKo04,FaChKo04,RiRiKoTrHe02,TrKo00}.
-
-\item [Attenuation] ~\\
-\cite{SaKoTr10,KoTr02a,KoTr99}.
-
-\item [Topography] ~\\
-\cite{LeKoHuTr09,LeChKoHuTr09,LeChLiKoHuTr08,GoAmTaCaSmSaMaKo09,WiKoScTr04}.
-
+\item [{Anisotropy}] ~\\
+ \citet{ChTr07,JiTsKoTr05,ChFaKo04,FaChKo04,RiRiKoTrHe02,TrKo00}.
+\item [{Attenuation}] ~\\
+ \citet{SaKoTr10,KoTr02a,KoTr99}.
+\item [{Topography}] ~\\
+ \citet{LeKoHuTr09,LeChKoHuTr09,LeChLiKoHuTr08,GoAmTaCaSmSaMaKo09,WiKoScTr04}.
 \end{description}
-
-The corresponding Bib\TeX{} entries may be found
-in file \texttt{doc/USER\_MANUAL/bibliography.bib}.
+The corresponding Bib\TeX{} entries may be found in file \texttt{doc/USER\_MANUAL/bibliography.bib}.
 
@@
 Grant No. ANR-05-CIGC-002, and European FP6 Marie Curie International
-Reintegration Grant No. MIRG-CT-2005-017461.
-Any opinions, findings, and conclusions or recommendations expressed in this material are
+Reintegration Grant No. MIRG-CT-2005-017461. Any opinions, findings,
+and conclusions or recommendations expressed in this material are
 those of the authors and do not necessarily reflect the views of the
@@
 
-The SPECFEM3D Cartesian software package comes in a gzipped tar ball. In the
-directory in which you want to install the package, type
+The SPECFEM3D Cartesian software package comes in a gzipped tar ball.
+In the directory in which you want to install the package, type 
 \begin{lyxcode}
@@
 \end{lyxcode}
-The directory \texttt{\small SPECFEM3D/} will then contain
-the source code.\\
-
-We recommend that you add {\texttt{ulimit -S -s unlimited}} to your {\texttt{.bash\_profile}} file and/or {\texttt{limit stacksize unlimited }} to your {\texttt{.cshrc}} file to suppress any potential limit to the size of the Unix stack.\\
-
-Then, to configure the software for your system, run the
-\texttt{configure} shell script. This script will attempt to guess
-the appropriate configuration values for your system. However, at
-a minimum, it is recommended that you explicitly specify the appropriate
-command names for your Fortran compiler and MPI package (another option is to define FC, CC and MPIF90 in your .bash\_profile 
-or your .cshrc file):
-
+The directory \texttt{\small SPECFEM3D/}{\small{} will then contain
+the source code.}\\
+{\small \par}
+
+We recommend that you add {\texttt{ulimit -S -s unlimited}} to your
+{\texttt{.bash\_profile}} file and/or {\texttt{limit stacksize
+unlimited }} to your {\texttt{.cshrc}} file to suppress any potential
+limit to the size of the Unix stack.\\
+
+Then, to configure the software for your system, run the \texttt{configure}
+shell script. This script will attempt to guess the appropriate configuration
+values for your system. However, at a minimum, it is recommended that
+you explicitly specify the appropriate command names for your Fortran
+compiler and MPI package (another option is to define FC, CC and MPIF90
+in your .bash\_profile or your .cshrc file):
 \begin{lyxcode}
@@
 \end{lyxcode}
-
-Before running the \texttt{configure} script, you should probably edit file \texttt{flags.guess} to make sure that it contains the best compiler options for your system.
-Known issues or things to check are:
-
+Before running the \texttt{configure} script, you should probably
+edit file \texttt{flags.guess} to make sure that it contains the best
+compiler options for your system. Known issues or things to check
+are:
 \begin{description}
-\item [{\texttt{Intel ifort compiler}}] See if you need to add \texttt{-assume byterecl} for your machine.
-\item [{\texttt{IBM compiler}}] See if you need to add \texttt{-qsave} or \texttt{-qnosave} for your machine.
+\item [{\texttt{Intel ifort compiler}}] See if you need to add \texttt{-assume
+byterecl} for your machine. 
+\item [{\texttt{IBM compiler}}] See if you need to add \texttt{-qsave}
+or \texttt{-qnosave} for your machine. 
 \item [{\texttt{Mac OS}}] You will probably need to install \texttt{XCODE}.
-In addition, the \texttt{clock\_gettime} routine, which is used by the \texttt{SCOTCH} library that we use, does not exist in Mac OS.
-You will need to replace it with \texttt{clock\_get\_time} if you want to use \texttt{SCOTCH}.
+In addition, the \texttt{clock\_gettime} routine, which is used by
+the \texttt{SCOTCH} library that we use, does not exist in Mac OS.
+You will need to replace it with \texttt{clock\_get\_time} if you
+want to use \texttt{SCOTCH}. 
 \end{description}
-
-When compiling on an IBM machine with the \texttt{xlf} and \texttt{xlc} compilers, we suggest running the \texttt{configure} script
-with the following options:\\
-
-\noindent\texttt{./configure FC=xlf90\_r MPIFC=mpif90 CC=xlc\_r CFLAGS="-O3 -q64" FCFLAGS="-O3 -q64" -with-scotch-dir=...}.\\
+When compiling on an IBM machine with the \texttt{xlf} and \texttt{xlc}
+compilers, we suggest running the \texttt{configure} script with the
+following options:\\
+
+\noindent \texttt{./configure FC=xlf90\_r MPIFC=mpif90 CC=xlc\_r CFLAGS=\textquotedbl{}-O3
+-q64\textquotedbl{} FCFLAGS=\textquotedbl{}-O3 -q64\textquotedbl{}
+-with-scotch-dir=...}.\\
 
@@
 
-If you run very large meshes on a relatively small number
-of processors, the memory size needed on each processor might become
-greater than 2 gigabytes, which is the upper limit for 32-bit addressing.
-In this case, on some compilers you may need to add \texttt{``-mcmodel=medium}'' or \texttt{``-mcmodel=medium -shared-intel}''
-to the configure options of CFLAGS, FCFLAGS and LDFLAGS otherwise the compiler will display an error
-message (for instance \texttt{``relocation truncated to fit: R\_X86\_64\_PC32 against .bss''} or something similar);
-on an IBM machine with the \texttt{xlf} and \texttt{xlc} compilers, using \texttt{-q64} is usually sufficient.\\
-
-Note that we use CUBIT to create meshes of hexahedra, but other packages can be used as well, for instance GiD from \url{http://gid.cimne.upc.es} or Gmsh from \url{http://geuz.org/gmsh} \citep{GeRe09}.
-Even mesh creation packages that generate tetrahedra, for instance TetGen from \url{http://tetgen.berlios.de}, can be used because each tetrahedron can then easily be decomposed into four hexahedra as shown in the picture of the TetGen logo at \url{http://tetgen.berlios.de/figs/Delaunay-Voronoi-3D.gif}; while this approach does not generate hexahedra of optimal quality, it can ease mesh creation in some situations and it has been shown that the spectral-element method can very accurately handle distorted mesh elements \citep{OlSe11}.
-
-The SPECFEM3D Cartesian software package relies on the SCOTCH library to partition meshes created with CUBIT.
-METIS \citep{KaKu98a,KaKu98c,KaKu98b} can also be used instead of SCOTCH if you prefer, by editing file src/decompose\_mesh/decompose\_mesh.F90 and uncommenting the flag USE\_METIS\_INSTEAD\_OF\_SCOTCH.
-You will also then need to install and compile Metis version 4.0 (do *NOT* install Metis version 5.0, which has incompatible function calls) and edit \texttt{src/decompose\_mesh/Makefile.in} and uncomment the METIS link flag in that file before running \texttt{configure}.\\
-
-The SCOTCH library \citep{PeRo96}
-provides efficent static mapping, graph and mesh partitioning routines. SCOTCH is a free software package developed by
-Fran\c {c}ois Pellegrini et al. from LaBRI and INRIA in Bordeaux, France, downloadable from the web page \url{https://gforge.inria.fr/projects/scotch/}.
-In case no SCOTCH libraries can be found on the system, the configuration will bundle the version provided with the source code for compilation.
-The path to an existing SCOTCH installation can to be set explicitly with the option \texttt{-{}-with-scotch-dir}. Just as an example:
+If you run very large meshes on a relatively small number of processors,
+the memory size needed on each processor might become greater than
+2 gigabytes, which is the upper limit for 32-bit addressing. In this
+case, on some compilers you may need to add \texttt{``-mcmodel=medium}''
+or \texttt{``-mcmodel=medium -shared-intel}'' to the configure options
+of CFLAGS, FCFLAGS and LDFLAGS otherwise the compiler will display
+an error message (for instance \texttt{``relocation truncated to
+fit: R\_X86\_64\_PC32 against .bss''} or something similar); on an
+IBM machine with the \texttt{xlf} and \texttt{xlc} compilers, using
+\texttt{-q64} is usually sufficient.\\
+
+Note that we use CUBIT to create meshes of hexahedra, but other packages
+can be used as well, for instance GiD from \url{http://gid.cimne.upc.es}
+or Gmsh from \url{http://geuz.org/gmsh} \citep{GeRe09}. Even mesh
+creation packages that generate tetrahedra, for instance TetGen from
+\url{http://tetgen.berlios.de}, can be used because each tetrahedron
+can then easily be decomposed into four hexahedra as shown in the
+picture of the TetGen logo at \url{http://tetgen.berlios.de/figs/Delaunay-Voronoi-3D.gif};
+while this approach does not generate hexahedra of optimal quality,
+it can ease mesh creation in some situations and it has been shown
+that the spectral-element method can very accurately handle distorted
+mesh elements \citep{OlSe11}.
+
+The SPECFEM3D Cartesian software package relies on the SCOTCH library
+to partition meshes created with CUBIT. METIS \citep{KaKu98a,KaKu98c,KaKu98b}
+can also be used instead of SCOTCH if you prefer, by editing file
+src/decompose\_mesh/decompose\_mesh.F90 and uncommenting the flag
+USE\_METIS\_INSTEAD\_OF\_SCOTCH. You will also then need to install
+and compile Metis version 4.0 (do {*}NOT{*} install Metis version
+5.0, which has incompatible function calls) and edit \texttt{src/decompose\_mesh/Makefile.in}
+and uncomment the METIS link flag in that file before running \texttt{configure}.\\
+
+The SCOTCH library \citep{PeRo96} provides efficent static mapping,
+graph and mesh partitioning routines. SCOTCH is a free software package
+developed by François Pellegrini et al. from LaBRI and INRIA in Bordeaux,
+France, downloadable from the web page \url{https://gforge.inria.fr/projects/scotch/}.
+In case no SCOTCH libraries can be found on the system, the configuration
+will bundle the version provided with the source code for compilation.
+The path to an existing SCOTCH installation can to be set explicitly
+with the option \texttt{-{}-with-scotch-dir}. Just as an example: 
 \begin{lyxcode}
@@
 \end{lyxcode}
-
-If you use the Intel ifort compiler to compile the code, we recommend that you use the Intel icc C compiler to compile Scotch, i.e., use:
+If you use the Intel ifort compiler to compile the code, we recommend
+that you use the Intel icc C compiler to compile Scotch, i.e., use: 
 \begin{lyxcode}
@@
 \end{lyxcode}
-
 To compile a serial version of the code for small meshes that fits
@@
 A summary of the most important configuration variables follows.
-
 \begin{description}
@@
 \end{description}
-The configuration script automatically creates for each executable a corresponding
-\texttt{Makefile} in the \texttt{src/} subdirectoy.
+The configuration script automatically creates for each executable
+a corresponding \texttt{Makefile} in the \texttt{src/} subdirectoy.
 The \texttt{Makefile} contains a number of suggested entries for various
@@
 system:
-
 \begin{description}
@@
 on most cluster applications. For reasons of speed, the (parallel)
-distributed database generator typically writes a (parallel) database for the solver on the
-local disks of the compute nodes. Some systems have no local disks,
-e.g., BlueGene or the Earth Simulator, and other systems have a fast
-parallel file system, in which case this flag should be set to \texttt{.true.}.
-Note that this flag is not used by the database generator or the solver; it is
-only used for some of the post-processing.
+distributed database generator typically writes a (parallel) database
+for the solver on the local disks of the compute nodes. Some systems
+have no local disks, e.g., BlueGene or the Earth Simulator, and other
+systems have a fast parallel file system, in which case this flag
+should be set to \texttt{.true.}. Note that this flag is not used
+by the database generator or the solver; it is only used for some
+of the post-processing. 
 \end{description}
-The package can run either
-in single or in double precision mode. The default is single precision
-because for almost all calculations performed using the spectral-element method
-using single precision is sufficient and gives the same results (i.e. the same seismograms);
-and the single precision code is faster and requires exactly half as much memory.
-Select your preference by selecting the appropriate setting in the \texttt{constants.h} file:
-
+The package can run either in single or in double precision mode.
+The default is single precision because for almost all calculations
+performed using the spectral-element method using single precision
+is sufficient and gives the same results (i.e. the same seismograms);
+and the single precision code is faster and requires exactly half
+as much memory. Select your preference by selecting the appropriate
+setting in the \texttt{constants.h} file:
 \begin{description}
@@
 In the \texttt{precision.h} file:
-
 \begin{description}
@@
 \end{description}
-On many current processors (e.g., Intel, AMD, IBM Power), single precision calculations
-are significantly faster; the difference can typically be 10\%
-to 25\%. It is therefore better to use single precision.
-What you can do once for the physical problem you want to study is run the same calculation in single precision
-and in double precision on your system and compare the seismograms.
-If they are identical (and in most cases they will), you can select single precision for your future runs.\\
+On many current processors (e.g., Intel, AMD, IBM Power), single precision
+calculations are significantly faster; the difference can typically
+be 10\% to 25\%. It is therefore better to use single precision. What
+you can do once for the physical problem you want to study is run
+the same calculation in single precision and in double precision on
+your system and compare the seismograms. If they are identical (and
+in most cases they will), you can select single precision for your
+future runs.\\
 
@@
 
-OpenMP support can be enabled in addition to MPI. However, in many cases performance will not improve
-because our pure MPI implementation is already heavily optimized and thus the resulting code will in fact
-be slightly slower. A possible exception could be IBM BlueGene-type architectures.
-
-To enable OpenMP, uncomment the OpenMP compiler option in two lines in file \texttt{src/specfem3D/Makefile.in}
-(before running \texttt{configure}) and also uncomment the \texttt{\#define USE\_OPENMP} statement in file\\
+OpenMP support can be enabled in addition to MPI. However, in many
+cases performance will not improve because our pure MPI implementation
+is already heavily optimized and thus the resulting code will in fact
+be slightly slower. A possible exception could be IBM BlueGene-type
+architectures.
+
+To enable OpenMP, uncomment the OpenMP compiler option in two lines
+in file \texttt{src/specfem3D/Makefile.in} (before running \texttt{configure})
+and also uncomment the \texttt{\#define USE\_OPENMP} statement in
+file\\
  \texttt{src/specfem3D/specfem3D.F90}.
@@
 environment variable can set the scheduling policy of that DO-loop.
-Tests performed by Marcin Zielinski at SARA (The Netherlands) showed that often
-the best scheduling policy is DYNAMIC with the size of the chunk equal to the number of
-OpenMP threads, but most preferably being twice as the number of
-OpenMP threads (thus chunk size = 8 for 4 OpenMP threads etc).
-If OMP\_SCHEDULE is not set or is empty, the DO-loop will assume generic
-scheduling policy, which will slow down the job quite a bit.
+Tests performed by Marcin Zielinski at SARA (The Netherlands) showed
+that often the best scheduling policy is DYNAMIC with the size of
+the chunk equal to the number of OpenMP threads, but most preferably
+being twice as the number of OpenMP threads (thus chunk size = 8 for
+4 OpenMP threads etc). If OMP\_SCHEDULE is not set or is empty, the
+DO-loop will assume generic scheduling policy, which will slow down
+the job quite a bit.
 
@@
 
-\noindent
-Edit file \texttt{flags.guess} and put this for \texttt{FLAGS\_CHECK}:\\
-
-\noindent
-\texttt{-g -qfullpath -O2 -qsave -qstrict -qtune=qp -qarch=qp -qcache=auto -qhalt=w}\\
-\noindent
-\texttt{-qfree=f90 -qsuffix=f=f90 -qlanglvl=95pure -Q -Q+rank,swap\_all -Wl,-relax}\\
-
-\noindent
-The most relevant are the -qarch and -qtune flags, otherwise if these flags are set to ``auto'' then they are wrongly assigned to
-the architecture of the frond-end node, which is different from that on the compute nodes.
-You will need to set these flags to the right architecture for your BlueGene compute nodes, which is not necessarily ``qp'';
-ask your system administrator.
-On some machines if is necessary to use -O2 in these flags instead of -O3 due to a compiler bug of the XLF version installed.
-We thus suggest to first try -O3, and then if the code does not compile or does not run fine then switch back to -O2.
-The debug flags (-g, -qfullpath) do not influence performance but are useful to get at least some insights in case of problems.\\
-
-\noindent
-Before running \texttt{configure}, select the XL Fortran compiler by typing \texttt{module load bgq-xl/1.0}
-or \texttt{module load bgq-xl} (another, less efficient option is to load the GNU compilers using \texttt{module load bgq-gnu/4.4.6} or similar).\\
-
-\noindent
-Then, to configure the code, type this:\\
-
-\noindent
-\texttt{./configure FC=bgxlf90\_r MPIFC=mpixlf90\_r CC=bgxlc\_r LOCAL\_PATH\_IS\_ALSO\_GLOBAL=true}\\
-
-In order for the SCOTCH domain decomposer to compile, on some (but not all) Blue Gene systems you may need to
-run \texttt{configure} with \texttt{CC=gcc} instead of \texttt{CC=bgxlc\_r}.
-
-\noindent
-\underline{Older installation instruction for IBM BlueGene, from 2011:}\\
-
-\noindent
-To compile the code on an IBM BlueGene, Laurent L\'eger from IDRIS, France, suggests the following: compile the code with
-
-\noindent
-\texttt{   FLAGS\_CHECK="-O3 -qsave -qstrict -qtune=auto -qarch=450d -qcache=auto \\}
-\texttt{   -qfree=f90 -qsuffix=f=f90 -g -qlanglvl=95pure -qhalt=w -Q -Q+rank,swap\_all -Wl,-relax"}
-
-\noindent
-Option "-Wl,-relax" must be added on many (but not all) BlueGene systems to be able to link the binaries \texttt{xmeshfem3D}
-and \texttt{xspecfem3D} because the final link step is done by the GNU \texttt{ld} linker even if
-one uses \texttt{FC=bgxlf90\_r, MPIFC=mpixlf90\_r} and \texttt{CC=bgxlc\_r} to create all the object files.
-On the contrary, on some BlueGene systems that use the native AIX linker option "-Wl,-relax" can lead to problems and must be suppressed from \texttt{flags.guess}.
-
-\noindent
-Also, \texttt{AR=ar, ARFLAGS=cru} and \texttt{RANLIB=ranlib} are hardwired in all \texttt{Makefile.in} files by default, but to cross-compile on BlueGene/P one needs to change
-these values to \texttt{AR=bgar, ARFLAGS=cru} and \texttt{RANLIB=bgranlib}. Thus the easiest thing to do is to modify
-all \texttt{Makefile.in} files and the \texttt{configure} script to set them automatically by \texttt{configure}.
-One then just needs to pass the right commands to the \texttt{configure} script:
-
-\noindent
-\texttt{   ./configure --prefix=/path/to/SPECFEM3DG\_SP --host=Babel --build=BGP \\}
-\texttt{      FC=bgxlf90\_r MPIFC=mpixlf90\_r CC=bgxlc\_r AR=bgar ARFLAGS=cru \\}
+\noindent Edit file \texttt{flags.guess} and put this for \texttt{FLAGS\_CHECK}:\\
+
+\noindent \texttt{-g -qfullpath -O2 -qsave -qstrict -qtune=qp -qarch=qp
+-qcache=auto -qhalt=w}\\
+ \texttt{-qfree=f90 -qsuffix=f=f90 -qlanglvl=95pure -Q -Q+rank,swap\_all
+-Wl,-relax}\\
+
+\noindent The most relevant are the -qarch and -qtune flags, otherwise
+if these flags are set to ``auto'' then they are wrongly assigned
+to the architecture of the frond-end node, which is different from
+that on the compute nodes. You will need to set these flags to the
+right architecture for your BlueGene compute nodes, which is not necessarily
+``qp''; ask your system administrator. On some machines if is necessary
+to use -O2 in these flags instead of -O3 due to a compiler bug of
+the XLF version installed. We thus suggest to first try -O3, and then
+if the code does not compile or does not run fine then switch back
+to -O2. The debug flags (-g, -qfullpath) do not influence performance
+but are useful to get at least some insights in case of problems.\\
+
+\noindent Before running \texttt{configure}, select the XL Fortran
+compiler by typing \texttt{module load bgq-xl/1.0} or \texttt{module
+load bgq-xl} (another, less efficient option is to load the GNU compilers
+using \texttt{module load bgq-gnu/4.4.6} or similar).\\
+
+\noindent Then, to configure the code, type this:\\
+
+\noindent \texttt{./configure FC=bgxlf90\_r MPIFC=mpixlf90\_r CC=bgxlc\_r
+LOCAL\_PATH\_IS\_ALSO\_GLOBAL=true}\\
+
+In order for the SCOTCH domain decomposer to compile, on some (but
+not all) Blue Gene systems you may need to run \texttt{configure}
+with \texttt{CC=gcc} instead of \texttt{CC=bgxlc\_r}.
+
+\noindent \underline{Older installation instruction for IBM BlueGene, from 2011:}\\
+
+\noindent To compile the code on an IBM BlueGene, Laurent Léger from
+IDRIS, France, suggests the following: compile the code with
+
+\noindent \texttt{ FLAGS\_CHECK=\textquotedbl{}-O3 -qsave -qstrict
+-qtune=auto -qarch=450d -qcache=auto }~\\
+ \texttt{ -qfree=f90 -qsuffix=f=f90 -g -qlanglvl=95pure -qhalt=w -Q
+-Q+rank,swap\_all -Wl,-relax\textquotedbl{}}
+
+\noindent Option \textquotedbl{}-Wl,-relax\textquotedbl{} must be
+added on many (but not all) BlueGene systems to be able to link the
+binaries \texttt{xmeshfem3D} and \texttt{xspecfem3D} because the final
+link step is done by the GNU \texttt{ld} linker even if one uses \texttt{FC=bgxlf90\_r,
+MPIFC=mpixlf90\_r} and \texttt{CC=bgxlc\_r} to create all the object
+files. On the contrary, on some BlueGene systems that use the native
+AIX linker option \textquotedbl{}-Wl,-relax\textquotedbl{} can lead
+to problems and must be suppressed from \texttt{flags.guess}.
+
+\noindent Also, \texttt{AR=ar, ARFLAGS=cru} and \texttt{RANLIB=ranlib}
+are hardwired in all \texttt{Makefile.in} files by default, but to
+cross-compile on BlueGene/P one needs to change these values to \texttt{AR=bgar,
+ARFLAGS=cru} and \texttt{RANLIB=bgranlib}. Thus the easiest thing
+to do is to modify all \texttt{Makefile.in} files and the \texttt{configure}
+script to set them automatically by \texttt{configure}. One then just
+needs to pass the right commands to the \texttt{configure} script:
+
+\noindent \texttt{ ./configure -{}-prefix=/path/to/SPECFEM3DG\_SP
+-{}-host=Babel -{}-build=BGP }~\\
+ \texttt{ FC=bgxlf90\_r MPIFC=mpixlf90\_r CC=bgxlc\_r AR=bgar ARFLAGS=cru
+}~\\
  \texttt{ RANLIB=bgranlib LOCAL\_PATH\_IS\_ALSO\_GLOBAL=false}
 
-\noindent
-This trick can be useful for all hosts on which one needs to cross-compile.
-
-\noindent
-On BlueGene, one also needs to run the \texttt{xcreate\_header\_file} binary file manually rather than in the Makefile:
-
-\noindent
-   \texttt{bgrun -np 1 -mode VN -exe ./bin/xcreate\_header\_file}
+\noindent This trick can be useful for all hosts on which one needs
+to cross-compile.
+
+\noindent On BlueGene, one also needs to run the \texttt{xcreate\_header\_file}
+binary file manually rather than in the Makefile:
+
+\noindent \texttt{bgrun -np 1 -mode VN -exe ./bin/xcreate\_header\_file}
 
@@
 
-Packages such as \texttt{doxywizard} can be used to visualize the subroutine calling tree of the source code.
-\texttt{Doxywizard} is a GUI front-end for configuring and running \texttt{doxygen}.
+Packages such as \texttt{doxywizard} can be used to visualize the
+subroutine calling tree of the source code. \texttt{Doxywizard} is
+a GUI front-end for configuring and running \texttt{doxygen}.
 
@@
 
-The first step in running a spectral-element simulation consists of constructing a high-quality mesh for the region under consideration.
-We provide two possibilities to do so: (1) relying on the external, hexahedral mesher CUBIT, or (2) using the provided, internal mesher \texttt{xmeshfem3D}.
-In the following, we explain these two approaches.
+The first step in running a spectral-element simulation consists of
+constructing a high-quality mesh for the region under consideration.
+We provide two possibilities to do so: (1) relying on the external,
+hexahedral mesher CUBIT, or (2) using the provided, internal mesher
+\texttt{xmeshfem3D}. In the following, we explain these two approaches.
 
@@
 
-CUBIT is a meshing tool suite for the creation of finite-element meshes for arbitrarily shaped models.
-It has been developed and maintained at Sandia National Laboratories and can be purchased for a small academic institutional fee at \url{http://cubit.sandia.gov}.
-Our experience showed that using CUBIT greatly facilitates and speeds up the generation and preparation of hexahedral, conforming meshes for a variety of geophysical models with increasing complexity.
+CUBIT is a meshing tool suite for the creation of finite-element meshes
+for arbitrarily shaped models. It has been developed and maintained
+at Sandia National Laboratories and can be purchased for a small academic
+institutional fee at \url{http://cubit.sandia.gov}. Our experience
+showed that using CUBIT greatly facilitates and speeds up the generation
+and preparation of hexahedral, conforming meshes for a variety of
+geophysical models with increasing complexity.
 
@@
 \noindent \begin{centering}
-\includegraphics[width=.6\textwidth]{figures/mount-cubit.jpg}
+\includegraphics[width=0.6\textwidth]{figures/mount-cubit} 
 \par\end{centering}
 
-\caption{Example of the graphical user interface of CUBIT. The hexahedral mesh shown in the main display consists of a hexahedral discretization of a single volume with topography.}
+\caption{Example of the graphical user interface of CUBIT. The hexahedral mesh
+shown in the main display consists of a hexahedral discretization
+of a single volume with topography.}
 
@@
 
-The basic steps in creating a load-balanced, partitioned mesh with CUBIT are:
+The basic steps in creating a load-balanced, partitioned mesh with
+CUBIT are: 
 \begin{description}
-\item [1.] setting up a hexahedral mesh with CUBIT,
-\item [2.] exporting the CUBIT mesh into a SPECFEM3D Cartesian file format and
-\item [3.] partitioning the SPECFEM3D Cartesian mesh files for a chosen number of cores.
+\item [{1.}] setting up a hexahedral mesh with CUBIT, 
+\item [{2.}] exporting the CUBIT mesh into a SPECFEM3D Cartesian file format
+and 
+\item [{3.}] partitioning the SPECFEM3D Cartesian mesh files for a chosen
+number of cores. 
 \end{description}
-
-Examples are provided in the SPECFEM3D Cartesian package in the subdirectory \texttt{examples/}.
-We strongly encourage you to contribute your own example to this package by contacting the CIG Computational Seismology Mailing List  \urlwithparentheses{cig-seismo@geodynamics.org}.
+Examples are provided in the SPECFEM3D Cartesian package in the subdirectory
+\texttt{examples/}. We strongly encourage you to contribute your own
+example to this package by contacting the CIG Computational Seismology
+Mailing List \urlwithparentheses{cig-seismo@geodynamics.org}.
 
@@
 
-For the installation and handling of the CUBIT meshing tool suite, please refer to the CUBIT user manual and documentation.
-In order to give you a basic understanding of how to use CUBIT for our purposes, examples are provided in the SPECFEM3D Cartesian package in the subdirectory \texttt{examples/}:
+For the installation and handling of the CUBIT meshing tool suite,
+please refer to the CUBIT user manual and documentation. In order
+to give you a basic understanding of how to use CUBIT for our purposes,
+examples are provided in the SPECFEM3D Cartesian package in the subdirectory
+\texttt{examples/}: 
 \begin{description}
-\item [{\texttt{homogeneous\_halfspace}}] Creates a single block model and assigns elastic material parameters.
-\item [{\texttt{layered\_halfspace}}]  Combines two different, elastic material volumes and creates a refinement layer between the two.
-This example can be compared for validation against the solutions provided in subdirectory ~\\
+\item [{\texttt{homogeneous\_halfspace}}] Creates a single block model
+and assigns elastic material parameters. 
+\item [{\texttt{layered\_halfspace}}] Combines two different, elastic material
+volumes and creates a refinement layer between the two. This example
+can be compared for validation against the solutions provided in subdirectory
+~\\
  \texttt{VALIDATION\_3D\_SEM\_SIMPLER\_LAYER\_SOURCE\_DEPTH/}. 
-\item [{\texttt{waterlayered\_halfspace}}] Combines an acoustic and elastic material volume as in a schematic marine survey example.
-\item [{\texttt{tomographic\_model}}] Creates a single block model whose material properties will have to be read in from a tomographic model file during the databases creation by \texttt{xgenerate\_databases}.
+\item [{\texttt{waterlayered\_halfspace}}] Combines an acoustic and elastic
+material volume as in a schematic marine survey example. 
+\item [{\texttt{tomographic\_model}}] Creates a single block model whose
+material properties will have to be read in from a tomographic model
+file during the databases creation by \texttt{xgenerate\_databases}. 
 \end{description}
-
 \begin{figure}[htbp]
 \noindent \begin{centering}
-\includegraphics[width=.45\textwidth]{figures/example-homogeneous.jpg}
-\includegraphics[width=.45\textwidth]{figures/example-2layers.jpg} \\
-\includegraphics[width=.45\textwidth]{figures/example-water.jpg}
-\includegraphics[width=.45\textwidth]{figures/example-tomo.jpg}
+\includegraphics[width=0.45\textwidth]{figures/example-homogeneous}
+\includegraphics[width=0.45\textwidth]{figures/example-2layers} \\
+ \includegraphics[width=0.45\textwidth]{figures/example-water} \includegraphics[width=0.45\textwidth]{figures/example-tomo} 
 \par\end{centering}
 
-\caption{Screenshots of the CUBIT examples provided in subdirectory  \texttt{examples/}: homogeneous halfspace (top-left), layered halfspace (top-right), water layered halfspace (bottom-left) and tomographic model (bottom-right).}
+\caption{Screenshots of the CUBIT examples provided in subdirectory \texttt{examples/}:
+homogeneous halfspace (top-left), layered halfspace (top-right), water
+layered halfspace (bottom-left) and tomographic model (bottom-right).}
 
@@
 
-In each example subdirectory you will find a \texttt{README} file, which explains in a step-by-step tutorial the workflow for the example.
-Please feel free to contribute your own example to this package by contacting the CIG Computational Seismology Mailing List  \urlwithparentheses{cig-seismo@geodynamics.org}.
+In each example subdirectory you will find a \texttt{README} file,
+which explains in a step-by-step tutorial the workflow for the example.
+Please feel free to contribute your own example to this package by
+contacting the CIG Computational Seismology Mailing List \urlwithparentheses{cig-seismo@geodynamics.org}.
 
 \subsection{Exporting the Mesh with \texttt{cubit2specfem3d.py} }
+
 \label{subsec:Exporting-the-Mesh}
 
-Once the geometric model volumes in CUBIT are meshed, you prepare the model for exportation with the definition of material blocks and boundary surfaces.
-Thus, prior to exporting the mesh, you need to define blocks specifying the materials and absorbing boundaries in CUBIT.
-This process could be done automatically using the script \texttt{boundary\_definition.py} if the mesh meets some conditions or manually, following the block convention:
-
+Once the geometric model volumes in CUBIT are meshed, you prepare
+the model for exportation with the definition of material blocks and
+boundary surfaces. Thus, prior to exporting the mesh, you need to
+define blocks specifying the materials and absorbing boundaries in
+CUBIT. This process could be done automatically using the script \texttt{boundary\_definition.py}
+if the mesh meets some conditions or manually, following the block
+convention:
 \begin{description}
-\item [material\_name] Each material should have a specific block defined by a unique name.
-The name convention of the material is to start with either 'elastic' or 'acoustic'.
-It must be then followed by a unique identifier, e.g.
-'elastic 1', 'elastic 2', etc.
-The additional attributes to the block define the material description.
+\item [{material\_name}] Each material should have a specific block defined
+by a unique name. The name convention of the material is to start
+with either 'elastic' or 'acoustic'. It must be then followed by a
+unique identifier, e.g. 'elastic 1', 'elastic 2', etc. The additional
+attributes to the block define the material description.
 
@@
 \begin{description}
-\item [material\_id] An integer value which is unique for this material.
-\item [Vp] P-wave speed of the material (given in m/s).
-\item [Vs] S-wave speed of the material (given in m/s).
-\item [rho] density of the material (given in kg/m$^3$).
-\item [Q] quality factor to use in case of a simulation with attenuation turned on.
-It should be between 1 and 9000.
-In case no attenuation information is available, it can be set to zero.
-Please note that your Vp- and Vs-speeds are given for a reference frequency.
-To change this reference frequency, you change the value of
-\texttt{ATTENUATION\_f0\_REFERENCE} in the main constants file \texttt{constants.h} found in subdirectory \texttt{src/shared/}.
-\item [anisotropic\_flag] Flag describing the anisotropic model to use in case an anisotropic simulation should be conducted.
-See the file \texttt{model\_aniso.f90} in subdirectory \texttt{src/generate\_databases/} for an implementation of the anisotropic models.
-In case no anisotropy is available, it can be set to zero.
+\item [{material\_id}] An integer value which is unique for this material. 
+\item [{Vp}] P-wave speed of the material (given in m/s). 
+\item [{Vs}] S-wave speed of the material (given in m/s). 
+\item [{rho}] density of the material (given in kg/m$^{3}$). 
+\item [{Q}] quality factor to use in case of a simulation with attenuation
+turned on. It should be between 1 and 9000. In case no attenuation
+information is available, it can be set to zero. Please note that
+your Vp- and Vs-speeds are given for a reference frequency. To change
+this reference frequency, you change the value of \texttt{ATTENUATION\_f0\_REFERENCE}
+in the main constants file \texttt{constants.h} found in subdirectory
+\texttt{src/shared/}. 
+\item [{anisotropic\_flag}] Flag describing the anisotropic model to use
+in case an anisotropic simulation should be conducted. See the file
+\texttt{model\_aniso.f90} in subdirectory \texttt{src/generate\_databases/}
+for an implementation of the anisotropic models. In case no anisotropy
+is available, it can be set to zero. 
 \end{description}
-Note that this material block has to be defined using all the volumes which belong to this elastic material.
-For volumes belonging to another, different material, you will need to define a new material block.
+
+Note that this material block has to be defined using all the volumes
+which belong to this elastic material. For volumes belonging to another,
+different material, you will need to define a new material block.
 
@@
 \begin{description}
-\item [material\_id] An integer value which is unique for this material.
-\item [Vp] P-wave speed of the material (given in m/s).
-\item [0] S-wave speed of the material is ignored.
-\item [rho] density of the material (given in kg/m$^3$).
+\item [{material\_id}] An integer value which is unique for this material. 
+\item [{Vp}] P-wave speed of the material (given in m/s). 
+\item [{0}] S-wave speed of the material is ignored. 
+\item [{rho}] density of the material (given in kg/m$^{3}$). 
 \end{description}
-
-\item [face\_topo] Block definition for the surface which defines the free surface (which can have topography).
-The name of this block must be 'face\_topo', the block has to be defined using all the surfaces which constitute the complete free surface of the model.
-
-\item [face\_abs\_xmin] Block definition for the faces on the absorbing boundaries, one block for each surface with x=Xmin.
-\item [face\_abs\_xmax] Block definition for the faces on the absorbing boundaries, one block for each surface with x=Xmax.
-\item [face\_abs\_ymin] Block definition for the faces on the absorbing boundaries, one block for each surface with y=Ymin.
-\item [face\_abs\_ymax] Block definition for the faces on the absorbing boundaries, one block for each surface with y=Ymax.
-\item [face\_abs\_bottom] Block definition for the faces on the absorbing boundaries, one block for each surface with z=bottom.
+\item [{face\_topo}] Block definition for the surface which defines the
+free surface (which can have topography). The name of this block must
+be 'face\_topo', the block has to be defined using all the surfaces
+which constitute the complete free surface of the model.
+\item [{face\_abs\_xmin}] Block definition for the faces on the absorbing
+boundaries, one block for each surface with x=Xmin. 
+\item [{face\_abs\_xmax}] Block definition for the faces on the absorbing
+boundaries, one block for each surface with x=Xmax. 
+\item [{face\_abs\_ymin}] Block definition for the faces on the absorbing
+boundaries, one block for each surface with y=Ymin. 
+\item [{face\_abs\_ymax}] Block definition for the faces on the absorbing
+boundaries, one block for each surface with y=Ymax. 
+\item [{face\_abs\_bottom}] Block definition for the faces on the absorbing
+boundaries, one block for each surface with z=bottom. 
 \end{description}
-
 \begin{figure}[htbp]
 \noindent \begin{centering}
-\includegraphics[width=.45\textwidth]{figures/mount-surface.jpg}
-\includegraphics[width=.45\textwidth]{figures/mount-abs.jpg}
+\includegraphics[width=0.45\textwidth]{figures/mount-surface} \includegraphics[width=0.45\textwidth]{figures/mount-abs} 
 \par\end{centering}
 
-\caption{Example of the block definitions for the free surface 'face\_topo' (left) and the absorbing boundaries, defined in a single block 'face\_abs\_xmin' (right) in CUBIT.}
+\caption{Example of the block definitions for the free surface 'face\_topo'
+(left) and the absorbing boundaries, defined in a single block 'face\_abs\_xmin'
+(right) in CUBIT.}
 
@@
 
-Optionally, instead of specifying for each surface at the boundaries a single block like mentioned above, you can also specify a single block for all boundary surfaces and name it as one of the absorbing blocks above, e.g.
-'face\_abs\_xmin'.
-
-After the block definitions are done, you export the mesh using the script \texttt{cubit2specfem3d.py} provided in each of the example directories.
-If the export was successful, you should find the following files in a subdirectory \texttt{MESH/}:
-
+Optionally, instead of specifying for each surface at the boundaries
+a single block like mentioned above, you can also specify a single
+block for all boundary surfaces and name it as one of the absorbing
+blocks above, e.g. 'face\_abs\_xmin'.
+
+After the block definitions are done, you export the mesh using the
+script \texttt{cubit2specfem3d.py} provided in each of the example
+directories. If the export was successful, you should find the following
+files in a subdirectory \texttt{MESH/}:
 \begin{description}
-\item [absorbing\_cpml\_file] {\bf (only needed in case of C-PML absorbing conditions)}
-Contains on the first line the total number of C-PML spectral elements in the mesh,
-and then on the following line the list of all these C-PML elements with two numbers per line:
-first the spectral element number, and then a C-PML flag indicating to which C-PML layer(s)
-that element belongs, according to the following convention:
+\item [{absorbing\_cpml\_file}] \textbf{(only needed in case of C-PML absorbing
+conditions)} Contains on the first line the total number of C-PML
+spectral elements in the mesh, and then on the following line the
+list of all these C-PML elements with two numbers per line: first
+the spectral element number, and then a C-PML flag indicating to which
+C-PML layer(s) that element belongs, according to the following convention:
 
 \begin{itemize}
-\item Flag = 1 : element belongs to a X CPML layer only (either in Xmin or in Xmax),
-\item Flag = 2 : element belongs to a Y CPML layer only (either in Ymin or in Ymax),
-\item Flag = 3 : element belongs to a Z CPML layer only (either in Zmin or in Zmax),
-\item Flag = 4 : element belongs to a X CPML layer and also to a Y CPML layer,
-\item Flag = 5 : element belongs to a X CPML layer and also to a Z CPML layer,
-\item Flag = 6 : element belongs to a Y CPML layer and also to a Z CPML layer,
-\item Flag = 7 : element belongs to a X, to a Y and to a Z CPML layer, i.e., it belongs to a CPML corner.
+\item Flag = 1 : element belongs to a X CPML layer only (either in Xmin
+or in Xmax), 
+\item Flag = 2 : element belongs to a Y CPML layer only (either in Ymin
+or in Ymax), 
+\item Flag = 3 : element belongs to a Z CPML layer only (either in Zmin
+or in Zmax), 
+\item Flag = 4 : element belongs to a X CPML layer and also to a Y CPML
+layer, 
+\item Flag = 5 : element belongs to a X CPML layer and also to a Z CPML
+layer, 
+\item Flag = 6 : element belongs to a Y CPML layer and also to a Z CPML
+layer, 
+\item Flag = 7 : element belongs to a X, to a Y and to a Z CPML layer, i.e.,
+it belongs to a CPML corner. 
 \end{itemize}
 
-Note that it does not matter whether an element belongs to a Xmin or to a Xmax CPML, the flag
-is the same in both cases; the same is true for Ymin and Ymax, and also Zmin and Zmax.
-
-When you have an existing CUBIT (or similar) mesh stored in SPECFEM3D format,
-i.e., if you have existing "nodes\_coords\_file" and "mesh\_file" files but do not know how to assign
-CPML flags to them, we have created a small serial Fortran program that will do that automatically
-for you, i.e., which will create the "absorbing\_cpml\_file" for you.
-That program is
-utils/CPML/convert\_external\_layers\_of\_a\_given\_mesh\_to\_CPML\_layers.f90, and a small Makefile
-is provided in that directory (utils/CPML).
-
-IMPORTANT: it is your responsibility to make sure that in the input CUBIT (or similar) mesh
-that this code will read in SPECFEM3D format from files "nodes\_coords\_file" and "mesh\_file"
-you have created layers of elements that constitute a layer of constant thickness aligned
-with the coordinate grid axes (X, Y and/or Z), so that this code can assign CPML flags to them.
-This code does NOT check that (because it cannot, in any easy way).
-The mesh inside these CPML layers does not need to be structured nor regular, any non-structured
-mesh is fine as long as it has flat PML inner and outer faces, parallel to the axes, and thus
-of a constant thickness.
-The thickness can be different for the X, Y and Z sides. But for X it must not vary,
-for Y it must not vary, and for Z it must not vary.
-If you do not know the exact thickness,
-you can use a slightly LARGER value
-in this code (say 2\% to 5\% more) and this code will fix that and will adjust it; never use
-a SMALLER value otherwise this code will miss some CPML elements.
-
-Note: in a future release we will remove the constraint of having CPML layers aligned with the
-coordinate axes; we will allow for meshes that are titled by any constant angle in the horizontal plane.
-However this is not implemented yet.
-
-Note: in the case of fluid-solid layers in contact, the fluid-solid interface currently needs to be flat and horizontal
-inside the CPML layer (i.e., bathymetry should become flat and horizontal when entering the CPML); this (small) constraint
-will probably remain in the code for a while because it makes fluid-solid matching inside the CPML much easier.
-
-\item [materials\_file] Contains the material associations for each element. The format is:
-\begin{lyxcode}
-\texttt{element\_ID} \texttt{material\_ID}
-\end{lyxcode}
-where \texttt{element\_ID} is the element identifier and \texttt{material\_ID} is a unique identifier,
-positive (for materials taken from this list of materials, i.e. for which each spectral
-element has constant material properties taken from this list) or negative (for tomographic models,
-i.e. for spectral element whose real velocities and density will be assigned later by calling an external function
-to define model variations, for instance in the case of tomographic models; in such a case, material properties
-can vary inside each spectral element, i.e. be different at each of its Gauss-Lobatto-Legendre grid points).
-
-\item [nummaterial\_velocity\_file] Defines the material properties.
+Note that it does not matter whether an element belongs to a Xmin
+or to a Xmax CPML, the flag is the same in both cases; the same is
+true for Ymin and Ymax, and also Zmin and Zmax.
+
+When you have an existing CUBIT (or similar) mesh stored in SPECFEM3D
+format, i.e., if you have existing \textquotedbl{}nodes\_coords\_file\textquotedbl{}
+and \textquotedbl{}mesh\_file\textquotedbl{} files but do not know
+how to assign CPML flags to them, we have created a small serial Fortran
+program that will do that automatically for you, i.e., which will
+create the \textquotedbl{}absorbing\_cpml\_file\textquotedbl{} for
+you. That program is utils/CPML/convert\_external\_layers\_of\_a\_given\_mesh\_to\_CPML\_layers.f90,
+and a small Makefile is provided in that directory (utils/CPML).
+
+IMPORTANT: it is your responsibility to make sure that in the input
+CUBIT (or similar) mesh that this code will read in SPECFEM3D format
+from files \textquotedbl{}nodes\_coords\_file\textquotedbl{} and \textquotedbl{}mesh\_file\textquotedbl{}
+you have created layers of elements that constitute a layer of constant
+thickness aligned with the coordinate grid axes (X, Y and/or Z), so
+that this code can assign CPML flags to them. This code does NOT check
+that (because it cannot, in any easy way). The mesh inside these CPML
+layers does not need to be structured nor regular, any non-structured
+mesh is fine as long as it has flat PML inner and outer faces, parallel
+to the axes, and thus of a constant thickness. The thickness can be
+different for the X, Y and Z sides. But for X it must not vary, for
+Y it must not vary, and for Z it must not vary. If you do not know
+the exact thickness, you can use a slightly LARGER value in this code
+(say 2\% to 5\% more) and this code will fix that and will adjust
+it; never use a SMALLER value otherwise this code will miss some CPML
+elements.
+
+Note: in a future release we will remove the constraint of having
+CPML layers aligned with the coordinate axes; we will allow for meshes
+that are titled by any constant angle in the horizontal plane. However
+this is not implemented yet.
+
+Note: in the case of fluid-solid layers in contact, the fluid-solid
+interface currently needs to be flat and horizontal inside the CPML
+layer (i.e., bathymetry should become flat and horizontal when entering
+the CPML); this (small) constraint will probably remain in the code
+for a while because it makes fluid-solid matching inside the CPML
+much easier.
+
+\item [{materials\_file}] Contains the material associations for each element.
+The format is:
+
+\begin{lyxcode}
+element\_ID~material\_ID
+\end{lyxcode}
+
+where \texttt{element\_ID} is the element identifier and \texttt{material\_ID}
+is a unique identifier, positive (for materials taken from this list
+of materials, i.e. for which each spectral element has constant material
+properties taken from this list) or negative (for tomographic models,
+i.e. for spectral element whose real velocities and density will be
+assigned later by calling an external function to define model variations,
+for instance in the case of tomographic models; in such a case, material
+properties can vary inside each spectral element, i.e. be different
+at each of its Gauss-Lobatto-Legendre grid points).
+
+\item [{nummaterial\_velocity\_file}] Defines the material properties.
+
 \begin{itemize}
-\item For classical materials (i.e., spectral elements for which the velocity and density
-model will not be assigned by calling an external function to define for instance a tomographic model), the format is:
-\begin{lyxcode}
-\texttt{domain\_ID} \texttt{material\_ID}  \texttt{rho}  \texttt{vp}  \texttt{vs}  \texttt{Qkappa}  \texttt{Qmu} \texttt{anisotropy\_flag}
-\end{lyxcode}
-where \textbf{\texttt{domain\_ID} is 1 for acoustic and 2 for elastic or viscoelastic materials,}
-\texttt{material\_ID} a unique identifier,  \texttt{rho} the density in $kg \, m^{-3}$,  \texttt{vp} the P-wave speed in $m \, s^{-1}$,
-\texttt{vs}  the S-wave speed in $m \, s^{-1}$, \texttt{Q} the quality factor and \texttt{anisotropy\_flag} an identifier for anisotropic models.
-Note that the \texttt{Qkappa} value is ignored by the code unless \texttt{FULL\_ATTENUATION\_SOLID} is set.
-Note also that both \texttt{Qkappa} and \texttt{Qmu} are ignored by the code unless \texttt{ATTENUATION} is set.
-If you want a model with no \texttt{Qmu} attenuation, both set \texttt{ATTENUATION} to \texttt{.false.} in the \texttt{Par\_file} and set \texttt{Qmu} to 9999 here.
-If you want a model with no \texttt{Qkappa} attenuation, both set \texttt{FULL\_ATTENUATION\_SOLID} to \texttt{.false.} in the \texttt{Par\_file} and set \texttt{Qkappa} to 9999 here.
-%%magnoni
-\item For tomographic velocity models, please read Chapter \ref{cha:-Changing-the} and Section \ref{sec:Using-tomographic} `Using external tomographic Earth models' for further details.
-%
+\item For classical materials (i.e., spectral elements for which the velocity
+and density model will not be assigned by calling an external function
+to define for instance a tomographic model), the format is:
+
+\begin{lyxcode}
+domain\_ID~material\_ID~rho~vp~vs~Qkappa~Qmu~anisotropy\_flag
+\end{lyxcode}
+
+where \texttt{\textbf{domain\_ID}}\textbf{ is 1 for acoustic and 2
+for elastic or viscoelastic materials,} \texttt{material\_ID} a unique
+identifier, \texttt{rho} the density in $kg\, m^{-3}$, \texttt{vp}
+the P-wave speed in $m\, s^{-1}$, \texttt{vs} the S-wave speed in
+$m\, s^{-1}$, \texttt{Q} the quality factor and \texttt{anisotropy\_flag}
+an identifier for anisotropic models. Note that the \texttt{Qkappa}
+value is ignored by the code unless \texttt{FULL\_ATTENUATION\_SOLID}
+is set. Note also that both \texttt{Qkappa} and \texttt{Qmu} are ignored
+by the code unless \texttt{ATTENUATION} is set. If you want a model
+with no \texttt{Qmu} attenuation, both set \texttt{ATTENUATION} to
+\texttt{.false.} in the \texttt{Par\_file} and set \texttt{Qmu} to
+9999 here. If you want a model with no \texttt{Qkappa} attenuation,
+both set \texttt{FULL\_ATTENUATION\_SOLID} to \texttt{.false.} in
+the \texttt{Par\_file} and set \texttt{Qkappa} to 9999 here. %%magnoni
+
+\item For tomographic velocity models, please read Chapter \ref{cha:-Changing-the}
+and Section \ref{sec:Using-tomographic} `Using external tomographic
+Earth models' for further details. 
 \end{itemize}
-
-\item [nodes\_coords\_file] Contains the point locations in Cartesian coordinates of the mesh element corners.
-\item [mesh\_file] Contains the mesh element connectivity. The hexahedral elements can have 8 or 27 nodes.\\
-See picture doc/mesh\_numbering\_convention/numbering\_convention\_27\_nodes.jpg to see\\
-in which (standard) order the points must be cited. In the case of 8 nodes, just include the first 8 points.
-\item [free\_or\_absorbing\_surface\_file\_zmax] Contains the free surface connectivity or \\
-the surface connectivity of the absorbing boundary surface at the top (Zmax), \\
-depending on whether the top surface is defined as free or absorbing in DATA/Par\_file. \\
-You should put both the surface of acoustic regions and of elastic regions in that file;
-that is, list all the element faces that constitute the surface of the model in that file.
-\item [absorbing\_surface\_file\_xmax] Contains the surface connectivity of the absorbing boundary surface at Xmax \\(also needed in the case of C-PML absorbing conditions,
-in order for the code to be able to impose Dirichlet conditions on their outer edge).
-\item [absorbing\_surface\_file\_xmin] Contains the surface connectivity of the absorbing boundary surface at Xmin \\(also needed in the case of C-PML absorbing conditions,
-in order for the code to be able to impose Dirichlet conditions on their outer edge).
-\item [absorbing\_surface\_file\_ymax] Contains the surface connectivity of the absorbing boundary surface at Ymax \\(also needed in the case of C-PML absorbing conditions,
-in order for the code to be able to impose Dirichlet conditions on their outer edge).
-\item [absorbing\_surface\_file\_ymin] Contains the surface connectivity of the absorbing boundary surface at Ymin \\(also needed in the case of C-PML absorbing conditions,
-in order for the code to be able to impose Dirichlet conditions on their outer edge).
-\item [absorbing\_surface\_file\_bottom] Contains the surface connectivity of the absorbing boundary surface at the bottom (Zmin) \\(also needed in the case of C-PML absorbing conditions,
-in order for the code to be able to impose Dirichlet conditions on their outer edge).
+\item [{nodes\_coords\_file}] Contains the point locations in Cartesian
+coordinates of the mesh element corners. 
+\item [{mesh\_file}] Contains the mesh element connectivity. The hexahedral
+elements can have 8 or 27 nodes.\\
+ See picture doc/mesh\_numbering\_convention/numbering\_convention\_27\_nodes.jpg
+to see\\
+ in which (standard) order the points must be cited. In the case of
+8 nodes, just include the first 8 points. 
+\item [{free\_or\_absorbing\_surface\_file\_zmax}] Contains the free surface
+connectivity or \\
+ the surface connectivity of the absorbing boundary surface at the
+top (Zmax), \\
+ depending on whether the top surface is defined as free or absorbing
+in DATA/Par\_file. \\
+ You should put both the surface of acoustic regions and of elastic
+regions in that file; that is, list all the element faces that constitute
+the surface of the model in that file. 
+\item [{absorbing\_surface\_file\_xmax}] Contains the surface connectivity
+of the absorbing boundary surface at Xmax \\
+(also needed in the case of C-PML absorbing conditions, in order for
+the code to be able to impose Dirichlet conditions on their outer
+edge). 
+\item [{absorbing\_surface\_file\_xmin}] Contains the surface connectivity
+of the absorbing boundary surface at Xmin \\
+(also needed in the case of C-PML absorbing conditions, in order for
+the code to be able to impose Dirichlet conditions on their outer
+edge). 
+\item [{absorbing\_surface\_file\_ymax}] Contains the surface connectivity
+of the absorbing boundary surface at Ymax \\
+(also needed in the case of C-PML absorbing conditions, in order for
+the code to be able to impose Dirichlet conditions on their outer
+edge). 
+\item [{absorbing\_surface\_file\_ymin}] Contains the surface connectivity
+of the absorbing boundary surface at Ymin \\
+(also needed in the case of C-PML absorbing conditions, in order for
+the code to be able to impose Dirichlet conditions on their outer
+edge). 
+\item [{absorbing\_surface\_file\_bottom}] Contains the surface connectivity
+of the absorbing boundary surface at the bottom (Zmin) \\
+(also needed in the case of C-PML absorbing conditions, in order for
+the code to be able to impose Dirichlet conditions on their outer
+edge). 
 \end{description}
-
 These mesh files are needed as input files for the partitioner \texttt{xdecompose\_mesh}
-to load-balance the mesh. Please see the next section for further details.
+to load-balance the mesh. Please see the next section for further
+details.
 
 \subsubsection*{Checking the mesh quality}
-The quality of the mesh may be inspected more precisely based upon the serial code
-in the file  \texttt{check\_mesh\_quality\_}~\\
+
+The quality of the mesh may be inspected more precisely based upon
+the serial code in the file \texttt{check\_mesh\_quality\_}~\\
  \texttt{CUBIT\_Abaqus.f90} located in the directory \texttt{src/check\_mesh\_quality\_CUBIT\_Abaqus/}.
@@
 
-Prior to running and compiling this code, you have to export your mesh in CUBIT to an ABAQUS (.inp) format.
-For example, export mesh block IDs belonging to volumes in order to check the quality of the hexahedral elements.
-You also have to determine a number of parameters of your mesh, such as the number of nodes and number of elements
-and modify the header of the \texttt{check\_mesh\_quality\_CUBIT\_Abaqus.f90} source file. Then,
-in the main directory, type
-\begin{lyxcode}
-{\small make xcheck\_mesh\_quality\_CUBIT\_Abaqus}{\small \par}
+Prior to running and compiling this code, you have to export your
+mesh in CUBIT to an ABAQUS (.inp) format. For example, export mesh
+block IDs belonging to volumes in order to check the quality of the
+hexahedral elements. You also have to determine a number of parameters
+of your mesh, such as the number of nodes and number of elements and
+modify the header of the \texttt{check\_mesh\_quality\_CUBIT\_Abaqus.f90}
+source file. Then, in the main directory, type 
+\begin{lyxcode}
+{\small make~xcheck\_mesh\_quality\_CUBIT\_Abaqus}{\small \par}
 \end{lyxcode}
@@
 \end{lyxcode}
-to generate an
-%not supported: AVS output file (\texttt{\small AVS\_meshquality.inp} in AVS UCD format) or
-OpenDX output file (\texttt{\small DX\_mesh\_quality.dx})
+to generate an %not supported: AVS output file (\texttt{\small AVS\_meshquality.inp} in AVS UCD format) or
+OpenDX output file (\texttt{\small DX\_mesh\_quality.dx}{\small )
 that can be used to investigate mesh quality, e.g. skewness of elements,
-and a Gnuplot histogram (\texttt{\small mesh\_quality\_histogram.txt})
-that can be plotted with gnuplot (type
-`\texttt{\small gnuplot plot\_mesh\_quality\_histogram.gnu}'). The
-histogram is also printed to the screen.
-Analyze that skewness histogram of mesh elements to make sure no element has a skewness above approximately 0.75, otherwise the mesh is of poor quality (and if even a single element has a skewness value above 0.80, then you must definitely improve the mesh).
-If you want to start designing
+and a Gnuplot histogram (}\texttt{\small mesh\_quality\_histogram.txt}{\small )
+that can be plotted with gnuplot (type `}\texttt{\small gnuplot plot\_mesh\_quality\_histogram.gnu}{\small ').
+The histogram is also printed to the screen. Analyze that skewness
+histogram of mesh elements to make sure no element has a skewness
+above approximately 0.75, otherwise the mesh is of poor quality (and
+if even a single element has a skewness value above 0.80, then you
+must definitely improve the mesh). If you want to start designing
 your own meshes, this tool is useful for viewing your creations. You
 are striving for meshes with elements with `cube-like' dimensions,
-e.g., the mesh should contain no very elongated or skewed elements.
+e.g., the mesh should contain no very elongated or skewed elements.}{\small \par}
 
@@
 
-The SPECFEM3D Cartesian software package performs large scale simulations in a parallel 'Single Process Multiple Data' way.
-The spectral-element mesh created with CUBIT needs to be distributed on the processors. This partitioning is executed once and for all prior to
-the execution of the solver so it is referred to as a static mapping.
-
-An efficient partitioning is important because it leverages the overall running time of the application.
-It amounts to balance the number of elements in each slice while minimizing the communication costs
-resulting from the placement of adjacent elements on different processors. \texttt{decompose\_mesh} depends on the SCOTCH library \citep{PeRo96},
-which provides efficent static mapping, graph and mesh partitioning routines. SCOTCH is a free software package developed by
-Fran\c {c}ois Pellegrini et al. from LaBRI and INRIA in Bordeaux, France, downloadable from the web page \url{https://gforge.inria.fr/projects/scotch/}.
+The SPECFEM3D Cartesian software package performs large scale simulations
+in a parallel 'Single Process Multiple Data' way. The spectral-element
+mesh created with CUBIT needs to be distributed on the processors.
+This partitioning is executed once and for all prior to the execution
+of the solver so it is referred to as a static mapping.
+
+An efficient partitioning is important because it leverages the overall
+running time of the application. It amounts to balance the number
+of elements in each slice while minimizing the communication costs
+resulting from the placement of adjacent elements on different processors.
+\texttt{decompose\_mesh} depends on the SCOTCH library \citep{PeRo96},
+which provides efficent static mapping, graph and mesh partitioning
+routines. SCOTCH is a free software package developed by François
+Pellegrini et al. from LaBRI and INRIA in Bordeaux, France, downloadable
+from the web page \url{https://gforge.inria.fr/projects/scotch/}.
 \\
 
-In most cases, the configuration with \texttt{./configure~FC=ifort} should be sufficient.
-During the configuration process, the script tries to find existing SCOTCH installations.
-In case your system has no pre-existing SCOTCH installation,
-we provide the source code of SCOTCH, which is released open source under the French CeCILL-C version 1 license,
-in directory
-\texttt{src/decompose\_mesh/scotch\_5.1.12b}.
-This version gets bundled with the compilation of the SPECFEM3D Cartesian package if no libraries could have been found.
-If this automatic compilation of the SCOTCH libraries fails,
-please refer to file INSTALL.txt in that directory to see further details how to compile it on your system.
-In case you want to use a pre-existing installation, make sure you have correctly specified the path
-of the SCOTCH library when using the option \texttt{-{}-with-scotch-dir} with the \texttt{./configure} script.
-In the future you should be able to find more recent versions at  \url{http://www.labri.fr/perso/pelegrin/scotch/scotch\_en.html} .
-\\
+In most cases, the configuration with \texttt{./configure~FC=ifort}
+should be sufficient. During the configuration process, the script
+tries to find existing SCOTCH installations. In case your system has
+no pre-existing SCOTCH installation, we provide the source code of
+SCOTCH, which is released open source under the French CeCILL-C version
+1 license, in directory \texttt{src/decompose\_mesh/scotch\_5.1.12b}.
+This version gets bundled with the compilation of the SPECFEM3D Cartesian
+package if no libraries could have been found. If this automatic compilation
+of the SCOTCH libraries fails, please refer to file INSTALL.txt in
+that directory to see further details how to compile it on your system.
+In case you want to use a pre-existing installation, make sure you
+have correctly specified the path of the SCOTCH library when using
+the option \texttt{-{}-with-scotch-dir} with the \texttt{./configure}
+script. In the future you should be able to find more recent versions
+at \url{http://www.labri.fr/perso/pelegrin/scotch/scotch_en.html}
+. \\
 
@@
 \noindent \begin{centering}
-\includegraphics[width=.45\textwidth]{figures/mount-partitions.jpg}
-\includegraphics[width=.45\textwidth]{figures/mount-partitions2.jpg}
+\includegraphics[width=0.45\textwidth]{figures/mount-partitions}
+\includegraphics[width=0.45\textwidth]{figures/mount-partitions2} 
 \par\end{centering}
 
-\caption{Example of a mesh partitioning onto four cores.
-Each single core partition is colored differently.
-The executable \texttt{xdecompose\_mesh} can equally distribute the mesh on any arbitrary number of cores.
-Domain decomposition is explained in detail in \cite{MaKoBlLe08}, and excellent scaling up to 150,000 processor cores in shown for instance in
-\cite{CaKoLaTiMiLeSnTr08,KoLaMi08a,MaKoBlLe08,KoErGoMi10,Kom11}.}
+\caption{Example of a mesh partitioning onto four cores. Each single core partition
+is colored differently. The executable \texttt{xdecompose\_mesh} can
+equally distribute the mesh on any arbitrary number of cores. Domain
+decomposition is explained in detail in \citet{MaKoBlLe08}, and excellent
+scaling up to 150,000 processor cores in shown for instance in \citet{CaKoLaTiMiLeSnTr08,KoLaMi08a,MaKoBlLe08,KoErGoMi10,Kom11}.}
 
@@
 
-When you are ready to compile,
-in the main directory type
-\begin{lyxcode}
-{\small make xdecompose\_mesh}{\small \par}
-\end{lyxcode}
-If all paths and flags have been set correctly, the executable \texttt{bin/xdecompose\_mesh} should be produced.
-
-The partitioning is done in serial for now (in the next release we will provide a parallel version of that code),
-the synopsis is:
-\begin{lyxcode}
-./xdecompose\_mesh  nparts  input\_directory output\_directory
+When you are ready to compile, in the main directory type 
+\begin{lyxcode}
+{\small make~xdecompose\_mesh}{\small \par}
+\end{lyxcode}
+If all paths and flags have been set correctly, the executable \texttt{bin/xdecompose\_mesh}
+should be produced.
+
+The partitioning is done in serial for now (in the next release we
+will provide a parallel version of that code), the synopsis is: 
+\begin{lyxcode}
+./xdecompose\_mesh~nparts~input\_directory~output\_directory
 \end{lyxcode}
@@
 \begin{itemize}
-     \item \texttt{nparts} is the number of partitions, i.e., the number of cores for the parallel simulations,
-     \item \texttt{input\_directory} is the directory which holds all the files generated by the Python script \texttt{cubit2specfem3d.py} explained in the previous Section~\ref{subsec:Exporting-the-Mesh}, e.g.
-\texttt{MESH/}, and
-     \item \texttt{output\_directory} is the directory for the output of this partitioner which stores ACII-format files named like \texttt{proc******\_Database} for each partition.
-These files will be needed for creating the distributed databases, and have to reside in the directory  \texttt{LOCAL\_PATH} specified in the main \texttt{Par\_file}, e.g.
-in directory \texttt{OUTPUT\_FILES/DATABASES\_MPI}.
-Please see Chapter~\ref{cha:Creating-Distributed-Databases} for further details.
+\item \texttt{nparts} is the number of partitions, i.e., the number of cores
+for the parallel simulations, 
+\item \texttt{input\_directory} is the directory which holds all the files
+generated by the Python script \texttt{cubit2specfem3d.py} explained
+in the previous Section~\ref{subsec:Exporting-the-Mesh}, e.g. \texttt{MESH/},
+and 
+\item \texttt{output\_directory} is the directory for the output of this
+partitioner which stores ACII-format files named like \texttt{proc{*}{*}{*}{*}{*}{*}\_Database}
+for each partition. These files will be needed for creating the distributed
+databases, and have to reside in the directory \texttt{LOCAL\_PATH}
+specified in the main \texttt{Par\_file}, e.g. in directory \texttt{OUTPUT\_FILES/DATABASES\_MPI}.
+Please see Chapter~\ref{cha:Creating-Distributed-Databases} for
+further details. 
 \end{itemize}
-Note that all the files generated by the Python script  \texttt{cubit2specfem3d.py} must be placed in the \texttt{input\_directory} folder before running the program.
+Note that all the files generated by the Python script \texttt{cubit2specfem3d.py}
+must be placed in the \texttt{input\_directory} folder before running
+the program.
 
@@
 
-In case you successfully ran the configuration script, you are also ready to compile the internal mesher.
-This is an alternative to CUBIT
-for the mesh generation of relatively simple geological models. The mesher is no longer
-dedicated to Southern California and more flexiblity is provided in this version of the package.
+In case you successfully ran the configuration script, you are also
+ready to compile the internal mesher. This is an alternative to CUBIT
+for the mesh generation of relatively simple geological models. The
+mesher is no longer dedicated to Southern California and more flexiblity
+is provided in this version of the package.
 
@@
 of elements. In this example we use $5^{2}=25$ processors. }
-
 \end{figure}
@@
 \begin{lyxcode}
-{\small make xmeshfem3D}{\small \par}
-\end{lyxcode}
-If all paths and flags
-have been set correctly, the mesher should now compile and produce
-the executable \texttt{bin/xmeshfem3D}. Please note that
-\texttt{xmeshfem3D} must be called directly from the \texttt{bin/} directory,
-as most of the binaries of the package.
+{\small make~xmeshfem3D}{\small \par}
+\end{lyxcode}
+If all paths and flags have been set correctly, the mesher should
+now compile and produce the executable \texttt{bin/xmeshfem3D}. Please
+note that \texttt{xmeshfem3D} must be called directly from the \texttt{bin/}
+directory, as most of the binaries of the package.
 
@@
 Before running the mesher, a number of parameters need to be set in
-the \texttt{Mesh\_Par\_file}. This requires a basic understanding of how
-the SEM is implemented, and we encourage you to read \citet{KoVi98,KoTr99}
+the \texttt{Mesh\_Par\_file}. This requires a basic understanding
+of how the SEM is implemented, and we encourage you to read \citet{KoVi98,KoTr99}
 and \citet{KoLiTrSuStSh04}.
@@
 benchmarks on rectangular models, turn the UTM projection off by using
-the flag \texttt{\small SUPPRESS\_UTM\_PROJECTION}, in which case
-all `longitude' parameters simply refer to the $x$~axis, and all
-`latitude' parameters simply refer to the $y$~axis. To run the mesher
-for a global simulation, the following parameters need to be set in
-the \texttt{Mesh\_Par\_file}:
-
+the flag \texttt{\small SUPPRESS\_UTM\_PROJECTION}{\small , in which
+case all `longitude' parameters simply refer to the $x$~axis, and
+all `latitude' parameters simply refer to the $y$~axis. To run the
+mesher for a global simulation, the following parameters need to be
+set in the }\texttt{\small Mesh\_Par\_file}{\small :}{\small \par}
 \begin{description}
@@
 \item [{\texttt{SUPPRESS\_UTM\_PROJECTION}}] set to be \texttt{.false.}
-when your model range is specified in geographical coordinates,
-and needs to be \texttt{.true.} when your model is specified in
-Cartesian coordinates. \noun{UTM projection zone in which your simulation
-region resides.}
-\item [{\texttt{INTERFACES\_FILE }}] File which contains the description of the topography and
-of the interfaces between the different layers of the model, if any.
-The number of spectral elements in the vertical direction within each layer is also defined in this file.
-
+when your model range is specified in geographical coordinates, and
+needs to be \texttt{.true.} when your model is specified in Cartesian
+coordinates. \noun{UTM projection zone in which your simulation region
+resides.} 
+\item [{\texttt{INTERFACES\_FILE }}] File which contains the description
+of the topography and of the interfaces between the different layers
+of the model, if any. The number of spectral elements in the vertical
+direction within each layer is also defined in this file.
 \item [{$\nexxi$}] The number of spectral elements along one side of the
@@
 \begin{equation}
-\mbox{shortest period (s)}=(288/\nexxi)\times2.\label{eq:shortest_period}\end{equation}
+\mbox{shortest period (s)}=(288/\nexxi)\times2.\label{eq:shortest_period}
+\end{equation}
  The number of grid points in each orthogonal direction of the reference
@@
 where $c\ge1$ is a positive integer. 
-\item [{\texttt{USE\_REGULAR\_MESH}}] set to be \texttt{.true.} if you want a perfectly regular
-mesh or \texttt{.false.} if you want to add doubling horizontal layers to coarsen the mesh. In this case, you
-also need to provide additional information by setting up the next three parameters.
-\item [{\texttt{NDOUBLINGS}}] The number of horizontal doubling layers. Must be set to \texttt{1} or \texttt{2}
-if \texttt{USE\_REGULAR\_MESH} is set to \texttt{.true.}.
-\item [{\texttt{NZ\_DOUBLING\_1}}] The position of the first doubling layer (only interpreted
-if \texttt{USE\_REGULAR\_MESH} is set to \texttt{.true.}).
-\item [{\texttt{NZ\_DOUBLING\_2}}] The position of the second doubling layer (only interpreted
-if \texttt{USE\_REGULAR\_MESH} is set to \texttt{.true.} and if \texttt{NDOUBLINGS} is set to \texttt{2}).
-\item [{\texttt{CREATE\_ABAQUS\_FILES}}] Set this flag to \texttt{.true.} to save Abaqus FEA \urlwithparentheses{www.simulia.com}
-mesh files for subsequent viewing. Turning the flag on generates files in the \texttt{LOCAL\_PATH} directory.
-See Section~\ref{sec:Mesh-graphics} for a discussion of mesh viewing features.
-\item [{\texttt{CREATE\_DX\_FILES}}] Set this flag to \texttt{.true.} to save OpenDX \urlwithparentheses{www.opendx.org}
-mesh files for subsequent viewing.
+\item [{\texttt{USE\_REGULAR\_MESH}}] set to be \texttt{.true.} if you
+want a perfectly regular mesh or \texttt{.false.} if you want to add
+doubling horizontal layers to coarsen the mesh. In this case, you
+also need to provide additional information by setting up the next
+three parameters. 
+\item [{\texttt{NDOUBLINGS}}] The number of horizontal doubling layers.
+Must be set to \texttt{1} or \texttt{2} if \texttt{USE\_REGULAR\_MESH}
+is set to \texttt{.true.}. 
+\item [{\texttt{NZ\_DOUBLING\_1}}] The position of the first doubling layer
+(only interpreted if \texttt{USE\_REGULAR\_MESH} is set to \texttt{.true.}). 
+\item [{\texttt{NZ\_DOUBLING\_2}}] The position of the second doubling
+layer (only interpreted if \texttt{USE\_REGULAR\_MESH} is set to \texttt{.true.}
+and if \texttt{NDOUBLINGS} is set to \texttt{2}). 
+\item [{\texttt{CREATE\_ABAQUS\_FILES}}] Set this flag to \texttt{.true.}
+to save Abaqus FEA \urlwithparentheses{www.simulia.com} mesh files
+for subsequent viewing. Turning the flag on generates files in the
+\texttt{LOCAL\_PATH} directory. See Section~\ref{sec:Mesh-graphics}
+for a discussion of mesh viewing features. 
+\item [{\texttt{CREATE\_DX\_FILES}}] Set this flag to \texttt{.true.} to
+save OpenDX \urlwithparentheses{www.opendx.org} mesh files for subsequent
+viewing. 
 \item [{\texttt{LOCAL\_PATH}}] Directory in which the partitions generated
@@
 mesher finishes, you can log in to one of the compute nodes and view
-the contents of the \texttt{LOCAL\_PATH} directory to see the
-files generated by the mesher.
-These files will be needed for creating the distributed databases, and have to reside in the directory  \texttt{LOCAL\_PATH} specified in the main \texttt{Par\_file}, e.g.
-in directory \texttt{OUTPUT\_FILES/DATABASES\_MPI}.
-Please see Chapter~\ref{cha:Creating-Distributed-Databases} for further details.
-\item [{\texttt{NMATERIALS}}] The number of different materials in your model. In the following lines, each material needs to be defined as :
-\begin{lyxcode}
-\texttt{material\_ID}  \texttt{rho}  \texttt{vp}  \texttt{vs}  \texttt{Q}  \texttt{anisotropy\_flag} \texttt{domain\_ID}
-\end{lyxcode}
+the contents of the \texttt{LOCAL\_PATH} directory to see the files
+generated by the mesher. These files will be needed for creating the
+distributed databases, and have to reside in the directory \texttt{LOCAL\_PATH}
+specified in the main \texttt{Par\_file}, e.g. in directory \texttt{OUTPUT\_FILES/DATABASES\_MPI}.
+Please see Chapter~\ref{cha:Creating-Distributed-Databases} for
+further details. 
+\item [{\texttt{NMATERIALS}}] The number of different materials in your
+model. In the following lines, each material needs to be defined as
+:
+
+\begin{lyxcode}
+material\_ID~rho~vp~vs~Q~anisotropy\_flag~domain\_ID
+\end{lyxcode}
+
 where 
@@
 \item \texttt{Q} : quality factor (0=no attenuation) 
-     \item \texttt{anisotropy\_flag}  : 0=no anisotropy / 1,2,.. check with implementation in \texttt{aniso\_model.f90}
+\item \texttt{anisotropy\_flag} : 0=no anisotropy / 1,2,.. check with implementation
+in \texttt{aniso\_model.f90} 
 \item \texttt{domain\_id} : 1=acoustic / 2=elastic 
 \end{itemize}
-\item [{\texttt{NREGIONS}}] The number of regions in the mesh.
-In the following lines, because the mesh is regular or 'almost regular', each region is defined as :
-\begin{lyxcode}
-  \texttt{NEX\_XI\_BEGIN}  \texttt{NEX\_XI\_END}  \texttt{NEX\_ETA\_BEGIN}  \texttt{NEX\_ETA\_END}  \texttt{NZ\_BEGIN}  \texttt{NZ\_END}  \texttt{material\_ID}
+\item [{\texttt{NREGIONS}}] The number of regions in the mesh. In the following
+lines, because the mesh is regular or 'almost regular', each region
+is defined as :
+
+\begin{lyxcode}
+NEX\_XI\_BEGIN~NEX\_XI\_END~NEX\_ETA\_BEGIN~NEX\_ETA\_END~NZ\_BEGIN~NZ\_END~material\_ID
 \end{lyxcode}
 \end{description}
-
-The \texttt{INTERFACES\_FILE} parameter of \texttt{Mesh\_Par\_File} defines the file which contains the settings of the topography grid and of the interfaces grids.
-Topography is defined as a set of elevation values on a regular 2D grid.
-It is also possible to
-define interfaces between the layers of the model in the same way.
-The file needs to define several parameters:
+The \texttt{INTERFACES\_FILE} parameter of \texttt{Mesh\_Par\_File}
+defines the file which contains the settings of the topography grid
+and of the interfaces grids. Topography is defined as a set of elevation
+values on a regular 2D grid. It is also possible to define interfaces
+between the layers of the model in the same way. The file needs to
+define several parameters: 
 \begin{itemize}
-\item The number of interfaces, including the topography. This needs to be set at the first line. Then, from the bottom
-to the top of the model, you need to define the grids with:
+\item The number of interfaces, including the topography. This needs to
+be set at the first line. Then, from the bottom to the top of the
+model, you need to define the grids with: 
 \item \texttt{SUPPRESS\_UTM\_PROJECTION} flag as described previously, 
@@
 \item minimal $x$ and $y$ coordinates (LONG\_MIN and LAT\_MIN), 
-\item spacing between points along $x$ and $y$ (SPACING\_XI and SPACING\_ETA) and
-\item the name of the file which contains the elevation values (in $y$.$x$ increasing order).
+\item spacing between points along $x$ and $y$ (SPACING\_XI and SPACING\_ETA)
+and 
+\item the name of the file which contains the elevation values (in $y$.$x$
+increasing order). 
 \end{itemize}
-At the end of this file, you simply need to set the number of spectral elements in the vertical direction for each layer. We provide a few models in the {\texttt{examples/}} directory.  \\
-
-Finally, depending on your system, you might need to provide a file that tells MPI what compute nodes
-to use for the simulations. The file must have a number of entries
-(one entry per line) at least equal to the number of processors needed
-for the run. A sample file is provided in the file \texttt{mymachines}.
-This file is not used by the mesher or solver, but is required by
-the \texttt{go\_mesher} and \texttt{go\_solver} default job submission
-scripts. See Chapter~\ref{cha:Scheduler} for information about running
-the code on a system with a scheduler, e.g., LSF.
+At the end of this file, you simply need to set the number of spectral
+elements in the vertical direction for each layer. We provide a few
+models in the {\texttt{examples/}} directory. \\
+
+Finally, depending on your system, you might need to provide a file
+that tells MPI what compute nodes to use for the simulations. The
+file must have a number of entries (one entry per line) at least equal
+to the number of processors needed for the run. A sample file is provided
+in the file \texttt{mymachines}. This file is not used by the mesher
+or solver, but is required by the \texttt{go\_mesher} and \texttt{go\_solver}
+default job submission scripts. See Chapter~\ref{cha:Scheduler}
+for information about running the code on a system with a scheduler,
+e.g., LSF.
 
@@
 in \texttt{output\_mesher.txt}; this file provides lots of details
-about the mesh that was generated. Please note that the mesher suggests a time step \texttt{DT} to run
-the solver with. The mesher output file also contains a table about the quality of the mesh to indicate
-possible problems with the distortions of elements.
-Alternatively, output can be directed
-to the screen instead by uncommenting a line in \texttt{constants.h}:
+about the mesh that was generated. Please note that the mesher suggests
+a time step \texttt{DT} to run the solver with. The mesher output
+file also contains a table about the quality of the mesh to indicate
+possible problems with the distortions of elements. Alternatively,
+output can be directed to the screen instead by uncommenting a line
+in \texttt{constants.h}: 
 \begin{lyxcode}
@@
 \end{lyxcode}
-
-To control the quality of the mesh, check the standard output (either on the screen or in the \texttt{OUTPUT\_FILES} directory in \texttt{output\_mesher.txt}) and analyze the skewness histogram of mesh elements to make sure no element has a skewness above approximately 0.75, otherwise the mesh is of poor quality (and if even a single element has a skewness value above 0.80, then you must definitely improve the mesh).
-To draw the skewness histogram on the screen, type \texttt{gnuplot plot\_mesh\_quality\_histogram.gnu}.
-
-\chapter{\label{cha:Creating-Distributed-Databases}Creating the Distributed Databases}
-
-After using \texttt{xmeshfem3D} or \texttt{xdecompose\_mesh}, the next step in the workflow is to compile
-\texttt{xgenerate\_}~\\
-\texttt{databases}. This program is going to create all the missing information needed by the SEM solver.
+To control the quality of the mesh, check the standard output (either
+on the screen or in the \texttt{OUTPUT\_FILES} directory in \texttt{output\_mesher.txt})
+and analyze the skewness histogram of mesh elements to make sure no
+element has a skewness above approximately 0.75, otherwise the mesh
+is of poor quality (and if even a single element has a skewness value
+above 0.80, then you must definitely improve the mesh). To draw the
+skewness histogram on the screen, type \texttt{gnuplot plot\_mesh\_quality\_histogram.gnu}.
+
+\chapter{\label{cha:Creating-Distributed-Databases}Creating the Distributed
+Databases}
+
+After using \texttt{xmeshfem3D} or \texttt{xdecompose\_mesh}, the
+next step in the workflow is to compile \texttt{xgenerate\_}~\\
+ \texttt{databases}. This program is going to create all the missing
+information needed by the SEM solver.
 
@@
 \noindent \begin{centering}
-\includegraphics[width=.6\textwidth]{figures/workflow.pdf}
+\includegraphics[width=0.6\textwidth]{figures/workflow} 
 \par\end{centering}
 
-\caption{Schematic workflow for a SPECFEM3D Cartesian simulation. The executable \texttt{xgenerate\_databases} creates the GLL mesh points and assigns specific model parameters.}
+\caption{Schematic workflow for a SPECFEM3D Cartesian simulation. The executable
+\texttt{xgenerate\_databases} creates the GLL mesh points and assigns
+specific model parameters.}
 
@@
 \begin{lyxcode}
-{\small make xgenerate\_databases}{\small \par}
-\end{lyxcode}
-Input for the program is provided through
-the main parameter file \texttt{Par\_file}, which resides in the subdirectory \texttt{DATA}. Please note that
-\texttt{xgenerate\_databases} must be called directly from the \texttt{bin/} directory, as most of the binaries of the package.
+{\small make~xgenerate\_databases}{\small \par}
+\end{lyxcode}
+Input for the program is provided through the main parameter file
+\texttt{Par\_file}, which resides in the subdirectory \texttt{DATA}.
+Please note that \texttt{xgenerate\_databases} must be called directly
+from the \texttt{bin/} directory, as most of the binaries of the package.
 
@@
 
-Before running \texttt{xgenerate\_databases}, a number of parameters need to be set in the main parameter \texttt{Par\_file}
-located in the subdirectory \texttt{DATA}:
-
+Before running \texttt{xgenerate\_databases}, a number of parameters
+need to be set in the main parameter \texttt{Par\_file} located in
+the subdirectory \texttt{DATA}:
 \begin{description}
@@
 region resides.} 
-\item [{\texttt{NPROC}}] The number of MPI processors, each one is assigned one slice of the whole mesh.
-\item [{\texttt{NSTEP}}] The number of time steps of the simulation. This controls the
-length of the numerical simulation, i.e., twice the number of time steps
-requires twice as much CPU time. This feature is not used at the time
-of generating the distributed databases but is required for the solver, i.e., you may change this
-parameter after running  \texttt{xgenerate\_databases}.
-\item [{\texttt{DT}}] The length of each time step in seconds.
-This feature is not used at the time of generating the distributed databases but is required for the solver.
-Please see also Section~\ref{sec:Choosing-the-Time-Step} for further details.
-\item [{\texttt{NGNOD}}] The number of nodes for 2D and 3D shape functions for hexahedra. We use either
-8-node mesh elements (bricks) or 27-node elements. If you use the internal mesher, the only option is 8-node bricks
-(27-node elements are not supported). \texttt{CUBIT} does not support HEX27 elements either (it can generate them,
-but they are flat, i.e. identical to HEX8). To generate HEX27 elements with curvature properly taken into account,
-you can use \texttt{Gmsh} \url{http://geuz.org/gmsh/}
+\item [{\texttt{NPROC}}] The number of MPI processors, each one is assigned
+one slice of the whole mesh. 
+\item [{\texttt{NSTEP}}] The number of time steps of the simulation. This
+controls the length of the numerical simulation, i.e., twice the number
+of time steps requires twice as much CPU time. This feature is not
+used at the time of generating the distributed databases but is required
+for the solver, i.e., you may change this parameter after running
+\texttt{xgenerate\_databases}. 
+\item [{\texttt{DT}}] The length of each time step in seconds. This feature
+is not used at the time of generating the distributed databases but
+is required for the solver. Please see also Section~\ref{sec:Choosing-the-Time-Step}
+for further details. 
+\item [{\texttt{NGNOD}}] The number of nodes for 2D and 3D shape functions
+for hexahedra. We use either 8-node mesh elements (bricks) or 27-node
+elements. If you use the internal mesher, the only option is 8-node
+bricks (27-node elements are not supported). \texttt{CUBIT} does not
+support HEX27 elements either (it can generate them, but they are
+flat, i.e. identical to HEX8). To generate HEX27 elements with curvature
+properly taken into account, you can use \texttt{Gmsh} \url{http://geuz.org/gmsh/} 
 \item [{\texttt{MODEL}}] Must be set to one of the following:
-%
+
 \begin{description}
-\item [{\textmd{Models defined by mesh parameters:}}]~
+\item [{\textmd{Models defined by mesh parameters:}}] ~
+
 \begin{description}
-\item [{\texttt{default}}] Uses model parameters as defined by meshing procedures described
-in the previous Chapter~\ref{cha:Mesh-Generation}.
+\item [{\texttt{default}}] Uses model parameters as defined by meshing
+procedures described in the previous Chapter~\ref{cha:Mesh-Generation}. 
 \end{description}
 \end{description}
-%
+\end{description}
+
 \begin{description}
-\item [{\textmd{1D~models~with~real~structure:}}]~
+\item [{\textmd{1D~models~with~real~structure:}}] ~
+
 \begin{description}
-\item [{\texttt{1D\_prem}}] Isotropic version of the spherically
-symmetric Preliminary Reference Earth Model (PREM) \citep{DzAn81}.
-\item [{\texttt{1D\_socal}}] A standard isotropic 1D model for Southern California.
+\item [{\texttt{1D\_prem}}] Isotropic version of the spherically symmetric
+Preliminary Reference Earth Model (PREM) \citep{DzAn81}.
+
+\begin{description}
+\item [{\texttt{1D\_socal}}] A standard isotropic 1D model for Southern
+California. 
 \item [{\texttt{1D\_cascadia}}] Isotropic 1D profile for the Cascadia region. 
@@
 \end{description}
-%
+\end{description}
+
 \begin{description}
-\item [{\textmd{Fully~3D~models:}}]~
+\item [{\textmd{Fully~3D~models:}}] ~
+
 \begin{description}
 \item [{\texttt{aniso}}] For a user-specified fully anisotropic model.
-Parameters are set up in routines located in file \texttt{model\_aniso.f90} in directory \texttt{src/generate\_databases/}.
-See Chapter~\ref{cha:-Changing-the} for a discussion on how to specify your own 3D model.
-\item [{\texttt{external}}] For a user-specified isotropic model which uses externally defined model parameters.
-Uses external model definitions set up in routines located in file \texttt{model\_external\_values.f90} in directory \texttt{src/generate\_databases/}.
-Please modify these generic template routines to use your own model definitions.
-\item [{\texttt{gll}}] For a user-specified isotropic model which uses external binary files for $v_p$, $v_s$ and $\rho$.
-Binary files are given in the same format as when outputted by the \texttt{xgenerate\_databases} executable when using option \texttt{SAVE\_MESH\_FILES}.
-These binary files define the model parameters on all GLL points which can be used for iterative inversion procedures.
-\item [{\texttt{salton\_trough}}] A 3D $V_p$ model for Southern California.
-Users must provide the corresponding data file \texttt{regrid3\_vel\_p.bin} in directory \texttt{DATA/st\_3D\_block\_harvard/}.
-\item [{\texttt{tomo}}] For a user-specified 3D isotropic model which uses a tomographic model file \texttt{tomographic\_model.xyz}
-in directory \texttt{DATA}. See Section \ref{sec:Using-tomographic},
-for a discussion on how to specify your own 3D tomographic model.
+Parameters are set up in routines located in file \texttt{model\_aniso.f90}
+in directory \texttt{src/generate\_databases/}. See Chapter~\ref{cha:-Changing-the}
+for a discussion on how to specify your own 3D model.
+
+\begin{description}
+\item [{\texttt{external}}] For a user-specified isotropic model which
+uses externally defined model parameters. Uses external model definitions
+set up in routines located in file \texttt{model\_external\_values.f90}
+in directory \texttt{src/generate\_databases/}. Please modify these
+generic template routines to use your own model definitions. 
+\item [{\texttt{gll}}] For a user-specified isotropic model which uses
+external binary files for $v_{p}$, $v_{s}$ and $\rho$. Binary files
+are given in the same format as when outputted by the \texttt{xgenerate\_databases}
+executable when using option \texttt{SAVE\_MESH\_FILES}. These binary
+files define the model parameters on all GLL points which can be used
+for iterative inversion procedures. 
+\item [{\texttt{salton\_trough}}] A 3D $V_{p}$ model for Southern California.
+Users must provide the corresponding data file \texttt{regrid3\_vel\_p.bin}
+in directory \texttt{DATA/st\_3D\_block\_harvard/}. 
+\item [{\texttt{tomo}}] For a user-specified 3D isotropic model which uses
+a tomographic model file \texttt{tomographic\_model.xyz} in directory
+\texttt{DATA}. See Section \ref{sec:Using-tomographic}, for a discussion
+on how to specify your own 3D tomographic model. 
 \end{description}
 \end{description}
-%
-%
-\item [{\texttt{APPROXIMATE\_OCEAN\_LOAD}}] Set to \texttt{.true.} if the effect of the oceans
-on seismic wave propagation should be incorporated based upon the
-(rough) approximate treatment discussed in \citet{KoTr02b}. This feature
-is inexpensive from a numerical perspective, both in terms of memory
-requirements and CPU time. This approximation is accurate at periods
-of roughly 20~s and longer. At shorter periods the effect of water
-phases/reverberations is not taken into account, even when the flag is on.
-If you want to model the effect of a fluid-solid model at short periods,
-then set this flag to \texttt{.false.} and mesh the fluid layer explicitly in your mesher,
-so that it is computed accurately and without this approximation.
-\item [{\texttt{TOPOGRAPHY}}] This feature is only effective if \texttt{APPROXIMATE\_OCEAN\_LOAD} is set to \texttt{.true.}.
-Set to \texttt{.true.} if topography and
-bathymetry should be read in based upon the topography file specified in the main
-constants file \texttt{constants.h} found in subdirectory \texttt{src/shared/} to evaluate
-elevations. If not set, elevations will be read from the numerical mesh.
+\item [{\texttt{APPROXIMATE\_OCEAN\_LOAD}}] Set to \texttt{.true.} if the
+effect of the oceans on seismic wave propagation should be incorporated
+based upon the (rough) approximate treatment discussed in \citet{KoTr02b}.
+This feature is inexpensive from a numerical perspective, both in
+terms of memory requirements and CPU time. This approximation is accurate
+at periods of roughly 20~s and longer. At shorter periods the effect
+of water phases/reverberations is not taken into account, even when
+the flag is on. If you want to model the effect of a fluid-solid model
+at short periods, then set this flag to \texttt{.false.} and mesh
+the fluid layer explicitly in your mesher, so that it is computed
+accurately and without this approximation. 
+\item [{\texttt{TOPOGRAPHY}}] This feature is only effective if \texttt{APPROXIMATE\_OCEAN\_LOAD}
+is set to \texttt{.true.}. Set to \texttt{.true.} if topography and
+bathymetry should be read in based upon the topography file specified
+in the main constants file \texttt{constants.h} found in subdirectory
+\texttt{src/shared/} to evaluate elevations. If not set, elevations
+will be read from the numerical mesh. 
 \item [{\texttt{ATTENUATION}}] Set to \texttt{.true.} if attenuation should
@@
 expensive. See \citet{KoTr99,KoTr02a} for a discussion on the implementation
-of attenuation based upon standard linear solids.
-Please note that the Vp- and Vs-velocities of your model are given for a reference frequency.
-To change this reference frequency, you change the value of
-\texttt{ATTENUATION\_f0\_REFERENCE} in the main constants file \texttt{constants.h} found in subdirectory \texttt{src/shared/}.
-\item [{\texttt{FULL\_ATTENUATION\_SOLID}}] Set to \texttt{.true.} to add $Q_\kappa$ attenuation to the simulations in addition to $Q_\mu$
-attenuation when the \texttt{ATTENUATION} flag above is on.
-Should be off in almost all cases for seismic wave propagation, since $Q_\kappa$ is almost always negligible in Earth models.
-However, this parameter can be useful for instance for ocean acoustics simulations, when $Q_\kappa$ is not negligible in the solid
-layer at the sea bottom.
-\item [{\texttt{ANISOTROPY}}] Set to \texttt{.true.} if you
-want to use an anisotropy model. Please see the file \texttt{model\_aniso.f90} in subdirectory \texttt{src/generate\_databases/} for the current implementation of anisotropic models.
-\item [{\texttt{TOMOGRAPHY\_PATH}}] Directory in which the tomography files are stored for using external
-tomographic Earth models (please read Chapter \ref{cha:-Changing-the} and Section \ref{sec:Using-tomographic}
+of attenuation based upon standard linear solids. Please note that
+the Vp- and Vs-velocities of your model are given for a reference
+frequency. To change this reference frequency, you change the value
+of \texttt{ATTENUATION\_f0\_REFERENCE} in the main constants file
+\texttt{constants.h} found in subdirectory \texttt{src/shared/}. 
+\item [{\texttt{FULL\_ATTENUATION\_SOLID}}] Set to \texttt{.true.} to add
+$Q_{\kappa}$ attenuation to the simulations in addition to $Q_{\mu}$
+attenuation when the \texttt{ATTENUATION} flag above is on. Should
+be off in almost all cases for seismic wave propagation, since $Q_{\kappa}$
+is almost always negligible in Earth models. However, this parameter
+can be useful for instance for ocean acoustics simulations, when $Q_{\kappa}$
+is not negligible in the solid layer at the sea bottom. 
+\item [{\texttt{ANISOTROPY}}] Set to \texttt{.true.} if you want to use
+an anisotropy model. Please see the file \texttt{model\_aniso.f90}
+in subdirectory \texttt{src/generate\_databases/} for the current
+implementation of anisotropic models. 
+\item [{\texttt{TOMOGRAPHY\_PATH}}] Directory in which the tomography files
+are stored for using external tomographic Earth models (please read
+Chapter \ref{cha:-Changing-the} and Section \ref{sec:Using-tomographic}
 `Using external tomographic Earth models' for further details.). 
 \item [{\texttt{USE\_OLSEN\_ATTENUATION}}] Set to \texttt{.true.} if you
-want to use the attenuation model that scaled from the S-wave speed model
-using Olsen's empirical relation (see \citet{OlDaBr03}).
-\item [{\texttt{OLSEN\_ATTENUATION\_RATIO}}] Determines the Olsen's constant in Olsen's empirical relation (see \citet{OlDaBr03}).
-\item [{\texttt{PML\_CONDITIONS}}] Set to \texttt{.true.} to turn on C-PML boundary conditions for a regional simulation. Both fluids and elastic solids are supported.
-\item [{\texttt{PML\_INSTEAD\_OF\_FREE\_SURFACE}}] Set to \texttt{.true.} to turn on C-PML boundary conditions on the top surface
-instead of the usual free surface.
-\item [{\texttt{f0\_FOR\_PML}}] Determines the dominant frequency that will be used in the calculation of PML damping profiles; this should be
-set to the same (or similar) dominant frequency as that of the source that you will use in your simulation. If you plan to use a Dirac source,
-then use the dominant frequency of the source wavelet with which you plan to convolve your seismograms later on in post-processing.
-\item [{\texttt{STACEY\_ABSORBING\_CONDITIONS}}] Set to \texttt{.true.} to turn
-on Clayton-Enquist absorbing boundary conditions (see \citet{KoTr99}). In almost all cases it is much better to use CPML absorbing
-layers (see the options above) and leave this flag to \texttt{.false.}.
-\item [{\texttt{STACEY\_INSTEAD\_OF\_FREE\_SURFACE}}] Set to \texttt{.true.} to turn on
-absorbing boundary conditions on the top surface which by default constitutes a free surface of the model.
+want to use the attenuation model that scaled from the S-wave speed
+model using Olsen's empirical relation (see \citet{OlDaBr03}). 
+\item [{\texttt{OLSEN\_ATTENUATION\_RATIO}}] Determines the Olsen's constant
+in Olsen's empirical relation (see \citet{OlDaBr03}). 
+\item [{\texttt{PML\_CONDITIONS}}] Set to \texttt{.true.} to turn on C-PML
+boundary conditions for a regional simulation. Both fluids and elastic
+solids are supported. 
+\item [{\texttt{PML\_INSTEAD\_OF\_FREE\_SURFACE}}] Set to \texttt{.true.}
+to turn on C-PML boundary conditions on the top surface instead of
+the usual free surface. 
+\item [{\texttt{f0\_FOR\_PML}}] Determines the dominant frequency that
+will be used in the calculation of PML damping profiles; this should
+be set to the same (or similar) dominant frequency as that of the
+source that you will use in your simulation. If you plan to use a
+Dirac source, then use the dominant frequency of the source wavelet
+with which you plan to convolve your seismograms later on in post-processing. 
+\item [{\texttt{STACEY\_ABSORBING\_CONDITIONS}}] Set to \texttt{.true.}
+to turn on Clayton-Enquist absorbing boundary conditions (see \citet{KoTr99}).
+In almost all cases it is much better to use CPML absorbing layers
+(see the options above) and leave this flag to \texttt{.false.}. 
+\item [{\texttt{STACEY\_INSTEAD\_OF\_FREE\_SURFACE}}] Set to \texttt{.true.}
+to turn on absorbing boundary conditions on the top surface which
+by default constitutes a free surface of the model. 
 \item [{\texttt{CREATE\_SHAKEMAP}}] Set this flag to \texttt{.true.} to
-create a ShakeMap\textregistered, i.e., a peak ground velocity map
-of the maximum absolute value of the two horizontal components of the velocity vector.
+create a ShakeMap\textregistered{}, i.e., a peak ground velocity map
+of the maximum absolute value of the two horizontal components of
+the velocity vector. 
 \item [{\texttt{MOVIE\_SURFACE}}] Set to \texttt{.false.}, unless you want
@@
 Turning this option on generates large output files. See Section~\ref{sec:Movies}
-for a discussion on the generation of movies. This feature is only relevant for the solver.
-\item [{\texttt{MOVIE\_TYPE}}] Set this flag to 1 to show the top
-  surface (tomography + oceans) only, to 2 to show all external faces
-  of the mesh (i.e. topography + vertical edges + bottom) in shakemaps and surface movies.
+for a discussion on the generation of movies. This feature is only
+relevant for the solver. 
+\item [{\texttt{MOVIE\_TYPE}}] Set this flag to 1 to show the top surface
+(tomography + oceans) only, to 2 to show all external faces of the
+mesh (i.e. topography + vertical edges + bottom) in shakemaps and
+surface movies. 
 \item [{\texttt{MOVIE\_VOLUME}}] Set to \texttt{.false.}, unless you want
@@
 Turning this option on generates huge output files. See Section~\ref{sec:Movies}
-for a discussion on the generation of movies. This feature is only relevant for the solver.
+for a discussion on the generation of movies. This feature is only
+relevant for the solver. 
 \item [{\texttt{SAVE\_DISPLACEMENT}}] Set this flag to \texttt{.true.}
@@
 generation of movies. This feature is only relevant for the solver. 
-\item [{\texttt{HDUR\_MOVIE}}] Determines the half duration of the source time function for the movie simulations.
-When this parameter is set to be 0, a default half duration that corresponds to the accuracy of the simulation is provided.
-Otherwise, it adds this half duration to the half duration specified in the source file \texttt{CMTSOLUTION},
+\item [{\texttt{HDUR\_MOVIE}}] Determines the half duration of the source
+time function for the movie simulations. When this parameter is set
+to be 0, a default half duration that corresponds to the accuracy
+of the simulation is provided. Otherwise, it adds this half duration
+to the half duration specified in the source file \texttt{CMTSOLUTION},
 thus simulates longer periods to make the movie images look smoother. 
 \item [{\texttt{SAVE\_MESH\_FILES}}] Set this flag to \texttt{.true.} to
-save ParaView \urlwithparentheses{www.paraview.org}
-mesh files for subsequent viewing. Turning the flag on generates large
-(distributed) files in the \texttt{LOCAL\_PATH} directory. See Section~\ref{sec:Mesh-graphics}
+save ParaView \urlwithparentheses{www.paraview.org} mesh files for
+subsequent viewing. Turning the flag on generates large (distributed)
+files in the \texttt{LOCAL\_PATH} directory. See Section~\ref{sec:Mesh-graphics}
 for a discussion of mesh viewing features. 
 \item [{\texttt{LOCAL\_PATH}}] Directory in which the distributed databases
-will be written. Generally one uses a directory on the
-local disk of the compute nodes, although on some machines these databases
-are written on a parallel (global) file system (see also the earlier
-discussion of the \texttt{LOCAL\_PATH\_IS\_ALSO\_GLOBAL} flag in Chapter~\ref{cha:Getting-Started}).
-\texttt{xgenerate\_databases} generates the necessary databases in parallel, one set
-for each of the \texttt{NPROC} slices that constitutes
-the mesh (see Figure~\ref{fig:mount.partitions} and Figure~\ref{fig:For-parallel-computing}). After
-the executable finishes, you can log in to one of the compute nodes and view
-the contents of the \texttt{LOCAL\_PATH} directory to see the (many)
-files generated by \texttt{xgenerate\_databases}.
-Please note that the \texttt{LOCAL\_PATH} directory should already contain the output files of
-the partitioner, i.e. from \texttt{xdecompose\_mesh} or \texttt{xmeshfem3D}.
+will be written. Generally one uses a directory on the local disk
+of the compute nodes, although on some machines these databases are
+written on a parallel (global) file system (see also the earlier discussion
+of the \texttt{LOCAL\_PATH\_IS\_ALSO\_GLOBAL} flag in Chapter~\ref{cha:Getting-Started}).
+\texttt{xgenerate\_databases} generates the necessary databases in
+parallel, one set for each of the \texttt{NPROC} slices that constitutes
+the mesh (see Figure~\ref{fig:mount.partitions} and Figure~\ref{fig:For-parallel-computing}).
+After the executable finishes, you can log in to one of the compute
+nodes and view the contents of the \texttt{LOCAL\_PATH} directory
+to see the (many) files generated by \texttt{xgenerate\_databases}.
+Please note that the \texttt{LOCAL\_PATH} directory should already
+contain the output files of the partitioner, i.e. from \texttt{xdecompose\_mesh}
+or \texttt{xmeshfem3D}. 
 \item [{\texttt{NTSTEP\_BETWEEN\_OUTPUT\_INFO}}] This parameter specifies
@@
 \texttt{NTSTEP\_BETWEEN\_OUTPUT\_SEISMOS} to a relatively high value
-to avoid writing to the seismograms too often. This feature is only relevant for the solver.
-\item [{\texttt{USE\_FORCE\_POINT\_SOURCE}}] Turn this flag on to use a (tilted) \texttt{FORCESOLUTION} force
-point source instead of a \texttt{CMTSOLUTION} moment-tensor source.
-When the force source does not fall exactly at a grid point, the solver interpolates the force between grid points
-using Lagrange interpolants. This can be useful e.g. for oil industry foothills simulations
-in which the source is a vertical force, normal force, tilted force, or an impact etc.
-Note that in the \texttt{FORCESOLUTION} file, you will need to edit
-the East, North and vertical components of an arbitrary (non-unitary)
-direction vector of the force vector; thus refer to
-Appendix~\ref{cha:Coordinates} for the orientation of the reference
-frame. This vector is made unitary internally in the solver and thus
-only its direction matters here; its norm is ignored and the norm of
-the force used is the factor force source times the source time
-function.
-\item [{\texttt{USE\_RICKER\_TIME\_FUNCTION}}] Turn this flag on to
-use a Ricker source time function instead of the source time functions set by default
-to represent a (tilted) \texttt{FORCESOLUTION} force point source or a \texttt{CMTSOLUTION} moment-tensor source.
-Originally, if a \texttt{CMTSOLUTION} moment-tensor source is used, a (pseudo) Heaviside step
-function with a very short half duration is defined for elastic cases to represent the permanent slip on the fault while
-in the acoustic case a Gaussian source time function with a similarly short half duration is defined to physically describe
-actions within the fluid. Otherwise, if a \texttt{FORCESOLUTION} force source is used, a (pseudo) Dirac delta source time
-function is defined by default. Any other source-time function may then be obtained by convolution.
+to avoid writing to the seismograms too often. This feature is only
+relevant for the solver. 
+\item [{\texttt{USE\_FORCE\_POINT\_SOURCE}}] Turn this flag on to use a
+(tilted) \texttt{FORCESOLUTION} force point source instead of a \texttt{CMTSOLUTION}
+moment-tensor source. When the force source does not fall exactly
+at a grid point, the solver interpolates the force between grid points
+using Lagrange interpolants. This can be useful e.g. for oil industry
+foothills simulations in which the source is a vertical force, normal
+force, tilted force, or an impact etc. Note that in the \texttt{FORCESOLUTION}
+file, you will need to edit the East, North and vertical components
+of an arbitrary (non-unitary) direction vector of the force vector;
+thus refer to Appendix~\ref{cha:Coordinates} for the orientation
+of the reference frame. This vector is made unitary internally in
+the solver and thus only its direction matters here; its norm is ignored
+and the norm of the force used is the factor force source times the
+source time function. 
+\item [{\texttt{USE\_RICKER\_TIME\_FUNCTION}}] Turn this flag on to use
+a Ricker source time function instead of the source time functions
+set by default to represent a (tilted) \texttt{FORCESOLUTION} force
+point source or a \texttt{CMTSOLUTION} moment-tensor source. Originally,
+if a \texttt{CMTSOLUTION} moment-tensor source is used, a (pseudo)
+Heaviside step function with a very short half duration is defined
+for elastic cases to represent the permanent slip on the fault while
+in the acoustic case a Gaussian source time function with a similarly
+short half duration is defined to physically describe actions within
+the fluid. Otherwise, if a \texttt{FORCESOLUTION} force source is
+used, a (pseudo) Dirac delta source time function is defined by default.
+Any other source-time function may then be obtained by convolution. 
 \item [{\texttt{PRINT\_SOURCE\_TIME\_FUNCTION}}] Turn this flag on to print
@@
 \end{description}
-
-If you use PML, the mesh elements that belong to the PML layers can be acoustic or elastic, but not viscoelastic nor poroelastic.
-Then, when defining your model, you should define these absorbing elements as either acoustic or elastic.
-In you forget to do that, the code will fix the problem by automatically converting the viscoelastic or poroelastic PML
-elements to elastic. This means that strictly speaking the PML layer will not be perfectly matched any more, since the physical
-model will change from viscoelastic or poroelastic to elastic at the entrance of the PML, but in practice this is sufficient and
-produces only tiny / negligible spurious reflections.
+If you use PML, the mesh elements that belong to the PML layers can
+be acoustic or elastic, but not viscoelastic nor poroelastic. Then,
+when defining your model, you should define these absorbing elements
+as either acoustic or elastic. In you forget to do that, the code
+will fix the problem by automatically converting the viscoelastic
+or poroelastic PML elements to elastic. This means that strictly speaking
+the PML layer will not be perfectly matched any more, since the physical
+model will change from viscoelastic or poroelastic to elastic at the
+entrance of the PML, but in practice this is sufficient and produces
+only tiny / negligible spurious reflections.
 
 If you use PML and an external tomographic velocity and density model,
-you should be careful because mathematically a PML cannot handle heterogeneities along the
-normal to the PML edge inside the PML layer. This comes from the fact that the damping profile
-that is defined assumes a constant velocity and density model along the normal
-direction.
-
-Thus, you need to modify your velocity and density model in order for it to be 1D inside
-the PML, as shown in Figure~\ref{fig:modify_external_velocity_model_to_use_PML}.
+you should be careful because mathematically a PML cannot handle heterogeneities
+along the normal to the PML edge inside the PML layer. This comes
+from the fact that the damping profile that is defined assumes a constant
+velocity and density model along the normal direction.
+
+Thus, you need to modify your velocity and density model in order
+for it to be 1D inside the PML, as shown in Figure~\ref{fig:modify_external_velocity_model_to_use_PML}.
 
@@
 
-To summarize, only use a 3D velocity and density model inside the physical region, and in
-all the PML layers extend it by continuity from its values along the
-inner PML edge.
+To summarize, only use a 3D velocity and density model inside the
+physical region, and in all the PML layers extend it by continuity
+from its values along the inner PML edge.
 
@@
 \noindent \begin{centering}
-\includegraphics[width=3in]{figures/how_to_use_PML_when_a_tomographic_velocity_model_is_used.png}
+\includegraphics[width=3in]{figures/how_to_use_PML_when_a_tomographic_velocity_model_is_used} 
 \par\end{centering}
-\caption{How to modify your external 3D velocity and density model in order to use PML.
-Such a modification is not needed when using Stacey absorbing boundary conditions (but such conditions
-are significantly less efficient).}
+
+\caption{How to modify your external 3D velocity and density model in order
+to use PML. Such a modification is not needed when using Stacey absorbing
+boundary conditions (but such conditions are significantly less efficient).}
+
 \label{fig:modify_external_velocity_model_to_use_PML} 
 \end{figure}
+
 %%
@@
 
-The parameter \texttt{DT} sets the length of each time step in seconds. The value of this parameter is crucial for the stability of
-the spectral-element simulation.
-Your time step \texttt{DT} will depend on the minimum ratio between the distance $h$ of neighboring mesh points and the wave speeds $v$ defined in your model.
-The condition for the time step $\Delta t$ is:
-
-\begin{lyxcode}
-$\Delta t < C ~\mathrm{min}_\Omega (~ h / v ~)$
-\end{lyxcode}
-where $C$ is the so-called Courant number and $\Omega$ denotes the model volume.
-The distance $h$ depends on the mesh element size and the number of
-GLL points \texttt{NGLL} specified in the main constants file \texttt{constants.h} located in the \texttt{src/shared/} subdirectory.
-The wave speed $v$ is determined based on your model's P- (or S-) wave speed values.
-
-The database generator \texttt{xgenerate\_databases}, as well as the internal mesher \texttt{xmeshfem3D}, are trying
-to evaluate the value of $\Delta t$ for empirically chosen Courant numbers $C \sim 0.3$.
-If you used the mesher \texttt{xmeshfem3D} to generate your mesh, you should set the value suggested in \texttt{OUTPUT\_FILES/output\_mesher.txt} file, which is created after the mesher completed.
-In case you used CUBIT to create the mesh, you might use an arbitrary value when running \texttt{xgenerate\_databases} and then use the value suggested in the ~\\
-\texttt{OUTPUT\_FILES/output\_mesher.txt} file after the database generation completed.
-Note that the implemented Newmark time scheme uses this time step globally, thus your simulations become more expensive for very small mesh elements in high wave-speed regions.
-Please be aware of this restriction when constructing your mesh in Chapter~\ref{cha:Mesh-Generation}.
+The parameter \texttt{DT} sets the length of each time step in seconds.
+The value of this parameter is crucial for the stability of the spectral-element
+simulation. Your time step \texttt{DT} will depend on the minimum
+ratio between the distance $h$ of neighboring mesh points and the
+wave speeds $v$ defined in your model. The condition for the time
+step $\Delta t$ is:
+\begin{lyxcode}
+$\Delta t<C~\mathrm{min}_{\Omega}(~h/v~)$
+\end{lyxcode}
+where $C$ is the so-called Courant number and $\Omega$ denotes the
+model volume. The distance $h$ depends on the mesh element size and
+the number of GLL points \texttt{NGLL} specified in the main constants
+file \texttt{constants.h} located in the \texttt{src/shared/} subdirectory.
+The wave speed $v$ is determined based on your model's P- (or S-)
+wave speed values.
+
+The database generator \texttt{xgenerate\_databases}, as well as the
+internal mesher \texttt{xmeshfem3D}, are trying to evaluate the value
+of $\Delta t$ for empirically chosen Courant numbers $C\sim0.3$.
+If you used the mesher \texttt{xmeshfem3D} to generate your mesh,
+you should set the value suggested in \texttt{OUTPUT\_FILES/output\_mesher.txt}
+file, which is created after the mesher completed. In case you used
+CUBIT to create the mesh, you might use an arbitrary value when running
+\texttt{xgenerate\_databases} and then use the value suggested in
+the ~\\
+ \texttt{OUTPUT\_FILES/output\_mesher.txt} file after the database
+generation completed. Note that the implemented Newmark time scheme
+uses this time step globally, thus your simulations become more expensive
+for very small mesh elements in high wave-speed regions. Please be
+aware of this restriction when constructing your mesh in Chapter~\ref{cha:Mesh-Generation}.
 
@@
 
-Now that you have successfully generated the databases, you are ready to compile
-the solver. In the main directory, type
-\begin{lyxcode}
-{\small make xspecfem3D}{\small \par}
+Now that you have successfully generated the databases, you are ready
+to compile the solver. In the main directory, type 
+\begin{lyxcode}
+{\small make~xspecfem3D}{\small \par}
 \end{lyxcode}
@@
 \begin{description}
-\item [{\texttt{Par\_file}}] the main parameter file which was discussed in detail in the previous
-Chapter~\ref{cha:Creating-Distributed-Databases},
-\item [{\texttt{CMTSOLUTION}} or {\texttt{FORCESOLUTION}}] the earthquake source parameter file or the force source parameter file, and
+\item [{\texttt{Par\_file}}] the main parameter file which was discussed
+in detail in the previous Chapter~\ref{cha:Creating-Distributed-Databases}, 
+\item [{\texttt{CMTSOLUTION}} or {\texttt{FORCESOLUTION}}] the earthquake
+source parameter file or the force source parameter file, and 
 \item [{\texttt{STATIONS}}] the stations file. 
 \end{description}
-
-Most parameters in the \texttt{Par\_file}
-should be set prior to running the databases generation. Only the following parameters
-may be changed after running \texttt{xgenerate\_databases}:
-
+Most parameters in the \texttt{Par\_file} should be set prior to running
+the databases generation. Only the following parameters may be changed
+after running \texttt{xgenerate\_databases}:
 \begin{itemize}
@@
 \item the time step parameters \texttt{NSTEP} and \texttt{DT} 
-\item the absorbing boundary control parameter \texttt{PML\_CONDITIONS} on condition that the\\
-  \texttt{PML\_INSTEAD\_OF\_FREE\_SURFACE} flag remains unmodified after running the databases generation.
+\item the absorbing boundary control parameter \texttt{PML\_CONDITIONS}
+on condition that the\\
+ \texttt{PML\_INSTEAD\_OF\_FREE\_SURFACE} flag remains unmodified
+after running the databases generation. 
 \item the movie control parameters \texttt{MOVIE\_SURFACE}, \texttt{MOVIE\_VOLUME},
 and \texttt{NTSTEPS\_BETWEEN\_FRAMES} 
-\item the ShakeMap\textregistered option \texttt{CREATE\_SHAKEMAP}
+\item the ShakeMap\textregistered{}option \texttt{CREATE\_SHAKEMAP} 
 \item the output information parameters \texttt{MOVIE\_TYPE}, \texttt{NTSTEP\_BETWEEN\_OUTPUT\_INFO}
@@
 Any other change to the \texttt{Par\_file} implies rerunning both
-the database generator \texttt{xgenerate\_databases} and the solver \texttt{xspecfem3D}.
+the database generator \texttt{xgenerate\_databases} and the solver
+\texttt{xspecfem3D}.
 
 For any particular earthquake, the \texttt{CMTSOLUTION} file that
-represents the point source may be obtained directly from the Harvard Centroid-Moment Tensor (CMT) web page \urlwithparentheses{www.seismology.harvard.edu}.
+represents the point source may be obtained directly from the Harvard
+Centroid-Moment Tensor (CMT) web page \urlwithparentheses{www.seismology.harvard.edu}.
 It looks like this:
-
-\begin{lyxcode}
-{\small }%
+\begin{lyxcode}
 \begin{figure}[H]
 \noindent \begin{centering}
-{\small \includegraphics[width=1\textwidth]{figures/Hollywood_CMT} }
-\par\end{centering}{\small \par}
+{\small \includegraphics[width=1\textwidth]{figures/Hollywood_CMT}
+} 
+\par\end{centering}
 
@@
 format from the Harvard CMT catalog. \textbf{M} is the moment tensor,
-$M_{0}${\small{} }is the seismic moment, and $M_{w}$ is the moment
+$M_{0}${\small {} }is the seismic moment, and $M_{w}$ is the moment
 magnitude.}
-
 \end{figure}
-{\small \par}
+
 \end{lyxcode}
 The \texttt{CMTSOLUTION} file should be edited in the following way:
-
 \begin{itemize}
@@
 script \texttt{convolve\_source\_timefunction.csh} for this purpose,
-or alternatively use signal-processing software packages such as SAC \urlwithparentheses{www.llnl.gov/sac}.
-Type
-\begin{lyxcode}
-{\small make xconvolve\_source\_timefunction}{\small \par}
-\end{lyxcode}
+or alternatively use signal-processing software packages such as SAC
+\urlwithparentheses{www.llnl.gov/sac}. Type
+
+\begin{lyxcode}
+{\small make~xconvolve\_source\_timefunction}{\small \par}
+\end{lyxcode}
+
 to compile the code and then set the parameter \texttt{hdur} in \texttt{convolve\_source\_timefunction.csh}
@@
 \end{lyxcode}
+
 where $t_{\mathrm{pde}}$ is the time given in the first line of the
@@
 \end{itemize}
-%
 \begin{figure}
 \noindent \begin{centering}
-\includegraphics[width=2.5in]{figures/gauss_vs_triangle_mod.pdf}
+\includegraphics[width=2.5in]{figures/gauss_vs_triangle_mod} 
 \par\end{centering}
-\caption{Comparison of the shape of a triangle and the Gaussian function actually used.}
+
+\caption{Comparison of the shape of a triangle and the Gaussian function actually
+used.}
+
 \label{fig:gauss.vs.triangle} 
@@
 
-If you know the earthquake source in strike/dip/rake format rather than in \texttt{CMTSOLUTION} format,
-use the C code \texttt{SPECFEM3D\_GLOBE/utils/strike\_dip\_rake\_to\_CMTSOLUTION.c} to convert it.
-The conversion formulas are given for instance in \cite{AkRi80}.
-Note that the \cite{AkRi80} convention is slightly different from the Harvard \texttt{CMTSOLUTION} convention
-(the sign of some components is different). The C code outputs both.
+If you know the earthquake source in strike/dip/rake format rather
+than in \texttt{CMTSOLUTION} format, use the C code \texttt{SPECFEM3D\_GLOBE/utils/strike\_dip\_rake\_to\_CMTSOLUTION.c}
+to convert it. The conversion formulas are given for instance in \citet{AkRi80}.
+Note that the \citet{AkRi80} convention is slightly different from
+the Harvard \texttt{CMTSOLUTION} convention (the sign of some components
+is different). The C code outputs both.
 
@@
 file adheres to the Harvard CMT conventions (see Appendix~\ref{cha:Coordinates}).
-Note that the first line in the \texttt{CMTSOLUTION} file is the Preliminary Determination of Earthquakes (PDE) solution performed by the USGS NEIC, which is used as a seed for the Harvard CMT inversion.
-The PDE solution is based upon P waves and often gives the hypocenter of the earthquake, i.e., the rupture initiation point, whereas the CMT solution gives the `centroid location', which is the location with dominant moment release.
-The PDE solution is not used by our software package but must be present anyway in the first line of the file.
+Note that the first line in the \texttt{CMTSOLUTION} file is the Preliminary
+Determination of Earthquakes (PDE) solution performed by the USGS
+NEIC, which is used as a seed for the Harvard CMT inversion. The PDE
+solution is based upon P waves and often gives the hypocenter of the
+earthquake, i.e., the rupture initiation point, whereas the CMT solution
+gives the `centroid location', which is the location with dominant
+moment release. The PDE solution is not used by our software package
+but must be present anyway in the first line of the file.
 
@@
 
-\noindent The \texttt{FORCESOLUTION} file should be edited in the following way:
-
+\noindent The \texttt{FORCESOLUTION} file should be edited in the
+following way:
 \begin{itemize}
@@
 in the post-processing (see Section \ref{sec:Process-data-and-syn}). 
-\item Set the \texttt{half duration} parameter of the Ricker source
-  time function when {\texttt{USE\_RICKER\_TIME\_FUNCTION}} is turned
-  on in the main parameter file {\texttt{Par\_file}}. In case that the
+\item Set the \texttt{half duration} parameter of the Ricker source time
+function when {\texttt{USE\_RICKER\_TIME\_FUNCTION}} is turned on
+in the main parameter file {\texttt{Par\_file}}. In case that the
 solver uses a (pseudo) Dirac delta source time function to represent
 a force point source, a very short \texttt{half duration} parameter
-  (hdur = 5 * DT) is automatically set by default.
-\item Set the latitude, longitude, depth of the source in geographical
-  coordinates.
+(hdur = 5 {*} DT) is automatically set by default. 
+\item Set the latitude, longitude, depth of the source in geographical coordinates. 
 \item Set the magnitude of the force source. 
-\item Set the components of a (non-unitary) direction vector for the
-  force source in the East/North/Vertical basis (see Appendix A for
-  the orientation of the reference frame).
+\item Set the components of a (non-unitary) direction vector for the force
+source in the East/North/Vertical basis (see Appendix A for the orientation
+of the reference frame). 
 \end{itemize}
-
 \noindent Where necessary, set a \texttt{FORCESOLUTION} file in the
-same   way   you    configure   a   \texttt{CMTSOLUTION}   file   with
-$N_{\mathrm{sources}}$   entries,  one   for   each  subevent   (i.e.,
-concatenate  $N_{\mathrm{sources}}$ \texttt{FORCESOLUTION} files  to a
-single   \texttt{FORCESOLUTION}  file).  At   least  one   entry  (not
-necessarily the first) must have a  zero \texttt{time shift}, and all the other
-entries must have non-negative \texttt{time  shift}. Each subevent can have its
-own half latitude, longitude, depth, \texttt{half duration} and force parameters.
-
-{\small }%
+same way you configure a \texttt{CMTSOLUTION} file with $N_{\mathrm{sources}}$
+entries, one for each subevent (i.e., concatenate $N_{\mathrm{sources}}$
+\texttt{FORCESOLUTION} files to a single \texttt{FORCESOLUTION} file).
+At least one entry (not necessarily the first) must have a zero \texttt{time
+shift}, and all the other entries must have non-negative \texttt{time
+shift}. Each subevent can have its own half latitude, longitude, depth,
+\texttt{half duration} and force parameters.
+
 \begin{figure}[H]
 \noindent \begin{centering}
-{\small \includegraphics[width=5in]{figures/source_timing.pdf} }
-\par\end{centering}{\small \par}
+{\small \includegraphics[width=5in]{figures/source_timing} } 
+\par\end{centering}
 
@@
 \end{figure}
-{\small \par}
 
@@
 
-{\small }%
 \begin{figure}[H]
@@
 {\small \includegraphics{figures/STATIONS_basin_explained} } 
-\par\end{centering}{\small \par}
-
-\caption{Sample \texttt{STATIONS} file. Station latitude and longitude should
+\par\end{centering}
+
+\label{fig:Sample-STATIONS-file.}\caption{Sample \texttt{STATIONS} file. Station latitude and longitude should
 be provided in geographical coordinates. The width of the station
@@
 in the \texttt{constants.h} file).}
-
 \end{figure}
-{\small \par}
 
 Each line represents one station in the following format:
-
 \begin{lyxcode}
@@
 \end{lyxcode}
-The solver \texttt{xspecfem3D} filters the list of stations in file \texttt{DATA/STATIONS}
-to exclude stations that are not located within the region given in
-the \texttt{Par\_file} (between \texttt{LATITUDE\_MIN} and \texttt{LATITUDE\_MAX}
-and between \texttt{LONGITUDE\_MIN} and \texttt{LONGITUDE\_MAX}).
-The filtered file is called \texttt{DATA/STATIONS\_FILTERED}.
+The solver \texttt{xspecfem3D} filters the list of stations in file
+\texttt{DATA/STATIONS} to exclude stations that are not located within
+the region given in the \texttt{Par\_file} (between \texttt{LATITUDE\_MIN}
+and \texttt{LATITUDE\_MAX} and between \texttt{LONGITUDE\_MIN} and
+\texttt{LONGITUDE\_MAX}). The filtered file is called \texttt{DATA/STATIONS\_FILTERED}.
 
@@
 the screen instead by uncommenting a line in \texttt{constants.h}:
-
 \begin{lyxcode}
@@
 While the solver is running, its progress may be tracked by monitoring
-the `\texttt{\small timestamp{*}}' files in the ~\\
-\texttt{\small OUTPUT\_FILES} directory.
-These tiny files look something like this:
-
+the `\texttt{\small timestamp{*}}{\small ' files in the ~}\\
+{\small{} }\texttt{\small OUTPUT\_FILES}{\small{} directory. These tiny
+files look something like this:}{\small \par}
 \begin{lyxcode}
@@
 \end{lyxcode}
-The \texttt{\small timestamp{*}} files provide the \texttt{Mean elapsed
-time per time step in seconds}, which may be used to assess performance
-on various machines (assuming you are the only user on a node), as
-well as the \texttt{Max norm displacement vector U in all slices~(m)}.
-If something is wrong with the model, the mesh, or the source, you
-will see the code become unstable through exponentially growing values
-of the displacement and fluid potential with time, and ultimately
-the run will be terminated by the program. You can control the rate
-at which the timestamp files are written based upon the parameter
-\texttt{NTSTEP\_BETWEEN\_OUTPUT\_INFO} in the \texttt{Par\_file}.
-
-Having set the \texttt{Par\_file} parameters, and having provided the
-\texttt{CMTSOLUTION} (or the \texttt{FORCESOLUTION}) and
-\texttt{STATIONS} files, you are now ready to launch the solver! This
-is most easily accomplished based upon the \texttt{go\_solver} script
-(See Chapter \ref{cha:Scheduler} for information about running through
-a scheduler, e.g., LSF). You may need to edit the last command at the
-end of the script that invokes the \texttt{mpirun} command. The
-\texttt{runall} script compiles and runs both
-\texttt{xgenerate\_databases} and \texttt{xspecfem3D} in
-sequence. This is a safe approach that ensures using the correct
+The \texttt{\small timestamp{*}}{\small{} files provide the }\texttt{\small Mean
+elapsed time per time step in seconds}{\small , which may be used
+to assess performance on various machines (assuming you are the only
+user on a node), as well as the }\texttt{\small Max norm displacement
+vector U in all slices~(m)}{\small . If something is wrong with the
+model, the mesh, or the source, you will see the code become unstable
+through exponentially growing values of the displacement and fluid
+potential with time, and ultimately the run will be terminated by
+the program. You can control the rate at which the timestamp files
+are written based upon the parameter }\texttt{\small NTSTEP\_BETWEEN\_OUTPUT\_INFO}{\small{}
+in the }\texttt{\small Par\_file}{\small .}{\small \par}
+
+Having set the \texttt{Par\_file} parameters, and having provided
+the \texttt{CMTSOLUTION} (or the \texttt{FORCESOLUTION}) and \texttt{STATIONS}
+files, you are now ready to launch the solver! This is most easily
+accomplished based upon the \texttt{go\_solver} script (See Chapter
+\ref{cha:Scheduler} for information about running through a scheduler,
+e.g., LSF). You may need to edit the last command at the end of the
+script that invokes the \texttt{mpirun} command. The \texttt{runall}
+script compiles and runs both \texttt{xgenerate\_databases} and \texttt{xspecfem3D}
+in sequence. This is a safe approach that ensures using the correct
 combination of distributed database output and solver input.
@@
 
-For the same model, one can rerun the solver for different events by
-simply changing the \texttt{CMTSOLUTION} or \texttt{FORCESOLUTION}
+For the same model, one can rerun the solver for different events
+by simply changing the \texttt{CMTSOLUTION} or \texttt{FORCESOLUTION}
 file, or for different stations by changing the \texttt{STATIONS}
 file. There is no need to rerun the \texttt{xgenerate\_databases}
-executable. Of course it is best to include as many stations as
-possible, since this does not add to the cost of the simulation.
-
-\chapter{\label{cha:Fault-Kinematics-Dynamics}Fault Kinematics and Dynamics}
-
-Fault is a discontinuity in the continuum that needs special treatment
-with two nodes at the same locations. These are called split nodes
-that allow to capture deformation on the fault plane termed slip.
-Fault motion determines the deformation of the rest of the bulk. This
-fault motion can be of two types either physics based spontaneous
-rupture propagation or a prescribed rupture description. The former
-is known as fault dynamics while the later is fault kinematics. We
-describe here first the common features to both fault dynamics and
-kinematics and then give detailed description of parameters required
-for each of them separately.
-
-Input files : Two additional input files are required in \texttt{DATA}
-directory for faults with split nodes\\
-
-\texttt{\textbf{Par\_file\_faults}}\texttt{ - }specialized parameters
-required by fault modules (\emph{if this file doesn't exist, code
-assumes that there are no faults})
-
-\texttt{\textbf{FAULT\_STATIONS}}\texttt{ - }stations on the fault
-where timeseries output of slip, sliprate and traction is required
-\texttt{}~\\
-
-Currently faults can only be run with \texttt{xdecompose\_mesh }and\texttt{
-CUBIT. xmeshfem3D }isn't yet ready to handle faults.\\
-
-Important module related to fault :
-
-Decomposing mesh stage : $\mathtt{src/decompose\_mesh\_SCOTCH/fault\_scotch.f90}$
-
-Generating databases stage : $\mathtt{src/generate\_databases/fault\_generate\_databases.f90}$
-
-Solver stage : $\mathtt{src/specfem3D/fault\_solver\_common.f90}$
-
-{\small \qquad{}}$\mathtt{src/specfem3D/fault\_solver\_kinematic.f90}$
-
-{\small \qquad{}}$\mathtt{src/specfem3D/fault\_solver\_dynamic.f90}$\\
-
-The features and the format of inputs and outputs of fault routines
-are subject to change.\\
-
-\section{Mesh Generation with Split Nodes}
+executable. Of course it is best to include as many stations as possible,
+since this does not add to the cost of the simulation.
+
+\chapter{\label{cha:Fault-Kinematics-Dynamics}Kinematic and dynamic fault
+sources}
+
+SPECFEM3D can handle finite fault sources of two kinds: 
+\begin{enumerate}
+\item \emph{Kinematic}: the spatio-temporal distribution of slip rate is
+prescribed along the fault surface
+\item \emph{Dynamic}: a friction law and initial stresses are prescribed
+on the fault, a spontaneous rupture process is computed.
+\end{enumerate}
+
+\section{\label{sec:Mesh-Generation-with}Mesh Generation with Split Nodes}
 
@@
 This approach is known as \textquotedbl{}split nodes\textquotedbl{}.
-To facilitate the mesh generation with split nodes in Cubit, we need
+Currently faults can only be run with \texttt{xdecompose\_mesh }and\texttt{
+CUBIT. xmeshfem3D }is not yet ready to handle faults.\\
+To facilitate the mesh generation with split nodes in CUBIT, we need
 to separate the two fault surfaces by a small distance, effectively
-creating a tiny opening of the fault. Note that only the interior
+creating a tiny opening of the fault (Figure \ref{fig:examples.splitnodes},
+\ref{fig:examples.splitnodes-surfacetrace}). Note that only the interior
 of the fault must be opened, its edges must remain closed (except
@@
 
-Here is an example Cubit script to generate a mesh with split nodes
-for a buried vertical strike-slip fault:\\
-reset \\
-brick x 10 y 10 z 10 \\
-webcut volume all with plane xplane \\
-webcut volume all with plane yplane \\
-webcut volume all with plane xplane offset 3 \\
-webcut volume all with plane zplane offset 3 \\
-webcut volume all with plane zplane offset -3 \\
-imprint all \\
-merge all \\
-unmerge surf 160 \\
-mesh vol all \\
-set node constraint off \\
-node in surf 168 move X 0 Y 0.01 Z 0 \\
-node in surf 160 move X 0 Y -0.01 Z 0\\
-
+Here is an example CUBIT script to generate a mesh with split nodes
+for a buried vertical strike-slip fault:
+\begin{lyxcode}
+~reset~~\\
+~brick~x~10~y~10~z~10~~\\
+~webcut~volume~all~with~plane~xplane~~\\
+~webcut~volume~all~with~plane~yplane~~\\
+~webcut~volume~all~with~plane~xplane~offset~3~~\\
+~webcut~volume~all~with~plane~zplane~offset~3~~\\
+~webcut~volume~all~with~plane~zplane~offset~-3~~\\
+~imprint~all~~\\
+~merge~all~~\\
+~unmerge~surf~160~~\\
+~mesh~vol~all~~\\
+~set~node~constraint~off~~\\
+~node~in~surf~168~move~X~0~Y~0.01~Z~0~~\\
+~node~in~surf~160~move~X~0~Y~-0.01~Z~0~\\
+
+\end{lyxcode}
 \begin{figure}[htbp]
 \noindent \begin{centering}
-\includegraphics[scale=0.3]{figures/faultmesh}
-\\
-\includegraphics[scale=0.3]{figures/surf168}
-\includegraphics[scale=0.3]{figures/splitnodes}
+\includegraphics[scale=0.3]{figures/faultmesh} \\
+ \includegraphics[scale=0.3]{figures/surf168} \includegraphics[scale=0.3]{figures/splitnodes} 
 \par\end{centering}
-\caption{Screenshots of the CUBIT example for embedded fault with split nodes:
-entire mesh decomposed into several volumes bounding the fault (top),
-fault surface 168 is selected (bottom-left) and split nodes of the
-fault illustrating only interior nodes of fault are split while those
-on the edges of fault surface are touching each other (bottom-right).}
+
+\caption{Screenshots of the CUBIT example for the embedded fault with split
+nodes: The entire mesh is decomposed into several volumes bounding
+the fault (top), the fault surface 168 is shown as orange squares
+(middle) and split nodes of the fault shown as orange dots(bottom).
+Note that only interior nodes of the fault are split while those on
+the edges of the fault surface are touching each other.}
+
 \label{fig:examples.splitnodes} 
@@
 
-The Cubit scripts ({*}.jou and {*}.py) in the directory $\mathtt{EXAMPLES}$
+\begin{lyxcode}
+
+\end{lyxcode}
+\begin{figure}[htbp]
+\noindent \begin{centering}
+\includegraphics[scale=0.55]{figures/splitnodes_surfacetrace} \\
+
+\par\end{centering}
+
+\caption{Screenshots of a CUBIT example showing split nodes for a fault reaching
+the surface. Surface trace of the fault is shown in orange. Note that
+the edges of the fault are not split while the interior nodes are
+offset by a small distance on either side of the fault}
+
+\label{fig:examples.splitnodes-surfacetrace} 
+\end{figure}
+
+The CUBIT scripts ({*}.jou and {*}.py) in the directory $\mathtt{examples}$
 generate more complicated meshes. The {*}.py files are Python scripts
-that execute Cubit commands and use the Cubit-python interface for
+that execute CUBIT commands and use the CUBIT-python interface for
 SPECFEM3D (see next section). The Python language allows to define
 and manipulate variables to parameterize the mesh. Alternatively,
-the Python script can call a Cubit journal file ({*}.jou), which looks
+the Python script can call a CUBIT journal file ({*}.jou), which looks
 like the example above. Variables can be defined and manipulated there
-using the $\mathtt{APREPRO}$ language built in Cubit.\\
-
-\texttt{Note}: you should avoid gaps in the list of indices of mesh
-objects with the following Cubit command:
-
-compress ids hex face edge node')
-
-(otherwise you will get a segmentation fault during domain decomposition)\\
-
-\section{Cubit-Python Scripts for Faults}
-
-The mesh generated in Cubit needs to be processed and exported in
+using the $\mathtt{APREPRO}$ language built in CUBIT.\\
+
+Note that you should avoid gaps in the list of indices of mesh objects
+with the following CUBIT command: 
+\begin{lyxcode}
+$\mathtt{compress\: ids\: hex\: face\: edge\: node}$
+\end{lyxcode}
+(otherwise you will get a segmentation fault during domain decomposition)
+
+\section{CUBIT-Python Scripts for Faults}
+
+The mesh generated in CUBIT needs to be processed and exported in
 a format compatible with SPECFEM3D. This is achieved in the Python
-scripts by calling the Python-Cubit interface functions defined in
+scripts by calling the Python-CUBIT interface functions defined in
 the CUBIT directory: 
 \begin{enumerate}
-\item Function \textquotedbl{}$\mathtt{define\_bc}$\textquotedbl{} (or
-similar ones) must be called to set up the absorbing boundaries database
-\item Function \textquotedbl{}$\mathtt{fault\_input}$\textquotedbl{} must
-be called once for each fault to set up the fault database
-\item Function \textquotedbl{}$\mathtt{cubit2specfem3d.export2SESAME}$\textquotedbl{}
-must be called at the very end of the script to export the mesh in
-a SPECFEM3D format.
+\item Function $\mathtt{define\_bc}$ (or $\mathtt{boundary\_definition.py}$)
+must be called to set up the absorbing boundaries database 
+\item Function $\mathtt{fault\_input}$ must be called once for each fault
+to set up the fault database 
+\item Function $\mathtt{cubit2specfem3d.export2SESAME}$ must be called
+at the very end of the script to export the mesh in a SPECFEM3D format. 
 \end{enumerate}
-The functions in \#1 and \#3 are for bulk and are not documented here.
-We focus here on \#2:\\
-
-\textbf{Function}: $\mathtt{fault\_input}$
-
-\textbf{Purpose}: export a fault mesh database from Cubit to a SPECFEM3D-compliant
+The functions in \#1 and \#3 are for the bulk and are documented in
+Section \ref{subsec:Exporting-the-Mesh}. We focus here on \#2:
+\begin{description}
+\item [{Function:}] $\mathtt{fault\_input}$
+\item [{Purpose:}] export a fault mesh database from CUBIT to a SPECFEM3D-compliant
 file
-
-\textbf{Syntax}: fault\_input(id\_fault, ids\_surf\_1, ids\_surf\_2)\\
-
-\textbf{Inputs}: id\_fault integer index assigned to the fault.
-
-(The user must number all the faults, starting at 1, with unit increments)
-\begin{lyxlist}{-}
-\item [{ids\_surf\_1}] list of Cubit ids of all surfaces that form side
-1 of the fault
-\item [{ids\_surf\_2}] list of Cubit ids of all surfaces that form side
-2 of the fault
-\end{lyxlist}
-(The user must decide which side of the fault is side 1. This choice
-affects the sign conventions of fault quantities as explained in a
-later section)\\
-
-\textbf{Outputs}: file \textquotedbl{}fault\_file\_X.dat\textquotedbl{},
-where X is the fault id (id\_fault).\\
-
-\textbf{Example}: For the example in the previous section:\\
-
-$\hphantom{}$A1 = {[}168{]}
-
-$\hphantom{}$A2 = {[}160{]}
-
-$\hphantom{}$faultA = fault\_input(1,A1,A2)\\
-
-\section{Running a Simulation }
-\begin{enumerate}
-\item Copy the input files to the work directory:
-
->\textcompwordmark{}> cd \textasciitilde{}/$\mathtt{SPECFEM3D}$$\mathtt{}$
-
->\textcompwordmark{}> cp -r $\mathtt{EXAMPLES/splay\_faults/DATA}$
-./
-
-\item Create a mesh with CUBIT: (If you have not installed CUBIT yet skip
-this step and use the mesh files included with the examples.)
-
-See the previous section \textquotedbl{}Mesh Generation with Split
-Nodes\textquotedbl{} or move to the directory CUBIT, open the cubit
-GUI, run a CUBIT script: >\textcompwordmark{}> cd \textasciitilde{}$\mathtt{/SPECFEM3D/CUBIT}$
->\textcompwordmark{}> cubit In CUBIT's menu \textquotedbl{}Tools\textquotedbl{},
-select \textquotedbl{}Play Journal File\textquotedbl{} and select
-a script file, for instance: $\mathtt{EXAMPLES/splay\_faults/splay\_faults.py}$
-
-This creates several mesh files in directory $\mathtt{CUBIT/MESH/}$:
-absorbing\_surface\_file\_{*} (5 files) free\_or\_absorbing\_surface\_file\_zmax materials\_file
-mesh\_file nodes\_coords\_file nummaterial\_velocity\_file
-
-The CUBIT graphics window should show the mesh (e.g. $\mathtt{EXAMPLES/splay_{f}aults/splay_{f}aults.jpg}$)
-You can zoom, pan or rotate using mouse gestures. To look at the block
-of your interest, set \textquotedbl{}visibility off\textquotedbl{}
-in the other blocks.
-
-To run CUBIT from different working directories it is convenient to
-include the path to $\mathtt{SPECFEM3D/CUBIT}$ in an environment
-variable $\mathtt{PYTHONPATH}$
-
-\item Partition the mesh with the domain decomposition software SCOTCH.
-
->\textcompwordmark{}> cd \textasciitilde{}/$\mathtt{SPECFEM3D/decompose\_mesh\_SCOTH}$
-
->\textcompwordmark{}> ./$\mathtt{decompose\_mesh\_SCOTCH}$ 'nproc'
-../$\mathtt{CUBIT/MESH/}$ ../$\mathtt{DATABASES\_MPI}$ /
-
-where 'nproc' is the number of processors that you will use in the
-simulation. The second argument (../$\mathtt{CUBIT/MESH/}$ in this
-example) is the path to the directory containing the mesh files that
-were generated by CUBIT. This creates one mesh file proc000{*}{*}{*}\_Database
-per processor in directory $\mathtt{DATABASES\_MPI}$ . These files
-can be large, so you might want to place $\mathtt{DATABASES\_MPI}$
-in a scratch disk and specify the path accordingly when executing
-$x\mathtt{decompose\_mesh\_SCOTCH}$.
-
-\item Edit the file $\mathtt{DATA/Par_{f}ile}$: (copy an example here,
-see III.2)
-
-$\mathtt{LOCAL\_PATH}$= should be the path to $\mathtt{DATABASES\_MPI}$
-
-$\mathtt{NPROC}$= number of processors. The same as the number of
-partitions in SCOTCH (step 3).
-
-\item Generate databases:
-
->\textcompwordmark{}> cd \textasciitilde{}/$\mathtt{SPECFEM3D}$
-
->\textcompwordmark{}> mpirun -np nproc ./$\mathtt{xgenerate_{d}atabases}$
-
-or
-
->\textcompwordmark{}> $\mathtt{./run/run.xdatabases}$
-
-or
-
->\textcompwordmark{}> qsub -l nodes=... -l walltime=... go\_mesher
-(the script go\_mesher is in $\mathtt{utils/}$)
-
-This creates several binary mesh files for each processor (proc000{*}{*}{*}.bin)
-in directory $\mathtt{DATABASES\_MPI}$.
-
-\item If this is the first time you run a simulation with this mesh, set
-the parameter DT (the time step) DATA/Par\_file to a value smaller
-than the \textquotedbl{}Maximum suggested time step\textquotedbl{}
-in $\mathtt{OUTPUT\_FILES/output\_mesher.txt}$. Consider the note
-in section IV about the effect of Kelvin-Voigt damping on the critical
-timestep. Edit the variables $\mathtt{NSTEP}$ and $\mathtt{NTSTEP\_*}$
-in $\mathtt{DATA/Par\_file}$ accordingly.
-\item Run the solver:
-
->\textcompwordmark{}> mpirun -np nproc ./$\mathtt{xspecfem3D}$
-
-or
-
->\textcompwordmark{}> $\mathtt{./run/run.xspecfem3d}$
-
-or
-
->\textcompwordmark{}> qsub -l nodes=... -l walltime=... go\_solver
-(the script go\_solver is in $\mathtt{utils/}$)
-
-\end{enumerate}
+\item [{Syntax:}] fault\_input(id\_fault, ids\_surf\_1, ids\_surf\_2)
+\item [{Inputs:}]~\end{description}
+\begin{itemize}
+\item id\_fault integer index assigned to the fault. (The user must number
+all the faults, starting at 1, with unit increments) 
+\item ids\_surf\_1 list of CUBIT ids of all surfaces that form side 1 of
+the fault. 
+\item ids\_surf\_2 list of CUBIT ids of all surfaces that form side 2 of
+the fault. (The user must decide which side of the fault is side 1.
+This choice affects the sign conventions of fault quantities as explained
+in Section \ref{sec:Sign-Convention-for}).\end{itemize}
+\begin{description}
+\item [{Outputs:}] file $\mathtt{fault\_file\_X.dat}$, where X is the
+fault id (id\_fault).
+\item [{Example:}] For the example in Section \ref{sec:Mesh-Generation-with}:\end{description}
+\begin{lyxcode}
+$\hphantom{}$A1~=~{[}168{]}
+
+$\hphantom{}$A2~=~{[}160{]}
+
+$\hphantom{}$fault\_input(1,A1,A2)
+\end{lyxcode}
 
@@
 \item TPV16, heterogenous initial stress 
+\end{itemize}
+and
+\begin{itemize}
 \item Splay fault models from Wendt et al. (2009)\\
@@
 \end{itemize}
-To run the examples:
-\begin{enumerate}
-\item Read the documents in the directory $\mathtt{EXAMPLES/*/description}$.
+Read the documents in the directory $\mathtt{examples/*/description}$.
 They contain a description of the example and additional instructions
-to run it.
-\item Replace the contents of directory $\mathtt{DATA/*}$ by one of the
-$\mathtt{EXAMPLES/tpv*/DATA}$, for instance:
-
->\textcompwordmark{}> cp -r $\mathtt{EXAMPLES/tpv5/DATA}$ ./
-
-\item Follow all the steps in section II above.
-\item Visualize the results with the matlab scripts in the directory $\mathtt{EXAMPLES/*/post}$
-\end{enumerate}
-
-\section{Sign Convention for Fault Quantities}
+to run it. Visualize the results with the Matlab scripts in the directory
+$\mathtt{examples/*/post}$
+
+\section{\label{sec:Sign-Convention-for}Sign Convention for Fault Quantities}
 
@@
 The local coordinate system on the fault is defined as the right-handed
-coordinate system defined by (strike, dip, normal), where \textquotedbl{}normal\textquotedbl{}
+coordinate system defined by (strike, dip, normal), where \textquotedbl{}\emph{normal}\textquotedbl{}
 is the normal vector outgoing from side 1, \textquotedbl{}\emph{dip}\textquotedbl{}
@@
 
-\textbf{DATA/Par\_file}
-\begin{lyxlist}{-}
-\item [{\textbf{DATA/STATIONS}}] List of stations outside the fault
+Three additional input files are required in the \texttt{DATA} directory
+for dynamic and kinematic faults. They are $\mathtt{Par\_file\_faults}$,
+$\mathtt{FAULT\_STATIONS}$ and $\mathtt{input\_file.txt}$ (optional).
+If the former does not exist, the code assumes that there are no faults.\\
+
+\begin{lyxlist}{00.00.0000}
 \item [{\textbf{DATA/Par\_file\_faults}}] contains parameters of the fault.
@@
 
-\end{lyxlist}
 \&\textbf{BEGIN\_FAULT} /
@@
 
-\begin{lyxlist}{-}
+\begin{lyxlist}{00.00.0000}
 \item [{\&\textbf{INIT\_STRESS}}] input block sets the initial fault stresses
@@
 
-S1 = constant background value of along-strike shear stress (positive
+\begin{description}
+\item [{S1}] = constant background value of along-strike shear stress (positive
 in the usual strike direction)
-
-S2 = constant background value of along-dip shear (positive is down-dip,
-normal faulting)
-
-S3 = constant background value of normal stress (negative in compresion)
-
-n1 = number of heterogeneous items for along-strike shear stress {[}default
+\item [{S2}] = constant background value of along-dip shear (positive is
+down-dip, normal faulting)
+\item [{S3}] = constant background value of normal stress (negative in
+compression)
+\item [{n1}] = number of heterogeneous items for along-strike shear stress
+{[}default is 0{]}
+\item [{n2}] = number of heterogeneous items for along-dip shear stress
+{[}default is 0{]}
+\item [{n3}] = number of heterogeneous items for normal stress {[}default
 is 0{]}
-
-n2 = number of heterogeneous items for along-dip shear stress {[}default
-is 0{]}
-
-n3 = number of heterogeneous items for normal stress {[}default is
-0{]}\\
-
+\end{description}
 \item [{\&\textbf{SWF}}] input block sets the slip-weakening friction parameters
@@
 
-mus = constant background value of static friction coefficient
-
-mud = constant background value of dynamic friction coefficient
-
-dc = constant background value of critical slip-weakening distance
-
-nmus = number of heterogeneous items for static friction coefficient
+\begin{description}
+\item [{mus}] = constant background value of static friction coefficient
+\item [{mud}] = constant background value of dynamic friction coefficient
+\item [{dc}] = constant background value of critical slip-weakening distance
+\item [{nmus}] = number of heterogeneous items for static friction coefficient
 {[}default is 0{]}
-
-nmud = number of heterogeneous items for dynamic friction coefficient
+\item [{nmud}] = number of heterogeneous items for dynamic friction coefficient
 {[}default is 0{]}
-
-ndc = number of heterogeneous items for critical slip-weakening distance
-{[}default is 0{]}\\
-
+\item [{ndc}] = number of heterogeneous items for critical slip-weakening
+distance {[}default is 0{]}\\
+
+\end{description}
 \item [{\&\textbf{DIST2D}}] input blocks modify (overwrite) the value of
@@
 
+\end{lyxlist}
+\end{lyxlist}
+
+\begin{lyxlist}{00.00.0000}
 \item [{\textbf{DATA/FAULT\_STATIONS}}] Stations in the fault plane.
 
-Line 1: number of stations.
-
-Line 2 to end: 5 columns: X, Y, Z (-depth), station name, fault-id
+\textbf{Line 1} : number of stations.
+
+\textbf{Line 2 to end}: 5 columns: X, Y, Z (-depth), station name,
+fault-id
 
@@
 
+\end{lyxlist}
+
+\begin{lyxlist}{00.00.0000}
 \item [{\textbf{DATA/input\_file.txt}}] Heterogeneous stresses and friction
-parameters Documented in page 10 of EXAMPLES/tpv16.crack/description/TPV16\_17\_Description\_v03.pdf
-
-To activate this feature, in fault\_solver.f90 set TPV16 = .true.
-then re-compile the code: cd SPECFEM3D make
-
-\item [{Heterogeneous}] velocity models can be given in a regular grid.
-
-In the mesh generation Python file (.py) or in the CUBIT journal file
-(.jou), set the material 1st attribute to -1:
-
-\qquad{}block {*} attribute index 1 -1
-
-In module model\_tomography.f90:
-
-\qquad{}Set the parameter TOMO\_FILENAME to the name of the file
-containing the velocity model.
-
-Re-compile the code:
-
-\qquad{}cd SPECFEM3D
-
-\qquad{}make \\
-
-The format of the velocity file is:
-\begin{description}
-\item [{Line}] 1: ORIG\_X, ORIG\_Y, ORIG\_Z, END\_X, END\_Y, END\_Z (coordinates
-of the two extreme corners of the box)
-\item [{Line}] 2: SPACING\_X, SPACING\_Y, SPACING\_Z (regular grid spacing
-in each direction)
-\item [{Line}] 3: NX, NY, NZ (number of grid point in each direction)
-\item [{Line}] 4: VP\_MIN, VP\_MAX, VS\_MIN, VS\_MAX, RHO\_MIN, RHO\_MAX
-(min and max values of P wave speed, S wave speed and density)
-\item [{Line}] 5 to 4+NX{*}NY{*}NZ: x\_tomo,y\_tomo,z\_tomo,vp\_tomo,vs\_tomo,rho\_tomo
-(position, P wave speed, S wave speed, density)\\
-
-\end{description}
+parameters are documented in page 10 of
+
+\begin{lyxlist}{00.00.0000}
+\item [{$\mathtt{examples/tpv16/description/TPV16\_17\_Description\_v03.pdf}$}]~
 \end{lyxlist}
-Several files are generated automatically by xgenerate\_databases
-in directory DATABASES\_MPI and do not need to be modified by the
-user.
-
-\section{How to set the Kelvin-Voigt Damping Parameter }
+
+To activate this feature, in fault\_solver\_dynamic.f90 set \textbf{TPV16}
+= .true. then re-compile the code.
+
+\end{lyxlist}
+
+\section{\label{sec:Setting-the-Kelvin-Voigt}Setting the Kelvin-Voigt Damping
+Parameter }
 
@@
 is to damp spurious oscillations generated by the fault slip at frequencies
-that are too high to be resolved by the mesh. The viscosity \textquotedbl{}\emph{eta}\textquotedbl{}
+that are too high to be resolved by the mesh. The viscosity $\mathtt{eta}$
 (in seconds) depends on the size of the elements on the fault. Here
@@
 \item Determine the average linear size of the elements on the fault plane,
-\textquotedbl{}\emph{h\_fault}\textquotedbl{}. Usually this value
-is prescribed by the user during mesh generation. Otherwise it can
-be found by inspection of the mesh inside the Cubit GUI.
-\item Use the matlab function $\mathtt{utils/critical_{t}imestep.m}$ to
-compute dtc\_fault = critical\_timestep(cp,h\_fault,ngll) This is
-the critical time step in an elastic medium for a hypothetical element
-of cubic shape with size equal to h\_fault.
-\item Set eta in Par\_file\_faults to (0.1 to 0.3){*}dtc\_fault. A larger
-eta damps high-frequencies more aggresively but it might also affect
-lower frequencies and rupture speed.
-
+$\mathtt{h\_fault}$. Usually this value is prescribed by the user
+during mesh generation. Otherwise it can be found by inspection of
+the mesh inside the CUBIT GUI. 
+\item Use the Matlab function $\mathtt{utils/critical\_timestep.m}$ to
+compute\\
+ $\mathtt{dtc\_fault}=\mathtt{critical\_timestep\left(c_{p},h\_fault,ngll\right)}$.
+\\
+This is the critical time step in an elastic medium for a hypothetical
+element of cubic shape with size equal to $\mathtt{h\_fault}$. 
+\item Set $\mathtt{eta}$ in $\mathtt{Par\_file\_faults}$ to (0.1 to 0.3)$\ensuremath{\times}\mathtt{dtc\_fault}$.
+A larger $\mathtt{eta}$ damps high-frequencies more aggressively
+but it might also affect lower frequencies and rupture speed. 
+\end{enumerate}
 Viscosity reduces numerical stability: the critical timestep in a
@@
 in a purely elastic simulation. Here is how to set the time step accordingly:
-
-\item Run a test simulation without viscosity (eta=0 and only a few time
-steps)
+\begin{enumerate}
+\item Run a test simulation without viscosity ($\mathtt{eta}$=0 and only
+a few time steps) 
 \item Look for the \textquotedbl{}maximum suggested time step\textquotedbl{}
-in $\mathtt{OUTPUT_{F}ILES/output_{m}esher.txt}$ This is the critical
-timestep of a purely elastic simulation, \textquotedbl{}dtc\_bulk\textquotedbl{}.
+in $\mathtt{OUTPUT\_FILES/output\_mesher.txt}$. This is the critical
+timestep of a purely elastic simulation, $\mathtt{dtc\_bulk}$. 
 \item Reset the timestep of the simulation with a Kelvin-Voigt material
-to a value smaller than dtc\_kv = eta{*}( sqrt(1+dtc\_bulk\textasciicircum{}2/eta\textasciicircum{}2)-1
-)
+to a value smaller than\\
+ $\mathtt{dtc\_kv}=\mathtt{eta}\left(\sqrt{1+\mathtt{dtc\_bulk^{2}}/\mathtt{eta^{2}}}-1\right)$ 
 \end{enumerate}
-Note that in general dtc\_bulk is smaller than dtc\_fault, because
-elements off the fault might be smaller or more distorted than element
-faces on the fault.
+Note that in general $\mathtt{dtc\_bulk}$ is smaller than $\mathtt{dtc\_fault}$,
+because elements off the fault might be smaller or more distorted
+than element faces on the fault.
 
@@
 
-Several output files are saved in \textasciitilde{}/$\mathtt{SPECFEM3D/OUTPUT_{F}ILES}$:\\
+Several output files are saved in $\mathtt{OUTPUT\_FILES}$:\\
 
 \begin{enumerate}
-\item Seismograms for each station on the fault plane given in $\mathtt{DATA/FAULT_{S}TATIONS}$.
+\item Seismograms for each station on the fault plane given in $\mathtt{DATA/FAUL\_STATIONS}$.
 One output file is generated for each station, named after the station.
@@
 
-\# Column \#1 = Time (s)
-
-\# Column \#2 = horizontal right-lateral slip (m)
-
-\# Column \#3 = horizontal right-lateral slip rate (m/s)
-
-\# Column \#4 = horizontal right-lateral shear stress (MPa)
-
-\# Column \#5 = vertical up-dip slip (m)
-
-\# Column \#6 = vertical up-dip slip rate (m/s)
-
-\# Column \#7 = vertical up-dip shear stress (MPa)
-
-\# Column \#8 = normal stress (MPa)\\
+Column 1 = Time (s)
+
+Column 2 = horizontal right-lateral slip (m)
+
+Column 3 = horizontal right-lateral slip rate (m/s)
+
+Column 4 = horizontal right-lateral shear stress (MPa)
+
+Column 5 = vertical up-dip slip (m)
+
+Column 6 = vertical up-dip slip rate (m/s)
+
+Column 7 = vertical up-dip shear stress (MPa)
+
+Column 8 = normal stress (MPa)\\
 
@@
 in $\mathtt{DATA/STATIONS}$. 
-\item Rupture time files are named Rupture\_time\_FAULT-id. One file is
-generated for each fault. The files are ascii and start with a header
-(12 lines long) followed by a data block with the following format,
-one line per fault node:
-
-\# Column \#1 = horizontal coordinate, distance along strike (m)
-
-\# Column \#2 = vertical coordinate, distance down-dip (m)
-
-\# Column \#3 = rupture time (s)
+\item Rupture time files are named $\mathtt{Rupture\_time\_FAULT}$-$\mathtt{id}$.
+One file is generated for each fault. The files are ascii and start
+with a header (12 lines long) followed by a data block with the following
+format, one line per fault node:
+
+Column 1 = horizontal coordinate, distance along strike (m)
+
+Column 2 = vertical coordinate, distance down-dip (m)
+
+Column 3 = rupture time (s)
 
 \item Fault quantities (slip, slip rate, stresses, etc) at regular times
-are stored in binary data files called Snapshot\#it\#.bin, where \#it\#
-is the timestep number. These can be read in Matlab with the function
-$\mathtt{Post-processing/FSEM3D\_snapshot.m}$
+are stored in binary data files called $\mathtt{Snapshot\#it\#.bin}$,
+where \#it\# is the timestep number. These can be read in Matlab with
+the function $\mathtt{utils/FSEM3D\_snapshot.m}$ 
 \end{enumerate}
@@
 Some Matlab functions for post-processing and visualization are included
-in directory Post-processing. The functions are internally documented
+in directory $\mathtt{utils}$. The functions are internally documented
 (see their matlab help).\\
@@
 
-The directories$\mathtt{EXAMPLES/*/post}$ contain additional Matlab
+The directories $\mathtt{examples/*/post}$ contain additional Matlab
 scripts to generate figures specific to each example.
@@
 
-\section{\label{sec:Adjoint-simulation-sources}Adjoint Simulations for Sources Only (not for the Model)}
+\section{\label{sec:Adjoint-simulation-sources}Adjoint Simulations for Sources
+Only (not for the Model)}
 
@@
 from the adjoint seismograms recorded at the original earthquake location.
-
 \begin{enumerate}
@@
 1} and \texttt{SAVE\_FORWARD = .false.}). You can automatically set
-these two variables using the \texttt{\small utils/change\_simulation\_type.pl}
-script:
+these two variables using the \texttt{\small utils/change\_simulation\_type.pl}{\small{}
+script:}{\small \par}
 
@@
 \end{lyxcode}
+
 and then collect the recorded seismograms at all the stations given
@@
 adjoint sources need to be computed from the original recorded seismograms
-for the selected stations and saved in the \texttt{OUTPUT\_FILES/SEM/} directory
-with the format \texttt{STA.NT.BX?.adj}, where \texttt{STA}, \texttt{NT}
-are the station name and network code given in the \texttt{DATA/STATIONS\_ADJOINT}
+for the selected stations and saved in the \texttt{OUTPUT\_FILES/SEM/}
+directory with the format \texttt{STA.NT.BX?.adj}, where \texttt{STA},
+\texttt{NT} are the station name and network code given in the \texttt{DATA/STATIONS\_ADJOINT}
 file, and \texttt{BX?} represents the component name of a particular
-adjoint seismogram. Please note that the band code can change depending on
-your sampling rate (see Appendix~\ref{cha:channel-codes} for further details).
+adjoint seismogram. Please note that the band code can change depending
+on your sampling rate (see Appendix~\ref{cha:channel-codes} for
+further details). 
 \item The adjoint seismograms are in the same format as the original seismogram
@@
 seismograms from \texttt{LOCAL\_PATH}.
-
 \begin{itemize}
@@
 Finite-frequency sensitivity kernels are computed in two successive
-simulations (please refer to \citet{LiTr06} and \citet{TrKoLi08} for details).
-
+simulations (please refer to \citet{LiTr06} and \citet{TrKoLi08}
+for details).
 \begin{enumerate}
@@
 
-Prepare the \texttt{\small CMTSOLUTION} and \texttt{\small STATIONS}
-files, set the parameters \texttt{\small SIMULATION\_TYPE}{\small{}
-}\texttt{\small =}{\small{} }\texttt{\small 1} and \texttt{\small SAVE\_FORWARD
-=}{\small{} }\texttt{\small .true.} in the \texttt{Par\_file} (\texttt{change\_simulation\_type
--F}), and run the solver.
-
+Prepare the \texttt{\small CMTSOLUTION}{\small{} and }\texttt{\small STATIONS}{\small{}
+files, set the parameters }\texttt{\small SIMULATION\_TYPE}{\small {}
+}\texttt{\small =}{\small {} }\texttt{\small 1}{\small{} and }\texttt{\small SAVE\_FORWARD
+=}{\small {} }\texttt{\small .true.}{\small{} in the }\texttt{\small Par\_file}{\small{}
+(}\texttt{\small change\_simulation\_type -F}{\small ), and run the
+solver.}{\small \par}
 \begin{itemize}
@@
 with an adjoint wavefield with the proper frequency content. 
-\item This forward simulation differs from the regular simulations (\texttt{\small SIMULATION\_TYPE}{\small{}
-}\texttt{\small =}{\small{} }\texttt{\small 1} and \texttt{\small SAVE\_FORWARD}{\small{}
-}\texttt{\small =}{\small{} }\texttt{\small .false.}) described in
-the previous chapters in that the state variables for the last time
-step of the simulation, including wavefields of the displacement,
-velocity, acceleration, etc., are saved to the \texttt{LOCAL\_PATH}
-to be used for the subsequent simulation.
+\item This forward simulation differs from the regular simulations (\texttt{\small SIMULATION\_TYPE}{\small {}
+}\texttt{\small =}{\small {} }\texttt{\small 1}{\small{} and }\texttt{\small SAVE\_FORWARD}{\small {}
+}\texttt{\small =}{\small {} }\texttt{\small .false.}{\small ) described
+in the previous chapters in that the state variables for the last
+time step of the simulation, including wavefields of the displacement,
+velocity, acceleration, etc., are saved to the }\texttt{\small LOCAL\_PATH}{\small{}
+to be used for the subsequent simulation. }{\small \par}
 \item For regional simulations, the files recording the absorbing boundary
@@
 in the Section \ref{enu:Prepare-the-adjoint}.
-
 \begin{itemize}
@@
 in the \texttt{STATIONS\_ADJOINT} list, we supply a sample program
-in \texttt{utils/adjoint\_sources/traveltime/xcreate\_adjsrc\_traveltime} to cut a certain portion of the original
-displacement seismograms and convert it into the proper adjoint source
-to compute the finite-frequency kernel.
-\begin{lyxcode}
-xcreate\_adjsrc\_traveltime~t1~t2~ifile{[}0-5]~E/N/Z-ascii-files~{[}baz]
-\end{lyxcode}
+in \texttt{utils/adjoint\_sources/traveltime/xcreate\_adjsrc\_traveltime}
+to cut a certain portion of the original displacement seismograms
+and convert it into the proper adjoint source to compute the finite-frequency
+kernel.
+
+\begin{lyxcode}
+xcreate\_adjsrc\_traveltime~t1~t2~ifile{[}0-5{]}~E/N/Z-ascii-files~{[}baz{]}
+\end{lyxcode}
+
 where \texttt{t1} and \texttt{t2} are the start and end time of the
@@
 
-\item Similarly, a sample program to compute adjoint sources for amplitude finite-frequency kernels
-may be found in \texttt{utils/adjoint\_sources/amplitude} and used in the same way as described for
-traveltime measurements
-\begin{lyxcode}
-xcreate\_adjsrc\_amplitude~t1~t2~ifile{[}0-5]~E/N/Z-ascii-files~{[}baz].
-\end{lyxcode}
-
+\item Similarly, a sample program to compute adjoint sources for amplitude
+finite-frequency kernels may be found in \texttt{utils/adjoint\_sources/amplitude}
+and used in the same way as described for traveltime measurements
+
+\begin{lyxcode}
+xcreate\_adjsrc\_amplitude~t1~t2~ifile{[}0-5{]}~E/N/Z-ascii-files~{[}baz{]}.
+\end{lyxcode}
 \end{itemize}
-
 \item \textbf{Run the kernel simulation}
@@
 With the successful forward simulation and the adjoint source ready
-in the \texttt{OUTPUT\_FILES/SEM/} directory, set \texttt{SIMULATION\_TYPE = 3} and \texttt{SAVE\_FORWARD
-= .false.} in the \texttt{Par\_file} (you can use \texttt{change\_simulation\_type.pl -b}),
-and rerun the solver.
-
+in the \texttt{OUTPUT\_FILES/SEM/} directory, set \texttt{SIMULATION\_TYPE
+= 3} and \texttt{SAVE\_FORWARD = .false.} in the \texttt{Par\_file}
+(you can use \texttt{change\_simulation\_type.pl -b}), and rerun the
+solver.
 \begin{itemize}
@@
 \end{itemize}
-
 \item \textbf{Run the anisotropic kernel simulation}
@@
 compute the kernels for the 21 independent components $C_{IJ},\, I,J=1,...,6$
-(using Voigt's notation) of the elastic tensor in the cartesian coordinate system. 
-This is done by setting \texttt{ANISOTROPIC\_KL} \texttt{=} \texttt{.true.} 
-in \texttt{constants.h} before compiling the package. The definition of the parameters 
-$C_{IJ}$ in terms of the corresponding components $c_{ijkl},ijkl,i,j,k,l=1,2,3$ of 
-the elastic tensor in cartesian coordinates follows \citet{ChTr07}. The 21 anisotropic 
-kernels are saved in the \texttt{LOCAL\_PATH} in one file with the name of 
-\texttt{proc??????\_cijkl\_kernel.bin} (with \texttt{proc??????} 
+(using Voigt's notation) of the elastic tensor in the cartesian coordinate
+system. This is done by setting \texttt{ANISOTROPIC\_KL} \texttt{=}
+\texttt{.true.} in \texttt{constants.h} before compiling the package.
+The definition of the parameters $C_{IJ}$ in terms of the corresponding
+components $c_{ijkl},ijkl,i,j,k,l=1,2,3$ of the elastic tensor in
+cartesian coordinates follows \citet{ChTr07}. The 21 anisotropic
+kernels are saved in the \texttt{LOCAL\_PATH} in one file with the
+name of \texttt{proc??????\_cijkl\_kernel.bin} (with \texttt{proc??????}
 the processor number). The output kernels correspond to absolute perturbations
 $\delta C_{IJ}$ of the elastic parameters and their unit is in $s/GPa/km^{3}$.
-For consistency, the output density kernels with this option turned on 
-are for a perturbation $\delta\rho$ (and not $\frac{\delta\rho}{\rho}$) and their unit
-is in s / (kg/m$^{3}$) / km$^{3}$. These `primary' anisotropic kernels
-can then be combined to obtain the kernels for different parameterizations of anisotropy. 
-This can be done, for example, when combining the kernel files from slices into one mesh file 
-(see Section~\ref{sec:Finite-Frequency-Kernels}).
-
-If \texttt{ANISOTROPIC\_KL} \texttt{=} \texttt{.true.} by additionally setting \texttt{ANISOTROPIC\_KL} 
-\texttt{=} \texttt{.true.} in \texttt{constants.h} the package will save anisotropic kernels paramterized
-as velocities related to transverse isotropy based on the the Chen and Tromp parameters \citet{ChTr07}.
-The kernels are saved as relative perturbations for horizontal and vertical P and S velocities, $\alpha_{v},
-\alpha_{h}, \beta_{v}, \beta_{h}$. Explicit relations can be found in appendix B. of \citet{SiLiTrTr07b}
+For consistency, the output density kernels with this option turned
+on are for a perturbation $\delta\rho$ (and not $\frac{\delta\rho}{\rho}$)
+and their unit is in s / (kg/m$^{3}$) / km$^{3}$. These `primary'
+anisotropic kernels can then be combined to obtain the kernels for
+different parameterizations of anisotropy. This can be done, for example,
+when combining the kernel files from slices into one mesh file (see
+Section~\ref{sec:Finite-Frequency-Kernels}).
+
+If \texttt{ANISOTROPIC\_KL} \texttt{=} \texttt{.true.} by additionally
+setting \texttt{ANISOTROPIC\_KL} \texttt{=} \texttt{.true.} in \texttt{constants.h}
+the package will save anisotropic kernels paramterized as velocities
+related to transverse isotropy based on the the Chen and Tromp parameters
+\citet{ChTr07}. The kernels are saved as relative perturbations for
+horizontal and vertical P and S velocities, $\alpha_{v},\alpha_{h},\beta_{v},\beta_{h}$.
+Explicit relations can be found in appendix B. of \citet{SiLiTrTr07b}
 
@@
 \end{enumerate}
-
 In general, the three steps need to be run sequentially to assure
@@
 
-Besides earthquake simulations, SPECFEM3D Cartesian includes functionality for seismic noise tomography as well.
-In order to proceed successfully in this chapter, it is critical that you have already familiarized yourself
-with procedures for meshing (Chapter \ref{cha:Mesh-Generation}),
+Besides earthquake simulations, SPECFEM3D Cartesian includes functionality
+for seismic noise tomography as well. In order to proceed successfully
+in this chapter, it is critical that you have already familiarized
+yourself with procedures for meshing (Chapter \ref{cha:Mesh-Generation}),
 creating distributed databases (Chapter \ref{cha:Creating-Distributed-Databases}),
@@
 and adjoint simulations (Chapter \ref{cha:Adjoint-Simulations}).
-Also, make sure you read the article `Noise cross-correlation sensitivity kernels' \citep{trompetal2010},
-in order to understand noise simulations from a theoretical perspective.
+Also, make sure you read the article `Noise cross-correlation sensitivity
+kernels' \citep{trompetal2010}, in order to understand noise simulations
+from a theoretical perspective.
 
@@
 
-As usual, the three main input files are crucial: \texttt{Par\_file}, \texttt{CMTSOLUTION} and \texttt{STATIONS}.
-Unless otherwise specified, those input files should be located in directory \texttt{DATA/}.\\
-
-\texttt{CMTSOLUTION} is required for all simulations.
-At a first glance, it may seem unexpected to have it here,
-since the noise simulations should have nothing to do with the earthquake -- \texttt{CMTSOLUTION}.
+As usual, the three main input files are crucial: \texttt{Par\_file},
+\texttt{CMTSOLUTION} and \texttt{STATIONS}. Unless otherwise specified,
+those input files should be located in directory \texttt{DATA/}.\\
+
+\texttt{CMTSOLUTION} is required for all simulations. At a first glance,
+it may seem unexpected to have it here, since the noise simulations
+should have nothing to do with the earthquake -- \texttt{CMTSOLUTION}.
 However, for noise simulations, it is critical to have no earthquakes.
-In other words, the moment tensor specified in \texttt{CMTSOLUTION} must be set to zero manually!\\
+In other words, the moment tensor specified in \texttt{CMTSOLUTION}
+must be set to zero manually!\\
 
 \texttt{STATIONS} remains the same as in previous earthquake simulations,
-except that the order of receivers listed in \texttt{STATIONS} is now important.
-The order will be used to determine the `master' receiver,
+except that the order of receivers listed in \texttt{STATIONS} is
+now important. The order will be used to determine the `master' receiver,
 i.e., the one that simultaneously cross correlates with the others.\\
 
-\texttt{Par\_file} also requires careful attention.
-A parameter called \texttt{NOISE\_TOMOGRAPHY} has been added which specifies the type of simulation to be run.
-\texttt{NOISE\_TOMOGRAPHY} is an integer with possible values 0, 1, 2 and 3.
-For example, when \texttt{NOISE\_TOMOGRAPHY} equals 0, a regular earthquake simulation will be run.
-When it is 1/2/3, you are about to run
-step 1/2/3 of the noise simulations respectively.
-Should you be confused by the three steps, refer to \citet{trompetal2010} for details.\\
+\texttt{Par\_file} also requires careful attention. A parameter called
+\texttt{NOISE\_TOMOGRAPHY} has been added which specifies the type
+of simulation to be run. \texttt{NOISE\_TOMOGRAPHY} is an integer
+with possible values 0, 1, 2 and 3. For example, when \texttt{NOISE\_TOMOGRAPHY}
+equals 0, a regular earthquake simulation will be run. When it is
+1/2/3, you are about to run step 1/2/3 of the noise simulations respectively.
+Should you be confused by the three steps, refer to \citet{trompetal2010}
+for details.\\
 
 Another change to \texttt{Par\_file} involves the parameter \texttt{NSTEP}.
-While for regular earthquake simulations this parameter specifies the length of synthetic seismograms generated,
-for noise simulations it specifies the length of the seismograms used to compute cross correlations.
+While for regular earthquake simulations this parameter specifies
+the length of synthetic seismograms generated, for noise simulations
+it specifies the length of the seismograms used to compute cross correlations.
 The actual cross correlations are thus twice this length, i.e., $2~\mathrm{NSTEP}-1$.
-The code automatically makes the modification accordingly, if \texttt{NOISE\_TOMOGRAPHY} is not zero.\\
-
-There are other parameters in \texttt{Par\_file} which should be given specific values.
-For instance, since the first two steps for calculating noise sensitivity kernels correspond to
-forward simulations, \texttt{SIMULATION\_TYPE} must be 1 when \texttt{NOISE\_TOMOGRAPHY} equals 1 or 2.
-Also, we have to reconstruct the ensemble forward wavefields in adjoint simulations,
-therefore we need to set \texttt{SAVE\_FORWARD} to \texttt{.true.} for the second step,
-i.e., when \texttt{NOISE\_TOMOGRAPHY} equals 2.
-The third step is for kernel constructions. Hence \texttt{SIMULATION\_TYPE} should be 3, whereas
-\texttt{SAVE\_FORWARD} must be \texttt{.false.}.\\
-
-Finally, for most system architectures,
-please make sure that \texttt{LOCAL\_PATH} in \texttt{Par\_file} is in fact local, not globally shared.
-Because we have to save the wavefields at the earth's surface at every time step,
-it is quite problematic to have a globally shared \texttt{LOCAL\_PATH},
+The code automatically makes the modification accordingly, if \texttt{NOISE\_TOMOGRAPHY}
+is not zero.\\
+
+There are other parameters in \texttt{Par\_file} which should be given
+specific values. For instance, since the first two steps for calculating
+noise sensitivity kernels correspond to forward simulations, \texttt{SIMULATION\_TYPE}
+must be 1 when \texttt{NOISE\_TOMOGRAPHY} equals 1 or 2. Also, we
+have to reconstruct the ensemble forward wavefields in adjoint simulations,
+therefore we need to set \texttt{SAVE\_FORWARD} to \texttt{.true.}
+for the second step, i.e., when \texttt{NOISE\_TOMOGRAPHY} equals
+2. The third step is for kernel constructions. Hence \texttt{SIMULATION\_TYPE}
+should be 3, whereas \texttt{SAVE\_FORWARD} must be \texttt{.false.}.\\
+
+Finally, for most system architectures, please make sure that \texttt{LOCAL\_PATH}
+in \texttt{Par\_file} is in fact local, not globally shared. Because
+we have to save the wavefields at the earth's surface at every time
+step, it is quite problematic to have a globally shared \texttt{LOCAL\_PATH},
 in terms of both disk storage and I/O speed.
@@
 
-Proper parameters in those parameter files are not enough for noise simulations to run.
-We have more parameters to specify:
-for example, the ensemble-averaged noise spectrum, the noise distribution etc.
-However, since there are a few `new' files, it is better to introduce them sequentially.
-In this section, standard procedures for noise simulations are described.
+Proper parameters in those parameter files are not enough for noise
+simulations to run. We have more parameters to specify: for example,
+the ensemble-averaged noise spectrum, the noise distribution etc.
+However, since there are a few `new' files, it is better to introduce
+them sequentially. In this section, standard procedures for noise
+simulations are described.
 
 \subsection{Pre-simulation}
-
 \begin{itemize}
-
-\item
-As usual, we first configure the software package using:
+\item As usual, we first configure the software package using:
 
@@
 
-\texttt{./configure FC=ifort MPIFC=mpif90 --with-scotch-dir=/opt/scotch}\\
-
-\item
-Next, we need to compile the source code using:
+\texttt{./configure FC=ifort MPIFC=mpif90 -{}-with-scotch-dir=/opt/scotch}\\
+
+\item Next, we need to compile the source code using:
 
@@
 
-\item
-Before we can run noise simulations, we have to specify the noise statistics,
-e.g., the ensemble-averaged noise spectrum.
-Matlab scripts are provided to help you to generate the necessary file.
+\item Before we can run noise simulations, we have to specify the noise
+statistics, e.g., the ensemble-averaged noise spectrum. Matlab scripts
+are provided to help you to generate the necessary file.
 
@@
 
-\texttt{DT} is given in \texttt{Par\_file},
-but \texttt{NSTEP} is NOT the one specified in \texttt{Par\_file}.
-Instead, you have to feed $2~\mathrm{NSTEP}-1$ to account for the doubled length of cross correlations.
-\texttt{Tmin} and \texttt{Tmax} correspond to the period range you are interested in,
-whereas \texttt{NOISE\_MODEL} denotes the noise model you will be using.
-Details can be found in the Matlab script.
-
-After running the Matlab script, you will be given the following information (plus a figure in Matlab):
-
-\texttt{*************************************************************}\\
+\texttt{DT} is given in \texttt{Par\_file}, but \texttt{NSTEP} is
+NOT the one specified in \texttt{Par\_file}. Instead, you have to
+feed $2~\mathrm{NSTEP}-1$ to account for the doubled length of cross
+correlations. \texttt{Tmin} and \texttt{Tmax} correspond to the period
+range you are interested in, whereas \texttt{NOISE\_MODEL} denotes
+the noise model you will be using. Details can be found in the Matlab
+script.
+
+After running the Matlab script, you will be given the following information
+(plus a figure in Matlab):
+
+\texttt{{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}}\\
  \texttt{the source time function has been saved in:}\\
-\texttt{/data2/yangl/3D\_NOISE/S\_squared} (note this path must be different)\\
+ \texttt{/data2/yangl/3D\_NOISE/S\_squared} (note this path must be
+different)\\
  \texttt{S\_squared should be put into directory:}\\
-\texttt{OUTPUT\_FILES/NOISE\_TOMOGRAPHY/ in the SPECFEM3D Cartesian package}\\
+ \texttt{OUTPUT\_FILES/NOISE\_TOMOGRAPHY/ in the SPECFEM3D Cartesian
+package}\\
 
@@
 
-One may choose a flat noise spectrum rather than Peterson's noise model.
-This can be done easily by modifying the Matlab script a little.
-
-\item
-Create a new directory in the SPECFEM3D Cartesian package, name it as \texttt{OUTPUT\_FILES/NOISE\_TOMOGRAPHY/}.
-We will add some parameter files later in this folder.
-
-\item
-Put the Matlab-generated-file \texttt{S\_squared} in \texttt{OUTPUT\_FILES/NOISE\_TOMOGRAPHY/}.
-
-That's to say, you will have a file \texttt{OUTPUT\_FILES/NOISE\_TOMOGRAPHY/S\_squared} in the SPECFEM3D Cartesian package.
-
-\item
-Create a file called \texttt{OUTPUT\_FILES/NOISE\_TOMOGRAPHY/irec\_master\_noise}. Note that this file is located
-in directory \texttt{OUTPUT\_FILES/NOISE\_TOMOGRAPHY/} as well. In general, all noise simulation related parameter
-files go into that directory.
-\texttt{irec\_master\_noise} contains only one integer, which is the ID
-of the `master' receiver. For example, if this file contains 5, it means that the fifth receiver
-listed in \texttt{DATA/STATIONS} becomes the `master'. That's why we mentioned previously that
-the order of receivers in \texttt{DATA/STATIONS} is important.
-
-Note that in regional simulations, the \texttt{DATA/STATIONS}
-might contain receivers which are outside of our computational domains.
-Therefore, the integer in \texttt{irec\_master\_noise}
-is actually the ID in \texttt{DATA/STATIONS\_FILTERED}
-(which is generated by \texttt{bin/xgenerate\_databases}).
-
-\item
-Create a file called \texttt{OUTPUT\_FILES/NOISE\_TOMOGRAPHY/nu\_master}. This file holds three numbers,
-forming a (unit) vector. It describes which component we are cross-correlating at the `master'
-receiver, i.e., ${\hat{\bf \nu}}^{\alpha}$ in \citet{trompetal2010}.
-The three numbers correspond to E/N/Z components respectively.
-Most often, the vertical component is used, and in those cases the three numbers should be 0, 0 and 1.
-
-\item
-Describe the noise direction and distributions in \texttt{src/shared/noise\_tomography.f90}.
+One may choose a flat noise spectrum rather than Peterson's noise
+model. This can be done easily by modifying the Matlab script a little.
+
+\item Create a new directory in the SPECFEM3D Cartesian package, name it
+as \texttt{OUTPUT\_FILES/NOISE\_TOMOGRAPHY/}. We will add some parameter
+files later in this folder.
+\item Put the Matlab-generated-file \texttt{S\_squared} in \texttt{OUTPUT\_FILES/NOISE\_TOMOGRAPHY/}.
+
+That's to say, you will have a file \texttt{OUTPUT\_FILES/NOISE\_TOMOGRAPHY/S\_squared}
+in the SPECFEM3D Cartesian package.
+
+\item Create a file called \texttt{OUTPUT\_FILES/NOISE\_TOMOGRAPHY/irec\_master\_noise}.
+Note that this file is located in directory \texttt{OUTPUT\_FILES/NOISE\_TOMOGRAPHY/}
+as well. In general, all noise simulation related parameter files
+go into that directory. \texttt{irec\_master\_noise} contains only
+one integer, which is the ID of the `master' receiver. For example,
+if this file contains 5, it means that the fifth receiver listed in
+\texttt{DATA/STATIONS} becomes the `master'. That's why we mentioned
+previously that the order of receivers in \texttt{DATA/STATIONS} is
+important.
+
+Note that in regional simulations, the \texttt{DATA/STATIONS} might
+contain receivers which are outside of our computational domains.
+Therefore, the integer in \texttt{irec\_master\_noise} is actually
+the ID in \texttt{DATA/STATIONS\_FILTERED} (which is generated by
+\texttt{bin/xgenerate\_databases}).
+
+\item Create a file called \texttt{OUTPUT\_FILES/NOISE\_TOMOGRAPHY/nu\_master}.
+This file holds three numbers, forming a (unit) vector. It describes
+which component we are cross-correlating at the `master' receiver,
+i.e., ${\hat{{\bf \nu}}}^{\alpha}$ in \citet{trompetal2010}. The
+three numbers correspond to E/N/Z components respectively. Most often,
+the vertical component is used, and in those cases the three numbers
+should be 0, 0 and 1.
+\item Describe the noise direction and distributions in \texttt{src/shared/noise\_tomography.f90}.
 Search for a subroutine called \texttt{noise\_distribution\_direction}
-in \texttt{noise\_tomography.f90}. It is actually located at the very beginning
-of \texttt{noise\_tomography.f90}. The default assumes vertical noise and a uniform distribution
-across the whole free surface of the model. It should be quite self-explanatory for modifications.
+in \texttt{noise\_tomography.f90}. It is actually located at the very
+beginning of \texttt{noise\_tomography.f90}. The default assumes vertical
+noise and a uniform distribution across the whole free surface of
+the model. It should be quite self-explanatory for modifications.
 Should you modify this part, you have to re-compile the source code.
-
 \end{itemize}
@@
 With all of the above done, we can finally launch our simulations.
-Again, please make sure that the \texttt{LOCAL\_PATH} in \texttt{DATA/Par\_file} is not globally shared.
-It is quite problematic to have a globally shared \texttt{LOCAL\_PATH}, in terms of both disk storage and
-speed of I/O (we have to save the wavefields at the earth's surface at every time step).
-
-As discussed in \citet{trompetal2010}, it takes three steps/simulations to obtain one contribution
-of the ensemble sensitivity kernels:
-
+Again, please make sure that the \texttt{LOCAL\_PATH} in \texttt{DATA/Par\_file}
+is not globally shared. It is quite problematic to have a globally
+shared \texttt{LOCAL\_PATH}, in terms of both disk storage and speed
+of I/O (we have to save the wavefields at the earth's surface at every
+time step).
+
+As discussed in \citet{trompetal2010}, it takes three steps/simulations
+to obtain one contribution of the ensemble sensitivity kernels:
 \begin{itemize}
-\item
-Step 1: simulation for generating wavefields\\
+\item Step 1: simulation for generating wavefields\\
  \texttt{SIMULATION\_TYPE=1}\\
@@
  \texttt{SAVE\_FORWARD} not used, can be either .true. or .false.
-
-\item
-Step 2: simulation for ensemble forward wavefields\\
+\item Step 2: simulation for ensemble forward wavefields\\
  \texttt{SIMULATION\_TYPE=1}\\
@@
  \texttt{SAVE\_FORWARD=.true.}
-
-\item
-Step 3: simulation for ensemble adjoint wavefields and sensitivity kernels\\
+\item Step 3: simulation for ensemble adjoint wavefields and sensitivity
+kernels\\
  \texttt{SIMULATION\_TYPE=3}\\
@@
 
-Note Step 3 is an adjoint simulation, please refer to previous chapters on how to prepare adjoint sources
-and other necessary files, as well as how adjoint simulations work.
+Note Step 3 is an adjoint simulation, please refer to previous chapters
+on how to prepare adjoint sources and other necessary files, as well
+as how adjoint simulations work.
 
 \end{itemize}
-
-Note that it is better to run the three steps continuously within the same job,
-otherwise you have to collect the saved surface movies from the old nodes/CPUs to the new nodes/CPUs.
-This process varies from cluster to cluster and thus cannot be discussed here.
-Please ask your cluster administrator for information/configuration of the cluster you are using.\\
+Note that it is better to run the three steps continuously within
+the same job, otherwise you have to collect the saved surface movies
+from the old nodes/CPUs to the new nodes/CPUs. This process varies
+from cluster to cluster and thus cannot be discussed here. Please
+ask your cluster administrator for information/configuration of the
+cluster you are using.\\
 
@@
 
-After those simulations, you have all stuff you need, either in the \texttt{OUTPUT\_FILES/} or
-in the directory specified by \texttt{LOCAL\_PATH} in \texttt{DATA/Par\_file}
-(which are most probably on local nodes).
-Collect whatever you want from the local nodes to your workstation, and then visualize them.
-This process is the same as what you may have done for regular earthquake simulations.
-Refer to other chapters if you have problems.\\
-
-Simply speaking, two outputs are the most interesting: the simulated ensemble cross correlations and
-one contribution of the ensemble sensitivity kernels.\\
-
-The simulated ensemble cross correlations are obtained after the second simulation (Step 2).
-Seismograms in \texttt{OUTPUT\_FILES/} are actually the simulated ensemble cross correlations.
-Collect them immediately after Step 2, or the Step 3 will overwrite them.
-Note that we have a `master' receiver specified by \texttt{irec\_master\_noise},
-the seismogram at one station corresponds to the cross correlation between that station and the `master'.
-Since the seismograms have three components, we may obtain cross correlations
-for different components as well, not necessarily the cross correlations between vertical components.\\
-
-One contribution of the ensemble cross-correlation sensitivity kernels are obtained after Step 3,
-stored in the \texttt{DATA/LOCAL\_PATH} on local nodes.
-The ensemble kernel files are named the same as regular earthquake kernels.\\
-
-You need to run another three simulations to get the other contribution of the ensemble kernels,
-using different forward and adjoint sources given in \citet{trompetal2010}.
+After those simulations, you have all stuff you need, either in the
+\texttt{OUTPUT\_FILES/} or in the directory specified by \texttt{LOCAL\_PATH}
+in \texttt{DATA/Par\_file} (which are most probably on local nodes).
+Collect whatever you want from the local nodes to your workstation,
+and then visualize them. This process is the same as what you may
+have done for regular earthquake simulations. Refer to other chapters
+if you have problems.\\
+
+Simply speaking, two outputs are the most interesting: the simulated
+ensemble cross correlations and one contribution of the ensemble sensitivity
+kernels.\\
+
+The simulated ensemble cross correlations are obtained after the second
+simulation (Step 2). Seismograms in \texttt{OUTPUT\_FILES/} are actually
+the simulated ensemble cross correlations. Collect them immediately
+after Step 2, or the Step 3 will overwrite them. Note that we have
+a `master' receiver specified by \texttt{irec\_master\_noise}, the
+seismogram at one station corresponds to the cross correlation between
+that station and the `master'. Since the seismograms have three components,
+we may obtain cross correlations for different components as well,
+not necessarily the cross correlations between vertical components.\\
+
+One contribution of the ensemble cross-correlation sensitivity kernels
+are obtained after Step 3, stored in the \texttt{DATA/LOCAL\_PATH}
+on local nodes. The ensemble kernel files are named the same as regular
+earthquake kernels.\\
+
+You need to run another three simulations to get the other contribution
+of the ensemble kernels, using different forward and adjoint sources
+given in \citet{trompetal2010}.
 
@@
 
-In order to illustrate noise simulations in an easy way, one example is provided in \texttt{examples/noise\_tomography/}.
-See \texttt{examples/noise\_tomography/README} for explanations. \\
+In order to illustrate noise simulations in an easy way, one example
+is provided in \texttt{examples/noise\_tomography/}. See \texttt{examples/noise\_tomography/README}
+for explanations. \\
 
@@
 
-If your workstation is suitable, you can run the example in \texttt{examples/noise\_tomography/} using:\\
+If your workstation is suitable, you can run the example in \texttt{examples/noise\_tomography/}
+using:\\
 
@@
 
-Even if this script does not work on your workstation, the procedure it describes is universal.
-You may review the whole process described in the last section by following the commands in \texttt{pre-processing.sh},
+Even if this script does not work on your workstation, the procedure
+it describes is universal. You may review the whole process described
+in the last section by following the commands in \texttt{pre-processing.sh},
 which should contain enough explanations for all the commands.
@@
 
-In case you used the internal mesher \texttt{xmeshfem3D} to create and partition your mesh,
-you can output mesh files in ABAQUS (.INP) and DX (.dx) format to visualize them. For this, you must set either the
-flag \texttt{CREATE\_DX\_FILES} or \texttt{CREATE\_ABAQUS\_FILES} to \texttt{.true.}
-in the mesher's parameter file \texttt{Mesh\_Par\_file}
-prior to running the mesher (see Chapter~\ref{cha:Running-the-Mesher-Meshfem3D} for details).
-You can then use AVS \urlwithparentheses{www.avs.com} or OpenDX \urlwithparentheses{www.opendx.org}
-to visualize the mesh and MPI partition (slices).
+In case you used the internal mesher \texttt{xmeshfem3D} to create
+and partition your mesh, you can output mesh files in ABAQUS (.INP)
+and DX (.dx) format to visualize them. For this, you must set either
+the flag \texttt{CREATE\_DX\_FILES} or \texttt{CREATE\_ABAQUS\_FILES}
+to \texttt{.true.} in the mesher's parameter file \texttt{Mesh\_Par\_file}
+prior to running the mesher (see Chapter~\ref{cha:Running-the-Mesher-Meshfem3D}
+for details). You can then use AVS \urlwithparentheses{www.avs.com}
+or OpenDX \urlwithparentheses{www.opendx.org} to visualize the mesh
+and MPI partition (slices).
 
@@
 \noindent \begin{centering}
-\includegraphics[width=.49\textwidth]{figures/vtk_mesh_vp.jpg}
-\includegraphics[width=.49\textwidth]{figures/vtk_mesh_vs.jpg}
+\includegraphics[width=0.49\textwidth]{figures/vtk_mesh_vp} \includegraphics[width=0.49\textwidth]{figures/vtk_mesh_vs} 
 \par\end{centering}
 
-\caption{Visualization using Paraview of VTK files created by \texttt{xgenerate\_databases} showing P- and S-wave velocities assigned to the mesh points.
-The mesh was created by \texttt{xmeshfem3D} for 4 processors.}
+\caption{Visualization using Paraview of VTK files created by \texttt{xgenerate\_databases}
+showing P- and S-wave velocities assigned to the mesh points. The
+mesh was created by \texttt{xmeshfem3D} for 4 processors.}
 
@@
 
-You have also the option to visualize the distributed databases produced by \texttt{xgenerate\_databases}
-using Paraview \urlwithparentheses{www.paraview.org}. For this, you must set the flag \texttt{SAVE\_MESH\_FILES} to \texttt{.true.}
-in the main parameter file \texttt{Par\_file} (see Chapter~\ref{cha:Main-Parameter} for details).
-This will create VTK files for each single partition. You can then use Paraview \urlwithparentheses{www.paraview.org} to visualized these partitions.
+You have also the option to visualize the distributed databases produced
+by \texttt{xgenerate\_databases} using Paraview \urlwithparentheses{www.paraview.org}.
+For this, you must set the flag \texttt{SAVE\_MESH\_FILES} to \texttt{.true.}
+in the main parameter file \texttt{Par\_file} (see Chapter~\ref{cha:Main-Parameter}
+for details). This will create VTK files for each single partition.
+You can then use Paraview \urlwithparentheses{www.paraview.org}
+to visualized these partitions.
 
@@
 To make a surface or volume movie of the simulation, set parameters
-\texttt{MOVIE\_SURFACE}, \texttt{MOVIE\_VOLUME}, \texttt{MOVIE\_TYPE}, and \texttt{NTSTEP\_BETWEEN\_FRAMES}
-in the \texttt{Par\_file}. Turning on the movie flags, in particular
-\texttt{MOVIE\_VOLUME}, produces large output files. \texttt{MOVIE\_VOLUME}
-files are saved in the \texttt{LOCAL\_PATH} directory, whereas \texttt{MOVIE\_SURFACE}
-output files are saved in the \texttt{OUTPUT\_FILES} directory. We
-save the displacement field if the parameter \texttt{SAVE\_DISPLACEMENT} is set, otherwise the velocity field is saved.
-The look of a movie is determined by the
-half-duration of the source. The half-duration should be large enough
-so that the movie does not contain frequencies that are not resolved
-by the mesh, i.e., it should not contain numerical noise. This can
-be accomplished by selecting a CMT \texttt{HALF\_DURATION} > 1.1 $\times$
-smallest period (see figure \ref{fig:CMTSOLUTION-file}). When \texttt{MOVIE\_SURFACE}
-= .\texttt{true.}, the half duration of each source in the \texttt{CMTSOLUTION}
-file is replaced by
-
+\texttt{MOVIE\_SURFACE}, \texttt{MOVIE\_VOLUME}, \texttt{MOVIE\_TYPE},
+and \texttt{NTSTEP\_BETWEEN\_FRAMES} in the \texttt{Par\_file}. Turning
+on the movie flags, in particular \texttt{MOVIE\_VOLUME}, produces
+large output files. \texttt{MOVIE\_VOLUME} files are saved in the
+\texttt{LOCAL\_PATH} directory, whereas \texttt{MOVIE\_SURFACE} output
+files are saved in the \texttt{OUTPUT\_FILES} directory. We save the
+displacement field if the parameter \texttt{SAVE\_DISPLACEMENT} is
+set, otherwise the velocity field is saved. The look of a movie is
+determined by the half-duration of the source. The half-duration should
+be large enough so that the movie does not contain frequencies that
+are not resolved by the mesh, i.e., it should not contain numerical
+noise. This can be accomplished by selecting a CMT \texttt{HALF\_DURATION}
+> 1.1 $\times$ smallest period (see figure \ref{fig:CMTSOLUTION-file}).
+When \texttt{MOVIE\_SURFACE} = .\texttt{true.}, the half duration
+of each source in the \texttt{CMTSOLUTION} file is replaced by
 \begin{quote}
 \[
-\sqrt{(}\mathrm{\mathtt{HALF\_DURATIO}\mathtt{N}^{2}}+\mathrm{\mathtt{HDUR\_MOVI}\mathtt{E}^{2}})\]
+\sqrt{(}\mathrm{\mathtt{HALF\_DURATIO}\mathtt{N}^{2}}+\mathrm{\mathtt{HDUR\_MOVI}\mathtt{E}^{2}})
+\]
  \textbf{NOTE:} If \texttt{HDUR\_MOVIE} is set to 0.0, the code will
@@
 usual, for a point source one can set \texttt{half duration} in the
-\texttt{CMTSOLUTION} file to be 0.0 and \texttt{HDUR\_MOVIE} = 0.0 in the \texttt{Par\_file} to get
-the highest frequencies resolved by the simulation, but for a finite
-source one would keep all the \texttt{half durations} as prescribed
-by the finite source model and set \texttt{HDUR\_MOVIE} = 0.0.
+\texttt{CMTSOLUTION} file to be 0.0 and \texttt{HDUR\_MOVIE} = 0.0
+in the \texttt{Par\_file} to get the highest frequencies resolved
+by the simulation, but for a finite source one would keep all the
+\texttt{half durations} as prescribed by the finite source model and
+set \texttt{HDUR\_MOVIE} = 0.0. 
 \end{quote}
@@
 flag turned on, the code outputs \texttt{moviedata??????} files in
-the \texttt{OUTPUT\_FILES} directory. There are several flags
-in the main parameter file \texttt{Par\_file} that control the output of these movie data files
-(see section \ref{cha:Main-Parameter} for details):
-\texttt{\small NTSTEP\_BETWEEN\_FRAMES} to set the timesteps between frames,
-\texttt{\small SAVE\_DISPLACEMENT} to save displacement instead of velocity,
-\texttt{USE\_HIGHRES\_FOR\_MOVIES} to save values at all GLL point instead of element edges.
-In order to output additionally shakemaps, you would set
-the parameter \texttt{CREATE\_SHAKEMAP} to \texttt{.true.}.
-
-The files are in a fairly complicated
-binary format, but there is a program provided to convert the
-output into more user friendly formats:
+the \texttt{OUTPUT\_FILES} directory. There are several flags in the
+main parameter file \texttt{Par\_file} that control the output of
+these movie data files (see section \ref{cha:Main-Parameter} for
+details): \texttt{\small NTSTEP\_BETWEEN\_FRAMES}{\small{} to set the
+timesteps between frames, }\texttt{\small SAVE\_DISPLACEMENT}{\small{}
+to save displacement instead of velocity, }\texttt{\small USE\_HIGHRES\_FOR\_MOVIES}{\small{}
+to save values at all GLL point instead of element edges. In order
+to output additionally shakemaps, you would set the parameter }\texttt{\small CREATE\_SHAKEMAP}{\small{}
+to }\texttt{\small .true.}{\small .}{\small \par}
+
+The files are in a fairly complicated binary format, but there is
+a program provided to convert the output into more user friendly formats: 
 \begin{description}
-\item [\texttt{xcreate\_movie\_shakemap\_AVS\_DX\_GMT}]
-From \texttt{create\_movie\_shakemap\_AVS\_DX\_GMT.f90}, it
-outputs data in ASCII, OpenDX, or AVS format (also readable in ParaView).
-Before compiling the code, make sure you have the file \texttt{surface\_from\_mesher.h}
-in the  \texttt{OUTPUT\_FILES/} directory. This file will be created by the solver run.
-Then type
-\begin{lyxcode}
-{\footnotesize make xcreate\_movie\_shakemap\_AVS\_DX\_GMT}{\footnotesize \par}
-\end{lyxcode}
+\item [{\texttt{xcreate\_movie\_shakemap\_AVS\_DX\_GMT}}] From \texttt{create\_movie\_shakemap\_AVS\_DX\_GMT.f90},
+it outputs data in ASCII, OpenDX, or AVS format (also readable in
+ParaView). Before compiling the code, make sure you have the file
+\texttt{surface\_from\_mesher.h} in the \texttt{OUTPUT\_FILES/} directory.
+This file will be created by the solver run. Then type
+
+\begin{lyxcode}
+{\footnotesize make~xcreate\_movie\_shakemap\_AVS\_DX\_GMT}{\footnotesize \par}
+\end{lyxcode}
+
 and run the executable \texttt{xcreate\_movie\_shakemap\_AVS\_DX\_GMT}
-in the \texttt{bin/} subdirectory. It will create visualization files in your format of choice. The code
-will prompt the user for input parameters.
+in the \texttt{bin/} subdirectory. It will create visualization files
+in your format of choice. The code will prompt the user for input
+parameters.
 
@@
 \noindent \begin{centering}
-\includegraphics[width=.32\textwidth]{figures/movie_surf_1.jpg}
-\includegraphics[width=.32\textwidth]{figures/movie_surf_2.jpg}
-\includegraphics[width=.32\textwidth]{figures/movie_surf_3.jpg}
+\includegraphics[width=0.32\textwidth]{figures/movie_surf_1} \includegraphics[width=0.32\textwidth]{figures/movie_surf_2}
+\includegraphics[width=0.32\textwidth]{figures/movie_surf_3} 
 \par\end{centering}
 
-\caption{Visualization using AVS files created by \texttt{xcreate\_movie\_shakemap\_AVS\_DX\_GMT} showing
-movie snapshots of vertical velocity components at different times.}
+\caption{Visualization using AVS files created by \texttt{xcreate\_movie\_shakemap\_AVS\_DX\_GMT}
+showing movie snapshots of vertical velocity components at different
+times.}
 
@@
 \end{description}
-
-The \texttt{SPECFEM3D Cartesian} code is running in near real-time to produce
-animations of southern California earthquakes via the web; see Southern California ShakeMovie\textregistered \urlwithparentheses{www.shakemovie.caltech.edu}.
+The \texttt{SPECFEM3D Cartesian} code is running in near real-time
+to produce animations of southern California earthquakes via the web;
+see Southern California ShakeMovie\textregistered{}\urlwithparentheses{www.shakemovie.caltech.edu}.
 
@@
 
-When running xspecfem3D with the \texttt{\small MOVIE\_VOLUME} flag
-turned on, the code outputs several files in \texttt{\small LOCAL\_PATH} specified in the main
-\texttt{Par\_file}, e.g. in directory \texttt{OUTPUT\_FILES/DATABASES\_MPI}.
-%As the files can be very large, there are several flags in the \texttt{\small Par\_file}
+When running xspecfem3D with the \texttt{\small MOVIE\_VOLUME}{\small{}
+flag turned on, the code outputs several files in }\texttt{\small LOCAL\_PATH}{\small{}
+specified in the main }\texttt{\small Par\_file}{\small , e.g. in
+directory }\texttt{\small OUTPUT\_FILES/DATABASES\_MPI}{\small . %As the files can be very large, there are several flags in the \texttt{\small Par\_file}
 %that control the region in space and time that is saved. These are:
 %\texttt{\small NTSTEP\_BETWEEN\_FRAMES}
-The output is saved by each processor
-at the time interval specified by \texttt{\small NTSTEP\_BETWEEN\_FRAMES}.
-For all domains, the velocity field is output to files:
-\texttt{\small proc??????\_velocity\_X\_it??????.bin},
-\texttt{\small proc??????\_velocity\_Y\_it??????.bin} and
-\texttt{\small proc??????\_velocity\_Z\_it??????.bin}.
-For elastic domains, the divergence and curl taken from the velocity field, i.e. $\nabla \cdot \bf{v}$
-and $\nabla \times \bf{v}$, get stored as well:
-\texttt{\small proc??????\_div\_it??????.bin},
-\texttt{\small proc??????\_curl\_X\_t??????.bin},
-\texttt{\small proc??????\_curl\_Y\_it??????.bin} and
-\texttt{\small proc??????\_curl\_Z\_it??????.bin}.
-The files denoted
-\texttt{\small proc??????\_div\_}~\\
-\texttt{\small glob\_it??????.bin} and
-\texttt{\small proc??????\_curl\_glob\_it??????.bin}
-are stored on the global points only, all the other arrays
-are stored on all GLL points. Note that the components X/Y/Z can change to E/N/Z according to
-the \texttt{SUPPRESS\_UTM\_PROJECTION} flag (see also Appendix~\ref{cha:Coordinates} and \ref{cha:channel-codes}).
+The output is saved by each processor at the time interval specified
+by }\texttt{\small NTSTEP\_BETWEEN\_FRAMES}{\small . For all domains,
+the velocity field is output to files: }\texttt{\small proc??????\_velocity\_X\_it??????.bin}{\small ,
+}\texttt{\small proc??????\_velocity\_Y\_it??????.bin}{\small{} and
+}\texttt{\small proc??????\_velocity\_Z\_it??????.bin}{\small . For
+elastic domains, the divergence and curl taken from the velocity field,
+i.e. $\nabla\cdot{\bf {v}}$ and $\nabla\times{\bf {v}}$, get stored
+as well: }\texttt{\small proc??????\_div\_it??????.bin}{\small , }\texttt{\small proc??????\_curl\_X\_t??????.bin}{\small ,
+}\texttt{\small proc??????\_curl\_Y\_it??????.bin}{\small{} and }\texttt{\small proc??????\_curl\_Z\_it??????.bin}{\small .
+The files denoted }\texttt{\small proc??????\_div\_}{\small ~}\\
+{\small{} }\texttt{\small glob\_it??????.bin}{\small{} and }\texttt{\small proc??????\_curl\_glob\_it??????.bin}{\small{}
+are stored on the global points only, all the other arrays are stored
+on all GLL points. Note that the components X/Y/Z can change to E/N/Z
+according to the }\texttt{\small SUPPRESS\_UTM\_PROJECTION}{\small{}
+flag (see also Appendix~\ref{cha:Coordinates} and \ref{cha:channel-codes}).}{\small \par}
 
@@
 \noindent \begin{centering}
-\includegraphics[width=.32\textwidth]{figures/movie_volume_1.jpg}
-\includegraphics[width=.32\textwidth]{figures/movie_volume_2.jpg}
-\includegraphics[width=.32\textwidth]{figures/movie_volume_3.jpg}
+\includegraphics[width=0.32\textwidth]{figures/movie_volume_1} \includegraphics[width=0.32\textwidth]{figures/movie_volume_2}
+\includegraphics[width=0.32\textwidth]{figures/movie_volume_3} 
 \par\end{centering}
 
-\caption{Paraview visualization using movie volume files (converted by \texttt{xcombine\_vol\_data} and \texttt{mesh2vtu.pl}) and showing
-snapshots of vertical velocity components at different times.}
+\caption{Paraview visualization using movie volume files (converted by \texttt{xcombine\_vol\_data}
+and \texttt{mesh2vtu.pl}) and showing snapshots of vertical velocity
+components at different times.}
 
@@
 To visualize these files, we use an auxilliary program \texttt{combine\_vol\_data.f90}
-to combine the data from all slices into one mesh file. To compile it in the root directory, type:
+to combine the data from all slices into one mesh file. To compile
+it in the root directory, type: 
 \begin{lyxcode}
@@
 \end{lyxcode}
-which will create the executable \texttt{xcombine\_vol\_data} in the directory \texttt{bin/}. To output the
-usage of this executable, type `./bin/xcombine\_vol\_data` without arguments.
-As an example, to run the executable you would use
+which will create the executable \texttt{xcombine\_vol\_data} in the
+directory \texttt{bin/}. To output the usage of this executable, type
+`./bin/xcombine\_vol\_data` without arguments. As an example, to run
+the executable you would use 
 \begin{lyxcode}
 {\footnotesize cd~bin/~}{\footnotesize \par}
-{\footnotesize ./xcombine\_vol\_data~0~3~velocity\_Z\_it000400~../OUTPUT\_FILES/DATABASES\_MPI~../OUTPUT\_FILES~0~}{\footnotesize \par}
-\end{lyxcode}
-to create a low-resolution mesh file. The output mesh file will have the name \texttt{velocity\_Z\_it000400.mesh}.
-We next convert the \texttt{.mesh} file into the VTU (Unstructured
-grid file) format which can be viewed in ParaView. For this task, you can use and modify the script \texttt{mesh2vtu.pl}
-located in directory \texttt{utils/Visualization/Paraview/}, for example:
+
+~{\footnotesize ./xcombine\_vol\_data~0~3~velocity\_Z\_it000400~../OUTPUT\_FILES/DATABASES\_MPI~../OUTPUT\_FILES~0~}{\footnotesize \par}
+\end{lyxcode}
+to create a low-resolution mesh file. The output mesh file will have
+the name \texttt{velocity\_Z\_it000400.mesh}. We next convert the
+\texttt{.mesh} file into the VTU (Unstructured grid file) format which
+can be viewed in ParaView. For this task, you can use and modify the
+script \texttt{mesh2vtu.pl} located in directory \texttt{utils/Visualization/Paraview/},
+for example: 
 \begin{lyxcode}
@@
 Notice that this Perl script uses a program \texttt{mesh2vtu} in the
-\texttt{utils/Visualization/Paraview/mesh2vtu} directory, which further uses the VTK \urlwithparentheses{http://www.vtk.org/}
-run-time library for its execution. Therefore, make sure you have
-them properly set in the script according to your system.
+\texttt{utils/Visualization/Paraview/mesh2vtu} directory, which further
+uses the VTK \urlwithparentheses{http://www.vtk.org/} run-time library
+for its execution. Therefore, make sure you have them properly set
+in the script according to your system.
 
@@
 programs.
-
 \begin{enumerate}
@@
 
-A Perl script \texttt{slice\_number.pl} located in directory \texttt{utils/Visualization/Paraview/} can
-help to figure out the slice numbers that lie along the great circle
-path. It applies to meshes created with the internal mesher \texttt{xmeshfem3D}.
-
+A Perl script \texttt{slice\_number.pl} located in directory \texttt{utils/Visualization/Paraview/}
+can help to figure out the slice numbers that lie along the great
+circle path. It applies to meshes created with the internal mesher
+\texttt{xmeshfem3D}.
 \begin{enumerate}
-\item On machines where you have access to the script, copy the \texttt{Mesh\_Par\_file}, and
-\texttt{output\_solver} files, and run:
+\item On machines where you have access to the script, copy the \texttt{Mesh\_Par\_file},
+and \texttt{output\_solver} files, and run:
 
@@
 \end{lyxcode}
+
 which will generate a \texttt{slices\_file}.
@@
 kernel files from the given slices.
-
 \begin{enumerate}
@@
 \begin{lyxcode}
-{\small utils/copy\_database.pl~slice\_file~lsf\_machine\_file~filename~{[}jobid]}{\small \par}
-\end{lyxcode}
-where \texttt{\small lsf\_machine\_file} is the machine file generated
-by the LSF scheduler, \texttt{\small filename} is the kernel name
-(e.g., \texttt{\small rho\_kernel}, \texttt{\small alpha\_kernel}
-and \texttt{\small beta\_kernel}), and the optional \texttt{\small jobid}
-is the name of the subdirectory under \texttt{\small LOCAL\_PATH}
-where all the kernel files are stored.
+{\small utils/copy\_database.pl~slice\_file~lsf\_machine\_file~filename~{[}jobid{]}}{\small \par}
+\end{lyxcode}
+
+where \texttt{\small lsf\_machine\_file}{\small{} is the machine file
+generated by the LSF scheduler, }\texttt{\small filename}{\small{}
+is the kernel name (e.g., }\texttt{\small rho\_kernel}{\small , }\texttt{\small alpha\_kernel}{\small{}
+and }\texttt{\small beta\_kernel}{\small ), and the optional }\texttt{\small jobid}{\small{}
+is the name of the subdirectory under }\texttt{\small LOCAL\_PATH}{\small{}
+where all the kernel files are stored.}{\small \par}
 
@@
 
-We use an auxilliary program \texttt{combine\_vol\_data.f90}
-to combine the kernel files from all slices into one mesh file.
-
+We use an auxilliary program \texttt{combine\_vol\_data.f90} to combine
+the kernel files from all slices into one mesh file.
 \begin{enumerate}
@@
 \end{lyxcode}
+
 where \texttt{input\_dir} is the directory where all the individual
@@
 \item We next convert the \texttt{.mesh} file into the VTU (Unstructured
-grid file) format which can be viewed in ParaView. For this task, you can use and modify the script \texttt{mesh2vtu.pl}
-located in directory \texttt{utils/Visualization/Paraview/}, for example:
+grid file) format which can be viewed in ParaView. For this task,
+you can use and modify the script \texttt{mesh2vtu.pl} located in
+directory \texttt{utils/Visualization/Paraview/}, for example:
 
@@
 \item Notice that this Perl script uses a program \texttt{mesh2vtu} in the
-\texttt{utils/Visualization/Paraview/mesh2vtu} directory, which further uses the VTK \urlwithparentheses{http://www.vtk.org/}
-run-time library for its execution. Therefore, make sure you have
-them properly set in the script according to your system.
+\texttt{utils/Visualization/Paraview/mesh2vtu} directory, which further
+uses the VTK \urlwithparentheses{http://www.vtk.org/} run-time library
+for its execution. Therefore, make sure you have them properly set
+in the script according to your system. 
 \end{enumerate}
@@
 In the case of a single source and a single receiver, the simulation
-also generates the file \texttt{sr.vtk} located in the
-\texttt{OUTPUT\_FILES/} directory to describe
-the source and receiver locations, which can also be viewed in Paraview
-in the next step.
+also generates the file \texttt{sr.vtk} located in the \texttt{OUTPUT\_FILES/}
+directory to describe the source and receiver locations, which can
+also be viewed in Paraview in the next step.
 
@@
 Finally, we can view the mesh in ParaView \urlwithparentheses{www.paraview.org}.
-
 \begin{enumerate}
@@
 \textsf{Glyph}, and represent the points by `\textsf{spheres}'. 
-\item For more information about ParaView, see the ParaView Users Guide \urlwithparentheses{www.paraview.org/files/v1.6/ParaViewUsersGuide.PDF}.
+\item For more information about ParaView, see the ParaView Users Guide
+\urlwithparentheses{www.paraview.org/files/v1.6/ParaViewUsersGuide.PDF}. 
 \end{enumerate}
 \end{enumerate}
-%
 \begin{figure}[H]
@@
 that uses a scheduler:
-
 \begin{itemize}
@@
 by the scheduler, based on availability. Therefore, in order for the
-\texttt{xgenerate\_databases} and the \texttt{xspecfem3D} executables (or between successive runs of the solver) to
-have access to the same database files (if they are stored on hard
-drives local to the nodes on which the code is run), they must be
-launched in sequence as a single job.
+\texttt{xgenerate\_databases} and the \texttt{xspecfem3D} executables
+(or between successive runs of the solver) to have access to the same
+database files (if they are stored on hard drives local to the nodes
+on which the code is run), they must be launched in sequence as a
+single job. 
 \item On some systems, the nodes to which running jobs are assigned are
 not configured for compilation. It may therefore be necessary to pre-compile
-both the \texttt{xgenerate\_databases} and the \texttt{xspecfem3D} executables.
+both the \texttt{xgenerate\_databases} and the \texttt{xspecfem3D}
+executables. 
 \item One feature of schedulers/queuing systems is that they allow submission
@@
 \end{itemize}
-Examples of job scripts can be found in the \texttt{\small utils/Cluster/}
+Examples of job scripts can be found in the \texttt{\small utils/Cluster/}{\small{}
 directory and can straightforwardly be modified and adapted to meet
-more specific running needs.
+more specific running needs.}{\small \par}
 
@@
 Caltech 1024-node cluster, CITerra, under the LSF scheduling system.
-We consider the submission of a regular forward simulation using the internal mesher
-to create mesh partitions. The two
-main scripts are \texttt{\small run\_lsf.bash}, which compiles the
-Fortran code and submits the job to the scheduler, and \texttt{\small go\_mesher\_solver\_lsf\_basin.forward},
-which contains the instructions that make up
-the job itself. These scripts can be found in \texttt{\small utils/Cluster/lsf/} directory
+We consider the submission of a regular forward simulation using the
+internal mesher to create mesh partitions. The two main scripts are
+\texttt{\small run\_lsf.bash}{\small , which compiles the Fortran
+code and submits the job to the scheduler, and }\texttt{\small go\_mesher\_solver\_lsf\_basin.forward}{\small ,
+which contains the instructions that make up the job itself. These
+scripts can be found in }\texttt{\small utils/Cluster/lsf/}{\small{}
+directory}{\small \par}
 
@@
 This script first sets the job queue to be `normal'. It then compiles
-the mesher, database generator and solver together, figures out the number of processors
-required for this simulation from the \texttt{DATA/Par\_file}, and
-submits the LSF job.
-
+the mesher, database generator and solver together, figures out the
+number of processors required for this simulation from the \texttt{DATA/Par\_file},
+and submits the LSF job.
 \begin{lyxcode}
@@
 
-\#~use~the~normal~queue~unless~otherwise~directed~queue=\char`\"{}-q~normal\char`\"{}~
-
-if~{[}~\$\#~-eq~1~];~then
-
-~~~~~~~~echo~\char`\"{}Setting~the~queue~to~\$1\char`\"{}
-
-~~~~~~~~queue=\char`\"{}-q~\$1\char`\"{}~
-
-fi~\\
-~\\
-\#~compile~the~mesher~and~the~solver~
-
-d=`date'~echo~\char`\"{}Starting~compilation~\$d\char`\"{}~
+\#~use~the~normal~queue~unless~otherwise~directed~queue=\textquotedbl{}-q~normal\textquotedbl{}~
+
+if~{[}~\$\#~-eq~1~{]};~then
+
+~~~~~~~~echo~\textquotedbl{}Setting~the~queue~to~\$1\textquotedbl{}
+
+~~~~~~~~queue=\textquotedbl{}-q~\$1\textquotedbl{}~
+
+fi~~\\
+~~~\\
+~\#~compile~the~mesher~and~the~solver~
+
+d=`date'~echo~\textquotedbl{}Starting~compilation~\$d\textquotedbl{}~
 
@@
 
-echo~\char`\"{}Finished~compilation~\$d\char`\"{}~\\
-~\\
-\#~get~total~number~of~nodes~needed~for~solver~
-
-NPROC=`grep~NPROC~DATA/Par\_file~|~cut~-c~34-~'~\\
-~\\
+echo~\textquotedbl{}Finished~compilation~\$d\textquotedbl{}~~\\
+~~~\\
+~\#~get~total~number~of~nodes~needed~for~solver~
+
+NPROC=`grep~NPROC~DATA/Par\_file~|~cut~-c~34-~'~~\\
+~~~\\
 
@@
 
-NPROC\_XI=`grep~NPROC\_XI~DATA/meshfem3D\_files/Mesh\_Par\_file~|~cut~-c~34-~'~
-~\\
-
-NPROC\_ETA=`grep~NPROC\_ETA~DATA/meshfem3D\_files/Mesh\_Par\_file~|~cut~-c~34-~'~
-~\\
-\#~total~number~of~nodes~is~the~product~of~the~values~read~
-
-numnodes=\$((~\$NPROC\_XI~*~\$NPROC\_ETA~))~\\
-~\\
+NPROC\_XI=`grep~NPROC\_XI~DATA/meshfem3D\_files/Mesh\_Par\_file~|~cut~-c~34-~'~~~~\\
+
+NPROC\_ETA=`grep~NPROC\_ETA~DATA/meshfem3D\_files/Mesh\_Par\_file~|~cut~-c~34-~'~~~~\\
+~\#~total~number~of~nodes~is~the~product~of~the~values~read~
+
+numnodes=\$((~\$NPROC\_XI~{*}~\$NPROC\_ETA~))~~\\
+~~~\\
 
@@
 
-if~{[}~\$numnodes~-neq~\$NPROC~];~then
-
-~~~~~~~~echo~\char`\"{}error number of procs mismatch\char`\"{}
+if~{[}~\$numnodes~-neq~\$NPROC~{]};~then
+
+~~~~~~~~echo~\textquotedbl{}error~number~of~procs~mismatch\textquotedbl{}
 
@@
 
-fi~\\
-
-~\\
-echo~\char`\"{}Submitting~job\char`\"{}~
+fi~~\\
+
+~~\\
+~echo~\textquotedbl{}Submitting~job\textquotedbl{}~
 
@@
 compute nodes to the job, the \texttt{run\_lsf.bash} commands used
-to run the mesher, database generator and the solver, and calls to scripts that collect
-the output seismograms from the compute nodes and perform clean-up
-operations.
-
+to run the mesher, database generator and the solver, and calls to
+scripts that collect the output seismograms from the compute nodes
+and perform clean-up operations.
 \begin{enumerate}
@@
 output from \texttt{stdout} into ~\\
-\texttt{\small OUTPUT\_FILES/\%J.o},
-where \texttt{\%J} is short-hand for the job-ID; it also tells the
-scheduler what version of \texttt{mpich} to use (\texttt{mpich\_gm})
-and how to name this job (\texttt{go\_mesher\_solver\_lsf}).
+ \texttt{\small OUTPUT\_FILES/\%J.o}{\small , where }\texttt{\small \%J}{\small{}
+is short-hand for the job-ID; it also tells the scheduler what version
+of }\texttt{\small mpich}{\small{} to use (}\texttt{\small mpich\_gm}{\small )
+and how to name this job (}\texttt{\small go\_mesher\_solver\_lsf}{\small ). }{\small \par}
 \item The script then creates a list of the nodes allocated to this job
@@
 and parsing the output into a one-column list using the Perl script
-\texttt{utils/Cluster/lsf/remap\_lsf\_machines.pl}. It then creates a set of scratch
-directories on these nodes (\texttt{\small /scratch/}~\\
-\texttt{\small \$USER/DATABASES\_MPI}) to be used as the \texttt{LOCAL\_PATH}
-for temporary storage of the database files. The scratch directories
-are created using \texttt{shmux}, a shell multiplexor that can execute
-the same commands on many hosts in parallel. \texttt{shmux} is available
-from Shmux \urlwithparentheses{web.taranis.org/shmux/}. Make sure that the \texttt{LOCAL\_PATH}
-parameter in \texttt{DATA/Par\_file} is also set properly.
-\item The next portion of the script launches the mesher, database generator and then the solver
-using \texttt{run\_lsf.bash}.
+\texttt{utils/Cluster/lsf/remap\_lsf\_machines.pl}. It then creates
+a set of scratch directories on these nodes (\texttt{\small /scratch/}{\small ~}\\
+{\small{} }\texttt{\small \$USER/DATABASES\_MPI}{\small ) to be used
+as the }\texttt{\small LOCAL\_PATH}{\small{} for temporary storage
+of the database files. The scratch directories are created using }\texttt{\small shmux}{\small ,
+a shell multiplexor that can execute the same commands on many hosts
+in parallel. }\texttt{\small shmux}{\small{} is available from Shmux
+\urlwithparentheses{web.taranis.org/shmux/}. Make sure that the
+}\texttt{\small LOCAL\_PATH}{\small{} parameter in }\texttt{\small DATA/Par\_file}{\small{}
+is also set properly. }{\small \par}
+\item The next portion of the script launches the mesher, database generator
+and then the solver using \texttt{run\_lsf.bash}. 
 \item The final portion of the script collects the seismograms and performs
 clean up on the nodes, using the Perl scripts \texttt{collect\_seismo\_lsf\_multi.pl}
-and \texttt{cleanmulti.pl}.
-\end{enumerate}
+and \texttt{cleanmulti.pl}. \end{enumerate}
 \begin{lyxcode}
@@
 
-\#BSUB~-J~go\_mesher\_solver\_lsf~\\
-~\\
-\#~set~up~local~scratch~directories
+\#BSUB~-J~go\_mesher\_solver\_lsf~~\\
+~~~\\
+~\#~set~up~local~scratch~directories
 
@@
 
-echo~\char`\"{}\$LSB\_MCPU\_HOSTS\char`\"{}~>~OUTPUT\_FILES/lsf\_machines~
-
-echo~\char`\"{}\$LSB\_JOBID\char`\"{}~>~OUTPUT\_FILES/jobid
-
-remap\_lsf\_machines.pl~OUTPUT\_FILES/lsf\_machines~> OUTPUT\_FILES/machines
-
-shmux~-M50~-Sall~-c~\char`\"{}rm~-r~-f~/scratch/\$USER;~\textbackslash{}
-
-~~~~~~mkdir~-p~/scratch/\$USER;~mkdir~-p~\$BASEMPIDIR\char`\"{}~\textbackslash{}
-
-~~~~~~-~<~OUTPUT\_FILES/machines~>/dev/null~\\
-~\\
-\#~run~the~specfem~program
-
-current\_pwd=\$PWD \\
-
-cd bin/ \\
-
-run\_lsf.bash~-{}-gm-no-shmem~-{}-gm-copy-env~\$current\_pwd/xmeshfem3D~\\
-
-run\_lsf.bash~-{}-gm-no-shmem~-{}-gm-copy-env~\$current\_pwd/xgenerate\_databases~\\
-
-run\_lsf.bash~-{}-gm-no-shmem~-{}-gm-copy-env~\$current\_pwd/xspecfem3D~\\
-~\\
-\#~collect~seismograms~and~clean~up
-
-cd current\_pwd/~ \\
+echo~\textquotedbl{}\$LSB\_MCPU\_HOSTS\textquotedbl{}~>~OUTPUT\_FILES/lsf\_machines~
+
+echo~\textquotedbl{}\$LSB\_JOBID\textquotedbl{}~>~OUTPUT\_FILES/jobid
+
+remap\_lsf\_machines.pl~OUTPUT\_FILES/lsf\_machines~>~OUTPUT\_FILES/machines
+
+shmux~-M50~-Sall~-c~\textquotedbl{}rm~-r~-f~/scratch/\$USER;~\textbackslash{}
+
+~~~~~~mkdir~-p~/scratch/\$USER;~mkdir~-p~\$BASEMPIDIR\textquotedbl{}~\textbackslash{}
+
+~~~~~~-~<~OUTPUT\_FILES/machines~>/dev/null~~\\
+~~~\\
+~\#~run~the~specfem~program
+
+current\_pwd=\$PWD~~\\
+
+cd~bin/~~\\
+
+run\_lsf.bash~-{}-gm-no-shmem~-{}-gm-copy-env~\$current\_pwd/xmeshfem3D~~\\
+
+run\_lsf.bash~-{}-gm-no-shmem~-{}-gm-copy-env~\$current\_pwd/xgenerate\_databases~~\\
+
+run\_lsf.bash~-{}-gm-no-shmem~-{}-gm-copy-env~\$current\_pwd/xspecfem3D~~\\
+~~~\\
+~\#~collect~seismograms~and~clean~up
+
+cd~current\_pwd/~~~\\
 
@@
 
-In this section we explain how to change the velocity model used for your simulations.
-These changes involve contributing specific subroutines
-that replace existing subroutines in the \texttt{SPECFEM3D Cartesian} package.
-Note that \texttt{SPECFEM3D Cartesian} can handle Earth models with material properties that vary within each spectral element.
+In this section we explain how to change the velocity model used for
+your simulations. These changes involve contributing specific subroutines
+that replace existing subroutines in the \texttt{SPECFEM3D Cartesian}
+package. Note that \texttt{SPECFEM3D Cartesian} can handle Earth models
+with material properties that vary within each spectral element.
 
 %% magnoni 6/12
-\section{{\normalsize \label{sec:Using-tomographic}}Using external tomographic Earth models}
-
-To implement your own external tomographic model(s), you must provide your own external tomography file(s),
-and choose between two possible options:\\
-\indent (1) set in the \texttt{Par\_file} the model parameter \texttt{MODEL = tomo}, \\
+
+\section{{\normalsize \label{sec:Using-tomographic}}Using external tomographic
+Earth models}
+
+To implement your own external tomographic model(s), you must provide
+your own external tomography file(s), and choose between two possible
+options:\\
+ \indent (1) set in the \texttt{Par\_file} the model parameter \texttt{MODEL
+= tomo}, \\
  or for more user control: \\
-\indent (2) set in the \texttt{Par\_file} the model parameter \texttt{MODEL = default}, define the negative \texttt{material\_ID} identifier for each element in the file \texttt{MESH/materials\_file}
-and use the following format in the file \texttt{MESH/nummaterial\_velocity\_file}
+ \indent (2) set in the \texttt{Par\_file} the model parameter \texttt{MODEL
+= default}, define the negative \texttt{material\_ID} identifier for
+each element in the file \texttt{MESH/materials\_file} and use the
+following format in the file \texttt{MESH/nummaterial\_velocity\_file}
 when using CUBIT to construct your mesh (see also section \ref{subsec:Exporting-the-Mesh}): 
 \begin{lyxcode}
-\texttt{domain\_ID} \texttt{material\_ID} \texttt{tomography} \texttt{elastic} \texttt{file\_name} 1
+domain\_ID~material\_ID~tomography~elastic~file\_name~1
 \end{lyxcode}
 where: \\
-\indent \texttt{domain\_ID} is 1 for acoustic or 2 for elastic materials, \\
-\indent \texttt{material\_ID} a negative, unique identifier (i.e., -1,-2,...), \\
-\indent \texttt{tomography} keyword for tomographic material definition, \\
-\indent \texttt{elastic} keyword for elastic material definition, \\
-\indent \texttt{file\_name} the name of the tomography file and 1 a positive unique identifier.\\
-
-The external tomographic model is represented by a grid of points with assigned material properties and homogeneous resolution along each spatial direction x, y and z.
-The ASCII file \texttt{file\_name} that describe the tomography should be located in the \texttt{TOMOGRAPHY\_PATH} directory, set in the \texttt{Par\_file}.
-The format of the file, as read from \texttt{model\_tomography.f90} located in the \texttt{src/generate\_databases} directory, looks like Figure \ref{fig:tomography_file},
-where
-%
+ \indent \texttt{domain\_ID} is 1 for acoustic or 2 for elastic materials,
+\\
+ \indent \texttt{material\_ID} a negative, unique identifier (i.e.,
+-1,-2,...), \\
+ \indent \texttt{tomography} keyword for tomographic material definition,
+\\
+ \indent \texttt{elastic} keyword for elastic material definition,
+\\
+ \indent \texttt{file\_name} the name of the tomography file and
+1 a positive unique identifier.\\
+
+The external tomographic model is represented by a grid of points
+with assigned material properties and homogeneous resolution along
+each spatial direction x, y and z. The ASCII file \texttt{file\_name}
+that describe the tomography should be located in the \texttt{TOMOGRAPHY\_PATH}
+directory, set in the \texttt{Par\_file}. The format of the file,
+as read from \texttt{model\_tomography.f90} located in the \texttt{src/generate\_databases}
+directory, looks like Figure \ref{fig:tomography_file}, where 
 \begin{description}
-\item [{\texttt{ORIG\_X}, \texttt{END\_X}}] are, respectively, the coordinates of the initial and final tomographic grid points along the x direction (in the mesh units, e.g., $m$);
-\item [{\texttt{ORIG\_Y}, \texttt{END\_Y}}] respectively the coordinates of the initial and final tomographic grid points along the y direction (in the mesh units, e.g., $m$);
-\item [{\texttt{ORIG\_Z}, \texttt{END\_Z}}] respectively the coordinates of the initial and final tomographic grid points along the z direction (in the mesh units, e.g., $m$);
-\item [\texttt{SPACING\_X}, \texttt{SPACING\_Y}, \texttt{SPACING\_Z}] the spacing between the tomographic grid points along the x, y and z directions, respectively (in the mesh units, e.g., $m$);
-\item [\texttt{NX}, \texttt{NY}, \texttt{NZ}] the number of grid points along the spatial directions x, y and z, respectively; \texttt{NX} is given by [(\texttt{END\_X} - \texttt{ORIG\_X})/\texttt{SPACING\_X}]+1; \texttt{NY} and \texttt{NZ} are the same as \texttt{NX}, but for the y and z directions, respectively;
-\item [\texttt{VP\_MIN}, \texttt{VP\_MAX}, \texttt{VS\_MIN}, \texttt{VS\_MAX}, \texttt{RHO\_MIN}, \texttt{RHO\_MAX}] the minimum and maximum values of the wave speed \texttt{vp} and \texttt{vs} (in $m \, s^{-1}$) and of the density \texttt{rho} (in $kg \, m^{-3}$); these values could be the actual limits of the tomographic parameters in the grid or the minimum and maximum values to which we force the cut of velocity and density in the model.
+\item [{\texttt{ORIG\_X}, \texttt{END\_X}}] are, respectively, the coordinates
+of the initial and final tomographic grid points along the x direction
+(in the mesh units, e.g., $m$); 
+\item [{\texttt{ORIG\_Y}, \texttt{END\_Y}}] respectively the coordinates
+of the initial and final tomographic grid points along the y direction
+(in the mesh units, e.g., $m$); 
+\item [{\texttt{ORIG\_Z}, \texttt{END\_Z}}] respectively the coordinates
+of the initial and final tomographic grid points along the z direction
+(in the mesh units, e.g., $m$); 
+\item [{\texttt{SPACING\_X}, \texttt{SPACING\_Y}, \texttt{SPACING\_Z}}] the
+spacing between the tomographic grid points along the x, y and z directions,
+respectively (in the mesh units, e.g., $m$); 
+\item [{\texttt{NX}, \texttt{NY}, \texttt{NZ}}] the number of grid points
+along the spatial directions x, y and z, respectively; \texttt{NX}
+is given by {[}(\texttt{END\_X} - \texttt{ORIG\_X})/\texttt{SPACING\_X}{]}+1;
+\texttt{NY} and \texttt{NZ} are the same as \texttt{NX}, but for the
+y and z directions, respectively; 
+\item [{\texttt{VP\_MIN}, \texttt{VP\_MAX}, \texttt{VS\_MIN}, \texttt{VS\_MAX}, \texttt{RHO\_MIN}, \texttt{RHO\_MAX}}] the
+minimum and maximum values of the wave speed \texttt{vp} and \texttt{vs}
+(in $m\, s^{-1}$) and of the density \texttt{rho} (in $kg\, m^{-3}$);
+these values could be the actual limits of the tomographic parameters
+in the grid or the minimum and maximum values to which we force the
+cut of velocity and density in the model. 
 \end{description}
-%
-After the first four lines, in the file \texttt{file\_name} the tomographic grid points are listed with the corresponding values of \texttt{vp}, \texttt{vs} and \texttt{rho}, scanning the grid along the x coordinate (from \texttt{ORIG\_X} to \texttt{END\_X} with step of \texttt{SPACING\_X}) for each given y (from \texttt{ORIG\_Y} to \texttt{END\_Y}, with step of \texttt{SPACING\_Y}) and each given z (from \texttt{ORIG\_Z} to \texttt{END\_Z}, with step of \texttt{SPACING\_Z}).
-%fig
+After the first four lines, in the file \texttt{file\_name} the tomographic
+grid points are listed with the corresponding values of \texttt{vp},
+\texttt{vs} and \texttt{rho}, scanning the grid along the x coordinate
+(from \texttt{ORIG\_X} to \texttt{END\_X} with step of \texttt{SPACING\_X})
+for each given y (from \texttt{ORIG\_Y} to \texttt{END\_Y}, with step
+of \texttt{SPACING\_Y}) and each given z (from \texttt{ORIG\_Z} to
+\texttt{END\_Z}, with step of \texttt{SPACING\_Z}). %fig
 \begin{figure}[htbp]
 \noindent \begin{centering}
-\includegraphics[width=1.0\textwidth]{figures/tomo_file.jpg}
+\includegraphics[width=1\textwidth]{figures/tomo_file} 
 \par\end{centering}
-\caption{Tomography file \texttt{file\_name} that describes an external Earth model.
-The coordinates x, y and z of the grid, their limits ORIG\_X, ORIG\_Y, ORIG\_Z, END\_X, END\_Y, END\_Z and the grid spacings SPACING\_X, SPACING\_Y, SPACING\_Z should be in the units of the constructed mesh (e.g., UTM coordinates in meters).}
+
+\caption{Tomography file \texttt{file\_name} that describes an external Earth
+model. The coordinates x, y and z of the grid, their limits ORIG\_X,
+ORIG\_Y, ORIG\_Z, END\_X, END\_Y, END\_Z and the grid spacings SPACING\_X,
+SPACING\_Y, SPACING\_Z should be in the units of the constructed mesh
+(e.g., UTM coordinates in meters).}
+
 \label{fig:tomography_file} 
@@
 
-The user can implement his own interpolation algorithm for the tomography model by changing the routine \texttt{model\_tomography.f90} located in the \texttt{src/generate\_databases/} directory.
-Moreover, for models that involve both fully defined materials and a tomography description, the \texttt{nummaterial\_velocity\_file}
-has multiple lines each with the corresponding suitable format described above.
+The user can implement his own interpolation algorithm for the tomography
+model by changing the routine \texttt{model\_tomography.f90} located
+in the \texttt{src/generate\_databases/} directory. Moreover, for
+models that involve both fully defined materials and a tomography
+description, the \texttt{nummaterial\_velocity\_file} has multiple
+lines each with the corresponding suitable format described above.
 
@@
 
-To use your own external model, you can set in the \texttt{Par\_file} the model parameter \texttt{MODEL = external}.
-Three-dimensional acoustic and/or (an)elastic (attenuation) models may then be superimposed
-onto the mesh based upon your subroutine in
-\texttt{model\_external\_values.f90} located in directory \texttt{src/generate\_databases}.
-The call to this routine would be as follows
-
-\begin{lyxcode}
-call~model\_external\_values(xmesh,~ymesh,~zmesh, ~\& \\
-~~~~~~~~~~rho,~vp,~vs,~qmu\_atten,~iflag\_aniso,~idomain\_id)
-\end{lyxcode}
-%
+To use your own external model, you can set in the \texttt{Par\_file}
+the model parameter \texttt{MODEL = external}. Three-dimensional acoustic
+and/or (an)elastic (attenuation) models may then be superimposed onto
+the mesh based upon your subroutine in \texttt{model\_external\_values.f90}
+located in directory \texttt{src/generate\_databases}. The call to
+this routine would be as follows
+\begin{lyxcode}
+call~model\_external\_values(xmesh,~ymesh,~zmesh,~~\&~~\\
+~~~~~~~~~~~rho,~vp,~vs,~qmu\_atten,~iflag\_aniso,~idomain\_id)
+\end{lyxcode}
 Input to this routine consists of: 
@@
 \end{description}
-%
 Output to this routine consists of: 
 \begin{description}
-\item [{\texttt{rho,vp,vs}}] isotropic model parameters for density $\rho$ ($kg/m^3$), $v_p$ ($m/s$) and $v_s$ ($m/s$)
+\item [{\texttt{rho,vp,vs}}] isotropic model parameters for density $\rho$
+($kg/m^{3}$), $v_{p}$ ($m/s$) and $v_{s}$ ($m/s$) 
 \item [{\texttt{qmu\_atten}}] Shear wave quality factor: $0<Q_{\mu}<9000$ 
-\item [{\texttt{iflag\_aniso}}] anisotropic model flag, $0$ indicating no anisotropy or $1$ using anisotropic model parameters as defined in routine file \texttt{model\_aniso.f90}
-\item [{\texttt{idomain\_id}}] domain identifier, $1$ for acoustic, $2$ for elastic, $3$ for poroelastic materials.
+\item [{\texttt{iflag\_aniso}}] anisotropic model flag, $0$ indicating
+no anisotropy or $1$ using anisotropic model parameters as defined
+in routine file \texttt{model\_aniso.f90} 
+\item [{\texttt{idomain\_id}}] domain identifier, $1$ for acoustic, $2$
+for elastic, $3$ for poroelastic materials. 
 \end{description}
-
 Note that the resolution and maximum value of anelastic models are
@@
 
-To use your anisotropic models, you can either set in the \texttt{Par\_file} the model parameter \texttt{ANISOTROPY} to \texttt{.true.} or set the model parameter \texttt{MODEL = aniso}.
-Three-dimensional anisotropic models may be superimposed on
-the mesh based upon the subroutine in file \texttt{model\_aniso.f90}
-located in directory \texttt{src/generate\_databases}.
+To use your anisotropic models, you can either set in the \texttt{Par\_file}
+the model parameter \texttt{ANISOTROPY} to \texttt{.true.} or set
+the model parameter \texttt{MODEL = aniso}. Three-dimensional anisotropic
+models may be superimposed on the mesh based upon the subroutine in
+file \texttt{model\_aniso.f90} located in directory \texttt{src/generate\_databases}.
 The call to this subroutine is of the form
-
-\begin{lyxcode}
-call~model\_aniso(iflag\_aniso,~rho,~vp,~vs,~\& \\
-~~~~~~~~~~c11,c12,c13,c14,c15,c16,c22,c23,c24,c25,c26,~\& \\
-~~~~~~~~~~c33,c34,c35,c36,c44,c45,c46,c55,c56,c66)~
-\end{lyxcode}
-%
+\begin{lyxcode}
+call~model\_aniso(iflag\_aniso,~rho,~vp,~vs,~\&~~\\
+~~~~~~~~~~~c11,c12,c13,c14,c15,c16,c22,c23,c24,c25,c26,~\&~~\\
+~~~~~~~~~~~c33,c34,c35,c36,c44,c45,c46,c55,c56,c66)~
+\end{lyxcode}
 Input to this routine consists of: 
 \begin{description}
-\item [{\texttt{iflag\_aniso}}] flag indicating the type of the anisotropic model, i.e. $0$ for no superimposed anisotropy, $1$ or $2$ for generic pre-defined anisotropic models.
-\item [{\texttt{rho,vp,vs}}] reference isotropic model parameters for density $\rho$, $v_p$ and $v_s$.
+\item [{\texttt{iflag\_aniso}}] flag indicating the type of the anisotropic
+model, i.e. $0$ for no superimposed anisotropy, $1$ or $2$ for
+generic pre-defined anisotropic models. 
+\item [{\texttt{rho,vp,vs}}] reference isotropic model parameters for density
+$\rho$, $v_{p}$ and $v_{s}$. 
 \end{description}
-%
 Output from the routine consists of the following non-dimensional
@@
 \end{description}
-
-You can replace the \texttt{model\_aniso.f90} file by
-your own version \textit{provided you do not change the call structure
-of the routine}, i.e., the new routine should take exactly the same
-input and produce the required relative output.
+You can replace the \texttt{model\_aniso.f90} file by your own version
+\textit{provided you do not change the call structure of the routine},
+i.e., the new routine should take exactly the same input and produce
+the required relative output.
 
@@
 systems, for example, the path of the executable programs. Here we
-only list a few of the available scripts and provide a brief description, and
-you can either refer to the related sections for detailed usage or,
-in a lot of cases, type the script/program name without arguments
+only list a few of the available scripts and provide a brief description,
+and you can either refer to the related sections for detailed usage
+or, in a lot of cases, type the script/program name without arguments
 for its usage.
@@
 For such comparisons, the following steps are recommended:
-
 \begin{enumerate}
-\item Make sure that both synthetic and observed seismograms have the correct station/event and timing information.
-\item Convolve synthetic seismograms with a source time function with the half duration specified in the \texttt{CMTSOLUTION} file, provided, as recommended, you used a zero half duration in the SEM simulations.
-\item Resample both observed and synthetic seismograms to a common sampling rate.
+\item Make sure that both synthetic and observed seismograms have the correct
+station/event and timing information. 
+\item Convolve synthetic seismograms with a source time function with the
+half duration specified in the \texttt{CMTSOLUTION} file, provided,
+as recommended, you used a zero half duration in the SEM simulations. 
+\item Resample both observed and synthetic seismograms to a common sampling
+rate. 
 \item Cut the records using the same window. 
 \item Remove the trend and mean from the records and taper them. 
-\item Remove the instrument response from the observed seismograms (recommended) or convolve the synthetic seismograms with the instrument response.
-\item Make sure that you apply the same filters to both observed and synthetic seismograms. Preferably, avoid filtering your records more than once.
+\item Remove the instrument response from the observed seismograms (recommended)
+or convolve the synthetic seismograms with the instrument response. 
+\item Make sure that you apply the same filters to both observed and synthetic
+seismograms. Preferably, avoid filtering your records more than once. 
 \item Now, you are ready to compare your synthetic and observed seismograms. 
 \end{enumerate}
-
-\noindent
-We generally use the following scripts provided in the \texttt{utils/seis\_process/} directory:
+\noindent We generally use the following scripts provided in the \texttt{utils/seis\_process/}
+directory:
 
@@
 without any argument. An example of the usage of the script:
-
-\begin{lyxcode}
-{\footnotesize process\_data.pl~-m~CMTSOLUTION~-s~1.0~-l~0/4000~-i~-f~-t~40/500~-p~-x~bp~DATA/1999.330{*}.BH?.SAC}
-{\footnotesize \par}
-\end{lyxcode}
-
-\noindent which has resampled the SAC files to a sampling rate of 1 seconds, cut
-them between 0 and 4000 seconds, transfered them into displacement
-records and filtered them between 40 and 500 seconds,
-picked the first P and S arrivals, and added suffix `\texttt{bp}' to the file names.
+\begin{lyxcode}
+{\footnotesize process\_data.pl~-m~CMTSOLUTION~-s~1.0~-l~0/4000~-i~-f~-t~40/500~-p~-x~bp~DATA/1999.330{*}.BH?.SAC}~
+\end{lyxcode}
+\noindent which has resampled the SAC files to a sampling rate of
+1 seconds, cut them between 0 and 4000 seconds, transfered them into
+displacement records and filtered them between 40 and 500 seconds,
+picked the first P and S arrivals, and added suffix `\texttt{bp}'
+to the file names.
 
@@
 \begin{lyxcode}
-{\footnotesize process\_syn.pl~-m~CMTSOLUTION~-h~-a~STATIONS~-s~1.0~-l~0/4000~-f~-t~40/500~-p~-x~bp~SEM/{*}.BX?.semd}
-{\footnotesize \par}
+{\footnotesize process\_syn.pl~-m~CMTSOLUTION~-h~-a~STATIONS~-s~1.0~-l~0/4000~-f~-t~40/500~-p~-x~bp~SEM/{*}.BX?.semd}~
 \end{lyxcode}
@@
 from the \texttt{CMTSOLUTION} file, convert the synthetics into SAC
-format, add event and station information into the SAC headers, resample the SAC files with a sampling rate of 1 seconds, cut them between 0 and 4000 seconds, filter them between 40 and
-500 seconds with the same filter used for the observed data, pick the first P and S arrivals, and add the suffix `\texttt{bp}'
-to the file names.
+format, add event and station information into the SAC headers, resample
+the SAC files with a sampling rate of 1 seconds, cut them between
+0 and 4000 seconds, filter them between 40 and 500 seconds with the
+same filter used for the observed data, pick the first P and S arrivals,
+and add the suffix `\texttt{bp}' to the file names.
 
@@
 
-The original data and synthetics have three components: vertical (BHZ resp. BXZ),
-north (BHN resp. BXN) and east (BHE resp. BXE). However, for most seismology applications,
-transverse and radial components are also desirable. Therefore, we
-need to rotate the horizontal components of both the data and the
-synthetics to the transverse and radial direction, and \texttt{\small rotate.pl}
-can be used to accomplish this:
-
+The original data and synthetics have three components: vertical (BHZ
+resp. BXZ), north (BHN resp. BXN) and east (BHE resp. BXE). However,
+for most seismology applications, transverse and radial components
+are also desirable. Therefore, we need to rotate the horizontal components
+of both the data and the synthetics to the transverse and radial direction,
+and \texttt{\small rotate.pl}{\small{} can be used to accomplish this:}{\small \par}
 \begin{lyxcode}
@@
 The forward and adjoint simulations generate synthetic seismograms
-in the \texttt{OUTPUT\_FILES/} directory by default. For the forward simulation, the files
-are named like \texttt{STA.NT.BX?.semd} for two-column time series, or
-\texttt{STA.NT.BX?.semd.sac} for ASCII SAC format, where STA and NT
-are the station name and network code, and \texttt{BX?} stands for
-the component name. Please see the Appendix~\ref{cha:Coordinates} and \ref{cha:channel-codes}
-for further details.
-
-The adjont simulations generate synthetic seismograms
-with the name \texttt{S?????.NT.S??.sem} (refer to Section \ref{sec:Adjoint-simulation-sources}
+in the \texttt{OUTPUT\_FILES/} directory by default. For the forward
+simulation, the files are named like \texttt{STA.NT.BX?.semd} for
+two-column time series, or \texttt{STA.NT.BX?.semd.sac} for ASCII
+SAC format, where STA and NT are the station name and network code,
+and \texttt{BX?} stands for the component name. Please see the Appendix~\ref{cha:Coordinates}
+and \ref{cha:channel-codes} for further details.
+
+The adjont simulations generate synthetic seismograms with the name
+\texttt{S?????.NT.S??.sem} (refer to Section \ref{sec:Adjoint-simulation-sources}
 for details). The kernel simulations output the back-reconstructed
@@
 
-You do have further options to change this default output behavior, given in the main constants file
-\texttt{constants.h} located in \texttt{src/shared/} directory:
+You do have further options to change this default output behavior,
+given in the main constants file \texttt{constants.h} located in \texttt{src/shared/}
+directory: 
 \begin{description}
-\item [{\texttt{SEISMOGRAMS\_BINARY}}] set to \texttt{.true.} to have seismograms written out in binary format.
-\item [{\texttt{WRITE\_SEISMOGRAMS\_BY\_MASTER}}] Set to \texttt{.true.} to have only the master process writing out seismograms.
-This can be useful on a cluster, where only the master process node has access to the output directory.
-\item [{\texttt{USE\_OUTPUT\_FILES\_PATH}}] Set to \texttt{.false.} to have the seismograms output to \texttt{LOCAL\_PATH} directory specified in the main parameter file \texttt{DATA/Par\_file}.
-In this case, you could collect the synthetics onto the frontend using the \texttt{collect\_seismo\_lsf\_multi.pl} script
-located in the \texttt{utils/Cluster/lsf/} directory. The usage of the script would be e.g.:
+\item [{\texttt{SEISMOGRAMS\_BINARY}}] set to \texttt{.true.} to have seismograms
+written out in binary format. 
+\item [{\texttt{WRITE\_SEISMOGRAMS\_BY\_MASTER}}] Set to \texttt{.true.}
+to have only the master process writing out seismograms. This can
+be useful on a cluster, where only the master process node has access
+to the output directory. 
+\item [{\texttt{USE\_OUTPUT\_FILES\_PATH}}] Set to \texttt{.false.} to
+have the seismograms output to \texttt{LOCAL\_PATH} directory specified
+in the main parameter file \texttt{DATA/Par\_file}. In this case,
+you could collect the synthetics onto the frontend using the \texttt{collect\_seismo\_lsf\_multi.pl}
+script located in the \texttt{utils/Cluster/lsf/} directory. The usage
+of the script would be e.g.:
 
@@
 \end{lyxcode}
-
 \end{description}
-where  \texttt{machines} is a file containing the node names and  \texttt{DATA/Par\_file} the parameter file used to extract the \texttt{LOCAL\_PATH} directory used for the simulation.
+where \texttt{machines} is a file containing the node names and \texttt{DATA/Par\_file}
+the parameter file used to extract the \texttt{LOCAL\_PATH} directory
+used for the simulation.
 
@@
 need to clean the local scratch disk for the next simulation. This
-is especially important in the case of kernel simulation,
-where very large files are generated for the absorbing boundaries
-to help with the reconstruction of the regular forward wavefield.
-A sample script is provided in \texttt{utils/}:
-
+is especially important in the case of kernel simulation, where very
+large files are generated for the absorbing boundaries to help with
+the reconstruction of the regular forward wavefield. A sample script
+is provided in \texttt{utils/}:
 \begin{lyxcode}
@@
 With the movie data saved in \texttt{OUTPUT\_FILES/} at the end of
-a movie simulation (\texttt{\small MOVIE\_SURFACE=.true.}), you can
-run the \texttt{`create\_movie\_shakemap\_AVS\_DX\_GMT}' code to convert these binary
-movie data into GMT xyz files for futher processing. A sample script
-\texttt{movie2gif.gmt.pl} is provided to do this conversion, and then
-plot the movie snapshots in GMT, for example:
-
+a movie simulation (\texttt{\small MOVIE\_SURFACE=.true.}{\small ),
+you can run the }\texttt{\small `create\_movie\_shakemap\_AVS\_DX\_GMT}{\small '
+code to convert these binary movie data into GMT xyz files for futher
+processing. A sample script }\texttt{\small movie2gif.gmt.pl}{\small{}
+is provided to do this conversion, and then plot the movie snapshots
+in GMT, for example:}{\small \par}
 \begin{lyxcode}
@@
 of a shakemap simulation ~\\
-(\texttt{CREATE\_SHAKEMAP=.true.}), you can
-also run \texttt{`create\_movie\_shakemap\_AVS\_DX\_GMT}' code to convert the binary
-shakemap data into GMT xyz files. A sample script \texttt{plot\_shakemap.gmt.pl}
-is provided to do this conversion, and then plot the shakemaps in
-GMT, for example:
-
+ (\texttt{CREATE\_SHAKEMAP=.true.}), you can also run \texttt{`create\_movie\_shakemap\_AVS\_DX\_GMT}'
+code to convert the binary shakemap data into GMT xyz files. A sample
+script \texttt{plot\_shakemap.gmt.pl} is provided to do this conversion,
+and then plot the shakemaps in GMT, for example:
 \begin{lyxcode}
@@
 
-One of the fundamental reasons for computing sensitivity kernels (Section~\ref{sec:Adjoint-simulation-finite}) is to use them within a tomographic inversion.
-In other words, use recorded seismograms, make measurements with synthetic seismograms, and use the misfit between the two to iteratively improve the model described by (at least) $V_{\rm p}$, $V_{\rm s}$, and $\rho$ as a function of space.
-
-Whatever misfit function you use for the tomographic inversion (several examples in \cite{TrTaLi05} and \cite{TrKoLi08}), you will weight the sensitivity kernels with measurements.
-Furthermore, you will use as many measurements (stations, components, time windows) as possible per event; hence, we call these composite kernels ``event kernels,'' which are volumetric fields representing the gradient of the misfit function with respect to one of the variables (\eg $V_{\rm s}$).
-The basic features of an adjoint-based tomographic inversion were illustrated in \citet{TaLiTr07} and \citet{TrKoLi08} using a conjugate-gradient algorithm; there are dozens of versions of gradient-based inversion algorithms that could alternatively be used.
-The tomographic inversion of \citet{TaLiMaTr09,TaLiMaTr2010} used SPECFEM3D Cartesian as well as several additional components which are also stored on the CIG svn server, described next.
-
-The directory containing utilities for tomographic inversion using SPECFEM3D Cartesian (or other packages that evaluate misfit functions and gradients) is here on the CIG svn server:
-%
-\begin{verbatim}
-/cig/seismo/3D/ADJOINT_TOMO/
+One of the fundamental reasons for computing sensitivity kernels (Section~\ref{sec:Adjoint-simulation-finite})
+is to use them within a tomographic inversion. In other words, use
+recorded seismograms, make measurements with synthetic seismograms,
+and use the misfit between the two to iteratively improve the model
+described by (at least) $V_{{\rm p}}$, $V_{{\rm s}}$, and $\rho$
+as a function of space.
+
+Whatever misfit function you use for the tomographic inversion (several
+examples in \citet{TrTaLi05} and \citet{TrKoLi08}), you will weight
+the sensitivity kernels with measurements. Furthermore, you will use
+as many measurements (stations, components, time windows) as possible
+per event; hence, we call these composite kernels ``event kernels,''
+which are volumetric fields representing the gradient of the misfit
+function with respect to one of the variables (\eg $V_{{\rm s}}$).
+The basic features of an adjoint-based tomographic inversion were
+illustrated in \citet{TaLiTr07} and \citet{TrKoLi08} using a conjugate-gradient
+algorithm; there are dozens of versions of gradient-based inversion
+algorithms that could alternatively be used. The tomographic inversion
+of \citet{TaLiMaTr09,TaLiMaTr2010} used SPECFEM3D Cartesian as well
+as several additional components which are also stored on the CIG
+svn server, described next.
+
+The directory containing utilities for tomographic inversion using
+SPECFEM3D Cartesian (or other packages that evaluate misfit functions
+and gradients) is here on the CIG svn server: \begin{verbatim} /cig/seismo/3D/ADJOINTTOMO/
 flexwin/ -- FLEXWIN algorithm for automated picking of time windows
-   measure_adj/ -- reads FLEXWIN output file and makes measurements,
-                   with the option for computing adjoint sources
-   iterate_adj/ -- various tools for iterative inversion
-                   (requires pre-computed "event kernels")
-\end{verbatim}
-%
-This directory also contains a brief \verb+README+ file indicating the role of the three subdirectories, \verb+flexwin+ \citep{Maggi2009}, \verb+measure_adj+, and \verb+iterate_adj+.
-The components for making the model update are there; however, there are no explicit rules for computing the model update, just as with any optimization problem.
-There are options for computing a conjugate gradient step, as well as a source subspace projection step.
-The workflow could use substantial ``stitching together,'' and we welcome suggestions;
-undoubtedly, the tomographic inversion procedure will improve with time.
-{\bf This is a work in progress.}
+measureadj/ -- reads FLEXWIN output file and makes measurements, with
+the option for computing adjoint sources iterateadj/ -- various tools
+for iterative inversion (requires pre-computed \textquotedbl{}event
+kernels\textquotedbl{}) \end{verbatim} This directory also contains
+a brief \verb+README+ file indicating the role of the three subdirectories,
+\verb+flexwin+ \citep{Maggi2009}, \verb+measure_adj+, and \verb+iterate_adj+.
+The components for making the model update are there; however, there
+are no explicit rules for computing the model update, just as with
+any optimization problem. There are options for computing a conjugate
+gradient step, as well as a source subspace projection step. The workflow
+could use substantial ``stitching together,'' and we welcome suggestions;
+undoubtedly, the tomographic inversion procedure will improve with
+time. \textbf{This is a work in progress.}
 
@@
 e-mail to the CIG Computational Seismology Mailing List \urlwithparentheses{cig-seismo@geodynamics.org}
-or Jeroen Tromp \urlwithparentheses{jtromp-AT-princeton.edu}, and/or use our online
-bug tracking system Roundup \urlwithparentheses{www.geodynamics.org/roundup}.
+or Jeroen Tromp \urlwithparentheses{jtromp-AT-princeton.edu}, and/or
+use our online bug tracking system Roundup \urlwithparentheses{www.geodynamics.org/roundup}.
 
 \addcontentsline{toc}{chapter}{Notes \& Acknowledgments}
+
 \chapter*{Notes \& Acknowledgments}
@@
 are based in part on software libraries from the Massachusetts Institute
-of Technology, Department of Mechanical Engineering (Cambridge, Massachusetts, USA).
-The non-structured global numbering software was provided by Paul
-F. Fischer (Brown University, Providence, Rhode Island, USA, now at Argonne National Laboratory, USA).
-
-OpenDX \urlwithparentheses{http://www.opendx.org} is open-source based on IBM Data
-Explorer, AVS \urlwithparentheses{http://www.avs.com} is a trademark of Advanced
-Visualization Systems, and ParaView \urlwithparentheses{http://www.paraview.org}
+of Technology, Department of Mechanical Engineering (Cambridge, Massachusetts,
+USA). The non-structured global numbering software was provided by
+Paul F. Fischer (Brown University, Providence, Rhode Island, USA,
+now at Argonne National Laboratory, USA).
+
+OpenDX \urlwithparentheses{http://www.opendx.org} is open-source
+based on IBM Data Explorer, AVS \urlwithparentheses{http://www.avs.com}
+is a trademark of Advanced Visualization Systems, and ParaView \urlwithparentheses{http://www.paraview.org}
 is an open-source visualization platform.
@@
 Please e-mail your feedback, questions, comments, and suggestions
-to Jeroen Tromp \urlwithparentheses{jtromp-AT-princeton.edu} or to the CIG Computational Seismology Mailing List \urlwithparentheses{cig-seismo@geodynamics.org}.
+to Jeroen Tromp \urlwithparentheses{jtromp-AT-princeton.edu} or
+to the CIG Computational Seismology Mailing List \urlwithparentheses{cig-seismo@geodynamics.org}.
 
 \addcontentsline{toc}{chapter}{Copyright}
+
 \chapter*{Copyright}
@@
 
-$\copyright$ Princeton University and CNRS / INRIA / University of Pau, July 2012
+$\copyright$ Princeton University and CNRS / INRIA / University of
+Pau, July 2012
 
@@
 
- MPI v 2.1, July 2012:
- Max Rietmann, Peter Messmer, Daniel Peter, Dimitri Komatitsch, Joseph Charles, Zhinan Xie: support for CUDA GPUs,
-better CFL stability for the Stacey absorbing conditions. \\
-
- MPI v. 2.0, November 2010:
- Dimitri Komatitsch, Nicolas Le Goff, Roland Martin and Pieyre Le Loher, University of Pau, France,
- Daniel Peter, Jeroen Tromp and the Princeton group of developers, Princeton University, USA,
- and Emanuele Casarotti, INGV Roma, Italy:
-  support for CUBIT meshes decomposed by SCOTCH;
-  much faster solver using Michel Deville's inlined matrix products.\\
-
- MPI v. 1.4 Dimitri Komatitsch, University of Pau, Qinya Liu and others, Caltech, September 2006:
-  better adjoint and kernel calculations, faster and better I/Os
-  on very large systems, many small improvements and bug fixes.\\
-
- MPI v. 1.3 Dimitri Komatitsch, University of Pau, and Qinya Liu, Caltech, July 2005:
-  serial version, regular mesh, adjoint and kernel calculations, ParaView support.\\
-
- MPI v. 1.2 Min Chen and Dimitri Komatitsch, Caltech, July 2004:
-  full anisotropy, volume movie.\\
-
- MPI v. 1.1 Dimitri Komatitsch, Caltech, October 2002: Zhu's Moho map, scaling
-  of $V_s$ with depth, Hauksson's regional model, attenuation, oceans, movies.\\
-
- MPI v. 1.0 Dimitri Komatitsch, Caltech, USA, May 2002: first MPI version based on global code.\\
-
- Dimitri Komatitsch, IPG Paris, France, December 1996: first 3-D solver for the CM-5 Connection Machine,
-    parallelized on 128 processors using Connection Machine Fortran.\\
-
-\addcontentsline{toc}{chapter}{References}
-\bibliography{bibliography}
+MPI v 2.1, July 2012: Max Rietmann, Peter Messmer, Daniel Peter, Dimitri
+Komatitsch, Joseph Charles, Zhinan Xie: support for CUDA GPUs, better
+CFL stability for the Stacey absorbing conditions. \\
+
+MPI v. 2.0, November 2010: Dimitri Komatitsch, Nicolas Le Goff, Roland
+Martin and Pieyre Le Loher, University of Pau, France, Daniel Peter,
+Jeroen Tromp and the Princeton group of developers, Princeton University,
+USA, and Emanuele Casarotti, INGV Roma, Italy: support for CUBIT meshes
+decomposed by SCOTCH; much faster solver using Michel Deville's inlined
+matrix products.\\
+
+MPI v. 1.4 Dimitri Komatitsch, University of Pau, Qinya Liu and others,
+Caltech, September 2006: better adjoint and kernel calculations, faster
+and better I/Os on very large systems, many small improvements and
+bug fixes.\\
+
+MPI v. 1.3 Dimitri Komatitsch, University of Pau, and Qinya Liu, Caltech,
+July 2005: serial version, regular mesh, adjoint and kernel calculations,
+ParaView support.\\
+
+MPI v. 1.2 Min Chen and Dimitri Komatitsch, Caltech, July 2004: full
+anisotropy, volume movie.\\
+
+MPI v. 1.1 Dimitri Komatitsch, Caltech, October 2002: Zhu's Moho map,
+scaling of $V_{s}$ with depth, Hauksson's regional model, attenuation,
+oceans, movies.\\
+
+MPI v. 1.0 Dimitri Komatitsch, Caltech, USA, May 2002: first MPI version
+based on global code.\\
+
+Dimitri Komatitsch, IPG Paris, France, December 1996: first 3-D solver
+for the CM-5 Connection Machine, parallelized on 128 processors using
+Connection Machine Fortran.\\
+
+\addcontentsline{toc}{chapter}{References} \bibliography{bibliography}
 
 \appendix
+%dummy comment inserted by tex2lyx to ensure that this paragraph is not empty
 
@@
 frame:
-
 \begin{itemize}
@@
 The Aki \& Richards convention is
-
 \begin{itemize}
@@
 and the Harvard CMT convention is
-
 \begin{itemize}
@@
 The given locations for sources and receiver locations thus may get
-converted.  Sources and receiver locations are read in from the
-\texttt{CMTSOLUTION} (or \texttt{FORCESOLUTION}) and \texttt{STATIONS}
-files.  Note that e.g. the \texttt{CMTSOLUTION} and
-\texttt{FORCESOLUTION} files denotes the location by
-"longitude/latitude/depth". We read in longitude as $x$ coordinate,
-latitude as $y$ coordinate.
+converted. Sources and receiver locations are read in from the \texttt{CMTSOLUTION}
+(or \texttt{FORCESOLUTION}) and \texttt{STATIONS} files. Note that
+e.g. the \texttt{CMTSOLUTION} and \texttt{FORCESOLUTION} files denotes
+the location by \textquotedbl{}longitude/latitude/depth\textquotedbl{}.
+We read in longitude as $x$ coordinate, latitude as $y$ coordinate.
 
@@
 in the main parameter file (see Chapter~\ref{cha:Creating-Distributed-Databases}),
-the $x$/$y$ coordinates have to be given in degrees and are converted to Cartesian coordinates
-using a UTM conversion with the specified UTM zone.
-
-The value for depth (given in $km$ in \texttt{CMTSOLUTION} and
-\texttt{FORCESOLUTION}) or burial depth (given in $m$ in
-\texttt{STATIONS}) is evaluated with respect to the surface of the
-mesh at the specified $x$/$y$ location to find the corresponding $z$
-coordinate.  It is possible to use this depth value directly as $z$
-coordinate by changing the flag \texttt{USE\_SOURCES\_RECVS\_Z} to
-\texttt{.true.} in the file \texttt{constants.h} located in the
-\texttt{src/shared/} subdirectory.
+the $x$/$y$ coordinates have to be given in degrees and are converted
+to Cartesian coordinates using a UTM conversion with the specified
+UTM zone.
+
+The value for depth (given in $km$ in \texttt{CMTSOLUTION} and \texttt{FORCESOLUTION})
+or burial depth (given in $m$ in \texttt{STATIONS}) is evaluated
+with respect to the surface of the mesh at the specified $x$/$y$
+location to find the corresponding $z$ coordinate. It is possible
+to use this depth value directly as $z$ coordinate by changing the
+flag \texttt{USE\_SOURCES\_RECVS\_Z} to \texttt{.true.} in the file
+\texttt{constants.h} located in the \texttt{src/shared/} subdirectory.
 
@@
 
-The seismogram output directions are given in Cartesian $x$/$y$/$z$ directions and not rotated any further.
-Changing flags in \texttt{constants.h} in the \texttt{src/shared/} subdirectory only rotates the seismogram outputs
-if receivers are forced to be located at the surface (\texttt{RECVS\_CAN\_BE\_BURIED\_EXT\_MESH} set to \texttt{.false.}) and the normal to the surface at the receiver location should be used (\texttt{EXT\_MESH\_RECV\_NORMAL} set to
-\texttt{.true.}) as vertical.
-In this case, the outputs are rotated to have the vertical component normal to the surface of the mesh, $x$ and $y$ directions are somewhat arbitrary orthogonal directions along the surface.
-
-For the labeling of the seismogram channels, see Appendix~\ref{cha:channel-codes}. Additionally, we add labels to the synthetics using the following convention:
-For a receiver station located in an
+The seismogram output directions are given in Cartesian $x$/$y$/$z$
+directions and not rotated any further. Changing flags in \texttt{constants.h}
+in the \texttt{src/shared/} subdirectory only rotates the seismogram
+outputs if receivers are forced to be located at the surface (\texttt{RECVS\_CAN\_BE\_BURIED\_EXT\_MESH}
+set to \texttt{.false.}) and the normal to the surface at the receiver
+location should be used (\texttt{EXT\_MESH\_RECV\_NORMAL} set to \texttt{.true.})
+as vertical. In this case, the outputs are rotated to have the vertical
+component normal to the surface of the mesh, $x$ and $y$ directions
+are somewhat arbitrary orthogonal directions along the surface.
+
+For the labeling of the seismogram channels, see Appendix~\ref{cha:channel-codes}.
+Additionally, we add labels to the synthetics using the following
+convention: For a receiver station located in an 
 \begin{description}
-\item[elastic domain:] ~\\
+\item [{elastic domain:}] ~\\
+
 \begin{itemize}
@@
 \end{itemize}
-\item[acoustic domain:] ~\\
-(please note that receiver stations in acoustic domains must be buried. This is due to the free surface condition which enforces a zero pressure at the surface.)
+\item [{acoustic domain:}] ~\\
+ (please note that receiver stations in acoustic domains must be buried.
+This is due to the free surface condition which enforces a zero pressure
+at the surface.)
+
 \begin{itemize}
@@
 \item \texttt{semv} for the velocity vector 
-\item \texttt{sema} for pressure which will be stored in the vertical component \texttt{Z}.
-Note that pressure in the acoustic media is isotropic and thus the pressure record would be the same in the other two component directions.
-We therefore use the other two seismogram components to store the acoustic potential in component \texttt{X} (or \texttt{N}) and the first time derivative of the acoustic potential in component \texttt{Y} (or \texttt{E}).
+\item \texttt{sema} for pressure which will be stored in the vertical component
+\texttt{Z}. Note that pressure in the acoustic media is isotropic
+and thus the pressure record would be the same in the other two component
+directions. We therefore use the other two seismogram components to
+store the acoustic potential in component \texttt{X} (or \texttt{N})
+and the first time derivative of the acoustic potential in component
+\texttt{Y} (or \texttt{E}). 
 \end{itemize}
 \end{description}
-The seismograms are by default written out in ASCII-format to the \texttt{OUTPUT\_FILES/} subdirectory by each parallel process.
-You can change this behavior by changing the following flags in the \texttt{constants.h} file located in the \texttt{src/shared/} subdirectory:
+The seismograms are by default written out in ASCII-format to the
+\texttt{OUTPUT\_FILES/} subdirectory by each parallel process. You
+can change this behavior by changing the following flags in the \texttt{constants.h}
+file located in the \texttt{src/shared/} subdirectory: 
 \begin{description}
-\item [{\texttt{SEISMOGRAMS\_BINARY}}] set to \texttt{.true.} to have seismograms written out in binary format.
-\item [{\texttt{WRITE\_SEISMOGRAMS\_BY\_MASTER}}] Set to \texttt{.true.} to have only the master process writing out seismograms.
-This can be useful on a cluster, where only the master process node has access to the output directory.
-\item [{\texttt{USE\_OUTPUT\_FILES\_PATH}}] Set to \texttt{.false.} to have the seismograms output to \texttt{LOCAL\_PATH} directory specified in the main parameter file \texttt{DATA/Par\_file}.
+\item [{\texttt{SEISMOGRAMS\_BINARY}}] set to \texttt{.true.} to have seismograms
+written out in binary format. 
+\item [{\texttt{WRITE\_SEISMOGRAMS\_BY\_MASTER}}] Set to \texttt{.true.}
+to have only the master process writing out seismograms. This can
+be useful on a cluster, where only the master process node has access
+to the output directory. 
+\item [{\texttt{USE\_OUTPUT\_FILES\_PATH}}] Set to \texttt{.false.} to
+have the seismograms output to \texttt{LOCAL\_PATH} directory specified
+in the main parameter file \texttt{DATA/Par\_file}. 
 \end{description}
@@
 
-Seismic networks, such as the Global Seismographic Network (GSN), generally involve various types of instruments with different bandwidths, sampling properties and component configurations.
-There are standards to name channel codes depending on instrument properties.
-IRIS  \texttt{(www.iris.edu)} uses SEED format for channel codes, which are represented by three letters, such as \texttt{LHN}, \texttt{BHZ}, etc.
-In older versions of the SPECFEM3D Cartesian package, a common format was used for the channel codes of all seismograms, which was \texttt{BHE/BHN/BHZ} for three components.
-To avoid confusion when comparison are made to observed data, we are now using the FDSN convention (\url{http://www.fdsn.org/}) for SEM seismograms.
-In the following, we give a brief explanation of the FDSN convention used by IRIS, and how it is adopted in SEM seismograms.
-Please visit \texttt{www.iris.edu/manuals/SEED\_appA.htm} for further information.\\
-
-\noindent \textbf{\texttt{Band code:}} The first letter in the channel code denotes the band code of seismograms, which depends on the response band and the sampling rate of instruments.
-The list of band codes used by IRIS is shown in Figure \ref{fig:IRIS_band_codes}.
-The sampling rate of SEM synthetics is controlled by the resolution of simulations rather than instrument properties.
-However, for consistency, we follow the FDSN convention for SEM seismograms governed by their sampling rate.
-For SEM synthetics, we consider band codes for which $dt \leq 1$ s.
-IRIS also considers the response band of instruments.
-For instance, short-period and broad-band seismograms with the same sampling rate correspond to different band codes, such as S and B, respectively.
-In such cases, we consider SEM seismograms as broad band, ignoring the corner period ($\geq 10$ s) of the response band of instruments (note that at these resolutions, the minimum period in the SEM synthetics will be less than $10$ s).
-Accordingly, when
-you run a simulation the band code will be chosen depending on the resolution of the synthetics, and channel codes of SEM seismograms will start with either \texttt{L}, \texttt{M}, \texttt{B}, \texttt{H}, \texttt{C} or \texttt{F}, shown by red color in the figure.\\
+Seismic networks, such as the Global Seismographic Network (GSN),
+generally involve various types of instruments with different bandwidths,
+sampling properties and component configurations. There are standards
+to name channel codes depending on instrument properties. IRIS \texttt{(www.iris.edu)}
+uses SEED format for channel codes, which are represented by three
+letters, such as \texttt{LHN}, \texttt{BHZ}, etc. In older versions
+of the SPECFEM3D Cartesian package, a common format was used for the
+channel codes of all seismograms, which was \texttt{BHE/BHN/BHZ} for
+three components. To avoid confusion when comparison are made to observed
+data, we are now using the FDSN convention (\url{http://www.fdsn.org/})
+for SEM seismograms. In the following, we give a brief explanation
+of the FDSN convention used by IRIS, and how it is adopted in SEM
+seismograms. Please visit \texttt{www.iris.edu/manuals/SEED\_appA.htm}
+for further information.\\
+
+\noindent \texttt{\textbf{Band code:}} The first letter in the channel
+code denotes the band code of seismograms, which depends on the response
+band and the sampling rate of instruments. The list of band codes
+used by IRIS is shown in Figure \ref{fig:IRIS_band_codes}. The sampling
+rate of SEM synthetics is controlled by the resolution of simulations
+rather than instrument properties. However, for consistency, we follow
+the FDSN convention for SEM seismograms governed by their sampling
+rate. For SEM synthetics, we consider band codes for which $dt\leq1$
+s. IRIS also considers the response band of instruments. For instance,
+short-period and broad-band seismograms with the same sampling rate
+correspond to different band codes, such as S and B, respectively.
+In such cases, we consider SEM seismograms as broad band, ignoring
+the corner period ($\geq10$ s) of the response band of instruments
+(note that at these resolutions, the minimum period in the SEM synthetics
+will be less than $10$ s). Accordingly, when you run a simulation
+the band code will be chosen depending on the resolution of the synthetics,
+and channel codes of SEM seismograms will start with either \texttt{L},
+\texttt{M}, \texttt{B}, \texttt{H}, \texttt{C} or \texttt{F}, shown
+by red color in the figure.\\
 
@@
 \noindent \begin{centering}
-\includegraphics[scale=0.6]{figures/IRIS_band_codes.pdf}
+\includegraphics[scale=0.6]{figures/IRIS_band_codes} 
 \par\end{centering}
 
-\caption{The band code convention is based on the sampling rate and the response band of instruments.
-Please visit \texttt{www.iris.edu/manuals/SEED\_appA.htm} for further information.
-Grey rows show the relative band-code range in SPECFEM3D Cartesian, and the band codes used to name SEM seismograms are denoted in red.}
+\caption{The band code convention is based on the sampling rate and the response
+band of instruments. Please visit \texttt{www.iris.edu/manuals/SEED\_appA.htm}
+for further information. Grey rows show the relative band-code range
+in SPECFEM3D Cartesian, and the band codes used to name SEM seismograms
+are denoted in red.}
 
@@
 
-\noindent \textbf{\texttt{Instrument code:}} The second letter in the channel code corresponds to instrument codes, which specify the family to which the sensor belongs.
-For instance, \texttt{H} and \texttt{L} are used for high-gain and low-gain seismometers, respectively.
-The instrument code of SEM seismograms will always be \texttt{X}, as assigned by FDSN for synthetic seismograms.
-\\
-
-\noindent \textbf{\texttt{Orientation code:}} The third letter in channel codes is an orientation code, which generally describes the physical configuration of the components of instrument packages.
-SPECFEM3D Cartesian uses the traditional orientation code  \texttt{E/N/Z} (East-West, North-South, Vertical) for three components when a UTM projection is used.
-If the UTM conversion is suppressed, i.e.
-the flag \texttt{SUPPRESS\_UTM\_PROJECTION} is set to \texttt{.true.}, then the three components are labelled \texttt{X/Y/Z} according to the Cartesian reference frame. \\
-
-\noindent \textbf{\texttt{EXAMPLE:}} The sampling rate is given by \texttt{DT} in the main parameter file \texttt{Par\_file} located in the \texttt{DATA/} subdirectory.
-Depending on the resolution of your simulations, if you choose a sampling rate greater than $0.01$ s and less than $1$ s, a seismogram recording displacements on the vertical component of a station \texttt{ASBS} (network \texttt{AZ}) will be named \texttt{ASBS.AZ.MXZ.semd.sac}, whereas it will be \texttt{ASBS.AZ.BXZ.semd.sac}, if the sampling rate is greater than 0.0125 and less equal to 0.1 s.
+\noindent \texttt{\textbf{Instrument code:}} The second letter in
+the channel code corresponds to instrument codes, which specify the
+family to which the sensor belongs. For instance, \texttt{H} and \texttt{L}
+are used for high-gain and low-gain seismometers, respectively. The
+instrument code of SEM seismograms will always be \texttt{X}, as assigned
+by FDSN for synthetic seismograms. \\
+
+\noindent \texttt{\textbf{Orientation code:}} The third letter in
+channel codes is an orientation code, which generally describes the
+physical configuration of the components of instrument packages. SPECFEM3D
+Cartesian uses the traditional orientation code \texttt{E/N/Z} (East-West,
+North-South, Vertical) for three components when a UTM projection
+is used. If the UTM conversion is suppressed, i.e. the flag \texttt{SUPPRESS\_UTM\_PROJECTION}
+is set to \texttt{.true.}, then the three components are labelled
+\texttt{X/Y/Z} according to the Cartesian reference frame. \\
+
+\noindent \texttt{\textbf{EXAMPLE:}} The sampling rate is given by
+\texttt{DT} in the main parameter file \texttt{Par\_file} located
+in the \texttt{DATA/} subdirectory. Depending on the resolution of
+your simulations, if you choose a sampling rate greater than $0.01$
+s and less than $1$ s, a seismogram recording displacements on the
+vertical component of a station \texttt{ASBS} (network \texttt{AZ})
+will be named \texttt{ASBS.AZ.MXZ.semd.sac}, whereas it will be \texttt{ASBS.AZ.BXZ.semd.sac},
+if the sampling rate is greater than 0.0125 and less equal to 0.1
+s.
 
@@
 \section*{FAQ}
-
 \begin{description}
-\item [configuration fails:]
+\item [{configuration fails:}]~
 
@@
 
-   Please make sure that you have a working installation of a Fortran compiler,
-   a C compiler and an MPI implementation. You should be able to compile this
-   little program code:
-
-\begin{lyxcode}
-
-      program main ~\\
-        include 'mpif.h' ~\\
-        integer, parameter :: CUSTOM\_MPI\_TYPE = MPI\_REAL ~\\
-        integer ier ~\\
-        call MPI\_INIT(ier) ~\\
-        call MPI\_BARRIER(MPI\_COMM\_WORLD,ier) ~\\
-        call MPI\_FINALIZE(ier) ~\\
-      end
-
-\end{lyxcode}
-
-\item [compilation fails stating:] ~\\
-\begin{lyxcode}
-    ...  ~\\
-    obj/program\_generate\_databases.o: In function `MAIN\_\_':  ~\\
-    program\_generate\_databases.f90:(.text+0x14): undefined reference to `\_gfortran\_set\_std'  ~\\
-    ...  ~\\
+Please make sure that you have a working installation of a Fortran
+compiler, a C compiler and an MPI implementation. You should be able
+to compile this little program code:
+\begin{lyxcode}
+program~main~~~\\
+~include~'mpif.h'~~~\\
+~integer,~parameter~::~CUSTOM\_MPI\_TYPE~=~MPI\_REAL~~~\\
+~integer~ier~~~\\
+~call~MPI\_INIT(ier)~~~\\
+~call~MPI\_BARRIER(MPI\_COMM\_WORLD,ier)~~~\\
+~call~MPI\_FINALIZE(ier)~~~\\
+~end
+\end{lyxcode}
+\item [{compilation fails stating:}] ~\\
+
+\begin{lyxcode}
+...~~~\\
+~obj/program\_generate\_databases.o:~In~function~`MAIN\_\_':~~~\\
+~program\_generate\_databases.f90:(.text+0x14):~undefined~reference~to~`\_gfortran\_set\_std'~~~\\
+~...~~~\\
+
 \end{lyxcode}
@@
 
-  Normally, this message will appear when you are mixing two different Fortran
-  compilers. That is, using e.g. gfortran to compile non-MPI files
-  and mpif90, wrapper provided for e.g. ifort, to compile MPI-files.
+Normally, this message will appear when you are mixing two different
+Fortran compilers. That is, using e.g. gfortran to compile non-MPI
+files and mpif90, wrapper provided for e.g. ifort, to compile MPI-files.
 
@@
 
-\item [after executing \texttt{xmeshfem3D} I've got elements with skewness of 81\% percent, what does this mean:]
-Look at the skewness table printed in the \texttt{output\_mesher.txt} file after executing \texttt{xmeshfem3D} for the
-example given in \texttt{examples/meshfem3D\_examples/simple\_model/}:
-\begin{lyxcode}
-...~\\
-histogram of skewness (0. good - 1. bad): ~\\
- 0.0000000E+00  -   5.0000001E-02       27648     81.81818      \% ~\\
- 5.0000001E-02  -   0.1000000               0    0.0000000E+00  \% ~\\
-... ~\\
-\end{lyxcode}
-The first line means that you have 27,648 elements with a skewness value between 0 and 0.05 (which means the element is basically not skewed, just plain regular hexahedral element).
-The total number of elements you have in this mesh is (see in the \texttt{output\_mesher.txt} file a bit further down):
-\begin{lyxcode}
-...~\\
-total number of elements in entire mesh:        33792 ~\\
-... ~\\
-\end{lyxcode}
-which gives you that:  27,648 / 33,792 $\sim$ 81.8 \% of all elements are not skewed, i.e. regular elements. a fantastic value :)
-
-The histogram lists for this mesh also some stronger skewed elements, for example the worst ones belong to:
-\begin{lyxcode}
-...~\\
-0.6000000      -   0.6500000            2048     6.060606      \% ~\\
-... ~\\
-\end{lyxcode}
-
-about 6 \% of all elements have distortions with a skewness value between 0.6 and 0.65.
-The skewness values give you a hint of how good your mesh is.
-In an ideal world, you would want to have no distortions, just like the 81\% from above.
-Those elements give you the best approximate values by the GLL quadrature used in the spectral-element method.
-However, having weakly distorted elements is still fine and the solutions are still accurate enough.
-So empirically, values up to around 0.7 are tolerable, above that you should consider remeshing...
-
-To give you an idea why some of the elements are distorted, see the following figure \ref{fig:mesh.vp} of the mesh you obtain in the example \texttt{examples/meshfem3D\_examples/simple\_model/}.
+\item [{after executing \texttt{xmeshfem3D} I've got elements with skewness of 81\% percent, what does this mean:}] Look
+at the skewness table printed in the \texttt{output\_mesher.txt} file
+after executing \texttt{xmeshfem3D} for the example given in \texttt{examples/meshfem3D\_examples/simple\_model/}:
+
+\begin{lyxcode}
+...~~\\
+~histogram~of~skewness~(0.~good~-~1.~bad):~~~\\
+~0.0000000E+00~-~5.0000001E-02~27648~81.81818~\%~~~\\
+~5.0000001E-02~-~0.1000000~0~0.0000000E+00~\%~~~\\
+~...~~~\\
+
+\end{lyxcode}
+
+The first line means that you have 27,648 elements with a skewness
+value between 0 and 0.05 (which means the element is basically not
+skewed, just plain regular hexahedral element). The total number of
+elements you have in this mesh is (see in the \texttt{output\_mesher.txt}
+file a bit further down): 
+\begin{lyxcode}
+...~~\\
+~total~number~of~elements~in~entire~mesh:~33792~~~\\
+~...~~~\\
+
+\end{lyxcode}
+
+which gives you that: 27,648 / 33,792 $\sim$ 81.8 \% of all elements
+are not skewed, i.e. regular elements. a fantastic value :)
+
+The histogram lists for this mesh also some stronger skewed elements,
+for example the worst ones belong to: 
+\begin{lyxcode}
+...~~\\
+~0.6000000~-~0.6500000~2048~6.060606~\%~~~\\
+~...~~~\\
+
+\end{lyxcode}
+
+about 6 \% of all elements have distortions with a skewness value
+between 0.6 and 0.65. The skewness values give you a hint of how good
+your mesh is. In an ideal world, you would want to have no distortions,
+just like the 81\% from above. Those elements give you the best approximate
+values by the GLL quadrature used in the spectral-element method.
+However, having weakly distorted elements is still fine and the solutions
+are still accurate enough. So empirically, values up to around 0.7
+are tolerable, above that you should consider remeshing...
+
+To give you an idea why some of the elements are distorted, see the
+following figure \ref{fig:mesh.vp} of the mesh you obtain in the
+example \texttt{examples/meshfem3D\_examples/simple\_model/}. 
 \begin{figure}[htbp]
 \noindent \begin{centering}
-\includegraphics[width=.9\textwidth]{figures/mesh_vp.jpg}
+\includegraphics[width=0.9\textwidth]{figures/mesh_vp} 
 \par\end{centering}
-\caption{Paraview visualization using the mesh vtk-files for the example given in \texttt{examples/meshfem3D\_examples/simple\_model/}.}
+
+\caption{Paraview visualization using the mesh vtk-files for the example given
+in \texttt{examples/meshfem3D\_examples/simple\_model/}.}
+
 \label{fig:mesh.vp} 
 \end{figure}
-You will see that the mesh contains a doubling layer, where we stitch elements together such that the size of two elements will transition to the size of one element (very useful to keep the ratio of wavespeed / element\_size about constant).
-Those elements in this doubling layer have higher skewness values and make up those 6 \% in the histogram.
-
-\item [the code gives following error message "need at least one receiver":]
-This means that no stations given in the input file \texttt{DATA/STATIONS} could be located within the dimensions of the mesh.
-This can happen for example when the mesh was created with the in-house mesher \texttt{xmeshfem3D} while using the Universal Transverse Mercator (UTM) projection but the simulation with \texttt{xspecfem3D} was suppressing this projection from latitude/longitude to x/y/z coordinates.
-
-In such cases, try to change your \texttt{DATA/Par\_file} and set e.g.:
-\begin{lyxcode}
-SUPPRESS\_UTM\_PROJECTION         = .false. ~\\
-\end{lyxcode}
+
+You will see that the mesh contains a doubling layer, where we stitch
+elements together such that the size of two elements will transition
+to the size of one element (very useful to keep the ratio of wavespeed
+/ element\_size about constant). Those elements in this doubling layer
+have higher skewness values and make up those 6 \% in the histogram.
+
+\item [{the code gives following error message "need at least one receiver":}] This
+means that no stations given in the input file \texttt{DATA/STATIONS}
+could be located within the dimensions of the mesh. This can happen
+for example when the mesh was created with the in-house mesher \texttt{xmeshfem3D}
+while using the Universal Transverse Mercator (UTM) projection but
+the simulation with \texttt{xspecfem3D} was suppressing this projection
+from latitude/longitude to x/y/z coordinates.
+
+In such cases, try to change your \texttt{DATA/Par\_file} and set
+e.g.: 
+\begin{lyxcode}
+SUPPRESS\_UTM\_PROJECTION~=~.false.~~~\\
+
+\end{lyxcode}
+
 to be the same in \texttt{Mesh\_Par\_file} and \texttt{Par\_file}.
-This flag should be identical when using the in-house mesher \texttt{xmeshfem3D}, \texttt{xgenerate\_databases} and \texttt{xspecfem3D} together to run simulations.
-
-The flag determines if the coordinates you specify for your source and station locations are given as lat/lon degrees and
-must be converted to UTM coordinates.
-As an example, if you use \texttt{.false.} within \texttt{Mesh\_Par\_file} then you create a mesh with \texttt{xmeshfem3D} using the UTM projection from lat/lon as input format to UTM projected coordinates to store the mesh point positions, which is fine.
-The error then may occur if in the \texttt{Par\_file} you have this set to \texttt{.true.} so that the \texttt{xgenerate\_databases} and \texttt{xspecfem3D} suppress the UTM projection and assume that all coordinates you use now for source and receiver locations are given in meters (that is, converted) already.
-So it won't find the specified locations in the used mesh.
-As a solutions, just change the flag in \texttt{Par\_file} to be the same as in \texttt{Mesh\_Par\_file} and rerun \texttt{xgenerate\_databases} and \texttt{xspecfem3D} to make sure that your simulation works fine.
-
-\item [I get the following error message "forward simulation became unstable and blew up":]
-In most cases this means that your time step size \texttt{DT} is chosen too big.
-Look at your files \texttt{output\_mesher.txt} or \texttt{output\_solver.txt} created in the folder \texttt{OUTPUT\_FILES}.
-In these output files, find the section:
-\begin{lyxcode}
-... ~\\
-********************************************* ~\\
-*** Verification of simulation parameters *** ~\\
-********************************************* ~\\
-... ~\\
-*** Minimum period resolved =    4.308774 ~\\
-*** Maximum suggested time step =   6.8863556E-02 ~\\
-... ~\\
-... ~\\
-\end{lyxcode}
-
-then change \texttt{DT} in the \texttt{DATA/Par\_file} to be somewhere close to the maximum suggested time step. In the example above: \\
+This flag should be identical when using the in-house mesher \texttt{xmeshfem3D},
+\texttt{xgenerate\_databases} and \texttt{xspecfem3D} together to
+run simulations.
+
+The flag determines if the coordinates you specify for your source
+and station locations are given as lat/lon degrees and must be converted
+to UTM coordinates. As an example, if you use \texttt{.false.} within
+\texttt{Mesh\_Par\_file} then you create a mesh with \texttt{xmeshfem3D}
+using the UTM projection from lat/lon as input format to UTM projected
+coordinates to store the mesh point positions, which is fine. The
+error then may occur if in the \texttt{Par\_file} you have this set
+to \texttt{.true.} so that the \texttt{xgenerate\_databases} and \texttt{xspecfem3D}
+suppress the UTM projection and assume that all coordinates you use
+now for source and receiver locations are given in meters (that is,
+converted) already. So it won't find the specified locations in the
+used mesh. As a solutions, just change the flag in \texttt{Par\_file}
+to be the same as in \texttt{Mesh\_Par\_file} and rerun \texttt{xgenerate\_databases}
+and \texttt{xspecfem3D} to make sure that your simulation works fine.
+
+\item [{I get the following error message "forward simulation became unstable and blew up":}] In
+most cases this means that your time step size \texttt{DT} is chosen
+too big. Look at your files \texttt{output\_mesher.txt} or \texttt{output\_solver.txt}
+created in the folder \texttt{OUTPUT\_FILES}. In these output files,
+find the section:
+
+\begin{lyxcode}
+...~~~\\
+~{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}~~~\\
+~{*}{*}{*}~Verification~of~simulation~parameters~{*}{*}{*}~~~\\
+~{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}~~~\\
+~...~~~\\
+~{*}{*}{*}~Minimum~period~resolved~=~4.308774~~~\\
+~{*}{*}{*}~Maximum~suggested~time~step~=~6.8863556E-02~~~\\
+~...~~~\\
+~...~~~\\
+
+\end{lyxcode}
+
+then change \texttt{DT} in the \texttt{DATA/Par\_file} to be somewhere
+close to the maximum suggested time step. In the example above: \\
  \texttt{DT = 0.05d0} \\
-would (most probably) work fine. It could be also bigger than the 0.068 s suggested. This depends a bit on the distortions of
-your mesh elements. The more regular they are, the bigger you can choose \texttt{DT}. Just play with this value a bit and see
-when the simulation becomes stable ...
+ would (most probably) work fine. It could be also bigger than the
+0.068 s suggested. This depends a bit on the distortions of your mesh
+elements. The more regular they are, the bigger you can choose \texttt{DT}.
+Just play with this value a bit and see when the simulation becomes
+stable ...
 
@@
 We protect your rights with two steps:
-
 \begin{enumerate}
@@
 AND MODIFICATION }
-
 \begin{itemize}
-
-\item[0.]This License applies to any program or other work which
-contains a notice placed by the copyright holder saying it may be
-distributed under the terms of this General Public License. The ``Program''
+\item 0.This License applies to any program or other work which contains
+a notice placed by the copyright holder saying it may be distributed
+under the terms of this General Public License. The ``Program''
 below refers to any such program or work, and a ``work based on the
@@
  \\
-Activities other than copying, distribution and modification are not
-covered by this License; they are outside its scope. The act of running
-the Program is not restricted, and the output from the Program is
-covered only if its contents constitute a work based on the Program
+ Activities other than copying, distribution and modification are
+not covered by this License; they are outside its scope. The act of
+running the Program is not restricted, and the output from the Program
+is covered only if its contents constitute a work based on the Program
 (independent of having been made by running the Program). Whether
-that is true depends on what the Program does.
-
-\end{itemize}
-
+that is true depends on what the Program does.\end{itemize}
 \begin{enumerate}
@@
 \end{enumerate}
+
 These requirements apply to the modified work as a whole. If identifiable
@@
 \end{enumerate}
+
 The source code for a work means the preferred form of the work for
@@
 \subsection*{NO WARRANTY }
-
 \begin{itemize}
-
-\item[11.]BECAUSE THE PROGRAM IS LICENSED FREE OF CHARGE, THERE IS
-NO WARRANTY FOR THE PROGRAM, TO THE EXTENT PERMITTED BY APPLICABLE
-LAW. EXCEPT WHEN OTHERWISE STATED IN WRITING THE COPYRIGHT HOLDERS
-AND/OR OTHER PARTIES PROVIDE THE PROGRAM ``AS IS'' WITHOUT WARRANTY
-OF ANY KIND, EITHER EXPRESSED OR IMPLIED, INCLUDING, BUT NOT LIMITED
-TO, THE IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR
-PURPOSE. THE ENTIRE RISK AS TO THE QUALITY AND PERFORMANCE OF THE
-PROGRAM IS WITH YOU. SHOULD THE PROGRAM PROVE DEFECTIVE, YOU ASSUME
-THE COST OF ALL NECESSARY SERVICING, REPAIR OR CORRECTION.
-
-\item[12.]IN NO EVENT UNLESS REQUIRED BY APPLICABLE LAW OR AGREED
-TO IN WRITING WILL ANY COPYRIGHT HOLDER, OR ANY OTHER PARTY WHO MAY
-MODIFY AND/OR REDISTRIBUTE THE PROGRAM AS PERMITTED ABOVE, BE LIABLE
-TO YOU FOR DAMAGES, INCLUDING ANY GENERAL, SPECIAL, INCIDENTAL OR
-CONSEQUENTIAL DAMAGES ARISING OUT OF THE USE OR INABILITY TO USE THE
-PROGRAM (INCLUDING BUT NOT LIMITED TO LOSS OF DATA OR DATA BEING RENDERED
-INACCURATE OR LOSSES SUSTAINED BY YOU OR THIRD PARTIES OR A FAILURE
-OF THE PROGRAM TO OPERATE WITH ANY OTHER PROGRAMS), EVEN IF SUCH HOLDER
-OR OTHER PARTY HAS BEEN ADVISED OF THE POSSIBILITY OF SUCH DAMAGES.
-
+\item 11.BECAUSE THE PROGRAM IS LICENSED FREE OF CHARGE, THERE IS NO WARRANTY
+FOR THE PROGRAM, TO THE EXTENT PERMITTED BY APPLICABLE LAW. EXCEPT
+WHEN OTHERWISE STATED IN WRITING THE COPYRIGHT HOLDERS AND/OR OTHER
+PARTIES PROVIDE THE PROGRAM ``AS IS'' WITHOUT WARRANTY OF ANY KIND,
+EITHER EXPRESSED OR IMPLIED, INCLUDING, BUT NOT LIMITED TO, THE IMPLIED
+WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE.
+THE ENTIRE RISK AS TO THE QUALITY AND PERFORMANCE OF THE PROGRAM IS
+WITH YOU. SHOULD THE PROGRAM PROVE DEFECTIVE, YOU ASSUME THE COST
+OF ALL NECESSARY SERVICING, REPAIR OR CORRECTION.
+\item 12.IN NO EVENT UNLESS REQUIRED BY APPLICABLE LAW OR AGREED TO IN WRITING
+WILL ANY COPYRIGHT HOLDER, OR ANY OTHER PARTY WHO MAY MODIFY AND/OR
+REDISTRIBUTE THE PROGRAM AS PERMITTED ABOVE, BE LIABLE TO YOU FOR
+DAMAGES, INCLUDING ANY GENERAL, SPECIAL, INCIDENTAL OR CONSEQUENTIAL
+DAMAGES ARISING OUT OF THE USE OR INABILITY TO USE THE PROGRAM (INCLUDING
+BUT NOT LIMITED TO LOSS OF DATA OR DATA BEING RENDERED INACCURATE
+OR LOSSES SUSTAINED BY YOU OR THIRD PARTIES OR A FAILURE OF THE PROGRAM
+TO OPERATE WITH ANY OTHER PROGRAMS), EVEN IF SUCH HOLDER OR OTHER
+PARTY HAS BEEN ADVISED OF THE POSSIBILITY OF SUCH DAMAGES.
 \end{itemize}
@@
 convey the exclusion of warranty; and each file should have at least
-the ``copyright'' line and a pointer to where the full notice is found.
-For example:
-
+the ``copyright'' line and a pointer to where the full notice is
+found. For example:
 \begin{quote}
@@
 this when it starts in an interactive mode:
-
 \begin{quote}
-Gnomovision version 69, Copyright $\copyright$ year name of author Gnomovision
-comes with ABSOLUTELY NO WARRANTY; for details type `show w'. This
-is free software, and you are welcome to redistribute it under certain
-conditions; type `show c' for details.
+Gnomovision version 69, Copyright $\copyright$ year name of author
+Gnomovision comes with ABSOLUTELY NO WARRANTY; for details type `show
+w'. This is free software, and you are welcome to redistribute it
+under certain conditions; type `show c' for details. 
 \end{quote}
@@
 You should also get your employer (if you work as a programmer) or
-your school, if any, to sign a ``copyright disclaimer'' for the program,
-if necessary. Here is a sample; alter the names:
-
+your school, if any, to sign a ``copyright disclaimer'' for the
+program, if necessary. Here is a sample; alter the names:
 \begin{quote}

### 2013-06-24 doc/USER_MANUAL/manual_SPECFEM3D_Cartesian.tex
added a reference about the dispersion analysis of the SEM

@@
 
-It has very good accuracy and convergence properties \citep{MaPa89,SePr94,DeFiMu02,Coh02,DeSe07,SeOl08}.
+It has very good accuracy and convergence properties \citep{MaPa89,SePr94,DeFiMu02,Coh02,DeSe07,SeOl08,MeStTh12}.
 The spectral element approach admits spectral rates of convergence
@@
 and in particular \citet{LeKoHuTr09,LeChKoHuTr09,LeChLiKoHuTr08,GoAmTaCaSmSaMaKo09,WiKoScTr04,KoLiTrSuStSh04}.
-A detailed theoretical analysis of the dispersion and stability properties
-of the SEM is available in \citet{Coh02}, \citet{DeSe07}, \citet{SeOl07}
-and \citet{SeOl08}.\\
+A detailed theoretical analysis of the dispersion
+and stability properties of the SEM is available in \citet{Coh02}, \citet{DeSe07}, \citet{SeOl07}, \citet{SeOl08} and \citet{MeStTh12}.\\
 

### 2013-06-24 doc/USER_MANUAL/manual_SPECFEM3D_Cartesian.tex
added three Discontinuous Galerkin references

@@
 it is then close to a particular case of the discontinuous Galerkin
-technique \citep{ReHi73,Arn82,FaRi99,HuHuRa99,CoKaSh00,GiHeWa02,RiWh03,MoRi05,GrScSc06,AiMoMu06,BeLaPi06,DuKa06,DeSeWh08,PuAmKa09,WiStBuGh10,DeSe10,EtChViGl10},
+technique \citep{ReHi73,LeRa74,Arn82,JoPi86,BoMaHe91,FaRi99,HuHuRa99,CoKaSh00,GiHeWa02,RiWh03,MoRi05,GrScSc06,AiMoMu06,BeLaPi06,DuKa06,DeSeWh08,PuAmKa09,WiStBuGh10,DeSe10,EtChViGl10},
 with optimized efficiency because of its tensorized basis functions

### 2013-07-06 doc/USER_MANUAL/manual_SPECFEM3D_Cartesian.tex
** Fix process.sh and update README for examples/layered_halfspace ** Update manual: fix small formatting problems; add a note on 'decompose_mesh has to be run in bin/'

@@
 script \texttt{cubit2specfem3d.py} provided in each of the example
-directories. If the export was successful, you should find the following
+directories (linked to the common script \texttt{CUBIT/cubit2specfem3d.py}). If the export was successful, you should find the following
 files in a subdirectory \texttt{MESH/}:
@@
 The partitioning is done in serial for now (in the next release we
-will provide a parallel version of that code), the synopsis is: 
+will provide a parallel version of that code). It needs to be run in the \texttt{bin/} directory because it expects the \texttt{../DATA/Par\_file}. The synopsis is: 
 \begin{lyxcode}
@@
 \end{description}
-\end{description}
-\end{description}
-
-\begin{description}
+
 \item [{\textmd{1D~models~with~real~structure:}}] ~
@@
 Preliminary Reference Earth Model (PREM) \citep{DzAn81}. 
-
-\begin{description}
 \item [{\texttt{1D\_socal}}] A standard isotropic 1D model for Southern
@@
 \end{description}
-\end{description}
-\end{description}
-
-\begin{description}
+
 \item [{\textmd{Fully~3D~models:}}] ~
@@
 
-\begin{description}
 \item [{\texttt{external}}] For a user-specified isotropic model which
@@
 Adjoint simulations are generally performed for two distinct applications.
-First, they can be used for earthquake source inversions, especially
+First, they can be used in point source moment-tensor inversions, or source imaging for
 earthquakes with large ruptures such as the Lander's earthquake \citep{WaHe94}.
@@
 
-In the case where a specific misfit function is minimized to invert
-for the earthquake source parameters, the gradient of the misfit function
+When a specific misfit function between data and synthetics is minimized to invert
+for earthquake source parameters, the gradient of the misfit function
 with respect to these source parameters can be computed by placing
-time-reversed seismograms at the receivers and using them as sources
-in an adjoint simulation, and then the value of the gradient is obtained
+time-reversed seismograms at the receivers as virtual sources
+in an adjoint simulation. Then the value of the gradient is obtained
 from the adjoint seismograms recorded at the original earthquake location.

### 2013-07-07 doc/USER_MANUAL/manual_SPECFEM3D_Cartesian.tex
Add kernel visualization png file for examples/homogenous_halfspace.

@@
  depending on whether the top surface is defined as free or absorbing
-in DATA/Par\_file. \\
+(\texttt{STACEY\_INSTEAD\_OF\_FREE\_SURFACE} in \texttt{DATA/Par\_file}). \\
  You should put both the surface of acoustic regions and of elastic

### 2013-07-08 doc/USER_MANUAL/manual_SPECFEM3D_Cartesian.tex
moved flag USE_BEST_LOCATION to the "constants.h.in" file to make it easier to change it if/when needed, and added a line about this to the users manual.

@@
 source time function. 
+When using this option, by default the code can locate the force source
+anywhere between mesh points in order to honor its exact location; this
+is more precise than using the closest GLL mesh point, but it is also a bit slower.
+If needed, you can change that default behavior and force the code to use the closest
+GLL mesh point instead by setting flag \texttt{USE\_BEST\_LOCATION} to \texttt{.false.} instead of \texttt{.true.} in file
+\texttt{src/shared/constants.h.in} and running the \texttt{configure} script again and recompiling the code.
 \item [{\texttt{USE\_RICKER\_TIME\_FUNCTION}}] Turn this flag on to use

### 2013-07-08 doc/USER_MANUAL/manual_SPECFEM3D_Cartesian.tex
new cover page with more logos for the users manual

@@
 
-\author{$\copyright$ Princeton University (USA) and\\
- CNRS / INRIA / University of Pau (France)\\
+\author{$\copyright$ Princeton University (USA),\\
+CNRS and University of Marseille (France),\\
+and ETH Z\"urich (Switzerland)\\
 Version 2.1}
@@
 \\
+
+\section*{Current and past main participants or main sponsors}
+
+\begin{figure}[htbp]
+\noindent \begin{centering}
+\includegraphics[width=0.125\textwidth]{figures/logo_cnrs}\vspace*{2truemm}
+\includegraphics[width=0.125\textwidth]{figures/logo_princeton}\vspace*{2truemm}
+\includegraphics[width=0.15\textwidth]{figures/logo_aix_marseille_universite}\vspace*{2truemm}
+\includegraphics[width=0.15\textwidth]{figures/logo_ETH}\vspace*{2truemm}
+\includegraphics[width=0.15\textwidth]{figures/logo_inria}\vspace*{2truemm}
+\includegraphics[width=0.15\textwidth]{figures/logo_UPPA}
+\par\end{centering}
+
+\noindent \begin{centering}
+\includegraphics[width=0.15\textwidth]{figures/logo_NVIDIA}\vspace*{2truemm}
+\includegraphics[width=0.125\textwidth]{figures/logo_IUF}\vspace*{2truemm}
+\includegraphics[width=0.112\textwidth]{figures/logo_Univ_Toulouse}\vspace*{2truemm}
+\includegraphics[width=0.135\textwidth]{figures/logo_Caltech}\vspace*{2truemm}
+\includegraphics[width=0.135\textwidth]{figures/logo_Harvard}\vspace*{2truemm}
+\includegraphics[width=0.135\textwidth]{figures/logo_IPGP}
+\par\end{centering}
+\end{figure}
 

### 2013-07-08 doc/USER_MANUAL/manual_SPECFEM3D_Cartesian.tex
added two more logos

@@
 \includegraphics[width=0.125\textwidth]{figures/logo_princeton}\vspace*{2truemm}
-\includegraphics[width=0.15\textwidth]{figures/logo_aix_marseille_universite}\vspace*{2truemm}
+\includegraphics[width=0.15\textwidth]{figures/logo_aix_marseille_universite}\vspace*{0.02truemm}
+\includegraphics[width=0.12\textwidth]{figures/logo_CSC_China}\vspace*{0.02truemm}
 \includegraphics[width=0.15\textwidth]{figures/logo_ETH}\vspace*{2truemm}
 \includegraphics[width=0.15\textwidth]{figures/logo_inria}\vspace*{2truemm}
-\includegraphics[width=0.15\textwidth]{figures/logo_UPPA}
+\includegraphics[width=0.14\textwidth]{figures/logo_UPPA}
 \par\end{centering}
@@
 \includegraphics[width=0.125\textwidth]{figures/logo_IUF}\vspace*{2truemm}
-\includegraphics[width=0.112\textwidth]{figures/logo_Univ_Toulouse}\vspace*{2truemm}
 \includegraphics[width=0.135\textwidth]{figures/logo_Caltech}\vspace*{2truemm}
 \includegraphics[width=0.135\textwidth]{figures/logo_Harvard}\vspace*{2truemm}
-\includegraphics[width=0.135\textwidth]{figures/logo_IPGP}
+\includegraphics[width=0.135\textwidth]{figures/logo_IPGP}\vspace*{2truemm}
+\includegraphics[width=0.112\textwidth]{figures/logo_TOTAL}\vspace*{2truemm}
+\includegraphics[width=0.112\textwidth]{figures/logo_Univ_Toulouse}
 \par\end{centering}

### 2013-07-08 doc/USER_MANUAL/manual_SPECFEM3D_Cartesian.tex
two more logos (again)
added one more logo
better copyright information

@@
 
-\author{$\copyright$ Princeton University (USA),\\
-CNRS and University of Marseille (France),\\
+\author{$\copyright$ Princeton University (USA), CNRS and University of Marseille (France),\\
 and ETH Z\"urich (Switzerland)\\
@@
 \includegraphics[width=0.135\textwidth]{figures/logo_IPGP}\vspace*{2truemm}
+\par\end{centering}
+
+\noindent \begin{centering}
+\includegraphics[width=0.119\textwidth]{figures/logo_INGV}\vspace*{2truemm}
+\includegraphics[width=0.119\textwidth]{figures/logo_University_of_Toronto}\vspace*{2truemm}
+\includegraphics[width=0.112\textwidth]{figures/logo_Univ_Toulouse}\vspace*{2truemm}
 \includegraphics[width=0.112\textwidth]{figures/logo_TOTAL}\vspace*{2truemm}
-\includegraphics[width=0.112\textwidth]{figures/logo_Univ_Toulouse}
+\includegraphics[width=0.130\textwidth]{figures/logo_Fairbanks}
 \par\end{centering}

### 2013-07-25 doc/USER_MANUAL/manual_SPECFEM3D_Cartesian.tex
added replace_use_mpi_with_include_mpif_dot_h.pl

@@
 
+If your compiler has problems with the \texttt{use mpi} statements that are used in the code, use the script called 
+\texttt{replace\_use\_mpi\_with\_include\_mpif\_dot\_h.pl} in the root directory to replace all of them with \texttt{include 'mpif.h'} automatically.
+
 \section{Adding OpenMP support in addition to MPI}
+
+NOTE FROM JULY 2013: OpenMP support is maybe / probably not maintained any more.
+Thus the section below is maybe obsolete.
 

### 2013-08-29 doc/USER_MANUAL/manual_SPECFEM3D_Cartesian.tex
changed OUTPUT_FILES/SEM to SEM

@@
 
-Note that we use CUBIT to create meshes of hexahedra, but other packages
+Note that we use CUBIT (now called Trelis) to create meshes of hexahedra, but other packages
 can be used as well, for instance GiD from \url{http://gid.cimne.upc.es}
@@
 adjoint sources need to be computed from the original recorded seismograms
-for the selected stations and saved in the \texttt{OUTPUT\_FILES/SEM/}
-directory with the format \texttt{STA.NT.BX?.adj}, where \texttt{STA},
+for the selected stations and saved in a sub-directory called \texttt{SEM/}
+in the root directory of the code, with the format \texttt{STA.NT.BX?.adj}, where \texttt{STA},
 \texttt{NT} are the station name and network code given in the \texttt{DATA/STATIONS\_ADJOINT}
@@
 With the successful forward simulation and the adjoint source ready
-in the \texttt{OUTPUT\_FILES/SEM/} directory, set \texttt{SIMULATION\_TYPE
+in the \texttt{SEM/} directory, set \texttt{SIMULATION\_TYPE
 = 3} and \texttt{SAVE\_FORWARD = .false.} in the \texttt{Par\_file}
@@
 
-mkdir~-p~OUTPUT\_FILES/SEM
-
-cd~OUTPUT\_FILES/SEM/~
+mkdir~-p~SEM
+
+cd~SEM/~
 

### 2013-09-20 doc/USER_MANUAL/manual_SPECFEM3D_Cartesian.tex
added Ainsworth
renamed CUBIT directory to CUBIT_GEOCUBIT

@@
 
-It has very good accuracy and convergence properties \citep{MaPa89,SePr94,DeFiMu02,Coh02,DeSe07,SeOl08,MeStTh12}.
+It has very good accuracy and convergence properties \citep{MaPa89,SePr94,DeFiMu02,Coh02,DeSe07,SeOl08,AiWa09,AiWa10,MeStTh12}.
 The spectral element approach admits spectral rates of convergence
@@
 script \texttt{cubit2specfem3d.py} provided in each of the example
-directories (linked to the common script \texttt{CUBIT/cubit2specfem3d.py}). If the export was successful, you should find the following
+directories (linked to the common script \texttt{CUBIT\_GEOCUBIT/cubit2specfem3d.py}). If the export was successful, you should find the following
 files in a subdirectory \texttt{MESH/}:

### 2013-09-20 CUBIT_GEOCUBIT/doc/geocubit_manual.tex
moved the final SVN version of GEOCUBIT at CIG to the SPECFEM3D_Cartesian package, to make it easier to maintain and to avoid having two different versions, none of them working fine in all cases and for all users

% Preview source code

%% LyX 2.0.0beta3 created this file.  For more info, see http://www.lyx.org/.
%% Do not edit unless you really know what you are doing.
\documentclass[a4paper,11pt,english]{article}
\setlength{\textwidth}{450pt}
\setlength{\oddsidemargin}{0pt}
\setlength{\marginparwidth}{0pt}
\setcounter{secnumdepth}{3}
\setcounter{tocdepth}{3}
\usepackage[T1]{fontenc}
\usepackage[latin9]{inputenc}
\usepackage{babel}
\usepackage{setspace}
\usepackage{listings}
\usepackage{graphicx,subfigure}
\usepackage{pdfsync}
\usepackage{colortbl}
\usepackage[unicode=true,pdfusetitle,
 bookmarks=true,bookmarksnumbered=true,bookmarksopen=false,
 breaklinks=false,pdfborder={0 0 0},backref=section,colorlinks=true,pdfpagemode=None]
 {hyperref}
\hypersetup{
 pdftex,pdfstartview=FitH,pdfpagelayout=SinglePage}

\newenvironment{lyxcode}
{\par\begin{list}{}{
\setlength{\rightmargin}{\leftmargin}
\setlength{\listparindent}{0pt}% needed for AMS classes
\setlength{\itemsep}{0pt}
\setlength{\parsep}{0pt}
\normalfont\ttfamily}% 
\item[]}
{\end{list}}

\begin{document}

\title{GEOCUBIT USER MANUAL}
\author{Emanuele Casarotti}

\maketitle

\tableofcontents
\newpage{}

\section{WHY}

GEOCUBIT is a python library wrapping around the \href{http://cubit.sandia.gov} {CUBIT} Python Interface and it aims to facilitate the meshing
process in some common problems in seismic wave propagation.\\ 
In particular, it is focused on the meshing requests of SPECFEM3D and it is helpful for some tedious tasks as:\\

\begin{itemize}
\item Creation of geophysical surfaces and volumes (ex. topography).
\item Mesh of layered volumes with hexahedral.
\item Creation of an anisotropic mesh suitable for cases where some alluvial
basin (or slow velocity zones) are present.
\item It can be used as serial or parallel process. The parallel meshing
capabilities are fundamental for large geophysical problems (ex. mesh
of Southern California using SRTM topography).
\end{itemize}
GEOCUBIT can be use inside the graphical interface of \href{http://cubit.sandia.gov} {CUBIT} (i.e.
as python object in the script tab) or as unix command. We refer to the \href{http://cubit.sandia.gov} {CUBIT} help appendix for more information about the python interface.

\newpage{}
\section{FIRST STEPS}

\subsection{Requirements}

The minimum requirements for using GEOCUBIT are:

\begin{itemize}
\item \href{http://cubit.sandia.gov} {CUBIT 12.2}
\item \href{http://downloads.sourceforge.net/numpy} {numpy 1.0+}
\item \href{http://www.python.org} {python 2.5}
 (strictly! it depends on the cubit library that refers to this version of python)
\end{itemize}
\raggedright{and  for using the parallel meshing capabilities:}
\begin{itemize}
\item \href{http://downloads.sourceforge.net/pympi/pyMPI-2.4b4.tar.gz} {pyMPI}
\end{itemize}

\subsection{Installation}

For installing, download the code and type in the GEOCUBIT directory:
\begin{lyxcode}
python2.5 setup.py install~
\end{lyxcode}

Check that the following variables are set:

\begin{lyxcode}
CUBITDIR=/usr/local/CUBIT\\
CUBITLIB=\$CUBITDIR/bin:\$CUBITDIR/structure:\$CUBITDIR/components\\
PYTHONPATH=\$CUBITDIR/components/cubit:\$CUBITDIR/structure:\$CUBITDIR/bin\\
LD\_LIBRARY\_PATH=\$CUBITDIR/bin\\
PATH=\$CUBITDIR/bin:\$CUBITDIR/components/cubit \\
\end{lyxcode}

\newpage{}
\section{USING GEOCUBIT AT COMMAND LINE:}

\subsection{UTILITIES}
\begin{itemize}
\item checking the configuration of the libraries and dependencies:\\
\begin{lyxcode}
GEOCUBIT.py --chklib
\end{lyxcode}

\item checking the parameter file:\\
\begin{lyxcode}
GEOCUBIT.py --chkcfg --cfg=[file]
\end{lyxcode}
\end{itemize}

\subsection{CREATE GEOMETRY}
SURFACES

\begin{itemize}
\item creating acis surfaces using a parameter file:\\
\begin{lyxcode}
 GEOCUBIT.py --build\_surface --cfg=[filename]
\end{lyxcode}
\end{itemize}

\begin{itemize}
\item creating a surface from regular ascii grid or ascii lines that define a "skin":\\
\begin{lyxcode}
 GEOCUBIT.py --surface=[file] (--regulargrid=[opts] | --skin=[opts])
\end{lyxcode}
\end{itemize}

VOLUMES
\begin{itemize}
\item **serial** command: creating cubit volumes using a parameter file:\\
\begin{lyxcode}
GEOCUBIT.py --surface=[file] (--regulargrid=[opts] | --skin=[opts])
\end{lyxcode}
\end{itemize}

\begin{itemize}
\item **parallel** command: creating cubit volumes using a parameter file:\\
\begin{lyxcode}
mpirun -n [numproc] pyMPI GEOCUBIT.py --build\_volume --cfg=[file]
\end{lyxcode}

\begin{figure}\begin{center}
    \subfigure[]{\includegraphics[scale=.5]{media/geocubitmanual_grid2.jpg}
    \label{fig:vol}}%
    \subfigure[]{\includegraphics[scale=.9]{media/geocubitmanual_grid.jpg}
    \label{fig:table}}%
\caption{a) Volume divided in slices. b) Parallel map of the IDs for volume in Figure~\ref{fig:vol} divided in (16 slices, 4 along x-axis and 4 along y-axis).}
\label{fig:grid}
\end{center}
\end{figure}

In this parallel application, the volume is separated in N slices, each one is assigned at a single process and one file for each slice is created (see Figure~\ref{fig:grid} as example).
\end{itemize}

\subsection{MESHING}
\begin{itemize}
\item **serial** command: building a volume and meshing it.\\
\begin{lyxcode}
GEOCUBIT.py --build\_volume --mesh --cfg=[file] (--id\_proc=n)
\end{lyxcode}
\end{itemize}

\begin{itemize}
\item **serial** command : meshing a volume\\
\begin{lyxcode}
GEOCUBIT.py --mesh --cfg=[file] (--id\_proc=[n, def=0])
\end{lyxcode}
\footnotesize note: without the --build\_volume flag the script recall an old "geometry\_vol\_[id\_proc].cub" file\\
\end{itemize}

\begin{itemize}
\item **parallel** command: meshing a volume from a parameter file\\
\begin{lyxcode}
mpirun -n [nproc] pyMPI GEOCUBIT.py --build\_volume --mesh --cfg=[file]
\end{lyxcode}
\end{itemize}

\subsection{FINALIZING AND EXPORTING}

It is possible to collect, merge, set the absorbing boundary conditions and to export the resulting mesh in a format readable by SPECFEM3D.
\begin{itemize}
\item
Collecting some cubit files, setting the absorbing boundary conditions and merging in a single free mesh cubitfile
\begin{lyxcode}
GEOCUBIT.py --collect --merge --meshfiles=[files] --cpux=N --cpuy=N (--rangecpux=[cpuxmin,cpuxmax] --rangecpuy=[cpuymin,cpuymax])
\end{lyxcode}
\texttt{cpuy, cpux} set the number of cpus used for creating the mesh files; \texttt{rangecpux} and \texttt{rangecpuy} set the range of slices that are used in the mesh.\\
Following the example in Figure~\ref{fig:grid}, the parameters \texttt{--cpux=4 --cpuy=4 --rangecpux=1,2 --rangecpuy=2,3} select the slices with id 10 11 14 15. Only these slices are going to be collect and merged with the appropriated absorbing boundary conditions.

\item Collecting a single free mesh cubitfile and refine the hex inside some curves (ex. alluvial basin, Figure~\ref{fig:refcurve})
\begin{lyxcode}
GEOCUBIT.py --collect --meshfiles=[list of files] --curverefining=[file]\\
\end{lyxcode}
\footnotesize note: the curves must be stored in acis format (sat) and must be closed.

\end{itemize}

\begin{figure}\center
\includegraphics[scale=.37]{media/refcurve_1.png}%
\includegraphics[scale=.303]{media/refcurve_2.png}%
\caption{}

\label{fig:refcurve}
\end{figure}

\begin{itemize}
\item Exporting a cubit mesh file in a SPECFEM3D mesh
\begin{lyxcode}
GEOCUBIT.py --export2SPECFEM3D=[output directory] --meshfiles=[filename] (--listblock=[list of cubit blocks] --listflag=[list of specfem3d flags])
\end{lyxcode}
\footnotesize If required, it is possible to assign personalized flags to each block using --listblock and the correspondent 	listflag (for example: --listblock=3,4 --listflag=1,-1)
\end{itemize}

\newpage{}
\section{GEOCUBIT and CUBIT GRAPHICAL INTERFACE}
In the python script tab of CUBIT GUI, type:

\begin{lyxcode}
>f="volume.cfg" $\to$  \textrm{\small{\textit{store the name of the configuration file, see next section}}}\\
\end{lyxcode}

\begin{lyxcode}
>from geocubitlib import volumes $\to$  \textrm{\small{\textit{load the geocubit modules}}}\\
>from geocubitlib import mesh\_volume\\
>from geocubitlib import exportlib\\
\end{lyxcode}

\begin{lyxcode}
volumes.volumes(f) $\to$  \textrm{\small{\textit{create the volumes, the parameters are stored in the cfg file}}}\\
mesh\_volume.mesh(f) $\to$  \textrm{\small{\textit{mesh the volume}}}\\
\end{lyxcode}\begin{lyxcode}
exportlib.collect() $\to$  \textrm{\small{\textit{set the boundary conditions, see SPECFEM3D manual}}}\\
exportlib.e2SEM(outdir="./output") $\to$  \textrm{\small{\textit{save the mesh in the SPECFEM3D format}}}\\
\end{lyxcode}
See the media/iterative.mov for a live example.

\section{EXAMPLES of CONFIGURATION FILES}
In the current section a few configuration files will be described. Please, see the files in the examples directory of the GEOCUBIT distribution for reference.\\

\subsection{GENERAL OPTIONS}

A cfg file has a format similar to ini windows files.
They are divided in \texttt{[sections]}. There is no order in the position or in the name of the sections.

Generally, the parameters that control the general options are in the first section, called \texttt{[cubit.options]} and \texttt{[simulation.cpu\_parameters]}. For example:

\begin{lyxcode}
[cubit.options]\\
cubit\_info=off $\to$  \textrm{\small{\textit{turn on/off the information of the cubit command}}}\\
echo\_info=off $\to$   \textrm{\small{\textit{turn on/off the echo of the cubit command}}}\\
cubit\_info=off $\to$   \textrm{\small{\textit{turn on/off the cubit journaling}}}\\
echo\_info=off $\to$   \textrm{\small{\textit{turn on/off the cubit error journaling}}}\\
working\_dir=tmp $\to$   \textrm{\small{\textit{set the working directory}}}\\
output\_dir=output $\to$   \textrm{\small{\textit{set the output directory}}}\\
save\_geometry\_cubit = True $\to$   \textrm{\small{\textit{true if it saves the geometry files}}}\\
save\_surface\_cubit = False $\to$   \textrm{\small{\textit{true if it saves the surfaces in a cubit files}}}\\
export\_exodus\_mesh = True $\to$   \textrm{\small{\textit{true if it saves the mesh also in exodus format. The default is the cubit format (cub), that contains both the geometry and the meshing information. In case of complex geometry, the cub file could be enormous and a more light exodus file become important.}}}\\
\end{lyxcode}
\begin{lyxcode}
[simulation.cpu\_parameters]\\
nodes = 1  $\to$   \textrm{\small{\textit{number of nodes/process}}}\\

\end{lyxcode}

\footnotesize note: Usually the *\_info options are turn off by default for improving the performances.\normalsize\\

\subsection{CREATING A TOPOGRAPHIC SURFACE}
\footnotesize See stromboli.cfg, execute with: \verb!GEOCUBIT.py --build\_surface --cfg=./example/stromboli.cfg. !\normalsize\\
~\\
GEOCUBIT is able to create a topographic surface based upon the CUBIT command \verb!create skin curve! and \verb!create surface net! (Figure~\ref{fig:surfacetype}). See the CUBIT manual for details.

\begin{lyxcode}
[geometry.surfaces]
nsurf = 2 $\to$   \textrm{\small{\textit{number of surfaces that will be created}}}\\
\end{lyxcode}
\begin{lyxcode}
[surface\textbf{1}.parameters]\\
name=example/data/stromboli.xyz  $\to$   \textrm{\small{\textit{the name of the file defining the surface}}}\\
surf\_type=skin $\to$   \textrm{\small{\textit{the type of surface:} \texttt{skin} \textit{, the file is a sequence of parallel points that span the surface in the order describer by the following direction parameters} \texttt{regular\_grid} \emph{the points are structured distritibuted. see Cubit Manual for details(skin surface and net surface) }}}\\
unit\_surf= utm $\to$   \textrm{\small{\textit{unit of the points,} \texttt{utm} \emph{or} \texttt{geo} \emph{(geographical) coordinates}}}\\
directionx = 0 $\to$   \textrm{\small{\textit{if} \texttt{surf\_type=skin}\emph{:} \texttt{directionx} \emph{and} \texttt{directiony} \emph{define how the points span the surface}}}\\
directiony = 1\\
step = 1 $\to$   \textrm{\small{\textit{the script creates a vertex each} \texttt{step} \emph{points: in this case} \texttt{step=1} \textit{means a vertex for all the points}}}\\
\end{lyxcode}

\begin{lyxcode}
[surface2.parameters] $\to$ \textrm{\small{\emph{note the different number of surface in the section name}}}
name=./examples/surfaces/topo.dat $\to$   \textrm{\small{\textit{the name of the file defining the surface}}}\\
surf\_type=regular\_grid $\to$   \textrm{\small{\textit{the type of surface:} \texttt{skin} \textit{, the file is a sequence of parallel points that span the surface in the order describer by the following direction parameters} \texttt{regular\_grid} \emph{the points are structured distritibuted. See see Cubit Manual for details (skin surface and net surface) }}}\\
unit\_surf= geo $\to$   \textrm{\small{\textit{unit of the points,} \texttt{utm} \emph{or} \texttt{geo} \emph{(geographical) coordinates}}}\\
nx=5 $\to$   \textrm{\small{\textit{In case of} \texttt{regular\_grid}: \texttt{nx} \emph{and} \texttt{ny} \emph{are the number of points along x and y direction.}}}\\
ny=5
\end{lyxcode}

\begin{figure}\begin{center}
    \subfigure[]{\includegraphics[scale=.4]{media/skinsurface.png}
    \label{fig:skin}}%
    \subfigure[]{\includegraphics[scale=.4]{media/netsurface.png}
    \label{fig:net}}%
    \subfigure[]{\includegraphics[scale=.25]{media/strombolisurface.png}
    \label{fig:stromboli}}%
\caption{GEOCUBIT creates a topographic surface in 2 ways: a) skin topography, \texttt{directionx=0} and \texttt{directiony=1}, b) topography from a net of structured distritibuted points, \texttt{nx=4} and \texttt{ny=4}, c) Topography of Stromboli volcano (Italy) using the net surface method.}
\label{fig:surfacetype}
\end{center}
\end{figure}

\subsection{CREATING and MESHING A VOLUME}
\footnotesize see volume.cfg, for executioning: \verb!GEOCUBIT.py --build\_volume --mesh --cfg=./example/volume.cfg !\normalsize\\
~\\
The example is for a volume with topography with dimension of the hexahedra increasing with depth by means of a refinement layer (Figure~\ref{fig:volume}).

\begin{figure}\begin{center}
\includegraphics[scale=.4]{media/geometryvolume.png}%
\includegraphics[scale=.377]{media/meshvolume.png}%
\caption{Mesh of a simple volume}

\label{fig:volume}
\end{center}
\end{figure}

\begin{lyxcode}
[geometry.volumes]\\
volume\_type = layercake\_volume\_ascii\_regulargrid\_regularmap\\
latitude\_min                    = 13.879\\
latitude\_max                    = 14.279\\
longitude\_min                   = 40.619\\
longitude\_max                   = 40.969\\
nx = 5 \\
ny = 5\\
unit                            = geo\\
\end{lyxcode}

\begin{lyxcode}
[geometry.volumes.layercake]\\
nz = 2\\
bottomflat = True\\
depth\_bottom = -7000\\
filename = ./example/topo.dat\\
geometry\_format=ascii\\
\end{lyxcode}

\texttt{nz} is the number of the horizontal surfaces in the volume (in this case: topography and bottom).\\
\texttt{bottomflat} is True if the bottom is flat.\\
\texttt{depth\_bottom} is the depth of the bottom.\\
\texttt{filename} is a list of files. Each one defines a surface (in this case there is only the topography file in a regular\_grid format).\\
\texttt{geometry\_format} is set to ascii since the definition of the surfaces comes from ascii files (structured xyz points). \\

\begin{lyxcode}
[meshing]\\
map\_meshing\_type=regularmap\\
iv\_interval=5, $\to$   \textrm{\small{\textit{if only one value is present, append the comma}}}\\
size=2000\\
or\_mesh\_scheme=map\\
ntripl=1\\
tripl=2, $\to$   \textrm{\small{\textit{if only one value is present, append the comma}}}\\
smoothing=False\\
\end{lyxcode}

The \texttt{meshing} section contains the parameters request for the meshing process.
\texttt{map\_meshing\_type} set the meshing scheme and is regularmap by default (other schemes are in preliminary phase).\\
\texttt{iv\_interval} set the number of "horizontal" hex sheets for each layer.
\texttt{size} is the dimension of hexahedral (horizontally)\\
\texttt{or\_mesh\_scheme} is the meshing scheme for the topography (\texttt{map} or \texttt{pave} are possible, see the CUBIT manual for more information).\\
\texttt{ntripl} is the number of tripling layer in the mesh (in this case 1).\\
\texttt{tripl} means in this case that the the refinement layer is located at the second surface (the topography). The surfaces are ordered from the bottom (surface 1) to the top (surface 2).\\
\texttt{smoothing} performes the smoothing command in Cubit.\\

\subsection{LAYERED VOLUME and MESH FOR CENTRAL ITALY (PARALLEL)}

\footnotesize In the example abruzzo.cfg, the layers are 2. For execution type: \texttt{mpirun -n 150 GEOCUBIT.py --build\_volume --mesh --cfg=./example/abruzzo.cfg}.\normalsize\\
~\\

Comparing the previous cfg files, there are few modification:

\begin{lyxcode}
$\cdots $\\
nz = 3 \\
$\cdots $\\
filename = example/data/moho\_int.xyz,example/data/topo.xyz\\
$\cdots $\\
iv\_interval=8,8 (one value for each layer)\\
$\cdots $\\
refinement\_depth=1,\\
\end{lyxcode}

The volume has \texttt{nz=3} interfaces: bottom, moho, and topography. The bottom is flat and  his z-coordinate position is set by \texttt{depth\_bottom}. 
The name of the files that storage the data for the other interfaces are listed in \texttt{filename}.
The interfaces defines 2 geological layers. Each layer has \texttt{iv\_interval=8} "horizontal" hex sheets. There is only a refinement and it is set in the \texttt{refinement\_depth=1} hex sheet, i.e just below the topography (\texttt{refinement\_depth=8} means just below the moho).

\subsection{CREATION OF A REGULAR MESH}

\footnotesize In the example grid.cfg, the layers are 3. For execution type: \texttt{GEOCUBIT.py --build\_volume --mesh --cfg=./example/grid.cfg}.\normalsize\\
~\\

In this example we create a simple layercake box \texttt{geometry\_format=regmesh} with 3 layers at depth defined by \texttt{zdepth=-7000,-3000,-600,0}. The initial mesh has hex with horizontal size \texttt{size=2000} and the numbers of vertical hex sheets is iv\_interval=3,1,1 respectively for the bottom, middle and top layer. We include a refinement layer (\texttt{ntripl=1}) at \texttt{tripl=2,} interface (the second from the bottom). Since the refinement occurs on the vertical direction, the vertical dimensions of the hexes in the top layer is too small and the quality of the mesh decreases. \texttt{coarsening\_top\_layer=True} remesh the top layer and the number of vertical hex sheet in the vertical is defined by \texttt{actual\_vertical\_interval\_top\_layer=1}. (see Figure~\ref{fig:regrid})

\begin{figure}\begin{center}
    \subfigure[]{\includegraphics[scale=.2]{media/regrid.png}
    \label{fig:regrid_nocoarse}}%
    \subfigure[]{\includegraphics[scale=.47]{media/regrid_coarse.png}
    \label{fig:regrid_coarse}}%
\caption{Regular mesh with one tripling layer, a) mesh with coarsening\_top\_layer=False, b) mesh with \texttt{coarsening\_top\_layer=True}, the top layer is remeshed with \texttt{actual\_vertical\_interval\_top\_layer=1} vertical hex sheet.}
\label{fig:regrid}
\end{center}
\end{figure}

\subsection{TODO: MESH for SOUTHERN CALIFORNIA}
Mesh with high number of hexes and high resolution topography. Mesh in Parallel.

TO DO

\subsection{TODO MESH for SANTA MONICA OVERTRUST}
Using part of the mesh for Southern California, stitching together several slices.

TO DO

\subsection{TODO Grouping the hexes with different period resolved}

Checking the stability
For execution type: \texttt{GEOCUBIT.py --meshfiles=[filename] --stability (--tomofile=[tomographic file]/--vp=[vp text value] --vs=[vs text value])}.

or

in the script tab of the CUBIT GUI, type:

\begin{lyxcode}
>from geocubitlib.hex_metric import SEM_stability_3D $\to$  \textrm{\small{\textit{load the geocubit modules}}}\\
>mesh=SEM_stability_3D()\\
>mesh.check_simulation_parameter(vp_static=vp,vs_static=vs) $\to$  \textrm{\small{\textit{text the mesh against some homogeneous velocity model}}}\\
or\\
>mesh.check_simulation_parameter(tomofile=tomofile) $\to$  \textrm{\small{\textit{text the mesh against a tomographic file}}}\\
mesh.group_timestep() $\to$  \textrm{\small{\textit{grouping hex with similar timestep}}}\\
mesh.group_period() $\to$  \textrm{\small{\textit{grouping hex with similar period}}}\\
\end{lyxcode}

TO DO: picture

\section{TODO SPECFEM3D MESH FORMAT}

TO DO

\section{TODO NOTES ABOUT THE DIMENSION OF THE HEXES, THE MINIMUM PERIOD RESOLVED AND THE TIME STEP OF THE SIMULATION}

TO DO
\bibliography{bibliography.bib}
\end{document}

### 2013-09-20 CUBIT_GEOCUBIT/GEOCUBIT.py
moved the final SVN version of GEOCUBIT at CIG to the SPECFEM3D_Cartesian package, to make it easier to maintain and to avoid having two different versions, none of them working fine in all cases and for all users

@@ -1,0 +1,73 @@
+#!/usr/bin/env python
+"""
+GEOCUBIT.py
+this file is part of GEOCUBIT
+
+Created by Emanuele Casarotti 
+Copyright (c) 2011 Istituto Nazionale di Geofisica e Vulcanologia 
+
+---------------------------------------------------------------------------
+ GEOCUBIT is free software: you can redistribute it and/or modify
+ it under the terms of the GNU General Public License as published by
+ the Free Software Foundation, either version 3 of the License, or
+ (at your option) any later version.
+
+ GEOCUBIT is distributed in the hope that it will be useful,
+ but WITHOUT ANY WARRANTY; without even the implied warranty of
+ MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ GNU General Public License for more details.
+
+ You should have received a copy of the GNU General Public License
+ along with GEOCUBIT.  If not, see <http://www.gnu.org/licenses/>.
+
+---------------------------------------------------------------------------
+
+GEOCUBIT requires:
+
+- CUBIT 12.2 - www.cubit.sandia.gov
+- python 2.5 
+- numpy 1.0+  - http://downloads.sourceforge.net/numpy
+
+  -- optional for parallel mesh
+- pympi - http://downloads.sourceforge.net/pympi
+
+"""
+import geocubitlib.menu as menu
+import geocubitlib.start as start
+mpiflag,iproc,numproc,mpi   = start.start_mpi()
+if menu.build_surface or menu.build_volume or menu.meshing:
+    cubit                   = start.start_cubit(init=True)
+else:
+    cubit                   = start.start_cubit()
+
+if __name__ == '__main__':
+    
+    #GEOMETRY
+    if menu.build_surface:              
+        from geocubitlib import surfaces
+        surfaces.surfaces()                 
+    if menu.build_volume:                   
+        from geocubitlib import volumes     
+        volumes.volumes()                   
+               
+    #MESHING
+    if menu.meshing:
+        from geocubitlib import mesh_volume
+        mesh_volume.mesh()             
+    
+    #EXPORT
+    if menu.collect:
+        from geocubitlib.exportlib import collect
+        try:
+            output=menu.output
+        except:
+            output='totalmesh_merged'
[…]

### 2013-09-20 CUBIT_GEOCUBIT/geocubitlib/dev/material.py

@@ -1,0 +1,80 @@
+#############################################################################
+# material.py                                                    
+# this file is part of GEOCUBIT                                             #
+#                                                                           #
+# Created by Emanuele Casarotti                                             #
+# Copyright (c) 2008 Istituto Nazionale di Geofisica e Vulcanologia         #
+#                                                                           #
+#############################################################################
+#                                                                           #
+# GEOCUBIT is free software: you can redistribute it and/or modify          #
+# it under the terms of the GNU General Public License as published by      #
+# the Free Software Foundation, either version 3 of the License, or         #
+# (at your option) any later version.                                       #
+#                                                                           #
+# GEOCUBIT is distributed in the hope that it will be useful,               #
+# but WITHOUT ANY WARRANTY; without even the implied warranty of            #
+# MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the             #
+# GNU General Public License for more details.                              #
+#                                                                           #
+# You should have received a copy of the GNU General Public License         #
+# along with GEOCUBIT.  If not, see <http://www.gnu.org/licenses/>.         #
+#                                                                           #
+#############################################################################
+
+
+
+class material():
+    def __init__(self,name,vp,vs,rho,q,ani):
+        self.name=name
+        self.vp=vp
+        self.vs=vs
+        self.rho=rho
+        self.qp=q
+        self.qs=ani
+    def __repr__(self):
+        print self.name
+        print 'vp',self.vp
+        print 'vs',self.vs
+        print 'rho',self.rho
+        print 'Qp',self.qp
+        print 'Qs',self.qs
+        return
+    def __str__(self):
+        txt=self.name+' vp '+str(self.vp)+ ' vs '+str(self.vs)+ ' rho '+str(self.rho)+ ' Qp '+str(self.qp)+ ' Qs '+str(self.qs)
+        return txt
+
+
+def read_material_archive(archivefile='material_archive.dat'):
+    archive=open(archivefile,'r')
+    archive_dict={}
+    for record in archive:
+        m=material(*record.split())
+        archive_dict[m.name]=m
+    return archive_dict
+
+
+def block_description(archivefile='material_archive.dat'):
+    try:
+        import initializing as initializing
[…]

### 2013-09-20 doc/GJI_2011_paper_about_the_code/paper_GJI_2011_about_the_code.tex
converted all the export2SESAME() names in CUBIT scripts to the new GEOCUBIT name, which is export2SPECFEM3D()

--- 2012-05-18 doc/GJI_2011_paper_about_the_code/paper_sesame_2.0.tex
@@
 %% ------------------------------------------------------------------------ %%
-\title[SPECFEM3D Version~2.0 `Sesame']{Forward and adjoint simulations of seismic wave propagation on fully unstructured hexahedral meshes}
+\title[SPECFEM3D Version~2.0]{Forward and adjoint simulations of seismic wave propagation on fully unstructured hexahedral meshes}
 
@@
 We present various examples of fully unstructured meshes, snapshots of wavefields and finite-frequency kernels
-generated by Version~2.0 `Sesame' of our widely used open source spectral-element package \href{http://www.geodynamics.org/cig/software/specfem3d}{SPECFEM3D}.
+generated by Version~2.0 of our widely used open source spectral-element package \href{http://www.geodynamics.org/cig/software/specfem3d}{SPECFEM3D}.
  \end{summary}
@@
 %% ------------------------------------------------------------------------ %%
-We present a new software package, \href{http://www.geodynamics.org/cig/software/specfem3d}{SPECFEM3D} Version~2.0 `Sesame',
+We present a new software package, \href{http://www.geodynamics.org/cig/software/specfem3d}{SPECFEM3D} Version~2.0,
 capable of simulating forward and adjoint seismic wave propagation on fully unstructured hexahedral meshes of
@@
 
-The purpose of this article is to present forward and adjoint simulations in various 3D models using the new software package, \href{http://www.geodynamics.org/cig/software/specfem3d}{SPECFEM3D} Version~2.0 `Sesame', thereby illustrating its current capabilities.
+The purpose of this article is to present forward and adjoint simulations in various 3D models using the new software package, \href{http://www.geodynamics.org/cig/software/specfem3d}{SPECFEM3D} Version~2.0, thereby illustrating its current capabilities.
 The original \href{http://www.geodynamics.org/cig/software/specfem3d}{SPECFEM3D} package for local simulations was extended, improved and optimized in various ways.
-The Version~2.0 `Sesame' release includes a more flexible internal mesher and accommodates more powerful external meshers,
+The Version~2.0 release includes a more flexible internal mesher and accommodates more powerful external meshers,
 such as \href{http://cubit.sandia.gov}{CUBIT}
@@
 For more technical details, the reader is referred to \citet{KoTr99}.
-\href{http://www.geodynamics.org/cig/software/specfem3d}{SPECFEM3D} Version~2.0 `Sesame' implements wave propagation in coupled (an)elastic and acoustic materials
+\href{http://www.geodynamics.org/cig/software/specfem3d}{SPECFEM3D} Version~2.0 implements wave propagation in coupled (an)elastic and acoustic materials
 on local scales.
@@
 Our new open source software package,
-\href{http://www.geodynamics.org/cig/software/specfem3d}{SPECFEM3D} Version~2.0 `Sesame',
+\href{http://www.geodynamics.org/cig/software/specfem3d}{SPECFEM3D} Version~2.0,
 performs acoustic and (an)elastic simulations of seismic wave propagation in complex geological models.
@@
 Princeton Institute for Computational Science \& Engineering (USA).
-The open source spectral-element software package \href{http://www.geodynamics.org/cig/software/specfem3d}{SPECFEM3D} Version~2.0 `Sesame' used
+The open source spectral-element software package \href{http://www.geodynamics.org/cig/software/specfem3d}{SPECFEM3D} Version~2.0 used
 for this article is available for download via the Computational Infrastructure for Geodynamics
@@
 \end{center}
-\caption{Workflow for running spectral-element simulations with \href{http://www.geodynamics.org/cig/software/specfem3d}{SPECFEM3D} Version~2.0 `Sesame'.
+\caption{Workflow for running spectral-element simulations with \href{http://www.geodynamics.org/cig/software/specfem3d}{SPECFEM3D} Version~2.0.
 }

### 2013-09-20 Mount_StHelens/mesh_mount_stl.py
converted all the export2SESAME() names in CUBIT scripts to the new GEOCUBIT name, which is export2SPECFEM3D()

@@ -187,10 +187,10 @@ cubit.cmd('save as "meshing.cub" overwrite')
 # cleanup
 os.system('rm -f topo_brick.stl topo_vol.stl topo_vol2.stl')
 
-#### Export to SESAME format using cubit2specfem3d.py of GEOCUBIT
+#### Export to SPECFEM3D format using cubit2specfem3d.py of GEOCUBIT
 
 os.system('mkdir -p MESH')
-cubit2specfem3d.export2SESAME('MESH')
+cubit2specfem3d.export2SPECFEM3D('MESH')
 
 # all files needed by SCOTCH are now in directory MESH
 

### 2013-09-20 doc/USER_MANUAL/manual_SPECFEM3D_Cartesian.tex
converted all the export2SESAME() names in CUBIT scripts to the new GEOCUBIT name, which is export2SPECFEM3D()
updated the manual regarding how to use "claro" and also "run_cubit2specfem3d.py"; also fixed a typo in a GEOCUBIT script

@@
 contacting the CIG Computational Seismology Mailing List \urlwithparentheses{cig-seismo@geodynamics.org}.
+
+In most cases, to re-create the meshes for the examples given, just type "claro ./block\_mesh.py" or similar 
+from the command line ("claro" is the command to run CUBIT from the command line).
 
@@
 details.
+
+In directory "CUBIT\_GEOCUBIT" we provide two scripts that can help doing the above tasks of exporting a CUBIT mesh
+to SPECFEM3D format automatically for you: "run\_cubit2specfem3d.py" and "run\_boundary\_definition.py".
+Just edit them to indicate the path to your local installation of CUBIT and also the name of the *.cub existing CUBIT mesh file
+that you want to export to SPECFEM3D format. These scripts will do the conversion for you automatically except assigning material properties
+to the different mesh layers. To do so, you will then need to edit the file 
+called "nummaterial\_velocity\_file" that will have just been created and change it from the prototype created:
+
+    0 1 vol1 --> syntax: \#material\_domain\_id \#material\_id \#rho \#vp \#vs \#Q\_kappa \#Q\_mu \#anisotropy
+    0 2 vol2 --> syntax: \#material\_domain\_id \#material\_id \#rho \#vp \#vs \#Q\_kappa \#Q\_mu \#anisotropy
+
+(where "vol1" and "vol2" here represent the volume labels that you have set while creating the mesh in CUBIT)
+to for instance
+
+    2 1 1500 2300 1800 9999.0 9999.0 0
+    2 2 1600 2500 20000 9999.0 9999.0 0
 
@@
 to set up the fault database
-\item Function $\mathtt{cubit2specfem3d.export2SESAME}$ must be called
+\item Function $\mathtt{cubit2specfem3d.export2SPECFEM3D}$ must be called
 at the very end of the script to export the mesh in a SPECFEM3D format.

### 2013-09-20 doc/USER_MANUAL/manual_SPECFEM3D_Cartesian.tex
deleted CUBIT_GEOCUBIT/run_boundary_definition.py, since the same thing is done better by run_cubit2specfem3d.py

@@
 
-\subsection{Exporting the Mesh with \texttt{cubit2specfem3d.py} }
+\subsection{Exporting the Mesh with \texttt{run\_cubit2specfem3d.py} }
 
@@
 define blocks specifying the materials and absorbing boundaries in
-CUBIT. This process could be done automatically using the script \texttt{boundary\_definition.py}
+CUBIT. This process could be done automatically using the script \texttt{run\_cubit2specfem3d.py}
 if the mesh meets some conditions or manually, following the block
@@
 
-In directory "CUBIT\_GEOCUBIT" we provide two scripts that can help doing the above tasks of exporting a CUBIT mesh
-to SPECFEM3D format automatically for you: "run\_cubit2specfem3d.py" and "run\_boundary\_definition.py".
+In directory "CUBIT\_GEOCUBIT" we provide a script that can help doing the above tasks of exporting a CUBIT mesh
+to SPECFEM3D format automatically for you: "run\_cubit2specfem3d.py".
 Just edit them to indicate the path to your local installation of CUBIT and also the name of the *.cub existing CUBIT mesh file

### 2013-09-26 CUBIT_GEOCUBIT/doc/geocubit_manual.tex
modified CUBIT_GEOCUBIT/ to match clone from github on 25-Sept-2013

@@
 %% Do not edit unless you really know what you are doing.
-\documentclass[a4paper,11pt,english]{article}
-\setlength{\textwidth}{450pt}
-\setlength{\oddsidemargin}{0pt}
-\setlength{\marginparwidth}{0pt}
+\documentclass[a4paper,11pt,oneside,english]{article}
 \setcounter{secnumdepth}{3}
@@
 \usepackage[T1]{fontenc}
-\usepackage[latin9]{inputenc}
+\usepackage[utf8]{inputenc}
+\usepackage[margin=2.5cm]{geometry}
 \usepackage{babel}
-\usepackage{setspace}
 \usepackage{listings}
-\usepackage{graphicx,subfigure}
-\usepackage{pdfsync}
+\usepackage{graphicx,subcaption}
 \usepackage{colortbl}
@@
  bookmarks=true,bookmarksnumbered=true,bookmarksopen=false,
- breaklinks=false,pdfborder={0 0 0},backref=section,colorlinks=true,pdfpagemode=None]
+ breaklinks=false,pdfborder={0 0 0},backref=section,colorlinks=true]
  {hyperref}
-\hypersetup{
- pdftex,pdfstartview=FitH,pdfpagelayout=SinglePage}
+\hypersetup{pdfstartview=FitH,pdfpagelayout=SinglePage}
 
-\newenvironment{lyxcode}
-{\par\begin{list}{}{
-\setlength{\rightmargin}{\leftmargin}
-\setlength{\listparindent}{0pt}% needed for AMS classes
-\setlength{\itemsep}{0pt}
-\setlength{\parsep}{0pt}
-\normalfont\ttfamily}% 
-\item[]}
-{\end{list}}
+\newcommand{\cubit}{\textsc{Cubit}}
+\newcommand{\geocubit}{\textsc{GeoCubit}}
+\newcommand{\specfem}{\textsc{Specfem3D}}
+
+\definecolor{codebg}{RGB}{238,238,238}
+\definecolor{codeframe}{RGB}{204,204,204}
+\lstset{
+  % The bash style works pretty well for cfg files as well.
+  language=bash,
+  basicstyle=\small\ttfamily,
+  breakatwhitespace=true,
+  breaklines=true,
+  escapechar=\@,
+  backgroundcolor=\color{codebg},
+  frame=single,
+  framesep=10pt,
+  rulecolor=\color{codeframe}
+}
 
@@
 
-\title{GEOCUBIT USER MANUAL}
+\title{\geocubit\ User Manual}
 \author{Emanuele Casarotti}
@@
 \tableofcontents
-\newpage{}
 
-\section{WHY}
+\clearpage
+\section{Why?}
 
-GEOCUBIT is a python library wrapping around the \href{http://cubit.sandia.gov} {CUBIT} Python Interface and it aims to facilitate the meshing
-process in some common problems in seismic wave propagation.\\ 
-In particular, it is focused on the meshing requests of SPECFEM3D and it is helpful for some tedious tasks as:\\
-
+\geocubit\ is a Python library wrapping around the \href{http://cubit.sandia.gov} {\cubit}
+Python Interface and it aims to facilitate the meshing
+process in some common problems in seismic wave propagation.
+In particular, it is focused on the meshing requests of \specfem\ and it is helpful for such tedious tasks as:
+%
 \begin{itemize}
@@
 basin (or slow velocity zones) are present.
-\item It can be used as serial or parallel process. The parallel meshing
+\item It can be used as a serial or parallel process. The parallel meshing
 capabilities are fundamental for large geophysical problems (ex. mesh
@@
 \end{itemize}
-GEOCUBIT can be use inside the graphical interface of \href{http://cubit.sandia.gov} {CUBIT} (i.e.
-as python object in the script tab) or as unix command. We refer to the \href{http://cubit.sandia.gov} {CUBIT} help appendix for more information about the python interface.
 
-\newpage{}
-\section{FIRST STEPS}
+\geocubit\ can be used inside the graphical interface of \cubit\ (i.e.
+as a Python object in the script tab) or as unix command. We refer to the
+\cubit\ help appendix for more information about the Python interface.
+
+\clearpage
+\section{First Steps}
 
@@
 
-The minimum requirements for using GEOCUBIT are:
-
+The minimum requirements for using \geocubit\ are:
+%
 \begin{itemize}
-\item \href{http://cubit.sandia.gov} {CUBIT 12.2}
-\item \href{http://downloads.sourceforge.net/numpy} {numpy 1.0+}
-\item \href{http://www.python.org} {python 2.5}
- (strictly! it depends on the cubit library that refers to this version of python)
+\item \href{http://cubit.sandia.gov} {\cubit\ 12.2}
+\item \href{http://downloads.sourceforge.net/numpy} {NumPy 1.0+}
+\item \href{http://www.python.org} {Python 2.5}
+ (strictly! it depends on the \cubit\ library that refers to this version of Python)
 \end{itemize}
-\raggedright{and  for using the parallel meshing capabilities:}
+%
+and for using the parallel meshing capabilities:
+%
 \begin{itemize}
@@
 For installing, download the code and type in the GEOCUBIT directory:
-\begin{lyxcode}
-python2.5 setup.py install~
-\end{lyxcode}
+%
+\begin{lstlisting}
+python2.5 setup.py install
+\end{lstlisting}
+%
+and check that the following variables are set:
+%
+\begin{lstlisting}
+CUBITDIR="/usr/local/CUBIT"
+CUBITLIB="$CUBITDIR/bin:$CUBITDIR/structure:$CUBITDIR/components"
+PYTHONPATH="$CUBITDIR/components/cubit:$CUBITDIR/structure:$CUBITDIR/bin"
+LD_LIBRARY_PATH="$CUBITDIR/bin"
+PATH="$CUBITDIR/bin:$CUBITDIR/components/cubit"
+\end{lstlisting}
 
-Check that the following variables are set:
+\clearpage
+\section{Using \geocubit\ from the Command Line}
 
-\begin{lyxcode}
-CUBITDIR=/usr/local/CUBIT\\
-CUBITLIB=\$CUBITDIR/bin:\$CUBITDIR/structure:\$CUBITDIR/components\\
-PYTHONPATH=\$CUBITDIR/components/cubit:\$CUBITDIR/structure:\$CUBITDIR/bin\\
-LD\_LIBRARY\_PATH=\$CUBITDIR/bin\\
-PATH=\$CUBITDIR/bin:\$CUBITDIR/components/cubit \\
-\end{lyxcode}
+\subsection{Utilities}
 
-\newpage{}
-\section{USING GEOCUBIT AT COMMAND LINE:}
+\begin{itemize}
+\item Checking the configuration of the libraries and dependencies:
+\begin{lstlisting}
+GEOCUBIT.py --chklib
+\end{lstlisting}
 
-\subsection{UTILITIES}
-\begin{itemize}
-\item checking the configuration of the libraries and dependencies:\\
-\begin{lyxcode}
-GEOCUBIT.py --chklib
-\end{lyxcode}
-
-\item checking the parameter file:\\
-\begin{lyxcode}
+\item Checking the parameter file:
+\begin{lstlisting}
 GEOCUBIT.py --chkcfg --cfg=[file]
-\end{lyxcode}
+\end{lstlisting}
 \end{itemize}
 
-\subsection{CREATE GEOMETRY}
-SURFACES
+\subsection{Create Geometry}
+
+\subsubsection{Surfaces}
 
 \begin{itemize}
-\item creating acis surfaces using a parameter file:\\
-\begin{lyxcode}
- GEOCUBIT.py --build\_surface --cfg=[filename]
-\end{lyxcode}
+\item Creating acis surfaces using a parameter file:
+\begin{lstlisting}
+ GEOCUBIT.py --build_surface --cfg=[filename]
+\end{lstlisting}
 \end{itemize}
@@
 \begin{itemize}
-\item creating a surface from regular ascii grid or ascii lines that define a "skin":\\
-\begin{lyxcode}
+\item Creating a surface from regular ascii grid or ascii lines that define a "skin":
+\begin{lstlisting}
  GEOCUBIT.py --surface=[file] (--regulargrid=[opts] | --skin=[opts])
-\end{lyxcode}
+\end{lstlisting}
 \end{itemize}
 
-VOLUMES
+\subsubsection{Volumes}
+
 \begin{itemize}
-\item **serial** command: creating cubit volumes using a parameter file:\\
-\begin{lyxcode}
+\item \emph{Serial} command: creating \cubit\ volumes using a parameter file:
+\begin{lstlisting}
 GEOCUBIT.py --surface=[file] (--regulargrid=[opts] | --skin=[opts])
-\end{lyxcode}
+\end{lstlisting}
 \end{itemize}
@@
 \begin{itemize}
-\item **parallel** command: creating cubit volumes using a parameter file:\\
-\begin{lyxcode}
-mpirun -n [numproc] pyMPI GEOCUBIT.py --build\_volume --cfg=[file]
-\end{lyxcode}
+\item \emph{Parallel} command: creating \cubit\ volumes using a parameter file:
+\begin{lstlisting}
+mpirun -n [numproc] pyMPI GEOCUBIT.py --build_volume --cfg=[file]
+\end{lstlisting}
 
-\begin{figure}\begin{center}
-    \subfigure[]{\includegraphics[scale=.5]{media/geocubitmanual_grid2.jpg}
-    \label{fig:vol}}%
-    \subfigure[]{\includegraphics[scale=.9]{media/geocubitmanual_grid.jpg}
-    \label{fig:table}}%
-\caption{a) Volume divided in slices. b) Parallel map of the IDs for volume in Figure~\ref{fig:vol} divided in (16 slices, 4 along x-axis and 4 along y-axis).}
+\begin{figure}[htbp]
+  \centering
+  \begin{subfigure}{0.45\textwidth}
+    \includegraphics[width=\textwidth,keepaspectratio]{media/geocubitmanual_grid2.jpg}
+    \caption{Volume divided in slices.}
+    \label{fig:vol}
+  \end{subfigure}
+  \begin{subfigure}{0.45\textwidth}
+    \includegraphics[width=\textwidth,keepaspectratio]{media/geocubitmanual_grid.jpg}
+    \caption{Parallel map of the IDs for volume in \autoref{fig:vol} divided in 16 slices (4 along x-axis and 4 along y-axis).}
+    \label{fig:table}
+  \end{subfigure}
+  \caption{Volume slicing.}
   \label{fig:grid}
-\end{center}
 \end{figure}
 
-In this parallel application, the volume is separated in N slices, each one is assigned at a single process and one file for each slice is created (see Figure~\ref{fig:grid} as example).
+In this parallel application, the volume is separated in $N$ slices, each one is assigned at a single process and one file for each slice is created (see \autoref{fig:grid} as example).
 \end{itemize}
 
-\subsection{MESHING}
+\subsection{Meshing}
+
 \begin{itemize}
-\item **serial** command: building a volume and meshing it.\\
-\begin{lyxcode}
-GEOCUBIT.py --build\_volume --mesh --cfg=[file] (--id\_proc=n)
-\end{lyxcode}
+\item \emph{Serial} command: building a volume and meshing it.
+\begin{lstlisting}
+GEOCUBIT.py --build_volume --mesh --cfg=[file] (--id_proc=n)
+\end{lstlisting}
+
+\item \emph{Serial} command: meshing a volume
+\begin{lstlisting}
+GEOCUBIT.py --mesh --cfg=[file] (--id_proc=[n, def=0])
+\end{lstlisting}
+
+{\footnotesize note: without the \lstinline|--build_volume| flag the script recall an old "\lstinline|geometry_vol_[id_proc].cub|" file}
+
+\item \emph{Parallel} command: meshing a volume from a parameter file
+\begin{lstlisting}
+mpirun -n [nproc] pyMPI GEOCUBIT.py --build_volume --mesh --cfg=[file]
+\end{lstlisting}
 \end{itemize}
 
-\begin{itemize}
-\item **serial** command : meshing a volume\\
-\begin{lyxcode}
-GEOCUBIT.py --mesh --cfg=[file] (--id\_proc=[n, def=0])
-\end{lyxcode}
-\footnotesize note: without the --build\_volume flag the script recall an old "geometry\_vol\_[id\_proc].cub" file\\
-\end{itemize}
+\subsection{Finalizing and Exporting}
 
-\begin{itemize}
-\item **parallel** command: meshing a volume from a parameter file\\
-\begin{lyxcode}
-mpirun -n [nproc] pyMPI GEOCUBIT.py --build\_volume --mesh --cfg=[file]
-\end{lyxcode}
-\end{itemize}
-
-\subsection{FINALIZING AND EXPORTING}
-
-It is possible to collect, merge, set the absorbing boundary conditions and to export the resulting mesh in a format readable by SPECFEM3D.
+It is possible to collect, merge, set the absorbing boundary conditions and to export the resulting mesh in a format readable by \specfem.
 \begin{itemize}
 \item
-Collecting some cubit files, setting the absorbing boundary conditions and merging in a single free mesh cubitfile
-\begin{lyxcode}
+Collecting some \cubit\ files, setting the absorbing boundary conditions and merging in a single free mesh cubitfile
+\begin{lstlisting}
 GEOCUBIT.py --collect --merge --meshfiles=[files] --cpux=N --cpuy=N (--rangecpux=[cpuxmin,cpuxmax] --rangecpuy=[cpuymin,cpuymax])
-\end{lyxcode}
-\texttt{cpuy, cpux} set the number of cpus used for creating the mesh files; \texttt{rangecpux} and \texttt{rangecpuy} set the range of slices that are used in the mesh.\\
-Following the example in Figure~\ref{fig:grid}, the parameters \texttt{--cpux=4 --cpuy=4 --rangecpux=1,2 --rangecpuy=2,3} select the slices with id 10 11 14 15.
-Only these slices are going to be collect and merged with the appropriated absorbing boundary conditions.
+\end{lstlisting}
 
-\item Collecting a single free mesh cubitfile and refine the hex inside some curves (ex. alluvial basin, Figure~\ref{fig:refcurve})
-\begin{lyxcode}
-GEOCUBIT.py --collect --meshfiles=[list of files] --curverefining=[file]\\
-\end{lyxcode}
-\footnotesize note: the curves must be stored in acis format (sat) and must be closed.
+\lstinline{cpux, cpuy} set the number of cpus used for creating the mesh files; \lstinline{rangecpux} and \lstinline{rangecpuy} set the range of slices that are used in the mesh.
 
-\end{itemize}
+Following the example in \autoref{fig:grid}, the parameters \lstinline{--cpux=4 --cpuy=4 --rangecpux=1,2 --rangecpuy=2,3} select the slices with id $10, 11, 14, 15$.
+Only these slices are going to be collected and merged with the appropriate absorbing boundary conditions.
 
-\begin{figure}\center
-\includegraphics[scale=.37]{media/refcurve_1.png}%
-\includegraphics[scale=.303]{media/refcurve_2.png}%
-\caption{}
+\item Collecting a single free mesh cubitfile and refine the hex inside some curves (ex. alluvial basin, \autoref{fig:refcurve})
+\begin{lstlisting}
+GEOCUBIT.py --collect --meshfiles=[list of files] --curverefining=[file]
+\end{lstlisting}
+
+{\footnotesize note: the curves must be stored in acis format (sat) and must be closed.}
+
+\begin{figure}[htbp]
+\center
+\includegraphics[height=0.4\textheight,keepaspectratio]{media/refcurve_1.png}%
+\includegraphics[height=0.4\textheight,keepaspectratio]{media/refcurve_2.png}%
+\caption{An alluvial basin example.}
 
@@
 
-\begin{itemize}
-\item Exporting a cubit mesh file in a SPECFEM3D mesh
-\begin{lyxcode}
-GEOCUBIT.py --export2SPECFEM3D=[output directory] --meshfiles=[filename] (--listblock=[list of cubit blocks] --listflag=[list of specfem3d flags])
-\end{lyxcode}
-\footnotesize If required, it is possible to assign personalized flags to each block using --listblock and the correspondent 	listflag (for example: --listblock=3,4 --listflag=1,-1)
+\item Exporting a \cubit\ mesh file to a \specfem\ mesh
+\begin{lstlisting}
+GEOCUBIT.py --export2SPECFEM3D=[output directory] --meshfiles=[filename] (--listblock=[list of @\cubit@ blocks] --listflag=[list of specfem3d flags])
+\end{lstlisting}
+
+If required, it is possible to assign personalized flags to each block using \lstinline{--listblock} and the corresponding listflag (for example: \lstinline{--listblock=3,4 --listflag=1,-1})
 \end{itemize}
 
-\newpage{}
-\section{GEOCUBIT and CUBIT GRAPHICAL INTERFACE}
-In the python script tab of CUBIT GUI, type:
+\clearpage
+\section{\geocubit\ and \cubit\ Graphical Interface}
 
-\begin{lyxcode}
->f="volume.cfg" $\to$  \textrm{\small{\textit{store the name of the configuration file, see next section}}}\\
-\end{lyxcode}
+In the Python script tab of \cubit\ GUI, type:
 
-\begin{lyxcode}
->from geocubitlib import volumes $\to$  \textrm{\small{\textit{load the geocubit modules}}}\\
->from geocubitlib import mesh\_volume\\
->from geocubitlib import exportlib\\
-\end{lyxcode}
+\begin{lstlisting}
+>f="volume.cfg" # store the name of the configuration file, see next section
 
-\begin{lyxcode}
-volumes.volumes(f) $\to$  \textrm{\small{\textit{create the volumes, the parameters are stored in the cfg file}}}\\
-mesh\_volume.mesh(f) $\to$  \textrm{\small{\textit{mesh the volume}}}\\
-\end{lyxcode}\begin{lyxcode}
-exportlib.collect() $\to$  \textrm{\small{\textit{set the boundary conditions, see SPECFEM3D manual}}}\\
-exportlib.e2SEM(outdir="./output") $\to$  \textrm{\small{\textit{save the mesh in the SPECFEM3D format}}}\\
-\end{lyxcode}
-See the media/iterative.mov for a live example.
+>from geocubitlib import volumes # load the @\geocubit@ modules
+>from geocubitlib import mesh_volume
+>from geocubitlib import exportlib
 
-\section{EXAMPLES of CONFIGURATION FILES}
-In the current section a few configuration files will be described. Please, see the files in the examples directory of the GEOCUBIT distribution for reference.\\
+volumes.volumes(f)  # create the volumes, the parameters are stored in the cfg file
+mesh_volume.mesh(f) # mesh the volume
 
-\subsection{GENERAL OPTIONS}
+exportlib.collect()                # set the boundary conditions, see @\specfem@ manual
+exportlib.e2SEM(outdir="./output") # save the mesh in the @\specfem@ format
+\end{lstlisting}
+
+See the \lstinline|media/iterative.mov| for a live example.
+
+\section{Examples of Configuration Files}
+
+In the current section a few configuration files will be described. Please see the files in the examples directory of the \geocubit\ distribution for reference.
+
+\subsection{General Options}
 
 A cfg file has a format similar to ini windows files.
-They are divided in \texttt{[sections]}. There is no order in the position or in the name of the sections.
+They are divided in \lstinline{[sections]}. There is no order in the position or in the name of the sections.
 
-Generally, the parameters that control the general options are in the first section, called \texttt{[cubit.options]} and \texttt{[simulation.cpu\_parameters]}. For example:
+Generally, the parameters that control the general options are in the first section, called \lstinline{[cubit.options]} and \lstinline{[simulation.cpu_parameters]}. For example:
 
-\begin{lyxcode}
-[cubit.options]\\
-cubit\_info=off $\to$  \textrm{\small{\textit{turn on/off the information of the cubit command}}}\\
-echo\_info=off $\to$   \textrm{\small{\textit{turn on/off the echo of the cubit command}}}\\
-cubit\_info=off $\to$   \textrm{\small{\textit{turn on/off the cubit journaling}}}\\
-echo\_info=off $\to$   \textrm{\small{\textit{turn on/off the cubit error journaling}}}\\
-working\_dir=tmp $\to$   \textrm{\small{\textit{set the working directory}}}\\
-output\_dir=output $\to$   \textrm{\small{\textit{set the output directory}}}\\
-save\_geometry\_cubit = True $\to$   \textrm{\small{\textit{true if it saves the geometry files}}}\\
-save\_surface\_cubit = False $\to$   \textrm{\small{\textit{true if it saves the surfaces in a cubit files}}}\\
-export\_exodus\_mesh = True $\to$   \textrm{\small{\textit{true if it saves the mesh also in exodus format.
-The default is the cubit format (cub), that contains both the geometry and the meshing information.
-In case of complex geometry, the cub file could be enormous and a more light exodus file become important.}}}\\
-\end{lyxcode}
-\begin{lyxcode}
-[simulation.cpu\_parameters]\\
-nodes = 1  $\to$   \textrm{\small{\textit{number of nodes/process}}}\\
+\begin{lstlisting}
+[cubit.options]
+cubit_info=off             # turn on/off the information of the @\cubit@ command
+echo_info=off              # turn on/off the echo of the @\cubit@ command
+cubit_info=off             # turn on/off the @\cubit@ journaling
+echo_info=off              # turn on/off the @\cubit@ error journaling
+working_dir=tmp            # set the working directory
+output_dir=output          # set the output directory
+save_geometry_cubit = True # true if it saves the geometry files
+save_surface_cubit = False # true if it saves the surfaces in a @\cubit@ files
+export_exodus_mesh = True  # true if it saves the mesh also in exodus format.
+The default is the @\cubit@ format (cub), that contains both the geometry and the meshing information.
+In case of complex geometry, the cub file could be enormous and a more light exodus file become important.
 
-\end{lyxcode}
+[simulation.cpu_parameters]
+nodes = 1  # number of nodes/process
+\end{lstlisting}
 
-\footnotesize note: Usually the *\_info options are turn off by default for improving the performances.\normalsize\\
+{\footnotesize note: Usually the \lstinline|*_info| options are turned off by default for improving the performance.}
 
-\subsection{CREATING A TOPOGRAPHIC SURFACE}
-\footnotesize See stromboli.cfg, execute with: \verb!GEOCUBIT.py --build\_surface --cfg=./example/stromboli.cfg. !\normalsize\\
-~\\
-GEOCUBIT is able to create a topographic surface based upon the CUBIT command \verb!create skin curve! and \verb!create surface net! (Figure~\ref{fig:surfacetype}). See the CUBIT manual for details.
+\subsection{Creating a Topographic Surface}
 
-\begin{lyxcode}
+{\footnotesize See stromboli.cfg, execute with: \lstinline|GEOCUBIT.py --build_surface --cfg=./example/stromboli.cfg|.}
+
+\geocubit\ is able to create a topographic surface based upon the \cubit\ command \lstinline|create skin curve| and \lstinline|create surface net| (\autoref{fig:surfacetype}).
+See the \cubit\ manual for details.
+
+\begin{lstlisting}
 [geometry.surfaces]
-nsurf = 2 $\to$   \textrm{\small{\textit{number of surfaces that will be created}}}\\
-\end{lyxcode}
-\begin{lyxcode}
-[surface\textbf{1}.parameters]\\
-name=example/data/stromboli.xyz  $\to$   \textrm{\small{\textit{the name of the file defining the surface}}}\\
-surf\_type=skin $\to$   \textrm{\small{\textit{the type of surface:} \texttt{skin} \textit{, the file is a sequence of parallel points that span the surface in the order describer by the following direction parameters} \texttt{regular\_grid} \emph{the points are structured distritibuted.
-see Cubit Manual for details(skin surface and net surface) }}}\\
-unit\_surf= utm $\to$   \textrm{\small{\textit{unit of the points,} \texttt{utm} \emph{or} \texttt{geo} \emph{(geographical) coordinates}}}\\
-directionx = 0 $\to$   \textrm{\small{\textit{if} \texttt{surf\_type=skin}\emph{:} \texttt{directionx} \emph{and} \texttt{directiony} \emph{define how the points span the surface}}}\\
-directiony = 1\\
-step = 1 $\to$   \textrm{\small{\textit{the script creates a vertex each} \texttt{step} \emph{points: in this case} \texttt{step=1} \textit{means a vertex for all the points}}}\\
-\end{lyxcode}
+nsurf = 2 # number of surfaces that will be created
 
-\begin{lyxcode}
-[surface2.parameters] $\to$ \textrm{\small{\emph{note the different number of surface in the section name}}}
-name=./examples/surfaces/topo.dat $\to$   \textrm{\small{\textit{the name of the file defining the surface}}}\\
-surf\_type=regular\_grid $\to$   \textrm{\small{\textit{the type of surface:} \texttt{skin} \textit{, the file is a sequence of parallel points that span the surface in the order describer by the following direction parameters} \texttt{regular\_grid} \emph{the points are structured distritibuted.
-See see Cubit Manual for details (skin surface and net surface) }}}\\
-unit\_surf= geo $\to$   \textrm{\small{\textit{unit of the points,} \texttt{utm} \emph{or} \texttt{geo} \emph{(geographical) coordinates}}}\\
-nx=5 $\to$   \textrm{\small{\textit{In case of} \texttt{regular\_grid}: \texttt{nx} \emph{and} \texttt{ny} \emph{are the number of points along x and y direction.}}}\\
+[surface1.parameters]
+name=example/data/stromboli.xyz # the name of the file defining the surface
+surf_type=skin # the type of surface: @\texttt{skin}@: the file is a sequence of parallel points that span the surface in the order described by the following direction parameters; @\texttt{regular\_grid}@: the points are structured distributed.
+See @\cubit@ Manual for details (skin surface and net surface)
+unit_surf = utm # unit of the points, @\texttt{utm}@ or @\texttt{geo}@ (geographical) coordinates
+directionx = 0 # if surf_type=skin: @\texttt{directionx}@ and @\texttt{directiony}@ define how the points span the surface
+directiony = 1
+step = 1 # the script creates a vertex each @\texttt{step}@ points: in this case @\texttt{step=1}@ means a vertex for all the points
+
+[surface2.parameters] # note the different number of surface in the section name
+name=./examples/surfaces/topo.dat # the name of the file defining the surface
+surf_type=regular_grid # the type of surface: @\texttt{skin}@: the file is a sequence of parallel points that span the surface in the order described by the following direction parameters; @\texttt{regular\_grid}@: the points are structured distributed.
+See see @\cubit@ Manual for details (skin surface and net surface)
+unit_surf= geo # unit of the points, @\texttt{utm}@ or @\texttt{geo}@ (geographical) coordinates
+nx=5 # In case of @\texttt{regular\_grid}@: @\texttt{nx}@ and @\texttt{ny}@ are the number of points along x and y direction.
 ny=5
-\end{lyxcode}
+\end{lstlisting}
 
-\begin{figure}\begin{center}
-    \subfigure[]{\includegraphics[scale=.4]{media/skinsurface.png}
-    \label{fig:skin}}%
-    \subfigure[]{\includegraphics[scale=.4]{media/netsurface.png}
-    \label{fig:net}}%
-    \subfigure[]{\includegraphics[scale=.25]{media/strombolisurface.png}
-    \label{fig:stromboli}}%
-\caption{GEOCUBIT creates a topographic surface in 2 ways: a) skin topography, \texttt{directionx=0} and \texttt{directiony=1}, b) topography from a net of structured distritibuted points, \texttt{nx=4} and \texttt{ny=4}, c) Topography of Stromboli volcano (Italy) using the net surface method.}
+\begin{figure}[htbp]
+  \centering
+  \begin{subfigure}{0.4\textwidth}
+    \includegraphics[width=\textwidth,keepaspectratio]{media/skinsurface.png}
+    \caption{Skin topography, \texttt{directionx=0} and \texttt{directiony=1}}
+    \label{fig:skin}
+  \end{subfigure}
+  \begin{subfigure}{0.4\textwidth}
+    \includegraphics[width=\textwidth,keepaspectratio]{media/netsurface.png}
+    \caption{Topography from a net of structured distributed points, \texttt{nx=4} and \texttt{ny=4}}
+    \label{fig:net}
+  \end{subfigure}
+
+  \begin{subfigure}{0.45\textwidth}
+    \includegraphics[width=\textwidth,keepaspectratio]{media/strombolisurface.png}
+    \caption{Topography of Stromboli volcano (Italy) using the net surface method.}
+    \label{fig:stromboli}
+  \end{subfigure}
+  \caption{Two methods of creating a topographic surface in \geocubit.}
   \label{fig:surfacetype}
-\end{center}
 \end{figure}
 
-\subsection{CREATING and MESHING A VOLUME}
-\footnotesize see volume.cfg, for executioning: \verb!GEOCUBIT.py --build\_volume --mesh --cfg=./example/volume.cfg !\normalsize\\
-~\\
-The example is for a volume with topography with dimension of the hexahedra increasing with depth by means of a refinement layer (Figure~\ref{fig:volume}).
+\subsection{Creating and Meshing a Volume}
 
-\begin{figure}\begin{center}
-\includegraphics[scale=.4]{media/geometryvolume.png}%
-\includegraphics[scale=.377]{media/meshvolume.png}%
+{\footnotesize see volume.cfg, for executioning: \lstinline|GEOCUBIT.py --build_volume --mesh --cfg=./example/volume.cfg|}
+
+The example is for a volume with topography with dimension of the hexahedra increasing with depth by means of a refinement layer (\autoref{fig:volume}).
+
+\begin{figure}[htbp]
+\centering
+\includegraphics[height=0.24\textheight,keepaspectratio]{media/geometryvolume.png}%
+\includegraphics[height=0.24\textheight,keepaspectratio]{media/meshvolume.png}%
 \caption{Mesh of a simple volume}
@@
 \label{fig:volume}
-\end{center}
 \end{figure}
 
-\begin{lyxcode}
-[geometry.volumes]\\
-volume\_type = layercake\_volume\_ascii\_regulargrid\_regularmap\\
-latitude\_min                    = 13.879\\
-latitude\_max                    = 14.279\\
-longitude\_min                   = 40.619\\
-longitude\_max                   = 40.969\\
-nx = 5 \\
-ny = 5\\
-unit                            = geo\\
-\end{lyxcode}
+\begin{lstlisting}
+[geometry.volumes]
+volume_type = layercake_volume_ascii_regulargrid_regularmap
+latitude_min                    = 13.879
+latitude_max                    = 14.279
+longitude_min                   = 40.619
+longitude_max                   = 40.969
+nx = 5
+ny = 5
+unit                            = geo
 
-\begin{lyxcode}
-[geometry.volumes.layercake]\\
-nz = 2\\
-bottomflat = True\\
-depth\_bottom = -7000\\
-filename = ./example/topo.dat\\
-geometry\_format=ascii\\
-\end{lyxcode}
+[geometry.volumes.layercake]
+nz = 2
+bottomflat = True
+depth_bottom = -7000
+filename = ./example/topo.dat
+geometry_format=ascii
+\end{lstlisting}
 
-\texttt{nz} is the number of the horizontal surfaces in the volume (in this case: topography and bottom).\\
-\texttt{bottomflat} is True if the bottom is flat.\\
-\texttt{depth\_bottom} is the depth of the bottom.\\
-\texttt{filename} is a list of files. Each one defines a surface (in this case there is only the topography file in a regular\_grid format).\\
-\texttt{geometry\_format} is set to ascii since the definition of the surfaces comes from ascii files (structured xyz points). \\
+\begin{description}
+\item[nz] is the number of the horizontal surfaces in the volume (in this case: topography and bottom).
+\item[bottomflat] is \lstinline|True| if the bottom is flat.
+\item[depth\_bottom] is the depth of the bottom.
+\item[filename] is a list of files. Each one defines a surface (in this case there is only the topography file in a \lstinline|regular_grid| format).
+\item[geometry\_format] is set to \lstinline|ascii| since the definition of the surfaces comes from ASCII files (structured xyz points).
+\end{description}
 
-\begin{lyxcode}
-[meshing]\\
-map\_meshing\_type=regularmap\\
-iv\_interval=5, $\to$   \textrm{\small{\textit{if only one value is present, append the comma}}}\\
-size=2000\\
-or\_mesh\_scheme=map\\
-ntripl=1\\
-tripl=2, $\to$   \textrm{\small{\textit{if only one value is present, append the comma}}}\\
-smoothing=False\\
-\end{lyxcode}
+\begin{lstlisting}
+[meshing]
+map_meshing_type=regularmap
+iv_interval=5, # if only one value is present, append the comma
+size=2000
+or_mesh_scheme=map
+ntripl=1
+tripl=2, # if only one value is present, append the comma
+smoothing=False
+\end{lstlisting}
 
-The \texttt{meshing} section contains the parameters request for the meshing process.
-\texttt{map\_meshing\_type} set the meshing scheme and is regularmap by default (other schemes are in preliminary phase).\\
-\texttt{iv\_interval} set the number of "horizontal" hex sheets for each layer.
-\texttt{size} is the dimension of hexahedral (horizontally)\\
-\texttt{or\_mesh\_scheme} is the meshing scheme for the topography (\texttt{map} or \texttt{pave} are possible, see the CUBIT manual for more information).\\
-\texttt{ntripl} is the number of tripling layer in the mesh (in this case 1).\\
-\texttt{tripl} means in this case that the the refinement layer is located at the second surface (the topography). The surfaces are ordered from the bottom (surface 1) to the top (surface 2).\\
-\texttt{smoothing} performes the smoothing command in Cubit.\\
+The \lstinline|[meshing]| section contains the parameters requested for the meshing process.
 
-\subsection{LAYERED VOLUME and MESH FOR CENTRAL ITALY (PARALLEL)}
+\begin{description}
+\item[map\_meshing\_type] sets the meshing scheme and is \lstinline|regularmap| by default (other schemes are in preliminary phase).
+\item[iv\_interval] sets the number of "horizontal" hex sheets for each layer.
+\item[size] is the dimension of hexahedral (horizontally).
+\item[or\_mesh\_scheme] is the meshing scheme for the topography (\lstinline|map| or \lstinline|pave| are possible, see the \cubit\ manual for more information).
+\item[ntripl] is the number of tripling layer in the mesh (in this case 1).
+\item[tripl] means in this case that the refinement layer is located at the second surface (the topography). The surfaces are ordered from the bottom (surface 1) to the top (surface 2).
+\item[smoothing] performs the smoothing command in \cubit.
+\end{description}
 
-\footnotesize In the example abruzzo.cfg, the layers are 2. For execution type: \texttt{mpirun -n 150 GEOCUBIT.py --build\_volume --mesh --cfg=./example/abruzzo.cfg}.\normalsize\\
-~\\
+\subsection{Layered Volume and Mesh for Central Italy (in Parallel)}
 
-Comparing the previous cfg files, there are few modification:
+In the example abruzzo.cfg, the layers are 2. For execution, type: \lstinline|mpirun -n 150 GEOCUBIT.py --build_volume --mesh --cfg=./example/abruzzo.cfg|.
 
-\begin{lyxcode}
-$\cdots $\\
-nz = 3 \\
-$\cdots $\\
-filename = example/data/moho\_int.xyz,example/data/topo.xyz\\
-$\cdots $\\
-iv\_interval=8,8 (one value for each layer)\\
-$\cdots $\\
-refinement\_depth=1,\\
-\end{lyxcode}
+Comparing the previous cfg files, there are few modifications:
+%
+\begin{lstlisting}
+. . .
+nz = 3
+. . .
+filename = example/data/moho_int.xyz,example/data/topo.xyz
+. . .
+iv_interval=8,8 (one value for each layer)
+. . .
+refinement_depth=1,
+\end{lstlisting}
 
-The volume has \texttt{nz=3} interfaces: bottom, moho, and topography. The bottom is flat and  his z-coordinate position is set by \texttt{depth\_bottom}. 
+The volume has \lstinline|nz=3| interfaces: bottom, moho, and topography. The bottom is flat and its z-coordinate position is set by \lstinline|depth_bottom|.
 The name of the files that storage the data for the other interfaces are listed in \texttt{filename}.
 The interfaces defines 2 geological layers.
-Each layer has \texttt{iv\_interval=8} "horizontal" hex sheets.
-There is only a refinement and it is set in the \texttt{refinement\_depth=1} hex sheet, i.e just below the topography (\texttt{refinement\_depth=8} means just below the moho).
+Each layer has \lstinline|iv_interval=8| "horizontal" hex sheets.
+There is only a refinement and it is set in the \lstinline|refinement_depth=1| hex sheet, i.e just below the topography (\lstinline|refinement_depth=8| means just below the moho).
 
-\subsection{CREATION OF A REGULAR MESH}
+\subsection{Creation of a Regular Mesh}
 
-\footnotesize In the example grid.cfg, the layers are 3. For execution type: \texttt{GEOCUBIT.py --build\_volume --mesh --cfg=./example/grid.cfg}.\normalsize\\
-~\\
+In the example grid.cfg, the layers are 3. For execution, type: \lstinline|GEOCUBIT.py --build_volume --mesh --cfg=./example/grid.cfg|.
 
-In this example we create a simple layercake box \texttt{geometry\_format=regmesh} with 3 layers at depth defined by \texttt{zdepth=-7000,-3000,-600,0}.
-The initial mesh has hex with horizontal size \texttt{size=2000} and the numbers of vertical hex sheets is iv\_interval=3,1,1 respectively for the bottom, middle and top layer.
-We include a refinement layer (\texttt{ntripl=1}) at \texttt{tripl=2,} interface (the second from the bottom).
+In this example we create a simple layercake box \lstinline|geometry_format=regmesh| with 3 layers at depth defined by \lstinline|zdepth=-7000,-3000,-600,0|.
+The initial mesh has hex with horizontal size \lstinline|size=2000| and the numbers of vertical hex sheets is \lstinline|iv_interval=3,1,1| respectively for the bottom, middle and top layer.
+We include a refinement layer (\lstinline|ntripl=1|) at \lstinline|tripl=2,| interface (the second from the bottom).
 Since the refinement occurs on the vertical direction, the vertical dimensions of the hexes in the top layer is too small and the quality of the mesh decreases.
-\texttt{coarsening\_top\_layer=True} remesh the top layer and the number of vertical hex sheet in the vertical is defined by \texttt{actual\_vertical\_interval\_top\_layer=1}.
-(see Figure~\ref{fig:regrid})
+\lstinline|coarsening_top_layer=True| remesh the top layer and the number of vertical hex sheet in the vertical is defined by \lstinline|actual_vertical_interval_top_layer=1|.
+(see \autoref{fig:regrid})
 
-\begin{figure}\begin{center}
-    \subfigure[]{\includegraphics[scale=.2]{media/regrid.png}
-    \label{fig:regrid_nocoarse}}%
-    \subfigure[]{\includegraphics[scale=.47]{media/regrid_coarse.png}
-    \label{fig:regrid_coarse}}%
-\caption{Regular mesh with one tripling layer, a) mesh with coarsening\_top\_layer=False, b) mesh with \texttt{coarsening\_top\_layer=True}, the top layer is remeshed with \texttt{actual\_vertical\_interval\_top\_layer=1} vertical hex sheet.}
+\begin{figure}[htbp]
+  \centering
+  \begin{subfigure}{0.45\textwidth}
+    \includegraphics[width=\textwidth,keepaspectratio]{media/regrid.png}
+    \caption{Mesh with \lstinline|coarsening_top_layer=False|}
+    \label{fig:regrid_nocoarse}
+  \end{subfigure}
+  \begin{subfigure}{0.45\textwidth}
+    \includegraphics[width=\textwidth,keepaspectratio]{media/regrid_coarse.png}
+    \caption{\raggedright Mesh with \lstinline|coarsening_top_layer=True|. The top layer is remeshed with \lstinline|actual_vertical_interval_top_layer=1| vertical hex sheet.}
+    \label{fig:regrid_coarse}
+  \end{subfigure}
+  \caption{Regular mesh with one tripling layer showing the effect of \lstinline|coarsening_top_layer|.}
   \label{fig:regrid}
-\end{center}
 \end{figure}
 
-\subsection{TODO: MESH for SOUTHERN CALIFORNIA}
+\subsection{TODO: Mesh for Southern California}
 Mesh with high number of hexes and high resolution topography. Mesh in Parallel.
@@
 
-\subsection{TODO MESH for SANTA MONICA OVERTRUST}
+\subsection{TODO: Mesh for Santa Monica Overtrust}
 Using part of the mesh for Southern California, stitching together several slices.
@@
 
-\subsection{TODO Grouping the hexes with different period resolved}
+\subsection{TODO: Grouping Hexes with Different Resolved Periods}
 
 Checking the stability
-For execution type: \texttt{GEOCUBIT.py --meshfiles=[filename] --stability (--tomofile=[tomographic file]/--vp=[vp text value] --vs=[vs text value])}.
+For execution type: \lstinline|GEOCUBIT.py --meshfiles=[filename] --stability (--tomofile=[tomographic file]/--vp=[vp text value] --vs=[vs text value])|.
 
-or
+Or in the script tab of the \cubit\ GUI, type:
 
-in the script tab of the CUBIT GUI, type:
-
-\begin{lyxcode}
->from geocubitlib.hex_metric import SEM_stability_3D $\to$  \textrm{\small{\textit{load the geocubit modules}}}\\
->mesh=SEM_stability_3D()\\
->mesh.check_simulation_parameter(vp_static=vp,vs_static=vs) $\to$  \textrm{\small{\textit{text the mesh against some homogeneous velocity model}}}\\
-or\\
->mesh.check_simulation_parameter(tomofile=tomofile) $\to$  \textrm{\small{\textit{text the mesh against a tomographic file}}}\\
-mesh.group_timestep() $\to$  \textrm{\small{\textit{grouping hex with similar timestep}}}\\
-mesh.group_period() $\to$  \textrm{\small{\textit{grouping hex with similar period}}}\\
-\end{lyxcode}
+\begin{lstlisting}
+>from geocubitlib.hex_metric import SEM_stability_3D # load the @\geocubit@ modules
+>mesh=SEM_stability_3D()
+>mesh.check_simulation_parameter(vp_static=vp,vs_static=vs) # test the mesh against some homogeneous velocity model
+# or
+>mesh.check_simulation_parameter(tomofile=tomofile) # test the mesh against a tomographic file
+mesh.group_timestep() # grouping hex with similar timestep
+mesh.group_period() # grouping hex with similar period
+\end{lstlisting}
 
@@
 
-\section{TODO SPECFEM3D MESH FORMAT}
+\section{TODO: \specfem\ Mesh Format}
 
@@
 
-\section{TODO NOTES ABOUT THE DIMENSION OF THE HEXES, THE MINIMUM PERIOD RESOLVED AND THE TIME STEP OF THE SIMULATION}
+\section{TODO: Notes about the Dimension of the Hexes, the Minimum Period Resolved and the Time Step of the Simulation}
 

### 2013-09-26 CUBIT_GEOCUBIT/geocubitlib/save_fault_nodes_elements.py
modified CUBIT_GEOCUBIT/ to match clone from github on 25-Sept-2013

@@ -1,84 +1,0 @@
-#!python
-# Opening fault cracks
-# Input : surface up and down.
-import cubit 
-
-class fault_input:
-   def __init__(self,id,surface_u,surface_d): 
-       self.id = id
-       self.surface_u = surface_u
-       self.surface_d = surface_d
-       self.name = 'MESH/fault_file_'+str(id)+'.dat'
-
-       quads_Aup,quads_Adp = save_cracks(self.name,self.surface_u,self.surface_d)
-       #Unpacking list.
-       quads_Au=unpack_list(quads_Aup)
-       quads_Ad=unpack_list(quads_Adp)
-
-       print 'len(Au):',len(quads_Au)
-       print 'len(Ad):',len(quads_Ad)
-
-       if not (len(quads_Au)==len(quads_Ad)):
-           print 'Number of elements for each fauld side up and down do not concide'
-           sys.exit('goodbye')
-           
-       save_elements_nodes(self.name,quads_Au,quads_Ad)
-
-
-def save_cracks(name,list_surface_up,list_surface_down):
-   quads_fault_up = [] 
-   quads_fault_down = []
-   for surface in list_surface_up   :
-       quads_fault = cubit.get_surface_quads(surface)
-       quads_fault_up.append(quads_fault)
-   for surface in list_surface_down :
-       quads_fault = cubit.get_surface_quads(surface)
-       quads_fault_down.append(quads_fault)
-    # TO DO : stop python properly in case fault nodes at both sides
-    #         do not match.
-    #   if len(quads_fault_u) != len(quads_fault_d): stop
-    #
-    # SAVING FAULT ELEMENTS AND NODES
-   return quads_fault_up,quads_fault_down
-
-def unpack_list(fault_list):
-    list_fault = []
-    for i in range(0,len(fault_list)):
-        el=list(fault_list[i])
-        for j in el:
-            list_fault.append(j)
-    return list_fault 
-          
-def save_elements_nodes(name,quads_fault_u,quads_fault_d):
-   fault_file = open(name,'w')
-   txt =''
-   list_hex=cubit.parse_cubit_list('hex','all')
-   txt='%10i %10i\n' % (len(quads_fault_u),len(quads_fault_d))
-   fault_file.write(txt)
-   
-   dic_quads_fault_u = dict(zip(quads_fault_u,quads_fault_u)) 
-   dic_quads_fault_d = dict(zip(quads_fault_d,quads_fault_d)) 
-   
[…]

### 2013-09-26 doc/USER_MANUAL/manual_SPECFEM3D_Cartesian.tex
removed the sentence about -mcmodel from the manual

@@
 in order to turn underflow trapping off.\\
-
-If you run very large meshes on a relatively small number of processors,
-the memory size needed on each processor might become greater than
-2 gigabytes, which is the upper limit for 32-bit addressing. In this
-case, on some compilers you may need to add \texttt{``-mcmodel=medium}''
-or \texttt{``-mcmodel=medium -shared-intel}'' to the configure options
-of CFLAGS, FCFLAGS and LDFLAGS otherwise the compiler will display
-an error message (for instance \texttt{``relocation truncated to
-fit: R\_X86\_64\_PC32 against .bss''} or something similar); on an
-IBM machine with the \texttt{xlf} and \texttt{xlc} compilers, using
-\texttt{-q64} is usually sufficient.\\
 

### 2014-01-24 doc/USER_MANUAL/manual_SPECFEM3D_Cartesian.tex

@@
+% !TEX encoding = IsoLatin
 %% LyX 2.0.2 created this file.  For more info, see http://www.lyx.org/.
@@
 Percy Galvez, Dominik Göddeke, Vala Hjörleifsdóttir, Sue Kientz, Dimitri
-Komatitsch, Jesús Labarta, Nicolas Le Goff, Pieyre Le Loher, Qinya
-Liu, Yang Luo, Alessia Maggi, Federica Magnoni, Roland Martin, René
-Matzen, Dennis McRitchie, Matthias Meschede, Peter Messmer, David
-Michéa, Surendra Nadh Somala, Tarje Nissen-Meyer, Daniel Peter, Max
+Komatitsch, Jesús Labarta, Nicolas Le Goff, Pieyre Le Loher, Matthieu
+Lefebvre, Qinya Liu, Yang Luo, Alessia Maggi, Federica Magnoni, Roland
+Martin, René Matzen, Dennis McRitchie, Matthias Meschede, Peter Messmer, 
+David Michéa, Surendra Nadh Somala, Tarje Nissen-Meyer, Daniel Peter, Max
 Rietmann, Elliott Sales de Andrade, Brian Savage, Bernhard Schuberth,
@@
 a GUI front-end for configuring and running \texttt{doxygen}.
+
+\section{Using the ADIOS library for I/O}
+Regular POSIX I/O can be problematic when dealing with large simulations one large
+clusters (typically more than $10,000$ processes). SPECFEM3D use the ADIOS library~\cite{Liu2013}
+to deal transparently take advantage of advanced paralllel file system features. To enable
+ADIOS, the following steps should be done:
+\begin{enumerate}
+\item Install ADIOS (available from \url{https://www.olcf.ornl.gov/center-projects/adios/}). 
+Make sure that your environment variables reference it. 
+\item You may want to change ADIOS related values in the \texttt{constants.h.in} file.
+The default values probably suit most cases.
+\item Configure using the \texttt{--with-adios} flag.
+\end{enumerate}
+ADIOS is currently only usable for meshfem3D generated mesh (i.e. not for meshes generated
+with CUBTI). Additional control parameters are discussed in section~\ref{cha:Main-Parameter}.
 
@@
 boundary conditions for a regional simulation. Both fluids and elastic
-solids are supported.
+solids are supported. Note that xmeshfem3d generated meshes do not support C-PML yet.
 \item [{\texttt{PML\_INSTEAD\_OF\_FREE\_SURFACE}}] Set to \texttt{.true.}
@@
 \item [{\texttt{GPU\_MODE}}] Turn this flag on to use GPUs.
+
+\item [\texttt{ADIOS\_ENABLED}] Turn this flag on to enable ADIOS. If set to \texttt{.false.}, subsequent ADIOS 
+parameters will not be considered.
+\item [\texttt{ADIOS\_FOR\_DATABASES}] Turn this flag on to use ADIOS for xmeshfem3D output and 
+xgenerate\_database input.
+\item [\texttt{ADIOS\_FOR\_MESH}]  Turn this flag on to use ADIOS for generated databases.
+\item [\texttt{ADIOS\_FOR\_FORWARD\_ARRAYS}] Turn this flag on to read and write forward arrays using ADIOS.
+\item [\texttt{ADIOS\_FOR\_KERNELS}] Turn this flag on to produce ADIOS kernels that can later be visualized with the ADIOS version of combine\_vol\_data.
+
 \end{description}

### 2014-02-08 doc/USER_MANUAL/manual_SPECFEM3D_Cartesian.tex
fixed overflow of degrees-of-freedom counters when more that 2 gigapoints; added comment about Zlib -lz to the users manual

@@
 \end{lyxcode}
+
+When compiling the SCOTCH source code, if you get a message such as: "ld: cannot find -lz",
+the Zlib compression development library is probably missing on your machine and you will need to install it or ask your system administrator to 
+do so. On Linux machines the package is often called "zlib1g-dev" or similar. (thus "sudo apt-get install zlib1g-dev" would install it)
+
 To compile a serial version of the code for small meshes that fits

### 2014-04-28 doc/USER_MANUAL/manual_SPECFEM3D_Cartesian.tex
updated all the copyright header files

@@
 \includegraphics[width=0.130\textwidth]{figures/logo_Fairbanks}
+\hspace*{3mm}\includegraphics[width=0.130\textwidth]{figures/logo_fondation_Del_Duca}
 \par\end{centering}
@@
 suited to parallel implementation on very large supercomputers \citep{KoTsChTr03,TsKoChTr03,KoLaMi08a,CaKoLaTiMiLeSnTr08,KoViCh10}
-as well as on clusters of GPU accelerating graphics cards \citep{KoMiEr09,KoErGoMi10,Kom11}.
+as well as on clusters of GPU accelerating graphics cards \citep{Kom11,MiKo10,KoMiEr09,KoErGoMi10}.
 Tensor products inside each element can be optimized to reach very
@@
 
+It reached the sustained one petaflop performance level for the first time in February 2013
+on the Blue Waters Cray supercomputer at the National Center for Supercomputing Applications (NCSA), located at the University of Illinois at Urbana-Champaign (USA).\\
+
 This new release of the code (version 2.1) includes Convolution or
@@
 (C-PML or ADE-PML) \citep{MaKoEz08,MaKoGe08,MaKo09,MaKoGeBr10,KoMa07}.
-It also includes support for GPU graphics card acceleration \citep{KoMiEr09,KoErGoMi10,MiKo10,Kom11}.\\
+It also includes support for GPU graphics card acceleration \citep{Kom11,MiKo10,KoMiEr09,KoErGoMi10}.\\
 
@@
  Computing on GPU graphics cards for acoustic or seismic wave propagation
-applications is described in detail in \citet{KoMiEr09,KoErGoMi10,MiKo10,Kom11}.
+applications is described in detail in \citet{Kom11,MiKo10,KoMiEr09,KoErGoMi10}.
 \end{description}
@@
 
-The SPECFEM3D Cartesian software package comes in a gzipped tar ball.
-In the directory in which you want to install the package, type
-\begin{lyxcode}
-{\small tar~-zxvf~SPECFEM3D\_V2.1.0\_Cartesian.tar.gz}{\small \par}
-\end{lyxcode}
-The directory \texttt{\small SPECFEM3D/}{\small{} will then contain
-the source code.}\\
+To download the SPECFEM3D\_Cartesian software package, type this:
+\begin{lyxcode}
+git clone --recursive --branch devel https://github.com/geodynamics/specfem3d.git
+\end{lyxcode}
 {\small \par}
@@
 To report bugs or suggest improvements to the code, please send an
-e-mail to the CIG Computational Seismology Mailing List \urlwithparentheses{cig-seismo@geodynamics.org}
-or Jeroen Tromp \urlwithparentheses{jtromp-AT-princeton.edu}, and/or
-use our online bug tracking system Roundup \urlwithparentheses{www.geodynamics.org/roundup}.
+e-mail to the CIG Computational Seismology Mailing List \urlwithparentheses{cig-seismo@geodynamics.org}.
 
@@
 Please e-mail your feedback, questions, comments, and suggestions
-to Jeroen Tromp \urlwithparentheses{jtromp-AT-princeton.edu} or
 to the CIG Computational Seismology Mailing List \urlwithparentheses{cig-seismo@geodynamics.org}.
@@
 
-Princeton University, USA, and CNRS / INRIA / University of Pau, France
-
-$\copyright$ Princeton University and CNRS / INRIA / University of
-Pau, July 2012
+Princeton University, USA, and CNRS / University of Marseille, France
+
+$\copyright$ Princeton University and CNRS / University of Marseille, July 2012
 

### 2014-05-01 doc/USER_MANUAL/manual_SPECFEM3D_Cartesian.tex

@@
 in order to turn underflow trapping off.\\
+
+You can add -DFORCE\_VECTORIZATION to the compiler options in \texttt{flags.guess} (for all compilers, except for IBM for which the syntax is -WF,-DFORCE\_VECTORIZATION )
+to speed up the code in the fluid (acoustic) parts (only; FORCE\_VECTORIZATION support for elastic parts has been discontinued in the source code).
+This works fine if (and only if) your computer always allocates a contiguous memory block for each allocatable array;
+this is the case for most machines and most compilers, but not all. For more details see https://github.com/geodynamics/specfem3d/issues/81 .
+To check if that option works fine on your machine, run the code with and without it for a model containing a significant fluid layer
+(or entirely fluid) and make sure the seismograms are identical.\\
 

### 2014-06-02 doc/USER_MANUAL/manual_SPECFEM3D_Cartesian.tex
added section "multiple external tomography files for SPECFEM3D" from Carl Tape

@@
 
+%==============
+
+\subsubsection*{Example: External tomography file with variable discretization intervals}
+
+The example in Figure~\ref{fig:tomography_file} is for an external tomography file with uniform spacing in the $x$, $y$, and $z$ directions.
+(Note that $\Delta x$, $\Delta y$, and $\Delta z$ need not be the same as in the example.) In that example, the  \texttt{nummaterial\_velocity\_file} is
+%
+\begin{lyxcode}
+2  -1 tomography elastic tomography\_model.xyz 1
+\end{lyxcode}
+%
+and the file \texttt{tomography\_model.xyz} will need to reside in \texttt{DATA/}.
+All entries in the second column of \texttt{materials\_file} will be \texttt{-1}, which means that each element in the mesh will be interpolated accordind to the values in \texttt{tomography\_model.xyz}.
+
+In some cases it may be desirable to have an external tomography model that is described in more than one file.
+For example, in cases like southern California, the length scale of variation in the structure of the wave speed model is much shorter in the sedimentary basin models within the upper 15~km.
+Therefore one might want to use an external tomography file that is sampled with, say, $\Delta x = 1000$~m, $\Delta y = 1000$~m, and $\Delta z = 250$~m in the uppermost 15~km, and then use $\Delta x = 2000$~m, $\Delta y = 2000$~m, and $\Delta z = 1000$~m below a depth of 15~km.
+If these intervals are chosen appropriately, then it will result in a pair of external tomography file that is much smaller than the alternative of having a single file with the fine discretization.
+In this case  \texttt{nummaterial\_velocity\_file} is
+%
+\begin{lyxcode}
+2  -1 tomography elastic file\_above\_15km.xyz 1 \\
+2  -2 tomography elastic file\_below\_15km.xyz 1
+\end{lyxcode}
+%
+and the files \texttt{file\_above\_15km.xyz} and \texttt{file\_below\_15km.xyz} will need to reside in \texttt{DATA/}.
+All entries in the second column of \texttt{materials\_file} will need to be either \texttt{-1} or \texttt{-2}, depending on whether the element is above or below 15~km depth.
+In this sense, the construction of the mesh and the discretization of the external model will generally be done in tandem.
+
+Other more complicated discretizations can be used following the same procedure. (In particular, there is no limit on the number of different external files that are used.)
+
+%=============
+
 \section{{\normalsize \label{sec:Anelastic-Models}}External (an)elastic Models}

### 2014-06-02 doc/USER_MANUAL/manual_SPECFEM3D_Cartesian.tex

@@
 
-Packages such as \texttt{doxywizard} can be used to visualize the
-subroutine calling tree of the source code. \texttt{Doxywizard} is
-a GUI front-end for configuring and running \texttt{doxygen}.
+\noindent
+Packages such as \texttt{Doxywizard} can be used to visualize the calling tree of the
+subroutines of the  source code. \texttt{Doxywizard} is
+a GUI front-end for configuring and running \texttt{Doxygen}.
+
+\bigskip
+
+\noindent To do your own call graphs, you can follow these simple steps below.
+
+\begin{enumerate}
+\setcounter{enumi}{-1}
+\item Install \texttt{Doxygen} \underline{and} \texttt{graphviz} (the two are usually in the package manager of classic Linux distribution).
+\item Run in the terminal : \texttt{doxygen -g}, which creates a \texttt{Doxyfile} that tells doxygen what you want it to do.
+\item Edit the Doxyfile. Two Doxyfile-type files have been already committed in the directory \\
+\noindent \texttt{specfem3d/doc/Call\_trees}:
+
+\begin{itemize}
+\item[\textbullet] \texttt{Doxyfile\_truncated\_call\_tree} will generate call graphs with maximum 3 or 4 levels of tree structure,
+\item[\textbullet] \texttt{Doxyfile\_complete\_call\_tree} will generate call graphs with complete tree structure.
+\end{itemize}
+
+The important entries in the Doxyfile are:
+
+\begin{description}
+\item [{\texttt{PROJECT\_NAME}}] 
+\item [{\texttt{OPTIMIZE\_FOR\_FORTRAN}}] Set to YES
+\item [{\texttt{EXTRACT\_ALL}}] Set to YES
+\item [{\texttt{EXTRACT\_PRIVATE}}] Set to YES
+\item [{\texttt{EXTRACT\_STATIC}}] Set to YES
+\item [{\texttt{INPUT}}] From the directory \texttt{specfem3d/doc/Call\_trees}, it is \texttt{"../../src/"}
+\item [{\texttt{FILE\_PATTERNS}}] In SPECFEM case, it is \texttt{*.f90* *.F90* *.c* *.cu* *.h*}
+\item [{\texttt{HAVE\_DOT}}] Set to YES
+\item [{\texttt{CALL\_GRAPH}}] Set to YES
+\item [{\texttt{CALLER\_GRAPH}}] Set to YES
+\item [{\texttt{DOT\_PATH}}] The path where is located the dot program graphviz (if it is not in your \$PATH)
+\item [{\texttt{RECURSIVE}}] This tag can be used to turn specify whether or not subdirectories should be searched for input files as well. In the case of SPECFEM, set to YES.
+\item [{\texttt{EXCLUDE}}] Here, you can exclude:  \\
+\texttt{"../../src/specfem3D/older\_not\_maintained\_partial\_OpenMP\_port"}\\
+\texttt{"../../src/decompose\_mesh/scotch"}\\
+\texttt{"../../src/decompose\_mesh/scotch\_5.1.12b"}
+\item [{\texttt{DOT\_GRAPH\_MAX\_NODES}}] to set the maximum number of nodes that will be shown in the graph.
+If the number of nodes in a graph becomes larger than this value, doxygen will truncate the graph, which is visualized by representing a node as a red box.
+Minimum value: 0, maximum value: 10000, default value: 50.
+\item [{\texttt{MAX\_DOT\_GRAPH\_DEPTH}}] to set the maximum depth of the graphs generated by dot.
+A depth value of 3 means that only nodes reachable from the root by following a path via at most 3 edges will be shown.
+Using a depth of 0 means no depth restriction.
+Minimum value: 0, maximum value: 1000, default value: 0.
+
+\end{description}
+
+\item Run : \texttt{doxygen Doxyfile}, HTML and LaTeX files created by default in \texttt{html} and \texttt{latex} subdirectories.
+
+\item To see the call trees, you have to open the file \texttt{html/index.html} in your \underline{browser}.
+You will have many informations about each subroutines of SPECFEM (not only call graphs), you can click on every boxes / subroutines.
+It show you the call, and, the caller graph of each subroutine : the subroutines called by the concerned subroutine, and the previous subroutines who call this subroutine (the previous path), respectively.
+In the case of a truncated calling tree, the boxes with a red border indicates a node that has \underline{more} arrows than are shown (in other words: the graph is truncated with respect to this node).
+
+\end{enumerate}
+
+\medskip
+
+\noindent Finally, some useful links:
+
+\begin{itemize}
+\item[\textbullet] a good and short summary for the basic utilisation of Doxygen:
+
+\texttt{http://www.softeng.rl.ac.uk/blog/2010/jan/30/callgraph-fortran-doxygen/} ,
+
+\item[\textbullet] to configure the diagrams :
+
+\texttt{http://www.stack.nl/~dimitri/doxygen/manual/diagrams.html} ,
+
+\item[\textbullet] the complete alphabetical index of the tags in Doxyfile:
+
+\texttt{http://www.stack.nl/~dimitri/doxygen/manual/config.html} ,
+
+\item[\textbullet] more generally, the Doxygen manual:
+
+\texttt{http://www.stack.nl/~dimitri/doxygen/manual/index.html} .
+
+\end{itemize}
+
+\medskip
 

### 2014-06-06 doc/USER_MANUAL/manual_SPECFEM3D_Cartesian.tex
added instructions for Python paths and GeoCUBIT to the manual

@@
 
-In most cases, to re-create the meshes for the examples given, just type "claro ./block\_mesh.py" or similar 
+In some cases, to re-create the meshes for the examples given, just type "claro ./create\_mesh.py" or similar 
 from the command line ("claro" is the command to run CUBIT from the command line).
+
+IMPORTANT note: in order to correctly set up GEOCUBIT and run the examples, please read the file called "examples/README";
+in particular, please make sure you correctly set up the Python paths as indicated in that file.
 

### 2014-06-12 doc/USER_MANUAL/manual_SPECFEM3D_Cartesian.tex
added a link to the developers' wiki

@@
 with CUBTI). Additional control parameters are discussed in section~\ref{cha:Main-Parameter}.
+
+\section{Becoming a developer of the code, or making small modifications in the source code}
+
+If you want to develop new features in the code, and/or if you want to make small changes, improvements, or bug fixes, you are very welcome to contribute.
+To do so, i.e.
+to access the development branch of the source code with read/write access (in a safe way, no need to worry too much about breaking the package, there is a robot called BuildBot that is in charge of checking and validating all new contributions and changes), please visit this Web page: \url{https://github.com/geodynamics/specfem3d/wiki/Using-Hub}.
 

### 2014-06-19 doc/USER_MANUAL/manual_SPECFEM3D_Cartesian.tex

@@
 Examples are provided in the SPECFEM3D Cartesian package in the subdirectory
-\texttt{examples/}. We strongly encourage you to contribute your own
+\texttt{EXAMPLES/}. We strongly encourage you to contribute your own
 example to this package by contacting the CIG Computational Seismology
@@
 examples are provided in the SPECFEM3D Cartesian package in the subdirectory
-\texttt{examples/}:
+\texttt{EXAMPLES/}:
 \begin{description}
@@
 
-\caption{Screenshots of the CUBIT examples provided in subdirectory \texttt{examples/}:
+\caption{Screenshots of the CUBIT examples provided in subdirectory \texttt{EXAMPLES/}:
 homogeneous halfspace (top-left), layered halfspace (top-right), water
@@
 
-IMPORTANT note: in order to correctly set up GEOCUBIT and run the examples, please read the file called "examples/README";
+IMPORTANT note: in order to correctly set up GEOCUBIT and run the examples, please read the file called "EXAMPLES/README";
 in particular, please make sure you correctly set up the Python paths as indicated in that file.
@@
 elements in the vertical direction for each layer. We provide a few
-models in the {\texttt{examples/}} directory. \\
+models in the {\texttt{EXAMPLES/}} directory. \\
 
@@
 
-The CUBIT scripts ({*}.jou and {*}.py) in the directory $\mathtt{examples}$
+The CUBIT scripts ({*}.jou and {*}.py) in the directory $\mathtt{EXAMPLES}$
 generate more complicated meshes. The {*}.py files are Python scripts
@@
 \end{itemize}
-Read the documents in the directory $\mathtt{examples/*/description}$.
+Read the documents in the directory $\mathtt{EXAMPLES/*/description}$.
 They contain a description of the example and additional instructions
 to run it. Visualize the results with the Matlab scripts in the directory
-$\mathtt{examples/*/post}$
+$\mathtt{EXAMPLES/*/post}$
 
@@
 \begin{lyxlist}{00.00.0000}
-\item [{$\mathtt{examples/tpv16/description/TPV16\_17\_Description\_v03.pdf}$}]~
+\item [{$\mathtt{EXAMPLES/tpv16/description/TPV16\_17\_Description\_v03.pdf}$}]~
 \end{lyxlist}
@@
 
-The directories $\mathtt{examples/*/post}$ contain additional Matlab
+The directories $\mathtt{EXAMPLES/*/post}$ contain additional Matlab
 scripts to generate figures specific to each example.
@@
 
-\texttt{examples/noise\_tomography/NOISE\_TOMOGRAPHY.m} (main program)\\
- \texttt{examples/noise\_tomography/PetersonNoiseModel.m}
+\texttt{EXAMPLES/noise\_tomography/NOISE\_TOMOGRAPHY.m} (main program)\\
+ \texttt{EXAMPLES/noise\_tomography/PetersonNoiseModel.m}
 
@@
 In order to illustrate noise simulations in an easy way, one example
-is provided in \texttt{examples/noise\_tomography/}. See \texttt{examples/noise\_tomography/README}
+is provided in \texttt{EXAMPLES/noise\_tomography/}. See \texttt{EXAMPLES/noise\_tomography/README}
 for explanations. \\
@@
 
-If your workstation is suitable, you can run the example in \texttt{examples/noise\_tomography/}
+If your workstation is suitable, you can run the example in \texttt{EXAMPLES/noise\_tomography/}
 using:\\
@@
 at the skewness table printed in the \texttt{output\_mesher.txt} file
-after executing \texttt{xmeshfem3D} for the example given in \texttt{examples/meshfem3D\_examples/simple\_model/}:
+after executing \texttt{xmeshfem3D} for the example given in \texttt{EXAMPLES/meshfem3D\_examples/simple\_model/}:
 
@@
 following figure \ref{fig:mesh.vp} of the mesh you obtain in the
-example \texttt{examples/meshfem3D\_examples/simple\_model/}.
+example \texttt{EXAMPLES/meshfem3D\_examples/simple\_model/}.
 \begin{figure}[htbp]
@@
 \caption{Paraview visualization using the mesh vtk-files for the example given
-in \texttt{examples/meshfem3D\_examples/simple\_model/}.}
+in \texttt{EXAMPLES/meshfem3D\_examples/simple\_model/}.}
 

### 2014-07-01 doc/USER_MANUAL/manual_SPECFEM3D_Cartesian.tex
added Can Kangchen Bai to the list of developers

@@
 
-\newcommand{\toall}[1]{\textbf{*** All: #1 ***}}
-\newcommand{\tojeroen}[1]{\textbf{*** Jeroen: #1 ***}}
-\newcommand{\tobrian}[1]{\textbf{*** Brian: #1 ***}}
-\newcommand{\tovala}[1]{\textbf{*** Vala: #1 ***}}
-\newcommand{\tovalabrian}[1]{\textbf{*** Vala \& Brian: #1 ***}}
-\newcommand{\tovalaqinya}[1]{\textbf{*** Vala \& Qinya: #1 ***}}
-\newcommand{\toqinya}[1]{\textbf{*** Qinya: #1 ***}}
-\newcommand{\tomin}[1]{\textbf{*** Min: #1 ***}}
-\newcommand{\toalessia}[1]{\textbf{*** Alessia: #1 ***}}
-\newcommand{\todimitri}[1]{\textbf{*** Dimitri: #1 ***}}
-
 \newcommand{\nexxi}{\mbox{\texttt{NEX\_XI\/}}}
@@
 
+%%%%%%%% please do NOT merge the lines of the paragraph below, it makes it difficult to insert new names in alphabetical order; leave a single name per line
+%%%%%%%% please do NOT merge the lines of the paragraph below, it makes it difficult to insert new names in alphabetical order; leave a single name per line
+%%%%%%%% please do NOT merge the lines of the paragraph below, it makes it difficult to insert new names in alphabetical order; leave a single name per line
 Since then it has been developed and maintained by a development team:
-in alphabetical order, Jean-Paul (Pablo) Ampuero, Piero Basini, Céline
-Blitz, Ebru Bozda\u{g}, Emanuele Casarotti, Joseph Charles, Min Chen,
-Percy Galvez, Dominik Göddeke, Vala Hjörleifsdóttir, Sue Kientz, Dimitri
-Komatitsch, Jesús Labarta, Nicolas Le Goff, Pieyre Le Loher, Matthieu
-Lefebvre, Qinya Liu, Yang Luo, Alessia Maggi, Federica Magnoni, Roland
-Martin, René Matzen, Dennis McRitchie, Matthias Meschede, Peter Messmer, 
-David Michéa, Surendra Nadh Somala, Tarje Nissen-Meyer, Daniel Peter, Max
-Rietmann, Elliott Sales de Andrade, Brian Savage, Bernhard Schuberth,
-Anne Sieminski, Leif Strand, Carl Tape, Jeroen Tromp, Jean-Pierre
-Vilotte, Zhinan Xie, Hejun Zhu.\\
+in alphabetical order,
+Jean-Paul (Pablo) Ampuero,
+Can Kangchen Bai,
+Piero Basini,
+Céline Blitz,
+Ebru Bozda\u{g},
+Emanuele Casarotti,
+Joseph Charles,
+Min Chen,
+Percy Galvez,
+Dominik Göddeke,
+Vala Hjörleifsdóttir,
+Sue Kientz,
+Dimitri Komatitsch,
+Jesús Labarta,
+Nicolas Le Goff,
+Pieyre Le Loher,
+Matthieu Lefebvre,
+Qinya Liu,
+Yang Luo,
+Alessia Maggi,
+Federica Magnoni,
+Roland Martin,
+René Matzen,
+Dennis McRitchie,
+Matthias Meschede,
+Peter Messmer, 
+David Michéa,
+Surendra Nadh Somala,
+Tarje Nissen-Meyer,
+Daniel Peter,
+Max Rietmann,
+Elliott Sales de Andrade,
+Brian Savage,
+Bernhard Schuberth,
+Anne Sieminski,
+Leif Strand,
+Carl Tape,
+Jeroen Tromp,
+Jean-Pierre Vilotte,
+Zhinan Xie,
+Hejun Zhu.\\
 

### 2014-07-01 doc/USER_MANUAL/manual_SPECFEM3D_Cartesian.tex

@@
 Jean-Paul (Pablo) Ampuero,
-Can Kangchen Bai,
+Kangchen Bai,
 Piero Basini,

### 2014-07-06 doc/USER_MANUAL/manual_SPECFEM3D_Cartesian.tex
acknowledge Geza Seriani and Enrico Priolo in the users manual

@@
 (IPGP) in Paris, France from 1995 to 1997 and then by Dimitri Komatitsch
-and Jeroen Tromp at Harvard University and Caltech, USA, starting
-in 1998. The story started on March 28, 1995, when Prof. Yvon Maday
-from CNRS and University of Paris, France, gave a lecture to Dimitri
-Komatitsch and Jean-Pierre Vilotte at IPG about the nice properties
-of the spectral-element method that he had used for other equations.
-We are deeply indebted and thankful to him for that.
+and Jeroen Tromp at Harvard University and Caltech, USA, starting in 1998.
+The story started on March 28, 1995, when Prof. Yvon Maday from CNRS and University of Paris, France, gave a lecture to
+Dimitri Komatitsch and Jean-Pierre Vilotte at IPG about the nice properties of the Legendre spectral-element method with diagonal mass matrix that he had used for
+other equations. We are deeply indebted and thankful to him for that.
+That followed a visit by Dimitri Komatitsch to OGS (Istituto Nazionale di Oceanografia e di Geofisica Sperimentale) in Trieste, Italy, in February 1995
+to meet with G\'eza Seriani and Enrico Priolo, who introduced him to their 2D Chebyshev version of the spectral-element method with a non-diagonal mass matrix.
+We are deeply indebted and thankful to them for that.
 

### 2014-07-17 doc/USER_MANUAL/manual_SPECFEM3D_Cartesian.tex
added the framework for the future option NUMBER_OF_SIMULTANEOUS_RUNS that I will add to setup/constants.h.in (off by default, i.e. equal to 1 by default) to have the ability to run several independent earthquakes in parallel from the same big job

@@
 
+We have also added the ability to run several calculations (several earthquakes)
+in an embarrassingly-parallel fashion from within the same run;
+this can be useful when using a very large supercomputer to compute
+many earthquakes in a catalog, in which case it can be better from
+a batch job submission point of few to start fewer and much larger jobs,
+each of them computing several earthquakes in parallel.
+To turn that option on, set parameter NUMBER\_OF\_SIMULTANEOUS\_RUNS
+to a value greater than 1 in file setup/constants.h.in before
+configuring and compiling the code.
+When that option is on, of course the number of processor cores used to start
+the code in the batch system must be a multiple of NUMBER\_OF\_SIMULTANEOUS\_RUNS,
+all the individual runs must use the same number of processor cores
+(which as usual is NPROC in the input file DATA/Par\_file,
+and thus the total number of processor cores to request from the batch system
+should be NUMBER\_OF\_SIMULTANEOUS\_RUNS * NPROC.
+All the runs to perform must be placed in directories called run0001, run0002, run0003 and so on (with exactly four digits).
+
 \chapter{\label{cha:Fault-Kinematics-Dynamics}Kinematic and dynamic fault
@@
 specified in the main }\texttt{\small Par\_file}{\small , e.g. in
-directory }\texttt{\small OUTPUT\_FILES/DATABASES\_MPI}{\small . %As the files can be very large, there are several flags in the \texttt{\small Par\_file}
-%that control the region in space and time that is saved. These are:
-%\texttt{\small NTSTEP\_BETWEEN\_FRAMES}
+directory }\texttt{\small OUTPUT\_FILES/DATABASES\_MPI}{\small .
 The output is saved by each processor at the time interval specified
@@
 
-We describe here in some detail a job submission procedure for the
-Caltech 1024-node cluster, CITerra, under the LSF scheduling system.
-We consider the submission of a regular forward simulation using the
-internal mesher to create mesh partitions. The two main scripts are
-\texttt{\small run\_lsf.bash}{\small , which compiles the Fortran
-code and submits the job to the scheduler, and }\texttt{\small go\_mesher\_solver\_lsf\_basin.forward}{\small ,
-which contains the instructions that make up the job itself. These
-scripts can be found in }\texttt{\small utils/Cluster/lsf/}{\small{}
-directory}{\small \par}
-
-\section{Job submission \texttt{run\_lsf.bash}}
-
-This script first sets the job queue to be `normal'. It then compiles
-the mesher, database generator and solver together, figures out the
-number of processors required for this simulation from the \texttt{DATA/Par\_file},
-and submits the LSF job.
-\begin{lyxcode}
-\#!/bin/bash
-
-\#~use~the~normal~queue~unless~otherwise~directed~queue=\textquotedbl{}-q~normal\textquotedbl{}~
-
-if~{[}~\$\#~-eq~1~{]};~then
-
-~~~~~~~~echo~\textquotedbl{}Setting~the~queue~to~\$1\textquotedbl{}
-
-~~~~~~~~queue=\textquotedbl{}-q~\$1\textquotedbl{}~
-
-fi~~\\
-~~~\\
-~\#~compile~the~mesher~and~the~solver~
-
-d=`date'~echo~\textquotedbl{}Starting~compilation~\$d\textquotedbl{}~
-
-make~clean~
-
-make~xmeshfem3D~
-
-make~xgenerate\_databases~
-
-make~xspecfem3D~
-
-d=`date'~
-
-echo~\textquotedbl{}Finished~compilation~\$d\textquotedbl{}~~\\
-~~~\\
-~\#~get~total~number~of~nodes~needed~for~solver~
-
-NPROC=`grep~NPROC~DATA/Par\_file~|~cut~-c~34-~'~~\\
-~~~\\
-
-\#~compute~total~number~of~nodes~needed~for~mesher~
-
-NPROC\_XI=`grep~NPROC\_XI~DATA/meshfem3D\_files/Mesh\_Par\_file~|~cut~-c~34-~'~~~~\\
-
-NPROC\_ETA=`grep~NPROC\_ETA~DATA/meshfem3D\_files/Mesh\_Par\_file~|~cut~-c~34-~'~~~~\\
-~\#~total~number~of~nodes~is~the~product~of~the~values~read~
-
-numnodes=\$((~\$NPROC\_XI~{*}~\$NPROC\_ETA~))~~\\
-~~~\\
-
-\#~checks~total~number~of~nodes~
-
-if~{[}~\$numnodes~-neq~\$NPROC~{]};~then
-
-~~~~~~~~echo~\textquotedbl{}error~number~of~procs~mismatch\textquotedbl{}
-
-~~~~~~~~exit~
-
-fi~~\\
-
-~~\\
-~echo~\textquotedbl{}Submitting~job\textquotedbl{}~
-
-bsub~\$queue~-n~\$numnodes~-W~60~-K~<go\_mesher\_solver\_lsf.forward~
-\end{lyxcode}
-
-\section{Job script \texttt{go\_mesher\_solver\_lsf.forward}}
-
-This script describes the job itself, including setup steps that can
-only be done once the scheduler has assigned a job-ID and a set of
-compute nodes to the job, the \texttt{run\_lsf.bash} commands used
-to run the mesher, database generator and the solver, and calls to
-scripts that collect the output seismograms from the compute nodes
-and perform clean-up operations.
-\begin{enumerate}
-\item First the script directs the scheduler to save its own output and
-output from \texttt{stdout} into ~\\
- \texttt{\small OUTPUT\_FILES/\%J.o}{\small , where }\texttt{\small \%J}{\small{}
-is short-hand for the job-ID; it also tells the scheduler what version
-of }\texttt{\small mpich}{\small{} to use (}\texttt{\small mpich\_gm}{\small )
-and how to name this job (}\texttt{\small go\_mesher\_solver\_lsf}{\small ). }{\small \par}
-\item The script then creates a list of the nodes allocated to this job
-by echoing the value of a dynamically set environment variable \texttt{LSB\_MCPU\_HOSTS}
-and parsing the output into a one-column list using the Perl script
-\texttt{utils/Cluster/lsf/remap\_lsf\_machines.pl}. It then creates
-a set of scratch directories on these nodes (\texttt{\small /scratch/}{\small ~}\\
-{\small{} }\texttt{\small \$USER/DATABASES\_MPI}{\small ) to be used
-as the }\texttt{\small LOCAL\_PATH}{\small{} for temporary storage
-of the database files. The scratch directories are created using }\texttt{\small shmux}{\small ,
-a shell multiplexor that can execute the same commands on many hosts
-in parallel. }\texttt{\small shmux}{\small{} is available from Shmux
-\urlwithparentheses{web.taranis.org/shmux/}. Make sure that the
-}\texttt{\small LOCAL\_PATH}{\small{} parameter in }\texttt{\small DATA/Par\_file}{\small{}
-is also set properly. }{\small \par}
-\item The next portion of the script launches the mesher, database generator
-and then the solver using \texttt{run\_lsf.bash}.
-\item The final portion of the script collects the seismograms and performs
-clean up on the nodes, using the Perl scripts \texttt{collect\_seismo\_lsf\_multi.pl}
-and \texttt{cleanmulti.pl}. \end{enumerate}
-\begin{lyxcode}
-\#!/bin/bash~-v~
-
-\#BSUB~-o~OUTPUT\_FILES/\%J.o~
-
-\#BSUB~-a~mpich\_gm~
-
-\#BSUB~-J~go\_mesher\_solver\_lsf~~\\
-~~~\\
-~\#~set~up~local~scratch~directories
-
-BASEMPIDIR=/scratch/\$USER/DATABASES\_MPI
-
-mkdir~-p~OUTPUT\_FILES~
-
-echo~\textquotedbl{}\$LSB\_MCPU\_HOSTS\textquotedbl{}~>~OUTPUT\_FILES/lsf\_machines~
-
-echo~\textquotedbl{}\$LSB\_JOBID\textquotedbl{}~>~OUTPUT\_FILES/jobid
-
-remap\_lsf\_machines.pl~OUTPUT\_FILES/lsf\_machines~>~OUTPUT\_FILES/machines
-
-shmux~-M50~-Sall~-c~\textquotedbl{}rm~-r~-f~/scratch/\$USER;~\textbackslash{}
-
-~~~~~~mkdir~-p~/scratch/\$USER;~mkdir~-p~\$BASEMPIDIR\textquotedbl{}~\textbackslash{}
-
-~~~~~~-~<~OUTPUT\_FILES/machines~>/dev/null~~\\
-~~~\\
-~\#~run~the~specfem~program
-
-current\_pwd=\$PWD~~\\
-
-cd~bin/~~\\
-
-run\_lsf.bash~-{}-gm-no-shmem~-{}-gm-copy-env~\$current\_pwd/xmeshfem3D~~\\
-
-run\_lsf.bash~-{}-gm-no-shmem~-{}-gm-copy-env~\$current\_pwd/xgenerate\_databases~~\\
-
-run\_lsf.bash~-{}-gm-no-shmem~-{}-gm-copy-env~\$current\_pwd/xspecfem3D~~\\
-~~~\\
-~\#~collect~seismograms~and~clean~up
-
-cd~current\_pwd/~~~\\
-
-mkdir~-p~SEM
-
-cd~SEM/~
-
-collect\_seismo.pl~../OUTPUT\_FILES/lsf\_machines
-
-cleanbase.pl~../OUTPUT\_FILES/machines
-\end{lyxcode}
-
 \chapter{{\normalsize \label{cha:-Changing-the}} Changing the Model}
@@
 
-%% magnoni 6/12
-
-\section{{\normalsize \label{sec:Using-tomographic}}Using external tomographic
-Earth models}
+\section{{\normalsize \label{sec:Using-tomographic}}Using external tomographic Earth models}
 

### 2014-07-18 doc/USER_MANUAL/manual_SPECFEM3D_Cartesian.tex

@@
 \usepackage[T1]{fontenc}
-\usepackage[latin1]{inputenc}
+%% DK DK we do not accept accented letters, use the old LaTeX syntax
+%% DK DK for them instead: \'e \`e  \"e \^e and so on.
+%% DK DK \usepackage[latin1]{inputenc}
 \usepackage[letterpaper]{geometry}
@@
 Piero Basini,
-Céline Blitz,
+C\'eline Blitz,
 Ebru Bozda\u{g},
@@
 Roland Martin,
-René Matzen,
+Ren\'e Matzen,
 Dennis McRitchie,
@@
 Peter Messmer,
-David Michéa,
+David Mich\'ea,
 Surendra Nadh Somala,
@@
 graph and mesh partitioning routines. SCOTCH is a free software package
-developed by François Pellegrini et al. from LaBRI and INRIA in Bordeaux,
+developed by Fran\c{c}ois Pellegrini et al. from LaBRI and INRIA in Bordeaux,
 France, downloadable from the web page \url{https://gforge.inria.fr/projects/scotch/}.
@@
 
-\noindent To compile the code on an IBM BlueGene, Laurent Léger from
+\noindent To compile the code on an IBM BlueGene, Laurent L\'eger from
 IDRIS, France, suggests the following: compile the code with
@@
 which provides efficent static mapping, graph and mesh partitioning
-routines. SCOTCH is a free software package developed by François
+routines. SCOTCH is a free software package developed by Fran\c{c}ois
 Pellegrini et al. from LaBRI and INRIA in Bordeaux, France, downloadable

### 2014-07-22 doc/USER_MANUAL/manual_SPECFEM3D_Cartesian.tex

@@
 many earthquakes in a catalog, in which case it can be better from
-a batch job submission point of few to start fewer and much larger jobs,
+a batch job submission point of view to start fewer and much larger jobs,
 each of them computing several earthquakes in parallel.
@@
 the code in the batch system must be a multiple of NUMBER\_OF\_SIMULTANEOUS\_RUNS,
-all the individual runs must use the same number of processor cores
-(which as usual is NPROC in the input file DATA/Par\_file,
+all the individual runs must use the same number of processor cores,
+which as usual is NPROC in the input file DATA/Par\_file,
 and thus the total number of processor cores to request from the batch system
-should be NUMBER\_OF\_SIMULTANEOUS\_RUNS * NPROC.
-All the runs to perform must be placed in directories called run0001, run0002, run0003 and so on (with exactly four digits).
+should be NUMBER\_OF\_SIMULTANEOUS\_RUNS $\times$ NPROC.
+All the runs to perform must be placed in directories called run0001, run0002, run0003 and so on (with exactly four digits)
+and you must create a link from the root directory of the code
+to the first copy of the executable programs by typing " ln -s run0001/bin bin ".
 

### 2014-07-31 ADJOINT_TOMOGRAPHY_TOOLS/flexwin/USER_MANUAL/manual_tuning.tex
added ADJOINT_TOMOGRAPHY_TOOLS (moved from the old and unused SVN repository to here)

\chapter{Tuning FLEXWIN for your seismograms\label{sec:tuning}}

FLEXWIN is adapted to your specific problem by modifying the values of the parameters in Table~\ref{tb:params}, and the functional form of those parameters that are time-dependent. We consider the algorithm to be correctly adapted when false positives (windows around undesirable features of the seismogram) are minimized, and true positives (window around desirable features) are maximized. The choice of what makes an adequate set of windows remains subjective, as it depends strongly on the quality of the input model, the quality of the data, and the region of the Earth that the tomographic inversion aims to constrain.

The base values of the various parameters are set in the {\tt PAR\_FILE}, which is read at run time.  Examples of base parameter values for the three tomographic scenarios discussed by \cite{MaggiEtal2009} can be found in Table~\ref{tb:example_params}.  The functional forms of the time dependent parameters may be adjusted by modifying {\tt user\_parameters.f90} (see next section), and re-compiling the code.

\begin{table}[h]
\begin{tabular}{lp{0.8\linewidth}}
\hline
\multicolumn{2}{l}{Standard tuning parameters:} \\[5pt]
$T_{0,1}$     & bandpass filter corner periods \\
$r_{P,A}$     & signal to noise ratios for whole waveform \\
$r_0(t)$      & signal to noise ratios single windows \\
$w_E(t)$      & water level on short-term:long-term ratio \\
$\mathrm{CC}_0(t)$          & acceptance level for normalized cross-correlation\\
$\Delta\tau_0(t)$  & acceptance level for time lag \\
$\Delta\ln{A}_0(t)$   & acceptance level for amplitude ratio \\ 
$\Delta\tau_{\rm ref}$ & reference time lag \\
$\Delta\ln{A}_{\rm ref}$ & reference amplitude ratio \\
\hline
\multicolumn{2}{l}{Fine tuning parameters:} \\ [5pt]
$c_0$ & for rejection of internal minima \\
$c_1$ & for rejection of short windows \\
$c_2$ & for rejection of un-prominent windows \\
$c_{3a,b}$  & for rejection of multiple distinct arrivals \\
$c_{4a,b}$  & for curtailing of windows with emergent starts and/or codas \\
$w_{\mathrm{CC}}\quad w_{\rm len}\quad w_{\rm nwin}$ & for selection of best non-overlapping window combination \\
\hline
\end{tabular}
\caption{\label{tb:params}
Overview of standard tuning parameters, and of fine
tuning parameters.  Values are defined in a parameter file, and the
time dependence of those that depend on time is described by user-defined functions.
} 
\end{table}

\pagebreak
\section{User parameters}
The main user parameters in the {\tt PAR\_FILE} are:
\begin{description}
\item[{\tt WIN\_MIN\_PERIOD}]Corresponds to $T_0$ in Table~\ref{tb:params}, and is the short wavelength cut-off for the band-pass filter applied to the raw synthetic and observed seismograms.
\item[{\tt WIN\_MAX\_PERIOD}]Corresponds to $T_1$ in Table~\ref{tb:params}, and is the long wavelength cut-off for the band-pass filter applied to the raw synthetic and observed seismograms.
\item[{\tt SNR\_INTEGRATE\_BASE}]Corresponds to $r_P$ in Table~\ref{tb:params}, and is the minimum signal to noise ratio on the power of the observed seismogram for windowing to continue.
\item[{\tt SNR\_MAX\_BASE}]Corresponds to $r_A$ in Table~\ref{tb:params}, and is the minimum signal to noise ratio on the modulus of the observed seismogram for windowing to continue.
\item[{\tt WINDOW\_S2N\_BASE}]Corresponds to $r_0$ in Table~\ref{tb:params}, and is the minimum signal to noise ratio for a window on the observed seismogram to be acceptable.
\item[{\tt STALTA\_BASE}]Corresponds to $w_E$ in Table~\ref{tb:params}, and is the water level to be applied to the synthetic short-term/long-term average waveform in order to generate candidate time windows.  See Figure~\ref{fg:win_composite}a.
\item[{\tt CC\_BASE}]Corresponds to $CC_0$ in Table~\ref{tb:params}, and is the minimum normalized cross-correlation value between synthetic and observed seismogram for a window to be acceptable.
\item[{\tt TSHIFT\_BASE}]Corresponds to $\Delta\tau_0$ in Table~\ref{tb:params}, and is the maximum cross-correlation lag (in seconds) between synthetic and observed seismogram for a window to be acceptable.
\item[{\tt DLNA\_BASE}]Corresponds to $\Delta\ln{A}_0$ in Table~\ref{tb:params}, and is the maximum amplitude ratio ($\Delta\ln{A}$ or $\Delta A/A$) between synthetic and observed seismogram for a window to be acceptable.
\item[{\tt TSHIFT\_REFERENCE}]Corresponds to $\Delta\tau_{\rm ref}$ in Table~\ref{tb:params}, and allows for a systematic traveltime bias in the synthetics.
\item[{\tt DLNA\_REFERENCE}]Corresponds to $\Delta\ln{A}_{\rm ref}$ in Table~\ref{tb:params}, and allows for a systematic amplitude bias in the synthetics.
\item[{\tt C\_0}]Corresponds to $C_0$ in Table~\ref{tb:params}, and is expressed as a multiple of $w_E$.  No window may contain a local minimum in its STA:LTA waveform that falls below the local value of $C_0 w_E$.  See Figure~\ref{fg:win_composite}b.
\item[{\tt C\_1}]Corresponds to $C_1$ in Table~\ref{tb:params}, and is expressed as a multiple of $T_0$.  No window may be shorter than $C_1 T_0$.
\item[{\tt C\_2}]Corresponds to $C_2$ in Table~\ref{tb:params}, and is expressed as a multiple of $w_E$.  A window whose seed maximum on the STA:LTA waveform rises less than $C_2 w_E$ above either of its adjacent minima is rejected.  See Figure~\ref{fg:win_composite}c.
\item[{\tt C\_{3a}}]Corresponds to $C_{3a}$ in Table~\ref{tb:params}, and is expressed as a fraction.  It regulates the acceptable height ratio between local maxima in a given window on the STA:LTA waveform.  See Figure~\ref{fg:separation}.
\item[{\tt C\_{3b}}]Corresponds to $C_{3b}$ in Table~\ref{tb:params}, and is expressed as a multiple of $T_0$.  It regulates the acceptable time separation between local maxima in a given window on the STA:LTA waveform.  See Figure~\ref{fg:separation}.
\item[{\tt C\_{4a}}]Corresponds to $C_{4a}$ in Table~\ref{tb:params}, and is expressed as a multiple of $T_0$.  It limits the length of a window before its first local maximum in STA:LTA. 
\item[{\tt C\_{4b}}]Corresponds to $C_{4b}$ in Table~\ref{tb:params}, and is expressed as a multiple of $T_0$.  It limits the length of a window beyond its last local maximum in STA:LTA.  See Figure~\ref{fg:win_composite}d.
\item[{\tt WEIGHT\_AVERAGE\_CC}]Corresponds to $w_{CC}$ in Table~\ref{tb:params}, and is the weight given to the average cross-correlation value in the process of resolving window overlaps. See Figure~\ref{fg:phaseD}.
\item[{\tt WEIGHT\_SPACE\_COVERAGE}]Corresponds to $w_{\rm len}$ in Table~\ref{tb:params}, and is the weight given to the time span covered by windows in the process of resolving window overlaps. 
\item[{\tt WEIGHT\_N\_WINDOWS}]Corresponds to $w_{\rm nwin}$ in Table~\ref{tb:params}, and is the weight given to the total number of windows in the process of resolving window overlaps. 
\end{description}

\begin{table}
\begin{center}
\begin{tabular}{lcccccc}
\hline
			& Global	& \multicolumn{2}{c}{Japan}	& \multicolumn{3}{c}{S. California} \\
\hline
$T_{0,1}$		& 50, 150	& 24, 120 	& 6, 30		& 6, 30		& 3, 30		& 2, 30		\\
$r_{P,A}$		& 3.5, 3.0	& 3.5, 3.0	& 3.5, 3.0	& 3.0, 2.5	& 2.5, 3.5	& 2.5, 3.5	\\
$r_0$			& 2.5		& 1.5		& 3.0		& 3.0		& 4.0		& 4.0		\\
$w_E$			& 0.08		& 0.11		& 0.12		& 0.18		& 0.11		& 0.07		\\
$\mathrm{CC}_0$		& 0.85		& 0.70		& 0.73		& 0.71		& 0.80		& 0.85		\\
$\Delta\tau_0$		& 15		& 12.0		& 3.0		& 8.0		& 4.0		& 3.0		\\
$\Delta\ln{A}_0$	& 1.0 		& 1.0		& 1.5		& 1.5		& 1.0		& 1.0		\\ 
$\Delta\tau_{\rm ref}$	& 0.0		& 0.0		& 0.0		& 4.0		& 2.0 		& 1.0		\\
$\Delta\ln{A}_{\rm ref}$& 0.0		& 0.0		& 0.0		& 0.0		& 0.0		& 0.0		\\
\hline
$c_0$			& 0.7		& 0.7		& 0.7		& 0.7		& 1.3		& 1.0		\\
$c_1$			& 4.0		& 3.0		& 3.0		& 2.0		& 4.0		& 5.0		\\
$c_2$			& 0.3		& 0.0		& 0.6		& 0.0		& 0.0		& 0.0		\\
$c_{3a,b}$		& 1.0, 2.0	& 1.0, 2.0	& 1.0, 2.0	& 3.0, 2.0	& 4.0, 2.5	& 4.0, 2.5	\\
$c_{4a,b}$		& 3.0, 10.0	& 3.0, 25.0	& 3.0, 12.0	& 2.5, 12.0	& 2.0, 6.0	& 2.0, 6.0	\\
$w_{\mathrm{CC}}, w_{\rm len}, w_{\rm nwin}$
			& 1, 1, 1 	& 1, 1, 1	& 1, 1, 1	& 0.5,1.0,0.7	& 0.70,0.25,0.05 & 1,1,1	\\
\hline
\end{tabular}
\caption{\label{tb:example_params}
Values of standard and fine-tuning parameters for the three seismological
scenarios discussed \cite{MaggiEtal2009}.  This table is identical to Table~3 of that study.
} 
\end{center}
\end{table}

\begin{figure}
\center \includegraphics[width=6in]{window_composite.pdf}
\caption{\label{fg:win_composite}
(a)~Window creation process.  The thick black line represents the STA:LTA
waveform $E(t)$, and the thick horizontal dashed line its water level $w_E(t)$.
Local maxima are indicated by alternating red and blue dots, windows are
indicated by two-headed horizontal arrows.  The time of the local maximum used
as the window seed $t_M$ is denoted by the position of the dot. Only windows for the fourth local maximum are shown.  (b)~Rejection of candidate windows based on the amplitude of the local minima.  The two deep
local minima indicated by the grey arrows form virtual barriers. All candidate
windows that cross these barriers are rejected.
(c)~Rejection of candidate
windows based on the prominence of the seed maximum.  The local maxima
indicated by the grey arrows are too low compared to the local minima
adjacent to them.  All windows that have these local maxima as their seed are
rejected (black crosses over the window segments below the time series).
(d)~Shortening of long coda windows.  The grey bar indicates the maximum coda
duration $c_{4b} T_0$.  Note that after the rejection based on prominence represented in (c) and before shortening of long coda windows represented in (d), the algorithm rejects candidate windows based on the separation of distinct phases, a process that is illustrated in Figure~\ref{fg:separation}.
}
\end{figure}

\begin{figure}
\center \includegraphics[width=5in]{window_rejection_separation.pdf}
\caption{\label{fg:separation}
Rejection of candidate windows based on the separation of distinct phases.
(a)~Heights of pairs of local maxima above their intervening minimum.
(b)~The black line represents the limiting value of $h/h_M$.  Vertical bars represent
$h/h_M$ for each pair of maxima.  Their position along the horizontal axis is
given by the time separation $\Delta T$ between the maxima of each pair.  The
color of the bar is given by the color of the seed maximum corresponding to $h_M$.  Bars whose height
exceeds the line represent windows to be rejected.
(c)~The windows that have been rejected by this criterion are indicated by black
crosses.
}
\end{figure}

\begin{figure}
\center \includegraphics[width=5in]{window_overlap.pdf}
\caption{\label{fg:phaseD}
The selection of the best non-overlapping window
combinations.  Each grey box represents a distinct group of windows.
Non-overlapping subsets of windows are shown on separate lines.  Only one
line from within each group will be chosen, the one corresponding to the
highest weighted score.  The resulting optimal set
of data windows is shown by thick arrows.}
\end{figure}

\section{Time dependence of user parameters}
A subset of the FLEXWIN parameters from Table~\ref{tb:params} are time-dependent (where time is measured along the seismogram).  This feature enables the user to exercise fine control of the windowing algorithm.  The user can modulate the time-dependence of these parameters by editing the {\tt set\_up\_criteria\_arrays} subroutine in the {\tt user\_functions.f90} file.  This subroutine is called after the seismograms have been read in, and the following variables have been set: 
\begin{description}
\item[{\tt npts, dt, b}]  Number of points, time step, time of first point with respect to the reference time of both seismograms.  The observed and synthetic seismograms should have identical values of these three quantities.
\item[{\tt evla, evlo, evdp, stla, stlo}] Event latitude, event longitude, event depth (km), station latitude, station longitude, read from the observed seismogram.
\item[{\tt azimuth, backazimuth, dist\_deg, dist\_km}]  Calculated from the event and station locations above.
\item[{\tt kstnm, knetwk, kcmpnm}] Station name, network name, component name, read from the observed seismogram.
\item[{\tt num\_phases, ph\_names, ph\_times}] Number, names and arrival times of standard seismic phases calculated through IASPEI91 using the event depth and epicentral distance.
\end{description}

The {\tt set\_up\_criteria\_arrays} subroutine first sets up non time-modulated versions of the time-dependent parameters using a simple loop over time:
{\small
\begin{verbatim}
! -----------------------------------------------------------------
! This is the basic version of the subroutine - no variation with time
! -----------------------------------------------------------------
  do i = 1, npts
    time=b+(i-1)*dt
    DLNA_LIMIT(i)=DLNA_BASE
    CC_LIMIT(i)=CC_BASE
    TSHIFT_LIMIT(i)=TSHIFT_BASE
    STALTA_W_LEVEL(i)=STALTA_BASE
    S2N_LIMIT(i)=WINDOW_AMP_BASE
  enddo
\end{verbatim}
}
It is then up to the user to modulate these values for the specific problem at hand.  For example, should the user want to discourage the algorithm from picking windows beyond the end of the first surface wave-train $R1$, the following lines should be added to the {\tt user\_functions.f90} file to raise the water level on the STA:LTA waveform by a factor of ten:
{\small
\begin{verbatim}
 ! --------------------------------
 ! Set approximate end of rayleigh wave arrival
 R_vel=3.2
 R_time=dist_km/R_vel
 ! --------------------------------
 ! modulate criteria in time
 do i = 1, npts
   time=b+(i-1)*dt
   if (time.gt.R_time) then
     STALTA_W_LEVEL(i)=STALTA_BASE*10.0     
   endif
 enddo
\end{verbatim}
}
Should the user want the algorithm to be less stringent in its requirements for the cross-correlation fit for the surface wave portion of the seismograms, and allow a greater travel-time lag for deeper events, the required lines could be:
{\small
\begin{verbatim}
 ! --------------------------------
 ! Set approximate end of rayleigh wave arrival
 R_vel=3.2
 R_time=dist_km/R_vel
 ! --------------------------------
 ! Set approximate start of love wave arrival
 Q_vel=4.2
 Q_time=dist_km/Q_vel
 ! --------------------------------
 ! modulate criteria in time
 do i = 1, npts
   time=b+(i-1)*dt
   ! if we are in the surface wave times, then make the cross-correlation
   ! criterion less severe
   if (time.gt.Q_time .and. time.lt.R_time) then
     CC_LIMIT(i)=0.9*CC_LIMIT(i)
   endif
   ! --------------------------------
   ! modulate criteria according to event depth
   !
   ! if an intermediate depth event
   if (evdp.ge.70 .and. evdp.lt.300) then
     TSHIFT_LIMIT(i)=TSHIFT_BASE*1.4
   ! if a deep event
   elseif (evdp.ge.300) then
     TSHIFT_LIMIT(i)=TSHIFT_BASE*1.7
   endif
 enddo
 \end{verbatim}
}
 The above examples illustrate the power of the {\tt user\_functions.f90} file.  The user can choose to include/exclude any portion of the seismogram, and to make the rejection criteria for windows more or less stringent on any other portion of the seismogram.  All the seismogram-dependent variables whose values are known when the {\tt set\_up\_criteria\_arrays} subroutine is executed may be used to inform these choices, leading to an infinite number of windowing possibilities.  The careful user will use knowledge of the properties of the observed data set, the limitations of the synthetic waveforms, and the final use to which the selected windows will be put in order to tailor the subroutine to the needs of each study.

 For a given set of data and synthetics, the {\tt PAR\_FILE} and {\tt user\_functions.f90} files uniquely determine the windowing results.

\subsection{Examples of user functions\label{ap:user_fn}}

%As concrete examples of how the time dependence of the tuning parameters can be exploited, we present here the functional forms of the time dependencies used for the three example tomographic scenarios (global, Japan and southern California) described in \cite{MaggiEtal2009}.
Here we present the time dependencies of tuning parameters used for three tomographic scenarios \citep{MaggiEtal2009}: global, Japan and southern California.
In each example we use predicted arrival times derived from 1D Earth models to help modulate certain parameters. Note, however, that the actual selection of individual windows is based on the details of the waveforms, and not on information from 1D Earth models.

\subsubsection{Global scenario\label{ap:user_global}}

In the following, $h$ indicates earthquake depth, $t_Q$ indicates the approximate start of the Love wave predicted by a group wave speed of 4.2~\kmps, and $t_R$ indicates the approximate end of the Rayleigh wave predicted by a group wave speed of 3.2~\kmps. In order to reduce the number of windows picked beyond R1, and to ensure that those selected beyond R1 are a very good match to the synthetic waveform, we raise the water level on the STA:LTA waveform and impose stricter criteria on the signal-to-noise ratio and the waveform similarity after the approximate end of the surface-wave arrivals.  We allow greater flexibility in cross-correlation time lag $\Delta\tau$ for intermediate depth and deep earthquakes.  We lower the cross-correlation value criterion for surface-waves in order to retain windows with a slight mismatch in dispersion characteristics.

We therefore use the following time modulations:
\begin{align}
w_E(t) & =
  \begin{cases}
    w_E \text{$t \leq t_R$} ,\\
    2 w_E \text{$t > t_R$},
  \end{cases}
\\
r_0(t) & = 
  \begin{cases}
    r_0 & \text{$t \leq t_R$}, \\
    10r_0 & \text{$t > t_R$} ,
  \end{cases}
\\
\mathrm{CC}_0(t) & = 
  \begin{cases}
    \mathrm{CC}_0 & \text{$t \leq t_R$}, \\
    0.9 \mathrm{CC}_0 & \text{$t_Q < t \leq t_R$}, \\
    0.95 & \text{$t > t_R$} ,
  \end{cases}
\\
\Delta\tau_0(t) & = 
  \begin{cases}
    \begin{cases}
      \tau_0 & \text{$t \leq t_R$}, \\
     \tau_0/3 & \text{$t > t_R$} ,
    \end{cases}
    & \text{$h \leq$ 70~km} \\
    1.4\tau_0 & \text{70~km $< h <$ 300~km}, \\
    1.7\tau_0 & \text{$h \geq$ 300~km},
  \end{cases}
  \\
\Delta \ln A_0(t) & = 
  \begin{cases}
    \Delta \ln A_0 & \text{$t \leq t_R$}, \\
    \Delta \ln A_0/3 & \text{$t > t_R$} .
  \end{cases}
\end{align}

%--------------------------

\subsubsection{Japan scenario\label{ap:user_japan}}
In the following, $t_P$ and $t_S$ denote the start of the time windows for $P$- and $S$ waves, as predicted by the 1-D IASPEI91 model \citep{KennettEngdahl1991}, and $t_{R1}$ indicates the end of the surface-wave time window.  For the \trange{24}{120} data, we consider the waveform between the start of the $P$ wave to the end of the surface-wave.  We therefore modulate $w_E(t)$ as follows:

%
\begin{align}
w_E(t) & =
  \begin{cases}
    10 w_E & \text{$t < t_P$}, \\
    w_E & \text{$t_P \le t \leq t_{R1}$}, \\
    10 w_E & \text{$t > t_{R1}$}.
  \end{cases}
\end{align}

For the \trange{6}{30} data, the fit between the synthetic and observed surface-waves is expected to be poor, as the 3D model used to calculate the synthetics cannot produce the required complexity. We therefore want to concentrate on body-wave arrivals only, and avoid surface-wave windows altogether by modulating $w_E(t)$ as follows:
%
\begin{align}
w_E(t) & =
  \begin{cases}
    10 w_E & \text{$t < t_P$}, \\
    w_E & \text{$t_P \le t \leq t_S$}, \\
    10 w_E & \text{$t > t_S$}.
  \end{cases}
\end{align}

We use constant values of $r_0(t)=r_0$, $\mathrm{CC}_0(t)=\mathrm{CC}_0$ and $\Delta \ln A_0(t)=\Delta \ln A_0$ for both period ranges.  In order to allow greater flexibility in cross-correlation time lag $\Delta\tau$ for intermediate depth and deep earthquakes we use:

\begin{align}
\Delta\tau_0(t) & = 
  \begin{cases}
    0.08 \text{$t_P$} & \text{$h \leq$ 70~km}, \\
    \max(0.05 \text{$t_P$}, 1.4\tau_0) & \text{70~km $< h <$ 300~km}, \\
    \max(0.05 \text{$t_P$}, 1.7\tau_0) & \text{$h \geq$ 300~km}.
  \end{cases}
\end{align}
%--------------------------

\subsubsection{Southern California scenario\label{ap:user_socal}}

In the following, $t_P$ and $t_S$ denote the start of the time windows for the crustal P wave and the crustal S wave, computed from a 1D layered model appropriate to Southern California \citep{Wald95}.  The start and end times for the surface-wave time window, $t_{R0}$ and $t_{R1}$, as well as the criteria for the time shifts $\Delta\tau_0(t)$, are derived from formulas in \cite{KomatitschEtal2004}. The source-receiver distance (in km) is denoted by $\Delta$.

%CHT modified

For the \trange{6}{30} and \trange{3}{30} data, we use constant values of $r_0(t)=r_0$, $\mathrm{CC}_0(t)=\mathrm{CC}_0$, $\Delta\tau_0(t)=\Delta\tau_0$, and $\Delta \ln A_0(t)=\Delta \ln A_0$. We exclude any arrivals before the $P$ wave and after the Rayleigh wave. This is achieved by the box-car function for $w_E(t)$:  
%
\begin{align}
w_E(t) & =
  \begin{cases}
    10 w_E & \text{$t < t_P$}, \\
    w_E & \text{$t_P \le t \leq t_{R1}$}, \\
    10 w_E & \text{$t > t_{R1}$},
  \end{cases}
\end{align}
%For the \trange{6}{30} data, we exclude any arrivals before the $P$ wave and reduce the number of windows picked beyond R1 by modulating $w_E(t)$.  We use constant values of $r_0(t)=r_0$, $\mathrm{CC}_0(t)=\mathrm{CC}_0$ and $\Delta \ln A_0(t)=\Delta \ln A_0$, but modulate the cross-correlation time lag criterion so that it is less strict at larger epicentral distances and for surface-waves.  We therefore use:  
%
%\begin{align}
%w_E(t) & =
%  \begin{cases}
%    10 w_E & \text{$t < t_P$}, \\
%    w_E & \text{$t_P \le t \leq t_{R1}$}, \\
%    2 w_E & \text{$t > t_{R1}$},
%  \end{cases}
%\\
%\Delta\tau_0(t) & = 
%  \begin{cases}
%    3.0 + \Delta/80.0 & \text{$t \le t_{R0}$}, \\
%    3.0 + \Delta/50.0 & \text{$t > t_{R0}$},
%  \end{cases}
%\end{align}

For the \trange{2}{30} data, we avoid selecting surface-wave arrivals as the 3D model used to calculate the synthetics cannot produce the required complexity. The water-level criteria then becomes:

\begin{align}
w_E(t) & =
  \begin{cases}
    10 w_E & \text{$t < t_P$}, \\
    w_E & \text{$t_P \le t \leq t_S$}, \\
    10 w_E & \text{$t > t_S$}.
  \end{cases}
%\\
%\Delta\tau_0(t) & = \Delta\tau_0.
\end{align}

%-----------------------

\subsection{Time intervals for signal-to-noise calculations}

If the parameter {\tt DATA$\_$QUALITY = .true.} in the {\tt PAR$\_$FILE}, then the {\tt user\_functions.f90} file requires the specification of four times that define the ``noise'' and ``signal'' time intervals.  For example, the default {\tt user\_functions.f90} file contains these lines:
%
\begin{verbatim}
  ! these values will be used for signal2noise calculations
  if (DATA_QUALITY) then
    noise_start=b
    noise_end=ph_times(1)-WIN_MAX_PERIOD
    signal_start=noise_end
    signal_end=b+(npts-1)*dt
  endif
\end{verbatim}
%
The use of all four variables allows provides the flexibility of choosing these time intervals.  (For example, the start of the noise does not need to be the beginning of the seismogram. Or the start of the signal does not have to coincide with the end of the noise.)

%-----------------------

\section{Tuning considerations}
FLEXWIN is not a black-box application and should not be applied blindly
to any given dataset or tomographic scenario.  The data windowing required by
any given problem will differ depending on the inversion method, the scale of
the problem (local, regional, global), the quality of the data set, the quality of the model,
and the accuracy of the method used to calculate the synthetic seismograms.  The user
must configure and tune the algorithm for the given problem.  Here we
discuss general considerations the user should bear in mind.

We suggest the following as a practical starting sequence for tuning the algorithm.
Keep in mind that this process may need to be repeated and refined several times before
converging on the optimal set of parameters for a given problem and data-set.

$T_{0,1}$ : In setting the corner periods of the bandpass filter, the
user is deciding on the frequency content of the information to be used in the
tomographic problem.  Values of these corner periods should reflect the
information content of the data, the quality of the Earth model and the
accuracy of the simulation used to generate the synthetic seismogram.  The
frequency content in the data depends on the spectral characteristics of the
source, on the instrument responses, and on the attenuation
characteristics of the medium. As $T_{0,1}$ depend on the source and station
characteristics, which may be heterogeneous in any given data-set, these filter
periods can be modified dynamically by constructing an appropriate user
function (e.g. {\em if station is in list of stations with instrument X then
reset T0 and T1 to new values}).

$r_{P,A}$ : In setting the signal-to-noise ratios for the entire seismogram, the
user is applying a simple quality control on the data.  Note that these criteria 
are applied after filtering.  No windows will be defined on data that fail this
quality control.

$w_E(t)$ : For a constant signal the short-term average long-term average ratio, $E(t)$,
converges to a constant value when
the length of the time-series is greater than the effective averaging length of
the long-term average.  This value is 0.08 for the short-term average long-term average ratio used in FLEXWIN (it has a small dependence on $T_0$, which can be ignored in most applications).  We suggest the user start with a constant
level for $w_E(t)$ equal to this convergence value.  The time dependence of
$w_E(t)$ should then be adjusted to exclude those portions of the waveform the
user is not interested in, by raising $w_E(t)$ (e.g. to exclude the fundamental
mode surface-wave: {\em if t $>$ fundamental mode surface-wave arrival time then set $w_E(t)=1$}).  
We suggest finer adjustments to $w_E(t)$ be made after $r0(t)$,
$CC_0(t)$, $\Delta T_0(t)$ and $\Delta \ln A_0(t)$ have been configured.

$r_0(t)$, $\mathrm{CC}_0(t)$, $\Delta \tau_{\rm ref}$, $\Delta
\tau_0(t)$, $\Delta \ln A_{\rm ref}$ and $\Delta \ln A_0(t)$ : These parameters ---
window signal-to-noise ratio, normalized cross-correlation value between
observed and synthetic seismograms, cross-correlation time lag, and amplitude
ratio --- control the degree of well-behavedness of the data within accepted
windows.  The user first sets constant values for these four parameters, then
adds a time dependence if required.  Considerations that should be taken into
account include the quality of the Earth model used to calculate the synthetic
seismograms, the frequency range, the dispersed nature of certain arrivals (e.g.
{\em for t corresponding to the group velocities of surface-waves, reduce
$CC_0(t)$}), and {\em a priori} preferences for picking certain small-amplitude seismic phases
(e.g. {\em for t close to the expected arrival for $P_{\rm diff}$, reduce $r_0(t)$}).  
$\Delta \tau_{\rm ref}$ and $\Delta \ln A_{\rm ref}$ should be set to zero at first, and only
reset if the synthetics contain a systematic bias in traveltimes or amplitudes.

$c_{0-4}$ : These parameters control the process by which the suite of all possible data windows is pared down using criteria on the shape of the STA:LTA $E(t)$ waveform alone.  We suggest the user start by setting these values to those used in our global example (see Table~\ref{tb:example_params}).  Subsequent minimal tuning should be performed by running the algorithm on a subset of the data and closely examining the lists of windows rejected at each stage to make sure the user agrees with the choices made by the algorithm.

$w_{\mathrm{CC}}$, $w_{\rm len}$ and $w_{\rm nwin}$ : These parameters control the overlap resolution stage of the algorithm.  Values of $w_{\mathrm{CC}}= w_{\rm len} = w_{\rm nwin} = 1$ should be reasonable for most applications.

The objective of the tuning process summarized here should be to maximize the selection of windows around desirable features in the seismogram, while minimizing the selection of undesirable features.
Of course, the desirability or undesirability of a given feature is subjective, and it depends on how the user subsequently intends to use the information contained within the data windows.

### 2014-07-31 ADJOINT_TOMOGRAPHY_TOOLS/ADJOINT_TOMOGRAPHY_TOOLKIT/SHARE_FILES/HEADER_FILES/values_from_mesher.h

@@ -1,0 +1,189 @@
+ 
+ !
+ ! this is the parameter file for static compilation of the solver
+ !
+ ! mesh statistics:
+ ! ---------------
+ !
+ !
+ ! number of chunks =            1
+ !
+ ! these statistics do not include the central cube
+ !
+ ! number of processors =          100
+ !
+ ! maximum number of points per region =       475797
+ !
+ ! on NEC SX, make sure "loopcnt=" parameter
+ ! in Makefile is greater than max vector length =      1427391
+ !
+ ! total elements per slice =         7821
+ ! total points per slice =       528135
+ !
+ ! total for full 6-chunk mesh:
+ ! ---------------------------
+ !
+ ! exact total number of spectral elements in entire mesh = 
+ !    4692600.00000000     
+ ! approximate total number of points in entire mesh = 
+ !    316881000.000000     
+ ! approximate total number of degrees of freedom in entire mesh = 
+ !    892096200.000000     
+ !
+ ! position of the mesh chunk at the surface:
+ ! -----------------------------------------
+ !
+ ! angular size in first direction in degrees =    68.00000    
+ ! angular size in second direction in degrees =    68.00000    
+ !
+ ! longitude of center in degrees =    7.000000    
+ ! latitude of center in degrees =    52.00000    
+ !
+ ! angle of rotation of the first chunk =    45.00000    
+ !
+ ! corner            1
+ ! longitude in degrees =    6.99999999999999     
+ ! latitude in degrees =    8.40726023480213     
+ !
+ ! corner            2
+ ! longitude in degrees =    64.1611658328499     
+ ! latitude in degrees =    34.9449365466497     
+ !
+ ! corner            3
+ ! longitude in degrees =   -50.1611658328499     
+ ! latitude in degrees =    34.9449365466497     
+ !
+ ! corner            4
+ ! longitude in degrees =   -173.000000000000     
+ ! latitude in degrees =    84.3892903527184     
+ !
+ ! resolution of the mesh at the surface:
+ ! -------------------------------------
+ !
[…]

### 2014-07-31 ADJOINT_TOMOGRAPHY_TOOLS/flexwin/io_subs.f90

@@ -1,0 +1,442 @@
+!----------------------------------------------------------------------
+
+  subroutine write_seis_gmt(basename)
+  use seismo_variables
+  implicit none
+
+  character*120 :: basename
+  character*240 :: file_obs, file_syn
+  double precision :: max_value_obs, max_value_syn, max_value
+
+  integer :: i,ier
+
+  ! create the filenames
+  file_obs=trim(basename)//'.obs'
+  file_syn=trim(basename)//'.syn'
+
+  ! find the maximum value
+  !max_value_obs = max (maxval(obs_lp), abs(minval(obs_lp)) )
+  !max_value_syn = max (maxval(synt_lp), abs(minval(synt_lp)) )
+  max_value_obs = maxval( abs(obs_lp) )
+  max_value_syn = maxval( abs(synt_lp) )
+  max_value = max(max_value_obs, max_value_syn)
+
+  ! open the files for writing
+  open(unit=11, file=file_obs,iostat=ier)
+  if( ier /= 0 ) then
+    print*,'Error opening: ',trim(file_obs)
+    stop
+  endif
+  open(unit=12, file=file_syn,iostat=ier)
+  if( ier /= 0 ) then
+    print*,'Error opening: ',trim(file_syn)
+    stop
+  endif
+
+  ! write the header - observed
+  write(11,'("# NPTS = ",i10)') npts
+  write(11,'("# PLOT_MAX = ",e12.6)') max_value
+  write(11,'("# T_START = ",f10.2)') b
+  write(11,'("# T_END = ",f10.2)') b+(npts-1)*dt
+
+  ! write the header - synthetic
+  write(12,'("# NPTS = ",i10)') npts
+  write(12,'("# PLOT_MAX = ",e12.6)') max_value
+  write(12,'("# T_START = ",f10.2)') b
+  write(12,'("# T_END = ",f10.2)') b+(npts-1)*dt
+
+  ! write the time series
+  do i = 1, npts
+    write(11,'(f10.2,2x,e12.6)') b+(i-1)*dt, obs_lp(i)
+    write(12,'(f10.2,2x,e12.6)') b+(i-1)*dt, synt_lp(i)
+  end do
+
+  ! close the files
+  close(11)
+  close(12)
+
+  end subroutine
+
+!----------------------------------------------------------------------
+
+  subroutine write_env_gmt(basename)
+  use seismo_variables
+  implicit none
+
+  character*120 :: basename
+  character*240 :: file_obs_env, file_syn_env
[…]

### 2014-07-31 ADJOINT_TOMOGRAPHY_TOOLS/iterate_adj/model_plot/gmt/scripts/ps2eps.pl

@@ -1,0 +1,37 @@
+#!/usr/bin/perl -w
+
+$ivert = 0;
+$irun = 1;
+$stip = sprintf("%2.2i",$irun);
+
+#$ftag = "vert_${stip}_xc";
+#$ftag = "horz_xc_vb_m16_m00";
+$ftag = "horz_xc_vs_m01_m00";
+
+@epsfiles = glob("${ftag}*.ps");
+
+ $neps = @epsfiles;
+ for ($k = 1; $k <= $neps; $k = $k+1) {
+   $efile = $epsfiles[$k-1];
+   `ps2eps -f $efile`;
+ }
+
+if ($irun==1 && $ivert==1) {
+  `ps2ps vert_01_xc_HEC_CI_282_9968977_vs_m16_m00_003.ps temp.ps ; ps2eps temp.ps ; mv temp.eps vert_01_xc_HEC_CI_282_9968977_vs_m16_m00_003.eps`;
+  `ps2ps vert_01_xc_LCG_CI_122_10215753_vs_m16_m00_021.ps temp.ps ; ps2eps temp.ps ; mv temp.eps vert_01_xc_LCG_CI_122_10215753_vs_m16_m00_021.eps`;
+  `ps2ps vert_01_xc_LGU_CI_353_10097009_vs_m16_m00_067.ps temp.ps ; ps2eps temp.ps ; mv temp.eps vert_01_xc_LGU_CI_353_10097009_vs_m16_m00_067.eps`;
+  `ps2ps vert_01_xc_BZN_AZ_306_14138080_vs_m16_m00_013.ps temp.ps ; ps2eps temp.ps ; mv temp.eps vert_01_xc_BZN_AZ_306_14138080_vs_m16_m00_013.eps`;
+  `ps2ps vert_01_xc_EDW2_CI_134_14236768_vs_m16_m00_011.ps temp.ps ; ps2eps temp.ps ; mv temp.eps vert_01_xc_EDW2_CI_134_14236768_vs_m16_m00_011.eps`;
+  `ps2ps vert_01_xc_ISA_CI_160_14383980_vs_m16_m00_049.ps temp.ps ; ps2eps temp.ps ; mv temp.eps vert_01_xc_ISA_CI_160_14383980_vs_m16_m00_049.eps`;
[…]

### 2014-07-31 ADJOINT_TOMOGRAPHY_TOOLS/iterate_adj/model_plot/matlab/matlab_scripts/raypath.m

@@ -1,0 +1,71 @@
+%
+% function [rdist,lon,lat] = raypath(lat1,lon1,preseg,postseg,dx)
+% CARL TAPE, 05-Dec-2008
+% printed xxx
+%
+% This function returns the values of points on a great-circle "ray path",
+% given a starting point and a finishing point.  The ray path is
+% constructed from two segments, each extending from the starting point.
+% The spacing between points will not necessarily be the same for the two
+% segments, and the finishing point will not necessarily be exactly one of
+% the digitized points on the ray path.
+%
+% calls xxx
+% called by xxx
+%
+
+function [lon,lat,rdist,azis] = raypath(lon1,lat1,lon2,lat2,preseg,postseg,dx)
+
+% distance and azimuth from point 1 to point 2
+[rng, az] = distance(lat1,lon1,lat2,lon2);
+
+% endpoints of the ray path, including the extensions
+%[latA,lonA] = reckon(lat1,lon1,km2deg(preseg),wrapTo360(az+180) );
+%[latB,lonB] = reckon(lat1,lon1,rng + km2deg(postseg),wrapTo360(az+180) );
+
+% lengths of two segments and number of legs per segment
+dist1 = preseg;
+dist2 = postseg+deg2km(rng);
+n1 = ceil(dist1/dx)+1;
+n2 = ceil(dist2/dx)+1;
+
+% two segments
+[lats1,lons1] = track1(lat1,lon1,wrapTo360(az+180),km2deg(dist1),[],[],n1);
+[lats2,lons2] = track1(lat1,lon1,az,km2deg(dist2),[],[],n2);
+npt1 = length(lats1);
+npt2 = length(lats2);
+npt = npt1 + npt2;
+
+% connect everything
+lon = zeros(npt,1);
+lat = zeros(npt,1);
+lon = [flipud(lons1) ; lons2(2:end) ];
+lat = [flipud(lats1) ; lats2(2:end) ];
+n = length(lon);
+
+[ddist,azis] = distance(lat1*ones(n,1),lon1*ones(n,1),lat,lon);
+rdist = deg2km(ddist);
+rdist(1:npt1) = -rdist(1:npt1);
+
+%------------------------------------
+
+% EXAMPLE
+if 0==1
+    lon1 = -120; lat1 = 36;
+    lon2 = -117; lat2 = 33;
+    preseg = 50; postseg = 50;
+    dx = 10;
+    
[…]

### 2014-07-31 ADJOINT_TOMOGRAPHY_TOOLS/measure_adj/UTIL/pssac2/pssac2.c

@@ -1,0 +1,939 @@
+/*--------------------------------------------------------------------
+ *    The GMT-system:	@(#)pssac.c	2.13  2/2/95
+ *
+ *    Copyright (c) 1991 by P. Wessel and W. H. F. Smith
+ *    See README file for copying and redistribution conditions.
+ *--------------------------------------------------------------------
+ * pssac will read sacfiles and plot them
+ * PostScript code is written to stdout.
+ *
+ * Revision:	Lupei Zhu from psxy
+ * Modifications: Upgraded to new version of GMT (Brian Savage)
+ *                Small updates
+ *                Added ability to plot seismograms on maps
+ *  2004 August 3   - Added Body wave scaling using JB Tables
+ *  2006 December 5 - Will only use the absolute value of the distance/gcarc
+ *                    Useful for negative distances
+ */
+
+#include <stdio.h>
+#include <string.h>
+#include <fcntl.h>
+#include <math.h>
+
+/* GMT and Postscript */
+#include <gmt.h>
+#include <pslib.h>
+
+/* Local include files */
+#include "sac.h"
+#include "jb.h"
+
+#define PI M_PI
+#define LINE_SIZE 10512
+
+#define GMT_VERSION_MAJOR  4
+#define GMT_VERSION_MINOR  3
+
+#define PSSAC "pssac2"
+
+void GMT_history(int argc, char *argv[]);
+
+static void
+usage(void) {
+  fprintf (stderr,"%s: Plot seismograms\n", PSSAC);
+  fprintf(stderr,"Usage: %s <infiles> -C<cut0/cut1> -J<params> -R<w/e/s/n>\n", PSSAC);
+  fprintf(stderr,"      [-B<tickinfo>] [-E(k|d|a|n|b)(t(0-9,-5(b),-3(o),-2(a))|vel)] [-K] [-V]\n");
+  fprintf(stderr,"      [-Msize/alpha] [-Gr/g/b/c/t1/t2] [-O] [-P] [-Sshift] [-U[<label>]]\n");
+  fprintf(stderr,"      [-W<pen>] [-#<ncopies>] [-X<x_shift>] [-Y<y_shift>] [-r]\n");
+  fprintf(stderr,"\n");
+  fprintf(stderr,"      sacfiles may be given on either the command line or\n");
+  fprintf(stderr,"      if none are given they will be expected using standard in (stdin)\n");
[…]

### 2014-07-31 ADJOINT_TOMOGRAPHY_TOOLS/measure_adj/ma_variables.f90

@@ -1,0 +1,56 @@
+module ma_variables
+
+  use ma_constants
+!
+! multi-taper measurements
+!
+! Ying Zhou: The fit between the recovered data and the data can be improved
+! by either increasing the window width (HWIN above) or by decreasing NPI.
+! In her experience, NPI = 2.5 is good for noisy data.
+! For synthetic data, we can use a lower NPI.
+! number of tapers should be fixed as twice NPI -- see Latex notes
+!
+! See write_par_file.pl and measure_adj.f90
+
+  character(len=150) :: OUT_DIR
+
+  double precision :: TLONG, TSHORT
+  double precision :: WTR, NPI, DT_FAC, ERR_FAC, DT_MAX_SCALE, NCYCLE_IN_WINDOW
+  !double precision :: BEFORE_QUALITY, AFTER_QUALITY, BEFORE_TSHIFT, AFTER_TSHIFT
+  double precision :: TSHIFT_MIN, TSHIFT_MAX, DLNA_MIN, DLNA_MAX, CC_MIN
+  double precision :: DT_SIGMA_MIN, DLNA_SIGMA_MIN
+
+  integer :: ntaper, ipwr_t, ipwr_w, ERROR_TYPE
+  integer :: imeas0, imeas, itaper, is_mtm0, is_mtm
+
+  logical :: DISPLAY_DETAILS,OUTPUT_MEASUREMENT_FILES,RUN_BANDPASS,COMPUTE_ADJOINT_SOURCE,USE_PHYSICAL_DISPERSION
+
+end module ma_variables
+
+
+module ma_weighting
+
+! module for weighting/normalizing measurements
+
+  logical,parameter :: DO_WEIGHTING = .false.
+
+  ! transverse, radial and vertical weights
+  double precision :: weight_T, weight_R, weight_Z
+  ! body waves: number of picks on vertical, radial and transverse component
+  double precision :: num_P_SV_V,num_P_SV_R,num_SH_T
+  ! surface waves: number of pick on vertical, radial and transverse
+  double precision :: num_Rayleigh_V,num_Rayleigh_R,num_Love_T
+
+  ! typical surface wave speed in km/s, to calculate surface wave arrival times
+  ! Love waves faster than Rayleigh
+  double precision, parameter :: surface_vel = 4.0
+
+  ! wave type pick
+  integer, parameter :: P_SV_V = 1
+  integer, parameter :: P_SV_R = 2
[…]

### 2014-08-01 ADJOINT_TOMOGRAPHY_TOOLS/flexwin/USER_MANUAL/manual_other.tex

\chapter{Miscellaneous}

\section{Bug reports and suggestions for improvements}
To report bugs or suggest improvements to the code, please send an email to the CIG Computational Seismology Mailing List (cig-seismo@geodynamics.org) or Alessia Maggi (alessia@unistra.fr), and/or use our online bug tracking system Roundup (www.geodynamics.org/roundup).

\section{Notes and Acknowledgments}
The main developers of the FLEXWIN source code are Alessia Maggi and Carl Tape.  The following individuals (listed in alphabetical order) have also contributed to the development of the source code: Daniel Chao, Min Chen, Vala Hjorleifsdottir, Qinya Liu, Jeroen Tromp.  The following individuals (listed in alphabetical order) contributed to this manual: Sue Kientz, Alessia Maggi, Carl Tape.

The FLEXWIN code makes use of filtering and enveloping algorithms that are part of SAC (Seismic Analysis Code, Lawerence Livermore National Laboratory) provided for free to IRIS members.  We thank Brian Savage for adding interfaces to these algorithms in recent SAC distributions.

We acknowledge support by the National Science Foundation under grant EAR-0711177.
Daniel Chao received additional support from a California Institute of Technology Summer Undergraduate Reseach Fellowship.

\section{License}
\singlespacing
\setlength\parskip{8pt}

{\center \Large \bf CeCILL FREE SOFTWARE LICENSE AGREEMENT}

\subsection*{Notice}

This Agreement is a Free Software license agreement that is the result of discussions between its authors in order to ensure compliance with the two main principles guiding its drafting:

\begin{itemize}
\item firstly, compliance with the principles governing the distribution of Free Software: access to source code, broad rights granted to users,
\item secondly, the election of a governing law, French law, with which it is conformant, both as regards the law of torts and intellectual property law, and the protection that it offers to both authors and holders of the economic rights over software.
\end{itemize}

The authors of the CeCILL\footnote{CeCILL stands for Ce(a) C(nrs) I(nria) L(ogiciel) L(ibre)
Version 2.0 dated 2006-09-05.} license are:

Commissariat à l'Energie Atomique - CEA, a public scientific, technical and industrial research establishment, having its principal place of business at 25 rue Leblanc, immeuble Le Ponant D, 75015 Paris, France.

Centre National de la Recherche Scientifique - CNRS, a public scientific and technological establishment, having its principal place of business at 3 rue Michel-Ange, 75794 Paris cedex 16, France.

Institut National de Recherche en Informatique et en Automatique - INRIA, a public scientific and technological establishment, having its principal place of business at Domaine de Voluceau, Rocquencourt, BP 105, 78153 Le Chesnay cedex, France.

\subsection*{Preamble}

The purpose of this Free Software license agreement is to grant users the right to modify and redistribute the software governed by this license within the framework of an open source distribution model.

The exercising of these rights is conditional upon certain obligations for users so as to preserve this status for all subsequent redistributions.

In consideration of access to the source code and the rights to copy, modify and redistribute granted by the license, users are provided only with a limited warranty and the software's author, the holder of the economic rights, and the successive licensors only have limited liability.

In this respect, the risks associated with loading, using, modifying and/or developing or reproducing the software by the user are brought to the user's attention, given its Free Software status, which may make it complicated to use, with the result that its use is reserved for developers and experienced professionals having in-depth computer knowledge. Users are therefore encouraged to load and test the suitability of the software as regards their requirements in conditions enabling the security of their systems and/or data to be ensured and, more generally, to use and operate it in the same conditions of security. This Agreement may be freely reproduced and published, provided it is not altered, and that no provisions are either added or removed herefrom.

This Agreement may apply to any or all software for which the holder of the economic rights decides to submit the use thereof to its provisions.

\subsection*{Article 1 - DEFINITIONS}

For the purpose of this Agreement, when the following expressions commence with a capital letter, they shall have the following meaning:
\begin{description}
\item[Agreement:] means this license agreement, and its possible subsequent versions and annexes.
\item[Software:] means the software in its Object Code and/or Source Code form and, where applicable, its documentation, "as is" when the Licensee accepts the Agreement.
\item[Initial Software:] means the Software in its Source Code and possibly its Object Code form and, where applicable, its documentation, "as is" when it is first distributed under the terms and conditions of the Agreement.\
\item[Modified Software:] means the Software modified by at least one Contribution.
\item[Source Code:] means all the Software's instructions and program lines to which access is required so as to modify the Software.
\item[Object Code:] means the binary files originating from the compilation of the Source Code.
\item[Holder:] means the holder(s) of the economic rights over the Initial Software.
\item[Licensee:] means the Software user(s) having accepted the Agreement.
\item[Contributor:] means a Licensee having made at least one Contribution.
\item[Licensor:] means the Holder, or any other individual or legal entity, who distributes the Software under the Agreement.
\item[Contribution:] means any or all modifications, corrections, translations, adaptations and/or new functions integrated into the Software by any or all Contributors, as well as any or all Internal Modules.
\item[Module:] means a set of sources files including their documentation that enables supplementary functions or services in addition to those offered by the Software.
\item[External Module:] means any or all Modules, not derived from the Software, so that this Module and the Software run in separate address spaces, with one calling the other when they are run.
\item[Internal Module:] means any or all Module, connected to the Software so that they both execute in the same address space.
\item[GNU GPL:] means the GNU General Public License version 2 or any subsequent version, as published by the Free Software Foundation Inc.
\item[Parties:] mean both the Licensee and the Licensor.
\end{description}

These expressions may be used both in singular and plural form.

\subsection*{Article 2 - PURPOSE}

The purpose of the Agreement is the grant by the Licensor to the Licensee of a non-exclusive, transferable and worldwide license for the Software as set forth in Article 5 hereinafter for the whole term of the protection granted by the rights over said Software.

\subsection*{Article 3 - ACCEPTANCE}

3.1 The Licensee shall be deemed as having accepted the terms and conditions of this Agreement upon the occurrence of the first of the following events:\\
(i) loading the Software by any or all means, notably, by downloading from a remote server, or by loading from a physical medium;\\
(ii) the first time the Licensee exercises any of the rights granted hereunder.

3.2 One copy of the Agreement, containing a notice relating to the characteristics of the Software, to the limited warranty, and to the fact that its use is restricted to experienced users has been provided to the Licensee prior to its acceptance as set forth in Article 3.1 hereinabove, and the Licensee hereby acknowledges that it has read and understood it.

\subsection*{Article 4 - EFFECTIVE DATE AND TERM}

4.1 EFFECTIVE DATE

The Agreement shall become effective on the date when it is accepted by the Licensee as set forth in Article 3.1.

4.2 TERM

The Agreement shall remain in force for the entire legal term of protection of the economic rights over the Software.

\subsection*{Article 5 - SCOPE OF RIGHTS GRANTED}

The Licensor hereby grants to the Licensee, who accepts, the following rights over the Software for any or all use, and for the term of the Agreement, on the basis of the terms and conditions set forth hereinafter.

Besides, if the Licensor owns or comes to own one or more patents protecting all or part of the functions of the Software or of its components, the Licensor undertakes not to enforce the rights granted by these patents against successive Licensees using, exploiting or modifying the Software. If these patents are transferred, the Licensor undertakes to have the transferees subscribe to the obligations set forth in this paragraph.

5.1 RIGHT OF USE

The Licensee is authorized to use the Software, without any limitation as to its fields of application, with it being hereinafter specified that this comprises:\\
1. permanent or temporary reproduction of all or part of the Software by any or all means and in any or all form.\\
2. loading, displaying, running, or storing the Software on any or all medium.\\
3. entitlement to observe, study or test its operation so as to determine the ideas and principles behind any or all constituent elements of said Software. This shall apply when the Licensee carries out any or all loading, displaying, running, transmission or storage operation as regards the Software, that it is entitled to carry out hereunder.

5.2 ENTITLEMENT TO MAKE CONTRIBUTIONS

The right to make Contributions includes the right to translate, adapt, arrange, or make any or all modifications to the Software, and the right to reproduce the resulting software.

The Licensee is authorized to make any or all Contributions to the Software provided that it includes an explicit notice that it is the author of said Contribution and indicates the date of the creation thereof.

5.3 RIGHT OF DISTRIBUTION

In particular, the right of distribution includes the right to publish, transmit and communicate the Software to the general public on any or all medium, and by any or all means, and the right to market, either in consideration of a fee, or free of charge, one or more copies of the Software by any means.

The Licensee is further authorized to distribute copies of the modified or unmodified Software to third parties according to the terms and conditions set forth hereinafter.

5.3.1 DISTRIBUTION OF SOFTWARE WITHOUT MODIFICATION

The Licensee is authorized to distribute true copies of the Software in Source Code or Object Code form, provided that said distribution complies with all the provisions of the Agreement and is accompanied by:\\
1. a copy of the Agreement,\\
2. a notice relating to the limitation of both the Licensor's warranty and liability as set forth in Articles 8 and 9,\\
and that, in the event that only the Object Code of the Software is redistributed, the Licensee allows future Licensees unhindered access to the full Source Code of the Software by indicating how to access it, it being understood that the additional cost of acquiring the Source Code shall not exceed the cost of transferring the data.

5.3.2 DISTRIBUTION OF MODIFIED SOFTWARE

When the Licensee makes a Contribution to the Software, the terms and conditions for the distribution of the resulting Modified Software become subject to all the provisions of this Agreement.

The Licensee is authorized to distribute the Modified Software, in source code or object code form, provided that said distribution complies with all the provisions of the Agreement and is accompanied by: \\
1. a copy of the Agreement,
2. a notice relating to the limitation of both the Licensor's warranty and liability as set forth in Articles 8 and 9, \\
and that, in the event that only the object code of the Modified Software is redistributed, the Licensee allows future Licensees unhindered access to the full source code of the Modified Software by indicating how to access it, it being understood that the additional cost of acquiring the source code shall not exceed the cost of transferring the data.

5.3.3 DISTRIBUTION OF EXTERNAL MODULES

When the Licensee has developed an External Module, the terms and conditions of this Agreement do not apply to said External Module, that may be distributed under a separate license agreement.

5.3.4 COMPATIBILITY WITH THE GNU GPL

The Licensee can include a code that is subject to the provisions of one of the versions of the GNU GPL in the Modified or unmodified Software, and distribute that entire code under the terms of the same version of the GNU GPL.

The Licensee can include the Modified or unmodified Software in a code that is subject to the provisions of one of the versions of the GNU GPL, and distribute that entire code under the terms of the same version of the GNU GPL.

\subsection*{Article 6 - INTELLECTUAL PROPERTY}

6.1 OVER THE INITIAL SOFTWARE

The Holder owns the economic rights over the Initial Software. Any or all use of the Initial Software is subject to compliance with the terms and conditions under which the Holder has elected to distribute its work and no one shall be entitled to modify the terms and conditions for the distribution of said Initial Software.

The Holder undertakes that the Initial Software will remain ruled at least by this Agreement, for the duration set forth in Article 4.2.

6.2 OVER THE CONTRIBUTIONS

The Licensee who develops a Contribution is the owner of the intellectual property rights over this Contribution as defined by applicable law.

6.3 OVER THE EXTERNAL MODULES

The Licensee who develops an External Module is the owner of the intellectual property rights over this External Module as defined by applicable law and is free to choose the type of agreement that shall govern its distribution.

6.4 JOINT PROVISIONS

The Licensee expressly undertakes: \\
1. not to remove, or modify, in any manner, the intellectual property notices attached to the Software; \\
2. to reproduce said notices, in an identical manner, in the copies of the Software modified or not.

The Licensee undertakes not to directly or indirectly infringe the intellectual property rights of the Holder and/or Contributors on the Software and to take, where applicable, vis-à-vis its staff, any and all measures required to ensure respect of said intellectual property rights of the Holder and/or Contributors.

\subsection*{Article 7 - RELATED SERVICES}

7.1 Under no circumstances shall the Agreement oblige the Licensor to provide technical assistance or maintenance services for the Software.

However, the Licensor is entitled to offer this type of services. The terms and conditions of such technical assistance, and/or such maintenance, shall be set forth in a separate instrument. Only the Licensor offering said maintenance and/or technical assistance services shall incur liability therefor.

7.2 Similarly, any Licensor is entitled to offer to its licensees, under its sole responsibility, a warranty, that shall only be binding upon itself, for the redistribution of the Software and/or the Modified Software, under terms and conditions that it is free to decide. Said warranty, and the financial terms and conditions of its application, shall be subject of a separate instrument executed between the Licensor and the Licensee.

\subsection*{Article 8 - LIABILITY}

8.1 Subject to the provisions of Article 8.2, the Licensee shall be entitled to claim compensation for any direct loss it may have suffered from the Software as a result of a fault on the part of the relevant Licensor, subject to providing evidence thereof.

8.2 The Licensor's liability is limited to the commitments made under this Agreement and shall not be incurred as a result of in particular: (i) loss due the Licensee's total or partial failure to fulfill its obligations, (ii) direct or consequential loss that is suffered by the Licensee due to the use or performance of the Software, and (iii) more generally, any consequential loss. In particular the Parties expressly agree that any or all pecuniary or business loss (i.e. loss of data, loss of profits, operating loss, loss of customers or orders, opportunity cost, any disturbance to business activities) or any or all legal proceedings instituted against the Licensee by a third party, shall constitute consequential loss and shall not provide entitlement to any or all compensation from the Licensor.

\subsection*{Article 9 - WARRANTY}

9.1 The Licensee acknowledges that the scientific and technical state-of-the-art when the Software was distributed did not enable all possible uses to be tested and verified, nor for the presence of possible defects to be detected. In this respect, the Licensee's attention has been drawn to the risks associated with loading, using, modifying and/or developing and reproducing the Software which are reserved for experienced users.

The Licensee shall be responsible for verifying, by any or all means, the suitability of the product for its requirements, its good working order, and for ensuring that it shall not cause damage to either persons or properties.

9.2 The Licensor hereby represents, in good faith, that it is entitled to grant all the rights over the Software (including in particular the rights set forth in Article 5).

9.3 The Licensee acknowledges that the Software is supplied "as is" by the Licensor without any other express or tacit warranty, other than that provided for in Article 9.2 and, in particular, without any warranty as to its commercial value, its secured, safe, innovative or relevant nature.

Specifically, the Licensor does not warrant that the Software is free from any error, that it will operate without interruption, that it will be compatible with the Licensee's own equipment and software configuration, nor that it will meet the Licensee's requirements.

9.4 The Licensor does not either expressly or tacitly warrant that the Software does not infringe any third party intellectual property right relating to a patent, software or any other property right. Therefore, the Licensor disclaims any and all liability towards the Licensee arising out of any or all proceedings for infringement that may be instituted in respect of the use, modification and redistribution of the Software. Nevertheless, should such proceedings be instituted against the Licensee, the Licensor shall provide it with technical and legal assistance for its defense. Such technical and legal assistance shall be decided on a case-by-case basis between the relevant Licensor and the Licensee pursuant to a memorandum of understanding. The Licensor disclaims any and all liability as regards the Licensee's use of the name of the Software. No warranty is given as regards the existence of prior rights over the name of the Software or as regards the existence of a trademark.

\subsection*{Article 10 - TERMINATION}

10.1 In the event of a breach by the Licensee of its obligations hereunder, the Licensor may automatically terminate this Agreement thirty (30) days after notice has been sent to the Licensee and has remained ineffective.

10.2 A Licensee whose Agreement is terminated shall no longer be authorized to use, modify or distribute the Software. However, any licenses that it may have granted prior to termination of the Agreement shall remain valid subject to their having been granted in compliance with the terms and conditions hereof.

\subsection*{Article 11 - MISCELLANEOUS}

11.1 EXCUSABLE EVENTS

Neither Party shall be liable for any or all delay, or failure to perform the Agreement, that may be attributable to an event of force majeure, an act of God or an outside cause, such as defective functioning or interruptions of the electricity or telecommunications networks, network paralysis following a virus attack, intervention by government authorities, natural disasters, water damage, earthquakes, fire, explosions, strikes and labor unrest, war, etc.

11.2 Any failure by either Party, on one or more occasions, to invoke one or more of the provisions hereof, shall under no circumstances be interpreted as being a waiver by the interested Party of its right to invoke said provision(s) subsequently.

11.3 The Agreement cancels and replaces any or all previous agreements, whether written or oral, between the Parties and having the same purpose, and constitutes the entirety of the agreement between said Parties concerning said purpose. No supplement or modification to the terms and conditions hereof shall be effective as between the Parties unless it is made in writing and signed by their duly authorized representatives.

11.4 In the event that one or more of the provisions hereof were to conflict with a current or future applicable act or legislative text, said act or legislative text shall prevail, and the Parties shall make the necessary amendments so as to comply with said act or legislative text. All other provisions shall remain effective. Similarly, invalidity of a provision of the Agreement, for any reason whatsoever, shall not cause the Agreement as a whole to be invalid.

11.5 LANGUAGE

The Agreement is drafted in both French and English and both versions are deemed authentic.

\subsection*{Article 12 - NEW VERSIONS OF THE AGREEMENT}

12.1 Any person is authorized to duplicate and distribute copies of this Agreement.

12.2 So as to ensure coherence, the wording of this Agreement is protected and may only be modified by the authors of the License, who reserve the right to periodically publish updates or new versions of the Agreement, each with a separate number. These subsequent versions may address new issues encountered by Free Software.

12.3 Any Software distributed under a given version of the Agreement may only be subsequently distributed under the same version of the Agreement or a subsequent version, subject to the provisions of Article 5.3.4.

\subsection*{Article 13 - GOVERNING LAW AND JURISDICTION}

13.1 The Agreement is governed by French law. The Parties agree to endeavor to seek an amicable solution to any disagreements or disputes that may arise during the performance of the Agreement.

13.2 Failing an amicable solution within two (2) months as from their occurrence, and unless emergency proceedings are necessary, the disagreements or disputes shall be referred to the Paris Courts having jurisdiction, by the more diligent Party.

### 2014-08-06 ADJOINT_TOMOGRAPHY_TOOLS/ADJOINT_TOMOGRAPHY_TOOLKIT/HowtoUseToolkit.tex

\documentclass[11pt]{article}
\usepackage{geometry}                % See geometry.pdf to learn the layout options. There are lots.
\geometry{letterpaper}                   % ... or a4paper or a5paper or ...
%\geometry{landscape}                % Activate for for rotated page geometry
%\usepackage[parfill]{parskip}    % Activate to begin paragraphs with an empty line rather than an indent
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{epstopdf}
\DeclareGraphicsRule{.tif}{png}{.png}{`convert #1 `dirname #1`/`basename #1 .tif`.png}

\title{How to use the tomography toolkit of SPECFEM3D}
\author{The Author}
%\date{}                                           % Activate to display a given date or no date

\begin{document}
\maketitle
%\section{}
%\subsection{}

\section{Future plan}
\begin{enumerate}
\item check in the latest version of measurement code, add a flag to decide whether new or old strategy you use
\item implement full attenuation for the reconstructed original wavefields
\item export models or gradients on a regular grid instead of a unstructured mesh
\item add Qinya's GUI for flexwin window selection part
\item add Qinya's CMT source inversion package to the SOURCE\_INVERSION directory
\item start to prepare a document for this toolkit
\item start to think about one example for the inversion (may be a subset of dataset)
\item start to use routine to interpolate back and forth between the SEM mesh and a regular mesh
\item start to incorporate noise tomography in the toolkit
\end{enumerate}

\section{Modifications of measurement code}
We changed the scheme about how to generate adjoint sources.
Right now, we use synthetic seismograms with physical dispersion to construct adjoint source for banana-dougnut kernels.
Frequency-dependent phase and amplitude anomalies are measured based on data and synthetics with full attenuation.
Thus, we changed the measurement code to read in three traces: data, synthetics with full attenuation and synthetics with physical dispersion.
One more forward simulation is required for each source.

\begin{enumerate}
\item adjoint sources for BD kernels: data + synthetics with physical dispersion
\item measurements: data + synthetics with full attenuation.
\end{enumerate}

1D test to take care of full attenuation is being designed.

\section{Source correction for origin time and scalar moment}
We use a 2D grid search to make correction for origin time and scalar moment.
Basic idea is that we perturbate synthetic seismograms by using origin time shift $\delta t_0$ and scalar moment change $\delta M_0$.
Then calculate misfit functions (cross-correlation phase and amplitude anomalies between data and synthetics) for each pair of ($\delta t_0$, $\delta M_0$) in a 2D searching domain, select the one with minimal misfit value. Then use this origin time and scalar moment to correct
your synthetic seismograms.

\begin{enumerate}
\item 2D grid search for origin time and scalar moment
\item correct your synthetic seismograms using new origin time and scalar moment
\item reprocessing your data and synthetic seismograms by using new source files.
\end{enumerate}

\section{Basic structure of adjoint toolkit}

\subsection{FORWARD\_ADJOINT}
This directory is used to launch forward and adjoint simulations, and process data, it includes:
\begin{enumerate}
\item XSRC\_SEM, specfem package for forward and adjoint simulations
\item XSRC\_FLEXWIN, flexwin for window selection
\item XSRC\_MEASURE\_ADJ, generate adjoint sources and make measurements
\item PERL\_CENTER, processing scripts for data and synthetics
\item xsubmit\_forward\_simulation.sh, a script used to submit forward simulations
\item xsubmit\_adjoint\_simulation.sh, a script used to submit adjoint simulations
\item XSHELL\_process\_dat.sh, a script used to process data and synthetics
\end{enumerate}

\subsection{ITERATION\_UPDATE}
This directory is used to perform gradient summation, precondition, smoothing and update models, it includes:
\begin{enumerate}
\item X01\_SRC\_SUM\_KERNELS, sum event kernels
\item X02\_SRC\_PRECOND\_KERNELS, precondition summed kernels
\item X03\_SRC\_SMOOTH\_KERNELS, smooth kernels
\item X04a\_SRC\_DIRECTION\_CG, calculate search directions using conjugate gradient method; DK DK comment from Dimitri Komatitsch: why not use BiCGStab here
instead of a classical CG; usually it is much better than simple CG; however it seems that Yang Luo tried BiCGStab already (?) and found it less efficient than L-BFGS anyway
\item X04b\_SRC\_DIRECTION\_SD, calculate search directions using steepest descent method
\item X04c\_SRC\_DIRECTION\_LBFGS, calculate search directions using L-BFGS method
\item X05\_SRC\_UPDATE\_MODELS, update model parameters based on search directions
\end{enumerate}

In Step 4 above, one should either 04a, 04b or 04c; in most cases, the best choice is 04c, which approximates the Hessian and thus 
converges much faster.

From Wikipedia http://en.wikipedia.org/wiki/Limited-memory\_BFGS, here is a brief description of L-BFGS:

The ''limited-memory BFGS'' (''L-BFGS'' or ''LM-BFGS'') algorithm is a member of the broad family of quasi-Newton optimization methods that uses a limited memory variation of the Broyden-Fletcher-Goldfarb-Shanno (BFGS) update to approximate the inverse Hessian matrix (denoted by $H_k$). Unlike the original BFGS method which stores a dense $ n \times n $ approximation, L-BFGS stores only a few vectors that represent the approximation implicitly. Due to its moderate memory requirement, the L-BFGS method is particularly well suited for optimization problems with a large number of variables.  L-BFGS never explicitly forms or stores $H_k$.  Instead, it maintains a history of the past $m\,\!$ updates of the position  $x\,\!$ and gradient  $\nabla f(x)$, where generally the history  $m\,\!$ can be short, often less than 10.  These updates are used to implicitly do operations requiring the $H_k$-vector product.  While strictly, a straightforward BFGS implementation at the  $i\,\!$-th iteration would represent the inverse Hessian approximation as informed by all updates on  $0 \ldots i-1$, L-BFGS does quite well using updates from only the most recent iterations  $i-m \ldots i-1$.

\subsection{MODEL\_VISUALIZATION}
This directory is used to extract xyz files from your models or gradients, and plot them based on GMT.
\begin{enumerate}
\item SRC\_GEN\_XYZ\_HORIZ\_VERT, matlab scripts used to generate xyz files (matlab scripts will be transfered to fortran code soon)
\item SRC\_MODEL\_SLICE\_HORIZ, subroutines used to extract xyz files from models for horizontal cross sections
\item SRC\_MODEL\_SLICE\_VERT, subroutines used to extract xyz files from modes for vertical cross sections
\end{enumerate}

\subsection{SHARE\_FILES}
This directory contains several common files which are used by other directories, such as FORWARD\_ADJOINT, ITERATION\_UPDATE.
\begin{enumerate}
\item CMTSOLUTION\_CENTER, contains all CMTSOLUTIONs
\item EVENTID\_CENTER, contains eventid used to submit jobs
\item HEADER\_FILES, contains header files generated by specfem package
\item STATIONS\_CENTER, contains all station files
\end{enumerate}

\subsection{SOURCE\_INVERSION}
This directory is used to perform source correction for origin time and scalar moment.
\begin{enumerate}
\item SRC\_GRIDSEARCH\_INITIALTIME\_MOMENT, perform 2D grid search for origin time and scalar moment
\item SRC\_CORRECT\_INITIALTIME\_MOMENT, correct synthetics based on grid search
\end{enumerate}

\subsection{XUTIL}
This directory contains useful scripts, such as checking scripts to check whether your jobs are finished or not.

\section{Other issues}
\subsection{name convention for data and synthetic seismograms}
At current stage, we rename all data and synthetics by using convention such as Network.Station.Comp.sac, which are easy to manipulate.

\subsection{name convention for scripts}
We use XSHELL\_**.sh as a main shell script which will submit XPBS\_**.sh on the cluster, such as in FORWARD\_ADJOINT, we use XSHELL\_process\_data.sh to submit XPBS\_process\_dat.sh iteratively.

\section{Suggestion by Dimitri Komatitsch, November 2012}

Below is some feedback (in no particular order) from us at CNRS:

- move the whole 'ADJOINT\_TOMO' set of tools to the SVN trunk of SPECFEM3D, and maintain it there instead of in an independent directory (CIG can do that for you using 'svn move', I guess you need root access to do that from one level above the different code directories); that point is very important, since having everything in the same SVN branch will be far more flexible; also considering that in 2013 or 2014 we will probably get rid of the GLOBE solver and permanently merge GLOBE into Cartesian (i.e. keep the GLOBE mesher only and just make its output compatible with the Cartesian solver); thus by then we would have everything in the same directory

- about Section 1: clarify how you generate "synthetics with physical dispersion but without full attenuation; One more forward run is required for each source" because currently Par\_file contains an ATTENUATION flag that can be on or off, but on means full attenuation;
thus not clear how the third option is implemented

- change the title of the document to "How to use the tomography toolkit of SPECFEM3D" or similar

- Section 2: type the comments made by Carl, Hejun and Jeroen about why source correction is that critical (there was a discussion about this yesterday, let us thus summarize it there in a few sentences);
also clarify what you mean by "3/ reprocessing your data and synthetic seismograms is necessary": does one only need to update the SAC headers? if so, how? (is there a script for that)

- section 3.2: Rename items 4, 5 and 6 to 4a, 4b and 4c and clarify that a single option should be selected; mentioned that Jeroen said that ultimately using L-BFGS is better; briefly explain why (i.e., briefly summarize the drawbacks of the first two options)

- clarify somewhere if and when/why a pseudo-Hessian calculation could or should be done; explain how to use the internal flag "APPROXIMATE\_HESS\_KL" of the code

- mention somewhere that Vadim and Sébastien Chevrot will commit their routines to interpolate back and forth between the SEM mesh and a regular mesh; mention that we should talk to Qinya about this, since she was in the process of implementing that as well (I think that Vadim and Sébastien's routine will be sufficient, but we should doublecheck with Qinya if she had something different (more powerful? or faster?) in mind

- what about boundary kernels, as developed by Qinya a few years ago?
we forgot to talk about this yesterday, and they are not mentioned in the document; I guess they are inactive by default, but can one use them, and if so how? and are there scripts to use for that?
(if not, i.e. if Qinya has some local scripts that have not been committed to SVN, we should probably commit them, even if they are not very flexible yet; scripts and code that are kept outside of SVN tend to quickly vanish ; if some scripts from Qinya are available, can somebody add a few lines in the document about how to use them?

- mention that Piero will update the document to explain how to use the toolkit for noise tomography

- In Section 1, mention that Zhinan and I will do some 1D tests in January (using SPECFEM1D) to try to undo full attenuation in the 1D case; in the case of the structural dynamics equation Zhinan managed to do it in a stable way (i.e. he managed to undo the -C*v damping term in a stable way; and that term is not that different from memory variables with a constant Q; thus there is hope); mention that this is already in the to-do list, but less urgent/important than the rest I guess

- In Section 3.3, mention that I think that it would be better to convert the Matlab scripts back to Fortran, or else use Python or similar; this way we avoid depending on a commercial package. That is probably not critical (at least for our groups), but at the same time if these scripts are simple and easy to export to Fortran, there is no reason not to do it (and the sooner the better). Mention maybe that if possible we should try to keep the whole toolkit independent of commercial packages, so that all users worldwide can use it easily

- in Section 4.1, I think your convention for seismogram names : Network.Station is better than the old convention, which was Station.Network. This way it is easier to handle all the stations from a given network. Please just confirm that you swapped that purposely, i.e. that it is not a typo (if not, I would suggest we switch anyway, since that is a very good idea)
Below is some feedback (in no particular order) from us at CNRS, which you can cut and paste in your file before uploading it to SVN.

- move the whole 'ADJOINT\_TOMO' set of tools to the SVN trunk of SPECFEM3D, and maintain it there instead of in an independent directory (CIG can do that for you using 'svn move', I guess you need root access to do that from one level above the different code directories); that point is very important, since having everything in the same SVN branch will be far more flexible; also considering that in 2013 or 2014 we will probably get rid of the GLOBE solver and permanently merge GLOBE into Cartesian (i.e. keep the GLOBE mesher only and just make its output compatible with the Cartesian solver); thus by then we would have everything in the same directory

- about Section 1: clarify how you generate "synthetics with physical dispersion but without full attenuation; One more forward run is required for each source" because currently Par\_file contains an ATTENUATION flag that can be on or off, but on means full attenuation;
thus not clear how the third option is implemented

- change the title of the document to "How to use the tomography toolkit of SPECFEM3D" or similar

- Section 2: type the comments made by Carl, Hejun and Jeroen about why source correction is that critical (there was a discussion about this yesterday, let us thus summarize it there in a few sentences);
also clarify what you mean by "3/ reprocessing your data and synthetic seismograms is necessary": does one only need to update the SAC headers? if so, how? (is there a script for that)

- section 3.2: Rename items 4, 5 and 6 to 4a, 4b and 4c and clarify that a single option should be selected; mentioned that Jeroen said that ultimately using L-BFGS is better; briefly explain why (i.e., briefly summarize the drawbacks of the first two options)

- clarify somewhere if and when/why a pseudo-Hessian calculation could or should be done; explain how to use the internal flag "APPROXIMATE\_HESS\_KL" of the code

- mention somewhere that Vadim and Sébastien Chevrot will commit their routines to interpolate back and forth between the SEM mesh and a regular mesh; mention that we should talk to Qinya about this, since she was in the process of implementing that as well (I think that Vadim and Sébastien's routine will be sufficient, but we should doublecheck with Qinya if she had something different (more powerful? or faster?) in mind

- what about boundary kernels, as developed by Qinya a few years ago?
we forgot to talk about this yesterday, and they are not mentioned in the document; I guess they are inactive by default, but can one use them, and if so how? and are there scripts to use for that?
(if not, i.e. if Qinya has some local scripts that have not been committed to SVN, we should probably commit them, even if they are not very flexible yet; scripts and code that are kept outside of SVN tend to quickly vanish ; if some scripts from Qinya are available, can somebody add a few lines in the document about how to use them?

- mention that Piero will update the document to explain how to use the toolkit for noise tomography

- In Section 1, mention that Zhinan and I will do some 1D tests in January (using SPECFEM1D) to try to undo full attenuation in the 1D case; in the case of the structural dynamics equation Zhinan managed to do it in a stable way (i.e. he managed to undo the -C*v damping term in a stable way; and that term is not that different from memory variables with a constant Q; thus there is hope); mention that this is already in the to-do list, but less urgent/important than the rest I guess

- In Section 3.3, mention that I think that it would be better to convert the Matlab scripts back to Fortran, or else use Python or similar; this way we avoid depending on a commercial package. That is probably not critical (at least for our groups), but at the same time if these scripts are simple and easy to export to Fortran, there is no reason not to do it (and the sooner the better). Mention maybe that if possible we should try to keep the whole toolkit independent of commercial packages, so that all users worldwide can use it easily

- in Section 4.1, I think your convention for seismogram names : Network.Station is better than the old convention, which was Station.Network. This way it is easier to handle all the stations from a given network. Please just confirm that you swapped that purposely, i.e. that it is not a typo (if not, I would suggest we switch anyway, since that is a very good idea)

\end{document}

### 2014-08-06 ADJOINT_TOMOGRAPHY_TOOLS/flexwin/USER_MANUAL/manual_introduction.tex

\chapter{Introduction}

The FLEXWIN software package automates the time-window selection problem for
seismologists.  It operates on pairs of observed and synthetic single component
seismograms, defining windows that cover as much of a given seismogram as
possible, while avoiding portions of the waveform that are dominated by noise.

FLEXWIN selects time windows on the synthetic seismogram within which the waveform
contains a distinct energy arrival, then requires an adequate correspondence
between observed and synthetic waveforms within these windows.

There is no restriction on the type of simulation used to
generate the synthetics. Realistic Earth models and more complete wave
propagation theories yield waveforms that are more similar to the observed
seismograms, and thereby promote the selection of measurement windows covering
more of the available data.  The input seismograms can be measures of
displacement, velocity or acceleration.  There is no requirement
for horizontal signals to be rotated into radial and transverse directions.

FLEXWIN is a configurable data selection process that can be adapted to
different tomographic scenarios by tuning a handful of parameters.  Although
the algorithm was designed for use in 3D-3D adjoint tomography, its inherent
flexibility should make it useful in many data-selection applications.

For a detailed introduction to FLEXWIN as applied to seismic tomography, please consult \cite{MaggiEtal2009}.  If you use FLEXWIN for your own research, please cite \cite{MaggiEtal2009}.

### 2014-08-12 doc/USER_MANUAL/manual_SPECFEM3D_Cartesian.tex
switched to the new version of src/shared/utm_geo.f90 in order to have support for the Southern hemisphere. Also switched to the WGS84 ellipsoid instead of Clarke 1866. Updated the users manual accordingly.

@@
  \texttt{PROJECTION} is \texttt{.false.}.
+Use a negative zone number for the Southern hemisphere:
+the Northern hemisphere corresponds to zones +1 to +60,
+the Southern hemisphere to zones -1 to -60.
+We use the WGS84 (World Geodetic System 1984) reference ellipsoid for the UTM projection. If you prefer to use the Clarke 1866 ellipsoid,
+edit file \texttt{src/shared/utm\_geo.f90}, uncomment that ellipsoid and recompile the code.
+From http://en.wikipedia.org/wiki/Universal\_Transverse\_Mercator\_coordinate\_system :
+The Universal Transverse Mercator coordinate system was developed by the United States Army Corps of Engineers in the 1940s.
+The system was based on an ellipsoidal model of Earth. For areas within the contiguous United States
+the Clarke Ellipsoid of 1866 was used. For the remaining areas of Earth, including Hawaii, the International Ellipsoid was used.
+The WGS84 ellipsoid is now generally used to model the Earth in the UTM coordinate system,
+which means that current UTM northing at a given point can be 200+ meters different from the old one.
+For different geographic regions, other datum systems (e.g.: ED50, NAD83) can be used.
 \item [{\texttt{SUPPRESS\_UTM\_PROJECTION}}] set to be \texttt{.false.}

### 2014-08-16 doc/USER_MANUAL/manual_SPECFEM3D_Cartesian.tex
added a warning about versions dot zero of a compiler

@@
 \begin{description}
-\item [{\texttt{Intel ifort compiler}}] See if you need to add \texttt{-assume
-byterecl} for your machine.
+\item [{\texttt{Intel ifort compiler}}] See if you need to add \texttt{-assume byterecl} for your machine.
+\textbf{In the case of that compiler, we have noticed that versions dot zero sometimes have bugs or issues that can lead to wrong results when running the code, thus we \emph{strongly} recommend using versions dot one or above (for instance version 13.1 instead of 13.0, version 14.1 instead of 14.0 and so on)}.
 \item [{\texttt{IBM compiler}}] See if you need to add \texttt{-qsave}
@@
 the \texttt{SCOTCH} library that we use, does not exist in Mac OS.
-You will need to replace it with \texttt{clock\_get\_time} if you
+You may need to replace it with \texttt{clock\_get\_time} if you
 want to use \texttt{SCOTCH}.

### 2014-09-16 doc/USER_MANUAL/manual_SPECFEM3D_Cartesian.tex
added doc/SEM_class_from_Yvon_Maday_to_Dimitri_Komatitsch_and_Jean_Pierre_Vilotte_04apr1995.pdf

@@
 and Jeroen Tromp at Harvard University and Caltech, USA, starting in 1998.
-The story started on March 28, 1995, when Prof. Yvon Maday from CNRS and University of Paris, France, gave a lecture to
+The story started on April 4, 1995, when Prof. Yvon Maday from CNRS and University of Paris, France, gave a lecture to
 Dimitri Komatitsch and Jean-Pierre Vilotte at IPG about the nice properties of the Legendre spectral-element method with diagonal mass matrix that he had used for
@@
 directory \texttt{bin/}. To output the usage of this executable, type
-`./bin/xcombine\_vol\_data` without arguments. As an example, to run
-the executable you would use
-\begin{lyxcode}
-{\footnotesize cd~bin/~}{\footnotesize \par}
-
-~{\footnotesize ./xcombine\_vol\_data~0~3~velocity\_Z\_it000400~../OUTPUT\_FILES/DATABASES\_MPI~../OUTPUT\_FILES~0~}{\footnotesize \par}
-\end{lyxcode}
-to create a low-resolution mesh file. The output mesh file will have
-the name \texttt{velocity\_Z\_it000400.mesh}. We next convert the
-\texttt{.mesh} file into the VTU (Unstructured grid file) format which
-can be viewed in ParaView. For this task, you can use and modify the
-script \texttt{mesh2vtu.pl} located in directory \texttt{utils/Visualization/Paraview/},
-for example:
+\begin{lyxcode}
+{\footnotesize ./bin/xcombine\_vol\_data~}{\footnotesize \par}
+\end{lyxcode}
+without arguments.
+
+\medskip
+
+\texttt{xcombine\_vol\_data} will combine the data on the different processors (located in \texttt{OUTPUT\_FILES/DATABASES\_MPI}), for a given quantity and a given iteration, to one file.
+For example, if you want to combine \texttt{velocity\_Z}, for the iteration 400 on 4 processors, i.e.
+if you want to combine these files :
+
+\medskip
+
+\texttt{proc000000\_velocity\_Z\_it000400.bin }
+
+\texttt{proc000001\_velocity\_Z\_it000400.bin }
+
+\texttt{proc000002\_velocity\_Z\_it000400.bin }
+
+\texttt{proc000003\_velocity\_Z\_it000400.bin}
+
+\medskip
+
+You have to go in the directory of the concerned example (where are the directories \texttt{DATA}, \texttt{OUTPUT\_FILES}, etc.).
+Then, you can launch \texttt{xcombine\_vol\_data}, specifying the path where it is located.
+As an example, if the \texttt{DATA} and \texttt{OUTPUT\_FILES} directories of the concerned example are in the root specfem3d directory, \texttt{xcombine\_vol\_data} is in \texttt{./bin}, and you have to type:
+
+\begin{lyxcode}
+~{\footnotesize ./bin/xcombine\_vol\_data~0~3~velocity\_Z\_it000400~./OUTPUT\_FILES/DATABASES\_MPI~./OUTPUT\_FILES~0~}{\footnotesize \par}
+\end{lyxcode}
+
+Here, $0$ is the number of the first processor, $3$ the number of the last one, \texttt{velocity\_Z\_it000400} the name of the files we want to combine without the prefix \texttt{proc000000*\_}, \texttt{./OUTPUT\_FILES/DATABASES\_MPI} the directory where the files are located, \texttt{./OUTPUT\_FILES} the directory where the combined file will be stored, and $0$ is the parameter to create a low-resolution mesh file ($1$ for a high-resolution).
+
+\medskip
+
+At the beginning of the source file \texttt{combine\_vol\_data.f90} (located in \texttt{src/auxiliaries/}), there is a logical constant parameter \texttt{USE\_VTK\_OUTPUT} whose default value is \texttt{.true.}, that which will provide the output combined file directly in the VTK format and you can use Paraview to visualize it.
+
+\medskip
+
+If the value of this parameter is \texttt{.false.}, the output mesh file will have the name \texttt{velocity\_Z\_it000400.mesh}. Then we next have to convert the
+\texttt{.mesh} file into the VTU (Unstructured grid file) format which can be viewed in ParaView. For this task, you can use and modify the
+script \texttt{mesh2vtu.pl} located in directory \texttt{utils/Visualization/Paraview/}, for example:
 \begin{lyxcode}
@@
 \end{lyxcode}
-Notice that this Perl script uses a program \texttt{mesh2vtu} in the
-\texttt{utils/Visualization/Paraview/mesh2vtu} directory, which further
-uses the VTK \urlwithparentheses{http://www.vtk.org/} run-time library
-for its execution. Therefore, make sure you have them properly set
+Notice that this Perl script uses a program \texttt{mesh2vtu} in the \texttt{utils/Visualization/Paraview/mesh2vtu} directory, which further
+uses the VTK \urlwithparentheses{http://www.vtk.org/} run-time library for its execution. Therefore, make sure you have them properly set
 in the script according to your system.
+
+\bigskip
+
+Then, to do a movie with several iterations, you have to repeat this process for each iteration you want to put in your movie.
 

### 2014-10-03 doc/USER_MANUAL/manual_SPECFEM3D_Cartesian.tex
replaced latitude with latorUTM and longitude with longorUTM in CMTSOLUTION and FORCESOLUTION (as mentioned by Alexis Bottero the original names did not make sense in the cause of SUPPRESS_UTM_PROJECTION = .true. because the values are then given in UTM rather than longitude and latitude). The code still accepts the old format for back-compatibility reasons. That is also why I kept the same length as before for the new names.
Adds note on MPI-IO

@@
 \end{lyxcode}
+Note that MPI should have been installed with MPI-IO enabled.
+
 Before running the \texttt{configure} script, you should probably
@@
 \begin{itemize}
+\item Set the latitude or UTM $x$ coordinate, longitude or UTM $y$ coordinate, depth of the source (in km).
 \item Set the \texttt{time shift} parameter equal to $0.0$ (the solver
@@
 (hdur = 5 {*} DT) is automatically set by default.
-\item Set the latitude, longitude, depth of the source in geographical coordinates.
+\item Set the latitude or UTM $x$ coordinate, longitude or UTM $y$ coordinate, depth of the source (in km).
 \item Set the magnitude of the force source.

### 2014-10-03 src/specfem3D/initialize_simulation.f90
merged Clement Durochat's modifications to merge Vadim Monteiller's mesher for a chunk of the Earth (for coupling with DSM)

@@ -52,8 +52,8 @@
                         NTSTEP_BETWEEN_READ_ADJSRC,NOISE_TOMOGRAPHY, &
                         USE_FORCE_POINT_SOURCE,STACEY_INSTEAD_OF_FREE_SURFACE, &
                         USE_RICKER_TIME_FUNCTION,OLSEN_ATTENUATION_RATIO,PML_CONDITIONS, &
-                        PML_INSTEAD_OF_FREE_SURFACE,f0_FOR_PML,IMODEL, SEP_MODEL_DIRECTORY, &
-                        FULL_ATTENUATION_SOLID,TRACTION_PATH,COUPLE_WITH_DSM)
+                        PML_INSTEAD_OF_FREE_SURFACE,f0_FOR_PML,IMODEL,SEP_MODEL_DIRECTORY, &
+                        FULL_ATTENUATION_SOLID,TRACTION_PATH,COUPLE_WITH_DSM,MESH_A_CHUNK_OF_THE_EARTH)
 
   call read_adios_parameters(ADIOS_ENABLED, ADIOS_FOR_DATABASES,       &
                              ADIOS_FOR_MESH, ADIOS_FOR_FORWARD_ARRAYS, &

### 2014-10-07 doc/USER_MANUAL/manual_SPECFEM3D_Cartesian.tex
merged Clement Durochat's modifications to merge Vadim Monteiller's mesher for a chunk of the Earth (for coupling with DSM)
Adds a section about SEP models in the manual

@@
 \end{lyxcode}
-Note that MPI should have been installed with MPI-IO enabled.
+Note that MPI must be installed with MPI-IO enabled because parts of SPECFEM3D perform I/Os through MPI-IO.
 
@@
 
+\section{{\normalsize \label{sec:Using-SEP}}Using external SEP models.}
+
+SEP models consists on two files per variables, one header file (\texttt{.H}) 
+and one binary file (\texttt{.H$@$}). 
+Among other information, the header file indicates, for each dimension, 
+the offset (\texttt{o1}, \texttt{o2} and \texttt{o3} fields), 
+the spatial step size (\texttt{d1}, \texttt{d2} and \texttt{d3} fields)
+and the number of samples (\texttt{n1}, \texttt{n2} and \texttt{n3} necessary to read
+the linked binary file (\texttt{in} field).
+
+SEP model reading occures during database generation. It it the user responsability
+to ensure that proper interfaces between acoustic and elastic domains are set up
+during the meshing phase. For more information, refer to 
+section~\ref{cha:Running-the-Mesher-Meshfem3D}.
+
+In \texttt{DATA/Par\_file}, \texttt{\textbf{MODEL}} should be set to \texttt{\textbf{sep}},
+and \texttt{\textbf{SEP\_MODEL\_DIRECTORY}} should point to wherever your sep model files reside.
+
+Note that in order not to surcharge the parameter file, SEP header files
+should be named \texttt{vp.H}, \texttt{vs.H} and \texttt{rho.H}. The \texttt{in} field
+in these binary files can be any path as long as it is relative to
+the previously set \texttt{SEP\_MODEL\_DIRECTORY}.
+We also assume that any dimensional value inside SEP header files is given in meters and
+that binary files are single precision floating point numbers stored in small endian format.
+There is currently only support for \texttt{vp}, \texttt{vs} and \texttt{rho} variables.
+
+An example on how to use SEP models is given in:
+\texttt{EXAMPLES/meshfem3D\_examples/sep\_bathymetry}.
+
 \chapter{Post-Processing Scripts}

### 2014-10-08 doc/USER_MANUAL/manual_SPECFEM3D_Cartesian.tex

--- 2014-10-03 doc/USER_MANUAL/manual_SPECFEM3D_Cartesian.tex
@@
 
+\section{{\normalsize \label{sec:Using-SEP}}Using external SEP models.}
+
+Stanford Exploration Project (SEP) models consists of two files per variables, one header file (\texttt{.H}) 
+and one binary file (\texttt{.H$@$}). 
+Among other information, the header file indicates, for each dimension, 
+the offset (\texttt{o1}, \texttt{o2} and \texttt{o3} fields), 
+the spatial step size (\texttt{d1}, \texttt{d2} and \texttt{d3} fields)
+and the number of samples (\texttt{n1}, \texttt{n2} and \texttt{n3}) necessary to read
+the linked binary file (\texttt{in} field).
+
+SEP model reading occurs during database generation. It it the user responsibility
+to ensure that proper interfaces between acoustic and elastic domains are set up
+during the meshing phase. For more information, refer to 
+section~\ref{cha:Running-the-Mesher-Meshfem3D}.
+
+In \texttt{DATA/Par\_file}, \texttt{\textbf{MODEL}} should be set to \texttt{\textbf{sep}},
+and \texttt{\textbf{SEP\_MODEL\_DIRECTORY}} should point to wherever your sep model files reside.
+
+Note that in order not to overload the parameter file, SEP header files
+should be named \texttt{vp.H}, \texttt{vs.H} and \texttt{rho.H}. The \texttt{in} field
+in these binary files can be any path as long as it is relative to
+the previously set \texttt{SEP\_MODEL\_DIRECTORY}.
+We also assume that any dimensional value inside SEP header files is given in meters and
+that binary files are single precision floating point numbers stored in little endian format.
+There is currently only support for \texttt{vp}, \texttt{vs} and \texttt{rho} variables.
+
+An example demonstrating the use SEP models is given in:
+\texttt{EXAMPLES/meshfem3D\_examples/sep\_bathymetry}.
+
 \chapter{Post-Processing Scripts}

### 2014-11-11 ADJOINT_TOMOGRAPHY_TOOLS/flexwin/USER_MANUAL/manual_technical.tex

\chapter{Getting started\label{sec:install}}

Here is where you find basic information for obtaining and installing the FLEXWIN package.
For details of how to tune the algorithm to your seismograms, see chapter~\ref{sec:tuning}.

\section{System requirements}
In order to install and run, FLEXWIN requires:
\begin{itemize}
\item UNIX operating system (Linux, Solaris, MacOS \ldots)
\item GNU make
\item a fortran compiler (gfortran, ifort, etc...) 
\item other packages : SAC (Seismic Analysis Code, available from IRIS); GMT (Generic Mapping Tools) for the plotting scripts
\end{itemize}

FLEXWIN requires the following libraries external to the package in order to
compile and run: {\tt libsacio.a} and {\tt libsac.a}. Both libraries
are distributed by IRIS as part of the SAC package (version 101.2 and above).
The IRIS download site (as of 30-March-2009) is here:
\url{http://www.iris.edu/software/sac/sac.request.htm}.
(To check your version, type sac.)

\section{Obtaining the code}

The code is available as a gzipped tarball from CIG (Computational Infrastructure for Geodynamics, \url{http://www.geodynamics.org}). The tarball is unpacked by typing {\tt tar xvzf flexwin.tgz}.

The package contains the flexwin code and documentation, as well as a set of
test data, examples of user files for different scenarios, and a set of utility
scripts that may be useful for running flexwin on large datasets.

\section{Compilation}

If your compiler of choice is gfortran, then you should be able to use the
{\tt make\_gfortran} makefiles with only minor modifications (notably you may need to
change the search path for the {\tt libsacio.a} library).  If you prefer another
compiler, you should modify the OPT and FC lines in the makefiles accordingly. We tested the code using gfortran version 4.1.2
(To check your version, type {\tt gfortran --version}.)

{\bf Important note}: All the code is compiled with the -m32 option, which makes
32bit binaries.  This option is currently required to enable compatibility with
SAC.  Future versions of the SAC distribution may no longer require this
compatibility flag.

Steps to compile the flexwin package:
\begin{enumerate}
\item Compile {\tt libtau.a} and create {\tt iasp91.hed} and {\tt iasp91.tbl}.  In the {\tt flexwin/ttimes\_mod} directory type: {\tt make -f make\_gfortran}.  This will compile {\tt libtau.a}, and two programs, {\tt remodl} and {\tt setbrn}.  The makefile will also run {\tt remodl} and {\tt setbrn} to create the {\tt iasp91.hed}and {\tt iasp91.tbl} files.  You should then type {\tt make -f make\_gfortran install} to install the iasp91 files.
\item Compile flexwin.  Edit the {\tt make\_gfortran} file in the flexwin root directory to ensure the {\tt SACLIBDIR} environment variable points to the location of your SAC libraries (by default {\tt \$SACHOME/lib}).  Then type {\tt make -f make\_gfortran}.
\end{enumerate}

You should end up with the {\tt flexwin} executable.  The program requires the {\tt iasp91.hed} and {\tt iasp91.tbl} files (or symbolic links to them) to be present in the directory from which the code is launched.

\section{Running the Test case}

You should test your compiled code on the {\tt test\_data} dataset provided.  In the
{\tt flexwin/test\_data} directory, type {\tt ./flexwin < input.test}.  The results of your
run will be found in the {\tt MEASURE} subdirectory, and
should match those found in the {\tt MEASURE.orig}
subdirectory.

You can also test the basic plotting script by running {\tt ./plot\_seismos\_gmt.sh
MEASURE/ABKT.II.LHZ}, whose output will be
{\tt MEASURE/ABKT.II.LHZ.seis.eps}.
file.  Your result should be identical to what is shown in Figure~\ref{fg:test_data}.

\begin{figure}
\center \includegraphics[width=7in]{ABKT_II_LHZ_seis.pdf}
\caption{\label{fg:test_data}
Windowing results for the test data set, plotted using the {\tt ./plot\_seismos\_gmt.sh} script.
}
\end{figure}

\section{Running FLEXWIN}

In general, flexwin is run as follows: {\tt ./flexwin < input} where the {\tt input} file is formatted as follows:
{\small
\begin{verbatim}
327
RAW_DATA/9627721.CI.ADO.BHR.sac.d.fil
SYNTH/ADO.CI.BHR.new.fil
MEASURE/ADO.CI.BHR
RAW_DATA/9627721.CI.ADO.BHT.sac.d.fil
SYNTH/ADO.CI.BHT.new.fil
MEASURE/ADO.CI.BHT
RAW_DATA/9627721.CI.ADO.BHZ.sac.d.fil
SYNTH/ADO.CI.BHZ.new.fil
MEASURE/ADO.CI.BHZ
....
\end{verbatim}
}

i.e. the number of traces to be measured, followed by (in order) the path to
the raw data sac file, the path to the synthetic sac file and the path and
basename for the (many!) output files for that trace.

\section{Output files}
Most output files are in ascii.  All file names start with the basename given
in the input file for that trace:
\begin{description}
\item[basename.obs]	ascii observed seismogram (filtered) 
\item[basename.syn]	ascii synthetic seismogram (filtered)
\item[basename.obs\_lp.sac] sac observed seismogram (filtered)
\item[basename.syn\_lp.sac] sac synthetic seismogram (filtered)
\item[basename.env.obs]	ascii envelope of observed seismogram (filtered)
\item[basename.env.syn]	ascii envelope of synthetic seismogram (filtered)
\item[basename.win]		list of windows with theoretical phase arrival times
\item[basename.win.qual]	list of windows with Tshift,CC,dlnA values 
\item[basename.phases]		theoretical arrival times of phases
\item[basename.stalta]		STA:LTA timeseries used to select the windows, and
the time-dependent values of the STA:LTA water level, the cross-correlation
limit $CC_0$, the time-lag limit $\Delta\tau_0$, amplitude ratio limit
$\Delta\ln{A}_0$ and the window signal to noise limit $r_0$.
\item[basename.info]	information on the path and some statistics
\end{description}
For more details about the file formats, read the {\tt write\_}
subroutines in {\tt io\_subs.f90}.

\section{Scripts}
Several plotting routines ({\tt plot\_*.sh}) are provided in the {\tt scripts} subdirectory as examples for plotting seismograms, measurements and adjoint sources.  All plotting is
done using GMT (Generic Mapping Tools).  These scripts will need to be modified to suit your particular plotting needs.

The script {\tt extract\_event\_windowing\_stats.sh} extracts statistical
information on the window selection process, on the measurements.  Again,
you can use use this script as a template for your own information
extraction needs.

\section{Pre-processing suggestions}

Pre-processing is a subtle procedure that can affect the selection of windows.
Here we list some suggestions for pre-processing:
%
\begin{enumerate}
\item Interpolate raw data and ``raw'' synthetics using the same time-step.

\item Cut the data and synthetic seismograms based on the data record.  If the data record starts before the synthetic record (which is common for local earthquakes), then consider padding zeros before the synthetic record; information prior to the origin time (and P arrival) is useful in assessing the signal-to-noise ratio in the observed records.

\item Bandpass both the data and synthetics over the desired period range, rather than doing this within FLEXWIN, thereby limiting the number of filtering operations.  However, for initial determination of the bandpass range of interest, it is simpler to experiment using the FLEXWIN parameters {\tt WIN\_MIN\_PERIOD} and {\tt WIN\_MAX\_PERIOD}.
\end{enumerate}

%-------------------------------------------------

### 2014-11-12 utils/ADJOINT_TOMOGRAPHY_TOOLS/iterate_adj/model_plot/gmt/xc_and_seis.tex
move ADJOINT_TOMOGRAPHY_TOOLS/ to utils/ to reflect the developmental (or obsolete) state of these tools; other efforts are ongoing outside of SPECFEM3D; see also src/tomography/

\documentclass[pdf,mpa]{prosper}
\usepackage{portland}
\begin{document}
\begin{slide}{}
\vspace{-1.7cm}
\fcolorbox{white}{white}{
\hspace{-1cm} 
\begin{tabular}{c}
\includegraphics[width=4.5cm,angle=-90]{/home/carltape/ADJOINT_TOMO/ADJOINT_TOMO_OUTPUT/model_plot_gmt_OUTPUT/VERT_01/vert_01_xc_LGU_CI_206_12659440_vs_m16_m00_077.eps}
\\
\includegraphics[width=3.8cm,angle=-90]{/home/carltape/ADJOINT_TOMO/ADJOINT_TOMO_OUTPUT/misfit_plot_OUTPUT/T006_T030_win1_1D1/12659440_T006_T030_LGU_CI_m16_seis_3.eps}
\end{tabular}
}
\end{slide}
\end{document}

### 2014-11-13 utils/ADJOINT_TOMOGRAPHY_TOOLS/measure_adj/USER_MANUAL/measure_adj_manual.tex

\documentclass[11pt,titlepage,fleqn]{article}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{latexsym}
\usepackage[round]{natbib}
\usepackage{xspace}
%\usepackage{epsfig}
\usepackage{graphicx}
\usepackage{bm}
\usepackage[usenames]{color}

%\usepackage{fancybox}  # do not use when showing Table of Contents
%\usepackage{fancyhdr}
%\pagestyle{fancy}

%-----------------------------------------------------------
%       SPACING COMMANDS (Latex Companion, p. 52)
%-----------------------------------------------------------

\usepackage{setspace}

\renewcommand{\baselinestretch}{1.2}

\textwidth 460pt
\textheight 690pt
\oddsidemargin 0pt
\evensidemargin 0pt

% see Latex Companion, p. 85
\voffset     -50pt
\topmargin     0pt
\headsep      20pt
\headheight   15pt
\headheight    0pt
\footskip     30pt
\hoffset       0pt

\include{newcommands}
\newcommand{\eg}{e.g.,}
\newcommand{\refSec}[1]{section \ref{#1}}
\newcommand{\vala}{Hj\"{o}rleifsd\'{o}ttir}

\graphicspath{
  {./figures/}
}

%-----------------------------------------------------------
\begin{document}
%-----------------------------------------------------------

\begin{center}
{\bf \Large \verb+measure_adj+: A software package for making measurements and computing adjoint sources for seismic tomography}
\end{center}

\vspace{0.4cm}
\begin{spacing}{1.0}
\noindent Primary Developers: Qinya Liu, Carl Tape, Ying Zhou

\noindent Author of Manual: Carl Tape, Qinya Liu

\noindent \today

\noindent
\begin{verbatim}
SVN directory    : /seismo/3D/measure_adj/USER_MANUAL/
This file        : measure_adj_manual.pdf
Additional notes : multitaper_notes.pdf, multitaper_vala.pdf
\end{verbatim}

\tableofcontents

\end{spacing}

%-------------------------------------------------------------------

\pagebreak
\section{Introduction}

{\bf NOTE: This is a preliminary user manual with many modifications to come.}

The open-source software package \verb+measure_adj+ is designed for making measurements for seismic tomography. Several misfit functions are available, such as a simple waveform difference, a simple cross-correlation traveltime difference, or a frequency dependent multitaper traveltime difference. For each misfit function, the code also provides adjoint sources, which are needed for tomographic inversions using adjoint methods \cite[\eg][]{Tromp2005,LiuTromp2006,Tape2007}. This code was used to make measurements and adjoint sources for the (large-scale) tomographic inversion of \citet{Tape2009} for the southern California crust.

MEAS--ADJ is designed to complement FLEXWIN \citep{Maggi2009}, which is designed to pick time windows for measurement.  In particular, FLEXWIN has the option of providing a \verb+MEASUREMENT.WINDOWS+ file that can be directly read into \verb+measure_adj+.  However, FLEXWIN is not necessary to run \verb+measure_adj+.

There are two additional sets of notes on multitaper measurements and adjoint sources, \verb+multitaper_notes.pdf+ and \verb+multitaper_vala.pdf+. Some sections of \verb+multitaper_notes.pdf+ were published as ``Multitaper measurements for adjoint tomography'' in \citet[][Appendix~C]{Tape2009phd}.

Please email Carl Tape (\verb+carltape@fas.harvard.edu+) with suggestions or corrections.

%-------------------------------------------------------------------

\section{Getting started}

%-------------

\subsection{System requirements}

In order to install and run, \verb+measure_adj+ program requires:
%
\begin{itemize}
\item UNIX operating system (Linux, Solaris, MacOS \ldots)
\item GNU make
\item a fortran compiler (gfortran, ifort, etc...) 
\item other packages : SAC (Seismic Analysis Code, available from IRIS); GMT (Generic Mapping Tools) for the plotting scripts
\end{itemize}

MEAS--ADJ requires the following libraries external to the package in order to compile and run: {\tt libsacio.a} and {\tt libsac.a}. Both libraries are distributed by IRIS as part of the SAC package (version 101.2 and above). The IRIS download site (as of 21-Oct-2009) is here: \verb+http://www.iris.edu/software/sac/sac.request.htm+. (To check your version, type sac.)

%-------------

\section{Obtaining the code}

%The code is available as a gzipped tarball from CIG (Computational Infrastructure for Geodynamics, \url{http://www.geodynamics.org}). The tarball is unpacked by typing {\tt tar xvzf flexwin.tgz}.

(Use SVN for now.)

The package contains the MEAS--ADJ code and documentation, as well as a set of test data, examples of user files for different scenarios, and a set of utility scripts that may be useful for running \verb+measure_adj+ on large datasets.

%-------------

\section{Compilation}

If your compiler of choice is gfortran, then you should be able to use the \verb+Makefile+ with only minor modifications (notably you may need to change the search path for the {\tt libsacio.a} library).  If you prefer another compiler, you should modify the OPT and FC lines in the makefiles accordingly. We tested the code using gfortran version 4.3.3. (To check your version, type \verb+gfortran --version+.)

Steps to compile the MEAS--ADJ package:
%
\begin{enumerate}
\item \verb+make clean+
\item \verb+make+
\end{enumerate}

You should end up with the \verb+measure_adj+ executable.

%-------------

\subsection{Running the test case}

After compiling, execute the following two commands:
%
\begin{verbatim}
measure_adj
csh -f run_mt_measure_adj.csh
\end{verbatim}
%
If all goes well, you should generate the figure
%
\begin{verbatim}
PLOTS/9818433_T006_T030_MPM_CI_m16_mtm_win_adj.pdf
\end{verbatim}

A set of auxilliary programs/files are required for the successful run of
post-processing and plotting scripts, including
\begin{itemize}
 \item Southern California faults xy files in \verb+PLOTS/plot_win_adj.pl+:
\begin{verbatim}
   $dir0 = "/opt/seismo/datalib/SC/";
   $fault_file = "$dir0/jennings.xy";
\end{verbatim}
\item Dr. Lupei Zhu's \verb+saclst+ program 
\item Dr. Lupei Zhu's \verb+pssac2+ plotting program (\verb+pssac+ may work as well). 
\end{itemize}
%-------------------------------------------------------------------

\pagebreak
\section{Running MEAS--ADJ}

Two input files are read in when \verb+mt_measure_adj+ is executed.

%-------------

\subsection{Input file 1: {\tt MEASUREMENT.WINDOWS}}

\begin{verbatim}
3
DATA/9818433.CI.MPM.BHR.sac.d.T006_T030
SYN/MPM.CI.BHR.semd.sac.m16.T006_T030
           3
   11.2770    58.4770
   58.4770    81.3770
   81.3770   101.6770
DATA/9818433.CI.MPM.BHT.sac.d.T006_T030
SYN/MPM.CI.BHT.semd.sac.m16.T006_T030
           1
   49.3270    95.3270
DATA/9818433.CI.MPM.BHZ.sac.d.T006_T030
SYN/MPM.CI.BHZ.semd.sac.m16.T006_T030
           1
   60.9770    99.3770
\end{verbatim}
%
The first line contains the number of pairs of records to be read in.  Each pair of files is followed by the number of windows within which measurements will be made, followed by the time intervals for each window.
Note that data and synthetics require to have exactly the same \verb+t0, dt, npts+ for the code to run through. Therefore 
no interpolation of \verb+dt+ is really necessary at any point.

%-------------

\pagebreak
\subsection{Input file 2: {\tt MEASUREMENT.PAR}}

\begin{verbatim}
  -0.585 0.0110   18200  # tstart, DT, npts: time vector for simulations
                      7  # imeas (1-8; see manual)
                     BH  # channel of synthetics: BH or LH 
      30.000      6.000  # TLONG and TSHORT: band-pass periods for records
                .false.  # RUN_BANDPASS: use band-pass on records
                 .true.  # DISPLAY_DETAILS
                 .true.  # OUTPUT_MEASUREMENT_FILES
                 .true.  # COMPUTE_ADJOINT_SOURCE
     -4.5000     4.5000  # TSHIFT_MIN; TSHIFT_MAX
     -1.5000     1.5000  # DLNA_MIN; DLNA_MAX
                  0.690  # CC_MIN
                      1  # ERROR_TYPE -- 0 none; 1 CC, MT-CC; 2 MT-jack-knife (all for MTM!)
                  1.000  # DT_SIGMA_MIN: water level of traveltime uncertainty
                  0.500  # DLNA_SIGMA_MIN: water level of dlnA uncertainty
                      1  # ITAPER:1 multi-taper; 2 cosine; 3 boxcar taper type for freq-dep. meas.
            0.020  2.50  # WTR, NPI (ntaper = 2*NPI)
                  2.000  # DT_FAC (following 4 pars are for MTM --> CC check)
                  2.500  # ERR_FAC
                  3.500  # DT_MAX_SCALE
                  1.500  # NCYCLE_IN_WINDOW
\end{verbatim}
%
The names below correspond to the variable names in \verb+mt_measure_adj+ (see subroutine \verb+read_par_file+ in \verb+mt_sub.f90+).
%
\begin{itemize}
\item Time vector: \verb+tt+ (start time), \verb+dtt+ (time step), \verb+nn+ (number of points). These values are particularly important if adjoint sources are requested. Based on the convention in \verb+SPECFEM3D+, for example, it is best to select the start time to be $t_0 = -1.5 \tau_{\rm h}$, where $\tau_{\rm h}$ is the half duration of the CMT source. The time step and the number of time steps should exactly match the time file of the \verb+SPECFEM3D+ synthetics.

\item Option for measurement: \verb+imeas0+ = 1--8 (see \refSec{sec:meas}).

\item Channel: \verb+chan+ is BH or LH (based on options for SEM synthetics).

\item Bandpass range: \verb+TLONG+ (long-period), \verb+TSHORT+ short-period

\item \verb+RUN_BANDPASS+: whether to bandpass the seismograms.

\item \verb+OUTPUT_MEASUREMENT_FILES+: whether to output measurement files for each window.

\item Limits on cross-correlation measurement: \verb+TSHIFT_MIN+; \verb+TSHIFT_MAX+. These can optionally be read in from the \verb+PAR_FILE+ used in the FLEXWIN algorithm.

\item Limits on amplitude measurement: \verb+DLNA_MIN+; \verb+DLNA_MAX+. These can optionally be read in from the \verb+PAR_FILE+ used in the FLEXWIN algorithm.

\item Limits on the cross-correlation measurement: \verb+CC_MIN+. This can optionally be read in from the \verb+PAR_FILE+ used in the FLEXWIN algorithm.

\item \verb+ERROR_TYPE+ = 0--2: Type of estimated errors associated with \blue{multi-taper} measurement (and adjoint source). No error (0), cross-correlation error estimate (1), multitaper jack-knife error estimate (2).  The cross-correlation error estimate is presented in \citet[][Appendix~A]{Tape2010}.  For imeas=5,6 (cc), \verb+sigma_dt/dlnA_cc+
is automatically used. For \verb+ERROR_TYPE = 0+, all \verb+sigma_tau+ and \verb+sigma_dlnA+ are set to be $1$ for multitaper adjoint sources.

\item \verb+DT_SIGMA_MIN+: water-level minimum cross-correlation traveltime difference uncertainty estimate.

\item \verb+DLNA_SIGMA_MIN+: water-level minimum amplitude difference uncertainty estimate.

\item Multitaper parameter, \verb+ITAPER+: Type of taper to use in constructing the transfer function between synthetics and data. Taper options are multitaper (1), cosine taper (2), and boxcar taper (3).  For the single-taper options (2--3) the transfer function is not used, as the adjoint sources are constructed directly from the synthetic seismograms. For the multitaper option, the number of tapers is fixed to be twice \verb+NPI+ (see \verb+multitaper_vala.pdf+).

\item Multitaper parameters, \verb+WTR+, \verb+NPI+ (see \verb+multitaper_vala.pdf+). Note \verb+WTR+ is also used to
determine \verb+i_right+ corresponding to the maximum frequency of valid frequency-dependent measurements. It requires 
that the power at this frequency is above $10$ times the \verb+WTR*max_syn_power+.

\item Multitaper parameter, \verb+DT_FAC+ (see \refSec{sec:MTparm}).

\item Multitaper parameter, \verb+ERR_FAC+ (see \refSec{sec:MTparm}).

\item Multitaper parameter, \verb+DT_MAX_SCALE+ (see \refSec{sec:MTparm}).

\item Multitaper parameter, \verb+NCYCLE_IN_WINDOW+, (see \refSec{sec:MTparm}).
\end{itemize}

%-------------

\subsubsection{Multitaper parameters}
\label{sec:MTparm}

The selection of the multitaper parameters \verb+DT_FAC+, \verb+ERR_FAC+, \verb+DT_MAX_SCALE+, and \verb+NCYCLE_IN_WINDOW+ are not easy for each particular dataset.  These parameters determine whether a multitaper measurement is reasonable enough to retain (otherwise revert to a cross-correlation measurement). The key subroutine is \verb+mt_measure_select.f90+.

MTMs are rejected (\verb+ user_trace = .false.+) based on:
\begin{itemize}
\item number of cycles in the window
\item number of frequency points given within \verb+[fstart,fend]+. Even when \verb+RUN_BANDPASS+ is turned off, the \verb+TLONG+ and \verb+TSHORT+ are still converted to \verb+fstart,fend+. They are best set to the values used in \verb+FLEXWIN+ 
\item \verb+tshift_cc <= dt+: too small a time shift
\item \verb+ abs(dtau_w(j)) > Tvec(j)/DT_FAC+: $\tau$ at a specific frequency is too high
\item \verb+ err_dt(j) > Tvec(j)/ERR_FAC+: error for a specific frequency is too high 
\item \verb+ abs(dtau_w(j)) > DT_MAX_SCALE*abs(tshift))+ $\tau$ at specific frequency deviates too much for global \verb+tshift_cc+
\end{itemize}
Note that the cross-correlation measurements are then run through \verb+cc_measure_select()+ to determine if they are
usable or not (\verb+tshift, dlnA = 0+).

Two example sets of parameters are given 
%
\begin{verbatim}
! ********* adjust following for global traces ************************
  ! ratio of current period with respect to dt and err_dt measurements
  double precision, parameter :: DT_FAC = 1.0  
  double precision, parameter :: ERR_FAC = 8.0
  ! max time shift allowed at all freq should be DT_MAX_SCALE * Tshift_xc
  double precision, parameter :: DT_MAX_SCALE = 5.0
  ! few cycles for surfaces waves
  double precision, parameter :: NCYCLE_IN_WINDOW = 2.0 ! window length > 2.*50
!***********************************************************************

! ********* adjust following for socal 3-30s traces ********************
  ! ratio of current period with respect to dt and err_dt measurements
  double precision, parameter :: DT_FAC = 2.0
  double precision, parameter :: ERR_FAC = 2.5 
  ! max time shift allowed at all freq should be DT_MAX_SCALE * Tshift_xc
  double precision, parameter :: DT_MAX_SCALE = 3.5
  ! use 3 cycles for surfaces waves
  double precision, parameter :: NCYCLE_IN_WINDOW = 1.5
!***********************************************************************
\end{verbatim}

%-------------

\subsection{Output files}
There are two output files related to measurements in the local run directory:
%
\begin{enumerate}

% unit 13
\item \verb+window_chi+, a comprehensive output file of misfit values, with one line per window.
\begin{verbatim}
! KEY: write misfit function values to file (for each window)
! Here are some columns of values in window_chi (add 8 for the actual column number)
!  1: MT-TT chi (imeas=7),    2: MT-dlnA chi (imeas=8)
!  3: XC-TT chi (imeas=5),    4: XC-dlnA chi (imeas=6)
!  5: MT-TT meas (freq-average),   6: MT-dlnA meas,   7: XC-TT meas,   8: XC-dlnA meas
!  9: MT-TT error, 10: MT-dlnA error, 11: XC-TT error, 12: XC-dlnA error
! WINDOW     : 13: data power, 14: syn power, 15: (data-syn) power, 16: window duration
! FULL RECORD: 17: data power, 18: syn power, 19: (data-syn) power, 20: record duration
! WINDOW     : 21: tr_chi,       22: am_chi
! Example of a reduced file: awk '{print $2,$3,$4,$5,$6,$29}' window_chi > window_tr_chi

write(13,'(a14,a8,a3,a5,i4,i4,2e14.6,20e14.6,2e14.6,2f14.6)') &
   file_prefix0,sta,net,chan_syn,j,imeas,&
   tstart,tend,window_chi(:),tr_chi,am_chi,T_pmax_dat,T_pmax_syn

! misfit function value
if(is_mtm==1) window_chi(1) = 0.5 * 2.0 * df * sum( (dtau_w(1:i_right))**2 * wp_taper(1:i_right) ) ! tr_chi
if(is_mtm==1) window_chi(2) = 0.5 * 2.0 * df * sum( (dlnA_w(1:i_right))**2 * wq_taper(1:i_right) ) ! amp_chi
window_chi(3) = 0.5 * (tshift/sigma_dt_cc)**2 ! tr_chi
window_chi(4) = 0.5 * (dlnA/sigma_dlnA_cc)**2 ! amp_chi
! cc/averaged-mt tshift/dlnA measurement
if(is_mtm==1) window_chi(5)  = sum( dtau_w(1:i_right) * w_taper(1:i_right) ) / sum(w_taper(1:i_right) )
if(is_mtm==1) window_chi(6)  = sum( dlnA_w(1:i_right) * w_taper(1:i_right) ) / sum(w_taper(1:i_right) )
window_chi(7) = tshift
window_chi(8) = dlnA
! estimated measurement uncertainties
if(is_mtm==1) window_chi(9) = sigma_dt
if(is_mtm==1) window_chi(10) = sigma_dlnA
window_chi(11) = sigma_dt_cc
window_chi(12) = sigma_dlnA_cc
! for normalization, divide by duration of window
window_chi(13) = 0.5 * sum(dat_dtw(:)**2)
window_chi(14) = 0.5 * sum(syn_dtw(:)**2)
window_chi(15) = 0.5 * sum((dat_dtw-syn_dtw)**2) ! tr/amp_chi for imeas=1/2
window_chi(16) = nlen*dt
window_chi(17) = 0.5 * sum( data**2 ) ! power of entire trace
window_chi(18) = 0.5 * sum( syn**2 )
window_chi(19) = 0.5 * sum( (data-syn)**2 )
window_chi(20) = npts*dt
\end{verbatim}
% unit 12
\item \verb+window_index+, an abbreviated output file with the indexing for each window. The columns are 
\begin{verbatim}
write(12,'(a3,a8,a5,a5,3i5,2f12.3)') net,sta,chan_syn,chan_dat,nwin,ipair,j,tstart,tend
\end{verbatim}
where \verb+nwin+ is the overall window counter, \verb+ipair+ is the seismogram (pair) counter, and \verb+j+ is the local window counter. For example, for the test case run:
%
\begin{verbatim}
CI MPM     BHR  BHR      1    1    1      11.277      58.477
CI MPM     BHR  BHR      2    1    2      58.477      81.377
CI MPM     BHR  BHR      3    1    3      81.377     101.677
CI MPM     BHT  BHT      4    2    1      49.327      95.327
CI MPM     BHZ  BHZ      5    3    1      60.977      99.377
\end{verbatim}
%
\end{enumerate}

With \verb+COMPUTE_ADJOINT_SOURCE = .true.+,  
\begin{enumerate}
\item adjoint source files (e.g., \verb+MPM.CI.BHZ.iker07.adj+) will appear
in \verb+OUTPUT_FILES/+. These are ascii files with the time colume defined by \verb+tstart,DT,npts+ in the parameter file.
(if \verb+DO_RAY_DENSITY_SOURCE = .true.+, adjoint file names are \verb+MPM.CI.BHZ.iker7.density.adj+).

\item \verb+window_chi_sum+ in local directory, the sum of misfit values (CC or MTM) of all windows, with weights taken into account if \verb+DO_WEIGHTING = .true.+. This scalar value literally corresponds to all the adjoint sources produced in  \verb+OUTPUT_FILES/+. If input includes all windows for a particular event, this scalar value is the event misfit.
More comprehensive sum of misfit can be found in the script
\begin{verbatim}
/ADJOINT_TOMO/iterate_adj/matlab/compute_misfit.m
\end{verbatim}
% added by LQY
\end{enumerate}

With \verb+OUTPUT_MEASUREMENT_FILES = .true.+ in \verb+MEASUREMENT.PAR+, you should find numerous output files in \verb+OUTPUT_FILES/+.
Other files include (\verb+prefix=MPM.CI.BHZ.01.mtm+, \verb+prefix0=MPM.CI.BHZ+)
\begin{verbatim}
prefix.recon_syn_cc.sac             # syn_dtw_cc
prefix.recon_syn_cc_dt.sac          # syn_dtw_cc_dt
prefix.dt/dlnA_average/cc           # dtau/dlnA_meas, dtaul/dlnA_sigma ('average' means 'mtm')
prefix.ph/abs/dlnA/ph_cor/dt        # transfer fun.:  phi/abs/log(abs)/phi(corr)/dtau(idf_new:idf_new:i_right)       
                                    #                 tshift/dlnA from cc are added to dtau/dlnA_wt
prefix.recon_syn.sac                # recon. syn with both dtau(om) and dlnA(om) applied for all imeas >=3 cases.
prefix.recon_syn_dt.sac             # reconstructed syn with only dtau(om) applied
prefix.err_ph/abs/dt/dlnA[_full]    # err_phi/abs/dt/dlnA_mtm(idf_new:idf_new:i_right)
prefix.freq_limits                  # fstart/fend (input),df, fstart/fend(adjusted for given window)
prefix0.recon.sac                   # recon. syn after applying CC or MTM for the entire data trace
\end{verbatim}

IF \verb+DISPLAY_MORE_DETAILS = .true.+ in the parameter file, following files will be written for every window pair
\begin{verbatim}
prefix.obs (original data/syn)
prefix.syn
prefix.dat.sac (time windowed data/syn)
prefix.syn.sac 
prefix.obs.power % abs(dat_dtwo(1:i_right)), i.e., dat_dtw_cc
prefix.syn.power % abs(syn_dtwo(1:i_right)), i.e., syn_dtw
\end{verbatim}

%-------------

\subsection{Scripts}

The three scripts in the run directory are
%
\begin{enumerate}
\item \verb+write_par_file.pl+. Write the file \verb+MEASUREMENT.PAR+ from a given set of parameters. This is called within the run scripts in \verb+scripts_tomo+.

\item \verb+prepare_adj_src.pl+. This script reads in a set of adjoint sources made from Z-R-T records and outputs a set of adjoint sources in Z-E-N that can be used in SPECFEM3D. 
\begin{verbatim}
prepare_adj_src.pl -m CMT -z BH -s STATION -o OUTDIR -i [rotation] all_adj_ZRT_files
prepare_adj_src.pl -m CMTSOLUTION_9818433 -s PLOTS/STATIONS_TOMO -o ADJOINT_SOURCES \
                   OUTPUT_FILES/*adj
\end{verbatim}
It loops over all the stations with adjoint sources present in \verb+OUTPUT_FILES+, and write an associate \verb+STATION_ADJOINT+ to local dir. It also rotates adjoint sources to E-N-Z, and deposits them into \verb+ADJOINT_SOURCES+ dir.
The key is that it will create an all-zeros record if no measurement is made on a particular component (say, Z), but IS made on another component (say, R or T). It uses the following fortran
program:
\begin{verbatim}
rotate_adj_src baz(radian!) zfile tfile rfile efile nfile
\end{verbatim}
The perl script also outputs \verb+STATIONS_ADJOINT+ file for all the stations with adjoint sources.
%{\bf NOTE: We need to eliminate the dependence on the library /opt/seismo-util/lib/perl.} --- LQY: done

\item \verb+combine_adj_src.pl+. This script combines two sets of adjoint sources (for example, two different bandpassed versions), and outputs the new adjoint sources, along with a new \verb+STATIONS_ADJOINT+ file to be used.
\begin{verbatim}
Usage: combine_adj_src.pl DIR_1 DIR_2 DIR_NEW imeas1 imeas2 chan
\end{verbatim}
Adjoint sources from \verb+dir1+ and \verb+dir2+ with the same station/net/component names are added up; station files 
from \verb+dir1/STATIONS_ADJOINT+ and \verb+dir2/STATIONS_ADJOINT+ are combined into \verb+dir_new/STATIONS_ADJOINT+.
\end{enumerate}
%

Scripts in the \verb+PLOT/+ directory:
%
\begin{itemize}
\item The primary plotting script is \verb+PLOTS/plot_win_adj.pl+.
\begin{verbatim}
plot_win_adj_all.pl -m CMTFILE -a STATION_FILE -n chan -b iboth -l tmin/tmax \
    -k imeas/iadj -d data_dir -s syn_dir -c recon_dir -w winfile -i smodel -j Tmin/Tmax
plot_win_adj_all.pl -l -10/200 -m ../CMTSOLUTION_9818433 -n BH -b 0 -k 7/1 \
    -a STATIONS_ADJOINT -d DATA -s SYN -c RECON -w MEASUREMENT.WINDOWS -i m16 -j 6/30
\end{verbatim}
which uses another script
\begin{verbatim}
plot_win_adj.pl -m $cmtfile -n $sta/$net/$chan -b $iboth -l $tcuts -k $opt_k \
   -a $station_list -d $data_dir -s $syn_dir -c $recon_dir -w $winfile -i $smodel -j $Ts
plot_win_adj.pl -m ../CMTSOLUTION_9818433 -n MPM/CI/BH -b 0 -l -10/200 -k 7/1 \
   -a STATIONS_ADJOINT -d DATA -s SYN -c RECON -w MEASUREMENT.WINDOWS -i m16 -j 6/30
\end{verbatim}
gives plots of detailed information on data, syn, reconstructed syn and corresponding adjoint sources, as well as 
source/station distributions on a map (Figure \ref{fig:iker07} for MTM, \verb+imeas=7+).
Some tweaking is probably necessary before successful application to a different dataset, including: setting
evid in CMT file (correspond to data name evid.net.sta), setting corresponding kstnam,knetwk,kcmpnm for both
data and synthetics, setting o header properly, and all synthetics need to be present (data/recon-syn not needed).

\item  plot the statistics of window files (check file structure before use)
\begin{verbatim}
plot_win_stats_all.pl Tmin/Tmax model iplot [plot histogram or not]
plot_win_stats.pl Tmin/Tmax eid name meas_file sta_file eid_text cmtall_psmeca
\end{verbatim}

\item organize \verb+plot_win_adj_all.pl+ output by station or event (???)
\begin{verbatim}
make_pdf_by_event.pl
make_pdf_by_station.pl
\end{verbatim}

\item For large datasets, we have included an additional set of scripts in the directory \verb+scripts_tomo+. Each user will have to make some modifications to these scripts.

\item For plotting output files for individual measurements, we have included a set of scripts in \verb+scripts_meas+.  These scripts have not been extensively tested.
\end{itemize}

%-------------------------------------------------------------------

%\pagebreak
\section{Measurement options}
\label{sec:meas}

There are several choices of measurements (or definitions of misfit functions), and each choice leads to a different adjoint source, as illustrated in \citet{Tromp2005}. The user must specify a value of \verb+imeas+ in \verb+MEASUREMENT.PAR+ with the following options:
%
\begin{enumerate}
\item \verb+imeas = 1+, normalized waveform difference. Adjoint source is constructed from the {\em data only}, with the form $-d(t)/\|d(t)\|^2$.
\eqa
\phi &=& \frac{1}{2}\frac{\int [d(t)-s(t)]^2\,dt }{\int d^2(t)\,dt} \quad \text{at}\quad s(t)=0 \nn\\
f^\dagger(t) &=& -\frac{d(t)}{\|d(t)\|^2} \nn
\ena
If \verb+NO_WAVEFORM_DIFFERENCE = .true.+, this becomes (why???)
\eq
f^\dagger(t) = d(t) \nn
\en

\item \verb+imeas = 2+, waveform difference, $s(t) - d(t)$.
\eqa
\phi &=&  \frac{1}{2}\int [d(t)-s(t)]^2\,dt  \nn\\
f^\dagger(t) &=& s(t)-d(t) \nn
\ena

\item \verb+imeas = 3+, cross-correlation traveltime for a (banana-doughtnut) sensitivity kernel. The measurement between data and synthetics {\em is not used} in constructing the adjoint source.
\eqa
\phi &=& T_{syn}  \nn\\
\delta \phi &=& \delta{T_{syn}} = - \frac{\int w(t) \dot{s}(t) \delta s(t)\,dt}{\int w(t) \dot{s}^2(t)\,dt} 
\quad\quad\quad\text{TTL (43)}\nn \\
f^\dagger(t) &=& -\frac{w(t)\dot{s}(t)}{\int w(t) \dot{s}^2(t)\,dt} \nn
\ena
where $w(t)$ is the time window over which the measurement is made. 
\red{Note the $-$ sign!.}

\item \verb+imeas = 4+, amplitude difference for a (banana-doughtnut) sensitivity kernel. The measurement between data and synthetics {\em is not used} in constructing the adjoint source. 
\eqa
\phi &=&  \ln A_{syn}  \nn\\
\delta \phi &=& \delta ln A_{syn} = \frac{\int w(t)s(t)\delta s(t)\,dt}{\int w(t) s^2(t)\,dt} 
\quad\quad\quad\text{TTL (64)}
\nn \\
f^\dagger(t) &=& \frac{w(t) s(t)}{\int w(t) s^2(t)\,dt} \nn
\ena

\item \verb+imeas = 5+, cross-correlation traveltime difference for an event kernel. 
\eqa
\phi &=& \frac{1}{2}\sum_{rp}\left[ \frac{T_{obs}-T_{syn}}{\sigma_{\Delta T}} \right]^2  \nn\\
\delta \phi &=& - \sum_{rp} \frac{T_{obs}-T_{syn}}{\sigma^2_{\Delta T}}\, \delta{T_{syn}} = \sum_{rp} \frac{ \Delta T_{syn}}{\sigma^2_{\Delta T}}
 \frac{\int w(t) \dot{s}(t) \delta s(t)\,dt}{\int w(t) \dot{s}^2(t)\,dt} \nn \\
f^\dagger(t) &=& \sum_{rp} \frac{w(t)\dot{s}(t)}{\int w(t) \dot{s}^2(t)\,dt}  \frac{ \Delta T_{syn}}{\sigma^2_{\Delta T}} \nn
\ena
where the traveltime \textit{delay} of observed data w.r.t. synthetics $\Delta T_{syn}= T_{obs}-T_{syn}$ {\em is used}  in constructing the adjoint source. The summation is performed over receivers ($r$) and phases ($p$).

Note the cross-correlation related measurements, including \verb+tshift, dlnA, cc_max+ 
given by \verb+compute_syn_cc()+ correspond to
\eq
cc_i = \frac{\sum_j s_j \red{d_{j+i}}}{\sum_j s_j s_j},\quad \Delta T = i_{max}*dt,
\quad
\Delta \ln A = \frac{1}{2} \ln \left[\frac{\int d^2(t) dt}{\int s^2(t) dt}\right],
\en
and average error estimates are given by
\eq
\sigma_{\Delta T} = \sqrt{\frac{\int[d(t)-s_c(t)]^2\,dt}{\int \dot{s}^2_c(t)\,dt}},\quad 
\sigma_{\Delta lnA} = \sqrt{\frac{\int[d(t)-s_c(t)]^2\,dt}{\int s^2_{ct}(t)\,dt}}
\en
where $s_c(t)$ is the reconstructed synthetics after applying $\Delta T$ and $\Delta \ln A$ corrections.

\item \verb+imeas = 6+, amplitude difference for an event kernel. As 
\eq
d(t) \sim \exp(\Delta \ln A)\, s(t) \sim (1+\Delta \ln A_{syn})\, s(t) \nn,
\en
the misfit of amplitude anomaly may be defined by
\eqa
\phi &=& \frac{1}{2}\sum_{rp}\left[\frac{\ln A_{obs} - \ln A_{syn}}{\sigma_{\Delta lnA}}\right]^2 
= \sum_{rp}\frac{1}{2}\left[\frac{\Delta \ln A}{\sigma_{\Delta lnA}}\right]^2  \nn\\
\delta \phi &=& -\sum_{rp}\frac{\Delta \ln A}{\sigma^2_{\Delta lnA}}\,\delta ln A = -\sum_{rp} \frac{\Delta \ln A}{\sigma^2_{\Delta lnA}} \frac{\int w(t)s(t)\delta s(t)\,dt}{\int w(t) s^2(t)\,dt} \nn \\
f^\dagger(t) &=&  -\sum_{rp}\frac{w(t) s(t)}{\int w(t) s^2(t)\,dt} \frac{\Delta \ln A}{\sigma^2_{\Delta lnA}}\nn
\ena
where the measurement amplitude anomaly between data and synthetics  $\Delta \ln A = \ln A_{obs}-\ln A_{syn} \sim A_{obs}/A_{syn} -1$ {\em is used} in constructing the adjoint source.

\item \verb+imeas = 7+, multitaper traveltime difference for an event kernel. The measurement between data and synthetics {\em is used} in constructing the adjoint source. See \verb+multitaper_notes.pdf+.
\eqa
&&\phi_P(\bbm)= \frac{1}{2}\sum_{rp} \int W_{P_{rp}}(\om) \left[\tau_{rp}^{obs}(\om) - \tau_{rp}(\om,\bbm)\right]^2\, d\om \nn \\
&&\delta \phi_P = -\sum_{rp}\int W_{P_{rp}}(\om) \Delta \tau_{rp}(\om, \bbm) \delta \tau_{rp}(\om,\bbm)\,d\om \nn \\ 
&&f^\dagger_P(t) = \sum_{rp}\sum_j h_j(t) P_j(t), \quad P_j(\om)=W_{P_{rp}}(\om) \Delta \tau_{rp}(\om) p_j(\om), \nn \\
&& \quad\quad\quad p_j(\om)=\frac{i\om s_j(\om)}{\sum_k |(i\om) s_k(\om)|^2}, \quad s_j(t)=s(t,\bbm)h_j(t)\nn
\ena
where $\Delta \tau_{rp}(\om,\bbm)=\tau^{obs}_{rp}(\om)-\tau_{rp}(\om,\bbm)$ is the frequency-dependent traveltime (\textit{delay}) measurements,
\eq
W_{P_{rp}}(\om)=\frac{W_{rp}(\om)}{\sigma_{P_{rp}}^2(\om)\int\,W_{rp}(\om)d\om}\nn
\en
is the frequency domain taper scaled by error estimates, and $h_j(t)$ is the $j$'th time-domain single taper applied to synthetics.
The frequency domain taper between $[f_1,f_2]$ may be defined by
\eq
W_{rp}(f)=1-\left[\cos\frac{\pi(f-f_1)}{f_2-f_1}\right]^\gamma, \quad \gamma = 10
\en

\item \verb+imeas = 8+, multitaper amplitude difference for an event kernel. The measurement between data and synthetics {\em is used} in constructing the adjoint source. See \verb+multitaper_notes.pdf+.
\eqa
\phi_Q(\bbm)&=& \frac{1}{2}\sum_{rp} \int W_{Q_{rp}}(\om) \left[ \ln A_{rp}^{obs}(\om) - \ln A_{rp}(\om,\bbm)\right]^2\, d\om \nn \\
\delta \phi_Q &=& -\sum_{rp}\int W_{Q_{rp}}(\om) \Delta \ln A_{rp}(\om, \bbm) \delta \ln A_{rp}(\om,\bbm)\,d\om \nn \\
f^\dagger_Q(t) &=& \sum_{rp}\sum_j h_j(t) Q_j(t), \quad  Q_j(\om)=W_{Q_{rp}}(\om) \Delta \ln A_{Q_{rp}}(\om) q_j(\om), \nn\\
&&\quad\quad q_j(\om)= -\frac{s_j}{\sum_k |s_k|^2}= i\om p_j(\om)\nn 
\ena
where $\Delta \ln A_{rp}(\om,\bbm)=\ln A^{obs}_{rp}(\om)-\ln A_{rp}(\om,\bbm)$ is the frequency-dependent amplitude anomaly measurements, and
\eq
W_{Q_{rp}}(\om)=\frac{W_{rp}(\om)}{\sigma_{Q_{rp}}^2(\om)\int\,W_{rp}(\om)d\om}\nn
\en
Both $\phi_P$ and $\phi_Q$ are dimensionaless numbers, and the adjoint sources $f^\dagger_P(t)$ and $f^\dagger_Q(t)$  have the units of $1/[\text{syn} \cdot \text{time}]$.

\end{enumerate}

\section{More Notes}
\subsection{NO\_WAVEFORM\_DIFFERENCE}
Carl: please add more here.

\subsection{DO\_WEIGHTING}
It is also possible to weigh windowed measurements of different categories if weighting is set in \verb+ma_weighting+ module (\verb+DO_WEIGHTING+):
\eq
\Phi = \sum_\alpha \, W_\alpha \, \sum_i \phi_{\alpha,i} 
\en
For instance, weights can be given to windows in \verb+P_SV/SH/Rayleigh/Love+ - \verb+Z/R/T+, 6 different categories of measurements separately. They are only used to change the traveltime adjoint sources (and corresponding $\phi$ values) at this
point.
\red{How about weighting related to the noise-level of the data trace itself?}

\subsection{Time-domain taper}
Time domain tapers are first applied to windowed data and synthetics (no matter what other options are)
\eq
w(t) = 1- \cos^a (2 \pi t/T), \quad 0\le t \le T
\en
where $a$ is an even integer (i.e., $a=10$) or a Welch window
\eq
w(t) = 1-\frac{(t-T/2)^2}{(T/2)^2}
\en

\subsection{Reonstructed synthetics}
\eq
\tilde{s}(\om)=s(\om) e^{\Delta \ln A(\om)- i \om \Delta\tau(\om)} 
\en
with the FFT convention in D\&T (p109)

\subsection{Post-processing of adjoint sources}
Note since data and synthetics have actually been pre-filtered, therefore the adjoint source that satisfies
\eq
\delta \phi = \int f^\dagger (t) s^f(t) dt = \frac{1}{2\pi} \int f^\dagger(\om) B^*(\om)s^*(\om)\,d\om 
=\int F^{-1}[f^\dagger(\om)B^*(\om)](t) s(t)\,dt
\en
i.e., the adjoint source also needs to be filtered by the same band-pass filter to produce the exact Fr\'{e}chet
derivatives.

\subsection{DO\_RAY\_DENSITY\_SOURCE}
Both \verb+sigma_tau+ and \verb+sigma_dlnA+, as well as measurements \verb+Delta_dtau/DlnA+ are set to be 1, 
and carried all the way to the adjoint source \verb+tr/amp_adj_src(:)+. \verb+DO_RAY_DENSITY_SOURCE+ automatically 
sets \verb+Error_type=0+. It provides a good indication of the volumetric coverage of the model domain by the
given measurement sets.

\subsection{MTM}
\verb+dtau_mtm(:)+ starts from $i > 1$, i.e., ignore the time shift at very long periods, and equivalently,
the first point of \verb+err_dt(1) = LARGE+ so that it does not contribute to adjoint sources and chi values.

\subsection{Questions}
Should we weigh data by their noise level (before the first arrival)?
%-------------------------------------------------------------------

\pagebreak
\section{Miscellaneous}

\subsection{Bug reports and suggestions for improvements}

To report bugs or suggest improvements to the code, please send an email to the CIG Computational Seismology Mailing List (cig-seismo@geodynamics.org) or Carl Tape (carltape@fas.harvard.edu), and/or use our online bug tracking system Roundup (www.geodynamics.org/roundup).

%-------------

\subsection{Notes and Acknowledgments}

This software package initiated as a effort of Jeroen Tromp's research group at Caltech starting in 2006.
The main developers of the MEAS-ADJ source code are Qinya Liu, Carl Tape, and Ying Zhou.
The multitaper measurement capability originated from codes used by Ying Zhou \citep[][]{YZhou2004,YZhou2005}.
The multitaper adjoint sources were implemented by Carl Tape \citep[][Appendix~C]{Tape2009phd}.
The organizational structure of the package was made by Qinya Liu, with adaptations by Carl Tape.
The following individuals have also contributed to the development of the source code or related scripts: Vala \vala, Min Chen.
The following individuals contributed to this manual: Carl Tape.

The \verb+measure_adj+ code makes use of filtering and enveloping algorithms that are part of SAC (Seismic Analysis Code, Lawerence Livermore National Laboratory) provided for free to IRIS members.  We thank Brian Savage for adding interfaces to these algorithms in recent SAC distributions.

We acknowledge support by the National Science Foundation under grant EAR-0711177.

%-------------

\subsection{License}

(See FLEXWIN manual for possible options.)

%-------------------------------------------------------------------
\pagebreak
%\small
\begin{spacing}{1}
\addcontentsline{toc}{section}{References}
\bibliographystyle{gji}
%\bibliography{preamble,REFERENCES,refs_carl,refs_socal}
\bibliography{measure}
\end{spacing}
%\normalsize
%-------------------------------------------------------------------

\begin{figure}
\includegraphics[width=17cm]{9818433_T006_T030_MPM_CI_m16_win.eps}
\caption[]
{{
Test case seismograms with {\tt COMPUTE$\_$ADJOINT$\_$SOURCE = .false.}.  Plot shows a three-component seismogram: Z = vertical, R = radial, T = transverse.  Black is the observed records, red is the synthetics, and blue is the reconstructed synthetics, made by applying the $\Delta T$ and $\Delta \ln A$ measurements within each window to the synthetics.  No adjoint sources are plotted.
\label{fig:noadj}
}}
\end{figure}

\begin{figure}
\includegraphics[width=17cm]{9818433_T006_T030_MPM_CI_m16_iker01_win_adj.eps}
\caption[]
{{
{\em Left:} Data (black), synthetics (red), and measurement windows.
{\em Right:} Adjoint sources constructed from the data alone ({\tt imeas = 1}), $-d(t)/\|d(t)\|^2$.
\label{fig:iker01}
}}
\end{figure}

\begin{figure}
\includegraphics[width=17cm]{9818433_T006_T030_MPM_CI_m16_iker02_win_adj.eps}
\caption[]
{{
{\em Left:} Data (black), synthetics (red), and measurement windows.
{\em Right:} Adjoint sources for a waveform difference measurement ({\tt imeas = 2}), $\bs - \bd$.
\label{fig:iker02}
}}
\end{figure}

\begin{figure}
\includegraphics[width=17cm]{9818433_T006_T030_MPM_CI_m16_iker03_win_adj.eps}
\caption[]
{{
{\em Left:} Data (black), synthetics (red), and measurement windows.
{\em Right:} Adjoint sources for a sensitivity kernel based on a cross-correlation traveltime difference, $\Delta T$ ({\tt imeas = 3}).
\label{fig:iker03}
}}
\end{figure}

\begin{figure}
\includegraphics[width=17cm]{9818433_T006_T030_MPM_CI_m16_iker04_win_adj.eps}
\caption[]
{{
{\em Left:} Data (black), synthetics (red), and measurement windows.
{\em Right:} Adjoint sources for a sensitivity kernel based on an amplitude difference, $\Delta \ln A$ ({\tt imeas = 4}).
\label{fig:iker04}
}}
\end{figure}

\begin{figure}
\includegraphics[width=17cm]{9818433_T006_T030_MPM_CI_m16_iker05_win_adj.eps}
\caption[]
{{
{\em Left:} Data (black), synthetics (red), and measurement windows.
{\em Right:} Adjoint sources for a cross-correlation traveltime difference, $\Delta T$ ({\tt imeas = 5}).
\label{fig:iker05}
}}
\end{figure}

\begin{figure}
\includegraphics[width=17cm]{9818433_T006_T030_MPM_CI_m16_iker06_win_adj.eps}
\caption[]
{{
{\em Left:} Data (black), synthetics (red), and measurement windows.
{\em Right:} Adjoint sources for an amplitude difference, $\Delta \ln A$ ({\tt imeas = 6}).
\label{fig:iker06}
}}
\end{figure}

\begin{figure}
\includegraphics[width=17cm]{9818433_T006_T030_MPM_CI_m16_iker07_win_adj.eps}
\caption[]
{{
{\em Left:} Data (black), synthetics (red), and measurement windows.
{\em Right:} Adjoint sources for a multitaper traveltime difference ({\tt imeas = 7}).
\label{fig:iker07}
}}
\end{figure}

\begin{figure}
\includegraphics[width=17cm]{9818433_T006_T030_MPM_CI_m16_iker08_win_adj.eps}
\caption[]
{{
{\em Left:} Data (black), synthetics (red), and measurement windows.
{\em Right:} Adjoint sources for a multitaper amplitude difference ({\tt imeas = 8}).
\label{fig:iker08}
}}
\end{figure}

%-------------------------------------------------------------------
\end{document}
%-------------------------------------------------------------------

### 2014-12-01 doc/USER_MANUAL/manual_SPECFEM3D_Cartesian.tex
fixed the hdur issue for FORCESOLUTION (the code was correct but hdur was used instead of f0 and standing for f0, which led to confusion). Also fixed a small error detected by Vadim Monteiller in a print statement in the acoustic case. Note that in DATA/FORCESOLUTION one now needs to enter f0 instead of hdur, and have f0: at the beginning of that input line instead of hdur:

@@
 \item [{\texttt{Mac OS}}] You will probably need to install \texttt{XCODE}.
-In addition, the \texttt{clock\_gettime} routine, which is used by
-the \texttt{SCOTCH} library that we use, does not exist in Mac OS.
-You may need to replace it with \texttt{clock\_get\_time} if you
-want to use \texttt{SCOTCH}.
 \end{description}
@@
 file, you will need to edit the East, North and vertical components
-of an arbitrary (non-unitary) direction vector of the force vector;
+of an arbitrary (not necessarily unitary, the code will normalize it automatically) direction vector of the force vector;
 thus refer to Appendix~\ref{cha:Coordinates} for the orientation
@@
 in the post-processing (see Section \ref{sec:Process-data-and-syn}).
-\item Set the \texttt{half duration} parameter of the Ricker source time
+\item Set the \texttt{f0} parameter (the dominant frequency) of the Ricker source time
 function when {\texttt{USE\_RICKER\_TIME\_FUNCTION}} is turned on
@@
 solver uses a (pseudo) Dirac delta source time function to represent
-a force point source, a very short \texttt{half duration} parameter
-(hdur = 5 {*} DT) is automatically set by default.
+a force point source, a very short duration of five time steps is automatically set by default.
 \item Set the latitude or UTM $x$ coordinate, longitude or UTM $y$ coordinate, depth of the source (in km).

### 2014-12-01 src/generate_databases/pml_set_local_dampingcoeff.f90

@@ -52,8 +52,6 @@ subroutine pml_set_local_dampingcoeff(myrank,xstore,ystore,zstore)
   ! local parameters
   integer :: i,j,k,ispec,iglob,ispec_CPML,ier
 
-  ! JC JC: Remove the parameter definition here and make the calculation of ALPHA_MAX_PML automatic
-  !        by recovering the value of hdur in FORCESOLUTION/CMTSOLUTION
   real(kind=CUSTOM_REAL) :: ALPHA_MAX_PML
   real(kind=CUSTOM_REAL), parameter :: K_MAX_PML = 1.0d0,K_MIN_PML=1.0d0
   real(kind=CUSTOM_REAL) :: pml_damping_profile_l,dist,vp

### 2014-12-23 doc/USER_MANUAL/manual_SPECFEM3D_Cartesian.tex
updates manual

@@
 
-You can add -DFORCE\_VECTORIZATION to the compiler options in \texttt{flags.guess} (for all compilers, except for IBM for which the syntax is -WF,-DFORCE\_VECTORIZATION )
-to speed up the code in the fluid (acoustic) parts (only; FORCE\_VECTORIZATION support for elastic parts has been discontinued in the source code).
+You can add \texttt{--enable-vectorization} to the configuration options to speed up the code in the fluid (acoustic) and elastic parts.
 This works fine if (and only if) your computer always allocates a contiguous memory block for each allocatable array;
-this is the case for most machines and most compilers, but not all. For more details see https://github.com/geodynamics/specfem3d/issues/81 .
-To check if that option works fine on your machine, run the code with and without it for a model containing a significant fluid layer
-(or entirely fluid) and make sure the seismograms are identical.\\
+this is the case for most machines and most compilers, but not all. To disable this feature, use option \texttt{--disable-vectorization}. 
+For more details see \href{https://github.com/geodynamics/specfem3d/issues/81}{github.com/geodynamics/specfem3d/issues/81} .
+To check if that option works fine on your machine, run the code with and without it for an acoustic/elastic model and make sure the seismograms are identical.\\
 
@@
 
-\texttt{http://www.softeng.rl.ac.uk/blog/2010/jan/30/callgraph-fortran-doxygen/} ,
+\url{http://www.softeng.rl.ac.uk/blog/2010/jan/30/callgraph-fortran-doxygen/},
 
@@
 
-\texttt{http://www.stack.nl/~dimitri/doxygen/manual/diagrams.html} ,
+\url{http://www.stack.nl/~dimitri/doxygen/manual/diagrams.html},
 
@@
 
-\texttt{http://www.stack.nl/~dimitri/doxygen/manual/config.html} ,
+\url{http://www.stack.nl/~dimitri/doxygen/manual/config.html},
 
@@
 
-\texttt{http://www.stack.nl/~dimitri/doxygen/manual/index.html} .
+\url{http://www.stack.nl/~dimitri/doxygen/manual/index.html}.
 
@@
 
+\noindent
 The basic steps in creating a load-balanced, partitioned mesh with
@@
 script \texttt{cubit2specfem3d.py} provided in each of the example
-directories (linked to the common script \texttt{CUBIT\_GEOCUBIT/cubit2specfem3d.py}). If the export was successful, you should find the following
+directories (linked to the common script \texttt{CUBIT\_GEOCUBIT/cubit2specfem3d.py}). 
+If the export was successful, you should find the following
 files in a subdirectory \texttt{MESH/}:
@@
 modify the header of the \texttt{check\_mesh\_quality\_CUBIT\_Abaqus.f90}
-source file. Then, in the main directory, type
+source file in directory \texttt{src/check\_mesh\_quality\_CUBIT\_Abaqus/}.
+
+\noindent
+Then, in the main directory, type
 \begin{lyxcode}
@@
 The partitioning is done in serial for now (in the next release we
-will provide a parallel version of that code). It needs to be run in the \texttt{bin/} directory because it expects the \texttt{../DATA/Par\_file}. The synopsis is:
-\begin{lyxcode}
-./xdecompose\_mesh~nparts~input\_directory~output\_directory
+will provide a parallel version of that code). It needs to be run in the main directory because it expects the \texttt{./DATA/Par\_file}. The synopsis is:
+\begin{lyxcode}
+./bin/xdecompose\_mesh~nparts~input\_directory~output\_directory
 \end{lyxcode}
@@
 generated by the Python script \texttt{cubit2specfem3d.py} explained
-in the previous Section~\ref{subsec:Exporting-the-Mesh}, e.g. \texttt{MESH/},
+in the previous Section~\ref{subsec:Exporting-the-Mesh}, e.g. \texttt{./MESH/},
 and
@@
 databases, and have to reside in the directory \texttt{LOCAL\_PATH}
-specified in the main \texttt{Par\_file}, e.g. in directory \texttt{OUTPUT\_FILES/DATABASES\_MPI}.
+specified in the main \texttt{Par\_file}, e.g. in directory \texttt{./OUTPUT\_FILES/DATABASES\_MPI}.
 Please see Chapter~\ref{cha:Creating-Distributed-Databases} for
@@
 
+\noindent
 In the main directory, type
@@
 now compile and produce the executable \texttt{bin/xmeshfem3D}. Please
-note that \texttt{xmeshfem3D} must be called directly from the \texttt{bin/}
+note that \texttt{xmeshfem3D} must be called directly from the main
 directory, as most of the binaries of the package.
@@
 Please note that \texttt{xgenerate\_databases} must be called directly
-from the \texttt{bin/} directory, as most of the binaries of the package.
+from the main directory, as most of the binaries of the package.
 
@@
 Please note that \texttt{xspecfem3D} must be called directly from
-the \texttt{bin/} directory, as most of the binaries of the package.
-
+the main directory, as most of the binaries of the package.\\
+
+\noindent
 The solver needs three input files in the \texttt{DATA} directory
@@
 Centroid-Moment Tensor (CMT) web page \urlwithparentheses{www.seismology.harvard.edu}.
-It looks like this:
-\begin{lyxcode}
-\begin{figure}[H]
-\noindent \begin{centering}
-{\small \includegraphics[width=1\textwidth]{figures/Hollywood_CMT}
+It looks like the example shown in Fig. \ref{fig:CMTSOLUTION-file}.
+%\begin{lyxcode}
+\begin{figure}[htp]
+\noindent 
+\begin{centering}
+{\small \includegraphics[width=0.8\textwidth]{figures/Hollywood_CMT}
 }
-\par\end{centering}
-
-\caption{\label{fig:CMTSOLUTION-file}\texttt{CMTSOLUTION} file based on the
+\par
+\end{centering}
+\caption{\texttt{CMTSOLUTION} file based on the
 format from the Harvard CMT catalog. \textbf{M} is the moment tensor,
@@
 magnitude.}
+\label{fig:CMTSOLUTION-file}
 \end{figure}
-
-\end{lyxcode}
+%\end{lyxcode}
 The \texttt{CMTSOLUTION} file should be edited in the following way:
@@
 \end{itemize}
-\begin{figure}
+\begin{figure}[htp]
 \noindent \begin{centering}
-\includegraphics[width=2.5in]{figures/gauss_vs_triangle_mod}
+\includegraphics[width=3.5in]{figures/gauss_vs_triangle_mod}
 \par\end{centering}
@@
 each of them computing several earthquakes in parallel.
-To turn that option on, set parameter NUMBER\_OF\_SIMULTANEOUS\_RUNS
+To turn that option on, set parameter \texttt{NUMBER\_OF\_SIMULTANEOUS\_RUNS}
 to a value greater than 1 in file setup/constants.h.in before
@@
 When that option is on, of course the number of processor cores used to start
-the code in the batch system must be a multiple of NUMBER\_OF\_SIMULTANEOUS\_RUNS,
+the code in the batch system must be a multiple of \texttt{NUMBER\_OF\_SIMULTANEOUS\_RUNS},
 all the individual runs must use the same number of processor cores,
-which as usual is NPROC in the input file DATA/Par\_file,
+which as usual is \texttt{NPROC} in the input file \texttt{DATA/Par\_file},
 and thus the total number of processor cores to request from the batch system
-should be NUMBER\_OF\_SIMULTANEOUS\_RUNS $\times$ NPROC.
-All the runs to perform must be placed in directories called run0001, run0002, run0003 and so on (with exactly four digits)
+should be \texttt{NUMBER\_OF\_SIMULTANEOUS\_RUNS $\times$ NPROC}.
+All the runs to perform must be placed in directories called \texttt{run0001}, \texttt{run0002}, \texttt{run0003} and so on (with exactly four digits)
 and you must create a link from the root directory of the code
-to the first copy of the executable programs by typing " ln -s run0001/bin bin ".
+to the first copy of the executable programs by typing \texttt{ln -s run0001/bin bin}.
 
@@
 This approach is known as \textquotedbl{}split nodes\textquotedbl{}.
-Currently faults can only be run with \texttt{xdecompose\_mesh }and\texttt{
-CUBIT. xmeshfem3D }is not yet ready to handle faults.\\
+Currently faults can only be run with \texttt{xdecompose\_mesh} and
+\texttt{CUBIT}. \texttt{xmeshfem3D} is not yet ready to handle faults.\\
 To facilitate the mesh generation with split nodes in CUBIT, we need
@@
 
-The CUBIT scripts ({*}.jou and {*}.py) in the directory $\mathtt{EXAMPLES}$
+The CUBIT scripts ({*}.jou and {*}.py) in the directory \texttt{EXAMPLES}
 generate more complicated meshes. The {*}.py files are Python scripts
@@
 \end{itemize}
-Read the documents in the directory $\mathtt{EXAMPLES/*/description}$.
+Read the documents in the directory \texttt{EXAMPLES/*/description}.
 They contain a description of the example and additional instructions
@@
 
-Three additional input files are required in the \texttt{DATA} directory
+Three additional input files are required in the \texttt{DATA/} directory
 for dynamic and kinematic faults. They are $\mathtt{Par\_file\_faults}$,
@@
 \item Look for the \textquotedbl{}maximum suggested time step\textquotedbl{}
-in $\mathtt{OUTPUT\_FILES/output\_mesher.txt}$. This is the critical
+in \texttt{OUTPUT\_FILES/output\_mesher.txt}. This is the critical
 timestep of a purely elastic simulation, $\mathtt{dtc\_bulk}$.
@@
 
-Several output files are saved in $\mathtt{OUTPUT\_FILES}$:\\
+Several output files are saved in \texttt{OUTPUT\_FILES/}:\\
 
 \begin{enumerate}
-\item Seismograms for each station on the fault plane given in $\mathtt{DATA/FAUL\_STATIONS}$.
+\item Seismograms for each station on the fault plane given in \texttt{DATA/FAUL\_STATIONS}.
 One output file is generated for each station, named after the station.
@@
 \item Seismograms at stations in the bulk (out of the fault plane) given
-in $\mathtt{DATA/STATIONS}$.
+in \texttt{DATA/STATIONS}.
 \item Rupture time files are named $\mathtt{Rupture\_time\_FAULT}$-$\mathtt{id}$.
@@
 
-The directories $\mathtt{EXAMPLES/*/post}$ contain additional Matlab
+The directories \texttt{EXAMPLES/*/post} contain additional Matlab
 scripts to generate figures specific to each example.
@@
 for the selected stations and saved in a sub-directory called \texttt{SEM/}
-in the root directory of the code, with the format \texttt{STA.NT.BX?.adj}, where \texttt{STA},
-\texttt{NT} are the station name and network code given in the \texttt{DATA/STATIONS\_ADJOINT}
+in the root directory of the code, with the format \texttt{NT.STA.BX?.adj}, where \texttt{NT},
+\texttt{STA} are the network code and station name given in the \texttt{DATA/STATIONS\_ADJOINT}
 file, and \texttt{BX?} represents the component name of a particular
@@
 \item The adjoint seismograms are in the same format as the original seismogram
-(\texttt{STA.NT.BX?.sem?}), with the same start time, time interval
+(\texttt{NT.STA.BX?.sem?}), with the same start time, time interval
 and record length.
@@
 earthquake sources given by the \texttt{DATA/CMTSOLUTION} file, and
-have names of the form \texttt{S?????.NT.S??.sem} for the six-component
+have names of the form \texttt{NT.S?????.S??.sem} for the six-component
 strain tensor (\texttt{SNN,SEE,SZZ,SNE,SNZ,SEZ}) at these locations,
 and ~\\
- \texttt{S?????.NT.BX?.sem} for the three-component displacements
+ \texttt{NT.S?????.BX?.sem} for the three-component displacements
 (\texttt{BXN,BXE,BXZ}) recorded at these locations.
@@
 and run the executable \texttt{xcreate\_movie\_shakemap\_AVS\_DX\_GMT}
-in the \texttt{bin/} subdirectory. It will create visualization files
+in the main directory. It will create visualization files
 in your format of choice. The code will prompt the user for input
@@
 \end{lyxcode}
-which will create the executable \texttt{xcombine\_vol\_data} in the
-directory \texttt{bin/}. To output the usage of this executable, type
+which will create the executable \texttt{bin/xcombine\_vol\_data}. 
+To output the usage of this executable, type
 \begin{lyxcode}
@@
 in the \texttt{OUTPUT\_FILES/} directory by default. For the forward
-simulation, the files are named like \texttt{STA.NT.BX?.semd} for
-two-column time series, or \texttt{STA.NT.BX?.semd.sac} for ASCII
-SAC format, where STA and NT are the station name and network code,
+simulation, the files are named like \texttt{NT.STA.BX?.semd} for
+two-column time series, or \texttt{NT.STA.BX?.semd.sac} for ASCII
+SAC format, where NT and STA are the network code and station name,
 and \texttt{BX?} stands for the component name. Please see the Appendix~\ref{cha:Coordinates}
@@
 The adjont simulations generate synthetic seismograms with the name
-\texttt{S?????.NT.S??.sem} (refer to Section \ref{sec:Adjoint-simulation-sources}
+\texttt{NT.S?????.S??.sem} (refer to Section \ref{sec:Adjoint-simulation-sources}
 for details). The kernel simulations output the back-reconstructed
-synthetic seismogram in the name \texttt{STA.NT.BX?.semd}, mainly
+synthetic seismogram in the name \texttt{NT.STA.BX?.semd}, mainly
 for the purpose of checking the accuracy of the reconstruction. Refer
@@
 vertical component of a station \texttt{ASBS} (network \texttt{AZ})
-will be named \texttt{ASBS.AZ.MXZ.semd.sac}, whereas it will be \texttt{ASBS.AZ.BXZ.semd.sac},
+will be named \texttt{AZ.ASBS.MXZ.semd.sac}, whereas it will be \texttt{AZ.ASBS.BXZ.semd.sac},
 if the sampling rate is greater than 0.0125 and less equal to 0.1

### 2014-12-23 doc/USER_MANUAL/manual_SPECFEM3D_Cartesian.tex
updates SPECFEM3D version number to version 3.0

@@
 and ETH Z\"urich (Switzerland)\\
-Version 2.1}
+Version 3.0}
 
@@
 
-This new release of the code (version 2.1) includes Convolution or
+This new release of the code includes Convolution or
 Auxiliary Differential Equation Perfectly Matched absorbing Layers
@@
 \end{description}
-\noindent If you use this new version 2.1, which has non blocking
+\noindent If you use this new version, which has non blocking
 MPI for much better performance for medium or large runs, please cite
@@
 
-The directory containing utilities for tomographic inversion using
+The directory containing external utilities for tomographic inversion using
 SPECFEM3D Cartesian (or other packages that evaluate misfit functions
-and gradients) is here on the CIG svn server: \begin{verbatim} /cig/seismo/3D/ADJOINTTOMO/
+and gradients) is in directory \texttt{utils/ADJOINT\_TOMOGRAPHY\_TOOLS/}:
+\begin{verbatim}
 flexwin/     -- FLEXWIN algorithm for automated picking of time windows
-measureadj/ -- reads FLEXWIN output file and makes measurements, with
-the option for computing adjoint sources iterateadj/ -- various tools
-for iterative inversion (requires pre-computed \textquotedbl{}event
-kernels\textquotedbl{}) \end{verbatim} This directory also contains
+measure_adj/ -- reads FLEXWIN output file and makes measurements,
+                with the option for computing adjoint sources
+iterate_adj/ -- various tools for iterative inversion
+                (requires pre-computed "event kernels")
+\end{verbatim}
+This directory also contains
 a brief \verb+README+ file indicating the role of the three subdirectories,
@@
 
-MPI v 2.1, July 2012: Max Rietmann, Peter Messmer, Daniel Peter, Dimitri
+MPI v. 3.0, December 2014: many developers.
+Convolutional PML, LDDRK time scheme, bulk attenuation support, simultaneous MPI runs,
+ADIOS file I/O support, coupling with external codes, new seismogram names,
+Deville routines for additional GLL degrees, tomography tools, unit/regression test framework,
+improved CUDA GPUs performance, additonal GEOCUBIT support, better make compilation,
+git versioning system. \\
+
+MPI v. 2.1, July 2012: Max Rietmann, Peter Messmer, Daniel Peter, Dimitri
 Komatitsch, Joseph Charles, Zhinan Xie: support for CUDA GPUs, better

### 2015-01-10 doc/USER_MANUAL/manual_SPECFEM3D_Cartesian.tex
fixed a typo

@@
 (SEM). The SEM is a continuous Galerkin technique \citep{TrKoLi08,PeKoLuMaLeCaLeMaLiBlNiBaTr11},
-which can easily be made discontinous \citep{BeMaPa94,Ch00,KoWoHu02,ChCaVi03,LaWaBe05,Kop06,WiStBuGh10,AcKo11};
+which can easily be made discontinuous \citep{BeMaPa94,Ch00,KoWoHu02,ChCaVi03,LaWaBe05,Kop06,WiStBuGh10,AcKo11};
 it is then close to a particular case of the discontinuous Galerkin

### 2015-01-11 doc/USER_MANUAL/manual_SPECFEM3D_Cartesian.tex
added Christina Morency's section about poroelastic runs to the manual, and fixed a problem on the cover

@@
 \end{description}
+
+The present version of SPECFEM can handle fully saturated porous simulations, Christina Morency implemented Biot equation. 
+But the code cannot calculate partially saturated cases in its state.
+Christina Morency is presently working on a dual porosity, dual permeability formulation, type Pride and Berryman, for an other project, but it will not be available for some time.\\
+
+The way we prescribe material property for porous material in SPECFEM3D is as follow:
+We use a file name nummaterial\_poroelastic\_file, which is located in the directory MESH, the format is as follow:\\
+ rhos rhof phi c kxx kxy kxz kyy kyz kzz Ks Kf Kfr etaf mufr\\
+where\\ 
+rho\_s = solid density\\
+rho\_f = fluid density\\
+phi = porosity\\
+tort = tortuosity\\
+kxx = xx component of permeability tensor\\
+kxy = xy,yx components of permeability tensor\\
+kyy = yy component of permeability tensor\\
+kxz = xz,zx components of permeability tensor\\
+kzz = zz component of permeability tensor\\
+kappa\_s = solid bulk modulus\\
+kappa\_f= fluid bulk modulus\\
+kappa\_fr= frame bulk modulus\\
+eta\_f = fluid viscosity\\
+mu\_fr = frame shear modulus.\\
+
+Using an external mesh (for instance coming from CUBIT/TRELIS), poroelastic materials have the ID number 3, while 1 is elastic, 2 is acoustic ?
+see the example in the package: examples/homogeneous\_poroelastic.
+
 If you use PML, the mesh elements that belong to the PML layers can

### 2015-01-20 doc/USER_MANUAL/manual_SPECFEM3D_Cartesian.tex
more logos

@@
 The cover graphic of the manual was created by Santiago Lombeyda from
-Caltech's Center for Advanced Computing Research (CACR) \urlwithparentheses{http://www.cacr.caltech.edu}.
-\\
-
-\section*{Current and past main participants or main sponsors}
-
+Caltech's Center for Advanced Computing Research (CACR) \urlwithparentheses{http://www.cacr.caltech.edu}.\\
+
+The code is released open-source under the GNU version 2+ license, see the license at the end of this manual.\\
+
+\newpage{}
+\section*{Current and past main participants or main sponsors (in no particular order)}
+%
 \begin{figure}[htbp]
@@
 \includegraphics[width=0.15\textwidth]{figures/logo_aix_marseille_universite}\vspace*{0.02truemm}
+\includegraphics[width=0.15\textwidth]{figures/logo_ETH}\vspace*{2truemm}
 \includegraphics[width=0.12\textwidth]{figures/logo_CSC_China}\vspace*{0.02truemm}
-\includegraphics[width=0.15\textwidth]{figures/logo_ETH}\vspace*{2truemm}
 \includegraphics[width=0.15\textwidth]{figures/logo_inria}\vspace*{2truemm}
@@
 \par\end{centering}
-
+%
+\vspace*{-2truemm}
+%
 \noindent \begin{centering}
@@
 \includegraphics[width=0.135\textwidth]{figures/logo_IPGP}\vspace*{2truemm}
+\includegraphics[width=0.112\textwidth]{figures/logo_ANR}\vspace*{2truemm}
 \par\end{centering}
-
+%
+\vspace*{-2truemm}
+%
 \noindent \begin{centering}
+\includegraphics[width=0.190\textwidth]{figures/logo_European_Union}\vspace*{2truemm}
+\includegraphics[width=0.112\textwidth]{figures/logo_GENCI}\vspace*{2truemm}
+\includegraphics[width=0.112\textwidth]{figures/logo_PRACE}\vspace*{2truemm}
+\includegraphics[width=0.112\textwidth]{figures/logo_CINES}\vspace*{2truemm}
+\includegraphics[width=0.130\textwidth]{figures/logo_Oak_Ridge}\vspace*{2truemm}
+\hspace*{3mm}\includegraphics[width=0.130\textwidth]{figures/logo_fondation_Del_Duca}
+\par\end{centering}
+%
+\vspace*{-2truemm}
+%
+\noindent \begin{centering}
+\includegraphics[width=0.112\textwidth]{figures/logo_University_of_Toronto}\vspace*{2truemm}
 \includegraphics[width=0.119\textwidth]{figures/logo_INGV}\vspace*{2truemm}
-\includegraphics[width=0.119\textwidth]{figures/logo_University_of_Toronto}\vspace*{2truemm}
 \includegraphics[width=0.112\textwidth]{figures/logo_Univ_Toulouse}\vspace*{2truemm}
 \includegraphics[width=0.112\textwidth]{figures/logo_TOTAL}\vspace*{2truemm}
-\includegraphics[width=0.130\textwidth]{figures/logo_Fairbanks}
-\hspace*{3mm}\includegraphics[width=0.130\textwidth]{figures/logo_fondation_Del_Duca}
+\includegraphics[width=0.130\textwidth]{figures/logo_Fairbanks}\vspace*{2truemm}
+\includegraphics[width=0.140\textwidth]{figures/logo_Intel_Exascale_Labs}\vspace*{2truemm}
+\includegraphics[width=0.130\textwidth]{figures/logo_Maison_Simulation}
 \par\end{centering}

### 2015-01-21 doc/USER_MANUAL/manual_SPECFEM3D_Cartesian.tex

@@
 \newpage{}
-\section*{Current and past main participants or main sponsors (in no particular order)}
+\section*{Current and past main participants or main sponsors of the SPECFEM project (in no particular order)}
 %
@@
 \includegraphics[width=0.112\textwidth]{figures/logo_ANR}\vspace*{2truemm}
+\includegraphics[width=0.112\textwidth]{figures/logo_NSF}\vspace*{2truemm}
 \par\end{centering}
@@
 \hspace*{3mm}\includegraphics[width=0.130\textwidth]{figures/logo_fondation_Del_Duca}
+\includegraphics[width=0.130\textwidth]{figures/logo_CIG}\vspace*{2truemm}
 \par\end{centering}
@@
 \includegraphics[width=0.112\textwidth]{figures/logo_University_of_Toronto}\vspace*{2truemm}
-\includegraphics[width=0.119\textwidth]{figures/logo_INGV}\vspace*{2truemm}
-\includegraphics[width=0.112\textwidth]{figures/logo_Univ_Toulouse}\vspace*{2truemm}
+\includegraphics[width=0.109\textwidth]{figures/logo_INGV}\vspace*{2truemm}
+\includegraphics[width=0.102\textwidth]{figures/logo_Univ_Toulouse}\vspace*{2truemm}
 \includegraphics[width=0.112\textwidth]{figures/logo_TOTAL}\vspace*{2truemm}
 \includegraphics[width=0.130\textwidth]{figures/logo_Fairbanks}\vspace*{2truemm}
+\includegraphics[width=0.120\textwidth]{figures/logo_CINECA}\vspace*{2truemm}
 \includegraphics[width=0.140\textwidth]{figures/logo_Intel_Exascale_Labs}\vspace*{2truemm}

### 2015-01-22 doc/GJI_2011_paper_about_the_code/paper_GJI_2011_about_the_code.tex
ran the code beautifier, which now removes accented letters and cleans text and LaTeX files

@@
 together with modern computers to enhance
-the quality of images of Earth’s interior and the earthquake rupture process.
+the quality of images of Earth's interior and the earthquake rupture process.
 Most traditional tomographic methods utilize traveltime or phase information measured by comparing data with simulations,

### 2015-01-22 utils/ADJOINT_TOMOGRAPHY_TOOLS/ADJOINT_TOMOGRAPHY_TOOLKIT/HowtoUseToolkit.tex
ran the code beautifier, which now removes accented letters and cleans text and LaTeX files

--- 2014-08-06 ADJOINT_TOMOGRAPHY_TOOLS/ADJOINT_TOMOGRAPHY_TOOLKIT/HowtoUseToolkit.tex
@@
 
-- mention somewhere that Vadim and Sébastien Chevrot will commit their routines to interpolate back and forth between the SEM mesh and a regular mesh; mention that we should talk to Qinya about this, since she was in the process of implementing that as well (I think that Vadim and Sébastien's routine will be sufficient, but we should doublecheck with Qinya if she had something different (more powerful?
+- mention somewhere that Vadim and Sebastien Chevrot will commit their routines to interpolate back and forth between the SEM mesh and a regular mesh; mention that we should talk to Qinya about this, since she was in the process of implementing that as well (I think that Vadim and Sebastien's routine will be sufficient, but we should doublecheck with Qinya if she had something different (more powerful?
 or faster?) in mind

### 2015-01-22 utils/DSM_FOR_SPECFEM3D/OLD--VM_mesher_now_handled_by_meshfem3d/mesh_chunk.f90

@@ -6,7 +6,7 @@
 !  Vadim Monteiler, Fevrier 2013
 !
 !
-! J'ai une convention propre pour le mapping de la sphere cubique (a compléter ... ) !!!!!!
+! J'ai une convention propre pour le mapping de la sphere cubique (a completer ... ) !!!!!!
 !
 program mesh_chunk
   implicit none
@@ -40,7 +40,7 @@ program mesh_chunk
 
   ! nb couches dans modele iasp91 ou ak135 ou prem
   integer nlayer
-  parameter (nlayer=12) ! 1 couche de plus que le modèle
+  parameter (nlayer=12) ! 1 couche de plus que le modele
   double precision zlayer(nlayer),vpv(nlayer,4),vsv(nlayer,4),density(nlayer,4)
   double precision, dimension(:,:), allocatable :: ProfForGemini
   double precision Z_DEPTH_BLOCK,UTM_X_MIN,UTM_X_MAX
@@ -57,7 +57,7 @@ program mesh_chunk
   character(len=250) model1D_file
 
   !! CONVENTION : (lon,lat) -> (xi,eta) (k=6 avec -z pour le mapping sphere cubique (cf Chervot 2012)
-   ! on definit le maillage d'un chunk dans la sphère cubique-------------------------------------------------------
+   ! on definit le maillage d'un chunk dans la sphere cubique-------------------------------------------------------
   PI = 3.141592653589793d0
   deg2rad = 3.141592653589793d0/180.d0
   R_EARTH=6371000.d0
@@ -75,7 +75,7 @@ program mesh_chunk
      chunk_azi=0.d0 !90.d0 !80.d0 !10.d0  !* deg2rad
      ! depth
      chunk_depth = 1000.d0 *1000.d0 ! 250.d0 * 1000.d0
-     ! nb d'éléments
+     ! nb d'elements
      nel_lat = 20 !120
      nel_lon = 40 !96
      nel_depth = 20 !100
@@ -99,7 +99,7 @@ program mesh_chunk
      ANGULAR_WIDTH_XI_RAD = deg2rad * ANGULAR_WIDTH_XI_RAD
      ANGULAR_WIDTH_ETA_RAD = deg2rad * ANGULAR_WIDTH_ETA_RAD
      chunk_depth = chunk_depth * 1000.d0
-  endif   !这里是读入chunk的尺寸和网点间隔，需要ParFileMeshChunk文件
[…]

### 2015-01-22 doc/USER_MANUAL/manual_SPECFEM3D_Cartesian.tex

@@
 Percy Galvez,
-Dominik Göddeke,
-Vala Hjörleifsdóttir,
+Dominik G\"oddeke,
+Vala Hj\"orleifsd\'ottir,
 Sue Kientz,
 Dimitri Komatitsch,
-Jesús Labarta,
+Jes\'us Labarta,
 Nicolas Le Goff,

### 2015-01-27 doc/USER_MANUAL/manual_SPECFEM3D_Cartesian.tex
added more details about PML to the manual (about regular elements inside the PMLs for better time stability)

@@
 
+If you use PML and an external mesh (created using an external meshing tool 
+such as CUBIT/TRELIS or similar), try to have elements inside the PML as regular as possible, 
+i.e. ideally non-deformed cubes obtained by `extrusion' of regular surface mesh elements meshing the
+outer edges of the computational domain without PML; by doing so, the PMLs obtained will be far more stable
+in time (PML being weakly unstable from a mathematical point of view, very deformed mesh elements
+inside the PMLs can trigger instabilities much more quickly).
+
 If you use PML and an external tomographic velocity and density model,

### 2015-03-05 doc/USER_MANUAL/manual_SPECFEM3D_Cartesian.tex
Buildbot ran the code beautifier.

@@
 C\'eline Blitz,
+Alexis Bottero,
 Ebru Bozda\u{g},
@@
 Min Chen,
+Paul Cristini,
 Percy Galvez,
@@
 Zhinan Xie,
+Chang-Hua Zhang,
 Hejun Zhu.\\

### 2015-03-08 doc/USER_MANUAL/manual_SPECFEM3D_Cartesian.tex
updated the manual

@@
 ready to compile the internal mesher. This is an alternative to CUBIT
-for the mesh generation of relatively simple geological models. The
-mesher is no longer dedicated to Southern California and more flexiblity
-is provided in this version of the package.
+for the mesh generation of relatively simple models.
 
@@
 file \texttt{Mesh\_Par\_file}, which resides in the subdirectory \texttt{DATA/meshfem3D\_files/}.
+(to see how to use it, see the EXAMPLES specific to the internal mesher in directory \texttt{EXAMPLES/meshfem3D\_examples/}.
 Before running the mesher, a number of parameters need to be set in

### 2015-03-18 doc/USER_MANUAL/manual_SPECFEM3D_Cartesian.tex
added NUMBER_OF_SIMULTANEOUS_RUNS to the users manual

@@
 This feature is only relevant for the solver.
+
+\item [{\texttt{NUMBER\_OF\_SIMULTANEOUS\_RUNS}}] adds the ability to run several calculations (several earthquakes)
+in an embarrassingly-parallel fashion from within the same run;
+this can be useful when using a very large supercomputer to compute
+many earthquakes in a catalog, in which case it can be better from
+a batch job submission point of view to start fewer and much larger jobs,
+each of them computing several earthquakes in parallel.
+To turn that option on, set parameter NUMBER\_OF\_SIMULTANEOUS\_RUNS to a value greater than 1.
+To implement that, we create NUMBER\_OF\_SIMULTANEOUS\_RUNS MPI sub-communicators,
+each of them being labeled "my\_local\_mpi\_comm\_world", and we use them
+in all the routines in "src/shared/parallel.f90", except in MPI\_ABORT() because in that case
+we need to kill the entire run.
+When that option is on, of course the number of processor cores used to start
+the code in the batch system must be a multiple of NUMBER\_OF\_SIMULTANEOUS\_RUNS,
+all the individual runs must use the same number of processor cores,
+which as usual is NPROC in the Par\_file,
+and thus the total number of processor cores to request from the batch system
+should be NUMBER\_OF\_SIMULTANEOUS\_RUNS * NPROC.
+All the runs to perform must be placed in directories called run0001, run0002, run0003 and so on
+(with exactly four digits).
+
+Imagine you have 10 independent calculations to do, each of them on 100 cores; you have three options:
+
+1/ submit 10 jobs to the batch system
+
+2/ submit a single job on 1000 cores to the batch, and in that script create a sub-array of jobs to start 10 jobs,
+each running on 100 cores (see e.g. http://www.schedmd.com/slurmdocs/job\_array.html )
+
+3/ submit a single job on 1000 cores to the batch, start SPECFEM3D on 1000 cores, create 10 sub-communicators,
+cd into one of 10 subdirectories (called e.g. run0001, run0002,... run0010) depending on the sub-communicator
+your MPI rank belongs to, and run normally on 100 cores using that sub-communicator.
+
+The option NUMBER\_OF\_SIMULTANEOUS\_RUNS implements 3/.
+
+\item [{\texttt{BROADCAST\_SAME\_MESH\_AND\_MODEL}}]: 
+if we perform simultaneous runs in parallel, if only the source and receivers vary between these runs
+but not the mesh nor the model (velocity and density) then we can also read the mesh and model files
+from a single run in the beginning and broadcast them to all the others; for a large number of simultaneous
+runs for instance when solving inverse problems iteratively this can DRASTICALLY reduce I/Os to disk in the solver
+(by a factor equal to NUMBER\_OF\_SIMULTANEOUS\_RUNS), and reducing I/Os is crucial in the case of huge runs.
+Thus, always set this option to .true. if the mesh and the model are the same for all simultaneous runs.
+In that case there is no need to duplicate the mesh and model file database (the content of the DATABASES\_MPI
+directories) in each of the run0001, run0002,... directories, it is sufficient to have one in run0001
+and the code will broadcast it to the others).
+
+TODO / future work to do: currently the BROADCAST\_SAME\_MESH\_AND\_MODEL option assumes to have the (master) mesh files in run0001/DATABASES\_MPI or run0001/OUTPUT\_FILES/DATABASES\_MPI.
+However, for adjoint runs you still need a DATABASES\_MPI folder in each of the sub-runs directories, e.g.
+run0002/DATABASES\_MPI, etc.
+to store the forward wavefields, kernels etc.
+of each sub-run.
+This would not be needed for forward simulations.
+
+TODO / future work to do: the sensitivity kernel summing and smoothing tools in directory src/tomography are currently not ported to this new option to do many runs simultaneously, only the solver (src/specfem3d) is.
+Thus these tools should work, but in their current version will need to be run for each simulation result independently.
+More precisely, the current kernel summing and smoothing routines work fine, with the exception that you need to move out the mesh files (and also the parameters).
+This works because these routines consider multiple runs by design.
+You simply have to provide them the directories where the kernels are.
+
 \item [{\texttt{GPU\_MODE}}] Turn this flag on to use GPUs.

### 2015-03-28 doc/USER_MANUAL/manual_SPECFEM3D_Cartesian.tex
Buildbot ran the code beautifier.
added a chapter about inverse problems to the users manual
renamed Roland_Sylvain integrals to gravity integrals

@@
 
+%------------------------------------------------------------------------------------------------%
+
+\chapter{Doing tomography, i.e., updating the model based on the sensitivity kernels obtained}
+
+%------------------------------------------------------------------------------------------------%
+
+UNDER CONSTRUCTION (March 2015). Will be added around April / May 2015.
+
+%------------------------------------------------------------------------------------------------%
+
 \chapter{Noise Cross-correlation Simulations}
+
+%------------------------------------------------------------------------------------------------%
 
@@
 
+%------------------------------------------------------------------------------------------------%
+
+\chapter{Gravity integral calculations for the gravity field of an Earth model}
+
+%------------------------------------------------------------------------------------------------%
+
+SPECFEM can now compute the gravity field as well as its derivatives (i.e., gravity gradiometry)
+generated by any given 3D Earth model at a given elevation, i.e., on a given observation surface.
+That feature is still experimental but should work just fine.\\
+
+To see how it is implemented and to use it, type this in the root directory of the code:\\
+
+\texttt{grep -i GRAVITY\_INTEGRALS src/*/*}
+
+%------------------------------------------------------------------------------------------------%
+
 \chapter{Graphics}
+
+%------------------------------------------------------------------------------------------------%
 

### 2015-04-18 doc/USER_MANUAL/manual_SPECFEM3D_Cartesian.tex
cleaned the demo codes and mentioned them in the manual

@@
 The software package SPECFEM3D Cartesian simulates seismic wave propagation
-at the local or regional scale based upon the spectral-element method
+at the local or regional scale and performs full waveform imaging (FWI) or adjoint tomography based upon the spectral-element method
 (SEM). The SEM is a continuous Galerkin technique \citep{TrKoLi08,PeKoLuMaLeCaLeMaLiBlNiBaTr11},
@@
 
+\chapter{\label{cha:developers}Information for developers of the code, and for people who want to learn how the technique works}
+
+You can get a very simple 1D version of a demo code (there is one in Fortran and one in Python):\\
+
+git clone --recursive https://github.com/geodynamics/specfem1d.git\\
+
+\noindent We also have simple 3D demo source codes that implement the SEM in a single, small program, in directory\\
+utils/small\_SEM\_solvers\_in\_Fortran\_and\_C\_without\_MPI\_to\_learn of the specfem3d package.
+They are useful to learn how the spectral-element method works, and how to write or modify a code to implement it.
+Also useful to test new ideas by modifying these simple codes to run some tests.\\
+
+\noindent For information on how to contribute to the code, i.e., for how to make your modifications, additions or improvements part of the
+official package, see https://github.com/geodynamics/specfem3d/wiki .
+
+%============================================================
+
 \chapter*{\label{cha:Bug-Reports-and}Bug Reports and Suggestions for Improvements}

### 2015-04-18 utils/small_SEM_solvers_in_Fortran_and_C_without_MPI_to_learn/mesher_for_serial/get_ellipticity.f90

@@ -1,0 +1,65 @@
+!=====================================================================
+!
+!          S p e c f e m 3 D  G l o b e  V e r s i o n  4 . 0
+!          --------------------------------------------------
+!
+!          Main authors: Dimitri Komatitsch and Jeroen Tromp
+!    Seismological Laboratory, California Institute of Technology, USA
+!             and University of Pau / CNRS / INRIA, France
+! (c) California Institute of Technology and University of Pau / CNRS / INRIA
+!                            February 2008
+!
+! This program is free software; you can redistribute it and/or modify
+! it under the terms of the GNU General Public License as published by
+! the Free Software Foundation; either version 2 of the License, or
+! (at your option) any later version.
+!
+! This program is distributed in the hope that it will be useful,
+! but WITHOUT ANY WARRANTY; without even the implied warranty of
+! MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+! GNU General Public License for more details.
+!
+! You should have received a copy of the GNU General Public License along
+! with this program; if not, write to the Free Software Foundation, Inc.,
+! 51 Franklin Street, Fifth Floor, Boston, MA 02110-1301 USA.
+!
+!=====================================================================
+
+  subroutine get_ellipticity(xelm,yelm,zelm,nspl,rspl,espl,espl2)
+
+  implicit none
+
+  include "constants.h"
+
+  integer nspl
+  double precision xelm(NGNOD)
+  double precision yelm(NGNOD)
+  double precision zelm(NGNOD)
+  double precision rspl(NR),espl(NR),espl2(NR)
+
+  integer ia
+
+  double precision ell
+  double precision r,theta,phi,factor
+  double precision cost,p20
+
+  do ia=1,NGNOD
+
+  call xyz_2_rthetaphi_dble(xelm(ia),yelm(ia),zelm(ia),r,theta,phi)
+
+  cost=dcos(theta)
+  p20=0.5d0*(3.0d0*cost*cost-1.0d0)
+
+! get ellipticity using spline evaluation
+  call spline_evaluation(rspl,espl,espl2,nspl,r,ell)
+
+  factor=ONE-(TWO/3.0d0)*ell*p20
+
+  xelm(ia)=xelm(ia)*factor
+  yelm(ia)=yelm(ia)*factor
+  zelm(ia)=zelm(ia)*factor
+
+  enddo
+
+  end subroutine get_ellipticity
+

### 2015-04-18 utils/small_SEM_solvers_in_Fortran_and_C_without_MPI_to_learn/older_slower_versions_not_so_good_do_not_use/serial_specfem3D_22dec2008_NGLOB.f90

@@ -1,0 +1,479 @@
+!=====================================================================
+!
+!          S p e c f e m 3 D  G l o b e  V e r s i o n  4 . 0
+!          --------------------------------------------------
+!
+!          Main authors: Dimitri Komatitsch and Jeroen Tromp
+!    Seismological Laboratory, California Institute of Technology, USA
+!             and University of Pau / CNRS / INRIA, France
+! (c) California Institute of Technology and University of Pau / CNRS / INRIA
+!                            February 2008
+!
+! This program is free software; you can redistribute it and/or modify
+! it under the terms of the GNU General Public License as published by
+! the Free Software Foundation; either version 2 of the License, or
+! (at your option) any later version.
+!
+! This program is distributed in the hope that it will be useful,
+! but WITHOUT ANY WARRANTY; without even the implied warranty of
+! MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+! GNU General Public License for more details.
+!
+! You should have received a copy of the GNU General Public License along
+! with this program; if not, write to the Free Software Foundation, Inc.,
+! 51 Franklin Street, Fifth Floor, Boston, MA 02110-1301 USA.
+!
+!=====================================================================
+
+  program serial_specfem3D
+
+  implicit none
+
+!!!!!!!!!!
+!!!!!!!!!! All the arrays below use static memory allocation,
+!!!!!!!!!! using constant sizes defined in values_from_mesher.h.
+!!!!!!!!!! This is done purposely to improve performance (Fortran compilers
+!!!!!!!!!! can optimize much more when the size of the loops and arrays
+!!!!!!!!!! is known at compile time).
+!!!!!!!!!! NGLLX, NGLLY and NGLLZ are set equal to 5,
+!!!!!!!!!! therefore each element contains NGLLX * NGLLY * NGLLZ = 125 points.
+!!!!!!!!!!
+
+!!!!!!!!!!
[…]

### 2015-04-18 utils/solver_classical_serial_Fortran_or_C_no_MPI/mesher_for_serial/get_ellipticity.f90

@@ -1,65 +1,0 @@
-!=====================================================================
-!
-!          S p e c f e m 3 D  G l o b e  V e r s i o n  4 . 0
-!          --------------------------------------------------
-!
-!          Main authors: Dimitri Komatitsch and Jeroen Tromp
-!    Seismological Laboratory, California Institute of Technology, USA
-!             and University of Pau / CNRS / INRIA, France
-! (c) California Institute of Technology and University of Pau / CNRS / INRIA
-!                            February 2008
-!
-! This program is free software; you can redistribute it and/or modify
-! it under the terms of the GNU General Public License as published by
-! the Free Software Foundation; either version 2 of the License, or
-! (at your option) any later version.
-!
-! This program is distributed in the hope that it will be useful,
-! but WITHOUT ANY WARRANTY; without even the implied warranty of
-! MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
-! GNU General Public License for more details.
-!
-! You should have received a copy of the GNU General Public License along
-! with this program; if not, write to the Free Software Foundation, Inc.,
-! 51 Franklin Street, Fifth Floor, Boston, MA 02110-1301 USA.
-!
-!=====================================================================
-
-  subroutine get_ellipticity(xelm,yelm,zelm,nspl,rspl,espl,espl2)
-
-  implicit none
-
-  include "constants.h"
-
-  integer nspl
-  double precision xelm(NGNOD)
-  double precision yelm(NGNOD)
-  double precision zelm(NGNOD)
-  double precision rspl(NR),espl(NR),espl2(NR)
-
-  integer ia
-
-  double precision ell
-  double precision r,theta,phi,factor
-  double precision cost,p20
-
-  do ia=1,NGNOD
-
-  call xyz_2_rthetaphi_dble(xelm(ia),yelm(ia),zelm(ia),r,theta,phi)
-
-  cost=dcos(theta)
-  p20=0.5d0*(3.0d0*cost*cost-1.0d0)
-
-! get ellipticity using spline evaluation
-  call spline_evaluation(rspl,espl,espl2,nspl,r,ell)
-
-  factor=ONE-(TWO/3.0d0)*ell*p20
-
-  xelm(ia)=xelm(ia)*factor
-  yelm(ia)=yelm(ia)*factor
-  zelm(ia)=zelm(ia)*factor
-
-  enddo
-
-  end subroutine get_ellipticity
-

### 2015-04-18 utils/solver_classical_serial_Fortran_or_C_no_MPI/mesher_for_serial/memory_eval.f90

@@ -1,358 +1,0 @@
-!=====================================================================
-!
-!          S p e c f e m 3 D  G l o b e  V e r s i o n  4 . 0
-!          --------------------------------------------------
-!
-!          Main authors: Dimitri Komatitsch and Jeroen Tromp
-!    Seismological Laboratory, California Institute of Technology, USA
-!             and University of Pau / CNRS / INRIA, France
-! (c) California Institute of Technology and University of Pau / CNRS / INRIA
-!                            February 2008
-!
-! This program is free software; you can redistribute it and/or modify
-! it under the terms of the GNU General Public License as published by
-! the Free Software Foundation; either version 2 of the License, or
-! (at your option) any later version.
-!
-! This program is distributed in the hope that it will be useful,
-! but WITHOUT ANY WARRANTY; without even the implied warranty of
-! MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
-! GNU General Public License for more details.
-!
-! You should have received a copy of the GNU General Public License along
-! with this program; if not, write to the Free Software Foundation, Inc.,
-! 51 Franklin Street, Fifth Floor, Boston, MA 02110-1301 USA.
-!
-!=====================================================================
-
-! compute the approximate amount of static memory needed to run the solver
-
-  subroutine memory_eval(OCEANS,ABSORBING_CONDITIONS,ATTENUATION,ANISOTROPIC_3D_MANTLE,&
[…]

### 2015-04-18 utils/solver_classical_serial_Fortran_or_C_no_MPI/mesher_for_serial/model_jp1d.f90

@@ -1,204 +1,0 @@
-!=====================================================================
-!
-!          S p e c f e m 3 D  G l o b e  V e r s i o n  4 . 0
-!          --------------------------------------------------
-!
-!          Main authors: Dimitri Komatitsch and Jeroen Tromp
-!    Seismological Laboratory, California Institute of Technology, USA
-!             and University of Pau / CNRS / INRIA, France
-! (c) California Institute of Technology and University of Pau / CNRS / INRIA
-!                            February 2008
-!
-! This program is free software; you can redistribute it and/or modify
-! it under the terms of the GNU General Public License as published by
-! the Free Software Foundation; either version 2 of the License, or
-! (at your option) any later version.
-!
-! This program is distributed in the hope that it will be useful,
-! but WITHOUT ANY WARRANTY; without even the implied warranty of
-! MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
-! GNU General Public License for more details.
-!
-! You should have received a copy of the GNU General Public License along
-! with this program; if not, write to the Free Software Foundation, Inc.,
-! 51 Franklin Street, Fifth Floor, Boston, MA 02110-1301 USA.
-!
-!=====================================================================
-
-subroutine model_jp1d(myrank,x,rho,vp,vs,Qkappa,Qmu,idoubling, &
-     check_doubling_flag,RICB,RCMB,RTOPDDOUBLEPRIME, &
-     R670,R220,R771,R400,R80,RMOHO,RMIDDLE_CRUST)
-
-  implicit none
-
-  include "constants.h"
-
-  ! given a normalized radius x, gives the non-dimesionalized density rho,
-! speeds vp and vs, and the quality factors Qkappa and Qmu
-
-  logical check_doubling_flag
-  integer idoubling,myrank
-
[…]

### 2015-04-18 utils/solver_classical_serial_Fortran_or_C_no_MPI/mesher_for_serial/rthetaphi_xyz.f90

@@ -1,119 +1,0 @@
-!=====================================================================
-!
-!          S p e c f e m 3 D  G l o b e  V e r s i o n  4 . 0
-!          --------------------------------------------------
-!
-!          Main authors: Dimitri Komatitsch and Jeroen Tromp
-!    Seismological Laboratory, California Institute of Technology, USA
-!             and University of Pau / CNRS / INRIA, France
-! (c) California Institute of Technology and University of Pau / CNRS / INRIA
-!                            February 2008
-!
-! This program is free software; you can redistribute it and/or modify
-! it under the terms of the GNU General Public License as published by
-! the Free Software Foundation; either version 2 of the License, or
-! (at your option) any later version.
-!
-! This program is distributed in the hope that it will be useful,
-! but WITHOUT ANY WARRANTY; without even the implied warranty of
-! MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
-! GNU General Public License for more details.
-!
-! You should have received a copy of the GNU General Public License along
-! with this program; if not, write to the Free Software Foundation, Inc.,
-! 51 Franklin Street, Fifth Floor, Boston, MA 02110-1301 USA.
-!
-!=====================================================================
-
-  subroutine xyz_2_rthetaphi(x,y,z,r,theta,phi)
-
-! convert x y z to r theta phi, single precision call
-
-  implicit none
-
-  include "constants.h"
-
-  real(kind=CUSTOM_REAL) x,y,z,r,theta,phi
-  double precision xmesh,ymesh,zmesh
-
-! distinguish between single and double precision for reals
-  if(CUSTOM_REAL == SIZE_REAL) then
-
-    xmesh = dble(x)
-    ymesh = dble(y)
-    zmesh = dble(z)
-
[…]

### 2015-04-18 utils/solver_classical_serial_Fortran_or_C_no_MPI/serial_specfem3D_single_with_Deville.c

@@ -1,716 +1,0 @@
-
-/*
-!=====================================================================
-!
-!          S p e c f e m 3 D  G l o b e  V e r s i o n  4 . 0
-!          --------------------------------------------------
-!
-!          Main authors: Dimitri Komatitsch and Jeroen Tromp
-!    Seismological Laboratory, California Institute of Technology, USA
-!             and University of Pau / CNRS / INRIA, France
-! (c) California Institute of Technology and University of Pau / CNRS / INRIA
-!                            February 2008
-!
-! This program is free software; you can redistribute it and/or modify
-! it under the terms of the GNU General Public License as published by
-! the Free Software Foundation; either version 2 of the License, or
-! (at your option) any later version.
-!
-! This program is distributed in the hope that it will be useful,
-! but WITHOUT ANY WARRANTY; without even the implied warranty of
-! MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
-! GNU General Public License for more details.
-!
-! You should have received a copy of the GNU General Public License along
-! with this program; if not, write to the Free Software Foundation, Inc.,
-! 51 Franklin Street, Fifth Floor, Boston, MA 02110-1301 USA.
-!
-!=====================================================================
-*/
-
-//
-// All the arrays below use static memory allocation,
-// using constant sizes defined in values_from_mesher.h.
-// This is done purposely to improve performance (Fortran compilers
-// can optimize much more when the size of the loops and arrays
-// is known at compile time).
-// NGLLX, NGLLY and NGLLZ are set equal to 5,
-// therefore each element contains NGLLX * NGLLY * NGLLZ = 125 points.
-//
-
-//
-// All the calculations are done in single precision.
[…]

### 2015-04-18 doc/USER_MANUAL/manual_SPECFEM3D_Cartesian.tex
merged the two tomography sections of the manual

@@
 
-\chapter{Doing tomography, i.e., updating the model based on the sensitivity kernels obtained}
+\chapter{\label{cha:tomo}Doing tomography, i.e., updating the model based on the sensitivity kernels obtained}
 
@@
 UNDER CONSTRUCTION (March 2015). Will be added around April / May 2015.
+
+\section{Tomographic inversion / imaging using the sensitivity kernels obtained}
+
+OLD VERSION OF THE SECTION, WILL SOON BE IMPROVED AND ENRICHED.\\
+
+One of the fundamental reasons for computing sensitivity kernels (Section~\ref{sec:Adjoint-simulation-finite})
+is to use them within a tomographic inversion. In other words, use
+recorded seismograms, make measurements with synthetic seismograms,
+and use the misfit between the two to iteratively improve the model
+described by (at least) $V_{{\rm p}}$, $V_{{\rm s}}$, and $\rho$
+as a function of space.
+
+Whatever misfit function you use for the tomographic inversion (you can find several
+examples in \citet{TrKoLi08} and \citet{TrTaLi05}), you will weight
+the sensitivity kernels with measurements. Furthermore, you will use
+as many measurements (stations, components, time windows) as possible
+per event; hence, we call these composite kernels ``event kernels,''
+which are volumetric fields representing the gradient of the misfit
+function with respect to one of the variables (\eg $V_{{\rm s}}$).
+The basic features of an adjoint-based tomographic inversion were
+illustrated in \citet{TaLiTr07} and \citet{TrKoLi08} using a conjugate-gradient
+algorithm; there are dozens of versions of gradient-based inversion
+algorithms that could alternatively be used. The tomographic inversion
+of \citet{TaLiMaTr09,TaLiMaTr2010} used SPECFEM3D Cartesian as well
+as several additional components which are also stored on the CIG
+svn server, described next.
+
+The directory containing external utilities for tomographic inversion using
+SPECFEM3D Cartesian (or other packages that evaluate misfit functions
+and gradients) is in directory \texttt{utils/ADJOINT\_TOMOGRAPHY\_TOOLS/}:
+\begin{verbatim}
+flexwin/     -- FLEXWIN algorithm for automated picking of time windows
+measure_adj/ -- reads FLEXWIN output file and makes measurements,
+                with the option for computing adjoint sources
+iterate_adj/ -- various tools for iterative inversion
+                (requires pre-computed "event kernels")
+\end{verbatim}
+This directory also contains
+a brief \verb+README+ file indicating the role of the three subdirectories,
+\verb+flexwin+ \citep{Maggi2009}, \verb+measure_adj+, and \verb+iterate_adj+.
+The components for making the model update are there; however, there
+are no explicit rules for computing the model update, just as with
+any optimization problem. There are options for computing a conjugate
+gradient step, as well as a source subspace projection step.
+
+The best single file to read is probably: \verb+ADJOINT_TOMO/iterate_adj/cluster/README+.
 
@@
 
-\chapter{\label{cha:tomo}Tomographic inversion using sensitivity kernels}
-
-One of the fundamental reasons for computing sensitivity kernels (Section~\ref{sec:Adjoint-simulation-finite})
-is to use them within a tomographic inversion. In other words, use
-recorded seismograms, make measurements with synthetic seismograms,
-and use the misfit between the two to iteratively improve the model
-described by (at least) $V_{{\rm p}}$, $V_{{\rm s}}$, and $\rho$
-as a function of space.
-
-Whatever misfit function you use for the tomographic inversion (several
-examples in \citet{TrTaLi05} and \citet{TrKoLi08}), you will weight
-the sensitivity kernels with measurements. Furthermore, you will use
-as many measurements (stations, components, time windows) as possible
-per event; hence, we call these composite kernels ``event kernels,''
-which are volumetric fields representing the gradient of the misfit
-function with respect to one of the variables (\eg $V_{{\rm s}}$).
-The basic features of an adjoint-based tomographic inversion were
-illustrated in \citet{TaLiTr07} and \citet{TrKoLi08} using a conjugate-gradient
-algorithm; there are dozens of versions of gradient-based inversion
-algorithms that could alternatively be used. The tomographic inversion
-of \citet{TaLiMaTr09,TaLiMaTr2010} used SPECFEM3D Cartesian as well
-as several additional components which are also stored on the CIG
-svn server, described next.
-
-The directory containing external utilities for tomographic inversion using
-SPECFEM3D Cartesian (or other packages that evaluate misfit functions
-and gradients) is in directory \texttt{utils/ADJOINT\_TOMOGRAPHY\_TOOLS/}:
-\begin{verbatim}
-flexwin/     -- FLEXWIN algorithm for automated picking of time windows
-measure_adj/ -- reads FLEXWIN output file and makes measurements,
-                with the option for computing adjoint sources
-iterate_adj/ -- various tools for iterative inversion
-                (requires pre-computed "event kernels")
-\end{verbatim}
-This directory also contains
-a brief \verb+README+ file indicating the role of the three subdirectories,
-\verb+flexwin+ \citep{Maggi2009}, \verb+measure_adj+, and \verb+iterate_adj+.
-The components for making the model update are there; however, there
-are no explicit rules for computing the model update, just as with
-any optimization problem. There are options for computing a conjugate
-gradient step, as well as a source subspace projection step. The workflow
-could use substantial ``stitching together,'' and we welcome suggestions;
-undoubtedly, the tomographic inversion procedure will improve with
-time. \textbf{This is a work in progress.}
-
-The best single file to read is probably: \verb+ADJOINT_TOMO/iterate_adj/cluster/README+.
-
 %============================================================

### 2015-04-19 doc/USER_MANUAL/manual_SPECFEM3D_Cartesian.tex
added a FWI section to the users manual

@@
 \newcommand{\red}[1]{\textcolor{Red}{#1}}
+
+% some of Dimitri's aliases for math symbols
+\newcommand{\Grad}{\mbox{\boldmath $\bf \nabla$}}
+\newcommand{\Div}{\Grad \cdot }
+\newcommand{\Laplacian}{\nabla^2}
+\newcommand{\pressure}{p}
+\newcommand{\Int}{\displaystyle\int}
+\newcommand{\Sum}{\displaystyle\sum}
+\renewcommand{\epsilon}{\varepsilon}
+\newcommand{\myiomega}{\mathbf{i} \, \omega}
+\newcommand{\p}{\partial}
+\newcommand{\boldu}{\mathbf{u}}
+\newcommand{\boldv}{\mathbf{v}}
+\newcommand{\boldw}{\mathbf{w}}
+\newcommand{\boldx}{\mathbf{x}}
+\newcommand{\boldy}{\mathbf{y}}
+\newcommand{\boldm}{\mathbf{m}}
+\newcommand{\boldp}{\mathbf{p}}
+\newcommand{\boldg}{\mathbf{g}}
+\newcommand{\bolds}{\mathbf{s}}
+\newcommand{\boldn}{\mathbf{\hat{n}}}
+\newcommand{\boldd}{\mathbf{d}}
+\newcommand{\boldf}{\mathbf{f}}
+\newcommand{\boldB}{\mathbf{B}}
+\newcommand{\boldc}{\mathbf{c}}
+\newcommand{\boldC}{\mathbf{C}}
+\newcommand{\boldE}{\mathbf{E}}
+\newcommand{\boldH}{\mathbf{H}}
+\newcommand{\boldI}{\mathbf{I}}
+\newcommand{\boldsigma}{\mbox{\boldmath $\bf \sigma$}}
+\newcommand{\boldxi}{\mbox{\boldmath $\bf \xi$}}
+\newcommand{\boldvarepsilon}{\mbox{\boldmath $\bf \varepsilon$}}
+\newcommand{\bolddelta}{\mbox{\boldmath $\bf \delta$}}
+\newcommand{\boldtau}{\mbox{\boldmath $\bf \tau$}}
+\newcommand{\Boxmath}{\Lambda}
+\newcommand{\rmd}{\, \mathrm{d}}
+\newcommand{\pv}{\ ,}
+\newcommand{\pp}{\ .}
 
@@
 
+\item [{\texttt{USE\_FAILSAFE\_MECHANISM}}]:
+if one or a few of these simultaneous runs fail, kill all the runs or let the others finish using a fail-safe mechanism
+(in most cases, should be set to true).
+
 TODO / future work to do: currently the BROADCAST\_SAME\_MESH\_AND\_MODEL option assumes to have the (master) mesh files in run0001/DATABASES\_MPI or run0001/OUTPUT\_FILES/DATABASES\_MPI.
@@
 
-\section{Post-processing and Visualization }
+\section{Post-processing and Visualization}
 
@@
 
-\chapter{\label{cha:tomo}Doing tomography, i.e., updating the model based on the sensitivity kernels obtained}
+\chapter[Doing tomography based on the sensitivity kernels obtained]{\label{cha:tomo}Doing tomography, i.e., updating the model based on the sensitivity kernels obtained}
 
@@
 
-UNDER CONSTRUCTION (March 2015). Will be added around April / May 2015.
-
-\section{Tomographic inversion / imaging using the sensitivity kernels obtained}
-
-OLD VERSION OF THE SECTION, WILL SOON BE IMPROVED AND ENRICHED.\\
+UNDER CONSTRUCTION (April 2015). Will be added around June 2015.
+
+\section{Tomographic full waveform inversion (FWI) / imaging using the sensitivity kernels obtained}
 
 One of the fundamental reasons for computing sensitivity kernels (Section~\ref{sec:Adjoint-simulation-finite})
-is to use them within a tomographic inversion. In other words, use
+is to use them within a tomographic inversion, i.e., perform imaging. In other words, use
 recorded seismograms, make measurements with synthetic seismograms,
 and use the misfit between the two to iteratively improve the model
-described by (at least) $V_{{\rm p}}$, $V_{{\rm s}}$, and $\rho$
-as a function of space.
+described by (at least) $V_{{\rm p}}$, $V_{{\rm s}}$, and $\rho$ as a function of space.
+
+As explained for instance in \cite{MoChKoWa15},
+full waveform inversion means that one considers the observed seismograms (possibly filtered) as the
+basic observables that one wants to fit. One thus searches for the model that minimizes the mean squared difference between observed
+and synthetic seismograms. In other words, the goal is to find a structural model that can explain a larger portion
+of seismological records, and not simply the phase of a few seismic arrivals.
 
@@
 as many measurements (stations, components, time windows) as possible
-per event; hence, we call these composite kernels ``event kernels,''
+per event; hence, we call these composite kernels ``event kernels'',
 which are volumetric fields representing the gradient of the misfit
@@
 The basic features of an adjoint-based tomographic inversion were
-illustrated in \citet{TaLiTr07} and \citet{TrKoLi08} using a conjugate-gradient
-algorithm; there are dozens of versions of gradient-based inversion
+illustrated in \citet{TrKoLi08} and \citet{TaLiTr07} using a conjugate-gradient algorithm.
+Other, more powerful techniques such as the Limited-Memory Broyden-Fletcher-Goldfarb-Shanno (L-BFGS) method can now be used,
+as illustrated for instance in \cite{MoChKoWa15}.
+
+\subsection{Principle}
+
+We want to minimize the classical waveform misfit function:
+\begin{equation}
+\chi \left( \boldm \right) = \sum_{s=1}^N\sum_{r=1}^M \int_0^T \frac{1}{2} \|\boldu(\boldx_r, \boldx_s; t) - \boldd(\boldx_r, \boldx_s; t)\| ^{2} \, dt .
+\label{misfit}
+\end{equation}
+This functional quantifies the $L^2$ difference between the observed waveforms $\boldd(\boldx_r, \boldx_s; t)$ at receivers $\boldx_r$, $r = 1, ..., M$
+produced by sources at $\boldx_s$, $s=1,...,N$, and the corresponding synthetic seismograms $\boldu(\boldx_r, \boldx_s; t)$ computed in model $\boldm$.
+While this misfit function is indeed classical, it is worth mentioning that in the case of noisy real data other norms could be used,
+since in the oil industry for instance it is known that the $L^1$ norm
+(\cite{CrPiNoMcTa90,Brossier_2010_WDR}), hybrid $L^1-L^2$ norms (\cite{Bube_1997_HL1}), Hubert norm (\cite{Ha_2009_WIU}),
+Student-t distribution (\cite{ArVaHe11,Jeong_2015}) etc... can be more robust that the $L^2$ norm used here in the context of synthetic data with no noise.
+In the vicinity of $\boldm$, the misfit function can be expanded into a Taylor series:
+\begin{equation}
+\chi(\boldm+ \delta\boldm) \approx \chi(\boldm)+\boldg(\boldm)\cdot\delta\boldm
++\delta\boldm\cdot\mathbf{H}(\boldm)\cdot\delta\boldm \, ,
+\label{local_misfit}
+\end{equation}
+where $\boldg(\boldm)$ is the gradient of the waveform misfit function:
+\begin{equation}
+\boldg(\boldm) = \frac{\partial\chi(\boldm)}{\partial\boldm} \, ,
+\end{equation}
+and $\mathbf{H}(\boldm)$ the Hessian:
+\begin{equation}
+\mathbf{H}(\boldm) = \frac{\partial^2\chi(\boldm)}{\partial\boldm^2} \, .
+\end{equation}
+In the following, for simplicity the dependence of the gradient and Hessian on the model will be implicitly assumed and omitted in the notations.
+The nearest minimum of $\chi$ in (\ref{local_misfit}) with respect to the model perturbation $\delta \boldm$ is reached for
+\begin{equation}
+\delta \boldm = -\mathbf{H}^{-1} \cdot \boldg \, .
+\label{Newton}
+\end{equation}
+The local minimum of (\ref{misfit}) is thus given by perturbing the model in the direction of the gradient preconditioned
+by the inverse Hessian.
+
+\subsection{Computation of the gradient based on the adjoint method}
+
+A direct method to compute the gradient is to take the derivative of (\ref{misfit}) with respect to model parameters:
+\begin{equation}
+\frac{\partial\chi(\boldm)}{\partial\boldm} = -\sum_{s=1}^N\sum_{r=1}^M \int_0^T
+\frac{\partial \boldu(\boldx_r, \boldx_s; t)}{\partial\boldm}\cdot\left[\boldu(\boldx_r, \boldx_s; t) - \boldd(\boldx_r, \boldx_s; t)\right] \, dt \, .
+\label{calc_gradient}
+\end{equation}
+This equation can be reformulated as the matrix-vector product:
+\begin{equation}
+\boldg = -\mathbf{J}^* \cdot \delta \mathbf{d} \, ,
+\label{Jacobian}
+\end{equation}
+where $\mathbf{J}^*$ is the adjoint of the Jacobian matrix of the forward problem that contains the Fr\'echet derivatives of the
+data with respect to model parameters, and $\delta \mathbf{d}$ is the vector that contains the data residuals.
+The determination of $\mathbf{J}$ would require computing the Fr\'echet derivatives for each time step in the time
+window considered and for all the source-station pairs, which is completely prohibitive on current supercomputers
+(let us note that this situation may change one day).
+However, it is possible to obtain this gradient without computing the Jacobian matrix explicitly.
+The approach to determine the gradient without computing the Fr\'echet derivatives was introduced in nonlinear optimization
+by \cite{Cha74} working with J. L. Lions, and later applied to seismic exploration problems by
+\cite{BaChLa77}, \cite{BaChHeLa82}, \cite{Lai83} and \cite{Tar84}. The idea is to resort to the adjoint state,
+which corresponds to the wavefield emitted and back-propagated from the receivers \cite[e.g.,][]{TrTaLi05,TrKoLi08,Plessix_2006_RAS,FiBuIg06}.
+
+Let us give an outline of the theory to compute the gradient based on the adjoint method, and refer the reader to e.g. \cite{TrTaLi05}
+and \cite{TrKoLi08} for further details. The perturbation of the misfit function can be expressed as:
+\begin{equation}
+\delta \chi (\boldm) = \sum_{s=1}^N\sum_{r=1}^M \int_0^T \left[\boldu(\boldx_r, \boldx_s; t) - \boldd(\boldx_r, \boldx_s; t)\right]\cdot \delta
+\boldu(\boldx_r, \boldx_s; t) \, dt \, ,
+\label{dchi}
+\end{equation}
+where $\delta\boldu$ is the perturbation of displacement given by the first-order Born approximation \cite[e.g.,][]{Hud77}:
+\begin{eqnarray}
+\delta \boldu (\boldx_r, \boldx_s; t) & = & -\int_0^t \int_V
+\left[\delta\rho(\boldx)\mathbf{G}(\boldx_r,\boldx;t-t')\cdot\partial^2_{t'}\boldu(\boldx,\boldx_s;t') \right. \nonumber \\
+& + & \left. \nabla \mathbf{G}(\boldx_r,\boldx;t-t'):\delta \boldc(\boldx): \nabla\boldu(\boldx;t') \right] \, d^3\boldx \, dt' \, .
+\label{Born}
+\end{eqnarray}
+In this expression, $\mathbf{G}$ is the Green's tensor, $\delta\rho$ the perturbation of density, $\delta c$ the perturbation of the fourth-order elasticity
+tensor,
+and a colon denotes a double tensor contraction operation.
+Inserting (\ref{Born}) into (\ref{dchi}) we obtain
+\begin{eqnarray}
+\delta \chi (\boldm) & = & -\sum_{s=1}^N\sum_{r=1}^M \int_0^T \left[\boldu(\boldx_r, \boldx_s; t) - \boldd(\boldx_r, \boldx_s; t)\right] \int_0^t \int_V
+\left[\delta\rho(\boldx)\mathbf{G}(\boldx_r,\boldx;t-t')\cdot\partial^2_{t'}\boldu(\boldx,\boldx_s;t') \right. \nonumber \\
+& + & \left. \nabla \mathbf{G}(\boldx_r,\boldx;t-t'):\delta \boldc(\boldx):\nabla\boldu(\boldx,t') \right] \, d^3\boldx \, dt' \, dt \,.
+\end{eqnarray}
+Defining the waveform adjoint source for each source $\boldx_s$
+\begin{equation}
+\boldf^\dagger(\boldx,\boldx_s;t) = \sum_{r=1}^M  \left[\boldu(\boldx_r, \boldx_s; T-t) - \boldd(\boldx_r, \boldx_s; T-t)\right] \delta(\boldx-\boldx_r) \, ,
+\end{equation}
+and the corresponding adjoint wavefield
+\begin{equation}
+\boldu^\dagger(\boldx,\boldx_s;t) = \int_0^{t'}\int_V\mathbf{G}(\boldx,\boldx';t'-t)\cdot\boldf^\dagger(\boldx',\boldx_s;t)\, d^3\boldx' \, dt \, ,
+\end{equation}
+the perturbation of the misfit function may be expressed as:
+\begin{eqnarray}
+\delta\chi(\boldm) & = & - \sum_{s=1}^N\nonumber \int_V \int_0^T \left[\delta\rho \, \boldu^\dagger(\boldx,\boldx_s;T-t)\cdot
+\partial^2_{t}\boldu(\boldx,\boldx_s;t) \right. \nonumber \\
+& +  & \left. \nabla\boldu^\dagger(\boldx,\boldx_s;T-t):\delta \boldc: \nabla \boldu(\boldx,\boldx_s;t) \right] \, d^3\boldx \, dt \, .
+\label{dchi2}
+\end{eqnarray}
+At this point, we make some assumptions on the nature of the elasticity tensor.
+A general fourth-order elasticity tensor is described by 21 elastic parameters, a very large number that makes its complete characterization
+way beyond the reach of any tomographic approach. For the time being, let us thus consider isotropic elasticity tensors,
+described by the two Lam\'e parameters $\lambda$ and $\mu$:
+\begin{equation}
+c_{ijkl} = \lambda \delta_{ij}\delta_{kl}+\mu(\delta_{ik}\delta_{jl}+\delta_{il}\delta_{jk}) \, .
+\end{equation}
+In this case, (\ref{dchi2}) can be written as:
+\begin{eqnarray}
+\delta\chi(\boldm) & = & - \sum_{s=1}^N\nonumber \int_V \left[ K_\rho(\boldx,\boldx_s)\delta\ln\rho(\boldx)
++K_\lambda(\boldx,\boldx_s)\delta\ln\lambda(\boldx) \right. \nonumber \\
+& + & \left. K_\mu(\boldx,\boldx_s)\delta\ln\mu(\boldx)\right] \, d^3\boldx \, ,
+\label{dchi2_1}
+\end{eqnarray}
+where ln() is the natural logarithm and where the Fr\'echet derivatives with respect to the density and Lam\'e parameters are given by:
+\begin{eqnarray}
+K_\rho(\boldx,\boldx_s) & = & - \int_0^T \rho(\boldx) \boldu^\dagger(\boldx,\boldx_s;T-t)\cdot \partial^2_{t}\boldu(\boldx,\boldx_s;t) \, dt \\
+K_\lambda(\boldx,\boldx_s) & = & - \int_0^T \lambda(\boldx) \nabla\cdot\boldu^\dagger(\boldx,\boldx_s;T-t) \nabla\cdot\boldu(\boldx,\boldx_s;t) \, dt \\
+K_\mu(\boldx,\boldx_s) & = & -2 \int_0^T \mu(\boldx) (\boldx) \nabla\boldu^\dagger(\boldx,\boldx_s;T-t) :\nabla\boldu(\boldx,\boldx_s;t)\, dt \, .
+\end{eqnarray}
+Since the propagation of seismic waves mainly depends on compressional wave speed $\alpha$ and shear wave speed $\beta$,
+but also because these seismic velocities are easier to interpret, tomographic models are usually described based on these
+two parameters. With this new parametrisation, the perturbation of the misfit function may be written as:
+\begin{eqnarray}
+\delta\chi(\boldm) & = & - \sum_{s=1}^N\nonumber \int_V \left[ K'_\rho(\boldx,\boldx_s)\delta\ln\rho(\boldx)
++K'_\alpha(\boldx,\boldx_s)\delta\ln\alpha(\boldx) \right. \nonumber \\
+& + & \left. K'_\beta(\boldx,\boldx_s)\delta\ln\beta(\boldx)\right] \, d^3\boldx \, ,
+\end{eqnarray}
+where
+\begin{eqnarray}
+K'_\rho(\boldx,\boldx_s) & = &  K_\rho(\boldx,\boldx_s)+K_\lambda(\boldx,\boldx_s)+K_\mu(\boldx,\boldx_s) \\
+K'_\alpha(\boldx,\boldx_s) & = &  2 \left(\frac{\lambda+2\mu}{\lambda}\right) K_\lambda(\boldx,\boldx_s) \\
+K'_\beta(\boldx,\boldx_s) & = & 2 K_\mu -\frac{4\mu}{\lambda} K_\lambda (\boldx,\boldx_s) \, .
+\end{eqnarray}
+
+As can be seen from these expressions, the principle of the adjoint-state method is to correlate two wavefields:
+the direct (i.e. forward) field that propagates from the source to the receivers, and the adjoint field that propagates from all the
+receivers backward in time. The same approach can be followed for any type of seismic observable (phase, amplitude,
+envelope, time series...), provided the appropriate adjoint source is used \cite[]{TrTaLi05,TrKoLi08}.
+For example, for the cross-correlated traveltime of a seismic
+phase, the adjoint source is defined as the velocity of that synthetic phase weighted by the travel-time residual.
+
+Computing the gradient based on the adjoint-state method requires performing two simulations per source (forward and adjoint
+fields) regardless of the type of observable. However, to define the adjoint field one
+must know the adjoint source, and that source is computed from the results of the forward simulation. One must therefore perform
+the forward simulation before the adjoint simulation.
+A straightforward solution for time-domain methods would be to store the whole forward field to disk at
+each time step during the forward run and then read it back during the adjoint simulation to calculate the interaction of these two fields.
+In 2-D this is feasible but in the 3-D case for very short seismic periods and without lossy compression, downsampling,
+or a large amount of disk or memory checkpointing \cite[e.g.,][]{FiKeIg09,RuHaPuGu14,CyShWi15}
+the amount of disk storage required would currently be too large.
+However let us note again that this situation will change in the future.
+In the mean time, a standard possible solution is to perform three simulations per source \cite[]{TrKoLi08,PeKoLuMaLeCaLeMaLiBlNiBaTr11},
+i.e., perform the forward calculation twice, once to compute the adjoint sources and once again
+at the same time as the adjoint simulation to correlate the two fields and sum their interaction on the fly over all the time steps.
+Doing so for an elastic Earth, one only needs a small amount of disk storage to store the last time
+step of the forward run, which is then used as an initial condition to redo the forward run backwards,
+as well as the field on the outer edges of the mesh for each time steps in order to be able to undo the absorbing boundary conditions.
+
+\section{OLD VERSION OF THE SECTION, WILL SOON BE IMPROVED AND ENRICHED}
+
+There are dozens of versions of gradient-based inversion
 algorithms that could alternatively be used. The tomographic inversion

### 2015-04-24 doc/USER_MANUAL/02_getting_started.tex
updates user manual to use facilitate the generation of a markdown version to the github wiki

--- 2015-04-18 doc/USER_MANUAL/manual_SPECFEM3D_Cartesian.tex
@@
-% !TEX encoding = IsoLatin
-%% LyX 2.0.2 created this file.  For more info, see http://www.lyx.org/.
-%% Do not edit unless you really know what you are doing.
-\documentclass[oneside,english]{book}
-\usepackage[T1]{fontenc}
-%% DK DK we do not accept accented letters, use the old LaTeX syntax
-%% DK DK for them instead: \'e \`e  \"e \^e and so on.
-%% DK DK \usepackage[latin1]{inputenc}
-\usepackage[letterpaper]{geometry}
-\geometry{verbose,tmargin=1in,bmargin=1in,lmargin=1in,rmargin=1in}
-\setcounter{secnumdepth}{3}
-\setcounter{tocdepth}{3}
-\usepackage{babel}
-\usepackage{float}
-\usepackage{textcomp}
-\usepackage{url}
-\usepackage{amsmath}
-\usepackage{amssymb}
-\usepackage{graphicx}
-\usepackage[authoryear]{natbib}
-\usepackage[unicode=true,pdfusetitle,
- bookmarks=true,bookmarksnumbered=true,bookmarksopen=false,
- breaklinks=false,pdfborder={0 0 0},backref=section,colorlinks=false,pdfpagemode=None]
- {hyperref}
-\hypersetup{
- pdftex,pdfstartview=FitH,pdfpagelayout=SinglePage}
-\usepackage{breakurl}
-
-\makeatletter
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% LyX specific LaTeX commands.
-\newcommand{\noun}[1]{\textsc{#1}}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Textclass specific LaTeX commands.
-\newenvironment{lyxcode}
-{\par\begin{list}{}{
-\setlength{\rightmargin}{\leftmargin}
-\setlength{\listparindent}{0pt}% needed for AMS classes
-\raggedright
-\setlength{\itemsep}{0pt}
-\setlength{\parsep}{0pt}
-\normalfont\ttfamily}%
- \item[]}
-{\end{list}}
-\newenvironment{lyxlist}[1]
-{\begin{list}{}
-{\settowidth{\labelwidth}{#1}
- \setlength{\leftmargin}{\labelwidth}
- \addtolength{\leftmargin}{\labelsep}
- \renewcommand{\makelabel}[1]{##1\hfil}}}
-{\end{list}}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% User specified LaTeX commands.
-%\renewcommand{\baselinestretch}{1.5}
-
-% figures
-\usepackage[dvips]{epsfig}
-
-% fonts
-\usepackage{times}
-
-% hyperlinks to sections and references
-
-\newcommand{\urlwithparentheses}[1]{(\url{#1})}
-
-% biblio GJI
-\bibliographystyle{abbrvnat}
-
-\newcommand{\nexxi}{\mbox{\texttt{NEX\_XI\/}}}
-\newcommand{\nexeta}{\mbox{\texttt{NEX\_ETA\/}}}
-\newcommand{\nprocxi}{\mbox{\texttt{NPROC\_XI\/}}}
-\newcommand{\nproceta}{\mbox{\texttt{NPROC\_ETA\/}}}
-\newcommand{\nchunks}{\mbox{\texttt{NCHUNKS\/}}}
-\newcommand{\eg}{e.g.,}
-
-% colors for the text
-\@ifundefined{definecolor}
- {\usepackage{color}}{}
-\newcommand{\red}[1]{\textcolor{Red}{#1}}
-
-\usepackage{babel}
-
-\makeatother
-
-\begin{document}
-\begin{center}
-\thispagestyle{empty}\vspace*{-1.8truecm} %
-\makebox[1\textwidth]{%
-\includegraphics[width=0.83\paperwidth]{figures/specfem_3d_Cartesian-cover} %
-}
-\par\end{center}
-
-\title{\textbf{SPECFEM3D Cartesian}\\
- \textbf{User Manual}}
-
-\author{$\copyright$ Princeton University (USA), CNRS and University of Marseille (France),\\
-and ETH Z\"urich (Switzerland)\\
-Version 3.0}
-
-\date{\today}
-
-\maketitle
-
-\section*{Authors}
-
-\noindent The SPECFEM3D Cartesian package was first developed by Dimitri
-Komatitsch and Jean-Pierre Vilotte at Institut de Physique du Globe
-(IPGP) in Paris, France from 1995 to 1997 and then by Dimitri Komatitsch
-and Jeroen Tromp at Harvard University and Caltech, USA, starting in 1998.
-The story started on April 4, 1995, when Prof. Yvon Maday from CNRS and University of Paris, France, gave a lecture to
-Dimitri Komatitsch and Jean-Pierre Vilotte at IPG about the nice properties of the Legendre spectral-element method with diagonal mass matrix that he had used for
-other equations. We are deeply indebted and thankful to him for that.
-That followed a visit by Dimitri Komatitsch to OGS (Istituto Nazionale di Oceanografia e di Geofisica Sperimentale) in Trieste, Italy, in February 1995
-to meet with G\'eza Seriani and Enrico Priolo, who introduced him to their 2D Chebyshev version of the spectral-element method with a non-diagonal mass matrix.
-We are deeply indebted and thankful to them for that.
-
-%%%%%%%% please do NOT merge the lines of the paragraph below, it makes it difficult to insert new names in alphabetical order; leave a single name per line
-%%%%%%%% please do NOT merge the lines of the paragraph below, it makes it difficult to insert new names in alphabetical order; leave a single name per line
-%%%%%%%% please do NOT merge the lines of the paragraph below, it makes it difficult to insert new names in alphabetical order; leave a single name per line
-Since then it has been developed and maintained by a development team:
-in alphabetical order,
-Jean-Paul (Pablo) Ampuero,
-Kangchen Bai,
-Piero Basini,
-C\'eline Blitz,
-Alexis Bottero,
-Ebru Bozda\u{g},
-Emanuele Casarotti,
-Joseph Charles,
-Min Chen,
-Paul Cristini,
-Percy Galvez,
-Dominik G\"oddeke,
-Vala Hj\"orleifsd\'ottir,
-Sue Kientz,
-Dimitri Komatitsch,
-Jes\'us Labarta,
-Nicolas Le Goff,
-Pieyre Le Loher,
-Matthieu Lefebvre,
-Qinya Liu,
-Yang Luo,
-Alessia Maggi,
-Federica Magnoni,
-Roland Martin,
-Ren\'e Matzen,
-Dennis McRitchie,
-Matthias Meschede,
-Peter Messmer,
-David Mich\'ea,
-Surendra Nadh Somala,
-Tarje Nissen-Meyer,
-Daniel Peter,
-Max Rietmann,
-Elliott Sales de Andrade,
-Brian Savage,
-Bernhard Schuberth,
-Anne Sieminski,
-Leif Strand,
-Carl Tape,
-Jeroen Tromp,
-Jean-Pierre Vilotte,
-Zhinan Xie,
-Chang-Hua Zhang,
-Hejun Zhu.\\
-
-The cover graphic of the manual was created by Santiago Lombeyda from
-Caltech's Center for Advanced Computing Research (CACR) \urlwithparentheses{http://www.cacr.caltech.edu}.\\
-
-The code is released open-source under the GNU version 2+ license, see the license at the end of this manual.\\
-
-\newpage{}
-\section*{Current and past main participants or main sponsors of the SPECFEM project (in no particular order)}
-%
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=0.125\textwidth]{figures/logo_cnrs}\vspace*{2truemm}
-\includegraphics[width=0.125\textwidth]{figures/logo_princeton}\vspace*{2truemm}
-\includegraphics[width=0.15\textwidth]{figures/logo_aix_marseille_universite}\vspace*{0.02truemm}
-\includegraphics[width=0.15\textwidth]{figures/logo_ETH}\vspace*{2truemm}
-\includegraphics[width=0.12\textwidth]{figures/logo_CSC_China}\vspace*{0.02truemm}
-\includegraphics[width=0.15\textwidth]{figures/logo_inria}\vspace*{2truemm}
-\includegraphics[width=0.14\textwidth]{figures/logo_UPPA}
-\par\end{centering}
-%
-\vspace*{-2truemm}
-%
-\noindent \begin{centering}
-\includegraphics[width=0.15\textwidth]{figures/logo_NVIDIA}\vspace*{2truemm}
-\includegraphics[width=0.125\textwidth]{figures/logo_IUF}\vspace*{2truemm}
-\includegraphics[width=0.135\textwidth]{figures/logo_Caltech}\vspace*{2truemm}
-\includegraphics[width=0.135\textwidth]{figures/logo_Harvard}\vspace*{2truemm}
-\includegraphics[width=0.135\textwidth]{figures/logo_IPGP}\vspace*{2truemm}
-\includegraphics[width=0.112\textwidth]{figures/logo_ANR}\vspace*{2truemm}
-\includegraphics[width=0.112\textwidth]{figures/logo_NSF}\vspace*{2truemm}
-\par\end{centering}
-%
-\vspace*{-2truemm}
-%
-\noindent \begin{centering}
-\includegraphics[width=0.190\textwidth]{figures/logo_European_Union}\vspace*{2truemm}
-\includegraphics[width=0.112\textwidth]{figures/logo_GENCI}\vspace*{2truemm}
-\includegraphics[width=0.112\textwidth]{figures/logo_PRACE}\vspace*{2truemm}
-\includegraphics[width=0.112\textwidth]{figures/logo_CINES}\vspace*{2truemm}
-\includegraphics[width=0.130\textwidth]{figures/logo_Oak_Ridge}\vspace*{2truemm}
-\hspace*{3mm}\includegraphics[width=0.130\textwidth]{figures/logo_fondation_Del_Duca}
-\includegraphics[width=0.130\textwidth]{figures/logo_CIG}\vspace*{2truemm}
-\par\end{centering}
-%
-\vspace*{-2truemm}
-%
-\noindent \begin{centering}
-\includegraphics[width=0.112\textwidth]{figures/logo_University_of_Toronto}\vspace*{2truemm}
-\includegraphics[width=0.109\textwidth]{figures/logo_INGV}\vspace*{2truemm}
-\includegraphics[width=0.102\textwidth]{figures/logo_Univ_Toulouse}\vspace*{2truemm}
-\includegraphics[width=0.112\textwidth]{figures/logo_TOTAL}\vspace*{2truemm}
-\includegraphics[width=0.130\textwidth]{figures/logo_Fairbanks}\vspace*{2truemm}
-\includegraphics[width=0.120\textwidth]{figures/logo_CINECA}\vspace*{2truemm}
-\includegraphics[width=0.140\textwidth]{figures/logo_Intel_Exascale_Labs}\vspace*{2truemm}
-\includegraphics[width=0.130\textwidth]{figures/logo_Maison_Simulation}
-\par\end{centering}
-\end{figure}
-
-\newpage{}
-
-\tableofcontents{}
-
-\chapter{Introduction}
-
-The software package SPECFEM3D Cartesian simulates seismic wave propagation
-at the local or regional scale and performs full waveform imaging (FWI) or adjoint tomography based upon the spectral-element method
-(SEM). The SEM is a continuous Galerkin technique \citep{TrKoLi08,PeKoLuMaLeCaLeMaLiBlNiBaTr11},
-which can easily be made discontinuous \citep{BeMaPa94,Ch00,KoWoHu02,ChCaVi03,LaWaBe05,Kop06,WiStBuGh10,AcKo11};
-it is then close to a particular case of the discontinuous Galerkin
-technique \citep{ReHi73,LeRa74,Arn82,JoPi86,BoMaHe91,FaRi99,HuHuRa99,CoKaSh00,GiHeWa02,RiWh03,MoRi05,GrScSc06,AiMoMu06,BeLaPi06,DuKa06,DeSeWh08,PuAmKa09,WiStBuGh10,DeSe10,EtChViGl10},
-with optimized efficiency because of its tensorized basis functions
-\citep{WiStBuGh10,AcKo11}. In particular, it can accurately handle
-very distorted mesh elements \citep{OlSe11}.\\
-
-It has very good accuracy and convergence properties \citep{MaPa89,SePr94,DeFiMu02,Coh02,DeSe07,SeOl08,AiWa09,AiWa10,MeStTh12}.
-The spectral element approach admits spectral rates of convergence
-and allows exploiting $hp$-convergence schemes. It is also very well
-suited to parallel implementation on very large supercomputers \citep{KoTsChTr03,TsKoChTr03,KoLaMi08a,CaKoLaTiMiLeSnTr08,KoViCh10}
-as well as on clusters of GPU accelerating graphics cards \citep{Kom11,MiKo10,KoMiEr09,KoErGoMi10}.
-Tensor products inside each element can be optimized to reach very
-high efficiency \citep{DeFiMu02}, and mesh point and element numbering
-can be optimized to reduce processor cache misses and improve cache
-reuse \citep{KoLaMi08a}. The SEM can also handle triangular (in 2D)
-or tetrahedral (in 3D) elements \citep{WinBoyd96,TaWi00,KoMaTrTaWi01,Coh02,MeViSa06}
-as well as mixed meshes, although with increased cost and reduced
-accuracy in these elements, as in the discontinuous Galerkin method.\\
-
-Note that in many geological models in the context of seismic wave
-propagation studies (except for instance for fault dynamic rupture
-studies, in which very high frequencies or supershear rupture need
-to be modeled near the fault, see e.g. \citet{BeGlCrViPi07,BeGlCrVi09,PuAmKa09,TaCrEtViBeSa10})
-a continuous formulation is sufficient because material property contrasts
-are not drastic and thus conforming mesh doubling bricks can efficiently
-handle mesh size variations \citep{KoTr02a,KoLiTrSuStSh04,LeChLiKoHuTr08,LeChKoHuTr09,LeKoHuTr09}.\\
-
-For a detailed introduction to the SEM as applied to regional seismic
-wave propagation, please consult \citet{PeKoLuMaLeCaLeMaLiBlNiBaTr11,TrKoLi08,KoVi98,KoTr99,ChKoViCaVaFe07}
-and in particular \citet{LeKoHuTr09,LeChKoHuTr09,LeChLiKoHuTr08,GoAmTaCaSmSaMaKo09,WiKoScTr04,KoLiTrSuStSh04}.
-A detailed theoretical analysis of the dispersion
-and stability properties of the SEM is available in \citet{Coh02}, \citet{DeSe07}, \citet{SeOl07}, \citet{SeOl08} and \citet{MeStTh12}.\\
-
-Effects due to lateral variations in compressional-wave speed, shear-wave
-speed, density, a 3D crustal model, topography and bathymetry are
-included. The package can accommodate full 21-parameter anisotropy
-(see~\citet{ChTr07}) as well as lateral variations in attenuation
-\citep{SaKoTr10}. Adjoint capabilities and finite-frequency kernel
-simulations are included \citep{TrKoLi08,PeKoLuMaLeCaLeMaLiBlNiBaTr11,LiTr06,FiIgBuKe09,ViOp09}.\\
-
-The SEM was originally developed in computational fluid dynamics \citep{Pat84,MaPa89}
-and has been successfully adapted to address problems in seismic wave
-propagation. Early seismic wave propagation applications of the SEM,
-utilizing Legendre basis functions and a perfectly diagonal mass matrix,
-include \citet{CoJoTo93}, \citet{Kom97}, \citet{FaMaPaQu97}, \citet{CaGa97},
-\citet{KoVi98} and \citet{KoTr99}, whereas applications involving
-Chebyshev basis functions and a nondiagonal mass matrix include \citet{SePr94},
-\citet{PrCaSe94} and \citet{SePrPr95}.\\
-
-All SPECFEM3D software is written in Fortran2003 with full portability
-in mind, and conforms strictly to the Fortran2003 standard. It uses
-no obsolete or obsolescent features of Fortran. The package uses parallel
-programming based upon the Message Passing Interface (MPI) \citep{GrLuSk94,Pac97}.\\
-
-SPECFEM3D won the Gordon Bell award for best performance at the SuperComputing~2003
-conference in Phoenix, Arizona (USA) (see \citet{KoTsChTr03} and
-\url{www.sc-conference.org/sc2003/nrfinalaward.html}). It was a finalist
-again in 2008 for a run at 0.16 petaflops (sustained) on 149,784 processors
-of the `Jaguar' Cray XT5 system at Oak Ridge National Laboratories
-(USA) \citep{CaKoLaTiMiLeSnTr08}. It also won the BULL Joseph Fourier
-supercomputing award in 2010.\\
-
-It reached the sustained one petaflop performance level for the first time in February 2013
-on the Blue Waters Cray supercomputer at the National Center for Supercomputing Applications (NCSA), located at the University of Illinois at Urbana-Champaign (USA).\\
-
-This new release of the code includes Convolution or
-Auxiliary Differential Equation Perfectly Matched absorbing Layers
-(C-PML or ADE-PML) \citep{MaKoEz08,MaKoGe08,MaKo09,MaKoGeBr10,KoMa07}.
-It also includes support for GPU graphics card acceleration \citep{Kom11,MiKo10,KoMiEr09,KoErGoMi10}.\\
-
-The next release of the code will use the PT-SCOTCH parallel and threaded
-version of SCOTCH for mesh partitioning instead of the serial version.\\
-
-SPECFEM3D Cartesian includes coupled fluid-solid domains and adjoint
-capabilities, which enables one to address seismological inverse problems,
-but for linear rheologies only so far. To accommodate visco-plastic
-or non-linear rheologies, the reader can refer to the \href{http://geoelse.stru.polimi.it/}{GeoELSE}software
-package \citep{CaGa97,StPaIg09}.
-
-\section{Citation}
-
-If you use SPECFEM3D Cartesian for your own research, please cite
-at least one of the following articles:
-\begin{description}
-\item [{Numerical simulations in general}] ~\\
- Forward and adjoint simulations are described in detail in \citet{TrKoLi08,PeKoLuMaLeCaLeMaLiBlNiBaTr11,VaCaSaKoVi99,KoMiEr09,KoErGoMi10,ChKoViCaVaFe07,MaKoDi09,KoViCh10,CaKoLaTiMiLeSnTr08,TrKoHjLiZhPeBoMcFrTrHu10,KoRiTr02,KoTr02a,KoTr02b,KoTr99}
-or \citet{KoVi98}. Additional aspects of adjoint simulations are
-described in \citet{TrTaLi05,LiTr06,TrKoLi08,LiTr08,TrKoHjLiZhPeBoMcFrTrHu10,PeKoLuMaLeCaLeMaLiBlNiBaTr11}.
-Domain decomposition is explained in detail in \citet{MaKoBlLe08},
-and excellent scaling up to 150,000 processor cores is shown for instance
-in \citet{CaKoLaTiMiLeSnTr08,KoLaMi08a,MaKoBlLe08,KoErGoMi10,Kom11},
-\item [{GPU computing}] ~\\
- Computing on GPU graphics cards for acoustic or seismic wave propagation
-applications is described in detail in \citet{Kom11,MiKo10,KoMiEr09,KoErGoMi10}.
-\end{description}
-\noindent If you use this new version, which has non blocking
-MPI for much better performance for medium or large runs, please cite
-at least one of these six articles, in which results of non blocking
-MPI runs are presented: \citet{PeKoLuMaLeCaLeMaLiBlNiBaTr11,KoErGoMi10,KoViCh10,Kom11,CaKoLaTiMiLeSnTr08,MaKoBlLe08}.\\
-
-\noindent If you work on geophysical applications, you may be interested
-in citing some of these application articles as well, among others:
-\begin{description}
-\item [{Southern California simulations}] ~\\
- \citet{KoLiTrSuStSh04,KrJiKoTr06a,KrJiKoTr06b}.
-
-If you use the 3D southern California model, please cite \citet{SuSh03}
-(Los Angeles model), \citet{lovelyetal06} (Salton Trough), and \citet{hauksson2000}
-(southern California). The Moho map was determined by \citet{zhu&kanamori2000}.
-The 1D SoCal model was developed by \citet{DrHe90}.
-
-\item [{Anisotropy}] ~\\
- \citet{ChTr07,JiTsKoTr05,ChFaKo04,FaChKo04,RiRiKoTrHe02,TrKo00}.
-\item [{Attenuation}] ~\\
- \citet{SaKoTr10,KoTr02a,KoTr99}.
-\item [{Topography}] ~\\
- \citet{LeKoHuTr09,LeChKoHuTr09,LeChLiKoHuTr08,GoAmTaCaSmSaMaKo09,WiKoScTr04}.
-\end{description}
-The corresponding Bib\TeX{} entries may be found in file \texttt{doc/USER\_MANUAL/bibliography.bib}.
-
-\section{Support}
-
-This material is based upon work supported by the USA National Science
-Foundation under Grants No. EAR-0406751 and EAR-0711177, by the French
-CNRS, French INRIA Sud-Ouest MAGIQUE-3D, French ANR NUMASIS under
-Grant No. ANR-05-CIGC-002, and European FP6 Marie Curie International
-Reintegration Grant No. MIRG-CT-2005-017461. Any opinions, findings,
-and conclusions or recommendations expressed in this material are
-those of the authors and do not necessarily reflect the views of the
-USA National Science Foundation, CNRS, INRIA, ANR or the European
-Marie Curie program.
-
-\chapter{\label{cha:Getting-Started}Getting Started}
+\chapter{Getting Started}\label{cha:Getting-Started}
 
@@
 \section{Using the ADIOS library for I/O}
+
 Regular POSIX I/O can be problematic when dealing with large simulations one large
@@
 to access the development branch of the source code with read/write access (in a safe way, no need to worry too much about breaking the package, there is a robot called BuildBot that is in charge of checking and validating all new contributions and changes), please visit this Web page: \url{https://github.com/geodynamics/specfem3d/wiki/Using-Hub}.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\chapter{\label{cha:Mesh-Generation}Mesh Generation}
-
-The first step in running a spectral-element simulation consists of
-constructing a high-quality mesh for the region under consideration.
-We provide two possibilities to do so: (1) relying on the external,
-hexahedral mesher CUBIT, or (2) using the provided, internal mesher
-\texttt{xmeshfem3D}. In the following, we explain these two approaches.
-
-\section{\label{cha:Running-the-Mesher-CUBIT}Meshing with \texttt{CUBIT}}
-
-CUBIT is a meshing tool suite for the creation of finite-element meshes
-for arbitrarily shaped models. It has been developed and maintained
-at Sandia National Laboratories and can be purchased for a small academic
-institutional fee at \url{http://cubit.sandia.gov}. Our experience
-showed that using CUBIT greatly facilitates and speeds up the generation
-and preparation of hexahedral, conforming meshes for a variety of
-geophysical models with increasing complexity.
-
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=0.6\textwidth]{figures/mount-cubit}
-\par\end{centering}
-
-\caption{Example of the graphical user interface of CUBIT. The hexahedral mesh
-shown in the main display consists of a hexahedral discretization
-of a single volume with topography.}
-
-\label{fig:mount.cubit}
-\end{figure}
-
-\noindent
-The basic steps in creating a load-balanced, partitioned mesh with
-CUBIT are:
-\begin{description}
-\item [{1.}] setting up a hexahedral mesh with CUBIT,
-\item [{2.}] exporting the CUBIT mesh into a SPECFEM3D Cartesian file format
-and
-\item [{3.}] partitioning the SPECFEM3D Cartesian mesh files for a chosen
-number of cores.
-\end{description}
-Examples are provided in the SPECFEM3D Cartesian package in the subdirectory
-\texttt{EXAMPLES/}. We strongly encourage you to contribute your own
-example to this package by contacting the CIG Computational Seismology
-Mailing List \urlwithparentheses{cig-seismo@geodynamics.org}.
-
-\subsection{Creating the Mesh with CUBIT}
-
-For the installation and handling of the CUBIT meshing tool suite,
-please refer to the CUBIT user manual and documentation. In order
-to give you a basic understanding of how to use CUBIT for our purposes,
-examples are provided in the SPECFEM3D Cartesian package in the subdirectory
-\texttt{EXAMPLES/}:
-\begin{description}
-\item [{\texttt{homogeneous\_halfspace}}] Creates a single block model
-and assigns elastic material parameters.
-\item [{\texttt{layered\_halfspace}}] Combines two different, elastic material
-volumes and creates a refinement layer between the two. This example
-can be compared for validation against the solutions provided in subdirectory
-~\\
- \texttt{VALIDATION\_3D\_SEM\_SIMPLER\_LAYER\_SOURCE\_DEPTH/}.
-\item [{\texttt{waterlayered\_halfspace}}] Combines an acoustic and elastic
-material volume as in a schematic marine survey example.
-\item [{\texttt{tomographic\_model}}] Creates a single block model whose
-material properties will have to be read in from a tomographic model
-file during the databases creation by \texttt{xgenerate\_databases}.
-\end{description}
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=0.45\textwidth]{figures/example-homogeneous}
-\includegraphics[width=0.45\textwidth]{figures/example-2layers} \\
- \includegraphics[width=0.45\textwidth]{figures/example-water} \includegraphics[width=0.45\textwidth]{figures/example-tomo}
-\par\end{centering}
-
-\caption{Screenshots of the CUBIT examples provided in subdirectory \texttt{EXAMPLES/}:
-homogeneous halfspace (top-left), layered halfspace (top-right), water
-layered halfspace (bottom-left) and tomographic model (bottom-right).}
-
-\label{fig:examples.cubit}
-\end{figure}
-
-In each example subdirectory you will find a \texttt{README} file,
-which explains in a step-by-step tutorial the workflow for the example.
-Please feel free to contribute your own example to this package by
-contacting the CIG Computational Seismology Mailing List \urlwithparentheses{cig-seismo@geodynamics.org}.
-
-In some cases, to re-create the meshes for the examples given, just type "claro ./create\_mesh.py" or similar
-from the command line ("claro" is the command to run CUBIT from the command line).
-
-IMPORTANT note: in order to correctly set up GEOCUBIT and run the examples, please read the file called "EXAMPLES/README";
-in particular, please make sure you correctly set up the Python paths as indicated in that file.
-
-\subsection{Exporting the Mesh with \texttt{run\_cubit2specfem3d.py} }
-
-\label{subsec:Exporting-the-Mesh}
-
-Once the geometric model volumes in CUBIT are meshed, you prepare
-the model for exportation with the definition of material blocks and
-boundary surfaces. Thus, prior to exporting the mesh, you need to
-define blocks specifying the materials and absorbing boundaries in
-CUBIT. This process could be done automatically using the script \texttt{run\_cubit2specfem3d.py}
-if the mesh meets some conditions or manually, following the block
-convention:
-\begin{description}
-\item [{material\_name}] Each material should have a specific block defined
-by a unique name. The name convention of the material is to start
-with either 'elastic' or 'acoustic'. It must be then followed by a
-unique identifier, e.g. 'elastic 1', 'elastic 2', etc. The additional
-attributes to the block define the material description.
-
-For an elastic material:
-\begin{description}
-\item [{material\_id}] An integer value which is unique for this material.
-\item [{Vp}] P-wave speed of the material (given in m/s).
-\item [{Vs}] S-wave speed of the material (given in m/s).
-\item [{rho}] density of the material (given in kg/m$^{3}$).
-\item [{Q}] quality factor to use in case of a simulation with attenuation
-turned on. It should be between 1 and 9000. In case no attenuation
-information is available, it can be set to zero. Please note that
-your Vp- and Vs-speeds are given for a reference frequency. To change
-this reference frequency, you change the value of \texttt{ATTENUATION\_f0\_REFERENCE}
-in the main constants file \texttt{constants.h} found in subdirectory
-\texttt{src/shared/}.
-\item [{anisotropic\_flag}] Flag describing the anisotropic model to use
-in case an anisotropic simulation should be conducted. See the file
-\texttt{model\_aniso.f90} in subdirectory \texttt{src/generate\_databases/}
-for an implementation of the anisotropic models. In case no anisotropy
-is available, it can be set to zero.
-\end{description}
-
-Note that this material block has to be defined using all the volumes
-which belong to this elastic material. For volumes belonging to another,
-different material, you will need to define a new material block.
-
-For an acoustic material:
-\begin{description}
-\item [{material\_id}] An integer value which is unique for this material.
-\item [{Vp}] P-wave speed of the material (given in m/s).
-\item [{0}] S-wave speed of the material is ignored.
-\item [{rho}] density of the material (given in kg/m$^{3}$).
-\end{description}
-\item [{face\_topo}] Block definition for the surface which defines the
-free surface (which can have topography). The name of this block must
-be 'face\_topo', the block has to be defined using all the surfaces
-which constitute the complete free surface of the model.
-\item [{face\_abs\_xmin}] Block definition for the faces on the absorbing
-boundaries, one block for each surface with x=Xmin.
-\item [{face\_abs\_xmax}] Block definition for the faces on the absorbing
-boundaries, one block for each surface with x=Xmax.
-\item [{face\_abs\_ymin}] Block definition for the faces on the absorbing
-boundaries, one block for each surface with y=Ymin.
-\item [{face\_abs\_ymax}] Block definition for the faces on the absorbing
-boundaries, one block for each surface with y=Ymax.
-\item [{face\_abs\_bottom}] Block definition for the faces on the absorbing
-boundaries, one block for each surface with z=bottom.
-\end{description}
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=0.45\textwidth]{figures/mount-surface} \includegraphics[width=0.45\textwidth]{figures/mount-abs}
-\par\end{centering}
-
-\caption{Example of the block definitions for the free surface 'face\_topo'
-(left) and the absorbing boundaries, defined in a single block 'face\_abs\_xmin'
-(right) in CUBIT.}
-
-\label{fig:mount.abs}
-\end{figure}
-
-Optionally, instead of specifying for each surface at the boundaries
-a single block like mentioned above, you can also specify a single
-block for all boundary surfaces and name it as one of the absorbing
-blocks above, e.g. 'face\_abs\_xmin'.
-
-After the block definitions are done, you export the mesh using the
-script \texttt{cubit2specfem3d.py} provided in each of the example
-directories (linked to the common script \texttt{CUBIT\_GEOCUBIT/cubit2specfem3d.py}).
-If the export was successful, you should find the following
-files in a subdirectory \texttt{MESH/}:
-\begin{description}
-\item [{absorbing\_cpml\_file}] \textbf{(only needed in case of C-PML absorbing
-conditions)} Contains on the first line the total number of C-PML
-spectral elements in the mesh, and then on the following line the
-list of all these C-PML elements with two numbers per line: first
-the spectral element number, and then a C-PML flag indicating to which
-C-PML layer(s) that element belongs, according to the following convention:
-
-\begin{itemize}
-\item Flag = 1 : element belongs to a X CPML layer only (either in Xmin
-or in Xmax),
-\item Flag = 2 : element belongs to a Y CPML layer only (either in Ymin
-or in Ymax),
-\item Flag = 3 : element belongs to a Z CPML layer only (either in Zmin
-or in Zmax),
-\item Flag = 4 : element belongs to a X CPML layer and also to a Y CPML
-layer,
-\item Flag = 5 : element belongs to a X CPML layer and also to a Z CPML
-layer,
-\item Flag = 6 : element belongs to a Y CPML layer and also to a Z CPML
-layer,
-\item Flag = 7 : element belongs to a X, to a Y and to a Z CPML layer, i.e.,
-it belongs to a CPML corner.
-\end{itemize}
-
-Note that it does not matter whether an element belongs to a Xmin
-or to a Xmax CPML, the flag is the same in both cases; the same is
-true for Ymin and Ymax, and also Zmin and Zmax.
-
-When you have an existing CUBIT (or similar) mesh stored in SPECFEM3D
-format, i.e., if you have existing \textquotedbl{}nodes\_coords\_file\textquotedbl{}
-and \textquotedbl{}mesh\_file\textquotedbl{} files but do not know
-how to assign CPML flags to them, we have created a small serial Fortran
-program that will do that automatically for you, i.e., which will
-create the \textquotedbl{}absorbing\_cpml\_file\textquotedbl{} for
-you. That program is utils/CPML/convert\_external\_layers\_of\_a\_given\_mesh\_to\_CPML\_layers.f90,
-and a small Makefile is provided in that directory (utils/CPML).
-
-IMPORTANT: it is your responsibility to make sure that in the input
-CUBIT (or similar) mesh that this code will read in SPECFEM3D format
-from files \textquotedbl{}nodes\_coords\_file\textquotedbl{} and \textquotedbl{}mesh\_file\textquotedbl{}
-you have created layers of elements that constitute a layer of constant
-thickness aligned with the coordinate grid axes (X, Y and/or Z), so
-that this code can assign CPML flags to them. This code does NOT check
-that (because it cannot, in any easy way). The mesh inside these CPML
-layers does not need to be structured nor regular, any non-structured
-mesh is fine as long as it has flat PML inner and outer faces, parallel
-to the axes, and thus of a constant thickness. The thickness can be
-different for the X, Y and Z sides. But for X it must not vary, for
-Y it must not vary, and for Z it must not vary. If you do not know
-the exact thickness, you can use a slightly LARGER value in this code
-(say 2\% to 5\% more) and this code will fix that and will adjust
-it; never use a SMALLER value otherwise this code will miss some CPML
-elements.
-
-Note: in a future release we will remove the constraint of having
-CPML layers aligned with the coordinate axes; we will allow for meshes
-that are titled by any constant angle in the horizontal plane. However
-this is not implemented yet.
-
-Note: in the case of fluid-solid layers in contact, the fluid-solid
-interface currently needs to be flat and horizontal inside the CPML
-layer (i.e., bathymetry should become flat and horizontal when entering
-the CPML); this (small) constraint will probably remain in the code
-for a while because it makes fluid-solid matching inside the CPML
-much easier.
-
-\item [{materials\_file}] Contains the material associations for each element.
-The format is:
-
-\begin{lyxcode}
-element\_ID~material\_ID
-\end{lyxcode}
-
-where \texttt{element\_ID} is the element identifier and \texttt{material\_ID}
-is a unique identifier, positive (for materials taken from this list
-of materials, i.e. for which each spectral element has constant material
-properties taken from this list) or negative (for tomographic models,
-i.e. for spectral element whose real velocities and density will be
-assigned later by calling an external function to define model variations,
-for instance in the case of tomographic models; in such a case, material
-properties can vary inside each spectral element, i.e. be different
-at each of its Gauss-Lobatto-Legendre grid points).
-
-\item [{nummaterial\_velocity\_file}] Defines the material properties.
-
-\begin{itemize}
-\item For classical materials (i.e., spectral elements for which the velocity
-and density model will not be assigned by calling an external function
-to define for instance a tomographic model), the format is:
-
-\begin{lyxcode}
-domain\_ID~material\_ID~rho~vp~vs~Qkappa~Qmu~anisotropy\_flag
-\end{lyxcode}
-
-where \texttt{\textbf{domain\_ID}}\textbf{ is 1 for acoustic and 2
-for elastic or viscoelastic materials,} \texttt{material\_ID} a unique
-identifier, \texttt{rho} the density in $kg\, m^{-3}$, \texttt{vp}
-the P-wave speed in $m\, s^{-1}$, \texttt{vs} the S-wave speed in
-$m\, s^{-1}$, \texttt{Q} the quality factor and \texttt{anisotropy\_flag}
-an identifier for anisotropic models. Note that the \texttt{Qkappa}
-value is ignored by the code unless \texttt{FULL\_ATTENUATION\_SOLID}
-is set. Note also that both \texttt{Qkappa} and \texttt{Qmu} are ignored
-by the code unless \texttt{ATTENUATION} is set. If you want a model
-with no \texttt{Qmu} attenuation, both set \texttt{ATTENUATION} to
-\texttt{.false.} in the \texttt{Par\_file} and set \texttt{Qmu} to
-9999 here. If you want a model with no \texttt{Qkappa} attenuation,
-both set \texttt{FULL\_ATTENUATION\_SOLID} to \texttt{.false.} in
-the \texttt{Par\_file} and set \texttt{Qkappa} to 9999 here. %%magnoni
-
-\item For tomographic velocity models, please read Chapter \ref{cha:-Changing-the}
-and Section \ref{sec:Using-tomographic} `Using external tomographic
-Earth models' for further details.
-\end{itemize}
-\item [{nodes\_coords\_file}] Contains the point locations in Cartesian
-coordinates of the mesh element corners.
-\item [{mesh\_file}] Contains the mesh element connectivity. The hexahedral
-elements can have 8 or 27 nodes.\\
- See picture doc/mesh\_numbering\_convention/numbering\_convention\_27\_nodes.jpg
-to see\\
- in which (standard) order the points must be cited. In the case of
-8 nodes, just include the first 8 points.
-\item [{free\_or\_absorbing\_surface\_file\_zmax}] Contains the free surface
-connectivity or \\
- the surface connectivity of the absorbing boundary surface at the
-top (Zmax), \\
- depending on whether the top surface is defined as free or absorbing
-(\texttt{STACEY\_INSTEAD\_OF\_FREE\_SURFACE} in \texttt{DATA/Par\_file}). \\
- You should put both the surface of acoustic regions and of elastic
-regions in that file; that is, list all the element faces that constitute
-the surface of the model in that file.
-\item [{absorbing\_surface\_file\_xmax}] Contains the surface connectivity
-of the absorbing boundary surface at Xmax \\
-(also needed in the case of C-PML absorbing conditions, in order for
-the code to be able to impose Dirichlet conditions on their outer
-edge).
-\item [{absorbing\_surface\_file\_xmin}] Contains the surface connectivity
-of the absorbing boundary surface at Xmin \\
-(also needed in the case of C-PML absorbing conditions, in order for
-the code to be able to impose Dirichlet conditions on their outer
-edge).
-\item [{absorbing\_surface\_file\_ymax}] Contains the surface connectivity
-of the absorbing boundary surface at Ymax \\
-(also needed in the case of C-PML absorbing conditions, in order for
-the code to be able to impose Dirichlet conditions on their outer
-edge).
-\item [{absorbing\_surface\_file\_ymin}] Contains the surface connectivity
-of the absorbing boundary surface at Ymin \\
-(also needed in the case of C-PML absorbing conditions, in order for
-the code to be able to impose Dirichlet conditions on their outer
-edge).
-\item [{absorbing\_surface\_file\_bottom}] Contains the surface connectivity
-of the absorbing boundary surface at the bottom (Zmin) \\
-(also needed in the case of C-PML absorbing conditions, in order for
-the code to be able to impose Dirichlet conditions on their outer
-edge).
-\end{description}
-These mesh files are needed as input files for the partitioner \texttt{xdecompose\_mesh}
-to load-balance the mesh. Please see the next section for further
-details.
-
-In directory "CUBIT\_GEOCUBIT" we provide a script that can help doing the above tasks of exporting a CUBIT mesh
-to SPECFEM3D format automatically for you: "run\_cubit2specfem3d.py".
-Just edit them to indicate the path to your local installation of CUBIT and also the name of the *.cub existing CUBIT mesh file
-that you want to export to SPECFEM3D format. These scripts will do the conversion for you automatically except assigning material properties
-to the different mesh layers. To do so, you will then need to edit the file
-called "nummaterial\_velocity\_file" that will have just been created and change it from the prototype created:
-
-    0 1 vol1 --> syntax: \#material\_domain\_id \#material\_id \#rho \#vp \#vs \#Q\_kappa \#Q\_mu \#anisotropy
-    0 2 vol2 --> syntax: \#material\_domain\_id \#material\_id \#rho \#vp \#vs \#Q\_kappa \#Q\_mu \#anisotropy
-
-(where "vol1" and "vol2" here represent the volume labels that you have set while creating the mesh in CUBIT)
-to for instance
-
-    2 1 1500 2300 1800 9999.0 9999.0 0
-    2 2 1600 2500 20000 9999.0 9999.0 0
-
-\subsubsection*{Checking the mesh quality}
-
-The quality of the mesh may be inspected more precisely based upon
-the serial code in the file \texttt{check\_mesh\_quality\_}~\\
- \texttt{CUBIT\_Abaqus.f90} located in the directory \texttt{src/check\_mesh\_quality\_CUBIT\_Abaqus/}.
-Running this code is optional because no information needed by the
-solver is generated.
-
-Prior to running and compiling this code, you have to export your
-mesh in CUBIT to an ABAQUS (.inp) format. For example, export mesh
-block IDs belonging to volumes in order to check the quality of the
-hexahedral elements. You also have to determine a number of parameters
-of your mesh, such as the number of nodes and number of elements and
-modify the header of the \texttt{check\_mesh\_quality\_CUBIT\_Abaqus.f90}
-source file in directory \texttt{src/check\_mesh\_quality\_CUBIT\_Abaqus/}.
-
-\noindent
-Then, in the main directory, type
-\begin{lyxcode}
-{\small make~xcheck\_mesh\_quality\_CUBIT\_Abaqus}{\small \par}
-\end{lyxcode}
-and use
-\begin{lyxcode}
-{\small ./bin/xcheck\_mesh\_quality\_CUBIT\_Abaqus}{\small \par}
-\end{lyxcode}
-to generate an %not supported: AVS output file (\texttt{\small AVS\_meshquality.inp} in AVS UCD format) or
-OpenDX output file (\texttt{\small DX\_mesh\_quality.dx}{\small )
-that can be used to investigate mesh quality, e.g. skewness of elements,
-and a Gnuplot histogram (}\texttt{\small mesh\_quality\_histogram.txt}{\small )
-that can be plotted with gnuplot (type `}\texttt{\small gnuplot plot\_mesh\_quality\_histogram.gnu}{\small ').
-The histogram is also printed to the screen. Analyze that skewness
-histogram of mesh elements to make sure no element has a skewness
-above approximately 0.75, otherwise the mesh is of poor quality (and
-if even a single element has a skewness value above 0.80, then you
-must definitely improve the mesh). If you want to start designing
-your own meshes, this tool is useful for viewing your creations. You
-are striving for meshes with elements with `cube-like' dimensions,
-e.g., the mesh should contain no very elongated or skewed elements.}{\small \par}
-
-\subsection{Partitioning the Mesh with \texttt{xdecompose\_mesh}}
-
-The SPECFEM3D Cartesian software package performs large scale simulations
-in a parallel 'Single Process Multiple Data' way. The spectral-element
-mesh created with CUBIT needs to be distributed on the processors.
-This partitioning is executed once and for all prior to the execution
-of the solver so it is referred to as a static mapping.
-
-An efficient partitioning is important because it leverages the overall
-running time of the application. It amounts to balance the number
-of elements in each slice while minimizing the communication costs
-resulting from the placement of adjacent elements on different processors.
-\texttt{decompose\_mesh} depends on the SCOTCH library \citep{PeRo96},
-which provides efficent static mapping, graph and mesh partitioning
-routines. SCOTCH is a free software package developed by Fran\c{c}ois
-Pellegrini et al. from LaBRI and INRIA in Bordeaux, France, downloadable
-from the web page \url{https://gforge.inria.fr/projects/scotch/}.
-\\
-
-In most cases, the configuration with \texttt{./configure~FC=ifort}
-should be sufficient. During the configuration process, the script
-tries to find existing SCOTCH installations. In case your system has
-no pre-existing SCOTCH installation, we provide the source code of
-SCOTCH, which is released open source under the French CeCILL-C version
-1 license, in directory \texttt{src/decompose\_mesh/scotch\_5.1.12b}.
-This version gets bundled with the compilation of the SPECFEM3D Cartesian
-package if no libraries could have been found. If this automatic compilation
-of the SCOTCH libraries fails, please refer to file INSTALL.txt in
-that directory to see further details how to compile it on your system.
-In case you want to use a pre-existing installation, make sure you
-have correctly specified the path of the SCOTCH library when using
-the option \texttt{-{}-with-scotch-dir} with the \texttt{./configure}
-script. In the future you should be able to find more recent versions
-at \url{http://www.labri.fr/perso/pelegrin/scotch/scotch_en.html}
-. \\
-
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=0.45\textwidth]{figures/mount-partitions}
-\includegraphics[width=0.45\textwidth]{figures/mount-partitions2}
-\par\end{centering}
-
-\caption{Example of a mesh partitioning onto four cores. Each single core partition
-is colored differently. The executable \texttt{xdecompose\_mesh} can
-equally distribute the mesh on any arbitrary number of cores. Domain
-decomposition is explained in detail in \citet{MaKoBlLe08}, and excellent
-scaling up to 150,000 processor cores in shown for instance in \citet{CaKoLaTiMiLeSnTr08,KoLaMi08a,MaKoBlLe08,KoErGoMi10,Kom11}.}
-
-\label{fig:mount.partitions}
-\end{figure}
-
-When you are ready to compile, in the main directory type
-\begin{lyxcode}
-{\small make~xdecompose\_mesh}{\small \par}
-\end{lyxcode}
-If all paths and flags have been set correctly, the executable \texttt{bin/xdecompose\_mesh}
-should be produced.
-
-The partitioning is done in serial for now (in the next release we
-will provide a parallel version of that code). It needs to be run in the main directory because it expects the \texttt{./DATA/Par\_file}. The synopsis is:
-\begin{lyxcode}
-./bin/xdecompose\_mesh~nparts~input\_directory~output\_directory
-\end{lyxcode}
-where
-\begin{itemize}
-\item \texttt{nparts} is the number of partitions, i.e., the number of cores
-for the parallel simulations,
-\item \texttt{input\_directory} is the directory which holds all the files
-generated by the Python script \texttt{cubit2specfem3d.py} explained
-in the previous Section~\ref{subsec:Exporting-the-Mesh}, e.g. \texttt{./MESH/},
-and
-\item \texttt{output\_directory} is the directory for the output of this
-partitioner which stores ACII-format files named like \texttt{proc{*}{*}{*}{*}{*}{*}\_Database}
-for each partition. These files will be needed for creating the distributed
-databases, and have to reside in the directory \texttt{LOCAL\_PATH}
-specified in the main \texttt{Par\_file}, e.g. in directory \texttt{./OUTPUT\_FILES/DATABASES\_MPI}.
-Please see Chapter~\ref{cha:Creating-Distributed-Databases} for
-further details.
-\end{itemize}
-Note that all the files generated by the Python script \texttt{cubit2specfem3d.py}
-must be placed in the \texttt{input\_directory} folder before running
-the program.
-
-\section{\label{cha:Running-the-Mesher-Meshfem3D}Meshing with \texttt{xmeshfem3D}}
-
-In case you successfully ran the configuration script, you are also
-ready to compile the internal mesher. This is an alternative to CUBIT
-for the mesh generation of relatively simple models.
-
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[scale=0.5]{figures/socal_map_mpi}
-\par\end{centering}
-
-\caption{\label{fig:For-parallel-computing}For parallel computing purposes,
-the model block is subdivided in $\nprocxi\times\nproceta$ slices
-of elements. In this example we use $5^{2}=25$ processors. }
-\end{figure}
-
-\noindent
-In the main directory, type
-\begin{lyxcode}
-{\small make~xmeshfem3D}{\small \par}
-\end{lyxcode}
-If all paths and flags have been set correctly, the mesher should
-now compile and produce the executable \texttt{bin/xmeshfem3D}. Please
-note that \texttt{xmeshfem3D} must be called directly from the main
-directory, as most of the binaries of the package.
-
-Input for the mesh generation program is provided through the parameter
-file \texttt{Mesh\_Par\_file}, which resides in the subdirectory \texttt{DATA/meshfem3D\_files/}.
-(to see how to use it, see the EXAMPLES specific to the internal mesher in directory \texttt{EXAMPLES/meshfem3D\_examples/}.
-Before running the mesher, a number of parameters need to be set in
-the \texttt{Mesh\_Par\_file}. This requires a basic understanding
-of how the SEM is implemented, and we encourage you to read \citet{KoVi98,KoTr99}
-and \citet{KoLiTrSuStSh04}.
-
-The mesher and the solver use UTM coordinates internally, therefore
-you need to define the zone number for the UTM projection (e.g., zone
-11 for Los Angeles). Use decimal values for latitude and longitude
-(no minutes/seconds). These values are approximate; the mesher will
-round them off to define a square mesh in UTM coordinates. When running
-benchmarks on rectangular models, turn the UTM projection off by using
-the flag \texttt{\small SUPPRESS\_UTM\_PROJECTION}{\small , in which
-case all `longitude' parameters simply refer to the $x$~axis, and
-all `latitude' parameters simply refer to the $y$~axis. To run the
-mesher for a global simulation, the following parameters need to be
-set in the }\texttt{\small Mesh\_Par\_file}{\small :}{\small \par}
-\begin{description}
-\item [{\texttt{LATITUDE\_MIN}}] Minimum latitude in the block (negative
-for South).
-\item [{\texttt{LATITUDE\_MAX}}] Maximum latitude in the block.
-\item [{\texttt{LONGITUDE\_MIN}}] Minimum longitude in the block (negative
-for West).
-\item [{\texttt{LONGITUDE\_MAX}}] Maximum longitude in the block.
-\item [{\texttt{DEPTH\_BLOCK\_KM}}] Depth of bottom of mesh in kilometers.
-\item [{\texttt{\noun{UTM\_PROJECTION\_ZONE}}}] UTM projection zone in
-which your model resides, only valid when \texttt{SUPPRESS\_UTM\_}~\\
- \texttt{PROJECTION} is \texttt{.false.}.
-Use a negative zone number for the Southern hemisphere:
-the Northern hemisphere corresponds to zones +1 to +60,
-the Southern hemisphere to zones -1 to -60.
-We use the WGS84 (World Geodetic System 1984) reference ellipsoid for the UTM projection. If you prefer to use the Clarke 1866 ellipsoid,
-edit file \texttt{src/shared/utm\_geo.f90}, uncomment that ellipsoid and recompile the code.
-From http://en.wikipedia.org/wiki/Universal\_Transverse\_Mercator\_coordinate\_system :
-The Universal Transverse Mercator coordinate system was developed by the United States Army Corps of Engineers in the 1940s.
-The system was based on an ellipsoidal model of Earth. For areas within the contiguous United States
-the Clarke Ellipsoid of 1866 was used. For the remaining areas of Earth, including Hawaii, the International Ellipsoid was used.
-The WGS84 ellipsoid is now generally used to model the Earth in the UTM coordinate system,
-which means that current UTM northing at a given point can be 200+ meters different from the old one.
-For different geographic regions, other datum systems (e.g.: ED50, NAD83) can be used.
-\item [{\texttt{SUPPRESS\_UTM\_PROJECTION}}] set to be \texttt{.false.}
-when your model range is specified in geographical coordinates, and
-needs to be \texttt{.true.} when your model is specified in Cartesian
-coordinates. \noun{UTM projection zone in which your simulation region
-resides.}
-\item [{\texttt{INTERFACES\_FILE }}] File which contains the description
-of the topography and of the interfaces between the different layers
-of the model, if any. The number of spectral elements in the vertical
-direction within each layer is also defined in this file.
-\item [{$\nexxi$}] The number of spectral elements along one side of the
-block. This number \textit{must} be 8~$\times$~a multiple of $\nprocxi$
-defined below. Based upon benchmarks against semi-analytical discrete
-wavenumber synthetic seismograms \citep{KoLiTrSuStSh04}, determined
-that a $\nexxi=288$ run is accurate to a shortest period of roughly
-2~s. Therefore, since accuracy is determined by the number of grid
-points per shortest wavelength, for any particular value of $\nexxi$
-the simulation will be accurate to a shortest period determined by
-\begin{equation}
-\mbox{shortest period (s)}=(288/\nexxi)\times2.\label{eq:shortest_period}
-\end{equation}
- The number of grid points in each orthogonal direction of the reference
-element, i.e., the number of Gauss-Lobatto-Legendre points, is determined
-by \texttt{NGLLX} in the \texttt{constants.h} file. We generally use
-$\mbox{\texttt{NGLLX\/}}=5$, for a total of $5^{3}=125$ points per
-elements. We suggest not to change this value.
-\item [{$\nexeta$}] The number of spectral elements along the other side
-of the block. This number \textit{must} be 8~$\times$~a multiple
-of $\nproceta$ defined below.
-\item [{$\nprocxi$}] The number of processors or slices along one side
-of the block (see Figure~\ref{fig:For-parallel-computing}); we must
-have $\nexxi=8\times c\times\nprocxi$, where $c\ge1$ is a positive
-integer.
-\item [{$\nproceta$}] The number of processors or slices along the other
-side of the block; we must have $\nexeta=8\times c\times\nproceta$,
-where $c\ge1$ is a positive integer.
-\item [{\texttt{USE\_REGULAR\_MESH}}] set to be \texttt{.true.} if you
-want a perfectly regular mesh or \texttt{.false.} if you want to add
-doubling horizontal layers to coarsen the mesh. In this case, you
-also need to provide additional information by setting up the next
-three parameters.
-\item [{\texttt{NDOUBLINGS}}] The number of horizontal doubling layers.
-Must be set to \texttt{1} or \texttt{2} if \texttt{USE\_REGULAR\_MESH}
-is set to \texttt{.true.}.
-\item [{\texttt{NZ\_DOUBLING\_1}}] The position of the first doubling layer
-(only interpreted if \texttt{USE\_REGULAR\_MESH} is set to \texttt{.true.}).
-\item [{\texttt{NZ\_DOUBLING\_2}}] The position of the second doubling
-layer (only interpreted if \texttt{USE\_REGULAR\_MESH} is set to \texttt{.true.}
-and if \texttt{NDOUBLINGS} is set to \texttt{2}).
-\item [{\texttt{CREATE\_ABAQUS\_FILES}}] Set this flag to \texttt{.true.}
-to save Abaqus FEA \urlwithparentheses{www.simulia.com} mesh files
-for subsequent viewing. Turning the flag on generates files in the
-\texttt{LOCAL\_PATH} directory. See Section~\ref{sec:Mesh-graphics}
-for a discussion of mesh viewing features.
-\item [{\texttt{CREATE\_DX\_FILES}}] Set this flag to \texttt{.true.} to
-save OpenDX \urlwithparentheses{www.opendx.org} mesh files for subsequent
-viewing.
-\item [{\texttt{LOCAL\_PATH}}] Directory in which the partitions generated
-by the mesher will be written. Generally one uses a directory on the
-local disk of the compute nodes, although on some machines these partitions
-are written on a parallel (global) file system (see also the earlier
-discussion of the \texttt{LOCAL\_PATH\_IS\_ALSO\_GLOBAL} flag in Chapter~\ref{cha:Getting-Started}).
-The mesher generates the necessary partitions in parallel, one set
-for each of the $\nprocxi\times\nproceta$ slices that constitutes
-the mesh (see Figure~\ref{fig:For-parallel-computing}). After the
-mesher finishes, you can log in to one of the compute nodes and view
-the contents of the \texttt{LOCAL\_PATH} directory to see the files
-generated by the mesher. These files will be needed for creating the
-distributed databases, and have to reside in the directory \texttt{LOCAL\_PATH}
-specified in the main \texttt{Par\_file}, e.g. in directory \texttt{OUTPUT\_FILES/DATABASES\_MPI}.
-Please see Chapter~\ref{cha:Creating-Distributed-Databases} for
-further details.
-\item [{\texttt{NMATERIALS}}] The number of different materials in your
-model. In the following lines, each material needs to be defined as
-:
-
-\begin{lyxcode}
-material\_ID~rho~vp~vs~Q~anisotropy\_flag~domain\_ID
-\end{lyxcode}
-
-where
-\begin{itemize}
-\item \texttt{Q} : quality factor (0=no attenuation)
-\item \texttt{anisotropy\_flag} : 0=no anisotropy / 1,2,.. check with implementation
-in \texttt{aniso\_model.f90}
-\item \texttt{domain\_id} : 1=acoustic / 2=elastic
-\end{itemize}
-\item [{\texttt{NREGIONS}}] The number of regions in the mesh. In the following
-lines, because the mesh is regular or 'almost regular', each region
-is defined as :
-
-\begin{lyxcode}
-NEX\_XI\_BEGIN~NEX\_XI\_END~NEX\_ETA\_BEGIN~NEX\_ETA\_END~NZ\_BEGIN~NZ\_END~material\_ID
-\end{lyxcode}
-\end{description}
-The \texttt{INTERFACES\_FILE} parameter of \texttt{Mesh\_Par\_File}
-defines the file which contains the settings of the topography grid
-and of the interfaces grids. Topography is defined as a set of elevation
-values on a regular 2D grid. It is also possible to define interfaces
-between the layers of the model in the same way. The file needs to
-define several parameters:
-\begin{itemize}
-\item The number of interfaces, including the topography. This needs to
-be set at the first line. Then, from the bottom to the top of the
-model, you need to define the grids with:
-\item \texttt{SUPPRESS\_UTM\_PROJECTION} flag as described previously,
-\item number of points along $x$ and $y$ direction (NXI and NETA),
-\item minimal $x$ and $y$ coordinates (LONG\_MIN and LAT\_MIN),
-\item spacing between points along $x$ and $y$ (SPACING\_XI and SPACING\_ETA)
-and
-\item the name of the file which contains the elevation values (in $y$.$x$
-increasing order).
-\end{itemize}
-At the end of this file, you simply need to set the number of spectral
-elements in the vertical direction for each layer. We provide a few
-models in the {\texttt{EXAMPLES/}} directory. \\
-
-Finally, depending on your system, you might need to provide a file
-that tells MPI what compute nodes to use for the simulations. The
-file must have a number of entries (one entry per line) at least equal
-to the number of processors needed for the run. A sample file is provided
-in the file \texttt{mymachines}. This file is not used by the mesher
-or solver, but is required by the \texttt{go\_mesher} and \texttt{go\_solver}
-default job submission scripts. See Chapter~\ref{cha:Scheduler}
-for information about running the code on a system with a scheduler,
-e.g., LSF.
-
-Now that you have set the appropriate parameters in the \texttt{Mesh\_Par\_file}
-and have compiled the mesher, you are ready to launch it! This is
-most easily accomplished based upon the \texttt{go\_mesher} script.
-When you run on a PC cluster, the script assumes that the nodes are
-named n001, n002, etc. If this is not the case, change the \texttt{tr
--d `n'} line in the script. You may also need to edit the last command
-at the end of the script that invokes the \texttt{mpirun} command.
-See Chapter~\ref{cha:Scheduler} for information about running the
-code on a system with a scheduler, e.g., LSF.
-
-Mesher output is provided in the \texttt{OUTPUT\_FILES} directory
-in \texttt{output\_mesher.txt}; this file provides lots of details
-about the mesh that was generated. Please note that the mesher suggests
-a time step \texttt{DT} to run the solver with. The mesher output
-file also contains a table about the quality of the mesh to indicate
-possible problems with the distortions of elements. Alternatively,
-output can be directed to the screen instead by uncommenting a line
-in \texttt{constants.h}:
-\begin{lyxcode}
-!~uncomment~this~to~write~messages~to~the~screen~
-
-!~integer,~parameter~::~IMAIN~=~ISTANDARD\_OUTPUT
-\end{lyxcode}
-To control the quality of the mesh, check the standard output (either
-on the screen or in the \texttt{OUTPUT\_FILES} directory in \texttt{output\_mesher.txt})
-and analyze the skewness histogram of mesh elements to make sure no
-element has a skewness above approximately 0.75, otherwise the mesh
-is of poor quality (and if even a single element has a skewness value
-above 0.80, then you must definitely improve the mesh). To draw the
-skewness histogram on the screen, type \texttt{gnuplot plot\_mesh\_quality\_histogram.gnu}.
-
-\chapter{\label{cha:Creating-Distributed-Databases}Creating the Distributed
-Databases}
-
-After using \texttt{xmeshfem3D} or \texttt{xdecompose\_mesh}, the
-next step in the workflow is to compile \texttt{xgenerate\_}~\\
- \texttt{databases}. This program is going to create all the missing
-information needed by the SEM solver.
-
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=0.6\textwidth]{figures/workflow}
-\par\end{centering}
-
-\caption{Schematic workflow for a SPECFEM3D Cartesian simulation. The executable
-\texttt{xgenerate\_databases} creates the GLL mesh points and assigns
-specific model parameters.}
-
-\label{fig:workflow.databases}
-\end{figure}
-
-\noindent
-In the main directory, type
-\begin{lyxcode}
-{\small make~xgenerate\_databases}{\small \par}
-\end{lyxcode}
-Input for the program is provided through the main parameter file
-\texttt{Par\_file}, which resides in the subdirectory \texttt{DATA}.
-Please note that \texttt{xgenerate\_databases} must be called directly
-from the main directory, as most of the binaries of the package.
-
-\section{\label{cha:Main-Parameter}Main parameter file \texttt{Par\_file}}
-
-Before running \texttt{xgenerate\_databases}, a number of parameters
-need to be set in the main parameter \texttt{Par\_file} located in
-the subdirectory \texttt{DATA}:
-\begin{description}
-\item [{\texttt{SIMULATION\_TYPE}}] is set to 1 for forward simulations,
-2 for adjoint simulations (see Section \ref{sec:Adjoint-simulation-finite})
-and 3 for kernel simulations (see Section \ref{sec:Finite-Frequency-Kernels}).
-\item [{\texttt{SAVE\_FORWARD}}] is only set to \texttt{.true.} for a forward
-simulation with the last frame of the simulation saved, as part of
-the finite-frequency kernel calculations (see Section \ref{sec:Finite-Frequency-Kernels}).
-For a regular forward simulation, leave \texttt{SIMULATION\_TYPE}
-and \texttt{SAVE\_FORWARD} at their default values.
-\item [{\texttt{\noun{UTM\_PROJECTION\_ZONE}}}] UTM projection zone in
-which your model resides, only valid when \texttt{SUPPRESS\_UTM\_}~\\
- \texttt{PROJECTION} is \texttt{.false.}.
-\item [{\texttt{SUPPRESS\_UTM\_PROJECTION}}] set to be \texttt{.false.}
-when your model range is specified in the geographical coordinates,
-and needs to be \texttt{.true.} when your model is specified in a
-cartesian coordinates. \noun{UTM projection zone in which your simulation
-region resides.}
-\item [{\texttt{NPROC}}] The number of MPI processors, each one is assigned
-one slice of the whole mesh.
-\item [{\texttt{NSTEP}}] The number of time steps of the simulation. This
-controls the length of the numerical simulation, i.e., twice the number
-of time steps requires twice as much CPU time. This feature is not
-used at the time of generating the distributed databases but is required
-for the solver, i.e., you may change this parameter after running
-\texttt{xgenerate\_databases}.
-\item [{\texttt{DT}}] The length of each time step in seconds. This feature
-is not used at the time of generating the distributed databases but
-is required for the solver. Please see also Section~\ref{sec:Choosing-the-Time-Step}
-for further details.
-\item [{\texttt{NGNOD}}] The number of nodes for 2D and 3D shape functions
-for hexahedra. We use either 8-node mesh elements (bricks) or 27-node
-elements. If you use the internal mesher, the only option is 8-node
-bricks (27-node elements are not supported). \texttt{CUBIT} does not
-support HEX27 elements either (it can generate them, but they are
-flat, i.e. identical to HEX8). To generate HEX27 elements with curvature
-properly taken into account, you can use \texttt{Gmsh} \url{http://geuz.org/gmsh/}
-\item [{\texttt{MODEL}}] Must be set to one of the following:
-
-\begin{description}
-\item [{\textmd{Models defined by mesh parameters:}}] ~
-
-\begin{description}
-\item [{\texttt{default}}] Uses model parameters as defined by meshing
-procedures described in the previous Chapter~\ref{cha:Mesh-Generation}.
-\end{description}
-
-\item [{\textmd{1D~models~with~real~structure:}}] ~
-
-\begin{description}
-\item [{\texttt{1D\_prem}}] Isotropic version of the spherically symmetric
-Preliminary Reference Earth Model (PREM) \citep{DzAn81}.
-\item [{\texttt{1D\_socal}}] A standard isotropic 1D model for Southern
-California.
-\item [{\texttt{1D\_cascadia}}] Isotropic 1D profile for the Cascadia region.
-\end{description}
-
-\item [{\textmd{Fully~3D~models:}}] ~
-
-\begin{description}
-\item [{\texttt{aniso}}] For a user-specified fully anisotropic model.
-Parameters are set up in routines located in file \texttt{model\_aniso.f90}
-in directory \texttt{src/generate\_databases/}. See Chapter~\ref{cha:-Changing-the}
-for a discussion on how to specify your own 3D model.
-
-\item [{\texttt{external}}] For a user-specified isotropic model which
-uses externally defined model parameters. Uses external model definitions
-set up in routines located in file \texttt{model\_external\_values.f90}
-in directory \texttt{src/generate\_databases/}. Please modify these
-generic template routines to use your own model definitions.
-\item [{\texttt{gll}}] For a user-specified isotropic model which uses
-external binary files for $v_{p}$, $v_{s}$ and $\rho$. Binary files
-are given in the same format as when outputted by the \texttt{xgenerate\_databases}
-executable when using option \texttt{SAVE\_MESH\_FILES}. These binary
-files define the model parameters on all GLL points which can be used
-for iterative inversion procedures.
-\item [{\texttt{salton\_trough}}] A 3D $V_{p}$ model for Southern California.
-Users must provide the corresponding data file \texttt{regrid3\_vel\_p.bin}
-in directory \texttt{DATA/st\_3D\_block\_harvard/}.
-\item [{\texttt{tomo}}] For a user-specified 3D isotropic model which uses
-a tomographic model file \texttt{tomographic\_model.xyz} in directory
-\texttt{DATA}. See Section \ref{sec:Using-tomographic}, for a discussion
-on how to specify your own 3D tomographic model.
-\end{description}
-\end{description}
-\item [{\texttt{APPROXIMATE\_OCEAN\_LOAD}}] Set to \texttt{.true.} if the
-effect of the oceans on seismic wave propagation should be incorporated
-based upon the (rough) approximate treatment discussed in \citet{KoTr02b}.
-This feature is inexpensive from a numerical perspective, both in
-terms of memory requirements and CPU time. This approximation is accurate
-at periods of roughly 20~s and longer. At shorter periods the effect
-of water phases/reverberations is not taken into account, even when
-the flag is on. If you want to model the effect of a fluid-solid model
-at short periods, then set this flag to \texttt{.false.} and mesh
-the fluid layer explicitly in your mesher, so that it is computed
-accurately and without this approximation.
-\item [{\texttt{TOPOGRAPHY}}] This feature is only effective if \texttt{APPROXIMATE\_OCEAN\_LOAD}
-is set to \texttt{.true.}. Set to \texttt{.true.} if topography and
-bathymetry should be read in based upon the topography file specified
-in the main constants file \texttt{constants.h} found in subdirectory
-\texttt{src/shared/} to evaluate elevations. If not set, elevations
-will be read from the numerical mesh.
-\item [{\texttt{ATTENUATION}}] Set to \texttt{.true.} if attenuation should
-be incorporated. Turning this feature on increases the memory requirements
-significantly (roughly by a factor of~1.5), and is numerically fairly
-expensive. See \citet{KoTr99,KoTr02a} for a discussion on the implementation
-of attenuation based upon standard linear solids. Please note that
-the Vp- and Vs-velocities of your model are given for a reference
-frequency. To change this reference frequency, you change the value
-of \texttt{ATTENUATION\_f0\_REFERENCE} in the main constants file
-\texttt{constants.h} found in subdirectory \texttt{src/shared/}.
-\item [{\texttt{FULL\_ATTENUATION\_SOLID}}] Set to \texttt{.true.} to add
-$Q_{\kappa}$ attenuation to the simulations in addition to $Q_{\mu}$
-attenuation when the \texttt{ATTENUATION} flag above is on. Should
-be off in almost all cases for seismic wave propagation, since $Q_{\kappa}$
-is almost always negligible in Earth models. However, this parameter
-can be useful for instance for ocean acoustics simulations, when $Q_{\kappa}$
-is not negligible in the solid layer at the sea bottom.
-\item [{\texttt{ANISOTROPY}}] Set to \texttt{.true.} if you want to use
-an anisotropy model. Please see the file \texttt{model\_aniso.f90}
-in subdirectory \texttt{src/generate\_databases/} for the current
-implementation of anisotropic models.
-\item [{\texttt{TOMOGRAPHY\_PATH}}] Directory in which the tomography files
-are stored for using external tomographic Earth models (please read
-Chapter \ref{cha:-Changing-the} and Section \ref{sec:Using-tomographic}
-`Using external tomographic Earth models' for further details.).
-\item [{\texttt{USE\_OLSEN\_ATTENUATION}}] Set to \texttt{.true.} if you
-want to use the attenuation model that scaled from the S-wave speed
-model using Olsen's empirical relation (see \citet{OlDaBr03}).
-\item [{\texttt{OLSEN\_ATTENUATION\_RATIO}}] Determines the Olsen's constant
-in Olsen's empirical relation (see \citet{OlDaBr03}).
-\item [{\texttt{PML\_CONDITIONS}}] Set to \texttt{.true.} to turn on C-PML
-boundary conditions for a regional simulation. Both fluids and elastic
-solids are supported. Note that xmeshfem3d generated meshes do not support C-PML yet.
-\item [{\texttt{PML\_INSTEAD\_OF\_FREE\_SURFACE}}] Set to \texttt{.true.}
-to turn on C-PML boundary conditions on the top surface instead of
-the usual free surface.
-\item [{\texttt{f0\_FOR\_PML}}] Determines the dominant frequency that
-will be used in the calculation of PML damping profiles; this should
-be set to the same (or similar) dominant frequency as that of the
-source that you will use in your simulation. If you plan to use a
-Dirac source, then use the dominant frequency of the source wavelet
-with which you plan to convolve your seismograms later on in post-processing.
-\item [{\texttt{STACEY\_ABSORBING\_CONDITIONS}}] Set to \texttt{.true.}
-to turn on Clayton-Enquist absorbing boundary conditions (see \citet{KoTr99}).
-In almost all cases it is much better to use CPML absorbing layers
-(see the options above) and leave this flag to \texttt{.false.}.
-\item [{\texttt{STACEY\_INSTEAD\_OF\_FREE\_SURFACE}}] Set to \texttt{.true.}
-to turn on absorbing boundary conditions on the top surface which
-by default constitutes a free surface of the model.
-\item [{\texttt{CREATE\_SHAKEMAP}}] Set this flag to \texttt{.true.} to
-create a ShakeMap\textregistered{}, i.e., a peak ground velocity map
-of the maximum absolute value of the two horizontal components of
-the velocity vector.
-\item [{\texttt{MOVIE\_SURFACE}}] Set to \texttt{.false.}, unless you want
-to create a movie of seismic wave propagation on the Earth's surface.
-Turning this option on generates large output files. See Section~\ref{sec:Movies}
-for a discussion on the generation of movies. This feature is only
-relevant for the solver.
-\item [{\texttt{MOVIE\_TYPE}}] Set this flag to 1 to show the top surface
-(tomography + oceans) only, to 2 to show all external faces of the
-mesh (i.e. topography + vertical edges + bottom) in shakemaps and
-surface movies.
-\item [{\texttt{MOVIE\_VOLUME}}] Set to \texttt{.false.}, unless you want
-to create a movie of seismic wave propagation in the Earth's interior.
-Turning this option on generates huge output files. See Section~\ref{sec:Movies}
-for a discussion on the generation of movies. This feature is only
-relevant for the solver.
-\item [{\texttt{SAVE\_DISPLACEMENT}}] Set this flag to \texttt{.true.}
-if you want to save the displacement instead of velocity for the movie
-frames.
-\item [{\texttt{USE\_HIGHRES\_FOR\_MOVIES}}] Set this flag to \texttt{.true.}
-if you want to save the values at all the NGLL grid points for the
-movie frames.
-\item [{\texttt{NTSTEP\_BETWEEN\_FRAMES}}] Determines the number of timesteps
-between movie frames. Typically you want to save a snapshot every
-100 timesteps. The smaller you make this number the more output will
-be generated! See Section~\ref{sec:Movies} for a discussion on the
-generation of movies. This feature is only relevant for the solver.
-\item [{\texttt{HDUR\_MOVIE}}] Determines the half duration of the source
-time function for the movie simulations. When this parameter is set
-to be 0, a default half duration that corresponds to the accuracy
-of the simulation is provided. Otherwise, it adds this half duration
-to the half duration specified in the source file \texttt{CMTSOLUTION},
-thus simulates longer periods to make the movie images look smoother.
-\item [{\texttt{SAVE\_MESH\_FILES}}] Set this flag to \texttt{.true.} to
-save ParaView \urlwithparentheses{www.paraview.org} mesh files for
-subsequent viewing. Turning the flag on generates large (distributed)
-files in the \texttt{LOCAL\_PATH} directory. See Section~\ref{sec:Mesh-graphics}
-for a discussion of mesh viewing features.
-\item [{\texttt{LOCAL\_PATH}}] Directory in which the distributed databases
-will be written. Generally one uses a directory on the local disk
-of the compute nodes, although on some machines these databases are
-written on a parallel (global) file system (see also the earlier discussion
-of the \texttt{LOCAL\_PATH\_IS\_ALSO\_GLOBAL} flag in Chapter~\ref{cha:Getting-Started}).
-\texttt{xgenerate\_databases} generates the necessary databases in
-parallel, one set for each of the \texttt{NPROC} slices that constitutes
-the mesh (see Figure~\ref{fig:mount.partitions} and Figure~\ref{fig:For-parallel-computing}).
-After the executable finishes, you can log in to one of the compute
-nodes and view the contents of the \texttt{LOCAL\_PATH} directory
-to see the (many) files generated by \texttt{xgenerate\_databases}.
-Please note that the \texttt{LOCAL\_PATH} directory should already
-contain the output files of the partitioner, i.e. from \texttt{xdecompose\_mesh}
-or \texttt{xmeshfem3D}.
-\item [{\texttt{NTSTEP\_BETWEEN\_OUTPUT\_INFO}}] This parameter specifies
-the interval at which basic information about a run is written to
-the file system (\texttt{timestamp{*}} files in the \texttt{OUTPUT\_FILES}
-directory). If you have access to a fast machine, set \texttt{NTSTEP\_BETWEEN\_OUTPUT\_INFO}
-to a relatively high value (e.g., at least 100, or even 1000 or more)
-to avoid writing output text files too often. This feature is not
-used at the time of meshing. One can set this parameter to a larger
-value than the number of time steps to avoid writing output during
-the run.
-\item [{\texttt{NTSTEP\_BETWEEN\_OUTPUT\_SEISMOS}}] This parameter specifies
-the interval at which synthetic seismograms are written in the \texttt{LOCAL\_PATH}
-directory. If a run crashes, you may still find usable (but shorter
-than requested) seismograms in this directory. On a fast machine set
-\texttt{NTSTEP\_BETWEEN\_OUTPUT\_SEISMOS} to a relatively high value
-to avoid writing to the seismograms too often. This feature is only
-relevant for the solver.
-\item [{\texttt{USE\_FORCE\_POINT\_SOURCE}}] Turn this flag on to use a
-(tilted) \texttt{FORCESOLUTION} force point source instead of a \texttt{CMTSOLUTION}
-moment-tensor source. When the force source does not fall exactly
-at a grid point, the solver interpolates the force between grid points
-using Lagrange interpolants. This can be useful e.g. for oil industry
-foothills simulations in which the source is a vertical force, normal
-force, tilted force, or an impact etc. Note that in the \texttt{FORCESOLUTION}
-file, you will need to edit the East, North and vertical components
-of an arbitrary (not necessarily unitary, the code will normalize it automatically) direction vector of the force vector;
-thus refer to Appendix~\ref{cha:Coordinates} for the orientation
-of the reference frame. This vector is made unitary internally in
-the solver and thus only its direction matters here; its norm is ignored
-and the norm of the force used is the factor force source times the
-source time function.
-When using this option, by default the code can locate the force source
-anywhere between mesh points in order to honor its exact location; this
-is more precise than using the closest GLL mesh point, but it is also a bit slower.
-If needed, you can change that default behavior and force the code to use the closest
-GLL mesh point instead by setting flag \texttt{USE\_BEST\_LOCATION} to \texttt{.false.} instead of \texttt{.true.} in file
-\texttt{src/shared/constants.h.in} and running the \texttt{configure} script again and recompiling the code.
-\item [{\texttt{USE\_RICKER\_TIME\_FUNCTION}}] Turn this flag on to use
-a Ricker source time function instead of the source time functions
-set by default to represent a (tilted) \texttt{FORCESOLUTION} force
-point source or a \texttt{CMTSOLUTION} moment-tensor source. Originally,
-if a \texttt{CMTSOLUTION} moment-tensor source is used, a (pseudo)
-Heaviside step function with a very short half duration is defined
-for elastic cases to represent the permanent slip on the fault while
-in the acoustic case a Gaussian source time function with a similarly
-short half duration is defined to physically describe actions within
-the fluid. Otherwise, if a \texttt{FORCESOLUTION} force source is
-used, a (pseudo) Dirac delta source time function is defined by default.
-Any other source-time function may then be obtained by convolution.
-\item [{\texttt{PRINT\_SOURCE\_TIME\_FUNCTION}}] Turn this flag on to print
-information about the source time function in the file \texttt{OUTPUT\_FILES/plot\_source\_time\_function.txt}.
-This feature is only relevant for the solver.
-
-\item [{\texttt{NUMBER\_OF\_SIMULTANEOUS\_RUNS}}] adds the ability to run several calculations (several earthquakes)
-in an embarrassingly-parallel fashion from within the same run;
-this can be useful when using a very large supercomputer to compute
-many earthquakes in a catalog, in which case it can be better from
-a batch job submission point of view to start fewer and much larger jobs,
-each of them computing several earthquakes in parallel.
-To turn that option on, set parameter NUMBER\_OF\_SIMULTANEOUS\_RUNS to a value greater than 1.
-To implement that, we create NUMBER\_OF\_SIMULTANEOUS\_RUNS MPI sub-communicators,
-each of them being labeled "my\_local\_mpi\_comm\_world", and we use them
-in all the routines in "src/shared/parallel.f90", except in MPI\_ABORT() because in that case
-we need to kill the entire run.
-When that option is on, of course the number of processor cores used to start
-the code in the batch system must be a multiple of NUMBER\_OF\_SIMULTANEOUS\_RUNS,
-all the individual runs must use the same number of processor cores,
-which as usual is NPROC in the Par\_file,
-and thus the total number of processor cores to request from the batch system
-should be NUMBER\_OF\_SIMULTANEOUS\_RUNS * NPROC.
-All the runs to perform must be placed in directories called run0001, run0002, run0003 and so on
-(with exactly four digits).
-
-Imagine you have 10 independent calculations to do, each of them on 100 cores; you have three options:
-
-1/ submit 10 jobs to the batch system
-
-2/ submit a single job on 1000 cores to the batch, and in that script create a sub-array of jobs to start 10 jobs,
-each running on 100 cores (see e.g. http://www.schedmd.com/slurmdocs/job\_array.html )
-
-3/ submit a single job on 1000 cores to the batch, start SPECFEM3D on 1000 cores, create 10 sub-communicators,
-cd into one of 10 subdirectories (called e.g. run0001, run0002,... run0010) depending on the sub-communicator
-your MPI rank belongs to, and run normally on 100 cores using that sub-communicator.
-
-The option NUMBER\_OF\_SIMULTANEOUS\_RUNS implements 3/.
-
-\item [{\texttt{BROADCAST\_SAME\_MESH\_AND\_MODEL}}]:
-if we perform simultaneous runs in parallel, if only the source and receivers vary between these runs
-but not the mesh nor the model (velocity and density) then we can also read the mesh and model files
-from a single run in the beginning and broadcast them to all the others; for a large number of simultaneous
-runs for instance when solving inverse problems iteratively this can DRASTICALLY reduce I/Os to disk in the solver
-(by a factor equal to NUMBER\_OF\_SIMULTANEOUS\_RUNS), and reducing I/Os is crucial in the case of huge runs.
-Thus, always set this option to .true. if the mesh and the model are the same for all simultaneous runs.
-In that case there is no need to duplicate the mesh and model file database (the content of the DATABASES\_MPI
-directories) in each of the run0001, run0002,... directories, it is sufficient to have one in run0001
-and the code will broadcast it to the others).
-
-TODO / future work to do: currently the BROADCAST\_SAME\_MESH\_AND\_MODEL option assumes to have the (master) mesh files in run0001/DATABASES\_MPI or run0001/OUTPUT\_FILES/DATABASES\_MPI.
-However, for adjoint runs you still need a DATABASES\_MPI folder in each of the sub-runs directories, e.g.
-run0002/DATABASES\_MPI, etc.
-to store the forward wavefields, kernels etc.
-of each sub-run.
-This would not be needed for forward simulations.
-
-TODO / future work to do: the sensitivity kernel summing and smoothing tools in directory src/tomography are currently not ported to this new option to do many runs simultaneously, only the solver (src/specfem3d) is.
-Thus these tools should work, but in their current version will need to be run for each simulation result independently.
-More precisely, the current kernel summing and smoothing routines work fine, with the exception that you need to move out the mesh files (and also the parameters).
-This works because these routines consider multiple runs by design.
-You simply have to provide them the directories where the kernels are.
-
-\item [{\texttt{GPU\_MODE}}] Turn this flag on to use GPUs.
-
-\item [\texttt{ADIOS\_ENABLED}] Turn this flag on to enable ADIOS. If set to \texttt{.false.}, subsequent ADIOS
-parameters will not be considered.
-\item [\texttt{ADIOS\_FOR\_DATABASES}] Turn this flag on to use ADIOS for xmeshfem3D output and
-xgenerate\_database input.
-\item [\texttt{ADIOS\_FOR\_MESH}]  Turn this flag on to use ADIOS for generated databases.
-\item [\texttt{ADIOS\_FOR\_FORWARD\_ARRAYS}] Turn this flag on to read and write forward arrays using ADIOS.
-\item [\texttt{ADIOS\_FOR\_KERNELS}] Turn this flag on to produce ADIOS kernels that can later be visualized with the ADIOS version of combine\_vol\_data.
-
-\end{description}
-
-The present version of SPECFEM can handle fully saturated porous simulations, Christina Morency implemented Biot equation.
-But the code cannot calculate partially saturated cases in its state.
-Christina Morency is presently working on a dual porosity, dual permeability formulation, type Pride and Berryman, for an other project, but it will not be available for some time.\\
-
-The way we prescribe material property for porous material in SPECFEM3D is as follow:
-We use a file name nummaterial\_poroelastic\_file, which is located in the directory MESH, the format is as follow:\\
- rhos rhof phi c kxx kxy kxz kyy kyz kzz Ks Kf Kfr etaf mufr\\
-where\\
-rho\_s = solid density\\
-rho\_f = fluid density\\
-phi = porosity\\
-tort = tortuosity\\
-kxx = xx component of permeability tensor\\
-kxy = xy,yx components of permeability tensor\\
-kyy = yy component of permeability tensor\\
-kxz = xz,zx components of permeability tensor\\
-kzz = zz component of permeability tensor\\
-kappa\_s = solid bulk modulus\\
-kappa\_f= fluid bulk modulus\\
-kappa\_fr= frame bulk modulus\\
-eta\_f = fluid viscosity\\
-mu\_fr = frame shear modulus.\\
-
-Using an external mesh (for instance coming from CUBIT/TRELIS), poroelastic materials have the ID number 3, while 1 is elastic, 2 is acoustic ?
-see the example in the package: examples/homogeneous\_poroelastic.
-
-If you use PML, the mesh elements that belong to the PML layers can
-be acoustic or elastic, but not viscoelastic nor poroelastic. Then,
-when defining your model, you should define these absorbing elements
-as either acoustic or elastic. In you forget to do that, the code
-will fix the problem by automatically converting the viscoelastic
-or poroelastic PML elements to elastic. This means that strictly speaking
-the PML layer will not be perfectly matched any more, since the physical
-model will change from viscoelastic or poroelastic to elastic at the
-entrance of the PML, but in practice this is sufficient and produces
-only tiny / negligible spurious reflections.
-
-If you use PML and an external mesh (created using an external meshing tool
-such as CUBIT/TRELIS or similar), try to have elements inside the PML as regular as possible,
-i.e. ideally non-deformed cubes obtained by `extrusion' of regular surface mesh elements meshing the
-outer edges of the computational domain without PML; by doing so, the PMLs obtained will be far more stable
-in time (PML being weakly unstable from a mathematical point of view, very deformed mesh elements
-inside the PMLs can trigger instabilities much more quickly).
-
-If you use PML and an external tomographic velocity and density model,
-you should be careful because mathematically a PML cannot handle heterogeneities
-along the normal to the PML edge inside the PML layer. This comes
-from the fact that the damping profile that is defined assumes a constant
-velocity and density model along the normal direction.
-
-Thus, you need to modify your velocity and density model in order
-for it to be 1D inside the PML, as shown in Figure~\ref{fig:modify_external_velocity_model_to_use_PML}.
-
-This applies to the bottom layer as well; there you should make sure
-that your model is 1D and thus constant along the vertical direction.
-
-To summarize, only use a 3D velocity and density model inside the
-physical region, and in all the PML layers extend it by continuity
-from its values along the inner PML edge.
-
-%%
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=3in]{figures/how_to_use_PML_when_a_tomographic_velocity_model_is_used}
-\par\end{centering}
-
-\caption{How to modify your external 3D velocity and density model in order
-to use PML. Such a modification is not needed when using Stacey absorbing
-boundary conditions (but such conditions are significantly less efficient).}
-
-\label{fig:modify_external_velocity_model_to_use_PML}
-\end{figure}
-
-%%
-
-\section{\label{sec:Choosing-the-Time-Step}Choosing the time step \texttt{DT}}
-
-The parameter \texttt{DT} sets the length of each time step in seconds.
-The value of this parameter is crucial for the stability of the spectral-element
-simulation. Your time step \texttt{DT} will depend on the minimum
-ratio between the distance $h$ of neighboring mesh points and the
-wave speeds $v$ defined in your model. The condition for the time
-step $\Delta t$ is:
-\begin{lyxcode}
-$\Delta t<C~\mathrm{min}_{\Omega}(~h/v~)$
-\end{lyxcode}
-where $C$ is the so-called Courant number and $\Omega$ denotes the
-model volume. The distance $h$ depends on the mesh element size and
-the number of GLL points \texttt{NGLL} specified in the main constants
-file \texttt{constants.h} located in the \texttt{src/shared/} subdirectory.
-The wave speed $v$ is determined based on your model's P- (or S-)
-wave speed values.
-
-The database generator \texttt{xgenerate\_databases}, as well as the
-internal mesher \texttt{xmeshfem3D}, are trying to evaluate the value
-of $\Delta t$ for empirically chosen Courant numbers $C\sim0.3$.
-If you used the mesher \texttt{xmeshfem3D} to generate your mesh,
-you should set the value suggested in \texttt{OUTPUT\_FILES/output\_mesher.txt}
-file, which is created after the mesher completed. In case you used
-CUBIT to create the mesh, you might use an arbitrary value when running
-\texttt{xgenerate\_databases} and then use the value suggested in
-the ~\\
- \texttt{OUTPUT\_FILES/output\_mesher.txt} file after the database
-generation completed. Note that the implemented Newmark time scheme
-uses this time step globally, thus your simulations become more expensive
-for very small mesh elements in high wave-speed regions. Please be
-aware of this restriction when constructing your mesh in Chapter~\ref{cha:Mesh-Generation}.
-
-\chapter{\label{cha:Running-the-Solver}Running the Solver \texttt{xspecfem3D}}
-
-Now that you have successfully generated the databases, you are ready
-to compile the solver. In the main directory, type
-\begin{lyxcode}
-{\small make~xspecfem3D}{\small \par}
-\end{lyxcode}
-Please note that \texttt{xspecfem3D} must be called directly from
-the main directory, as most of the binaries of the package.\\
-
-\noindent
-The solver needs three input files in the \texttt{DATA} directory
-to run:
-\begin{description}
-\item [{\texttt{Par\_file}}] the main parameter file which was discussed
-in detail in the previous Chapter~\ref{cha:Creating-Distributed-Databases},
-\item [{\texttt{CMTSOLUTION}} or {\texttt{FORCESOLUTION}}] the earthquake
-source parameter file or the force source parameter file, and
-\item [{\texttt{STATIONS}}] the stations file.
-\end{description}
-Most parameters in the \texttt{Par\_file} should be set prior to running
-the databases generation. Only the following parameters may be changed
-after running \texttt{xgenerate\_databases}:
-\begin{itemize}
-\item the simulation type control parameters: \texttt{SIMULATION\_TYPE}
-and \texttt{SAVE\_FORWARD}
-\item the time step parameters \texttt{NSTEP} and \texttt{DT}
-\item the absorbing boundary control parameter \texttt{PML\_CONDITIONS}
-on condition that the\\
- \texttt{PML\_INSTEAD\_OF\_FREE\_SURFACE} flag remains unmodified
-after running the databases generation.
-\item the movie control parameters \texttt{MOVIE\_SURFACE}, \texttt{MOVIE\_VOLUME},
-and \texttt{NTSTEPS\_BETWEEN\_FRAMES}
-\item the ShakeMap\textregistered{}option \texttt{CREATE\_SHAKEMAP}
-\item the output information parameters \texttt{MOVIE\_TYPE}, \texttt{NTSTEP\_BETWEEN\_OUTPUT\_INFO}
-and \texttt{NTSTEP\_BETWEEN\_OUTPUT\_}~\\
- \texttt{SEISMOS}
-\item the \texttt{PRINT\_SOURCE\_TIME\_FUNCTION} flags
-\end{itemize}
-Any other change to the \texttt{Par\_file} implies rerunning both
-the database generator \texttt{xgenerate\_databases} and the solver
-\texttt{xspecfem3D}.
-
-For any particular earthquake, the \texttt{CMTSOLUTION} file that
-represents the point source may be obtained directly from the Harvard
-Centroid-Moment Tensor (CMT) web page \urlwithparentheses{www.seismology.harvard.edu}.
-It looks like the example shown in Fig. \ref{fig:CMTSOLUTION-file}.
-%\begin{lyxcode}
-\begin{figure}[htp]
-\noindent
-\begin{centering}
-{\small \includegraphics[width=0.8\textwidth]{figures/Hollywood_CMT}
-}
-\par
-\end{centering}
-\caption{\texttt{CMTSOLUTION} file based on the
-format from the Harvard CMT catalog. \textbf{M} is the moment tensor,
-$M_{0}${\small {} }is the seismic moment, and $M_{w}$ is the moment
-magnitude.}
-\label{fig:CMTSOLUTION-file}
-\end{figure}
-%\end{lyxcode}
-The \texttt{CMTSOLUTION} file should be edited in the following way:
-\begin{itemize}
-\item Set the latitude or UTM $x$ coordinate, longitude or UTM $y$ coordinate, depth of the source (in km).
-\item Set the \texttt{time shift} parameter equal to $0.0$ (the solver
-will not run otherwise.) The time shift parameter would simply apply
-an overall time shift to the synthetics, something that can be done
-in the post-processing (see Section \ref{sec:Process-data-and-syn}).
-\item For point-source simulations (see finite sources, page \pageref{To-simulate-a})
-we recommend setting the source half-duration parameter \texttt{half
-duration} equal to zero, which corresponds to simulating a step source-time
-function, i.e., a moment-rate function that is a delta function. If
-\texttt{half duration} is not set to zero, the code will use a Gaussian
-(i.e., a signal with a shape similar to a `smoothed triangle', as
-explained in \citet{KoTr02a} and shown in Fig~\ref{fig:gauss.vs.triangle})
-source-time function with half-width \texttt{half duration}. We prefer
-to run the solver with \texttt{half duration} set to zero and convolve
-the resulting synthetic seismograms in post-processing after the run,
-because this way it is easy to use a variety of source-time functions
-(see Section \ref{sec:Process-data-and-syn}). \citet{KoTr02a} determined
-that the noise generated in the simulation by using a step source
-time function may be safely filtered out afterward based upon a convolution
-with the desired source time function and/or low-pass filtering. Use
-the serial code \texttt{convolve\_source\_timefunction.f90} and the
-script \texttt{convolve\_source\_timefunction.csh} for this purpose,
-or alternatively use signal-processing software packages such as SAC
-\urlwithparentheses{www.llnl.gov/sac}. Type
-
-\begin{lyxcode}
-{\small make~xconvolve\_source\_timefunction}{\small \par}
-\end{lyxcode}
-
-to compile the code and then set the parameter \texttt{hdur} in \texttt{convolve\_source\_timefunction.csh}
-to the desired half-duration.
-
-\item The zero time of the simulation corresponds to the center of the triangle/Gaussian,
-or the centroid time of the earthquake. The start time of the simulation
-is $t=-1.5*\texttt{half duration}$ (the 1.5 is to make sure the moment
-rate function is very close to zero when starting the simulation).
-To convert to absolute time $t_{\mathrm{abs}}$, set
-
-\begin{lyxcode}
-$t_{\mathrm{abs}}=t_{\mathrm{pde}}+\texttt{time shift}+t_{\mathrm{synthetic}}$
-\end{lyxcode}
-
-where $t_{\mathrm{pde}}$ is the time given in the first line of the
-\texttt{CMTSOLUTION}, \texttt{time shift} is the corresponding value
-from the original \texttt{CMTSOLUTION} file and $t_{\mathrm{synthetic}}$
-is the time in the first column of the output seismogram.
-
-\end{itemize}
-\begin{figure}[htp]
-\noindent \begin{centering}
-\includegraphics[width=3.5in]{figures/gauss_vs_triangle_mod}
-\par\end{centering}
-
-\caption{Comparison of the shape of a triangle and the Gaussian function actually
-used.}
-
-\label{fig:gauss.vs.triangle}
-\end{figure}
-
-If you know the earthquake source in strike/dip/rake format rather
-than in \texttt{CMTSOLUTION} format, use the C code \texttt{SPECFEM3D\_GLOBE/utils/strike\_dip\_rake\_to\_CMTSOLUTION.c}
-to convert it. The conversion formulas are given for instance in \citet{AkRi80}.
-Note that the \citet{AkRi80} convention is slightly different from
-the Harvard \texttt{CMTSOLUTION} convention (the sign of some components
-is different). The C code outputs both.
-
-Centroid latitude and longitude should be provided in geographical
-coordinates. The code converts these coordinates to geocentric coordinates~\citep{DaTr98}.
-Of course you may provide your own source representations by designing
-your own \texttt{CMTSOLUTION} file. Just make sure that the resulting
-file adheres to the Harvard CMT conventions (see Appendix~\ref{cha:Coordinates}).
-Note that the first line in the \texttt{CMTSOLUTION} file is the Preliminary
-Determination of Earthquakes (PDE) solution performed by the USGS
-NEIC, which is used as a seed for the Harvard CMT inversion. The PDE
-solution is based upon P waves and often gives the hypocenter of the
-earthquake, i.e., the rupture initiation point, whereas the CMT solution
-gives the `centroid location', which is the location with dominant
-moment release. The PDE solution is not used by our software package
-but must be present anyway in the first line of the file.
-
-\label{To-simulate-a}To simulate a kinematic rupture, i.e., a finite-source
-event, represented in terms of $N_{\mathrm{sources}}$ point sources,
-provide a \texttt{CMTSOLUTION} file that has $N_{\mathrm{sources}}$
-entries, one for each subevent (i.e., concatenate $N_{\mathrm{sources}}$
-\texttt{CMTSOLUTION} files to a single \texttt{CMTSOLUTION} file).
-At least one entry (not necessarily the first) must have a zero \texttt{time
-shift}, and all the other entries must have non-negative \texttt{time
-shift}. Each subevent can have its own half duration, latitude, longitude,
-depth, and moment tensor (effectively, the local moment-density tensor).
-
-Note that the zero in the synthetics does NOT represent the hypocentral
-time or centroid time in general, but the timing of the \textit{center}
-of the source triangle with zero \texttt{time shift} (Fig~\ref{fig:source_timing}).
-
-Although it is convenient to think of each source as a triangle, in
-the simulation they are actually Gaussians (as they have better frequency
-characteristics). The relationship between the triangle and the gaussian
-used is shown in Fig~\ref{fig:gauss.vs.triangle}. For finite fault
-simulations it is usually not advisable to use a zero half duration
-and convolve afterwards, since the half duration is generally fixed
-by the finite fault model.
-
-\vspace{1cm}
-
-\noindent The \texttt{FORCESOLUTION} file should be edited in the
-following way:
-\begin{itemize}
-\item Set the \texttt{time shift} parameter equal to $0.0$ (the solver
-will not run otherwise.) The time shift parameter would simply apply
-an overall time shift to the synthetics, something that can be done
-in the post-processing (see Section \ref{sec:Process-data-and-syn}).
-\item Set the \texttt{f0} parameter (the dominant frequency) of the Ricker source time
-function when {\texttt{USE\_RICKER\_TIME\_FUNCTION}} is turned on
-in the main parameter file {\texttt{Par\_file}}. In case that the
-solver uses a (pseudo) Dirac delta source time function to represent
-a force point source, a very short duration of five time steps is automatically set by default.
-\item Set the latitude or UTM $x$ coordinate, longitude or UTM $y$ coordinate, depth of the source (in km).
-\item Set the magnitude of the force source.
-\item Set the components of a (non-unitary) direction vector for the force
-source in the East/North/Vertical basis (see Appendix A for the orientation
-of the reference frame).
-\end{itemize}
-\noindent Where necessary, set a \texttt{FORCESOLUTION} file in the
-same way you configure a \texttt{CMTSOLUTION} file with $N_{\mathrm{sources}}$
-entries, one for each subevent (i.e., concatenate $N_{\mathrm{sources}}$
-\texttt{FORCESOLUTION} files to a single \texttt{FORCESOLUTION} file).
-At least one entry (not necessarily the first) must have a zero \texttt{time
-shift}, and all the other entries must have non-negative \texttt{time
-shift}. Each subevent can have its own half latitude, longitude, depth,
-\texttt{half duration} and force parameters.
-
-\begin{figure}[H]
-\noindent \begin{centering}
-{\small \includegraphics[width=5in]{figures/source_timing} }
-\par\end{centering}
-
-\caption{Example of timing for three sources. The center of the first source
-triangle is defined to be time zero. Note that this is NOT in general
-the hypocentral time, or the start time of the source (marked as tstart).
-The parameter \texttt{time shift} in the \texttt{CMTSOLUTION} file
-would be t1(=0), t2, t3 in this case, and the parameter \texttt{half
-duration} would be hdur1, hdur2, hdur3 for the sources 1, 2, 3 respectively.}
-
-{\small \label{fig:source_timing} }
-\end{figure}
-
-The solver can calculate seismograms at any number of stations for
-basically the same numerical cost, so the user is encouraged to include
-as many stations as conceivably useful in the \texttt{STATIONS} file,
-which looks like this:
-
-\begin{figure}[H]
-\noindent \begin{centering}
-{\small \includegraphics{figures/STATIONS_basin_explained} }
-\par\end{centering}
-
-\label{fig:Sample-STATIONS-file.}\caption{Sample \texttt{STATIONS} file. Station latitude and longitude should
-be provided in geographical coordinates. The width of the station
-label should be no more than 32 characters (see \texttt{MAX\_LENGTH\_STATION\_NAME}
-in the \texttt{constants.h} file), and the network label should be
-no more than 8 characters (see \texttt{MAX\_LENGTH\_NETWORK\_NAME}
-in the \texttt{constants.h} file).}
-\end{figure}
-
-Each line represents one station in the following format:
-\begin{lyxcode}
-{\small Station~Network~Latitude~(degrees)~Longitude~(degrees)~Elevation~(m)~burial~(m)~}{\small \par}
-\end{lyxcode}
-The solver \texttt{xspecfem3D} filters the list of stations in file
-\texttt{DATA/STATIONS} to exclude stations that are not located within
-the region given in the \texttt{Par\_file} (between \texttt{LATITUDE\_MIN}
-and \texttt{LATITUDE\_MAX} and between \texttt{LONGITUDE\_MIN} and
-\texttt{LONGITUDE\_MAX}). The filtered file is called \texttt{DATA/STATIONS\_FILTERED}.
-
-Solver output is provided in the \texttt{OUTPUT\_FILES} directory
-in the \texttt{output\_solver.txt} file. Output can be directed to
-the screen instead by uncommenting a line in \texttt{constants.h}:
-\begin{lyxcode}
-!~uncomment~this~to~write~messages~to~the~screen~
-
-!~integer,~parameter~::~IMAIN~=~ISTANDARD\_OUTPUT~
-\end{lyxcode}
-On PC clusters the seismogram files are generally written to the local
-disks (the path \texttt{LOCAL\_PATH} in the \texttt{Par\_file}) and
-need to be gathered at the end of the simulation.
-
-While the solver is running, its progress may be tracked by monitoring
-the `\texttt{\small timestamp{*}}{\small ' files in the ~}\\
-{\small{} }\texttt{\small OUTPUT\_FILES}{\small{} directory. These tiny
-files look something like this:}{\small \par}
-\begin{lyxcode}
-Time~step~\#~~~~~~~~~~10000~
-
-Time:~~~~~108.4890~~~~~~seconds~
-
-Elapsed~time~in~seconds~=~~~~1153.28696703911~
-
-Elapsed~time~in~hh:mm:ss~=~~~~~0~h~19~m~13~s~
-
-Mean~elapsed~time~per~time~step~in~seconds~=~~~~~0.115328696703911~
-
-Max~norm~displacement~vector~U~in~all~slices~(m)~=~~~~1.0789589E-02~
-\end{lyxcode}
-The \texttt{\small timestamp{*}}{\small{} files provide the }\texttt{\small Mean
-elapsed time per time step in seconds}{\small , which may be used
-to assess performance on various machines (assuming you are the only
-user on a node), as well as the }\texttt{\small Max norm displacement
-vector U in all slices~(m)}{\small . If something is wrong with the
-model, the mesh, or the source, you will see the code become unstable
-through exponentially growing values of the displacement and fluid
-potential with time, and ultimately the run will be terminated by
-the program. You can control the rate at which the timestamp files
-are written based upon the parameter }\texttt{\small NTSTEP\_BETWEEN\_OUTPUT\_INFO}{\small{}
-in the }\texttt{\small Par\_file}{\small .}{\small \par}
-
-Having set the \texttt{Par\_file} parameters, and having provided
-the \texttt{CMTSOLUTION} (or the \texttt{FORCESOLUTION}) and \texttt{STATIONS}
-files, you are now ready to launch the solver! This is most easily
-accomplished based upon the \texttt{go\_solver} script (See Chapter
-\ref{cha:Scheduler} for information about running through a scheduler,
-e.g., LSF). You may need to edit the last command at the end of the
-script that invokes the \texttt{mpirun} command. The \texttt{runall}
-script compiles and runs both \texttt{xgenerate\_databases} and \texttt{xspecfem3D}
-in sequence. This is a safe approach that ensures using the correct
-combination of distributed database output and solver input.
-
-It is important to realize that the CPU and memory requirements of
-the solver are closely tied to choices about attenuation (\texttt{ATTENUATION})
-and the nature of the model (i.e., isotropic models are cheaper than
-anisotropic models). We encourage you to run a variety of simulations
-with various flags turned on or off to develop a sense for what is
-involved.
-
-For the same model, one can rerun the solver for different events
-by simply changing the \texttt{CMTSOLUTION} or \texttt{FORCESOLUTION}
-file, or for different stations by changing the \texttt{STATIONS}
-file. There is no need to rerun the \texttt{xgenerate\_databases}
-executable. Of course it is best to include as many stations as possible,
-since this does not add to the cost of the simulation.
-
-We have also added the ability to run several calculations (several earthquakes)
-in an embarrassingly-parallel fashion from within the same run;
-this can be useful when using a very large supercomputer to compute
-many earthquakes in a catalog, in which case it can be better from
-a batch job submission point of view to start fewer and much larger jobs,
-each of them computing several earthquakes in parallel.
-To turn that option on, set parameter \texttt{NUMBER\_OF\_SIMULTANEOUS\_RUNS}
-to a value greater than 1 in file setup/constants.h.in before
-configuring and compiling the code.
-When that option is on, of course the number of processor cores used to start
-the code in the batch system must be a multiple of \texttt{NUMBER\_OF\_SIMULTANEOUS\_RUNS},
-all the individual runs must use the same number of processor cores,
-which as usual is \texttt{NPROC} in the input file \texttt{DATA/Par\_file},
-and thus the total number of processor cores to request from the batch system
-should be \texttt{NUMBER\_OF\_SIMULTANEOUS\_RUNS $\times$ NPROC}.
-All the runs to perform must be placed in directories called \texttt{run0001}, \texttt{run0002}, \texttt{run0003} and so on (with exactly four digits)
-and you must create a link from the root directory of the code
-to the first copy of the executable programs by typing \texttt{ln -s run0001/bin bin}.
-
-\chapter{\label{cha:Fault-Kinematics-Dynamics}Kinematic and dynamic fault
-sources}
-
-SPECFEM3D can handle finite fault sources of two kinds:
-\begin{enumerate}
-\item \emph{Kinematic}: the spatio-temporal distribution of slip rate is
-prescribed along the fault surface
-\item \emph{Dynamic}: a friction law and initial stresses are prescribed
-on the fault, a spontaneous rupture process is computed.
-\end{enumerate}
-
-\section{\label{sec:Mesh-Generation-with}Mesh Generation with Split Nodes}
-
-Faults need to be handled in a special way during mesh generation.
-A fault surface must lie at the interface between elements (the mesh
-must honor the fault surfaces). Moreover, a fault is made of two surfaces
-in contact. Each of these two surfaces needs a separate set of nodes.
-This approach is known as \textquotedbl{}split nodes\textquotedbl{}.
-Currently faults can only be run with \texttt{xdecompose\_mesh} and
-\texttt{CUBIT}. \texttt{xmeshfem3D} is not yet ready to handle faults.\\
-To facilitate the mesh generation with split nodes in CUBIT, we need
-to separate the two fault surfaces by a small distance, effectively
-creating a tiny opening of the fault (Figure \ref{fig:examples.splitnodes},
-\ref{fig:examples.splitnodes-surfacetrace}). Note that only the interior
-of the fault must be opened, its edges must remain closed (except
-the edge on the free surface). The fault is automatically closed later
-by SPECFEM3D.\\
-
-Here is an example CUBIT script to generate a mesh with split nodes
-for a buried vertical strike-slip fault:
-\begin{lyxcode}
-~reset~~\\
-~brick~x~10~y~10~z~10~~\\
-~webcut~volume~all~with~plane~xplane~~\\
-~webcut~volume~all~with~plane~yplane~~\\
-~webcut~volume~all~with~plane~xplane~offset~3~~\\
-~webcut~volume~all~with~plane~zplane~offset~3~~\\
-~webcut~volume~all~with~plane~zplane~offset~-3~~\\
-~imprint~all~~\\
-~merge~all~~\\
-~unmerge~surf~160~~\\
-~mesh~vol~all~~\\
-~set~node~constraint~off~~\\
-~node~in~surf~168~move~X~0~Y~0.01~Z~0~~\\
-~node~in~surf~160~move~X~0~Y~-0.01~Z~0~\\
-
-\end{lyxcode}
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[scale=0.3]{figures/faultmesh} \\
- \includegraphics[scale=0.3]{figures/surf168} \includegraphics[scale=0.3]{figures/splitnodes}
-\par\end{centering}
-
-\caption{Screenshots of the CUBIT example for the embedded fault with split
-nodes: The entire mesh is decomposed into several volumes bounding
-the fault (top), the fault surface 168 is shown as orange squares
-(middle) and split nodes of the fault shown as orange dots(bottom).
-Note that only interior nodes of the fault are split while those on
-the edges of the fault surface are touching each other.}
-
-\label{fig:examples.splitnodes}
-\end{figure}
-
-\begin{lyxcode}
-
-\end{lyxcode}
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[scale=0.55]{figures/splitnodes_surfacetrace} \\
-
-\par\end{centering}
-
-\caption{Screenshots of a CUBIT example showing split nodes for a fault reaching
-the surface. Surface trace of the fault is shown in orange. Note that
-the edges of the fault are not split while the interior nodes are
-offset by a small distance on either side of the fault}
-
-\label{fig:examples.splitnodes-surfacetrace}
-\end{figure}
-
-The CUBIT scripts ({*}.jou and {*}.py) in the directory \texttt{EXAMPLES}
-generate more complicated meshes. The {*}.py files are Python scripts
-that execute CUBIT commands and use the CUBIT-python interface for
-SPECFEM3D (see next section). The Python language allows to define
-and manipulate variables to parameterize the mesh. Alternatively,
-the Python script can call a CUBIT journal file ({*}.jou), which looks
-like the example above. Variables can be defined and manipulated there
-using the $\mathtt{APREPRO}$ language built in CUBIT.\\
-
-Note that you should avoid gaps in the list of indices of mesh objects
-with the following CUBIT command:
-\begin{lyxcode}
-$\mathtt{compress\: ids\: hex\: face\: edge\: node}$
-\end{lyxcode}
-(otherwise you will get a segmentation fault during domain decomposition)
-
-\section{CUBIT-Python Scripts for Faults}
-
-The mesh generated in CUBIT needs to be processed and exported in
-a format compatible with SPECFEM3D. This is achieved in the Python
-scripts by calling the Python-CUBIT interface functions defined in
-the CUBIT directory:
-\begin{enumerate}
-\item Function $\mathtt{define\_bc}$ (or $\mathtt{boundary\_definition.py}$)
-must be called to set up the absorbing boundaries database
-\item Function $\mathtt{fault\_input}$ must be called once for each fault
-to set up the fault database
-\item Function $\mathtt{cubit2specfem3d.export2SPECFEM3D}$ must be called
-at the very end of the script to export the mesh in a SPECFEM3D format.
-\end{enumerate}
-The functions in \#1 and \#3 are for the bulk and are documented in
-Section \ref{subsec:Exporting-the-Mesh}. We focus here on \#2:
-\begin{description}
-\item [{Function:}] $\mathtt{fault\_input}$
-\item [{Purpose:}] export a fault mesh database from CUBIT to a SPECFEM3D-compliant
-file
-\item [{Syntax:}] fault\_input(id\_fault, ids\_surf\_1, ids\_surf\_2)
-\item [{Inputs:}]~\end{description}
-\begin{itemize}
-\item id\_fault integer index assigned to the fault. (The user must number
-all the faults, starting at 1, with unit increments)
-\item ids\_surf\_1 list of CUBIT ids of all surfaces that form side 1 of
-the fault.
-\item ids\_surf\_2 list of CUBIT ids of all surfaces that form side 2 of
-the fault. (The user must decide which side of the fault is side 1.
-This choice affects the sign conventions of fault quantities as explained
-in Section \ref{sec:Sign-Convention-for}).\end{itemize}
-\begin{description}
-\item [{Outputs:}] file $\mathtt{fault\_file\_X.dat}$, where X is the
-fault id (id\_fault).
-\item [{Example:}] For the example in Section \ref{sec:Mesh-Generation-with}:\end{description}
-\begin{lyxcode}
-$\hphantom{}$A1~=~{[}168{]}
-
-$\hphantom{}$A2~=~{[}160{]}
-
-$\hphantom{}$fault\_input(1,A1,A2)
-\end{lyxcode}
-
-\section{Examples}
-
-The package includes examples, the SCEC benchmark problems:
-\begin{itemize}
-\item TPV5, a planar vertical strike-slip fault
-\item TPV14 and TPV15, vertical strike-slip fault system with a fault branch
-\item TPV102 and TPV103, rate-and-state benchmarks
-\item TPV22 and TPV23, step-over benchmarks
-\item TPV16, heterogenous initial stress
-\end{itemize}
-and
-\begin{itemize}
-\item Splay fault models from Wendt et al. (2009)\\
-
-\end{itemize}
-Read the documents in the directory \texttt{EXAMPLES/*/description}.
-They contain a description of the example and additional instructions
-to run it. Visualize the results with the Matlab scripts in the directory
-$\mathtt{EXAMPLES/*/post}$
-
-\section{\label{sec:Sign-Convention-for}Sign Convention for Fault Quantities}
-
-During mesh generation, the fault is defined by two surfaces in contact.
-Let's denote as \textquotedbl{}\emph{side 1}\textquotedbl{} the SECOND
-surface declared by the user in the call to the python function \textquotedbl{}\emph{fault\_input}\textquotedbl{},
-and the FIRST surface as \textquotedbl{}\emph{side 2}\textquotedbl{}.
-The local coordinate system on the fault is defined as the right-handed
-coordinate system defined by (strike, dip, normal), where \textquotedbl{}\emph{normal}\textquotedbl{}
-is the normal vector outgoing from side 1, \textquotedbl{}\emph{dip}\textquotedbl{}
-is parallel to the along-dip direction pointing downwards, and \textquotedbl{}\emph{strike}\textquotedbl{}
-is the horizontal along-strike vector such that the system is right-handed.\\
-
-Slip is defined as displacement on side 2 minus displacement on side
-1. In the local coordinate system on the fault, positive along-strike
-slip is right-lateral and positive along-dip slip is thrust if side
-1 is on the hanging wall (normal faulting if side 1 is on the foot
-wall).\\
-
-Traction is defined as the stress induced on side 1 by side 2, which
-is the stress tensor times the normal vector outgoing from side 1.
-In the local coordinate system on the fault, the normal traction is
-negative in compression, positive along-strike traction generates
-right-lateral slip and positive along-dip traction generates thrust
-slip if side 1 is on the hanging wall (normal faulting if side 1 is
-on the foot wall).
-
-\section{Input Files}
-
-Three additional input files are required in the \texttt{DATA/} directory
-for dynamic and kinematic faults. They are $\mathtt{Par\_file\_faults}$,
-$\mathtt{FAULT\_STATIONS}$ and $\mathtt{input\_file.txt}$ (optional).
-If the former does not exist, the code assumes that there are no faults.\\
-
-\begin{lyxlist}{00.00.0000}
-\item [{\textbf{DATA/Par\_file\_faults}}] contains parameters of the fault.
-The first part of this file has a strict format:
-
-\begin{description}
-\item [{Line}] 1: Number of faults (NF)
-\item [{Lines}] 2 to NF+1: Kelvin Voigt damping (in seconds) for each fault.
-(See below how to set this parameter)
-\item [{Line}] NF+2: Type of simulation (1=dynamic , 2 = kinematic)
-\item [{Line}] NF+3: Number of time steps between updates of the time series
-outputs at selected fault points (see DATA/FAULT\_STATIONS), usually
-a large number (100s or 1000s). Note that the sampling rate of the
-time series is usually much higher.
-\item [{Line}] NF+4: Number of time steps between fault snapshot outputs
-(quantities at every fault point exported at regular times), usually
-a large number (100s or 1000s).
-\item [{Line}] NF+5: Slip velocity threshold below which frictional healing
-is set (friction coefficient is reset to its static value). If this
-value is negative healing is disabled.
-\item [{Line}] NF+6: Slip velocity threshold to define the rupture front.
-Only used for outputs.
-\end{description}
-
-The rest of this file is made of namelist input blocks (see \textquotedbl{}namelist\textquotedbl{}
-in a Fortran 9x manual). The input for each fault has the following
-sequence (arguments in {[}brackets{]} are optional):
-
-\&\textbf{BEGIN\_FAULT} /
-
-\&\textbf{INIT\_STRESS} S1, S2, S3 {[},n1, n2, n3{]} /
-
-followed by (n1+n2+n3) \&\textbf{DIST2D} blocks
-
-\&\textbf{SWF} mus, mud, dc {[}, nmus, nmud, ndc{]} /
-
-followed by (nmus+nmud+ndc) \&\textbf{DIST2D} blocks\\
-
-\begin{lyxlist}{00.00.0000}
-\item [{\&\textbf{INIT\_STRESS}}] input block sets the initial fault stresses
-relative to the foot-wall side of the fault. Initial stresses are
-composed of a constant background value possibly overwritten in prescribed
-regions by heterogeneous distributions (see \&\textbf{DIST2D} blocks
-below):
-
-\begin{description}
-\item [{S1}] = constant background value of along-strike shear stress (positive
-in the usual strike direction)
-\item [{S2}] = constant background value of along-dip shear (positive is
-down-dip, normal faulting)
-\item [{S3}] = constant background value of normal stress (negative in
-compression)
-\item [{n1}] = number of heterogeneous items for along-strike shear stress
-{[}default is 0{]}
-\item [{n2}] = number of heterogeneous items for along-dip shear stress
-{[}default is 0{]}
-\item [{n3}] = number of heterogeneous items for normal stress {[}default
-is 0{]}
-\end{description}
-\item [{\&\textbf{SWF}}] input block sets the slip-weakening friction parameters
-of the fault:
-
-\begin{description}
-\item [{mus}] = constant background value of static friction coefficient
-\item [{mud}] = constant background value of dynamic friction coefficient
-\item [{dc}] = constant background value of critical slip-weakening distance
-\item [{nmus}] = number of heterogeneous items for static friction coefficient
-{[}default is 0{]}
-\item [{nmud}] = number of heterogeneous items for dynamic friction coefficient
-{[}default is 0{]}
-\item [{ndc}] = number of heterogeneous items for critical slip-weakening
-distance {[}default is 0{]}\\
-
-\end{description}
-\item [{\&\textbf{DIST2D}}] input blocks modify (overwrite) the value of
-a fault parameter by a heterogeneous spatial distribution:
-
-\&\textbf{DIST2D} shape='$\mathtt{square}$', val, xc, yc, zc, l /
-
-$\;$sets a constant value (val) within a cube with center (xc,yc,zc)
-and edge size l.
-
-\&\textbf{DIST2D} shape='$\mathtt{rectangle}$', val, xc, yc, zc,
-lx, ly, lz /
-
-$\;$sets a constant value (val) within a rectangular prism with center
-(xc,yc,zc) and edge sizes (lx,ly,lz).
-
-\&\textbf{DIST2D} shape='$\mathtt{rectangle-taper}$', val, valh,
-xc, yc, zc, lx, ly, lz /
-
-$\;$sets a vertical linear gradient within a rectangular prism with
-center (xc,yc,zc) and edge sizes (lx,ly,lz). Values vary linearly
-as a function of vertical position z between value val at z = zc-lz/2
-and value valh at z = zc+lz/2 .
-
-\&\textbf{DIST2D} shape='$\mathtt{circular}$', val, xc, yc, zc, r
-/
-
-$\;$sets a constant value (val) within a sphere with center (xc,yc,zc)
-and radius r. \\
-
-\end{lyxlist}
-\end{lyxlist}
-
-\begin{lyxlist}{00.00.0000}
-\item [{\textbf{DATA/FAULT\_STATIONS}}] Stations in the fault plane.
-
-\textbf{Line 1} : number of stations.
-
-\textbf{Line 2 to end}: 5 columns: X, Y, Z (-depth), station name,
-fault-id
-
-The fault-id identifies the fault that contains the station. It is
-the index of appearance in the faults list after line 2 of Par\_file\_faults
-
-\end{lyxlist}
-
-\begin{lyxlist}{00.00.0000}
-\item [{\textbf{DATA/input\_file.txt}}] Heterogeneous stresses and friction
-parameters are documented in page 10 of
-
-\begin{lyxlist}{00.00.0000}
-\item [{$\mathtt{EXAMPLES/tpv16/description/TPV16\_17\_Description\_v03.pdf}$}]~
-\end{lyxlist}
-
-To activate this feature, in fault\_solver\_dynamic.f90 set \textbf{TPV16}
-= .true. then re-compile the code.
-
-\end{lyxlist}
-
-\section{\label{sec:Setting-the-Kelvin-Voigt}Setting the Kelvin-Voigt Damping
-Parameter }
-
-The purpose of the Kelvin-Voigt viscosity in the dynamic fault solver
-is to damp spurious oscillations generated by the fault slip at frequencies
-that are too high to be resolved by the mesh. The viscosity $\mathtt{eta}$
-(in seconds) depends on the size of the elements on the fault. Here
-is how to set it:
-\begin{enumerate}
-\item Determine the average linear size of the elements on the fault plane,
-$\mathtt{h\_fault}$. Usually this value is prescribed by the user
-during mesh generation. Otherwise it can be found by inspection of
-the mesh inside the CUBIT GUI.
-\item Use the Matlab function $\mathtt{utils/critical\_timestep.m}$ to
-compute\\
- $\mathtt{dtc\_fault}=\mathtt{critical\_timestep\left(c_{p},h\_fault,ngll\right)}$.
-\\
-This is the critical time step in an elastic medium for a hypothetical
-element of cubic shape with size equal to $\mathtt{h\_fault}$.
-\item Set $\mathtt{eta}$ in $\mathtt{Par\_file\_faults}$ to (0.1 to 0.3)$\ensuremath{\times}\mathtt{dtc\_fault}$.
-A larger $\mathtt{eta}$ damps high-frequencies more aggressively
-but it might also affect lower frequencies and rupture speed.
-\end{enumerate}
-Viscosity reduces numerical stability: the critical timestep in a
-simulation with Kelvin-Voigt damping needs to be smaller than that
-in a purely elastic simulation. Here is how to set the time step accordingly:
-\begin{enumerate}
-\item Run a test simulation without viscosity ($\mathtt{eta}$=0 and only
-a few time steps)
-\item Look for the \textquotedbl{}maximum suggested time step\textquotedbl{}
-in \texttt{OUTPUT\_FILES/output\_mesher.txt}. This is the critical
-timestep of a purely elastic simulation, $\mathtt{dtc\_bulk}$.
-\item Reset the timestep of the simulation with a Kelvin-Voigt material
-to a value smaller than\\
- $\mathtt{dtc\_kv}=\mathtt{eta}\left(\sqrt{1+\mathtt{dtc\_bulk^{2}}/\mathtt{eta^{2}}}-1\right)$
-\end{enumerate}
-Note that in general $\mathtt{dtc\_bulk}$ is smaller than $\mathtt{dtc\_fault}$,
-because elements off the fault might be smaller or more distorted
-than element faces on the fault.
-
-\section{Output Files }
-
-Several output files are saved in \texttt{OUTPUT\_FILES/}:\\
-
-\begin{enumerate}
-\item Seismograms for each station on the fault plane given in \texttt{DATA/FAUL\_STATIONS}.
-One output file is generated for each station, named after the station.
-The files are ascii and start with a header (22 lines long) followed
-by a data block with the following format, one line per time sample:\\
-
-Column 1 = Time (s)
-
-Column 2 = horizontal right-lateral slip (m)
-
-Column 3 = horizontal right-lateral slip rate (m/s)
-
-Column 4 = horizontal right-lateral shear stress (MPa)
-
-Column 5 = vertical up-dip slip (m)
-
-Column 6 = vertical up-dip slip rate (m/s)
-
-Column 7 = vertical up-dip shear stress (MPa)
-
-Column 8 = normal stress (MPa)\\
-
-The stresses are relative to the footwall side of the fault (this
-convention controls their sign, but not their amplitude). Slip is
-defined as displacement of the hanging wall relative to the footwall.
-
-\item Seismograms at stations in the bulk (out of the fault plane) given
-in \texttt{DATA/STATIONS}.
-\item Rupture time files are named $\mathtt{Rupture\_time\_FAULT}$-$\mathtt{id}$.
-One file is generated for each fault. The files are ascii and start
-with a header (12 lines long) followed by a data block with the following
-format, one line per fault node:
-
-Column 1 = horizontal coordinate, distance along strike (m)
-
-Column 2 = vertical coordinate, distance down-dip (m)
-
-Column 3 = rupture time (s)
-
-\item Fault quantities (slip, slip rate, stresses, etc) at regular times
-are stored in binary data files called $\mathtt{Snapshot\#it\#.bin}$,
-where \#it\# is the timestep number. These can be read in Matlab with
-the function $\mathtt{utils/FSEM3D\_snapshot.m}$
-\end{enumerate}
-
-\section{Post-processing and Visualization }
-
-Some Matlab functions for post-processing and visualization are included
-in directory $\mathtt{utils}$. The functions are internally documented
-(see their matlab help).\\
-
-\texttt{FSEM3D\_snapshot} reads a fault data snapshot\\
-
-The directories \texttt{EXAMPLES/*/post} contain additional Matlab
-scripts to generate figures specific to each example.
-
-\chapter{\label{cha:Adjoint-Simulations}Adjoint Simulations}
-
-Adjoint simulations are generally performed for two distinct applications.
-First, they can be used in point source moment-tensor inversions, or source imaging for
-earthquakes with large ruptures such as the Lander's earthquake \citep{WaHe94}.
-Second, they can be used to generate finite-frequency sensitivity
-kernels that are a critical part of tomographic inversions based upon
-3D reference models \citep{TrTaLi05,LiTr06,TrKoLi08,LiTr08}. In either
-case, source parameter or velocity structure updates are sought to
-minimize a specific misfit function (e.g., waveform or traveltime
-differences), and the adjoint simulation provides a means of computing
-the gradient of the misfit function and further reducing it in successive
-iterations. Applications and procedures pertaining to source studies
-and finite-frequency kernels are discussed in Sections~\ref{sec:Adjoint-simulation-sources}
-and \ref{sec:Adjoint-simulation-finite}, respectively. The two related
-parameters in the \texttt{Par\_file} are \texttt{SIMULATION\_TYPE}
-(1, 2 or 3) and the \texttt{SAVE\_FORWARD} (boolean).
-
-\section{\label{sec:Adjoint-simulation-sources}Adjoint Simulations for Sources
-Only (not for the Model)}
-
-When a specific misfit function between data and synthetics is minimized to invert
-for earthquake source parameters, the gradient of the misfit function
-with respect to these source parameters can be computed by placing
-time-reversed seismograms at the receivers as virtual sources
-in an adjoint simulation. Then the value of the gradient is obtained
-from the adjoint seismograms recorded at the original earthquake location.
-\begin{enumerate}
-\item \textbf{Prepare the adjoint sources} \label{enu:Prepare-the-adjoint}
-
-\begin{enumerate}
-\item First, run a regular forward simlation (\texttt{SIMULATION\_TYPE =
-1} and \texttt{SAVE\_FORWARD = .false.}). You can automatically set
-these two variables using the \texttt{\small utils/change\_simulation\_type.pl}{\small{}
-script:}{\small \par}
-
-\begin{lyxcode}
-utils/change\_simulation\_type.pl~-f~
-\end{lyxcode}
-
-and then collect the recorded seismograms at all the stations given
-in \texttt{DATA/STATIONS}.
-
-\item Then select the stations for which you want to compute the time-reversed
-adjoint sources and run the adjoint simulation, and compile them into
-the \texttt{DATA/STATIONS\_ADJOINT} file, which has the same format
-as the regular \texttt{DATA/STATIONS} file.
-
-\begin{itemize}
-\item Depending on what type of misfit function is used for the source inversion,
-adjoint sources need to be computed from the original recorded seismograms
-for the selected stations and saved in a sub-directory called \texttt{SEM/}
-in the root directory of the code, with the format \texttt{NT.STA.BX?.adj}, where \texttt{NT},
-\texttt{STA} are the network code and station name given in the \texttt{DATA/STATIONS\_ADJOINT}
-file, and \texttt{BX?} represents the component name of a particular
-adjoint seismogram. Please note that the band code can change depending
-on your sampling rate (see Appendix~\ref{cha:channel-codes} for
-further details).
-\item The adjoint seismograms are in the same format as the original seismogram
-(\texttt{NT.STA.BX?.sem?}), with the same start time, time interval
-and record length.
-\end{itemize}
-\item Notice that even if you choose to time reverse only one component
-from one specific station, you still need to supply all three components
-because the code is expecting them (you can set the other two components
-to be zero).
-\item Also note that since time-reversal is done in the code itself, no
-explicit time-reversing is needed for the preparation of the adjoint
-sources, i.e., the adjoint sources are in the same forward time sense
-as the original recorded seismograms.
-\end{enumerate}
-\item \textbf{Set the related parameters and run the adjoint simulation}\\
- In the \texttt{DATA/Par\_file}, set the two related parameters to
-be \texttt{SIMULATION\_TYPE = 2} and \texttt{SAVE\_FORWARD = .false.}.
-More conveniently, use the scripts \texttt{utils/change\_simulation\_type.pl}
-to modify the \texttt{Par\_file} automatically (\texttt{change\_simulation\_type.pl
--a}). Then run the solver to launch the adjoint simulation.
-\item \textbf{Collect the seismograms at the original source location}
-
-After the adjoint simulation has completed successfully, collect the
-seismograms from \texttt{LOCAL\_PATH}.
-\begin{itemize}
-\item These adjoint seismograms are recorded at the locations of the original
-earthquake sources given by the \texttt{DATA/CMTSOLUTION} file, and
-have names of the form \texttt{NT.S?????.S??.sem} for the six-component
-strain tensor (\texttt{SNN,SEE,SZZ,SNE,SNZ,SEZ}) at these locations,
-and ~\\
- \texttt{NT.S?????.BX?.sem} for the three-component displacements
-(\texttt{BXN,BXE,BXZ}) recorded at these locations.
-\item \texttt{S?????} denotes the source number; for example, if the original
-\texttt{CMTSOLUTION} provides only a point source, then the seismograms
-collected will start with \texttt{S00001}.
-\item These adjoint seismograms provide critical information for the computation
-of the gradient of the misfit function.
-\end{itemize}
-\end{enumerate}
-
-\section{\label{sec:Adjoint-simulation-finite}Adjoint Simulations for Finite-Frequency
-Kernels (Kernel Simulation)}
-
-Finite-frequency sensitivity kernels are computed in two successive
-simulations (please refer to \citet{LiTr06} and \citet{TrKoLi08}
-for details).
-\begin{enumerate}
-\item \textbf{Run a forward simulation with the state variables saved at
-the end of the simulation}
-
-Prepare the \texttt{\small CMTSOLUTION}{\small{} and }\texttt{\small STATIONS}{\small{}
-files, set the parameters }\texttt{\small SIMULATION\_TYPE}{\small {}
-}\texttt{\small =}{\small {} }\texttt{\small 1}{\small{} and }\texttt{\small SAVE\_FORWARD
-=}{\small {} }\texttt{\small .true.}{\small{} in the }\texttt{\small Par\_file}{\small{}
-(}\texttt{\small change\_simulation\_type -F}{\small ), and run the
-solver.}{\small \par}
-\begin{itemize}
-\item Notice that attenuation is not implemented yet for the computation
-of finite-frequency kernels; therefore set \texttt{ATTENUATION = .false.}
-in the \texttt{Par\_file}.
-\item We also suggest you modify the half duration of the \texttt{CMTSOLUTION}
-to be similar to the accuracy of the simulation (see Equation \ref{eq:shortest_period})
-to avoid too much high-frequency noise in the forward wavefield, although
-theoretically the high-frequency noise should be eliminated when convolved
-with an adjoint wavefield with the proper frequency content.
-\item This forward simulation differs from the regular simulations (\texttt{\small SIMULATION\_TYPE}{\small {}
-}\texttt{\small =}{\small {} }\texttt{\small 1}{\small{} and }\texttt{\small SAVE\_FORWARD}{\small {}
-}\texttt{\small =}{\small {} }\texttt{\small .false.}{\small ) described
-in the previous chapters in that the state variables for the last
-time step of the simulation, including wavefields of the displacement,
-velocity, acceleration, etc., are saved to the }\texttt{\small LOCAL\_PATH}{\small{}
-to be used for the subsequent simulation. }{\small \par}
-\item For regional simulations, the files recording the absorbing boundary
-contribution are also written to the \texttt{LOCAL\_PATH} when \texttt{SAVE\_FORWARD
-= .true.}.
-\end{itemize}
-\item \textbf{Prepare the adjoint sources}
-
-The adjoint sources need to be prepared the same way as described
-in the Section \ref{enu:Prepare-the-adjoint}.
-\begin{itemize}
-\item In the case of travel-time finite-frequency kernel for one source-receiver
-pair, i.e., point source from the \texttt{CMTSOLUTION}, and one station
-in the \texttt{STATIONS\_ADJOINT} list, we supply a sample program
-in \texttt{utils/adjoint\_sources/traveltime/xcreate\_adjsrc\_traveltime}
-to cut a certain portion of the original displacement seismograms
-and convert it into the proper adjoint source to compute the finite-frequency
-kernel.
-
-\begin{lyxcode}
-xcreate\_adjsrc\_traveltime~t1~t2~ifile{[}0-5{]}~E/N/Z-ascii-files~{[}baz{]}
-\end{lyxcode}
-
-where \texttt{t1} and \texttt{t2} are the start and end time of the
-portion you are interested in, \texttt{ifile} denotes the component
-of the seismograms to be used (0 for all three components, 1 for East,
-2 for North, and 3 for vertical, 4 for transverse, and 5 for radial
-component), \texttt{E/N/Z-ascii-files} indicate the three-component
-displacement seismograms in the right order, and \texttt{baz} is the
-back-azimuth of the station. Note that \texttt{baz} is only supplied
-when \texttt{ifile} = 4 or 5.
-
-\item Similarly, a sample program to compute adjoint sources for amplitude
-finite-frequency kernels may be found in \texttt{utils/adjoint\_sources/amplitude}
-and used in the same way as described for traveltime measurements
-
-\begin{lyxcode}
-xcreate\_adjsrc\_amplitude~t1~t2~ifile{[}0-5{]}~E/N/Z-ascii-files~{[}baz{]}.
-\end{lyxcode}
-\end{itemize}
-\item \textbf{Run the kernel simulation}
-
-With the successful forward simulation and the adjoint source ready
-in the \texttt{SEM/} directory, set \texttt{SIMULATION\_TYPE
-= 3} and \texttt{SAVE\_FORWARD = .false.} in the \texttt{Par\_file}
-(you can use \texttt{change\_simulation\_type.pl -b}), and rerun the
-solver.
-\begin{itemize}
-\item The adjoint simulation is launched together with the back reconstruction
-of the original forward wavefield from the state variables saved from
-the previous forward simulation, and the finite-frequency kernels
-are computed by the interaction of the reconstructed forward wavefield
-and the adjoint wavefield.
-\item The back-reconstructed seismograms at the original station locations
-are saved to the \texttt{LOCAL\_PATH} at the end of the kernel simulations,
-and can be collected to the local disk.
-\item These back-constructed seismograms can be compared with the time-reversed
-original seismograms to assess the accuracy of the backward reconstruction,
-and they should match very well.
-\item The arrays for density, P-wave speed and S-wave speed kernels are
-also saved in the \texttt{LOCAL\_PATH} with the names \texttt{proc??????\_rho(alpha,beta)\_kernel.bin},
-where \texttt{proc??????} represents the processor number, \texttt{rho(alpha,beta)}
-are the different types of kernels.
-\end{itemize}
-\item \textbf{Run the anisotropic kernel simulation}
-
-Instead of the kernels for the isotropic wave speeds, you can also
-compute the kernels for the 21 independent components $C_{IJ},\, I,J=1,...,6$
-(using Voigt's notation) of the elastic tensor in the cartesian coordinate
-system. This is done by setting \texttt{ANISOTROPIC\_KL} \texttt{=}
-\texttt{.true.} in \texttt{constants.h} before compiling the package.
-The definition of the parameters $C_{IJ}$ in terms of the corresponding
-components $c_{ijkl},ijkl,i,j,k,l=1,2,3$ of the elastic tensor in
-cartesian coordinates follows \citet{ChTr07}. The 21 anisotropic
-kernels are saved in the \texttt{LOCAL\_PATH} in one file with the
-name of \texttt{proc??????\_cijkl\_kernel.bin} (with \texttt{proc??????}
-the processor number). The output kernels correspond to absolute perturbations
-$\delta C_{IJ}$ of the elastic parameters and their unit is in $s/GPa/km^{3}$.
-For consistency, the output density kernels with this option turned
-on are for a perturbation $\delta\rho$ (and not $\frac{\delta\rho}{\rho}$)
-and their unit is in s / (kg/m$^{3}$) / km$^{3}$. These `primary'
-anisotropic kernels can then be combined to obtain the kernels for
-different parameterizations of anisotropy. This can be done, for example,
-when combining the kernel files from slices into one mesh file (see
-Section~\ref{sec:Finite-Frequency-Kernels}).
-
-If \texttt{ANISOTROPIC\_KL} \texttt{=} \texttt{.true.} by additionally
-setting \texttt{ANISOTROPIC\_KL} \texttt{=} \texttt{.true.} in \texttt{constants.h}
-the package will save anisotropic kernels paramterized as velocities
-related to transverse isotropy based on the the Chen and Tromp parameters
-\citet{ChTr07}. The kernels are saved as relative perturbations for
-horizontal and vertical P and S velocities, $\alpha_{v},\alpha_{h},\beta_{v},\beta_{h}$.
-Explicit relations can be found in appendix B. of \citet{SiLiTrTr07b}
-
-The anisotropic kernels are only currently available for CPU mode.
-
-\end{enumerate}
-In general, the three steps need to be run sequentially to assure
-proper access to the necessary files. If the simulations are run through
-some cluster scheduling system (e.g., LSF), and the forward simulation
-and the subsequent kernel simulations cannot be assigned to the same
-set of computer nodes, the kernel simulation will not be able to access
-the database files saved by the forward simulation. Solutions for
-this dilemma are provided in Chapter~\ref{cha:Scheduler}. Visualization
-of the finite-frequency kernels is discussed in Section~\ref{sec:Finite-Frequency-Kernels}.
-
-%------------------------------------------------------------------------------------------------%
-
-\chapter{\label{cha:tomo}Doing tomography, i.e., updating the model based on the sensitivity kernels obtained}
-
-%------------------------------------------------------------------------------------------------%
-
-UNDER CONSTRUCTION (March 2015). Will be added around April / May 2015.
-
-\section{Tomographic inversion / imaging using the sensitivity kernels obtained}
-
-OLD VERSION OF THE SECTION, WILL SOON BE IMPROVED AND ENRICHED.\\
-
-One of the fundamental reasons for computing sensitivity kernels (Section~\ref{sec:Adjoint-simulation-finite})
-is to use them within a tomographic inversion. In other words, use
-recorded seismograms, make measurements with synthetic seismograms,
-and use the misfit between the two to iteratively improve the model
-described by (at least) $V_{{\rm p}}$, $V_{{\rm s}}$, and $\rho$
-as a function of space.
-
-Whatever misfit function you use for the tomographic inversion (you can find several
-examples in \citet{TrKoLi08} and \citet{TrTaLi05}), you will weight
-the sensitivity kernels with measurements. Furthermore, you will use
-as many measurements (stations, components, time windows) as possible
-per event; hence, we call these composite kernels ``event kernels,''
-which are volumetric fields representing the gradient of the misfit
-function with respect to one of the variables (\eg $V_{{\rm s}}$).
-The basic features of an adjoint-based tomographic inversion were
-illustrated in \citet{TaLiTr07} and \citet{TrKoLi08} using a conjugate-gradient
-algorithm; there are dozens of versions of gradient-based inversion
-algorithms that could alternatively be used. The tomographic inversion
-of \citet{TaLiMaTr09,TaLiMaTr2010} used SPECFEM3D Cartesian as well
-as several additional components which are also stored on the CIG
-svn server, described next.
-
-The directory containing external utilities for tomographic inversion using
-SPECFEM3D Cartesian (or other packages that evaluate misfit functions
-and gradients) is in directory \texttt{utils/ADJOINT\_TOMOGRAPHY\_TOOLS/}:
-\begin{verbatim}
-flexwin/     -- FLEXWIN algorithm for automated picking of time windows
-measure_adj/ -- reads FLEXWIN output file and makes measurements,
-                with the option for computing adjoint sources
-iterate_adj/ -- various tools for iterative inversion
-                (requires pre-computed "event kernels")
-\end{verbatim}
-This directory also contains
-a brief \verb+README+ file indicating the role of the three subdirectories,
-\verb+flexwin+ \citep{Maggi2009}, \verb+measure_adj+, and \verb+iterate_adj+.
-The components for making the model update are there; however, there
-are no explicit rules for computing the model update, just as with
-any optimization problem. There are options for computing a conjugate
-gradient step, as well as a source subspace projection step.
-
-The best single file to read is probably: \verb+ADJOINT_TOMO/iterate_adj/cluster/README+.
-
-%------------------------------------------------------------------------------------------------%
-
-\chapter{Noise Cross-correlation Simulations}
-
-%------------------------------------------------------------------------------------------------%
-
-Besides earthquake simulations, SPECFEM3D Cartesian includes functionality
-for seismic noise tomography as well. In order to proceed successfully
-in this chapter, it is critical that you have already familiarized
-yourself with procedures for meshing (Chapter \ref{cha:Mesh-Generation}),
-creating distributed databases (Chapter \ref{cha:Creating-Distributed-Databases}),
-running earthquake simulations (Chapters \ref{cha:Running-the-Solver})
-and adjoint simulations (Chapter \ref{cha:Adjoint-Simulations}).
-Also, make sure you read the article `Noise cross-correlation sensitivity
-kernels' \citep{trompetal2010}, in order to understand noise simulations
-from a theoretical perspective.
-
-\section{Input Parameter Files}
-
-As usual, the three main input files are crucial: \texttt{Par\_file},
-\texttt{CMTSOLUTION} and \texttt{STATIONS}. Unless otherwise specified,
-those input files should be located in directory \texttt{DATA/}.\\
-
-\texttt{CMTSOLUTION} is required for all simulations. At a first glance,
-it may seem unexpected to have it here, since the noise simulations
-should have nothing to do with the earthquake -- \texttt{CMTSOLUTION}.
-However, for noise simulations, it is critical to have no earthquakes.
-In other words, the moment tensor specified in \texttt{CMTSOLUTION}
-must be set to zero manually!\\
-
-\texttt{STATIONS} remains the same as in previous earthquake simulations,
-except that the order of receivers listed in \texttt{STATIONS} is
-now important. The order will be used to determine the `master' receiver,
-i.e., the one that simultaneously cross correlates with the others.\\
-
-\texttt{Par\_file} also requires careful attention. A parameter called
-\texttt{NOISE\_TOMOGRAPHY} has been added which specifies the type
-of simulation to be run. \texttt{NOISE\_TOMOGRAPHY} is an integer
-with possible values 0, 1, 2 and 3. For example, when \texttt{NOISE\_TOMOGRAPHY}
-equals 0, a regular earthquake simulation will be run. When it is
-1/2/3, you are about to run step 1/2/3 of the noise simulations respectively.
-Should you be confused by the three steps, refer to \citet{trompetal2010}
-for details.\\
-
-Another change to \texttt{Par\_file} involves the parameter \texttt{NSTEP}.
-While for regular earthquake simulations this parameter specifies
-the length of synthetic seismograms generated, for noise simulations
-it specifies the length of the seismograms used to compute cross correlations.
-The actual cross correlations are thus twice this length, i.e., $2~\mathrm{NSTEP}-1$.
-The code automatically makes the modification accordingly, if \texttt{NOISE\_TOMOGRAPHY}
-is not zero.\\
-
-There are other parameters in \texttt{Par\_file} which should be given
-specific values. For instance, since the first two steps for calculating
-noise sensitivity kernels correspond to forward simulations, \texttt{SIMULATION\_TYPE}
-must be 1 when \texttt{NOISE\_TOMOGRAPHY} equals 1 or 2. Also, we
-have to reconstruct the ensemble forward wavefields in adjoint simulations,
-therefore we need to set \texttt{SAVE\_FORWARD} to \texttt{.true.}
-for the second step, i.e., when \texttt{NOISE\_TOMOGRAPHY} equals
-2. The third step is for kernel constructions. Hence \texttt{SIMULATION\_TYPE}
-should be 3, whereas \texttt{SAVE\_FORWARD} must be \texttt{.false.}.\\
-
-Finally, for most system architectures, please make sure that \texttt{LOCAL\_PATH}
-in \texttt{Par\_file} is in fact local, not globally shared. Because
-we have to save the wavefields at the earth's surface at every time
-step, it is quite problematic to have a globally shared \texttt{LOCAL\_PATH},
-in terms of both disk storage and I/O speed.
-
-\section{Noise Simulations: Step by Step}
-
-Proper parameters in those parameter files are not enough for noise
-simulations to run. We have more parameters to specify: for example,
-the ensemble-averaged noise spectrum, the noise distribution etc.
-However, since there are a few `new' files, it is better to introduce
-them sequentially. In this section, standard procedures for noise
-simulations are described.
-
-\subsection{Pre-simulation}
-\begin{itemize}
-\item As usual, we first configure the software package using:
-
-\texttt{./configure FC=ifort MPIFC=mpif90}
-
-Use the following if SCOTCH is needed:
-
-\texttt{./configure FC=ifort MPIFC=mpif90 -{}-with-scotch-dir=/opt/scotch}\\
-
-\item Next, we need to compile the source code using:
-
-\texttt{make xgenerate\_databases}
-
-\texttt{make xspecfem3D} \\
-
-\item Before we can run noise simulations, we have to specify the noise
-statistics, e.g., the ensemble-averaged noise spectrum. Matlab scripts
-are provided to help you to generate the necessary file.
-
-\texttt{EXAMPLES/noise\_tomography/NOISE\_TOMOGRAPHY.m} (main program)\\
- \texttt{EXAMPLES/noise\_tomography/PetersonNoiseModel.m}
-
-In Matlab, simply run:
-
-\texttt{NOISE\_TOMOGRAPHY(NSTEP, DT, Tmin, Tmax, NOISE\_MODEL)}\\
-
-\texttt{DT} is given in \texttt{Par\_file}, but \texttt{NSTEP} is
-NOT the one specified in \texttt{Par\_file}. Instead, you have to
-feed $2~\mathrm{NSTEP}-1$ to account for the doubled length of cross
-correlations. \texttt{Tmin} and \texttt{Tmax} correspond to the period
-range you are interested in, whereas \texttt{NOISE\_MODEL} denotes
-the noise model you will be using. Details can be found in the Matlab
-script.
-
-After running the Matlab script, you will be given the following information
-(plus a figure in Matlab):
-
-\texttt{{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}}\\
- \texttt{the source time function has been saved in:}\\
- \texttt{/data2/yangl/3D\_NOISE/S\_squared} (note this path must be
-different)\\
- \texttt{S\_squared should be put into directory:}\\
- \texttt{OUTPUT\_FILES/NOISE\_TOMOGRAPHY/ in the SPECFEM3D Cartesian
-package}\\
-
-In other words, the Matlab script creates a file called \texttt{S\_squared},
-which is the first `new' input file we encounter for noise simulations.
-
-One may choose a flat noise spectrum rather than Peterson's noise
-model. This can be done easily by modifying the Matlab script a little.
-
-\item Create a new directory in the SPECFEM3D Cartesian package, name it
-as \texttt{OUTPUT\_FILES/NOISE\_TOMOGRAPHY/}. We will add some parameter
-files later in this folder.
-\item Put the Matlab-generated-file \texttt{S\_squared} in \texttt{OUTPUT\_FILES/NOISE\_TOMOGRAPHY/}.
-
-That's to say, you will have a file \texttt{OUTPUT\_FILES/NOISE\_TOMOGRAPHY/S\_squared}
-in the SPECFEM3D Cartesian package.
-
-\item Create a file called \texttt{OUTPUT\_FILES/NOISE\_TOMOGRAPHY/irec\_master\_noise}.
-Note that this file is located in directory \texttt{OUTPUT\_FILES/NOISE\_TOMOGRAPHY/}
-as well. In general, all noise simulation related parameter files
-go into that directory. \texttt{irec\_master\_noise} contains only
-one integer, which is the ID of the `master' receiver. For example,
-if this file contains 5, it means that the fifth receiver listed in
-\texttt{DATA/STATIONS} becomes the `master'. That's why we mentioned
-previously that the order of receivers in \texttt{DATA/STATIONS} is
-important.
-
-Note that in regional simulations, the \texttt{DATA/STATIONS} might
-contain receivers which are outside of our computational domains.
-Therefore, the integer in \texttt{irec\_master\_noise} is actually
-the ID in \texttt{DATA/STATIONS\_FILTERED} (which is generated by
-\texttt{bin/xgenerate\_databases}).
-
-\item Create a file called \texttt{OUTPUT\_FILES/NOISE\_TOMOGRAPHY/nu\_master}.
-This file holds three numbers, forming a (unit) vector. It describes
-which component we are cross-correlating at the `master' receiver,
-i.e., ${\hat{{\bf \nu}}}^{\alpha}$ in \citet{trompetal2010}. The
-three numbers correspond to E/N/Z components respectively. Most often,
-the vertical component is used, and in those cases the three numbers
-should be 0, 0 and 1.
-\item Describe the noise direction and distributions in \texttt{src/shared/noise\_tomography.f90}.
-Search for a subroutine called \texttt{noise\_distribution\_direction}
-in \texttt{noise\_tomography.f90}. It is actually located at the very
-beginning of \texttt{noise\_tomography.f90}. The default assumes vertical
-noise and a uniform distribution across the whole free surface of
-the model. It should be quite self-explanatory for modifications.
-Should you modify this part, you have to re-compile the source code.
-\end{itemize}
-
-\subsection{Simulations}
-
-With all of the above done, we can finally launch our simulations.
-Again, please make sure that the \texttt{LOCAL\_PATH} in \texttt{DATA/Par\_file}
-is not globally shared. It is quite problematic to have a globally
-shared \texttt{LOCAL\_PATH}, in terms of both disk storage and speed
-of I/O (we have to save the wavefields at the earth's surface at every
-time step).
-
-As discussed in \citet{trompetal2010}, it takes three steps/simulations
-to obtain one contribution of the ensemble sensitivity kernels:
-\begin{itemize}
-\item Step 1: simulation for generating wavefields\\
- \texttt{SIMULATION\_TYPE=1}\\
- \texttt{NOISE\_TOMOGRAPHY=1}\\
- \texttt{SAVE\_FORWARD} not used, can be either .true. or .false.
-\item Step 2: simulation for ensemble forward wavefields\\
- \texttt{SIMULATION\_TYPE=1}\\
- \texttt{NOISE\_TOMOGRAPHY=2}\\
- \texttt{SAVE\_FORWARD=.true.}
-\item Step 3: simulation for ensemble adjoint wavefields and sensitivity
-kernels\\
- \texttt{SIMULATION\_TYPE=3}\\
- \texttt{NOISE\_TOMOGRAPHY=3}\\
- \texttt{SAVE\_FORWARD=.false.}
-
-Note Step 3 is an adjoint simulation, please refer to previous chapters
-on how to prepare adjoint sources and other necessary files, as well
-as how adjoint simulations work.
-
-\end{itemize}
-Note that it is better to run the three steps continuously within
-the same job, otherwise you have to collect the saved surface movies
-from the old nodes/CPUs to the new nodes/CPUs. This process varies
-from cluster to cluster and thus cannot be discussed here. Please
-ask your cluster administrator for information/configuration of the
-cluster you are using.\\
-
-\subsection{Post-simulation}
-
-After those simulations, you have all stuff you need, either in the
-\texttt{OUTPUT\_FILES/} or in the directory specified by \texttt{LOCAL\_PATH}
-in \texttt{DATA/Par\_file} (which are most probably on local nodes).
-Collect whatever you want from the local nodes to your workstation,
-and then visualize them. This process is the same as what you may
-have done for regular earthquake simulations. Refer to other chapters
-if you have problems.\\
-
-Simply speaking, two outputs are the most interesting: the simulated
-ensemble cross correlations and one contribution of the ensemble sensitivity
-kernels.\\
-
-The simulated ensemble cross correlations are obtained after the second
-simulation (Step 2). Seismograms in \texttt{OUTPUT\_FILES/} are actually
-the simulated ensemble cross correlations. Collect them immediately
-after Step 2, or the Step 3 will overwrite them. Note that we have
-a `master' receiver specified by \texttt{irec\_master\_noise}, the
-seismogram at one station corresponds to the cross correlation between
-that station and the `master'. Since the seismograms have three components,
-we may obtain cross correlations for different components as well,
-not necessarily the cross correlations between vertical components.\\
-
-One contribution of the ensemble cross-correlation sensitivity kernels
-are obtained after Step 3, stored in the \texttt{DATA/LOCAL\_PATH}
-on local nodes. The ensemble kernel files are named the same as regular
-earthquake kernels.\\
-
-You need to run another three simulations to get the other contribution
-of the ensemble kernels, using different forward and adjoint sources
-given in \citet{trompetal2010}.
-
-\section{Example}
-
-In order to illustrate noise simulations in an easy way, one example
-is provided in \texttt{EXAMPLES/noise\_tomography/}. See \texttt{EXAMPLES/noise\_tomography/README}
-for explanations. \\
-
-Note, however, that they are created for a specific workstation (CLOVER@PRINCETON),
-which has at least 4 cores with `mpif90' working properly. \\
-
-If your workstation is suitable, you can run the example in \texttt{EXAMPLES/noise\_tomography/}
-using:\\
-
-\texttt{./pre-processing.sh}\\
-
-Even if this script does not work on your workstation, the procedure
-it describes is universal. You may review the whole process described
-in the last section by following the commands in \texttt{pre-processing.sh},
-which should contain enough explanations for all the commands.
-
-%------------------------------------------------------------------------------------------------%
-
-\chapter{Gravity integral calculations for the gravity field of an Earth model}
-
-%------------------------------------------------------------------------------------------------%
-
-SPECFEM can now compute the gravity field as well as its derivatives (i.e., gravity gradiometry)
-generated by any given 3D Earth model at a given elevation, i.e., on a given observation surface.
-That feature is still experimental but should work just fine.\\
-
-To see how it is implemented and to use it, type this in the root directory of the code:\\
-
-\texttt{grep -i GRAVITY\_INTEGRALS src/*/*}
-
-%------------------------------------------------------------------------------------------------%
-
-\chapter{Graphics}
-
-%------------------------------------------------------------------------------------------------%
-
-\section{\label{sec:Mesh-graphics}Meshes}
-
-In case you used the internal mesher \texttt{xmeshfem3D} to create
-and partition your mesh, you can output mesh files in ABAQUS (.INP)
-and DX (.dx) format to visualize them. For this, you must set either
-the flag \texttt{CREATE\_DX\_FILES} or \texttt{CREATE\_ABAQUS\_FILES}
-to \texttt{.true.} in the mesher's parameter file \texttt{Mesh\_Par\_file}
-prior to running the mesher (see Chapter~\ref{cha:Running-the-Mesher-Meshfem3D}
-for details). You can then use AVS \urlwithparentheses{www.avs.com}
-or OpenDX \urlwithparentheses{www.opendx.org} to visualize the mesh
-and MPI partition (slices).
-
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=0.49\textwidth]{figures/vtk_mesh_vp} \includegraphics[width=0.49\textwidth]{figures/vtk_mesh_vs}
-\par\end{centering}
-
-\caption{Visualization using Paraview of VTK files created by \texttt{xgenerate\_databases}
-showing P- and S-wave velocities assigned to the mesh points. The
-mesh was created by \texttt{xmeshfem3D} for 4 processors.}
-
-\label{fig:vtk.mesh}
-\end{figure}
-
-You have also the option to visualize the distributed databases produced
-by \texttt{xgenerate\_databases} using Paraview \urlwithparentheses{www.paraview.org}.
-For this, you must set the flag \texttt{SAVE\_MESH\_FILES} to \texttt{.true.}
-in the main parameter file \texttt{Par\_file} (see Chapter~\ref{cha:Main-Parameter}
-for details). This will create VTK files for each single partition.
-You can then use Paraview \urlwithparentheses{www.paraview.org}
-to visualized these partitions.
-
-\section{\label{sec:Movies}Movies}
-
-To make a surface or volume movie of the simulation, set parameters
-\texttt{MOVIE\_SURFACE}, \texttt{MOVIE\_VOLUME}, \texttt{MOVIE\_TYPE},
-and \texttt{NTSTEP\_BETWEEN\_FRAMES} in the \texttt{Par\_file}. Turning
-on the movie flags, in particular \texttt{MOVIE\_VOLUME}, produces
-large output files. \texttt{MOVIE\_VOLUME} files are saved in the
-\texttt{LOCAL\_PATH} directory, whereas \texttt{MOVIE\_SURFACE} output
-files are saved in the \texttt{OUTPUT\_FILES} directory. We save the
-displacement field if the parameter \texttt{SAVE\_DISPLACEMENT} is
-set, otherwise the velocity field is saved. The look of a movie is
-determined by the half-duration of the source. The half-duration should
-be large enough so that the movie does not contain frequencies that
-are not resolved by the mesh, i.e., it should not contain numerical
-noise. This can be accomplished by selecting a CMT \texttt{HALF\_DURATION}
-> 1.1 $\times$ smallest period (see figure \ref{fig:CMTSOLUTION-file}).
-When \texttt{MOVIE\_SURFACE} = .\texttt{true.}, the half duration
-of each source in the \texttt{CMTSOLUTION} file is replaced by
-\begin{quote}
-\[
-\sqrt{(}\mathrm{\mathtt{HALF\_DURATIO}\mathtt{N}^{2}}+\mathrm{\mathtt{HDUR\_MOVI}\mathtt{E}^{2}})
-\]
- \textbf{NOTE:} If \texttt{HDUR\_MOVIE} is set to 0.0, the code will
-select the appropriate value of 1.1 $\times$ smallest period. As
-usual, for a point source one can set \texttt{half duration} in the
-\texttt{CMTSOLUTION} file to be 0.0 and \texttt{HDUR\_MOVIE} = 0.0
-in the \texttt{Par\_file} to get the highest frequencies resolved
-by the simulation, but for a finite source one would keep all the
-\texttt{half durations} as prescribed by the finite source model and
-set \texttt{HDUR\_MOVIE} = 0.0.
-\end{quote}
-
-\subsection{Movie Surface}
-
-When running \texttt{xspecfem3D} with the \texttt{MOVIE\_SURFACE}
-flag turned on, the code outputs \texttt{moviedata??????} files in
-the \texttt{OUTPUT\_FILES} directory. There are several flags in the
-main parameter file \texttt{Par\_file} that control the output of
-these movie data files (see section \ref{cha:Main-Parameter} for
-details): \texttt{\small NTSTEP\_BETWEEN\_FRAMES}{\small{} to set the
-timesteps between frames, }\texttt{\small SAVE\_DISPLACEMENT}{\small{}
-to save displacement instead of velocity, }\texttt{\small USE\_HIGHRES\_FOR\_MOVIES}{\small{}
-to save values at all GLL point instead of element edges. In order
-to output additionally shakemaps, you would set the parameter }\texttt{\small CREATE\_SHAKEMAP}{\small{}
-to }\texttt{\small .true.}{\small .}{\small \par}
-
-The files are in a fairly complicated binary format, but there is
-a program provided to convert the output into more user friendly formats:
-\begin{description}
-\item [{\texttt{xcreate\_movie\_shakemap\_AVS\_DX\_GMT}}] From \texttt{create\_movie\_shakemap\_AVS\_DX\_GMT.f90},
-it outputs data in ASCII, OpenDX, or AVS format (also readable in
-ParaView). Before compiling the code, make sure you have the file
-\texttt{surface\_from\_mesher.h} in the \texttt{OUTPUT\_FILES/} directory.
-This file will be created by the solver run. Then type
-
-\begin{lyxcode}
-{\footnotesize make~xcreate\_movie\_shakemap\_AVS\_DX\_GMT}{\footnotesize \par}
-\end{lyxcode}
-
-and run the executable \texttt{xcreate\_movie\_shakemap\_AVS\_DX\_GMT}
-in the main directory. It will create visualization files
-in your format of choice. The code will prompt the user for input
-parameters.
-
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=0.32\textwidth]{figures/movie_surf_1} \includegraphics[width=0.32\textwidth]{figures/movie_surf_2}
-\includegraphics[width=0.32\textwidth]{figures/movie_surf_3}
-\par\end{centering}
-
-\caption{Visualization using AVS files created by \texttt{xcreate\_movie\_shakemap\_AVS\_DX\_GMT}
-showing movie snapshots of vertical velocity components at different
-times.}
-
-\label{fig:movie.surf}
-\end{figure}
-
-\end{description}
-The \texttt{SPECFEM3D Cartesian} code is running in near real-time
-to produce animations of southern California earthquakes via the web;
-see Southern California ShakeMovie\textregistered{}\urlwithparentheses{www.shakemovie.caltech.edu}.
-
-\subsection{\label{sub:Movie-Volume}Movie Volume}
-
-When running xspecfem3D with the \texttt{\small MOVIE\_VOLUME}{\small{}
-flag turned on, the code outputs several files in }\texttt{\small LOCAL\_PATH}{\small{}
-specified in the main }\texttt{\small Par\_file}{\small , e.g. in
-directory }\texttt{\small OUTPUT\_FILES/DATABASES\_MPI}{\small .
-The output is saved by each processor at the time interval specified
-by }\texttt{\small NTSTEP\_BETWEEN\_FRAMES}{\small . For all domains,
-the velocity field is output to files: }\texttt{\small proc??????\_velocity\_X\_it??????.bin}{\small ,
-}\texttt{\small proc??????\_velocity\_Y\_it??????.bin}{\small{} and
-}\texttt{\small proc??????\_velocity\_Z\_it??????.bin}{\small . For
-elastic domains, the divergence and curl taken from the velocity field,
-i.e. $\nabla\cdot{\bf {v}}$ and $\nabla\times{\bf {v}}$, get stored
-as well: }\texttt{\small proc??????\_div\_it??????.bin}{\small , }\texttt{\small proc??????\_curl\_X\_t??????.bin}{\small ,
-}\texttt{\small proc??????\_curl\_Y\_it??????.bin}{\small{} and }\texttt{\small proc??????\_curl\_Z\_it??????.bin}{\small .
-The files denoted }\texttt{\small proc??????\_div\_}{\small ~}\\
-{\small{} }\texttt{\small glob\_it??????.bin}{\small{} and }\texttt{\small proc??????\_curl\_glob\_it??????.bin}{\small{}
-are stored on the global points only, all the other arrays are stored
-on all GLL points. Note that the components X/Y/Z can change to E/N/Z
-according to the }\texttt{\small SUPPRESS\_UTM\_PROJECTION}{\small{}
-flag (see also Appendix~\ref{cha:Coordinates} and \ref{cha:channel-codes}).}{\small \par}
-
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=0.32\textwidth]{figures/movie_volume_1} \includegraphics[width=0.32\textwidth]{figures/movie_volume_2}
-\includegraphics[width=0.32\textwidth]{figures/movie_volume_3}
-\par\end{centering}
-
-\caption{Paraview visualization using movie volume files (converted by \texttt{xcombine\_vol\_data}
-and \texttt{mesh2vtu.pl}) and showing snapshots of vertical velocity
-components at different times.}
-
-\label{fig:movie.volume}
-\end{figure}
-
-To visualize these files, we use an auxilliary program \texttt{combine\_vol\_data.f90}
-to combine the data from all slices into one mesh file. To compile
-it in the root directory, type:
-\begin{lyxcode}
-{\footnotesize make~xcombine\_vol\_data~}{\footnotesize \par}
-\end{lyxcode}
-which will create the executable \texttt{bin/xcombine\_vol\_data}.
-To output the usage of this executable, type
-\begin{lyxcode}
-{\footnotesize ./bin/xcombine\_vol\_data~}{\footnotesize \par}
-\end{lyxcode}
-without arguments.
-
-\medskip
-
-\texttt{xcombine\_vol\_data} will combine the data on the different processors (located in \texttt{OUTPUT\_FILES/DATABASES\_MPI}), for a given quantity and a given iteration, to one file.
-For example, if you want to combine \texttt{velocity\_Z}, for the iteration 400 on 4 processors, i.e.
-if you want to combine these files :
-
-\medskip
-
-\texttt{proc000000\_velocity\_Z\_it000400.bin }
-
-\texttt{proc000001\_velocity\_Z\_it000400.bin }
-
-\texttt{proc000002\_velocity\_Z\_it000400.bin }
-
-\texttt{proc000003\_velocity\_Z\_it000400.bin}
-
-\medskip
-
-You have to go in the directory of the concerned example (where are the directories \texttt{DATA}, \texttt{OUTPUT\_FILES}, etc.).
-Then, you can launch \texttt{xcombine\_vol\_data}, specifying the path where it is located.
-As an example, if the \texttt{DATA} and \texttt{OUTPUT\_FILES} directories of the concerned example are in the root specfem3d directory, \texttt{xcombine\_vol\_data} is in \texttt{./bin}, and you have to type:
-
-\begin{lyxcode}
-~{\footnotesize ./bin/xcombine\_vol\_data~0~3~velocity\_Z\_it000400~./OUTPUT\_FILES/DATABASES\_MPI~./OUTPUT\_FILES~0~}{\footnotesize \par}
-\end{lyxcode}
-
-Here, $0$ is the number of the first processor, $3$ the number of the last one, \texttt{velocity\_Z\_it000400} the name of the files we want to combine without the prefix \texttt{proc000000*\_}, \texttt{./OUTPUT\_FILES/DATABASES\_MPI} the directory where the files are located, \texttt{./OUTPUT\_FILES} the directory where the combined file will be stored, and $0$ is the parameter to create a low-resolution mesh file ($1$ for a high-resolution).
-
-\medskip
-
-At the beginning of the source file \texttt{combine\_vol\_data.f90} (located in \texttt{src/auxiliaries/}), there is a logical constant parameter \texttt{USE\_VTK\_OUTPUT} whose default value is \texttt{.true.}, that which will provide the output combined file directly in the VTK format and you can use Paraview to visualize it.
-
-\medskip
-
-If the value of this parameter is \texttt{.false.}, the output mesh file will have the name \texttt{velocity\_Z\_it000400.mesh}. Then we next have to convert the
-\texttt{.mesh} file into the VTU (Unstructured grid file) format which can be viewed in ParaView. For this task, you can use and modify the
-script \texttt{mesh2vtu.pl} located in directory \texttt{utils/Visualization/Paraview/}, for example:
-\begin{lyxcode}
-{\footnotesize mesh2vtu.pl~-i~velocity\_Z\_it000400.mesh~-o~velocity\_Z\_it000400.vtu}{\footnotesize \par}
-\end{lyxcode}
-Notice that this Perl script uses a program \texttt{mesh2vtu} in the \texttt{utils/Visualization/Paraview/mesh2vtu} directory, which further
-uses the VTK \urlwithparentheses{http://www.vtk.org/} run-time library for its execution. Therefore, make sure you have them properly set
-in the script according to your system.
-
-\bigskip
-
-Then, to do a movie with several iterations, you have to repeat this process for each iteration you want to put in your movie.
-
-\section{\label{sec:Finite-Frequency-Kernels}Finite-Frequency Kernels}
-
-The finite-frequency kernels computed as explained in Section \ref{sec:Adjoint-simulation-finite}
-are saved in the \texttt{LOCAL\_PATH} at the end of the simulation.
-Therefore, we first need to collect these files on the front end,
-combine them into one mesh file, and visualize them with some auxilliary
-programs.
-\begin{enumerate}
-\item \textbf{Create slice files}
-
-We will only discuss the case of one source-receiver pair, i.e., the
-so-called banana-doughnut kernels. Although it is possible to collect
-the kernel files from all slices on the front end, it usually takes
-up too much storage space (at least tens of gigabytes). Since the
-sensitivity kernels are the strongest along the source-receiver great
-circle path, it is sufficient to collect only the slices that are
-along or close to the great circle path.
-
-A Perl script \texttt{slice\_number.pl} located in directory \texttt{utils/Visualization/Paraview/}
-can help to figure out the slice numbers that lie along the great
-circle path. It applies to meshes created with the internal mesher
-\texttt{xmeshfem3D}.
-\begin{enumerate}
-\item On machines where you have access to the script, copy the \texttt{Mesh\_Par\_file},
-and \texttt{output\_solver} files, and run:
-
-\begin{lyxcode}
-{\small slice\_number.pl~Mesh\_Par\_file~output\_solver.txt~slice\_file}{\small \par}
-\end{lyxcode}
-
-which will generate a \texttt{slices\_file}.
-
-\item For cases with multiple sources and multiple receivers, you need to
-provide a slice file before proceeding to the next step.
-\end{enumerate}
-\item \textbf{Collect the kernel files}
-
-After obtaining the slice files, you can collect the corresponding
-kernel files from the given slices.
-\begin{enumerate}
-\item You can use or modify the script \texttt{utils/copy\_basin\_database.pl}
-to accomplish this:
-
-\begin{lyxcode}
-{\small utils/copy\_database.pl~slice\_file~lsf\_machine\_file~filename~{[}jobid{]}}{\small \par}
-\end{lyxcode}
-
-where \texttt{\small lsf\_machine\_file}{\small{} is the machine file
-generated by the LSF scheduler, }\texttt{\small filename}{\small{}
-is the kernel name (e.g., }\texttt{\small rho\_kernel}{\small , }\texttt{\small alpha\_kernel}{\small{}
-and }\texttt{\small beta\_kernel}{\small ), and the optional }\texttt{\small jobid}{\small{}
-is the name of the subdirectory under }\texttt{\small LOCAL\_PATH}{\small{}
-where all the kernel files are stored.}{\small \par}
-
-\item After executing this script, all the necessary mesh topology files
-as well as the kernel array files are collected to the local directory
-of the front end.
-\end{enumerate}
-\item \textbf{Combine kernel files into one mesh file}
-
-We use an auxilliary program \texttt{combine\_vol\_data.f90} to combine
-the kernel files from all slices into one mesh file.
-\begin{enumerate}
-\item Compile it in the root directory:
-
-\begin{lyxcode}
-{\footnotesize make~xcombine\_vol\_data~}{\footnotesize \par}
-
-{\footnotesize ./bin/xcombine\_vol\_data~slice\_list~filename~input\_dir~output\_dir~high/low-resolution~}{\footnotesize \par}
-\end{lyxcode}
-
-where \texttt{input\_dir} is the directory where all the individual
-kernel files are stored, and \texttt{output\_dir} is where the mesh
-file will be written.
-
-\item Use 1 for a high-resolution mesh, outputting all the GLL points to
-the mesh file, or use 0 for low resolution, outputting only the corner
-points of the elements to the mesh file.
-\item The output mesh file will have the name \texttt{filename\_rho(alpha,beta).mesh}
-\end{enumerate}
-\item \textbf{Convert mesh files into .vtu files}
-
-\begin{enumerate}
-\item We next convert the \texttt{.mesh} file into the VTU (Unstructured
-grid file) format which can be viewed in ParaView. For this task,
-you can use and modify the script \texttt{mesh2vtu.pl} located in
-directory \texttt{utils/Visualization/Paraview/}, for example:
-
-\begin{lyxcode}
-{\footnotesize mesh2vtu.pl~-i~file.mesh~-o~file.vtu~}{\footnotesize \par}
-\end{lyxcode}
-\item Notice that this Perl script uses a program \texttt{mesh2vtu} in the
-\texttt{utils/Visualization/Paraview/mesh2vtu} directory, which further
-uses the VTK \urlwithparentheses{http://www.vtk.org/} run-time library
-for its execution. Therefore, make sure you have them properly set
-in the script according to your system.
-\end{enumerate}
-\item \textbf{Copy over the source and receiver .vtk file}
-
-In the case of a single source and a single receiver, the simulation
-also generates the file \texttt{sr.vtk} located in the \texttt{OUTPUT\_FILES/}
-directory to describe the source and receiver locations, which can
-also be viewed in Paraview in the next step.
-
-\item \textbf{View the mesh in ParaView}
-
-Finally, we can view the mesh in ParaView \urlwithparentheses{www.paraview.org}.
-\begin{enumerate}
-\item Open ParaView.
-\item From the top menu, \textsf{File} $\rightarrow$\textsf{Open data},
-select \texttt{file.vtu}, and click the \textsf{Accept} button.
-
-\begin{itemize}
-\item If the mesh file is of moderate size, it shows up on the screen; otherwise,
-only the bounding box is shown.
-\end{itemize}
-\item Click \textsf{Display Tab} $\rightarrow$ \textsf{Display Style} $\rightarrow$
-\textsf{Representation} and select \textsf{wireframe of surface} to
-display it.
-\item To create a cross-section of the volumetric mesh, choose \textsf{Filter}
-$\rightarrow$ \textsf{cut}, and under \textsf{Parameters Tab}, choose
-\textsf{Cut Function} $\rightarrow$ \textsf{plane}.
-\item Fill in center and normal information given by the \texttt{global\_slice\_number.pl}
-script (either from the standard output or from \texttt{normal\_plane.txt}
-file).
-\item To change the color scale, go to \textsf{Display Tab} $\rightarrow$
-\textsf{Color} $\rightarrow$ \textsf{Edit Color Map} and reselect
-lower and upper limits, or change the color scheme.
-\item Now load in the source and receiver location file by \textsf{File}
-$\rightarrow$ \textsf{Open data}, select \texttt{sr.vt}k, and click
-the \textsf{Accept} button. Choose \textsf{Filter} $\rightarrow$
-\textsf{Glyph}, and represent the points by `\textsf{spheres}'.
-\item For more information about ParaView, see the ParaView Users Guide
-\urlwithparentheses{www.paraview.org/files/v1.6/ParaViewUsersGuide.PDF}.
-\end{enumerate}
-\end{enumerate}
-\begin{figure}[H]
-\noindent \begin{centering}
-\includegraphics[scale=0.6]{figures/3D-S-Kernel}
-\par\end{centering}
-
-\caption{(a) Top Panel: Vertical source-receiver cross-section of the S-wave
-finite-frequency sensitivity kernel $K_{\beta}$ for station GSC at
-an epicentral distance of 176 km from the September 3, 2002, Yorba
-Linda earthquake. Lower Panel: Vertical source-receiver cross-section
-of the 3D S-wave speed model used for the spectral-element simulations
-\citep{KoLiTrSuStSh04}. (b) The same as (a) but for station HEC at
-an epicentral distance of 165 km \citep{LiTr06}.}
-
-\label{figure:P-wave-speed-finite-frequency}
-\end{figure}
-
-\chapter{\label{cha:Scheduler}Running through a Scheduler}
-
-The code is usually run on large parallel machines, often PC clusters,
-most of which use schedulers, i.e., queuing or batch management systems
-to manage the running of jobs from a large number of users. The following
-considerations need to be taken into account when running on a system
-that uses a scheduler:
-\begin{itemize}
-\item The processors/nodes to be used for each run are assigned dynamically
-by the scheduler, based on availability. Therefore, in order for the
-\texttt{xgenerate\_databases} and the \texttt{xspecfem3D} executables
-(or between successive runs of the solver) to have access to the same
-database files (if they are stored on hard drives local to the nodes
-on which the code is run), they must be launched in sequence as a
-single job.
-\item On some systems, the nodes to which running jobs are assigned are
-not configured for compilation. It may therefore be necessary to pre-compile
-both the \texttt{xgenerate\_databases} and the \texttt{xspecfem3D}
-executables.
-\item One feature of schedulers/queuing systems is that they allow submission
-of multiple jobs in a ``launch and forget'' mode. In order to take
-advantage of this property, care needs to be taken that output and
-intermediate files from separate jobs do not overwrite each other,
-or otherwise interfere with other running jobs.
-\end{itemize}
-Examples of job scripts can be found in the \texttt{\small utils/Cluster/}{\small{}
-directory and can straightforwardly be modified and adapted to meet
-more specific running needs.}{\small \par}
-
-\chapter{{\normalsize \label{cha:-Changing-the}} Changing the Model}
-
-In this section we explain how to change the velocity model used for
-your simulations. These changes involve contributing specific subroutines
-that replace existing subroutines in the \texttt{SPECFEM3D Cartesian}
-package. Note that \texttt{SPECFEM3D Cartesian} can handle Earth models
-with material properties that vary within each spectral element.
-
-\section{{\normalsize \label{sec:Using-tomographic}}Using external tomographic Earth models}
-
-To implement your own external tomographic model(s), you must provide
-your own external tomography file(s), and choose between two possible
-options:\\
- \indent (1) set in the \texttt{Par\_file} the model parameter \texttt{MODEL
-= tomo}, \\
- or for more user control: \\
- \indent (2) set in the \texttt{Par\_file} the model parameter \texttt{MODEL
-= default}, define the negative \texttt{material\_ID} identifier for
-each element in the file \texttt{MESH/materials\_file} and use the
-following format in the file \texttt{MESH/nummaterial\_velocity\_file}
-when using CUBIT to construct your mesh (see also section \ref{subsec:Exporting-the-Mesh}):
-\begin{lyxcode}
-domain\_ID~material\_ID~tomography~elastic~file\_name~1
-\end{lyxcode}
-where: \\
- \indent \texttt{domain\_ID} is 1 for acoustic or 2 for elastic materials,
-\\
- \indent \texttt{material\_ID} a negative, unique identifier (i.e.,
--1,-2,...), \\
- \indent \texttt{tomography} keyword for tomographic material definition,
-\\
- \indent \texttt{elastic} keyword for elastic material definition,
-\\
- \indent \texttt{file\_name} the name of the tomography file and
-1 a positive unique identifier.\\
-
-The external tomographic model is represented by a grid of points
-with assigned material properties and homogeneous resolution along
-each spatial direction x, y and z. The ASCII file \texttt{file\_name}
-that describe the tomography should be located in the \texttt{TOMOGRAPHY\_PATH}
-directory, set in the \texttt{Par\_file}. The format of the file,
-as read from \texttt{model\_tomography.f90} located in the \texttt{src/generate\_databases}
-directory, looks like Figure \ref{fig:tomography_file}, where
-\begin{description}
-\item [{\texttt{ORIG\_X}, \texttt{END\_X}}] are, respectively, the coordinates
-of the initial and final tomographic grid points along the x direction
-(in the mesh units, e.g., $m$);
-\item [{\texttt{ORIG\_Y}, \texttt{END\_Y}}] respectively the coordinates
-of the initial and final tomographic grid points along the y direction
-(in the mesh units, e.g., $m$);
-\item [{\texttt{ORIG\_Z}, \texttt{END\_Z}}] respectively the coordinates
-of the initial and final tomographic grid points along the z direction
-(in the mesh units, e.g., $m$);
-\item [{\texttt{SPACING\_X}, \texttt{SPACING\_Y}, \texttt{SPACING\_Z}}] the
-spacing between the tomographic grid points along the x, y and z directions,
-respectively (in the mesh units, e.g., $m$);
-\item [{\texttt{NX}, \texttt{NY}, \texttt{NZ}}] the number of grid points
-along the spatial directions x, y and z, respectively; \texttt{NX}
-is given by {[}(\texttt{END\_X} - \texttt{ORIG\_X})/\texttt{SPACING\_X}{]}+1;
-\texttt{NY} and \texttt{NZ} are the same as \texttt{NX}, but for the
-y and z directions, respectively;
-\item [{\texttt{VP\_MIN}, \texttt{VP\_MAX}, \texttt{VS\_MIN}, \texttt{VS\_MAX}, \texttt{RHO\_MIN}, \texttt{RHO\_MAX}}] the
-minimum and maximum values of the wave speed \texttt{vp} and \texttt{vs}
-(in $m\, s^{-1}$) and of the density \texttt{rho} (in $kg\, m^{-3}$);
-these values could be the actual limits of the tomographic parameters
-in the grid or the minimum and maximum values to which we force the
-cut of velocity and density in the model.
-\end{description}
-After the first four lines, in the file \texttt{file\_name} the tomographic
-grid points are listed with the corresponding values of \texttt{vp},
-\texttt{vs} and \texttt{rho}, scanning the grid along the x coordinate
-(from \texttt{ORIG\_X} to \texttt{END\_X} with step of \texttt{SPACING\_X})
-for each given y (from \texttt{ORIG\_Y} to \texttt{END\_Y}, with step
-of \texttt{SPACING\_Y}) and each given z (from \texttt{ORIG\_Z} to
-\texttt{END\_Z}, with step of \texttt{SPACING\_Z}). %fig
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=1\textwidth]{figures/tomo_file}
-\par\end{centering}
-
-\caption{Tomography file \texttt{file\_name} that describes an external Earth
-model. The coordinates x, y and z of the grid, their limits ORIG\_X,
-ORIG\_Y, ORIG\_Z, END\_X, END\_Y, END\_Z and the grid spacings SPACING\_X,
-SPACING\_Y, SPACING\_Z should be in the units of the constructed mesh
-(e.g., UTM coordinates in meters).}
-
-\label{fig:tomography_file}
-\end{figure}
-
-The user can implement his own interpolation algorithm for the tomography
-model by changing the routine \texttt{model\_tomography.f90} located
-in the \texttt{src/generate\_databases/} directory. Moreover, for
-models that involve both fully defined materials and a tomography
-description, the \texttt{nummaterial\_velocity\_file} has multiple
-lines each with the corresponding suitable format described above.
-
-%==============
-
-\subsubsection*{Example: External tomography file with variable discretization intervals}
-
-The example in Figure~\ref{fig:tomography_file} is for an external tomography file with uniform spacing in the $x$, $y$, and $z$ directions.
-(Note that $\Delta x$, $\Delta y$, and $\Delta z$ need not be the same as in the example.) In that example, the  \texttt{nummaterial\_velocity\_file} is
-%
-\begin{lyxcode}
-2  -1 tomography elastic tomography\_model.xyz 1
-\end{lyxcode}
-%
-and the file \texttt{tomography\_model.xyz} will need to reside in \texttt{DATA/}.
-All entries in the second column of \texttt{materials\_file} will be \texttt{-1}, which means that each element in the mesh will be interpolated accordind to the values in \texttt{tomography\_model.xyz}.
-
-In some cases it may be desirable to have an external tomography model that is described in more than one file.
-For example, in cases like southern California, the length scale of variation in the structure of the wave speed model is much shorter in the sedimentary basin models within the upper 15~km.
-Therefore one might want to use an external tomography file that is sampled with, say, $\Delta x = 1000$~m, $\Delta y = 1000$~m, and $\Delta z = 250$~m in the uppermost 15~km, and then use $\Delta x = 2000$~m, $\Delta y = 2000$~m, and $\Delta z = 1000$~m below a depth of 15~km.
-If these intervals are chosen appropriately, then it will result in a pair of external tomography file that is much smaller than the alternative of having a single file with the fine discretization.
-In this case  \texttt{nummaterial\_velocity\_file} is
-%
-\begin{lyxcode}
-2  -1 tomography elastic file\_above\_15km.xyz 1 \\
-2  -2 tomography elastic file\_below\_15km.xyz 1
-\end{lyxcode}
-%
-and the files \texttt{file\_above\_15km.xyz} and \texttt{file\_below\_15km.xyz} will need to reside in \texttt{DATA/}.
-All entries in the second column of \texttt{materials\_file} will need to be either \texttt{-1} or \texttt{-2}, depending on whether the element is above or below 15~km depth.
-In this sense, the construction of the mesh and the discretization of the external model will generally be done in tandem.
-
-Other more complicated discretizations can be used following the same procedure. (In particular, there is no limit on the number of different external files that are used.)
-
-%=============
-
-\section{{\normalsize \label{sec:Anelastic-Models}}External (an)elastic Models}
-
-To use your own external model, you can set in the \texttt{Par\_file}
-the model parameter \texttt{MODEL = external}. Three-dimensional acoustic
-and/or (an)elastic (attenuation) models may then be superimposed onto
-the mesh based upon your subroutine in \texttt{model\_external\_values.f90}
-located in directory \texttt{src/generate\_databases}. The call to
-this routine would be as follows
-\begin{lyxcode}
-call~model\_external\_values(xmesh,~ymesh,~zmesh,~~\&~~\\
-~~~~~~~~~~~rho,~vp,~vs,~qmu\_atten,~iflag\_aniso,~idomain\_id)
-\end{lyxcode}
-Input to this routine consists of:
-\begin{description}
-\item [{\texttt{xmesh,ymesh,zmesh}}] location of mesh point
-\end{description}
-Output to this routine consists of:
-\begin{description}
-\item [{\texttt{rho,vp,vs}}] isotropic model parameters for density $\rho$
-($kg/m^{3}$), $v_{p}$ ($m/s$) and $v_{s}$ ($m/s$)
-\item [{\texttt{qmu\_atten}}] Shear wave quality factor: $0<Q_{\mu}<9000$
-\item [{\texttt{iflag\_aniso}}] anisotropic model flag, $0$ indicating
-no anisotropy or $1$ using anisotropic model parameters as defined
-in routine file \texttt{model\_aniso.f90}
-\item [{\texttt{idomain\_id}}] domain identifier, $1$ for acoustic, $2$
-for elastic, $3$ for poroelastic materials.
-\end{description}
-Note that the resolution and maximum value of anelastic models are
-truncated. This speeds the construction of the standard linear solids
-during the meshing stage. To change the resolution, currently at one
-significant figure following the decimal, or the maximum value (9000),
-consult \texttt{shared/constants.h}.
-
-\section{{\normalsize \label{sec:Anisotropic-Models}}Anisotropic Models}
-
-To use your anisotropic models, you can either set in the \texttt{Par\_file}
-the model parameter \texttt{ANISOTROPY} to \texttt{.true.} or set
-the model parameter \texttt{MODEL = aniso}. Three-dimensional anisotropic
-models may be superimposed on the mesh based upon the subroutine in
-file \texttt{model\_aniso.f90} located in directory \texttt{src/generate\_databases}.
-The call to this subroutine is of the form
-\begin{lyxcode}
-call~model\_aniso(iflag\_aniso,~rho,~vp,~vs,~\&~~\\
-~~~~~~~~~~~c11,c12,c13,c14,c15,c16,c22,c23,c24,c25,c26,~\&~~\\
-~~~~~~~~~~~c33,c34,c35,c36,c44,c45,c46,c55,c56,c66)~
-\end{lyxcode}
-Input to this routine consists of:
-\begin{description}
-\item [{\texttt{iflag\_aniso}}] flag indicating the type of the anisotropic
-model, i.e. $0$ for no superimposed anisotropy, $1$ or $2$ for
-generic pre-defined anisotropic models.
-\item [{\texttt{rho,vp,vs}}] reference isotropic model parameters for density
-$\rho$, $v_{p}$ and $v_{s}$.
-\end{description}
-Output from the routine consists of the following non-dimensional
-model parameters:
-\begin{description}
-\item [{\texttt{c11},}] \textbf{$\cdots$,} \texttt{\textbf{c66}} 21~dimensionalized
-anisotropic elastic parameters.
-\end{description}
-You can replace the \texttt{model\_aniso.f90} file by your own version
-\textit{provided you do not change the call structure of the routine},
-i.e., the new routine should take exactly the same input and produce
-the required relative output.
-
-\section{{\normalsize \label{sec:Using-SEP}}Using external SEP models.}
-
-Stanford Exploration Project (SEP) models consists of two files per variables, one header file (\texttt{.H})
-and one binary file (\texttt{.H$@$}).
-Among other information, the header file indicates, for each dimension,
-the offset (\texttt{o1}, \texttt{o2} and \texttt{o3} fields),
-the spatial step size (\texttt{d1}, \texttt{d2} and \texttt{d3} fields)
-and the number of samples (\texttt{n1}, \texttt{n2} and \texttt{n3}) necessary to read
-the linked binary file (\texttt{in} field).
-
-SEP model reading occurs during database generation. It it the user responsibility
-to ensure that proper interfaces between acoustic and elastic domains are set up
-during the meshing phase. For more information, refer to
-section~\ref{cha:Running-the-Mesher-Meshfem3D}.
-
-In \texttt{DATA/Par\_file}, \texttt{\textbf{MODEL}} should be set to \texttt{\textbf{sep}},
-and \texttt{\textbf{SEP\_MODEL\_DIRECTORY}} should point to wherever your sep model files reside.
-
-Note that in order not to overload the parameter file, SEP header files
-should be named \texttt{vp.H}, \texttt{vs.H} and \texttt{rho.H}. The \texttt{in} field
-in these binary files can be any path as long as it is relative to
-the previously set \texttt{SEP\_MODEL\_DIRECTORY}.
-We also assume that any dimensional value inside SEP header files is given in meters and
-that binary files are single precision floating point numbers stored in little endian format.
-There is currently only support for \texttt{vp}, \texttt{vs} and \texttt{rho} variables.
-
-An example demonstrating the use SEP models is given in:
-\texttt{EXAMPLES/meshfem3D\_examples/sep\_bathymetry}.
-
-\chapter{Post-Processing Scripts}
-
-Several post-processing scripts/programs are provided in the \texttt{utils/}
-directory, and most of them need to be adjusted when used on different
-systems, for example, the path of the executable programs. Here we
-only list a few of the available scripts and provide a brief description,
-and you can either refer to the related sections for detailed usage
-or, in a lot of cases, type the script/program name without arguments
-for its usage.
-
-\section{Process Data and Synthetics\label{sec:Process-data-and-syn}}
-
-In many cases, the SEM synthetics are calculated and compared to data
-seismograms recorded at seismic stations. Since the SEM synthetics
-are accurate for a certain frequency range, both the original data
-and the synthetics need to be processed before a comparison can be
-made.
-
-For such comparisons, the following steps are recommended:
-\begin{enumerate}
-\item Make sure that both synthetic and observed seismograms have the correct
-station/event and timing information.
-\item Convolve synthetic seismograms with a source time function with the
-half duration specified in the \texttt{CMTSOLUTION} file, provided,
-as recommended, you used a zero half duration in the SEM simulations.
-\item Resample both observed and synthetic seismograms to a common sampling
-rate.
-\item Cut the records using the same window.
-\item Remove the trend and mean from the records and taper them.
-\item Remove the instrument response from the observed seismograms (recommended)
-or convolve the synthetic seismograms with the instrument response.
-\item Make sure that you apply the same filters to both observed and synthetic
-seismograms. Preferably, avoid filtering your records more than once.
-\item Now, you are ready to compare your synthetic and observed seismograms.
-\end{enumerate}
-\noindent We generally use the following scripts provided in the \texttt{utils/seis\_process/}
-directory:
-
-\subsection{Data processing script \texttt{process\_data.pl}}
-
-This script cuts a given portion of the original data, filters it,
-transfers the data into a displacement record, and picks the first
-P and S arrivals. For more functionality, type `\texttt{process\_data.pl}'
-without any argument. An example of the usage of the script:
-\begin{lyxcode}
-{\footnotesize process\_data.pl~-m~CMTSOLUTION~-s~1.0~-l~0/4000~-i~-f~-t~40/500~-p~-x~bp~DATA/1999.330{*}.BH?.SAC}~
-\end{lyxcode}
-\noindent which has resampled the SAC files to a sampling rate of
-1 seconds, cut them between 0 and 4000 seconds, transfered them into
-displacement records and filtered them between 40 and 500 seconds,
-picked the first P and S arrivals, and added suffix `\texttt{bp}'
-to the file names.
-
-Note that all of the scripts in this section actually use the SAC
-and/or IASP91 to do the core operations; therefore make sure that
-the SAC and IASP91 packages are installed properly on your system,
-and that all the environment variables are set properly before running
-these scripts.
-
-\subsection{Synthetics processing script \texttt{process\_syn.pl}}
-
-This script converts the synthetic output from the SEM code from ASCII
-to SAC format, and performs similar operations as `\texttt{process\_data.pl}'.
-An example of the usage of the script:
-\begin{lyxcode}
-{\footnotesize process\_syn.pl~-m~CMTSOLUTION~-h~-a~STATIONS~-s~1.0~-l~0/4000~-f~-t~40/500~-p~-x~bp~SEM/{*}.BX?.semd}~
-\end{lyxcode}
-which will convolve the synthetics with a triangular source-time function
-from the \texttt{CMTSOLUTION} file, convert the synthetics into SAC
-format, add event and station information into the SAC headers, resample
-the SAC files with a sampling rate of 1 seconds, cut them between
-0 and 4000 seconds, filter them between 40 and 500 seconds with the
-same filter used for the observed data, pick the first P and S arrivals,
-and add the suffix `\texttt{bp}' to the file names.
-
-More options are available for this script, such as adding time shift
-to the origin time of the synthetics, convolving the synthetics with
-a triangular source time function with a given half duration, etc.
-Type \texttt{process\_syn.pl} without any argument for a detailed
-usage.
-
-In order to convert between SAC format and ASCII files, useful scripts
-are provided in the subdirectories ~\\
- \texttt{utils/sac2000\_alpha\_convert/} and \texttt{utils/seis\_process/asc2sac/}.
-
-\subsection{Script \texttt{rotate.pl}}
-
-The original data and synthetics have three components: vertical (BHZ
-resp. BXZ), north (BHN resp. BXN) and east (BHE resp. BXE). However,
-for most seismology applications, transverse and radial components
-are also desirable. Therefore, we need to rotate the horizontal components
-of both the data and the synthetics to the transverse and radial direction,
-and \texttt{\small rotate.pl}{\small{} can be used to accomplish this:}{\small \par}
-\begin{lyxcode}
-rotate.pl~-l~0~-L~180~-d~DATA/{*}.BHE.SAC.bp~
-
-rotate.pl~-l~0~-L~180~SEM/{*}.BXE.semd.sac.bp~
-\end{lyxcode}
-where the first command performs rotation on the SAC data obtained
-through Seismogram Transfer Program (STP) \urlwithparentheses{http://www.data.scec.org/STP/stp.html},
-while the second command rotates the processed SEM synthetics.
-
-\section{Collect Synthetic Seismograms}
-
-The forward and adjoint simulations generate synthetic seismograms
-in the \texttt{OUTPUT\_FILES/} directory by default. For the forward
-simulation, the files are named like \texttt{NT.STA.BX?.semd} for
-two-column time series, or \texttt{NT.STA.BX?.semd.sac} for ASCII
-SAC format, where NT and STA are the network code and station name,
-and \texttt{BX?} stands for the component name. Please see the Appendix~\ref{cha:Coordinates}
-and \ref{cha:channel-codes} for further details.
-
-The adjont simulations generate synthetic seismograms with the name
-\texttt{NT.S?????.S??.sem} (refer to Section \ref{sec:Adjoint-simulation-sources}
-for details). The kernel simulations output the back-reconstructed
-synthetic seismogram in the name \texttt{NT.STA.BX?.semd}, mainly
-for the purpose of checking the accuracy of the reconstruction. Refer
-to Section \ref{sec:Adjoint-simulation-finite} for further details.
-
-You do have further options to change this default output behavior,
-given in the main constants file \texttt{constants.h} located in \texttt{src/shared/}
-directory:
-\begin{description}
-\item [{\texttt{SEISMOGRAMS\_BINARY}}] set to \texttt{.true.} to have seismograms
-written out in binary format.
-\item [{\texttt{WRITE\_SEISMOGRAMS\_BY\_MASTER}}] Set to \texttt{.true.}
-to have only the master process writing out seismograms. This can
-be useful on a cluster, where only the master process node has access
-to the output directory.
-\item [{\texttt{USE\_OUTPUT\_FILES\_PATH}}] Set to \texttt{.false.} to
-have the seismograms output to \texttt{LOCAL\_PATH} directory specified
-in the main parameter file \texttt{DATA/Par\_file}. In this case,
-you could collect the synthetics onto the frontend using the \texttt{collect\_seismo\_lsf\_multi.pl}
-script located in the \texttt{utils/Cluster/lsf/} directory. The usage
-of the script would be e.g.:
-
-\begin{lyxcode}
-collect\_seismo.pl~machines~DATA/Par\_file
-\end{lyxcode}
-\end{description}
-where \texttt{machines} is a file containing the node names and \texttt{DATA/Par\_file}
-the parameter file used to extract the \texttt{LOCAL\_PATH} directory
-used for the simulation.
-
-\section{Clean Local Database}
-
-After all the simulations are done, the seismograms are collected,
-and the useful database files are copied to the frontend, you may
-need to clean the local scratch disk for the next simulation. This
-is especially important in the case of kernel simulation, where very
-large files are generated for the absorbing boundaries to help with
-the reconstruction of the regular forward wavefield. A sample script
-is provided in \texttt{utils/}:
-\begin{lyxcode}
-cleanbase.pl~machines
-\end{lyxcode}
-where \texttt{machines} is a file containing the node names.
-
-\section{Plot Movie Snapshots and Synthetic Shakemaps}
-
-\subsection{Script \texttt{movie2gif.gmt.pl}}
-
-With the movie data saved in \texttt{OUTPUT\_FILES/} at the end of
-a movie simulation (\texttt{\small MOVIE\_SURFACE=.true.}{\small ),
-you can run the }\texttt{\small `create\_movie\_shakemap\_AVS\_DX\_GMT}{\small '
-code to convert these binary movie data into GMT xyz files for futher
-processing. A sample script }\texttt{\small movie2gif.gmt.pl}{\small{}
-is provided to do this conversion, and then plot the movie snapshots
-in GMT, for example:}{\small \par}
-\begin{lyxcode}
-movie2gif.gmt.pl~-m~CMTSOLUTION~-g~-f~1/40~-n~-2~-p~
-\end{lyxcode}
-which for the first through the 40th movie frame, converts the \texttt{moviedata}
-files into GMT xyz files, interpolates them using the 'nearneighbor'
-command in GMT, and plots them on a 2D topography map. Note that `\texttt{-2}'
-and `\texttt{-p}' are both optional.
-
-\subsection{Script \texttt{plot\_shakemap.gmt.pl}}
-
-With the shakemap data saved in \texttt{OUTPUT\_FILES/} at the end
-of a shakemap simulation ~\\
- (\texttt{CREATE\_SHAKEMAP=.true.}), you can also run \texttt{`create\_movie\_shakemap\_AVS\_DX\_GMT}'
-code to convert the binary shakemap data into GMT xyz files. A sample
-script \texttt{plot\_shakemap.gmt.pl} is provided to do this conversion,
-and then plot the shakemaps in GMT, for example:
-\begin{lyxcode}
-plot\_shakemap.gmt.pl~data\_dir~type(1,2,3)~CMTSOLUTION~
-\end{lyxcode}
-where \texttt{type=1} for a displacement shakemap, \texttt{2} for
-velocity, and \texttt{3} for acceleration.
-
-\section{Map Local Database}
-
-A sample program \texttt{remap\_database} is provided to map the local
-database from a set of machines to another set of machines. This is
-especially useful when you want to run mesher and solver, or different
-types of solvers separately through a scheduler (refer to Chapter~\ref{cha:Scheduler}).
-\begin{lyxcode}
-run\_lsf.bash~-{}-gm-no-shmem~-{}-gm-copy-env~remap\_database~old\_machines~150
-\end{lyxcode}
-where \texttt{old\_machines} is the LSF machine file used in the previous
-simulation, and \texttt{150} is the number of processors in total.
-
-%============================================================
-
-\chapter{\label{cha:developers}Information for developers of the code, and for people who want to learn how the technique works}
-
-You can get a very simple 1D version of a demo code (there is one in Fortran and one in Python):\\
-
-git clone --recursive https://github.com/geodynamics/specfem1d.git\\
-
-\noindent We also have simple 3D demo source codes that implement the SEM in a single, small program, in directory\\
-utils/small\_SEM\_solvers\_in\_Fortran\_and\_C\_without\_MPI\_to\_learn of the specfem3d package.
-They are useful to learn how the spectral-element method works, and how to write or modify a code to implement it.
-Also useful to test new ideas by modifying these simple codes to run some tests.\\
-
-\noindent For information on how to contribute to the code, i.e., for how to make your modifications, additions or improvements part of the
-official package, see https://github.com/geodynamics/specfem3d/wiki .
-
-%============================================================
-
-\chapter*{\label{cha:Bug-Reports-and}Bug Reports and Suggestions for Improvements}
-
-To report bugs or suggest improvements to the code, please send an
-e-mail to the CIG Computational Seismology Mailing List \urlwithparentheses{cig-seismo@geodynamics.org}.
-
-\addcontentsline{toc}{chapter}{Notes \& Acknowledgments}
-
-\chapter*{Notes \& Acknowledgments}
-
-In order to keep the software package thread-safe in case a multithreaded
-implementation of MPI is used, developers should not add modules or
-common blocks to the source code but rather use regular subroutine
-arguments (which can be grouped in ``derived types'' if needed for
-clarity).
-
-The Gauss-Lobatto-Legendre subroutines in \texttt{gll\_library.f90}
-are based in part on software libraries from the Massachusetts Institute
-of Technology, Department of Mechanical Engineering (Cambridge, Massachusetts,
-USA). The non-structured global numbering software was provided by
-Paul F. Fischer (Brown University, Providence, Rhode Island, USA,
-now at Argonne National Laboratory, USA).
-
-OpenDX \urlwithparentheses{http://www.opendx.org} is open-source
-based on IBM Data Explorer, AVS \urlwithparentheses{http://www.avs.com}
-is a trademark of Advanced Visualization Systems, and ParaView \urlwithparentheses{http://www.paraview.org}
-is an open-source visualization platform.
-
-Please e-mail your feedback, questions, comments, and suggestions
-to the CIG Computational Seismology Mailing List \urlwithparentheses{cig-seismo@geodynamics.org}.
-
-\addcontentsline{toc}{chapter}{Copyright}
-
-\chapter*{Copyright}
-
-Main historical authors: Dimitri Komatitsch and Jeroen Tromp
-
-Princeton University, USA, and CNRS / University of Marseille, France
-
-$\copyright$ Princeton University and CNRS / University of Marseille, July 2012
-
-This program is free software; you can redistribute it and/or modify
-it under the terms of the GNU General Public License as published
-by the Free Software Foundation (see Appendix \ref{cha:License}).\\
-
-\textbf{\underline{Evolution of the code:}}\\
-
-MPI v. 3.0, December 2014: many developers.
-Convolutional PML, LDDRK time scheme, bulk attenuation support, simultaneous MPI runs,
-ADIOS file I/O support, coupling with external codes, new seismogram names,
-Deville routines for additional GLL degrees, tomography tools, unit/regression test framework,
-improved CUDA GPUs performance, additonal GEOCUBIT support, better make compilation,
-git versioning system. \\
-
-MPI v. 2.1, July 2012: Max Rietmann, Peter Messmer, Daniel Peter, Dimitri
-Komatitsch, Joseph Charles, Zhinan Xie: support for CUDA GPUs, better
-CFL stability for the Stacey absorbing conditions. \\
-
-MPI v. 2.0, November 2010: Dimitri Komatitsch, Nicolas Le Goff, Roland
-Martin and Pieyre Le Loher, University of Pau, France, Daniel Peter,
-Jeroen Tromp and the Princeton group of developers, Princeton University,
-USA, and Emanuele Casarotti, INGV Roma, Italy: support for CUBIT meshes
-decomposed by SCOTCH; much faster solver using Michel Deville's inlined
-matrix products.\\
-
-MPI v. 1.4 Dimitri Komatitsch, University of Pau, Qinya Liu and others,
-Caltech, September 2006: better adjoint and kernel calculations, faster
-and better I/Os on very large systems, many small improvements and
-bug fixes.\\
-
-MPI v. 1.3 Dimitri Komatitsch, University of Pau, and Qinya Liu, Caltech,
-July 2005: serial version, regular mesh, adjoint and kernel calculations,
-ParaView support.\\
-
-MPI v. 1.2 Min Chen and Dimitri Komatitsch, Caltech, July 2004: full
-anisotropy, volume movie.\\
-
-MPI v. 1.1 Dimitri Komatitsch, Caltech, October 2002: Zhu's Moho map,
-scaling of $V_{s}$ with depth, Hauksson's regional model, attenuation,
-oceans, movies.\\
-
-MPI v. 1.0 Dimitri Komatitsch, Caltech, USA, May 2002: first MPI version
-based on global code.\\
-
-Dimitri Komatitsch, IPG Paris, France, December 1996: first 3-D solver
-for the CM-5 Connection Machine, parallelized on 128 processors using
-Connection Machine Fortran.\\
-
-\addcontentsline{toc}{chapter}{References} \bibliography{bibliography}
-
-\appendix
-%dummy comment inserted by tex2lyx to ensure that this paragraph is not empty
-
-\chapter{\label{cha:Coordinates}Reference Frame Convention}
-
-The code uses the following convention for the Cartesian reference
-frame:
-\begin{itemize}
-\item the $x$ axis points East
-\item the $y$ axis points North
-\item the $z$ axis points up
-\end{itemize}
-Note that this convention is different from both the \citet{AkRi80}
-convention and the Harvard Centroid-Moment Tensor (CMT) convention.
-The Aki \& Richards convention is
-\begin{itemize}
-\item the $x$ axis points North
-\item the $y$ axis points East
-\item the $z$ axis points down
-\end{itemize}
-and the Harvard CMT convention is
-\begin{itemize}
-\item the $x$ axis points South
-\item the $y$ axis points East
-\item the $z$ axis points up
-\end{itemize}
-
-\subsection*{Source and receiver locations}
-
-The SPECFEM3D Cartesian software code internally uses Cartesian coordinates.
-The given locations for sources and receiver locations thus may get
-converted. Sources and receiver locations are read in from the \texttt{CMTSOLUTION}
-(or \texttt{FORCESOLUTION}) and \texttt{STATIONS} files. Note that
-e.g. the \texttt{CMTSOLUTION} and \texttt{FORCESOLUTION} files denotes
-the location by \textquotedbl{}longitude/latitude/depth\textquotedbl{}.
-We read in longitude as $x$ coordinate, latitude as $y$ coordinate.
-
-In case the flag \texttt{SUPPRESS\_UTM\_PROJECTION} is set to \texttt{.false.}
-in the main parameter file (see Chapter~\ref{cha:Creating-Distributed-Databases}),
-the $x$/$y$ coordinates have to be given in degrees and are converted
-to Cartesian coordinates using a UTM conversion with the specified
-UTM zone.
-
-The value for depth (given in $km$ in \texttt{CMTSOLUTION} and \texttt{FORCESOLUTION})
-or burial depth (given in $m$ in \texttt{STATIONS}) is evaluated
-with respect to the surface of the mesh at the specified $x$/$y$
-location to find the corresponding $z$ coordinate. It is possible
-to use this depth value directly as $z$ coordinate by changing the
-flag \texttt{USE\_SOURCES\_RECVS\_Z} to \texttt{.true.} in the file
-\texttt{constants.h} located in the \texttt{src/shared/} subdirectory.
-
-\subsection*{Seismogram outputs}
-
-The seismogram output directions are given in Cartesian $x$/$y$/$z$
-directions and not rotated any further. Changing flags in \texttt{constants.h}
-in the \texttt{src/shared/} subdirectory only rotates the seismogram
-outputs if receivers are forced to be located at the surface (\texttt{RECVS\_CAN\_BE\_BURIED\_EXT\_MESH}
-set to \texttt{.false.}) and the normal to the surface at the receiver
-location should be used (\texttt{EXT\_MESH\_RECV\_NORMAL} set to \texttt{.true.})
-as vertical. In this case, the outputs are rotated to have the vertical
-component normal to the surface of the mesh, $x$ and $y$ directions
-are somewhat arbitrary orthogonal directions along the surface.
-
-For the labeling of the seismogram channels, see Appendix~\ref{cha:channel-codes}.
-Additionally, we add labels to the synthetics using the following
-convention: For a receiver station located in an
-\begin{description}
-\item [{elastic domain:}] ~\\
-
-\begin{itemize}
-\item \texttt{semd} for the displacement vector
-\item \texttt{semv} for the velocity vector
-\item \texttt{sema} for the acceleration vector
-\end{itemize}
-\item [{acoustic domain:}] ~\\
- (please note that receiver stations in acoustic domains must be buried.
-This is due to the free surface condition which enforces a zero pressure
-at the surface.)
-
-\begin{itemize}
-\item \texttt{semd} for the displacement vector
-\item \texttt{semv} for the velocity vector
-\item \texttt{sema} for pressure which will be stored in the vertical component
-\texttt{Z}. Note that pressure in the acoustic media is isotropic
-and thus the pressure record would be the same in the other two component
-directions. We therefore use the other two seismogram components to
-store the acoustic potential in component \texttt{X} (or \texttt{N})
-and the first time derivative of the acoustic potential in component
-\texttt{Y} (or \texttt{E}).
-\end{itemize}
-\end{description}
-The seismograms are by default written out in ASCII-format to the
-\texttt{OUTPUT\_FILES/} subdirectory by each parallel process. You
-can change this behavior by changing the following flags in the \texttt{constants.h}
-file located in the \texttt{src/shared/} subdirectory:
-\begin{description}
-\item [{\texttt{SEISMOGRAMS\_BINARY}}] set to \texttt{.true.} to have seismograms
-written out in binary format.
-\item [{\texttt{WRITE\_SEISMOGRAMS\_BY\_MASTER}}] Set to \texttt{.true.}
-to have only the master process writing out seismograms. This can
-be useful on a cluster, where only the master process node has access
-to the output directory.
-\item [{\texttt{USE\_OUTPUT\_FILES\_PATH}}] Set to \texttt{.false.} to
-have the seismograms output to \texttt{LOCAL\_PATH} directory specified
-in the main parameter file \texttt{DATA/Par\_file}.
-\end{description}
-
-\chapter{\label{cha:channel-codes}Channel Codes of Seismograms}
-
-Seismic networks, such as the Global Seismographic Network (GSN),
-generally involve various types of instruments with different bandwidths,
-sampling properties and component configurations. There are standards
-to name channel codes depending on instrument properties. IRIS \texttt{(www.iris.edu)}
-uses SEED format for channel codes, which are represented by three
-letters, such as \texttt{LHN}, \texttt{BHZ}, etc. In older versions
-of the SPECFEM3D Cartesian package, a common format was used for the
-channel codes of all seismograms, which was \texttt{BHE/BHN/BHZ} for
-three components. To avoid confusion when comparison are made to observed
-data, we are now using the FDSN convention (\url{http://www.fdsn.org/})
-for SEM seismograms. In the following, we give a brief explanation
-of the FDSN convention used by IRIS, and how it is adopted in SEM
-seismograms. Please visit \texttt{www.iris.edu/manuals/SEED\_appA.htm}
-for further information.\\
-
-\noindent \texttt{\textbf{Band code:}} The first letter in the channel
-code denotes the band code of seismograms, which depends on the response
-band and the sampling rate of instruments. The list of band codes
-used by IRIS is shown in Figure \ref{fig:IRIS_band_codes}. The sampling
-rate of SEM synthetics is controlled by the resolution of simulations
-rather than instrument properties. However, for consistency, we follow
-the FDSN convention for SEM seismograms governed by their sampling
-rate. For SEM synthetics, we consider band codes for which $dt\leq1$
-s. IRIS also considers the response band of instruments. For instance,
-short-period and broad-band seismograms with the same sampling rate
-correspond to different band codes, such as S and B, respectively.
-In such cases, we consider SEM seismograms as broad band, ignoring
-the corner period ($\geq10$ s) of the response band of instruments
-(note that at these resolutions, the minimum period in the SEM synthetics
-will be less than $10$ s). Accordingly, when you run a simulation
-the band code will be chosen depending on the resolution of the synthetics,
-and channel codes of SEM seismograms will start with either \texttt{L},
-\texttt{M}, \texttt{B}, \texttt{H}, \texttt{C} or \texttt{F}, shown
-by red color in the figure.\\
-
-\begin{figure}[ht]
-\noindent \begin{centering}
-\includegraphics[scale=0.6]{figures/IRIS_band_codes}
-\par\end{centering}
-
-\caption{The band code convention is based on the sampling rate and the response
-band of instruments. Please visit \texttt{www.iris.edu/manuals/SEED\_appA.htm}
-for further information. Grey rows show the relative band-code range
-in SPECFEM3D Cartesian, and the band codes used to name SEM seismograms
-are denoted in red.}
-
-\label{fig:IRIS_band_codes}
-\end{figure}
-
-\noindent \texttt{\textbf{Instrument code:}} The second letter in
-the channel code corresponds to instrument codes, which specify the
-family to which the sensor belongs. For instance, \texttt{H} and \texttt{L}
-are used for high-gain and low-gain seismometers, respectively. The
-instrument code of SEM seismograms will always be \texttt{X}, as assigned
-by FDSN for synthetic seismograms. \\
-
-\noindent \texttt{\textbf{Orientation code:}} The third letter in
-channel codes is an orientation code, which generally describes the
-physical configuration of the components of instrument packages. SPECFEM3D
-Cartesian uses the traditional orientation code \texttt{E/N/Z} (East-West,
-North-South, Vertical) for three components when a UTM projection
-is used. If the UTM conversion is suppressed, i.e. the flag \texttt{SUPPRESS\_UTM\_PROJECTION}
-is set to \texttt{.true.}, then the three components are labelled
-\texttt{X/Y/Z} according to the Cartesian reference frame. \\
-
-\noindent \texttt{\textbf{EXAMPLE:}} The sampling rate is given by
-\texttt{DT} in the main parameter file \texttt{Par\_file} located
-in the \texttt{DATA/} subdirectory. Depending on the resolution of
-your simulations, if you choose a sampling rate greater than $0.01$
-s and less than $1$ s, a seismogram recording displacements on the
-vertical component of a station \texttt{ASBS} (network \texttt{AZ})
-will be named \texttt{AZ.ASBS.MXZ.semd.sac}, whereas it will be \texttt{AZ.ASBS.BXZ.semd.sac},
-if the sampling rate is greater than 0.0125 and less equal to 0.1
-s.
-
-\chapter{\label{cha:Troubleshooting}Troubleshooting}
-
-\section*{FAQ}
-\begin{description}
-\item [{configuration fails:}]~
-
-Examine the log file 'config.log'. It contains detailed informations.
-In many cases, the paths to these specific compiler commands F90,
-CC and MPIF90 will not be correct if `./configure` fails.
-
-Please make sure that you have a working installation of a Fortran
-compiler, a C compiler and an MPI implementation. You should be able
-to compile this little program code:
-\begin{lyxcode}
-program~main~~~\\
-~include~'mpif.h'~~~\\
-~integer,~parameter~::~CUSTOM\_MPI\_TYPE~=~MPI\_REAL~~~\\
-~integer~ier~~~\\
-~call~MPI\_INIT(ier)~~~\\
-~call~MPI\_BARRIER(MPI\_COMM\_WORLD,ier)~~~\\
-~call~MPI\_FINALIZE(ier)~~~\\
-~end
-\end{lyxcode}
-\item [{compilation fails stating:}] ~\\
-
-\begin{lyxcode}
-...~~~\\
-~obj/program\_generate\_databases.o:~In~function~`MAIN\_\_':~~~\\
-~program\_generate\_databases.f90:(.text+0x14):~undefined~reference~to~`\_gfortran\_set\_std'~~~\\
-~...~~~\\
-
-\end{lyxcode}
-
-Make sure you're pointing to the right 'mpif90' wrapper command.
-
-Normally, this message will appear when you are mixing two different
-Fortran compilers. That is, using e.g. gfortran to compile non-MPI
-files and mpif90, wrapper provided for e.g. ifort, to compile MPI-files.
-
-fix: e.g. specify > ./configure FC=gfortran MPIF90=/usr/local/openmpi-gfortran/bin/mpif90
-
-\item [{after executing \texttt{xmeshfem3D} I've got elements with skewness of 81\% percent, what does this mean:}] Look
-at the skewness table printed in the \texttt{output\_mesher.txt} file
-after executing \texttt{xmeshfem3D} for the example given in \texttt{EXAMPLES/meshfem3D\_examples/simple\_model/}:
-
-\begin{lyxcode}
-...~~\\
-~histogram~of~skewness~(0.~good~-~1.~bad):~~~\\
-~0.0000000E+00~-~5.0000001E-02~27648~81.81818~\%~~~\\
-~5.0000001E-02~-~0.1000000~0~0.0000000E+00~\%~~~\\
-~...~~~\\
-
-\end{lyxcode}
-
-The first line means that you have 27,648 elements with a skewness
-value between 0 and 0.05 (which means the element is basically not
-skewed, just plain regular hexahedral element). The total number of
-elements you have in this mesh is (see in the \texttt{output\_mesher.txt}
-file a bit further down):
-\begin{lyxcode}
-...~~\\
-~total~number~of~elements~in~entire~mesh:~33792~~~\\
-~...~~~\\
-
-\end{lyxcode}
-
-which gives you that: 27,648 / 33,792 $\sim$ 81.8 \% of all elements
-are not skewed, i.e. regular elements. a fantastic value :)
-
-The histogram lists for this mesh also some stronger skewed elements,
-for example the worst ones belong to:
-\begin{lyxcode}
-...~~\\
-~0.6000000~-~0.6500000~2048~6.060606~\%~~~\\
-~...~~~\\
-
-\end{lyxcode}
-
-about 6 \% of all elements have distortions with a skewness value
-between 0.6 and 0.65. The skewness values give you a hint of how good
-your mesh is. In an ideal world, you would want to have no distortions,
-just like the 81\% from above. Those elements give you the best approximate
-values by the GLL quadrature used in the spectral-element method.
-However, having weakly distorted elements is still fine and the solutions
-are still accurate enough. So empirically, values up to around 0.7
-are tolerable, above that you should consider remeshing...
-
-To give you an idea why some of the elements are distorted, see the
-following figure \ref{fig:mesh.vp} of the mesh you obtain in the
-example \texttt{EXAMPLES/meshfem3D\_examples/simple\_model/}.
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=0.9\textwidth]{figures/mesh_vp}
-\par\end{centering}
-
-\caption{Paraview visualization using the mesh vtk-files for the example given
-in \texttt{EXAMPLES/meshfem3D\_examples/simple\_model/}.}
-
-\label{fig:mesh.vp}
-\end{figure}
-
-You will see that the mesh contains a doubling layer, where we stitch
-elements together such that the size of two elements will transition
-to the size of one element (very useful to keep the ratio of wavespeed
-/ element\_size about constant). Those elements in this doubling layer
-have higher skewness values and make up those 6 \% in the histogram.
-
-\item [{the code gives following error message "need at least one receiver":}] This
-means that no stations given in the input file \texttt{DATA/STATIONS}
-could be located within the dimensions of the mesh. This can happen
-for example when the mesh was created with the in-house mesher \texttt{xmeshfem3D}
-while using the Universal Transverse Mercator (UTM) projection but
-the simulation with \texttt{xspecfem3D} was suppressing this projection
-from latitude/longitude to x/y/z coordinates.
-
-In such cases, try to change your \texttt{DATA/Par\_file} and set
-e.g.:
-\begin{lyxcode}
-SUPPRESS\_UTM\_PROJECTION~=~.false.~~~\\
-
-\end{lyxcode}
-
-to be the same in \texttt{Mesh\_Par\_file} and \texttt{Par\_file}.
-This flag should be identical when using the in-house mesher \texttt{xmeshfem3D},
-\texttt{xgenerate\_databases} and \texttt{xspecfem3D} together to
-run simulations.
-
-The flag determines if the coordinates you specify for your source
-and station locations are given as lat/lon degrees and must be converted
-to UTM coordinates. As an example, if you use \texttt{.false.} within
-\texttt{Mesh\_Par\_file} then you create a mesh with \texttt{xmeshfem3D}
-using the UTM projection from lat/lon as input format to UTM projected
-coordinates to store the mesh point positions, which is fine. The
-error then may occur if in the \texttt{Par\_file} you have this set
-to \texttt{.true.} so that the \texttt{xgenerate\_databases} and \texttt{xspecfem3D}
-suppress the UTM projection and assume that all coordinates you use
-now for source and receiver locations are given in meters (that is,
-converted) already. So it won't find the specified locations in the
-used mesh. As a solutions, just change the flag in \texttt{Par\_file}
-to be the same as in \texttt{Mesh\_Par\_file} and rerun \texttt{xgenerate\_databases}
-and \texttt{xspecfem3D} to make sure that your simulation works fine.
-
-\item [{I get the following error message "forward simulation became unstable and blew up":}] In
-most cases this means that your time step size \texttt{DT} is chosen
-too big. Look at your files \texttt{output\_mesher.txt} or \texttt{output\_solver.txt}
-created in the folder \texttt{OUTPUT\_FILES}. In these output files,
-find the section:
-
-\begin{lyxcode}
-...~~~\\
-~{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}~~~\\
-~{*}{*}{*}~Verification~of~simulation~parameters~{*}{*}{*}~~~\\
-~{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}~~~\\
-~...~~~\\
-~{*}{*}{*}~Minimum~period~resolved~=~4.308774~~~\\
-~{*}{*}{*}~Maximum~suggested~time~step~=~6.8863556E-02~~~\\
-~...~~~\\
-~...~~~\\
-
-\end{lyxcode}
-
-then change \texttt{DT} in the \texttt{DATA/Par\_file} to be somewhere
-close to the maximum suggested time step. In the example above: \\
- \texttt{DT = 0.05d0} \\
- would (most probably) work fine. It could be also bigger than the
-0.068 s suggested. This depends a bit on the distortions of your mesh
-elements. The more regular they are, the bigger you can choose \texttt{DT}.
-Just play with this value a bit and see when the simulation becomes
-stable ...
-
-\end{description}
-
-\chapter{\label{cha:License}License }
-
-\textbf{GNU GENERAL PUBLIC LICENSE Version 2, June 1991. Copyright
-(C) 1989, 1991 Free Software Foundation, Inc. 59 Temple Place, Suite
-330, Boston, MA 02111-1307 USA} \\
- Everyone is permitted to copy and distribute verbatim copies of this
-license document, but changing it is not allowed.
-
-\section*{Preamble}
-
-The licenses for most software are designed to take away your freedom
-to share and change it. By contrast, the GNU General Public License
-is intended to guarantee your freedom to share and change free software
--- to make sure the software is free for all its users. This General
-Public License applies to most of the Free Software Foundation's software
-and to any other program whose authors commit to using it. (Some other
-Free Software Foundation software is covered by the GNU Library General
-Public License instead.) You can apply it to your programs, too.
-
-When we speak of free software, we are referring to freedom, not price.
-Our General Public Licenses are designed to make sure that you have
-the freedom to distribute copies of free software (and charge for
-this service if you wish), that you receive source code or can get
-it if you want it, that you can change the software or use pieces
-of it in new free programs; and that you know you can do these things.
-
-To protect your rights, we need to make restrictions that forbid anyone
-to deny you these rights or to ask you to surrender the rights. These
-restrictions translate to certain responsibilities for you if you
-distribute copies of the software, or if you modify it.
-
-For example, if you distribute copies of such a program, whether gratis
-or for a fee, you must give the recipients all the rights that you
-have. You must make sure that they, too, receive or can get the source
-code. And you must show them these terms so they know their rights.
-
-We protect your rights with two steps:
-\begin{enumerate}
-\item Copyright the software, and
-\item Offer you this license which gives you legal permission to copy, distribute
-and/or modify the software.
-\end{enumerate}
-Also, for each author's protection and ours, we want to make certain
-that everyone understands that there is no warranty for this free
-software. If the software is modified by someone else and passed on,
-we want its recipients to know that what they have is not the original,
-so that any problems introduced by others will not reflect on the
-original authors' reputations.
-
-Finally, any free program is threatened constantly by software patents.
-We wish to avoid the danger that redistributors of a free program
-will individually obtain patent licenses, in effect making the program
-proprietary. To prevent this, we have made it clear that any patent
-must be licensed for everyone's free use or not licensed at all.
-
-The precise terms and conditions for copying, distribution and modification
-follow.
-
-\section*{GNU GENERAL PUBLIC LICENSE TERMS AND CONDITIONS FOR COPYING, DISTRIBUTION
-AND MODIFICATION }
-\begin{itemize}
-\item 0.This License applies to any program or other work which contains
-a notice placed by the copyright holder saying it may be distributed
-under the terms of this General Public License. The ``Program''
-below refers to any such program or work, and a ``work based on the
-Program'' means either the Program or any derivative work under copyright
-law: that is to say, a work containing the Program or a portion of
-it, either verbatim or with modifications and/or translated into another
-language. (Hereinafter, translation is included without limitation
-in the term ``modification.'') Each licensee is addressed as ``you.''\\
- \\
- Activities other than copying, distribution and modification are
-not covered by this License; they are outside its scope. The act of
-running the Program is not restricted, and the output from the Program
-is covered only if its contents constitute a work based on the Program
-(independent of having been made by running the Program). Whether
-that is true depends on what the Program does.\end{itemize}
-\begin{enumerate}
-\item You may copy and distribute verbatim copies of the Program's source
-code as you receive it, in any medium, provided that you conspicuously
-and appropriately publish on each copy an appropriate copyright notice
-and disclaimer of warranty; keep intact all the notices that refer
-to this License and to the absence of any warranty; and give any other
-recipients of the Program a copy of this License along with the Program.
-
-You may charge a fee for the physical act of transferring a copy,
-and you may at your option offer warranty protection in exchange for
-a fee.
-
-\item You may modify your copy or copies of the Program or any portion of
-it, thus forming a work based on the Program, and copy and distribute
-such modifications or work under the terms of Section 1 above, provided
-that you also meet all of these conditions:
-
-\begin{enumerate}
-\item You must cause the modified files to carry prominent notices stating
-that you changed the files and the date of any change.
-\item You must cause any work that you distribute or publish, that in whole
-or in part contains or is derived from the Program or any part thereof,
-to be licensed as a whole at no charge to all third parties under
-the terms of this License.
-\item If the modified program normally reads commands interactively when
-run, you must cause it, when started running for such interactive
-use in the most ordinary way, to print or display an announcement
-including an appropriate copyright notice and a notice that there
-is no warranty (or else, saying that you provide a warranty) and that
-users may redistribute the program under these conditions, and telling
-the user how to view a copy of this License. (Exception: if the Program
-itself is interactive but does not normally print such an announcement,
-your work based on the Program is not required to print an announcement.)
-\end{enumerate}
-
-These requirements apply to the modified work as a whole. If identifiable
-sections of that work are not derived from the Program, and can be
-reasonably considered independent and separate works in themselves,
-then this License, and its terms, do not apply to those sections when
-you distribute them as separate works. But when you distribute the
-same sections as part of a whole which is a work based on the Program,
-the distribution of the whole must be on the terms of this License,
-whose permissions for other licensees extend to the entire whole,
-and thus to each and every part regardless of who wrote it.
-
-Thus, it is not the intent of this section to claim rights or contest
-your rights to work written entirely by you; rather, the intent is
-to exercise the right to control the distribution of derivative or
-collective works based on the Program.
-
-In addition, mere aggregation of another work not based on the Program
-with the Program (or with a work based on the Program) on a volume
-of a storage or distribution medium does not bring the other work
-under the scope of this License.
-
-\item You may copy and distribute the Program (or a work based on it, under
-Section 2) in object code or executable form under the terms of Sections
-1 and 2 above provided that you also do one of the following:
-
-\begin{enumerate}
-\item Accompany it with the complete corresponding machine-readable source
-code, which must be distributed under the terms of Sections 1 and
-2 above on a medium customarily used for software interchange; or,
-\item Accompany it with a written offer, valid for at least three years,
-to give any third party, for a charge no more than your cost of physically
-performing source distribution, a complete machine-readable copy of
-the corresponding source code, to be distributed under the terms of
-Sections 1 and 2 above on a medium customarily used for software interchange;
-or,
-\item Accompany it with the information you received as to the offer to
-distribute corresponding source code. (This alternative is allowed
-only for noncommercial distribution and only if you received the program
-in object code or executable form with such an offer, in accord with
-Subsection b above.)
-\end{enumerate}
-
-The source code for a work means the preferred form of the work for
-making modifications to it. For an executable work, complete source
-code means all the source code for all modules it contains, plus any
-associated interface definition files, plus the scripts used to control
-compilation and installation of the executable. However, as a special
-exception, the source code distributed need not include anything that
-is normally distributed (in either source or binary form) with the
-major components (compiler, kernel, and so on) of the operating system
-on which the executable runs, unless that component itself accompanies
-the executable.
-
-If distribution of executable or object code is made by offering access
-to copy from a designated place, then offering equivalent access to
-copy the source code from the same place counts as distribution of
-the source code, even though third parties are not compelled to copy
-the source along with the object code.
-
-\item You may not copy, modify, sublicense, or distribute the Program except
-as expressly provided under this License. Any attempt otherwise to
-copy, modify, sublicense or distribute the Program is void, and will
-automatically terminate your rights under this License. However, parties
-who have received copies, or rights, from you under this License will
-not have their licenses terminated so long as such parties remain
-in full compliance.
-\item You are not required to accept this License, since you have not signed
-it. However, nothing else grants you permission to modify or distribute
-the Program or its derivative works. These actions are prohibited
-by law if you do not accept this License. Therefore, by modifying
-or distributing the Program (or any work based on the Program), you
-indicate your acceptance of this License to do so, and all its terms
-and conditions for copying, distributing or modifying the Program
-or works based on it.
-\item Each time you redistribute the Program (or any work based on the Program),
-the recipient automatically receives a license from the original licensor
-to copy, distribute or modify the Program subject to these terms and
-conditions. You may not impose any further restrictions on the recipients'
-exercise of the rights granted herein. You are not responsible for
-enforcing compliance by third parties to this License.
-\item If, as a consequence of a court judgment or allegation of patent infringement
-or for any other reason (not limited to patent issues), conditions
-are imposed on you (whether by court order, agreement or otherwise)
-that contradict the conditions of this License, they do not excuse
-you from the conditions of this License. If you cannot distribute
-so as to satisfy simultaneously your obligations under this License
-and any other pertinent obligations, then as a consequence you may
-not distribute the Program at all. For example, if a patent license
-would not permit royalty-free redistribution of the Program by all
-those who receive copies directly or indirectly through you, then
-the only way you could satisfy both it and this License would be to
-refrain entirely from distribution of the Program.
-
-If any portion of this section is held invalid or unenforceable under
-any particular circumstance, the balance of the section is intended
-to apply and the section as a whole is intended to apply in other
-circumstances.
-
-It is not the purpose of this section to induce you to infringe any
-patents or other property right claims or to contest validity of any
-such claims; this section has the sole purpose of protecting the integrity
-of the free software distribution system, which is implemented by
-public license practices. Many people have made generous contributions
-to the wide range of software distributed through that system in reliance
-on consistent application of that system; it is up to the author/donor
-to decide if he or she is willing to distribute software through any
-other system and a licensee cannot impose that choice.
-
-This section is intended to make thoroughly clear what is believed
-to be a consequence of the rest of this License.
-
-\item If the distribution and/or use of the Program is restricted in certain
-countries either by patents or by copyrighted interfaces, the original
-copyright holder who places the Program under this License may add
-an explicit geographical distribution limitation excluding those countries,
-so that distribution is permitted only in or among countries not thus
-excluded. In such case, this License incorporates the limitation as
-if written in the body of this License.
-\item The Free Software Foundation may publish revised and/or new versions
-of the General Public License from time to time. Such new versions
-will be similar in spirit to the present version, but may differ in
-detail to address new problems or concerns.
-
-Each version is given a distinguishing version number. If the Program
-specifies a version number of this License which applies to it and
-``any later version,'' you have the option of following the terms
-and conditions either of that version or of any later version published
-by the Free Software Foundation. If the Program does not specify a
-version number of this License, you may choose any version ever published
-by the Free Software Foundation.
-
-\item If you wish to incorporate parts of the Program into other free programs
-whose distribution conditions are different, write to the author to
-ask for permission. For software which is copyrighted by the Free
-Software Foundation, write to the Free Software Foundation; we sometimes
-make exceptions for this. Our decision will be guided by the two goals
-of preserving the free status of all derivatives of our free software
-and of promoting the sharing and reuse of software generally.
-\end{enumerate}
-
-\subsection*{NO WARRANTY }
-\begin{itemize}
-\item 11.BECAUSE THE PROGRAM IS LICENSED FREE OF CHARGE, THERE IS NO WARRANTY
-FOR THE PROGRAM, TO THE EXTENT PERMITTED BY APPLICABLE LAW. EXCEPT
-WHEN OTHERWISE STATED IN WRITING THE COPYRIGHT HOLDERS AND/OR OTHER
-PARTIES PROVIDE THE PROGRAM ``AS IS'' WITHOUT WARRANTY OF ANY KIND,
-EITHER EXPRESSED OR IMPLIED, INCLUDING, BUT NOT LIMITED TO, THE IMPLIED
-WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE.
-THE ENTIRE RISK AS TO THE QUALITY AND PERFORMANCE OF THE PROGRAM IS
-WITH YOU. SHOULD THE PROGRAM PROVE DEFECTIVE, YOU ASSUME THE COST
-OF ALL NECESSARY SERVICING, REPAIR OR CORRECTION.
-\item 12.IN NO EVENT UNLESS REQUIRED BY APPLICABLE LAW OR AGREED TO IN WRITING
-WILL ANY COPYRIGHT HOLDER, OR ANY OTHER PARTY WHO MAY MODIFY AND/OR
-REDISTRIBUTE THE PROGRAM AS PERMITTED ABOVE, BE LIABLE TO YOU FOR
-DAMAGES, INCLUDING ANY GENERAL, SPECIAL, INCIDENTAL OR CONSEQUENTIAL
-DAMAGES ARISING OUT OF THE USE OR INABILITY TO USE THE PROGRAM (INCLUDING
-BUT NOT LIMITED TO LOSS OF DATA OR DATA BEING RENDERED INACCURATE
-OR LOSSES SUSTAINED BY YOU OR THIRD PARTIES OR A FAILURE OF THE PROGRAM
-TO OPERATE WITH ANY OTHER PROGRAMS), EVEN IF SUCH HOLDER OR OTHER
-PARTY HAS BEEN ADVISED OF THE POSSIBILITY OF SUCH DAMAGES.
-\end{itemize}
-
-\section*{END OF TERMS AND CONDITIONS }
-
-\subsection*{How to Apply These Terms to Your New Programs}
-
-If you develop a new program, and you want it to be of the greatest
-possible use to the public, the best way to achieve this is to make
-it free software which everyone can redistribute and change under
-these terms.
-
-To do so, attach the following notices to the program. It is safest
-to attach them to the start of each source file to most effectively
-convey the exclusion of warranty; and each file should have at least
-the ``copyright'' line and a pointer to where the full notice is
-found. For example:
-\begin{quote}
-One line to give the program's name and a brief idea of what it does.
-Copyright {\footnotesize $\copyright$ (}year) (name of author)
-
-This program is free software; you can redistribute it and/or modify
-it under the terms of the GNU General Public License as published
-by the Free Software Foundation; either version 2 of the License,
-or (at your option) any later version.
-
-This program is distributed in the hope that it will be useful, but
-WITHOUT ANY WARRANTY; without even the implied warranty of MERCHANTABILITY
-or FITNESS FOR A PARTICULAR PURPOSE. See the GNU General Public License
-for more details.
-
-You should have received a copy of the GNU General Public License
-along with this program; if not, write to the Free Software Foundation,
-Inc., 59 Temple Place, Suite 330, Boston, MA 02111-1307 USA
-\end{quote}
-Also add information on how to contact you by electronic and paper
-mail.
-
-If the program is interactive, make it output a short notice like
-this when it starts in an interactive mode:
-\begin{quote}
-Gnomovision version 69, Copyright $\copyright$ year name of author
-Gnomovision comes with ABSOLUTELY NO WARRANTY; for details type `show
-w'. This is free software, and you are welcome to redistribute it
-under certain conditions; type `show c' for details.
-\end{quote}
-The hypothetical commands `show w' and `show c' should show the appropriate
-parts of the General Public License. Of course, the commands you use
-may be called something other than `show w' and `show c'; they could
-even be mouse-clicks or menu items -- whatever suits your program.
-
-You should also get your employer (if you work as a programmer) or
-your school, if any, to sign a ``copyright disclaimer'' for the
-program, if necessary. Here is a sample; alter the names:
-\begin{quote}
-Yoyodyne, Inc., hereby disclaims all copyright interest in the program
-`Gnomovision' (which makes passes at compilers) written by James Hacker.
-
-(signature of Ty Coon)\\
- 1 April 1989 \\
- Ty Coon, President of Vice
-\end{quote}
-This General Public License does not permit incorporating your program
-into proprietary programs. If your program is a subroutine library,
-you may consider it more useful to permit linking proprietary applications
-with the library. If this is what you want to do, use the GNU Library
-General Public License instead of this License.
-\end{document}

### 2015-04-24 doc/USER_MANUAL/07_adjoint_simulations.tex
updates user manual to use facilitate the generation of a markdown version to the github wiki

--- 2015-04-18 doc/USER_MANUAL/manual_SPECFEM3D_Cartesian.tex
@@
-% !TEX encoding = IsoLatin
-%% LyX 2.0.2 created this file.  For more info, see http://www.lyx.org/.
-%% Do not edit unless you really know what you are doing.
-\documentclass[oneside,english]{book}
-\usepackage[T1]{fontenc}
-%% DK DK we do not accept accented letters, use the old LaTeX syntax
-%% DK DK for them instead: \'e \`e  \"e \^e and so on.
-%% DK DK \usepackage[latin1]{inputenc}
-\usepackage[letterpaper]{geometry}
-\geometry{verbose,tmargin=1in,bmargin=1in,lmargin=1in,rmargin=1in}
-\setcounter{secnumdepth}{3}
-\setcounter{tocdepth}{3}
-\usepackage{babel}
-\usepackage{float}
-\usepackage{textcomp}
-\usepackage{url}
-\usepackage{amsmath}
-\usepackage{amssymb}
-\usepackage{graphicx}
-\usepackage[authoryear]{natbib}
-\usepackage[unicode=true,pdfusetitle,
- bookmarks=true,bookmarksnumbered=true,bookmarksopen=false,
- breaklinks=false,pdfborder={0 0 0},backref=section,colorlinks=false,pdfpagemode=None]
- {hyperref}
-\hypersetup{
- pdftex,pdfstartview=FitH,pdfpagelayout=SinglePage}
-\usepackage{breakurl}
-
-\makeatletter
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% LyX specific LaTeX commands.
-\newcommand{\noun}[1]{\textsc{#1}}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Textclass specific LaTeX commands.
-\newenvironment{lyxcode}
-{\par\begin{list}{}{
-\setlength{\rightmargin}{\leftmargin}
-\setlength{\listparindent}{0pt}% needed for AMS classes
-\raggedright
-\setlength{\itemsep}{0pt}
-\setlength{\parsep}{0pt}
-\normalfont\ttfamily}%
- \item[]}
-{\end{list}}
-\newenvironment{lyxlist}[1]
-{\begin{list}{}
-{\settowidth{\labelwidth}{#1}
- \setlength{\leftmargin}{\labelwidth}
- \addtolength{\leftmargin}{\labelsep}
- \renewcommand{\makelabel}[1]{##1\hfil}}}
-{\end{list}}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% User specified LaTeX commands.
-%\renewcommand{\baselinestretch}{1.5}
-
-% figures
-\usepackage[dvips]{epsfig}
-
-% fonts
-\usepackage{times}
-
-% hyperlinks to sections and references
-
-\newcommand{\urlwithparentheses}[1]{(\url{#1})}
-
-% biblio GJI
-\bibliographystyle{abbrvnat}
-
-\newcommand{\nexxi}{\mbox{\texttt{NEX\_XI\/}}}
-\newcommand{\nexeta}{\mbox{\texttt{NEX\_ETA\/}}}
-\newcommand{\nprocxi}{\mbox{\texttt{NPROC\_XI\/}}}
-\newcommand{\nproceta}{\mbox{\texttt{NPROC\_ETA\/}}}
-\newcommand{\nchunks}{\mbox{\texttt{NCHUNKS\/}}}
-\newcommand{\eg}{e.g.,}
-
-% colors for the text
-\@ifundefined{definecolor}
- {\usepackage{color}}{}
-\newcommand{\red}[1]{\textcolor{Red}{#1}}
-
-\usepackage{babel}
-
-\makeatother
-
-\begin{document}
-\begin{center}
-\thispagestyle{empty}\vspace*{-1.8truecm} %
-\makebox[1\textwidth]{%
-\includegraphics[width=0.83\paperwidth]{figures/specfem_3d_Cartesian-cover} %
-}
-\par\end{center}
-
-\title{\textbf{SPECFEM3D Cartesian}\\
- \textbf{User Manual}}
-
-\author{$\copyright$ Princeton University (USA), CNRS and University of Marseille (France),\\
-and ETH Z\"urich (Switzerland)\\
-Version 3.0}
-
-\date{\today}
-
-\maketitle
-
-\section*{Authors}
-
-\noindent The SPECFEM3D Cartesian package was first developed by Dimitri
-Komatitsch and Jean-Pierre Vilotte at Institut de Physique du Globe
-(IPGP) in Paris, France from 1995 to 1997 and then by Dimitri Komatitsch
-and Jeroen Tromp at Harvard University and Caltech, USA, starting in 1998.
-The story started on April 4, 1995, when Prof. Yvon Maday from CNRS and University of Paris, France, gave a lecture to
-Dimitri Komatitsch and Jean-Pierre Vilotte at IPG about the nice properties of the Legendre spectral-element method with diagonal mass matrix that he had used for
-other equations. We are deeply indebted and thankful to him for that.
-That followed a visit by Dimitri Komatitsch to OGS (Istituto Nazionale di Oceanografia e di Geofisica Sperimentale) in Trieste, Italy, in February 1995
-to meet with G\'eza Seriani and Enrico Priolo, who introduced him to their 2D Chebyshev version of the spectral-element method with a non-diagonal mass matrix.
-We are deeply indebted and thankful to them for that.
-
-%%%%%%%% please do NOT merge the lines of the paragraph below, it makes it difficult to insert new names in alphabetical order; leave a single name per line
-%%%%%%%% please do NOT merge the lines of the paragraph below, it makes it difficult to insert new names in alphabetical order; leave a single name per line
-%%%%%%%% please do NOT merge the lines of the paragraph below, it makes it difficult to insert new names in alphabetical order; leave a single name per line
-Since then it has been developed and maintained by a development team:
-in alphabetical order,
-Jean-Paul (Pablo) Ampuero,
-Kangchen Bai,
-Piero Basini,
-C\'eline Blitz,
-Alexis Bottero,
-Ebru Bozda\u{g},
-Emanuele Casarotti,
-Joseph Charles,
-Min Chen,
-Paul Cristini,
-Percy Galvez,
-Dominik G\"oddeke,
-Vala Hj\"orleifsd\'ottir,
-Sue Kientz,
-Dimitri Komatitsch,
-Jes\'us Labarta,
-Nicolas Le Goff,
-Pieyre Le Loher,
-Matthieu Lefebvre,
-Qinya Liu,
-Yang Luo,
-Alessia Maggi,
-Federica Magnoni,
-Roland Martin,
-Ren\'e Matzen,
-Dennis McRitchie,
-Matthias Meschede,
-Peter Messmer,
-David Mich\'ea,
-Surendra Nadh Somala,
-Tarje Nissen-Meyer,
-Daniel Peter,
-Max Rietmann,
-Elliott Sales de Andrade,
-Brian Savage,
-Bernhard Schuberth,
-Anne Sieminski,
-Leif Strand,
-Carl Tape,
-Jeroen Tromp,
-Jean-Pierre Vilotte,
-Zhinan Xie,
-Chang-Hua Zhang,
-Hejun Zhu.\\
-
-The cover graphic of the manual was created by Santiago Lombeyda from
-Caltech's Center for Advanced Computing Research (CACR) \urlwithparentheses{http://www.cacr.caltech.edu}.\\
-
-The code is released open-source under the GNU version 2+ license, see the license at the end of this manual.\\
-
-\newpage{}
-\section*{Current and past main participants or main sponsors of the SPECFEM project (in no particular order)}
-%
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=0.125\textwidth]{figures/logo_cnrs}\vspace*{2truemm}
-\includegraphics[width=0.125\textwidth]{figures/logo_princeton}\vspace*{2truemm}
-\includegraphics[width=0.15\textwidth]{figures/logo_aix_marseille_universite}\vspace*{0.02truemm}
-\includegraphics[width=0.15\textwidth]{figures/logo_ETH}\vspace*{2truemm}
-\includegraphics[width=0.12\textwidth]{figures/logo_CSC_China}\vspace*{0.02truemm}
-\includegraphics[width=0.15\textwidth]{figures/logo_inria}\vspace*{2truemm}
-\includegraphics[width=0.14\textwidth]{figures/logo_UPPA}
-\par\end{centering}
-%
-\vspace*{-2truemm}
-%
-\noindent \begin{centering}
-\includegraphics[width=0.15\textwidth]{figures/logo_NVIDIA}\vspace*{2truemm}
-\includegraphics[width=0.125\textwidth]{figures/logo_IUF}\vspace*{2truemm}
-\includegraphics[width=0.135\textwidth]{figures/logo_Caltech}\vspace*{2truemm}
-\includegraphics[width=0.135\textwidth]{figures/logo_Harvard}\vspace*{2truemm}
-\includegraphics[width=0.135\textwidth]{figures/logo_IPGP}\vspace*{2truemm}
-\includegraphics[width=0.112\textwidth]{figures/logo_ANR}\vspace*{2truemm}
-\includegraphics[width=0.112\textwidth]{figures/logo_NSF}\vspace*{2truemm}
-\par\end{centering}
-%
-\vspace*{-2truemm}
-%
-\noindent \begin{centering}
-\includegraphics[width=0.190\textwidth]{figures/logo_European_Union}\vspace*{2truemm}
-\includegraphics[width=0.112\textwidth]{figures/logo_GENCI}\vspace*{2truemm}
-\includegraphics[width=0.112\textwidth]{figures/logo_PRACE}\vspace*{2truemm}
-\includegraphics[width=0.112\textwidth]{figures/logo_CINES}\vspace*{2truemm}
-\includegraphics[width=0.130\textwidth]{figures/logo_Oak_Ridge}\vspace*{2truemm}
-\hspace*{3mm}\includegraphics[width=0.130\textwidth]{figures/logo_fondation_Del_Duca}
-\includegraphics[width=0.130\textwidth]{figures/logo_CIG}\vspace*{2truemm}
-\par\end{centering}
-%
-\vspace*{-2truemm}
-%
-\noindent \begin{centering}
-\includegraphics[width=0.112\textwidth]{figures/logo_University_of_Toronto}\vspace*{2truemm}
-\includegraphics[width=0.109\textwidth]{figures/logo_INGV}\vspace*{2truemm}
-\includegraphics[width=0.102\textwidth]{figures/logo_Univ_Toulouse}\vspace*{2truemm}
-\includegraphics[width=0.112\textwidth]{figures/logo_TOTAL}\vspace*{2truemm}
-\includegraphics[width=0.130\textwidth]{figures/logo_Fairbanks}\vspace*{2truemm}
-\includegraphics[width=0.120\textwidth]{figures/logo_CINECA}\vspace*{2truemm}
-\includegraphics[width=0.140\textwidth]{figures/logo_Intel_Exascale_Labs}\vspace*{2truemm}
-\includegraphics[width=0.130\textwidth]{figures/logo_Maison_Simulation}
-\par\end{centering}
-\end{figure}
-
-\newpage{}
-
-\tableofcontents{}
-
-\chapter{Introduction}
-
-The software package SPECFEM3D Cartesian simulates seismic wave propagation
-at the local or regional scale and performs full waveform imaging (FWI) or adjoint tomography based upon the spectral-element method
-(SEM). The SEM is a continuous Galerkin technique \citep{TrKoLi08,PeKoLuMaLeCaLeMaLiBlNiBaTr11},
-which can easily be made discontinuous \citep{BeMaPa94,Ch00,KoWoHu02,ChCaVi03,LaWaBe05,Kop06,WiStBuGh10,AcKo11};
-it is then close to a particular case of the discontinuous Galerkin
-technique \citep{ReHi73,LeRa74,Arn82,JoPi86,BoMaHe91,FaRi99,HuHuRa99,CoKaSh00,GiHeWa02,RiWh03,MoRi05,GrScSc06,AiMoMu06,BeLaPi06,DuKa06,DeSeWh08,PuAmKa09,WiStBuGh10,DeSe10,EtChViGl10},
-with optimized efficiency because of its tensorized basis functions
-\citep{WiStBuGh10,AcKo11}. In particular, it can accurately handle
-very distorted mesh elements \citep{OlSe11}.\\
-
-It has very good accuracy and convergence properties \citep{MaPa89,SePr94,DeFiMu02,Coh02,DeSe07,SeOl08,AiWa09,AiWa10,MeStTh12}.
-The spectral element approach admits spectral rates of convergence
-and allows exploiting $hp$-convergence schemes. It is also very well
-suited to parallel implementation on very large supercomputers \citep{KoTsChTr03,TsKoChTr03,KoLaMi08a,CaKoLaTiMiLeSnTr08,KoViCh10}
-as well as on clusters of GPU accelerating graphics cards \citep{Kom11,MiKo10,KoMiEr09,KoErGoMi10}.
-Tensor products inside each element can be optimized to reach very
-high efficiency \citep{DeFiMu02}, and mesh point and element numbering
-can be optimized to reduce processor cache misses and improve cache
-reuse \citep{KoLaMi08a}. The SEM can also handle triangular (in 2D)
-or tetrahedral (in 3D) elements \citep{WinBoyd96,TaWi00,KoMaTrTaWi01,Coh02,MeViSa06}
-as well as mixed meshes, although with increased cost and reduced
-accuracy in these elements, as in the discontinuous Galerkin method.\\
-
-Note that in many geological models in the context of seismic wave
-propagation studies (except for instance for fault dynamic rupture
-studies, in which very high frequencies or supershear rupture need
-to be modeled near the fault, see e.g. \citet{BeGlCrViPi07,BeGlCrVi09,PuAmKa09,TaCrEtViBeSa10})
-a continuous formulation is sufficient because material property contrasts
-are not drastic and thus conforming mesh doubling bricks can efficiently
-handle mesh size variations \citep{KoTr02a,KoLiTrSuStSh04,LeChLiKoHuTr08,LeChKoHuTr09,LeKoHuTr09}.\\
-
-For a detailed introduction to the SEM as applied to regional seismic
-wave propagation, please consult \citet{PeKoLuMaLeCaLeMaLiBlNiBaTr11,TrKoLi08,KoVi98,KoTr99,ChKoViCaVaFe07}
-and in particular \citet{LeKoHuTr09,LeChKoHuTr09,LeChLiKoHuTr08,GoAmTaCaSmSaMaKo09,WiKoScTr04,KoLiTrSuStSh04}.
-A detailed theoretical analysis of the dispersion
-and stability properties of the SEM is available in \citet{Coh02}, \citet{DeSe07}, \citet{SeOl07}, \citet{SeOl08} and \citet{MeStTh12}.\\
-
-Effects due to lateral variations in compressional-wave speed, shear-wave
-speed, density, a 3D crustal model, topography and bathymetry are
-included. The package can accommodate full 21-parameter anisotropy
-(see~\citet{ChTr07}) as well as lateral variations in attenuation
-\citep{SaKoTr10}. Adjoint capabilities and finite-frequency kernel
-simulations are included \citep{TrKoLi08,PeKoLuMaLeCaLeMaLiBlNiBaTr11,LiTr06,FiIgBuKe09,ViOp09}.\\
-
-The SEM was originally developed in computational fluid dynamics \citep{Pat84,MaPa89}
-and has been successfully adapted to address problems in seismic wave
-propagation. Early seismic wave propagation applications of the SEM,
-utilizing Legendre basis functions and a perfectly diagonal mass matrix,
-include \citet{CoJoTo93}, \citet{Kom97}, \citet{FaMaPaQu97}, \citet{CaGa97},
-\citet{KoVi98} and \citet{KoTr99}, whereas applications involving
-Chebyshev basis functions and a nondiagonal mass matrix include \citet{SePr94},
-\citet{PrCaSe94} and \citet{SePrPr95}.\\
-
-All SPECFEM3D software is written in Fortran2003 with full portability
-in mind, and conforms strictly to the Fortran2003 standard. It uses
-no obsolete or obsolescent features of Fortran. The package uses parallel
-programming based upon the Message Passing Interface (MPI) \citep{GrLuSk94,Pac97}.\\
-
-SPECFEM3D won the Gordon Bell award for best performance at the SuperComputing~2003
-conference in Phoenix, Arizona (USA) (see \citet{KoTsChTr03} and
-\url{www.sc-conference.org/sc2003/nrfinalaward.html}). It was a finalist
-again in 2008 for a run at 0.16 petaflops (sustained) on 149,784 processors
-of the `Jaguar' Cray XT5 system at Oak Ridge National Laboratories
-(USA) \citep{CaKoLaTiMiLeSnTr08}. It also won the BULL Joseph Fourier
-supercomputing award in 2010.\\
-
-It reached the sustained one petaflop performance level for the first time in February 2013
-on the Blue Waters Cray supercomputer at the National Center for Supercomputing Applications (NCSA), located at the University of Illinois at Urbana-Champaign (USA).\\
-
-This new release of the code includes Convolution or
-Auxiliary Differential Equation Perfectly Matched absorbing Layers
-(C-PML or ADE-PML) \citep{MaKoEz08,MaKoGe08,MaKo09,MaKoGeBr10,KoMa07}.
-It also includes support for GPU graphics card acceleration \citep{Kom11,MiKo10,KoMiEr09,KoErGoMi10}.\\
-
-The next release of the code will use the PT-SCOTCH parallel and threaded
-version of SCOTCH for mesh partitioning instead of the serial version.\\
-
-SPECFEM3D Cartesian includes coupled fluid-solid domains and adjoint
-capabilities, which enables one to address seismological inverse problems,
-but for linear rheologies only so far. To accommodate visco-plastic
-or non-linear rheologies, the reader can refer to the \href{http://geoelse.stru.polimi.it/}{GeoELSE}software
-package \citep{CaGa97,StPaIg09}.
-
-\section{Citation}
-
-If you use SPECFEM3D Cartesian for your own research, please cite
-at least one of the following articles:
-\begin{description}
-\item [{Numerical simulations in general}] ~\\
- Forward and adjoint simulations are described in detail in \citet{TrKoLi08,PeKoLuMaLeCaLeMaLiBlNiBaTr11,VaCaSaKoVi99,KoMiEr09,KoErGoMi10,ChKoViCaVaFe07,MaKoDi09,KoViCh10,CaKoLaTiMiLeSnTr08,TrKoHjLiZhPeBoMcFrTrHu10,KoRiTr02,KoTr02a,KoTr02b,KoTr99}
-or \citet{KoVi98}. Additional aspects of adjoint simulations are
-described in \citet{TrTaLi05,LiTr06,TrKoLi08,LiTr08,TrKoHjLiZhPeBoMcFrTrHu10,PeKoLuMaLeCaLeMaLiBlNiBaTr11}.
-Domain decomposition is explained in detail in \citet{MaKoBlLe08},
-and excellent scaling up to 150,000 processor cores is shown for instance
-in \citet{CaKoLaTiMiLeSnTr08,KoLaMi08a,MaKoBlLe08,KoErGoMi10,Kom11},
-\item [{GPU computing}] ~\\
- Computing on GPU graphics cards for acoustic or seismic wave propagation
-applications is described in detail in \citet{Kom11,MiKo10,KoMiEr09,KoErGoMi10}.
-\end{description}
-\noindent If you use this new version, which has non blocking
-MPI for much better performance for medium or large runs, please cite
-at least one of these six articles, in which results of non blocking
-MPI runs are presented: \citet{PeKoLuMaLeCaLeMaLiBlNiBaTr11,KoErGoMi10,KoViCh10,Kom11,CaKoLaTiMiLeSnTr08,MaKoBlLe08}.\\
-
-\noindent If you work on geophysical applications, you may be interested
-in citing some of these application articles as well, among others:
-\begin{description}
-\item [{Southern California simulations}] ~\\
- \citet{KoLiTrSuStSh04,KrJiKoTr06a,KrJiKoTr06b}.
-
-If you use the 3D southern California model, please cite \citet{SuSh03}
-(Los Angeles model), \citet{lovelyetal06} (Salton Trough), and \citet{hauksson2000}
-(southern California). The Moho map was determined by \citet{zhu&kanamori2000}.
-The 1D SoCal model was developed by \citet{DrHe90}.
-
-\item [{Anisotropy}] ~\\
- \citet{ChTr07,JiTsKoTr05,ChFaKo04,FaChKo04,RiRiKoTrHe02,TrKo00}.
-\item [{Attenuation}] ~\\
- \citet{SaKoTr10,KoTr02a,KoTr99}.
-\item [{Topography}] ~\\
- \citet{LeKoHuTr09,LeChKoHuTr09,LeChLiKoHuTr08,GoAmTaCaSmSaMaKo09,WiKoScTr04}.
-\end{description}
-The corresponding Bib\TeX{} entries may be found in file \texttt{doc/USER\_MANUAL/bibliography.bib}.
-
-\section{Support}
-
-This material is based upon work supported by the USA National Science
-Foundation under Grants No. EAR-0406751 and EAR-0711177, by the French
-CNRS, French INRIA Sud-Ouest MAGIQUE-3D, French ANR NUMASIS under
-Grant No. ANR-05-CIGC-002, and European FP6 Marie Curie International
-Reintegration Grant No. MIRG-CT-2005-017461. Any opinions, findings,
-and conclusions or recommendations expressed in this material are
-those of the authors and do not necessarily reflect the views of the
-USA National Science Foundation, CNRS, INRIA, ANR or the European
-Marie Curie program.
-
-\chapter{\label{cha:Getting-Started}Getting Started}
-
-To download the SPECFEM3D\_Cartesian software package, type this:
-\begin{lyxcode}
-git clone --recursive --branch devel https://github.com/geodynamics/specfem3d.git
-\end{lyxcode}
-{\small \par}
-
-We recommend that you add {\texttt{ulimit -S -s unlimited}} to your
-{\texttt{.bash\_profile}} file and/or {\texttt{limit stacksize
-unlimited }} to your {\texttt{.cshrc}} file to suppress any potential
-limit to the size of the Unix stack.\\
-
-Then, to configure the software for your system, run the \texttt{configure}
-shell script. This script will attempt to guess the appropriate configuration
-values for your system. However, at a minimum, it is recommended that
-you explicitly specify the appropriate command names for your Fortran
-compiler and MPI package (another option is to define FC, CC and MPIF90
-in your .bash\_profile or your .cshrc file):
-\begin{lyxcode}
-./configure~FC=ifort~MPIFC=mpif90
-\end{lyxcode}
-Note that MPI must be installed with MPI-IO enabled because parts of SPECFEM3D perform I/Os through MPI-IO.
-
-Before running the \texttt{configure} script, you should probably
-edit file \texttt{flags.guess} to make sure that it contains the best
-compiler options for your system. Known issues or things to check
-are:
-\begin{description}
-\item [{\texttt{Intel ifort compiler}}] See if you need to add \texttt{-assume byterecl} for your machine.
-\textbf{In the case of that compiler, we have noticed that versions dot zero sometimes have bugs or issues that can lead to wrong results when running the code, thus we \emph{strongly} recommend using versions dot one or above (for instance version 13.1 instead of 13.0, version 14.1 instead of 14.0 and so on)}.
-\item [{\texttt{IBM compiler}}] See if you need to add \texttt{-qsave}
-or \texttt{-qnosave} for your machine.
-\item [{\texttt{Mac OS}}] You will probably need to install \texttt{XCODE}.
-\end{description}
-When compiling on an IBM machine with the \texttt{xlf} and \texttt{xlc}
-compilers, we suggest running the \texttt{configure} script with the
-following options:\\
-
-\noindent \texttt{./configure FC=xlf90\_r MPIFC=mpif90 CC=xlc\_r CFLAGS=\textquotedbl{}-O3
--q64\textquotedbl{} FCFLAGS=\textquotedbl{}-O3 -q64\textquotedbl{}
--with-scotch-dir=...}.\\
-
-On SGI systems, \texttt{flags.guess} automatically informs \texttt{configure}
-to insert `\texttt{`TRAP\_FPE=OFF}'' into the generated \texttt{Makefile}
-in order to turn underflow trapping off.\\
-
-You can add \texttt{--enable-vectorization} to the configuration options to speed up the code in the fluid (acoustic) and elastic parts.
-This works fine if (and only if) your computer always allocates a contiguous memory block for each allocatable array;
-this is the case for most machines and most compilers, but not all. To disable this feature, use option \texttt{--disable-vectorization}.
-For more details see \href{https://github.com/geodynamics/specfem3d/issues/81}{github.com/geodynamics/specfem3d/issues/81} .
-To check if that option works fine on your machine, run the code with and without it for an acoustic/elastic model and make sure the seismograms are identical.\\
-
-Note that we use CUBIT (now called Trelis) to create meshes of hexahedra, but other packages
-can be used as well, for instance GiD from \url{http://gid.cimne.upc.es}
-or Gmsh from \url{http://geuz.org/gmsh} \citep{GeRe09}. Even mesh
-creation packages that generate tetrahedra, for instance TetGen from
-\url{http://tetgen.berlios.de}, can be used because each tetrahedron
-can then easily be decomposed into four hexahedra as shown in the
-picture of the TetGen logo at \url{http://tetgen.berlios.de/figs/Delaunay-Voronoi-3D.gif};
-while this approach does not generate hexahedra of optimal quality,
-it can ease mesh creation in some situations and it has been shown
-that the spectral-element method can very accurately handle distorted
-mesh elements \citep{OlSe11}.
-
-The SPECFEM3D Cartesian software package relies on the SCOTCH library
-to partition meshes created with CUBIT. METIS \citep{KaKu98a,KaKu98c,KaKu98b}
-can also be used instead of SCOTCH if you prefer, by editing file
-src/decompose\_mesh/decompose\_mesh.F90 and uncommenting the flag
-USE\_METIS\_INSTEAD\_OF\_SCOTCH. You will also then need to install
-and compile Metis version 4.0 (do {*}NOT{*} install Metis version
-5.0, which has incompatible function calls) and edit \texttt{src/decompose\_mesh/Makefile.in}
-and uncomment the METIS link flag in that file before running \texttt{configure}.\\
-
-The SCOTCH library \citep{PeRo96} provides efficent static mapping,
-graph and mesh partitioning routines. SCOTCH is a free software package
-developed by Fran\c{c}ois Pellegrini et al. from LaBRI and INRIA in Bordeaux,
-France, downloadable from the web page \url{https://gforge.inria.fr/projects/scotch/}.
-In case no SCOTCH libraries can be found on the system, the configuration
-will bundle the version provided with the source code for compilation.
-The path to an existing SCOTCH installation can to be set explicitly
-with the option \texttt{-{}-with-scotch-dir}. Just as an example:
-\begin{lyxcode}
-./configure~FC=ifort~MPIFC=mpif90~-{}-with-scotch-dir=/opt/scotch
-\end{lyxcode}
-If you use the Intel ifort compiler to compile the code, we recommend
-that you use the Intel icc C compiler to compile Scotch, i.e., use:
-\begin{lyxcode}
-./configure~CC=icc~FC=ifort~MPIFC=mpif90
-\end{lyxcode}
-
-When compiling the SCOTCH source code, if you get a message such as: "ld: cannot find -lz",
-the Zlib compression development library is probably missing on your machine and you will need to install it or ask your system administrator to
-do so. On Linux machines the package is often called "zlib1g-dev" or similar. (thus "sudo apt-get install zlib1g-dev" would install it)
-
-To compile a serial version of the code for small meshes that fits
-on one compute node and can therefore be run serially, run \texttt{configure}
-with the \texttt{-{}-without-mpi} option to suppress all calls to
-MPI.
-
-A summary of the most important configuration variables follows.
-\begin{description}
-\item [{\texttt{F90}}] Path to the Fortran compiler.
-\item [{\texttt{MPIF90}}] Path to MPI Fortran.
-\item [{\texttt{MPI\_FLAGS}}] Some systems require this flag to link to
-MPI libraries.
-\item [{\texttt{FLAGS\_CHECK}}] Compiler flags.
-\end{description}
-The configuration script automatically creates for each executable
-a corresponding \texttt{Makefile} in the \texttt{src/} subdirectoy.
-The \texttt{Makefile} contains a number of suggested entries for various
-compilers, e.g., Portland, Intel, Absoft, NAG, and Lahey. The software
-has run on a wide variety of compute platforms, e.g., various PC clusters
-and machines from Sun, SGI, IBM, Compaq, and NEC. Select the compiler
-you wish to use on your system and choose the related optimization
-flags. Note that the default flags in the \texttt{Makefile} are undoubtedly
-not optimal for your system, so we encourage you to experiment with
-these flags and to solicit advice from your systems administrator.
-Selecting the right compiler and optimization flags can make a tremendous
-difference in terms of performance. We welcome feedback on your experience
-with various compilers and flags.
-
-Now that you have set the compiler information, you need to select
-a number of flags in the \texttt{constants.h} file depending on your
-system:
-\begin{description}
-\item [{\texttt{LOCAL\_PATH\_IS\_ALSO\_GLOBAL}}] Set to \texttt{.false.}
-on most cluster applications. For reasons of speed, the (parallel)
-distributed database generator typically writes a (parallel) database
-for the solver on the local disks of the compute nodes. Some systems
-have no local disks, e.g., BlueGene or the Earth Simulator, and other
-systems have a fast parallel file system, in which case this flag
-should be set to \texttt{.true.}. Note that this flag is not used
-by the database generator or the solver; it is only used for some
-of the post-processing.
-\end{description}
-The package can run either in single or in double precision mode.
-The default is single precision because for almost all calculations
-performed using the spectral-element method using single precision
-is sufficient and gives the same results (i.e. the same seismograms);
-and the single precision code is faster and requires exactly half
-as much memory. Select your preference by selecting the appropriate
-setting in the \texttt{constants.h} file:
-\begin{description}
-\item [{\texttt{CUSTOM\_REAL}}] Set to \texttt{SIZE\_REAL} for single precision
-and \texttt{SIZE\_DOUBLE} for double precision.
-\end{description}
-In the \texttt{precision.h} file:
-\begin{description}
-\item [{\texttt{CUSTOM\_MPI\_TYPE}}] Set to \texttt{MPI\_REAL} for single
-precision and \texttt{MPI\_DOUBLE\_PRECISION} for double precision.
-\end{description}
-On many current processors (e.g., Intel, AMD, IBM Power), single precision
-calculations are significantly faster; the difference can typically
-be 10\% to 25\%. It is therefore better to use single precision. What
-you can do once for the physical problem you want to study is run
-the same calculation in single precision and in double precision on
-your system and compare the seismograms. If they are identical (and
-in most cases they will), you can select single precision for your
-future runs.\\
-
-If your compiler has problems with the \texttt{use mpi} statements that are used in the code, use the script called
-\texttt{replace\_use\_mpi\_with\_include\_mpif\_dot\_h.pl} in the root directory to replace all of them with \texttt{include 'mpif.h'} automatically.
-
-\section{Adding OpenMP support in addition to MPI}
-
-NOTE FROM JULY 2013: OpenMP support is maybe / probably not maintained any more.
-Thus the section below is maybe obsolete.
-
-OpenMP support can be enabled in addition to MPI. However, in many
-cases performance will not improve because our pure MPI implementation
-is already heavily optimized and thus the resulting code will in fact
-be slightly slower. A possible exception could be IBM BlueGene-type
-architectures.
-
-To enable OpenMP, uncomment the OpenMP compiler option in two lines
-in file \texttt{src/specfem3D/Makefile.in} (before running \texttt{configure})
-and also uncomment the \texttt{\#define USE\_OPENMP} statement in
-file\\
- \texttt{src/specfem3D/specfem3D.F90}.
-
-The DO-loop using OpenMP threads has a SCHEDULE property. The OMP\_SCHEDULE
-environment variable can set the scheduling policy of that DO-loop.
-Tests performed by Marcin Zielinski at SARA (The Netherlands) showed
-that often the best scheduling policy is DYNAMIC with the size of
-the chunk equal to the number of OpenMP threads, but most preferably
-being twice as the number of OpenMP threads (thus chunk size = 8 for
-4 OpenMP threads etc). If OMP\_SCHEDULE is not set or is empty, the
-DO-loop will assume generic scheduling policy, which will slow down
-the job quite a bit.
-
-\section{Compiling on an IBM BlueGene}
-
-\underline{More recent installation instruction for IBM BlueGene, from April 2013:}\\
-
-\noindent Edit file \texttt{flags.guess} and put this for \texttt{FLAGS\_CHECK}:\\
-
-\noindent \texttt{-g -qfullpath -O2 -qsave -qstrict -qtune=qp -qarch=qp
--qcache=auto -qhalt=w}\\
- \texttt{-qfree=f90 -qsuffix=f=f90 -qlanglvl=95pure -Q -Q+rank,swap\_all
--Wl,-relax}\\
-
-\noindent The most relevant are the -qarch and -qtune flags, otherwise
-if these flags are set to ``auto'' then they are wrongly assigned
-to the architecture of the frond-end node, which is different from
-that on the compute nodes. You will need to set these flags to the
-right architecture for your BlueGene compute nodes, which is not necessarily
-``qp''; ask your system administrator. On some machines if is necessary
-to use -O2 in these flags instead of -O3 due to a compiler bug of
-the XLF version installed. We thus suggest to first try -O3, and then
-if the code does not compile or does not run fine then switch back
-to -O2. The debug flags (-g, -qfullpath) do not influence performance
-but are useful to get at least some insights in case of problems.\\
-
-\noindent Before running \texttt{configure}, select the XL Fortran
-compiler by typing \texttt{module load bgq-xl/1.0} or \texttt{module
-load bgq-xl} (another, less efficient option is to load the GNU compilers
-using \texttt{module load bgq-gnu/4.4.6} or similar).\\
-
-\noindent Then, to configure the code, type this:\\
-
-\noindent \texttt{./configure FC=bgxlf90\_r MPIFC=mpixlf90\_r CC=bgxlc\_r
-LOCAL\_PATH\_IS\_ALSO\_GLOBAL=true}\\
-
-In order for the SCOTCH domain decomposer to compile, on some (but
-not all) Blue Gene systems you may need to run \texttt{configure}
-with \texttt{CC=gcc} instead of \texttt{CC=bgxlc\_r}.
-
-\noindent \underline{Older installation instruction for IBM BlueGene, from 2011:}\\
-
-\noindent To compile the code on an IBM BlueGene, Laurent L\'eger from
-IDRIS, France, suggests the following: compile the code with
-
-\noindent \texttt{ FLAGS\_CHECK=\textquotedbl{}-O3 -qsave -qstrict
--qtune=auto -qarch=450d -qcache=auto }~\\
- \texttt{ -qfree=f90 -qsuffix=f=f90 -g -qlanglvl=95pure -qhalt=w -Q
--Q+rank,swap\_all -Wl,-relax\textquotedbl{}}
-
-\noindent Option \textquotedbl{}-Wl,-relax\textquotedbl{} must be
-added on many (but not all) BlueGene systems to be able to link the
-binaries \texttt{xmeshfem3D} and \texttt{xspecfem3D} because the final
-link step is done by the GNU \texttt{ld} linker even if one uses \texttt{FC=bgxlf90\_r,
-MPIFC=mpixlf90\_r} and \texttt{CC=bgxlc\_r} to create all the object
-files. On the contrary, on some BlueGene systems that use the native
-AIX linker option \textquotedbl{}-Wl,-relax\textquotedbl{} can lead
-to problems and must be suppressed from \texttt{flags.guess}.
-
-\noindent Also, \texttt{AR=ar, ARFLAGS=cru} and \texttt{RANLIB=ranlib}
-are hardwired in all \texttt{Makefile.in} files by default, but to
-cross-compile on BlueGene/P one needs to change these values to \texttt{AR=bgar,
-ARFLAGS=cru} and \texttt{RANLIB=bgranlib}. Thus the easiest thing
-to do is to modify all \texttt{Makefile.in} files and the \texttt{configure}
-script to set them automatically by \texttt{configure}. One then just
-needs to pass the right commands to the \texttt{configure} script:
-
-\noindent \texttt{ ./configure -{}-prefix=/path/to/SPECFEM3DG\_SP
--{}-host=Babel -{}-build=BGP }~\\
- \texttt{ FC=bgxlf90\_r MPIFC=mpixlf90\_r CC=bgxlc\_r AR=bgar ARFLAGS=cru
-}~\\
- \texttt{ RANLIB=bgranlib LOCAL\_PATH\_IS\_ALSO\_GLOBAL=false}
-
-\noindent This trick can be useful for all hosts on which one needs
-to cross-compile.
-
-\noindent On BlueGene, one also needs to run the \texttt{xcreate\_header\_file}
-binary file manually rather than in the Makefile:
-
-\noindent \texttt{bgrun -np 1 -mode VN -exe ./bin/xcreate\_header\_file}
-
-\section{Visualizing the subroutine calling tree of the source code}
-
-\noindent
-Packages such as \texttt{Doxywizard} can be used to visualize the calling tree of the
-subroutines of the  source code. \texttt{Doxywizard} is
-a GUI front-end for configuring and running \texttt{Doxygen}.
-
-\bigskip
-
-\noindent To do your own call graphs, you can follow these simple steps below.
-
-\begin{enumerate}
-\setcounter{enumi}{-1}
-\item Install \texttt{Doxygen} \underline{and} \texttt{graphviz} (the two are usually in the package manager of classic Linux distribution).
-\item Run in the terminal : \texttt{doxygen -g}, which creates a \texttt{Doxyfile} that tells doxygen what you want it to do.
-\item Edit the Doxyfile. Two Doxyfile-type files have been already committed in the directory \\
-\noindent \texttt{specfem3d/doc/Call\_trees}:
-
-\begin{itemize}
-\item[\textbullet] \texttt{Doxyfile\_truncated\_call\_tree} will generate call graphs with maximum 3 or 4 levels of tree structure,
-\item[\textbullet] \texttt{Doxyfile\_complete\_call\_tree} will generate call graphs with complete tree structure.
-\end{itemize}
-
-The important entries in the Doxyfile are:
-
-\begin{description}
-\item [{\texttt{PROJECT\_NAME}}]
-\item [{\texttt{OPTIMIZE\_FOR\_FORTRAN}}] Set to YES
-\item [{\texttt{EXTRACT\_ALL}}] Set to YES
-\item [{\texttt{EXTRACT\_PRIVATE}}] Set to YES
-\item [{\texttt{EXTRACT\_STATIC}}] Set to YES
-\item [{\texttt{INPUT}}] From the directory \texttt{specfem3d/doc/Call\_trees}, it is \texttt{"../../src/"}
-\item [{\texttt{FILE\_PATTERNS}}] In SPECFEM case, it is \texttt{*.f90* *.F90* *.c* *.cu* *.h*}
-\item [{\texttt{HAVE\_DOT}}] Set to YES
-\item [{\texttt{CALL\_GRAPH}}] Set to YES
-\item [{\texttt{CALLER\_GRAPH}}] Set to YES
-\item [{\texttt{DOT\_PATH}}] The path where is located the dot program graphviz (if it is not in your \$PATH)
-\item [{\texttt{RECURSIVE}}] This tag can be used to turn specify whether or not subdirectories should be searched for input files as well. In the case of SPECFEM, set to YES.
-\item [{\texttt{EXCLUDE}}] Here, you can exclude:  \\
-\texttt{"../../src/specfem3D/older\_not\_maintained\_partial\_OpenMP\_port"}\\
-\texttt{"../../src/decompose\_mesh/scotch"}\\
-\texttt{"../../src/decompose\_mesh/scotch\_5.1.12b"}
-\item [{\texttt{DOT\_GRAPH\_MAX\_NODES}}] to set the maximum number of nodes that will be shown in the graph.
-If the number of nodes in a graph becomes larger than this value, doxygen will truncate the graph, which is visualized by representing a node as a red box.
-Minimum value: 0, maximum value: 10000, default value: 50.
-\item [{\texttt{MAX\_DOT\_GRAPH\_DEPTH}}] to set the maximum depth of the graphs generated by dot.
-A depth value of 3 means that only nodes reachable from the root by following a path via at most 3 edges will be shown.
-Using a depth of 0 means no depth restriction.
-Minimum value: 0, maximum value: 1000, default value: 0.
-
-\end{description}
-
-\item Run : \texttt{doxygen Doxyfile}, HTML and LaTeX files created by default in \texttt{html} and \texttt{latex} subdirectories.
-
-\item To see the call trees, you have to open the file \texttt{html/index.html} in your \underline{browser}.
-You will have many informations about each subroutines of SPECFEM (not only call graphs), you can click on every boxes / subroutines.
-It show you the call, and, the caller graph of each subroutine : the subroutines called by the concerned subroutine, and the previous subroutines who call this subroutine (the previous path), respectively.
-In the case of a truncated calling tree, the boxes with a red border indicates a node that has \underline{more} arrows than are shown (in other words: the graph is truncated with respect to this node).
-
-\end{enumerate}
-
-\medskip
-
-\noindent Finally, some useful links:
-
-\begin{itemize}
-\item[\textbullet] a good and short summary for the basic utilisation of Doxygen:
-
-\url{http://www.softeng.rl.ac.uk/blog/2010/jan/30/callgraph-fortran-doxygen/},
-
-\item[\textbullet] to configure the diagrams :
-
-\url{http://www.stack.nl/~dimitri/doxygen/manual/diagrams.html},
-
-\item[\textbullet] the complete alphabetical index of the tags in Doxyfile:
-
-\url{http://www.stack.nl/~dimitri/doxygen/manual/config.html},
-
-\item[\textbullet] more generally, the Doxygen manual:
-
-\url{http://www.stack.nl/~dimitri/doxygen/manual/index.html}.
-
-\end{itemize}
-
-\medskip
-
-\section{Using the ADIOS library for I/O}
-Regular POSIX I/O can be problematic when dealing with large simulations one large
-clusters (typically more than $10,000$ processes). SPECFEM3D use the ADIOS library~\cite{Liu2013}
-to deal transparently take advantage of advanced paralllel file system features. To enable
-ADIOS, the following steps should be done:
-\begin{enumerate}
-\item Install ADIOS (available from \url{https://www.olcf.ornl.gov/center-projects/adios/}).
-Make sure that your environment variables reference it.
-\item You may want to change ADIOS related values in the \texttt{constants.h.in} file.
-The default values probably suit most cases.
-\item Configure using the \texttt{--with-adios} flag.
-\end{enumerate}
-ADIOS is currently only usable for meshfem3D generated mesh (i.e. not for meshes generated
-with CUBTI). Additional control parameters are discussed in section~\ref{cha:Main-Parameter}.
-
-\section{Becoming a developer of the code, or making small modifications in the source code}
-
-If you want to develop new features in the code, and/or if you want to make small changes, improvements, or bug fixes, you are very welcome to contribute.
-To do so, i.e.
-to access the development branch of the source code with read/write access (in a safe way, no need to worry too much about breaking the package, there is a robot called BuildBot that is in charge of checking and validating all new contributions and changes), please visit this Web page: \url{https://github.com/geodynamics/specfem3d/wiki/Using-Hub}.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\chapter{\label{cha:Mesh-Generation}Mesh Generation}
-
-The first step in running a spectral-element simulation consists of
-constructing a high-quality mesh for the region under consideration.
-We provide two possibilities to do so: (1) relying on the external,
-hexahedral mesher CUBIT, or (2) using the provided, internal mesher
-\texttt{xmeshfem3D}. In the following, we explain these two approaches.
-
-\section{\label{cha:Running-the-Mesher-CUBIT}Meshing with \texttt{CUBIT}}
-
-CUBIT is a meshing tool suite for the creation of finite-element meshes
-for arbitrarily shaped models. It has been developed and maintained
-at Sandia National Laboratories and can be purchased for a small academic
-institutional fee at \url{http://cubit.sandia.gov}. Our experience
-showed that using CUBIT greatly facilitates and speeds up the generation
-and preparation of hexahedral, conforming meshes for a variety of
-geophysical models with increasing complexity.
-
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=0.6\textwidth]{figures/mount-cubit}
-\par\end{centering}
-
-\caption{Example of the graphical user interface of CUBIT. The hexahedral mesh
-shown in the main display consists of a hexahedral discretization
-of a single volume with topography.}
-
-\label{fig:mount.cubit}
-\end{figure}
-
-\noindent
-The basic steps in creating a load-balanced, partitioned mesh with
-CUBIT are:
-\begin{description}
-\item [{1.}] setting up a hexahedral mesh with CUBIT,
-\item [{2.}] exporting the CUBIT mesh into a SPECFEM3D Cartesian file format
-and
-\item [{3.}] partitioning the SPECFEM3D Cartesian mesh files for a chosen
-number of cores.
-\end{description}
-Examples are provided in the SPECFEM3D Cartesian package in the subdirectory
-\texttt{EXAMPLES/}. We strongly encourage you to contribute your own
-example to this package by contacting the CIG Computational Seismology
-Mailing List \urlwithparentheses{cig-seismo@geodynamics.org}.
-
-\subsection{Creating the Mesh with CUBIT}
-
-For the installation and handling of the CUBIT meshing tool suite,
-please refer to the CUBIT user manual and documentation. In order
-to give you a basic understanding of how to use CUBIT for our purposes,
-examples are provided in the SPECFEM3D Cartesian package in the subdirectory
-\texttt{EXAMPLES/}:
-\begin{description}
-\item [{\texttt{homogeneous\_halfspace}}] Creates a single block model
-and assigns elastic material parameters.
-\item [{\texttt{layered\_halfspace}}] Combines two different, elastic material
-volumes and creates a refinement layer between the two. This example
-can be compared for validation against the solutions provided in subdirectory
-~\\
- \texttt{VALIDATION\_3D\_SEM\_SIMPLER\_LAYER\_SOURCE\_DEPTH/}.
-\item [{\texttt{waterlayered\_halfspace}}] Combines an acoustic and elastic
-material volume as in a schematic marine survey example.
-\item [{\texttt{tomographic\_model}}] Creates a single block model whose
-material properties will have to be read in from a tomographic model
-file during the databases creation by \texttt{xgenerate\_databases}.
-\end{description}
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=0.45\textwidth]{figures/example-homogeneous}
-\includegraphics[width=0.45\textwidth]{figures/example-2layers} \\
- \includegraphics[width=0.45\textwidth]{figures/example-water} \includegraphics[width=0.45\textwidth]{figures/example-tomo}
-\par\end{centering}
-
-\caption{Screenshots of the CUBIT examples provided in subdirectory \texttt{EXAMPLES/}:
-homogeneous halfspace (top-left), layered halfspace (top-right), water
-layered halfspace (bottom-left) and tomographic model (bottom-right).}
-
-\label{fig:examples.cubit}
-\end{figure}
-
-In each example subdirectory you will find a \texttt{README} file,
-which explains in a step-by-step tutorial the workflow for the example.
-Please feel free to contribute your own example to this package by
-contacting the CIG Computational Seismology Mailing List \urlwithparentheses{cig-seismo@geodynamics.org}.
-
-In some cases, to re-create the meshes for the examples given, just type "claro ./create\_mesh.py" or similar
-from the command line ("claro" is the command to run CUBIT from the command line).
-
-IMPORTANT note: in order to correctly set up GEOCUBIT and run the examples, please read the file called "EXAMPLES/README";
-in particular, please make sure you correctly set up the Python paths as indicated in that file.
-
-\subsection{Exporting the Mesh with \texttt{run\_cubit2specfem3d.py} }
-
-\label{subsec:Exporting-the-Mesh}
-
-Once the geometric model volumes in CUBIT are meshed, you prepare
-the model for exportation with the definition of material blocks and
-boundary surfaces. Thus, prior to exporting the mesh, you need to
-define blocks specifying the materials and absorbing boundaries in
-CUBIT. This process could be done automatically using the script \texttt{run\_cubit2specfem3d.py}
-if the mesh meets some conditions or manually, following the block
-convention:
-\begin{description}
-\item [{material\_name}] Each material should have a specific block defined
-by a unique name. The name convention of the material is to start
-with either 'elastic' or 'acoustic'. It must be then followed by a
-unique identifier, e.g. 'elastic 1', 'elastic 2', etc. The additional
-attributes to the block define the material description.
-
-For an elastic material:
-\begin{description}
-\item [{material\_id}] An integer value which is unique for this material.
-\item [{Vp}] P-wave speed of the material (given in m/s).
-\item [{Vs}] S-wave speed of the material (given in m/s).
-\item [{rho}] density of the material (given in kg/m$^{3}$).
-\item [{Q}] quality factor to use in case of a simulation with attenuation
-turned on. It should be between 1 and 9000. In case no attenuation
-information is available, it can be set to zero. Please note that
-your Vp- and Vs-speeds are given for a reference frequency. To change
-this reference frequency, you change the value of \texttt{ATTENUATION\_f0\_REFERENCE}
-in the main constants file \texttt{constants.h} found in subdirectory
-\texttt{src/shared/}.
-\item [{anisotropic\_flag}] Flag describing the anisotropic model to use
-in case an anisotropic simulation should be conducted. See the file
-\texttt{model\_aniso.f90} in subdirectory \texttt{src/generate\_databases/}
-for an implementation of the anisotropic models. In case no anisotropy
-is available, it can be set to zero.
-\end{description}
-
-Note that this material block has to be defined using all the volumes
-which belong to this elastic material. For volumes belonging to another,
-different material, you will need to define a new material block.
-
-For an acoustic material:
-\begin{description}
-\item [{material\_id}] An integer value which is unique for this material.
-\item [{Vp}] P-wave speed of the material (given in m/s).
-\item [{0}] S-wave speed of the material is ignored.
-\item [{rho}] density of the material (given in kg/m$^{3}$).
-\end{description}
-\item [{face\_topo}] Block definition for the surface which defines the
-free surface (which can have topography). The name of this block must
-be 'face\_topo', the block has to be defined using all the surfaces
-which constitute the complete free surface of the model.
-\item [{face\_abs\_xmin}] Block definition for the faces on the absorbing
-boundaries, one block for each surface with x=Xmin.
-\item [{face\_abs\_xmax}] Block definition for the faces on the absorbing
-boundaries, one block for each surface with x=Xmax.
-\item [{face\_abs\_ymin}] Block definition for the faces on the absorbing
-boundaries, one block for each surface with y=Ymin.
-\item [{face\_abs\_ymax}] Block definition for the faces on the absorbing
-boundaries, one block for each surface with y=Ymax.
-\item [{face\_abs\_bottom}] Block definition for the faces on the absorbing
-boundaries, one block for each surface with z=bottom.
-\end{description}
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=0.45\textwidth]{figures/mount-surface} \includegraphics[width=0.45\textwidth]{figures/mount-abs}
-\par\end{centering}
-
-\caption{Example of the block definitions for the free surface 'face\_topo'
-(left) and the absorbing boundaries, defined in a single block 'face\_abs\_xmin'
-(right) in CUBIT.}
-
-\label{fig:mount.abs}
-\end{figure}
-
-Optionally, instead of specifying for each surface at the boundaries
-a single block like mentioned above, you can also specify a single
-block for all boundary surfaces and name it as one of the absorbing
-blocks above, e.g. 'face\_abs\_xmin'.
-
-After the block definitions are done, you export the mesh using the
-script \texttt{cubit2specfem3d.py} provided in each of the example
-directories (linked to the common script \texttt{CUBIT\_GEOCUBIT/cubit2specfem3d.py}).
-If the export was successful, you should find the following
-files in a subdirectory \texttt{MESH/}:
-\begin{description}
-\item [{absorbing\_cpml\_file}] \textbf{(only needed in case of C-PML absorbing
-conditions)} Contains on the first line the total number of C-PML
-spectral elements in the mesh, and then on the following line the
-list of all these C-PML elements with two numbers per line: first
-the spectral element number, and then a C-PML flag indicating to which
-C-PML layer(s) that element belongs, according to the following convention:
-
-\begin{itemize}
-\item Flag = 1 : element belongs to a X CPML layer only (either in Xmin
-or in Xmax),
-\item Flag = 2 : element belongs to a Y CPML layer only (either in Ymin
-or in Ymax),
-\item Flag = 3 : element belongs to a Z CPML layer only (either in Zmin
-or in Zmax),
-\item Flag = 4 : element belongs to a X CPML layer and also to a Y CPML
-layer,
-\item Flag = 5 : element belongs to a X CPML layer and also to a Z CPML
-layer,
-\item Flag = 6 : element belongs to a Y CPML layer and also to a Z CPML
-layer,
-\item Flag = 7 : element belongs to a X, to a Y and to a Z CPML layer, i.e.,
-it belongs to a CPML corner.
-\end{itemize}
-
-Note that it does not matter whether an element belongs to a Xmin
-or to a Xmax CPML, the flag is the same in both cases; the same is
-true for Ymin and Ymax, and also Zmin and Zmax.
-
-When you have an existing CUBIT (or similar) mesh stored in SPECFEM3D
-format, i.e., if you have existing \textquotedbl{}nodes\_coords\_file\textquotedbl{}
-and \textquotedbl{}mesh\_file\textquotedbl{} files but do not know
-how to assign CPML flags to them, we have created a small serial Fortran
-program that will do that automatically for you, i.e., which will
-create the \textquotedbl{}absorbing\_cpml\_file\textquotedbl{} for
-you. That program is utils/CPML/convert\_external\_layers\_of\_a\_given\_mesh\_to\_CPML\_layers.f90,
-and a small Makefile is provided in that directory (utils/CPML).
-
-IMPORTANT: it is your responsibility to make sure that in the input
-CUBIT (or similar) mesh that this code will read in SPECFEM3D format
-from files \textquotedbl{}nodes\_coords\_file\textquotedbl{} and \textquotedbl{}mesh\_file\textquotedbl{}
-you have created layers of elements that constitute a layer of constant
-thickness aligned with the coordinate grid axes (X, Y and/or Z), so
-that this code can assign CPML flags to them. This code does NOT check
-that (because it cannot, in any easy way). The mesh inside these CPML
-layers does not need to be structured nor regular, any non-structured
-mesh is fine as long as it has flat PML inner and outer faces, parallel
-to the axes, and thus of a constant thickness. The thickness can be
-different for the X, Y and Z sides. But for X it must not vary, for
-Y it must not vary, and for Z it must not vary. If you do not know
-the exact thickness, you can use a slightly LARGER value in this code
-(say 2\% to 5\% more) and this code will fix that and will adjust
-it; never use a SMALLER value otherwise this code will miss some CPML
-elements.
-
-Note: in a future release we will remove the constraint of having
-CPML layers aligned with the coordinate axes; we will allow for meshes
-that are titled by any constant angle in the horizontal plane. However
-this is not implemented yet.
-
-Note: in the case of fluid-solid layers in contact, the fluid-solid
-interface currently needs to be flat and horizontal inside the CPML
-layer (i.e., bathymetry should become flat and horizontal when entering
-the CPML); this (small) constraint will probably remain in the code
-for a while because it makes fluid-solid matching inside the CPML
-much easier.
-
-\item [{materials\_file}] Contains the material associations for each element.
-The format is:
-
-\begin{lyxcode}
-element\_ID~material\_ID
-\end{lyxcode}
-
-where \texttt{element\_ID} is the element identifier and \texttt{material\_ID}
-is a unique identifier, positive (for materials taken from this list
-of materials, i.e. for which each spectral element has constant material
-properties taken from this list) or negative (for tomographic models,
-i.e. for spectral element whose real velocities and density will be
-assigned later by calling an external function to define model variations,
-for instance in the case of tomographic models; in such a case, material
-properties can vary inside each spectral element, i.e. be different
-at each of its Gauss-Lobatto-Legendre grid points).
-
-\item [{nummaterial\_velocity\_file}] Defines the material properties.
-
-\begin{itemize}
-\item For classical materials (i.e., spectral elements for which the velocity
-and density model will not be assigned by calling an external function
-to define for instance a tomographic model), the format is:
-
-\begin{lyxcode}
-domain\_ID~material\_ID~rho~vp~vs~Qkappa~Qmu~anisotropy\_flag
-\end{lyxcode}
-
-where \texttt{\textbf{domain\_ID}}\textbf{ is 1 for acoustic and 2
-for elastic or viscoelastic materials,} \texttt{material\_ID} a unique
-identifier, \texttt{rho} the density in $kg\, m^{-3}$, \texttt{vp}
-the P-wave speed in $m\, s^{-1}$, \texttt{vs} the S-wave speed in
-$m\, s^{-1}$, \texttt{Q} the quality factor and \texttt{anisotropy\_flag}
-an identifier for anisotropic models. Note that the \texttt{Qkappa}
-value is ignored by the code unless \texttt{FULL\_ATTENUATION\_SOLID}
-is set. Note also that both \texttt{Qkappa} and \texttt{Qmu} are ignored
-by the code unless \texttt{ATTENUATION} is set. If you want a model
-with no \texttt{Qmu} attenuation, both set \texttt{ATTENUATION} to
-\texttt{.false.} in the \texttt{Par\_file} and set \texttt{Qmu} to
-9999 here. If you want a model with no \texttt{Qkappa} attenuation,
-both set \texttt{FULL\_ATTENUATION\_SOLID} to \texttt{.false.} in
-the \texttt{Par\_file} and set \texttt{Qkappa} to 9999 here. %%magnoni
-
-\item For tomographic velocity models, please read Chapter \ref{cha:-Changing-the}
-and Section \ref{sec:Using-tomographic} `Using external tomographic
-Earth models' for further details.
-\end{itemize}
-\item [{nodes\_coords\_file}] Contains the point locations in Cartesian
-coordinates of the mesh element corners.
-\item [{mesh\_file}] Contains the mesh element connectivity. The hexahedral
-elements can have 8 or 27 nodes.\\
- See picture doc/mesh\_numbering\_convention/numbering\_convention\_27\_nodes.jpg
-to see\\
- in which (standard) order the points must be cited. In the case of
-8 nodes, just include the first 8 points.
-\item [{free\_or\_absorbing\_surface\_file\_zmax}] Contains the free surface
-connectivity or \\
- the surface connectivity of the absorbing boundary surface at the
-top (Zmax), \\
- depending on whether the top surface is defined as free or absorbing
-(\texttt{STACEY\_INSTEAD\_OF\_FREE\_SURFACE} in \texttt{DATA/Par\_file}). \\
- You should put both the surface of acoustic regions and of elastic
-regions in that file; that is, list all the element faces that constitute
-the surface of the model in that file.
-\item [{absorbing\_surface\_file\_xmax}] Contains the surface connectivity
-of the absorbing boundary surface at Xmax \\
-(also needed in the case of C-PML absorbing conditions, in order for
-the code to be able to impose Dirichlet conditions on their outer
-edge).
-\item [{absorbing\_surface\_file\_xmin}] Contains the surface connectivity
-of the absorbing boundary surface at Xmin \\
-(also needed in the case of C-PML absorbing conditions, in order for
-the code to be able to impose Dirichlet conditions on their outer
-edge).
-\item [{absorbing\_surface\_file\_ymax}] Contains the surface connectivity
-of the absorbing boundary surface at Ymax \\
-(also needed in the case of C-PML absorbing conditions, in order for
-the code to be able to impose Dirichlet conditions on their outer
-edge).
-\item [{absorbing\_surface\_file\_ymin}] Contains the surface connectivity
-of the absorbing boundary surface at Ymin \\
-(also needed in the case of C-PML absorbing conditions, in order for
-the code to be able to impose Dirichlet conditions on their outer
-edge).
-\item [{absorbing\_surface\_file\_bottom}] Contains the surface connectivity
-of the absorbing boundary surface at the bottom (Zmin) \\
-(also needed in the case of C-PML absorbing conditions, in order for
-the code to be able to impose Dirichlet conditions on their outer
-edge).
-\end{description}
-These mesh files are needed as input files for the partitioner \texttt{xdecompose\_mesh}
-to load-balance the mesh. Please see the next section for further
-details.
-
-In directory "CUBIT\_GEOCUBIT" we provide a script that can help doing the above tasks of exporting a CUBIT mesh
-to SPECFEM3D format automatically for you: "run\_cubit2specfem3d.py".
-Just edit them to indicate the path to your local installation of CUBIT and also the name of the *.cub existing CUBIT mesh file
-that you want to export to SPECFEM3D format. These scripts will do the conversion for you automatically except assigning material properties
-to the different mesh layers. To do so, you will then need to edit the file
-called "nummaterial\_velocity\_file" that will have just been created and change it from the prototype created:
-
-    0 1 vol1 --> syntax: \#material\_domain\_id \#material\_id \#rho \#vp \#vs \#Q\_kappa \#Q\_mu \#anisotropy
-    0 2 vol2 --> syntax: \#material\_domain\_id \#material\_id \#rho \#vp \#vs \#Q\_kappa \#Q\_mu \#anisotropy
-
-(where "vol1" and "vol2" here represent the volume labels that you have set while creating the mesh in CUBIT)
-to for instance
-
-    2 1 1500 2300 1800 9999.0 9999.0 0
-    2 2 1600 2500 20000 9999.0 9999.0 0
-
-\subsubsection*{Checking the mesh quality}
-
-The quality of the mesh may be inspected more precisely based upon
-the serial code in the file \texttt{check\_mesh\_quality\_}~\\
- \texttt{CUBIT\_Abaqus.f90} located in the directory \texttt{src/check\_mesh\_quality\_CUBIT\_Abaqus/}.
-Running this code is optional because no information needed by the
-solver is generated.
-
-Prior to running and compiling this code, you have to export your
-mesh in CUBIT to an ABAQUS (.inp) format. For example, export mesh
-block IDs belonging to volumes in order to check the quality of the
-hexahedral elements. You also have to determine a number of parameters
-of your mesh, such as the number of nodes and number of elements and
-modify the header of the \texttt{check\_mesh\_quality\_CUBIT\_Abaqus.f90}
-source file in directory \texttt{src/check\_mesh\_quality\_CUBIT\_Abaqus/}.
-
-\noindent
-Then, in the main directory, type
-\begin{lyxcode}
-{\small make~xcheck\_mesh\_quality\_CUBIT\_Abaqus}{\small \par}
-\end{lyxcode}
-and use
-\begin{lyxcode}
-{\small ./bin/xcheck\_mesh\_quality\_CUBIT\_Abaqus}{\small \par}
-\end{lyxcode}
-to generate an %not supported: AVS output file (\texttt{\small AVS\_meshquality.inp} in AVS UCD format) or
-OpenDX output file (\texttt{\small DX\_mesh\_quality.dx}{\small )
-that can be used to investigate mesh quality, e.g. skewness of elements,
-and a Gnuplot histogram (}\texttt{\small mesh\_quality\_histogram.txt}{\small )
-that can be plotted with gnuplot (type `}\texttt{\small gnuplot plot\_mesh\_quality\_histogram.gnu}{\small ').
-The histogram is also printed to the screen. Analyze that skewness
-histogram of mesh elements to make sure no element has a skewness
-above approximately 0.75, otherwise the mesh is of poor quality (and
-if even a single element has a skewness value above 0.80, then you
-must definitely improve the mesh). If you want to start designing
-your own meshes, this tool is useful for viewing your creations. You
-are striving for meshes with elements with `cube-like' dimensions,
-e.g., the mesh should contain no very elongated or skewed elements.}{\small \par}
-
-\subsection{Partitioning the Mesh with \texttt{xdecompose\_mesh}}
-
-The SPECFEM3D Cartesian software package performs large scale simulations
-in a parallel 'Single Process Multiple Data' way. The spectral-element
-mesh created with CUBIT needs to be distributed on the processors.
-This partitioning is executed once and for all prior to the execution
-of the solver so it is referred to as a static mapping.
-
-An efficient partitioning is important because it leverages the overall
-running time of the application. It amounts to balance the number
-of elements in each slice while minimizing the communication costs
-resulting from the placement of adjacent elements on different processors.
-\texttt{decompose\_mesh} depends on the SCOTCH library \citep{PeRo96},
-which provides efficent static mapping, graph and mesh partitioning
-routines. SCOTCH is a free software package developed by Fran\c{c}ois
-Pellegrini et al. from LaBRI and INRIA in Bordeaux, France, downloadable
-from the web page \url{https://gforge.inria.fr/projects/scotch/}.
-\\
-
-In most cases, the configuration with \texttt{./configure~FC=ifort}
-should be sufficient. During the configuration process, the script
-tries to find existing SCOTCH installations. In case your system has
-no pre-existing SCOTCH installation, we provide the source code of
-SCOTCH, which is released open source under the French CeCILL-C version
-1 license, in directory \texttt{src/decompose\_mesh/scotch\_5.1.12b}.
-This version gets bundled with the compilation of the SPECFEM3D Cartesian
-package if no libraries could have been found. If this automatic compilation
-of the SCOTCH libraries fails, please refer to file INSTALL.txt in
-that directory to see further details how to compile it on your system.
-In case you want to use a pre-existing installation, make sure you
-have correctly specified the path of the SCOTCH library when using
-the option \texttt{-{}-with-scotch-dir} with the \texttt{./configure}
-script. In the future you should be able to find more recent versions
-at \url{http://www.labri.fr/perso/pelegrin/scotch/scotch_en.html}
-. \\
-
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=0.45\textwidth]{figures/mount-partitions}
-\includegraphics[width=0.45\textwidth]{figures/mount-partitions2}
-\par\end{centering}
-
-\caption{Example of a mesh partitioning onto four cores. Each single core partition
-is colored differently. The executable \texttt{xdecompose\_mesh} can
-equally distribute the mesh on any arbitrary number of cores. Domain
-decomposition is explained in detail in \citet{MaKoBlLe08}, and excellent
-scaling up to 150,000 processor cores in shown for instance in \citet{CaKoLaTiMiLeSnTr08,KoLaMi08a,MaKoBlLe08,KoErGoMi10,Kom11}.}
-
-\label{fig:mount.partitions}
-\end{figure}
-
-When you are ready to compile, in the main directory type
-\begin{lyxcode}
-{\small make~xdecompose\_mesh}{\small \par}
-\end{lyxcode}
-If all paths and flags have been set correctly, the executable \texttt{bin/xdecompose\_mesh}
-should be produced.
-
-The partitioning is done in serial for now (in the next release we
-will provide a parallel version of that code). It needs to be run in the main directory because it expects the \texttt{./DATA/Par\_file}. The synopsis is:
-\begin{lyxcode}
-./bin/xdecompose\_mesh~nparts~input\_directory~output\_directory
-\end{lyxcode}
-where
-\begin{itemize}
-\item \texttt{nparts} is the number of partitions, i.e., the number of cores
-for the parallel simulations,
-\item \texttt{input\_directory} is the directory which holds all the files
-generated by the Python script \texttt{cubit2specfem3d.py} explained
-in the previous Section~\ref{subsec:Exporting-the-Mesh}, e.g. \texttt{./MESH/},
-and
-\item \texttt{output\_directory} is the directory for the output of this
-partitioner which stores ACII-format files named like \texttt{proc{*}{*}{*}{*}{*}{*}\_Database}
-for each partition. These files will be needed for creating the distributed
-databases, and have to reside in the directory \texttt{LOCAL\_PATH}
-specified in the main \texttt{Par\_file}, e.g. in directory \texttt{./OUTPUT\_FILES/DATABASES\_MPI}.
-Please see Chapter~\ref{cha:Creating-Distributed-Databases} for
-further details.
-\end{itemize}
-Note that all the files generated by the Python script \texttt{cubit2specfem3d.py}
-must be placed in the \texttt{input\_directory} folder before running
-the program.
-
-\section{\label{cha:Running-the-Mesher-Meshfem3D}Meshing with \texttt{xmeshfem3D}}
-
-In case you successfully ran the configuration script, you are also
-ready to compile the internal mesher. This is an alternative to CUBIT
-for the mesh generation of relatively simple models.
-
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[scale=0.5]{figures/socal_map_mpi}
-\par\end{centering}
-
-\caption{\label{fig:For-parallel-computing}For parallel computing purposes,
-the model block is subdivided in $\nprocxi\times\nproceta$ slices
-of elements. In this example we use $5^{2}=25$ processors. }
-\end{figure}
-
-\noindent
-In the main directory, type
-\begin{lyxcode}
-{\small make~xmeshfem3D}{\small \par}
-\end{lyxcode}
-If all paths and flags have been set correctly, the mesher should
-now compile and produce the executable \texttt{bin/xmeshfem3D}. Please
-note that \texttt{xmeshfem3D} must be called directly from the main
-directory, as most of the binaries of the package.
-
-Input for the mesh generation program is provided through the parameter
-file \texttt{Mesh\_Par\_file}, which resides in the subdirectory \texttt{DATA/meshfem3D\_files/}.
-(to see how to use it, see the EXAMPLES specific to the internal mesher in directory \texttt{EXAMPLES/meshfem3D\_examples/}.
-Before running the mesher, a number of parameters need to be set in
-the \texttt{Mesh\_Par\_file}. This requires a basic understanding
-of how the SEM is implemented, and we encourage you to read \citet{KoVi98,KoTr99}
-and \citet{KoLiTrSuStSh04}.
-
-The mesher and the solver use UTM coordinates internally, therefore
-you need to define the zone number for the UTM projection (e.g., zone
-11 for Los Angeles). Use decimal values for latitude and longitude
-(no minutes/seconds). These values are approximate; the mesher will
-round them off to define a square mesh in UTM coordinates. When running
-benchmarks on rectangular models, turn the UTM projection off by using
-the flag \texttt{\small SUPPRESS\_UTM\_PROJECTION}{\small , in which
-case all `longitude' parameters simply refer to the $x$~axis, and
-all `latitude' parameters simply refer to the $y$~axis. To run the
-mesher for a global simulation, the following parameters need to be
-set in the }\texttt{\small Mesh\_Par\_file}{\small :}{\small \par}
-\begin{description}
-\item [{\texttt{LATITUDE\_MIN}}] Minimum latitude in the block (negative
-for South).
-\item [{\texttt{LATITUDE\_MAX}}] Maximum latitude in the block.
-\item [{\texttt{LONGITUDE\_MIN}}] Minimum longitude in the block (negative
-for West).
-\item [{\texttt{LONGITUDE\_MAX}}] Maximum longitude in the block.
-\item [{\texttt{DEPTH\_BLOCK\_KM}}] Depth of bottom of mesh in kilometers.
-\item [{\texttt{\noun{UTM\_PROJECTION\_ZONE}}}] UTM projection zone in
-which your model resides, only valid when \texttt{SUPPRESS\_UTM\_}~\\
- \texttt{PROJECTION} is \texttt{.false.}.
-Use a negative zone number for the Southern hemisphere:
-the Northern hemisphere corresponds to zones +1 to +60,
-the Southern hemisphere to zones -1 to -60.
-We use the WGS84 (World Geodetic System 1984) reference ellipsoid for the UTM projection. If you prefer to use the Clarke 1866 ellipsoid,
-edit file \texttt{src/shared/utm\_geo.f90}, uncomment that ellipsoid and recompile the code.
-From http://en.wikipedia.org/wiki/Universal\_Transverse\_Mercator\_coordinate\_system :
-The Universal Transverse Mercator coordinate system was developed by the United States Army Corps of Engineers in the 1940s.
-The system was based on an ellipsoidal model of Earth. For areas within the contiguous United States
-the Clarke Ellipsoid of 1866 was used. For the remaining areas of Earth, including Hawaii, the International Ellipsoid was used.
-The WGS84 ellipsoid is now generally used to model the Earth in the UTM coordinate system,
-which means that current UTM northing at a given point can be 200+ meters different from the old one.
-For different geographic regions, other datum systems (e.g.: ED50, NAD83) can be used.
-\item [{\texttt{SUPPRESS\_UTM\_PROJECTION}}] set to be \texttt{.false.}
-when your model range is specified in geographical coordinates, and
-needs to be \texttt{.true.} when your model is specified in Cartesian
-coordinates. \noun{UTM projection zone in which your simulation region
-resides.}
-\item [{\texttt{INTERFACES\_FILE }}] File which contains the description
-of the topography and of the interfaces between the different layers
-of the model, if any. The number of spectral elements in the vertical
-direction within each layer is also defined in this file.
-\item [{$\nexxi$}] The number of spectral elements along one side of the
-block. This number \textit{must} be 8~$\times$~a multiple of $\nprocxi$
-defined below. Based upon benchmarks against semi-analytical discrete
-wavenumber synthetic seismograms \citep{KoLiTrSuStSh04}, determined
-that a $\nexxi=288$ run is accurate to a shortest period of roughly
-2~s. Therefore, since accuracy is determined by the number of grid
-points per shortest wavelength, for any particular value of $\nexxi$
-the simulation will be accurate to a shortest period determined by
-\begin{equation}
-\mbox{shortest period (s)}=(288/\nexxi)\times2.\label{eq:shortest_period}
-\end{equation}
- The number of grid points in each orthogonal direction of the reference
-element, i.e., the number of Gauss-Lobatto-Legendre points, is determined
-by \texttt{NGLLX} in the \texttt{constants.h} file. We generally use
-$\mbox{\texttt{NGLLX\/}}=5$, for a total of $5^{3}=125$ points per
-elements. We suggest not to change this value.
-\item [{$\nexeta$}] The number of spectral elements along the other side
-of the block. This number \textit{must} be 8~$\times$~a multiple
-of $\nproceta$ defined below.
-\item [{$\nprocxi$}] The number of processors or slices along one side
-of the block (see Figure~\ref{fig:For-parallel-computing}); we must
-have $\nexxi=8\times c\times\nprocxi$, where $c\ge1$ is a positive
-integer.
-\item [{$\nproceta$}] The number of processors or slices along the other
-side of the block; we must have $\nexeta=8\times c\times\nproceta$,
-where $c\ge1$ is a positive integer.
-\item [{\texttt{USE\_REGULAR\_MESH}}] set to be \texttt{.true.} if you
-want a perfectly regular mesh or \texttt{.false.} if you want to add
-doubling horizontal layers to coarsen the mesh. In this case, you
-also need to provide additional information by setting up the next
-three parameters.
-\item [{\texttt{NDOUBLINGS}}] The number of horizontal doubling layers.
-Must be set to \texttt{1} or \texttt{2} if \texttt{USE\_REGULAR\_MESH}
-is set to \texttt{.true.}.
-\item [{\texttt{NZ\_DOUBLING\_1}}] The position of the first doubling layer
-(only interpreted if \texttt{USE\_REGULAR\_MESH} is set to \texttt{.true.}).
-\item [{\texttt{NZ\_DOUBLING\_2}}] The position of the second doubling
-layer (only interpreted if \texttt{USE\_REGULAR\_MESH} is set to \texttt{.true.}
-and if \texttt{NDOUBLINGS} is set to \texttt{2}).
-\item [{\texttt{CREATE\_ABAQUS\_FILES}}] Set this flag to \texttt{.true.}
-to save Abaqus FEA \urlwithparentheses{www.simulia.com} mesh files
-for subsequent viewing. Turning the flag on generates files in the
-\texttt{LOCAL\_PATH} directory. See Section~\ref{sec:Mesh-graphics}
-for a discussion of mesh viewing features.
-\item [{\texttt{CREATE\_DX\_FILES}}] Set this flag to \texttt{.true.} to
-save OpenDX \urlwithparentheses{www.opendx.org} mesh files for subsequent
-viewing.
-\item [{\texttt{LOCAL\_PATH}}] Directory in which the partitions generated
-by the mesher will be written. Generally one uses a directory on the
-local disk of the compute nodes, although on some machines these partitions
-are written on a parallel (global) file system (see also the earlier
-discussion of the \texttt{LOCAL\_PATH\_IS\_ALSO\_GLOBAL} flag in Chapter~\ref{cha:Getting-Started}).
-The mesher generates the necessary partitions in parallel, one set
-for each of the $\nprocxi\times\nproceta$ slices that constitutes
-the mesh (see Figure~\ref{fig:For-parallel-computing}). After the
-mesher finishes, you can log in to one of the compute nodes and view
-the contents of the \texttt{LOCAL\_PATH} directory to see the files
-generated by the mesher. These files will be needed for creating the
-distributed databases, and have to reside in the directory \texttt{LOCAL\_PATH}
-specified in the main \texttt{Par\_file}, e.g. in directory \texttt{OUTPUT\_FILES/DATABASES\_MPI}.
-Please see Chapter~\ref{cha:Creating-Distributed-Databases} for
-further details.
-\item [{\texttt{NMATERIALS}}] The number of different materials in your
-model. In the following lines, each material needs to be defined as
-:
-
-\begin{lyxcode}
-material\_ID~rho~vp~vs~Q~anisotropy\_flag~domain\_ID
-\end{lyxcode}
-
-where
-\begin{itemize}
-\item \texttt{Q} : quality factor (0=no attenuation)
-\item \texttt{anisotropy\_flag} : 0=no anisotropy / 1,2,.. check with implementation
-in \texttt{aniso\_model.f90}
-\item \texttt{domain\_id} : 1=acoustic / 2=elastic
-\end{itemize}
-\item [{\texttt{NREGIONS}}] The number of regions in the mesh. In the following
-lines, because the mesh is regular or 'almost regular', each region
-is defined as :
-
-\begin{lyxcode}
-NEX\_XI\_BEGIN~NEX\_XI\_END~NEX\_ETA\_BEGIN~NEX\_ETA\_END~NZ\_BEGIN~NZ\_END~material\_ID
-\end{lyxcode}
-\end{description}
-The \texttt{INTERFACES\_FILE} parameter of \texttt{Mesh\_Par\_File}
-defines the file which contains the settings of the topography grid
-and of the interfaces grids. Topography is defined as a set of elevation
-values on a regular 2D grid. It is also possible to define interfaces
-between the layers of the model in the same way. The file needs to
-define several parameters:
-\begin{itemize}
-\item The number of interfaces, including the topography. This needs to
-be set at the first line. Then, from the bottom to the top of the
-model, you need to define the grids with:
-\item \texttt{SUPPRESS\_UTM\_PROJECTION} flag as described previously,
-\item number of points along $x$ and $y$ direction (NXI and NETA),
-\item minimal $x$ and $y$ coordinates (LONG\_MIN and LAT\_MIN),
-\item spacing between points along $x$ and $y$ (SPACING\_XI and SPACING\_ETA)
-and
-\item the name of the file which contains the elevation values (in $y$.$x$
-increasing order).
-\end{itemize}
-At the end of this file, you simply need to set the number of spectral
-elements in the vertical direction for each layer. We provide a few
-models in the {\texttt{EXAMPLES/}} directory. \\
-
-Finally, depending on your system, you might need to provide a file
-that tells MPI what compute nodes to use for the simulations. The
-file must have a number of entries (one entry per line) at least equal
-to the number of processors needed for the run. A sample file is provided
-in the file \texttt{mymachines}. This file is not used by the mesher
-or solver, but is required by the \texttt{go\_mesher} and \texttt{go\_solver}
-default job submission scripts. See Chapter~\ref{cha:Scheduler}
-for information about running the code on a system with a scheduler,
-e.g., LSF.
-
-Now that you have set the appropriate parameters in the \texttt{Mesh\_Par\_file}
-and have compiled the mesher, you are ready to launch it! This is
-most easily accomplished based upon the \texttt{go\_mesher} script.
-When you run on a PC cluster, the script assumes that the nodes are
-named n001, n002, etc. If this is not the case, change the \texttt{tr
--d `n'} line in the script. You may also need to edit the last command
-at the end of the script that invokes the \texttt{mpirun} command.
-See Chapter~\ref{cha:Scheduler} for information about running the
-code on a system with a scheduler, e.g., LSF.
-
-Mesher output is provided in the \texttt{OUTPUT\_FILES} directory
-in \texttt{output\_mesher.txt}; this file provides lots of details
-about the mesh that was generated. Please note that the mesher suggests
-a time step \texttt{DT} to run the solver with. The mesher output
-file also contains a table about the quality of the mesh to indicate
-possible problems with the distortions of elements. Alternatively,
-output can be directed to the screen instead by uncommenting a line
-in \texttt{constants.h}:
-\begin{lyxcode}
-!~uncomment~this~to~write~messages~to~the~screen~
-
-!~integer,~parameter~::~IMAIN~=~ISTANDARD\_OUTPUT
-\end{lyxcode}
-To control the quality of the mesh, check the standard output (either
-on the screen or in the \texttt{OUTPUT\_FILES} directory in \texttt{output\_mesher.txt})
-and analyze the skewness histogram of mesh elements to make sure no
-element has a skewness above approximately 0.75, otherwise the mesh
-is of poor quality (and if even a single element has a skewness value
-above 0.80, then you must definitely improve the mesh). To draw the
-skewness histogram on the screen, type \texttt{gnuplot plot\_mesh\_quality\_histogram.gnu}.
-
-\chapter{\label{cha:Creating-Distributed-Databases}Creating the Distributed
-Databases}
-
-After using \texttt{xmeshfem3D} or \texttt{xdecompose\_mesh}, the
-next step in the workflow is to compile \texttt{xgenerate\_}~\\
- \texttt{databases}. This program is going to create all the missing
-information needed by the SEM solver.
-
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=0.6\textwidth]{figures/workflow}
-\par\end{centering}
-
-\caption{Schematic workflow for a SPECFEM3D Cartesian simulation. The executable
-\texttt{xgenerate\_databases} creates the GLL mesh points and assigns
-specific model parameters.}
-
-\label{fig:workflow.databases}
-\end{figure}
-
-\noindent
-In the main directory, type
-\begin{lyxcode}
-{\small make~xgenerate\_databases}{\small \par}
-\end{lyxcode}
-Input for the program is provided through the main parameter file
-\texttt{Par\_file}, which resides in the subdirectory \texttt{DATA}.
-Please note that \texttt{xgenerate\_databases} must be called directly
-from the main directory, as most of the binaries of the package.
-
-\section{\label{cha:Main-Parameter}Main parameter file \texttt{Par\_file}}
-
-Before running \texttt{xgenerate\_databases}, a number of parameters
-need to be set in the main parameter \texttt{Par\_file} located in
-the subdirectory \texttt{DATA}:
-\begin{description}
-\item [{\texttt{SIMULATION\_TYPE}}] is set to 1 for forward simulations,
-2 for adjoint simulations (see Section \ref{sec:Adjoint-simulation-finite})
-and 3 for kernel simulations (see Section \ref{sec:Finite-Frequency-Kernels}).
-\item [{\texttt{SAVE\_FORWARD}}] is only set to \texttt{.true.} for a forward
-simulation with the last frame of the simulation saved, as part of
-the finite-frequency kernel calculations (see Section \ref{sec:Finite-Frequency-Kernels}).
-For a regular forward simulation, leave \texttt{SIMULATION\_TYPE}
-and \texttt{SAVE\_FORWARD} at their default values.
-\item [{\texttt{\noun{UTM\_PROJECTION\_ZONE}}}] UTM projection zone in
-which your model resides, only valid when \texttt{SUPPRESS\_UTM\_}~\\
- \texttt{PROJECTION} is \texttt{.false.}.
-\item [{\texttt{SUPPRESS\_UTM\_PROJECTION}}] set to be \texttt{.false.}
-when your model range is specified in the geographical coordinates,
-and needs to be \texttt{.true.} when your model is specified in a
-cartesian coordinates. \noun{UTM projection zone in which your simulation
-region resides.}
-\item [{\texttt{NPROC}}] The number of MPI processors, each one is assigned
-one slice of the whole mesh.
-\item [{\texttt{NSTEP}}] The number of time steps of the simulation. This
-controls the length of the numerical simulation, i.e., twice the number
-of time steps requires twice as much CPU time. This feature is not
-used at the time of generating the distributed databases but is required
-for the solver, i.e., you may change this parameter after running
-\texttt{xgenerate\_databases}.
-\item [{\texttt{DT}}] The length of each time step in seconds. This feature
-is not used at the time of generating the distributed databases but
-is required for the solver. Please see also Section~\ref{sec:Choosing-the-Time-Step}
-for further details.
-\item [{\texttt{NGNOD}}] The number of nodes for 2D and 3D shape functions
-for hexahedra. We use either 8-node mesh elements (bricks) or 27-node
-elements. If you use the internal mesher, the only option is 8-node
-bricks (27-node elements are not supported). \texttt{CUBIT} does not
-support HEX27 elements either (it can generate them, but they are
-flat, i.e. identical to HEX8). To generate HEX27 elements with curvature
-properly taken into account, you can use \texttt{Gmsh} \url{http://geuz.org/gmsh/}
-\item [{\texttt{MODEL}}] Must be set to one of the following:
-
-\begin{description}
-\item [{\textmd{Models defined by mesh parameters:}}] ~
-
-\begin{description}
-\item [{\texttt{default}}] Uses model parameters as defined by meshing
-procedures described in the previous Chapter~\ref{cha:Mesh-Generation}.
-\end{description}
-
-\item [{\textmd{1D~models~with~real~structure:}}] ~
-
-\begin{description}
-\item [{\texttt{1D\_prem}}] Isotropic version of the spherically symmetric
-Preliminary Reference Earth Model (PREM) \citep{DzAn81}.
-\item [{\texttt{1D\_socal}}] A standard isotropic 1D model for Southern
-California.
-\item [{\texttt{1D\_cascadia}}] Isotropic 1D profile for the Cascadia region.
-\end{description}
-
-\item [{\textmd{Fully~3D~models:}}] ~
-
-\begin{description}
-\item [{\texttt{aniso}}] For a user-specified fully anisotropic model.
-Parameters are set up in routines located in file \texttt{model\_aniso.f90}
-in directory \texttt{src/generate\_databases/}. See Chapter~\ref{cha:-Changing-the}
-for a discussion on how to specify your own 3D model.
-
-\item [{\texttt{external}}] For a user-specified isotropic model which
-uses externally defined model parameters. Uses external model definitions
-set up in routines located in file \texttt{model\_external\_values.f90}
-in directory \texttt{src/generate\_databases/}. Please modify these
-generic template routines to use your own model definitions.
-\item [{\texttt{gll}}] For a user-specified isotropic model which uses
-external binary files for $v_{p}$, $v_{s}$ and $\rho$. Binary files
-are given in the same format as when outputted by the \texttt{xgenerate\_databases}
-executable when using option \texttt{SAVE\_MESH\_FILES}. These binary
-files define the model parameters on all GLL points which can be used
-for iterative inversion procedures.
-\item [{\texttt{salton\_trough}}] A 3D $V_{p}$ model for Southern California.
-Users must provide the corresponding data file \texttt{regrid3\_vel\_p.bin}
-in directory \texttt{DATA/st\_3D\_block\_harvard/}.
-\item [{\texttt{tomo}}] For a user-specified 3D isotropic model which uses
-a tomographic model file \texttt{tomographic\_model.xyz} in directory
-\texttt{DATA}. See Section \ref{sec:Using-tomographic}, for a discussion
-on how to specify your own 3D tomographic model.
-\end{description}
-\end{description}
-\item [{\texttt{APPROXIMATE\_OCEAN\_LOAD}}] Set to \texttt{.true.} if the
-effect of the oceans on seismic wave propagation should be incorporated
-based upon the (rough) approximate treatment discussed in \citet{KoTr02b}.
-This feature is inexpensive from a numerical perspective, both in
-terms of memory requirements and CPU time. This approximation is accurate
-at periods of roughly 20~s and longer. At shorter periods the effect
-of water phases/reverberations is not taken into account, even when
-the flag is on. If you want to model the effect of a fluid-solid model
-at short periods, then set this flag to \texttt{.false.} and mesh
-the fluid layer explicitly in your mesher, so that it is computed
-accurately and without this approximation.
-\item [{\texttt{TOPOGRAPHY}}] This feature is only effective if \texttt{APPROXIMATE\_OCEAN\_LOAD}
-is set to \texttt{.true.}. Set to \texttt{.true.} if topography and
-bathymetry should be read in based upon the topography file specified
-in the main constants file \texttt{constants.h} found in subdirectory
-\texttt{src/shared/} to evaluate elevations. If not set, elevations
-will be read from the numerical mesh.
-\item [{\texttt{ATTENUATION}}] Set to \texttt{.true.} if attenuation should
-be incorporated. Turning this feature on increases the memory requirements
-significantly (roughly by a factor of~1.5), and is numerically fairly
-expensive. See \citet{KoTr99,KoTr02a} for a discussion on the implementation
-of attenuation based upon standard linear solids. Please note that
-the Vp- and Vs-velocities of your model are given for a reference
-frequency. To change this reference frequency, you change the value
-of \texttt{ATTENUATION\_f0\_REFERENCE} in the main constants file
-\texttt{constants.h} found in subdirectory \texttt{src/shared/}.
-\item [{\texttt{FULL\_ATTENUATION\_SOLID}}] Set to \texttt{.true.} to add
-$Q_{\kappa}$ attenuation to the simulations in addition to $Q_{\mu}$
-attenuation when the \texttt{ATTENUATION} flag above is on. Should
-be off in almost all cases for seismic wave propagation, since $Q_{\kappa}$
-is almost always negligible in Earth models. However, this parameter
-can be useful for instance for ocean acoustics simulations, when $Q_{\kappa}$
-is not negligible in the solid layer at the sea bottom.
-\item [{\texttt{ANISOTROPY}}] Set to \texttt{.true.} if you want to use
-an anisotropy model. Please see the file \texttt{model\_aniso.f90}
-in subdirectory \texttt{src/generate\_databases/} for the current
-implementation of anisotropic models.
-\item [{\texttt{TOMOGRAPHY\_PATH}}] Directory in which the tomography files
-are stored for using external tomographic Earth models (please read
-Chapter \ref{cha:-Changing-the} and Section \ref{sec:Using-tomographic}
-`Using external tomographic Earth models' for further details.).
-\item [{\texttt{USE\_OLSEN\_ATTENUATION}}] Set to \texttt{.true.} if you
-want to use the attenuation model that scaled from the S-wave speed
-model using Olsen's empirical relation (see \citet{OlDaBr03}).
-\item [{\texttt{OLSEN\_ATTENUATION\_RATIO}}] Determines the Olsen's constant
-in Olsen's empirical relation (see \citet{OlDaBr03}).
-\item [{\texttt{PML\_CONDITIONS}}] Set to \texttt{.true.} to turn on C-PML
-boundary conditions for a regional simulation. Both fluids and elastic
-solids are supported. Note that xmeshfem3d generated meshes do not support C-PML yet.
-\item [{\texttt{PML\_INSTEAD\_OF\_FREE\_SURFACE}}] Set to \texttt{.true.}
-to turn on C-PML boundary conditions on the top surface instead of
-the usual free surface.
-\item [{\texttt{f0\_FOR\_PML}}] Determines the dominant frequency that
-will be used in the calculation of PML damping profiles; this should
-be set to the same (or similar) dominant frequency as that of the
-source that you will use in your simulation. If you plan to use a
-Dirac source, then use the dominant frequency of the source wavelet
-with which you plan to convolve your seismograms later on in post-processing.
-\item [{\texttt{STACEY\_ABSORBING\_CONDITIONS}}] Set to \texttt{.true.}
-to turn on Clayton-Enquist absorbing boundary conditions (see \citet{KoTr99}).
-In almost all cases it is much better to use CPML absorbing layers
-(see the options above) and leave this flag to \texttt{.false.}.
-\item [{\texttt{STACEY\_INSTEAD\_OF\_FREE\_SURFACE}}] Set to \texttt{.true.}
-to turn on absorbing boundary conditions on the top surface which
-by default constitutes a free surface of the model.
-\item [{\texttt{CREATE\_SHAKEMAP}}] Set this flag to \texttt{.true.} to
-create a ShakeMap\textregistered{}, i.e., a peak ground velocity map
-of the maximum absolute value of the two horizontal components of
-the velocity vector.
-\item [{\texttt{MOVIE\_SURFACE}}] Set to \texttt{.false.}, unless you want
-to create a movie of seismic wave propagation on the Earth's surface.
-Turning this option on generates large output files. See Section~\ref{sec:Movies}
-for a discussion on the generation of movies. This feature is only
-relevant for the solver.
-\item [{\texttt{MOVIE\_TYPE}}] Set this flag to 1 to show the top surface
-(tomography + oceans) only, to 2 to show all external faces of the
-mesh (i.e. topography + vertical edges + bottom) in shakemaps and
-surface movies.
-\item [{\texttt{MOVIE\_VOLUME}}] Set to \texttt{.false.}, unless you want
-to create a movie of seismic wave propagation in the Earth's interior.
-Turning this option on generates huge output files. See Section~\ref{sec:Movies}
-for a discussion on the generation of movies. This feature is only
-relevant for the solver.
-\item [{\texttt{SAVE\_DISPLACEMENT}}] Set this flag to \texttt{.true.}
-if you want to save the displacement instead of velocity for the movie
-frames.
-\item [{\texttt{USE\_HIGHRES\_FOR\_MOVIES}}] Set this flag to \texttt{.true.}
-if you want to save the values at all the NGLL grid points for the
-movie frames.
-\item [{\texttt{NTSTEP\_BETWEEN\_FRAMES}}] Determines the number of timesteps
-between movie frames. Typically you want to save a snapshot every
-100 timesteps. The smaller you make this number the more output will
-be generated! See Section~\ref{sec:Movies} for a discussion on the
-generation of movies. This feature is only relevant for the solver.
-\item [{\texttt{HDUR\_MOVIE}}] Determines the half duration of the source
-time function for the movie simulations. When this parameter is set
-to be 0, a default half duration that corresponds to the accuracy
-of the simulation is provided. Otherwise, it adds this half duration
-to the half duration specified in the source file \texttt{CMTSOLUTION},
-thus simulates longer periods to make the movie images look smoother.
-\item [{\texttt{SAVE\_MESH\_FILES}}] Set this flag to \texttt{.true.} to
-save ParaView \urlwithparentheses{www.paraview.org} mesh files for
-subsequent viewing. Turning the flag on generates large (distributed)
-files in the \texttt{LOCAL\_PATH} directory. See Section~\ref{sec:Mesh-graphics}
-for a discussion of mesh viewing features.
-\item [{\texttt{LOCAL\_PATH}}] Directory in which the distributed databases
-will be written. Generally one uses a directory on the local disk
-of the compute nodes, although on some machines these databases are
-written on a parallel (global) file system (see also the earlier discussion
-of the \texttt{LOCAL\_PATH\_IS\_ALSO\_GLOBAL} flag in Chapter~\ref{cha:Getting-Started}).
-\texttt{xgenerate\_databases} generates the necessary databases in
-parallel, one set for each of the \texttt{NPROC} slices that constitutes
-the mesh (see Figure~\ref{fig:mount.partitions} and Figure~\ref{fig:For-parallel-computing}).
-After the executable finishes, you can log in to one of the compute
-nodes and view the contents of the \texttt{LOCAL\_PATH} directory
-to see the (many) files generated by \texttt{xgenerate\_databases}.
-Please note that the \texttt{LOCAL\_PATH} directory should already
-contain the output files of the partitioner, i.e. from \texttt{xdecompose\_mesh}
-or \texttt{xmeshfem3D}.
-\item [{\texttt{NTSTEP\_BETWEEN\_OUTPUT\_INFO}}] This parameter specifies
-the interval at which basic information about a run is written to
-the file system (\texttt{timestamp{*}} files in the \texttt{OUTPUT\_FILES}
-directory). If you have access to a fast machine, set \texttt{NTSTEP\_BETWEEN\_OUTPUT\_INFO}
-to a relatively high value (e.g., at least 100, or even 1000 or more)
-to avoid writing output text files too often. This feature is not
-used at the time of meshing. One can set this parameter to a larger
-value than the number of time steps to avoid writing output during
-the run.
-\item [{\texttt{NTSTEP\_BETWEEN\_OUTPUT\_SEISMOS}}] This parameter specifies
-the interval at which synthetic seismograms are written in the \texttt{LOCAL\_PATH}
-directory. If a run crashes, you may still find usable (but shorter
-than requested) seismograms in this directory. On a fast machine set
-\texttt{NTSTEP\_BETWEEN\_OUTPUT\_SEISMOS} to a relatively high value
-to avoid writing to the seismograms too often. This feature is only
-relevant for the solver.
-\item [{\texttt{USE\_FORCE\_POINT\_SOURCE}}] Turn this flag on to use a
-(tilted) \texttt{FORCESOLUTION} force point source instead of a \texttt{CMTSOLUTION}
-moment-tensor source. When the force source does not fall exactly
-at a grid point, the solver interpolates the force between grid points
-using Lagrange interpolants. This can be useful e.g. for oil industry
-foothills simulations in which the source is a vertical force, normal
-force, tilted force, or an impact etc. Note that in the \texttt{FORCESOLUTION}
-file, you will need to edit the East, North and vertical components
-of an arbitrary (not necessarily unitary, the code will normalize it automatically) direction vector of the force vector;
-thus refer to Appendix~\ref{cha:Coordinates} for the orientation
-of the reference frame. This vector is made unitary internally in
-the solver and thus only its direction matters here; its norm is ignored
-and the norm of the force used is the factor force source times the
-source time function.
-When using this option, by default the code can locate the force source
-anywhere between mesh points in order to honor its exact location; this
-is more precise than using the closest GLL mesh point, but it is also a bit slower.
-If needed, you can change that default behavior and force the code to use the closest
-GLL mesh point instead by setting flag \texttt{USE\_BEST\_LOCATION} to \texttt{.false.} instead of \texttt{.true.} in file
-\texttt{src/shared/constants.h.in} and running the \texttt{configure} script again and recompiling the code.
-\item [{\texttt{USE\_RICKER\_TIME\_FUNCTION}}] Turn this flag on to use
-a Ricker source time function instead of the source time functions
-set by default to represent a (tilted) \texttt{FORCESOLUTION} force
-point source or a \texttt{CMTSOLUTION} moment-tensor source. Originally,
-if a \texttt{CMTSOLUTION} moment-tensor source is used, a (pseudo)
-Heaviside step function with a very short half duration is defined
-for elastic cases to represent the permanent slip on the fault while
-in the acoustic case a Gaussian source time function with a similarly
-short half duration is defined to physically describe actions within
-the fluid. Otherwise, if a \texttt{FORCESOLUTION} force source is
-used, a (pseudo) Dirac delta source time function is defined by default.
-Any other source-time function may then be obtained by convolution.
-\item [{\texttt{PRINT\_SOURCE\_TIME\_FUNCTION}}] Turn this flag on to print
-information about the source time function in the file \texttt{OUTPUT\_FILES/plot\_source\_time\_function.txt}.
-This feature is only relevant for the solver.
-
-\item [{\texttt{NUMBER\_OF\_SIMULTANEOUS\_RUNS}}] adds the ability to run several calculations (several earthquakes)
-in an embarrassingly-parallel fashion from within the same run;
-this can be useful when using a very large supercomputer to compute
-many earthquakes in a catalog, in which case it can be better from
-a batch job submission point of view to start fewer and much larger jobs,
-each of them computing several earthquakes in parallel.
-To turn that option on, set parameter NUMBER\_OF\_SIMULTANEOUS\_RUNS to a value greater than 1.
-To implement that, we create NUMBER\_OF\_SIMULTANEOUS\_RUNS MPI sub-communicators,
-each of them being labeled "my\_local\_mpi\_comm\_world", and we use them
-in all the routines in "src/shared/parallel.f90", except in MPI\_ABORT() because in that case
-we need to kill the entire run.
-When that option is on, of course the number of processor cores used to start
-the code in the batch system must be a multiple of NUMBER\_OF\_SIMULTANEOUS\_RUNS,
-all the individual runs must use the same number of processor cores,
-which as usual is NPROC in the Par\_file,
-and thus the total number of processor cores to request from the batch system
-should be NUMBER\_OF\_SIMULTANEOUS\_RUNS * NPROC.
-All the runs to perform must be placed in directories called run0001, run0002, run0003 and so on
-(with exactly four digits).
-
-Imagine you have 10 independent calculations to do, each of them on 100 cores; you have three options:
-
-1/ submit 10 jobs to the batch system
-
-2/ submit a single job on 1000 cores to the batch, and in that script create a sub-array of jobs to start 10 jobs,
-each running on 100 cores (see e.g. http://www.schedmd.com/slurmdocs/job\_array.html )
-
-3/ submit a single job on 1000 cores to the batch, start SPECFEM3D on 1000 cores, create 10 sub-communicators,
-cd into one of 10 subdirectories (called e.g. run0001, run0002,... run0010) depending on the sub-communicator
-your MPI rank belongs to, and run normally on 100 cores using that sub-communicator.
-
-The option NUMBER\_OF\_SIMULTANEOUS\_RUNS implements 3/.
-
-\item [{\texttt{BROADCAST\_SAME\_MESH\_AND\_MODEL}}]:
-if we perform simultaneous runs in parallel, if only the source and receivers vary between these runs
-but not the mesh nor the model (velocity and density) then we can also read the mesh and model files
-from a single run in the beginning and broadcast them to all the others; for a large number of simultaneous
-runs for instance when solving inverse problems iteratively this can DRASTICALLY reduce I/Os to disk in the solver
-(by a factor equal to NUMBER\_OF\_SIMULTANEOUS\_RUNS), and reducing I/Os is crucial in the case of huge runs.
-Thus, always set this option to .true. if the mesh and the model are the same for all simultaneous runs.
-In that case there is no need to duplicate the mesh and model file database (the content of the DATABASES\_MPI
-directories) in each of the run0001, run0002,... directories, it is sufficient to have one in run0001
-and the code will broadcast it to the others).
-
-TODO / future work to do: currently the BROADCAST\_SAME\_MESH\_AND\_MODEL option assumes to have the (master) mesh files in run0001/DATABASES\_MPI or run0001/OUTPUT\_FILES/DATABASES\_MPI.
-However, for adjoint runs you still need a DATABASES\_MPI folder in each of the sub-runs directories, e.g.
-run0002/DATABASES\_MPI, etc.
-to store the forward wavefields, kernels etc.
-of each sub-run.
-This would not be needed for forward simulations.
-
-TODO / future work to do: the sensitivity kernel summing and smoothing tools in directory src/tomography are currently not ported to this new option to do many runs simultaneously, only the solver (src/specfem3d) is.
-Thus these tools should work, but in their current version will need to be run for each simulation result independently.
-More precisely, the current kernel summing and smoothing routines work fine, with the exception that you need to move out the mesh files (and also the parameters).
-This works because these routines consider multiple runs by design.
-You simply have to provide them the directories where the kernels are.
-
-\item [{\texttt{GPU\_MODE}}] Turn this flag on to use GPUs.
-
-\item [\texttt{ADIOS\_ENABLED}] Turn this flag on to enable ADIOS. If set to \texttt{.false.}, subsequent ADIOS
-parameters will not be considered.
-\item [\texttt{ADIOS\_FOR\_DATABASES}] Turn this flag on to use ADIOS for xmeshfem3D output and
-xgenerate\_database input.
-\item [\texttt{ADIOS\_FOR\_MESH}]  Turn this flag on to use ADIOS for generated databases.
-\item [\texttt{ADIOS\_FOR\_FORWARD\_ARRAYS}] Turn this flag on to read and write forward arrays using ADIOS.
-\item [\texttt{ADIOS\_FOR\_KERNELS}] Turn this flag on to produce ADIOS kernels that can later be visualized with the ADIOS version of combine\_vol\_data.
-
-\end{description}
-
-The present version of SPECFEM can handle fully saturated porous simulations, Christina Morency implemented Biot equation.
-But the code cannot calculate partially saturated cases in its state.
-Christina Morency is presently working on a dual porosity, dual permeability formulation, type Pride and Berryman, for an other project, but it will not be available for some time.\\
-
-The way we prescribe material property for porous material in SPECFEM3D is as follow:
-We use a file name nummaterial\_poroelastic\_file, which is located in the directory MESH, the format is as follow:\\
- rhos rhof phi c kxx kxy kxz kyy kyz kzz Ks Kf Kfr etaf mufr\\
-where\\
-rho\_s = solid density\\
-rho\_f = fluid density\\
-phi = porosity\\
-tort = tortuosity\\
-kxx = xx component of permeability tensor\\
-kxy = xy,yx components of permeability tensor\\
-kyy = yy component of permeability tensor\\
-kxz = xz,zx components of permeability tensor\\
-kzz = zz component of permeability tensor\\
-kappa\_s = solid bulk modulus\\
-kappa\_f= fluid bulk modulus\\
-kappa\_fr= frame bulk modulus\\
-eta\_f = fluid viscosity\\
-mu\_fr = frame shear modulus.\\
-
-Using an external mesh (for instance coming from CUBIT/TRELIS), poroelastic materials have the ID number 3, while 1 is elastic, 2 is acoustic ?
-see the example in the package: examples/homogeneous\_poroelastic.
-
-If you use PML, the mesh elements that belong to the PML layers can
-be acoustic or elastic, but not viscoelastic nor poroelastic. Then,
-when defining your model, you should define these absorbing elements
-as either acoustic or elastic. In you forget to do that, the code
-will fix the problem by automatically converting the viscoelastic
-or poroelastic PML elements to elastic. This means that strictly speaking
-the PML layer will not be perfectly matched any more, since the physical
-model will change from viscoelastic or poroelastic to elastic at the
-entrance of the PML, but in practice this is sufficient and produces
-only tiny / negligible spurious reflections.
-
-If you use PML and an external mesh (created using an external meshing tool
-such as CUBIT/TRELIS or similar), try to have elements inside the PML as regular as possible,
-i.e. ideally non-deformed cubes obtained by `extrusion' of regular surface mesh elements meshing the
-outer edges of the computational domain without PML; by doing so, the PMLs obtained will be far more stable
-in time (PML being weakly unstable from a mathematical point of view, very deformed mesh elements
-inside the PMLs can trigger instabilities much more quickly).
-
-If you use PML and an external tomographic velocity and density model,
-you should be careful because mathematically a PML cannot handle heterogeneities
-along the normal to the PML edge inside the PML layer. This comes
-from the fact that the damping profile that is defined assumes a constant
-velocity and density model along the normal direction.
-
-Thus, you need to modify your velocity and density model in order
-for it to be 1D inside the PML, as shown in Figure~\ref{fig:modify_external_velocity_model_to_use_PML}.
-
-This applies to the bottom layer as well; there you should make sure
-that your model is 1D and thus constant along the vertical direction.
-
-To summarize, only use a 3D velocity and density model inside the
-physical region, and in all the PML layers extend it by continuity
-from its values along the inner PML edge.
-
-%%
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=3in]{figures/how_to_use_PML_when_a_tomographic_velocity_model_is_used}
-\par\end{centering}
-
-\caption{How to modify your external 3D velocity and density model in order
-to use PML. Such a modification is not needed when using Stacey absorbing
-boundary conditions (but such conditions are significantly less efficient).}
-
-\label{fig:modify_external_velocity_model_to_use_PML}
-\end{figure}
-
-%%
-
-\section{\label{sec:Choosing-the-Time-Step}Choosing the time step \texttt{DT}}
-
-The parameter \texttt{DT} sets the length of each time step in seconds.
-The value of this parameter is crucial for the stability of the spectral-element
-simulation. Your time step \texttt{DT} will depend on the minimum
-ratio between the distance $h$ of neighboring mesh points and the
-wave speeds $v$ defined in your model. The condition for the time
-step $\Delta t$ is:
-\begin{lyxcode}
-$\Delta t<C~\mathrm{min}_{\Omega}(~h/v~)$
-\end{lyxcode}
-where $C$ is the so-called Courant number and $\Omega$ denotes the
-model volume. The distance $h$ depends on the mesh element size and
-the number of GLL points \texttt{NGLL} specified in the main constants
-file \texttt{constants.h} located in the \texttt{src/shared/} subdirectory.
-The wave speed $v$ is determined based on your model's P- (or S-)
-wave speed values.
-
-The database generator \texttt{xgenerate\_databases}, as well as the
-internal mesher \texttt{xmeshfem3D}, are trying to evaluate the value
-of $\Delta t$ for empirically chosen Courant numbers $C\sim0.3$.
-If you used the mesher \texttt{xmeshfem3D} to generate your mesh,
-you should set the value suggested in \texttt{OUTPUT\_FILES/output\_mesher.txt}
-file, which is created after the mesher completed. In case you used
-CUBIT to create the mesh, you might use an arbitrary value when running
-\texttt{xgenerate\_databases} and then use the value suggested in
-the ~\\
- \texttt{OUTPUT\_FILES/output\_mesher.txt} file after the database
-generation completed. Note that the implemented Newmark time scheme
-uses this time step globally, thus your simulations become more expensive
-for very small mesh elements in high wave-speed regions. Please be
-aware of this restriction when constructing your mesh in Chapter~\ref{cha:Mesh-Generation}.
-
-\chapter{\label{cha:Running-the-Solver}Running the Solver \texttt{xspecfem3D}}
-
-Now that you have successfully generated the databases, you are ready
-to compile the solver. In the main directory, type
-\begin{lyxcode}
-{\small make~xspecfem3D}{\small \par}
-\end{lyxcode}
-Please note that \texttt{xspecfem3D} must be called directly from
-the main directory, as most of the binaries of the package.\\
-
-\noindent
-The solver needs three input files in the \texttt{DATA} directory
-to run:
-\begin{description}
-\item [{\texttt{Par\_file}}] the main parameter file which was discussed
-in detail in the previous Chapter~\ref{cha:Creating-Distributed-Databases},
-\item [{\texttt{CMTSOLUTION}} or {\texttt{FORCESOLUTION}}] the earthquake
-source parameter file or the force source parameter file, and
-\item [{\texttt{STATIONS}}] the stations file.
-\end{description}
-Most parameters in the \texttt{Par\_file} should be set prior to running
-the databases generation. Only the following parameters may be changed
-after running \texttt{xgenerate\_databases}:
-\begin{itemize}
-\item the simulation type control parameters: \texttt{SIMULATION\_TYPE}
-and \texttt{SAVE\_FORWARD}
-\item the time step parameters \texttt{NSTEP} and \texttt{DT}
-\item the absorbing boundary control parameter \texttt{PML\_CONDITIONS}
-on condition that the\\
- \texttt{PML\_INSTEAD\_OF\_FREE\_SURFACE} flag remains unmodified
-after running the databases generation.
-\item the movie control parameters \texttt{MOVIE\_SURFACE}, \texttt{MOVIE\_VOLUME},
-and \texttt{NTSTEPS\_BETWEEN\_FRAMES}
-\item the ShakeMap\textregistered{}option \texttt{CREATE\_SHAKEMAP}
-\item the output information parameters \texttt{MOVIE\_TYPE}, \texttt{NTSTEP\_BETWEEN\_OUTPUT\_INFO}
-and \texttt{NTSTEP\_BETWEEN\_OUTPUT\_}~\\
- \texttt{SEISMOS}
-\item the \texttt{PRINT\_SOURCE\_TIME\_FUNCTION} flags
-\end{itemize}
-Any other change to the \texttt{Par\_file} implies rerunning both
-the database generator \texttt{xgenerate\_databases} and the solver
-\texttt{xspecfem3D}.
-
-For any particular earthquake, the \texttt{CMTSOLUTION} file that
-represents the point source may be obtained directly from the Harvard
-Centroid-Moment Tensor (CMT) web page \urlwithparentheses{www.seismology.harvard.edu}.
-It looks like the example shown in Fig. \ref{fig:CMTSOLUTION-file}.
-%\begin{lyxcode}
-\begin{figure}[htp]
-\noindent
-\begin{centering}
-{\small \includegraphics[width=0.8\textwidth]{figures/Hollywood_CMT}
-}
-\par
-\end{centering}
-\caption{\texttt{CMTSOLUTION} file based on the
-format from the Harvard CMT catalog. \textbf{M} is the moment tensor,
-$M_{0}${\small {} }is the seismic moment, and $M_{w}$ is the moment
-magnitude.}
-\label{fig:CMTSOLUTION-file}
-\end{figure}
-%\end{lyxcode}
-The \texttt{CMTSOLUTION} file should be edited in the following way:
-\begin{itemize}
-\item Set the latitude or UTM $x$ coordinate, longitude or UTM $y$ coordinate, depth of the source (in km).
-\item Set the \texttt{time shift} parameter equal to $0.0$ (the solver
-will not run otherwise.) The time shift parameter would simply apply
-an overall time shift to the synthetics, something that can be done
-in the post-processing (see Section \ref{sec:Process-data-and-syn}).
-\item For point-source simulations (see finite sources, page \pageref{To-simulate-a})
-we recommend setting the source half-duration parameter \texttt{half
-duration} equal to zero, which corresponds to simulating a step source-time
-function, i.e., a moment-rate function that is a delta function. If
-\texttt{half duration} is not set to zero, the code will use a Gaussian
-(i.e., a signal with a shape similar to a `smoothed triangle', as
-explained in \citet{KoTr02a} and shown in Fig~\ref{fig:gauss.vs.triangle})
-source-time function with half-width \texttt{half duration}. We prefer
-to run the solver with \texttt{half duration} set to zero and convolve
-the resulting synthetic seismograms in post-processing after the run,
-because this way it is easy to use a variety of source-time functions
-(see Section \ref{sec:Process-data-and-syn}). \citet{KoTr02a} determined
-that the noise generated in the simulation by using a step source
-time function may be safely filtered out afterward based upon a convolution
-with the desired source time function and/or low-pass filtering. Use
-the serial code \texttt{convolve\_source\_timefunction.f90} and the
-script \texttt{convolve\_source\_timefunction.csh} for this purpose,
-or alternatively use signal-processing software packages such as SAC
-\urlwithparentheses{www.llnl.gov/sac}. Type
-
-\begin{lyxcode}
-{\small make~xconvolve\_source\_timefunction}{\small \par}
-\end{lyxcode}
-
-to compile the code and then set the parameter \texttt{hdur} in \texttt{convolve\_source\_timefunction.csh}
-to the desired half-duration.
-
-\item The zero time of the simulation corresponds to the center of the triangle/Gaussian,
-or the centroid time of the earthquake. The start time of the simulation
-is $t=-1.5*\texttt{half duration}$ (the 1.5 is to make sure the moment
-rate function is very close to zero when starting the simulation).
-To convert to absolute time $t_{\mathrm{abs}}$, set
-
-\begin{lyxcode}
-$t_{\mathrm{abs}}=t_{\mathrm{pde}}+\texttt{time shift}+t_{\mathrm{synthetic}}$
-\end{lyxcode}
-
-where $t_{\mathrm{pde}}$ is the time given in the first line of the
-\texttt{CMTSOLUTION}, \texttt{time shift} is the corresponding value
-from the original \texttt{CMTSOLUTION} file and $t_{\mathrm{synthetic}}$
-is the time in the first column of the output seismogram.
-
-\end{itemize}
-\begin{figure}[htp]
-\noindent \begin{centering}
-\includegraphics[width=3.5in]{figures/gauss_vs_triangle_mod}
-\par\end{centering}
-
-\caption{Comparison of the shape of a triangle and the Gaussian function actually
-used.}
-
-\label{fig:gauss.vs.triangle}
-\end{figure}
-
-If you know the earthquake source in strike/dip/rake format rather
-than in \texttt{CMTSOLUTION} format, use the C code \texttt{SPECFEM3D\_GLOBE/utils/strike\_dip\_rake\_to\_CMTSOLUTION.c}
-to convert it. The conversion formulas are given for instance in \citet{AkRi80}.
-Note that the \citet{AkRi80} convention is slightly different from
-the Harvard \texttt{CMTSOLUTION} convention (the sign of some components
-is different). The C code outputs both.
-
-Centroid latitude and longitude should be provided in geographical
-coordinates. The code converts these coordinates to geocentric coordinates~\citep{DaTr98}.
-Of course you may provide your own source representations by designing
-your own \texttt{CMTSOLUTION} file. Just make sure that the resulting
-file adheres to the Harvard CMT conventions (see Appendix~\ref{cha:Coordinates}).
-Note that the first line in the \texttt{CMTSOLUTION} file is the Preliminary
-Determination of Earthquakes (PDE) solution performed by the USGS
-NEIC, which is used as a seed for the Harvard CMT inversion. The PDE
-solution is based upon P waves and often gives the hypocenter of the
-earthquake, i.e., the rupture initiation point, whereas the CMT solution
-gives the `centroid location', which is the location with dominant
-moment release. The PDE solution is not used by our software package
-but must be present anyway in the first line of the file.
-
-\label{To-simulate-a}To simulate a kinematic rupture, i.e., a finite-source
-event, represented in terms of $N_{\mathrm{sources}}$ point sources,
-provide a \texttt{CMTSOLUTION} file that has $N_{\mathrm{sources}}$
-entries, one for each subevent (i.e., concatenate $N_{\mathrm{sources}}$
-\texttt{CMTSOLUTION} files to a single \texttt{CMTSOLUTION} file).
-At least one entry (not necessarily the first) must have a zero \texttt{time
-shift}, and all the other entries must have non-negative \texttt{time
-shift}. Each subevent can have its own half duration, latitude, longitude,
-depth, and moment tensor (effectively, the local moment-density tensor).
-
-Note that the zero in the synthetics does NOT represent the hypocentral
-time or centroid time in general, but the timing of the \textit{center}
-of the source triangle with zero \texttt{time shift} (Fig~\ref{fig:source_timing}).
-
-Although it is convenient to think of each source as a triangle, in
-the simulation they are actually Gaussians (as they have better frequency
-characteristics). The relationship between the triangle and the gaussian
-used is shown in Fig~\ref{fig:gauss.vs.triangle}. For finite fault
-simulations it is usually not advisable to use a zero half duration
-and convolve afterwards, since the half duration is generally fixed
-by the finite fault model.
-
-\vspace{1cm}
-
-\noindent The \texttt{FORCESOLUTION} file should be edited in the
-following way:
-\begin{itemize}
-\item Set the \texttt{time shift} parameter equal to $0.0$ (the solver
-will not run otherwise.) The time shift parameter would simply apply
-an overall time shift to the synthetics, something that can be done
-in the post-processing (see Section \ref{sec:Process-data-and-syn}).
-\item Set the \texttt{f0} parameter (the dominant frequency) of the Ricker source time
-function when {\texttt{USE\_RICKER\_TIME\_FUNCTION}} is turned on
-in the main parameter file {\texttt{Par\_file}}. In case that the
-solver uses a (pseudo) Dirac delta source time function to represent
-a force point source, a very short duration of five time steps is automatically set by default.
-\item Set the latitude or UTM $x$ coordinate, longitude or UTM $y$ coordinate, depth of the source (in km).
-\item Set the magnitude of the force source.
-\item Set the components of a (non-unitary) direction vector for the force
-source in the East/North/Vertical basis (see Appendix A for the orientation
-of the reference frame).
-\end{itemize}
-\noindent Where necessary, set a \texttt{FORCESOLUTION} file in the
-same way you configure a \texttt{CMTSOLUTION} file with $N_{\mathrm{sources}}$
-entries, one for each subevent (i.e., concatenate $N_{\mathrm{sources}}$
-\texttt{FORCESOLUTION} files to a single \texttt{FORCESOLUTION} file).
-At least one entry (not necessarily the first) must have a zero \texttt{time
-shift}, and all the other entries must have non-negative \texttt{time
-shift}. Each subevent can have its own half latitude, longitude, depth,
-\texttt{half duration} and force parameters.
-
-\begin{figure}[H]
-\noindent \begin{centering}
-{\small \includegraphics[width=5in]{figures/source_timing} }
-\par\end{centering}
-
-\caption{Example of timing for three sources. The center of the first source
-triangle is defined to be time zero. Note that this is NOT in general
-the hypocentral time, or the start time of the source (marked as tstart).
-The parameter \texttt{time shift} in the \texttt{CMTSOLUTION} file
-would be t1(=0), t2, t3 in this case, and the parameter \texttt{half
-duration} would be hdur1, hdur2, hdur3 for the sources 1, 2, 3 respectively.}
-
-{\small \label{fig:source_timing} }
-\end{figure}
-
-The solver can calculate seismograms at any number of stations for
-basically the same numerical cost, so the user is encouraged to include
-as many stations as conceivably useful in the \texttt{STATIONS} file,
-which looks like this:
-
-\begin{figure}[H]
-\noindent \begin{centering}
-{\small \includegraphics{figures/STATIONS_basin_explained} }
-\par\end{centering}
-
-\label{fig:Sample-STATIONS-file.}\caption{Sample \texttt{STATIONS} file. Station latitude and longitude should
-be provided in geographical coordinates. The width of the station
-label should be no more than 32 characters (see \texttt{MAX\_LENGTH\_STATION\_NAME}
-in the \texttt{constants.h} file), and the network label should be
-no more than 8 characters (see \texttt{MAX\_LENGTH\_NETWORK\_NAME}
-in the \texttt{constants.h} file).}
-\end{figure}
-
-Each line represents one station in the following format:
-\begin{lyxcode}
-{\small Station~Network~Latitude~(degrees)~Longitude~(degrees)~Elevation~(m)~burial~(m)~}{\small \par}
-\end{lyxcode}
-The solver \texttt{xspecfem3D} filters the list of stations in file
-\texttt{DATA/STATIONS} to exclude stations that are not located within
-the region given in the \texttt{Par\_file} (between \texttt{LATITUDE\_MIN}
-and \texttt{LATITUDE\_MAX} and between \texttt{LONGITUDE\_MIN} and
-\texttt{LONGITUDE\_MAX}). The filtered file is called \texttt{DATA/STATIONS\_FILTERED}.
-
-Solver output is provided in the \texttt{OUTPUT\_FILES} directory
-in the \texttt{output\_solver.txt} file. Output can be directed to
-the screen instead by uncommenting a line in \texttt{constants.h}:
-\begin{lyxcode}
-!~uncomment~this~to~write~messages~to~the~screen~
-
-!~integer,~parameter~::~IMAIN~=~ISTANDARD\_OUTPUT~
-\end{lyxcode}
-On PC clusters the seismogram files are generally written to the local
-disks (the path \texttt{LOCAL\_PATH} in the \texttt{Par\_file}) and
-need to be gathered at the end of the simulation.
-
-While the solver is running, its progress may be tracked by monitoring
-the `\texttt{\small timestamp{*}}{\small ' files in the ~}\\
-{\small{} }\texttt{\small OUTPUT\_FILES}{\small{} directory. These tiny
-files look something like this:}{\small \par}
-\begin{lyxcode}
-Time~step~\#~~~~~~~~~~10000~
-
-Time:~~~~~108.4890~~~~~~seconds~
-
-Elapsed~time~in~seconds~=~~~~1153.28696703911~
-
-Elapsed~time~in~hh:mm:ss~=~~~~~0~h~19~m~13~s~
-
-Mean~elapsed~time~per~time~step~in~seconds~=~~~~~0.115328696703911~
-
-Max~norm~displacement~vector~U~in~all~slices~(m)~=~~~~1.0789589E-02~
-\end{lyxcode}
-The \texttt{\small timestamp{*}}{\small{} files provide the }\texttt{\small Mean
-elapsed time per time step in seconds}{\small , which may be used
-to assess performance on various machines (assuming you are the only
-user on a node), as well as the }\texttt{\small Max norm displacement
-vector U in all slices~(m)}{\small . If something is wrong with the
-model, the mesh, or the source, you will see the code become unstable
-through exponentially growing values of the displacement and fluid
-potential with time, and ultimately the run will be terminated by
-the program. You can control the rate at which the timestamp files
-are written based upon the parameter }\texttt{\small NTSTEP\_BETWEEN\_OUTPUT\_INFO}{\small{}
-in the }\texttt{\small Par\_file}{\small .}{\small \par}
-
-Having set the \texttt{Par\_file} parameters, and having provided
-the \texttt{CMTSOLUTION} (or the \texttt{FORCESOLUTION}) and \texttt{STATIONS}
-files, you are now ready to launch the solver! This is most easily
-accomplished based upon the \texttt{go\_solver} script (See Chapter
-\ref{cha:Scheduler} for information about running through a scheduler,
-e.g., LSF). You may need to edit the last command at the end of the
-script that invokes the \texttt{mpirun} command. The \texttt{runall}
-script compiles and runs both \texttt{xgenerate\_databases} and \texttt{xspecfem3D}
-in sequence. This is a safe approach that ensures using the correct
-combination of distributed database output and solver input.
-
-It is important to realize that the CPU and memory requirements of
-the solver are closely tied to choices about attenuation (\texttt{ATTENUATION})
-and the nature of the model (i.e., isotropic models are cheaper than
-anisotropic models). We encourage you to run a variety of simulations
-with various flags turned on or off to develop a sense for what is
-involved.
-
-For the same model, one can rerun the solver for different events
-by simply changing the \texttt{CMTSOLUTION} or \texttt{FORCESOLUTION}
-file, or for different stations by changing the \texttt{STATIONS}
-file. There is no need to rerun the \texttt{xgenerate\_databases}
-executable. Of course it is best to include as many stations as possible,
-since this does not add to the cost of the simulation.
-
-We have also added the ability to run several calculations (several earthquakes)
-in an embarrassingly-parallel fashion from within the same run;
-this can be useful when using a very large supercomputer to compute
-many earthquakes in a catalog, in which case it can be better from
-a batch job submission point of view to start fewer and much larger jobs,
-each of them computing several earthquakes in parallel.
-To turn that option on, set parameter \texttt{NUMBER\_OF\_SIMULTANEOUS\_RUNS}
-to a value greater than 1 in file setup/constants.h.in before
-configuring and compiling the code.
-When that option is on, of course the number of processor cores used to start
-the code in the batch system must be a multiple of \texttt{NUMBER\_OF\_SIMULTANEOUS\_RUNS},
-all the individual runs must use the same number of processor cores,
-which as usual is \texttt{NPROC} in the input file \texttt{DATA/Par\_file},
-and thus the total number of processor cores to request from the batch system
-should be \texttt{NUMBER\_OF\_SIMULTANEOUS\_RUNS $\times$ NPROC}.
-All the runs to perform must be placed in directories called \texttt{run0001}, \texttt{run0002}, \texttt{run0003} and so on (with exactly four digits)
-and you must create a link from the root directory of the code
-to the first copy of the executable programs by typing \texttt{ln -s run0001/bin bin}.
-
-\chapter{\label{cha:Fault-Kinematics-Dynamics}Kinematic and dynamic fault
-sources}
-
-SPECFEM3D can handle finite fault sources of two kinds:
-\begin{enumerate}
-\item \emph{Kinematic}: the spatio-temporal distribution of slip rate is
-prescribed along the fault surface
-\item \emph{Dynamic}: a friction law and initial stresses are prescribed
-on the fault, a spontaneous rupture process is computed.
-\end{enumerate}
-
-\section{\label{sec:Mesh-Generation-with}Mesh Generation with Split Nodes}
-
-Faults need to be handled in a special way during mesh generation.
-A fault surface must lie at the interface between elements (the mesh
-must honor the fault surfaces). Moreover, a fault is made of two surfaces
-in contact. Each of these two surfaces needs a separate set of nodes.
-This approach is known as \textquotedbl{}split nodes\textquotedbl{}.
-Currently faults can only be run with \texttt{xdecompose\_mesh} and
-\texttt{CUBIT}. \texttt{xmeshfem3D} is not yet ready to handle faults.\\
-To facilitate the mesh generation with split nodes in CUBIT, we need
-to separate the two fault surfaces by a small distance, effectively
-creating a tiny opening of the fault (Figure \ref{fig:examples.splitnodes},
-\ref{fig:examples.splitnodes-surfacetrace}). Note that only the interior
-of the fault must be opened, its edges must remain closed (except
-the edge on the free surface). The fault is automatically closed later
-by SPECFEM3D.\\
-
-Here is an example CUBIT script to generate a mesh with split nodes
-for a buried vertical strike-slip fault:
-\begin{lyxcode}
-~reset~~\\
-~brick~x~10~y~10~z~10~~\\
-~webcut~volume~all~with~plane~xplane~~\\
-~webcut~volume~all~with~plane~yplane~~\\
-~webcut~volume~all~with~plane~xplane~offset~3~~\\
-~webcut~volume~all~with~plane~zplane~offset~3~~\\
-~webcut~volume~all~with~plane~zplane~offset~-3~~\\
-~imprint~all~~\\
-~merge~all~~\\
-~unmerge~surf~160~~\\
-~mesh~vol~all~~\\
-~set~node~constraint~off~~\\
-~node~in~surf~168~move~X~0~Y~0.01~Z~0~~\\
-~node~in~surf~160~move~X~0~Y~-0.01~Z~0~\\
-
-\end{lyxcode}
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[scale=0.3]{figures/faultmesh} \\
- \includegraphics[scale=0.3]{figures/surf168} \includegraphics[scale=0.3]{figures/splitnodes}
-\par\end{centering}
-
-\caption{Screenshots of the CUBIT example for the embedded fault with split
-nodes: The entire mesh is decomposed into several volumes bounding
-the fault (top), the fault surface 168 is shown as orange squares
-(middle) and split nodes of the fault shown as orange dots(bottom).
-Note that only interior nodes of the fault are split while those on
-the edges of the fault surface are touching each other.}
-
-\label{fig:examples.splitnodes}
-\end{figure}
-
-\begin{lyxcode}
-
-\end{lyxcode}
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[scale=0.55]{figures/splitnodes_surfacetrace} \\
-
-\par\end{centering}
-
-\caption{Screenshots of a CUBIT example showing split nodes for a fault reaching
-the surface. Surface trace of the fault is shown in orange. Note that
-the edges of the fault are not split while the interior nodes are
-offset by a small distance on either side of the fault}
-
-\label{fig:examples.splitnodes-surfacetrace}
-\end{figure}
-
-The CUBIT scripts ({*}.jou and {*}.py) in the directory \texttt{EXAMPLES}
-generate more complicated meshes. The {*}.py files are Python scripts
-that execute CUBIT commands and use the CUBIT-python interface for
-SPECFEM3D (see next section). The Python language allows to define
-and manipulate variables to parameterize the mesh. Alternatively,
-the Python script can call a CUBIT journal file ({*}.jou), which looks
-like the example above. Variables can be defined and manipulated there
-using the $\mathtt{APREPRO}$ language built in CUBIT.\\
-
-Note that you should avoid gaps in the list of indices of mesh objects
-with the following CUBIT command:
-\begin{lyxcode}
-$\mathtt{compress\: ids\: hex\: face\: edge\: node}$
-\end{lyxcode}
-(otherwise you will get a segmentation fault during domain decomposition)
-
-\section{CUBIT-Python Scripts for Faults}
-
-The mesh generated in CUBIT needs to be processed and exported in
-a format compatible with SPECFEM3D. This is achieved in the Python
-scripts by calling the Python-CUBIT interface functions defined in
-the CUBIT directory:
-\begin{enumerate}
-\item Function $\mathtt{define\_bc}$ (or $\mathtt{boundary\_definition.py}$)
-must be called to set up the absorbing boundaries database
-\item Function $\mathtt{fault\_input}$ must be called once for each fault
-to set up the fault database
-\item Function $\mathtt{cubit2specfem3d.export2SPECFEM3D}$ must be called
-at the very end of the script to export the mesh in a SPECFEM3D format.
-\end{enumerate}
-The functions in \#1 and \#3 are for the bulk and are documented in
-Section \ref{subsec:Exporting-the-Mesh}. We focus here on \#2:
-\begin{description}
-\item [{Function:}] $\mathtt{fault\_input}$
-\item [{Purpose:}] export a fault mesh database from CUBIT to a SPECFEM3D-compliant
-file
-\item [{Syntax:}] fault\_input(id\_fault, ids\_surf\_1, ids\_surf\_2)
-\item [{Inputs:}]~\end{description}
-\begin{itemize}
-\item id\_fault integer index assigned to the fault. (The user must number
-all the faults, starting at 1, with unit increments)
-\item ids\_surf\_1 list of CUBIT ids of all surfaces that form side 1 of
-the fault.
-\item ids\_surf\_2 list of CUBIT ids of all surfaces that form side 2 of
-the fault. (The user must decide which side of the fault is side 1.
-This choice affects the sign conventions of fault quantities as explained
-in Section \ref{sec:Sign-Convention-for}).\end{itemize}
-\begin{description}
-\item [{Outputs:}] file $\mathtt{fault\_file\_X.dat}$, where X is the
-fault id (id\_fault).
-\item [{Example:}] For the example in Section \ref{sec:Mesh-Generation-with}:\end{description}
-\begin{lyxcode}
-$\hphantom{}$A1~=~{[}168{]}
-
-$\hphantom{}$A2~=~{[}160{]}
-
-$\hphantom{}$fault\_input(1,A1,A2)
-\end{lyxcode}
-
-\section{Examples}
-
-The package includes examples, the SCEC benchmark problems:
-\begin{itemize}
-\item TPV5, a planar vertical strike-slip fault
-\item TPV14 and TPV15, vertical strike-slip fault system with a fault branch
-\item TPV102 and TPV103, rate-and-state benchmarks
-\item TPV22 and TPV23, step-over benchmarks
-\item TPV16, heterogenous initial stress
-\end{itemize}
-and
-\begin{itemize}
-\item Splay fault models from Wendt et al. (2009)\\
-
-\end{itemize}
-Read the documents in the directory \texttt{EXAMPLES/*/description}.
-They contain a description of the example and additional instructions
-to run it. Visualize the results with the Matlab scripts in the directory
-$\mathtt{EXAMPLES/*/post}$
-
-\section{\label{sec:Sign-Convention-for}Sign Convention for Fault Quantities}
-
-During mesh generation, the fault is defined by two surfaces in contact.
-Let's denote as \textquotedbl{}\emph{side 1}\textquotedbl{} the SECOND
-surface declared by the user in the call to the python function \textquotedbl{}\emph{fault\_input}\textquotedbl{},
-and the FIRST surface as \textquotedbl{}\emph{side 2}\textquotedbl{}.
-The local coordinate system on the fault is defined as the right-handed
-coordinate system defined by (strike, dip, normal), where \textquotedbl{}\emph{normal}\textquotedbl{}
-is the normal vector outgoing from side 1, \textquotedbl{}\emph{dip}\textquotedbl{}
-is parallel to the along-dip direction pointing downwards, and \textquotedbl{}\emph{strike}\textquotedbl{}
-is the horizontal along-strike vector such that the system is right-handed.\\
-
-Slip is defined as displacement on side 2 minus displacement on side
-1. In the local coordinate system on the fault, positive along-strike
-slip is right-lateral and positive along-dip slip is thrust if side
-1 is on the hanging wall (normal faulting if side 1 is on the foot
-wall).\\
-
-Traction is defined as the stress induced on side 1 by side 2, which
-is the stress tensor times the normal vector outgoing from side 1.
-In the local coordinate system on the fault, the normal traction is
-negative in compression, positive along-strike traction generates
-right-lateral slip and positive along-dip traction generates thrust
-slip if side 1 is on the hanging wall (normal faulting if side 1 is
-on the foot wall).
-
-\section{Input Files}
-
-Three additional input files are required in the \texttt{DATA/} directory
-for dynamic and kinematic faults. They are $\mathtt{Par\_file\_faults}$,
-$\mathtt{FAULT\_STATIONS}$ and $\mathtt{input\_file.txt}$ (optional).
-If the former does not exist, the code assumes that there are no faults.\\
-
-\begin{lyxlist}{00.00.0000}
-\item [{\textbf{DATA/Par\_file\_faults}}] contains parameters of the fault.
-The first part of this file has a strict format:
-
-\begin{description}
-\item [{Line}] 1: Number of faults (NF)
-\item [{Lines}] 2 to NF+1: Kelvin Voigt damping (in seconds) for each fault.
-(See below how to set this parameter)
-\item [{Line}] NF+2: Type of simulation (1=dynamic , 2 = kinematic)
-\item [{Line}] NF+3: Number of time steps between updates of the time series
-outputs at selected fault points (see DATA/FAULT\_STATIONS), usually
-a large number (100s or 1000s). Note that the sampling rate of the
-time series is usually much higher.
-\item [{Line}] NF+4: Number of time steps between fault snapshot outputs
-(quantities at every fault point exported at regular times), usually
-a large number (100s or 1000s).
-\item [{Line}] NF+5: Slip velocity threshold below which frictional healing
-is set (friction coefficient is reset to its static value). If this
-value is negative healing is disabled.
-\item [{Line}] NF+6: Slip velocity threshold to define the rupture front.
-Only used for outputs.
-\end{description}
-
-The rest of this file is made of namelist input blocks (see \textquotedbl{}namelist\textquotedbl{}
-in a Fortran 9x manual). The input for each fault has the following
-sequence (arguments in {[}brackets{]} are optional):
-
-\&\textbf{BEGIN\_FAULT} /
-
-\&\textbf{INIT\_STRESS} S1, S2, S3 {[},n1, n2, n3{]} /
-
-followed by (n1+n2+n3) \&\textbf{DIST2D} blocks
-
-\&\textbf{SWF} mus, mud, dc {[}, nmus, nmud, ndc{]} /
-
-followed by (nmus+nmud+ndc) \&\textbf{DIST2D} blocks\\
-
-\begin{lyxlist}{00.00.0000}
-\item [{\&\textbf{INIT\_STRESS}}] input block sets the initial fault stresses
-relative to the foot-wall side of the fault. Initial stresses are
-composed of a constant background value possibly overwritten in prescribed
-regions by heterogeneous distributions (see \&\textbf{DIST2D} blocks
-below):
-
-\begin{description}
-\item [{S1}] = constant background value of along-strike shear stress (positive
-in the usual strike direction)
-\item [{S2}] = constant background value of along-dip shear (positive is
-down-dip, normal faulting)
-\item [{S3}] = constant background value of normal stress (negative in
-compression)
-\item [{n1}] = number of heterogeneous items for along-strike shear stress
-{[}default is 0{]}
-\item [{n2}] = number of heterogeneous items for along-dip shear stress
-{[}default is 0{]}
-\item [{n3}] = number of heterogeneous items for normal stress {[}default
-is 0{]}
-\end{description}
-\item [{\&\textbf{SWF}}] input block sets the slip-weakening friction parameters
-of the fault:
-
-\begin{description}
-\item [{mus}] = constant background value of static friction coefficient
-\item [{mud}] = constant background value of dynamic friction coefficient
-\item [{dc}] = constant background value of critical slip-weakening distance
-\item [{nmus}] = number of heterogeneous items for static friction coefficient
-{[}default is 0{]}
-\item [{nmud}] = number of heterogeneous items for dynamic friction coefficient
-{[}default is 0{]}
-\item [{ndc}] = number of heterogeneous items for critical slip-weakening
-distance {[}default is 0{]}\\
-
-\end{description}
-\item [{\&\textbf{DIST2D}}] input blocks modify (overwrite) the value of
-a fault parameter by a heterogeneous spatial distribution:
-
-\&\textbf{DIST2D} shape='$\mathtt{square}$', val, xc, yc, zc, l /
-
-$\;$sets a constant value (val) within a cube with center (xc,yc,zc)
-and edge size l.
-
-\&\textbf{DIST2D} shape='$\mathtt{rectangle}$', val, xc, yc, zc,
-lx, ly, lz /
-
-$\;$sets a constant value (val) within a rectangular prism with center
-(xc,yc,zc) and edge sizes (lx,ly,lz).
-
-\&\textbf{DIST2D} shape='$\mathtt{rectangle-taper}$', val, valh,
-xc, yc, zc, lx, ly, lz /
-
-$\;$sets a vertical linear gradient within a rectangular prism with
-center (xc,yc,zc) and edge sizes (lx,ly,lz). Values vary linearly
-as a function of vertical position z between value val at z = zc-lz/2
-and value valh at z = zc+lz/2 .
-
-\&\textbf{DIST2D} shape='$\mathtt{circular}$', val, xc, yc, zc, r
-/
-
-$\;$sets a constant value (val) within a sphere with center (xc,yc,zc)
-and radius r. \\
-
-\end{lyxlist}
-\end{lyxlist}
-
-\begin{lyxlist}{00.00.0000}
-\item [{\textbf{DATA/FAULT\_STATIONS}}] Stations in the fault plane.
-
-\textbf{Line 1} : number of stations.
-
-\textbf{Line 2 to end}: 5 columns: X, Y, Z (-depth), station name,
-fault-id
-
-The fault-id identifies the fault that contains the station. It is
-the index of appearance in the faults list after line 2 of Par\_file\_faults
-
-\end{lyxlist}
-
-\begin{lyxlist}{00.00.0000}
-\item [{\textbf{DATA/input\_file.txt}}] Heterogeneous stresses and friction
-parameters are documented in page 10 of
-
-\begin{lyxlist}{00.00.0000}
-\item [{$\mathtt{EXAMPLES/tpv16/description/TPV16\_17\_Description\_v03.pdf}$}]~
-\end{lyxlist}
-
-To activate this feature, in fault\_solver\_dynamic.f90 set \textbf{TPV16}
-= .true. then re-compile the code.
-
-\end{lyxlist}
-
-\section{\label{sec:Setting-the-Kelvin-Voigt}Setting the Kelvin-Voigt Damping
-Parameter }
-
-The purpose of the Kelvin-Voigt viscosity in the dynamic fault solver
-is to damp spurious oscillations generated by the fault slip at frequencies
-that are too high to be resolved by the mesh. The viscosity $\mathtt{eta}$
-(in seconds) depends on the size of the elements on the fault. Here
-is how to set it:
-\begin{enumerate}
-\item Determine the average linear size of the elements on the fault plane,
-$\mathtt{h\_fault}$. Usually this value is prescribed by the user
-during mesh generation. Otherwise it can be found by inspection of
-the mesh inside the CUBIT GUI.
-\item Use the Matlab function $\mathtt{utils/critical\_timestep.m}$ to
-compute\\
- $\mathtt{dtc\_fault}=\mathtt{critical\_timestep\left(c_{p},h\_fault,ngll\right)}$.
-\\
-This is the critical time step in an elastic medium for a hypothetical
-element of cubic shape with size equal to $\mathtt{h\_fault}$.
-\item Set $\mathtt{eta}$ in $\mathtt{Par\_file\_faults}$ to (0.1 to 0.3)$\ensuremath{\times}\mathtt{dtc\_fault}$.
-A larger $\mathtt{eta}$ damps high-frequencies more aggressively
-but it might also affect lower frequencies and rupture speed.
-\end{enumerate}
-Viscosity reduces numerical stability: the critical timestep in a
-simulation with Kelvin-Voigt damping needs to be smaller than that
-in a purely elastic simulation. Here is how to set the time step accordingly:
-\begin{enumerate}
-\item Run a test simulation without viscosity ($\mathtt{eta}$=0 and only
-a few time steps)
-\item Look for the \textquotedbl{}maximum suggested time step\textquotedbl{}
-in \texttt{OUTPUT\_FILES/output\_mesher.txt}. This is the critical
-timestep of a purely elastic simulation, $\mathtt{dtc\_bulk}$.
-\item Reset the timestep of the simulation with a Kelvin-Voigt material
-to a value smaller than\\
- $\mathtt{dtc\_kv}=\mathtt{eta}\left(\sqrt{1+\mathtt{dtc\_bulk^{2}}/\mathtt{eta^{2}}}-1\right)$
-\end{enumerate}
-Note that in general $\mathtt{dtc\_bulk}$ is smaller than $\mathtt{dtc\_fault}$,
-because elements off the fault might be smaller or more distorted
-than element faces on the fault.
-
-\section{Output Files }
-
-Several output files are saved in \texttt{OUTPUT\_FILES/}:\\
-
-\begin{enumerate}
-\item Seismograms for each station on the fault plane given in \texttt{DATA/FAUL\_STATIONS}.
-One output file is generated for each station, named after the station.
-The files are ascii and start with a header (22 lines long) followed
-by a data block with the following format, one line per time sample:\\
-
-Column 1 = Time (s)
-
-Column 2 = horizontal right-lateral slip (m)
-
-Column 3 = horizontal right-lateral slip rate (m/s)
-
-Column 4 = horizontal right-lateral shear stress (MPa)
-
-Column 5 = vertical up-dip slip (m)
-
-Column 6 = vertical up-dip slip rate (m/s)
-
-Column 7 = vertical up-dip shear stress (MPa)
-
-Column 8 = normal stress (MPa)\\
-
-The stresses are relative to the footwall side of the fault (this
-convention controls their sign, but not their amplitude). Slip is
-defined as displacement of the hanging wall relative to the footwall.
-
-\item Seismograms at stations in the bulk (out of the fault plane) given
-in \texttt{DATA/STATIONS}.
-\item Rupture time files are named $\mathtt{Rupture\_time\_FAULT}$-$\mathtt{id}$.
-One file is generated for each fault. The files are ascii and start
-with a header (12 lines long) followed by a data block with the following
-format, one line per fault node:
-
-Column 1 = horizontal coordinate, distance along strike (m)
-
-Column 2 = vertical coordinate, distance down-dip (m)
-
-Column 3 = rupture time (s)
-
-\item Fault quantities (slip, slip rate, stresses, etc) at regular times
-are stored in binary data files called $\mathtt{Snapshot\#it\#.bin}$,
-where \#it\# is the timestep number. These can be read in Matlab with
-the function $\mathtt{utils/FSEM3D\_snapshot.m}$
-\end{enumerate}
-
-\section{Post-processing and Visualization }
-
-Some Matlab functions for post-processing and visualization are included
-in directory $\mathtt{utils}$. The functions are internally documented
-(see their matlab help).\\
-
-\texttt{FSEM3D\_snapshot} reads a fault data snapshot\\
-
-The directories \texttt{EXAMPLES/*/post} contain additional Matlab
-scripts to generate figures specific to each example.
-
-\chapter{\label{cha:Adjoint-Simulations}Adjoint Simulations}
+\chapter{Adjoint Simulations}\label{cha:Adjoint-Simulations}
 
@@
 
-\section{\label{sec:Adjoint-simulation-sources}Adjoint Simulations for Sources
-Only (not for the Model)}
+\section{Adjoint Simulations for Sources Only (not for the Model)}\label{sec:Adjoint-simulation-sources}
 
@@
 
-\section{\label{sec:Adjoint-simulation-finite}Adjoint Simulations for Finite-Frequency
-Kernels (Kernel Simulation)}
+\section{Adjoint Simulations for Finite-Frequency Kernels (Kernel Simulation)}\label{sec:Adjoint-simulation-finite}
 
@@
 of the finite-frequency kernels is discussed in Section~\ref{sec:Finite-Frequency-Kernels}.
-
-%------------------------------------------------------------------------------------------------%
-
-\chapter{\label{cha:tomo}Doing tomography, i.e., updating the model based on the sensitivity kernels obtained}
-
-%------------------------------------------------------------------------------------------------%
-
-UNDER CONSTRUCTION (March 2015). Will be added around April / May 2015.
-
-\section{Tomographic inversion / imaging using the sensitivity kernels obtained}
-
-OLD VERSION OF THE SECTION, WILL SOON BE IMPROVED AND ENRICHED.\\
-
-One of the fundamental reasons for computing sensitivity kernels (Section~\ref{sec:Adjoint-simulation-finite})
-is to use them within a tomographic inversion. In other words, use
-recorded seismograms, make measurements with synthetic seismograms,
-and use the misfit between the two to iteratively improve the model
-described by (at least) $V_{{\rm p}}$, $V_{{\rm s}}$, and $\rho$
-as a function of space.
-
-Whatever misfit function you use for the tomographic inversion (you can find several
-examples in \citet{TrKoLi08} and \citet{TrTaLi05}), you will weight
-the sensitivity kernels with measurements. Furthermore, you will use
-as many measurements (stations, components, time windows) as possible
-per event; hence, we call these composite kernels ``event kernels,''
-which are volumetric fields representing the gradient of the misfit
-function with respect to one of the variables (\eg $V_{{\rm s}}$).
-The basic features of an adjoint-based tomographic inversion were
-illustrated in \citet{TaLiTr07} and \citet{TrKoLi08} using a conjugate-gradient
-algorithm; there are dozens of versions of gradient-based inversion
-algorithms that could alternatively be used. The tomographic inversion
-of \citet{TaLiMaTr09,TaLiMaTr2010} used SPECFEM3D Cartesian as well
-as several additional components which are also stored on the CIG
-svn server, described next.
-
-The directory containing external utilities for tomographic inversion using
-SPECFEM3D Cartesian (or other packages that evaluate misfit functions
-and gradients) is in directory \texttt{utils/ADJOINT\_TOMOGRAPHY\_TOOLS/}:
-\begin{verbatim}
-flexwin/     -- FLEXWIN algorithm for automated picking of time windows
-measure_adj/ -- reads FLEXWIN output file and makes measurements,
-                with the option for computing adjoint sources
-iterate_adj/ -- various tools for iterative inversion
-                (requires pre-computed "event kernels")
-\end{verbatim}
-This directory also contains
-a brief \verb+README+ file indicating the role of the three subdirectories,
-\verb+flexwin+ \citep{Maggi2009}, \verb+measure_adj+, and \verb+iterate_adj+.
-The components for making the model update are there; however, there
-are no explicit rules for computing the model update, just as with
-any optimization problem. There are options for computing a conjugate
-gradient step, as well as a source subspace projection step.
-
-The best single file to read is probably: \verb+ADJOINT_TOMO/iterate_adj/cluster/README+.
-
-%------------------------------------------------------------------------------------------------%
-
-\chapter{Noise Cross-correlation Simulations}
-
-%------------------------------------------------------------------------------------------------%
-
-Besides earthquake simulations, SPECFEM3D Cartesian includes functionality
-for seismic noise tomography as well. In order to proceed successfully
-in this chapter, it is critical that you have already familiarized
-yourself with procedures for meshing (Chapter \ref{cha:Mesh-Generation}),
-creating distributed databases (Chapter \ref{cha:Creating-Distributed-Databases}),
-running earthquake simulations (Chapters \ref{cha:Running-the-Solver})
-and adjoint simulations (Chapter \ref{cha:Adjoint-Simulations}).
-Also, make sure you read the article `Noise cross-correlation sensitivity
-kernels' \citep{trompetal2010}, in order to understand noise simulations
-from a theoretical perspective.
-
-\section{Input Parameter Files}
-
-As usual, the three main input files are crucial: \texttt{Par\_file},
-\texttt{CMTSOLUTION} and \texttt{STATIONS}. Unless otherwise specified,
-those input files should be located in directory \texttt{DATA/}.\\
-
-\texttt{CMTSOLUTION} is required for all simulations. At a first glance,
-it may seem unexpected to have it here, since the noise simulations
-should have nothing to do with the earthquake -- \texttt{CMTSOLUTION}.
-However, for noise simulations, it is critical to have no earthquakes.
-In other words, the moment tensor specified in \texttt{CMTSOLUTION}
-must be set to zero manually!\\
-
-\texttt{STATIONS} remains the same as in previous earthquake simulations,
-except that the order of receivers listed in \texttt{STATIONS} is
-now important. The order will be used to determine the `master' receiver,
-i.e., the one that simultaneously cross correlates with the others.\\
-
-\texttt{Par\_file} also requires careful attention. A parameter called
-\texttt{NOISE\_TOMOGRAPHY} has been added which specifies the type
-of simulation to be run. \texttt{NOISE\_TOMOGRAPHY} is an integer
-with possible values 0, 1, 2 and 3. For example, when \texttt{NOISE\_TOMOGRAPHY}
-equals 0, a regular earthquake simulation will be run. When it is
-1/2/3, you are about to run step 1/2/3 of the noise simulations respectively.
-Should you be confused by the three steps, refer to \citet{trompetal2010}
-for details.\\
-
-Another change to \texttt{Par\_file} involves the parameter \texttt{NSTEP}.
-While for regular earthquake simulations this parameter specifies
-the length of synthetic seismograms generated, for noise simulations
-it specifies the length of the seismograms used to compute cross correlations.
-The actual cross correlations are thus twice this length, i.e., $2~\mathrm{NSTEP}-1$.
-The code automatically makes the modification accordingly, if \texttt{NOISE\_TOMOGRAPHY}
-is not zero.\\
-
-There are other parameters in \texttt{Par\_file} which should be given
-specific values. For instance, since the first two steps for calculating
-noise sensitivity kernels correspond to forward simulations, \texttt{SIMULATION\_TYPE}
-must be 1 when \texttt{NOISE\_TOMOGRAPHY} equals 1 or 2. Also, we
-have to reconstruct the ensemble forward wavefields in adjoint simulations,
-therefore we need to set \texttt{SAVE\_FORWARD} to \texttt{.true.}
-for the second step, i.e., when \texttt{NOISE\_TOMOGRAPHY} equals
-2. The third step is for kernel constructions. Hence \texttt{SIMULATION\_TYPE}
-should be 3, whereas \texttt{SAVE\_FORWARD} must be \texttt{.false.}.\\
-
-Finally, for most system architectures, please make sure that \texttt{LOCAL\_PATH}
-in \texttt{Par\_file} is in fact local, not globally shared. Because
-we have to save the wavefields at the earth's surface at every time
-step, it is quite problematic to have a globally shared \texttt{LOCAL\_PATH},
-in terms of both disk storage and I/O speed.
-
-\section{Noise Simulations: Step by Step}
-
-Proper parameters in those parameter files are not enough for noise
-simulations to run. We have more parameters to specify: for example,
-the ensemble-averaged noise spectrum, the noise distribution etc.
-However, since there are a few `new' files, it is better to introduce
-them sequentially. In this section, standard procedures for noise
-simulations are described.
-
-\subsection{Pre-simulation}
-\begin{itemize}
-\item As usual, we first configure the software package using:
-
-\texttt{./configure FC=ifort MPIFC=mpif90}
-
-Use the following if SCOTCH is needed:
-
-\texttt{./configure FC=ifort MPIFC=mpif90 -{}-with-scotch-dir=/opt/scotch}\\
-
-\item Next, we need to compile the source code using:
-
-\texttt{make xgenerate\_databases}
-
-\texttt{make xspecfem3D} \\
-
-\item Before we can run noise simulations, we have to specify the noise
-statistics, e.g., the ensemble-averaged noise spectrum. Matlab scripts
-are provided to help you to generate the necessary file.
-
-\texttt{EXAMPLES/noise\_tomography/NOISE\_TOMOGRAPHY.m} (main program)\\
- \texttt{EXAMPLES/noise\_tomography/PetersonNoiseModel.m}
-
-In Matlab, simply run:
-
-\texttt{NOISE\_TOMOGRAPHY(NSTEP, DT, Tmin, Tmax, NOISE\_MODEL)}\\
-
-\texttt{DT} is given in \texttt{Par\_file}, but \texttt{NSTEP} is
-NOT the one specified in \texttt{Par\_file}. Instead, you have to
-feed $2~\mathrm{NSTEP}-1$ to account for the doubled length of cross
-correlations. \texttt{Tmin} and \texttt{Tmax} correspond to the period
-range you are interested in, whereas \texttt{NOISE\_MODEL} denotes
-the noise model you will be using. Details can be found in the Matlab
-script.
-
-After running the Matlab script, you will be given the following information
-(plus a figure in Matlab):
-
-\texttt{{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}}\\
- \texttt{the source time function has been saved in:}\\
- \texttt{/data2/yangl/3D\_NOISE/S\_squared} (note this path must be
-different)\\
- \texttt{S\_squared should be put into directory:}\\
- \texttt{OUTPUT\_FILES/NOISE\_TOMOGRAPHY/ in the SPECFEM3D Cartesian
-package}\\
-
-In other words, the Matlab script creates a file called \texttt{S\_squared},
-which is the first `new' input file we encounter for noise simulations.
-
-One may choose a flat noise spectrum rather than Peterson's noise
-model. This can be done easily by modifying the Matlab script a little.
-
-\item Create a new directory in the SPECFEM3D Cartesian package, name it
-as \texttt{OUTPUT\_FILES/NOISE\_TOMOGRAPHY/}. We will add some parameter
-files later in this folder.
-\item Put the Matlab-generated-file \texttt{S\_squared} in \texttt{OUTPUT\_FILES/NOISE\_TOMOGRAPHY/}.
-
-That's to say, you will have a file \texttt{OUTPUT\_FILES/NOISE\_TOMOGRAPHY/S\_squared}
-in the SPECFEM3D Cartesian package.
-
-\item Create a file called \texttt{OUTPUT\_FILES/NOISE\_TOMOGRAPHY/irec\_master\_noise}.
-Note that this file is located in directory \texttt{OUTPUT\_FILES/NOISE\_TOMOGRAPHY/}
-as well. In general, all noise simulation related parameter files
-go into that directory. \texttt{irec\_master\_noise} contains only
-one integer, which is the ID of the `master' receiver. For example,
-if this file contains 5, it means that the fifth receiver listed in
-\texttt{DATA/STATIONS} becomes the `master'. That's why we mentioned
-previously that the order of receivers in \texttt{DATA/STATIONS} is
-important.
-
-Note that in regional simulations, the \texttt{DATA/STATIONS} might
-contain receivers which are outside of our computational domains.
-Therefore, the integer in \texttt{irec\_master\_noise} is actually
-the ID in \texttt{DATA/STATIONS\_FILTERED} (which is generated by
-\texttt{bin/xgenerate\_databases}).
-
-\item Create a file called \texttt{OUTPUT\_FILES/NOISE\_TOMOGRAPHY/nu\_master}.
-This file holds three numbers, forming a (unit) vector. It describes
-which component we are cross-correlating at the `master' receiver,
-i.e., ${\hat{{\bf \nu}}}^{\alpha}$ in \citet{trompetal2010}. The
-three numbers correspond to E/N/Z components respectively. Most often,
-the vertical component is used, and in those cases the three numbers
-should be 0, 0 and 1.
-\item Describe the noise direction and distributions in \texttt{src/shared/noise\_tomography.f90}.
-Search for a subroutine called \texttt{noise\_distribution\_direction}
-in \texttt{noise\_tomography.f90}. It is actually located at the very
-beginning of \texttt{noise\_tomography.f90}. The default assumes vertical
-noise and a uniform distribution across the whole free surface of
-the model. It should be quite self-explanatory for modifications.
-Should you modify this part, you have to re-compile the source code.
-\end{itemize}
-
-\subsection{Simulations}
-
-With all of the above done, we can finally launch our simulations.
-Again, please make sure that the \texttt{LOCAL\_PATH} in \texttt{DATA/Par\_file}
-is not globally shared. It is quite problematic to have a globally
-shared \texttt{LOCAL\_PATH}, in terms of both disk storage and speed
-of I/O (we have to save the wavefields at the earth's surface at every
-time step).
-
-As discussed in \citet{trompetal2010}, it takes three steps/simulations
-to obtain one contribution of the ensemble sensitivity kernels:
-\begin{itemize}
-\item Step 1: simulation for generating wavefields\\
- \texttt{SIMULATION\_TYPE=1}\\
- \texttt{NOISE\_TOMOGRAPHY=1}\\
- \texttt{SAVE\_FORWARD} not used, can be either .true. or .false.
-\item Step 2: simulation for ensemble forward wavefields\\
- \texttt{SIMULATION\_TYPE=1}\\
- \texttt{NOISE\_TOMOGRAPHY=2}\\
- \texttt{SAVE\_FORWARD=.true.}
-\item Step 3: simulation for ensemble adjoint wavefields and sensitivity
-kernels\\
- \texttt{SIMULATION\_TYPE=3}\\
- \texttt{NOISE\_TOMOGRAPHY=3}\\
- \texttt{SAVE\_FORWARD=.false.}
-
-Note Step 3 is an adjoint simulation, please refer to previous chapters
-on how to prepare adjoint sources and other necessary files, as well
-as how adjoint simulations work.
-
-\end{itemize}
-Note that it is better to run the three steps continuously within
-the same job, otherwise you have to collect the saved surface movies
-from the old nodes/CPUs to the new nodes/CPUs. This process varies
-from cluster to cluster and thus cannot be discussed here. Please
-ask your cluster administrator for information/configuration of the
-cluster you are using.\\
-
-\subsection{Post-simulation}
-
-After those simulations, you have all stuff you need, either in the
-\texttt{OUTPUT\_FILES/} or in the directory specified by \texttt{LOCAL\_PATH}
-in \texttt{DATA/Par\_file} (which are most probably on local nodes).
-Collect whatever you want from the local nodes to your workstation,
-and then visualize them. This process is the same as what you may
-have done for regular earthquake simulations. Refer to other chapters
-if you have problems.\\
-
-Simply speaking, two outputs are the most interesting: the simulated
-ensemble cross correlations and one contribution of the ensemble sensitivity
-kernels.\\
-
-The simulated ensemble cross correlations are obtained after the second
-simulation (Step 2). Seismograms in \texttt{OUTPUT\_FILES/} are actually
-the simulated ensemble cross correlations. Collect them immediately
-after Step 2, or the Step 3 will overwrite them. Note that we have
-a `master' receiver specified by \texttt{irec\_master\_noise}, the
-seismogram at one station corresponds to the cross correlation between
-that station and the `master'. Since the seismograms have three components,
-we may obtain cross correlations for different components as well,
-not necessarily the cross correlations between vertical components.\\
-
-One contribution of the ensemble cross-correlation sensitivity kernels
-are obtained after Step 3, stored in the \texttt{DATA/LOCAL\_PATH}
-on local nodes. The ensemble kernel files are named the same as regular
-earthquake kernels.\\
-
-You need to run another three simulations to get the other contribution
-of the ensemble kernels, using different forward and adjoint sources
-given in \citet{trompetal2010}.
-
-\section{Example}
-
-In order to illustrate noise simulations in an easy way, one example
-is provided in \texttt{EXAMPLES/noise\_tomography/}. See \texttt{EXAMPLES/noise\_tomography/README}
-for explanations. \\
-
-Note, however, that they are created for a specific workstation (CLOVER@PRINCETON),
-which has at least 4 cores with `mpif90' working properly. \\
-
-If your workstation is suitable, you can run the example in \texttt{EXAMPLES/noise\_tomography/}
-using:\\
-
-\texttt{./pre-processing.sh}\\
-
-Even if this script does not work on your workstation, the procedure
-it describes is universal. You may review the whole process described
-in the last section by following the commands in \texttt{pre-processing.sh},
-which should contain enough explanations for all the commands.
-
-%------------------------------------------------------------------------------------------------%
-
-\chapter{Gravity integral calculations for the gravity field of an Earth model}
-
-%------------------------------------------------------------------------------------------------%
-
-SPECFEM can now compute the gravity field as well as its derivatives (i.e., gravity gradiometry)
-generated by any given 3D Earth model at a given elevation, i.e., on a given observation surface.
-That feature is still experimental but should work just fine.\\
-
-To see how it is implemented and to use it, type this in the root directory of the code:\\
-
-\texttt{grep -i GRAVITY\_INTEGRALS src/*/*}
-
-%------------------------------------------------------------------------------------------------%
-
-\chapter{Graphics}
-
-%------------------------------------------------------------------------------------------------%
-
-\section{\label{sec:Mesh-graphics}Meshes}
-
-In case you used the internal mesher \texttt{xmeshfem3D} to create
-and partition your mesh, you can output mesh files in ABAQUS (.INP)
-and DX (.dx) format to visualize them. For this, you must set either
-the flag \texttt{CREATE\_DX\_FILES} or \texttt{CREATE\_ABAQUS\_FILES}
-to \texttt{.true.} in the mesher's parameter file \texttt{Mesh\_Par\_file}
-prior to running the mesher (see Chapter~\ref{cha:Running-the-Mesher-Meshfem3D}
-for details). You can then use AVS \urlwithparentheses{www.avs.com}
-or OpenDX \urlwithparentheses{www.opendx.org} to visualize the mesh
-and MPI partition (slices).
-
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=0.49\textwidth]{figures/vtk_mesh_vp} \includegraphics[width=0.49\textwidth]{figures/vtk_mesh_vs}
-\par\end{centering}
-
-\caption{Visualization using Paraview of VTK files created by \texttt{xgenerate\_databases}
-showing P- and S-wave velocities assigned to the mesh points. The
-mesh was created by \texttt{xmeshfem3D} for 4 processors.}
-
-\label{fig:vtk.mesh}
-\end{figure}
-
-You have also the option to visualize the distributed databases produced
-by \texttt{xgenerate\_databases} using Paraview \urlwithparentheses{www.paraview.org}.
-For this, you must set the flag \texttt{SAVE\_MESH\_FILES} to \texttt{.true.}
-in the main parameter file \texttt{Par\_file} (see Chapter~\ref{cha:Main-Parameter}
-for details). This will create VTK files for each single partition.
-You can then use Paraview \urlwithparentheses{www.paraview.org}
-to visualized these partitions.
-
-\section{\label{sec:Movies}Movies}
-
-To make a surface or volume movie of the simulation, set parameters
-\texttt{MOVIE\_SURFACE}, \texttt{MOVIE\_VOLUME}, \texttt{MOVIE\_TYPE},
-and \texttt{NTSTEP\_BETWEEN\_FRAMES} in the \texttt{Par\_file}. Turning
-on the movie flags, in particular \texttt{MOVIE\_VOLUME}, produces
-large output files. \texttt{MOVIE\_VOLUME} files are saved in the
-\texttt{LOCAL\_PATH} directory, whereas \texttt{MOVIE\_SURFACE} output
-files are saved in the \texttt{OUTPUT\_FILES} directory. We save the
-displacement field if the parameter \texttt{SAVE\_DISPLACEMENT} is
-set, otherwise the velocity field is saved. The look of a movie is
-determined by the half-duration of the source. The half-duration should
-be large enough so that the movie does not contain frequencies that
-are not resolved by the mesh, i.e., it should not contain numerical
-noise. This can be accomplished by selecting a CMT \texttt{HALF\_DURATION}
-> 1.1 $\times$ smallest period (see figure \ref{fig:CMTSOLUTION-file}).
-When \texttt{MOVIE\_SURFACE} = .\texttt{true.}, the half duration
-of each source in the \texttt{CMTSOLUTION} file is replaced by
-\begin{quote}
-\[
-\sqrt{(}\mathrm{\mathtt{HALF\_DURATIO}\mathtt{N}^{2}}+\mathrm{\mathtt{HDUR\_MOVI}\mathtt{E}^{2}})
-\]
- \textbf{NOTE:} If \texttt{HDUR\_MOVIE} is set to 0.0, the code will
-select the appropriate value of 1.1 $\times$ smallest period. As
-usual, for a point source one can set \texttt{half duration} in the
-\texttt{CMTSOLUTION} file to be 0.0 and \texttt{HDUR\_MOVIE} = 0.0
-in the \texttt{Par\_file} to get the highest frequencies resolved
-by the simulation, but for a finite source one would keep all the
-\texttt{half durations} as prescribed by the finite source model and
-set \texttt{HDUR\_MOVIE} = 0.0.
-\end{quote}
-
-\subsection{Movie Surface}
-
-When running \texttt{xspecfem3D} with the \texttt{MOVIE\_SURFACE}
-flag turned on, the code outputs \texttt{moviedata??????} files in
-the \texttt{OUTPUT\_FILES} directory. There are several flags in the
-main parameter file \texttt{Par\_file} that control the output of
-these movie data files (see section \ref{cha:Main-Parameter} for
-details): \texttt{\small NTSTEP\_BETWEEN\_FRAMES}{\small{} to set the
-timesteps between frames, }\texttt{\small SAVE\_DISPLACEMENT}{\small{}
-to save displacement instead of velocity, }\texttt{\small USE\_HIGHRES\_FOR\_MOVIES}{\small{}
-to save values at all GLL point instead of element edges. In order
-to output additionally shakemaps, you would set the parameter }\texttt{\small CREATE\_SHAKEMAP}{\small{}
-to }\texttt{\small .true.}{\small .}{\small \par}
-
-The files are in a fairly complicated binary format, but there is
-a program provided to convert the output into more user friendly formats:
-\begin{description}
-\item [{\texttt{xcreate\_movie\_shakemap\_AVS\_DX\_GMT}}] From \texttt{create\_movie\_shakemap\_AVS\_DX\_GMT.f90},
-it outputs data in ASCII, OpenDX, or AVS format (also readable in
-ParaView). Before compiling the code, make sure you have the file
-\texttt{surface\_from\_mesher.h} in the \texttt{OUTPUT\_FILES/} directory.
-This file will be created by the solver run. Then type
-
-\begin{lyxcode}
-{\footnotesize make~xcreate\_movie\_shakemap\_AVS\_DX\_GMT}{\footnotesize \par}
-\end{lyxcode}
-
-and run the executable \texttt{xcreate\_movie\_shakemap\_AVS\_DX\_GMT}
-in the main directory. It will create visualization files
-in your format of choice. The code will prompt the user for input
-parameters.
-
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=0.32\textwidth]{figures/movie_surf_1} \includegraphics[width=0.32\textwidth]{figures/movie_surf_2}
-\includegraphics[width=0.32\textwidth]{figures/movie_surf_3}
-\par\end{centering}
-
-\caption{Visualization using AVS files created by \texttt{xcreate\_movie\_shakemap\_AVS\_DX\_GMT}
-showing movie snapshots of vertical velocity components at different
-times.}
-
-\label{fig:movie.surf}
-\end{figure}
-
-\end{description}
-The \texttt{SPECFEM3D Cartesian} code is running in near real-time
-to produce animations of southern California earthquakes via the web;
-see Southern California ShakeMovie\textregistered{}\urlwithparentheses{www.shakemovie.caltech.edu}.
-
-\subsection{\label{sub:Movie-Volume}Movie Volume}
-
-When running xspecfem3D with the \texttt{\small MOVIE\_VOLUME}{\small{}
-flag turned on, the code outputs several files in }\texttt{\small LOCAL\_PATH}{\small{}
-specified in the main }\texttt{\small Par\_file}{\small , e.g. in
-directory }\texttt{\small OUTPUT\_FILES/DATABASES\_MPI}{\small .
-The output is saved by each processor at the time interval specified
-by }\texttt{\small NTSTEP\_BETWEEN\_FRAMES}{\small . For all domains,
-the velocity field is output to files: }\texttt{\small proc??????\_velocity\_X\_it??????.bin}{\small ,
-}\texttt{\small proc??????\_velocity\_Y\_it??????.bin}{\small{} and
-}\texttt{\small proc??????\_velocity\_Z\_it??????.bin}{\small . For
-elastic domains, the divergence and curl taken from the velocity field,
-i.e. $\nabla\cdot{\bf {v}}$ and $\nabla\times{\bf {v}}$, get stored
-as well: }\texttt{\small proc??????\_div\_it??????.bin}{\small , }\texttt{\small proc??????\_curl\_X\_t??????.bin}{\small ,
-}\texttt{\small proc??????\_curl\_Y\_it??????.bin}{\small{} and }\texttt{\small proc??????\_curl\_Z\_it??????.bin}{\small .
-The files denoted }\texttt{\small proc??????\_div\_}{\small ~}\\
-{\small{} }\texttt{\small glob\_it??????.bin}{\small{} and }\texttt{\small proc??????\_curl\_glob\_it??????.bin}{\small{}
-are stored on the global points only, all the other arrays are stored
-on all GLL points. Note that the components X/Y/Z can change to E/N/Z
-according to the }\texttt{\small SUPPRESS\_UTM\_PROJECTION}{\small{}
-flag (see also Appendix~\ref{cha:Coordinates} and \ref{cha:channel-codes}).}{\small \par}
-
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=0.32\textwidth]{figures/movie_volume_1} \includegraphics[width=0.32\textwidth]{figures/movie_volume_2}
-\includegraphics[width=0.32\textwidth]{figures/movie_volume_3}
-\par\end{centering}
-
-\caption{Paraview visualization using movie volume files (converted by \texttt{xcombine\_vol\_data}
-and \texttt{mesh2vtu.pl}) and showing snapshots of vertical velocity
-components at different times.}
-
-\label{fig:movie.volume}
-\end{figure}
-
-To visualize these files, we use an auxilliary program \texttt{combine\_vol\_data.f90}
-to combine the data from all slices into one mesh file. To compile
-it in the root directory, type:
-\begin{lyxcode}
-{\footnotesize make~xcombine\_vol\_data~}{\footnotesize \par}
-\end{lyxcode}
-which will create the executable \texttt{bin/xcombine\_vol\_data}.
-To output the usage of this executable, type
-\begin{lyxcode}
-{\footnotesize ./bin/xcombine\_vol\_data~}{\footnotesize \par}
-\end{lyxcode}
-without arguments.
-
-\medskip
-
-\texttt{xcombine\_vol\_data} will combine the data on the different processors (located in \texttt{OUTPUT\_FILES/DATABASES\_MPI}), for a given quantity and a given iteration, to one file.
-For example, if you want to combine \texttt{velocity\_Z}, for the iteration 400 on 4 processors, i.e.
-if you want to combine these files :
-
-\medskip
-
-\texttt{proc000000\_velocity\_Z\_it000400.bin }
-
-\texttt{proc000001\_velocity\_Z\_it000400.bin }
-
-\texttt{proc000002\_velocity\_Z\_it000400.bin }
-
-\texttt{proc000003\_velocity\_Z\_it000400.bin}
-
-\medskip
-
-You have to go in the directory of the concerned example (where are the directories \texttt{DATA}, \texttt{OUTPUT\_FILES}, etc.).
-Then, you can launch \texttt{xcombine\_vol\_data}, specifying the path where it is located.
-As an example, if the \texttt{DATA} and \texttt{OUTPUT\_FILES} directories of the concerned example are in the root specfem3d directory, \texttt{xcombine\_vol\_data} is in \texttt{./bin}, and you have to type:
-
-\begin{lyxcode}
-~{\footnotesize ./bin/xcombine\_vol\_data~0~3~velocity\_Z\_it000400~./OUTPUT\_FILES/DATABASES\_MPI~./OUTPUT\_FILES~0~}{\footnotesize \par}
-\end{lyxcode}
-
-Here, $0$ is the number of the first processor, $3$ the number of the last one, \texttt{velocity\_Z\_it000400} the name of the files we want to combine without the prefix \texttt{proc000000*\_}, \texttt{./OUTPUT\_FILES/DATABASES\_MPI} the directory where the files are located, \texttt{./OUTPUT\_FILES} the directory where the combined file will be stored, and $0$ is the parameter to create a low-resolution mesh file ($1$ for a high-resolution).
-
-\medskip
-
-At the beginning of the source file \texttt{combine\_vol\_data.f90} (located in \texttt{src/auxiliaries/}), there is a logical constant parameter \texttt{USE\_VTK\_OUTPUT} whose default value is \texttt{.true.}, that which will provide the output combined file directly in the VTK format and you can use Paraview to visualize it.
-
-\medskip
-
-If the value of this parameter is \texttt{.false.}, the output mesh file will have the name \texttt{velocity\_Z\_it000400.mesh}. Then we next have to convert the
-\texttt{.mesh} file into the VTU (Unstructured grid file) format which can be viewed in ParaView. For this task, you can use and modify the
-script \texttt{mesh2vtu.pl} located in directory \texttt{utils/Visualization/Paraview/}, for example:
-\begin{lyxcode}
-{\footnotesize mesh2vtu.pl~-i~velocity\_Z\_it000400.mesh~-o~velocity\_Z\_it000400.vtu}{\footnotesize \par}
-\end{lyxcode}
-Notice that this Perl script uses a program \texttt{mesh2vtu} in the \texttt{utils/Visualization/Paraview/mesh2vtu} directory, which further
-uses the VTK \urlwithparentheses{http://www.vtk.org/} run-time library for its execution. Therefore, make sure you have them properly set
-in the script according to your system.
-
-\bigskip
-
-Then, to do a movie with several iterations, you have to repeat this process for each iteration you want to put in your movie.
-
-\section{\label{sec:Finite-Frequency-Kernels}Finite-Frequency Kernels}
-
-The finite-frequency kernels computed as explained in Section \ref{sec:Adjoint-simulation-finite}
-are saved in the \texttt{LOCAL\_PATH} at the end of the simulation.
-Therefore, we first need to collect these files on the front end,
-combine them into one mesh file, and visualize them with some auxilliary
-programs.
-\begin{enumerate}
-\item \textbf{Create slice files}
-
-We will only discuss the case of one source-receiver pair, i.e., the
-so-called banana-doughnut kernels. Although it is possible to collect
-the kernel files from all slices on the front end, it usually takes
-up too much storage space (at least tens of gigabytes). Since the
-sensitivity kernels are the strongest along the source-receiver great
-circle path, it is sufficient to collect only the slices that are
-along or close to the great circle path.
-
-A Perl script \texttt{slice\_number.pl} located in directory \texttt{utils/Visualization/Paraview/}
-can help to figure out the slice numbers that lie along the great
-circle path. It applies to meshes created with the internal mesher
-\texttt{xmeshfem3D}.
-\begin{enumerate}
-\item On machines where you have access to the script, copy the \texttt{Mesh\_Par\_file},
-and \texttt{output\_solver} files, and run:
-
-\begin{lyxcode}
-{\small slice\_number.pl~Mesh\_Par\_file~output\_solver.txt~slice\_file}{\small \par}
-\end{lyxcode}
-
-which will generate a \texttt{slices\_file}.
-
-\item For cases with multiple sources and multiple receivers, you need to
-provide a slice file before proceeding to the next step.
-\end{enumerate}
-\item \textbf{Collect the kernel files}
-
-After obtaining the slice files, you can collect the corresponding
-kernel files from the given slices.
-\begin{enumerate}
-\item You can use or modify the script \texttt{utils/copy\_basin\_database.pl}
-to accomplish this:
-
-\begin{lyxcode}
-{\small utils/copy\_database.pl~slice\_file~lsf\_machine\_file~filename~{[}jobid{]}}{\small \par}
-\end{lyxcode}
-
-where \texttt{\small lsf\_machine\_file}{\small{} is the machine file
-generated by the LSF scheduler, }\texttt{\small filename}{\small{}
-is the kernel name (e.g., }\texttt{\small rho\_kernel}{\small , }\texttt{\small alpha\_kernel}{\small{}
-and }\texttt{\small beta\_kernel}{\small ), and the optional }\texttt{\small jobid}{\small{}
-is the name of the subdirectory under }\texttt{\small LOCAL\_PATH}{\small{}
-where all the kernel files are stored.}{\small \par}
-
-\item After executing this script, all the necessary mesh topology files
-as well as the kernel array files are collected to the local directory
-of the front end.
-\end{enumerate}
-\item \textbf{Combine kernel files into one mesh file}
-
-We use an auxilliary program \texttt{combine\_vol\_data.f90} to combine
-the kernel files from all slices into one mesh file.
-\begin{enumerate}
-\item Compile it in the root directory:
-
-\begin{lyxcode}
-{\footnotesize make~xcombine\_vol\_data~}{\footnotesize \par}
-
-{\footnotesize ./bin/xcombine\_vol\_data~slice\_list~filename~input\_dir~output\_dir~high/low-resolution~}{\footnotesize \par}
-\end{lyxcode}
-
-where \texttt{input\_dir} is the directory where all the individual
-kernel files are stored, and \texttt{output\_dir} is where the mesh
-file will be written.
-
-\item Use 1 for a high-resolution mesh, outputting all the GLL points to
-the mesh file, or use 0 for low resolution, outputting only the corner
-points of the elements to the mesh file.
-\item The output mesh file will have the name \texttt{filename\_rho(alpha,beta).mesh}
-\end{enumerate}
-\item \textbf{Convert mesh files into .vtu files}
-
-\begin{enumerate}
-\item We next convert the \texttt{.mesh} file into the VTU (Unstructured
-grid file) format which can be viewed in ParaView. For this task,
-you can use and modify the script \texttt{mesh2vtu.pl} located in
-directory \texttt{utils/Visualization/Paraview/}, for example:
-
-\begin{lyxcode}
-{\footnotesize mesh2vtu.pl~-i~file.mesh~-o~file.vtu~}{\footnotesize \par}
-\end{lyxcode}
-\item Notice that this Perl script uses a program \texttt{mesh2vtu} in the
-\texttt{utils/Visualization/Paraview/mesh2vtu} directory, which further
-uses the VTK \urlwithparentheses{http://www.vtk.org/} run-time library
-for its execution. Therefore, make sure you have them properly set
-in the script according to your system.
-\end{enumerate}
-\item \textbf{Copy over the source and receiver .vtk file}
-
-In the case of a single source and a single receiver, the simulation
-also generates the file \texttt{sr.vtk} located in the \texttt{OUTPUT\_FILES/}
-directory to describe the source and receiver locations, which can
-also be viewed in Paraview in the next step.
-
-\item \textbf{View the mesh in ParaView}
-
-Finally, we can view the mesh in ParaView \urlwithparentheses{www.paraview.org}.
-\begin{enumerate}
-\item Open ParaView.
-\item From the top menu, \textsf{File} $\rightarrow$\textsf{Open data},
-select \texttt{file.vtu}, and click the \textsf{Accept} button.
-
-\begin{itemize}
-\item If the mesh file is of moderate size, it shows up on the screen; otherwise,
-only the bounding box is shown.
-\end{itemize}
-\item Click \textsf{Display Tab} $\rightarrow$ \textsf{Display Style} $\rightarrow$
-\textsf{Representation} and select \textsf{wireframe of surface} to
-display it.
-\item To create a cross-section of the volumetric mesh, choose \textsf{Filter}
-$\rightarrow$ \textsf{cut}, and under \textsf{Parameters Tab}, choose
-\textsf{Cut Function} $\rightarrow$ \textsf{plane}.
-\item Fill in center and normal information given by the \texttt{global\_slice\_number.pl}
-script (either from the standard output or from \texttt{normal\_plane.txt}
-file).
-\item To change the color scale, go to \textsf{Display Tab} $\rightarrow$
-\textsf{Color} $\rightarrow$ \textsf{Edit Color Map} and reselect
-lower and upper limits, or change the color scheme.
-\item Now load in the source and receiver location file by \textsf{File}
-$\rightarrow$ \textsf{Open data}, select \texttt{sr.vt}k, and click
-the \textsf{Accept} button. Choose \textsf{Filter} $\rightarrow$
-\textsf{Glyph}, and represent the points by `\textsf{spheres}'.
-\item For more information about ParaView, see the ParaView Users Guide
-\urlwithparentheses{www.paraview.org/files/v1.6/ParaViewUsersGuide.PDF}.
-\end{enumerate}
-\end{enumerate}
-\begin{figure}[H]
-\noindent \begin{centering}
-\includegraphics[scale=0.6]{figures/3D-S-Kernel}
-\par\end{centering}
-
-\caption{(a) Top Panel: Vertical source-receiver cross-section of the S-wave
-finite-frequency sensitivity kernel $K_{\beta}$ for station GSC at
-an epicentral distance of 176 km from the September 3, 2002, Yorba
-Linda earthquake. Lower Panel: Vertical source-receiver cross-section
-of the 3D S-wave speed model used for the spectral-element simulations
-\citep{KoLiTrSuStSh04}. (b) The same as (a) but for station HEC at
-an epicentral distance of 165 km \citep{LiTr06}.}
-
-\label{figure:P-wave-speed-finite-frequency}
-\end{figure}
-
-\chapter{\label{cha:Scheduler}Running through a Scheduler}
-
-The code is usually run on large parallel machines, often PC clusters,
-most of which use schedulers, i.e., queuing or batch management systems
-to manage the running of jobs from a large number of users. The following
-considerations need to be taken into account when running on a system
-that uses a scheduler:
-\begin{itemize}
-\item The processors/nodes to be used for each run are assigned dynamically
-by the scheduler, based on availability. Therefore, in order for the
-\texttt{xgenerate\_databases} and the \texttt{xspecfem3D} executables
-(or between successive runs of the solver) to have access to the same
-database files (if they are stored on hard drives local to the nodes
-on which the code is run), they must be launched in sequence as a
-single job.
-\item On some systems, the nodes to which running jobs are assigned are
-not configured for compilation. It may therefore be necessary to pre-compile
-both the \texttt{xgenerate\_databases} and the \texttt{xspecfem3D}
-executables.
-\item One feature of schedulers/queuing systems is that they allow submission
-of multiple jobs in a ``launch and forget'' mode. In order to take
-advantage of this property, care needs to be taken that output and
-intermediate files from separate jobs do not overwrite each other,
-or otherwise interfere with other running jobs.
-\end{itemize}
-Examples of job scripts can be found in the \texttt{\small utils/Cluster/}{\small{}
-directory and can straightforwardly be modified and adapted to meet
-more specific running needs.}{\small \par}
-
-\chapter{{\normalsize \label{cha:-Changing-the}} Changing the Model}
-
-In this section we explain how to change the velocity model used for
-your simulations. These changes involve contributing specific subroutines
-that replace existing subroutines in the \texttt{SPECFEM3D Cartesian}
-package. Note that \texttt{SPECFEM3D Cartesian} can handle Earth models
-with material properties that vary within each spectral element.
-
-\section{{\normalsize \label{sec:Using-tomographic}}Using external tomographic Earth models}
-
-To implement your own external tomographic model(s), you must provide
-your own external tomography file(s), and choose between two possible
-options:\\
- \indent (1) set in the \texttt{Par\_file} the model parameter \texttt{MODEL
-= tomo}, \\
- or for more user control: \\
- \indent (2) set in the \texttt{Par\_file} the model parameter \texttt{MODEL
-= default}, define the negative \texttt{material\_ID} identifier for
-each element in the file \texttt{MESH/materials\_file} and use the
-following format in the file \texttt{MESH/nummaterial\_velocity\_file}
-when using CUBIT to construct your mesh (see also section \ref{subsec:Exporting-the-Mesh}):
-\begin{lyxcode}
-domain\_ID~material\_ID~tomography~elastic~file\_name~1
-\end{lyxcode}
-where: \\
- \indent \texttt{domain\_ID} is 1 for acoustic or 2 for elastic materials,
-\\
- \indent \texttt{material\_ID} a negative, unique identifier (i.e.,
--1,-2,...), \\
- \indent \texttt{tomography} keyword for tomographic material definition,
-\\
- \indent \texttt{elastic} keyword for elastic material definition,
-\\
- \indent \texttt{file\_name} the name of the tomography file and
-1 a positive unique identifier.\\
-
-The external tomographic model is represented by a grid of points
-with assigned material properties and homogeneous resolution along
-each spatial direction x, y and z. The ASCII file \texttt{file\_name}
-that describe the tomography should be located in the \texttt{TOMOGRAPHY\_PATH}
-directory, set in the \texttt{Par\_file}. The format of the file,
-as read from \texttt{model\_tomography.f90} located in the \texttt{src/generate\_databases}
-directory, looks like Figure \ref{fig:tomography_file}, where
-\begin{description}
-\item [{\texttt{ORIG\_X}, \texttt{END\_X}}] are, respectively, the coordinates
-of the initial and final tomographic grid points along the x direction
-(in the mesh units, e.g., $m$);
-\item [{\texttt{ORIG\_Y}, \texttt{END\_Y}}] respectively the coordinates
-of the initial and final tomographic grid points along the y direction
-(in the mesh units, e.g., $m$);
-\item [{\texttt{ORIG\_Z}, \texttt{END\_Z}}] respectively the coordinates
-of the initial and final tomographic grid points along the z direction
-(in the mesh units, e.g., $m$);
-\item [{\texttt{SPACING\_X}, \texttt{SPACING\_Y}, \texttt{SPACING\_Z}}] the
-spacing between the tomographic grid points along the x, y and z directions,
-respectively (in the mesh units, e.g., $m$);
-\item [{\texttt{NX}, \texttt{NY}, \texttt{NZ}}] the number of grid points
-along the spatial directions x, y and z, respectively; \texttt{NX}
-is given by {[}(\texttt{END\_X} - \texttt{ORIG\_X})/\texttt{SPACING\_X}{]}+1;
-\texttt{NY} and \texttt{NZ} are the same as \texttt{NX}, but for the
-y and z directions, respectively;
-\item [{\texttt{VP\_MIN}, \texttt{VP\_MAX}, \texttt{VS\_MIN}, \texttt{VS\_MAX}, \texttt{RHO\_MIN}, \texttt{RHO\_MAX}}] the
-minimum and maximum values of the wave speed \texttt{vp} and \texttt{vs}
-(in $m\, s^{-1}$) and of the density \texttt{rho} (in $kg\, m^{-3}$);
-these values could be the actual limits of the tomographic parameters
-in the grid or the minimum and maximum values to which we force the
-cut of velocity and density in the model.
-\end{description}
-After the first four lines, in the file \texttt{file\_name} the tomographic
-grid points are listed with the corresponding values of \texttt{vp},
-\texttt{vs} and \texttt{rho}, scanning the grid along the x coordinate
-(from \texttt{ORIG\_X} to \texttt{END\_X} with step of \texttt{SPACING\_X})
-for each given y (from \texttt{ORIG\_Y} to \texttt{END\_Y}, with step
-of \texttt{SPACING\_Y}) and each given z (from \texttt{ORIG\_Z} to
-\texttt{END\_Z}, with step of \texttt{SPACING\_Z}). %fig
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=1\textwidth]{figures/tomo_file}
-\par\end{centering}
-
-\caption{Tomography file \texttt{file\_name} that describes an external Earth
-model. The coordinates x, y and z of the grid, their limits ORIG\_X,
-ORIG\_Y, ORIG\_Z, END\_X, END\_Y, END\_Z and the grid spacings SPACING\_X,
-SPACING\_Y, SPACING\_Z should be in the units of the constructed mesh
-(e.g., UTM coordinates in meters).}
-
-\label{fig:tomography_file}
-\end{figure}
-
-The user can implement his own interpolation algorithm for the tomography
-model by changing the routine \texttt{model\_tomography.f90} located
-in the \texttt{src/generate\_databases/} directory. Moreover, for
-models that involve both fully defined materials and a tomography
-description, the \texttt{nummaterial\_velocity\_file} has multiple
-lines each with the corresponding suitable format described above.
-
-%==============
-
-\subsubsection*{Example: External tomography file with variable discretization intervals}
-
-The example in Figure~\ref{fig:tomography_file} is for an external tomography file with uniform spacing in the $x$, $y$, and $z$ directions.
-(Note that $\Delta x$, $\Delta y$, and $\Delta z$ need not be the same as in the example.) In that example, the  \texttt{nummaterial\_velocity\_file} is
-%
-\begin{lyxcode}
-2  -1 tomography elastic tomography\_model.xyz 1
-\end{lyxcode}
-%
-and the file \texttt{tomography\_model.xyz} will need to reside in \texttt{DATA/}.
-All entries in the second column of \texttt{materials\_file} will be \texttt{-1}, which means that each element in the mesh will be interpolated accordind to the values in \texttt{tomography\_model.xyz}.
-
-In some cases it may be desirable to have an external tomography model that is described in more than one file.
-For example, in cases like southern California, the length scale of variation in the structure of the wave speed model is much shorter in the sedimentary basin models within the upper 15~km.
-Therefore one might want to use an external tomography file that is sampled with, say, $\Delta x = 1000$~m, $\Delta y = 1000$~m, and $\Delta z = 250$~m in the uppermost 15~km, and then use $\Delta x = 2000$~m, $\Delta y = 2000$~m, and $\Delta z = 1000$~m below a depth of 15~km.
-If these intervals are chosen appropriately, then it will result in a pair of external tomography file that is much smaller than the alternative of having a single file with the fine discretization.
-In this case  \texttt{nummaterial\_velocity\_file} is
-%
-\begin{lyxcode}
-2  -1 tomography elastic file\_above\_15km.xyz 1 \\
-2  -2 tomography elastic file\_below\_15km.xyz 1
-\end{lyxcode}
-%
-and the files \texttt{file\_above\_15km.xyz} and \texttt{file\_below\_15km.xyz} will need to reside in \texttt{DATA/}.
-All entries in the second column of \texttt{materials\_file} will need to be either \texttt{-1} or \texttt{-2}, depending on whether the element is above or below 15~km depth.
-In this sense, the construction of the mesh and the discretization of the external model will generally be done in tandem.
-
-Other more complicated discretizations can be used following the same procedure. (In particular, there is no limit on the number of different external files that are used.)
-
-%=============
-
-\section{{\normalsize \label{sec:Anelastic-Models}}External (an)elastic Models}
-
-To use your own external model, you can set in the \texttt{Par\_file}
-the model parameter \texttt{MODEL = external}. Three-dimensional acoustic
-and/or (an)elastic (attenuation) models may then be superimposed onto
-the mesh based upon your subroutine in \texttt{model\_external\_values.f90}
-located in directory \texttt{src/generate\_databases}. The call to
-this routine would be as follows
-\begin{lyxcode}
-call~model\_external\_values(xmesh,~ymesh,~zmesh,~~\&~~\\
-~~~~~~~~~~~rho,~vp,~vs,~qmu\_atten,~iflag\_aniso,~idomain\_id)
-\end{lyxcode}
-Input to this routine consists of:
-\begin{description}
-\item [{\texttt{xmesh,ymesh,zmesh}}] location of mesh point
-\end{description}
-Output to this routine consists of:
-\begin{description}
-\item [{\texttt{rho,vp,vs}}] isotropic model parameters for density $\rho$
-($kg/m^{3}$), $v_{p}$ ($m/s$) and $v_{s}$ ($m/s$)
-\item [{\texttt{qmu\_atten}}] Shear wave quality factor: $0<Q_{\mu}<9000$
-\item [{\texttt{iflag\_aniso}}] anisotropic model flag, $0$ indicating
-no anisotropy or $1$ using anisotropic model parameters as defined
-in routine file \texttt{model\_aniso.f90}
-\item [{\texttt{idomain\_id}}] domain identifier, $1$ for acoustic, $2$
-for elastic, $3$ for poroelastic materials.
-\end{description}
-Note that the resolution and maximum value of anelastic models are
-truncated. This speeds the construction of the standard linear solids
-during the meshing stage. To change the resolution, currently at one
-significant figure following the decimal, or the maximum value (9000),
-consult \texttt{shared/constants.h}.
-
-\section{{\normalsize \label{sec:Anisotropic-Models}}Anisotropic Models}
-
-To use your anisotropic models, you can either set in the \texttt{Par\_file}
-the model parameter \texttt{ANISOTROPY} to \texttt{.true.} or set
-the model parameter \texttt{MODEL = aniso}. Three-dimensional anisotropic
-models may be superimposed on the mesh based upon the subroutine in
-file \texttt{model\_aniso.f90} located in directory \texttt{src/generate\_databases}.
-The call to this subroutine is of the form
-\begin{lyxcode}
-call~model\_aniso(iflag\_aniso,~rho,~vp,~vs,~\&~~\\
-~~~~~~~~~~~c11,c12,c13,c14,c15,c16,c22,c23,c24,c25,c26,~\&~~\\
-~~~~~~~~~~~c33,c34,c35,c36,c44,c45,c46,c55,c56,c66)~
-\end{lyxcode}
-Input to this routine consists of:
-\begin{description}
-\item [{\texttt{iflag\_aniso}}] flag indicating the type of the anisotropic
-model, i.e. $0$ for no superimposed anisotropy, $1$ or $2$ for
-generic pre-defined anisotropic models.
-\item [{\texttt{rho,vp,vs}}] reference isotropic model parameters for density
-$\rho$, $v_{p}$ and $v_{s}$.
-\end{description}
-Output from the routine consists of the following non-dimensional
-model parameters:
-\begin{description}
-\item [{\texttt{c11},}] \textbf{$\cdots$,} \texttt{\textbf{c66}} 21~dimensionalized
-anisotropic elastic parameters.
-\end{description}
-You can replace the \texttt{model\_aniso.f90} file by your own version
-\textit{provided you do not change the call structure of the routine},
-i.e., the new routine should take exactly the same input and produce
-the required relative output.
-
-\section{{\normalsize \label{sec:Using-SEP}}Using external SEP models.}
-
-Stanford Exploration Project (SEP) models consists of two files per variables, one header file (\texttt{.H})
-and one binary file (\texttt{.H$@$}).
-Among other information, the header file indicates, for each dimension,
-the offset (\texttt{o1}, \texttt{o2} and \texttt{o3} fields),
-the spatial step size (\texttt{d1}, \texttt{d2} and \texttt{d3} fields)
-and the number of samples (\texttt{n1}, \texttt{n2} and \texttt{n3}) necessary to read
-the linked binary file (\texttt{in} field).
-
-SEP model reading occurs during database generation. It it the user responsibility
-to ensure that proper interfaces between acoustic and elastic domains are set up
-during the meshing phase. For more information, refer to
-section~\ref{cha:Running-the-Mesher-Meshfem3D}.
-
-In \texttt{DATA/Par\_file}, \texttt{\textbf{MODEL}} should be set to \texttt{\textbf{sep}},
-and \texttt{\textbf{SEP\_MODEL\_DIRECTORY}} should point to wherever your sep model files reside.
-
-Note that in order not to overload the parameter file, SEP header files
-should be named \texttt{vp.H}, \texttt{vs.H} and \texttt{rho.H}. The \texttt{in} field
-in these binary files can be any path as long as it is relative to
-the previously set \texttt{SEP\_MODEL\_DIRECTORY}.
-We also assume that any dimensional value inside SEP header files is given in meters and
-that binary files are single precision floating point numbers stored in little endian format.
-There is currently only support for \texttt{vp}, \texttt{vs} and \texttt{rho} variables.
-
-An example demonstrating the use SEP models is given in:
-\texttt{EXAMPLES/meshfem3D\_examples/sep\_bathymetry}.
-
-\chapter{Post-Processing Scripts}
-
-Several post-processing scripts/programs are provided in the \texttt{utils/}
-directory, and most of them need to be adjusted when used on different
-systems, for example, the path of the executable programs. Here we
-only list a few of the available scripts and provide a brief description,
-and you can either refer to the related sections for detailed usage
-or, in a lot of cases, type the script/program name without arguments
-for its usage.
-
-\section{Process Data and Synthetics\label{sec:Process-data-and-syn}}
-
-In many cases, the SEM synthetics are calculated and compared to data
-seismograms recorded at seismic stations. Since the SEM synthetics
-are accurate for a certain frequency range, both the original data
-and the synthetics need to be processed before a comparison can be
-made.
-
-For such comparisons, the following steps are recommended:
-\begin{enumerate}
-\item Make sure that both synthetic and observed seismograms have the correct
-station/event and timing information.
-\item Convolve synthetic seismograms with a source time function with the
-half duration specified in the \texttt{CMTSOLUTION} file, provided,
-as recommended, you used a zero half duration in the SEM simulations.
-\item Resample both observed and synthetic seismograms to a common sampling
-rate.
-\item Cut the records using the same window.
-\item Remove the trend and mean from the records and taper them.
-\item Remove the instrument response from the observed seismograms (recommended)
-or convolve the synthetic seismograms with the instrument response.
-\item Make sure that you apply the same filters to both observed and synthetic
-seismograms. Preferably, avoid filtering your records more than once.
-\item Now, you are ready to compare your synthetic and observed seismograms.
-\end{enumerate}
-\noindent We generally use the following scripts provided in the \texttt{utils/seis\_process/}
-directory:
-
-\subsection{Data processing script \texttt{process\_data.pl}}
-
-This script cuts a given portion of the original data, filters it,
-transfers the data into a displacement record, and picks the first
-P and S arrivals. For more functionality, type `\texttt{process\_data.pl}'
-without any argument. An example of the usage of the script:
-\begin{lyxcode}
-{\footnotesize process\_data.pl~-m~CMTSOLUTION~-s~1.0~-l~0/4000~-i~-f~-t~40/500~-p~-x~bp~DATA/1999.330{*}.BH?.SAC}~
-\end{lyxcode}
-\noindent which has resampled the SAC files to a sampling rate of
-1 seconds, cut them between 0 and 4000 seconds, transfered them into
-displacement records and filtered them between 40 and 500 seconds,
-picked the first P and S arrivals, and added suffix `\texttt{bp}'
-to the file names.
-
-Note that all of the scripts in this section actually use the SAC
-and/or IASP91 to do the core operations; therefore make sure that
-the SAC and IASP91 packages are installed properly on your system,
-and that all the environment variables are set properly before running
-these scripts.
-
-\subsection{Synthetics processing script \texttt{process\_syn.pl}}
-
-This script converts the synthetic output from the SEM code from ASCII
-to SAC format, and performs similar operations as `\texttt{process\_data.pl}'.
-An example of the usage of the script:
-\begin{lyxcode}
-{\footnotesize process\_syn.pl~-m~CMTSOLUTION~-h~-a~STATIONS~-s~1.0~-l~0/4000~-f~-t~40/500~-p~-x~bp~SEM/{*}.BX?.semd}~
-\end{lyxcode}
-which will convolve the synthetics with a triangular source-time function
-from the \texttt{CMTSOLUTION} file, convert the synthetics into SAC
-format, add event and station information into the SAC headers, resample
-the SAC files with a sampling rate of 1 seconds, cut them between
-0 and 4000 seconds, filter them between 40 and 500 seconds with the
-same filter used for the observed data, pick the first P and S arrivals,
-and add the suffix `\texttt{bp}' to the file names.
-
-More options are available for this script, such as adding time shift
-to the origin time of the synthetics, convolving the synthetics with
-a triangular source time function with a given half duration, etc.
-Type \texttt{process\_syn.pl} without any argument for a detailed
-usage.
-
-In order to convert between SAC format and ASCII files, useful scripts
-are provided in the subdirectories ~\\
- \texttt{utils/sac2000\_alpha\_convert/} and \texttt{utils/seis\_process/asc2sac/}.
-
-\subsection{Script \texttt{rotate.pl}}
-
-The original data and synthetics have three components: vertical (BHZ
-resp. BXZ), north (BHN resp. BXN) and east (BHE resp. BXE). However,
-for most seismology applications, transverse and radial components
-are also desirable. Therefore, we need to rotate the horizontal components
-of both the data and the synthetics to the transverse and radial direction,
-and \texttt{\small rotate.pl}{\small{} can be used to accomplish this:}{\small \par}
-\begin{lyxcode}
-rotate.pl~-l~0~-L~180~-d~DATA/{*}.BHE.SAC.bp~
-
-rotate.pl~-l~0~-L~180~SEM/{*}.BXE.semd.sac.bp~
-\end{lyxcode}
-where the first command performs rotation on the SAC data obtained
-through Seismogram Transfer Program (STP) \urlwithparentheses{http://www.data.scec.org/STP/stp.html},
-while the second command rotates the processed SEM synthetics.
-
-\section{Collect Synthetic Seismograms}
-
-The forward and adjoint simulations generate synthetic seismograms
-in the \texttt{OUTPUT\_FILES/} directory by default. For the forward
-simulation, the files are named like \texttt{NT.STA.BX?.semd} for
-two-column time series, or \texttt{NT.STA.BX?.semd.sac} for ASCII
-SAC format, where NT and STA are the network code and station name,
-and \texttt{BX?} stands for the component name. Please see the Appendix~\ref{cha:Coordinates}
-and \ref{cha:channel-codes} for further details.
-
-The adjont simulations generate synthetic seismograms with the name
-\texttt{NT.S?????.S??.sem} (refer to Section \ref{sec:Adjoint-simulation-sources}
-for details). The kernel simulations output the back-reconstructed
-synthetic seismogram in the name \texttt{NT.STA.BX?.semd}, mainly
-for the purpose of checking the accuracy of the reconstruction. Refer
-to Section \ref{sec:Adjoint-simulation-finite} for further details.
-
-You do have further options to change this default output behavior,
-given in the main constants file \texttt{constants.h} located in \texttt{src/shared/}
-directory:
-\begin{description}
-\item [{\texttt{SEISMOGRAMS\_BINARY}}] set to \texttt{.true.} to have seismograms
-written out in binary format.
-\item [{\texttt{WRITE\_SEISMOGRAMS\_BY\_MASTER}}] Set to \texttt{.true.}
-to have only the master process writing out seismograms. This can
-be useful on a cluster, where only the master process node has access
-to the output directory.
-\item [{\texttt{USE\_OUTPUT\_FILES\_PATH}}] Set to \texttt{.false.} to
-have the seismograms output to \texttt{LOCAL\_PATH} directory specified
-in the main parameter file \texttt{DATA/Par\_file}. In this case,
-you could collect the synthetics onto the frontend using the \texttt{collect\_seismo\_lsf\_multi.pl}
-script located in the \texttt{utils/Cluster/lsf/} directory. The usage
-of the script would be e.g.:
-
-\begin{lyxcode}
-collect\_seismo.pl~machines~DATA/Par\_file
-\end{lyxcode}
-\end{description}
-where \texttt{machines} is a file containing the node names and \texttt{DATA/Par\_file}
-the parameter file used to extract the \texttt{LOCAL\_PATH} directory
-used for the simulation.
-
-\section{Clean Local Database}
-
-After all the simulations are done, the seismograms are collected,
-and the useful database files are copied to the frontend, you may
-need to clean the local scratch disk for the next simulation. This
-is especially important in the case of kernel simulation, where very
-large files are generated for the absorbing boundaries to help with
-the reconstruction of the regular forward wavefield. A sample script
-is provided in \texttt{utils/}:
-\begin{lyxcode}
-cleanbase.pl~machines
-\end{lyxcode}
-where \texttt{machines} is a file containing the node names.
-
-\section{Plot Movie Snapshots and Synthetic Shakemaps}
-
-\subsection{Script \texttt{movie2gif.gmt.pl}}
-
-With the movie data saved in \texttt{OUTPUT\_FILES/} at the end of
-a movie simulation (\texttt{\small MOVIE\_SURFACE=.true.}{\small ),
-you can run the }\texttt{\small `create\_movie\_shakemap\_AVS\_DX\_GMT}{\small '
-code to convert these binary movie data into GMT xyz files for futher
-processing. A sample script }\texttt{\small movie2gif.gmt.pl}{\small{}
-is provided to do this conversion, and then plot the movie snapshots
-in GMT, for example:}{\small \par}
-\begin{lyxcode}
-movie2gif.gmt.pl~-m~CMTSOLUTION~-g~-f~1/40~-n~-2~-p~
-\end{lyxcode}
-which for the first through the 40th movie frame, converts the \texttt{moviedata}
-files into GMT xyz files, interpolates them using the 'nearneighbor'
-command in GMT, and plots them on a 2D topography map. Note that `\texttt{-2}'
-and `\texttt{-p}' are both optional.
-
-\subsection{Script \texttt{plot\_shakemap.gmt.pl}}
-
-With the shakemap data saved in \texttt{OUTPUT\_FILES/} at the end
-of a shakemap simulation ~\\
- (\texttt{CREATE\_SHAKEMAP=.true.}), you can also run \texttt{`create\_movie\_shakemap\_AVS\_DX\_GMT}'
-code to convert the binary shakemap data into GMT xyz files. A sample
-script \texttt{plot\_shakemap.gmt.pl} is provided to do this conversion,
-and then plot the shakemaps in GMT, for example:
-\begin{lyxcode}
-plot\_shakemap.gmt.pl~data\_dir~type(1,2,3)~CMTSOLUTION~
-\end{lyxcode}
-where \texttt{type=1} for a displacement shakemap, \texttt{2} for
-velocity, and \texttt{3} for acceleration.
-
-\section{Map Local Database}
-
-A sample program \texttt{remap\_database} is provided to map the local
-database from a set of machines to another set of machines. This is
-especially useful when you want to run mesher and solver, or different
-types of solvers separately through a scheduler (refer to Chapter~\ref{cha:Scheduler}).
-\begin{lyxcode}
-run\_lsf.bash~-{}-gm-no-shmem~-{}-gm-copy-env~remap\_database~old\_machines~150
-\end{lyxcode}
-where \texttt{old\_machines} is the LSF machine file used in the previous
-simulation, and \texttt{150} is the number of processors in total.
-
-%============================================================
-
-\chapter{\label{cha:developers}Information for developers of the code, and for people who want to learn how the technique works}
-
-You can get a very simple 1D version of a demo code (there is one in Fortran and one in Python):\\
-
-git clone --recursive https://github.com/geodynamics/specfem1d.git\\
-
-\noindent We also have simple 3D demo source codes that implement the SEM in a single, small program, in directory\\
-utils/small\_SEM\_solvers\_in\_Fortran\_and\_C\_without\_MPI\_to\_learn of the specfem3d package.
-They are useful to learn how the spectral-element method works, and how to write or modify a code to implement it.
-Also useful to test new ideas by modifying these simple codes to run some tests.\\
-
-\noindent For information on how to contribute to the code, i.e., for how to make your modifications, additions or improvements part of the
-official package, see https://github.com/geodynamics/specfem3d/wiki .
-
-%============================================================
-
-\chapter*{\label{cha:Bug-Reports-and}Bug Reports and Suggestions for Improvements}
-
-To report bugs or suggest improvements to the code, please send an
-e-mail to the CIG Computational Seismology Mailing List \urlwithparentheses{cig-seismo@geodynamics.org}.
-
-\addcontentsline{toc}{chapter}{Notes \& Acknowledgments}
-
-\chapter*{Notes \& Acknowledgments}
-
-In order to keep the software package thread-safe in case a multithreaded
-implementation of MPI is used, developers should not add modules or
-common blocks to the source code but rather use regular subroutine
-arguments (which can be grouped in ``derived types'' if needed for
-clarity).
-
-The Gauss-Lobatto-Legendre subroutines in \texttt{gll\_library.f90}
-are based in part on software libraries from the Massachusetts Institute
-of Technology, Department of Mechanical Engineering (Cambridge, Massachusetts,
-USA). The non-structured global numbering software was provided by
-Paul F. Fischer (Brown University, Providence, Rhode Island, USA,
-now at Argonne National Laboratory, USA).
-
-OpenDX \urlwithparentheses{http://www.opendx.org} is open-source
-based on IBM Data Explorer, AVS \urlwithparentheses{http://www.avs.com}
-is a trademark of Advanced Visualization Systems, and ParaView \urlwithparentheses{http://www.paraview.org}
-is an open-source visualization platform.
-
-Please e-mail your feedback, questions, comments, and suggestions
-to the CIG Computational Seismology Mailing List \urlwithparentheses{cig-seismo@geodynamics.org}.
-
-\addcontentsline{toc}{chapter}{Copyright}
-
-\chapter*{Copyright}
-
-Main historical authors: Dimitri Komatitsch and Jeroen Tromp
-
-Princeton University, USA, and CNRS / University of Marseille, France
-
-$\copyright$ Princeton University and CNRS / University of Marseille, July 2012
-
-This program is free software; you can redistribute it and/or modify
-it under the terms of the GNU General Public License as published
-by the Free Software Foundation (see Appendix \ref{cha:License}).\\
-
-\textbf{\underline{Evolution of the code:}}\\
-
-MPI v. 3.0, December 2014: many developers.
-Convolutional PML, LDDRK time scheme, bulk attenuation support, simultaneous MPI runs,
-ADIOS file I/O support, coupling with external codes, new seismogram names,
-Deville routines for additional GLL degrees, tomography tools, unit/regression test framework,
-improved CUDA GPUs performance, additonal GEOCUBIT support, better make compilation,
-git versioning system. \\
-
-MPI v. 2.1, July 2012: Max Rietmann, Peter Messmer, Daniel Peter, Dimitri
-Komatitsch, Joseph Charles, Zhinan Xie: support for CUDA GPUs, better
-CFL stability for the Stacey absorbing conditions. \\
-
-MPI v. 2.0, November 2010: Dimitri Komatitsch, Nicolas Le Goff, Roland
-Martin and Pieyre Le Loher, University of Pau, France, Daniel Peter,
-Jeroen Tromp and the Princeton group of developers, Princeton University,
-USA, and Emanuele Casarotti, INGV Roma, Italy: support for CUBIT meshes
-decomposed by SCOTCH; much faster solver using Michel Deville's inlined
-matrix products.\\
-
-MPI v. 1.4 Dimitri Komatitsch, University of Pau, Qinya Liu and others,
-Caltech, September 2006: better adjoint and kernel calculations, faster
-and better I/Os on very large systems, many small improvements and
-bug fixes.\\
-
-MPI v. 1.3 Dimitri Komatitsch, University of Pau, and Qinya Liu, Caltech,
-July 2005: serial version, regular mesh, adjoint and kernel calculations,
-ParaView support.\\
-
-MPI v. 1.2 Min Chen and Dimitri Komatitsch, Caltech, July 2004: full
-anisotropy, volume movie.\\
-
-MPI v. 1.1 Dimitri Komatitsch, Caltech, October 2002: Zhu's Moho map,
-scaling of $V_{s}$ with depth, Hauksson's regional model, attenuation,
-oceans, movies.\\
-
-MPI v. 1.0 Dimitri Komatitsch, Caltech, USA, May 2002: first MPI version
-based on global code.\\
-
-Dimitri Komatitsch, IPG Paris, France, December 1996: first 3-D solver
-for the CM-5 Connection Machine, parallelized on 128 processors using
-Connection Machine Fortran.\\
-
-\addcontentsline{toc}{chapter}{References} \bibliography{bibliography}
-
-\appendix
-%dummy comment inserted by tex2lyx to ensure that this paragraph is not empty
-
-\chapter{\label{cha:Coordinates}Reference Frame Convention}
-
-The code uses the following convention for the Cartesian reference
-frame:
-\begin{itemize}
-\item the $x$ axis points East
-\item the $y$ axis points North
-\item the $z$ axis points up
-\end{itemize}
-Note that this convention is different from both the \citet{AkRi80}
-convention and the Harvard Centroid-Moment Tensor (CMT) convention.
-The Aki \& Richards convention is
-\begin{itemize}
-\item the $x$ axis points North
-\item the $y$ axis points East
-\item the $z$ axis points down
-\end{itemize}
-and the Harvard CMT convention is
-\begin{itemize}
-\item the $x$ axis points South
-\item the $y$ axis points East
-\item the $z$ axis points up
-\end{itemize}
-
-\subsection*{Source and receiver locations}
-
-The SPECFEM3D Cartesian software code internally uses Cartesian coordinates.
-The given locations for sources and receiver locations thus may get
-converted. Sources and receiver locations are read in from the \texttt{CMTSOLUTION}
-(or \texttt{FORCESOLUTION}) and \texttt{STATIONS} files. Note that
-e.g. the \texttt{CMTSOLUTION} and \texttt{FORCESOLUTION} files denotes
-the location by \textquotedbl{}longitude/latitude/depth\textquotedbl{}.
-We read in longitude as $x$ coordinate, latitude as $y$ coordinate.
-
-In case the flag \texttt{SUPPRESS\_UTM\_PROJECTION} is set to \texttt{.false.}
-in the main parameter file (see Chapter~\ref{cha:Creating-Distributed-Databases}),
-the $x$/$y$ coordinates have to be given in degrees and are converted
-to Cartesian coordinates using a UTM conversion with the specified
-UTM zone.
-
-The value for depth (given in $km$ in \texttt{CMTSOLUTION} and \texttt{FORCESOLUTION})
-or burial depth (given in $m$ in \texttt{STATIONS}) is evaluated
-with respect to the surface of the mesh at the specified $x$/$y$
-location to find the corresponding $z$ coordinate. It is possible
-to use this depth value directly as $z$ coordinate by changing the
-flag \texttt{USE\_SOURCES\_RECVS\_Z} to \texttt{.true.} in the file
-\texttt{constants.h} located in the \texttt{src/shared/} subdirectory.
-
-\subsection*{Seismogram outputs}
-
-The seismogram output directions are given in Cartesian $x$/$y$/$z$
-directions and not rotated any further. Changing flags in \texttt{constants.h}
-in the \texttt{src/shared/} subdirectory only rotates the seismogram
-outputs if receivers are forced to be located at the surface (\texttt{RECVS\_CAN\_BE\_BURIED\_EXT\_MESH}
-set to \texttt{.false.}) and the normal to the surface at the receiver
-location should be used (\texttt{EXT\_MESH\_RECV\_NORMAL} set to \texttt{.true.})
-as vertical. In this case, the outputs are rotated to have the vertical
-component normal to the surface of the mesh, $x$ and $y$ directions
-are somewhat arbitrary orthogonal directions along the surface.
-
-For the labeling of the seismogram channels, see Appendix~\ref{cha:channel-codes}.
-Additionally, we add labels to the synthetics using the following
-convention: For a receiver station located in an
-\begin{description}
-\item [{elastic domain:}] ~\\
-
-\begin{itemize}
-\item \texttt{semd} for the displacement vector
-\item \texttt{semv} for the velocity vector
-\item \texttt{sema} for the acceleration vector
-\end{itemize}
-\item [{acoustic domain:}] ~\\
- (please note that receiver stations in acoustic domains must be buried.
-This is due to the free surface condition which enforces a zero pressure
-at the surface.)
-
-\begin{itemize}
-\item \texttt{semd} for the displacement vector
-\item \texttt{semv} for the velocity vector
-\item \texttt{sema} for pressure which will be stored in the vertical component
-\texttt{Z}. Note that pressure in the acoustic media is isotropic
-and thus the pressure record would be the same in the other two component
-directions. We therefore use the other two seismogram components to
-store the acoustic potential in component \texttt{X} (or \texttt{N})
-and the first time derivative of the acoustic potential in component
-\texttt{Y} (or \texttt{E}).
-\end{itemize}
-\end{description}
-The seismograms are by default written out in ASCII-format to the
-\texttt{OUTPUT\_FILES/} subdirectory by each parallel process. You
-can change this behavior by changing the following flags in the \texttt{constants.h}
-file located in the \texttt{src/shared/} subdirectory:
-\begin{description}
-\item [{\texttt{SEISMOGRAMS\_BINARY}}] set to \texttt{.true.} to have seismograms
-written out in binary format.
-\item [{\texttt{WRITE\_SEISMOGRAMS\_BY\_MASTER}}] Set to \texttt{.true.}
-to have only the master process writing out seismograms. This can
-be useful on a cluster, where only the master process node has access
-to the output directory.
-\item [{\texttt{USE\_OUTPUT\_FILES\_PATH}}] Set to \texttt{.false.} to
-have the seismograms output to \texttt{LOCAL\_PATH} directory specified
-in the main parameter file \texttt{DATA/Par\_file}.
-\end{description}
-
-\chapter{\label{cha:channel-codes}Channel Codes of Seismograms}
-
-Seismic networks, such as the Global Seismographic Network (GSN),
-generally involve various types of instruments with different bandwidths,
-sampling properties and component configurations. There are standards
-to name channel codes depending on instrument properties. IRIS \texttt{(www.iris.edu)}
-uses SEED format for channel codes, which are represented by three
-letters, such as \texttt{LHN}, \texttt{BHZ}, etc. In older versions
-of the SPECFEM3D Cartesian package, a common format was used for the
-channel codes of all seismograms, which was \texttt{BHE/BHN/BHZ} for
-three components. To avoid confusion when comparison are made to observed
-data, we are now using the FDSN convention (\url{http://www.fdsn.org/})
-for SEM seismograms. In the following, we give a brief explanation
-of the FDSN convention used by IRIS, and how it is adopted in SEM
-seismograms. Please visit \texttt{www.iris.edu/manuals/SEED\_appA.htm}
-for further information.\\
-
-\noindent \texttt{\textbf{Band code:}} The first letter in the channel
-code denotes the band code of seismograms, which depends on the response
-band and the sampling rate of instruments. The list of band codes
-used by IRIS is shown in Figure \ref{fig:IRIS_band_codes}. The sampling
-rate of SEM synthetics is controlled by the resolution of simulations
-rather than instrument properties. However, for consistency, we follow
-the FDSN convention for SEM seismograms governed by their sampling
-rate. For SEM synthetics, we consider band codes for which $dt\leq1$
-s. IRIS also considers the response band of instruments. For instance,
-short-period and broad-band seismograms with the same sampling rate
-correspond to different band codes, such as S and B, respectively.
-In such cases, we consider SEM seismograms as broad band, ignoring
-the corner period ($\geq10$ s) of the response band of instruments
-(note that at these resolutions, the minimum period in the SEM synthetics
-will be less than $10$ s). Accordingly, when you run a simulation
-the band code will be chosen depending on the resolution of the synthetics,
-and channel codes of SEM seismograms will start with either \texttt{L},
-\texttt{M}, \texttt{B}, \texttt{H}, \texttt{C} or \texttt{F}, shown
-by red color in the figure.\\
-
-\begin{figure}[ht]
-\noindent \begin{centering}
-\includegraphics[scale=0.6]{figures/IRIS_band_codes}
-\par\end{centering}
-
-\caption{The band code convention is based on the sampling rate and the response
-band of instruments. Please visit \texttt{www.iris.edu/manuals/SEED\_appA.htm}
-for further information. Grey rows show the relative band-code range
-in SPECFEM3D Cartesian, and the band codes used to name SEM seismograms
-are denoted in red.}
-
-\label{fig:IRIS_band_codes}
-\end{figure}
-
-\noindent \texttt{\textbf{Instrument code:}} The second letter in
-the channel code corresponds to instrument codes, which specify the
-family to which the sensor belongs. For instance, \texttt{H} and \texttt{L}
-are used for high-gain and low-gain seismometers, respectively. The
-instrument code of SEM seismograms will always be \texttt{X}, as assigned
-by FDSN for synthetic seismograms. \\
-
-\noindent \texttt{\textbf{Orientation code:}} The third letter in
-channel codes is an orientation code, which generally describes the
-physical configuration of the components of instrument packages. SPECFEM3D
-Cartesian uses the traditional orientation code \texttt{E/N/Z} (East-West,
-North-South, Vertical) for three components when a UTM projection
-is used. If the UTM conversion is suppressed, i.e. the flag \texttt{SUPPRESS\_UTM\_PROJECTION}
-is set to \texttt{.true.}, then the three components are labelled
-\texttt{X/Y/Z} according to the Cartesian reference frame. \\
-
-\noindent \texttt{\textbf{EXAMPLE:}} The sampling rate is given by
-\texttt{DT} in the main parameter file \texttt{Par\_file} located
-in the \texttt{DATA/} subdirectory. Depending on the resolution of
-your simulations, if you choose a sampling rate greater than $0.01$
-s and less than $1$ s, a seismogram recording displacements on the
-vertical component of a station \texttt{ASBS} (network \texttt{AZ})
-will be named \texttt{AZ.ASBS.MXZ.semd.sac}, whereas it will be \texttt{AZ.ASBS.BXZ.semd.sac},
-if the sampling rate is greater than 0.0125 and less equal to 0.1
-s.
-
-\chapter{\label{cha:Troubleshooting}Troubleshooting}
-
-\section*{FAQ}
-\begin{description}
-\item [{configuration fails:}]~
-
-Examine the log file 'config.log'. It contains detailed informations.
-In many cases, the paths to these specific compiler commands F90,
-CC and MPIF90 will not be correct if `./configure` fails.
-
-Please make sure that you have a working installation of a Fortran
-compiler, a C compiler and an MPI implementation. You should be able
-to compile this little program code:
-\begin{lyxcode}
-program~main~~~\\
-~include~'mpif.h'~~~\\
-~integer,~parameter~::~CUSTOM\_MPI\_TYPE~=~MPI\_REAL~~~\\
-~integer~ier~~~\\
-~call~MPI\_INIT(ier)~~~\\
-~call~MPI\_BARRIER(MPI\_COMM\_WORLD,ier)~~~\\
-~call~MPI\_FINALIZE(ier)~~~\\
-~end
-\end{lyxcode}
-\item [{compilation fails stating:}] ~\\
-
-\begin{lyxcode}
-...~~~\\
-~obj/program\_generate\_databases.o:~In~function~`MAIN\_\_':~~~\\
-~program\_generate\_databases.f90:(.text+0x14):~undefined~reference~to~`\_gfortran\_set\_std'~~~\\
-~...~~~\\
-
-\end{lyxcode}
-
-Make sure you're pointing to the right 'mpif90' wrapper command.
-
-Normally, this message will appear when you are mixing two different
-Fortran compilers. That is, using e.g. gfortran to compile non-MPI
-files and mpif90, wrapper provided for e.g. ifort, to compile MPI-files.
-
-fix: e.g. specify > ./configure FC=gfortran MPIF90=/usr/local/openmpi-gfortran/bin/mpif90
-
-\item [{after executing \texttt{xmeshfem3D} I've got elements with skewness of 81\% percent, what does this mean:}] Look
-at the skewness table printed in the \texttt{output\_mesher.txt} file
-after executing \texttt{xmeshfem3D} for the example given in \texttt{EXAMPLES/meshfem3D\_examples/simple\_model/}:
-
-\begin{lyxcode}
-...~~\\
-~histogram~of~skewness~(0.~good~-~1.~bad):~~~\\
-~0.0000000E+00~-~5.0000001E-02~27648~81.81818~\%~~~\\
-~5.0000001E-02~-~0.1000000~0~0.0000000E+00~\%~~~\\
-~...~~~\\
-
-\end{lyxcode}
-
-The first line means that you have 27,648 elements with a skewness
-value between 0 and 0.05 (which means the element is basically not
-skewed, just plain regular hexahedral element). The total number of
-elements you have in this mesh is (see in the \texttt{output\_mesher.txt}
-file a bit further down):
-\begin{lyxcode}
-...~~\\
-~total~number~of~elements~in~entire~mesh:~33792~~~\\
-~...~~~\\
-
-\end{lyxcode}
-
-which gives you that: 27,648 / 33,792 $\sim$ 81.8 \% of all elements
-are not skewed, i.e. regular elements. a fantastic value :)
-
-The histogram lists for this mesh also some stronger skewed elements,
-for example the worst ones belong to:
-\begin{lyxcode}
-...~~\\
-~0.6000000~-~0.6500000~2048~6.060606~\%~~~\\
-~...~~~\\
-
-\end{lyxcode}
-
-about 6 \% of all elements have distortions with a skewness value
-between 0.6 and 0.65. The skewness values give you a hint of how good
-your mesh is. In an ideal world, you would want to have no distortions,
-just like the 81\% from above. Those elements give you the best approximate
-values by the GLL quadrature used in the spectral-element method.
-However, having weakly distorted elements is still fine and the solutions
-are still accurate enough. So empirically, values up to around 0.7
-are tolerable, above that you should consider remeshing...
-
-To give you an idea why some of the elements are distorted, see the
-following figure \ref{fig:mesh.vp} of the mesh you obtain in the
-example \texttt{EXAMPLES/meshfem3D\_examples/simple\_model/}.
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=0.9\textwidth]{figures/mesh_vp}
-\par\end{centering}
-
-\caption{Paraview visualization using the mesh vtk-files for the example given
-in \texttt{EXAMPLES/meshfem3D\_examples/simple\_model/}.}
-
-\label{fig:mesh.vp}
-\end{figure}
-
-You will see that the mesh contains a doubling layer, where we stitch
-elements together such that the size of two elements will transition
-to the size of one element (very useful to keep the ratio of wavespeed
-/ element\_size about constant). Those elements in this doubling layer
-have higher skewness values and make up those 6 \% in the histogram.
-
-\item [{the code gives following error message "need at least one receiver":}] This
-means that no stations given in the input file \texttt{DATA/STATIONS}
-could be located within the dimensions of the mesh. This can happen
-for example when the mesh was created with the in-house mesher \texttt{xmeshfem3D}
-while using the Universal Transverse Mercator (UTM) projection but
-the simulation with \texttt{xspecfem3D} was suppressing this projection
-from latitude/longitude to x/y/z coordinates.
-
-In such cases, try to change your \texttt{DATA/Par\_file} and set
-e.g.:
-\begin{lyxcode}
-SUPPRESS\_UTM\_PROJECTION~=~.false.~~~\\
-
-\end{lyxcode}
-
-to be the same in \texttt{Mesh\_Par\_file} and \texttt{Par\_file}.
-This flag should be identical when using the in-house mesher \texttt{xmeshfem3D},
-\texttt{xgenerate\_databases} and \texttt{xspecfem3D} together to
-run simulations.
-
-The flag determines if the coordinates you specify for your source
-and station locations are given as lat/lon degrees and must be converted
-to UTM coordinates. As an example, if you use \texttt{.false.} within
-\texttt{Mesh\_Par\_file} then you create a mesh with \texttt{xmeshfem3D}
-using the UTM projection from lat/lon as input format to UTM projected
-coordinates to store the mesh point positions, which is fine. The
-error then may occur if in the \texttt{Par\_file} you have this set
-to \texttt{.true.} so that the \texttt{xgenerate\_databases} and \texttt{xspecfem3D}
-suppress the UTM projection and assume that all coordinates you use
-now for source and receiver locations are given in meters (that is,
-converted) already. So it won't find the specified locations in the
-used mesh. As a solutions, just change the flag in \texttt{Par\_file}
-to be the same as in \texttt{Mesh\_Par\_file} and rerun \texttt{xgenerate\_databases}
-and \texttt{xspecfem3D} to make sure that your simulation works fine.
-
-\item [{I get the following error message "forward simulation became unstable and blew up":}] In
-most cases this means that your time step size \texttt{DT} is chosen
-too big. Look at your files \texttt{output\_mesher.txt} or \texttt{output\_solver.txt}
-created in the folder \texttt{OUTPUT\_FILES}. In these output files,
-find the section:
-
-\begin{lyxcode}
-...~~~\\
-~{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}~~~\\
-~{*}{*}{*}~Verification~of~simulation~parameters~{*}{*}{*}~~~\\
-~{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}~~~\\
-~...~~~\\
-~{*}{*}{*}~Minimum~period~resolved~=~4.308774~~~\\
-~{*}{*}{*}~Maximum~suggested~time~step~=~6.8863556E-02~~~\\
-~...~~~\\
-~...~~~\\
-
-\end{lyxcode}
-
-then change \texttt{DT} in the \texttt{DATA/Par\_file} to be somewhere
-close to the maximum suggested time step. In the example above: \\
- \texttt{DT = 0.05d0} \\
- would (most probably) work fine. It could be also bigger than the
-0.068 s suggested. This depends a bit on the distortions of your mesh
-elements. The more regular they are, the bigger you can choose \texttt{DT}.
-Just play with this value a bit and see when the simulation becomes
-stable ...
-
-\end{description}
-
-\chapter{\label{cha:License}License }
-
-\textbf{GNU GENERAL PUBLIC LICENSE Version 2, June 1991. Copyright
-(C) 1989, 1991 Free Software Foundation, Inc. 59 Temple Place, Suite
-330, Boston, MA 02111-1307 USA} \\
- Everyone is permitted to copy and distribute verbatim copies of this
-license document, but changing it is not allowed.
-
-\section*{Preamble}
-
-The licenses for most software are designed to take away your freedom
-to share and change it. By contrast, the GNU General Public License
-is intended to guarantee your freedom to share and change free software
--- to make sure the software is free for all its users. This General
-Public License applies to most of the Free Software Foundation's software
-and to any other program whose authors commit to using it. (Some other
-Free Software Foundation software is covered by the GNU Library General
-Public License instead.) You can apply it to your programs, too.
-
-When we speak of free software, we are referring to freedom, not price.
-Our General Public Licenses are designed to make sure that you have
-the freedom to distribute copies of free software (and charge for
-this service if you wish), that you receive source code or can get
-it if you want it, that you can change the software or use pieces
-of it in new free programs; and that you know you can do these things.
-
-To protect your rights, we need to make restrictions that forbid anyone
-to deny you these rights or to ask you to surrender the rights. These
-restrictions translate to certain responsibilities for you if you
-distribute copies of the software, or if you modify it.
-
-For example, if you distribute copies of such a program, whether gratis
-or for a fee, you must give the recipients all the rights that you
-have. You must make sure that they, too, receive or can get the source
-code. And you must show them these terms so they know their rights.
-
-We protect your rights with two steps:
-\begin{enumerate}
-\item Copyright the software, and
-\item Offer you this license which gives you legal permission to copy, distribute
-and/or modify the software.
-\end{enumerate}
-Also, for each author's protection and ours, we want to make certain
-that everyone understands that there is no warranty for this free
-software. If the software is modified by someone else and passed on,
-we want its recipients to know that what they have is not the original,
-so that any problems introduced by others will not reflect on the
-original authors' reputations.
-
-Finally, any free program is threatened constantly by software patents.
-We wish to avoid the danger that redistributors of a free program
-will individually obtain patent licenses, in effect making the program
-proprietary. To prevent this, we have made it clear that any patent
-must be licensed for everyone's free use or not licensed at all.
-
-The precise terms and conditions for copying, distribution and modification
-follow.
-
-\section*{GNU GENERAL PUBLIC LICENSE TERMS AND CONDITIONS FOR COPYING, DISTRIBUTION
-AND MODIFICATION }
-\begin{itemize}
-\item 0.This License applies to any program or other work which contains
-a notice placed by the copyright holder saying it may be distributed
-under the terms of this General Public License. The ``Program''
-below refers to any such program or work, and a ``work based on the
-Program'' means either the Program or any derivative work under copyright
-law: that is to say, a work containing the Program or a portion of
-it, either verbatim or with modifications and/or translated into another
-language. (Hereinafter, translation is included without limitation
-in the term ``modification.'') Each licensee is addressed as ``you.''\\
- \\
- Activities other than copying, distribution and modification are
-not covered by this License; they are outside its scope. The act of
-running the Program is not restricted, and the output from the Program
-is covered only if its contents constitute a work based on the Program
-(independent of having been made by running the Program). Whether
-that is true depends on what the Program does.\end{itemize}
-\begin{enumerate}
-\item You may copy and distribute verbatim copies of the Program's source
-code as you receive it, in any medium, provided that you conspicuously
-and appropriately publish on each copy an appropriate copyright notice
-and disclaimer of warranty; keep intact all the notices that refer
-to this License and to the absence of any warranty; and give any other
-recipients of the Program a copy of this License along with the Program.
-
-You may charge a fee for the physical act of transferring a copy,
-and you may at your option offer warranty protection in exchange for
-a fee.
-
-\item You may modify your copy or copies of the Program or any portion of
-it, thus forming a work based on the Program, and copy and distribute
-such modifications or work under the terms of Section 1 above, provided
-that you also meet all of these conditions:
-
-\begin{enumerate}
-\item You must cause the modified files to carry prominent notices stating
-that you changed the files and the date of any change.
-\item You must cause any work that you distribute or publish, that in whole
-or in part contains or is derived from the Program or any part thereof,
-to be licensed as a whole at no charge to all third parties under
-the terms of this License.
-\item If the modified program normally reads commands interactively when
-run, you must cause it, when started running for such interactive
-use in the most ordinary way, to print or display an announcement
-including an appropriate copyright notice and a notice that there
-is no warranty (or else, saying that you provide a warranty) and that
-users may redistribute the program under these conditions, and telling
-the user how to view a copy of this License. (Exception: if the Program
-itself is interactive but does not normally print such an announcement,
-your work based on the Program is not required to print an announcement.)
-\end{enumerate}
-
-These requirements apply to the modified work as a whole. If identifiable
-sections of that work are not derived from the Program, and can be
-reasonably considered independent and separate works in themselves,
-then this License, and its terms, do not apply to those sections when
-you distribute them as separate works. But when you distribute the
-same sections as part of a whole which is a work based on the Program,
-the distribution of the whole must be on the terms of this License,
-whose permissions for other licensees extend to the entire whole,
-and thus to each and every part regardless of who wrote it.
-
-Thus, it is not the intent of this section to claim rights or contest
-your rights to work written entirely by you; rather, the intent is
-to exercise the right to control the distribution of derivative or
-collective works based on the Program.
-
-In addition, mere aggregation of another work not based on the Program
-with the Program (or with a work based on the Program) on a volume
-of a storage or distribution medium does not bring the other work
-under the scope of this License.
-
-\item You may copy and distribute the Program (or a work based on it, under
-Section 2) in object code or executable form under the terms of Sections
-1 and 2 above provided that you also do one of the following:
-
-\begin{enumerate}
-\item Accompany it with the complete corresponding machine-readable source
-code, which must be distributed under the terms of Sections 1 and
-2 above on a medium customarily used for software interchange; or,
-\item Accompany it with a written offer, valid for at least three years,
-to give any third party, for a charge no more than your cost of physically
-performing source distribution, a complete machine-readable copy of
-the corresponding source code, to be distributed under the terms of
-Sections 1 and 2 above on a medium customarily used for software interchange;
-or,
-\item Accompany it with the information you received as to the offer to
-distribute corresponding source code. (This alternative is allowed
-only for noncommercial distribution and only if you received the program
-in object code or executable form with such an offer, in accord with
-Subsection b above.)
-\end{enumerate}
-
-The source code for a work means the preferred form of the work for
-making modifications to it. For an executable work, complete source
-code means all the source code for all modules it contains, plus any
-associated interface definition files, plus the scripts used to control
-compilation and installation of the executable. However, as a special
-exception, the source code distributed need not include anything that
-is normally distributed (in either source or binary form) with the
-major components (compiler, kernel, and so on) of the operating system
-on which the executable runs, unless that component itself accompanies
-the executable.
-
-If distribution of executable or object code is made by offering access
-to copy from a designated place, then offering equivalent access to
-copy the source code from the same place counts as distribution of
-the source code, even though third parties are not compelled to copy
-the source along with the object code.
-
-\item You may not copy, modify, sublicense, or distribute the Program except
-as expressly provided under this License. Any attempt otherwise to
-copy, modify, sublicense or distribute the Program is void, and will
-automatically terminate your rights under this License. However, parties
-who have received copies, or rights, from you under this License will
-not have their licenses terminated so long as such parties remain
-in full compliance.
-\item You are not required to accept this License, since you have not signed
-it. However, nothing else grants you permission to modify or distribute
-the Program or its derivative works. These actions are prohibited
-by law if you do not accept this License. Therefore, by modifying
-or distributing the Program (or any work based on the Program), you
-indicate your acceptance of this License to do so, and all its terms
-and conditions for copying, distributing or modifying the Program
-or works based on it.
-\item Each time you redistribute the Program (or any work based on the Program),
-the recipient automatically receives a license from the original licensor
-to copy, distribute or modify the Program subject to these terms and
-conditions. You may not impose any further restrictions on the recipients'
-exercise of the rights granted herein. You are not responsible for
-enforcing compliance by third parties to this License.
-\item If, as a consequence of a court judgment or allegation of patent infringement
-or for any other reason (not limited to patent issues), conditions
-are imposed on you (whether by court order, agreement or otherwise)
-that contradict the conditions of this License, they do not excuse
-you from the conditions of this License. If you cannot distribute
-so as to satisfy simultaneously your obligations under this License
-and any other pertinent obligations, then as a consequence you may
-not distribute the Program at all. For example, if a patent license
-would not permit royalty-free redistribution of the Program by all
-those who receive copies directly or indirectly through you, then
-the only way you could satisfy both it and this License would be to
-refrain entirely from distribution of the Program.
-
-If any portion of this section is held invalid or unenforceable under
-any particular circumstance, the balance of the section is intended
-to apply and the section as a whole is intended to apply in other
-circumstances.
-
-It is not the purpose of this section to induce you to infringe any
-patents or other property right claims or to contest validity of any
-such claims; this section has the sole purpose of protecting the integrity
-of the free software distribution system, which is implemented by
-public license practices. Many people have made generous contributions
-to the wide range of software distributed through that system in reliance
-on consistent application of that system; it is up to the author/donor
-to decide if he or she is willing to distribute software through any
-other system and a licensee cannot impose that choice.
-
-This section is intended to make thoroughly clear what is believed
-to be a consequence of the rest of this License.
-
-\item If the distribution and/or use of the Program is restricted in certain
-countries either by patents or by copyrighted interfaces, the original
-copyright holder who places the Program under this License may add
-an explicit geographical distribution limitation excluding those countries,
-so that distribution is permitted only in or among countries not thus
-excluded. In such case, this License incorporates the limitation as
-if written in the body of this License.
-\item The Free Software Foundation may publish revised and/or new versions
-of the General Public License from time to time. Such new versions
-will be similar in spirit to the present version, but may differ in
-detail to address new problems or concerns.
-
-Each version is given a distinguishing version number. If the Program
-specifies a version number of this License which applies to it and
-``any later version,'' you have the option of following the terms
-and conditions either of that version or of any later version published
-by the Free Software Foundation. If the Program does not specify a
-version number of this License, you may choose any version ever published
-by the Free Software Foundation.
-
-\item If you wish to incorporate parts of the Program into other free programs
-whose distribution conditions are different, write to the author to
-ask for permission. For software which is copyrighted by the Free
-Software Foundation, write to the Free Software Foundation; we sometimes
-make exceptions for this. Our decision will be guided by the two goals
-of preserving the free status of all derivatives of our free software
-and of promoting the sharing and reuse of software generally.
-\end{enumerate}
-
-\subsection*{NO WARRANTY }
-\begin{itemize}
-\item 11.BECAUSE THE PROGRAM IS LICENSED FREE OF CHARGE, THERE IS NO WARRANTY
-FOR THE PROGRAM, TO THE EXTENT PERMITTED BY APPLICABLE LAW. EXCEPT
-WHEN OTHERWISE STATED IN WRITING THE COPYRIGHT HOLDERS AND/OR OTHER
-PARTIES PROVIDE THE PROGRAM ``AS IS'' WITHOUT WARRANTY OF ANY KIND,
-EITHER EXPRESSED OR IMPLIED, INCLUDING, BUT NOT LIMITED TO, THE IMPLIED
-WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE.
-THE ENTIRE RISK AS TO THE QUALITY AND PERFORMANCE OF THE PROGRAM IS
-WITH YOU. SHOULD THE PROGRAM PROVE DEFECTIVE, YOU ASSUME THE COST
-OF ALL NECESSARY SERVICING, REPAIR OR CORRECTION.
-\item 12.IN NO EVENT UNLESS REQUIRED BY APPLICABLE LAW OR AGREED TO IN WRITING
-WILL ANY COPYRIGHT HOLDER, OR ANY OTHER PARTY WHO MAY MODIFY AND/OR
-REDISTRIBUTE THE PROGRAM AS PERMITTED ABOVE, BE LIABLE TO YOU FOR
-DAMAGES, INCLUDING ANY GENERAL, SPECIAL, INCIDENTAL OR CONSEQUENTIAL
-DAMAGES ARISING OUT OF THE USE OR INABILITY TO USE THE PROGRAM (INCLUDING
-BUT NOT LIMITED TO LOSS OF DATA OR DATA BEING RENDERED INACCURATE
-OR LOSSES SUSTAINED BY YOU OR THIRD PARTIES OR A FAILURE OF THE PROGRAM
-TO OPERATE WITH ANY OTHER PROGRAMS), EVEN IF SUCH HOLDER OR OTHER
-PARTY HAS BEEN ADVISED OF THE POSSIBILITY OF SUCH DAMAGES.
-\end{itemize}
-
-\section*{END OF TERMS AND CONDITIONS }
-
-\subsection*{How to Apply These Terms to Your New Programs}
-
-If you develop a new program, and you want it to be of the greatest
-possible use to the public, the best way to achieve this is to make
-it free software which everyone can redistribute and change under
-these terms.
-
-To do so, attach the following notices to the program. It is safest
-to attach them to the start of each source file to most effectively
-convey the exclusion of warranty; and each file should have at least
-the ``copyright'' line and a pointer to where the full notice is
-found. For example:
-\begin{quote}
-One line to give the program's name and a brief idea of what it does.
-Copyright {\footnotesize $\copyright$ (}year) (name of author)
-
-This program is free software; you can redistribute it and/or modify
-it under the terms of the GNU General Public License as published
-by the Free Software Foundation; either version 2 of the License,
-or (at your option) any later version.
-
-This program is distributed in the hope that it will be useful, but
-WITHOUT ANY WARRANTY; without even the implied warranty of MERCHANTABILITY
-or FITNESS FOR A PARTICULAR PURPOSE. See the GNU General Public License
-for more details.
-
-You should have received a copy of the GNU General Public License
-along with this program; if not, write to the Free Software Foundation,
-Inc., 59 Temple Place, Suite 330, Boston, MA 02111-1307 USA
-\end{quote}
-Also add information on how to contact you by electronic and paper
-mail.
-
-If the program is interactive, make it output a short notice like
-this when it starts in an interactive mode:
-\begin{quote}
-Gnomovision version 69, Copyright $\copyright$ year name of author
-Gnomovision comes with ABSOLUTELY NO WARRANTY; for details type `show
-w'. This is free software, and you are welcome to redistribute it
-under certain conditions; type `show c' for details.
-\end{quote}
-The hypothetical commands `show w' and `show c' should show the appropriate
-parts of the General Public License. Of course, the commands you use
-may be called something other than `show w' and `show c'; they could
-even be mouse-clicks or menu items -- whatever suits your program.
-
-You should also get your employer (if you work as a programmer) or
-your school, if any, to sign a ``copyright disclaimer'' for the
-program, if necessary. Here is a sample; alter the names:
-\begin{quote}
-Yoyodyne, Inc., hereby disclaims all copyright interest in the program
-`Gnomovision' (which makes passes at compilers) written by James Hacker.
-
-(signature of Ty Coon)\\
- 1 April 1989 \\
- Ty Coon, President of Vice
-\end{quote}
-This General Public License does not permit incorporating your program
-into proprietary programs. If your program is a subroutine library,
-you may consider it more useful to permit linking proprietary applications
-with the library. If this is what you want to do, use the GNU Library
-General Public License instead of this License.
-\end{document}

### 2015-04-24 doc/USER_MANUAL/09_noise_simulations.tex
updates user manual to use facilitate the generation of a markdown version to the github wiki

--- 2015-04-18 doc/USER_MANUAL/manual_SPECFEM3D_Cartesian.tex
@@
-% !TEX encoding = IsoLatin
-%% LyX 2.0.2 created this file.  For more info, see http://www.lyx.org/.
-%% Do not edit unless you really know what you are doing.
-\documentclass[oneside,english]{book}
-\usepackage[T1]{fontenc}
-%% DK DK we do not accept accented letters, use the old LaTeX syntax
-%% DK DK for them instead: \'e \`e  \"e \^e and so on.
-%% DK DK \usepackage[latin1]{inputenc}
-\usepackage[letterpaper]{geometry}
-\geometry{verbose,tmargin=1in,bmargin=1in,lmargin=1in,rmargin=1in}
-\setcounter{secnumdepth}{3}
-\setcounter{tocdepth}{3}
-\usepackage{babel}
-\usepackage{float}
-\usepackage{textcomp}
-\usepackage{url}
-\usepackage{amsmath}
-\usepackage{amssymb}
-\usepackage{graphicx}
-\usepackage[authoryear]{natbib}
-\usepackage[unicode=true,pdfusetitle,
- bookmarks=true,bookmarksnumbered=true,bookmarksopen=false,
- breaklinks=false,pdfborder={0 0 0},backref=section,colorlinks=false,pdfpagemode=None]
- {hyperref}
-\hypersetup{
- pdftex,pdfstartview=FitH,pdfpagelayout=SinglePage}
-\usepackage{breakurl}
-
-\makeatletter
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% LyX specific LaTeX commands.
-\newcommand{\noun}[1]{\textsc{#1}}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Textclass specific LaTeX commands.
-\newenvironment{lyxcode}
-{\par\begin{list}{}{
-\setlength{\rightmargin}{\leftmargin}
-\setlength{\listparindent}{0pt}% needed for AMS classes
-\raggedright
-\setlength{\itemsep}{0pt}
-\setlength{\parsep}{0pt}
-\normalfont\ttfamily}%
- \item[]}
-{\end{list}}
-\newenvironment{lyxlist}[1]
-{\begin{list}{}
-{\settowidth{\labelwidth}{#1}
- \setlength{\leftmargin}{\labelwidth}
- \addtolength{\leftmargin}{\labelsep}
- \renewcommand{\makelabel}[1]{##1\hfil}}}
-{\end{list}}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% User specified LaTeX commands.
-%\renewcommand{\baselinestretch}{1.5}
-
-% figures
-\usepackage[dvips]{epsfig}
-
-% fonts
-\usepackage{times}
-
-% hyperlinks to sections and references
-
-\newcommand{\urlwithparentheses}[1]{(\url{#1})}
-
-% biblio GJI
-\bibliographystyle{abbrvnat}
-
-\newcommand{\nexxi}{\mbox{\texttt{NEX\_XI\/}}}
-\newcommand{\nexeta}{\mbox{\texttt{NEX\_ETA\/}}}
-\newcommand{\nprocxi}{\mbox{\texttt{NPROC\_XI\/}}}
-\newcommand{\nproceta}{\mbox{\texttt{NPROC\_ETA\/}}}
-\newcommand{\nchunks}{\mbox{\texttt{NCHUNKS\/}}}
-\newcommand{\eg}{e.g.,}
-
-% colors for the text
-\@ifundefined{definecolor}
- {\usepackage{color}}{}
-\newcommand{\red}[1]{\textcolor{Red}{#1}}
-
-\usepackage{babel}
-
-\makeatother
-
-\begin{document}
-\begin{center}
-\thispagestyle{empty}\vspace*{-1.8truecm} %
-\makebox[1\textwidth]{%
-\includegraphics[width=0.83\paperwidth]{figures/specfem_3d_Cartesian-cover} %
-}
-\par\end{center}
-
-\title{\textbf{SPECFEM3D Cartesian}\\
- \textbf{User Manual}}
-
-\author{$\copyright$ Princeton University (USA), CNRS and University of Marseille (France),\\
-and ETH Z\"urich (Switzerland)\\
-Version 3.0}
-
-\date{\today}
-
-\maketitle
-
-\section*{Authors}
-
-\noindent The SPECFEM3D Cartesian package was first developed by Dimitri
-Komatitsch and Jean-Pierre Vilotte at Institut de Physique du Globe
-(IPGP) in Paris, France from 1995 to 1997 and then by Dimitri Komatitsch
-and Jeroen Tromp at Harvard University and Caltech, USA, starting in 1998.
-The story started on April 4, 1995, when Prof. Yvon Maday from CNRS and University of Paris, France, gave a lecture to
-Dimitri Komatitsch and Jean-Pierre Vilotte at IPG about the nice properties of the Legendre spectral-element method with diagonal mass matrix that he had used for
-other equations. We are deeply indebted and thankful to him for that.
-That followed a visit by Dimitri Komatitsch to OGS (Istituto Nazionale di Oceanografia e di Geofisica Sperimentale) in Trieste, Italy, in February 1995
-to meet with G\'eza Seriani and Enrico Priolo, who introduced him to their 2D Chebyshev version of the spectral-element method with a non-diagonal mass matrix.
-We are deeply indebted and thankful to them for that.
-
-%%%%%%%% please do NOT merge the lines of the paragraph below, it makes it difficult to insert new names in alphabetical order; leave a single name per line
-%%%%%%%% please do NOT merge the lines of the paragraph below, it makes it difficult to insert new names in alphabetical order; leave a single name per line
-%%%%%%%% please do NOT merge the lines of the paragraph below, it makes it difficult to insert new names in alphabetical order; leave a single name per line
-Since then it has been developed and maintained by a development team:
-in alphabetical order,
-Jean-Paul (Pablo) Ampuero,
-Kangchen Bai,
-Piero Basini,
-C\'eline Blitz,
-Alexis Bottero,
-Ebru Bozda\u{g},
-Emanuele Casarotti,
-Joseph Charles,
-Min Chen,
-Paul Cristini,
-Percy Galvez,
-Dominik G\"oddeke,
-Vala Hj\"orleifsd\'ottir,
-Sue Kientz,
-Dimitri Komatitsch,
-Jes\'us Labarta,
-Nicolas Le Goff,
-Pieyre Le Loher,
-Matthieu Lefebvre,
-Qinya Liu,
-Yang Luo,
-Alessia Maggi,
-Federica Magnoni,
-Roland Martin,
-Ren\'e Matzen,
-Dennis McRitchie,
-Matthias Meschede,
-Peter Messmer,
-David Mich\'ea,
-Surendra Nadh Somala,
-Tarje Nissen-Meyer,
-Daniel Peter,
-Max Rietmann,
-Elliott Sales de Andrade,
-Brian Savage,
-Bernhard Schuberth,
-Anne Sieminski,
-Leif Strand,
-Carl Tape,
-Jeroen Tromp,
-Jean-Pierre Vilotte,
-Zhinan Xie,
-Chang-Hua Zhang,
-Hejun Zhu.\\
-
-The cover graphic of the manual was created by Santiago Lombeyda from
-Caltech's Center for Advanced Computing Research (CACR) \urlwithparentheses{http://www.cacr.caltech.edu}.\\
-
-The code is released open-source under the GNU version 2+ license, see the license at the end of this manual.\\
-
-\newpage{}
-\section*{Current and past main participants or main sponsors of the SPECFEM project (in no particular order)}
-%
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=0.125\textwidth]{figures/logo_cnrs}\vspace*{2truemm}
-\includegraphics[width=0.125\textwidth]{figures/logo_princeton}\vspace*{2truemm}
-\includegraphics[width=0.15\textwidth]{figures/logo_aix_marseille_universite}\vspace*{0.02truemm}
-\includegraphics[width=0.15\textwidth]{figures/logo_ETH}\vspace*{2truemm}
-\includegraphics[width=0.12\textwidth]{figures/logo_CSC_China}\vspace*{0.02truemm}
-\includegraphics[width=0.15\textwidth]{figures/logo_inria}\vspace*{2truemm}
-\includegraphics[width=0.14\textwidth]{figures/logo_UPPA}
-\par\end{centering}
-%
-\vspace*{-2truemm}
-%
-\noindent \begin{centering}
-\includegraphics[width=0.15\textwidth]{figures/logo_NVIDIA}\vspace*{2truemm}
-\includegraphics[width=0.125\textwidth]{figures/logo_IUF}\vspace*{2truemm}
-\includegraphics[width=0.135\textwidth]{figures/logo_Caltech}\vspace*{2truemm}
-\includegraphics[width=0.135\textwidth]{figures/logo_Harvard}\vspace*{2truemm}
-\includegraphics[width=0.135\textwidth]{figures/logo_IPGP}\vspace*{2truemm}
-\includegraphics[width=0.112\textwidth]{figures/logo_ANR}\vspace*{2truemm}
-\includegraphics[width=0.112\textwidth]{figures/logo_NSF}\vspace*{2truemm}
-\par\end{centering}
-%
-\vspace*{-2truemm}
-%
-\noindent \begin{centering}
-\includegraphics[width=0.190\textwidth]{figures/logo_European_Union}\vspace*{2truemm}
-\includegraphics[width=0.112\textwidth]{figures/logo_GENCI}\vspace*{2truemm}
-\includegraphics[width=0.112\textwidth]{figures/logo_PRACE}\vspace*{2truemm}
-\includegraphics[width=0.112\textwidth]{figures/logo_CINES}\vspace*{2truemm}
-\includegraphics[width=0.130\textwidth]{figures/logo_Oak_Ridge}\vspace*{2truemm}
-\hspace*{3mm}\includegraphics[width=0.130\textwidth]{figures/logo_fondation_Del_Duca}
-\includegraphics[width=0.130\textwidth]{figures/logo_CIG}\vspace*{2truemm}
-\par\end{centering}
-%
-\vspace*{-2truemm}
-%
-\noindent \begin{centering}
-\includegraphics[width=0.112\textwidth]{figures/logo_University_of_Toronto}\vspace*{2truemm}
-\includegraphics[width=0.109\textwidth]{figures/logo_INGV}\vspace*{2truemm}
-\includegraphics[width=0.102\textwidth]{figures/logo_Univ_Toulouse}\vspace*{2truemm}
-\includegraphics[width=0.112\textwidth]{figures/logo_TOTAL}\vspace*{2truemm}
-\includegraphics[width=0.130\textwidth]{figures/logo_Fairbanks}\vspace*{2truemm}
-\includegraphics[width=0.120\textwidth]{figures/logo_CINECA}\vspace*{2truemm}
-\includegraphics[width=0.140\textwidth]{figures/logo_Intel_Exascale_Labs}\vspace*{2truemm}
-\includegraphics[width=0.130\textwidth]{figures/logo_Maison_Simulation}
-\par\end{centering}
-\end{figure}
-
-\newpage{}
-
-\tableofcontents{}
-
-\chapter{Introduction}
-
-The software package SPECFEM3D Cartesian simulates seismic wave propagation
-at the local or regional scale and performs full waveform imaging (FWI) or adjoint tomography based upon the spectral-element method
-(SEM). The SEM is a continuous Galerkin technique \citep{TrKoLi08,PeKoLuMaLeCaLeMaLiBlNiBaTr11},
-which can easily be made discontinuous \citep{BeMaPa94,Ch00,KoWoHu02,ChCaVi03,LaWaBe05,Kop06,WiStBuGh10,AcKo11};
-it is then close to a particular case of the discontinuous Galerkin
-technique \citep{ReHi73,LeRa74,Arn82,JoPi86,BoMaHe91,FaRi99,HuHuRa99,CoKaSh00,GiHeWa02,RiWh03,MoRi05,GrScSc06,AiMoMu06,BeLaPi06,DuKa06,DeSeWh08,PuAmKa09,WiStBuGh10,DeSe10,EtChViGl10},
-with optimized efficiency because of its tensorized basis functions
-\citep{WiStBuGh10,AcKo11}. In particular, it can accurately handle
-very distorted mesh elements \citep{OlSe11}.\\
-
-It has very good accuracy and convergence properties \citep{MaPa89,SePr94,DeFiMu02,Coh02,DeSe07,SeOl08,AiWa09,AiWa10,MeStTh12}.
-The spectral element approach admits spectral rates of convergence
-and allows exploiting $hp$-convergence schemes. It is also very well
-suited to parallel implementation on very large supercomputers \citep{KoTsChTr03,TsKoChTr03,KoLaMi08a,CaKoLaTiMiLeSnTr08,KoViCh10}
-as well as on clusters of GPU accelerating graphics cards \citep{Kom11,MiKo10,KoMiEr09,KoErGoMi10}.
-Tensor products inside each element can be optimized to reach very
-high efficiency \citep{DeFiMu02}, and mesh point and element numbering
-can be optimized to reduce processor cache misses and improve cache
-reuse \citep{KoLaMi08a}. The SEM can also handle triangular (in 2D)
-or tetrahedral (in 3D) elements \citep{WinBoyd96,TaWi00,KoMaTrTaWi01,Coh02,MeViSa06}
-as well as mixed meshes, although with increased cost and reduced
-accuracy in these elements, as in the discontinuous Galerkin method.\\
-
-Note that in many geological models in the context of seismic wave
-propagation studies (except for instance for fault dynamic rupture
-studies, in which very high frequencies or supershear rupture need
-to be modeled near the fault, see e.g. \citet{BeGlCrViPi07,BeGlCrVi09,PuAmKa09,TaCrEtViBeSa10})
-a continuous formulation is sufficient because material property contrasts
-are not drastic and thus conforming mesh doubling bricks can efficiently
-handle mesh size variations \citep{KoTr02a,KoLiTrSuStSh04,LeChLiKoHuTr08,LeChKoHuTr09,LeKoHuTr09}.\\
-
-For a detailed introduction to the SEM as applied to regional seismic
-wave propagation, please consult \citet{PeKoLuMaLeCaLeMaLiBlNiBaTr11,TrKoLi08,KoVi98,KoTr99,ChKoViCaVaFe07}
-and in particular \citet{LeKoHuTr09,LeChKoHuTr09,LeChLiKoHuTr08,GoAmTaCaSmSaMaKo09,WiKoScTr04,KoLiTrSuStSh04}.
-A detailed theoretical analysis of the dispersion
-and stability properties of the SEM is available in \citet{Coh02}, \citet{DeSe07}, \citet{SeOl07}, \citet{SeOl08} and \citet{MeStTh12}.\\
-
-Effects due to lateral variations in compressional-wave speed, shear-wave
-speed, density, a 3D crustal model, topography and bathymetry are
-included. The package can accommodate full 21-parameter anisotropy
-(see~\citet{ChTr07}) as well as lateral variations in attenuation
-\citep{SaKoTr10}. Adjoint capabilities and finite-frequency kernel
-simulations are included \citep{TrKoLi08,PeKoLuMaLeCaLeMaLiBlNiBaTr11,LiTr06,FiIgBuKe09,ViOp09}.\\
-
-The SEM was originally developed in computational fluid dynamics \citep{Pat84,MaPa89}
-and has been successfully adapted to address problems in seismic wave
-propagation. Early seismic wave propagation applications of the SEM,
-utilizing Legendre basis functions and a perfectly diagonal mass matrix,
-include \citet{CoJoTo93}, \citet{Kom97}, \citet{FaMaPaQu97}, \citet{CaGa97},
-\citet{KoVi98} and \citet{KoTr99}, whereas applications involving
-Chebyshev basis functions and a nondiagonal mass matrix include \citet{SePr94},
-\citet{PrCaSe94} and \citet{SePrPr95}.\\
-
-All SPECFEM3D software is written in Fortran2003 with full portability
-in mind, and conforms strictly to the Fortran2003 standard. It uses
-no obsolete or obsolescent features of Fortran. The package uses parallel
-programming based upon the Message Passing Interface (MPI) \citep{GrLuSk94,Pac97}.\\
-
-SPECFEM3D won the Gordon Bell award for best performance at the SuperComputing~2003
-conference in Phoenix, Arizona (USA) (see \citet{KoTsChTr03} and
-\url{www.sc-conference.org/sc2003/nrfinalaward.html}). It was a finalist
-again in 2008 for a run at 0.16 petaflops (sustained) on 149,784 processors
-of the `Jaguar' Cray XT5 system at Oak Ridge National Laboratories
-(USA) \citep{CaKoLaTiMiLeSnTr08}. It also won the BULL Joseph Fourier
-supercomputing award in 2010.\\
-
-It reached the sustained one petaflop performance level for the first time in February 2013
-on the Blue Waters Cray supercomputer at the National Center for Supercomputing Applications (NCSA), located at the University of Illinois at Urbana-Champaign (USA).\\
-
-This new release of the code includes Convolution or
-Auxiliary Differential Equation Perfectly Matched absorbing Layers
-(C-PML or ADE-PML) \citep{MaKoEz08,MaKoGe08,MaKo09,MaKoGeBr10,KoMa07}.
-It also includes support for GPU graphics card acceleration \citep{Kom11,MiKo10,KoMiEr09,KoErGoMi10}.\\
-
-The next release of the code will use the PT-SCOTCH parallel and threaded
-version of SCOTCH for mesh partitioning instead of the serial version.\\
-
-SPECFEM3D Cartesian includes coupled fluid-solid domains and adjoint
-capabilities, which enables one to address seismological inverse problems,
-but for linear rheologies only so far. To accommodate visco-plastic
-or non-linear rheologies, the reader can refer to the \href{http://geoelse.stru.polimi.it/}{GeoELSE}software
-package \citep{CaGa97,StPaIg09}.
-
-\section{Citation}
-
-If you use SPECFEM3D Cartesian for your own research, please cite
-at least one of the following articles:
-\begin{description}
-\item [{Numerical simulations in general}] ~\\
- Forward and adjoint simulations are described in detail in \citet{TrKoLi08,PeKoLuMaLeCaLeMaLiBlNiBaTr11,VaCaSaKoVi99,KoMiEr09,KoErGoMi10,ChKoViCaVaFe07,MaKoDi09,KoViCh10,CaKoLaTiMiLeSnTr08,TrKoHjLiZhPeBoMcFrTrHu10,KoRiTr02,KoTr02a,KoTr02b,KoTr99}
-or \citet{KoVi98}. Additional aspects of adjoint simulations are
-described in \citet{TrTaLi05,LiTr06,TrKoLi08,LiTr08,TrKoHjLiZhPeBoMcFrTrHu10,PeKoLuMaLeCaLeMaLiBlNiBaTr11}.
-Domain decomposition is explained in detail in \citet{MaKoBlLe08},
-and excellent scaling up to 150,000 processor cores is shown for instance
-in \citet{CaKoLaTiMiLeSnTr08,KoLaMi08a,MaKoBlLe08,KoErGoMi10,Kom11},
-\item [{GPU computing}] ~\\
- Computing on GPU graphics cards for acoustic or seismic wave propagation
-applications is described in detail in \citet{Kom11,MiKo10,KoMiEr09,KoErGoMi10}.
-\end{description}
-\noindent If you use this new version, which has non blocking
-MPI for much better performance for medium or large runs, please cite
-at least one of these six articles, in which results of non blocking
-MPI runs are presented: \citet{PeKoLuMaLeCaLeMaLiBlNiBaTr11,KoErGoMi10,KoViCh10,Kom11,CaKoLaTiMiLeSnTr08,MaKoBlLe08}.\\
-
-\noindent If you work on geophysical applications, you may be interested
-in citing some of these application articles as well, among others:
-\begin{description}
-\item [{Southern California simulations}] ~\\
- \citet{KoLiTrSuStSh04,KrJiKoTr06a,KrJiKoTr06b}.
-
-If you use the 3D southern California model, please cite \citet{SuSh03}
-(Los Angeles model), \citet{lovelyetal06} (Salton Trough), and \citet{hauksson2000}
-(southern California). The Moho map was determined by \citet{zhu&kanamori2000}.
-The 1D SoCal model was developed by \citet{DrHe90}.
-
-\item [{Anisotropy}] ~\\
- \citet{ChTr07,JiTsKoTr05,ChFaKo04,FaChKo04,RiRiKoTrHe02,TrKo00}.
-\item [{Attenuation}] ~\\
- \citet{SaKoTr10,KoTr02a,KoTr99}.
-\item [{Topography}] ~\\
- \citet{LeKoHuTr09,LeChKoHuTr09,LeChLiKoHuTr08,GoAmTaCaSmSaMaKo09,WiKoScTr04}.
-\end{description}
-The corresponding Bib\TeX{} entries may be found in file \texttt{doc/USER\_MANUAL/bibliography.bib}.
-
-\section{Support}
-
-This material is based upon work supported by the USA National Science
-Foundation under Grants No. EAR-0406751 and EAR-0711177, by the French
-CNRS, French INRIA Sud-Ouest MAGIQUE-3D, French ANR NUMASIS under
-Grant No. ANR-05-CIGC-002, and European FP6 Marie Curie International
-Reintegration Grant No. MIRG-CT-2005-017461. Any opinions, findings,
-and conclusions or recommendations expressed in this material are
-those of the authors and do not necessarily reflect the views of the
-USA National Science Foundation, CNRS, INRIA, ANR or the European
-Marie Curie program.
-
-\chapter{\label{cha:Getting-Started}Getting Started}
-
-To download the SPECFEM3D\_Cartesian software package, type this:
-\begin{lyxcode}
-git clone --recursive --branch devel https://github.com/geodynamics/specfem3d.git
-\end{lyxcode}
-{\small \par}
-
-We recommend that you add {\texttt{ulimit -S -s unlimited}} to your
-{\texttt{.bash\_profile}} file and/or {\texttt{limit stacksize
-unlimited }} to your {\texttt{.cshrc}} file to suppress any potential
-limit to the size of the Unix stack.\\
-
-Then, to configure the software for your system, run the \texttt{configure}
-shell script. This script will attempt to guess the appropriate configuration
-values for your system. However, at a minimum, it is recommended that
-you explicitly specify the appropriate command names for your Fortran
-compiler and MPI package (another option is to define FC, CC and MPIF90
-in your .bash\_profile or your .cshrc file):
-\begin{lyxcode}
-./configure~FC=ifort~MPIFC=mpif90
-\end{lyxcode}
-Note that MPI must be installed with MPI-IO enabled because parts of SPECFEM3D perform I/Os through MPI-IO.
-
-Before running the \texttt{configure} script, you should probably
-edit file \texttt{flags.guess} to make sure that it contains the best
-compiler options for your system. Known issues or things to check
-are:
-\begin{description}
-\item [{\texttt{Intel ifort compiler}}] See if you need to add \texttt{-assume byterecl} for your machine.
-\textbf{In the case of that compiler, we have noticed that versions dot zero sometimes have bugs or issues that can lead to wrong results when running the code, thus we \emph{strongly} recommend using versions dot one or above (for instance version 13.1 instead of 13.0, version 14.1 instead of 14.0 and so on)}.
-\item [{\texttt{IBM compiler}}] See if you need to add \texttt{-qsave}
-or \texttt{-qnosave} for your machine.
-\item [{\texttt{Mac OS}}] You will probably need to install \texttt{XCODE}.
-\end{description}
-When compiling on an IBM machine with the \texttt{xlf} and \texttt{xlc}
-compilers, we suggest running the \texttt{configure} script with the
-following options:\\
-
-\noindent \texttt{./configure FC=xlf90\_r MPIFC=mpif90 CC=xlc\_r CFLAGS=\textquotedbl{}-O3
--q64\textquotedbl{} FCFLAGS=\textquotedbl{}-O3 -q64\textquotedbl{}
--with-scotch-dir=...}.\\
-
-On SGI systems, \texttt{flags.guess} automatically informs \texttt{configure}
-to insert `\texttt{`TRAP\_FPE=OFF}'' into the generated \texttt{Makefile}
-in order to turn underflow trapping off.\\
-
-You can add \texttt{--enable-vectorization} to the configuration options to speed up the code in the fluid (acoustic) and elastic parts.
-This works fine if (and only if) your computer always allocates a contiguous memory block for each allocatable array;
-this is the case for most machines and most compilers, but not all. To disable this feature, use option \texttt{--disable-vectorization}.
-For more details see \href{https://github.com/geodynamics/specfem3d/issues/81}{github.com/geodynamics/specfem3d/issues/81} .
-To check if that option works fine on your machine, run the code with and without it for an acoustic/elastic model and make sure the seismograms are identical.\\
-
-Note that we use CUBIT (now called Trelis) to create meshes of hexahedra, but other packages
-can be used as well, for instance GiD from \url{http://gid.cimne.upc.es}
-or Gmsh from \url{http://geuz.org/gmsh} \citep{GeRe09}. Even mesh
-creation packages that generate tetrahedra, for instance TetGen from
-\url{http://tetgen.berlios.de}, can be used because each tetrahedron
-can then easily be decomposed into four hexahedra as shown in the
-picture of the TetGen logo at \url{http://tetgen.berlios.de/figs/Delaunay-Voronoi-3D.gif};
-while this approach does not generate hexahedra of optimal quality,
-it can ease mesh creation in some situations and it has been shown
-that the spectral-element method can very accurately handle distorted
-mesh elements \citep{OlSe11}.
-
-The SPECFEM3D Cartesian software package relies on the SCOTCH library
-to partition meshes created with CUBIT. METIS \citep{KaKu98a,KaKu98c,KaKu98b}
-can also be used instead of SCOTCH if you prefer, by editing file
-src/decompose\_mesh/decompose\_mesh.F90 and uncommenting the flag
-USE\_METIS\_INSTEAD\_OF\_SCOTCH. You will also then need to install
-and compile Metis version 4.0 (do {*}NOT{*} install Metis version
-5.0, which has incompatible function calls) and edit \texttt{src/decompose\_mesh/Makefile.in}
-and uncomment the METIS link flag in that file before running \texttt{configure}.\\
-
-The SCOTCH library \citep{PeRo96} provides efficent static mapping,
-graph and mesh partitioning routines. SCOTCH is a free software package
-developed by Fran\c{c}ois Pellegrini et al. from LaBRI and INRIA in Bordeaux,
-France, downloadable from the web page \url{https://gforge.inria.fr/projects/scotch/}.
-In case no SCOTCH libraries can be found on the system, the configuration
-will bundle the version provided with the source code for compilation.
-The path to an existing SCOTCH installation can to be set explicitly
-with the option \texttt{-{}-with-scotch-dir}. Just as an example:
-\begin{lyxcode}
-./configure~FC=ifort~MPIFC=mpif90~-{}-with-scotch-dir=/opt/scotch
-\end{lyxcode}
-If you use the Intel ifort compiler to compile the code, we recommend
-that you use the Intel icc C compiler to compile Scotch, i.e., use:
-\begin{lyxcode}
-./configure~CC=icc~FC=ifort~MPIFC=mpif90
-\end{lyxcode}
-
-When compiling the SCOTCH source code, if you get a message such as: "ld: cannot find -lz",
-the Zlib compression development library is probably missing on your machine and you will need to install it or ask your system administrator to
-do so. On Linux machines the package is often called "zlib1g-dev" or similar. (thus "sudo apt-get install zlib1g-dev" would install it)
-
-To compile a serial version of the code for small meshes that fits
-on one compute node and can therefore be run serially, run \texttt{configure}
-with the \texttt{-{}-without-mpi} option to suppress all calls to
-MPI.
-
-A summary of the most important configuration variables follows.
-\begin{description}
-\item [{\texttt{F90}}] Path to the Fortran compiler.
-\item [{\texttt{MPIF90}}] Path to MPI Fortran.
-\item [{\texttt{MPI\_FLAGS}}] Some systems require this flag to link to
-MPI libraries.
-\item [{\texttt{FLAGS\_CHECK}}] Compiler flags.
-\end{description}
-The configuration script automatically creates for each executable
-a corresponding \texttt{Makefile} in the \texttt{src/} subdirectoy.
-The \texttt{Makefile} contains a number of suggested entries for various
-compilers, e.g., Portland, Intel, Absoft, NAG, and Lahey. The software
-has run on a wide variety of compute platforms, e.g., various PC clusters
-and machines from Sun, SGI, IBM, Compaq, and NEC. Select the compiler
-you wish to use on your system and choose the related optimization
-flags. Note that the default flags in the \texttt{Makefile} are undoubtedly
-not optimal for your system, so we encourage you to experiment with
-these flags and to solicit advice from your systems administrator.
-Selecting the right compiler and optimization flags can make a tremendous
-difference in terms of performance. We welcome feedback on your experience
-with various compilers and flags.
-
-Now that you have set the compiler information, you need to select
-a number of flags in the \texttt{constants.h} file depending on your
-system:
-\begin{description}
-\item [{\texttt{LOCAL\_PATH\_IS\_ALSO\_GLOBAL}}] Set to \texttt{.false.}
-on most cluster applications. For reasons of speed, the (parallel)
-distributed database generator typically writes a (parallel) database
-for the solver on the local disks of the compute nodes. Some systems
-have no local disks, e.g., BlueGene or the Earth Simulator, and other
-systems have a fast parallel file system, in which case this flag
-should be set to \texttt{.true.}. Note that this flag is not used
-by the database generator or the solver; it is only used for some
-of the post-processing.
-\end{description}
-The package can run either in single or in double precision mode.
-The default is single precision because for almost all calculations
-performed using the spectral-element method using single precision
-is sufficient and gives the same results (i.e. the same seismograms);
-and the single precision code is faster and requires exactly half
-as much memory. Select your preference by selecting the appropriate
-setting in the \texttt{constants.h} file:
-\begin{description}
-\item [{\texttt{CUSTOM\_REAL}}] Set to \texttt{SIZE\_REAL} for single precision
-and \texttt{SIZE\_DOUBLE} for double precision.
-\end{description}
-In the \texttt{precision.h} file:
-\begin{description}
-\item [{\texttt{CUSTOM\_MPI\_TYPE}}] Set to \texttt{MPI\_REAL} for single
-precision and \texttt{MPI\_DOUBLE\_PRECISION} for double precision.
-\end{description}
-On many current processors (e.g., Intel, AMD, IBM Power), single precision
-calculations are significantly faster; the difference can typically
-be 10\% to 25\%. It is therefore better to use single precision. What
-you can do once for the physical problem you want to study is run
-the same calculation in single precision and in double precision on
-your system and compare the seismograms. If they are identical (and
-in most cases they will), you can select single precision for your
-future runs.\\
-
-If your compiler has problems with the \texttt{use mpi} statements that are used in the code, use the script called
-\texttt{replace\_use\_mpi\_with\_include\_mpif\_dot\_h.pl} in the root directory to replace all of them with \texttt{include 'mpif.h'} automatically.
-
-\section{Adding OpenMP support in addition to MPI}
-
-NOTE FROM JULY 2013: OpenMP support is maybe / probably not maintained any more.
-Thus the section below is maybe obsolete.
-
-OpenMP support can be enabled in addition to MPI. However, in many
-cases performance will not improve because our pure MPI implementation
-is already heavily optimized and thus the resulting code will in fact
-be slightly slower. A possible exception could be IBM BlueGene-type
-architectures.
-
-To enable OpenMP, uncomment the OpenMP compiler option in two lines
-in file \texttt{src/specfem3D/Makefile.in} (before running \texttt{configure})
-and also uncomment the \texttt{\#define USE\_OPENMP} statement in
-file\\
- \texttt{src/specfem3D/specfem3D.F90}.
-
-The DO-loop using OpenMP threads has a SCHEDULE property. The OMP\_SCHEDULE
-environment variable can set the scheduling policy of that DO-loop.
-Tests performed by Marcin Zielinski at SARA (The Netherlands) showed
-that often the best scheduling policy is DYNAMIC with the size of
-the chunk equal to the number of OpenMP threads, but most preferably
-being twice as the number of OpenMP threads (thus chunk size = 8 for
-4 OpenMP threads etc). If OMP\_SCHEDULE is not set or is empty, the
-DO-loop will assume generic scheduling policy, which will slow down
-the job quite a bit.
-
-\section{Compiling on an IBM BlueGene}
-
-\underline{More recent installation instruction for IBM BlueGene, from April 2013:}\\
-
-\noindent Edit file \texttt{flags.guess} and put this for \texttt{FLAGS\_CHECK}:\\
-
-\noindent \texttt{-g -qfullpath -O2 -qsave -qstrict -qtune=qp -qarch=qp
--qcache=auto -qhalt=w}\\
- \texttt{-qfree=f90 -qsuffix=f=f90 -qlanglvl=95pure -Q -Q+rank,swap\_all
--Wl,-relax}\\
-
-\noindent The most relevant are the -qarch and -qtune flags, otherwise
-if these flags are set to ``auto'' then they are wrongly assigned
-to the architecture of the frond-end node, which is different from
-that on the compute nodes. You will need to set these flags to the
-right architecture for your BlueGene compute nodes, which is not necessarily
-``qp''; ask your system administrator. On some machines if is necessary
-to use -O2 in these flags instead of -O3 due to a compiler bug of
-the XLF version installed. We thus suggest to first try -O3, and then
-if the code does not compile or does not run fine then switch back
-to -O2. The debug flags (-g, -qfullpath) do not influence performance
-but are useful to get at least some insights in case of problems.\\
-
-\noindent Before running \texttt{configure}, select the XL Fortran
-compiler by typing \texttt{module load bgq-xl/1.0} or \texttt{module
-load bgq-xl} (another, less efficient option is to load the GNU compilers
-using \texttt{module load bgq-gnu/4.4.6} or similar).\\
-
-\noindent Then, to configure the code, type this:\\
-
-\noindent \texttt{./configure FC=bgxlf90\_r MPIFC=mpixlf90\_r CC=bgxlc\_r
-LOCAL\_PATH\_IS\_ALSO\_GLOBAL=true}\\
-
-In order for the SCOTCH domain decomposer to compile, on some (but
-not all) Blue Gene systems you may need to run \texttt{configure}
-with \texttt{CC=gcc} instead of \texttt{CC=bgxlc\_r}.
-
-\noindent \underline{Older installation instruction for IBM BlueGene, from 2011:}\\
-
-\noindent To compile the code on an IBM BlueGene, Laurent L\'eger from
-IDRIS, France, suggests the following: compile the code with
-
-\noindent \texttt{ FLAGS\_CHECK=\textquotedbl{}-O3 -qsave -qstrict
--qtune=auto -qarch=450d -qcache=auto }~\\
- \texttt{ -qfree=f90 -qsuffix=f=f90 -g -qlanglvl=95pure -qhalt=w -Q
--Q+rank,swap\_all -Wl,-relax\textquotedbl{}}
-
-\noindent Option \textquotedbl{}-Wl,-relax\textquotedbl{} must be
-added on many (but not all) BlueGene systems to be able to link the
-binaries \texttt{xmeshfem3D} and \texttt{xspecfem3D} because the final
-link step is done by the GNU \texttt{ld} linker even if one uses \texttt{FC=bgxlf90\_r,
-MPIFC=mpixlf90\_r} and \texttt{CC=bgxlc\_r} to create all the object
-files. On the contrary, on some BlueGene systems that use the native
-AIX linker option \textquotedbl{}-Wl,-relax\textquotedbl{} can lead
-to problems and must be suppressed from \texttt{flags.guess}.
-
-\noindent Also, \texttt{AR=ar, ARFLAGS=cru} and \texttt{RANLIB=ranlib}
-are hardwired in all \texttt{Makefile.in} files by default, but to
-cross-compile on BlueGene/P one needs to change these values to \texttt{AR=bgar,
-ARFLAGS=cru} and \texttt{RANLIB=bgranlib}. Thus the easiest thing
-to do is to modify all \texttt{Makefile.in} files and the \texttt{configure}
-script to set them automatically by \texttt{configure}. One then just
-needs to pass the right commands to the \texttt{configure} script:
-
-\noindent \texttt{ ./configure -{}-prefix=/path/to/SPECFEM3DG\_SP
--{}-host=Babel -{}-build=BGP }~\\
- \texttt{ FC=bgxlf90\_r MPIFC=mpixlf90\_r CC=bgxlc\_r AR=bgar ARFLAGS=cru
-}~\\
- \texttt{ RANLIB=bgranlib LOCAL\_PATH\_IS\_ALSO\_GLOBAL=false}
-
-\noindent This trick can be useful for all hosts on which one needs
-to cross-compile.
-
-\noindent On BlueGene, one also needs to run the \texttt{xcreate\_header\_file}
-binary file manually rather than in the Makefile:
-
-\noindent \texttt{bgrun -np 1 -mode VN -exe ./bin/xcreate\_header\_file}
-
-\section{Visualizing the subroutine calling tree of the source code}
-
-\noindent
-Packages such as \texttt{Doxywizard} can be used to visualize the calling tree of the
-subroutines of the  source code. \texttt{Doxywizard} is
-a GUI front-end for configuring and running \texttt{Doxygen}.
-
-\bigskip
-
-\noindent To do your own call graphs, you can follow these simple steps below.
-
-\begin{enumerate}
-\setcounter{enumi}{-1}
-\item Install \texttt{Doxygen} \underline{and} \texttt{graphviz} (the two are usually in the package manager of classic Linux distribution).
-\item Run in the terminal : \texttt{doxygen -g}, which creates a \texttt{Doxyfile} that tells doxygen what you want it to do.
-\item Edit the Doxyfile. Two Doxyfile-type files have been already committed in the directory \\
-\noindent \texttt{specfem3d/doc/Call\_trees}:
-
-\begin{itemize}
-\item[\textbullet] \texttt{Doxyfile\_truncated\_call\_tree} will generate call graphs with maximum 3 or 4 levels of tree structure,
-\item[\textbullet] \texttt{Doxyfile\_complete\_call\_tree} will generate call graphs with complete tree structure.
-\end{itemize}
-
-The important entries in the Doxyfile are:
-
-\begin{description}
-\item [{\texttt{PROJECT\_NAME}}]
-\item [{\texttt{OPTIMIZE\_FOR\_FORTRAN}}] Set to YES
-\item [{\texttt{EXTRACT\_ALL}}] Set to YES
-\item [{\texttt{EXTRACT\_PRIVATE}}] Set to YES
-\item [{\texttt{EXTRACT\_STATIC}}] Set to YES
-\item [{\texttt{INPUT}}] From the directory \texttt{specfem3d/doc/Call\_trees}, it is \texttt{"../../src/"}
-\item [{\texttt{FILE\_PATTERNS}}] In SPECFEM case, it is \texttt{*.f90* *.F90* *.c* *.cu* *.h*}
-\item [{\texttt{HAVE\_DOT}}] Set to YES
-\item [{\texttt{CALL\_GRAPH}}] Set to YES
-\item [{\texttt{CALLER\_GRAPH}}] Set to YES
-\item [{\texttt{DOT\_PATH}}] The path where is located the dot program graphviz (if it is not in your \$PATH)
-\item [{\texttt{RECURSIVE}}] This tag can be used to turn specify whether or not subdirectories should be searched for input files as well. In the case of SPECFEM, set to YES.
-\item [{\texttt{EXCLUDE}}] Here, you can exclude:  \\
-\texttt{"../../src/specfem3D/older\_not\_maintained\_partial\_OpenMP\_port"}\\
-\texttt{"../../src/decompose\_mesh/scotch"}\\
-\texttt{"../../src/decompose\_mesh/scotch\_5.1.12b"}
-\item [{\texttt{DOT\_GRAPH\_MAX\_NODES}}] to set the maximum number of nodes that will be shown in the graph.
-If the number of nodes in a graph becomes larger than this value, doxygen will truncate the graph, which is visualized by representing a node as a red box.
-Minimum value: 0, maximum value: 10000, default value: 50.
-\item [{\texttt{MAX\_DOT\_GRAPH\_DEPTH}}] to set the maximum depth of the graphs generated by dot.
-A depth value of 3 means that only nodes reachable from the root by following a path via at most 3 edges will be shown.
-Using a depth of 0 means no depth restriction.
-Minimum value: 0, maximum value: 1000, default value: 0.
-
-\end{description}
-
-\item Run : \texttt{doxygen Doxyfile}, HTML and LaTeX files created by default in \texttt{html} and \texttt{latex} subdirectories.
-
-\item To see the call trees, you have to open the file \texttt{html/index.html} in your \underline{browser}.
-You will have many informations about each subroutines of SPECFEM (not only call graphs), you can click on every boxes / subroutines.
-It show you the call, and, the caller graph of each subroutine : the subroutines called by the concerned subroutine, and the previous subroutines who call this subroutine (the previous path), respectively.
-In the case of a truncated calling tree, the boxes with a red border indicates a node that has \underline{more} arrows than are shown (in other words: the graph is truncated with respect to this node).
-
-\end{enumerate}
-
-\medskip
-
-\noindent Finally, some useful links:
-
-\begin{itemize}
-\item[\textbullet] a good and short summary for the basic utilisation of Doxygen:
-
-\url{http://www.softeng.rl.ac.uk/blog/2010/jan/30/callgraph-fortran-doxygen/},
-
-\item[\textbullet] to configure the diagrams :
-
-\url{http://www.stack.nl/~dimitri/doxygen/manual/diagrams.html},
-
-\item[\textbullet] the complete alphabetical index of the tags in Doxyfile:
-
-\url{http://www.stack.nl/~dimitri/doxygen/manual/config.html},
-
-\item[\textbullet] more generally, the Doxygen manual:
-
-\url{http://www.stack.nl/~dimitri/doxygen/manual/index.html}.
-
-\end{itemize}
-
-\medskip
-
-\section{Using the ADIOS library for I/O}
-Regular POSIX I/O can be problematic when dealing with large simulations one large
-clusters (typically more than $10,000$ processes). SPECFEM3D use the ADIOS library~\cite{Liu2013}
-to deal transparently take advantage of advanced paralllel file system features. To enable
-ADIOS, the following steps should be done:
-\begin{enumerate}
-\item Install ADIOS (available from \url{https://www.olcf.ornl.gov/center-projects/adios/}).
-Make sure that your environment variables reference it.
-\item You may want to change ADIOS related values in the \texttt{constants.h.in} file.
-The default values probably suit most cases.
-\item Configure using the \texttt{--with-adios} flag.
-\end{enumerate}
-ADIOS is currently only usable for meshfem3D generated mesh (i.e. not for meshes generated
-with CUBTI). Additional control parameters are discussed in section~\ref{cha:Main-Parameter}.
-
-\section{Becoming a developer of the code, or making small modifications in the source code}
-
-If you want to develop new features in the code, and/or if you want to make small changes, improvements, or bug fixes, you are very welcome to contribute.
-To do so, i.e.
-to access the development branch of the source code with read/write access (in a safe way, no need to worry too much about breaking the package, there is a robot called BuildBot that is in charge of checking and validating all new contributions and changes), please visit this Web page: \url{https://github.com/geodynamics/specfem3d/wiki/Using-Hub}.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\chapter{\label{cha:Mesh-Generation}Mesh Generation}
-
-The first step in running a spectral-element simulation consists of
-constructing a high-quality mesh for the region under consideration.
-We provide two possibilities to do so: (1) relying on the external,
-hexahedral mesher CUBIT, or (2) using the provided, internal mesher
-\texttt{xmeshfem3D}. In the following, we explain these two approaches.
-
-\section{\label{cha:Running-the-Mesher-CUBIT}Meshing with \texttt{CUBIT}}
-
-CUBIT is a meshing tool suite for the creation of finite-element meshes
-for arbitrarily shaped models. It has been developed and maintained
-at Sandia National Laboratories and can be purchased for a small academic
-institutional fee at \url{http://cubit.sandia.gov}. Our experience
-showed that using CUBIT greatly facilitates and speeds up the generation
-and preparation of hexahedral, conforming meshes for a variety of
-geophysical models with increasing complexity.
-
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=0.6\textwidth]{figures/mount-cubit}
-\par\end{centering}
-
-\caption{Example of the graphical user interface of CUBIT. The hexahedral mesh
-shown in the main display consists of a hexahedral discretization
-of a single volume with topography.}
-
-\label{fig:mount.cubit}
-\end{figure}
-
-\noindent
-The basic steps in creating a load-balanced, partitioned mesh with
-CUBIT are:
-\begin{description}
-\item [{1.}] setting up a hexahedral mesh with CUBIT,
-\item [{2.}] exporting the CUBIT mesh into a SPECFEM3D Cartesian file format
-and
-\item [{3.}] partitioning the SPECFEM3D Cartesian mesh files for a chosen
-number of cores.
-\end{description}
-Examples are provided in the SPECFEM3D Cartesian package in the subdirectory
-\texttt{EXAMPLES/}. We strongly encourage you to contribute your own
-example to this package by contacting the CIG Computational Seismology
-Mailing List \urlwithparentheses{cig-seismo@geodynamics.org}.
-
-\subsection{Creating the Mesh with CUBIT}
-
-For the installation and handling of the CUBIT meshing tool suite,
-please refer to the CUBIT user manual and documentation. In order
-to give you a basic understanding of how to use CUBIT for our purposes,
-examples are provided in the SPECFEM3D Cartesian package in the subdirectory
-\texttt{EXAMPLES/}:
-\begin{description}
-\item [{\texttt{homogeneous\_halfspace}}] Creates a single block model
-and assigns elastic material parameters.
-\item [{\texttt{layered\_halfspace}}] Combines two different, elastic material
-volumes and creates a refinement layer between the two. This example
-can be compared for validation against the solutions provided in subdirectory
-~\\
- \texttt{VALIDATION\_3D\_SEM\_SIMPLER\_LAYER\_SOURCE\_DEPTH/}.
-\item [{\texttt{waterlayered\_halfspace}}] Combines an acoustic and elastic
-material volume as in a schematic marine survey example.
-\item [{\texttt{tomographic\_model}}] Creates a single block model whose
-material properties will have to be read in from a tomographic model
-file during the databases creation by \texttt{xgenerate\_databases}.
-\end{description}
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=0.45\textwidth]{figures/example-homogeneous}
-\includegraphics[width=0.45\textwidth]{figures/example-2layers} \\
- \includegraphics[width=0.45\textwidth]{figures/example-water} \includegraphics[width=0.45\textwidth]{figures/example-tomo}
-\par\end{centering}
-
-\caption{Screenshots of the CUBIT examples provided in subdirectory \texttt{EXAMPLES/}:
-homogeneous halfspace (top-left), layered halfspace (top-right), water
-layered halfspace (bottom-left) and tomographic model (bottom-right).}
-
-\label{fig:examples.cubit}
-\end{figure}
-
-In each example subdirectory you will find a \texttt{README} file,
-which explains in a step-by-step tutorial the workflow for the example.
-Please feel free to contribute your own example to this package by
-contacting the CIG Computational Seismology Mailing List \urlwithparentheses{cig-seismo@geodynamics.org}.
-
-In some cases, to re-create the meshes for the examples given, just type "claro ./create\_mesh.py" or similar
-from the command line ("claro" is the command to run CUBIT from the command line).
-
-IMPORTANT note: in order to correctly set up GEOCUBIT and run the examples, please read the file called "EXAMPLES/README";
-in particular, please make sure you correctly set up the Python paths as indicated in that file.
-
-\subsection{Exporting the Mesh with \texttt{run\_cubit2specfem3d.py} }
-
-\label{subsec:Exporting-the-Mesh}
-
-Once the geometric model volumes in CUBIT are meshed, you prepare
-the model for exportation with the definition of material blocks and
-boundary surfaces. Thus, prior to exporting the mesh, you need to
-define blocks specifying the materials and absorbing boundaries in
-CUBIT. This process could be done automatically using the script \texttt{run\_cubit2specfem3d.py}
-if the mesh meets some conditions or manually, following the block
-convention:
-\begin{description}
-\item [{material\_name}] Each material should have a specific block defined
-by a unique name. The name convention of the material is to start
-with either 'elastic' or 'acoustic'. It must be then followed by a
-unique identifier, e.g. 'elastic 1', 'elastic 2', etc. The additional
-attributes to the block define the material description.
-
-For an elastic material:
-\begin{description}
-\item [{material\_id}] An integer value which is unique for this material.
-\item [{Vp}] P-wave speed of the material (given in m/s).
-\item [{Vs}] S-wave speed of the material (given in m/s).
-\item [{rho}] density of the material (given in kg/m$^{3}$).
-\item [{Q}] quality factor to use in case of a simulation with attenuation
-turned on. It should be between 1 and 9000. In case no attenuation
-information is available, it can be set to zero. Please note that
-your Vp- and Vs-speeds are given for a reference frequency. To change
-this reference frequency, you change the value of \texttt{ATTENUATION\_f0\_REFERENCE}
-in the main constants file \texttt{constants.h} found in subdirectory
-\texttt{src/shared/}.
-\item [{anisotropic\_flag}] Flag describing the anisotropic model to use
-in case an anisotropic simulation should be conducted. See the file
-\texttt{model\_aniso.f90} in subdirectory \texttt{src/generate\_databases/}
-for an implementation of the anisotropic models. In case no anisotropy
-is available, it can be set to zero.
-\end{description}
-
-Note that this material block has to be defined using all the volumes
-which belong to this elastic material. For volumes belonging to another,
-different material, you will need to define a new material block.
-
-For an acoustic material:
-\begin{description}
-\item [{material\_id}] An integer value which is unique for this material.
-\item [{Vp}] P-wave speed of the material (given in m/s).
-\item [{0}] S-wave speed of the material is ignored.
-\item [{rho}] density of the material (given in kg/m$^{3}$).
-\end{description}
-\item [{face\_topo}] Block definition for the surface which defines the
-free surface (which can have topography). The name of this block must
-be 'face\_topo', the block has to be defined using all the surfaces
-which constitute the complete free surface of the model.
-\item [{face\_abs\_xmin}] Block definition for the faces on the absorbing
-boundaries, one block for each surface with x=Xmin.
-\item [{face\_abs\_xmax}] Block definition for the faces on the absorbing
-boundaries, one block for each surface with x=Xmax.
-\item [{face\_abs\_ymin}] Block definition for the faces on the absorbing
-boundaries, one block for each surface with y=Ymin.
-\item [{face\_abs\_ymax}] Block definition for the faces on the absorbing
-boundaries, one block for each surface with y=Ymax.
-\item [{face\_abs\_bottom}] Block definition for the faces on the absorbing
-boundaries, one block for each surface with z=bottom.
-\end{description}
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=0.45\textwidth]{figures/mount-surface} \includegraphics[width=0.45\textwidth]{figures/mount-abs}
-\par\end{centering}
-
-\caption{Example of the block definitions for the free surface 'face\_topo'
-(left) and the absorbing boundaries, defined in a single block 'face\_abs\_xmin'
-(right) in CUBIT.}
-
-\label{fig:mount.abs}
-\end{figure}
-
-Optionally, instead of specifying for each surface at the boundaries
-a single block like mentioned above, you can also specify a single
-block for all boundary surfaces and name it as one of the absorbing
-blocks above, e.g. 'face\_abs\_xmin'.
-
-After the block definitions are done, you export the mesh using the
-script \texttt{cubit2specfem3d.py} provided in each of the example
-directories (linked to the common script \texttt{CUBIT\_GEOCUBIT/cubit2specfem3d.py}).
-If the export was successful, you should find the following
-files in a subdirectory \texttt{MESH/}:
-\begin{description}
-\item [{absorbing\_cpml\_file}] \textbf{(only needed in case of C-PML absorbing
-conditions)} Contains on the first line the total number of C-PML
-spectral elements in the mesh, and then on the following line the
-list of all these C-PML elements with two numbers per line: first
-the spectral element number, and then a C-PML flag indicating to which
-C-PML layer(s) that element belongs, according to the following convention:
-
-\begin{itemize}
-\item Flag = 1 : element belongs to a X CPML layer only (either in Xmin
-or in Xmax),
-\item Flag = 2 : element belongs to a Y CPML layer only (either in Ymin
-or in Ymax),
-\item Flag = 3 : element belongs to a Z CPML layer only (either in Zmin
-or in Zmax),
-\item Flag = 4 : element belongs to a X CPML layer and also to a Y CPML
-layer,
-\item Flag = 5 : element belongs to a X CPML layer and also to a Z CPML
-layer,
-\item Flag = 6 : element belongs to a Y CPML layer and also to a Z CPML
-layer,
-\item Flag = 7 : element belongs to a X, to a Y and to a Z CPML layer, i.e.,
-it belongs to a CPML corner.
-\end{itemize}
-
-Note that it does not matter whether an element belongs to a Xmin
-or to a Xmax CPML, the flag is the same in both cases; the same is
-true for Ymin and Ymax, and also Zmin and Zmax.
-
-When you have an existing CUBIT (or similar) mesh stored in SPECFEM3D
-format, i.e., if you have existing \textquotedbl{}nodes\_coords\_file\textquotedbl{}
-and \textquotedbl{}mesh\_file\textquotedbl{} files but do not know
-how to assign CPML flags to them, we have created a small serial Fortran
-program that will do that automatically for you, i.e., which will
-create the \textquotedbl{}absorbing\_cpml\_file\textquotedbl{} for
-you. That program is utils/CPML/convert\_external\_layers\_of\_a\_given\_mesh\_to\_CPML\_layers.f90,
-and a small Makefile is provided in that directory (utils/CPML).
-
-IMPORTANT: it is your responsibility to make sure that in the input
-CUBIT (or similar) mesh that this code will read in SPECFEM3D format
-from files \textquotedbl{}nodes\_coords\_file\textquotedbl{} and \textquotedbl{}mesh\_file\textquotedbl{}
-you have created layers of elements that constitute a layer of constant
-thickness aligned with the coordinate grid axes (X, Y and/or Z), so
-that this code can assign CPML flags to them. This code does NOT check
-that (because it cannot, in any easy way). The mesh inside these CPML
-layers does not need to be structured nor regular, any non-structured
-mesh is fine as long as it has flat PML inner and outer faces, parallel
-to the axes, and thus of a constant thickness. The thickness can be
-different for the X, Y and Z sides. But for X it must not vary, for
-Y it must not vary, and for Z it must not vary. If you do not know
-the exact thickness, you can use a slightly LARGER value in this code
-(say 2\% to 5\% more) and this code will fix that and will adjust
-it; never use a SMALLER value otherwise this code will miss some CPML
-elements.
-
-Note: in a future release we will remove the constraint of having
-CPML layers aligned with the coordinate axes; we will allow for meshes
-that are titled by any constant angle in the horizontal plane. However
-this is not implemented yet.
-
-Note: in the case of fluid-solid layers in contact, the fluid-solid
-interface currently needs to be flat and horizontal inside the CPML
-layer (i.e., bathymetry should become flat and horizontal when entering
-the CPML); this (small) constraint will probably remain in the code
-for a while because it makes fluid-solid matching inside the CPML
-much easier.
-
-\item [{materials\_file}] Contains the material associations for each element.
-The format is:
-
-\begin{lyxcode}
-element\_ID~material\_ID
-\end{lyxcode}
-
-where \texttt{element\_ID} is the element identifier and \texttt{material\_ID}
-is a unique identifier, positive (for materials taken from this list
-of materials, i.e. for which each spectral element has constant material
-properties taken from this list) or negative (for tomographic models,
-i.e. for spectral element whose real velocities and density will be
-assigned later by calling an external function to define model variations,
-for instance in the case of tomographic models; in such a case, material
-properties can vary inside each spectral element, i.e. be different
-at each of its Gauss-Lobatto-Legendre grid points).
-
-\item [{nummaterial\_velocity\_file}] Defines the material properties.
-
-\begin{itemize}
-\item For classical materials (i.e., spectral elements for which the velocity
-and density model will not be assigned by calling an external function
-to define for instance a tomographic model), the format is:
-
-\begin{lyxcode}
-domain\_ID~material\_ID~rho~vp~vs~Qkappa~Qmu~anisotropy\_flag
-\end{lyxcode}
-
-where \texttt{\textbf{domain\_ID}}\textbf{ is 1 for acoustic and 2
-for elastic or viscoelastic materials,} \texttt{material\_ID} a unique
-identifier, \texttt{rho} the density in $kg\, m^{-3}$, \texttt{vp}
-the P-wave speed in $m\, s^{-1}$, \texttt{vs} the S-wave speed in
-$m\, s^{-1}$, \texttt{Q} the quality factor and \texttt{anisotropy\_flag}
-an identifier for anisotropic models. Note that the \texttt{Qkappa}
-value is ignored by the code unless \texttt{FULL\_ATTENUATION\_SOLID}
-is set. Note also that both \texttt{Qkappa} and \texttt{Qmu} are ignored
-by the code unless \texttt{ATTENUATION} is set. If you want a model
-with no \texttt{Qmu} attenuation, both set \texttt{ATTENUATION} to
-\texttt{.false.} in the \texttt{Par\_file} and set \texttt{Qmu} to
-9999 here. If you want a model with no \texttt{Qkappa} attenuation,
-both set \texttt{FULL\_ATTENUATION\_SOLID} to \texttt{.false.} in
-the \texttt{Par\_file} and set \texttt{Qkappa} to 9999 here. %%magnoni
-
-\item For tomographic velocity models, please read Chapter \ref{cha:-Changing-the}
-and Section \ref{sec:Using-tomographic} `Using external tomographic
-Earth models' for further details.
-\end{itemize}
-\item [{nodes\_coords\_file}] Contains the point locations in Cartesian
-coordinates of the mesh element corners.
-\item [{mesh\_file}] Contains the mesh element connectivity. The hexahedral
-elements can have 8 or 27 nodes.\\
- See picture doc/mesh\_numbering\_convention/numbering\_convention\_27\_nodes.jpg
-to see\\
- in which (standard) order the points must be cited. In the case of
-8 nodes, just include the first 8 points.
-\item [{free\_or\_absorbing\_surface\_file\_zmax}] Contains the free surface
-connectivity or \\
- the surface connectivity of the absorbing boundary surface at the
-top (Zmax), \\
- depending on whether the top surface is defined as free or absorbing
-(\texttt{STACEY\_INSTEAD\_OF\_FREE\_SURFACE} in \texttt{DATA/Par\_file}). \\
- You should put both the surface of acoustic regions and of elastic
-regions in that file; that is, list all the element faces that constitute
-the surface of the model in that file.
-\item [{absorbing\_surface\_file\_xmax}] Contains the surface connectivity
-of the absorbing boundary surface at Xmax \\
-(also needed in the case of C-PML absorbing conditions, in order for
-the code to be able to impose Dirichlet conditions on their outer
-edge).
-\item [{absorbing\_surface\_file\_xmin}] Contains the surface connectivity
-of the absorbing boundary surface at Xmin \\
-(also needed in the case of C-PML absorbing conditions, in order for
-the code to be able to impose Dirichlet conditions on their outer
-edge).
-\item [{absorbing\_surface\_file\_ymax}] Contains the surface connectivity
-of the absorbing boundary surface at Ymax \\
-(also needed in the case of C-PML absorbing conditions, in order for
-the code to be able to impose Dirichlet conditions on their outer
-edge).
-\item [{absorbing\_surface\_file\_ymin}] Contains the surface connectivity
-of the absorbing boundary surface at Ymin \\
-(also needed in the case of C-PML absorbing conditions, in order for
-the code to be able to impose Dirichlet conditions on their outer
-edge).
-\item [{absorbing\_surface\_file\_bottom}] Contains the surface connectivity
-of the absorbing boundary surface at the bottom (Zmin) \\
-(also needed in the case of C-PML absorbing conditions, in order for
-the code to be able to impose Dirichlet conditions on their outer
-edge).
-\end{description}
-These mesh files are needed as input files for the partitioner \texttt{xdecompose\_mesh}
-to load-balance the mesh. Please see the next section for further
-details.
-
-In directory "CUBIT\_GEOCUBIT" we provide a script that can help doing the above tasks of exporting a CUBIT mesh
-to SPECFEM3D format automatically for you: "run\_cubit2specfem3d.py".
-Just edit them to indicate the path to your local installation of CUBIT and also the name of the *.cub existing CUBIT mesh file
-that you want to export to SPECFEM3D format. These scripts will do the conversion for you automatically except assigning material properties
-to the different mesh layers. To do so, you will then need to edit the file
-called "nummaterial\_velocity\_file" that will have just been created and change it from the prototype created:
-
-    0 1 vol1 --> syntax: \#material\_domain\_id \#material\_id \#rho \#vp \#vs \#Q\_kappa \#Q\_mu \#anisotropy
-    0 2 vol2 --> syntax: \#material\_domain\_id \#material\_id \#rho \#vp \#vs \#Q\_kappa \#Q\_mu \#anisotropy
-
-(where "vol1" and "vol2" here represent the volume labels that you have set while creating the mesh in CUBIT)
-to for instance
-
-    2 1 1500 2300 1800 9999.0 9999.0 0
-    2 2 1600 2500 20000 9999.0 9999.0 0
-
-\subsubsection*{Checking the mesh quality}
-
-The quality of the mesh may be inspected more precisely based upon
-the serial code in the file \texttt{check\_mesh\_quality\_}~\\
- \texttt{CUBIT\_Abaqus.f90} located in the directory \texttt{src/check\_mesh\_quality\_CUBIT\_Abaqus/}.
-Running this code is optional because no information needed by the
-solver is generated.
-
-Prior to running and compiling this code, you have to export your
-mesh in CUBIT to an ABAQUS (.inp) format. For example, export mesh
-block IDs belonging to volumes in order to check the quality of the
-hexahedral elements. You also have to determine a number of parameters
-of your mesh, such as the number of nodes and number of elements and
-modify the header of the \texttt{check\_mesh\_quality\_CUBIT\_Abaqus.f90}
-source file in directory \texttt{src/check\_mesh\_quality\_CUBIT\_Abaqus/}.
-
-\noindent
-Then, in the main directory, type
-\begin{lyxcode}
-{\small make~xcheck\_mesh\_quality\_CUBIT\_Abaqus}{\small \par}
-\end{lyxcode}
-and use
-\begin{lyxcode}
-{\small ./bin/xcheck\_mesh\_quality\_CUBIT\_Abaqus}{\small \par}
-\end{lyxcode}
-to generate an %not supported: AVS output file (\texttt{\small AVS\_meshquality.inp} in AVS UCD format) or
-OpenDX output file (\texttt{\small DX\_mesh\_quality.dx}{\small )
-that can be used to investigate mesh quality, e.g. skewness of elements,
-and a Gnuplot histogram (}\texttt{\small mesh\_quality\_histogram.txt}{\small )
-that can be plotted with gnuplot (type `}\texttt{\small gnuplot plot\_mesh\_quality\_histogram.gnu}{\small ').
-The histogram is also printed to the screen. Analyze that skewness
-histogram of mesh elements to make sure no element has a skewness
-above approximately 0.75, otherwise the mesh is of poor quality (and
-if even a single element has a skewness value above 0.80, then you
-must definitely improve the mesh). If you want to start designing
-your own meshes, this tool is useful for viewing your creations. You
-are striving for meshes with elements with `cube-like' dimensions,
-e.g., the mesh should contain no very elongated or skewed elements.}{\small \par}
-
-\subsection{Partitioning the Mesh with \texttt{xdecompose\_mesh}}
-
-The SPECFEM3D Cartesian software package performs large scale simulations
-in a parallel 'Single Process Multiple Data' way. The spectral-element
-mesh created with CUBIT needs to be distributed on the processors.
-This partitioning is executed once and for all prior to the execution
-of the solver so it is referred to as a static mapping.
-
-An efficient partitioning is important because it leverages the overall
-running time of the application. It amounts to balance the number
-of elements in each slice while minimizing the communication costs
-resulting from the placement of adjacent elements on different processors.
-\texttt{decompose\_mesh} depends on the SCOTCH library \citep{PeRo96},
-which provides efficent static mapping, graph and mesh partitioning
-routines. SCOTCH is a free software package developed by Fran\c{c}ois
-Pellegrini et al. from LaBRI and INRIA in Bordeaux, France, downloadable
-from the web page \url{https://gforge.inria.fr/projects/scotch/}.
-\\
-
-In most cases, the configuration with \texttt{./configure~FC=ifort}
-should be sufficient. During the configuration process, the script
-tries to find existing SCOTCH installations. In case your system has
-no pre-existing SCOTCH installation, we provide the source code of
-SCOTCH, which is released open source under the French CeCILL-C version
-1 license, in directory \texttt{src/decompose\_mesh/scotch\_5.1.12b}.
-This version gets bundled with the compilation of the SPECFEM3D Cartesian
-package if no libraries could have been found. If this automatic compilation
-of the SCOTCH libraries fails, please refer to file INSTALL.txt in
-that directory to see further details how to compile it on your system.
-In case you want to use a pre-existing installation, make sure you
-have correctly specified the path of the SCOTCH library when using
-the option \texttt{-{}-with-scotch-dir} with the \texttt{./configure}
-script. In the future you should be able to find more recent versions
-at \url{http://www.labri.fr/perso/pelegrin/scotch/scotch_en.html}
-. \\
-
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=0.45\textwidth]{figures/mount-partitions}
-\includegraphics[width=0.45\textwidth]{figures/mount-partitions2}
-\par\end{centering}
-
-\caption{Example of a mesh partitioning onto four cores. Each single core partition
-is colored differently. The executable \texttt{xdecompose\_mesh} can
-equally distribute the mesh on any arbitrary number of cores. Domain
-decomposition is explained in detail in \citet{MaKoBlLe08}, and excellent
-scaling up to 150,000 processor cores in shown for instance in \citet{CaKoLaTiMiLeSnTr08,KoLaMi08a,MaKoBlLe08,KoErGoMi10,Kom11}.}
-
-\label{fig:mount.partitions}
-\end{figure}
-
-When you are ready to compile, in the main directory type
-\begin{lyxcode}
-{\small make~xdecompose\_mesh}{\small \par}
-\end{lyxcode}
-If all paths and flags have been set correctly, the executable \texttt{bin/xdecompose\_mesh}
-should be produced.
-
-The partitioning is done in serial for now (in the next release we
-will provide a parallel version of that code). It needs to be run in the main directory because it expects the \texttt{./DATA/Par\_file}. The synopsis is:
-\begin{lyxcode}
-./bin/xdecompose\_mesh~nparts~input\_directory~output\_directory
-\end{lyxcode}
-where
-\begin{itemize}
-\item \texttt{nparts} is the number of partitions, i.e., the number of cores
-for the parallel simulations,
-\item \texttt{input\_directory} is the directory which holds all the files
-generated by the Python script \texttt{cubit2specfem3d.py} explained
-in the previous Section~\ref{subsec:Exporting-the-Mesh}, e.g. \texttt{./MESH/},
-and
-\item \texttt{output\_directory} is the directory for the output of this
-partitioner which stores ACII-format files named like \texttt{proc{*}{*}{*}{*}{*}{*}\_Database}
-for each partition. These files will be needed for creating the distributed
-databases, and have to reside in the directory \texttt{LOCAL\_PATH}
-specified in the main \texttt{Par\_file}, e.g. in directory \texttt{./OUTPUT\_FILES/DATABASES\_MPI}.
-Please see Chapter~\ref{cha:Creating-Distributed-Databases} for
-further details.
-\end{itemize}
-Note that all the files generated by the Python script \texttt{cubit2specfem3d.py}
-must be placed in the \texttt{input\_directory} folder before running
-the program.
-
-\section{\label{cha:Running-the-Mesher-Meshfem3D}Meshing with \texttt{xmeshfem3D}}
-
-In case you successfully ran the configuration script, you are also
-ready to compile the internal mesher. This is an alternative to CUBIT
-for the mesh generation of relatively simple models.
-
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[scale=0.5]{figures/socal_map_mpi}
-\par\end{centering}
-
-\caption{\label{fig:For-parallel-computing}For parallel computing purposes,
-the model block is subdivided in $\nprocxi\times\nproceta$ slices
-of elements. In this example we use $5^{2}=25$ processors. }
-\end{figure}
-
-\noindent
-In the main directory, type
-\begin{lyxcode}
-{\small make~xmeshfem3D}{\small \par}
-\end{lyxcode}
-If all paths and flags have been set correctly, the mesher should
-now compile and produce the executable \texttt{bin/xmeshfem3D}. Please
-note that \texttt{xmeshfem3D} must be called directly from the main
-directory, as most of the binaries of the package.
-
-Input for the mesh generation program is provided through the parameter
-file \texttt{Mesh\_Par\_file}, which resides in the subdirectory \texttt{DATA/meshfem3D\_files/}.
-(to see how to use it, see the EXAMPLES specific to the internal mesher in directory \texttt{EXAMPLES/meshfem3D\_examples/}.
-Before running the mesher, a number of parameters need to be set in
-the \texttt{Mesh\_Par\_file}. This requires a basic understanding
-of how the SEM is implemented, and we encourage you to read \citet{KoVi98,KoTr99}
-and \citet{KoLiTrSuStSh04}.
-
-The mesher and the solver use UTM coordinates internally, therefore
-you need to define the zone number for the UTM projection (e.g., zone
-11 for Los Angeles). Use decimal values for latitude and longitude
-(no minutes/seconds). These values are approximate; the mesher will
-round them off to define a square mesh in UTM coordinates. When running
-benchmarks on rectangular models, turn the UTM projection off by using
-the flag \texttt{\small SUPPRESS\_UTM\_PROJECTION}{\small , in which
-case all `longitude' parameters simply refer to the $x$~axis, and
-all `latitude' parameters simply refer to the $y$~axis. To run the
-mesher for a global simulation, the following parameters need to be
-set in the }\texttt{\small Mesh\_Par\_file}{\small :}{\small \par}
-\begin{description}
-\item [{\texttt{LATITUDE\_MIN}}] Minimum latitude in the block (negative
-for South).
-\item [{\texttt{LATITUDE\_MAX}}] Maximum latitude in the block.
-\item [{\texttt{LONGITUDE\_MIN}}] Minimum longitude in the block (negative
-for West).
-\item [{\texttt{LONGITUDE\_MAX}}] Maximum longitude in the block.
-\item [{\texttt{DEPTH\_BLOCK\_KM}}] Depth of bottom of mesh in kilometers.
-\item [{\texttt{\noun{UTM\_PROJECTION\_ZONE}}}] UTM projection zone in
-which your model resides, only valid when \texttt{SUPPRESS\_UTM\_}~\\
- \texttt{PROJECTION} is \texttt{.false.}.
-Use a negative zone number for the Southern hemisphere:
-the Northern hemisphere corresponds to zones +1 to +60,
-the Southern hemisphere to zones -1 to -60.
-We use the WGS84 (World Geodetic System 1984) reference ellipsoid for the UTM projection. If you prefer to use the Clarke 1866 ellipsoid,
-edit file \texttt{src/shared/utm\_geo.f90}, uncomment that ellipsoid and recompile the code.
-From http://en.wikipedia.org/wiki/Universal\_Transverse\_Mercator\_coordinate\_system :
-The Universal Transverse Mercator coordinate system was developed by the United States Army Corps of Engineers in the 1940s.
-The system was based on an ellipsoidal model of Earth. For areas within the contiguous United States
-the Clarke Ellipsoid of 1866 was used. For the remaining areas of Earth, including Hawaii, the International Ellipsoid was used.
-The WGS84 ellipsoid is now generally used to model the Earth in the UTM coordinate system,
-which means that current UTM northing at a given point can be 200+ meters different from the old one.
-For different geographic regions, other datum systems (e.g.: ED50, NAD83) can be used.
-\item [{\texttt{SUPPRESS\_UTM\_PROJECTION}}] set to be \texttt{.false.}
-when your model range is specified in geographical coordinates, and
-needs to be \texttt{.true.} when your model is specified in Cartesian
-coordinates. \noun{UTM projection zone in which your simulation region
-resides.}
-\item [{\texttt{INTERFACES\_FILE }}] File which contains the description
-of the topography and of the interfaces between the different layers
-of the model, if any. The number of spectral elements in the vertical
-direction within each layer is also defined in this file.
-\item [{$\nexxi$}] The number of spectral elements along one side of the
-block. This number \textit{must} be 8~$\times$~a multiple of $\nprocxi$
-defined below. Based upon benchmarks against semi-analytical discrete
-wavenumber synthetic seismograms \citep{KoLiTrSuStSh04}, determined
-that a $\nexxi=288$ run is accurate to a shortest period of roughly
-2~s. Therefore, since accuracy is determined by the number of grid
-points per shortest wavelength, for any particular value of $\nexxi$
-the simulation will be accurate to a shortest period determined by
-\begin{equation}
-\mbox{shortest period (s)}=(288/\nexxi)\times2.\label{eq:shortest_period}
-\end{equation}
- The number of grid points in each orthogonal direction of the reference
-element, i.e., the number of Gauss-Lobatto-Legendre points, is determined
-by \texttt{NGLLX} in the \texttt{constants.h} file. We generally use
-$\mbox{\texttt{NGLLX\/}}=5$, for a total of $5^{3}=125$ points per
-elements. We suggest not to change this value.
-\item [{$\nexeta$}] The number of spectral elements along the other side
-of the block. This number \textit{must} be 8~$\times$~a multiple
-of $\nproceta$ defined below.
-\item [{$\nprocxi$}] The number of processors or slices along one side
-of the block (see Figure~\ref{fig:For-parallel-computing}); we must
-have $\nexxi=8\times c\times\nprocxi$, where $c\ge1$ is a positive
-integer.
-\item [{$\nproceta$}] The number of processors or slices along the other
-side of the block; we must have $\nexeta=8\times c\times\nproceta$,
-where $c\ge1$ is a positive integer.
-\item [{\texttt{USE\_REGULAR\_MESH}}] set to be \texttt{.true.} if you
-want a perfectly regular mesh or \texttt{.false.} if you want to add
-doubling horizontal layers to coarsen the mesh. In this case, you
-also need to provide additional information by setting up the next
-three parameters.
-\item [{\texttt{NDOUBLINGS}}] The number of horizontal doubling layers.
-Must be set to \texttt{1} or \texttt{2} if \texttt{USE\_REGULAR\_MESH}
-is set to \texttt{.true.}.
-\item [{\texttt{NZ\_DOUBLING\_1}}] The position of the first doubling layer
-(only interpreted if \texttt{USE\_REGULAR\_MESH} is set to \texttt{.true.}).
-\item [{\texttt{NZ\_DOUBLING\_2}}] The position of the second doubling
-layer (only interpreted if \texttt{USE\_REGULAR\_MESH} is set to \texttt{.true.}
-and if \texttt{NDOUBLINGS} is set to \texttt{2}).
-\item [{\texttt{CREATE\_ABAQUS\_FILES}}] Set this flag to \texttt{.true.}
-to save Abaqus FEA \urlwithparentheses{www.simulia.com} mesh files
-for subsequent viewing. Turning the flag on generates files in the
-\texttt{LOCAL\_PATH} directory. See Section~\ref{sec:Mesh-graphics}
-for a discussion of mesh viewing features.
-\item [{\texttt{CREATE\_DX\_FILES}}] Set this flag to \texttt{.true.} to
-save OpenDX \urlwithparentheses{www.opendx.org} mesh files for subsequent
-viewing.
-\item [{\texttt{LOCAL\_PATH}}] Directory in which the partitions generated
-by the mesher will be written. Generally one uses a directory on the
-local disk of the compute nodes, although on some machines these partitions
-are written on a parallel (global) file system (see also the earlier
-discussion of the \texttt{LOCAL\_PATH\_IS\_ALSO\_GLOBAL} flag in Chapter~\ref{cha:Getting-Started}).
-The mesher generates the necessary partitions in parallel, one set
-for each of the $\nprocxi\times\nproceta$ slices that constitutes
-the mesh (see Figure~\ref{fig:For-parallel-computing}). After the
-mesher finishes, you can log in to one of the compute nodes and view
-the contents of the \texttt{LOCAL\_PATH} directory to see the files
-generated by the mesher. These files will be needed for creating the
-distributed databases, and have to reside in the directory \texttt{LOCAL\_PATH}
-specified in the main \texttt{Par\_file}, e.g. in directory \texttt{OUTPUT\_FILES/DATABASES\_MPI}.
-Please see Chapter~\ref{cha:Creating-Distributed-Databases} for
-further details.
-\item [{\texttt{NMATERIALS}}] The number of different materials in your
-model. In the following lines, each material needs to be defined as
-:
-
-\begin{lyxcode}
-material\_ID~rho~vp~vs~Q~anisotropy\_flag~domain\_ID
-\end{lyxcode}
-
-where
-\begin{itemize}
-\item \texttt{Q} : quality factor (0=no attenuation)
-\item \texttt{anisotropy\_flag} : 0=no anisotropy / 1,2,.. check with implementation
-in \texttt{aniso\_model.f90}
-\item \texttt{domain\_id} : 1=acoustic / 2=elastic
-\end{itemize}
-\item [{\texttt{NREGIONS}}] The number of regions in the mesh. In the following
-lines, because the mesh is regular or 'almost regular', each region
-is defined as :
-
-\begin{lyxcode}
-NEX\_XI\_BEGIN~NEX\_XI\_END~NEX\_ETA\_BEGIN~NEX\_ETA\_END~NZ\_BEGIN~NZ\_END~material\_ID
-\end{lyxcode}
-\end{description}
-The \texttt{INTERFACES\_FILE} parameter of \texttt{Mesh\_Par\_File}
-defines the file which contains the settings of the topography grid
-and of the interfaces grids. Topography is defined as a set of elevation
-values on a regular 2D grid. It is also possible to define interfaces
-between the layers of the model in the same way. The file needs to
-define several parameters:
-\begin{itemize}
-\item The number of interfaces, including the topography. This needs to
-be set at the first line. Then, from the bottom to the top of the
-model, you need to define the grids with:
-\item \texttt{SUPPRESS\_UTM\_PROJECTION} flag as described previously,
-\item number of points along $x$ and $y$ direction (NXI and NETA),
-\item minimal $x$ and $y$ coordinates (LONG\_MIN and LAT\_MIN),
-\item spacing between points along $x$ and $y$ (SPACING\_XI and SPACING\_ETA)
-and
-\item the name of the file which contains the elevation values (in $y$.$x$
-increasing order).
-\end{itemize}
-At the end of this file, you simply need to set the number of spectral
-elements in the vertical direction for each layer. We provide a few
-models in the {\texttt{EXAMPLES/}} directory. \\
-
-Finally, depending on your system, you might need to provide a file
-that tells MPI what compute nodes to use for the simulations. The
-file must have a number of entries (one entry per line) at least equal
-to the number of processors needed for the run. A sample file is provided
-in the file \texttt{mymachines}. This file is not used by the mesher
-or solver, but is required by the \texttt{go\_mesher} and \texttt{go\_solver}
-default job submission scripts. See Chapter~\ref{cha:Scheduler}
-for information about running the code on a system with a scheduler,
-e.g., LSF.
-
-Now that you have set the appropriate parameters in the \texttt{Mesh\_Par\_file}
-and have compiled the mesher, you are ready to launch it! This is
-most easily accomplished based upon the \texttt{go\_mesher} script.
-When you run on a PC cluster, the script assumes that the nodes are
-named n001, n002, etc. If this is not the case, change the \texttt{tr
--d `n'} line in the script. You may also need to edit the last command
-at the end of the script that invokes the \texttt{mpirun} command.
-See Chapter~\ref{cha:Scheduler} for information about running the
-code on a system with a scheduler, e.g., LSF.
-
-Mesher output is provided in the \texttt{OUTPUT\_FILES} directory
-in \texttt{output\_mesher.txt}; this file provides lots of details
-about the mesh that was generated. Please note that the mesher suggests
-a time step \texttt{DT} to run the solver with. The mesher output
-file also contains a table about the quality of the mesh to indicate
-possible problems with the distortions of elements. Alternatively,
-output can be directed to the screen instead by uncommenting a line
-in \texttt{constants.h}:
-\begin{lyxcode}
-!~uncomment~this~to~write~messages~to~the~screen~
-
-!~integer,~parameter~::~IMAIN~=~ISTANDARD\_OUTPUT
-\end{lyxcode}
-To control the quality of the mesh, check the standard output (either
-on the screen or in the \texttt{OUTPUT\_FILES} directory in \texttt{output\_mesher.txt})
-and analyze the skewness histogram of mesh elements to make sure no
-element has a skewness above approximately 0.75, otherwise the mesh
-is of poor quality (and if even a single element has a skewness value
-above 0.80, then you must definitely improve the mesh). To draw the
-skewness histogram on the screen, type \texttt{gnuplot plot\_mesh\_quality\_histogram.gnu}.
-
-\chapter{\label{cha:Creating-Distributed-Databases}Creating the Distributed
-Databases}
-
-After using \texttt{xmeshfem3D} or \texttt{xdecompose\_mesh}, the
-next step in the workflow is to compile \texttt{xgenerate\_}~\\
- \texttt{databases}. This program is going to create all the missing
-information needed by the SEM solver.
-
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=0.6\textwidth]{figures/workflow}
-\par\end{centering}
-
-\caption{Schematic workflow for a SPECFEM3D Cartesian simulation. The executable
-\texttt{xgenerate\_databases} creates the GLL mesh points and assigns
-specific model parameters.}
-
-\label{fig:workflow.databases}
-\end{figure}
-
-\noindent
-In the main directory, type
-\begin{lyxcode}
-{\small make~xgenerate\_databases}{\small \par}
-\end{lyxcode}
-Input for the program is provided through the main parameter file
-\texttt{Par\_file}, which resides in the subdirectory \texttt{DATA}.
-Please note that \texttt{xgenerate\_databases} must be called directly
-from the main directory, as most of the binaries of the package.
-
-\section{\label{cha:Main-Parameter}Main parameter file \texttt{Par\_file}}
-
-Before running \texttt{xgenerate\_databases}, a number of parameters
-need to be set in the main parameter \texttt{Par\_file} located in
-the subdirectory \texttt{DATA}:
-\begin{description}
-\item [{\texttt{SIMULATION\_TYPE}}] is set to 1 for forward simulations,
-2 for adjoint simulations (see Section \ref{sec:Adjoint-simulation-finite})
-and 3 for kernel simulations (see Section \ref{sec:Finite-Frequency-Kernels}).
-\item [{\texttt{SAVE\_FORWARD}}] is only set to \texttt{.true.} for a forward
-simulation with the last frame of the simulation saved, as part of
-the finite-frequency kernel calculations (see Section \ref{sec:Finite-Frequency-Kernels}).
-For a regular forward simulation, leave \texttt{SIMULATION\_TYPE}
-and \texttt{SAVE\_FORWARD} at their default values.
-\item [{\texttt{\noun{UTM\_PROJECTION\_ZONE}}}] UTM projection zone in
-which your model resides, only valid when \texttt{SUPPRESS\_UTM\_}~\\
- \texttt{PROJECTION} is \texttt{.false.}.
-\item [{\texttt{SUPPRESS\_UTM\_PROJECTION}}] set to be \texttt{.false.}
-when your model range is specified in the geographical coordinates,
-and needs to be \texttt{.true.} when your model is specified in a
-cartesian coordinates. \noun{UTM projection zone in which your simulation
-region resides.}
-\item [{\texttt{NPROC}}] The number of MPI processors, each one is assigned
-one slice of the whole mesh.
-\item [{\texttt{NSTEP}}] The number of time steps of the simulation. This
-controls the length of the numerical simulation, i.e., twice the number
-of time steps requires twice as much CPU time. This feature is not
-used at the time of generating the distributed databases but is required
-for the solver, i.e., you may change this parameter after running
-\texttt{xgenerate\_databases}.
-\item [{\texttt{DT}}] The length of each time step in seconds. This feature
-is not used at the time of generating the distributed databases but
-is required for the solver. Please see also Section~\ref{sec:Choosing-the-Time-Step}
-for further details.
-\item [{\texttt{NGNOD}}] The number of nodes for 2D and 3D shape functions
-for hexahedra. We use either 8-node mesh elements (bricks) or 27-node
-elements. If you use the internal mesher, the only option is 8-node
-bricks (27-node elements are not supported). \texttt{CUBIT} does not
-support HEX27 elements either (it can generate them, but they are
-flat, i.e. identical to HEX8). To generate HEX27 elements with curvature
-properly taken into account, you can use \texttt{Gmsh} \url{http://geuz.org/gmsh/}
-\item [{\texttt{MODEL}}] Must be set to one of the following:
-
-\begin{description}
-\item [{\textmd{Models defined by mesh parameters:}}] ~
-
-\begin{description}
-\item [{\texttt{default}}] Uses model parameters as defined by meshing
-procedures described in the previous Chapter~\ref{cha:Mesh-Generation}.
-\end{description}
-
-\item [{\textmd{1D~models~with~real~structure:}}] ~
-
-\begin{description}
-\item [{\texttt{1D\_prem}}] Isotropic version of the spherically symmetric
-Preliminary Reference Earth Model (PREM) \citep{DzAn81}.
-\item [{\texttt{1D\_socal}}] A standard isotropic 1D model for Southern
-California.
-\item [{\texttt{1D\_cascadia}}] Isotropic 1D profile for the Cascadia region.
-\end{description}
-
-\item [{\textmd{Fully~3D~models:}}] ~
-
-\begin{description}
-\item [{\texttt{aniso}}] For a user-specified fully anisotropic model.
-Parameters are set up in routines located in file \texttt{model\_aniso.f90}
-in directory \texttt{src/generate\_databases/}. See Chapter~\ref{cha:-Changing-the}
-for a discussion on how to specify your own 3D model.
-
-\item [{\texttt{external}}] For a user-specified isotropic model which
-uses externally defined model parameters. Uses external model definitions
-set up in routines located in file \texttt{model\_external\_values.f90}
-in directory \texttt{src/generate\_databases/}. Please modify these
-generic template routines to use your own model definitions.
-\item [{\texttt{gll}}] For a user-specified isotropic model which uses
-external binary files for $v_{p}$, $v_{s}$ and $\rho$. Binary files
-are given in the same format as when outputted by the \texttt{xgenerate\_databases}
-executable when using option \texttt{SAVE\_MESH\_FILES}. These binary
-files define the model parameters on all GLL points which can be used
-for iterative inversion procedures.
-\item [{\texttt{salton\_trough}}] A 3D $V_{p}$ model for Southern California.
-Users must provide the corresponding data file \texttt{regrid3\_vel\_p.bin}
-in directory \texttt{DATA/st\_3D\_block\_harvard/}.
-\item [{\texttt{tomo}}] For a user-specified 3D isotropic model which uses
-a tomographic model file \texttt{tomographic\_model.xyz} in directory
-\texttt{DATA}. See Section \ref{sec:Using-tomographic}, for a discussion
-on how to specify your own 3D tomographic model.
-\end{description}
-\end{description}
-\item [{\texttt{APPROXIMATE\_OCEAN\_LOAD}}] Set to \texttt{.true.} if the
-effect of the oceans on seismic wave propagation should be incorporated
-based upon the (rough) approximate treatment discussed in \citet{KoTr02b}.
-This feature is inexpensive from a numerical perspective, both in
-terms of memory requirements and CPU time. This approximation is accurate
-at periods of roughly 20~s and longer. At shorter periods the effect
-of water phases/reverberations is not taken into account, even when
-the flag is on. If you want to model the effect of a fluid-solid model
-at short periods, then set this flag to \texttt{.false.} and mesh
-the fluid layer explicitly in your mesher, so that it is computed
-accurately and without this approximation.
-\item [{\texttt{TOPOGRAPHY}}] This feature is only effective if \texttt{APPROXIMATE\_OCEAN\_LOAD}
-is set to \texttt{.true.}. Set to \texttt{.true.} if topography and
-bathymetry should be read in based upon the topography file specified
-in the main constants file \texttt{constants.h} found in subdirectory
-\texttt{src/shared/} to evaluate elevations. If not set, elevations
-will be read from the numerical mesh.
-\item [{\texttt{ATTENUATION}}] Set to \texttt{.true.} if attenuation should
-be incorporated. Turning this feature on increases the memory requirements
-significantly (roughly by a factor of~1.5), and is numerically fairly
-expensive. See \citet{KoTr99,KoTr02a} for a discussion on the implementation
-of attenuation based upon standard linear solids. Please note that
-the Vp- and Vs-velocities of your model are given for a reference
-frequency. To change this reference frequency, you change the value
-of \texttt{ATTENUATION\_f0\_REFERENCE} in the main constants file
-\texttt{constants.h} found in subdirectory \texttt{src/shared/}.
-\item [{\texttt{FULL\_ATTENUATION\_SOLID}}] Set to \texttt{.true.} to add
-$Q_{\kappa}$ attenuation to the simulations in addition to $Q_{\mu}$
-attenuation when the \texttt{ATTENUATION} flag above is on. Should
-be off in almost all cases for seismic wave propagation, since $Q_{\kappa}$
-is almost always negligible in Earth models. However, this parameter
-can be useful for instance for ocean acoustics simulations, when $Q_{\kappa}$
-is not negligible in the solid layer at the sea bottom.
-\item [{\texttt{ANISOTROPY}}] Set to \texttt{.true.} if you want to use
-an anisotropy model. Please see the file \texttt{model\_aniso.f90}
-in subdirectory \texttt{src/generate\_databases/} for the current
-implementation of anisotropic models.
-\item [{\texttt{TOMOGRAPHY\_PATH}}] Directory in which the tomography files
-are stored for using external tomographic Earth models (please read
-Chapter \ref{cha:-Changing-the} and Section \ref{sec:Using-tomographic}
-`Using external tomographic Earth models' for further details.).
-\item [{\texttt{USE\_OLSEN\_ATTENUATION}}] Set to \texttt{.true.} if you
-want to use the attenuation model that scaled from the S-wave speed
-model using Olsen's empirical relation (see \citet{OlDaBr03}).
-\item [{\texttt{OLSEN\_ATTENUATION\_RATIO}}] Determines the Olsen's constant
-in Olsen's empirical relation (see \citet{OlDaBr03}).
-\item [{\texttt{PML\_CONDITIONS}}] Set to \texttt{.true.} to turn on C-PML
-boundary conditions for a regional simulation. Both fluids and elastic
-solids are supported. Note that xmeshfem3d generated meshes do not support C-PML yet.
-\item [{\texttt{PML\_INSTEAD\_OF\_FREE\_SURFACE}}] Set to \texttt{.true.}
-to turn on C-PML boundary conditions on the top surface instead of
-the usual free surface.
-\item [{\texttt{f0\_FOR\_PML}}] Determines the dominant frequency that
-will be used in the calculation of PML damping profiles; this should
-be set to the same (or similar) dominant frequency as that of the
-source that you will use in your simulation. If you plan to use a
-Dirac source, then use the dominant frequency of the source wavelet
-with which you plan to convolve your seismograms later on in post-processing.
-\item [{\texttt{STACEY\_ABSORBING\_CONDITIONS}}] Set to \texttt{.true.}
-to turn on Clayton-Enquist absorbing boundary conditions (see \citet{KoTr99}).
-In almost all cases it is much better to use CPML absorbing layers
-(see the options above) and leave this flag to \texttt{.false.}.
-\item [{\texttt{STACEY\_INSTEAD\_OF\_FREE\_SURFACE}}] Set to \texttt{.true.}
-to turn on absorbing boundary conditions on the top surface which
-by default constitutes a free surface of the model.
-\item [{\texttt{CREATE\_SHAKEMAP}}] Set this flag to \texttt{.true.} to
-create a ShakeMap\textregistered{}, i.e., a peak ground velocity map
-of the maximum absolute value of the two horizontal components of
-the velocity vector.
-\item [{\texttt{MOVIE\_SURFACE}}] Set to \texttt{.false.}, unless you want
-to create a movie of seismic wave propagation on the Earth's surface.
-Turning this option on generates large output files. See Section~\ref{sec:Movies}
-for a discussion on the generation of movies. This feature is only
-relevant for the solver.
-\item [{\texttt{MOVIE\_TYPE}}] Set this flag to 1 to show the top surface
-(tomography + oceans) only, to 2 to show all external faces of the
-mesh (i.e. topography + vertical edges + bottom) in shakemaps and
-surface movies.
-\item [{\texttt{MOVIE\_VOLUME}}] Set to \texttt{.false.}, unless you want
-to create a movie of seismic wave propagation in the Earth's interior.
-Turning this option on generates huge output files. See Section~\ref{sec:Movies}
-for a discussion on the generation of movies. This feature is only
-relevant for the solver.
-\item [{\texttt{SAVE\_DISPLACEMENT}}] Set this flag to \texttt{.true.}
-if you want to save the displacement instead of velocity for the movie
-frames.
-\item [{\texttt{USE\_HIGHRES\_FOR\_MOVIES}}] Set this flag to \texttt{.true.}
-if you want to save the values at all the NGLL grid points for the
-movie frames.
-\item [{\texttt{NTSTEP\_BETWEEN\_FRAMES}}] Determines the number of timesteps
-between movie frames. Typically you want to save a snapshot every
-100 timesteps. The smaller you make this number the more output will
-be generated! See Section~\ref{sec:Movies} for a discussion on the
-generation of movies. This feature is only relevant for the solver.
-\item [{\texttt{HDUR\_MOVIE}}] Determines the half duration of the source
-time function for the movie simulations. When this parameter is set
-to be 0, a default half duration that corresponds to the accuracy
-of the simulation is provided. Otherwise, it adds this half duration
-to the half duration specified in the source file \texttt{CMTSOLUTION},
-thus simulates longer periods to make the movie images look smoother.
-\item [{\texttt{SAVE\_MESH\_FILES}}] Set this flag to \texttt{.true.} to
-save ParaView \urlwithparentheses{www.paraview.org} mesh files for
-subsequent viewing. Turning the flag on generates large (distributed)
-files in the \texttt{LOCAL\_PATH} directory. See Section~\ref{sec:Mesh-graphics}
-for a discussion of mesh viewing features.
-\item [{\texttt{LOCAL\_PATH}}] Directory in which the distributed databases
-will be written. Generally one uses a directory on the local disk
-of the compute nodes, although on some machines these databases are
-written on a parallel (global) file system (see also the earlier discussion
-of the \texttt{LOCAL\_PATH\_IS\_ALSO\_GLOBAL} flag in Chapter~\ref{cha:Getting-Started}).
-\texttt{xgenerate\_databases} generates the necessary databases in
-parallel, one set for each of the \texttt{NPROC} slices that constitutes
-the mesh (see Figure~\ref{fig:mount.partitions} and Figure~\ref{fig:For-parallel-computing}).
-After the executable finishes, you can log in to one of the compute
-nodes and view the contents of the \texttt{LOCAL\_PATH} directory
-to see the (many) files generated by \texttt{xgenerate\_databases}.
-Please note that the \texttt{LOCAL\_PATH} directory should already
-contain the output files of the partitioner, i.e. from \texttt{xdecompose\_mesh}
-or \texttt{xmeshfem3D}.
-\item [{\texttt{NTSTEP\_BETWEEN\_OUTPUT\_INFO}}] This parameter specifies
-the interval at which basic information about a run is written to
-the file system (\texttt{timestamp{*}} files in the \texttt{OUTPUT\_FILES}
-directory). If you have access to a fast machine, set \texttt{NTSTEP\_BETWEEN\_OUTPUT\_INFO}
-to a relatively high value (e.g., at least 100, or even 1000 or more)
-to avoid writing output text files too often. This feature is not
-used at the time of meshing. One can set this parameter to a larger
-value than the number of time steps to avoid writing output during
-the run.
-\item [{\texttt{NTSTEP\_BETWEEN\_OUTPUT\_SEISMOS}}] This parameter specifies
-the interval at which synthetic seismograms are written in the \texttt{LOCAL\_PATH}
-directory. If a run crashes, you may still find usable (but shorter
-than requested) seismograms in this directory. On a fast machine set
-\texttt{NTSTEP\_BETWEEN\_OUTPUT\_SEISMOS} to a relatively high value
-to avoid writing to the seismograms too often. This feature is only
-relevant for the solver.
-\item [{\texttt{USE\_FORCE\_POINT\_SOURCE}}] Turn this flag on to use a
-(tilted) \texttt{FORCESOLUTION} force point source instead of a \texttt{CMTSOLUTION}
-moment-tensor source. When the force source does not fall exactly
-at a grid point, the solver interpolates the force between grid points
-using Lagrange interpolants. This can be useful e.g. for oil industry
-foothills simulations in which the source is a vertical force, normal
-force, tilted force, or an impact etc. Note that in the \texttt{FORCESOLUTION}
-file, you will need to edit the East, North and vertical components
-of an arbitrary (not necessarily unitary, the code will normalize it automatically) direction vector of the force vector;
-thus refer to Appendix~\ref{cha:Coordinates} for the orientation
-of the reference frame. This vector is made unitary internally in
-the solver and thus only its direction matters here; its norm is ignored
-and the norm of the force used is the factor force source times the
-source time function.
-When using this option, by default the code can locate the force source
-anywhere between mesh points in order to honor its exact location; this
-is more precise than using the closest GLL mesh point, but it is also a bit slower.
-If needed, you can change that default behavior and force the code to use the closest
-GLL mesh point instead by setting flag \texttt{USE\_BEST\_LOCATION} to \texttt{.false.} instead of \texttt{.true.} in file
-\texttt{src/shared/constants.h.in} and running the \texttt{configure} script again and recompiling the code.
-\item [{\texttt{USE\_RICKER\_TIME\_FUNCTION}}] Turn this flag on to use
-a Ricker source time function instead of the source time functions
-set by default to represent a (tilted) \texttt{FORCESOLUTION} force
-point source or a \texttt{CMTSOLUTION} moment-tensor source. Originally,
-if a \texttt{CMTSOLUTION} moment-tensor source is used, a (pseudo)
-Heaviside step function with a very short half duration is defined
-for elastic cases to represent the permanent slip on the fault while
-in the acoustic case a Gaussian source time function with a similarly
-short half duration is defined to physically describe actions within
-the fluid. Otherwise, if a \texttt{FORCESOLUTION} force source is
-used, a (pseudo) Dirac delta source time function is defined by default.
-Any other source-time function may then be obtained by convolution.
-\item [{\texttt{PRINT\_SOURCE\_TIME\_FUNCTION}}] Turn this flag on to print
-information about the source time function in the file \texttt{OUTPUT\_FILES/plot\_source\_time\_function.txt}.
-This feature is only relevant for the solver.
-
-\item [{\texttt{NUMBER\_OF\_SIMULTANEOUS\_RUNS}}] adds the ability to run several calculations (several earthquakes)
-in an embarrassingly-parallel fashion from within the same run;
-this can be useful when using a very large supercomputer to compute
-many earthquakes in a catalog, in which case it can be better from
-a batch job submission point of view to start fewer and much larger jobs,
-each of them computing several earthquakes in parallel.
-To turn that option on, set parameter NUMBER\_OF\_SIMULTANEOUS\_RUNS to a value greater than 1.
-To implement that, we create NUMBER\_OF\_SIMULTANEOUS\_RUNS MPI sub-communicators,
-each of them being labeled "my\_local\_mpi\_comm\_world", and we use them
-in all the routines in "src/shared/parallel.f90", except in MPI\_ABORT() because in that case
-we need to kill the entire run.
-When that option is on, of course the number of processor cores used to start
-the code in the batch system must be a multiple of NUMBER\_OF\_SIMULTANEOUS\_RUNS,
-all the individual runs must use the same number of processor cores,
-which as usual is NPROC in the Par\_file,
-and thus the total number of processor cores to request from the batch system
-should be NUMBER\_OF\_SIMULTANEOUS\_RUNS * NPROC.
-All the runs to perform must be placed in directories called run0001, run0002, run0003 and so on
-(with exactly four digits).
-
-Imagine you have 10 independent calculations to do, each of them on 100 cores; you have three options:
-
-1/ submit 10 jobs to the batch system
-
-2/ submit a single job on 1000 cores to the batch, and in that script create a sub-array of jobs to start 10 jobs,
-each running on 100 cores (see e.g. http://www.schedmd.com/slurmdocs/job\_array.html )
-
-3/ submit a single job on 1000 cores to the batch, start SPECFEM3D on 1000 cores, create 10 sub-communicators,
-cd into one of 10 subdirectories (called e.g. run0001, run0002,... run0010) depending on the sub-communicator
-your MPI rank belongs to, and run normally on 100 cores using that sub-communicator.
-
-The option NUMBER\_OF\_SIMULTANEOUS\_RUNS implements 3/.
-
-\item [{\texttt{BROADCAST\_SAME\_MESH\_AND\_MODEL}}]:
-if we perform simultaneous runs in parallel, if only the source and receivers vary between these runs
-but not the mesh nor the model (velocity and density) then we can also read the mesh and model files
-from a single run in the beginning and broadcast them to all the others; for a large number of simultaneous
-runs for instance when solving inverse problems iteratively this can DRASTICALLY reduce I/Os to disk in the solver
-(by a factor equal to NUMBER\_OF\_SIMULTANEOUS\_RUNS), and reducing I/Os is crucial in the case of huge runs.
-Thus, always set this option to .true. if the mesh and the model are the same for all simultaneous runs.
-In that case there is no need to duplicate the mesh and model file database (the content of the DATABASES\_MPI
-directories) in each of the run0001, run0002,... directories, it is sufficient to have one in run0001
-and the code will broadcast it to the others).
-
-TODO / future work to do: currently the BROADCAST\_SAME\_MESH\_AND\_MODEL option assumes to have the (master) mesh files in run0001/DATABASES\_MPI or run0001/OUTPUT\_FILES/DATABASES\_MPI.
-However, for adjoint runs you still need a DATABASES\_MPI folder in each of the sub-runs directories, e.g.
-run0002/DATABASES\_MPI, etc.
-to store the forward wavefields, kernels etc.
-of each sub-run.
-This would not be needed for forward simulations.
-
-TODO / future work to do: the sensitivity kernel summing and smoothing tools in directory src/tomography are currently not ported to this new option to do many runs simultaneously, only the solver (src/specfem3d) is.
-Thus these tools should work, but in their current version will need to be run for each simulation result independently.
-More precisely, the current kernel summing and smoothing routines work fine, with the exception that you need to move out the mesh files (and also the parameters).
-This works because these routines consider multiple runs by design.
-You simply have to provide them the directories where the kernels are.
-
-\item [{\texttt{GPU\_MODE}}] Turn this flag on to use GPUs.
-
-\item [\texttt{ADIOS\_ENABLED}] Turn this flag on to enable ADIOS. If set to \texttt{.false.}, subsequent ADIOS
-parameters will not be considered.
-\item [\texttt{ADIOS\_FOR\_DATABASES}] Turn this flag on to use ADIOS for xmeshfem3D output and
-xgenerate\_database input.
-\item [\texttt{ADIOS\_FOR\_MESH}]  Turn this flag on to use ADIOS for generated databases.
-\item [\texttt{ADIOS\_FOR\_FORWARD\_ARRAYS}] Turn this flag on to read and write forward arrays using ADIOS.
-\item [\texttt{ADIOS\_FOR\_KERNELS}] Turn this flag on to produce ADIOS kernels that can later be visualized with the ADIOS version of combine\_vol\_data.
-
-\end{description}
-
-The present version of SPECFEM can handle fully saturated porous simulations, Christina Morency implemented Biot equation.
-But the code cannot calculate partially saturated cases in its state.
-Christina Morency is presently working on a dual porosity, dual permeability formulation, type Pride and Berryman, for an other project, but it will not be available for some time.\\
-
-The way we prescribe material property for porous material in SPECFEM3D is as follow:
-We use a file name nummaterial\_poroelastic\_file, which is located in the directory MESH, the format is as follow:\\
- rhos rhof phi c kxx kxy kxz kyy kyz kzz Ks Kf Kfr etaf mufr\\
-where\\
-rho\_s = solid density\\
-rho\_f = fluid density\\
-phi = porosity\\
-tort = tortuosity\\
-kxx = xx component of permeability tensor\\
-kxy = xy,yx components of permeability tensor\\
-kyy = yy component of permeability tensor\\
-kxz = xz,zx components of permeability tensor\\
-kzz = zz component of permeability tensor\\
-kappa\_s = solid bulk modulus\\
-kappa\_f= fluid bulk modulus\\
-kappa\_fr= frame bulk modulus\\
-eta\_f = fluid viscosity\\
-mu\_fr = frame shear modulus.\\
-
-Using an external mesh (for instance coming from CUBIT/TRELIS), poroelastic materials have the ID number 3, while 1 is elastic, 2 is acoustic ?
-see the example in the package: examples/homogeneous\_poroelastic.
-
-If you use PML, the mesh elements that belong to the PML layers can
-be acoustic or elastic, but not viscoelastic nor poroelastic. Then,
-when defining your model, you should define these absorbing elements
-as either acoustic or elastic. In you forget to do that, the code
-will fix the problem by automatically converting the viscoelastic
-or poroelastic PML elements to elastic. This means that strictly speaking
-the PML layer will not be perfectly matched any more, since the physical
-model will change from viscoelastic or poroelastic to elastic at the
-entrance of the PML, but in practice this is sufficient and produces
-only tiny / negligible spurious reflections.
-
-If you use PML and an external mesh (created using an external meshing tool
-such as CUBIT/TRELIS or similar), try to have elements inside the PML as regular as possible,
-i.e. ideally non-deformed cubes obtained by `extrusion' of regular surface mesh elements meshing the
-outer edges of the computational domain without PML; by doing so, the PMLs obtained will be far more stable
-in time (PML being weakly unstable from a mathematical point of view, very deformed mesh elements
-inside the PMLs can trigger instabilities much more quickly).
-
-If you use PML and an external tomographic velocity and density model,
-you should be careful because mathematically a PML cannot handle heterogeneities
-along the normal to the PML edge inside the PML layer. This comes
-from the fact that the damping profile that is defined assumes a constant
-velocity and density model along the normal direction.
-
-Thus, you need to modify your velocity and density model in order
-for it to be 1D inside the PML, as shown in Figure~\ref{fig:modify_external_velocity_model_to_use_PML}.
-
-This applies to the bottom layer as well; there you should make sure
-that your model is 1D and thus constant along the vertical direction.
-
-To summarize, only use a 3D velocity and density model inside the
-physical region, and in all the PML layers extend it by continuity
-from its values along the inner PML edge.
-
-%%
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=3in]{figures/how_to_use_PML_when_a_tomographic_velocity_model_is_used}
-\par\end{centering}
-
-\caption{How to modify your external 3D velocity and density model in order
-to use PML. Such a modification is not needed when using Stacey absorbing
-boundary conditions (but such conditions are significantly less efficient).}
-
-\label{fig:modify_external_velocity_model_to_use_PML}
-\end{figure}
-
-%%
-
-\section{\label{sec:Choosing-the-Time-Step}Choosing the time step \texttt{DT}}
-
-The parameter \texttt{DT} sets the length of each time step in seconds.
-The value of this parameter is crucial for the stability of the spectral-element
-simulation. Your time step \texttt{DT} will depend on the minimum
-ratio between the distance $h$ of neighboring mesh points and the
-wave speeds $v$ defined in your model. The condition for the time
-step $\Delta t$ is:
-\begin{lyxcode}
-$\Delta t<C~\mathrm{min}_{\Omega}(~h/v~)$
-\end{lyxcode}
-where $C$ is the so-called Courant number and $\Omega$ denotes the
-model volume. The distance $h$ depends on the mesh element size and
-the number of GLL points \texttt{NGLL} specified in the main constants
-file \texttt{constants.h} located in the \texttt{src/shared/} subdirectory.
-The wave speed $v$ is determined based on your model's P- (or S-)
-wave speed values.
-
-The database generator \texttt{xgenerate\_databases}, as well as the
-internal mesher \texttt{xmeshfem3D}, are trying to evaluate the value
-of $\Delta t$ for empirically chosen Courant numbers $C\sim0.3$.
-If you used the mesher \texttt{xmeshfem3D} to generate your mesh,
-you should set the value suggested in \texttt{OUTPUT\_FILES/output\_mesher.txt}
-file, which is created after the mesher completed. In case you used
-CUBIT to create the mesh, you might use an arbitrary value when running
-\texttt{xgenerate\_databases} and then use the value suggested in
-the ~\\
- \texttt{OUTPUT\_FILES/output\_mesher.txt} file after the database
-generation completed. Note that the implemented Newmark time scheme
-uses this time step globally, thus your simulations become more expensive
-for very small mesh elements in high wave-speed regions. Please be
-aware of this restriction when constructing your mesh in Chapter~\ref{cha:Mesh-Generation}.
-
-\chapter{\label{cha:Running-the-Solver}Running the Solver \texttt{xspecfem3D}}
-
-Now that you have successfully generated the databases, you are ready
-to compile the solver. In the main directory, type
-\begin{lyxcode}
-{\small make~xspecfem3D}{\small \par}
-\end{lyxcode}
-Please note that \texttt{xspecfem3D} must be called directly from
-the main directory, as most of the binaries of the package.\\
-
-\noindent
-The solver needs three input files in the \texttt{DATA} directory
-to run:
-\begin{description}
-\item [{\texttt{Par\_file}}] the main parameter file which was discussed
-in detail in the previous Chapter~\ref{cha:Creating-Distributed-Databases},
-\item [{\texttt{CMTSOLUTION}} or {\texttt{FORCESOLUTION}}] the earthquake
-source parameter file or the force source parameter file, and
-\item [{\texttt{STATIONS}}] the stations file.
-\end{description}
-Most parameters in the \texttt{Par\_file} should be set prior to running
-the databases generation. Only the following parameters may be changed
-after running \texttt{xgenerate\_databases}:
-\begin{itemize}
-\item the simulation type control parameters: \texttt{SIMULATION\_TYPE}
-and \texttt{SAVE\_FORWARD}
-\item the time step parameters \texttt{NSTEP} and \texttt{DT}
-\item the absorbing boundary control parameter \texttt{PML\_CONDITIONS}
-on condition that the\\
- \texttt{PML\_INSTEAD\_OF\_FREE\_SURFACE} flag remains unmodified
-after running the databases generation.
-\item the movie control parameters \texttt{MOVIE\_SURFACE}, \texttt{MOVIE\_VOLUME},
-and \texttt{NTSTEPS\_BETWEEN\_FRAMES}
-\item the ShakeMap\textregistered{}option \texttt{CREATE\_SHAKEMAP}
-\item the output information parameters \texttt{MOVIE\_TYPE}, \texttt{NTSTEP\_BETWEEN\_OUTPUT\_INFO}
-and \texttt{NTSTEP\_BETWEEN\_OUTPUT\_}~\\
- \texttt{SEISMOS}
-\item the \texttt{PRINT\_SOURCE\_TIME\_FUNCTION} flags
-\end{itemize}
-Any other change to the \texttt{Par\_file} implies rerunning both
-the database generator \texttt{xgenerate\_databases} and the solver
-\texttt{xspecfem3D}.
-
-For any particular earthquake, the \texttt{CMTSOLUTION} file that
-represents the point source may be obtained directly from the Harvard
-Centroid-Moment Tensor (CMT) web page \urlwithparentheses{www.seismology.harvard.edu}.
-It looks like the example shown in Fig. \ref{fig:CMTSOLUTION-file}.
-%\begin{lyxcode}
-\begin{figure}[htp]
-\noindent
-\begin{centering}
-{\small \includegraphics[width=0.8\textwidth]{figures/Hollywood_CMT}
-}
-\par
-\end{centering}
-\caption{\texttt{CMTSOLUTION} file based on the
-format from the Harvard CMT catalog. \textbf{M} is the moment tensor,
-$M_{0}${\small {} }is the seismic moment, and $M_{w}$ is the moment
-magnitude.}
-\label{fig:CMTSOLUTION-file}
-\end{figure}
-%\end{lyxcode}
-The \texttt{CMTSOLUTION} file should be edited in the following way:
-\begin{itemize}
-\item Set the latitude or UTM $x$ coordinate, longitude or UTM $y$ coordinate, depth of the source (in km).
-\item Set the \texttt{time shift} parameter equal to $0.0$ (the solver
-will not run otherwise.) The time shift parameter would simply apply
-an overall time shift to the synthetics, something that can be done
-in the post-processing (see Section \ref{sec:Process-data-and-syn}).
-\item For point-source simulations (see finite sources, page \pageref{To-simulate-a})
-we recommend setting the source half-duration parameter \texttt{half
-duration} equal to zero, which corresponds to simulating a step source-time
-function, i.e., a moment-rate function that is a delta function. If
-\texttt{half duration} is not set to zero, the code will use a Gaussian
-(i.e., a signal with a shape similar to a `smoothed triangle', as
-explained in \citet{KoTr02a} and shown in Fig~\ref{fig:gauss.vs.triangle})
-source-time function with half-width \texttt{half duration}. We prefer
-to run the solver with \texttt{half duration} set to zero and convolve
-the resulting synthetic seismograms in post-processing after the run,
-because this way it is easy to use a variety of source-time functions
-(see Section \ref{sec:Process-data-and-syn}). \citet{KoTr02a} determined
-that the noise generated in the simulation by using a step source
-time function may be safely filtered out afterward based upon a convolution
-with the desired source time function and/or low-pass filtering. Use
-the serial code \texttt{convolve\_source\_timefunction.f90} and the
-script \texttt{convolve\_source\_timefunction.csh} for this purpose,
-or alternatively use signal-processing software packages such as SAC
-\urlwithparentheses{www.llnl.gov/sac}. Type
-
-\begin{lyxcode}
-{\small make~xconvolve\_source\_timefunction}{\small \par}
-\end{lyxcode}
-
-to compile the code and then set the parameter \texttt{hdur} in \texttt{convolve\_source\_timefunction.csh}
-to the desired half-duration.
-
-\item The zero time of the simulation corresponds to the center of the triangle/Gaussian,
-or the centroid time of the earthquake. The start time of the simulation
-is $t=-1.5*\texttt{half duration}$ (the 1.5 is to make sure the moment
-rate function is very close to zero when starting the simulation).
-To convert to absolute time $t_{\mathrm{abs}}$, set
-
-\begin{lyxcode}
-$t_{\mathrm{abs}}=t_{\mathrm{pde}}+\texttt{time shift}+t_{\mathrm{synthetic}}$
-\end{lyxcode}
-
-where $t_{\mathrm{pde}}$ is the time given in the first line of the
-\texttt{CMTSOLUTION}, \texttt{time shift} is the corresponding value
-from the original \texttt{CMTSOLUTION} file and $t_{\mathrm{synthetic}}$
-is the time in the first column of the output seismogram.
-
-\end{itemize}
-\begin{figure}[htp]
-\noindent \begin{centering}
-\includegraphics[width=3.5in]{figures/gauss_vs_triangle_mod}
-\par\end{centering}
-
-\caption{Comparison of the shape of a triangle and the Gaussian function actually
-used.}
-
-\label{fig:gauss.vs.triangle}
-\end{figure}
-
-If you know the earthquake source in strike/dip/rake format rather
-than in \texttt{CMTSOLUTION} format, use the C code \texttt{SPECFEM3D\_GLOBE/utils/strike\_dip\_rake\_to\_CMTSOLUTION.c}
-to convert it. The conversion formulas are given for instance in \citet{AkRi80}.
-Note that the \citet{AkRi80} convention is slightly different from
-the Harvard \texttt{CMTSOLUTION} convention (the sign of some components
-is different). The C code outputs both.
-
-Centroid latitude and longitude should be provided in geographical
-coordinates. The code converts these coordinates to geocentric coordinates~\citep{DaTr98}.
-Of course you may provide your own source representations by designing
-your own \texttt{CMTSOLUTION} file. Just make sure that the resulting
-file adheres to the Harvard CMT conventions (see Appendix~\ref{cha:Coordinates}).
-Note that the first line in the \texttt{CMTSOLUTION} file is the Preliminary
-Determination of Earthquakes (PDE) solution performed by the USGS
-NEIC, which is used as a seed for the Harvard CMT inversion. The PDE
-solution is based upon P waves and often gives the hypocenter of the
-earthquake, i.e., the rupture initiation point, whereas the CMT solution
-gives the `centroid location', which is the location with dominant
-moment release. The PDE solution is not used by our software package
-but must be present anyway in the first line of the file.
-
-\label{To-simulate-a}To simulate a kinematic rupture, i.e., a finite-source
-event, represented in terms of $N_{\mathrm{sources}}$ point sources,
-provide a \texttt{CMTSOLUTION} file that has $N_{\mathrm{sources}}$
-entries, one for each subevent (i.e., concatenate $N_{\mathrm{sources}}$
-\texttt{CMTSOLUTION} files to a single \texttt{CMTSOLUTION} file).
-At least one entry (not necessarily the first) must have a zero \texttt{time
-shift}, and all the other entries must have non-negative \texttt{time
-shift}. Each subevent can have its own half duration, latitude, longitude,
-depth, and moment tensor (effectively, the local moment-density tensor).
-
-Note that the zero in the synthetics does NOT represent the hypocentral
-time or centroid time in general, but the timing of the \textit{center}
-of the source triangle with zero \texttt{time shift} (Fig~\ref{fig:source_timing}).
-
-Although it is convenient to think of each source as a triangle, in
-the simulation they are actually Gaussians (as they have better frequency
-characteristics). The relationship between the triangle and the gaussian
-used is shown in Fig~\ref{fig:gauss.vs.triangle}. For finite fault
-simulations it is usually not advisable to use a zero half duration
-and convolve afterwards, since the half duration is generally fixed
-by the finite fault model.
-
-\vspace{1cm}
-
-\noindent The \texttt{FORCESOLUTION} file should be edited in the
-following way:
-\begin{itemize}
-\item Set the \texttt{time shift} parameter equal to $0.0$ (the solver
-will not run otherwise.) The time shift parameter would simply apply
-an overall time shift to the synthetics, something that can be done
-in the post-processing (see Section \ref{sec:Process-data-and-syn}).
-\item Set the \texttt{f0} parameter (the dominant frequency) of the Ricker source time
-function when {\texttt{USE\_RICKER\_TIME\_FUNCTION}} is turned on
-in the main parameter file {\texttt{Par\_file}}. In case that the
-solver uses a (pseudo) Dirac delta source time function to represent
-a force point source, a very short duration of five time steps is automatically set by default.
-\item Set the latitude or UTM $x$ coordinate, longitude or UTM $y$ coordinate, depth of the source (in km).
-\item Set the magnitude of the force source.
-\item Set the components of a (non-unitary) direction vector for the force
-source in the East/North/Vertical basis (see Appendix A for the orientation
-of the reference frame).
-\end{itemize}
-\noindent Where necessary, set a \texttt{FORCESOLUTION} file in the
-same way you configure a \texttt{CMTSOLUTION} file with $N_{\mathrm{sources}}$
-entries, one for each subevent (i.e., concatenate $N_{\mathrm{sources}}$
-\texttt{FORCESOLUTION} files to a single \texttt{FORCESOLUTION} file).
-At least one entry (not necessarily the first) must have a zero \texttt{time
-shift}, and all the other entries must have non-negative \texttt{time
-shift}. Each subevent can have its own half latitude, longitude, depth,
-\texttt{half duration} and force parameters.
-
-\begin{figure}[H]
-\noindent \begin{centering}
-{\small \includegraphics[width=5in]{figures/source_timing} }
-\par\end{centering}
-
-\caption{Example of timing for three sources. The center of the first source
-triangle is defined to be time zero. Note that this is NOT in general
-the hypocentral time, or the start time of the source (marked as tstart).
-The parameter \texttt{time shift} in the \texttt{CMTSOLUTION} file
-would be t1(=0), t2, t3 in this case, and the parameter \texttt{half
-duration} would be hdur1, hdur2, hdur3 for the sources 1, 2, 3 respectively.}
-
-{\small \label{fig:source_timing} }
-\end{figure}
-
-The solver can calculate seismograms at any number of stations for
-basically the same numerical cost, so the user is encouraged to include
-as many stations as conceivably useful in the \texttt{STATIONS} file,
-which looks like this:
-
-\begin{figure}[H]
-\noindent \begin{centering}
-{\small \includegraphics{figures/STATIONS_basin_explained} }
-\par\end{centering}
-
-\label{fig:Sample-STATIONS-file.}\caption{Sample \texttt{STATIONS} file. Station latitude and longitude should
-be provided in geographical coordinates. The width of the station
-label should be no more than 32 characters (see \texttt{MAX\_LENGTH\_STATION\_NAME}
-in the \texttt{constants.h} file), and the network label should be
-no more than 8 characters (see \texttt{MAX\_LENGTH\_NETWORK\_NAME}
-in the \texttt{constants.h} file).}
-\end{figure}
-
-Each line represents one station in the following format:
-\begin{lyxcode}
-{\small Station~Network~Latitude~(degrees)~Longitude~(degrees)~Elevation~(m)~burial~(m)~}{\small \par}
-\end{lyxcode}
-The solver \texttt{xspecfem3D} filters the list of stations in file
-\texttt{DATA/STATIONS} to exclude stations that are not located within
-the region given in the \texttt{Par\_file} (between \texttt{LATITUDE\_MIN}
-and \texttt{LATITUDE\_MAX} and between \texttt{LONGITUDE\_MIN} and
-\texttt{LONGITUDE\_MAX}). The filtered file is called \texttt{DATA/STATIONS\_FILTERED}.
-
-Solver output is provided in the \texttt{OUTPUT\_FILES} directory
-in the \texttt{output\_solver.txt} file. Output can be directed to
-the screen instead by uncommenting a line in \texttt{constants.h}:
-\begin{lyxcode}
-!~uncomment~this~to~write~messages~to~the~screen~
-
-!~integer,~parameter~::~IMAIN~=~ISTANDARD\_OUTPUT~
-\end{lyxcode}
-On PC clusters the seismogram files are generally written to the local
-disks (the path \texttt{LOCAL\_PATH} in the \texttt{Par\_file}) and
-need to be gathered at the end of the simulation.
-
-While the solver is running, its progress may be tracked by monitoring
-the `\texttt{\small timestamp{*}}{\small ' files in the ~}\\
-{\small{} }\texttt{\small OUTPUT\_FILES}{\small{} directory. These tiny
-files look something like this:}{\small \par}
-\begin{lyxcode}
-Time~step~\#~~~~~~~~~~10000~
-
-Time:~~~~~108.4890~~~~~~seconds~
-
-Elapsed~time~in~seconds~=~~~~1153.28696703911~
-
-Elapsed~time~in~hh:mm:ss~=~~~~~0~h~19~m~13~s~
-
-Mean~elapsed~time~per~time~step~in~seconds~=~~~~~0.115328696703911~
-
-Max~norm~displacement~vector~U~in~all~slices~(m)~=~~~~1.0789589E-02~
-\end{lyxcode}
-The \texttt{\small timestamp{*}}{\small{} files provide the }\texttt{\small Mean
-elapsed time per time step in seconds}{\small , which may be used
-to assess performance on various machines (assuming you are the only
-user on a node), as well as the }\texttt{\small Max norm displacement
-vector U in all slices~(m)}{\small . If something is wrong with the
-model, the mesh, or the source, you will see the code become unstable
-through exponentially growing values of the displacement and fluid
-potential with time, and ultimately the run will be terminated by
-the program. You can control the rate at which the timestamp files
-are written based upon the parameter }\texttt{\small NTSTEP\_BETWEEN\_OUTPUT\_INFO}{\small{}
-in the }\texttt{\small Par\_file}{\small .}{\small \par}
-
-Having set the \texttt{Par\_file} parameters, and having provided
-the \texttt{CMTSOLUTION} (or the \texttt{FORCESOLUTION}) and \texttt{STATIONS}
-files, you are now ready to launch the solver! This is most easily
-accomplished based upon the \texttt{go\_solver} script (See Chapter
-\ref{cha:Scheduler} for information about running through a scheduler,
-e.g., LSF). You may need to edit the last command at the end of the
-script that invokes the \texttt{mpirun} command. The \texttt{runall}
-script compiles and runs both \texttt{xgenerate\_databases} and \texttt{xspecfem3D}
-in sequence. This is a safe approach that ensures using the correct
-combination of distributed database output and solver input.
-
-It is important to realize that the CPU and memory requirements of
-the solver are closely tied to choices about attenuation (\texttt{ATTENUATION})
-and the nature of the model (i.e., isotropic models are cheaper than
-anisotropic models). We encourage you to run a variety of simulations
-with various flags turned on or off to develop a sense for what is
-involved.
-
-For the same model, one can rerun the solver for different events
-by simply changing the \texttt{CMTSOLUTION} or \texttt{FORCESOLUTION}
-file, or for different stations by changing the \texttt{STATIONS}
-file. There is no need to rerun the \texttt{xgenerate\_databases}
-executable. Of course it is best to include as many stations as possible,
-since this does not add to the cost of the simulation.
-
-We have also added the ability to run several calculations (several earthquakes)
-in an embarrassingly-parallel fashion from within the same run;
-this can be useful when using a very large supercomputer to compute
-many earthquakes in a catalog, in which case it can be better from
-a batch job submission point of view to start fewer and much larger jobs,
-each of them computing several earthquakes in parallel.
-To turn that option on, set parameter \texttt{NUMBER\_OF\_SIMULTANEOUS\_RUNS}
-to a value greater than 1 in file setup/constants.h.in before
-configuring and compiling the code.
-When that option is on, of course the number of processor cores used to start
-the code in the batch system must be a multiple of \texttt{NUMBER\_OF\_SIMULTANEOUS\_RUNS},
-all the individual runs must use the same number of processor cores,
-which as usual is \texttt{NPROC} in the input file \texttt{DATA/Par\_file},
-and thus the total number of processor cores to request from the batch system
-should be \texttt{NUMBER\_OF\_SIMULTANEOUS\_RUNS $\times$ NPROC}.
-All the runs to perform must be placed in directories called \texttt{run0001}, \texttt{run0002}, \texttt{run0003} and so on (with exactly four digits)
-and you must create a link from the root directory of the code
-to the first copy of the executable programs by typing \texttt{ln -s run0001/bin bin}.
-
-\chapter{\label{cha:Fault-Kinematics-Dynamics}Kinematic and dynamic fault
-sources}
-
-SPECFEM3D can handle finite fault sources of two kinds:
-\begin{enumerate}
-\item \emph{Kinematic}: the spatio-temporal distribution of slip rate is
-prescribed along the fault surface
-\item \emph{Dynamic}: a friction law and initial stresses are prescribed
-on the fault, a spontaneous rupture process is computed.
-\end{enumerate}
-
-\section{\label{sec:Mesh-Generation-with}Mesh Generation with Split Nodes}
-
-Faults need to be handled in a special way during mesh generation.
-A fault surface must lie at the interface between elements (the mesh
-must honor the fault surfaces). Moreover, a fault is made of two surfaces
-in contact. Each of these two surfaces needs a separate set of nodes.
-This approach is known as \textquotedbl{}split nodes\textquotedbl{}.
-Currently faults can only be run with \texttt{xdecompose\_mesh} and
-\texttt{CUBIT}. \texttt{xmeshfem3D} is not yet ready to handle faults.\\
-To facilitate the mesh generation with split nodes in CUBIT, we need
-to separate the two fault surfaces by a small distance, effectively
-creating a tiny opening of the fault (Figure \ref{fig:examples.splitnodes},
-\ref{fig:examples.splitnodes-surfacetrace}). Note that only the interior
-of the fault must be opened, its edges must remain closed (except
-the edge on the free surface). The fault is automatically closed later
-by SPECFEM3D.\\
-
-Here is an example CUBIT script to generate a mesh with split nodes
-for a buried vertical strike-slip fault:
-\begin{lyxcode}
-~reset~~\\
-~brick~x~10~y~10~z~10~~\\
-~webcut~volume~all~with~plane~xplane~~\\
-~webcut~volume~all~with~plane~yplane~~\\
-~webcut~volume~all~with~plane~xplane~offset~3~~\\
-~webcut~volume~all~with~plane~zplane~offset~3~~\\
-~webcut~volume~all~with~plane~zplane~offset~-3~~\\
-~imprint~all~~\\
-~merge~all~~\\
-~unmerge~surf~160~~\\
-~mesh~vol~all~~\\
-~set~node~constraint~off~~\\
-~node~in~surf~168~move~X~0~Y~0.01~Z~0~~\\
-~node~in~surf~160~move~X~0~Y~-0.01~Z~0~\\
-
-\end{lyxcode}
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[scale=0.3]{figures/faultmesh} \\
- \includegraphics[scale=0.3]{figures/surf168} \includegraphics[scale=0.3]{figures/splitnodes}
-\par\end{centering}
-
-\caption{Screenshots of the CUBIT example for the embedded fault with split
-nodes: The entire mesh is decomposed into several volumes bounding
-the fault (top), the fault surface 168 is shown as orange squares
-(middle) and split nodes of the fault shown as orange dots(bottom).
-Note that only interior nodes of the fault are split while those on
-the edges of the fault surface are touching each other.}
-
-\label{fig:examples.splitnodes}
-\end{figure}
-
-\begin{lyxcode}
-
-\end{lyxcode}
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[scale=0.55]{figures/splitnodes_surfacetrace} \\
-
-\par\end{centering}
-
-\caption{Screenshots of a CUBIT example showing split nodes for a fault reaching
-the surface. Surface trace of the fault is shown in orange. Note that
-the edges of the fault are not split while the interior nodes are
-offset by a small distance on either side of the fault}
-
-\label{fig:examples.splitnodes-surfacetrace}
-\end{figure}
-
-The CUBIT scripts ({*}.jou and {*}.py) in the directory \texttt{EXAMPLES}
-generate more complicated meshes. The {*}.py files are Python scripts
-that execute CUBIT commands and use the CUBIT-python interface for
-SPECFEM3D (see next section). The Python language allows to define
-and manipulate variables to parameterize the mesh. Alternatively,
-the Python script can call a CUBIT journal file ({*}.jou), which looks
-like the example above. Variables can be defined and manipulated there
-using the $\mathtt{APREPRO}$ language built in CUBIT.\\
-
-Note that you should avoid gaps in the list of indices of mesh objects
-with the following CUBIT command:
-\begin{lyxcode}
-$\mathtt{compress\: ids\: hex\: face\: edge\: node}$
-\end{lyxcode}
-(otherwise you will get a segmentation fault during domain decomposition)
-
-\section{CUBIT-Python Scripts for Faults}
-
-The mesh generated in CUBIT needs to be processed and exported in
-a format compatible with SPECFEM3D. This is achieved in the Python
-scripts by calling the Python-CUBIT interface functions defined in
-the CUBIT directory:
-\begin{enumerate}
-\item Function $\mathtt{define\_bc}$ (or $\mathtt{boundary\_definition.py}$)
-must be called to set up the absorbing boundaries database
-\item Function $\mathtt{fault\_input}$ must be called once for each fault
-to set up the fault database
-\item Function $\mathtt{cubit2specfem3d.export2SPECFEM3D}$ must be called
-at the very end of the script to export the mesh in a SPECFEM3D format.
-\end{enumerate}
-The functions in \#1 and \#3 are for the bulk and are documented in
-Section \ref{subsec:Exporting-the-Mesh}. We focus here on \#2:
-\begin{description}
-\item [{Function:}] $\mathtt{fault\_input}$
-\item [{Purpose:}] export a fault mesh database from CUBIT to a SPECFEM3D-compliant
-file
-\item [{Syntax:}] fault\_input(id\_fault, ids\_surf\_1, ids\_surf\_2)
-\item [{Inputs:}]~\end{description}
-\begin{itemize}
-\item id\_fault integer index assigned to the fault. (The user must number
-all the faults, starting at 1, with unit increments)
-\item ids\_surf\_1 list of CUBIT ids of all surfaces that form side 1 of
-the fault.
-\item ids\_surf\_2 list of CUBIT ids of all surfaces that form side 2 of
-the fault. (The user must decide which side of the fault is side 1.
-This choice affects the sign conventions of fault quantities as explained
-in Section \ref{sec:Sign-Convention-for}).\end{itemize}
-\begin{description}
-\item [{Outputs:}] file $\mathtt{fault\_file\_X.dat}$, where X is the
-fault id (id\_fault).
-\item [{Example:}] For the example in Section \ref{sec:Mesh-Generation-with}:\end{description}
-\begin{lyxcode}
-$\hphantom{}$A1~=~{[}168{]}
-
-$\hphantom{}$A2~=~{[}160{]}
-
-$\hphantom{}$fault\_input(1,A1,A2)
-\end{lyxcode}
-
-\section{Examples}
-
-The package includes examples, the SCEC benchmark problems:
-\begin{itemize}
-\item TPV5, a planar vertical strike-slip fault
-\item TPV14 and TPV15, vertical strike-slip fault system with a fault branch
-\item TPV102 and TPV103, rate-and-state benchmarks
-\item TPV22 and TPV23, step-over benchmarks
-\item TPV16, heterogenous initial stress
-\end{itemize}
-and
-\begin{itemize}
-\item Splay fault models from Wendt et al. (2009)\\
-
-\end{itemize}
-Read the documents in the directory \texttt{EXAMPLES/*/description}.
-They contain a description of the example and additional instructions
-to run it. Visualize the results with the Matlab scripts in the directory
-$\mathtt{EXAMPLES/*/post}$
-
-\section{\label{sec:Sign-Convention-for}Sign Convention for Fault Quantities}
-
-During mesh generation, the fault is defined by two surfaces in contact.
-Let's denote as \textquotedbl{}\emph{side 1}\textquotedbl{} the SECOND
-surface declared by the user in the call to the python function \textquotedbl{}\emph{fault\_input}\textquotedbl{},
-and the FIRST surface as \textquotedbl{}\emph{side 2}\textquotedbl{}.
-The local coordinate system on the fault is defined as the right-handed
-coordinate system defined by (strike, dip, normal), where \textquotedbl{}\emph{normal}\textquotedbl{}
-is the normal vector outgoing from side 1, \textquotedbl{}\emph{dip}\textquotedbl{}
-is parallel to the along-dip direction pointing downwards, and \textquotedbl{}\emph{strike}\textquotedbl{}
-is the horizontal along-strike vector such that the system is right-handed.\\
-
-Slip is defined as displacement on side 2 minus displacement on side
-1. In the local coordinate system on the fault, positive along-strike
-slip is right-lateral and positive along-dip slip is thrust if side
-1 is on the hanging wall (normal faulting if side 1 is on the foot
-wall).\\
-
-Traction is defined as the stress induced on side 1 by side 2, which
-is the stress tensor times the normal vector outgoing from side 1.
-In the local coordinate system on the fault, the normal traction is
-negative in compression, positive along-strike traction generates
-right-lateral slip and positive along-dip traction generates thrust
-slip if side 1 is on the hanging wall (normal faulting if side 1 is
-on the foot wall).
-
-\section{Input Files}
-
-Three additional input files are required in the \texttt{DATA/} directory
-for dynamic and kinematic faults. They are $\mathtt{Par\_file\_faults}$,
-$\mathtt{FAULT\_STATIONS}$ and $\mathtt{input\_file.txt}$ (optional).
-If the former does not exist, the code assumes that there are no faults.\\
-
-\begin{lyxlist}{00.00.0000}
-\item [{\textbf{DATA/Par\_file\_faults}}] contains parameters of the fault.
-The first part of this file has a strict format:
-
-\begin{description}
-\item [{Line}] 1: Number of faults (NF)
-\item [{Lines}] 2 to NF+1: Kelvin Voigt damping (in seconds) for each fault.
-(See below how to set this parameter)
-\item [{Line}] NF+2: Type of simulation (1=dynamic , 2 = kinematic)
-\item [{Line}] NF+3: Number of time steps between updates of the time series
-outputs at selected fault points (see DATA/FAULT\_STATIONS), usually
-a large number (100s or 1000s). Note that the sampling rate of the
-time series is usually much higher.
-\item [{Line}] NF+4: Number of time steps between fault snapshot outputs
-(quantities at every fault point exported at regular times), usually
-a large number (100s or 1000s).
-\item [{Line}] NF+5: Slip velocity threshold below which frictional healing
-is set (friction coefficient is reset to its static value). If this
-value is negative healing is disabled.
-\item [{Line}] NF+6: Slip velocity threshold to define the rupture front.
-Only used for outputs.
-\end{description}
-
-The rest of this file is made of namelist input blocks (see \textquotedbl{}namelist\textquotedbl{}
-in a Fortran 9x manual). The input for each fault has the following
-sequence (arguments in {[}brackets{]} are optional):
-
-\&\textbf{BEGIN\_FAULT} /
-
-\&\textbf{INIT\_STRESS} S1, S2, S3 {[},n1, n2, n3{]} /
-
-followed by (n1+n2+n3) \&\textbf{DIST2D} blocks
-
-\&\textbf{SWF} mus, mud, dc {[}, nmus, nmud, ndc{]} /
-
-followed by (nmus+nmud+ndc) \&\textbf{DIST2D} blocks\\
-
-\begin{lyxlist}{00.00.0000}
-\item [{\&\textbf{INIT\_STRESS}}] input block sets the initial fault stresses
-relative to the foot-wall side of the fault. Initial stresses are
-composed of a constant background value possibly overwritten in prescribed
-regions by heterogeneous distributions (see \&\textbf{DIST2D} blocks
-below):
-
-\begin{description}
-\item [{S1}] = constant background value of along-strike shear stress (positive
-in the usual strike direction)
-\item [{S2}] = constant background value of along-dip shear (positive is
-down-dip, normal faulting)
-\item [{S3}] = constant background value of normal stress (negative in
-compression)
-\item [{n1}] = number of heterogeneous items for along-strike shear stress
-{[}default is 0{]}
-\item [{n2}] = number of heterogeneous items for along-dip shear stress
-{[}default is 0{]}
-\item [{n3}] = number of heterogeneous items for normal stress {[}default
-is 0{]}
-\end{description}
-\item [{\&\textbf{SWF}}] input block sets the slip-weakening friction parameters
-of the fault:
-
-\begin{description}
-\item [{mus}] = constant background value of static friction coefficient
-\item [{mud}] = constant background value of dynamic friction coefficient
-\item [{dc}] = constant background value of critical slip-weakening distance
-\item [{nmus}] = number of heterogeneous items for static friction coefficient
-{[}default is 0{]}
-\item [{nmud}] = number of heterogeneous items for dynamic friction coefficient
-{[}default is 0{]}
-\item [{ndc}] = number of heterogeneous items for critical slip-weakening
-distance {[}default is 0{]}\\
-
-\end{description}
-\item [{\&\textbf{DIST2D}}] input blocks modify (overwrite) the value of
-a fault parameter by a heterogeneous spatial distribution:
-
-\&\textbf{DIST2D} shape='$\mathtt{square}$', val, xc, yc, zc, l /
-
-$\;$sets a constant value (val) within a cube with center (xc,yc,zc)
-and edge size l.
-
-\&\textbf{DIST2D} shape='$\mathtt{rectangle}$', val, xc, yc, zc,
-lx, ly, lz /
-
-$\;$sets a constant value (val) within a rectangular prism with center
-(xc,yc,zc) and edge sizes (lx,ly,lz).
-
-\&\textbf{DIST2D} shape='$\mathtt{rectangle-taper}$', val, valh,
-xc, yc, zc, lx, ly, lz /
-
-$\;$sets a vertical linear gradient within a rectangular prism with
-center (xc,yc,zc) and edge sizes (lx,ly,lz). Values vary linearly
-as a function of vertical position z between value val at z = zc-lz/2
-and value valh at z = zc+lz/2 .
-
-\&\textbf{DIST2D} shape='$\mathtt{circular}$', val, xc, yc, zc, r
-/
-
-$\;$sets a constant value (val) within a sphere with center (xc,yc,zc)
-and radius r. \\
-
-\end{lyxlist}
-\end{lyxlist}
-
-\begin{lyxlist}{00.00.0000}
-\item [{\textbf{DATA/FAULT\_STATIONS}}] Stations in the fault plane.
-
-\textbf{Line 1} : number of stations.
-
-\textbf{Line 2 to end}: 5 columns: X, Y, Z (-depth), station name,
-fault-id
-
-The fault-id identifies the fault that contains the station. It is
-the index of appearance in the faults list after line 2 of Par\_file\_faults
-
-\end{lyxlist}
-
-\begin{lyxlist}{00.00.0000}
-\item [{\textbf{DATA/input\_file.txt}}] Heterogeneous stresses and friction
-parameters are documented in page 10 of
-
-\begin{lyxlist}{00.00.0000}
-\item [{$\mathtt{EXAMPLES/tpv16/description/TPV16\_17\_Description\_v03.pdf}$}]~
-\end{lyxlist}
-
-To activate this feature, in fault\_solver\_dynamic.f90 set \textbf{TPV16}
-= .true. then re-compile the code.
-
-\end{lyxlist}
-
-\section{\label{sec:Setting-the-Kelvin-Voigt}Setting the Kelvin-Voigt Damping
-Parameter }
-
-The purpose of the Kelvin-Voigt viscosity in the dynamic fault solver
-is to damp spurious oscillations generated by the fault slip at frequencies
-that are too high to be resolved by the mesh. The viscosity $\mathtt{eta}$
-(in seconds) depends on the size of the elements on the fault. Here
-is how to set it:
-\begin{enumerate}
-\item Determine the average linear size of the elements on the fault plane,
-$\mathtt{h\_fault}$. Usually this value is prescribed by the user
-during mesh generation. Otherwise it can be found by inspection of
-the mesh inside the CUBIT GUI.
-\item Use the Matlab function $\mathtt{utils/critical\_timestep.m}$ to
-compute\\
- $\mathtt{dtc\_fault}=\mathtt{critical\_timestep\left(c_{p},h\_fault,ngll\right)}$.
-\\
-This is the critical time step in an elastic medium for a hypothetical
-element of cubic shape with size equal to $\mathtt{h\_fault}$.
-\item Set $\mathtt{eta}$ in $\mathtt{Par\_file\_faults}$ to (0.1 to 0.3)$\ensuremath{\times}\mathtt{dtc\_fault}$.
-A larger $\mathtt{eta}$ damps high-frequencies more aggressively
-but it might also affect lower frequencies and rupture speed.
-\end{enumerate}
-Viscosity reduces numerical stability: the critical timestep in a
-simulation with Kelvin-Voigt damping needs to be smaller than that
-in a purely elastic simulation. Here is how to set the time step accordingly:
-\begin{enumerate}
-\item Run a test simulation without viscosity ($\mathtt{eta}$=0 and only
-a few time steps)
-\item Look for the \textquotedbl{}maximum suggested time step\textquotedbl{}
-in \texttt{OUTPUT\_FILES/output\_mesher.txt}. This is the critical
-timestep of a purely elastic simulation, $\mathtt{dtc\_bulk}$.
-\item Reset the timestep of the simulation with a Kelvin-Voigt material
-to a value smaller than\\
- $\mathtt{dtc\_kv}=\mathtt{eta}\left(\sqrt{1+\mathtt{dtc\_bulk^{2}}/\mathtt{eta^{2}}}-1\right)$
-\end{enumerate}
-Note that in general $\mathtt{dtc\_bulk}$ is smaller than $\mathtt{dtc\_fault}$,
-because elements off the fault might be smaller or more distorted
-than element faces on the fault.
-
-\section{Output Files }
-
-Several output files are saved in \texttt{OUTPUT\_FILES/}:\\
-
-\begin{enumerate}
-\item Seismograms for each station on the fault plane given in \texttt{DATA/FAUL\_STATIONS}.
-One output file is generated for each station, named after the station.
-The files are ascii and start with a header (22 lines long) followed
-by a data block with the following format, one line per time sample:\\
-
-Column 1 = Time (s)
-
-Column 2 = horizontal right-lateral slip (m)
-
-Column 3 = horizontal right-lateral slip rate (m/s)
-
-Column 4 = horizontal right-lateral shear stress (MPa)
-
-Column 5 = vertical up-dip slip (m)
-
-Column 6 = vertical up-dip slip rate (m/s)
-
-Column 7 = vertical up-dip shear stress (MPa)
-
-Column 8 = normal stress (MPa)\\
-
-The stresses are relative to the footwall side of the fault (this
-convention controls their sign, but not their amplitude). Slip is
-defined as displacement of the hanging wall relative to the footwall.
-
-\item Seismograms at stations in the bulk (out of the fault plane) given
-in \texttt{DATA/STATIONS}.
-\item Rupture time files are named $\mathtt{Rupture\_time\_FAULT}$-$\mathtt{id}$.
-One file is generated for each fault. The files are ascii and start
-with a header (12 lines long) followed by a data block with the following
-format, one line per fault node:
-
-Column 1 = horizontal coordinate, distance along strike (m)
-
-Column 2 = vertical coordinate, distance down-dip (m)
-
-Column 3 = rupture time (s)
-
-\item Fault quantities (slip, slip rate, stresses, etc) at regular times
-are stored in binary data files called $\mathtt{Snapshot\#it\#.bin}$,
-where \#it\# is the timestep number. These can be read in Matlab with
-the function $\mathtt{utils/FSEM3D\_snapshot.m}$
-\end{enumerate}
-
-\section{Post-processing and Visualization }
-
-Some Matlab functions for post-processing and visualization are included
-in directory $\mathtt{utils}$. The functions are internally documented
-(see their matlab help).\\
-
-\texttt{FSEM3D\_snapshot} reads a fault data snapshot\\
-
-The directories \texttt{EXAMPLES/*/post} contain additional Matlab
-scripts to generate figures specific to each example.
-
-\chapter{\label{cha:Adjoint-Simulations}Adjoint Simulations}
-
-Adjoint simulations are generally performed for two distinct applications.
-First, they can be used in point source moment-tensor inversions, or source imaging for
-earthquakes with large ruptures such as the Lander's earthquake \citep{WaHe94}.
-Second, they can be used to generate finite-frequency sensitivity
-kernels that are a critical part of tomographic inversions based upon
-3D reference models \citep{TrTaLi05,LiTr06,TrKoLi08,LiTr08}. In either
-case, source parameter or velocity structure updates are sought to
-minimize a specific misfit function (e.g., waveform or traveltime
-differences), and the adjoint simulation provides a means of computing
-the gradient of the misfit function and further reducing it in successive
-iterations. Applications and procedures pertaining to source studies
-and finite-frequency kernels are discussed in Sections~\ref{sec:Adjoint-simulation-sources}
-and \ref{sec:Adjoint-simulation-finite}, respectively. The two related
-parameters in the \texttt{Par\_file} are \texttt{SIMULATION\_TYPE}
-(1, 2 or 3) and the \texttt{SAVE\_FORWARD} (boolean).
-
-\section{\label{sec:Adjoint-simulation-sources}Adjoint Simulations for Sources
-Only (not for the Model)}
-
-When a specific misfit function between data and synthetics is minimized to invert
-for earthquake source parameters, the gradient of the misfit function
-with respect to these source parameters can be computed by placing
-time-reversed seismograms at the receivers as virtual sources
-in an adjoint simulation. Then the value of the gradient is obtained
-from the adjoint seismograms recorded at the original earthquake location.
-\begin{enumerate}
-\item \textbf{Prepare the adjoint sources} \label{enu:Prepare-the-adjoint}
-
-\begin{enumerate}
-\item First, run a regular forward simlation (\texttt{SIMULATION\_TYPE =
-1} and \texttt{SAVE\_FORWARD = .false.}). You can automatically set
-these two variables using the \texttt{\small utils/change\_simulation\_type.pl}{\small{}
-script:}{\small \par}
-
-\begin{lyxcode}
-utils/change\_simulation\_type.pl~-f~
-\end{lyxcode}
-
-and then collect the recorded seismograms at all the stations given
-in \texttt{DATA/STATIONS}.
-
-\item Then select the stations for which you want to compute the time-reversed
-adjoint sources and run the adjoint simulation, and compile them into
-the \texttt{DATA/STATIONS\_ADJOINT} file, which has the same format
-as the regular \texttt{DATA/STATIONS} file.
-
-\begin{itemize}
-\item Depending on what type of misfit function is used for the source inversion,
-adjoint sources need to be computed from the original recorded seismograms
-for the selected stations and saved in a sub-directory called \texttt{SEM/}
-in the root directory of the code, with the format \texttt{NT.STA.BX?.adj}, where \texttt{NT},
-\texttt{STA} are the network code and station name given in the \texttt{DATA/STATIONS\_ADJOINT}
-file, and \texttt{BX?} represents the component name of a particular
-adjoint seismogram. Please note that the band code can change depending
-on your sampling rate (see Appendix~\ref{cha:channel-codes} for
-further details).
-\item The adjoint seismograms are in the same format as the original seismogram
-(\texttt{NT.STA.BX?.sem?}), with the same start time, time interval
-and record length.
-\end{itemize}
-\item Notice that even if you choose to time reverse only one component
-from one specific station, you still need to supply all three components
-because the code is expecting them (you can set the other two components
-to be zero).
-\item Also note that since time-reversal is done in the code itself, no
-explicit time-reversing is needed for the preparation of the adjoint
-sources, i.e., the adjoint sources are in the same forward time sense
-as the original recorded seismograms.
-\end{enumerate}
-\item \textbf{Set the related parameters and run the adjoint simulation}\\
- In the \texttt{DATA/Par\_file}, set the two related parameters to
-be \texttt{SIMULATION\_TYPE = 2} and \texttt{SAVE\_FORWARD = .false.}.
-More conveniently, use the scripts \texttt{utils/change\_simulation\_type.pl}
-to modify the \texttt{Par\_file} automatically (\texttt{change\_simulation\_type.pl
--a}). Then run the solver to launch the adjoint simulation.
-\item \textbf{Collect the seismograms at the original source location}
-
-After the adjoint simulation has completed successfully, collect the
-seismograms from \texttt{LOCAL\_PATH}.
-\begin{itemize}
-\item These adjoint seismograms are recorded at the locations of the original
-earthquake sources given by the \texttt{DATA/CMTSOLUTION} file, and
-have names of the form \texttt{NT.S?????.S??.sem} for the six-component
-strain tensor (\texttt{SNN,SEE,SZZ,SNE,SNZ,SEZ}) at these locations,
-and ~\\
- \texttt{NT.S?????.BX?.sem} for the three-component displacements
-(\texttt{BXN,BXE,BXZ}) recorded at these locations.
-\item \texttt{S?????} denotes the source number; for example, if the original
-\texttt{CMTSOLUTION} provides only a point source, then the seismograms
-collected will start with \texttt{S00001}.
-\item These adjoint seismograms provide critical information for the computation
-of the gradient of the misfit function.
-\end{itemize}
-\end{enumerate}
-
-\section{\label{sec:Adjoint-simulation-finite}Adjoint Simulations for Finite-Frequency
-Kernels (Kernel Simulation)}
-
-Finite-frequency sensitivity kernels are computed in two successive
-simulations (please refer to \citet{LiTr06} and \citet{TrKoLi08}
-for details).
-\begin{enumerate}
-\item \textbf{Run a forward simulation with the state variables saved at
-the end of the simulation}
-
-Prepare the \texttt{\small CMTSOLUTION}{\small{} and }\texttt{\small STATIONS}{\small{}
-files, set the parameters }\texttt{\small SIMULATION\_TYPE}{\small {}
-}\texttt{\small =}{\small {} }\texttt{\small 1}{\small{} and }\texttt{\small SAVE\_FORWARD
-=}{\small {} }\texttt{\small .true.}{\small{} in the }\texttt{\small Par\_file}{\small{}
-(}\texttt{\small change\_simulation\_type -F}{\small ), and run the
-solver.}{\small \par}
-\begin{itemize}
-\item Notice that attenuation is not implemented yet for the computation
-of finite-frequency kernels; therefore set \texttt{ATTENUATION = .false.}
-in the \texttt{Par\_file}.
-\item We also suggest you modify the half duration of the \texttt{CMTSOLUTION}
-to be similar to the accuracy of the simulation (see Equation \ref{eq:shortest_period})
-to avoid too much high-frequency noise in the forward wavefield, although
-theoretically the high-frequency noise should be eliminated when convolved
-with an adjoint wavefield with the proper frequency content.
-\item This forward simulation differs from the regular simulations (\texttt{\small SIMULATION\_TYPE}{\small {}
-}\texttt{\small =}{\small {} }\texttt{\small 1}{\small{} and }\texttt{\small SAVE\_FORWARD}{\small {}
-}\texttt{\small =}{\small {} }\texttt{\small .false.}{\small ) described
-in the previous chapters in that the state variables for the last
-time step of the simulation, including wavefields of the displacement,
-velocity, acceleration, etc., are saved to the }\texttt{\small LOCAL\_PATH}{\small{}
-to be used for the subsequent simulation. }{\small \par}
-\item For regional simulations, the files recording the absorbing boundary
-contribution are also written to the \texttt{LOCAL\_PATH} when \texttt{SAVE\_FORWARD
-= .true.}.
-\end{itemize}
-\item \textbf{Prepare the adjoint sources}
-
-The adjoint sources need to be prepared the same way as described
-in the Section \ref{enu:Prepare-the-adjoint}.
-\begin{itemize}
-\item In the case of travel-time finite-frequency kernel for one source-receiver
-pair, i.e., point source from the \texttt{CMTSOLUTION}, and one station
-in the \texttt{STATIONS\_ADJOINT} list, we supply a sample program
-in \texttt{utils/adjoint\_sources/traveltime/xcreate\_adjsrc\_traveltime}
-to cut a certain portion of the original displacement seismograms
-and convert it into the proper adjoint source to compute the finite-frequency
-kernel.
-
-\begin{lyxcode}
-xcreate\_adjsrc\_traveltime~t1~t2~ifile{[}0-5{]}~E/N/Z-ascii-files~{[}baz{]}
-\end{lyxcode}
-
-where \texttt{t1} and \texttt{t2} are the start and end time of the
-portion you are interested in, \texttt{ifile} denotes the component
-of the seismograms to be used (0 for all three components, 1 for East,
-2 for North, and 3 for vertical, 4 for transverse, and 5 for radial
-component), \texttt{E/N/Z-ascii-files} indicate the three-component
-displacement seismograms in the right order, and \texttt{baz} is the
-back-azimuth of the station. Note that \texttt{baz} is only supplied
-when \texttt{ifile} = 4 or 5.
-
-\item Similarly, a sample program to compute adjoint sources for amplitude
-finite-frequency kernels may be found in \texttt{utils/adjoint\_sources/amplitude}
-and used in the same way as described for traveltime measurements
-
-\begin{lyxcode}
-xcreate\_adjsrc\_amplitude~t1~t2~ifile{[}0-5{]}~E/N/Z-ascii-files~{[}baz{]}.
-\end{lyxcode}
-\end{itemize}
-\item \textbf{Run the kernel simulation}
-
-With the successful forward simulation and the adjoint source ready
-in the \texttt{SEM/} directory, set \texttt{SIMULATION\_TYPE
-= 3} and \texttt{SAVE\_FORWARD = .false.} in the \texttt{Par\_file}
-(you can use \texttt{change\_simulation\_type.pl -b}), and rerun the
-solver.
-\begin{itemize}
-\item The adjoint simulation is launched together with the back reconstruction
-of the original forward wavefield from the state variables saved from
-the previous forward simulation, and the finite-frequency kernels
-are computed by the interaction of the reconstructed forward wavefield
-and the adjoint wavefield.
-\item The back-reconstructed seismograms at the original station locations
-are saved to the \texttt{LOCAL\_PATH} at the end of the kernel simulations,
-and can be collected to the local disk.
-\item These back-constructed seismograms can be compared with the time-reversed
-original seismograms to assess the accuracy of the backward reconstruction,
-and they should match very well.
-\item The arrays for density, P-wave speed and S-wave speed kernels are
-also saved in the \texttt{LOCAL\_PATH} with the names \texttt{proc??????\_rho(alpha,beta)\_kernel.bin},
-where \texttt{proc??????} represents the processor number, \texttt{rho(alpha,beta)}
-are the different types of kernels.
-\end{itemize}
-\item \textbf{Run the anisotropic kernel simulation}
-
-Instead of the kernels for the isotropic wave speeds, you can also
-compute the kernels for the 21 independent components $C_{IJ},\, I,J=1,...,6$
-(using Voigt's notation) of the elastic tensor in the cartesian coordinate
-system. This is done by setting \texttt{ANISOTROPIC\_KL} \texttt{=}
-\texttt{.true.} in \texttt{constants.h} before compiling the package.
-The definition of the parameters $C_{IJ}$ in terms of the corresponding
-components $c_{ijkl},ijkl,i,j,k,l=1,2,3$ of the elastic tensor in
-cartesian coordinates follows \citet{ChTr07}. The 21 anisotropic
-kernels are saved in the \texttt{LOCAL\_PATH} in one file with the
-name of \texttt{proc??????\_cijkl\_kernel.bin} (with \texttt{proc??????}
-the processor number). The output kernels correspond to absolute perturbations
-$\delta C_{IJ}$ of the elastic parameters and their unit is in $s/GPa/km^{3}$.
-For consistency, the output density kernels with this option turned
-on are for a perturbation $\delta\rho$ (and not $\frac{\delta\rho}{\rho}$)
-and their unit is in s / (kg/m$^{3}$) / km$^{3}$. These `primary'
-anisotropic kernels can then be combined to obtain the kernels for
-different parameterizations of anisotropy. This can be done, for example,
-when combining the kernel files from slices into one mesh file (see
-Section~\ref{sec:Finite-Frequency-Kernels}).
-
-If \texttt{ANISOTROPIC\_KL} \texttt{=} \texttt{.true.} by additionally
-setting \texttt{ANISOTROPIC\_KL} \texttt{=} \texttt{.true.} in \texttt{constants.h}
-the package will save anisotropic kernels paramterized as velocities
-related to transverse isotropy based on the the Chen and Tromp parameters
-\citet{ChTr07}. The kernels are saved as relative perturbations for
-horizontal and vertical P and S velocities, $\alpha_{v},\alpha_{h},\beta_{v},\beta_{h}$.
-Explicit relations can be found in appendix B. of \citet{SiLiTrTr07b}
-
-The anisotropic kernels are only currently available for CPU mode.
-
-\end{enumerate}
-In general, the three steps need to be run sequentially to assure
-proper access to the necessary files. If the simulations are run through
-some cluster scheduling system (e.g., LSF), and the forward simulation
-and the subsequent kernel simulations cannot be assigned to the same
-set of computer nodes, the kernel simulation will not be able to access
-the database files saved by the forward simulation. Solutions for
-this dilemma are provided in Chapter~\ref{cha:Scheduler}. Visualization
-of the finite-frequency kernels is discussed in Section~\ref{sec:Finite-Frequency-Kernels}.
-
-%------------------------------------------------------------------------------------------------%
-
-\chapter{\label{cha:tomo}Doing tomography, i.e., updating the model based on the sensitivity kernels obtained}
-
-%------------------------------------------------------------------------------------------------%
-
-UNDER CONSTRUCTION (March 2015). Will be added around April / May 2015.
-
-\section{Tomographic inversion / imaging using the sensitivity kernels obtained}
-
-OLD VERSION OF THE SECTION, WILL SOON BE IMPROVED AND ENRICHED.\\
-
-One of the fundamental reasons for computing sensitivity kernels (Section~\ref{sec:Adjoint-simulation-finite})
-is to use them within a tomographic inversion. In other words, use
-recorded seismograms, make measurements with synthetic seismograms,
-and use the misfit between the two to iteratively improve the model
-described by (at least) $V_{{\rm p}}$, $V_{{\rm s}}$, and $\rho$
-as a function of space.
-
-Whatever misfit function you use for the tomographic inversion (you can find several
-examples in \citet{TrKoLi08} and \citet{TrTaLi05}), you will weight
-the sensitivity kernels with measurements. Furthermore, you will use
-as many measurements (stations, components, time windows) as possible
-per event; hence, we call these composite kernels ``event kernels,''
-which are volumetric fields representing the gradient of the misfit
-function with respect to one of the variables (\eg $V_{{\rm s}}$).
-The basic features of an adjoint-based tomographic inversion were
-illustrated in \citet{TaLiTr07} and \citet{TrKoLi08} using a conjugate-gradient
-algorithm; there are dozens of versions of gradient-based inversion
-algorithms that could alternatively be used. The tomographic inversion
-of \citet{TaLiMaTr09,TaLiMaTr2010} used SPECFEM3D Cartesian as well
-as several additional components which are also stored on the CIG
-svn server, described next.
-
-The directory containing external utilities for tomographic inversion using
-SPECFEM3D Cartesian (or other packages that evaluate misfit functions
-and gradients) is in directory \texttt{utils/ADJOINT\_TOMOGRAPHY\_TOOLS/}:
-\begin{verbatim}
-flexwin/     -- FLEXWIN algorithm for automated picking of time windows
-measure_adj/ -- reads FLEXWIN output file and makes measurements,
-                with the option for computing adjoint sources
-iterate_adj/ -- various tools for iterative inversion
-                (requires pre-computed "event kernels")
-\end{verbatim}
-This directory also contains
-a brief \verb+README+ file indicating the role of the three subdirectories,
-\verb+flexwin+ \citep{Maggi2009}, \verb+measure_adj+, and \verb+iterate_adj+.
-The components for making the model update are there; however, there
-are no explicit rules for computing the model update, just as with
-any optimization problem. There are options for computing a conjugate
-gradient step, as well as a source subspace projection step.
-
-The best single file to read is probably: \verb+ADJOINT_TOMO/iterate_adj/cluster/README+.
-
-%------------------------------------------------------------------------------------------------%
-
 \chapter{Noise Cross-correlation Simulations}
-
-%------------------------------------------------------------------------------------------------%
 
@@
 which should contain enough explanations for all the commands.
-
-%------------------------------------------------------------------------------------------------%
-
-\chapter{Gravity integral calculations for the gravity field of an Earth model}
-
-%------------------------------------------------------------------------------------------------%
-
-SPECFEM can now compute the gravity field as well as its derivatives (i.e., gravity gradiometry)
-generated by any given 3D Earth model at a given elevation, i.e., on a given observation surface.
-That feature is still experimental but should work just fine.\\
-
-To see how it is implemented and to use it, type this in the root directory of the code:\\
-
-\texttt{grep -i GRAVITY\_INTEGRALS src/*/*}
-
-%------------------------------------------------------------------------------------------------%
-
-\chapter{Graphics}
-
-%------------------------------------------------------------------------------------------------%
-
-\section{\label{sec:Mesh-graphics}Meshes}
-
-In case you used the internal mesher \texttt{xmeshfem3D} to create
-and partition your mesh, you can output mesh files in ABAQUS (.INP)
-and DX (.dx) format to visualize them. For this, you must set either
-the flag \texttt{CREATE\_DX\_FILES} or \texttt{CREATE\_ABAQUS\_FILES}
-to \texttt{.true.} in the mesher's parameter file \texttt{Mesh\_Par\_file}
-prior to running the mesher (see Chapter~\ref{cha:Running-the-Mesher-Meshfem3D}
-for details). You can then use AVS \urlwithparentheses{www.avs.com}
-or OpenDX \urlwithparentheses{www.opendx.org} to visualize the mesh
-and MPI partition (slices).
-
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=0.49\textwidth]{figures/vtk_mesh_vp} \includegraphics[width=0.49\textwidth]{figures/vtk_mesh_vs}
-\par\end{centering}
-
-\caption{Visualization using Paraview of VTK files created by \texttt{xgenerate\_databases}
-showing P- and S-wave velocities assigned to the mesh points. The
-mesh was created by \texttt{xmeshfem3D} for 4 processors.}
-
-\label{fig:vtk.mesh}
-\end{figure}
-
-You have also the option to visualize the distributed databases produced
-by \texttt{xgenerate\_databases} using Paraview \urlwithparentheses{www.paraview.org}.
-For this, you must set the flag \texttt{SAVE\_MESH\_FILES} to \texttt{.true.}
-in the main parameter file \texttt{Par\_file} (see Chapter~\ref{cha:Main-Parameter}
-for details). This will create VTK files for each single partition.
-You can then use Paraview \urlwithparentheses{www.paraview.org}
-to visualized these partitions.
-
-\section{\label{sec:Movies}Movies}
-
-To make a surface or volume movie of the simulation, set parameters
-\texttt{MOVIE\_SURFACE}, \texttt{MOVIE\_VOLUME}, \texttt{MOVIE\_TYPE},
-and \texttt{NTSTEP\_BETWEEN\_FRAMES} in the \texttt{Par\_file}. Turning
-on the movie flags, in particular \texttt{MOVIE\_VOLUME}, produces
-large output files. \texttt{MOVIE\_VOLUME} files are saved in the
-\texttt{LOCAL\_PATH} directory, whereas \texttt{MOVIE\_SURFACE} output
-files are saved in the \texttt{OUTPUT\_FILES} directory. We save the
-displacement field if the parameter \texttt{SAVE\_DISPLACEMENT} is
-set, otherwise the velocity field is saved. The look of a movie is
-determined by the half-duration of the source. The half-duration should
-be large enough so that the movie does not contain frequencies that
-are not resolved by the mesh, i.e., it should not contain numerical
-noise. This can be accomplished by selecting a CMT \texttt{HALF\_DURATION}
-> 1.1 $\times$ smallest period (see figure \ref{fig:CMTSOLUTION-file}).
-When \texttt{MOVIE\_SURFACE} = .\texttt{true.}, the half duration
-of each source in the \texttt{CMTSOLUTION} file is replaced by
-\begin{quote}
-\[
-\sqrt{(}\mathrm{\mathtt{HALF\_DURATIO}\mathtt{N}^{2}}+\mathrm{\mathtt{HDUR\_MOVI}\mathtt{E}^{2}})
-\]
- \textbf{NOTE:} If \texttt{HDUR\_MOVIE} is set to 0.0, the code will
-select the appropriate value of 1.1 $\times$ smallest period. As
-usual, for a point source one can set \texttt{half duration} in the
-\texttt{CMTSOLUTION} file to be 0.0 and \texttt{HDUR\_MOVIE} = 0.0
-in the \texttt{Par\_file} to get the highest frequencies resolved
-by the simulation, but for a finite source one would keep all the
-\texttt{half durations} as prescribed by the finite source model and
-set \texttt{HDUR\_MOVIE} = 0.0.
-\end{quote}
-
-\subsection{Movie Surface}
-
-When running \texttt{xspecfem3D} with the \texttt{MOVIE\_SURFACE}
-flag turned on, the code outputs \texttt{moviedata??????} files in
-the \texttt{OUTPUT\_FILES} directory. There are several flags in the
-main parameter file \texttt{Par\_file} that control the output of
-these movie data files (see section \ref{cha:Main-Parameter} for
-details): \texttt{\small NTSTEP\_BETWEEN\_FRAMES}{\small{} to set the
-timesteps between frames, }\texttt{\small SAVE\_DISPLACEMENT}{\small{}
-to save displacement instead of velocity, }\texttt{\small USE\_HIGHRES\_FOR\_MOVIES}{\small{}
-to save values at all GLL point instead of element edges. In order
-to output additionally shakemaps, you would set the parameter }\texttt{\small CREATE\_SHAKEMAP}{\small{}
-to }\texttt{\small .true.}{\small .}{\small \par}
-
-The files are in a fairly complicated binary format, but there is
-a program provided to convert the output into more user friendly formats:
-\begin{description}
-\item [{\texttt{xcreate\_movie\_shakemap\_AVS\_DX\_GMT}}] From \texttt{create\_movie\_shakemap\_AVS\_DX\_GMT.f90},
-it outputs data in ASCII, OpenDX, or AVS format (also readable in
-ParaView). Before compiling the code, make sure you have the file
-\texttt{surface\_from\_mesher.h} in the \texttt{OUTPUT\_FILES/} directory.
-This file will be created by the solver run. Then type
-
-\begin{lyxcode}
-{\footnotesize make~xcreate\_movie\_shakemap\_AVS\_DX\_GMT}{\footnotesize \par}
-\end{lyxcode}
-
-and run the executable \texttt{xcreate\_movie\_shakemap\_AVS\_DX\_GMT}
-in the main directory. It will create visualization files
-in your format of choice. The code will prompt the user for input
-parameters.
-
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=0.32\textwidth]{figures/movie_surf_1} \includegraphics[width=0.32\textwidth]{figures/movie_surf_2}
-\includegraphics[width=0.32\textwidth]{figures/movie_surf_3}
-\par\end{centering}
-
-\caption{Visualization using AVS files created by \texttt{xcreate\_movie\_shakemap\_AVS\_DX\_GMT}
-showing movie snapshots of vertical velocity components at different
-times.}
-
-\label{fig:movie.surf}
-\end{figure}
-
-\end{description}
-The \texttt{SPECFEM3D Cartesian} code is running in near real-time
-to produce animations of southern California earthquakes via the web;
-see Southern California ShakeMovie\textregistered{}\urlwithparentheses{www.shakemovie.caltech.edu}.
-
-\subsection{\label{sub:Movie-Volume}Movie Volume}
-
-When running xspecfem3D with the \texttt{\small MOVIE\_VOLUME}{\small{}
-flag turned on, the code outputs several files in }\texttt{\small LOCAL\_PATH}{\small{}
-specified in the main }\texttt{\small Par\_file}{\small , e.g. in
-directory }\texttt{\small OUTPUT\_FILES/DATABASES\_MPI}{\small .
-The output is saved by each processor at the time interval specified
-by }\texttt{\small NTSTEP\_BETWEEN\_FRAMES}{\small . For all domains,
-the velocity field is output to files: }\texttt{\small proc??????\_velocity\_X\_it??????.bin}{\small ,
-}\texttt{\small proc??????\_velocity\_Y\_it??????.bin}{\small{} and
-}\texttt{\small proc??????\_velocity\_Z\_it??????.bin}{\small . For
-elastic domains, the divergence and curl taken from the velocity field,
-i.e. $\nabla\cdot{\bf {v}}$ and $\nabla\times{\bf {v}}$, get stored
-as well: }\texttt{\small proc??????\_div\_it??????.bin}{\small , }\texttt{\small proc??????\_curl\_X\_t??????.bin}{\small ,
-}\texttt{\small proc??????\_curl\_Y\_it??????.bin}{\small{} and }\texttt{\small proc??????\_curl\_Z\_it??????.bin}{\small .
-The files denoted }\texttt{\small proc??????\_div\_}{\small ~}\\
-{\small{} }\texttt{\small glob\_it??????.bin}{\small{} and }\texttt{\small proc??????\_curl\_glob\_it??????.bin}{\small{}
-are stored on the global points only, all the other arrays are stored
-on all GLL points. Note that the components X/Y/Z can change to E/N/Z
-according to the }\texttt{\small SUPPRESS\_UTM\_PROJECTION}{\small{}
-flag (see also Appendix~\ref{cha:Coordinates} and \ref{cha:channel-codes}).}{\small \par}
-
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=0.32\textwidth]{figures/movie_volume_1} \includegraphics[width=0.32\textwidth]{figures/movie_volume_2}
-\includegraphics[width=0.32\textwidth]{figures/movie_volume_3}
-\par\end{centering}
-
-\caption{Paraview visualization using movie volume files (converted by \texttt{xcombine\_vol\_data}
-and \texttt{mesh2vtu.pl}) and showing snapshots of vertical velocity
-components at different times.}
-
-\label{fig:movie.volume}
-\end{figure}
-
-To visualize these files, we use an auxilliary program \texttt{combine\_vol\_data.f90}
-to combine the data from all slices into one mesh file. To compile
-it in the root directory, type:
-\begin{lyxcode}
-{\footnotesize make~xcombine\_vol\_data~}{\footnotesize \par}
-\end{lyxcode}
-which will create the executable \texttt{bin/xcombine\_vol\_data}.
-To output the usage of this executable, type
-\begin{lyxcode}
-{\footnotesize ./bin/xcombine\_vol\_data~}{\footnotesize \par}
-\end{lyxcode}
-without arguments.
-
-\medskip
-
-\texttt{xcombine\_vol\_data} will combine the data on the different processors (located in \texttt{OUTPUT\_FILES/DATABASES\_MPI}), for a given quantity and a given iteration, to one file.
-For example, if you want to combine \texttt{velocity\_Z}, for the iteration 400 on 4 processors, i.e.
-if you want to combine these files :
-
-\medskip
-
-\texttt{proc000000\_velocity\_Z\_it000400.bin }
-
-\texttt{proc000001\_velocity\_Z\_it000400.bin }
-
-\texttt{proc000002\_velocity\_Z\_it000400.bin }
-
-\texttt{proc000003\_velocity\_Z\_it000400.bin}
-
-\medskip
-
-You have to go in the directory of the concerned example (where are the directories \texttt{DATA}, \texttt{OUTPUT\_FILES}, etc.).
-Then, you can launch \texttt{xcombine\_vol\_data}, specifying the path where it is located.
-As an example, if the \texttt{DATA} and \texttt{OUTPUT\_FILES} directories of the concerned example are in the root specfem3d directory, \texttt{xcombine\_vol\_data} is in \texttt{./bin}, and you have to type:
-
-\begin{lyxcode}
-~{\footnotesize ./bin/xcombine\_vol\_data~0~3~velocity\_Z\_it000400~./OUTPUT\_FILES/DATABASES\_MPI~./OUTPUT\_FILES~0~}{\footnotesize \par}
-\end{lyxcode}
-
-Here, $0$ is the number of the first processor, $3$ the number of the last one, \texttt{velocity\_Z\_it000400} the name of the files we want to combine without the prefix \texttt{proc000000*\_}, \texttt{./OUTPUT\_FILES/DATABASES\_MPI} the directory where the files are located, \texttt{./OUTPUT\_FILES} the directory where the combined file will be stored, and $0$ is the parameter to create a low-resolution mesh file ($1$ for a high-resolution).
-
-\medskip
-
-At the beginning of the source file \texttt{combine\_vol\_data.f90} (located in \texttt{src/auxiliaries/}), there is a logical constant parameter \texttt{USE\_VTK\_OUTPUT} whose default value is \texttt{.true.}, that which will provide the output combined file directly in the VTK format and you can use Paraview to visualize it.
-
-\medskip
-
-If the value of this parameter is \texttt{.false.}, the output mesh file will have the name \texttt{velocity\_Z\_it000400.mesh}. Then we next have to convert the
-\texttt{.mesh} file into the VTU (Unstructured grid file) format which can be viewed in ParaView. For this task, you can use and modify the
-script \texttt{mesh2vtu.pl} located in directory \texttt{utils/Visualization/Paraview/}, for example:
-\begin{lyxcode}
-{\footnotesize mesh2vtu.pl~-i~velocity\_Z\_it000400.mesh~-o~velocity\_Z\_it000400.vtu}{\footnotesize \par}
-\end{lyxcode}
-Notice that this Perl script uses a program \texttt{mesh2vtu} in the \texttt{utils/Visualization/Paraview/mesh2vtu} directory, which further
-uses the VTK \urlwithparentheses{http://www.vtk.org/} run-time library for its execution. Therefore, make sure you have them properly set
-in the script according to your system.
-
-\bigskip
-
-Then, to do a movie with several iterations, you have to repeat this process for each iteration you want to put in your movie.
-
-\section{\label{sec:Finite-Frequency-Kernels}Finite-Frequency Kernels}
-
-The finite-frequency kernels computed as explained in Section \ref{sec:Adjoint-simulation-finite}
-are saved in the \texttt{LOCAL\_PATH} at the end of the simulation.
-Therefore, we first need to collect these files on the front end,
-combine them into one mesh file, and visualize them with some auxilliary
-programs.
-\begin{enumerate}
-\item \textbf{Create slice files}
-
-We will only discuss the case of one source-receiver pair, i.e., the
-so-called banana-doughnut kernels. Although it is possible to collect
-the kernel files from all slices on the front end, it usually takes
-up too much storage space (at least tens of gigabytes). Since the
-sensitivity kernels are the strongest along the source-receiver great
-circle path, it is sufficient to collect only the slices that are
-along or close to the great circle path.
-
-A Perl script \texttt{slice\_number.pl} located in directory \texttt{utils/Visualization/Paraview/}
-can help to figure out the slice numbers that lie along the great
-circle path. It applies to meshes created with the internal mesher
-\texttt{xmeshfem3D}.
-\begin{enumerate}
-\item On machines where you have access to the script, copy the \texttt{Mesh\_Par\_file},
-and \texttt{output\_solver} files, and run:
-
-\begin{lyxcode}
-{\small slice\_number.pl~Mesh\_Par\_file~output\_solver.txt~slice\_file}{\small \par}
-\end{lyxcode}
-
-which will generate a \texttt{slices\_file}.
-
-\item For cases with multiple sources and multiple receivers, you need to
-provide a slice file before proceeding to the next step.
-\end{enumerate}
-\item \textbf{Collect the kernel files}
-
-After obtaining the slice files, you can collect the corresponding
-kernel files from the given slices.
-\begin{enumerate}
-\item You can use or modify the script \texttt{utils/copy\_basin\_database.pl}
-to accomplish this:
-
-\begin{lyxcode}
-{\small utils/copy\_database.pl~slice\_file~lsf\_machine\_file~filename~{[}jobid{]}}{\small \par}
-\end{lyxcode}
-
-where \texttt{\small lsf\_machine\_file}{\small{} is the machine file
-generated by the LSF scheduler, }\texttt{\small filename}{\small{}
-is the kernel name (e.g., }\texttt{\small rho\_kernel}{\small , }\texttt{\small alpha\_kernel}{\small{}
-and }\texttt{\small beta\_kernel}{\small ), and the optional }\texttt{\small jobid}{\small{}
-is the name of the subdirectory under }\texttt{\small LOCAL\_PATH}{\small{}
-where all the kernel files are stored.}{\small \par}
-
-\item After executing this script, all the necessary mesh topology files
-as well as the kernel array files are collected to the local directory
-of the front end.
-\end{enumerate}
-\item \textbf{Combine kernel files into one mesh file}
-
-We use an auxilliary program \texttt{combine\_vol\_data.f90} to combine
-the kernel files from all slices into one mesh file.
-\begin{enumerate}
-\item Compile it in the root directory:
-
-\begin{lyxcode}
-{\footnotesize make~xcombine\_vol\_data~}{\footnotesize \par}
-
-{\footnotesize ./bin/xcombine\_vol\_data~slice\_list~filename~input\_dir~output\_dir~high/low-resolution~}{\footnotesize \par}
-\end{lyxcode}
-
-where \texttt{input\_dir} is the directory where all the individual
-kernel files are stored, and \texttt{output\_dir} is where the mesh
-file will be written.
-
-\item Use 1 for a high-resolution mesh, outputting all the GLL points to
-the mesh file, or use 0 for low resolution, outputting only the corner
-points of the elements to the mesh file.
-\item The output mesh file will have the name \texttt{filename\_rho(alpha,beta).mesh}
-\end{enumerate}
-\item \textbf{Convert mesh files into .vtu files}
-
-\begin{enumerate}
-\item We next convert the \texttt{.mesh} file into the VTU (Unstructured
-grid file) format which can be viewed in ParaView. For this task,
-you can use and modify the script \texttt{mesh2vtu.pl} located in
-directory \texttt{utils/Visualization/Paraview/}, for example:
-
-\begin{lyxcode}
-{\footnotesize mesh2vtu.pl~-i~file.mesh~-o~file.vtu~}{\footnotesize \par}
-\end{lyxcode}
-\item Notice that this Perl script uses a program \texttt{mesh2vtu} in the
-\texttt{utils/Visualization/Paraview/mesh2vtu} directory, which further
-uses the VTK \urlwithparentheses{http://www.vtk.org/} run-time library
-for its execution. Therefore, make sure you have them properly set
-in the script according to your system.
-\end{enumerate}
-\item \textbf{Copy over the source and receiver .vtk file}
-
-In the case of a single source and a single receiver, the simulation
-also generates the file \texttt{sr.vtk} located in the \texttt{OUTPUT\_FILES/}
-directory to describe the source and receiver locations, which can
-also be viewed in Paraview in the next step.
-
-\item \textbf{View the mesh in ParaView}
-
-Finally, we can view the mesh in ParaView \urlwithparentheses{www.paraview.org}.
-\begin{enumerate}
-\item Open ParaView.
-\item From the top menu, \textsf{File} $\rightarrow$\textsf{Open data},
-select \texttt{file.vtu}, and click the \textsf{Accept} button.
-
-\begin{itemize}
-\item If the mesh file is of moderate size, it shows up on the screen; otherwise,
-only the bounding box is shown.
-\end{itemize}
-\item Click \textsf{Display Tab} $\rightarrow$ \textsf{Display Style} $\rightarrow$
-\textsf{Representation} and select \textsf{wireframe of surface} to
-display it.
-\item To create a cross-section of the volumetric mesh, choose \textsf{Filter}
-$\rightarrow$ \textsf{cut}, and under \textsf{Parameters Tab}, choose
-\textsf{Cut Function} $\rightarrow$ \textsf{plane}.
-\item Fill in center and normal information given by the \texttt{global\_slice\_number.pl}
-script (either from the standard output or from \texttt{normal\_plane.txt}
-file).
-\item To change the color scale, go to \textsf{Display Tab} $\rightarrow$
-\textsf{Color} $\rightarrow$ \textsf{Edit Color Map} and reselect
-lower and upper limits, or change the color scheme.
-\item Now load in the source and receiver location file by \textsf{File}
-$\rightarrow$ \textsf{Open data}, select \texttt{sr.vt}k, and click
-the \textsf{Accept} button. Choose \textsf{Filter} $\rightarrow$
-\textsf{Glyph}, and represent the points by `\textsf{spheres}'.
-\item For more information about ParaView, see the ParaView Users Guide
-\urlwithparentheses{www.paraview.org/files/v1.6/ParaViewUsersGuide.PDF}.
-\end{enumerate}
-\end{enumerate}
-\begin{figure}[H]
-\noindent \begin{centering}
-\includegraphics[scale=0.6]{figures/3D-S-Kernel}
-\par\end{centering}
-
-\caption{(a) Top Panel: Vertical source-receiver cross-section of the S-wave
-finite-frequency sensitivity kernel $K_{\beta}$ for station GSC at
-an epicentral distance of 176 km from the September 3, 2002, Yorba
-Linda earthquake. Lower Panel: Vertical source-receiver cross-section
-of the 3D S-wave speed model used for the spectral-element simulations
-\citep{KoLiTrSuStSh04}. (b) The same as (a) but for station HEC at
-an epicentral distance of 165 km \citep{LiTr06}.}
-
-\label{figure:P-wave-speed-finite-frequency}
-\end{figure}
-
-\chapter{\label{cha:Scheduler}Running through a Scheduler}
-
-The code is usually run on large parallel machines, often PC clusters,
-most of which use schedulers, i.e., queuing or batch management systems
-to manage the running of jobs from a large number of users. The following
-considerations need to be taken into account when running on a system
-that uses a scheduler:
-\begin{itemize}
-\item The processors/nodes to be used for each run are assigned dynamically
-by the scheduler, based on availability. Therefore, in order for the
-\texttt{xgenerate\_databases} and the \texttt{xspecfem3D} executables
-(or between successive runs of the solver) to have access to the same
-database files (if they are stored on hard drives local to the nodes
-on which the code is run), they must be launched in sequence as a
-single job.
-\item On some systems, the nodes to which running jobs are assigned are
-not configured for compilation. It may therefore be necessary to pre-compile
-both the \texttt{xgenerate\_databases} and the \texttt{xspecfem3D}
-executables.
-\item One feature of schedulers/queuing systems is that they allow submission
-of multiple jobs in a ``launch and forget'' mode. In order to take
-advantage of this property, care needs to be taken that output and
-intermediate files from separate jobs do not overwrite each other,
-or otherwise interfere with other running jobs.
-\end{itemize}
-Examples of job scripts can be found in the \texttt{\small utils/Cluster/}{\small{}
-directory and can straightforwardly be modified and adapted to meet
-more specific running needs.}{\small \par}
-
-\chapter{{\normalsize \label{cha:-Changing-the}} Changing the Model}
-
-In this section we explain how to change the velocity model used for
-your simulations. These changes involve contributing specific subroutines
-that replace existing subroutines in the \texttt{SPECFEM3D Cartesian}
-package. Note that \texttt{SPECFEM3D Cartesian} can handle Earth models
-with material properties that vary within each spectral element.
-
-\section{{\normalsize \label{sec:Using-tomographic}}Using external tomographic Earth models}
-
-To implement your own external tomographic model(s), you must provide
-your own external tomography file(s), and choose between two possible
-options:\\
- \indent (1) set in the \texttt{Par\_file} the model parameter \texttt{MODEL
-= tomo}, \\
- or for more user control: \\
- \indent (2) set in the \texttt{Par\_file} the model parameter \texttt{MODEL
-= default}, define the negative \texttt{material\_ID} identifier for
-each element in the file \texttt{MESH/materials\_file} and use the
-following format in the file \texttt{MESH/nummaterial\_velocity\_file}
-when using CUBIT to construct your mesh (see also section \ref{subsec:Exporting-the-Mesh}):
-\begin{lyxcode}
-domain\_ID~material\_ID~tomography~elastic~file\_name~1
-\end{lyxcode}
-where: \\
- \indent \texttt{domain\_ID} is 1 for acoustic or 2 for elastic materials,
-\\
- \indent \texttt{material\_ID} a negative, unique identifier (i.e.,
--1,-2,...), \\
- \indent \texttt{tomography} keyword for tomographic material definition,
-\\
- \indent \texttt{elastic} keyword for elastic material definition,
-\\
- \indent \texttt{file\_name} the name of the tomography file and
-1 a positive unique identifier.\\
-
-The external tomographic model is represented by a grid of points
-with assigned material properties and homogeneous resolution along
-each spatial direction x, y and z. The ASCII file \texttt{file\_name}
-that describe the tomography should be located in the \texttt{TOMOGRAPHY\_PATH}
-directory, set in the \texttt{Par\_file}. The format of the file,
-as read from \texttt{model\_tomography.f90} located in the \texttt{src/generate\_databases}
-directory, looks like Figure \ref{fig:tomography_file}, where
-\begin{description}
-\item [{\texttt{ORIG\_X}, \texttt{END\_X}}] are, respectively, the coordinates
-of the initial and final tomographic grid points along the x direction
-(in the mesh units, e.g., $m$);
-\item [{\texttt{ORIG\_Y}, \texttt{END\_Y}}] respectively the coordinates
-of the initial and final tomographic grid points along the y direction
-(in the mesh units, e.g., $m$);
-\item [{\texttt{ORIG\_Z}, \texttt{END\_Z}}] respectively the coordinates
-of the initial and final tomographic grid points along the z direction
-(in the mesh units, e.g., $m$);
-\item [{\texttt{SPACING\_X}, \texttt{SPACING\_Y}, \texttt{SPACING\_Z}}] the
-spacing between the tomographic grid points along the x, y and z directions,
-respectively (in the mesh units, e.g., $m$);
-\item [{\texttt{NX}, \texttt{NY}, \texttt{NZ}}] the number of grid points
-along the spatial directions x, y and z, respectively; \texttt{NX}
-is given by {[}(\texttt{END\_X} - \texttt{ORIG\_X})/\texttt{SPACING\_X}{]}+1;
-\texttt{NY} and \texttt{NZ} are the same as \texttt{NX}, but for the
-y and z directions, respectively;
-\item [{\texttt{VP\_MIN}, \texttt{VP\_MAX}, \texttt{VS\_MIN}, \texttt{VS\_MAX}, \texttt{RHO\_MIN}, \texttt{RHO\_MAX}}] the
-minimum and maximum values of the wave speed \texttt{vp} and \texttt{vs}
-(in $m\, s^{-1}$) and of the density \texttt{rho} (in $kg\, m^{-3}$);
-these values could be the actual limits of the tomographic parameters
-in the grid or the minimum and maximum values to which we force the
-cut of velocity and density in the model.
-\end{description}
-After the first four lines, in the file \texttt{file\_name} the tomographic
-grid points are listed with the corresponding values of \texttt{vp},
-\texttt{vs} and \texttt{rho}, scanning the grid along the x coordinate
-(from \texttt{ORIG\_X} to \texttt{END\_X} with step of \texttt{SPACING\_X})
-for each given y (from \texttt{ORIG\_Y} to \texttt{END\_Y}, with step
-of \texttt{SPACING\_Y}) and each given z (from \texttt{ORIG\_Z} to
-\texttt{END\_Z}, with step of \texttt{SPACING\_Z}). %fig
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=1\textwidth]{figures/tomo_file}
-\par\end{centering}
-
-\caption{Tomography file \texttt{file\_name} that describes an external Earth
-model. The coordinates x, y and z of the grid, their limits ORIG\_X,
-ORIG\_Y, ORIG\_Z, END\_X, END\_Y, END\_Z and the grid spacings SPACING\_X,
-SPACING\_Y, SPACING\_Z should be in the units of the constructed mesh
-(e.g., UTM coordinates in meters).}
-
-\label{fig:tomography_file}
-\end{figure}
-
-The user can implement his own interpolation algorithm for the tomography
-model by changing the routine \texttt{model\_tomography.f90} located
-in the \texttt{src/generate\_databases/} directory. Moreover, for
-models that involve both fully defined materials and a tomography
-description, the \texttt{nummaterial\_velocity\_file} has multiple
-lines each with the corresponding suitable format described above.
-
-%==============
-
-\subsubsection*{Example: External tomography file with variable discretization intervals}
-
-The example in Figure~\ref{fig:tomography_file} is for an external tomography file with uniform spacing in the $x$, $y$, and $z$ directions.
-(Note that $\Delta x$, $\Delta y$, and $\Delta z$ need not be the same as in the example.) In that example, the  \texttt{nummaterial\_velocity\_file} is
-%
-\begin{lyxcode}
-2  -1 tomography elastic tomography\_model.xyz 1
-\end{lyxcode}
-%
-and the file \texttt{tomography\_model.xyz} will need to reside in \texttt{DATA/}.
-All entries in the second column of \texttt{materials\_file} will be \texttt{-1}, which means that each element in the mesh will be interpolated accordind to the values in \texttt{tomography\_model.xyz}.
-
-In some cases it may be desirable to have an external tomography model that is described in more than one file.
-For example, in cases like southern California, the length scale of variation in the structure of the wave speed model is much shorter in the sedimentary basin models within the upper 15~km.
-Therefore one might want to use an external tomography file that is sampled with, say, $\Delta x = 1000$~m, $\Delta y = 1000$~m, and $\Delta z = 250$~m in the uppermost 15~km, and then use $\Delta x = 2000$~m, $\Delta y = 2000$~m, and $\Delta z = 1000$~m below a depth of 15~km.
-If these intervals are chosen appropriately, then it will result in a pair of external tomography file that is much smaller than the alternative of having a single file with the fine discretization.
-In this case  \texttt{nummaterial\_velocity\_file} is
-%
-\begin{lyxcode}
-2  -1 tomography elastic file\_above\_15km.xyz 1 \\
-2  -2 tomography elastic file\_below\_15km.xyz 1
-\end{lyxcode}
-%
-and the files \texttt{file\_above\_15km.xyz} and \texttt{file\_below\_15km.xyz} will need to reside in \texttt{DATA/}.
-All entries in the second column of \texttt{materials\_file} will need to be either \texttt{-1} or \texttt{-2}, depending on whether the element is above or below 15~km depth.
-In this sense, the construction of the mesh and the discretization of the external model will generally be done in tandem.
-
-Other more complicated discretizations can be used following the same procedure. (In particular, there is no limit on the number of different external files that are used.)
-
-%=============
-
-\section{{\normalsize \label{sec:Anelastic-Models}}External (an)elastic Models}
-
-To use your own external model, you can set in the \texttt{Par\_file}
-the model parameter \texttt{MODEL = external}. Three-dimensional acoustic
-and/or (an)elastic (attenuation) models may then be superimposed onto
-the mesh based upon your subroutine in \texttt{model\_external\_values.f90}
-located in directory \texttt{src/generate\_databases}. The call to
-this routine would be as follows
-\begin{lyxcode}
-call~model\_external\_values(xmesh,~ymesh,~zmesh,~~\&~~\\
-~~~~~~~~~~~rho,~vp,~vs,~qmu\_atten,~iflag\_aniso,~idomain\_id)
-\end{lyxcode}
-Input to this routine consists of:
-\begin{description}
-\item [{\texttt{xmesh,ymesh,zmesh}}] location of mesh point
-\end{description}
-Output to this routine consists of:
-\begin{description}
-\item [{\texttt{rho,vp,vs}}] isotropic model parameters for density $\rho$
-($kg/m^{3}$), $v_{p}$ ($m/s$) and $v_{s}$ ($m/s$)
-\item [{\texttt{qmu\_atten}}] Shear wave quality factor: $0<Q_{\mu}<9000$
-\item [{\texttt{iflag\_aniso}}] anisotropic model flag, $0$ indicating
-no anisotropy or $1$ using anisotropic model parameters as defined
-in routine file \texttt{model\_aniso.f90}
-\item [{\texttt{idomain\_id}}] domain identifier, $1$ for acoustic, $2$
-for elastic, $3$ for poroelastic materials.
-\end{description}
-Note that the resolution and maximum value of anelastic models are
-truncated. This speeds the construction of the standard linear solids
-during the meshing stage. To change the resolution, currently at one
-significant figure following the decimal, or the maximum value (9000),
-consult \texttt{shared/constants.h}.
-
-\section{{\normalsize \label{sec:Anisotropic-Models}}Anisotropic Models}
-
-To use your anisotropic models, you can either set in the \texttt{Par\_file}
-the model parameter \texttt{ANISOTROPY} to \texttt{.true.} or set
-the model parameter \texttt{MODEL = aniso}. Three-dimensional anisotropic
-models may be superimposed on the mesh based upon the subroutine in
-file \texttt{model\_aniso.f90} located in directory \texttt{src/generate\_databases}.
-The call to this subroutine is of the form
-\begin{lyxcode}
-call~model\_aniso(iflag\_aniso,~rho,~vp,~vs,~\&~~\\
-~~~~~~~~~~~c11,c12,c13,c14,c15,c16,c22,c23,c24,c25,c26,~\&~~\\
-~~~~~~~~~~~c33,c34,c35,c36,c44,c45,c46,c55,c56,c66)~
-\end{lyxcode}
-Input to this routine consists of:
-\begin{description}
-\item [{\texttt{iflag\_aniso}}] flag indicating the type of the anisotropic
-model, i.e. $0$ for no superimposed anisotropy, $1$ or $2$ for
-generic pre-defined anisotropic models.
-\item [{\texttt{rho,vp,vs}}] reference isotropic model parameters for density
-$\rho$, $v_{p}$ and $v_{s}$.
-\end{description}
-Output from the routine consists of the following non-dimensional
-model parameters:
-\begin{description}
-\item [{\texttt{c11},}] \textbf{$\cdots$,} \texttt{\textbf{c66}} 21~dimensionalized
-anisotropic elastic parameters.
-\end{description}
-You can replace the \texttt{model\_aniso.f90} file by your own version
-\textit{provided you do not change the call structure of the routine},
-i.e., the new routine should take exactly the same input and produce
-the required relative output.
-
-\section{{\normalsize \label{sec:Using-SEP}}Using external SEP models.}
-
-Stanford Exploration Project (SEP) models consists of two files per variables, one header file (\texttt{.H})
-and one binary file (\texttt{.H$@$}).
-Among other information, the header file indicates, for each dimension,
-the offset (\texttt{o1}, \texttt{o2} and \texttt{o3} fields),
-the spatial step size (\texttt{d1}, \texttt{d2} and \texttt{d3} fields)
-and the number of samples (\texttt{n1}, \texttt{n2} and \texttt{n3}) necessary to read
-the linked binary file (\texttt{in} field).
-
-SEP model reading occurs during database generation. It it the user responsibility
-to ensure that proper interfaces between acoustic and elastic domains are set up
-during the meshing phase. For more information, refer to
-section~\ref{cha:Running-the-Mesher-Meshfem3D}.
-
-In \texttt{DATA/Par\_file}, \texttt{\textbf{MODEL}} should be set to \texttt{\textbf{sep}},
-and \texttt{\textbf{SEP\_MODEL\_DIRECTORY}} should point to wherever your sep model files reside.
-
-Note that in order not to overload the parameter file, SEP header files
-should be named \texttt{vp.H}, \texttt{vs.H} and \texttt{rho.H}. The \texttt{in} field
-in these binary files can be any path as long as it is relative to
-the previously set \texttt{SEP\_MODEL\_DIRECTORY}.
-We also assume that any dimensional value inside SEP header files is given in meters and
-that binary files are single precision floating point numbers stored in little endian format.
-There is currently only support for \texttt{vp}, \texttt{vs} and \texttt{rho} variables.
-
-An example demonstrating the use SEP models is given in:
-\texttt{EXAMPLES/meshfem3D\_examples/sep\_bathymetry}.
-
-\chapter{Post-Processing Scripts}
-
-Several post-processing scripts/programs are provided in the \texttt{utils/}
-directory, and most of them need to be adjusted when used on different
-systems, for example, the path of the executable programs. Here we
-only list a few of the available scripts and provide a brief description,
-and you can either refer to the related sections for detailed usage
-or, in a lot of cases, type the script/program name without arguments
-for its usage.
-
-\section{Process Data and Synthetics\label{sec:Process-data-and-syn}}
-
-In many cases, the SEM synthetics are calculated and compared to data
-seismograms recorded at seismic stations. Since the SEM synthetics
-are accurate for a certain frequency range, both the original data
-and the synthetics need to be processed before a comparison can be
-made.
-
-For such comparisons, the following steps are recommended:
-\begin{enumerate}
-\item Make sure that both synthetic and observed seismograms have the correct
-station/event and timing information.
-\item Convolve synthetic seismograms with a source time function with the
-half duration specified in the \texttt{CMTSOLUTION} file, provided,
-as recommended, you used a zero half duration in the SEM simulations.
-\item Resample both observed and synthetic seismograms to a common sampling
-rate.
-\item Cut the records using the same window.
-\item Remove the trend and mean from the records and taper them.
-\item Remove the instrument response from the observed seismograms (recommended)
-or convolve the synthetic seismograms with the instrument response.
-\item Make sure that you apply the same filters to both observed and synthetic
-seismograms. Preferably, avoid filtering your records more than once.
-\item Now, you are ready to compare your synthetic and observed seismograms.
-\end{enumerate}
-\noindent We generally use the following scripts provided in the \texttt{utils/seis\_process/}
-directory:
-
-\subsection{Data processing script \texttt{process\_data.pl}}
-
-This script cuts a given portion of the original data, filters it,
-transfers the data into a displacement record, and picks the first
-P and S arrivals. For more functionality, type `\texttt{process\_data.pl}'
-without any argument. An example of the usage of the script:
-\begin{lyxcode}
-{\footnotesize process\_data.pl~-m~CMTSOLUTION~-s~1.0~-l~0/4000~-i~-f~-t~40/500~-p~-x~bp~DATA/1999.330{*}.BH?.SAC}~
-\end{lyxcode}
-\noindent which has resampled the SAC files to a sampling rate of
-1 seconds, cut them between 0 and 4000 seconds, transfered them into
-displacement records and filtered them between 40 and 500 seconds,
-picked the first P and S arrivals, and added suffix `\texttt{bp}'
-to the file names.
-
-Note that all of the scripts in this section actually use the SAC
-and/or IASP91 to do the core operations; therefore make sure that
-the SAC and IASP91 packages are installed properly on your system,
-and that all the environment variables are set properly before running
-these scripts.
-
-\subsection{Synthetics processing script \texttt{process\_syn.pl}}
-
-This script converts the synthetic output from the SEM code from ASCII
-to SAC format, and performs similar operations as `\texttt{process\_data.pl}'.
-An example of the usage of the script:
-\begin{lyxcode}
-{\footnotesize process\_syn.pl~-m~CMTSOLUTION~-h~-a~STATIONS~-s~1.0~-l~0/4000~-f~-t~40/500~-p~-x~bp~SEM/{*}.BX?.semd}~
-\end{lyxcode}
-which will convolve the synthetics with a triangular source-time function
-from the \texttt{CMTSOLUTION} file, convert the synthetics into SAC
-format, add event and station information into the SAC headers, resample
-the SAC files with a sampling rate of 1 seconds, cut them between
-0 and 4000 seconds, filter them between 40 and 500 seconds with the
-same filter used for the observed data, pick the first P and S arrivals,
-and add the suffix `\texttt{bp}' to the file names.
-
-More options are available for this script, such as adding time shift
-to the origin time of the synthetics, convolving the synthetics with
-a triangular source time function with a given half duration, etc.
-Type \texttt{process\_syn.pl} without any argument for a detailed
-usage.
-
-In order to convert between SAC format and ASCII files, useful scripts
-are provided in the subdirectories ~\\
- \texttt{utils/sac2000\_alpha\_convert/} and \texttt{utils/seis\_process/asc2sac/}.
-
-\subsection{Script \texttt{rotate.pl}}
-
-The original data and synthetics have three components: vertical (BHZ
-resp. BXZ), north (BHN resp. BXN) and east (BHE resp. BXE). However,
-for most seismology applications, transverse and radial components
-are also desirable. Therefore, we need to rotate the horizontal components
-of both the data and the synthetics to the transverse and radial direction,
-and \texttt{\small rotate.pl}{\small{} can be used to accomplish this:}{\small \par}
-\begin{lyxcode}
-rotate.pl~-l~0~-L~180~-d~DATA/{*}.BHE.SAC.bp~
-
-rotate.pl~-l~0~-L~180~SEM/{*}.BXE.semd.sac.bp~
-\end{lyxcode}
-where the first command performs rotation on the SAC data obtained
-through Seismogram Transfer Program (STP) \urlwithparentheses{http://www.data.scec.org/STP/stp.html},
-while the second command rotates the processed SEM synthetics.
-
-\section{Collect Synthetic Seismograms}
-
-The forward and adjoint simulations generate synthetic seismograms
-in the \texttt{OUTPUT\_FILES/} directory by default. For the forward
-simulation, the files are named like \texttt{NT.STA.BX?.semd} for
-two-column time series, or \texttt{NT.STA.BX?.semd.sac} for ASCII
-SAC format, where NT and STA are the network code and station name,
-and \texttt{BX?} stands for the component name. Please see the Appendix~\ref{cha:Coordinates}
-and \ref{cha:channel-codes} for further details.
-
-The adjont simulations generate synthetic seismograms with the name
-\texttt{NT.S?????.S??.sem} (refer to Section \ref{sec:Adjoint-simulation-sources}
-for details). The kernel simulations output the back-reconstructed
-synthetic seismogram in the name \texttt{NT.STA.BX?.semd}, mainly
-for the purpose of checking the accuracy of the reconstruction. Refer
-to Section \ref{sec:Adjoint-simulation-finite} for further details.
-
-You do have further options to change this default output behavior,
-given in the main constants file \texttt{constants.h} located in \texttt{src/shared/}
-directory:
-\begin{description}
-\item [{\texttt{SEISMOGRAMS\_BINARY}}] set to \texttt{.true.} to have seismograms
-written out in binary format.
-\item [{\texttt{WRITE\_SEISMOGRAMS\_BY\_MASTER}}] Set to \texttt{.true.}
-to have only the master process writing out seismograms. This can
-be useful on a cluster, where only the master process node has access
-to the output directory.
-\item [{\texttt{USE\_OUTPUT\_FILES\_PATH}}] Set to \texttt{.false.} to
-have the seismograms output to \texttt{LOCAL\_PATH} directory specified
-in the main parameter file \texttt{DATA/Par\_file}. In this case,
-you could collect the synthetics onto the frontend using the \texttt{collect\_seismo\_lsf\_multi.pl}
-script located in the \texttt{utils/Cluster/lsf/} directory. The usage
-of the script would be e.g.:
-
-\begin{lyxcode}
-collect\_seismo.pl~machines~DATA/Par\_file
-\end{lyxcode}
-\end{description}
-where \texttt{machines} is a file containing the node names and \texttt{DATA/Par\_file}
-the parameter file used to extract the \texttt{LOCAL\_PATH} directory
-used for the simulation.
-
-\section{Clean Local Database}
-
-After all the simulations are done, the seismograms are collected,
-and the useful database files are copied to the frontend, you may
-need to clean the local scratch disk for the next simulation. This
-is especially important in the case of kernel simulation, where very
-large files are generated for the absorbing boundaries to help with
-the reconstruction of the regular forward wavefield. A sample script
-is provided in \texttt{utils/}:
-\begin{lyxcode}
-cleanbase.pl~machines
-\end{lyxcode}
-where \texttt{machines} is a file containing the node names.
-
-\section{Plot Movie Snapshots and Synthetic Shakemaps}
-
-\subsection{Script \texttt{movie2gif.gmt.pl}}
-
-With the movie data saved in \texttt{OUTPUT\_FILES/} at the end of
-a movie simulation (\texttt{\small MOVIE\_SURFACE=.true.}{\small ),
-you can run the }\texttt{\small `create\_movie\_shakemap\_AVS\_DX\_GMT}{\small '
-code to convert these binary movie data into GMT xyz files for futher
-processing. A sample script }\texttt{\small movie2gif.gmt.pl}{\small{}
-is provided to do this conversion, and then plot the movie snapshots
-in GMT, for example:}{\small \par}
-\begin{lyxcode}
-movie2gif.gmt.pl~-m~CMTSOLUTION~-g~-f~1/40~-n~-2~-p~
-\end{lyxcode}
-which for the first through the 40th movie frame, converts the \texttt{moviedata}
-files into GMT xyz files, interpolates them using the 'nearneighbor'
-command in GMT, and plots them on a 2D topography map. Note that `\texttt{-2}'
-and `\texttt{-p}' are both optional.
-
-\subsection{Script \texttt{plot\_shakemap.gmt.pl}}
-
-With the shakemap data saved in \texttt{OUTPUT\_FILES/} at the end
-of a shakemap simulation ~\\
- (\texttt{CREATE\_SHAKEMAP=.true.}), you can also run \texttt{`create\_movie\_shakemap\_AVS\_DX\_GMT}'
-code to convert the binary shakemap data into GMT xyz files. A sample
-script \texttt{plot\_shakemap.gmt.pl} is provided to do this conversion,
-and then plot the shakemaps in GMT, for example:
-\begin{lyxcode}
-plot\_shakemap.gmt.pl~data\_dir~type(1,2,3)~CMTSOLUTION~
-\end{lyxcode}
-where \texttt{type=1} for a displacement shakemap, \texttt{2} for
-velocity, and \texttt{3} for acceleration.
-
-\section{Map Local Database}
-
-A sample program \texttt{remap\_database} is provided to map the local
-database from a set of machines to another set of machines. This is
-especially useful when you want to run mesher and solver, or different
-types of solvers separately through a scheduler (refer to Chapter~\ref{cha:Scheduler}).
-\begin{lyxcode}
-run\_lsf.bash~-{}-gm-no-shmem~-{}-gm-copy-env~remap\_database~old\_machines~150
-\end{lyxcode}
-where \texttt{old\_machines} is the LSF machine file used in the previous
-simulation, and \texttt{150} is the number of processors in total.
-
-%============================================================
-
-\chapter{\label{cha:developers}Information for developers of the code, and for people who want to learn how the technique works}
-
-You can get a very simple 1D version of a demo code (there is one in Fortran and one in Python):\\
-
-git clone --recursive https://github.com/geodynamics/specfem1d.git\\
-
-\noindent We also have simple 3D demo source codes that implement the SEM in a single, small program, in directory\\
-utils/small\_SEM\_solvers\_in\_Fortran\_and\_C\_without\_MPI\_to\_learn of the specfem3d package.
-They are useful to learn how the spectral-element method works, and how to write or modify a code to implement it.
-Also useful to test new ideas by modifying these simple codes to run some tests.\\
-
-\noindent For information on how to contribute to the code, i.e., for how to make your modifications, additions or improvements part of the
-official package, see https://github.com/geodynamics/specfem3d/wiki .
-
-%============================================================
-
-\chapter*{\label{cha:Bug-Reports-and}Bug Reports and Suggestions for Improvements}
-
-To report bugs or suggest improvements to the code, please send an
-e-mail to the CIG Computational Seismology Mailing List \urlwithparentheses{cig-seismo@geodynamics.org}.
-
-\addcontentsline{toc}{chapter}{Notes \& Acknowledgments}
-
-\chapter*{Notes \& Acknowledgments}
-
-In order to keep the software package thread-safe in case a multithreaded
-implementation of MPI is used, developers should not add modules or
-common blocks to the source code but rather use regular subroutine
-arguments (which can be grouped in ``derived types'' if needed for
-clarity).
-
-The Gauss-Lobatto-Legendre subroutines in \texttt{gll\_library.f90}
-are based in part on software libraries from the Massachusetts Institute
-of Technology, Department of Mechanical Engineering (Cambridge, Massachusetts,
-USA). The non-structured global numbering software was provided by
-Paul F. Fischer (Brown University, Providence, Rhode Island, USA,
-now at Argonne National Laboratory, USA).
-
-OpenDX \urlwithparentheses{http://www.opendx.org} is open-source
-based on IBM Data Explorer, AVS \urlwithparentheses{http://www.avs.com}
-is a trademark of Advanced Visualization Systems, and ParaView \urlwithparentheses{http://www.paraview.org}
-is an open-source visualization platform.
-
-Please e-mail your feedback, questions, comments, and suggestions
-to the CIG Computational Seismology Mailing List \urlwithparentheses{cig-seismo@geodynamics.org}.
-
-\addcontentsline{toc}{chapter}{Copyright}
-
-\chapter*{Copyright}
-
-Main historical authors: Dimitri Komatitsch and Jeroen Tromp
-
-Princeton University, USA, and CNRS / University of Marseille, France
-
-$\copyright$ Princeton University and CNRS / University of Marseille, July 2012
-
-This program is free software; you can redistribute it and/or modify
-it under the terms of the GNU General Public License as published
-by the Free Software Foundation (see Appendix \ref{cha:License}).\\
-
-\textbf{\underline{Evolution of the code:}}\\
-
-MPI v. 3.0, December 2014: many developers.
-Convolutional PML, LDDRK time scheme, bulk attenuation support, simultaneous MPI runs,
-ADIOS file I/O support, coupling with external codes, new seismogram names,
-Deville routines for additional GLL degrees, tomography tools, unit/regression test framework,
-improved CUDA GPUs performance, additonal GEOCUBIT support, better make compilation,
-git versioning system. \\
-
-MPI v. 2.1, July 2012: Max Rietmann, Peter Messmer, Daniel Peter, Dimitri
-Komatitsch, Joseph Charles, Zhinan Xie: support for CUDA GPUs, better
-CFL stability for the Stacey absorbing conditions. \\
-
-MPI v. 2.0, November 2010: Dimitri Komatitsch, Nicolas Le Goff, Roland
-Martin and Pieyre Le Loher, University of Pau, France, Daniel Peter,
-Jeroen Tromp and the Princeton group of developers, Princeton University,
-USA, and Emanuele Casarotti, INGV Roma, Italy: support for CUBIT meshes
-decomposed by SCOTCH; much faster solver using Michel Deville's inlined
-matrix products.\\
-
-MPI v. 1.4 Dimitri Komatitsch, University of Pau, Qinya Liu and others,
-Caltech, September 2006: better adjoint and kernel calculations, faster
-and better I/Os on very large systems, many small improvements and
-bug fixes.\\
-
-MPI v. 1.3 Dimitri Komatitsch, University of Pau, and Qinya Liu, Caltech,
-July 2005: serial version, regular mesh, adjoint and kernel calculations,
-ParaView support.\\
-
-MPI v. 1.2 Min Chen and Dimitri Komatitsch, Caltech, July 2004: full
-anisotropy, volume movie.\\
-
-MPI v. 1.1 Dimitri Komatitsch, Caltech, October 2002: Zhu's Moho map,
-scaling of $V_{s}$ with depth, Hauksson's regional model, attenuation,
-oceans, movies.\\
-
-MPI v. 1.0 Dimitri Komatitsch, Caltech, USA, May 2002: first MPI version
-based on global code.\\
-
-Dimitri Komatitsch, IPG Paris, France, December 1996: first 3-D solver
-for the CM-5 Connection Machine, parallelized on 128 processors using
-Connection Machine Fortran.\\
-
-\addcontentsline{toc}{chapter}{References} \bibliography{bibliography}
-
-\appendix
-%dummy comment inserted by tex2lyx to ensure that this paragraph is not empty
-
-\chapter{\label{cha:Coordinates}Reference Frame Convention}
-
-The code uses the following convention for the Cartesian reference
-frame:
-\begin{itemize}
-\item the $x$ axis points East
-\item the $y$ axis points North
-\item the $z$ axis points up
-\end{itemize}
-Note that this convention is different from both the \citet{AkRi80}
-convention and the Harvard Centroid-Moment Tensor (CMT) convention.
-The Aki \& Richards convention is
-\begin{itemize}
-\item the $x$ axis points North
-\item the $y$ axis points East
-\item the $z$ axis points down
-\end{itemize}
-and the Harvard CMT convention is
-\begin{itemize}
-\item the $x$ axis points South
-\item the $y$ axis points East
-\item the $z$ axis points up
-\end{itemize}
-
-\subsection*{Source and receiver locations}
-
-The SPECFEM3D Cartesian software code internally uses Cartesian coordinates.
-The given locations for sources and receiver locations thus may get
-converted. Sources and receiver locations are read in from the \texttt{CMTSOLUTION}
-(or \texttt{FORCESOLUTION}) and \texttt{STATIONS} files. Note that
-e.g. the \texttt{CMTSOLUTION} and \texttt{FORCESOLUTION} files denotes
-the location by \textquotedbl{}longitude/latitude/depth\textquotedbl{}.
-We read in longitude as $x$ coordinate, latitude as $y$ coordinate.
-
-In case the flag \texttt{SUPPRESS\_UTM\_PROJECTION} is set to \texttt{.false.}
-in the main parameter file (see Chapter~\ref{cha:Creating-Distributed-Databases}),
-the $x$/$y$ coordinates have to be given in degrees and are converted
-to Cartesian coordinates using a UTM conversion with the specified
-UTM zone.
-
-The value for depth (given in $km$ in \texttt{CMTSOLUTION} and \texttt{FORCESOLUTION})
-or burial depth (given in $m$ in \texttt{STATIONS}) is evaluated
-with respect to the surface of the mesh at the specified $x$/$y$
-location to find the corresponding $z$ coordinate. It is possible
-to use this depth value directly as $z$ coordinate by changing the
-flag \texttt{USE\_SOURCES\_RECVS\_Z} to \texttt{.true.} in the file
-\texttt{constants.h} located in the \texttt{src/shared/} subdirectory.
-
-\subsection*{Seismogram outputs}
-
-The seismogram output directions are given in Cartesian $x$/$y$/$z$
-directions and not rotated any further. Changing flags in \texttt{constants.h}
-in the \texttt{src/shared/} subdirectory only rotates the seismogram
-outputs if receivers are forced to be located at the surface (\texttt{RECVS\_CAN\_BE\_BURIED\_EXT\_MESH}
-set to \texttt{.false.}) and the normal to the surface at the receiver
-location should be used (\texttt{EXT\_MESH\_RECV\_NORMAL} set to \texttt{.true.})
-as vertical. In this case, the outputs are rotated to have the vertical
-component normal to the surface of the mesh, $x$ and $y$ directions
-are somewhat arbitrary orthogonal directions along the surface.
-
-For the labeling of the seismogram channels, see Appendix~\ref{cha:channel-codes}.
-Additionally, we add labels to the synthetics using the following
-convention: For a receiver station located in an
-\begin{description}
-\item [{elastic domain:}] ~\\
-
-\begin{itemize}
-\item \texttt{semd} for the displacement vector
-\item \texttt{semv} for the velocity vector
-\item \texttt{sema} for the acceleration vector
-\end{itemize}
-\item [{acoustic domain:}] ~\\
- (please note that receiver stations in acoustic domains must be buried.
-This is due to the free surface condition which enforces a zero pressure
-at the surface.)
-
-\begin{itemize}
-\item \texttt{semd} for the displacement vector
-\item \texttt{semv} for the velocity vector
-\item \texttt{sema} for pressure which will be stored in the vertical component
-\texttt{Z}. Note that pressure in the acoustic media is isotropic
-and thus the pressure record would be the same in the other two component
-directions. We therefore use the other two seismogram components to
-store the acoustic potential in component \texttt{X} (or \texttt{N})
-and the first time derivative of the acoustic potential in component
-\texttt{Y} (or \texttt{E}).
-\end{itemize}
-\end{description}
-The seismograms are by default written out in ASCII-format to the
-\texttt{OUTPUT\_FILES/} subdirectory by each parallel process. You
-can change this behavior by changing the following flags in the \texttt{constants.h}
-file located in the \texttt{src/shared/} subdirectory:
-\begin{description}
-\item [{\texttt{SEISMOGRAMS\_BINARY}}] set to \texttt{.true.} to have seismograms
-written out in binary format.
-\item [{\texttt{WRITE\_SEISMOGRAMS\_BY\_MASTER}}] Set to \texttt{.true.}
-to have only the master process writing out seismograms. This can
-be useful on a cluster, where only the master process node has access
-to the output directory.
-\item [{\texttt{USE\_OUTPUT\_FILES\_PATH}}] Set to \texttt{.false.} to
-have the seismograms output to \texttt{LOCAL\_PATH} directory specified
-in the main parameter file \texttt{DATA/Par\_file}.
-\end{description}
-
-\chapter{\label{cha:channel-codes}Channel Codes of Seismograms}
-
-Seismic networks, such as the Global Seismographic Network (GSN),
-generally involve various types of instruments with different bandwidths,
-sampling properties and component configurations. There are standards
-to name channel codes depending on instrument properties. IRIS \texttt{(www.iris.edu)}
-uses SEED format for channel codes, which are represented by three
-letters, such as \texttt{LHN}, \texttt{BHZ}, etc. In older versions
-of the SPECFEM3D Cartesian package, a common format was used for the
-channel codes of all seismograms, which was \texttt{BHE/BHN/BHZ} for
-three components. To avoid confusion when comparison are made to observed
-data, we are now using the FDSN convention (\url{http://www.fdsn.org/})
-for SEM seismograms. In the following, we give a brief explanation
-of the FDSN convention used by IRIS, and how it is adopted in SEM
-seismograms. Please visit \texttt{www.iris.edu/manuals/SEED\_appA.htm}
-for further information.\\
-
-\noindent \texttt{\textbf{Band code:}} The first letter in the channel
-code denotes the band code of seismograms, which depends on the response
-band and the sampling rate of instruments. The list of band codes
-used by IRIS is shown in Figure \ref{fig:IRIS_band_codes}. The sampling
-rate of SEM synthetics is controlled by the resolution of simulations
-rather than instrument properties. However, for consistency, we follow
-the FDSN convention for SEM seismograms governed by their sampling
-rate. For SEM synthetics, we consider band codes for which $dt\leq1$
-s. IRIS also considers the response band of instruments. For instance,
-short-period and broad-band seismograms with the same sampling rate
-correspond to different band codes, such as S and B, respectively.
-In such cases, we consider SEM seismograms as broad band, ignoring
-the corner period ($\geq10$ s) of the response band of instruments
-(note that at these resolutions, the minimum period in the SEM synthetics
-will be less than $10$ s). Accordingly, when you run a simulation
-the band code will be chosen depending on the resolution of the synthetics,
-and channel codes of SEM seismograms will start with either \texttt{L},
-\texttt{M}, \texttt{B}, \texttt{H}, \texttt{C} or \texttt{F}, shown
-by red color in the figure.\\
-
-\begin{figure}[ht]
-\noindent \begin{centering}
-\includegraphics[scale=0.6]{figures/IRIS_band_codes}
-\par\end{centering}
-
-\caption{The band code convention is based on the sampling rate and the response
-band of instruments. Please visit \texttt{www.iris.edu/manuals/SEED\_appA.htm}
-for further information. Grey rows show the relative band-code range
-in SPECFEM3D Cartesian, and the band codes used to name SEM seismograms
-are denoted in red.}
-
-\label{fig:IRIS_band_codes}
-\end{figure}
-
-\noindent \texttt{\textbf{Instrument code:}} The second letter in
-the channel code corresponds to instrument codes, which specify the
-family to which the sensor belongs. For instance, \texttt{H} and \texttt{L}
-are used for high-gain and low-gain seismometers, respectively. The
-instrument code of SEM seismograms will always be \texttt{X}, as assigned
-by FDSN for synthetic seismograms. \\
-
-\noindent \texttt{\textbf{Orientation code:}} The third letter in
-channel codes is an orientation code, which generally describes the
-physical configuration of the components of instrument packages. SPECFEM3D
-Cartesian uses the traditional orientation code \texttt{E/N/Z} (East-West,
-North-South, Vertical) for three components when a UTM projection
-is used. If the UTM conversion is suppressed, i.e. the flag \texttt{SUPPRESS\_UTM\_PROJECTION}
-is set to \texttt{.true.}, then the three components are labelled
-\texttt{X/Y/Z} according to the Cartesian reference frame. \\
-
-\noindent \texttt{\textbf{EXAMPLE:}} The sampling rate is given by
-\texttt{DT} in the main parameter file \texttt{Par\_file} located
-in the \texttt{DATA/} subdirectory. Depending on the resolution of
-your simulations, if you choose a sampling rate greater than $0.01$
-s and less than $1$ s, a seismogram recording displacements on the
-vertical component of a station \texttt{ASBS} (network \texttt{AZ})
-will be named \texttt{AZ.ASBS.MXZ.semd.sac}, whereas it will be \texttt{AZ.ASBS.BXZ.semd.sac},
-if the sampling rate is greater than 0.0125 and less equal to 0.1
-s.
-
-\chapter{\label{cha:Troubleshooting}Troubleshooting}
-
-\section*{FAQ}
-\begin{description}
-\item [{configuration fails:}]~
-
-Examine the log file 'config.log'. It contains detailed informations.
-In many cases, the paths to these specific compiler commands F90,
-CC and MPIF90 will not be correct if `./configure` fails.
-
-Please make sure that you have a working installation of a Fortran
-compiler, a C compiler and an MPI implementation. You should be able
-to compile this little program code:
-\begin{lyxcode}
-program~main~~~\\
-~include~'mpif.h'~~~\\
-~integer,~parameter~::~CUSTOM\_MPI\_TYPE~=~MPI\_REAL~~~\\
-~integer~ier~~~\\
-~call~MPI\_INIT(ier)~~~\\
-~call~MPI\_BARRIER(MPI\_COMM\_WORLD,ier)~~~\\
-~call~MPI\_FINALIZE(ier)~~~\\
-~end
-\end{lyxcode}
-\item [{compilation fails stating:}] ~\\
-
-\begin{lyxcode}
-...~~~\\
-~obj/program\_generate\_databases.o:~In~function~`MAIN\_\_':~~~\\
-~program\_generate\_databases.f90:(.text+0x14):~undefined~reference~to~`\_gfortran\_set\_std'~~~\\
-~...~~~\\
-
-\end{lyxcode}
-
-Make sure you're pointing to the right 'mpif90' wrapper command.
-
-Normally, this message will appear when you are mixing two different
-Fortran compilers. That is, using e.g. gfortran to compile non-MPI
-files and mpif90, wrapper provided for e.g. ifort, to compile MPI-files.
-
-fix: e.g. specify > ./configure FC=gfortran MPIF90=/usr/local/openmpi-gfortran/bin/mpif90
-
-\item [{after executing \texttt{xmeshfem3D} I've got elements with skewness of 81\% percent, what does this mean:}] Look
-at the skewness table printed in the \texttt{output\_mesher.txt} file
-after executing \texttt{xmeshfem3D} for the example given in \texttt{EXAMPLES/meshfem3D\_examples/simple\_model/}:
-
-\begin{lyxcode}
-...~~\\
-~histogram~of~skewness~(0.~good~-~1.~bad):~~~\\
-~0.0000000E+00~-~5.0000001E-02~27648~81.81818~\%~~~\\
-~5.0000001E-02~-~0.1000000~0~0.0000000E+00~\%~~~\\
-~...~~~\\
-
-\end{lyxcode}
-
-The first line means that you have 27,648 elements with a skewness
-value between 0 and 0.05 (which means the element is basically not
-skewed, just plain regular hexahedral element). The total number of
-elements you have in this mesh is (see in the \texttt{output\_mesher.txt}
-file a bit further down):
-\begin{lyxcode}
-...~~\\
-~total~number~of~elements~in~entire~mesh:~33792~~~\\
-~...~~~\\
-
-\end{lyxcode}
-
-which gives you that: 27,648 / 33,792 $\sim$ 81.8 \% of all elements
-are not skewed, i.e. regular elements. a fantastic value :)
-
-The histogram lists for this mesh also some stronger skewed elements,
-for example the worst ones belong to:
-\begin{lyxcode}
-...~~\\
-~0.6000000~-~0.6500000~2048~6.060606~\%~~~\\
-~...~~~\\
-
-\end{lyxcode}
-
-about 6 \% of all elements have distortions with a skewness value
-between 0.6 and 0.65. The skewness values give you a hint of how good
-your mesh is. In an ideal world, you would want to have no distortions,
-just like the 81\% from above. Those elements give you the best approximate
-values by the GLL quadrature used in the spectral-element method.
-However, having weakly distorted elements is still fine and the solutions
-are still accurate enough. So empirically, values up to around 0.7
-are tolerable, above that you should consider remeshing...
-
-To give you an idea why some of the elements are distorted, see the
-following figure \ref{fig:mesh.vp} of the mesh you obtain in the
-example \texttt{EXAMPLES/meshfem3D\_examples/simple\_model/}.
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=0.9\textwidth]{figures/mesh_vp}
-\par\end{centering}
-
-\caption{Paraview visualization using the mesh vtk-files for the example given
-in \texttt{EXAMPLES/meshfem3D\_examples/simple\_model/}.}
-
-\label{fig:mesh.vp}
-\end{figure}
-
-You will see that the mesh contains a doubling layer, where we stitch
-elements together such that the size of two elements will transition
-to the size of one element (very useful to keep the ratio of wavespeed
-/ element\_size about constant). Those elements in this doubling layer
-have higher skewness values and make up those 6 \% in the histogram.
-
-\item [{the code gives following error message "need at least one receiver":}] This
-means that no stations given in the input file \texttt{DATA/STATIONS}
-could be located within the dimensions of the mesh. This can happen
-for example when the mesh was created with the in-house mesher \texttt{xmeshfem3D}
-while using the Universal Transverse Mercator (UTM) projection but
-the simulation with \texttt{xspecfem3D} was suppressing this projection
-from latitude/longitude to x/y/z coordinates.
-
-In such cases, try to change your \texttt{DATA/Par\_file} and set
-e.g.:
-\begin{lyxcode}
-SUPPRESS\_UTM\_PROJECTION~=~.false.~~~\\
-
-\end{lyxcode}
-
-to be the same in \texttt{Mesh\_Par\_file} and \texttt{Par\_file}.
-This flag should be identical when using the in-house mesher \texttt{xmeshfem3D},
-\texttt{xgenerate\_databases} and \texttt{xspecfem3D} together to
-run simulations.
-
-The flag determines if the coordinates you specify for your source
-and station locations are given as lat/lon degrees and must be converted
-to UTM coordinates. As an example, if you use \texttt{.false.} within
-\texttt{Mesh\_Par\_file} then you create a mesh with \texttt{xmeshfem3D}
-using the UTM projection from lat/lon as input format to UTM projected
-coordinates to store the mesh point positions, which is fine. The
-error then may occur if in the \texttt{Par\_file} you have this set
-to \texttt{.true.} so that the \texttt{xgenerate\_databases} and \texttt{xspecfem3D}
-suppress the UTM projection and assume that all coordinates you use
-now for source and receiver locations are given in meters (that is,
-converted) already. So it won't find the specified locations in the
-used mesh. As a solutions, just change the flag in \texttt{Par\_file}
-to be the same as in \texttt{Mesh\_Par\_file} and rerun \texttt{xgenerate\_databases}
-and \texttt{xspecfem3D} to make sure that your simulation works fine.
-
-\item [{I get the following error message "forward simulation became unstable and blew up":}] In
-most cases this means that your time step size \texttt{DT} is chosen
-too big. Look at your files \texttt{output\_mesher.txt} or \texttt{output\_solver.txt}
-created in the folder \texttt{OUTPUT\_FILES}. In these output files,
-find the section:
-
-\begin{lyxcode}
-...~~~\\
-~{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}~~~\\
-~{*}{*}{*}~Verification~of~simulation~parameters~{*}{*}{*}~~~\\
-~{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}~~~\\
-~...~~~\\
-~{*}{*}{*}~Minimum~period~resolved~=~4.308774~~~\\
-~{*}{*}{*}~Maximum~suggested~time~step~=~6.8863556E-02~~~\\
-~...~~~\\
-~...~~~\\
-
-\end{lyxcode}
-
-then change \texttt{DT} in the \texttt{DATA/Par\_file} to be somewhere
-close to the maximum suggested time step. In the example above: \\
- \texttt{DT = 0.05d0} \\
- would (most probably) work fine. It could be also bigger than the
-0.068 s suggested. This depends a bit on the distortions of your mesh
-elements. The more regular they are, the bigger you can choose \texttt{DT}.
-Just play with this value a bit and see when the simulation becomes
-stable ...
-
-\end{description}
-
-\chapter{\label{cha:License}License }
-
-\textbf{GNU GENERAL PUBLIC LICENSE Version 2, June 1991. Copyright
-(C) 1989, 1991 Free Software Foundation, Inc. 59 Temple Place, Suite
-330, Boston, MA 02111-1307 USA} \\
- Everyone is permitted to copy and distribute verbatim copies of this
-license document, but changing it is not allowed.
-
-\section*{Preamble}
-
-The licenses for most software are designed to take away your freedom
-to share and change it. By contrast, the GNU General Public License
-is intended to guarantee your freedom to share and change free software
--- to make sure the software is free for all its users. This General
-Public License applies to most of the Free Software Foundation's software
-and to any other program whose authors commit to using it. (Some other
-Free Software Foundation software is covered by the GNU Library General
-Public License instead.) You can apply it to your programs, too.
-
-When we speak of free software, we are referring to freedom, not price.
-Our General Public Licenses are designed to make sure that you have
-the freedom to distribute copies of free software (and charge for
-this service if you wish), that you receive source code or can get
-it if you want it, that you can change the software or use pieces
-of it in new free programs; and that you know you can do these things.
-
-To protect your rights, we need to make restrictions that forbid anyone
-to deny you these rights or to ask you to surrender the rights. These
-restrictions translate to certain responsibilities for you if you
-distribute copies of the software, or if you modify it.
-
-For example, if you distribute copies of such a program, whether gratis
-or for a fee, you must give the recipients all the rights that you
-have. You must make sure that they, too, receive or can get the source
-code. And you must show them these terms so they know their rights.
-
-We protect your rights with two steps:
-\begin{enumerate}
-\item Copyright the software, and
-\item Offer you this license which gives you legal permission to copy, distribute
-and/or modify the software.
-\end{enumerate}
-Also, for each author's protection and ours, we want to make certain
-that everyone understands that there is no warranty for this free
-software. If the software is modified by someone else and passed on,
-we want its recipients to know that what they have is not the original,
-so that any problems introduced by others will not reflect on the
-original authors' reputations.
-
-Finally, any free program is threatened constantly by software patents.
-We wish to avoid the danger that redistributors of a free program
-will individually obtain patent licenses, in effect making the program
-proprietary. To prevent this, we have made it clear that any patent
-must be licensed for everyone's free use or not licensed at all.
-
-The precise terms and conditions for copying, distribution and modification
-follow.
-
-\section*{GNU GENERAL PUBLIC LICENSE TERMS AND CONDITIONS FOR COPYING, DISTRIBUTION
-AND MODIFICATION }
-\begin{itemize}
-\item 0.This License applies to any program or other work which contains
-a notice placed by the copyright holder saying it may be distributed
-under the terms of this General Public License. The ``Program''
-below refers to any such program or work, and a ``work based on the
-Program'' means either the Program or any derivative work under copyright
-law: that is to say, a work containing the Program or a portion of
-it, either verbatim or with modifications and/or translated into another
-language. (Hereinafter, translation is included without limitation
-in the term ``modification.'') Each licensee is addressed as ``you.''\\
- \\
- Activities other than copying, distribution and modification are
-not covered by this License; they are outside its scope. The act of
-running the Program is not restricted, and the output from the Program
-is covered only if its contents constitute a work based on the Program
-(independent of having been made by running the Program). Whether
-that is true depends on what the Program does.\end{itemize}
-\begin{enumerate}
-\item You may copy and distribute verbatim copies of the Program's source
-code as you receive it, in any medium, provided that you conspicuously
-and appropriately publish on each copy an appropriate copyright notice
-and disclaimer of warranty; keep intact all the notices that refer
-to this License and to the absence of any warranty; and give any other
-recipients of the Program a copy of this License along with the Program.
-
-You may charge a fee for the physical act of transferring a copy,
-and you may at your option offer warranty protection in exchange for
-a fee.
-
-\item You may modify your copy or copies of the Program or any portion of
-it, thus forming a work based on the Program, and copy and distribute
-such modifications or work under the terms of Section 1 above, provided
-that you also meet all of these conditions:
-
-\begin{enumerate}
-\item You must cause the modified files to carry prominent notices stating
-that you changed the files and the date of any change.
-\item You must cause any work that you distribute or publish, that in whole
-or in part contains or is derived from the Program or any part thereof,
-to be licensed as a whole at no charge to all third parties under
-the terms of this License.
-\item If the modified program normally reads commands interactively when
-run, you must cause it, when started running for such interactive
-use in the most ordinary way, to print or display an announcement
-including an appropriate copyright notice and a notice that there
-is no warranty (or else, saying that you provide a warranty) and that
-users may redistribute the program under these conditions, and telling
-the user how to view a copy of this License. (Exception: if the Program
-itself is interactive but does not normally print such an announcement,
-your work based on the Program is not required to print an announcement.)
-\end{enumerate}
-
-These requirements apply to the modified work as a whole. If identifiable
-sections of that work are not derived from the Program, and can be
-reasonably considered independent and separate works in themselves,
-then this License, and its terms, do not apply to those sections when
-you distribute them as separate works. But when you distribute the
-same sections as part of a whole which is a work based on the Program,
-the distribution of the whole must be on the terms of this License,
-whose permissions for other licensees extend to the entire whole,
-and thus to each and every part regardless of who wrote it.
-
-Thus, it is not the intent of this section to claim rights or contest
-your rights to work written entirely by you; rather, the intent is
-to exercise the right to control the distribution of derivative or
-collective works based on the Program.
-
-In addition, mere aggregation of another work not based on the Program
-with the Program (or with a work based on the Program) on a volume
-of a storage or distribution medium does not bring the other work
-under the scope of this License.
-
-\item You may copy and distribute the Program (or a work based on it, under
-Section 2) in object code or executable form under the terms of Sections
-1 and 2 above provided that you also do one of the following:
-
-\begin{enumerate}
-\item Accompany it with the complete corresponding machine-readable source
-code, which must be distributed under the terms of Sections 1 and
-2 above on a medium customarily used for software interchange; or,
-\item Accompany it with a written offer, valid for at least three years,
-to give any third party, for a charge no more than your cost of physically
-performing source distribution, a complete machine-readable copy of
-the corresponding source code, to be distributed under the terms of
-Sections 1 and 2 above on a medium customarily used for software interchange;
-or,
-\item Accompany it with the information you received as to the offer to
-distribute corresponding source code. (This alternative is allowed
-only for noncommercial distribution and only if you received the program
-in object code or executable form with such an offer, in accord with
-Subsection b above.)
-\end{enumerate}
-
-The source code for a work means the preferred form of the work for
-making modifications to it. For an executable work, complete source
-code means all the source code for all modules it contains, plus any
-associated interface definition files, plus the scripts used to control
-compilation and installation of the executable. However, as a special
-exception, the source code distributed need not include anything that
-is normally distributed (in either source or binary form) with the
-major components (compiler, kernel, and so on) of the operating system
-on which the executable runs, unless that component itself accompanies
-the executable.
-
-If distribution of executable or object code is made by offering access
-to copy from a designated place, then offering equivalent access to
-copy the source code from the same place counts as distribution of
-the source code, even though third parties are not compelled to copy
-the source along with the object code.
-
-\item You may not copy, modify, sublicense, or distribute the Program except
-as expressly provided under this License. Any attempt otherwise to
-copy, modify, sublicense or distribute the Program is void, and will
-automatically terminate your rights under this License. However, parties
-who have received copies, or rights, from you under this License will
-not have their licenses terminated so long as such parties remain
-in full compliance.
-\item You are not required to accept this License, since you have not signed
-it. However, nothing else grants you permission to modify or distribute
-the Program or its derivative works. These actions are prohibited
-by law if you do not accept this License. Therefore, by modifying
-or distributing the Program (or any work based on the Program), you
-indicate your acceptance of this License to do so, and all its terms
-and conditions for copying, distributing or modifying the Program
-or works based on it.
-\item Each time you redistribute the Program (or any work based on the Program),
-the recipient automatically receives a license from the original licensor
-to copy, distribute or modify the Program subject to these terms and
-conditions. You may not impose any further restrictions on the recipients'
-exercise of the rights granted herein. You are not responsible for
-enforcing compliance by third parties to this License.
-\item If, as a consequence of a court judgment or allegation of patent infringement
-or for any other reason (not limited to patent issues), conditions
-are imposed on you (whether by court order, agreement or otherwise)
-that contradict the conditions of this License, they do not excuse
-you from the conditions of this License. If you cannot distribute
-so as to satisfy simultaneously your obligations under this License
-and any other pertinent obligations, then as a consequence you may
-not distribute the Program at all. For example, if a patent license
-would not permit royalty-free redistribution of the Program by all
-those who receive copies directly or indirectly through you, then
-the only way you could satisfy both it and this License would be to
-refrain entirely from distribution of the Program.
-
-If any portion of this section is held invalid or unenforceable under
-any particular circumstance, the balance of the section is intended
-to apply and the section as a whole is intended to apply in other
-circumstances.
-
-It is not the purpose of this section to induce you to infringe any
-patents or other property right claims or to contest validity of any
-such claims; this section has the sole purpose of protecting the integrity
-of the free software distribution system, which is implemented by
-public license practices. Many people have made generous contributions
-to the wide range of software distributed through that system in reliance
-on consistent application of that system; it is up to the author/donor
-to decide if he or she is willing to distribute software through any
-other system and a licensee cannot impose that choice.
-
-This section is intended to make thoroughly clear what is believed
-to be a consequence of the rest of this License.
-
-\item If the distribution and/or use of the Program is restricted in certain
-countries either by patents or by copyrighted interfaces, the original
-copyright holder who places the Program under this License may add
-an explicit geographical distribution limitation excluding those countries,
-so that distribution is permitted only in or among countries not thus
-excluded. In such case, this License incorporates the limitation as
-if written in the body of this License.
-\item The Free Software Foundation may publish revised and/or new versions
-of the General Public License from time to time. Such new versions
-will be similar in spirit to the present version, but may differ in
-detail to address new problems or concerns.
-
-Each version is given a distinguishing version number. If the Program
-specifies a version number of this License which applies to it and
-``any later version,'' you have the option of following the terms
-and conditions either of that version or of any later version published
-by the Free Software Foundation. If the Program does not specify a
-version number of this License, you may choose any version ever published
-by the Free Software Foundation.
-
-\item If you wish to incorporate parts of the Program into other free programs
-whose distribution conditions are different, write to the author to
-ask for permission. For software which is copyrighted by the Free
-Software Foundation, write to the Free Software Foundation; we sometimes
-make exceptions for this. Our decision will be guided by the two goals
-of preserving the free status of all derivatives of our free software
-and of promoting the sharing and reuse of software generally.
-\end{enumerate}
-
-\subsection*{NO WARRANTY }
-\begin{itemize}
-\item 11.BECAUSE THE PROGRAM IS LICENSED FREE OF CHARGE, THERE IS NO WARRANTY
-FOR THE PROGRAM, TO THE EXTENT PERMITTED BY APPLICABLE LAW. EXCEPT
-WHEN OTHERWISE STATED IN WRITING THE COPYRIGHT HOLDERS AND/OR OTHER
-PARTIES PROVIDE THE PROGRAM ``AS IS'' WITHOUT WARRANTY OF ANY KIND,
-EITHER EXPRESSED OR IMPLIED, INCLUDING, BUT NOT LIMITED TO, THE IMPLIED
-WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE.
-THE ENTIRE RISK AS TO THE QUALITY AND PERFORMANCE OF THE PROGRAM IS
-WITH YOU. SHOULD THE PROGRAM PROVE DEFECTIVE, YOU ASSUME THE COST
-OF ALL NECESSARY SERVICING, REPAIR OR CORRECTION.
-\item 12.IN NO EVENT UNLESS REQUIRED BY APPLICABLE LAW OR AGREED TO IN WRITING
-WILL ANY COPYRIGHT HOLDER, OR ANY OTHER PARTY WHO MAY MODIFY AND/OR
-REDISTRIBUTE THE PROGRAM AS PERMITTED ABOVE, BE LIABLE TO YOU FOR
-DAMAGES, INCLUDING ANY GENERAL, SPECIAL, INCIDENTAL OR CONSEQUENTIAL
-DAMAGES ARISING OUT OF THE USE OR INABILITY TO USE THE PROGRAM (INCLUDING
-BUT NOT LIMITED TO LOSS OF DATA OR DATA BEING RENDERED INACCURATE
-OR LOSSES SUSTAINED BY YOU OR THIRD PARTIES OR A FAILURE OF THE PROGRAM
-TO OPERATE WITH ANY OTHER PROGRAMS), EVEN IF SUCH HOLDER OR OTHER
-PARTY HAS BEEN ADVISED OF THE POSSIBILITY OF SUCH DAMAGES.
-\end{itemize}
-
-\section*{END OF TERMS AND CONDITIONS }
-
-\subsection*{How to Apply These Terms to Your New Programs}
-
-If you develop a new program, and you want it to be of the greatest
-possible use to the public, the best way to achieve this is to make
-it free software which everyone can redistribute and change under
-these terms.
-
-To do so, attach the following notices to the program. It is safest
-to attach them to the start of each source file to most effectively
-convey the exclusion of warranty; and each file should have at least
-the ``copyright'' line and a pointer to where the full notice is
-found. For example:
-\begin{quote}
-One line to give the program's name and a brief idea of what it does.
-Copyright {\footnotesize $\copyright$ (}year) (name of author)
-
-This program is free software; you can redistribute it and/or modify
-it under the terms of the GNU General Public License as published
-by the Free Software Foundation; either version 2 of the License,
-or (at your option) any later version.
-
-This program is distributed in the hope that it will be useful, but
-WITHOUT ANY WARRANTY; without even the implied warranty of MERCHANTABILITY
-or FITNESS FOR A PARTICULAR PURPOSE. See the GNU General Public License
-for more details.
-
-You should have received a copy of the GNU General Public License
-along with this program; if not, write to the Free Software Foundation,
-Inc., 59 Temple Place, Suite 330, Boston, MA 02111-1307 USA
-\end{quote}
-Also add information on how to contact you by electronic and paper
-mail.
-
-If the program is interactive, make it output a short notice like
-this when it starts in an interactive mode:
-\begin{quote}
-Gnomovision version 69, Copyright $\copyright$ year name of author
-Gnomovision comes with ABSOLUTELY NO WARRANTY; for details type `show
-w'. This is free software, and you are welcome to redistribute it
-under certain conditions; type `show c' for details.
-\end{quote}
-The hypothetical commands `show w' and `show c' should show the appropriate
-parts of the General Public License. Of course, the commands you use
-may be called something other than `show w' and `show c'; they could
-even be mouse-clicks or menu items -- whatever suits your program.
-
-You should also get your employer (if you work as a programmer) or
-your school, if any, to sign a ``copyright disclaimer'' for the
-program, if necessary. Here is a sample; alter the names:
-\begin{quote}
-Yoyodyne, Inc., hereby disclaims all copyright interest in the program
-`Gnomovision' (which makes passes at compilers) written by James Hacker.
-
-(signature of Ty Coon)\\
- 1 April 1989 \\
- Ty Coon, President of Vice
-\end{quote}
-This General Public License does not permit incorporating your program
-into proprietary programs. If your program is a subroutine library,
-you may consider it more useful to permit linking proprietary applications
-with the library. If this is what you want to do, use the GNU Library
-General Public License instead of this License.
-\end{document}

### 2015-04-24 doc/USER_MANUAL/10_gravity_calculations.tex
updates user manual to use facilitate the generation of a markdown version to the github wiki

--- 2015-04-18 doc/USER_MANUAL/manual_SPECFEM3D_Cartesian.tex
@@
-% !TEX encoding = IsoLatin
-%% LyX 2.0.2 created this file.  For more info, see http://www.lyx.org/.
-%% Do not edit unless you really know what you are doing.
-\documentclass[oneside,english]{book}
-\usepackage[T1]{fontenc}
-%% DK DK we do not accept accented letters, use the old LaTeX syntax
-%% DK DK for them instead: \'e \`e  \"e \^e and so on.
-%% DK DK \usepackage[latin1]{inputenc}
-\usepackage[letterpaper]{geometry}
-\geometry{verbose,tmargin=1in,bmargin=1in,lmargin=1in,rmargin=1in}
-\setcounter{secnumdepth}{3}
-\setcounter{tocdepth}{3}
-\usepackage{babel}
-\usepackage{float}
-\usepackage{textcomp}
-\usepackage{url}
-\usepackage{amsmath}
-\usepackage{amssymb}
-\usepackage{graphicx}
-\usepackage[authoryear]{natbib}
-\usepackage[unicode=true,pdfusetitle,
- bookmarks=true,bookmarksnumbered=true,bookmarksopen=false,
- breaklinks=false,pdfborder={0 0 0},backref=section,colorlinks=false,pdfpagemode=None]
- {hyperref}
-\hypersetup{
- pdftex,pdfstartview=FitH,pdfpagelayout=SinglePage}
-\usepackage{breakurl}
-
-\makeatletter
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% LyX specific LaTeX commands.
-\newcommand{\noun}[1]{\textsc{#1}}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Textclass specific LaTeX commands.
-\newenvironment{lyxcode}
-{\par\begin{list}{}{
-\setlength{\rightmargin}{\leftmargin}
-\setlength{\listparindent}{0pt}% needed for AMS classes
-\raggedright
-\setlength{\itemsep}{0pt}
-\setlength{\parsep}{0pt}
-\normalfont\ttfamily}%
- \item[]}
-{\end{list}}
-\newenvironment{lyxlist}[1]
-{\begin{list}{}
-{\settowidth{\labelwidth}{#1}
- \setlength{\leftmargin}{\labelwidth}
- \addtolength{\leftmargin}{\labelsep}
- \renewcommand{\makelabel}[1]{##1\hfil}}}
-{\end{list}}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% User specified LaTeX commands.
-%\renewcommand{\baselinestretch}{1.5}
-
-% figures
-\usepackage[dvips]{epsfig}
-
-% fonts
-\usepackage{times}
-
-% hyperlinks to sections and references
-
-\newcommand{\urlwithparentheses}[1]{(\url{#1})}
-
-% biblio GJI
-\bibliographystyle{abbrvnat}
-
-\newcommand{\nexxi}{\mbox{\texttt{NEX\_XI\/}}}
-\newcommand{\nexeta}{\mbox{\texttt{NEX\_ETA\/}}}
-\newcommand{\nprocxi}{\mbox{\texttt{NPROC\_XI\/}}}
-\newcommand{\nproceta}{\mbox{\texttt{NPROC\_ETA\/}}}
-\newcommand{\nchunks}{\mbox{\texttt{NCHUNKS\/}}}
-\newcommand{\eg}{e.g.,}
-
-% colors for the text
-\@ifundefined{definecolor}
- {\usepackage{color}}{}
-\newcommand{\red}[1]{\textcolor{Red}{#1}}
-
-\usepackage{babel}
-
-\makeatother
-
-\begin{document}
-\begin{center}
-\thispagestyle{empty}\vspace*{-1.8truecm} %
-\makebox[1\textwidth]{%
-\includegraphics[width=0.83\paperwidth]{figures/specfem_3d_Cartesian-cover} %
-}
-\par\end{center}
-
-\title{\textbf{SPECFEM3D Cartesian}\\
- \textbf{User Manual}}
-
-\author{$\copyright$ Princeton University (USA), CNRS and University of Marseille (France),\\
-and ETH Z\"urich (Switzerland)\\
-Version 3.0}
-
-\date{\today}
-
-\maketitle
-
-\section*{Authors}
-
-\noindent The SPECFEM3D Cartesian package was first developed by Dimitri
-Komatitsch and Jean-Pierre Vilotte at Institut de Physique du Globe
-(IPGP) in Paris, France from 1995 to 1997 and then by Dimitri Komatitsch
-and Jeroen Tromp at Harvard University and Caltech, USA, starting in 1998.
-The story started on April 4, 1995, when Prof. Yvon Maday from CNRS and University of Paris, France, gave a lecture to
-Dimitri Komatitsch and Jean-Pierre Vilotte at IPG about the nice properties of the Legendre spectral-element method with diagonal mass matrix that he had used for
-other equations. We are deeply indebted and thankful to him for that.
-That followed a visit by Dimitri Komatitsch to OGS (Istituto Nazionale di Oceanografia e di Geofisica Sperimentale) in Trieste, Italy, in February 1995
-to meet with G\'eza Seriani and Enrico Priolo, who introduced him to their 2D Chebyshev version of the spectral-element method with a non-diagonal mass matrix.
-We are deeply indebted and thankful to them for that.
-
-%%%%%%%% please do NOT merge the lines of the paragraph below, it makes it difficult to insert new names in alphabetical order; leave a single name per line
-%%%%%%%% please do NOT merge the lines of the paragraph below, it makes it difficult to insert new names in alphabetical order; leave a single name per line
-%%%%%%%% please do NOT merge the lines of the paragraph below, it makes it difficult to insert new names in alphabetical order; leave a single name per line
-Since then it has been developed and maintained by a development team:
-in alphabetical order,
-Jean-Paul (Pablo) Ampuero,
-Kangchen Bai,
-Piero Basini,
-C\'eline Blitz,
-Alexis Bottero,
-Ebru Bozda\u{g},
-Emanuele Casarotti,
-Joseph Charles,
-Min Chen,
-Paul Cristini,
-Percy Galvez,
-Dominik G\"oddeke,
-Vala Hj\"orleifsd\'ottir,
-Sue Kientz,
-Dimitri Komatitsch,
-Jes\'us Labarta,
-Nicolas Le Goff,
-Pieyre Le Loher,
-Matthieu Lefebvre,
-Qinya Liu,
-Yang Luo,
-Alessia Maggi,
-Federica Magnoni,
-Roland Martin,
-Ren\'e Matzen,
-Dennis McRitchie,
-Matthias Meschede,
-Peter Messmer,
-David Mich\'ea,
-Surendra Nadh Somala,
-Tarje Nissen-Meyer,
-Daniel Peter,
-Max Rietmann,
-Elliott Sales de Andrade,
-Brian Savage,
-Bernhard Schuberth,
-Anne Sieminski,
-Leif Strand,
-Carl Tape,
-Jeroen Tromp,
-Jean-Pierre Vilotte,
-Zhinan Xie,
-Chang-Hua Zhang,
-Hejun Zhu.\\
-
-The cover graphic of the manual was created by Santiago Lombeyda from
-Caltech's Center for Advanced Computing Research (CACR) \urlwithparentheses{http://www.cacr.caltech.edu}.\\
-
-The code is released open-source under the GNU version 2+ license, see the license at the end of this manual.\\
-
-\newpage{}
-\section*{Current and past main participants or main sponsors of the SPECFEM project (in no particular order)}
-%
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=0.125\textwidth]{figures/logo_cnrs}\vspace*{2truemm}
-\includegraphics[width=0.125\textwidth]{figures/logo_princeton}\vspace*{2truemm}
-\includegraphics[width=0.15\textwidth]{figures/logo_aix_marseille_universite}\vspace*{0.02truemm}
-\includegraphics[width=0.15\textwidth]{figures/logo_ETH}\vspace*{2truemm}
-\includegraphics[width=0.12\textwidth]{figures/logo_CSC_China}\vspace*{0.02truemm}
-\includegraphics[width=0.15\textwidth]{figures/logo_inria}\vspace*{2truemm}
-\includegraphics[width=0.14\textwidth]{figures/logo_UPPA}
-\par\end{centering}
-%
-\vspace*{-2truemm}
-%
-\noindent \begin{centering}
-\includegraphics[width=0.15\textwidth]{figures/logo_NVIDIA}\vspace*{2truemm}
-\includegraphics[width=0.125\textwidth]{figures/logo_IUF}\vspace*{2truemm}
-\includegraphics[width=0.135\textwidth]{figures/logo_Caltech}\vspace*{2truemm}
-\includegraphics[width=0.135\textwidth]{figures/logo_Harvard}\vspace*{2truemm}
-\includegraphics[width=0.135\textwidth]{figures/logo_IPGP}\vspace*{2truemm}
-\includegraphics[width=0.112\textwidth]{figures/logo_ANR}\vspace*{2truemm}
-\includegraphics[width=0.112\textwidth]{figures/logo_NSF}\vspace*{2truemm}
-\par\end{centering}
-%
-\vspace*{-2truemm}
-%
-\noindent \begin{centering}
-\includegraphics[width=0.190\textwidth]{figures/logo_European_Union}\vspace*{2truemm}
-\includegraphics[width=0.112\textwidth]{figures/logo_GENCI}\vspace*{2truemm}
-\includegraphics[width=0.112\textwidth]{figures/logo_PRACE}\vspace*{2truemm}
-\includegraphics[width=0.112\textwidth]{figures/logo_CINES}\vspace*{2truemm}
-\includegraphics[width=0.130\textwidth]{figures/logo_Oak_Ridge}\vspace*{2truemm}
-\hspace*{3mm}\includegraphics[width=0.130\textwidth]{figures/logo_fondation_Del_Duca}
-\includegraphics[width=0.130\textwidth]{figures/logo_CIG}\vspace*{2truemm}
-\par\end{centering}
-%
-\vspace*{-2truemm}
-%
-\noindent \begin{centering}
-\includegraphics[width=0.112\textwidth]{figures/logo_University_of_Toronto}\vspace*{2truemm}
-\includegraphics[width=0.109\textwidth]{figures/logo_INGV}\vspace*{2truemm}
-\includegraphics[width=0.102\textwidth]{figures/logo_Univ_Toulouse}\vspace*{2truemm}
-\includegraphics[width=0.112\textwidth]{figures/logo_TOTAL}\vspace*{2truemm}
-\includegraphics[width=0.130\textwidth]{figures/logo_Fairbanks}\vspace*{2truemm}
-\includegraphics[width=0.120\textwidth]{figures/logo_CINECA}\vspace*{2truemm}
-\includegraphics[width=0.140\textwidth]{figures/logo_Intel_Exascale_Labs}\vspace*{2truemm}
-\includegraphics[width=0.130\textwidth]{figures/logo_Maison_Simulation}
-\par\end{centering}
-\end{figure}
-
-\newpage{}
-
-\tableofcontents{}
-
-\chapter{Introduction}
-
-The software package SPECFEM3D Cartesian simulates seismic wave propagation
-at the local or regional scale and performs full waveform imaging (FWI) or adjoint tomography based upon the spectral-element method
-(SEM). The SEM is a continuous Galerkin technique \citep{TrKoLi08,PeKoLuMaLeCaLeMaLiBlNiBaTr11},
-which can easily be made discontinuous \citep{BeMaPa94,Ch00,KoWoHu02,ChCaVi03,LaWaBe05,Kop06,WiStBuGh10,AcKo11};
-it is then close to a particular case of the discontinuous Galerkin
-technique \citep{ReHi73,LeRa74,Arn82,JoPi86,BoMaHe91,FaRi99,HuHuRa99,CoKaSh00,GiHeWa02,RiWh03,MoRi05,GrScSc06,AiMoMu06,BeLaPi06,DuKa06,DeSeWh08,PuAmKa09,WiStBuGh10,DeSe10,EtChViGl10},
-with optimized efficiency because of its tensorized basis functions
-\citep{WiStBuGh10,AcKo11}. In particular, it can accurately handle
-very distorted mesh elements \citep{OlSe11}.\\
-
-It has very good accuracy and convergence properties \citep{MaPa89,SePr94,DeFiMu02,Coh02,DeSe07,SeOl08,AiWa09,AiWa10,MeStTh12}.
-The spectral element approach admits spectral rates of convergence
-and allows exploiting $hp$-convergence schemes. It is also very well
-suited to parallel implementation on very large supercomputers \citep{KoTsChTr03,TsKoChTr03,KoLaMi08a,CaKoLaTiMiLeSnTr08,KoViCh10}
-as well as on clusters of GPU accelerating graphics cards \citep{Kom11,MiKo10,KoMiEr09,KoErGoMi10}.
-Tensor products inside each element can be optimized to reach very
-high efficiency \citep{DeFiMu02}, and mesh point and element numbering
-can be optimized to reduce processor cache misses and improve cache
-reuse \citep{KoLaMi08a}. The SEM can also handle triangular (in 2D)
-or tetrahedral (in 3D) elements \citep{WinBoyd96,TaWi00,KoMaTrTaWi01,Coh02,MeViSa06}
-as well as mixed meshes, although with increased cost and reduced
-accuracy in these elements, as in the discontinuous Galerkin method.\\
-
-Note that in many geological models in the context of seismic wave
-propagation studies (except for instance for fault dynamic rupture
-studies, in which very high frequencies or supershear rupture need
-to be modeled near the fault, see e.g. \citet{BeGlCrViPi07,BeGlCrVi09,PuAmKa09,TaCrEtViBeSa10})
-a continuous formulation is sufficient because material property contrasts
-are not drastic and thus conforming mesh doubling bricks can efficiently
-handle mesh size variations \citep{KoTr02a,KoLiTrSuStSh04,LeChLiKoHuTr08,LeChKoHuTr09,LeKoHuTr09}.\\
-
-For a detailed introduction to the SEM as applied to regional seismic
-wave propagation, please consult \citet{PeKoLuMaLeCaLeMaLiBlNiBaTr11,TrKoLi08,KoVi98,KoTr99,ChKoViCaVaFe07}
-and in particular \citet{LeKoHuTr09,LeChKoHuTr09,LeChLiKoHuTr08,GoAmTaCaSmSaMaKo09,WiKoScTr04,KoLiTrSuStSh04}.
-A detailed theoretical analysis of the dispersion
-and stability properties of the SEM is available in \citet{Coh02}, \citet{DeSe07}, \citet{SeOl07}, \citet{SeOl08} and \citet{MeStTh12}.\\
-
-Effects due to lateral variations in compressional-wave speed, shear-wave
-speed, density, a 3D crustal model, topography and bathymetry are
-included. The package can accommodate full 21-parameter anisotropy
-(see~\citet{ChTr07}) as well as lateral variations in attenuation
-\citep{SaKoTr10}. Adjoint capabilities and finite-frequency kernel
-simulations are included \citep{TrKoLi08,PeKoLuMaLeCaLeMaLiBlNiBaTr11,LiTr06,FiIgBuKe09,ViOp09}.\\
-
-The SEM was originally developed in computational fluid dynamics \citep{Pat84,MaPa89}
-and has been successfully adapted to address problems in seismic wave
-propagation. Early seismic wave propagation applications of the SEM,
-utilizing Legendre basis functions and a perfectly diagonal mass matrix,
-include \citet{CoJoTo93}, \citet{Kom97}, \citet{FaMaPaQu97}, \citet{CaGa97},
-\citet{KoVi98} and \citet{KoTr99}, whereas applications involving
-Chebyshev basis functions and a nondiagonal mass matrix include \citet{SePr94},
-\citet{PrCaSe94} and \citet{SePrPr95}.\\
-
-All SPECFEM3D software is written in Fortran2003 with full portability
-in mind, and conforms strictly to the Fortran2003 standard. It uses
-no obsolete or obsolescent features of Fortran. The package uses parallel
-programming based upon the Message Passing Interface (MPI) \citep{GrLuSk94,Pac97}.\\
-
-SPECFEM3D won the Gordon Bell award for best performance at the SuperComputing~2003
-conference in Phoenix, Arizona (USA) (see \citet{KoTsChTr03} and
-\url{www.sc-conference.org/sc2003/nrfinalaward.html}). It was a finalist
-again in 2008 for a run at 0.16 petaflops (sustained) on 149,784 processors
-of the `Jaguar' Cray XT5 system at Oak Ridge National Laboratories
-(USA) \citep{CaKoLaTiMiLeSnTr08}. It also won the BULL Joseph Fourier
-supercomputing award in 2010.\\
-
-It reached the sustained one petaflop performance level for the first time in February 2013
-on the Blue Waters Cray supercomputer at the National Center for Supercomputing Applications (NCSA), located at the University of Illinois at Urbana-Champaign (USA).\\
-
-This new release of the code includes Convolution or
-Auxiliary Differential Equation Perfectly Matched absorbing Layers
-(C-PML or ADE-PML) \citep{MaKoEz08,MaKoGe08,MaKo09,MaKoGeBr10,KoMa07}.
-It also includes support for GPU graphics card acceleration \citep{Kom11,MiKo10,KoMiEr09,KoErGoMi10}.\\
-
-The next release of the code will use the PT-SCOTCH parallel and threaded
-version of SCOTCH for mesh partitioning instead of the serial version.\\
-
-SPECFEM3D Cartesian includes coupled fluid-solid domains and adjoint
-capabilities, which enables one to address seismological inverse problems,
-but for linear rheologies only so far. To accommodate visco-plastic
-or non-linear rheologies, the reader can refer to the \href{http://geoelse.stru.polimi.it/}{GeoELSE}software
-package \citep{CaGa97,StPaIg09}.
-
-\section{Citation}
-
-If you use SPECFEM3D Cartesian for your own research, please cite
-at least one of the following articles:
-\begin{description}
-\item [{Numerical simulations in general}] ~\\
- Forward and adjoint simulations are described in detail in \citet{TrKoLi08,PeKoLuMaLeCaLeMaLiBlNiBaTr11,VaCaSaKoVi99,KoMiEr09,KoErGoMi10,ChKoViCaVaFe07,MaKoDi09,KoViCh10,CaKoLaTiMiLeSnTr08,TrKoHjLiZhPeBoMcFrTrHu10,KoRiTr02,KoTr02a,KoTr02b,KoTr99}
-or \citet{KoVi98}. Additional aspects of adjoint simulations are
-described in \citet{TrTaLi05,LiTr06,TrKoLi08,LiTr08,TrKoHjLiZhPeBoMcFrTrHu10,PeKoLuMaLeCaLeMaLiBlNiBaTr11}.
-Domain decomposition is explained in detail in \citet{MaKoBlLe08},
-and excellent scaling up to 150,000 processor cores is shown for instance
-in \citet{CaKoLaTiMiLeSnTr08,KoLaMi08a,MaKoBlLe08,KoErGoMi10,Kom11},
-\item [{GPU computing}] ~\\
- Computing on GPU graphics cards for acoustic or seismic wave propagation
-applications is described in detail in \citet{Kom11,MiKo10,KoMiEr09,KoErGoMi10}.
-\end{description}
-\noindent If you use this new version, which has non blocking
-MPI for much better performance for medium or large runs, please cite
-at least one of these six articles, in which results of non blocking
-MPI runs are presented: \citet{PeKoLuMaLeCaLeMaLiBlNiBaTr11,KoErGoMi10,KoViCh10,Kom11,CaKoLaTiMiLeSnTr08,MaKoBlLe08}.\\
-
-\noindent If you work on geophysical applications, you may be interested
-in citing some of these application articles as well, among others:
-\begin{description}
-\item [{Southern California simulations}] ~\\
- \citet{KoLiTrSuStSh04,KrJiKoTr06a,KrJiKoTr06b}.
-
-If you use the 3D southern California model, please cite \citet{SuSh03}
-(Los Angeles model), \citet{lovelyetal06} (Salton Trough), and \citet{hauksson2000}
-(southern California). The Moho map was determined by \citet{zhu&kanamori2000}.
-The 1D SoCal model was developed by \citet{DrHe90}.
-
-\item [{Anisotropy}] ~\\
- \citet{ChTr07,JiTsKoTr05,ChFaKo04,FaChKo04,RiRiKoTrHe02,TrKo00}.
-\item [{Attenuation}] ~\\
- \citet{SaKoTr10,KoTr02a,KoTr99}.
-\item [{Topography}] ~\\
- \citet{LeKoHuTr09,LeChKoHuTr09,LeChLiKoHuTr08,GoAmTaCaSmSaMaKo09,WiKoScTr04}.
-\end{description}
-The corresponding Bib\TeX{} entries may be found in file \texttt{doc/USER\_MANUAL/bibliography.bib}.
-
-\section{Support}
-
-This material is based upon work supported by the USA National Science
-Foundation under Grants No. EAR-0406751 and EAR-0711177, by the French
-CNRS, French INRIA Sud-Ouest MAGIQUE-3D, French ANR NUMASIS under
-Grant No. ANR-05-CIGC-002, and European FP6 Marie Curie International
-Reintegration Grant No. MIRG-CT-2005-017461. Any opinions, findings,
-and conclusions or recommendations expressed in this material are
-those of the authors and do not necessarily reflect the views of the
-USA National Science Foundation, CNRS, INRIA, ANR or the European
-Marie Curie program.
-
-\chapter{\label{cha:Getting-Started}Getting Started}
-
-To download the SPECFEM3D\_Cartesian software package, type this:
-\begin{lyxcode}
-git clone --recursive --branch devel https://github.com/geodynamics/specfem3d.git
-\end{lyxcode}
-{\small \par}
-
-We recommend that you add {\texttt{ulimit -S -s unlimited}} to your
-{\texttt{.bash\_profile}} file and/or {\texttt{limit stacksize
-unlimited }} to your {\texttt{.cshrc}} file to suppress any potential
-limit to the size of the Unix stack.\\
-
-Then, to configure the software for your system, run the \texttt{configure}
-shell script. This script will attempt to guess the appropriate configuration
-values for your system. However, at a minimum, it is recommended that
-you explicitly specify the appropriate command names for your Fortran
-compiler and MPI package (another option is to define FC, CC and MPIF90
-in your .bash\_profile or your .cshrc file):
-\begin{lyxcode}
-./configure~FC=ifort~MPIFC=mpif90
-\end{lyxcode}
-Note that MPI must be installed with MPI-IO enabled because parts of SPECFEM3D perform I/Os through MPI-IO.
-
-Before running the \texttt{configure} script, you should probably
-edit file \texttt{flags.guess} to make sure that it contains the best
-compiler options for your system. Known issues or things to check
-are:
-\begin{description}
-\item [{\texttt{Intel ifort compiler}}] See if you need to add \texttt{-assume byterecl} for your machine.
-\textbf{In the case of that compiler, we have noticed that versions dot zero sometimes have bugs or issues that can lead to wrong results when running the code, thus we \emph{strongly} recommend using versions dot one or above (for instance version 13.1 instead of 13.0, version 14.1 instead of 14.0 and so on)}.
-\item [{\texttt{IBM compiler}}] See if you need to add \texttt{-qsave}
-or \texttt{-qnosave} for your machine.
-\item [{\texttt{Mac OS}}] You will probably need to install \texttt{XCODE}.
-\end{description}
-When compiling on an IBM machine with the \texttt{xlf} and \texttt{xlc}
-compilers, we suggest running the \texttt{configure} script with the
-following options:\\
-
-\noindent \texttt{./configure FC=xlf90\_r MPIFC=mpif90 CC=xlc\_r CFLAGS=\textquotedbl{}-O3
--q64\textquotedbl{} FCFLAGS=\textquotedbl{}-O3 -q64\textquotedbl{}
--with-scotch-dir=...}.\\
-
-On SGI systems, \texttt{flags.guess} automatically informs \texttt{configure}
-to insert `\texttt{`TRAP\_FPE=OFF}'' into the generated \texttt{Makefile}
-in order to turn underflow trapping off.\\
-
-You can add \texttt{--enable-vectorization} to the configuration options to speed up the code in the fluid (acoustic) and elastic parts.
-This works fine if (and only if) your computer always allocates a contiguous memory block for each allocatable array;
-this is the case for most machines and most compilers, but not all. To disable this feature, use option \texttt{--disable-vectorization}.
-For more details see \href{https://github.com/geodynamics/specfem3d/issues/81}{github.com/geodynamics/specfem3d/issues/81} .
-To check if that option works fine on your machine, run the code with and without it for an acoustic/elastic model and make sure the seismograms are identical.\\
-
-Note that we use CUBIT (now called Trelis) to create meshes of hexahedra, but other packages
-can be used as well, for instance GiD from \url{http://gid.cimne.upc.es}
-or Gmsh from \url{http://geuz.org/gmsh} \citep{GeRe09}. Even mesh
-creation packages that generate tetrahedra, for instance TetGen from
-\url{http://tetgen.berlios.de}, can be used because each tetrahedron
-can then easily be decomposed into four hexahedra as shown in the
-picture of the TetGen logo at \url{http://tetgen.berlios.de/figs/Delaunay-Voronoi-3D.gif};
-while this approach does not generate hexahedra of optimal quality,
-it can ease mesh creation in some situations and it has been shown
-that the spectral-element method can very accurately handle distorted
-mesh elements \citep{OlSe11}.
-
-The SPECFEM3D Cartesian software package relies on the SCOTCH library
-to partition meshes created with CUBIT. METIS \citep{KaKu98a,KaKu98c,KaKu98b}
-can also be used instead of SCOTCH if you prefer, by editing file
-src/decompose\_mesh/decompose\_mesh.F90 and uncommenting the flag
-USE\_METIS\_INSTEAD\_OF\_SCOTCH. You will also then need to install
-and compile Metis version 4.0 (do {*}NOT{*} install Metis version
-5.0, which has incompatible function calls) and edit \texttt{src/decompose\_mesh/Makefile.in}
-and uncomment the METIS link flag in that file before running \texttt{configure}.\\
-
-The SCOTCH library \citep{PeRo96} provides efficent static mapping,
-graph and mesh partitioning routines. SCOTCH is a free software package
-developed by Fran\c{c}ois Pellegrini et al. from LaBRI and INRIA in Bordeaux,
-France, downloadable from the web page \url{https://gforge.inria.fr/projects/scotch/}.
-In case no SCOTCH libraries can be found on the system, the configuration
-will bundle the version provided with the source code for compilation.
-The path to an existing SCOTCH installation can to be set explicitly
-with the option \texttt{-{}-with-scotch-dir}. Just as an example:
-\begin{lyxcode}
-./configure~FC=ifort~MPIFC=mpif90~-{}-with-scotch-dir=/opt/scotch
-\end{lyxcode}
-If you use the Intel ifort compiler to compile the code, we recommend
-that you use the Intel icc C compiler to compile Scotch, i.e., use:
-\begin{lyxcode}
-./configure~CC=icc~FC=ifort~MPIFC=mpif90
-\end{lyxcode}
-
-When compiling the SCOTCH source code, if you get a message such as: "ld: cannot find -lz",
-the Zlib compression development library is probably missing on your machine and you will need to install it or ask your system administrator to
-do so. On Linux machines the package is often called "zlib1g-dev" or similar. (thus "sudo apt-get install zlib1g-dev" would install it)
-
-To compile a serial version of the code for small meshes that fits
-on one compute node and can therefore be run serially, run \texttt{configure}
-with the \texttt{-{}-without-mpi} option to suppress all calls to
-MPI.
-
-A summary of the most important configuration variables follows.
-\begin{description}
-\item [{\texttt{F90}}] Path to the Fortran compiler.
-\item [{\texttt{MPIF90}}] Path to MPI Fortran.
-\item [{\texttt{MPI\_FLAGS}}] Some systems require this flag to link to
-MPI libraries.
-\item [{\texttt{FLAGS\_CHECK}}] Compiler flags.
-\end{description}
-The configuration script automatically creates for each executable
-a corresponding \texttt{Makefile} in the \texttt{src/} subdirectoy.
-The \texttt{Makefile} contains a number of suggested entries for various
-compilers, e.g., Portland, Intel, Absoft, NAG, and Lahey. The software
-has run on a wide variety of compute platforms, e.g., various PC clusters
-and machines from Sun, SGI, IBM, Compaq, and NEC. Select the compiler
-you wish to use on your system and choose the related optimization
-flags. Note that the default flags in the \texttt{Makefile} are undoubtedly
-not optimal for your system, so we encourage you to experiment with
-these flags and to solicit advice from your systems administrator.
-Selecting the right compiler and optimization flags can make a tremendous
-difference in terms of performance. We welcome feedback on your experience
-with various compilers and flags.
-
-Now that you have set the compiler information, you need to select
-a number of flags in the \texttt{constants.h} file depending on your
-system:
-\begin{description}
-\item [{\texttt{LOCAL\_PATH\_IS\_ALSO\_GLOBAL}}] Set to \texttt{.false.}
-on most cluster applications. For reasons of speed, the (parallel)
-distributed database generator typically writes a (parallel) database
-for the solver on the local disks of the compute nodes. Some systems
-have no local disks, e.g., BlueGene or the Earth Simulator, and other
-systems have a fast parallel file system, in which case this flag
-should be set to \texttt{.true.}. Note that this flag is not used
-by the database generator or the solver; it is only used for some
-of the post-processing.
-\end{description}
-The package can run either in single or in double precision mode.
-The default is single precision because for almost all calculations
-performed using the spectral-element method using single precision
-is sufficient and gives the same results (i.e. the same seismograms);
-and the single precision code is faster and requires exactly half
-as much memory. Select your preference by selecting the appropriate
-setting in the \texttt{constants.h} file:
-\begin{description}
-\item [{\texttt{CUSTOM\_REAL}}] Set to \texttt{SIZE\_REAL} for single precision
-and \texttt{SIZE\_DOUBLE} for double precision.
-\end{description}
-In the \texttt{precision.h} file:
-\begin{description}
-\item [{\texttt{CUSTOM\_MPI\_TYPE}}] Set to \texttt{MPI\_REAL} for single
-precision and \texttt{MPI\_DOUBLE\_PRECISION} for double precision.
-\end{description}
-On many current processors (e.g., Intel, AMD, IBM Power), single precision
-calculations are significantly faster; the difference can typically
-be 10\% to 25\%. It is therefore better to use single precision. What
-you can do once for the physical problem you want to study is run
-the same calculation in single precision and in double precision on
-your system and compare the seismograms. If they are identical (and
-in most cases they will), you can select single precision for your
-future runs.\\
-
-If your compiler has problems with the \texttt{use mpi} statements that are used in the code, use the script called
-\texttt{replace\_use\_mpi\_with\_include\_mpif\_dot\_h.pl} in the root directory to replace all of them with \texttt{include 'mpif.h'} automatically.
-
-\section{Adding OpenMP support in addition to MPI}
-
-NOTE FROM JULY 2013: OpenMP support is maybe / probably not maintained any more.
-Thus the section below is maybe obsolete.
-
-OpenMP support can be enabled in addition to MPI. However, in many
-cases performance will not improve because our pure MPI implementation
-is already heavily optimized and thus the resulting code will in fact
-be slightly slower. A possible exception could be IBM BlueGene-type
-architectures.
-
-To enable OpenMP, uncomment the OpenMP compiler option in two lines
-in file \texttt{src/specfem3D/Makefile.in} (before running \texttt{configure})
-and also uncomment the \texttt{\#define USE\_OPENMP} statement in
-file\\
- \texttt{src/specfem3D/specfem3D.F90}.
-
-The DO-loop using OpenMP threads has a SCHEDULE property. The OMP\_SCHEDULE
-environment variable can set the scheduling policy of that DO-loop.
-Tests performed by Marcin Zielinski at SARA (The Netherlands) showed
-that often the best scheduling policy is DYNAMIC with the size of
-the chunk equal to the number of OpenMP threads, but most preferably
-being twice as the number of OpenMP threads (thus chunk size = 8 for
-4 OpenMP threads etc). If OMP\_SCHEDULE is not set or is empty, the
-DO-loop will assume generic scheduling policy, which will slow down
-the job quite a bit.
-
-\section{Compiling on an IBM BlueGene}
-
-\underline{More recent installation instruction for IBM BlueGene, from April 2013:}\\
-
-\noindent Edit file \texttt{flags.guess} and put this for \texttt{FLAGS\_CHECK}:\\
-
-\noindent \texttt{-g -qfullpath -O2 -qsave -qstrict -qtune=qp -qarch=qp
--qcache=auto -qhalt=w}\\
- \texttt{-qfree=f90 -qsuffix=f=f90 -qlanglvl=95pure -Q -Q+rank,swap\_all
--Wl,-relax}\\
-
-\noindent The most relevant are the -qarch and -qtune flags, otherwise
-if these flags are set to ``auto'' then they are wrongly assigned
-to the architecture of the frond-end node, which is different from
-that on the compute nodes. You will need to set these flags to the
-right architecture for your BlueGene compute nodes, which is not necessarily
-``qp''; ask your system administrator. On some machines if is necessary
-to use -O2 in these flags instead of -O3 due to a compiler bug of
-the XLF version installed. We thus suggest to first try -O3, and then
-if the code does not compile or does not run fine then switch back
-to -O2. The debug flags (-g, -qfullpath) do not influence performance
-but are useful to get at least some insights in case of problems.\\
-
-\noindent Before running \texttt{configure}, select the XL Fortran
-compiler by typing \texttt{module load bgq-xl/1.0} or \texttt{module
-load bgq-xl} (another, less efficient option is to load the GNU compilers
-using \texttt{module load bgq-gnu/4.4.6} or similar).\\
-
-\noindent Then, to configure the code, type this:\\
-
-\noindent \texttt{./configure FC=bgxlf90\_r MPIFC=mpixlf90\_r CC=bgxlc\_r
-LOCAL\_PATH\_IS\_ALSO\_GLOBAL=true}\\
-
-In order for the SCOTCH domain decomposer to compile, on some (but
-not all) Blue Gene systems you may need to run \texttt{configure}
-with \texttt{CC=gcc} instead of \texttt{CC=bgxlc\_r}.
-
-\noindent \underline{Older installation instruction for IBM BlueGene, from 2011:}\\
-
-\noindent To compile the code on an IBM BlueGene, Laurent L\'eger from
-IDRIS, France, suggests the following: compile the code with
-
-\noindent \texttt{ FLAGS\_CHECK=\textquotedbl{}-O3 -qsave -qstrict
--qtune=auto -qarch=450d -qcache=auto }~\\
- \texttt{ -qfree=f90 -qsuffix=f=f90 -g -qlanglvl=95pure -qhalt=w -Q
--Q+rank,swap\_all -Wl,-relax\textquotedbl{}}
-
-\noindent Option \textquotedbl{}-Wl,-relax\textquotedbl{} must be
-added on many (but not all) BlueGene systems to be able to link the
-binaries \texttt{xmeshfem3D} and \texttt{xspecfem3D} because the final
-link step is done by the GNU \texttt{ld} linker even if one uses \texttt{FC=bgxlf90\_r,
-MPIFC=mpixlf90\_r} and \texttt{CC=bgxlc\_r} to create all the object
-files. On the contrary, on some BlueGene systems that use the native
-AIX linker option \textquotedbl{}-Wl,-relax\textquotedbl{} can lead
-to problems and must be suppressed from \texttt{flags.guess}.
-
-\noindent Also, \texttt{AR=ar, ARFLAGS=cru} and \texttt{RANLIB=ranlib}
-are hardwired in all \texttt{Makefile.in} files by default, but to
-cross-compile on BlueGene/P one needs to change these values to \texttt{AR=bgar,
-ARFLAGS=cru} and \texttt{RANLIB=bgranlib}. Thus the easiest thing
-to do is to modify all \texttt{Makefile.in} files and the \texttt{configure}
-script to set them automatically by \texttt{configure}. One then just
-needs to pass the right commands to the \texttt{configure} script:
-
-\noindent \texttt{ ./configure -{}-prefix=/path/to/SPECFEM3DG\_SP
--{}-host=Babel -{}-build=BGP }~\\
- \texttt{ FC=bgxlf90\_r MPIFC=mpixlf90\_r CC=bgxlc\_r AR=bgar ARFLAGS=cru
-}~\\
- \texttt{ RANLIB=bgranlib LOCAL\_PATH\_IS\_ALSO\_GLOBAL=false}
-
-\noindent This trick can be useful for all hosts on which one needs
-to cross-compile.
-
-\noindent On BlueGene, one also needs to run the \texttt{xcreate\_header\_file}
-binary file manually rather than in the Makefile:
-
-\noindent \texttt{bgrun -np 1 -mode VN -exe ./bin/xcreate\_header\_file}
-
-\section{Visualizing the subroutine calling tree of the source code}
-
-\noindent
-Packages such as \texttt{Doxywizard} can be used to visualize the calling tree of the
-subroutines of the  source code. \texttt{Doxywizard} is
-a GUI front-end for configuring and running \texttt{Doxygen}.
-
-\bigskip
-
-\noindent To do your own call graphs, you can follow these simple steps below.
-
-\begin{enumerate}
-\setcounter{enumi}{-1}
-\item Install \texttt{Doxygen} \underline{and} \texttt{graphviz} (the two are usually in the package manager of classic Linux distribution).
-\item Run in the terminal : \texttt{doxygen -g}, which creates a \texttt{Doxyfile} that tells doxygen what you want it to do.
-\item Edit the Doxyfile. Two Doxyfile-type files have been already committed in the directory \\
-\noindent \texttt{specfem3d/doc/Call\_trees}:
-
-\begin{itemize}
-\item[\textbullet] \texttt{Doxyfile\_truncated\_call\_tree} will generate call graphs with maximum 3 or 4 levels of tree structure,
-\item[\textbullet] \texttt{Doxyfile\_complete\_call\_tree} will generate call graphs with complete tree structure.
-\end{itemize}
-
-The important entries in the Doxyfile are:
-
-\begin{description}
-\item [{\texttt{PROJECT\_NAME}}]
-\item [{\texttt{OPTIMIZE\_FOR\_FORTRAN}}] Set to YES
-\item [{\texttt{EXTRACT\_ALL}}] Set to YES
-\item [{\texttt{EXTRACT\_PRIVATE}}] Set to YES
-\item [{\texttt{EXTRACT\_STATIC}}] Set to YES
-\item [{\texttt{INPUT}}] From the directory \texttt{specfem3d/doc/Call\_trees}, it is \texttt{"../../src/"}
-\item [{\texttt{FILE\_PATTERNS}}] In SPECFEM case, it is \texttt{*.f90* *.F90* *.c* *.cu* *.h*}
-\item [{\texttt{HAVE\_DOT}}] Set to YES
-\item [{\texttt{CALL\_GRAPH}}] Set to YES
-\item [{\texttt{CALLER\_GRAPH}}] Set to YES
-\item [{\texttt{DOT\_PATH}}] The path where is located the dot program graphviz (if it is not in your \$PATH)
-\item [{\texttt{RECURSIVE}}] This tag can be used to turn specify whether or not subdirectories should be searched for input files as well. In the case of SPECFEM, set to YES.
-\item [{\texttt{EXCLUDE}}] Here, you can exclude:  \\
-\texttt{"../../src/specfem3D/older\_not\_maintained\_partial\_OpenMP\_port"}\\
-\texttt{"../../src/decompose\_mesh/scotch"}\\
-\texttt{"../../src/decompose\_mesh/scotch\_5.1.12b"}
-\item [{\texttt{DOT\_GRAPH\_MAX\_NODES}}] to set the maximum number of nodes that will be shown in the graph.
-If the number of nodes in a graph becomes larger than this value, doxygen will truncate the graph, which is visualized by representing a node as a red box.
-Minimum value: 0, maximum value: 10000, default value: 50.
-\item [{\texttt{MAX\_DOT\_GRAPH\_DEPTH}}] to set the maximum depth of the graphs generated by dot.
-A depth value of 3 means that only nodes reachable from the root by following a path via at most 3 edges will be shown.
-Using a depth of 0 means no depth restriction.
-Minimum value: 0, maximum value: 1000, default value: 0.
-
-\end{description}
-
-\item Run : \texttt{doxygen Doxyfile}, HTML and LaTeX files created by default in \texttt{html} and \texttt{latex} subdirectories.
-
-\item To see the call trees, you have to open the file \texttt{html/index.html} in your \underline{browser}.
-You will have many informations about each subroutines of SPECFEM (not only call graphs), you can click on every boxes / subroutines.
-It show you the call, and, the caller graph of each subroutine : the subroutines called by the concerned subroutine, and the previous subroutines who call this subroutine (the previous path), respectively.
-In the case of a truncated calling tree, the boxes with a red border indicates a node that has \underline{more} arrows than are shown (in other words: the graph is truncated with respect to this node).
-
-\end{enumerate}
-
-\medskip
-
-\noindent Finally, some useful links:
-
-\begin{itemize}
-\item[\textbullet] a good and short summary for the basic utilisation of Doxygen:
-
-\url{http://www.softeng.rl.ac.uk/blog/2010/jan/30/callgraph-fortran-doxygen/},
-
-\item[\textbullet] to configure the diagrams :
-
-\url{http://www.stack.nl/~dimitri/doxygen/manual/diagrams.html},
-
-\item[\textbullet] the complete alphabetical index of the tags in Doxyfile:
-
-\url{http://www.stack.nl/~dimitri/doxygen/manual/config.html},
-
-\item[\textbullet] more generally, the Doxygen manual:
-
-\url{http://www.stack.nl/~dimitri/doxygen/manual/index.html}.
-
-\end{itemize}
-
-\medskip
-
-\section{Using the ADIOS library for I/O}
-Regular POSIX I/O can be problematic when dealing with large simulations one large
-clusters (typically more than $10,000$ processes). SPECFEM3D use the ADIOS library~\cite{Liu2013}
-to deal transparently take advantage of advanced paralllel file system features. To enable
-ADIOS, the following steps should be done:
-\begin{enumerate}
-\item Install ADIOS (available from \url{https://www.olcf.ornl.gov/center-projects/adios/}).
-Make sure that your environment variables reference it.
-\item You may want to change ADIOS related values in the \texttt{constants.h.in} file.
-The default values probably suit most cases.
-\item Configure using the \texttt{--with-adios} flag.
-\end{enumerate}
-ADIOS is currently only usable for meshfem3D generated mesh (i.e. not for meshes generated
-with CUBTI). Additional control parameters are discussed in section~\ref{cha:Main-Parameter}.
-
-\section{Becoming a developer of the code, or making small modifications in the source code}
-
-If you want to develop new features in the code, and/or if you want to make small changes, improvements, or bug fixes, you are very welcome to contribute.
-To do so, i.e.
-to access the development branch of the source code with read/write access (in a safe way, no need to worry too much about breaking the package, there is a robot called BuildBot that is in charge of checking and validating all new contributions and changes), please visit this Web page: \url{https://github.com/geodynamics/specfem3d/wiki/Using-Hub}.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\chapter{\label{cha:Mesh-Generation}Mesh Generation}
-
-The first step in running a spectral-element simulation consists of
-constructing a high-quality mesh for the region under consideration.
-We provide two possibilities to do so: (1) relying on the external,
-hexahedral mesher CUBIT, or (2) using the provided, internal mesher
-\texttt{xmeshfem3D}. In the following, we explain these two approaches.
-
-\section{\label{cha:Running-the-Mesher-CUBIT}Meshing with \texttt{CUBIT}}
-
-CUBIT is a meshing tool suite for the creation of finite-element meshes
-for arbitrarily shaped models. It has been developed and maintained
-at Sandia National Laboratories and can be purchased for a small academic
-institutional fee at \url{http://cubit.sandia.gov}. Our experience
-showed that using CUBIT greatly facilitates and speeds up the generation
-and preparation of hexahedral, conforming meshes for a variety of
-geophysical models with increasing complexity.
-
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=0.6\textwidth]{figures/mount-cubit}
-\par\end{centering}
-
-\caption{Example of the graphical user interface of CUBIT. The hexahedral mesh
-shown in the main display consists of a hexahedral discretization
-of a single volume with topography.}
-
-\label{fig:mount.cubit}
-\end{figure}
-
-\noindent
-The basic steps in creating a load-balanced, partitioned mesh with
-CUBIT are:
-\begin{description}
-\item [{1.}] setting up a hexahedral mesh with CUBIT,
-\item [{2.}] exporting the CUBIT mesh into a SPECFEM3D Cartesian file format
-and
-\item [{3.}] partitioning the SPECFEM3D Cartesian mesh files for a chosen
-number of cores.
-\end{description}
-Examples are provided in the SPECFEM3D Cartesian package in the subdirectory
-\texttt{EXAMPLES/}. We strongly encourage you to contribute your own
-example to this package by contacting the CIG Computational Seismology
-Mailing List \urlwithparentheses{cig-seismo@geodynamics.org}.
-
-\subsection{Creating the Mesh with CUBIT}
-
-For the installation and handling of the CUBIT meshing tool suite,
-please refer to the CUBIT user manual and documentation. In order
-to give you a basic understanding of how to use CUBIT for our purposes,
-examples are provided in the SPECFEM3D Cartesian package in the subdirectory
-\texttt{EXAMPLES/}:
-\begin{description}
-\item [{\texttt{homogeneous\_halfspace}}] Creates a single block model
-and assigns elastic material parameters.
-\item [{\texttt{layered\_halfspace}}] Combines two different, elastic material
-volumes and creates a refinement layer between the two. This example
-can be compared for validation against the solutions provided in subdirectory
-~\\
- \texttt{VALIDATION\_3D\_SEM\_SIMPLER\_LAYER\_SOURCE\_DEPTH/}.
-\item [{\texttt{waterlayered\_halfspace}}] Combines an acoustic and elastic
-material volume as in a schematic marine survey example.
-\item [{\texttt{tomographic\_model}}] Creates a single block model whose
-material properties will have to be read in from a tomographic model
-file during the databases creation by \texttt{xgenerate\_databases}.
-\end{description}
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=0.45\textwidth]{figures/example-homogeneous}
-\includegraphics[width=0.45\textwidth]{figures/example-2layers} \\
- \includegraphics[width=0.45\textwidth]{figures/example-water} \includegraphics[width=0.45\textwidth]{figures/example-tomo}
-\par\end{centering}
-
-\caption{Screenshots of the CUBIT examples provided in subdirectory \texttt{EXAMPLES/}:
-homogeneous halfspace (top-left), layered halfspace (top-right), water
-layered halfspace (bottom-left) and tomographic model (bottom-right).}
-
-\label{fig:examples.cubit}
-\end{figure}
-
-In each example subdirectory you will find a \texttt{README} file,
-which explains in a step-by-step tutorial the workflow for the example.
-Please feel free to contribute your own example to this package by
-contacting the CIG Computational Seismology Mailing List \urlwithparentheses{cig-seismo@geodynamics.org}.
-
-In some cases, to re-create the meshes for the examples given, just type "claro ./create\_mesh.py" or similar
-from the command line ("claro" is the command to run CUBIT from the command line).
-
-IMPORTANT note: in order to correctly set up GEOCUBIT and run the examples, please read the file called "EXAMPLES/README";
-in particular, please make sure you correctly set up the Python paths as indicated in that file.
-
-\subsection{Exporting the Mesh with \texttt{run\_cubit2specfem3d.py} }
-
-\label{subsec:Exporting-the-Mesh}
-
-Once the geometric model volumes in CUBIT are meshed, you prepare
-the model for exportation with the definition of material blocks and
-boundary surfaces. Thus, prior to exporting the mesh, you need to
-define blocks specifying the materials and absorbing boundaries in
-CUBIT. This process could be done automatically using the script \texttt{run\_cubit2specfem3d.py}
-if the mesh meets some conditions or manually, following the block
-convention:
-\begin{description}
-\item [{material\_name}] Each material should have a specific block defined
-by a unique name. The name convention of the material is to start
-with either 'elastic' or 'acoustic'. It must be then followed by a
-unique identifier, e.g. 'elastic 1', 'elastic 2', etc. The additional
-attributes to the block define the material description.
-
-For an elastic material:
-\begin{description}
-\item [{material\_id}] An integer value which is unique for this material.
-\item [{Vp}] P-wave speed of the material (given in m/s).
-\item [{Vs}] S-wave speed of the material (given in m/s).
-\item [{rho}] density of the material (given in kg/m$^{3}$).
-\item [{Q}] quality factor to use in case of a simulation with attenuation
-turned on. It should be between 1 and 9000. In case no attenuation
-information is available, it can be set to zero. Please note that
-your Vp- and Vs-speeds are given for a reference frequency. To change
-this reference frequency, you change the value of \texttt{ATTENUATION\_f0\_REFERENCE}
-in the main constants file \texttt{constants.h} found in subdirectory
-\texttt{src/shared/}.
-\item [{anisotropic\_flag}] Flag describing the anisotropic model to use
-in case an anisotropic simulation should be conducted. See the file
-\texttt{model\_aniso.f90} in subdirectory \texttt{src/generate\_databases/}
-for an implementation of the anisotropic models. In case no anisotropy
-is available, it can be set to zero.
-\end{description}
-
-Note that this material block has to be defined using all the volumes
-which belong to this elastic material. For volumes belonging to another,
-different material, you will need to define a new material block.
-
-For an acoustic material:
-\begin{description}
-\item [{material\_id}] An integer value which is unique for this material.
-\item [{Vp}] P-wave speed of the material (given in m/s).
-\item [{0}] S-wave speed of the material is ignored.
-\item [{rho}] density of the material (given in kg/m$^{3}$).
-\end{description}
-\item [{face\_topo}] Block definition for the surface which defines the
-free surface (which can have topography). The name of this block must
-be 'face\_topo', the block has to be defined using all the surfaces
-which constitute the complete free surface of the model.
-\item [{face\_abs\_xmin}] Block definition for the faces on the absorbing
-boundaries, one block for each surface with x=Xmin.
-\item [{face\_abs\_xmax}] Block definition for the faces on the absorbing
-boundaries, one block for each surface with x=Xmax.
-\item [{face\_abs\_ymin}] Block definition for the faces on the absorbing
-boundaries, one block for each surface with y=Ymin.
-\item [{face\_abs\_ymax}] Block definition for the faces on the absorbing
-boundaries, one block for each surface with y=Ymax.
-\item [{face\_abs\_bottom}] Block definition for the faces on the absorbing
-boundaries, one block for each surface with z=bottom.
-\end{description}
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=0.45\textwidth]{figures/mount-surface} \includegraphics[width=0.45\textwidth]{figures/mount-abs}
-\par\end{centering}
-
-\caption{Example of the block definitions for the free surface 'face\_topo'
-(left) and the absorbing boundaries, defined in a single block 'face\_abs\_xmin'
-(right) in CUBIT.}
-
-\label{fig:mount.abs}
-\end{figure}
-
-Optionally, instead of specifying for each surface at the boundaries
-a single block like mentioned above, you can also specify a single
-block for all boundary surfaces and name it as one of the absorbing
-blocks above, e.g. 'face\_abs\_xmin'.
-
-After the block definitions are done, you export the mesh using the
-script \texttt{cubit2specfem3d.py} provided in each of the example
-directories (linked to the common script \texttt{CUBIT\_GEOCUBIT/cubit2specfem3d.py}).
-If the export was successful, you should find the following
-files in a subdirectory \texttt{MESH/}:
-\begin{description}
-\item [{absorbing\_cpml\_file}] \textbf{(only needed in case of C-PML absorbing
-conditions)} Contains on the first line the total number of C-PML
-spectral elements in the mesh, and then on the following line the
-list of all these C-PML elements with two numbers per line: first
-the spectral element number, and then a C-PML flag indicating to which
-C-PML layer(s) that element belongs, according to the following convention:
-
-\begin{itemize}
-\item Flag = 1 : element belongs to a X CPML layer only (either in Xmin
-or in Xmax),
-\item Flag = 2 : element belongs to a Y CPML layer only (either in Ymin
-or in Ymax),
-\item Flag = 3 : element belongs to a Z CPML layer only (either in Zmin
-or in Zmax),
-\item Flag = 4 : element belongs to a X CPML layer and also to a Y CPML
-layer,
-\item Flag = 5 : element belongs to a X CPML layer and also to a Z CPML
-layer,
-\item Flag = 6 : element belongs to a Y CPML layer and also to a Z CPML
-layer,
-\item Flag = 7 : element belongs to a X, to a Y and to a Z CPML layer, i.e.,
-it belongs to a CPML corner.
-\end{itemize}
-
-Note that it does not matter whether an element belongs to a Xmin
-or to a Xmax CPML, the flag is the same in both cases; the same is
-true for Ymin and Ymax, and also Zmin and Zmax.
-
-When you have an existing CUBIT (or similar) mesh stored in SPECFEM3D
-format, i.e., if you have existing \textquotedbl{}nodes\_coords\_file\textquotedbl{}
-and \textquotedbl{}mesh\_file\textquotedbl{} files but do not know
-how to assign CPML flags to them, we have created a small serial Fortran
-program that will do that automatically for you, i.e., which will
-create the \textquotedbl{}absorbing\_cpml\_file\textquotedbl{} for
-you. That program is utils/CPML/convert\_external\_layers\_of\_a\_given\_mesh\_to\_CPML\_layers.f90,
-and a small Makefile is provided in that directory (utils/CPML).
-
-IMPORTANT: it is your responsibility to make sure that in the input
-CUBIT (or similar) mesh that this code will read in SPECFEM3D format
-from files \textquotedbl{}nodes\_coords\_file\textquotedbl{} and \textquotedbl{}mesh\_file\textquotedbl{}
-you have created layers of elements that constitute a layer of constant
-thickness aligned with the coordinate grid axes (X, Y and/or Z), so
-that this code can assign CPML flags to them. This code does NOT check
-that (because it cannot, in any easy way). The mesh inside these CPML
-layers does not need to be structured nor regular, any non-structured
-mesh is fine as long as it has flat PML inner and outer faces, parallel
-to the axes, and thus of a constant thickness. The thickness can be
-different for the X, Y and Z sides. But for X it must not vary, for
-Y it must not vary, and for Z it must not vary. If you do not know
-the exact thickness, you can use a slightly LARGER value in this code
-(say 2\% to 5\% more) and this code will fix that and will adjust
-it; never use a SMALLER value otherwise this code will miss some CPML
-elements.
-
-Note: in a future release we will remove the constraint of having
-CPML layers aligned with the coordinate axes; we will allow for meshes
-that are titled by any constant angle in the horizontal plane. However
-this is not implemented yet.
-
-Note: in the case of fluid-solid layers in contact, the fluid-solid
-interface currently needs to be flat and horizontal inside the CPML
-layer (i.e., bathymetry should become flat and horizontal when entering
-the CPML); this (small) constraint will probably remain in the code
-for a while because it makes fluid-solid matching inside the CPML
-much easier.
-
-\item [{materials\_file}] Contains the material associations for each element.
-The format is:
-
-\begin{lyxcode}
-element\_ID~material\_ID
-\end{lyxcode}
-
-where \texttt{element\_ID} is the element identifier and \texttt{material\_ID}
-is a unique identifier, positive (for materials taken from this list
-of materials, i.e. for which each spectral element has constant material
-properties taken from this list) or negative (for tomographic models,
-i.e. for spectral element whose real velocities and density will be
-assigned later by calling an external function to define model variations,
-for instance in the case of tomographic models; in such a case, material
-properties can vary inside each spectral element, i.e. be different
-at each of its Gauss-Lobatto-Legendre grid points).
-
-\item [{nummaterial\_velocity\_file}] Defines the material properties.
-
-\begin{itemize}
-\item For classical materials (i.e., spectral elements for which the velocity
-and density model will not be assigned by calling an external function
-to define for instance a tomographic model), the format is:
-
-\begin{lyxcode}
-domain\_ID~material\_ID~rho~vp~vs~Qkappa~Qmu~anisotropy\_flag
-\end{lyxcode}
-
-where \texttt{\textbf{domain\_ID}}\textbf{ is 1 for acoustic and 2
-for elastic or viscoelastic materials,} \texttt{material\_ID} a unique
-identifier, \texttt{rho} the density in $kg\, m^{-3}$, \texttt{vp}
-the P-wave speed in $m\, s^{-1}$, \texttt{vs} the S-wave speed in
-$m\, s^{-1}$, \texttt{Q} the quality factor and \texttt{anisotropy\_flag}
-an identifier for anisotropic models. Note that the \texttt{Qkappa}
-value is ignored by the code unless \texttt{FULL\_ATTENUATION\_SOLID}
-is set. Note also that both \texttt{Qkappa} and \texttt{Qmu} are ignored
-by the code unless \texttt{ATTENUATION} is set. If you want a model
-with no \texttt{Qmu} attenuation, both set \texttt{ATTENUATION} to
-\texttt{.false.} in the \texttt{Par\_file} and set \texttt{Qmu} to
-9999 here. If you want a model with no \texttt{Qkappa} attenuation,
-both set \texttt{FULL\_ATTENUATION\_SOLID} to \texttt{.false.} in
-the \texttt{Par\_file} and set \texttt{Qkappa} to 9999 here. %%magnoni
-
-\item For tomographic velocity models, please read Chapter \ref{cha:-Changing-the}
-and Section \ref{sec:Using-tomographic} `Using external tomographic
-Earth models' for further details.
-\end{itemize}
-\item [{nodes\_coords\_file}] Contains the point locations in Cartesian
-coordinates of the mesh element corners.
-\item [{mesh\_file}] Contains the mesh element connectivity. The hexahedral
-elements can have 8 or 27 nodes.\\
- See picture doc/mesh\_numbering\_convention/numbering\_convention\_27\_nodes.jpg
-to see\\
- in which (standard) order the points must be cited. In the case of
-8 nodes, just include the first 8 points.
-\item [{free\_or\_absorbing\_surface\_file\_zmax}] Contains the free surface
-connectivity or \\
- the surface connectivity of the absorbing boundary surface at the
-top (Zmax), \\
- depending on whether the top surface is defined as free or absorbing
-(\texttt{STACEY\_INSTEAD\_OF\_FREE\_SURFACE} in \texttt{DATA/Par\_file}). \\
- You should put both the surface of acoustic regions and of elastic
-regions in that file; that is, list all the element faces that constitute
-the surface of the model in that file.
-\item [{absorbing\_surface\_file\_xmax}] Contains the surface connectivity
-of the absorbing boundary surface at Xmax \\
-(also needed in the case of C-PML absorbing conditions, in order for
-the code to be able to impose Dirichlet conditions on their outer
-edge).
-\item [{absorbing\_surface\_file\_xmin}] Contains the surface connectivity
-of the absorbing boundary surface at Xmin \\
-(also needed in the case of C-PML absorbing conditions, in order for
-the code to be able to impose Dirichlet conditions on their outer
-edge).
-\item [{absorbing\_surface\_file\_ymax}] Contains the surface connectivity
-of the absorbing boundary surface at Ymax \\
-(also needed in the case of C-PML absorbing conditions, in order for
-the code to be able to impose Dirichlet conditions on their outer
-edge).
-\item [{absorbing\_surface\_file\_ymin}] Contains the surface connectivity
-of the absorbing boundary surface at Ymin \\
-(also needed in the case of C-PML absorbing conditions, in order for
-the code to be able to impose Dirichlet conditions on their outer
-edge).
-\item [{absorbing\_surface\_file\_bottom}] Contains the surface connectivity
-of the absorbing boundary surface at the bottom (Zmin) \\
-(also needed in the case of C-PML absorbing conditions, in order for
-the code to be able to impose Dirichlet conditions on their outer
-edge).
-\end{description}
-These mesh files are needed as input files for the partitioner \texttt{xdecompose\_mesh}
-to load-balance the mesh. Please see the next section for further
-details.
-
-In directory "CUBIT\_GEOCUBIT" we provide a script that can help doing the above tasks of exporting a CUBIT mesh
-to SPECFEM3D format automatically for you: "run\_cubit2specfem3d.py".
-Just edit them to indicate the path to your local installation of CUBIT and also the name of the *.cub existing CUBIT mesh file
-that you want to export to SPECFEM3D format. These scripts will do the conversion for you automatically except assigning material properties
-to the different mesh layers. To do so, you will then need to edit the file
-called "nummaterial\_velocity\_file" that will have just been created and change it from the prototype created:
-
-    0 1 vol1 --> syntax: \#material\_domain\_id \#material\_id \#rho \#vp \#vs \#Q\_kappa \#Q\_mu \#anisotropy
-    0 2 vol2 --> syntax: \#material\_domain\_id \#material\_id \#rho \#vp \#vs \#Q\_kappa \#Q\_mu \#anisotropy
-
-(where "vol1" and "vol2" here represent the volume labels that you have set while creating the mesh in CUBIT)
-to for instance
-
-    2 1 1500 2300 1800 9999.0 9999.0 0
-    2 2 1600 2500 20000 9999.0 9999.0 0
-
-\subsubsection*{Checking the mesh quality}
-
-The quality of the mesh may be inspected more precisely based upon
-the serial code in the file \texttt{check\_mesh\_quality\_}~\\
- \texttt{CUBIT\_Abaqus.f90} located in the directory \texttt{src/check\_mesh\_quality\_CUBIT\_Abaqus/}.
-Running this code is optional because no information needed by the
-solver is generated.
-
-Prior to running and compiling this code, you have to export your
-mesh in CUBIT to an ABAQUS (.inp) format. For example, export mesh
-block IDs belonging to volumes in order to check the quality of the
-hexahedral elements. You also have to determine a number of parameters
-of your mesh, such as the number of nodes and number of elements and
-modify the header of the \texttt{check\_mesh\_quality\_CUBIT\_Abaqus.f90}
-source file in directory \texttt{src/check\_mesh\_quality\_CUBIT\_Abaqus/}.
-
-\noindent
-Then, in the main directory, type
-\begin{lyxcode}
-{\small make~xcheck\_mesh\_quality\_CUBIT\_Abaqus}{\small \par}
-\end{lyxcode}
-and use
-\begin{lyxcode}
-{\small ./bin/xcheck\_mesh\_quality\_CUBIT\_Abaqus}{\small \par}
-\end{lyxcode}
-to generate an %not supported: AVS output file (\texttt{\small AVS\_meshquality.inp} in AVS UCD format) or
-OpenDX output file (\texttt{\small DX\_mesh\_quality.dx}{\small )
-that can be used to investigate mesh quality, e.g. skewness of elements,
-and a Gnuplot histogram (}\texttt{\small mesh\_quality\_histogram.txt}{\small )
-that can be plotted with gnuplot (type `}\texttt{\small gnuplot plot\_mesh\_quality\_histogram.gnu}{\small ').
-The histogram is also printed to the screen. Analyze that skewness
-histogram of mesh elements to make sure no element has a skewness
-above approximately 0.75, otherwise the mesh is of poor quality (and
-if even a single element has a skewness value above 0.80, then you
-must definitely improve the mesh). If you want to start designing
-your own meshes, this tool is useful for viewing your creations. You
-are striving for meshes with elements with `cube-like' dimensions,
-e.g., the mesh should contain no very elongated or skewed elements.}{\small \par}
-
-\subsection{Partitioning the Mesh with \texttt{xdecompose\_mesh}}
-
-The SPECFEM3D Cartesian software package performs large scale simulations
-in a parallel 'Single Process Multiple Data' way. The spectral-element
-mesh created with CUBIT needs to be distributed on the processors.
-This partitioning is executed once and for all prior to the execution
-of the solver so it is referred to as a static mapping.
-
-An efficient partitioning is important because it leverages the overall
-running time of the application. It amounts to balance the number
-of elements in each slice while minimizing the communication costs
-resulting from the placement of adjacent elements on different processors.
-\texttt{decompose\_mesh} depends on the SCOTCH library \citep{PeRo96},
-which provides efficent static mapping, graph and mesh partitioning
-routines. SCOTCH is a free software package developed by Fran\c{c}ois
-Pellegrini et al. from LaBRI and INRIA in Bordeaux, France, downloadable
-from the web page \url{https://gforge.inria.fr/projects/scotch/}.
-\\
-
-In most cases, the configuration with \texttt{./configure~FC=ifort}
-should be sufficient. During the configuration process, the script
-tries to find existing SCOTCH installations. In case your system has
-no pre-existing SCOTCH installation, we provide the source code of
-SCOTCH, which is released open source under the French CeCILL-C version
-1 license, in directory \texttt{src/decompose\_mesh/scotch\_5.1.12b}.
-This version gets bundled with the compilation of the SPECFEM3D Cartesian
-package if no libraries could have been found. If this automatic compilation
-of the SCOTCH libraries fails, please refer to file INSTALL.txt in
-that directory to see further details how to compile it on your system.
-In case you want to use a pre-existing installation, make sure you
-have correctly specified the path of the SCOTCH library when using
-the option \texttt{-{}-with-scotch-dir} with the \texttt{./configure}
-script. In the future you should be able to find more recent versions
-at \url{http://www.labri.fr/perso/pelegrin/scotch/scotch_en.html}
-. \\
-
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=0.45\textwidth]{figures/mount-partitions}
-\includegraphics[width=0.45\textwidth]{figures/mount-partitions2}
-\par\end{centering}
-
-\caption{Example of a mesh partitioning onto four cores. Each single core partition
-is colored differently. The executable \texttt{xdecompose\_mesh} can
-equally distribute the mesh on any arbitrary number of cores. Domain
-decomposition is explained in detail in \citet{MaKoBlLe08}, and excellent
-scaling up to 150,000 processor cores in shown for instance in \citet{CaKoLaTiMiLeSnTr08,KoLaMi08a,MaKoBlLe08,KoErGoMi10,Kom11}.}
-
-\label{fig:mount.partitions}
-\end{figure}
-
-When you are ready to compile, in the main directory type
-\begin{lyxcode}
-{\small make~xdecompose\_mesh}{\small \par}
-\end{lyxcode}
-If all paths and flags have been set correctly, the executable \texttt{bin/xdecompose\_mesh}
-should be produced.
-
-The partitioning is done in serial for now (in the next release we
-will provide a parallel version of that code). It needs to be run in the main directory because it expects the \texttt{./DATA/Par\_file}. The synopsis is:
-\begin{lyxcode}
-./bin/xdecompose\_mesh~nparts~input\_directory~output\_directory
-\end{lyxcode}
-where
-\begin{itemize}
-\item \texttt{nparts} is the number of partitions, i.e., the number of cores
-for the parallel simulations,
-\item \texttt{input\_directory} is the directory which holds all the files
-generated by the Python script \texttt{cubit2specfem3d.py} explained
-in the previous Section~\ref{subsec:Exporting-the-Mesh}, e.g. \texttt{./MESH/},
-and
-\item \texttt{output\_directory} is the directory for the output of this
-partitioner which stores ACII-format files named like \texttt{proc{*}{*}{*}{*}{*}{*}\_Database}
-for each partition. These files will be needed for creating the distributed
-databases, and have to reside in the directory \texttt{LOCAL\_PATH}
-specified in the main \texttt{Par\_file}, e.g. in directory \texttt{./OUTPUT\_FILES/DATABASES\_MPI}.
-Please see Chapter~\ref{cha:Creating-Distributed-Databases} for
-further details.
-\end{itemize}
-Note that all the files generated by the Python script \texttt{cubit2specfem3d.py}
-must be placed in the \texttt{input\_directory} folder before running
-the program.
-
-\section{\label{cha:Running-the-Mesher-Meshfem3D}Meshing with \texttt{xmeshfem3D}}
-
-In case you successfully ran the configuration script, you are also
-ready to compile the internal mesher. This is an alternative to CUBIT
-for the mesh generation of relatively simple models.
-
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[scale=0.5]{figures/socal_map_mpi}
-\par\end{centering}
-
-\caption{\label{fig:For-parallel-computing}For parallel computing purposes,
-the model block is subdivided in $\nprocxi\times\nproceta$ slices
-of elements. In this example we use $5^{2}=25$ processors. }
-\end{figure}
-
-\noindent
-In the main directory, type
-\begin{lyxcode}
-{\small make~xmeshfem3D}{\small \par}
-\end{lyxcode}
-If all paths and flags have been set correctly, the mesher should
-now compile and produce the executable \texttt{bin/xmeshfem3D}. Please
-note that \texttt{xmeshfem3D} must be called directly from the main
-directory, as most of the binaries of the package.
-
-Input for the mesh generation program is provided through the parameter
-file \texttt{Mesh\_Par\_file}, which resides in the subdirectory \texttt{DATA/meshfem3D\_files/}.
-(to see how to use it, see the EXAMPLES specific to the internal mesher in directory \texttt{EXAMPLES/meshfem3D\_examples/}.
-Before running the mesher, a number of parameters need to be set in
-the \texttt{Mesh\_Par\_file}. This requires a basic understanding
-of how the SEM is implemented, and we encourage you to read \citet{KoVi98,KoTr99}
-and \citet{KoLiTrSuStSh04}.
-
-The mesher and the solver use UTM coordinates internally, therefore
-you need to define the zone number for the UTM projection (e.g., zone
-11 for Los Angeles). Use decimal values for latitude and longitude
-(no minutes/seconds). These values are approximate; the mesher will
-round them off to define a square mesh in UTM coordinates. When running
-benchmarks on rectangular models, turn the UTM projection off by using
-the flag \texttt{\small SUPPRESS\_UTM\_PROJECTION}{\small , in which
-case all `longitude' parameters simply refer to the $x$~axis, and
-all `latitude' parameters simply refer to the $y$~axis. To run the
-mesher for a global simulation, the following parameters need to be
-set in the }\texttt{\small Mesh\_Par\_file}{\small :}{\small \par}
-\begin{description}
-\item [{\texttt{LATITUDE\_MIN}}] Minimum latitude in the block (negative
-for South).
-\item [{\texttt{LATITUDE\_MAX}}] Maximum latitude in the block.
-\item [{\texttt{LONGITUDE\_MIN}}] Minimum longitude in the block (negative
-for West).
-\item [{\texttt{LONGITUDE\_MAX}}] Maximum longitude in the block.
-\item [{\texttt{DEPTH\_BLOCK\_KM}}] Depth of bottom of mesh in kilometers.
-\item [{\texttt{\noun{UTM\_PROJECTION\_ZONE}}}] UTM projection zone in
-which your model resides, only valid when \texttt{SUPPRESS\_UTM\_}~\\
- \texttt{PROJECTION} is \texttt{.false.}.
-Use a negative zone number for the Southern hemisphere:
-the Northern hemisphere corresponds to zones +1 to +60,
-the Southern hemisphere to zones -1 to -60.
-We use the WGS84 (World Geodetic System 1984) reference ellipsoid for the UTM projection. If you prefer to use the Clarke 1866 ellipsoid,
-edit file \texttt{src/shared/utm\_geo.f90}, uncomment that ellipsoid and recompile the code.
-From http://en.wikipedia.org/wiki/Universal\_Transverse\_Mercator\_coordinate\_system :
-The Universal Transverse Mercator coordinate system was developed by the United States Army Corps of Engineers in the 1940s.
-The system was based on an ellipsoidal model of Earth. For areas within the contiguous United States
-the Clarke Ellipsoid of 1866 was used. For the remaining areas of Earth, including Hawaii, the International Ellipsoid was used.
-The WGS84 ellipsoid is now generally used to model the Earth in the UTM coordinate system,
-which means that current UTM northing at a given point can be 200+ meters different from the old one.
-For different geographic regions, other datum systems (e.g.: ED50, NAD83) can be used.
-\item [{\texttt{SUPPRESS\_UTM\_PROJECTION}}] set to be \texttt{.false.}
-when your model range is specified in geographical coordinates, and
-needs to be \texttt{.true.} when your model is specified in Cartesian
-coordinates. \noun{UTM projection zone in which your simulation region
-resides.}
-\item [{\texttt{INTERFACES\_FILE }}] File which contains the description
-of the topography and of the interfaces between the different layers
-of the model, if any. The number of spectral elements in the vertical
-direction within each layer is also defined in this file.
-\item [{$\nexxi$}] The number of spectral elements along one side of the
-block. This number \textit{must} be 8~$\times$~a multiple of $\nprocxi$
-defined below. Based upon benchmarks against semi-analytical discrete
-wavenumber synthetic seismograms \citep{KoLiTrSuStSh04}, determined
-that a $\nexxi=288$ run is accurate to a shortest period of roughly
-2~s. Therefore, since accuracy is determined by the number of grid
-points per shortest wavelength, for any particular value of $\nexxi$
-the simulation will be accurate to a shortest period determined by
-\begin{equation}
-\mbox{shortest period (s)}=(288/\nexxi)\times2.\label{eq:shortest_period}
-\end{equation}
- The number of grid points in each orthogonal direction of the reference
-element, i.e., the number of Gauss-Lobatto-Legendre points, is determined
-by \texttt{NGLLX} in the \texttt{constants.h} file. We generally use
-$\mbox{\texttt{NGLLX\/}}=5$, for a total of $5^{3}=125$ points per
-elements. We suggest not to change this value.
-\item [{$\nexeta$}] The number of spectral elements along the other side
-of the block. This number \textit{must} be 8~$\times$~a multiple
-of $\nproceta$ defined below.
-\item [{$\nprocxi$}] The number of processors or slices along one side
-of the block (see Figure~\ref{fig:For-parallel-computing}); we must
-have $\nexxi=8\times c\times\nprocxi$, where $c\ge1$ is a positive
-integer.
-\item [{$\nproceta$}] The number of processors or slices along the other
-side of the block; we must have $\nexeta=8\times c\times\nproceta$,
-where $c\ge1$ is a positive integer.
-\item [{\texttt{USE\_REGULAR\_MESH}}] set to be \texttt{.true.} if you
-want a perfectly regular mesh or \texttt{.false.} if you want to add
-doubling horizontal layers to coarsen the mesh. In this case, you
-also need to provide additional information by setting up the next
-three parameters.
-\item [{\texttt{NDOUBLINGS}}] The number of horizontal doubling layers.
-Must be set to \texttt{1} or \texttt{2} if \texttt{USE\_REGULAR\_MESH}
-is set to \texttt{.true.}.
-\item [{\texttt{NZ\_DOUBLING\_1}}] The position of the first doubling layer
-(only interpreted if \texttt{USE\_REGULAR\_MESH} is set to \texttt{.true.}).
-\item [{\texttt{NZ\_DOUBLING\_2}}] The position of the second doubling
-layer (only interpreted if \texttt{USE\_REGULAR\_MESH} is set to \texttt{.true.}
-and if \texttt{NDOUBLINGS} is set to \texttt{2}).
-\item [{\texttt{CREATE\_ABAQUS\_FILES}}] Set this flag to \texttt{.true.}
-to save Abaqus FEA \urlwithparentheses{www.simulia.com} mesh files
-for subsequent viewing. Turning the flag on generates files in the
-\texttt{LOCAL\_PATH} directory. See Section~\ref{sec:Mesh-graphics}
-for a discussion of mesh viewing features.
-\item [{\texttt{CREATE\_DX\_FILES}}] Set this flag to \texttt{.true.} to
-save OpenDX \urlwithparentheses{www.opendx.org} mesh files for subsequent
-viewing.
-\item [{\texttt{LOCAL\_PATH}}] Directory in which the partitions generated
-by the mesher will be written. Generally one uses a directory on the
-local disk of the compute nodes, although on some machines these partitions
-are written on a parallel (global) file system (see also the earlier
-discussion of the \texttt{LOCAL\_PATH\_IS\_ALSO\_GLOBAL} flag in Chapter~\ref{cha:Getting-Started}).
-The mesher generates the necessary partitions in parallel, one set
-for each of the $\nprocxi\times\nproceta$ slices that constitutes
-the mesh (see Figure~\ref{fig:For-parallel-computing}). After the
-mesher finishes, you can log in to one of the compute nodes and view
-the contents of the \texttt{LOCAL\_PATH} directory to see the files
-generated by the mesher. These files will be needed for creating the
-distributed databases, and have to reside in the directory \texttt{LOCAL\_PATH}
-specified in the main \texttt{Par\_file}, e.g. in directory \texttt{OUTPUT\_FILES/DATABASES\_MPI}.
-Please see Chapter~\ref{cha:Creating-Distributed-Databases} for
-further details.
-\item [{\texttt{NMATERIALS}}] The number of different materials in your
-model. In the following lines, each material needs to be defined as
-:
-
-\begin{lyxcode}
-material\_ID~rho~vp~vs~Q~anisotropy\_flag~domain\_ID
-\end{lyxcode}
-
-where
-\begin{itemize}
-\item \texttt{Q} : quality factor (0=no attenuation)
-\item \texttt{anisotropy\_flag} : 0=no anisotropy / 1,2,.. check with implementation
-in \texttt{aniso\_model.f90}
-\item \texttt{domain\_id} : 1=acoustic / 2=elastic
-\end{itemize}
-\item [{\texttt{NREGIONS}}] The number of regions in the mesh. In the following
-lines, because the mesh is regular or 'almost regular', each region
-is defined as :
-
-\begin{lyxcode}
-NEX\_XI\_BEGIN~NEX\_XI\_END~NEX\_ETA\_BEGIN~NEX\_ETA\_END~NZ\_BEGIN~NZ\_END~material\_ID
-\end{lyxcode}
-\end{description}
-The \texttt{INTERFACES\_FILE} parameter of \texttt{Mesh\_Par\_File}
-defines the file which contains the settings of the topography grid
-and of the interfaces grids. Topography is defined as a set of elevation
-values on a regular 2D grid. It is also possible to define interfaces
-between the layers of the model in the same way. The file needs to
-define several parameters:
-\begin{itemize}
-\item The number of interfaces, including the topography. This needs to
-be set at the first line. Then, from the bottom to the top of the
-model, you need to define the grids with:
-\item \texttt{SUPPRESS\_UTM\_PROJECTION} flag as described previously,
-\item number of points along $x$ and $y$ direction (NXI and NETA),
-\item minimal $x$ and $y$ coordinates (LONG\_MIN and LAT\_MIN),
-\item spacing between points along $x$ and $y$ (SPACING\_XI and SPACING\_ETA)
-and
-\item the name of the file which contains the elevation values (in $y$.$x$
-increasing order).
-\end{itemize}
-At the end of this file, you simply need to set the number of spectral
-elements in the vertical direction for each layer. We provide a few
-models in the {\texttt{EXAMPLES/}} directory. \\
-
-Finally, depending on your system, you might need to provide a file
-that tells MPI what compute nodes to use for the simulations. The
-file must have a number of entries (one entry per line) at least equal
-to the number of processors needed for the run. A sample file is provided
-in the file \texttt{mymachines}. This file is not used by the mesher
-or solver, but is required by the \texttt{go\_mesher} and \texttt{go\_solver}
-default job submission scripts. See Chapter~\ref{cha:Scheduler}
-for information about running the code on a system with a scheduler,
-e.g., LSF.
-
-Now that you have set the appropriate parameters in the \texttt{Mesh\_Par\_file}
-and have compiled the mesher, you are ready to launch it! This is
-most easily accomplished based upon the \texttt{go\_mesher} script.
-When you run on a PC cluster, the script assumes that the nodes are
-named n001, n002, etc. If this is not the case, change the \texttt{tr
--d `n'} line in the script. You may also need to edit the last command
-at the end of the script that invokes the \texttt{mpirun} command.
-See Chapter~\ref{cha:Scheduler} for information about running the
-code on a system with a scheduler, e.g., LSF.
-
-Mesher output is provided in the \texttt{OUTPUT\_FILES} directory
-in \texttt{output\_mesher.txt}; this file provides lots of details
-about the mesh that was generated. Please note that the mesher suggests
-a time step \texttt{DT} to run the solver with. The mesher output
-file also contains a table about the quality of the mesh to indicate
-possible problems with the distortions of elements. Alternatively,
-output can be directed to the screen instead by uncommenting a line
-in \texttt{constants.h}:
-\begin{lyxcode}
-!~uncomment~this~to~write~messages~to~the~screen~
-
-!~integer,~parameter~::~IMAIN~=~ISTANDARD\_OUTPUT
-\end{lyxcode}
-To control the quality of the mesh, check the standard output (either
-on the screen or in the \texttt{OUTPUT\_FILES} directory in \texttt{output\_mesher.txt})
-and analyze the skewness histogram of mesh elements to make sure no
-element has a skewness above approximately 0.75, otherwise the mesh
-is of poor quality (and if even a single element has a skewness value
-above 0.80, then you must definitely improve the mesh). To draw the
-skewness histogram on the screen, type \texttt{gnuplot plot\_mesh\_quality\_histogram.gnu}.
-
-\chapter{\label{cha:Creating-Distributed-Databases}Creating the Distributed
-Databases}
-
-After using \texttt{xmeshfem3D} or \texttt{xdecompose\_mesh}, the
-next step in the workflow is to compile \texttt{xgenerate\_}~\\
- \texttt{databases}. This program is going to create all the missing
-information needed by the SEM solver.
-
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=0.6\textwidth]{figures/workflow}
-\par\end{centering}
-
-\caption{Schematic workflow for a SPECFEM3D Cartesian simulation. The executable
-\texttt{xgenerate\_databases} creates the GLL mesh points and assigns
-specific model parameters.}
-
-\label{fig:workflow.databases}
-\end{figure}
-
-\noindent
-In the main directory, type
-\begin{lyxcode}
-{\small make~xgenerate\_databases}{\small \par}
-\end{lyxcode}
-Input for the program is provided through the main parameter file
-\texttt{Par\_file}, which resides in the subdirectory \texttt{DATA}.
-Please note that \texttt{xgenerate\_databases} must be called directly
-from the main directory, as most of the binaries of the package.
-
-\section{\label{cha:Main-Parameter}Main parameter file \texttt{Par\_file}}
-
-Before running \texttt{xgenerate\_databases}, a number of parameters
-need to be set in the main parameter \texttt{Par\_file} located in
-the subdirectory \texttt{DATA}:
-\begin{description}
-\item [{\texttt{SIMULATION\_TYPE}}] is set to 1 for forward simulations,
-2 for adjoint simulations (see Section \ref{sec:Adjoint-simulation-finite})
-and 3 for kernel simulations (see Section \ref{sec:Finite-Frequency-Kernels}).
-\item [{\texttt{SAVE\_FORWARD}}] is only set to \texttt{.true.} for a forward
-simulation with the last frame of the simulation saved, as part of
-the finite-frequency kernel calculations (see Section \ref{sec:Finite-Frequency-Kernels}).
-For a regular forward simulation, leave \texttt{SIMULATION\_TYPE}
-and \texttt{SAVE\_FORWARD} at their default values.
-\item [{\texttt{\noun{UTM\_PROJECTION\_ZONE}}}] UTM projection zone in
-which your model resides, only valid when \texttt{SUPPRESS\_UTM\_}~\\
- \texttt{PROJECTION} is \texttt{.false.}.
-\item [{\texttt{SUPPRESS\_UTM\_PROJECTION}}] set to be \texttt{.false.}
-when your model range is specified in the geographical coordinates,
-and needs to be \texttt{.true.} when your model is specified in a
-cartesian coordinates. \noun{UTM projection zone in which your simulation
-region resides.}
-\item [{\texttt{NPROC}}] The number of MPI processors, each one is assigned
-one slice of the whole mesh.
-\item [{\texttt{NSTEP}}] The number of time steps of the simulation. This
-controls the length of the numerical simulation, i.e., twice the number
-of time steps requires twice as much CPU time. This feature is not
-used at the time of generating the distributed databases but is required
-for the solver, i.e., you may change this parameter after running
-\texttt{xgenerate\_databases}.
-\item [{\texttt{DT}}] The length of each time step in seconds. This feature
-is not used at the time of generating the distributed databases but
-is required for the solver. Please see also Section~\ref{sec:Choosing-the-Time-Step}
-for further details.
-\item [{\texttt{NGNOD}}] The number of nodes for 2D and 3D shape functions
-for hexahedra. We use either 8-node mesh elements (bricks) or 27-node
-elements. If you use the internal mesher, the only option is 8-node
-bricks (27-node elements are not supported). \texttt{CUBIT} does not
-support HEX27 elements either (it can generate them, but they are
-flat, i.e. identical to HEX8). To generate HEX27 elements with curvature
-properly taken into account, you can use \texttt{Gmsh} \url{http://geuz.org/gmsh/}
-\item [{\texttt{MODEL}}] Must be set to one of the following:
-
-\begin{description}
-\item [{\textmd{Models defined by mesh parameters:}}] ~
-
-\begin{description}
-\item [{\texttt{default}}] Uses model parameters as defined by meshing
-procedures described in the previous Chapter~\ref{cha:Mesh-Generation}.
-\end{description}
-
-\item [{\textmd{1D~models~with~real~structure:}}] ~
-
-\begin{description}
-\item [{\texttt{1D\_prem}}] Isotropic version of the spherically symmetric
-Preliminary Reference Earth Model (PREM) \citep{DzAn81}.
-\item [{\texttt{1D\_socal}}] A standard isotropic 1D model for Southern
-California.
-\item [{\texttt{1D\_cascadia}}] Isotropic 1D profile for the Cascadia region.
-\end{description}
-
-\item [{\textmd{Fully~3D~models:}}] ~
-
-\begin{description}
-\item [{\texttt{aniso}}] For a user-specified fully anisotropic model.
-Parameters are set up in routines located in file \texttt{model\_aniso.f90}
-in directory \texttt{src/generate\_databases/}. See Chapter~\ref{cha:-Changing-the}
-for a discussion on how to specify your own 3D model.
-
-\item [{\texttt{external}}] For a user-specified isotropic model which
-uses externally defined model parameters. Uses external model definitions
-set up in routines located in file \texttt{model\_external\_values.f90}
-in directory \texttt{src/generate\_databases/}. Please modify these
-generic template routines to use your own model definitions.
-\item [{\texttt{gll}}] For a user-specified isotropic model which uses
-external binary files for $v_{p}$, $v_{s}$ and $\rho$. Binary files
-are given in the same format as when outputted by the \texttt{xgenerate\_databases}
-executable when using option \texttt{SAVE\_MESH\_FILES}. These binary
-files define the model parameters on all GLL points which can be used
-for iterative inversion procedures.
-\item [{\texttt{salton\_trough}}] A 3D $V_{p}$ model for Southern California.
-Users must provide the corresponding data file \texttt{regrid3\_vel\_p.bin}
-in directory \texttt{DATA/st\_3D\_block\_harvard/}.
-\item [{\texttt{tomo}}] For a user-specified 3D isotropic model which uses
-a tomographic model file \texttt{tomographic\_model.xyz} in directory
-\texttt{DATA}. See Section \ref{sec:Using-tomographic}, for a discussion
-on how to specify your own 3D tomographic model.
-\end{description}
-\end{description}
-\item [{\texttt{APPROXIMATE\_OCEAN\_LOAD}}] Set to \texttt{.true.} if the
-effect of the oceans on seismic wave propagation should be incorporated
-based upon the (rough) approximate treatment discussed in \citet{KoTr02b}.
-This feature is inexpensive from a numerical perspective, both in
-terms of memory requirements and CPU time. This approximation is accurate
-at periods of roughly 20~s and longer. At shorter periods the effect
-of water phases/reverberations is not taken into account, even when
-the flag is on. If you want to model the effect of a fluid-solid model
-at short periods, then set this flag to \texttt{.false.} and mesh
-the fluid layer explicitly in your mesher, so that it is computed
-accurately and without this approximation.
-\item [{\texttt{TOPOGRAPHY}}] This feature is only effective if \texttt{APPROXIMATE\_OCEAN\_LOAD}
-is set to \texttt{.true.}. Set to \texttt{.true.} if topography and
-bathymetry should be read in based upon the topography file specified
-in the main constants file \texttt{constants.h} found in subdirectory
-\texttt{src/shared/} to evaluate elevations. If not set, elevations
-will be read from the numerical mesh.
-\item [{\texttt{ATTENUATION}}] Set to \texttt{.true.} if attenuation should
-be incorporated. Turning this feature on increases the memory requirements
-significantly (roughly by a factor of~1.5), and is numerically fairly
-expensive. See \citet{KoTr99,KoTr02a} for a discussion on the implementation
-of attenuation based upon standard linear solids. Please note that
-the Vp- and Vs-velocities of your model are given for a reference
-frequency. To change this reference frequency, you change the value
-of \texttt{ATTENUATION\_f0\_REFERENCE} in the main constants file
-\texttt{constants.h} found in subdirectory \texttt{src/shared/}.
-\item [{\texttt{FULL\_ATTENUATION\_SOLID}}] Set to \texttt{.true.} to add
-$Q_{\kappa}$ attenuation to the simulations in addition to $Q_{\mu}$
-attenuation when the \texttt{ATTENUATION} flag above is on. Should
-be off in almost all cases for seismic wave propagation, since $Q_{\kappa}$
-is almost always negligible in Earth models. However, this parameter
-can be useful for instance for ocean acoustics simulations, when $Q_{\kappa}$
-is not negligible in the solid layer at the sea bottom.
-\item [{\texttt{ANISOTROPY}}] Set to \texttt{.true.} if you want to use
-an anisotropy model. Please see the file \texttt{model\_aniso.f90}
-in subdirectory \texttt{src/generate\_databases/} for the current
-implementation of anisotropic models.
-\item [{\texttt{TOMOGRAPHY\_PATH}}] Directory in which the tomography files
-are stored for using external tomographic Earth models (please read
-Chapter \ref{cha:-Changing-the} and Section \ref{sec:Using-tomographic}
-`Using external tomographic Earth models' for further details.).
-\item [{\texttt{USE\_OLSEN\_ATTENUATION}}] Set to \texttt{.true.} if you
-want to use the attenuation model that scaled from the S-wave speed
-model using Olsen's empirical relation (see \citet{OlDaBr03}).
-\item [{\texttt{OLSEN\_ATTENUATION\_RATIO}}] Determines the Olsen's constant
-in Olsen's empirical relation (see \citet{OlDaBr03}).
-\item [{\texttt{PML\_CONDITIONS}}] Set to \texttt{.true.} to turn on C-PML
-boundary conditions for a regional simulation. Both fluids and elastic
-solids are supported. Note that xmeshfem3d generated meshes do not support C-PML yet.
-\item [{\texttt{PML\_INSTEAD\_OF\_FREE\_SURFACE}}] Set to \texttt{.true.}
-to turn on C-PML boundary conditions on the top surface instead of
-the usual free surface.
-\item [{\texttt{f0\_FOR\_PML}}] Determines the dominant frequency that
-will be used in the calculation of PML damping profiles; this should
-be set to the same (or similar) dominant frequency as that of the
-source that you will use in your simulation. If you plan to use a
-Dirac source, then use the dominant frequency of the source wavelet
-with which you plan to convolve your seismograms later on in post-processing.
-\item [{\texttt{STACEY\_ABSORBING\_CONDITIONS}}] Set to \texttt{.true.}
-to turn on Clayton-Enquist absorbing boundary conditions (see \citet{KoTr99}).
-In almost all cases it is much better to use CPML absorbing layers
-(see the options above) and leave this flag to \texttt{.false.}.
-\item [{\texttt{STACEY\_INSTEAD\_OF\_FREE\_SURFACE}}] Set to \texttt{.true.}
-to turn on absorbing boundary conditions on the top surface which
-by default constitutes a free surface of the model.
-\item [{\texttt{CREATE\_SHAKEMAP}}] Set this flag to \texttt{.true.} to
-create a ShakeMap\textregistered{}, i.e., a peak ground velocity map
-of the maximum absolute value of the two horizontal components of
-the velocity vector.
-\item [{\texttt{MOVIE\_SURFACE}}] Set to \texttt{.false.}, unless you want
-to create a movie of seismic wave propagation on the Earth's surface.
-Turning this option on generates large output files. See Section~\ref{sec:Movies}
-for a discussion on the generation of movies. This feature is only
-relevant for the solver.
-\item [{\texttt{MOVIE\_TYPE}}] Set this flag to 1 to show the top surface
-(tomography + oceans) only, to 2 to show all external faces of the
-mesh (i.e. topography + vertical edges + bottom) in shakemaps and
-surface movies.
-\item [{\texttt{MOVIE\_VOLUME}}] Set to \texttt{.false.}, unless you want
-to create a movie of seismic wave propagation in the Earth's interior.
-Turning this option on generates huge output files. See Section~\ref{sec:Movies}
-for a discussion on the generation of movies. This feature is only
-relevant for the solver.
-\item [{\texttt{SAVE\_DISPLACEMENT}}] Set this flag to \texttt{.true.}
-if you want to save the displacement instead of velocity for the movie
-frames.
-\item [{\texttt{USE\_HIGHRES\_FOR\_MOVIES}}] Set this flag to \texttt{.true.}
-if you want to save the values at all the NGLL grid points for the
-movie frames.
-\item [{\texttt{NTSTEP\_BETWEEN\_FRAMES}}] Determines the number of timesteps
-between movie frames. Typically you want to save a snapshot every
-100 timesteps. The smaller you make this number the more output will
-be generated! See Section~\ref{sec:Movies} for a discussion on the
-generation of movies. This feature is only relevant for the solver.
-\item [{\texttt{HDUR\_MOVIE}}] Determines the half duration of the source
-time function for the movie simulations. When this parameter is set
-to be 0, a default half duration that corresponds to the accuracy
-of the simulation is provided. Otherwise, it adds this half duration
-to the half duration specified in the source file \texttt{CMTSOLUTION},
-thus simulates longer periods to make the movie images look smoother.
-\item [{\texttt{SAVE\_MESH\_FILES}}] Set this flag to \texttt{.true.} to
-save ParaView \urlwithparentheses{www.paraview.org} mesh files for
-subsequent viewing. Turning the flag on generates large (distributed)
-files in the \texttt{LOCAL\_PATH} directory. See Section~\ref{sec:Mesh-graphics}
-for a discussion of mesh viewing features.
-\item [{\texttt{LOCAL\_PATH}}] Directory in which the distributed databases
-will be written. Generally one uses a directory on the local disk
-of the compute nodes, although on some machines these databases are
-written on a parallel (global) file system (see also the earlier discussion
-of the \texttt{LOCAL\_PATH\_IS\_ALSO\_GLOBAL} flag in Chapter~\ref{cha:Getting-Started}).
-\texttt{xgenerate\_databases} generates the necessary databases in
-parallel, one set for each of the \texttt{NPROC} slices that constitutes
-the mesh (see Figure~\ref{fig:mount.partitions} and Figure~\ref{fig:For-parallel-computing}).
-After the executable finishes, you can log in to one of the compute
-nodes and view the contents of the \texttt{LOCAL\_PATH} directory
-to see the (many) files generated by \texttt{xgenerate\_databases}.
-Please note that the \texttt{LOCAL\_PATH} directory should already
-contain the output files of the partitioner, i.e. from \texttt{xdecompose\_mesh}
-or \texttt{xmeshfem3D}.
-\item [{\texttt{NTSTEP\_BETWEEN\_OUTPUT\_INFO}}] This parameter specifies
-the interval at which basic information about a run is written to
-the file system (\texttt{timestamp{*}} files in the \texttt{OUTPUT\_FILES}
-directory). If you have access to a fast machine, set \texttt{NTSTEP\_BETWEEN\_OUTPUT\_INFO}
-to a relatively high value (e.g., at least 100, or even 1000 or more)
-to avoid writing output text files too often. This feature is not
-used at the time of meshing. One can set this parameter to a larger
-value than the number of time steps to avoid writing output during
-the run.
-\item [{\texttt{NTSTEP\_BETWEEN\_OUTPUT\_SEISMOS}}] This parameter specifies
-the interval at which synthetic seismograms are written in the \texttt{LOCAL\_PATH}
-directory. If a run crashes, you may still find usable (but shorter
-than requested) seismograms in this directory. On a fast machine set
-\texttt{NTSTEP\_BETWEEN\_OUTPUT\_SEISMOS} to a relatively high value
-to avoid writing to the seismograms too often. This feature is only
-relevant for the solver.
-\item [{\texttt{USE\_FORCE\_POINT\_SOURCE}}] Turn this flag on to use a
-(tilted) \texttt{FORCESOLUTION} force point source instead of a \texttt{CMTSOLUTION}
-moment-tensor source. When the force source does not fall exactly
-at a grid point, the solver interpolates the force between grid points
-using Lagrange interpolants. This can be useful e.g. for oil industry
-foothills simulations in which the source is a vertical force, normal
-force, tilted force, or an impact etc. Note that in the \texttt{FORCESOLUTION}
-file, you will need to edit the East, North and vertical components
-of an arbitrary (not necessarily unitary, the code will normalize it automatically) direction vector of the force vector;
-thus refer to Appendix~\ref{cha:Coordinates} for the orientation
-of the reference frame. This vector is made unitary internally in
-the solver and thus only its direction matters here; its norm is ignored
-and the norm of the force used is the factor force source times the
-source time function.
-When using this option, by default the code can locate the force source
-anywhere between mesh points in order to honor its exact location; this
-is more precise than using the closest GLL mesh point, but it is also a bit slower.
-If needed, you can change that default behavior and force the code to use the closest
-GLL mesh point instead by setting flag \texttt{USE\_BEST\_LOCATION} to \texttt{.false.} instead of \texttt{.true.} in file
-\texttt{src/shared/constants.h.in} and running the \texttt{configure} script again and recompiling the code.
-\item [{\texttt{USE\_RICKER\_TIME\_FUNCTION}}] Turn this flag on to use
-a Ricker source time function instead of the source time functions
-set by default to represent a (tilted) \texttt{FORCESOLUTION} force
-point source or a \texttt{CMTSOLUTION} moment-tensor source. Originally,
-if a \texttt{CMTSOLUTION} moment-tensor source is used, a (pseudo)
-Heaviside step function with a very short half duration is defined
-for elastic cases to represent the permanent slip on the fault while
-in the acoustic case a Gaussian source time function with a similarly
-short half duration is defined to physically describe actions within
-the fluid. Otherwise, if a \texttt{FORCESOLUTION} force source is
-used, a (pseudo) Dirac delta source time function is defined by default.
-Any other source-time function may then be obtained by convolution.
-\item [{\texttt{PRINT\_SOURCE\_TIME\_FUNCTION}}] Turn this flag on to print
-information about the source time function in the file \texttt{OUTPUT\_FILES/plot\_source\_time\_function.txt}.
-This feature is only relevant for the solver.
-
-\item [{\texttt{NUMBER\_OF\_SIMULTANEOUS\_RUNS}}] adds the ability to run several calculations (several earthquakes)
-in an embarrassingly-parallel fashion from within the same run;
-this can be useful when using a very large supercomputer to compute
-many earthquakes in a catalog, in which case it can be better from
-a batch job submission point of view to start fewer and much larger jobs,
-each of them computing several earthquakes in parallel.
-To turn that option on, set parameter NUMBER\_OF\_SIMULTANEOUS\_RUNS to a value greater than 1.
-To implement that, we create NUMBER\_OF\_SIMULTANEOUS\_RUNS MPI sub-communicators,
-each of them being labeled "my\_local\_mpi\_comm\_world", and we use them
-in all the routines in "src/shared/parallel.f90", except in MPI\_ABORT() because in that case
-we need to kill the entire run.
-When that option is on, of course the number of processor cores used to start
-the code in the batch system must be a multiple of NUMBER\_OF\_SIMULTANEOUS\_RUNS,
-all the individual runs must use the same number of processor cores,
-which as usual is NPROC in the Par\_file,
-and thus the total number of processor cores to request from the batch system
-should be NUMBER\_OF\_SIMULTANEOUS\_RUNS * NPROC.
-All the runs to perform must be placed in directories called run0001, run0002, run0003 and so on
-(with exactly four digits).
-
-Imagine you have 10 independent calculations to do, each of them on 100 cores; you have three options:
-
-1/ submit 10 jobs to the batch system
-
-2/ submit a single job on 1000 cores to the batch, and in that script create a sub-array of jobs to start 10 jobs,
-each running on 100 cores (see e.g. http://www.schedmd.com/slurmdocs/job\_array.html )
-
-3/ submit a single job on 1000 cores to the batch, start SPECFEM3D on 1000 cores, create 10 sub-communicators,
-cd into one of 10 subdirectories (called e.g. run0001, run0002,... run0010) depending on the sub-communicator
-your MPI rank belongs to, and run normally on 100 cores using that sub-communicator.
-
-The option NUMBER\_OF\_SIMULTANEOUS\_RUNS implements 3/.
-
-\item [{\texttt{BROADCAST\_SAME\_MESH\_AND\_MODEL}}]:
-if we perform simultaneous runs in parallel, if only the source and receivers vary between these runs
-but not the mesh nor the model (velocity and density) then we can also read the mesh and model files
-from a single run in the beginning and broadcast them to all the others; for a large number of simultaneous
-runs for instance when solving inverse problems iteratively this can DRASTICALLY reduce I/Os to disk in the solver
-(by a factor equal to NUMBER\_OF\_SIMULTANEOUS\_RUNS), and reducing I/Os is crucial in the case of huge runs.
-Thus, always set this option to .true. if the mesh and the model are the same for all simultaneous runs.
-In that case there is no need to duplicate the mesh and model file database (the content of the DATABASES\_MPI
-directories) in each of the run0001, run0002,... directories, it is sufficient to have one in run0001
-and the code will broadcast it to the others).
-
-TODO / future work to do: currently the BROADCAST\_SAME\_MESH\_AND\_MODEL option assumes to have the (master) mesh files in run0001/DATABASES\_MPI or run0001/OUTPUT\_FILES/DATABASES\_MPI.
-However, for adjoint runs you still need a DATABASES\_MPI folder in each of the sub-runs directories, e.g.
-run0002/DATABASES\_MPI, etc.
-to store the forward wavefields, kernels etc.
-of each sub-run.
-This would not be needed for forward simulations.
-
-TODO / future work to do: the sensitivity kernel summing and smoothing tools in directory src/tomography are currently not ported to this new option to do many runs simultaneously, only the solver (src/specfem3d) is.
-Thus these tools should work, but in their current version will need to be run for each simulation result independently.
-More precisely, the current kernel summing and smoothing routines work fine, with the exception that you need to move out the mesh files (and also the parameters).
-This works because these routines consider multiple runs by design.
-You simply have to provide them the directories where the kernels are.
-
-\item [{\texttt{GPU\_MODE}}] Turn this flag on to use GPUs.
-
-\item [\texttt{ADIOS\_ENABLED}] Turn this flag on to enable ADIOS. If set to \texttt{.false.}, subsequent ADIOS
-parameters will not be considered.
-\item [\texttt{ADIOS\_FOR\_DATABASES}] Turn this flag on to use ADIOS for xmeshfem3D output and
-xgenerate\_database input.
-\item [\texttt{ADIOS\_FOR\_MESH}]  Turn this flag on to use ADIOS for generated databases.
-\item [\texttt{ADIOS\_FOR\_FORWARD\_ARRAYS}] Turn this flag on to read and write forward arrays using ADIOS.
-\item [\texttt{ADIOS\_FOR\_KERNELS}] Turn this flag on to produce ADIOS kernels that can later be visualized with the ADIOS version of combine\_vol\_data.
-
-\end{description}
-
-The present version of SPECFEM can handle fully saturated porous simulations, Christina Morency implemented Biot equation.
-But the code cannot calculate partially saturated cases in its state.
-Christina Morency is presently working on a dual porosity, dual permeability formulation, type Pride and Berryman, for an other project, but it will not be available for some time.\\
-
-The way we prescribe material property for porous material in SPECFEM3D is as follow:
-We use a file name nummaterial\_poroelastic\_file, which is located in the directory MESH, the format is as follow:\\
- rhos rhof phi c kxx kxy kxz kyy kyz kzz Ks Kf Kfr etaf mufr\\
-where\\
-rho\_s = solid density\\
-rho\_f = fluid density\\
-phi = porosity\\
-tort = tortuosity\\
-kxx = xx component of permeability tensor\\
-kxy = xy,yx components of permeability tensor\\
-kyy = yy component of permeability tensor\\
-kxz = xz,zx components of permeability tensor\\
-kzz = zz component of permeability tensor\\
-kappa\_s = solid bulk modulus\\
-kappa\_f= fluid bulk modulus\\
-kappa\_fr= frame bulk modulus\\
-eta\_f = fluid viscosity\\
-mu\_fr = frame shear modulus.\\
-
-Using an external mesh (for instance coming from CUBIT/TRELIS), poroelastic materials have the ID number 3, while 1 is elastic, 2 is acoustic ?
-see the example in the package: examples/homogeneous\_poroelastic.
-
-If you use PML, the mesh elements that belong to the PML layers can
-be acoustic or elastic, but not viscoelastic nor poroelastic. Then,
-when defining your model, you should define these absorbing elements
-as either acoustic or elastic. In you forget to do that, the code
-will fix the problem by automatically converting the viscoelastic
-or poroelastic PML elements to elastic. This means that strictly speaking
-the PML layer will not be perfectly matched any more, since the physical
-model will change from viscoelastic or poroelastic to elastic at the
-entrance of the PML, but in practice this is sufficient and produces
-only tiny / negligible spurious reflections.
-
-If you use PML and an external mesh (created using an external meshing tool
-such as CUBIT/TRELIS or similar), try to have elements inside the PML as regular as possible,
-i.e. ideally non-deformed cubes obtained by `extrusion' of regular surface mesh elements meshing the
-outer edges of the computational domain without PML; by doing so, the PMLs obtained will be far more stable
-in time (PML being weakly unstable from a mathematical point of view, very deformed mesh elements
-inside the PMLs can trigger instabilities much more quickly).
-
-If you use PML and an external tomographic velocity and density model,
-you should be careful because mathematically a PML cannot handle heterogeneities
-along the normal to the PML edge inside the PML layer. This comes
-from the fact that the damping profile that is defined assumes a constant
-velocity and density model along the normal direction.
-
-Thus, you need to modify your velocity and density model in order
-for it to be 1D inside the PML, as shown in Figure~\ref{fig:modify_external_velocity_model_to_use_PML}.
-
-This applies to the bottom layer as well; there you should make sure
-that your model is 1D and thus constant along the vertical direction.
-
-To summarize, only use a 3D velocity and density model inside the
-physical region, and in all the PML layers extend it by continuity
-from its values along the inner PML edge.
-
-%%
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=3in]{figures/how_to_use_PML_when_a_tomographic_velocity_model_is_used}
-\par\end{centering}
-
-\caption{How to modify your external 3D velocity and density model in order
-to use PML. Such a modification is not needed when using Stacey absorbing
-boundary conditions (but such conditions are significantly less efficient).}
-
-\label{fig:modify_external_velocity_model_to_use_PML}
-\end{figure}
-
-%%
-
-\section{\label{sec:Choosing-the-Time-Step}Choosing the time step \texttt{DT}}
-
-The parameter \texttt{DT} sets the length of each time step in seconds.
-The value of this parameter is crucial for the stability of the spectral-element
-simulation. Your time step \texttt{DT} will depend on the minimum
-ratio between the distance $h$ of neighboring mesh points and the
-wave speeds $v$ defined in your model. The condition for the time
-step $\Delta t$ is:
-\begin{lyxcode}
-$\Delta t<C~\mathrm{min}_{\Omega}(~h/v~)$
-\end{lyxcode}
-where $C$ is the so-called Courant number and $\Omega$ denotes the
-model volume. The distance $h$ depends on the mesh element size and
-the number of GLL points \texttt{NGLL} specified in the main constants
-file \texttt{constants.h} located in the \texttt{src/shared/} subdirectory.
-The wave speed $v$ is determined based on your model's P- (or S-)
-wave speed values.
-
-The database generator \texttt{xgenerate\_databases}, as well as the
-internal mesher \texttt{xmeshfem3D}, are trying to evaluate the value
-of $\Delta t$ for empirically chosen Courant numbers $C\sim0.3$.
-If you used the mesher \texttt{xmeshfem3D} to generate your mesh,
-you should set the value suggested in \texttt{OUTPUT\_FILES/output\_mesher.txt}
-file, which is created after the mesher completed. In case you used
-CUBIT to create the mesh, you might use an arbitrary value when running
-\texttt{xgenerate\_databases} and then use the value suggested in
-the ~\\
- \texttt{OUTPUT\_FILES/output\_mesher.txt} file after the database
-generation completed. Note that the implemented Newmark time scheme
-uses this time step globally, thus your simulations become more expensive
-for very small mesh elements in high wave-speed regions. Please be
-aware of this restriction when constructing your mesh in Chapter~\ref{cha:Mesh-Generation}.
-
-\chapter{\label{cha:Running-the-Solver}Running the Solver \texttt{xspecfem3D}}
-
-Now that you have successfully generated the databases, you are ready
-to compile the solver. In the main directory, type
-\begin{lyxcode}
-{\small make~xspecfem3D}{\small \par}
-\end{lyxcode}
-Please note that \texttt{xspecfem3D} must be called directly from
-the main directory, as most of the binaries of the package.\\
-
-\noindent
-The solver needs three input files in the \texttt{DATA} directory
-to run:
-\begin{description}
-\item [{\texttt{Par\_file}}] the main parameter file which was discussed
-in detail in the previous Chapter~\ref{cha:Creating-Distributed-Databases},
-\item [{\texttt{CMTSOLUTION}} or {\texttt{FORCESOLUTION}}] the earthquake
-source parameter file or the force source parameter file, and
-\item [{\texttt{STATIONS}}] the stations file.
-\end{description}
-Most parameters in the \texttt{Par\_file} should be set prior to running
-the databases generation. Only the following parameters may be changed
-after running \texttt{xgenerate\_databases}:
-\begin{itemize}
-\item the simulation type control parameters: \texttt{SIMULATION\_TYPE}
-and \texttt{SAVE\_FORWARD}
-\item the time step parameters \texttt{NSTEP} and \texttt{DT}
-\item the absorbing boundary control parameter \texttt{PML\_CONDITIONS}
-on condition that the\\
- \texttt{PML\_INSTEAD\_OF\_FREE\_SURFACE} flag remains unmodified
-after running the databases generation.
-\item the movie control parameters \texttt{MOVIE\_SURFACE}, \texttt{MOVIE\_VOLUME},
-and \texttt{NTSTEPS\_BETWEEN\_FRAMES}
-\item the ShakeMap\textregistered{}option \texttt{CREATE\_SHAKEMAP}
-\item the output information parameters \texttt{MOVIE\_TYPE}, \texttt{NTSTEP\_BETWEEN\_OUTPUT\_INFO}
-and \texttt{NTSTEP\_BETWEEN\_OUTPUT\_}~\\
- \texttt{SEISMOS}
-\item the \texttt{PRINT\_SOURCE\_TIME\_FUNCTION} flags
-\end{itemize}
-Any other change to the \texttt{Par\_file} implies rerunning both
-the database generator \texttt{xgenerate\_databases} and the solver
-\texttt{xspecfem3D}.
-
-For any particular earthquake, the \texttt{CMTSOLUTION} file that
-represents the point source may be obtained directly from the Harvard
-Centroid-Moment Tensor (CMT) web page \urlwithparentheses{www.seismology.harvard.edu}.
-It looks like the example shown in Fig. \ref{fig:CMTSOLUTION-file}.
-%\begin{lyxcode}
-\begin{figure}[htp]
-\noindent
-\begin{centering}
-{\small \includegraphics[width=0.8\textwidth]{figures/Hollywood_CMT}
-}
-\par
-\end{centering}
-\caption{\texttt{CMTSOLUTION} file based on the
-format from the Harvard CMT catalog. \textbf{M} is the moment tensor,
-$M_{0}${\small {} }is the seismic moment, and $M_{w}$ is the moment
-magnitude.}
-\label{fig:CMTSOLUTION-file}
-\end{figure}
-%\end{lyxcode}
-The \texttt{CMTSOLUTION} file should be edited in the following way:
-\begin{itemize}
-\item Set the latitude or UTM $x$ coordinate, longitude or UTM $y$ coordinate, depth of the source (in km).
-\item Set the \texttt{time shift} parameter equal to $0.0$ (the solver
-will not run otherwise.) The time shift parameter would simply apply
-an overall time shift to the synthetics, something that can be done
-in the post-processing (see Section \ref{sec:Process-data-and-syn}).
-\item For point-source simulations (see finite sources, page \pageref{To-simulate-a})
-we recommend setting the source half-duration parameter \texttt{half
-duration} equal to zero, which corresponds to simulating a step source-time
-function, i.e., a moment-rate function that is a delta function. If
-\texttt{half duration} is not set to zero, the code will use a Gaussian
-(i.e., a signal with a shape similar to a `smoothed triangle', as
-explained in \citet{KoTr02a} and shown in Fig~\ref{fig:gauss.vs.triangle})
-source-time function with half-width \texttt{half duration}. We prefer
-to run the solver with \texttt{half duration} set to zero and convolve
-the resulting synthetic seismograms in post-processing after the run,
-because this way it is easy to use a variety of source-time functions
-(see Section \ref{sec:Process-data-and-syn}). \citet{KoTr02a} determined
-that the noise generated in the simulation by using a step source
-time function may be safely filtered out afterward based upon a convolution
-with the desired source time function and/or low-pass filtering. Use
-the serial code \texttt{convolve\_source\_timefunction.f90} and the
-script \texttt{convolve\_source\_timefunction.csh} for this purpose,
-or alternatively use signal-processing software packages such as SAC
-\urlwithparentheses{www.llnl.gov/sac}. Type
-
-\begin{lyxcode}
-{\small make~xconvolve\_source\_timefunction}{\small \par}
-\end{lyxcode}
-
-to compile the code and then set the parameter \texttt{hdur} in \texttt{convolve\_source\_timefunction.csh}
-to the desired half-duration.
-
-\item The zero time of the simulation corresponds to the center of the triangle/Gaussian,
-or the centroid time of the earthquake. The start time of the simulation
-is $t=-1.5*\texttt{half duration}$ (the 1.5 is to make sure the moment
-rate function is very close to zero when starting the simulation).
-To convert to absolute time $t_{\mathrm{abs}}$, set
-
-\begin{lyxcode}
-$t_{\mathrm{abs}}=t_{\mathrm{pde}}+\texttt{time shift}+t_{\mathrm{synthetic}}$
-\end{lyxcode}
-
-where $t_{\mathrm{pde}}$ is the time given in the first line of the
-\texttt{CMTSOLUTION}, \texttt{time shift} is the corresponding value
-from the original \texttt{CMTSOLUTION} file and $t_{\mathrm{synthetic}}$
-is the time in the first column of the output seismogram.
-
-\end{itemize}
-\begin{figure}[htp]
-\noindent \begin{centering}
-\includegraphics[width=3.5in]{figures/gauss_vs_triangle_mod}
-\par\end{centering}
-
-\caption{Comparison of the shape of a triangle and the Gaussian function actually
-used.}
-
-\label{fig:gauss.vs.triangle}
-\end{figure}
-
-If you know the earthquake source in strike/dip/rake format rather
-than in \texttt{CMTSOLUTION} format, use the C code \texttt{SPECFEM3D\_GLOBE/utils/strike\_dip\_rake\_to\_CMTSOLUTION.c}
-to convert it. The conversion formulas are given for instance in \citet{AkRi80}.
-Note that the \citet{AkRi80} convention is slightly different from
-the Harvard \texttt{CMTSOLUTION} convention (the sign of some components
-is different). The C code outputs both.
-
-Centroid latitude and longitude should be provided in geographical
-coordinates. The code converts these coordinates to geocentric coordinates~\citep{DaTr98}.
-Of course you may provide your own source representations by designing
-your own \texttt{CMTSOLUTION} file. Just make sure that the resulting
-file adheres to the Harvard CMT conventions (see Appendix~\ref{cha:Coordinates}).
-Note that the first line in the \texttt{CMTSOLUTION} file is the Preliminary
-Determination of Earthquakes (PDE) solution performed by the USGS
-NEIC, which is used as a seed for the Harvard CMT inversion. The PDE
-solution is based upon P waves and often gives the hypocenter of the
-earthquake, i.e., the rupture initiation point, whereas the CMT solution
-gives the `centroid location', which is the location with dominant
-moment release. The PDE solution is not used by our software package
-but must be present anyway in the first line of the file.
-
-\label{To-simulate-a}To simulate a kinematic rupture, i.e., a finite-source
-event, represented in terms of $N_{\mathrm{sources}}$ point sources,
-provide a \texttt{CMTSOLUTION} file that has $N_{\mathrm{sources}}$
-entries, one for each subevent (i.e., concatenate $N_{\mathrm{sources}}$
-\texttt{CMTSOLUTION} files to a single \texttt{CMTSOLUTION} file).
-At least one entry (not necessarily the first) must have a zero \texttt{time
-shift}, and all the other entries must have non-negative \texttt{time
-shift}. Each subevent can have its own half duration, latitude, longitude,
-depth, and moment tensor (effectively, the local moment-density tensor).
-
-Note that the zero in the synthetics does NOT represent the hypocentral
-time or centroid time in general, but the timing of the \textit{center}
-of the source triangle with zero \texttt{time shift} (Fig~\ref{fig:source_timing}).
-
-Although it is convenient to think of each source as a triangle, in
-the simulation they are actually Gaussians (as they have better frequency
-characteristics). The relationship between the triangle and the gaussian
-used is shown in Fig~\ref{fig:gauss.vs.triangle}. For finite fault
-simulations it is usually not advisable to use a zero half duration
-and convolve afterwards, since the half duration is generally fixed
-by the finite fault model.
-
-\vspace{1cm}
-
-\noindent The \texttt{FORCESOLUTION} file should be edited in the
-following way:
-\begin{itemize}
-\item Set the \texttt{time shift} parameter equal to $0.0$ (the solver
-will not run otherwise.) The time shift parameter would simply apply
-an overall time shift to the synthetics, something that can be done
-in the post-processing (see Section \ref{sec:Process-data-and-syn}).
-\item Set the \texttt{f0} parameter (the dominant frequency) of the Ricker source time
-function when {\texttt{USE\_RICKER\_TIME\_FUNCTION}} is turned on
-in the main parameter file {\texttt{Par\_file}}. In case that the
-solver uses a (pseudo) Dirac delta source time function to represent
-a force point source, a very short duration of five time steps is automatically set by default.
-\item Set the latitude or UTM $x$ coordinate, longitude or UTM $y$ coordinate, depth of the source (in km).
-\item Set the magnitude of the force source.
-\item Set the components of a (non-unitary) direction vector for the force
-source in the East/North/Vertical basis (see Appendix A for the orientation
-of the reference frame).
-\end{itemize}
-\noindent Where necessary, set a \texttt{FORCESOLUTION} file in the
-same way you configure a \texttt{CMTSOLUTION} file with $N_{\mathrm{sources}}$
-entries, one for each subevent (i.e., concatenate $N_{\mathrm{sources}}$
-\texttt{FORCESOLUTION} files to a single \texttt{FORCESOLUTION} file).
-At least one entry (not necessarily the first) must have a zero \texttt{time
-shift}, and all the other entries must have non-negative \texttt{time
-shift}. Each subevent can have its own half latitude, longitude, depth,
-\texttt{half duration} and force parameters.
-
-\begin{figure}[H]
-\noindent \begin{centering}
-{\small \includegraphics[width=5in]{figures/source_timing} }
-\par\end{centering}
-
-\caption{Example of timing for three sources. The center of the first source
-triangle is defined to be time zero. Note that this is NOT in general
-the hypocentral time, or the start time of the source (marked as tstart).
-The parameter \texttt{time shift} in the \texttt{CMTSOLUTION} file
-would be t1(=0), t2, t3 in this case, and the parameter \texttt{half
-duration} would be hdur1, hdur2, hdur3 for the sources 1, 2, 3 respectively.}
-
-{\small \label{fig:source_timing} }
-\end{figure}
-
-The solver can calculate seismograms at any number of stations for
-basically the same numerical cost, so the user is encouraged to include
-as many stations as conceivably useful in the \texttt{STATIONS} file,
-which looks like this:
-
-\begin{figure}[H]
-\noindent \begin{centering}
-{\small \includegraphics{figures/STATIONS_basin_explained} }
-\par\end{centering}
-
-\label{fig:Sample-STATIONS-file.}\caption{Sample \texttt{STATIONS} file. Station latitude and longitude should
-be provided in geographical coordinates. The width of the station
-label should be no more than 32 characters (see \texttt{MAX\_LENGTH\_STATION\_NAME}
-in the \texttt{constants.h} file), and the network label should be
-no more than 8 characters (see \texttt{MAX\_LENGTH\_NETWORK\_NAME}
-in the \texttt{constants.h} file).}
-\end{figure}
-
-Each line represents one station in the following format:
-\begin{lyxcode}
-{\small Station~Network~Latitude~(degrees)~Longitude~(degrees)~Elevation~(m)~burial~(m)~}{\small \par}
-\end{lyxcode}
-The solver \texttt{xspecfem3D} filters the list of stations in file
-\texttt{DATA/STATIONS} to exclude stations that are not located within
-the region given in the \texttt{Par\_file} (between \texttt{LATITUDE\_MIN}
-and \texttt{LATITUDE\_MAX} and between \texttt{LONGITUDE\_MIN} and
-\texttt{LONGITUDE\_MAX}). The filtered file is called \texttt{DATA/STATIONS\_FILTERED}.
-
-Solver output is provided in the \texttt{OUTPUT\_FILES} directory
-in the \texttt{output\_solver.txt} file. Output can be directed to
-the screen instead by uncommenting a line in \texttt{constants.h}:
-\begin{lyxcode}
-!~uncomment~this~to~write~messages~to~the~screen~
-
-!~integer,~parameter~::~IMAIN~=~ISTANDARD\_OUTPUT~
-\end{lyxcode}
-On PC clusters the seismogram files are generally written to the local
-disks (the path \texttt{LOCAL\_PATH} in the \texttt{Par\_file}) and
-need to be gathered at the end of the simulation.
-
-While the solver is running, its progress may be tracked by monitoring
-the `\texttt{\small timestamp{*}}{\small ' files in the ~}\\
-{\small{} }\texttt{\small OUTPUT\_FILES}{\small{} directory. These tiny
-files look something like this:}{\small \par}
-\begin{lyxcode}
-Time~step~\#~~~~~~~~~~10000~
-
-Time:~~~~~108.4890~~~~~~seconds~
-
-Elapsed~time~in~seconds~=~~~~1153.28696703911~
-
-Elapsed~time~in~hh:mm:ss~=~~~~~0~h~19~m~13~s~
-
-Mean~elapsed~time~per~time~step~in~seconds~=~~~~~0.115328696703911~
-
-Max~norm~displacement~vector~U~in~all~slices~(m)~=~~~~1.0789589E-02~
-\end{lyxcode}
-The \texttt{\small timestamp{*}}{\small{} files provide the }\texttt{\small Mean
-elapsed time per time step in seconds}{\small , which may be used
-to assess performance on various machines (assuming you are the only
-user on a node), as well as the }\texttt{\small Max norm displacement
-vector U in all slices~(m)}{\small . If something is wrong with the
-model, the mesh, or the source, you will see the code become unstable
-through exponentially growing values of the displacement and fluid
-potential with time, and ultimately the run will be terminated by
-the program. You can control the rate at which the timestamp files
-are written based upon the parameter }\texttt{\small NTSTEP\_BETWEEN\_OUTPUT\_INFO}{\small{}
-in the }\texttt{\small Par\_file}{\small .}{\small \par}
-
-Having set the \texttt{Par\_file} parameters, and having provided
-the \texttt{CMTSOLUTION} (or the \texttt{FORCESOLUTION}) and \texttt{STATIONS}
-files, you are now ready to launch the solver! This is most easily
-accomplished based upon the \texttt{go\_solver} script (See Chapter
-\ref{cha:Scheduler} for information about running through a scheduler,
-e.g., LSF). You may need to edit the last command at the end of the
-script that invokes the \texttt{mpirun} command. The \texttt{runall}
-script compiles and runs both \texttt{xgenerate\_databases} and \texttt{xspecfem3D}
-in sequence. This is a safe approach that ensures using the correct
-combination of distributed database output and solver input.
-
-It is important to realize that the CPU and memory requirements of
-the solver are closely tied to choices about attenuation (\texttt{ATTENUATION})
-and the nature of the model (i.e., isotropic models are cheaper than
-anisotropic models). We encourage you to run a variety of simulations
-with various flags turned on or off to develop a sense for what is
-involved.
-
-For the same model, one can rerun the solver for different events
-by simply changing the \texttt{CMTSOLUTION} or \texttt{FORCESOLUTION}
-file, or for different stations by changing the \texttt{STATIONS}
-file. There is no need to rerun the \texttt{xgenerate\_databases}
-executable. Of course it is best to include as many stations as possible,
-since this does not add to the cost of the simulation.
-
-We have also added the ability to run several calculations (several earthquakes)
-in an embarrassingly-parallel fashion from within the same run;
-this can be useful when using a very large supercomputer to compute
-many earthquakes in a catalog, in which case it can be better from
-a batch job submission point of view to start fewer and much larger jobs,
-each of them computing several earthquakes in parallel.
-To turn that option on, set parameter \texttt{NUMBER\_OF\_SIMULTANEOUS\_RUNS}
-to a value greater than 1 in file setup/constants.h.in before
-configuring and compiling the code.
-When that option is on, of course the number of processor cores used to start
-the code in the batch system must be a multiple of \texttt{NUMBER\_OF\_SIMULTANEOUS\_RUNS},
-all the individual runs must use the same number of processor cores,
-which as usual is \texttt{NPROC} in the input file \texttt{DATA/Par\_file},
-and thus the total number of processor cores to request from the batch system
-should be \texttt{NUMBER\_OF\_SIMULTANEOUS\_RUNS $\times$ NPROC}.
-All the runs to perform must be placed in directories called \texttt{run0001}, \texttt{run0002}, \texttt{run0003} and so on (with exactly four digits)
-and you must create a link from the root directory of the code
-to the first copy of the executable programs by typing \texttt{ln -s run0001/bin bin}.
-
-\chapter{\label{cha:Fault-Kinematics-Dynamics}Kinematic and dynamic fault
-sources}
-
-SPECFEM3D can handle finite fault sources of two kinds:
-\begin{enumerate}
-\item \emph{Kinematic}: the spatio-temporal distribution of slip rate is
-prescribed along the fault surface
-\item \emph{Dynamic}: a friction law and initial stresses are prescribed
-on the fault, a spontaneous rupture process is computed.
-\end{enumerate}
-
-\section{\label{sec:Mesh-Generation-with}Mesh Generation with Split Nodes}
-
-Faults need to be handled in a special way during mesh generation.
-A fault surface must lie at the interface between elements (the mesh
-must honor the fault surfaces). Moreover, a fault is made of two surfaces
-in contact. Each of these two surfaces needs a separate set of nodes.
-This approach is known as \textquotedbl{}split nodes\textquotedbl{}.
-Currently faults can only be run with \texttt{xdecompose\_mesh} and
-\texttt{CUBIT}. \texttt{xmeshfem3D} is not yet ready to handle faults.\\
-To facilitate the mesh generation with split nodes in CUBIT, we need
-to separate the two fault surfaces by a small distance, effectively
-creating a tiny opening of the fault (Figure \ref{fig:examples.splitnodes},
-\ref{fig:examples.splitnodes-surfacetrace}). Note that only the interior
-of the fault must be opened, its edges must remain closed (except
-the edge on the free surface). The fault is automatically closed later
-by SPECFEM3D.\\
-
-Here is an example CUBIT script to generate a mesh with split nodes
-for a buried vertical strike-slip fault:
-\begin{lyxcode}
-~reset~~\\
-~brick~x~10~y~10~z~10~~\\
-~webcut~volume~all~with~plane~xplane~~\\
-~webcut~volume~all~with~plane~yplane~~\\
-~webcut~volume~all~with~plane~xplane~offset~3~~\\
-~webcut~volume~all~with~plane~zplane~offset~3~~\\
-~webcut~volume~all~with~plane~zplane~offset~-3~~\\
-~imprint~all~~\\
-~merge~all~~\\
-~unmerge~surf~160~~\\
-~mesh~vol~all~~\\
-~set~node~constraint~off~~\\
-~node~in~surf~168~move~X~0~Y~0.01~Z~0~~\\
-~node~in~surf~160~move~X~0~Y~-0.01~Z~0~\\
-
-\end{lyxcode}
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[scale=0.3]{figures/faultmesh} \\
- \includegraphics[scale=0.3]{figures/surf168} \includegraphics[scale=0.3]{figures/splitnodes}
-\par\end{centering}
-
-\caption{Screenshots of the CUBIT example for the embedded fault with split
-nodes: The entire mesh is decomposed into several volumes bounding
-the fault (top), the fault surface 168 is shown as orange squares
-(middle) and split nodes of the fault shown as orange dots(bottom).
-Note that only interior nodes of the fault are split while those on
-the edges of the fault surface are touching each other.}
-
-\label{fig:examples.splitnodes}
-\end{figure}
-
-\begin{lyxcode}
-
-\end{lyxcode}
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[scale=0.55]{figures/splitnodes_surfacetrace} \\
-
-\par\end{centering}
-
-\caption{Screenshots of a CUBIT example showing split nodes for a fault reaching
-the surface. Surface trace of the fault is shown in orange. Note that
-the edges of the fault are not split while the interior nodes are
-offset by a small distance on either side of the fault}
-
-\label{fig:examples.splitnodes-surfacetrace}
-\end{figure}
-
-The CUBIT scripts ({*}.jou and {*}.py) in the directory \texttt{EXAMPLES}
-generate more complicated meshes. The {*}.py files are Python scripts
-that execute CUBIT commands and use the CUBIT-python interface for
-SPECFEM3D (see next section). The Python language allows to define
-and manipulate variables to parameterize the mesh. Alternatively,
-the Python script can call a CUBIT journal file ({*}.jou), which looks
-like the example above. Variables can be defined and manipulated there
-using the $\mathtt{APREPRO}$ language built in CUBIT.\\
-
-Note that you should avoid gaps in the list of indices of mesh objects
-with the following CUBIT command:
-\begin{lyxcode}
-$\mathtt{compress\: ids\: hex\: face\: edge\: node}$
-\end{lyxcode}
-(otherwise you will get a segmentation fault during domain decomposition)
-
-\section{CUBIT-Python Scripts for Faults}
-
-The mesh generated in CUBIT needs to be processed and exported in
-a format compatible with SPECFEM3D. This is achieved in the Python
-scripts by calling the Python-CUBIT interface functions defined in
-the CUBIT directory:
-\begin{enumerate}
-\item Function $\mathtt{define\_bc}$ (or $\mathtt{boundary\_definition.py}$)
-must be called to set up the absorbing boundaries database
-\item Function $\mathtt{fault\_input}$ must be called once for each fault
-to set up the fault database
-\item Function $\mathtt{cubit2specfem3d.export2SPECFEM3D}$ must be called
-at the very end of the script to export the mesh in a SPECFEM3D format.
-\end{enumerate}
-The functions in \#1 and \#3 are for the bulk and are documented in
-Section \ref{subsec:Exporting-the-Mesh}. We focus here on \#2:
-\begin{description}
-\item [{Function:}] $\mathtt{fault\_input}$
-\item [{Purpose:}] export a fault mesh database from CUBIT to a SPECFEM3D-compliant
-file
-\item [{Syntax:}] fault\_input(id\_fault, ids\_surf\_1, ids\_surf\_2)
-\item [{Inputs:}]~\end{description}
-\begin{itemize}
-\item id\_fault integer index assigned to the fault. (The user must number
-all the faults, starting at 1, with unit increments)
-\item ids\_surf\_1 list of CUBIT ids of all surfaces that form side 1 of
-the fault.
-\item ids\_surf\_2 list of CUBIT ids of all surfaces that form side 2 of
-the fault. (The user must decide which side of the fault is side 1.
-This choice affects the sign conventions of fault quantities as explained
-in Section \ref{sec:Sign-Convention-for}).\end{itemize}
-\begin{description}
-\item [{Outputs:}] file $\mathtt{fault\_file\_X.dat}$, where X is the
-fault id (id\_fault).
-\item [{Example:}] For the example in Section \ref{sec:Mesh-Generation-with}:\end{description}
-\begin{lyxcode}
-$\hphantom{}$A1~=~{[}168{]}
-
-$\hphantom{}$A2~=~{[}160{]}
-
-$\hphantom{}$fault\_input(1,A1,A2)
-\end{lyxcode}
-
-\section{Examples}
-
-The package includes examples, the SCEC benchmark problems:
-\begin{itemize}
-\item TPV5, a planar vertical strike-slip fault
-\item TPV14 and TPV15, vertical strike-slip fault system with a fault branch
-\item TPV102 and TPV103, rate-and-state benchmarks
-\item TPV22 and TPV23, step-over benchmarks
-\item TPV16, heterogenous initial stress
-\end{itemize}
-and
-\begin{itemize}
-\item Splay fault models from Wendt et al. (2009)\\
-
-\end{itemize}
-Read the documents in the directory \texttt{EXAMPLES/*/description}.
-They contain a description of the example and additional instructions
-to run it. Visualize the results with the Matlab scripts in the directory
-$\mathtt{EXAMPLES/*/post}$
-
-\section{\label{sec:Sign-Convention-for}Sign Convention for Fault Quantities}
-
-During mesh generation, the fault is defined by two surfaces in contact.
-Let's denote as \textquotedbl{}\emph{side 1}\textquotedbl{} the SECOND
-surface declared by the user in the call to the python function \textquotedbl{}\emph{fault\_input}\textquotedbl{},
-and the FIRST surface as \textquotedbl{}\emph{side 2}\textquotedbl{}.
-The local coordinate system on the fault is defined as the right-handed
-coordinate system defined by (strike, dip, normal), where \textquotedbl{}\emph{normal}\textquotedbl{}
-is the normal vector outgoing from side 1, \textquotedbl{}\emph{dip}\textquotedbl{}
-is parallel to the along-dip direction pointing downwards, and \textquotedbl{}\emph{strike}\textquotedbl{}
-is the horizontal along-strike vector such that the system is right-handed.\\
-
-Slip is defined as displacement on side 2 minus displacement on side
-1. In the local coordinate system on the fault, positive along-strike
-slip is right-lateral and positive along-dip slip is thrust if side
-1 is on the hanging wall (normal faulting if side 1 is on the foot
-wall).\\
-
-Traction is defined as the stress induced on side 1 by side 2, which
-is the stress tensor times the normal vector outgoing from side 1.
-In the local coordinate system on the fault, the normal traction is
-negative in compression, positive along-strike traction generates
-right-lateral slip and positive along-dip traction generates thrust
-slip if side 1 is on the hanging wall (normal faulting if side 1 is
-on the foot wall).
-
-\section{Input Files}
-
-Three additional input files are required in the \texttt{DATA/} directory
-for dynamic and kinematic faults. They are $\mathtt{Par\_file\_faults}$,
-$\mathtt{FAULT\_STATIONS}$ and $\mathtt{input\_file.txt}$ (optional).
-If the former does not exist, the code assumes that there are no faults.\\
-
-\begin{lyxlist}{00.00.0000}
-\item [{\textbf{DATA/Par\_file\_faults}}] contains parameters of the fault.
-The first part of this file has a strict format:
-
-\begin{description}
-\item [{Line}] 1: Number of faults (NF)
-\item [{Lines}] 2 to NF+1: Kelvin Voigt damping (in seconds) for each fault.
-(See below how to set this parameter)
-\item [{Line}] NF+2: Type of simulation (1=dynamic , 2 = kinematic)
-\item [{Line}] NF+3: Number of time steps between updates of the time series
-outputs at selected fault points (see DATA/FAULT\_STATIONS), usually
-a large number (100s or 1000s). Note that the sampling rate of the
-time series is usually much higher.
-\item [{Line}] NF+4: Number of time steps between fault snapshot outputs
-(quantities at every fault point exported at regular times), usually
-a large number (100s or 1000s).
-\item [{Line}] NF+5: Slip velocity threshold below which frictional healing
-is set (friction coefficient is reset to its static value). If this
-value is negative healing is disabled.
-\item [{Line}] NF+6: Slip velocity threshold to define the rupture front.
-Only used for outputs.
-\end{description}
-
-The rest of this file is made of namelist input blocks (see \textquotedbl{}namelist\textquotedbl{}
-in a Fortran 9x manual). The input for each fault has the following
-sequence (arguments in {[}brackets{]} are optional):
-
-\&\textbf{BEGIN\_FAULT} /
-
-\&\textbf{INIT\_STRESS} S1, S2, S3 {[},n1, n2, n3{]} /
-
-followed by (n1+n2+n3) \&\textbf{DIST2D} blocks
-
-\&\textbf{SWF} mus, mud, dc {[}, nmus, nmud, ndc{]} /
-
-followed by (nmus+nmud+ndc) \&\textbf{DIST2D} blocks\\
-
-\begin{lyxlist}{00.00.0000}
-\item [{\&\textbf{INIT\_STRESS}}] input block sets the initial fault stresses
-relative to the foot-wall side of the fault. Initial stresses are
-composed of a constant background value possibly overwritten in prescribed
-regions by heterogeneous distributions (see \&\textbf{DIST2D} blocks
-below):
-
-\begin{description}
-\item [{S1}] = constant background value of along-strike shear stress (positive
-in the usual strike direction)
-\item [{S2}] = constant background value of along-dip shear (positive is
-down-dip, normal faulting)
-\item [{S3}] = constant background value of normal stress (negative in
-compression)
-\item [{n1}] = number of heterogeneous items for along-strike shear stress
-{[}default is 0{]}
-\item [{n2}] = number of heterogeneous items for along-dip shear stress
-{[}default is 0{]}
-\item [{n3}] = number of heterogeneous items for normal stress {[}default
-is 0{]}
-\end{description}
-\item [{\&\textbf{SWF}}] input block sets the slip-weakening friction parameters
-of the fault:
-
-\begin{description}
-\item [{mus}] = constant background value of static friction coefficient
-\item [{mud}] = constant background value of dynamic friction coefficient
-\item [{dc}] = constant background value of critical slip-weakening distance
-\item [{nmus}] = number of heterogeneous items for static friction coefficient
-{[}default is 0{]}
-\item [{nmud}] = number of heterogeneous items for dynamic friction coefficient
-{[}default is 0{]}
-\item [{ndc}] = number of heterogeneous items for critical slip-weakening
-distance {[}default is 0{]}\\
-
-\end{description}
-\item [{\&\textbf{DIST2D}}] input blocks modify (overwrite) the value of
-a fault parameter by a heterogeneous spatial distribution:
-
-\&\textbf{DIST2D} shape='$\mathtt{square}$', val, xc, yc, zc, l /
-
-$\;$sets a constant value (val) within a cube with center (xc,yc,zc)
-and edge size l.
-
-\&\textbf{DIST2D} shape='$\mathtt{rectangle}$', val, xc, yc, zc,
-lx, ly, lz /
-
-$\;$sets a constant value (val) within a rectangular prism with center
-(xc,yc,zc) and edge sizes (lx,ly,lz).
-
-\&\textbf{DIST2D} shape='$\mathtt{rectangle-taper}$', val, valh,
-xc, yc, zc, lx, ly, lz /
-
-$\;$sets a vertical linear gradient within a rectangular prism with
-center (xc,yc,zc) and edge sizes (lx,ly,lz). Values vary linearly
-as a function of vertical position z between value val at z = zc-lz/2
-and value valh at z = zc+lz/2 .
-
-\&\textbf{DIST2D} shape='$\mathtt{circular}$', val, xc, yc, zc, r
-/
-
-$\;$sets a constant value (val) within a sphere with center (xc,yc,zc)
-and radius r. \\
-
-\end{lyxlist}
-\end{lyxlist}
-
-\begin{lyxlist}{00.00.0000}
-\item [{\textbf{DATA/FAULT\_STATIONS}}] Stations in the fault plane.
-
-\textbf{Line 1} : number of stations.
-
-\textbf{Line 2 to end}: 5 columns: X, Y, Z (-depth), station name,
-fault-id
-
-The fault-id identifies the fault that contains the station. It is
-the index of appearance in the faults list after line 2 of Par\_file\_faults
-
-\end{lyxlist}
-
-\begin{lyxlist}{00.00.0000}
-\item [{\textbf{DATA/input\_file.txt}}] Heterogeneous stresses and friction
-parameters are documented in page 10 of
-
-\begin{lyxlist}{00.00.0000}
-\item [{$\mathtt{EXAMPLES/tpv16/description/TPV16\_17\_Description\_v03.pdf}$}]~
-\end{lyxlist}
-
-To activate this feature, in fault\_solver\_dynamic.f90 set \textbf{TPV16}
-= .true. then re-compile the code.
-
-\end{lyxlist}
-
-\section{\label{sec:Setting-the-Kelvin-Voigt}Setting the Kelvin-Voigt Damping
-Parameter }
-
-The purpose of the Kelvin-Voigt viscosity in the dynamic fault solver
-is to damp spurious oscillations generated by the fault slip at frequencies
-that are too high to be resolved by the mesh. The viscosity $\mathtt{eta}$
-(in seconds) depends on the size of the elements on the fault. Here
-is how to set it:
-\begin{enumerate}
-\item Determine the average linear size of the elements on the fault plane,
-$\mathtt{h\_fault}$. Usually this value is prescribed by the user
-during mesh generation. Otherwise it can be found by inspection of
-the mesh inside the CUBIT GUI.
-\item Use the Matlab function $\mathtt{utils/critical\_timestep.m}$ to
-compute\\
- $\mathtt{dtc\_fault}=\mathtt{critical\_timestep\left(c_{p},h\_fault,ngll\right)}$.
-\\
-This is the critical time step in an elastic medium for a hypothetical
-element of cubic shape with size equal to $\mathtt{h\_fault}$.
-\item Set $\mathtt{eta}$ in $\mathtt{Par\_file\_faults}$ to (0.1 to 0.3)$\ensuremath{\times}\mathtt{dtc\_fault}$.
-A larger $\mathtt{eta}$ damps high-frequencies more aggressively
-but it might also affect lower frequencies and rupture speed.
-\end{enumerate}
-Viscosity reduces numerical stability: the critical timestep in a
-simulation with Kelvin-Voigt damping needs to be smaller than that
-in a purely elastic simulation. Here is how to set the time step accordingly:
-\begin{enumerate}
-\item Run a test simulation without viscosity ($\mathtt{eta}$=0 and only
-a few time steps)
-\item Look for the \textquotedbl{}maximum suggested time step\textquotedbl{}
-in \texttt{OUTPUT\_FILES/output\_mesher.txt}. This is the critical
-timestep of a purely elastic simulation, $\mathtt{dtc\_bulk}$.
-\item Reset the timestep of the simulation with a Kelvin-Voigt material
-to a value smaller than\\
- $\mathtt{dtc\_kv}=\mathtt{eta}\left(\sqrt{1+\mathtt{dtc\_bulk^{2}}/\mathtt{eta^{2}}}-1\right)$
-\end{enumerate}
-Note that in general $\mathtt{dtc\_bulk}$ is smaller than $\mathtt{dtc\_fault}$,
-because elements off the fault might be smaller or more distorted
-than element faces on the fault.
-
-\section{Output Files }
-
-Several output files are saved in \texttt{OUTPUT\_FILES/}:\\
-
-\begin{enumerate}
-\item Seismograms for each station on the fault plane given in \texttt{DATA/FAUL\_STATIONS}.
-One output file is generated for each station, named after the station.
-The files are ascii and start with a header (22 lines long) followed
-by a data block with the following format, one line per time sample:\\
-
-Column 1 = Time (s)
-
-Column 2 = horizontal right-lateral slip (m)
-
-Column 3 = horizontal right-lateral slip rate (m/s)
-
-Column 4 = horizontal right-lateral shear stress (MPa)
-
-Column 5 = vertical up-dip slip (m)
-
-Column 6 = vertical up-dip slip rate (m/s)
-
-Column 7 = vertical up-dip shear stress (MPa)
-
-Column 8 = normal stress (MPa)\\
-
-The stresses are relative to the footwall side of the fault (this
-convention controls their sign, but not their amplitude). Slip is
-defined as displacement of the hanging wall relative to the footwall.
-
-\item Seismograms at stations in the bulk (out of the fault plane) given
-in \texttt{DATA/STATIONS}.
-\item Rupture time files are named $\mathtt{Rupture\_time\_FAULT}$-$\mathtt{id}$.
-One file is generated for each fault. The files are ascii and start
-with a header (12 lines long) followed by a data block with the following
-format, one line per fault node:
-
-Column 1 = horizontal coordinate, distance along strike (m)
-
-Column 2 = vertical coordinate, distance down-dip (m)
-
-Column 3 = rupture time (s)
-
-\item Fault quantities (slip, slip rate, stresses, etc) at regular times
-are stored in binary data files called $\mathtt{Snapshot\#it\#.bin}$,
-where \#it\# is the timestep number. These can be read in Matlab with
-the function $\mathtt{utils/FSEM3D\_snapshot.m}$
-\end{enumerate}
-
-\section{Post-processing and Visualization }
-
-Some Matlab functions for post-processing and visualization are included
-in directory $\mathtt{utils}$. The functions are internally documented
-(see their matlab help).\\
-
-\texttt{FSEM3D\_snapshot} reads a fault data snapshot\\
-
-The directories \texttt{EXAMPLES/*/post} contain additional Matlab
-scripts to generate figures specific to each example.
-
-\chapter{\label{cha:Adjoint-Simulations}Adjoint Simulations}
-
-Adjoint simulations are generally performed for two distinct applications.
-First, they can be used in point source moment-tensor inversions, or source imaging for
-earthquakes with large ruptures such as the Lander's earthquake \citep{WaHe94}.
-Second, they can be used to generate finite-frequency sensitivity
-kernels that are a critical part of tomographic inversions based upon
-3D reference models \citep{TrTaLi05,LiTr06,TrKoLi08,LiTr08}. In either
-case, source parameter or velocity structure updates are sought to
-minimize a specific misfit function (e.g., waveform or traveltime
-differences), and the adjoint simulation provides a means of computing
-the gradient of the misfit function and further reducing it in successive
-iterations. Applications and procedures pertaining to source studies
-and finite-frequency kernels are discussed in Sections~\ref{sec:Adjoint-simulation-sources}
-and \ref{sec:Adjoint-simulation-finite}, respectively. The two related
-parameters in the \texttt{Par\_file} are \texttt{SIMULATION\_TYPE}
-(1, 2 or 3) and the \texttt{SAVE\_FORWARD} (boolean).
-
-\section{\label{sec:Adjoint-simulation-sources}Adjoint Simulations for Sources
-Only (not for the Model)}
-
-When a specific misfit function between data and synthetics is minimized to invert
-for earthquake source parameters, the gradient of the misfit function
-with respect to these source parameters can be computed by placing
-time-reversed seismograms at the receivers as virtual sources
-in an adjoint simulation. Then the value of the gradient is obtained
-from the adjoint seismograms recorded at the original earthquake location.
-\begin{enumerate}
-\item \textbf{Prepare the adjoint sources} \label{enu:Prepare-the-adjoint}
-
-\begin{enumerate}
-\item First, run a regular forward simlation (\texttt{SIMULATION\_TYPE =
-1} and \texttt{SAVE\_FORWARD = .false.}). You can automatically set
-these two variables using the \texttt{\small utils/change\_simulation\_type.pl}{\small{}
-script:}{\small \par}
-
-\begin{lyxcode}
-utils/change\_simulation\_type.pl~-f~
-\end{lyxcode}
-
-and then collect the recorded seismograms at all the stations given
-in \texttt{DATA/STATIONS}.
-
-\item Then select the stations for which you want to compute the time-reversed
-adjoint sources and run the adjoint simulation, and compile them into
-the \texttt{DATA/STATIONS\_ADJOINT} file, which has the same format
-as the regular \texttt{DATA/STATIONS} file.
-
-\begin{itemize}
-\item Depending on what type of misfit function is used for the source inversion,
-adjoint sources need to be computed from the original recorded seismograms
-for the selected stations and saved in a sub-directory called \texttt{SEM/}
-in the root directory of the code, with the format \texttt{NT.STA.BX?.adj}, where \texttt{NT},
-\texttt{STA} are the network code and station name given in the \texttt{DATA/STATIONS\_ADJOINT}
-file, and \texttt{BX?} represents the component name of a particular
-adjoint seismogram. Please note that the band code can change depending
-on your sampling rate (see Appendix~\ref{cha:channel-codes} for
-further details).
-\item The adjoint seismograms are in the same format as the original seismogram
-(\texttt{NT.STA.BX?.sem?}), with the same start time, time interval
-and record length.
-\end{itemize}
-\item Notice that even if you choose to time reverse only one component
-from one specific station, you still need to supply all three components
-because the code is expecting them (you can set the other two components
-to be zero).
-\item Also note that since time-reversal is done in the code itself, no
-explicit time-reversing is needed for the preparation of the adjoint
-sources, i.e., the adjoint sources are in the same forward time sense
-as the original recorded seismograms.
-\end{enumerate}
-\item \textbf{Set the related parameters and run the adjoint simulation}\\
- In the \texttt{DATA/Par\_file}, set the two related parameters to
-be \texttt{SIMULATION\_TYPE = 2} and \texttt{SAVE\_FORWARD = .false.}.
-More conveniently, use the scripts \texttt{utils/change\_simulation\_type.pl}
-to modify the \texttt{Par\_file} automatically (\texttt{change\_simulation\_type.pl
--a}). Then run the solver to launch the adjoint simulation.
-\item \textbf{Collect the seismograms at the original source location}
-
-After the adjoint simulation has completed successfully, collect the
-seismograms from \texttt{LOCAL\_PATH}.
-\begin{itemize}
-\item These adjoint seismograms are recorded at the locations of the original
-earthquake sources given by the \texttt{DATA/CMTSOLUTION} file, and
-have names of the form \texttt{NT.S?????.S??.sem} for the six-component
-strain tensor (\texttt{SNN,SEE,SZZ,SNE,SNZ,SEZ}) at these locations,
-and ~\\
- \texttt{NT.S?????.BX?.sem} for the three-component displacements
-(\texttt{BXN,BXE,BXZ}) recorded at these locations.
-\item \texttt{S?????} denotes the source number; for example, if the original
-\texttt{CMTSOLUTION} provides only a point source, then the seismograms
-collected will start with \texttt{S00001}.
-\item These adjoint seismograms provide critical information for the computation
-of the gradient of the misfit function.
-\end{itemize}
-\end{enumerate}
-
-\section{\label{sec:Adjoint-simulation-finite}Adjoint Simulations for Finite-Frequency
-Kernels (Kernel Simulation)}
-
-Finite-frequency sensitivity kernels are computed in two successive
-simulations (please refer to \citet{LiTr06} and \citet{TrKoLi08}
-for details).
-\begin{enumerate}
-\item \textbf{Run a forward simulation with the state variables saved at
-the end of the simulation}
-
-Prepare the \texttt{\small CMTSOLUTION}{\small{} and }\texttt{\small STATIONS}{\small{}
-files, set the parameters }\texttt{\small SIMULATION\_TYPE}{\small {}
-}\texttt{\small =}{\small {} }\texttt{\small 1}{\small{} and }\texttt{\small SAVE\_FORWARD
-=}{\small {} }\texttt{\small .true.}{\small{} in the }\texttt{\small Par\_file}{\small{}
-(}\texttt{\small change\_simulation\_type -F}{\small ), and run the
-solver.}{\small \par}
-\begin{itemize}
-\item Notice that attenuation is not implemented yet for the computation
-of finite-frequency kernels; therefore set \texttt{ATTENUATION = .false.}
-in the \texttt{Par\_file}.
-\item We also suggest you modify the half duration of the \texttt{CMTSOLUTION}
-to be similar to the accuracy of the simulation (see Equation \ref{eq:shortest_period})
-to avoid too much high-frequency noise in the forward wavefield, although
-theoretically the high-frequency noise should be eliminated when convolved
-with an adjoint wavefield with the proper frequency content.
-\item This forward simulation differs from the regular simulations (\texttt{\small SIMULATION\_TYPE}{\small {}
-}\texttt{\small =}{\small {} }\texttt{\small 1}{\small{} and }\texttt{\small SAVE\_FORWARD}{\small {}
-}\texttt{\small =}{\small {} }\texttt{\small .false.}{\small ) described
-in the previous chapters in that the state variables for the last
-time step of the simulation, including wavefields of the displacement,
-velocity, acceleration, etc., are saved to the }\texttt{\small LOCAL\_PATH}{\small{}
-to be used for the subsequent simulation. }{\small \par}
-\item For regional simulations, the files recording the absorbing boundary
-contribution are also written to the \texttt{LOCAL\_PATH} when \texttt{SAVE\_FORWARD
-= .true.}.
-\end{itemize}
-\item \textbf{Prepare the adjoint sources}
-
-The adjoint sources need to be prepared the same way as described
-in the Section \ref{enu:Prepare-the-adjoint}.
-\begin{itemize}
-\item In the case of travel-time finite-frequency kernel for one source-receiver
-pair, i.e., point source from the \texttt{CMTSOLUTION}, and one station
-in the \texttt{STATIONS\_ADJOINT} list, we supply a sample program
-in \texttt{utils/adjoint\_sources/traveltime/xcreate\_adjsrc\_traveltime}
-to cut a certain portion of the original displacement seismograms
-and convert it into the proper adjoint source to compute the finite-frequency
-kernel.
-
-\begin{lyxcode}
-xcreate\_adjsrc\_traveltime~t1~t2~ifile{[}0-5{]}~E/N/Z-ascii-files~{[}baz{]}
-\end{lyxcode}
-
-where \texttt{t1} and \texttt{t2} are the start and end time of the
-portion you are interested in, \texttt{ifile} denotes the component
-of the seismograms to be used (0 for all three components, 1 for East,
-2 for North, and 3 for vertical, 4 for transverse, and 5 for radial
-component), \texttt{E/N/Z-ascii-files} indicate the three-component
-displacement seismograms in the right order, and \texttt{baz} is the
-back-azimuth of the station. Note that \texttt{baz} is only supplied
-when \texttt{ifile} = 4 or 5.
-
-\item Similarly, a sample program to compute adjoint sources for amplitude
-finite-frequency kernels may be found in \texttt{utils/adjoint\_sources/amplitude}
-and used in the same way as described for traveltime measurements
-
-\begin{lyxcode}
-xcreate\_adjsrc\_amplitude~t1~t2~ifile{[}0-5{]}~E/N/Z-ascii-files~{[}baz{]}.
-\end{lyxcode}
-\end{itemize}
-\item \textbf{Run the kernel simulation}
-
-With the successful forward simulation and the adjoint source ready
-in the \texttt{SEM/} directory, set \texttt{SIMULATION\_TYPE
-= 3} and \texttt{SAVE\_FORWARD = .false.} in the \texttt{Par\_file}
-(you can use \texttt{change\_simulation\_type.pl -b}), and rerun the
-solver.
-\begin{itemize}
-\item The adjoint simulation is launched together with the back reconstruction
-of the original forward wavefield from the state variables saved from
-the previous forward simulation, and the finite-frequency kernels
-are computed by the interaction of the reconstructed forward wavefield
-and the adjoint wavefield.
-\item The back-reconstructed seismograms at the original station locations
-are saved to the \texttt{LOCAL\_PATH} at the end of the kernel simulations,
-and can be collected to the local disk.
-\item These back-constructed seismograms can be compared with the time-reversed
-original seismograms to assess the accuracy of the backward reconstruction,
-and they should match very well.
-\item The arrays for density, P-wave speed and S-wave speed kernels are
-also saved in the \texttt{LOCAL\_PATH} with the names \texttt{proc??????\_rho(alpha,beta)\_kernel.bin},
-where \texttt{proc??????} represents the processor number, \texttt{rho(alpha,beta)}
-are the different types of kernels.
-\end{itemize}
-\item \textbf{Run the anisotropic kernel simulation}
-
-Instead of the kernels for the isotropic wave speeds, you can also
-compute the kernels for the 21 independent components $C_{IJ},\, I,J=1,...,6$
-(using Voigt's notation) of the elastic tensor in the cartesian coordinate
-system. This is done by setting \texttt{ANISOTROPIC\_KL} \texttt{=}
-\texttt{.true.} in \texttt{constants.h} before compiling the package.
-The definition of the parameters $C_{IJ}$ in terms of the corresponding
-components $c_{ijkl},ijkl,i,j,k,l=1,2,3$ of the elastic tensor in
-cartesian coordinates follows \citet{ChTr07}. The 21 anisotropic
-kernels are saved in the \texttt{LOCAL\_PATH} in one file with the
-name of \texttt{proc??????\_cijkl\_kernel.bin} (with \texttt{proc??????}
-the processor number). The output kernels correspond to absolute perturbations
-$\delta C_{IJ}$ of the elastic parameters and their unit is in $s/GPa/km^{3}$.
-For consistency, the output density kernels with this option turned
-on are for a perturbation $\delta\rho$ (and not $\frac{\delta\rho}{\rho}$)
-and their unit is in s / (kg/m$^{3}$) / km$^{3}$. These `primary'
-anisotropic kernels can then be combined to obtain the kernels for
-different parameterizations of anisotropy. This can be done, for example,
-when combining the kernel files from slices into one mesh file (see
-Section~\ref{sec:Finite-Frequency-Kernels}).
-
-If \texttt{ANISOTROPIC\_KL} \texttt{=} \texttt{.true.} by additionally
-setting \texttt{ANISOTROPIC\_KL} \texttt{=} \texttt{.true.} in \texttt{constants.h}
-the package will save anisotropic kernels paramterized as velocities
-related to transverse isotropy based on the the Chen and Tromp parameters
-\citet{ChTr07}. The kernels are saved as relative perturbations for
-horizontal and vertical P and S velocities, $\alpha_{v},\alpha_{h},\beta_{v},\beta_{h}$.
-Explicit relations can be found in appendix B. of \citet{SiLiTrTr07b}
-
-The anisotropic kernels are only currently available for CPU mode.
-
-\end{enumerate}
-In general, the three steps need to be run sequentially to assure
-proper access to the necessary files. If the simulations are run through
-some cluster scheduling system (e.g., LSF), and the forward simulation
-and the subsequent kernel simulations cannot be assigned to the same
-set of computer nodes, the kernel simulation will not be able to access
-the database files saved by the forward simulation. Solutions for
-this dilemma are provided in Chapter~\ref{cha:Scheduler}. Visualization
-of the finite-frequency kernels is discussed in Section~\ref{sec:Finite-Frequency-Kernels}.
-
-%------------------------------------------------------------------------------------------------%
-
-\chapter{\label{cha:tomo}Doing tomography, i.e., updating the model based on the sensitivity kernels obtained}
-
-%------------------------------------------------------------------------------------------------%
-
-UNDER CONSTRUCTION (March 2015). Will be added around April / May 2015.
-
-\section{Tomographic inversion / imaging using the sensitivity kernels obtained}
-
-OLD VERSION OF THE SECTION, WILL SOON BE IMPROVED AND ENRICHED.\\
-
-One of the fundamental reasons for computing sensitivity kernels (Section~\ref{sec:Adjoint-simulation-finite})
-is to use them within a tomographic inversion. In other words, use
-recorded seismograms, make measurements with synthetic seismograms,
-and use the misfit between the two to iteratively improve the model
-described by (at least) $V_{{\rm p}}$, $V_{{\rm s}}$, and $\rho$
-as a function of space.
-
-Whatever misfit function you use for the tomographic inversion (you can find several
-examples in \citet{TrKoLi08} and \citet{TrTaLi05}), you will weight
-the sensitivity kernels with measurements. Furthermore, you will use
-as many measurements (stations, components, time windows) as possible
-per event; hence, we call these composite kernels ``event kernels,''
-which are volumetric fields representing the gradient of the misfit
-function with respect to one of the variables (\eg $V_{{\rm s}}$).
-The basic features of an adjoint-based tomographic inversion were
-illustrated in \citet{TaLiTr07} and \citet{TrKoLi08} using a conjugate-gradient
-algorithm; there are dozens of versions of gradient-based inversion
-algorithms that could alternatively be used. The tomographic inversion
-of \citet{TaLiMaTr09,TaLiMaTr2010} used SPECFEM3D Cartesian as well
-as several additional components which are also stored on the CIG
-svn server, described next.
-
-The directory containing external utilities for tomographic inversion using
-SPECFEM3D Cartesian (or other packages that evaluate misfit functions
-and gradients) is in directory \texttt{utils/ADJOINT\_TOMOGRAPHY\_TOOLS/}:
-\begin{verbatim}
-flexwin/     -- FLEXWIN algorithm for automated picking of time windows
-measure_adj/ -- reads FLEXWIN output file and makes measurements,
-                with the option for computing adjoint sources
-iterate_adj/ -- various tools for iterative inversion
-                (requires pre-computed "event kernels")
-\end{verbatim}
-This directory also contains
-a brief \verb+README+ file indicating the role of the three subdirectories,
-\verb+flexwin+ \citep{Maggi2009}, \verb+measure_adj+, and \verb+iterate_adj+.
-The components for making the model update are there; however, there
-are no explicit rules for computing the model update, just as with
-any optimization problem. There are options for computing a conjugate
-gradient step, as well as a source subspace projection step.
-
-The best single file to read is probably: \verb+ADJOINT_TOMO/iterate_adj/cluster/README+.
-
-%------------------------------------------------------------------------------------------------%
-
-\chapter{Noise Cross-correlation Simulations}
-
-%------------------------------------------------------------------------------------------------%
-
-Besides earthquake simulations, SPECFEM3D Cartesian includes functionality
-for seismic noise tomography as well. In order to proceed successfully
-in this chapter, it is critical that you have already familiarized
-yourself with procedures for meshing (Chapter \ref{cha:Mesh-Generation}),
-creating distributed databases (Chapter \ref{cha:Creating-Distributed-Databases}),
-running earthquake simulations (Chapters \ref{cha:Running-the-Solver})
-and adjoint simulations (Chapter \ref{cha:Adjoint-Simulations}).
-Also, make sure you read the article `Noise cross-correlation sensitivity
-kernels' \citep{trompetal2010}, in order to understand noise simulations
-from a theoretical perspective.
-
-\section{Input Parameter Files}
-
-As usual, the three main input files are crucial: \texttt{Par\_file},
-\texttt{CMTSOLUTION} and \texttt{STATIONS}. Unless otherwise specified,
-those input files should be located in directory \texttt{DATA/}.\\
-
-\texttt{CMTSOLUTION} is required for all simulations. At a first glance,
-it may seem unexpected to have it here, since the noise simulations
-should have nothing to do with the earthquake -- \texttt{CMTSOLUTION}.
-However, for noise simulations, it is critical to have no earthquakes.
-In other words, the moment tensor specified in \texttt{CMTSOLUTION}
-must be set to zero manually!\\
-
-\texttt{STATIONS} remains the same as in previous earthquake simulations,
-except that the order of receivers listed in \texttt{STATIONS} is
-now important. The order will be used to determine the `master' receiver,
-i.e., the one that simultaneously cross correlates with the others.\\
-
-\texttt{Par\_file} also requires careful attention. A parameter called
-\texttt{NOISE\_TOMOGRAPHY} has been added which specifies the type
-of simulation to be run. \texttt{NOISE\_TOMOGRAPHY} is an integer
-with possible values 0, 1, 2 and 3. For example, when \texttt{NOISE\_TOMOGRAPHY}
-equals 0, a regular earthquake simulation will be run. When it is
-1/2/3, you are about to run step 1/2/3 of the noise simulations respectively.
-Should you be confused by the three steps, refer to \citet{trompetal2010}
-for details.\\
-
-Another change to \texttt{Par\_file} involves the parameter \texttt{NSTEP}.
-While for regular earthquake simulations this parameter specifies
-the length of synthetic seismograms generated, for noise simulations
-it specifies the length of the seismograms used to compute cross correlations.
-The actual cross correlations are thus twice this length, i.e., $2~\mathrm{NSTEP}-1$.
-The code automatically makes the modification accordingly, if \texttt{NOISE\_TOMOGRAPHY}
-is not zero.\\
-
-There are other parameters in \texttt{Par\_file} which should be given
-specific values. For instance, since the first two steps for calculating
-noise sensitivity kernels correspond to forward simulations, \texttt{SIMULATION\_TYPE}
-must be 1 when \texttt{NOISE\_TOMOGRAPHY} equals 1 or 2. Also, we
-have to reconstruct the ensemble forward wavefields in adjoint simulations,
-therefore we need to set \texttt{SAVE\_FORWARD} to \texttt{.true.}
-for the second step, i.e., when \texttt{NOISE\_TOMOGRAPHY} equals
-2. The third step is for kernel constructions. Hence \texttt{SIMULATION\_TYPE}
-should be 3, whereas \texttt{SAVE\_FORWARD} must be \texttt{.false.}.\\
-
-Finally, for most system architectures, please make sure that \texttt{LOCAL\_PATH}
-in \texttt{Par\_file} is in fact local, not globally shared. Because
-we have to save the wavefields at the earth's surface at every time
-step, it is quite problematic to have a globally shared \texttt{LOCAL\_PATH},
-in terms of both disk storage and I/O speed.
-
-\section{Noise Simulations: Step by Step}
-
-Proper parameters in those parameter files are not enough for noise
-simulations to run. We have more parameters to specify: for example,
-the ensemble-averaged noise spectrum, the noise distribution etc.
-However, since there are a few `new' files, it is better to introduce
-them sequentially. In this section, standard procedures for noise
-simulations are described.
-
-\subsection{Pre-simulation}
-\begin{itemize}
-\item As usual, we first configure the software package using:
-
-\texttt{./configure FC=ifort MPIFC=mpif90}
-
-Use the following if SCOTCH is needed:
-
-\texttt{./configure FC=ifort MPIFC=mpif90 -{}-with-scotch-dir=/opt/scotch}\\
-
-\item Next, we need to compile the source code using:
-
-\texttt{make xgenerate\_databases}
-
-\texttt{make xspecfem3D} \\
-
-\item Before we can run noise simulations, we have to specify the noise
-statistics, e.g., the ensemble-averaged noise spectrum. Matlab scripts
-are provided to help you to generate the necessary file.
-
-\texttt{EXAMPLES/noise\_tomography/NOISE\_TOMOGRAPHY.m} (main program)\\
- \texttt{EXAMPLES/noise\_tomography/PetersonNoiseModel.m}
-
-In Matlab, simply run:
-
-\texttt{NOISE\_TOMOGRAPHY(NSTEP, DT, Tmin, Tmax, NOISE\_MODEL)}\\
-
-\texttt{DT} is given in \texttt{Par\_file}, but \texttt{NSTEP} is
-NOT the one specified in \texttt{Par\_file}. Instead, you have to
-feed $2~\mathrm{NSTEP}-1$ to account for the doubled length of cross
-correlations. \texttt{Tmin} and \texttt{Tmax} correspond to the period
-range you are interested in, whereas \texttt{NOISE\_MODEL} denotes
-the noise model you will be using. Details can be found in the Matlab
-script.
-
-After running the Matlab script, you will be given the following information
-(plus a figure in Matlab):
-
-\texttt{{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}}\\
- \texttt{the source time function has been saved in:}\\
- \texttt{/data2/yangl/3D\_NOISE/S\_squared} (note this path must be
-different)\\
- \texttt{S\_squared should be put into directory:}\\
- \texttt{OUTPUT\_FILES/NOISE\_TOMOGRAPHY/ in the SPECFEM3D Cartesian
-package}\\
-
-In other words, the Matlab script creates a file called \texttt{S\_squared},
-which is the first `new' input file we encounter for noise simulations.
-
-One may choose a flat noise spectrum rather than Peterson's noise
-model. This can be done easily by modifying the Matlab script a little.
-
-\item Create a new directory in the SPECFEM3D Cartesian package, name it
-as \texttt{OUTPUT\_FILES/NOISE\_TOMOGRAPHY/}. We will add some parameter
-files later in this folder.
-\item Put the Matlab-generated-file \texttt{S\_squared} in \texttt{OUTPUT\_FILES/NOISE\_TOMOGRAPHY/}.
-
-That's to say, you will have a file \texttt{OUTPUT\_FILES/NOISE\_TOMOGRAPHY/S\_squared}
-in the SPECFEM3D Cartesian package.
-
-\item Create a file called \texttt{OUTPUT\_FILES/NOISE\_TOMOGRAPHY/irec\_master\_noise}.
-Note that this file is located in directory \texttt{OUTPUT\_FILES/NOISE\_TOMOGRAPHY/}
-as well. In general, all noise simulation related parameter files
-go into that directory. \texttt{irec\_master\_noise} contains only
-one integer, which is the ID of the `master' receiver. For example,
-if this file contains 5, it means that the fifth receiver listed in
-\texttt{DATA/STATIONS} becomes the `master'. That's why we mentioned
-previously that the order of receivers in \texttt{DATA/STATIONS} is
-important.
-
-Note that in regional simulations, the \texttt{DATA/STATIONS} might
-contain receivers which are outside of our computational domains.
-Therefore, the integer in \texttt{irec\_master\_noise} is actually
-the ID in \texttt{DATA/STATIONS\_FILTERED} (which is generated by
-\texttt{bin/xgenerate\_databases}).
-
-\item Create a file called \texttt{OUTPUT\_FILES/NOISE\_TOMOGRAPHY/nu\_master}.
-This file holds three numbers, forming a (unit) vector. It describes
-which component we are cross-correlating at the `master' receiver,
-i.e., ${\hat{{\bf \nu}}}^{\alpha}$ in \citet{trompetal2010}. The
-three numbers correspond to E/N/Z components respectively. Most often,
-the vertical component is used, and in those cases the three numbers
-should be 0, 0 and 1.
-\item Describe the noise direction and distributions in \texttt{src/shared/noise\_tomography.f90}.
-Search for a subroutine called \texttt{noise\_distribution\_direction}
-in \texttt{noise\_tomography.f90}. It is actually located at the very
-beginning of \texttt{noise\_tomography.f90}. The default assumes vertical
-noise and a uniform distribution across the whole free surface of
-the model. It should be quite self-explanatory for modifications.
-Should you modify this part, you have to re-compile the source code.
-\end{itemize}
-
-\subsection{Simulations}
-
-With all of the above done, we can finally launch our simulations.
-Again, please make sure that the \texttt{LOCAL\_PATH} in \texttt{DATA/Par\_file}
-is not globally shared. It is quite problematic to have a globally
-shared \texttt{LOCAL\_PATH}, in terms of both disk storage and speed
-of I/O (we have to save the wavefields at the earth's surface at every
-time step).
-
-As discussed in \citet{trompetal2010}, it takes three steps/simulations
-to obtain one contribution of the ensemble sensitivity kernels:
-\begin{itemize}
-\item Step 1: simulation for generating wavefields\\
- \texttt{SIMULATION\_TYPE=1}\\
- \texttt{NOISE\_TOMOGRAPHY=1}\\
- \texttt{SAVE\_FORWARD} not used, can be either .true. or .false.
-\item Step 2: simulation for ensemble forward wavefields\\
- \texttt{SIMULATION\_TYPE=1}\\
- \texttt{NOISE\_TOMOGRAPHY=2}\\
- \texttt{SAVE\_FORWARD=.true.}
-\item Step 3: simulation for ensemble adjoint wavefields and sensitivity
-kernels\\
- \texttt{SIMULATION\_TYPE=3}\\
- \texttt{NOISE\_TOMOGRAPHY=3}\\
- \texttt{SAVE\_FORWARD=.false.}
-
-Note Step 3 is an adjoint simulation, please refer to previous chapters
-on how to prepare adjoint sources and other necessary files, as well
-as how adjoint simulations work.
-
-\end{itemize}
-Note that it is better to run the three steps continuously within
-the same job, otherwise you have to collect the saved surface movies
-from the old nodes/CPUs to the new nodes/CPUs. This process varies
-from cluster to cluster and thus cannot be discussed here. Please
-ask your cluster administrator for information/configuration of the
-cluster you are using.\\
-
-\subsection{Post-simulation}
-
-After those simulations, you have all stuff you need, either in the
-\texttt{OUTPUT\_FILES/} or in the directory specified by \texttt{LOCAL\_PATH}
-in \texttt{DATA/Par\_file} (which are most probably on local nodes).
-Collect whatever you want from the local nodes to your workstation,
-and then visualize them. This process is the same as what you may
-have done for regular earthquake simulations. Refer to other chapters
-if you have problems.\\
-
-Simply speaking, two outputs are the most interesting: the simulated
-ensemble cross correlations and one contribution of the ensemble sensitivity
-kernels.\\
-
-The simulated ensemble cross correlations are obtained after the second
-simulation (Step 2). Seismograms in \texttt{OUTPUT\_FILES/} are actually
-the simulated ensemble cross correlations. Collect them immediately
-after Step 2, or the Step 3 will overwrite them. Note that we have
-a `master' receiver specified by \texttt{irec\_master\_noise}, the
-seismogram at one station corresponds to the cross correlation between
-that station and the `master'. Since the seismograms have three components,
-we may obtain cross correlations for different components as well,
-not necessarily the cross correlations between vertical components.\\
-
-One contribution of the ensemble cross-correlation sensitivity kernels
-are obtained after Step 3, stored in the \texttt{DATA/LOCAL\_PATH}
-on local nodes. The ensemble kernel files are named the same as regular
-earthquake kernels.\\
-
-You need to run another three simulations to get the other contribution
-of the ensemble kernels, using different forward and adjoint sources
-given in \citet{trompetal2010}.
-
-\section{Example}
-
-In order to illustrate noise simulations in an easy way, one example
-is provided in \texttt{EXAMPLES/noise\_tomography/}. See \texttt{EXAMPLES/noise\_tomography/README}
-for explanations. \\
-
-Note, however, that they are created for a specific workstation (CLOVER@PRINCETON),
-which has at least 4 cores with `mpif90' working properly. \\
-
-If your workstation is suitable, you can run the example in \texttt{EXAMPLES/noise\_tomography/}
-using:\\
-
-\texttt{./pre-processing.sh}\\
-
-Even if this script does not work on your workstation, the procedure
-it describes is universal. You may review the whole process described
-in the last section by following the commands in \texttt{pre-processing.sh},
-which should contain enough explanations for all the commands.
-
-%------------------------------------------------------------------------------------------------%
-
 \chapter{Gravity integral calculations for the gravity field of an Earth model}
-
-%------------------------------------------------------------------------------------------------%
 
@@
 \texttt{grep -i GRAVITY\_INTEGRALS src/*/*}
-
-%------------------------------------------------------------------------------------------------%
-
-\chapter{Graphics}
-
-%------------------------------------------------------------------------------------------------%
-
-\section{\label{sec:Mesh-graphics}Meshes}
-
-In case you used the internal mesher \texttt{xmeshfem3D} to create
-and partition your mesh, you can output mesh files in ABAQUS (.INP)
-and DX (.dx) format to visualize them. For this, you must set either
-the flag \texttt{CREATE\_DX\_FILES} or \texttt{CREATE\_ABAQUS\_FILES}
-to \texttt{.true.} in the mesher's parameter file \texttt{Mesh\_Par\_file}
-prior to running the mesher (see Chapter~\ref{cha:Running-the-Mesher-Meshfem3D}
-for details). You can then use AVS \urlwithparentheses{www.avs.com}
-or OpenDX \urlwithparentheses{www.opendx.org} to visualize the mesh
-and MPI partition (slices).
-
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=0.49\textwidth]{figures/vtk_mesh_vp} \includegraphics[width=0.49\textwidth]{figures/vtk_mesh_vs}
-\par\end{centering}
-
-\caption{Visualization using Paraview of VTK files created by \texttt{xgenerate\_databases}
-showing P- and S-wave velocities assigned to the mesh points. The
-mesh was created by \texttt{xmeshfem3D} for 4 processors.}
-
-\label{fig:vtk.mesh}
-\end{figure}
-
-You have also the option to visualize the distributed databases produced
-by \texttt{xgenerate\_databases} using Paraview \urlwithparentheses{www.paraview.org}.
-For this, you must set the flag \texttt{SAVE\_MESH\_FILES} to \texttt{.true.}
-in the main parameter file \texttt{Par\_file} (see Chapter~\ref{cha:Main-Parameter}
-for details). This will create VTK files for each single partition.
-You can then use Paraview \urlwithparentheses{www.paraview.org}
-to visualized these partitions.
-
-\section{\label{sec:Movies}Movies}
-
-To make a surface or volume movie of the simulation, set parameters
-\texttt{MOVIE\_SURFACE}, \texttt{MOVIE\_VOLUME}, \texttt{MOVIE\_TYPE},
-and \texttt{NTSTEP\_BETWEEN\_FRAMES} in the \texttt{Par\_file}. Turning
-on the movie flags, in particular \texttt{MOVIE\_VOLUME}, produces
-large output files. \texttt{MOVIE\_VOLUME} files are saved in the
-\texttt{LOCAL\_PATH} directory, whereas \texttt{MOVIE\_SURFACE} output
-files are saved in the \texttt{OUTPUT\_FILES} directory. We save the
-displacement field if the parameter \texttt{SAVE\_DISPLACEMENT} is
-set, otherwise the velocity field is saved. The look of a movie is
-determined by the half-duration of the source. The half-duration should
-be large enough so that the movie does not contain frequencies that
-are not resolved by the mesh, i.e., it should not contain numerical
-noise. This can be accomplished by selecting a CMT \texttt{HALF\_DURATION}
-> 1.1 $\times$ smallest period (see figure \ref{fig:CMTSOLUTION-file}).
-When \texttt{MOVIE\_SURFACE} = .\texttt{true.}, the half duration
-of each source in the \texttt{CMTSOLUTION} file is replaced by
-\begin{quote}
-\[
-\sqrt{(}\mathrm{\mathtt{HALF\_DURATIO}\mathtt{N}^{2}}+\mathrm{\mathtt{HDUR\_MOVI}\mathtt{E}^{2}})
-\]
- \textbf{NOTE:} If \texttt{HDUR\_MOVIE} is set to 0.0, the code will
-select the appropriate value of 1.1 $\times$ smallest period. As
-usual, for a point source one can set \texttt{half duration} in the
-\texttt{CMTSOLUTION} file to be 0.0 and \texttt{HDUR\_MOVIE} = 0.0
-in the \texttt{Par\_file} to get the highest frequencies resolved
-by the simulation, but for a finite source one would keep all the
-\texttt{half durations} as prescribed by the finite source model and
-set \texttt{HDUR\_MOVIE} = 0.0.
-\end{quote}
-
-\subsection{Movie Surface}
-
-When running \texttt{xspecfem3D} with the \texttt{MOVIE\_SURFACE}
-flag turned on, the code outputs \texttt{moviedata??????} files in
-the \texttt{OUTPUT\_FILES} directory. There are several flags in the
-main parameter file \texttt{Par\_file} that control the output of
-these movie data files (see section \ref{cha:Main-Parameter} for
-details): \texttt{\small NTSTEP\_BETWEEN\_FRAMES}{\small{} to set the
-timesteps between frames, }\texttt{\small SAVE\_DISPLACEMENT}{\small{}
-to save displacement instead of velocity, }\texttt{\small USE\_HIGHRES\_FOR\_MOVIES}{\small{}
-to save values at all GLL point instead of element edges. In order
-to output additionally shakemaps, you would set the parameter }\texttt{\small CREATE\_SHAKEMAP}{\small{}
-to }\texttt{\small .true.}{\small .}{\small \par}
-
-The files are in a fairly complicated binary format, but there is
-a program provided to convert the output into more user friendly formats:
-\begin{description}
-\item [{\texttt{xcreate\_movie\_shakemap\_AVS\_DX\_GMT}}] From \texttt{create\_movie\_shakemap\_AVS\_DX\_GMT.f90},
-it outputs data in ASCII, OpenDX, or AVS format (also readable in
-ParaView). Before compiling the code, make sure you have the file
-\texttt{surface\_from\_mesher.h} in the \texttt{OUTPUT\_FILES/} directory.
-This file will be created by the solver run. Then type
-
-\begin{lyxcode}
-{\footnotesize make~xcreate\_movie\_shakemap\_AVS\_DX\_GMT}{\footnotesize \par}
-\end{lyxcode}
-
-and run the executable \texttt{xcreate\_movie\_shakemap\_AVS\_DX\_GMT}
-in the main directory. It will create visualization files
-in your format of choice. The code will prompt the user for input
-parameters.
-
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=0.32\textwidth]{figures/movie_surf_1} \includegraphics[width=0.32\textwidth]{figures/movie_surf_2}
-\includegraphics[width=0.32\textwidth]{figures/movie_surf_3}
-\par\end{centering}
-
-\caption{Visualization using AVS files created by \texttt{xcreate\_movie\_shakemap\_AVS\_DX\_GMT}
-showing movie snapshots of vertical velocity components at different
-times.}
-
-\label{fig:movie.surf}
-\end{figure}
-
-\end{description}
-The \texttt{SPECFEM3D Cartesian} code is running in near real-time
-to produce animations of southern California earthquakes via the web;
-see Southern California ShakeMovie\textregistered{}\urlwithparentheses{www.shakemovie.caltech.edu}.
-
-\subsection{\label{sub:Movie-Volume}Movie Volume}
-
-When running xspecfem3D with the \texttt{\small MOVIE\_VOLUME}{\small{}
-flag turned on, the code outputs several files in }\texttt{\small LOCAL\_PATH}{\small{}
-specified in the main }\texttt{\small Par\_file}{\small , e.g. in
-directory }\texttt{\small OUTPUT\_FILES/DATABASES\_MPI}{\small .
-The output is saved by each processor at the time interval specified
-by }\texttt{\small NTSTEP\_BETWEEN\_FRAMES}{\small . For all domains,
-the velocity field is output to files: }\texttt{\small proc??????\_velocity\_X\_it??????.bin}{\small ,
-}\texttt{\small proc??????\_velocity\_Y\_it??????.bin}{\small{} and
-}\texttt{\small proc??????\_velocity\_Z\_it??????.bin}{\small . For
-elastic domains, the divergence and curl taken from the velocity field,
-i.e. $\nabla\cdot{\bf {v}}$ and $\nabla\times{\bf {v}}$, get stored
-as well: }\texttt{\small proc??????\_div\_it??????.bin}{\small , }\texttt{\small proc??????\_curl\_X\_t??????.bin}{\small ,
-}\texttt{\small proc??????\_curl\_Y\_it??????.bin}{\small{} and }\texttt{\small proc??????\_curl\_Z\_it??????.bin}{\small .
-The files denoted }\texttt{\small proc??????\_div\_}{\small ~}\\
-{\small{} }\texttt{\small glob\_it??????.bin}{\small{} and }\texttt{\small proc??????\_curl\_glob\_it??????.bin}{\small{}
-are stored on the global points only, all the other arrays are stored
-on all GLL points. Note that the components X/Y/Z can change to E/N/Z
-according to the }\texttt{\small SUPPRESS\_UTM\_PROJECTION}{\small{}
-flag (see also Appendix~\ref{cha:Coordinates} and \ref{cha:channel-codes}).}{\small \par}
-
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=0.32\textwidth]{figures/movie_volume_1} \includegraphics[width=0.32\textwidth]{figures/movie_volume_2}
-\includegraphics[width=0.32\textwidth]{figures/movie_volume_3}
-\par\end{centering}
-
-\caption{Paraview visualization using movie volume files (converted by \texttt{xcombine\_vol\_data}
-and \texttt{mesh2vtu.pl}) and showing snapshots of vertical velocity
-components at different times.}
-
-\label{fig:movie.volume}
-\end{figure}
-
-To visualize these files, we use an auxilliary program \texttt{combine\_vol\_data.f90}
-to combine the data from all slices into one mesh file. To compile
-it in the root directory, type:
-\begin{lyxcode}
-{\footnotesize make~xcombine\_vol\_data~}{\footnotesize \par}
-\end{lyxcode}
-which will create the executable \texttt{bin/xcombine\_vol\_data}.
-To output the usage of this executable, type
-\begin{lyxcode}
-{\footnotesize ./bin/xcombine\_vol\_data~}{\footnotesize \par}
-\end{lyxcode}
-without arguments.
-
-\medskip
-
-\texttt{xcombine\_vol\_data} will combine the data on the different processors (located in \texttt{OUTPUT\_FILES/DATABASES\_MPI}), for a given quantity and a given iteration, to one file.
-For example, if you want to combine \texttt{velocity\_Z}, for the iteration 400 on 4 processors, i.e.
-if you want to combine these files :
-
-\medskip
-
-\texttt{proc000000\_velocity\_Z\_it000400.bin }
-
-\texttt{proc000001\_velocity\_Z\_it000400.bin }
-
-\texttt{proc000002\_velocity\_Z\_it000400.bin }
-
-\texttt{proc000003\_velocity\_Z\_it000400.bin}
-
-\medskip
-
-You have to go in the directory of the concerned example (where are the directories \texttt{DATA}, \texttt{OUTPUT\_FILES}, etc.).
-Then, you can launch \texttt{xcombine\_vol\_data}, specifying the path where it is located.
-As an example, if the \texttt{DATA} and \texttt{OUTPUT\_FILES} directories of the concerned example are in the root specfem3d directory, \texttt{xcombine\_vol\_data} is in \texttt{./bin}, and you have to type:
-
-\begin{lyxcode}
-~{\footnotesize ./bin/xcombine\_vol\_data~0~3~velocity\_Z\_it000400~./OUTPUT\_FILES/DATABASES\_MPI~./OUTPUT\_FILES~0~}{\footnotesize \par}
-\end{lyxcode}
-
-Here, $0$ is the number of the first processor, $3$ the number of the last one, \texttt{velocity\_Z\_it000400} the name of the files we want to combine without the prefix \texttt{proc000000*\_}, \texttt{./OUTPUT\_FILES/DATABASES\_MPI} the directory where the files are located, \texttt{./OUTPUT\_FILES} the directory where the combined file will be stored, and $0$ is the parameter to create a low-resolution mesh file ($1$ for a high-resolution).
-
-\medskip
-
-At the beginning of the source file \texttt{combine\_vol\_data.f90} (located in \texttt{src/auxiliaries/}), there is a logical constant parameter \texttt{USE\_VTK\_OUTPUT} whose default value is \texttt{.true.}, that which will provide the output combined file directly in the VTK format and you can use Paraview to visualize it.
-
-\medskip
-
-If the value of this parameter is \texttt{.false.}, the output mesh file will have the name \texttt{velocity\_Z\_it000400.mesh}. Then we next have to convert the
-\texttt{.mesh} file into the VTU (Unstructured grid file) format which can be viewed in ParaView. For this task, you can use and modify the
-script \texttt{mesh2vtu.pl} located in directory \texttt{utils/Visualization/Paraview/}, for example:
-\begin{lyxcode}
-{\footnotesize mesh2vtu.pl~-i~velocity\_Z\_it000400.mesh~-o~velocity\_Z\_it000400.vtu}{\footnotesize \par}
-\end{lyxcode}
-Notice that this Perl script uses a program \texttt{mesh2vtu} in the \texttt{utils/Visualization/Paraview/mesh2vtu} directory, which further
-uses the VTK \urlwithparentheses{http://www.vtk.org/} run-time library for its execution. Therefore, make sure you have them properly set
-in the script according to your system.
-
-\bigskip
-
-Then, to do a movie with several iterations, you have to repeat this process for each iteration you want to put in your movie.
-
-\section{\label{sec:Finite-Frequency-Kernels}Finite-Frequency Kernels}
-
-The finite-frequency kernels computed as explained in Section \ref{sec:Adjoint-simulation-finite}
-are saved in the \texttt{LOCAL\_PATH} at the end of the simulation.
-Therefore, we first need to collect these files on the front end,
-combine them into one mesh file, and visualize them with some auxilliary
-programs.
-\begin{enumerate}
-\item \textbf{Create slice files}
-
-We will only discuss the case of one source-receiver pair, i.e., the
-so-called banana-doughnut kernels. Although it is possible to collect
-the kernel files from all slices on the front end, it usually takes
-up too much storage space (at least tens of gigabytes). Since the
-sensitivity kernels are the strongest along the source-receiver great
-circle path, it is sufficient to collect only the slices that are
-along or close to the great circle path.
-
-A Perl script \texttt{slice\_number.pl} located in directory \texttt{utils/Visualization/Paraview/}
-can help to figure out the slice numbers that lie along the great
-circle path. It applies to meshes created with the internal mesher
-\texttt{xmeshfem3D}.
-\begin{enumerate}
-\item On machines where you have access to the script, copy the \texttt{Mesh\_Par\_file},
-and \texttt{output\_solver} files, and run:
-
-\begin{lyxcode}
-{\small slice\_number.pl~Mesh\_Par\_file~output\_solver.txt~slice\_file}{\small \par}
-\end{lyxcode}
-
-which will generate a \texttt{slices\_file}.
-
-\item For cases with multiple sources and multiple receivers, you need to
-provide a slice file before proceeding to the next step.
-\end{enumerate}
-\item \textbf{Collect the kernel files}
-
-After obtaining the slice files, you can collect the corresponding
-kernel files from the given slices.
-\begin{enumerate}
-\item You can use or modify the script \texttt{utils/copy\_basin\_database.pl}
-to accomplish this:
-
-\begin{lyxcode}
-{\small utils/copy\_database.pl~slice\_file~lsf\_machine\_file~filename~{[}jobid{]}}{\small \par}
-\end{lyxcode}
-
-where \texttt{\small lsf\_machine\_file}{\small{} is the machine file
-generated by the LSF scheduler, }\texttt{\small filename}{\small{}
-is the kernel name (e.g., }\texttt{\small rho\_kernel}{\small , }\texttt{\small alpha\_kernel}{\small{}
-and }\texttt{\small beta\_kernel}{\small ), and the optional }\texttt{\small jobid}{\small{}
-is the name of the subdirectory under }\texttt{\small LOCAL\_PATH}{\small{}
-where all the kernel files are stored.}{\small \par}
-
-\item After executing this script, all the necessary mesh topology files
-as well as the kernel array files are collected to the local directory
-of the front end.
-\end{enumerate}
-\item \textbf{Combine kernel files into one mesh file}
-
-We use an auxilliary program \texttt{combine\_vol\_data.f90} to combine
-the kernel files from all slices into one mesh file.
-\begin{enumerate}
-\item Compile it in the root directory:
-
-\begin{lyxcode}
-{\footnotesize make~xcombine\_vol\_data~}{\footnotesize \par}
-
-{\footnotesize ./bin/xcombine\_vol\_data~slice\_list~filename~input\_dir~output\_dir~high/low-resolution~}{\footnotesize \par}
-\end{lyxcode}
-
-where \texttt{input\_dir} is the directory where all the individual
-kernel files are stored, and \texttt{output\_dir} is where the mesh
-file will be written.
-
-\item Use 1 for a high-resolution mesh, outputting all the GLL points to
-the mesh file, or use 0 for low resolution, outputting only the corner
-points of the elements to the mesh file.
-\item The output mesh file will have the name \texttt{filename\_rho(alpha,beta).mesh}
-\end{enumerate}
-\item \textbf{Convert mesh files into .vtu files}
-
-\begin{enumerate}
-\item We next convert the \texttt{.mesh} file into the VTU (Unstructured
-grid file) format which can be viewed in ParaView. For this task,
-you can use and modify the script \texttt{mesh2vtu.pl} located in
-directory \texttt{utils/Visualization/Paraview/}, for example:
-
-\begin{lyxcode}
-{\footnotesize mesh2vtu.pl~-i~file.mesh~-o~file.vtu~}{\footnotesize \par}
-\end{lyxcode}
-\item Notice that this Perl script uses a program \texttt{mesh2vtu} in the
-\texttt{utils/Visualization/Paraview/mesh2vtu} directory, which further
-uses the VTK \urlwithparentheses{http://www.vtk.org/} run-time library
-for its execution. Therefore, make sure you have them properly set
-in the script according to your system.
-\end{enumerate}
-\item \textbf{Copy over the source and receiver .vtk file}
-
-In the case of a single source and a single receiver, the simulation
-also generates the file \texttt{sr.vtk} located in the \texttt{OUTPUT\_FILES/}
-directory to describe the source and receiver locations, which can
-also be viewed in Paraview in the next step.
-
-\item \textbf{View the mesh in ParaView}
-
-Finally, we can view the mesh in ParaView \urlwithparentheses{www.paraview.org}.
-\begin{enumerate}
-\item Open ParaView.
-\item From the top menu, \textsf{File} $\rightarrow$\textsf{Open data},
-select \texttt{file.vtu}, and click the \textsf{Accept} button.
-
-\begin{itemize}
-\item If the mesh file is of moderate size, it shows up on the screen; otherwise,
-only the bounding box is shown.
-\end{itemize}
-\item Click \textsf{Display Tab} $\rightarrow$ \textsf{Display Style} $\rightarrow$
-\textsf{Representation} and select \textsf{wireframe of surface} to
-display it.
-\item To create a cross-section of the volumetric mesh, choose \textsf{Filter}
-$\rightarrow$ \textsf{cut}, and under \textsf{Parameters Tab}, choose
-\textsf{Cut Function} $\rightarrow$ \textsf{plane}.
-\item Fill in center and normal information given by the \texttt{global\_slice\_number.pl}
-script (either from the standard output or from \texttt{normal\_plane.txt}
-file).
-\item To change the color scale, go to \textsf{Display Tab} $\rightarrow$
-\textsf{Color} $\rightarrow$ \textsf{Edit Color Map} and reselect
-lower and upper limits, or change the color scheme.
-\item Now load in the source and receiver location file by \textsf{File}
-$\rightarrow$ \textsf{Open data}, select \texttt{sr.vt}k, and click
-the \textsf{Accept} button. Choose \textsf{Filter} $\rightarrow$
-\textsf{Glyph}, and represent the points by `\textsf{spheres}'.
-\item For more information about ParaView, see the ParaView Users Guide
-\urlwithparentheses{www.paraview.org/files/v1.6/ParaViewUsersGuide.PDF}.
-\end{enumerate}
-\end{enumerate}
-\begin{figure}[H]
-\noindent \begin{centering}
-\includegraphics[scale=0.6]{figures/3D-S-Kernel}
-\par\end{centering}
-
-\caption{(a) Top Panel: Vertical source-receiver cross-section of the S-wave
-finite-frequency sensitivity kernel $K_{\beta}$ for station GSC at
-an epicentral distance of 176 km from the September 3, 2002, Yorba
-Linda earthquake. Lower Panel: Vertical source-receiver cross-section
-of the 3D S-wave speed model used for the spectral-element simulations
-\citep{KoLiTrSuStSh04}. (b) The same as (a) but for station HEC at
-an epicentral distance of 165 km \citep{LiTr06}.}
-
-\label{figure:P-wave-speed-finite-frequency}
-\end{figure}
-
-\chapter{\label{cha:Scheduler}Running through a Scheduler}
-
-The code is usually run on large parallel machines, often PC clusters,
-most of which use schedulers, i.e., queuing or batch management systems
-to manage the running of jobs from a large number of users. The following
-considerations need to be taken into account when running on a system
-that uses a scheduler:
-\begin{itemize}
-\item The processors/nodes to be used for each run are assigned dynamically
-by the scheduler, based on availability. Therefore, in order for the
-\texttt{xgenerate\_databases} and the \texttt{xspecfem3D} executables
-(or between successive runs of the solver) to have access to the same
-database files (if they are stored on hard drives local to the nodes
-on which the code is run), they must be launched in sequence as a
-single job.
-\item On some systems, the nodes to which running jobs are assigned are
-not configured for compilation. It may therefore be necessary to pre-compile
-both the \texttt{xgenerate\_databases} and the \texttt{xspecfem3D}
-executables.
-\item One feature of schedulers/queuing systems is that they allow submission
-of multiple jobs in a ``launch and forget'' mode. In order to take
-advantage of this property, care needs to be taken that output and
-intermediate files from separate jobs do not overwrite each other,
-or otherwise interfere with other running jobs.
-\end{itemize}
-Examples of job scripts can be found in the \texttt{\small utils/Cluster/}{\small{}
-directory and can straightforwardly be modified and adapted to meet
-more specific running needs.}{\small \par}
-
-\chapter{{\normalsize \label{cha:-Changing-the}} Changing the Model}
-
-In this section we explain how to change the velocity model used for
-your simulations. These changes involve contributing specific subroutines
-that replace existing subroutines in the \texttt{SPECFEM3D Cartesian}
-package. Note that \texttt{SPECFEM3D Cartesian} can handle Earth models
-with material properties that vary within each spectral element.
-
-\section{{\normalsize \label{sec:Using-tomographic}}Using external tomographic Earth models}
-
-To implement your own external tomographic model(s), you must provide
-your own external tomography file(s), and choose between two possible
-options:\\
- \indent (1) set in the \texttt{Par\_file} the model parameter \texttt{MODEL
-= tomo}, \\
- or for more user control: \\
- \indent (2) set in the \texttt{Par\_file} the model parameter \texttt{MODEL
-= default}, define the negative \texttt{material\_ID} identifier for
-each element in the file \texttt{MESH/materials\_file} and use the
-following format in the file \texttt{MESH/nummaterial\_velocity\_file}
-when using CUBIT to construct your mesh (see also section \ref{subsec:Exporting-the-Mesh}):
-\begin{lyxcode}
-domain\_ID~material\_ID~tomography~elastic~file\_name~1
-\end{lyxcode}
-where: \\
- \indent \texttt{domain\_ID} is 1 for acoustic or 2 for elastic materials,
-\\
- \indent \texttt{material\_ID} a negative, unique identifier (i.e.,
--1,-2,...), \\
- \indent \texttt{tomography} keyword for tomographic material definition,
-\\
- \indent \texttt{elastic} keyword for elastic material definition,
-\\
- \indent \texttt{file\_name} the name of the tomography file and
-1 a positive unique identifier.\\
-
-The external tomographic model is represented by a grid of points
-with assigned material properties and homogeneous resolution along
-each spatial direction x, y and z. The ASCII file \texttt{file\_name}
-that describe the tomography should be located in the \texttt{TOMOGRAPHY\_PATH}
-directory, set in the \texttt{Par\_file}. The format of the file,
-as read from \texttt{model\_tomography.f90} located in the \texttt{src/generate\_databases}
-directory, looks like Figure \ref{fig:tomography_file}, where
-\begin{description}
-\item [{\texttt{ORIG\_X}, \texttt{END\_X}}] are, respectively, the coordinates
-of the initial and final tomographic grid points along the x direction
-(in the mesh units, e.g., $m$);
-\item [{\texttt{ORIG\_Y}, \texttt{END\_Y}}] respectively the coordinates
-of the initial and final tomographic grid points along the y direction
-(in the mesh units, e.g., $m$);
-\item [{\texttt{ORIG\_Z}, \texttt{END\_Z}}] respectively the coordinates
-of the initial and final tomographic grid points along the z direction
-(in the mesh units, e.g., $m$);
-\item [{\texttt{SPACING\_X}, \texttt{SPACING\_Y}, \texttt{SPACING\_Z}}] the
-spacing between the tomographic grid points along the x, y and z directions,
-respectively (in the mesh units, e.g., $m$);
-\item [{\texttt{NX}, \texttt{NY}, \texttt{NZ}}] the number of grid points
-along the spatial directions x, y and z, respectively; \texttt{NX}
-is given by {[}(\texttt{END\_X} - \texttt{ORIG\_X})/\texttt{SPACING\_X}{]}+1;
-\texttt{NY} and \texttt{NZ} are the same as \texttt{NX}, but for the
-y and z directions, respectively;
-\item [{\texttt{VP\_MIN}, \texttt{VP\_MAX}, \texttt{VS\_MIN}, \texttt{VS\_MAX}, \texttt{RHO\_MIN}, \texttt{RHO\_MAX}}] the
-minimum and maximum values of the wave speed \texttt{vp} and \texttt{vs}
-(in $m\, s^{-1}$) and of the density \texttt{rho} (in $kg\, m^{-3}$);
-these values could be the actual limits of the tomographic parameters
-in the grid or the minimum and maximum values to which we force the
-cut of velocity and density in the model.
-\end{description}
-After the first four lines, in the file \texttt{file\_name} the tomographic
-grid points are listed with the corresponding values of \texttt{vp},
-\texttt{vs} and \texttt{rho}, scanning the grid along the x coordinate
-(from \texttt{ORIG\_X} to \texttt{END\_X} with step of \texttt{SPACING\_X})
-for each given y (from \texttt{ORIG\_Y} to \texttt{END\_Y}, with step
-of \texttt{SPACING\_Y}) and each given z (from \texttt{ORIG\_Z} to
-\texttt{END\_Z}, with step of \texttt{SPACING\_Z}). %fig
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=1\textwidth]{figures/tomo_file}
-\par\end{centering}
-
-\caption{Tomography file \texttt{file\_name} that describes an external Earth
-model. The coordinates x, y and z of the grid, their limits ORIG\_X,
-ORIG\_Y, ORIG\_Z, END\_X, END\_Y, END\_Z and the grid spacings SPACING\_X,
-SPACING\_Y, SPACING\_Z should be in the units of the constructed mesh
-(e.g., UTM coordinates in meters).}
-
-\label{fig:tomography_file}
-\end{figure}
-
-The user can implement his own interpolation algorithm for the tomography
-model by changing the routine \texttt{model\_tomography.f90} located
-in the \texttt{src/generate\_databases/} directory. Moreover, for
-models that involve both fully defined materials and a tomography
-description, the \texttt{nummaterial\_velocity\_file} has multiple
-lines each with the corresponding suitable format described above.
-
-%==============
-
-\subsubsection*{Example: External tomography file with variable discretization intervals}
-
-The example in Figure~\ref{fig:tomography_file} is for an external tomography file with uniform spacing in the $x$, $y$, and $z$ directions.
-(Note that $\Delta x$, $\Delta y$, and $\Delta z$ need not be the same as in the example.) In that example, the  \texttt{nummaterial\_velocity\_file} is
-%
-\begin{lyxcode}
-2  -1 tomography elastic tomography\_model.xyz 1
-\end{lyxcode}
-%
-and the file \texttt{tomography\_model.xyz} will need to reside in \texttt{DATA/}.
-All entries in the second column of \texttt{materials\_file} will be \texttt{-1}, which means that each element in the mesh will be interpolated accordind to the values in \texttt{tomography\_model.xyz}.
-
-In some cases it may be desirable to have an external tomography model that is described in more than one file.
-For example, in cases like southern California, the length scale of variation in the structure of the wave speed model is much shorter in the sedimentary basin models within the upper 15~km.
-Therefore one might want to use an external tomography file that is sampled with, say, $\Delta x = 1000$~m, $\Delta y = 1000$~m, and $\Delta z = 250$~m in the uppermost 15~km, and then use $\Delta x = 2000$~m, $\Delta y = 2000$~m, and $\Delta z = 1000$~m below a depth of 15~km.
-If these intervals are chosen appropriately, then it will result in a pair of external tomography file that is much smaller than the alternative of having a single file with the fine discretization.
-In this case  \texttt{nummaterial\_velocity\_file} is
-%
-\begin{lyxcode}
-2  -1 tomography elastic file\_above\_15km.xyz 1 \\
-2  -2 tomography elastic file\_below\_15km.xyz 1
-\end{lyxcode}
-%
-and the files \texttt{file\_above\_15km.xyz} and \texttt{file\_below\_15km.xyz} will need to reside in \texttt{DATA/}.
-All entries in the second column of \texttt{materials\_file} will need to be either \texttt{-1} or \texttt{-2}, depending on whether the element is above or below 15~km depth.
-In this sense, the construction of the mesh and the discretization of the external model will generally be done in tandem.
-
-Other more complicated discretizations can be used following the same procedure. (In particular, there is no limit on the number of different external files that are used.)
-
-%=============
-
-\section{{\normalsize \label{sec:Anelastic-Models}}External (an)elastic Models}
-
-To use your own external model, you can set in the \texttt{Par\_file}
-the model parameter \texttt{MODEL = external}. Three-dimensional acoustic
-and/or (an)elastic (attenuation) models may then be superimposed onto
-the mesh based upon your subroutine in \texttt{model\_external\_values.f90}
-located in directory \texttt{src/generate\_databases}. The call to
-this routine would be as follows
-\begin{lyxcode}
-call~model\_external\_values(xmesh,~ymesh,~zmesh,~~\&~~\\
-~~~~~~~~~~~rho,~vp,~vs,~qmu\_atten,~iflag\_aniso,~idomain\_id)
-\end{lyxcode}
-Input to this routine consists of:
-\begin{description}
-\item [{\texttt{xmesh,ymesh,zmesh}}] location of mesh point
-\end{description}
-Output to this routine consists of:
-\begin{description}
-\item [{\texttt{rho,vp,vs}}] isotropic model parameters for density $\rho$
-($kg/m^{3}$), $v_{p}$ ($m/s$) and $v_{s}$ ($m/s$)
-\item [{\texttt{qmu\_atten}}] Shear wave quality factor: $0<Q_{\mu}<9000$
-\item [{\texttt{iflag\_aniso}}] anisotropic model flag, $0$ indicating
-no anisotropy or $1$ using anisotropic model parameters as defined
-in routine file \texttt{model\_aniso.f90}
-\item [{\texttt{idomain\_id}}] domain identifier, $1$ for acoustic, $2$
-for elastic, $3$ for poroelastic materials.
-\end{description}
-Note that the resolution and maximum value of anelastic models are
-truncated. This speeds the construction of the standard linear solids
-during the meshing stage. To change the resolution, currently at one
-significant figure following the decimal, or the maximum value (9000),
-consult \texttt{shared/constants.h}.
-
-\section{{\normalsize \label{sec:Anisotropic-Models}}Anisotropic Models}
-
-To use your anisotropic models, you can either set in the \texttt{Par\_file}
-the model parameter \texttt{ANISOTROPY} to \texttt{.true.} or set
-the model parameter \texttt{MODEL = aniso}. Three-dimensional anisotropic
-models may be superimposed on the mesh based upon the subroutine in
-file \texttt{model\_aniso.f90} located in directory \texttt{src/generate\_databases}.
-The call to this subroutine is of the form
-\begin{lyxcode}
-call~model\_aniso(iflag\_aniso,~rho,~vp,~vs,~\&~~\\
-~~~~~~~~~~~c11,c12,c13,c14,c15,c16,c22,c23,c24,c25,c26,~\&~~\\
-~~~~~~~~~~~c33,c34,c35,c36,c44,c45,c46,c55,c56,c66)~
-\end{lyxcode}
-Input to this routine consists of:
-\begin{description}
-\item [{\texttt{iflag\_aniso}}] flag indicating the type of the anisotropic
-model, i.e. $0$ for no superimposed anisotropy, $1$ or $2$ for
-generic pre-defined anisotropic models.
-\item [{\texttt{rho,vp,vs}}] reference isotropic model parameters for density
-$\rho$, $v_{p}$ and $v_{s}$.
-\end{description}
-Output from the routine consists of the following non-dimensional
-model parameters:
-\begin{description}
-\item [{\texttt{c11},}] \textbf{$\cdots$,} \texttt{\textbf{c66}} 21~dimensionalized
-anisotropic elastic parameters.
-\end{description}
-You can replace the \texttt{model\_aniso.f90} file by your own version
-\textit{provided you do not change the call structure of the routine},
-i.e., the new routine should take exactly the same input and produce
-the required relative output.
-
-\section{{\normalsize \label{sec:Using-SEP}}Using external SEP models.}
-
-Stanford Exploration Project (SEP) models consists of two files per variables, one header file (\texttt{.H})
-and one binary file (\texttt{.H$@$}).
-Among other information, the header file indicates, for each dimension,
-the offset (\texttt{o1}, \texttt{o2} and \texttt{o3} fields),
-the spatial step size (\texttt{d1}, \texttt{d2} and \texttt{d3} fields)
-and the number of samples (\texttt{n1}, \texttt{n2} and \texttt{n3}) necessary to read
-the linked binary file (\texttt{in} field).
-
-SEP model reading occurs during database generation. It it the user responsibility
-to ensure that proper interfaces between acoustic and elastic domains are set up
-during the meshing phase. For more information, refer to
-section~\ref{cha:Running-the-Mesher-Meshfem3D}.
-
-In \texttt{DATA/Par\_file}, \texttt{\textbf{MODEL}} should be set to \texttt{\textbf{sep}},
-and \texttt{\textbf{SEP\_MODEL\_DIRECTORY}} should point to wherever your sep model files reside.
-
-Note that in order not to overload the parameter file, SEP header files
-should be named \texttt{vp.H}, \texttt{vs.H} and \texttt{rho.H}. The \texttt{in} field
-in these binary files can be any path as long as it is relative to
-the previously set \texttt{SEP\_MODEL\_DIRECTORY}.
-We also assume that any dimensional value inside SEP header files is given in meters and
-that binary files are single precision floating point numbers stored in little endian format.
-There is currently only support for \texttt{vp}, \texttt{vs} and \texttt{rho} variables.
-
-An example demonstrating the use SEP models is given in:
-\texttt{EXAMPLES/meshfem3D\_examples/sep\_bathymetry}.
-
-\chapter{Post-Processing Scripts}
-
-Several post-processing scripts/programs are provided in the \texttt{utils/}
-directory, and most of them need to be adjusted when used on different
-systems, for example, the path of the executable programs. Here we
-only list a few of the available scripts and provide a brief description,
-and you can either refer to the related sections for detailed usage
-or, in a lot of cases, type the script/program name without arguments
-for its usage.
-
-\section{Process Data and Synthetics\label{sec:Process-data-and-syn}}
-
-In many cases, the SEM synthetics are calculated and compared to data
-seismograms recorded at seismic stations. Since the SEM synthetics
-are accurate for a certain frequency range, both the original data
-and the synthetics need to be processed before a comparison can be
-made.
-
-For such comparisons, the following steps are recommended:
-\begin{enumerate}
-\item Make sure that both synthetic and observed seismograms have the correct
-station/event and timing information.
-\item Convolve synthetic seismograms with a source time function with the
-half duration specified in the \texttt{CMTSOLUTION} file, provided,
-as recommended, you used a zero half duration in the SEM simulations.
-\item Resample both observed and synthetic seismograms to a common sampling
-rate.
-\item Cut the records using the same window.
-\item Remove the trend and mean from the records and taper them.
-\item Remove the instrument response from the observed seismograms (recommended)
-or convolve the synthetic seismograms with the instrument response.
-\item Make sure that you apply the same filters to both observed and synthetic
-seismograms. Preferably, avoid filtering your records more than once.
-\item Now, you are ready to compare your synthetic and observed seismograms.
-\end{enumerate}
-\noindent We generally use the following scripts provided in the \texttt{utils/seis\_process/}
-directory:
-
-\subsection{Data processing script \texttt{process\_data.pl}}
-
-This script cuts a given portion of the original data, filters it,
-transfers the data into a displacement record, and picks the first
-P and S arrivals. For more functionality, type `\texttt{process\_data.pl}'
-without any argument. An example of the usage of the script:
-\begin{lyxcode}
-{\footnotesize process\_data.pl~-m~CMTSOLUTION~-s~1.0~-l~0/4000~-i~-f~-t~40/500~-p~-x~bp~DATA/1999.330{*}.BH?.SAC}~
-\end{lyxcode}
-\noindent which has resampled the SAC files to a sampling rate of
-1 seconds, cut them between 0 and 4000 seconds, transfered them into
-displacement records and filtered them between 40 and 500 seconds,
-picked the first P and S arrivals, and added suffix `\texttt{bp}'
-to the file names.
-
-Note that all of the scripts in this section actually use the SAC
-and/or IASP91 to do the core operations; therefore make sure that
-the SAC and IASP91 packages are installed properly on your system,
-and that all the environment variables are set properly before running
-these scripts.
-
-\subsection{Synthetics processing script \texttt{process\_syn.pl}}
-
-This script converts the synthetic output from the SEM code from ASCII
-to SAC format, and performs similar operations as `\texttt{process\_data.pl}'.
-An example of the usage of the script:
-\begin{lyxcode}
-{\footnotesize process\_syn.pl~-m~CMTSOLUTION~-h~-a~STATIONS~-s~1.0~-l~0/4000~-f~-t~40/500~-p~-x~bp~SEM/{*}.BX?.semd}~
-\end{lyxcode}
-which will convolve the synthetics with a triangular source-time function
-from the \texttt{CMTSOLUTION} file, convert the synthetics into SAC
-format, add event and station information into the SAC headers, resample
-the SAC files with a sampling rate of 1 seconds, cut them between
-0 and 4000 seconds, filter them between 40 and 500 seconds with the
-same filter used for the observed data, pick the first P and S arrivals,
-and add the suffix `\texttt{bp}' to the file names.
-
-More options are available for this script, such as adding time shift
-to the origin time of the synthetics, convolving the synthetics with
-a triangular source time function with a given half duration, etc.
-Type \texttt{process\_syn.pl} without any argument for a detailed
-usage.
-
-In order to convert between SAC format and ASCII files, useful scripts
-are provided in the subdirectories ~\\
- \texttt{utils/sac2000\_alpha\_convert/} and \texttt{utils/seis\_process/asc2sac/}.
-
-\subsection{Script \texttt{rotate.pl}}
-
-The original data and synthetics have three components: vertical (BHZ
-resp. BXZ), north (BHN resp. BXN) and east (BHE resp. BXE). However,
-for most seismology applications, transverse and radial components
-are also desirable. Therefore, we need to rotate the horizontal components
-of both the data and the synthetics to the transverse and radial direction,
-and \texttt{\small rotate.pl}{\small{} can be used to accomplish this:}{\small \par}
-\begin{lyxcode}
-rotate.pl~-l~0~-L~180~-d~DATA/{*}.BHE.SAC.bp~
-
-rotate.pl~-l~0~-L~180~SEM/{*}.BXE.semd.sac.bp~
-\end{lyxcode}
-where the first command performs rotation on the SAC data obtained
-through Seismogram Transfer Program (STP) \urlwithparentheses{http://www.data.scec.org/STP/stp.html},
-while the second command rotates the processed SEM synthetics.
-
-\section{Collect Synthetic Seismograms}
-
-The forward and adjoint simulations generate synthetic seismograms
-in the \texttt{OUTPUT\_FILES/} directory by default. For the forward
-simulation, the files are named like \texttt{NT.STA.BX?.semd} for
-two-column time series, or \texttt{NT.STA.BX?.semd.sac} for ASCII
-SAC format, where NT and STA are the network code and station name,
-and \texttt{BX?} stands for the component name. Please see the Appendix~\ref{cha:Coordinates}
-and \ref{cha:channel-codes} for further details.
-
-The adjont simulations generate synthetic seismograms with the name
-\texttt{NT.S?????.S??.sem} (refer to Section \ref{sec:Adjoint-simulation-sources}
-for details). The kernel simulations output the back-reconstructed
-synthetic seismogram in the name \texttt{NT.STA.BX?.semd}, mainly
-for the purpose of checking the accuracy of the reconstruction. Refer
-to Section \ref{sec:Adjoint-simulation-finite} for further details.
-
-You do have further options to change this default output behavior,
-given in the main constants file \texttt{constants.h} located in \texttt{src/shared/}
-directory:
-\begin{description}
-\item [{\texttt{SEISMOGRAMS\_BINARY}}] set to \texttt{.true.} to have seismograms
-written out in binary format.
-\item [{\texttt{WRITE\_SEISMOGRAMS\_BY\_MASTER}}] Set to \texttt{.true.}
-to have only the master process writing out seismograms. This can
-be useful on a cluster, where only the master process node has access
-to the output directory.
-\item [{\texttt{USE\_OUTPUT\_FILES\_PATH}}] Set to \texttt{.false.} to
-have the seismograms output to \texttt{LOCAL\_PATH} directory specified
-in the main parameter file \texttt{DATA/Par\_file}. In this case,
-you could collect the synthetics onto the frontend using the \texttt{collect\_seismo\_lsf\_multi.pl}
-script located in the \texttt{utils/Cluster/lsf/} directory. The usage
-of the script would be e.g.:
-
-\begin{lyxcode}
-collect\_seismo.pl~machines~DATA/Par\_file
-\end{lyxcode}
-\end{description}
-where \texttt{machines} is a file containing the node names and \texttt{DATA/Par\_file}
-the parameter file used to extract the \texttt{LOCAL\_PATH} directory
-used for the simulation.
-
-\section{Clean Local Database}
-
-After all the simulations are done, the seismograms are collected,
-and the useful database files are copied to the frontend, you may
-need to clean the local scratch disk for the next simulation. This
-is especially important in the case of kernel simulation, where very
-large files are generated for the absorbing boundaries to help with
-the reconstruction of the regular forward wavefield. A sample script
-is provided in \texttt{utils/}:
-\begin{lyxcode}
-cleanbase.pl~machines
-\end{lyxcode}
-where \texttt{machines} is a file containing the node names.
-
-\section{Plot Movie Snapshots and Synthetic Shakemaps}
-
-\subsection{Script \texttt{movie2gif.gmt.pl}}
-
-With the movie data saved in \texttt{OUTPUT\_FILES/} at the end of
-a movie simulation (\texttt{\small MOVIE\_SURFACE=.true.}{\small ),
-you can run the }\texttt{\small `create\_movie\_shakemap\_AVS\_DX\_GMT}{\small '
-code to convert these binary movie data into GMT xyz files for futher
-processing. A sample script }\texttt{\small movie2gif.gmt.pl}{\small{}
-is provided to do this conversion, and then plot the movie snapshots
-in GMT, for example:}{\small \par}
-\begin{lyxcode}
-movie2gif.gmt.pl~-m~CMTSOLUTION~-g~-f~1/40~-n~-2~-p~
-\end{lyxcode}
-which for the first through the 40th movie frame, converts the \texttt{moviedata}
-files into GMT xyz files, interpolates them using the 'nearneighbor'
-command in GMT, and plots them on a 2D topography map. Note that `\texttt{-2}'
-and `\texttt{-p}' are both optional.
-
-\subsection{Script \texttt{plot\_shakemap.gmt.pl}}
-
-With the shakemap data saved in \texttt{OUTPUT\_FILES/} at the end
-of a shakemap simulation ~\\
- (\texttt{CREATE\_SHAKEMAP=.true.}), you can also run \texttt{`create\_movie\_shakemap\_AVS\_DX\_GMT}'
-code to convert the binary shakemap data into GMT xyz files. A sample
-script \texttt{plot\_shakemap.gmt.pl} is provided to do this conversion,
-and then plot the shakemaps in GMT, for example:
-\begin{lyxcode}
-plot\_shakemap.gmt.pl~data\_dir~type(1,2,3)~CMTSOLUTION~
-\end{lyxcode}
-where \texttt{type=1} for a displacement shakemap, \texttt{2} for
-velocity, and \texttt{3} for acceleration.
-
-\section{Map Local Database}
-
-A sample program \texttt{remap\_database} is provided to map the local
-database from a set of machines to another set of machines. This is
-especially useful when you want to run mesher and solver, or different
-types of solvers separately through a scheduler (refer to Chapter~\ref{cha:Scheduler}).
-\begin{lyxcode}
-run\_lsf.bash~-{}-gm-no-shmem~-{}-gm-copy-env~remap\_database~old\_machines~150
-\end{lyxcode}
-where \texttt{old\_machines} is the LSF machine file used in the previous
-simulation, and \texttt{150} is the number of processors in total.
-
-%============================================================
-
-\chapter{\label{cha:developers}Information for developers of the code, and for people who want to learn how the technique works}
-
-You can get a very simple 1D version of a demo code (there is one in Fortran and one in Python):\\
-
-git clone --recursive https://github.com/geodynamics/specfem1d.git\\
-
-\noindent We also have simple 3D demo source codes that implement the SEM in a single, small program, in directory\\
-utils/small\_SEM\_solvers\_in\_Fortran\_and\_C\_without\_MPI\_to\_learn of the specfem3d package.
-They are useful to learn how the spectral-element method works, and how to write or modify a code to implement it.
-Also useful to test new ideas by modifying these simple codes to run some tests.\\
-
-\noindent For information on how to contribute to the code, i.e., for how to make your modifications, additions or improvements part of the
-official package, see https://github.com/geodynamics/specfem3d/wiki .
-
-%============================================================
-
-\chapter*{\label{cha:Bug-Reports-and}Bug Reports and Suggestions for Improvements}
-
-To report bugs or suggest improvements to the code, please send an
-e-mail to the CIG Computational Seismology Mailing List \urlwithparentheses{cig-seismo@geodynamics.org}.
-
-\addcontentsline{toc}{chapter}{Notes \& Acknowledgments}
-
-\chapter*{Notes \& Acknowledgments}
-
-In order to keep the software package thread-safe in case a multithreaded
-implementation of MPI is used, developers should not add modules or
-common blocks to the source code but rather use regular subroutine
-arguments (which can be grouped in ``derived types'' if needed for
-clarity).
-
-The Gauss-Lobatto-Legendre subroutines in \texttt{gll\_library.f90}
-are based in part on software libraries from the Massachusetts Institute
-of Technology, Department of Mechanical Engineering (Cambridge, Massachusetts,
-USA). The non-structured global numbering software was provided by
-Paul F. Fischer (Brown University, Providence, Rhode Island, USA,
-now at Argonne National Laboratory, USA).
-
-OpenDX \urlwithparentheses{http://www.opendx.org} is open-source
-based on IBM Data Explorer, AVS \urlwithparentheses{http://www.avs.com}
-is a trademark of Advanced Visualization Systems, and ParaView \urlwithparentheses{http://www.paraview.org}
-is an open-source visualization platform.
-
-Please e-mail your feedback, questions, comments, and suggestions
-to the CIG Computational Seismology Mailing List \urlwithparentheses{cig-seismo@geodynamics.org}.
-
-\addcontentsline{toc}{chapter}{Copyright}
-
-\chapter*{Copyright}
-
-Main historical authors: Dimitri Komatitsch and Jeroen Tromp
-
-Princeton University, USA, and CNRS / University of Marseille, France
-
-$\copyright$ Princeton University and CNRS / University of Marseille, July 2012
-
-This program is free software; you can redistribute it and/or modify
-it under the terms of the GNU General Public License as published
-by the Free Software Foundation (see Appendix \ref{cha:License}).\\
-
-\textbf{\underline{Evolution of the code:}}\\
-
-MPI v. 3.0, December 2014: many developers.
-Convolutional PML, LDDRK time scheme, bulk attenuation support, simultaneous MPI runs,
-ADIOS file I/O support, coupling with external codes, new seismogram names,
-Deville routines for additional GLL degrees, tomography tools, unit/regression test framework,
-improved CUDA GPUs performance, additonal GEOCUBIT support, better make compilation,
-git versioning system. \\
-
-MPI v. 2.1, July 2012: Max Rietmann, Peter Messmer, Daniel Peter, Dimitri
-Komatitsch, Joseph Charles, Zhinan Xie: support for CUDA GPUs, better
-CFL stability for the Stacey absorbing conditions. \\
-
-MPI v. 2.0, November 2010: Dimitri Komatitsch, Nicolas Le Goff, Roland
-Martin and Pieyre Le Loher, University of Pau, France, Daniel Peter,
-Jeroen Tromp and the Princeton group of developers, Princeton University,
-USA, and Emanuele Casarotti, INGV Roma, Italy: support for CUBIT meshes
-decomposed by SCOTCH; much faster solver using Michel Deville's inlined
-matrix products.\\
-
-MPI v. 1.4 Dimitri Komatitsch, University of Pau, Qinya Liu and others,
-Caltech, September 2006: better adjoint and kernel calculations, faster
-and better I/Os on very large systems, many small improvements and
-bug fixes.\\
-
-MPI v. 1.3 Dimitri Komatitsch, University of Pau, and Qinya Liu, Caltech,
-July 2005: serial version, regular mesh, adjoint and kernel calculations,
-ParaView support.\\
-
-MPI v. 1.2 Min Chen and Dimitri Komatitsch, Caltech, July 2004: full
-anisotropy, volume movie.\\
-
-MPI v. 1.1 Dimitri Komatitsch, Caltech, October 2002: Zhu's Moho map,
-scaling of $V_{s}$ with depth, Hauksson's regional model, attenuation,
-oceans, movies.\\
-
-MPI v. 1.0 Dimitri Komatitsch, Caltech, USA, May 2002: first MPI version
-based on global code.\\
-
-Dimitri Komatitsch, IPG Paris, France, December 1996: first 3-D solver
-for the CM-5 Connection Machine, parallelized on 128 processors using
-Connection Machine Fortran.\\
-
-\addcontentsline{toc}{chapter}{References} \bibliography{bibliography}
-
-\appendix
-%dummy comment inserted by tex2lyx to ensure that this paragraph is not empty
-
-\chapter{\label{cha:Coordinates}Reference Frame Convention}
-
-The code uses the following convention for the Cartesian reference
-frame:
-\begin{itemize}
-\item the $x$ axis points East
-\item the $y$ axis points North
-\item the $z$ axis points up
-\end{itemize}
-Note that this convention is different from both the \citet{AkRi80}
-convention and the Harvard Centroid-Moment Tensor (CMT) convention.
-The Aki \& Richards convention is
-\begin{itemize}
-\item the $x$ axis points North
-\item the $y$ axis points East
-\item the $z$ axis points down
-\end{itemize}
-and the Harvard CMT convention is
-\begin{itemize}
-\item the $x$ axis points South
-\item the $y$ axis points East
-\item the $z$ axis points up
-\end{itemize}
-
-\subsection*{Source and receiver locations}
-
-The SPECFEM3D Cartesian software code internally uses Cartesian coordinates.
-The given locations for sources and receiver locations thus may get
-converted. Sources and receiver locations are read in from the \texttt{CMTSOLUTION}
-(or \texttt{FORCESOLUTION}) and \texttt{STATIONS} files. Note that
-e.g. the \texttt{CMTSOLUTION} and \texttt{FORCESOLUTION} files denotes
-the location by \textquotedbl{}longitude/latitude/depth\textquotedbl{}.
-We read in longitude as $x$ coordinate, latitude as $y$ coordinate.
-
-In case the flag \texttt{SUPPRESS\_UTM\_PROJECTION} is set to \texttt{.false.}
-in the main parameter file (see Chapter~\ref{cha:Creating-Distributed-Databases}),
-the $x$/$y$ coordinates have to be given in degrees and are converted
-to Cartesian coordinates using a UTM conversion with the specified
-UTM zone.
-
-The value for depth (given in $km$ in \texttt{CMTSOLUTION} and \texttt{FORCESOLUTION})
-or burial depth (given in $m$ in \texttt{STATIONS}) is evaluated
-with respect to the surface of the mesh at the specified $x$/$y$
-location to find the corresponding $z$ coordinate. It is possible
-to use this depth value directly as $z$ coordinate by changing the
-flag \texttt{USE\_SOURCES\_RECVS\_Z} to \texttt{.true.} in the file
-\texttt{constants.h} located in the \texttt{src/shared/} subdirectory.
-
-\subsection*{Seismogram outputs}
-
-The seismogram output directions are given in Cartesian $x$/$y$/$z$
-directions and not rotated any further. Changing flags in \texttt{constants.h}
-in the \texttt{src/shared/} subdirectory only rotates the seismogram
-outputs if receivers are forced to be located at the surface (\texttt{RECVS\_CAN\_BE\_BURIED\_EXT\_MESH}
-set to \texttt{.false.}) and the normal to the surface at the receiver
-location should be used (\texttt{EXT\_MESH\_RECV\_NORMAL} set to \texttt{.true.})
-as vertical. In this case, the outputs are rotated to have the vertical
-component normal to the surface of the mesh, $x$ and $y$ directions
-are somewhat arbitrary orthogonal directions along the surface.
-
-For the labeling of the seismogram channels, see Appendix~\ref{cha:channel-codes}.
-Additionally, we add labels to the synthetics using the following
-convention: For a receiver station located in an
-\begin{description}
-\item [{elastic domain:}] ~\\
-
-\begin{itemize}
-\item \texttt{semd} for the displacement vector
-\item \texttt{semv} for the velocity vector
-\item \texttt{sema} for the acceleration vector
-\end{itemize}
-\item [{acoustic domain:}] ~\\
- (please note that receiver stations in acoustic domains must be buried.
-This is due to the free surface condition which enforces a zero pressure
-at the surface.)
-
-\begin{itemize}
-\item \texttt{semd} for the displacement vector
-\item \texttt{semv} for the velocity vector
-\item \texttt{sema} for pressure which will be stored in the vertical component
-\texttt{Z}. Note that pressure in the acoustic media is isotropic
-and thus the pressure record would be the same in the other two component
-directions. We therefore use the other two seismogram components to
-store the acoustic potential in component \texttt{X} (or \texttt{N})
-and the first time derivative of the acoustic potential in component
-\texttt{Y} (or \texttt{E}).
-\end{itemize}
-\end{description}
-The seismograms are by default written out in ASCII-format to the
-\texttt{OUTPUT\_FILES/} subdirectory by each parallel process. You
-can change this behavior by changing the following flags in the \texttt{constants.h}
-file located in the \texttt{src/shared/} subdirectory:
-\begin{description}
-\item [{\texttt{SEISMOGRAMS\_BINARY}}] set to \texttt{.true.} to have seismograms
-written out in binary format.
-\item [{\texttt{WRITE\_SEISMOGRAMS\_BY\_MASTER}}] Set to \texttt{.true.}
-to have only the master process writing out seismograms. This can
-be useful on a cluster, where only the master process node has access
-to the output directory.
-\item [{\texttt{USE\_OUTPUT\_FILES\_PATH}}] Set to \texttt{.false.} to
-have the seismograms output to \texttt{LOCAL\_PATH} directory specified
-in the main parameter file \texttt{DATA/Par\_file}.
-\end{description}
-
-\chapter{\label{cha:channel-codes}Channel Codes of Seismograms}
-
-Seismic networks, such as the Global Seismographic Network (GSN),
-generally involve various types of instruments with different bandwidths,
-sampling properties and component configurations. There are standards
-to name channel codes depending on instrument properties. IRIS \texttt{(www.iris.edu)}
-uses SEED format for channel codes, which are represented by three
-letters, such as \texttt{LHN}, \texttt{BHZ}, etc. In older versions
-of the SPECFEM3D Cartesian package, a common format was used for the
-channel codes of all seismograms, which was \texttt{BHE/BHN/BHZ} for
-three components. To avoid confusion when comparison are made to observed
-data, we are now using the FDSN convention (\url{http://www.fdsn.org/})
-for SEM seismograms. In the following, we give a brief explanation
-of the FDSN convention used by IRIS, and how it is adopted in SEM
-seismograms. Please visit \texttt{www.iris.edu/manuals/SEED\_appA.htm}
-for further information.\\
-
-\noindent \texttt{\textbf{Band code:}} The first letter in the channel
-code denotes the band code of seismograms, which depends on the response
-band and the sampling rate of instruments. The list of band codes
-used by IRIS is shown in Figure \ref{fig:IRIS_band_codes}. The sampling
-rate of SEM synthetics is controlled by the resolution of simulations
-rather than instrument properties. However, for consistency, we follow
-the FDSN convention for SEM seismograms governed by their sampling
-rate. For SEM synthetics, we consider band codes for which $dt\leq1$
-s. IRIS also considers the response band of instruments. For instance,
-short-period and broad-band seismograms with the same sampling rate
-correspond to different band codes, such as S and B, respectively.
-In such cases, we consider SEM seismograms as broad band, ignoring
-the corner period ($\geq10$ s) of the response band of instruments
-(note that at these resolutions, the minimum period in the SEM synthetics
-will be less than $10$ s). Accordingly, when you run a simulation
-the band code will be chosen depending on the resolution of the synthetics,
-and channel codes of SEM seismograms will start with either \texttt{L},
-\texttt{M}, \texttt{B}, \texttt{H}, \texttt{C} or \texttt{F}, shown
-by red color in the figure.\\
-
-\begin{figure}[ht]
-\noindent \begin{centering}
-\includegraphics[scale=0.6]{figures/IRIS_band_codes}
-\par\end{centering}
-
-\caption{The band code convention is based on the sampling rate and the response
-band of instruments. Please visit \texttt{www.iris.edu/manuals/SEED\_appA.htm}
-for further information. Grey rows show the relative band-code range
-in SPECFEM3D Cartesian, and the band codes used to name SEM seismograms
-are denoted in red.}
-
-\label{fig:IRIS_band_codes}
-\end{figure}
-
-\noindent \texttt{\textbf{Instrument code:}} The second letter in
-the channel code corresponds to instrument codes, which specify the
-family to which the sensor belongs. For instance, \texttt{H} and \texttt{L}
-are used for high-gain and low-gain seismometers, respectively. The
-instrument code of SEM seismograms will always be \texttt{X}, as assigned
-by FDSN for synthetic seismograms. \\
-
-\noindent \texttt{\textbf{Orientation code:}} The third letter in
-channel codes is an orientation code, which generally describes the
-physical configuration of the components of instrument packages. SPECFEM3D
-Cartesian uses the traditional orientation code \texttt{E/N/Z} (East-West,
-North-South, Vertical) for three components when a UTM projection
-is used. If the UTM conversion is suppressed, i.e. the flag \texttt{SUPPRESS\_UTM\_PROJECTION}
-is set to \texttt{.true.}, then the three components are labelled
-\texttt{X/Y/Z} according to the Cartesian reference frame. \\
-
-\noindent \texttt{\textbf{EXAMPLE:}} The sampling rate is given by
-\texttt{DT} in the main parameter file \texttt{Par\_file} located
-in the \texttt{DATA/} subdirectory. Depending on the resolution of
-your simulations, if you choose a sampling rate greater than $0.01$
-s and less than $1$ s, a seismogram recording displacements on the
-vertical component of a station \texttt{ASBS} (network \texttt{AZ})
-will be named \texttt{AZ.ASBS.MXZ.semd.sac}, whereas it will be \texttt{AZ.ASBS.BXZ.semd.sac},
-if the sampling rate is greater than 0.0125 and less equal to 0.1
-s.
-
-\chapter{\label{cha:Troubleshooting}Troubleshooting}
-
-\section*{FAQ}
-\begin{description}
-\item [{configuration fails:}]~
-
-Examine the log file 'config.log'. It contains detailed informations.
-In many cases, the paths to these specific compiler commands F90,
-CC and MPIF90 will not be correct if `./configure` fails.
-
-Please make sure that you have a working installation of a Fortran
-compiler, a C compiler and an MPI implementation. You should be able
-to compile this little program code:
-\begin{lyxcode}
-program~main~~~\\
-~include~'mpif.h'~~~\\
-~integer,~parameter~::~CUSTOM\_MPI\_TYPE~=~MPI\_REAL~~~\\
-~integer~ier~~~\\
-~call~MPI\_INIT(ier)~~~\\
-~call~MPI\_BARRIER(MPI\_COMM\_WORLD,ier)~~~\\
-~call~MPI\_FINALIZE(ier)~~~\\
-~end
-\end{lyxcode}
-\item [{compilation fails stating:}] ~\\
-
-\begin{lyxcode}
-...~~~\\
-~obj/program\_generate\_databases.o:~In~function~`MAIN\_\_':~~~\\
-~program\_generate\_databases.f90:(.text+0x14):~undefined~reference~to~`\_gfortran\_set\_std'~~~\\
-~...~~~\\
-
-\end{lyxcode}
-
-Make sure you're pointing to the right 'mpif90' wrapper command.
-
-Normally, this message will appear when you are mixing two different
-Fortran compilers. That is, using e.g. gfortran to compile non-MPI
-files and mpif90, wrapper provided for e.g. ifort, to compile MPI-files.
-
-fix: e.g. specify > ./configure FC=gfortran MPIF90=/usr/local/openmpi-gfortran/bin/mpif90
-
-\item [{after executing \texttt{xmeshfem3D} I've got elements with skewness of 81\% percent, what does this mean:}] Look
-at the skewness table printed in the \texttt{output\_mesher.txt} file
-after executing \texttt{xmeshfem3D} for the example given in \texttt{EXAMPLES/meshfem3D\_examples/simple\_model/}:
-
-\begin{lyxcode}
-...~~\\
-~histogram~of~skewness~(0.~good~-~1.~bad):~~~\\
-~0.0000000E+00~-~5.0000001E-02~27648~81.81818~\%~~~\\
-~5.0000001E-02~-~0.1000000~0~0.0000000E+00~\%~~~\\
-~...~~~\\
-
-\end{lyxcode}
-
-The first line means that you have 27,648 elements with a skewness
-value between 0 and 0.05 (which means the element is basically not
-skewed, just plain regular hexahedral element). The total number of
-elements you have in this mesh is (see in the \texttt{output\_mesher.txt}
-file a bit further down):
-\begin{lyxcode}
-...~~\\
-~total~number~of~elements~in~entire~mesh:~33792~~~\\
-~...~~~\\
-
-\end{lyxcode}
-
-which gives you that: 27,648 / 33,792 $\sim$ 81.8 \% of all elements
-are not skewed, i.e. regular elements. a fantastic value :)
-
-The histogram lists for this mesh also some stronger skewed elements,
-for example the worst ones belong to:
-\begin{lyxcode}
-...~~\\
-~0.6000000~-~0.6500000~2048~6.060606~\%~~~\\
-~...~~~\\
-
-\end{lyxcode}
-
-about 6 \% of all elements have distortions with a skewness value
-between 0.6 and 0.65. The skewness values give you a hint of how good
-your mesh is. In an ideal world, you would want to have no distortions,
-just like the 81\% from above. Those elements give you the best approximate
-values by the GLL quadrature used in the spectral-element method.
-However, having weakly distorted elements is still fine and the solutions
-are still accurate enough. So empirically, values up to around 0.7
-are tolerable, above that you should consider remeshing...
-
-To give you an idea why some of the elements are distorted, see the
-following figure \ref{fig:mesh.vp} of the mesh you obtain in the
-example \texttt{EXAMPLES/meshfem3D\_examples/simple\_model/}.
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=0.9\textwidth]{figures/mesh_vp}
-\par\end{centering}
-
-\caption{Paraview visualization using the mesh vtk-files for the example given
-in \texttt{EXAMPLES/meshfem3D\_examples/simple\_model/}.}
-
-\label{fig:mesh.vp}
-\end{figure}
-
-You will see that the mesh contains a doubling layer, where we stitch
-elements together such that the size of two elements will transition
-to the size of one element (very useful to keep the ratio of wavespeed
-/ element\_size about constant). Those elements in this doubling layer
-have higher skewness values and make up those 6 \% in the histogram.
-
-\item [{the code gives following error message "need at least one receiver":}] This
-means that no stations given in the input file \texttt{DATA/STATIONS}
-could be located within the dimensions of the mesh. This can happen
-for example when the mesh was created with the in-house mesher \texttt{xmeshfem3D}
-while using the Universal Transverse Mercator (UTM) projection but
-the simulation with \texttt{xspecfem3D} was suppressing this projection
-from latitude/longitude to x/y/z coordinates.
-
-In such cases, try to change your \texttt{DATA/Par\_file} and set
-e.g.:
-\begin{lyxcode}
-SUPPRESS\_UTM\_PROJECTION~=~.false.~~~\\
-
-\end{lyxcode}
-
-to be the same in \texttt{Mesh\_Par\_file} and \texttt{Par\_file}.
-This flag should be identical when using the in-house mesher \texttt{xmeshfem3D},
-\texttt{xgenerate\_databases} and \texttt{xspecfem3D} together to
-run simulations.
-
-The flag determines if the coordinates you specify for your source
-and station locations are given as lat/lon degrees and must be converted
-to UTM coordinates. As an example, if you use \texttt{.false.} within
-\texttt{Mesh\_Par\_file} then you create a mesh with \texttt{xmeshfem3D}
-using the UTM projection from lat/lon as input format to UTM projected
-coordinates to store the mesh point positions, which is fine. The
-error then may occur if in the \texttt{Par\_file} you have this set
-to \texttt{.true.} so that the \texttt{xgenerate\_databases} and \texttt{xspecfem3D}
-suppress the UTM projection and assume that all coordinates you use
-now for source and receiver locations are given in meters (that is,
-converted) already. So it won't find the specified locations in the
-used mesh. As a solutions, just change the flag in \texttt{Par\_file}
-to be the same as in \texttt{Mesh\_Par\_file} and rerun \texttt{xgenerate\_databases}
-and \texttt{xspecfem3D} to make sure that your simulation works fine.
-
-\item [{I get the following error message "forward simulation became unstable and blew up":}] In
-most cases this means that your time step size \texttt{DT} is chosen
-too big. Look at your files \texttt{output\_mesher.txt} or \texttt{output\_solver.txt}
-created in the folder \texttt{OUTPUT\_FILES}. In these output files,
-find the section:
-
-\begin{lyxcode}
-...~~~\\
-~{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}~~~\\
-~{*}{*}{*}~Verification~of~simulation~parameters~{*}{*}{*}~~~\\
-~{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}~~~\\
-~...~~~\\
-~{*}{*}{*}~Minimum~period~resolved~=~4.308774~~~\\
-~{*}{*}{*}~Maximum~suggested~time~step~=~6.8863556E-02~~~\\
-~...~~~\\
-~...~~~\\
-
-\end{lyxcode}
-
-then change \texttt{DT} in the \texttt{DATA/Par\_file} to be somewhere
-close to the maximum suggested time step. In the example above: \\
- \texttt{DT = 0.05d0} \\
- would (most probably) work fine. It could be also bigger than the
-0.068 s suggested. This depends a bit on the distortions of your mesh
-elements. The more regular they are, the bigger you can choose \texttt{DT}.
-Just play with this value a bit and see when the simulation becomes
-stable ...
-
-\end{description}
-
-\chapter{\label{cha:License}License }
-
-\textbf{GNU GENERAL PUBLIC LICENSE Version 2, June 1991. Copyright
-(C) 1989, 1991 Free Software Foundation, Inc. 59 Temple Place, Suite
-330, Boston, MA 02111-1307 USA} \\
- Everyone is permitted to copy and distribute verbatim copies of this
-license document, but changing it is not allowed.
-
-\section*{Preamble}
-
-The licenses for most software are designed to take away your freedom
-to share and change it. By contrast, the GNU General Public License
-is intended to guarantee your freedom to share and change free software
--- to make sure the software is free for all its users. This General
-Public License applies to most of the Free Software Foundation's software
-and to any other program whose authors commit to using it. (Some other
-Free Software Foundation software is covered by the GNU Library General
-Public License instead.) You can apply it to your programs, too.
-
-When we speak of free software, we are referring to freedom, not price.
-Our General Public Licenses are designed to make sure that you have
-the freedom to distribute copies of free software (and charge for
-this service if you wish), that you receive source code or can get
-it if you want it, that you can change the software or use pieces
-of it in new free programs; and that you know you can do these things.
-
-To protect your rights, we need to make restrictions that forbid anyone
-to deny you these rights or to ask you to surrender the rights. These
-restrictions translate to certain responsibilities for you if you
-distribute copies of the software, or if you modify it.
-
-For example, if you distribute copies of such a program, whether gratis
-or for a fee, you must give the recipients all the rights that you
-have. You must make sure that they, too, receive or can get the source
-code. And you must show them these terms so they know their rights.
-
-We protect your rights with two steps:
-\begin{enumerate}
-\item Copyright the software, and
-\item Offer you this license which gives you legal permission to copy, distribute
-and/or modify the software.
-\end{enumerate}
-Also, for each author's protection and ours, we want to make certain
-that everyone understands that there is no warranty for this free
-software. If the software is modified by someone else and passed on,
-we want its recipients to know that what they have is not the original,
-so that any problems introduced by others will not reflect on the
-original authors' reputations.
-
-Finally, any free program is threatened constantly by software patents.
-We wish to avoid the danger that redistributors of a free program
-will individually obtain patent licenses, in effect making the program
-proprietary. To prevent this, we have made it clear that any patent
-must be licensed for everyone's free use or not licensed at all.
-
-The precise terms and conditions for copying, distribution and modification
-follow.
-
-\section*{GNU GENERAL PUBLIC LICENSE TERMS AND CONDITIONS FOR COPYING, DISTRIBUTION
-AND MODIFICATION }
-\begin{itemize}
-\item 0.This License applies to any program or other work which contains
-a notice placed by the copyright holder saying it may be distributed
-under the terms of this General Public License. The ``Program''
-below refers to any such program or work, and a ``work based on the
-Program'' means either the Program or any derivative work under copyright
-law: that is to say, a work containing the Program or a portion of
-it, either verbatim or with modifications and/or translated into another
-language. (Hereinafter, translation is included without limitation
-in the term ``modification.'') Each licensee is addressed as ``you.''\\
- \\
- Activities other than copying, distribution and modification are
-not covered by this License; they are outside its scope. The act of
-running the Program is not restricted, and the output from the Program
-is covered only if its contents constitute a work based on the Program
-(independent of having been made by running the Program). Whether
-that is true depends on what the Program does.\end{itemize}
-\begin{enumerate}
-\item You may copy and distribute verbatim copies of the Program's source
-code as you receive it, in any medium, provided that you conspicuously
-and appropriately publish on each copy an appropriate copyright notice
-and disclaimer of warranty; keep intact all the notices that refer
-to this License and to the absence of any warranty; and give any other
-recipients of the Program a copy of this License along with the Program.
-
-You may charge a fee for the physical act of transferring a copy,
-and you may at your option offer warranty protection in exchange for
-a fee.
-
-\item You may modify your copy or copies of the Program or any portion of
-it, thus forming a work based on the Program, and copy and distribute
-such modifications or work under the terms of Section 1 above, provided
-that you also meet all of these conditions:
-
-\begin{enumerate}
-\item You must cause the modified files to carry prominent notices stating
-that you changed the files and the date of any change.
-\item You must cause any work that you distribute or publish, that in whole
-or in part contains or is derived from the Program or any part thereof,
-to be licensed as a whole at no charge to all third parties under
-the terms of this License.
-\item If the modified program normally reads commands interactively when
-run, you must cause it, when started running for such interactive
-use in the most ordinary way, to print or display an announcement
-including an appropriate copyright notice and a notice that there
-is no warranty (or else, saying that you provide a warranty) and that
-users may redistribute the program under these conditions, and telling
-the user how to view a copy of this License. (Exception: if the Program
-itself is interactive but does not normally print such an announcement,
-your work based on the Program is not required to print an announcement.)
-\end{enumerate}
-
-These requirements apply to the modified work as a whole. If identifiable
-sections of that work are not derived from the Program, and can be
-reasonably considered independent and separate works in themselves,
-then this License, and its terms, do not apply to those sections when
-you distribute them as separate works. But when you distribute the
-same sections as part of a whole which is a work based on the Program,
-the distribution of the whole must be on the terms of this License,
-whose permissions for other licensees extend to the entire whole,
-and thus to each and every part regardless of who wrote it.
-
-Thus, it is not the intent of this section to claim rights or contest
-your rights to work written entirely by you; rather, the intent is
-to exercise the right to control the distribution of derivative or
-collective works based on the Program.
-
-In addition, mere aggregation of another work not based on the Program
-with the Program (or with a work based on the Program) on a volume
-of a storage or distribution medium does not bring the other work
-under the scope of this License.
-
-\item You may copy and distribute the Program (or a work based on it, under
-Section 2) in object code or executable form under the terms of Sections
-1 and 2 above provided that you also do one of the following:
-
-\begin{enumerate}
-\item Accompany it with the complete corresponding machine-readable source
-code, which must be distributed under the terms of Sections 1 and
-2 above on a medium customarily used for software interchange; or,
-\item Accompany it with a written offer, valid for at least three years,
-to give any third party, for a charge no more than your cost of physically
-performing source distribution, a complete machine-readable copy of
-the corresponding source code, to be distributed under the terms of
-Sections 1 and 2 above on a medium customarily used for software interchange;
-or,
-\item Accompany it with the information you received as to the offer to
-distribute corresponding source code. (This alternative is allowed
-only for noncommercial distribution and only if you received the program
-in object code or executable form with such an offer, in accord with
-Subsection b above.)
-\end{enumerate}
-
-The source code for a work means the preferred form of the work for
-making modifications to it. For an executable work, complete source
-code means all the source code for all modules it contains, plus any
-associated interface definition files, plus the scripts used to control
-compilation and installation of the executable. However, as a special
-exception, the source code distributed need not include anything that
-is normally distributed (in either source or binary form) with the
-major components (compiler, kernel, and so on) of the operating system
-on which the executable runs, unless that component itself accompanies
-the executable.
-
-If distribution of executable or object code is made by offering access
-to copy from a designated place, then offering equivalent access to
-copy the source code from the same place counts as distribution of
-the source code, even though third parties are not compelled to copy
-the source along with the object code.
-
-\item You may not copy, modify, sublicense, or distribute the Program except
-as expressly provided under this License. Any attempt otherwise to
-copy, modify, sublicense or distribute the Program is void, and will
-automatically terminate your rights under this License. However, parties
-who have received copies, or rights, from you under this License will
-not have their licenses terminated so long as such parties remain
-in full compliance.
-\item You are not required to accept this License, since you have not signed
-it. However, nothing else grants you permission to modify or distribute
-the Program or its derivative works. These actions are prohibited
-by law if you do not accept this License. Therefore, by modifying
-or distributing the Program (or any work based on the Program), you
-indicate your acceptance of this License to do so, and all its terms
-and conditions for copying, distributing or modifying the Program
-or works based on it.
-\item Each time you redistribute the Program (or any work based on the Program),
-the recipient automatically receives a license from the original licensor
-to copy, distribute or modify the Program subject to these terms and
-conditions. You may not impose any further restrictions on the recipients'
-exercise of the rights granted herein. You are not responsible for
-enforcing compliance by third parties to this License.
-\item If, as a consequence of a court judgment or allegation of patent infringement
-or for any other reason (not limited to patent issues), conditions
-are imposed on you (whether by court order, agreement or otherwise)
-that contradict the conditions of this License, they do not excuse
-you from the conditions of this License. If you cannot distribute
-so as to satisfy simultaneously your obligations under this License
-and any other pertinent obligations, then as a consequence you may
-not distribute the Program at all. For example, if a patent license
-would not permit royalty-free redistribution of the Program by all
-those who receive copies directly or indirectly through you, then
-the only way you could satisfy both it and this License would be to
-refrain entirely from distribution of the Program.
-
-If any portion of this section is held invalid or unenforceable under
-any particular circumstance, the balance of the section is intended
-to apply and the section as a whole is intended to apply in other
-circumstances.
-
-It is not the purpose of this section to induce you to infringe any
-patents or other property right claims or to contest validity of any
-such claims; this section has the sole purpose of protecting the integrity
-of the free software distribution system, which is implemented by
-public license practices. Many people have made generous contributions
-to the wide range of software distributed through that system in reliance
-on consistent application of that system; it is up to the author/donor
-to decide if he or she is willing to distribute software through any
-other system and a licensee cannot impose that choice.
-
-This section is intended to make thoroughly clear what is believed
-to be a consequence of the rest of this License.
-
-\item If the distribution and/or use of the Program is restricted in certain
-countries either by patents or by copyrighted interfaces, the original
-copyright holder who places the Program under this License may add
-an explicit geographical distribution limitation excluding those countries,
-so that distribution is permitted only in or among countries not thus
-excluded. In such case, this License incorporates the limitation as
-if written in the body of this License.
-\item The Free Software Foundation may publish revised and/or new versions
-of the General Public License from time to time. Such new versions
-will be similar in spirit to the present version, but may differ in
-detail to address new problems or concerns.
-
-Each version is given a distinguishing version number. If the Program
-specifies a version number of this License which applies to it and
-``any later version,'' you have the option of following the terms
-and conditions either of that version or of any later version published
-by the Free Software Foundation. If the Program does not specify a
-version number of this License, you may choose any version ever published
-by the Free Software Foundation.
-
-\item If you wish to incorporate parts of the Program into other free programs
-whose distribution conditions are different, write to the author to
-ask for permission. For software which is copyrighted by the Free
-Software Foundation, write to the Free Software Foundation; we sometimes
-make exceptions for this. Our decision will be guided by the two goals
-of preserving the free status of all derivatives of our free software
-and of promoting the sharing and reuse of software generally.
-\end{enumerate}
-
-\subsection*{NO WARRANTY }
-\begin{itemize}
-\item 11.BECAUSE THE PROGRAM IS LICENSED FREE OF CHARGE, THERE IS NO WARRANTY
-FOR THE PROGRAM, TO THE EXTENT PERMITTED BY APPLICABLE LAW. EXCEPT
-WHEN OTHERWISE STATED IN WRITING THE COPYRIGHT HOLDERS AND/OR OTHER
-PARTIES PROVIDE THE PROGRAM ``AS IS'' WITHOUT WARRANTY OF ANY KIND,
-EITHER EXPRESSED OR IMPLIED, INCLUDING, BUT NOT LIMITED TO, THE IMPLIED
-WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE.
-THE ENTIRE RISK AS TO THE QUALITY AND PERFORMANCE OF THE PROGRAM IS
-WITH YOU. SHOULD THE PROGRAM PROVE DEFECTIVE, YOU ASSUME THE COST
-OF ALL NECESSARY SERVICING, REPAIR OR CORRECTION.
-\item 12.IN NO EVENT UNLESS REQUIRED BY APPLICABLE LAW OR AGREED TO IN WRITING
-WILL ANY COPYRIGHT HOLDER, OR ANY OTHER PARTY WHO MAY MODIFY AND/OR
-REDISTRIBUTE THE PROGRAM AS PERMITTED ABOVE, BE LIABLE TO YOU FOR
-DAMAGES, INCLUDING ANY GENERAL, SPECIAL, INCIDENTAL OR CONSEQUENTIAL
-DAMAGES ARISING OUT OF THE USE OR INABILITY TO USE THE PROGRAM (INCLUDING
-BUT NOT LIMITED TO LOSS OF DATA OR DATA BEING RENDERED INACCURATE
-OR LOSSES SUSTAINED BY YOU OR THIRD PARTIES OR A FAILURE OF THE PROGRAM
-TO OPERATE WITH ANY OTHER PROGRAMS), EVEN IF SUCH HOLDER OR OTHER
-PARTY HAS BEEN ADVISED OF THE POSSIBILITY OF SUCH DAMAGES.
-\end{itemize}
-
-\section*{END OF TERMS AND CONDITIONS }
-
-\subsection*{How to Apply These Terms to Your New Programs}
-
-If you develop a new program, and you want it to be of the greatest
-possible use to the public, the best way to achieve this is to make
-it free software which everyone can redistribute and change under
-these terms.
-
-To do so, attach the following notices to the program. It is safest
-to attach them to the start of each source file to most effectively
-convey the exclusion of warranty; and each file should have at least
-the ``copyright'' line and a pointer to where the full notice is
-found. For example:
-\begin{quote}
-One line to give the program's name and a brief idea of what it does.
-Copyright {\footnotesize $\copyright$ (}year) (name of author)
-
-This program is free software; you can redistribute it and/or modify
-it under the terms of the GNU General Public License as published
-by the Free Software Foundation; either version 2 of the License,
-or (at your option) any later version.
-
-This program is distributed in the hope that it will be useful, but
-WITHOUT ANY WARRANTY; without even the implied warranty of MERCHANTABILITY
-or FITNESS FOR A PARTICULAR PURPOSE. See the GNU General Public License
-for more details.
-
-You should have received a copy of the GNU General Public License
-along with this program; if not, write to the Free Software Foundation,
-Inc., 59 Temple Place, Suite 330, Boston, MA 02111-1307 USA
-\end{quote}
-Also add information on how to contact you by electronic and paper
-mail.
-
-If the program is interactive, make it output a short notice like
-this when it starts in an interactive mode:
-\begin{quote}
-Gnomovision version 69, Copyright $\copyright$ year name of author
-Gnomovision comes with ABSOLUTELY NO WARRANTY; for details type `show
-w'. This is free software, and you are welcome to redistribute it
-under certain conditions; type `show c' for details.
-\end{quote}
-The hypothetical commands `show w' and `show c' should show the appropriate
-parts of the General Public License. Of course, the commands you use
-may be called something other than `show w' and `show c'; they could
-even be mouse-clicks or menu items -- whatever suits your program.
-
-You should also get your employer (if you work as a programmer) or
-your school, if any, to sign a ``copyright disclaimer'' for the
-program, if necessary. Here is a sample; alter the names:
-\begin{quote}
-Yoyodyne, Inc., hereby disclaims all copyright interest in the program
-`Gnomovision' (which makes passes at compilers) written by James Hacker.
-
-(signature of Ty Coon)\\
- 1 April 1989 \\
- Ty Coon, President of Vice
-\end{quote}
-This General Public License does not permit incorporating your program
-into proprietary programs. If your program is a subroutine library,
-you may consider it more useful to permit linking proprietary applications
-with the library. If this is what you want to do, use the GNU Library
-General Public License instead of this License.
-\end{document}

### 2015-04-24 doc/USER_MANUAL/12_running_scheduler.tex
updates user manual to use facilitate the generation of a markdown version to the github wiki

--- 2015-04-18 doc/USER_MANUAL/manual_SPECFEM3D_Cartesian.tex
@@
-% !TEX encoding = IsoLatin
-%% LyX 2.0.2 created this file.  For more info, see http://www.lyx.org/.
-%% Do not edit unless you really know what you are doing.
-\documentclass[oneside,english]{book}
-\usepackage[T1]{fontenc}
-%% DK DK we do not accept accented letters, use the old LaTeX syntax
-%% DK DK for them instead: \'e \`e  \"e \^e and so on.
-%% DK DK \usepackage[latin1]{inputenc}
-\usepackage[letterpaper]{geometry}
-\geometry{verbose,tmargin=1in,bmargin=1in,lmargin=1in,rmargin=1in}
-\setcounter{secnumdepth}{3}
-\setcounter{tocdepth}{3}
-\usepackage{babel}
-\usepackage{float}
-\usepackage{textcomp}
-\usepackage{url}
-\usepackage{amsmath}
-\usepackage{amssymb}
-\usepackage{graphicx}
-\usepackage[authoryear]{natbib}
-\usepackage[unicode=true,pdfusetitle,
- bookmarks=true,bookmarksnumbered=true,bookmarksopen=false,
- breaklinks=false,pdfborder={0 0 0},backref=section,colorlinks=false,pdfpagemode=None]
- {hyperref}
-\hypersetup{
- pdftex,pdfstartview=FitH,pdfpagelayout=SinglePage}
-\usepackage{breakurl}
-
-\makeatletter
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% LyX specific LaTeX commands.
-\newcommand{\noun}[1]{\textsc{#1}}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Textclass specific LaTeX commands.
-\newenvironment{lyxcode}
-{\par\begin{list}{}{
-\setlength{\rightmargin}{\leftmargin}
-\setlength{\listparindent}{0pt}% needed for AMS classes
-\raggedright
-\setlength{\itemsep}{0pt}
-\setlength{\parsep}{0pt}
-\normalfont\ttfamily}%
- \item[]}
-{\end{list}}
-\newenvironment{lyxlist}[1]
-{\begin{list}{}
-{\settowidth{\labelwidth}{#1}
- \setlength{\leftmargin}{\labelwidth}
- \addtolength{\leftmargin}{\labelsep}
- \renewcommand{\makelabel}[1]{##1\hfil}}}
-{\end{list}}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% User specified LaTeX commands.
-%\renewcommand{\baselinestretch}{1.5}
-
-% figures
-\usepackage[dvips]{epsfig}
-
-% fonts
-\usepackage{times}
-
-% hyperlinks to sections and references
-
-\newcommand{\urlwithparentheses}[1]{(\url{#1})}
-
-% biblio GJI
-\bibliographystyle{abbrvnat}
-
-\newcommand{\nexxi}{\mbox{\texttt{NEX\_XI\/}}}
-\newcommand{\nexeta}{\mbox{\texttt{NEX\_ETA\/}}}
-\newcommand{\nprocxi}{\mbox{\texttt{NPROC\_XI\/}}}
-\newcommand{\nproceta}{\mbox{\texttt{NPROC\_ETA\/}}}
-\newcommand{\nchunks}{\mbox{\texttt{NCHUNKS\/}}}
-\newcommand{\eg}{e.g.,}
-
-% colors for the text
-\@ifundefined{definecolor}
- {\usepackage{color}}{}
-\newcommand{\red}[1]{\textcolor{Red}{#1}}
-
-\usepackage{babel}
-
-\makeatother
-
-\begin{document}
-\begin{center}
-\thispagestyle{empty}\vspace*{-1.8truecm} %
-\makebox[1\textwidth]{%
-\includegraphics[width=0.83\paperwidth]{figures/specfem_3d_Cartesian-cover} %
-}
-\par\end{center}
-
-\title{\textbf{SPECFEM3D Cartesian}\\
- \textbf{User Manual}}
-
-\author{$\copyright$ Princeton University (USA), CNRS and University of Marseille (France),\\
-and ETH Z\"urich (Switzerland)\\
-Version 3.0}
-
-\date{\today}
-
-\maketitle
-
-\section*{Authors}
-
-\noindent The SPECFEM3D Cartesian package was first developed by Dimitri
-Komatitsch and Jean-Pierre Vilotte at Institut de Physique du Globe
-(IPGP) in Paris, France from 1995 to 1997 and then by Dimitri Komatitsch
-and Jeroen Tromp at Harvard University and Caltech, USA, starting in 1998.
-The story started on April 4, 1995, when Prof. Yvon Maday from CNRS and University of Paris, France, gave a lecture to
-Dimitri Komatitsch and Jean-Pierre Vilotte at IPG about the nice properties of the Legendre spectral-element method with diagonal mass matrix that he had used for
-other equations. We are deeply indebted and thankful to him for that.
-That followed a visit by Dimitri Komatitsch to OGS (Istituto Nazionale di Oceanografia e di Geofisica Sperimentale) in Trieste, Italy, in February 1995
-to meet with G\'eza Seriani and Enrico Priolo, who introduced him to their 2D Chebyshev version of the spectral-element method with a non-diagonal mass matrix.
-We are deeply indebted and thankful to them for that.
-
-%%%%%%%% please do NOT merge the lines of the paragraph below, it makes it difficult to insert new names in alphabetical order; leave a single name per line
-%%%%%%%% please do NOT merge the lines of the paragraph below, it makes it difficult to insert new names in alphabetical order; leave a single name per line
-%%%%%%%% please do NOT merge the lines of the paragraph below, it makes it difficult to insert new names in alphabetical order; leave a single name per line
-Since then it has been developed and maintained by a development team:
-in alphabetical order,
-Jean-Paul (Pablo) Ampuero,
-Kangchen Bai,
-Piero Basini,
-C\'eline Blitz,
-Alexis Bottero,
-Ebru Bozda\u{g},
-Emanuele Casarotti,
-Joseph Charles,
-Min Chen,
-Paul Cristini,
-Percy Galvez,
-Dominik G\"oddeke,
-Vala Hj\"orleifsd\'ottir,
-Sue Kientz,
-Dimitri Komatitsch,
-Jes\'us Labarta,
-Nicolas Le Goff,
-Pieyre Le Loher,
-Matthieu Lefebvre,
-Qinya Liu,
-Yang Luo,
-Alessia Maggi,
-Federica Magnoni,
-Roland Martin,
-Ren\'e Matzen,
-Dennis McRitchie,
-Matthias Meschede,
-Peter Messmer,
-David Mich\'ea,
-Surendra Nadh Somala,
-Tarje Nissen-Meyer,
-Daniel Peter,
-Max Rietmann,
-Elliott Sales de Andrade,
-Brian Savage,
-Bernhard Schuberth,
-Anne Sieminski,
-Leif Strand,
-Carl Tape,
-Jeroen Tromp,
-Jean-Pierre Vilotte,
-Zhinan Xie,
-Chang-Hua Zhang,
-Hejun Zhu.\\
-
-The cover graphic of the manual was created by Santiago Lombeyda from
-Caltech's Center for Advanced Computing Research (CACR) \urlwithparentheses{http://www.cacr.caltech.edu}.\\
-
-The code is released open-source under the GNU version 2+ license, see the license at the end of this manual.\\
-
-\newpage{}
-\section*{Current and past main participants or main sponsors of the SPECFEM project (in no particular order)}
-%
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=0.125\textwidth]{figures/logo_cnrs}\vspace*{2truemm}
-\includegraphics[width=0.125\textwidth]{figures/logo_princeton}\vspace*{2truemm}
-\includegraphics[width=0.15\textwidth]{figures/logo_aix_marseille_universite}\vspace*{0.02truemm}
-\includegraphics[width=0.15\textwidth]{figures/logo_ETH}\vspace*{2truemm}
-\includegraphics[width=0.12\textwidth]{figures/logo_CSC_China}\vspace*{0.02truemm}
-\includegraphics[width=0.15\textwidth]{figures/logo_inria}\vspace*{2truemm}
-\includegraphics[width=0.14\textwidth]{figures/logo_UPPA}
-\par\end{centering}
-%
-\vspace*{-2truemm}
-%
-\noindent \begin{centering}
-\includegraphics[width=0.15\textwidth]{figures/logo_NVIDIA}\vspace*{2truemm}
-\includegraphics[width=0.125\textwidth]{figures/logo_IUF}\vspace*{2truemm}
-\includegraphics[width=0.135\textwidth]{figures/logo_Caltech}\vspace*{2truemm}
-\includegraphics[width=0.135\textwidth]{figures/logo_Harvard}\vspace*{2truemm}
-\includegraphics[width=0.135\textwidth]{figures/logo_IPGP}\vspace*{2truemm}
-\includegraphics[width=0.112\textwidth]{figures/logo_ANR}\vspace*{2truemm}
-\includegraphics[width=0.112\textwidth]{figures/logo_NSF}\vspace*{2truemm}
-\par\end{centering}
-%
-\vspace*{-2truemm}
-%
-\noindent \begin{centering}
-\includegraphics[width=0.190\textwidth]{figures/logo_European_Union}\vspace*{2truemm}
-\includegraphics[width=0.112\textwidth]{figures/logo_GENCI}\vspace*{2truemm}
-\includegraphics[width=0.112\textwidth]{figures/logo_PRACE}\vspace*{2truemm}
-\includegraphics[width=0.112\textwidth]{figures/logo_CINES}\vspace*{2truemm}
-\includegraphics[width=0.130\textwidth]{figures/logo_Oak_Ridge}\vspace*{2truemm}
-\hspace*{3mm}\includegraphics[width=0.130\textwidth]{figures/logo_fondation_Del_Duca}
-\includegraphics[width=0.130\textwidth]{figures/logo_CIG}\vspace*{2truemm}
-\par\end{centering}
-%
-\vspace*{-2truemm}
-%
-\noindent \begin{centering}
-\includegraphics[width=0.112\textwidth]{figures/logo_University_of_Toronto}\vspace*{2truemm}
-\includegraphics[width=0.109\textwidth]{figures/logo_INGV}\vspace*{2truemm}
-\includegraphics[width=0.102\textwidth]{figures/logo_Univ_Toulouse}\vspace*{2truemm}
-\includegraphics[width=0.112\textwidth]{figures/logo_TOTAL}\vspace*{2truemm}
-\includegraphics[width=0.130\textwidth]{figures/logo_Fairbanks}\vspace*{2truemm}
-\includegraphics[width=0.120\textwidth]{figures/logo_CINECA}\vspace*{2truemm}
-\includegraphics[width=0.140\textwidth]{figures/logo_Intel_Exascale_Labs}\vspace*{2truemm}
-\includegraphics[width=0.130\textwidth]{figures/logo_Maison_Simulation}
-\par\end{centering}
-\end{figure}
-
-\newpage{}
-
-\tableofcontents{}
-
-\chapter{Introduction}
-
-The software package SPECFEM3D Cartesian simulates seismic wave propagation
-at the local or regional scale and performs full waveform imaging (FWI) or adjoint tomography based upon the spectral-element method
-(SEM). The SEM is a continuous Galerkin technique \citep{TrKoLi08,PeKoLuMaLeCaLeMaLiBlNiBaTr11},
-which can easily be made discontinuous \citep{BeMaPa94,Ch00,KoWoHu02,ChCaVi03,LaWaBe05,Kop06,WiStBuGh10,AcKo11};
-it is then close to a particular case of the discontinuous Galerkin
-technique \citep{ReHi73,LeRa74,Arn82,JoPi86,BoMaHe91,FaRi99,HuHuRa99,CoKaSh00,GiHeWa02,RiWh03,MoRi05,GrScSc06,AiMoMu06,BeLaPi06,DuKa06,DeSeWh08,PuAmKa09,WiStBuGh10,DeSe10,EtChViGl10},
-with optimized efficiency because of its tensorized basis functions
-\citep{WiStBuGh10,AcKo11}. In particular, it can accurately handle
-very distorted mesh elements \citep{OlSe11}.\\
-
-It has very good accuracy and convergence properties \citep{MaPa89,SePr94,DeFiMu02,Coh02,DeSe07,SeOl08,AiWa09,AiWa10,MeStTh12}.
-The spectral element approach admits spectral rates of convergence
-and allows exploiting $hp$-convergence schemes. It is also very well
-suited to parallel implementation on very large supercomputers \citep{KoTsChTr03,TsKoChTr03,KoLaMi08a,CaKoLaTiMiLeSnTr08,KoViCh10}
-as well as on clusters of GPU accelerating graphics cards \citep{Kom11,MiKo10,KoMiEr09,KoErGoMi10}.
-Tensor products inside each element can be optimized to reach very
-high efficiency \citep{DeFiMu02}, and mesh point and element numbering
-can be optimized to reduce processor cache misses and improve cache
-reuse \citep{KoLaMi08a}. The SEM can also handle triangular (in 2D)
-or tetrahedral (in 3D) elements \citep{WinBoyd96,TaWi00,KoMaTrTaWi01,Coh02,MeViSa06}
-as well as mixed meshes, although with increased cost and reduced
-accuracy in these elements, as in the discontinuous Galerkin method.\\
-
-Note that in many geological models in the context of seismic wave
-propagation studies (except for instance for fault dynamic rupture
-studies, in which very high frequencies or supershear rupture need
-to be modeled near the fault, see e.g. \citet{BeGlCrViPi07,BeGlCrVi09,PuAmKa09,TaCrEtViBeSa10})
-a continuous formulation is sufficient because material property contrasts
-are not drastic and thus conforming mesh doubling bricks can efficiently
-handle mesh size variations \citep{KoTr02a,KoLiTrSuStSh04,LeChLiKoHuTr08,LeChKoHuTr09,LeKoHuTr09}.\\
-
-For a detailed introduction to the SEM as applied to regional seismic
-wave propagation, please consult \citet{PeKoLuMaLeCaLeMaLiBlNiBaTr11,TrKoLi08,KoVi98,KoTr99,ChKoViCaVaFe07}
-and in particular \citet{LeKoHuTr09,LeChKoHuTr09,LeChLiKoHuTr08,GoAmTaCaSmSaMaKo09,WiKoScTr04,KoLiTrSuStSh04}.
-A detailed theoretical analysis of the dispersion
-and stability properties of the SEM is available in \citet{Coh02}, \citet{DeSe07}, \citet{SeOl07}, \citet{SeOl08} and \citet{MeStTh12}.\\
-
-Effects due to lateral variations in compressional-wave speed, shear-wave
-speed, density, a 3D crustal model, topography and bathymetry are
-included. The package can accommodate full 21-parameter anisotropy
-(see~\citet{ChTr07}) as well as lateral variations in attenuation
-\citep{SaKoTr10}. Adjoint capabilities and finite-frequency kernel
-simulations are included \citep{TrKoLi08,PeKoLuMaLeCaLeMaLiBlNiBaTr11,LiTr06,FiIgBuKe09,ViOp09}.\\
-
-The SEM was originally developed in computational fluid dynamics \citep{Pat84,MaPa89}
-and has been successfully adapted to address problems in seismic wave
-propagation. Early seismic wave propagation applications of the SEM,
-utilizing Legendre basis functions and a perfectly diagonal mass matrix,
-include \citet{CoJoTo93}, \citet{Kom97}, \citet{FaMaPaQu97}, \citet{CaGa97},
-\citet{KoVi98} and \citet{KoTr99}, whereas applications involving
-Chebyshev basis functions and a nondiagonal mass matrix include \citet{SePr94},
-\citet{PrCaSe94} and \citet{SePrPr95}.\\
-
-All SPECFEM3D software is written in Fortran2003 with full portability
-in mind, and conforms strictly to the Fortran2003 standard. It uses
-no obsolete or obsolescent features of Fortran. The package uses parallel
-programming based upon the Message Passing Interface (MPI) \citep{GrLuSk94,Pac97}.\\
-
-SPECFEM3D won the Gordon Bell award for best performance at the SuperComputing~2003
-conference in Phoenix, Arizona (USA) (see \citet{KoTsChTr03} and
-\url{www.sc-conference.org/sc2003/nrfinalaward.html}). It was a finalist
-again in 2008 for a run at 0.16 petaflops (sustained) on 149,784 processors
-of the `Jaguar' Cray XT5 system at Oak Ridge National Laboratories
-(USA) \citep{CaKoLaTiMiLeSnTr08}. It also won the BULL Joseph Fourier
-supercomputing award in 2010.\\
-
-It reached the sustained one petaflop performance level for the first time in February 2013
-on the Blue Waters Cray supercomputer at the National Center for Supercomputing Applications (NCSA), located at the University of Illinois at Urbana-Champaign (USA).\\
-
-This new release of the code includes Convolution or
-Auxiliary Differential Equation Perfectly Matched absorbing Layers
-(C-PML or ADE-PML) \citep{MaKoEz08,MaKoGe08,MaKo09,MaKoGeBr10,KoMa07}.
-It also includes support for GPU graphics card acceleration \citep{Kom11,MiKo10,KoMiEr09,KoErGoMi10}.\\
-
-The next release of the code will use the PT-SCOTCH parallel and threaded
-version of SCOTCH for mesh partitioning instead of the serial version.\\
-
-SPECFEM3D Cartesian includes coupled fluid-solid domains and adjoint
-capabilities, which enables one to address seismological inverse problems,
-but for linear rheologies only so far. To accommodate visco-plastic
-or non-linear rheologies, the reader can refer to the \href{http://geoelse.stru.polimi.it/}{GeoELSE}software
-package \citep{CaGa97,StPaIg09}.
-
-\section{Citation}
-
-If you use SPECFEM3D Cartesian for your own research, please cite
-at least one of the following articles:
-\begin{description}
-\item [{Numerical simulations in general}] ~\\
- Forward and adjoint simulations are described in detail in \citet{TrKoLi08,PeKoLuMaLeCaLeMaLiBlNiBaTr11,VaCaSaKoVi99,KoMiEr09,KoErGoMi10,ChKoViCaVaFe07,MaKoDi09,KoViCh10,CaKoLaTiMiLeSnTr08,TrKoHjLiZhPeBoMcFrTrHu10,KoRiTr02,KoTr02a,KoTr02b,KoTr99}
-or \citet{KoVi98}. Additional aspects of adjoint simulations are
-described in \citet{TrTaLi05,LiTr06,TrKoLi08,LiTr08,TrKoHjLiZhPeBoMcFrTrHu10,PeKoLuMaLeCaLeMaLiBlNiBaTr11}.
-Domain decomposition is explained in detail in \citet{MaKoBlLe08},
-and excellent scaling up to 150,000 processor cores is shown for instance
-in \citet{CaKoLaTiMiLeSnTr08,KoLaMi08a,MaKoBlLe08,KoErGoMi10,Kom11},
-\item [{GPU computing}] ~\\
- Computing on GPU graphics cards for acoustic or seismic wave propagation
-applications is described in detail in \citet{Kom11,MiKo10,KoMiEr09,KoErGoMi10}.
-\end{description}
-\noindent If you use this new version, which has non blocking
-MPI for much better performance for medium or large runs, please cite
-at least one of these six articles, in which results of non blocking
-MPI runs are presented: \citet{PeKoLuMaLeCaLeMaLiBlNiBaTr11,KoErGoMi10,KoViCh10,Kom11,CaKoLaTiMiLeSnTr08,MaKoBlLe08}.\\
-
-\noindent If you work on geophysical applications, you may be interested
-in citing some of these application articles as well, among others:
-\begin{description}
-\item [{Southern California simulations}] ~\\
- \citet{KoLiTrSuStSh04,KrJiKoTr06a,KrJiKoTr06b}.
-
-If you use the 3D southern California model, please cite \citet{SuSh03}
-(Los Angeles model), \citet{lovelyetal06} (Salton Trough), and \citet{hauksson2000}
-(southern California). The Moho map was determined by \citet{zhu&kanamori2000}.
-The 1D SoCal model was developed by \citet{DrHe90}.
-
-\item [{Anisotropy}] ~\\
- \citet{ChTr07,JiTsKoTr05,ChFaKo04,FaChKo04,RiRiKoTrHe02,TrKo00}.
-\item [{Attenuation}] ~\\
- \citet{SaKoTr10,KoTr02a,KoTr99}.
-\item [{Topography}] ~\\
- \citet{LeKoHuTr09,LeChKoHuTr09,LeChLiKoHuTr08,GoAmTaCaSmSaMaKo09,WiKoScTr04}.
-\end{description}
-The corresponding Bib\TeX{} entries may be found in file \texttt{doc/USER\_MANUAL/bibliography.bib}.
-
-\section{Support}
-
-This material is based upon work supported by the USA National Science
-Foundation under Grants No. EAR-0406751 and EAR-0711177, by the French
-CNRS, French INRIA Sud-Ouest MAGIQUE-3D, French ANR NUMASIS under
-Grant No. ANR-05-CIGC-002, and European FP6 Marie Curie International
-Reintegration Grant No. MIRG-CT-2005-017461. Any opinions, findings,
-and conclusions or recommendations expressed in this material are
-those of the authors and do not necessarily reflect the views of the
-USA National Science Foundation, CNRS, INRIA, ANR or the European
-Marie Curie program.
-
-\chapter{\label{cha:Getting-Started}Getting Started}
-
-To download the SPECFEM3D\_Cartesian software package, type this:
-\begin{lyxcode}
-git clone --recursive --branch devel https://github.com/geodynamics/specfem3d.git
-\end{lyxcode}
-{\small \par}
-
-We recommend that you add {\texttt{ulimit -S -s unlimited}} to your
-{\texttt{.bash\_profile}} file and/or {\texttt{limit stacksize
-unlimited }} to your {\texttt{.cshrc}} file to suppress any potential
-limit to the size of the Unix stack.\\
-
-Then, to configure the software for your system, run the \texttt{configure}
-shell script. This script will attempt to guess the appropriate configuration
-values for your system. However, at a minimum, it is recommended that
-you explicitly specify the appropriate command names for your Fortran
-compiler and MPI package (another option is to define FC, CC and MPIF90
-in your .bash\_profile or your .cshrc file):
-\begin{lyxcode}
-./configure~FC=ifort~MPIFC=mpif90
-\end{lyxcode}
-Note that MPI must be installed with MPI-IO enabled because parts of SPECFEM3D perform I/Os through MPI-IO.
-
-Before running the \texttt{configure} script, you should probably
-edit file \texttt{flags.guess} to make sure that it contains the best
-compiler options for your system. Known issues or things to check
-are:
-\begin{description}
-\item [{\texttt{Intel ifort compiler}}] See if you need to add \texttt{-assume byterecl} for your machine.
-\textbf{In the case of that compiler, we have noticed that versions dot zero sometimes have bugs or issues that can lead to wrong results when running the code, thus we \emph{strongly} recommend using versions dot one or above (for instance version 13.1 instead of 13.0, version 14.1 instead of 14.0 and so on)}.
-\item [{\texttt{IBM compiler}}] See if you need to add \texttt{-qsave}
-or \texttt{-qnosave} for your machine.
-\item [{\texttt{Mac OS}}] You will probably need to install \texttt{XCODE}.
-\end{description}
-When compiling on an IBM machine with the \texttt{xlf} and \texttt{xlc}
-compilers, we suggest running the \texttt{configure} script with the
-following options:\\
-
-\noindent \texttt{./configure FC=xlf90\_r MPIFC=mpif90 CC=xlc\_r CFLAGS=\textquotedbl{}-O3
--q64\textquotedbl{} FCFLAGS=\textquotedbl{}-O3 -q64\textquotedbl{}
--with-scotch-dir=...}.\\
-
-On SGI systems, \texttt{flags.guess} automatically informs \texttt{configure}
-to insert `\texttt{`TRAP\_FPE=OFF}'' into the generated \texttt{Makefile}
-in order to turn underflow trapping off.\\
-
-You can add \texttt{--enable-vectorization} to the configuration options to speed up the code in the fluid (acoustic) and elastic parts.
-This works fine if (and only if) your computer always allocates a contiguous memory block for each allocatable array;
-this is the case for most machines and most compilers, but not all. To disable this feature, use option \texttt{--disable-vectorization}.
-For more details see \href{https://github.com/geodynamics/specfem3d/issues/81}{github.com/geodynamics/specfem3d/issues/81} .
-To check if that option works fine on your machine, run the code with and without it for an acoustic/elastic model and make sure the seismograms are identical.\\
-
-Note that we use CUBIT (now called Trelis) to create meshes of hexahedra, but other packages
-can be used as well, for instance GiD from \url{http://gid.cimne.upc.es}
-or Gmsh from \url{http://geuz.org/gmsh} \citep{GeRe09}. Even mesh
-creation packages that generate tetrahedra, for instance TetGen from
-\url{http://tetgen.berlios.de}, can be used because each tetrahedron
-can then easily be decomposed into four hexahedra as shown in the
-picture of the TetGen logo at \url{http://tetgen.berlios.de/figs/Delaunay-Voronoi-3D.gif};
-while this approach does not generate hexahedra of optimal quality,
-it can ease mesh creation in some situations and it has been shown
-that the spectral-element method can very accurately handle distorted
-mesh elements \citep{OlSe11}.
-
-The SPECFEM3D Cartesian software package relies on the SCOTCH library
-to partition meshes created with CUBIT. METIS \citep{KaKu98a,KaKu98c,KaKu98b}
-can also be used instead of SCOTCH if you prefer, by editing file
-src/decompose\_mesh/decompose\_mesh.F90 and uncommenting the flag
-USE\_METIS\_INSTEAD\_OF\_SCOTCH. You will also then need to install
-and compile Metis version 4.0 (do {*}NOT{*} install Metis version
-5.0, which has incompatible function calls) and edit \texttt{src/decompose\_mesh/Makefile.in}
-and uncomment the METIS link flag in that file before running \texttt{configure}.\\
-
-The SCOTCH library \citep{PeRo96} provides efficent static mapping,
-graph and mesh partitioning routines. SCOTCH is a free software package
-developed by Fran\c{c}ois Pellegrini et al. from LaBRI and INRIA in Bordeaux,
-France, downloadable from the web page \url{https://gforge.inria.fr/projects/scotch/}.
-In case no SCOTCH libraries can be found on the system, the configuration
-will bundle the version provided with the source code for compilation.
-The path to an existing SCOTCH installation can to be set explicitly
-with the option \texttt{-{}-with-scotch-dir}. Just as an example:
-\begin{lyxcode}
-./configure~FC=ifort~MPIFC=mpif90~-{}-with-scotch-dir=/opt/scotch
-\end{lyxcode}
-If you use the Intel ifort compiler to compile the code, we recommend
-that you use the Intel icc C compiler to compile Scotch, i.e., use:
-\begin{lyxcode}
-./configure~CC=icc~FC=ifort~MPIFC=mpif90
-\end{lyxcode}
-
-When compiling the SCOTCH source code, if you get a message such as: "ld: cannot find -lz",
-the Zlib compression development library is probably missing on your machine and you will need to install it or ask your system administrator to
-do so. On Linux machines the package is often called "zlib1g-dev" or similar. (thus "sudo apt-get install zlib1g-dev" would install it)
-
-To compile a serial version of the code for small meshes that fits
-on one compute node and can therefore be run serially, run \texttt{configure}
-with the \texttt{-{}-without-mpi} option to suppress all calls to
-MPI.
-
-A summary of the most important configuration variables follows.
-\begin{description}
-\item [{\texttt{F90}}] Path to the Fortran compiler.
-\item [{\texttt{MPIF90}}] Path to MPI Fortran.
-\item [{\texttt{MPI\_FLAGS}}] Some systems require this flag to link to
-MPI libraries.
-\item [{\texttt{FLAGS\_CHECK}}] Compiler flags.
-\end{description}
-The configuration script automatically creates for each executable
-a corresponding \texttt{Makefile} in the \texttt{src/} subdirectoy.
-The \texttt{Makefile} contains a number of suggested entries for various
-compilers, e.g., Portland, Intel, Absoft, NAG, and Lahey. The software
-has run on a wide variety of compute platforms, e.g., various PC clusters
-and machines from Sun, SGI, IBM, Compaq, and NEC. Select the compiler
-you wish to use on your system and choose the related optimization
-flags. Note that the default flags in the \texttt{Makefile} are undoubtedly
-not optimal for your system, so we encourage you to experiment with
-these flags and to solicit advice from your systems administrator.
-Selecting the right compiler and optimization flags can make a tremendous
-difference in terms of performance. We welcome feedback on your experience
-with various compilers and flags.
-
-Now that you have set the compiler information, you need to select
-a number of flags in the \texttt{constants.h} file depending on your
-system:
-\begin{description}
-\item [{\texttt{LOCAL\_PATH\_IS\_ALSO\_GLOBAL}}] Set to \texttt{.false.}
-on most cluster applications. For reasons of speed, the (parallel)
-distributed database generator typically writes a (parallel) database
-for the solver on the local disks of the compute nodes. Some systems
-have no local disks, e.g., BlueGene or the Earth Simulator, and other
-systems have a fast parallel file system, in which case this flag
-should be set to \texttt{.true.}. Note that this flag is not used
-by the database generator or the solver; it is only used for some
-of the post-processing.
-\end{description}
-The package can run either in single or in double precision mode.
-The default is single precision because for almost all calculations
-performed using the spectral-element method using single precision
-is sufficient and gives the same results (i.e. the same seismograms);
-and the single precision code is faster and requires exactly half
-as much memory. Select your preference by selecting the appropriate
-setting in the \texttt{constants.h} file:
-\begin{description}
-\item [{\texttt{CUSTOM\_REAL}}] Set to \texttt{SIZE\_REAL} for single precision
-and \texttt{SIZE\_DOUBLE} for double precision.
-\end{description}
-In the \texttt{precision.h} file:
-\begin{description}
-\item [{\texttt{CUSTOM\_MPI\_TYPE}}] Set to \texttt{MPI\_REAL} for single
-precision and \texttt{MPI\_DOUBLE\_PRECISION} for double precision.
-\end{description}
-On many current processors (e.g., Intel, AMD, IBM Power), single precision
-calculations are significantly faster; the difference can typically
-be 10\% to 25\%. It is therefore better to use single precision. What
-you can do once for the physical problem you want to study is run
-the same calculation in single precision and in double precision on
-your system and compare the seismograms. If they are identical (and
-in most cases they will), you can select single precision for your
-future runs.\\
-
-If your compiler has problems with the \texttt{use mpi} statements that are used in the code, use the script called
-\texttt{replace\_use\_mpi\_with\_include\_mpif\_dot\_h.pl} in the root directory to replace all of them with \texttt{include 'mpif.h'} automatically.
-
-\section{Adding OpenMP support in addition to MPI}
-
-NOTE FROM JULY 2013: OpenMP support is maybe / probably not maintained any more.
-Thus the section below is maybe obsolete.
-
-OpenMP support can be enabled in addition to MPI. However, in many
-cases performance will not improve because our pure MPI implementation
-is already heavily optimized and thus the resulting code will in fact
-be slightly slower. A possible exception could be IBM BlueGene-type
-architectures.
-
-To enable OpenMP, uncomment the OpenMP compiler option in two lines
-in file \texttt{src/specfem3D/Makefile.in} (before running \texttt{configure})
-and also uncomment the \texttt{\#define USE\_OPENMP} statement in
-file\\
- \texttt{src/specfem3D/specfem3D.F90}.
-
-The DO-loop using OpenMP threads has a SCHEDULE property. The OMP\_SCHEDULE
-environment variable can set the scheduling policy of that DO-loop.
-Tests performed by Marcin Zielinski at SARA (The Netherlands) showed
-that often the best scheduling policy is DYNAMIC with the size of
-the chunk equal to the number of OpenMP threads, but most preferably
-being twice as the number of OpenMP threads (thus chunk size = 8 for
-4 OpenMP threads etc). If OMP\_SCHEDULE is not set or is empty, the
-DO-loop will assume generic scheduling policy, which will slow down
-the job quite a bit.
-
-\section{Compiling on an IBM BlueGene}
-
-\underline{More recent installation instruction for IBM BlueGene, from April 2013:}\\
-
-\noindent Edit file \texttt{flags.guess} and put this for \texttt{FLAGS\_CHECK}:\\
-
-\noindent \texttt{-g -qfullpath -O2 -qsave -qstrict -qtune=qp -qarch=qp
--qcache=auto -qhalt=w}\\
- \texttt{-qfree=f90 -qsuffix=f=f90 -qlanglvl=95pure -Q -Q+rank,swap\_all
--Wl,-relax}\\
-
-\noindent The most relevant are the -qarch and -qtune flags, otherwise
-if these flags are set to ``auto'' then they are wrongly assigned
-to the architecture of the frond-end node, which is different from
-that on the compute nodes. You will need to set these flags to the
-right architecture for your BlueGene compute nodes, which is not necessarily
-``qp''; ask your system administrator. On some machines if is necessary
-to use -O2 in these flags instead of -O3 due to a compiler bug of
-the XLF version installed. We thus suggest to first try -O3, and then
-if the code does not compile or does not run fine then switch back
-to -O2. The debug flags (-g, -qfullpath) do not influence performance
-but are useful to get at least some insights in case of problems.\\
-
-\noindent Before running \texttt{configure}, select the XL Fortran
-compiler by typing \texttt{module load bgq-xl/1.0} or \texttt{module
-load bgq-xl} (another, less efficient option is to load the GNU compilers
-using \texttt{module load bgq-gnu/4.4.6} or similar).\\
-
-\noindent Then, to configure the code, type this:\\
-
-\noindent \texttt{./configure FC=bgxlf90\_r MPIFC=mpixlf90\_r CC=bgxlc\_r
-LOCAL\_PATH\_IS\_ALSO\_GLOBAL=true}\\
-
-In order for the SCOTCH domain decomposer to compile, on some (but
-not all) Blue Gene systems you may need to run \texttt{configure}
-with \texttt{CC=gcc} instead of \texttt{CC=bgxlc\_r}.
-
-\noindent \underline{Older installation instruction for IBM BlueGene, from 2011:}\\
-
-\noindent To compile the code on an IBM BlueGene, Laurent L\'eger from
-IDRIS, France, suggests the following: compile the code with
-
-\noindent \texttt{ FLAGS\_CHECK=\textquotedbl{}-O3 -qsave -qstrict
--qtune=auto -qarch=450d -qcache=auto }~\\
- \texttt{ -qfree=f90 -qsuffix=f=f90 -g -qlanglvl=95pure -qhalt=w -Q
--Q+rank,swap\_all -Wl,-relax\textquotedbl{}}
-
-\noindent Option \textquotedbl{}-Wl,-relax\textquotedbl{} must be
-added on many (but not all) BlueGene systems to be able to link the
-binaries \texttt{xmeshfem3D} and \texttt{xspecfem3D} because the final
-link step is done by the GNU \texttt{ld} linker even if one uses \texttt{FC=bgxlf90\_r,
-MPIFC=mpixlf90\_r} and \texttt{CC=bgxlc\_r} to create all the object
-files. On the contrary, on some BlueGene systems that use the native
-AIX linker option \textquotedbl{}-Wl,-relax\textquotedbl{} can lead
-to problems and must be suppressed from \texttt{flags.guess}.
-
-\noindent Also, \texttt{AR=ar, ARFLAGS=cru} and \texttt{RANLIB=ranlib}
-are hardwired in all \texttt{Makefile.in} files by default, but to
-cross-compile on BlueGene/P one needs to change these values to \texttt{AR=bgar,
-ARFLAGS=cru} and \texttt{RANLIB=bgranlib}. Thus the easiest thing
-to do is to modify all \texttt{Makefile.in} files and the \texttt{configure}
-script to set them automatically by \texttt{configure}. One then just
-needs to pass the right commands to the \texttt{configure} script:
-
-\noindent \texttt{ ./configure -{}-prefix=/path/to/SPECFEM3DG\_SP
--{}-host=Babel -{}-build=BGP }~\\
- \texttt{ FC=bgxlf90\_r MPIFC=mpixlf90\_r CC=bgxlc\_r AR=bgar ARFLAGS=cru
-}~\\
- \texttt{ RANLIB=bgranlib LOCAL\_PATH\_IS\_ALSO\_GLOBAL=false}
-
-\noindent This trick can be useful for all hosts on which one needs
-to cross-compile.
-
-\noindent On BlueGene, one also needs to run the \texttt{xcreate\_header\_file}
-binary file manually rather than in the Makefile:
-
-\noindent \texttt{bgrun -np 1 -mode VN -exe ./bin/xcreate\_header\_file}
-
-\section{Visualizing the subroutine calling tree of the source code}
-
-\noindent
-Packages such as \texttt{Doxywizard} can be used to visualize the calling tree of the
-subroutines of the  source code. \texttt{Doxywizard} is
-a GUI front-end for configuring and running \texttt{Doxygen}.
-
-\bigskip
-
-\noindent To do your own call graphs, you can follow these simple steps below.
-
-\begin{enumerate}
-\setcounter{enumi}{-1}
-\item Install \texttt{Doxygen} \underline{and} \texttt{graphviz} (the two are usually in the package manager of classic Linux distribution).
-\item Run in the terminal : \texttt{doxygen -g}, which creates a \texttt{Doxyfile} that tells doxygen what you want it to do.
-\item Edit the Doxyfile. Two Doxyfile-type files have been already committed in the directory \\
-\noindent \texttt{specfem3d/doc/Call\_trees}:
-
-\begin{itemize}
-\item[\textbullet] \texttt{Doxyfile\_truncated\_call\_tree} will generate call graphs with maximum 3 or 4 levels of tree structure,
-\item[\textbullet] \texttt{Doxyfile\_complete\_call\_tree} will generate call graphs with complete tree structure.
-\end{itemize}
-
-The important entries in the Doxyfile are:
-
-\begin{description}
-\item [{\texttt{PROJECT\_NAME}}]
-\item [{\texttt{OPTIMIZE\_FOR\_FORTRAN}}] Set to YES
-\item [{\texttt{EXTRACT\_ALL}}] Set to YES
-\item [{\texttt{EXTRACT\_PRIVATE}}] Set to YES
-\item [{\texttt{EXTRACT\_STATIC}}] Set to YES
-\item [{\texttt{INPUT}}] From the directory \texttt{specfem3d/doc/Call\_trees}, it is \texttt{"../../src/"}
-\item [{\texttt{FILE\_PATTERNS}}] In SPECFEM case, it is \texttt{*.f90* *.F90* *.c* *.cu* *.h*}
-\item [{\texttt{HAVE\_DOT}}] Set to YES
-\item [{\texttt{CALL\_GRAPH}}] Set to YES
-\item [{\texttt{CALLER\_GRAPH}}] Set to YES
-\item [{\texttt{DOT\_PATH}}] The path where is located the dot program graphviz (if it is not in your \$PATH)
-\item [{\texttt{RECURSIVE}}] This tag can be used to turn specify whether or not subdirectories should be searched for input files as well. In the case of SPECFEM, set to YES.
-\item [{\texttt{EXCLUDE}}] Here, you can exclude:  \\
-\texttt{"../../src/specfem3D/older\_not\_maintained\_partial\_OpenMP\_port"}\\
-\texttt{"../../src/decompose\_mesh/scotch"}\\
-\texttt{"../../src/decompose\_mesh/scotch\_5.1.12b"}
-\item [{\texttt{DOT\_GRAPH\_MAX\_NODES}}] to set the maximum number of nodes that will be shown in the graph.
-If the number of nodes in a graph becomes larger than this value, doxygen will truncate the graph, which is visualized by representing a node as a red box.
-Minimum value: 0, maximum value: 10000, default value: 50.
-\item [{\texttt{MAX\_DOT\_GRAPH\_DEPTH}}] to set the maximum depth of the graphs generated by dot.
-A depth value of 3 means that only nodes reachable from the root by following a path via at most 3 edges will be shown.
-Using a depth of 0 means no depth restriction.
-Minimum value: 0, maximum value: 1000, default value: 0.
-
-\end{description}
-
-\item Run : \texttt{doxygen Doxyfile}, HTML and LaTeX files created by default in \texttt{html} and \texttt{latex} subdirectories.
-
-\item To see the call trees, you have to open the file \texttt{html/index.html} in your \underline{browser}.
-You will have many informations about each subroutines of SPECFEM (not only call graphs), you can click on every boxes / subroutines.
-It show you the call, and, the caller graph of each subroutine : the subroutines called by the concerned subroutine, and the previous subroutines who call this subroutine (the previous path), respectively.
-In the case of a truncated calling tree, the boxes with a red border indicates a node that has \underline{more} arrows than are shown (in other words: the graph is truncated with respect to this node).
-
-\end{enumerate}
-
-\medskip
-
-\noindent Finally, some useful links:
-
-\begin{itemize}
-\item[\textbullet] a good and short summary for the basic utilisation of Doxygen:
-
-\url{http://www.softeng.rl.ac.uk/blog/2010/jan/30/callgraph-fortran-doxygen/},
-
-\item[\textbullet] to configure the diagrams :
-
-\url{http://www.stack.nl/~dimitri/doxygen/manual/diagrams.html},
-
-\item[\textbullet] the complete alphabetical index of the tags in Doxyfile:
-
-\url{http://www.stack.nl/~dimitri/doxygen/manual/config.html},
-
-\item[\textbullet] more generally, the Doxygen manual:
-
-\url{http://www.stack.nl/~dimitri/doxygen/manual/index.html}.
-
-\end{itemize}
-
-\medskip
-
-\section{Using the ADIOS library for I/O}
-Regular POSIX I/O can be problematic when dealing with large simulations one large
-clusters (typically more than $10,000$ processes). SPECFEM3D use the ADIOS library~\cite{Liu2013}
-to deal transparently take advantage of advanced paralllel file system features. To enable
-ADIOS, the following steps should be done:
-\begin{enumerate}
-\item Install ADIOS (available from \url{https://www.olcf.ornl.gov/center-projects/adios/}).
-Make sure that your environment variables reference it.
-\item You may want to change ADIOS related values in the \texttt{constants.h.in} file.
-The default values probably suit most cases.
-\item Configure using the \texttt{--with-adios} flag.
-\end{enumerate}
-ADIOS is currently only usable for meshfem3D generated mesh (i.e. not for meshes generated
-with CUBTI). Additional control parameters are discussed in section~\ref{cha:Main-Parameter}.
-
-\section{Becoming a developer of the code, or making small modifications in the source code}
-
-If you want to develop new features in the code, and/or if you want to make small changes, improvements, or bug fixes, you are very welcome to contribute.
-To do so, i.e.
-to access the development branch of the source code with read/write access (in a safe way, no need to worry too much about breaking the package, there is a robot called BuildBot that is in charge of checking and validating all new contributions and changes), please visit this Web page: \url{https://github.com/geodynamics/specfem3d/wiki/Using-Hub}.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\chapter{\label{cha:Mesh-Generation}Mesh Generation}
-
-The first step in running a spectral-element simulation consists of
-constructing a high-quality mesh for the region under consideration.
-We provide two possibilities to do so: (1) relying on the external,
-hexahedral mesher CUBIT, or (2) using the provided, internal mesher
-\texttt{xmeshfem3D}. In the following, we explain these two approaches.
-
-\section{\label{cha:Running-the-Mesher-CUBIT}Meshing with \texttt{CUBIT}}
-
-CUBIT is a meshing tool suite for the creation of finite-element meshes
-for arbitrarily shaped models. It has been developed and maintained
-at Sandia National Laboratories and can be purchased for a small academic
-institutional fee at \url{http://cubit.sandia.gov}. Our experience
-showed that using CUBIT greatly facilitates and speeds up the generation
-and preparation of hexahedral, conforming meshes for a variety of
-geophysical models with increasing complexity.
-
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=0.6\textwidth]{figures/mount-cubit}
-\par\end{centering}
-
-\caption{Example of the graphical user interface of CUBIT. The hexahedral mesh
-shown in the main display consists of a hexahedral discretization
-of a single volume with topography.}
-
-\label{fig:mount.cubit}
-\end{figure}
-
-\noindent
-The basic steps in creating a load-balanced, partitioned mesh with
-CUBIT are:
-\begin{description}
-\item [{1.}] setting up a hexahedral mesh with CUBIT,
-\item [{2.}] exporting the CUBIT mesh into a SPECFEM3D Cartesian file format
-and
-\item [{3.}] partitioning the SPECFEM3D Cartesian mesh files for a chosen
-number of cores.
-\end{description}
-Examples are provided in the SPECFEM3D Cartesian package in the subdirectory
-\texttt{EXAMPLES/}. We strongly encourage you to contribute your own
-example to this package by contacting the CIG Computational Seismology
-Mailing List \urlwithparentheses{cig-seismo@geodynamics.org}.
-
-\subsection{Creating the Mesh with CUBIT}
-
-For the installation and handling of the CUBIT meshing tool suite,
-please refer to the CUBIT user manual and documentation. In order
-to give you a basic understanding of how to use CUBIT for our purposes,
-examples are provided in the SPECFEM3D Cartesian package in the subdirectory
-\texttt{EXAMPLES/}:
-\begin{description}
-\item [{\texttt{homogeneous\_halfspace}}] Creates a single block model
-and assigns elastic material parameters.
-\item [{\texttt{layered\_halfspace}}] Combines two different, elastic material
-volumes and creates a refinement layer between the two. This example
-can be compared for validation against the solutions provided in subdirectory
-~\\
- \texttt{VALIDATION\_3D\_SEM\_SIMPLER\_LAYER\_SOURCE\_DEPTH/}.
-\item [{\texttt{waterlayered\_halfspace}}] Combines an acoustic and elastic
-material volume as in a schematic marine survey example.
-\item [{\texttt{tomographic\_model}}] Creates a single block model whose
-material properties will have to be read in from a tomographic model
-file during the databases creation by \texttt{xgenerate\_databases}.
-\end{description}
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=0.45\textwidth]{figures/example-homogeneous}
-\includegraphics[width=0.45\textwidth]{figures/example-2layers} \\
- \includegraphics[width=0.45\textwidth]{figures/example-water} \includegraphics[width=0.45\textwidth]{figures/example-tomo}
-\par\end{centering}
-
-\caption{Screenshots of the CUBIT examples provided in subdirectory \texttt{EXAMPLES/}:
-homogeneous halfspace (top-left), layered halfspace (top-right), water
-layered halfspace (bottom-left) and tomographic model (bottom-right).}
-
-\label{fig:examples.cubit}
-\end{figure}
-
-In each example subdirectory you will find a \texttt{README} file,
-which explains in a step-by-step tutorial the workflow for the example.
-Please feel free to contribute your own example to this package by
-contacting the CIG Computational Seismology Mailing List \urlwithparentheses{cig-seismo@geodynamics.org}.
-
-In some cases, to re-create the meshes for the examples given, just type "claro ./create\_mesh.py" or similar
-from the command line ("claro" is the command to run CUBIT from the command line).
-
-IMPORTANT note: in order to correctly set up GEOCUBIT and run the examples, please read the file called "EXAMPLES/README";
-in particular, please make sure you correctly set up the Python paths as indicated in that file.
-
-\subsection{Exporting the Mesh with \texttt{run\_cubit2specfem3d.py} }
-
-\label{subsec:Exporting-the-Mesh}
-
-Once the geometric model volumes in CUBIT are meshed, you prepare
-the model for exportation with the definition of material blocks and
-boundary surfaces. Thus, prior to exporting the mesh, you need to
-define blocks specifying the materials and absorbing boundaries in
-CUBIT. This process could be done automatically using the script \texttt{run\_cubit2specfem3d.py}
-if the mesh meets some conditions or manually, following the block
-convention:
-\begin{description}
-\item [{material\_name}] Each material should have a specific block defined
-by a unique name. The name convention of the material is to start
-with either 'elastic' or 'acoustic'. It must be then followed by a
-unique identifier, e.g. 'elastic 1', 'elastic 2', etc. The additional
-attributes to the block define the material description.
-
-For an elastic material:
-\begin{description}
-\item [{material\_id}] An integer value which is unique for this material.
-\item [{Vp}] P-wave speed of the material (given in m/s).
-\item [{Vs}] S-wave speed of the material (given in m/s).
-\item [{rho}] density of the material (given in kg/m$^{3}$).
-\item [{Q}] quality factor to use in case of a simulation with attenuation
-turned on. It should be between 1 and 9000. In case no attenuation
-information is available, it can be set to zero. Please note that
-your Vp- and Vs-speeds are given for a reference frequency. To change
-this reference frequency, you change the value of \texttt{ATTENUATION\_f0\_REFERENCE}
-in the main constants file \texttt{constants.h} found in subdirectory
-\texttt{src/shared/}.
-\item [{anisotropic\_flag}] Flag describing the anisotropic model to use
-in case an anisotropic simulation should be conducted. See the file
-\texttt{model\_aniso.f90} in subdirectory \texttt{src/generate\_databases/}
-for an implementation of the anisotropic models. In case no anisotropy
-is available, it can be set to zero.
-\end{description}
-
-Note that this material block has to be defined using all the volumes
-which belong to this elastic material. For volumes belonging to another,
-different material, you will need to define a new material block.
-
-For an acoustic material:
-\begin{description}
-\item [{material\_id}] An integer value which is unique for this material.
-\item [{Vp}] P-wave speed of the material (given in m/s).
-\item [{0}] S-wave speed of the material is ignored.
-\item [{rho}] density of the material (given in kg/m$^{3}$).
-\end{description}
-\item [{face\_topo}] Block definition for the surface which defines the
-free surface (which can have topography). The name of this block must
-be 'face\_topo', the block has to be defined using all the surfaces
-which constitute the complete free surface of the model.
-\item [{face\_abs\_xmin}] Block definition for the faces on the absorbing
-boundaries, one block for each surface with x=Xmin.
-\item [{face\_abs\_xmax}] Block definition for the faces on the absorbing
-boundaries, one block for each surface with x=Xmax.
-\item [{face\_abs\_ymin}] Block definition for the faces on the absorbing
-boundaries, one block for each surface with y=Ymin.
-\item [{face\_abs\_ymax}] Block definition for the faces on the absorbing
-boundaries, one block for each surface with y=Ymax.
-\item [{face\_abs\_bottom}] Block definition for the faces on the absorbing
-boundaries, one block for each surface with z=bottom.
-\end{description}
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=0.45\textwidth]{figures/mount-surface} \includegraphics[width=0.45\textwidth]{figures/mount-abs}
-\par\end{centering}
-
-\caption{Example of the block definitions for the free surface 'face\_topo'
-(left) and the absorbing boundaries, defined in a single block 'face\_abs\_xmin'
-(right) in CUBIT.}
-
-\label{fig:mount.abs}
-\end{figure}
-
-Optionally, instead of specifying for each surface at the boundaries
-a single block like mentioned above, you can also specify a single
-block for all boundary surfaces and name it as one of the absorbing
-blocks above, e.g. 'face\_abs\_xmin'.
-
-After the block definitions are done, you export the mesh using the
-script \texttt{cubit2specfem3d.py} provided in each of the example
-directories (linked to the common script \texttt{CUBIT\_GEOCUBIT/cubit2specfem3d.py}).
-If the export was successful, you should find the following
-files in a subdirectory \texttt{MESH/}:
-\begin{description}
-\item [{absorbing\_cpml\_file}] \textbf{(only needed in case of C-PML absorbing
-conditions)} Contains on the first line the total number of C-PML
-spectral elements in the mesh, and then on the following line the
-list of all these C-PML elements with two numbers per line: first
-the spectral element number, and then a C-PML flag indicating to which
-C-PML layer(s) that element belongs, according to the following convention:
-
-\begin{itemize}
-\item Flag = 1 : element belongs to a X CPML layer only (either in Xmin
-or in Xmax),
-\item Flag = 2 : element belongs to a Y CPML layer only (either in Ymin
-or in Ymax),
-\item Flag = 3 : element belongs to a Z CPML layer only (either in Zmin
-or in Zmax),
-\item Flag = 4 : element belongs to a X CPML layer and also to a Y CPML
-layer,
-\item Flag = 5 : element belongs to a X CPML layer and also to a Z CPML
-layer,
-\item Flag = 6 : element belongs to a Y CPML layer and also to a Z CPML
-layer,
-\item Flag = 7 : element belongs to a X, to a Y and to a Z CPML layer, i.e.,
-it belongs to a CPML corner.
-\end{itemize}
-
-Note that it does not matter whether an element belongs to a Xmin
-or to a Xmax CPML, the flag is the same in both cases; the same is
-true for Ymin and Ymax, and also Zmin and Zmax.
-
-When you have an existing CUBIT (or similar) mesh stored in SPECFEM3D
-format, i.e., if you have existing \textquotedbl{}nodes\_coords\_file\textquotedbl{}
-and \textquotedbl{}mesh\_file\textquotedbl{} files but do not know
-how to assign CPML flags to them, we have created a small serial Fortran
-program that will do that automatically for you, i.e., which will
-create the \textquotedbl{}absorbing\_cpml\_file\textquotedbl{} for
-you. That program is utils/CPML/convert\_external\_layers\_of\_a\_given\_mesh\_to\_CPML\_layers.f90,
-and a small Makefile is provided in that directory (utils/CPML).
-
-IMPORTANT: it is your responsibility to make sure that in the input
-CUBIT (or similar) mesh that this code will read in SPECFEM3D format
-from files \textquotedbl{}nodes\_coords\_file\textquotedbl{} and \textquotedbl{}mesh\_file\textquotedbl{}
-you have created layers of elements that constitute a layer of constant
-thickness aligned with the coordinate grid axes (X, Y and/or Z), so
-that this code can assign CPML flags to them. This code does NOT check
-that (because it cannot, in any easy way). The mesh inside these CPML
-layers does not need to be structured nor regular, any non-structured
-mesh is fine as long as it has flat PML inner and outer faces, parallel
-to the axes, and thus of a constant thickness. The thickness can be
-different for the X, Y and Z sides. But for X it must not vary, for
-Y it must not vary, and for Z it must not vary. If you do not know
-the exact thickness, you can use a slightly LARGER value in this code
-(say 2\% to 5\% more) and this code will fix that and will adjust
-it; never use a SMALLER value otherwise this code will miss some CPML
-elements.
-
-Note: in a future release we will remove the constraint of having
-CPML layers aligned with the coordinate axes; we will allow for meshes
-that are titled by any constant angle in the horizontal plane. However
-this is not implemented yet.
-
-Note: in the case of fluid-solid layers in contact, the fluid-solid
-interface currently needs to be flat and horizontal inside the CPML
-layer (i.e., bathymetry should become flat and horizontal when entering
-the CPML); this (small) constraint will probably remain in the code
-for a while because it makes fluid-solid matching inside the CPML
-much easier.
-
-\item [{materials\_file}] Contains the material associations for each element.
-The format is:
-
-\begin{lyxcode}
-element\_ID~material\_ID
-\end{lyxcode}
-
-where \texttt{element\_ID} is the element identifier and \texttt{material\_ID}
-is a unique identifier, positive (for materials taken from this list
-of materials, i.e. for which each spectral element has constant material
-properties taken from this list) or negative (for tomographic models,
-i.e. for spectral element whose real velocities and density will be
-assigned later by calling an external function to define model variations,
-for instance in the case of tomographic models; in such a case, material
-properties can vary inside each spectral element, i.e. be different
-at each of its Gauss-Lobatto-Legendre grid points).
-
-\item [{nummaterial\_velocity\_file}] Defines the material properties.
-
-\begin{itemize}
-\item For classical materials (i.e., spectral elements for which the velocity
-and density model will not be assigned by calling an external function
-to define for instance a tomographic model), the format is:
-
-\begin{lyxcode}
-domain\_ID~material\_ID~rho~vp~vs~Qkappa~Qmu~anisotropy\_flag
-\end{lyxcode}
-
-where \texttt{\textbf{domain\_ID}}\textbf{ is 1 for acoustic and 2
-for elastic or viscoelastic materials,} \texttt{material\_ID} a unique
-identifier, \texttt{rho} the density in $kg\, m^{-3}$, \texttt{vp}
-the P-wave speed in $m\, s^{-1}$, \texttt{vs} the S-wave speed in
-$m\, s^{-1}$, \texttt{Q} the quality factor and \texttt{anisotropy\_flag}
-an identifier for anisotropic models. Note that the \texttt{Qkappa}
-value is ignored by the code unless \texttt{FULL\_ATTENUATION\_SOLID}
-is set. Note also that both \texttt{Qkappa} and \texttt{Qmu} are ignored
-by the code unless \texttt{ATTENUATION} is set. If you want a model
-with no \texttt{Qmu} attenuation, both set \texttt{ATTENUATION} to
-\texttt{.false.} in the \texttt{Par\_file} and set \texttt{Qmu} to
-9999 here. If you want a model with no \texttt{Qkappa} attenuation,
-both set \texttt{FULL\_ATTENUATION\_SOLID} to \texttt{.false.} in
-the \texttt{Par\_file} and set \texttt{Qkappa} to 9999 here. %%magnoni
-
-\item For tomographic velocity models, please read Chapter \ref{cha:-Changing-the}
-and Section \ref{sec:Using-tomographic} `Using external tomographic
-Earth models' for further details.
-\end{itemize}
-\item [{nodes\_coords\_file}] Contains the point locations in Cartesian
-coordinates of the mesh element corners.
-\item [{mesh\_file}] Contains the mesh element connectivity. The hexahedral
-elements can have 8 or 27 nodes.\\
- See picture doc/mesh\_numbering\_convention/numbering\_convention\_27\_nodes.jpg
-to see\\
- in which (standard) order the points must be cited. In the case of
-8 nodes, just include the first 8 points.
-\item [{free\_or\_absorbing\_surface\_file\_zmax}] Contains the free surface
-connectivity or \\
- the surface connectivity of the absorbing boundary surface at the
-top (Zmax), \\
- depending on whether the top surface is defined as free or absorbing
-(\texttt{STACEY\_INSTEAD\_OF\_FREE\_SURFACE} in \texttt{DATA/Par\_file}). \\
- You should put both the surface of acoustic regions and of elastic
-regions in that file; that is, list all the element faces that constitute
-the surface of the model in that file.
-\item [{absorbing\_surface\_file\_xmax}] Contains the surface connectivity
-of the absorbing boundary surface at Xmax \\
-(also needed in the case of C-PML absorbing conditions, in order for
-the code to be able to impose Dirichlet conditions on their outer
-edge).
-\item [{absorbing\_surface\_file\_xmin}] Contains the surface connectivity
-of the absorbing boundary surface at Xmin \\
-(also needed in the case of C-PML absorbing conditions, in order for
-the code to be able to impose Dirichlet conditions on their outer
-edge).
-\item [{absorbing\_surface\_file\_ymax}] Contains the surface connectivity
-of the absorbing boundary surface at Ymax \\
-(also needed in the case of C-PML absorbing conditions, in order for
-the code to be able to impose Dirichlet conditions on their outer
-edge).
-\item [{absorbing\_surface\_file\_ymin}] Contains the surface connectivity
-of the absorbing boundary surface at Ymin \\
-(also needed in the case of C-PML absorbing conditions, in order for
-the code to be able to impose Dirichlet conditions on their outer
-edge).
-\item [{absorbing\_surface\_file\_bottom}] Contains the surface connectivity
-of the absorbing boundary surface at the bottom (Zmin) \\
-(also needed in the case of C-PML absorbing conditions, in order for
-the code to be able to impose Dirichlet conditions on their outer
-edge).
-\end{description}
-These mesh files are needed as input files for the partitioner \texttt{xdecompose\_mesh}
-to load-balance the mesh. Please see the next section for further
-details.
-
-In directory "CUBIT\_GEOCUBIT" we provide a script that can help doing the above tasks of exporting a CUBIT mesh
-to SPECFEM3D format automatically for you: "run\_cubit2specfem3d.py".
-Just edit them to indicate the path to your local installation of CUBIT and also the name of the *.cub existing CUBIT mesh file
-that you want to export to SPECFEM3D format. These scripts will do the conversion for you automatically except assigning material properties
-to the different mesh layers. To do so, you will then need to edit the file
-called "nummaterial\_velocity\_file" that will have just been created and change it from the prototype created:
-
-    0 1 vol1 --> syntax: \#material\_domain\_id \#material\_id \#rho \#vp \#vs \#Q\_kappa \#Q\_mu \#anisotropy
-    0 2 vol2 --> syntax: \#material\_domain\_id \#material\_id \#rho \#vp \#vs \#Q\_kappa \#Q\_mu \#anisotropy
-
-(where "vol1" and "vol2" here represent the volume labels that you have set while creating the mesh in CUBIT)
-to for instance
-
-    2 1 1500 2300 1800 9999.0 9999.0 0
-    2 2 1600 2500 20000 9999.0 9999.0 0
-
-\subsubsection*{Checking the mesh quality}
-
-The quality of the mesh may be inspected more precisely based upon
-the serial code in the file \texttt{check\_mesh\_quality\_}~\\
- \texttt{CUBIT\_Abaqus.f90} located in the directory \texttt{src/check\_mesh\_quality\_CUBIT\_Abaqus/}.
-Running this code is optional because no information needed by the
-solver is generated.
-
-Prior to running and compiling this code, you have to export your
-mesh in CUBIT to an ABAQUS (.inp) format. For example, export mesh
-block IDs belonging to volumes in order to check the quality of the
-hexahedral elements. You also have to determine a number of parameters
-of your mesh, such as the number of nodes and number of elements and
-modify the header of the \texttt{check\_mesh\_quality\_CUBIT\_Abaqus.f90}
-source file in directory \texttt{src/check\_mesh\_quality\_CUBIT\_Abaqus/}.
-
-\noindent
-Then, in the main directory, type
-\begin{lyxcode}
-{\small make~xcheck\_mesh\_quality\_CUBIT\_Abaqus}{\small \par}
-\end{lyxcode}
-and use
-\begin{lyxcode}
-{\small ./bin/xcheck\_mesh\_quality\_CUBIT\_Abaqus}{\small \par}
-\end{lyxcode}
-to generate an %not supported: AVS output file (\texttt{\small AVS\_meshquality.inp} in AVS UCD format) or
-OpenDX output file (\texttt{\small DX\_mesh\_quality.dx}{\small )
-that can be used to investigate mesh quality, e.g. skewness of elements,
-and a Gnuplot histogram (}\texttt{\small mesh\_quality\_histogram.txt}{\small )
-that can be plotted with gnuplot (type `}\texttt{\small gnuplot plot\_mesh\_quality\_histogram.gnu}{\small ').
-The histogram is also printed to the screen. Analyze that skewness
-histogram of mesh elements to make sure no element has a skewness
-above approximately 0.75, otherwise the mesh is of poor quality (and
-if even a single element has a skewness value above 0.80, then you
-must definitely improve the mesh). If you want to start designing
-your own meshes, this tool is useful for viewing your creations. You
-are striving for meshes with elements with `cube-like' dimensions,
-e.g., the mesh should contain no very elongated or skewed elements.}{\small \par}
-
-\subsection{Partitioning the Mesh with \texttt{xdecompose\_mesh}}
-
-The SPECFEM3D Cartesian software package performs large scale simulations
-in a parallel 'Single Process Multiple Data' way. The spectral-element
-mesh created with CUBIT needs to be distributed on the processors.
-This partitioning is executed once and for all prior to the execution
-of the solver so it is referred to as a static mapping.
-
-An efficient partitioning is important because it leverages the overall
-running time of the application. It amounts to balance the number
-of elements in each slice while minimizing the communication costs
-resulting from the placement of adjacent elements on different processors.
-\texttt{decompose\_mesh} depends on the SCOTCH library \citep{PeRo96},
-which provides efficent static mapping, graph and mesh partitioning
-routines. SCOTCH is a free software package developed by Fran\c{c}ois
-Pellegrini et al. from LaBRI and INRIA in Bordeaux, France, downloadable
-from the web page \url{https://gforge.inria.fr/projects/scotch/}.
-\\
-
-In most cases, the configuration with \texttt{./configure~FC=ifort}
-should be sufficient. During the configuration process, the script
-tries to find existing SCOTCH installations. In case your system has
-no pre-existing SCOTCH installation, we provide the source code of
-SCOTCH, which is released open source under the French CeCILL-C version
-1 license, in directory \texttt{src/decompose\_mesh/scotch\_5.1.12b}.
-This version gets bundled with the compilation of the SPECFEM3D Cartesian
-package if no libraries could have been found. If this automatic compilation
-of the SCOTCH libraries fails, please refer to file INSTALL.txt in
-that directory to see further details how to compile it on your system.
-In case you want to use a pre-existing installation, make sure you
-have correctly specified the path of the SCOTCH library when using
-the option \texttt{-{}-with-scotch-dir} with the \texttt{./configure}
-script. In the future you should be able to find more recent versions
-at \url{http://www.labri.fr/perso/pelegrin/scotch/scotch_en.html}
-. \\
-
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=0.45\textwidth]{figures/mount-partitions}
-\includegraphics[width=0.45\textwidth]{figures/mount-partitions2}
-\par\end{centering}
-
-\caption{Example of a mesh partitioning onto four cores. Each single core partition
-is colored differently. The executable \texttt{xdecompose\_mesh} can
-equally distribute the mesh on any arbitrary number of cores. Domain
-decomposition is explained in detail in \citet{MaKoBlLe08}, and excellent
-scaling up to 150,000 processor cores in shown for instance in \citet{CaKoLaTiMiLeSnTr08,KoLaMi08a,MaKoBlLe08,KoErGoMi10,Kom11}.}
-
-\label{fig:mount.partitions}
-\end{figure}
-
-When you are ready to compile, in the main directory type
-\begin{lyxcode}
-{\small make~xdecompose\_mesh}{\small \par}
-\end{lyxcode}
-If all paths and flags have been set correctly, the executable \texttt{bin/xdecompose\_mesh}
-should be produced.
-
-The partitioning is done in serial for now (in the next release we
-will provide a parallel version of that code). It needs to be run in the main directory because it expects the \texttt{./DATA/Par\_file}. The synopsis is:
-\begin{lyxcode}
-./bin/xdecompose\_mesh~nparts~input\_directory~output\_directory
-\end{lyxcode}
-where
-\begin{itemize}
-\item \texttt{nparts} is the number of partitions, i.e., the number of cores
-for the parallel simulations,
-\item \texttt{input\_directory} is the directory which holds all the files
-generated by the Python script \texttt{cubit2specfem3d.py} explained
-in the previous Section~\ref{subsec:Exporting-the-Mesh}, e.g. \texttt{./MESH/},
-and
-\item \texttt{output\_directory} is the directory for the output of this
-partitioner which stores ACII-format files named like \texttt{proc{*}{*}{*}{*}{*}{*}\_Database}
-for each partition. These files will be needed for creating the distributed
-databases, and have to reside in the directory \texttt{LOCAL\_PATH}
-specified in the main \texttt{Par\_file}, e.g. in directory \texttt{./OUTPUT\_FILES/DATABASES\_MPI}.
-Please see Chapter~\ref{cha:Creating-Distributed-Databases} for
-further details.
-\end{itemize}
-Note that all the files generated by the Python script \texttt{cubit2specfem3d.py}
-must be placed in the \texttt{input\_directory} folder before running
-the program.
-
-\section{\label{cha:Running-the-Mesher-Meshfem3D}Meshing with \texttt{xmeshfem3D}}
-
-In case you successfully ran the configuration script, you are also
-ready to compile the internal mesher. This is an alternative to CUBIT
-for the mesh generation of relatively simple models.
-
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[scale=0.5]{figures/socal_map_mpi}
-\par\end{centering}
-
-\caption{\label{fig:For-parallel-computing}For parallel computing purposes,
-the model block is subdivided in $\nprocxi\times\nproceta$ slices
-of elements. In this example we use $5^{2}=25$ processors. }
-\end{figure}
-
-\noindent
-In the main directory, type
-\begin{lyxcode}
-{\small make~xmeshfem3D}{\small \par}
-\end{lyxcode}
-If all paths and flags have been set correctly, the mesher should
-now compile and produce the executable \texttt{bin/xmeshfem3D}. Please
-note that \texttt{xmeshfem3D} must be called directly from the main
-directory, as most of the binaries of the package.
-
-Input for the mesh generation program is provided through the parameter
-file \texttt{Mesh\_Par\_file}, which resides in the subdirectory \texttt{DATA/meshfem3D\_files/}.
-(to see how to use it, see the EXAMPLES specific to the internal mesher in directory \texttt{EXAMPLES/meshfem3D\_examples/}.
-Before running the mesher, a number of parameters need to be set in
-the \texttt{Mesh\_Par\_file}. This requires a basic understanding
-of how the SEM is implemented, and we encourage you to read \citet{KoVi98,KoTr99}
-and \citet{KoLiTrSuStSh04}.
-
-The mesher and the solver use UTM coordinates internally, therefore
-you need to define the zone number for the UTM projection (e.g., zone
-11 for Los Angeles). Use decimal values for latitude and longitude
-(no minutes/seconds). These values are approximate; the mesher will
-round them off to define a square mesh in UTM coordinates. When running
-benchmarks on rectangular models, turn the UTM projection off by using
-the flag \texttt{\small SUPPRESS\_UTM\_PROJECTION}{\small , in which
-case all `longitude' parameters simply refer to the $x$~axis, and
-all `latitude' parameters simply refer to the $y$~axis. To run the
-mesher for a global simulation, the following parameters need to be
-set in the }\texttt{\small Mesh\_Par\_file}{\small :}{\small \par}
-\begin{description}
-\item [{\texttt{LATITUDE\_MIN}}] Minimum latitude in the block (negative
-for South).
-\item [{\texttt{LATITUDE\_MAX}}] Maximum latitude in the block.
-\item [{\texttt{LONGITUDE\_MIN}}] Minimum longitude in the block (negative
-for West).
-\item [{\texttt{LONGITUDE\_MAX}}] Maximum longitude in the block.
-\item [{\texttt{DEPTH\_BLOCK\_KM}}] Depth of bottom of mesh in kilometers.
-\item [{\texttt{\noun{UTM\_PROJECTION\_ZONE}}}] UTM projection zone in
-which your model resides, only valid when \texttt{SUPPRESS\_UTM\_}~\\
- \texttt{PROJECTION} is \texttt{.false.}.
-Use a negative zone number for the Southern hemisphere:
-the Northern hemisphere corresponds to zones +1 to +60,
-the Southern hemisphere to zones -1 to -60.
-We use the WGS84 (World Geodetic System 1984) reference ellipsoid for the UTM projection. If you prefer to use the Clarke 1866 ellipsoid,
-edit file \texttt{src/shared/utm\_geo.f90}, uncomment that ellipsoid and recompile the code.
-From http://en.wikipedia.org/wiki/Universal\_Transverse\_Mercator\_coordinate\_system :
-The Universal Transverse Mercator coordinate system was developed by the United States Army Corps of Engineers in the 1940s.
-The system was based on an ellipsoidal model of Earth. For areas within the contiguous United States
-the Clarke Ellipsoid of 1866 was used. For the remaining areas of Earth, including Hawaii, the International Ellipsoid was used.
-The WGS84 ellipsoid is now generally used to model the Earth in the UTM coordinate system,
-which means that current UTM northing at a given point can be 200+ meters different from the old one.
-For different geographic regions, other datum systems (e.g.: ED50, NAD83) can be used.
-\item [{\texttt{SUPPRESS\_UTM\_PROJECTION}}] set to be \texttt{.false.}
-when your model range is specified in geographical coordinates, and
-needs to be \texttt{.true.} when your model is specified in Cartesian
-coordinates. \noun{UTM projection zone in which your simulation region
-resides.}
-\item [{\texttt{INTERFACES\_FILE }}] File which contains the description
-of the topography and of the interfaces between the different layers
-of the model, if any. The number of spectral elements in the vertical
-direction within each layer is also defined in this file.
-\item [{$\nexxi$}] The number of spectral elements along one side of the
-block. This number \textit{must} be 8~$\times$~a multiple of $\nprocxi$
-defined below. Based upon benchmarks against semi-analytical discrete
-wavenumber synthetic seismograms \citep{KoLiTrSuStSh04}, determined
-that a $\nexxi=288$ run is accurate to a shortest period of roughly
-2~s. Therefore, since accuracy is determined by the number of grid
-points per shortest wavelength, for any particular value of $\nexxi$
-the simulation will be accurate to a shortest period determined by
-\begin{equation}
-\mbox{shortest period (s)}=(288/\nexxi)\times2.\label{eq:shortest_period}
-\end{equation}
- The number of grid points in each orthogonal direction of the reference
-element, i.e., the number of Gauss-Lobatto-Legendre points, is determined
-by \texttt{NGLLX} in the \texttt{constants.h} file. We generally use
-$\mbox{\texttt{NGLLX\/}}=5$, for a total of $5^{3}=125$ points per
-elements. We suggest not to change this value.
-\item [{$\nexeta$}] The number of spectral elements along the other side
-of the block. This number \textit{must} be 8~$\times$~a multiple
-of $\nproceta$ defined below.
-\item [{$\nprocxi$}] The number of processors or slices along one side
-of the block (see Figure~\ref{fig:For-parallel-computing}); we must
-have $\nexxi=8\times c\times\nprocxi$, where $c\ge1$ is a positive
-integer.
-\item [{$\nproceta$}] The number of processors or slices along the other
-side of the block; we must have $\nexeta=8\times c\times\nproceta$,
-where $c\ge1$ is a positive integer.
-\item [{\texttt{USE\_REGULAR\_MESH}}] set to be \texttt{.true.} if you
-want a perfectly regular mesh or \texttt{.false.} if you want to add
-doubling horizontal layers to coarsen the mesh. In this case, you
-also need to provide additional information by setting up the next
-three parameters.
-\item [{\texttt{NDOUBLINGS}}] The number of horizontal doubling layers.
-Must be set to \texttt{1} or \texttt{2} if \texttt{USE\_REGULAR\_MESH}
-is set to \texttt{.true.}.
-\item [{\texttt{NZ\_DOUBLING\_1}}] The position of the first doubling layer
-(only interpreted if \texttt{USE\_REGULAR\_MESH} is set to \texttt{.true.}).
-\item [{\texttt{NZ\_DOUBLING\_2}}] The position of the second doubling
-layer (only interpreted if \texttt{USE\_REGULAR\_MESH} is set to \texttt{.true.}
-and if \texttt{NDOUBLINGS} is set to \texttt{2}).
-\item [{\texttt{CREATE\_ABAQUS\_FILES}}] Set this flag to \texttt{.true.}
-to save Abaqus FEA \urlwithparentheses{www.simulia.com} mesh files
-for subsequent viewing. Turning the flag on generates files in the
-\texttt{LOCAL\_PATH} directory. See Section~\ref{sec:Mesh-graphics}
-for a discussion of mesh viewing features.
-\item [{\texttt{CREATE\_DX\_FILES}}] Set this flag to \texttt{.true.} to
-save OpenDX \urlwithparentheses{www.opendx.org} mesh files for subsequent
-viewing.
-\item [{\texttt{LOCAL\_PATH}}] Directory in which the partitions generated
-by the mesher will be written. Generally one uses a directory on the
-local disk of the compute nodes, although on some machines these partitions
-are written on a parallel (global) file system (see also the earlier
-discussion of the \texttt{LOCAL\_PATH\_IS\_ALSO\_GLOBAL} flag in Chapter~\ref{cha:Getting-Started}).
-The mesher generates the necessary partitions in parallel, one set
-for each of the $\nprocxi\times\nproceta$ slices that constitutes
-the mesh (see Figure~\ref{fig:For-parallel-computing}). After the
-mesher finishes, you can log in to one of the compute nodes and view
-the contents of the \texttt{LOCAL\_PATH} directory to see the files
-generated by the mesher. These files will be needed for creating the
-distributed databases, and have to reside in the directory \texttt{LOCAL\_PATH}
-specified in the main \texttt{Par\_file}, e.g. in directory \texttt{OUTPUT\_FILES/DATABASES\_MPI}.
-Please see Chapter~\ref{cha:Creating-Distributed-Databases} for
-further details.
-\item [{\texttt{NMATERIALS}}] The number of different materials in your
-model. In the following lines, each material needs to be defined as
-:
-
-\begin{lyxcode}
-material\_ID~rho~vp~vs~Q~anisotropy\_flag~domain\_ID
-\end{lyxcode}
-
-where
-\begin{itemize}
-\item \texttt{Q} : quality factor (0=no attenuation)
-\item \texttt{anisotropy\_flag} : 0=no anisotropy / 1,2,.. check with implementation
-in \texttt{aniso\_model.f90}
-\item \texttt{domain\_id} : 1=acoustic / 2=elastic
-\end{itemize}
-\item [{\texttt{NREGIONS}}] The number of regions in the mesh. In the following
-lines, because the mesh is regular or 'almost regular', each region
-is defined as :
-
-\begin{lyxcode}
-NEX\_XI\_BEGIN~NEX\_XI\_END~NEX\_ETA\_BEGIN~NEX\_ETA\_END~NZ\_BEGIN~NZ\_END~material\_ID
-\end{lyxcode}
-\end{description}
-The \texttt{INTERFACES\_FILE} parameter of \texttt{Mesh\_Par\_File}
-defines the file which contains the settings of the topography grid
-and of the interfaces grids. Topography is defined as a set of elevation
-values on a regular 2D grid. It is also possible to define interfaces
-between the layers of the model in the same way. The file needs to
-define several parameters:
-\begin{itemize}
-\item The number of interfaces, including the topography. This needs to
-be set at the first line. Then, from the bottom to the top of the
-model, you need to define the grids with:
-\item \texttt{SUPPRESS\_UTM\_PROJECTION} flag as described previously,
-\item number of points along $x$ and $y$ direction (NXI and NETA),
-\item minimal $x$ and $y$ coordinates (LONG\_MIN and LAT\_MIN),
-\item spacing between points along $x$ and $y$ (SPACING\_XI and SPACING\_ETA)
-and
-\item the name of the file which contains the elevation values (in $y$.$x$
-increasing order).
-\end{itemize}
-At the end of this file, you simply need to set the number of spectral
-elements in the vertical direction for each layer. We provide a few
-models in the {\texttt{EXAMPLES/}} directory. \\
-
-Finally, depending on your system, you might need to provide a file
-that tells MPI what compute nodes to use for the simulations. The
-file must have a number of entries (one entry per line) at least equal
-to the number of processors needed for the run. A sample file is provided
-in the file \texttt{mymachines}. This file is not used by the mesher
-or solver, but is required by the \texttt{go\_mesher} and \texttt{go\_solver}
-default job submission scripts. See Chapter~\ref{cha:Scheduler}
-for information about running the code on a system with a scheduler,
-e.g., LSF.
-
-Now that you have set the appropriate parameters in the \texttt{Mesh\_Par\_file}
-and have compiled the mesher, you are ready to launch it! This is
-most easily accomplished based upon the \texttt{go\_mesher} script.
-When you run on a PC cluster, the script assumes that the nodes are
-named n001, n002, etc. If this is not the case, change the \texttt{tr
--d `n'} line in the script. You may also need to edit the last command
-at the end of the script that invokes the \texttt{mpirun} command.
-See Chapter~\ref{cha:Scheduler} for information about running the
-code on a system with a scheduler, e.g., LSF.
-
-Mesher output is provided in the \texttt{OUTPUT\_FILES} directory
-in \texttt{output\_mesher.txt}; this file provides lots of details
-about the mesh that was generated. Please note that the mesher suggests
-a time step \texttt{DT} to run the solver with. The mesher output
-file also contains a table about the quality of the mesh to indicate
-possible problems with the distortions of elements. Alternatively,
-output can be directed to the screen instead by uncommenting a line
-in \texttt{constants.h}:
-\begin{lyxcode}
-!~uncomment~this~to~write~messages~to~the~screen~
-
-!~integer,~parameter~::~IMAIN~=~ISTANDARD\_OUTPUT
-\end{lyxcode}
-To control the quality of the mesh, check the standard output (either
-on the screen or in the \texttt{OUTPUT\_FILES} directory in \texttt{output\_mesher.txt})
-and analyze the skewness histogram of mesh elements to make sure no
-element has a skewness above approximately 0.75, otherwise the mesh
-is of poor quality (and if even a single element has a skewness value
-above 0.80, then you must definitely improve the mesh). To draw the
-skewness histogram on the screen, type \texttt{gnuplot plot\_mesh\_quality\_histogram.gnu}.
-
-\chapter{\label{cha:Creating-Distributed-Databases}Creating the Distributed
-Databases}
-
-After using \texttt{xmeshfem3D} or \texttt{xdecompose\_mesh}, the
-next step in the workflow is to compile \texttt{xgenerate\_}~\\
- \texttt{databases}. This program is going to create all the missing
-information needed by the SEM solver.
-
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=0.6\textwidth]{figures/workflow}
-\par\end{centering}
-
-\caption{Schematic workflow for a SPECFEM3D Cartesian simulation. The executable
-\texttt{xgenerate\_databases} creates the GLL mesh points and assigns
-specific model parameters.}
-
-\label{fig:workflow.databases}
-\end{figure}
-
-\noindent
-In the main directory, type
-\begin{lyxcode}
-{\small make~xgenerate\_databases}{\small \par}
-\end{lyxcode}
-Input for the program is provided through the main parameter file
-\texttt{Par\_file}, which resides in the subdirectory \texttt{DATA}.
-Please note that \texttt{xgenerate\_databases} must be called directly
-from the main directory, as most of the binaries of the package.
-
-\section{\label{cha:Main-Parameter}Main parameter file \texttt{Par\_file}}
-
-Before running \texttt{xgenerate\_databases}, a number of parameters
-need to be set in the main parameter \texttt{Par\_file} located in
-the subdirectory \texttt{DATA}:
-\begin{description}
-\item [{\texttt{SIMULATION\_TYPE}}] is set to 1 for forward simulations,
-2 for adjoint simulations (see Section \ref{sec:Adjoint-simulation-finite})
-and 3 for kernel simulations (see Section \ref{sec:Finite-Frequency-Kernels}).
-\item [{\texttt{SAVE\_FORWARD}}] is only set to \texttt{.true.} for a forward
-simulation with the last frame of the simulation saved, as part of
-the finite-frequency kernel calculations (see Section \ref{sec:Finite-Frequency-Kernels}).
-For a regular forward simulation, leave \texttt{SIMULATION\_TYPE}
-and \texttt{SAVE\_FORWARD} at their default values.
-\item [{\texttt{\noun{UTM\_PROJECTION\_ZONE}}}] UTM projection zone in
-which your model resides, only valid when \texttt{SUPPRESS\_UTM\_}~\\
- \texttt{PROJECTION} is \texttt{.false.}.
-\item [{\texttt{SUPPRESS\_UTM\_PROJECTION}}] set to be \texttt{.false.}
-when your model range is specified in the geographical coordinates,
-and needs to be \texttt{.true.} when your model is specified in a
-cartesian coordinates. \noun{UTM projection zone in which your simulation
-region resides.}
-\item [{\texttt{NPROC}}] The number of MPI processors, each one is assigned
-one slice of the whole mesh.
-\item [{\texttt{NSTEP}}] The number of time steps of the simulation. This
-controls the length of the numerical simulation, i.e., twice the number
-of time steps requires twice as much CPU time. This feature is not
-used at the time of generating the distributed databases but is required
-for the solver, i.e., you may change this parameter after running
-\texttt{xgenerate\_databases}.
-\item [{\texttt{DT}}] The length of each time step in seconds. This feature
-is not used at the time of generating the distributed databases but
-is required for the solver. Please see also Section~\ref{sec:Choosing-the-Time-Step}
-for further details.
-\item [{\texttt{NGNOD}}] The number of nodes for 2D and 3D shape functions
-for hexahedra. We use either 8-node mesh elements (bricks) or 27-node
-elements. If you use the internal mesher, the only option is 8-node
-bricks (27-node elements are not supported). \texttt{CUBIT} does not
-support HEX27 elements either (it can generate them, but they are
-flat, i.e. identical to HEX8). To generate HEX27 elements with curvature
-properly taken into account, you can use \texttt{Gmsh} \url{http://geuz.org/gmsh/}
-\item [{\texttt{MODEL}}] Must be set to one of the following:
-
-\begin{description}
-\item [{\textmd{Models defined by mesh parameters:}}] ~
-
-\begin{description}
-\item [{\texttt{default}}] Uses model parameters as defined by meshing
-procedures described in the previous Chapter~\ref{cha:Mesh-Generation}.
-\end{description}
-
-\item [{\textmd{1D~models~with~real~structure:}}] ~
-
-\begin{description}
-\item [{\texttt{1D\_prem}}] Isotropic version of the spherically symmetric
-Preliminary Reference Earth Model (PREM) \citep{DzAn81}.
-\item [{\texttt{1D\_socal}}] A standard isotropic 1D model for Southern
-California.
-\item [{\texttt{1D\_cascadia}}] Isotropic 1D profile for the Cascadia region.
-\end{description}
-
-\item [{\textmd{Fully~3D~models:}}] ~
-
-\begin{description}
-\item [{\texttt{aniso}}] For a user-specified fully anisotropic model.
-Parameters are set up in routines located in file \texttt{model\_aniso.f90}
-in directory \texttt{src/generate\_databases/}. See Chapter~\ref{cha:-Changing-the}
-for a discussion on how to specify your own 3D model.
-
-\item [{\texttt{external}}] For a user-specified isotropic model which
-uses externally defined model parameters. Uses external model definitions
-set up in routines located in file \texttt{model\_external\_values.f90}
-in directory \texttt{src/generate\_databases/}. Please modify these
-generic template routines to use your own model definitions.
-\item [{\texttt{gll}}] For a user-specified isotropic model which uses
-external binary files for $v_{p}$, $v_{s}$ and $\rho$. Binary files
-are given in the same format as when outputted by the \texttt{xgenerate\_databases}
-executable when using option \texttt{SAVE\_MESH\_FILES}. These binary
-files define the model parameters on all GLL points which can be used
-for iterative inversion procedures.
-\item [{\texttt{salton\_trough}}] A 3D $V_{p}$ model for Southern California.
-Users must provide the corresponding data file \texttt{regrid3\_vel\_p.bin}
-in directory \texttt{DATA/st\_3D\_block\_harvard/}.
-\item [{\texttt{tomo}}] For a user-specified 3D isotropic model which uses
-a tomographic model file \texttt{tomographic\_model.xyz} in directory
-\texttt{DATA}. See Section \ref{sec:Using-tomographic}, for a discussion
-on how to specify your own 3D tomographic model.
-\end{description}
-\end{description}
-\item [{\texttt{APPROXIMATE\_OCEAN\_LOAD}}] Set to \texttt{.true.} if the
-effect of the oceans on seismic wave propagation should be incorporated
-based upon the (rough) approximate treatment discussed in \citet{KoTr02b}.
-This feature is inexpensive from a numerical perspective, both in
-terms of memory requirements and CPU time. This approximation is accurate
-at periods of roughly 20~s and longer. At shorter periods the effect
-of water phases/reverberations is not taken into account, even when
-the flag is on. If you want to model the effect of a fluid-solid model
-at short periods, then set this flag to \texttt{.false.} and mesh
-the fluid layer explicitly in your mesher, so that it is computed
-accurately and without this approximation.
-\item [{\texttt{TOPOGRAPHY}}] This feature is only effective if \texttt{APPROXIMATE\_OCEAN\_LOAD}
-is set to \texttt{.true.}. Set to \texttt{.true.} if topography and
-bathymetry should be read in based upon the topography file specified
-in the main constants file \texttt{constants.h} found in subdirectory
-\texttt{src/shared/} to evaluate elevations. If not set, elevations
-will be read from the numerical mesh.
-\item [{\texttt{ATTENUATION}}] Set to \texttt{.true.} if attenuation should
-be incorporated. Turning this feature on increases the memory requirements
-significantly (roughly by a factor of~1.5), and is numerically fairly
-expensive. See \citet{KoTr99,KoTr02a} for a discussion on the implementation
-of attenuation based upon standard linear solids. Please note that
-the Vp- and Vs-velocities of your model are given for a reference
-frequency. To change this reference frequency, you change the value
-of \texttt{ATTENUATION\_f0\_REFERENCE} in the main constants file
-\texttt{constants.h} found in subdirectory \texttt{src/shared/}.
-\item [{\texttt{FULL\_ATTENUATION\_SOLID}}] Set to \texttt{.true.} to add
-$Q_{\kappa}$ attenuation to the simulations in addition to $Q_{\mu}$
-attenuation when the \texttt{ATTENUATION} flag above is on. Should
-be off in almost all cases for seismic wave propagation, since $Q_{\kappa}$
-is almost always negligible in Earth models. However, this parameter
-can be useful for instance for ocean acoustics simulations, when $Q_{\kappa}$
-is not negligible in the solid layer at the sea bottom.
-\item [{\texttt{ANISOTROPY}}] Set to \texttt{.true.} if you want to use
-an anisotropy model. Please see the file \texttt{model\_aniso.f90}
-in subdirectory \texttt{src/generate\_databases/} for the current
-implementation of anisotropic models.
-\item [{\texttt{TOMOGRAPHY\_PATH}}] Directory in which the tomography files
-are stored for using external tomographic Earth models (please read
-Chapter \ref{cha:-Changing-the} and Section \ref{sec:Using-tomographic}
-`Using external tomographic Earth models' for further details.).
-\item [{\texttt{USE\_OLSEN\_ATTENUATION}}] Set to \texttt{.true.} if you
-want to use the attenuation model that scaled from the S-wave speed
-model using Olsen's empirical relation (see \citet{OlDaBr03}).
-\item [{\texttt{OLSEN\_ATTENUATION\_RATIO}}] Determines the Olsen's constant
-in Olsen's empirical relation (see \citet{OlDaBr03}).
-\item [{\texttt{PML\_CONDITIONS}}] Set to \texttt{.true.} to turn on C-PML
-boundary conditions for a regional simulation. Both fluids and elastic
-solids are supported. Note that xmeshfem3d generated meshes do not support C-PML yet.
-\item [{\texttt{PML\_INSTEAD\_OF\_FREE\_SURFACE}}] Set to \texttt{.true.}
-to turn on C-PML boundary conditions on the top surface instead of
-the usual free surface.
-\item [{\texttt{f0\_FOR\_PML}}] Determines the dominant frequency that
-will be used in the calculation of PML damping profiles; this should
-be set to the same (or similar) dominant frequency as that of the
-source that you will use in your simulation. If you plan to use a
-Dirac source, then use the dominant frequency of the source wavelet
-with which you plan to convolve your seismograms later on in post-processing.
-\item [{\texttt{STACEY\_ABSORBING\_CONDITIONS}}] Set to \texttt{.true.}
-to turn on Clayton-Enquist absorbing boundary conditions (see \citet{KoTr99}).
-In almost all cases it is much better to use CPML absorbing layers
-(see the options above) and leave this flag to \texttt{.false.}.
-\item [{\texttt{STACEY\_INSTEAD\_OF\_FREE\_SURFACE}}] Set to \texttt{.true.}
-to turn on absorbing boundary conditions on the top surface which
-by default constitutes a free surface of the model.
-\item [{\texttt{CREATE\_SHAKEMAP}}] Set this flag to \texttt{.true.} to
-create a ShakeMap\textregistered{}, i.e., a peak ground velocity map
-of the maximum absolute value of the two horizontal components of
-the velocity vector.
-\item [{\texttt{MOVIE\_SURFACE}}] Set to \texttt{.false.}, unless you want
-to create a movie of seismic wave propagation on the Earth's surface.
-Turning this option on generates large output files. See Section~\ref{sec:Movies}
-for a discussion on the generation of movies. This feature is only
-relevant for the solver.
-\item [{\texttt{MOVIE\_TYPE}}] Set this flag to 1 to show the top surface
-(tomography + oceans) only, to 2 to show all external faces of the
-mesh (i.e. topography + vertical edges + bottom) in shakemaps and
-surface movies.
-\item [{\texttt{MOVIE\_VOLUME}}] Set to \texttt{.false.}, unless you want
-to create a movie of seismic wave propagation in the Earth's interior.
-Turning this option on generates huge output files. See Section~\ref{sec:Movies}
-for a discussion on the generation of movies. This feature is only
-relevant for the solver.
-\item [{\texttt{SAVE\_DISPLACEMENT}}] Set this flag to \texttt{.true.}
-if you want to save the displacement instead of velocity for the movie
-frames.
-\item [{\texttt{USE\_HIGHRES\_FOR\_MOVIES}}] Set this flag to \texttt{.true.}
-if you want to save the values at all the NGLL grid points for the
-movie frames.
-\item [{\texttt{NTSTEP\_BETWEEN\_FRAMES}}] Determines the number of timesteps
-between movie frames. Typically you want to save a snapshot every
-100 timesteps. The smaller you make this number the more output will
-be generated! See Section~\ref{sec:Movies} for a discussion on the
-generation of movies. This feature is only relevant for the solver.
-\item [{\texttt{HDUR\_MOVIE}}] Determines the half duration of the source
-time function for the movie simulations. When this parameter is set
-to be 0, a default half duration that corresponds to the accuracy
-of the simulation is provided. Otherwise, it adds this half duration
-to the half duration specified in the source file \texttt{CMTSOLUTION},
-thus simulates longer periods to make the movie images look smoother.
-\item [{\texttt{SAVE\_MESH\_FILES}}] Set this flag to \texttt{.true.} to
-save ParaView \urlwithparentheses{www.paraview.org} mesh files for
-subsequent viewing. Turning the flag on generates large (distributed)
-files in the \texttt{LOCAL\_PATH} directory. See Section~\ref{sec:Mesh-graphics}
-for a discussion of mesh viewing features.
-\item [{\texttt{LOCAL\_PATH}}] Directory in which the distributed databases
-will be written. Generally one uses a directory on the local disk
-of the compute nodes, although on some machines these databases are
-written on a parallel (global) file system (see also the earlier discussion
-of the \texttt{LOCAL\_PATH\_IS\_ALSO\_GLOBAL} flag in Chapter~\ref{cha:Getting-Started}).
-\texttt{xgenerate\_databases} generates the necessary databases in
-parallel, one set for each of the \texttt{NPROC} slices that constitutes
-the mesh (see Figure~\ref{fig:mount.partitions} and Figure~\ref{fig:For-parallel-computing}).
-After the executable finishes, you can log in to one of the compute
-nodes and view the contents of the \texttt{LOCAL\_PATH} directory
-to see the (many) files generated by \texttt{xgenerate\_databases}.
-Please note that the \texttt{LOCAL\_PATH} directory should already
-contain the output files of the partitioner, i.e. from \texttt{xdecompose\_mesh}
-or \texttt{xmeshfem3D}.
-\item [{\texttt{NTSTEP\_BETWEEN\_OUTPUT\_INFO}}] This parameter specifies
-the interval at which basic information about a run is written to
-the file system (\texttt{timestamp{*}} files in the \texttt{OUTPUT\_FILES}
-directory). If you have access to a fast machine, set \texttt{NTSTEP\_BETWEEN\_OUTPUT\_INFO}
-to a relatively high value (e.g., at least 100, or even 1000 or more)
-to avoid writing output text files too often. This feature is not
-used at the time of meshing. One can set this parameter to a larger
-value than the number of time steps to avoid writing output during
-the run.
-\item [{\texttt{NTSTEP\_BETWEEN\_OUTPUT\_SEISMOS}}] This parameter specifies
-the interval at which synthetic seismograms are written in the \texttt{LOCAL\_PATH}
-directory. If a run crashes, you may still find usable (but shorter
-than requested) seismograms in this directory. On a fast machine set
-\texttt{NTSTEP\_BETWEEN\_OUTPUT\_SEISMOS} to a relatively high value
-to avoid writing to the seismograms too often. This feature is only
-relevant for the solver.
-\item [{\texttt{USE\_FORCE\_POINT\_SOURCE}}] Turn this flag on to use a
-(tilted) \texttt{FORCESOLUTION} force point source instead of a \texttt{CMTSOLUTION}
-moment-tensor source. When the force source does not fall exactly
-at a grid point, the solver interpolates the force between grid points
-using Lagrange interpolants. This can be useful e.g. for oil industry
-foothills simulations in which the source is a vertical force, normal
-force, tilted force, or an impact etc. Note that in the \texttt{FORCESOLUTION}
-file, you will need to edit the East, North and vertical components
-of an arbitrary (not necessarily unitary, the code will normalize it automatically) direction vector of the force vector;
-thus refer to Appendix~\ref{cha:Coordinates} for the orientation
-of the reference frame. This vector is made unitary internally in
-the solver and thus only its direction matters here; its norm is ignored
-and the norm of the force used is the factor force source times the
-source time function.
-When using this option, by default the code can locate the force source
-anywhere between mesh points in order to honor its exact location; this
-is more precise than using the closest GLL mesh point, but it is also a bit slower.
-If needed, you can change that default behavior and force the code to use the closest
-GLL mesh point instead by setting flag \texttt{USE\_BEST\_LOCATION} to \texttt{.false.} instead of \texttt{.true.} in file
-\texttt{src/shared/constants.h.in} and running the \texttt{configure} script again and recompiling the code.
-\item [{\texttt{USE\_RICKER\_TIME\_FUNCTION}}] Turn this flag on to use
-a Ricker source time function instead of the source time functions
-set by default to represent a (tilted) \texttt{FORCESOLUTION} force
-point source or a \texttt{CMTSOLUTION} moment-tensor source. Originally,
-if a \texttt{CMTSOLUTION} moment-tensor source is used, a (pseudo)
-Heaviside step function with a very short half duration is defined
-for elastic cases to represent the permanent slip on the fault while
-in the acoustic case a Gaussian source time function with a similarly
-short half duration is defined to physically describe actions within
-the fluid. Otherwise, if a \texttt{FORCESOLUTION} force source is
-used, a (pseudo) Dirac delta source time function is defined by default.
-Any other source-time function may then be obtained by convolution.
-\item [{\texttt{PRINT\_SOURCE\_TIME\_FUNCTION}}] Turn this flag on to print
-information about the source time function in the file \texttt{OUTPUT\_FILES/plot\_source\_time\_function.txt}.
-This feature is only relevant for the solver.
-
-\item [{\texttt{NUMBER\_OF\_SIMULTANEOUS\_RUNS}}] adds the ability to run several calculations (several earthquakes)
-in an embarrassingly-parallel fashion from within the same run;
-this can be useful when using a very large supercomputer to compute
-many earthquakes in a catalog, in which case it can be better from
-a batch job submission point of view to start fewer and much larger jobs,
-each of them computing several earthquakes in parallel.
-To turn that option on, set parameter NUMBER\_OF\_SIMULTANEOUS\_RUNS to a value greater than 1.
-To implement that, we create NUMBER\_OF\_SIMULTANEOUS\_RUNS MPI sub-communicators,
-each of them being labeled "my\_local\_mpi\_comm\_world", and we use them
-in all the routines in "src/shared/parallel.f90", except in MPI\_ABORT() because in that case
-we need to kill the entire run.
-When that option is on, of course the number of processor cores used to start
-the code in the batch system must be a multiple of NUMBER\_OF\_SIMULTANEOUS\_RUNS,
-all the individual runs must use the same number of processor cores,
-which as usual is NPROC in the Par\_file,
-and thus the total number of processor cores to request from the batch system
-should be NUMBER\_OF\_SIMULTANEOUS\_RUNS * NPROC.
-All the runs to perform must be placed in directories called run0001, run0002, run0003 and so on
-(with exactly four digits).
-
-Imagine you have 10 independent calculations to do, each of them on 100 cores; you have three options:
-
-1/ submit 10 jobs to the batch system
-
-2/ submit a single job on 1000 cores to the batch, and in that script create a sub-array of jobs to start 10 jobs,
-each running on 100 cores (see e.g. http://www.schedmd.com/slurmdocs/job\_array.html )
-
-3/ submit a single job on 1000 cores to the batch, start SPECFEM3D on 1000 cores, create 10 sub-communicators,
-cd into one of 10 subdirectories (called e.g. run0001, run0002,... run0010) depending on the sub-communicator
-your MPI rank belongs to, and run normally on 100 cores using that sub-communicator.
-
-The option NUMBER\_OF\_SIMULTANEOUS\_RUNS implements 3/.
-
-\item [{\texttt{BROADCAST\_SAME\_MESH\_AND\_MODEL}}]:
-if we perform simultaneous runs in parallel, if only the source and receivers vary between these runs
-but not the mesh nor the model (velocity and density) then we can also read the mesh and model files
-from a single run in the beginning and broadcast them to all the others; for a large number of simultaneous
-runs for instance when solving inverse problems iteratively this can DRASTICALLY reduce I/Os to disk in the solver
-(by a factor equal to NUMBER\_OF\_SIMULTANEOUS\_RUNS), and reducing I/Os is crucial in the case of huge runs.
-Thus, always set this option to .true. if the mesh and the model are the same for all simultaneous runs.
-In that case there is no need to duplicate the mesh and model file database (the content of the DATABASES\_MPI
-directories) in each of the run0001, run0002,... directories, it is sufficient to have one in run0001
-and the code will broadcast it to the others).
-
-TODO / future work to do: currently the BROADCAST\_SAME\_MESH\_AND\_MODEL option assumes to have the (master) mesh files in run0001/DATABASES\_MPI or run0001/OUTPUT\_FILES/DATABASES\_MPI.
-However, for adjoint runs you still need a DATABASES\_MPI folder in each of the sub-runs directories, e.g.
-run0002/DATABASES\_MPI, etc.
-to store the forward wavefields, kernels etc.
-of each sub-run.
-This would not be needed for forward simulations.
-
-TODO / future work to do: the sensitivity kernel summing and smoothing tools in directory src/tomography are currently not ported to this new option to do many runs simultaneously, only the solver (src/specfem3d) is.
-Thus these tools should work, but in their current version will need to be run for each simulation result independently.
-More precisely, the current kernel summing and smoothing routines work fine, with the exception that you need to move out the mesh files (and also the parameters).
-This works because these routines consider multiple runs by design.
-You simply have to provide them the directories where the kernels are.
-
-\item [{\texttt{GPU\_MODE}}] Turn this flag on to use GPUs.
-
-\item [\texttt{ADIOS\_ENABLED}] Turn this flag on to enable ADIOS. If set to \texttt{.false.}, subsequent ADIOS
-parameters will not be considered.
-\item [\texttt{ADIOS\_FOR\_DATABASES}] Turn this flag on to use ADIOS for xmeshfem3D output and
-xgenerate\_database input.
-\item [\texttt{ADIOS\_FOR\_MESH}]  Turn this flag on to use ADIOS for generated databases.
-\item [\texttt{ADIOS\_FOR\_FORWARD\_ARRAYS}] Turn this flag on to read and write forward arrays using ADIOS.
-\item [\texttt{ADIOS\_FOR\_KERNELS}] Turn this flag on to produce ADIOS kernels that can later be visualized with the ADIOS version of combine\_vol\_data.
-
-\end{description}
-
-The present version of SPECFEM can handle fully saturated porous simulations, Christina Morency implemented Biot equation.
-But the code cannot calculate partially saturated cases in its state.
-Christina Morency is presently working on a dual porosity, dual permeability formulation, type Pride and Berryman, for an other project, but it will not be available for some time.\\
-
-The way we prescribe material property for porous material in SPECFEM3D is as follow:
-We use a file name nummaterial\_poroelastic\_file, which is located in the directory MESH, the format is as follow:\\
- rhos rhof phi c kxx kxy kxz kyy kyz kzz Ks Kf Kfr etaf mufr\\
-where\\
-rho\_s = solid density\\
-rho\_f = fluid density\\
-phi = porosity\\
-tort = tortuosity\\
-kxx = xx component of permeability tensor\\
-kxy = xy,yx components of permeability tensor\\
-kyy = yy component of permeability tensor\\
-kxz = xz,zx components of permeability tensor\\
-kzz = zz component of permeability tensor\\
-kappa\_s = solid bulk modulus\\
-kappa\_f= fluid bulk modulus\\
-kappa\_fr= frame bulk modulus\\
-eta\_f = fluid viscosity\\
-mu\_fr = frame shear modulus.\\
-
-Using an external mesh (for instance coming from CUBIT/TRELIS), poroelastic materials have the ID number 3, while 1 is elastic, 2 is acoustic ?
-see the example in the package: examples/homogeneous\_poroelastic.
-
-If you use PML, the mesh elements that belong to the PML layers can
-be acoustic or elastic, but not viscoelastic nor poroelastic. Then,
-when defining your model, you should define these absorbing elements
-as either acoustic or elastic. In you forget to do that, the code
-will fix the problem by automatically converting the viscoelastic
-or poroelastic PML elements to elastic. This means that strictly speaking
-the PML layer will not be perfectly matched any more, since the physical
-model will change from viscoelastic or poroelastic to elastic at the
-entrance of the PML, but in practice this is sufficient and produces
-only tiny / negligible spurious reflections.
-
-If you use PML and an external mesh (created using an external meshing tool
-such as CUBIT/TRELIS or similar), try to have elements inside the PML as regular as possible,
-i.e. ideally non-deformed cubes obtained by `extrusion' of regular surface mesh elements meshing the
-outer edges of the computational domain without PML; by doing so, the PMLs obtained will be far more stable
-in time (PML being weakly unstable from a mathematical point of view, very deformed mesh elements
-inside the PMLs can trigger instabilities much more quickly).
-
-If you use PML and an external tomographic velocity and density model,
-you should be careful because mathematically a PML cannot handle heterogeneities
-along the normal to the PML edge inside the PML layer. This comes
-from the fact that the damping profile that is defined assumes a constant
-velocity and density model along the normal direction.
-
-Thus, you need to modify your velocity and density model in order
-for it to be 1D inside the PML, as shown in Figure~\ref{fig:modify_external_velocity_model_to_use_PML}.
-
-This applies to the bottom layer as well; there you should make sure
-that your model is 1D and thus constant along the vertical direction.
-
-To summarize, only use a 3D velocity and density model inside the
-physical region, and in all the PML layers extend it by continuity
-from its values along the inner PML edge.
-
-%%
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=3in]{figures/how_to_use_PML_when_a_tomographic_velocity_model_is_used}
-\par\end{centering}
-
-\caption{How to modify your external 3D velocity and density model in order
-to use PML. Such a modification is not needed when using Stacey absorbing
-boundary conditions (but such conditions are significantly less efficient).}
-
-\label{fig:modify_external_velocity_model_to_use_PML}
-\end{figure}
-
-%%
-
-\section{\label{sec:Choosing-the-Time-Step}Choosing the time step \texttt{DT}}
-
-The parameter \texttt{DT} sets the length of each time step in seconds.
-The value of this parameter is crucial for the stability of the spectral-element
-simulation. Your time step \texttt{DT} will depend on the minimum
-ratio between the distance $h$ of neighboring mesh points and the
-wave speeds $v$ defined in your model. The condition for the time
-step $\Delta t$ is:
-\begin{lyxcode}
-$\Delta t<C~\mathrm{min}_{\Omega}(~h/v~)$
-\end{lyxcode}
-where $C$ is the so-called Courant number and $\Omega$ denotes the
-model volume. The distance $h$ depends on the mesh element size and
-the number of GLL points \texttt{NGLL} specified in the main constants
-file \texttt{constants.h} located in the \texttt{src/shared/} subdirectory.
-The wave speed $v$ is determined based on your model's P- (or S-)
-wave speed values.
-
-The database generator \texttt{xgenerate\_databases}, as well as the
-internal mesher \texttt{xmeshfem3D}, are trying to evaluate the value
-of $\Delta t$ for empirically chosen Courant numbers $C\sim0.3$.
-If you used the mesher \texttt{xmeshfem3D} to generate your mesh,
-you should set the value suggested in \texttt{OUTPUT\_FILES/output\_mesher.txt}
-file, which is created after the mesher completed. In case you used
-CUBIT to create the mesh, you might use an arbitrary value when running
-\texttt{xgenerate\_databases} and then use the value suggested in
-the ~\\
- \texttt{OUTPUT\_FILES/output\_mesher.txt} file after the database
-generation completed. Note that the implemented Newmark time scheme
-uses this time step globally, thus your simulations become more expensive
-for very small mesh elements in high wave-speed regions. Please be
-aware of this restriction when constructing your mesh in Chapter~\ref{cha:Mesh-Generation}.
-
-\chapter{\label{cha:Running-the-Solver}Running the Solver \texttt{xspecfem3D}}
-
-Now that you have successfully generated the databases, you are ready
-to compile the solver. In the main directory, type
-\begin{lyxcode}
-{\small make~xspecfem3D}{\small \par}
-\end{lyxcode}
-Please note that \texttt{xspecfem3D} must be called directly from
-the main directory, as most of the binaries of the package.\\
-
-\noindent
-The solver needs three input files in the \texttt{DATA} directory
-to run:
-\begin{description}
-\item [{\texttt{Par\_file}}] the main parameter file which was discussed
-in detail in the previous Chapter~\ref{cha:Creating-Distributed-Databases},
-\item [{\texttt{CMTSOLUTION}} or {\texttt{FORCESOLUTION}}] the earthquake
-source parameter file or the force source parameter file, and
-\item [{\texttt{STATIONS}}] the stations file.
-\end{description}
-Most parameters in the \texttt{Par\_file} should be set prior to running
-the databases generation. Only the following parameters may be changed
-after running \texttt{xgenerate\_databases}:
-\begin{itemize}
-\item the simulation type control parameters: \texttt{SIMULATION\_TYPE}
-and \texttt{SAVE\_FORWARD}
-\item the time step parameters \texttt{NSTEP} and \texttt{DT}
-\item the absorbing boundary control parameter \texttt{PML\_CONDITIONS}
-on condition that the\\
- \texttt{PML\_INSTEAD\_OF\_FREE\_SURFACE} flag remains unmodified
-after running the databases generation.
-\item the movie control parameters \texttt{MOVIE\_SURFACE}, \texttt{MOVIE\_VOLUME},
-and \texttt{NTSTEPS\_BETWEEN\_FRAMES}
-\item the ShakeMap\textregistered{}option \texttt{CREATE\_SHAKEMAP}
-\item the output information parameters \texttt{MOVIE\_TYPE}, \texttt{NTSTEP\_BETWEEN\_OUTPUT\_INFO}
-and \texttt{NTSTEP\_BETWEEN\_OUTPUT\_}~\\
- \texttt{SEISMOS}
-\item the \texttt{PRINT\_SOURCE\_TIME\_FUNCTION} flags
-\end{itemize}
-Any other change to the \texttt{Par\_file} implies rerunning both
-the database generator \texttt{xgenerate\_databases} and the solver
-\texttt{xspecfem3D}.
-
-For any particular earthquake, the \texttt{CMTSOLUTION} file that
-represents the point source may be obtained directly from the Harvard
-Centroid-Moment Tensor (CMT) web page \urlwithparentheses{www.seismology.harvard.edu}.
-It looks like the example shown in Fig. \ref{fig:CMTSOLUTION-file}.
-%\begin{lyxcode}
-\begin{figure}[htp]
-\noindent
-\begin{centering}
-{\small \includegraphics[width=0.8\textwidth]{figures/Hollywood_CMT}
-}
-\par
-\end{centering}
-\caption{\texttt{CMTSOLUTION} file based on the
-format from the Harvard CMT catalog. \textbf{M} is the moment tensor,
-$M_{0}${\small {} }is the seismic moment, and $M_{w}$ is the moment
-magnitude.}
-\label{fig:CMTSOLUTION-file}
-\end{figure}
-%\end{lyxcode}
-The \texttt{CMTSOLUTION} file should be edited in the following way:
-\begin{itemize}
-\item Set the latitude or UTM $x$ coordinate, longitude or UTM $y$ coordinate, depth of the source (in km).
-\item Set the \texttt{time shift} parameter equal to $0.0$ (the solver
-will not run otherwise.) The time shift parameter would simply apply
-an overall time shift to the synthetics, something that can be done
-in the post-processing (see Section \ref{sec:Process-data-and-syn}).
-\item For point-source simulations (see finite sources, page \pageref{To-simulate-a})
-we recommend setting the source half-duration parameter \texttt{half
-duration} equal to zero, which corresponds to simulating a step source-time
-function, i.e., a moment-rate function that is a delta function. If
-\texttt{half duration} is not set to zero, the code will use a Gaussian
-(i.e., a signal with a shape similar to a `smoothed triangle', as
-explained in \citet{KoTr02a} and shown in Fig~\ref{fig:gauss.vs.triangle})
-source-time function with half-width \texttt{half duration}. We prefer
-to run the solver with \texttt{half duration} set to zero and convolve
-the resulting synthetic seismograms in post-processing after the run,
-because this way it is easy to use a variety of source-time functions
-(see Section \ref{sec:Process-data-and-syn}). \citet{KoTr02a} determined
-that the noise generated in the simulation by using a step source
-time function may be safely filtered out afterward based upon a convolution
-with the desired source time function and/or low-pass filtering. Use
-the serial code \texttt{convolve\_source\_timefunction.f90} and the
-script \texttt{convolve\_source\_timefunction.csh} for this purpose,
-or alternatively use signal-processing software packages such as SAC
-\urlwithparentheses{www.llnl.gov/sac}. Type
-
-\begin{lyxcode}
-{\small make~xconvolve\_source\_timefunction}{\small \par}
-\end{lyxcode}
-
-to compile the code and then set the parameter \texttt{hdur} in \texttt{convolve\_source\_timefunction.csh}
-to the desired half-duration.
-
-\item The zero time of the simulation corresponds to the center of the triangle/Gaussian,
-or the centroid time of the earthquake. The start time of the simulation
-is $t=-1.5*\texttt{half duration}$ (the 1.5 is to make sure the moment
-rate function is very close to zero when starting the simulation).
-To convert to absolute time $t_{\mathrm{abs}}$, set
-
-\begin{lyxcode}
-$t_{\mathrm{abs}}=t_{\mathrm{pde}}+\texttt{time shift}+t_{\mathrm{synthetic}}$
-\end{lyxcode}
-
-where $t_{\mathrm{pde}}$ is the time given in the first line of the
-\texttt{CMTSOLUTION}, \texttt{time shift} is the corresponding value
-from the original \texttt{CMTSOLUTION} file and $t_{\mathrm{synthetic}}$
-is the time in the first column of the output seismogram.
-
-\end{itemize}
-\begin{figure}[htp]
-\noindent \begin{centering}
-\includegraphics[width=3.5in]{figures/gauss_vs_triangle_mod}
-\par\end{centering}
-
-\caption{Comparison of the shape of a triangle and the Gaussian function actually
-used.}
-
-\label{fig:gauss.vs.triangle}
-\end{figure}
-
-If you know the earthquake source in strike/dip/rake format rather
-than in \texttt{CMTSOLUTION} format, use the C code \texttt{SPECFEM3D\_GLOBE/utils/strike\_dip\_rake\_to\_CMTSOLUTION.c}
-to convert it. The conversion formulas are given for instance in \citet{AkRi80}.
-Note that the \citet{AkRi80} convention is slightly different from
-the Harvard \texttt{CMTSOLUTION} convention (the sign of some components
-is different). The C code outputs both.
-
-Centroid latitude and longitude should be provided in geographical
-coordinates. The code converts these coordinates to geocentric coordinates~\citep{DaTr98}.
-Of course you may provide your own source representations by designing
-your own \texttt{CMTSOLUTION} file. Just make sure that the resulting
-file adheres to the Harvard CMT conventions (see Appendix~\ref{cha:Coordinates}).
-Note that the first line in the \texttt{CMTSOLUTION} file is the Preliminary
-Determination of Earthquakes (PDE) solution performed by the USGS
-NEIC, which is used as a seed for the Harvard CMT inversion. The PDE
-solution is based upon P waves and often gives the hypocenter of the
-earthquake, i.e., the rupture initiation point, whereas the CMT solution
-gives the `centroid location', which is the location with dominant
-moment release. The PDE solution is not used by our software package
-but must be present anyway in the first line of the file.
-
-\label{To-simulate-a}To simulate a kinematic rupture, i.e., a finite-source
-event, represented in terms of $N_{\mathrm{sources}}$ point sources,
-provide a \texttt{CMTSOLUTION} file that has $N_{\mathrm{sources}}$
-entries, one for each subevent (i.e., concatenate $N_{\mathrm{sources}}$
-\texttt{CMTSOLUTION} files to a single \texttt{CMTSOLUTION} file).
-At least one entry (not necessarily the first) must have a zero \texttt{time
-shift}, and all the other entries must have non-negative \texttt{time
-shift}. Each subevent can have its own half duration, latitude, longitude,
-depth, and moment tensor (effectively, the local moment-density tensor).
-
-Note that the zero in the synthetics does NOT represent the hypocentral
-time or centroid time in general, but the timing of the \textit{center}
-of the source triangle with zero \texttt{time shift} (Fig~\ref{fig:source_timing}).
-
-Although it is convenient to think of each source as a triangle, in
-the simulation they are actually Gaussians (as they have better frequency
-characteristics). The relationship between the triangle and the gaussian
-used is shown in Fig~\ref{fig:gauss.vs.triangle}. For finite fault
-simulations it is usually not advisable to use a zero half duration
-and convolve afterwards, since the half duration is generally fixed
-by the finite fault model.
-
-\vspace{1cm}
-
-\noindent The \texttt{FORCESOLUTION} file should be edited in the
-following way:
-\begin{itemize}
-\item Set the \texttt{time shift} parameter equal to $0.0$ (the solver
-will not run otherwise.) The time shift parameter would simply apply
-an overall time shift to the synthetics, something that can be done
-in the post-processing (see Section \ref{sec:Process-data-and-syn}).
-\item Set the \texttt{f0} parameter (the dominant frequency) of the Ricker source time
-function when {\texttt{USE\_RICKER\_TIME\_FUNCTION}} is turned on
-in the main parameter file {\texttt{Par\_file}}. In case that the
-solver uses a (pseudo) Dirac delta source time function to represent
-a force point source, a very short duration of five time steps is automatically set by default.
-\item Set the latitude or UTM $x$ coordinate, longitude or UTM $y$ coordinate, depth of the source (in km).
-\item Set the magnitude of the force source.
-\item Set the components of a (non-unitary) direction vector for the force
-source in the East/North/Vertical basis (see Appendix A for the orientation
-of the reference frame).
-\end{itemize}
-\noindent Where necessary, set a \texttt{FORCESOLUTION} file in the
-same way you configure a \texttt{CMTSOLUTION} file with $N_{\mathrm{sources}}$
-entries, one for each subevent (i.e., concatenate $N_{\mathrm{sources}}$
-\texttt{FORCESOLUTION} files to a single \texttt{FORCESOLUTION} file).
-At least one entry (not necessarily the first) must have a zero \texttt{time
-shift}, and all the other entries must have non-negative \texttt{time
-shift}. Each subevent can have its own half latitude, longitude, depth,
-\texttt{half duration} and force parameters.
-
-\begin{figure}[H]
-\noindent \begin{centering}
-{\small \includegraphics[width=5in]{figures/source_timing} }
-\par\end{centering}
-
-\caption{Example of timing for three sources. The center of the first source
-triangle is defined to be time zero. Note that this is NOT in general
-the hypocentral time, or the start time of the source (marked as tstart).
-The parameter \texttt{time shift} in the \texttt{CMTSOLUTION} file
-would be t1(=0), t2, t3 in this case, and the parameter \texttt{half
-duration} would be hdur1, hdur2, hdur3 for the sources 1, 2, 3 respectively.}
-
-{\small \label{fig:source_timing} }
-\end{figure}
-
-The solver can calculate seismograms at any number of stations for
-basically the same numerical cost, so the user is encouraged to include
-as many stations as conceivably useful in the \texttt{STATIONS} file,
-which looks like this:
-
-\begin{figure}[H]
-\noindent \begin{centering}
-{\small \includegraphics{figures/STATIONS_basin_explained} }
-\par\end{centering}
-
-\label{fig:Sample-STATIONS-file.}\caption{Sample \texttt{STATIONS} file. Station latitude and longitude should
-be provided in geographical coordinates. The width of the station
-label should be no more than 32 characters (see \texttt{MAX\_LENGTH\_STATION\_NAME}
-in the \texttt{constants.h} file), and the network label should be
-no more than 8 characters (see \texttt{MAX\_LENGTH\_NETWORK\_NAME}
-in the \texttt{constants.h} file).}
-\end{figure}
-
-Each line represents one station in the following format:
-\begin{lyxcode}
-{\small Station~Network~Latitude~(degrees)~Longitude~(degrees)~Elevation~(m)~burial~(m)~}{\small \par}
-\end{lyxcode}
-The solver \texttt{xspecfem3D} filters the list of stations in file
-\texttt{DATA/STATIONS} to exclude stations that are not located within
-the region given in the \texttt{Par\_file} (between \texttt{LATITUDE\_MIN}
-and \texttt{LATITUDE\_MAX} and between \texttt{LONGITUDE\_MIN} and
-\texttt{LONGITUDE\_MAX}). The filtered file is called \texttt{DATA/STATIONS\_FILTERED}.
-
-Solver output is provided in the \texttt{OUTPUT\_FILES} directory
-in the \texttt{output\_solver.txt} file. Output can be directed to
-the screen instead by uncommenting a line in \texttt{constants.h}:
-\begin{lyxcode}
-!~uncomment~this~to~write~messages~to~the~screen~
-
-!~integer,~parameter~::~IMAIN~=~ISTANDARD\_OUTPUT~
-\end{lyxcode}
-On PC clusters the seismogram files are generally written to the local
-disks (the path \texttt{LOCAL\_PATH} in the \texttt{Par\_file}) and
-need to be gathered at the end of the simulation.
-
-While the solver is running, its progress may be tracked by monitoring
-the `\texttt{\small timestamp{*}}{\small ' files in the ~}\\
-{\small{} }\texttt{\small OUTPUT\_FILES}{\small{} directory. These tiny
-files look something like this:}{\small \par}
-\begin{lyxcode}
-Time~step~\#~~~~~~~~~~10000~
-
-Time:~~~~~108.4890~~~~~~seconds~
-
-Elapsed~time~in~seconds~=~~~~1153.28696703911~
-
-Elapsed~time~in~hh:mm:ss~=~~~~~0~h~19~m~13~s~
-
-Mean~elapsed~time~per~time~step~in~seconds~=~~~~~0.115328696703911~
-
-Max~norm~displacement~vector~U~in~all~slices~(m)~=~~~~1.0789589E-02~
-\end{lyxcode}
-The \texttt{\small timestamp{*}}{\small{} files provide the }\texttt{\small Mean
-elapsed time per time step in seconds}{\small , which may be used
-to assess performance on various machines (assuming you are the only
-user on a node), as well as the }\texttt{\small Max norm displacement
-vector U in all slices~(m)}{\small . If something is wrong with the
-model, the mesh, or the source, you will see the code become unstable
-through exponentially growing values of the displacement and fluid
-potential with time, and ultimately the run will be terminated by
-the program. You can control the rate at which the timestamp files
-are written based upon the parameter }\texttt{\small NTSTEP\_BETWEEN\_OUTPUT\_INFO}{\small{}
-in the }\texttt{\small Par\_file}{\small .}{\small \par}
-
-Having set the \texttt{Par\_file} parameters, and having provided
-the \texttt{CMTSOLUTION} (or the \texttt{FORCESOLUTION}) and \texttt{STATIONS}
-files, you are now ready to launch the solver! This is most easily
-accomplished based upon the \texttt{go\_solver} script (See Chapter
-\ref{cha:Scheduler} for information about running through a scheduler,
-e.g., LSF). You may need to edit the last command at the end of the
-script that invokes the \texttt{mpirun} command. The \texttt{runall}
-script compiles and runs both \texttt{xgenerate\_databases} and \texttt{xspecfem3D}
-in sequence. This is a safe approach that ensures using the correct
-combination of distributed database output and solver input.
-
-It is important to realize that the CPU and memory requirements of
-the solver are closely tied to choices about attenuation (\texttt{ATTENUATION})
-and the nature of the model (i.e., isotropic models are cheaper than
-anisotropic models). We encourage you to run a variety of simulations
-with various flags turned on or off to develop a sense for what is
-involved.
-
-For the same model, one can rerun the solver for different events
-by simply changing the \texttt{CMTSOLUTION} or \texttt{FORCESOLUTION}
-file, or for different stations by changing the \texttt{STATIONS}
-file. There is no need to rerun the \texttt{xgenerate\_databases}
-executable. Of course it is best to include as many stations as possible,
-since this does not add to the cost of the simulation.
-
-We have also added the ability to run several calculations (several earthquakes)
-in an embarrassingly-parallel fashion from within the same run;
-this can be useful when using a very large supercomputer to compute
-many earthquakes in a catalog, in which case it can be better from
-a batch job submission point of view to start fewer and much larger jobs,
-each of them computing several earthquakes in parallel.
-To turn that option on, set parameter \texttt{NUMBER\_OF\_SIMULTANEOUS\_RUNS}
-to a value greater than 1 in file setup/constants.h.in before
-configuring and compiling the code.
-When that option is on, of course the number of processor cores used to start
-the code in the batch system must be a multiple of \texttt{NUMBER\_OF\_SIMULTANEOUS\_RUNS},
-all the individual runs must use the same number of processor cores,
-which as usual is \texttt{NPROC} in the input file \texttt{DATA/Par\_file},
-and thus the total number of processor cores to request from the batch system
-should be \texttt{NUMBER\_OF\_SIMULTANEOUS\_RUNS $\times$ NPROC}.
-All the runs to perform must be placed in directories called \texttt{run0001}, \texttt{run0002}, \texttt{run0003} and so on (with exactly four digits)
-and you must create a link from the root directory of the code
-to the first copy of the executable programs by typing \texttt{ln -s run0001/bin bin}.
-
-\chapter{\label{cha:Fault-Kinematics-Dynamics}Kinematic and dynamic fault
-sources}
-
-SPECFEM3D can handle finite fault sources of two kinds:
-\begin{enumerate}
-\item \emph{Kinematic}: the spatio-temporal distribution of slip rate is
-prescribed along the fault surface
-\item \emph{Dynamic}: a friction law and initial stresses are prescribed
-on the fault, a spontaneous rupture process is computed.
-\end{enumerate}
-
-\section{\label{sec:Mesh-Generation-with}Mesh Generation with Split Nodes}
-
-Faults need to be handled in a special way during mesh generation.
-A fault surface must lie at the interface between elements (the mesh
-must honor the fault surfaces). Moreover, a fault is made of two surfaces
-in contact. Each of these two surfaces needs a separate set of nodes.
-This approach is known as \textquotedbl{}split nodes\textquotedbl{}.
-Currently faults can only be run with \texttt{xdecompose\_mesh} and
-\texttt{CUBIT}. \texttt{xmeshfem3D} is not yet ready to handle faults.\\
-To facilitate the mesh generation with split nodes in CUBIT, we need
-to separate the two fault surfaces by a small distance, effectively
-creating a tiny opening of the fault (Figure \ref{fig:examples.splitnodes},
-\ref{fig:examples.splitnodes-surfacetrace}). Note that only the interior
-of the fault must be opened, its edges must remain closed (except
-the edge on the free surface). The fault is automatically closed later
-by SPECFEM3D.\\
-
-Here is an example CUBIT script to generate a mesh with split nodes
-for a buried vertical strike-slip fault:
-\begin{lyxcode}
-~reset~~\\
-~brick~x~10~y~10~z~10~~\\
-~webcut~volume~all~with~plane~xplane~~\\
-~webcut~volume~all~with~plane~yplane~~\\
-~webcut~volume~all~with~plane~xplane~offset~3~~\\
-~webcut~volume~all~with~plane~zplane~offset~3~~\\
-~webcut~volume~all~with~plane~zplane~offset~-3~~\\
-~imprint~all~~\\
-~merge~all~~\\
-~unmerge~surf~160~~\\
-~mesh~vol~all~~\\
-~set~node~constraint~off~~\\
-~node~in~surf~168~move~X~0~Y~0.01~Z~0~~\\
-~node~in~surf~160~move~X~0~Y~-0.01~Z~0~\\
-
-\end{lyxcode}
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[scale=0.3]{figures/faultmesh} \\
- \includegraphics[scale=0.3]{figures/surf168} \includegraphics[scale=0.3]{figures/splitnodes}
-\par\end{centering}
-
-\caption{Screenshots of the CUBIT example for the embedded fault with split
-nodes: The entire mesh is decomposed into several volumes bounding
-the fault (top), the fault surface 168 is shown as orange squares
-(middle) and split nodes of the fault shown as orange dots(bottom).
-Note that only interior nodes of the fault are split while those on
-the edges of the fault surface are touching each other.}
-
-\label{fig:examples.splitnodes}
-\end{figure}
-
-\begin{lyxcode}
-
-\end{lyxcode}
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[scale=0.55]{figures/splitnodes_surfacetrace} \\
-
-\par\end{centering}
-
-\caption{Screenshots of a CUBIT example showing split nodes for a fault reaching
-the surface. Surface trace of the fault is shown in orange. Note that
-the edges of the fault are not split while the interior nodes are
-offset by a small distance on either side of the fault}
-
-\label{fig:examples.splitnodes-surfacetrace}
-\end{figure}
-
-The CUBIT scripts ({*}.jou and {*}.py) in the directory \texttt{EXAMPLES}
-generate more complicated meshes. The {*}.py files are Python scripts
-that execute CUBIT commands and use the CUBIT-python interface for
-SPECFEM3D (see next section). The Python language allows to define
-and manipulate variables to parameterize the mesh. Alternatively,
-the Python script can call a CUBIT journal file ({*}.jou), which looks
-like the example above. Variables can be defined and manipulated there
-using the $\mathtt{APREPRO}$ language built in CUBIT.\\
-
-Note that you should avoid gaps in the list of indices of mesh objects
-with the following CUBIT command:
-\begin{lyxcode}
-$\mathtt{compress\: ids\: hex\: face\: edge\: node}$
-\end{lyxcode}
-(otherwise you will get a segmentation fault during domain decomposition)
-
-\section{CUBIT-Python Scripts for Faults}
-
-The mesh generated in CUBIT needs to be processed and exported in
-a format compatible with SPECFEM3D. This is achieved in the Python
-scripts by calling the Python-CUBIT interface functions defined in
-the CUBIT directory:
-\begin{enumerate}
-\item Function $\mathtt{define\_bc}$ (or $\mathtt{boundary\_definition.py}$)
-must be called to set up the absorbing boundaries database
-\item Function $\mathtt{fault\_input}$ must be called once for each fault
-to set up the fault database
-\item Function $\mathtt{cubit2specfem3d.export2SPECFEM3D}$ must be called
-at the very end of the script to export the mesh in a SPECFEM3D format.
-\end{enumerate}
-The functions in \#1 and \#3 are for the bulk and are documented in
-Section \ref{subsec:Exporting-the-Mesh}. We focus here on \#2:
-\begin{description}
-\item [{Function:}] $\mathtt{fault\_input}$
-\item [{Purpose:}] export a fault mesh database from CUBIT to a SPECFEM3D-compliant
-file
-\item [{Syntax:}] fault\_input(id\_fault, ids\_surf\_1, ids\_surf\_2)
-\item [{Inputs:}]~\end{description}
-\begin{itemize}
-\item id\_fault integer index assigned to the fault. (The user must number
-all the faults, starting at 1, with unit increments)
-\item ids\_surf\_1 list of CUBIT ids of all surfaces that form side 1 of
-the fault.
-\item ids\_surf\_2 list of CUBIT ids of all surfaces that form side 2 of
-the fault. (The user must decide which side of the fault is side 1.
-This choice affects the sign conventions of fault quantities as explained
-in Section \ref{sec:Sign-Convention-for}).\end{itemize}
-\begin{description}
-\item [{Outputs:}] file $\mathtt{fault\_file\_X.dat}$, where X is the
-fault id (id\_fault).
-\item [{Example:}] For the example in Section \ref{sec:Mesh-Generation-with}:\end{description}
-\begin{lyxcode}
-$\hphantom{}$A1~=~{[}168{]}
-
-$\hphantom{}$A2~=~{[}160{]}
-
-$\hphantom{}$fault\_input(1,A1,A2)
-\end{lyxcode}
-
-\section{Examples}
-
-The package includes examples, the SCEC benchmark problems:
-\begin{itemize}
-\item TPV5, a planar vertical strike-slip fault
-\item TPV14 and TPV15, vertical strike-slip fault system with a fault branch
-\item TPV102 and TPV103, rate-and-state benchmarks
-\item TPV22 and TPV23, step-over benchmarks
-\item TPV16, heterogenous initial stress
-\end{itemize}
-and
-\begin{itemize}
-\item Splay fault models from Wendt et al. (2009)\\
-
-\end{itemize}
-Read the documents in the directory \texttt{EXAMPLES/*/description}.
-They contain a description of the example and additional instructions
-to run it. Visualize the results with the Matlab scripts in the directory
-$\mathtt{EXAMPLES/*/post}$
-
-\section{\label{sec:Sign-Convention-for}Sign Convention for Fault Quantities}
-
-During mesh generation, the fault is defined by two surfaces in contact.
-Let's denote as \textquotedbl{}\emph{side 1}\textquotedbl{} the SECOND
-surface declared by the user in the call to the python function \textquotedbl{}\emph{fault\_input}\textquotedbl{},
-and the FIRST surface as \textquotedbl{}\emph{side 2}\textquotedbl{}.
-The local coordinate system on the fault is defined as the right-handed
-coordinate system defined by (strike, dip, normal), where \textquotedbl{}\emph{normal}\textquotedbl{}
-is the normal vector outgoing from side 1, \textquotedbl{}\emph{dip}\textquotedbl{}
-is parallel to the along-dip direction pointing downwards, and \textquotedbl{}\emph{strike}\textquotedbl{}
-is the horizontal along-strike vector such that the system is right-handed.\\
-
-Slip is defined as displacement on side 2 minus displacement on side
-1. In the local coordinate system on the fault, positive along-strike
-slip is right-lateral and positive along-dip slip is thrust if side
-1 is on the hanging wall (normal faulting if side 1 is on the foot
-wall).\\
-
-Traction is defined as the stress induced on side 1 by side 2, which
-is the stress tensor times the normal vector outgoing from side 1.
-In the local coordinate system on the fault, the normal traction is
-negative in compression, positive along-strike traction generates
-right-lateral slip and positive along-dip traction generates thrust
-slip if side 1 is on the hanging wall (normal faulting if side 1 is
-on the foot wall).
-
-\section{Input Files}
-
-Three additional input files are required in the \texttt{DATA/} directory
-for dynamic and kinematic faults. They are $\mathtt{Par\_file\_faults}$,
-$\mathtt{FAULT\_STATIONS}$ and $\mathtt{input\_file.txt}$ (optional).
-If the former does not exist, the code assumes that there are no faults.\\
-
-\begin{lyxlist}{00.00.0000}
-\item [{\textbf{DATA/Par\_file\_faults}}] contains parameters of the fault.
-The first part of this file has a strict format:
-
-\begin{description}
-\item [{Line}] 1: Number of faults (NF)
-\item [{Lines}] 2 to NF+1: Kelvin Voigt damping (in seconds) for each fault.
-(See below how to set this parameter)
-\item [{Line}] NF+2: Type of simulation (1=dynamic , 2 = kinematic)
-\item [{Line}] NF+3: Number of time steps between updates of the time series
-outputs at selected fault points (see DATA/FAULT\_STATIONS), usually
-a large number (100s or 1000s). Note that the sampling rate of the
-time series is usually much higher.
-\item [{Line}] NF+4: Number of time steps between fault snapshot outputs
-(quantities at every fault point exported at regular times), usually
-a large number (100s or 1000s).
-\item [{Line}] NF+5: Slip velocity threshold below which frictional healing
-is set (friction coefficient is reset to its static value). If this
-value is negative healing is disabled.
-\item [{Line}] NF+6: Slip velocity threshold to define the rupture front.
-Only used for outputs.
-\end{description}
-
-The rest of this file is made of namelist input blocks (see \textquotedbl{}namelist\textquotedbl{}
-in a Fortran 9x manual). The input for each fault has the following
-sequence (arguments in {[}brackets{]} are optional):
-
-\&\textbf{BEGIN\_FAULT} /
-
-\&\textbf{INIT\_STRESS} S1, S2, S3 {[},n1, n2, n3{]} /
-
-followed by (n1+n2+n3) \&\textbf{DIST2D} blocks
-
-\&\textbf{SWF} mus, mud, dc {[}, nmus, nmud, ndc{]} /
-
-followed by (nmus+nmud+ndc) \&\textbf{DIST2D} blocks\\
-
-\begin{lyxlist}{00.00.0000}
-\item [{\&\textbf{INIT\_STRESS}}] input block sets the initial fault stresses
-relative to the foot-wall side of the fault. Initial stresses are
-composed of a constant background value possibly overwritten in prescribed
-regions by heterogeneous distributions (see \&\textbf{DIST2D} blocks
-below):
-
-\begin{description}
-\item [{S1}] = constant background value of along-strike shear stress (positive
-in the usual strike direction)
-\item [{S2}] = constant background value of along-dip shear (positive is
-down-dip, normal faulting)
-\item [{S3}] = constant background value of normal stress (negative in
-compression)
-\item [{n1}] = number of heterogeneous items for along-strike shear stress
-{[}default is 0{]}
-\item [{n2}] = number of heterogeneous items for along-dip shear stress
-{[}default is 0{]}
-\item [{n3}] = number of heterogeneous items for normal stress {[}default
-is 0{]}
-\end{description}
-\item [{\&\textbf{SWF}}] input block sets the slip-weakening friction parameters
-of the fault:
-
-\begin{description}
-\item [{mus}] = constant background value of static friction coefficient
-\item [{mud}] = constant background value of dynamic friction coefficient
-\item [{dc}] = constant background value of critical slip-weakening distance
-\item [{nmus}] = number of heterogeneous items for static friction coefficient
-{[}default is 0{]}
-\item [{nmud}] = number of heterogeneous items for dynamic friction coefficient
-{[}default is 0{]}
-\item [{ndc}] = number of heterogeneous items for critical slip-weakening
-distance {[}default is 0{]}\\
-
-\end{description}
-\item [{\&\textbf{DIST2D}}] input blocks modify (overwrite) the value of
-a fault parameter by a heterogeneous spatial distribution:
-
-\&\textbf{DIST2D} shape='$\mathtt{square}$', val, xc, yc, zc, l /
-
-$\;$sets a constant value (val) within a cube with center (xc,yc,zc)
-and edge size l.
-
-\&\textbf{DIST2D} shape='$\mathtt{rectangle}$', val, xc, yc, zc,
-lx, ly, lz /
-
-$\;$sets a constant value (val) within a rectangular prism with center
-(xc,yc,zc) and edge sizes (lx,ly,lz).
-
-\&\textbf{DIST2D} shape='$\mathtt{rectangle-taper}$', val, valh,
-xc, yc, zc, lx, ly, lz /
-
-$\;$sets a vertical linear gradient within a rectangular prism with
-center (xc,yc,zc) and edge sizes (lx,ly,lz). Values vary linearly
-as a function of vertical position z between value val at z = zc-lz/2
-and value valh at z = zc+lz/2 .
-
-\&\textbf{DIST2D} shape='$\mathtt{circular}$', val, xc, yc, zc, r
-/
-
-$\;$sets a constant value (val) within a sphere with center (xc,yc,zc)
-and radius r. \\
-
-\end{lyxlist}
-\end{lyxlist}
-
-\begin{lyxlist}{00.00.0000}
-\item [{\textbf{DATA/FAULT\_STATIONS}}] Stations in the fault plane.
-
-\textbf{Line 1} : number of stations.
-
-\textbf{Line 2 to end}: 5 columns: X, Y, Z (-depth), station name,
-fault-id
-
-The fault-id identifies the fault that contains the station. It is
-the index of appearance in the faults list after line 2 of Par\_file\_faults
-
-\end{lyxlist}
-
-\begin{lyxlist}{00.00.0000}
-\item [{\textbf{DATA/input\_file.txt}}] Heterogeneous stresses and friction
-parameters are documented in page 10 of
-
-\begin{lyxlist}{00.00.0000}
-\item [{$\mathtt{EXAMPLES/tpv16/description/TPV16\_17\_Description\_v03.pdf}$}]~
-\end{lyxlist}
-
-To activate this feature, in fault\_solver\_dynamic.f90 set \textbf{TPV16}
-= .true. then re-compile the code.
-
-\end{lyxlist}
-
-\section{\label{sec:Setting-the-Kelvin-Voigt}Setting the Kelvin-Voigt Damping
-Parameter }
-
-The purpose of the Kelvin-Voigt viscosity in the dynamic fault solver
-is to damp spurious oscillations generated by the fault slip at frequencies
-that are too high to be resolved by the mesh. The viscosity $\mathtt{eta}$
-(in seconds) depends on the size of the elements on the fault. Here
-is how to set it:
-\begin{enumerate}
-\item Determine the average linear size of the elements on the fault plane,
-$\mathtt{h\_fault}$. Usually this value is prescribed by the user
-during mesh generation. Otherwise it can be found by inspection of
-the mesh inside the CUBIT GUI.
-\item Use the Matlab function $\mathtt{utils/critical\_timestep.m}$ to
-compute\\
- $\mathtt{dtc\_fault}=\mathtt{critical\_timestep\left(c_{p},h\_fault,ngll\right)}$.
-\\
-This is the critical time step in an elastic medium for a hypothetical
-element of cubic shape with size equal to $\mathtt{h\_fault}$.
-\item Set $\mathtt{eta}$ in $\mathtt{Par\_file\_faults}$ to (0.1 to 0.3)$\ensuremath{\times}\mathtt{dtc\_fault}$.
-A larger $\mathtt{eta}$ damps high-frequencies more aggressively
-but it might also affect lower frequencies and rupture speed.
-\end{enumerate}
-Viscosity reduces numerical stability: the critical timestep in a
-simulation with Kelvin-Voigt damping needs to be smaller than that
-in a purely elastic simulation. Here is how to set the time step accordingly:
-\begin{enumerate}
-\item Run a test simulation without viscosity ($\mathtt{eta}$=0 and only
-a few time steps)
-\item Look for the \textquotedbl{}maximum suggested time step\textquotedbl{}
-in \texttt{OUTPUT\_FILES/output\_mesher.txt}. This is the critical
-timestep of a purely elastic simulation, $\mathtt{dtc\_bulk}$.
-\item Reset the timestep of the simulation with a Kelvin-Voigt material
-to a value smaller than\\
- $\mathtt{dtc\_kv}=\mathtt{eta}\left(\sqrt{1+\mathtt{dtc\_bulk^{2}}/\mathtt{eta^{2}}}-1\right)$
-\end{enumerate}
-Note that in general $\mathtt{dtc\_bulk}$ is smaller than $\mathtt{dtc\_fault}$,
-because elements off the fault might be smaller or more distorted
-than element faces on the fault.
-
-\section{Output Files }
-
-Several output files are saved in \texttt{OUTPUT\_FILES/}:\\
-
-\begin{enumerate}
-\item Seismograms for each station on the fault plane given in \texttt{DATA/FAUL\_STATIONS}.
-One output file is generated for each station, named after the station.
-The files are ascii and start with a header (22 lines long) followed
-by a data block with the following format, one line per time sample:\\
-
-Column 1 = Time (s)
-
-Column 2 = horizontal right-lateral slip (m)
-
-Column 3 = horizontal right-lateral slip rate (m/s)
-
-Column 4 = horizontal right-lateral shear stress (MPa)
-
-Column 5 = vertical up-dip slip (m)
-
-Column 6 = vertical up-dip slip rate (m/s)
-
-Column 7 = vertical up-dip shear stress (MPa)
-
-Column 8 = normal stress (MPa)\\
-
-The stresses are relative to the footwall side of the fault (this
-convention controls their sign, but not their amplitude). Slip is
-defined as displacement of the hanging wall relative to the footwall.
-
-\item Seismograms at stations in the bulk (out of the fault plane) given
-in \texttt{DATA/STATIONS}.
-\item Rupture time files are named $\mathtt{Rupture\_time\_FAULT}$-$\mathtt{id}$.
-One file is generated for each fault. The files are ascii and start
-with a header (12 lines long) followed by a data block with the following
-format, one line per fault node:
-
-Column 1 = horizontal coordinate, distance along strike (m)
-
-Column 2 = vertical coordinate, distance down-dip (m)
-
-Column 3 = rupture time (s)
-
-\item Fault quantities (slip, slip rate, stresses, etc) at regular times
-are stored in binary data files called $\mathtt{Snapshot\#it\#.bin}$,
-where \#it\# is the timestep number. These can be read in Matlab with
-the function $\mathtt{utils/FSEM3D\_snapshot.m}$
-\end{enumerate}
-
-\section{Post-processing and Visualization }
-
-Some Matlab functions for post-processing and visualization are included
-in directory $\mathtt{utils}$. The functions are internally documented
-(see their matlab help).\\
-
-\texttt{FSEM3D\_snapshot} reads a fault data snapshot\\
-
-The directories \texttt{EXAMPLES/*/post} contain additional Matlab
-scripts to generate figures specific to each example.
-
-\chapter{\label{cha:Adjoint-Simulations}Adjoint Simulations}
-
-Adjoint simulations are generally performed for two distinct applications.
-First, they can be used in point source moment-tensor inversions, or source imaging for
-earthquakes with large ruptures such as the Lander's earthquake \citep{WaHe94}.
-Second, they can be used to generate finite-frequency sensitivity
-kernels that are a critical part of tomographic inversions based upon
-3D reference models \citep{TrTaLi05,LiTr06,TrKoLi08,LiTr08}. In either
-case, source parameter or velocity structure updates are sought to
-minimize a specific misfit function (e.g., waveform or traveltime
-differences), and the adjoint simulation provides a means of computing
-the gradient of the misfit function and further reducing it in successive
-iterations. Applications and procedures pertaining to source studies
-and finite-frequency kernels are discussed in Sections~\ref{sec:Adjoint-simulation-sources}
-and \ref{sec:Adjoint-simulation-finite}, respectively. The two related
-parameters in the \texttt{Par\_file} are \texttt{SIMULATION\_TYPE}
-(1, 2 or 3) and the \texttt{SAVE\_FORWARD} (boolean).
-
-\section{\label{sec:Adjoint-simulation-sources}Adjoint Simulations for Sources
-Only (not for the Model)}
-
-When a specific misfit function between data and synthetics is minimized to invert
-for earthquake source parameters, the gradient of the misfit function
-with respect to these source parameters can be computed by placing
-time-reversed seismograms at the receivers as virtual sources
-in an adjoint simulation. Then the value of the gradient is obtained
-from the adjoint seismograms recorded at the original earthquake location.
-\begin{enumerate}
-\item \textbf{Prepare the adjoint sources} \label{enu:Prepare-the-adjoint}
-
-\begin{enumerate}
-\item First, run a regular forward simlation (\texttt{SIMULATION\_TYPE =
-1} and \texttt{SAVE\_FORWARD = .false.}). You can automatically set
-these two variables using the \texttt{\small utils/change\_simulation\_type.pl}{\small{}
-script:}{\small \par}
-
-\begin{lyxcode}
-utils/change\_simulation\_type.pl~-f~
-\end{lyxcode}
-
-and then collect the recorded seismograms at all the stations given
-in \texttt{DATA/STATIONS}.
-
-\item Then select the stations for which you want to compute the time-reversed
-adjoint sources and run the adjoint simulation, and compile them into
-the \texttt{DATA/STATIONS\_ADJOINT} file, which has the same format
-as the regular \texttt{DATA/STATIONS} file.
-
-\begin{itemize}
-\item Depending on what type of misfit function is used for the source inversion,
-adjoint sources need to be computed from the original recorded seismograms
-for the selected stations and saved in a sub-directory called \texttt{SEM/}
-in the root directory of the code, with the format \texttt{NT.STA.BX?.adj}, where \texttt{NT},
-\texttt{STA} are the network code and station name given in the \texttt{DATA/STATIONS\_ADJOINT}
-file, and \texttt{BX?} represents the component name of a particular
-adjoint seismogram. Please note that the band code can change depending
-on your sampling rate (see Appendix~\ref{cha:channel-codes} for
-further details).
-\item The adjoint seismograms are in the same format as the original seismogram
-(\texttt{NT.STA.BX?.sem?}), with the same start time, time interval
-and record length.
-\end{itemize}
-\item Notice that even if you choose to time reverse only one component
-from one specific station, you still need to supply all three components
-because the code is expecting them (you can set the other two components
-to be zero).
-\item Also note that since time-reversal is done in the code itself, no
-explicit time-reversing is needed for the preparation of the adjoint
-sources, i.e., the adjoint sources are in the same forward time sense
-as the original recorded seismograms.
-\end{enumerate}
-\item \textbf{Set the related parameters and run the adjoint simulation}\\
- In the \texttt{DATA/Par\_file}, set the two related parameters to
-be \texttt{SIMULATION\_TYPE = 2} and \texttt{SAVE\_FORWARD = .false.}.
-More conveniently, use the scripts \texttt{utils/change\_simulation\_type.pl}
-to modify the \texttt{Par\_file} automatically (\texttt{change\_simulation\_type.pl
--a}). Then run the solver to launch the adjoint simulation.
-\item \textbf{Collect the seismograms at the original source location}
-
-After the adjoint simulation has completed successfully, collect the
-seismograms from \texttt{LOCAL\_PATH}.
-\begin{itemize}
-\item These adjoint seismograms are recorded at the locations of the original
-earthquake sources given by the \texttt{DATA/CMTSOLUTION} file, and
-have names of the form \texttt{NT.S?????.S??.sem} for the six-component
-strain tensor (\texttt{SNN,SEE,SZZ,SNE,SNZ,SEZ}) at these locations,
-and ~\\
- \texttt{NT.S?????.BX?.sem} for the three-component displacements
-(\texttt{BXN,BXE,BXZ}) recorded at these locations.
-\item \texttt{S?????} denotes the source number; for example, if the original
-\texttt{CMTSOLUTION} provides only a point source, then the seismograms
-collected will start with \texttt{S00001}.
-\item These adjoint seismograms provide critical information for the computation
-of the gradient of the misfit function.
-\end{itemize}
-\end{enumerate}
-
-\section{\label{sec:Adjoint-simulation-finite}Adjoint Simulations for Finite-Frequency
-Kernels (Kernel Simulation)}
-
-Finite-frequency sensitivity kernels are computed in two successive
-simulations (please refer to \citet{LiTr06} and \citet{TrKoLi08}
-for details).
-\begin{enumerate}
-\item \textbf{Run a forward simulation with the state variables saved at
-the end of the simulation}
-
-Prepare the \texttt{\small CMTSOLUTION}{\small{} and }\texttt{\small STATIONS}{\small{}
-files, set the parameters }\texttt{\small SIMULATION\_TYPE}{\small {}
-}\texttt{\small =}{\small {} }\texttt{\small 1}{\small{} and }\texttt{\small SAVE\_FORWARD
-=}{\small {} }\texttt{\small .true.}{\small{} in the }\texttt{\small Par\_file}{\small{}
-(}\texttt{\small change\_simulation\_type -F}{\small ), and run the
-solver.}{\small \par}
-\begin{itemize}
-\item Notice that attenuation is not implemented yet for the computation
-of finite-frequency kernels; therefore set \texttt{ATTENUATION = .false.}
-in the \texttt{Par\_file}.
-\item We also suggest you modify the half duration of the \texttt{CMTSOLUTION}
-to be similar to the accuracy of the simulation (see Equation \ref{eq:shortest_period})
-to avoid too much high-frequency noise in the forward wavefield, although
-theoretically the high-frequency noise should be eliminated when convolved
-with an adjoint wavefield with the proper frequency content.
-\item This forward simulation differs from the regular simulations (\texttt{\small SIMULATION\_TYPE}{\small {}
-}\texttt{\small =}{\small {} }\texttt{\small 1}{\small{} and }\texttt{\small SAVE\_FORWARD}{\small {}
-}\texttt{\small =}{\small {} }\texttt{\small .false.}{\small ) described
-in the previous chapters in that the state variables for the last
-time step of the simulation, including wavefields of the displacement,
-velocity, acceleration, etc., are saved to the }\texttt{\small LOCAL\_PATH}{\small{}
-to be used for the subsequent simulation. }{\small \par}
-\item For regional simulations, the files recording the absorbing boundary
-contribution are also written to the \texttt{LOCAL\_PATH} when \texttt{SAVE\_FORWARD
-= .true.}.
-\end{itemize}
-\item \textbf{Prepare the adjoint sources}
-
-The adjoint sources need to be prepared the same way as described
-in the Section \ref{enu:Prepare-the-adjoint}.
-\begin{itemize}
-\item In the case of travel-time finite-frequency kernel for one source-receiver
-pair, i.e., point source from the \texttt{CMTSOLUTION}, and one station
-in the \texttt{STATIONS\_ADJOINT} list, we supply a sample program
-in \texttt{utils/adjoint\_sources/traveltime/xcreate\_adjsrc\_traveltime}
-to cut a certain portion of the original displacement seismograms
-and convert it into the proper adjoint source to compute the finite-frequency
-kernel.
-
-\begin{lyxcode}
-xcreate\_adjsrc\_traveltime~t1~t2~ifile{[}0-5{]}~E/N/Z-ascii-files~{[}baz{]}
-\end{lyxcode}
-
-where \texttt{t1} and \texttt{t2} are the start and end time of the
-portion you are interested in, \texttt{ifile} denotes the component
-of the seismograms to be used (0 for all three components, 1 for East,
-2 for North, and 3 for vertical, 4 for transverse, and 5 for radial
-component), \texttt{E/N/Z-ascii-files} indicate the three-component
-displacement seismograms in the right order, and \texttt{baz} is the
-back-azimuth of the station. Note that \texttt{baz} is only supplied
-when \texttt{ifile} = 4 or 5.
-
-\item Similarly, a sample program to compute adjoint sources for amplitude
-finite-frequency kernels may be found in \texttt{utils/adjoint\_sources/amplitude}
-and used in the same way as described for traveltime measurements
-
-\begin{lyxcode}
-xcreate\_adjsrc\_amplitude~t1~t2~ifile{[}0-5{]}~E/N/Z-ascii-files~{[}baz{]}.
-\end{lyxcode}
-\end{itemize}
-\item \textbf{Run the kernel simulation}
-
-With the successful forward simulation and the adjoint source ready
-in the \texttt{SEM/} directory, set \texttt{SIMULATION\_TYPE
-= 3} and \texttt{SAVE\_FORWARD = .false.} in the \texttt{Par\_file}
-(you can use \texttt{change\_simulation\_type.pl -b}), and rerun the
-solver.
-\begin{itemize}
-\item The adjoint simulation is launched together with the back reconstruction
-of the original forward wavefield from the state variables saved from
-the previous forward simulation, and the finite-frequency kernels
-are computed by the interaction of the reconstructed forward wavefield
-and the adjoint wavefield.
-\item The back-reconstructed seismograms at the original station locations
-are saved to the \texttt{LOCAL\_PATH} at the end of the kernel simulations,
-and can be collected to the local disk.
-\item These back-constructed seismograms can be compared with the time-reversed
-original seismograms to assess the accuracy of the backward reconstruction,
-and they should match very well.
-\item The arrays for density, P-wave speed and S-wave speed kernels are
-also saved in the \texttt{LOCAL\_PATH} with the names \texttt{proc??????\_rho(alpha,beta)\_kernel.bin},
-where \texttt{proc??????} represents the processor number, \texttt{rho(alpha,beta)}
-are the different types of kernels.
-\end{itemize}
-\item \textbf{Run the anisotropic kernel simulation}
-
-Instead of the kernels for the isotropic wave speeds, you can also
-compute the kernels for the 21 independent components $C_{IJ},\, I,J=1,...,6$
-(using Voigt's notation) of the elastic tensor in the cartesian coordinate
-system. This is done by setting \texttt{ANISOTROPIC\_KL} \texttt{=}
-\texttt{.true.} in \texttt{constants.h} before compiling the package.
-The definition of the parameters $C_{IJ}$ in terms of the corresponding
-components $c_{ijkl},ijkl,i,j,k,l=1,2,3$ of the elastic tensor in
-cartesian coordinates follows \citet{ChTr07}. The 21 anisotropic
-kernels are saved in the \texttt{LOCAL\_PATH} in one file with the
-name of \texttt{proc??????\_cijkl\_kernel.bin} (with \texttt{proc??????}
-the processor number). The output kernels correspond to absolute perturbations
-$\delta C_{IJ}$ of the elastic parameters and their unit is in $s/GPa/km^{3}$.
-For consistency, the output density kernels with this option turned
-on are for a perturbation $\delta\rho$ (and not $\frac{\delta\rho}{\rho}$)
-and their unit is in s / (kg/m$^{3}$) / km$^{3}$. These `primary'
-anisotropic kernels can then be combined to obtain the kernels for
-different parameterizations of anisotropy. This can be done, for example,
-when combining the kernel files from slices into one mesh file (see
-Section~\ref{sec:Finite-Frequency-Kernels}).
-
-If \texttt{ANISOTROPIC\_KL} \texttt{=} \texttt{.true.} by additionally
-setting \texttt{ANISOTROPIC\_KL} \texttt{=} \texttt{.true.} in \texttt{constants.h}
-the package will save anisotropic kernels paramterized as velocities
-related to transverse isotropy based on the the Chen and Tromp parameters
-\citet{ChTr07}. The kernels are saved as relative perturbations for
-horizontal and vertical P and S velocities, $\alpha_{v},\alpha_{h},\beta_{v},\beta_{h}$.
-Explicit relations can be found in appendix B. of \citet{SiLiTrTr07b}
-
-The anisotropic kernels are only currently available for CPU mode.
-
-\end{enumerate}
-In general, the three steps need to be run sequentially to assure
-proper access to the necessary files. If the simulations are run through
-some cluster scheduling system (e.g., LSF), and the forward simulation
-and the subsequent kernel simulations cannot be assigned to the same
-set of computer nodes, the kernel simulation will not be able to access
-the database files saved by the forward simulation. Solutions for
-this dilemma are provided in Chapter~\ref{cha:Scheduler}. Visualization
-of the finite-frequency kernels is discussed in Section~\ref{sec:Finite-Frequency-Kernels}.
-
-%------------------------------------------------------------------------------------------------%
-
-\chapter{\label{cha:tomo}Doing tomography, i.e., updating the model based on the sensitivity kernels obtained}
-
-%------------------------------------------------------------------------------------------------%
-
-UNDER CONSTRUCTION (March 2015). Will be added around April / May 2015.
-
-\section{Tomographic inversion / imaging using the sensitivity kernels obtained}
-
-OLD VERSION OF THE SECTION, WILL SOON BE IMPROVED AND ENRICHED.\\
-
-One of the fundamental reasons for computing sensitivity kernels (Section~\ref{sec:Adjoint-simulation-finite})
-is to use them within a tomographic inversion. In other words, use
-recorded seismograms, make measurements with synthetic seismograms,
-and use the misfit between the two to iteratively improve the model
-described by (at least) $V_{{\rm p}}$, $V_{{\rm s}}$, and $\rho$
-as a function of space.
-
-Whatever misfit function you use for the tomographic inversion (you can find several
-examples in \citet{TrKoLi08} and \citet{TrTaLi05}), you will weight
-the sensitivity kernels with measurements. Furthermore, you will use
-as many measurements (stations, components, time windows) as possible
-per event; hence, we call these composite kernels ``event kernels,''
-which are volumetric fields representing the gradient of the misfit
-function with respect to one of the variables (\eg $V_{{\rm s}}$).
-The basic features of an adjoint-based tomographic inversion were
-illustrated in \citet{TaLiTr07} and \citet{TrKoLi08} using a conjugate-gradient
-algorithm; there are dozens of versions of gradient-based inversion
-algorithms that could alternatively be used. The tomographic inversion
-of \citet{TaLiMaTr09,TaLiMaTr2010} used SPECFEM3D Cartesian as well
-as several additional components which are also stored on the CIG
-svn server, described next.
-
-The directory containing external utilities for tomographic inversion using
-SPECFEM3D Cartesian (or other packages that evaluate misfit functions
-and gradients) is in directory \texttt{utils/ADJOINT\_TOMOGRAPHY\_TOOLS/}:
-\begin{verbatim}
-flexwin/     -- FLEXWIN algorithm for automated picking of time windows
-measure_adj/ -- reads FLEXWIN output file and makes measurements,
-                with the option for computing adjoint sources
-iterate_adj/ -- various tools for iterative inversion
-                (requires pre-computed "event kernels")
-\end{verbatim}
-This directory also contains
-a brief \verb+README+ file indicating the role of the three subdirectories,
-\verb+flexwin+ \citep{Maggi2009}, \verb+measure_adj+, and \verb+iterate_adj+.
-The components for making the model update are there; however, there
-are no explicit rules for computing the model update, just as with
-any optimization problem. There are options for computing a conjugate
-gradient step, as well as a source subspace projection step.
-
-The best single file to read is probably: \verb+ADJOINT_TOMO/iterate_adj/cluster/README+.
-
-%------------------------------------------------------------------------------------------------%
-
-\chapter{Noise Cross-correlation Simulations}
-
-%------------------------------------------------------------------------------------------------%
-
-Besides earthquake simulations, SPECFEM3D Cartesian includes functionality
-for seismic noise tomography as well. In order to proceed successfully
-in this chapter, it is critical that you have already familiarized
-yourself with procedures for meshing (Chapter \ref{cha:Mesh-Generation}),
-creating distributed databases (Chapter \ref{cha:Creating-Distributed-Databases}),
-running earthquake simulations (Chapters \ref{cha:Running-the-Solver})
-and adjoint simulations (Chapter \ref{cha:Adjoint-Simulations}).
-Also, make sure you read the article `Noise cross-correlation sensitivity
-kernels' \citep{trompetal2010}, in order to understand noise simulations
-from a theoretical perspective.
-
-\section{Input Parameter Files}
-
-As usual, the three main input files are crucial: \texttt{Par\_file},
-\texttt{CMTSOLUTION} and \texttt{STATIONS}. Unless otherwise specified,
-those input files should be located in directory \texttt{DATA/}.\\
-
-\texttt{CMTSOLUTION} is required for all simulations. At a first glance,
-it may seem unexpected to have it here, since the noise simulations
-should have nothing to do with the earthquake -- \texttt{CMTSOLUTION}.
-However, for noise simulations, it is critical to have no earthquakes.
-In other words, the moment tensor specified in \texttt{CMTSOLUTION}
-must be set to zero manually!\\
-
-\texttt{STATIONS} remains the same as in previous earthquake simulations,
-except that the order of receivers listed in \texttt{STATIONS} is
-now important. The order will be used to determine the `master' receiver,
-i.e., the one that simultaneously cross correlates with the others.\\
-
-\texttt{Par\_file} also requires careful attention. A parameter called
-\texttt{NOISE\_TOMOGRAPHY} has been added which specifies the type
-of simulation to be run. \texttt{NOISE\_TOMOGRAPHY} is an integer
-with possible values 0, 1, 2 and 3. For example, when \texttt{NOISE\_TOMOGRAPHY}
-equals 0, a regular earthquake simulation will be run. When it is
-1/2/3, you are about to run step 1/2/3 of the noise simulations respectively.
-Should you be confused by the three steps, refer to \citet{trompetal2010}
-for details.\\
-
-Another change to \texttt{Par\_file} involves the parameter \texttt{NSTEP}.
-While for regular earthquake simulations this parameter specifies
-the length of synthetic seismograms generated, for noise simulations
-it specifies the length of the seismograms used to compute cross correlations.
-The actual cross correlations are thus twice this length, i.e., $2~\mathrm{NSTEP}-1$.
-The code automatically makes the modification accordingly, if \texttt{NOISE\_TOMOGRAPHY}
-is not zero.\\
-
-There are other parameters in \texttt{Par\_file} which should be given
-specific values. For instance, since the first two steps for calculating
-noise sensitivity kernels correspond to forward simulations, \texttt{SIMULATION\_TYPE}
-must be 1 when \texttt{NOISE\_TOMOGRAPHY} equals 1 or 2. Also, we
-have to reconstruct the ensemble forward wavefields in adjoint simulations,
-therefore we need to set \texttt{SAVE\_FORWARD} to \texttt{.true.}
-for the second step, i.e., when \texttt{NOISE\_TOMOGRAPHY} equals
-2. The third step is for kernel constructions. Hence \texttt{SIMULATION\_TYPE}
-should be 3, whereas \texttt{SAVE\_FORWARD} must be \texttt{.false.}.\\
-
-Finally, for most system architectures, please make sure that \texttt{LOCAL\_PATH}
-in \texttt{Par\_file} is in fact local, not globally shared. Because
-we have to save the wavefields at the earth's surface at every time
-step, it is quite problematic to have a globally shared \texttt{LOCAL\_PATH},
-in terms of both disk storage and I/O speed.
-
-\section{Noise Simulations: Step by Step}
-
-Proper parameters in those parameter files are not enough for noise
-simulations to run. We have more parameters to specify: for example,
-the ensemble-averaged noise spectrum, the noise distribution etc.
-However, since there are a few `new' files, it is better to introduce
-them sequentially. In this section, standard procedures for noise
-simulations are described.
-
-\subsection{Pre-simulation}
-\begin{itemize}
-\item As usual, we first configure the software package using:
-
-\texttt{./configure FC=ifort MPIFC=mpif90}
-
-Use the following if SCOTCH is needed:
-
-\texttt{./configure FC=ifort MPIFC=mpif90 -{}-with-scotch-dir=/opt/scotch}\\
-
-\item Next, we need to compile the source code using:
-
-\texttt{make xgenerate\_databases}
-
-\texttt{make xspecfem3D} \\
-
-\item Before we can run noise simulations, we have to specify the noise
-statistics, e.g., the ensemble-averaged noise spectrum. Matlab scripts
-are provided to help you to generate the necessary file.
-
-\texttt{EXAMPLES/noise\_tomography/NOISE\_TOMOGRAPHY.m} (main program)\\
- \texttt{EXAMPLES/noise\_tomography/PetersonNoiseModel.m}
-
-In Matlab, simply run:
-
-\texttt{NOISE\_TOMOGRAPHY(NSTEP, DT, Tmin, Tmax, NOISE\_MODEL)}\\
-
-\texttt{DT} is given in \texttt{Par\_file}, but \texttt{NSTEP} is
-NOT the one specified in \texttt{Par\_file}. Instead, you have to
-feed $2~\mathrm{NSTEP}-1$ to account for the doubled length of cross
-correlations. \texttt{Tmin} and \texttt{Tmax} correspond to the period
-range you are interested in, whereas \texttt{NOISE\_MODEL} denotes
-the noise model you will be using. Details can be found in the Matlab
-script.
-
-After running the Matlab script, you will be given the following information
-(plus a figure in Matlab):
-
-\texttt{{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}}\\
- \texttt{the source time function has been saved in:}\\
- \texttt{/data2/yangl/3D\_NOISE/S\_squared} (note this path must be
-different)\\
- \texttt{S\_squared should be put into directory:}\\
- \texttt{OUTPUT\_FILES/NOISE\_TOMOGRAPHY/ in the SPECFEM3D Cartesian
-package}\\
-
-In other words, the Matlab script creates a file called \texttt{S\_squared},
-which is the first `new' input file we encounter for noise simulations.
-
-One may choose a flat noise spectrum rather than Peterson's noise
-model. This can be done easily by modifying the Matlab script a little.
-
-\item Create a new directory in the SPECFEM3D Cartesian package, name it
-as \texttt{OUTPUT\_FILES/NOISE\_TOMOGRAPHY/}. We will add some parameter
-files later in this folder.
-\item Put the Matlab-generated-file \texttt{S\_squared} in \texttt{OUTPUT\_FILES/NOISE\_TOMOGRAPHY/}.
-
-That's to say, you will have a file \texttt{OUTPUT\_FILES/NOISE\_TOMOGRAPHY/S\_squared}
-in the SPECFEM3D Cartesian package.
-
-\item Create a file called \texttt{OUTPUT\_FILES/NOISE\_TOMOGRAPHY/irec\_master\_noise}.
-Note that this file is located in directory \texttt{OUTPUT\_FILES/NOISE\_TOMOGRAPHY/}
-as well. In general, all noise simulation related parameter files
-go into that directory. \texttt{irec\_master\_noise} contains only
-one integer, which is the ID of the `master' receiver. For example,
-if this file contains 5, it means that the fifth receiver listed in
-\texttt{DATA/STATIONS} becomes the `master'. That's why we mentioned
-previously that the order of receivers in \texttt{DATA/STATIONS} is
-important.
-
-Note that in regional simulations, the \texttt{DATA/STATIONS} might
-contain receivers which are outside of our computational domains.
-Therefore, the integer in \texttt{irec\_master\_noise} is actually
-the ID in \texttt{DATA/STATIONS\_FILTERED} (which is generated by
-\texttt{bin/xgenerate\_databases}).
-
-\item Create a file called \texttt{OUTPUT\_FILES/NOISE\_TOMOGRAPHY/nu\_master}.
-This file holds three numbers, forming a (unit) vector. It describes
-which component we are cross-correlating at the `master' receiver,
-i.e., ${\hat{{\bf \nu}}}^{\alpha}$ in \citet{trompetal2010}. The
-three numbers correspond to E/N/Z components respectively. Most often,
-the vertical component is used, and in those cases the three numbers
-should be 0, 0 and 1.
-\item Describe the noise direction and distributions in \texttt{src/shared/noise\_tomography.f90}.
-Search for a subroutine called \texttt{noise\_distribution\_direction}
-in \texttt{noise\_tomography.f90}. It is actually located at the very
-beginning of \texttt{noise\_tomography.f90}. The default assumes vertical
-noise and a uniform distribution across the whole free surface of
-the model. It should be quite self-explanatory for modifications.
-Should you modify this part, you have to re-compile the source code.
-\end{itemize}
-
-\subsection{Simulations}
-
-With all of the above done, we can finally launch our simulations.
-Again, please make sure that the \texttt{LOCAL\_PATH} in \texttt{DATA/Par\_file}
-is not globally shared. It is quite problematic to have a globally
-shared \texttt{LOCAL\_PATH}, in terms of both disk storage and speed
-of I/O (we have to save the wavefields at the earth's surface at every
-time step).
-
-As discussed in \citet{trompetal2010}, it takes three steps/simulations
-to obtain one contribution of the ensemble sensitivity kernels:
-\begin{itemize}
-\item Step 1: simulation for generating wavefields\\
- \texttt{SIMULATION\_TYPE=1}\\
- \texttt{NOISE\_TOMOGRAPHY=1}\\
- \texttt{SAVE\_FORWARD} not used, can be either .true. or .false.
-\item Step 2: simulation for ensemble forward wavefields\\
- \texttt{SIMULATION\_TYPE=1}\\
- \texttt{NOISE\_TOMOGRAPHY=2}\\
- \texttt{SAVE\_FORWARD=.true.}
-\item Step 3: simulation for ensemble adjoint wavefields and sensitivity
-kernels\\
- \texttt{SIMULATION\_TYPE=3}\\
- \texttt{NOISE\_TOMOGRAPHY=3}\\
- \texttt{SAVE\_FORWARD=.false.}
-
-Note Step 3 is an adjoint simulation, please refer to previous chapters
-on how to prepare adjoint sources and other necessary files, as well
-as how adjoint simulations work.
-
-\end{itemize}
-Note that it is better to run the three steps continuously within
-the same job, otherwise you have to collect the saved surface movies
-from the old nodes/CPUs to the new nodes/CPUs. This process varies
-from cluster to cluster and thus cannot be discussed here. Please
-ask your cluster administrator for information/configuration of the
-cluster you are using.\\
-
-\subsection{Post-simulation}
-
-After those simulations, you have all stuff you need, either in the
-\texttt{OUTPUT\_FILES/} or in the directory specified by \texttt{LOCAL\_PATH}
-in \texttt{DATA/Par\_file} (which are most probably on local nodes).
-Collect whatever you want from the local nodes to your workstation,
-and then visualize them. This process is the same as what you may
-have done for regular earthquake simulations. Refer to other chapters
-if you have problems.\\
-
-Simply speaking, two outputs are the most interesting: the simulated
-ensemble cross correlations and one contribution of the ensemble sensitivity
-kernels.\\
-
-The simulated ensemble cross correlations are obtained after the second
-simulation (Step 2). Seismograms in \texttt{OUTPUT\_FILES/} are actually
-the simulated ensemble cross correlations. Collect them immediately
-after Step 2, or the Step 3 will overwrite them. Note that we have
-a `master' receiver specified by \texttt{irec\_master\_noise}, the
-seismogram at one station corresponds to the cross correlation between
-that station and the `master'. Since the seismograms have three components,
-we may obtain cross correlations for different components as well,
-not necessarily the cross correlations between vertical components.\\
-
-One contribution of the ensemble cross-correlation sensitivity kernels
-are obtained after Step 3, stored in the \texttt{DATA/LOCAL\_PATH}
-on local nodes. The ensemble kernel files are named the same as regular
-earthquake kernels.\\
-
-You need to run another three simulations to get the other contribution
-of the ensemble kernels, using different forward and adjoint sources
-given in \citet{trompetal2010}.
-
-\section{Example}
-
-In order to illustrate noise simulations in an easy way, one example
-is provided in \texttt{EXAMPLES/noise\_tomography/}. See \texttt{EXAMPLES/noise\_tomography/README}
-for explanations. \\
-
-Note, however, that they are created for a specific workstation (CLOVER@PRINCETON),
-which has at least 4 cores with `mpif90' working properly. \\
-
-If your workstation is suitable, you can run the example in \texttt{EXAMPLES/noise\_tomography/}
-using:\\
-
-\texttt{./pre-processing.sh}\\
-
-Even if this script does not work on your workstation, the procedure
-it describes is universal. You may review the whole process described
-in the last section by following the commands in \texttt{pre-processing.sh},
-which should contain enough explanations for all the commands.
-
-%------------------------------------------------------------------------------------------------%
-
-\chapter{Gravity integral calculations for the gravity field of an Earth model}
-
-%------------------------------------------------------------------------------------------------%
-
-SPECFEM can now compute the gravity field as well as its derivatives (i.e., gravity gradiometry)
-generated by any given 3D Earth model at a given elevation, i.e., on a given observation surface.
-That feature is still experimental but should work just fine.\\
-
-To see how it is implemented and to use it, type this in the root directory of the code:\\
-
-\texttt{grep -i GRAVITY\_INTEGRALS src/*/*}
-
-%------------------------------------------------------------------------------------------------%
-
-\chapter{Graphics}
-
-%------------------------------------------------------------------------------------------------%
-
-\section{\label{sec:Mesh-graphics}Meshes}
-
-In case you used the internal mesher \texttt{xmeshfem3D} to create
-and partition your mesh, you can output mesh files in ABAQUS (.INP)
-and DX (.dx) format to visualize them. For this, you must set either
-the flag \texttt{CREATE\_DX\_FILES} or \texttt{CREATE\_ABAQUS\_FILES}
-to \texttt{.true.} in the mesher's parameter file \texttt{Mesh\_Par\_file}
-prior to running the mesher (see Chapter~\ref{cha:Running-the-Mesher-Meshfem3D}
-for details). You can then use AVS \urlwithparentheses{www.avs.com}
-or OpenDX \urlwithparentheses{www.opendx.org} to visualize the mesh
-and MPI partition (slices).
-
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=0.49\textwidth]{figures/vtk_mesh_vp} \includegraphics[width=0.49\textwidth]{figures/vtk_mesh_vs}
-\par\end{centering}
-
-\caption{Visualization using Paraview of VTK files created by \texttt{xgenerate\_databases}
-showing P- and S-wave velocities assigned to the mesh points. The
-mesh was created by \texttt{xmeshfem3D} for 4 processors.}
-
-\label{fig:vtk.mesh}
-\end{figure}
-
-You have also the option to visualize the distributed databases produced
-by \texttt{xgenerate\_databases} using Paraview \urlwithparentheses{www.paraview.org}.
-For this, you must set the flag \texttt{SAVE\_MESH\_FILES} to \texttt{.true.}
-in the main parameter file \texttt{Par\_file} (see Chapter~\ref{cha:Main-Parameter}
-for details). This will create VTK files for each single partition.
-You can then use Paraview \urlwithparentheses{www.paraview.org}
-to visualized these partitions.
-
-\section{\label{sec:Movies}Movies}
-
-To make a surface or volume movie of the simulation, set parameters
-\texttt{MOVIE\_SURFACE}, \texttt{MOVIE\_VOLUME}, \texttt{MOVIE\_TYPE},
-and \texttt{NTSTEP\_BETWEEN\_FRAMES} in the \texttt{Par\_file}. Turning
-on the movie flags, in particular \texttt{MOVIE\_VOLUME}, produces
-large output files. \texttt{MOVIE\_VOLUME} files are saved in the
-\texttt{LOCAL\_PATH} directory, whereas \texttt{MOVIE\_SURFACE} output
-files are saved in the \texttt{OUTPUT\_FILES} directory. We save the
-displacement field if the parameter \texttt{SAVE\_DISPLACEMENT} is
-set, otherwise the velocity field is saved. The look of a movie is
-determined by the half-duration of the source. The half-duration should
-be large enough so that the movie does not contain frequencies that
-are not resolved by the mesh, i.e., it should not contain numerical
-noise. This can be accomplished by selecting a CMT \texttt{HALF\_DURATION}
-> 1.1 $\times$ smallest period (see figure \ref{fig:CMTSOLUTION-file}).
-When \texttt{MOVIE\_SURFACE} = .\texttt{true.}, the half duration
-of each source in the \texttt{CMTSOLUTION} file is replaced by
-\begin{quote}
-\[
-\sqrt{(}\mathrm{\mathtt{HALF\_DURATIO}\mathtt{N}^{2}}+\mathrm{\mathtt{HDUR\_MOVI}\mathtt{E}^{2}})
-\]
- \textbf{NOTE:} If \texttt{HDUR\_MOVIE} is set to 0.0, the code will
-select the appropriate value of 1.1 $\times$ smallest period. As
-usual, for a point source one can set \texttt{half duration} in the
-\texttt{CMTSOLUTION} file to be 0.0 and \texttt{HDUR\_MOVIE} = 0.0
-in the \texttt{Par\_file} to get the highest frequencies resolved
-by the simulation, but for a finite source one would keep all the
-\texttt{half durations} as prescribed by the finite source model and
-set \texttt{HDUR\_MOVIE} = 0.0.
-\end{quote}
-
-\subsection{Movie Surface}
-
-When running \texttt{xspecfem3D} with the \texttt{MOVIE\_SURFACE}
-flag turned on, the code outputs \texttt{moviedata??????} files in
-the \texttt{OUTPUT\_FILES} directory. There are several flags in the
-main parameter file \texttt{Par\_file} that control the output of
-these movie data files (see section \ref{cha:Main-Parameter} for
-details): \texttt{\small NTSTEP\_BETWEEN\_FRAMES}{\small{} to set the
-timesteps between frames, }\texttt{\small SAVE\_DISPLACEMENT}{\small{}
-to save displacement instead of velocity, }\texttt{\small USE\_HIGHRES\_FOR\_MOVIES}{\small{}
-to save values at all GLL point instead of element edges. In order
-to output additionally shakemaps, you would set the parameter }\texttt{\small CREATE\_SHAKEMAP}{\small{}
-to }\texttt{\small .true.}{\small .}{\small \par}
-
-The files are in a fairly complicated binary format, but there is
-a program provided to convert the output into more user friendly formats:
-\begin{description}
-\item [{\texttt{xcreate\_movie\_shakemap\_AVS\_DX\_GMT}}] From \texttt{create\_movie\_shakemap\_AVS\_DX\_GMT.f90},
-it outputs data in ASCII, OpenDX, or AVS format (also readable in
-ParaView). Before compiling the code, make sure you have the file
-\texttt{surface\_from\_mesher.h} in the \texttt{OUTPUT\_FILES/} directory.
-This file will be created by the solver run. Then type
-
-\begin{lyxcode}
-{\footnotesize make~xcreate\_movie\_shakemap\_AVS\_DX\_GMT}{\footnotesize \par}
-\end{lyxcode}
-
-and run the executable \texttt{xcreate\_movie\_shakemap\_AVS\_DX\_GMT}
-in the main directory. It will create visualization files
-in your format of choice. The code will prompt the user for input
-parameters.
-
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=0.32\textwidth]{figures/movie_surf_1} \includegraphics[width=0.32\textwidth]{figures/movie_surf_2}
-\includegraphics[width=0.32\textwidth]{figures/movie_surf_3}
-\par\end{centering}
-
-\caption{Visualization using AVS files created by \texttt{xcreate\_movie\_shakemap\_AVS\_DX\_GMT}
-showing movie snapshots of vertical velocity components at different
-times.}
-
-\label{fig:movie.surf}
-\end{figure}
-
-\end{description}
-The \texttt{SPECFEM3D Cartesian} code is running in near real-time
-to produce animations of southern California earthquakes via the web;
-see Southern California ShakeMovie\textregistered{}\urlwithparentheses{www.shakemovie.caltech.edu}.
-
-\subsection{\label{sub:Movie-Volume}Movie Volume}
-
-When running xspecfem3D with the \texttt{\small MOVIE\_VOLUME}{\small{}
-flag turned on, the code outputs several files in }\texttt{\small LOCAL\_PATH}{\small{}
-specified in the main }\texttt{\small Par\_file}{\small , e.g. in
-directory }\texttt{\small OUTPUT\_FILES/DATABASES\_MPI}{\small .
-The output is saved by each processor at the time interval specified
-by }\texttt{\small NTSTEP\_BETWEEN\_FRAMES}{\small . For all domains,
-the velocity field is output to files: }\texttt{\small proc??????\_velocity\_X\_it??????.bin}{\small ,
-}\texttt{\small proc??????\_velocity\_Y\_it??????.bin}{\small{} and
-}\texttt{\small proc??????\_velocity\_Z\_it??????.bin}{\small . For
-elastic domains, the divergence and curl taken from the velocity field,
-i.e. $\nabla\cdot{\bf {v}}$ and $\nabla\times{\bf {v}}$, get stored
-as well: }\texttt{\small proc??????\_div\_it??????.bin}{\small , }\texttt{\small proc??????\_curl\_X\_t??????.bin}{\small ,
-}\texttt{\small proc??????\_curl\_Y\_it??????.bin}{\small{} and }\texttt{\small proc??????\_curl\_Z\_it??????.bin}{\small .
-The files denoted }\texttt{\small proc??????\_div\_}{\small ~}\\
-{\small{} }\texttt{\small glob\_it??????.bin}{\small{} and }\texttt{\small proc??????\_curl\_glob\_it??????.bin}{\small{}
-are stored on the global points only, all the other arrays are stored
-on all GLL points. Note that the components X/Y/Z can change to E/N/Z
-according to the }\texttt{\small SUPPRESS\_UTM\_PROJECTION}{\small{}
-flag (see also Appendix~\ref{cha:Coordinates} and \ref{cha:channel-codes}).}{\small \par}
-
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=0.32\textwidth]{figures/movie_volume_1} \includegraphics[width=0.32\textwidth]{figures/movie_volume_2}
-\includegraphics[width=0.32\textwidth]{figures/movie_volume_3}
-\par\end{centering}
-
-\caption{Paraview visualization using movie volume files (converted by \texttt{xcombine\_vol\_data}
-and \texttt{mesh2vtu.pl}) and showing snapshots of vertical velocity
-components at different times.}
-
-\label{fig:movie.volume}
-\end{figure}
-
-To visualize these files, we use an auxilliary program \texttt{combine\_vol\_data.f90}
-to combine the data from all slices into one mesh file. To compile
-it in the root directory, type:
-\begin{lyxcode}
-{\footnotesize make~xcombine\_vol\_data~}{\footnotesize \par}
-\end{lyxcode}
-which will create the executable \texttt{bin/xcombine\_vol\_data}.
-To output the usage of this executable, type
-\begin{lyxcode}
-{\footnotesize ./bin/xcombine\_vol\_data~}{\footnotesize \par}
-\end{lyxcode}
-without arguments.
-
-\medskip
-
-\texttt{xcombine\_vol\_data} will combine the data on the different processors (located in \texttt{OUTPUT\_FILES/DATABASES\_MPI}), for a given quantity and a given iteration, to one file.
-For example, if you want to combine \texttt{velocity\_Z}, for the iteration 400 on 4 processors, i.e.
-if you want to combine these files :
-
-\medskip
-
-\texttt{proc000000\_velocity\_Z\_it000400.bin }
-
-\texttt{proc000001\_velocity\_Z\_it000400.bin }
-
-\texttt{proc000002\_velocity\_Z\_it000400.bin }
-
-\texttt{proc000003\_velocity\_Z\_it000400.bin}
-
-\medskip
-
-You have to go in the directory of the concerned example (where are the directories \texttt{DATA}, \texttt{OUTPUT\_FILES}, etc.).
-Then, you can launch \texttt{xcombine\_vol\_data}, specifying the path where it is located.
-As an example, if the \texttt{DATA} and \texttt{OUTPUT\_FILES} directories of the concerned example are in the root specfem3d directory, \texttt{xcombine\_vol\_data} is in \texttt{./bin}, and you have to type:
-
-\begin{lyxcode}
-~{\footnotesize ./bin/xcombine\_vol\_data~0~3~velocity\_Z\_it000400~./OUTPUT\_FILES/DATABASES\_MPI~./OUTPUT\_FILES~0~}{\footnotesize \par}
-\end{lyxcode}
-
-Here, $0$ is the number of the first processor, $3$ the number of the last one, \texttt{velocity\_Z\_it000400} the name of the files we want to combine without the prefix \texttt{proc000000*\_}, \texttt{./OUTPUT\_FILES/DATABASES\_MPI} the directory where the files are located, \texttt{./OUTPUT\_FILES} the directory where the combined file will be stored, and $0$ is the parameter to create a low-resolution mesh file ($1$ for a high-resolution).
-
-\medskip
-
-At the beginning of the source file \texttt{combine\_vol\_data.f90} (located in \texttt{src/auxiliaries/}), there is a logical constant parameter \texttt{USE\_VTK\_OUTPUT} whose default value is \texttt{.true.}, that which will provide the output combined file directly in the VTK format and you can use Paraview to visualize it.
-
-\medskip
-
-If the value of this parameter is \texttt{.false.}, the output mesh file will have the name \texttt{velocity\_Z\_it000400.mesh}. Then we next have to convert the
-\texttt{.mesh} file into the VTU (Unstructured grid file) format which can be viewed in ParaView. For this task, you can use and modify the
-script \texttt{mesh2vtu.pl} located in directory \texttt{utils/Visualization/Paraview/}, for example:
-\begin{lyxcode}
-{\footnotesize mesh2vtu.pl~-i~velocity\_Z\_it000400.mesh~-o~velocity\_Z\_it000400.vtu}{\footnotesize \par}
-\end{lyxcode}
-Notice that this Perl script uses a program \texttt{mesh2vtu} in the \texttt{utils/Visualization/Paraview/mesh2vtu} directory, which further
-uses the VTK \urlwithparentheses{http://www.vtk.org/} run-time library for its execution. Therefore, make sure you have them properly set
-in the script according to your system.
-
-\bigskip
-
-Then, to do a movie with several iterations, you have to repeat this process for each iteration you want to put in your movie.
-
-\section{\label{sec:Finite-Frequency-Kernels}Finite-Frequency Kernels}
-
-The finite-frequency kernels computed as explained in Section \ref{sec:Adjoint-simulation-finite}
-are saved in the \texttt{LOCAL\_PATH} at the end of the simulation.
-Therefore, we first need to collect these files on the front end,
-combine them into one mesh file, and visualize them with some auxilliary
-programs.
-\begin{enumerate}
-\item \textbf{Create slice files}
-
-We will only discuss the case of one source-receiver pair, i.e., the
-so-called banana-doughnut kernels. Although it is possible to collect
-the kernel files from all slices on the front end, it usually takes
-up too much storage space (at least tens of gigabytes). Since the
-sensitivity kernels are the strongest along the source-receiver great
-circle path, it is sufficient to collect only the slices that are
-along or close to the great circle path.
-
-A Perl script \texttt{slice\_number.pl} located in directory \texttt{utils/Visualization/Paraview/}
-can help to figure out the slice numbers that lie along the great
-circle path. It applies to meshes created with the internal mesher
-\texttt{xmeshfem3D}.
-\begin{enumerate}
-\item On machines where you have access to the script, copy the \texttt{Mesh\_Par\_file},
-and \texttt{output\_solver} files, and run:
-
-\begin{lyxcode}
-{\small slice\_number.pl~Mesh\_Par\_file~output\_solver.txt~slice\_file}{\small \par}
-\end{lyxcode}
-
-which will generate a \texttt{slices\_file}.
-
-\item For cases with multiple sources and multiple receivers, you need to
-provide a slice file before proceeding to the next step.
-\end{enumerate}
-\item \textbf{Collect the kernel files}
-
-After obtaining the slice files, you can collect the corresponding
-kernel files from the given slices.
-\begin{enumerate}
-\item You can use or modify the script \texttt{utils/copy\_basin\_database.pl}
-to accomplish this:
-
-\begin{lyxcode}
-{\small utils/copy\_database.pl~slice\_file~lsf\_machine\_file~filename~{[}jobid{]}}{\small \par}
-\end{lyxcode}
-
-where \texttt{\small lsf\_machine\_file}{\small{} is the machine file
-generated by the LSF scheduler, }\texttt{\small filename}{\small{}
-is the kernel name (e.g., }\texttt{\small rho\_kernel}{\small , }\texttt{\small alpha\_kernel}{\small{}
-and }\texttt{\small beta\_kernel}{\small ), and the optional }\texttt{\small jobid}{\small{}
-is the name of the subdirectory under }\texttt{\small LOCAL\_PATH}{\small{}
-where all the kernel files are stored.}{\small \par}
-
-\item After executing this script, all the necessary mesh topology files
-as well as the kernel array files are collected to the local directory
-of the front end.
-\end{enumerate}
-\item \textbf{Combine kernel files into one mesh file}
-
-We use an auxilliary program \texttt{combine\_vol\_data.f90} to combine
-the kernel files from all slices into one mesh file.
-\begin{enumerate}
-\item Compile it in the root directory:
-
-\begin{lyxcode}
-{\footnotesize make~xcombine\_vol\_data~}{\footnotesize \par}
-
-{\footnotesize ./bin/xcombine\_vol\_data~slice\_list~filename~input\_dir~output\_dir~high/low-resolution~}{\footnotesize \par}
-\end{lyxcode}
-
-where \texttt{input\_dir} is the directory where all the individual
-kernel files are stored, and \texttt{output\_dir} is where the mesh
-file will be written.
-
-\item Use 1 for a high-resolution mesh, outputting all the GLL points to
-the mesh file, or use 0 for low resolution, outputting only the corner
-points of the elements to the mesh file.
-\item The output mesh file will have the name \texttt{filename\_rho(alpha,beta).mesh}
-\end{enumerate}
-\item \textbf{Convert mesh files into .vtu files}
-
-\begin{enumerate}
-\item We next convert the \texttt{.mesh} file into the VTU (Unstructured
-grid file) format which can be viewed in ParaView. For this task,
-you can use and modify the script \texttt{mesh2vtu.pl} located in
-directory \texttt{utils/Visualization/Paraview/}, for example:
-
-\begin{lyxcode}
-{\footnotesize mesh2vtu.pl~-i~file.mesh~-o~file.vtu~}{\footnotesize \par}
-\end{lyxcode}
-\item Notice that this Perl script uses a program \texttt{mesh2vtu} in the
-\texttt{utils/Visualization/Paraview/mesh2vtu} directory, which further
-uses the VTK \urlwithparentheses{http://www.vtk.org/} run-time library
-for its execution. Therefore, make sure you have them properly set
-in the script according to your system.
-\end{enumerate}
-\item \textbf{Copy over the source and receiver .vtk file}
-
-In the case of a single source and a single receiver, the simulation
-also generates the file \texttt{sr.vtk} located in the \texttt{OUTPUT\_FILES/}
-directory to describe the source and receiver locations, which can
-also be viewed in Paraview in the next step.
-
-\item \textbf{View the mesh in ParaView}
-
-Finally, we can view the mesh in ParaView \urlwithparentheses{www.paraview.org}.
-\begin{enumerate}
-\item Open ParaView.
-\item From the top menu, \textsf{File} $\rightarrow$\textsf{Open data},
-select \texttt{file.vtu}, and click the \textsf{Accept} button.
-
-\begin{itemize}
-\item If the mesh file is of moderate size, it shows up on the screen; otherwise,
-only the bounding box is shown.
-\end{itemize}
-\item Click \textsf{Display Tab} $\rightarrow$ \textsf{Display Style} $\rightarrow$
-\textsf{Representation} and select \textsf{wireframe of surface} to
-display it.
-\item To create a cross-section of the volumetric mesh, choose \textsf{Filter}
-$\rightarrow$ \textsf{cut}, and under \textsf{Parameters Tab}, choose
-\textsf{Cut Function} $\rightarrow$ \textsf{plane}.
-\item Fill in center and normal information given by the \texttt{global\_slice\_number.pl}
-script (either from the standard output or from \texttt{normal\_plane.txt}
-file).
-\item To change the color scale, go to \textsf{Display Tab} $\rightarrow$
-\textsf{Color} $\rightarrow$ \textsf{Edit Color Map} and reselect
-lower and upper limits, or change the color scheme.
-\item Now load in the source and receiver location file by \textsf{File}
-$\rightarrow$ \textsf{Open data}, select \texttt{sr.vt}k, and click
-the \textsf{Accept} button. Choose \textsf{Filter} $\rightarrow$
-\textsf{Glyph}, and represent the points by `\textsf{spheres}'.
-\item For more information about ParaView, see the ParaView Users Guide
-\urlwithparentheses{www.paraview.org/files/v1.6/ParaViewUsersGuide.PDF}.
-\end{enumerate}
-\end{enumerate}
-\begin{figure}[H]
-\noindent \begin{centering}
-\includegraphics[scale=0.6]{figures/3D-S-Kernel}
-\par\end{centering}
-
-\caption{(a) Top Panel: Vertical source-receiver cross-section of the S-wave
-finite-frequency sensitivity kernel $K_{\beta}$ for station GSC at
-an epicentral distance of 176 km from the September 3, 2002, Yorba
-Linda earthquake. Lower Panel: Vertical source-receiver cross-section
-of the 3D S-wave speed model used for the spectral-element simulations
-\citep{KoLiTrSuStSh04}. (b) The same as (a) but for station HEC at
-an epicentral distance of 165 km \citep{LiTr06}.}
-
-\label{figure:P-wave-speed-finite-frequency}
-\end{figure}
-
-\chapter{\label{cha:Scheduler}Running through a Scheduler}
+\chapter{Running through a Scheduler}\label{cha:Scheduler}
 
@@
 more specific running needs.}{\small \par}
-
-\chapter{{\normalsize \label{cha:-Changing-the}} Changing the Model}
-
-In this section we explain how to change the velocity model used for
-your simulations. These changes involve contributing specific subroutines
-that replace existing subroutines in the \texttt{SPECFEM3D Cartesian}
-package. Note that \texttt{SPECFEM3D Cartesian} can handle Earth models
-with material properties that vary within each spectral element.
-
-\section{{\normalsize \label{sec:Using-tomographic}}Using external tomographic Earth models}
-
-To implement your own external tomographic model(s), you must provide
-your own external tomography file(s), and choose between two possible
-options:\\
- \indent (1) set in the \texttt{Par\_file} the model parameter \texttt{MODEL
-= tomo}, \\
- or for more user control: \\
- \indent (2) set in the \texttt{Par\_file} the model parameter \texttt{MODEL
-= default}, define the negative \texttt{material\_ID} identifier for
-each element in the file \texttt{MESH/materials\_file} and use the
-following format in the file \texttt{MESH/nummaterial\_velocity\_file}
-when using CUBIT to construct your mesh (see also section \ref{subsec:Exporting-the-Mesh}):
-\begin{lyxcode}
-domain\_ID~material\_ID~tomography~elastic~file\_name~1
-\end{lyxcode}
-where: \\
- \indent \texttt{domain\_ID} is 1 for acoustic or 2 for elastic materials,
-\\
- \indent \texttt{material\_ID} a negative, unique identifier (i.e.,
--1,-2,...), \\
- \indent \texttt{tomography} keyword for tomographic material definition,
-\\
- \indent \texttt{elastic} keyword for elastic material definition,
-\\
- \indent \texttt{file\_name} the name of the tomography file and
-1 a positive unique identifier.\\
-
-The external tomographic model is represented by a grid of points
-with assigned material properties and homogeneous resolution along
-each spatial direction x, y and z. The ASCII file \texttt{file\_name}
-that describe the tomography should be located in the \texttt{TOMOGRAPHY\_PATH}
-directory, set in the \texttt{Par\_file}. The format of the file,
-as read from \texttt{model\_tomography.f90} located in the \texttt{src/generate\_databases}
-directory, looks like Figure \ref{fig:tomography_file}, where
-\begin{description}
-\item [{\texttt{ORIG\_X}, \texttt{END\_X}}] are, respectively, the coordinates
-of the initial and final tomographic grid points along the x direction
-(in the mesh units, e.g., $m$);
-\item [{\texttt{ORIG\_Y}, \texttt{END\_Y}}] respectively the coordinates
-of the initial and final tomographic grid points along the y direction
-(in the mesh units, e.g., $m$);
-\item [{\texttt{ORIG\_Z}, \texttt{END\_Z}}] respectively the coordinates
-of the initial and final tomographic grid points along the z direction
-(in the mesh units, e.g., $m$);
-\item [{\texttt{SPACING\_X}, \texttt{SPACING\_Y}, \texttt{SPACING\_Z}}] the
-spacing between the tomographic grid points along the x, y and z directions,
-respectively (in the mesh units, e.g., $m$);
-\item [{\texttt{NX}, \texttt{NY}, \texttt{NZ}}] the number of grid points
-along the spatial directions x, y and z, respectively; \texttt{NX}
-is given by {[}(\texttt{END\_X} - \texttt{ORIG\_X})/\texttt{SPACING\_X}{]}+1;
-\texttt{NY} and \texttt{NZ} are the same as \texttt{NX}, but for the
-y and z directions, respectively;
-\item [{\texttt{VP\_MIN}, \texttt{VP\_MAX}, \texttt{VS\_MIN}, \texttt{VS\_MAX}, \texttt{RHO\_MIN}, \texttt{RHO\_MAX}}] the
-minimum and maximum values of the wave speed \texttt{vp} and \texttt{vs}
-(in $m\, s^{-1}$) and of the density \texttt{rho} (in $kg\, m^{-3}$);
-these values could be the actual limits of the tomographic parameters
-in the grid or the minimum and maximum values to which we force the
-cut of velocity and density in the model.
-\end{description}
-After the first four lines, in the file \texttt{file\_name} the tomographic
-grid points are listed with the corresponding values of \texttt{vp},
-\texttt{vs} and \texttt{rho}, scanning the grid along the x coordinate
-(from \texttt{ORIG\_X} to \texttt{END\_X} with step of \texttt{SPACING\_X})
-for each given y (from \texttt{ORIG\_Y} to \texttt{END\_Y}, with step
-of \texttt{SPACING\_Y}) and each given z (from \texttt{ORIG\_Z} to
-\texttt{END\_Z}, with step of \texttt{SPACING\_Z}). %fig
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=1\textwidth]{figures/tomo_file}
-\par\end{centering}
-
-\caption{Tomography file \texttt{file\_name} that describes an external Earth
-model. The coordinates x, y and z of the grid, their limits ORIG\_X,
-ORIG\_Y, ORIG\_Z, END\_X, END\_Y, END\_Z and the grid spacings SPACING\_X,
-SPACING\_Y, SPACING\_Z should be in the units of the constructed mesh
-(e.g., UTM coordinates in meters).}
-
-\label{fig:tomography_file}
-\end{figure}
-
-The user can implement his own interpolation algorithm for the tomography
-model by changing the routine \texttt{model\_tomography.f90} located
-in the \texttt{src/generate\_databases/} directory. Moreover, for
-models that involve both fully defined materials and a tomography
-description, the \texttt{nummaterial\_velocity\_file} has multiple
-lines each with the corresponding suitable format described above.
-
-%==============
-
-\subsubsection*{Example: External tomography file with variable discretization intervals}
-
-The example in Figure~\ref{fig:tomography_file} is for an external tomography file with uniform spacing in the $x$, $y$, and $z$ directions.
-(Note that $\Delta x$, $\Delta y$, and $\Delta z$ need not be the same as in the example.) In that example, the  \texttt{nummaterial\_velocity\_file} is
-%
-\begin{lyxcode}
-2  -1 tomography elastic tomography\_model.xyz 1
-\end{lyxcode}
-%
-and the file \texttt{tomography\_model.xyz} will need to reside in \texttt{DATA/}.
-All entries in the second column of \texttt{materials\_file} will be \texttt{-1}, which means that each element in the mesh will be interpolated accordind to the values in \texttt{tomography\_model.xyz}.
-
-In some cases it may be desirable to have an external tomography model that is described in more than one file.
-For example, in cases like southern California, the length scale of variation in the structure of the wave speed model is much shorter in the sedimentary basin models within the upper 15~km.
-Therefore one might want to use an external tomography file that is sampled with, say, $\Delta x = 1000$~m, $\Delta y = 1000$~m, and $\Delta z = 250$~m in the uppermost 15~km, and then use $\Delta x = 2000$~m, $\Delta y = 2000$~m, and $\Delta z = 1000$~m below a depth of 15~km.
-If these intervals are chosen appropriately, then it will result in a pair of external tomography file that is much smaller than the alternative of having a single file with the fine discretization.
-In this case  \texttt{nummaterial\_velocity\_file} is
-%
-\begin{lyxcode}
-2  -1 tomography elastic file\_above\_15km.xyz 1 \\
-2  -2 tomography elastic file\_below\_15km.xyz 1
-\end{lyxcode}
-%
-and the files \texttt{file\_above\_15km.xyz} and \texttt{file\_below\_15km.xyz} will need to reside in \texttt{DATA/}.
-All entries in the second column of \texttt{materials\_file} will need to be either \texttt{-1} or \texttt{-2}, depending on whether the element is above or below 15~km depth.
-In this sense, the construction of the mesh and the discretization of the external model will generally be done in tandem.
-
-Other more complicated discretizations can be used following the same procedure. (In particular, there is no limit on the number of different external files that are used.)
-
-%=============
-
-\section{{\normalsize \label{sec:Anelastic-Models}}External (an)elastic Models}
-
-To use your own external model, you can set in the \texttt{Par\_file}
-the model parameter \texttt{MODEL = external}. Three-dimensional acoustic
-and/or (an)elastic (attenuation) models may then be superimposed onto
-the mesh based upon your subroutine in \texttt{model\_external\_values.f90}
-located in directory \texttt{src/generate\_databases}. The call to
-this routine would be as follows
-\begin{lyxcode}
-call~model\_external\_values(xmesh,~ymesh,~zmesh,~~\&~~\\
-~~~~~~~~~~~rho,~vp,~vs,~qmu\_atten,~iflag\_aniso,~idomain\_id)
-\end{lyxcode}
-Input to this routine consists of:
-\begin{description}
-\item [{\texttt{xmesh,ymesh,zmesh}}] location of mesh point
-\end{description}
-Output to this routine consists of:
-\begin{description}
-\item [{\texttt{rho,vp,vs}}] isotropic model parameters for density $\rho$
-($kg/m^{3}$), $v_{p}$ ($m/s$) and $v_{s}$ ($m/s$)
-\item [{\texttt{qmu\_atten}}] Shear wave quality factor: $0<Q_{\mu}<9000$
-\item [{\texttt{iflag\_aniso}}] anisotropic model flag, $0$ indicating
-no anisotropy or $1$ using anisotropic model parameters as defined
-in routine file \texttt{model\_aniso.f90}
-\item [{\texttt{idomain\_id}}] domain identifier, $1$ for acoustic, $2$
-for elastic, $3$ for poroelastic materials.
-\end{description}
-Note that the resolution and maximum value of anelastic models are
-truncated. This speeds the construction of the standard linear solids
-during the meshing stage. To change the resolution, currently at one
-significant figure following the decimal, or the maximum value (9000),
-consult \texttt{shared/constants.h}.
-
-\section{{\normalsize \label{sec:Anisotropic-Models}}Anisotropic Models}
-
-To use your anisotropic models, you can either set in the \texttt{Par\_file}
-the model parameter \texttt{ANISOTROPY} to \texttt{.true.} or set
-the model parameter \texttt{MODEL = aniso}. Three-dimensional anisotropic
-models may be superimposed on the mesh based upon the subroutine in
-file \texttt{model\_aniso.f90} located in directory \texttt{src/generate\_databases}.
-The call to this subroutine is of the form
-\begin{lyxcode}
-call~model\_aniso(iflag\_aniso,~rho,~vp,~vs,~\&~~\\
-~~~~~~~~~~~c11,c12,c13,c14,c15,c16,c22,c23,c24,c25,c26,~\&~~\\
-~~~~~~~~~~~c33,c34,c35,c36,c44,c45,c46,c55,c56,c66)~
-\end{lyxcode}
-Input to this routine consists of:
-\begin{description}
-\item [{\texttt{iflag\_aniso}}] flag indicating the type of the anisotropic
-model, i.e. $0$ for no superimposed anisotropy, $1$ or $2$ for
-generic pre-defined anisotropic models.
-\item [{\texttt{rho,vp,vs}}] reference isotropic model parameters for density
-$\rho$, $v_{p}$ and $v_{s}$.
-\end{description}
-Output from the routine consists of the following non-dimensional
-model parameters:
-\begin{description}
-\item [{\texttt{c11},}] \textbf{$\cdots$,} \texttt{\textbf{c66}} 21~dimensionalized
-anisotropic elastic parameters.
-\end{description}
-You can replace the \texttt{model\_aniso.f90} file by your own version
-\textit{provided you do not change the call structure of the routine},
-i.e., the new routine should take exactly the same input and produce
-the required relative output.
-
-\section{{\normalsize \label{sec:Using-SEP}}Using external SEP models.}
-
-Stanford Exploration Project (SEP) models consists of two files per variables, one header file (\texttt{.H})
-and one binary file (\texttt{.H$@$}).
-Among other information, the header file indicates, for each dimension,
-the offset (\texttt{o1}, \texttt{o2} and \texttt{o3} fields),
-the spatial step size (\texttt{d1}, \texttt{d2} and \texttt{d3} fields)
-and the number of samples (\texttt{n1}, \texttt{n2} and \texttt{n3}) necessary to read
-the linked binary file (\texttt{in} field).
-
-SEP model reading occurs during database generation. It it the user responsibility
-to ensure that proper interfaces between acoustic and elastic domains are set up
-during the meshing phase. For more information, refer to
-section~\ref{cha:Running-the-Mesher-Meshfem3D}.
-
-In \texttt{DATA/Par\_file}, \texttt{\textbf{MODEL}} should be set to \texttt{\textbf{sep}},
-and \texttt{\textbf{SEP\_MODEL\_DIRECTORY}} should point to wherever your sep model files reside.
-
-Note that in order not to overload the parameter file, SEP header files
-should be named \texttt{vp.H}, \texttt{vs.H} and \texttt{rho.H}. The \texttt{in} field
-in these binary files can be any path as long as it is relative to
-the previously set \texttt{SEP\_MODEL\_DIRECTORY}.
-We also assume that any dimensional value inside SEP header files is given in meters and
-that binary files are single precision floating point numbers stored in little endian format.
-There is currently only support for \texttt{vp}, \texttt{vs} and \texttt{rho} variables.
-
-An example demonstrating the use SEP models is given in:
-\texttt{EXAMPLES/meshfem3D\_examples/sep\_bathymetry}.
-
-\chapter{Post-Processing Scripts}
-
-Several post-processing scripts/programs are provided in the \texttt{utils/}
-directory, and most of them need to be adjusted when used on different
-systems, for example, the path of the executable programs. Here we
-only list a few of the available scripts and provide a brief description,
-and you can either refer to the related sections for detailed usage
-or, in a lot of cases, type the script/program name without arguments
-for its usage.
-
-\section{Process Data and Synthetics\label{sec:Process-data-and-syn}}
-
-In many cases, the SEM synthetics are calculated and compared to data
-seismograms recorded at seismic stations. Since the SEM synthetics
-are accurate for a certain frequency range, both the original data
-and the synthetics need to be processed before a comparison can be
-made.
-
-For such comparisons, the following steps are recommended:
-\begin{enumerate}
-\item Make sure that both synthetic and observed seismograms have the correct
-station/event and timing information.
-\item Convolve synthetic seismograms with a source time function with the
-half duration specified in the \texttt{CMTSOLUTION} file, provided,
-as recommended, you used a zero half duration in the SEM simulations.
-\item Resample both observed and synthetic seismograms to a common sampling
-rate.
-\item Cut the records using the same window.
-\item Remove the trend and mean from the records and taper them.
-\item Remove the instrument response from the observed seismograms (recommended)
-or convolve the synthetic seismograms with the instrument response.
-\item Make sure that you apply the same filters to both observed and synthetic
-seismograms. Preferably, avoid filtering your records more than once.
-\item Now, you are ready to compare your synthetic and observed seismograms.
-\end{enumerate}
-\noindent We generally use the following scripts provided in the \texttt{utils/seis\_process/}
-directory:
-
-\subsection{Data processing script \texttt{process\_data.pl}}
-
-This script cuts a given portion of the original data, filters it,
-transfers the data into a displacement record, and picks the first
-P and S arrivals. For more functionality, type `\texttt{process\_data.pl}'
-without any argument. An example of the usage of the script:
-\begin{lyxcode}
-{\footnotesize process\_data.pl~-m~CMTSOLUTION~-s~1.0~-l~0/4000~-i~-f~-t~40/500~-p~-x~bp~DATA/1999.330{*}.BH?.SAC}~
-\end{lyxcode}
-\noindent which has resampled the SAC files to a sampling rate of
-1 seconds, cut them between 0 and 4000 seconds, transfered them into
-displacement records and filtered them between 40 and 500 seconds,
-picked the first P and S arrivals, and added suffix `\texttt{bp}'
-to the file names.
-
-Note that all of the scripts in this section actually use the SAC
-and/or IASP91 to do the core operations; therefore make sure that
-the SAC and IASP91 packages are installed properly on your system,
-and that all the environment variables are set properly before running
-these scripts.
-
-\subsection{Synthetics processing script \texttt{process\_syn.pl}}
-
-This script converts the synthetic output from the SEM code from ASCII
-to SAC format, and performs similar operations as `\texttt{process\_data.pl}'.
-An example of the usage of the script:
-\begin{lyxcode}
-{\footnotesize process\_syn.pl~-m~CMTSOLUTION~-h~-a~STATIONS~-s~1.0~-l~0/4000~-f~-t~40/500~-p~-x~bp~SEM/{*}.BX?.semd}~
-\end{lyxcode}
-which will convolve the synthetics with a triangular source-time function
-from the \texttt{CMTSOLUTION} file, convert the synthetics into SAC
-format, add event and station information into the SAC headers, resample
-the SAC files with a sampling rate of 1 seconds, cut them between
-0 and 4000 seconds, filter them between 40 and 500 seconds with the
-same filter used for the observed data, pick the first P and S arrivals,
-and add the suffix `\texttt{bp}' to the file names.
-
-More options are available for this script, such as adding time shift
-to the origin time of the synthetics, convolving the synthetics with
-a triangular source time function with a given half duration, etc.
-Type \texttt{process\_syn.pl} without any argument for a detailed
-usage.
-
-In order to convert between SAC format and ASCII files, useful scripts
-are provided in the subdirectories ~\\
- \texttt{utils/sac2000\_alpha\_convert/} and \texttt{utils/seis\_process/asc2sac/}.
-
-\subsection{Script \texttt{rotate.pl}}
-
-The original data and synthetics have three components: vertical (BHZ
-resp. BXZ), north (BHN resp. BXN) and east (BHE resp. BXE). However,
-for most seismology applications, transverse and radial components
-are also desirable. Therefore, we need to rotate the horizontal components
-of both the data and the synthetics to the transverse and radial direction,
-and \texttt{\small rotate.pl}{\small{} can be used to accomplish this:}{\small \par}
-\begin{lyxcode}
-rotate.pl~-l~0~-L~180~-d~DATA/{*}.BHE.SAC.bp~
-
-rotate.pl~-l~0~-L~180~SEM/{*}.BXE.semd.sac.bp~
-\end{lyxcode}
-where the first command performs rotation on the SAC data obtained
-through Seismogram Transfer Program (STP) \urlwithparentheses{http://www.data.scec.org/STP/stp.html},
-while the second command rotates the processed SEM synthetics.
-
-\section{Collect Synthetic Seismograms}
-
-The forward and adjoint simulations generate synthetic seismograms
-in the \texttt{OUTPUT\_FILES/} directory by default. For the forward
-simulation, the files are named like \texttt{NT.STA.BX?.semd} for
-two-column time series, or \texttt{NT.STA.BX?.semd.sac} for ASCII
-SAC format, where NT and STA are the network code and station name,
-and \texttt{BX?} stands for the component name. Please see the Appendix~\ref{cha:Coordinates}
-and \ref{cha:channel-codes} for further details.
-
-The adjont simulations generate synthetic seismograms with the name
-\texttt{NT.S?????.S??.sem} (refer to Section \ref{sec:Adjoint-simulation-sources}
-for details). The kernel simulations output the back-reconstructed
-synthetic seismogram in the name \texttt{NT.STA.BX?.semd}, mainly
-for the purpose of checking the accuracy of the reconstruction. Refer
-to Section \ref{sec:Adjoint-simulation-finite} for further details.
-
-You do have further options to change this default output behavior,
-given in the main constants file \texttt{constants.h} located in \texttt{src/shared/}
-directory:
-\begin{description}
-\item [{\texttt{SEISMOGRAMS\_BINARY}}] set to \texttt{.true.} to have seismograms
-written out in binary format.
-\item [{\texttt{WRITE\_SEISMOGRAMS\_BY\_MASTER}}] Set to \texttt{.true.}
-to have only the master process writing out seismograms. This can
-be useful on a cluster, where only the master process node has access
-to the output directory.
-\item [{\texttt{USE\_OUTPUT\_FILES\_PATH}}] Set to \texttt{.false.} to
-have the seismograms output to \texttt{LOCAL\_PATH} directory specified
-in the main parameter file \texttt{DATA/Par\_file}. In this case,
-you could collect the synthetics onto the frontend using the \texttt{collect\_seismo\_lsf\_multi.pl}
-script located in the \texttt{utils/Cluster/lsf/} directory. The usage
-of the script would be e.g.:
-
-\begin{lyxcode}
-collect\_seismo.pl~machines~DATA/Par\_file
-\end{lyxcode}
-\end{description}
-where \texttt{machines} is a file containing the node names and \texttt{DATA/Par\_file}
-the parameter file used to extract the \texttt{LOCAL\_PATH} directory
-used for the simulation.
-
-\section{Clean Local Database}
-
-After all the simulations are done, the seismograms are collected,
-and the useful database files are copied to the frontend, you may
-need to clean the local scratch disk for the next simulation. This
-is especially important in the case of kernel simulation, where very
-large files are generated for the absorbing boundaries to help with
-the reconstruction of the regular forward wavefield. A sample script
-is provided in \texttt{utils/}:
-\begin{lyxcode}
-cleanbase.pl~machines
-\end{lyxcode}
-where \texttt{machines} is a file containing the node names.
-
-\section{Plot Movie Snapshots and Synthetic Shakemaps}
-
-\subsection{Script \texttt{movie2gif.gmt.pl}}
-
-With the movie data saved in \texttt{OUTPUT\_FILES/} at the end of
-a movie simulation (\texttt{\small MOVIE\_SURFACE=.true.}{\small ),
-you can run the }\texttt{\small `create\_movie\_shakemap\_AVS\_DX\_GMT}{\small '
-code to convert these binary movie data into GMT xyz files for futher
-processing. A sample script }\texttt{\small movie2gif.gmt.pl}{\small{}
-is provided to do this conversion, and then plot the movie snapshots
-in GMT, for example:}{\small \par}
-\begin{lyxcode}
-movie2gif.gmt.pl~-m~CMTSOLUTION~-g~-f~1/40~-n~-2~-p~
-\end{lyxcode}
-which for the first through the 40th movie frame, converts the \texttt{moviedata}
-files into GMT xyz files, interpolates them using the 'nearneighbor'
-command in GMT, and plots them on a 2D topography map. Note that `\texttt{-2}'
-and `\texttt{-p}' are both optional.
-
-\subsection{Script \texttt{plot\_shakemap.gmt.pl}}
-
-With the shakemap data saved in \texttt{OUTPUT\_FILES/} at the end
-of a shakemap simulation ~\\
- (\texttt{CREATE\_SHAKEMAP=.true.}), you can also run \texttt{`create\_movie\_shakemap\_AVS\_DX\_GMT}'
-code to convert the binary shakemap data into GMT xyz files. A sample
-script \texttt{plot\_shakemap.gmt.pl} is provided to do this conversion,
-and then plot the shakemaps in GMT, for example:
-\begin{lyxcode}
-plot\_shakemap.gmt.pl~data\_dir~type(1,2,3)~CMTSOLUTION~
-\end{lyxcode}
-where \texttt{type=1} for a displacement shakemap, \texttt{2} for
-velocity, and \texttt{3} for acceleration.
-
-\section{Map Local Database}
-
-A sample program \texttt{remap\_database} is provided to map the local
-database from a set of machines to another set of machines. This is
-especially useful when you want to run mesher and solver, or different
-types of solvers separately through a scheduler (refer to Chapter~\ref{cha:Scheduler}).
-\begin{lyxcode}
-run\_lsf.bash~-{}-gm-no-shmem~-{}-gm-copy-env~remap\_database~old\_machines~150
-\end{lyxcode}
-where \texttt{old\_machines} is the LSF machine file used in the previous
-simulation, and \texttt{150} is the number of processors in total.
-
-%============================================================
-
-\chapter{\label{cha:developers}Information for developers of the code, and for people who want to learn how the technique works}
-
-You can get a very simple 1D version of a demo code (there is one in Fortran and one in Python):\\
-
-git clone --recursive https://github.com/geodynamics/specfem1d.git\\
-
-\noindent We also have simple 3D demo source codes that implement the SEM in a single, small program, in directory\\
-utils/small\_SEM\_solvers\_in\_Fortran\_and\_C\_without\_MPI\_to\_learn of the specfem3d package.
-They are useful to learn how the spectral-element method works, and how to write or modify a code to implement it.
-Also useful to test new ideas by modifying these simple codes to run some tests.\\
-
-\noindent For information on how to contribute to the code, i.e., for how to make your modifications, additions or improvements part of the
-official package, see https://github.com/geodynamics/specfem3d/wiki .
-
-%============================================================
-
-\chapter*{\label{cha:Bug-Reports-and}Bug Reports and Suggestions for Improvements}
-
-To report bugs or suggest improvements to the code, please send an
-e-mail to the CIG Computational Seismology Mailing List \urlwithparentheses{cig-seismo@geodynamics.org}.
-
-\addcontentsline{toc}{chapter}{Notes \& Acknowledgments}
-
-\chapter*{Notes \& Acknowledgments}
-
-In order to keep the software package thread-safe in case a multithreaded
-implementation of MPI is used, developers should not add modules or
-common blocks to the source code but rather use regular subroutine
-arguments (which can be grouped in ``derived types'' if needed for
-clarity).
-
-The Gauss-Lobatto-Legendre subroutines in \texttt{gll\_library.f90}
-are based in part on software libraries from the Massachusetts Institute
-of Technology, Department of Mechanical Engineering (Cambridge, Massachusetts,
-USA). The non-structured global numbering software was provided by
-Paul F. Fischer (Brown University, Providence, Rhode Island, USA,
-now at Argonne National Laboratory, USA).
-
-OpenDX \urlwithparentheses{http://www.opendx.org} is open-source
-based on IBM Data Explorer, AVS \urlwithparentheses{http://www.avs.com}
-is a trademark of Advanced Visualization Systems, and ParaView \urlwithparentheses{http://www.paraview.org}
-is an open-source visualization platform.
-
-Please e-mail your feedback, questions, comments, and suggestions
-to the CIG Computational Seismology Mailing List \urlwithparentheses{cig-seismo@geodynamics.org}.
-
-\addcontentsline{toc}{chapter}{Copyright}
-
-\chapter*{Copyright}
-
-Main historical authors: Dimitri Komatitsch and Jeroen Tromp
-
-Princeton University, USA, and CNRS / University of Marseille, France
-
-$\copyright$ Princeton University and CNRS / University of Marseille, July 2012
-
-This program is free software; you can redistribute it and/or modify
-it under the terms of the GNU General Public License as published
-by the Free Software Foundation (see Appendix \ref{cha:License}).\\
-
-\textbf{\underline{Evolution of the code:}}\\
-
-MPI v. 3.0, December 2014: many developers.
-Convolutional PML, LDDRK time scheme, bulk attenuation support, simultaneous MPI runs,
-ADIOS file I/O support, coupling with external codes, new seismogram names,
-Deville routines for additional GLL degrees, tomography tools, unit/regression test framework,
-improved CUDA GPUs performance, additonal GEOCUBIT support, better make compilation,
-git versioning system. \\
-
-MPI v. 2.1, July 2012: Max Rietmann, Peter Messmer, Daniel Peter, Dimitri
-Komatitsch, Joseph Charles, Zhinan Xie: support for CUDA GPUs, better
-CFL stability for the Stacey absorbing conditions. \\
-
-MPI v. 2.0, November 2010: Dimitri Komatitsch, Nicolas Le Goff, Roland
-Martin and Pieyre Le Loher, University of Pau, France, Daniel Peter,
-Jeroen Tromp and the Princeton group of developers, Princeton University,
-USA, and Emanuele Casarotti, INGV Roma, Italy: support for CUBIT meshes
-decomposed by SCOTCH; much faster solver using Michel Deville's inlined
-matrix products.\\
-
-MPI v. 1.4 Dimitri Komatitsch, University of Pau, Qinya Liu and others,
-Caltech, September 2006: better adjoint and kernel calculations, faster
-and better I/Os on very large systems, many small improvements and
-bug fixes.\\
-
-MPI v. 1.3 Dimitri Komatitsch, University of Pau, and Qinya Liu, Caltech,
-July 2005: serial version, regular mesh, adjoint and kernel calculations,
-ParaView support.\\
-
-MPI v. 1.2 Min Chen and Dimitri Komatitsch, Caltech, July 2004: full
-anisotropy, volume movie.\\
-
-MPI v. 1.1 Dimitri Komatitsch, Caltech, October 2002: Zhu's Moho map,
-scaling of $V_{s}$ with depth, Hauksson's regional model, attenuation,
-oceans, movies.\\
-
-MPI v. 1.0 Dimitri Komatitsch, Caltech, USA, May 2002: first MPI version
-based on global code.\\
-
-Dimitri Komatitsch, IPG Paris, France, December 1996: first 3-D solver
-for the CM-5 Connection Machine, parallelized on 128 processors using
-Connection Machine Fortran.\\
-
-\addcontentsline{toc}{chapter}{References} \bibliography{bibliography}
-
-\appendix
-%dummy comment inserted by tex2lyx to ensure that this paragraph is not empty
-
-\chapter{\label{cha:Coordinates}Reference Frame Convention}
-
-The code uses the following convention for the Cartesian reference
-frame:
-\begin{itemize}
-\item the $x$ axis points East
-\item the $y$ axis points North
-\item the $z$ axis points up
-\end{itemize}
-Note that this convention is different from both the \citet{AkRi80}
-convention and the Harvard Centroid-Moment Tensor (CMT) convention.
-The Aki \& Richards convention is
-\begin{itemize}
-\item the $x$ axis points North
-\item the $y$ axis points East
-\item the $z$ axis points down
-\end{itemize}
-and the Harvard CMT convention is
-\begin{itemize}
-\item the $x$ axis points South
-\item the $y$ axis points East
-\item the $z$ axis points up
-\end{itemize}
-
-\subsection*{Source and receiver locations}
-
-The SPECFEM3D Cartesian software code internally uses Cartesian coordinates.
-The given locations for sources and receiver locations thus may get
-converted. Sources and receiver locations are read in from the \texttt{CMTSOLUTION}
-(or \texttt{FORCESOLUTION}) and \texttt{STATIONS} files. Note that
-e.g. the \texttt{CMTSOLUTION} and \texttt{FORCESOLUTION} files denotes
-the location by \textquotedbl{}longitude/latitude/depth\textquotedbl{}.
-We read in longitude as $x$ coordinate, latitude as $y$ coordinate.
-
-In case the flag \texttt{SUPPRESS\_UTM\_PROJECTION} is set to \texttt{.false.}
-in the main parameter file (see Chapter~\ref{cha:Creating-Distributed-Databases}),
-the $x$/$y$ coordinates have to be given in degrees and are converted
-to Cartesian coordinates using a UTM conversion with the specified
-UTM zone.
-
-The value for depth (given in $km$ in \texttt{CMTSOLUTION} and \texttt{FORCESOLUTION})
-or burial depth (given in $m$ in \texttt{STATIONS}) is evaluated
-with respect to the surface of the mesh at the specified $x$/$y$
-location to find the corresponding $z$ coordinate. It is possible
-to use this depth value directly as $z$ coordinate by changing the
-flag \texttt{USE\_SOURCES\_RECVS\_Z} to \texttt{.true.} in the file
-\texttt{constants.h} located in the \texttt{src/shared/} subdirectory.
-
-\subsection*{Seismogram outputs}
-
-The seismogram output directions are given in Cartesian $x$/$y$/$z$
-directions and not rotated any further. Changing flags in \texttt{constants.h}
-in the \texttt{src/shared/} subdirectory only rotates the seismogram
-outputs if receivers are forced to be located at the surface (\texttt{RECVS\_CAN\_BE\_BURIED\_EXT\_MESH}
-set to \texttt{.false.}) and the normal to the surface at the receiver
-location should be used (\texttt{EXT\_MESH\_RECV\_NORMAL} set to \texttt{.true.})
-as vertical. In this case, the outputs are rotated to have the vertical
-component normal to the surface of the mesh, $x$ and $y$ directions
-are somewhat arbitrary orthogonal directions along the surface.
-
-For the labeling of the seismogram channels, see Appendix~\ref{cha:channel-codes}.
-Additionally, we add labels to the synthetics using the following
-convention: For a receiver station located in an
-\begin{description}
-\item [{elastic domain:}] ~\\
-
-\begin{itemize}
-\item \texttt{semd} for the displacement vector
-\item \texttt{semv} for the velocity vector
-\item \texttt{sema} for the acceleration vector
-\end{itemize}
-\item [{acoustic domain:}] ~\\
- (please note that receiver stations in acoustic domains must be buried.
-This is due to the free surface condition which enforces a zero pressure
-at the surface.)
-
-\begin{itemize}
-\item \texttt{semd} for the displacement vector
-\item \texttt{semv} for the velocity vector
-\item \texttt{sema} for pressure which will be stored in the vertical component
-\texttt{Z}. Note that pressure in the acoustic media is isotropic
-and thus the pressure record would be the same in the other two component
-directions. We therefore use the other two seismogram components to
-store the acoustic potential in component \texttt{X} (or \texttt{N})
-and the first time derivative of the acoustic potential in component
-\texttt{Y} (or \texttt{E}).
-\end{itemize}
-\end{description}
-The seismograms are by default written out in ASCII-format to the
-\texttt{OUTPUT\_FILES/} subdirectory by each parallel process. You
-can change this behavior by changing the following flags in the \texttt{constants.h}
-file located in the \texttt{src/shared/} subdirectory:
-\begin{description}
-\item [{\texttt{SEISMOGRAMS\_BINARY}}] set to \texttt{.true.} to have seismograms
-written out in binary format.
-\item [{\texttt{WRITE\_SEISMOGRAMS\_BY\_MASTER}}] Set to \texttt{.true.}
-to have only the master process writing out seismograms. This can
-be useful on a cluster, where only the master process node has access
-to the output directory.
-\item [{\texttt{USE\_OUTPUT\_FILES\_PATH}}] Set to \texttt{.false.} to
-have the seismograms output to \texttt{LOCAL\_PATH} directory specified
-in the main parameter file \texttt{DATA/Par\_file}.
-\end{description}
-
-\chapter{\label{cha:channel-codes}Channel Codes of Seismograms}
-
-Seismic networks, such as the Global Seismographic Network (GSN),
-generally involve various types of instruments with different bandwidths,
-sampling properties and component configurations. There are standards
-to name channel codes depending on instrument properties. IRIS \texttt{(www.iris.edu)}
-uses SEED format for channel codes, which are represented by three
-letters, such as \texttt{LHN}, \texttt{BHZ}, etc. In older versions
-of the SPECFEM3D Cartesian package, a common format was used for the
-channel codes of all seismograms, which was \texttt{BHE/BHN/BHZ} for
-three components. To avoid confusion when comparison are made to observed
-data, we are now using the FDSN convention (\url{http://www.fdsn.org/})
-for SEM seismograms. In the following, we give a brief explanation
-of the FDSN convention used by IRIS, and how it is adopted in SEM
-seismograms. Please visit \texttt{www.iris.edu/manuals/SEED\_appA.htm}
-for further information.\\
-
-\noindent \texttt{\textbf{Band code:}} The first letter in the channel
-code denotes the band code of seismograms, which depends on the response
-band and the sampling rate of instruments. The list of band codes
-used by IRIS is shown in Figure \ref{fig:IRIS_band_codes}. The sampling
-rate of SEM synthetics is controlled by the resolution of simulations
-rather than instrument properties. However, for consistency, we follow
-the FDSN convention for SEM seismograms governed by their sampling
-rate. For SEM synthetics, we consider band codes for which $dt\leq1$
-s. IRIS also considers the response band of instruments. For instance,
-short-period and broad-band seismograms with the same sampling rate
-correspond to different band codes, such as S and B, respectively.
-In such cases, we consider SEM seismograms as broad band, ignoring
-the corner period ($\geq10$ s) of the response band of instruments
-(note that at these resolutions, the minimum period in the SEM synthetics
-will be less than $10$ s). Accordingly, when you run a simulation
-the band code will be chosen depending on the resolution of the synthetics,
-and channel codes of SEM seismograms will start with either \texttt{L},
-\texttt{M}, \texttt{B}, \texttt{H}, \texttt{C} or \texttt{F}, shown
-by red color in the figure.\\
-
-\begin{figure}[ht]
-\noindent \begin{centering}
-\includegraphics[scale=0.6]{figures/IRIS_band_codes}
-\par\end{centering}
-
-\caption{The band code convention is based on the sampling rate and the response
-band of instruments. Please visit \texttt{www.iris.edu/manuals/SEED\_appA.htm}
-for further information. Grey rows show the relative band-code range
-in SPECFEM3D Cartesian, and the band codes used to name SEM seismograms
-are denoted in red.}
-
-\label{fig:IRIS_band_codes}
-\end{figure}
-
-\noindent \texttt{\textbf{Instrument code:}} The second letter in
-the channel code corresponds to instrument codes, which specify the
-family to which the sensor belongs. For instance, \texttt{H} and \texttt{L}
-are used for high-gain and low-gain seismometers, respectively. The
-instrument code of SEM seismograms will always be \texttt{X}, as assigned
-by FDSN for synthetic seismograms. \\
-
-\noindent \texttt{\textbf{Orientation code:}} The third letter in
-channel codes is an orientation code, which generally describes the
-physical configuration of the components of instrument packages. SPECFEM3D
-Cartesian uses the traditional orientation code \texttt{E/N/Z} (East-West,
-North-South, Vertical) for three components when a UTM projection
-is used. If the UTM conversion is suppressed, i.e. the flag \texttt{SUPPRESS\_UTM\_PROJECTION}
-is set to \texttt{.true.}, then the three components are labelled
-\texttt{X/Y/Z} according to the Cartesian reference frame. \\
-
-\noindent \texttt{\textbf{EXAMPLE:}} The sampling rate is given by
-\texttt{DT} in the main parameter file \texttt{Par\_file} located
-in the \texttt{DATA/} subdirectory. Depending on the resolution of
-your simulations, if you choose a sampling rate greater than $0.01$
-s and less than $1$ s, a seismogram recording displacements on the
-vertical component of a station \texttt{ASBS} (network \texttt{AZ})
-will be named \texttt{AZ.ASBS.MXZ.semd.sac}, whereas it will be \texttt{AZ.ASBS.BXZ.semd.sac},
-if the sampling rate is greater than 0.0125 and less equal to 0.1
-s.
-
-\chapter{\label{cha:Troubleshooting}Troubleshooting}
-
-\section*{FAQ}
-\begin{description}
-\item [{configuration fails:}]~
-
-Examine the log file 'config.log'. It contains detailed informations.
-In many cases, the paths to these specific compiler commands F90,
-CC and MPIF90 will not be correct if `./configure` fails.
-
-Please make sure that you have a working installation of a Fortran
-compiler, a C compiler and an MPI implementation. You should be able
-to compile this little program code:
-\begin{lyxcode}
-program~main~~~\\
-~include~'mpif.h'~~~\\
-~integer,~parameter~::~CUSTOM\_MPI\_TYPE~=~MPI\_REAL~~~\\
-~integer~ier~~~\\
-~call~MPI\_INIT(ier)~~~\\
-~call~MPI\_BARRIER(MPI\_COMM\_WORLD,ier)~~~\\
-~call~MPI\_FINALIZE(ier)~~~\\
-~end
-\end{lyxcode}
-\item [{compilation fails stating:}] ~\\
-
-\begin{lyxcode}
-...~~~\\
-~obj/program\_generate\_databases.o:~In~function~`MAIN\_\_':~~~\\
-~program\_generate\_databases.f90:(.text+0x14):~undefined~reference~to~`\_gfortran\_set\_std'~~~\\
-~...~~~\\
-
-\end{lyxcode}
-
-Make sure you're pointing to the right 'mpif90' wrapper command.
-
-Normally, this message will appear when you are mixing two different
-Fortran compilers. That is, using e.g. gfortran to compile non-MPI
-files and mpif90, wrapper provided for e.g. ifort, to compile MPI-files.
-
-fix: e.g. specify > ./configure FC=gfortran MPIF90=/usr/local/openmpi-gfortran/bin/mpif90
-
-\item [{after executing \texttt{xmeshfem3D} I've got elements with skewness of 81\% percent, what does this mean:}] Look
-at the skewness table printed in the \texttt{output\_mesher.txt} file
-after executing \texttt{xmeshfem3D} for the example given in \texttt{EXAMPLES/meshfem3D\_examples/simple\_model/}:
-
-\begin{lyxcode}
-...~~\\
-~histogram~of~skewness~(0.~good~-~1.~bad):~~~\\
-~0.0000000E+00~-~5.0000001E-02~27648~81.81818~\%~~~\\
-~5.0000001E-02~-~0.1000000~0~0.0000000E+00~\%~~~\\
-~...~~~\\
-
-\end{lyxcode}
-
-The first line means that you have 27,648 elements with a skewness
-value between 0 and 0.05 (which means the element is basically not
-skewed, just plain regular hexahedral element). The total number of
-elements you have in this mesh is (see in the \texttt{output\_mesher.txt}
-file a bit further down):
-\begin{lyxcode}
-...~~\\
-~total~number~of~elements~in~entire~mesh:~33792~~~\\
-~...~~~\\
-
-\end{lyxcode}
-
-which gives you that: 27,648 / 33,792 $\sim$ 81.8 \% of all elements
-are not skewed, i.e. regular elements. a fantastic value :)
-
-The histogram lists for this mesh also some stronger skewed elements,
-for example the worst ones belong to:
-\begin{lyxcode}
-...~~\\
-~0.6000000~-~0.6500000~2048~6.060606~\%~~~\\
-~...~~~\\
-
-\end{lyxcode}
-
-about 6 \% of all elements have distortions with a skewness value
-between 0.6 and 0.65. The skewness values give you a hint of how good
-your mesh is. In an ideal world, you would want to have no distortions,
-just like the 81\% from above. Those elements give you the best approximate
-values by the GLL quadrature used in the spectral-element method.
-However, having weakly distorted elements is still fine and the solutions
-are still accurate enough. So empirically, values up to around 0.7
-are tolerable, above that you should consider remeshing...
-
-To give you an idea why some of the elements are distorted, see the
-following figure \ref{fig:mesh.vp} of the mesh you obtain in the
-example \texttt{EXAMPLES/meshfem3D\_examples/simple\_model/}.
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=0.9\textwidth]{figures/mesh_vp}
-\par\end{centering}
-
-\caption{Paraview visualization using the mesh vtk-files for the example given
-in \texttt{EXAMPLES/meshfem3D\_examples/simple\_model/}.}
-
-\label{fig:mesh.vp}
-\end{figure}
-
-You will see that the mesh contains a doubling layer, where we stitch
-elements together such that the size of two elements will transition
-to the size of one element (very useful to keep the ratio of wavespeed
-/ element\_size about constant). Those elements in this doubling layer
-have higher skewness values and make up those 6 \% in the histogram.
-
-\item [{the code gives following error message "need at least one receiver":}] This
-means that no stations given in the input file \texttt{DATA/STATIONS}
-could be located within the dimensions of the mesh. This can happen
-for example when the mesh was created with the in-house mesher \texttt{xmeshfem3D}
-while using the Universal Transverse Mercator (UTM) projection but
-the simulation with \texttt{xspecfem3D} was suppressing this projection
-from latitude/longitude to x/y/z coordinates.
-
-In such cases, try to change your \texttt{DATA/Par\_file} and set
-e.g.:
-\begin{lyxcode}
-SUPPRESS\_UTM\_PROJECTION~=~.false.~~~\\
-
-\end{lyxcode}
-
-to be the same in \texttt{Mesh\_Par\_file} and \texttt{Par\_file}.
-This flag should be identical when using the in-house mesher \texttt{xmeshfem3D},
-\texttt{xgenerate\_databases} and \texttt{xspecfem3D} together to
-run simulations.
-
-The flag determines if the coordinates you specify for your source
-and station locations are given as lat/lon degrees and must be converted
-to UTM coordinates. As an example, if you use \texttt{.false.} within
-\texttt{Mesh\_Par\_file} then you create a mesh with \texttt{xmeshfem3D}
-using the UTM projection from lat/lon as input format to UTM projected
-coordinates to store the mesh point positions, which is fine. The
-error then may occur if in the \texttt{Par\_file} you have this set
-to \texttt{.true.} so that the \texttt{xgenerate\_databases} and \texttt{xspecfem3D}
-suppress the UTM projection and assume that all coordinates you use
-now for source and receiver locations are given in meters (that is,
-converted) already. So it won't find the specified locations in the
-used mesh. As a solutions, just change the flag in \texttt{Par\_file}
-to be the same as in \texttt{Mesh\_Par\_file} and rerun \texttt{xgenerate\_databases}
-and \texttt{xspecfem3D} to make sure that your simulation works fine.
-
-\item [{I get the following error message "forward simulation became unstable and blew up":}] In
-most cases this means that your time step size \texttt{DT} is chosen
-too big. Look at your files \texttt{output\_mesher.txt} or \texttt{output\_solver.txt}
-created in the folder \texttt{OUTPUT\_FILES}. In these output files,
-find the section:
-
-\begin{lyxcode}
-...~~~\\
-~{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}~~~\\
-~{*}{*}{*}~Verification~of~simulation~parameters~{*}{*}{*}~~~\\
-~{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}~~~\\
-~...~~~\\
-~{*}{*}{*}~Minimum~period~resolved~=~4.308774~~~\\
-~{*}{*}{*}~Maximum~suggested~time~step~=~6.8863556E-02~~~\\
-~...~~~\\
-~...~~~\\
-
-\end{lyxcode}
-
-then change \texttt{DT} in the \texttt{DATA/Par\_file} to be somewhere
-close to the maximum suggested time step. In the example above: \\
- \texttt{DT = 0.05d0} \\
- would (most probably) work fine. It could be also bigger than the
-0.068 s suggested. This depends a bit on the distortions of your mesh
-elements. The more regular they are, the bigger you can choose \texttt{DT}.
-Just play with this value a bit and see when the simulation becomes
-stable ...
-
-\end{description}
-
-\chapter{\label{cha:License}License }
-
-\textbf{GNU GENERAL PUBLIC LICENSE Version 2, June 1991. Copyright
-(C) 1989, 1991 Free Software Foundation, Inc. 59 Temple Place, Suite
-330, Boston, MA 02111-1307 USA} \\
- Everyone is permitted to copy and distribute verbatim copies of this
-license document, but changing it is not allowed.
-
-\section*{Preamble}
-
-The licenses for most software are designed to take away your freedom
-to share and change it. By contrast, the GNU General Public License
-is intended to guarantee your freedom to share and change free software
--- to make sure the software is free for all its users. This General
-Public License applies to most of the Free Software Foundation's software
-and to any other program whose authors commit to using it. (Some other
-Free Software Foundation software is covered by the GNU Library General
-Public License instead.) You can apply it to your programs, too.
-
-When we speak of free software, we are referring to freedom, not price.
-Our General Public Licenses are designed to make sure that you have
-the freedom to distribute copies of free software (and charge for
-this service if you wish), that you receive source code or can get
-it if you want it, that you can change the software or use pieces
-of it in new free programs; and that you know you can do these things.
-
-To protect your rights, we need to make restrictions that forbid anyone
-to deny you these rights or to ask you to surrender the rights. These
-restrictions translate to certain responsibilities for you if you
-distribute copies of the software, or if you modify it.
-
-For example, if you distribute copies of such a program, whether gratis
-or for a fee, you must give the recipients all the rights that you
-have. You must make sure that they, too, receive or can get the source
-code. And you must show them these terms so they know their rights.
-
-We protect your rights with two steps:
-\begin{enumerate}
-\item Copyright the software, and
-\item Offer you this license which gives you legal permission to copy, distribute
-and/or modify the software.
-\end{enumerate}
-Also, for each author's protection and ours, we want to make certain
-that everyone understands that there is no warranty for this free
-software. If the software is modified by someone else and passed on,
-we want its recipients to know that what they have is not the original,
-so that any problems introduced by others will not reflect on the
-original authors' reputations.
-
-Finally, any free program is threatened constantly by software patents.
-We wish to avoid the danger that redistributors of a free program
-will individually obtain patent licenses, in effect making the program
-proprietary. To prevent this, we have made it clear that any patent
-must be licensed for everyone's free use or not licensed at all.
-
-The precise terms and conditions for copying, distribution and modification
-follow.
-
-\section*{GNU GENERAL PUBLIC LICENSE TERMS AND CONDITIONS FOR COPYING, DISTRIBUTION
-AND MODIFICATION }
-\begin{itemize}
-\item 0.This License applies to any program or other work which contains
-a notice placed by the copyright holder saying it may be distributed
-under the terms of this General Public License. The ``Program''
-below refers to any such program or work, and a ``work based on the
-Program'' means either the Program or any derivative work under copyright
-law: that is to say, a work containing the Program or a portion of
-it, either verbatim or with modifications and/or translated into another
-language. (Hereinafter, translation is included without limitation
-in the term ``modification.'') Each licensee is addressed as ``you.''\\
- \\
- Activities other than copying, distribution and modification are
-not covered by this License; they are outside its scope. The act of
-running the Program is not restricted, and the output from the Program
-is covered only if its contents constitute a work based on the Program
-(independent of having been made by running the Program). Whether
-that is true depends on what the Program does.\end{itemize}
-\begin{enumerate}
-\item You may copy and distribute verbatim copies of the Program's source
-code as you receive it, in any medium, provided that you conspicuously
-and appropriately publish on each copy an appropriate copyright notice
-and disclaimer of warranty; keep intact all the notices that refer
-to this License and to the absence of any warranty; and give any other
-recipients of the Program a copy of this License along with the Program.
-
-You may charge a fee for the physical act of transferring a copy,
-and you may at your option offer warranty protection in exchange for
-a fee.
-
-\item You may modify your copy or copies of the Program or any portion of
-it, thus forming a work based on the Program, and copy and distribute
-such modifications or work under the terms of Section 1 above, provided
-that you also meet all of these conditions:
-
-\begin{enumerate}
-\item You must cause the modified files to carry prominent notices stating
-that you changed the files and the date of any change.
-\item You must cause any work that you distribute or publish, that in whole
-or in part contains or is derived from the Program or any part thereof,
-to be licensed as a whole at no charge to all third parties under
-the terms of this License.
-\item If the modified program normally reads commands interactively when
-run, you must cause it, when started running for such interactive
-use in the most ordinary way, to print or display an announcement
-including an appropriate copyright notice and a notice that there
-is no warranty (or else, saying that you provide a warranty) and that
-users may redistribute the program under these conditions, and telling
-the user how to view a copy of this License. (Exception: if the Program
-itself is interactive but does not normally print such an announcement,
-your work based on the Program is not required to print an announcement.)
-\end{enumerate}
-
-These requirements apply to the modified work as a whole. If identifiable
-sections of that work are not derived from the Program, and can be
-reasonably considered independent and separate works in themselves,
-then this License, and its terms, do not apply to those sections when
-you distribute them as separate works. But when you distribute the
-same sections as part of a whole which is a work based on the Program,
-the distribution of the whole must be on the terms of this License,
-whose permissions for other licensees extend to the entire whole,
-and thus to each and every part regardless of who wrote it.
-
-Thus, it is not the intent of this section to claim rights or contest
-your rights to work written entirely by you; rather, the intent is
-to exercise the right to control the distribution of derivative or
-collective works based on the Program.
-
-In addition, mere aggregation of another work not based on the Program
-with the Program (or with a work based on the Program) on a volume
-of a storage or distribution medium does not bring the other work
-under the scope of this License.
-
-\item You may copy and distribute the Program (or a work based on it, under
-Section 2) in object code or executable form under the terms of Sections
-1 and 2 above provided that you also do one of the following:
-
-\begin{enumerate}
-\item Accompany it with the complete corresponding machine-readable source
-code, which must be distributed under the terms of Sections 1 and
-2 above on a medium customarily used for software interchange; or,
-\item Accompany it with a written offer, valid for at least three years,
-to give any third party, for a charge no more than your cost of physically
-performing source distribution, a complete machine-readable copy of
-the corresponding source code, to be distributed under the terms of
-Sections 1 and 2 above on a medium customarily used for software interchange;
-or,
-\item Accompany it with the information you received as to the offer to
-distribute corresponding source code. (This alternative is allowed
-only for noncommercial distribution and only if you received the program
-in object code or executable form with such an offer, in accord with
-Subsection b above.)
-\end{enumerate}
-
-The source code for a work means the preferred form of the work for
-making modifications to it. For an executable work, complete source
-code means all the source code for all modules it contains, plus any
-associated interface definition files, plus the scripts used to control
-compilation and installation of the executable. However, as a special
-exception, the source code distributed need not include anything that
-is normally distributed (in either source or binary form) with the
-major components (compiler, kernel, and so on) of the operating system
-on which the executable runs, unless that component itself accompanies
-the executable.
-
-If distribution of executable or object code is made by offering access
-to copy from a designated place, then offering equivalent access to
-copy the source code from the same place counts as distribution of
-the source code, even though third parties are not compelled to copy
-the source along with the object code.
-
-\item You may not copy, modify, sublicense, or distribute the Program except
-as expressly provided under this License. Any attempt otherwise to
-copy, modify, sublicense or distribute the Program is void, and will
-automatically terminate your rights under this License. However, parties
-who have received copies, or rights, from you under this License will
-not have their licenses terminated so long as such parties remain
-in full compliance.
-\item You are not required to accept this License, since you have not signed
-it. However, nothing else grants you permission to modify or distribute
-the Program or its derivative works. These actions are prohibited
-by law if you do not accept this License. Therefore, by modifying
-or distributing the Program (or any work based on the Program), you
-indicate your acceptance of this License to do so, and all its terms
-and conditions for copying, distributing or modifying the Program
-or works based on it.
-\item Each time you redistribute the Program (or any work based on the Program),
-the recipient automatically receives a license from the original licensor
-to copy, distribute or modify the Program subject to these terms and
-conditions. You may not impose any further restrictions on the recipients'
-exercise of the rights granted herein. You are not responsible for
-enforcing compliance by third parties to this License.
-\item If, as a consequence of a court judgment or allegation of patent infringement
-or for any other reason (not limited to patent issues), conditions
-are imposed on you (whether by court order, agreement or otherwise)
-that contradict the conditions of this License, they do not excuse
-you from the conditions of this License. If you cannot distribute
-so as to satisfy simultaneously your obligations under this License
-and any other pertinent obligations, then as a consequence you may
-not distribute the Program at all. For example, if a patent license
-would not permit royalty-free redistribution of the Program by all
-those who receive copies directly or indirectly through you, then
-the only way you could satisfy both it and this License would be to
-refrain entirely from distribution of the Program.
-
-If any portion of this section is held invalid or unenforceable under
-any particular circumstance, the balance of the section is intended
-to apply and the section as a whole is intended to apply in other
-circumstances.
-
-It is not the purpose of this section to induce you to infringe any
-patents or other property right claims or to contest validity of any
-such claims; this section has the sole purpose of protecting the integrity
-of the free software distribution system, which is implemented by
-public license practices. Many people have made generous contributions
-to the wide range of software distributed through that system in reliance
-on consistent application of that system; it is up to the author/donor
-to decide if he or she is willing to distribute software through any
-other system and a licensee cannot impose that choice.
-
-This section is intended to make thoroughly clear what is believed
-to be a consequence of the rest of this License.
-
-\item If the distribution and/or use of the Program is restricted in certain
-countries either by patents or by copyrighted interfaces, the original
-copyright holder who places the Program under this License may add
-an explicit geographical distribution limitation excluding those countries,
-so that distribution is permitted only in or among countries not thus
-excluded. In such case, this License incorporates the limitation as
-if written in the body of this License.
-\item The Free Software Foundation may publish revised and/or new versions
-of the General Public License from time to time. Such new versions
-will be similar in spirit to the present version, but may differ in
-detail to address new problems or concerns.
-
-Each version is given a distinguishing version number. If the Program
-specifies a version number of this License which applies to it and
-``any later version,'' you have the option of following the terms
-and conditions either of that version or of any later version published
-by the Free Software Foundation. If the Program does not specify a
-version number of this License, you may choose any version ever published
-by the Free Software Foundation.
-
-\item If you wish to incorporate parts of the Program into other free programs
-whose distribution conditions are different, write to the author to
-ask for permission. For software which is copyrighted by the Free
-Software Foundation, write to the Free Software Foundation; we sometimes
-make exceptions for this. Our decision will be guided by the two goals
-of preserving the free status of all derivatives of our free software
-and of promoting the sharing and reuse of software generally.
-\end{enumerate}
-
-\subsection*{NO WARRANTY }
-\begin{itemize}
-\item 11.BECAUSE THE PROGRAM IS LICENSED FREE OF CHARGE, THERE IS NO WARRANTY
-FOR THE PROGRAM, TO THE EXTENT PERMITTED BY APPLICABLE LAW. EXCEPT
-WHEN OTHERWISE STATED IN WRITING THE COPYRIGHT HOLDERS AND/OR OTHER
-PARTIES PROVIDE THE PROGRAM ``AS IS'' WITHOUT WARRANTY OF ANY KIND,
-EITHER EXPRESSED OR IMPLIED, INCLUDING, BUT NOT LIMITED TO, THE IMPLIED
-WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE.
-THE ENTIRE RISK AS TO THE QUALITY AND PERFORMANCE OF THE PROGRAM IS
-WITH YOU. SHOULD THE PROGRAM PROVE DEFECTIVE, YOU ASSUME THE COST
-OF ALL NECESSARY SERVICING, REPAIR OR CORRECTION.
-\item 12.IN NO EVENT UNLESS REQUIRED BY APPLICABLE LAW OR AGREED TO IN WRITING
-WILL ANY COPYRIGHT HOLDER, OR ANY OTHER PARTY WHO MAY MODIFY AND/OR
-REDISTRIBUTE THE PROGRAM AS PERMITTED ABOVE, BE LIABLE TO YOU FOR
-DAMAGES, INCLUDING ANY GENERAL, SPECIAL, INCIDENTAL OR CONSEQUENTIAL
-DAMAGES ARISING OUT OF THE USE OR INABILITY TO USE THE PROGRAM (INCLUDING
-BUT NOT LIMITED TO LOSS OF DATA OR DATA BEING RENDERED INACCURATE
-OR LOSSES SUSTAINED BY YOU OR THIRD PARTIES OR A FAILURE OF THE PROGRAM
-TO OPERATE WITH ANY OTHER PROGRAMS), EVEN IF SUCH HOLDER OR OTHER
-PARTY HAS BEEN ADVISED OF THE POSSIBILITY OF SUCH DAMAGES.
-\end{itemize}
-
-\section*{END OF TERMS AND CONDITIONS }
-
-\subsection*{How to Apply These Terms to Your New Programs}
-
-If you develop a new program, and you want it to be of the greatest
-possible use to the public, the best way to achieve this is to make
-it free software which everyone can redistribute and change under
-these terms.
-
-To do so, attach the following notices to the program. It is safest
-to attach them to the start of each source file to most effectively
-convey the exclusion of warranty; and each file should have at least
-the ``copyright'' line and a pointer to where the full notice is
-found. For example:
-\begin{quote}
-One line to give the program's name and a brief idea of what it does.
-Copyright {\footnotesize $\copyright$ (}year) (name of author)
-
-This program is free software; you can redistribute it and/or modify
-it under the terms of the GNU General Public License as published
-by the Free Software Foundation; either version 2 of the License,
-or (at your option) any later version.
-
-This program is distributed in the hope that it will be useful, but
-WITHOUT ANY WARRANTY; without even the implied warranty of MERCHANTABILITY
-or FITNESS FOR A PARTICULAR PURPOSE. See the GNU General Public License
-for more details.
-
-You should have received a copy of the GNU General Public License
-along with this program; if not, write to the Free Software Foundation,
-Inc., 59 Temple Place, Suite 330, Boston, MA 02111-1307 USA
-\end{quote}
-Also add information on how to contact you by electronic and paper
-mail.
-
-If the program is interactive, make it output a short notice like
-this when it starts in an interactive mode:
-\begin{quote}
-Gnomovision version 69, Copyright $\copyright$ year name of author
-Gnomovision comes with ABSOLUTELY NO WARRANTY; for details type `show
-w'. This is free software, and you are welcome to redistribute it
-under certain conditions; type `show c' for details.
-\end{quote}
-The hypothetical commands `show w' and `show c' should show the appropriate
-parts of the General Public License. Of course, the commands you use
-may be called something other than `show w' and `show c'; they could
-even be mouse-clicks or menu items -- whatever suits your program.
-
-You should also get your employer (if you work as a programmer) or
-your school, if any, to sign a ``copyright disclaimer'' for the
-program, if necessary. Here is a sample; alter the names:
-\begin{quote}
-Yoyodyne, Inc., hereby disclaims all copyright interest in the program
-`Gnomovision' (which makes passes at compilers) written by James Hacker.
-
-(signature of Ty Coon)\\
- 1 April 1989 \\
- Ty Coon, President of Vice
-\end{quote}
-This General Public License does not permit incorporating your program
-into proprietary programs. If your program is a subroutine library,
-you may consider it more useful to permit linking proprietary applications
-with the library. If this is what you want to do, use the GNU Library
-General Public License instead of this License.
-\end{document}

### 2015-04-24 doc/USER_MANUAL/13_changing_the_model.tex
updates user manual to use facilitate the generation of a markdown version to the github wiki

--- 2015-04-18 doc/USER_MANUAL/manual_SPECFEM3D_Cartesian.tex
@@
-% !TEX encoding = IsoLatin
-%% LyX 2.0.2 created this file.  For more info, see http://www.lyx.org/.
-%% Do not edit unless you really know what you are doing.
-\documentclass[oneside,english]{book}
-\usepackage[T1]{fontenc}
-%% DK DK we do not accept accented letters, use the old LaTeX syntax
-%% DK DK for them instead: \'e \`e  \"e \^e and so on.
-%% DK DK \usepackage[latin1]{inputenc}
-\usepackage[letterpaper]{geometry}
-\geometry{verbose,tmargin=1in,bmargin=1in,lmargin=1in,rmargin=1in}
-\setcounter{secnumdepth}{3}
-\setcounter{tocdepth}{3}
-\usepackage{babel}
-\usepackage{float}
-\usepackage{textcomp}
-\usepackage{url}
-\usepackage{amsmath}
-\usepackage{amssymb}
-\usepackage{graphicx}
-\usepackage[authoryear]{natbib}
-\usepackage[unicode=true,pdfusetitle,
- bookmarks=true,bookmarksnumbered=true,bookmarksopen=false,
- breaklinks=false,pdfborder={0 0 0},backref=section,colorlinks=false,pdfpagemode=None]
- {hyperref}
-\hypersetup{
- pdftex,pdfstartview=FitH,pdfpagelayout=SinglePage}
-\usepackage{breakurl}
-
-\makeatletter
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% LyX specific LaTeX commands.
-\newcommand{\noun}[1]{\textsc{#1}}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Textclass specific LaTeX commands.
-\newenvironment{lyxcode}
-{\par\begin{list}{}{
-\setlength{\rightmargin}{\leftmargin}
-\setlength{\listparindent}{0pt}% needed for AMS classes
-\raggedright
-\setlength{\itemsep}{0pt}
-\setlength{\parsep}{0pt}
-\normalfont\ttfamily}%
- \item[]}
-{\end{list}}
-\newenvironment{lyxlist}[1]
-{\begin{list}{}
-{\settowidth{\labelwidth}{#1}
- \setlength{\leftmargin}{\labelwidth}
- \addtolength{\leftmargin}{\labelsep}
- \renewcommand{\makelabel}[1]{##1\hfil}}}
-{\end{list}}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% User specified LaTeX commands.
-%\renewcommand{\baselinestretch}{1.5}
-
-% figures
-\usepackage[dvips]{epsfig}
-
-% fonts
-\usepackage{times}
-
-% hyperlinks to sections and references
-
-\newcommand{\urlwithparentheses}[1]{(\url{#1})}
-
-% biblio GJI
-\bibliographystyle{abbrvnat}
-
-\newcommand{\nexxi}{\mbox{\texttt{NEX\_XI\/}}}
-\newcommand{\nexeta}{\mbox{\texttt{NEX\_ETA\/}}}
-\newcommand{\nprocxi}{\mbox{\texttt{NPROC\_XI\/}}}
-\newcommand{\nproceta}{\mbox{\texttt{NPROC\_ETA\/}}}
-\newcommand{\nchunks}{\mbox{\texttt{NCHUNKS\/}}}
-\newcommand{\eg}{e.g.,}
-
-% colors for the text
-\@ifundefined{definecolor}
- {\usepackage{color}}{}
-\newcommand{\red}[1]{\textcolor{Red}{#1}}
-
-\usepackage{babel}
-
-\makeatother
-
-\begin{document}
-\begin{center}
-\thispagestyle{empty}\vspace*{-1.8truecm} %
-\makebox[1\textwidth]{%
-\includegraphics[width=0.83\paperwidth]{figures/specfem_3d_Cartesian-cover} %
-}
-\par\end{center}
-
-\title{\textbf{SPECFEM3D Cartesian}\\
- \textbf{User Manual}}
-
-\author{$\copyright$ Princeton University (USA), CNRS and University of Marseille (France),\\
-and ETH Z\"urich (Switzerland)\\
-Version 3.0}
-
-\date{\today}
-
-\maketitle
-
-\section*{Authors}
-
-\noindent The SPECFEM3D Cartesian package was first developed by Dimitri
-Komatitsch and Jean-Pierre Vilotte at Institut de Physique du Globe
-(IPGP) in Paris, France from 1995 to 1997 and then by Dimitri Komatitsch
-and Jeroen Tromp at Harvard University and Caltech, USA, starting in 1998.
-The story started on April 4, 1995, when Prof. Yvon Maday from CNRS and University of Paris, France, gave a lecture to
-Dimitri Komatitsch and Jean-Pierre Vilotte at IPG about the nice properties of the Legendre spectral-element method with diagonal mass matrix that he had used for
-other equations. We are deeply indebted and thankful to him for that.
-That followed a visit by Dimitri Komatitsch to OGS (Istituto Nazionale di Oceanografia e di Geofisica Sperimentale) in Trieste, Italy, in February 1995
-to meet with G\'eza Seriani and Enrico Priolo, who introduced him to their 2D Chebyshev version of the spectral-element method with a non-diagonal mass matrix.
-We are deeply indebted and thankful to them for that.
-
-%%%%%%%% please do NOT merge the lines of the paragraph below, it makes it difficult to insert new names in alphabetical order; leave a single name per line
-%%%%%%%% please do NOT merge the lines of the paragraph below, it makes it difficult to insert new names in alphabetical order; leave a single name per line
-%%%%%%%% please do NOT merge the lines of the paragraph below, it makes it difficult to insert new names in alphabetical order; leave a single name per line
-Since then it has been developed and maintained by a development team:
-in alphabetical order,
-Jean-Paul (Pablo) Ampuero,
-Kangchen Bai,
-Piero Basini,
-C\'eline Blitz,
-Alexis Bottero,
-Ebru Bozda\u{g},
-Emanuele Casarotti,
-Joseph Charles,
-Min Chen,
-Paul Cristini,
-Percy Galvez,
-Dominik G\"oddeke,
-Vala Hj\"orleifsd\'ottir,
-Sue Kientz,
-Dimitri Komatitsch,
-Jes\'us Labarta,
-Nicolas Le Goff,
-Pieyre Le Loher,
-Matthieu Lefebvre,
-Qinya Liu,
-Yang Luo,
-Alessia Maggi,
-Federica Magnoni,
-Roland Martin,
-Ren\'e Matzen,
-Dennis McRitchie,
-Matthias Meschede,
-Peter Messmer,
-David Mich\'ea,
-Surendra Nadh Somala,
-Tarje Nissen-Meyer,
-Daniel Peter,
-Max Rietmann,
-Elliott Sales de Andrade,
-Brian Savage,
-Bernhard Schuberth,
-Anne Sieminski,
-Leif Strand,
-Carl Tape,
-Jeroen Tromp,
-Jean-Pierre Vilotte,
-Zhinan Xie,
-Chang-Hua Zhang,
-Hejun Zhu.\\
-
-The cover graphic of the manual was created by Santiago Lombeyda from
-Caltech's Center for Advanced Computing Research (CACR) \urlwithparentheses{http://www.cacr.caltech.edu}.\\
-
-The code is released open-source under the GNU version 2+ license, see the license at the end of this manual.\\
-
-\newpage{}
-\section*{Current and past main participants or main sponsors of the SPECFEM project (in no particular order)}
-%
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=0.125\textwidth]{figures/logo_cnrs}\vspace*{2truemm}
-\includegraphics[width=0.125\textwidth]{figures/logo_princeton}\vspace*{2truemm}
-\includegraphics[width=0.15\textwidth]{figures/logo_aix_marseille_universite}\vspace*{0.02truemm}
-\includegraphics[width=0.15\textwidth]{figures/logo_ETH}\vspace*{2truemm}
-\includegraphics[width=0.12\textwidth]{figures/logo_CSC_China}\vspace*{0.02truemm}
-\includegraphics[width=0.15\textwidth]{figures/logo_inria}\vspace*{2truemm}
-\includegraphics[width=0.14\textwidth]{figures/logo_UPPA}
-\par\end{centering}
-%
-\vspace*{-2truemm}
-%
-\noindent \begin{centering}
-\includegraphics[width=0.15\textwidth]{figures/logo_NVIDIA}\vspace*{2truemm}
-\includegraphics[width=0.125\textwidth]{figures/logo_IUF}\vspace*{2truemm}
-\includegraphics[width=0.135\textwidth]{figures/logo_Caltech}\vspace*{2truemm}
-\includegraphics[width=0.135\textwidth]{figures/logo_Harvard}\vspace*{2truemm}
-\includegraphics[width=0.135\textwidth]{figures/logo_IPGP}\vspace*{2truemm}
-\includegraphics[width=0.112\textwidth]{figures/logo_ANR}\vspace*{2truemm}
-\includegraphics[width=0.112\textwidth]{figures/logo_NSF}\vspace*{2truemm}
-\par\end{centering}
-%
-\vspace*{-2truemm}
-%
-\noindent \begin{centering}
-\includegraphics[width=0.190\textwidth]{figures/logo_European_Union}\vspace*{2truemm}
-\includegraphics[width=0.112\textwidth]{figures/logo_GENCI}\vspace*{2truemm}
-\includegraphics[width=0.112\textwidth]{figures/logo_PRACE}\vspace*{2truemm}
-\includegraphics[width=0.112\textwidth]{figures/logo_CINES}\vspace*{2truemm}
-\includegraphics[width=0.130\textwidth]{figures/logo_Oak_Ridge}\vspace*{2truemm}
-\hspace*{3mm}\includegraphics[width=0.130\textwidth]{figures/logo_fondation_Del_Duca}
-\includegraphics[width=0.130\textwidth]{figures/logo_CIG}\vspace*{2truemm}
-\par\end{centering}
-%
-\vspace*{-2truemm}
-%
-\noindent \begin{centering}
-\includegraphics[width=0.112\textwidth]{figures/logo_University_of_Toronto}\vspace*{2truemm}
-\includegraphics[width=0.109\textwidth]{figures/logo_INGV}\vspace*{2truemm}
-\includegraphics[width=0.102\textwidth]{figures/logo_Univ_Toulouse}\vspace*{2truemm}
-\includegraphics[width=0.112\textwidth]{figures/logo_TOTAL}\vspace*{2truemm}
-\includegraphics[width=0.130\textwidth]{figures/logo_Fairbanks}\vspace*{2truemm}
-\includegraphics[width=0.120\textwidth]{figures/logo_CINECA}\vspace*{2truemm}
-\includegraphics[width=0.140\textwidth]{figures/logo_Intel_Exascale_Labs}\vspace*{2truemm}
-\includegraphics[width=0.130\textwidth]{figures/logo_Maison_Simulation}
-\par\end{centering}
-\end{figure}
-
-\newpage{}
-
-\tableofcontents{}
-
-\chapter{Introduction}
-
-The software package SPECFEM3D Cartesian simulates seismic wave propagation
-at the local or regional scale and performs full waveform imaging (FWI) or adjoint tomography based upon the spectral-element method
-(SEM). The SEM is a continuous Galerkin technique \citep{TrKoLi08,PeKoLuMaLeCaLeMaLiBlNiBaTr11},
-which can easily be made discontinuous \citep{BeMaPa94,Ch00,KoWoHu02,ChCaVi03,LaWaBe05,Kop06,WiStBuGh10,AcKo11};
-it is then close to a particular case of the discontinuous Galerkin
-technique \citep{ReHi73,LeRa74,Arn82,JoPi86,BoMaHe91,FaRi99,HuHuRa99,CoKaSh00,GiHeWa02,RiWh03,MoRi05,GrScSc06,AiMoMu06,BeLaPi06,DuKa06,DeSeWh08,PuAmKa09,WiStBuGh10,DeSe10,EtChViGl10},
-with optimized efficiency because of its tensorized basis functions
-\citep{WiStBuGh10,AcKo11}. In particular, it can accurately handle
-very distorted mesh elements \citep{OlSe11}.\\
-
-It has very good accuracy and convergence properties \citep{MaPa89,SePr94,DeFiMu02,Coh02,DeSe07,SeOl08,AiWa09,AiWa10,MeStTh12}.
-The spectral element approach admits spectral rates of convergence
-and allows exploiting $hp$-convergence schemes. It is also very well
-suited to parallel implementation on very large supercomputers \citep{KoTsChTr03,TsKoChTr03,KoLaMi08a,CaKoLaTiMiLeSnTr08,KoViCh10}
-as well as on clusters of GPU accelerating graphics cards \citep{Kom11,MiKo10,KoMiEr09,KoErGoMi10}.
-Tensor products inside each element can be optimized to reach very
-high efficiency \citep{DeFiMu02}, and mesh point and element numbering
-can be optimized to reduce processor cache misses and improve cache
-reuse \citep{KoLaMi08a}. The SEM can also handle triangular (in 2D)
-or tetrahedral (in 3D) elements \citep{WinBoyd96,TaWi00,KoMaTrTaWi01,Coh02,MeViSa06}
-as well as mixed meshes, although with increased cost and reduced
-accuracy in these elements, as in the discontinuous Galerkin method.\\
-
-Note that in many geological models in the context of seismic wave
-propagation studies (except for instance for fault dynamic rupture
-studies, in which very high frequencies or supershear rupture need
-to be modeled near the fault, see e.g. \citet{BeGlCrViPi07,BeGlCrVi09,PuAmKa09,TaCrEtViBeSa10})
-a continuous formulation is sufficient because material property contrasts
-are not drastic and thus conforming mesh doubling bricks can efficiently
-handle mesh size variations \citep{KoTr02a,KoLiTrSuStSh04,LeChLiKoHuTr08,LeChKoHuTr09,LeKoHuTr09}.\\
-
-For a detailed introduction to the SEM as applied to regional seismic
-wave propagation, please consult \citet{PeKoLuMaLeCaLeMaLiBlNiBaTr11,TrKoLi08,KoVi98,KoTr99,ChKoViCaVaFe07}
-and in particular \citet{LeKoHuTr09,LeChKoHuTr09,LeChLiKoHuTr08,GoAmTaCaSmSaMaKo09,WiKoScTr04,KoLiTrSuStSh04}.
-A detailed theoretical analysis of the dispersion
-and stability properties of the SEM is available in \citet{Coh02}, \citet{DeSe07}, \citet{SeOl07}, \citet{SeOl08} and \citet{MeStTh12}.\\
-
-Effects due to lateral variations in compressional-wave speed, shear-wave
-speed, density, a 3D crustal model, topography and bathymetry are
-included. The package can accommodate full 21-parameter anisotropy
-(see~\citet{ChTr07}) as well as lateral variations in attenuation
-\citep{SaKoTr10}. Adjoint capabilities and finite-frequency kernel
-simulations are included \citep{TrKoLi08,PeKoLuMaLeCaLeMaLiBlNiBaTr11,LiTr06,FiIgBuKe09,ViOp09}.\\
-
-The SEM was originally developed in computational fluid dynamics \citep{Pat84,MaPa89}
-and has been successfully adapted to address problems in seismic wave
-propagation. Early seismic wave propagation applications of the SEM,
-utilizing Legendre basis functions and a perfectly diagonal mass matrix,
-include \citet{CoJoTo93}, \citet{Kom97}, \citet{FaMaPaQu97}, \citet{CaGa97},
-\citet{KoVi98} and \citet{KoTr99}, whereas applications involving
-Chebyshev basis functions and a nondiagonal mass matrix include \citet{SePr94},
-\citet{PrCaSe94} and \citet{SePrPr95}.\\
-
-All SPECFEM3D software is written in Fortran2003 with full portability
-in mind, and conforms strictly to the Fortran2003 standard. It uses
-no obsolete or obsolescent features of Fortran. The package uses parallel
-programming based upon the Message Passing Interface (MPI) \citep{GrLuSk94,Pac97}.\\
-
-SPECFEM3D won the Gordon Bell award for best performance at the SuperComputing~2003
-conference in Phoenix, Arizona (USA) (see \citet{KoTsChTr03} and
-\url{www.sc-conference.org/sc2003/nrfinalaward.html}). It was a finalist
-again in 2008 for a run at 0.16 petaflops (sustained) on 149,784 processors
-of the `Jaguar' Cray XT5 system at Oak Ridge National Laboratories
-(USA) \citep{CaKoLaTiMiLeSnTr08}. It also won the BULL Joseph Fourier
-supercomputing award in 2010.\\
-
-It reached the sustained one petaflop performance level for the first time in February 2013
-on the Blue Waters Cray supercomputer at the National Center for Supercomputing Applications (NCSA), located at the University of Illinois at Urbana-Champaign (USA).\\
-
-This new release of the code includes Convolution or
-Auxiliary Differential Equation Perfectly Matched absorbing Layers
-(C-PML or ADE-PML) \citep{MaKoEz08,MaKoGe08,MaKo09,MaKoGeBr10,KoMa07}.
-It also includes support for GPU graphics card acceleration \citep{Kom11,MiKo10,KoMiEr09,KoErGoMi10}.\\
-
-The next release of the code will use the PT-SCOTCH parallel and threaded
-version of SCOTCH for mesh partitioning instead of the serial version.\\
-
-SPECFEM3D Cartesian includes coupled fluid-solid domains and adjoint
-capabilities, which enables one to address seismological inverse problems,
-but for linear rheologies only so far. To accommodate visco-plastic
-or non-linear rheologies, the reader can refer to the \href{http://geoelse.stru.polimi.it/}{GeoELSE}software
-package \citep{CaGa97,StPaIg09}.
-
-\section{Citation}
-
-If you use SPECFEM3D Cartesian for your own research, please cite
-at least one of the following articles:
-\begin{description}
-\item [{Numerical simulations in general}] ~\\
- Forward and adjoint simulations are described in detail in \citet{TrKoLi08,PeKoLuMaLeCaLeMaLiBlNiBaTr11,VaCaSaKoVi99,KoMiEr09,KoErGoMi10,ChKoViCaVaFe07,MaKoDi09,KoViCh10,CaKoLaTiMiLeSnTr08,TrKoHjLiZhPeBoMcFrTrHu10,KoRiTr02,KoTr02a,KoTr02b,KoTr99}
-or \citet{KoVi98}. Additional aspects of adjoint simulations are
-described in \citet{TrTaLi05,LiTr06,TrKoLi08,LiTr08,TrKoHjLiZhPeBoMcFrTrHu10,PeKoLuMaLeCaLeMaLiBlNiBaTr11}.
-Domain decomposition is explained in detail in \citet{MaKoBlLe08},
-and excellent scaling up to 150,000 processor cores is shown for instance
-in \citet{CaKoLaTiMiLeSnTr08,KoLaMi08a,MaKoBlLe08,KoErGoMi10,Kom11},
-\item [{GPU computing}] ~\\
- Computing on GPU graphics cards for acoustic or seismic wave propagation
-applications is described in detail in \citet{Kom11,MiKo10,KoMiEr09,KoErGoMi10}.
-\end{description}
-\noindent If you use this new version, which has non blocking
-MPI for much better performance for medium or large runs, please cite
-at least one of these six articles, in which results of non blocking
-MPI runs are presented: \citet{PeKoLuMaLeCaLeMaLiBlNiBaTr11,KoErGoMi10,KoViCh10,Kom11,CaKoLaTiMiLeSnTr08,MaKoBlLe08}.\\
-
-\noindent If you work on geophysical applications, you may be interested
-in citing some of these application articles as well, among others:
-\begin{description}
-\item [{Southern California simulations}] ~\\
- \citet{KoLiTrSuStSh04,KrJiKoTr06a,KrJiKoTr06b}.
-
-If you use the 3D southern California model, please cite \citet{SuSh03}
-(Los Angeles model), \citet{lovelyetal06} (Salton Trough), and \citet{hauksson2000}
-(southern California). The Moho map was determined by \citet{zhu&kanamori2000}.
-The 1D SoCal model was developed by \citet{DrHe90}.
-
-\item [{Anisotropy}] ~\\
- \citet{ChTr07,JiTsKoTr05,ChFaKo04,FaChKo04,RiRiKoTrHe02,TrKo00}.
-\item [{Attenuation}] ~\\
- \citet{SaKoTr10,KoTr02a,KoTr99}.
-\item [{Topography}] ~\\
- \citet{LeKoHuTr09,LeChKoHuTr09,LeChLiKoHuTr08,GoAmTaCaSmSaMaKo09,WiKoScTr04}.
-\end{description}
-The corresponding Bib\TeX{} entries may be found in file \texttt{doc/USER\_MANUAL/bibliography.bib}.
-
-\section{Support}
-
-This material is based upon work supported by the USA National Science
-Foundation under Grants No. EAR-0406751 and EAR-0711177, by the French
-CNRS, French INRIA Sud-Ouest MAGIQUE-3D, French ANR NUMASIS under
-Grant No. ANR-05-CIGC-002, and European FP6 Marie Curie International
-Reintegration Grant No. MIRG-CT-2005-017461. Any opinions, findings,
-and conclusions or recommendations expressed in this material are
-those of the authors and do not necessarily reflect the views of the
-USA National Science Foundation, CNRS, INRIA, ANR or the European
-Marie Curie program.
-
-\chapter{\label{cha:Getting-Started}Getting Started}
-
-To download the SPECFEM3D\_Cartesian software package, type this:
-\begin{lyxcode}
-git clone --recursive --branch devel https://github.com/geodynamics/specfem3d.git
-\end{lyxcode}
-{\small \par}
-
-We recommend that you add {\texttt{ulimit -S -s unlimited}} to your
-{\texttt{.bash\_profile}} file and/or {\texttt{limit stacksize
-unlimited }} to your {\texttt{.cshrc}} file to suppress any potential
-limit to the size of the Unix stack.\\
-
-Then, to configure the software for your system, run the \texttt{configure}
-shell script. This script will attempt to guess the appropriate configuration
-values for your system. However, at a minimum, it is recommended that
-you explicitly specify the appropriate command names for your Fortran
-compiler and MPI package (another option is to define FC, CC and MPIF90
-in your .bash\_profile or your .cshrc file):
-\begin{lyxcode}
-./configure~FC=ifort~MPIFC=mpif90
-\end{lyxcode}
-Note that MPI must be installed with MPI-IO enabled because parts of SPECFEM3D perform I/Os through MPI-IO.
-
-Before running the \texttt{configure} script, you should probably
-edit file \texttt{flags.guess} to make sure that it contains the best
-compiler options for your system. Known issues or things to check
-are:
-\begin{description}
-\item [{\texttt{Intel ifort compiler}}] See if you need to add \texttt{-assume byterecl} for your machine.
-\textbf{In the case of that compiler, we have noticed that versions dot zero sometimes have bugs or issues that can lead to wrong results when running the code, thus we \emph{strongly} recommend using versions dot one or above (for instance version 13.1 instead of 13.0, version 14.1 instead of 14.0 and so on)}.
-\item [{\texttt{IBM compiler}}] See if you need to add \texttt{-qsave}
-or \texttt{-qnosave} for your machine.
-\item [{\texttt{Mac OS}}] You will probably need to install \texttt{XCODE}.
-\end{description}
-When compiling on an IBM machine with the \texttt{xlf} and \texttt{xlc}
-compilers, we suggest running the \texttt{configure} script with the
-following options:\\
-
-\noindent \texttt{./configure FC=xlf90\_r MPIFC=mpif90 CC=xlc\_r CFLAGS=\textquotedbl{}-O3
--q64\textquotedbl{} FCFLAGS=\textquotedbl{}-O3 -q64\textquotedbl{}
--with-scotch-dir=...}.\\
-
-On SGI systems, \texttt{flags.guess} automatically informs \texttt{configure}
-to insert `\texttt{`TRAP\_FPE=OFF}'' into the generated \texttt{Makefile}
-in order to turn underflow trapping off.\\
-
-You can add \texttt{--enable-vectorization} to the configuration options to speed up the code in the fluid (acoustic) and elastic parts.
-This works fine if (and only if) your computer always allocates a contiguous memory block for each allocatable array;
-this is the case for most machines and most compilers, but not all. To disable this feature, use option \texttt{--disable-vectorization}.
-For more details see \href{https://github.com/geodynamics/specfem3d/issues/81}{github.com/geodynamics/specfem3d/issues/81} .
-To check if that option works fine on your machine, run the code with and without it for an acoustic/elastic model and make sure the seismograms are identical.\\
-
-Note that we use CUBIT (now called Trelis) to create meshes of hexahedra, but other packages
-can be used as well, for instance GiD from \url{http://gid.cimne.upc.es}
-or Gmsh from \url{http://geuz.org/gmsh} \citep{GeRe09}. Even mesh
-creation packages that generate tetrahedra, for instance TetGen from
-\url{http://tetgen.berlios.de}, can be used because each tetrahedron
-can then easily be decomposed into four hexahedra as shown in the
-picture of the TetGen logo at \url{http://tetgen.berlios.de/figs/Delaunay-Voronoi-3D.gif};
-while this approach does not generate hexahedra of optimal quality,
-it can ease mesh creation in some situations and it has been shown
-that the spectral-element method can very accurately handle distorted
-mesh elements \citep{OlSe11}.
-
-The SPECFEM3D Cartesian software package relies on the SCOTCH library
-to partition meshes created with CUBIT. METIS \citep{KaKu98a,KaKu98c,KaKu98b}
-can also be used instead of SCOTCH if you prefer, by editing file
-src/decompose\_mesh/decompose\_mesh.F90 and uncommenting the flag
-USE\_METIS\_INSTEAD\_OF\_SCOTCH. You will also then need to install
-and compile Metis version 4.0 (do {*}NOT{*} install Metis version
-5.0, which has incompatible function calls) and edit \texttt{src/decompose\_mesh/Makefile.in}
-and uncomment the METIS link flag in that file before running \texttt{configure}.\\
-
-The SCOTCH library \citep{PeRo96} provides efficent static mapping,
-graph and mesh partitioning routines. SCOTCH is a free software package
-developed by Fran\c{c}ois Pellegrini et al. from LaBRI and INRIA in Bordeaux,
-France, downloadable from the web page \url{https://gforge.inria.fr/projects/scotch/}.
-In case no SCOTCH libraries can be found on the system, the configuration
-will bundle the version provided with the source code for compilation.
-The path to an existing SCOTCH installation can to be set explicitly
-with the option \texttt{-{}-with-scotch-dir}. Just as an example:
-\begin{lyxcode}
-./configure~FC=ifort~MPIFC=mpif90~-{}-with-scotch-dir=/opt/scotch
-\end{lyxcode}
-If you use the Intel ifort compiler to compile the code, we recommend
-that you use the Intel icc C compiler to compile Scotch, i.e., use:
-\begin{lyxcode}
-./configure~CC=icc~FC=ifort~MPIFC=mpif90
-\end{lyxcode}
-
-When compiling the SCOTCH source code, if you get a message such as: "ld: cannot find -lz",
-the Zlib compression development library is probably missing on your machine and you will need to install it or ask your system administrator to
-do so. On Linux machines the package is often called "zlib1g-dev" or similar. (thus "sudo apt-get install zlib1g-dev" would install it)
-
-To compile a serial version of the code for small meshes that fits
-on one compute node and can therefore be run serially, run \texttt{configure}
-with the \texttt{-{}-without-mpi} option to suppress all calls to
-MPI.
-
-A summary of the most important configuration variables follows.
-\begin{description}
-\item [{\texttt{F90}}] Path to the Fortran compiler.
-\item [{\texttt{MPIF90}}] Path to MPI Fortran.
-\item [{\texttt{MPI\_FLAGS}}] Some systems require this flag to link to
-MPI libraries.
-\item [{\texttt{FLAGS\_CHECK}}] Compiler flags.
-\end{description}
-The configuration script automatically creates for each executable
-a corresponding \texttt{Makefile} in the \texttt{src/} subdirectoy.
-The \texttt{Makefile} contains a number of suggested entries for various
-compilers, e.g., Portland, Intel, Absoft, NAG, and Lahey. The software
-has run on a wide variety of compute platforms, e.g., various PC clusters
-and machines from Sun, SGI, IBM, Compaq, and NEC. Select the compiler
-you wish to use on your system and choose the related optimization
-flags. Note that the default flags in the \texttt{Makefile} are undoubtedly
-not optimal for your system, so we encourage you to experiment with
-these flags and to solicit advice from your systems administrator.
-Selecting the right compiler and optimization flags can make a tremendous
-difference in terms of performance. We welcome feedback on your experience
-with various compilers and flags.
-
-Now that you have set the compiler information, you need to select
-a number of flags in the \texttt{constants.h} file depending on your
-system:
-\begin{description}
-\item [{\texttt{LOCAL\_PATH\_IS\_ALSO\_GLOBAL}}] Set to \texttt{.false.}
-on most cluster applications. For reasons of speed, the (parallel)
-distributed database generator typically writes a (parallel) database
-for the solver on the local disks of the compute nodes. Some systems
-have no local disks, e.g., BlueGene or the Earth Simulator, and other
-systems have a fast parallel file system, in which case this flag
-should be set to \texttt{.true.}. Note that this flag is not used
-by the database generator or the solver; it is only used for some
-of the post-processing.
-\end{description}
-The package can run either in single or in double precision mode.
-The default is single precision because for almost all calculations
-performed using the spectral-element method using single precision
-is sufficient and gives the same results (i.e. the same seismograms);
-and the single precision code is faster and requires exactly half
-as much memory. Select your preference by selecting the appropriate
-setting in the \texttt{constants.h} file:
-\begin{description}
-\item [{\texttt{CUSTOM\_REAL}}] Set to \texttt{SIZE\_REAL} for single precision
-and \texttt{SIZE\_DOUBLE} for double precision.
-\end{description}
-In the \texttt{precision.h} file:
-\begin{description}
-\item [{\texttt{CUSTOM\_MPI\_TYPE}}] Set to \texttt{MPI\_REAL} for single
-precision and \texttt{MPI\_DOUBLE\_PRECISION} for double precision.
-\end{description}
-On many current processors (e.g., Intel, AMD, IBM Power), single precision
-calculations are significantly faster; the difference can typically
-be 10\% to 25\%. It is therefore better to use single precision. What
-you can do once for the physical problem you want to study is run
-the same calculation in single precision and in double precision on
-your system and compare the seismograms. If they are identical (and
-in most cases they will), you can select single precision for your
-future runs.\\
-
-If your compiler has problems with the \texttt{use mpi} statements that are used in the code, use the script called
-\texttt{replace\_use\_mpi\_with\_include\_mpif\_dot\_h.pl} in the root directory to replace all of them with \texttt{include 'mpif.h'} automatically.
-
-\section{Adding OpenMP support in addition to MPI}
-
-NOTE FROM JULY 2013: OpenMP support is maybe / probably not maintained any more.
-Thus the section below is maybe obsolete.
-
-OpenMP support can be enabled in addition to MPI. However, in many
-cases performance will not improve because our pure MPI implementation
-is already heavily optimized and thus the resulting code will in fact
-be slightly slower. A possible exception could be IBM BlueGene-type
-architectures.
-
-To enable OpenMP, uncomment the OpenMP compiler option in two lines
-in file \texttt{src/specfem3D/Makefile.in} (before running \texttt{configure})
-and also uncomment the \texttt{\#define USE\_OPENMP} statement in
-file\\
- \texttt{src/specfem3D/specfem3D.F90}.
-
-The DO-loop using OpenMP threads has a SCHEDULE property. The OMP\_SCHEDULE
-environment variable can set the scheduling policy of that DO-loop.
-Tests performed by Marcin Zielinski at SARA (The Netherlands) showed
-that often the best scheduling policy is DYNAMIC with the size of
-the chunk equal to the number of OpenMP threads, but most preferably
-being twice as the number of OpenMP threads (thus chunk size = 8 for
-4 OpenMP threads etc). If OMP\_SCHEDULE is not set or is empty, the
-DO-loop will assume generic scheduling policy, which will slow down
-the job quite a bit.
-
-\section{Compiling on an IBM BlueGene}
-
-\underline{More recent installation instruction for IBM BlueGene, from April 2013:}\\
-
-\noindent Edit file \texttt{flags.guess} and put this for \texttt{FLAGS\_CHECK}:\\
-
-\noindent \texttt{-g -qfullpath -O2 -qsave -qstrict -qtune=qp -qarch=qp
--qcache=auto -qhalt=w}\\
- \texttt{-qfree=f90 -qsuffix=f=f90 -qlanglvl=95pure -Q -Q+rank,swap\_all
--Wl,-relax}\\
-
-\noindent The most relevant are the -qarch and -qtune flags, otherwise
-if these flags are set to ``auto'' then they are wrongly assigned
-to the architecture of the frond-end node, which is different from
-that on the compute nodes. You will need to set these flags to the
-right architecture for your BlueGene compute nodes, which is not necessarily
-``qp''; ask your system administrator. On some machines if is necessary
-to use -O2 in these flags instead of -O3 due to a compiler bug of
-the XLF version installed. We thus suggest to first try -O3, and then
-if the code does not compile or does not run fine then switch back
-to -O2. The debug flags (-g, -qfullpath) do not influence performance
-but are useful to get at least some insights in case of problems.\\
-
-\noindent Before running \texttt{configure}, select the XL Fortran
-compiler by typing \texttt{module load bgq-xl/1.0} or \texttt{module
-load bgq-xl} (another, less efficient option is to load the GNU compilers
-using \texttt{module load bgq-gnu/4.4.6} or similar).\\
-
-\noindent Then, to configure the code, type this:\\
-
-\noindent \texttt{./configure FC=bgxlf90\_r MPIFC=mpixlf90\_r CC=bgxlc\_r
-LOCAL\_PATH\_IS\_ALSO\_GLOBAL=true}\\
-
-In order for the SCOTCH domain decomposer to compile, on some (but
-not all) Blue Gene systems you may need to run \texttt{configure}
-with \texttt{CC=gcc} instead of \texttt{CC=bgxlc\_r}.
-
-\noindent \underline{Older installation instruction for IBM BlueGene, from 2011:}\\
-
-\noindent To compile the code on an IBM BlueGene, Laurent L\'eger from
-IDRIS, France, suggests the following: compile the code with
-
-\noindent \texttt{ FLAGS\_CHECK=\textquotedbl{}-O3 -qsave -qstrict
--qtune=auto -qarch=450d -qcache=auto }~\\
- \texttt{ -qfree=f90 -qsuffix=f=f90 -g -qlanglvl=95pure -qhalt=w -Q
--Q+rank,swap\_all -Wl,-relax\textquotedbl{}}
-
-\noindent Option \textquotedbl{}-Wl,-relax\textquotedbl{} must be
-added on many (but not all) BlueGene systems to be able to link the
-binaries \texttt{xmeshfem3D} and \texttt{xspecfem3D} because the final
-link step is done by the GNU \texttt{ld} linker even if one uses \texttt{FC=bgxlf90\_r,
-MPIFC=mpixlf90\_r} and \texttt{CC=bgxlc\_r} to create all the object
-files. On the contrary, on some BlueGene systems that use the native
-AIX linker option \textquotedbl{}-Wl,-relax\textquotedbl{} can lead
-to problems and must be suppressed from \texttt{flags.guess}.
-
-\noindent Also, \texttt{AR=ar, ARFLAGS=cru} and \texttt{RANLIB=ranlib}
-are hardwired in all \texttt{Makefile.in} files by default, but to
-cross-compile on BlueGene/P one needs to change these values to \texttt{AR=bgar,
-ARFLAGS=cru} and \texttt{RANLIB=bgranlib}. Thus the easiest thing
-to do is to modify all \texttt{Makefile.in} files and the \texttt{configure}
-script to set them automatically by \texttt{configure}. One then just
-needs to pass the right commands to the \texttt{configure} script:
-
-\noindent \texttt{ ./configure -{}-prefix=/path/to/SPECFEM3DG\_SP
--{}-host=Babel -{}-build=BGP }~\\
- \texttt{ FC=bgxlf90\_r MPIFC=mpixlf90\_r CC=bgxlc\_r AR=bgar ARFLAGS=cru
-}~\\
- \texttt{ RANLIB=bgranlib LOCAL\_PATH\_IS\_ALSO\_GLOBAL=false}
-
-\noindent This trick can be useful for all hosts on which one needs
-to cross-compile.
-
-\noindent On BlueGene, one also needs to run the \texttt{xcreate\_header\_file}
-binary file manually rather than in the Makefile:
-
-\noindent \texttt{bgrun -np 1 -mode VN -exe ./bin/xcreate\_header\_file}
-
-\section{Visualizing the subroutine calling tree of the source code}
-
-\noindent
-Packages such as \texttt{Doxywizard} can be used to visualize the calling tree of the
-subroutines of the  source code. \texttt{Doxywizard} is
-a GUI front-end for configuring and running \texttt{Doxygen}.
-
-\bigskip
-
-\noindent To do your own call graphs, you can follow these simple steps below.
-
-\begin{enumerate}
-\setcounter{enumi}{-1}
-\item Install \texttt{Doxygen} \underline{and} \texttt{graphviz} (the two are usually in the package manager of classic Linux distribution).
-\item Run in the terminal : \texttt{doxygen -g}, which creates a \texttt{Doxyfile} that tells doxygen what you want it to do.
-\item Edit the Doxyfile. Two Doxyfile-type files have been already committed in the directory \\
-\noindent \texttt{specfem3d/doc/Call\_trees}:
-
-\begin{itemize}
-\item[\textbullet] \texttt{Doxyfile\_truncated\_call\_tree} will generate call graphs with maximum 3 or 4 levels of tree structure,
-\item[\textbullet] \texttt{Doxyfile\_complete\_call\_tree} will generate call graphs with complete tree structure.
-\end{itemize}
-
-The important entries in the Doxyfile are:
-
-\begin{description}
-\item [{\texttt{PROJECT\_NAME}}]
-\item [{\texttt{OPTIMIZE\_FOR\_FORTRAN}}] Set to YES
-\item [{\texttt{EXTRACT\_ALL}}] Set to YES
-\item [{\texttt{EXTRACT\_PRIVATE}}] Set to YES
-\item [{\texttt{EXTRACT\_STATIC}}] Set to YES
-\item [{\texttt{INPUT}}] From the directory \texttt{specfem3d/doc/Call\_trees}, it is \texttt{"../../src/"}
-\item [{\texttt{FILE\_PATTERNS}}] In SPECFEM case, it is \texttt{*.f90* *.F90* *.c* *.cu* *.h*}
-\item [{\texttt{HAVE\_DOT}}] Set to YES
-\item [{\texttt{CALL\_GRAPH}}] Set to YES
-\item [{\texttt{CALLER\_GRAPH}}] Set to YES
-\item [{\texttt{DOT\_PATH}}] The path where is located the dot program graphviz (if it is not in your \$PATH)
-\item [{\texttt{RECURSIVE}}] This tag can be used to turn specify whether or not subdirectories should be searched for input files as well. In the case of SPECFEM, set to YES.
-\item [{\texttt{EXCLUDE}}] Here, you can exclude:  \\
-\texttt{"../../src/specfem3D/older\_not\_maintained\_partial\_OpenMP\_port"}\\
-\texttt{"../../src/decompose\_mesh/scotch"}\\
-\texttt{"../../src/decompose\_mesh/scotch\_5.1.12b"}
-\item [{\texttt{DOT\_GRAPH\_MAX\_NODES}}] to set the maximum number of nodes that will be shown in the graph.
-If the number of nodes in a graph becomes larger than this value, doxygen will truncate the graph, which is visualized by representing a node as a red box.
-Minimum value: 0, maximum value: 10000, default value: 50.
-\item [{\texttt{MAX\_DOT\_GRAPH\_DEPTH}}] to set the maximum depth of the graphs generated by dot.
-A depth value of 3 means that only nodes reachable from the root by following a path via at most 3 edges will be shown.
-Using a depth of 0 means no depth restriction.
-Minimum value: 0, maximum value: 1000, default value: 0.
-
-\end{description}
-
-\item Run : \texttt{doxygen Doxyfile}, HTML and LaTeX files created by default in \texttt{html} and \texttt{latex} subdirectories.
-
-\item To see the call trees, you have to open the file \texttt{html/index.html} in your \underline{browser}.
-You will have many informations about each subroutines of SPECFEM (not only call graphs), you can click on every boxes / subroutines.
-It show you the call, and, the caller graph of each subroutine : the subroutines called by the concerned subroutine, and the previous subroutines who call this subroutine (the previous path), respectively.
-In the case of a truncated calling tree, the boxes with a red border indicates a node that has \underline{more} arrows than are shown (in other words: the graph is truncated with respect to this node).
-
-\end{enumerate}
-
-\medskip
-
-\noindent Finally, some useful links:
-
-\begin{itemize}
-\item[\textbullet] a good and short summary for the basic utilisation of Doxygen:
-
-\url{http://www.softeng.rl.ac.uk/blog/2010/jan/30/callgraph-fortran-doxygen/},
-
-\item[\textbullet] to configure the diagrams :
-
-\url{http://www.stack.nl/~dimitri/doxygen/manual/diagrams.html},
-
-\item[\textbullet] the complete alphabetical index of the tags in Doxyfile:
-
-\url{http://www.stack.nl/~dimitri/doxygen/manual/config.html},
-
-\item[\textbullet] more generally, the Doxygen manual:
-
-\url{http://www.stack.nl/~dimitri/doxygen/manual/index.html}.
-
-\end{itemize}
-
-\medskip
-
-\section{Using the ADIOS library for I/O}
-Regular POSIX I/O can be problematic when dealing with large simulations one large
-clusters (typically more than $10,000$ processes). SPECFEM3D use the ADIOS library~\cite{Liu2013}
-to deal transparently take advantage of advanced paralllel file system features. To enable
-ADIOS, the following steps should be done:
-\begin{enumerate}
-\item Install ADIOS (available from \url{https://www.olcf.ornl.gov/center-projects/adios/}).
-Make sure that your environment variables reference it.
-\item You may want to change ADIOS related values in the \texttt{constants.h.in} file.
-The default values probably suit most cases.
-\item Configure using the \texttt{--with-adios} flag.
-\end{enumerate}
-ADIOS is currently only usable for meshfem3D generated mesh (i.e. not for meshes generated
-with CUBTI). Additional control parameters are discussed in section~\ref{cha:Main-Parameter}.
-
-\section{Becoming a developer of the code, or making small modifications in the source code}
-
-If you want to develop new features in the code, and/or if you want to make small changes, improvements, or bug fixes, you are very welcome to contribute.
-To do so, i.e.
-to access the development branch of the source code with read/write access (in a safe way, no need to worry too much about breaking the package, there is a robot called BuildBot that is in charge of checking and validating all new contributions and changes), please visit this Web page: \url{https://github.com/geodynamics/specfem3d/wiki/Using-Hub}.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\chapter{\label{cha:Mesh-Generation}Mesh Generation}
-
-The first step in running a spectral-element simulation consists of
-constructing a high-quality mesh for the region under consideration.
-We provide two possibilities to do so: (1) relying on the external,
-hexahedral mesher CUBIT, or (2) using the provided, internal mesher
-\texttt{xmeshfem3D}. In the following, we explain these two approaches.
-
-\section{\label{cha:Running-the-Mesher-CUBIT}Meshing with \texttt{CUBIT}}
-
-CUBIT is a meshing tool suite for the creation of finite-element meshes
-for arbitrarily shaped models. It has been developed and maintained
-at Sandia National Laboratories and can be purchased for a small academic
-institutional fee at \url{http://cubit.sandia.gov}. Our experience
-showed that using CUBIT greatly facilitates and speeds up the generation
-and preparation of hexahedral, conforming meshes for a variety of
-geophysical models with increasing complexity.
-
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=0.6\textwidth]{figures/mount-cubit}
-\par\end{centering}
-
-\caption{Example of the graphical user interface of CUBIT. The hexahedral mesh
-shown in the main display consists of a hexahedral discretization
-of a single volume with topography.}
-
-\label{fig:mount.cubit}
-\end{figure}
-
-\noindent
-The basic steps in creating a load-balanced, partitioned mesh with
-CUBIT are:
-\begin{description}
-\item [{1.}] setting up a hexahedral mesh with CUBIT,
-\item [{2.}] exporting the CUBIT mesh into a SPECFEM3D Cartesian file format
-and
-\item [{3.}] partitioning the SPECFEM3D Cartesian mesh files for a chosen
-number of cores.
-\end{description}
-Examples are provided in the SPECFEM3D Cartesian package in the subdirectory
-\texttt{EXAMPLES/}. We strongly encourage you to contribute your own
-example to this package by contacting the CIG Computational Seismology
-Mailing List \urlwithparentheses{cig-seismo@geodynamics.org}.
-
-\subsection{Creating the Mesh with CUBIT}
-
-For the installation and handling of the CUBIT meshing tool suite,
-please refer to the CUBIT user manual and documentation. In order
-to give you a basic understanding of how to use CUBIT for our purposes,
-examples are provided in the SPECFEM3D Cartesian package in the subdirectory
-\texttt{EXAMPLES/}:
-\begin{description}
-\item [{\texttt{homogeneous\_halfspace}}] Creates a single block model
-and assigns elastic material parameters.
-\item [{\texttt{layered\_halfspace}}] Combines two different, elastic material
-volumes and creates a refinement layer between the two. This example
-can be compared for validation against the solutions provided in subdirectory
-~\\
- \texttt{VALIDATION\_3D\_SEM\_SIMPLER\_LAYER\_SOURCE\_DEPTH/}.
-\item [{\texttt{waterlayered\_halfspace}}] Combines an acoustic and elastic
-material volume as in a schematic marine survey example.
-\item [{\texttt{tomographic\_model}}] Creates a single block model whose
-material properties will have to be read in from a tomographic model
-file during the databases creation by \texttt{xgenerate\_databases}.
-\end{description}
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=0.45\textwidth]{figures/example-homogeneous}
-\includegraphics[width=0.45\textwidth]{figures/example-2layers} \\
- \includegraphics[width=0.45\textwidth]{figures/example-water} \includegraphics[width=0.45\textwidth]{figures/example-tomo}
-\par\end{centering}
-
-\caption{Screenshots of the CUBIT examples provided in subdirectory \texttt{EXAMPLES/}:
-homogeneous halfspace (top-left), layered halfspace (top-right), water
-layered halfspace (bottom-left) and tomographic model (bottom-right).}
-
-\label{fig:examples.cubit}
-\end{figure}
-
-In each example subdirectory you will find a \texttt{README} file,
-which explains in a step-by-step tutorial the workflow for the example.
-Please feel free to contribute your own example to this package by
-contacting the CIG Computational Seismology Mailing List \urlwithparentheses{cig-seismo@geodynamics.org}.
-
-In some cases, to re-create the meshes for the examples given, just type "claro ./create\_mesh.py" or similar
-from the command line ("claro" is the command to run CUBIT from the command line).
-
-IMPORTANT note: in order to correctly set up GEOCUBIT and run the examples, please read the file called "EXAMPLES/README";
-in particular, please make sure you correctly set up the Python paths as indicated in that file.
-
-\subsection{Exporting the Mesh with \texttt{run\_cubit2specfem3d.py} }
-
-\label{subsec:Exporting-the-Mesh}
-
-Once the geometric model volumes in CUBIT are meshed, you prepare
-the model for exportation with the definition of material blocks and
-boundary surfaces. Thus, prior to exporting the mesh, you need to
-define blocks specifying the materials and absorbing boundaries in
-CUBIT. This process could be done automatically using the script \texttt{run\_cubit2specfem3d.py}
-if the mesh meets some conditions or manually, following the block
-convention:
-\begin{description}
-\item [{material\_name}] Each material should have a specific block defined
-by a unique name. The name convention of the material is to start
-with either 'elastic' or 'acoustic'. It must be then followed by a
-unique identifier, e.g. 'elastic 1', 'elastic 2', etc. The additional
-attributes to the block define the material description.
-
-For an elastic material:
-\begin{description}
-\item [{material\_id}] An integer value which is unique for this material.
-\item [{Vp}] P-wave speed of the material (given in m/s).
-\item [{Vs}] S-wave speed of the material (given in m/s).
-\item [{rho}] density of the material (given in kg/m$^{3}$).
-\item [{Q}] quality factor to use in case of a simulation with attenuation
-turned on. It should be between 1 and 9000. In case no attenuation
-information is available, it can be set to zero. Please note that
-your Vp- and Vs-speeds are given for a reference frequency. To change
-this reference frequency, you change the value of \texttt{ATTENUATION\_f0\_REFERENCE}
-in the main constants file \texttt{constants.h} found in subdirectory
-\texttt{src/shared/}.
-\item [{anisotropic\_flag}] Flag describing the anisotropic model to use
-in case an anisotropic simulation should be conducted. See the file
-\texttt{model\_aniso.f90} in subdirectory \texttt{src/generate\_databases/}
-for an implementation of the anisotropic models. In case no anisotropy
-is available, it can be set to zero.
-\end{description}
-
-Note that this material block has to be defined using all the volumes
-which belong to this elastic material. For volumes belonging to another,
-different material, you will need to define a new material block.
-
-For an acoustic material:
-\begin{description}
-\item [{material\_id}] An integer value which is unique for this material.
-\item [{Vp}] P-wave speed of the material (given in m/s).
-\item [{0}] S-wave speed of the material is ignored.
-\item [{rho}] density of the material (given in kg/m$^{3}$).
-\end{description}
-\item [{face\_topo}] Block definition for the surface which defines the
-free surface (which can have topography). The name of this block must
-be 'face\_topo', the block has to be defined using all the surfaces
-which constitute the complete free surface of the model.
-\item [{face\_abs\_xmin}] Block definition for the faces on the absorbing
-boundaries, one block for each surface with x=Xmin.
-\item [{face\_abs\_xmax}] Block definition for the faces on the absorbing
-boundaries, one block for each surface with x=Xmax.
-\item [{face\_abs\_ymin}] Block definition for the faces on the absorbing
-boundaries, one block for each surface with y=Ymin.
-\item [{face\_abs\_ymax}] Block definition for the faces on the absorbing
-boundaries, one block for each surface with y=Ymax.
-\item [{face\_abs\_bottom}] Block definition for the faces on the absorbing
-boundaries, one block for each surface with z=bottom.
-\end{description}
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=0.45\textwidth]{figures/mount-surface} \includegraphics[width=0.45\textwidth]{figures/mount-abs}
-\par\end{centering}
-
-\caption{Example of the block definitions for the free surface 'face\_topo'
-(left) and the absorbing boundaries, defined in a single block 'face\_abs\_xmin'
-(right) in CUBIT.}
-
-\label{fig:mount.abs}
-\end{figure}
-
-Optionally, instead of specifying for each surface at the boundaries
-a single block like mentioned above, you can also specify a single
-block for all boundary surfaces and name it as one of the absorbing
-blocks above, e.g. 'face\_abs\_xmin'.
-
-After the block definitions are done, you export the mesh using the
-script \texttt{cubit2specfem3d.py} provided in each of the example
-directories (linked to the common script \texttt{CUBIT\_GEOCUBIT/cubit2specfem3d.py}).
-If the export was successful, you should find the following
-files in a subdirectory \texttt{MESH/}:
-\begin{description}
-\item [{absorbing\_cpml\_file}] \textbf{(only needed in case of C-PML absorbing
-conditions)} Contains on the first line the total number of C-PML
-spectral elements in the mesh, and then on the following line the
-list of all these C-PML elements with two numbers per line: first
-the spectral element number, and then a C-PML flag indicating to which
-C-PML layer(s) that element belongs, according to the following convention:
-
-\begin{itemize}
-\item Flag = 1 : element belongs to a X CPML layer only (either in Xmin
-or in Xmax),
-\item Flag = 2 : element belongs to a Y CPML layer only (either in Ymin
-or in Ymax),
-\item Flag = 3 : element belongs to a Z CPML layer only (either in Zmin
-or in Zmax),
-\item Flag = 4 : element belongs to a X CPML layer and also to a Y CPML
-layer,
-\item Flag = 5 : element belongs to a X CPML layer and also to a Z CPML
-layer,
-\item Flag = 6 : element belongs to a Y CPML layer and also to a Z CPML
-layer,
-\item Flag = 7 : element belongs to a X, to a Y and to a Z CPML layer, i.e.,
-it belongs to a CPML corner.
-\end{itemize}
-
-Note that it does not matter whether an element belongs to a Xmin
-or to a Xmax CPML, the flag is the same in both cases; the same is
-true for Ymin and Ymax, and also Zmin and Zmax.
-
-When you have an existing CUBIT (or similar) mesh stored in SPECFEM3D
-format, i.e., if you have existing \textquotedbl{}nodes\_coords\_file\textquotedbl{}
-and \textquotedbl{}mesh\_file\textquotedbl{} files but do not know
-how to assign CPML flags to them, we have created a small serial Fortran
-program that will do that automatically for you, i.e., which will
-create the \textquotedbl{}absorbing\_cpml\_file\textquotedbl{} for
-you. That program is utils/CPML/convert\_external\_layers\_of\_a\_given\_mesh\_to\_CPML\_layers.f90,
-and a small Makefile is provided in that directory (utils/CPML).
-
-IMPORTANT: it is your responsibility to make sure that in the input
-CUBIT (or similar) mesh that this code will read in SPECFEM3D format
-from files \textquotedbl{}nodes\_coords\_file\textquotedbl{} and \textquotedbl{}mesh\_file\textquotedbl{}
-you have created layers of elements that constitute a layer of constant
-thickness aligned with the coordinate grid axes (X, Y and/or Z), so
-that this code can assign CPML flags to them. This code does NOT check
-that (because it cannot, in any easy way). The mesh inside these CPML
-layers does not need to be structured nor regular, any non-structured
-mesh is fine as long as it has flat PML inner and outer faces, parallel
-to the axes, and thus of a constant thickness. The thickness can be
-different for the X, Y and Z sides. But for X it must not vary, for
-Y it must not vary, and for Z it must not vary. If you do not know
-the exact thickness, you can use a slightly LARGER value in this code
-(say 2\% to 5\% more) and this code will fix that and will adjust
-it; never use a SMALLER value otherwise this code will miss some CPML
-elements.
-
-Note: in a future release we will remove the constraint of having
-CPML layers aligned with the coordinate axes; we will allow for meshes
-that are titled by any constant angle in the horizontal plane. However
-this is not implemented yet.
-
-Note: in the case of fluid-solid layers in contact, the fluid-solid
-interface currently needs to be flat and horizontal inside the CPML
-layer (i.e., bathymetry should become flat and horizontal when entering
-the CPML); this (small) constraint will probably remain in the code
-for a while because it makes fluid-solid matching inside the CPML
-much easier.
-
-\item [{materials\_file}] Contains the material associations for each element.
-The format is:
-
-\begin{lyxcode}
-element\_ID~material\_ID
-\end{lyxcode}
-
-where \texttt{element\_ID} is the element identifier and \texttt{material\_ID}
-is a unique identifier, positive (for materials taken from this list
-of materials, i.e. for which each spectral element has constant material
-properties taken from this list) or negative (for tomographic models,
-i.e. for spectral element whose real velocities and density will be
-assigned later by calling an external function to define model variations,
-for instance in the case of tomographic models; in such a case, material
-properties can vary inside each spectral element, i.e. be different
-at each of its Gauss-Lobatto-Legendre grid points).
-
-\item [{nummaterial\_velocity\_file}] Defines the material properties.
-
-\begin{itemize}
-\item For classical materials (i.e., spectral elements for which the velocity
-and density model will not be assigned by calling an external function
-to define for instance a tomographic model), the format is:
-
-\begin{lyxcode}
-domain\_ID~material\_ID~rho~vp~vs~Qkappa~Qmu~anisotropy\_flag
-\end{lyxcode}
-
-where \texttt{\textbf{domain\_ID}}\textbf{ is 1 for acoustic and 2
-for elastic or viscoelastic materials,} \texttt{material\_ID} a unique
-identifier, \texttt{rho} the density in $kg\, m^{-3}$, \texttt{vp}
-the P-wave speed in $m\, s^{-1}$, \texttt{vs} the S-wave speed in
-$m\, s^{-1}$, \texttt{Q} the quality factor and \texttt{anisotropy\_flag}
-an identifier for anisotropic models. Note that the \texttt{Qkappa}
-value is ignored by the code unless \texttt{FULL\_ATTENUATION\_SOLID}
-is set. Note also that both \texttt{Qkappa} and \texttt{Qmu} are ignored
-by the code unless \texttt{ATTENUATION} is set. If you want a model
-with no \texttt{Qmu} attenuation, both set \texttt{ATTENUATION} to
-\texttt{.false.} in the \texttt{Par\_file} and set \texttt{Qmu} to
-9999 here. If you want a model with no \texttt{Qkappa} attenuation,
-both set \texttt{FULL\_ATTENUATION\_SOLID} to \texttt{.false.} in
-the \texttt{Par\_file} and set \texttt{Qkappa} to 9999 here. %%magnoni
-
-\item For tomographic velocity models, please read Chapter \ref{cha:-Changing-the}
-and Section \ref{sec:Using-tomographic} `Using external tomographic
-Earth models' for further details.
-\end{itemize}
-\item [{nodes\_coords\_file}] Contains the point locations in Cartesian
-coordinates of the mesh element corners.
-\item [{mesh\_file}] Contains the mesh element connectivity. The hexahedral
-elements can have 8 or 27 nodes.\\
- See picture doc/mesh\_numbering\_convention/numbering\_convention\_27\_nodes.jpg
-to see\\
- in which (standard) order the points must be cited. In the case of
-8 nodes, just include the first 8 points.
-\item [{free\_or\_absorbing\_surface\_file\_zmax}] Contains the free surface
-connectivity or \\
- the surface connectivity of the absorbing boundary surface at the
-top (Zmax), \\
- depending on whether the top surface is defined as free or absorbing
-(\texttt{STACEY\_INSTEAD\_OF\_FREE\_SURFACE} in \texttt{DATA/Par\_file}). \\
- You should put both the surface of acoustic regions and of elastic
-regions in that file; that is, list all the element faces that constitute
-the surface of the model in that file.
-\item [{absorbing\_surface\_file\_xmax}] Contains the surface connectivity
-of the absorbing boundary surface at Xmax \\
-(also needed in the case of C-PML absorbing conditions, in order for
-the code to be able to impose Dirichlet conditions on their outer
-edge).
-\item [{absorbing\_surface\_file\_xmin}] Contains the surface connectivity
-of the absorbing boundary surface at Xmin \\
-(also needed in the case of C-PML absorbing conditions, in order for
-the code to be able to impose Dirichlet conditions on their outer
-edge).
-\item [{absorbing\_surface\_file\_ymax}] Contains the surface connectivity
-of the absorbing boundary surface at Ymax \\
-(also needed in the case of C-PML absorbing conditions, in order for
-the code to be able to impose Dirichlet conditions on their outer
-edge).
-\item [{absorbing\_surface\_file\_ymin}] Contains the surface connectivity
-of the absorbing boundary surface at Ymin \\
-(also needed in the case of C-PML absorbing conditions, in order for
-the code to be able to impose Dirichlet conditions on their outer
-edge).
-\item [{absorbing\_surface\_file\_bottom}] Contains the surface connectivity
-of the absorbing boundary surface at the bottom (Zmin) \\
-(also needed in the case of C-PML absorbing conditions, in order for
-the code to be able to impose Dirichlet conditions on their outer
-edge).
-\end{description}
-These mesh files are needed as input files for the partitioner \texttt{xdecompose\_mesh}
-to load-balance the mesh. Please see the next section for further
-details.
-
-In directory "CUBIT\_GEOCUBIT" we provide a script that can help doing the above tasks of exporting a CUBIT mesh
-to SPECFEM3D format automatically for you: "run\_cubit2specfem3d.py".
-Just edit them to indicate the path to your local installation of CUBIT and also the name of the *.cub existing CUBIT mesh file
-that you want to export to SPECFEM3D format. These scripts will do the conversion for you automatically except assigning material properties
-to the different mesh layers. To do so, you will then need to edit the file
-called "nummaterial\_velocity\_file" that will have just been created and change it from the prototype created:
-
-    0 1 vol1 --> syntax: \#material\_domain\_id \#material\_id \#rho \#vp \#vs \#Q\_kappa \#Q\_mu \#anisotropy
-    0 2 vol2 --> syntax: \#material\_domain\_id \#material\_id \#rho \#vp \#vs \#Q\_kappa \#Q\_mu \#anisotropy
-
-(where "vol1" and "vol2" here represent the volume labels that you have set while creating the mesh in CUBIT)
-to for instance
-
-    2 1 1500 2300 1800 9999.0 9999.0 0
-    2 2 1600 2500 20000 9999.0 9999.0 0
-
-\subsubsection*{Checking the mesh quality}
-
-The quality of the mesh may be inspected more precisely based upon
-the serial code in the file \texttt{check\_mesh\_quality\_}~\\
- \texttt{CUBIT\_Abaqus.f90} located in the directory \texttt{src/check\_mesh\_quality\_CUBIT\_Abaqus/}.
-Running this code is optional because no information needed by the
-solver is generated.
-
-Prior to running and compiling this code, you have to export your
-mesh in CUBIT to an ABAQUS (.inp) format. For example, export mesh
-block IDs belonging to volumes in order to check the quality of the
-hexahedral elements. You also have to determine a number of parameters
-of your mesh, such as the number of nodes and number of elements and
-modify the header of the \texttt{check\_mesh\_quality\_CUBIT\_Abaqus.f90}
-source file in directory \texttt{src/check\_mesh\_quality\_CUBIT\_Abaqus/}.
-
-\noindent
-Then, in the main directory, type
-\begin{lyxcode}
-{\small make~xcheck\_mesh\_quality\_CUBIT\_Abaqus}{\small \par}
-\end{lyxcode}
-and use
-\begin{lyxcode}
-{\small ./bin/xcheck\_mesh\_quality\_CUBIT\_Abaqus}{\small \par}
-\end{lyxcode}
-to generate an %not supported: AVS output file (\texttt{\small AVS\_meshquality.inp} in AVS UCD format) or
-OpenDX output file (\texttt{\small DX\_mesh\_quality.dx}{\small )
-that can be used to investigate mesh quality, e.g. skewness of elements,
-and a Gnuplot histogram (}\texttt{\small mesh\_quality\_histogram.txt}{\small )
-that can be plotted with gnuplot (type `}\texttt{\small gnuplot plot\_mesh\_quality\_histogram.gnu}{\small ').
-The histogram is also printed to the screen. Analyze that skewness
-histogram of mesh elements to make sure no element has a skewness
-above approximately 0.75, otherwise the mesh is of poor quality (and
-if even a single element has a skewness value above 0.80, then you
-must definitely improve the mesh). If you want to start designing
-your own meshes, this tool is useful for viewing your creations. You
-are striving for meshes with elements with `cube-like' dimensions,
-e.g., the mesh should contain no very elongated or skewed elements.}{\small \par}
-
-\subsection{Partitioning the Mesh with \texttt{xdecompose\_mesh}}
-
-The SPECFEM3D Cartesian software package performs large scale simulations
-in a parallel 'Single Process Multiple Data' way. The spectral-element
-mesh created with CUBIT needs to be distributed on the processors.
-This partitioning is executed once and for all prior to the execution
-of the solver so it is referred to as a static mapping.
-
-An efficient partitioning is important because it leverages the overall
-running time of the application. It amounts to balance the number
-of elements in each slice while minimizing the communication costs
-resulting from the placement of adjacent elements on different processors.
-\texttt{decompose\_mesh} depends on the SCOTCH library \citep{PeRo96},
-which provides efficent static mapping, graph and mesh partitioning
-routines. SCOTCH is a free software package developed by Fran\c{c}ois
-Pellegrini et al. from LaBRI and INRIA in Bordeaux, France, downloadable
-from the web page \url{https://gforge.inria.fr/projects/scotch/}.
-\\
-
-In most cases, the configuration with \texttt{./configure~FC=ifort}
-should be sufficient. During the configuration process, the script
-tries to find existing SCOTCH installations. In case your system has
-no pre-existing SCOTCH installation, we provide the source code of
-SCOTCH, which is released open source under the French CeCILL-C version
-1 license, in directory \texttt{src/decompose\_mesh/scotch\_5.1.12b}.
-This version gets bundled with the compilation of the SPECFEM3D Cartesian
-package if no libraries could have been found. If this automatic compilation
-of the SCOTCH libraries fails, please refer to file INSTALL.txt in
-that directory to see further details how to compile it on your system.
-In case you want to use a pre-existing installation, make sure you
-have correctly specified the path of the SCOTCH library when using
-the option \texttt{-{}-with-scotch-dir} with the \texttt{./configure}
-script. In the future you should be able to find more recent versions
-at \url{http://www.labri.fr/perso/pelegrin/scotch/scotch_en.html}
-. \\
-
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=0.45\textwidth]{figures/mount-partitions}
-\includegraphics[width=0.45\textwidth]{figures/mount-partitions2}
-\par\end{centering}
-
-\caption{Example of a mesh partitioning onto four cores. Each single core partition
-is colored differently. The executable \texttt{xdecompose\_mesh} can
-equally distribute the mesh on any arbitrary number of cores. Domain
-decomposition is explained in detail in \citet{MaKoBlLe08}, and excellent
-scaling up to 150,000 processor cores in shown for instance in \citet{CaKoLaTiMiLeSnTr08,KoLaMi08a,MaKoBlLe08,KoErGoMi10,Kom11}.}
-
-\label{fig:mount.partitions}
-\end{figure}
-
-When you are ready to compile, in the main directory type
-\begin{lyxcode}
-{\small make~xdecompose\_mesh}{\small \par}
-\end{lyxcode}
-If all paths and flags have been set correctly, the executable \texttt{bin/xdecompose\_mesh}
-should be produced.
-
-The partitioning is done in serial for now (in the next release we
-will provide a parallel version of that code). It needs to be run in the main directory because it expects the \texttt{./DATA/Par\_file}. The synopsis is:
-\begin{lyxcode}
-./bin/xdecompose\_mesh~nparts~input\_directory~output\_directory
-\end{lyxcode}
-where
-\begin{itemize}
-\item \texttt{nparts} is the number of partitions, i.e., the number of cores
-for the parallel simulations,
-\item \texttt{input\_directory} is the directory which holds all the files
-generated by the Python script \texttt{cubit2specfem3d.py} explained
-in the previous Section~\ref{subsec:Exporting-the-Mesh}, e.g. \texttt{./MESH/},
-and
-\item \texttt{output\_directory} is the directory for the output of this
-partitioner which stores ACII-format files named like \texttt{proc{*}{*}{*}{*}{*}{*}\_Database}
-for each partition. These files will be needed for creating the distributed
-databases, and have to reside in the directory \texttt{LOCAL\_PATH}
-specified in the main \texttt{Par\_file}, e.g. in directory \texttt{./OUTPUT\_FILES/DATABASES\_MPI}.
-Please see Chapter~\ref{cha:Creating-Distributed-Databases} for
-further details.
-\end{itemize}
-Note that all the files generated by the Python script \texttt{cubit2specfem3d.py}
-must be placed in the \texttt{input\_directory} folder before running
-the program.
-
-\section{\label{cha:Running-the-Mesher-Meshfem3D}Meshing with \texttt{xmeshfem3D}}
-
-In case you successfully ran the configuration script, you are also
-ready to compile the internal mesher. This is an alternative to CUBIT
-for the mesh generation of relatively simple models.
-
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[scale=0.5]{figures/socal_map_mpi}
-\par\end{centering}
-
-\caption{\label{fig:For-parallel-computing}For parallel computing purposes,
-the model block is subdivided in $\nprocxi\times\nproceta$ slices
-of elements. In this example we use $5^{2}=25$ processors. }
-\end{figure}
-
-\noindent
-In the main directory, type
-\begin{lyxcode}
-{\small make~xmeshfem3D}{\small \par}
-\end{lyxcode}
-If all paths and flags have been set correctly, the mesher should
-now compile and produce the executable \texttt{bin/xmeshfem3D}. Please
-note that \texttt{xmeshfem3D} must be called directly from the main
-directory, as most of the binaries of the package.
-
-Input for the mesh generation program is provided through the parameter
-file \texttt{Mesh\_Par\_file}, which resides in the subdirectory \texttt{DATA/meshfem3D\_files/}.
-(to see how to use it, see the EXAMPLES specific to the internal mesher in directory \texttt{EXAMPLES/meshfem3D\_examples/}.
-Before running the mesher, a number of parameters need to be set in
-the \texttt{Mesh\_Par\_file}. This requires a basic understanding
-of how the SEM is implemented, and we encourage you to read \citet{KoVi98,KoTr99}
-and \citet{KoLiTrSuStSh04}.
-
-The mesher and the solver use UTM coordinates internally, therefore
-you need to define the zone number for the UTM projection (e.g., zone
-11 for Los Angeles). Use decimal values for latitude and longitude
-(no minutes/seconds). These values are approximate; the mesher will
-round them off to define a square mesh in UTM coordinates. When running
-benchmarks on rectangular models, turn the UTM projection off by using
-the flag \texttt{\small SUPPRESS\_UTM\_PROJECTION}{\small , in which
-case all `longitude' parameters simply refer to the $x$~axis, and
-all `latitude' parameters simply refer to the $y$~axis. To run the
-mesher for a global simulation, the following parameters need to be
-set in the }\texttt{\small Mesh\_Par\_file}{\small :}{\small \par}
-\begin{description}
-\item [{\texttt{LATITUDE\_MIN}}] Minimum latitude in the block (negative
-for South).
-\item [{\texttt{LATITUDE\_MAX}}] Maximum latitude in the block.
-\item [{\texttt{LONGITUDE\_MIN}}] Minimum longitude in the block (negative
-for West).
-\item [{\texttt{LONGITUDE\_MAX}}] Maximum longitude in the block.
-\item [{\texttt{DEPTH\_BLOCK\_KM}}] Depth of bottom of mesh in kilometers.
-\item [{\texttt{\noun{UTM\_PROJECTION\_ZONE}}}] UTM projection zone in
-which your model resides, only valid when \texttt{SUPPRESS\_UTM\_}~\\
- \texttt{PROJECTION} is \texttt{.false.}.
-Use a negative zone number for the Southern hemisphere:
-the Northern hemisphere corresponds to zones +1 to +60,
-the Southern hemisphere to zones -1 to -60.
-We use the WGS84 (World Geodetic System 1984) reference ellipsoid for the UTM projection. If you prefer to use the Clarke 1866 ellipsoid,
-edit file \texttt{src/shared/utm\_geo.f90}, uncomment that ellipsoid and recompile the code.
-From http://en.wikipedia.org/wiki/Universal\_Transverse\_Mercator\_coordinate\_system :
-The Universal Transverse Mercator coordinate system was developed by the United States Army Corps of Engineers in the 1940s.
-The system was based on an ellipsoidal model of Earth. For areas within the contiguous United States
-the Clarke Ellipsoid of 1866 was used. For the remaining areas of Earth, including Hawaii, the International Ellipsoid was used.
-The WGS84 ellipsoid is now generally used to model the Earth in the UTM coordinate system,
-which means that current UTM northing at a given point can be 200+ meters different from the old one.
-For different geographic regions, other datum systems (e.g.: ED50, NAD83) can be used.
-\item [{\texttt{SUPPRESS\_UTM\_PROJECTION}}] set to be \texttt{.false.}
-when your model range is specified in geographical coordinates, and
-needs to be \texttt{.true.} when your model is specified in Cartesian
-coordinates. \noun{UTM projection zone in which your simulation region
-resides.}
-\item [{\texttt{INTERFACES\_FILE }}] File which contains the description
-of the topography and of the interfaces between the different layers
-of the model, if any. The number of spectral elements in the vertical
-direction within each layer is also defined in this file.
-\item [{$\nexxi$}] The number of spectral elements along one side of the
-block. This number \textit{must} be 8~$\times$~a multiple of $\nprocxi$
-defined below. Based upon benchmarks against semi-analytical discrete
-wavenumber synthetic seismograms \citep{KoLiTrSuStSh04}, determined
-that a $\nexxi=288$ run is accurate to a shortest period of roughly
-2~s. Therefore, since accuracy is determined by the number of grid
-points per shortest wavelength, for any particular value of $\nexxi$
-the simulation will be accurate to a shortest period determined by
-\begin{equation}
-\mbox{shortest period (s)}=(288/\nexxi)\times2.\label{eq:shortest_period}
-\end{equation}
- The number of grid points in each orthogonal direction of the reference
-element, i.e., the number of Gauss-Lobatto-Legendre points, is determined
-by \texttt{NGLLX} in the \texttt{constants.h} file. We generally use
-$\mbox{\texttt{NGLLX\/}}=5$, for a total of $5^{3}=125$ points per
-elements. We suggest not to change this value.
-\item [{$\nexeta$}] The number of spectral elements along the other side
-of the block. This number \textit{must} be 8~$\times$~a multiple
-of $\nproceta$ defined below.
-\item [{$\nprocxi$}] The number of processors or slices along one side
-of the block (see Figure~\ref{fig:For-parallel-computing}); we must
-have $\nexxi=8\times c\times\nprocxi$, where $c\ge1$ is a positive
-integer.
-\item [{$\nproceta$}] The number of processors or slices along the other
-side of the block; we must have $\nexeta=8\times c\times\nproceta$,
-where $c\ge1$ is a positive integer.
-\item [{\texttt{USE\_REGULAR\_MESH}}] set to be \texttt{.true.} if you
-want a perfectly regular mesh or \texttt{.false.} if you want to add
-doubling horizontal layers to coarsen the mesh. In this case, you
-also need to provide additional information by setting up the next
-three parameters.
-\item [{\texttt{NDOUBLINGS}}] The number of horizontal doubling layers.
-Must be set to \texttt{1} or \texttt{2} if \texttt{USE\_REGULAR\_MESH}
-is set to \texttt{.true.}.
-\item [{\texttt{NZ\_DOUBLING\_1}}] The position of the first doubling layer
-(only interpreted if \texttt{USE\_REGULAR\_MESH} is set to \texttt{.true.}).
-\item [{\texttt{NZ\_DOUBLING\_2}}] The position of the second doubling
-layer (only interpreted if \texttt{USE\_REGULAR\_MESH} is set to \texttt{.true.}
-and if \texttt{NDOUBLINGS} is set to \texttt{2}).
-\item [{\texttt{CREATE\_ABAQUS\_FILES}}] Set this flag to \texttt{.true.}
-to save Abaqus FEA \urlwithparentheses{www.simulia.com} mesh files
-for subsequent viewing. Turning the flag on generates files in the
-\texttt{LOCAL\_PATH} directory. See Section~\ref{sec:Mesh-graphics}
-for a discussion of mesh viewing features.
-\item [{\texttt{CREATE\_DX\_FILES}}] Set this flag to \texttt{.true.} to
-save OpenDX \urlwithparentheses{www.opendx.org} mesh files for subsequent
-viewing.
-\item [{\texttt{LOCAL\_PATH}}] Directory in which the partitions generated
-by the mesher will be written. Generally one uses a directory on the
-local disk of the compute nodes, although on some machines these partitions
-are written on a parallel (global) file system (see also the earlier
-discussion of the \texttt{LOCAL\_PATH\_IS\_ALSO\_GLOBAL} flag in Chapter~\ref{cha:Getting-Started}).
-The mesher generates the necessary partitions in parallel, one set
-for each of the $\nprocxi\times\nproceta$ slices that constitutes
-the mesh (see Figure~\ref{fig:For-parallel-computing}). After the
-mesher finishes, you can log in to one of the compute nodes and view
-the contents of the \texttt{LOCAL\_PATH} directory to see the files
-generated by the mesher. These files will be needed for creating the
-distributed databases, and have to reside in the directory \texttt{LOCAL\_PATH}
-specified in the main \texttt{Par\_file}, e.g. in directory \texttt{OUTPUT\_FILES/DATABASES\_MPI}.
-Please see Chapter~\ref{cha:Creating-Distributed-Databases} for
-further details.
-\item [{\texttt{NMATERIALS}}] The number of different materials in your
-model. In the following lines, each material needs to be defined as
-:
-
-\begin{lyxcode}
-material\_ID~rho~vp~vs~Q~anisotropy\_flag~domain\_ID
-\end{lyxcode}
-
-where
-\begin{itemize}
-\item \texttt{Q} : quality factor (0=no attenuation)
-\item \texttt{anisotropy\_flag} : 0=no anisotropy / 1,2,.. check with implementation
-in \texttt{aniso\_model.f90}
-\item \texttt{domain\_id} : 1=acoustic / 2=elastic
-\end{itemize}
-\item [{\texttt{NREGIONS}}] The number of regions in the mesh. In the following
-lines, because the mesh is regular or 'almost regular', each region
-is defined as :
-
-\begin{lyxcode}
-NEX\_XI\_BEGIN~NEX\_XI\_END~NEX\_ETA\_BEGIN~NEX\_ETA\_END~NZ\_BEGIN~NZ\_END~material\_ID
-\end{lyxcode}
-\end{description}
-The \texttt{INTERFACES\_FILE} parameter of \texttt{Mesh\_Par\_File}
-defines the file which contains the settings of the topography grid
-and of the interfaces grids. Topography is defined as a set of elevation
-values on a regular 2D grid. It is also possible to define interfaces
-between the layers of the model in the same way. The file needs to
-define several parameters:
-\begin{itemize}
-\item The number of interfaces, including the topography. This needs to
-be set at the first line. Then, from the bottom to the top of the
-model, you need to define the grids with:
-\item \texttt{SUPPRESS\_UTM\_PROJECTION} flag as described previously,
-\item number of points along $x$ and $y$ direction (NXI and NETA),
-\item minimal $x$ and $y$ coordinates (LONG\_MIN and LAT\_MIN),
-\item spacing between points along $x$ and $y$ (SPACING\_XI and SPACING\_ETA)
-and
-\item the name of the file which contains the elevation values (in $y$.$x$
-increasing order).
-\end{itemize}
-At the end of this file, you simply need to set the number of spectral
-elements in the vertical direction for each layer. We provide a few
-models in the {\texttt{EXAMPLES/}} directory. \\
-
-Finally, depending on your system, you might need to provide a file
-that tells MPI what compute nodes to use for the simulations. The
-file must have a number of entries (one entry per line) at least equal
-to the number of processors needed for the run. A sample file is provided
-in the file \texttt{mymachines}. This file is not used by the mesher
-or solver, but is required by the \texttt{go\_mesher} and \texttt{go\_solver}
-default job submission scripts. See Chapter~\ref{cha:Scheduler}
-for information about running the code on a system with a scheduler,
-e.g., LSF.
-
-Now that you have set the appropriate parameters in the \texttt{Mesh\_Par\_file}
-and have compiled the mesher, you are ready to launch it! This is
-most easily accomplished based upon the \texttt{go\_mesher} script.
-When you run on a PC cluster, the script assumes that the nodes are
-named n001, n002, etc. If this is not the case, change the \texttt{tr
--d `n'} line in the script. You may also need to edit the last command
-at the end of the script that invokes the \texttt{mpirun} command.
-See Chapter~\ref{cha:Scheduler} for information about running the
-code on a system with a scheduler, e.g., LSF.
-
-Mesher output is provided in the \texttt{OUTPUT\_FILES} directory
-in \texttt{output\_mesher.txt}; this file provides lots of details
-about the mesh that was generated. Please note that the mesher suggests
-a time step \texttt{DT} to run the solver with. The mesher output
-file also contains a table about the quality of the mesh to indicate
-possible problems with the distortions of elements. Alternatively,
-output can be directed to the screen instead by uncommenting a line
-in \texttt{constants.h}:
-\begin{lyxcode}
-!~uncomment~this~to~write~messages~to~the~screen~
-
-!~integer,~parameter~::~IMAIN~=~ISTANDARD\_OUTPUT
-\end{lyxcode}
-To control the quality of the mesh, check the standard output (either
-on the screen or in the \texttt{OUTPUT\_FILES} directory in \texttt{output\_mesher.txt})
-and analyze the skewness histogram of mesh elements to make sure no
-element has a skewness above approximately 0.75, otherwise the mesh
-is of poor quality (and if even a single element has a skewness value
-above 0.80, then you must definitely improve the mesh). To draw the
-skewness histogram on the screen, type \texttt{gnuplot plot\_mesh\_quality\_histogram.gnu}.
-
-\chapter{\label{cha:Creating-Distributed-Databases}Creating the Distributed
-Databases}
-
-After using \texttt{xmeshfem3D} or \texttt{xdecompose\_mesh}, the
-next step in the workflow is to compile \texttt{xgenerate\_}~\\
- \texttt{databases}. This program is going to create all the missing
-information needed by the SEM solver.
-
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=0.6\textwidth]{figures/workflow}
-\par\end{centering}
-
-\caption{Schematic workflow for a SPECFEM3D Cartesian simulation. The executable
-\texttt{xgenerate\_databases} creates the GLL mesh points and assigns
-specific model parameters.}
-
-\label{fig:workflow.databases}
-\end{figure}
-
-\noindent
-In the main directory, type
-\begin{lyxcode}
-{\small make~xgenerate\_databases}{\small \par}
-\end{lyxcode}
-Input for the program is provided through the main parameter file
-\texttt{Par\_file}, which resides in the subdirectory \texttt{DATA}.
-Please note that \texttt{xgenerate\_databases} must be called directly
-from the main directory, as most of the binaries of the package.
-
-\section{\label{cha:Main-Parameter}Main parameter file \texttt{Par\_file}}
-
-Before running \texttt{xgenerate\_databases}, a number of parameters
-need to be set in the main parameter \texttt{Par\_file} located in
-the subdirectory \texttt{DATA}:
-\begin{description}
-\item [{\texttt{SIMULATION\_TYPE}}] is set to 1 for forward simulations,
-2 for adjoint simulations (see Section \ref{sec:Adjoint-simulation-finite})
-and 3 for kernel simulations (see Section \ref{sec:Finite-Frequency-Kernels}).
-\item [{\texttt{SAVE\_FORWARD}}] is only set to \texttt{.true.} for a forward
-simulation with the last frame of the simulation saved, as part of
-the finite-frequency kernel calculations (see Section \ref{sec:Finite-Frequency-Kernels}).
-For a regular forward simulation, leave \texttt{SIMULATION\_TYPE}
-and \texttt{SAVE\_FORWARD} at their default values.
-\item [{\texttt{\noun{UTM\_PROJECTION\_ZONE}}}] UTM projection zone in
-which your model resides, only valid when \texttt{SUPPRESS\_UTM\_}~\\
- \texttt{PROJECTION} is \texttt{.false.}.
-\item [{\texttt{SUPPRESS\_UTM\_PROJECTION}}] set to be \texttt{.false.}
-when your model range is specified in the geographical coordinates,
-and needs to be \texttt{.true.} when your model is specified in a
-cartesian coordinates. \noun{UTM projection zone in which your simulation
-region resides.}
-\item [{\texttt{NPROC}}] The number of MPI processors, each one is assigned
-one slice of the whole mesh.
-\item [{\texttt{NSTEP}}] The number of time steps of the simulation. This
-controls the length of the numerical simulation, i.e., twice the number
-of time steps requires twice as much CPU time. This feature is not
-used at the time of generating the distributed databases but is required
-for the solver, i.e., you may change this parameter after running
-\texttt{xgenerate\_databases}.
-\item [{\texttt{DT}}] The length of each time step in seconds. This feature
-is not used at the time of generating the distributed databases but
-is required for the solver. Please see also Section~\ref{sec:Choosing-the-Time-Step}
-for further details.
-\item [{\texttt{NGNOD}}] The number of nodes for 2D and 3D shape functions
-for hexahedra. We use either 8-node mesh elements (bricks) or 27-node
-elements. If you use the internal mesher, the only option is 8-node
-bricks (27-node elements are not supported). \texttt{CUBIT} does not
-support HEX27 elements either (it can generate them, but they are
-flat, i.e. identical to HEX8). To generate HEX27 elements with curvature
-properly taken into account, you can use \texttt{Gmsh} \url{http://geuz.org/gmsh/}
-\item [{\texttt{MODEL}}] Must be set to one of the following:
-
-\begin{description}
-\item [{\textmd{Models defined by mesh parameters:}}] ~
-
-\begin{description}
-\item [{\texttt{default}}] Uses model parameters as defined by meshing
-procedures described in the previous Chapter~\ref{cha:Mesh-Generation}.
-\end{description}
-
-\item [{\textmd{1D~models~with~real~structure:}}] ~
-
-\begin{description}
-\item [{\texttt{1D\_prem}}] Isotropic version of the spherically symmetric
-Preliminary Reference Earth Model (PREM) \citep{DzAn81}.
-\item [{\texttt{1D\_socal}}] A standard isotropic 1D model for Southern
-California.
-\item [{\texttt{1D\_cascadia}}] Isotropic 1D profile for the Cascadia region.
-\end{description}
-
-\item [{\textmd{Fully~3D~models:}}] ~
-
-\begin{description}
-\item [{\texttt{aniso}}] For a user-specified fully anisotropic model.
-Parameters are set up in routines located in file \texttt{model\_aniso.f90}
-in directory \texttt{src/generate\_databases/}. See Chapter~\ref{cha:-Changing-the}
-for a discussion on how to specify your own 3D model.
-
-\item [{\texttt{external}}] For a user-specified isotropic model which
-uses externally defined model parameters. Uses external model definitions
-set up in routines located in file \texttt{model\_external\_values.f90}
-in directory \texttt{src/generate\_databases/}. Please modify these
-generic template routines to use your own model definitions.
-\item [{\texttt{gll}}] For a user-specified isotropic model which uses
-external binary files for $v_{p}$, $v_{s}$ and $\rho$. Binary files
-are given in the same format as when outputted by the \texttt{xgenerate\_databases}
-executable when using option \texttt{SAVE\_MESH\_FILES}. These binary
-files define the model parameters on all GLL points which can be used
-for iterative inversion procedures.
-\item [{\texttt{salton\_trough}}] A 3D $V_{p}$ model for Southern California.
-Users must provide the corresponding data file \texttt{regrid3\_vel\_p.bin}
-in directory \texttt{DATA/st\_3D\_block\_harvard/}.
-\item [{\texttt{tomo}}] For a user-specified 3D isotropic model which uses
-a tomographic model file \texttt{tomographic\_model.xyz} in directory
-\texttt{DATA}. See Section \ref{sec:Using-tomographic}, for a discussion
-on how to specify your own 3D tomographic model.
-\end{description}
-\end{description}
-\item [{\texttt{APPROXIMATE\_OCEAN\_LOAD}}] Set to \texttt{.true.} if the
-effect of the oceans on seismic wave propagation should be incorporated
-based upon the (rough) approximate treatment discussed in \citet{KoTr02b}.
-This feature is inexpensive from a numerical perspective, both in
-terms of memory requirements and CPU time. This approximation is accurate
-at periods of roughly 20~s and longer. At shorter periods the effect
-of water phases/reverberations is not taken into account, even when
-the flag is on. If you want to model the effect of a fluid-solid model
-at short periods, then set this flag to \texttt{.false.} and mesh
-the fluid layer explicitly in your mesher, so that it is computed
-accurately and without this approximation.
-\item [{\texttt{TOPOGRAPHY}}] This feature is only effective if \texttt{APPROXIMATE\_OCEAN\_LOAD}
-is set to \texttt{.true.}. Set to \texttt{.true.} if topography and
-bathymetry should be read in based upon the topography file specified
-in the main constants file \texttt{constants.h} found in subdirectory
-\texttt{src/shared/} to evaluate elevations. If not set, elevations
-will be read from the numerical mesh.
-\item [{\texttt{ATTENUATION}}] Set to \texttt{.true.} if attenuation should
-be incorporated. Turning this feature on increases the memory requirements
-significantly (roughly by a factor of~1.5), and is numerically fairly
-expensive. See \citet{KoTr99,KoTr02a} for a discussion on the implementation
-of attenuation based upon standard linear solids. Please note that
-the Vp- and Vs-velocities of your model are given for a reference
-frequency. To change this reference frequency, you change the value
-of \texttt{ATTENUATION\_f0\_REFERENCE} in the main constants file
-\texttt{constants.h} found in subdirectory \texttt{src/shared/}.
-\item [{\texttt{FULL\_ATTENUATION\_SOLID}}] Set to \texttt{.true.} to add
-$Q_{\kappa}$ attenuation to the simulations in addition to $Q_{\mu}$
-attenuation when the \texttt{ATTENUATION} flag above is on. Should
-be off in almost all cases for seismic wave propagation, since $Q_{\kappa}$
-is almost always negligible in Earth models. However, this parameter
-can be useful for instance for ocean acoustics simulations, when $Q_{\kappa}$
-is not negligible in the solid layer at the sea bottom.
-\item [{\texttt{ANISOTROPY}}] Set to \texttt{.true.} if you want to use
-an anisotropy model. Please see the file \texttt{model\_aniso.f90}
-in subdirectory \texttt{src/generate\_databases/} for the current
-implementation of anisotropic models.
-\item [{\texttt{TOMOGRAPHY\_PATH}}] Directory in which the tomography files
-are stored for using external tomographic Earth models (please read
-Chapter \ref{cha:-Changing-the} and Section \ref{sec:Using-tomographic}
-`Using external tomographic Earth models' for further details.).
-\item [{\texttt{USE\_OLSEN\_ATTENUATION}}] Set to \texttt{.true.} if you
-want to use the attenuation model that scaled from the S-wave speed
-model using Olsen's empirical relation (see \citet{OlDaBr03}).
-\item [{\texttt{OLSEN\_ATTENUATION\_RATIO}}] Determines the Olsen's constant
-in Olsen's empirical relation (see \citet{OlDaBr03}).
-\item [{\texttt{PML\_CONDITIONS}}] Set to \texttt{.true.} to turn on C-PML
-boundary conditions for a regional simulation. Both fluids and elastic
-solids are supported. Note that xmeshfem3d generated meshes do not support C-PML yet.
-\item [{\texttt{PML\_INSTEAD\_OF\_FREE\_SURFACE}}] Set to \texttt{.true.}
-to turn on C-PML boundary conditions on the top surface instead of
-the usual free surface.
-\item [{\texttt{f0\_FOR\_PML}}] Determines the dominant frequency that
-will be used in the calculation of PML damping profiles; this should
-be set to the same (or similar) dominant frequency as that of the
-source that you will use in your simulation. If you plan to use a
-Dirac source, then use the dominant frequency of the source wavelet
-with which you plan to convolve your seismograms later on in post-processing.
-\item [{\texttt{STACEY\_ABSORBING\_CONDITIONS}}] Set to \texttt{.true.}
-to turn on Clayton-Enquist absorbing boundary conditions (see \citet{KoTr99}).
-In almost all cases it is much better to use CPML absorbing layers
-(see the options above) and leave this flag to \texttt{.false.}.
-\item [{\texttt{STACEY\_INSTEAD\_OF\_FREE\_SURFACE}}] Set to \texttt{.true.}
-to turn on absorbing boundary conditions on the top surface which
-by default constitutes a free surface of the model.
-\item [{\texttt{CREATE\_SHAKEMAP}}] Set this flag to \texttt{.true.} to
-create a ShakeMap\textregistered{}, i.e., a peak ground velocity map
-of the maximum absolute value of the two horizontal components of
-the velocity vector.
-\item [{\texttt{MOVIE\_SURFACE}}] Set to \texttt{.false.}, unless you want
-to create a movie of seismic wave propagation on the Earth's surface.
-Turning this option on generates large output files. See Section~\ref{sec:Movies}
-for a discussion on the generation of movies. This feature is only
-relevant for the solver.
-\item [{\texttt{MOVIE\_TYPE}}] Set this flag to 1 to show the top surface
-(tomography + oceans) only, to 2 to show all external faces of the
-mesh (i.e. topography + vertical edges + bottom) in shakemaps and
-surface movies.
-\item [{\texttt{MOVIE\_VOLUME}}] Set to \texttt{.false.}, unless you want
-to create a movie of seismic wave propagation in the Earth's interior.
-Turning this option on generates huge output files. See Section~\ref{sec:Movies}
-for a discussion on the generation of movies. This feature is only
-relevant for the solver.
-\item [{\texttt{SAVE\_DISPLACEMENT}}] Set this flag to \texttt{.true.}
-if you want to save the displacement instead of velocity for the movie
-frames.
-\item [{\texttt{USE\_HIGHRES\_FOR\_MOVIES}}] Set this flag to \texttt{.true.}
-if you want to save the values at all the NGLL grid points for the
-movie frames.
-\item [{\texttt{NTSTEP\_BETWEEN\_FRAMES}}] Determines the number of timesteps
-between movie frames. Typically you want to save a snapshot every
-100 timesteps. The smaller you make this number the more output will
-be generated! See Section~\ref{sec:Movies} for a discussion on the
-generation of movies. This feature is only relevant for the solver.
-\item [{\texttt{HDUR\_MOVIE}}] Determines the half duration of the source
-time function for the movie simulations. When this parameter is set
-to be 0, a default half duration that corresponds to the accuracy
-of the simulation is provided. Otherwise, it adds this half duration
-to the half duration specified in the source file \texttt{CMTSOLUTION},
-thus simulates longer periods to make the movie images look smoother.
-\item [{\texttt{SAVE\_MESH\_FILES}}] Set this flag to \texttt{.true.} to
-save ParaView \urlwithparentheses{www.paraview.org} mesh files for
-subsequent viewing. Turning the flag on generates large (distributed)
-files in the \texttt{LOCAL\_PATH} directory. See Section~\ref{sec:Mesh-graphics}
-for a discussion of mesh viewing features.
-\item [{\texttt{LOCAL\_PATH}}] Directory in which the distributed databases
-will be written. Generally one uses a directory on the local disk
-of the compute nodes, although on some machines these databases are
-written on a parallel (global) file system (see also the earlier discussion
-of the \texttt{LOCAL\_PATH\_IS\_ALSO\_GLOBAL} flag in Chapter~\ref{cha:Getting-Started}).
-\texttt{xgenerate\_databases} generates the necessary databases in
-parallel, one set for each of the \texttt{NPROC} slices that constitutes
-the mesh (see Figure~\ref{fig:mount.partitions} and Figure~\ref{fig:For-parallel-computing}).
-After the executable finishes, you can log in to one of the compute
-nodes and view the contents of the \texttt{LOCAL\_PATH} directory
-to see the (many) files generated by \texttt{xgenerate\_databases}.
-Please note that the \texttt{LOCAL\_PATH} directory should already
-contain the output files of the partitioner, i.e. from \texttt{xdecompose\_mesh}
-or \texttt{xmeshfem3D}.
-\item [{\texttt{NTSTEP\_BETWEEN\_OUTPUT\_INFO}}] This parameter specifies
-the interval at which basic information about a run is written to
-the file system (\texttt{timestamp{*}} files in the \texttt{OUTPUT\_FILES}
-directory). If you have access to a fast machine, set \texttt{NTSTEP\_BETWEEN\_OUTPUT\_INFO}
-to a relatively high value (e.g., at least 100, or even 1000 or more)
-to avoid writing output text files too often. This feature is not
-used at the time of meshing. One can set this parameter to a larger
-value than the number of time steps to avoid writing output during
-the run.
-\item [{\texttt{NTSTEP\_BETWEEN\_OUTPUT\_SEISMOS}}] This parameter specifies
-the interval at which synthetic seismograms are written in the \texttt{LOCAL\_PATH}
-directory. If a run crashes, you may still find usable (but shorter
-than requested) seismograms in this directory. On a fast machine set
-\texttt{NTSTEP\_BETWEEN\_OUTPUT\_SEISMOS} to a relatively high value
-to avoid writing to the seismograms too often. This feature is only
-relevant for the solver.
-\item [{\texttt{USE\_FORCE\_POINT\_SOURCE}}] Turn this flag on to use a
-(tilted) \texttt{FORCESOLUTION} force point source instead of a \texttt{CMTSOLUTION}
-moment-tensor source. When the force source does not fall exactly
-at a grid point, the solver interpolates the force between grid points
-using Lagrange interpolants. This can be useful e.g. for oil industry
-foothills simulations in which the source is a vertical force, normal
-force, tilted force, or an impact etc. Note that in the \texttt{FORCESOLUTION}
-file, you will need to edit the East, North and vertical components
-of an arbitrary (not necessarily unitary, the code will normalize it automatically) direction vector of the force vector;
-thus refer to Appendix~\ref{cha:Coordinates} for the orientation
-of the reference frame. This vector is made unitary internally in
-the solver and thus only its direction matters here; its norm is ignored
-and the norm of the force used is the factor force source times the
-source time function.
-When using this option, by default the code can locate the force source
-anywhere between mesh points in order to honor its exact location; this
-is more precise than using the closest GLL mesh point, but it is also a bit slower.
-If needed, you can change that default behavior and force the code to use the closest
-GLL mesh point instead by setting flag \texttt{USE\_BEST\_LOCATION} to \texttt{.false.} instead of \texttt{.true.} in file
-\texttt{src/shared/constants.h.in} and running the \texttt{configure} script again and recompiling the code.
-\item [{\texttt{USE\_RICKER\_TIME\_FUNCTION}}] Turn this flag on to use
-a Ricker source time function instead of the source time functions
-set by default to represent a (tilted) \texttt{FORCESOLUTION} force
-point source or a \texttt{CMTSOLUTION} moment-tensor source. Originally,
-if a \texttt{CMTSOLUTION} moment-tensor source is used, a (pseudo)
-Heaviside step function with a very short half duration is defined
-for elastic cases to represent the permanent slip on the fault while
-in the acoustic case a Gaussian source time function with a similarly
-short half duration is defined to physically describe actions within
-the fluid. Otherwise, if a \texttt{FORCESOLUTION} force source is
-used, a (pseudo) Dirac delta source time function is defined by default.
-Any other source-time function may then be obtained by convolution.
-\item [{\texttt{PRINT\_SOURCE\_TIME\_FUNCTION}}] Turn this flag on to print
-information about the source time function in the file \texttt{OUTPUT\_FILES/plot\_source\_time\_function.txt}.
-This feature is only relevant for the solver.
-
-\item [{\texttt{NUMBER\_OF\_SIMULTANEOUS\_RUNS}}] adds the ability to run several calculations (several earthquakes)
-in an embarrassingly-parallel fashion from within the same run;
-this can be useful when using a very large supercomputer to compute
-many earthquakes in a catalog, in which case it can be better from
-a batch job submission point of view to start fewer and much larger jobs,
-each of them computing several earthquakes in parallel.
-To turn that option on, set parameter NUMBER\_OF\_SIMULTANEOUS\_RUNS to a value greater than 1.
-To implement that, we create NUMBER\_OF\_SIMULTANEOUS\_RUNS MPI sub-communicators,
-each of them being labeled "my\_local\_mpi\_comm\_world", and we use them
-in all the routines in "src/shared/parallel.f90", except in MPI\_ABORT() because in that case
-we need to kill the entire run.
-When that option is on, of course the number of processor cores used to start
-the code in the batch system must be a multiple of NUMBER\_OF\_SIMULTANEOUS\_RUNS,
-all the individual runs must use the same number of processor cores,
-which as usual is NPROC in the Par\_file,
-and thus the total number of processor cores to request from the batch system
-should be NUMBER\_OF\_SIMULTANEOUS\_RUNS * NPROC.
-All the runs to perform must be placed in directories called run0001, run0002, run0003 and so on
-(with exactly four digits).
-
-Imagine you have 10 independent calculations to do, each of them on 100 cores; you have three options:
-
-1/ submit 10 jobs to the batch system
-
-2/ submit a single job on 1000 cores to the batch, and in that script create a sub-array of jobs to start 10 jobs,
-each running on 100 cores (see e.g. http://www.schedmd.com/slurmdocs/job\_array.html )
-
-3/ submit a single job on 1000 cores to the batch, start SPECFEM3D on 1000 cores, create 10 sub-communicators,
-cd into one of 10 subdirectories (called e.g. run0001, run0002,... run0010) depending on the sub-communicator
-your MPI rank belongs to, and run normally on 100 cores using that sub-communicator.
-
-The option NUMBER\_OF\_SIMULTANEOUS\_RUNS implements 3/.
-
-\item [{\texttt{BROADCAST\_SAME\_MESH\_AND\_MODEL}}]:
-if we perform simultaneous runs in parallel, if only the source and receivers vary between these runs
-but not the mesh nor the model (velocity and density) then we can also read the mesh and model files
-from a single run in the beginning and broadcast them to all the others; for a large number of simultaneous
-runs for instance when solving inverse problems iteratively this can DRASTICALLY reduce I/Os to disk in the solver
-(by a factor equal to NUMBER\_OF\_SIMULTANEOUS\_RUNS), and reducing I/Os is crucial in the case of huge runs.
-Thus, always set this option to .true. if the mesh and the model are the same for all simultaneous runs.
-In that case there is no need to duplicate the mesh and model file database (the content of the DATABASES\_MPI
-directories) in each of the run0001, run0002,... directories, it is sufficient to have one in run0001
-and the code will broadcast it to the others).
-
-TODO / future work to do: currently the BROADCAST\_SAME\_MESH\_AND\_MODEL option assumes to have the (master) mesh files in run0001/DATABASES\_MPI or run0001/OUTPUT\_FILES/DATABASES\_MPI.
-However, for adjoint runs you still need a DATABASES\_MPI folder in each of the sub-runs directories, e.g.
-run0002/DATABASES\_MPI, etc.
-to store the forward wavefields, kernels etc.
-of each sub-run.
-This would not be needed for forward simulations.
-
-TODO / future work to do: the sensitivity kernel summing and smoothing tools in directory src/tomography are currently not ported to this new option to do many runs simultaneously, only the solver (src/specfem3d) is.
-Thus these tools should work, but in their current version will need to be run for each simulation result independently.
-More precisely, the current kernel summing and smoothing routines work fine, with the exception that you need to move out the mesh files (and also the parameters).
-This works because these routines consider multiple runs by design.
-You simply have to provide them the directories where the kernels are.
-
-\item [{\texttt{GPU\_MODE}}] Turn this flag on to use GPUs.
-
-\item [\texttt{ADIOS\_ENABLED}] Turn this flag on to enable ADIOS. If set to \texttt{.false.}, subsequent ADIOS
-parameters will not be considered.
-\item [\texttt{ADIOS\_FOR\_DATABASES}] Turn this flag on to use ADIOS for xmeshfem3D output and
-xgenerate\_database input.
-\item [\texttt{ADIOS\_FOR\_MESH}]  Turn this flag on to use ADIOS for generated databases.
-\item [\texttt{ADIOS\_FOR\_FORWARD\_ARRAYS}] Turn this flag on to read and write forward arrays using ADIOS.
-\item [\texttt{ADIOS\_FOR\_KERNELS}] Turn this flag on to produce ADIOS kernels that can later be visualized with the ADIOS version of combine\_vol\_data.
-
-\end{description}
-
-The present version of SPECFEM can handle fully saturated porous simulations, Christina Morency implemented Biot equation.
-But the code cannot calculate partially saturated cases in its state.
-Christina Morency is presently working on a dual porosity, dual permeability formulation, type Pride and Berryman, for an other project, but it will not be available for some time.\\
-
-The way we prescribe material property for porous material in SPECFEM3D is as follow:
-We use a file name nummaterial\_poroelastic\_file, which is located in the directory MESH, the format is as follow:\\
- rhos rhof phi c kxx kxy kxz kyy kyz kzz Ks Kf Kfr etaf mufr\\
-where\\
-rho\_s = solid density\\
-rho\_f = fluid density\\
-phi = porosity\\
-tort = tortuosity\\
-kxx = xx component of permeability tensor\\
-kxy = xy,yx components of permeability tensor\\
-kyy = yy component of permeability tensor\\
-kxz = xz,zx components of permeability tensor\\
-kzz = zz component of permeability tensor\\
-kappa\_s = solid bulk modulus\\
-kappa\_f= fluid bulk modulus\\
-kappa\_fr= frame bulk modulus\\
-eta\_f = fluid viscosity\\
-mu\_fr = frame shear modulus.\\
-
-Using an external mesh (for instance coming from CUBIT/TRELIS), poroelastic materials have the ID number 3, while 1 is elastic, 2 is acoustic ?
-see the example in the package: examples/homogeneous\_poroelastic.
-
-If you use PML, the mesh elements that belong to the PML layers can
-be acoustic or elastic, but not viscoelastic nor poroelastic. Then,
-when defining your model, you should define these absorbing elements
-as either acoustic or elastic. In you forget to do that, the code
-will fix the problem by automatically converting the viscoelastic
-or poroelastic PML elements to elastic. This means that strictly speaking
-the PML layer will not be perfectly matched any more, since the physical
-model will change from viscoelastic or poroelastic to elastic at the
-entrance of the PML, but in practice this is sufficient and produces
-only tiny / negligible spurious reflections.
-
-If you use PML and an external mesh (created using an external meshing tool
-such as CUBIT/TRELIS or similar), try to have elements inside the PML as regular as possible,
-i.e. ideally non-deformed cubes obtained by `extrusion' of regular surface mesh elements meshing the
-outer edges of the computational domain without PML; by doing so, the PMLs obtained will be far more stable
-in time (PML being weakly unstable from a mathematical point of view, very deformed mesh elements
-inside the PMLs can trigger instabilities much more quickly).
-
-If you use PML and an external tomographic velocity and density model,
-you should be careful because mathematically a PML cannot handle heterogeneities
-along the normal to the PML edge inside the PML layer. This comes
-from the fact that the damping profile that is defined assumes a constant
-velocity and density model along the normal direction.
-
-Thus, you need to modify your velocity and density model in order
-for it to be 1D inside the PML, as shown in Figure~\ref{fig:modify_external_velocity_model_to_use_PML}.
-
-This applies to the bottom layer as well; there you should make sure
-that your model is 1D and thus constant along the vertical direction.
-
-To summarize, only use a 3D velocity and density model inside the
-physical region, and in all the PML layers extend it by continuity
-from its values along the inner PML edge.
-
-%%
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=3in]{figures/how_to_use_PML_when_a_tomographic_velocity_model_is_used}
-\par\end{centering}
-
-\caption{How to modify your external 3D velocity and density model in order
-to use PML. Such a modification is not needed when using Stacey absorbing
-boundary conditions (but such conditions are significantly less efficient).}
-
-\label{fig:modify_external_velocity_model_to_use_PML}
-\end{figure}
-
-%%
-
-\section{\label{sec:Choosing-the-Time-Step}Choosing the time step \texttt{DT}}
-
-The parameter \texttt{DT} sets the length of each time step in seconds.
-The value of this parameter is crucial for the stability of the spectral-element
-simulation. Your time step \texttt{DT} will depend on the minimum
-ratio between the distance $h$ of neighboring mesh points and the
-wave speeds $v$ defined in your model. The condition for the time
-step $\Delta t$ is:
-\begin{lyxcode}
-$\Delta t<C~\mathrm{min}_{\Omega}(~h/v~)$
-\end{lyxcode}
-where $C$ is the so-called Courant number and $\Omega$ denotes the
-model volume. The distance $h$ depends on the mesh element size and
-the number of GLL points \texttt{NGLL} specified in the main constants
-file \texttt{constants.h} located in the \texttt{src/shared/} subdirectory.
-The wave speed $v$ is determined based on your model's P- (or S-)
-wave speed values.
-
-The database generator \texttt{xgenerate\_databases}, as well as the
-internal mesher \texttt{xmeshfem3D}, are trying to evaluate the value
-of $\Delta t$ for empirically chosen Courant numbers $C\sim0.3$.
-If you used the mesher \texttt{xmeshfem3D} to generate your mesh,
-you should set the value suggested in \texttt{OUTPUT\_FILES/output\_mesher.txt}
-file, which is created after the mesher completed. In case you used
-CUBIT to create the mesh, you might use an arbitrary value when running
-\texttt{xgenerate\_databases} and then use the value suggested in
-the ~\\
- \texttt{OUTPUT\_FILES/output\_mesher.txt} file after the database
-generation completed. Note that the implemented Newmark time scheme
-uses this time step globally, thus your simulations become more expensive
-for very small mesh elements in high wave-speed regions. Please be
-aware of this restriction when constructing your mesh in Chapter~\ref{cha:Mesh-Generation}.
-
-\chapter{\label{cha:Running-the-Solver}Running the Solver \texttt{xspecfem3D}}
-
-Now that you have successfully generated the databases, you are ready
-to compile the solver. In the main directory, type
-\begin{lyxcode}
-{\small make~xspecfem3D}{\small \par}
-\end{lyxcode}
-Please note that \texttt{xspecfem3D} must be called directly from
-the main directory, as most of the binaries of the package.\\
-
-\noindent
-The solver needs three input files in the \texttt{DATA} directory
-to run:
-\begin{description}
-\item [{\texttt{Par\_file}}] the main parameter file which was discussed
-in detail in the previous Chapter~\ref{cha:Creating-Distributed-Databases},
-\item [{\texttt{CMTSOLUTION}} or {\texttt{FORCESOLUTION}}] the earthquake
-source parameter file or the force source parameter file, and
-\item [{\texttt{STATIONS}}] the stations file.
-\end{description}
-Most parameters in the \texttt{Par\_file} should be set prior to running
-the databases generation. Only the following parameters may be changed
-after running \texttt{xgenerate\_databases}:
-\begin{itemize}
-\item the simulation type control parameters: \texttt{SIMULATION\_TYPE}
-and \texttt{SAVE\_FORWARD}
-\item the time step parameters \texttt{NSTEP} and \texttt{DT}
-\item the absorbing boundary control parameter \texttt{PML\_CONDITIONS}
-on condition that the\\
- \texttt{PML\_INSTEAD\_OF\_FREE\_SURFACE} flag remains unmodified
-after running the databases generation.
-\item the movie control parameters \texttt{MOVIE\_SURFACE}, \texttt{MOVIE\_VOLUME},
-and \texttt{NTSTEPS\_BETWEEN\_FRAMES}
-\item the ShakeMap\textregistered{}option \texttt{CREATE\_SHAKEMAP}
-\item the output information parameters \texttt{MOVIE\_TYPE}, \texttt{NTSTEP\_BETWEEN\_OUTPUT\_INFO}
-and \texttt{NTSTEP\_BETWEEN\_OUTPUT\_}~\\
- \texttt{SEISMOS}
-\item the \texttt{PRINT\_SOURCE\_TIME\_FUNCTION} flags
-\end{itemize}
-Any other change to the \texttt{Par\_file} implies rerunning both
-the database generator \texttt{xgenerate\_databases} and the solver
-\texttt{xspecfem3D}.
-
-For any particular earthquake, the \texttt{CMTSOLUTION} file that
-represents the point source may be obtained directly from the Harvard
-Centroid-Moment Tensor (CMT) web page \urlwithparentheses{www.seismology.harvard.edu}.
-It looks like the example shown in Fig. \ref{fig:CMTSOLUTION-file}.
-%\begin{lyxcode}
-\begin{figure}[htp]
-\noindent
-\begin{centering}
-{\small \includegraphics[width=0.8\textwidth]{figures/Hollywood_CMT}
-}
-\par
-\end{centering}
-\caption{\texttt{CMTSOLUTION} file based on the
-format from the Harvard CMT catalog. \textbf{M} is the moment tensor,
-$M_{0}${\small {} }is the seismic moment, and $M_{w}$ is the moment
-magnitude.}
-\label{fig:CMTSOLUTION-file}
-\end{figure}
-%\end{lyxcode}
-The \texttt{CMTSOLUTION} file should be edited in the following way:
-\begin{itemize}
-\item Set the latitude or UTM $x$ coordinate, longitude or UTM $y$ coordinate, depth of the source (in km).
-\item Set the \texttt{time shift} parameter equal to $0.0$ (the solver
-will not run otherwise.) The time shift parameter would simply apply
-an overall time shift to the synthetics, something that can be done
-in the post-processing (see Section \ref{sec:Process-data-and-syn}).
-\item For point-source simulations (see finite sources, page \pageref{To-simulate-a})
-we recommend setting the source half-duration parameter \texttt{half
-duration} equal to zero, which corresponds to simulating a step source-time
-function, i.e., a moment-rate function that is a delta function. If
-\texttt{half duration} is not set to zero, the code will use a Gaussian
-(i.e., a signal with a shape similar to a `smoothed triangle', as
-explained in \citet{KoTr02a} and shown in Fig~\ref{fig:gauss.vs.triangle})
-source-time function with half-width \texttt{half duration}. We prefer
-to run the solver with \texttt{half duration} set to zero and convolve
-the resulting synthetic seismograms in post-processing after the run,
-because this way it is easy to use a variety of source-time functions
-(see Section \ref{sec:Process-data-and-syn}). \citet{KoTr02a} determined
-that the noise generated in the simulation by using a step source
-time function may be safely filtered out afterward based upon a convolution
-with the desired source time function and/or low-pass filtering. Use
-the serial code \texttt{convolve\_source\_timefunction.f90} and the
-script \texttt{convolve\_source\_timefunction.csh} for this purpose,
-or alternatively use signal-processing software packages such as SAC
-\urlwithparentheses{www.llnl.gov/sac}. Type
-
-\begin{lyxcode}
-{\small make~xconvolve\_source\_timefunction}{\small \par}
-\end{lyxcode}
-
-to compile the code and then set the parameter \texttt{hdur} in \texttt{convolve\_source\_timefunction.csh}
-to the desired half-duration.
-
-\item The zero time of the simulation corresponds to the center of the triangle/Gaussian,
-or the centroid time of the earthquake. The start time of the simulation
-is $t=-1.5*\texttt{half duration}$ (the 1.5 is to make sure the moment
-rate function is very close to zero when starting the simulation).
-To convert to absolute time $t_{\mathrm{abs}}$, set
-
-\begin{lyxcode}
-$t_{\mathrm{abs}}=t_{\mathrm{pde}}+\texttt{time shift}+t_{\mathrm{synthetic}}$
-\end{lyxcode}
-
-where $t_{\mathrm{pde}}$ is the time given in the first line of the
-\texttt{CMTSOLUTION}, \texttt{time shift} is the corresponding value
-from the original \texttt{CMTSOLUTION} file and $t_{\mathrm{synthetic}}$
-is the time in the first column of the output seismogram.
-
-\end{itemize}
-\begin{figure}[htp]
-\noindent \begin{centering}
-\includegraphics[width=3.5in]{figures/gauss_vs_triangle_mod}
-\par\end{centering}
-
-\caption{Comparison of the shape of a triangle and the Gaussian function actually
-used.}
-
-\label{fig:gauss.vs.triangle}
-\end{figure}
-
-If you know the earthquake source in strike/dip/rake format rather
-than in \texttt{CMTSOLUTION} format, use the C code \texttt{SPECFEM3D\_GLOBE/utils/strike\_dip\_rake\_to\_CMTSOLUTION.c}
-to convert it. The conversion formulas are given for instance in \citet{AkRi80}.
-Note that the \citet{AkRi80} convention is slightly different from
-the Harvard \texttt{CMTSOLUTION} convention (the sign of some components
-is different). The C code outputs both.
-
-Centroid latitude and longitude should be provided in geographical
-coordinates. The code converts these coordinates to geocentric coordinates~\citep{DaTr98}.
-Of course you may provide your own source representations by designing
-your own \texttt{CMTSOLUTION} file. Just make sure that the resulting
-file adheres to the Harvard CMT conventions (see Appendix~\ref{cha:Coordinates}).
-Note that the first line in the \texttt{CMTSOLUTION} file is the Preliminary
-Determination of Earthquakes (PDE) solution performed by the USGS
-NEIC, which is used as a seed for the Harvard CMT inversion. The PDE
-solution is based upon P waves and often gives the hypocenter of the
-earthquake, i.e., the rupture initiation point, whereas the CMT solution
-gives the `centroid location', which is the location with dominant
-moment release. The PDE solution is not used by our software package
-but must be present anyway in the first line of the file.
-
-\label{To-simulate-a}To simulate a kinematic rupture, i.e., a finite-source
-event, represented in terms of $N_{\mathrm{sources}}$ point sources,
-provide a \texttt{CMTSOLUTION} file that has $N_{\mathrm{sources}}$
-entries, one for each subevent (i.e., concatenate $N_{\mathrm{sources}}$
-\texttt{CMTSOLUTION} files to a single \texttt{CMTSOLUTION} file).
-At least one entry (not necessarily the first) must have a zero \texttt{time
-shift}, and all the other entries must have non-negative \texttt{time
-shift}. Each subevent can have its own half duration, latitude, longitude,
-depth, and moment tensor (effectively, the local moment-density tensor).
-
-Note that the zero in the synthetics does NOT represent the hypocentral
-time or centroid time in general, but the timing of the \textit{center}
-of the source triangle with zero \texttt{time shift} (Fig~\ref{fig:source_timing}).
-
-Although it is convenient to think of each source as a triangle, in
-the simulation they are actually Gaussians (as they have better frequency
-characteristics). The relationship between the triangle and the gaussian
-used is shown in Fig~\ref{fig:gauss.vs.triangle}. For finite fault
-simulations it is usually not advisable to use a zero half duration
-and convolve afterwards, since the half duration is generally fixed
-by the finite fault model.
-
-\vspace{1cm}
-
-\noindent The \texttt{FORCESOLUTION} file should be edited in the
-following way:
-\begin{itemize}
-\item Set the \texttt{time shift} parameter equal to $0.0$ (the solver
-will not run otherwise.) The time shift parameter would simply apply
-an overall time shift to the synthetics, something that can be done
-in the post-processing (see Section \ref{sec:Process-data-and-syn}).
-\item Set the \texttt{f0} parameter (the dominant frequency) of the Ricker source time
-function when {\texttt{USE\_RICKER\_TIME\_FUNCTION}} is turned on
-in the main parameter file {\texttt{Par\_file}}. In case that the
-solver uses a (pseudo) Dirac delta source time function to represent
-a force point source, a very short duration of five time steps is automatically set by default.
-\item Set the latitude or UTM $x$ coordinate, longitude or UTM $y$ coordinate, depth of the source (in km).
-\item Set the magnitude of the force source.
-\item Set the components of a (non-unitary) direction vector for the force
-source in the East/North/Vertical basis (see Appendix A for the orientation
-of the reference frame).
-\end{itemize}
-\noindent Where necessary, set a \texttt{FORCESOLUTION} file in the
-same way you configure a \texttt{CMTSOLUTION} file with $N_{\mathrm{sources}}$
-entries, one for each subevent (i.e., concatenate $N_{\mathrm{sources}}$
-\texttt{FORCESOLUTION} files to a single \texttt{FORCESOLUTION} file).
-At least one entry (not necessarily the first) must have a zero \texttt{time
-shift}, and all the other entries must have non-negative \texttt{time
-shift}. Each subevent can have its own half latitude, longitude, depth,
-\texttt{half duration} and force parameters.
-
-\begin{figure}[H]
-\noindent \begin{centering}
-{\small \includegraphics[width=5in]{figures/source_timing} }
-\par\end{centering}
-
-\caption{Example of timing for three sources. The center of the first source
-triangle is defined to be time zero. Note that this is NOT in general
-the hypocentral time, or the start time of the source (marked as tstart).
-The parameter \texttt{time shift} in the \texttt{CMTSOLUTION} file
-would be t1(=0), t2, t3 in this case, and the parameter \texttt{half
-duration} would be hdur1, hdur2, hdur3 for the sources 1, 2, 3 respectively.}
-
-{\small \label{fig:source_timing} }
-\end{figure}
-
-The solver can calculate seismograms at any number of stations for
-basically the same numerical cost, so the user is encouraged to include
-as many stations as conceivably useful in the \texttt{STATIONS} file,
-which looks like this:
-
-\begin{figure}[H]
-\noindent \begin{centering}
-{\small \includegraphics{figures/STATIONS_basin_explained} }
-\par\end{centering}
-
-\label{fig:Sample-STATIONS-file.}\caption{Sample \texttt{STATIONS} file. Station latitude and longitude should
-be provided in geographical coordinates. The width of the station
-label should be no more than 32 characters (see \texttt{MAX\_LENGTH\_STATION\_NAME}
-in the \texttt{constants.h} file), and the network label should be
-no more than 8 characters (see \texttt{MAX\_LENGTH\_NETWORK\_NAME}
-in the \texttt{constants.h} file).}
-\end{figure}
-
-Each line represents one station in the following format:
-\begin{lyxcode}
-{\small Station~Network~Latitude~(degrees)~Longitude~(degrees)~Elevation~(m)~burial~(m)~}{\small \par}
-\end{lyxcode}
-The solver \texttt{xspecfem3D} filters the list of stations in file
-\texttt{DATA/STATIONS} to exclude stations that are not located within
-the region given in the \texttt{Par\_file} (between \texttt{LATITUDE\_MIN}
-and \texttt{LATITUDE\_MAX} and between \texttt{LONGITUDE\_MIN} and
-\texttt{LONGITUDE\_MAX}). The filtered file is called \texttt{DATA/STATIONS\_FILTERED}.
-
-Solver output is provided in the \texttt{OUTPUT\_FILES} directory
-in the \texttt{output\_solver.txt} file. Output can be directed to
-the screen instead by uncommenting a line in \texttt{constants.h}:
-\begin{lyxcode}
-!~uncomment~this~to~write~messages~to~the~screen~
-
-!~integer,~parameter~::~IMAIN~=~ISTANDARD\_OUTPUT~
-\end{lyxcode}
-On PC clusters the seismogram files are generally written to the local
-disks (the path \texttt{LOCAL\_PATH} in the \texttt{Par\_file}) and
-need to be gathered at the end of the simulation.
-
-While the solver is running, its progress may be tracked by monitoring
-the `\texttt{\small timestamp{*}}{\small ' files in the ~}\\
-{\small{} }\texttt{\small OUTPUT\_FILES}{\small{} directory. These tiny
-files look something like this:}{\small \par}
-\begin{lyxcode}
-Time~step~\#~~~~~~~~~~10000~
-
-Time:~~~~~108.4890~~~~~~seconds~
-
-Elapsed~time~in~seconds~=~~~~1153.28696703911~
-
-Elapsed~time~in~hh:mm:ss~=~~~~~0~h~19~m~13~s~
-
-Mean~elapsed~time~per~time~step~in~seconds~=~~~~~0.115328696703911~
-
-Max~norm~displacement~vector~U~in~all~slices~(m)~=~~~~1.0789589E-02~
-\end{lyxcode}
-The \texttt{\small timestamp{*}}{\small{} files provide the }\texttt{\small Mean
-elapsed time per time step in seconds}{\small , which may be used
-to assess performance on various machines (assuming you are the only
-user on a node), as well as the }\texttt{\small Max norm displacement
-vector U in all slices~(m)}{\small . If something is wrong with the
-model, the mesh, or the source, you will see the code become unstable
-through exponentially growing values of the displacement and fluid
-potential with time, and ultimately the run will be terminated by
-the program. You can control the rate at which the timestamp files
-are written based upon the parameter }\texttt{\small NTSTEP\_BETWEEN\_OUTPUT\_INFO}{\small{}
-in the }\texttt{\small Par\_file}{\small .}{\small \par}
-
-Having set the \texttt{Par\_file} parameters, and having provided
-the \texttt{CMTSOLUTION} (or the \texttt{FORCESOLUTION}) and \texttt{STATIONS}
-files, you are now ready to launch the solver! This is most easily
-accomplished based upon the \texttt{go\_solver} script (See Chapter
-\ref{cha:Scheduler} for information about running through a scheduler,
-e.g., LSF). You may need to edit the last command at the end of the
-script that invokes the \texttt{mpirun} command. The \texttt{runall}
-script compiles and runs both \texttt{xgenerate\_databases} and \texttt{xspecfem3D}
-in sequence. This is a safe approach that ensures using the correct
-combination of distributed database output and solver input.
-
-It is important to realize that the CPU and memory requirements of
-the solver are closely tied to choices about attenuation (\texttt{ATTENUATION})
-and the nature of the model (i.e., isotropic models are cheaper than
-anisotropic models). We encourage you to run a variety of simulations
-with various flags turned on or off to develop a sense for what is
-involved.
-
-For the same model, one can rerun the solver for different events
-by simply changing the \texttt{CMTSOLUTION} or \texttt{FORCESOLUTION}
-file, or for different stations by changing the \texttt{STATIONS}
-file. There is no need to rerun the \texttt{xgenerate\_databases}
-executable. Of course it is best to include as many stations as possible,
-since this does not add to the cost of the simulation.
-
-We have also added the ability to run several calculations (several earthquakes)
-in an embarrassingly-parallel fashion from within the same run;
-this can be useful when using a very large supercomputer to compute
-many earthquakes in a catalog, in which case it can be better from
-a batch job submission point of view to start fewer and much larger jobs,
-each of them computing several earthquakes in parallel.
-To turn that option on, set parameter \texttt{NUMBER\_OF\_SIMULTANEOUS\_RUNS}
-to a value greater than 1 in file setup/constants.h.in before
-configuring and compiling the code.
-When that option is on, of course the number of processor cores used to start
-the code in the batch system must be a multiple of \texttt{NUMBER\_OF\_SIMULTANEOUS\_RUNS},
-all the individual runs must use the same number of processor cores,
-which as usual is \texttt{NPROC} in the input file \texttt{DATA/Par\_file},
-and thus the total number of processor cores to request from the batch system
-should be \texttt{NUMBER\_OF\_SIMULTANEOUS\_RUNS $\times$ NPROC}.
-All the runs to perform must be placed in directories called \texttt{run0001}, \texttt{run0002}, \texttt{run0003} and so on (with exactly four digits)
-and you must create a link from the root directory of the code
-to the first copy of the executable programs by typing \texttt{ln -s run0001/bin bin}.
-
-\chapter{\label{cha:Fault-Kinematics-Dynamics}Kinematic and dynamic fault
-sources}
-
-SPECFEM3D can handle finite fault sources of two kinds:
-\begin{enumerate}
-\item \emph{Kinematic}: the spatio-temporal distribution of slip rate is
-prescribed along the fault surface
-\item \emph{Dynamic}: a friction law and initial stresses are prescribed
-on the fault, a spontaneous rupture process is computed.
-\end{enumerate}
-
-\section{\label{sec:Mesh-Generation-with}Mesh Generation with Split Nodes}
-
-Faults need to be handled in a special way during mesh generation.
-A fault surface must lie at the interface between elements (the mesh
-must honor the fault surfaces). Moreover, a fault is made of two surfaces
-in contact. Each of these two surfaces needs a separate set of nodes.
-This approach is known as \textquotedbl{}split nodes\textquotedbl{}.
-Currently faults can only be run with \texttt{xdecompose\_mesh} and
-\texttt{CUBIT}. \texttt{xmeshfem3D} is not yet ready to handle faults.\\
-To facilitate the mesh generation with split nodes in CUBIT, we need
-to separate the two fault surfaces by a small distance, effectively
-creating a tiny opening of the fault (Figure \ref{fig:examples.splitnodes},
-\ref{fig:examples.splitnodes-surfacetrace}). Note that only the interior
-of the fault must be opened, its edges must remain closed (except
-the edge on the free surface). The fault is automatically closed later
-by SPECFEM3D.\\
-
-Here is an example CUBIT script to generate a mesh with split nodes
-for a buried vertical strike-slip fault:
-\begin{lyxcode}
-~reset~~\\
-~brick~x~10~y~10~z~10~~\\
-~webcut~volume~all~with~plane~xplane~~\\
-~webcut~volume~all~with~plane~yplane~~\\
-~webcut~volume~all~with~plane~xplane~offset~3~~\\
-~webcut~volume~all~with~plane~zplane~offset~3~~\\
-~webcut~volume~all~with~plane~zplane~offset~-3~~\\
-~imprint~all~~\\
-~merge~all~~\\
-~unmerge~surf~160~~\\
-~mesh~vol~all~~\\
-~set~node~constraint~off~~\\
-~node~in~surf~168~move~X~0~Y~0.01~Z~0~~\\
-~node~in~surf~160~move~X~0~Y~-0.01~Z~0~\\
-
-\end{lyxcode}
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[scale=0.3]{figures/faultmesh} \\
- \includegraphics[scale=0.3]{figures/surf168} \includegraphics[scale=0.3]{figures/splitnodes}
-\par\end{centering}
-
-\caption{Screenshots of the CUBIT example for the embedded fault with split
-nodes: The entire mesh is decomposed into several volumes bounding
-the fault (top), the fault surface 168 is shown as orange squares
-(middle) and split nodes of the fault shown as orange dots(bottom).
-Note that only interior nodes of the fault are split while those on
-the edges of the fault surface are touching each other.}
-
-\label{fig:examples.splitnodes}
-\end{figure}
-
-\begin{lyxcode}
-
-\end{lyxcode}
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[scale=0.55]{figures/splitnodes_surfacetrace} \\
-
-\par\end{centering}
-
-\caption{Screenshots of a CUBIT example showing split nodes for a fault reaching
-the surface. Surface trace of the fault is shown in orange. Note that
-the edges of the fault are not split while the interior nodes are
-offset by a small distance on either side of the fault}
-
-\label{fig:examples.splitnodes-surfacetrace}
-\end{figure}
-
-The CUBIT scripts ({*}.jou and {*}.py) in the directory \texttt{EXAMPLES}
-generate more complicated meshes. The {*}.py files are Python scripts
-that execute CUBIT commands and use the CUBIT-python interface for
-SPECFEM3D (see next section). The Python language allows to define
-and manipulate variables to parameterize the mesh. Alternatively,
-the Python script can call a CUBIT journal file ({*}.jou), which looks
-like the example above. Variables can be defined and manipulated there
-using the $\mathtt{APREPRO}$ language built in CUBIT.\\
-
-Note that you should avoid gaps in the list of indices of mesh objects
-with the following CUBIT command:
-\begin{lyxcode}
-$\mathtt{compress\: ids\: hex\: face\: edge\: node}$
-\end{lyxcode}
-(otherwise you will get a segmentation fault during domain decomposition)
-
-\section{CUBIT-Python Scripts for Faults}
-
-The mesh generated in CUBIT needs to be processed and exported in
-a format compatible with SPECFEM3D. This is achieved in the Python
-scripts by calling the Python-CUBIT interface functions defined in
-the CUBIT directory:
-\begin{enumerate}
-\item Function $\mathtt{define\_bc}$ (or $\mathtt{boundary\_definition.py}$)
-must be called to set up the absorbing boundaries database
-\item Function $\mathtt{fault\_input}$ must be called once for each fault
-to set up the fault database
-\item Function $\mathtt{cubit2specfem3d.export2SPECFEM3D}$ must be called
-at the very end of the script to export the mesh in a SPECFEM3D format.
-\end{enumerate}
-The functions in \#1 and \#3 are for the bulk and are documented in
-Section \ref{subsec:Exporting-the-Mesh}. We focus here on \#2:
-\begin{description}
-\item [{Function:}] $\mathtt{fault\_input}$
-\item [{Purpose:}] export a fault mesh database from CUBIT to a SPECFEM3D-compliant
-file
-\item [{Syntax:}] fault\_input(id\_fault, ids\_surf\_1, ids\_surf\_2)
-\item [{Inputs:}]~\end{description}
-\begin{itemize}
-\item id\_fault integer index assigned to the fault. (The user must number
-all the faults, starting at 1, with unit increments)
-\item ids\_surf\_1 list of CUBIT ids of all surfaces that form side 1 of
-the fault.
-\item ids\_surf\_2 list of CUBIT ids of all surfaces that form side 2 of
-the fault. (The user must decide which side of the fault is side 1.
-This choice affects the sign conventions of fault quantities as explained
-in Section \ref{sec:Sign-Convention-for}).\end{itemize}
-\begin{description}
-\item [{Outputs:}] file $\mathtt{fault\_file\_X.dat}$, where X is the
-fault id (id\_fault).
-\item [{Example:}] For the example in Section \ref{sec:Mesh-Generation-with}:\end{description}
-\begin{lyxcode}
-$\hphantom{}$A1~=~{[}168{]}
-
-$\hphantom{}$A2~=~{[}160{]}
-
-$\hphantom{}$fault\_input(1,A1,A2)
-\end{lyxcode}
-
-\section{Examples}
-
-The package includes examples, the SCEC benchmark problems:
-\begin{itemize}
-\item TPV5, a planar vertical strike-slip fault
-\item TPV14 and TPV15, vertical strike-slip fault system with a fault branch
-\item TPV102 and TPV103, rate-and-state benchmarks
-\item TPV22 and TPV23, step-over benchmarks
-\item TPV16, heterogenous initial stress
-\end{itemize}
-and
-\begin{itemize}
-\item Splay fault models from Wendt et al. (2009)\\
-
-\end{itemize}
-Read the documents in the directory \texttt{EXAMPLES/*/description}.
-They contain a description of the example and additional instructions
-to run it. Visualize the results with the Matlab scripts in the directory
-$\mathtt{EXAMPLES/*/post}$
-
-\section{\label{sec:Sign-Convention-for}Sign Convention for Fault Quantities}
-
-During mesh generation, the fault is defined by two surfaces in contact.
-Let's denote as \textquotedbl{}\emph{side 1}\textquotedbl{} the SECOND
-surface declared by the user in the call to the python function \textquotedbl{}\emph{fault\_input}\textquotedbl{},
-and the FIRST surface as \textquotedbl{}\emph{side 2}\textquotedbl{}.
-The local coordinate system on the fault is defined as the right-handed
-coordinate system defined by (strike, dip, normal), where \textquotedbl{}\emph{normal}\textquotedbl{}
-is the normal vector outgoing from side 1, \textquotedbl{}\emph{dip}\textquotedbl{}
-is parallel to the along-dip direction pointing downwards, and \textquotedbl{}\emph{strike}\textquotedbl{}
-is the horizontal along-strike vector such that the system is right-handed.\\
-
-Slip is defined as displacement on side 2 minus displacement on side
-1. In the local coordinate system on the fault, positive along-strike
-slip is right-lateral and positive along-dip slip is thrust if side
-1 is on the hanging wall (normal faulting if side 1 is on the foot
-wall).\\
-
-Traction is defined as the stress induced on side 1 by side 2, which
-is the stress tensor times the normal vector outgoing from side 1.
-In the local coordinate system on the fault, the normal traction is
-negative in compression, positive along-strike traction generates
-right-lateral slip and positive along-dip traction generates thrust
-slip if side 1 is on the hanging wall (normal faulting if side 1 is
-on the foot wall).
-
-\section{Input Files}
-
-Three additional input files are required in the \texttt{DATA/} directory
-for dynamic and kinematic faults. They are $\mathtt{Par\_file\_faults}$,
-$\mathtt{FAULT\_STATIONS}$ and $\mathtt{input\_file.txt}$ (optional).
-If the former does not exist, the code assumes that there are no faults.\\
-
-\begin{lyxlist}{00.00.0000}
-\item [{\textbf{DATA/Par\_file\_faults}}] contains parameters of the fault.
-The first part of this file has a strict format:
-
-\begin{description}
-\item [{Line}] 1: Number of faults (NF)
-\item [{Lines}] 2 to NF+1: Kelvin Voigt damping (in seconds) for each fault.
-(See below how to set this parameter)
-\item [{Line}] NF+2: Type of simulation (1=dynamic , 2 = kinematic)
-\item [{Line}] NF+3: Number of time steps between updates of the time series
-outputs at selected fault points (see DATA/FAULT\_STATIONS), usually
-a large number (100s or 1000s). Note that the sampling rate of the
-time series is usually much higher.
-\item [{Line}] NF+4: Number of time steps between fault snapshot outputs
-(quantities at every fault point exported at regular times), usually
-a large number (100s or 1000s).
-\item [{Line}] NF+5: Slip velocity threshold below which frictional healing
-is set (friction coefficient is reset to its static value). If this
-value is negative healing is disabled.
-\item [{Line}] NF+6: Slip velocity threshold to define the rupture front.
-Only used for outputs.
-\end{description}
-
-The rest of this file is made of namelist input blocks (see \textquotedbl{}namelist\textquotedbl{}
-in a Fortran 9x manual). The input for each fault has the following
-sequence (arguments in {[}brackets{]} are optional):
-
-\&\textbf{BEGIN\_FAULT} /
-
-\&\textbf{INIT\_STRESS} S1, S2, S3 {[},n1, n2, n3{]} /
-
-followed by (n1+n2+n3) \&\textbf{DIST2D} blocks
-
-\&\textbf{SWF} mus, mud, dc {[}, nmus, nmud, ndc{]} /
-
-followed by (nmus+nmud+ndc) \&\textbf{DIST2D} blocks\\
-
-\begin{lyxlist}{00.00.0000}
-\item [{\&\textbf{INIT\_STRESS}}] input block sets the initial fault stresses
-relative to the foot-wall side of the fault. Initial stresses are
-composed of a constant background value possibly overwritten in prescribed
-regions by heterogeneous distributions (see \&\textbf{DIST2D} blocks
-below):
-
-\begin{description}
-\item [{S1}] = constant background value of along-strike shear stress (positive
-in the usual strike direction)
-\item [{S2}] = constant background value of along-dip shear (positive is
-down-dip, normal faulting)
-\item [{S3}] = constant background value of normal stress (negative in
-compression)
-\item [{n1}] = number of heterogeneous items for along-strike shear stress
-{[}default is 0{]}
-\item [{n2}] = number of heterogeneous items for along-dip shear stress
-{[}default is 0{]}
-\item [{n3}] = number of heterogeneous items for normal stress {[}default
-is 0{]}
-\end{description}
-\item [{\&\textbf{SWF}}] input block sets the slip-weakening friction parameters
-of the fault:
-
-\begin{description}
-\item [{mus}] = constant background value of static friction coefficient
-\item [{mud}] = constant background value of dynamic friction coefficient
-\item [{dc}] = constant background value of critical slip-weakening distance
-\item [{nmus}] = number of heterogeneous items for static friction coefficient
-{[}default is 0{]}
-\item [{nmud}] = number of heterogeneous items for dynamic friction coefficient
-{[}default is 0{]}
-\item [{ndc}] = number of heterogeneous items for critical slip-weakening
-distance {[}default is 0{]}\\
-
-\end{description}
-\item [{\&\textbf{DIST2D}}] input blocks modify (overwrite) the value of
-a fault parameter by a heterogeneous spatial distribution:
-
-\&\textbf{DIST2D} shape='$\mathtt{square}$', val, xc, yc, zc, l /
-
-$\;$sets a constant value (val) within a cube with center (xc,yc,zc)
-and edge size l.
-
-\&\textbf{DIST2D} shape='$\mathtt{rectangle}$', val, xc, yc, zc,
-lx, ly, lz /
-
-$\;$sets a constant value (val) within a rectangular prism with center
-(xc,yc,zc) and edge sizes (lx,ly,lz).
-
-\&\textbf{DIST2D} shape='$\mathtt{rectangle-taper}$', val, valh,
-xc, yc, zc, lx, ly, lz /
-
-$\;$sets a vertical linear gradient within a rectangular prism with
-center (xc,yc,zc) and edge sizes (lx,ly,lz). Values vary linearly
-as a function of vertical position z between value val at z = zc-lz/2
-and value valh at z = zc+lz/2 .
-
-\&\textbf{DIST2D} shape='$\mathtt{circular}$', val, xc, yc, zc, r
-/
-
-$\;$sets a constant value (val) within a sphere with center (xc,yc,zc)
-and radius r. \\
-
-\end{lyxlist}
-\end{lyxlist}
-
-\begin{lyxlist}{00.00.0000}
-\item [{\textbf{DATA/FAULT\_STATIONS}}] Stations in the fault plane.
-
-\textbf{Line 1} : number of stations.
-
-\textbf{Line 2 to end}: 5 columns: X, Y, Z (-depth), station name,
-fault-id
-
-The fault-id identifies the fault that contains the station. It is
-the index of appearance in the faults list after line 2 of Par\_file\_faults
-
-\end{lyxlist}
-
-\begin{lyxlist}{00.00.0000}
-\item [{\textbf{DATA/input\_file.txt}}] Heterogeneous stresses and friction
-parameters are documented in page 10 of
-
-\begin{lyxlist}{00.00.0000}
-\item [{$\mathtt{EXAMPLES/tpv16/description/TPV16\_17\_Description\_v03.pdf}$}]~
-\end{lyxlist}
-
-To activate this feature, in fault\_solver\_dynamic.f90 set \textbf{TPV16}
-= .true. then re-compile the code.
-
-\end{lyxlist}
-
-\section{\label{sec:Setting-the-Kelvin-Voigt}Setting the Kelvin-Voigt Damping
-Parameter }
-
-The purpose of the Kelvin-Voigt viscosity in the dynamic fault solver
-is to damp spurious oscillations generated by the fault slip at frequencies
-that are too high to be resolved by the mesh. The viscosity $\mathtt{eta}$
-(in seconds) depends on the size of the elements on the fault. Here
-is how to set it:
-\begin{enumerate}
-\item Determine the average linear size of the elements on the fault plane,
-$\mathtt{h\_fault}$. Usually this value is prescribed by the user
-during mesh generation. Otherwise it can be found by inspection of
-the mesh inside the CUBIT GUI.
-\item Use the Matlab function $\mathtt{utils/critical\_timestep.m}$ to
-compute\\
- $\mathtt{dtc\_fault}=\mathtt{critical\_timestep\left(c_{p},h\_fault,ngll\right)}$.
-\\
-This is the critical time step in an elastic medium for a hypothetical
-element of cubic shape with size equal to $\mathtt{h\_fault}$.
-\item Set $\mathtt{eta}$ in $\mathtt{Par\_file\_faults}$ to (0.1 to 0.3)$\ensuremath{\times}\mathtt{dtc\_fault}$.
-A larger $\mathtt{eta}$ damps high-frequencies more aggressively
-but it might also affect lower frequencies and rupture speed.
-\end{enumerate}
-Viscosity reduces numerical stability: the critical timestep in a
-simulation with Kelvin-Voigt damping needs to be smaller than that
-in a purely elastic simulation. Here is how to set the time step accordingly:
-\begin{enumerate}
-\item Run a test simulation without viscosity ($\mathtt{eta}$=0 and only
-a few time steps)
-\item Look for the \textquotedbl{}maximum suggested time step\textquotedbl{}
-in \texttt{OUTPUT\_FILES/output\_mesher.txt}. This is the critical
-timestep of a purely elastic simulation, $\mathtt{dtc\_bulk}$.
-\item Reset the timestep of the simulation with a Kelvin-Voigt material
-to a value smaller than\\
- $\mathtt{dtc\_kv}=\mathtt{eta}\left(\sqrt{1+\mathtt{dtc\_bulk^{2}}/\mathtt{eta^{2}}}-1\right)$
-\end{enumerate}
-Note that in general $\mathtt{dtc\_bulk}$ is smaller than $\mathtt{dtc\_fault}$,
-because elements off the fault might be smaller or more distorted
-than element faces on the fault.
-
-\section{Output Files }
-
-Several output files are saved in \texttt{OUTPUT\_FILES/}:\\
-
-\begin{enumerate}
-\item Seismograms for each station on the fault plane given in \texttt{DATA/FAUL\_STATIONS}.
-One output file is generated for each station, named after the station.
-The files are ascii and start with a header (22 lines long) followed
-by a data block with the following format, one line per time sample:\\
-
-Column 1 = Time (s)
-
-Column 2 = horizontal right-lateral slip (m)
-
-Column 3 = horizontal right-lateral slip rate (m/s)
-
-Column 4 = horizontal right-lateral shear stress (MPa)
-
-Column 5 = vertical up-dip slip (m)
-
-Column 6 = vertical up-dip slip rate (m/s)
-
-Column 7 = vertical up-dip shear stress (MPa)
-
-Column 8 = normal stress (MPa)\\
-
-The stresses are relative to the footwall side of the fault (this
-convention controls their sign, but not their amplitude). Slip is
-defined as displacement of the hanging wall relative to the footwall.
-
-\item Seismograms at stations in the bulk (out of the fault plane) given
-in \texttt{DATA/STATIONS}.
-\item Rupture time files are named $\mathtt{Rupture\_time\_FAULT}$-$\mathtt{id}$.
-One file is generated for each fault. The files are ascii and start
-with a header (12 lines long) followed by a data block with the following
-format, one line per fault node:
-
-Column 1 = horizontal coordinate, distance along strike (m)
-
-Column 2 = vertical coordinate, distance down-dip (m)
-
-Column 3 = rupture time (s)
-
-\item Fault quantities (slip, slip rate, stresses, etc) at regular times
-are stored in binary data files called $\mathtt{Snapshot\#it\#.bin}$,
-where \#it\# is the timestep number. These can be read in Matlab with
-the function $\mathtt{utils/FSEM3D\_snapshot.m}$
-\end{enumerate}
-
-\section{Post-processing and Visualization }
-
-Some Matlab functions for post-processing and visualization are included
-in directory $\mathtt{utils}$. The functions are internally documented
-(see their matlab help).\\
-
-\texttt{FSEM3D\_snapshot} reads a fault data snapshot\\
-
-The directories \texttt{EXAMPLES/*/post} contain additional Matlab
-scripts to generate figures specific to each example.
-
-\chapter{\label{cha:Adjoint-Simulations}Adjoint Simulations}
-
-Adjoint simulations are generally performed for two distinct applications.
-First, they can be used in point source moment-tensor inversions, or source imaging for
-earthquakes with large ruptures such as the Lander's earthquake \citep{WaHe94}.
-Second, they can be used to generate finite-frequency sensitivity
-kernels that are a critical part of tomographic inversions based upon
-3D reference models \citep{TrTaLi05,LiTr06,TrKoLi08,LiTr08}. In either
-case, source parameter or velocity structure updates are sought to
-minimize a specific misfit function (e.g., waveform or traveltime
-differences), and the adjoint simulation provides a means of computing
-the gradient of the misfit function and further reducing it in successive
-iterations. Applications and procedures pertaining to source studies
-and finite-frequency kernels are discussed in Sections~\ref{sec:Adjoint-simulation-sources}
-and \ref{sec:Adjoint-simulation-finite}, respectively. The two related
-parameters in the \texttt{Par\_file} are \texttt{SIMULATION\_TYPE}
-(1, 2 or 3) and the \texttt{SAVE\_FORWARD} (boolean).
-
-\section{\label{sec:Adjoint-simulation-sources}Adjoint Simulations for Sources
-Only (not for the Model)}
-
-When a specific misfit function between data and synthetics is minimized to invert
-for earthquake source parameters, the gradient of the misfit function
-with respect to these source parameters can be computed by placing
-time-reversed seismograms at the receivers as virtual sources
-in an adjoint simulation. Then the value of the gradient is obtained
-from the adjoint seismograms recorded at the original earthquake location.
-\begin{enumerate}
-\item \textbf{Prepare the adjoint sources} \label{enu:Prepare-the-adjoint}
-
-\begin{enumerate}
-\item First, run a regular forward simlation (\texttt{SIMULATION\_TYPE =
-1} and \texttt{SAVE\_FORWARD = .false.}). You can automatically set
-these two variables using the \texttt{\small utils/change\_simulation\_type.pl}{\small{}
-script:}{\small \par}
-
-\begin{lyxcode}
-utils/change\_simulation\_type.pl~-f~
-\end{lyxcode}
-
-and then collect the recorded seismograms at all the stations given
-in \texttt{DATA/STATIONS}.
-
-\item Then select the stations for which you want to compute the time-reversed
-adjoint sources and run the adjoint simulation, and compile them into
-the \texttt{DATA/STATIONS\_ADJOINT} file, which has the same format
-as the regular \texttt{DATA/STATIONS} file.
-
-\begin{itemize}
-\item Depending on what type of misfit function is used for the source inversion,
-adjoint sources need to be computed from the original recorded seismograms
-for the selected stations and saved in a sub-directory called \texttt{SEM/}
-in the root directory of the code, with the format \texttt{NT.STA.BX?.adj}, where \texttt{NT},
-\texttt{STA} are the network code and station name given in the \texttt{DATA/STATIONS\_ADJOINT}
-file, and \texttt{BX?} represents the component name of a particular
-adjoint seismogram. Please note that the band code can change depending
-on your sampling rate (see Appendix~\ref{cha:channel-codes} for
-further details).
-\item The adjoint seismograms are in the same format as the original seismogram
-(\texttt{NT.STA.BX?.sem?}), with the same start time, time interval
-and record length.
-\end{itemize}
-\item Notice that even if you choose to time reverse only one component
-from one specific station, you still need to supply all three components
-because the code is expecting them (you can set the other two components
-to be zero).
-\item Also note that since time-reversal is done in the code itself, no
-explicit time-reversing is needed for the preparation of the adjoint
-sources, i.e., the adjoint sources are in the same forward time sense
-as the original recorded seismograms.
-\end{enumerate}
-\item \textbf{Set the related parameters and run the adjoint simulation}\\
- In the \texttt{DATA/Par\_file}, set the two related parameters to
-be \texttt{SIMULATION\_TYPE = 2} and \texttt{SAVE\_FORWARD = .false.}.
-More conveniently, use the scripts \texttt{utils/change\_simulation\_type.pl}
-to modify the \texttt{Par\_file} automatically (\texttt{change\_simulation\_type.pl
--a}). Then run the solver to launch the adjoint simulation.
-\item \textbf{Collect the seismograms at the original source location}
-
-After the adjoint simulation has completed successfully, collect the
-seismograms from \texttt{LOCAL\_PATH}.
-\begin{itemize}
-\item These adjoint seismograms are recorded at the locations of the original
-earthquake sources given by the \texttt{DATA/CMTSOLUTION} file, and
-have names of the form \texttt{NT.S?????.S??.sem} for the six-component
-strain tensor (\texttt{SNN,SEE,SZZ,SNE,SNZ,SEZ}) at these locations,
-and ~\\
- \texttt{NT.S?????.BX?.sem} for the three-component displacements
-(\texttt{BXN,BXE,BXZ}) recorded at these locations.
-\item \texttt{S?????} denotes the source number; for example, if the original
-\texttt{CMTSOLUTION} provides only a point source, then the seismograms
-collected will start with \texttt{S00001}.
-\item These adjoint seismograms provide critical information for the computation
-of the gradient of the misfit function.
-\end{itemize}
-\end{enumerate}
-
-\section{\label{sec:Adjoint-simulation-finite}Adjoint Simulations for Finite-Frequency
-Kernels (Kernel Simulation)}
-
-Finite-frequency sensitivity kernels are computed in two successive
-simulations (please refer to \citet{LiTr06} and \citet{TrKoLi08}
-for details).
-\begin{enumerate}
-\item \textbf{Run a forward simulation with the state variables saved at
-the end of the simulation}
-
-Prepare the \texttt{\small CMTSOLUTION}{\small{} and }\texttt{\small STATIONS}{\small{}
-files, set the parameters }\texttt{\small SIMULATION\_TYPE}{\small {}
-}\texttt{\small =}{\small {} }\texttt{\small 1}{\small{} and }\texttt{\small SAVE\_FORWARD
-=}{\small {} }\texttt{\small .true.}{\small{} in the }\texttt{\small Par\_file}{\small{}
-(}\texttt{\small change\_simulation\_type -F}{\small ), and run the
-solver.}{\small \par}
-\begin{itemize}
-\item Notice that attenuation is not implemented yet for the computation
-of finite-frequency kernels; therefore set \texttt{ATTENUATION = .false.}
-in the \texttt{Par\_file}.
-\item We also suggest you modify the half duration of the \texttt{CMTSOLUTION}
-to be similar to the accuracy of the simulation (see Equation \ref{eq:shortest_period})
-to avoid too much high-frequency noise in the forward wavefield, although
-theoretically the high-frequency noise should be eliminated when convolved
-with an adjoint wavefield with the proper frequency content.
-\item This forward simulation differs from the regular simulations (\texttt{\small SIMULATION\_TYPE}{\small {}
-}\texttt{\small =}{\small {} }\texttt{\small 1}{\small{} and }\texttt{\small SAVE\_FORWARD}{\small {}
-}\texttt{\small =}{\small {} }\texttt{\small .false.}{\small ) described
-in the previous chapters in that the state variables for the last
-time step of the simulation, including wavefields of the displacement,
-velocity, acceleration, etc., are saved to the }\texttt{\small LOCAL\_PATH}{\small{}
-to be used for the subsequent simulation. }{\small \par}
-\item For regional simulations, the files recording the absorbing boundary
-contribution are also written to the \texttt{LOCAL\_PATH} when \texttt{SAVE\_FORWARD
-= .true.}.
-\end{itemize}
-\item \textbf{Prepare the adjoint sources}
-
-The adjoint sources need to be prepared the same way as described
-in the Section \ref{enu:Prepare-the-adjoint}.
-\begin{itemize}
-\item In the case of travel-time finite-frequency kernel for one source-receiver
-pair, i.e., point source from the \texttt{CMTSOLUTION}, and one station
-in the \texttt{STATIONS\_ADJOINT} list, we supply a sample program
-in \texttt{utils/adjoint\_sources/traveltime/xcreate\_adjsrc\_traveltime}
-to cut a certain portion of the original displacement seismograms
-and convert it into the proper adjoint source to compute the finite-frequency
-kernel.
-
-\begin{lyxcode}
-xcreate\_adjsrc\_traveltime~t1~t2~ifile{[}0-5{]}~E/N/Z-ascii-files~{[}baz{]}
-\end{lyxcode}
-
-where \texttt{t1} and \texttt{t2} are the start and end time of the
-portion you are interested in, \texttt{ifile} denotes the component
-of the seismograms to be used (0 for all three components, 1 for East,
-2 for North, and 3 for vertical, 4 for transverse, and 5 for radial
-component), \texttt{E/N/Z-ascii-files} indicate the three-component
-displacement seismograms in the right order, and \texttt{baz} is the
-back-azimuth of the station. Note that \texttt{baz} is only supplied
-when \texttt{ifile} = 4 or 5.
-
-\item Similarly, a sample program to compute adjoint sources for amplitude
-finite-frequency kernels may be found in \texttt{utils/adjoint\_sources/amplitude}
-and used in the same way as described for traveltime measurements
-
-\begin{lyxcode}
-xcreate\_adjsrc\_amplitude~t1~t2~ifile{[}0-5{]}~E/N/Z-ascii-files~{[}baz{]}.
-\end{lyxcode}
-\end{itemize}
-\item \textbf{Run the kernel simulation}
-
-With the successful forward simulation and the adjoint source ready
-in the \texttt{SEM/} directory, set \texttt{SIMULATION\_TYPE
-= 3} and \texttt{SAVE\_FORWARD = .false.} in the \texttt{Par\_file}
-(you can use \texttt{change\_simulation\_type.pl -b}), and rerun the
-solver.
-\begin{itemize}
-\item The adjoint simulation is launched together with the back reconstruction
-of the original forward wavefield from the state variables saved from
-the previous forward simulation, and the finite-frequency kernels
-are computed by the interaction of the reconstructed forward wavefield
-and the adjoint wavefield.
-\item The back-reconstructed seismograms at the original station locations
-are saved to the \texttt{LOCAL\_PATH} at the end of the kernel simulations,
-and can be collected to the local disk.
-\item These back-constructed seismograms can be compared with the time-reversed
-original seismograms to assess the accuracy of the backward reconstruction,
-and they should match very well.
-\item The arrays for density, P-wave speed and S-wave speed kernels are
-also saved in the \texttt{LOCAL\_PATH} with the names \texttt{proc??????\_rho(alpha,beta)\_kernel.bin},
-where \texttt{proc??????} represents the processor number, \texttt{rho(alpha,beta)}
-are the different types of kernels.
-\end{itemize}
-\item \textbf{Run the anisotropic kernel simulation}
-
-Instead of the kernels for the isotropic wave speeds, you can also
-compute the kernels for the 21 independent components $C_{IJ},\, I,J=1,...,6$
-(using Voigt's notation) of the elastic tensor in the cartesian coordinate
-system. This is done by setting \texttt{ANISOTROPIC\_KL} \texttt{=}
-\texttt{.true.} in \texttt{constants.h} before compiling the package.
-The definition of the parameters $C_{IJ}$ in terms of the corresponding
-components $c_{ijkl},ijkl,i,j,k,l=1,2,3$ of the elastic tensor in
-cartesian coordinates follows \citet{ChTr07}. The 21 anisotropic
-kernels are saved in the \texttt{LOCAL\_PATH} in one file with the
-name of \texttt{proc??????\_cijkl\_kernel.bin} (with \texttt{proc??????}
-the processor number). The output kernels correspond to absolute perturbations
-$\delta C_{IJ}$ of the elastic parameters and their unit is in $s/GPa/km^{3}$.
-For consistency, the output density kernels with this option turned
-on are for a perturbation $\delta\rho$ (and not $\frac{\delta\rho}{\rho}$)
-and their unit is in s / (kg/m$^{3}$) / km$^{3}$. These `primary'
-anisotropic kernels can then be combined to obtain the kernels for
-different parameterizations of anisotropy. This can be done, for example,
-when combining the kernel files from slices into one mesh file (see
-Section~\ref{sec:Finite-Frequency-Kernels}).
-
-If \texttt{ANISOTROPIC\_KL} \texttt{=} \texttt{.true.} by additionally
-setting \texttt{ANISOTROPIC\_KL} \texttt{=} \texttt{.true.} in \texttt{constants.h}
-the package will save anisotropic kernels paramterized as velocities
-related to transverse isotropy based on the the Chen and Tromp parameters
-\citet{ChTr07}. The kernels are saved as relative perturbations for
-horizontal and vertical P and S velocities, $\alpha_{v},\alpha_{h},\beta_{v},\beta_{h}$.
-Explicit relations can be found in appendix B. of \citet{SiLiTrTr07b}
-
-The anisotropic kernels are only currently available for CPU mode.
-
-\end{enumerate}
-In general, the three steps need to be run sequentially to assure
-proper access to the necessary files. If the simulations are run through
-some cluster scheduling system (e.g., LSF), and the forward simulation
-and the subsequent kernel simulations cannot be assigned to the same
-set of computer nodes, the kernel simulation will not be able to access
-the database files saved by the forward simulation. Solutions for
-this dilemma are provided in Chapter~\ref{cha:Scheduler}. Visualization
-of the finite-frequency kernels is discussed in Section~\ref{sec:Finite-Frequency-Kernels}.
-
-%------------------------------------------------------------------------------------------------%
-
-\chapter{\label{cha:tomo}Doing tomography, i.e., updating the model based on the sensitivity kernels obtained}
-
-%------------------------------------------------------------------------------------------------%
-
-UNDER CONSTRUCTION (March 2015). Will be added around April / May 2015.
-
-\section{Tomographic inversion / imaging using the sensitivity kernels obtained}
-
-OLD VERSION OF THE SECTION, WILL SOON BE IMPROVED AND ENRICHED.\\
-
-One of the fundamental reasons for computing sensitivity kernels (Section~\ref{sec:Adjoint-simulation-finite})
-is to use them within a tomographic inversion. In other words, use
-recorded seismograms, make measurements with synthetic seismograms,
-and use the misfit between the two to iteratively improve the model
-described by (at least) $V_{{\rm p}}$, $V_{{\rm s}}$, and $\rho$
-as a function of space.
-
-Whatever misfit function you use for the tomographic inversion (you can find several
-examples in \citet{TrKoLi08} and \citet{TrTaLi05}), you will weight
-the sensitivity kernels with measurements. Furthermore, you will use
-as many measurements (stations, components, time windows) as possible
-per event; hence, we call these composite kernels ``event kernels,''
-which are volumetric fields representing the gradient of the misfit
-function with respect to one of the variables (\eg $V_{{\rm s}}$).
-The basic features of an adjoint-based tomographic inversion were
-illustrated in \citet{TaLiTr07} and \citet{TrKoLi08} using a conjugate-gradient
-algorithm; there are dozens of versions of gradient-based inversion
-algorithms that could alternatively be used. The tomographic inversion
-of \citet{TaLiMaTr09,TaLiMaTr2010} used SPECFEM3D Cartesian as well
-as several additional components which are also stored on the CIG
-svn server, described next.
-
-The directory containing external utilities for tomographic inversion using
-SPECFEM3D Cartesian (or other packages that evaluate misfit functions
-and gradients) is in directory \texttt{utils/ADJOINT\_TOMOGRAPHY\_TOOLS/}:
-\begin{verbatim}
-flexwin/     -- FLEXWIN algorithm for automated picking of time windows
-measure_adj/ -- reads FLEXWIN output file and makes measurements,
-                with the option for computing adjoint sources
-iterate_adj/ -- various tools for iterative inversion
-                (requires pre-computed "event kernels")
-\end{verbatim}
-This directory also contains
-a brief \verb+README+ file indicating the role of the three subdirectories,
-\verb+flexwin+ \citep{Maggi2009}, \verb+measure_adj+, and \verb+iterate_adj+.
-The components for making the model update are there; however, there
-are no explicit rules for computing the model update, just as with
-any optimization problem. There are options for computing a conjugate
-gradient step, as well as a source subspace projection step.
-
-The best single file to read is probably: \verb+ADJOINT_TOMO/iterate_adj/cluster/README+.
-
-%------------------------------------------------------------------------------------------------%
-
-\chapter{Noise Cross-correlation Simulations}
-
-%------------------------------------------------------------------------------------------------%
-
-Besides earthquake simulations, SPECFEM3D Cartesian includes functionality
-for seismic noise tomography as well. In order to proceed successfully
-in this chapter, it is critical that you have already familiarized
-yourself with procedures for meshing (Chapter \ref{cha:Mesh-Generation}),
-creating distributed databases (Chapter \ref{cha:Creating-Distributed-Databases}),
-running earthquake simulations (Chapters \ref{cha:Running-the-Solver})
-and adjoint simulations (Chapter \ref{cha:Adjoint-Simulations}).
-Also, make sure you read the article `Noise cross-correlation sensitivity
-kernels' \citep{trompetal2010}, in order to understand noise simulations
-from a theoretical perspective.
-
-\section{Input Parameter Files}
-
-As usual, the three main input files are crucial: \texttt{Par\_file},
-\texttt{CMTSOLUTION} and \texttt{STATIONS}. Unless otherwise specified,
-those input files should be located in directory \texttt{DATA/}.\\
-
-\texttt{CMTSOLUTION} is required for all simulations. At a first glance,
-it may seem unexpected to have it here, since the noise simulations
-should have nothing to do with the earthquake -- \texttt{CMTSOLUTION}.
-However, for noise simulations, it is critical to have no earthquakes.
-In other words, the moment tensor specified in \texttt{CMTSOLUTION}
-must be set to zero manually!\\
-
-\texttt{STATIONS} remains the same as in previous earthquake simulations,
-except that the order of receivers listed in \texttt{STATIONS} is
-now important. The order will be used to determine the `master' receiver,
-i.e., the one that simultaneously cross correlates with the others.\\
-
-\texttt{Par\_file} also requires careful attention. A parameter called
-\texttt{NOISE\_TOMOGRAPHY} has been added which specifies the type
-of simulation to be run. \texttt{NOISE\_TOMOGRAPHY} is an integer
-with possible values 0, 1, 2 and 3. For example, when \texttt{NOISE\_TOMOGRAPHY}
-equals 0, a regular earthquake simulation will be run. When it is
-1/2/3, you are about to run step 1/2/3 of the noise simulations respectively.
-Should you be confused by the three steps, refer to \citet{trompetal2010}
-for details.\\
-
-Another change to \texttt{Par\_file} involves the parameter \texttt{NSTEP}.
-While for regular earthquake simulations this parameter specifies
-the length of synthetic seismograms generated, for noise simulations
-it specifies the length of the seismograms used to compute cross correlations.
-The actual cross correlations are thus twice this length, i.e., $2~\mathrm{NSTEP}-1$.
-The code automatically makes the modification accordingly, if \texttt{NOISE\_TOMOGRAPHY}
-is not zero.\\
-
-There are other parameters in \texttt{Par\_file} which should be given
-specific values. For instance, since the first two steps for calculating
-noise sensitivity kernels correspond to forward simulations, \texttt{SIMULATION\_TYPE}
-must be 1 when \texttt{NOISE\_TOMOGRAPHY} equals 1 or 2. Also, we
-have to reconstruct the ensemble forward wavefields in adjoint simulations,
-therefore we need to set \texttt{SAVE\_FORWARD} to \texttt{.true.}
-for the second step, i.e., when \texttt{NOISE\_TOMOGRAPHY} equals
-2. The third step is for kernel constructions. Hence \texttt{SIMULATION\_TYPE}
-should be 3, whereas \texttt{SAVE\_FORWARD} must be \texttt{.false.}.\\
-
-Finally, for most system architectures, please make sure that \texttt{LOCAL\_PATH}
-in \texttt{Par\_file} is in fact local, not globally shared. Because
-we have to save the wavefields at the earth's surface at every time
-step, it is quite problematic to have a globally shared \texttt{LOCAL\_PATH},
-in terms of both disk storage and I/O speed.
-
-\section{Noise Simulations: Step by Step}
-
-Proper parameters in those parameter files are not enough for noise
-simulations to run. We have more parameters to specify: for example,
-the ensemble-averaged noise spectrum, the noise distribution etc.
-However, since there are a few `new' files, it is better to introduce
-them sequentially. In this section, standard procedures for noise
-simulations are described.
-
-\subsection{Pre-simulation}
-\begin{itemize}
-\item As usual, we first configure the software package using:
-
-\texttt{./configure FC=ifort MPIFC=mpif90}
-
-Use the following if SCOTCH is needed:
-
-\texttt{./configure FC=ifort MPIFC=mpif90 -{}-with-scotch-dir=/opt/scotch}\\
-
-\item Next, we need to compile the source code using:
-
-\texttt{make xgenerate\_databases}
-
-\texttt{make xspecfem3D} \\
-
-\item Before we can run noise simulations, we have to specify the noise
-statistics, e.g., the ensemble-averaged noise spectrum. Matlab scripts
-are provided to help you to generate the necessary file.
-
-\texttt{EXAMPLES/noise\_tomography/NOISE\_TOMOGRAPHY.m} (main program)\\
- \texttt{EXAMPLES/noise\_tomography/PetersonNoiseModel.m}
-
-In Matlab, simply run:
-
-\texttt{NOISE\_TOMOGRAPHY(NSTEP, DT, Tmin, Tmax, NOISE\_MODEL)}\\
-
-\texttt{DT} is given in \texttt{Par\_file}, but \texttt{NSTEP} is
-NOT the one specified in \texttt{Par\_file}. Instead, you have to
-feed $2~\mathrm{NSTEP}-1$ to account for the doubled length of cross
-correlations. \texttt{Tmin} and \texttt{Tmax} correspond to the period
-range you are interested in, whereas \texttt{NOISE\_MODEL} denotes
-the noise model you will be using. Details can be found in the Matlab
-script.
-
-After running the Matlab script, you will be given the following information
-(plus a figure in Matlab):
-
-\texttt{{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}}\\
- \texttt{the source time function has been saved in:}\\
- \texttt{/data2/yangl/3D\_NOISE/S\_squared} (note this path must be
-different)\\
- \texttt{S\_squared should be put into directory:}\\
- \texttt{OUTPUT\_FILES/NOISE\_TOMOGRAPHY/ in the SPECFEM3D Cartesian
-package}\\
-
-In other words, the Matlab script creates a file called \texttt{S\_squared},
-which is the first `new' input file we encounter for noise simulations.
-
-One may choose a flat noise spectrum rather than Peterson's noise
-model. This can be done easily by modifying the Matlab script a little.
-
-\item Create a new directory in the SPECFEM3D Cartesian package, name it
-as \texttt{OUTPUT\_FILES/NOISE\_TOMOGRAPHY/}. We will add some parameter
-files later in this folder.
-\item Put the Matlab-generated-file \texttt{S\_squared} in \texttt{OUTPUT\_FILES/NOISE\_TOMOGRAPHY/}.
-
-That's to say, you will have a file \texttt{OUTPUT\_FILES/NOISE\_TOMOGRAPHY/S\_squared}
-in the SPECFEM3D Cartesian package.
-
-\item Create a file called \texttt{OUTPUT\_FILES/NOISE\_TOMOGRAPHY/irec\_master\_noise}.
-Note that this file is located in directory \texttt{OUTPUT\_FILES/NOISE\_TOMOGRAPHY/}
-as well. In general, all noise simulation related parameter files
-go into that directory. \texttt{irec\_master\_noise} contains only
-one integer, which is the ID of the `master' receiver. For example,
-if this file contains 5, it means that the fifth receiver listed in
-\texttt{DATA/STATIONS} becomes the `master'. That's why we mentioned
-previously that the order of receivers in \texttt{DATA/STATIONS} is
-important.
-
-Note that in regional simulations, the \texttt{DATA/STATIONS} might
-contain receivers which are outside of our computational domains.
-Therefore, the integer in \texttt{irec\_master\_noise} is actually
-the ID in \texttt{DATA/STATIONS\_FILTERED} (which is generated by
-\texttt{bin/xgenerate\_databases}).
-
-\item Create a file called \texttt{OUTPUT\_FILES/NOISE\_TOMOGRAPHY/nu\_master}.
-This file holds three numbers, forming a (unit) vector. It describes
-which component we are cross-correlating at the `master' receiver,
-i.e., ${\hat{{\bf \nu}}}^{\alpha}$ in \citet{trompetal2010}. The
-three numbers correspond to E/N/Z components respectively. Most often,
-the vertical component is used, and in those cases the three numbers
-should be 0, 0 and 1.
-\item Describe the noise direction and distributions in \texttt{src/shared/noise\_tomography.f90}.
-Search for a subroutine called \texttt{noise\_distribution\_direction}
-in \texttt{noise\_tomography.f90}. It is actually located at the very
-beginning of \texttt{noise\_tomography.f90}. The default assumes vertical
-noise and a uniform distribution across the whole free surface of
-the model. It should be quite self-explanatory for modifications.
-Should you modify this part, you have to re-compile the source code.
-\end{itemize}
-
-\subsection{Simulations}
-
-With all of the above done, we can finally launch our simulations.
-Again, please make sure that the \texttt{LOCAL\_PATH} in \texttt{DATA/Par\_file}
-is not globally shared. It is quite problematic to have a globally
-shared \texttt{LOCAL\_PATH}, in terms of both disk storage and speed
-of I/O (we have to save the wavefields at the earth's surface at every
-time step).
-
-As discussed in \citet{trompetal2010}, it takes three steps/simulations
-to obtain one contribution of the ensemble sensitivity kernels:
-\begin{itemize}
-\item Step 1: simulation for generating wavefields\\
- \texttt{SIMULATION\_TYPE=1}\\
- \texttt{NOISE\_TOMOGRAPHY=1}\\
- \texttt{SAVE\_FORWARD} not used, can be either .true. or .false.
-\item Step 2: simulation for ensemble forward wavefields\\
- \texttt{SIMULATION\_TYPE=1}\\
- \texttt{NOISE\_TOMOGRAPHY=2}\\
- \texttt{SAVE\_FORWARD=.true.}
-\item Step 3: simulation for ensemble adjoint wavefields and sensitivity
-kernels\\
- \texttt{SIMULATION\_TYPE=3}\\
- \texttt{NOISE\_TOMOGRAPHY=3}\\
- \texttt{SAVE\_FORWARD=.false.}
-
-Note Step 3 is an adjoint simulation, please refer to previous chapters
-on how to prepare adjoint sources and other necessary files, as well
-as how adjoint simulations work.
-
-\end{itemize}
-Note that it is better to run the three steps continuously within
-the same job, otherwise you have to collect the saved surface movies
-from the old nodes/CPUs to the new nodes/CPUs. This process varies
-from cluster to cluster and thus cannot be discussed here. Please
-ask your cluster administrator for information/configuration of the
-cluster you are using.\\
-
-\subsection{Post-simulation}
-
-After those simulations, you have all stuff you need, either in the
-\texttt{OUTPUT\_FILES/} or in the directory specified by \texttt{LOCAL\_PATH}
-in \texttt{DATA/Par\_file} (which are most probably on local nodes).
-Collect whatever you want from the local nodes to your workstation,
-and then visualize them. This process is the same as what you may
-have done for regular earthquake simulations. Refer to other chapters
-if you have problems.\\
-
-Simply speaking, two outputs are the most interesting: the simulated
-ensemble cross correlations and one contribution of the ensemble sensitivity
-kernels.\\
-
-The simulated ensemble cross correlations are obtained after the second
-simulation (Step 2). Seismograms in \texttt{OUTPUT\_FILES/} are actually
-the simulated ensemble cross correlations. Collect them immediately
-after Step 2, or the Step 3 will overwrite them. Note that we have
-a `master' receiver specified by \texttt{irec\_master\_noise}, the
-seismogram at one station corresponds to the cross correlation between
-that station and the `master'. Since the seismograms have three components,
-we may obtain cross correlations for different components as well,
-not necessarily the cross correlations between vertical components.\\
-
-One contribution of the ensemble cross-correlation sensitivity kernels
-are obtained after Step 3, stored in the \texttt{DATA/LOCAL\_PATH}
-on local nodes. The ensemble kernel files are named the same as regular
-earthquake kernels.\\
-
-You need to run another three simulations to get the other contribution
-of the ensemble kernels, using different forward and adjoint sources
-given in \citet{trompetal2010}.
-
-\section{Example}
-
-In order to illustrate noise simulations in an easy way, one example
-is provided in \texttt{EXAMPLES/noise\_tomography/}. See \texttt{EXAMPLES/noise\_tomography/README}
-for explanations. \\
-
-Note, however, that they are created for a specific workstation (CLOVER@PRINCETON),
-which has at least 4 cores with `mpif90' working properly. \\
-
-If your workstation is suitable, you can run the example in \texttt{EXAMPLES/noise\_tomography/}
-using:\\
-
-\texttt{./pre-processing.sh}\\
-
-Even if this script does not work on your workstation, the procedure
-it describes is universal. You may review the whole process described
-in the last section by following the commands in \texttt{pre-processing.sh},
-which should contain enough explanations for all the commands.
-
-%------------------------------------------------------------------------------------------------%
-
-\chapter{Gravity integral calculations for the gravity field of an Earth model}
-
-%------------------------------------------------------------------------------------------------%
-
-SPECFEM can now compute the gravity field as well as its derivatives (i.e., gravity gradiometry)
-generated by any given 3D Earth model at a given elevation, i.e., on a given observation surface.
-That feature is still experimental but should work just fine.\\
-
-To see how it is implemented and to use it, type this in the root directory of the code:\\
-
-\texttt{grep -i GRAVITY\_INTEGRALS src/*/*}
-
-%------------------------------------------------------------------------------------------------%
-
-\chapter{Graphics}
-
-%------------------------------------------------------------------------------------------------%
-
-\section{\label{sec:Mesh-graphics}Meshes}
-
-In case you used the internal mesher \texttt{xmeshfem3D} to create
-and partition your mesh, you can output mesh files in ABAQUS (.INP)
-and DX (.dx) format to visualize them. For this, you must set either
-the flag \texttt{CREATE\_DX\_FILES} or \texttt{CREATE\_ABAQUS\_FILES}
-to \texttt{.true.} in the mesher's parameter file \texttt{Mesh\_Par\_file}
-prior to running the mesher (see Chapter~\ref{cha:Running-the-Mesher-Meshfem3D}
-for details). You can then use AVS \urlwithparentheses{www.avs.com}
-or OpenDX \urlwithparentheses{www.opendx.org} to visualize the mesh
-and MPI partition (slices).
-
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=0.49\textwidth]{figures/vtk_mesh_vp} \includegraphics[width=0.49\textwidth]{figures/vtk_mesh_vs}
-\par\end{centering}
-
-\caption{Visualization using Paraview of VTK files created by \texttt{xgenerate\_databases}
-showing P- and S-wave velocities assigned to the mesh points. The
-mesh was created by \texttt{xmeshfem3D} for 4 processors.}
-
-\label{fig:vtk.mesh}
-\end{figure}
-
-You have also the option to visualize the distributed databases produced
-by \texttt{xgenerate\_databases} using Paraview \urlwithparentheses{www.paraview.org}.
-For this, you must set the flag \texttt{SAVE\_MESH\_FILES} to \texttt{.true.}
-in the main parameter file \texttt{Par\_file} (see Chapter~\ref{cha:Main-Parameter}
-for details). This will create VTK files for each single partition.
-You can then use Paraview \urlwithparentheses{www.paraview.org}
-to visualized these partitions.
-
-\section{\label{sec:Movies}Movies}
-
-To make a surface or volume movie of the simulation, set parameters
-\texttt{MOVIE\_SURFACE}, \texttt{MOVIE\_VOLUME}, \texttt{MOVIE\_TYPE},
-and \texttt{NTSTEP\_BETWEEN\_FRAMES} in the \texttt{Par\_file}. Turning
-on the movie flags, in particular \texttt{MOVIE\_VOLUME}, produces
-large output files. \texttt{MOVIE\_VOLUME} files are saved in the
-\texttt{LOCAL\_PATH} directory, whereas \texttt{MOVIE\_SURFACE} output
-files are saved in the \texttt{OUTPUT\_FILES} directory. We save the
-displacement field if the parameter \texttt{SAVE\_DISPLACEMENT} is
-set, otherwise the velocity field is saved. The look of a movie is
-determined by the half-duration of the source. The half-duration should
-be large enough so that the movie does not contain frequencies that
-are not resolved by the mesh, i.e., it should not contain numerical
-noise. This can be accomplished by selecting a CMT \texttt{HALF\_DURATION}
-> 1.1 $\times$ smallest period (see figure \ref{fig:CMTSOLUTION-file}).
-When \texttt{MOVIE\_SURFACE} = .\texttt{true.}, the half duration
-of each source in the \texttt{CMTSOLUTION} file is replaced by
-\begin{quote}
-\[
-\sqrt{(}\mathrm{\mathtt{HALF\_DURATIO}\mathtt{N}^{2}}+\mathrm{\mathtt{HDUR\_MOVI}\mathtt{E}^{2}})
-\]
- \textbf{NOTE:} If \texttt{HDUR\_MOVIE} is set to 0.0, the code will
-select the appropriate value of 1.1 $\times$ smallest period. As
-usual, for a point source one can set \texttt{half duration} in the
-\texttt{CMTSOLUTION} file to be 0.0 and \texttt{HDUR\_MOVIE} = 0.0
-in the \texttt{Par\_file} to get the highest frequencies resolved
-by the simulation, but for a finite source one would keep all the
-\texttt{half durations} as prescribed by the finite source model and
-set \texttt{HDUR\_MOVIE} = 0.0.
-\end{quote}
-
-\subsection{Movie Surface}
-
-When running \texttt{xspecfem3D} with the \texttt{MOVIE\_SURFACE}
-flag turned on, the code outputs \texttt{moviedata??????} files in
-the \texttt{OUTPUT\_FILES} directory. There are several flags in the
-main parameter file \texttt{Par\_file} that control the output of
-these movie data files (see section \ref{cha:Main-Parameter} for
-details): \texttt{\small NTSTEP\_BETWEEN\_FRAMES}{\small{} to set the
-timesteps between frames, }\texttt{\small SAVE\_DISPLACEMENT}{\small{}
-to save displacement instead of velocity, }\texttt{\small USE\_HIGHRES\_FOR\_MOVIES}{\small{}
-to save values at all GLL point instead of element edges. In order
-to output additionally shakemaps, you would set the parameter }\texttt{\small CREATE\_SHAKEMAP}{\small{}
-to }\texttt{\small .true.}{\small .}{\small \par}
-
-The files are in a fairly complicated binary format, but there is
-a program provided to convert the output into more user friendly formats:
-\begin{description}
-\item [{\texttt{xcreate\_movie\_shakemap\_AVS\_DX\_GMT}}] From \texttt{create\_movie\_shakemap\_AVS\_DX\_GMT.f90},
-it outputs data in ASCII, OpenDX, or AVS format (also readable in
-ParaView). Before compiling the code, make sure you have the file
-\texttt{surface\_from\_mesher.h} in the \texttt{OUTPUT\_FILES/} directory.
-This file will be created by the solver run. Then type
-
-\begin{lyxcode}
-{\footnotesize make~xcreate\_movie\_shakemap\_AVS\_DX\_GMT}{\footnotesize \par}
-\end{lyxcode}
-
-and run the executable \texttt{xcreate\_movie\_shakemap\_AVS\_DX\_GMT}
-in the main directory. It will create visualization files
-in your format of choice. The code will prompt the user for input
-parameters.
-
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=0.32\textwidth]{figures/movie_surf_1} \includegraphics[width=0.32\textwidth]{figures/movie_surf_2}
-\includegraphics[width=0.32\textwidth]{figures/movie_surf_3}
-\par\end{centering}
-
-\caption{Visualization using AVS files created by \texttt{xcreate\_movie\_shakemap\_AVS\_DX\_GMT}
-showing movie snapshots of vertical velocity components at different
-times.}
-
-\label{fig:movie.surf}
-\end{figure}
-
-\end{description}
-The \texttt{SPECFEM3D Cartesian} code is running in near real-time
-to produce animations of southern California earthquakes via the web;
-see Southern California ShakeMovie\textregistered{}\urlwithparentheses{www.shakemovie.caltech.edu}.
-
-\subsection{\label{sub:Movie-Volume}Movie Volume}
-
-When running xspecfem3D with the \texttt{\small MOVIE\_VOLUME}{\small{}
-flag turned on, the code outputs several files in }\texttt{\small LOCAL\_PATH}{\small{}
-specified in the main }\texttt{\small Par\_file}{\small , e.g. in
-directory }\texttt{\small OUTPUT\_FILES/DATABASES\_MPI}{\small .
-The output is saved by each processor at the time interval specified
-by }\texttt{\small NTSTEP\_BETWEEN\_FRAMES}{\small . For all domains,
-the velocity field is output to files: }\texttt{\small proc??????\_velocity\_X\_it??????.bin}{\small ,
-}\texttt{\small proc??????\_velocity\_Y\_it??????.bin}{\small{} and
-}\texttt{\small proc??????\_velocity\_Z\_it??????.bin}{\small . For
-elastic domains, the divergence and curl taken from the velocity field,
-i.e. $\nabla\cdot{\bf {v}}$ and $\nabla\times{\bf {v}}$, get stored
-as well: }\texttt{\small proc??????\_div\_it??????.bin}{\small , }\texttt{\small proc??????\_curl\_X\_t??????.bin}{\small ,
-}\texttt{\small proc??????\_curl\_Y\_it??????.bin}{\small{} and }\texttt{\small proc??????\_curl\_Z\_it??????.bin}{\small .
-The files denoted }\texttt{\small proc??????\_div\_}{\small ~}\\
-{\small{} }\texttt{\small glob\_it??????.bin}{\small{} and }\texttt{\small proc??????\_curl\_glob\_it??????.bin}{\small{}
-are stored on the global points only, all the other arrays are stored
-on all GLL points. Note that the components X/Y/Z can change to E/N/Z
-according to the }\texttt{\small SUPPRESS\_UTM\_PROJECTION}{\small{}
-flag (see also Appendix~\ref{cha:Coordinates} and \ref{cha:channel-codes}).}{\small \par}
-
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=0.32\textwidth]{figures/movie_volume_1} \includegraphics[width=0.32\textwidth]{figures/movie_volume_2}
-\includegraphics[width=0.32\textwidth]{figures/movie_volume_3}
-\par\end{centering}
-
-\caption{Paraview visualization using movie volume files (converted by \texttt{xcombine\_vol\_data}
-and \texttt{mesh2vtu.pl}) and showing snapshots of vertical velocity
-components at different times.}
-
-\label{fig:movie.volume}
-\end{figure}
-
-To visualize these files, we use an auxilliary program \texttt{combine\_vol\_data.f90}
-to combine the data from all slices into one mesh file. To compile
-it in the root directory, type:
-\begin{lyxcode}
-{\footnotesize make~xcombine\_vol\_data~}{\footnotesize \par}
-\end{lyxcode}
-which will create the executable \texttt{bin/xcombine\_vol\_data}.
-To output the usage of this executable, type
-\begin{lyxcode}
-{\footnotesize ./bin/xcombine\_vol\_data~}{\footnotesize \par}
-\end{lyxcode}
-without arguments.
-
-\medskip
-
-\texttt{xcombine\_vol\_data} will combine the data on the different processors (located in \texttt{OUTPUT\_FILES/DATABASES\_MPI}), for a given quantity and a given iteration, to one file.
-For example, if you want to combine \texttt{velocity\_Z}, for the iteration 400 on 4 processors, i.e.
-if you want to combine these files :
-
-\medskip
-
-\texttt{proc000000\_velocity\_Z\_it000400.bin }
-
-\texttt{proc000001\_velocity\_Z\_it000400.bin }
-
-\texttt{proc000002\_velocity\_Z\_it000400.bin }
-
-\texttt{proc000003\_velocity\_Z\_it000400.bin}
-
-\medskip
-
-You have to go in the directory of the concerned example (where are the directories \texttt{DATA}, \texttt{OUTPUT\_FILES}, etc.).
-Then, you can launch \texttt{xcombine\_vol\_data}, specifying the path where it is located.
-As an example, if the \texttt{DATA} and \texttt{OUTPUT\_FILES} directories of the concerned example are in the root specfem3d directory, \texttt{xcombine\_vol\_data} is in \texttt{./bin}, and you have to type:
-
-\begin{lyxcode}
-~{\footnotesize ./bin/xcombine\_vol\_data~0~3~velocity\_Z\_it000400~./OUTPUT\_FILES/DATABASES\_MPI~./OUTPUT\_FILES~0~}{\footnotesize \par}
-\end{lyxcode}
-
-Here, $0$ is the number of the first processor, $3$ the number of the last one, \texttt{velocity\_Z\_it000400} the name of the files we want to combine without the prefix \texttt{proc000000*\_}, \texttt{./OUTPUT\_FILES/DATABASES\_MPI} the directory where the files are located, \texttt{./OUTPUT\_FILES} the directory where the combined file will be stored, and $0$ is the parameter to create a low-resolution mesh file ($1$ for a high-resolution).
-
-\medskip
-
-At the beginning of the source file \texttt{combine\_vol\_data.f90} (located in \texttt{src/auxiliaries/}), there is a logical constant parameter \texttt{USE\_VTK\_OUTPUT} whose default value is \texttt{.true.}, that which will provide the output combined file directly in the VTK format and you can use Paraview to visualize it.
-
-\medskip
-
-If the value of this parameter is \texttt{.false.}, the output mesh file will have the name \texttt{velocity\_Z\_it000400.mesh}. Then we next have to convert the
-\texttt{.mesh} file into the VTU (Unstructured grid file) format which can be viewed in ParaView. For this task, you can use and modify the
-script \texttt{mesh2vtu.pl} located in directory \texttt{utils/Visualization/Paraview/}, for example:
-\begin{lyxcode}
-{\footnotesize mesh2vtu.pl~-i~velocity\_Z\_it000400.mesh~-o~velocity\_Z\_it000400.vtu}{\footnotesize \par}
-\end{lyxcode}
-Notice that this Perl script uses a program \texttt{mesh2vtu} in the \texttt{utils/Visualization/Paraview/mesh2vtu} directory, which further
-uses the VTK \urlwithparentheses{http://www.vtk.org/} run-time library for its execution. Therefore, make sure you have them properly set
-in the script according to your system.
-
-\bigskip
-
-Then, to do a movie with several iterations, you have to repeat this process for each iteration you want to put in your movie.
-
-\section{\label{sec:Finite-Frequency-Kernels}Finite-Frequency Kernels}
-
-The finite-frequency kernels computed as explained in Section \ref{sec:Adjoint-simulation-finite}
-are saved in the \texttt{LOCAL\_PATH} at the end of the simulation.
-Therefore, we first need to collect these files on the front end,
-combine them into one mesh file, and visualize them with some auxilliary
-programs.
-\begin{enumerate}
-\item \textbf{Create slice files}
-
-We will only discuss the case of one source-receiver pair, i.e., the
-so-called banana-doughnut kernels. Although it is possible to collect
-the kernel files from all slices on the front end, it usually takes
-up too much storage space (at least tens of gigabytes). Since the
-sensitivity kernels are the strongest along the source-receiver great
-circle path, it is sufficient to collect only the slices that are
-along or close to the great circle path.
-
-A Perl script \texttt{slice\_number.pl} located in directory \texttt{utils/Visualization/Paraview/}
-can help to figure out the slice numbers that lie along the great
-circle path. It applies to meshes created with the internal mesher
-\texttt{xmeshfem3D}.
-\begin{enumerate}
-\item On machines where you have access to the script, copy the \texttt{Mesh\_Par\_file},
-and \texttt{output\_solver} files, and run:
-
-\begin{lyxcode}
-{\small slice\_number.pl~Mesh\_Par\_file~output\_solver.txt~slice\_file}{\small \par}
-\end{lyxcode}
-
-which will generate a \texttt{slices\_file}.
-
-\item For cases with multiple sources and multiple receivers, you need to
-provide a slice file before proceeding to the next step.
-\end{enumerate}
-\item \textbf{Collect the kernel files}
-
-After obtaining the slice files, you can collect the corresponding
-kernel files from the given slices.
-\begin{enumerate}
-\item You can use or modify the script \texttt{utils/copy\_basin\_database.pl}
-to accomplish this:
-
-\begin{lyxcode}
-{\small utils/copy\_database.pl~slice\_file~lsf\_machine\_file~filename~{[}jobid{]}}{\small \par}
-\end{lyxcode}
-
-where \texttt{\small lsf\_machine\_file}{\small{} is the machine file
-generated by the LSF scheduler, }\texttt{\small filename}{\small{}
-is the kernel name (e.g., }\texttt{\small rho\_kernel}{\small , }\texttt{\small alpha\_kernel}{\small{}
-and }\texttt{\small beta\_kernel}{\small ), and the optional }\texttt{\small jobid}{\small{}
-is the name of the subdirectory under }\texttt{\small LOCAL\_PATH}{\small{}
-where all the kernel files are stored.}{\small \par}
-
-\item After executing this script, all the necessary mesh topology files
-as well as the kernel array files are collected to the local directory
-of the front end.
-\end{enumerate}
-\item \textbf{Combine kernel files into one mesh file}
-
-We use an auxilliary program \texttt{combine\_vol\_data.f90} to combine
-the kernel files from all slices into one mesh file.
-\begin{enumerate}
-\item Compile it in the root directory:
-
-\begin{lyxcode}
-{\footnotesize make~xcombine\_vol\_data~}{\footnotesize \par}
-
-{\footnotesize ./bin/xcombine\_vol\_data~slice\_list~filename~input\_dir~output\_dir~high/low-resolution~}{\footnotesize \par}
-\end{lyxcode}
-
-where \texttt{input\_dir} is the directory where all the individual
-kernel files are stored, and \texttt{output\_dir} is where the mesh
-file will be written.
-
-\item Use 1 for a high-resolution mesh, outputting all the GLL points to
-the mesh file, or use 0 for low resolution, outputting only the corner
-points of the elements to the mesh file.
-\item The output mesh file will have the name \texttt{filename\_rho(alpha,beta).mesh}
-\end{enumerate}
-\item \textbf{Convert mesh files into .vtu files}
-
-\begin{enumerate}
-\item We next convert the \texttt{.mesh} file into the VTU (Unstructured
-grid file) format which can be viewed in ParaView. For this task,
-you can use and modify the script \texttt{mesh2vtu.pl} located in
-directory \texttt{utils/Visualization/Paraview/}, for example:
-
-\begin{lyxcode}
-{\footnotesize mesh2vtu.pl~-i~file.mesh~-o~file.vtu~}{\footnotesize \par}
-\end{lyxcode}
-\item Notice that this Perl script uses a program \texttt{mesh2vtu} in the
-\texttt{utils/Visualization/Paraview/mesh2vtu} directory, which further
-uses the VTK \urlwithparentheses{http://www.vtk.org/} run-time library
-for its execution. Therefore, make sure you have them properly set
-in the script according to your system.
-\end{enumerate}
-\item \textbf{Copy over the source and receiver .vtk file}
-
-In the case of a single source and a single receiver, the simulation
-also generates the file \texttt{sr.vtk} located in the \texttt{OUTPUT\_FILES/}
-directory to describe the source and receiver locations, which can
-also be viewed in Paraview in the next step.
-
-\item \textbf{View the mesh in ParaView}
-
-Finally, we can view the mesh in ParaView \urlwithparentheses{www.paraview.org}.
-\begin{enumerate}
-\item Open ParaView.
-\item From the top menu, \textsf{File} $\rightarrow$\textsf{Open data},
-select \texttt{file.vtu}, and click the \textsf{Accept} button.
-
-\begin{itemize}
-\item If the mesh file is of moderate size, it shows up on the screen; otherwise,
-only the bounding box is shown.
-\end{itemize}
-\item Click \textsf{Display Tab} $\rightarrow$ \textsf{Display Style} $\rightarrow$
-\textsf{Representation} and select \textsf{wireframe of surface} to
-display it.
-\item To create a cross-section of the volumetric mesh, choose \textsf{Filter}
-$\rightarrow$ \textsf{cut}, and under \textsf{Parameters Tab}, choose
-\textsf{Cut Function} $\rightarrow$ \textsf{plane}.
-\item Fill in center and normal information given by the \texttt{global\_slice\_number.pl}
-script (either from the standard output or from \texttt{normal\_plane.txt}
-file).
-\item To change the color scale, go to \textsf{Display Tab} $\rightarrow$
-\textsf{Color} $\rightarrow$ \textsf{Edit Color Map} and reselect
-lower and upper limits, or change the color scheme.
-\item Now load in the source and receiver location file by \textsf{File}
-$\rightarrow$ \textsf{Open data}, select \texttt{sr.vt}k, and click
-the \textsf{Accept} button. Choose \textsf{Filter} $\rightarrow$
-\textsf{Glyph}, and represent the points by `\textsf{spheres}'.
-\item For more information about ParaView, see the ParaView Users Guide
-\urlwithparentheses{www.paraview.org/files/v1.6/ParaViewUsersGuide.PDF}.
-\end{enumerate}
-\end{enumerate}
-\begin{figure}[H]
-\noindent \begin{centering}
-\includegraphics[scale=0.6]{figures/3D-S-Kernel}
-\par\end{centering}
-
-\caption{(a) Top Panel: Vertical source-receiver cross-section of the S-wave
-finite-frequency sensitivity kernel $K_{\beta}$ for station GSC at
-an epicentral distance of 176 km from the September 3, 2002, Yorba
-Linda earthquake. Lower Panel: Vertical source-receiver cross-section
-of the 3D S-wave speed model used for the spectral-element simulations
-\citep{KoLiTrSuStSh04}. (b) The same as (a) but for station HEC at
-an epicentral distance of 165 km \citep{LiTr06}.}
-
-\label{figure:P-wave-speed-finite-frequency}
-\end{figure}
-
-\chapter{\label{cha:Scheduler}Running through a Scheduler}
-
-The code is usually run on large parallel machines, often PC clusters,
-most of which use schedulers, i.e., queuing or batch management systems
-to manage the running of jobs from a large number of users. The following
-considerations need to be taken into account when running on a system
-that uses a scheduler:
-\begin{itemize}
-\item The processors/nodes to be used for each run are assigned dynamically
-by the scheduler, based on availability. Therefore, in order for the
-\texttt{xgenerate\_databases} and the \texttt{xspecfem3D} executables
-(or between successive runs of the solver) to have access to the same
-database files (if they are stored on hard drives local to the nodes
-on which the code is run), they must be launched in sequence as a
-single job.
-\item On some systems, the nodes to which running jobs are assigned are
-not configured for compilation. It may therefore be necessary to pre-compile
-both the \texttt{xgenerate\_databases} and the \texttt{xspecfem3D}
-executables.
-\item One feature of schedulers/queuing systems is that they allow submission
-of multiple jobs in a ``launch and forget'' mode. In order to take
-advantage of this property, care needs to be taken that output and
-intermediate files from separate jobs do not overwrite each other,
-or otherwise interfere with other running jobs.
-\end{itemize}
-Examples of job scripts can be found in the \texttt{\small utils/Cluster/}{\small{}
-directory and can straightforwardly be modified and adapted to meet
-more specific running needs.}{\small \par}
-
-\chapter{{\normalsize \label{cha:-Changing-the}} Changing the Model}
+\chapter{Changing the Model}\label{cha:-Changing-the}
 
@@
 
-\section{{\normalsize \label{sec:Using-tomographic}}Using external tomographic Earth models}
+\section{Using external tomographic Earth models}\label{sec:Using-tomographic}
 
@@
 \noindent \begin{centering}
-\includegraphics[width=1\textwidth]{figures/tomo_file}
+\includegraphics[width=1\textwidth]{figures/tomo_file.jpg}
 \par\end{centering}
@@
 
-\section{{\normalsize \label{sec:Anelastic-Models}}External (an)elastic Models}
+\section{External (an)elastic Models}\label{sec:Anelastic-Models}
 
@@
 
-\section{{\normalsize \label{sec:Anisotropic-Models}}Anisotropic Models}
+\section{Anisotropic Models}\label{sec:Anisotropic-Models}
 
@@
 
-\section{{\normalsize \label{sec:Using-SEP}}Using external SEP models.}
+\section{Using external SEP models}\label{sec:Using-SEP}
 
@@
 \texttt{EXAMPLES/meshfem3D\_examples/sep\_bathymetry}.
-
-\chapter{Post-Processing Scripts}
-
-Several post-processing scripts/programs are provided in the \texttt{utils/}
-directory, and most of them need to be adjusted when used on different
-systems, for example, the path of the executable programs. Here we
-only list a few of the available scripts and provide a brief description,
-and you can either refer to the related sections for detailed usage
-or, in a lot of cases, type the script/program name without arguments
-for its usage.
-
-\section{Process Data and Synthetics\label{sec:Process-data-and-syn}}
-
-In many cases, the SEM synthetics are calculated and compared to data
-seismograms recorded at seismic stations. Since the SEM synthetics
-are accurate for a certain frequency range, both the original data
-and the synthetics need to be processed before a comparison can be
-made.
-
-For such comparisons, the following steps are recommended:
-\begin{enumerate}
-\item Make sure that both synthetic and observed seismograms have the correct
-station/event and timing information.
-\item Convolve synthetic seismograms with a source time function with the
-half duration specified in the \texttt{CMTSOLUTION} file, provided,
-as recommended, you used a zero half duration in the SEM simulations.
-\item Resample both observed and synthetic seismograms to a common sampling
-rate.
-\item Cut the records using the same window.
-\item Remove the trend and mean from the records and taper them.
-\item Remove the instrument response from the observed seismograms (recommended)
-or convolve the synthetic seismograms with the instrument response.
-\item Make sure that you apply the same filters to both observed and synthetic
-seismograms. Preferably, avoid filtering your records more than once.
-\item Now, you are ready to compare your synthetic and observed seismograms.
-\end{enumerate}
-\noindent We generally use the following scripts provided in the \texttt{utils/seis\_process/}
-directory:
-
-\subsection{Data processing script \texttt{process\_data.pl}}
-
-This script cuts a given portion of the original data, filters it,
-transfers the data into a displacement record, and picks the first
-P and S arrivals. For more functionality, type `\texttt{process\_data.pl}'
-without any argument. An example of the usage of the script:
-\begin{lyxcode}
-{\footnotesize process\_data.pl~-m~CMTSOLUTION~-s~1.0~-l~0/4000~-i~-f~-t~40/500~-p~-x~bp~DATA/1999.330{*}.BH?.SAC}~
-\end{lyxcode}
-\noindent which has resampled the SAC files to a sampling rate of
-1 seconds, cut them between 0 and 4000 seconds, transfered them into
-displacement records and filtered them between 40 and 500 seconds,
-picked the first P and S arrivals, and added suffix `\texttt{bp}'
-to the file names.
-
-Note that all of the scripts in this section actually use the SAC
-and/or IASP91 to do the core operations; therefore make sure that
-the SAC and IASP91 packages are installed properly on your system,
-and that all the environment variables are set properly before running
-these scripts.
-
-\subsection{Synthetics processing script \texttt{process\_syn.pl}}
-
-This script converts the synthetic output from the SEM code from ASCII
-to SAC format, and performs similar operations as `\texttt{process\_data.pl}'.
-An example of the usage of the script:
-\begin{lyxcode}
-{\footnotesize process\_syn.pl~-m~CMTSOLUTION~-h~-a~STATIONS~-s~1.0~-l~0/4000~-f~-t~40/500~-p~-x~bp~SEM/{*}.BX?.semd}~
-\end{lyxcode}
-which will convolve the synthetics with a triangular source-time function
-from the \texttt{CMTSOLUTION} file, convert the synthetics into SAC
-format, add event and station information into the SAC headers, resample
-the SAC files with a sampling rate of 1 seconds, cut them between
-0 and 4000 seconds, filter them between 40 and 500 seconds with the
-same filter used for the observed data, pick the first P and S arrivals,
-and add the suffix `\texttt{bp}' to the file names.
-
-More options are available for this script, such as adding time shift
-to the origin time of the synthetics, convolving the synthetics with
-a triangular source time function with a given half duration, etc.
-Type \texttt{process\_syn.pl} without any argument for a detailed
-usage.
-
-In order to convert between SAC format and ASCII files, useful scripts
-are provided in the subdirectories ~\\
- \texttt{utils/sac2000\_alpha\_convert/} and \texttt{utils/seis\_process/asc2sac/}.
-
-\subsection{Script \texttt{rotate.pl}}
-
-The original data and synthetics have three components: vertical (BHZ
-resp. BXZ), north (BHN resp. BXN) and east (BHE resp. BXE). However,
-for most seismology applications, transverse and radial components
-are also desirable. Therefore, we need to rotate the horizontal components
-of both the data and the synthetics to the transverse and radial direction,
-and \texttt{\small rotate.pl}{\small{} can be used to accomplish this:}{\small \par}
-\begin{lyxcode}
-rotate.pl~-l~0~-L~180~-d~DATA/{*}.BHE.SAC.bp~
-
-rotate.pl~-l~0~-L~180~SEM/{*}.BXE.semd.sac.bp~
-\end{lyxcode}
-where the first command performs rotation on the SAC data obtained
-through Seismogram Transfer Program (STP) \urlwithparentheses{http://www.data.scec.org/STP/stp.html},
-while the second command rotates the processed SEM synthetics.
-
-\section{Collect Synthetic Seismograms}
-
-The forward and adjoint simulations generate synthetic seismograms
-in the \texttt{OUTPUT\_FILES/} directory by default. For the forward
-simulation, the files are named like \texttt{NT.STA.BX?.semd} for
-two-column time series, or \texttt{NT.STA.BX?.semd.sac} for ASCII
-SAC format, where NT and STA are the network code and station name,
-and \texttt{BX?} stands for the component name. Please see the Appendix~\ref{cha:Coordinates}
-and \ref{cha:channel-codes} for further details.
-
-The adjont simulations generate synthetic seismograms with the name
-\texttt{NT.S?????.S??.sem} (refer to Section \ref{sec:Adjoint-simulation-sources}
-for details). The kernel simulations output the back-reconstructed
-synthetic seismogram in the name \texttt{NT.STA.BX?.semd}, mainly
-for the purpose of checking the accuracy of the reconstruction. Refer
-to Section \ref{sec:Adjoint-simulation-finite} for further details.
-
-You do have further options to change this default output behavior,
-given in the main constants file \texttt{constants.h} located in \texttt{src/shared/}
-directory:
-\begin{description}
-\item [{\texttt{SEISMOGRAMS\_BINARY}}] set to \texttt{.true.} to have seismograms
-written out in binary format.
-\item [{\texttt{WRITE\_SEISMOGRAMS\_BY\_MASTER}}] Set to \texttt{.true.}
-to have only the master process writing out seismograms. This can
-be useful on a cluster, where only the master process node has access
-to the output directory.
-\item [{\texttt{USE\_OUTPUT\_FILES\_PATH}}] Set to \texttt{.false.} to
-have the seismograms output to \texttt{LOCAL\_PATH} directory specified
-in the main parameter file \texttt{DATA/Par\_file}. In this case,
-you could collect the synthetics onto the frontend using the \texttt{collect\_seismo\_lsf\_multi.pl}
-script located in the \texttt{utils/Cluster/lsf/} directory. The usage
-of the script would be e.g.:
-
-\begin{lyxcode}
-collect\_seismo.pl~machines~DATA/Par\_file
-\end{lyxcode}
-\end{description}
-where \texttt{machines} is a file containing the node names and \texttt{DATA/Par\_file}
-the parameter file used to extract the \texttt{LOCAL\_PATH} directory
-used for the simulation.
-
-\section{Clean Local Database}
-
-After all the simulations are done, the seismograms are collected,
-and the useful database files are copied to the frontend, you may
-need to clean the local scratch disk for the next simulation. This
-is especially important in the case of kernel simulation, where very
-large files are generated for the absorbing boundaries to help with
-the reconstruction of the regular forward wavefield. A sample script
-is provided in \texttt{utils/}:
-\begin{lyxcode}
-cleanbase.pl~machines
-\end{lyxcode}
-where \texttt{machines} is a file containing the node names.
-
-\section{Plot Movie Snapshots and Synthetic Shakemaps}
-
-\subsection{Script \texttt{movie2gif.gmt.pl}}
-
-With the movie data saved in \texttt{OUTPUT\_FILES/} at the end of
-a movie simulation (\texttt{\small MOVIE\_SURFACE=.true.}{\small ),
-you can run the }\texttt{\small `create\_movie\_shakemap\_AVS\_DX\_GMT}{\small '
-code to convert these binary movie data into GMT xyz files for futher
-processing. A sample script }\texttt{\small movie2gif.gmt.pl}{\small{}
-is provided to do this conversion, and then plot the movie snapshots
-in GMT, for example:}{\small \par}
-\begin{lyxcode}
-movie2gif.gmt.pl~-m~CMTSOLUTION~-g~-f~1/40~-n~-2~-p~
-\end{lyxcode}
-which for the first through the 40th movie frame, converts the \texttt{moviedata}
-files into GMT xyz files, interpolates them using the 'nearneighbor'
-command in GMT, and plots them on a 2D topography map. Note that `\texttt{-2}'
-and `\texttt{-p}' are both optional.
-
-\subsection{Script \texttt{plot\_shakemap.gmt.pl}}
-
-With the shakemap data saved in \texttt{OUTPUT\_FILES/} at the end
-of a shakemap simulation ~\\
- (\texttt{CREATE\_SHAKEMAP=.true.}), you can also run \texttt{`create\_movie\_shakemap\_AVS\_DX\_GMT}'
-code to convert the binary shakemap data into GMT xyz files. A sample
-script \texttt{plot\_shakemap.gmt.pl} is provided to do this conversion,
-and then plot the shakemaps in GMT, for example:
-\begin{lyxcode}
-plot\_shakemap.gmt.pl~data\_dir~type(1,2,3)~CMTSOLUTION~
-\end{lyxcode}
-where \texttt{type=1} for a displacement shakemap, \texttt{2} for
-velocity, and \texttt{3} for acceleration.
-
-\section{Map Local Database}
-
-A sample program \texttt{remap\_database} is provided to map the local
-database from a set of machines to another set of machines. This is
-especially useful when you want to run mesher and solver, or different
-types of solvers separately through a scheduler (refer to Chapter~\ref{cha:Scheduler}).
-\begin{lyxcode}
-run\_lsf.bash~-{}-gm-no-shmem~-{}-gm-copy-env~remap\_database~old\_machines~150
-\end{lyxcode}
-where \texttt{old\_machines} is the LSF machine file used in the previous
-simulation, and \texttt{150} is the number of processors in total.
-
-%============================================================
-
-\chapter{\label{cha:developers}Information for developers of the code, and for people who want to learn how the technique works}
-
-You can get a very simple 1D version of a demo code (there is one in Fortran and one in Python):\\
-
-git clone --recursive https://github.com/geodynamics/specfem1d.git\\
-
-\noindent We also have simple 3D demo source codes that implement the SEM in a single, small program, in directory\\
-utils/small\_SEM\_solvers\_in\_Fortran\_and\_C\_without\_MPI\_to\_learn of the specfem3d package.
-They are useful to learn how the spectral-element method works, and how to write or modify a code to implement it.
-Also useful to test new ideas by modifying these simple codes to run some tests.\\
-
-\noindent For information on how to contribute to the code, i.e., for how to make your modifications, additions or improvements part of the
-official package, see https://github.com/geodynamics/specfem3d/wiki .
-
-%============================================================
-
-\chapter*{\label{cha:Bug-Reports-and}Bug Reports and Suggestions for Improvements}
-
-To report bugs or suggest improvements to the code, please send an
-e-mail to the CIG Computational Seismology Mailing List \urlwithparentheses{cig-seismo@geodynamics.org}.
-
-\addcontentsline{toc}{chapter}{Notes \& Acknowledgments}
-
-\chapter*{Notes \& Acknowledgments}
-
-In order to keep the software package thread-safe in case a multithreaded
-implementation of MPI is used, developers should not add modules or
-common blocks to the source code but rather use regular subroutine
-arguments (which can be grouped in ``derived types'' if needed for
-clarity).
-
-The Gauss-Lobatto-Legendre subroutines in \texttt{gll\_library.f90}
-are based in part on software libraries from the Massachusetts Institute
-of Technology, Department of Mechanical Engineering (Cambridge, Massachusetts,
-USA). The non-structured global numbering software was provided by
-Paul F. Fischer (Brown University, Providence, Rhode Island, USA,
-now at Argonne National Laboratory, USA).
-
-OpenDX \urlwithparentheses{http://www.opendx.org} is open-source
-based on IBM Data Explorer, AVS \urlwithparentheses{http://www.avs.com}
-is a trademark of Advanced Visualization Systems, and ParaView \urlwithparentheses{http://www.paraview.org}
-is an open-source visualization platform.
-
-Please e-mail your feedback, questions, comments, and suggestions
-to the CIG Computational Seismology Mailing List \urlwithparentheses{cig-seismo@geodynamics.org}.
-
-\addcontentsline{toc}{chapter}{Copyright}
-
-\chapter*{Copyright}
-
-Main historical authors: Dimitri Komatitsch and Jeroen Tromp
-
-Princeton University, USA, and CNRS / University of Marseille, France
-
-$\copyright$ Princeton University and CNRS / University of Marseille, July 2012
-
-This program is free software; you can redistribute it and/or modify
-it under the terms of the GNU General Public License as published
-by the Free Software Foundation (see Appendix \ref{cha:License}).\\
-
-\textbf{\underline{Evolution of the code:}}\\
-
-MPI v. 3.0, December 2014: many developers.
-Convolutional PML, LDDRK time scheme, bulk attenuation support, simultaneous MPI runs,
-ADIOS file I/O support, coupling with external codes, new seismogram names,
-Deville routines for additional GLL degrees, tomography tools, unit/regression test framework,
-improved CUDA GPUs performance, additonal GEOCUBIT support, better make compilation,
-git versioning system. \\
-
-MPI v. 2.1, July 2012: Max Rietmann, Peter Messmer, Daniel Peter, Dimitri
-Komatitsch, Joseph Charles, Zhinan Xie: support for CUDA GPUs, better
-CFL stability for the Stacey absorbing conditions. \\
-
-MPI v. 2.0, November 2010: Dimitri Komatitsch, Nicolas Le Goff, Roland
-Martin and Pieyre Le Loher, University of Pau, France, Daniel Peter,
-Jeroen Tromp and the Princeton group of developers, Princeton University,
-USA, and Emanuele Casarotti, INGV Roma, Italy: support for CUBIT meshes
-decomposed by SCOTCH; much faster solver using Michel Deville's inlined
-matrix products.\\
-
-MPI v. 1.4 Dimitri Komatitsch, University of Pau, Qinya Liu and others,
-Caltech, September 2006: better adjoint and kernel calculations, faster
-and better I/Os on very large systems, many small improvements and
-bug fixes.\\
-
-MPI v. 1.3 Dimitri Komatitsch, University of Pau, and Qinya Liu, Caltech,
-July 2005: serial version, regular mesh, adjoint and kernel calculations,
-ParaView support.\\
-
-MPI v. 1.2 Min Chen and Dimitri Komatitsch, Caltech, July 2004: full
-anisotropy, volume movie.\\
-
-MPI v. 1.1 Dimitri Komatitsch, Caltech, October 2002: Zhu's Moho map,
-scaling of $V_{s}$ with depth, Hauksson's regional model, attenuation,
-oceans, movies.\\
-
-MPI v. 1.0 Dimitri Komatitsch, Caltech, USA, May 2002: first MPI version
-based on global code.\\
-
-Dimitri Komatitsch, IPG Paris, France, December 1996: first 3-D solver
-for the CM-5 Connection Machine, parallelized on 128 processors using
-Connection Machine Fortran.\\
-
-\addcontentsline{toc}{chapter}{References} \bibliography{bibliography}
-
-\appendix
-%dummy comment inserted by tex2lyx to ensure that this paragraph is not empty
-
-\chapter{\label{cha:Coordinates}Reference Frame Convention}
-
-The code uses the following convention for the Cartesian reference
-frame:
-\begin{itemize}
-\item the $x$ axis points East
-\item the $y$ axis points North
-\item the $z$ axis points up
-\end{itemize}
-Note that this convention is different from both the \citet{AkRi80}
-convention and the Harvard Centroid-Moment Tensor (CMT) convention.
-The Aki \& Richards convention is
-\begin{itemize}
-\item the $x$ axis points North
-\item the $y$ axis points East
-\item the $z$ axis points down
-\end{itemize}
-and the Harvard CMT convention is
-\begin{itemize}
-\item the $x$ axis points South
-\item the $y$ axis points East
-\item the $z$ axis points up
-\end{itemize}
-
-\subsection*{Source and receiver locations}
-
-The SPECFEM3D Cartesian software code internally uses Cartesian coordinates.
-The given locations for sources and receiver locations thus may get
-converted. Sources and receiver locations are read in from the \texttt{CMTSOLUTION}
-(or \texttt{FORCESOLUTION}) and \texttt{STATIONS} files. Note that
-e.g. the \texttt{CMTSOLUTION} and \texttt{FORCESOLUTION} files denotes
-the location by \textquotedbl{}longitude/latitude/depth\textquotedbl{}.
-We read in longitude as $x$ coordinate, latitude as $y$ coordinate.
-
-In case the flag \texttt{SUPPRESS\_UTM\_PROJECTION} is set to \texttt{.false.}
-in the main parameter file (see Chapter~\ref{cha:Creating-Distributed-Databases}),
-the $x$/$y$ coordinates have to be given in degrees and are converted
-to Cartesian coordinates using a UTM conversion with the specified
-UTM zone.
-
-The value for depth (given in $km$ in \texttt{CMTSOLUTION} and \texttt{FORCESOLUTION})
-or burial depth (given in $m$ in \texttt{STATIONS}) is evaluated
-with respect to the surface of the mesh at the specified $x$/$y$
-location to find the corresponding $z$ coordinate. It is possible
-to use this depth value directly as $z$ coordinate by changing the
-flag \texttt{USE\_SOURCES\_RECVS\_Z} to \texttt{.true.} in the file
-\texttt{constants.h} located in the \texttt{src/shared/} subdirectory.
-
-\subsection*{Seismogram outputs}
-
-The seismogram output directions are given in Cartesian $x$/$y$/$z$
-directions and not rotated any further. Changing flags in \texttt{constants.h}
-in the \texttt{src/shared/} subdirectory only rotates the seismogram
-outputs if receivers are forced to be located at the surface (\texttt{RECVS\_CAN\_BE\_BURIED\_EXT\_MESH}
-set to \texttt{.false.}) and the normal to the surface at the receiver
-location should be used (\texttt{EXT\_MESH\_RECV\_NORMAL} set to \texttt{.true.})
-as vertical. In this case, the outputs are rotated to have the vertical
-component normal to the surface of the mesh, $x$ and $y$ directions
-are somewhat arbitrary orthogonal directions along the surface.
-
-For the labeling of the seismogram channels, see Appendix~\ref{cha:channel-codes}.
-Additionally, we add labels to the synthetics using the following
-convention: For a receiver station located in an
-\begin{description}
-\item [{elastic domain:}] ~\\
-
-\begin{itemize}
-\item \texttt{semd} for the displacement vector
-\item \texttt{semv} for the velocity vector
-\item \texttt{sema} for the acceleration vector
-\end{itemize}
-\item [{acoustic domain:}] ~\\
- (please note that receiver stations in acoustic domains must be buried.
-This is due to the free surface condition which enforces a zero pressure
-at the surface.)
-
-\begin{itemize}
-\item \texttt{semd} for the displacement vector
-\item \texttt{semv} for the velocity vector
-\item \texttt{sema} for pressure which will be stored in the vertical component
-\texttt{Z}. Note that pressure in the acoustic media is isotropic
-and thus the pressure record would be the same in the other two component
-directions. We therefore use the other two seismogram components to
-store the acoustic potential in component \texttt{X} (or \texttt{N})
-and the first time derivative of the acoustic potential in component
-\texttt{Y} (or \texttt{E}).
-\end{itemize}
-\end{description}
-The seismograms are by default written out in ASCII-format to the
-\texttt{OUTPUT\_FILES/} subdirectory by each parallel process. You
-can change this behavior by changing the following flags in the \texttt{constants.h}
-file located in the \texttt{src/shared/} subdirectory:
-\begin{description}
-\item [{\texttt{SEISMOGRAMS\_BINARY}}] set to \texttt{.true.} to have seismograms
-written out in binary format.
-\item [{\texttt{WRITE\_SEISMOGRAMS\_BY\_MASTER}}] Set to \texttt{.true.}
-to have only the master process writing out seismograms. This can
-be useful on a cluster, where only the master process node has access
-to the output directory.
-\item [{\texttt{USE\_OUTPUT\_FILES\_PATH}}] Set to \texttt{.false.} to
-have the seismograms output to \texttt{LOCAL\_PATH} directory specified
-in the main parameter file \texttt{DATA/Par\_file}.
-\end{description}
-
-\chapter{\label{cha:channel-codes}Channel Codes of Seismograms}
-
-Seismic networks, such as the Global Seismographic Network (GSN),
-generally involve various types of instruments with different bandwidths,
-sampling properties and component configurations. There are standards
-to name channel codes depending on instrument properties. IRIS \texttt{(www.iris.edu)}
-uses SEED format for channel codes, which are represented by three
-letters, such as \texttt{LHN}, \texttt{BHZ}, etc. In older versions
-of the SPECFEM3D Cartesian package, a common format was used for the
-channel codes of all seismograms, which was \texttt{BHE/BHN/BHZ} for
-three components. To avoid confusion when comparison are made to observed
-data, we are now using the FDSN convention (\url{http://www.fdsn.org/})
-for SEM seismograms. In the following, we give a brief explanation
-of the FDSN convention used by IRIS, and how it is adopted in SEM
-seismograms. Please visit \texttt{www.iris.edu/manuals/SEED\_appA.htm}
-for further information.\\
-
-\noindent \texttt{\textbf{Band code:}} The first letter in the channel
-code denotes the band code of seismograms, which depends on the response
-band and the sampling rate of instruments. The list of band codes
-used by IRIS is shown in Figure \ref{fig:IRIS_band_codes}. The sampling
-rate of SEM synthetics is controlled by the resolution of simulations
-rather than instrument properties. However, for consistency, we follow
-the FDSN convention for SEM seismograms governed by their sampling
-rate. For SEM synthetics, we consider band codes for which $dt\leq1$
-s. IRIS also considers the response band of instruments. For instance,
-short-period and broad-band seismograms with the same sampling rate
-correspond to different band codes, such as S and B, respectively.
-In such cases, we consider SEM seismograms as broad band, ignoring
-the corner period ($\geq10$ s) of the response band of instruments
-(note that at these resolutions, the minimum period in the SEM synthetics
-will be less than $10$ s). Accordingly, when you run a simulation
-the band code will be chosen depending on the resolution of the synthetics,
-and channel codes of SEM seismograms will start with either \texttt{L},
-\texttt{M}, \texttt{B}, \texttt{H}, \texttt{C} or \texttt{F}, shown
-by red color in the figure.\\
-
-\begin{figure}[ht]
-\noindent \begin{centering}
-\includegraphics[scale=0.6]{figures/IRIS_band_codes}
-\par\end{centering}
-
-\caption{The band code convention is based on the sampling rate and the response
-band of instruments. Please visit \texttt{www.iris.edu/manuals/SEED\_appA.htm}
-for further information. Grey rows show the relative band-code range
-in SPECFEM3D Cartesian, and the band codes used to name SEM seismograms
-are denoted in red.}
-
-\label{fig:IRIS_band_codes}
-\end{figure}
-
-\noindent \texttt{\textbf{Instrument code:}} The second letter in
-the channel code corresponds to instrument codes, which specify the
-family to which the sensor belongs. For instance, \texttt{H} and \texttt{L}
-are used for high-gain and low-gain seismometers, respectively. The
-instrument code of SEM seismograms will always be \texttt{X}, as assigned
-by FDSN for synthetic seismograms. \\
-
-\noindent \texttt{\textbf{Orientation code:}} The third letter in
-channel codes is an orientation code, which generally describes the
-physical configuration of the components of instrument packages. SPECFEM3D
-Cartesian uses the traditional orientation code \texttt{E/N/Z} (East-West,
-North-South, Vertical) for three components when a UTM projection
-is used. If the UTM conversion is suppressed, i.e. the flag \texttt{SUPPRESS\_UTM\_PROJECTION}
-is set to \texttt{.true.}, then the three components are labelled
-\texttt{X/Y/Z} according to the Cartesian reference frame. \\
-
-\noindent \texttt{\textbf{EXAMPLE:}} The sampling rate is given by
-\texttt{DT} in the main parameter file \texttt{Par\_file} located
-in the \texttt{DATA/} subdirectory. Depending on the resolution of
-your simulations, if you choose a sampling rate greater than $0.01$
-s and less than $1$ s, a seismogram recording displacements on the
-vertical component of a station \texttt{ASBS} (network \texttt{AZ})
-will be named \texttt{AZ.ASBS.MXZ.semd.sac}, whereas it will be \texttt{AZ.ASBS.BXZ.semd.sac},
-if the sampling rate is greater than 0.0125 and less equal to 0.1
-s.
-
-\chapter{\label{cha:Troubleshooting}Troubleshooting}
-
-\section*{FAQ}
-\begin{description}
-\item [{configuration fails:}]~
-
-Examine the log file 'config.log'. It contains detailed informations.
-In many cases, the paths to these specific compiler commands F90,
-CC and MPIF90 will not be correct if `./configure` fails.
-
-Please make sure that you have a working installation of a Fortran
-compiler, a C compiler and an MPI implementation. You should be able
-to compile this little program code:
-\begin{lyxcode}
-program~main~~~\\
-~include~'mpif.h'~~~\\
-~integer,~parameter~::~CUSTOM\_MPI\_TYPE~=~MPI\_REAL~~~\\
-~integer~ier~~~\\
-~call~MPI\_INIT(ier)~~~\\
-~call~MPI\_BARRIER(MPI\_COMM\_WORLD,ier)~~~\\
-~call~MPI\_FINALIZE(ier)~~~\\
-~end
-\end{lyxcode}
-\item [{compilation fails stating:}] ~\\
-
-\begin{lyxcode}
-...~~~\\
-~obj/program\_generate\_databases.o:~In~function~`MAIN\_\_':~~~\\
-~program\_generate\_databases.f90:(.text+0x14):~undefined~reference~to~`\_gfortran\_set\_std'~~~\\
-~...~~~\\
-
-\end{lyxcode}
-
-Make sure you're pointing to the right 'mpif90' wrapper command.
-
-Normally, this message will appear when you are mixing two different
-Fortran compilers. That is, using e.g. gfortran to compile non-MPI
-files and mpif90, wrapper provided for e.g. ifort, to compile MPI-files.
-
-fix: e.g. specify > ./configure FC=gfortran MPIF90=/usr/local/openmpi-gfortran/bin/mpif90
-
-\item [{after executing \texttt{xmeshfem3D} I've got elements with skewness of 81\% percent, what does this mean:}] Look
-at the skewness table printed in the \texttt{output\_mesher.txt} file
-after executing \texttt{xmeshfem3D} for the example given in \texttt{EXAMPLES/meshfem3D\_examples/simple\_model/}:
-
-\begin{lyxcode}
-...~~\\
-~histogram~of~skewness~(0.~good~-~1.~bad):~~~\\
-~0.0000000E+00~-~5.0000001E-02~27648~81.81818~\%~~~\\
-~5.0000001E-02~-~0.1000000~0~0.0000000E+00~\%~~~\\
-~...~~~\\
-
-\end{lyxcode}
-
-The first line means that you have 27,648 elements with a skewness
-value between 0 and 0.05 (which means the element is basically not
-skewed, just plain regular hexahedral element). The total number of
-elements you have in this mesh is (see in the \texttt{output\_mesher.txt}
-file a bit further down):
-\begin{lyxcode}
-...~~\\
-~total~number~of~elements~in~entire~mesh:~33792~~~\\
-~...~~~\\
-
-\end{lyxcode}
-
-which gives you that: 27,648 / 33,792 $\sim$ 81.8 \% of all elements
-are not skewed, i.e. regular elements. a fantastic value :)
-
-The histogram lists for this mesh also some stronger skewed elements,
-for example the worst ones belong to:
-\begin{lyxcode}
-...~~\\
-~0.6000000~-~0.6500000~2048~6.060606~\%~~~\\
-~...~~~\\
-
-\end{lyxcode}
-
-about 6 \% of all elements have distortions with a skewness value
-between 0.6 and 0.65. The skewness values give you a hint of how good
-your mesh is. In an ideal world, you would want to have no distortions,
-just like the 81\% from above. Those elements give you the best approximate
-values by the GLL quadrature used in the spectral-element method.
-However, having weakly distorted elements is still fine and the solutions
-are still accurate enough. So empirically, values up to around 0.7
-are tolerable, above that you should consider remeshing...
-
-To give you an idea why some of the elements are distorted, see the
-following figure \ref{fig:mesh.vp} of the mesh you obtain in the
-example \texttt{EXAMPLES/meshfem3D\_examples/simple\_model/}.
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=0.9\textwidth]{figures/mesh_vp}
-\par\end{centering}
-
-\caption{Paraview visualization using the mesh vtk-files for the example given
-in \texttt{EXAMPLES/meshfem3D\_examples/simple\_model/}.}
-
-\label{fig:mesh.vp}
-\end{figure}
-
-You will see that the mesh contains a doubling layer, where we stitch
-elements together such that the size of two elements will transition
-to the size of one element (very useful to keep the ratio of wavespeed
-/ element\_size about constant). Those elements in this doubling layer
-have higher skewness values and make up those 6 \% in the histogram.
-
-\item [{the code gives following error message "need at least one receiver":}] This
-means that no stations given in the input file \texttt{DATA/STATIONS}
-could be located within the dimensions of the mesh. This can happen
-for example when the mesh was created with the in-house mesher \texttt{xmeshfem3D}
-while using the Universal Transverse Mercator (UTM) projection but
-the simulation with \texttt{xspecfem3D} was suppressing this projection
-from latitude/longitude to x/y/z coordinates.
-
-In such cases, try to change your \texttt{DATA/Par\_file} and set
-e.g.:
-\begin{lyxcode}
-SUPPRESS\_UTM\_PROJECTION~=~.false.~~~\\
-
-\end{lyxcode}
-
-to be the same in \texttt{Mesh\_Par\_file} and \texttt{Par\_file}.
-This flag should be identical when using the in-house mesher \texttt{xmeshfem3D},
-\texttt{xgenerate\_databases} and \texttt{xspecfem3D} together to
-run simulations.
-
-The flag determines if the coordinates you specify for your source
-and station locations are given as lat/lon degrees and must be converted
-to UTM coordinates. As an example, if you use \texttt{.false.} within
-\texttt{Mesh\_Par\_file} then you create a mesh with \texttt{xmeshfem3D}
-using the UTM projection from lat/lon as input format to UTM projected
-coordinates to store the mesh point positions, which is fine. The
-error then may occur if in the \texttt{Par\_file} you have this set
-to \texttt{.true.} so that the \texttt{xgenerate\_databases} and \texttt{xspecfem3D}
-suppress the UTM projection and assume that all coordinates you use
-now for source and receiver locations are given in meters (that is,
-converted) already. So it won't find the specified locations in the
-used mesh. As a solutions, just change the flag in \texttt{Par\_file}
-to be the same as in \texttt{Mesh\_Par\_file} and rerun \texttt{xgenerate\_databases}
-and \texttt{xspecfem3D} to make sure that your simulation works fine.
-
-\item [{I get the following error message "forward simulation became unstable and blew up":}] In
-most cases this means that your time step size \texttt{DT} is chosen
-too big. Look at your files \texttt{output\_mesher.txt} or \texttt{output\_solver.txt}
-created in the folder \texttt{OUTPUT\_FILES}. In these output files,
-find the section:
-
-\begin{lyxcode}
-...~~~\\
-~{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}~~~\\
-~{*}{*}{*}~Verification~of~simulation~parameters~{*}{*}{*}~~~\\
-~{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}~~~\\
-~...~~~\\
-~{*}{*}{*}~Minimum~period~resolved~=~4.308774~~~\\
-~{*}{*}{*}~Maximum~suggested~time~step~=~6.8863556E-02~~~\\
-~...~~~\\
-~...~~~\\
-
-\end{lyxcode}
-
-then change \texttt{DT} in the \texttt{DATA/Par\_file} to be somewhere
-close to the maximum suggested time step. In the example above: \\
- \texttt{DT = 0.05d0} \\
- would (most probably) work fine. It could be also bigger than the
-0.068 s suggested. This depends a bit on the distortions of your mesh
-elements. The more regular they are, the bigger you can choose \texttt{DT}.
-Just play with this value a bit and see when the simulation becomes
-stable ...
-
-\end{description}
-
-\chapter{\label{cha:License}License }
-
-\textbf{GNU GENERAL PUBLIC LICENSE Version 2, June 1991. Copyright
-(C) 1989, 1991 Free Software Foundation, Inc. 59 Temple Place, Suite
-330, Boston, MA 02111-1307 USA} \\
- Everyone is permitted to copy and distribute verbatim copies of this
-license document, but changing it is not allowed.
-
-\section*{Preamble}
-
-The licenses for most software are designed to take away your freedom
-to share and change it. By contrast, the GNU General Public License
-is intended to guarantee your freedom to share and change free software
--- to make sure the software is free for all its users. This General
-Public License applies to most of the Free Software Foundation's software
-and to any other program whose authors commit to using it. (Some other
-Free Software Foundation software is covered by the GNU Library General
-Public License instead.) You can apply it to your programs, too.
-
-When we speak of free software, we are referring to freedom, not price.
-Our General Public Licenses are designed to make sure that you have
-the freedom to distribute copies of free software (and charge for
-this service if you wish), that you receive source code or can get
-it if you want it, that you can change the software or use pieces
-of it in new free programs; and that you know you can do these things.
-
-To protect your rights, we need to make restrictions that forbid anyone
-to deny you these rights or to ask you to surrender the rights. These
-restrictions translate to certain responsibilities for you if you
-distribute copies of the software, or if you modify it.
-
-For example, if you distribute copies of such a program, whether gratis
-or for a fee, you must give the recipients all the rights that you
-have. You must make sure that they, too, receive or can get the source
-code. And you must show them these terms so they know their rights.
-
-We protect your rights with two steps:
-\begin{enumerate}
-\item Copyright the software, and
-\item Offer you this license which gives you legal permission to copy, distribute
-and/or modify the software.
-\end{enumerate}
-Also, for each author's protection and ours, we want to make certain
-that everyone understands that there is no warranty for this free
-software. If the software is modified by someone else and passed on,
-we want its recipients to know that what they have is not the original,
-so that any problems introduced by others will not reflect on the
-original authors' reputations.
-
-Finally, any free program is threatened constantly by software patents.
-We wish to avoid the danger that redistributors of a free program
-will individually obtain patent licenses, in effect making the program
-proprietary. To prevent this, we have made it clear that any patent
-must be licensed for everyone's free use or not licensed at all.
-
-The precise terms and conditions for copying, distribution and modification
-follow.
-
-\section*{GNU GENERAL PUBLIC LICENSE TERMS AND CONDITIONS FOR COPYING, DISTRIBUTION
-AND MODIFICATION }
-\begin{itemize}
-\item 0.This License applies to any program or other work which contains
-a notice placed by the copyright holder saying it may be distributed
-under the terms of this General Public License. The ``Program''
-below refers to any such program or work, and a ``work based on the
-Program'' means either the Program or any derivative work under copyright
-law: that is to say, a work containing the Program or a portion of
-it, either verbatim or with modifications and/or translated into another
-language. (Hereinafter, translation is included without limitation
-in the term ``modification.'') Each licensee is addressed as ``you.''\\
- \\
- Activities other than copying, distribution and modification are
-not covered by this License; they are outside its scope. The act of
-running the Program is not restricted, and the output from the Program
-is covered only if its contents constitute a work based on the Program
-(independent of having been made by running the Program). Whether
-that is true depends on what the Program does.\end{itemize}
-\begin{enumerate}
-\item You may copy and distribute verbatim copies of the Program's source
-code as you receive it, in any medium, provided that you conspicuously
-and appropriately publish on each copy an appropriate copyright notice
-and disclaimer of warranty; keep intact all the notices that refer
-to this License and to the absence of any warranty; and give any other
-recipients of the Program a copy of this License along with the Program.
-
-You may charge a fee for the physical act of transferring a copy,
-and you may at your option offer warranty protection in exchange for
-a fee.
-
-\item You may modify your copy or copies of the Program or any portion of
-it, thus forming a work based on the Program, and copy and distribute
-such modifications or work under the terms of Section 1 above, provided
-that you also meet all of these conditions:
-
-\begin{enumerate}
-\item You must cause the modified files to carry prominent notices stating
-that you changed the files and the date of any change.
-\item You must cause any work that you distribute or publish, that in whole
-or in part contains or is derived from the Program or any part thereof,
-to be licensed as a whole at no charge to all third parties under
-the terms of this License.
-\item If the modified program normally reads commands interactively when
-run, you must cause it, when started running for such interactive
-use in the most ordinary way, to print or display an announcement
-including an appropriate copyright notice and a notice that there
-is no warranty (or else, saying that you provide a warranty) and that
-users may redistribute the program under these conditions, and telling
-the user how to view a copy of this License. (Exception: if the Program
-itself is interactive but does not normally print such an announcement,
-your work based on the Program is not required to print an announcement.)
-\end{enumerate}
-
-These requirements apply to the modified work as a whole. If identifiable
-sections of that work are not derived from the Program, and can be
-reasonably considered independent and separate works in themselves,
-then this License, and its terms, do not apply to those sections when
-you distribute them as separate works. But when you distribute the
-same sections as part of a whole which is a work based on the Program,
-the distribution of the whole must be on the terms of this License,
-whose permissions for other licensees extend to the entire whole,
-and thus to each and every part regardless of who wrote it.
-
-Thus, it is not the intent of this section to claim rights or contest
-your rights to work written entirely by you; rather, the intent is
-to exercise the right to control the distribution of derivative or
-collective works based on the Program.
-
-In addition, mere aggregation of another work not based on the Program
-with the Program (or with a work based on the Program) on a volume
-of a storage or distribution medium does not bring the other work
-under the scope of this License.
-
-\item You may copy and distribute the Program (or a work based on it, under
-Section 2) in object code or executable form under the terms of Sections
-1 and 2 above provided that you also do one of the following:
-
-\begin{enumerate}
-\item Accompany it with the complete corresponding machine-readable source
-code, which must be distributed under the terms of Sections 1 and
-2 above on a medium customarily used for software interchange; or,
-\item Accompany it with a written offer, valid for at least three years,
-to give any third party, for a charge no more than your cost of physically
-performing source distribution, a complete machine-readable copy of
-the corresponding source code, to be distributed under the terms of
-Sections 1 and 2 above on a medium customarily used for software interchange;
-or,
-\item Accompany it with the information you received as to the offer to
-distribute corresponding source code. (This alternative is allowed
-only for noncommercial distribution and only if you received the program
-in object code or executable form with such an offer, in accord with
-Subsection b above.)
-\end{enumerate}
-
-The source code for a work means the preferred form of the work for
-making modifications to it. For an executable work, complete source
-code means all the source code for all modules it contains, plus any
-associated interface definition files, plus the scripts used to control
-compilation and installation of the executable. However, as a special
-exception, the source code distributed need not include anything that
-is normally distributed (in either source or binary form) with the
-major components (compiler, kernel, and so on) of the operating system
-on which the executable runs, unless that component itself accompanies
-the executable.
-
-If distribution of executable or object code is made by offering access
-to copy from a designated place, then offering equivalent access to
-copy the source code from the same place counts as distribution of
-the source code, even though third parties are not compelled to copy
-the source along with the object code.
-
-\item You may not copy, modify, sublicense, or distribute the Program except
-as expressly provided under this License. Any attempt otherwise to
-copy, modify, sublicense or distribute the Program is void, and will
-automatically terminate your rights under this License. However, parties
-who have received copies, or rights, from you under this License will
-not have their licenses terminated so long as such parties remain
-in full compliance.
-\item You are not required to accept this License, since you have not signed
-it. However, nothing else grants you permission to modify or distribute
-the Program or its derivative works. These actions are prohibited
-by law if you do not accept this License. Therefore, by modifying
-or distributing the Program (or any work based on the Program), you
-indicate your acceptance of this License to do so, and all its terms
-and conditions for copying, distributing or modifying the Program
-or works based on it.
-\item Each time you redistribute the Program (or any work based on the Program),
-the recipient automatically receives a license from the original licensor
-to copy, distribute or modify the Program subject to these terms and
-conditions. You may not impose any further restrictions on the recipients'
-exercise of the rights granted herein. You are not responsible for
-enforcing compliance by third parties to this License.
-\item If, as a consequence of a court judgment or allegation of patent infringement
-or for any other reason (not limited to patent issues), conditions
-are imposed on you (whether by court order, agreement or otherwise)
-that contradict the conditions of this License, they do not excuse
-you from the conditions of this License. If you cannot distribute
-so as to satisfy simultaneously your obligations under this License
-and any other pertinent obligations, then as a consequence you may
-not distribute the Program at all. For example, if a patent license
-would not permit royalty-free redistribution of the Program by all
-those who receive copies directly or indirectly through you, then
-the only way you could satisfy both it and this License would be to
-refrain entirely from distribution of the Program.
-
-If any portion of this section is held invalid or unenforceable under
-any particular circumstance, the balance of the section is intended
-to apply and the section as a whole is intended to apply in other
-circumstances.
-
-It is not the purpose of this section to induce you to infringe any
-patents or other property right claims or to contest validity of any
-such claims; this section has the sole purpose of protecting the integrity
-of the free software distribution system, which is implemented by
-public license practices. Many people have made generous contributions
-to the wide range of software distributed through that system in reliance
-on consistent application of that system; it is up to the author/donor
-to decide if he or she is willing to distribute software through any
-other system and a licensee cannot impose that choice.
-
-This section is intended to make thoroughly clear what is believed
-to be a consequence of the rest of this License.
-
-\item If the distribution and/or use of the Program is restricted in certain
-countries either by patents or by copyrighted interfaces, the original
-copyright holder who places the Program under this License may add
-an explicit geographical distribution limitation excluding those countries,
-so that distribution is permitted only in or among countries not thus
-excluded. In such case, this License incorporates the limitation as
-if written in the body of this License.
-\item The Free Software Foundation may publish revised and/or new versions
-of the General Public License from time to time. Such new versions
-will be similar in spirit to the present version, but may differ in
-detail to address new problems or concerns.
-
-Each version is given a distinguishing version number. If the Program
-specifies a version number of this License which applies to it and
-``any later version,'' you have the option of following the terms
-and conditions either of that version or of any later version published
-by the Free Software Foundation. If the Program does not specify a
-version number of this License, you may choose any version ever published
-by the Free Software Foundation.
-
-\item If you wish to incorporate parts of the Program into other free programs
-whose distribution conditions are different, write to the author to
-ask for permission. For software which is copyrighted by the Free
-Software Foundation, write to the Free Software Foundation; we sometimes
-make exceptions for this. Our decision will be guided by the two goals
-of preserving the free status of all derivatives of our free software
-and of promoting the sharing and reuse of software generally.
-\end{enumerate}
-
-\subsection*{NO WARRANTY }
-\begin{itemize}
-\item 11.BECAUSE THE PROGRAM IS LICENSED FREE OF CHARGE, THERE IS NO WARRANTY
-FOR THE PROGRAM, TO THE EXTENT PERMITTED BY APPLICABLE LAW. EXCEPT
-WHEN OTHERWISE STATED IN WRITING THE COPYRIGHT HOLDERS AND/OR OTHER
-PARTIES PROVIDE THE PROGRAM ``AS IS'' WITHOUT WARRANTY OF ANY KIND,
-EITHER EXPRESSED OR IMPLIED, INCLUDING, BUT NOT LIMITED TO, THE IMPLIED
-WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE.
-THE ENTIRE RISK AS TO THE QUALITY AND PERFORMANCE OF THE PROGRAM IS
-WITH YOU. SHOULD THE PROGRAM PROVE DEFECTIVE, YOU ASSUME THE COST
-OF ALL NECESSARY SERVICING, REPAIR OR CORRECTION.
-\item 12.IN NO EVENT UNLESS REQUIRED BY APPLICABLE LAW OR AGREED TO IN WRITING
-WILL ANY COPYRIGHT HOLDER, OR ANY OTHER PARTY WHO MAY MODIFY AND/OR
-REDISTRIBUTE THE PROGRAM AS PERMITTED ABOVE, BE LIABLE TO YOU FOR
-DAMAGES, INCLUDING ANY GENERAL, SPECIAL, INCIDENTAL OR CONSEQUENTIAL
-DAMAGES ARISING OUT OF THE USE OR INABILITY TO USE THE PROGRAM (INCLUDING
-BUT NOT LIMITED TO LOSS OF DATA OR DATA BEING RENDERED INACCURATE
-OR LOSSES SUSTAINED BY YOU OR THIRD PARTIES OR A FAILURE OF THE PROGRAM
-TO OPERATE WITH ANY OTHER PROGRAMS), EVEN IF SUCH HOLDER OR OTHER
-PARTY HAS BEEN ADVISED OF THE POSSIBILITY OF SUCH DAMAGES.
-\end{itemize}
-
-\section*{END OF TERMS AND CONDITIONS }
-
-\subsection*{How to Apply These Terms to Your New Programs}
-
-If you develop a new program, and you want it to be of the greatest
-possible use to the public, the best way to achieve this is to make
-it free software which everyone can redistribute and change under
-these terms.
-
-To do so, attach the following notices to the program. It is safest
-to attach them to the start of each source file to most effectively
-convey the exclusion of warranty; and each file should have at least
-the ``copyright'' line and a pointer to where the full notice is
-found. For example:
-\begin{quote}
-One line to give the program's name and a brief idea of what it does.
-Copyright {\footnotesize $\copyright$ (}year) (name of author)
-
-This program is free software; you can redistribute it and/or modify
-it under the terms of the GNU General Public License as published
-by the Free Software Foundation; either version 2 of the License,
-or (at your option) any later version.
-
-This program is distributed in the hope that it will be useful, but
-WITHOUT ANY WARRANTY; without even the implied warranty of MERCHANTABILITY
-or FITNESS FOR A PARTICULAR PURPOSE. See the GNU General Public License
-for more details.
-
-You should have received a copy of the GNU General Public License
-along with this program; if not, write to the Free Software Foundation,
-Inc., 59 Temple Place, Suite 330, Boston, MA 02111-1307 USA
-\end{quote}
-Also add information on how to contact you by electronic and paper
-mail.
-
-If the program is interactive, make it output a short notice like
-this when it starts in an interactive mode:
-\begin{quote}
-Gnomovision version 69, Copyright $\copyright$ year name of author
-Gnomovision comes with ABSOLUTELY NO WARRANTY; for details type `show
-w'. This is free software, and you are welcome to redistribute it
-under certain conditions; type `show c' for details.
-\end{quote}
-The hypothetical commands `show w' and `show c' should show the appropriate
-parts of the General Public License. Of course, the commands you use
-may be called something other than `show w' and `show c'; they could
-even be mouse-clicks or menu items -- whatever suits your program.
-
-You should also get your employer (if you work as a programmer) or
-your school, if any, to sign a ``copyright disclaimer'' for the
-program, if necessary. Here is a sample; alter the names:
-\begin{quote}
-Yoyodyne, Inc., hereby disclaims all copyright interest in the program
-`Gnomovision' (which makes passes at compilers) written by James Hacker.
-
-(signature of Ty Coon)\\
- 1 April 1989 \\
- Ty Coon, President of Vice
-\end{quote}
-This General Public License does not permit incorporating your program
-into proprietary programs. If your program is a subroutine library,
-you may consider it more useful to permit linking proprietary applications
-with the library. If this is what you want to do, use the GNU Library
-General Public License instead of this License.
-\end{document}

### 2015-04-24 doc/USER_MANUAL/14_post_processing.tex
updates user manual to use facilitate the generation of a markdown version to the github wiki

--- 2015-04-18 doc/USER_MANUAL/manual_SPECFEM3D_Cartesian.tex
@@
-% !TEX encoding = IsoLatin
-%% LyX 2.0.2 created this file.  For more info, see http://www.lyx.org/.
-%% Do not edit unless you really know what you are doing.
-\documentclass[oneside,english]{book}
-\usepackage[T1]{fontenc}
-%% DK DK we do not accept accented letters, use the old LaTeX syntax
-%% DK DK for them instead: \'e \`e  \"e \^e and so on.
-%% DK DK \usepackage[latin1]{inputenc}
-\usepackage[letterpaper]{geometry}
-\geometry{verbose,tmargin=1in,bmargin=1in,lmargin=1in,rmargin=1in}
-\setcounter{secnumdepth}{3}
-\setcounter{tocdepth}{3}
-\usepackage{babel}
-\usepackage{float}
-\usepackage{textcomp}
-\usepackage{url}
-\usepackage{amsmath}
-\usepackage{amssymb}
-\usepackage{graphicx}
-\usepackage[authoryear]{natbib}
-\usepackage[unicode=true,pdfusetitle,
- bookmarks=true,bookmarksnumbered=true,bookmarksopen=false,
- breaklinks=false,pdfborder={0 0 0},backref=section,colorlinks=false,pdfpagemode=None]
- {hyperref}
-\hypersetup{
- pdftex,pdfstartview=FitH,pdfpagelayout=SinglePage}
-\usepackage{breakurl}
-
-\makeatletter
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% LyX specific LaTeX commands.
-\newcommand{\noun}[1]{\textsc{#1}}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Textclass specific LaTeX commands.
-\newenvironment{lyxcode}
-{\par\begin{list}{}{
-\setlength{\rightmargin}{\leftmargin}
-\setlength{\listparindent}{0pt}% needed for AMS classes
-\raggedright
-\setlength{\itemsep}{0pt}
-\setlength{\parsep}{0pt}
-\normalfont\ttfamily}%
- \item[]}
-{\end{list}}
-\newenvironment{lyxlist}[1]
-{\begin{list}{}
-{\settowidth{\labelwidth}{#1}
- \setlength{\leftmargin}{\labelwidth}
- \addtolength{\leftmargin}{\labelsep}
- \renewcommand{\makelabel}[1]{##1\hfil}}}
-{\end{list}}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% User specified LaTeX commands.
-%\renewcommand{\baselinestretch}{1.5}
-
-% figures
-\usepackage[dvips]{epsfig}
-
-% fonts
-\usepackage{times}
-
-% hyperlinks to sections and references
-
-\newcommand{\urlwithparentheses}[1]{(\url{#1})}
-
-% biblio GJI
-\bibliographystyle{abbrvnat}
-
-\newcommand{\nexxi}{\mbox{\texttt{NEX\_XI\/}}}
-\newcommand{\nexeta}{\mbox{\texttt{NEX\_ETA\/}}}
-\newcommand{\nprocxi}{\mbox{\texttt{NPROC\_XI\/}}}
-\newcommand{\nproceta}{\mbox{\texttt{NPROC\_ETA\/}}}
-\newcommand{\nchunks}{\mbox{\texttt{NCHUNKS\/}}}
-\newcommand{\eg}{e.g.,}
-
-% colors for the text
-\@ifundefined{definecolor}
- {\usepackage{color}}{}
-\newcommand{\red}[1]{\textcolor{Red}{#1}}
-
-\usepackage{babel}
-
-\makeatother
-
-\begin{document}
-\begin{center}
-\thispagestyle{empty}\vspace*{-1.8truecm} %
-\makebox[1\textwidth]{%
-\includegraphics[width=0.83\paperwidth]{figures/specfem_3d_Cartesian-cover} %
-}
-\par\end{center}
-
-\title{\textbf{SPECFEM3D Cartesian}\\
- \textbf{User Manual}}
-
-\author{$\copyright$ Princeton University (USA), CNRS and University of Marseille (France),\\
-and ETH Z\"urich (Switzerland)\\
-Version 3.0}
-
-\date{\today}
-
-\maketitle
-
-\section*{Authors}
-
-\noindent The SPECFEM3D Cartesian package was first developed by Dimitri
-Komatitsch and Jean-Pierre Vilotte at Institut de Physique du Globe
-(IPGP) in Paris, France from 1995 to 1997 and then by Dimitri Komatitsch
-and Jeroen Tromp at Harvard University and Caltech, USA, starting in 1998.
-The story started on April 4, 1995, when Prof. Yvon Maday from CNRS and University of Paris, France, gave a lecture to
-Dimitri Komatitsch and Jean-Pierre Vilotte at IPG about the nice properties of the Legendre spectral-element method with diagonal mass matrix that he had used for
-other equations. We are deeply indebted and thankful to him for that.
-That followed a visit by Dimitri Komatitsch to OGS (Istituto Nazionale di Oceanografia e di Geofisica Sperimentale) in Trieste, Italy, in February 1995
-to meet with G\'eza Seriani and Enrico Priolo, who introduced him to their 2D Chebyshev version of the spectral-element method with a non-diagonal mass matrix.
-We are deeply indebted and thankful to them for that.
-
-%%%%%%%% please do NOT merge the lines of the paragraph below, it makes it difficult to insert new names in alphabetical order; leave a single name per line
-%%%%%%%% please do NOT merge the lines of the paragraph below, it makes it difficult to insert new names in alphabetical order; leave a single name per line
-%%%%%%%% please do NOT merge the lines of the paragraph below, it makes it difficult to insert new names in alphabetical order; leave a single name per line
-Since then it has been developed and maintained by a development team:
-in alphabetical order,
-Jean-Paul (Pablo) Ampuero,
-Kangchen Bai,
-Piero Basini,
-C\'eline Blitz,
-Alexis Bottero,
-Ebru Bozda\u{g},
-Emanuele Casarotti,
-Joseph Charles,
-Min Chen,
-Paul Cristini,
-Percy Galvez,
-Dominik G\"oddeke,
-Vala Hj\"orleifsd\'ottir,
-Sue Kientz,
-Dimitri Komatitsch,
-Jes\'us Labarta,
-Nicolas Le Goff,
-Pieyre Le Loher,
-Matthieu Lefebvre,
-Qinya Liu,
-Yang Luo,
-Alessia Maggi,
-Federica Magnoni,
-Roland Martin,
-Ren\'e Matzen,
-Dennis McRitchie,
-Matthias Meschede,
-Peter Messmer,
-David Mich\'ea,
-Surendra Nadh Somala,
-Tarje Nissen-Meyer,
-Daniel Peter,
-Max Rietmann,
-Elliott Sales de Andrade,
-Brian Savage,
-Bernhard Schuberth,
-Anne Sieminski,
-Leif Strand,
-Carl Tape,
-Jeroen Tromp,
-Jean-Pierre Vilotte,
-Zhinan Xie,
-Chang-Hua Zhang,
-Hejun Zhu.\\
-
-The cover graphic of the manual was created by Santiago Lombeyda from
-Caltech's Center for Advanced Computing Research (CACR) \urlwithparentheses{http://www.cacr.caltech.edu}.\\
-
-The code is released open-source under the GNU version 2+ license, see the license at the end of this manual.\\
-
-\newpage{}
-\section*{Current and past main participants or main sponsors of the SPECFEM project (in no particular order)}
-%
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=0.125\textwidth]{figures/logo_cnrs}\vspace*{2truemm}
-\includegraphics[width=0.125\textwidth]{figures/logo_princeton}\vspace*{2truemm}
-\includegraphics[width=0.15\textwidth]{figures/logo_aix_marseille_universite}\vspace*{0.02truemm}
-\includegraphics[width=0.15\textwidth]{figures/logo_ETH}\vspace*{2truemm}
-\includegraphics[width=0.12\textwidth]{figures/logo_CSC_China}\vspace*{0.02truemm}
-\includegraphics[width=0.15\textwidth]{figures/logo_inria}\vspace*{2truemm}
-\includegraphics[width=0.14\textwidth]{figures/logo_UPPA}
-\par\end{centering}
-%
-\vspace*{-2truemm}
-%
-\noindent \begin{centering}
-\includegraphics[width=0.15\textwidth]{figures/logo_NVIDIA}\vspace*{2truemm}
-\includegraphics[width=0.125\textwidth]{figures/logo_IUF}\vspace*{2truemm}
-\includegraphics[width=0.135\textwidth]{figures/logo_Caltech}\vspace*{2truemm}
-\includegraphics[width=0.135\textwidth]{figures/logo_Harvard}\vspace*{2truemm}
-\includegraphics[width=0.135\textwidth]{figures/logo_IPGP}\vspace*{2truemm}
-\includegraphics[width=0.112\textwidth]{figures/logo_ANR}\vspace*{2truemm}
-\includegraphics[width=0.112\textwidth]{figures/logo_NSF}\vspace*{2truemm}
-\par\end{centering}
-%
-\vspace*{-2truemm}
-%
-\noindent \begin{centering}
-\includegraphics[width=0.190\textwidth]{figures/logo_European_Union}\vspace*{2truemm}
-\includegraphics[width=0.112\textwidth]{figures/logo_GENCI}\vspace*{2truemm}
-\includegraphics[width=0.112\textwidth]{figures/logo_PRACE}\vspace*{2truemm}
-\includegraphics[width=0.112\textwidth]{figures/logo_CINES}\vspace*{2truemm}
-\includegraphics[width=0.130\textwidth]{figures/logo_Oak_Ridge}\vspace*{2truemm}
-\hspace*{3mm}\includegraphics[width=0.130\textwidth]{figures/logo_fondation_Del_Duca}
-\includegraphics[width=0.130\textwidth]{figures/logo_CIG}\vspace*{2truemm}
-\par\end{centering}
-%
-\vspace*{-2truemm}
-%
-\noindent \begin{centering}
-\includegraphics[width=0.112\textwidth]{figures/logo_University_of_Toronto}\vspace*{2truemm}
-\includegraphics[width=0.109\textwidth]{figures/logo_INGV}\vspace*{2truemm}
-\includegraphics[width=0.102\textwidth]{figures/logo_Univ_Toulouse}\vspace*{2truemm}
-\includegraphics[width=0.112\textwidth]{figures/logo_TOTAL}\vspace*{2truemm}
-\includegraphics[width=0.130\textwidth]{figures/logo_Fairbanks}\vspace*{2truemm}
-\includegraphics[width=0.120\textwidth]{figures/logo_CINECA}\vspace*{2truemm}
-\includegraphics[width=0.140\textwidth]{figures/logo_Intel_Exascale_Labs}\vspace*{2truemm}
-\includegraphics[width=0.130\textwidth]{figures/logo_Maison_Simulation}
-\par\end{centering}
-\end{figure}
-
-\newpage{}
-
-\tableofcontents{}
-
-\chapter{Introduction}
-
-The software package SPECFEM3D Cartesian simulates seismic wave propagation
-at the local or regional scale and performs full waveform imaging (FWI) or adjoint tomography based upon the spectral-element method
-(SEM). The SEM is a continuous Galerkin technique \citep{TrKoLi08,PeKoLuMaLeCaLeMaLiBlNiBaTr11},
-which can easily be made discontinuous \citep{BeMaPa94,Ch00,KoWoHu02,ChCaVi03,LaWaBe05,Kop06,WiStBuGh10,AcKo11};
-it is then close to a particular case of the discontinuous Galerkin
-technique \citep{ReHi73,LeRa74,Arn82,JoPi86,BoMaHe91,FaRi99,HuHuRa99,CoKaSh00,GiHeWa02,RiWh03,MoRi05,GrScSc06,AiMoMu06,BeLaPi06,DuKa06,DeSeWh08,PuAmKa09,WiStBuGh10,DeSe10,EtChViGl10},
-with optimized efficiency because of its tensorized basis functions
-\citep{WiStBuGh10,AcKo11}. In particular, it can accurately handle
-very distorted mesh elements \citep{OlSe11}.\\
-
-It has very good accuracy and convergence properties \citep{MaPa89,SePr94,DeFiMu02,Coh02,DeSe07,SeOl08,AiWa09,AiWa10,MeStTh12}.
-The spectral element approach admits spectral rates of convergence
-and allows exploiting $hp$-convergence schemes. It is also very well
-suited to parallel implementation on very large supercomputers \citep{KoTsChTr03,TsKoChTr03,KoLaMi08a,CaKoLaTiMiLeSnTr08,KoViCh10}
-as well as on clusters of GPU accelerating graphics cards \citep{Kom11,MiKo10,KoMiEr09,KoErGoMi10}.
-Tensor products inside each element can be optimized to reach very
-high efficiency \citep{DeFiMu02}, and mesh point and element numbering
-can be optimized to reduce processor cache misses and improve cache
-reuse \citep{KoLaMi08a}. The SEM can also handle triangular (in 2D)
-or tetrahedral (in 3D) elements \citep{WinBoyd96,TaWi00,KoMaTrTaWi01,Coh02,MeViSa06}
-as well as mixed meshes, although with increased cost and reduced
-accuracy in these elements, as in the discontinuous Galerkin method.\\
-
-Note that in many geological models in the context of seismic wave
-propagation studies (except for instance for fault dynamic rupture
-studies, in which very high frequencies or supershear rupture need
-to be modeled near the fault, see e.g. \citet{BeGlCrViPi07,BeGlCrVi09,PuAmKa09,TaCrEtViBeSa10})
-a continuous formulation is sufficient because material property contrasts
-are not drastic and thus conforming mesh doubling bricks can efficiently
-handle mesh size variations \citep{KoTr02a,KoLiTrSuStSh04,LeChLiKoHuTr08,LeChKoHuTr09,LeKoHuTr09}.\\
-
-For a detailed introduction to the SEM as applied to regional seismic
-wave propagation, please consult \citet{PeKoLuMaLeCaLeMaLiBlNiBaTr11,TrKoLi08,KoVi98,KoTr99,ChKoViCaVaFe07}
-and in particular \citet{LeKoHuTr09,LeChKoHuTr09,LeChLiKoHuTr08,GoAmTaCaSmSaMaKo09,WiKoScTr04,KoLiTrSuStSh04}.
-A detailed theoretical analysis of the dispersion
-and stability properties of the SEM is available in \citet{Coh02}, \citet{DeSe07}, \citet{SeOl07}, \citet{SeOl08} and \citet{MeStTh12}.\\
-
-Effects due to lateral variations in compressional-wave speed, shear-wave
-speed, density, a 3D crustal model, topography and bathymetry are
-included. The package can accommodate full 21-parameter anisotropy
-(see~\citet{ChTr07}) as well as lateral variations in attenuation
-\citep{SaKoTr10}. Adjoint capabilities and finite-frequency kernel
-simulations are included \citep{TrKoLi08,PeKoLuMaLeCaLeMaLiBlNiBaTr11,LiTr06,FiIgBuKe09,ViOp09}.\\
-
-The SEM was originally developed in computational fluid dynamics \citep{Pat84,MaPa89}
-and has been successfully adapted to address problems in seismic wave
-propagation. Early seismic wave propagation applications of the SEM,
-utilizing Legendre basis functions and a perfectly diagonal mass matrix,
-include \citet{CoJoTo93}, \citet{Kom97}, \citet{FaMaPaQu97}, \citet{CaGa97},
-\citet{KoVi98} and \citet{KoTr99}, whereas applications involving
-Chebyshev basis functions and a nondiagonal mass matrix include \citet{SePr94},
-\citet{PrCaSe94} and \citet{SePrPr95}.\\
-
-All SPECFEM3D software is written in Fortran2003 with full portability
-in mind, and conforms strictly to the Fortran2003 standard. It uses
-no obsolete or obsolescent features of Fortran. The package uses parallel
-programming based upon the Message Passing Interface (MPI) \citep{GrLuSk94,Pac97}.\\
-
-SPECFEM3D won the Gordon Bell award for best performance at the SuperComputing~2003
-conference in Phoenix, Arizona (USA) (see \citet{KoTsChTr03} and
-\url{www.sc-conference.org/sc2003/nrfinalaward.html}). It was a finalist
-again in 2008 for a run at 0.16 petaflops (sustained) on 149,784 processors
-of the `Jaguar' Cray XT5 system at Oak Ridge National Laboratories
-(USA) \citep{CaKoLaTiMiLeSnTr08}. It also won the BULL Joseph Fourier
-supercomputing award in 2010.\\
-
-It reached the sustained one petaflop performance level for the first time in February 2013
-on the Blue Waters Cray supercomputer at the National Center for Supercomputing Applications (NCSA), located at the University of Illinois at Urbana-Champaign (USA).\\
-
-This new release of the code includes Convolution or
-Auxiliary Differential Equation Perfectly Matched absorbing Layers
-(C-PML or ADE-PML) \citep{MaKoEz08,MaKoGe08,MaKo09,MaKoGeBr10,KoMa07}.
-It also includes support for GPU graphics card acceleration \citep{Kom11,MiKo10,KoMiEr09,KoErGoMi10}.\\
-
-The next release of the code will use the PT-SCOTCH parallel and threaded
-version of SCOTCH for mesh partitioning instead of the serial version.\\
-
-SPECFEM3D Cartesian includes coupled fluid-solid domains and adjoint
-capabilities, which enables one to address seismological inverse problems,
-but for linear rheologies only so far. To accommodate visco-plastic
-or non-linear rheologies, the reader can refer to the \href{http://geoelse.stru.polimi.it/}{GeoELSE}software
-package \citep{CaGa97,StPaIg09}.
-
-\section{Citation}
-
-If you use SPECFEM3D Cartesian for your own research, please cite
-at least one of the following articles:
-\begin{description}
-\item [{Numerical simulations in general}] ~\\
- Forward and adjoint simulations are described in detail in \citet{TrKoLi08,PeKoLuMaLeCaLeMaLiBlNiBaTr11,VaCaSaKoVi99,KoMiEr09,KoErGoMi10,ChKoViCaVaFe07,MaKoDi09,KoViCh10,CaKoLaTiMiLeSnTr08,TrKoHjLiZhPeBoMcFrTrHu10,KoRiTr02,KoTr02a,KoTr02b,KoTr99}
-or \citet{KoVi98}. Additional aspects of adjoint simulations are
-described in \citet{TrTaLi05,LiTr06,TrKoLi08,LiTr08,TrKoHjLiZhPeBoMcFrTrHu10,PeKoLuMaLeCaLeMaLiBlNiBaTr11}.
-Domain decomposition is explained in detail in \citet{MaKoBlLe08},
-and excellent scaling up to 150,000 processor cores is shown for instance
-in \citet{CaKoLaTiMiLeSnTr08,KoLaMi08a,MaKoBlLe08,KoErGoMi10,Kom11},
-\item [{GPU computing}] ~\\
- Computing on GPU graphics cards for acoustic or seismic wave propagation
-applications is described in detail in \citet{Kom11,MiKo10,KoMiEr09,KoErGoMi10}.
-\end{description}
-\noindent If you use this new version, which has non blocking
-MPI for much better performance for medium or large runs, please cite
-at least one of these six articles, in which results of non blocking
-MPI runs are presented: \citet{PeKoLuMaLeCaLeMaLiBlNiBaTr11,KoErGoMi10,KoViCh10,Kom11,CaKoLaTiMiLeSnTr08,MaKoBlLe08}.\\
-
-\noindent If you work on geophysical applications, you may be interested
-in citing some of these application articles as well, among others:
-\begin{description}
-\item [{Southern California simulations}] ~\\
- \citet{KoLiTrSuStSh04,KrJiKoTr06a,KrJiKoTr06b}.
-
-If you use the 3D southern California model, please cite \citet{SuSh03}
-(Los Angeles model), \citet{lovelyetal06} (Salton Trough), and \citet{hauksson2000}
-(southern California). The Moho map was determined by \citet{zhu&kanamori2000}.
-The 1D SoCal model was developed by \citet{DrHe90}.
-
-\item [{Anisotropy}] ~\\
- \citet{ChTr07,JiTsKoTr05,ChFaKo04,FaChKo04,RiRiKoTrHe02,TrKo00}.
-\item [{Attenuation}] ~\\
- \citet{SaKoTr10,KoTr02a,KoTr99}.
-\item [{Topography}] ~\\
- \citet{LeKoHuTr09,LeChKoHuTr09,LeChLiKoHuTr08,GoAmTaCaSmSaMaKo09,WiKoScTr04}.
-\end{description}
-The corresponding Bib\TeX{} entries may be found in file \texttt{doc/USER\_MANUAL/bibliography.bib}.
-
-\section{Support}
-
-This material is based upon work supported by the USA National Science
-Foundation under Grants No. EAR-0406751 and EAR-0711177, by the French
-CNRS, French INRIA Sud-Ouest MAGIQUE-3D, French ANR NUMASIS under
-Grant No. ANR-05-CIGC-002, and European FP6 Marie Curie International
-Reintegration Grant No. MIRG-CT-2005-017461. Any opinions, findings,
-and conclusions or recommendations expressed in this material are
-those of the authors and do not necessarily reflect the views of the
-USA National Science Foundation, CNRS, INRIA, ANR or the European
-Marie Curie program.
-
-\chapter{\label{cha:Getting-Started}Getting Started}
-
-To download the SPECFEM3D\_Cartesian software package, type this:
-\begin{lyxcode}
-git clone --recursive --branch devel https://github.com/geodynamics/specfem3d.git
-\end{lyxcode}
-{\small \par}
-
-We recommend that you add {\texttt{ulimit -S -s unlimited}} to your
-{\texttt{.bash\_profile}} file and/or {\texttt{limit stacksize
-unlimited }} to your {\texttt{.cshrc}} file to suppress any potential
-limit to the size of the Unix stack.\\
-
-Then, to configure the software for your system, run the \texttt{configure}
-shell script. This script will attempt to guess the appropriate configuration
-values for your system. However, at a minimum, it is recommended that
-you explicitly specify the appropriate command names for your Fortran
-compiler and MPI package (another option is to define FC, CC and MPIF90
-in your .bash\_profile or your .cshrc file):
-\begin{lyxcode}
-./configure~FC=ifort~MPIFC=mpif90
-\end{lyxcode}
-Note that MPI must be installed with MPI-IO enabled because parts of SPECFEM3D perform I/Os through MPI-IO.
-
-Before running the \texttt{configure} script, you should probably
-edit file \texttt{flags.guess} to make sure that it contains the best
-compiler options for your system. Known issues or things to check
-are:
-\begin{description}
-\item [{\texttt{Intel ifort compiler}}] See if you need to add \texttt{-assume byterecl} for your machine.
-\textbf{In the case of that compiler, we have noticed that versions dot zero sometimes have bugs or issues that can lead to wrong results when running the code, thus we \emph{strongly} recommend using versions dot one or above (for instance version 13.1 instead of 13.0, version 14.1 instead of 14.0 and so on)}.
-\item [{\texttt{IBM compiler}}] See if you need to add \texttt{-qsave}
-or \texttt{-qnosave} for your machine.
-\item [{\texttt{Mac OS}}] You will probably need to install \texttt{XCODE}.
-\end{description}
-When compiling on an IBM machine with the \texttt{xlf} and \texttt{xlc}
-compilers, we suggest running the \texttt{configure} script with the
-following options:\\
-
-\noindent \texttt{./configure FC=xlf90\_r MPIFC=mpif90 CC=xlc\_r CFLAGS=\textquotedbl{}-O3
--q64\textquotedbl{} FCFLAGS=\textquotedbl{}-O3 -q64\textquotedbl{}
--with-scotch-dir=...}.\\
-
-On SGI systems, \texttt{flags.guess} automatically informs \texttt{configure}
-to insert `\texttt{`TRAP\_FPE=OFF}'' into the generated \texttt{Makefile}
-in order to turn underflow trapping off.\\
-
-You can add \texttt{--enable-vectorization} to the configuration options to speed up the code in the fluid (acoustic) and elastic parts.
-This works fine if (and only if) your computer always allocates a contiguous memory block for each allocatable array;
-this is the case for most machines and most compilers, but not all. To disable this feature, use option \texttt{--disable-vectorization}.
-For more details see \href{https://github.com/geodynamics/specfem3d/issues/81}{github.com/geodynamics/specfem3d/issues/81} .
-To check if that option works fine on your machine, run the code with and without it for an acoustic/elastic model and make sure the seismograms are identical.\\
-
-Note that we use CUBIT (now called Trelis) to create meshes of hexahedra, but other packages
-can be used as well, for instance GiD from \url{http://gid.cimne.upc.es}
-or Gmsh from \url{http://geuz.org/gmsh} \citep{GeRe09}. Even mesh
-creation packages that generate tetrahedra, for instance TetGen from
-\url{http://tetgen.berlios.de}, can be used because each tetrahedron
-can then easily be decomposed into four hexahedra as shown in the
-picture of the TetGen logo at \url{http://tetgen.berlios.de/figs/Delaunay-Voronoi-3D.gif};
-while this approach does not generate hexahedra of optimal quality,
-it can ease mesh creation in some situations and it has been shown
-that the spectral-element method can very accurately handle distorted
-mesh elements \citep{OlSe11}.
-
-The SPECFEM3D Cartesian software package relies on the SCOTCH library
-to partition meshes created with CUBIT. METIS \citep{KaKu98a,KaKu98c,KaKu98b}
-can also be used instead of SCOTCH if you prefer, by editing file
-src/decompose\_mesh/decompose\_mesh.F90 and uncommenting the flag
-USE\_METIS\_INSTEAD\_OF\_SCOTCH. You will also then need to install
-and compile Metis version 4.0 (do {*}NOT{*} install Metis version
-5.0, which has incompatible function calls) and edit \texttt{src/decompose\_mesh/Makefile.in}
-and uncomment the METIS link flag in that file before running \texttt{configure}.\\
-
-The SCOTCH library \citep{PeRo96} provides efficent static mapping,
-graph and mesh partitioning routines. SCOTCH is a free software package
-developed by Fran\c{c}ois Pellegrini et al. from LaBRI and INRIA in Bordeaux,
-France, downloadable from the web page \url{https://gforge.inria.fr/projects/scotch/}.
-In case no SCOTCH libraries can be found on the system, the configuration
-will bundle the version provided with the source code for compilation.
-The path to an existing SCOTCH installation can to be set explicitly
-with the option \texttt{-{}-with-scotch-dir}. Just as an example:
-\begin{lyxcode}
-./configure~FC=ifort~MPIFC=mpif90~-{}-with-scotch-dir=/opt/scotch
-\end{lyxcode}
-If you use the Intel ifort compiler to compile the code, we recommend
-that you use the Intel icc C compiler to compile Scotch, i.e., use:
-\begin{lyxcode}
-./configure~CC=icc~FC=ifort~MPIFC=mpif90
-\end{lyxcode}
-
-When compiling the SCOTCH source code, if you get a message such as: "ld: cannot find -lz",
-the Zlib compression development library is probably missing on your machine and you will need to install it or ask your system administrator to
-do so. On Linux machines the package is often called "zlib1g-dev" or similar. (thus "sudo apt-get install zlib1g-dev" would install it)
-
-To compile a serial version of the code for small meshes that fits
-on one compute node and can therefore be run serially, run \texttt{configure}
-with the \texttt{-{}-without-mpi} option to suppress all calls to
-MPI.
-
-A summary of the most important configuration variables follows.
-\begin{description}
-\item [{\texttt{F90}}] Path to the Fortran compiler.
-\item [{\texttt{MPIF90}}] Path to MPI Fortran.
-\item [{\texttt{MPI\_FLAGS}}] Some systems require this flag to link to
-MPI libraries.
-\item [{\texttt{FLAGS\_CHECK}}] Compiler flags.
-\end{description}
-The configuration script automatically creates for each executable
-a corresponding \texttt{Makefile} in the \texttt{src/} subdirectoy.
-The \texttt{Makefile} contains a number of suggested entries for various
-compilers, e.g., Portland, Intel, Absoft, NAG, and Lahey. The software
-has run on a wide variety of compute platforms, e.g., various PC clusters
-and machines from Sun, SGI, IBM, Compaq, and NEC. Select the compiler
-you wish to use on your system and choose the related optimization
-flags. Note that the default flags in the \texttt{Makefile} are undoubtedly
-not optimal for your system, so we encourage you to experiment with
-these flags and to solicit advice from your systems administrator.
-Selecting the right compiler and optimization flags can make a tremendous
-difference in terms of performance. We welcome feedback on your experience
-with various compilers and flags.
-
-Now that you have set the compiler information, you need to select
-a number of flags in the \texttt{constants.h} file depending on your
-system:
-\begin{description}
-\item [{\texttt{LOCAL\_PATH\_IS\_ALSO\_GLOBAL}}] Set to \texttt{.false.}
-on most cluster applications. For reasons of speed, the (parallel)
-distributed database generator typically writes a (parallel) database
-for the solver on the local disks of the compute nodes. Some systems
-have no local disks, e.g., BlueGene or the Earth Simulator, and other
-systems have a fast parallel file system, in which case this flag
-should be set to \texttt{.true.}. Note that this flag is not used
-by the database generator or the solver; it is only used for some
-of the post-processing.
-\end{description}
-The package can run either in single or in double precision mode.
-The default is single precision because for almost all calculations
-performed using the spectral-element method using single precision
-is sufficient and gives the same results (i.e. the same seismograms);
-and the single precision code is faster and requires exactly half
-as much memory. Select your preference by selecting the appropriate
-setting in the \texttt{constants.h} file:
-\begin{description}
-\item [{\texttt{CUSTOM\_REAL}}] Set to \texttt{SIZE\_REAL} for single precision
-and \texttt{SIZE\_DOUBLE} for double precision.
-\end{description}
-In the \texttt{precision.h} file:
-\begin{description}
-\item [{\texttt{CUSTOM\_MPI\_TYPE}}] Set to \texttt{MPI\_REAL} for single
-precision and \texttt{MPI\_DOUBLE\_PRECISION} for double precision.
-\end{description}
-On many current processors (e.g., Intel, AMD, IBM Power), single precision
-calculations are significantly faster; the difference can typically
-be 10\% to 25\%. It is therefore better to use single precision. What
-you can do once for the physical problem you want to study is run
-the same calculation in single precision and in double precision on
-your system and compare the seismograms. If they are identical (and
-in most cases they will), you can select single precision for your
-future runs.\\
-
-If your compiler has problems with the \texttt{use mpi} statements that are used in the code, use the script called
-\texttt{replace\_use\_mpi\_with\_include\_mpif\_dot\_h.pl} in the root directory to replace all of them with \texttt{include 'mpif.h'} automatically.
-
-\section{Adding OpenMP support in addition to MPI}
-
-NOTE FROM JULY 2013: OpenMP support is maybe / probably not maintained any more.
-Thus the section below is maybe obsolete.
-
-OpenMP support can be enabled in addition to MPI. However, in many
-cases performance will not improve because our pure MPI implementation
-is already heavily optimized and thus the resulting code will in fact
-be slightly slower. A possible exception could be IBM BlueGene-type
-architectures.
-
-To enable OpenMP, uncomment the OpenMP compiler option in two lines
-in file \texttt{src/specfem3D/Makefile.in} (before running \texttt{configure})
-and also uncomment the \texttt{\#define USE\_OPENMP} statement in
-file\\
- \texttt{src/specfem3D/specfem3D.F90}.
-
-The DO-loop using OpenMP threads has a SCHEDULE property. The OMP\_SCHEDULE
-environment variable can set the scheduling policy of that DO-loop.
-Tests performed by Marcin Zielinski at SARA (The Netherlands) showed
-that often the best scheduling policy is DYNAMIC with the size of
-the chunk equal to the number of OpenMP threads, but most preferably
-being twice as the number of OpenMP threads (thus chunk size = 8 for
-4 OpenMP threads etc). If OMP\_SCHEDULE is not set or is empty, the
-DO-loop will assume generic scheduling policy, which will slow down
-the job quite a bit.
-
-\section{Compiling on an IBM BlueGene}
-
-\underline{More recent installation instruction for IBM BlueGene, from April 2013:}\\
-
-\noindent Edit file \texttt{flags.guess} and put this for \texttt{FLAGS\_CHECK}:\\
-
-\noindent \texttt{-g -qfullpath -O2 -qsave -qstrict -qtune=qp -qarch=qp
--qcache=auto -qhalt=w}\\
- \texttt{-qfree=f90 -qsuffix=f=f90 -qlanglvl=95pure -Q -Q+rank,swap\_all
--Wl,-relax}\\
-
-\noindent The most relevant are the -qarch and -qtune flags, otherwise
-if these flags are set to ``auto'' then they are wrongly assigned
-to the architecture of the frond-end node, which is different from
-that on the compute nodes. You will need to set these flags to the
-right architecture for your BlueGene compute nodes, which is not necessarily
-``qp''; ask your system administrator. On some machines if is necessary
-to use -O2 in these flags instead of -O3 due to a compiler bug of
-the XLF version installed. We thus suggest to first try -O3, and then
-if the code does not compile or does not run fine then switch back
-to -O2. The debug flags (-g, -qfullpath) do not influence performance
-but are useful to get at least some insights in case of problems.\\
-
-\noindent Before running \texttt{configure}, select the XL Fortran
-compiler by typing \texttt{module load bgq-xl/1.0} or \texttt{module
-load bgq-xl} (another, less efficient option is to load the GNU compilers
-using \texttt{module load bgq-gnu/4.4.6} or similar).\\
-
-\noindent Then, to configure the code, type this:\\
-
-\noindent \texttt{./configure FC=bgxlf90\_r MPIFC=mpixlf90\_r CC=bgxlc\_r
-LOCAL\_PATH\_IS\_ALSO\_GLOBAL=true}\\
-
-In order for the SCOTCH domain decomposer to compile, on some (but
-not all) Blue Gene systems you may need to run \texttt{configure}
-with \texttt{CC=gcc} instead of \texttt{CC=bgxlc\_r}.
-
-\noindent \underline{Older installation instruction for IBM BlueGene, from 2011:}\\
-
-\noindent To compile the code on an IBM BlueGene, Laurent L\'eger from
-IDRIS, France, suggests the following: compile the code with
-
-\noindent \texttt{ FLAGS\_CHECK=\textquotedbl{}-O3 -qsave -qstrict
--qtune=auto -qarch=450d -qcache=auto }~\\
- \texttt{ -qfree=f90 -qsuffix=f=f90 -g -qlanglvl=95pure -qhalt=w -Q
--Q+rank,swap\_all -Wl,-relax\textquotedbl{}}
-
-\noindent Option \textquotedbl{}-Wl,-relax\textquotedbl{} must be
-added on many (but not all) BlueGene systems to be able to link the
-binaries \texttt{xmeshfem3D} and \texttt{xspecfem3D} because the final
-link step is done by the GNU \texttt{ld} linker even if one uses \texttt{FC=bgxlf90\_r,
-MPIFC=mpixlf90\_r} and \texttt{CC=bgxlc\_r} to create all the object
-files. On the contrary, on some BlueGene systems that use the native
-AIX linker option \textquotedbl{}-Wl,-relax\textquotedbl{} can lead
-to problems and must be suppressed from \texttt{flags.guess}.
-
-\noindent Also, \texttt{AR=ar, ARFLAGS=cru} and \texttt{RANLIB=ranlib}
-are hardwired in all \texttt{Makefile.in} files by default, but to
-cross-compile on BlueGene/P one needs to change these values to \texttt{AR=bgar,
-ARFLAGS=cru} and \texttt{RANLIB=bgranlib}. Thus the easiest thing
-to do is to modify all \texttt{Makefile.in} files and the \texttt{configure}
-script to set them automatically by \texttt{configure}. One then just
-needs to pass the right commands to the \texttt{configure} script:
-
-\noindent \texttt{ ./configure -{}-prefix=/path/to/SPECFEM3DG\_SP
--{}-host=Babel -{}-build=BGP }~\\
- \texttt{ FC=bgxlf90\_r MPIFC=mpixlf90\_r CC=bgxlc\_r AR=bgar ARFLAGS=cru
-}~\\
- \texttt{ RANLIB=bgranlib LOCAL\_PATH\_IS\_ALSO\_GLOBAL=false}
-
-\noindent This trick can be useful for all hosts on which one needs
-to cross-compile.
-
-\noindent On BlueGene, one also needs to run the \texttt{xcreate\_header\_file}
-binary file manually rather than in the Makefile:
-
-\noindent \texttt{bgrun -np 1 -mode VN -exe ./bin/xcreate\_header\_file}
-
-\section{Visualizing the subroutine calling tree of the source code}
-
-\noindent
-Packages such as \texttt{Doxywizard} can be used to visualize the calling tree of the
-subroutines of the  source code. \texttt{Doxywizard} is
-a GUI front-end for configuring and running \texttt{Doxygen}.
-
-\bigskip
-
-\noindent To do your own call graphs, you can follow these simple steps below.
-
-\begin{enumerate}
-\setcounter{enumi}{-1}
-\item Install \texttt{Doxygen} \underline{and} \texttt{graphviz} (the two are usually in the package manager of classic Linux distribution).
-\item Run in the terminal : \texttt{doxygen -g}, which creates a \texttt{Doxyfile} that tells doxygen what you want it to do.
-\item Edit the Doxyfile. Two Doxyfile-type files have been already committed in the directory \\
-\noindent \texttt{specfem3d/doc/Call\_trees}:
-
-\begin{itemize}
-\item[\textbullet] \texttt{Doxyfile\_truncated\_call\_tree} will generate call graphs with maximum 3 or 4 levels of tree structure,
-\item[\textbullet] \texttt{Doxyfile\_complete\_call\_tree} will generate call graphs with complete tree structure.
-\end{itemize}
-
-The important entries in the Doxyfile are:
-
-\begin{description}
-\item [{\texttt{PROJECT\_NAME}}]
-\item [{\texttt{OPTIMIZE\_FOR\_FORTRAN}}] Set to YES
-\item [{\texttt{EXTRACT\_ALL}}] Set to YES
-\item [{\texttt{EXTRACT\_PRIVATE}}] Set to YES
-\item [{\texttt{EXTRACT\_STATIC}}] Set to YES
-\item [{\texttt{INPUT}}] From the directory \texttt{specfem3d/doc/Call\_trees}, it is \texttt{"../../src/"}
-\item [{\texttt{FILE\_PATTERNS}}] In SPECFEM case, it is \texttt{*.f90* *.F90* *.c* *.cu* *.h*}
-\item [{\texttt{HAVE\_DOT}}] Set to YES
-\item [{\texttt{CALL\_GRAPH}}] Set to YES
-\item [{\texttt{CALLER\_GRAPH}}] Set to YES
-\item [{\texttt{DOT\_PATH}}] The path where is located the dot program graphviz (if it is not in your \$PATH)
-\item [{\texttt{RECURSIVE}}] This tag can be used to turn specify whether or not subdirectories should be searched for input files as well. In the case of SPECFEM, set to YES.
-\item [{\texttt{EXCLUDE}}] Here, you can exclude:  \\
-\texttt{"../../src/specfem3D/older\_not\_maintained\_partial\_OpenMP\_port"}\\
-\texttt{"../../src/decompose\_mesh/scotch"}\\
-\texttt{"../../src/decompose\_mesh/scotch\_5.1.12b"}
-\item [{\texttt{DOT\_GRAPH\_MAX\_NODES}}] to set the maximum number of nodes that will be shown in the graph.
-If the number of nodes in a graph becomes larger than this value, doxygen will truncate the graph, which is visualized by representing a node as a red box.
-Minimum value: 0, maximum value: 10000, default value: 50.
-\item [{\texttt{MAX\_DOT\_GRAPH\_DEPTH}}] to set the maximum depth of the graphs generated by dot.
-A depth value of 3 means that only nodes reachable from the root by following a path via at most 3 edges will be shown.
-Using a depth of 0 means no depth restriction.
-Minimum value: 0, maximum value: 1000, default value: 0.
-
-\end{description}
-
-\item Run : \texttt{doxygen Doxyfile}, HTML and LaTeX files created by default in \texttt{html} and \texttt{latex} subdirectories.
-
-\item To see the call trees, you have to open the file \texttt{html/index.html} in your \underline{browser}.
-You will have many informations about each subroutines of SPECFEM (not only call graphs), you can click on every boxes / subroutines.
-It show you the call, and, the caller graph of each subroutine : the subroutines called by the concerned subroutine, and the previous subroutines who call this subroutine (the previous path), respectively.
-In the case of a truncated calling tree, the boxes with a red border indicates a node that has \underline{more} arrows than are shown (in other words: the graph is truncated with respect to this node).
-
-\end{enumerate}
-
-\medskip
-
-\noindent Finally, some useful links:
-
-\begin{itemize}
-\item[\textbullet] a good and short summary for the basic utilisation of Doxygen:
-
-\url{http://www.softeng.rl.ac.uk/blog/2010/jan/30/callgraph-fortran-doxygen/},
-
-\item[\textbullet] to configure the diagrams :
-
-\url{http://www.stack.nl/~dimitri/doxygen/manual/diagrams.html},
-
-\item[\textbullet] the complete alphabetical index of the tags in Doxyfile:
-
-\url{http://www.stack.nl/~dimitri/doxygen/manual/config.html},
-
-\item[\textbullet] more generally, the Doxygen manual:
-
-\url{http://www.stack.nl/~dimitri/doxygen/manual/index.html}.
-
-\end{itemize}
-
-\medskip
-
-\section{Using the ADIOS library for I/O}
-Regular POSIX I/O can be problematic when dealing with large simulations one large
-clusters (typically more than $10,000$ processes). SPECFEM3D use the ADIOS library~\cite{Liu2013}
-to deal transparently take advantage of advanced paralllel file system features. To enable
-ADIOS, the following steps should be done:
-\begin{enumerate}
-\item Install ADIOS (available from \url{https://www.olcf.ornl.gov/center-projects/adios/}).
-Make sure that your environment variables reference it.
-\item You may want to change ADIOS related values in the \texttt{constants.h.in} file.
-The default values probably suit most cases.
-\item Configure using the \texttt{--with-adios} flag.
-\end{enumerate}
-ADIOS is currently only usable for meshfem3D generated mesh (i.e. not for meshes generated
-with CUBTI). Additional control parameters are discussed in section~\ref{cha:Main-Parameter}.
-
-\section{Becoming a developer of the code, or making small modifications in the source code}
-
-If you want to develop new features in the code, and/or if you want to make small changes, improvements, or bug fixes, you are very welcome to contribute.
-To do so, i.e.
-to access the development branch of the source code with read/write access (in a safe way, no need to worry too much about breaking the package, there is a robot called BuildBot that is in charge of checking and validating all new contributions and changes), please visit this Web page: \url{https://github.com/geodynamics/specfem3d/wiki/Using-Hub}.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\chapter{\label{cha:Mesh-Generation}Mesh Generation}
-
-The first step in running a spectral-element simulation consists of
-constructing a high-quality mesh for the region under consideration.
-We provide two possibilities to do so: (1) relying on the external,
-hexahedral mesher CUBIT, or (2) using the provided, internal mesher
-\texttt{xmeshfem3D}. In the following, we explain these two approaches.
-
-\section{\label{cha:Running-the-Mesher-CUBIT}Meshing with \texttt{CUBIT}}
-
-CUBIT is a meshing tool suite for the creation of finite-element meshes
-for arbitrarily shaped models. It has been developed and maintained
-at Sandia National Laboratories and can be purchased for a small academic
-institutional fee at \url{http://cubit.sandia.gov}. Our experience
-showed that using CUBIT greatly facilitates and speeds up the generation
-and preparation of hexahedral, conforming meshes for a variety of
-geophysical models with increasing complexity.
-
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=0.6\textwidth]{figures/mount-cubit}
-\par\end{centering}
-
-\caption{Example of the graphical user interface of CUBIT. The hexahedral mesh
-shown in the main display consists of a hexahedral discretization
-of a single volume with topography.}
-
-\label{fig:mount.cubit}
-\end{figure}
-
-\noindent
-The basic steps in creating a load-balanced, partitioned mesh with
-CUBIT are:
-\begin{description}
-\item [{1.}] setting up a hexahedral mesh with CUBIT,
-\item [{2.}] exporting the CUBIT mesh into a SPECFEM3D Cartesian file format
-and
-\item [{3.}] partitioning the SPECFEM3D Cartesian mesh files for a chosen
-number of cores.
-\end{description}
-Examples are provided in the SPECFEM3D Cartesian package in the subdirectory
-\texttt{EXAMPLES/}. We strongly encourage you to contribute your own
-example to this package by contacting the CIG Computational Seismology
-Mailing List \urlwithparentheses{cig-seismo@geodynamics.org}.
-
-\subsection{Creating the Mesh with CUBIT}
-
-For the installation and handling of the CUBIT meshing tool suite,
-please refer to the CUBIT user manual and documentation. In order
-to give you a basic understanding of how to use CUBIT for our purposes,
-examples are provided in the SPECFEM3D Cartesian package in the subdirectory
-\texttt{EXAMPLES/}:
-\begin{description}
-\item [{\texttt{homogeneous\_halfspace}}] Creates a single block model
-and assigns elastic material parameters.
-\item [{\texttt{layered\_halfspace}}] Combines two different, elastic material
-volumes and creates a refinement layer between the two. This example
-can be compared for validation against the solutions provided in subdirectory
-~\\
- \texttt{VALIDATION\_3D\_SEM\_SIMPLER\_LAYER\_SOURCE\_DEPTH/}.
-\item [{\texttt{waterlayered\_halfspace}}] Combines an acoustic and elastic
-material volume as in a schematic marine survey example.
-\item [{\texttt{tomographic\_model}}] Creates a single block model whose
-material properties will have to be read in from a tomographic model
-file during the databases creation by \texttt{xgenerate\_databases}.
-\end{description}
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=0.45\textwidth]{figures/example-homogeneous}
-\includegraphics[width=0.45\textwidth]{figures/example-2layers} \\
- \includegraphics[width=0.45\textwidth]{figures/example-water} \includegraphics[width=0.45\textwidth]{figures/example-tomo}
-\par\end{centering}
-
-\caption{Screenshots of the CUBIT examples provided in subdirectory \texttt{EXAMPLES/}:
-homogeneous halfspace (top-left), layered halfspace (top-right), water
-layered halfspace (bottom-left) and tomographic model (bottom-right).}
-
-\label{fig:examples.cubit}
-\end{figure}
-
-In each example subdirectory you will find a \texttt{README} file,
-which explains in a step-by-step tutorial the workflow for the example.
-Please feel free to contribute your own example to this package by
-contacting the CIG Computational Seismology Mailing List \urlwithparentheses{cig-seismo@geodynamics.org}.
-
-In some cases, to re-create the meshes for the examples given, just type "claro ./create\_mesh.py" or similar
-from the command line ("claro" is the command to run CUBIT from the command line).
-
-IMPORTANT note: in order to correctly set up GEOCUBIT and run the examples, please read the file called "EXAMPLES/README";
-in particular, please make sure you correctly set up the Python paths as indicated in that file.
-
-\subsection{Exporting the Mesh with \texttt{run\_cubit2specfem3d.py} }
-
-\label{subsec:Exporting-the-Mesh}
-
-Once the geometric model volumes in CUBIT are meshed, you prepare
-the model for exportation with the definition of material blocks and
-boundary surfaces. Thus, prior to exporting the mesh, you need to
-define blocks specifying the materials and absorbing boundaries in
-CUBIT. This process could be done automatically using the script \texttt{run\_cubit2specfem3d.py}
-if the mesh meets some conditions or manually, following the block
-convention:
-\begin{description}
-\item [{material\_name}] Each material should have a specific block defined
-by a unique name. The name convention of the material is to start
-with either 'elastic' or 'acoustic'. It must be then followed by a
-unique identifier, e.g. 'elastic 1', 'elastic 2', etc. The additional
-attributes to the block define the material description.
-
-For an elastic material:
-\begin{description}
-\item [{material\_id}] An integer value which is unique for this material.
-\item [{Vp}] P-wave speed of the material (given in m/s).
-\item [{Vs}] S-wave speed of the material (given in m/s).
-\item [{rho}] density of the material (given in kg/m$^{3}$).
-\item [{Q}] quality factor to use in case of a simulation with attenuation
-turned on. It should be between 1 and 9000. In case no attenuation
-information is available, it can be set to zero. Please note that
-your Vp- and Vs-speeds are given for a reference frequency. To change
-this reference frequency, you change the value of \texttt{ATTENUATION\_f0\_REFERENCE}
-in the main constants file \texttt{constants.h} found in subdirectory
-\texttt{src/shared/}.
-\item [{anisotropic\_flag}] Flag describing the anisotropic model to use
-in case an anisotropic simulation should be conducted. See the file
-\texttt{model\_aniso.f90} in subdirectory \texttt{src/generate\_databases/}
-for an implementation of the anisotropic models. In case no anisotropy
-is available, it can be set to zero.
-\end{description}
-
-Note that this material block has to be defined using all the volumes
-which belong to this elastic material. For volumes belonging to another,
-different material, you will need to define a new material block.
-
-For an acoustic material:
-\begin{description}
-\item [{material\_id}] An integer value which is unique for this material.
-\item [{Vp}] P-wave speed of the material (given in m/s).
-\item [{0}] S-wave speed of the material is ignored.
-\item [{rho}] density of the material (given in kg/m$^{3}$).
-\end{description}
-\item [{face\_topo}] Block definition for the surface which defines the
-free surface (which can have topography). The name of this block must
-be 'face\_topo', the block has to be defined using all the surfaces
-which constitute the complete free surface of the model.
-\item [{face\_abs\_xmin}] Block definition for the faces on the absorbing
-boundaries, one block for each surface with x=Xmin.
-\item [{face\_abs\_xmax}] Block definition for the faces on the absorbing
-boundaries, one block for each surface with x=Xmax.
-\item [{face\_abs\_ymin}] Block definition for the faces on the absorbing
-boundaries, one block for each surface with y=Ymin.
-\item [{face\_abs\_ymax}] Block definition for the faces on the absorbing
-boundaries, one block for each surface with y=Ymax.
-\item [{face\_abs\_bottom}] Block definition for the faces on the absorbing
-boundaries, one block for each surface with z=bottom.
-\end{description}
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=0.45\textwidth]{figures/mount-surface} \includegraphics[width=0.45\textwidth]{figures/mount-abs}
-\par\end{centering}
-
-\caption{Example of the block definitions for the free surface 'face\_topo'
-(left) and the absorbing boundaries, defined in a single block 'face\_abs\_xmin'
-(right) in CUBIT.}
-
-\label{fig:mount.abs}
-\end{figure}
-
-Optionally, instead of specifying for each surface at the boundaries
-a single block like mentioned above, you can also specify a single
-block for all boundary surfaces and name it as one of the absorbing
-blocks above, e.g. 'face\_abs\_xmin'.
-
-After the block definitions are done, you export the mesh using the
-script \texttt{cubit2specfem3d.py} provided in each of the example
-directories (linked to the common script \texttt{CUBIT\_GEOCUBIT/cubit2specfem3d.py}).
-If the export was successful, you should find the following
-files in a subdirectory \texttt{MESH/}:
-\begin{description}
-\item [{absorbing\_cpml\_file}] \textbf{(only needed in case of C-PML absorbing
-conditions)} Contains on the first line the total number of C-PML
-spectral elements in the mesh, and then on the following line the
-list of all these C-PML elements with two numbers per line: first
-the spectral element number, and then a C-PML flag indicating to which
-C-PML layer(s) that element belongs, according to the following convention:
-
-\begin{itemize}
-\item Flag = 1 : element belongs to a X CPML layer only (either in Xmin
-or in Xmax),
-\item Flag = 2 : element belongs to a Y CPML layer only (either in Ymin
-or in Ymax),
-\item Flag = 3 : element belongs to a Z CPML layer only (either in Zmin
-or in Zmax),
-\item Flag = 4 : element belongs to a X CPML layer and also to a Y CPML
-layer,
-\item Flag = 5 : element belongs to a X CPML layer and also to a Z CPML
-layer,
-\item Flag = 6 : element belongs to a Y CPML layer and also to a Z CPML
-layer,
-\item Flag = 7 : element belongs to a X, to a Y and to a Z CPML layer, i.e.,
-it belongs to a CPML corner.
-\end{itemize}
-
-Note that it does not matter whether an element belongs to a Xmin
-or to a Xmax CPML, the flag is the same in both cases; the same is
-true for Ymin and Ymax, and also Zmin and Zmax.
-
-When you have an existing CUBIT (or similar) mesh stored in SPECFEM3D
-format, i.e., if you have existing \textquotedbl{}nodes\_coords\_file\textquotedbl{}
-and \textquotedbl{}mesh\_file\textquotedbl{} files but do not know
-how to assign CPML flags to them, we have created a small serial Fortran
-program that will do that automatically for you, i.e., which will
-create the \textquotedbl{}absorbing\_cpml\_file\textquotedbl{} for
-you. That program is utils/CPML/convert\_external\_layers\_of\_a\_given\_mesh\_to\_CPML\_layers.f90,
-and a small Makefile is provided in that directory (utils/CPML).
-
-IMPORTANT: it is your responsibility to make sure that in the input
-CUBIT (or similar) mesh that this code will read in SPECFEM3D format
-from files \textquotedbl{}nodes\_coords\_file\textquotedbl{} and \textquotedbl{}mesh\_file\textquotedbl{}
-you have created layers of elements that constitute a layer of constant
-thickness aligned with the coordinate grid axes (X, Y and/or Z), so
-that this code can assign CPML flags to them. This code does NOT check
-that (because it cannot, in any easy way). The mesh inside these CPML
-layers does not need to be structured nor regular, any non-structured
-mesh is fine as long as it has flat PML inner and outer faces, parallel
-to the axes, and thus of a constant thickness. The thickness can be
-different for the X, Y and Z sides. But for X it must not vary, for
-Y it must not vary, and for Z it must not vary. If you do not know
-the exact thickness, you can use a slightly LARGER value in this code
-(say 2\% to 5\% more) and this code will fix that and will adjust
-it; never use a SMALLER value otherwise this code will miss some CPML
-elements.
-
-Note: in a future release we will remove the constraint of having
-CPML layers aligned with the coordinate axes; we will allow for meshes
-that are titled by any constant angle in the horizontal plane. However
-this is not implemented yet.
-
-Note: in the case of fluid-solid layers in contact, the fluid-solid
-interface currently needs to be flat and horizontal inside the CPML
-layer (i.e., bathymetry should become flat and horizontal when entering
-the CPML); this (small) constraint will probably remain in the code
-for a while because it makes fluid-solid matching inside the CPML
-much easier.
-
-\item [{materials\_file}] Contains the material associations for each element.
-The format is:
-
-\begin{lyxcode}
-element\_ID~material\_ID
-\end{lyxcode}
-
-where \texttt{element\_ID} is the element identifier and \texttt{material\_ID}
-is a unique identifier, positive (for materials taken from this list
-of materials, i.e. for which each spectral element has constant material
-properties taken from this list) or negative (for tomographic models,
-i.e. for spectral element whose real velocities and density will be
-assigned later by calling an external function to define model variations,
-for instance in the case of tomographic models; in such a case, material
-properties can vary inside each spectral element, i.e. be different
-at each of its Gauss-Lobatto-Legendre grid points).
-
-\item [{nummaterial\_velocity\_file}] Defines the material properties.
-
-\begin{itemize}
-\item For classical materials (i.e., spectral elements for which the velocity
-and density model will not be assigned by calling an external function
-to define for instance a tomographic model), the format is:
-
-\begin{lyxcode}
-domain\_ID~material\_ID~rho~vp~vs~Qkappa~Qmu~anisotropy\_flag
-\end{lyxcode}
-
-where \texttt{\textbf{domain\_ID}}\textbf{ is 1 for acoustic and 2
-for elastic or viscoelastic materials,} \texttt{material\_ID} a unique
-identifier, \texttt{rho} the density in $kg\, m^{-3}$, \texttt{vp}
-the P-wave speed in $m\, s^{-1}$, \texttt{vs} the S-wave speed in
-$m\, s^{-1}$, \texttt{Q} the quality factor and \texttt{anisotropy\_flag}
-an identifier for anisotropic models. Note that the \texttt{Qkappa}
-value is ignored by the code unless \texttt{FULL\_ATTENUATION\_SOLID}
-is set. Note also that both \texttt{Qkappa} and \texttt{Qmu} are ignored
-by the code unless \texttt{ATTENUATION} is set. If you want a model
-with no \texttt{Qmu} attenuation, both set \texttt{ATTENUATION} to
-\texttt{.false.} in the \texttt{Par\_file} and set \texttt{Qmu} to
-9999 here. If you want a model with no \texttt{Qkappa} attenuation,
-both set \texttt{FULL\_ATTENUATION\_SOLID} to \texttt{.false.} in
-the \texttt{Par\_file} and set \texttt{Qkappa} to 9999 here. %%magnoni
-
-\item For tomographic velocity models, please read Chapter \ref{cha:-Changing-the}
-and Section \ref{sec:Using-tomographic} `Using external tomographic
-Earth models' for further details.
-\end{itemize}
-\item [{nodes\_coords\_file}] Contains the point locations in Cartesian
-coordinates of the mesh element corners.
-\item [{mesh\_file}] Contains the mesh element connectivity. The hexahedral
-elements can have 8 or 27 nodes.\\
- See picture doc/mesh\_numbering\_convention/numbering\_convention\_27\_nodes.jpg
-to see\\
- in which (standard) order the points must be cited. In the case of
-8 nodes, just include the first 8 points.
-\item [{free\_or\_absorbing\_surface\_file\_zmax}] Contains the free surface
-connectivity or \\
- the surface connectivity of the absorbing boundary surface at the
-top (Zmax), \\
- depending on whether the top surface is defined as free or absorbing
-(\texttt{STACEY\_INSTEAD\_OF\_FREE\_SURFACE} in \texttt{DATA/Par\_file}). \\
- You should put both the surface of acoustic regions and of elastic
-regions in that file; that is, list all the element faces that constitute
-the surface of the model in that file.
-\item [{absorbing\_surface\_file\_xmax}] Contains the surface connectivity
-of the absorbing boundary surface at Xmax \\
-(also needed in the case of C-PML absorbing conditions, in order for
-the code to be able to impose Dirichlet conditions on their outer
-edge).
-\item [{absorbing\_surface\_file\_xmin}] Contains the surface connectivity
-of the absorbing boundary surface at Xmin \\
-(also needed in the case of C-PML absorbing conditions, in order for
-the code to be able to impose Dirichlet conditions on their outer
-edge).
-\item [{absorbing\_surface\_file\_ymax}] Contains the surface connectivity
-of the absorbing boundary surface at Ymax \\
-(also needed in the case of C-PML absorbing conditions, in order for
-the code to be able to impose Dirichlet conditions on their outer
-edge).
-\item [{absorbing\_surface\_file\_ymin}] Contains the surface connectivity
-of the absorbing boundary surface at Ymin \\
-(also needed in the case of C-PML absorbing conditions, in order for
-the code to be able to impose Dirichlet conditions on their outer
-edge).
-\item [{absorbing\_surface\_file\_bottom}] Contains the surface connectivity
-of the absorbing boundary surface at the bottom (Zmin) \\
-(also needed in the case of C-PML absorbing conditions, in order for
-the code to be able to impose Dirichlet conditions on their outer
-edge).
-\end{description}
-These mesh files are needed as input files for the partitioner \texttt{xdecompose\_mesh}
-to load-balance the mesh. Please see the next section for further
-details.
-
-In directory "CUBIT\_GEOCUBIT" we provide a script that can help doing the above tasks of exporting a CUBIT mesh
-to SPECFEM3D format automatically for you: "run\_cubit2specfem3d.py".
-Just edit them to indicate the path to your local installation of CUBIT and also the name of the *.cub existing CUBIT mesh file
-that you want to export to SPECFEM3D format. These scripts will do the conversion for you automatically except assigning material properties
-to the different mesh layers. To do so, you will then need to edit the file
-called "nummaterial\_velocity\_file" that will have just been created and change it from the prototype created:
-
-    0 1 vol1 --> syntax: \#material\_domain\_id \#material\_id \#rho \#vp \#vs \#Q\_kappa \#Q\_mu \#anisotropy
-    0 2 vol2 --> syntax: \#material\_domain\_id \#material\_id \#rho \#vp \#vs \#Q\_kappa \#Q\_mu \#anisotropy
-
-(where "vol1" and "vol2" here represent the volume labels that you have set while creating the mesh in CUBIT)
-to for instance
-
-    2 1 1500 2300 1800 9999.0 9999.0 0
-    2 2 1600 2500 20000 9999.0 9999.0 0
-
-\subsubsection*{Checking the mesh quality}
-
-The quality of the mesh may be inspected more precisely based upon
-the serial code in the file \texttt{check\_mesh\_quality\_}~\\
- \texttt{CUBIT\_Abaqus.f90} located in the directory \texttt{src/check\_mesh\_quality\_CUBIT\_Abaqus/}.
-Running this code is optional because no information needed by the
-solver is generated.
-
-Prior to running and compiling this code, you have to export your
-mesh in CUBIT to an ABAQUS (.inp) format. For example, export mesh
-block IDs belonging to volumes in order to check the quality of the
-hexahedral elements. You also have to determine a number of parameters
-of your mesh, such as the number of nodes and number of elements and
-modify the header of the \texttt{check\_mesh\_quality\_CUBIT\_Abaqus.f90}
-source file in directory \texttt{src/check\_mesh\_quality\_CUBIT\_Abaqus/}.
-
-\noindent
-Then, in the main directory, type
-\begin{lyxcode}
-{\small make~xcheck\_mesh\_quality\_CUBIT\_Abaqus}{\small \par}
-\end{lyxcode}
-and use
-\begin{lyxcode}
-{\small ./bin/xcheck\_mesh\_quality\_CUBIT\_Abaqus}{\small \par}
-\end{lyxcode}
-to generate an %not supported: AVS output file (\texttt{\small AVS\_meshquality.inp} in AVS UCD format) or
-OpenDX output file (\texttt{\small DX\_mesh\_quality.dx}{\small )
-that can be used to investigate mesh quality, e.g. skewness of elements,
-and a Gnuplot histogram (}\texttt{\small mesh\_quality\_histogram.txt}{\small )
-that can be plotted with gnuplot (type `}\texttt{\small gnuplot plot\_mesh\_quality\_histogram.gnu}{\small ').
-The histogram is also printed to the screen. Analyze that skewness
-histogram of mesh elements to make sure no element has a skewness
-above approximately 0.75, otherwise the mesh is of poor quality (and
-if even a single element has a skewness value above 0.80, then you
-must definitely improve the mesh). If you want to start designing
-your own meshes, this tool is useful for viewing your creations. You
-are striving for meshes with elements with `cube-like' dimensions,
-e.g., the mesh should contain no very elongated or skewed elements.}{\small \par}
-
-\subsection{Partitioning the Mesh with \texttt{xdecompose\_mesh}}
-
-The SPECFEM3D Cartesian software package performs large scale simulations
-in a parallel 'Single Process Multiple Data' way. The spectral-element
-mesh created with CUBIT needs to be distributed on the processors.
-This partitioning is executed once and for all prior to the execution
-of the solver so it is referred to as a static mapping.
-
-An efficient partitioning is important because it leverages the overall
-running time of the application. It amounts to balance the number
-of elements in each slice while minimizing the communication costs
-resulting from the placement of adjacent elements on different processors.
-\texttt{decompose\_mesh} depends on the SCOTCH library \citep{PeRo96},
-which provides efficent static mapping, graph and mesh partitioning
-routines. SCOTCH is a free software package developed by Fran\c{c}ois
-Pellegrini et al. from LaBRI and INRIA in Bordeaux, France, downloadable
-from the web page \url{https://gforge.inria.fr/projects/scotch/}.
-\\
-
-In most cases, the configuration with \texttt{./configure~FC=ifort}
-should be sufficient. During the configuration process, the script
-tries to find existing SCOTCH installations. In case your system has
-no pre-existing SCOTCH installation, we provide the source code of
-SCOTCH, which is released open source under the French CeCILL-C version
-1 license, in directory \texttt{src/decompose\_mesh/scotch\_5.1.12b}.
-This version gets bundled with the compilation of the SPECFEM3D Cartesian
-package if no libraries could have been found. If this automatic compilation
-of the SCOTCH libraries fails, please refer to file INSTALL.txt in
-that directory to see further details how to compile it on your system.
-In case you want to use a pre-existing installation, make sure you
-have correctly specified the path of the SCOTCH library when using
-the option \texttt{-{}-with-scotch-dir} with the \texttt{./configure}
-script. In the future you should be able to find more recent versions
-at \url{http://www.labri.fr/perso/pelegrin/scotch/scotch_en.html}
-. \\
-
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=0.45\textwidth]{figures/mount-partitions}
-\includegraphics[width=0.45\textwidth]{figures/mount-partitions2}
-\par\end{centering}
-
-\caption{Example of a mesh partitioning onto four cores. Each single core partition
-is colored differently. The executable \texttt{xdecompose\_mesh} can
-equally distribute the mesh on any arbitrary number of cores. Domain
-decomposition is explained in detail in \citet{MaKoBlLe08}, and excellent
-scaling up to 150,000 processor cores in shown for instance in \citet{CaKoLaTiMiLeSnTr08,KoLaMi08a,MaKoBlLe08,KoErGoMi10,Kom11}.}
-
-\label{fig:mount.partitions}
-\end{figure}
-
-When you are ready to compile, in the main directory type
-\begin{lyxcode}
-{\small make~xdecompose\_mesh}{\small \par}
-\end{lyxcode}
-If all paths and flags have been set correctly, the executable \texttt{bin/xdecompose\_mesh}
-should be produced.
-
-The partitioning is done in serial for now (in the next release we
-will provide a parallel version of that code). It needs to be run in the main directory because it expects the \texttt{./DATA/Par\_file}. The synopsis is:
-\begin{lyxcode}
-./bin/xdecompose\_mesh~nparts~input\_directory~output\_directory
-\end{lyxcode}
-where
-\begin{itemize}
-\item \texttt{nparts} is the number of partitions, i.e., the number of cores
-for the parallel simulations,
-\item \texttt{input\_directory} is the directory which holds all the files
-generated by the Python script \texttt{cubit2specfem3d.py} explained
-in the previous Section~\ref{subsec:Exporting-the-Mesh}, e.g. \texttt{./MESH/},
-and
-\item \texttt{output\_directory} is the directory for the output of this
-partitioner which stores ACII-format files named like \texttt{proc{*}{*}{*}{*}{*}{*}\_Database}
-for each partition. These files will be needed for creating the distributed
-databases, and have to reside in the directory \texttt{LOCAL\_PATH}
-specified in the main \texttt{Par\_file}, e.g. in directory \texttt{./OUTPUT\_FILES/DATABASES\_MPI}.
-Please see Chapter~\ref{cha:Creating-Distributed-Databases} for
-further details.
-\end{itemize}
-Note that all the files generated by the Python script \texttt{cubit2specfem3d.py}
-must be placed in the \texttt{input\_directory} folder before running
-the program.
-
-\section{\label{cha:Running-the-Mesher-Meshfem3D}Meshing with \texttt{xmeshfem3D}}
-
-In case you successfully ran the configuration script, you are also
-ready to compile the internal mesher. This is an alternative to CUBIT
-for the mesh generation of relatively simple models.
-
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[scale=0.5]{figures/socal_map_mpi}
-\par\end{centering}
-
-\caption{\label{fig:For-parallel-computing}For parallel computing purposes,
-the model block is subdivided in $\nprocxi\times\nproceta$ slices
-of elements. In this example we use $5^{2}=25$ processors. }
-\end{figure}
-
-\noindent
-In the main directory, type
-\begin{lyxcode}
-{\small make~xmeshfem3D}{\small \par}
-\end{lyxcode}
-If all paths and flags have been set correctly, the mesher should
-now compile and produce the executable \texttt{bin/xmeshfem3D}. Please
-note that \texttt{xmeshfem3D} must be called directly from the main
-directory, as most of the binaries of the package.
-
-Input for the mesh generation program is provided through the parameter
-file \texttt{Mesh\_Par\_file}, which resides in the subdirectory \texttt{DATA/meshfem3D\_files/}.
-(to see how to use it, see the EXAMPLES specific to the internal mesher in directory \texttt{EXAMPLES/meshfem3D\_examples/}.
-Before running the mesher, a number of parameters need to be set in
-the \texttt{Mesh\_Par\_file}. This requires a basic understanding
-of how the SEM is implemented, and we encourage you to read \citet{KoVi98,KoTr99}
-and \citet{KoLiTrSuStSh04}.
-
-The mesher and the solver use UTM coordinates internally, therefore
-you need to define the zone number for the UTM projection (e.g., zone
-11 for Los Angeles). Use decimal values for latitude and longitude
-(no minutes/seconds). These values are approximate; the mesher will
-round them off to define a square mesh in UTM coordinates. When running
-benchmarks on rectangular models, turn the UTM projection off by using
-the flag \texttt{\small SUPPRESS\_UTM\_PROJECTION}{\small , in which
-case all `longitude' parameters simply refer to the $x$~axis, and
-all `latitude' parameters simply refer to the $y$~axis. To run the
-mesher for a global simulation, the following parameters need to be
-set in the }\texttt{\small Mesh\_Par\_file}{\small :}{\small \par}
-\begin{description}
-\item [{\texttt{LATITUDE\_MIN}}] Minimum latitude in the block (negative
-for South).
-\item [{\texttt{LATITUDE\_MAX}}] Maximum latitude in the block.
-\item [{\texttt{LONGITUDE\_MIN}}] Minimum longitude in the block (negative
-for West).
-\item [{\texttt{LONGITUDE\_MAX}}] Maximum longitude in the block.
-\item [{\texttt{DEPTH\_BLOCK\_KM}}] Depth of bottom of mesh in kilometers.
-\item [{\texttt{\noun{UTM\_PROJECTION\_ZONE}}}] UTM projection zone in
-which your model resides, only valid when \texttt{SUPPRESS\_UTM\_}~\\
- \texttt{PROJECTION} is \texttt{.false.}.
-Use a negative zone number for the Southern hemisphere:
-the Northern hemisphere corresponds to zones +1 to +60,
-the Southern hemisphere to zones -1 to -60.
-We use the WGS84 (World Geodetic System 1984) reference ellipsoid for the UTM projection. If you prefer to use the Clarke 1866 ellipsoid,
-edit file \texttt{src/shared/utm\_geo.f90}, uncomment that ellipsoid and recompile the code.
-From http://en.wikipedia.org/wiki/Universal\_Transverse\_Mercator\_coordinate\_system :
-The Universal Transverse Mercator coordinate system was developed by the United States Army Corps of Engineers in the 1940s.
-The system was based on an ellipsoidal model of Earth. For areas within the contiguous United States
-the Clarke Ellipsoid of 1866 was used. For the remaining areas of Earth, including Hawaii, the International Ellipsoid was used.
-The WGS84 ellipsoid is now generally used to model the Earth in the UTM coordinate system,
-which means that current UTM northing at a given point can be 200+ meters different from the old one.
-For different geographic regions, other datum systems (e.g.: ED50, NAD83) can be used.
-\item [{\texttt{SUPPRESS\_UTM\_PROJECTION}}] set to be \texttt{.false.}
-when your model range is specified in geographical coordinates, and
-needs to be \texttt{.true.} when your model is specified in Cartesian
-coordinates. \noun{UTM projection zone in which your simulation region
-resides.}
-\item [{\texttt{INTERFACES\_FILE }}] File which contains the description
-of the topography and of the interfaces between the different layers
-of the model, if any. The number of spectral elements in the vertical
-direction within each layer is also defined in this file.
-\item [{$\nexxi$}] The number of spectral elements along one side of the
-block. This number \textit{must} be 8~$\times$~a multiple of $\nprocxi$
-defined below. Based upon benchmarks against semi-analytical discrete
-wavenumber synthetic seismograms \citep{KoLiTrSuStSh04}, determined
-that a $\nexxi=288$ run is accurate to a shortest period of roughly
-2~s. Therefore, since accuracy is determined by the number of grid
-points per shortest wavelength, for any particular value of $\nexxi$
-the simulation will be accurate to a shortest period determined by
-\begin{equation}
-\mbox{shortest period (s)}=(288/\nexxi)\times2.\label{eq:shortest_period}
-\end{equation}
- The number of grid points in each orthogonal direction of the reference
-element, i.e., the number of Gauss-Lobatto-Legendre points, is determined
-by \texttt{NGLLX} in the \texttt{constants.h} file. We generally use
-$\mbox{\texttt{NGLLX\/}}=5$, for a total of $5^{3}=125$ points per
-elements. We suggest not to change this value.
-\item [{$\nexeta$}] The number of spectral elements along the other side
-of the block. This number \textit{must} be 8~$\times$~a multiple
-of $\nproceta$ defined below.
-\item [{$\nprocxi$}] The number of processors or slices along one side
-of the block (see Figure~\ref{fig:For-parallel-computing}); we must
-have $\nexxi=8\times c\times\nprocxi$, where $c\ge1$ is a positive
-integer.
-\item [{$\nproceta$}] The number of processors or slices along the other
-side of the block; we must have $\nexeta=8\times c\times\nproceta$,
-where $c\ge1$ is a positive integer.
-\item [{\texttt{USE\_REGULAR\_MESH}}] set to be \texttt{.true.} if you
-want a perfectly regular mesh or \texttt{.false.} if you want to add
-doubling horizontal layers to coarsen the mesh. In this case, you
-also need to provide additional information by setting up the next
-three parameters.
-\item [{\texttt{NDOUBLINGS}}] The number of horizontal doubling layers.
-Must be set to \texttt{1} or \texttt{2} if \texttt{USE\_REGULAR\_MESH}
-is set to \texttt{.true.}.
-\item [{\texttt{NZ\_DOUBLING\_1}}] The position of the first doubling layer
-(only interpreted if \texttt{USE\_REGULAR\_MESH} is set to \texttt{.true.}).
-\item [{\texttt{NZ\_DOUBLING\_2}}] The position of the second doubling
-layer (only interpreted if \texttt{USE\_REGULAR\_MESH} is set to \texttt{.true.}
-and if \texttt{NDOUBLINGS} is set to \texttt{2}).
-\item [{\texttt{CREATE\_ABAQUS\_FILES}}] Set this flag to \texttt{.true.}
-to save Abaqus FEA \urlwithparentheses{www.simulia.com} mesh files
-for subsequent viewing. Turning the flag on generates files in the
-\texttt{LOCAL\_PATH} directory. See Section~\ref{sec:Mesh-graphics}
-for a discussion of mesh viewing features.
-\item [{\texttt{CREATE\_DX\_FILES}}] Set this flag to \texttt{.true.} to
-save OpenDX \urlwithparentheses{www.opendx.org} mesh files for subsequent
-viewing.
-\item [{\texttt{LOCAL\_PATH}}] Directory in which the partitions generated
-by the mesher will be written. Generally one uses a directory on the
-local disk of the compute nodes, although on some machines these partitions
-are written on a parallel (global) file system (see also the earlier
-discussion of the \texttt{LOCAL\_PATH\_IS\_ALSO\_GLOBAL} flag in Chapter~\ref{cha:Getting-Started}).
-The mesher generates the necessary partitions in parallel, one set
-for each of the $\nprocxi\times\nproceta$ slices that constitutes
-the mesh (see Figure~\ref{fig:For-parallel-computing}). After the
-mesher finishes, you can log in to one of the compute nodes and view
-the contents of the \texttt{LOCAL\_PATH} directory to see the files
-generated by the mesher. These files will be needed for creating the
-distributed databases, and have to reside in the directory \texttt{LOCAL\_PATH}
-specified in the main \texttt{Par\_file}, e.g. in directory \texttt{OUTPUT\_FILES/DATABASES\_MPI}.
-Please see Chapter~\ref{cha:Creating-Distributed-Databases} for
-further details.
-\item [{\texttt{NMATERIALS}}] The number of different materials in your
-model. In the following lines, each material needs to be defined as
-:
-
-\begin{lyxcode}
-material\_ID~rho~vp~vs~Q~anisotropy\_flag~domain\_ID
-\end{lyxcode}
-
-where
-\begin{itemize}
-\item \texttt{Q} : quality factor (0=no attenuation)
-\item \texttt{anisotropy\_flag} : 0=no anisotropy / 1,2,.. check with implementation
-in \texttt{aniso\_model.f90}
-\item \texttt{domain\_id} : 1=acoustic / 2=elastic
-\end{itemize}
-\item [{\texttt{NREGIONS}}] The number of regions in the mesh. In the following
-lines, because the mesh is regular or 'almost regular', each region
-is defined as :
-
-\begin{lyxcode}
-NEX\_XI\_BEGIN~NEX\_XI\_END~NEX\_ETA\_BEGIN~NEX\_ETA\_END~NZ\_BEGIN~NZ\_END~material\_ID
-\end{lyxcode}
-\end{description}
-The \texttt{INTERFACES\_FILE} parameter of \texttt{Mesh\_Par\_File}
-defines the file which contains the settings of the topography grid
-and of the interfaces grids. Topography is defined as a set of elevation
-values on a regular 2D grid. It is also possible to define interfaces
-between the layers of the model in the same way. The file needs to
-define several parameters:
-\begin{itemize}
-\item The number of interfaces, including the topography. This needs to
-be set at the first line. Then, from the bottom to the top of the
-model, you need to define the grids with:
-\item \texttt{SUPPRESS\_UTM\_PROJECTION} flag as described previously,
-\item number of points along $x$ and $y$ direction (NXI and NETA),
-\item minimal $x$ and $y$ coordinates (LONG\_MIN and LAT\_MIN),
-\item spacing between points along $x$ and $y$ (SPACING\_XI and SPACING\_ETA)
-and
-\item the name of the file which contains the elevation values (in $y$.$x$
-increasing order).
-\end{itemize}
-At the end of this file, you simply need to set the number of spectral
-elements in the vertical direction for each layer. We provide a few
-models in the {\texttt{EXAMPLES/}} directory. \\
-
-Finally, depending on your system, you might need to provide a file
-that tells MPI what compute nodes to use for the simulations. The
-file must have a number of entries (one entry per line) at least equal
-to the number of processors needed for the run. A sample file is provided
-in the file \texttt{mymachines}. This file is not used by the mesher
-or solver, but is required by the \texttt{go\_mesher} and \texttt{go\_solver}
-default job submission scripts. See Chapter~\ref{cha:Scheduler}
-for information about running the code on a system with a scheduler,
-e.g., LSF.
-
-Now that you have set the appropriate parameters in the \texttt{Mesh\_Par\_file}
-and have compiled the mesher, you are ready to launch it! This is
-most easily accomplished based upon the \texttt{go\_mesher} script.
-When you run on a PC cluster, the script assumes that the nodes are
-named n001, n002, etc. If this is not the case, change the \texttt{tr
--d `n'} line in the script. You may also need to edit the last command
-at the end of the script that invokes the \texttt{mpirun} command.
-See Chapter~\ref{cha:Scheduler} for information about running the
-code on a system with a scheduler, e.g., LSF.
-
-Mesher output is provided in the \texttt{OUTPUT\_FILES} directory
-in \texttt{output\_mesher.txt}; this file provides lots of details
-about the mesh that was generated. Please note that the mesher suggests
-a time step \texttt{DT} to run the solver with. The mesher output
-file also contains a table about the quality of the mesh to indicate
-possible problems with the distortions of elements. Alternatively,
-output can be directed to the screen instead by uncommenting a line
-in \texttt{constants.h}:
-\begin{lyxcode}
-!~uncomment~this~to~write~messages~to~the~screen~
-
-!~integer,~parameter~::~IMAIN~=~ISTANDARD\_OUTPUT
-\end{lyxcode}
-To control the quality of the mesh, check the standard output (either
-on the screen or in the \texttt{OUTPUT\_FILES} directory in \texttt{output\_mesher.txt})
-and analyze the skewness histogram of mesh elements to make sure no
-element has a skewness above approximately 0.75, otherwise the mesh
-is of poor quality (and if even a single element has a skewness value
-above 0.80, then you must definitely improve the mesh). To draw the
-skewness histogram on the screen, type \texttt{gnuplot plot\_mesh\_quality\_histogram.gnu}.
-
-\chapter{\label{cha:Creating-Distributed-Databases}Creating the Distributed
-Databases}
-
-After using \texttt{xmeshfem3D} or \texttt{xdecompose\_mesh}, the
-next step in the workflow is to compile \texttt{xgenerate\_}~\\
- \texttt{databases}. This program is going to create all the missing
-information needed by the SEM solver.
-
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=0.6\textwidth]{figures/workflow}
-\par\end{centering}
-
-\caption{Schematic workflow for a SPECFEM3D Cartesian simulation. The executable
-\texttt{xgenerate\_databases} creates the GLL mesh points and assigns
-specific model parameters.}
-
-\label{fig:workflow.databases}
-\end{figure}
-
-\noindent
-In the main directory, type
-\begin{lyxcode}
-{\small make~xgenerate\_databases}{\small \par}
-\end{lyxcode}
-Input for the program is provided through the main parameter file
-\texttt{Par\_file}, which resides in the subdirectory \texttt{DATA}.
-Please note that \texttt{xgenerate\_databases} must be called directly
-from the main directory, as most of the binaries of the package.
-
-\section{\label{cha:Main-Parameter}Main parameter file \texttt{Par\_file}}
-
-Before running \texttt{xgenerate\_databases}, a number of parameters
-need to be set in the main parameter \texttt{Par\_file} located in
-the subdirectory \texttt{DATA}:
-\begin{description}
-\item [{\texttt{SIMULATION\_TYPE}}] is set to 1 for forward simulations,
-2 for adjoint simulations (see Section \ref{sec:Adjoint-simulation-finite})
-and 3 for kernel simulations (see Section \ref{sec:Finite-Frequency-Kernels}).
-\item [{\texttt{SAVE\_FORWARD}}] is only set to \texttt{.true.} for a forward
-simulation with the last frame of the simulation saved, as part of
-the finite-frequency kernel calculations (see Section \ref{sec:Finite-Frequency-Kernels}).
-For a regular forward simulation, leave \texttt{SIMULATION\_TYPE}
-and \texttt{SAVE\_FORWARD} at their default values.
-\item [{\texttt{\noun{UTM\_PROJECTION\_ZONE}}}] UTM projection zone in
-which your model resides, only valid when \texttt{SUPPRESS\_UTM\_}~\\
- \texttt{PROJECTION} is \texttt{.false.}.
-\item [{\texttt{SUPPRESS\_UTM\_PROJECTION}}] set to be \texttt{.false.}
-when your model range is specified in the geographical coordinates,
-and needs to be \texttt{.true.} when your model is specified in a
-cartesian coordinates. \noun{UTM projection zone in which your simulation
-region resides.}
-\item [{\texttt{NPROC}}] The number of MPI processors, each one is assigned
-one slice of the whole mesh.
-\item [{\texttt{NSTEP}}] The number of time steps of the simulation. This
-controls the length of the numerical simulation, i.e., twice the number
-of time steps requires twice as much CPU time. This feature is not
-used at the time of generating the distributed databases but is required
-for the solver, i.e., you may change this parameter after running
-\texttt{xgenerate\_databases}.
-\item [{\texttt{DT}}] The length of each time step in seconds. This feature
-is not used at the time of generating the distributed databases but
-is required for the solver. Please see also Section~\ref{sec:Choosing-the-Time-Step}
-for further details.
-\item [{\texttt{NGNOD}}] The number of nodes for 2D and 3D shape functions
-for hexahedra. We use either 8-node mesh elements (bricks) or 27-node
-elements. If you use the internal mesher, the only option is 8-node
-bricks (27-node elements are not supported). \texttt{CUBIT} does not
-support HEX27 elements either (it can generate them, but they are
-flat, i.e. identical to HEX8). To generate HEX27 elements with curvature
-properly taken into account, you can use \texttt{Gmsh} \url{http://geuz.org/gmsh/}
-\item [{\texttt{MODEL}}] Must be set to one of the following:
-
-\begin{description}
-\item [{\textmd{Models defined by mesh parameters:}}] ~
-
-\begin{description}
-\item [{\texttt{default}}] Uses model parameters as defined by meshing
-procedures described in the previous Chapter~\ref{cha:Mesh-Generation}.
-\end{description}
-
-\item [{\textmd{1D~models~with~real~structure:}}] ~
-
-\begin{description}
-\item [{\texttt{1D\_prem}}] Isotropic version of the spherically symmetric
-Preliminary Reference Earth Model (PREM) \citep{DzAn81}.
-\item [{\texttt{1D\_socal}}] A standard isotropic 1D model for Southern
-California.
-\item [{\texttt{1D\_cascadia}}] Isotropic 1D profile for the Cascadia region.
-\end{description}
-
-\item [{\textmd{Fully~3D~models:}}] ~
-
-\begin{description}
-\item [{\texttt{aniso}}] For a user-specified fully anisotropic model.
-Parameters are set up in routines located in file \texttt{model\_aniso.f90}
-in directory \texttt{src/generate\_databases/}. See Chapter~\ref{cha:-Changing-the}
-for a discussion on how to specify your own 3D model.
-
-\item [{\texttt{external}}] For a user-specified isotropic model which
-uses externally defined model parameters. Uses external model definitions
-set up in routines located in file \texttt{model\_external\_values.f90}
-in directory \texttt{src/generate\_databases/}. Please modify these
-generic template routines to use your own model definitions.
-\item [{\texttt{gll}}] For a user-specified isotropic model which uses
-external binary files for $v_{p}$, $v_{s}$ and $\rho$. Binary files
-are given in the same format as when outputted by the \texttt{xgenerate\_databases}
-executable when using option \texttt{SAVE\_MESH\_FILES}. These binary
-files define the model parameters on all GLL points which can be used
-for iterative inversion procedures.
-\item [{\texttt{salton\_trough}}] A 3D $V_{p}$ model for Southern California.
-Users must provide the corresponding data file \texttt{regrid3\_vel\_p.bin}
-in directory \texttt{DATA/st\_3D\_block\_harvard/}.
-\item [{\texttt{tomo}}] For a user-specified 3D isotropic model which uses
-a tomographic model file \texttt{tomographic\_model.xyz} in directory
-\texttt{DATA}. See Section \ref{sec:Using-tomographic}, for a discussion
-on how to specify your own 3D tomographic model.
-\end{description}
-\end{description}
-\item [{\texttt{APPROXIMATE\_OCEAN\_LOAD}}] Set to \texttt{.true.} if the
-effect of the oceans on seismic wave propagation should be incorporated
-based upon the (rough) approximate treatment discussed in \citet{KoTr02b}.
-This feature is inexpensive from a numerical perspective, both in
-terms of memory requirements and CPU time. This approximation is accurate
-at periods of roughly 20~s and longer. At shorter periods the effect
-of water phases/reverberations is not taken into account, even when
-the flag is on. If you want to model the effect of a fluid-solid model
-at short periods, then set this flag to \texttt{.false.} and mesh
-the fluid layer explicitly in your mesher, so that it is computed
-accurately and without this approximation.
-\item [{\texttt{TOPOGRAPHY}}] This feature is only effective if \texttt{APPROXIMATE\_OCEAN\_LOAD}
-is set to \texttt{.true.}. Set to \texttt{.true.} if topography and
-bathymetry should be read in based upon the topography file specified
-in the main constants file \texttt{constants.h} found in subdirectory
-\texttt{src/shared/} to evaluate elevations. If not set, elevations
-will be read from the numerical mesh.
-\item [{\texttt{ATTENUATION}}] Set to \texttt{.true.} if attenuation should
-be incorporated. Turning this feature on increases the memory requirements
-significantly (roughly by a factor of~1.5), and is numerically fairly
-expensive. See \citet{KoTr99,KoTr02a} for a discussion on the implementation
-of attenuation based upon standard linear solids. Please note that
-the Vp- and Vs-velocities of your model are given for a reference
-frequency. To change this reference frequency, you change the value
-of \texttt{ATTENUATION\_f0\_REFERENCE} in the main constants file
-\texttt{constants.h} found in subdirectory \texttt{src/shared/}.
-\item [{\texttt{FULL\_ATTENUATION\_SOLID}}] Set to \texttt{.true.} to add
-$Q_{\kappa}$ attenuation to the simulations in addition to $Q_{\mu}$
-attenuation when the \texttt{ATTENUATION} flag above is on. Should
-be off in almost all cases for seismic wave propagation, since $Q_{\kappa}$
-is almost always negligible in Earth models. However, this parameter
-can be useful for instance for ocean acoustics simulations, when $Q_{\kappa}$
-is not negligible in the solid layer at the sea bottom.
-\item [{\texttt{ANISOTROPY}}] Set to \texttt{.true.} if you want to use
-an anisotropy model. Please see the file \texttt{model\_aniso.f90}
-in subdirectory \texttt{src/generate\_databases/} for the current
-implementation of anisotropic models.
-\item [{\texttt{TOMOGRAPHY\_PATH}}] Directory in which the tomography files
-are stored for using external tomographic Earth models (please read
-Chapter \ref{cha:-Changing-the} and Section \ref{sec:Using-tomographic}
-`Using external tomographic Earth models' for further details.).
-\item [{\texttt{USE\_OLSEN\_ATTENUATION}}] Set to \texttt{.true.} if you
-want to use the attenuation model that scaled from the S-wave speed
-model using Olsen's empirical relation (see \citet{OlDaBr03}).
-\item [{\texttt{OLSEN\_ATTENUATION\_RATIO}}] Determines the Olsen's constant
-in Olsen's empirical relation (see \citet{OlDaBr03}).
-\item [{\texttt{PML\_CONDITIONS}}] Set to \texttt{.true.} to turn on C-PML
-boundary conditions for a regional simulation. Both fluids and elastic
-solids are supported. Note that xmeshfem3d generated meshes do not support C-PML yet.
-\item [{\texttt{PML\_INSTEAD\_OF\_FREE\_SURFACE}}] Set to \texttt{.true.}
-to turn on C-PML boundary conditions on the top surface instead of
-the usual free surface.
-\item [{\texttt{f0\_FOR\_PML}}] Determines the dominant frequency that
-will be used in the calculation of PML damping profiles; this should
-be set to the same (or similar) dominant frequency as that of the
-source that you will use in your simulation. If you plan to use a
-Dirac source, then use the dominant frequency of the source wavelet
-with which you plan to convolve your seismograms later on in post-processing.
-\item [{\texttt{STACEY\_ABSORBING\_CONDITIONS}}] Set to \texttt{.true.}
-to turn on Clayton-Enquist absorbing boundary conditions (see \citet{KoTr99}).
-In almost all cases it is much better to use CPML absorbing layers
-(see the options above) and leave this flag to \texttt{.false.}.
-\item [{\texttt{STACEY\_INSTEAD\_OF\_FREE\_SURFACE}}] Set to \texttt{.true.}
-to turn on absorbing boundary conditions on the top surface which
-by default constitutes a free surface of the model.
-\item [{\texttt{CREATE\_SHAKEMAP}}] Set this flag to \texttt{.true.} to
-create a ShakeMap\textregistered{}, i.e., a peak ground velocity map
-of the maximum absolute value of the two horizontal components of
-the velocity vector.
-\item [{\texttt{MOVIE\_SURFACE}}] Set to \texttt{.false.}, unless you want
-to create a movie of seismic wave propagation on the Earth's surface.
-Turning this option on generates large output files. See Section~\ref{sec:Movies}
-for a discussion on the generation of movies. This feature is only
-relevant for the solver.
-\item [{\texttt{MOVIE\_TYPE}}] Set this flag to 1 to show the top surface
-(tomography + oceans) only, to 2 to show all external faces of the
-mesh (i.e. topography + vertical edges + bottom) in shakemaps and
-surface movies.
-\item [{\texttt{MOVIE\_VOLUME}}] Set to \texttt{.false.}, unless you want
-to create a movie of seismic wave propagation in the Earth's interior.
-Turning this option on generates huge output files. See Section~\ref{sec:Movies}
-for a discussion on the generation of movies. This feature is only
-relevant for the solver.
-\item [{\texttt{SAVE\_DISPLACEMENT}}] Set this flag to \texttt{.true.}
-if you want to save the displacement instead of velocity for the movie
-frames.
-\item [{\texttt{USE\_HIGHRES\_FOR\_MOVIES}}] Set this flag to \texttt{.true.}
-if you want to save the values at all the NGLL grid points for the
-movie frames.
-\item [{\texttt{NTSTEP\_BETWEEN\_FRAMES}}] Determines the number of timesteps
-between movie frames. Typically you want to save a snapshot every
-100 timesteps. The smaller you make this number the more output will
-be generated! See Section~\ref{sec:Movies} for a discussion on the
-generation of movies. This feature is only relevant for the solver.
-\item [{\texttt{HDUR\_MOVIE}}] Determines the half duration of the source
-time function for the movie simulations. When this parameter is set
-to be 0, a default half duration that corresponds to the accuracy
-of the simulation is provided. Otherwise, it adds this half duration
-to the half duration specified in the source file \texttt{CMTSOLUTION},
-thus simulates longer periods to make the movie images look smoother.
-\item [{\texttt{SAVE\_MESH\_FILES}}] Set this flag to \texttt{.true.} to
-save ParaView \urlwithparentheses{www.paraview.org} mesh files for
-subsequent viewing. Turning the flag on generates large (distributed)
-files in the \texttt{LOCAL\_PATH} directory. See Section~\ref{sec:Mesh-graphics}
-for a discussion of mesh viewing features.
-\item [{\texttt{LOCAL\_PATH}}] Directory in which the distributed databases
-will be written. Generally one uses a directory on the local disk
-of the compute nodes, although on some machines these databases are
-written on a parallel (global) file system (see also the earlier discussion
-of the \texttt{LOCAL\_PATH\_IS\_ALSO\_GLOBAL} flag in Chapter~\ref{cha:Getting-Started}).
-\texttt{xgenerate\_databases} generates the necessary databases in
-parallel, one set for each of the \texttt{NPROC} slices that constitutes
-the mesh (see Figure~\ref{fig:mount.partitions} and Figure~\ref{fig:For-parallel-computing}).
-After the executable finishes, you can log in to one of the compute
-nodes and view the contents of the \texttt{LOCAL\_PATH} directory
-to see the (many) files generated by \texttt{xgenerate\_databases}.
-Please note that the \texttt{LOCAL\_PATH} directory should already
-contain the output files of the partitioner, i.e. from \texttt{xdecompose\_mesh}
-or \texttt{xmeshfem3D}.
-\item [{\texttt{NTSTEP\_BETWEEN\_OUTPUT\_INFO}}] This parameter specifies
-the interval at which basic information about a run is written to
-the file system (\texttt{timestamp{*}} files in the \texttt{OUTPUT\_FILES}
-directory). If you have access to a fast machine, set \texttt{NTSTEP\_BETWEEN\_OUTPUT\_INFO}
-to a relatively high value (e.g., at least 100, or even 1000 or more)
-to avoid writing output text files too often. This feature is not
-used at the time of meshing. One can set this parameter to a larger
-value than the number of time steps to avoid writing output during
-the run.
-\item [{\texttt{NTSTEP\_BETWEEN\_OUTPUT\_SEISMOS}}] This parameter specifies
-the interval at which synthetic seismograms are written in the \texttt{LOCAL\_PATH}
-directory. If a run crashes, you may still find usable (but shorter
-than requested) seismograms in this directory. On a fast machine set
-\texttt{NTSTEP\_BETWEEN\_OUTPUT\_SEISMOS} to a relatively high value
-to avoid writing to the seismograms too often. This feature is only
-relevant for the solver.
-\item [{\texttt{USE\_FORCE\_POINT\_SOURCE}}] Turn this flag on to use a
-(tilted) \texttt{FORCESOLUTION} force point source instead of a \texttt{CMTSOLUTION}
-moment-tensor source. When the force source does not fall exactly
-at a grid point, the solver interpolates the force between grid points
-using Lagrange interpolants. This can be useful e.g. for oil industry
-foothills simulations in which the source is a vertical force, normal
-force, tilted force, or an impact etc. Note that in the \texttt{FORCESOLUTION}
-file, you will need to edit the East, North and vertical components
-of an arbitrary (not necessarily unitary, the code will normalize it automatically) direction vector of the force vector;
-thus refer to Appendix~\ref{cha:Coordinates} for the orientation
-of the reference frame. This vector is made unitary internally in
-the solver and thus only its direction matters here; its norm is ignored
-and the norm of the force used is the factor force source times the
-source time function.
-When using this option, by default the code can locate the force source
-anywhere between mesh points in order to honor its exact location; this
-is more precise than using the closest GLL mesh point, but it is also a bit slower.
-If needed, you can change that default behavior and force the code to use the closest
-GLL mesh point instead by setting flag \texttt{USE\_BEST\_LOCATION} to \texttt{.false.} instead of \texttt{.true.} in file
-\texttt{src/shared/constants.h.in} and running the \texttt{configure} script again and recompiling the code.
-\item [{\texttt{USE\_RICKER\_TIME\_FUNCTION}}] Turn this flag on to use
-a Ricker source time function instead of the source time functions
-set by default to represent a (tilted) \texttt{FORCESOLUTION} force
-point source or a \texttt{CMTSOLUTION} moment-tensor source. Originally,
-if a \texttt{CMTSOLUTION} moment-tensor source is used, a (pseudo)
-Heaviside step function with a very short half duration is defined
-for elastic cases to represent the permanent slip on the fault while
-in the acoustic case a Gaussian source time function with a similarly
-short half duration is defined to physically describe actions within
-the fluid. Otherwise, if a \texttt{FORCESOLUTION} force source is
-used, a (pseudo) Dirac delta source time function is defined by default.
-Any other source-time function may then be obtained by convolution.
-\item [{\texttt{PRINT\_SOURCE\_TIME\_FUNCTION}}] Turn this flag on to print
-information about the source time function in the file \texttt{OUTPUT\_FILES/plot\_source\_time\_function.txt}.
-This feature is only relevant for the solver.
-
-\item [{\texttt{NUMBER\_OF\_SIMULTANEOUS\_RUNS}}] adds the ability to run several calculations (several earthquakes)
-in an embarrassingly-parallel fashion from within the same run;
-this can be useful when using a very large supercomputer to compute
-many earthquakes in a catalog, in which case it can be better from
-a batch job submission point of view to start fewer and much larger jobs,
-each of them computing several earthquakes in parallel.
-To turn that option on, set parameter NUMBER\_OF\_SIMULTANEOUS\_RUNS to a value greater than 1.
-To implement that, we create NUMBER\_OF\_SIMULTANEOUS\_RUNS MPI sub-communicators,
-each of them being labeled "my\_local\_mpi\_comm\_world", and we use them
-in all the routines in "src/shared/parallel.f90", except in MPI\_ABORT() because in that case
-we need to kill the entire run.
-When that option is on, of course the number of processor cores used to start
-the code in the batch system must be a multiple of NUMBER\_OF\_SIMULTANEOUS\_RUNS,
-all the individual runs must use the same number of processor cores,
-which as usual is NPROC in the Par\_file,
-and thus the total number of processor cores to request from the batch system
-should be NUMBER\_OF\_SIMULTANEOUS\_RUNS * NPROC.
-All the runs to perform must be placed in directories called run0001, run0002, run0003 and so on
-(with exactly four digits).
-
-Imagine you have 10 independent calculations to do, each of them on 100 cores; you have three options:
-
-1/ submit 10 jobs to the batch system
-
-2/ submit a single job on 1000 cores to the batch, and in that script create a sub-array of jobs to start 10 jobs,
-each running on 100 cores (see e.g. http://www.schedmd.com/slurmdocs/job\_array.html )
-
-3/ submit a single job on 1000 cores to the batch, start SPECFEM3D on 1000 cores, create 10 sub-communicators,
-cd into one of 10 subdirectories (called e.g. run0001, run0002,... run0010) depending on the sub-communicator
-your MPI rank belongs to, and run normally on 100 cores using that sub-communicator.
-
-The option NUMBER\_OF\_SIMULTANEOUS\_RUNS implements 3/.
-
-\item [{\texttt{BROADCAST\_SAME\_MESH\_AND\_MODEL}}]:
-if we perform simultaneous runs in parallel, if only the source and receivers vary between these runs
-but not the mesh nor the model (velocity and density) then we can also read the mesh and model files
-from a single run in the beginning and broadcast them to all the others; for a large number of simultaneous
-runs for instance when solving inverse problems iteratively this can DRASTICALLY reduce I/Os to disk in the solver
-(by a factor equal to NUMBER\_OF\_SIMULTANEOUS\_RUNS), and reducing I/Os is crucial in the case of huge runs.
-Thus, always set this option to .true. if the mesh and the model are the same for all simultaneous runs.
-In that case there is no need to duplicate the mesh and model file database (the content of the DATABASES\_MPI
-directories) in each of the run0001, run0002,... directories, it is sufficient to have one in run0001
-and the code will broadcast it to the others).
-
-TODO / future work to do: currently the BROADCAST\_SAME\_MESH\_AND\_MODEL option assumes to have the (master) mesh files in run0001/DATABASES\_MPI or run0001/OUTPUT\_FILES/DATABASES\_MPI.
-However, for adjoint runs you still need a DATABASES\_MPI folder in each of the sub-runs directories, e.g.
-run0002/DATABASES\_MPI, etc.
-to store the forward wavefields, kernels etc.
-of each sub-run.
-This would not be needed for forward simulations.
-
-TODO / future work to do: the sensitivity kernel summing and smoothing tools in directory src/tomography are currently not ported to this new option to do many runs simultaneously, only the solver (src/specfem3d) is.
-Thus these tools should work, but in their current version will need to be run for each simulation result independently.
-More precisely, the current kernel summing and smoothing routines work fine, with the exception that you need to move out the mesh files (and also the parameters).
-This works because these routines consider multiple runs by design.
-You simply have to provide them the directories where the kernels are.
-
-\item [{\texttt{GPU\_MODE}}] Turn this flag on to use GPUs.
-
-\item [\texttt{ADIOS\_ENABLED}] Turn this flag on to enable ADIOS. If set to \texttt{.false.}, subsequent ADIOS
-parameters will not be considered.
-\item [\texttt{ADIOS\_FOR\_DATABASES}] Turn this flag on to use ADIOS for xmeshfem3D output and
-xgenerate\_database input.
-\item [\texttt{ADIOS\_FOR\_MESH}]  Turn this flag on to use ADIOS for generated databases.
-\item [\texttt{ADIOS\_FOR\_FORWARD\_ARRAYS}] Turn this flag on to read and write forward arrays using ADIOS.
-\item [\texttt{ADIOS\_FOR\_KERNELS}] Turn this flag on to produce ADIOS kernels that can later be visualized with the ADIOS version of combine\_vol\_data.
-
-\end{description}
-
-The present version of SPECFEM can handle fully saturated porous simulations, Christina Morency implemented Biot equation.
-But the code cannot calculate partially saturated cases in its state.
-Christina Morency is presently working on a dual porosity, dual permeability formulation, type Pride and Berryman, for an other project, but it will not be available for some time.\\
-
-The way we prescribe material property for porous material in SPECFEM3D is as follow:
-We use a file name nummaterial\_poroelastic\_file, which is located in the directory MESH, the format is as follow:\\
- rhos rhof phi c kxx kxy kxz kyy kyz kzz Ks Kf Kfr etaf mufr\\
-where\\
-rho\_s = solid density\\
-rho\_f = fluid density\\
-phi = porosity\\
-tort = tortuosity\\
-kxx = xx component of permeability tensor\\
-kxy = xy,yx components of permeability tensor\\
-kyy = yy component of permeability tensor\\
-kxz = xz,zx components of permeability tensor\\
-kzz = zz component of permeability tensor\\
-kappa\_s = solid bulk modulus\\
-kappa\_f= fluid bulk modulus\\
-kappa\_fr= frame bulk modulus\\
-eta\_f = fluid viscosity\\
-mu\_fr = frame shear modulus.\\
-
-Using an external mesh (for instance coming from CUBIT/TRELIS), poroelastic materials have the ID number 3, while 1 is elastic, 2 is acoustic ?
-see the example in the package: examples/homogeneous\_poroelastic.
-
-If you use PML, the mesh elements that belong to the PML layers can
-be acoustic or elastic, but not viscoelastic nor poroelastic. Then,
-when defining your model, you should define these absorbing elements
-as either acoustic or elastic. In you forget to do that, the code
-will fix the problem by automatically converting the viscoelastic
-or poroelastic PML elements to elastic. This means that strictly speaking
-the PML layer will not be perfectly matched any more, since the physical
-model will change from viscoelastic or poroelastic to elastic at the
-entrance of the PML, but in practice this is sufficient and produces
-only tiny / negligible spurious reflections.
-
-If you use PML and an external mesh (created using an external meshing tool
-such as CUBIT/TRELIS or similar), try to have elements inside the PML as regular as possible,
-i.e. ideally non-deformed cubes obtained by `extrusion' of regular surface mesh elements meshing the
-outer edges of the computational domain without PML; by doing so, the PMLs obtained will be far more stable
-in time (PML being weakly unstable from a mathematical point of view, very deformed mesh elements
-inside the PMLs can trigger instabilities much more quickly).
-
-If you use PML and an external tomographic velocity and density model,
-you should be careful because mathematically a PML cannot handle heterogeneities
-along the normal to the PML edge inside the PML layer. This comes
-from the fact that the damping profile that is defined assumes a constant
-velocity and density model along the normal direction.
-
-Thus, you need to modify your velocity and density model in order
-for it to be 1D inside the PML, as shown in Figure~\ref{fig:modify_external_velocity_model_to_use_PML}.
-
-This applies to the bottom layer as well; there you should make sure
-that your model is 1D and thus constant along the vertical direction.
-
-To summarize, only use a 3D velocity and density model inside the
-physical region, and in all the PML layers extend it by continuity
-from its values along the inner PML edge.
-
-%%
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=3in]{figures/how_to_use_PML_when_a_tomographic_velocity_model_is_used}
-\par\end{centering}
-
-\caption{How to modify your external 3D velocity and density model in order
-to use PML. Such a modification is not needed when using Stacey absorbing
-boundary conditions (but such conditions are significantly less efficient).}
-
-\label{fig:modify_external_velocity_model_to_use_PML}
-\end{figure}
-
-%%
-
-\section{\label{sec:Choosing-the-Time-Step}Choosing the time step \texttt{DT}}
-
-The parameter \texttt{DT} sets the length of each time step in seconds.
-The value of this parameter is crucial for the stability of the spectral-element
-simulation. Your time step \texttt{DT} will depend on the minimum
-ratio between the distance $h$ of neighboring mesh points and the
-wave speeds $v$ defined in your model. The condition for the time
-step $\Delta t$ is:
-\begin{lyxcode}
-$\Delta t<C~\mathrm{min}_{\Omega}(~h/v~)$
-\end{lyxcode}
-where $C$ is the so-called Courant number and $\Omega$ denotes the
-model volume. The distance $h$ depends on the mesh element size and
-the number of GLL points \texttt{NGLL} specified in the main constants
-file \texttt{constants.h} located in the \texttt{src/shared/} subdirectory.
-The wave speed $v$ is determined based on your model's P- (or S-)
-wave speed values.
-
-The database generator \texttt{xgenerate\_databases}, as well as the
-internal mesher \texttt{xmeshfem3D}, are trying to evaluate the value
-of $\Delta t$ for empirically chosen Courant numbers $C\sim0.3$.
-If you used the mesher \texttt{xmeshfem3D} to generate your mesh,
-you should set the value suggested in \texttt{OUTPUT\_FILES/output\_mesher.txt}
-file, which is created after the mesher completed. In case you used
-CUBIT to create the mesh, you might use an arbitrary value when running
-\texttt{xgenerate\_databases} and then use the value suggested in
-the ~\\
- \texttt{OUTPUT\_FILES/output\_mesher.txt} file after the database
-generation completed. Note that the implemented Newmark time scheme
-uses this time step globally, thus your simulations become more expensive
-for very small mesh elements in high wave-speed regions. Please be
-aware of this restriction when constructing your mesh in Chapter~\ref{cha:Mesh-Generation}.
-
-\chapter{\label{cha:Running-the-Solver}Running the Solver \texttt{xspecfem3D}}
-
-Now that you have successfully generated the databases, you are ready
-to compile the solver. In the main directory, type
-\begin{lyxcode}
-{\small make~xspecfem3D}{\small \par}
-\end{lyxcode}
-Please note that \texttt{xspecfem3D} must be called directly from
-the main directory, as most of the binaries of the package.\\
-
-\noindent
-The solver needs three input files in the \texttt{DATA} directory
-to run:
-\begin{description}
-\item [{\texttt{Par\_file}}] the main parameter file which was discussed
-in detail in the previous Chapter~\ref{cha:Creating-Distributed-Databases},
-\item [{\texttt{CMTSOLUTION}} or {\texttt{FORCESOLUTION}}] the earthquake
-source parameter file or the force source parameter file, and
-\item [{\texttt{STATIONS}}] the stations file.
-\end{description}
-Most parameters in the \texttt{Par\_file} should be set prior to running
-the databases generation. Only the following parameters may be changed
-after running \texttt{xgenerate\_databases}:
-\begin{itemize}
-\item the simulation type control parameters: \texttt{SIMULATION\_TYPE}
-and \texttt{SAVE\_FORWARD}
-\item the time step parameters \texttt{NSTEP} and \texttt{DT}
-\item the absorbing boundary control parameter \texttt{PML\_CONDITIONS}
-on condition that the\\
- \texttt{PML\_INSTEAD\_OF\_FREE\_SURFACE} flag remains unmodified
-after running the databases generation.
-\item the movie control parameters \texttt{MOVIE\_SURFACE}, \texttt{MOVIE\_VOLUME},
-and \texttt{NTSTEPS\_BETWEEN\_FRAMES}
-\item the ShakeMap\textregistered{}option \texttt{CREATE\_SHAKEMAP}
-\item the output information parameters \texttt{MOVIE\_TYPE}, \texttt{NTSTEP\_BETWEEN\_OUTPUT\_INFO}
-and \texttt{NTSTEP\_BETWEEN\_OUTPUT\_}~\\
- \texttt{SEISMOS}
-\item the \texttt{PRINT\_SOURCE\_TIME\_FUNCTION} flags
-\end{itemize}
-Any other change to the \texttt{Par\_file} implies rerunning both
-the database generator \texttt{xgenerate\_databases} and the solver
-\texttt{xspecfem3D}.
-
-For any particular earthquake, the \texttt{CMTSOLUTION} file that
-represents the point source may be obtained directly from the Harvard
-Centroid-Moment Tensor (CMT) web page \urlwithparentheses{www.seismology.harvard.edu}.
-It looks like the example shown in Fig. \ref{fig:CMTSOLUTION-file}.
-%\begin{lyxcode}
-\begin{figure}[htp]
-\noindent
-\begin{centering}
-{\small \includegraphics[width=0.8\textwidth]{figures/Hollywood_CMT}
-}
-\par
-\end{centering}
-\caption{\texttt{CMTSOLUTION} file based on the
-format from the Harvard CMT catalog. \textbf{M} is the moment tensor,
-$M_{0}${\small {} }is the seismic moment, and $M_{w}$ is the moment
-magnitude.}
-\label{fig:CMTSOLUTION-file}
-\end{figure}
-%\end{lyxcode}
-The \texttt{CMTSOLUTION} file should be edited in the following way:
-\begin{itemize}
-\item Set the latitude or UTM $x$ coordinate, longitude or UTM $y$ coordinate, depth of the source (in km).
-\item Set the \texttt{time shift} parameter equal to $0.0$ (the solver
-will not run otherwise.) The time shift parameter would simply apply
-an overall time shift to the synthetics, something that can be done
-in the post-processing (see Section \ref{sec:Process-data-and-syn}).
-\item For point-source simulations (see finite sources, page \pageref{To-simulate-a})
-we recommend setting the source half-duration parameter \texttt{half
-duration} equal to zero, which corresponds to simulating a step source-time
-function, i.e., a moment-rate function that is a delta function. If
-\texttt{half duration} is not set to zero, the code will use a Gaussian
-(i.e., a signal with a shape similar to a `smoothed triangle', as
-explained in \citet{KoTr02a} and shown in Fig~\ref{fig:gauss.vs.triangle})
-source-time function with half-width \texttt{half duration}. We prefer
-to run the solver with \texttt{half duration} set to zero and convolve
-the resulting synthetic seismograms in post-processing after the run,
-because this way it is easy to use a variety of source-time functions
-(see Section \ref{sec:Process-data-and-syn}). \citet{KoTr02a} determined
-that the noise generated in the simulation by using a step source
-time function may be safely filtered out afterward based upon a convolution
-with the desired source time function and/or low-pass filtering. Use
-the serial code \texttt{convolve\_source\_timefunction.f90} and the
-script \texttt{convolve\_source\_timefunction.csh} for this purpose,
-or alternatively use signal-processing software packages such as SAC
-\urlwithparentheses{www.llnl.gov/sac}. Type
-
-\begin{lyxcode}
-{\small make~xconvolve\_source\_timefunction}{\small \par}
-\end{lyxcode}
-
-to compile the code and then set the parameter \texttt{hdur} in \texttt{convolve\_source\_timefunction.csh}
-to the desired half-duration.
-
-\item The zero time of the simulation corresponds to the center of the triangle/Gaussian,
-or the centroid time of the earthquake. The start time of the simulation
-is $t=-1.5*\texttt{half duration}$ (the 1.5 is to make sure the moment
-rate function is very close to zero when starting the simulation).
-To convert to absolute time $t_{\mathrm{abs}}$, set
-
-\begin{lyxcode}
-$t_{\mathrm{abs}}=t_{\mathrm{pde}}+\texttt{time shift}+t_{\mathrm{synthetic}}$
-\end{lyxcode}
-
-where $t_{\mathrm{pde}}$ is the time given in the first line of the
-\texttt{CMTSOLUTION}, \texttt{time shift} is the corresponding value
-from the original \texttt{CMTSOLUTION} file and $t_{\mathrm{synthetic}}$
-is the time in the first column of the output seismogram.
-
-\end{itemize}
-\begin{figure}[htp]
-\noindent \begin{centering}
-\includegraphics[width=3.5in]{figures/gauss_vs_triangle_mod}
-\par\end{centering}
-
-\caption{Comparison of the shape of a triangle and the Gaussian function actually
-used.}
-
-\label{fig:gauss.vs.triangle}
-\end{figure}
-
-If you know the earthquake source in strike/dip/rake format rather
-than in \texttt{CMTSOLUTION} format, use the C code \texttt{SPECFEM3D\_GLOBE/utils/strike\_dip\_rake\_to\_CMTSOLUTION.c}
-to convert it. The conversion formulas are given for instance in \citet{AkRi80}.
-Note that the \citet{AkRi80} convention is slightly different from
-the Harvard \texttt{CMTSOLUTION} convention (the sign of some components
-is different). The C code outputs both.
-
-Centroid latitude and longitude should be provided in geographical
-coordinates. The code converts these coordinates to geocentric coordinates~\citep{DaTr98}.
-Of course you may provide your own source representations by designing
-your own \texttt{CMTSOLUTION} file. Just make sure that the resulting
-file adheres to the Harvard CMT conventions (see Appendix~\ref{cha:Coordinates}).
-Note that the first line in the \texttt{CMTSOLUTION} file is the Preliminary
-Determination of Earthquakes (PDE) solution performed by the USGS
-NEIC, which is used as a seed for the Harvard CMT inversion. The PDE
-solution is based upon P waves and often gives the hypocenter of the
-earthquake, i.e., the rupture initiation point, whereas the CMT solution
-gives the `centroid location', which is the location with dominant
-moment release. The PDE solution is not used by our software package
-but must be present anyway in the first line of the file.
-
-\label{To-simulate-a}To simulate a kinematic rupture, i.e., a finite-source
-event, represented in terms of $N_{\mathrm{sources}}$ point sources,
-provide a \texttt{CMTSOLUTION} file that has $N_{\mathrm{sources}}$
-entries, one for each subevent (i.e., concatenate $N_{\mathrm{sources}}$
-\texttt{CMTSOLUTION} files to a single \texttt{CMTSOLUTION} file).
-At least one entry (not necessarily the first) must have a zero \texttt{time
-shift}, and all the other entries must have non-negative \texttt{time
-shift}. Each subevent can have its own half duration, latitude, longitude,
-depth, and moment tensor (effectively, the local moment-density tensor).
-
-Note that the zero in the synthetics does NOT represent the hypocentral
-time or centroid time in general, but the timing of the \textit{center}
-of the source triangle with zero \texttt{time shift} (Fig~\ref{fig:source_timing}).
-
-Although it is convenient to think of each source as a triangle, in
-the simulation they are actually Gaussians (as they have better frequency
-characteristics). The relationship between the triangle and the gaussian
-used is shown in Fig~\ref{fig:gauss.vs.triangle}. For finite fault
-simulations it is usually not advisable to use a zero half duration
-and convolve afterwards, since the half duration is generally fixed
-by the finite fault model.
-
-\vspace{1cm}
-
-\noindent The \texttt{FORCESOLUTION} file should be edited in the
-following way:
-\begin{itemize}
-\item Set the \texttt{time shift} parameter equal to $0.0$ (the solver
-will not run otherwise.) The time shift parameter would simply apply
-an overall time shift to the synthetics, something that can be done
-in the post-processing (see Section \ref{sec:Process-data-and-syn}).
-\item Set the \texttt{f0} parameter (the dominant frequency) of the Ricker source time
-function when {\texttt{USE\_RICKER\_TIME\_FUNCTION}} is turned on
-in the main parameter file {\texttt{Par\_file}}. In case that the
-solver uses a (pseudo) Dirac delta source time function to represent
-a force point source, a very short duration of five time steps is automatically set by default.
-\item Set the latitude or UTM $x$ coordinate, longitude or UTM $y$ coordinate, depth of the source (in km).
-\item Set the magnitude of the force source.
-\item Set the components of a (non-unitary) direction vector for the force
-source in the East/North/Vertical basis (see Appendix A for the orientation
-of the reference frame).
-\end{itemize}
-\noindent Where necessary, set a \texttt{FORCESOLUTION} file in the
-same way you configure a \texttt{CMTSOLUTION} file with $N_{\mathrm{sources}}$
-entries, one for each subevent (i.e., concatenate $N_{\mathrm{sources}}$
-\texttt{FORCESOLUTION} files to a single \texttt{FORCESOLUTION} file).
-At least one entry (not necessarily the first) must have a zero \texttt{time
-shift}, and all the other entries must have non-negative \texttt{time
-shift}. Each subevent can have its own half latitude, longitude, depth,
-\texttt{half duration} and force parameters.
-
-\begin{figure}[H]
-\noindent \begin{centering}
-{\small \includegraphics[width=5in]{figures/source_timing} }
-\par\end{centering}
-
-\caption{Example of timing for three sources. The center of the first source
-triangle is defined to be time zero. Note that this is NOT in general
-the hypocentral time, or the start time of the source (marked as tstart).
-The parameter \texttt{time shift} in the \texttt{CMTSOLUTION} file
-would be t1(=0), t2, t3 in this case, and the parameter \texttt{half
-duration} would be hdur1, hdur2, hdur3 for the sources 1, 2, 3 respectively.}
-
-{\small \label{fig:source_timing} }
-\end{figure}
-
-The solver can calculate seismograms at any number of stations for
-basically the same numerical cost, so the user is encouraged to include
-as many stations as conceivably useful in the \texttt{STATIONS} file,
-which looks like this:
-
-\begin{figure}[H]
-\noindent \begin{centering}
-{\small \includegraphics{figures/STATIONS_basin_explained} }
-\par\end{centering}
-
-\label{fig:Sample-STATIONS-file.}\caption{Sample \texttt{STATIONS} file. Station latitude and longitude should
-be provided in geographical coordinates. The width of the station
-label should be no more than 32 characters (see \texttt{MAX\_LENGTH\_STATION\_NAME}
-in the \texttt{constants.h} file), and the network label should be
-no more than 8 characters (see \texttt{MAX\_LENGTH\_NETWORK\_NAME}
-in the \texttt{constants.h} file).}
-\end{figure}
-
-Each line represents one station in the following format:
-\begin{lyxcode}
-{\small Station~Network~Latitude~(degrees)~Longitude~(degrees)~Elevation~(m)~burial~(m)~}{\small \par}
-\end{lyxcode}
-The solver \texttt{xspecfem3D} filters the list of stations in file
-\texttt{DATA/STATIONS} to exclude stations that are not located within
-the region given in the \texttt{Par\_file} (between \texttt{LATITUDE\_MIN}
-and \texttt{LATITUDE\_MAX} and between \texttt{LONGITUDE\_MIN} and
-\texttt{LONGITUDE\_MAX}). The filtered file is called \texttt{DATA/STATIONS\_FILTERED}.
-
-Solver output is provided in the \texttt{OUTPUT\_FILES} directory
-in the \texttt{output\_solver.txt} file. Output can be directed to
-the screen instead by uncommenting a line in \texttt{constants.h}:
-\begin{lyxcode}
-!~uncomment~this~to~write~messages~to~the~screen~
-
-!~integer,~parameter~::~IMAIN~=~ISTANDARD\_OUTPUT~
-\end{lyxcode}
-On PC clusters the seismogram files are generally written to the local
-disks (the path \texttt{LOCAL\_PATH} in the \texttt{Par\_file}) and
-need to be gathered at the end of the simulation.
-
-While the solver is running, its progress may be tracked by monitoring
-the `\texttt{\small timestamp{*}}{\small ' files in the ~}\\
-{\small{} }\texttt{\small OUTPUT\_FILES}{\small{} directory. These tiny
-files look something like this:}{\small \par}
-\begin{lyxcode}
-Time~step~\#~~~~~~~~~~10000~
-
-Time:~~~~~108.4890~~~~~~seconds~
-
-Elapsed~time~in~seconds~=~~~~1153.28696703911~
-
-Elapsed~time~in~hh:mm:ss~=~~~~~0~h~19~m~13~s~
-
-Mean~elapsed~time~per~time~step~in~seconds~=~~~~~0.115328696703911~
-
-Max~norm~displacement~vector~U~in~all~slices~(m)~=~~~~1.0789589E-02~
-\end{lyxcode}
-The \texttt{\small timestamp{*}}{\small{} files provide the }\texttt{\small Mean
-elapsed time per time step in seconds}{\small , which may be used
-to assess performance on various machines (assuming you are the only
-user on a node), as well as the }\texttt{\small Max norm displacement
-vector U in all slices~(m)}{\small . If something is wrong with the
-model, the mesh, or the source, you will see the code become unstable
-through exponentially growing values of the displacement and fluid
-potential with time, and ultimately the run will be terminated by
-the program. You can control the rate at which the timestamp files
-are written based upon the parameter }\texttt{\small NTSTEP\_BETWEEN\_OUTPUT\_INFO}{\small{}
-in the }\texttt{\small Par\_file}{\small .}{\small \par}
-
-Having set the \texttt{Par\_file} parameters, and having provided
-the \texttt{CMTSOLUTION} (or the \texttt{FORCESOLUTION}) and \texttt{STATIONS}
-files, you are now ready to launch the solver! This is most easily
-accomplished based upon the \texttt{go\_solver} script (See Chapter
-\ref{cha:Scheduler} for information about running through a scheduler,
-e.g., LSF). You may need to edit the last command at the end of the
-script that invokes the \texttt{mpirun} command. The \texttt{runall}
-script compiles and runs both \texttt{xgenerate\_databases} and \texttt{xspecfem3D}
-in sequence. This is a safe approach that ensures using the correct
-combination of distributed database output and solver input.
-
-It is important to realize that the CPU and memory requirements of
-the solver are closely tied to choices about attenuation (\texttt{ATTENUATION})
-and the nature of the model (i.e., isotropic models are cheaper than
-anisotropic models). We encourage you to run a variety of simulations
-with various flags turned on or off to develop a sense for what is
-involved.
-
-For the same model, one can rerun the solver for different events
-by simply changing the \texttt{CMTSOLUTION} or \texttt{FORCESOLUTION}
-file, or for different stations by changing the \texttt{STATIONS}
-file. There is no need to rerun the \texttt{xgenerate\_databases}
-executable. Of course it is best to include as many stations as possible,
-since this does not add to the cost of the simulation.
-
-We have also added the ability to run several calculations (several earthquakes)
-in an embarrassingly-parallel fashion from within the same run;
-this can be useful when using a very large supercomputer to compute
-many earthquakes in a catalog, in which case it can be better from
-a batch job submission point of view to start fewer and much larger jobs,
-each of them computing several earthquakes in parallel.
-To turn that option on, set parameter \texttt{NUMBER\_OF\_SIMULTANEOUS\_RUNS}
-to a value greater than 1 in file setup/constants.h.in before
-configuring and compiling the code.
-When that option is on, of course the number of processor cores used to start
-the code in the batch system must be a multiple of \texttt{NUMBER\_OF\_SIMULTANEOUS\_RUNS},
-all the individual runs must use the same number of processor cores,
-which as usual is \texttt{NPROC} in the input file \texttt{DATA/Par\_file},
-and thus the total number of processor cores to request from the batch system
-should be \texttt{NUMBER\_OF\_SIMULTANEOUS\_RUNS $\times$ NPROC}.
-All the runs to perform must be placed in directories called \texttt{run0001}, \texttt{run0002}, \texttt{run0003} and so on (with exactly four digits)
-and you must create a link from the root directory of the code
-to the first copy of the executable programs by typing \texttt{ln -s run0001/bin bin}.
-
-\chapter{\label{cha:Fault-Kinematics-Dynamics}Kinematic and dynamic fault
-sources}
-
-SPECFEM3D can handle finite fault sources of two kinds:
-\begin{enumerate}
-\item \emph{Kinematic}: the spatio-temporal distribution of slip rate is
-prescribed along the fault surface
-\item \emph{Dynamic}: a friction law and initial stresses are prescribed
-on the fault, a spontaneous rupture process is computed.
-\end{enumerate}
-
-\section{\label{sec:Mesh-Generation-with}Mesh Generation with Split Nodes}
-
-Faults need to be handled in a special way during mesh generation.
-A fault surface must lie at the interface between elements (the mesh
-must honor the fault surfaces). Moreover, a fault is made of two surfaces
-in contact. Each of these two surfaces needs a separate set of nodes.
-This approach is known as \textquotedbl{}split nodes\textquotedbl{}.
-Currently faults can only be run with \texttt{xdecompose\_mesh} and
-\texttt{CUBIT}. \texttt{xmeshfem3D} is not yet ready to handle faults.\\
-To facilitate the mesh generation with split nodes in CUBIT, we need
-to separate the two fault surfaces by a small distance, effectively
-creating a tiny opening of the fault (Figure \ref{fig:examples.splitnodes},
-\ref{fig:examples.splitnodes-surfacetrace}). Note that only the interior
-of the fault must be opened, its edges must remain closed (except
-the edge on the free surface). The fault is automatically closed later
-by SPECFEM3D.\\
-
-Here is an example CUBIT script to generate a mesh with split nodes
-for a buried vertical strike-slip fault:
-\begin{lyxcode}
-~reset~~\\
-~brick~x~10~y~10~z~10~~\\
-~webcut~volume~all~with~plane~xplane~~\\
-~webcut~volume~all~with~plane~yplane~~\\
-~webcut~volume~all~with~plane~xplane~offset~3~~\\
-~webcut~volume~all~with~plane~zplane~offset~3~~\\
-~webcut~volume~all~with~plane~zplane~offset~-3~~\\
-~imprint~all~~\\
-~merge~all~~\\
-~unmerge~surf~160~~\\
-~mesh~vol~all~~\\
-~set~node~constraint~off~~\\
-~node~in~surf~168~move~X~0~Y~0.01~Z~0~~\\
-~node~in~surf~160~move~X~0~Y~-0.01~Z~0~\\
-
-\end{lyxcode}
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[scale=0.3]{figures/faultmesh} \\
- \includegraphics[scale=0.3]{figures/surf168} \includegraphics[scale=0.3]{figures/splitnodes}
-\par\end{centering}
-
-\caption{Screenshots of the CUBIT example for the embedded fault with split
-nodes: The entire mesh is decomposed into several volumes bounding
-the fault (top), the fault surface 168 is shown as orange squares
-(middle) and split nodes of the fault shown as orange dots(bottom).
-Note that only interior nodes of the fault are split while those on
-the edges of the fault surface are touching each other.}
-
-\label{fig:examples.splitnodes}
-\end{figure}
-
-\begin{lyxcode}
-
-\end{lyxcode}
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[scale=0.55]{figures/splitnodes_surfacetrace} \\
-
-\par\end{centering}
-
-\caption{Screenshots of a CUBIT example showing split nodes for a fault reaching
-the surface. Surface trace of the fault is shown in orange. Note that
-the edges of the fault are not split while the interior nodes are
-offset by a small distance on either side of the fault}
-
-\label{fig:examples.splitnodes-surfacetrace}
-\end{figure}
-
-The CUBIT scripts ({*}.jou and {*}.py) in the directory \texttt{EXAMPLES}
-generate more complicated meshes. The {*}.py files are Python scripts
-that execute CUBIT commands and use the CUBIT-python interface for
-SPECFEM3D (see next section). The Python language allows to define
-and manipulate variables to parameterize the mesh. Alternatively,
-the Python script can call a CUBIT journal file ({*}.jou), which looks
-like the example above. Variables can be defined and manipulated there
-using the $\mathtt{APREPRO}$ language built in CUBIT.\\
-
-Note that you should avoid gaps in the list of indices of mesh objects
-with the following CUBIT command:
-\begin{lyxcode}
-$\mathtt{compress\: ids\: hex\: face\: edge\: node}$
-\end{lyxcode}
-(otherwise you will get a segmentation fault during domain decomposition)
-
-\section{CUBIT-Python Scripts for Faults}
-
-The mesh generated in CUBIT needs to be processed and exported in
-a format compatible with SPECFEM3D. This is achieved in the Python
-scripts by calling the Python-CUBIT interface functions defined in
-the CUBIT directory:
-\begin{enumerate}
-\item Function $\mathtt{define\_bc}$ (or $\mathtt{boundary\_definition.py}$)
-must be called to set up the absorbing boundaries database
-\item Function $\mathtt{fault\_input}$ must be called once for each fault
-to set up the fault database
-\item Function $\mathtt{cubit2specfem3d.export2SPECFEM3D}$ must be called
-at the very end of the script to export the mesh in a SPECFEM3D format.
-\end{enumerate}
-The functions in \#1 and \#3 are for the bulk and are documented in
-Section \ref{subsec:Exporting-the-Mesh}. We focus here on \#2:
-\begin{description}
-\item [{Function:}] $\mathtt{fault\_input}$
-\item [{Purpose:}] export a fault mesh database from CUBIT to a SPECFEM3D-compliant
-file
-\item [{Syntax:}] fault\_input(id\_fault, ids\_surf\_1, ids\_surf\_2)
-\item [{Inputs:}]~\end{description}
-\begin{itemize}
-\item id\_fault integer index assigned to the fault. (The user must number
-all the faults, starting at 1, with unit increments)
-\item ids\_surf\_1 list of CUBIT ids of all surfaces that form side 1 of
-the fault.
-\item ids\_surf\_2 list of CUBIT ids of all surfaces that form side 2 of
-the fault. (The user must decide which side of the fault is side 1.
-This choice affects the sign conventions of fault quantities as explained
-in Section \ref{sec:Sign-Convention-for}).\end{itemize}
-\begin{description}
-\item [{Outputs:}] file $\mathtt{fault\_file\_X.dat}$, where X is the
-fault id (id\_fault).
-\item [{Example:}] For the example in Section \ref{sec:Mesh-Generation-with}:\end{description}
-\begin{lyxcode}
-$\hphantom{}$A1~=~{[}168{]}
-
-$\hphantom{}$A2~=~{[}160{]}
-
-$\hphantom{}$fault\_input(1,A1,A2)
-\end{lyxcode}
-
-\section{Examples}
-
-The package includes examples, the SCEC benchmark problems:
-\begin{itemize}
-\item TPV5, a planar vertical strike-slip fault
-\item TPV14 and TPV15, vertical strike-slip fault system with a fault branch
-\item TPV102 and TPV103, rate-and-state benchmarks
-\item TPV22 and TPV23, step-over benchmarks
-\item TPV16, heterogenous initial stress
-\end{itemize}
-and
-\begin{itemize}
-\item Splay fault models from Wendt et al. (2009)\\
-
-\end{itemize}
-Read the documents in the directory \texttt{EXAMPLES/*/description}.
-They contain a description of the example and additional instructions
-to run it. Visualize the results with the Matlab scripts in the directory
-$\mathtt{EXAMPLES/*/post}$
-
-\section{\label{sec:Sign-Convention-for}Sign Convention for Fault Quantities}
-
-During mesh generation, the fault is defined by two surfaces in contact.
-Let's denote as \textquotedbl{}\emph{side 1}\textquotedbl{} the SECOND
-surface declared by the user in the call to the python function \textquotedbl{}\emph{fault\_input}\textquotedbl{},
-and the FIRST surface as \textquotedbl{}\emph{side 2}\textquotedbl{}.
-The local coordinate system on the fault is defined as the right-handed
-coordinate system defined by (strike, dip, normal), where \textquotedbl{}\emph{normal}\textquotedbl{}
-is the normal vector outgoing from side 1, \textquotedbl{}\emph{dip}\textquotedbl{}
-is parallel to the along-dip direction pointing downwards, and \textquotedbl{}\emph{strike}\textquotedbl{}
-is the horizontal along-strike vector such that the system is right-handed.\\
-
-Slip is defined as displacement on side 2 minus displacement on side
-1. In the local coordinate system on the fault, positive along-strike
-slip is right-lateral and positive along-dip slip is thrust if side
-1 is on the hanging wall (normal faulting if side 1 is on the foot
-wall).\\
-
-Traction is defined as the stress induced on side 1 by side 2, which
-is the stress tensor times the normal vector outgoing from side 1.
-In the local coordinate system on the fault, the normal traction is
-negative in compression, positive along-strike traction generates
-right-lateral slip and positive along-dip traction generates thrust
-slip if side 1 is on the hanging wall (normal faulting if side 1 is
-on the foot wall).
-
-\section{Input Files}
-
-Three additional input files are required in the \texttt{DATA/} directory
-for dynamic and kinematic faults. They are $\mathtt{Par\_file\_faults}$,
-$\mathtt{FAULT\_STATIONS}$ and $\mathtt{input\_file.txt}$ (optional).
-If the former does not exist, the code assumes that there are no faults.\\
-
-\begin{lyxlist}{00.00.0000}
-\item [{\textbf{DATA/Par\_file\_faults}}] contains parameters of the fault.
-The first part of this file has a strict format:
-
-\begin{description}
-\item [{Line}] 1: Number of faults (NF)
-\item [{Lines}] 2 to NF+1: Kelvin Voigt damping (in seconds) for each fault.
-(See below how to set this parameter)
-\item [{Line}] NF+2: Type of simulation (1=dynamic , 2 = kinematic)
-\item [{Line}] NF+3: Number of time steps between updates of the time series
-outputs at selected fault points (see DATA/FAULT\_STATIONS), usually
-a large number (100s or 1000s). Note that the sampling rate of the
-time series is usually much higher.
-\item [{Line}] NF+4: Number of time steps between fault snapshot outputs
-(quantities at every fault point exported at regular times), usually
-a large number (100s or 1000s).
-\item [{Line}] NF+5: Slip velocity threshold below which frictional healing
-is set (friction coefficient is reset to its static value). If this
-value is negative healing is disabled.
-\item [{Line}] NF+6: Slip velocity threshold to define the rupture front.
-Only used for outputs.
-\end{description}
-
-The rest of this file is made of namelist input blocks (see \textquotedbl{}namelist\textquotedbl{}
-in a Fortran 9x manual). The input for each fault has the following
-sequence (arguments in {[}brackets{]} are optional):
-
-\&\textbf{BEGIN\_FAULT} /
-
-\&\textbf{INIT\_STRESS} S1, S2, S3 {[},n1, n2, n3{]} /
-
-followed by (n1+n2+n3) \&\textbf{DIST2D} blocks
-
-\&\textbf{SWF} mus, mud, dc {[}, nmus, nmud, ndc{]} /
-
-followed by (nmus+nmud+ndc) \&\textbf{DIST2D} blocks\\
-
-\begin{lyxlist}{00.00.0000}
-\item [{\&\textbf{INIT\_STRESS}}] input block sets the initial fault stresses
-relative to the foot-wall side of the fault. Initial stresses are
-composed of a constant background value possibly overwritten in prescribed
-regions by heterogeneous distributions (see \&\textbf{DIST2D} blocks
-below):
-
-\begin{description}
-\item [{S1}] = constant background value of along-strike shear stress (positive
-in the usual strike direction)
-\item [{S2}] = constant background value of along-dip shear (positive is
-down-dip, normal faulting)
-\item [{S3}] = constant background value of normal stress (negative in
-compression)
-\item [{n1}] = number of heterogeneous items for along-strike shear stress
-{[}default is 0{]}
-\item [{n2}] = number of heterogeneous items for along-dip shear stress
-{[}default is 0{]}
-\item [{n3}] = number of heterogeneous items for normal stress {[}default
-is 0{]}
-\end{description}
-\item [{\&\textbf{SWF}}] input block sets the slip-weakening friction parameters
-of the fault:
-
-\begin{description}
-\item [{mus}] = constant background value of static friction coefficient
-\item [{mud}] = constant background value of dynamic friction coefficient
-\item [{dc}] = constant background value of critical slip-weakening distance
-\item [{nmus}] = number of heterogeneous items for static friction coefficient
-{[}default is 0{]}
-\item [{nmud}] = number of heterogeneous items for dynamic friction coefficient
-{[}default is 0{]}
-\item [{ndc}] = number of heterogeneous items for critical slip-weakening
-distance {[}default is 0{]}\\
-
-\end{description}
-\item [{\&\textbf{DIST2D}}] input blocks modify (overwrite) the value of
-a fault parameter by a heterogeneous spatial distribution:
-
-\&\textbf{DIST2D} shape='$\mathtt{square}$', val, xc, yc, zc, l /
-
-$\;$sets a constant value (val) within a cube with center (xc,yc,zc)
-and edge size l.
-
-\&\textbf{DIST2D} shape='$\mathtt{rectangle}$', val, xc, yc, zc,
-lx, ly, lz /
-
-$\;$sets a constant value (val) within a rectangular prism with center
-(xc,yc,zc) and edge sizes (lx,ly,lz).
-
-\&\textbf{DIST2D} shape='$\mathtt{rectangle-taper}$', val, valh,
-xc, yc, zc, lx, ly, lz /
-
-$\;$sets a vertical linear gradient within a rectangular prism with
-center (xc,yc,zc) and edge sizes (lx,ly,lz). Values vary linearly
-as a function of vertical position z between value val at z = zc-lz/2
-and value valh at z = zc+lz/2 .
-
-\&\textbf{DIST2D} shape='$\mathtt{circular}$', val, xc, yc, zc, r
-/
-
-$\;$sets a constant value (val) within a sphere with center (xc,yc,zc)
-and radius r. \\
-
-\end{lyxlist}
-\end{lyxlist}
-
-\begin{lyxlist}{00.00.0000}
-\item [{\textbf{DATA/FAULT\_STATIONS}}] Stations in the fault plane.
-
-\textbf{Line 1} : number of stations.
-
-\textbf{Line 2 to end}: 5 columns: X, Y, Z (-depth), station name,
-fault-id
-
-The fault-id identifies the fault that contains the station. It is
-the index of appearance in the faults list after line 2 of Par\_file\_faults
-
-\end{lyxlist}
-
-\begin{lyxlist}{00.00.0000}
-\item [{\textbf{DATA/input\_file.txt}}] Heterogeneous stresses and friction
-parameters are documented in page 10 of
-
-\begin{lyxlist}{00.00.0000}
-\item [{$\mathtt{EXAMPLES/tpv16/description/TPV16\_17\_Description\_v03.pdf}$}]~
-\end{lyxlist}
-
-To activate this feature, in fault\_solver\_dynamic.f90 set \textbf{TPV16}
-= .true. then re-compile the code.
-
-\end{lyxlist}
-
-\section{\label{sec:Setting-the-Kelvin-Voigt}Setting the Kelvin-Voigt Damping
-Parameter }
-
-The purpose of the Kelvin-Voigt viscosity in the dynamic fault solver
-is to damp spurious oscillations generated by the fault slip at frequencies
-that are too high to be resolved by the mesh. The viscosity $\mathtt{eta}$
-(in seconds) depends on the size of the elements on the fault. Here
-is how to set it:
-\begin{enumerate}
-\item Determine the average linear size of the elements on the fault plane,
-$\mathtt{h\_fault}$. Usually this value is prescribed by the user
-during mesh generation. Otherwise it can be found by inspection of
-the mesh inside the CUBIT GUI.
-\item Use the Matlab function $\mathtt{utils/critical\_timestep.m}$ to
-compute\\
- $\mathtt{dtc\_fault}=\mathtt{critical\_timestep\left(c_{p},h\_fault,ngll\right)}$.
-\\
-This is the critical time step in an elastic medium for a hypothetical
-element of cubic shape with size equal to $\mathtt{h\_fault}$.
-\item Set $\mathtt{eta}$ in $\mathtt{Par\_file\_faults}$ to (0.1 to 0.3)$\ensuremath{\times}\mathtt{dtc\_fault}$.
-A larger $\mathtt{eta}$ damps high-frequencies more aggressively
-but it might also affect lower frequencies and rupture speed.
-\end{enumerate}
-Viscosity reduces numerical stability: the critical timestep in a
-simulation with Kelvin-Voigt damping needs to be smaller than that
-in a purely elastic simulation. Here is how to set the time step accordingly:
-\begin{enumerate}
-\item Run a test simulation without viscosity ($\mathtt{eta}$=0 and only
-a few time steps)
-\item Look for the \textquotedbl{}maximum suggested time step\textquotedbl{}
-in \texttt{OUTPUT\_FILES/output\_mesher.txt}. This is the critical
-timestep of a purely elastic simulation, $\mathtt{dtc\_bulk}$.
-\item Reset the timestep of the simulation with a Kelvin-Voigt material
-to a value smaller than\\
- $\mathtt{dtc\_kv}=\mathtt{eta}\left(\sqrt{1+\mathtt{dtc\_bulk^{2}}/\mathtt{eta^{2}}}-1\right)$
-\end{enumerate}
-Note that in general $\mathtt{dtc\_bulk}$ is smaller than $\mathtt{dtc\_fault}$,
-because elements off the fault might be smaller or more distorted
-than element faces on the fault.
-
-\section{Output Files }
-
-Several output files are saved in \texttt{OUTPUT\_FILES/}:\\
-
-\begin{enumerate}
-\item Seismograms for each station on the fault plane given in \texttt{DATA/FAUL\_STATIONS}.
-One output file is generated for each station, named after the station.
-The files are ascii and start with a header (22 lines long) followed
-by a data block with the following format, one line per time sample:\\
-
-Column 1 = Time (s)
-
-Column 2 = horizontal right-lateral slip (m)
-
-Column 3 = horizontal right-lateral slip rate (m/s)
-
-Column 4 = horizontal right-lateral shear stress (MPa)
-
-Column 5 = vertical up-dip slip (m)
-
-Column 6 = vertical up-dip slip rate (m/s)
-
-Column 7 = vertical up-dip shear stress (MPa)
-
-Column 8 = normal stress (MPa)\\
-
-The stresses are relative to the footwall side of the fault (this
-convention controls their sign, but not their amplitude). Slip is
-defined as displacement of the hanging wall relative to the footwall.
-
-\item Seismograms at stations in the bulk (out of the fault plane) given
-in \texttt{DATA/STATIONS}.
-\item Rupture time files are named $\mathtt{Rupture\_time\_FAULT}$-$\mathtt{id}$.
-One file is generated for each fault. The files are ascii and start
-with a header (12 lines long) followed by a data block with the following
-format, one line per fault node:
-
-Column 1 = horizontal coordinate, distance along strike (m)
-
-Column 2 = vertical coordinate, distance down-dip (m)
-
-Column 3 = rupture time (s)
-
-\item Fault quantities (slip, slip rate, stresses, etc) at regular times
-are stored in binary data files called $\mathtt{Snapshot\#it\#.bin}$,
-where \#it\# is the timestep number. These can be read in Matlab with
-the function $\mathtt{utils/FSEM3D\_snapshot.m}$
-\end{enumerate}
-
-\section{Post-processing and Visualization }
-
-Some Matlab functions for post-processing and visualization are included
-in directory $\mathtt{utils}$. The functions are internally documented
-(see their matlab help).\\
-
-\texttt{FSEM3D\_snapshot} reads a fault data snapshot\\
-
-The directories \texttt{EXAMPLES/*/post} contain additional Matlab
-scripts to generate figures specific to each example.
-
-\chapter{\label{cha:Adjoint-Simulations}Adjoint Simulations}
-
-Adjoint simulations are generally performed for two distinct applications.
-First, they can be used in point source moment-tensor inversions, or source imaging for
-earthquakes with large ruptures such as the Lander's earthquake \citep{WaHe94}.
-Second, they can be used to generate finite-frequency sensitivity
-kernels that are a critical part of tomographic inversions based upon
-3D reference models \citep{TrTaLi05,LiTr06,TrKoLi08,LiTr08}. In either
-case, source parameter or velocity structure updates are sought to
-minimize a specific misfit function (e.g., waveform or traveltime
-differences), and the adjoint simulation provides a means of computing
-the gradient of the misfit function and further reducing it in successive
-iterations. Applications and procedures pertaining to source studies
-and finite-frequency kernels are discussed in Sections~\ref{sec:Adjoint-simulation-sources}
-and \ref{sec:Adjoint-simulation-finite}, respectively. The two related
-parameters in the \texttt{Par\_file} are \texttt{SIMULATION\_TYPE}
-(1, 2 or 3) and the \texttt{SAVE\_FORWARD} (boolean).
-
-\section{\label{sec:Adjoint-simulation-sources}Adjoint Simulations for Sources
-Only (not for the Model)}
-
-When a specific misfit function between data and synthetics is minimized to invert
-for earthquake source parameters, the gradient of the misfit function
-with respect to these source parameters can be computed by placing
-time-reversed seismograms at the receivers as virtual sources
-in an adjoint simulation. Then the value of the gradient is obtained
-from the adjoint seismograms recorded at the original earthquake location.
-\begin{enumerate}
-\item \textbf{Prepare the adjoint sources} \label{enu:Prepare-the-adjoint}
-
-\begin{enumerate}
-\item First, run a regular forward simlation (\texttt{SIMULATION\_TYPE =
-1} and \texttt{SAVE\_FORWARD = .false.}). You can automatically set
-these two variables using the \texttt{\small utils/change\_simulation\_type.pl}{\small{}
-script:}{\small \par}
-
-\begin{lyxcode}
-utils/change\_simulation\_type.pl~-f~
-\end{lyxcode}
-
-and then collect the recorded seismograms at all the stations given
-in \texttt{DATA/STATIONS}.
-
-\item Then select the stations for which you want to compute the time-reversed
-adjoint sources and run the adjoint simulation, and compile them into
-the \texttt{DATA/STATIONS\_ADJOINT} file, which has the same format
-as the regular \texttt{DATA/STATIONS} file.
-
-\begin{itemize}
-\item Depending on what type of misfit function is used for the source inversion,
-adjoint sources need to be computed from the original recorded seismograms
-for the selected stations and saved in a sub-directory called \texttt{SEM/}
-in the root directory of the code, with the format \texttt{NT.STA.BX?.adj}, where \texttt{NT},
-\texttt{STA} are the network code and station name given in the \texttt{DATA/STATIONS\_ADJOINT}
-file, and \texttt{BX?} represents the component name of a particular
-adjoint seismogram. Please note that the band code can change depending
-on your sampling rate (see Appendix~\ref{cha:channel-codes} for
-further details).
-\item The adjoint seismograms are in the same format as the original seismogram
-(\texttt{NT.STA.BX?.sem?}), with the same start time, time interval
-and record length.
-\end{itemize}
-\item Notice that even if you choose to time reverse only one component
-from one specific station, you still need to supply all three components
-because the code is expecting them (you can set the other two components
-to be zero).
-\item Also note that since time-reversal is done in the code itself, no
-explicit time-reversing is needed for the preparation of the adjoint
-sources, i.e., the adjoint sources are in the same forward time sense
-as the original recorded seismograms.
-\end{enumerate}
-\item \textbf{Set the related parameters and run the adjoint simulation}\\
- In the \texttt{DATA/Par\_file}, set the two related parameters to
-be \texttt{SIMULATION\_TYPE = 2} and \texttt{SAVE\_FORWARD = .false.}.
-More conveniently, use the scripts \texttt{utils/change\_simulation\_type.pl}
-to modify the \texttt{Par\_file} automatically (\texttt{change\_simulation\_type.pl
--a}). Then run the solver to launch the adjoint simulation.
-\item \textbf{Collect the seismograms at the original source location}
-
-After the adjoint simulation has completed successfully, collect the
-seismograms from \texttt{LOCAL\_PATH}.
-\begin{itemize}
-\item These adjoint seismograms are recorded at the locations of the original
-earthquake sources given by the \texttt{DATA/CMTSOLUTION} file, and
-have names of the form \texttt{NT.S?????.S??.sem} for the six-component
-strain tensor (\texttt{SNN,SEE,SZZ,SNE,SNZ,SEZ}) at these locations,
-and ~\\
- \texttt{NT.S?????.BX?.sem} for the three-component displacements
-(\texttt{BXN,BXE,BXZ}) recorded at these locations.
-\item \texttt{S?????} denotes the source number; for example, if the original
-\texttt{CMTSOLUTION} provides only a point source, then the seismograms
-collected will start with \texttt{S00001}.
-\item These adjoint seismograms provide critical information for the computation
-of the gradient of the misfit function.
-\end{itemize}
-\end{enumerate}
-
-\section{\label{sec:Adjoint-simulation-finite}Adjoint Simulations for Finite-Frequency
-Kernels (Kernel Simulation)}
-
-Finite-frequency sensitivity kernels are computed in two successive
-simulations (please refer to \citet{LiTr06} and \citet{TrKoLi08}
-for details).
-\begin{enumerate}
-\item \textbf{Run a forward simulation with the state variables saved at
-the end of the simulation}
-
-Prepare the \texttt{\small CMTSOLUTION}{\small{} and }\texttt{\small STATIONS}{\small{}
-files, set the parameters }\texttt{\small SIMULATION\_TYPE}{\small {}
-}\texttt{\small =}{\small {} }\texttt{\small 1}{\small{} and }\texttt{\small SAVE\_FORWARD
-=}{\small {} }\texttt{\small .true.}{\small{} in the }\texttt{\small Par\_file}{\small{}
-(}\texttt{\small change\_simulation\_type -F}{\small ), and run the
-solver.}{\small \par}
-\begin{itemize}
-\item Notice that attenuation is not implemented yet for the computation
-of finite-frequency kernels; therefore set \texttt{ATTENUATION = .false.}
-in the \texttt{Par\_file}.
-\item We also suggest you modify the half duration of the \texttt{CMTSOLUTION}
-to be similar to the accuracy of the simulation (see Equation \ref{eq:shortest_period})
-to avoid too much high-frequency noise in the forward wavefield, although
-theoretically the high-frequency noise should be eliminated when convolved
-with an adjoint wavefield with the proper frequency content.
-\item This forward simulation differs from the regular simulations (\texttt{\small SIMULATION\_TYPE}{\small {}
-}\texttt{\small =}{\small {} }\texttt{\small 1}{\small{} and }\texttt{\small SAVE\_FORWARD}{\small {}
-}\texttt{\small =}{\small {} }\texttt{\small .false.}{\small ) described
-in the previous chapters in that the state variables for the last
-time step of the simulation, including wavefields of the displacement,
-velocity, acceleration, etc., are saved to the }\texttt{\small LOCAL\_PATH}{\small{}
-to be used for the subsequent simulation. }{\small \par}
-\item For regional simulations, the files recording the absorbing boundary
-contribution are also written to the \texttt{LOCAL\_PATH} when \texttt{SAVE\_FORWARD
-= .true.}.
-\end{itemize}
-\item \textbf{Prepare the adjoint sources}
-
-The adjoint sources need to be prepared the same way as described
-in the Section \ref{enu:Prepare-the-adjoint}.
-\begin{itemize}
-\item In the case of travel-time finite-frequency kernel for one source-receiver
-pair, i.e., point source from the \texttt{CMTSOLUTION}, and one station
-in the \texttt{STATIONS\_ADJOINT} list, we supply a sample program
-in \texttt{utils/adjoint\_sources/traveltime/xcreate\_adjsrc\_traveltime}
-to cut a certain portion of the original displacement seismograms
-and convert it into the proper adjoint source to compute the finite-frequency
-kernel.
-
-\begin{lyxcode}
-xcreate\_adjsrc\_traveltime~t1~t2~ifile{[}0-5{]}~E/N/Z-ascii-files~{[}baz{]}
-\end{lyxcode}
-
-where \texttt{t1} and \texttt{t2} are the start and end time of the
-portion you are interested in, \texttt{ifile} denotes the component
-of the seismograms to be used (0 for all three components, 1 for East,
-2 for North, and 3 for vertical, 4 for transverse, and 5 for radial
-component), \texttt{E/N/Z-ascii-files} indicate the three-component
-displacement seismograms in the right order, and \texttt{baz} is the
-back-azimuth of the station. Note that \texttt{baz} is only supplied
-when \texttt{ifile} = 4 or 5.
-
-\item Similarly, a sample program to compute adjoint sources for amplitude
-finite-frequency kernels may be found in \texttt{utils/adjoint\_sources/amplitude}
-and used in the same way as described for traveltime measurements
-
-\begin{lyxcode}
-xcreate\_adjsrc\_amplitude~t1~t2~ifile{[}0-5{]}~E/N/Z-ascii-files~{[}baz{]}.
-\end{lyxcode}
-\end{itemize}
-\item \textbf{Run the kernel simulation}
-
-With the successful forward simulation and the adjoint source ready
-in the \texttt{SEM/} directory, set \texttt{SIMULATION\_TYPE
-= 3} and \texttt{SAVE\_FORWARD = .false.} in the \texttt{Par\_file}
-(you can use \texttt{change\_simulation\_type.pl -b}), and rerun the
-solver.
-\begin{itemize}
-\item The adjoint simulation is launched together with the back reconstruction
-of the original forward wavefield from the state variables saved from
-the previous forward simulation, and the finite-frequency kernels
-are computed by the interaction of the reconstructed forward wavefield
-and the adjoint wavefield.
-\item The back-reconstructed seismograms at the original station locations
-are saved to the \texttt{LOCAL\_PATH} at the end of the kernel simulations,
-and can be collected to the local disk.
-\item These back-constructed seismograms can be compared with the time-reversed
-original seismograms to assess the accuracy of the backward reconstruction,
-and they should match very well.
-\item The arrays for density, P-wave speed and S-wave speed kernels are
-also saved in the \texttt{LOCAL\_PATH} with the names \texttt{proc??????\_rho(alpha,beta)\_kernel.bin},
-where \texttt{proc??????} represents the processor number, \texttt{rho(alpha,beta)}
-are the different types of kernels.
-\end{itemize}
-\item \textbf{Run the anisotropic kernel simulation}
-
-Instead of the kernels for the isotropic wave speeds, you can also
-compute the kernels for the 21 independent components $C_{IJ},\, I,J=1,...,6$
-(using Voigt's notation) of the elastic tensor in the cartesian coordinate
-system. This is done by setting \texttt{ANISOTROPIC\_KL} \texttt{=}
-\texttt{.true.} in \texttt{constants.h} before compiling the package.
-The definition of the parameters $C_{IJ}$ in terms of the corresponding
-components $c_{ijkl},ijkl,i,j,k,l=1,2,3$ of the elastic tensor in
-cartesian coordinates follows \citet{ChTr07}. The 21 anisotropic
-kernels are saved in the \texttt{LOCAL\_PATH} in one file with the
-name of \texttt{proc??????\_cijkl\_kernel.bin} (with \texttt{proc??????}
-the processor number). The output kernels correspond to absolute perturbations
-$\delta C_{IJ}$ of the elastic parameters and their unit is in $s/GPa/km^{3}$.
-For consistency, the output density kernels with this option turned
-on are for a perturbation $\delta\rho$ (and not $\frac{\delta\rho}{\rho}$)
-and their unit is in s / (kg/m$^{3}$) / km$^{3}$. These `primary'
-anisotropic kernels can then be combined to obtain the kernels for
-different parameterizations of anisotropy. This can be done, for example,
-when combining the kernel files from slices into one mesh file (see
-Section~\ref{sec:Finite-Frequency-Kernels}).
-
-If \texttt{ANISOTROPIC\_KL} \texttt{=} \texttt{.true.} by additionally
-setting \texttt{ANISOTROPIC\_KL} \texttt{=} \texttt{.true.} in \texttt{constants.h}
-the package will save anisotropic kernels paramterized as velocities
-related to transverse isotropy based on the the Chen and Tromp parameters
-\citet{ChTr07}. The kernels are saved as relative perturbations for
-horizontal and vertical P and S velocities, $\alpha_{v},\alpha_{h},\beta_{v},\beta_{h}$.
-Explicit relations can be found in appendix B. of \citet{SiLiTrTr07b}
-
-The anisotropic kernels are only currently available for CPU mode.
-
-\end{enumerate}
-In general, the three steps need to be run sequentially to assure
-proper access to the necessary files. If the simulations are run through
-some cluster scheduling system (e.g., LSF), and the forward simulation
-and the subsequent kernel simulations cannot be assigned to the same
-set of computer nodes, the kernel simulation will not be able to access
-the database files saved by the forward simulation. Solutions for
-this dilemma are provided in Chapter~\ref{cha:Scheduler}. Visualization
-of the finite-frequency kernels is discussed in Section~\ref{sec:Finite-Frequency-Kernels}.
-
-%------------------------------------------------------------------------------------------------%
-
-\chapter{\label{cha:tomo}Doing tomography, i.e., updating the model based on the sensitivity kernels obtained}
-
-%------------------------------------------------------------------------------------------------%
-
-UNDER CONSTRUCTION (March 2015). Will be added around April / May 2015.
-
-\section{Tomographic inversion / imaging using the sensitivity kernels obtained}
-
-OLD VERSION OF THE SECTION, WILL SOON BE IMPROVED AND ENRICHED.\\
-
-One of the fundamental reasons for computing sensitivity kernels (Section~\ref{sec:Adjoint-simulation-finite})
-is to use them within a tomographic inversion. In other words, use
-recorded seismograms, make measurements with synthetic seismograms,
-and use the misfit between the two to iteratively improve the model
-described by (at least) $V_{{\rm p}}$, $V_{{\rm s}}$, and $\rho$
-as a function of space.
-
-Whatever misfit function you use for the tomographic inversion (you can find several
-examples in \citet{TrKoLi08} and \citet{TrTaLi05}), you will weight
-the sensitivity kernels with measurements. Furthermore, you will use
-as many measurements (stations, components, time windows) as possible
-per event; hence, we call these composite kernels ``event kernels,''
-which are volumetric fields representing the gradient of the misfit
-function with respect to one of the variables (\eg $V_{{\rm s}}$).
-The basic features of an adjoint-based tomographic inversion were
-illustrated in \citet{TaLiTr07} and \citet{TrKoLi08} using a conjugate-gradient
-algorithm; there are dozens of versions of gradient-based inversion
-algorithms that could alternatively be used. The tomographic inversion
-of \citet{TaLiMaTr09,TaLiMaTr2010} used SPECFEM3D Cartesian as well
-as several additional components which are also stored on the CIG
-svn server, described next.
-
-The directory containing external utilities for tomographic inversion using
-SPECFEM3D Cartesian (or other packages that evaluate misfit functions
-and gradients) is in directory \texttt{utils/ADJOINT\_TOMOGRAPHY\_TOOLS/}:
-\begin{verbatim}
-flexwin/     -- FLEXWIN algorithm for automated picking of time windows
-measure_adj/ -- reads FLEXWIN output file and makes measurements,
-                with the option for computing adjoint sources
-iterate_adj/ -- various tools for iterative inversion
-                (requires pre-computed "event kernels")
-\end{verbatim}
-This directory also contains
-a brief \verb+README+ file indicating the role of the three subdirectories,
-\verb+flexwin+ \citep{Maggi2009}, \verb+measure_adj+, and \verb+iterate_adj+.
-The components for making the model update are there; however, there
-are no explicit rules for computing the model update, just as with
-any optimization problem. There are options for computing a conjugate
-gradient step, as well as a source subspace projection step.
-
-The best single file to read is probably: \verb+ADJOINT_TOMO/iterate_adj/cluster/README+.
-
-%------------------------------------------------------------------------------------------------%
-
-\chapter{Noise Cross-correlation Simulations}
-
-%------------------------------------------------------------------------------------------------%
-
-Besides earthquake simulations, SPECFEM3D Cartesian includes functionality
-for seismic noise tomography as well. In order to proceed successfully
-in this chapter, it is critical that you have already familiarized
-yourself with procedures for meshing (Chapter \ref{cha:Mesh-Generation}),
-creating distributed databases (Chapter \ref{cha:Creating-Distributed-Databases}),
-running earthquake simulations (Chapters \ref{cha:Running-the-Solver})
-and adjoint simulations (Chapter \ref{cha:Adjoint-Simulations}).
-Also, make sure you read the article `Noise cross-correlation sensitivity
-kernels' \citep{trompetal2010}, in order to understand noise simulations
-from a theoretical perspective.
-
-\section{Input Parameter Files}
-
-As usual, the three main input files are crucial: \texttt{Par\_file},
-\texttt{CMTSOLUTION} and \texttt{STATIONS}. Unless otherwise specified,
-those input files should be located in directory \texttt{DATA/}.\\
-
-\texttt{CMTSOLUTION} is required for all simulations. At a first glance,
-it may seem unexpected to have it here, since the noise simulations
-should have nothing to do with the earthquake -- \texttt{CMTSOLUTION}.
-However, for noise simulations, it is critical to have no earthquakes.
-In other words, the moment tensor specified in \texttt{CMTSOLUTION}
-must be set to zero manually!\\
-
-\texttt{STATIONS} remains the same as in previous earthquake simulations,
-except that the order of receivers listed in \texttt{STATIONS} is
-now important. The order will be used to determine the `master' receiver,
-i.e., the one that simultaneously cross correlates with the others.\\
-
-\texttt{Par\_file} also requires careful attention. A parameter called
-\texttt{NOISE\_TOMOGRAPHY} has been added which specifies the type
-of simulation to be run. \texttt{NOISE\_TOMOGRAPHY} is an integer
-with possible values 0, 1, 2 and 3. For example, when \texttt{NOISE\_TOMOGRAPHY}
-equals 0, a regular earthquake simulation will be run. When it is
-1/2/3, you are about to run step 1/2/3 of the noise simulations respectively.
-Should you be confused by the three steps, refer to \citet{trompetal2010}
-for details.\\
-
-Another change to \texttt{Par\_file} involves the parameter \texttt{NSTEP}.
-While for regular earthquake simulations this parameter specifies
-the length of synthetic seismograms generated, for noise simulations
-it specifies the length of the seismograms used to compute cross correlations.
-The actual cross correlations are thus twice this length, i.e., $2~\mathrm{NSTEP}-1$.
-The code automatically makes the modification accordingly, if \texttt{NOISE\_TOMOGRAPHY}
-is not zero.\\
-
-There are other parameters in \texttt{Par\_file} which should be given
-specific values. For instance, since the first two steps for calculating
-noise sensitivity kernels correspond to forward simulations, \texttt{SIMULATION\_TYPE}
-must be 1 when \texttt{NOISE\_TOMOGRAPHY} equals 1 or 2. Also, we
-have to reconstruct the ensemble forward wavefields in adjoint simulations,
-therefore we need to set \texttt{SAVE\_FORWARD} to \texttt{.true.}
-for the second step, i.e., when \texttt{NOISE\_TOMOGRAPHY} equals
-2. The third step is for kernel constructions. Hence \texttt{SIMULATION\_TYPE}
-should be 3, whereas \texttt{SAVE\_FORWARD} must be \texttt{.false.}.\\
-
-Finally, for most system architectures, please make sure that \texttt{LOCAL\_PATH}
-in \texttt{Par\_file} is in fact local, not globally shared. Because
-we have to save the wavefields at the earth's surface at every time
-step, it is quite problematic to have a globally shared \texttt{LOCAL\_PATH},
-in terms of both disk storage and I/O speed.
-
-\section{Noise Simulations: Step by Step}
-
-Proper parameters in those parameter files are not enough for noise
-simulations to run. We have more parameters to specify: for example,
-the ensemble-averaged noise spectrum, the noise distribution etc.
-However, since there are a few `new' files, it is better to introduce
-them sequentially. In this section, standard procedures for noise
-simulations are described.
-
-\subsection{Pre-simulation}
-\begin{itemize}
-\item As usual, we first configure the software package using:
-
-\texttt{./configure FC=ifort MPIFC=mpif90}
-
-Use the following if SCOTCH is needed:
-
-\texttt{./configure FC=ifort MPIFC=mpif90 -{}-with-scotch-dir=/opt/scotch}\\
-
-\item Next, we need to compile the source code using:
-
-\texttt{make xgenerate\_databases}
-
-\texttt{make xspecfem3D} \\
-
-\item Before we can run noise simulations, we have to specify the noise
-statistics, e.g., the ensemble-averaged noise spectrum. Matlab scripts
-are provided to help you to generate the necessary file.
-
-\texttt{EXAMPLES/noise\_tomography/NOISE\_TOMOGRAPHY.m} (main program)\\
- \texttt{EXAMPLES/noise\_tomography/PetersonNoiseModel.m}
-
-In Matlab, simply run:
-
-\texttt{NOISE\_TOMOGRAPHY(NSTEP, DT, Tmin, Tmax, NOISE\_MODEL)}\\
-
-\texttt{DT} is given in \texttt{Par\_file}, but \texttt{NSTEP} is
-NOT the one specified in \texttt{Par\_file}. Instead, you have to
-feed $2~\mathrm{NSTEP}-1$ to account for the doubled length of cross
-correlations. \texttt{Tmin} and \texttt{Tmax} correspond to the period
-range you are interested in, whereas \texttt{NOISE\_MODEL} denotes
-the noise model you will be using. Details can be found in the Matlab
-script.
-
-After running the Matlab script, you will be given the following information
-(plus a figure in Matlab):
-
-\texttt{{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}}\\
- \texttt{the source time function has been saved in:}\\
- \texttt{/data2/yangl/3D\_NOISE/S\_squared} (note this path must be
-different)\\
- \texttt{S\_squared should be put into directory:}\\
- \texttt{OUTPUT\_FILES/NOISE\_TOMOGRAPHY/ in the SPECFEM3D Cartesian
-package}\\
-
-In other words, the Matlab script creates a file called \texttt{S\_squared},
-which is the first `new' input file we encounter for noise simulations.
-
-One may choose a flat noise spectrum rather than Peterson's noise
-model. This can be done easily by modifying the Matlab script a little.
-
-\item Create a new directory in the SPECFEM3D Cartesian package, name it
-as \texttt{OUTPUT\_FILES/NOISE\_TOMOGRAPHY/}. We will add some parameter
-files later in this folder.
-\item Put the Matlab-generated-file \texttt{S\_squared} in \texttt{OUTPUT\_FILES/NOISE\_TOMOGRAPHY/}.
-
-That's to say, you will have a file \texttt{OUTPUT\_FILES/NOISE\_TOMOGRAPHY/S\_squared}
-in the SPECFEM3D Cartesian package.
-
-\item Create a file called \texttt{OUTPUT\_FILES/NOISE\_TOMOGRAPHY/irec\_master\_noise}.
-Note that this file is located in directory \texttt{OUTPUT\_FILES/NOISE\_TOMOGRAPHY/}
-as well. In general, all noise simulation related parameter files
-go into that directory. \texttt{irec\_master\_noise} contains only
-one integer, which is the ID of the `master' receiver. For example,
-if this file contains 5, it means that the fifth receiver listed in
-\texttt{DATA/STATIONS} becomes the `master'. That's why we mentioned
-previously that the order of receivers in \texttt{DATA/STATIONS} is
-important.
-
-Note that in regional simulations, the \texttt{DATA/STATIONS} might
-contain receivers which are outside of our computational domains.
-Therefore, the integer in \texttt{irec\_master\_noise} is actually
-the ID in \texttt{DATA/STATIONS\_FILTERED} (which is generated by
-\texttt{bin/xgenerate\_databases}).
-
-\item Create a file called \texttt{OUTPUT\_FILES/NOISE\_TOMOGRAPHY/nu\_master}.
-This file holds three numbers, forming a (unit) vector. It describes
-which component we are cross-correlating at the `master' receiver,
-i.e., ${\hat{{\bf \nu}}}^{\alpha}$ in \citet{trompetal2010}. The
-three numbers correspond to E/N/Z components respectively. Most often,
-the vertical component is used, and in those cases the three numbers
-should be 0, 0 and 1.
-\item Describe the noise direction and distributions in \texttt{src/shared/noise\_tomography.f90}.
-Search for a subroutine called \texttt{noise\_distribution\_direction}
-in \texttt{noise\_tomography.f90}. It is actually located at the very
-beginning of \texttt{noise\_tomography.f90}. The default assumes vertical
-noise and a uniform distribution across the whole free surface of
-the model. It should be quite self-explanatory for modifications.
-Should you modify this part, you have to re-compile the source code.
-\end{itemize}
-
-\subsection{Simulations}
-
-With all of the above done, we can finally launch our simulations.
-Again, please make sure that the \texttt{LOCAL\_PATH} in \texttt{DATA/Par\_file}
-is not globally shared. It is quite problematic to have a globally
-shared \texttt{LOCAL\_PATH}, in terms of both disk storage and speed
-of I/O (we have to save the wavefields at the earth's surface at every
-time step).
-
-As discussed in \citet{trompetal2010}, it takes three steps/simulations
-to obtain one contribution of the ensemble sensitivity kernels:
-\begin{itemize}
-\item Step 1: simulation for generating wavefields\\
- \texttt{SIMULATION\_TYPE=1}\\
- \texttt{NOISE\_TOMOGRAPHY=1}\\
- \texttt{SAVE\_FORWARD} not used, can be either .true. or .false.
-\item Step 2: simulation for ensemble forward wavefields\\
- \texttt{SIMULATION\_TYPE=1}\\
- \texttt{NOISE\_TOMOGRAPHY=2}\\
- \texttt{SAVE\_FORWARD=.true.}
-\item Step 3: simulation for ensemble adjoint wavefields and sensitivity
-kernels\\
- \texttt{SIMULATION\_TYPE=3}\\
- \texttt{NOISE\_TOMOGRAPHY=3}\\
- \texttt{SAVE\_FORWARD=.false.}
-
-Note Step 3 is an adjoint simulation, please refer to previous chapters
-on how to prepare adjoint sources and other necessary files, as well
-as how adjoint simulations work.
-
-\end{itemize}
-Note that it is better to run the three steps continuously within
-the same job, otherwise you have to collect the saved surface movies
-from the old nodes/CPUs to the new nodes/CPUs. This process varies
-from cluster to cluster and thus cannot be discussed here. Please
-ask your cluster administrator for information/configuration of the
-cluster you are using.\\
-
-\subsection{Post-simulation}
-
-After those simulations, you have all stuff you need, either in the
-\texttt{OUTPUT\_FILES/} or in the directory specified by \texttt{LOCAL\_PATH}
-in \texttt{DATA/Par\_file} (which are most probably on local nodes).
-Collect whatever you want from the local nodes to your workstation,
-and then visualize them. This process is the same as what you may
-have done for regular earthquake simulations. Refer to other chapters
-if you have problems.\\
-
-Simply speaking, two outputs are the most interesting: the simulated
-ensemble cross correlations and one contribution of the ensemble sensitivity
-kernels.\\
-
-The simulated ensemble cross correlations are obtained after the second
-simulation (Step 2). Seismograms in \texttt{OUTPUT\_FILES/} are actually
-the simulated ensemble cross correlations. Collect them immediately
-after Step 2, or the Step 3 will overwrite them. Note that we have
-a `master' receiver specified by \texttt{irec\_master\_noise}, the
-seismogram at one station corresponds to the cross correlation between
-that station and the `master'. Since the seismograms have three components,
-we may obtain cross correlations for different components as well,
-not necessarily the cross correlations between vertical components.\\
-
-One contribution of the ensemble cross-correlation sensitivity kernels
-are obtained after Step 3, stored in the \texttt{DATA/LOCAL\_PATH}
-on local nodes. The ensemble kernel files are named the same as regular
-earthquake kernels.\\
-
-You need to run another three simulations to get the other contribution
-of the ensemble kernels, using different forward and adjoint sources
-given in \citet{trompetal2010}.
-
-\section{Example}
-
-In order to illustrate noise simulations in an easy way, one example
-is provided in \texttt{EXAMPLES/noise\_tomography/}. See \texttt{EXAMPLES/noise\_tomography/README}
-for explanations. \\
-
-Note, however, that they are created for a specific workstation (CLOVER@PRINCETON),
-which has at least 4 cores with `mpif90' working properly. \\
-
-If your workstation is suitable, you can run the example in \texttt{EXAMPLES/noise\_tomography/}
-using:\\
-
-\texttt{./pre-processing.sh}\\
-
-Even if this script does not work on your workstation, the procedure
-it describes is universal. You may review the whole process described
-in the last section by following the commands in \texttt{pre-processing.sh},
-which should contain enough explanations for all the commands.
-
-%------------------------------------------------------------------------------------------------%
-
-\chapter{Gravity integral calculations for the gravity field of an Earth model}
-
-%------------------------------------------------------------------------------------------------%
-
-SPECFEM can now compute the gravity field as well as its derivatives (i.e., gravity gradiometry)
-generated by any given 3D Earth model at a given elevation, i.e., on a given observation surface.
-That feature is still experimental but should work just fine.\\
-
-To see how it is implemented and to use it, type this in the root directory of the code:\\
-
-\texttt{grep -i GRAVITY\_INTEGRALS src/*/*}
-
-%------------------------------------------------------------------------------------------------%
-
-\chapter{Graphics}
-
-%------------------------------------------------------------------------------------------------%
-
-\section{\label{sec:Mesh-graphics}Meshes}
-
-In case you used the internal mesher \texttt{xmeshfem3D} to create
-and partition your mesh, you can output mesh files in ABAQUS (.INP)
-and DX (.dx) format to visualize them. For this, you must set either
-the flag \texttt{CREATE\_DX\_FILES} or \texttt{CREATE\_ABAQUS\_FILES}
-to \texttt{.true.} in the mesher's parameter file \texttt{Mesh\_Par\_file}
-prior to running the mesher (see Chapter~\ref{cha:Running-the-Mesher-Meshfem3D}
-for details). You can then use AVS \urlwithparentheses{www.avs.com}
-or OpenDX \urlwithparentheses{www.opendx.org} to visualize the mesh
-and MPI partition (slices).
-
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=0.49\textwidth]{figures/vtk_mesh_vp} \includegraphics[width=0.49\textwidth]{figures/vtk_mesh_vs}
-\par\end{centering}
-
-\caption{Visualization using Paraview of VTK files created by \texttt{xgenerate\_databases}
-showing P- and S-wave velocities assigned to the mesh points. The
-mesh was created by \texttt{xmeshfem3D} for 4 processors.}
-
-\label{fig:vtk.mesh}
-\end{figure}
-
-You have also the option to visualize the distributed databases produced
-by \texttt{xgenerate\_databases} using Paraview \urlwithparentheses{www.paraview.org}.
-For this, you must set the flag \texttt{SAVE\_MESH\_FILES} to \texttt{.true.}
-in the main parameter file \texttt{Par\_file} (see Chapter~\ref{cha:Main-Parameter}
-for details). This will create VTK files for each single partition.
-You can then use Paraview \urlwithparentheses{www.paraview.org}
-to visualized these partitions.
-
-\section{\label{sec:Movies}Movies}
-
-To make a surface or volume movie of the simulation, set parameters
-\texttt{MOVIE\_SURFACE}, \texttt{MOVIE\_VOLUME}, \texttt{MOVIE\_TYPE},
-and \texttt{NTSTEP\_BETWEEN\_FRAMES} in the \texttt{Par\_file}. Turning
-on the movie flags, in particular \texttt{MOVIE\_VOLUME}, produces
-large output files. \texttt{MOVIE\_VOLUME} files are saved in the
-\texttt{LOCAL\_PATH} directory, whereas \texttt{MOVIE\_SURFACE} output
-files are saved in the \texttt{OUTPUT\_FILES} directory. We save the
-displacement field if the parameter \texttt{SAVE\_DISPLACEMENT} is
-set, otherwise the velocity field is saved. The look of a movie is
-determined by the half-duration of the source. The half-duration should
-be large enough so that the movie does not contain frequencies that
-are not resolved by the mesh, i.e., it should not contain numerical
-noise. This can be accomplished by selecting a CMT \texttt{HALF\_DURATION}
-> 1.1 $\times$ smallest period (see figure \ref{fig:CMTSOLUTION-file}).
-When \texttt{MOVIE\_SURFACE} = .\texttt{true.}, the half duration
-of each source in the \texttt{CMTSOLUTION} file is replaced by
-\begin{quote}
-\[
-\sqrt{(}\mathrm{\mathtt{HALF\_DURATIO}\mathtt{N}^{2}}+\mathrm{\mathtt{HDUR\_MOVI}\mathtt{E}^{2}})
-\]
- \textbf{NOTE:} If \texttt{HDUR\_MOVIE} is set to 0.0, the code will
-select the appropriate value of 1.1 $\times$ smallest period. As
-usual, for a point source one can set \texttt{half duration} in the
-\texttt{CMTSOLUTION} file to be 0.0 and \texttt{HDUR\_MOVIE} = 0.0
-in the \texttt{Par\_file} to get the highest frequencies resolved
-by the simulation, but for a finite source one would keep all the
-\texttt{half durations} as prescribed by the finite source model and
-set \texttt{HDUR\_MOVIE} = 0.0.
-\end{quote}
-
-\subsection{Movie Surface}
-
-When running \texttt{xspecfem3D} with the \texttt{MOVIE\_SURFACE}
-flag turned on, the code outputs \texttt{moviedata??????} files in
-the \texttt{OUTPUT\_FILES} directory. There are several flags in the
-main parameter file \texttt{Par\_file} that control the output of
-these movie data files (see section \ref{cha:Main-Parameter} for
-details): \texttt{\small NTSTEP\_BETWEEN\_FRAMES}{\small{} to set the
-timesteps between frames, }\texttt{\small SAVE\_DISPLACEMENT}{\small{}
-to save displacement instead of velocity, }\texttt{\small USE\_HIGHRES\_FOR\_MOVIES}{\small{}
-to save values at all GLL point instead of element edges. In order
-to output additionally shakemaps, you would set the parameter }\texttt{\small CREATE\_SHAKEMAP}{\small{}
-to }\texttt{\small .true.}{\small .}{\small \par}
-
-The files are in a fairly complicated binary format, but there is
-a program provided to convert the output into more user friendly formats:
-\begin{description}
-\item [{\texttt{xcreate\_movie\_shakemap\_AVS\_DX\_GMT}}] From \texttt{create\_movie\_shakemap\_AVS\_DX\_GMT.f90},
-it outputs data in ASCII, OpenDX, or AVS format (also readable in
-ParaView). Before compiling the code, make sure you have the file
-\texttt{surface\_from\_mesher.h} in the \texttt{OUTPUT\_FILES/} directory.
-This file will be created by the solver run. Then type
-
-\begin{lyxcode}
-{\footnotesize make~xcreate\_movie\_shakemap\_AVS\_DX\_GMT}{\footnotesize \par}
-\end{lyxcode}
-
-and run the executable \texttt{xcreate\_movie\_shakemap\_AVS\_DX\_GMT}
-in the main directory. It will create visualization files
-in your format of choice. The code will prompt the user for input
-parameters.
-
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=0.32\textwidth]{figures/movie_surf_1} \includegraphics[width=0.32\textwidth]{figures/movie_surf_2}
-\includegraphics[width=0.32\textwidth]{figures/movie_surf_3}
-\par\end{centering}
-
-\caption{Visualization using AVS files created by \texttt{xcreate\_movie\_shakemap\_AVS\_DX\_GMT}
-showing movie snapshots of vertical velocity components at different
-times.}
-
-\label{fig:movie.surf}
-\end{figure}
-
-\end{description}
-The \texttt{SPECFEM3D Cartesian} code is running in near real-time
-to produce animations of southern California earthquakes via the web;
-see Southern California ShakeMovie\textregistered{}\urlwithparentheses{www.shakemovie.caltech.edu}.
-
-\subsection{\label{sub:Movie-Volume}Movie Volume}
-
-When running xspecfem3D with the \texttt{\small MOVIE\_VOLUME}{\small{}
-flag turned on, the code outputs several files in }\texttt{\small LOCAL\_PATH}{\small{}
-specified in the main }\texttt{\small Par\_file}{\small , e.g. in
-directory }\texttt{\small OUTPUT\_FILES/DATABASES\_MPI}{\small .
-The output is saved by each processor at the time interval specified
-by }\texttt{\small NTSTEP\_BETWEEN\_FRAMES}{\small . For all domains,
-the velocity field is output to files: }\texttt{\small proc??????\_velocity\_X\_it??????.bin}{\small ,
-}\texttt{\small proc??????\_velocity\_Y\_it??????.bin}{\small{} and
-}\texttt{\small proc??????\_velocity\_Z\_it??????.bin}{\small . For
-elastic domains, the divergence and curl taken from the velocity field,
-i.e. $\nabla\cdot{\bf {v}}$ and $\nabla\times{\bf {v}}$, get stored
-as well: }\texttt{\small proc??????\_div\_it??????.bin}{\small , }\texttt{\small proc??????\_curl\_X\_t??????.bin}{\small ,
-}\texttt{\small proc??????\_curl\_Y\_it??????.bin}{\small{} and }\texttt{\small proc??????\_curl\_Z\_it??????.bin}{\small .
-The files denoted }\texttt{\small proc??????\_div\_}{\small ~}\\
-{\small{} }\texttt{\small glob\_it??????.bin}{\small{} and }\texttt{\small proc??????\_curl\_glob\_it??????.bin}{\small{}
-are stored on the global points only, all the other arrays are stored
-on all GLL points. Note that the components X/Y/Z can change to E/N/Z
-according to the }\texttt{\small SUPPRESS\_UTM\_PROJECTION}{\small{}
-flag (see also Appendix~\ref{cha:Coordinates} and \ref{cha:channel-codes}).}{\small \par}
-
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=0.32\textwidth]{figures/movie_volume_1} \includegraphics[width=0.32\textwidth]{figures/movie_volume_2}
-\includegraphics[width=0.32\textwidth]{figures/movie_volume_3}
-\par\end{centering}
-
-\caption{Paraview visualization using movie volume files (converted by \texttt{xcombine\_vol\_data}
-and \texttt{mesh2vtu.pl}) and showing snapshots of vertical velocity
-components at different times.}
-
-\label{fig:movie.volume}
-\end{figure}
-
-To visualize these files, we use an auxilliary program \texttt{combine\_vol\_data.f90}
-to combine the data from all slices into one mesh file. To compile
-it in the root directory, type:
-\begin{lyxcode}
-{\footnotesize make~xcombine\_vol\_data~}{\footnotesize \par}
-\end{lyxcode}
-which will create the executable \texttt{bin/xcombine\_vol\_data}.
-To output the usage of this executable, type
-\begin{lyxcode}
-{\footnotesize ./bin/xcombine\_vol\_data~}{\footnotesize \par}
-\end{lyxcode}
-without arguments.
-
-\medskip
-
-\texttt{xcombine\_vol\_data} will combine the data on the different processors (located in \texttt{OUTPUT\_FILES/DATABASES\_MPI}), for a given quantity and a given iteration, to one file.
-For example, if you want to combine \texttt{velocity\_Z}, for the iteration 400 on 4 processors, i.e.
-if you want to combine these files :
-
-\medskip
-
-\texttt{proc000000\_velocity\_Z\_it000400.bin }
-
-\texttt{proc000001\_velocity\_Z\_it000400.bin }
-
-\texttt{proc000002\_velocity\_Z\_it000400.bin }
-
-\texttt{proc000003\_velocity\_Z\_it000400.bin}
-
-\medskip
-
-You have to go in the directory of the concerned example (where are the directories \texttt{DATA}, \texttt{OUTPUT\_FILES}, etc.).
-Then, you can launch \texttt{xcombine\_vol\_data}, specifying the path where it is located.
-As an example, if the \texttt{DATA} and \texttt{OUTPUT\_FILES} directories of the concerned example are in the root specfem3d directory, \texttt{xcombine\_vol\_data} is in \texttt{./bin}, and you have to type:
-
-\begin{lyxcode}
-~{\footnotesize ./bin/xcombine\_vol\_data~0~3~velocity\_Z\_it000400~./OUTPUT\_FILES/DATABASES\_MPI~./OUTPUT\_FILES~0~}{\footnotesize \par}
-\end{lyxcode}
-
-Here, $0$ is the number of the first processor, $3$ the number of the last one, \texttt{velocity\_Z\_it000400} the name of the files we want to combine without the prefix \texttt{proc000000*\_}, \texttt{./OUTPUT\_FILES/DATABASES\_MPI} the directory where the files are located, \texttt{./OUTPUT\_FILES} the directory where the combined file will be stored, and $0$ is the parameter to create a low-resolution mesh file ($1$ for a high-resolution).
-
-\medskip
-
-At the beginning of the source file \texttt{combine\_vol\_data.f90} (located in \texttt{src/auxiliaries/}), there is a logical constant parameter \texttt{USE\_VTK\_OUTPUT} whose default value is \texttt{.true.}, that which will provide the output combined file directly in the VTK format and you can use Paraview to visualize it.
-
-\medskip
-
-If the value of this parameter is \texttt{.false.}, the output mesh file will have the name \texttt{velocity\_Z\_it000400.mesh}. Then we next have to convert the
-\texttt{.mesh} file into the VTU (Unstructured grid file) format which can be viewed in ParaView. For this task, you can use and modify the
-script \texttt{mesh2vtu.pl} located in directory \texttt{utils/Visualization/Paraview/}, for example:
-\begin{lyxcode}
-{\footnotesize mesh2vtu.pl~-i~velocity\_Z\_it000400.mesh~-o~velocity\_Z\_it000400.vtu}{\footnotesize \par}
-\end{lyxcode}
-Notice that this Perl script uses a program \texttt{mesh2vtu} in the \texttt{utils/Visualization/Paraview/mesh2vtu} directory, which further
-uses the VTK \urlwithparentheses{http://www.vtk.org/} run-time library for its execution. Therefore, make sure you have them properly set
-in the script according to your system.
-
-\bigskip
-
-Then, to do a movie with several iterations, you have to repeat this process for each iteration you want to put in your movie.
-
-\section{\label{sec:Finite-Frequency-Kernels}Finite-Frequency Kernels}
-
-The finite-frequency kernels computed as explained in Section \ref{sec:Adjoint-simulation-finite}
-are saved in the \texttt{LOCAL\_PATH} at the end of the simulation.
-Therefore, we first need to collect these files on the front end,
-combine them into one mesh file, and visualize them with some auxilliary
-programs.
-\begin{enumerate}
-\item \textbf{Create slice files}
-
-We will only discuss the case of one source-receiver pair, i.e., the
-so-called banana-doughnut kernels. Although it is possible to collect
-the kernel files from all slices on the front end, it usually takes
-up too much storage space (at least tens of gigabytes). Since the
-sensitivity kernels are the strongest along the source-receiver great
-circle path, it is sufficient to collect only the slices that are
-along or close to the great circle path.
-
-A Perl script \texttt{slice\_number.pl} located in directory \texttt{utils/Visualization/Paraview/}
-can help to figure out the slice numbers that lie along the great
-circle path. It applies to meshes created with the internal mesher
-\texttt{xmeshfem3D}.
-\begin{enumerate}
-\item On machines where you have access to the script, copy the \texttt{Mesh\_Par\_file},
-and \texttt{output\_solver} files, and run:
-
-\begin{lyxcode}
-{\small slice\_number.pl~Mesh\_Par\_file~output\_solver.txt~slice\_file}{\small \par}
-\end{lyxcode}
-
-which will generate a \texttt{slices\_file}.
-
-\item For cases with multiple sources and multiple receivers, you need to
-provide a slice file before proceeding to the next step.
-\end{enumerate}
-\item \textbf{Collect the kernel files}
-
-After obtaining the slice files, you can collect the corresponding
-kernel files from the given slices.
-\begin{enumerate}
-\item You can use or modify the script \texttt{utils/copy\_basin\_database.pl}
-to accomplish this:
-
-\begin{lyxcode}
-{\small utils/copy\_database.pl~slice\_file~lsf\_machine\_file~filename~{[}jobid{]}}{\small \par}
-\end{lyxcode}
-
-where \texttt{\small lsf\_machine\_file}{\small{} is the machine file
-generated by the LSF scheduler, }\texttt{\small filename}{\small{}
-is the kernel name (e.g., }\texttt{\small rho\_kernel}{\small , }\texttt{\small alpha\_kernel}{\small{}
-and }\texttt{\small beta\_kernel}{\small ), and the optional }\texttt{\small jobid}{\small{}
-is the name of the subdirectory under }\texttt{\small LOCAL\_PATH}{\small{}
-where all the kernel files are stored.}{\small \par}
-
-\item After executing this script, all the necessary mesh topology files
-as well as the kernel array files are collected to the local directory
-of the front end.
-\end{enumerate}
-\item \textbf{Combine kernel files into one mesh file}
-
-We use an auxilliary program \texttt{combine\_vol\_data.f90} to combine
-the kernel files from all slices into one mesh file.
-\begin{enumerate}
-\item Compile it in the root directory:
-
-\begin{lyxcode}
-{\footnotesize make~xcombine\_vol\_data~}{\footnotesize \par}
-
-{\footnotesize ./bin/xcombine\_vol\_data~slice\_list~filename~input\_dir~output\_dir~high/low-resolution~}{\footnotesize \par}
-\end{lyxcode}
-
-where \texttt{input\_dir} is the directory where all the individual
-kernel files are stored, and \texttt{output\_dir} is where the mesh
-file will be written.
-
-\item Use 1 for a high-resolution mesh, outputting all the GLL points to
-the mesh file, or use 0 for low resolution, outputting only the corner
-points of the elements to the mesh file.
-\item The output mesh file will have the name \texttt{filename\_rho(alpha,beta).mesh}
-\end{enumerate}
-\item \textbf{Convert mesh files into .vtu files}
-
-\begin{enumerate}
-\item We next convert the \texttt{.mesh} file into the VTU (Unstructured
-grid file) format which can be viewed in ParaView. For this task,
-you can use and modify the script \texttt{mesh2vtu.pl} located in
-directory \texttt{utils/Visualization/Paraview/}, for example:
-
-\begin{lyxcode}
-{\footnotesize mesh2vtu.pl~-i~file.mesh~-o~file.vtu~}{\footnotesize \par}
-\end{lyxcode}
-\item Notice that this Perl script uses a program \texttt{mesh2vtu} in the
-\texttt{utils/Visualization/Paraview/mesh2vtu} directory, which further
-uses the VTK \urlwithparentheses{http://www.vtk.org/} run-time library
-for its execution. Therefore, make sure you have them properly set
-in the script according to your system.
-\end{enumerate}
-\item \textbf{Copy over the source and receiver .vtk file}
-
-In the case of a single source and a single receiver, the simulation
-also generates the file \texttt{sr.vtk} located in the \texttt{OUTPUT\_FILES/}
-directory to describe the source and receiver locations, which can
-also be viewed in Paraview in the next step.
-
-\item \textbf{View the mesh in ParaView}
-
-Finally, we can view the mesh in ParaView \urlwithparentheses{www.paraview.org}.
-\begin{enumerate}
-\item Open ParaView.
-\item From the top menu, \textsf{File} $\rightarrow$\textsf{Open data},
-select \texttt{file.vtu}, and click the \textsf{Accept} button.
-
-\begin{itemize}
-\item If the mesh file is of moderate size, it shows up on the screen; otherwise,
-only the bounding box is shown.
-\end{itemize}
-\item Click \textsf{Display Tab} $\rightarrow$ \textsf{Display Style} $\rightarrow$
-\textsf{Representation} and select \textsf{wireframe of surface} to
-display it.
-\item To create a cross-section of the volumetric mesh, choose \textsf{Filter}
-$\rightarrow$ \textsf{cut}, and under \textsf{Parameters Tab}, choose
-\textsf{Cut Function} $\rightarrow$ \textsf{plane}.
-\item Fill in center and normal information given by the \texttt{global\_slice\_number.pl}
-script (either from the standard output or from \texttt{normal\_plane.txt}
-file).
-\item To change the color scale, go to \textsf{Display Tab} $\rightarrow$
-\textsf{Color} $\rightarrow$ \textsf{Edit Color Map} and reselect
-lower and upper limits, or change the color scheme.
-\item Now load in the source and receiver location file by \textsf{File}
-$\rightarrow$ \textsf{Open data}, select \texttt{sr.vt}k, and click
-the \textsf{Accept} button. Choose \textsf{Filter} $\rightarrow$
-\textsf{Glyph}, and represent the points by `\textsf{spheres}'.
-\item For more information about ParaView, see the ParaView Users Guide
-\urlwithparentheses{www.paraview.org/files/v1.6/ParaViewUsersGuide.PDF}.
-\end{enumerate}
-\end{enumerate}
-\begin{figure}[H]
-\noindent \begin{centering}
-\includegraphics[scale=0.6]{figures/3D-S-Kernel}
-\par\end{centering}
-
-\caption{(a) Top Panel: Vertical source-receiver cross-section of the S-wave
-finite-frequency sensitivity kernel $K_{\beta}$ for station GSC at
-an epicentral distance of 176 km from the September 3, 2002, Yorba
-Linda earthquake. Lower Panel: Vertical source-receiver cross-section
-of the 3D S-wave speed model used for the spectral-element simulations
-\citep{KoLiTrSuStSh04}. (b) The same as (a) but for station HEC at
-an epicentral distance of 165 km \citep{LiTr06}.}
-
-\label{figure:P-wave-speed-finite-frequency}
-\end{figure}
-
-\chapter{\label{cha:Scheduler}Running through a Scheduler}
-
-The code is usually run on large parallel machines, often PC clusters,
-most of which use schedulers, i.e., queuing or batch management systems
-to manage the running of jobs from a large number of users. The following
-considerations need to be taken into account when running on a system
-that uses a scheduler:
-\begin{itemize}
-\item The processors/nodes to be used for each run are assigned dynamically
-by the scheduler, based on availability. Therefore, in order for the
-\texttt{xgenerate\_databases} and the \texttt{xspecfem3D} executables
-(or between successive runs of the solver) to have access to the same
-database files (if they are stored on hard drives local to the nodes
-on which the code is run), they must be launched in sequence as a
-single job.
-\item On some systems, the nodes to which running jobs are assigned are
-not configured for compilation. It may therefore be necessary to pre-compile
-both the \texttt{xgenerate\_databases} and the \texttt{xspecfem3D}
-executables.
-\item One feature of schedulers/queuing systems is that they allow submission
-of multiple jobs in a ``launch and forget'' mode. In order to take
-advantage of this property, care needs to be taken that output and
-intermediate files from separate jobs do not overwrite each other,
-or otherwise interfere with other running jobs.
-\end{itemize}
-Examples of job scripts can be found in the \texttt{\small utils/Cluster/}{\small{}
-directory and can straightforwardly be modified and adapted to meet
-more specific running needs.}{\small \par}
-
-\chapter{{\normalsize \label{cha:-Changing-the}} Changing the Model}
-
-In this section we explain how to change the velocity model used for
-your simulations. These changes involve contributing specific subroutines
-that replace existing subroutines in the \texttt{SPECFEM3D Cartesian}
-package. Note that \texttt{SPECFEM3D Cartesian} can handle Earth models
-with material properties that vary within each spectral element.
-
-\section{{\normalsize \label{sec:Using-tomographic}}Using external tomographic Earth models}
-
-To implement your own external tomographic model(s), you must provide
-your own external tomography file(s), and choose between two possible
-options:\\
- \indent (1) set in the \texttt{Par\_file} the model parameter \texttt{MODEL
-= tomo}, \\
- or for more user control: \\
- \indent (2) set in the \texttt{Par\_file} the model parameter \texttt{MODEL
-= default}, define the negative \texttt{material\_ID} identifier for
-each element in the file \texttt{MESH/materials\_file} and use the
-following format in the file \texttt{MESH/nummaterial\_velocity\_file}
-when using CUBIT to construct your mesh (see also section \ref{subsec:Exporting-the-Mesh}):
-\begin{lyxcode}
-domain\_ID~material\_ID~tomography~elastic~file\_name~1
-\end{lyxcode}
-where: \\
- \indent \texttt{domain\_ID} is 1 for acoustic or 2 for elastic materials,
-\\
- \indent \texttt{material\_ID} a negative, unique identifier (i.e.,
--1,-2,...), \\
- \indent \texttt{tomography} keyword for tomographic material definition,
-\\
- \indent \texttt{elastic} keyword for elastic material definition,
-\\
- \indent \texttt{file\_name} the name of the tomography file and
-1 a positive unique identifier.\\
-
-The external tomographic model is represented by a grid of points
-with assigned material properties and homogeneous resolution along
-each spatial direction x, y and z. The ASCII file \texttt{file\_name}
-that describe the tomography should be located in the \texttt{TOMOGRAPHY\_PATH}
-directory, set in the \texttt{Par\_file}. The format of the file,
-as read from \texttt{model\_tomography.f90} located in the \texttt{src/generate\_databases}
-directory, looks like Figure \ref{fig:tomography_file}, where
-\begin{description}
-\item [{\texttt{ORIG\_X}, \texttt{END\_X}}] are, respectively, the coordinates
-of the initial and final tomographic grid points along the x direction
-(in the mesh units, e.g., $m$);
-\item [{\texttt{ORIG\_Y}, \texttt{END\_Y}}] respectively the coordinates
-of the initial and final tomographic grid points along the y direction
-(in the mesh units, e.g., $m$);
-\item [{\texttt{ORIG\_Z}, \texttt{END\_Z}}] respectively the coordinates
-of the initial and final tomographic grid points along the z direction
-(in the mesh units, e.g., $m$);
-\item [{\texttt{SPACING\_X}, \texttt{SPACING\_Y}, \texttt{SPACING\_Z}}] the
-spacing between the tomographic grid points along the x, y and z directions,
-respectively (in the mesh units, e.g., $m$);
-\item [{\texttt{NX}, \texttt{NY}, \texttt{NZ}}] the number of grid points
-along the spatial directions x, y and z, respectively; \texttt{NX}
-is given by {[}(\texttt{END\_X} - \texttt{ORIG\_X})/\texttt{SPACING\_X}{]}+1;
-\texttt{NY} and \texttt{NZ} are the same as \texttt{NX}, but for the
-y and z directions, respectively;
-\item [{\texttt{VP\_MIN}, \texttt{VP\_MAX}, \texttt{VS\_MIN}, \texttt{VS\_MAX}, \texttt{RHO\_MIN}, \texttt{RHO\_MAX}}] the
-minimum and maximum values of the wave speed \texttt{vp} and \texttt{vs}
-(in $m\, s^{-1}$) and of the density \texttt{rho} (in $kg\, m^{-3}$);
-these values could be the actual limits of the tomographic parameters
-in the grid or the minimum and maximum values to which we force the
-cut of velocity and density in the model.
-\end{description}
-After the first four lines, in the file \texttt{file\_name} the tomographic
-grid points are listed with the corresponding values of \texttt{vp},
-\texttt{vs} and \texttt{rho}, scanning the grid along the x coordinate
-(from \texttt{ORIG\_X} to \texttt{END\_X} with step of \texttt{SPACING\_X})
-for each given y (from \texttt{ORIG\_Y} to \texttt{END\_Y}, with step
-of \texttt{SPACING\_Y}) and each given z (from \texttt{ORIG\_Z} to
-\texttt{END\_Z}, with step of \texttt{SPACING\_Z}). %fig
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=1\textwidth]{figures/tomo_file}
-\par\end{centering}
-
-\caption{Tomography file \texttt{file\_name} that describes an external Earth
-model. The coordinates x, y and z of the grid, their limits ORIG\_X,
-ORIG\_Y, ORIG\_Z, END\_X, END\_Y, END\_Z and the grid spacings SPACING\_X,
-SPACING\_Y, SPACING\_Z should be in the units of the constructed mesh
-(e.g., UTM coordinates in meters).}
-
-\label{fig:tomography_file}
-\end{figure}
-
-The user can implement his own interpolation algorithm for the tomography
-model by changing the routine \texttt{model\_tomography.f90} located
-in the \texttt{src/generate\_databases/} directory. Moreover, for
-models that involve both fully defined materials and a tomography
-description, the \texttt{nummaterial\_velocity\_file} has multiple
-lines each with the corresponding suitable format described above.
-
-%==============
-
-\subsubsection*{Example: External tomography file with variable discretization intervals}
-
-The example in Figure~\ref{fig:tomography_file} is for an external tomography file with uniform spacing in the $x$, $y$, and $z$ directions.
-(Note that $\Delta x$, $\Delta y$, and $\Delta z$ need not be the same as in the example.) In that example, the  \texttt{nummaterial\_velocity\_file} is
-%
-\begin{lyxcode}
-2  -1 tomography elastic tomography\_model.xyz 1
-\end{lyxcode}
-%
-and the file \texttt{tomography\_model.xyz} will need to reside in \texttt{DATA/}.
-All entries in the second column of \texttt{materials\_file} will be \texttt{-1}, which means that each element in the mesh will be interpolated accordind to the values in \texttt{tomography\_model.xyz}.
-
-In some cases it may be desirable to have an external tomography model that is described in more than one file.
-For example, in cases like southern California, the length scale of variation in the structure of the wave speed model is much shorter in the sedimentary basin models within the upper 15~km.
-Therefore one might want to use an external tomography file that is sampled with, say, $\Delta x = 1000$~m, $\Delta y = 1000$~m, and $\Delta z = 250$~m in the uppermost 15~km, and then use $\Delta x = 2000$~m, $\Delta y = 2000$~m, and $\Delta z = 1000$~m below a depth of 15~km.
-If these intervals are chosen appropriately, then it will result in a pair of external tomography file that is much smaller than the alternative of having a single file with the fine discretization.
-In this case  \texttt{nummaterial\_velocity\_file} is
-%
-\begin{lyxcode}
-2  -1 tomography elastic file\_above\_15km.xyz 1 \\
-2  -2 tomography elastic file\_below\_15km.xyz 1
-\end{lyxcode}
-%
-and the files \texttt{file\_above\_15km.xyz} and \texttt{file\_below\_15km.xyz} will need to reside in \texttt{DATA/}.
-All entries in the second column of \texttt{materials\_file} will need to be either \texttt{-1} or \texttt{-2}, depending on whether the element is above or below 15~km depth.
-In this sense, the construction of the mesh and the discretization of the external model will generally be done in tandem.
-
-Other more complicated discretizations can be used following the same procedure. (In particular, there is no limit on the number of different external files that are used.)
-
-%=============
-
-\section{{\normalsize \label{sec:Anelastic-Models}}External (an)elastic Models}
-
-To use your own external model, you can set in the \texttt{Par\_file}
-the model parameter \texttt{MODEL = external}. Three-dimensional acoustic
-and/or (an)elastic (attenuation) models may then be superimposed onto
-the mesh based upon your subroutine in \texttt{model\_external\_values.f90}
-located in directory \texttt{src/generate\_databases}. The call to
-this routine would be as follows
-\begin{lyxcode}
-call~model\_external\_values(xmesh,~ymesh,~zmesh,~~\&~~\\
-~~~~~~~~~~~rho,~vp,~vs,~qmu\_atten,~iflag\_aniso,~idomain\_id)
-\end{lyxcode}
-Input to this routine consists of:
-\begin{description}
-\item [{\texttt{xmesh,ymesh,zmesh}}] location of mesh point
-\end{description}
-Output to this routine consists of:
-\begin{description}
-\item [{\texttt{rho,vp,vs}}] isotropic model parameters for density $\rho$
-($kg/m^{3}$), $v_{p}$ ($m/s$) and $v_{s}$ ($m/s$)
-\item [{\texttt{qmu\_atten}}] Shear wave quality factor: $0<Q_{\mu}<9000$
-\item [{\texttt{iflag\_aniso}}] anisotropic model flag, $0$ indicating
-no anisotropy or $1$ using anisotropic model parameters as defined
-in routine file \texttt{model\_aniso.f90}
-\item [{\texttt{idomain\_id}}] domain identifier, $1$ for acoustic, $2$
-for elastic, $3$ for poroelastic materials.
-\end{description}
-Note that the resolution and maximum value of anelastic models are
-truncated. This speeds the construction of the standard linear solids
-during the meshing stage. To change the resolution, currently at one
-significant figure following the decimal, or the maximum value (9000),
-consult \texttt{shared/constants.h}.
-
-\section{{\normalsize \label{sec:Anisotropic-Models}}Anisotropic Models}
-
-To use your anisotropic models, you can either set in the \texttt{Par\_file}
-the model parameter \texttt{ANISOTROPY} to \texttt{.true.} or set
-the model parameter \texttt{MODEL = aniso}. Three-dimensional anisotropic
-models may be superimposed on the mesh based upon the subroutine in
-file \texttt{model\_aniso.f90} located in directory \texttt{src/generate\_databases}.
-The call to this subroutine is of the form
-\begin{lyxcode}
-call~model\_aniso(iflag\_aniso,~rho,~vp,~vs,~\&~~\\
-~~~~~~~~~~~c11,c12,c13,c14,c15,c16,c22,c23,c24,c25,c26,~\&~~\\
-~~~~~~~~~~~c33,c34,c35,c36,c44,c45,c46,c55,c56,c66)~
-\end{lyxcode}
-Input to this routine consists of:
-\begin{description}
-\item [{\texttt{iflag\_aniso}}] flag indicating the type of the anisotropic
-model, i.e. $0$ for no superimposed anisotropy, $1$ or $2$ for
-generic pre-defined anisotropic models.
-\item [{\texttt{rho,vp,vs}}] reference isotropic model parameters for density
-$\rho$, $v_{p}$ and $v_{s}$.
-\end{description}
-Output from the routine consists of the following non-dimensional
-model parameters:
-\begin{description}
-\item [{\texttt{c11},}] \textbf{$\cdots$,} \texttt{\textbf{c66}} 21~dimensionalized
-anisotropic elastic parameters.
-\end{description}
-You can replace the \texttt{model\_aniso.f90} file by your own version
-\textit{provided you do not change the call structure of the routine},
-i.e., the new routine should take exactly the same input and produce
-the required relative output.
-
-\section{{\normalsize \label{sec:Using-SEP}}Using external SEP models.}
-
-Stanford Exploration Project (SEP) models consists of two files per variables, one header file (\texttt{.H})
-and one binary file (\texttt{.H$@$}).
-Among other information, the header file indicates, for each dimension,
-the offset (\texttt{o1}, \texttt{o2} and \texttt{o3} fields),
-the spatial step size (\texttt{d1}, \texttt{d2} and \texttt{d3} fields)
-and the number of samples (\texttt{n1}, \texttt{n2} and \texttt{n3}) necessary to read
-the linked binary file (\texttt{in} field).
-
-SEP model reading occurs during database generation. It it the user responsibility
-to ensure that proper interfaces between acoustic and elastic domains are set up
-during the meshing phase. For more information, refer to
-section~\ref{cha:Running-the-Mesher-Meshfem3D}.
-
-In \texttt{DATA/Par\_file}, \texttt{\textbf{MODEL}} should be set to \texttt{\textbf{sep}},
-and \texttt{\textbf{SEP\_MODEL\_DIRECTORY}} should point to wherever your sep model files reside.
-
-Note that in order not to overload the parameter file, SEP header files
-should be named \texttt{vp.H}, \texttt{vs.H} and \texttt{rho.H}. The \texttt{in} field
-in these binary files can be any path as long as it is relative to
-the previously set \texttt{SEP\_MODEL\_DIRECTORY}.
-We also assume that any dimensional value inside SEP header files is given in meters and
-that binary files are single precision floating point numbers stored in little endian format.
-There is currently only support for \texttt{vp}, \texttt{vs} and \texttt{rho} variables.
-
-An example demonstrating the use SEP models is given in:
-\texttt{EXAMPLES/meshfem3D\_examples/sep\_bathymetry}.
-
 \chapter{Post-Processing Scripts}
@@
 
-\section{Process Data and Synthetics\label{sec:Process-data-and-syn}}
+\section{Process Data and Synthetics}\label{sec:Process-data-and-syn}
 
@@
 simulation, and \texttt{150} is the number of processors in total.
-
-%============================================================
-
-\chapter{\label{cha:developers}Information for developers of the code, and for people who want to learn how the technique works}
-
-You can get a very simple 1D version of a demo code (there is one in Fortran and one in Python):\\
-
-git clone --recursive https://github.com/geodynamics/specfem1d.git\\
-
-\noindent We also have simple 3D demo source codes that implement the SEM in a single, small program, in directory\\
-utils/small\_SEM\_solvers\_in\_Fortran\_and\_C\_without\_MPI\_to\_learn of the specfem3d package.
-They are useful to learn how the spectral-element method works, and how to write or modify a code to implement it.
-Also useful to test new ideas by modifying these simple codes to run some tests.\\
-
-\noindent For information on how to contribute to the code, i.e., for how to make your modifications, additions or improvements part of the
-official package, see https://github.com/geodynamics/specfem3d/wiki .
-
-%============================================================
-
-\chapter*{\label{cha:Bug-Reports-and}Bug Reports and Suggestions for Improvements}
-
-To report bugs or suggest improvements to the code, please send an
-e-mail to the CIG Computational Seismology Mailing List \urlwithparentheses{cig-seismo@geodynamics.org}.
-
-\addcontentsline{toc}{chapter}{Notes \& Acknowledgments}
-
-\chapter*{Notes \& Acknowledgments}
-
-In order to keep the software package thread-safe in case a multithreaded
-implementation of MPI is used, developers should not add modules or
-common blocks to the source code but rather use regular subroutine
-arguments (which can be grouped in ``derived types'' if needed for
-clarity).
-
-The Gauss-Lobatto-Legendre subroutines in \texttt{gll\_library.f90}
-are based in part on software libraries from the Massachusetts Institute
-of Technology, Department of Mechanical Engineering (Cambridge, Massachusetts,
-USA). The non-structured global numbering software was provided by
-Paul F. Fischer (Brown University, Providence, Rhode Island, USA,
-now at Argonne National Laboratory, USA).
-
-OpenDX \urlwithparentheses{http://www.opendx.org} is open-source
-based on IBM Data Explorer, AVS \urlwithparentheses{http://www.avs.com}
-is a trademark of Advanced Visualization Systems, and ParaView \urlwithparentheses{http://www.paraview.org}
-is an open-source visualization platform.
-
-Please e-mail your feedback, questions, comments, and suggestions
-to the CIG Computational Seismology Mailing List \urlwithparentheses{cig-seismo@geodynamics.org}.
-
-\addcontentsline{toc}{chapter}{Copyright}
-
-\chapter*{Copyright}
-
-Main historical authors: Dimitri Komatitsch and Jeroen Tromp
-
-Princeton University, USA, and CNRS / University of Marseille, France
-
-$\copyright$ Princeton University and CNRS / University of Marseille, July 2012
-
-This program is free software; you can redistribute it and/or modify
-it under the terms of the GNU General Public License as published
-by the Free Software Foundation (see Appendix \ref{cha:License}).\\
-
-\textbf{\underline{Evolution of the code:}}\\
-
-MPI v. 3.0, December 2014: many developers.
-Convolutional PML, LDDRK time scheme, bulk attenuation support, simultaneous MPI runs,
-ADIOS file I/O support, coupling with external codes, new seismogram names,
-Deville routines for additional GLL degrees, tomography tools, unit/regression test framework,
-improved CUDA GPUs performance, additonal GEOCUBIT support, better make compilation,
-git versioning system. \\
-
-MPI v. 2.1, July 2012: Max Rietmann, Peter Messmer, Daniel Peter, Dimitri
-Komatitsch, Joseph Charles, Zhinan Xie: support for CUDA GPUs, better
-CFL stability for the Stacey absorbing conditions. \\
-
-MPI v. 2.0, November 2010: Dimitri Komatitsch, Nicolas Le Goff, Roland
-Martin and Pieyre Le Loher, University of Pau, France, Daniel Peter,
-Jeroen Tromp and the Princeton group of developers, Princeton University,
-USA, and Emanuele Casarotti, INGV Roma, Italy: support for CUBIT meshes
-decomposed by SCOTCH; much faster solver using Michel Deville's inlined
-matrix products.\\
-
-MPI v. 1.4 Dimitri Komatitsch, University of Pau, Qinya Liu and others,
-Caltech, September 2006: better adjoint and kernel calculations, faster
-and better I/Os on very large systems, many small improvements and
-bug fixes.\\
-
-MPI v. 1.3 Dimitri Komatitsch, University of Pau, and Qinya Liu, Caltech,
-July 2005: serial version, regular mesh, adjoint and kernel calculations,
-ParaView support.\\
-
-MPI v. 1.2 Min Chen and Dimitri Komatitsch, Caltech, July 2004: full
-anisotropy, volume movie.\\
-
-MPI v. 1.1 Dimitri Komatitsch, Caltech, October 2002: Zhu's Moho map,
-scaling of $V_{s}$ with depth, Hauksson's regional model, attenuation,
-oceans, movies.\\
-
-MPI v. 1.0 Dimitri Komatitsch, Caltech, USA, May 2002: first MPI version
-based on global code.\\
-
-Dimitri Komatitsch, IPG Paris, France, December 1996: first 3-D solver
-for the CM-5 Connection Machine, parallelized on 128 processors using
-Connection Machine Fortran.\\
-
-\addcontentsline{toc}{chapter}{References} \bibliography{bibliography}
-
-\appendix
-%dummy comment inserted by tex2lyx to ensure that this paragraph is not empty
-
-\chapter{\label{cha:Coordinates}Reference Frame Convention}
-
-The code uses the following convention for the Cartesian reference
-frame:
-\begin{itemize}
-\item the $x$ axis points East
-\item the $y$ axis points North
-\item the $z$ axis points up
-\end{itemize}
-Note that this convention is different from both the \citet{AkRi80}
-convention and the Harvard Centroid-Moment Tensor (CMT) convention.
-The Aki \& Richards convention is
-\begin{itemize}
-\item the $x$ axis points North
-\item the $y$ axis points East
-\item the $z$ axis points down
-\end{itemize}
-and the Harvard CMT convention is
-\begin{itemize}
-\item the $x$ axis points South
-\item the $y$ axis points East
-\item the $z$ axis points up
-\end{itemize}
-
-\subsection*{Source and receiver locations}
-
-The SPECFEM3D Cartesian software code internally uses Cartesian coordinates.
-The given locations for sources and receiver locations thus may get
-converted. Sources and receiver locations are read in from the \texttt{CMTSOLUTION}
-(or \texttt{FORCESOLUTION}) and \texttt{STATIONS} files. Note that
-e.g. the \texttt{CMTSOLUTION} and \texttt{FORCESOLUTION} files denotes
-the location by \textquotedbl{}longitude/latitude/depth\textquotedbl{}.
-We read in longitude as $x$ coordinate, latitude as $y$ coordinate.
-
-In case the flag \texttt{SUPPRESS\_UTM\_PROJECTION} is set to \texttt{.false.}
-in the main parameter file (see Chapter~\ref{cha:Creating-Distributed-Databases}),
-the $x$/$y$ coordinates have to be given in degrees and are converted
-to Cartesian coordinates using a UTM conversion with the specified
-UTM zone.
-
-The value for depth (given in $km$ in \texttt{CMTSOLUTION} and \texttt{FORCESOLUTION})
-or burial depth (given in $m$ in \texttt{STATIONS}) is evaluated
-with respect to the surface of the mesh at the specified $x$/$y$
-location to find the corresponding $z$ coordinate. It is possible
-to use this depth value directly as $z$ coordinate by changing the
-flag \texttt{USE\_SOURCES\_RECVS\_Z} to \texttt{.true.} in the file
-\texttt{constants.h} located in the \texttt{src/shared/} subdirectory.
-
-\subsection*{Seismogram outputs}
-
-The seismogram output directions are given in Cartesian $x$/$y$/$z$
-directions and not rotated any further. Changing flags in \texttt{constants.h}
-in the \texttt{src/shared/} subdirectory only rotates the seismogram
-outputs if receivers are forced to be located at the surface (\texttt{RECVS\_CAN\_BE\_BURIED\_EXT\_MESH}
-set to \texttt{.false.}) and the normal to the surface at the receiver
-location should be used (\texttt{EXT\_MESH\_RECV\_NORMAL} set to \texttt{.true.})
-as vertical. In this case, the outputs are rotated to have the vertical
-component normal to the surface of the mesh, $x$ and $y$ directions
-are somewhat arbitrary orthogonal directions along the surface.
-
-For the labeling of the seismogram channels, see Appendix~\ref{cha:channel-codes}.
-Additionally, we add labels to the synthetics using the following
-convention: For a receiver station located in an
-\begin{description}
-\item [{elastic domain:}] ~\\
-
-\begin{itemize}
-\item \texttt{semd} for the displacement vector
-\item \texttt{semv} for the velocity vector
-\item \texttt{sema} for the acceleration vector
-\end{itemize}
-\item [{acoustic domain:}] ~\\
- (please note that receiver stations in acoustic domains must be buried.
-This is due to the free surface condition which enforces a zero pressure
-at the surface.)
-
-\begin{itemize}
-\item \texttt{semd} for the displacement vector
-\item \texttt{semv} for the velocity vector
-\item \texttt{sema} for pressure which will be stored in the vertical component
-\texttt{Z}. Note that pressure in the acoustic media is isotropic
-and thus the pressure record would be the same in the other two component
-directions. We therefore use the other two seismogram components to
-store the acoustic potential in component \texttt{X} (or \texttt{N})
-and the first time derivative of the acoustic potential in component
-\texttt{Y} (or \texttt{E}).
-\end{itemize}
-\end{description}
-The seismograms are by default written out in ASCII-format to the
-\texttt{OUTPUT\_FILES/} subdirectory by each parallel process. You
-can change this behavior by changing the following flags in the \texttt{constants.h}
-file located in the \texttt{src/shared/} subdirectory:
-\begin{description}
-\item [{\texttt{SEISMOGRAMS\_BINARY}}] set to \texttt{.true.} to have seismograms
-written out in binary format.
-\item [{\texttt{WRITE\_SEISMOGRAMS\_BY\_MASTER}}] Set to \texttt{.true.}
-to have only the master process writing out seismograms. This can
-be useful on a cluster, where only the master process node has access
-to the output directory.
-\item [{\texttt{USE\_OUTPUT\_FILES\_PATH}}] Set to \texttt{.false.} to
-have the seismograms output to \texttt{LOCAL\_PATH} directory specified
-in the main parameter file \texttt{DATA/Par\_file}.
-\end{description}
-
-\chapter{\label{cha:channel-codes}Channel Codes of Seismograms}
-
-Seismic networks, such as the Global Seismographic Network (GSN),
-generally involve various types of instruments with different bandwidths,
-sampling properties and component configurations. There are standards
-to name channel codes depending on instrument properties. IRIS \texttt{(www.iris.edu)}
-uses SEED format for channel codes, which are represented by three
-letters, such as \texttt{LHN}, \texttt{BHZ}, etc. In older versions
-of the SPECFEM3D Cartesian package, a common format was used for the
-channel codes of all seismograms, which was \texttt{BHE/BHN/BHZ} for
-three components. To avoid confusion when comparison are made to observed
-data, we are now using the FDSN convention (\url{http://www.fdsn.org/})
-for SEM seismograms. In the following, we give a brief explanation
-of the FDSN convention used by IRIS, and how it is adopted in SEM
-seismograms. Please visit \texttt{www.iris.edu/manuals/SEED\_appA.htm}
-for further information.\\
-
-\noindent \texttt{\textbf{Band code:}} The first letter in the channel
-code denotes the band code of seismograms, which depends on the response
-band and the sampling rate of instruments. The list of band codes
-used by IRIS is shown in Figure \ref{fig:IRIS_band_codes}. The sampling
-rate of SEM synthetics is controlled by the resolution of simulations
-rather than instrument properties. However, for consistency, we follow
-the FDSN convention for SEM seismograms governed by their sampling
-rate. For SEM synthetics, we consider band codes for which $dt\leq1$
-s. IRIS also considers the response band of instruments. For instance,
-short-period and broad-band seismograms with the same sampling rate
-correspond to different band codes, such as S and B, respectively.
-In such cases, we consider SEM seismograms as broad band, ignoring
-the corner period ($\geq10$ s) of the response band of instruments
-(note that at these resolutions, the minimum period in the SEM synthetics
-will be less than $10$ s). Accordingly, when you run a simulation
-the band code will be chosen depending on the resolution of the synthetics,
-and channel codes of SEM seismograms will start with either \texttt{L},
-\texttt{M}, \texttt{B}, \texttt{H}, \texttt{C} or \texttt{F}, shown
-by red color in the figure.\\
-
-\begin{figure}[ht]
-\noindent \begin{centering}
-\includegraphics[scale=0.6]{figures/IRIS_band_codes}
-\par\end{centering}
-
-\caption{The band code convention is based on the sampling rate and the response
-band of instruments. Please visit \texttt{www.iris.edu/manuals/SEED\_appA.htm}
-for further information. Grey rows show the relative band-code range
-in SPECFEM3D Cartesian, and the band codes used to name SEM seismograms
-are denoted in red.}
-
-\label{fig:IRIS_band_codes}
-\end{figure}
-
-\noindent \texttt{\textbf{Instrument code:}} The second letter in
-the channel code corresponds to instrument codes, which specify the
-family to which the sensor belongs. For instance, \texttt{H} and \texttt{L}
-are used for high-gain and low-gain seismometers, respectively. The
-instrument code of SEM seismograms will always be \texttt{X}, as assigned
-by FDSN for synthetic seismograms. \\
-
-\noindent \texttt{\textbf{Orientation code:}} The third letter in
-channel codes is an orientation code, which generally describes the
-physical configuration of the components of instrument packages. SPECFEM3D
-Cartesian uses the traditional orientation code \texttt{E/N/Z} (East-West,
-North-South, Vertical) for three components when a UTM projection
-is used. If the UTM conversion is suppressed, i.e. the flag \texttt{SUPPRESS\_UTM\_PROJECTION}
-is set to \texttt{.true.}, then the three components are labelled
-\texttt{X/Y/Z} according to the Cartesian reference frame. \\
-
-\noindent \texttt{\textbf{EXAMPLE:}} The sampling rate is given by
-\texttt{DT} in the main parameter file \texttt{Par\_file} located
-in the \texttt{DATA/} subdirectory. Depending on the resolution of
-your simulations, if you choose a sampling rate greater than $0.01$
-s and less than $1$ s, a seismogram recording displacements on the
-vertical component of a station \texttt{ASBS} (network \texttt{AZ})
-will be named \texttt{AZ.ASBS.MXZ.semd.sac}, whereas it will be \texttt{AZ.ASBS.BXZ.semd.sac},
-if the sampling rate is greater than 0.0125 and less equal to 0.1
-s.
-
-\chapter{\label{cha:Troubleshooting}Troubleshooting}
-
-\section*{FAQ}
-\begin{description}
-\item [{configuration fails:}]~
-
-Examine the log file 'config.log'. It contains detailed informations.
-In many cases, the paths to these specific compiler commands F90,
-CC and MPIF90 will not be correct if `./configure` fails.
-
-Please make sure that you have a working installation of a Fortran
-compiler, a C compiler and an MPI implementation. You should be able
-to compile this little program code:
-\begin{lyxcode}
-program~main~~~\\
-~include~'mpif.h'~~~\\
-~integer,~parameter~::~CUSTOM\_MPI\_TYPE~=~MPI\_REAL~~~\\
-~integer~ier~~~\\
-~call~MPI\_INIT(ier)~~~\\
-~call~MPI\_BARRIER(MPI\_COMM\_WORLD,ier)~~~\\
-~call~MPI\_FINALIZE(ier)~~~\\
-~end
-\end{lyxcode}
-\item [{compilation fails stating:}] ~\\
-
-\begin{lyxcode}
-...~~~\\
-~obj/program\_generate\_databases.o:~In~function~`MAIN\_\_':~~~\\
-~program\_generate\_databases.f90:(.text+0x14):~undefined~reference~to~`\_gfortran\_set\_std'~~~\\
-~...~~~\\
-
-\end{lyxcode}
-
-Make sure you're pointing to the right 'mpif90' wrapper command.
-
-Normally, this message will appear when you are mixing two different
-Fortran compilers. That is, using e.g. gfortran to compile non-MPI
-files and mpif90, wrapper provided for e.g. ifort, to compile MPI-files.
-
-fix: e.g. specify > ./configure FC=gfortran MPIF90=/usr/local/openmpi-gfortran/bin/mpif90
-
-\item [{after executing \texttt{xmeshfem3D} I've got elements with skewness of 81\% percent, what does this mean:}] Look
-at the skewness table printed in the \texttt{output\_mesher.txt} file
-after executing \texttt{xmeshfem3D} for the example given in \texttt{EXAMPLES/meshfem3D\_examples/simple\_model/}:
-
-\begin{lyxcode}
-...~~\\
-~histogram~of~skewness~(0.~good~-~1.~bad):~~~\\
-~0.0000000E+00~-~5.0000001E-02~27648~81.81818~\%~~~\\
-~5.0000001E-02~-~0.1000000~0~0.0000000E+00~\%~~~\\
-~...~~~\\
-
-\end{lyxcode}
-
-The first line means that you have 27,648 elements with a skewness
-value between 0 and 0.05 (which means the element is basically not
-skewed, just plain regular hexahedral element). The total number of
-elements you have in this mesh is (see in the \texttt{output\_mesher.txt}
-file a bit further down):
-\begin{lyxcode}
-...~~\\
-~total~number~of~elements~in~entire~mesh:~33792~~~\\
-~...~~~\\
-
-\end{lyxcode}
-
-which gives you that: 27,648 / 33,792 $\sim$ 81.8 \% of all elements
-are not skewed, i.e. regular elements. a fantastic value :)
-
-The histogram lists for this mesh also some stronger skewed elements,
-for example the worst ones belong to:
-\begin{lyxcode}
-...~~\\
-~0.6000000~-~0.6500000~2048~6.060606~\%~~~\\
-~...~~~\\
-
-\end{lyxcode}
-
-about 6 \% of all elements have distortions with a skewness value
-between 0.6 and 0.65. The skewness values give you a hint of how good
-your mesh is. In an ideal world, you would want to have no distortions,
-just like the 81\% from above. Those elements give you the best approximate
-values by the GLL quadrature used in the spectral-element method.
-However, having weakly distorted elements is still fine and the solutions
-are still accurate enough. So empirically, values up to around 0.7
-are tolerable, above that you should consider remeshing...
-
-To give you an idea why some of the elements are distorted, see the
-following figure \ref{fig:mesh.vp} of the mesh you obtain in the
-example \texttt{EXAMPLES/meshfem3D\_examples/simple\_model/}.
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=0.9\textwidth]{figures/mesh_vp}
-\par\end{centering}
-
-\caption{Paraview visualization using the mesh vtk-files for the example given
-in \texttt{EXAMPLES/meshfem3D\_examples/simple\_model/}.}
-
-\label{fig:mesh.vp}
-\end{figure}
-
-You will see that the mesh contains a doubling layer, where we stitch
-elements together such that the size of two elements will transition
-to the size of one element (very useful to keep the ratio of wavespeed
-/ element\_size about constant). Those elements in this doubling layer
-have higher skewness values and make up those 6 \% in the histogram.
-
-\item [{the code gives following error message "need at least one receiver":}] This
-means that no stations given in the input file \texttt{DATA/STATIONS}
-could be located within the dimensions of the mesh. This can happen
-for example when the mesh was created with the in-house mesher \texttt{xmeshfem3D}
-while using the Universal Transverse Mercator (UTM) projection but
-the simulation with \texttt{xspecfem3D} was suppressing this projection
-from latitude/longitude to x/y/z coordinates.
-
-In such cases, try to change your \texttt{DATA/Par\_file} and set
-e.g.:
-\begin{lyxcode}
-SUPPRESS\_UTM\_PROJECTION~=~.false.~~~\\
-
-\end{lyxcode}
-
-to be the same in \texttt{Mesh\_Par\_file} and \texttt{Par\_file}.
-This flag should be identical when using the in-house mesher \texttt{xmeshfem3D},
-\texttt{xgenerate\_databases} and \texttt{xspecfem3D} together to
-run simulations.
-
-The flag determines if the coordinates you specify for your source
-and station locations are given as lat/lon degrees and must be converted
-to UTM coordinates. As an example, if you use \texttt{.false.} within
-\texttt{Mesh\_Par\_file} then you create a mesh with \texttt{xmeshfem3D}
-using the UTM projection from lat/lon as input format to UTM projected
-coordinates to store the mesh point positions, which is fine. The
-error then may occur if in the \texttt{Par\_file} you have this set
-to \texttt{.true.} so that the \texttt{xgenerate\_databases} and \texttt{xspecfem3D}
-suppress the UTM projection and assume that all coordinates you use
-now for source and receiver locations are given in meters (that is,
-converted) already. So it won't find the specified locations in the
-used mesh. As a solutions, just change the flag in \texttt{Par\_file}
-to be the same as in \texttt{Mesh\_Par\_file} and rerun \texttt{xgenerate\_databases}
-and \texttt{xspecfem3D} to make sure that your simulation works fine.
-
-\item [{I get the following error message "forward simulation became unstable and blew up":}] In
-most cases this means that your time step size \texttt{DT} is chosen
-too big. Look at your files \texttt{output\_mesher.txt} or \texttt{output\_solver.txt}
-created in the folder \texttt{OUTPUT\_FILES}. In these output files,
-find the section:
-
-\begin{lyxcode}
-...~~~\\
-~{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}~~~\\
-~{*}{*}{*}~Verification~of~simulation~parameters~{*}{*}{*}~~~\\
-~{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}~~~\\
-~...~~~\\
-~{*}{*}{*}~Minimum~period~resolved~=~4.308774~~~\\
-~{*}{*}{*}~Maximum~suggested~time~step~=~6.8863556E-02~~~\\
-~...~~~\\
-~...~~~\\
-
-\end{lyxcode}
-
-then change \texttt{DT} in the \texttt{DATA/Par\_file} to be somewhere
-close to the maximum suggested time step. In the example above: \\
- \texttt{DT = 0.05d0} \\
- would (most probably) work fine. It could be also bigger than the
-0.068 s suggested. This depends a bit on the distortions of your mesh
-elements. The more regular they are, the bigger you can choose \texttt{DT}.
-Just play with this value a bit and see when the simulation becomes
-stable ...
-
-\end{description}
-
-\chapter{\label{cha:License}License }
-
-\textbf{GNU GENERAL PUBLIC LICENSE Version 2, June 1991. Copyright
-(C) 1989, 1991 Free Software Foundation, Inc. 59 Temple Place, Suite
-330, Boston, MA 02111-1307 USA} \\
- Everyone is permitted to copy and distribute verbatim copies of this
-license document, but changing it is not allowed.
-
-\section*{Preamble}
-
-The licenses for most software are designed to take away your freedom
-to share and change it. By contrast, the GNU General Public License
-is intended to guarantee your freedom to share and change free software
--- to make sure the software is free for all its users. This General
-Public License applies to most of the Free Software Foundation's software
-and to any other program whose authors commit to using it. (Some other
-Free Software Foundation software is covered by the GNU Library General
-Public License instead.) You can apply it to your programs, too.
-
-When we speak of free software, we are referring to freedom, not price.
-Our General Public Licenses are designed to make sure that you have
-the freedom to distribute copies of free software (and charge for
-this service if you wish), that you receive source code or can get
-it if you want it, that you can change the software or use pieces
-of it in new free programs; and that you know you can do these things.
-
-To protect your rights, we need to make restrictions that forbid anyone
-to deny you these rights or to ask you to surrender the rights. These
-restrictions translate to certain responsibilities for you if you
-distribute copies of the software, or if you modify it.
-
-For example, if you distribute copies of such a program, whether gratis
-or for a fee, you must give the recipients all the rights that you
-have. You must make sure that they, too, receive or can get the source
-code. And you must show them these terms so they know their rights.
-
-We protect your rights with two steps:
-\begin{enumerate}
-\item Copyright the software, and
-\item Offer you this license which gives you legal permission to copy, distribute
-and/or modify the software.
-\end{enumerate}
-Also, for each author's protection and ours, we want to make certain
-that everyone understands that there is no warranty for this free
-software. If the software is modified by someone else and passed on,
-we want its recipients to know that what they have is not the original,
-so that any problems introduced by others will not reflect on the
-original authors' reputations.
-
-Finally, any free program is threatened constantly by software patents.
-We wish to avoid the danger that redistributors of a free program
-will individually obtain patent licenses, in effect making the program
-proprietary. To prevent this, we have made it clear that any patent
-must be licensed for everyone's free use or not licensed at all.
-
-The precise terms and conditions for copying, distribution and modification
-follow.
-
-\section*{GNU GENERAL PUBLIC LICENSE TERMS AND CONDITIONS FOR COPYING, DISTRIBUTION
-AND MODIFICATION }
-\begin{itemize}
-\item 0.This License applies to any program or other work which contains
-a notice placed by the copyright holder saying it may be distributed
-under the terms of this General Public License. The ``Program''
-below refers to any such program or work, and a ``work based on the
-Program'' means either the Program or any derivative work under copyright
-law: that is to say, a work containing the Program or a portion of
-it, either verbatim or with modifications and/or translated into another
-language. (Hereinafter, translation is included without limitation
-in the term ``modification.'') Each licensee is addressed as ``you.''\\
- \\
- Activities other than copying, distribution and modification are
-not covered by this License; they are outside its scope. The act of
-running the Program is not restricted, and the output from the Program
-is covered only if its contents constitute a work based on the Program
-(independent of having been made by running the Program). Whether
-that is true depends on what the Program does.\end{itemize}
-\begin{enumerate}
-\item You may copy and distribute verbatim copies of the Program's source
-code as you receive it, in any medium, provided that you conspicuously
-and appropriately publish on each copy an appropriate copyright notice
-and disclaimer of warranty; keep intact all the notices that refer
-to this License and to the absence of any warranty; and give any other
-recipients of the Program a copy of this License along with the Program.
-
-You may charge a fee for the physical act of transferring a copy,
-and you may at your option offer warranty protection in exchange for
-a fee.
-
-\item You may modify your copy or copies of the Program or any portion of
-it, thus forming a work based on the Program, and copy and distribute
-such modifications or work under the terms of Section 1 above, provided
-that you also meet all of these conditions:
-
-\begin{enumerate}
-\item You must cause the modified files to carry prominent notices stating
-that you changed the files and the date of any change.
-\item You must cause any work that you distribute or publish, that in whole
-or in part contains or is derived from the Program or any part thereof,
-to be licensed as a whole at no charge to all third parties under
-the terms of this License.
-\item If the modified program normally reads commands interactively when
-run, you must cause it, when started running for such interactive
-use in the most ordinary way, to print or display an announcement
-including an appropriate copyright notice and a notice that there
-is no warranty (or else, saying that you provide a warranty) and that
-users may redistribute the program under these conditions, and telling
-the user how to view a copy of this License. (Exception: if the Program
-itself is interactive but does not normally print such an announcement,
-your work based on the Program is not required to print an announcement.)
-\end{enumerate}
-
-These requirements apply to the modified work as a whole. If identifiable
-sections of that work are not derived from the Program, and can be
-reasonably considered independent and separate works in themselves,
-then this License, and its terms, do not apply to those sections when
-you distribute them as separate works. But when you distribute the
-same sections as part of a whole which is a work based on the Program,
-the distribution of the whole must be on the terms of this License,
-whose permissions for other licensees extend to the entire whole,
-and thus to each and every part regardless of who wrote it.
-
-Thus, it is not the intent of this section to claim rights or contest
-your rights to work written entirely by you; rather, the intent is
-to exercise the right to control the distribution of derivative or
-collective works based on the Program.
-
-In addition, mere aggregation of another work not based on the Program
-with the Program (or with a work based on the Program) on a volume
-of a storage or distribution medium does not bring the other work
-under the scope of this License.
-
-\item You may copy and distribute the Program (or a work based on it, under
-Section 2) in object code or executable form under the terms of Sections
-1 and 2 above provided that you also do one of the following:
-
-\begin{enumerate}
-\item Accompany it with the complete corresponding machine-readable source
-code, which must be distributed under the terms of Sections 1 and
-2 above on a medium customarily used for software interchange; or,
-\item Accompany it with a written offer, valid for at least three years,
-to give any third party, for a charge no more than your cost of physically
-performing source distribution, a complete machine-readable copy of
-the corresponding source code, to be distributed under the terms of
-Sections 1 and 2 above on a medium customarily used for software interchange;
-or,
-\item Accompany it with the information you received as to the offer to
-distribute corresponding source code. (This alternative is allowed
-only for noncommercial distribution and only if you received the program
-in object code or executable form with such an offer, in accord with
-Subsection b above.)
-\end{enumerate}
-
-The source code for a work means the preferred form of the work for
-making modifications to it. For an executable work, complete source
-code means all the source code for all modules it contains, plus any
-associated interface definition files, plus the scripts used to control
-compilation and installation of the executable. However, as a special
-exception, the source code distributed need not include anything that
-is normally distributed (in either source or binary form) with the
-major components (compiler, kernel, and so on) of the operating system
-on which the executable runs, unless that component itself accompanies
-the executable.
-
-If distribution of executable or object code is made by offering access
-to copy from a designated place, then offering equivalent access to
-copy the source code from the same place counts as distribution of
-the source code, even though third parties are not compelled to copy
-the source along with the object code.
-
-\item You may not copy, modify, sublicense, or distribute the Program except
-as expressly provided under this License. Any attempt otherwise to
-copy, modify, sublicense or distribute the Program is void, and will
-automatically terminate your rights under this License. However, parties
-who have received copies, or rights, from you under this License will
-not have their licenses terminated so long as such parties remain
-in full compliance.
-\item You are not required to accept this License, since you have not signed
-it. However, nothing else grants you permission to modify or distribute
-the Program or its derivative works. These actions are prohibited
-by law if you do not accept this License. Therefore, by modifying
-or distributing the Program (or any work based on the Program), you
-indicate your acceptance of this License to do so, and all its terms
-and conditions for copying, distributing or modifying the Program
-or works based on it.
-\item Each time you redistribute the Program (or any work based on the Program),
-the recipient automatically receives a license from the original licensor
-to copy, distribute or modify the Program subject to these terms and
-conditions. You may not impose any further restrictions on the recipients'
-exercise of the rights granted herein. You are not responsible for
-enforcing compliance by third parties to this License.
-\item If, as a consequence of a court judgment or allegation of patent infringement
-or for any other reason (not limited to patent issues), conditions
-are imposed on you (whether by court order, agreement or otherwise)
-that contradict the conditions of this License, they do not excuse
-you from the conditions of this License. If you cannot distribute
-so as to satisfy simultaneously your obligations under this License
-and any other pertinent obligations, then as a consequence you may
-not distribute the Program at all. For example, if a patent license
-would not permit royalty-free redistribution of the Program by all
-those who receive copies directly or indirectly through you, then
-the only way you could satisfy both it and this License would be to
-refrain entirely from distribution of the Program.
-
-If any portion of this section is held invalid or unenforceable under
-any particular circumstance, the balance of the section is intended
-to apply and the section as a whole is intended to apply in other
-circumstances.
-
-It is not the purpose of this section to induce you to infringe any
-patents or other property right claims or to contest validity of any
-such claims; this section has the sole purpose of protecting the integrity
-of the free software distribution system, which is implemented by
-public license practices. Many people have made generous contributions
-to the wide range of software distributed through that system in reliance
-on consistent application of that system; it is up to the author/donor
-to decide if he or she is willing to distribute software through any
-other system and a licensee cannot impose that choice.
-
-This section is intended to make thoroughly clear what is believed
-to be a consequence of the rest of this License.
-
-\item If the distribution and/or use of the Program is restricted in certain
-countries either by patents or by copyrighted interfaces, the original
-copyright holder who places the Program under this License may add
-an explicit geographical distribution limitation excluding those countries,
-so that distribution is permitted only in or among countries not thus
-excluded. In such case, this License incorporates the limitation as
-if written in the body of this License.
-\item The Free Software Foundation may publish revised and/or new versions
-of the General Public License from time to time. Such new versions
-will be similar in spirit to the present version, but may differ in
-detail to address new problems or concerns.
-
-Each version is given a distinguishing version number. If the Program
-specifies a version number of this License which applies to it and
-``any later version,'' you have the option of following the terms
-and conditions either of that version or of any later version published
-by the Free Software Foundation. If the Program does not specify a
-version number of this License, you may choose any version ever published
-by the Free Software Foundation.
-
-\item If you wish to incorporate parts of the Program into other free programs
-whose distribution conditions are different, write to the author to
-ask for permission. For software which is copyrighted by the Free
-Software Foundation, write to the Free Software Foundation; we sometimes
-make exceptions for this. Our decision will be guided by the two goals
-of preserving the free status of all derivatives of our free software
-and of promoting the sharing and reuse of software generally.
-\end{enumerate}
-
-\subsection*{NO WARRANTY }
-\begin{itemize}
-\item 11.BECAUSE THE PROGRAM IS LICENSED FREE OF CHARGE, THERE IS NO WARRANTY
-FOR THE PROGRAM, TO THE EXTENT PERMITTED BY APPLICABLE LAW. EXCEPT
-WHEN OTHERWISE STATED IN WRITING THE COPYRIGHT HOLDERS AND/OR OTHER
-PARTIES PROVIDE THE PROGRAM ``AS IS'' WITHOUT WARRANTY OF ANY KIND,
-EITHER EXPRESSED OR IMPLIED, INCLUDING, BUT NOT LIMITED TO, THE IMPLIED
-WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE.
-THE ENTIRE RISK AS TO THE QUALITY AND PERFORMANCE OF THE PROGRAM IS
-WITH YOU. SHOULD THE PROGRAM PROVE DEFECTIVE, YOU ASSUME THE COST
-OF ALL NECESSARY SERVICING, REPAIR OR CORRECTION.
-\item 12.IN NO EVENT UNLESS REQUIRED BY APPLICABLE LAW OR AGREED TO IN WRITING
-WILL ANY COPYRIGHT HOLDER, OR ANY OTHER PARTY WHO MAY MODIFY AND/OR
-REDISTRIBUTE THE PROGRAM AS PERMITTED ABOVE, BE LIABLE TO YOU FOR
-DAMAGES, INCLUDING ANY GENERAL, SPECIAL, INCIDENTAL OR CONSEQUENTIAL
-DAMAGES ARISING OUT OF THE USE OR INABILITY TO USE THE PROGRAM (INCLUDING
-BUT NOT LIMITED TO LOSS OF DATA OR DATA BEING RENDERED INACCURATE
-OR LOSSES SUSTAINED BY YOU OR THIRD PARTIES OR A FAILURE OF THE PROGRAM
-TO OPERATE WITH ANY OTHER PROGRAMS), EVEN IF SUCH HOLDER OR OTHER
-PARTY HAS BEEN ADVISED OF THE POSSIBILITY OF SUCH DAMAGES.
-\end{itemize}
-
-\section*{END OF TERMS AND CONDITIONS }
-
-\subsection*{How to Apply These Terms to Your New Programs}
-
-If you develop a new program, and you want it to be of the greatest
-possible use to the public, the best way to achieve this is to make
-it free software which everyone can redistribute and change under
-these terms.
-
-To do so, attach the following notices to the program. It is safest
-to attach them to the start of each source file to most effectively
-convey the exclusion of warranty; and each file should have at least
-the ``copyright'' line and a pointer to where the full notice is
-found. For example:
-\begin{quote}
-One line to give the program's name and a brief idea of what it does.
-Copyright {\footnotesize $\copyright$ (}year) (name of author)
-
-This program is free software; you can redistribute it and/or modify
-it under the terms of the GNU General Public License as published
-by the Free Software Foundation; either version 2 of the License,
-or (at your option) any later version.
-
-This program is distributed in the hope that it will be useful, but
-WITHOUT ANY WARRANTY; without even the implied warranty of MERCHANTABILITY
-or FITNESS FOR A PARTICULAR PURPOSE. See the GNU General Public License
-for more details.
-
-You should have received a copy of the GNU General Public License
-along with this program; if not, write to the Free Software Foundation,
-Inc., 59 Temple Place, Suite 330, Boston, MA 02111-1307 USA
-\end{quote}
-Also add information on how to contact you by electronic and paper
-mail.
-
-If the program is interactive, make it output a short notice like
-this when it starts in an interactive mode:
-\begin{quote}
-Gnomovision version 69, Copyright $\copyright$ year name of author
-Gnomovision comes with ABSOLUTELY NO WARRANTY; for details type `show
-w'. This is free software, and you are welcome to redistribute it
-under certain conditions; type `show c' for details.
-\end{quote}
-The hypothetical commands `show w' and `show c' should show the appropriate
-parts of the General Public License. Of course, the commands you use
-may be called something other than `show w' and `show c'; they could
-even be mouse-clicks or menu items -- whatever suits your program.
-
-You should also get your employer (if you work as a programmer) or
-your school, if any, to sign a ``copyright disclaimer'' for the
-program, if necessary. Here is a sample; alter the names:
-\begin{quote}
-Yoyodyne, Inc., hereby disclaims all copyright interest in the program
-`Gnomovision' (which makes passes at compilers) written by James Hacker.
-
-(signature of Ty Coon)\\
- 1 April 1989 \\
- Ty Coon, President of Vice
-\end{quote}
-This General Public License does not permit incorporating your program
-into proprietary programs. If your program is a subroutine library,
-you may consider it more useful to permit linking proprietary applications
-with the library. If this is what you want to do, use the GNU Library
-General Public License instead of this License.
-\end{document}

### 2015-04-24 doc/USER_MANUAL/A_reference_frame.tex
updates user manual to use facilitate the generation of a markdown version to the github wiki

--- 2015-04-18 doc/USER_MANUAL/manual_SPECFEM3D_Cartesian.tex
@@
-% !TEX encoding = IsoLatin
-%% LyX 2.0.2 created this file.  For more info, see http://www.lyx.org/.
-%% Do not edit unless you really know what you are doing.
-\documentclass[oneside,english]{book}
-\usepackage[T1]{fontenc}
-%% DK DK we do not accept accented letters, use the old LaTeX syntax
-%% DK DK for them instead: \'e \`e  \"e \^e and so on.
-%% DK DK \usepackage[latin1]{inputenc}
-\usepackage[letterpaper]{geometry}
-\geometry{verbose,tmargin=1in,bmargin=1in,lmargin=1in,rmargin=1in}
-\setcounter{secnumdepth}{3}
-\setcounter{tocdepth}{3}
-\usepackage{babel}
-\usepackage{float}
-\usepackage{textcomp}
-\usepackage{url}
-\usepackage{amsmath}
-\usepackage{amssymb}
-\usepackage{graphicx}
-\usepackage[authoryear]{natbib}
-\usepackage[unicode=true,pdfusetitle,
- bookmarks=true,bookmarksnumbered=true,bookmarksopen=false,
- breaklinks=false,pdfborder={0 0 0},backref=section,colorlinks=false,pdfpagemode=None]
- {hyperref}
-\hypersetup{
- pdftex,pdfstartview=FitH,pdfpagelayout=SinglePage}
-\usepackage{breakurl}
-
-\makeatletter
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% LyX specific LaTeX commands.
-\newcommand{\noun}[1]{\textsc{#1}}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Textclass specific LaTeX commands.
-\newenvironment{lyxcode}
-{\par\begin{list}{}{
-\setlength{\rightmargin}{\leftmargin}
-\setlength{\listparindent}{0pt}% needed for AMS classes
-\raggedright
-\setlength{\itemsep}{0pt}
-\setlength{\parsep}{0pt}
-\normalfont\ttfamily}%
- \item[]}
-{\end{list}}
-\newenvironment{lyxlist}[1]
-{\begin{list}{}
-{\settowidth{\labelwidth}{#1}
- \setlength{\leftmargin}{\labelwidth}
- \addtolength{\leftmargin}{\labelsep}
- \renewcommand{\makelabel}[1]{##1\hfil}}}
-{\end{list}}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% User specified LaTeX commands.
-%\renewcommand{\baselinestretch}{1.5}
-
-% figures
-\usepackage[dvips]{epsfig}
-
-% fonts
-\usepackage{times}
-
-% hyperlinks to sections and references
-
-\newcommand{\urlwithparentheses}[1]{(\url{#1})}
-
-% biblio GJI
-\bibliographystyle{abbrvnat}
-
-\newcommand{\nexxi}{\mbox{\texttt{NEX\_XI\/}}}
-\newcommand{\nexeta}{\mbox{\texttt{NEX\_ETA\/}}}
-\newcommand{\nprocxi}{\mbox{\texttt{NPROC\_XI\/}}}
-\newcommand{\nproceta}{\mbox{\texttt{NPROC\_ETA\/}}}
-\newcommand{\nchunks}{\mbox{\texttt{NCHUNKS\/}}}
-\newcommand{\eg}{e.g.,}
-
-% colors for the text
-\@ifundefined{definecolor}
- {\usepackage{color}}{}
-\newcommand{\red}[1]{\textcolor{Red}{#1}}
-
-\usepackage{babel}
-
-\makeatother
-
-\begin{document}
-\begin{center}
-\thispagestyle{empty}\vspace*{-1.8truecm} %
-\makebox[1\textwidth]{%
-\includegraphics[width=0.83\paperwidth]{figures/specfem_3d_Cartesian-cover} %
-}
-\par\end{center}
-
-\title{\textbf{SPECFEM3D Cartesian}\\
- \textbf{User Manual}}
-
-\author{$\copyright$ Princeton University (USA), CNRS and University of Marseille (France),\\
-and ETH Z\"urich (Switzerland)\\
-Version 3.0}
-
-\date{\today}
-
-\maketitle
-
-\section*{Authors}
-
-\noindent The SPECFEM3D Cartesian package was first developed by Dimitri
-Komatitsch and Jean-Pierre Vilotte at Institut de Physique du Globe
-(IPGP) in Paris, France from 1995 to 1997 and then by Dimitri Komatitsch
-and Jeroen Tromp at Harvard University and Caltech, USA, starting in 1998.
-The story started on April 4, 1995, when Prof. Yvon Maday from CNRS and University of Paris, France, gave a lecture to
-Dimitri Komatitsch and Jean-Pierre Vilotte at IPG about the nice properties of the Legendre spectral-element method with diagonal mass matrix that he had used for
-other equations. We are deeply indebted and thankful to him for that.
-That followed a visit by Dimitri Komatitsch to OGS (Istituto Nazionale di Oceanografia e di Geofisica Sperimentale) in Trieste, Italy, in February 1995
-to meet with G\'eza Seriani and Enrico Priolo, who introduced him to their 2D Chebyshev version of the spectral-element method with a non-diagonal mass matrix.
-We are deeply indebted and thankful to them for that.
-
-%%%%%%%% please do NOT merge the lines of the paragraph below, it makes it difficult to insert new names in alphabetical order; leave a single name per line
-%%%%%%%% please do NOT merge the lines of the paragraph below, it makes it difficult to insert new names in alphabetical order; leave a single name per line
-%%%%%%%% please do NOT merge the lines of the paragraph below, it makes it difficult to insert new names in alphabetical order; leave a single name per line
-Since then it has been developed and maintained by a development team:
-in alphabetical order,
-Jean-Paul (Pablo) Ampuero,
-Kangchen Bai,
-Piero Basini,
-C\'eline Blitz,
-Alexis Bottero,
-Ebru Bozda\u{g},
-Emanuele Casarotti,
-Joseph Charles,
-Min Chen,
-Paul Cristini,
-Percy Galvez,
-Dominik G\"oddeke,
-Vala Hj\"orleifsd\'ottir,
-Sue Kientz,
-Dimitri Komatitsch,
-Jes\'us Labarta,
-Nicolas Le Goff,
-Pieyre Le Loher,
-Matthieu Lefebvre,
-Qinya Liu,
-Yang Luo,
-Alessia Maggi,
-Federica Magnoni,
-Roland Martin,
-Ren\'e Matzen,
-Dennis McRitchie,
-Matthias Meschede,
-Peter Messmer,
-David Mich\'ea,
-Surendra Nadh Somala,
-Tarje Nissen-Meyer,
-Daniel Peter,
-Max Rietmann,
-Elliott Sales de Andrade,
-Brian Savage,
-Bernhard Schuberth,
-Anne Sieminski,
-Leif Strand,
-Carl Tape,
-Jeroen Tromp,
-Jean-Pierre Vilotte,
-Zhinan Xie,
-Chang-Hua Zhang,
-Hejun Zhu.\\
-
-The cover graphic of the manual was created by Santiago Lombeyda from
-Caltech's Center for Advanced Computing Research (CACR) \urlwithparentheses{http://www.cacr.caltech.edu}.\\
-
-The code is released open-source under the GNU version 2+ license, see the license at the end of this manual.\\
-
-\newpage{}
-\section*{Current and past main participants or main sponsors of the SPECFEM project (in no particular order)}
-%
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=0.125\textwidth]{figures/logo_cnrs}\vspace*{2truemm}
-\includegraphics[width=0.125\textwidth]{figures/logo_princeton}\vspace*{2truemm}
-\includegraphics[width=0.15\textwidth]{figures/logo_aix_marseille_universite}\vspace*{0.02truemm}
-\includegraphics[width=0.15\textwidth]{figures/logo_ETH}\vspace*{2truemm}
-\includegraphics[width=0.12\textwidth]{figures/logo_CSC_China}\vspace*{0.02truemm}
-\includegraphics[width=0.15\textwidth]{figures/logo_inria}\vspace*{2truemm}
-\includegraphics[width=0.14\textwidth]{figures/logo_UPPA}
-\par\end{centering}
-%
-\vspace*{-2truemm}
-%
-\noindent \begin{centering}
-\includegraphics[width=0.15\textwidth]{figures/logo_NVIDIA}\vspace*{2truemm}
-\includegraphics[width=0.125\textwidth]{figures/logo_IUF}\vspace*{2truemm}
-\includegraphics[width=0.135\textwidth]{figures/logo_Caltech}\vspace*{2truemm}
-\includegraphics[width=0.135\textwidth]{figures/logo_Harvard}\vspace*{2truemm}
-\includegraphics[width=0.135\textwidth]{figures/logo_IPGP}\vspace*{2truemm}
-\includegraphics[width=0.112\textwidth]{figures/logo_ANR}\vspace*{2truemm}
-\includegraphics[width=0.112\textwidth]{figures/logo_NSF}\vspace*{2truemm}
-\par\end{centering}
-%
-\vspace*{-2truemm}
-%
-\noindent \begin{centering}
-\includegraphics[width=0.190\textwidth]{figures/logo_European_Union}\vspace*{2truemm}
-\includegraphics[width=0.112\textwidth]{figures/logo_GENCI}\vspace*{2truemm}
-\includegraphics[width=0.112\textwidth]{figures/logo_PRACE}\vspace*{2truemm}
-\includegraphics[width=0.112\textwidth]{figures/logo_CINES}\vspace*{2truemm}
-\includegraphics[width=0.130\textwidth]{figures/logo_Oak_Ridge}\vspace*{2truemm}
-\hspace*{3mm}\includegraphics[width=0.130\textwidth]{figures/logo_fondation_Del_Duca}
-\includegraphics[width=0.130\textwidth]{figures/logo_CIG}\vspace*{2truemm}
-\par\end{centering}
-%
-\vspace*{-2truemm}
-%
-\noindent \begin{centering}
-\includegraphics[width=0.112\textwidth]{figures/logo_University_of_Toronto}\vspace*{2truemm}
-\includegraphics[width=0.109\textwidth]{figures/logo_INGV}\vspace*{2truemm}
-\includegraphics[width=0.102\textwidth]{figures/logo_Univ_Toulouse}\vspace*{2truemm}
-\includegraphics[width=0.112\textwidth]{figures/logo_TOTAL}\vspace*{2truemm}
-\includegraphics[width=0.130\textwidth]{figures/logo_Fairbanks}\vspace*{2truemm}
-\includegraphics[width=0.120\textwidth]{figures/logo_CINECA}\vspace*{2truemm}
-\includegraphics[width=0.140\textwidth]{figures/logo_Intel_Exascale_Labs}\vspace*{2truemm}
-\includegraphics[width=0.130\textwidth]{figures/logo_Maison_Simulation}
-\par\end{centering}
-\end{figure}
-
-\newpage{}
-
-\tableofcontents{}
-
-\chapter{Introduction}
-
-The software package SPECFEM3D Cartesian simulates seismic wave propagation
-at the local or regional scale and performs full waveform imaging (FWI) or adjoint tomography based upon the spectral-element method
-(SEM). The SEM is a continuous Galerkin technique \citep{TrKoLi08,PeKoLuMaLeCaLeMaLiBlNiBaTr11},
-which can easily be made discontinuous \citep{BeMaPa94,Ch00,KoWoHu02,ChCaVi03,LaWaBe05,Kop06,WiStBuGh10,AcKo11};
-it is then close to a particular case of the discontinuous Galerkin
-technique \citep{ReHi73,LeRa74,Arn82,JoPi86,BoMaHe91,FaRi99,HuHuRa99,CoKaSh00,GiHeWa02,RiWh03,MoRi05,GrScSc06,AiMoMu06,BeLaPi06,DuKa06,DeSeWh08,PuAmKa09,WiStBuGh10,DeSe10,EtChViGl10},
-with optimized efficiency because of its tensorized basis functions
-\citep{WiStBuGh10,AcKo11}. In particular, it can accurately handle
-very distorted mesh elements \citep{OlSe11}.\\
-
-It has very good accuracy and convergence properties \citep{MaPa89,SePr94,DeFiMu02,Coh02,DeSe07,SeOl08,AiWa09,AiWa10,MeStTh12}.
-The spectral element approach admits spectral rates of convergence
-and allows exploiting $hp$-convergence schemes. It is also very well
-suited to parallel implementation on very large supercomputers \citep{KoTsChTr03,TsKoChTr03,KoLaMi08a,CaKoLaTiMiLeSnTr08,KoViCh10}
-as well as on clusters of GPU accelerating graphics cards \citep{Kom11,MiKo10,KoMiEr09,KoErGoMi10}.
-Tensor products inside each element can be optimized to reach very
-high efficiency \citep{DeFiMu02}, and mesh point and element numbering
-can be optimized to reduce processor cache misses and improve cache
-reuse \citep{KoLaMi08a}. The SEM can also handle triangular (in 2D)
-or tetrahedral (in 3D) elements \citep{WinBoyd96,TaWi00,KoMaTrTaWi01,Coh02,MeViSa06}
-as well as mixed meshes, although with increased cost and reduced
-accuracy in these elements, as in the discontinuous Galerkin method.\\
-
-Note that in many geological models in the context of seismic wave
-propagation studies (except for instance for fault dynamic rupture
-studies, in which very high frequencies or supershear rupture need
-to be modeled near the fault, see e.g. \citet{BeGlCrViPi07,BeGlCrVi09,PuAmKa09,TaCrEtViBeSa10})
-a continuous formulation is sufficient because material property contrasts
-are not drastic and thus conforming mesh doubling bricks can efficiently
-handle mesh size variations \citep{KoTr02a,KoLiTrSuStSh04,LeChLiKoHuTr08,LeChKoHuTr09,LeKoHuTr09}.\\
-
-For a detailed introduction to the SEM as applied to regional seismic
-wave propagation, please consult \citet{PeKoLuMaLeCaLeMaLiBlNiBaTr11,TrKoLi08,KoVi98,KoTr99,ChKoViCaVaFe07}
-and in particular \citet{LeKoHuTr09,LeChKoHuTr09,LeChLiKoHuTr08,GoAmTaCaSmSaMaKo09,WiKoScTr04,KoLiTrSuStSh04}.
-A detailed theoretical analysis of the dispersion
-and stability properties of the SEM is available in \citet{Coh02}, \citet{DeSe07}, \citet{SeOl07}, \citet{SeOl08} and \citet{MeStTh12}.\\
-
-Effects due to lateral variations in compressional-wave speed, shear-wave
-speed, density, a 3D crustal model, topography and bathymetry are
-included. The package can accommodate full 21-parameter anisotropy
-(see~\citet{ChTr07}) as well as lateral variations in attenuation
-\citep{SaKoTr10}. Adjoint capabilities and finite-frequency kernel
-simulations are included \citep{TrKoLi08,PeKoLuMaLeCaLeMaLiBlNiBaTr11,LiTr06,FiIgBuKe09,ViOp09}.\\
-
-The SEM was originally developed in computational fluid dynamics \citep{Pat84,MaPa89}
-and has been successfully adapted to address problems in seismic wave
-propagation. Early seismic wave propagation applications of the SEM,
-utilizing Legendre basis functions and a perfectly diagonal mass matrix,
-include \citet{CoJoTo93}, \citet{Kom97}, \citet{FaMaPaQu97}, \citet{CaGa97},
-\citet{KoVi98} and \citet{KoTr99}, whereas applications involving
-Chebyshev basis functions and a nondiagonal mass matrix include \citet{SePr94},
-\citet{PrCaSe94} and \citet{SePrPr95}.\\
-
-All SPECFEM3D software is written in Fortran2003 with full portability
-in mind, and conforms strictly to the Fortran2003 standard. It uses
-no obsolete or obsolescent features of Fortran. The package uses parallel
-programming based upon the Message Passing Interface (MPI) \citep{GrLuSk94,Pac97}.\\
-
-SPECFEM3D won the Gordon Bell award for best performance at the SuperComputing~2003
-conference in Phoenix, Arizona (USA) (see \citet{KoTsChTr03} and
-\url{www.sc-conference.org/sc2003/nrfinalaward.html}). It was a finalist
-again in 2008 for a run at 0.16 petaflops (sustained) on 149,784 processors
-of the `Jaguar' Cray XT5 system at Oak Ridge National Laboratories
-(USA) \citep{CaKoLaTiMiLeSnTr08}. It also won the BULL Joseph Fourier
-supercomputing award in 2010.\\
-
-It reached the sustained one petaflop performance level for the first time in February 2013
-on the Blue Waters Cray supercomputer at the National Center for Supercomputing Applications (NCSA), located at the University of Illinois at Urbana-Champaign (USA).\\
-
-This new release of the code includes Convolution or
-Auxiliary Differential Equation Perfectly Matched absorbing Layers
-(C-PML or ADE-PML) \citep{MaKoEz08,MaKoGe08,MaKo09,MaKoGeBr10,KoMa07}.
-It also includes support for GPU graphics card acceleration \citep{Kom11,MiKo10,KoMiEr09,KoErGoMi10}.\\
-
-The next release of the code will use the PT-SCOTCH parallel and threaded
-version of SCOTCH for mesh partitioning instead of the serial version.\\
-
-SPECFEM3D Cartesian includes coupled fluid-solid domains and adjoint
-capabilities, which enables one to address seismological inverse problems,
-but for linear rheologies only so far. To accommodate visco-plastic
-or non-linear rheologies, the reader can refer to the \href{http://geoelse.stru.polimi.it/}{GeoELSE}software
-package \citep{CaGa97,StPaIg09}.
-
-\section{Citation}
-
-If you use SPECFEM3D Cartesian for your own research, please cite
-at least one of the following articles:
-\begin{description}
-\item [{Numerical simulations in general}] ~\\
- Forward and adjoint simulations are described in detail in \citet{TrKoLi08,PeKoLuMaLeCaLeMaLiBlNiBaTr11,VaCaSaKoVi99,KoMiEr09,KoErGoMi10,ChKoViCaVaFe07,MaKoDi09,KoViCh10,CaKoLaTiMiLeSnTr08,TrKoHjLiZhPeBoMcFrTrHu10,KoRiTr02,KoTr02a,KoTr02b,KoTr99}
-or \citet{KoVi98}. Additional aspects of adjoint simulations are
-described in \citet{TrTaLi05,LiTr06,TrKoLi08,LiTr08,TrKoHjLiZhPeBoMcFrTrHu10,PeKoLuMaLeCaLeMaLiBlNiBaTr11}.
-Domain decomposition is explained in detail in \citet{MaKoBlLe08},
-and excellent scaling up to 150,000 processor cores is shown for instance
-in \citet{CaKoLaTiMiLeSnTr08,KoLaMi08a,MaKoBlLe08,KoErGoMi10,Kom11},
-\item [{GPU computing}] ~\\
- Computing on GPU graphics cards for acoustic or seismic wave propagation
-applications is described in detail in \citet{Kom11,MiKo10,KoMiEr09,KoErGoMi10}.
-\end{description}
-\noindent If you use this new version, which has non blocking
-MPI for much better performance for medium or large runs, please cite
-at least one of these six articles, in which results of non blocking
-MPI runs are presented: \citet{PeKoLuMaLeCaLeMaLiBlNiBaTr11,KoErGoMi10,KoViCh10,Kom11,CaKoLaTiMiLeSnTr08,MaKoBlLe08}.\\
-
-\noindent If you work on geophysical applications, you may be interested
-in citing some of these application articles as well, among others:
-\begin{description}
-\item [{Southern California simulations}] ~\\
- \citet{KoLiTrSuStSh04,KrJiKoTr06a,KrJiKoTr06b}.
-
-If you use the 3D southern California model, please cite \citet{SuSh03}
-(Los Angeles model), \citet{lovelyetal06} (Salton Trough), and \citet{hauksson2000}
-(southern California). The Moho map was determined by \citet{zhu&kanamori2000}.
-The 1D SoCal model was developed by \citet{DrHe90}.
-
-\item [{Anisotropy}] ~\\
- \citet{ChTr07,JiTsKoTr05,ChFaKo04,FaChKo04,RiRiKoTrHe02,TrKo00}.
-\item [{Attenuation}] ~\\
- \citet{SaKoTr10,KoTr02a,KoTr99}.
-\item [{Topography}] ~\\
- \citet{LeKoHuTr09,LeChKoHuTr09,LeChLiKoHuTr08,GoAmTaCaSmSaMaKo09,WiKoScTr04}.
-\end{description}
-The corresponding Bib\TeX{} entries may be found in file \texttt{doc/USER\_MANUAL/bibliography.bib}.
-
-\section{Support}
-
-This material is based upon work supported by the USA National Science
-Foundation under Grants No. EAR-0406751 and EAR-0711177, by the French
-CNRS, French INRIA Sud-Ouest MAGIQUE-3D, French ANR NUMASIS under
-Grant No. ANR-05-CIGC-002, and European FP6 Marie Curie International
-Reintegration Grant No. MIRG-CT-2005-017461. Any opinions, findings,
-and conclusions or recommendations expressed in this material are
-those of the authors and do not necessarily reflect the views of the
-USA National Science Foundation, CNRS, INRIA, ANR or the European
-Marie Curie program.
-
-\chapter{\label{cha:Getting-Started}Getting Started}
-
-To download the SPECFEM3D\_Cartesian software package, type this:
-\begin{lyxcode}
-git clone --recursive --branch devel https://github.com/geodynamics/specfem3d.git
-\end{lyxcode}
-{\small \par}
-
-We recommend that you add {\texttt{ulimit -S -s unlimited}} to your
-{\texttt{.bash\_profile}} file and/or {\texttt{limit stacksize
-unlimited }} to your {\texttt{.cshrc}} file to suppress any potential
-limit to the size of the Unix stack.\\
-
-Then, to configure the software for your system, run the \texttt{configure}
-shell script. This script will attempt to guess the appropriate configuration
-values for your system. However, at a minimum, it is recommended that
-you explicitly specify the appropriate command names for your Fortran
-compiler and MPI package (another option is to define FC, CC and MPIF90
-in your .bash\_profile or your .cshrc file):
-\begin{lyxcode}
-./configure~FC=ifort~MPIFC=mpif90
-\end{lyxcode}
-Note that MPI must be installed with MPI-IO enabled because parts of SPECFEM3D perform I/Os through MPI-IO.
-
-Before running the \texttt{configure} script, you should probably
-edit file \texttt{flags.guess} to make sure that it contains the best
-compiler options for your system. Known issues or things to check
-are:
-\begin{description}
-\item [{\texttt{Intel ifort compiler}}] See if you need to add \texttt{-assume byterecl} for your machine.
-\textbf{In the case of that compiler, we have noticed that versions dot zero sometimes have bugs or issues that can lead to wrong results when running the code, thus we \emph{strongly} recommend using versions dot one or above (for instance version 13.1 instead of 13.0, version 14.1 instead of 14.0 and so on)}.
-\item [{\texttt{IBM compiler}}] See if you need to add \texttt{-qsave}
-or \texttt{-qnosave} for your machine.
-\item [{\texttt{Mac OS}}] You will probably need to install \texttt{XCODE}.
-\end{description}
-When compiling on an IBM machine with the \texttt{xlf} and \texttt{xlc}
-compilers, we suggest running the \texttt{configure} script with the
-following options:\\
-
-\noindent \texttt{./configure FC=xlf90\_r MPIFC=mpif90 CC=xlc\_r CFLAGS=\textquotedbl{}-O3
--q64\textquotedbl{} FCFLAGS=\textquotedbl{}-O3 -q64\textquotedbl{}
--with-scotch-dir=...}.\\
-
-On SGI systems, \texttt{flags.guess} automatically informs \texttt{configure}
-to insert `\texttt{`TRAP\_FPE=OFF}'' into the generated \texttt{Makefile}
-in order to turn underflow trapping off.\\
-
-You can add \texttt{--enable-vectorization} to the configuration options to speed up the code in the fluid (acoustic) and elastic parts.
-This works fine if (and only if) your computer always allocates a contiguous memory block for each allocatable array;
-this is the case for most machines and most compilers, but not all. To disable this feature, use option \texttt{--disable-vectorization}.
-For more details see \href{https://github.com/geodynamics/specfem3d/issues/81}{github.com/geodynamics/specfem3d/issues/81} .
-To check if that option works fine on your machine, run the code with and without it for an acoustic/elastic model and make sure the seismograms are identical.\\
-
-Note that we use CUBIT (now called Trelis) to create meshes of hexahedra, but other packages
-can be used as well, for instance GiD from \url{http://gid.cimne.upc.es}
-or Gmsh from \url{http://geuz.org/gmsh} \citep{GeRe09}. Even mesh
-creation packages that generate tetrahedra, for instance TetGen from
-\url{http://tetgen.berlios.de}, can be used because each tetrahedron
-can then easily be decomposed into four hexahedra as shown in the
-picture of the TetGen logo at \url{http://tetgen.berlios.de/figs/Delaunay-Voronoi-3D.gif};
-while this approach does not generate hexahedra of optimal quality,
-it can ease mesh creation in some situations and it has been shown
-that the spectral-element method can very accurately handle distorted
-mesh elements \citep{OlSe11}.
-
-The SPECFEM3D Cartesian software package relies on the SCOTCH library
-to partition meshes created with CUBIT. METIS \citep{KaKu98a,KaKu98c,KaKu98b}
-can also be used instead of SCOTCH if you prefer, by editing file
-src/decompose\_mesh/decompose\_mesh.F90 and uncommenting the flag
-USE\_METIS\_INSTEAD\_OF\_SCOTCH. You will also then need to install
-and compile Metis version 4.0 (do {*}NOT{*} install Metis version
-5.0, which has incompatible function calls) and edit \texttt{src/decompose\_mesh/Makefile.in}
-and uncomment the METIS link flag in that file before running \texttt{configure}.\\
-
-The SCOTCH library \citep{PeRo96} provides efficent static mapping,
-graph and mesh partitioning routines. SCOTCH is a free software package
-developed by Fran\c{c}ois Pellegrini et al. from LaBRI and INRIA in Bordeaux,
-France, downloadable from the web page \url{https://gforge.inria.fr/projects/scotch/}.
-In case no SCOTCH libraries can be found on the system, the configuration
-will bundle the version provided with the source code for compilation.
-The path to an existing SCOTCH installation can to be set explicitly
-with the option \texttt{-{}-with-scotch-dir}. Just as an example:
-\begin{lyxcode}
-./configure~FC=ifort~MPIFC=mpif90~-{}-with-scotch-dir=/opt/scotch
-\end{lyxcode}
-If you use the Intel ifort compiler to compile the code, we recommend
-that you use the Intel icc C compiler to compile Scotch, i.e., use:
-\begin{lyxcode}
-./configure~CC=icc~FC=ifort~MPIFC=mpif90
-\end{lyxcode}
-
-When compiling the SCOTCH source code, if you get a message such as: "ld: cannot find -lz",
-the Zlib compression development library is probably missing on your machine and you will need to install it or ask your system administrator to
-do so. On Linux machines the package is often called "zlib1g-dev" or similar. (thus "sudo apt-get install zlib1g-dev" would install it)
-
-To compile a serial version of the code for small meshes that fits
-on one compute node and can therefore be run serially, run \texttt{configure}
-with the \texttt{-{}-without-mpi} option to suppress all calls to
-MPI.
-
-A summary of the most important configuration variables follows.
-\begin{description}
-\item [{\texttt{F90}}] Path to the Fortran compiler.
-\item [{\texttt{MPIF90}}] Path to MPI Fortran.
-\item [{\texttt{MPI\_FLAGS}}] Some systems require this flag to link to
-MPI libraries.
-\item [{\texttt{FLAGS\_CHECK}}] Compiler flags.
-\end{description}
-The configuration script automatically creates for each executable
-a corresponding \texttt{Makefile} in the \texttt{src/} subdirectoy.
-The \texttt{Makefile} contains a number of suggested entries for various
-compilers, e.g., Portland, Intel, Absoft, NAG, and Lahey. The software
-has run on a wide variety of compute platforms, e.g., various PC clusters
-and machines from Sun, SGI, IBM, Compaq, and NEC. Select the compiler
-you wish to use on your system and choose the related optimization
-flags. Note that the default flags in the \texttt{Makefile} are undoubtedly
-not optimal for your system, so we encourage you to experiment with
-these flags and to solicit advice from your systems administrator.
-Selecting the right compiler and optimization flags can make a tremendous
-difference in terms of performance. We welcome feedback on your experience
-with various compilers and flags.
-
-Now that you have set the compiler information, you need to select
-a number of flags in the \texttt{constants.h} file depending on your
-system:
-\begin{description}
-\item [{\texttt{LOCAL\_PATH\_IS\_ALSO\_GLOBAL}}] Set to \texttt{.false.}
-on most cluster applications. For reasons of speed, the (parallel)
-distributed database generator typically writes a (parallel) database
-for the solver on the local disks of the compute nodes. Some systems
-have no local disks, e.g., BlueGene or the Earth Simulator, and other
-systems have a fast parallel file system, in which case this flag
-should be set to \texttt{.true.}. Note that this flag is not used
-by the database generator or the solver; it is only used for some
-of the post-processing.
-\end{description}
-The package can run either in single or in double precision mode.
-The default is single precision because for almost all calculations
-performed using the spectral-element method using single precision
-is sufficient and gives the same results (i.e. the same seismograms);
-and the single precision code is faster and requires exactly half
-as much memory. Select your preference by selecting the appropriate
-setting in the \texttt{constants.h} file:
-\begin{description}
-\item [{\texttt{CUSTOM\_REAL}}] Set to \texttt{SIZE\_REAL} for single precision
-and \texttt{SIZE\_DOUBLE} for double precision.
-\end{description}
-In the \texttt{precision.h} file:
-\begin{description}
-\item [{\texttt{CUSTOM\_MPI\_TYPE}}] Set to \texttt{MPI\_REAL} for single
-precision and \texttt{MPI\_DOUBLE\_PRECISION} for double precision.
-\end{description}
-On many current processors (e.g., Intel, AMD, IBM Power), single precision
-calculations are significantly faster; the difference can typically
-be 10\% to 25\%. It is therefore better to use single precision. What
-you can do once for the physical problem you want to study is run
-the same calculation in single precision and in double precision on
-your system and compare the seismograms. If they are identical (and
-in most cases they will), you can select single precision for your
-future runs.\\
-
-If your compiler has problems with the \texttt{use mpi} statements that are used in the code, use the script called
-\texttt{replace\_use\_mpi\_with\_include\_mpif\_dot\_h.pl} in the root directory to replace all of them with \texttt{include 'mpif.h'} automatically.
-
-\section{Adding OpenMP support in addition to MPI}
-
-NOTE FROM JULY 2013: OpenMP support is maybe / probably not maintained any more.
-Thus the section below is maybe obsolete.
-
-OpenMP support can be enabled in addition to MPI. However, in many
-cases performance will not improve because our pure MPI implementation
-is already heavily optimized and thus the resulting code will in fact
-be slightly slower. A possible exception could be IBM BlueGene-type
-architectures.
-
-To enable OpenMP, uncomment the OpenMP compiler option in two lines
-in file \texttt{src/specfem3D/Makefile.in} (before running \texttt{configure})
-and also uncomment the \texttt{\#define USE\_OPENMP} statement in
-file\\
- \texttt{src/specfem3D/specfem3D.F90}.
-
-The DO-loop using OpenMP threads has a SCHEDULE property. The OMP\_SCHEDULE
-environment variable can set the scheduling policy of that DO-loop.
-Tests performed by Marcin Zielinski at SARA (The Netherlands) showed
-that often the best scheduling policy is DYNAMIC with the size of
-the chunk equal to the number of OpenMP threads, but most preferably
-being twice as the number of OpenMP threads (thus chunk size = 8 for
-4 OpenMP threads etc). If OMP\_SCHEDULE is not set or is empty, the
-DO-loop will assume generic scheduling policy, which will slow down
-the job quite a bit.
-
-\section{Compiling on an IBM BlueGene}
-
-\underline{More recent installation instruction for IBM BlueGene, from April 2013:}\\
-
-\noindent Edit file \texttt{flags.guess} and put this for \texttt{FLAGS\_CHECK}:\\
-
-\noindent \texttt{-g -qfullpath -O2 -qsave -qstrict -qtune=qp -qarch=qp
--qcache=auto -qhalt=w}\\
- \texttt{-qfree=f90 -qsuffix=f=f90 -qlanglvl=95pure -Q -Q+rank,swap\_all
--Wl,-relax}\\
-
-\noindent The most relevant are the -qarch and -qtune flags, otherwise
-if these flags are set to ``auto'' then they are wrongly assigned
-to the architecture of the frond-end node, which is different from
-that on the compute nodes. You will need to set these flags to the
-right architecture for your BlueGene compute nodes, which is not necessarily
-``qp''; ask your system administrator. On some machines if is necessary
-to use -O2 in these flags instead of -O3 due to a compiler bug of
-the XLF version installed. We thus suggest to first try -O3, and then
-if the code does not compile or does not run fine then switch back
-to -O2. The debug flags (-g, -qfullpath) do not influence performance
-but are useful to get at least some insights in case of problems.\\
-
-\noindent Before running \texttt{configure}, select the XL Fortran
-compiler by typing \texttt{module load bgq-xl/1.0} or \texttt{module
-load bgq-xl} (another, less efficient option is to load the GNU compilers
-using \texttt{module load bgq-gnu/4.4.6} or similar).\\
-
-\noindent Then, to configure the code, type this:\\
-
-\noindent \texttt{./configure FC=bgxlf90\_r MPIFC=mpixlf90\_r CC=bgxlc\_r
-LOCAL\_PATH\_IS\_ALSO\_GLOBAL=true}\\
-
-In order for the SCOTCH domain decomposer to compile, on some (but
-not all) Blue Gene systems you may need to run \texttt{configure}
-with \texttt{CC=gcc} instead of \texttt{CC=bgxlc\_r}.
-
-\noindent \underline{Older installation instruction for IBM BlueGene, from 2011:}\\
-
-\noindent To compile the code on an IBM BlueGene, Laurent L\'eger from
-IDRIS, France, suggests the following: compile the code with
-
-\noindent \texttt{ FLAGS\_CHECK=\textquotedbl{}-O3 -qsave -qstrict
--qtune=auto -qarch=450d -qcache=auto }~\\
- \texttt{ -qfree=f90 -qsuffix=f=f90 -g -qlanglvl=95pure -qhalt=w -Q
--Q+rank,swap\_all -Wl,-relax\textquotedbl{}}
-
-\noindent Option \textquotedbl{}-Wl,-relax\textquotedbl{} must be
-added on many (but not all) BlueGene systems to be able to link the
-binaries \texttt{xmeshfem3D} and \texttt{xspecfem3D} because the final
-link step is done by the GNU \texttt{ld} linker even if one uses \texttt{FC=bgxlf90\_r,
-MPIFC=mpixlf90\_r} and \texttt{CC=bgxlc\_r} to create all the object
-files. On the contrary, on some BlueGene systems that use the native
-AIX linker option \textquotedbl{}-Wl,-relax\textquotedbl{} can lead
-to problems and must be suppressed from \texttt{flags.guess}.
-
-\noindent Also, \texttt{AR=ar, ARFLAGS=cru} and \texttt{RANLIB=ranlib}
-are hardwired in all \texttt{Makefile.in} files by default, but to
-cross-compile on BlueGene/P one needs to change these values to \texttt{AR=bgar,
-ARFLAGS=cru} and \texttt{RANLIB=bgranlib}. Thus the easiest thing
-to do is to modify all \texttt{Makefile.in} files and the \texttt{configure}
-script to set them automatically by \texttt{configure}. One then just
-needs to pass the right commands to the \texttt{configure} script:
-
-\noindent \texttt{ ./configure -{}-prefix=/path/to/SPECFEM3DG\_SP
--{}-host=Babel -{}-build=BGP }~\\
- \texttt{ FC=bgxlf90\_r MPIFC=mpixlf90\_r CC=bgxlc\_r AR=bgar ARFLAGS=cru
-}~\\
- \texttt{ RANLIB=bgranlib LOCAL\_PATH\_IS\_ALSO\_GLOBAL=false}
-
-\noindent This trick can be useful for all hosts on which one needs
-to cross-compile.
-
-\noindent On BlueGene, one also needs to run the \texttt{xcreate\_header\_file}
-binary file manually rather than in the Makefile:
-
-\noindent \texttt{bgrun -np 1 -mode VN -exe ./bin/xcreate\_header\_file}
-
-\section{Visualizing the subroutine calling tree of the source code}
-
-\noindent
-Packages such as \texttt{Doxywizard} can be used to visualize the calling tree of the
-subroutines of the  source code. \texttt{Doxywizard} is
-a GUI front-end for configuring and running \texttt{Doxygen}.
-
-\bigskip
-
-\noindent To do your own call graphs, you can follow these simple steps below.
-
-\begin{enumerate}
-\setcounter{enumi}{-1}
-\item Install \texttt{Doxygen} \underline{and} \texttt{graphviz} (the two are usually in the package manager of classic Linux distribution).
-\item Run in the terminal : \texttt{doxygen -g}, which creates a \texttt{Doxyfile} that tells doxygen what you want it to do.
-\item Edit the Doxyfile. Two Doxyfile-type files have been already committed in the directory \\
-\noindent \texttt{specfem3d/doc/Call\_trees}:
-
-\begin{itemize}
-\item[\textbullet] \texttt{Doxyfile\_truncated\_call\_tree} will generate call graphs with maximum 3 or 4 levels of tree structure,
-\item[\textbullet] \texttt{Doxyfile\_complete\_call\_tree} will generate call graphs with complete tree structure.
-\end{itemize}
-
-The important entries in the Doxyfile are:
-
-\begin{description}
-\item [{\texttt{PROJECT\_NAME}}]
-\item [{\texttt{OPTIMIZE\_FOR\_FORTRAN}}] Set to YES
-\item [{\texttt{EXTRACT\_ALL}}] Set to YES
-\item [{\texttt{EXTRACT\_PRIVATE}}] Set to YES
-\item [{\texttt{EXTRACT\_STATIC}}] Set to YES
-\item [{\texttt{INPUT}}] From the directory \texttt{specfem3d/doc/Call\_trees}, it is \texttt{"../../src/"}
-\item [{\texttt{FILE\_PATTERNS}}] In SPECFEM case, it is \texttt{*.f90* *.F90* *.c* *.cu* *.h*}
-\item [{\texttt{HAVE\_DOT}}] Set to YES
-\item [{\texttt{CALL\_GRAPH}}] Set to YES
-\item [{\texttt{CALLER\_GRAPH}}] Set to YES
-\item [{\texttt{DOT\_PATH}}] The path where is located the dot program graphviz (if it is not in your \$PATH)
-\item [{\texttt{RECURSIVE}}] This tag can be used to turn specify whether or not subdirectories should be searched for input files as well. In the case of SPECFEM, set to YES.
-\item [{\texttt{EXCLUDE}}] Here, you can exclude:  \\
-\texttt{"../../src/specfem3D/older\_not\_maintained\_partial\_OpenMP\_port"}\\
-\texttt{"../../src/decompose\_mesh/scotch"}\\
-\texttt{"../../src/decompose\_mesh/scotch\_5.1.12b"}
-\item [{\texttt{DOT\_GRAPH\_MAX\_NODES}}] to set the maximum number of nodes that will be shown in the graph.
-If the number of nodes in a graph becomes larger than this value, doxygen will truncate the graph, which is visualized by representing a node as a red box.
-Minimum value: 0, maximum value: 10000, default value: 50.
-\item [{\texttt{MAX\_DOT\_GRAPH\_DEPTH}}] to set the maximum depth of the graphs generated by dot.
-A depth value of 3 means that only nodes reachable from the root by following a path via at most 3 edges will be shown.
-Using a depth of 0 means no depth restriction.
-Minimum value: 0, maximum value: 1000, default value: 0.
-
-\end{description}
-
-\item Run : \texttt{doxygen Doxyfile}, HTML and LaTeX files created by default in \texttt{html} and \texttt{latex} subdirectories.
-
-\item To see the call trees, you have to open the file \texttt{html/index.html} in your \underline{browser}.
-You will have many informations about each subroutines of SPECFEM (not only call graphs), you can click on every boxes / subroutines.
-It show you the call, and, the caller graph of each subroutine : the subroutines called by the concerned subroutine, and the previous subroutines who call this subroutine (the previous path), respectively.
-In the case of a truncated calling tree, the boxes with a red border indicates a node that has \underline{more} arrows than are shown (in other words: the graph is truncated with respect to this node).
-
-\end{enumerate}
-
-\medskip
-
-\noindent Finally, some useful links:
-
-\begin{itemize}
-\item[\textbullet] a good and short summary for the basic utilisation of Doxygen:
-
-\url{http://www.softeng.rl.ac.uk/blog/2010/jan/30/callgraph-fortran-doxygen/},
-
-\item[\textbullet] to configure the diagrams :
-
-\url{http://www.stack.nl/~dimitri/doxygen/manual/diagrams.html},
-
-\item[\textbullet] the complete alphabetical index of the tags in Doxyfile:
-
-\url{http://www.stack.nl/~dimitri/doxygen/manual/config.html},
-
-\item[\textbullet] more generally, the Doxygen manual:
-
-\url{http://www.stack.nl/~dimitri/doxygen/manual/index.html}.
-
-\end{itemize}
-
-\medskip
-
-\section{Using the ADIOS library for I/O}
-Regular POSIX I/O can be problematic when dealing with large simulations one large
-clusters (typically more than $10,000$ processes). SPECFEM3D use the ADIOS library~\cite{Liu2013}
-to deal transparently take advantage of advanced paralllel file system features. To enable
-ADIOS, the following steps should be done:
-\begin{enumerate}
-\item Install ADIOS (available from \url{https://www.olcf.ornl.gov/center-projects/adios/}).
-Make sure that your environment variables reference it.
-\item You may want to change ADIOS related values in the \texttt{constants.h.in} file.
-The default values probably suit most cases.
-\item Configure using the \texttt{--with-adios} flag.
-\end{enumerate}
-ADIOS is currently only usable for meshfem3D generated mesh (i.e. not for meshes generated
-with CUBTI). Additional control parameters are discussed in section~\ref{cha:Main-Parameter}.
-
-\section{Becoming a developer of the code, or making small modifications in the source code}
-
-If you want to develop new features in the code, and/or if you want to make small changes, improvements, or bug fixes, you are very welcome to contribute.
-To do so, i.e.
-to access the development branch of the source code with read/write access (in a safe way, no need to worry too much about breaking the package, there is a robot called BuildBot that is in charge of checking and validating all new contributions and changes), please visit this Web page: \url{https://github.com/geodynamics/specfem3d/wiki/Using-Hub}.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\chapter{\label{cha:Mesh-Generation}Mesh Generation}
-
-The first step in running a spectral-element simulation consists of
-constructing a high-quality mesh for the region under consideration.
-We provide two possibilities to do so: (1) relying on the external,
-hexahedral mesher CUBIT, or (2) using the provided, internal mesher
-\texttt{xmeshfem3D}. In the following, we explain these two approaches.
-
-\section{\label{cha:Running-the-Mesher-CUBIT}Meshing with \texttt{CUBIT}}
-
-CUBIT is a meshing tool suite for the creation of finite-element meshes
-for arbitrarily shaped models. It has been developed and maintained
-at Sandia National Laboratories and can be purchased for a small academic
-institutional fee at \url{http://cubit.sandia.gov}. Our experience
-showed that using CUBIT greatly facilitates and speeds up the generation
-and preparation of hexahedral, conforming meshes for a variety of
-geophysical models with increasing complexity.
-
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=0.6\textwidth]{figures/mount-cubit}
-\par\end{centering}
-
-\caption{Example of the graphical user interface of CUBIT. The hexahedral mesh
-shown in the main display consists of a hexahedral discretization
-of a single volume with topography.}
-
-\label{fig:mount.cubit}
-\end{figure}
-
-\noindent
-The basic steps in creating a load-balanced, partitioned mesh with
-CUBIT are:
-\begin{description}
-\item [{1.}] setting up a hexahedral mesh with CUBIT,
-\item [{2.}] exporting the CUBIT mesh into a SPECFEM3D Cartesian file format
-and
-\item [{3.}] partitioning the SPECFEM3D Cartesian mesh files for a chosen
-number of cores.
-\end{description}
-Examples are provided in the SPECFEM3D Cartesian package in the subdirectory
-\texttt{EXAMPLES/}. We strongly encourage you to contribute your own
-example to this package by contacting the CIG Computational Seismology
-Mailing List \urlwithparentheses{cig-seismo@geodynamics.org}.
-
-\subsection{Creating the Mesh with CUBIT}
-
-For the installation and handling of the CUBIT meshing tool suite,
-please refer to the CUBIT user manual and documentation. In order
-to give you a basic understanding of how to use CUBIT for our purposes,
-examples are provided in the SPECFEM3D Cartesian package in the subdirectory
-\texttt{EXAMPLES/}:
-\begin{description}
-\item [{\texttt{homogeneous\_halfspace}}] Creates a single block model
-and assigns elastic material parameters.
-\item [{\texttt{layered\_halfspace}}] Combines two different, elastic material
-volumes and creates a refinement layer between the two. This example
-can be compared for validation against the solutions provided in subdirectory
-~\\
- \texttt{VALIDATION\_3D\_SEM\_SIMPLER\_LAYER\_SOURCE\_DEPTH/}.
-\item [{\texttt{waterlayered\_halfspace}}] Combines an acoustic and elastic
-material volume as in a schematic marine survey example.
-\item [{\texttt{tomographic\_model}}] Creates a single block model whose
-material properties will have to be read in from a tomographic model
-file during the databases creation by \texttt{xgenerate\_databases}.
-\end{description}
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=0.45\textwidth]{figures/example-homogeneous}
-\includegraphics[width=0.45\textwidth]{figures/example-2layers} \\
- \includegraphics[width=0.45\textwidth]{figures/example-water} \includegraphics[width=0.45\textwidth]{figures/example-tomo}
-\par\end{centering}
-
-\caption{Screenshots of the CUBIT examples provided in subdirectory \texttt{EXAMPLES/}:
-homogeneous halfspace (top-left), layered halfspace (top-right), water
-layered halfspace (bottom-left) and tomographic model (bottom-right).}
-
-\label{fig:examples.cubit}
-\end{figure}
-
-In each example subdirectory you will find a \texttt{README} file,
-which explains in a step-by-step tutorial the workflow for the example.
-Please feel free to contribute your own example to this package by
-contacting the CIG Computational Seismology Mailing List \urlwithparentheses{cig-seismo@geodynamics.org}.
-
-In some cases, to re-create the meshes for the examples given, just type "claro ./create\_mesh.py" or similar
-from the command line ("claro" is the command to run CUBIT from the command line).
-
-IMPORTANT note: in order to correctly set up GEOCUBIT and run the examples, please read the file called "EXAMPLES/README";
-in particular, please make sure you correctly set up the Python paths as indicated in that file.
-
-\subsection{Exporting the Mesh with \texttt{run\_cubit2specfem3d.py} }
-
-\label{subsec:Exporting-the-Mesh}
-
-Once the geometric model volumes in CUBIT are meshed, you prepare
-the model for exportation with the definition of material blocks and
-boundary surfaces. Thus, prior to exporting the mesh, you need to
-define blocks specifying the materials and absorbing boundaries in
-CUBIT. This process could be done automatically using the script \texttt{run\_cubit2specfem3d.py}
-if the mesh meets some conditions or manually, following the block
-convention:
-\begin{description}
-\item [{material\_name}] Each material should have a specific block defined
-by a unique name. The name convention of the material is to start
-with either 'elastic' or 'acoustic'. It must be then followed by a
-unique identifier, e.g. 'elastic 1', 'elastic 2', etc. The additional
-attributes to the block define the material description.
-
-For an elastic material:
-\begin{description}
-\item [{material\_id}] An integer value which is unique for this material.
-\item [{Vp}] P-wave speed of the material (given in m/s).
-\item [{Vs}] S-wave speed of the material (given in m/s).
-\item [{rho}] density of the material (given in kg/m$^{3}$).
-\item [{Q}] quality factor to use in case of a simulation with attenuation
-turned on. It should be between 1 and 9000. In case no attenuation
-information is available, it can be set to zero. Please note that
-your Vp- and Vs-speeds are given for a reference frequency. To change
-this reference frequency, you change the value of \texttt{ATTENUATION\_f0\_REFERENCE}
-in the main constants file \texttt{constants.h} found in subdirectory
-\texttt{src/shared/}.
-\item [{anisotropic\_flag}] Flag describing the anisotropic model to use
-in case an anisotropic simulation should be conducted. See the file
-\texttt{model\_aniso.f90} in subdirectory \texttt{src/generate\_databases/}
-for an implementation of the anisotropic models. In case no anisotropy
-is available, it can be set to zero.
-\end{description}
-
-Note that this material block has to be defined using all the volumes
-which belong to this elastic material. For volumes belonging to another,
-different material, you will need to define a new material block.
-
-For an acoustic material:
-\begin{description}
-\item [{material\_id}] An integer value which is unique for this material.
-\item [{Vp}] P-wave speed of the material (given in m/s).
-\item [{0}] S-wave speed of the material is ignored.
-\item [{rho}] density of the material (given in kg/m$^{3}$).
-\end{description}
-\item [{face\_topo}] Block definition for the surface which defines the
-free surface (which can have topography). The name of this block must
-be 'face\_topo', the block has to be defined using all the surfaces
-which constitute the complete free surface of the model.
-\item [{face\_abs\_xmin}] Block definition for the faces on the absorbing
-boundaries, one block for each surface with x=Xmin.
-\item [{face\_abs\_xmax}] Block definition for the faces on the absorbing
-boundaries, one block for each surface with x=Xmax.
-\item [{face\_abs\_ymin}] Block definition for the faces on the absorbing
-boundaries, one block for each surface with y=Ymin.
-\item [{face\_abs\_ymax}] Block definition for the faces on the absorbing
-boundaries, one block for each surface with y=Ymax.
-\item [{face\_abs\_bottom}] Block definition for the faces on the absorbing
-boundaries, one block for each surface with z=bottom.
-\end{description}
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=0.45\textwidth]{figures/mount-surface} \includegraphics[width=0.45\textwidth]{figures/mount-abs}
-\par\end{centering}
-
-\caption{Example of the block definitions for the free surface 'face\_topo'
-(left) and the absorbing boundaries, defined in a single block 'face\_abs\_xmin'
-(right) in CUBIT.}
-
-\label{fig:mount.abs}
-\end{figure}
-
-Optionally, instead of specifying for each surface at the boundaries
-a single block like mentioned above, you can also specify a single
-block for all boundary surfaces and name it as one of the absorbing
-blocks above, e.g. 'face\_abs\_xmin'.
-
-After the block definitions are done, you export the mesh using the
-script \texttt{cubit2specfem3d.py} provided in each of the example
-directories (linked to the common script \texttt{CUBIT\_GEOCUBIT/cubit2specfem3d.py}).
-If the export was successful, you should find the following
-files in a subdirectory \texttt{MESH/}:
-\begin{description}
-\item [{absorbing\_cpml\_file}] \textbf{(only needed in case of C-PML absorbing
-conditions)} Contains on the first line the total number of C-PML
-spectral elements in the mesh, and then on the following line the
-list of all these C-PML elements with two numbers per line: first
-the spectral element number, and then a C-PML flag indicating to which
-C-PML layer(s) that element belongs, according to the following convention:
-
-\begin{itemize}
-\item Flag = 1 : element belongs to a X CPML layer only (either in Xmin
-or in Xmax),
-\item Flag = 2 : element belongs to a Y CPML layer only (either in Ymin
-or in Ymax),
-\item Flag = 3 : element belongs to a Z CPML layer only (either in Zmin
-or in Zmax),
-\item Flag = 4 : element belongs to a X CPML layer and also to a Y CPML
-layer,
-\item Flag = 5 : element belongs to a X CPML layer and also to a Z CPML
-layer,
-\item Flag = 6 : element belongs to a Y CPML layer and also to a Z CPML
-layer,
-\item Flag = 7 : element belongs to a X, to a Y and to a Z CPML layer, i.e.,
-it belongs to a CPML corner.
-\end{itemize}
-
-Note that it does not matter whether an element belongs to a Xmin
-or to a Xmax CPML, the flag is the same in both cases; the same is
-true for Ymin and Ymax, and also Zmin and Zmax.
-
-When you have an existing CUBIT (or similar) mesh stored in SPECFEM3D
-format, i.e., if you have existing \textquotedbl{}nodes\_coords\_file\textquotedbl{}
-and \textquotedbl{}mesh\_file\textquotedbl{} files but do not know
-how to assign CPML flags to them, we have created a small serial Fortran
-program that will do that automatically for you, i.e., which will
-create the \textquotedbl{}absorbing\_cpml\_file\textquotedbl{} for
-you. That program is utils/CPML/convert\_external\_layers\_of\_a\_given\_mesh\_to\_CPML\_layers.f90,
-and a small Makefile is provided in that directory (utils/CPML).
-
-IMPORTANT: it is your responsibility to make sure that in the input
-CUBIT (or similar) mesh that this code will read in SPECFEM3D format
-from files \textquotedbl{}nodes\_coords\_file\textquotedbl{} and \textquotedbl{}mesh\_file\textquotedbl{}
-you have created layers of elements that constitute a layer of constant
-thickness aligned with the coordinate grid axes (X, Y and/or Z), so
-that this code can assign CPML flags to them. This code does NOT check
-that (because it cannot, in any easy way). The mesh inside these CPML
-layers does not need to be structured nor regular, any non-structured
-mesh is fine as long as it has flat PML inner and outer faces, parallel
-to the axes, and thus of a constant thickness. The thickness can be
-different for the X, Y and Z sides. But for X it must not vary, for
-Y it must not vary, and for Z it must not vary. If you do not know
-the exact thickness, you can use a slightly LARGER value in this code
-(say 2\% to 5\% more) and this code will fix that and will adjust
-it; never use a SMALLER value otherwise this code will miss some CPML
-elements.
-
-Note: in a future release we will remove the constraint of having
-CPML layers aligned with the coordinate axes; we will allow for meshes
-that are titled by any constant angle in the horizontal plane. However
-this is not implemented yet.
-
-Note: in the case of fluid-solid layers in contact, the fluid-solid
-interface currently needs to be flat and horizontal inside the CPML
-layer (i.e., bathymetry should become flat and horizontal when entering
-the CPML); this (small) constraint will probably remain in the code
-for a while because it makes fluid-solid matching inside the CPML
-much easier.
-
-\item [{materials\_file}] Contains the material associations for each element.
-The format is:
-
-\begin{lyxcode}
-element\_ID~material\_ID
-\end{lyxcode}
-
-where \texttt{element\_ID} is the element identifier and \texttt{material\_ID}
-is a unique identifier, positive (for materials taken from this list
-of materials, i.e. for which each spectral element has constant material
-properties taken from this list) or negative (for tomographic models,
-i.e. for spectral element whose real velocities and density will be
-assigned later by calling an external function to define model variations,
-for instance in the case of tomographic models; in such a case, material
-properties can vary inside each spectral element, i.e. be different
-at each of its Gauss-Lobatto-Legendre grid points).
-
-\item [{nummaterial\_velocity\_file}] Defines the material properties.
-
-\begin{itemize}
-\item For classical materials (i.e., spectral elements for which the velocity
-and density model will not be assigned by calling an external function
-to define for instance a tomographic model), the format is:
-
-\begin{lyxcode}
-domain\_ID~material\_ID~rho~vp~vs~Qkappa~Qmu~anisotropy\_flag
-\end{lyxcode}
-
-where \texttt{\textbf{domain\_ID}}\textbf{ is 1 for acoustic and 2
-for elastic or viscoelastic materials,} \texttt{material\_ID} a unique
-identifier, \texttt{rho} the density in $kg\, m^{-3}$, \texttt{vp}
-the P-wave speed in $m\, s^{-1}$, \texttt{vs} the S-wave speed in
-$m\, s^{-1}$, \texttt{Q} the quality factor and \texttt{anisotropy\_flag}
-an identifier for anisotropic models. Note that the \texttt{Qkappa}
-value is ignored by the code unless \texttt{FULL\_ATTENUATION\_SOLID}
-is set. Note also that both \texttt{Qkappa} and \texttt{Qmu} are ignored
-by the code unless \texttt{ATTENUATION} is set. If you want a model
-with no \texttt{Qmu} attenuation, both set \texttt{ATTENUATION} to
-\texttt{.false.} in the \texttt{Par\_file} and set \texttt{Qmu} to
-9999 here. If you want a model with no \texttt{Qkappa} attenuation,
-both set \texttt{FULL\_ATTENUATION\_SOLID} to \texttt{.false.} in
-the \texttt{Par\_file} and set \texttt{Qkappa} to 9999 here. %%magnoni
-
-\item For tomographic velocity models, please read Chapter \ref{cha:-Changing-the}
-and Section \ref{sec:Using-tomographic} `Using external tomographic
-Earth models' for further details.
-\end{itemize}
-\item [{nodes\_coords\_file}] Contains the point locations in Cartesian
-coordinates of the mesh element corners.
-\item [{mesh\_file}] Contains the mesh element connectivity. The hexahedral
-elements can have 8 or 27 nodes.\\
- See picture doc/mesh\_numbering\_convention/numbering\_convention\_27\_nodes.jpg
-to see\\
- in which (standard) order the points must be cited. In the case of
-8 nodes, just include the first 8 points.
-\item [{free\_or\_absorbing\_surface\_file\_zmax}] Contains the free surface
-connectivity or \\
- the surface connectivity of the absorbing boundary surface at the
-top (Zmax), \\
- depending on whether the top surface is defined as free or absorbing
-(\texttt{STACEY\_INSTEAD\_OF\_FREE\_SURFACE} in \texttt{DATA/Par\_file}). \\
- You should put both the surface of acoustic regions and of elastic
-regions in that file; that is, list all the element faces that constitute
-the surface of the model in that file.
-\item [{absorbing\_surface\_file\_xmax}] Contains the surface connectivity
-of the absorbing boundary surface at Xmax \\
-(also needed in the case of C-PML absorbing conditions, in order for
-the code to be able to impose Dirichlet conditions on their outer
-edge).
-\item [{absorbing\_surface\_file\_xmin}] Contains the surface connectivity
-of the absorbing boundary surface at Xmin \\
-(also needed in the case of C-PML absorbing conditions, in order for
-the code to be able to impose Dirichlet conditions on their outer
-edge).
-\item [{absorbing\_surface\_file\_ymax}] Contains the surface connectivity
-of the absorbing boundary surface at Ymax \\
-(also needed in the case of C-PML absorbing conditions, in order for
-the code to be able to impose Dirichlet conditions on their outer
-edge).
-\item [{absorbing\_surface\_file\_ymin}] Contains the surface connectivity
-of the absorbing boundary surface at Ymin \\
-(also needed in the case of C-PML absorbing conditions, in order for
-the code to be able to impose Dirichlet conditions on their outer
-edge).
-\item [{absorbing\_surface\_file\_bottom}] Contains the surface connectivity
-of the absorbing boundary surface at the bottom (Zmin) \\
-(also needed in the case of C-PML absorbing conditions, in order for
-the code to be able to impose Dirichlet conditions on their outer
-edge).
-\end{description}
-These mesh files are needed as input files for the partitioner \texttt{xdecompose\_mesh}
-to load-balance the mesh. Please see the next section for further
-details.
-
-In directory "CUBIT\_GEOCUBIT" we provide a script that can help doing the above tasks of exporting a CUBIT mesh
-to SPECFEM3D format automatically for you: "run\_cubit2specfem3d.py".
-Just edit them to indicate the path to your local installation of CUBIT and also the name of the *.cub existing CUBIT mesh file
-that you want to export to SPECFEM3D format. These scripts will do the conversion for you automatically except assigning material properties
-to the different mesh layers. To do so, you will then need to edit the file
-called "nummaterial\_velocity\_file" that will have just been created and change it from the prototype created:
-
-    0 1 vol1 --> syntax: \#material\_domain\_id \#material\_id \#rho \#vp \#vs \#Q\_kappa \#Q\_mu \#anisotropy
-    0 2 vol2 --> syntax: \#material\_domain\_id \#material\_id \#rho \#vp \#vs \#Q\_kappa \#Q\_mu \#anisotropy
-
-(where "vol1" and "vol2" here represent the volume labels that you have set while creating the mesh in CUBIT)
-to for instance
-
-    2 1 1500 2300 1800 9999.0 9999.0 0
-    2 2 1600 2500 20000 9999.0 9999.0 0
-
-\subsubsection*{Checking the mesh quality}
-
-The quality of the mesh may be inspected more precisely based upon
-the serial code in the file \texttt{check\_mesh\_quality\_}~\\
- \texttt{CUBIT\_Abaqus.f90} located in the directory \texttt{src/check\_mesh\_quality\_CUBIT\_Abaqus/}.
-Running this code is optional because no information needed by the
-solver is generated.
-
-Prior to running and compiling this code, you have to export your
-mesh in CUBIT to an ABAQUS (.inp) format. For example, export mesh
-block IDs belonging to volumes in order to check the quality of the
-hexahedral elements. You also have to determine a number of parameters
-of your mesh, such as the number of nodes and number of elements and
-modify the header of the \texttt{check\_mesh\_quality\_CUBIT\_Abaqus.f90}
-source file in directory \texttt{src/check\_mesh\_quality\_CUBIT\_Abaqus/}.
-
-\noindent
-Then, in the main directory, type
-\begin{lyxcode}
-{\small make~xcheck\_mesh\_quality\_CUBIT\_Abaqus}{\small \par}
-\end{lyxcode}
-and use
-\begin{lyxcode}
-{\small ./bin/xcheck\_mesh\_quality\_CUBIT\_Abaqus}{\small \par}
-\end{lyxcode}
-to generate an %not supported: AVS output file (\texttt{\small AVS\_meshquality.inp} in AVS UCD format) or
-OpenDX output file (\texttt{\small DX\_mesh\_quality.dx}{\small )
-that can be used to investigate mesh quality, e.g. skewness of elements,
-and a Gnuplot histogram (}\texttt{\small mesh\_quality\_histogram.txt}{\small )
-that can be plotted with gnuplot (type `}\texttt{\small gnuplot plot\_mesh\_quality\_histogram.gnu}{\small ').
-The histogram is also printed to the screen. Analyze that skewness
-histogram of mesh elements to make sure no element has a skewness
-above approximately 0.75, otherwise the mesh is of poor quality (and
-if even a single element has a skewness value above 0.80, then you
-must definitely improve the mesh). If you want to start designing
-your own meshes, this tool is useful for viewing your creations. You
-are striving for meshes with elements with `cube-like' dimensions,
-e.g., the mesh should contain no very elongated or skewed elements.}{\small \par}
-
-\subsection{Partitioning the Mesh with \texttt{xdecompose\_mesh}}
-
-The SPECFEM3D Cartesian software package performs large scale simulations
-in a parallel 'Single Process Multiple Data' way. The spectral-element
-mesh created with CUBIT needs to be distributed on the processors.
-This partitioning is executed once and for all prior to the execution
-of the solver so it is referred to as a static mapping.
-
-An efficient partitioning is important because it leverages the overall
-running time of the application. It amounts to balance the number
-of elements in each slice while minimizing the communication costs
-resulting from the placement of adjacent elements on different processors.
-\texttt{decompose\_mesh} depends on the SCOTCH library \citep{PeRo96},
-which provides efficent static mapping, graph and mesh partitioning
-routines. SCOTCH is a free software package developed by Fran\c{c}ois
-Pellegrini et al. from LaBRI and INRIA in Bordeaux, France, downloadable
-from the web page \url{https://gforge.inria.fr/projects/scotch/}.
-\\
-
-In most cases, the configuration with \texttt{./configure~FC=ifort}
-should be sufficient. During the configuration process, the script
-tries to find existing SCOTCH installations. In case your system has
-no pre-existing SCOTCH installation, we provide the source code of
-SCOTCH, which is released open source under the French CeCILL-C version
-1 license, in directory \texttt{src/decompose\_mesh/scotch\_5.1.12b}.
-This version gets bundled with the compilation of the SPECFEM3D Cartesian
-package if no libraries could have been found. If this automatic compilation
-of the SCOTCH libraries fails, please refer to file INSTALL.txt in
-that directory to see further details how to compile it on your system.
-In case you want to use a pre-existing installation, make sure you
-have correctly specified the path of the SCOTCH library when using
-the option \texttt{-{}-with-scotch-dir} with the \texttt{./configure}
-script. In the future you should be able to find more recent versions
-at \url{http://www.labri.fr/perso/pelegrin/scotch/scotch_en.html}
-. \\
-
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=0.45\textwidth]{figures/mount-partitions}
-\includegraphics[width=0.45\textwidth]{figures/mount-partitions2}
-\par\end{centering}
-
-\caption{Example of a mesh partitioning onto four cores. Each single core partition
-is colored differently. The executable \texttt{xdecompose\_mesh} can
-equally distribute the mesh on any arbitrary number of cores. Domain
-decomposition is explained in detail in \citet{MaKoBlLe08}, and excellent
-scaling up to 150,000 processor cores in shown for instance in \citet{CaKoLaTiMiLeSnTr08,KoLaMi08a,MaKoBlLe08,KoErGoMi10,Kom11}.}
-
-\label{fig:mount.partitions}
-\end{figure}
-
-When you are ready to compile, in the main directory type
-\begin{lyxcode}
-{\small make~xdecompose\_mesh}{\small \par}
-\end{lyxcode}
-If all paths and flags have been set correctly, the executable \texttt{bin/xdecompose\_mesh}
-should be produced.
-
-The partitioning is done in serial for now (in the next release we
-will provide a parallel version of that code). It needs to be run in the main directory because it expects the \texttt{./DATA/Par\_file}. The synopsis is:
-\begin{lyxcode}
-./bin/xdecompose\_mesh~nparts~input\_directory~output\_directory
-\end{lyxcode}
-where
-\begin{itemize}
-\item \texttt{nparts} is the number of partitions, i.e., the number of cores
-for the parallel simulations,
-\item \texttt{input\_directory} is the directory which holds all the files
-generated by the Python script \texttt{cubit2specfem3d.py} explained
-in the previous Section~\ref{subsec:Exporting-the-Mesh}, e.g. \texttt{./MESH/},
-and
-\item \texttt{output\_directory} is the directory for the output of this
-partitioner which stores ACII-format files named like \texttt{proc{*}{*}{*}{*}{*}{*}\_Database}
-for each partition. These files will be needed for creating the distributed
-databases, and have to reside in the directory \texttt{LOCAL\_PATH}
-specified in the main \texttt{Par\_file}, e.g. in directory \texttt{./OUTPUT\_FILES/DATABASES\_MPI}.
-Please see Chapter~\ref{cha:Creating-Distributed-Databases} for
-further details.
-\end{itemize}
-Note that all the files generated by the Python script \texttt{cubit2specfem3d.py}
-must be placed in the \texttt{input\_directory} folder before running
-the program.
-
-\section{\label{cha:Running-the-Mesher-Meshfem3D}Meshing with \texttt{xmeshfem3D}}
-
-In case you successfully ran the configuration script, you are also
-ready to compile the internal mesher. This is an alternative to CUBIT
-for the mesh generation of relatively simple models.
-
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[scale=0.5]{figures/socal_map_mpi}
-\par\end{centering}
-
-\caption{\label{fig:For-parallel-computing}For parallel computing purposes,
-the model block is subdivided in $\nprocxi\times\nproceta$ slices
-of elements. In this example we use $5^{2}=25$ processors. }
-\end{figure}
-
-\noindent
-In the main directory, type
-\begin{lyxcode}
-{\small make~xmeshfem3D}{\small \par}
-\end{lyxcode}
-If all paths and flags have been set correctly, the mesher should
-now compile and produce the executable \texttt{bin/xmeshfem3D}. Please
-note that \texttt{xmeshfem3D} must be called directly from the main
-directory, as most of the binaries of the package.
-
-Input for the mesh generation program is provided through the parameter
-file \texttt{Mesh\_Par\_file}, which resides in the subdirectory \texttt{DATA/meshfem3D\_files/}.
-(to see how to use it, see the EXAMPLES specific to the internal mesher in directory \texttt{EXAMPLES/meshfem3D\_examples/}.
-Before running the mesher, a number of parameters need to be set in
-the \texttt{Mesh\_Par\_file}. This requires a basic understanding
-of how the SEM is implemented, and we encourage you to read \citet{KoVi98,KoTr99}
-and \citet{KoLiTrSuStSh04}.
-
-The mesher and the solver use UTM coordinates internally, therefore
-you need to define the zone number for the UTM projection (e.g., zone
-11 for Los Angeles). Use decimal values for latitude and longitude
-(no minutes/seconds). These values are approximate; the mesher will
-round them off to define a square mesh in UTM coordinates. When running
-benchmarks on rectangular models, turn the UTM projection off by using
-the flag \texttt{\small SUPPRESS\_UTM\_PROJECTION}{\small , in which
-case all `longitude' parameters simply refer to the $x$~axis, and
-all `latitude' parameters simply refer to the $y$~axis. To run the
-mesher for a global simulation, the following parameters need to be
-set in the }\texttt{\small Mesh\_Par\_file}{\small :}{\small \par}
-\begin{description}
-\item [{\texttt{LATITUDE\_MIN}}] Minimum latitude in the block (negative
-for South).
-\item [{\texttt{LATITUDE\_MAX}}] Maximum latitude in the block.
-\item [{\texttt{LONGITUDE\_MIN}}] Minimum longitude in the block (negative
-for West).
-\item [{\texttt{LONGITUDE\_MAX}}] Maximum longitude in the block.
-\item [{\texttt{DEPTH\_BLOCK\_KM}}] Depth of bottom of mesh in kilometers.
-\item [{\texttt{\noun{UTM\_PROJECTION\_ZONE}}}] UTM projection zone in
-which your model resides, only valid when \texttt{SUPPRESS\_UTM\_}~\\
- \texttt{PROJECTION} is \texttt{.false.}.
-Use a negative zone number for the Southern hemisphere:
-the Northern hemisphere corresponds to zones +1 to +60,
-the Southern hemisphere to zones -1 to -60.
-We use the WGS84 (World Geodetic System 1984) reference ellipsoid for the UTM projection. If you prefer to use the Clarke 1866 ellipsoid,
-edit file \texttt{src/shared/utm\_geo.f90}, uncomment that ellipsoid and recompile the code.
-From http://en.wikipedia.org/wiki/Universal\_Transverse\_Mercator\_coordinate\_system :
-The Universal Transverse Mercator coordinate system was developed by the United States Army Corps of Engineers in the 1940s.
-The system was based on an ellipsoidal model of Earth. For areas within the contiguous United States
-the Clarke Ellipsoid of 1866 was used. For the remaining areas of Earth, including Hawaii, the International Ellipsoid was used.
-The WGS84 ellipsoid is now generally used to model the Earth in the UTM coordinate system,
-which means that current UTM northing at a given point can be 200+ meters different from the old one.
-For different geographic regions, other datum systems (e.g.: ED50, NAD83) can be used.
-\item [{\texttt{SUPPRESS\_UTM\_PROJECTION}}] set to be \texttt{.false.}
-when your model range is specified in geographical coordinates, and
-needs to be \texttt{.true.} when your model is specified in Cartesian
-coordinates. \noun{UTM projection zone in which your simulation region
-resides.}
-\item [{\texttt{INTERFACES\_FILE }}] File which contains the description
-of the topography and of the interfaces between the different layers
-of the model, if any. The number of spectral elements in the vertical
-direction within each layer is also defined in this file.
-\item [{$\nexxi$}] The number of spectral elements along one side of the
-block. This number \textit{must} be 8~$\times$~a multiple of $\nprocxi$
-defined below. Based upon benchmarks against semi-analytical discrete
-wavenumber synthetic seismograms \citep{KoLiTrSuStSh04}, determined
-that a $\nexxi=288$ run is accurate to a shortest period of roughly
-2~s. Therefore, since accuracy is determined by the number of grid
-points per shortest wavelength, for any particular value of $\nexxi$
-the simulation will be accurate to a shortest period determined by
-\begin{equation}
-\mbox{shortest period (s)}=(288/\nexxi)\times2.\label{eq:shortest_period}
-\end{equation}
- The number of grid points in each orthogonal direction of the reference
-element, i.e., the number of Gauss-Lobatto-Legendre points, is determined
-by \texttt{NGLLX} in the \texttt{constants.h} file. We generally use
-$\mbox{\texttt{NGLLX\/}}=5$, for a total of $5^{3}=125$ points per
-elements. We suggest not to change this value.
-\item [{$\nexeta$}] The number of spectral elements along the other side
-of the block. This number \textit{must} be 8~$\times$~a multiple
-of $\nproceta$ defined below.
-\item [{$\nprocxi$}] The number of processors or slices along one side
-of the block (see Figure~\ref{fig:For-parallel-computing}); we must
-have $\nexxi=8\times c\times\nprocxi$, where $c\ge1$ is a positive
-integer.
-\item [{$\nproceta$}] The number of processors or slices along the other
-side of the block; we must have $\nexeta=8\times c\times\nproceta$,
-where $c\ge1$ is a positive integer.
-\item [{\texttt{USE\_REGULAR\_MESH}}] set to be \texttt{.true.} if you
-want a perfectly regular mesh or \texttt{.false.} if you want to add
-doubling horizontal layers to coarsen the mesh. In this case, you
-also need to provide additional information by setting up the next
-three parameters.
-\item [{\texttt{NDOUBLINGS}}] The number of horizontal doubling layers.
-Must be set to \texttt{1} or \texttt{2} if \texttt{USE\_REGULAR\_MESH}
-is set to \texttt{.true.}.
-\item [{\texttt{NZ\_DOUBLING\_1}}] The position of the first doubling layer
-(only interpreted if \texttt{USE\_REGULAR\_MESH} is set to \texttt{.true.}).
-\item [{\texttt{NZ\_DOUBLING\_2}}] The position of the second doubling
-layer (only interpreted if \texttt{USE\_REGULAR\_MESH} is set to \texttt{.true.}
-and if \texttt{NDOUBLINGS} is set to \texttt{2}).
-\item [{\texttt{CREATE\_ABAQUS\_FILES}}] Set this flag to \texttt{.true.}
-to save Abaqus FEA \urlwithparentheses{www.simulia.com} mesh files
-for subsequent viewing. Turning the flag on generates files in the
-\texttt{LOCAL\_PATH} directory. See Section~\ref{sec:Mesh-graphics}
-for a discussion of mesh viewing features.
-\item [{\texttt{CREATE\_DX\_FILES}}] Set this flag to \texttt{.true.} to
-save OpenDX \urlwithparentheses{www.opendx.org} mesh files for subsequent
-viewing.
-\item [{\texttt{LOCAL\_PATH}}] Directory in which the partitions generated
-by the mesher will be written. Generally one uses a directory on the
-local disk of the compute nodes, although on some machines these partitions
-are written on a parallel (global) file system (see also the earlier
-discussion of the \texttt{LOCAL\_PATH\_IS\_ALSO\_GLOBAL} flag in Chapter~\ref{cha:Getting-Started}).
-The mesher generates the necessary partitions in parallel, one set
-for each of the $\nprocxi\times\nproceta$ slices that constitutes
-the mesh (see Figure~\ref{fig:For-parallel-computing}). After the
-mesher finishes, you can log in to one of the compute nodes and view
-the contents of the \texttt{LOCAL\_PATH} directory to see the files
-generated by the mesher. These files will be needed for creating the
-distributed databases, and have to reside in the directory \texttt{LOCAL\_PATH}
-specified in the main \texttt{Par\_file}, e.g. in directory \texttt{OUTPUT\_FILES/DATABASES\_MPI}.
-Please see Chapter~\ref{cha:Creating-Distributed-Databases} for
-further details.
-\item [{\texttt{NMATERIALS}}] The number of different materials in your
-model. In the following lines, each material needs to be defined as
-:
-
-\begin{lyxcode}
-material\_ID~rho~vp~vs~Q~anisotropy\_flag~domain\_ID
-\end{lyxcode}
-
-where
-\begin{itemize}
-\item \texttt{Q} : quality factor (0=no attenuation)
-\item \texttt{anisotropy\_flag} : 0=no anisotropy / 1,2,.. check with implementation
-in \texttt{aniso\_model.f90}
-\item \texttt{domain\_id} : 1=acoustic / 2=elastic
-\end{itemize}
-\item [{\texttt{NREGIONS}}] The number of regions in the mesh. In the following
-lines, because the mesh is regular or 'almost regular', each region
-is defined as :
-
-\begin{lyxcode}
-NEX\_XI\_BEGIN~NEX\_XI\_END~NEX\_ETA\_BEGIN~NEX\_ETA\_END~NZ\_BEGIN~NZ\_END~material\_ID
-\end{lyxcode}
-\end{description}
-The \texttt{INTERFACES\_FILE} parameter of \texttt{Mesh\_Par\_File}
-defines the file which contains the settings of the topography grid
-and of the interfaces grids. Topography is defined as a set of elevation
-values on a regular 2D grid. It is also possible to define interfaces
-between the layers of the model in the same way. The file needs to
-define several parameters:
-\begin{itemize}
-\item The number of interfaces, including the topography. This needs to
-be set at the first line. Then, from the bottom to the top of the
-model, you need to define the grids with:
-\item \texttt{SUPPRESS\_UTM\_PROJECTION} flag as described previously,
-\item number of points along $x$ and $y$ direction (NXI and NETA),
-\item minimal $x$ and $y$ coordinates (LONG\_MIN and LAT\_MIN),
-\item spacing between points along $x$ and $y$ (SPACING\_XI and SPACING\_ETA)
-and
-\item the name of the file which contains the elevation values (in $y$.$x$
-increasing order).
-\end{itemize}
-At the end of this file, you simply need to set the number of spectral
-elements in the vertical direction for each layer. We provide a few
-models in the {\texttt{EXAMPLES/}} directory. \\
-
-Finally, depending on your system, you might need to provide a file
-that tells MPI what compute nodes to use for the simulations. The
-file must have a number of entries (one entry per line) at least equal
-to the number of processors needed for the run. A sample file is provided
-in the file \texttt{mymachines}. This file is not used by the mesher
-or solver, but is required by the \texttt{go\_mesher} and \texttt{go\_solver}
-default job submission scripts. See Chapter~\ref{cha:Scheduler}
-for information about running the code on a system with a scheduler,
-e.g., LSF.
-
-Now that you have set the appropriate parameters in the \texttt{Mesh\_Par\_file}
-and have compiled the mesher, you are ready to launch it! This is
-most easily accomplished based upon the \texttt{go\_mesher} script.
-When you run on a PC cluster, the script assumes that the nodes are
-named n001, n002, etc. If this is not the case, change the \texttt{tr
--d `n'} line in the script. You may also need to edit the last command
-at the end of the script that invokes the \texttt{mpirun} command.
-See Chapter~\ref{cha:Scheduler} for information about running the
-code on a system with a scheduler, e.g., LSF.
-
-Mesher output is provided in the \texttt{OUTPUT\_FILES} directory
-in \texttt{output\_mesher.txt}; this file provides lots of details
-about the mesh that was generated. Please note that the mesher suggests
-a time step \texttt{DT} to run the solver with. The mesher output
-file also contains a table about the quality of the mesh to indicate
-possible problems with the distortions of elements. Alternatively,
-output can be directed to the screen instead by uncommenting a line
-in \texttt{constants.h}:
-\begin{lyxcode}
-!~uncomment~this~to~write~messages~to~the~screen~
-
-!~integer,~parameter~::~IMAIN~=~ISTANDARD\_OUTPUT
-\end{lyxcode}
-To control the quality of the mesh, check the standard output (either
-on the screen or in the \texttt{OUTPUT\_FILES} directory in \texttt{output\_mesher.txt})
-and analyze the skewness histogram of mesh elements to make sure no
-element has a skewness above approximately 0.75, otherwise the mesh
-is of poor quality (and if even a single element has a skewness value
-above 0.80, then you must definitely improve the mesh). To draw the
-skewness histogram on the screen, type \texttt{gnuplot plot\_mesh\_quality\_histogram.gnu}.
-
-\chapter{\label{cha:Creating-Distributed-Databases}Creating the Distributed
-Databases}
-
-After using \texttt{xmeshfem3D} or \texttt{xdecompose\_mesh}, the
-next step in the workflow is to compile \texttt{xgenerate\_}~\\
- \texttt{databases}. This program is going to create all the missing
-information needed by the SEM solver.
-
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=0.6\textwidth]{figures/workflow}
-\par\end{centering}
-
-\caption{Schematic workflow for a SPECFEM3D Cartesian simulation. The executable
-\texttt{xgenerate\_databases} creates the GLL mesh points and assigns
-specific model parameters.}
-
-\label{fig:workflow.databases}
-\end{figure}
-
-\noindent
-In the main directory, type
-\begin{lyxcode}
-{\small make~xgenerate\_databases}{\small \par}
-\end{lyxcode}
-Input for the program is provided through the main parameter file
-\texttt{Par\_file}, which resides in the subdirectory \texttt{DATA}.
-Please note that \texttt{xgenerate\_databases} must be called directly
-from the main directory, as most of the binaries of the package.
-
-\section{\label{cha:Main-Parameter}Main parameter file \texttt{Par\_file}}
-
-Before running \texttt{xgenerate\_databases}, a number of parameters
-need to be set in the main parameter \texttt{Par\_file} located in
-the subdirectory \texttt{DATA}:
-\begin{description}
-\item [{\texttt{SIMULATION\_TYPE}}] is set to 1 for forward simulations,
-2 for adjoint simulations (see Section \ref{sec:Adjoint-simulation-finite})
-and 3 for kernel simulations (see Section \ref{sec:Finite-Frequency-Kernels}).
-\item [{\texttt{SAVE\_FORWARD}}] is only set to \texttt{.true.} for a forward
-simulation with the last frame of the simulation saved, as part of
-the finite-frequency kernel calculations (see Section \ref{sec:Finite-Frequency-Kernels}).
-For a regular forward simulation, leave \texttt{SIMULATION\_TYPE}
-and \texttt{SAVE\_FORWARD} at their default values.
-\item [{\texttt{\noun{UTM\_PROJECTION\_ZONE}}}] UTM projection zone in
-which your model resides, only valid when \texttt{SUPPRESS\_UTM\_}~\\
- \texttt{PROJECTION} is \texttt{.false.}.
-\item [{\texttt{SUPPRESS\_UTM\_PROJECTION}}] set to be \texttt{.false.}
-when your model range is specified in the geographical coordinates,
-and needs to be \texttt{.true.} when your model is specified in a
-cartesian coordinates. \noun{UTM projection zone in which your simulation
-region resides.}
-\item [{\texttt{NPROC}}] The number of MPI processors, each one is assigned
-one slice of the whole mesh.
-\item [{\texttt{NSTEP}}] The number of time steps of the simulation. This
-controls the length of the numerical simulation, i.e., twice the number
-of time steps requires twice as much CPU time. This feature is not
-used at the time of generating the distributed databases but is required
-for the solver, i.e., you may change this parameter after running
-\texttt{xgenerate\_databases}.
-\item [{\texttt{DT}}] The length of each time step in seconds. This feature
-is not used at the time of generating the distributed databases but
-is required for the solver. Please see also Section~\ref{sec:Choosing-the-Time-Step}
-for further details.
-\item [{\texttt{NGNOD}}] The number of nodes for 2D and 3D shape functions
-for hexahedra. We use either 8-node mesh elements (bricks) or 27-node
-elements. If you use the internal mesher, the only option is 8-node
-bricks (27-node elements are not supported). \texttt{CUBIT} does not
-support HEX27 elements either (it can generate them, but they are
-flat, i.e. identical to HEX8). To generate HEX27 elements with curvature
-properly taken into account, you can use \texttt{Gmsh} \url{http://geuz.org/gmsh/}
-\item [{\texttt{MODEL}}] Must be set to one of the following:
-
-\begin{description}
-\item [{\textmd{Models defined by mesh parameters:}}] ~
-
-\begin{description}
-\item [{\texttt{default}}] Uses model parameters as defined by meshing
-procedures described in the previous Chapter~\ref{cha:Mesh-Generation}.
-\end{description}
-
-\item [{\textmd{1D~models~with~real~structure:}}] ~
-
-\begin{description}
-\item [{\texttt{1D\_prem}}] Isotropic version of the spherically symmetric
-Preliminary Reference Earth Model (PREM) \citep{DzAn81}.
-\item [{\texttt{1D\_socal}}] A standard isotropic 1D model for Southern
-California.
-\item [{\texttt{1D\_cascadia}}] Isotropic 1D profile for the Cascadia region.
-\end{description}
-
-\item [{\textmd{Fully~3D~models:}}] ~
-
-\begin{description}
-\item [{\texttt{aniso}}] For a user-specified fully anisotropic model.
-Parameters are set up in routines located in file \texttt{model\_aniso.f90}
-in directory \texttt{src/generate\_databases/}. See Chapter~\ref{cha:-Changing-the}
-for a discussion on how to specify your own 3D model.
-
-\item [{\texttt{external}}] For a user-specified isotropic model which
-uses externally defined model parameters. Uses external model definitions
-set up in routines located in file \texttt{model\_external\_values.f90}
-in directory \texttt{src/generate\_databases/}. Please modify these
-generic template routines to use your own model definitions.
-\item [{\texttt{gll}}] For a user-specified isotropic model which uses
-external binary files for $v_{p}$, $v_{s}$ and $\rho$. Binary files
-are given in the same format as when outputted by the \texttt{xgenerate\_databases}
-executable when using option \texttt{SAVE\_MESH\_FILES}. These binary
-files define the model parameters on all GLL points which can be used
-for iterative inversion procedures.
-\item [{\texttt{salton\_trough}}] A 3D $V_{p}$ model for Southern California.
-Users must provide the corresponding data file \texttt{regrid3\_vel\_p.bin}
-in directory \texttt{DATA/st\_3D\_block\_harvard/}.
-\item [{\texttt{tomo}}] For a user-specified 3D isotropic model which uses
-a tomographic model file \texttt{tomographic\_model.xyz} in directory
-\texttt{DATA}. See Section \ref{sec:Using-tomographic}, for a discussion
-on how to specify your own 3D tomographic model.
-\end{description}
-\end{description}
-\item [{\texttt{APPROXIMATE\_OCEAN\_LOAD}}] Set to \texttt{.true.} if the
-effect of the oceans on seismic wave propagation should be incorporated
-based upon the (rough) approximate treatment discussed in \citet{KoTr02b}.
-This feature is inexpensive from a numerical perspective, both in
-terms of memory requirements and CPU time. This approximation is accurate
-at periods of roughly 20~s and longer. At shorter periods the effect
-of water phases/reverberations is not taken into account, even when
-the flag is on. If you want to model the effect of a fluid-solid model
-at short periods, then set this flag to \texttt{.false.} and mesh
-the fluid layer explicitly in your mesher, so that it is computed
-accurately and without this approximation.
-\item [{\texttt{TOPOGRAPHY}}] This feature is only effective if \texttt{APPROXIMATE\_OCEAN\_LOAD}
-is set to \texttt{.true.}. Set to \texttt{.true.} if topography and
-bathymetry should be read in based upon the topography file specified
-in the main constants file \texttt{constants.h} found in subdirectory
-\texttt{src/shared/} to evaluate elevations. If not set, elevations
-will be read from the numerical mesh.
-\item [{\texttt{ATTENUATION}}] Set to \texttt{.true.} if attenuation should
-be incorporated. Turning this feature on increases the memory requirements
-significantly (roughly by a factor of~1.5), and is numerically fairly
-expensive. See \citet{KoTr99,KoTr02a} for a discussion on the implementation
-of attenuation based upon standard linear solids. Please note that
-the Vp- and Vs-velocities of your model are given for a reference
-frequency. To change this reference frequency, you change the value
-of \texttt{ATTENUATION\_f0\_REFERENCE} in the main constants file
-\texttt{constants.h} found in subdirectory \texttt{src/shared/}.
-\item [{\texttt{FULL\_ATTENUATION\_SOLID}}] Set to \texttt{.true.} to add
-$Q_{\kappa}$ attenuation to the simulations in addition to $Q_{\mu}$
-attenuation when the \texttt{ATTENUATION} flag above is on. Should
-be off in almost all cases for seismic wave propagation, since $Q_{\kappa}$
-is almost always negligible in Earth models. However, this parameter
-can be useful for instance for ocean acoustics simulations, when $Q_{\kappa}$
-is not negligible in the solid layer at the sea bottom.
-\item [{\texttt{ANISOTROPY}}] Set to \texttt{.true.} if you want to use
-an anisotropy model. Please see the file \texttt{model\_aniso.f90}
-in subdirectory \texttt{src/generate\_databases/} for the current
-implementation of anisotropic models.
-\item [{\texttt{TOMOGRAPHY\_PATH}}] Directory in which the tomography files
-are stored for using external tomographic Earth models (please read
-Chapter \ref{cha:-Changing-the} and Section \ref{sec:Using-tomographic}
-`Using external tomographic Earth models' for further details.).
-\item [{\texttt{USE\_OLSEN\_ATTENUATION}}] Set to \texttt{.true.} if you
-want to use the attenuation model that scaled from the S-wave speed
-model using Olsen's empirical relation (see \citet{OlDaBr03}).
-\item [{\texttt{OLSEN\_ATTENUATION\_RATIO}}] Determines the Olsen's constant
-in Olsen's empirical relation (see \citet{OlDaBr03}).
-\item [{\texttt{PML\_CONDITIONS}}] Set to \texttt{.true.} to turn on C-PML
-boundary conditions for a regional simulation. Both fluids and elastic
-solids are supported. Note that xmeshfem3d generated meshes do not support C-PML yet.
-\item [{\texttt{PML\_INSTEAD\_OF\_FREE\_SURFACE}}] Set to \texttt{.true.}
-to turn on C-PML boundary conditions on the top surface instead of
-the usual free surface.
-\item [{\texttt{f0\_FOR\_PML}}] Determines the dominant frequency that
-will be used in the calculation of PML damping profiles; this should
-be set to the same (or similar) dominant frequency as that of the
-source that you will use in your simulation. If you plan to use a
-Dirac source, then use the dominant frequency of the source wavelet
-with which you plan to convolve your seismograms later on in post-processing.
-\item [{\texttt{STACEY\_ABSORBING\_CONDITIONS}}] Set to \texttt{.true.}
-to turn on Clayton-Enquist absorbing boundary conditions (see \citet{KoTr99}).
-In almost all cases it is much better to use CPML absorbing layers
-(see the options above) and leave this flag to \texttt{.false.}.
-\item [{\texttt{STACEY\_INSTEAD\_OF\_FREE\_SURFACE}}] Set to \texttt{.true.}
-to turn on absorbing boundary conditions on the top surface which
-by default constitutes a free surface of the model.
-\item [{\texttt{CREATE\_SHAKEMAP}}] Set this flag to \texttt{.true.} to
-create a ShakeMap\textregistered{}, i.e., a peak ground velocity map
-of the maximum absolute value of the two horizontal components of
-the velocity vector.
-\item [{\texttt{MOVIE\_SURFACE}}] Set to \texttt{.false.}, unless you want
-to create a movie of seismic wave propagation on the Earth's surface.
-Turning this option on generates large output files. See Section~\ref{sec:Movies}
-for a discussion on the generation of movies. This feature is only
-relevant for the solver.
-\item [{\texttt{MOVIE\_TYPE}}] Set this flag to 1 to show the top surface
-(tomography + oceans) only, to 2 to show all external faces of the
-mesh (i.e. topography + vertical edges + bottom) in shakemaps and
-surface movies.
-\item [{\texttt{MOVIE\_VOLUME}}] Set to \texttt{.false.}, unless you want
-to create a movie of seismic wave propagation in the Earth's interior.
-Turning this option on generates huge output files. See Section~\ref{sec:Movies}
-for a discussion on the generation of movies. This feature is only
-relevant for the solver.
-\item [{\texttt{SAVE\_DISPLACEMENT}}] Set this flag to \texttt{.true.}
-if you want to save the displacement instead of velocity for the movie
-frames.
-\item [{\texttt{USE\_HIGHRES\_FOR\_MOVIES}}] Set this flag to \texttt{.true.}
-if you want to save the values at all the NGLL grid points for the
-movie frames.
-\item [{\texttt{NTSTEP\_BETWEEN\_FRAMES}}] Determines the number of timesteps
-between movie frames. Typically you want to save a snapshot every
-100 timesteps. The smaller you make this number the more output will
-be generated! See Section~\ref{sec:Movies} for a discussion on the
-generation of movies. This feature is only relevant for the solver.
-\item [{\texttt{HDUR\_MOVIE}}] Determines the half duration of the source
-time function for the movie simulations. When this parameter is set
-to be 0, a default half duration that corresponds to the accuracy
-of the simulation is provided. Otherwise, it adds this half duration
-to the half duration specified in the source file \texttt{CMTSOLUTION},
-thus simulates longer periods to make the movie images look smoother.
-\item [{\texttt{SAVE\_MESH\_FILES}}] Set this flag to \texttt{.true.} to
-save ParaView \urlwithparentheses{www.paraview.org} mesh files for
-subsequent viewing. Turning the flag on generates large (distributed)
-files in the \texttt{LOCAL\_PATH} directory. See Section~\ref{sec:Mesh-graphics}
-for a discussion of mesh viewing features.
-\item [{\texttt{LOCAL\_PATH}}] Directory in which the distributed databases
-will be written. Generally one uses a directory on the local disk
-of the compute nodes, although on some machines these databases are
-written on a parallel (global) file system (see also the earlier discussion
-of the \texttt{LOCAL\_PATH\_IS\_ALSO\_GLOBAL} flag in Chapter~\ref{cha:Getting-Started}).
-\texttt{xgenerate\_databases} generates the necessary databases in
-parallel, one set for each of the \texttt{NPROC} slices that constitutes
-the mesh (see Figure~\ref{fig:mount.partitions} and Figure~\ref{fig:For-parallel-computing}).
-After the executable finishes, you can log in to one of the compute
-nodes and view the contents of the \texttt{LOCAL\_PATH} directory
-to see the (many) files generated by \texttt{xgenerate\_databases}.
-Please note that the \texttt{LOCAL\_PATH} directory should already
-contain the output files of the partitioner, i.e. from \texttt{xdecompose\_mesh}
-or \texttt{xmeshfem3D}.
-\item [{\texttt{NTSTEP\_BETWEEN\_OUTPUT\_INFO}}] This parameter specifies
-the interval at which basic information about a run is written to
-the file system (\texttt{timestamp{*}} files in the \texttt{OUTPUT\_FILES}
-directory). If you have access to a fast machine, set \texttt{NTSTEP\_BETWEEN\_OUTPUT\_INFO}
-to a relatively high value (e.g., at least 100, or even 1000 or more)
-to avoid writing output text files too often. This feature is not
-used at the time of meshing. One can set this parameter to a larger
-value than the number of time steps to avoid writing output during
-the run.
-\item [{\texttt{NTSTEP\_BETWEEN\_OUTPUT\_SEISMOS}}] This parameter specifies
-the interval at which synthetic seismograms are written in the \texttt{LOCAL\_PATH}
-directory. If a run crashes, you may still find usable (but shorter
-than requested) seismograms in this directory. On a fast machine set
-\texttt{NTSTEP\_BETWEEN\_OUTPUT\_SEISMOS} to a relatively high value
-to avoid writing to the seismograms too often. This feature is only
-relevant for the solver.
-\item [{\texttt{USE\_FORCE\_POINT\_SOURCE}}] Turn this flag on to use a
-(tilted) \texttt{FORCESOLUTION} force point source instead of a \texttt{CMTSOLUTION}
-moment-tensor source. When the force source does not fall exactly
-at a grid point, the solver interpolates the force between grid points
-using Lagrange interpolants. This can be useful e.g. for oil industry
-foothills simulations in which the source is a vertical force, normal
-force, tilted force, or an impact etc. Note that in the \texttt{FORCESOLUTION}
-file, you will need to edit the East, North and vertical components
-of an arbitrary (not necessarily unitary, the code will normalize it automatically) direction vector of the force vector;
-thus refer to Appendix~\ref{cha:Coordinates} for the orientation
-of the reference frame. This vector is made unitary internally in
-the solver and thus only its direction matters here; its norm is ignored
-and the norm of the force used is the factor force source times the
-source time function.
-When using this option, by default the code can locate the force source
-anywhere between mesh points in order to honor its exact location; this
-is more precise than using the closest GLL mesh point, but it is also a bit slower.
-If needed, you can change that default behavior and force the code to use the closest
-GLL mesh point instead by setting flag \texttt{USE\_BEST\_LOCATION} to \texttt{.false.} instead of \texttt{.true.} in file
-\texttt{src/shared/constants.h.in} and running the \texttt{configure} script again and recompiling the code.
-\item [{\texttt{USE\_RICKER\_TIME\_FUNCTION}}] Turn this flag on to use
-a Ricker source time function instead of the source time functions
-set by default to represent a (tilted) \texttt{FORCESOLUTION} force
-point source or a \texttt{CMTSOLUTION} moment-tensor source. Originally,
-if a \texttt{CMTSOLUTION} moment-tensor source is used, a (pseudo)
-Heaviside step function with a very short half duration is defined
-for elastic cases to represent the permanent slip on the fault while
-in the acoustic case a Gaussian source time function with a similarly
-short half duration is defined to physically describe actions within
-the fluid. Otherwise, if a \texttt{FORCESOLUTION} force source is
-used, a (pseudo) Dirac delta source time function is defined by default.
-Any other source-time function may then be obtained by convolution.
-\item [{\texttt{PRINT\_SOURCE\_TIME\_FUNCTION}}] Turn this flag on to print
-information about the source time function in the file \texttt{OUTPUT\_FILES/plot\_source\_time\_function.txt}.
-This feature is only relevant for the solver.
-
-\item [{\texttt{NUMBER\_OF\_SIMULTANEOUS\_RUNS}}] adds the ability to run several calculations (several earthquakes)
-in an embarrassingly-parallel fashion from within the same run;
-this can be useful when using a very large supercomputer to compute
-many earthquakes in a catalog, in which case it can be better from
-a batch job submission point of view to start fewer and much larger jobs,
-each of them computing several earthquakes in parallel.
-To turn that option on, set parameter NUMBER\_OF\_SIMULTANEOUS\_RUNS to a value greater than 1.
-To implement that, we create NUMBER\_OF\_SIMULTANEOUS\_RUNS MPI sub-communicators,
-each of them being labeled "my\_local\_mpi\_comm\_world", and we use them
-in all the routines in "src/shared/parallel.f90", except in MPI\_ABORT() because in that case
-we need to kill the entire run.
-When that option is on, of course the number of processor cores used to start
-the code in the batch system must be a multiple of NUMBER\_OF\_SIMULTANEOUS\_RUNS,
-all the individual runs must use the same number of processor cores,
-which as usual is NPROC in the Par\_file,
-and thus the total number of processor cores to request from the batch system
-should be NUMBER\_OF\_SIMULTANEOUS\_RUNS * NPROC.
-All the runs to perform must be placed in directories called run0001, run0002, run0003 and so on
-(with exactly four digits).
-
-Imagine you have 10 independent calculations to do, each of them on 100 cores; you have three options:
-
-1/ submit 10 jobs to the batch system
-
-2/ submit a single job on 1000 cores to the batch, and in that script create a sub-array of jobs to start 10 jobs,
-each running on 100 cores (see e.g. http://www.schedmd.com/slurmdocs/job\_array.html )
-
-3/ submit a single job on 1000 cores to the batch, start SPECFEM3D on 1000 cores, create 10 sub-communicators,
-cd into one of 10 subdirectories (called e.g. run0001, run0002,... run0010) depending on the sub-communicator
-your MPI rank belongs to, and run normally on 100 cores using that sub-communicator.
-
-The option NUMBER\_OF\_SIMULTANEOUS\_RUNS implements 3/.
-
-\item [{\texttt{BROADCAST\_SAME\_MESH\_AND\_MODEL}}]:
-if we perform simultaneous runs in parallel, if only the source and receivers vary between these runs
-but not the mesh nor the model (velocity and density) then we can also read the mesh and model files
-from a single run in the beginning and broadcast them to all the others; for a large number of simultaneous
-runs for instance when solving inverse problems iteratively this can DRASTICALLY reduce I/Os to disk in the solver
-(by a factor equal to NUMBER\_OF\_SIMULTANEOUS\_RUNS), and reducing I/Os is crucial in the case of huge runs.
-Thus, always set this option to .true. if the mesh and the model are the same for all simultaneous runs.
-In that case there is no need to duplicate the mesh and model file database (the content of the DATABASES\_MPI
-directories) in each of the run0001, run0002,... directories, it is sufficient to have one in run0001
-and the code will broadcast it to the others).
-
-TODO / future work to do: currently the BROADCAST\_SAME\_MESH\_AND\_MODEL option assumes to have the (master) mesh files in run0001/DATABASES\_MPI or run0001/OUTPUT\_FILES/DATABASES\_MPI.
-However, for adjoint runs you still need a DATABASES\_MPI folder in each of the sub-runs directories, e.g.
-run0002/DATABASES\_MPI, etc.
-to store the forward wavefields, kernels etc.
-of each sub-run.
-This would not be needed for forward simulations.
-
-TODO / future work to do: the sensitivity kernel summing and smoothing tools in directory src/tomography are currently not ported to this new option to do many runs simultaneously, only the solver (src/specfem3d) is.
-Thus these tools should work, but in their current version will need to be run for each simulation result independently.
-More precisely, the current kernel summing and smoothing routines work fine, with the exception that you need to move out the mesh files (and also the parameters).
-This works because these routines consider multiple runs by design.
-You simply have to provide them the directories where the kernels are.
-
-\item [{\texttt{GPU\_MODE}}] Turn this flag on to use GPUs.
-
-\item [\texttt{ADIOS\_ENABLED}] Turn this flag on to enable ADIOS. If set to \texttt{.false.}, subsequent ADIOS
-parameters will not be considered.
-\item [\texttt{ADIOS\_FOR\_DATABASES}] Turn this flag on to use ADIOS for xmeshfem3D output and
-xgenerate\_database input.
-\item [\texttt{ADIOS\_FOR\_MESH}]  Turn this flag on to use ADIOS for generated databases.
-\item [\texttt{ADIOS\_FOR\_FORWARD\_ARRAYS}] Turn this flag on to read and write forward arrays using ADIOS.
-\item [\texttt{ADIOS\_FOR\_KERNELS}] Turn this flag on to produce ADIOS kernels that can later be visualized with the ADIOS version of combine\_vol\_data.
-
-\end{description}
-
-The present version of SPECFEM can handle fully saturated porous simulations, Christina Morency implemented Biot equation.
-But the code cannot calculate partially saturated cases in its state.
-Christina Morency is presently working on a dual porosity, dual permeability formulation, type Pride and Berryman, for an other project, but it will not be available for some time.\\
-
-The way we prescribe material property for porous material in SPECFEM3D is as follow:
-We use a file name nummaterial\_poroelastic\_file, which is located in the directory MESH, the format is as follow:\\
- rhos rhof phi c kxx kxy kxz kyy kyz kzz Ks Kf Kfr etaf mufr\\
-where\\
-rho\_s = solid density\\
-rho\_f = fluid density\\
-phi = porosity\\
-tort = tortuosity\\
-kxx = xx component of permeability tensor\\
-kxy = xy,yx components of permeability tensor\\
-kyy = yy component of permeability tensor\\
-kxz = xz,zx components of permeability tensor\\
-kzz = zz component of permeability tensor\\
-kappa\_s = solid bulk modulus\\
-kappa\_f= fluid bulk modulus\\
-kappa\_fr= frame bulk modulus\\
-eta\_f = fluid viscosity\\
-mu\_fr = frame shear modulus.\\
-
-Using an external mesh (for instance coming from CUBIT/TRELIS), poroelastic materials have the ID number 3, while 1 is elastic, 2 is acoustic ?
-see the example in the package: examples/homogeneous\_poroelastic.
-
-If you use PML, the mesh elements that belong to the PML layers can
-be acoustic or elastic, but not viscoelastic nor poroelastic. Then,
-when defining your model, you should define these absorbing elements
-as either acoustic or elastic. In you forget to do that, the code
-will fix the problem by automatically converting the viscoelastic
-or poroelastic PML elements to elastic. This means that strictly speaking
-the PML layer will not be perfectly matched any more, since the physical
-model will change from viscoelastic or poroelastic to elastic at the
-entrance of the PML, but in practice this is sufficient and produces
-only tiny / negligible spurious reflections.
-
-If you use PML and an external mesh (created using an external meshing tool
-such as CUBIT/TRELIS or similar), try to have elements inside the PML as regular as possible,
-i.e. ideally non-deformed cubes obtained by `extrusion' of regular surface mesh elements meshing the
-outer edges of the computational domain without PML; by doing so, the PMLs obtained will be far more stable
-in time (PML being weakly unstable from a mathematical point of view, very deformed mesh elements
-inside the PMLs can trigger instabilities much more quickly).
-
-If you use PML and an external tomographic velocity and density model,
-you should be careful because mathematically a PML cannot handle heterogeneities
-along the normal to the PML edge inside the PML layer. This comes
-from the fact that the damping profile that is defined assumes a constant
-velocity and density model along the normal direction.
-
-Thus, you need to modify your velocity and density model in order
-for it to be 1D inside the PML, as shown in Figure~\ref{fig:modify_external_velocity_model_to_use_PML}.
-
-This applies to the bottom layer as well; there you should make sure
-that your model is 1D and thus constant along the vertical direction.
-
-To summarize, only use a 3D velocity and density model inside the
-physical region, and in all the PML layers extend it by continuity
-from its values along the inner PML edge.
-
-%%
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=3in]{figures/how_to_use_PML_when_a_tomographic_velocity_model_is_used}
-\par\end{centering}
-
-\caption{How to modify your external 3D velocity and density model in order
-to use PML. Such a modification is not needed when using Stacey absorbing
-boundary conditions (but such conditions are significantly less efficient).}
-
-\label{fig:modify_external_velocity_model_to_use_PML}
-\end{figure}
-
-%%
-
-\section{\label{sec:Choosing-the-Time-Step}Choosing the time step \texttt{DT}}
-
-The parameter \texttt{DT} sets the length of each time step in seconds.
-The value of this parameter is crucial for the stability of the spectral-element
-simulation. Your time step \texttt{DT} will depend on the minimum
-ratio between the distance $h$ of neighboring mesh points and the
-wave speeds $v$ defined in your model. The condition for the time
-step $\Delta t$ is:
-\begin{lyxcode}
-$\Delta t<C~\mathrm{min}_{\Omega}(~h/v~)$
-\end{lyxcode}
-where $C$ is the so-called Courant number and $\Omega$ denotes the
-model volume. The distance $h$ depends on the mesh element size and
-the number of GLL points \texttt{NGLL} specified in the main constants
-file \texttt{constants.h} located in the \texttt{src/shared/} subdirectory.
-The wave speed $v$ is determined based on your model's P- (or S-)
-wave speed values.
-
-The database generator \texttt{xgenerate\_databases}, as well as the
-internal mesher \texttt{xmeshfem3D}, are trying to evaluate the value
-of $\Delta t$ for empirically chosen Courant numbers $C\sim0.3$.
-If you used the mesher \texttt{xmeshfem3D} to generate your mesh,
-you should set the value suggested in \texttt{OUTPUT\_FILES/output\_mesher.txt}
-file, which is created after the mesher completed. In case you used
-CUBIT to create the mesh, you might use an arbitrary value when running
-\texttt{xgenerate\_databases} and then use the value suggested in
-the ~\\
- \texttt{OUTPUT\_FILES/output\_mesher.txt} file after the database
-generation completed. Note that the implemented Newmark time scheme
-uses this time step globally, thus your simulations become more expensive
-for very small mesh elements in high wave-speed regions. Please be
-aware of this restriction when constructing your mesh in Chapter~\ref{cha:Mesh-Generation}.
-
-\chapter{\label{cha:Running-the-Solver}Running the Solver \texttt{xspecfem3D}}
-
-Now that you have successfully generated the databases, you are ready
-to compile the solver. In the main directory, type
-\begin{lyxcode}
-{\small make~xspecfem3D}{\small \par}
-\end{lyxcode}
-Please note that \texttt{xspecfem3D} must be called directly from
-the main directory, as most of the binaries of the package.\\
-
-\noindent
-The solver needs three input files in the \texttt{DATA} directory
-to run:
-\begin{description}
-\item [{\texttt{Par\_file}}] the main parameter file which was discussed
-in detail in the previous Chapter~\ref{cha:Creating-Distributed-Databases},
-\item [{\texttt{CMTSOLUTION}} or {\texttt{FORCESOLUTION}}] the earthquake
-source parameter file or the force source parameter file, and
-\item [{\texttt{STATIONS}}] the stations file.
-\end{description}
-Most parameters in the \texttt{Par\_file} should be set prior to running
-the databases generation. Only the following parameters may be changed
-after running \texttt{xgenerate\_databases}:
-\begin{itemize}
-\item the simulation type control parameters: \texttt{SIMULATION\_TYPE}
-and \texttt{SAVE\_FORWARD}
-\item the time step parameters \texttt{NSTEP} and \texttt{DT}
-\item the absorbing boundary control parameter \texttt{PML\_CONDITIONS}
-on condition that the\\
- \texttt{PML\_INSTEAD\_OF\_FREE\_SURFACE} flag remains unmodified
-after running the databases generation.
-\item the movie control parameters \texttt{MOVIE\_SURFACE}, \texttt{MOVIE\_VOLUME},
-and \texttt{NTSTEPS\_BETWEEN\_FRAMES}
-\item the ShakeMap\textregistered{}option \texttt{CREATE\_SHAKEMAP}
-\item the output information parameters \texttt{MOVIE\_TYPE}, \texttt{NTSTEP\_BETWEEN\_OUTPUT\_INFO}
-and \texttt{NTSTEP\_BETWEEN\_OUTPUT\_}~\\
- \texttt{SEISMOS}
-\item the \texttt{PRINT\_SOURCE\_TIME\_FUNCTION} flags
-\end{itemize}
-Any other change to the \texttt{Par\_file} implies rerunning both
-the database generator \texttt{xgenerate\_databases} and the solver
-\texttt{xspecfem3D}.
-
-For any particular earthquake, the \texttt{CMTSOLUTION} file that
-represents the point source may be obtained directly from the Harvard
-Centroid-Moment Tensor (CMT) web page \urlwithparentheses{www.seismology.harvard.edu}.
-It looks like the example shown in Fig. \ref{fig:CMTSOLUTION-file}.
-%\begin{lyxcode}
-\begin{figure}[htp]
-\noindent
-\begin{centering}
-{\small \includegraphics[width=0.8\textwidth]{figures/Hollywood_CMT}
-}
-\par
-\end{centering}
-\caption{\texttt{CMTSOLUTION} file based on the
-format from the Harvard CMT catalog. \textbf{M} is the moment tensor,
-$M_{0}${\small {} }is the seismic moment, and $M_{w}$ is the moment
-magnitude.}
-\label{fig:CMTSOLUTION-file}
-\end{figure}
-%\end{lyxcode}
-The \texttt{CMTSOLUTION} file should be edited in the following way:
-\begin{itemize}
-\item Set the latitude or UTM $x$ coordinate, longitude or UTM $y$ coordinate, depth of the source (in km).
-\item Set the \texttt{time shift} parameter equal to $0.0$ (the solver
-will not run otherwise.) The time shift parameter would simply apply
-an overall time shift to the synthetics, something that can be done
-in the post-processing (see Section \ref{sec:Process-data-and-syn}).
-\item For point-source simulations (see finite sources, page \pageref{To-simulate-a})
-we recommend setting the source half-duration parameter \texttt{half
-duration} equal to zero, which corresponds to simulating a step source-time
-function, i.e., a moment-rate function that is a delta function. If
-\texttt{half duration} is not set to zero, the code will use a Gaussian
-(i.e., a signal with a shape similar to a `smoothed triangle', as
-explained in \citet{KoTr02a} and shown in Fig~\ref{fig:gauss.vs.triangle})
-source-time function with half-width \texttt{half duration}. We prefer
-to run the solver with \texttt{half duration} set to zero and convolve
-the resulting synthetic seismograms in post-processing after the run,
-because this way it is easy to use a variety of source-time functions
-(see Section \ref{sec:Process-data-and-syn}). \citet{KoTr02a} determined
-that the noise generated in the simulation by using a step source
-time function may be safely filtered out afterward based upon a convolution
-with the desired source time function and/or low-pass filtering. Use
-the serial code \texttt{convolve\_source\_timefunction.f90} and the
-script \texttt{convolve\_source\_timefunction.csh} for this purpose,
-or alternatively use signal-processing software packages such as SAC
-\urlwithparentheses{www.llnl.gov/sac}. Type
-
-\begin{lyxcode}
-{\small make~xconvolve\_source\_timefunction}{\small \par}
-\end{lyxcode}
-
-to compile the code and then set the parameter \texttt{hdur} in \texttt{convolve\_source\_timefunction.csh}
-to the desired half-duration.
-
-\item The zero time of the simulation corresponds to the center of the triangle/Gaussian,
-or the centroid time of the earthquake. The start time of the simulation
-is $t=-1.5*\texttt{half duration}$ (the 1.5 is to make sure the moment
-rate function is very close to zero when starting the simulation).
-To convert to absolute time $t_{\mathrm{abs}}$, set
-
-\begin{lyxcode}
-$t_{\mathrm{abs}}=t_{\mathrm{pde}}+\texttt{time shift}+t_{\mathrm{synthetic}}$
-\end{lyxcode}
-
-where $t_{\mathrm{pde}}$ is the time given in the first line of the
-\texttt{CMTSOLUTION}, \texttt{time shift} is the corresponding value
-from the original \texttt{CMTSOLUTION} file and $t_{\mathrm{synthetic}}$
-is the time in the first column of the output seismogram.
-
-\end{itemize}
-\begin{figure}[htp]
-\noindent \begin{centering}
-\includegraphics[width=3.5in]{figures/gauss_vs_triangle_mod}
-\par\end{centering}
-
-\caption{Comparison of the shape of a triangle and the Gaussian function actually
-used.}
-
-\label{fig:gauss.vs.triangle}
-\end{figure}
-
-If you know the earthquake source in strike/dip/rake format rather
-than in \texttt{CMTSOLUTION} format, use the C code \texttt{SPECFEM3D\_GLOBE/utils/strike\_dip\_rake\_to\_CMTSOLUTION.c}
-to convert it. The conversion formulas are given for instance in \citet{AkRi80}.
-Note that the \citet{AkRi80} convention is slightly different from
-the Harvard \texttt{CMTSOLUTION} convention (the sign of some components
-is different). The C code outputs both.
-
-Centroid latitude and longitude should be provided in geographical
-coordinates. The code converts these coordinates to geocentric coordinates~\citep{DaTr98}.
-Of course you may provide your own source representations by designing
-your own \texttt{CMTSOLUTION} file. Just make sure that the resulting
-file adheres to the Harvard CMT conventions (see Appendix~\ref{cha:Coordinates}).
-Note that the first line in the \texttt{CMTSOLUTION} file is the Preliminary
-Determination of Earthquakes (PDE) solution performed by the USGS
-NEIC, which is used as a seed for the Harvard CMT inversion. The PDE
-solution is based upon P waves and often gives the hypocenter of the
-earthquake, i.e., the rupture initiation point, whereas the CMT solution
-gives the `centroid location', which is the location with dominant
-moment release. The PDE solution is not used by our software package
-but must be present anyway in the first line of the file.
-
-\label{To-simulate-a}To simulate a kinematic rupture, i.e., a finite-source
-event, represented in terms of $N_{\mathrm{sources}}$ point sources,
-provide a \texttt{CMTSOLUTION} file that has $N_{\mathrm{sources}}$
-entries, one for each subevent (i.e., concatenate $N_{\mathrm{sources}}$
-\texttt{CMTSOLUTION} files to a single \texttt{CMTSOLUTION} file).
-At least one entry (not necessarily the first) must have a zero \texttt{time
-shift}, and all the other entries must have non-negative \texttt{time
-shift}. Each subevent can have its own half duration, latitude, longitude,
-depth, and moment tensor (effectively, the local moment-density tensor).
-
-Note that the zero in the synthetics does NOT represent the hypocentral
-time or centroid time in general, but the timing of the \textit{center}
-of the source triangle with zero \texttt{time shift} (Fig~\ref{fig:source_timing}).
-
-Although it is convenient to think of each source as a triangle, in
-the simulation they are actually Gaussians (as they have better frequency
-characteristics). The relationship between the triangle and the gaussian
-used is shown in Fig~\ref{fig:gauss.vs.triangle}. For finite fault
-simulations it is usually not advisable to use a zero half duration
-and convolve afterwards, since the half duration is generally fixed
-by the finite fault model.
-
-\vspace{1cm}
-
-\noindent The \texttt{FORCESOLUTION} file should be edited in the
-following way:
-\begin{itemize}
-\item Set the \texttt{time shift} parameter equal to $0.0$ (the solver
-will not run otherwise.) The time shift parameter would simply apply
-an overall time shift to the synthetics, something that can be done
-in the post-processing (see Section \ref{sec:Process-data-and-syn}).
-\item Set the \texttt{f0} parameter (the dominant frequency) of the Ricker source time
-function when {\texttt{USE\_RICKER\_TIME\_FUNCTION}} is turned on
-in the main parameter file {\texttt{Par\_file}}. In case that the
-solver uses a (pseudo) Dirac delta source time function to represent
-a force point source, a very short duration of five time steps is automatically set by default.
-\item Set the latitude or UTM $x$ coordinate, longitude or UTM $y$ coordinate, depth of the source (in km).
-\item Set the magnitude of the force source.
-\item Set the components of a (non-unitary) direction vector for the force
-source in the East/North/Vertical basis (see Appendix A for the orientation
-of the reference frame).
-\end{itemize}
-\noindent Where necessary, set a \texttt{FORCESOLUTION} file in the
-same way you configure a \texttt{CMTSOLUTION} file with $N_{\mathrm{sources}}$
-entries, one for each subevent (i.e., concatenate $N_{\mathrm{sources}}$
-\texttt{FORCESOLUTION} files to a single \texttt{FORCESOLUTION} file).
-At least one entry (not necessarily the first) must have a zero \texttt{time
-shift}, and all the other entries must have non-negative \texttt{time
-shift}. Each subevent can have its own half latitude, longitude, depth,
-\texttt{half duration} and force parameters.
-
-\begin{figure}[H]
-\noindent \begin{centering}
-{\small \includegraphics[width=5in]{figures/source_timing} }
-\par\end{centering}
-
-\caption{Example of timing for three sources. The center of the first source
-triangle is defined to be time zero. Note that this is NOT in general
-the hypocentral time, or the start time of the source (marked as tstart).
-The parameter \texttt{time shift} in the \texttt{CMTSOLUTION} file
-would be t1(=0), t2, t3 in this case, and the parameter \texttt{half
-duration} would be hdur1, hdur2, hdur3 for the sources 1, 2, 3 respectively.}
-
-{\small \label{fig:source_timing} }
-\end{figure}
-
-The solver can calculate seismograms at any number of stations for
-basically the same numerical cost, so the user is encouraged to include
-as many stations as conceivably useful in the \texttt{STATIONS} file,
-which looks like this:
-
-\begin{figure}[H]
-\noindent \begin{centering}
-{\small \includegraphics{figures/STATIONS_basin_explained} }
-\par\end{centering}
-
-\label{fig:Sample-STATIONS-file.}\caption{Sample \texttt{STATIONS} file. Station latitude and longitude should
-be provided in geographical coordinates. The width of the station
-label should be no more than 32 characters (see \texttt{MAX\_LENGTH\_STATION\_NAME}
-in the \texttt{constants.h} file), and the network label should be
-no more than 8 characters (see \texttt{MAX\_LENGTH\_NETWORK\_NAME}
-in the \texttt{constants.h} file).}
-\end{figure}
-
-Each line represents one station in the following format:
-\begin{lyxcode}
-{\small Station~Network~Latitude~(degrees)~Longitude~(degrees)~Elevation~(m)~burial~(m)~}{\small \par}
-\end{lyxcode}
-The solver \texttt{xspecfem3D} filters the list of stations in file
-\texttt{DATA/STATIONS} to exclude stations that are not located within
-the region given in the \texttt{Par\_file} (between \texttt{LATITUDE\_MIN}
-and \texttt{LATITUDE\_MAX} and between \texttt{LONGITUDE\_MIN} and
-\texttt{LONGITUDE\_MAX}). The filtered file is called \texttt{DATA/STATIONS\_FILTERED}.
-
-Solver output is provided in the \texttt{OUTPUT\_FILES} directory
-in the \texttt{output\_solver.txt} file. Output can be directed to
-the screen instead by uncommenting a line in \texttt{constants.h}:
-\begin{lyxcode}
-!~uncomment~this~to~write~messages~to~the~screen~
-
-!~integer,~parameter~::~IMAIN~=~ISTANDARD\_OUTPUT~
-\end{lyxcode}
-On PC clusters the seismogram files are generally written to the local
-disks (the path \texttt{LOCAL\_PATH} in the \texttt{Par\_file}) and
-need to be gathered at the end of the simulation.
-
-While the solver is running, its progress may be tracked by monitoring
-the `\texttt{\small timestamp{*}}{\small ' files in the ~}\\
-{\small{} }\texttt{\small OUTPUT\_FILES}{\small{} directory. These tiny
-files look something like this:}{\small \par}
-\begin{lyxcode}
-Time~step~\#~~~~~~~~~~10000~
-
-Time:~~~~~108.4890~~~~~~seconds~
-
-Elapsed~time~in~seconds~=~~~~1153.28696703911~
-
-Elapsed~time~in~hh:mm:ss~=~~~~~0~h~19~m~13~s~
-
-Mean~elapsed~time~per~time~step~in~seconds~=~~~~~0.115328696703911~
-
-Max~norm~displacement~vector~U~in~all~slices~(m)~=~~~~1.0789589E-02~
-\end{lyxcode}
-The \texttt{\small timestamp{*}}{\small{} files provide the }\texttt{\small Mean
-elapsed time per time step in seconds}{\small , which may be used
-to assess performance on various machines (assuming you are the only
-user on a node), as well as the }\texttt{\small Max norm displacement
-vector U in all slices~(m)}{\small . If something is wrong with the
-model, the mesh, or the source, you will see the code become unstable
-through exponentially growing values of the displacement and fluid
-potential with time, and ultimately the run will be terminated by
-the program. You can control the rate at which the timestamp files
-are written based upon the parameter }\texttt{\small NTSTEP\_BETWEEN\_OUTPUT\_INFO}{\small{}
-in the }\texttt{\small Par\_file}{\small .}{\small \par}
-
-Having set the \texttt{Par\_file} parameters, and having provided
-the \texttt{CMTSOLUTION} (or the \texttt{FORCESOLUTION}) and \texttt{STATIONS}
-files, you are now ready to launch the solver! This is most easily
-accomplished based upon the \texttt{go\_solver} script (See Chapter
-\ref{cha:Scheduler} for information about running through a scheduler,
-e.g., LSF). You may need to edit the last command at the end of the
-script that invokes the \texttt{mpirun} command. The \texttt{runall}
-script compiles and runs both \texttt{xgenerate\_databases} and \texttt{xspecfem3D}
-in sequence. This is a safe approach that ensures using the correct
-combination of distributed database output and solver input.
-
-It is important to realize that the CPU and memory requirements of
-the solver are closely tied to choices about attenuation (\texttt{ATTENUATION})
-and the nature of the model (i.e., isotropic models are cheaper than
-anisotropic models). We encourage you to run a variety of simulations
-with various flags turned on or off to develop a sense for what is
-involved.
-
-For the same model, one can rerun the solver for different events
-by simply changing the \texttt{CMTSOLUTION} or \texttt{FORCESOLUTION}
-file, or for different stations by changing the \texttt{STATIONS}
-file. There is no need to rerun the \texttt{xgenerate\_databases}
-executable. Of course it is best to include as many stations as possible,
-since this does not add to the cost of the simulation.
-
-We have also added the ability to run several calculations (several earthquakes)
-in an embarrassingly-parallel fashion from within the same run;
-this can be useful when using a very large supercomputer to compute
-many earthquakes in a catalog, in which case it can be better from
-a batch job submission point of view to start fewer and much larger jobs,
-each of them computing several earthquakes in parallel.
-To turn that option on, set parameter \texttt{NUMBER\_OF\_SIMULTANEOUS\_RUNS}
-to a value greater than 1 in file setup/constants.h.in before
-configuring and compiling the code.
-When that option is on, of course the number of processor cores used to start
-the code in the batch system must be a multiple of \texttt{NUMBER\_OF\_SIMULTANEOUS\_RUNS},
-all the individual runs must use the same number of processor cores,
-which as usual is \texttt{NPROC} in the input file \texttt{DATA/Par\_file},
-and thus the total number of processor cores to request from the batch system
-should be \texttt{NUMBER\_OF\_SIMULTANEOUS\_RUNS $\times$ NPROC}.
-All the runs to perform must be placed in directories called \texttt{run0001}, \texttt{run0002}, \texttt{run0003} and so on (with exactly four digits)
-and you must create a link from the root directory of the code
-to the first copy of the executable programs by typing \texttt{ln -s run0001/bin bin}.
-
-\chapter{\label{cha:Fault-Kinematics-Dynamics}Kinematic and dynamic fault
-sources}
-
-SPECFEM3D can handle finite fault sources of two kinds:
-\begin{enumerate}
-\item \emph{Kinematic}: the spatio-temporal distribution of slip rate is
-prescribed along the fault surface
-\item \emph{Dynamic}: a friction law and initial stresses are prescribed
-on the fault, a spontaneous rupture process is computed.
-\end{enumerate}
-
-\section{\label{sec:Mesh-Generation-with}Mesh Generation with Split Nodes}
-
-Faults need to be handled in a special way during mesh generation.
-A fault surface must lie at the interface between elements (the mesh
-must honor the fault surfaces). Moreover, a fault is made of two surfaces
-in contact. Each of these two surfaces needs a separate set of nodes.
-This approach is known as \textquotedbl{}split nodes\textquotedbl{}.
-Currently faults can only be run with \texttt{xdecompose\_mesh} and
-\texttt{CUBIT}. \texttt{xmeshfem3D} is not yet ready to handle faults.\\
-To facilitate the mesh generation with split nodes in CUBIT, we need
-to separate the two fault surfaces by a small distance, effectively
-creating a tiny opening of the fault (Figure \ref{fig:examples.splitnodes},
-\ref{fig:examples.splitnodes-surfacetrace}). Note that only the interior
-of the fault must be opened, its edges must remain closed (except
-the edge on the free surface). The fault is automatically closed later
-by SPECFEM3D.\\
-
-Here is an example CUBIT script to generate a mesh with split nodes
-for a buried vertical strike-slip fault:
-\begin{lyxcode}
-~reset~~\\
-~brick~x~10~y~10~z~10~~\\
-~webcut~volume~all~with~plane~xplane~~\\
-~webcut~volume~all~with~plane~yplane~~\\
-~webcut~volume~all~with~plane~xplane~offset~3~~\\
-~webcut~volume~all~with~plane~zplane~offset~3~~\\
-~webcut~volume~all~with~plane~zplane~offset~-3~~\\
-~imprint~all~~\\
-~merge~all~~\\
-~unmerge~surf~160~~\\
-~mesh~vol~all~~\\
-~set~node~constraint~off~~\\
-~node~in~surf~168~move~X~0~Y~0.01~Z~0~~\\
-~node~in~surf~160~move~X~0~Y~-0.01~Z~0~\\
-
-\end{lyxcode}
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[scale=0.3]{figures/faultmesh} \\
- \includegraphics[scale=0.3]{figures/surf168} \includegraphics[scale=0.3]{figures/splitnodes}
-\par\end{centering}
-
-\caption{Screenshots of the CUBIT example for the embedded fault with split
-nodes: The entire mesh is decomposed into several volumes bounding
-the fault (top), the fault surface 168 is shown as orange squares
-(middle) and split nodes of the fault shown as orange dots(bottom).
-Note that only interior nodes of the fault are split while those on
-the edges of the fault surface are touching each other.}
-
-\label{fig:examples.splitnodes}
-\end{figure}
-
-\begin{lyxcode}
-
-\end{lyxcode}
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[scale=0.55]{figures/splitnodes_surfacetrace} \\
-
-\par\end{centering}
-
-\caption{Screenshots of a CUBIT example showing split nodes for a fault reaching
-the surface. Surface trace of the fault is shown in orange. Note that
-the edges of the fault are not split while the interior nodes are
-offset by a small distance on either side of the fault}
-
-\label{fig:examples.splitnodes-surfacetrace}
-\end{figure}
-
-The CUBIT scripts ({*}.jou and {*}.py) in the directory \texttt{EXAMPLES}
-generate more complicated meshes. The {*}.py files are Python scripts
-that execute CUBIT commands and use the CUBIT-python interface for
-SPECFEM3D (see next section). The Python language allows to define
-and manipulate variables to parameterize the mesh. Alternatively,
-the Python script can call a CUBIT journal file ({*}.jou), which looks
-like the example above. Variables can be defined and manipulated there
-using the $\mathtt{APREPRO}$ language built in CUBIT.\\
-
-Note that you should avoid gaps in the list of indices of mesh objects
-with the following CUBIT command:
-\begin{lyxcode}
-$\mathtt{compress\: ids\: hex\: face\: edge\: node}$
-\end{lyxcode}
-(otherwise you will get a segmentation fault during domain decomposition)
-
-\section{CUBIT-Python Scripts for Faults}
-
-The mesh generated in CUBIT needs to be processed and exported in
-a format compatible with SPECFEM3D. This is achieved in the Python
-scripts by calling the Python-CUBIT interface functions defined in
-the CUBIT directory:
-\begin{enumerate}
-\item Function $\mathtt{define\_bc}$ (or $\mathtt{boundary\_definition.py}$)
-must be called to set up the absorbing boundaries database
-\item Function $\mathtt{fault\_input}$ must be called once for each fault
-to set up the fault database
-\item Function $\mathtt{cubit2specfem3d.export2SPECFEM3D}$ must be called
-at the very end of the script to export the mesh in a SPECFEM3D format.
-\end{enumerate}
-The functions in \#1 and \#3 are for the bulk and are documented in
-Section \ref{subsec:Exporting-the-Mesh}. We focus here on \#2:
-\begin{description}
-\item [{Function:}] $\mathtt{fault\_input}$
-\item [{Purpose:}] export a fault mesh database from CUBIT to a SPECFEM3D-compliant
-file
-\item [{Syntax:}] fault\_input(id\_fault, ids\_surf\_1, ids\_surf\_2)
-\item [{Inputs:}]~\end{description}
-\begin{itemize}
-\item id\_fault integer index assigned to the fault. (The user must number
-all the faults, starting at 1, with unit increments)
-\item ids\_surf\_1 list of CUBIT ids of all surfaces that form side 1 of
-the fault.
-\item ids\_surf\_2 list of CUBIT ids of all surfaces that form side 2 of
-the fault. (The user must decide which side of the fault is side 1.
-This choice affects the sign conventions of fault quantities as explained
-in Section \ref{sec:Sign-Convention-for}).\end{itemize}
-\begin{description}
-\item [{Outputs:}] file $\mathtt{fault\_file\_X.dat}$, where X is the
-fault id (id\_fault).
-\item [{Example:}] For the example in Section \ref{sec:Mesh-Generation-with}:\end{description}
-\begin{lyxcode}
-$\hphantom{}$A1~=~{[}168{]}
-
-$\hphantom{}$A2~=~{[}160{]}
-
-$\hphantom{}$fault\_input(1,A1,A2)
-\end{lyxcode}
-
-\section{Examples}
-
-The package includes examples, the SCEC benchmark problems:
-\begin{itemize}
-\item TPV5, a planar vertical strike-slip fault
-\item TPV14 and TPV15, vertical strike-slip fault system with a fault branch
-\item TPV102 and TPV103, rate-and-state benchmarks
-\item TPV22 and TPV23, step-over benchmarks
-\item TPV16, heterogenous initial stress
-\end{itemize}
-and
-\begin{itemize}
-\item Splay fault models from Wendt et al. (2009)\\
-
-\end{itemize}
-Read the documents in the directory \texttt{EXAMPLES/*/description}.
-They contain a description of the example and additional instructions
-to run it. Visualize the results with the Matlab scripts in the directory
-$\mathtt{EXAMPLES/*/post}$
-
-\section{\label{sec:Sign-Convention-for}Sign Convention for Fault Quantities}
-
-During mesh generation, the fault is defined by two surfaces in contact.
-Let's denote as \textquotedbl{}\emph{side 1}\textquotedbl{} the SECOND
-surface declared by the user in the call to the python function \textquotedbl{}\emph{fault\_input}\textquotedbl{},
-and the FIRST surface as \textquotedbl{}\emph{side 2}\textquotedbl{}.
-The local coordinate system on the fault is defined as the right-handed
-coordinate system defined by (strike, dip, normal), where \textquotedbl{}\emph{normal}\textquotedbl{}
-is the normal vector outgoing from side 1, \textquotedbl{}\emph{dip}\textquotedbl{}
-is parallel to the along-dip direction pointing downwards, and \textquotedbl{}\emph{strike}\textquotedbl{}
-is the horizontal along-strike vector such that the system is right-handed.\\
-
-Slip is defined as displacement on side 2 minus displacement on side
-1. In the local coordinate system on the fault, positive along-strike
-slip is right-lateral and positive along-dip slip is thrust if side
-1 is on the hanging wall (normal faulting if side 1 is on the foot
-wall).\\
-
-Traction is defined as the stress induced on side 1 by side 2, which
-is the stress tensor times the normal vector outgoing from side 1.
-In the local coordinate system on the fault, the normal traction is
-negative in compression, positive along-strike traction generates
-right-lateral slip and positive along-dip traction generates thrust
-slip if side 1 is on the hanging wall (normal faulting if side 1 is
-on the foot wall).
-
-\section{Input Files}
-
-Three additional input files are required in the \texttt{DATA/} directory
-for dynamic and kinematic faults. They are $\mathtt{Par\_file\_faults}$,
-$\mathtt{FAULT\_STATIONS}$ and $\mathtt{input\_file.txt}$ (optional).
-If the former does not exist, the code assumes that there are no faults.\\
-
-\begin{lyxlist}{00.00.0000}
-\item [{\textbf{DATA/Par\_file\_faults}}] contains parameters of the fault.
-The first part of this file has a strict format:
-
-\begin{description}
-\item [{Line}] 1: Number of faults (NF)
-\item [{Lines}] 2 to NF+1: Kelvin Voigt damping (in seconds) for each fault.
-(See below how to set this parameter)
-\item [{Line}] NF+2: Type of simulation (1=dynamic , 2 = kinematic)
-\item [{Line}] NF+3: Number of time steps between updates of the time series
-outputs at selected fault points (see DATA/FAULT\_STATIONS), usually
-a large number (100s or 1000s). Note that the sampling rate of the
-time series is usually much higher.
-\item [{Line}] NF+4: Number of time steps between fault snapshot outputs
-(quantities at every fault point exported at regular times), usually
-a large number (100s or 1000s).
-\item [{Line}] NF+5: Slip velocity threshold below which frictional healing
-is set (friction coefficient is reset to its static value). If this
-value is negative healing is disabled.
-\item [{Line}] NF+6: Slip velocity threshold to define the rupture front.
-Only used for outputs.
-\end{description}
-
-The rest of this file is made of namelist input blocks (see \textquotedbl{}namelist\textquotedbl{}
-in a Fortran 9x manual). The input for each fault has the following
-sequence (arguments in {[}brackets{]} are optional):
-
-\&\textbf{BEGIN\_FAULT} /
-
-\&\textbf{INIT\_STRESS} S1, S2, S3 {[},n1, n2, n3{]} /
-
-followed by (n1+n2+n3) \&\textbf{DIST2D} blocks
-
-\&\textbf{SWF} mus, mud, dc {[}, nmus, nmud, ndc{]} /
-
-followed by (nmus+nmud+ndc) \&\textbf{DIST2D} blocks\\
-
-\begin{lyxlist}{00.00.0000}
-\item [{\&\textbf{INIT\_STRESS}}] input block sets the initial fault stresses
-relative to the foot-wall side of the fault. Initial stresses are
-composed of a constant background value possibly overwritten in prescribed
-regions by heterogeneous distributions (see \&\textbf{DIST2D} blocks
-below):
-
-\begin{description}
-\item [{S1}] = constant background value of along-strike shear stress (positive
-in the usual strike direction)
-\item [{S2}] = constant background value of along-dip shear (positive is
-down-dip, normal faulting)
-\item [{S3}] = constant background value of normal stress (negative in
-compression)
-\item [{n1}] = number of heterogeneous items for along-strike shear stress
-{[}default is 0{]}
-\item [{n2}] = number of heterogeneous items for along-dip shear stress
-{[}default is 0{]}
-\item [{n3}] = number of heterogeneous items for normal stress {[}default
-is 0{]}
-\end{description}
-\item [{\&\textbf{SWF}}] input block sets the slip-weakening friction parameters
-of the fault:
-
-\begin{description}
-\item [{mus}] = constant background value of static friction coefficient
-\item [{mud}] = constant background value of dynamic friction coefficient
-\item [{dc}] = constant background value of critical slip-weakening distance
-\item [{nmus}] = number of heterogeneous items for static friction coefficient
-{[}default is 0{]}
-\item [{nmud}] = number of heterogeneous items for dynamic friction coefficient
-{[}default is 0{]}
-\item [{ndc}] = number of heterogeneous items for critical slip-weakening
-distance {[}default is 0{]}\\
-
-\end{description}
-\item [{\&\textbf{DIST2D}}] input blocks modify (overwrite) the value of
-a fault parameter by a heterogeneous spatial distribution:
-
-\&\textbf{DIST2D} shape='$\mathtt{square}$', val, xc, yc, zc, l /
-
-$\;$sets a constant value (val) within a cube with center (xc,yc,zc)
-and edge size l.
-
-\&\textbf{DIST2D} shape='$\mathtt{rectangle}$', val, xc, yc, zc,
-lx, ly, lz /
-
-$\;$sets a constant value (val) within a rectangular prism with center
-(xc,yc,zc) and edge sizes (lx,ly,lz).
-
-\&\textbf{DIST2D} shape='$\mathtt{rectangle-taper}$', val, valh,
-xc, yc, zc, lx, ly, lz /
-
-$\;$sets a vertical linear gradient within a rectangular prism with
-center (xc,yc,zc) and edge sizes (lx,ly,lz). Values vary linearly
-as a function of vertical position z between value val at z = zc-lz/2
-and value valh at z = zc+lz/2 .
-
-\&\textbf{DIST2D} shape='$\mathtt{circular}$', val, xc, yc, zc, r
-/
-
-$\;$sets a constant value (val) within a sphere with center (xc,yc,zc)
-and radius r. \\
-
-\end{lyxlist}
-\end{lyxlist}
-
-\begin{lyxlist}{00.00.0000}
-\item [{\textbf{DATA/FAULT\_STATIONS}}] Stations in the fault plane.
-
-\textbf{Line 1} : number of stations.
-
-\textbf{Line 2 to end}: 5 columns: X, Y, Z (-depth), station name,
-fault-id
-
-The fault-id identifies the fault that contains the station. It is
-the index of appearance in the faults list after line 2 of Par\_file\_faults
-
-\end{lyxlist}
-
-\begin{lyxlist}{00.00.0000}
-\item [{\textbf{DATA/input\_file.txt}}] Heterogeneous stresses and friction
-parameters are documented in page 10 of
-
-\begin{lyxlist}{00.00.0000}
-\item [{$\mathtt{EXAMPLES/tpv16/description/TPV16\_17\_Description\_v03.pdf}$}]~
-\end{lyxlist}
-
-To activate this feature, in fault\_solver\_dynamic.f90 set \textbf{TPV16}
-= .true. then re-compile the code.
-
-\end{lyxlist}
-
-\section{\label{sec:Setting-the-Kelvin-Voigt}Setting the Kelvin-Voigt Damping
-Parameter }
-
-The purpose of the Kelvin-Voigt viscosity in the dynamic fault solver
-is to damp spurious oscillations generated by the fault slip at frequencies
-that are too high to be resolved by the mesh. The viscosity $\mathtt{eta}$
-(in seconds) depends on the size of the elements on the fault. Here
-is how to set it:
-\begin{enumerate}
-\item Determine the average linear size of the elements on the fault plane,
-$\mathtt{h\_fault}$. Usually this value is prescribed by the user
-during mesh generation. Otherwise it can be found by inspection of
-the mesh inside the CUBIT GUI.
-\item Use the Matlab function $\mathtt{utils/critical\_timestep.m}$ to
-compute\\
- $\mathtt{dtc\_fault}=\mathtt{critical\_timestep\left(c_{p},h\_fault,ngll\right)}$.
-\\
-This is the critical time step in an elastic medium for a hypothetical
-element of cubic shape with size equal to $\mathtt{h\_fault}$.
-\item Set $\mathtt{eta}$ in $\mathtt{Par\_file\_faults}$ to (0.1 to 0.3)$\ensuremath{\times}\mathtt{dtc\_fault}$.
-A larger $\mathtt{eta}$ damps high-frequencies more aggressively
-but it might also affect lower frequencies and rupture speed.
-\end{enumerate}
-Viscosity reduces numerical stability: the critical timestep in a
-simulation with Kelvin-Voigt damping needs to be smaller than that
-in a purely elastic simulation. Here is how to set the time step accordingly:
-\begin{enumerate}
-\item Run a test simulation without viscosity ($\mathtt{eta}$=0 and only
-a few time steps)
-\item Look for the \textquotedbl{}maximum suggested time step\textquotedbl{}
-in \texttt{OUTPUT\_FILES/output\_mesher.txt}. This is the critical
-timestep of a purely elastic simulation, $\mathtt{dtc\_bulk}$.
-\item Reset the timestep of the simulation with a Kelvin-Voigt material
-to a value smaller than\\
- $\mathtt{dtc\_kv}=\mathtt{eta}\left(\sqrt{1+\mathtt{dtc\_bulk^{2}}/\mathtt{eta^{2}}}-1\right)$
-\end{enumerate}
-Note that in general $\mathtt{dtc\_bulk}$ is smaller than $\mathtt{dtc\_fault}$,
-because elements off the fault might be smaller or more distorted
-than element faces on the fault.
-
-\section{Output Files }
-
-Several output files are saved in \texttt{OUTPUT\_FILES/}:\\
-
-\begin{enumerate}
-\item Seismograms for each station on the fault plane given in \texttt{DATA/FAUL\_STATIONS}.
-One output file is generated for each station, named after the station.
-The files are ascii and start with a header (22 lines long) followed
-by a data block with the following format, one line per time sample:\\
-
-Column 1 = Time (s)
-
-Column 2 = horizontal right-lateral slip (m)
-
-Column 3 = horizontal right-lateral slip rate (m/s)
-
-Column 4 = horizontal right-lateral shear stress (MPa)
-
-Column 5 = vertical up-dip slip (m)
-
-Column 6 = vertical up-dip slip rate (m/s)
-
-Column 7 = vertical up-dip shear stress (MPa)
-
-Column 8 = normal stress (MPa)\\
-
-The stresses are relative to the footwall side of the fault (this
-convention controls their sign, but not their amplitude). Slip is
-defined as displacement of the hanging wall relative to the footwall.
-
-\item Seismograms at stations in the bulk (out of the fault plane) given
-in \texttt{DATA/STATIONS}.
-\item Rupture time files are named $\mathtt{Rupture\_time\_FAULT}$-$\mathtt{id}$.
-One file is generated for each fault. The files are ascii and start
-with a header (12 lines long) followed by a data block with the following
-format, one line per fault node:
-
-Column 1 = horizontal coordinate, distance along strike (m)
-
-Column 2 = vertical coordinate, distance down-dip (m)
-
-Column 3 = rupture time (s)
-
-\item Fault quantities (slip, slip rate, stresses, etc) at regular times
-are stored in binary data files called $\mathtt{Snapshot\#it\#.bin}$,
-where \#it\# is the timestep number. These can be read in Matlab with
-the function $\mathtt{utils/FSEM3D\_snapshot.m}$
-\end{enumerate}
-
-\section{Post-processing and Visualization }
-
-Some Matlab functions for post-processing and visualization are included
-in directory $\mathtt{utils}$. The functions are internally documented
-(see their matlab help).\\
-
-\texttt{FSEM3D\_snapshot} reads a fault data snapshot\\
-
-The directories \texttt{EXAMPLES/*/post} contain additional Matlab
-scripts to generate figures specific to each example.
-
-\chapter{\label{cha:Adjoint-Simulations}Adjoint Simulations}
-
-Adjoint simulations are generally performed for two distinct applications.
-First, they can be used in point source moment-tensor inversions, or source imaging for
-earthquakes with large ruptures such as the Lander's earthquake \citep{WaHe94}.
-Second, they can be used to generate finite-frequency sensitivity
-kernels that are a critical part of tomographic inversions based upon
-3D reference models \citep{TrTaLi05,LiTr06,TrKoLi08,LiTr08}. In either
-case, source parameter or velocity structure updates are sought to
-minimize a specific misfit function (e.g., waveform or traveltime
-differences), and the adjoint simulation provides a means of computing
-the gradient of the misfit function and further reducing it in successive
-iterations. Applications and procedures pertaining to source studies
-and finite-frequency kernels are discussed in Sections~\ref{sec:Adjoint-simulation-sources}
-and \ref{sec:Adjoint-simulation-finite}, respectively. The two related
-parameters in the \texttt{Par\_file} are \texttt{SIMULATION\_TYPE}
-(1, 2 or 3) and the \texttt{SAVE\_FORWARD} (boolean).
-
-\section{\label{sec:Adjoint-simulation-sources}Adjoint Simulations for Sources
-Only (not for the Model)}
-
-When a specific misfit function between data and synthetics is minimized to invert
-for earthquake source parameters, the gradient of the misfit function
-with respect to these source parameters can be computed by placing
-time-reversed seismograms at the receivers as virtual sources
-in an adjoint simulation. Then the value of the gradient is obtained
-from the adjoint seismograms recorded at the original earthquake location.
-\begin{enumerate}
-\item \textbf{Prepare the adjoint sources} \label{enu:Prepare-the-adjoint}
-
-\begin{enumerate}
-\item First, run a regular forward simlation (\texttt{SIMULATION\_TYPE =
-1} and \texttt{SAVE\_FORWARD = .false.}). You can automatically set
-these two variables using the \texttt{\small utils/change\_simulation\_type.pl}{\small{}
-script:}{\small \par}
-
-\begin{lyxcode}
-utils/change\_simulation\_type.pl~-f~
-\end{lyxcode}
-
-and then collect the recorded seismograms at all the stations given
-in \texttt{DATA/STATIONS}.
-
-\item Then select the stations for which you want to compute the time-reversed
-adjoint sources and run the adjoint simulation, and compile them into
-the \texttt{DATA/STATIONS\_ADJOINT} file, which has the same format
-as the regular \texttt{DATA/STATIONS} file.
-
-\begin{itemize}
-\item Depending on what type of misfit function is used for the source inversion,
-adjoint sources need to be computed from the original recorded seismograms
-for the selected stations and saved in a sub-directory called \texttt{SEM/}
-in the root directory of the code, with the format \texttt{NT.STA.BX?.adj}, where \texttt{NT},
-\texttt{STA} are the network code and station name given in the \texttt{DATA/STATIONS\_ADJOINT}
-file, and \texttt{BX?} represents the component name of a particular
-adjoint seismogram. Please note that the band code can change depending
-on your sampling rate (see Appendix~\ref{cha:channel-codes} for
-further details).
-\item The adjoint seismograms are in the same format as the original seismogram
-(\texttt{NT.STA.BX?.sem?}), with the same start time, time interval
-and record length.
-\end{itemize}
-\item Notice that even if you choose to time reverse only one component
-from one specific station, you still need to supply all three components
-because the code is expecting them (you can set the other two components
-to be zero).
-\item Also note that since time-reversal is done in the code itself, no
-explicit time-reversing is needed for the preparation of the adjoint
-sources, i.e., the adjoint sources are in the same forward time sense
-as the original recorded seismograms.
-\end{enumerate}
-\item \textbf{Set the related parameters and run the adjoint simulation}\\
- In the \texttt{DATA/Par\_file}, set the two related parameters to
-be \texttt{SIMULATION\_TYPE = 2} and \texttt{SAVE\_FORWARD = .false.}.
-More conveniently, use the scripts \texttt{utils/change\_simulation\_type.pl}
-to modify the \texttt{Par\_file} automatically (\texttt{change\_simulation\_type.pl
--a}). Then run the solver to launch the adjoint simulation.
-\item \textbf{Collect the seismograms at the original source location}
-
-After the adjoint simulation has completed successfully, collect the
-seismograms from \texttt{LOCAL\_PATH}.
-\begin{itemize}
-\item These adjoint seismograms are recorded at the locations of the original
-earthquake sources given by the \texttt{DATA/CMTSOLUTION} file, and
-have names of the form \texttt{NT.S?????.S??.sem} for the six-component
-strain tensor (\texttt{SNN,SEE,SZZ,SNE,SNZ,SEZ}) at these locations,
-and ~\\
- \texttt{NT.S?????.BX?.sem} for the three-component displacements
-(\texttt{BXN,BXE,BXZ}) recorded at these locations.
-\item \texttt{S?????} denotes the source number; for example, if the original
-\texttt{CMTSOLUTION} provides only a point source, then the seismograms
-collected will start with \texttt{S00001}.
-\item These adjoint seismograms provide critical information for the computation
-of the gradient of the misfit function.
-\end{itemize}
-\end{enumerate}
-
-\section{\label{sec:Adjoint-simulation-finite}Adjoint Simulations for Finite-Frequency
-Kernels (Kernel Simulation)}
-
-Finite-frequency sensitivity kernels are computed in two successive
-simulations (please refer to \citet{LiTr06} and \citet{TrKoLi08}
-for details).
-\begin{enumerate}
-\item \textbf{Run a forward simulation with the state variables saved at
-the end of the simulation}
-
-Prepare the \texttt{\small CMTSOLUTION}{\small{} and }\texttt{\small STATIONS}{\small{}
-files, set the parameters }\texttt{\small SIMULATION\_TYPE}{\small {}
-}\texttt{\small =}{\small {} }\texttt{\small 1}{\small{} and }\texttt{\small SAVE\_FORWARD
-=}{\small {} }\texttt{\small .true.}{\small{} in the }\texttt{\small Par\_file}{\small{}
-(}\texttt{\small change\_simulation\_type -F}{\small ), and run the
-solver.}{\small \par}
-\begin{itemize}
-\item Notice that attenuation is not implemented yet for the computation
-of finite-frequency kernels; therefore set \texttt{ATTENUATION = .false.}
-in the \texttt{Par\_file}.
-\item We also suggest you modify the half duration of the \texttt{CMTSOLUTION}
-to be similar to the accuracy of the simulation (see Equation \ref{eq:shortest_period})
-to avoid too much high-frequency noise in the forward wavefield, although
-theoretically the high-frequency noise should be eliminated when convolved
-with an adjoint wavefield with the proper frequency content.
-\item This forward simulation differs from the regular simulations (\texttt{\small SIMULATION\_TYPE}{\small {}
-}\texttt{\small =}{\small {} }\texttt{\small 1}{\small{} and }\texttt{\small SAVE\_FORWARD}{\small {}
-}\texttt{\small =}{\small {} }\texttt{\small .false.}{\small ) described
-in the previous chapters in that the state variables for the last
-time step of the simulation, including wavefields of the displacement,
-velocity, acceleration, etc., are saved to the }\texttt{\small LOCAL\_PATH}{\small{}
-to be used for the subsequent simulation. }{\small \par}
-\item For regional simulations, the files recording the absorbing boundary
-contribution are also written to the \texttt{LOCAL\_PATH} when \texttt{SAVE\_FORWARD
-= .true.}.
-\end{itemize}
-\item \textbf{Prepare the adjoint sources}
-
-The adjoint sources need to be prepared the same way as described
-in the Section \ref{enu:Prepare-the-adjoint}.
-\begin{itemize}
-\item In the case of travel-time finite-frequency kernel for one source-receiver
-pair, i.e., point source from the \texttt{CMTSOLUTION}, and one station
-in the \texttt{STATIONS\_ADJOINT} list, we supply a sample program
-in \texttt{utils/adjoint\_sources/traveltime/xcreate\_adjsrc\_traveltime}
-to cut a certain portion of the original displacement seismograms
-and convert it into the proper adjoint source to compute the finite-frequency
-kernel.
-
-\begin{lyxcode}
-xcreate\_adjsrc\_traveltime~t1~t2~ifile{[}0-5{]}~E/N/Z-ascii-files~{[}baz{]}
-\end{lyxcode}
-
-where \texttt{t1} and \texttt{t2} are the start and end time of the
-portion you are interested in, \texttt{ifile} denotes the component
-of the seismograms to be used (0 for all three components, 1 for East,
-2 for North, and 3 for vertical, 4 for transverse, and 5 for radial
-component), \texttt{E/N/Z-ascii-files} indicate the three-component
-displacement seismograms in the right order, and \texttt{baz} is the
-back-azimuth of the station. Note that \texttt{baz} is only supplied
-when \texttt{ifile} = 4 or 5.
-
-\item Similarly, a sample program to compute adjoint sources for amplitude
-finite-frequency kernels may be found in \texttt{utils/adjoint\_sources/amplitude}
-and used in the same way as described for traveltime measurements
-
-\begin{lyxcode}
-xcreate\_adjsrc\_amplitude~t1~t2~ifile{[}0-5{]}~E/N/Z-ascii-files~{[}baz{]}.
-\end{lyxcode}
-\end{itemize}
-\item \textbf{Run the kernel simulation}
-
-With the successful forward simulation and the adjoint source ready
-in the \texttt{SEM/} directory, set \texttt{SIMULATION\_TYPE
-= 3} and \texttt{SAVE\_FORWARD = .false.} in the \texttt{Par\_file}
-(you can use \texttt{change\_simulation\_type.pl -b}), and rerun the
-solver.
-\begin{itemize}
-\item The adjoint simulation is launched together with the back reconstruction
-of the original forward wavefield from the state variables saved from
-the previous forward simulation, and the finite-frequency kernels
-are computed by the interaction of the reconstructed forward wavefield
-and the adjoint wavefield.
-\item The back-reconstructed seismograms at the original station locations
-are saved to the \texttt{LOCAL\_PATH} at the end of the kernel simulations,
-and can be collected to the local disk.
-\item These back-constructed seismograms can be compared with the time-reversed
-original seismograms to assess the accuracy of the backward reconstruction,
-and they should match very well.
-\item The arrays for density, P-wave speed and S-wave speed kernels are
-also saved in the \texttt{LOCAL\_PATH} with the names \texttt{proc??????\_rho(alpha,beta)\_kernel.bin},
-where \texttt{proc??????} represents the processor number, \texttt{rho(alpha,beta)}
-are the different types of kernels.
-\end{itemize}
-\item \textbf{Run the anisotropic kernel simulation}
-
-Instead of the kernels for the isotropic wave speeds, you can also
-compute the kernels for the 21 independent components $C_{IJ},\, I,J=1,...,6$
-(using Voigt's notation) of the elastic tensor in the cartesian coordinate
-system. This is done by setting \texttt{ANISOTROPIC\_KL} \texttt{=}
-\texttt{.true.} in \texttt{constants.h} before compiling the package.
-The definition of the parameters $C_{IJ}$ in terms of the corresponding
-components $c_{ijkl},ijkl,i,j,k,l=1,2,3$ of the elastic tensor in
-cartesian coordinates follows \citet{ChTr07}. The 21 anisotropic
-kernels are saved in the \texttt{LOCAL\_PATH} in one file with the
-name of \texttt{proc??????\_cijkl\_kernel.bin} (with \texttt{proc??????}
-the processor number). The output kernels correspond to absolute perturbations
-$\delta C_{IJ}$ of the elastic parameters and their unit is in $s/GPa/km^{3}$.
-For consistency, the output density kernels with this option turned
-on are for a perturbation $\delta\rho$ (and not $\frac{\delta\rho}{\rho}$)
-and their unit is in s / (kg/m$^{3}$) / km$^{3}$. These `primary'
-anisotropic kernels can then be combined to obtain the kernels for
-different parameterizations of anisotropy. This can be done, for example,
-when combining the kernel files from slices into one mesh file (see
-Section~\ref{sec:Finite-Frequency-Kernels}).
-
-If \texttt{ANISOTROPIC\_KL} \texttt{=} \texttt{.true.} by additionally
-setting \texttt{ANISOTROPIC\_KL} \texttt{=} \texttt{.true.} in \texttt{constants.h}
-the package will save anisotropic kernels paramterized as velocities
-related to transverse isotropy based on the the Chen and Tromp parameters
-\citet{ChTr07}. The kernels are saved as relative perturbations for
-horizontal and vertical P and S velocities, $\alpha_{v},\alpha_{h},\beta_{v},\beta_{h}$.
-Explicit relations can be found in appendix B. of \citet{SiLiTrTr07b}
-
-The anisotropic kernels are only currently available for CPU mode.
-
-\end{enumerate}
-In general, the three steps need to be run sequentially to assure
-proper access to the necessary files. If the simulations are run through
-some cluster scheduling system (e.g., LSF), and the forward simulation
-and the subsequent kernel simulations cannot be assigned to the same
-set of computer nodes, the kernel simulation will not be able to access
-the database files saved by the forward simulation. Solutions for
-this dilemma are provided in Chapter~\ref{cha:Scheduler}. Visualization
-of the finite-frequency kernels is discussed in Section~\ref{sec:Finite-Frequency-Kernels}.
-
-%------------------------------------------------------------------------------------------------%
-
-\chapter{\label{cha:tomo}Doing tomography, i.e., updating the model based on the sensitivity kernels obtained}
-
-%------------------------------------------------------------------------------------------------%
-
-UNDER CONSTRUCTION (March 2015). Will be added around April / May 2015.
-
-\section{Tomographic inversion / imaging using the sensitivity kernels obtained}
-
-OLD VERSION OF THE SECTION, WILL SOON BE IMPROVED AND ENRICHED.\\
-
-One of the fundamental reasons for computing sensitivity kernels (Section~\ref{sec:Adjoint-simulation-finite})
-is to use them within a tomographic inversion. In other words, use
-recorded seismograms, make measurements with synthetic seismograms,
-and use the misfit between the two to iteratively improve the model
-described by (at least) $V_{{\rm p}}$, $V_{{\rm s}}$, and $\rho$
-as a function of space.
-
-Whatever misfit function you use for the tomographic inversion (you can find several
-examples in \citet{TrKoLi08} and \citet{TrTaLi05}), you will weight
-the sensitivity kernels with measurements. Furthermore, you will use
-as many measurements (stations, components, time windows) as possible
-per event; hence, we call these composite kernels ``event kernels,''
-which are volumetric fields representing the gradient of the misfit
-function with respect to one of the variables (\eg $V_{{\rm s}}$).
-The basic features of an adjoint-based tomographic inversion were
-illustrated in \citet{TaLiTr07} and \citet{TrKoLi08} using a conjugate-gradient
-algorithm; there are dozens of versions of gradient-based inversion
-algorithms that could alternatively be used. The tomographic inversion
-of \citet{TaLiMaTr09,TaLiMaTr2010} used SPECFEM3D Cartesian as well
-as several additional components which are also stored on the CIG
-svn server, described next.
-
-The directory containing external utilities for tomographic inversion using
-SPECFEM3D Cartesian (or other packages that evaluate misfit functions
-and gradients) is in directory \texttt{utils/ADJOINT\_TOMOGRAPHY\_TOOLS/}:
-\begin{verbatim}
-flexwin/     -- FLEXWIN algorithm for automated picking of time windows
-measure_adj/ -- reads FLEXWIN output file and makes measurements,
-                with the option for computing adjoint sources
-iterate_adj/ -- various tools for iterative inversion
-                (requires pre-computed "event kernels")
-\end{verbatim}
-This directory also contains
-a brief \verb+README+ file indicating the role of the three subdirectories,
-\verb+flexwin+ \citep{Maggi2009}, \verb+measure_adj+, and \verb+iterate_adj+.
-The components for making the model update are there; however, there
-are no explicit rules for computing the model update, just as with
-any optimization problem. There are options for computing a conjugate
-gradient step, as well as a source subspace projection step.
-
-The best single file to read is probably: \verb+ADJOINT_TOMO/iterate_adj/cluster/README+.
-
-%------------------------------------------------------------------------------------------------%
-
-\chapter{Noise Cross-correlation Simulations}
-
-%------------------------------------------------------------------------------------------------%
-
-Besides earthquake simulations, SPECFEM3D Cartesian includes functionality
-for seismic noise tomography as well. In order to proceed successfully
-in this chapter, it is critical that you have already familiarized
-yourself with procedures for meshing (Chapter \ref{cha:Mesh-Generation}),
-creating distributed databases (Chapter \ref{cha:Creating-Distributed-Databases}),
-running earthquake simulations (Chapters \ref{cha:Running-the-Solver})
-and adjoint simulations (Chapter \ref{cha:Adjoint-Simulations}).
-Also, make sure you read the article `Noise cross-correlation sensitivity
-kernels' \citep{trompetal2010}, in order to understand noise simulations
-from a theoretical perspective.
-
-\section{Input Parameter Files}
-
-As usual, the three main input files are crucial: \texttt{Par\_file},
-\texttt{CMTSOLUTION} and \texttt{STATIONS}. Unless otherwise specified,
-those input files should be located in directory \texttt{DATA/}.\\
-
-\texttt{CMTSOLUTION} is required for all simulations. At a first glance,
-it may seem unexpected to have it here, since the noise simulations
-should have nothing to do with the earthquake -- \texttt{CMTSOLUTION}.
-However, for noise simulations, it is critical to have no earthquakes.
-In other words, the moment tensor specified in \texttt{CMTSOLUTION}
-must be set to zero manually!\\
-
-\texttt{STATIONS} remains the same as in previous earthquake simulations,
-except that the order of receivers listed in \texttt{STATIONS} is
-now important. The order will be used to determine the `master' receiver,
-i.e., the one that simultaneously cross correlates with the others.\\
-
-\texttt{Par\_file} also requires careful attention. A parameter called
-\texttt{NOISE\_TOMOGRAPHY} has been added which specifies the type
-of simulation to be run. \texttt{NOISE\_TOMOGRAPHY} is an integer
-with possible values 0, 1, 2 and 3. For example, when \texttt{NOISE\_TOMOGRAPHY}
-equals 0, a regular earthquake simulation will be run. When it is
-1/2/3, you are about to run step 1/2/3 of the noise simulations respectively.
-Should you be confused by the three steps, refer to \citet{trompetal2010}
-for details.\\
-
-Another change to \texttt{Par\_file} involves the parameter \texttt{NSTEP}.
-While for regular earthquake simulations this parameter specifies
-the length of synthetic seismograms generated, for noise simulations
-it specifies the length of the seismograms used to compute cross correlations.
-The actual cross correlations are thus twice this length, i.e., $2~\mathrm{NSTEP}-1$.
-The code automatically makes the modification accordingly, if \texttt{NOISE\_TOMOGRAPHY}
-is not zero.\\
-
-There are other parameters in \texttt{Par\_file} which should be given
-specific values. For instance, since the first two steps for calculating
-noise sensitivity kernels correspond to forward simulations, \texttt{SIMULATION\_TYPE}
-must be 1 when \texttt{NOISE\_TOMOGRAPHY} equals 1 or 2. Also, we
-have to reconstruct the ensemble forward wavefields in adjoint simulations,
-therefore we need to set \texttt{SAVE\_FORWARD} to \texttt{.true.}
-for the second step, i.e., when \texttt{NOISE\_TOMOGRAPHY} equals
-2. The third step is for kernel constructions. Hence \texttt{SIMULATION\_TYPE}
-should be 3, whereas \texttt{SAVE\_FORWARD} must be \texttt{.false.}.\\
-
-Finally, for most system architectures, please make sure that \texttt{LOCAL\_PATH}
-in \texttt{Par\_file} is in fact local, not globally shared. Because
-we have to save the wavefields at the earth's surface at every time
-step, it is quite problematic to have a globally shared \texttt{LOCAL\_PATH},
-in terms of both disk storage and I/O speed.
-
-\section{Noise Simulations: Step by Step}
-
-Proper parameters in those parameter files are not enough for noise
-simulations to run. We have more parameters to specify: for example,
-the ensemble-averaged noise spectrum, the noise distribution etc.
-However, since there are a few `new' files, it is better to introduce
-them sequentially. In this section, standard procedures for noise
-simulations are described.
-
-\subsection{Pre-simulation}
-\begin{itemize}
-\item As usual, we first configure the software package using:
-
-\texttt{./configure FC=ifort MPIFC=mpif90}
-
-Use the following if SCOTCH is needed:
-
-\texttt{./configure FC=ifort MPIFC=mpif90 -{}-with-scotch-dir=/opt/scotch}\\
-
-\item Next, we need to compile the source code using:
-
-\texttt{make xgenerate\_databases}
-
-\texttt{make xspecfem3D} \\
-
-\item Before we can run noise simulations, we have to specify the noise
-statistics, e.g., the ensemble-averaged noise spectrum. Matlab scripts
-are provided to help you to generate the necessary file.
-
-\texttt{EXAMPLES/noise\_tomography/NOISE\_TOMOGRAPHY.m} (main program)\\
- \texttt{EXAMPLES/noise\_tomography/PetersonNoiseModel.m}
-
-In Matlab, simply run:
-
-\texttt{NOISE\_TOMOGRAPHY(NSTEP, DT, Tmin, Tmax, NOISE\_MODEL)}\\
-
-\texttt{DT} is given in \texttt{Par\_file}, but \texttt{NSTEP} is
-NOT the one specified in \texttt{Par\_file}. Instead, you have to
-feed $2~\mathrm{NSTEP}-1$ to account for the doubled length of cross
-correlations. \texttt{Tmin} and \texttt{Tmax} correspond to the period
-range you are interested in, whereas \texttt{NOISE\_MODEL} denotes
-the noise model you will be using. Details can be found in the Matlab
-script.
-
-After running the Matlab script, you will be given the following information
-(plus a figure in Matlab):
-
-\texttt{{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}}\\
- \texttt{the source time function has been saved in:}\\
- \texttt{/data2/yangl/3D\_NOISE/S\_squared} (note this path must be
-different)\\
- \texttt{S\_squared should be put into directory:}\\
- \texttt{OUTPUT\_FILES/NOISE\_TOMOGRAPHY/ in the SPECFEM3D Cartesian
-package}\\
-
-In other words, the Matlab script creates a file called \texttt{S\_squared},
-which is the first `new' input file we encounter for noise simulations.
-
-One may choose a flat noise spectrum rather than Peterson's noise
-model. This can be done easily by modifying the Matlab script a little.
-
-\item Create a new directory in the SPECFEM3D Cartesian package, name it
-as \texttt{OUTPUT\_FILES/NOISE\_TOMOGRAPHY/}. We will add some parameter
-files later in this folder.
-\item Put the Matlab-generated-file \texttt{S\_squared} in \texttt{OUTPUT\_FILES/NOISE\_TOMOGRAPHY/}.
-
-That's to say, you will have a file \texttt{OUTPUT\_FILES/NOISE\_TOMOGRAPHY/S\_squared}
-in the SPECFEM3D Cartesian package.
-
-\item Create a file called \texttt{OUTPUT\_FILES/NOISE\_TOMOGRAPHY/irec\_master\_noise}.
-Note that this file is located in directory \texttt{OUTPUT\_FILES/NOISE\_TOMOGRAPHY/}
-as well. In general, all noise simulation related parameter files
-go into that directory. \texttt{irec\_master\_noise} contains only
-one integer, which is the ID of the `master' receiver. For example,
-if this file contains 5, it means that the fifth receiver listed in
-\texttt{DATA/STATIONS} becomes the `master'. That's why we mentioned
-previously that the order of receivers in \texttt{DATA/STATIONS} is
-important.
-
-Note that in regional simulations, the \texttt{DATA/STATIONS} might
-contain receivers which are outside of our computational domains.
-Therefore, the integer in \texttt{irec\_master\_noise} is actually
-the ID in \texttt{DATA/STATIONS\_FILTERED} (which is generated by
-\texttt{bin/xgenerate\_databases}).
-
-\item Create a file called \texttt{OUTPUT\_FILES/NOISE\_TOMOGRAPHY/nu\_master}.
-This file holds three numbers, forming a (unit) vector. It describes
-which component we are cross-correlating at the `master' receiver,
-i.e., ${\hat{{\bf \nu}}}^{\alpha}$ in \citet{trompetal2010}. The
-three numbers correspond to E/N/Z components respectively. Most often,
-the vertical component is used, and in those cases the three numbers
-should be 0, 0 and 1.
-\item Describe the noise direction and distributions in \texttt{src/shared/noise\_tomography.f90}.
-Search for a subroutine called \texttt{noise\_distribution\_direction}
-in \texttt{noise\_tomography.f90}. It is actually located at the very
-beginning of \texttt{noise\_tomography.f90}. The default assumes vertical
-noise and a uniform distribution across the whole free surface of
-the model. It should be quite self-explanatory for modifications.
-Should you modify this part, you have to re-compile the source code.
-\end{itemize}
-
-\subsection{Simulations}
-
-With all of the above done, we can finally launch our simulations.
-Again, please make sure that the \texttt{LOCAL\_PATH} in \texttt{DATA/Par\_file}
-is not globally shared. It is quite problematic to have a globally
-shared \texttt{LOCAL\_PATH}, in terms of both disk storage and speed
-of I/O (we have to save the wavefields at the earth's surface at every
-time step).
-
-As discussed in \citet{trompetal2010}, it takes three steps/simulations
-to obtain one contribution of the ensemble sensitivity kernels:
-\begin{itemize}
-\item Step 1: simulation for generating wavefields\\
- \texttt{SIMULATION\_TYPE=1}\\
- \texttt{NOISE\_TOMOGRAPHY=1}\\
- \texttt{SAVE\_FORWARD} not used, can be either .true. or .false.
-\item Step 2: simulation for ensemble forward wavefields\\
- \texttt{SIMULATION\_TYPE=1}\\
- \texttt{NOISE\_TOMOGRAPHY=2}\\
- \texttt{SAVE\_FORWARD=.true.}
-\item Step 3: simulation for ensemble adjoint wavefields and sensitivity
-kernels\\
- \texttt{SIMULATION\_TYPE=3}\\
- \texttt{NOISE\_TOMOGRAPHY=3}\\
- \texttt{SAVE\_FORWARD=.false.}
-
-Note Step 3 is an adjoint simulation, please refer to previous chapters
-on how to prepare adjoint sources and other necessary files, as well
-as how adjoint simulations work.
-
-\end{itemize}
-Note that it is better to run the three steps continuously within
-the same job, otherwise you have to collect the saved surface movies
-from the old nodes/CPUs to the new nodes/CPUs. This process varies
-from cluster to cluster and thus cannot be discussed here. Please
-ask your cluster administrator for information/configuration of the
-cluster you are using.\\
-
-\subsection{Post-simulation}
-
-After those simulations, you have all stuff you need, either in the
-\texttt{OUTPUT\_FILES/} or in the directory specified by \texttt{LOCAL\_PATH}
-in \texttt{DATA/Par\_file} (which are most probably on local nodes).
-Collect whatever you want from the local nodes to your workstation,
-and then visualize them. This process is the same as what you may
-have done for regular earthquake simulations. Refer to other chapters
-if you have problems.\\
-
-Simply speaking, two outputs are the most interesting: the simulated
-ensemble cross correlations and one contribution of the ensemble sensitivity
-kernels.\\
-
-The simulated ensemble cross correlations are obtained after the second
-simulation (Step 2). Seismograms in \texttt{OUTPUT\_FILES/} are actually
-the simulated ensemble cross correlations. Collect them immediately
-after Step 2, or the Step 3 will overwrite them. Note that we have
-a `master' receiver specified by \texttt{irec\_master\_noise}, the
-seismogram at one station corresponds to the cross correlation between
-that station and the `master'. Since the seismograms have three components,
-we may obtain cross correlations for different components as well,
-not necessarily the cross correlations between vertical components.\\
-
-One contribution of the ensemble cross-correlation sensitivity kernels
-are obtained after Step 3, stored in the \texttt{DATA/LOCAL\_PATH}
-on local nodes. The ensemble kernel files are named the same as regular
-earthquake kernels.\\
-
-You need to run another three simulations to get the other contribution
-of the ensemble kernels, using different forward and adjoint sources
-given in \citet{trompetal2010}.
-
-\section{Example}
-
-In order to illustrate noise simulations in an easy way, one example
-is provided in \texttt{EXAMPLES/noise\_tomography/}. See \texttt{EXAMPLES/noise\_tomography/README}
-for explanations. \\
-
-Note, however, that they are created for a specific workstation (CLOVER@PRINCETON),
-which has at least 4 cores with `mpif90' working properly. \\
-
-If your workstation is suitable, you can run the example in \texttt{EXAMPLES/noise\_tomography/}
-using:\\
-
-\texttt{./pre-processing.sh}\\
-
-Even if this script does not work on your workstation, the procedure
-it describes is universal. You may review the whole process described
-in the last section by following the commands in \texttt{pre-processing.sh},
-which should contain enough explanations for all the commands.
-
-%------------------------------------------------------------------------------------------------%
-
-\chapter{Gravity integral calculations for the gravity field of an Earth model}
-
-%------------------------------------------------------------------------------------------------%
-
-SPECFEM can now compute the gravity field as well as its derivatives (i.e., gravity gradiometry)
-generated by any given 3D Earth model at a given elevation, i.e., on a given observation surface.
-That feature is still experimental but should work just fine.\\
-
-To see how it is implemented and to use it, type this in the root directory of the code:\\
-
-\texttt{grep -i GRAVITY\_INTEGRALS src/*/*}
-
-%------------------------------------------------------------------------------------------------%
-
-\chapter{Graphics}
-
-%------------------------------------------------------------------------------------------------%
-
-\section{\label{sec:Mesh-graphics}Meshes}
-
-In case you used the internal mesher \texttt{xmeshfem3D} to create
-and partition your mesh, you can output mesh files in ABAQUS (.INP)
-and DX (.dx) format to visualize them. For this, you must set either
-the flag \texttt{CREATE\_DX\_FILES} or \texttt{CREATE\_ABAQUS\_FILES}
-to \texttt{.true.} in the mesher's parameter file \texttt{Mesh\_Par\_file}
-prior to running the mesher (see Chapter~\ref{cha:Running-the-Mesher-Meshfem3D}
-for details). You can then use AVS \urlwithparentheses{www.avs.com}
-or OpenDX \urlwithparentheses{www.opendx.org} to visualize the mesh
-and MPI partition (slices).
-
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=0.49\textwidth]{figures/vtk_mesh_vp} \includegraphics[width=0.49\textwidth]{figures/vtk_mesh_vs}
-\par\end{centering}
-
-\caption{Visualization using Paraview of VTK files created by \texttt{xgenerate\_databases}
-showing P- and S-wave velocities assigned to the mesh points. The
-mesh was created by \texttt{xmeshfem3D} for 4 processors.}
-
-\label{fig:vtk.mesh}
-\end{figure}
-
-You have also the option to visualize the distributed databases produced
-by \texttt{xgenerate\_databases} using Paraview \urlwithparentheses{www.paraview.org}.
-For this, you must set the flag \texttt{SAVE\_MESH\_FILES} to \texttt{.true.}
-in the main parameter file \texttt{Par\_file} (see Chapter~\ref{cha:Main-Parameter}
-for details). This will create VTK files for each single partition.
-You can then use Paraview \urlwithparentheses{www.paraview.org}
-to visualized these partitions.
-
-\section{\label{sec:Movies}Movies}
-
-To make a surface or volume movie of the simulation, set parameters
-\texttt{MOVIE\_SURFACE}, \texttt{MOVIE\_VOLUME}, \texttt{MOVIE\_TYPE},
-and \texttt{NTSTEP\_BETWEEN\_FRAMES} in the \texttt{Par\_file}. Turning
-on the movie flags, in particular \texttt{MOVIE\_VOLUME}, produces
-large output files. \texttt{MOVIE\_VOLUME} files are saved in the
-\texttt{LOCAL\_PATH} directory, whereas \texttt{MOVIE\_SURFACE} output
-files are saved in the \texttt{OUTPUT\_FILES} directory. We save the
-displacement field if the parameter \texttt{SAVE\_DISPLACEMENT} is
-set, otherwise the velocity field is saved. The look of a movie is
-determined by the half-duration of the source. The half-duration should
-be large enough so that the movie does not contain frequencies that
-are not resolved by the mesh, i.e., it should not contain numerical
-noise. This can be accomplished by selecting a CMT \texttt{HALF\_DURATION}
-> 1.1 $\times$ smallest period (see figure \ref{fig:CMTSOLUTION-file}).
-When \texttt{MOVIE\_SURFACE} = .\texttt{true.}, the half duration
-of each source in the \texttt{CMTSOLUTION} file is replaced by
-\begin{quote}
-\[
-\sqrt{(}\mathrm{\mathtt{HALF\_DURATIO}\mathtt{N}^{2}}+\mathrm{\mathtt{HDUR\_MOVI}\mathtt{E}^{2}})
-\]
- \textbf{NOTE:} If \texttt{HDUR\_MOVIE} is set to 0.0, the code will
-select the appropriate value of 1.1 $\times$ smallest period. As
-usual, for a point source one can set \texttt{half duration} in the
-\texttt{CMTSOLUTION} file to be 0.0 and \texttt{HDUR\_MOVIE} = 0.0
-in the \texttt{Par\_file} to get the highest frequencies resolved
-by the simulation, but for a finite source one would keep all the
-\texttt{half durations} as prescribed by the finite source model and
-set \texttt{HDUR\_MOVIE} = 0.0.
-\end{quote}
-
-\subsection{Movie Surface}
-
-When running \texttt{xspecfem3D} with the \texttt{MOVIE\_SURFACE}
-flag turned on, the code outputs \texttt{moviedata??????} files in
-the \texttt{OUTPUT\_FILES} directory. There are several flags in the
-main parameter file \texttt{Par\_file} that control the output of
-these movie data files (see section \ref{cha:Main-Parameter} for
-details): \texttt{\small NTSTEP\_BETWEEN\_FRAMES}{\small{} to set the
-timesteps between frames, }\texttt{\small SAVE\_DISPLACEMENT}{\small{}
-to save displacement instead of velocity, }\texttt{\small USE\_HIGHRES\_FOR\_MOVIES}{\small{}
-to save values at all GLL point instead of element edges. In order
-to output additionally shakemaps, you would set the parameter }\texttt{\small CREATE\_SHAKEMAP}{\small{}
-to }\texttt{\small .true.}{\small .}{\small \par}
-
-The files are in a fairly complicated binary format, but there is
-a program provided to convert the output into more user friendly formats:
-\begin{description}
-\item [{\texttt{xcreate\_movie\_shakemap\_AVS\_DX\_GMT}}] From \texttt{create\_movie\_shakemap\_AVS\_DX\_GMT.f90},
-it outputs data in ASCII, OpenDX, or AVS format (also readable in
-ParaView). Before compiling the code, make sure you have the file
-\texttt{surface\_from\_mesher.h} in the \texttt{OUTPUT\_FILES/} directory.
-This file will be created by the solver run. Then type
-
-\begin{lyxcode}
-{\footnotesize make~xcreate\_movie\_shakemap\_AVS\_DX\_GMT}{\footnotesize \par}
-\end{lyxcode}
-
-and run the executable \texttt{xcreate\_movie\_shakemap\_AVS\_DX\_GMT}
-in the main directory. It will create visualization files
-in your format of choice. The code will prompt the user for input
-parameters.
-
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=0.32\textwidth]{figures/movie_surf_1} \includegraphics[width=0.32\textwidth]{figures/movie_surf_2}
-\includegraphics[width=0.32\textwidth]{figures/movie_surf_3}
-\par\end{centering}
-
-\caption{Visualization using AVS files created by \texttt{xcreate\_movie\_shakemap\_AVS\_DX\_GMT}
-showing movie snapshots of vertical velocity components at different
-times.}
-
-\label{fig:movie.surf}
-\end{figure}
-
-\end{description}
-The \texttt{SPECFEM3D Cartesian} code is running in near real-time
-to produce animations of southern California earthquakes via the web;
-see Southern California ShakeMovie\textregistered{}\urlwithparentheses{www.shakemovie.caltech.edu}.
-
-\subsection{\label{sub:Movie-Volume}Movie Volume}
-
-When running xspecfem3D with the \texttt{\small MOVIE\_VOLUME}{\small{}
-flag turned on, the code outputs several files in }\texttt{\small LOCAL\_PATH}{\small{}
-specified in the main }\texttt{\small Par\_file}{\small , e.g. in
-directory }\texttt{\small OUTPUT\_FILES/DATABASES\_MPI}{\small .
-The output is saved by each processor at the time interval specified
-by }\texttt{\small NTSTEP\_BETWEEN\_FRAMES}{\small . For all domains,
-the velocity field is output to files: }\texttt{\small proc??????\_velocity\_X\_it??????.bin}{\small ,
-}\texttt{\small proc??????\_velocity\_Y\_it??????.bin}{\small{} and
-}\texttt{\small proc??????\_velocity\_Z\_it??????.bin}{\small . For
-elastic domains, the divergence and curl taken from the velocity field,
-i.e. $\nabla\cdot{\bf {v}}$ and $\nabla\times{\bf {v}}$, get stored
-as well: }\texttt{\small proc??????\_div\_it??????.bin}{\small , }\texttt{\small proc??????\_curl\_X\_t??????.bin}{\small ,
-}\texttt{\small proc??????\_curl\_Y\_it??????.bin}{\small{} and }\texttt{\small proc??????\_curl\_Z\_it??????.bin}{\small .
-The files denoted }\texttt{\small proc??????\_div\_}{\small ~}\\
-{\small{} }\texttt{\small glob\_it??????.bin}{\small{} and }\texttt{\small proc??????\_curl\_glob\_it??????.bin}{\small{}
-are stored on the global points only, all the other arrays are stored
-on all GLL points. Note that the components X/Y/Z can change to E/N/Z
-according to the }\texttt{\small SUPPRESS\_UTM\_PROJECTION}{\small{}
-flag (see also Appendix~\ref{cha:Coordinates} and \ref{cha:channel-codes}).}{\small \par}
-
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=0.32\textwidth]{figures/movie_volume_1} \includegraphics[width=0.32\textwidth]{figures/movie_volume_2}
-\includegraphics[width=0.32\textwidth]{figures/movie_volume_3}
-\par\end{centering}
-
-\caption{Paraview visualization using movie volume files (converted by \texttt{xcombine\_vol\_data}
-and \texttt{mesh2vtu.pl}) and showing snapshots of vertical velocity
-components at different times.}
-
-\label{fig:movie.volume}
-\end{figure}
-
-To visualize these files, we use an auxilliary program \texttt{combine\_vol\_data.f90}
-to combine the data from all slices into one mesh file. To compile
-it in the root directory, type:
-\begin{lyxcode}
-{\footnotesize make~xcombine\_vol\_data~}{\footnotesize \par}
-\end{lyxcode}
-which will create the executable \texttt{bin/xcombine\_vol\_data}.
-To output the usage of this executable, type
-\begin{lyxcode}
-{\footnotesize ./bin/xcombine\_vol\_data~}{\footnotesize \par}
-\end{lyxcode}
-without arguments.
-
-\medskip
-
-\texttt{xcombine\_vol\_data} will combine the data on the different processors (located in \texttt{OUTPUT\_FILES/DATABASES\_MPI}), for a given quantity and a given iteration, to one file.
-For example, if you want to combine \texttt{velocity\_Z}, for the iteration 400 on 4 processors, i.e.
-if you want to combine these files :
-
-\medskip
-
-\texttt{proc000000\_velocity\_Z\_it000400.bin }
-
-\texttt{proc000001\_velocity\_Z\_it000400.bin }
-
-\texttt{proc000002\_velocity\_Z\_it000400.bin }
-
-\texttt{proc000003\_velocity\_Z\_it000400.bin}
-
-\medskip
-
-You have to go in the directory of the concerned example (where are the directories \texttt{DATA}, \texttt{OUTPUT\_FILES}, etc.).
-Then, you can launch \texttt{xcombine\_vol\_data}, specifying the path where it is located.
-As an example, if the \texttt{DATA} and \texttt{OUTPUT\_FILES} directories of the concerned example are in the root specfem3d directory, \texttt{xcombine\_vol\_data} is in \texttt{./bin}, and you have to type:
-
-\begin{lyxcode}
-~{\footnotesize ./bin/xcombine\_vol\_data~0~3~velocity\_Z\_it000400~./OUTPUT\_FILES/DATABASES\_MPI~./OUTPUT\_FILES~0~}{\footnotesize \par}
-\end{lyxcode}
-
-Here, $0$ is the number of the first processor, $3$ the number of the last one, \texttt{velocity\_Z\_it000400} the name of the files we want to combine without the prefix \texttt{proc000000*\_}, \texttt{./OUTPUT\_FILES/DATABASES\_MPI} the directory where the files are located, \texttt{./OUTPUT\_FILES} the directory where the combined file will be stored, and $0$ is the parameter to create a low-resolution mesh file ($1$ for a high-resolution).
-
-\medskip
-
-At the beginning of the source file \texttt{combine\_vol\_data.f90} (located in \texttt{src/auxiliaries/}), there is a logical constant parameter \texttt{USE\_VTK\_OUTPUT} whose default value is \texttt{.true.}, that which will provide the output combined file directly in the VTK format and you can use Paraview to visualize it.
-
-\medskip
-
-If the value of this parameter is \texttt{.false.}, the output mesh file will have the name \texttt{velocity\_Z\_it000400.mesh}. Then we next have to convert the
-\texttt{.mesh} file into the VTU (Unstructured grid file) format which can be viewed in ParaView. For this task, you can use and modify the
-script \texttt{mesh2vtu.pl} located in directory \texttt{utils/Visualization/Paraview/}, for example:
-\begin{lyxcode}
-{\footnotesize mesh2vtu.pl~-i~velocity\_Z\_it000400.mesh~-o~velocity\_Z\_it000400.vtu}{\footnotesize \par}
-\end{lyxcode}
-Notice that this Perl script uses a program \texttt{mesh2vtu} in the \texttt{utils/Visualization/Paraview/mesh2vtu} directory, which further
-uses the VTK \urlwithparentheses{http://www.vtk.org/} run-time library for its execution. Therefore, make sure you have them properly set
-in the script according to your system.
-
-\bigskip
-
-Then, to do a movie with several iterations, you have to repeat this process for each iteration you want to put in your movie.
-
-\section{\label{sec:Finite-Frequency-Kernels}Finite-Frequency Kernels}
-
-The finite-frequency kernels computed as explained in Section \ref{sec:Adjoint-simulation-finite}
-are saved in the \texttt{LOCAL\_PATH} at the end of the simulation.
-Therefore, we first need to collect these files on the front end,
-combine them into one mesh file, and visualize them with some auxilliary
-programs.
-\begin{enumerate}
-\item \textbf{Create slice files}
-
-We will only discuss the case of one source-receiver pair, i.e., the
-so-called banana-doughnut kernels. Although it is possible to collect
-the kernel files from all slices on the front end, it usually takes
-up too much storage space (at least tens of gigabytes). Since the
-sensitivity kernels are the strongest along the source-receiver great
-circle path, it is sufficient to collect only the slices that are
-along or close to the great circle path.
-
-A Perl script \texttt{slice\_number.pl} located in directory \texttt{utils/Visualization/Paraview/}
-can help to figure out the slice numbers that lie along the great
-circle path. It applies to meshes created with the internal mesher
-\texttt{xmeshfem3D}.
-\begin{enumerate}
-\item On machines where you have access to the script, copy the \texttt{Mesh\_Par\_file},
-and \texttt{output\_solver} files, and run:
-
-\begin{lyxcode}
-{\small slice\_number.pl~Mesh\_Par\_file~output\_solver.txt~slice\_file}{\small \par}
-\end{lyxcode}
-
-which will generate a \texttt{slices\_file}.
-
-\item For cases with multiple sources and multiple receivers, you need to
-provide a slice file before proceeding to the next step.
-\end{enumerate}
-\item \textbf{Collect the kernel files}
-
-After obtaining the slice files, you can collect the corresponding
-kernel files from the given slices.
-\begin{enumerate}
-\item You can use or modify the script \texttt{utils/copy\_basin\_database.pl}
-to accomplish this:
-
-\begin{lyxcode}
-{\small utils/copy\_database.pl~slice\_file~lsf\_machine\_file~filename~{[}jobid{]}}{\small \par}
-\end{lyxcode}
-
-where \texttt{\small lsf\_machine\_file}{\small{} is the machine file
-generated by the LSF scheduler, }\texttt{\small filename}{\small{}
-is the kernel name (e.g., }\texttt{\small rho\_kernel}{\small , }\texttt{\small alpha\_kernel}{\small{}
-and }\texttt{\small beta\_kernel}{\small ), and the optional }\texttt{\small jobid}{\small{}
-is the name of the subdirectory under }\texttt{\small LOCAL\_PATH}{\small{}
-where all the kernel files are stored.}{\small \par}
-
-\item After executing this script, all the necessary mesh topology files
-as well as the kernel array files are collected to the local directory
-of the front end.
-\end{enumerate}
-\item \textbf{Combine kernel files into one mesh file}
-
-We use an auxilliary program \texttt{combine\_vol\_data.f90} to combine
-the kernel files from all slices into one mesh file.
-\begin{enumerate}
-\item Compile it in the root directory:
-
-\begin{lyxcode}
-{\footnotesize make~xcombine\_vol\_data~}{\footnotesize \par}
-
-{\footnotesize ./bin/xcombine\_vol\_data~slice\_list~filename~input\_dir~output\_dir~high/low-resolution~}{\footnotesize \par}
-\end{lyxcode}
-
-where \texttt{input\_dir} is the directory where all the individual
-kernel files are stored, and \texttt{output\_dir} is where the mesh
-file will be written.
-
-\item Use 1 for a high-resolution mesh, outputting all the GLL points to
-the mesh file, or use 0 for low resolution, outputting only the corner
-points of the elements to the mesh file.
-\item The output mesh file will have the name \texttt{filename\_rho(alpha,beta).mesh}
-\end{enumerate}
-\item \textbf{Convert mesh files into .vtu files}
-
-\begin{enumerate}
-\item We next convert the \texttt{.mesh} file into the VTU (Unstructured
-grid file) format which can be viewed in ParaView. For this task,
-you can use and modify the script \texttt{mesh2vtu.pl} located in
-directory \texttt{utils/Visualization/Paraview/}, for example:
-
-\begin{lyxcode}
-{\footnotesize mesh2vtu.pl~-i~file.mesh~-o~file.vtu~}{\footnotesize \par}
-\end{lyxcode}
-\item Notice that this Perl script uses a program \texttt{mesh2vtu} in the
-\texttt{utils/Visualization/Paraview/mesh2vtu} directory, which further
-uses the VTK \urlwithparentheses{http://www.vtk.org/} run-time library
-for its execution. Therefore, make sure you have them properly set
-in the script according to your system.
-\end{enumerate}
-\item \textbf{Copy over the source and receiver .vtk file}
-
-In the case of a single source and a single receiver, the simulation
-also generates the file \texttt{sr.vtk} located in the \texttt{OUTPUT\_FILES/}
-directory to describe the source and receiver locations, which can
-also be viewed in Paraview in the next step.
-
-\item \textbf{View the mesh in ParaView}
-
-Finally, we can view the mesh in ParaView \urlwithparentheses{www.paraview.org}.
-\begin{enumerate}
-\item Open ParaView.
-\item From the top menu, \textsf{File} $\rightarrow$\textsf{Open data},
-select \texttt{file.vtu}, and click the \textsf{Accept} button.
-
-\begin{itemize}
-\item If the mesh file is of moderate size, it shows up on the screen; otherwise,
-only the bounding box is shown.
-\end{itemize}
-\item Click \textsf{Display Tab} $\rightarrow$ \textsf{Display Style} $\rightarrow$
-\textsf{Representation} and select \textsf{wireframe of surface} to
-display it.
-\item To create a cross-section of the volumetric mesh, choose \textsf{Filter}
-$\rightarrow$ \textsf{cut}, and under \textsf{Parameters Tab}, choose
-\textsf{Cut Function} $\rightarrow$ \textsf{plane}.
-\item Fill in center and normal information given by the \texttt{global\_slice\_number.pl}
-script (either from the standard output or from \texttt{normal\_plane.txt}
-file).
-\item To change the color scale, go to \textsf{Display Tab} $\rightarrow$
-\textsf{Color} $\rightarrow$ \textsf{Edit Color Map} and reselect
-lower and upper limits, or change the color scheme.
-\item Now load in the source and receiver location file by \textsf{File}
-$\rightarrow$ \textsf{Open data}, select \texttt{sr.vt}k, and click
-the \textsf{Accept} button. Choose \textsf{Filter} $\rightarrow$
-\textsf{Glyph}, and represent the points by `\textsf{spheres}'.
-\item For more information about ParaView, see the ParaView Users Guide
-\urlwithparentheses{www.paraview.org/files/v1.6/ParaViewUsersGuide.PDF}.
-\end{enumerate}
-\end{enumerate}
-\begin{figure}[H]
-\noindent \begin{centering}
-\includegraphics[scale=0.6]{figures/3D-S-Kernel}
-\par\end{centering}
-
-\caption{(a) Top Panel: Vertical source-receiver cross-section of the S-wave
-finite-frequency sensitivity kernel $K_{\beta}$ for station GSC at
-an epicentral distance of 176 km from the September 3, 2002, Yorba
-Linda earthquake. Lower Panel: Vertical source-receiver cross-section
-of the 3D S-wave speed model used for the spectral-element simulations
-\citep{KoLiTrSuStSh04}. (b) The same as (a) but for station HEC at
-an epicentral distance of 165 km \citep{LiTr06}.}
-
-\label{figure:P-wave-speed-finite-frequency}
-\end{figure}
-
-\chapter{\label{cha:Scheduler}Running through a Scheduler}
-
-The code is usually run on large parallel machines, often PC clusters,
-most of which use schedulers, i.e., queuing or batch management systems
-to manage the running of jobs from a large number of users. The following
-considerations need to be taken into account when running on a system
-that uses a scheduler:
-\begin{itemize}
-\item The processors/nodes to be used for each run are assigned dynamically
-by the scheduler, based on availability. Therefore, in order for the
-\texttt{xgenerate\_databases} and the \texttt{xspecfem3D} executables
-(or between successive runs of the solver) to have access to the same
-database files (if they are stored on hard drives local to the nodes
-on which the code is run), they must be launched in sequence as a
-single job.
-\item On some systems, the nodes to which running jobs are assigned are
-not configured for compilation. It may therefore be necessary to pre-compile
-both the \texttt{xgenerate\_databases} and the \texttt{xspecfem3D}
-executables.
-\item One feature of schedulers/queuing systems is that they allow submission
-of multiple jobs in a ``launch and forget'' mode. In order to take
-advantage of this property, care needs to be taken that output and
-intermediate files from separate jobs do not overwrite each other,
-or otherwise interfere with other running jobs.
-\end{itemize}
-Examples of job scripts can be found in the \texttt{\small utils/Cluster/}{\small{}
-directory and can straightforwardly be modified and adapted to meet
-more specific running needs.}{\small \par}
-
-\chapter{{\normalsize \label{cha:-Changing-the}} Changing the Model}
-
-In this section we explain how to change the velocity model used for
-your simulations. These changes involve contributing specific subroutines
-that replace existing subroutines in the \texttt{SPECFEM3D Cartesian}
-package. Note that \texttt{SPECFEM3D Cartesian} can handle Earth models
-with material properties that vary within each spectral element.
-
-\section{{\normalsize \label{sec:Using-tomographic}}Using external tomographic Earth models}
-
-To implement your own external tomographic model(s), you must provide
-your own external tomography file(s), and choose between two possible
-options:\\
- \indent (1) set in the \texttt{Par\_file} the model parameter \texttt{MODEL
-= tomo}, \\
- or for more user control: \\
- \indent (2) set in the \texttt{Par\_file} the model parameter \texttt{MODEL
-= default}, define the negative \texttt{material\_ID} identifier for
-each element in the file \texttt{MESH/materials\_file} and use the
-following format in the file \texttt{MESH/nummaterial\_velocity\_file}
-when using CUBIT to construct your mesh (see also section \ref{subsec:Exporting-the-Mesh}):
-\begin{lyxcode}
-domain\_ID~material\_ID~tomography~elastic~file\_name~1
-\end{lyxcode}
-where: \\
- \indent \texttt{domain\_ID} is 1 for acoustic or 2 for elastic materials,
-\\
- \indent \texttt{material\_ID} a negative, unique identifier (i.e.,
--1,-2,...), \\
- \indent \texttt{tomography} keyword for tomographic material definition,
-\\
- \indent \texttt{elastic} keyword for elastic material definition,
-\\
- \indent \texttt{file\_name} the name of the tomography file and
-1 a positive unique identifier.\\
-
-The external tomographic model is represented by a grid of points
-with assigned material properties and homogeneous resolution along
-each spatial direction x, y and z. The ASCII file \texttt{file\_name}
-that describe the tomography should be located in the \texttt{TOMOGRAPHY\_PATH}
-directory, set in the \texttt{Par\_file}. The format of the file,
-as read from \texttt{model\_tomography.f90} located in the \texttt{src/generate\_databases}
-directory, looks like Figure \ref{fig:tomography_file}, where
-\begin{description}
-\item [{\texttt{ORIG\_X}, \texttt{END\_X}}] are, respectively, the coordinates
-of the initial and final tomographic grid points along the x direction
-(in the mesh units, e.g., $m$);
-\item [{\texttt{ORIG\_Y}, \texttt{END\_Y}}] respectively the coordinates
-of the initial and final tomographic grid points along the y direction
-(in the mesh units, e.g., $m$);
-\item [{\texttt{ORIG\_Z}, \texttt{END\_Z}}] respectively the coordinates
-of the initial and final tomographic grid points along the z direction
-(in the mesh units, e.g., $m$);
-\item [{\texttt{SPACING\_X}, \texttt{SPACING\_Y}, \texttt{SPACING\_Z}}] the
-spacing between the tomographic grid points along the x, y and z directions,
-respectively (in the mesh units, e.g., $m$);
-\item [{\texttt{NX}, \texttt{NY}, \texttt{NZ}}] the number of grid points
-along the spatial directions x, y and z, respectively; \texttt{NX}
-is given by {[}(\texttt{END\_X} - \texttt{ORIG\_X})/\texttt{SPACING\_X}{]}+1;
-\texttt{NY} and \texttt{NZ} are the same as \texttt{NX}, but for the
-y and z directions, respectively;
-\item [{\texttt{VP\_MIN}, \texttt{VP\_MAX}, \texttt{VS\_MIN}, \texttt{VS\_MAX}, \texttt{RHO\_MIN}, \texttt{RHO\_MAX}}] the
-minimum and maximum values of the wave speed \texttt{vp} and \texttt{vs}
-(in $m\, s^{-1}$) and of the density \texttt{rho} (in $kg\, m^{-3}$);
-these values could be the actual limits of the tomographic parameters
-in the grid or the minimum and maximum values to which we force the
-cut of velocity and density in the model.
-\end{description}
-After the first four lines, in the file \texttt{file\_name} the tomographic
-grid points are listed with the corresponding values of \texttt{vp},
-\texttt{vs} and \texttt{rho}, scanning the grid along the x coordinate
-(from \texttt{ORIG\_X} to \texttt{END\_X} with step of \texttt{SPACING\_X})
-for each given y (from \texttt{ORIG\_Y} to \texttt{END\_Y}, with step
-of \texttt{SPACING\_Y}) and each given z (from \texttt{ORIG\_Z} to
-\texttt{END\_Z}, with step of \texttt{SPACING\_Z}). %fig
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=1\textwidth]{figures/tomo_file}
-\par\end{centering}
-
-\caption{Tomography file \texttt{file\_name} that describes an external Earth
-model. The coordinates x, y and z of the grid, their limits ORIG\_X,
-ORIG\_Y, ORIG\_Z, END\_X, END\_Y, END\_Z and the grid spacings SPACING\_X,
-SPACING\_Y, SPACING\_Z should be in the units of the constructed mesh
-(e.g., UTM coordinates in meters).}
-
-\label{fig:tomography_file}
-\end{figure}
-
-The user can implement his own interpolation algorithm for the tomography
-model by changing the routine \texttt{model\_tomography.f90} located
-in the \texttt{src/generate\_databases/} directory. Moreover, for
-models that involve both fully defined materials and a tomography
-description, the \texttt{nummaterial\_velocity\_file} has multiple
-lines each with the corresponding suitable format described above.
-
-%==============
-
-\subsubsection*{Example: External tomography file with variable discretization intervals}
-
-The example in Figure~\ref{fig:tomography_file} is for an external tomography file with uniform spacing in the $x$, $y$, and $z$ directions.
-(Note that $\Delta x$, $\Delta y$, and $\Delta z$ need not be the same as in the example.) In that example, the  \texttt{nummaterial\_velocity\_file} is
-%
-\begin{lyxcode}
-2  -1 tomography elastic tomography\_model.xyz 1
-\end{lyxcode}
-%
-and the file \texttt{tomography\_model.xyz} will need to reside in \texttt{DATA/}.
-All entries in the second column of \texttt{materials\_file} will be \texttt{-1}, which means that each element in the mesh will be interpolated accordind to the values in \texttt{tomography\_model.xyz}.
-
-In some cases it may be desirable to have an external tomography model that is described in more than one file.
-For example, in cases like southern California, the length scale of variation in the structure of the wave speed model is much shorter in the sedimentary basin models within the upper 15~km.
-Therefore one might want to use an external tomography file that is sampled with, say, $\Delta x = 1000$~m, $\Delta y = 1000$~m, and $\Delta z = 250$~m in the uppermost 15~km, and then use $\Delta x = 2000$~m, $\Delta y = 2000$~m, and $\Delta z = 1000$~m below a depth of 15~km.
-If these intervals are chosen appropriately, then it will result in a pair of external tomography file that is much smaller than the alternative of having a single file with the fine discretization.
-In this case  \texttt{nummaterial\_velocity\_file} is
-%
-\begin{lyxcode}
-2  -1 tomography elastic file\_above\_15km.xyz 1 \\
-2  -2 tomography elastic file\_below\_15km.xyz 1
-\end{lyxcode}
-%
-and the files \texttt{file\_above\_15km.xyz} and \texttt{file\_below\_15km.xyz} will need to reside in \texttt{DATA/}.
-All entries in the second column of \texttt{materials\_file} will need to be either \texttt{-1} or \texttt{-2}, depending on whether the element is above or below 15~km depth.
-In this sense, the construction of the mesh and the discretization of the external model will generally be done in tandem.
-
-Other more complicated discretizations can be used following the same procedure. (In particular, there is no limit on the number of different external files that are used.)
-
-%=============
-
-\section{{\normalsize \label{sec:Anelastic-Models}}External (an)elastic Models}
-
-To use your own external model, you can set in the \texttt{Par\_file}
-the model parameter \texttt{MODEL = external}. Three-dimensional acoustic
-and/or (an)elastic (attenuation) models may then be superimposed onto
-the mesh based upon your subroutine in \texttt{model\_external\_values.f90}
-located in directory \texttt{src/generate\_databases}. The call to
-this routine would be as follows
-\begin{lyxcode}
-call~model\_external\_values(xmesh,~ymesh,~zmesh,~~\&~~\\
-~~~~~~~~~~~rho,~vp,~vs,~qmu\_atten,~iflag\_aniso,~idomain\_id)
-\end{lyxcode}
-Input to this routine consists of:
-\begin{description}
-\item [{\texttt{xmesh,ymesh,zmesh}}] location of mesh point
-\end{description}
-Output to this routine consists of:
-\begin{description}
-\item [{\texttt{rho,vp,vs}}] isotropic model parameters for density $\rho$
-($kg/m^{3}$), $v_{p}$ ($m/s$) and $v_{s}$ ($m/s$)
-\item [{\texttt{qmu\_atten}}] Shear wave quality factor: $0<Q_{\mu}<9000$
-\item [{\texttt{iflag\_aniso}}] anisotropic model flag, $0$ indicating
-no anisotropy or $1$ using anisotropic model parameters as defined
-in routine file \texttt{model\_aniso.f90}
-\item [{\texttt{idomain\_id}}] domain identifier, $1$ for acoustic, $2$
-for elastic, $3$ for poroelastic materials.
-\end{description}
-Note that the resolution and maximum value of anelastic models are
-truncated. This speeds the construction of the standard linear solids
-during the meshing stage. To change the resolution, currently at one
-significant figure following the decimal, or the maximum value (9000),
-consult \texttt{shared/constants.h}.
-
-\section{{\normalsize \label{sec:Anisotropic-Models}}Anisotropic Models}
-
-To use your anisotropic models, you can either set in the \texttt{Par\_file}
-the model parameter \texttt{ANISOTROPY} to \texttt{.true.} or set
-the model parameter \texttt{MODEL = aniso}. Three-dimensional anisotropic
-models may be superimposed on the mesh based upon the subroutine in
-file \texttt{model\_aniso.f90} located in directory \texttt{src/generate\_databases}.
-The call to this subroutine is of the form
-\begin{lyxcode}
-call~model\_aniso(iflag\_aniso,~rho,~vp,~vs,~\&~~\\
-~~~~~~~~~~~c11,c12,c13,c14,c15,c16,c22,c23,c24,c25,c26,~\&~~\\
-~~~~~~~~~~~c33,c34,c35,c36,c44,c45,c46,c55,c56,c66)~
-\end{lyxcode}
-Input to this routine consists of:
-\begin{description}
-\item [{\texttt{iflag\_aniso}}] flag indicating the type of the anisotropic
-model, i.e. $0$ for no superimposed anisotropy, $1$ or $2$ for
-generic pre-defined anisotropic models.
-\item [{\texttt{rho,vp,vs}}] reference isotropic model parameters for density
-$\rho$, $v_{p}$ and $v_{s}$.
-\end{description}
-Output from the routine consists of the following non-dimensional
-model parameters:
-\begin{description}
-\item [{\texttt{c11},}] \textbf{$\cdots$,} \texttt{\textbf{c66}} 21~dimensionalized
-anisotropic elastic parameters.
-\end{description}
-You can replace the \texttt{model\_aniso.f90} file by your own version
-\textit{provided you do not change the call structure of the routine},
-i.e., the new routine should take exactly the same input and produce
-the required relative output.
-
-\section{{\normalsize \label{sec:Using-SEP}}Using external SEP models.}
-
-Stanford Exploration Project (SEP) models consists of two files per variables, one header file (\texttt{.H})
-and one binary file (\texttt{.H$@$}).
-Among other information, the header file indicates, for each dimension,
-the offset (\texttt{o1}, \texttt{o2} and \texttt{o3} fields),
-the spatial step size (\texttt{d1}, \texttt{d2} and \texttt{d3} fields)
-and the number of samples (\texttt{n1}, \texttt{n2} and \texttt{n3}) necessary to read
-the linked binary file (\texttt{in} field).
-
-SEP model reading occurs during database generation. It it the user responsibility
-to ensure that proper interfaces between acoustic and elastic domains are set up
-during the meshing phase. For more information, refer to
-section~\ref{cha:Running-the-Mesher-Meshfem3D}.
-
-In \texttt{DATA/Par\_file}, \texttt{\textbf{MODEL}} should be set to \texttt{\textbf{sep}},
-and \texttt{\textbf{SEP\_MODEL\_DIRECTORY}} should point to wherever your sep model files reside.
-
-Note that in order not to overload the parameter file, SEP header files
-should be named \texttt{vp.H}, \texttt{vs.H} and \texttt{rho.H}. The \texttt{in} field
-in these binary files can be any path as long as it is relative to
-the previously set \texttt{SEP\_MODEL\_DIRECTORY}.
-We also assume that any dimensional value inside SEP header files is given in meters and
-that binary files are single precision floating point numbers stored in little endian format.
-There is currently only support for \texttt{vp}, \texttt{vs} and \texttt{rho} variables.
-
-An example demonstrating the use SEP models is given in:
-\texttt{EXAMPLES/meshfem3D\_examples/sep\_bathymetry}.
-
-\chapter{Post-Processing Scripts}
-
-Several post-processing scripts/programs are provided in the \texttt{utils/}
-directory, and most of them need to be adjusted when used on different
-systems, for example, the path of the executable programs. Here we
-only list a few of the available scripts and provide a brief description,
-and you can either refer to the related sections for detailed usage
-or, in a lot of cases, type the script/program name without arguments
-for its usage.
-
-\section{Process Data and Synthetics\label{sec:Process-data-and-syn}}
-
-In many cases, the SEM synthetics are calculated and compared to data
-seismograms recorded at seismic stations. Since the SEM synthetics
-are accurate for a certain frequency range, both the original data
-and the synthetics need to be processed before a comparison can be
-made.
-
-For such comparisons, the following steps are recommended:
-\begin{enumerate}
-\item Make sure that both synthetic and observed seismograms have the correct
-station/event and timing information.
-\item Convolve synthetic seismograms with a source time function with the
-half duration specified in the \texttt{CMTSOLUTION} file, provided,
-as recommended, you used a zero half duration in the SEM simulations.
-\item Resample both observed and synthetic seismograms to a common sampling
-rate.
-\item Cut the records using the same window.
-\item Remove the trend and mean from the records and taper them.
-\item Remove the instrument response from the observed seismograms (recommended)
-or convolve the synthetic seismograms with the instrument response.
-\item Make sure that you apply the same filters to both observed and synthetic
-seismograms. Preferably, avoid filtering your records more than once.
-\item Now, you are ready to compare your synthetic and observed seismograms.
-\end{enumerate}
-\noindent We generally use the following scripts provided in the \texttt{utils/seis\_process/}
-directory:
-
-\subsection{Data processing script \texttt{process\_data.pl}}
-
-This script cuts a given portion of the original data, filters it,
-transfers the data into a displacement record, and picks the first
-P and S arrivals. For more functionality, type `\texttt{process\_data.pl}'
-without any argument. An example of the usage of the script:
-\begin{lyxcode}
-{\footnotesize process\_data.pl~-m~CMTSOLUTION~-s~1.0~-l~0/4000~-i~-f~-t~40/500~-p~-x~bp~DATA/1999.330{*}.BH?.SAC}~
-\end{lyxcode}
-\noindent which has resampled the SAC files to a sampling rate of
-1 seconds, cut them between 0 and 4000 seconds, transfered them into
-displacement records and filtered them between 40 and 500 seconds,
-picked the first P and S arrivals, and added suffix `\texttt{bp}'
-to the file names.
-
-Note that all of the scripts in this section actually use the SAC
-and/or IASP91 to do the core operations; therefore make sure that
-the SAC and IASP91 packages are installed properly on your system,
-and that all the environment variables are set properly before running
-these scripts.
-
-\subsection{Synthetics processing script \texttt{process\_syn.pl}}
-
-This script converts the synthetic output from the SEM code from ASCII
-to SAC format, and performs similar operations as `\texttt{process\_data.pl}'.
-An example of the usage of the script:
-\begin{lyxcode}
-{\footnotesize process\_syn.pl~-m~CMTSOLUTION~-h~-a~STATIONS~-s~1.0~-l~0/4000~-f~-t~40/500~-p~-x~bp~SEM/{*}.BX?.semd}~
-\end{lyxcode}
-which will convolve the synthetics with a triangular source-time function
-from the \texttt{CMTSOLUTION} file, convert the synthetics into SAC
-format, add event and station information into the SAC headers, resample
-the SAC files with a sampling rate of 1 seconds, cut them between
-0 and 4000 seconds, filter them between 40 and 500 seconds with the
-same filter used for the observed data, pick the first P and S arrivals,
-and add the suffix `\texttt{bp}' to the file names.
-
-More options are available for this script, such as adding time shift
-to the origin time of the synthetics, convolving the synthetics with
-a triangular source time function with a given half duration, etc.
-Type \texttt{process\_syn.pl} without any argument for a detailed
-usage.
-
-In order to convert between SAC format and ASCII files, useful scripts
-are provided in the subdirectories ~\\
- \texttt{utils/sac2000\_alpha\_convert/} and \texttt{utils/seis\_process/asc2sac/}.
-
-\subsection{Script \texttt{rotate.pl}}
-
-The original data and synthetics have three components: vertical (BHZ
-resp. BXZ), north (BHN resp. BXN) and east (BHE resp. BXE). However,
-for most seismology applications, transverse and radial components
-are also desirable. Therefore, we need to rotate the horizontal components
-of both the data and the synthetics to the transverse and radial direction,
-and \texttt{\small rotate.pl}{\small{} can be used to accomplish this:}{\small \par}
-\begin{lyxcode}
-rotate.pl~-l~0~-L~180~-d~DATA/{*}.BHE.SAC.bp~
-
-rotate.pl~-l~0~-L~180~SEM/{*}.BXE.semd.sac.bp~
-\end{lyxcode}
-where the first command performs rotation on the SAC data obtained
-through Seismogram Transfer Program (STP) \urlwithparentheses{http://www.data.scec.org/STP/stp.html},
-while the second command rotates the processed SEM synthetics.
-
-\section{Collect Synthetic Seismograms}
-
-The forward and adjoint simulations generate synthetic seismograms
-in the \texttt{OUTPUT\_FILES/} directory by default. For the forward
-simulation, the files are named like \texttt{NT.STA.BX?.semd} for
-two-column time series, or \texttt{NT.STA.BX?.semd.sac} for ASCII
-SAC format, where NT and STA are the network code and station name,
-and \texttt{BX?} stands for the component name. Please see the Appendix~\ref{cha:Coordinates}
-and \ref{cha:channel-codes} for further details.
-
-The adjont simulations generate synthetic seismograms with the name
-\texttt{NT.S?????.S??.sem} (refer to Section \ref{sec:Adjoint-simulation-sources}
-for details). The kernel simulations output the back-reconstructed
-synthetic seismogram in the name \texttt{NT.STA.BX?.semd}, mainly
-for the purpose of checking the accuracy of the reconstruction. Refer
-to Section \ref{sec:Adjoint-simulation-finite} for further details.
-
-You do have further options to change this default output behavior,
-given in the main constants file \texttt{constants.h} located in \texttt{src/shared/}
-directory:
-\begin{description}
-\item [{\texttt{SEISMOGRAMS\_BINARY}}] set to \texttt{.true.} to have seismograms
-written out in binary format.
-\item [{\texttt{WRITE\_SEISMOGRAMS\_BY\_MASTER}}] Set to \texttt{.true.}
-to have only the master process writing out seismograms. This can
-be useful on a cluster, where only the master process node has access
-to the output directory.
-\item [{\texttt{USE\_OUTPUT\_FILES\_PATH}}] Set to \texttt{.false.} to
-have the seismograms output to \texttt{LOCAL\_PATH} directory specified
-in the main parameter file \texttt{DATA/Par\_file}. In this case,
-you could collect the synthetics onto the frontend using the \texttt{collect\_seismo\_lsf\_multi.pl}
-script located in the \texttt{utils/Cluster/lsf/} directory. The usage
-of the script would be e.g.:
-
-\begin{lyxcode}
-collect\_seismo.pl~machines~DATA/Par\_file
-\end{lyxcode}
-\end{description}
-where \texttt{machines} is a file containing the node names and \texttt{DATA/Par\_file}
-the parameter file used to extract the \texttt{LOCAL\_PATH} directory
-used for the simulation.
-
-\section{Clean Local Database}
-
-After all the simulations are done, the seismograms are collected,
-and the useful database files are copied to the frontend, you may
-need to clean the local scratch disk for the next simulation. This
-is especially important in the case of kernel simulation, where very
-large files are generated for the absorbing boundaries to help with
-the reconstruction of the regular forward wavefield. A sample script
-is provided in \texttt{utils/}:
-\begin{lyxcode}
-cleanbase.pl~machines
-\end{lyxcode}
-where \texttt{machines} is a file containing the node names.
-
-\section{Plot Movie Snapshots and Synthetic Shakemaps}
-
-\subsection{Script \texttt{movie2gif.gmt.pl}}
-
-With the movie data saved in \texttt{OUTPUT\_FILES/} at the end of
-a movie simulation (\texttt{\small MOVIE\_SURFACE=.true.}{\small ),
-you can run the }\texttt{\small `create\_movie\_shakemap\_AVS\_DX\_GMT}{\small '
-code to convert these binary movie data into GMT xyz files for futher
-processing. A sample script }\texttt{\small movie2gif.gmt.pl}{\small{}
-is provided to do this conversion, and then plot the movie snapshots
-in GMT, for example:}{\small \par}
-\begin{lyxcode}
-movie2gif.gmt.pl~-m~CMTSOLUTION~-g~-f~1/40~-n~-2~-p~
-\end{lyxcode}
-which for the first through the 40th movie frame, converts the \texttt{moviedata}
-files into GMT xyz files, interpolates them using the 'nearneighbor'
-command in GMT, and plots them on a 2D topography map. Note that `\texttt{-2}'
-and `\texttt{-p}' are both optional.
-
-\subsection{Script \texttt{plot\_shakemap.gmt.pl}}
-
-With the shakemap data saved in \texttt{OUTPUT\_FILES/} at the end
-of a shakemap simulation ~\\
- (\texttt{CREATE\_SHAKEMAP=.true.}), you can also run \texttt{`create\_movie\_shakemap\_AVS\_DX\_GMT}'
-code to convert the binary shakemap data into GMT xyz files. A sample
-script \texttt{plot\_shakemap.gmt.pl} is provided to do this conversion,
-and then plot the shakemaps in GMT, for example:
-\begin{lyxcode}
-plot\_shakemap.gmt.pl~data\_dir~type(1,2,3)~CMTSOLUTION~
-\end{lyxcode}
-where \texttt{type=1} for a displacement shakemap, \texttt{2} for
-velocity, and \texttt{3} for acceleration.
-
-\section{Map Local Database}
-
-A sample program \texttt{remap\_database} is provided to map the local
-database from a set of machines to another set of machines. This is
-especially useful when you want to run mesher and solver, or different
-types of solvers separately through a scheduler (refer to Chapter~\ref{cha:Scheduler}).
-\begin{lyxcode}
-run\_lsf.bash~-{}-gm-no-shmem~-{}-gm-copy-env~remap\_database~old\_machines~150
-\end{lyxcode}
-where \texttt{old\_machines} is the LSF machine file used in the previous
-simulation, and \texttt{150} is the number of processors in total.
-
-%============================================================
-
-\chapter{\label{cha:developers}Information for developers of the code, and for people who want to learn how the technique works}
-
-You can get a very simple 1D version of a demo code (there is one in Fortran and one in Python):\\
-
-git clone --recursive https://github.com/geodynamics/specfem1d.git\\
-
-\noindent We also have simple 3D demo source codes that implement the SEM in a single, small program, in directory\\
-utils/small\_SEM\_solvers\_in\_Fortran\_and\_C\_without\_MPI\_to\_learn of the specfem3d package.
-They are useful to learn how the spectral-element method works, and how to write or modify a code to implement it.
-Also useful to test new ideas by modifying these simple codes to run some tests.\\
-
-\noindent For information on how to contribute to the code, i.e., for how to make your modifications, additions or improvements part of the
-official package, see https://github.com/geodynamics/specfem3d/wiki .
-
-%============================================================
-
-\chapter*{\label{cha:Bug-Reports-and}Bug Reports and Suggestions for Improvements}
-
-To report bugs or suggest improvements to the code, please send an
-e-mail to the CIG Computational Seismology Mailing List \urlwithparentheses{cig-seismo@geodynamics.org}.
-
-\addcontentsline{toc}{chapter}{Notes \& Acknowledgments}
-
-\chapter*{Notes \& Acknowledgments}
-
-In order to keep the software package thread-safe in case a multithreaded
-implementation of MPI is used, developers should not add modules or
-common blocks to the source code but rather use regular subroutine
-arguments (which can be grouped in ``derived types'' if needed for
-clarity).
-
-The Gauss-Lobatto-Legendre subroutines in \texttt{gll\_library.f90}
-are based in part on software libraries from the Massachusetts Institute
-of Technology, Department of Mechanical Engineering (Cambridge, Massachusetts,
-USA). The non-structured global numbering software was provided by
-Paul F. Fischer (Brown University, Providence, Rhode Island, USA,
-now at Argonne National Laboratory, USA).
-
-OpenDX \urlwithparentheses{http://www.opendx.org} is open-source
-based on IBM Data Explorer, AVS \urlwithparentheses{http://www.avs.com}
-is a trademark of Advanced Visualization Systems, and ParaView \urlwithparentheses{http://www.paraview.org}
-is an open-source visualization platform.
-
-Please e-mail your feedback, questions, comments, and suggestions
-to the CIG Computational Seismology Mailing List \urlwithparentheses{cig-seismo@geodynamics.org}.
-
-\addcontentsline{toc}{chapter}{Copyright}
-
-\chapter*{Copyright}
-
-Main historical authors: Dimitri Komatitsch and Jeroen Tromp
-
-Princeton University, USA, and CNRS / University of Marseille, France
-
-$\copyright$ Princeton University and CNRS / University of Marseille, July 2012
-
-This program is free software; you can redistribute it and/or modify
-it under the terms of the GNU General Public License as published
-by the Free Software Foundation (see Appendix \ref{cha:License}).\\
-
-\textbf{\underline{Evolution of the code:}}\\
-
-MPI v. 3.0, December 2014: many developers.
-Convolutional PML, LDDRK time scheme, bulk attenuation support, simultaneous MPI runs,
-ADIOS file I/O support, coupling with external codes, new seismogram names,
-Deville routines for additional GLL degrees, tomography tools, unit/regression test framework,
-improved CUDA GPUs performance, additonal GEOCUBIT support, better make compilation,
-git versioning system. \\
-
-MPI v. 2.1, July 2012: Max Rietmann, Peter Messmer, Daniel Peter, Dimitri
-Komatitsch, Joseph Charles, Zhinan Xie: support for CUDA GPUs, better
-CFL stability for the Stacey absorbing conditions. \\
-
-MPI v. 2.0, November 2010: Dimitri Komatitsch, Nicolas Le Goff, Roland
-Martin and Pieyre Le Loher, University of Pau, France, Daniel Peter,
-Jeroen Tromp and the Princeton group of developers, Princeton University,
-USA, and Emanuele Casarotti, INGV Roma, Italy: support for CUBIT meshes
-decomposed by SCOTCH; much faster solver using Michel Deville's inlined
-matrix products.\\
-
-MPI v. 1.4 Dimitri Komatitsch, University of Pau, Qinya Liu and others,
-Caltech, September 2006: better adjoint and kernel calculations, faster
-and better I/Os on very large systems, many small improvements and
-bug fixes.\\
-
-MPI v. 1.3 Dimitri Komatitsch, University of Pau, and Qinya Liu, Caltech,
-July 2005: serial version, regular mesh, adjoint and kernel calculations,
-ParaView support.\\
-
-MPI v. 1.2 Min Chen and Dimitri Komatitsch, Caltech, July 2004: full
-anisotropy, volume movie.\\
-
-MPI v. 1.1 Dimitri Komatitsch, Caltech, October 2002: Zhu's Moho map,
-scaling of $V_{s}$ with depth, Hauksson's regional model, attenuation,
-oceans, movies.\\
-
-MPI v. 1.0 Dimitri Komatitsch, Caltech, USA, May 2002: first MPI version
-based on global code.\\
-
-Dimitri Komatitsch, IPG Paris, France, December 1996: first 3-D solver
-for the CM-5 Connection Machine, parallelized on 128 processors using
-Connection Machine Fortran.\\
-
-\addcontentsline{toc}{chapter}{References} \bibliography{bibliography}
-
-\appendix
-%dummy comment inserted by tex2lyx to ensure that this paragraph is not empty
-
-\chapter{\label{cha:Coordinates}Reference Frame Convention}
+\chapter{Reference Frame Convention}\label{cha:Coordinates}
 
@@
 \end{description}
-
-\chapter{\label{cha:channel-codes}Channel Codes of Seismograms}
-
-Seismic networks, such as the Global Seismographic Network (GSN),
-generally involve various types of instruments with different bandwidths,
-sampling properties and component configurations. There are standards
-to name channel codes depending on instrument properties. IRIS \texttt{(www.iris.edu)}
-uses SEED format for channel codes, which are represented by three
-letters, such as \texttt{LHN}, \texttt{BHZ}, etc. In older versions
-of the SPECFEM3D Cartesian package, a common format was used for the
-channel codes of all seismograms, which was \texttt{BHE/BHN/BHZ} for
-three components. To avoid confusion when comparison are made to observed
-data, we are now using the FDSN convention (\url{http://www.fdsn.org/})
-for SEM seismograms. In the following, we give a brief explanation
-of the FDSN convention used by IRIS, and how it is adopted in SEM
-seismograms. Please visit \texttt{www.iris.edu/manuals/SEED\_appA.htm}
-for further information.\\
-
-\noindent \texttt{\textbf{Band code:}} The first letter in the channel
-code denotes the band code of seismograms, which depends on the response
-band and the sampling rate of instruments. The list of band codes
-used by IRIS is shown in Figure \ref{fig:IRIS_band_codes}. The sampling
-rate of SEM synthetics is controlled by the resolution of simulations
-rather than instrument properties. However, for consistency, we follow
-the FDSN convention for SEM seismograms governed by their sampling
-rate. For SEM synthetics, we consider band codes for which $dt\leq1$
-s. IRIS also considers the response band of instruments. For instance,
-short-period and broad-band seismograms with the same sampling rate
-correspond to different band codes, such as S and B, respectively.
-In such cases, we consider SEM seismograms as broad band, ignoring
-the corner period ($\geq10$ s) of the response band of instruments
-(note that at these resolutions, the minimum period in the SEM synthetics
-will be less than $10$ s). Accordingly, when you run a simulation
-the band code will be chosen depending on the resolution of the synthetics,
-and channel codes of SEM seismograms will start with either \texttt{L},
-\texttt{M}, \texttt{B}, \texttt{H}, \texttt{C} or \texttt{F}, shown
-by red color in the figure.\\
-
-\begin{figure}[ht]
-\noindent \begin{centering}
-\includegraphics[scale=0.6]{figures/IRIS_band_codes}
-\par\end{centering}
-
-\caption{The band code convention is based on the sampling rate and the response
-band of instruments. Please visit \texttt{www.iris.edu/manuals/SEED\_appA.htm}
-for further information. Grey rows show the relative band-code range
-in SPECFEM3D Cartesian, and the band codes used to name SEM seismograms
-are denoted in red.}
-
-\label{fig:IRIS_band_codes}
-\end{figure}
-
-\noindent \texttt{\textbf{Instrument code:}} The second letter in
-the channel code corresponds to instrument codes, which specify the
-family to which the sensor belongs. For instance, \texttt{H} and \texttt{L}
-are used for high-gain and low-gain seismometers, respectively. The
-instrument code of SEM seismograms will always be \texttt{X}, as assigned
-by FDSN for synthetic seismograms. \\
-
-\noindent \texttt{\textbf{Orientation code:}} The third letter in
-channel codes is an orientation code, which generally describes the
-physical configuration of the components of instrument packages. SPECFEM3D
-Cartesian uses the traditional orientation code \texttt{E/N/Z} (East-West,
-North-South, Vertical) for three components when a UTM projection
-is used. If the UTM conversion is suppressed, i.e. the flag \texttt{SUPPRESS\_UTM\_PROJECTION}
-is set to \texttt{.true.}, then the three components are labelled
-\texttt{X/Y/Z} according to the Cartesian reference frame. \\
-
-\noindent \texttt{\textbf{EXAMPLE:}} The sampling rate is given by
-\texttt{DT} in the main parameter file \texttt{Par\_file} located
-in the \texttt{DATA/} subdirectory. Depending on the resolution of
-your simulations, if you choose a sampling rate greater than $0.01$
-s and less than $1$ s, a seismogram recording displacements on the
-vertical component of a station \texttt{ASBS} (network \texttt{AZ})
-will be named \texttt{AZ.ASBS.MXZ.semd.sac}, whereas it will be \texttt{AZ.ASBS.BXZ.semd.sac},
-if the sampling rate is greater than 0.0125 and less equal to 0.1
-s.
-
-\chapter{\label{cha:Troubleshooting}Troubleshooting}
-
-\section*{FAQ}
-\begin{description}
-\item [{configuration fails:}]~
-
-Examine the log file 'config.log'. It contains detailed informations.
-In many cases, the paths to these specific compiler commands F90,
-CC and MPIF90 will not be correct if `./configure` fails.
-
-Please make sure that you have a working installation of a Fortran
-compiler, a C compiler and an MPI implementation. You should be able
-to compile this little program code:
-\begin{lyxcode}
-program~main~~~\\
-~include~'mpif.h'~~~\\
-~integer,~parameter~::~CUSTOM\_MPI\_TYPE~=~MPI\_REAL~~~\\
-~integer~ier~~~\\
-~call~MPI\_INIT(ier)~~~\\
-~call~MPI\_BARRIER(MPI\_COMM\_WORLD,ier)~~~\\
-~call~MPI\_FINALIZE(ier)~~~\\
-~end
-\end{lyxcode}
-\item [{compilation fails stating:}] ~\\
-
-\begin{lyxcode}
-...~~~\\
-~obj/program\_generate\_databases.o:~In~function~`MAIN\_\_':~~~\\
-~program\_generate\_databases.f90:(.text+0x14):~undefined~reference~to~`\_gfortran\_set\_std'~~~\\
-~...~~~\\
-
-\end{lyxcode}
-
-Make sure you're pointing to the right 'mpif90' wrapper command.
-
-Normally, this message will appear when you are mixing two different
-Fortran compilers. That is, using e.g. gfortran to compile non-MPI
-files and mpif90, wrapper provided for e.g. ifort, to compile MPI-files.
-
-fix: e.g. specify > ./configure FC=gfortran MPIF90=/usr/local/openmpi-gfortran/bin/mpif90
-
-\item [{after executing \texttt{xmeshfem3D} I've got elements with skewness of 81\% percent, what does this mean:}] Look
-at the skewness table printed in the \texttt{output\_mesher.txt} file
-after executing \texttt{xmeshfem3D} for the example given in \texttt{EXAMPLES/meshfem3D\_examples/simple\_model/}:
-
-\begin{lyxcode}
-...~~\\
-~histogram~of~skewness~(0.~good~-~1.~bad):~~~\\
-~0.0000000E+00~-~5.0000001E-02~27648~81.81818~\%~~~\\
-~5.0000001E-02~-~0.1000000~0~0.0000000E+00~\%~~~\\
-~...~~~\\
-
-\end{lyxcode}
-
-The first line means that you have 27,648 elements with a skewness
-value between 0 and 0.05 (which means the element is basically not
-skewed, just plain regular hexahedral element). The total number of
-elements you have in this mesh is (see in the \texttt{output\_mesher.txt}
-file a bit further down):
-\begin{lyxcode}
-...~~\\
-~total~number~of~elements~in~entire~mesh:~33792~~~\\
-~...~~~\\
-
-\end{lyxcode}
-
-which gives you that: 27,648 / 33,792 $\sim$ 81.8 \% of all elements
-are not skewed, i.e. regular elements. a fantastic value :)
-
-The histogram lists for this mesh also some stronger skewed elements,
-for example the worst ones belong to:
-\begin{lyxcode}
-...~~\\
-~0.6000000~-~0.6500000~2048~6.060606~\%~~~\\
-~...~~~\\
-
-\end{lyxcode}
-
-about 6 \% of all elements have distortions with a skewness value
-between 0.6 and 0.65. The skewness values give you a hint of how good
-your mesh is. In an ideal world, you would want to have no distortions,
-just like the 81\% from above. Those elements give you the best approximate
-values by the GLL quadrature used in the spectral-element method.
-However, having weakly distorted elements is still fine and the solutions
-are still accurate enough. So empirically, values up to around 0.7
-are tolerable, above that you should consider remeshing...
-
-To give you an idea why some of the elements are distorted, see the
-following figure \ref{fig:mesh.vp} of the mesh you obtain in the
-example \texttt{EXAMPLES/meshfem3D\_examples/simple\_model/}.
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=0.9\textwidth]{figures/mesh_vp}
-\par\end{centering}
-
-\caption{Paraview visualization using the mesh vtk-files for the example given
-in \texttt{EXAMPLES/meshfem3D\_examples/simple\_model/}.}
-
-\label{fig:mesh.vp}
-\end{figure}
-
-You will see that the mesh contains a doubling layer, where we stitch
-elements together such that the size of two elements will transition
-to the size of one element (very useful to keep the ratio of wavespeed
-/ element\_size about constant). Those elements in this doubling layer
-have higher skewness values and make up those 6 \% in the histogram.
-
-\item [{the code gives following error message "need at least one receiver":}] This
-means that no stations given in the input file \texttt{DATA/STATIONS}
-could be located within the dimensions of the mesh. This can happen
-for example when the mesh was created with the in-house mesher \texttt{xmeshfem3D}
-while using the Universal Transverse Mercator (UTM) projection but
-the simulation with \texttt{xspecfem3D} was suppressing this projection
-from latitude/longitude to x/y/z coordinates.
-
-In such cases, try to change your \texttt{DATA/Par\_file} and set
-e.g.:
-\begin{lyxcode}
-SUPPRESS\_UTM\_PROJECTION~=~.false.~~~\\
-
-\end{lyxcode}
-
-to be the same in \texttt{Mesh\_Par\_file} and \texttt{Par\_file}.
-This flag should be identical when using the in-house mesher \texttt{xmeshfem3D},
-\texttt{xgenerate\_databases} and \texttt{xspecfem3D} together to
-run simulations.
-
-The flag determines if the coordinates you specify for your source
-and station locations are given as lat/lon degrees and must be converted
-to UTM coordinates. As an example, if you use \texttt{.false.} within
-\texttt{Mesh\_Par\_file} then you create a mesh with \texttt{xmeshfem3D}
-using the UTM projection from lat/lon as input format to UTM projected
-coordinates to store the mesh point positions, which is fine. The
-error then may occur if in the \texttt{Par\_file} you have this set
-to \texttt{.true.} so that the \texttt{xgenerate\_databases} and \texttt{xspecfem3D}
-suppress the UTM projection and assume that all coordinates you use
-now for source and receiver locations are given in meters (that is,
-converted) already. So it won't find the specified locations in the
-used mesh. As a solutions, just change the flag in \texttt{Par\_file}
-to be the same as in \texttt{Mesh\_Par\_file} and rerun \texttt{xgenerate\_databases}
-and \texttt{xspecfem3D} to make sure that your simulation works fine.
-
-\item [{I get the following error message "forward simulation became unstable and blew up":}] In
-most cases this means that your time step size \texttt{DT} is chosen
-too big. Look at your files \texttt{output\_mesher.txt} or \texttt{output\_solver.txt}
-created in the folder \texttt{OUTPUT\_FILES}. In these output files,
-find the section:
-
-\begin{lyxcode}
-...~~~\\
-~{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}~~~\\
-~{*}{*}{*}~Verification~of~simulation~parameters~{*}{*}{*}~~~\\
-~{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}~~~\\
-~...~~~\\
-~{*}{*}{*}~Minimum~period~resolved~=~4.308774~~~\\
-~{*}{*}{*}~Maximum~suggested~time~step~=~6.8863556E-02~~~\\
-~...~~~\\
-~...~~~\\
-
-\end{lyxcode}
-
-then change \texttt{DT} in the \texttt{DATA/Par\_file} to be somewhere
-close to the maximum suggested time step. In the example above: \\
- \texttt{DT = 0.05d0} \\
- would (most probably) work fine. It could be also bigger than the
-0.068 s suggested. This depends a bit on the distortions of your mesh
-elements. The more regular they are, the bigger you can choose \texttt{DT}.
-Just play with this value a bit and see when the simulation becomes
-stable ...
-
-\end{description}
-
-\chapter{\label{cha:License}License }
-
-\textbf{GNU GENERAL PUBLIC LICENSE Version 2, June 1991. Copyright
-(C) 1989, 1991 Free Software Foundation, Inc. 59 Temple Place, Suite
-330, Boston, MA 02111-1307 USA} \\
- Everyone is permitted to copy and distribute verbatim copies of this
-license document, but changing it is not allowed.
-
-\section*{Preamble}
-
-The licenses for most software are designed to take away your freedom
-to share and change it. By contrast, the GNU General Public License
-is intended to guarantee your freedom to share and change free software
--- to make sure the software is free for all its users. This General
-Public License applies to most of the Free Software Foundation's software
-and to any other program whose authors commit to using it. (Some other
-Free Software Foundation software is covered by the GNU Library General
-Public License instead.) You can apply it to your programs, too.
-
-When we speak of free software, we are referring to freedom, not price.
-Our General Public Licenses are designed to make sure that you have
-the freedom to distribute copies of free software (and charge for
-this service if you wish), that you receive source code or can get
-it if you want it, that you can change the software or use pieces
-of it in new free programs; and that you know you can do these things.
-
-To protect your rights, we need to make restrictions that forbid anyone
-to deny you these rights or to ask you to surrender the rights. These
-restrictions translate to certain responsibilities for you if you
-distribute copies of the software, or if you modify it.
-
-For example, if you distribute copies of such a program, whether gratis
-or for a fee, you must give the recipients all the rights that you
-have. You must make sure that they, too, receive or can get the source
-code. And you must show them these terms so they know their rights.
-
-We protect your rights with two steps:
-\begin{enumerate}
-\item Copyright the software, and
-\item Offer you this license which gives you legal permission to copy, distribute
-and/or modify the software.
-\end{enumerate}
-Also, for each author's protection and ours, we want to make certain
-that everyone understands that there is no warranty for this free
-software. If the software is modified by someone else and passed on,
-we want its recipients to know that what they have is not the original,
-so that any problems introduced by others will not reflect on the
-original authors' reputations.
-
-Finally, any free program is threatened constantly by software patents.
-We wish to avoid the danger that redistributors of a free program
-will individually obtain patent licenses, in effect making the program
-proprietary. To prevent this, we have made it clear that any patent
-must be licensed for everyone's free use or not licensed at all.
-
-The precise terms and conditions for copying, distribution and modification
-follow.
-
-\section*{GNU GENERAL PUBLIC LICENSE TERMS AND CONDITIONS FOR COPYING, DISTRIBUTION
-AND MODIFICATION }
-\begin{itemize}
-\item 0.This License applies to any program or other work which contains
-a notice placed by the copyright holder saying it may be distributed
-under the terms of this General Public License. The ``Program''
-below refers to any such program or work, and a ``work based on the
-Program'' means either the Program or any derivative work under copyright
-law: that is to say, a work containing the Program or a portion of
-it, either verbatim or with modifications and/or translated into another
-language. (Hereinafter, translation is included without limitation
-in the term ``modification.'') Each licensee is addressed as ``you.''\\
- \\
- Activities other than copying, distribution and modification are
-not covered by this License; they are outside its scope. The act of
-running the Program is not restricted, and the output from the Program
-is covered only if its contents constitute a work based on the Program
-(independent of having been made by running the Program). Whether
-that is true depends on what the Program does.\end{itemize}
-\begin{enumerate}
-\item You may copy and distribute verbatim copies of the Program's source
-code as you receive it, in any medium, provided that you conspicuously
-and appropriately publish on each copy an appropriate copyright notice
-and disclaimer of warranty; keep intact all the notices that refer
-to this License and to the absence of any warranty; and give any other
-recipients of the Program a copy of this License along with the Program.
-
-You may charge a fee for the physical act of transferring a copy,
-and you may at your option offer warranty protection in exchange for
-a fee.
-
-\item You may modify your copy or copies of the Program or any portion of
-it, thus forming a work based on the Program, and copy and distribute
-such modifications or work under the terms of Section 1 above, provided
-that you also meet all of these conditions:
-
-\begin{enumerate}
-\item You must cause the modified files to carry prominent notices stating
-that you changed the files and the date of any change.
-\item You must cause any work that you distribute or publish, that in whole
-or in part contains or is derived from the Program or any part thereof,
-to be licensed as a whole at no charge to all third parties under
-the terms of this License.
-\item If the modified program normally reads commands interactively when
-run, you must cause it, when started running for such interactive
-use in the most ordinary way, to print or display an announcement
-including an appropriate copyright notice and a notice that there
-is no warranty (or else, saying that you provide a warranty) and that
-users may redistribute the program under these conditions, and telling
-the user how to view a copy of this License. (Exception: if the Program
-itself is interactive but does not normally print such an announcement,
-your work based on the Program is not required to print an announcement.)
-\end{enumerate}
-
-These requirements apply to the modified work as a whole. If identifiable
-sections of that work are not derived from the Program, and can be
-reasonably considered independent and separate works in themselves,
-then this License, and its terms, do not apply to those sections when
-you distribute them as separate works. But when you distribute the
-same sections as part of a whole which is a work based on the Program,
-the distribution of the whole must be on the terms of this License,
-whose permissions for other licensees extend to the entire whole,
-and thus to each and every part regardless of who wrote it.
-
-Thus, it is not the intent of this section to claim rights or contest
-your rights to work written entirely by you; rather, the intent is
-to exercise the right to control the distribution of derivative or
-collective works based on the Program.
-
-In addition, mere aggregation of another work not based on the Program
-with the Program (or with a work based on the Program) on a volume
-of a storage or distribution medium does not bring the other work
-under the scope of this License.
-
-\item You may copy and distribute the Program (or a work based on it, under
-Section 2) in object code or executable form under the terms of Sections
-1 and 2 above provided that you also do one of the following:
-
-\begin{enumerate}
-\item Accompany it with the complete corresponding machine-readable source
-code, which must be distributed under the terms of Sections 1 and
-2 above on a medium customarily used for software interchange; or,
-\item Accompany it with a written offer, valid for at least three years,
-to give any third party, for a charge no more than your cost of physically
-performing source distribution, a complete machine-readable copy of
-the corresponding source code, to be distributed under the terms of
-Sections 1 and 2 above on a medium customarily used for software interchange;
-or,
-\item Accompany it with the information you received as to the offer to
-distribute corresponding source code. (This alternative is allowed
-only for noncommercial distribution and only if you received the program
-in object code or executable form with such an offer, in accord with
-Subsection b above.)
-\end{enumerate}
-
-The source code for a work means the preferred form of the work for
-making modifications to it. For an executable work, complete source
-code means all the source code for all modules it contains, plus any
-associated interface definition files, plus the scripts used to control
-compilation and installation of the executable. However, as a special
-exception, the source code distributed need not include anything that
-is normally distributed (in either source or binary form) with the
-major components (compiler, kernel, and so on) of the operating system
-on which the executable runs, unless that component itself accompanies
-the executable.
-
-If distribution of executable or object code is made by offering access
-to copy from a designated place, then offering equivalent access to
-copy the source code from the same place counts as distribution of
-the source code, even though third parties are not compelled to copy
-the source along with the object code.
-
-\item You may not copy, modify, sublicense, or distribute the Program except
-as expressly provided under this License. Any attempt otherwise to
-copy, modify, sublicense or distribute the Program is void, and will
-automatically terminate your rights under this License. However, parties
-who have received copies, or rights, from you under this License will
-not have their licenses terminated so long as such parties remain
-in full compliance.
-\item You are not required to accept this License, since you have not signed
-it. However, nothing else grants you permission to modify or distribute
-the Program or its derivative works. These actions are prohibited
-by law if you do not accept this License. Therefore, by modifying
-or distributing the Program (or any work based on the Program), you
-indicate your acceptance of this License to do so, and all its terms
-and conditions for copying, distributing or modifying the Program
-or works based on it.
-\item Each time you redistribute the Program (or any work based on the Program),
-the recipient automatically receives a license from the original licensor
-to copy, distribute or modify the Program subject to these terms and
-conditions. You may not impose any further restrictions on the recipients'
-exercise of the rights granted herein. You are not responsible for
-enforcing compliance by third parties to this License.
-\item If, as a consequence of a court judgment or allegation of patent infringement
-or for any other reason (not limited to patent issues), conditions
-are imposed on you (whether by court order, agreement or otherwise)
-that contradict the conditions of this License, they do not excuse
-you from the conditions of this License. If you cannot distribute
-so as to satisfy simultaneously your obligations under this License
-and any other pertinent obligations, then as a consequence you may
-not distribute the Program at all. For example, if a patent license
-would not permit royalty-free redistribution of the Program by all
-those who receive copies directly or indirectly through you, then
-the only way you could satisfy both it and this License would be to
-refrain entirely from distribution of the Program.
-
-If any portion of this section is held invalid or unenforceable under
-any particular circumstance, the balance of the section is intended
-to apply and the section as a whole is intended to apply in other
-circumstances.
-
-It is not the purpose of this section to induce you to infringe any
-patents or other property right claims or to contest validity of any
-such claims; this section has the sole purpose of protecting the integrity
-of the free software distribution system, which is implemented by
-public license practices. Many people have made generous contributions
-to the wide range of software distributed through that system in reliance
-on consistent application of that system; it is up to the author/donor
-to decide if he or she is willing to distribute software through any
-other system and a licensee cannot impose that choice.
-
-This section is intended to make thoroughly clear what is believed
-to be a consequence of the rest of this License.
-
-\item If the distribution and/or use of the Program is restricted in certain
-countries either by patents or by copyrighted interfaces, the original
-copyright holder who places the Program under this License may add
-an explicit geographical distribution limitation excluding those countries,
-so that distribution is permitted only in or among countries not thus
-excluded. In such case, this License incorporates the limitation as
-if written in the body of this License.
-\item The Free Software Foundation may publish revised and/or new versions
-of the General Public License from time to time. Such new versions
-will be similar in spirit to the present version, but may differ in
-detail to address new problems or concerns.
-
-Each version is given a distinguishing version number. If the Program
-specifies a version number of this License which applies to it and
-``any later version,'' you have the option of following the terms
-and conditions either of that version or of any later version published
-by the Free Software Foundation. If the Program does not specify a
-version number of this License, you may choose any version ever published
-by the Free Software Foundation.
-
-\item If you wish to incorporate parts of the Program into other free programs
-whose distribution conditions are different, write to the author to
-ask for permission. For software which is copyrighted by the Free
-Software Foundation, write to the Free Software Foundation; we sometimes
-make exceptions for this. Our decision will be guided by the two goals
-of preserving the free status of all derivatives of our free software
-and of promoting the sharing and reuse of software generally.
-\end{enumerate}
-
-\subsection*{NO WARRANTY }
-\begin{itemize}
-\item 11.BECAUSE THE PROGRAM IS LICENSED FREE OF CHARGE, THERE IS NO WARRANTY
-FOR THE PROGRAM, TO THE EXTENT PERMITTED BY APPLICABLE LAW. EXCEPT
-WHEN OTHERWISE STATED IN WRITING THE COPYRIGHT HOLDERS AND/OR OTHER
-PARTIES PROVIDE THE PROGRAM ``AS IS'' WITHOUT WARRANTY OF ANY KIND,
-EITHER EXPRESSED OR IMPLIED, INCLUDING, BUT NOT LIMITED TO, THE IMPLIED
-WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE.
-THE ENTIRE RISK AS TO THE QUALITY AND PERFORMANCE OF THE PROGRAM IS
-WITH YOU. SHOULD THE PROGRAM PROVE DEFECTIVE, YOU ASSUME THE COST
-OF ALL NECESSARY SERVICING, REPAIR OR CORRECTION.
-\item 12.IN NO EVENT UNLESS REQUIRED BY APPLICABLE LAW OR AGREED TO IN WRITING
-WILL ANY COPYRIGHT HOLDER, OR ANY OTHER PARTY WHO MAY MODIFY AND/OR
-REDISTRIBUTE THE PROGRAM AS PERMITTED ABOVE, BE LIABLE TO YOU FOR
-DAMAGES, INCLUDING ANY GENERAL, SPECIAL, INCIDENTAL OR CONSEQUENTIAL
-DAMAGES ARISING OUT OF THE USE OR INABILITY TO USE THE PROGRAM (INCLUDING
-BUT NOT LIMITED TO LOSS OF DATA OR DATA BEING RENDERED INACCURATE
-OR LOSSES SUSTAINED BY YOU OR THIRD PARTIES OR A FAILURE OF THE PROGRAM
-TO OPERATE WITH ANY OTHER PROGRAMS), EVEN IF SUCH HOLDER OR OTHER
-PARTY HAS BEEN ADVISED OF THE POSSIBILITY OF SUCH DAMAGES.
-\end{itemize}
-
-\section*{END OF TERMS AND CONDITIONS }
-
-\subsection*{How to Apply These Terms to Your New Programs}
-
-If you develop a new program, and you want it to be of the greatest
-possible use to the public, the best way to achieve this is to make
-it free software which everyone can redistribute and change under
-these terms.
-
-To do so, attach the following notices to the program. It is safest
-to attach them to the start of each source file to most effectively
-convey the exclusion of warranty; and each file should have at least
-the ``copyright'' line and a pointer to where the full notice is
-found. For example:
-\begin{quote}
-One line to give the program's name and a brief idea of what it does.
-Copyright {\footnotesize $\copyright$ (}year) (name of author)
-
-This program is free software; you can redistribute it and/or modify
-it under the terms of the GNU General Public License as published
-by the Free Software Foundation; either version 2 of the License,
-or (at your option) any later version.
-
-This program is distributed in the hope that it will be useful, but
-WITHOUT ANY WARRANTY; without even the implied warranty of MERCHANTABILITY
-or FITNESS FOR A PARTICULAR PURPOSE. See the GNU General Public License
-for more details.
-
-You should have received a copy of the GNU General Public License
-along with this program; if not, write to the Free Software Foundation,
-Inc., 59 Temple Place, Suite 330, Boston, MA 02111-1307 USA
-\end{quote}
-Also add information on how to contact you by electronic and paper
-mail.
-
-If the program is interactive, make it output a short notice like
-this when it starts in an interactive mode:
-\begin{quote}
-Gnomovision version 69, Copyright $\copyright$ year name of author
-Gnomovision comes with ABSOLUTELY NO WARRANTY; for details type `show
-w'. This is free software, and you are welcome to redistribute it
-under certain conditions; type `show c' for details.
-\end{quote}
-The hypothetical commands `show w' and `show c' should show the appropriate
-parts of the General Public License. Of course, the commands you use
-may be called something other than `show w' and `show c'; they could
-even be mouse-clicks or menu items -- whatever suits your program.
-
-You should also get your employer (if you work as a programmer) or
-your school, if any, to sign a ``copyright disclaimer'' for the
-program, if necessary. Here is a sample; alter the names:
-\begin{quote}
-Yoyodyne, Inc., hereby disclaims all copyright interest in the program
-`Gnomovision' (which makes passes at compilers) written by James Hacker.
-
-(signature of Ty Coon)\\
- 1 April 1989 \\
- Ty Coon, President of Vice
-\end{quote}
-This General Public License does not permit incorporating your program
-into proprietary programs. If your program is a subroutine library,
-you may consider it more useful to permit linking proprietary applications
-with the library. If this is what you want to do, use the GNU Library
-General Public License instead of this License.
-\end{document}

### 2015-04-24 doc/USER_MANUAL/B_channel_codes.tex
updates user manual to use facilitate the generation of a markdown version to the github wiki

--- 2015-04-18 doc/USER_MANUAL/manual_SPECFEM3D_Cartesian.tex
@@
-% !TEX encoding = IsoLatin
-%% LyX 2.0.2 created this file.  For more info, see http://www.lyx.org/.
-%% Do not edit unless you really know what you are doing.
-\documentclass[oneside,english]{book}
-\usepackage[T1]{fontenc}
-%% DK DK we do not accept accented letters, use the old LaTeX syntax
-%% DK DK for them instead: \'e \`e  \"e \^e and so on.
-%% DK DK \usepackage[latin1]{inputenc}
-\usepackage[letterpaper]{geometry}
-\geometry{verbose,tmargin=1in,bmargin=1in,lmargin=1in,rmargin=1in}
-\setcounter{secnumdepth}{3}
-\setcounter{tocdepth}{3}
-\usepackage{babel}
-\usepackage{float}
-\usepackage{textcomp}
-\usepackage{url}
-\usepackage{amsmath}
-\usepackage{amssymb}
-\usepackage{graphicx}
-\usepackage[authoryear]{natbib}
-\usepackage[unicode=true,pdfusetitle,
- bookmarks=true,bookmarksnumbered=true,bookmarksopen=false,
- breaklinks=false,pdfborder={0 0 0},backref=section,colorlinks=false,pdfpagemode=None]
- {hyperref}
-\hypersetup{
- pdftex,pdfstartview=FitH,pdfpagelayout=SinglePage}
-\usepackage{breakurl}
-
-\makeatletter
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% LyX specific LaTeX commands.
-\newcommand{\noun}[1]{\textsc{#1}}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Textclass specific LaTeX commands.
-\newenvironment{lyxcode}
-{\par\begin{list}{}{
-\setlength{\rightmargin}{\leftmargin}
-\setlength{\listparindent}{0pt}% needed for AMS classes
-\raggedright
-\setlength{\itemsep}{0pt}
-\setlength{\parsep}{0pt}
-\normalfont\ttfamily}%
- \item[]}
-{\end{list}}
-\newenvironment{lyxlist}[1]
-{\begin{list}{}
-{\settowidth{\labelwidth}{#1}
- \setlength{\leftmargin}{\labelwidth}
- \addtolength{\leftmargin}{\labelsep}
- \renewcommand{\makelabel}[1]{##1\hfil}}}
-{\end{list}}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% User specified LaTeX commands.
-%\renewcommand{\baselinestretch}{1.5}
-
-% figures
-\usepackage[dvips]{epsfig}
-
-% fonts
-\usepackage{times}
-
-% hyperlinks to sections and references
-
-\newcommand{\urlwithparentheses}[1]{(\url{#1})}
-
-% biblio GJI
-\bibliographystyle{abbrvnat}
-
-\newcommand{\nexxi}{\mbox{\texttt{NEX\_XI\/}}}
-\newcommand{\nexeta}{\mbox{\texttt{NEX\_ETA\/}}}
-\newcommand{\nprocxi}{\mbox{\texttt{NPROC\_XI\/}}}
-\newcommand{\nproceta}{\mbox{\texttt{NPROC\_ETA\/}}}
-\newcommand{\nchunks}{\mbox{\texttt{NCHUNKS\/}}}
-\newcommand{\eg}{e.g.,}
-
-% colors for the text
-\@ifundefined{definecolor}
- {\usepackage{color}}{}
-\newcommand{\red}[1]{\textcolor{Red}{#1}}
-
-\usepackage{babel}
-
-\makeatother
-
-\begin{document}
-\begin{center}
-\thispagestyle{empty}\vspace*{-1.8truecm} %
-\makebox[1\textwidth]{%
-\includegraphics[width=0.83\paperwidth]{figures/specfem_3d_Cartesian-cover} %
-}
-\par\end{center}
-
-\title{\textbf{SPECFEM3D Cartesian}\\
- \textbf{User Manual}}
-
-\author{$\copyright$ Princeton University (USA), CNRS and University of Marseille (France),\\
-and ETH Z\"urich (Switzerland)\\
-Version 3.0}
-
-\date{\today}
-
-\maketitle
-
-\section*{Authors}
-
-\noindent The SPECFEM3D Cartesian package was first developed by Dimitri
-Komatitsch and Jean-Pierre Vilotte at Institut de Physique du Globe
-(IPGP) in Paris, France from 1995 to 1997 and then by Dimitri Komatitsch
-and Jeroen Tromp at Harvard University and Caltech, USA, starting in 1998.
-The story started on April 4, 1995, when Prof. Yvon Maday from CNRS and University of Paris, France, gave a lecture to
-Dimitri Komatitsch and Jean-Pierre Vilotte at IPG about the nice properties of the Legendre spectral-element method with diagonal mass matrix that he had used for
-other equations. We are deeply indebted and thankful to him for that.
-That followed a visit by Dimitri Komatitsch to OGS (Istituto Nazionale di Oceanografia e di Geofisica Sperimentale) in Trieste, Italy, in February 1995
-to meet with G\'eza Seriani and Enrico Priolo, who introduced him to their 2D Chebyshev version of the spectral-element method with a non-diagonal mass matrix.
-We are deeply indebted and thankful to them for that.
-
-%%%%%%%% please do NOT merge the lines of the paragraph below, it makes it difficult to insert new names in alphabetical order; leave a single name per line
-%%%%%%%% please do NOT merge the lines of the paragraph below, it makes it difficult to insert new names in alphabetical order; leave a single name per line
-%%%%%%%% please do NOT merge the lines of the paragraph below, it makes it difficult to insert new names in alphabetical order; leave a single name per line
-Since then it has been developed and maintained by a development team:
-in alphabetical order,
-Jean-Paul (Pablo) Ampuero,
-Kangchen Bai,
-Piero Basini,
-C\'eline Blitz,
-Alexis Bottero,
-Ebru Bozda\u{g},
-Emanuele Casarotti,
-Joseph Charles,
-Min Chen,
-Paul Cristini,
-Percy Galvez,
-Dominik G\"oddeke,
-Vala Hj\"orleifsd\'ottir,
-Sue Kientz,
-Dimitri Komatitsch,
-Jes\'us Labarta,
-Nicolas Le Goff,
-Pieyre Le Loher,
-Matthieu Lefebvre,
-Qinya Liu,
-Yang Luo,
-Alessia Maggi,
-Federica Magnoni,
-Roland Martin,
-Ren\'e Matzen,
-Dennis McRitchie,
-Matthias Meschede,
-Peter Messmer,
-David Mich\'ea,
-Surendra Nadh Somala,
-Tarje Nissen-Meyer,
-Daniel Peter,
-Max Rietmann,
-Elliott Sales de Andrade,
-Brian Savage,
-Bernhard Schuberth,
-Anne Sieminski,
-Leif Strand,
-Carl Tape,
-Jeroen Tromp,
-Jean-Pierre Vilotte,
-Zhinan Xie,
-Chang-Hua Zhang,
-Hejun Zhu.\\
-
-The cover graphic of the manual was created by Santiago Lombeyda from
-Caltech's Center for Advanced Computing Research (CACR) \urlwithparentheses{http://www.cacr.caltech.edu}.\\
-
-The code is released open-source under the GNU version 2+ license, see the license at the end of this manual.\\
-
-\newpage{}
-\section*{Current and past main participants or main sponsors of the SPECFEM project (in no particular order)}
-%
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=0.125\textwidth]{figures/logo_cnrs}\vspace*{2truemm}
-\includegraphics[width=0.125\textwidth]{figures/logo_princeton}\vspace*{2truemm}
-\includegraphics[width=0.15\textwidth]{figures/logo_aix_marseille_universite}\vspace*{0.02truemm}
-\includegraphics[width=0.15\textwidth]{figures/logo_ETH}\vspace*{2truemm}
-\includegraphics[width=0.12\textwidth]{figures/logo_CSC_China}\vspace*{0.02truemm}
-\includegraphics[width=0.15\textwidth]{figures/logo_inria}\vspace*{2truemm}
-\includegraphics[width=0.14\textwidth]{figures/logo_UPPA}
-\par\end{centering}
-%
-\vspace*{-2truemm}
-%
-\noindent \begin{centering}
-\includegraphics[width=0.15\textwidth]{figures/logo_NVIDIA}\vspace*{2truemm}
-\includegraphics[width=0.125\textwidth]{figures/logo_IUF}\vspace*{2truemm}
-\includegraphics[width=0.135\textwidth]{figures/logo_Caltech}\vspace*{2truemm}
-\includegraphics[width=0.135\textwidth]{figures/logo_Harvard}\vspace*{2truemm}
-\includegraphics[width=0.135\textwidth]{figures/logo_IPGP}\vspace*{2truemm}
-\includegraphics[width=0.112\textwidth]{figures/logo_ANR}\vspace*{2truemm}
-\includegraphics[width=0.112\textwidth]{figures/logo_NSF}\vspace*{2truemm}
-\par\end{centering}
-%
-\vspace*{-2truemm}
-%
-\noindent \begin{centering}
-\includegraphics[width=0.190\textwidth]{figures/logo_European_Union}\vspace*{2truemm}
-\includegraphics[width=0.112\textwidth]{figures/logo_GENCI}\vspace*{2truemm}
-\includegraphics[width=0.112\textwidth]{figures/logo_PRACE}\vspace*{2truemm}
-\includegraphics[width=0.112\textwidth]{figures/logo_CINES}\vspace*{2truemm}
-\includegraphics[width=0.130\textwidth]{figures/logo_Oak_Ridge}\vspace*{2truemm}
-\hspace*{3mm}\includegraphics[width=0.130\textwidth]{figures/logo_fondation_Del_Duca}
-\includegraphics[width=0.130\textwidth]{figures/logo_CIG}\vspace*{2truemm}
-\par\end{centering}
-%
-\vspace*{-2truemm}
-%
-\noindent \begin{centering}
-\includegraphics[width=0.112\textwidth]{figures/logo_University_of_Toronto}\vspace*{2truemm}
-\includegraphics[width=0.109\textwidth]{figures/logo_INGV}\vspace*{2truemm}
-\includegraphics[width=0.102\textwidth]{figures/logo_Univ_Toulouse}\vspace*{2truemm}
-\includegraphics[width=0.112\textwidth]{figures/logo_TOTAL}\vspace*{2truemm}
-\includegraphics[width=0.130\textwidth]{figures/logo_Fairbanks}\vspace*{2truemm}
-\includegraphics[width=0.120\textwidth]{figures/logo_CINECA}\vspace*{2truemm}
-\includegraphics[width=0.140\textwidth]{figures/logo_Intel_Exascale_Labs}\vspace*{2truemm}
-\includegraphics[width=0.130\textwidth]{figures/logo_Maison_Simulation}
-\par\end{centering}
-\end{figure}
-
-\newpage{}
-
-\tableofcontents{}
-
-\chapter{Introduction}
-
-The software package SPECFEM3D Cartesian simulates seismic wave propagation
-at the local or regional scale and performs full waveform imaging (FWI) or adjoint tomography based upon the spectral-element method
-(SEM). The SEM is a continuous Galerkin technique \citep{TrKoLi08,PeKoLuMaLeCaLeMaLiBlNiBaTr11},
-which can easily be made discontinuous \citep{BeMaPa94,Ch00,KoWoHu02,ChCaVi03,LaWaBe05,Kop06,WiStBuGh10,AcKo11};
-it is then close to a particular case of the discontinuous Galerkin
-technique \citep{ReHi73,LeRa74,Arn82,JoPi86,BoMaHe91,FaRi99,HuHuRa99,CoKaSh00,GiHeWa02,RiWh03,MoRi05,GrScSc06,AiMoMu06,BeLaPi06,DuKa06,DeSeWh08,PuAmKa09,WiStBuGh10,DeSe10,EtChViGl10},
-with optimized efficiency because of its tensorized basis functions
-\citep{WiStBuGh10,AcKo11}. In particular, it can accurately handle
-very distorted mesh elements \citep{OlSe11}.\\
-
-It has very good accuracy and convergence properties \citep{MaPa89,SePr94,DeFiMu02,Coh02,DeSe07,SeOl08,AiWa09,AiWa10,MeStTh12}.
-The spectral element approach admits spectral rates of convergence
-and allows exploiting $hp$-convergence schemes. It is also very well
-suited to parallel implementation on very large supercomputers \citep{KoTsChTr03,TsKoChTr03,KoLaMi08a,CaKoLaTiMiLeSnTr08,KoViCh10}
-as well as on clusters of GPU accelerating graphics cards \citep{Kom11,MiKo10,KoMiEr09,KoErGoMi10}.
-Tensor products inside each element can be optimized to reach very
-high efficiency \citep{DeFiMu02}, and mesh point and element numbering
-can be optimized to reduce processor cache misses and improve cache
-reuse \citep{KoLaMi08a}. The SEM can also handle triangular (in 2D)
-or tetrahedral (in 3D) elements \citep{WinBoyd96,TaWi00,KoMaTrTaWi01,Coh02,MeViSa06}
-as well as mixed meshes, although with increased cost and reduced
-accuracy in these elements, as in the discontinuous Galerkin method.\\
-
-Note that in many geological models in the context of seismic wave
-propagation studies (except for instance for fault dynamic rupture
-studies, in which very high frequencies or supershear rupture need
-to be modeled near the fault, see e.g. \citet{BeGlCrViPi07,BeGlCrVi09,PuAmKa09,TaCrEtViBeSa10})
-a continuous formulation is sufficient because material property contrasts
-are not drastic and thus conforming mesh doubling bricks can efficiently
-handle mesh size variations \citep{KoTr02a,KoLiTrSuStSh04,LeChLiKoHuTr08,LeChKoHuTr09,LeKoHuTr09}.\\
-
-For a detailed introduction to the SEM as applied to regional seismic
-wave propagation, please consult \citet{PeKoLuMaLeCaLeMaLiBlNiBaTr11,TrKoLi08,KoVi98,KoTr99,ChKoViCaVaFe07}
-and in particular \citet{LeKoHuTr09,LeChKoHuTr09,LeChLiKoHuTr08,GoAmTaCaSmSaMaKo09,WiKoScTr04,KoLiTrSuStSh04}.
-A detailed theoretical analysis of the dispersion
-and stability properties of the SEM is available in \citet{Coh02}, \citet{DeSe07}, \citet{SeOl07}, \citet{SeOl08} and \citet{MeStTh12}.\\
-
-Effects due to lateral variations in compressional-wave speed, shear-wave
-speed, density, a 3D crustal model, topography and bathymetry are
-included. The package can accommodate full 21-parameter anisotropy
-(see~\citet{ChTr07}) as well as lateral variations in attenuation
-\citep{SaKoTr10}. Adjoint capabilities and finite-frequency kernel
-simulations are included \citep{TrKoLi08,PeKoLuMaLeCaLeMaLiBlNiBaTr11,LiTr06,FiIgBuKe09,ViOp09}.\\
-
-The SEM was originally developed in computational fluid dynamics \citep{Pat84,MaPa89}
-and has been successfully adapted to address problems in seismic wave
-propagation. Early seismic wave propagation applications of the SEM,
-utilizing Legendre basis functions and a perfectly diagonal mass matrix,
-include \citet{CoJoTo93}, \citet{Kom97}, \citet{FaMaPaQu97}, \citet{CaGa97},
-\citet{KoVi98} and \citet{KoTr99}, whereas applications involving
-Chebyshev basis functions and a nondiagonal mass matrix include \citet{SePr94},
-\citet{PrCaSe94} and \citet{SePrPr95}.\\
-
-All SPECFEM3D software is written in Fortran2003 with full portability
-in mind, and conforms strictly to the Fortran2003 standard. It uses
-no obsolete or obsolescent features of Fortran. The package uses parallel
-programming based upon the Message Passing Interface (MPI) \citep{GrLuSk94,Pac97}.\\
-
-SPECFEM3D won the Gordon Bell award for best performance at the SuperComputing~2003
-conference in Phoenix, Arizona (USA) (see \citet{KoTsChTr03} and
-\url{www.sc-conference.org/sc2003/nrfinalaward.html}). It was a finalist
-again in 2008 for a run at 0.16 petaflops (sustained) on 149,784 processors
-of the `Jaguar' Cray XT5 system at Oak Ridge National Laboratories
-(USA) \citep{CaKoLaTiMiLeSnTr08}. It also won the BULL Joseph Fourier
-supercomputing award in 2010.\\
-
-It reached the sustained one petaflop performance level for the first time in February 2013
-on the Blue Waters Cray supercomputer at the National Center for Supercomputing Applications (NCSA), located at the University of Illinois at Urbana-Champaign (USA).\\
-
-This new release of the code includes Convolution or
-Auxiliary Differential Equation Perfectly Matched absorbing Layers
-(C-PML or ADE-PML) \citep{MaKoEz08,MaKoGe08,MaKo09,MaKoGeBr10,KoMa07}.
-It also includes support for GPU graphics card acceleration \citep{Kom11,MiKo10,KoMiEr09,KoErGoMi10}.\\
-
-The next release of the code will use the PT-SCOTCH parallel and threaded
-version of SCOTCH for mesh partitioning instead of the serial version.\\
-
-SPECFEM3D Cartesian includes coupled fluid-solid domains and adjoint
-capabilities, which enables one to address seismological inverse problems,
-but for linear rheologies only so far. To accommodate visco-plastic
-or non-linear rheologies, the reader can refer to the \href{http://geoelse.stru.polimi.it/}{GeoELSE}software
-package \citep{CaGa97,StPaIg09}.
-
-\section{Citation}
-
-If you use SPECFEM3D Cartesian for your own research, please cite
-at least one of the following articles:
-\begin{description}
-\item [{Numerical simulations in general}] ~\\
- Forward and adjoint simulations are described in detail in \citet{TrKoLi08,PeKoLuMaLeCaLeMaLiBlNiBaTr11,VaCaSaKoVi99,KoMiEr09,KoErGoMi10,ChKoViCaVaFe07,MaKoDi09,KoViCh10,CaKoLaTiMiLeSnTr08,TrKoHjLiZhPeBoMcFrTrHu10,KoRiTr02,KoTr02a,KoTr02b,KoTr99}
-or \citet{KoVi98}. Additional aspects of adjoint simulations are
-described in \citet{TrTaLi05,LiTr06,TrKoLi08,LiTr08,TrKoHjLiZhPeBoMcFrTrHu10,PeKoLuMaLeCaLeMaLiBlNiBaTr11}.
-Domain decomposition is explained in detail in \citet{MaKoBlLe08},
-and excellent scaling up to 150,000 processor cores is shown for instance
-in \citet{CaKoLaTiMiLeSnTr08,KoLaMi08a,MaKoBlLe08,KoErGoMi10,Kom11},
-\item [{GPU computing}] ~\\
- Computing on GPU graphics cards for acoustic or seismic wave propagation
-applications is described in detail in \citet{Kom11,MiKo10,KoMiEr09,KoErGoMi10}.
-\end{description}
-\noindent If you use this new version, which has non blocking
-MPI for much better performance for medium or large runs, please cite
-at least one of these six articles, in which results of non blocking
-MPI runs are presented: \citet{PeKoLuMaLeCaLeMaLiBlNiBaTr11,KoErGoMi10,KoViCh10,Kom11,CaKoLaTiMiLeSnTr08,MaKoBlLe08}.\\
-
-\noindent If you work on geophysical applications, you may be interested
-in citing some of these application articles as well, among others:
-\begin{description}
-\item [{Southern California simulations}] ~\\
- \citet{KoLiTrSuStSh04,KrJiKoTr06a,KrJiKoTr06b}.
-
-If you use the 3D southern California model, please cite \citet{SuSh03}
-(Los Angeles model), \citet{lovelyetal06} (Salton Trough), and \citet{hauksson2000}
-(southern California). The Moho map was determined by \citet{zhu&kanamori2000}.
-The 1D SoCal model was developed by \citet{DrHe90}.
-
-\item [{Anisotropy}] ~\\
- \citet{ChTr07,JiTsKoTr05,ChFaKo04,FaChKo04,RiRiKoTrHe02,TrKo00}.
-\item [{Attenuation}] ~\\
- \citet{SaKoTr10,KoTr02a,KoTr99}.
-\item [{Topography}] ~\\
- \citet{LeKoHuTr09,LeChKoHuTr09,LeChLiKoHuTr08,GoAmTaCaSmSaMaKo09,WiKoScTr04}.
-\end{description}
-The corresponding Bib\TeX{} entries may be found in file \texttt{doc/USER\_MANUAL/bibliography.bib}.
-
-\section{Support}
-
-This material is based upon work supported by the USA National Science
-Foundation under Grants No. EAR-0406751 and EAR-0711177, by the French
-CNRS, French INRIA Sud-Ouest MAGIQUE-3D, French ANR NUMASIS under
-Grant No. ANR-05-CIGC-002, and European FP6 Marie Curie International
-Reintegration Grant No. MIRG-CT-2005-017461. Any opinions, findings,
-and conclusions or recommendations expressed in this material are
-those of the authors and do not necessarily reflect the views of the
-USA National Science Foundation, CNRS, INRIA, ANR or the European
-Marie Curie program.
-
-\chapter{\label{cha:Getting-Started}Getting Started}
-
-To download the SPECFEM3D\_Cartesian software package, type this:
-\begin{lyxcode}
-git clone --recursive --branch devel https://github.com/geodynamics/specfem3d.git
-\end{lyxcode}
-{\small \par}
-
-We recommend that you add {\texttt{ulimit -S -s unlimited}} to your
-{\texttt{.bash\_profile}} file and/or {\texttt{limit stacksize
-unlimited }} to your {\texttt{.cshrc}} file to suppress any potential
-limit to the size of the Unix stack.\\
-
-Then, to configure the software for your system, run the \texttt{configure}
-shell script. This script will attempt to guess the appropriate configuration
-values for your system. However, at a minimum, it is recommended that
-you explicitly specify the appropriate command names for your Fortran
-compiler and MPI package (another option is to define FC, CC and MPIF90
-in your .bash\_profile or your .cshrc file):
-\begin{lyxcode}
-./configure~FC=ifort~MPIFC=mpif90
-\end{lyxcode}
-Note that MPI must be installed with MPI-IO enabled because parts of SPECFEM3D perform I/Os through MPI-IO.
-
-Before running the \texttt{configure} script, you should probably
-edit file \texttt{flags.guess} to make sure that it contains the best
-compiler options for your system. Known issues or things to check
-are:
-\begin{description}
-\item [{\texttt{Intel ifort compiler}}] See if you need to add \texttt{-assume byterecl} for your machine.
-\textbf{In the case of that compiler, we have noticed that versions dot zero sometimes have bugs or issues that can lead to wrong results when running the code, thus we \emph{strongly} recommend using versions dot one or above (for instance version 13.1 instead of 13.0, version 14.1 instead of 14.0 and so on)}.
-\item [{\texttt{IBM compiler}}] See if you need to add \texttt{-qsave}
-or \texttt{-qnosave} for your machine.
-\item [{\texttt{Mac OS}}] You will probably need to install \texttt{XCODE}.
-\end{description}
-When compiling on an IBM machine with the \texttt{xlf} and \texttt{xlc}
-compilers, we suggest running the \texttt{configure} script with the
-following options:\\
-
-\noindent \texttt{./configure FC=xlf90\_r MPIFC=mpif90 CC=xlc\_r CFLAGS=\textquotedbl{}-O3
--q64\textquotedbl{} FCFLAGS=\textquotedbl{}-O3 -q64\textquotedbl{}
--with-scotch-dir=...}.\\
-
-On SGI systems, \texttt{flags.guess} automatically informs \texttt{configure}
-to insert `\texttt{`TRAP\_FPE=OFF}'' into the generated \texttt{Makefile}
-in order to turn underflow trapping off.\\
-
-You can add \texttt{--enable-vectorization} to the configuration options to speed up the code in the fluid (acoustic) and elastic parts.
-This works fine if (and only if) your computer always allocates a contiguous memory block for each allocatable array;
-this is the case for most machines and most compilers, but not all. To disable this feature, use option \texttt{--disable-vectorization}.
-For more details see \href{https://github.com/geodynamics/specfem3d/issues/81}{github.com/geodynamics/specfem3d/issues/81} .
-To check if that option works fine on your machine, run the code with and without it for an acoustic/elastic model and make sure the seismograms are identical.\\
-
-Note that we use CUBIT (now called Trelis) to create meshes of hexahedra, but other packages
-can be used as well, for instance GiD from \url{http://gid.cimne.upc.es}
-or Gmsh from \url{http://geuz.org/gmsh} \citep{GeRe09}. Even mesh
-creation packages that generate tetrahedra, for instance TetGen from
-\url{http://tetgen.berlios.de}, can be used because each tetrahedron
-can then easily be decomposed into four hexahedra as shown in the
-picture of the TetGen logo at \url{http://tetgen.berlios.de/figs/Delaunay-Voronoi-3D.gif};
-while this approach does not generate hexahedra of optimal quality,
-it can ease mesh creation in some situations and it has been shown
-that the spectral-element method can very accurately handle distorted
-mesh elements \citep{OlSe11}.
-
-The SPECFEM3D Cartesian software package relies on the SCOTCH library
-to partition meshes created with CUBIT. METIS \citep{KaKu98a,KaKu98c,KaKu98b}
-can also be used instead of SCOTCH if you prefer, by editing file
-src/decompose\_mesh/decompose\_mesh.F90 and uncommenting the flag
-USE\_METIS\_INSTEAD\_OF\_SCOTCH. You will also then need to install
-and compile Metis version 4.0 (do {*}NOT{*} install Metis version
-5.0, which has incompatible function calls) and edit \texttt{src/decompose\_mesh/Makefile.in}
-and uncomment the METIS link flag in that file before running \texttt{configure}.\\
-
-The SCOTCH library \citep{PeRo96} provides efficent static mapping,
-graph and mesh partitioning routines. SCOTCH is a free software package
-developed by Fran\c{c}ois Pellegrini et al. from LaBRI and INRIA in Bordeaux,
-France, downloadable from the web page \url{https://gforge.inria.fr/projects/scotch/}.
-In case no SCOTCH libraries can be found on the system, the configuration
-will bundle the version provided with the source code for compilation.
-The path to an existing SCOTCH installation can to be set explicitly
-with the option \texttt{-{}-with-scotch-dir}. Just as an example:
-\begin{lyxcode}
-./configure~FC=ifort~MPIFC=mpif90~-{}-with-scotch-dir=/opt/scotch
-\end{lyxcode}
-If you use the Intel ifort compiler to compile the code, we recommend
-that you use the Intel icc C compiler to compile Scotch, i.e., use:
-\begin{lyxcode}
-./configure~CC=icc~FC=ifort~MPIFC=mpif90
-\end{lyxcode}
-
-When compiling the SCOTCH source code, if you get a message such as: "ld: cannot find -lz",
-the Zlib compression development library is probably missing on your machine and you will need to install it or ask your system administrator to
-do so. On Linux machines the package is often called "zlib1g-dev" or similar. (thus "sudo apt-get install zlib1g-dev" would install it)
-
-To compile a serial version of the code for small meshes that fits
-on one compute node and can therefore be run serially, run \texttt{configure}
-with the \texttt{-{}-without-mpi} option to suppress all calls to
-MPI.
-
-A summary of the most important configuration variables follows.
-\begin{description}
-\item [{\texttt{F90}}] Path to the Fortran compiler.
-\item [{\texttt{MPIF90}}] Path to MPI Fortran.
-\item [{\texttt{MPI\_FLAGS}}] Some systems require this flag to link to
-MPI libraries.
-\item [{\texttt{FLAGS\_CHECK}}] Compiler flags.
-\end{description}
-The configuration script automatically creates for each executable
-a corresponding \texttt{Makefile} in the \texttt{src/} subdirectoy.
-The \texttt{Makefile} contains a number of suggested entries for various
-compilers, e.g., Portland, Intel, Absoft, NAG, and Lahey. The software
-has run on a wide variety of compute platforms, e.g., various PC clusters
-and machines from Sun, SGI, IBM, Compaq, and NEC. Select the compiler
-you wish to use on your system and choose the related optimization
-flags. Note that the default flags in the \texttt{Makefile} are undoubtedly
-not optimal for your system, so we encourage you to experiment with
-these flags and to solicit advice from your systems administrator.
-Selecting the right compiler and optimization flags can make a tremendous
-difference in terms of performance. We welcome feedback on your experience
-with various compilers and flags.
-
-Now that you have set the compiler information, you need to select
-a number of flags in the \texttt{constants.h} file depending on your
-system:
-\begin{description}
-\item [{\texttt{LOCAL\_PATH\_IS\_ALSO\_GLOBAL}}] Set to \texttt{.false.}
-on most cluster applications. For reasons of speed, the (parallel)
-distributed database generator typically writes a (parallel) database
-for the solver on the local disks of the compute nodes. Some systems
-have no local disks, e.g., BlueGene or the Earth Simulator, and other
-systems have a fast parallel file system, in which case this flag
-should be set to \texttt{.true.}. Note that this flag is not used
-by the database generator or the solver; it is only used for some
-of the post-processing.
-\end{description}
-The package can run either in single or in double precision mode.
-The default is single precision because for almost all calculations
-performed using the spectral-element method using single precision
-is sufficient and gives the same results (i.e. the same seismograms);
-and the single precision code is faster and requires exactly half
-as much memory. Select your preference by selecting the appropriate
-setting in the \texttt{constants.h} file:
-\begin{description}
-\item [{\texttt{CUSTOM\_REAL}}] Set to \texttt{SIZE\_REAL} for single precision
-and \texttt{SIZE\_DOUBLE} for double precision.
-\end{description}
-In the \texttt{precision.h} file:
-\begin{description}
-\item [{\texttt{CUSTOM\_MPI\_TYPE}}] Set to \texttt{MPI\_REAL} for single
-precision and \texttt{MPI\_DOUBLE\_PRECISION} for double precision.
-\end{description}
-On many current processors (e.g., Intel, AMD, IBM Power), single precision
-calculations are significantly faster; the difference can typically
-be 10\% to 25\%. It is therefore better to use single precision. What
-you can do once for the physical problem you want to study is run
-the same calculation in single precision and in double precision on
-your system and compare the seismograms. If they are identical (and
-in most cases they will), you can select single precision for your
-future runs.\\
-
-If your compiler has problems with the \texttt{use mpi} statements that are used in the code, use the script called
-\texttt{replace\_use\_mpi\_with\_include\_mpif\_dot\_h.pl} in the root directory to replace all of them with \texttt{include 'mpif.h'} automatically.
-
-\section{Adding OpenMP support in addition to MPI}
-
-NOTE FROM JULY 2013: OpenMP support is maybe / probably not maintained any more.
-Thus the section below is maybe obsolete.
-
-OpenMP support can be enabled in addition to MPI. However, in many
-cases performance will not improve because our pure MPI implementation
-is already heavily optimized and thus the resulting code will in fact
-be slightly slower. A possible exception could be IBM BlueGene-type
-architectures.
-
-To enable OpenMP, uncomment the OpenMP compiler option in two lines
-in file \texttt{src/specfem3D/Makefile.in} (before running \texttt{configure})
-and also uncomment the \texttt{\#define USE\_OPENMP} statement in
-file\\
- \texttt{src/specfem3D/specfem3D.F90}.
-
-The DO-loop using OpenMP threads has a SCHEDULE property. The OMP\_SCHEDULE
-environment variable can set the scheduling policy of that DO-loop.
-Tests performed by Marcin Zielinski at SARA (The Netherlands) showed
-that often the best scheduling policy is DYNAMIC with the size of
-the chunk equal to the number of OpenMP threads, but most preferably
-being twice as the number of OpenMP threads (thus chunk size = 8 for
-4 OpenMP threads etc). If OMP\_SCHEDULE is not set or is empty, the
-DO-loop will assume generic scheduling policy, which will slow down
-the job quite a bit.
-
-\section{Compiling on an IBM BlueGene}
-
-\underline{More recent installation instruction for IBM BlueGene, from April 2013:}\\
-
-\noindent Edit file \texttt{flags.guess} and put this for \texttt{FLAGS\_CHECK}:\\
-
-\noindent \texttt{-g -qfullpath -O2 -qsave -qstrict -qtune=qp -qarch=qp
--qcache=auto -qhalt=w}\\
- \texttt{-qfree=f90 -qsuffix=f=f90 -qlanglvl=95pure -Q -Q+rank,swap\_all
--Wl,-relax}\\
-
-\noindent The most relevant are the -qarch and -qtune flags, otherwise
-if these flags are set to ``auto'' then they are wrongly assigned
-to the architecture of the frond-end node, which is different from
-that on the compute nodes. You will need to set these flags to the
-right architecture for your BlueGene compute nodes, which is not necessarily
-``qp''; ask your system administrator. On some machines if is necessary
-to use -O2 in these flags instead of -O3 due to a compiler bug of
-the XLF version installed. We thus suggest to first try -O3, and then
-if the code does not compile or does not run fine then switch back
-to -O2. The debug flags (-g, -qfullpath) do not influence performance
-but are useful to get at least some insights in case of problems.\\
-
-\noindent Before running \texttt{configure}, select the XL Fortran
-compiler by typing \texttt{module load bgq-xl/1.0} or \texttt{module
-load bgq-xl} (another, less efficient option is to load the GNU compilers
-using \texttt{module load bgq-gnu/4.4.6} or similar).\\
-
-\noindent Then, to configure the code, type this:\\
-
-\noindent \texttt{./configure FC=bgxlf90\_r MPIFC=mpixlf90\_r CC=bgxlc\_r
-LOCAL\_PATH\_IS\_ALSO\_GLOBAL=true}\\
-
-In order for the SCOTCH domain decomposer to compile, on some (but
-not all) Blue Gene systems you may need to run \texttt{configure}
-with \texttt{CC=gcc} instead of \texttt{CC=bgxlc\_r}.
-
-\noindent \underline{Older installation instruction for IBM BlueGene, from 2011:}\\
-
-\noindent To compile the code on an IBM BlueGene, Laurent L\'eger from
-IDRIS, France, suggests the following: compile the code with
-
-\noindent \texttt{ FLAGS\_CHECK=\textquotedbl{}-O3 -qsave -qstrict
--qtune=auto -qarch=450d -qcache=auto }~\\
- \texttt{ -qfree=f90 -qsuffix=f=f90 -g -qlanglvl=95pure -qhalt=w -Q
--Q+rank,swap\_all -Wl,-relax\textquotedbl{}}
-
-\noindent Option \textquotedbl{}-Wl,-relax\textquotedbl{} must be
-added on many (but not all) BlueGene systems to be able to link the
-binaries \texttt{xmeshfem3D} and \texttt{xspecfem3D} because the final
-link step is done by the GNU \texttt{ld} linker even if one uses \texttt{FC=bgxlf90\_r,
-MPIFC=mpixlf90\_r} and \texttt{CC=bgxlc\_r} to create all the object
-files. On the contrary, on some BlueGene systems that use the native
-AIX linker option \textquotedbl{}-Wl,-relax\textquotedbl{} can lead
-to problems and must be suppressed from \texttt{flags.guess}.
-
-\noindent Also, \texttt{AR=ar, ARFLAGS=cru} and \texttt{RANLIB=ranlib}
-are hardwired in all \texttt{Makefile.in} files by default, but to
-cross-compile on BlueGene/P one needs to change these values to \texttt{AR=bgar,
-ARFLAGS=cru} and \texttt{RANLIB=bgranlib}. Thus the easiest thing
-to do is to modify all \texttt{Makefile.in} files and the \texttt{configure}
-script to set them automatically by \texttt{configure}. One then just
-needs to pass the right commands to the \texttt{configure} script:
-
-\noindent \texttt{ ./configure -{}-prefix=/path/to/SPECFEM3DG\_SP
--{}-host=Babel -{}-build=BGP }~\\
- \texttt{ FC=bgxlf90\_r MPIFC=mpixlf90\_r CC=bgxlc\_r AR=bgar ARFLAGS=cru
-}~\\
- \texttt{ RANLIB=bgranlib LOCAL\_PATH\_IS\_ALSO\_GLOBAL=false}
-
-\noindent This trick can be useful for all hosts on which one needs
-to cross-compile.
-
-\noindent On BlueGene, one also needs to run the \texttt{xcreate\_header\_file}
-binary file manually rather than in the Makefile:
-
-\noindent \texttt{bgrun -np 1 -mode VN -exe ./bin/xcreate\_header\_file}
-
-\section{Visualizing the subroutine calling tree of the source code}
-
-\noindent
-Packages such as \texttt{Doxywizard} can be used to visualize the calling tree of the
-subroutines of the  source code. \texttt{Doxywizard} is
-a GUI front-end for configuring and running \texttt{Doxygen}.
-
-\bigskip
-
-\noindent To do your own call graphs, you can follow these simple steps below.
-
-\begin{enumerate}
-\setcounter{enumi}{-1}
-\item Install \texttt{Doxygen} \underline{and} \texttt{graphviz} (the two are usually in the package manager of classic Linux distribution).
-\item Run in the terminal : \texttt{doxygen -g}, which creates a \texttt{Doxyfile} that tells doxygen what you want it to do.
-\item Edit the Doxyfile. Two Doxyfile-type files have been already committed in the directory \\
-\noindent \texttt{specfem3d/doc/Call\_trees}:
-
-\begin{itemize}
-\item[\textbullet] \texttt{Doxyfile\_truncated\_call\_tree} will generate call graphs with maximum 3 or 4 levels of tree structure,
-\item[\textbullet] \texttt{Doxyfile\_complete\_call\_tree} will generate call graphs with complete tree structure.
-\end{itemize}
-
-The important entries in the Doxyfile are:
-
-\begin{description}
-\item [{\texttt{PROJECT\_NAME}}]
-\item [{\texttt{OPTIMIZE\_FOR\_FORTRAN}}] Set to YES
-\item [{\texttt{EXTRACT\_ALL}}] Set to YES
-\item [{\texttt{EXTRACT\_PRIVATE}}] Set to YES
-\item [{\texttt{EXTRACT\_STATIC}}] Set to YES
-\item [{\texttt{INPUT}}] From the directory \texttt{specfem3d/doc/Call\_trees}, it is \texttt{"../../src/"}
-\item [{\texttt{FILE\_PATTERNS}}] In SPECFEM case, it is \texttt{*.f90* *.F90* *.c* *.cu* *.h*}
-\item [{\texttt{HAVE\_DOT}}] Set to YES
-\item [{\texttt{CALL\_GRAPH}}] Set to YES
-\item [{\texttt{CALLER\_GRAPH}}] Set to YES
-\item [{\texttt{DOT\_PATH}}] The path where is located the dot program graphviz (if it is not in your \$PATH)
-\item [{\texttt{RECURSIVE}}] This tag can be used to turn specify whether or not subdirectories should be searched for input files as well. In the case of SPECFEM, set to YES.
-\item [{\texttt{EXCLUDE}}] Here, you can exclude:  \\
-\texttt{"../../src/specfem3D/older\_not\_maintained\_partial\_OpenMP\_port"}\\
-\texttt{"../../src/decompose\_mesh/scotch"}\\
-\texttt{"../../src/decompose\_mesh/scotch\_5.1.12b"}
-\item [{\texttt{DOT\_GRAPH\_MAX\_NODES}}] to set the maximum number of nodes that will be shown in the graph.
-If the number of nodes in a graph becomes larger than this value, doxygen will truncate the graph, which is visualized by representing a node as a red box.
-Minimum value: 0, maximum value: 10000, default value: 50.
-\item [{\texttt{MAX\_DOT\_GRAPH\_DEPTH}}] to set the maximum depth of the graphs generated by dot.
-A depth value of 3 means that only nodes reachable from the root by following a path via at most 3 edges will be shown.
-Using a depth of 0 means no depth restriction.
-Minimum value: 0, maximum value: 1000, default value: 0.
-
-\end{description}
-
-\item Run : \texttt{doxygen Doxyfile}, HTML and LaTeX files created by default in \texttt{html} and \texttt{latex} subdirectories.
-
-\item To see the call trees, you have to open the file \texttt{html/index.html} in your \underline{browser}.
-You will have many informations about each subroutines of SPECFEM (not only call graphs), you can click on every boxes / subroutines.
-It show you the call, and, the caller graph of each subroutine : the subroutines called by the concerned subroutine, and the previous subroutines who call this subroutine (the previous path), respectively.
-In the case of a truncated calling tree, the boxes with a red border indicates a node that has \underline{more} arrows than are shown (in other words: the graph is truncated with respect to this node).
-
-\end{enumerate}
-
-\medskip
-
-\noindent Finally, some useful links:
-
-\begin{itemize}
-\item[\textbullet] a good and short summary for the basic utilisation of Doxygen:
-
-\url{http://www.softeng.rl.ac.uk/blog/2010/jan/30/callgraph-fortran-doxygen/},
-
-\item[\textbullet] to configure the diagrams :
-
-\url{http://www.stack.nl/~dimitri/doxygen/manual/diagrams.html},
-
-\item[\textbullet] the complete alphabetical index of the tags in Doxyfile:
-
-\url{http://www.stack.nl/~dimitri/doxygen/manual/config.html},
-
-\item[\textbullet] more generally, the Doxygen manual:
-
-\url{http://www.stack.nl/~dimitri/doxygen/manual/index.html}.
-
-\end{itemize}
-
-\medskip
-
-\section{Using the ADIOS library for I/O}
-Regular POSIX I/O can be problematic when dealing with large simulations one large
-clusters (typically more than $10,000$ processes). SPECFEM3D use the ADIOS library~\cite{Liu2013}
-to deal transparently take advantage of advanced paralllel file system features. To enable
-ADIOS, the following steps should be done:
-\begin{enumerate}
-\item Install ADIOS (available from \url{https://www.olcf.ornl.gov/center-projects/adios/}).
-Make sure that your environment variables reference it.
-\item You may want to change ADIOS related values in the \texttt{constants.h.in} file.
-The default values probably suit most cases.
-\item Configure using the \texttt{--with-adios} flag.
-\end{enumerate}
-ADIOS is currently only usable for meshfem3D generated mesh (i.e. not for meshes generated
-with CUBTI). Additional control parameters are discussed in section~\ref{cha:Main-Parameter}.
-
-\section{Becoming a developer of the code, or making small modifications in the source code}
-
-If you want to develop new features in the code, and/or if you want to make small changes, improvements, or bug fixes, you are very welcome to contribute.
-To do so, i.e.
-to access the development branch of the source code with read/write access (in a safe way, no need to worry too much about breaking the package, there is a robot called BuildBot that is in charge of checking and validating all new contributions and changes), please visit this Web page: \url{https://github.com/geodynamics/specfem3d/wiki/Using-Hub}.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\chapter{\label{cha:Mesh-Generation}Mesh Generation}
-
-The first step in running a spectral-element simulation consists of
-constructing a high-quality mesh for the region under consideration.
-We provide two possibilities to do so: (1) relying on the external,
-hexahedral mesher CUBIT, or (2) using the provided, internal mesher
-\texttt{xmeshfem3D}. In the following, we explain these two approaches.
-
-\section{\label{cha:Running-the-Mesher-CUBIT}Meshing with \texttt{CUBIT}}
-
-CUBIT is a meshing tool suite for the creation of finite-element meshes
-for arbitrarily shaped models. It has been developed and maintained
-at Sandia National Laboratories and can be purchased for a small academic
-institutional fee at \url{http://cubit.sandia.gov}. Our experience
-showed that using CUBIT greatly facilitates and speeds up the generation
-and preparation of hexahedral, conforming meshes for a variety of
-geophysical models with increasing complexity.
-
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=0.6\textwidth]{figures/mount-cubit}
-\par\end{centering}
-
-\caption{Example of the graphical user interface of CUBIT. The hexahedral mesh
-shown in the main display consists of a hexahedral discretization
-of a single volume with topography.}
-
-\label{fig:mount.cubit}
-\end{figure}
-
-\noindent
-The basic steps in creating a load-balanced, partitioned mesh with
-CUBIT are:
-\begin{description}
-\item [{1.}] setting up a hexahedral mesh with CUBIT,
-\item [{2.}] exporting the CUBIT mesh into a SPECFEM3D Cartesian file format
-and
-\item [{3.}] partitioning the SPECFEM3D Cartesian mesh files for a chosen
-number of cores.
-\end{description}
-Examples are provided in the SPECFEM3D Cartesian package in the subdirectory
-\texttt{EXAMPLES/}. We strongly encourage you to contribute your own
-example to this package by contacting the CIG Computational Seismology
-Mailing List \urlwithparentheses{cig-seismo@geodynamics.org}.
-
-\subsection{Creating the Mesh with CUBIT}
-
-For the installation and handling of the CUBIT meshing tool suite,
-please refer to the CUBIT user manual and documentation. In order
-to give you a basic understanding of how to use CUBIT for our purposes,
-examples are provided in the SPECFEM3D Cartesian package in the subdirectory
-\texttt{EXAMPLES/}:
-\begin{description}
-\item [{\texttt{homogeneous\_halfspace}}] Creates a single block model
-and assigns elastic material parameters.
-\item [{\texttt{layered\_halfspace}}] Combines two different, elastic material
-volumes and creates a refinement layer between the two. This example
-can be compared for validation against the solutions provided in subdirectory
-~\\
- \texttt{VALIDATION\_3D\_SEM\_SIMPLER\_LAYER\_SOURCE\_DEPTH/}.
-\item [{\texttt{waterlayered\_halfspace}}] Combines an acoustic and elastic
-material volume as in a schematic marine survey example.
-\item [{\texttt{tomographic\_model}}] Creates a single block model whose
-material properties will have to be read in from a tomographic model
-file during the databases creation by \texttt{xgenerate\_databases}.
-\end{description}
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=0.45\textwidth]{figures/example-homogeneous}
-\includegraphics[width=0.45\textwidth]{figures/example-2layers} \\
- \includegraphics[width=0.45\textwidth]{figures/example-water} \includegraphics[width=0.45\textwidth]{figures/example-tomo}
-\par\end{centering}
-
-\caption{Screenshots of the CUBIT examples provided in subdirectory \texttt{EXAMPLES/}:
-homogeneous halfspace (top-left), layered halfspace (top-right), water
-layered halfspace (bottom-left) and tomographic model (bottom-right).}
-
-\label{fig:examples.cubit}
-\end{figure}
-
-In each example subdirectory you will find a \texttt{README} file,
-which explains in a step-by-step tutorial the workflow for the example.
-Please feel free to contribute your own example to this package by
-contacting the CIG Computational Seismology Mailing List \urlwithparentheses{cig-seismo@geodynamics.org}.
-
-In some cases, to re-create the meshes for the examples given, just type "claro ./create\_mesh.py" or similar
-from the command line ("claro" is the command to run CUBIT from the command line).
-
-IMPORTANT note: in order to correctly set up GEOCUBIT and run the examples, please read the file called "EXAMPLES/README";
-in particular, please make sure you correctly set up the Python paths as indicated in that file.
-
-\subsection{Exporting the Mesh with \texttt{run\_cubit2specfem3d.py} }
-
-\label{subsec:Exporting-the-Mesh}
-
-Once the geometric model volumes in CUBIT are meshed, you prepare
-the model for exportation with the definition of material blocks and
-boundary surfaces. Thus, prior to exporting the mesh, you need to
-define blocks specifying the materials and absorbing boundaries in
-CUBIT. This process could be done automatically using the script \texttt{run\_cubit2specfem3d.py}
-if the mesh meets some conditions or manually, following the block
-convention:
-\begin{description}
-\item [{material\_name}] Each material should have a specific block defined
-by a unique name. The name convention of the material is to start
-with either 'elastic' or 'acoustic'. It must be then followed by a
-unique identifier, e.g. 'elastic 1', 'elastic 2', etc. The additional
-attributes to the block define the material description.
-
-For an elastic material:
-\begin{description}
-\item [{material\_id}] An integer value which is unique for this material.
-\item [{Vp}] P-wave speed of the material (given in m/s).
-\item [{Vs}] S-wave speed of the material (given in m/s).
-\item [{rho}] density of the material (given in kg/m$^{3}$).
-\item [{Q}] quality factor to use in case of a simulation with attenuation
-turned on. It should be between 1 and 9000. In case no attenuation
-information is available, it can be set to zero. Please note that
-your Vp- and Vs-speeds are given for a reference frequency. To change
-this reference frequency, you change the value of \texttt{ATTENUATION\_f0\_REFERENCE}
-in the main constants file \texttt{constants.h} found in subdirectory
-\texttt{src/shared/}.
-\item [{anisotropic\_flag}] Flag describing the anisotropic model to use
-in case an anisotropic simulation should be conducted. See the file
-\texttt{model\_aniso.f90} in subdirectory \texttt{src/generate\_databases/}
-for an implementation of the anisotropic models. In case no anisotropy
-is available, it can be set to zero.
-\end{description}
-
-Note that this material block has to be defined using all the volumes
-which belong to this elastic material. For volumes belonging to another,
-different material, you will need to define a new material block.
-
-For an acoustic material:
-\begin{description}
-\item [{material\_id}] An integer value which is unique for this material.
-\item [{Vp}] P-wave speed of the material (given in m/s).
-\item [{0}] S-wave speed of the material is ignored.
-\item [{rho}] density of the material (given in kg/m$^{3}$).
-\end{description}
-\item [{face\_topo}] Block definition for the surface which defines the
-free surface (which can have topography). The name of this block must
-be 'face\_topo', the block has to be defined using all the surfaces
-which constitute the complete free surface of the model.
-\item [{face\_abs\_xmin}] Block definition for the faces on the absorbing
-boundaries, one block for each surface with x=Xmin.
-\item [{face\_abs\_xmax}] Block definition for the faces on the absorbing
-boundaries, one block for each surface with x=Xmax.
-\item [{face\_abs\_ymin}] Block definition for the faces on the absorbing
-boundaries, one block for each surface with y=Ymin.
-\item [{face\_abs\_ymax}] Block definition for the faces on the absorbing
-boundaries, one block for each surface with y=Ymax.
-\item [{face\_abs\_bottom}] Block definition for the faces on the absorbing
-boundaries, one block for each surface with z=bottom.
-\end{description}
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=0.45\textwidth]{figures/mount-surface} \includegraphics[width=0.45\textwidth]{figures/mount-abs}
-\par\end{centering}
-
-\caption{Example of the block definitions for the free surface 'face\_topo'
-(left) and the absorbing boundaries, defined in a single block 'face\_abs\_xmin'
-(right) in CUBIT.}
-
-\label{fig:mount.abs}
-\end{figure}
-
-Optionally, instead of specifying for each surface at the boundaries
-a single block like mentioned above, you can also specify a single
-block for all boundary surfaces and name it as one of the absorbing
-blocks above, e.g. 'face\_abs\_xmin'.
-
-After the block definitions are done, you export the mesh using the
-script \texttt{cubit2specfem3d.py} provided in each of the example
-directories (linked to the common script \texttt{CUBIT\_GEOCUBIT/cubit2specfem3d.py}).
-If the export was successful, you should find the following
-files in a subdirectory \texttt{MESH/}:
-\begin{description}
-\item [{absorbing\_cpml\_file}] \textbf{(only needed in case of C-PML absorbing
-conditions)} Contains on the first line the total number of C-PML
-spectral elements in the mesh, and then on the following line the
-list of all these C-PML elements with two numbers per line: first
-the spectral element number, and then a C-PML flag indicating to which
-C-PML layer(s) that element belongs, according to the following convention:
-
-\begin{itemize}
-\item Flag = 1 : element belongs to a X CPML layer only (either in Xmin
-or in Xmax),
-\item Flag = 2 : element belongs to a Y CPML layer only (either in Ymin
-or in Ymax),
-\item Flag = 3 : element belongs to a Z CPML layer only (either in Zmin
-or in Zmax),
-\item Flag = 4 : element belongs to a X CPML layer and also to a Y CPML
-layer,
-\item Flag = 5 : element belongs to a X CPML layer and also to a Z CPML
-layer,
-\item Flag = 6 : element belongs to a Y CPML layer and also to a Z CPML
-layer,
-\item Flag = 7 : element belongs to a X, to a Y and to a Z CPML layer, i.e.,
-it belongs to a CPML corner.
-\end{itemize}
-
-Note that it does not matter whether an element belongs to a Xmin
-or to a Xmax CPML, the flag is the same in both cases; the same is
-true for Ymin and Ymax, and also Zmin and Zmax.
-
-When you have an existing CUBIT (or similar) mesh stored in SPECFEM3D
-format, i.e., if you have existing \textquotedbl{}nodes\_coords\_file\textquotedbl{}
-and \textquotedbl{}mesh\_file\textquotedbl{} files but do not know
-how to assign CPML flags to them, we have created a small serial Fortran
-program that will do that automatically for you, i.e., which will
-create the \textquotedbl{}absorbing\_cpml\_file\textquotedbl{} for
-you. That program is utils/CPML/convert\_external\_layers\_of\_a\_given\_mesh\_to\_CPML\_layers.f90,
-and a small Makefile is provided in that directory (utils/CPML).
-
-IMPORTANT: it is your responsibility to make sure that in the input
-CUBIT (or similar) mesh that this code will read in SPECFEM3D format
-from files \textquotedbl{}nodes\_coords\_file\textquotedbl{} and \textquotedbl{}mesh\_file\textquotedbl{}
-you have created layers of elements that constitute a layer of constant
-thickness aligned with the coordinate grid axes (X, Y and/or Z), so
-that this code can assign CPML flags to them. This code does NOT check
-that (because it cannot, in any easy way). The mesh inside these CPML
-layers does not need to be structured nor regular, any non-structured
-mesh is fine as long as it has flat PML inner and outer faces, parallel
-to the axes, and thus of a constant thickness. The thickness can be
-different for the X, Y and Z sides. But for X it must not vary, for
-Y it must not vary, and for Z it must not vary. If you do not know
-the exact thickness, you can use a slightly LARGER value in this code
-(say 2\% to 5\% more) and this code will fix that and will adjust
-it; never use a SMALLER value otherwise this code will miss some CPML
-elements.
-
-Note: in a future release we will remove the constraint of having
-CPML layers aligned with the coordinate axes; we will allow for meshes
-that are titled by any constant angle in the horizontal plane. However
-this is not implemented yet.
-
-Note: in the case of fluid-solid layers in contact, the fluid-solid
-interface currently needs to be flat and horizontal inside the CPML
-layer (i.e., bathymetry should become flat and horizontal when entering
-the CPML); this (small) constraint will probably remain in the code
-for a while because it makes fluid-solid matching inside the CPML
-much easier.
-
-\item [{materials\_file}] Contains the material associations for each element.
-The format is:
-
-\begin{lyxcode}
-element\_ID~material\_ID
-\end{lyxcode}
-
-where \texttt{element\_ID} is the element identifier and \texttt{material\_ID}
-is a unique identifier, positive (for materials taken from this list
-of materials, i.e. for which each spectral element has constant material
-properties taken from this list) or negative (for tomographic models,
-i.e. for spectral element whose real velocities and density will be
-assigned later by calling an external function to define model variations,
-for instance in the case of tomographic models; in such a case, material
-properties can vary inside each spectral element, i.e. be different
-at each of its Gauss-Lobatto-Legendre grid points).
-
-\item [{nummaterial\_velocity\_file}] Defines the material properties.
-
-\begin{itemize}
-\item For classical materials (i.e., spectral elements for which the velocity
-and density model will not be assigned by calling an external function
-to define for instance a tomographic model), the format is:
-
-\begin{lyxcode}
-domain\_ID~material\_ID~rho~vp~vs~Qkappa~Qmu~anisotropy\_flag
-\end{lyxcode}
-
-where \texttt{\textbf{domain\_ID}}\textbf{ is 1 for acoustic and 2
-for elastic or viscoelastic materials,} \texttt{material\_ID} a unique
-identifier, \texttt{rho} the density in $kg\, m^{-3}$, \texttt{vp}
-the P-wave speed in $m\, s^{-1}$, \texttt{vs} the S-wave speed in
-$m\, s^{-1}$, \texttt{Q} the quality factor and \texttt{anisotropy\_flag}
-an identifier for anisotropic models. Note that the \texttt{Qkappa}
-value is ignored by the code unless \texttt{FULL\_ATTENUATION\_SOLID}
-is set. Note also that both \texttt{Qkappa} and \texttt{Qmu} are ignored
-by the code unless \texttt{ATTENUATION} is set. If you want a model
-with no \texttt{Qmu} attenuation, both set \texttt{ATTENUATION} to
-\texttt{.false.} in the \texttt{Par\_file} and set \texttt{Qmu} to
-9999 here. If you want a model with no \texttt{Qkappa} attenuation,
-both set \texttt{FULL\_ATTENUATION\_SOLID} to \texttt{.false.} in
-the \texttt{Par\_file} and set \texttt{Qkappa} to 9999 here. %%magnoni
-
-\item For tomographic velocity models, please read Chapter \ref{cha:-Changing-the}
-and Section \ref{sec:Using-tomographic} `Using external tomographic
-Earth models' for further details.
-\end{itemize}
-\item [{nodes\_coords\_file}] Contains the point locations in Cartesian
-coordinates of the mesh element corners.
-\item [{mesh\_file}] Contains the mesh element connectivity. The hexahedral
-elements can have 8 or 27 nodes.\\
- See picture doc/mesh\_numbering\_convention/numbering\_convention\_27\_nodes.jpg
-to see\\
- in which (standard) order the points must be cited. In the case of
-8 nodes, just include the first 8 points.
-\item [{free\_or\_absorbing\_surface\_file\_zmax}] Contains the free surface
-connectivity or \\
- the surface connectivity of the absorbing boundary surface at the
-top (Zmax), \\
- depending on whether the top surface is defined as free or absorbing
-(\texttt{STACEY\_INSTEAD\_OF\_FREE\_SURFACE} in \texttt{DATA/Par\_file}). \\
- You should put both the surface of acoustic regions and of elastic
-regions in that file; that is, list all the element faces that constitute
-the surface of the model in that file.
-\item [{absorbing\_surface\_file\_xmax}] Contains the surface connectivity
-of the absorbing boundary surface at Xmax \\
-(also needed in the case of C-PML absorbing conditions, in order for
-the code to be able to impose Dirichlet conditions on their outer
-edge).
-\item [{absorbing\_surface\_file\_xmin}] Contains the surface connectivity
-of the absorbing boundary surface at Xmin \\
-(also needed in the case of C-PML absorbing conditions, in order for
-the code to be able to impose Dirichlet conditions on their outer
-edge).
-\item [{absorbing\_surface\_file\_ymax}] Contains the surface connectivity
-of the absorbing boundary surface at Ymax \\
-(also needed in the case of C-PML absorbing conditions, in order for
-the code to be able to impose Dirichlet conditions on their outer
-edge).
-\item [{absorbing\_surface\_file\_ymin}] Contains the surface connectivity
-of the absorbing boundary surface at Ymin \\
-(also needed in the case of C-PML absorbing conditions, in order for
-the code to be able to impose Dirichlet conditions on their outer
-edge).
-\item [{absorbing\_surface\_file\_bottom}] Contains the surface connectivity
-of the absorbing boundary surface at the bottom (Zmin) \\
-(also needed in the case of C-PML absorbing conditions, in order for
-the code to be able to impose Dirichlet conditions on their outer
-edge).
-\end{description}
-These mesh files are needed as input files for the partitioner \texttt{xdecompose\_mesh}
-to load-balance the mesh. Please see the next section for further
-details.
-
-In directory "CUBIT\_GEOCUBIT" we provide a script that can help doing the above tasks of exporting a CUBIT mesh
-to SPECFEM3D format automatically for you: "run\_cubit2specfem3d.py".
-Just edit them to indicate the path to your local installation of CUBIT and also the name of the *.cub existing CUBIT mesh file
-that you want to export to SPECFEM3D format. These scripts will do the conversion for you automatically except assigning material properties
-to the different mesh layers. To do so, you will then need to edit the file
-called "nummaterial\_velocity\_file" that will have just been created and change it from the prototype created:
-
-    0 1 vol1 --> syntax: \#material\_domain\_id \#material\_id \#rho \#vp \#vs \#Q\_kappa \#Q\_mu \#anisotropy
-    0 2 vol2 --> syntax: \#material\_domain\_id \#material\_id \#rho \#vp \#vs \#Q\_kappa \#Q\_mu \#anisotropy
-
-(where "vol1" and "vol2" here represent the volume labels that you have set while creating the mesh in CUBIT)
-to for instance
-
-    2 1 1500 2300 1800 9999.0 9999.0 0
-    2 2 1600 2500 20000 9999.0 9999.0 0
-
-\subsubsection*{Checking the mesh quality}
-
-The quality of the mesh may be inspected more precisely based upon
-the serial code in the file \texttt{check\_mesh\_quality\_}~\\
- \texttt{CUBIT\_Abaqus.f90} located in the directory \texttt{src/check\_mesh\_quality\_CUBIT\_Abaqus/}.
-Running this code is optional because no information needed by the
-solver is generated.
-
-Prior to running and compiling this code, you have to export your
-mesh in CUBIT to an ABAQUS (.inp) format. For example, export mesh
-block IDs belonging to volumes in order to check the quality of the
-hexahedral elements. You also have to determine a number of parameters
-of your mesh, such as the number of nodes and number of elements and
-modify the header of the \texttt{check\_mesh\_quality\_CUBIT\_Abaqus.f90}
-source file in directory \texttt{src/check\_mesh\_quality\_CUBIT\_Abaqus/}.
-
-\noindent
-Then, in the main directory, type
-\begin{lyxcode}
-{\small make~xcheck\_mesh\_quality\_CUBIT\_Abaqus}{\small \par}
-\end{lyxcode}
-and use
-\begin{lyxcode}
-{\small ./bin/xcheck\_mesh\_quality\_CUBIT\_Abaqus}{\small \par}
-\end{lyxcode}
-to generate an %not supported: AVS output file (\texttt{\small AVS\_meshquality.inp} in AVS UCD format) or
-OpenDX output file (\texttt{\small DX\_mesh\_quality.dx}{\small )
-that can be used to investigate mesh quality, e.g. skewness of elements,
-and a Gnuplot histogram (}\texttt{\small mesh\_quality\_histogram.txt}{\small )
-that can be plotted with gnuplot (type `}\texttt{\small gnuplot plot\_mesh\_quality\_histogram.gnu}{\small ').
-The histogram is also printed to the screen. Analyze that skewness
-histogram of mesh elements to make sure no element has a skewness
-above approximately 0.75, otherwise the mesh is of poor quality (and
-if even a single element has a skewness value above 0.80, then you
-must definitely improve the mesh). If you want to start designing
-your own meshes, this tool is useful for viewing your creations. You
-are striving for meshes with elements with `cube-like' dimensions,
-e.g., the mesh should contain no very elongated or skewed elements.}{\small \par}
-
-\subsection{Partitioning the Mesh with \texttt{xdecompose\_mesh}}
-
-The SPECFEM3D Cartesian software package performs large scale simulations
-in a parallel 'Single Process Multiple Data' way. The spectral-element
-mesh created with CUBIT needs to be distributed on the processors.
-This partitioning is executed once and for all prior to the execution
-of the solver so it is referred to as a static mapping.
-
-An efficient partitioning is important because it leverages the overall
-running time of the application. It amounts to balance the number
-of elements in each slice while minimizing the communication costs
-resulting from the placement of adjacent elements on different processors.
-\texttt{decompose\_mesh} depends on the SCOTCH library \citep{PeRo96},
-which provides efficent static mapping, graph and mesh partitioning
-routines. SCOTCH is a free software package developed by Fran\c{c}ois
-Pellegrini et al. from LaBRI and INRIA in Bordeaux, France, downloadable
-from the web page \url{https://gforge.inria.fr/projects/scotch/}.
-\\
-
-In most cases, the configuration with \texttt{./configure~FC=ifort}
-should be sufficient. During the configuration process, the script
-tries to find existing SCOTCH installations. In case your system has
-no pre-existing SCOTCH installation, we provide the source code of
-SCOTCH, which is released open source under the French CeCILL-C version
-1 license, in directory \texttt{src/decompose\_mesh/scotch\_5.1.12b}.
-This version gets bundled with the compilation of the SPECFEM3D Cartesian
-package if no libraries could have been found. If this automatic compilation
-of the SCOTCH libraries fails, please refer to file INSTALL.txt in
-that directory to see further details how to compile it on your system.
-In case you want to use a pre-existing installation, make sure you
-have correctly specified the path of the SCOTCH library when using
-the option \texttt{-{}-with-scotch-dir} with the \texttt{./configure}
-script. In the future you should be able to find more recent versions
-at \url{http://www.labri.fr/perso/pelegrin/scotch/scotch_en.html}
-. \\
-
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=0.45\textwidth]{figures/mount-partitions}
-\includegraphics[width=0.45\textwidth]{figures/mount-partitions2}
-\par\end{centering}
-
-\caption{Example of a mesh partitioning onto four cores. Each single core partition
-is colored differently. The executable \texttt{xdecompose\_mesh} can
-equally distribute the mesh on any arbitrary number of cores. Domain
-decomposition is explained in detail in \citet{MaKoBlLe08}, and excellent
-scaling up to 150,000 processor cores in shown for instance in \citet{CaKoLaTiMiLeSnTr08,KoLaMi08a,MaKoBlLe08,KoErGoMi10,Kom11}.}
-
-\label{fig:mount.partitions}
-\end{figure}
-
-When you are ready to compile, in the main directory type
-\begin{lyxcode}
-{\small make~xdecompose\_mesh}{\small \par}
-\end{lyxcode}
-If all paths and flags have been set correctly, the executable \texttt{bin/xdecompose\_mesh}
-should be produced.
-
-The partitioning is done in serial for now (in the next release we
-will provide a parallel version of that code). It needs to be run in the main directory because it expects the \texttt{./DATA/Par\_file}. The synopsis is:
-\begin{lyxcode}
-./bin/xdecompose\_mesh~nparts~input\_directory~output\_directory
-\end{lyxcode}
-where
-\begin{itemize}
-\item \texttt{nparts} is the number of partitions, i.e., the number of cores
-for the parallel simulations,
-\item \texttt{input\_directory} is the directory which holds all the files
-generated by the Python script \texttt{cubit2specfem3d.py} explained
-in the previous Section~\ref{subsec:Exporting-the-Mesh}, e.g. \texttt{./MESH/},
-and
-\item \texttt{output\_directory} is the directory for the output of this
-partitioner which stores ACII-format files named like \texttt{proc{*}{*}{*}{*}{*}{*}\_Database}
-for each partition. These files will be needed for creating the distributed
-databases, and have to reside in the directory \texttt{LOCAL\_PATH}
-specified in the main \texttt{Par\_file}, e.g. in directory \texttt{./OUTPUT\_FILES/DATABASES\_MPI}.
-Please see Chapter~\ref{cha:Creating-Distributed-Databases} for
-further details.
-\end{itemize}
-Note that all the files generated by the Python script \texttt{cubit2specfem3d.py}
-must be placed in the \texttt{input\_directory} folder before running
-the program.
-
-\section{\label{cha:Running-the-Mesher-Meshfem3D}Meshing with \texttt{xmeshfem3D}}
-
-In case you successfully ran the configuration script, you are also
-ready to compile the internal mesher. This is an alternative to CUBIT
-for the mesh generation of relatively simple models.
-
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[scale=0.5]{figures/socal_map_mpi}
-\par\end{centering}
-
-\caption{\label{fig:For-parallel-computing}For parallel computing purposes,
-the model block is subdivided in $\nprocxi\times\nproceta$ slices
-of elements. In this example we use $5^{2}=25$ processors. }
-\end{figure}
-
-\noindent
-In the main directory, type
-\begin{lyxcode}
-{\small make~xmeshfem3D}{\small \par}
-\end{lyxcode}
-If all paths and flags have been set correctly, the mesher should
-now compile and produce the executable \texttt{bin/xmeshfem3D}. Please
-note that \texttt{xmeshfem3D} must be called directly from the main
-directory, as most of the binaries of the package.
-
-Input for the mesh generation program is provided through the parameter
-file \texttt{Mesh\_Par\_file}, which resides in the subdirectory \texttt{DATA/meshfem3D\_files/}.
-(to see how to use it, see the EXAMPLES specific to the internal mesher in directory \texttt{EXAMPLES/meshfem3D\_examples/}.
-Before running the mesher, a number of parameters need to be set in
-the \texttt{Mesh\_Par\_file}. This requires a basic understanding
-of how the SEM is implemented, and we encourage you to read \citet{KoVi98,KoTr99}
-and \citet{KoLiTrSuStSh04}.
-
-The mesher and the solver use UTM coordinates internally, therefore
-you need to define the zone number for the UTM projection (e.g., zone
-11 for Los Angeles). Use decimal values for latitude and longitude
-(no minutes/seconds). These values are approximate; the mesher will
-round them off to define a square mesh in UTM coordinates. When running
-benchmarks on rectangular models, turn the UTM projection off by using
-the flag \texttt{\small SUPPRESS\_UTM\_PROJECTION}{\small , in which
-case all `longitude' parameters simply refer to the $x$~axis, and
-all `latitude' parameters simply refer to the $y$~axis. To run the
-mesher for a global simulation, the following parameters need to be
-set in the }\texttt{\small Mesh\_Par\_file}{\small :}{\small \par}
-\begin{description}
-\item [{\texttt{LATITUDE\_MIN}}] Minimum latitude in the block (negative
-for South).
-\item [{\texttt{LATITUDE\_MAX}}] Maximum latitude in the block.
-\item [{\texttt{LONGITUDE\_MIN}}] Minimum longitude in the block (negative
-for West).
-\item [{\texttt{LONGITUDE\_MAX}}] Maximum longitude in the block.
-\item [{\texttt{DEPTH\_BLOCK\_KM}}] Depth of bottom of mesh in kilometers.
-\item [{\texttt{\noun{UTM\_PROJECTION\_ZONE}}}] UTM projection zone in
-which your model resides, only valid when \texttt{SUPPRESS\_UTM\_}~\\
- \texttt{PROJECTION} is \texttt{.false.}.
-Use a negative zone number for the Southern hemisphere:
-the Northern hemisphere corresponds to zones +1 to +60,
-the Southern hemisphere to zones -1 to -60.
-We use the WGS84 (World Geodetic System 1984) reference ellipsoid for the UTM projection. If you prefer to use the Clarke 1866 ellipsoid,
-edit file \texttt{src/shared/utm\_geo.f90}, uncomment that ellipsoid and recompile the code.
-From http://en.wikipedia.org/wiki/Universal\_Transverse\_Mercator\_coordinate\_system :
-The Universal Transverse Mercator coordinate system was developed by the United States Army Corps of Engineers in the 1940s.
-The system was based on an ellipsoidal model of Earth. For areas within the contiguous United States
-the Clarke Ellipsoid of 1866 was used. For the remaining areas of Earth, including Hawaii, the International Ellipsoid was used.
-The WGS84 ellipsoid is now generally used to model the Earth in the UTM coordinate system,
-which means that current UTM northing at a given point can be 200+ meters different from the old one.
-For different geographic regions, other datum systems (e.g.: ED50, NAD83) can be used.
-\item [{\texttt{SUPPRESS\_UTM\_PROJECTION}}] set to be \texttt{.false.}
-when your model range is specified in geographical coordinates, and
-needs to be \texttt{.true.} when your model is specified in Cartesian
-coordinates. \noun{UTM projection zone in which your simulation region
-resides.}
-\item [{\texttt{INTERFACES\_FILE }}] File which contains the description
-of the topography and of the interfaces between the different layers
-of the model, if any. The number of spectral elements in the vertical
-direction within each layer is also defined in this file.
-\item [{$\nexxi$}] The number of spectral elements along one side of the
-block. This number \textit{must} be 8~$\times$~a multiple of $\nprocxi$
-defined below. Based upon benchmarks against semi-analytical discrete
-wavenumber synthetic seismograms \citep{KoLiTrSuStSh04}, determined
-that a $\nexxi=288$ run is accurate to a shortest period of roughly
-2~s. Therefore, since accuracy is determined by the number of grid
-points per shortest wavelength, for any particular value of $\nexxi$
-the simulation will be accurate to a shortest period determined by
-\begin{equation}
-\mbox{shortest period (s)}=(288/\nexxi)\times2.\label{eq:shortest_period}
-\end{equation}
- The number of grid points in each orthogonal direction of the reference
-element, i.e., the number of Gauss-Lobatto-Legendre points, is determined
-by \texttt{NGLLX} in the \texttt{constants.h} file. We generally use
-$\mbox{\texttt{NGLLX\/}}=5$, for a total of $5^{3}=125$ points per
-elements. We suggest not to change this value.
-\item [{$\nexeta$}] The number of spectral elements along the other side
-of the block. This number \textit{must} be 8~$\times$~a multiple
-of $\nproceta$ defined below.
-\item [{$\nprocxi$}] The number of processors or slices along one side
-of the block (see Figure~\ref{fig:For-parallel-computing}); we must
-have $\nexxi=8\times c\times\nprocxi$, where $c\ge1$ is a positive
-integer.
-\item [{$\nproceta$}] The number of processors or slices along the other
-side of the block; we must have $\nexeta=8\times c\times\nproceta$,
-where $c\ge1$ is a positive integer.
-\item [{\texttt{USE\_REGULAR\_MESH}}] set to be \texttt{.true.} if you
-want a perfectly regular mesh or \texttt{.false.} if you want to add
-doubling horizontal layers to coarsen the mesh. In this case, you
-also need to provide additional information by setting up the next
-three parameters.
-\item [{\texttt{NDOUBLINGS}}] The number of horizontal doubling layers.
-Must be set to \texttt{1} or \texttt{2} if \texttt{USE\_REGULAR\_MESH}
-is set to \texttt{.true.}.
-\item [{\texttt{NZ\_DOUBLING\_1}}] The position of the first doubling layer
-(only interpreted if \texttt{USE\_REGULAR\_MESH} is set to \texttt{.true.}).
-\item [{\texttt{NZ\_DOUBLING\_2}}] The position of the second doubling
-layer (only interpreted if \texttt{USE\_REGULAR\_MESH} is set to \texttt{.true.}
-and if \texttt{NDOUBLINGS} is set to \texttt{2}).
-\item [{\texttt{CREATE\_ABAQUS\_FILES}}] Set this flag to \texttt{.true.}
-to save Abaqus FEA \urlwithparentheses{www.simulia.com} mesh files
-for subsequent viewing. Turning the flag on generates files in the
-\texttt{LOCAL\_PATH} directory. See Section~\ref{sec:Mesh-graphics}
-for a discussion of mesh viewing features.
-\item [{\texttt{CREATE\_DX\_FILES}}] Set this flag to \texttt{.true.} to
-save OpenDX \urlwithparentheses{www.opendx.org} mesh files for subsequent
-viewing.
-\item [{\texttt{LOCAL\_PATH}}] Directory in which the partitions generated
-by the mesher will be written. Generally one uses a directory on the
-local disk of the compute nodes, although on some machines these partitions
-are written on a parallel (global) file system (see also the earlier
-discussion of the \texttt{LOCAL\_PATH\_IS\_ALSO\_GLOBAL} flag in Chapter~\ref{cha:Getting-Started}).
-The mesher generates the necessary partitions in parallel, one set
-for each of the $\nprocxi\times\nproceta$ slices that constitutes
-the mesh (see Figure~\ref{fig:For-parallel-computing}). After the
-mesher finishes, you can log in to one of the compute nodes and view
-the contents of the \texttt{LOCAL\_PATH} directory to see the files
-generated by the mesher. These files will be needed for creating the
-distributed databases, and have to reside in the directory \texttt{LOCAL\_PATH}
-specified in the main \texttt{Par\_file}, e.g. in directory \texttt{OUTPUT\_FILES/DATABASES\_MPI}.
-Please see Chapter~\ref{cha:Creating-Distributed-Databases} for
-further details.
-\item [{\texttt{NMATERIALS}}] The number of different materials in your
-model. In the following lines, each material needs to be defined as
-:
-
-\begin{lyxcode}
-material\_ID~rho~vp~vs~Q~anisotropy\_flag~domain\_ID
-\end{lyxcode}
-
-where
-\begin{itemize}
-\item \texttt{Q} : quality factor (0=no attenuation)
-\item \texttt{anisotropy\_flag} : 0=no anisotropy / 1,2,.. check with implementation
-in \texttt{aniso\_model.f90}
-\item \texttt{domain\_id} : 1=acoustic / 2=elastic
-\end{itemize}
-\item [{\texttt{NREGIONS}}] The number of regions in the mesh. In the following
-lines, because the mesh is regular or 'almost regular', each region
-is defined as :
-
-\begin{lyxcode}
-NEX\_XI\_BEGIN~NEX\_XI\_END~NEX\_ETA\_BEGIN~NEX\_ETA\_END~NZ\_BEGIN~NZ\_END~material\_ID
-\end{lyxcode}
-\end{description}
-The \texttt{INTERFACES\_FILE} parameter of \texttt{Mesh\_Par\_File}
-defines the file which contains the settings of the topography grid
-and of the interfaces grids. Topography is defined as a set of elevation
-values on a regular 2D grid. It is also possible to define interfaces
-between the layers of the model in the same way. The file needs to
-define several parameters:
-\begin{itemize}
-\item The number of interfaces, including the topography. This needs to
-be set at the first line. Then, from the bottom to the top of the
-model, you need to define the grids with:
-\item \texttt{SUPPRESS\_UTM\_PROJECTION} flag as described previously,
-\item number of points along $x$ and $y$ direction (NXI and NETA),
-\item minimal $x$ and $y$ coordinates (LONG\_MIN and LAT\_MIN),
-\item spacing between points along $x$ and $y$ (SPACING\_XI and SPACING\_ETA)
-and
-\item the name of the file which contains the elevation values (in $y$.$x$
-increasing order).
-\end{itemize}
-At the end of this file, you simply need to set the number of spectral
-elements in the vertical direction for each layer. We provide a few
-models in the {\texttt{EXAMPLES/}} directory. \\
-
-Finally, depending on your system, you might need to provide a file
-that tells MPI what compute nodes to use for the simulations. The
-file must have a number of entries (one entry per line) at least equal
-to the number of processors needed for the run. A sample file is provided
-in the file \texttt{mymachines}. This file is not used by the mesher
-or solver, but is required by the \texttt{go\_mesher} and \texttt{go\_solver}
-default job submission scripts. See Chapter~\ref{cha:Scheduler}
-for information about running the code on a system with a scheduler,
-e.g., LSF.
-
-Now that you have set the appropriate parameters in the \texttt{Mesh\_Par\_file}
-and have compiled the mesher, you are ready to launch it! This is
-most easily accomplished based upon the \texttt{go\_mesher} script.
-When you run on a PC cluster, the script assumes that the nodes are
-named n001, n002, etc. If this is not the case, change the \texttt{tr
--d `n'} line in the script. You may also need to edit the last command
-at the end of the script that invokes the \texttt{mpirun} command.
-See Chapter~\ref{cha:Scheduler} for information about running the
-code on a system with a scheduler, e.g., LSF.
-
-Mesher output is provided in the \texttt{OUTPUT\_FILES} directory
-in \texttt{output\_mesher.txt}; this file provides lots of details
-about the mesh that was generated. Please note that the mesher suggests
-a time step \texttt{DT} to run the solver with. The mesher output
-file also contains a table about the quality of the mesh to indicate
-possible problems with the distortions of elements. Alternatively,
-output can be directed to the screen instead by uncommenting a line
-in \texttt{constants.h}:
-\begin{lyxcode}
-!~uncomment~this~to~write~messages~to~the~screen~
-
-!~integer,~parameter~::~IMAIN~=~ISTANDARD\_OUTPUT
-\end{lyxcode}
-To control the quality of the mesh, check the standard output (either
-on the screen or in the \texttt{OUTPUT\_FILES} directory in \texttt{output\_mesher.txt})
-and analyze the skewness histogram of mesh elements to make sure no
-element has a skewness above approximately 0.75, otherwise the mesh
-is of poor quality (and if even a single element has a skewness value
-above 0.80, then you must definitely improve the mesh). To draw the
-skewness histogram on the screen, type \texttt{gnuplot plot\_mesh\_quality\_histogram.gnu}.
-
-\chapter{\label{cha:Creating-Distributed-Databases}Creating the Distributed
-Databases}
-
-After using \texttt{xmeshfem3D} or \texttt{xdecompose\_mesh}, the
-next step in the workflow is to compile \texttt{xgenerate\_}~\\
- \texttt{databases}. This program is going to create all the missing
-information needed by the SEM solver.
-
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=0.6\textwidth]{figures/workflow}
-\par\end{centering}
-
-\caption{Schematic workflow for a SPECFEM3D Cartesian simulation. The executable
-\texttt{xgenerate\_databases} creates the GLL mesh points and assigns
-specific model parameters.}
-
-\label{fig:workflow.databases}
-\end{figure}
-
-\noindent
-In the main directory, type
-\begin{lyxcode}
-{\small make~xgenerate\_databases}{\small \par}
-\end{lyxcode}
-Input for the program is provided through the main parameter file
-\texttt{Par\_file}, which resides in the subdirectory \texttt{DATA}.
-Please note that \texttt{xgenerate\_databases} must be called directly
-from the main directory, as most of the binaries of the package.
-
-\section{\label{cha:Main-Parameter}Main parameter file \texttt{Par\_file}}
-
-Before running \texttt{xgenerate\_databases}, a number of parameters
-need to be set in the main parameter \texttt{Par\_file} located in
-the subdirectory \texttt{DATA}:
-\begin{description}
-\item [{\texttt{SIMULATION\_TYPE}}] is set to 1 for forward simulations,
-2 for adjoint simulations (see Section \ref{sec:Adjoint-simulation-finite})
-and 3 for kernel simulations (see Section \ref{sec:Finite-Frequency-Kernels}).
-\item [{\texttt{SAVE\_FORWARD}}] is only set to \texttt{.true.} for a forward
-simulation with the last frame of the simulation saved, as part of
-the finite-frequency kernel calculations (see Section \ref{sec:Finite-Frequency-Kernels}).
-For a regular forward simulation, leave \texttt{SIMULATION\_TYPE}
-and \texttt{SAVE\_FORWARD} at their default values.
-\item [{\texttt{\noun{UTM\_PROJECTION\_ZONE}}}] UTM projection zone in
-which your model resides, only valid when \texttt{SUPPRESS\_UTM\_}~\\
- \texttt{PROJECTION} is \texttt{.false.}.
-\item [{\texttt{SUPPRESS\_UTM\_PROJECTION}}] set to be \texttt{.false.}
-when your model range is specified in the geographical coordinates,
-and needs to be \texttt{.true.} when your model is specified in a
-cartesian coordinates. \noun{UTM projection zone in which your simulation
-region resides.}
-\item [{\texttt{NPROC}}] The number of MPI processors, each one is assigned
-one slice of the whole mesh.
-\item [{\texttt{NSTEP}}] The number of time steps of the simulation. This
-controls the length of the numerical simulation, i.e., twice the number
-of time steps requires twice as much CPU time. This feature is not
-used at the time of generating the distributed databases but is required
-for the solver, i.e., you may change this parameter after running
-\texttt{xgenerate\_databases}.
-\item [{\texttt{DT}}] The length of each time step in seconds. This feature
-is not used at the time of generating the distributed databases but
-is required for the solver. Please see also Section~\ref{sec:Choosing-the-Time-Step}
-for further details.
-\item [{\texttt{NGNOD}}] The number of nodes for 2D and 3D shape functions
-for hexahedra. We use either 8-node mesh elements (bricks) or 27-node
-elements. If you use the internal mesher, the only option is 8-node
-bricks (27-node elements are not supported). \texttt{CUBIT} does not
-support HEX27 elements either (it can generate them, but they are
-flat, i.e. identical to HEX8). To generate HEX27 elements with curvature
-properly taken into account, you can use \texttt{Gmsh} \url{http://geuz.org/gmsh/}
-\item [{\texttt{MODEL}}] Must be set to one of the following:
-
-\begin{description}
-\item [{\textmd{Models defined by mesh parameters:}}] ~
-
-\begin{description}
-\item [{\texttt{default}}] Uses model parameters as defined by meshing
-procedures described in the previous Chapter~\ref{cha:Mesh-Generation}.
-\end{description}
-
-\item [{\textmd{1D~models~with~real~structure:}}] ~
-
-\begin{description}
-\item [{\texttt{1D\_prem}}] Isotropic version of the spherically symmetric
-Preliminary Reference Earth Model (PREM) \citep{DzAn81}.
-\item [{\texttt{1D\_socal}}] A standard isotropic 1D model for Southern
-California.
-\item [{\texttt{1D\_cascadia}}] Isotropic 1D profile for the Cascadia region.
-\end{description}
-
-\item [{\textmd{Fully~3D~models:}}] ~
-
-\begin{description}
-\item [{\texttt{aniso}}] For a user-specified fully anisotropic model.
-Parameters are set up in routines located in file \texttt{model\_aniso.f90}
-in directory \texttt{src/generate\_databases/}. See Chapter~\ref{cha:-Changing-the}
-for a discussion on how to specify your own 3D model.
-
-\item [{\texttt{external}}] For a user-specified isotropic model which
-uses externally defined model parameters. Uses external model definitions
-set up in routines located in file \texttt{model\_external\_values.f90}
-in directory \texttt{src/generate\_databases/}. Please modify these
-generic template routines to use your own model definitions.
-\item [{\texttt{gll}}] For a user-specified isotropic model which uses
-external binary files for $v_{p}$, $v_{s}$ and $\rho$. Binary files
-are given in the same format as when outputted by the \texttt{xgenerate\_databases}
-executable when using option \texttt{SAVE\_MESH\_FILES}. These binary
-files define the model parameters on all GLL points which can be used
-for iterative inversion procedures.
-\item [{\texttt{salton\_trough}}] A 3D $V_{p}$ model for Southern California.
-Users must provide the corresponding data file \texttt{regrid3\_vel\_p.bin}
-in directory \texttt{DATA/st\_3D\_block\_harvard/}.
-\item [{\texttt{tomo}}] For a user-specified 3D isotropic model which uses
-a tomographic model file \texttt{tomographic\_model.xyz} in directory
-\texttt{DATA}. See Section \ref{sec:Using-tomographic}, for a discussion
-on how to specify your own 3D tomographic model.
-\end{description}
-\end{description}
-\item [{\texttt{APPROXIMATE\_OCEAN\_LOAD}}] Set to \texttt{.true.} if the
-effect of the oceans on seismic wave propagation should be incorporated
-based upon the (rough) approximate treatment discussed in \citet{KoTr02b}.
-This feature is inexpensive from a numerical perspective, both in
-terms of memory requirements and CPU time. This approximation is accurate
-at periods of roughly 20~s and longer. At shorter periods the effect
-of water phases/reverberations is not taken into account, even when
-the flag is on. If you want to model the effect of a fluid-solid model
-at short periods, then set this flag to \texttt{.false.} and mesh
-the fluid layer explicitly in your mesher, so that it is computed
-accurately and without this approximation.
-\item [{\texttt{TOPOGRAPHY}}] This feature is only effective if \texttt{APPROXIMATE\_OCEAN\_LOAD}
-is set to \texttt{.true.}. Set to \texttt{.true.} if topography and
-bathymetry should be read in based upon the topography file specified
-in the main constants file \texttt{constants.h} found in subdirectory
-\texttt{src/shared/} to evaluate elevations. If not set, elevations
-will be read from the numerical mesh.
-\item [{\texttt{ATTENUATION}}] Set to \texttt{.true.} if attenuation should
-be incorporated. Turning this feature on increases the memory requirements
-significantly (roughly by a factor of~1.5), and is numerically fairly
-expensive. See \citet{KoTr99,KoTr02a} for a discussion on the implementation
-of attenuation based upon standard linear solids. Please note that
-the Vp- and Vs-velocities of your model are given for a reference
-frequency. To change this reference frequency, you change the value
-of \texttt{ATTENUATION\_f0\_REFERENCE} in the main constants file
-\texttt{constants.h} found in subdirectory \texttt{src/shared/}.
-\item [{\texttt{FULL\_ATTENUATION\_SOLID}}] Set to \texttt{.true.} to add
-$Q_{\kappa}$ attenuation to the simulations in addition to $Q_{\mu}$
-attenuation when the \texttt{ATTENUATION} flag above is on. Should
-be off in almost all cases for seismic wave propagation, since $Q_{\kappa}$
-is almost always negligible in Earth models. However, this parameter
-can be useful for instance for ocean acoustics simulations, when $Q_{\kappa}$
-is not negligible in the solid layer at the sea bottom.
-\item [{\texttt{ANISOTROPY}}] Set to \texttt{.true.} if you want to use
-an anisotropy model. Please see the file \texttt{model\_aniso.f90}
-in subdirectory \texttt{src/generate\_databases/} for the current
-implementation of anisotropic models.
-\item [{\texttt{TOMOGRAPHY\_PATH}}] Directory in which the tomography files
-are stored for using external tomographic Earth models (please read
-Chapter \ref{cha:-Changing-the} and Section \ref{sec:Using-tomographic}
-`Using external tomographic Earth models' for further details.).
-\item [{\texttt{USE\_OLSEN\_ATTENUATION}}] Set to \texttt{.true.} if you
-want to use the attenuation model that scaled from the S-wave speed
-model using Olsen's empirical relation (see \citet{OlDaBr03}).
-\item [{\texttt{OLSEN\_ATTENUATION\_RATIO}}] Determines the Olsen's constant
-in Olsen's empirical relation (see \citet{OlDaBr03}).
-\item [{\texttt{PML\_CONDITIONS}}] Set to \texttt{.true.} to turn on C-PML
-boundary conditions for a regional simulation. Both fluids and elastic
-solids are supported. Note that xmeshfem3d generated meshes do not support C-PML yet.
-\item [{\texttt{PML\_INSTEAD\_OF\_FREE\_SURFACE}}] Set to \texttt{.true.}
-to turn on C-PML boundary conditions on the top surface instead of
-the usual free surface.
-\item [{\texttt{f0\_FOR\_PML}}] Determines the dominant frequency that
-will be used in the calculation of PML damping profiles; this should
-be set to the same (or similar) dominant frequency as that of the
-source that you will use in your simulation. If you plan to use a
-Dirac source, then use the dominant frequency of the source wavelet
-with which you plan to convolve your seismograms later on in post-processing.
-\item [{\texttt{STACEY\_ABSORBING\_CONDITIONS}}] Set to \texttt{.true.}
-to turn on Clayton-Enquist absorbing boundary conditions (see \citet{KoTr99}).
-In almost all cases it is much better to use CPML absorbing layers
-(see the options above) and leave this flag to \texttt{.false.}.
-\item [{\texttt{STACEY\_INSTEAD\_OF\_FREE\_SURFACE}}] Set to \texttt{.true.}
-to turn on absorbing boundary conditions on the top surface which
-by default constitutes a free surface of the model.
-\item [{\texttt{CREATE\_SHAKEMAP}}] Set this flag to \texttt{.true.} to
-create a ShakeMap\textregistered{}, i.e., a peak ground velocity map
-of the maximum absolute value of the two horizontal components of
-the velocity vector.
-\item [{\texttt{MOVIE\_SURFACE}}] Set to \texttt{.false.}, unless you want
-to create a movie of seismic wave propagation on the Earth's surface.
-Turning this option on generates large output files. See Section~\ref{sec:Movies}
-for a discussion on the generation of movies. This feature is only
-relevant for the solver.
-\item [{\texttt{MOVIE\_TYPE}}] Set this flag to 1 to show the top surface
-(tomography + oceans) only, to 2 to show all external faces of the
-mesh (i.e. topography + vertical edges + bottom) in shakemaps and
-surface movies.
-\item [{\texttt{MOVIE\_VOLUME}}] Set to \texttt{.false.}, unless you want
-to create a movie of seismic wave propagation in the Earth's interior.
-Turning this option on generates huge output files. See Section~\ref{sec:Movies}
-for a discussion on the generation of movies. This feature is only
-relevant for the solver.
-\item [{\texttt{SAVE\_DISPLACEMENT}}] Set this flag to \texttt{.true.}
-if you want to save the displacement instead of velocity for the movie
-frames.
-\item [{\texttt{USE\_HIGHRES\_FOR\_MOVIES}}] Set this flag to \texttt{.true.}
-if you want to save the values at all the NGLL grid points for the
-movie frames.
-\item [{\texttt{NTSTEP\_BETWEEN\_FRAMES}}] Determines the number of timesteps
-between movie frames. Typically you want to save a snapshot every
-100 timesteps. The smaller you make this number the more output will
-be generated! See Section~\ref{sec:Movies} for a discussion on the
-generation of movies. This feature is only relevant for the solver.
-\item [{\texttt{HDUR\_MOVIE}}] Determines the half duration of the source
-time function for the movie simulations. When this parameter is set
-to be 0, a default half duration that corresponds to the accuracy
-of the simulation is provided. Otherwise, it adds this half duration
-to the half duration specified in the source file \texttt{CMTSOLUTION},
-thus simulates longer periods to make the movie images look smoother.
-\item [{\texttt{SAVE\_MESH\_FILES}}] Set this flag to \texttt{.true.} to
-save ParaView \urlwithparentheses{www.paraview.org} mesh files for
-subsequent viewing. Turning the flag on generates large (distributed)
-files in the \texttt{LOCAL\_PATH} directory. See Section~\ref{sec:Mesh-graphics}
-for a discussion of mesh viewing features.
-\item [{\texttt{LOCAL\_PATH}}] Directory in which the distributed databases
-will be written. Generally one uses a directory on the local disk
-of the compute nodes, although on some machines these databases are
-written on a parallel (global) file system (see also the earlier discussion
-of the \texttt{LOCAL\_PATH\_IS\_ALSO\_GLOBAL} flag in Chapter~\ref{cha:Getting-Started}).
-\texttt{xgenerate\_databases} generates the necessary databases in
-parallel, one set for each of the \texttt{NPROC} slices that constitutes
-the mesh (see Figure~\ref{fig:mount.partitions} and Figure~\ref{fig:For-parallel-computing}).
-After the executable finishes, you can log in to one of the compute
-nodes and view the contents of the \texttt{LOCAL\_PATH} directory
-to see the (many) files generated by \texttt{xgenerate\_databases}.
-Please note that the \texttt{LOCAL\_PATH} directory should already
-contain the output files of the partitioner, i.e. from \texttt{xdecompose\_mesh}
-or \texttt{xmeshfem3D}.
-\item [{\texttt{NTSTEP\_BETWEEN\_OUTPUT\_INFO}}] This parameter specifies
-the interval at which basic information about a run is written to
-the file system (\texttt{timestamp{*}} files in the \texttt{OUTPUT\_FILES}
-directory). If you have access to a fast machine, set \texttt{NTSTEP\_BETWEEN\_OUTPUT\_INFO}
-to a relatively high value (e.g., at least 100, or even 1000 or more)
-to avoid writing output text files too often. This feature is not
-used at the time of meshing. One can set this parameter to a larger
-value than the number of time steps to avoid writing output during
-the run.
-\item [{\texttt{NTSTEP\_BETWEEN\_OUTPUT\_SEISMOS}}] This parameter specifies
-the interval at which synthetic seismograms are written in the \texttt{LOCAL\_PATH}
-directory. If a run crashes, you may still find usable (but shorter
-than requested) seismograms in this directory. On a fast machine set
-\texttt{NTSTEP\_BETWEEN\_OUTPUT\_SEISMOS} to a relatively high value
-to avoid writing to the seismograms too often. This feature is only
-relevant for the solver.
-\item [{\texttt{USE\_FORCE\_POINT\_SOURCE}}] Turn this flag on to use a
-(tilted) \texttt{FORCESOLUTION} force point source instead of a \texttt{CMTSOLUTION}
-moment-tensor source. When the force source does not fall exactly
-at a grid point, the solver interpolates the force between grid points
-using Lagrange interpolants. This can be useful e.g. for oil industry
-foothills simulations in which the source is a vertical force, normal
-force, tilted force, or an impact etc. Note that in the \texttt{FORCESOLUTION}
-file, you will need to edit the East, North and vertical components
-of an arbitrary (not necessarily unitary, the code will normalize it automatically) direction vector of the force vector;
-thus refer to Appendix~\ref{cha:Coordinates} for the orientation
-of the reference frame. This vector is made unitary internally in
-the solver and thus only its direction matters here; its norm is ignored
-and the norm of the force used is the factor force source times the
-source time function.
-When using this option, by default the code can locate the force source
-anywhere between mesh points in order to honor its exact location; this
-is more precise than using the closest GLL mesh point, but it is also a bit slower.
-If needed, you can change that default behavior and force the code to use the closest
-GLL mesh point instead by setting flag \texttt{USE\_BEST\_LOCATION} to \texttt{.false.} instead of \texttt{.true.} in file
-\texttt{src/shared/constants.h.in} and running the \texttt{configure} script again and recompiling the code.
-\item [{\texttt{USE\_RICKER\_TIME\_FUNCTION}}] Turn this flag on to use
-a Ricker source time function instead of the source time functions
-set by default to represent a (tilted) \texttt{FORCESOLUTION} force
-point source or a \texttt{CMTSOLUTION} moment-tensor source. Originally,
-if a \texttt{CMTSOLUTION} moment-tensor source is used, a (pseudo)
-Heaviside step function with a very short half duration is defined
-for elastic cases to represent the permanent slip on the fault while
-in the acoustic case a Gaussian source time function with a similarly
-short half duration is defined to physically describe actions within
-the fluid. Otherwise, if a \texttt{FORCESOLUTION} force source is
-used, a (pseudo) Dirac delta source time function is defined by default.
-Any other source-time function may then be obtained by convolution.
-\item [{\texttt{PRINT\_SOURCE\_TIME\_FUNCTION}}] Turn this flag on to print
-information about the source time function in the file \texttt{OUTPUT\_FILES/plot\_source\_time\_function.txt}.
-This feature is only relevant for the solver.
-
-\item [{\texttt{NUMBER\_OF\_SIMULTANEOUS\_RUNS}}] adds the ability to run several calculations (several earthquakes)
-in an embarrassingly-parallel fashion from within the same run;
-this can be useful when using a very large supercomputer to compute
-many earthquakes in a catalog, in which case it can be better from
-a batch job submission point of view to start fewer and much larger jobs,
-each of them computing several earthquakes in parallel.
-To turn that option on, set parameter NUMBER\_OF\_SIMULTANEOUS\_RUNS to a value greater than 1.
-To implement that, we create NUMBER\_OF\_SIMULTANEOUS\_RUNS MPI sub-communicators,
-each of them being labeled "my\_local\_mpi\_comm\_world", and we use them
-in all the routines in "src/shared/parallel.f90", except in MPI\_ABORT() because in that case
-we need to kill the entire run.
-When that option is on, of course the number of processor cores used to start
-the code in the batch system must be a multiple of NUMBER\_OF\_SIMULTANEOUS\_RUNS,
-all the individual runs must use the same number of processor cores,
-which as usual is NPROC in the Par\_file,
-and thus the total number of processor cores to request from the batch system
-should be NUMBER\_OF\_SIMULTANEOUS\_RUNS * NPROC.
-All the runs to perform must be placed in directories called run0001, run0002, run0003 and so on
-(with exactly four digits).
-
-Imagine you have 10 independent calculations to do, each of them on 100 cores; you have three options:
-
-1/ submit 10 jobs to the batch system
-
-2/ submit a single job on 1000 cores to the batch, and in that script create a sub-array of jobs to start 10 jobs,
-each running on 100 cores (see e.g. http://www.schedmd.com/slurmdocs/job\_array.html )
-
-3/ submit a single job on 1000 cores to the batch, start SPECFEM3D on 1000 cores, create 10 sub-communicators,
-cd into one of 10 subdirectories (called e.g. run0001, run0002,... run0010) depending on the sub-communicator
-your MPI rank belongs to, and run normally on 100 cores using that sub-communicator.
-
-The option NUMBER\_OF\_SIMULTANEOUS\_RUNS implements 3/.
-
-\item [{\texttt{BROADCAST\_SAME\_MESH\_AND\_MODEL}}]:
-if we perform simultaneous runs in parallel, if only the source and receivers vary between these runs
-but not the mesh nor the model (velocity and density) then we can also read the mesh and model files
-from a single run in the beginning and broadcast them to all the others; for a large number of simultaneous
-runs for instance when solving inverse problems iteratively this can DRASTICALLY reduce I/Os to disk in the solver
-(by a factor equal to NUMBER\_OF\_SIMULTANEOUS\_RUNS), and reducing I/Os is crucial in the case of huge runs.
-Thus, always set this option to .true. if the mesh and the model are the same for all simultaneous runs.
-In that case there is no need to duplicate the mesh and model file database (the content of the DATABASES\_MPI
-directories) in each of the run0001, run0002,... directories, it is sufficient to have one in run0001
-and the code will broadcast it to the others).
-
-TODO / future work to do: currently the BROADCAST\_SAME\_MESH\_AND\_MODEL option assumes to have the (master) mesh files in run0001/DATABASES\_MPI or run0001/OUTPUT\_FILES/DATABASES\_MPI.
-However, for adjoint runs you still need a DATABASES\_MPI folder in each of the sub-runs directories, e.g.
-run0002/DATABASES\_MPI, etc.
-to store the forward wavefields, kernels etc.
-of each sub-run.
-This would not be needed for forward simulations.
-
-TODO / future work to do: the sensitivity kernel summing and smoothing tools in directory src/tomography are currently not ported to this new option to do many runs simultaneously, only the solver (src/specfem3d) is.
-Thus these tools should work, but in their current version will need to be run for each simulation result independently.
-More precisely, the current kernel summing and smoothing routines work fine, with the exception that you need to move out the mesh files (and also the parameters).
-This works because these routines consider multiple runs by design.
-You simply have to provide them the directories where the kernels are.
-
-\item [{\texttt{GPU\_MODE}}] Turn this flag on to use GPUs.
-
-\item [\texttt{ADIOS\_ENABLED}] Turn this flag on to enable ADIOS. If set to \texttt{.false.}, subsequent ADIOS
-parameters will not be considered.
-\item [\texttt{ADIOS\_FOR\_DATABASES}] Turn this flag on to use ADIOS for xmeshfem3D output and
-xgenerate\_database input.
-\item [\texttt{ADIOS\_FOR\_MESH}]  Turn this flag on to use ADIOS for generated databases.
-\item [\texttt{ADIOS\_FOR\_FORWARD\_ARRAYS}] Turn this flag on to read and write forward arrays using ADIOS.
-\item [\texttt{ADIOS\_FOR\_KERNELS}] Turn this flag on to produce ADIOS kernels that can later be visualized with the ADIOS version of combine\_vol\_data.
-
-\end{description}
-
-The present version of SPECFEM can handle fully saturated porous simulations, Christina Morency implemented Biot equation.
-But the code cannot calculate partially saturated cases in its state.
-Christina Morency is presently working on a dual porosity, dual permeability formulation, type Pride and Berryman, for an other project, but it will not be available for some time.\\
-
-The way we prescribe material property for porous material in SPECFEM3D is as follow:
-We use a file name nummaterial\_poroelastic\_file, which is located in the directory MESH, the format is as follow:\\
- rhos rhof phi c kxx kxy kxz kyy kyz kzz Ks Kf Kfr etaf mufr\\
-where\\
-rho\_s = solid density\\
-rho\_f = fluid density\\
-phi = porosity\\
-tort = tortuosity\\
-kxx = xx component of permeability tensor\\
-kxy = xy,yx components of permeability tensor\\
-kyy = yy component of permeability tensor\\
-kxz = xz,zx components of permeability tensor\\
-kzz = zz component of permeability tensor\\
-kappa\_s = solid bulk modulus\\
-kappa\_f= fluid bulk modulus\\
-kappa\_fr= frame bulk modulus\\
-eta\_f = fluid viscosity\\
-mu\_fr = frame shear modulus.\\
-
-Using an external mesh (for instance coming from CUBIT/TRELIS), poroelastic materials have the ID number 3, while 1 is elastic, 2 is acoustic ?
-see the example in the package: examples/homogeneous\_poroelastic.
-
-If you use PML, the mesh elements that belong to the PML layers can
-be acoustic or elastic, but not viscoelastic nor poroelastic. Then,
-when defining your model, you should define these absorbing elements
-as either acoustic or elastic. In you forget to do that, the code
-will fix the problem by automatically converting the viscoelastic
-or poroelastic PML elements to elastic. This means that strictly speaking
-the PML layer will not be perfectly matched any more, since the physical
-model will change from viscoelastic or poroelastic to elastic at the
-entrance of the PML, but in practice this is sufficient and produces
-only tiny / negligible spurious reflections.
-
-If you use PML and an external mesh (created using an external meshing tool
-such as CUBIT/TRELIS or similar), try to have elements inside the PML as regular as possible,
-i.e. ideally non-deformed cubes obtained by `extrusion' of regular surface mesh elements meshing the
-outer edges of the computational domain without PML; by doing so, the PMLs obtained will be far more stable
-in time (PML being weakly unstable from a mathematical point of view, very deformed mesh elements
-inside the PMLs can trigger instabilities much more quickly).
-
-If you use PML and an external tomographic velocity and density model,
-you should be careful because mathematically a PML cannot handle heterogeneities
-along the normal to the PML edge inside the PML layer. This comes
-from the fact that the damping profile that is defined assumes a constant
-velocity and density model along the normal direction.
-
-Thus, you need to modify your velocity and density model in order
-for it to be 1D inside the PML, as shown in Figure~\ref{fig:modify_external_velocity_model_to_use_PML}.
-
-This applies to the bottom layer as well; there you should make sure
-that your model is 1D and thus constant along the vertical direction.
-
-To summarize, only use a 3D velocity and density model inside the
-physical region, and in all the PML layers extend it by continuity
-from its values along the inner PML edge.
-
-%%
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=3in]{figures/how_to_use_PML_when_a_tomographic_velocity_model_is_used}
-\par\end{centering}
-
-\caption{How to modify your external 3D velocity and density model in order
-to use PML. Such a modification is not needed when using Stacey absorbing
-boundary conditions (but such conditions are significantly less efficient).}
-
-\label{fig:modify_external_velocity_model_to_use_PML}
-\end{figure}
-
-%%
-
-\section{\label{sec:Choosing-the-Time-Step}Choosing the time step \texttt{DT}}
-
-The parameter \texttt{DT} sets the length of each time step in seconds.
-The value of this parameter is crucial for the stability of the spectral-element
-simulation. Your time step \texttt{DT} will depend on the minimum
-ratio between the distance $h$ of neighboring mesh points and the
-wave speeds $v$ defined in your model. The condition for the time
-step $\Delta t$ is:
-\begin{lyxcode}
-$\Delta t<C~\mathrm{min}_{\Omega}(~h/v~)$
-\end{lyxcode}
-where $C$ is the so-called Courant number and $\Omega$ denotes the
-model volume. The distance $h$ depends on the mesh element size and
-the number of GLL points \texttt{NGLL} specified in the main constants
-file \texttt{constants.h} located in the \texttt{src/shared/} subdirectory.
-The wave speed $v$ is determined based on your model's P- (or S-)
-wave speed values.
-
-The database generator \texttt{xgenerate\_databases}, as well as the
-internal mesher \texttt{xmeshfem3D}, are trying to evaluate the value
-of $\Delta t$ for empirically chosen Courant numbers $C\sim0.3$.
-If you used the mesher \texttt{xmeshfem3D} to generate your mesh,
-you should set the value suggested in \texttt{OUTPUT\_FILES/output\_mesher.txt}
-file, which is created after the mesher completed. In case you used
-CUBIT to create the mesh, you might use an arbitrary value when running
-\texttt{xgenerate\_databases} and then use the value suggested in
-the ~\\
- \texttt{OUTPUT\_FILES/output\_mesher.txt} file after the database
-generation completed. Note that the implemented Newmark time scheme
-uses this time step globally, thus your simulations become more expensive
-for very small mesh elements in high wave-speed regions. Please be
-aware of this restriction when constructing your mesh in Chapter~\ref{cha:Mesh-Generation}.
-
-\chapter{\label{cha:Running-the-Solver}Running the Solver \texttt{xspecfem3D}}
-
-Now that you have successfully generated the databases, you are ready
-to compile the solver. In the main directory, type
-\begin{lyxcode}
-{\small make~xspecfem3D}{\small \par}
-\end{lyxcode}
-Please note that \texttt{xspecfem3D} must be called directly from
-the main directory, as most of the binaries of the package.\\
-
-\noindent
-The solver needs three input files in the \texttt{DATA} directory
-to run:
-\begin{description}
-\item [{\texttt{Par\_file}}] the main parameter file which was discussed
-in detail in the previous Chapter~\ref{cha:Creating-Distributed-Databases},
-\item [{\texttt{CMTSOLUTION}} or {\texttt{FORCESOLUTION}}] the earthquake
-source parameter file or the force source parameter file, and
-\item [{\texttt{STATIONS}}] the stations file.
-\end{description}
-Most parameters in the \texttt{Par\_file} should be set prior to running
-the databases generation. Only the following parameters may be changed
-after running \texttt{xgenerate\_databases}:
-\begin{itemize}
-\item the simulation type control parameters: \texttt{SIMULATION\_TYPE}
-and \texttt{SAVE\_FORWARD}
-\item the time step parameters \texttt{NSTEP} and \texttt{DT}
-\item the absorbing boundary control parameter \texttt{PML\_CONDITIONS}
-on condition that the\\
- \texttt{PML\_INSTEAD\_OF\_FREE\_SURFACE} flag remains unmodified
-after running the databases generation.
-\item the movie control parameters \texttt{MOVIE\_SURFACE}, \texttt{MOVIE\_VOLUME},
-and \texttt{NTSTEPS\_BETWEEN\_FRAMES}
-\item the ShakeMap\textregistered{}option \texttt{CREATE\_SHAKEMAP}
-\item the output information parameters \texttt{MOVIE\_TYPE}, \texttt{NTSTEP\_BETWEEN\_OUTPUT\_INFO}
-and \texttt{NTSTEP\_BETWEEN\_OUTPUT\_}~\\
- \texttt{SEISMOS}
-\item the \texttt{PRINT\_SOURCE\_TIME\_FUNCTION} flags
-\end{itemize}
-Any other change to the \texttt{Par\_file} implies rerunning both
-the database generator \texttt{xgenerate\_databases} and the solver
-\texttt{xspecfem3D}.
-
-For any particular earthquake, the \texttt{CMTSOLUTION} file that
-represents the point source may be obtained directly from the Harvard
-Centroid-Moment Tensor (CMT) web page \urlwithparentheses{www.seismology.harvard.edu}.
-It looks like the example shown in Fig. \ref{fig:CMTSOLUTION-file}.
-%\begin{lyxcode}
-\begin{figure}[htp]
-\noindent
-\begin{centering}
-{\small \includegraphics[width=0.8\textwidth]{figures/Hollywood_CMT}
-}
-\par
-\end{centering}
-\caption{\texttt{CMTSOLUTION} file based on the
-format from the Harvard CMT catalog. \textbf{M} is the moment tensor,
-$M_{0}${\small {} }is the seismic moment, and $M_{w}$ is the moment
-magnitude.}
-\label{fig:CMTSOLUTION-file}
-\end{figure}
-%\end{lyxcode}
-The \texttt{CMTSOLUTION} file should be edited in the following way:
-\begin{itemize}
-\item Set the latitude or UTM $x$ coordinate, longitude or UTM $y$ coordinate, depth of the source (in km).
-\item Set the \texttt{time shift} parameter equal to $0.0$ (the solver
-will not run otherwise.) The time shift parameter would simply apply
-an overall time shift to the synthetics, something that can be done
-in the post-processing (see Section \ref{sec:Process-data-and-syn}).
-\item For point-source simulations (see finite sources, page \pageref{To-simulate-a})
-we recommend setting the source half-duration parameter \texttt{half
-duration} equal to zero, which corresponds to simulating a step source-time
-function, i.e., a moment-rate function that is a delta function. If
-\texttt{half duration} is not set to zero, the code will use a Gaussian
-(i.e., a signal with a shape similar to a `smoothed triangle', as
-explained in \citet{KoTr02a} and shown in Fig~\ref{fig:gauss.vs.triangle})
-source-time function with half-width \texttt{half duration}. We prefer
-to run the solver with \texttt{half duration} set to zero and convolve
-the resulting synthetic seismograms in post-processing after the run,
-because this way it is easy to use a variety of source-time functions
-(see Section \ref{sec:Process-data-and-syn}). \citet{KoTr02a} determined
-that the noise generated in the simulation by using a step source
-time function may be safely filtered out afterward based upon a convolution
-with the desired source time function and/or low-pass filtering. Use
-the serial code \texttt{convolve\_source\_timefunction.f90} and the
-script \texttt{convolve\_source\_timefunction.csh} for this purpose,
-or alternatively use signal-processing software packages such as SAC
-\urlwithparentheses{www.llnl.gov/sac}. Type
-
-\begin{lyxcode}
-{\small make~xconvolve\_source\_timefunction}{\small \par}
-\end{lyxcode}
-
-to compile the code and then set the parameter \texttt{hdur} in \texttt{convolve\_source\_timefunction.csh}
-to the desired half-duration.
-
-\item The zero time of the simulation corresponds to the center of the triangle/Gaussian,
-or the centroid time of the earthquake. The start time of the simulation
-is $t=-1.5*\texttt{half duration}$ (the 1.5 is to make sure the moment
-rate function is very close to zero when starting the simulation).
-To convert to absolute time $t_{\mathrm{abs}}$, set
-
-\begin{lyxcode}
-$t_{\mathrm{abs}}=t_{\mathrm{pde}}+\texttt{time shift}+t_{\mathrm{synthetic}}$
-\end{lyxcode}
-
-where $t_{\mathrm{pde}}$ is the time given in the first line of the
-\texttt{CMTSOLUTION}, \texttt{time shift} is the corresponding value
-from the original \texttt{CMTSOLUTION} file and $t_{\mathrm{synthetic}}$
-is the time in the first column of the output seismogram.
-
-\end{itemize}
-\begin{figure}[htp]
-\noindent \begin{centering}
-\includegraphics[width=3.5in]{figures/gauss_vs_triangle_mod}
-\par\end{centering}
-
-\caption{Comparison of the shape of a triangle and the Gaussian function actually
-used.}
-
-\label{fig:gauss.vs.triangle}
-\end{figure}
-
-If you know the earthquake source in strike/dip/rake format rather
-than in \texttt{CMTSOLUTION} format, use the C code \texttt{SPECFEM3D\_GLOBE/utils/strike\_dip\_rake\_to\_CMTSOLUTION.c}
-to convert it. The conversion formulas are given for instance in \citet{AkRi80}.
-Note that the \citet{AkRi80} convention is slightly different from
-the Harvard \texttt{CMTSOLUTION} convention (the sign of some components
-is different). The C code outputs both.
-
-Centroid latitude and longitude should be provided in geographical
-coordinates. The code converts these coordinates to geocentric coordinates~\citep{DaTr98}.
-Of course you may provide your own source representations by designing
-your own \texttt{CMTSOLUTION} file. Just make sure that the resulting
-file adheres to the Harvard CMT conventions (see Appendix~\ref{cha:Coordinates}).
-Note that the first line in the \texttt{CMTSOLUTION} file is the Preliminary
-Determination of Earthquakes (PDE) solution performed by the USGS
-NEIC, which is used as a seed for the Harvard CMT inversion. The PDE
-solution is based upon P waves and often gives the hypocenter of the
-earthquake, i.e., the rupture initiation point, whereas the CMT solution
-gives the `centroid location', which is the location with dominant
-moment release. The PDE solution is not used by our software package
-but must be present anyway in the first line of the file.
-
-\label{To-simulate-a}To simulate a kinematic rupture, i.e., a finite-source
-event, represented in terms of $N_{\mathrm{sources}}$ point sources,
-provide a \texttt{CMTSOLUTION} file that has $N_{\mathrm{sources}}$
-entries, one for each subevent (i.e., concatenate $N_{\mathrm{sources}}$
-\texttt{CMTSOLUTION} files to a single \texttt{CMTSOLUTION} file).
-At least one entry (not necessarily the first) must have a zero \texttt{time
-shift}, and all the other entries must have non-negative \texttt{time
-shift}. Each subevent can have its own half duration, latitude, longitude,
-depth, and moment tensor (effectively, the local moment-density tensor).
-
-Note that the zero in the synthetics does NOT represent the hypocentral
-time or centroid time in general, but the timing of the \textit{center}
-of the source triangle with zero \texttt{time shift} (Fig~\ref{fig:source_timing}).
-
-Although it is convenient to think of each source as a triangle, in
-the simulation they are actually Gaussians (as they have better frequency
-characteristics). The relationship between the triangle and the gaussian
-used is shown in Fig~\ref{fig:gauss.vs.triangle}. For finite fault
-simulations it is usually not advisable to use a zero half duration
-and convolve afterwards, since the half duration is generally fixed
-by the finite fault model.
-
-\vspace{1cm}
-
-\noindent The \texttt{FORCESOLUTION} file should be edited in the
-following way:
-\begin{itemize}
-\item Set the \texttt{time shift} parameter equal to $0.0$ (the solver
-will not run otherwise.) The time shift parameter would simply apply
-an overall time shift to the synthetics, something that can be done
-in the post-processing (see Section \ref{sec:Process-data-and-syn}).
-\item Set the \texttt{f0} parameter (the dominant frequency) of the Ricker source time
-function when {\texttt{USE\_RICKER\_TIME\_FUNCTION}} is turned on
-in the main parameter file {\texttt{Par\_file}}. In case that the
-solver uses a (pseudo) Dirac delta source time function to represent
-a force point source, a very short duration of five time steps is automatically set by default.
-\item Set the latitude or UTM $x$ coordinate, longitude or UTM $y$ coordinate, depth of the source (in km).
-\item Set the magnitude of the force source.
-\item Set the components of a (non-unitary) direction vector for the force
-source in the East/North/Vertical basis (see Appendix A for the orientation
-of the reference frame).
-\end{itemize}
-\noindent Where necessary, set a \texttt{FORCESOLUTION} file in the
-same way you configure a \texttt{CMTSOLUTION} file with $N_{\mathrm{sources}}$
-entries, one for each subevent (i.e., concatenate $N_{\mathrm{sources}}$
-\texttt{FORCESOLUTION} files to a single \texttt{FORCESOLUTION} file).
-At least one entry (not necessarily the first) must have a zero \texttt{time
-shift}, and all the other entries must have non-negative \texttt{time
-shift}. Each subevent can have its own half latitude, longitude, depth,
-\texttt{half duration} and force parameters.
-
-\begin{figure}[H]
-\noindent \begin{centering}
-{\small \includegraphics[width=5in]{figures/source_timing} }
-\par\end{centering}
-
-\caption{Example of timing for three sources. The center of the first source
-triangle is defined to be time zero. Note that this is NOT in general
-the hypocentral time, or the start time of the source (marked as tstart).
-The parameter \texttt{time shift} in the \texttt{CMTSOLUTION} file
-would be t1(=0), t2, t3 in this case, and the parameter \texttt{half
-duration} would be hdur1, hdur2, hdur3 for the sources 1, 2, 3 respectively.}
-
-{\small \label{fig:source_timing} }
-\end{figure}
-
-The solver can calculate seismograms at any number of stations for
-basically the same numerical cost, so the user is encouraged to include
-as many stations as conceivably useful in the \texttt{STATIONS} file,
-which looks like this:
-
-\begin{figure}[H]
-\noindent \begin{centering}
-{\small \includegraphics{figures/STATIONS_basin_explained} }
-\par\end{centering}
-
-\label{fig:Sample-STATIONS-file.}\caption{Sample \texttt{STATIONS} file. Station latitude and longitude should
-be provided in geographical coordinates. The width of the station
-label should be no more than 32 characters (see \texttt{MAX\_LENGTH\_STATION\_NAME}
-in the \texttt{constants.h} file), and the network label should be
-no more than 8 characters (see \texttt{MAX\_LENGTH\_NETWORK\_NAME}
-in the \texttt{constants.h} file).}
-\end{figure}
-
-Each line represents one station in the following format:
-\begin{lyxcode}
-{\small Station~Network~Latitude~(degrees)~Longitude~(degrees)~Elevation~(m)~burial~(m)~}{\small \par}
-\end{lyxcode}
-The solver \texttt{xspecfem3D} filters the list of stations in file
-\texttt{DATA/STATIONS} to exclude stations that are not located within
-the region given in the \texttt{Par\_file} (between \texttt{LATITUDE\_MIN}
-and \texttt{LATITUDE\_MAX} and between \texttt{LONGITUDE\_MIN} and
-\texttt{LONGITUDE\_MAX}). The filtered file is called \texttt{DATA/STATIONS\_FILTERED}.
-
-Solver output is provided in the \texttt{OUTPUT\_FILES} directory
-in the \texttt{output\_solver.txt} file. Output can be directed to
-the screen instead by uncommenting a line in \texttt{constants.h}:
-\begin{lyxcode}
-!~uncomment~this~to~write~messages~to~the~screen~
-
-!~integer,~parameter~::~IMAIN~=~ISTANDARD\_OUTPUT~
-\end{lyxcode}
-On PC clusters the seismogram files are generally written to the local
-disks (the path \texttt{LOCAL\_PATH} in the \texttt{Par\_file}) and
-need to be gathered at the end of the simulation.
-
-While the solver is running, its progress may be tracked by monitoring
-the `\texttt{\small timestamp{*}}{\small ' files in the ~}\\
-{\small{} }\texttt{\small OUTPUT\_FILES}{\small{} directory. These tiny
-files look something like this:}{\small \par}
-\begin{lyxcode}
-Time~step~\#~~~~~~~~~~10000~
-
-Time:~~~~~108.4890~~~~~~seconds~
-
-Elapsed~time~in~seconds~=~~~~1153.28696703911~
-
-Elapsed~time~in~hh:mm:ss~=~~~~~0~h~19~m~13~s~
-
-Mean~elapsed~time~per~time~step~in~seconds~=~~~~~0.115328696703911~
-
-Max~norm~displacement~vector~U~in~all~slices~(m)~=~~~~1.0789589E-02~
-\end{lyxcode}
-The \texttt{\small timestamp{*}}{\small{} files provide the }\texttt{\small Mean
-elapsed time per time step in seconds}{\small , which may be used
-to assess performance on various machines (assuming you are the only
-user on a node), as well as the }\texttt{\small Max norm displacement
-vector U in all slices~(m)}{\small . If something is wrong with the
-model, the mesh, or the source, you will see the code become unstable
-through exponentially growing values of the displacement and fluid
-potential with time, and ultimately the run will be terminated by
-the program. You can control the rate at which the timestamp files
-are written based upon the parameter }\texttt{\small NTSTEP\_BETWEEN\_OUTPUT\_INFO}{\small{}
-in the }\texttt{\small Par\_file}{\small .}{\small \par}
-
-Having set the \texttt{Par\_file} parameters, and having provided
-the \texttt{CMTSOLUTION} (or the \texttt{FORCESOLUTION}) and \texttt{STATIONS}
-files, you are now ready to launch the solver! This is most easily
-accomplished based upon the \texttt{go\_solver} script (See Chapter
-\ref{cha:Scheduler} for information about running through a scheduler,
-e.g., LSF). You may need to edit the last command at the end of the
-script that invokes the \texttt{mpirun} command. The \texttt{runall}
-script compiles and runs both \texttt{xgenerate\_databases} and \texttt{xspecfem3D}
-in sequence. This is a safe approach that ensures using the correct
-combination of distributed database output and solver input.
-
-It is important to realize that the CPU and memory requirements of
-the solver are closely tied to choices about attenuation (\texttt{ATTENUATION})
-and the nature of the model (i.e., isotropic models are cheaper than
-anisotropic models). We encourage you to run a variety of simulations
-with various flags turned on or off to develop a sense for what is
-involved.
-
-For the same model, one can rerun the solver for different events
-by simply changing the \texttt{CMTSOLUTION} or \texttt{FORCESOLUTION}
-file, or for different stations by changing the \texttt{STATIONS}
-file. There is no need to rerun the \texttt{xgenerate\_databases}
-executable. Of course it is best to include as many stations as possible,
-since this does not add to the cost of the simulation.
-
-We have also added the ability to run several calculations (several earthquakes)
-in an embarrassingly-parallel fashion from within the same run;
-this can be useful when using a very large supercomputer to compute
-many earthquakes in a catalog, in which case it can be better from
-a batch job submission point of view to start fewer and much larger jobs,
-each of them computing several earthquakes in parallel.
-To turn that option on, set parameter \texttt{NUMBER\_OF\_SIMULTANEOUS\_RUNS}
-to a value greater than 1 in file setup/constants.h.in before
-configuring and compiling the code.
-When that option is on, of course the number of processor cores used to start
-the code in the batch system must be a multiple of \texttt{NUMBER\_OF\_SIMULTANEOUS\_RUNS},
-all the individual runs must use the same number of processor cores,
-which as usual is \texttt{NPROC} in the input file \texttt{DATA/Par\_file},
-and thus the total number of processor cores to request from the batch system
-should be \texttt{NUMBER\_OF\_SIMULTANEOUS\_RUNS $\times$ NPROC}.
-All the runs to perform must be placed in directories called \texttt{run0001}, \texttt{run0002}, \texttt{run0003} and so on (with exactly four digits)
-and you must create a link from the root directory of the code
-to the first copy of the executable programs by typing \texttt{ln -s run0001/bin bin}.
-
-\chapter{\label{cha:Fault-Kinematics-Dynamics}Kinematic and dynamic fault
-sources}
-
-SPECFEM3D can handle finite fault sources of two kinds:
-\begin{enumerate}
-\item \emph{Kinematic}: the spatio-temporal distribution of slip rate is
-prescribed along the fault surface
-\item \emph{Dynamic}: a friction law and initial stresses are prescribed
-on the fault, a spontaneous rupture process is computed.
-\end{enumerate}
-
-\section{\label{sec:Mesh-Generation-with}Mesh Generation with Split Nodes}
-
-Faults need to be handled in a special way during mesh generation.
-A fault surface must lie at the interface between elements (the mesh
-must honor the fault surfaces). Moreover, a fault is made of two surfaces
-in contact. Each of these two surfaces needs a separate set of nodes.
-This approach is known as \textquotedbl{}split nodes\textquotedbl{}.
-Currently faults can only be run with \texttt{xdecompose\_mesh} and
-\texttt{CUBIT}. \texttt{xmeshfem3D} is not yet ready to handle faults.\\
-To facilitate the mesh generation with split nodes in CUBIT, we need
-to separate the two fault surfaces by a small distance, effectively
-creating a tiny opening of the fault (Figure \ref{fig:examples.splitnodes},
-\ref{fig:examples.splitnodes-surfacetrace}). Note that only the interior
-of the fault must be opened, its edges must remain closed (except
-the edge on the free surface). The fault is automatically closed later
-by SPECFEM3D.\\
-
-Here is an example CUBIT script to generate a mesh with split nodes
-for a buried vertical strike-slip fault:
-\begin{lyxcode}
-~reset~~\\
-~brick~x~10~y~10~z~10~~\\
-~webcut~volume~all~with~plane~xplane~~\\
-~webcut~volume~all~with~plane~yplane~~\\
-~webcut~volume~all~with~plane~xplane~offset~3~~\\
-~webcut~volume~all~with~plane~zplane~offset~3~~\\
-~webcut~volume~all~with~plane~zplane~offset~-3~~\\
-~imprint~all~~\\
-~merge~all~~\\
-~unmerge~surf~160~~\\
-~mesh~vol~all~~\\
-~set~node~constraint~off~~\\
-~node~in~surf~168~move~X~0~Y~0.01~Z~0~~\\
-~node~in~surf~160~move~X~0~Y~-0.01~Z~0~\\
-
-\end{lyxcode}
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[scale=0.3]{figures/faultmesh} \\
- \includegraphics[scale=0.3]{figures/surf168} \includegraphics[scale=0.3]{figures/splitnodes}
-\par\end{centering}
-
-\caption{Screenshots of the CUBIT example for the embedded fault with split
-nodes: The entire mesh is decomposed into several volumes bounding
-the fault (top), the fault surface 168 is shown as orange squares
-(middle) and split nodes of the fault shown as orange dots(bottom).
-Note that only interior nodes of the fault are split while those on
-the edges of the fault surface are touching each other.}
-
-\label{fig:examples.splitnodes}
-\end{figure}
-
-\begin{lyxcode}
-
-\end{lyxcode}
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[scale=0.55]{figures/splitnodes_surfacetrace} \\
-
-\par\end{centering}
-
-\caption{Screenshots of a CUBIT example showing split nodes for a fault reaching
-the surface. Surface trace of the fault is shown in orange. Note that
-the edges of the fault are not split while the interior nodes are
-offset by a small distance on either side of the fault}
-
-\label{fig:examples.splitnodes-surfacetrace}
-\end{figure}
-
-The CUBIT scripts ({*}.jou and {*}.py) in the directory \texttt{EXAMPLES}
-generate more complicated meshes. The {*}.py files are Python scripts
-that execute CUBIT commands and use the CUBIT-python interface for
-SPECFEM3D (see next section). The Python language allows to define
-and manipulate variables to parameterize the mesh. Alternatively,
-the Python script can call a CUBIT journal file ({*}.jou), which looks
-like the example above. Variables can be defined and manipulated there
-using the $\mathtt{APREPRO}$ language built in CUBIT.\\
-
-Note that you should avoid gaps in the list of indices of mesh objects
-with the following CUBIT command:
-\begin{lyxcode}
-$\mathtt{compress\: ids\: hex\: face\: edge\: node}$
-\end{lyxcode}
-(otherwise you will get a segmentation fault during domain decomposition)
-
-\section{CUBIT-Python Scripts for Faults}
-
-The mesh generated in CUBIT needs to be processed and exported in
-a format compatible with SPECFEM3D. This is achieved in the Python
-scripts by calling the Python-CUBIT interface functions defined in
-the CUBIT directory:
-\begin{enumerate}
-\item Function $\mathtt{define\_bc}$ (or $\mathtt{boundary\_definition.py}$)
-must be called to set up the absorbing boundaries database
-\item Function $\mathtt{fault\_input}$ must be called once for each fault
-to set up the fault database
-\item Function $\mathtt{cubit2specfem3d.export2SPECFEM3D}$ must be called
-at the very end of the script to export the mesh in a SPECFEM3D format.
-\end{enumerate}
-The functions in \#1 and \#3 are for the bulk and are documented in
-Section \ref{subsec:Exporting-the-Mesh}. We focus here on \#2:
-\begin{description}
-\item [{Function:}] $\mathtt{fault\_input}$
-\item [{Purpose:}] export a fault mesh database from CUBIT to a SPECFEM3D-compliant
-file
-\item [{Syntax:}] fault\_input(id\_fault, ids\_surf\_1, ids\_surf\_2)
-\item [{Inputs:}]~\end{description}
-\begin{itemize}
-\item id\_fault integer index assigned to the fault. (The user must number
-all the faults, starting at 1, with unit increments)
-\item ids\_surf\_1 list of CUBIT ids of all surfaces that form side 1 of
-the fault.
-\item ids\_surf\_2 list of CUBIT ids of all surfaces that form side 2 of
-the fault. (The user must decide which side of the fault is side 1.
-This choice affects the sign conventions of fault quantities as explained
-in Section \ref{sec:Sign-Convention-for}).\end{itemize}
-\begin{description}
-\item [{Outputs:}] file $\mathtt{fault\_file\_X.dat}$, where X is the
-fault id (id\_fault).
-\item [{Example:}] For the example in Section \ref{sec:Mesh-Generation-with}:\end{description}
-\begin{lyxcode}
-$\hphantom{}$A1~=~{[}168{]}
-
-$\hphantom{}$A2~=~{[}160{]}
-
-$\hphantom{}$fault\_input(1,A1,A2)
-\end{lyxcode}
-
-\section{Examples}
-
-The package includes examples, the SCEC benchmark problems:
-\begin{itemize}
-\item TPV5, a planar vertical strike-slip fault
-\item TPV14 and TPV15, vertical strike-slip fault system with a fault branch
-\item TPV102 and TPV103, rate-and-state benchmarks
-\item TPV22 and TPV23, step-over benchmarks
-\item TPV16, heterogenous initial stress
-\end{itemize}
-and
-\begin{itemize}
-\item Splay fault models from Wendt et al. (2009)\\
-
-\end{itemize}
-Read the documents in the directory \texttt{EXAMPLES/*/description}.
-They contain a description of the example and additional instructions
-to run it. Visualize the results with the Matlab scripts in the directory
-$\mathtt{EXAMPLES/*/post}$
-
-\section{\label{sec:Sign-Convention-for}Sign Convention for Fault Quantities}
-
-During mesh generation, the fault is defined by two surfaces in contact.
-Let's denote as \textquotedbl{}\emph{side 1}\textquotedbl{} the SECOND
-surface declared by the user in the call to the python function \textquotedbl{}\emph{fault\_input}\textquotedbl{},
-and the FIRST surface as \textquotedbl{}\emph{side 2}\textquotedbl{}.
-The local coordinate system on the fault is defined as the right-handed
-coordinate system defined by (strike, dip, normal), where \textquotedbl{}\emph{normal}\textquotedbl{}
-is the normal vector outgoing from side 1, \textquotedbl{}\emph{dip}\textquotedbl{}
-is parallel to the along-dip direction pointing downwards, and \textquotedbl{}\emph{strike}\textquotedbl{}
-is the horizontal along-strike vector such that the system is right-handed.\\
-
-Slip is defined as displacement on side 2 minus displacement on side
-1. In the local coordinate system on the fault, positive along-strike
-slip is right-lateral and positive along-dip slip is thrust if side
-1 is on the hanging wall (normal faulting if side 1 is on the foot
-wall).\\
-
-Traction is defined as the stress induced on side 1 by side 2, which
-is the stress tensor times the normal vector outgoing from side 1.
-In the local coordinate system on the fault, the normal traction is
-negative in compression, positive along-strike traction generates
-right-lateral slip and positive along-dip traction generates thrust
-slip if side 1 is on the hanging wall (normal faulting if side 1 is
-on the foot wall).
-
-\section{Input Files}
-
-Three additional input files are required in the \texttt{DATA/} directory
-for dynamic and kinematic faults. They are $\mathtt{Par\_file\_faults}$,
-$\mathtt{FAULT\_STATIONS}$ and $\mathtt{input\_file.txt}$ (optional).
-If the former does not exist, the code assumes that there are no faults.\\
-
-\begin{lyxlist}{00.00.0000}
-\item [{\textbf{DATA/Par\_file\_faults}}] contains parameters of the fault.
-The first part of this file has a strict format:
-
-\begin{description}
-\item [{Line}] 1: Number of faults (NF)
-\item [{Lines}] 2 to NF+1: Kelvin Voigt damping (in seconds) for each fault.
-(See below how to set this parameter)
-\item [{Line}] NF+2: Type of simulation (1=dynamic , 2 = kinematic)
-\item [{Line}] NF+3: Number of time steps between updates of the time series
-outputs at selected fault points (see DATA/FAULT\_STATIONS), usually
-a large number (100s or 1000s). Note that the sampling rate of the
-time series is usually much higher.
-\item [{Line}] NF+4: Number of time steps between fault snapshot outputs
-(quantities at every fault point exported at regular times), usually
-a large number (100s or 1000s).
-\item [{Line}] NF+5: Slip velocity threshold below which frictional healing
-is set (friction coefficient is reset to its static value). If this
-value is negative healing is disabled.
-\item [{Line}] NF+6: Slip velocity threshold to define the rupture front.
-Only used for outputs.
-\end{description}
-
-The rest of this file is made of namelist input blocks (see \textquotedbl{}namelist\textquotedbl{}
-in a Fortran 9x manual). The input for each fault has the following
-sequence (arguments in {[}brackets{]} are optional):
-
-\&\textbf{BEGIN\_FAULT} /
-
-\&\textbf{INIT\_STRESS} S1, S2, S3 {[},n1, n2, n3{]} /
-
-followed by (n1+n2+n3) \&\textbf{DIST2D} blocks
-
-\&\textbf{SWF} mus, mud, dc {[}, nmus, nmud, ndc{]} /
-
-followed by (nmus+nmud+ndc) \&\textbf{DIST2D} blocks\\
-
-\begin{lyxlist}{00.00.0000}
-\item [{\&\textbf{INIT\_STRESS}}] input block sets the initial fault stresses
-relative to the foot-wall side of the fault. Initial stresses are
-composed of a constant background value possibly overwritten in prescribed
-regions by heterogeneous distributions (see \&\textbf{DIST2D} blocks
-below):
-
-\begin{description}
-\item [{S1}] = constant background value of along-strike shear stress (positive
-in the usual strike direction)
-\item [{S2}] = constant background value of along-dip shear (positive is
-down-dip, normal faulting)
-\item [{S3}] = constant background value of normal stress (negative in
-compression)
-\item [{n1}] = number of heterogeneous items for along-strike shear stress
-{[}default is 0{]}
-\item [{n2}] = number of heterogeneous items for along-dip shear stress
-{[}default is 0{]}
-\item [{n3}] = number of heterogeneous items for normal stress {[}default
-is 0{]}
-\end{description}
-\item [{\&\textbf{SWF}}] input block sets the slip-weakening friction parameters
-of the fault:
-
-\begin{description}
-\item [{mus}] = constant background value of static friction coefficient
-\item [{mud}] = constant background value of dynamic friction coefficient
-\item [{dc}] = constant background value of critical slip-weakening distance
-\item [{nmus}] = number of heterogeneous items for static friction coefficient
-{[}default is 0{]}
-\item [{nmud}] = number of heterogeneous items for dynamic friction coefficient
-{[}default is 0{]}
-\item [{ndc}] = number of heterogeneous items for critical slip-weakening
-distance {[}default is 0{]}\\
-
-\end{description}
-\item [{\&\textbf{DIST2D}}] input blocks modify (overwrite) the value of
-a fault parameter by a heterogeneous spatial distribution:
-
-\&\textbf{DIST2D} shape='$\mathtt{square}$', val, xc, yc, zc, l /
-
-$\;$sets a constant value (val) within a cube with center (xc,yc,zc)
-and edge size l.
-
-\&\textbf{DIST2D} shape='$\mathtt{rectangle}$', val, xc, yc, zc,
-lx, ly, lz /
-
-$\;$sets a constant value (val) within a rectangular prism with center
-(xc,yc,zc) and edge sizes (lx,ly,lz).
-
-\&\textbf{DIST2D} shape='$\mathtt{rectangle-taper}$', val, valh,
-xc, yc, zc, lx, ly, lz /
-
-$\;$sets a vertical linear gradient within a rectangular prism with
-center (xc,yc,zc) and edge sizes (lx,ly,lz). Values vary linearly
-as a function of vertical position z between value val at z = zc-lz/2
-and value valh at z = zc+lz/2 .
-
-\&\textbf{DIST2D} shape='$\mathtt{circular}$', val, xc, yc, zc, r
-/
-
-$\;$sets a constant value (val) within a sphere with center (xc,yc,zc)
-and radius r. \\
-
-\end{lyxlist}
-\end{lyxlist}
-
-\begin{lyxlist}{00.00.0000}
-\item [{\textbf{DATA/FAULT\_STATIONS}}] Stations in the fault plane.
-
-\textbf{Line 1} : number of stations.
-
-\textbf{Line 2 to end}: 5 columns: X, Y, Z (-depth), station name,
-fault-id
-
-The fault-id identifies the fault that contains the station. It is
-the index of appearance in the faults list after line 2 of Par\_file\_faults
-
-\end{lyxlist}
-
-\begin{lyxlist}{00.00.0000}
-\item [{\textbf{DATA/input\_file.txt}}] Heterogeneous stresses and friction
-parameters are documented in page 10 of
-
-\begin{lyxlist}{00.00.0000}
-\item [{$\mathtt{EXAMPLES/tpv16/description/TPV16\_17\_Description\_v03.pdf}$}]~
-\end{lyxlist}
-
-To activate this feature, in fault\_solver\_dynamic.f90 set \textbf{TPV16}
-= .true. then re-compile the code.
-
-\end{lyxlist}
-
-\section{\label{sec:Setting-the-Kelvin-Voigt}Setting the Kelvin-Voigt Damping
-Parameter }
-
-The purpose of the Kelvin-Voigt viscosity in the dynamic fault solver
-is to damp spurious oscillations generated by the fault slip at frequencies
-that are too high to be resolved by the mesh. The viscosity $\mathtt{eta}$
-(in seconds) depends on the size of the elements on the fault. Here
-is how to set it:
-\begin{enumerate}
-\item Determine the average linear size of the elements on the fault plane,
-$\mathtt{h\_fault}$. Usually this value is prescribed by the user
-during mesh generation. Otherwise it can be found by inspection of
-the mesh inside the CUBIT GUI.
-\item Use the Matlab function $\mathtt{utils/critical\_timestep.m}$ to
-compute\\
- $\mathtt{dtc\_fault}=\mathtt{critical\_timestep\left(c_{p},h\_fault,ngll\right)}$.
-\\
-This is the critical time step in an elastic medium for a hypothetical
-element of cubic shape with size equal to $\mathtt{h\_fault}$.
-\item Set $\mathtt{eta}$ in $\mathtt{Par\_file\_faults}$ to (0.1 to 0.3)$\ensuremath{\times}\mathtt{dtc\_fault}$.
-A larger $\mathtt{eta}$ damps high-frequencies more aggressively
-but it might also affect lower frequencies and rupture speed.
-\end{enumerate}
-Viscosity reduces numerical stability: the critical timestep in a
-simulation with Kelvin-Voigt damping needs to be smaller than that
-in a purely elastic simulation. Here is how to set the time step accordingly:
-\begin{enumerate}
-\item Run a test simulation without viscosity ($\mathtt{eta}$=0 and only
-a few time steps)
-\item Look for the \textquotedbl{}maximum suggested time step\textquotedbl{}
-in \texttt{OUTPUT\_FILES/output\_mesher.txt}. This is the critical
-timestep of a purely elastic simulation, $\mathtt{dtc\_bulk}$.
-\item Reset the timestep of the simulation with a Kelvin-Voigt material
-to a value smaller than\\
- $\mathtt{dtc\_kv}=\mathtt{eta}\left(\sqrt{1+\mathtt{dtc\_bulk^{2}}/\mathtt{eta^{2}}}-1\right)$
-\end{enumerate}
-Note that in general $\mathtt{dtc\_bulk}$ is smaller than $\mathtt{dtc\_fault}$,
-because elements off the fault might be smaller or more distorted
-than element faces on the fault.
-
-\section{Output Files }
-
-Several output files are saved in \texttt{OUTPUT\_FILES/}:\\
-
-\begin{enumerate}
-\item Seismograms for each station on the fault plane given in \texttt{DATA/FAUL\_STATIONS}.
-One output file is generated for each station, named after the station.
-The files are ascii and start with a header (22 lines long) followed
-by a data block with the following format, one line per time sample:\\
-
-Column 1 = Time (s)
-
-Column 2 = horizontal right-lateral slip (m)
-
-Column 3 = horizontal right-lateral slip rate (m/s)
-
-Column 4 = horizontal right-lateral shear stress (MPa)
-
-Column 5 = vertical up-dip slip (m)
-
-Column 6 = vertical up-dip slip rate (m/s)
-
-Column 7 = vertical up-dip shear stress (MPa)
-
-Column 8 = normal stress (MPa)\\
-
-The stresses are relative to the footwall side of the fault (this
-convention controls their sign, but not their amplitude). Slip is
-defined as displacement of the hanging wall relative to the footwall.
-
-\item Seismograms at stations in the bulk (out of the fault plane) given
-in \texttt{DATA/STATIONS}.
-\item Rupture time files are named $\mathtt{Rupture\_time\_FAULT}$-$\mathtt{id}$.
-One file is generated for each fault. The files are ascii and start
-with a header (12 lines long) followed by a data block with the following
-format, one line per fault node:
-
-Column 1 = horizontal coordinate, distance along strike (m)
-
-Column 2 = vertical coordinate, distance down-dip (m)
-
-Column 3 = rupture time (s)
-
-\item Fault quantities (slip, slip rate, stresses, etc) at regular times
-are stored in binary data files called $\mathtt{Snapshot\#it\#.bin}$,
-where \#it\# is the timestep number. These can be read in Matlab with
-the function $\mathtt{utils/FSEM3D\_snapshot.m}$
-\end{enumerate}
-
-\section{Post-processing and Visualization }
-
-Some Matlab functions for post-processing and visualization are included
-in directory $\mathtt{utils}$. The functions are internally documented
-(see their matlab help).\\
-
-\texttt{FSEM3D\_snapshot} reads a fault data snapshot\\
-
-The directories \texttt{EXAMPLES/*/post} contain additional Matlab
-scripts to generate figures specific to each example.
-
-\chapter{\label{cha:Adjoint-Simulations}Adjoint Simulations}
-
-Adjoint simulations are generally performed for two distinct applications.
-First, they can be used in point source moment-tensor inversions, or source imaging for
-earthquakes with large ruptures such as the Lander's earthquake \citep{WaHe94}.
-Second, they can be used to generate finite-frequency sensitivity
-kernels that are a critical part of tomographic inversions based upon
-3D reference models \citep{TrTaLi05,LiTr06,TrKoLi08,LiTr08}. In either
-case, source parameter or velocity structure updates are sought to
-minimize a specific misfit function (e.g., waveform or traveltime
-differences), and the adjoint simulation provides a means of computing
-the gradient of the misfit function and further reducing it in successive
-iterations. Applications and procedures pertaining to source studies
-and finite-frequency kernels are discussed in Sections~\ref{sec:Adjoint-simulation-sources}
-and \ref{sec:Adjoint-simulation-finite}, respectively. The two related
-parameters in the \texttt{Par\_file} are \texttt{SIMULATION\_TYPE}
-(1, 2 or 3) and the \texttt{SAVE\_FORWARD} (boolean).
-
-\section{\label{sec:Adjoint-simulation-sources}Adjoint Simulations for Sources
-Only (not for the Model)}
-
-When a specific misfit function between data and synthetics is minimized to invert
-for earthquake source parameters, the gradient of the misfit function
-with respect to these source parameters can be computed by placing
-time-reversed seismograms at the receivers as virtual sources
-in an adjoint simulation. Then the value of the gradient is obtained
-from the adjoint seismograms recorded at the original earthquake location.
-\begin{enumerate}
-\item \textbf{Prepare the adjoint sources} \label{enu:Prepare-the-adjoint}
-
-\begin{enumerate}
-\item First, run a regular forward simlation (\texttt{SIMULATION\_TYPE =
-1} and \texttt{SAVE\_FORWARD = .false.}). You can automatically set
-these two variables using the \texttt{\small utils/change\_simulation\_type.pl}{\small{}
-script:}{\small \par}
-
-\begin{lyxcode}
-utils/change\_simulation\_type.pl~-f~
-\end{lyxcode}
-
-and then collect the recorded seismograms at all the stations given
-in \texttt{DATA/STATIONS}.
-
-\item Then select the stations for which you want to compute the time-reversed
-adjoint sources and run the adjoint simulation, and compile them into
-the \texttt{DATA/STATIONS\_ADJOINT} file, which has the same format
-as the regular \texttt{DATA/STATIONS} file.
-
-\begin{itemize}
-\item Depending on what type of misfit function is used for the source inversion,
-adjoint sources need to be computed from the original recorded seismograms
-for the selected stations and saved in a sub-directory called \texttt{SEM/}
-in the root directory of the code, with the format \texttt{NT.STA.BX?.adj}, where \texttt{NT},
-\texttt{STA} are the network code and station name given in the \texttt{DATA/STATIONS\_ADJOINT}
-file, and \texttt{BX?} represents the component name of a particular
-adjoint seismogram. Please note that the band code can change depending
-on your sampling rate (see Appendix~\ref{cha:channel-codes} for
-further details).
-\item The adjoint seismograms are in the same format as the original seismogram
-(\texttt{NT.STA.BX?.sem?}), with the same start time, time interval
-and record length.
-\end{itemize}
-\item Notice that even if you choose to time reverse only one component
-from one specific station, you still need to supply all three components
-because the code is expecting them (you can set the other two components
-to be zero).
-\item Also note that since time-reversal is done in the code itself, no
-explicit time-reversing is needed for the preparation of the adjoint
-sources, i.e., the adjoint sources are in the same forward time sense
-as the original recorded seismograms.
-\end{enumerate}
-\item \textbf{Set the related parameters and run the adjoint simulation}\\
- In the \texttt{DATA/Par\_file}, set the two related parameters to
-be \texttt{SIMULATION\_TYPE = 2} and \texttt{SAVE\_FORWARD = .false.}.
-More conveniently, use the scripts \texttt{utils/change\_simulation\_type.pl}
-to modify the \texttt{Par\_file} automatically (\texttt{change\_simulation\_type.pl
--a}). Then run the solver to launch the adjoint simulation.
-\item \textbf{Collect the seismograms at the original source location}
-
-After the adjoint simulation has completed successfully, collect the
-seismograms from \texttt{LOCAL\_PATH}.
-\begin{itemize}
-\item These adjoint seismograms are recorded at the locations of the original
-earthquake sources given by the \texttt{DATA/CMTSOLUTION} file, and
-have names of the form \texttt{NT.S?????.S??.sem} for the six-component
-strain tensor (\texttt{SNN,SEE,SZZ,SNE,SNZ,SEZ}) at these locations,
-and ~\\
- \texttt{NT.S?????.BX?.sem} for the three-component displacements
-(\texttt{BXN,BXE,BXZ}) recorded at these locations.
-\item \texttt{S?????} denotes the source number; for example, if the original
-\texttt{CMTSOLUTION} provides only a point source, then the seismograms
-collected will start with \texttt{S00001}.
-\item These adjoint seismograms provide critical information for the computation
-of the gradient of the misfit function.
-\end{itemize}
-\end{enumerate}
-
-\section{\label{sec:Adjoint-simulation-finite}Adjoint Simulations for Finite-Frequency
-Kernels (Kernel Simulation)}
-
-Finite-frequency sensitivity kernels are computed in two successive
-simulations (please refer to \citet{LiTr06} and \citet{TrKoLi08}
-for details).
-\begin{enumerate}
-\item \textbf{Run a forward simulation with the state variables saved at
-the end of the simulation}
-
-Prepare the \texttt{\small CMTSOLUTION}{\small{} and }\texttt{\small STATIONS}{\small{}
-files, set the parameters }\texttt{\small SIMULATION\_TYPE}{\small {}
-}\texttt{\small =}{\small {} }\texttt{\small 1}{\small{} and }\texttt{\small SAVE\_FORWARD
-=}{\small {} }\texttt{\small .true.}{\small{} in the }\texttt{\small Par\_file}{\small{}
-(}\texttt{\small change\_simulation\_type -F}{\small ), and run the
-solver.}{\small \par}
-\begin{itemize}
-\item Notice that attenuation is not implemented yet for the computation
-of finite-frequency kernels; therefore set \texttt{ATTENUATION = .false.}
-in the \texttt{Par\_file}.
-\item We also suggest you modify the half duration of the \texttt{CMTSOLUTION}
-to be similar to the accuracy of the simulation (see Equation \ref{eq:shortest_period})
-to avoid too much high-frequency noise in the forward wavefield, although
-theoretically the high-frequency noise should be eliminated when convolved
-with an adjoint wavefield with the proper frequency content.
-\item This forward simulation differs from the regular simulations (\texttt{\small SIMULATION\_TYPE}{\small {}
-}\texttt{\small =}{\small {} }\texttt{\small 1}{\small{} and }\texttt{\small SAVE\_FORWARD}{\small {}
-}\texttt{\small =}{\small {} }\texttt{\small .false.}{\small ) described
-in the previous chapters in that the state variables for the last
-time step of the simulation, including wavefields of the displacement,
-velocity, acceleration, etc., are saved to the }\texttt{\small LOCAL\_PATH}{\small{}
-to be used for the subsequent simulation. }{\small \par}
-\item For regional simulations, the files recording the absorbing boundary
-contribution are also written to the \texttt{LOCAL\_PATH} when \texttt{SAVE\_FORWARD
-= .true.}.
-\end{itemize}
-\item \textbf{Prepare the adjoint sources}
-
-The adjoint sources need to be prepared the same way as described
-in the Section \ref{enu:Prepare-the-adjoint}.
-\begin{itemize}
-\item In the case of travel-time finite-frequency kernel for one source-receiver
-pair, i.e., point source from the \texttt{CMTSOLUTION}, and one station
-in the \texttt{STATIONS\_ADJOINT} list, we supply a sample program
-in \texttt{utils/adjoint\_sources/traveltime/xcreate\_adjsrc\_traveltime}
-to cut a certain portion of the original displacement seismograms
-and convert it into the proper adjoint source to compute the finite-frequency
-kernel.
-
-\begin{lyxcode}
-xcreate\_adjsrc\_traveltime~t1~t2~ifile{[}0-5{]}~E/N/Z-ascii-files~{[}baz{]}
-\end{lyxcode}
-
-where \texttt{t1} and \texttt{t2} are the start and end time of the
-portion you are interested in, \texttt{ifile} denotes the component
-of the seismograms to be used (0 for all three components, 1 for East,
-2 for North, and 3 for vertical, 4 for transverse, and 5 for radial
-component), \texttt{E/N/Z-ascii-files} indicate the three-component
-displacement seismograms in the right order, and \texttt{baz} is the
-back-azimuth of the station. Note that \texttt{baz} is only supplied
-when \texttt{ifile} = 4 or 5.
-
-\item Similarly, a sample program to compute adjoint sources for amplitude
-finite-frequency kernels may be found in \texttt{utils/adjoint\_sources/amplitude}
-and used in the same way as described for traveltime measurements
-
-\begin{lyxcode}
-xcreate\_adjsrc\_amplitude~t1~t2~ifile{[}0-5{]}~E/N/Z-ascii-files~{[}baz{]}.
-\end{lyxcode}
-\end{itemize}
-\item \textbf{Run the kernel simulation}
-
-With the successful forward simulation and the adjoint source ready
-in the \texttt{SEM/} directory, set \texttt{SIMULATION\_TYPE
-= 3} and \texttt{SAVE\_FORWARD = .false.} in the \texttt{Par\_file}
-(you can use \texttt{change\_simulation\_type.pl -b}), and rerun the
-solver.
-\begin{itemize}
-\item The adjoint simulation is launched together with the back reconstruction
-of the original forward wavefield from the state variables saved from
-the previous forward simulation, and the finite-frequency kernels
-are computed by the interaction of the reconstructed forward wavefield
-and the adjoint wavefield.
-\item The back-reconstructed seismograms at the original station locations
-are saved to the \texttt{LOCAL\_PATH} at the end of the kernel simulations,
-and can be collected to the local disk.
-\item These back-constructed seismograms can be compared with the time-reversed
-original seismograms to assess the accuracy of the backward reconstruction,
-and they should match very well.
-\item The arrays for density, P-wave speed and S-wave speed kernels are
-also saved in the \texttt{LOCAL\_PATH} with the names \texttt{proc??????\_rho(alpha,beta)\_kernel.bin},
-where \texttt{proc??????} represents the processor number, \texttt{rho(alpha,beta)}
-are the different types of kernels.
-\end{itemize}
-\item \textbf{Run the anisotropic kernel simulation}
-
-Instead of the kernels for the isotropic wave speeds, you can also
-compute the kernels for the 21 independent components $C_{IJ},\, I,J=1,...,6$
-(using Voigt's notation) of the elastic tensor in the cartesian coordinate
-system. This is done by setting \texttt{ANISOTROPIC\_KL} \texttt{=}
-\texttt{.true.} in \texttt{constants.h} before compiling the package.
-The definition of the parameters $C_{IJ}$ in terms of the corresponding
-components $c_{ijkl},ijkl,i,j,k,l=1,2,3$ of the elastic tensor in
-cartesian coordinates follows \citet{ChTr07}. The 21 anisotropic
-kernels are saved in the \texttt{LOCAL\_PATH} in one file with the
-name of \texttt{proc??????\_cijkl\_kernel.bin} (with \texttt{proc??????}
-the processor number). The output kernels correspond to absolute perturbations
-$\delta C_{IJ}$ of the elastic parameters and their unit is in $s/GPa/km^{3}$.
-For consistency, the output density kernels with this option turned
-on are for a perturbation $\delta\rho$ (and not $\frac{\delta\rho}{\rho}$)
-and their unit is in s / (kg/m$^{3}$) / km$^{3}$. These `primary'
-anisotropic kernels can then be combined to obtain the kernels for
-different parameterizations of anisotropy. This can be done, for example,
-when combining the kernel files from slices into one mesh file (see
-Section~\ref{sec:Finite-Frequency-Kernels}).
-
-If \texttt{ANISOTROPIC\_KL} \texttt{=} \texttt{.true.} by additionally
-setting \texttt{ANISOTROPIC\_KL} \texttt{=} \texttt{.true.} in \texttt{constants.h}
-the package will save anisotropic kernels paramterized as velocities
-related to transverse isotropy based on the the Chen and Tromp parameters
-\citet{ChTr07}. The kernels are saved as relative perturbations for
-horizontal and vertical P and S velocities, $\alpha_{v},\alpha_{h},\beta_{v},\beta_{h}$.
-Explicit relations can be found in appendix B. of \citet{SiLiTrTr07b}
-
-The anisotropic kernels are only currently available for CPU mode.
-
-\end{enumerate}
-In general, the three steps need to be run sequentially to assure
-proper access to the necessary files. If the simulations are run through
-some cluster scheduling system (e.g., LSF), and the forward simulation
-and the subsequent kernel simulations cannot be assigned to the same
-set of computer nodes, the kernel simulation will not be able to access
-the database files saved by the forward simulation. Solutions for
-this dilemma are provided in Chapter~\ref{cha:Scheduler}. Visualization
-of the finite-frequency kernels is discussed in Section~\ref{sec:Finite-Frequency-Kernels}.
-
-%------------------------------------------------------------------------------------------------%
-
-\chapter{\label{cha:tomo}Doing tomography, i.e., updating the model based on the sensitivity kernels obtained}
-
-%------------------------------------------------------------------------------------------------%
-
-UNDER CONSTRUCTION (March 2015). Will be added around April / May 2015.
-
-\section{Tomographic inversion / imaging using the sensitivity kernels obtained}
-
-OLD VERSION OF THE SECTION, WILL SOON BE IMPROVED AND ENRICHED.\\
-
-One of the fundamental reasons for computing sensitivity kernels (Section~\ref{sec:Adjoint-simulation-finite})
-is to use them within a tomographic inversion. In other words, use
-recorded seismograms, make measurements with synthetic seismograms,
-and use the misfit between the two to iteratively improve the model
-described by (at least) $V_{{\rm p}}$, $V_{{\rm s}}$, and $\rho$
-as a function of space.
-
-Whatever misfit function you use for the tomographic inversion (you can find several
-examples in \citet{TrKoLi08} and \citet{TrTaLi05}), you will weight
-the sensitivity kernels with measurements. Furthermore, you will use
-as many measurements (stations, components, time windows) as possible
-per event; hence, we call these composite kernels ``event kernels,''
-which are volumetric fields representing the gradient of the misfit
-function with respect to one of the variables (\eg $V_{{\rm s}}$).
-The basic features of an adjoint-based tomographic inversion were
-illustrated in \citet{TaLiTr07} and \citet{TrKoLi08} using a conjugate-gradient
-algorithm; there are dozens of versions of gradient-based inversion
-algorithms that could alternatively be used. The tomographic inversion
-of \citet{TaLiMaTr09,TaLiMaTr2010} used SPECFEM3D Cartesian as well
-as several additional components which are also stored on the CIG
-svn server, described next.
-
-The directory containing external utilities for tomographic inversion using
-SPECFEM3D Cartesian (or other packages that evaluate misfit functions
-and gradients) is in directory \texttt{utils/ADJOINT\_TOMOGRAPHY\_TOOLS/}:
-\begin{verbatim}
-flexwin/     -- FLEXWIN algorithm for automated picking of time windows
-measure_adj/ -- reads FLEXWIN output file and makes measurements,
-                with the option for computing adjoint sources
-iterate_adj/ -- various tools for iterative inversion
-                (requires pre-computed "event kernels")
-\end{verbatim}
-This directory also contains
-a brief \verb+README+ file indicating the role of the three subdirectories,
-\verb+flexwin+ \citep{Maggi2009}, \verb+measure_adj+, and \verb+iterate_adj+.
-The components for making the model update are there; however, there
-are no explicit rules for computing the model update, just as with
-any optimization problem. There are options for computing a conjugate
-gradient step, as well as a source subspace projection step.
-
-The best single file to read is probably: \verb+ADJOINT_TOMO/iterate_adj/cluster/README+.
-
-%------------------------------------------------------------------------------------------------%
-
-\chapter{Noise Cross-correlation Simulations}
-
-%------------------------------------------------------------------------------------------------%
-
-Besides earthquake simulations, SPECFEM3D Cartesian includes functionality
-for seismic noise tomography as well. In order to proceed successfully
-in this chapter, it is critical that you have already familiarized
-yourself with procedures for meshing (Chapter \ref{cha:Mesh-Generation}),
-creating distributed databases (Chapter \ref{cha:Creating-Distributed-Databases}),
-running earthquake simulations (Chapters \ref{cha:Running-the-Solver})
-and adjoint simulations (Chapter \ref{cha:Adjoint-Simulations}).
-Also, make sure you read the article `Noise cross-correlation sensitivity
-kernels' \citep{trompetal2010}, in order to understand noise simulations
-from a theoretical perspective.
-
-\section{Input Parameter Files}
-
-As usual, the three main input files are crucial: \texttt{Par\_file},
-\texttt{CMTSOLUTION} and \texttt{STATIONS}. Unless otherwise specified,
-those input files should be located in directory \texttt{DATA/}.\\
-
-\texttt{CMTSOLUTION} is required for all simulations. At a first glance,
-it may seem unexpected to have it here, since the noise simulations
-should have nothing to do with the earthquake -- \texttt{CMTSOLUTION}.
-However, for noise simulations, it is critical to have no earthquakes.
-In other words, the moment tensor specified in \texttt{CMTSOLUTION}
-must be set to zero manually!\\
-
-\texttt{STATIONS} remains the same as in previous earthquake simulations,
-except that the order of receivers listed in \texttt{STATIONS} is
-now important. The order will be used to determine the `master' receiver,
-i.e., the one that simultaneously cross correlates with the others.\\
-
-\texttt{Par\_file} also requires careful attention. A parameter called
-\texttt{NOISE\_TOMOGRAPHY} has been added which specifies the type
-of simulation to be run. \texttt{NOISE\_TOMOGRAPHY} is an integer
-with possible values 0, 1, 2 and 3. For example, when \texttt{NOISE\_TOMOGRAPHY}
-equals 0, a regular earthquake simulation will be run. When it is
-1/2/3, you are about to run step 1/2/3 of the noise simulations respectively.
-Should you be confused by the three steps, refer to \citet{trompetal2010}
-for details.\\
-
-Another change to \texttt{Par\_file} involves the parameter \texttt{NSTEP}.
-While for regular earthquake simulations this parameter specifies
-the length of synthetic seismograms generated, for noise simulations
-it specifies the length of the seismograms used to compute cross correlations.
-The actual cross correlations are thus twice this length, i.e., $2~\mathrm{NSTEP}-1$.
-The code automatically makes the modification accordingly, if \texttt{NOISE\_TOMOGRAPHY}
-is not zero.\\
-
-There are other parameters in \texttt{Par\_file} which should be given
-specific values. For instance, since the first two steps for calculating
-noise sensitivity kernels correspond to forward simulations, \texttt{SIMULATION\_TYPE}
-must be 1 when \texttt{NOISE\_TOMOGRAPHY} equals 1 or 2. Also, we
-have to reconstruct the ensemble forward wavefields in adjoint simulations,
-therefore we need to set \texttt{SAVE\_FORWARD} to \texttt{.true.}
-for the second step, i.e., when \texttt{NOISE\_TOMOGRAPHY} equals
-2. The third step is for kernel constructions. Hence \texttt{SIMULATION\_TYPE}
-should be 3, whereas \texttt{SAVE\_FORWARD} must be \texttt{.false.}.\\
-
-Finally, for most system architectures, please make sure that \texttt{LOCAL\_PATH}
-in \texttt{Par\_file} is in fact local, not globally shared. Because
-we have to save the wavefields at the earth's surface at every time
-step, it is quite problematic to have a globally shared \texttt{LOCAL\_PATH},
-in terms of both disk storage and I/O speed.
-
-\section{Noise Simulations: Step by Step}
-
-Proper parameters in those parameter files are not enough for noise
-simulations to run. We have more parameters to specify: for example,
-the ensemble-averaged noise spectrum, the noise distribution etc.
-However, since there are a few `new' files, it is better to introduce
-them sequentially. In this section, standard procedures for noise
-simulations are described.
-
-\subsection{Pre-simulation}
-\begin{itemize}
-\item As usual, we first configure the software package using:
-
-\texttt{./configure FC=ifort MPIFC=mpif90}
-
-Use the following if SCOTCH is needed:
-
-\texttt{./configure FC=ifort MPIFC=mpif90 -{}-with-scotch-dir=/opt/scotch}\\
-
-\item Next, we need to compile the source code using:
-
-\texttt{make xgenerate\_databases}
-
-\texttt{make xspecfem3D} \\
-
-\item Before we can run noise simulations, we have to specify the noise
-statistics, e.g., the ensemble-averaged noise spectrum. Matlab scripts
-are provided to help you to generate the necessary file.
-
-\texttt{EXAMPLES/noise\_tomography/NOISE\_TOMOGRAPHY.m} (main program)\\
- \texttt{EXAMPLES/noise\_tomography/PetersonNoiseModel.m}
-
-In Matlab, simply run:
-
-\texttt{NOISE\_TOMOGRAPHY(NSTEP, DT, Tmin, Tmax, NOISE\_MODEL)}\\
-
-\texttt{DT} is given in \texttt{Par\_file}, but \texttt{NSTEP} is
-NOT the one specified in \texttt{Par\_file}. Instead, you have to
-feed $2~\mathrm{NSTEP}-1$ to account for the doubled length of cross
-correlations. \texttt{Tmin} and \texttt{Tmax} correspond to the period
-range you are interested in, whereas \texttt{NOISE\_MODEL} denotes
-the noise model you will be using. Details can be found in the Matlab
-script.
-
-After running the Matlab script, you will be given the following information
-(plus a figure in Matlab):
-
-\texttt{{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}}\\
- \texttt{the source time function has been saved in:}\\
- \texttt{/data2/yangl/3D\_NOISE/S\_squared} (note this path must be
-different)\\
- \texttt{S\_squared should be put into directory:}\\
- \texttt{OUTPUT\_FILES/NOISE\_TOMOGRAPHY/ in the SPECFEM3D Cartesian
-package}\\
-
-In other words, the Matlab script creates a file called \texttt{S\_squared},
-which is the first `new' input file we encounter for noise simulations.
-
-One may choose a flat noise spectrum rather than Peterson's noise
-model. This can be done easily by modifying the Matlab script a little.
-
-\item Create a new directory in the SPECFEM3D Cartesian package, name it
-as \texttt{OUTPUT\_FILES/NOISE\_TOMOGRAPHY/}. We will add some parameter
-files later in this folder.
-\item Put the Matlab-generated-file \texttt{S\_squared} in \texttt{OUTPUT\_FILES/NOISE\_TOMOGRAPHY/}.
-
-That's to say, you will have a file \texttt{OUTPUT\_FILES/NOISE\_TOMOGRAPHY/S\_squared}
-in the SPECFEM3D Cartesian package.
-
-\item Create a file called \texttt{OUTPUT\_FILES/NOISE\_TOMOGRAPHY/irec\_master\_noise}.
-Note that this file is located in directory \texttt{OUTPUT\_FILES/NOISE\_TOMOGRAPHY/}
-as well. In general, all noise simulation related parameter files
-go into that directory. \texttt{irec\_master\_noise} contains only
-one integer, which is the ID of the `master' receiver. For example,
-if this file contains 5, it means that the fifth receiver listed in
-\texttt{DATA/STATIONS} becomes the `master'. That's why we mentioned
-previously that the order of receivers in \texttt{DATA/STATIONS} is
-important.
-
-Note that in regional simulations, the \texttt{DATA/STATIONS} might
-contain receivers which are outside of our computational domains.
-Therefore, the integer in \texttt{irec\_master\_noise} is actually
-the ID in \texttt{DATA/STATIONS\_FILTERED} (which is generated by
-\texttt{bin/xgenerate\_databases}).
-
-\item Create a file called \texttt{OUTPUT\_FILES/NOISE\_TOMOGRAPHY/nu\_master}.
-This file holds three numbers, forming a (unit) vector. It describes
-which component we are cross-correlating at the `master' receiver,
-i.e., ${\hat{{\bf \nu}}}^{\alpha}$ in \citet{trompetal2010}. The
-three numbers correspond to E/N/Z components respectively. Most often,
-the vertical component is used, and in those cases the three numbers
-should be 0, 0 and 1.
-\item Describe the noise direction and distributions in \texttt{src/shared/noise\_tomography.f90}.
-Search for a subroutine called \texttt{noise\_distribution\_direction}
-in \texttt{noise\_tomography.f90}. It is actually located at the very
-beginning of \texttt{noise\_tomography.f90}. The default assumes vertical
-noise and a uniform distribution across the whole free surface of
-the model. It should be quite self-explanatory for modifications.
-Should you modify this part, you have to re-compile the source code.
-\end{itemize}
-
-\subsection{Simulations}
-
-With all of the above done, we can finally launch our simulations.
-Again, please make sure that the \texttt{LOCAL\_PATH} in \texttt{DATA/Par\_file}
-is not globally shared. It is quite problematic to have a globally
-shared \texttt{LOCAL\_PATH}, in terms of both disk storage and speed
-of I/O (we have to save the wavefields at the earth's surface at every
-time step).
-
-As discussed in \citet{trompetal2010}, it takes three steps/simulations
-to obtain one contribution of the ensemble sensitivity kernels:
-\begin{itemize}
-\item Step 1: simulation for generating wavefields\\
- \texttt{SIMULATION\_TYPE=1}\\
- \texttt{NOISE\_TOMOGRAPHY=1}\\
- \texttt{SAVE\_FORWARD} not used, can be either .true. or .false.
-\item Step 2: simulation for ensemble forward wavefields\\
- \texttt{SIMULATION\_TYPE=1}\\
- \texttt{NOISE\_TOMOGRAPHY=2}\\
- \texttt{SAVE\_FORWARD=.true.}
-\item Step 3: simulation for ensemble adjoint wavefields and sensitivity
-kernels\\
- \texttt{SIMULATION\_TYPE=3}\\
- \texttt{NOISE\_TOMOGRAPHY=3}\\
- \texttt{SAVE\_FORWARD=.false.}
-
-Note Step 3 is an adjoint simulation, please refer to previous chapters
-on how to prepare adjoint sources and other necessary files, as well
-as how adjoint simulations work.
-
-\end{itemize}
-Note that it is better to run the three steps continuously within
-the same job, otherwise you have to collect the saved surface movies
-from the old nodes/CPUs to the new nodes/CPUs. This process varies
-from cluster to cluster and thus cannot be discussed here. Please
-ask your cluster administrator for information/configuration of the
-cluster you are using.\\
-
-\subsection{Post-simulation}
-
-After those simulations, you have all stuff you need, either in the
-\texttt{OUTPUT\_FILES/} or in the directory specified by \texttt{LOCAL\_PATH}
-in \texttt{DATA/Par\_file} (which are most probably on local nodes).
-Collect whatever you want from the local nodes to your workstation,
-and then visualize them. This process is the same as what you may
-have done for regular earthquake simulations. Refer to other chapters
-if you have problems.\\
-
-Simply speaking, two outputs are the most interesting: the simulated
-ensemble cross correlations and one contribution of the ensemble sensitivity
-kernels.\\
-
-The simulated ensemble cross correlations are obtained after the second
-simulation (Step 2). Seismograms in \texttt{OUTPUT\_FILES/} are actually
-the simulated ensemble cross correlations. Collect them immediately
-after Step 2, or the Step 3 will overwrite them. Note that we have
-a `master' receiver specified by \texttt{irec\_master\_noise}, the
-seismogram at one station corresponds to the cross correlation between
-that station and the `master'. Since the seismograms have three components,
-we may obtain cross correlations for different components as well,
-not necessarily the cross correlations between vertical components.\\
-
-One contribution of the ensemble cross-correlation sensitivity kernels
-are obtained after Step 3, stored in the \texttt{DATA/LOCAL\_PATH}
-on local nodes. The ensemble kernel files are named the same as regular
-earthquake kernels.\\
-
-You need to run another three simulations to get the other contribution
-of the ensemble kernels, using different forward and adjoint sources
-given in \citet{trompetal2010}.
-
-\section{Example}
-
-In order to illustrate noise simulations in an easy way, one example
-is provided in \texttt{EXAMPLES/noise\_tomography/}. See \texttt{EXAMPLES/noise\_tomography/README}
-for explanations. \\
-
-Note, however, that they are created for a specific workstation (CLOVER@PRINCETON),
-which has at least 4 cores with `mpif90' working properly. \\
-
-If your workstation is suitable, you can run the example in \texttt{EXAMPLES/noise\_tomography/}
-using:\\
-
-\texttt{./pre-processing.sh}\\
-
-Even if this script does not work on your workstation, the procedure
-it describes is universal. You may review the whole process described
-in the last section by following the commands in \texttt{pre-processing.sh},
-which should contain enough explanations for all the commands.
-
-%------------------------------------------------------------------------------------------------%
-
-\chapter{Gravity integral calculations for the gravity field of an Earth model}
-
-%------------------------------------------------------------------------------------------------%
-
-SPECFEM can now compute the gravity field as well as its derivatives (i.e., gravity gradiometry)
-generated by any given 3D Earth model at a given elevation, i.e., on a given observation surface.
-That feature is still experimental but should work just fine.\\
-
-To see how it is implemented and to use it, type this in the root directory of the code:\\
-
-\texttt{grep -i GRAVITY\_INTEGRALS src/*/*}
-
-%------------------------------------------------------------------------------------------------%
-
-\chapter{Graphics}
-
-%------------------------------------------------------------------------------------------------%
-
-\section{\label{sec:Mesh-graphics}Meshes}
-
-In case you used the internal mesher \texttt{xmeshfem3D} to create
-and partition your mesh, you can output mesh files in ABAQUS (.INP)
-and DX (.dx) format to visualize them. For this, you must set either
-the flag \texttt{CREATE\_DX\_FILES} or \texttt{CREATE\_ABAQUS\_FILES}
-to \texttt{.true.} in the mesher's parameter file \texttt{Mesh\_Par\_file}
-prior to running the mesher (see Chapter~\ref{cha:Running-the-Mesher-Meshfem3D}
-for details). You can then use AVS \urlwithparentheses{www.avs.com}
-or OpenDX \urlwithparentheses{www.opendx.org} to visualize the mesh
-and MPI partition (slices).
-
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=0.49\textwidth]{figures/vtk_mesh_vp} \includegraphics[width=0.49\textwidth]{figures/vtk_mesh_vs}
-\par\end{centering}
-
-\caption{Visualization using Paraview of VTK files created by \texttt{xgenerate\_databases}
-showing P- and S-wave velocities assigned to the mesh points. The
-mesh was created by \texttt{xmeshfem3D} for 4 processors.}
-
-\label{fig:vtk.mesh}
-\end{figure}
-
-You have also the option to visualize the distributed databases produced
-by \texttt{xgenerate\_databases} using Paraview \urlwithparentheses{www.paraview.org}.
-For this, you must set the flag \texttt{SAVE\_MESH\_FILES} to \texttt{.true.}
-in the main parameter file \texttt{Par\_file} (see Chapter~\ref{cha:Main-Parameter}
-for details). This will create VTK files for each single partition.
-You can then use Paraview \urlwithparentheses{www.paraview.org}
-to visualized these partitions.
-
-\section{\label{sec:Movies}Movies}
-
-To make a surface or volume movie of the simulation, set parameters
-\texttt{MOVIE\_SURFACE}, \texttt{MOVIE\_VOLUME}, \texttt{MOVIE\_TYPE},
-and \texttt{NTSTEP\_BETWEEN\_FRAMES} in the \texttt{Par\_file}. Turning
-on the movie flags, in particular \texttt{MOVIE\_VOLUME}, produces
-large output files. \texttt{MOVIE\_VOLUME} files are saved in the
-\texttt{LOCAL\_PATH} directory, whereas \texttt{MOVIE\_SURFACE} output
-files are saved in the \texttt{OUTPUT\_FILES} directory. We save the
-displacement field if the parameter \texttt{SAVE\_DISPLACEMENT} is
-set, otherwise the velocity field is saved. The look of a movie is
-determined by the half-duration of the source. The half-duration should
-be large enough so that the movie does not contain frequencies that
-are not resolved by the mesh, i.e., it should not contain numerical
-noise. This can be accomplished by selecting a CMT \texttt{HALF\_DURATION}
-> 1.1 $\times$ smallest period (see figure \ref{fig:CMTSOLUTION-file}).
-When \texttt{MOVIE\_SURFACE} = .\texttt{true.}, the half duration
-of each source in the \texttt{CMTSOLUTION} file is replaced by
-\begin{quote}
-\[
-\sqrt{(}\mathrm{\mathtt{HALF\_DURATIO}\mathtt{N}^{2}}+\mathrm{\mathtt{HDUR\_MOVI}\mathtt{E}^{2}})
-\]
- \textbf{NOTE:} If \texttt{HDUR\_MOVIE} is set to 0.0, the code will
-select the appropriate value of 1.1 $\times$ smallest period. As
-usual, for a point source one can set \texttt{half duration} in the
-\texttt{CMTSOLUTION} file to be 0.0 and \texttt{HDUR\_MOVIE} = 0.0
-in the \texttt{Par\_file} to get the highest frequencies resolved
-by the simulation, but for a finite source one would keep all the
-\texttt{half durations} as prescribed by the finite source model and
-set \texttt{HDUR\_MOVIE} = 0.0.
-\end{quote}
-
-\subsection{Movie Surface}
-
-When running \texttt{xspecfem3D} with the \texttt{MOVIE\_SURFACE}
-flag turned on, the code outputs \texttt{moviedata??????} files in
-the \texttt{OUTPUT\_FILES} directory. There are several flags in the
-main parameter file \texttt{Par\_file} that control the output of
-these movie data files (see section \ref{cha:Main-Parameter} for
-details): \texttt{\small NTSTEP\_BETWEEN\_FRAMES}{\small{} to set the
-timesteps between frames, }\texttt{\small SAVE\_DISPLACEMENT}{\small{}
-to save displacement instead of velocity, }\texttt{\small USE\_HIGHRES\_FOR\_MOVIES}{\small{}
-to save values at all GLL point instead of element edges. In order
-to output additionally shakemaps, you would set the parameter }\texttt{\small CREATE\_SHAKEMAP}{\small{}
-to }\texttt{\small .true.}{\small .}{\small \par}
-
-The files are in a fairly complicated binary format, but there is
-a program provided to convert the output into more user friendly formats:
-\begin{description}
-\item [{\texttt{xcreate\_movie\_shakemap\_AVS\_DX\_GMT}}] From \texttt{create\_movie\_shakemap\_AVS\_DX\_GMT.f90},
-it outputs data in ASCII, OpenDX, or AVS format (also readable in
-ParaView). Before compiling the code, make sure you have the file
-\texttt{surface\_from\_mesher.h} in the \texttt{OUTPUT\_FILES/} directory.
-This file will be created by the solver run. Then type
-
-\begin{lyxcode}
-{\footnotesize make~xcreate\_movie\_shakemap\_AVS\_DX\_GMT}{\footnotesize \par}
-\end{lyxcode}
-
-and run the executable \texttt{xcreate\_movie\_shakemap\_AVS\_DX\_GMT}
-in the main directory. It will create visualization files
-in your format of choice. The code will prompt the user for input
-parameters.
-
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=0.32\textwidth]{figures/movie_surf_1} \includegraphics[width=0.32\textwidth]{figures/movie_surf_2}
-\includegraphics[width=0.32\textwidth]{figures/movie_surf_3}
-\par\end{centering}
-
-\caption{Visualization using AVS files created by \texttt{xcreate\_movie\_shakemap\_AVS\_DX\_GMT}
-showing movie snapshots of vertical velocity components at different
-times.}
-
-\label{fig:movie.surf}
-\end{figure}
-
-\end{description}
-The \texttt{SPECFEM3D Cartesian} code is running in near real-time
-to produce animations of southern California earthquakes via the web;
-see Southern California ShakeMovie\textregistered{}\urlwithparentheses{www.shakemovie.caltech.edu}.
-
-\subsection{\label{sub:Movie-Volume}Movie Volume}
-
-When running xspecfem3D with the \texttt{\small MOVIE\_VOLUME}{\small{}
-flag turned on, the code outputs several files in }\texttt{\small LOCAL\_PATH}{\small{}
-specified in the main }\texttt{\small Par\_file}{\small , e.g. in
-directory }\texttt{\small OUTPUT\_FILES/DATABASES\_MPI}{\small .
-The output is saved by each processor at the time interval specified
-by }\texttt{\small NTSTEP\_BETWEEN\_FRAMES}{\small . For all domains,
-the velocity field is output to files: }\texttt{\small proc??????\_velocity\_X\_it??????.bin}{\small ,
-}\texttt{\small proc??????\_velocity\_Y\_it??????.bin}{\small{} and
-}\texttt{\small proc??????\_velocity\_Z\_it??????.bin}{\small . For
-elastic domains, the divergence and curl taken from the velocity field,
-i.e. $\nabla\cdot{\bf {v}}$ and $\nabla\times{\bf {v}}$, get stored
-as well: }\texttt{\small proc??????\_div\_it??????.bin}{\small , }\texttt{\small proc??????\_curl\_X\_t??????.bin}{\small ,
-}\texttt{\small proc??????\_curl\_Y\_it??????.bin}{\small{} and }\texttt{\small proc??????\_curl\_Z\_it??????.bin}{\small .
-The files denoted }\texttt{\small proc??????\_div\_}{\small ~}\\
-{\small{} }\texttt{\small glob\_it??????.bin}{\small{} and }\texttt{\small proc??????\_curl\_glob\_it??????.bin}{\small{}
-are stored on the global points only, all the other arrays are stored
-on all GLL points. Note that the components X/Y/Z can change to E/N/Z
-according to the }\texttt{\small SUPPRESS\_UTM\_PROJECTION}{\small{}
-flag (see also Appendix~\ref{cha:Coordinates} and \ref{cha:channel-codes}).}{\small \par}
-
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=0.32\textwidth]{figures/movie_volume_1} \includegraphics[width=0.32\textwidth]{figures/movie_volume_2}
-\includegraphics[width=0.32\textwidth]{figures/movie_volume_3}
-\par\end{centering}
-
-\caption{Paraview visualization using movie volume files (converted by \texttt{xcombine\_vol\_data}
-and \texttt{mesh2vtu.pl}) and showing snapshots of vertical velocity
-components at different times.}
-
-\label{fig:movie.volume}
-\end{figure}
-
-To visualize these files, we use an auxilliary program \texttt{combine\_vol\_data.f90}
-to combine the data from all slices into one mesh file. To compile
-it in the root directory, type:
-\begin{lyxcode}
-{\footnotesize make~xcombine\_vol\_data~}{\footnotesize \par}
-\end{lyxcode}
-which will create the executable \texttt{bin/xcombine\_vol\_data}.
-To output the usage of this executable, type
-\begin{lyxcode}
-{\footnotesize ./bin/xcombine\_vol\_data~}{\footnotesize \par}
-\end{lyxcode}
-without arguments.
-
-\medskip
-
-\texttt{xcombine\_vol\_data} will combine the data on the different processors (located in \texttt{OUTPUT\_FILES/DATABASES\_MPI}), for a given quantity and a given iteration, to one file.
-For example, if you want to combine \texttt{velocity\_Z}, for the iteration 400 on 4 processors, i.e.
-if you want to combine these files :
-
-\medskip
-
-\texttt{proc000000\_velocity\_Z\_it000400.bin }
-
-\texttt{proc000001\_velocity\_Z\_it000400.bin }
-
-\texttt{proc000002\_velocity\_Z\_it000400.bin }
-
-\texttt{proc000003\_velocity\_Z\_it000400.bin}
-
-\medskip
-
-You have to go in the directory of the concerned example (where are the directories \texttt{DATA}, \texttt{OUTPUT\_FILES}, etc.).
-Then, you can launch \texttt{xcombine\_vol\_data}, specifying the path where it is located.
-As an example, if the \texttt{DATA} and \texttt{OUTPUT\_FILES} directories of the concerned example are in the root specfem3d directory, \texttt{xcombine\_vol\_data} is in \texttt{./bin}, and you have to type:
-
-\begin{lyxcode}
-~{\footnotesize ./bin/xcombine\_vol\_data~0~3~velocity\_Z\_it000400~./OUTPUT\_FILES/DATABASES\_MPI~./OUTPUT\_FILES~0~}{\footnotesize \par}
-\end{lyxcode}
-
-Here, $0$ is the number of the first processor, $3$ the number of the last one, \texttt{velocity\_Z\_it000400} the name of the files we want to combine without the prefix \texttt{proc000000*\_}, \texttt{./OUTPUT\_FILES/DATABASES\_MPI} the directory where the files are located, \texttt{./OUTPUT\_FILES} the directory where the combined file will be stored, and $0$ is the parameter to create a low-resolution mesh file ($1$ for a high-resolution).
-
-\medskip
-
-At the beginning of the source file \texttt{combine\_vol\_data.f90} (located in \texttt{src/auxiliaries/}), there is a logical constant parameter \texttt{USE\_VTK\_OUTPUT} whose default value is \texttt{.true.}, that which will provide the output combined file directly in the VTK format and you can use Paraview to visualize it.
-
-\medskip
-
-If the value of this parameter is \texttt{.false.}, the output mesh file will have the name \texttt{velocity\_Z\_it000400.mesh}. Then we next have to convert the
-\texttt{.mesh} file into the VTU (Unstructured grid file) format which can be viewed in ParaView. For this task, you can use and modify the
-script \texttt{mesh2vtu.pl} located in directory \texttt{utils/Visualization/Paraview/}, for example:
-\begin{lyxcode}
-{\footnotesize mesh2vtu.pl~-i~velocity\_Z\_it000400.mesh~-o~velocity\_Z\_it000400.vtu}{\footnotesize \par}
-\end{lyxcode}
-Notice that this Perl script uses a program \texttt{mesh2vtu} in the \texttt{utils/Visualization/Paraview/mesh2vtu} directory, which further
-uses the VTK \urlwithparentheses{http://www.vtk.org/} run-time library for its execution. Therefore, make sure you have them properly set
-in the script according to your system.
-
-\bigskip
-
-Then, to do a movie with several iterations, you have to repeat this process for each iteration you want to put in your movie.
-
-\section{\label{sec:Finite-Frequency-Kernels}Finite-Frequency Kernels}
-
-The finite-frequency kernels computed as explained in Section \ref{sec:Adjoint-simulation-finite}
-are saved in the \texttt{LOCAL\_PATH} at the end of the simulation.
-Therefore, we first need to collect these files on the front end,
-combine them into one mesh file, and visualize them with some auxilliary
-programs.
-\begin{enumerate}
-\item \textbf{Create slice files}
-
-We will only discuss the case of one source-receiver pair, i.e., the
-so-called banana-doughnut kernels. Although it is possible to collect
-the kernel files from all slices on the front end, it usually takes
-up too much storage space (at least tens of gigabytes). Since the
-sensitivity kernels are the strongest along the source-receiver great
-circle path, it is sufficient to collect only the slices that are
-along or close to the great circle path.
-
-A Perl script \texttt{slice\_number.pl} located in directory \texttt{utils/Visualization/Paraview/}
-can help to figure out the slice numbers that lie along the great
-circle path. It applies to meshes created with the internal mesher
-\texttt{xmeshfem3D}.
-\begin{enumerate}
-\item On machines where you have access to the script, copy the \texttt{Mesh\_Par\_file},
-and \texttt{output\_solver} files, and run:
-
-\begin{lyxcode}
-{\small slice\_number.pl~Mesh\_Par\_file~output\_solver.txt~slice\_file}{\small \par}
-\end{lyxcode}
-
-which will generate a \texttt{slices\_file}.
-
-\item For cases with multiple sources and multiple receivers, you need to
-provide a slice file before proceeding to the next step.
-\end{enumerate}
-\item \textbf{Collect the kernel files}
-
-After obtaining the slice files, you can collect the corresponding
-kernel files from the given slices.
-\begin{enumerate}
-\item You can use or modify the script \texttt{utils/copy\_basin\_database.pl}
-to accomplish this:
-
-\begin{lyxcode}
-{\small utils/copy\_database.pl~slice\_file~lsf\_machine\_file~filename~{[}jobid{]}}{\small \par}
-\end{lyxcode}
-
-where \texttt{\small lsf\_machine\_file}{\small{} is the machine file
-generated by the LSF scheduler, }\texttt{\small filename}{\small{}
-is the kernel name (e.g., }\texttt{\small rho\_kernel}{\small , }\texttt{\small alpha\_kernel}{\small{}
-and }\texttt{\small beta\_kernel}{\small ), and the optional }\texttt{\small jobid}{\small{}
-is the name of the subdirectory under }\texttt{\small LOCAL\_PATH}{\small{}
-where all the kernel files are stored.}{\small \par}
-
-\item After executing this script, all the necessary mesh topology files
-as well as the kernel array files are collected to the local directory
-of the front end.
-\end{enumerate}
-\item \textbf{Combine kernel files into one mesh file}
-
-We use an auxilliary program \texttt{combine\_vol\_data.f90} to combine
-the kernel files from all slices into one mesh file.
-\begin{enumerate}
-\item Compile it in the root directory:
-
-\begin{lyxcode}
-{\footnotesize make~xcombine\_vol\_data~}{\footnotesize \par}
-
-{\footnotesize ./bin/xcombine\_vol\_data~slice\_list~filename~input\_dir~output\_dir~high/low-resolution~}{\footnotesize \par}
-\end{lyxcode}
-
-where \texttt{input\_dir} is the directory where all the individual
-kernel files are stored, and \texttt{output\_dir} is where the mesh
-file will be written.
-
-\item Use 1 for a high-resolution mesh, outputting all the GLL points to
-the mesh file, or use 0 for low resolution, outputting only the corner
-points of the elements to the mesh file.
-\item The output mesh file will have the name \texttt{filename\_rho(alpha,beta).mesh}
-\end{enumerate}
-\item \textbf{Convert mesh files into .vtu files}
-
-\begin{enumerate}
-\item We next convert the \texttt{.mesh} file into the VTU (Unstructured
-grid file) format which can be viewed in ParaView. For this task,
-you can use and modify the script \texttt{mesh2vtu.pl} located in
-directory \texttt{utils/Visualization/Paraview/}, for example:
-
-\begin{lyxcode}
-{\footnotesize mesh2vtu.pl~-i~file.mesh~-o~file.vtu~}{\footnotesize \par}
-\end{lyxcode}
-\item Notice that this Perl script uses a program \texttt{mesh2vtu} in the
-\texttt{utils/Visualization/Paraview/mesh2vtu} directory, which further
-uses the VTK \urlwithparentheses{http://www.vtk.org/} run-time library
-for its execution. Therefore, make sure you have them properly set
-in the script according to your system.
-\end{enumerate}
-\item \textbf{Copy over the source and receiver .vtk file}
-
-In the case of a single source and a single receiver, the simulation
-also generates the file \texttt{sr.vtk} located in the \texttt{OUTPUT\_FILES/}
-directory to describe the source and receiver locations, which can
-also be viewed in Paraview in the next step.
-
-\item \textbf{View the mesh in ParaView}
-
-Finally, we can view the mesh in ParaView \urlwithparentheses{www.paraview.org}.
-\begin{enumerate}
-\item Open ParaView.
-\item From the top menu, \textsf{File} $\rightarrow$\textsf{Open data},
-select \texttt{file.vtu}, and click the \textsf{Accept} button.
-
-\begin{itemize}
-\item If the mesh file is of moderate size, it shows up on the screen; otherwise,
-only the bounding box is shown.
-\end{itemize}
-\item Click \textsf{Display Tab} $\rightarrow$ \textsf{Display Style} $\rightarrow$
-\textsf{Representation} and select \textsf{wireframe of surface} to
-display it.
-\item To create a cross-section of the volumetric mesh, choose \textsf{Filter}
-$\rightarrow$ \textsf{cut}, and under \textsf{Parameters Tab}, choose
-\textsf{Cut Function} $\rightarrow$ \textsf{plane}.
-\item Fill in center and normal information given by the \texttt{global\_slice\_number.pl}
-script (either from the standard output or from \texttt{normal\_plane.txt}
-file).
-\item To change the color scale, go to \textsf{Display Tab} $\rightarrow$
-\textsf{Color} $\rightarrow$ \textsf{Edit Color Map} and reselect
-lower and upper limits, or change the color scheme.
-\item Now load in the source and receiver location file by \textsf{File}
-$\rightarrow$ \textsf{Open data}, select \texttt{sr.vt}k, and click
-the \textsf{Accept} button. Choose \textsf{Filter} $\rightarrow$
-\textsf{Glyph}, and represent the points by `\textsf{spheres}'.
-\item For more information about ParaView, see the ParaView Users Guide
-\urlwithparentheses{www.paraview.org/files/v1.6/ParaViewUsersGuide.PDF}.
-\end{enumerate}
-\end{enumerate}
-\begin{figure}[H]
-\noindent \begin{centering}
-\includegraphics[scale=0.6]{figures/3D-S-Kernel}
-\par\end{centering}
-
-\caption{(a) Top Panel: Vertical source-receiver cross-section of the S-wave
-finite-frequency sensitivity kernel $K_{\beta}$ for station GSC at
-an epicentral distance of 176 km from the September 3, 2002, Yorba
-Linda earthquake. Lower Panel: Vertical source-receiver cross-section
-of the 3D S-wave speed model used for the spectral-element simulations
-\citep{KoLiTrSuStSh04}. (b) The same as (a) but for station HEC at
-an epicentral distance of 165 km \citep{LiTr06}.}
-
-\label{figure:P-wave-speed-finite-frequency}
-\end{figure}
-
-\chapter{\label{cha:Scheduler}Running through a Scheduler}
-
-The code is usually run on large parallel machines, often PC clusters,
-most of which use schedulers, i.e., queuing or batch management systems
-to manage the running of jobs from a large number of users. The following
-considerations need to be taken into account when running on a system
-that uses a scheduler:
-\begin{itemize}
-\item The processors/nodes to be used for each run are assigned dynamically
-by the scheduler, based on availability. Therefore, in order for the
-\texttt{xgenerate\_databases} and the \texttt{xspecfem3D} executables
-(or between successive runs of the solver) to have access to the same
-database files (if they are stored on hard drives local to the nodes
-on which the code is run), they must be launched in sequence as a
-single job.
-\item On some systems, the nodes to which running jobs are assigned are
-not configured for compilation. It may therefore be necessary to pre-compile
-both the \texttt{xgenerate\_databases} and the \texttt{xspecfem3D}
-executables.
-\item One feature of schedulers/queuing systems is that they allow submission
-of multiple jobs in a ``launch and forget'' mode. In order to take
-advantage of this property, care needs to be taken that output and
-intermediate files from separate jobs do not overwrite each other,
-or otherwise interfere with other running jobs.
-\end{itemize}
-Examples of job scripts can be found in the \texttt{\small utils/Cluster/}{\small{}
-directory and can straightforwardly be modified and adapted to meet
-more specific running needs.}{\small \par}
-
-\chapter{{\normalsize \label{cha:-Changing-the}} Changing the Model}
-
-In this section we explain how to change the velocity model used for
-your simulations. These changes involve contributing specific subroutines
-that replace existing subroutines in the \texttt{SPECFEM3D Cartesian}
-package. Note that \texttt{SPECFEM3D Cartesian} can handle Earth models
-with material properties that vary within each spectral element.
-
-\section{{\normalsize \label{sec:Using-tomographic}}Using external tomographic Earth models}
-
-To implement your own external tomographic model(s), you must provide
-your own external tomography file(s), and choose between two possible
-options:\\
- \indent (1) set in the \texttt{Par\_file} the model parameter \texttt{MODEL
-= tomo}, \\
- or for more user control: \\
- \indent (2) set in the \texttt{Par\_file} the model parameter \texttt{MODEL
-= default}, define the negative \texttt{material\_ID} identifier for
-each element in the file \texttt{MESH/materials\_file} and use the
-following format in the file \texttt{MESH/nummaterial\_velocity\_file}
-when using CUBIT to construct your mesh (see also section \ref{subsec:Exporting-the-Mesh}):
-\begin{lyxcode}
-domain\_ID~material\_ID~tomography~elastic~file\_name~1
-\end{lyxcode}
-where: \\
- \indent \texttt{domain\_ID} is 1 for acoustic or 2 for elastic materials,
-\\
- \indent \texttt{material\_ID} a negative, unique identifier (i.e.,
--1,-2,...), \\
- \indent \texttt{tomography} keyword for tomographic material definition,
-\\
- \indent \texttt{elastic} keyword for elastic material definition,
-\\
- \indent \texttt{file\_name} the name of the tomography file and
-1 a positive unique identifier.\\
-
-The external tomographic model is represented by a grid of points
-with assigned material properties and homogeneous resolution along
-each spatial direction x, y and z. The ASCII file \texttt{file\_name}
-that describe the tomography should be located in the \texttt{TOMOGRAPHY\_PATH}
-directory, set in the \texttt{Par\_file}. The format of the file,
-as read from \texttt{model\_tomography.f90} located in the \texttt{src/generate\_databases}
-directory, looks like Figure \ref{fig:tomography_file}, where
-\begin{description}
-\item [{\texttt{ORIG\_X}, \texttt{END\_X}}] are, respectively, the coordinates
-of the initial and final tomographic grid points along the x direction
-(in the mesh units, e.g., $m$);
-\item [{\texttt{ORIG\_Y}, \texttt{END\_Y}}] respectively the coordinates
-of the initial and final tomographic grid points along the y direction
-(in the mesh units, e.g., $m$);
-\item [{\texttt{ORIG\_Z}, \texttt{END\_Z}}] respectively the coordinates
-of the initial and final tomographic grid points along the z direction
-(in the mesh units, e.g., $m$);
-\item [{\texttt{SPACING\_X}, \texttt{SPACING\_Y}, \texttt{SPACING\_Z}}] the
-spacing between the tomographic grid points along the x, y and z directions,
-respectively (in the mesh units, e.g., $m$);
-\item [{\texttt{NX}, \texttt{NY}, \texttt{NZ}}] the number of grid points
-along the spatial directions x, y and z, respectively; \texttt{NX}
-is given by {[}(\texttt{END\_X} - \texttt{ORIG\_X})/\texttt{SPACING\_X}{]}+1;
-\texttt{NY} and \texttt{NZ} are the same as \texttt{NX}, but for the
-y and z directions, respectively;
-\item [{\texttt{VP\_MIN}, \texttt{VP\_MAX}, \texttt{VS\_MIN}, \texttt{VS\_MAX}, \texttt{RHO\_MIN}, \texttt{RHO\_MAX}}] the
-minimum and maximum values of the wave speed \texttt{vp} and \texttt{vs}
-(in $m\, s^{-1}$) and of the density \texttt{rho} (in $kg\, m^{-3}$);
-these values could be the actual limits of the tomographic parameters
-in the grid or the minimum and maximum values to which we force the
-cut of velocity and density in the model.
-\end{description}
-After the first four lines, in the file \texttt{file\_name} the tomographic
-grid points are listed with the corresponding values of \texttt{vp},
-\texttt{vs} and \texttt{rho}, scanning the grid along the x coordinate
-(from \texttt{ORIG\_X} to \texttt{END\_X} with step of \texttt{SPACING\_X})
-for each given y (from \texttt{ORIG\_Y} to \texttt{END\_Y}, with step
-of \texttt{SPACING\_Y}) and each given z (from \texttt{ORIG\_Z} to
-\texttt{END\_Z}, with step of \texttt{SPACING\_Z}). %fig
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=1\textwidth]{figures/tomo_file}
-\par\end{centering}
-
-\caption{Tomography file \texttt{file\_name} that describes an external Earth
-model. The coordinates x, y and z of the grid, their limits ORIG\_X,
-ORIG\_Y, ORIG\_Z, END\_X, END\_Y, END\_Z and the grid spacings SPACING\_X,
-SPACING\_Y, SPACING\_Z should be in the units of the constructed mesh
-(e.g., UTM coordinates in meters).}
-
-\label{fig:tomography_file}
-\end{figure}
-
-The user can implement his own interpolation algorithm for the tomography
-model by changing the routine \texttt{model\_tomography.f90} located
-in the \texttt{src/generate\_databases/} directory. Moreover, for
-models that involve both fully defined materials and a tomography
-description, the \texttt{nummaterial\_velocity\_file} has multiple
-lines each with the corresponding suitable format described above.
-
-%==============
-
-\subsubsection*{Example: External tomography file with variable discretization intervals}
-
-The example in Figure~\ref{fig:tomography_file} is for an external tomography file with uniform spacing in the $x$, $y$, and $z$ directions.
-(Note that $\Delta x$, $\Delta y$, and $\Delta z$ need not be the same as in the example.) In that example, the  \texttt{nummaterial\_velocity\_file} is
-%
-\begin{lyxcode}
-2  -1 tomography elastic tomography\_model.xyz 1
-\end{lyxcode}
-%
-and the file \texttt{tomography\_model.xyz} will need to reside in \texttt{DATA/}.
-All entries in the second column of \texttt{materials\_file} will be \texttt{-1}, which means that each element in the mesh will be interpolated accordind to the values in \texttt{tomography\_model.xyz}.
-
-In some cases it may be desirable to have an external tomography model that is described in more than one file.
-For example, in cases like southern California, the length scale of variation in the structure of the wave speed model is much shorter in the sedimentary basin models within the upper 15~km.
-Therefore one might want to use an external tomography file that is sampled with, say, $\Delta x = 1000$~m, $\Delta y = 1000$~m, and $\Delta z = 250$~m in the uppermost 15~km, and then use $\Delta x = 2000$~m, $\Delta y = 2000$~m, and $\Delta z = 1000$~m below a depth of 15~km.
-If these intervals are chosen appropriately, then it will result in a pair of external tomography file that is much smaller than the alternative of having a single file with the fine discretization.
-In this case  \texttt{nummaterial\_velocity\_file} is
-%
-\begin{lyxcode}
-2  -1 tomography elastic file\_above\_15km.xyz 1 \\
-2  -2 tomography elastic file\_below\_15km.xyz 1
-\end{lyxcode}
-%
-and the files \texttt{file\_above\_15km.xyz} and \texttt{file\_below\_15km.xyz} will need to reside in \texttt{DATA/}.
-All entries in the second column of \texttt{materials\_file} will need to be either \texttt{-1} or \texttt{-2}, depending on whether the element is above or below 15~km depth.
-In this sense, the construction of the mesh and the discretization of the external model will generally be done in tandem.
-
-Other more complicated discretizations can be used following the same procedure. (In particular, there is no limit on the number of different external files that are used.)
-
-%=============
-
-\section{{\normalsize \label{sec:Anelastic-Models}}External (an)elastic Models}
-
-To use your own external model, you can set in the \texttt{Par\_file}
-the model parameter \texttt{MODEL = external}. Three-dimensional acoustic
-and/or (an)elastic (attenuation) models may then be superimposed onto
-the mesh based upon your subroutine in \texttt{model\_external\_values.f90}
-located in directory \texttt{src/generate\_databases}. The call to
-this routine would be as follows
-\begin{lyxcode}
-call~model\_external\_values(xmesh,~ymesh,~zmesh,~~\&~~\\
-~~~~~~~~~~~rho,~vp,~vs,~qmu\_atten,~iflag\_aniso,~idomain\_id)
-\end{lyxcode}
-Input to this routine consists of:
-\begin{description}
-\item [{\texttt{xmesh,ymesh,zmesh}}] location of mesh point
-\end{description}
-Output to this routine consists of:
-\begin{description}
-\item [{\texttt{rho,vp,vs}}] isotropic model parameters for density $\rho$
-($kg/m^{3}$), $v_{p}$ ($m/s$) and $v_{s}$ ($m/s$)
-\item [{\texttt{qmu\_atten}}] Shear wave quality factor: $0<Q_{\mu}<9000$
-\item [{\texttt{iflag\_aniso}}] anisotropic model flag, $0$ indicating
-no anisotropy or $1$ using anisotropic model parameters as defined
-in routine file \texttt{model\_aniso.f90}
-\item [{\texttt{idomain\_id}}] domain identifier, $1$ for acoustic, $2$
-for elastic, $3$ for poroelastic materials.
-\end{description}
-Note that the resolution and maximum value of anelastic models are
-truncated. This speeds the construction of the standard linear solids
-during the meshing stage. To change the resolution, currently at one
-significant figure following the decimal, or the maximum value (9000),
-consult \texttt{shared/constants.h}.
-
-\section{{\normalsize \label{sec:Anisotropic-Models}}Anisotropic Models}
-
-To use your anisotropic models, you can either set in the \texttt{Par\_file}
-the model parameter \texttt{ANISOTROPY} to \texttt{.true.} or set
-the model parameter \texttt{MODEL = aniso}. Three-dimensional anisotropic
-models may be superimposed on the mesh based upon the subroutine in
-file \texttt{model\_aniso.f90} located in directory \texttt{src/generate\_databases}.
-The call to this subroutine is of the form
-\begin{lyxcode}
-call~model\_aniso(iflag\_aniso,~rho,~vp,~vs,~\&~~\\
-~~~~~~~~~~~c11,c12,c13,c14,c15,c16,c22,c23,c24,c25,c26,~\&~~\\
-~~~~~~~~~~~c33,c34,c35,c36,c44,c45,c46,c55,c56,c66)~
-\end{lyxcode}
-Input to this routine consists of:
-\begin{description}
-\item [{\texttt{iflag\_aniso}}] flag indicating the type of the anisotropic
-model, i.e. $0$ for no superimposed anisotropy, $1$ or $2$ for
-generic pre-defined anisotropic models.
-\item [{\texttt{rho,vp,vs}}] reference isotropic model parameters for density
-$\rho$, $v_{p}$ and $v_{s}$.
-\end{description}
-Output from the routine consists of the following non-dimensional
-model parameters:
-\begin{description}
-\item [{\texttt{c11},}] \textbf{$\cdots$,} \texttt{\textbf{c66}} 21~dimensionalized
-anisotropic elastic parameters.
-\end{description}
-You can replace the \texttt{model\_aniso.f90} file by your own version
-\textit{provided you do not change the call structure of the routine},
-i.e., the new routine should take exactly the same input and produce
-the required relative output.
-
-\section{{\normalsize \label{sec:Using-SEP}}Using external SEP models.}
-
-Stanford Exploration Project (SEP) models consists of two files per variables, one header file (\texttt{.H})
-and one binary file (\texttt{.H$@$}).
-Among other information, the header file indicates, for each dimension,
-the offset (\texttt{o1}, \texttt{o2} and \texttt{o3} fields),
-the spatial step size (\texttt{d1}, \texttt{d2} and \texttt{d3} fields)
-and the number of samples (\texttt{n1}, \texttt{n2} and \texttt{n3}) necessary to read
-the linked binary file (\texttt{in} field).
-
-SEP model reading occurs during database generation. It it the user responsibility
-to ensure that proper interfaces between acoustic and elastic domains are set up
-during the meshing phase. For more information, refer to
-section~\ref{cha:Running-the-Mesher-Meshfem3D}.
-
-In \texttt{DATA/Par\_file}, \texttt{\textbf{MODEL}} should be set to \texttt{\textbf{sep}},
-and \texttt{\textbf{SEP\_MODEL\_DIRECTORY}} should point to wherever your sep model files reside.
-
-Note that in order not to overload the parameter file, SEP header files
-should be named \texttt{vp.H}, \texttt{vs.H} and \texttt{rho.H}. The \texttt{in} field
-in these binary files can be any path as long as it is relative to
-the previously set \texttt{SEP\_MODEL\_DIRECTORY}.
-We also assume that any dimensional value inside SEP header files is given in meters and
-that binary files are single precision floating point numbers stored in little endian format.
-There is currently only support for \texttt{vp}, \texttt{vs} and \texttt{rho} variables.
-
-An example demonstrating the use SEP models is given in:
-\texttt{EXAMPLES/meshfem3D\_examples/sep\_bathymetry}.
-
-\chapter{Post-Processing Scripts}
-
-Several post-processing scripts/programs are provided in the \texttt{utils/}
-directory, and most of them need to be adjusted when used on different
-systems, for example, the path of the executable programs. Here we
-only list a few of the available scripts and provide a brief description,
-and you can either refer to the related sections for detailed usage
-or, in a lot of cases, type the script/program name without arguments
-for its usage.
-
-\section{Process Data and Synthetics\label{sec:Process-data-and-syn}}
-
-In many cases, the SEM synthetics are calculated and compared to data
-seismograms recorded at seismic stations. Since the SEM synthetics
-are accurate for a certain frequency range, both the original data
-and the synthetics need to be processed before a comparison can be
-made.
-
-For such comparisons, the following steps are recommended:
-\begin{enumerate}
-\item Make sure that both synthetic and observed seismograms have the correct
-station/event and timing information.
-\item Convolve synthetic seismograms with a source time function with the
-half duration specified in the \texttt{CMTSOLUTION} file, provided,
-as recommended, you used a zero half duration in the SEM simulations.
-\item Resample both observed and synthetic seismograms to a common sampling
-rate.
-\item Cut the records using the same window.
-\item Remove the trend and mean from the records and taper them.
-\item Remove the instrument response from the observed seismograms (recommended)
-or convolve the synthetic seismograms with the instrument response.
-\item Make sure that you apply the same filters to both observed and synthetic
-seismograms. Preferably, avoid filtering your records more than once.
-\item Now, you are ready to compare your synthetic and observed seismograms.
-\end{enumerate}
-\noindent We generally use the following scripts provided in the \texttt{utils/seis\_process/}
-directory:
-
-\subsection{Data processing script \texttt{process\_data.pl}}
-
-This script cuts a given portion of the original data, filters it,
-transfers the data into a displacement record, and picks the first
-P and S arrivals. For more functionality, type `\texttt{process\_data.pl}'
-without any argument. An example of the usage of the script:
-\begin{lyxcode}
-{\footnotesize process\_data.pl~-m~CMTSOLUTION~-s~1.0~-l~0/4000~-i~-f~-t~40/500~-p~-x~bp~DATA/1999.330{*}.BH?.SAC}~
-\end{lyxcode}
-\noindent which has resampled the SAC files to a sampling rate of
-1 seconds, cut them between 0 and 4000 seconds, transfered them into
-displacement records and filtered them between 40 and 500 seconds,
-picked the first P and S arrivals, and added suffix `\texttt{bp}'
-to the file names.
-
-Note that all of the scripts in this section actually use the SAC
-and/or IASP91 to do the core operations; therefore make sure that
-the SAC and IASP91 packages are installed properly on your system,
-and that all the environment variables are set properly before running
-these scripts.
-
-\subsection{Synthetics processing script \texttt{process\_syn.pl}}
-
-This script converts the synthetic output from the SEM code from ASCII
-to SAC format, and performs similar operations as `\texttt{process\_data.pl}'.
-An example of the usage of the script:
-\begin{lyxcode}
-{\footnotesize process\_syn.pl~-m~CMTSOLUTION~-h~-a~STATIONS~-s~1.0~-l~0/4000~-f~-t~40/500~-p~-x~bp~SEM/{*}.BX?.semd}~
-\end{lyxcode}
-which will convolve the synthetics with a triangular source-time function
-from the \texttt{CMTSOLUTION} file, convert the synthetics into SAC
-format, add event and station information into the SAC headers, resample
-the SAC files with a sampling rate of 1 seconds, cut them between
-0 and 4000 seconds, filter them between 40 and 500 seconds with the
-same filter used for the observed data, pick the first P and S arrivals,
-and add the suffix `\texttt{bp}' to the file names.
-
-More options are available for this script, such as adding time shift
-to the origin time of the synthetics, convolving the synthetics with
-a triangular source time function with a given half duration, etc.
-Type \texttt{process\_syn.pl} without any argument for a detailed
-usage.
-
-In order to convert between SAC format and ASCII files, useful scripts
-are provided in the subdirectories ~\\
- \texttt{utils/sac2000\_alpha\_convert/} and \texttt{utils/seis\_process/asc2sac/}.
-
-\subsection{Script \texttt{rotate.pl}}
-
-The original data and synthetics have three components: vertical (BHZ
-resp. BXZ), north (BHN resp. BXN) and east (BHE resp. BXE). However,
-for most seismology applications, transverse and radial components
-are also desirable. Therefore, we need to rotate the horizontal components
-of both the data and the synthetics to the transverse and radial direction,
-and \texttt{\small rotate.pl}{\small{} can be used to accomplish this:}{\small \par}
-\begin{lyxcode}
-rotate.pl~-l~0~-L~180~-d~DATA/{*}.BHE.SAC.bp~
-
-rotate.pl~-l~0~-L~180~SEM/{*}.BXE.semd.sac.bp~
-\end{lyxcode}
-where the first command performs rotation on the SAC data obtained
-through Seismogram Transfer Program (STP) \urlwithparentheses{http://www.data.scec.org/STP/stp.html},
-while the second command rotates the processed SEM synthetics.
-
-\section{Collect Synthetic Seismograms}
-
-The forward and adjoint simulations generate synthetic seismograms
-in the \texttt{OUTPUT\_FILES/} directory by default. For the forward
-simulation, the files are named like \texttt{NT.STA.BX?.semd} for
-two-column time series, or \texttt{NT.STA.BX?.semd.sac} for ASCII
-SAC format, where NT and STA are the network code and station name,
-and \texttt{BX?} stands for the component name. Please see the Appendix~\ref{cha:Coordinates}
-and \ref{cha:channel-codes} for further details.
-
-The adjont simulations generate synthetic seismograms with the name
-\texttt{NT.S?????.S??.sem} (refer to Section \ref{sec:Adjoint-simulation-sources}
-for details). The kernel simulations output the back-reconstructed
-synthetic seismogram in the name \texttt{NT.STA.BX?.semd}, mainly
-for the purpose of checking the accuracy of the reconstruction. Refer
-to Section \ref{sec:Adjoint-simulation-finite} for further details.
-
-You do have further options to change this default output behavior,
-given in the main constants file \texttt{constants.h} located in \texttt{src/shared/}
-directory:
-\begin{description}
-\item [{\texttt{SEISMOGRAMS\_BINARY}}] set to \texttt{.true.} to have seismograms
-written out in binary format.
-\item [{\texttt{WRITE\_SEISMOGRAMS\_BY\_MASTER}}] Set to \texttt{.true.}
-to have only the master process writing out seismograms. This can
-be useful on a cluster, where only the master process node has access
-to the output directory.
-\item [{\texttt{USE\_OUTPUT\_FILES\_PATH}}] Set to \texttt{.false.} to
-have the seismograms output to \texttt{LOCAL\_PATH} directory specified
-in the main parameter file \texttt{DATA/Par\_file}. In this case,
-you could collect the synthetics onto the frontend using the \texttt{collect\_seismo\_lsf\_multi.pl}
-script located in the \texttt{utils/Cluster/lsf/} directory. The usage
-of the script would be e.g.:
-
-\begin{lyxcode}
-collect\_seismo.pl~machines~DATA/Par\_file
-\end{lyxcode}
-\end{description}
-where \texttt{machines} is a file containing the node names and \texttt{DATA/Par\_file}
-the parameter file used to extract the \texttt{LOCAL\_PATH} directory
-used for the simulation.
-
-\section{Clean Local Database}
-
-After all the simulations are done, the seismograms are collected,
-and the useful database files are copied to the frontend, you may
-need to clean the local scratch disk for the next simulation. This
-is especially important in the case of kernel simulation, where very
-large files are generated for the absorbing boundaries to help with
-the reconstruction of the regular forward wavefield. A sample script
-is provided in \texttt{utils/}:
-\begin{lyxcode}
-cleanbase.pl~machines
-\end{lyxcode}
-where \texttt{machines} is a file containing the node names.
-
-\section{Plot Movie Snapshots and Synthetic Shakemaps}
-
-\subsection{Script \texttt{movie2gif.gmt.pl}}
-
-With the movie data saved in \texttt{OUTPUT\_FILES/} at the end of
-a movie simulation (\texttt{\small MOVIE\_SURFACE=.true.}{\small ),
-you can run the }\texttt{\small `create\_movie\_shakemap\_AVS\_DX\_GMT}{\small '
-code to convert these binary movie data into GMT xyz files for futher
-processing. A sample script }\texttt{\small movie2gif.gmt.pl}{\small{}
-is provided to do this conversion, and then plot the movie snapshots
-in GMT, for example:}{\small \par}
-\begin{lyxcode}
-movie2gif.gmt.pl~-m~CMTSOLUTION~-g~-f~1/40~-n~-2~-p~
-\end{lyxcode}
-which for the first through the 40th movie frame, converts the \texttt{moviedata}
-files into GMT xyz files, interpolates them using the 'nearneighbor'
-command in GMT, and plots them on a 2D topography map. Note that `\texttt{-2}'
-and `\texttt{-p}' are both optional.
-
-\subsection{Script \texttt{plot\_shakemap.gmt.pl}}
-
-With the shakemap data saved in \texttt{OUTPUT\_FILES/} at the end
-of a shakemap simulation ~\\
- (\texttt{CREATE\_SHAKEMAP=.true.}), you can also run \texttt{`create\_movie\_shakemap\_AVS\_DX\_GMT}'
-code to convert the binary shakemap data into GMT xyz files. A sample
-script \texttt{plot\_shakemap.gmt.pl} is provided to do this conversion,
-and then plot the shakemaps in GMT, for example:
-\begin{lyxcode}
-plot\_shakemap.gmt.pl~data\_dir~type(1,2,3)~CMTSOLUTION~
-\end{lyxcode}
-where \texttt{type=1} for a displacement shakemap, \texttt{2} for
-velocity, and \texttt{3} for acceleration.
-
-\section{Map Local Database}
-
-A sample program \texttt{remap\_database} is provided to map the local
-database from a set of machines to another set of machines. This is
-especially useful when you want to run mesher and solver, or different
-types of solvers separately through a scheduler (refer to Chapter~\ref{cha:Scheduler}).
-\begin{lyxcode}
-run\_lsf.bash~-{}-gm-no-shmem~-{}-gm-copy-env~remap\_database~old\_machines~150
-\end{lyxcode}
-where \texttt{old\_machines} is the LSF machine file used in the previous
-simulation, and \texttt{150} is the number of processors in total.
-
-%============================================================
-
-\chapter{\label{cha:developers}Information for developers of the code, and for people who want to learn how the technique works}
-
-You can get a very simple 1D version of a demo code (there is one in Fortran and one in Python):\\
-
-git clone --recursive https://github.com/geodynamics/specfem1d.git\\
-
-\noindent We also have simple 3D demo source codes that implement the SEM in a single, small program, in directory\\
-utils/small\_SEM\_solvers\_in\_Fortran\_and\_C\_without\_MPI\_to\_learn of the specfem3d package.
-They are useful to learn how the spectral-element method works, and how to write or modify a code to implement it.
-Also useful to test new ideas by modifying these simple codes to run some tests.\\
-
-\noindent For information on how to contribute to the code, i.e., for how to make your modifications, additions or improvements part of the
-official package, see https://github.com/geodynamics/specfem3d/wiki .
-
-%============================================================
-
-\chapter*{\label{cha:Bug-Reports-and}Bug Reports and Suggestions for Improvements}
-
-To report bugs or suggest improvements to the code, please send an
-e-mail to the CIG Computational Seismology Mailing List \urlwithparentheses{cig-seismo@geodynamics.org}.
-
-\addcontentsline{toc}{chapter}{Notes \& Acknowledgments}
-
-\chapter*{Notes \& Acknowledgments}
-
-In order to keep the software package thread-safe in case a multithreaded
-implementation of MPI is used, developers should not add modules or
-common blocks to the source code but rather use regular subroutine
-arguments (which can be grouped in ``derived types'' if needed for
-clarity).
-
-The Gauss-Lobatto-Legendre subroutines in \texttt{gll\_library.f90}
-are based in part on software libraries from the Massachusetts Institute
-of Technology, Department of Mechanical Engineering (Cambridge, Massachusetts,
-USA). The non-structured global numbering software was provided by
-Paul F. Fischer (Brown University, Providence, Rhode Island, USA,
-now at Argonne National Laboratory, USA).
-
-OpenDX \urlwithparentheses{http://www.opendx.org} is open-source
-based on IBM Data Explorer, AVS \urlwithparentheses{http://www.avs.com}
-is a trademark of Advanced Visualization Systems, and ParaView \urlwithparentheses{http://www.paraview.org}
-is an open-source visualization platform.
-
-Please e-mail your feedback, questions, comments, and suggestions
-to the CIG Computational Seismology Mailing List \urlwithparentheses{cig-seismo@geodynamics.org}.
-
-\addcontentsline{toc}{chapter}{Copyright}
-
-\chapter*{Copyright}
-
-Main historical authors: Dimitri Komatitsch and Jeroen Tromp
-
-Princeton University, USA, and CNRS / University of Marseille, France
-
-$\copyright$ Princeton University and CNRS / University of Marseille, July 2012
-
-This program is free software; you can redistribute it and/or modify
-it under the terms of the GNU General Public License as published
-by the Free Software Foundation (see Appendix \ref{cha:License}).\\
-
-\textbf{\underline{Evolution of the code:}}\\
-
-MPI v. 3.0, December 2014: many developers.
-Convolutional PML, LDDRK time scheme, bulk attenuation support, simultaneous MPI runs,
-ADIOS file I/O support, coupling with external codes, new seismogram names,
-Deville routines for additional GLL degrees, tomography tools, unit/regression test framework,
-improved CUDA GPUs performance, additonal GEOCUBIT support, better make compilation,
-git versioning system. \\
-
-MPI v. 2.1, July 2012: Max Rietmann, Peter Messmer, Daniel Peter, Dimitri
-Komatitsch, Joseph Charles, Zhinan Xie: support for CUDA GPUs, better
-CFL stability for the Stacey absorbing conditions. \\
-
-MPI v. 2.0, November 2010: Dimitri Komatitsch, Nicolas Le Goff, Roland
-Martin and Pieyre Le Loher, University of Pau, France, Daniel Peter,
-Jeroen Tromp and the Princeton group of developers, Princeton University,
-USA, and Emanuele Casarotti, INGV Roma, Italy: support for CUBIT meshes
-decomposed by SCOTCH; much faster solver using Michel Deville's inlined
-matrix products.\\
-
-MPI v. 1.4 Dimitri Komatitsch, University of Pau, Qinya Liu and others,
-Caltech, September 2006: better adjoint and kernel calculations, faster
-and better I/Os on very large systems, many small improvements and
-bug fixes.\\
-
-MPI v. 1.3 Dimitri Komatitsch, University of Pau, and Qinya Liu, Caltech,
-July 2005: serial version, regular mesh, adjoint and kernel calculations,
-ParaView support.\\
-
-MPI v. 1.2 Min Chen and Dimitri Komatitsch, Caltech, July 2004: full
-anisotropy, volume movie.\\
-
-MPI v. 1.1 Dimitri Komatitsch, Caltech, October 2002: Zhu's Moho map,
-scaling of $V_{s}$ with depth, Hauksson's regional model, attenuation,
-oceans, movies.\\
-
-MPI v. 1.0 Dimitri Komatitsch, Caltech, USA, May 2002: first MPI version
-based on global code.\\
-
-Dimitri Komatitsch, IPG Paris, France, December 1996: first 3-D solver
-for the CM-5 Connection Machine, parallelized on 128 processors using
-Connection Machine Fortran.\\
-
-\addcontentsline{toc}{chapter}{References} \bibliography{bibliography}
-
-\appendix
-%dummy comment inserted by tex2lyx to ensure that this paragraph is not empty
-
-\chapter{\label{cha:Coordinates}Reference Frame Convention}
-
-The code uses the following convention for the Cartesian reference
-frame:
-\begin{itemize}
-\item the $x$ axis points East
-\item the $y$ axis points North
-\item the $z$ axis points up
-\end{itemize}
-Note that this convention is different from both the \citet{AkRi80}
-convention and the Harvard Centroid-Moment Tensor (CMT) convention.
-The Aki \& Richards convention is
-\begin{itemize}
-\item the $x$ axis points North
-\item the $y$ axis points East
-\item the $z$ axis points down
-\end{itemize}
-and the Harvard CMT convention is
-\begin{itemize}
-\item the $x$ axis points South
-\item the $y$ axis points East
-\item the $z$ axis points up
-\end{itemize}
-
-\subsection*{Source and receiver locations}
-
-The SPECFEM3D Cartesian software code internally uses Cartesian coordinates.
-The given locations for sources and receiver locations thus may get
-converted. Sources and receiver locations are read in from the \texttt{CMTSOLUTION}
-(or \texttt{FORCESOLUTION}) and \texttt{STATIONS} files. Note that
-e.g. the \texttt{CMTSOLUTION} and \texttt{FORCESOLUTION} files denotes
-the location by \textquotedbl{}longitude/latitude/depth\textquotedbl{}.
-We read in longitude as $x$ coordinate, latitude as $y$ coordinate.
-
-In case the flag \texttt{SUPPRESS\_UTM\_PROJECTION} is set to \texttt{.false.}
-in the main parameter file (see Chapter~\ref{cha:Creating-Distributed-Databases}),
-the $x$/$y$ coordinates have to be given in degrees and are converted
-to Cartesian coordinates using a UTM conversion with the specified
-UTM zone.
-
-The value for depth (given in $km$ in \texttt{CMTSOLUTION} and \texttt{FORCESOLUTION})
-or burial depth (given in $m$ in \texttt{STATIONS}) is evaluated
-with respect to the surface of the mesh at the specified $x$/$y$
-location to find the corresponding $z$ coordinate. It is possible
-to use this depth value directly as $z$ coordinate by changing the
-flag \texttt{USE\_SOURCES\_RECVS\_Z} to \texttt{.true.} in the file
-\texttt{constants.h} located in the \texttt{src/shared/} subdirectory.
-
-\subsection*{Seismogram outputs}
-
-The seismogram output directions are given in Cartesian $x$/$y$/$z$
-directions and not rotated any further. Changing flags in \texttt{constants.h}
-in the \texttt{src/shared/} subdirectory only rotates the seismogram
-outputs if receivers are forced to be located at the surface (\texttt{RECVS\_CAN\_BE\_BURIED\_EXT\_MESH}
-set to \texttt{.false.}) and the normal to the surface at the receiver
-location should be used (\texttt{EXT\_MESH\_RECV\_NORMAL} set to \texttt{.true.})
-as vertical. In this case, the outputs are rotated to have the vertical
-component normal to the surface of the mesh, $x$ and $y$ directions
-are somewhat arbitrary orthogonal directions along the surface.
-
-For the labeling of the seismogram channels, see Appendix~\ref{cha:channel-codes}.
-Additionally, we add labels to the synthetics using the following
-convention: For a receiver station located in an
-\begin{description}
-\item [{elastic domain:}] ~\\
-
-\begin{itemize}
-\item \texttt{semd} for the displacement vector
-\item \texttt{semv} for the velocity vector
-\item \texttt{sema} for the acceleration vector
-\end{itemize}
-\item [{acoustic domain:}] ~\\
- (please note that receiver stations in acoustic domains must be buried.
-This is due to the free surface condition which enforces a zero pressure
-at the surface.)
-
-\begin{itemize}
-\item \texttt{semd} for the displacement vector
-\item \texttt{semv} for the velocity vector
-\item \texttt{sema} for pressure which will be stored in the vertical component
-\texttt{Z}. Note that pressure in the acoustic media is isotropic
-and thus the pressure record would be the same in the other two component
-directions. We therefore use the other two seismogram components to
-store the acoustic potential in component \texttt{X} (or \texttt{N})
-and the first time derivative of the acoustic potential in component
-\texttt{Y} (or \texttt{E}).
-\end{itemize}
-\end{description}
-The seismograms are by default written out in ASCII-format to the
-\texttt{OUTPUT\_FILES/} subdirectory by each parallel process. You
-can change this behavior by changing the following flags in the \texttt{constants.h}
-file located in the \texttt{src/shared/} subdirectory:
-\begin{description}
-\item [{\texttt{SEISMOGRAMS\_BINARY}}] set to \texttt{.true.} to have seismograms
-written out in binary format.
-\item [{\texttt{WRITE\_SEISMOGRAMS\_BY\_MASTER}}] Set to \texttt{.true.}
-to have only the master process writing out seismograms. This can
-be useful on a cluster, where only the master process node has access
-to the output directory.
-\item [{\texttt{USE\_OUTPUT\_FILES\_PATH}}] Set to \texttt{.false.} to
-have the seismograms output to \texttt{LOCAL\_PATH} directory specified
-in the main parameter file \texttt{DATA/Par\_file}.
-\end{description}
-
-\chapter{\label{cha:channel-codes}Channel Codes of Seismograms}
+\chapter{Channel Codes of Seismograms}\label{cha:channel-codes}
 
@@
 \noindent \begin{centering}
-\includegraphics[scale=0.6]{figures/IRIS_band_codes}
+\includegraphics[scale=0.6]{figures/IRIS_band_codes.pdf}
 \par\end{centering}
@@
 s.
-
-\chapter{\label{cha:Troubleshooting}Troubleshooting}
-
-\section*{FAQ}
-\begin{description}
-\item [{configuration fails:}]~
-
-Examine the log file 'config.log'. It contains detailed informations.
-In many cases, the paths to these specific compiler commands F90,
-CC and MPIF90 will not be correct if `./configure` fails.
-
-Please make sure that you have a working installation of a Fortran
-compiler, a C compiler and an MPI implementation. You should be able
-to compile this little program code:
-\begin{lyxcode}
-program~main~~~\\
-~include~'mpif.h'~~~\\
-~integer,~parameter~::~CUSTOM\_MPI\_TYPE~=~MPI\_REAL~~~\\
-~integer~ier~~~\\
-~call~MPI\_INIT(ier)~~~\\
-~call~MPI\_BARRIER(MPI\_COMM\_WORLD,ier)~~~\\
-~call~MPI\_FINALIZE(ier)~~~\\
-~end
-\end{lyxcode}
-\item [{compilation fails stating:}] ~\\
-
-\begin{lyxcode}
-...~~~\\
-~obj/program\_generate\_databases.o:~In~function~`MAIN\_\_':~~~\\
-~program\_generate\_databases.f90:(.text+0x14):~undefined~reference~to~`\_gfortran\_set\_std'~~~\\
-~...~~~\\
-
-\end{lyxcode}
-
-Make sure you're pointing to the right 'mpif90' wrapper command.
-
-Normally, this message will appear when you are mixing two different
-Fortran compilers. That is, using e.g. gfortran to compile non-MPI
-files and mpif90, wrapper provided for e.g. ifort, to compile MPI-files.
-
-fix: e.g. specify > ./configure FC=gfortran MPIF90=/usr/local/openmpi-gfortran/bin/mpif90
-
-\item [{after executing \texttt{xmeshfem3D} I've got elements with skewness of 81\% percent, what does this mean:}] Look
-at the skewness table printed in the \texttt{output\_mesher.txt} file
-after executing \texttt{xmeshfem3D} for the example given in \texttt{EXAMPLES/meshfem3D\_examples/simple\_model/}:
-
-\begin{lyxcode}
-...~~\\
-~histogram~of~skewness~(0.~good~-~1.~bad):~~~\\
-~0.0000000E+00~-~5.0000001E-02~27648~81.81818~\%~~~\\
-~5.0000001E-02~-~0.1000000~0~0.0000000E+00~\%~~~\\
-~...~~~\\
-
-\end{lyxcode}
-
-The first line means that you have 27,648 elements with a skewness
-value between 0 and 0.05 (which means the element is basically not
-skewed, just plain regular hexahedral element). The total number of
-elements you have in this mesh is (see in the \texttt{output\_mesher.txt}
-file a bit further down):
-\begin{lyxcode}
-...~~\\
-~total~number~of~elements~in~entire~mesh:~33792~~~\\
-~...~~~\\
-
-\end{lyxcode}
-
-which gives you that: 27,648 / 33,792 $\sim$ 81.8 \% of all elements
-are not skewed, i.e. regular elements. a fantastic value :)
-
-The histogram lists for this mesh also some stronger skewed elements,
-for example the worst ones belong to:
-\begin{lyxcode}
-...~~\\
-~0.6000000~-~0.6500000~2048~6.060606~\%~~~\\
-~...~~~\\
-
-\end{lyxcode}
-
-about 6 \% of all elements have distortions with a skewness value
-between 0.6 and 0.65. The skewness values give you a hint of how good
-your mesh is. In an ideal world, you would want to have no distortions,
-just like the 81\% from above. Those elements give you the best approximate
-values by the GLL quadrature used in the spectral-element method.
-However, having weakly distorted elements is still fine and the solutions
-are still accurate enough. So empirically, values up to around 0.7
-are tolerable, above that you should consider remeshing...
-
-To give you an idea why some of the elements are distorted, see the
-following figure \ref{fig:mesh.vp} of the mesh you obtain in the
-example \texttt{EXAMPLES/meshfem3D\_examples/simple\_model/}.
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=0.9\textwidth]{figures/mesh_vp}
-\par\end{centering}
-
-\caption{Paraview visualization using the mesh vtk-files for the example given
-in \texttt{EXAMPLES/meshfem3D\_examples/simple\_model/}.}
-
-\label{fig:mesh.vp}
-\end{figure}
-
-You will see that the mesh contains a doubling layer, where we stitch
-elements together such that the size of two elements will transition
-to the size of one element (very useful to keep the ratio of wavespeed
-/ element\_size about constant). Those elements in this doubling layer
-have higher skewness values and make up those 6 \% in the histogram.
-
-\item [{the code gives following error message "need at least one receiver":}] This
-means that no stations given in the input file \texttt{DATA/STATIONS}
-could be located within the dimensions of the mesh. This can happen
-for example when the mesh was created with the in-house mesher \texttt{xmeshfem3D}
-while using the Universal Transverse Mercator (UTM) projection but
-the simulation with \texttt{xspecfem3D} was suppressing this projection
-from latitude/longitude to x/y/z coordinates.
-
-In such cases, try to change your \texttt{DATA/Par\_file} and set
-e.g.:
-\begin{lyxcode}
-SUPPRESS\_UTM\_PROJECTION~=~.false.~~~\\
-
-\end{lyxcode}
-
-to be the same in \texttt{Mesh\_Par\_file} and \texttt{Par\_file}.
-This flag should be identical when using the in-house mesher \texttt{xmeshfem3D},
-\texttt{xgenerate\_databases} and \texttt{xspecfem3D} together to
-run simulations.
-
-The flag determines if the coordinates you specify for your source
-and station locations are given as lat/lon degrees and must be converted
-to UTM coordinates. As an example, if you use \texttt{.false.} within
-\texttt{Mesh\_Par\_file} then you create a mesh with \texttt{xmeshfem3D}
-using the UTM projection from lat/lon as input format to UTM projected
-coordinates to store the mesh point positions, which is fine. The
-error then may occur if in the \texttt{Par\_file} you have this set
-to \texttt{.true.} so that the \texttt{xgenerate\_databases} and \texttt{xspecfem3D}
-suppress the UTM projection and assume that all coordinates you use
-now for source and receiver locations are given in meters (that is,
-converted) already. So it won't find the specified locations in the
-used mesh. As a solutions, just change the flag in \texttt{Par\_file}
-to be the same as in \texttt{Mesh\_Par\_file} and rerun \texttt{xgenerate\_databases}
-and \texttt{xspecfem3D} to make sure that your simulation works fine.
-
-\item [{I get the following error message "forward simulation became unstable and blew up":}] In
-most cases this means that your time step size \texttt{DT} is chosen
-too big. Look at your files \texttt{output\_mesher.txt} or \texttt{output\_solver.txt}
-created in the folder \texttt{OUTPUT\_FILES}. In these output files,
-find the section:
-
-\begin{lyxcode}
-...~~~\\
-~{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}~~~\\
-~{*}{*}{*}~Verification~of~simulation~parameters~{*}{*}{*}~~~\\
-~{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}~~~\\
-~...~~~\\
-~{*}{*}{*}~Minimum~period~resolved~=~4.308774~~~\\
-~{*}{*}{*}~Maximum~suggested~time~step~=~6.8863556E-02~~~\\
-~...~~~\\
-~...~~~\\
-
-\end{lyxcode}
-
-then change \texttt{DT} in the \texttt{DATA/Par\_file} to be somewhere
-close to the maximum suggested time step. In the example above: \\
- \texttt{DT = 0.05d0} \\
- would (most probably) work fine. It could be also bigger than the
-0.068 s suggested. This depends a bit on the distortions of your mesh
-elements. The more regular they are, the bigger you can choose \texttt{DT}.
-Just play with this value a bit and see when the simulation becomes
-stable ...
-
-\end{description}
-
-\chapter{\label{cha:License}License }
-
-\textbf{GNU GENERAL PUBLIC LICENSE Version 2, June 1991. Copyright
-(C) 1989, 1991 Free Software Foundation, Inc. 59 Temple Place, Suite
-330, Boston, MA 02111-1307 USA} \\
- Everyone is permitted to copy and distribute verbatim copies of this
-license document, but changing it is not allowed.
-
-\section*{Preamble}
-
-The licenses for most software are designed to take away your freedom
-to share and change it. By contrast, the GNU General Public License
-is intended to guarantee your freedom to share and change free software
--- to make sure the software is free for all its users. This General
-Public License applies to most of the Free Software Foundation's software
-and to any other program whose authors commit to using it. (Some other
-Free Software Foundation software is covered by the GNU Library General
-Public License instead.) You can apply it to your programs, too.
-
-When we speak of free software, we are referring to freedom, not price.
-Our General Public Licenses are designed to make sure that you have
-the freedom to distribute copies of free software (and charge for
-this service if you wish), that you receive source code or can get
-it if you want it, that you can change the software or use pieces
-of it in new free programs; and that you know you can do these things.
-
-To protect your rights, we need to make restrictions that forbid anyone
-to deny you these rights or to ask you to surrender the rights. These
-restrictions translate to certain responsibilities for you if you
-distribute copies of the software, or if you modify it.
-
-For example, if you distribute copies of such a program, whether gratis
-or for a fee, you must give the recipients all the rights that you
-have. You must make sure that they, too, receive or can get the source
-code. And you must show them these terms so they know their rights.
-
-We protect your rights with two steps:
-\begin{enumerate}
-\item Copyright the software, and
-\item Offer you this license which gives you legal permission to copy, distribute
-and/or modify the software.
-\end{enumerate}
-Also, for each author's protection and ours, we want to make certain
-that everyone understands that there is no warranty for this free
-software. If the software is modified by someone else and passed on,
-we want its recipients to know that what they have is not the original,
-so that any problems introduced by others will not reflect on the
-original authors' reputations.
-
-Finally, any free program is threatened constantly by software patents.
-We wish to avoid the danger that redistributors of a free program
-will individually obtain patent licenses, in effect making the program
-proprietary. To prevent this, we have made it clear that any patent
-must be licensed for everyone's free use or not licensed at all.
-
-The precise terms and conditions for copying, distribution and modification
-follow.
-
-\section*{GNU GENERAL PUBLIC LICENSE TERMS AND CONDITIONS FOR COPYING, DISTRIBUTION
-AND MODIFICATION }
-\begin{itemize}
-\item 0.This License applies to any program or other work which contains
-a notice placed by the copyright holder saying it may be distributed
-under the terms of this General Public License. The ``Program''
-below refers to any such program or work, and a ``work based on the
-Program'' means either the Program or any derivative work under copyright
-law: that is to say, a work containing the Program or a portion of
-it, either verbatim or with modifications and/or translated into another
-language. (Hereinafter, translation is included without limitation
-in the term ``modification.'') Each licensee is addressed as ``you.''\\
- \\
- Activities other than copying, distribution and modification are
-not covered by this License; they are outside its scope. The act of
-running the Program is not restricted, and the output from the Program
-is covered only if its contents constitute a work based on the Program
-(independent of having been made by running the Program). Whether
-that is true depends on what the Program does.\end{itemize}
-\begin{enumerate}
-\item You may copy and distribute verbatim copies of the Program's source
-code as you receive it, in any medium, provided that you conspicuously
-and appropriately publish on each copy an appropriate copyright notice
-and disclaimer of warranty; keep intact all the notices that refer
-to this License and to the absence of any warranty; and give any other
-recipients of the Program a copy of this License along with the Program.
-
-You may charge a fee for the physical act of transferring a copy,
-and you may at your option offer warranty protection in exchange for
-a fee.
-
-\item You may modify your copy or copies of the Program or any portion of
-it, thus forming a work based on the Program, and copy and distribute
-such modifications or work under the terms of Section 1 above, provided
-that you also meet all of these conditions:
-
-\begin{enumerate}
-\item You must cause the modified files to carry prominent notices stating
-that you changed the files and the date of any change.
-\item You must cause any work that you distribute or publish, that in whole
-or in part contains or is derived from the Program or any part thereof,
-to be licensed as a whole at no charge to all third parties under
-the terms of this License.
-\item If the modified program normally reads commands interactively when
-run, you must cause it, when started running for such interactive
-use in the most ordinary way, to print or display an announcement
-including an appropriate copyright notice and a notice that there
-is no warranty (or else, saying that you provide a warranty) and that
-users may redistribute the program under these conditions, and telling
-the user how to view a copy of this License. (Exception: if the Program
-itself is interactive but does not normally print such an announcement,
-your work based on the Program is not required to print an announcement.)
-\end{enumerate}
-
-These requirements apply to the modified work as a whole. If identifiable
-sections of that work are not derived from the Program, and can be
-reasonably considered independent and separate works in themselves,
-then this License, and its terms, do not apply to those sections when
-you distribute them as separate works. But when you distribute the
-same sections as part of a whole which is a work based on the Program,
-the distribution of the whole must be on the terms of this License,
-whose permissions for other licensees extend to the entire whole,
-and thus to each and every part regardless of who wrote it.
-
-Thus, it is not the intent of this section to claim rights or contest
-your rights to work written entirely by you; rather, the intent is
-to exercise the right to control the distribution of derivative or
-collective works based on the Program.
-
-In addition, mere aggregation of another work not based on the Program
-with the Program (or with a work based on the Program) on a volume
-of a storage or distribution medium does not bring the other work
-under the scope of this License.
-
-\item You may copy and distribute the Program (or a work based on it, under
-Section 2) in object code or executable form under the terms of Sections
-1 and 2 above provided that you also do one of the following:
-
-\begin{enumerate}
-\item Accompany it with the complete corresponding machine-readable source
-code, which must be distributed under the terms of Sections 1 and
-2 above on a medium customarily used for software interchange; or,
-\item Accompany it with a written offer, valid for at least three years,
-to give any third party, for a charge no more than your cost of physically
-performing source distribution, a complete machine-readable copy of
-the corresponding source code, to be distributed under the terms of
-Sections 1 and 2 above on a medium customarily used for software interchange;
-or,
-\item Accompany it with the information you received as to the offer to
-distribute corresponding source code. (This alternative is allowed
-only for noncommercial distribution and only if you received the program
-in object code or executable form with such an offer, in accord with
-Subsection b above.)
-\end{enumerate}
-
-The source code for a work means the preferred form of the work for
-making modifications to it. For an executable work, complete source
-code means all the source code for all modules it contains, plus any
-associated interface definition files, plus the scripts used to control
-compilation and installation of the executable. However, as a special
-exception, the source code distributed need not include anything that
-is normally distributed (in either source or binary form) with the
-major components (compiler, kernel, and so on) of the operating system
-on which the executable runs, unless that component itself accompanies
-the executable.
-
-If distribution of executable or object code is made by offering access
-to copy from a designated place, then offering equivalent access to
-copy the source code from the same place counts as distribution of
-the source code, even though third parties are not compelled to copy
-the source along with the object code.
-
-\item You may not copy, modify, sublicense, or distribute the Program except
-as expressly provided under this License. Any attempt otherwise to
-copy, modify, sublicense or distribute the Program is void, and will
-automatically terminate your rights under this License. However, parties
-who have received copies, or rights, from you under this License will
-not have their licenses terminated so long as such parties remain
-in full compliance.
-\item You are not required to accept this License, since you have not signed
-it. However, nothing else grants you permission to modify or distribute
-the Program or its derivative works. These actions are prohibited
-by law if you do not accept this License. Therefore, by modifying
-or distributing the Program (or any work based on the Program), you
-indicate your acceptance of this License to do so, and all its terms
-and conditions for copying, distributing or modifying the Program
-or works based on it.
-\item Each time you redistribute the Program (or any work based on the Program),
-the recipient automatically receives a license from the original licensor
-to copy, distribute or modify the Program subject to these terms and
-conditions. You may not impose any further restrictions on the recipients'
-exercise of the rights granted herein. You are not responsible for
-enforcing compliance by third parties to this License.
-\item If, as a consequence of a court judgment or allegation of patent infringement
-or for any other reason (not limited to patent issues), conditions
-are imposed on you (whether by court order, agreement or otherwise)
-that contradict the conditions of this License, they do not excuse
-you from the conditions of this License. If you cannot distribute
-so as to satisfy simultaneously your obligations under this License
-and any other pertinent obligations, then as a consequence you may
-not distribute the Program at all. For example, if a patent license
-would not permit royalty-free redistribution of the Program by all
-those who receive copies directly or indirectly through you, then
-the only way you could satisfy both it and this License would be to
-refrain entirely from distribution of the Program.
-
-If any portion of this section is held invalid or unenforceable under
-any particular circumstance, the balance of the section is intended
-to apply and the section as a whole is intended to apply in other
-circumstances.
-
-It is not the purpose of this section to induce you to infringe any
-patents or other property right claims or to contest validity of any
-such claims; this section has the sole purpose of protecting the integrity
-of the free software distribution system, which is implemented by
-public license practices. Many people have made generous contributions
-to the wide range of software distributed through that system in reliance
-on consistent application of that system; it is up to the author/donor
-to decide if he or she is willing to distribute software through any
-other system and a licensee cannot impose that choice.
-
-This section is intended to make thoroughly clear what is believed
-to be a consequence of the rest of this License.
-
-\item If the distribution and/or use of the Program is restricted in certain
-countries either by patents or by copyrighted interfaces, the original
-copyright holder who places the Program under this License may add
-an explicit geographical distribution limitation excluding those countries,
-so that distribution is permitted only in or among countries not thus
-excluded. In such case, this License incorporates the limitation as
-if written in the body of this License.
-\item The Free Software Foundation may publish revised and/or new versions
-of the General Public License from time to time. Such new versions
-will be similar in spirit to the present version, but may differ in
-detail to address new problems or concerns.
-
-Each version is given a distinguishing version number. If the Program
-specifies a version number of this License which applies to it and
-``any later version,'' you have the option of following the terms
-and conditions either of that version or of any later version published
-by the Free Software Foundation. If the Program does not specify a
-version number of this License, you may choose any version ever published
-by the Free Software Foundation.
-
-\item If you wish to incorporate parts of the Program into other free programs
-whose distribution conditions are different, write to the author to
-ask for permission. For software which is copyrighted by the Free
-Software Foundation, write to the Free Software Foundation; we sometimes
-make exceptions for this. Our decision will be guided by the two goals
-of preserving the free status of all derivatives of our free software
-and of promoting the sharing and reuse of software generally.
-\end{enumerate}
-
-\subsection*{NO WARRANTY }
-\begin{itemize}
-\item 11.BECAUSE THE PROGRAM IS LICENSED FREE OF CHARGE, THERE IS NO WARRANTY
-FOR THE PROGRAM, TO THE EXTENT PERMITTED BY APPLICABLE LAW. EXCEPT
-WHEN OTHERWISE STATED IN WRITING THE COPYRIGHT HOLDERS AND/OR OTHER
-PARTIES PROVIDE THE PROGRAM ``AS IS'' WITHOUT WARRANTY OF ANY KIND,
-EITHER EXPRESSED OR IMPLIED, INCLUDING, BUT NOT LIMITED TO, THE IMPLIED
-WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE.
-THE ENTIRE RISK AS TO THE QUALITY AND PERFORMANCE OF THE PROGRAM IS
-WITH YOU. SHOULD THE PROGRAM PROVE DEFECTIVE, YOU ASSUME THE COST
-OF ALL NECESSARY SERVICING, REPAIR OR CORRECTION.
-\item 12.IN NO EVENT UNLESS REQUIRED BY APPLICABLE LAW OR AGREED TO IN WRITING
-WILL ANY COPYRIGHT HOLDER, OR ANY OTHER PARTY WHO MAY MODIFY AND/OR
-REDISTRIBUTE THE PROGRAM AS PERMITTED ABOVE, BE LIABLE TO YOU FOR
-DAMAGES, INCLUDING ANY GENERAL, SPECIAL, INCIDENTAL OR CONSEQUENTIAL
-DAMAGES ARISING OUT OF THE USE OR INABILITY TO USE THE PROGRAM (INCLUDING
-BUT NOT LIMITED TO LOSS OF DATA OR DATA BEING RENDERED INACCURATE
-OR LOSSES SUSTAINED BY YOU OR THIRD PARTIES OR A FAILURE OF THE PROGRAM
-TO OPERATE WITH ANY OTHER PROGRAMS), EVEN IF SUCH HOLDER OR OTHER
-PARTY HAS BEEN ADVISED OF THE POSSIBILITY OF SUCH DAMAGES.
-\end{itemize}
-
-\section*{END OF TERMS AND CONDITIONS }
-
-\subsection*{How to Apply These Terms to Your New Programs}
-
-If you develop a new program, and you want it to be of the greatest
-possible use to the public, the best way to achieve this is to make
-it free software which everyone can redistribute and change under
-these terms.
-
-To do so, attach the following notices to the program. It is safest
-to attach them to the start of each source file to most effectively
-convey the exclusion of warranty; and each file should have at least
-the ``copyright'' line and a pointer to where the full notice is
-found. For example:
-\begin{quote}
-One line to give the program's name and a brief idea of what it does.
-Copyright {\footnotesize $\copyright$ (}year) (name of author)
-
-This program is free software; you can redistribute it and/or modify
-it under the terms of the GNU General Public License as published
-by the Free Software Foundation; either version 2 of the License,
-or (at your option) any later version.
-
-This program is distributed in the hope that it will be useful, but
-WITHOUT ANY WARRANTY; without even the implied warranty of MERCHANTABILITY
-or FITNESS FOR A PARTICULAR PURPOSE. See the GNU General Public License
-for more details.
-
-You should have received a copy of the GNU General Public License
-along with this program; if not, write to the Free Software Foundation,
-Inc., 59 Temple Place, Suite 330, Boston, MA 02111-1307 USA
-\end{quote}
-Also add information on how to contact you by electronic and paper
-mail.
-
-If the program is interactive, make it output a short notice like
-this when it starts in an interactive mode:
-\begin{quote}
-Gnomovision version 69, Copyright $\copyright$ year name of author
-Gnomovision comes with ABSOLUTELY NO WARRANTY; for details type `show
-w'. This is free software, and you are welcome to redistribute it
-under certain conditions; type `show c' for details.
-\end{quote}
-The hypothetical commands `show w' and `show c' should show the appropriate
-parts of the General Public License. Of course, the commands you use
-may be called something other than `show w' and `show c'; they could
-even be mouse-clicks or menu items -- whatever suits your program.
-
-You should also get your employer (if you work as a programmer) or
-your school, if any, to sign a ``copyright disclaimer'' for the
-program, if necessary. Here is a sample; alter the names:
-\begin{quote}
-Yoyodyne, Inc., hereby disclaims all copyright interest in the program
-`Gnomovision' (which makes passes at compilers) written by James Hacker.
-
-(signature of Ty Coon)\\
- 1 April 1989 \\
- Ty Coon, President of Vice
-\end{quote}
-This General Public License does not permit incorporating your program
-into proprietary programs. If your program is a subroutine library,
-you may consider it more useful to permit linking proprietary applications
-with the library. If this is what you want to do, use the GNU Library
-General Public License instead of this License.
-\end{document}

### 2015-04-24 doc/USER_MANUAL/manual_SPECFEM3D_Cartesian.tex
updates user manual to use facilitate the generation of a markdown version to the github wiki

--- 2015-04-18 doc/USER_MANUAL/manual_SPECFEM3D_Cartesian.tex
@@
 % figures
-\usepackage[dvips]{epsfig}
+%\usepackage[dvips]{epsfig}
 
@@
 
+% some of Dimitri's aliases for math symbols
+\newcommand{\Grad}{\mbox{\boldmath $\bf \nabla$}}
+\newcommand{\Div}{\Grad \cdot }
+\newcommand{\Laplacian}{\nabla^2}
+\newcommand{\pressure}{p}
+\newcommand{\Int}{\displaystyle\int}
+\newcommand{\Sum}{\displaystyle\sum}
+\renewcommand{\epsilon}{\varepsilon}
+\newcommand{\myiomega}{\mathbf{i} \, \omega}
+\newcommand{\p}{\partial}
+\newcommand{\boldu}{\mathbf{u}}
+\newcommand{\boldv}{\mathbf{v}}
+\newcommand{\boldw}{\mathbf{w}}
+\newcommand{\boldx}{\mathbf{x}}
+\newcommand{\boldy}{\mathbf{y}}
+\newcommand{\boldm}{\mathbf{m}}
+\newcommand{\boldp}{\mathbf{p}}
+\newcommand{\boldg}{\mathbf{g}}
+\newcommand{\bolds}{\mathbf{s}}
+\newcommand{\boldn}{\mathbf{\hat{n}}}
+\newcommand{\boldd}{\mathbf{d}}
+\newcommand{\boldf}{\mathbf{f}}
+\newcommand{\boldB}{\mathbf{B}}
+\newcommand{\boldc}{\mathbf{c}}
+\newcommand{\boldC}{\mathbf{C}}
+\newcommand{\boldE}{\mathbf{E}}
+\newcommand{\boldH}{\mathbf{H}}
+\newcommand{\boldI}{\mathbf{I}}
+\newcommand{\boldsigma}{\mbox{\boldmath $\bf \sigma$}}
+\newcommand{\boldxi}{\mbox{\boldmath $\bf \xi$}}
+\newcommand{\boldvarepsilon}{\mbox{\boldmath $\bf \varepsilon$}}
+\newcommand{\bolddelta}{\mbox{\boldmath $\bf \delta$}}
+\newcommand{\boldtau}{\mbox{\boldmath $\bf \tau$}}
+\newcommand{\Boxmath}{\Lambda}
+\newcommand{\rmd}{\, \mathrm{d}}
+\newcommand{\pv}{\ ,}
+\newcommand{\pp}{\ .}
+
 \usepackage{babel}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \begin{document}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% TITLE page
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{center}
@@
 \makebox[1\textwidth]{%
-\includegraphics[width=0.83\paperwidth]{figures/specfem_3d_Cartesian-cover} %
+\includegraphics[width=0.83\paperwidth]{figures/specfem_3d_Cartesian-cover.pdf} %
 }
@@
 
-\author{$\copyright$ Princeton University (USA), CNRS and University of Marseille (France),\\
+\author{\copyright Princeton University (USA), CNRS and University of Marseille (France),\\
 and ETH Z\"urich (Switzerland)\\
@@
 
-\section*{Authors}
-
-\noindent The SPECFEM3D Cartesian package was first developed by Dimitri
-Komatitsch and Jean-Pierre Vilotte at Institut de Physique du Globe
-(IPGP) in Paris, France from 1995 to 1997 and then by Dimitri Komatitsch
-and Jeroen Tromp at Harvard University and Caltech, USA, starting in 1998.
-The story started on April 4, 1995, when Prof. Yvon Maday from CNRS and University of Paris, France, gave a lecture to
-Dimitri Komatitsch and Jean-Pierre Vilotte at IPG about the nice properties of the Legendre spectral-element method with diagonal mass matrix that he had used for
-other equations. We are deeply indebted and thankful to him for that.
-That followed a visit by Dimitri Komatitsch to OGS (Istituto Nazionale di Oceanografia e di Geofisica Sperimentale) in Trieste, Italy, in February 1995
-to meet with G\'eza Seriani and Enrico Priolo, who introduced him to their 2D Chebyshev version of the spectral-element method with a non-diagonal mass matrix.
-We are deeply indebted and thankful to them for that.
-
-%%%%%%%% please do NOT merge the lines of the paragraph below, it makes it difficult to insert new names in alphabetical order; leave a single name per line
-%%%%%%%% please do NOT merge the lines of the paragraph below, it makes it difficult to insert new names in alphabetical order; leave a single name per line
-%%%%%%%% please do NOT merge the lines of the paragraph below, it makes it difficult to insert new names in alphabetical order; leave a single name per line
-Since then it has been developed and maintained by a development team:
-in alphabetical order,
-Jean-Paul (Pablo) Ampuero,
-Kangchen Bai,
-Piero Basini,
-C\'eline Blitz,
-Alexis Bottero,
-Ebru Bozda\u{g},
-Emanuele Casarotti,
-Joseph Charles,
-Min Chen,
-Paul Cristini,
-Percy Galvez,
-Dominik G\"oddeke,
-Vala Hj\"orleifsd\'ottir,
-Sue Kientz,
-Dimitri Komatitsch,
-Jes\'us Labarta,
-Nicolas Le Goff,
-Pieyre Le Loher,
-Matthieu Lefebvre,
-Qinya Liu,
-Yang Luo,
-Alessia Maggi,
-Federica Magnoni,
-Roland Martin,
-Ren\'e Matzen,
-Dennis McRitchie,
-Matthias Meschede,
-Peter Messmer,
-David Mich\'ea,
-Surendra Nadh Somala,
-Tarje Nissen-Meyer,
-Daniel Peter,
-Max Rietmann,
-Elliott Sales de Andrade,
-Brian Savage,
-Bernhard Schuberth,
-Anne Sieminski,
-Leif Strand,
-Carl Tape,
-Jeroen Tromp,
-Jean-Pierre Vilotte,
-Zhinan Xie,
-Chang-Hua Zhang,
-Hejun Zhu.\\
-
-The cover graphic of the manual was created by Santiago Lombeyda from
-Caltech's Center for Advanced Computing Research (CACR) \urlwithparentheses{http://www.cacr.caltech.edu}.\\
-
-The code is released open-source under the GNU version 2+ license, see the license at the end of this manual.\\
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\include{authors}
 
 \newpage{}
-\section*{Current and past main participants or main sponsors of the SPECFEM project (in no particular order)}
-%
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=0.125\textwidth]{figures/logo_cnrs}\vspace*{2truemm}
-\includegraphics[width=0.125\textwidth]{figures/logo_princeton}\vspace*{2truemm}
-\includegraphics[width=0.15\textwidth]{figures/logo_aix_marseille_universite}\vspace*{0.02truemm}
-\includegraphics[width=0.15\textwidth]{figures/logo_ETH}\vspace*{2truemm}
-\includegraphics[width=0.12\textwidth]{figures/logo_CSC_China}\vspace*{0.02truemm}
-\includegraphics[width=0.15\textwidth]{figures/logo_inria}\vspace*{2truemm}
-\includegraphics[width=0.14\textwidth]{figures/logo_UPPA}
-\par\end{centering}
-%
-\vspace*{-2truemm}
-%
-\noindent \begin{centering}
-\includegraphics[width=0.15\textwidth]{figures/logo_NVIDIA}\vspace*{2truemm}
-\includegraphics[width=0.125\textwidth]{figures/logo_IUF}\vspace*{2truemm}
-\includegraphics[width=0.135\textwidth]{figures/logo_Caltech}\vspace*{2truemm}
-\includegraphics[width=0.135\textwidth]{figures/logo_Harvard}\vspace*{2truemm}
-\includegraphics[width=0.135\textwidth]{figures/logo_IPGP}\vspace*{2truemm}
-\includegraphics[width=0.112\textwidth]{figures/logo_ANR}\vspace*{2truemm}
-\includegraphics[width=0.112\textwidth]{figures/logo_NSF}\vspace*{2truemm}
-\par\end{centering}
-%
-\vspace*{-2truemm}
-%
-\noindent \begin{centering}
-\includegraphics[width=0.190\textwidth]{figures/logo_European_Union}\vspace*{2truemm}
-\includegraphics[width=0.112\textwidth]{figures/logo_GENCI}\vspace*{2truemm}
-\includegraphics[width=0.112\textwidth]{figures/logo_PRACE}\vspace*{2truemm}
-\includegraphics[width=0.112\textwidth]{figures/logo_CINES}\vspace*{2truemm}
-\includegraphics[width=0.130\textwidth]{figures/logo_Oak_Ridge}\vspace*{2truemm}
-\hspace*{3mm}\includegraphics[width=0.130\textwidth]{figures/logo_fondation_Del_Duca}
-\includegraphics[width=0.130\textwidth]{figures/logo_CIG}\vspace*{2truemm}
-\par\end{centering}
-%
-\vspace*{-2truemm}
-%
-\noindent \begin{centering}
-\includegraphics[width=0.112\textwidth]{figures/logo_University_of_Toronto}\vspace*{2truemm}
-\includegraphics[width=0.109\textwidth]{figures/logo_INGV}\vspace*{2truemm}
-\includegraphics[width=0.102\textwidth]{figures/logo_Univ_Toulouse}\vspace*{2truemm}
-\includegraphics[width=0.112\textwidth]{figures/logo_TOTAL}\vspace*{2truemm}
-\includegraphics[width=0.130\textwidth]{figures/logo_Fairbanks}\vspace*{2truemm}
-\includegraphics[width=0.120\textwidth]{figures/logo_CINECA}\vspace*{2truemm}
-\includegraphics[width=0.140\textwidth]{figures/logo_Intel_Exascale_Labs}\vspace*{2truemm}
-\includegraphics[width=0.130\textwidth]{figures/logo_Maison_Simulation}
-\par\end{centering}
-\end{figure}
+
+\include{sponsors}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% Table of contents
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
-\chapter{Introduction}
-
-The software package SPECFEM3D Cartesian simulates seismic wave propagation
-at the local or regional scale and performs full waveform imaging (FWI) or adjoint tomography based upon the spectral-element method
-(SEM). The SEM is a continuous Galerkin technique \citep{TrKoLi08,PeKoLuMaLeCaLeMaLiBlNiBaTr11},
-which can easily be made discontinuous \citep{BeMaPa94,Ch00,KoWoHu02,ChCaVi03,LaWaBe05,Kop06,WiStBuGh10,AcKo11};
-it is then close to a particular case of the discontinuous Galerkin
-technique \citep{ReHi73,LeRa74,Arn82,JoPi86,BoMaHe91,FaRi99,HuHuRa99,CoKaSh00,GiHeWa02,RiWh03,MoRi05,GrScSc06,AiMoMu06,BeLaPi06,DuKa06,DeSeWh08,PuAmKa09,WiStBuGh10,DeSe10,EtChViGl10},
-with optimized efficiency because of its tensorized basis functions
-\citep{WiStBuGh10,AcKo11}. In particular, it can accurately handle
-very distorted mesh elements \citep{OlSe11}.\\
-
-It has very good accuracy and convergence properties \citep{MaPa89,SePr94,DeFiMu02,Coh02,DeSe07,SeOl08,AiWa09,AiWa10,MeStTh12}.
-The spectral element approach admits spectral rates of convergence
-and allows exploiting $hp$-convergence schemes. It is also very well
-suited to parallel implementation on very large supercomputers \citep{KoTsChTr03,TsKoChTr03,KoLaMi08a,CaKoLaTiMiLeSnTr08,KoViCh10}
-as well as on clusters of GPU accelerating graphics cards \citep{Kom11,MiKo10,KoMiEr09,KoErGoMi10}.
-Tensor products inside each element can be optimized to reach very
-high efficiency \citep{DeFiMu02}, and mesh point and element numbering
-can be optimized to reduce processor cache misses and improve cache
-reuse \citep{KoLaMi08a}. The SEM can also handle triangular (in 2D)
-or tetrahedral (in 3D) elements \citep{WinBoyd96,TaWi00,KoMaTrTaWi01,Coh02,MeViSa06}
-as well as mixed meshes, although with increased cost and reduced
-accuracy in these elements, as in the discontinuous Galerkin method.\\
-
-Note that in many geological models in the context of seismic wave
-propagation studies (except for instance for fault dynamic rupture
-studies, in which very high frequencies or supershear rupture need
-to be modeled near the fault, see e.g. \citet{BeGlCrViPi07,BeGlCrVi09,PuAmKa09,TaCrEtViBeSa10})
-a continuous formulation is sufficient because material property contrasts
-are not drastic and thus conforming mesh doubling bricks can efficiently
-handle mesh size variations \citep{KoTr02a,KoLiTrSuStSh04,LeChLiKoHuTr08,LeChKoHuTr09,LeKoHuTr09}.\\
-
-For a detailed introduction to the SEM as applied to regional seismic
-wave propagation, please consult \citet{PeKoLuMaLeCaLeMaLiBlNiBaTr11,TrKoLi08,KoVi98,KoTr99,ChKoViCaVaFe07}
-and in particular \citet{LeKoHuTr09,LeChKoHuTr09,LeChLiKoHuTr08,GoAmTaCaSmSaMaKo09,WiKoScTr04,KoLiTrSuStSh04}.
-A detailed theoretical analysis of the dispersion
-and stability properties of the SEM is available in \citet{Coh02}, \citet{DeSe07}, \citet{SeOl07}, \citet{SeOl08} and \citet{MeStTh12}.\\
-
-Effects due to lateral variations in compressional-wave speed, shear-wave
-speed, density, a 3D crustal model, topography and bathymetry are
-included. The package can accommodate full 21-parameter anisotropy
-(see~\citet{ChTr07}) as well as lateral variations in attenuation
-\citep{SaKoTr10}. Adjoint capabilities and finite-frequency kernel
-simulations are included \citep{TrKoLi08,PeKoLuMaLeCaLeMaLiBlNiBaTr11,LiTr06,FiIgBuKe09,ViOp09}.\\
-
-The SEM was originally developed in computational fluid dynamics \citep{Pat84,MaPa89}
-and has been successfully adapted to address problems in seismic wave
-propagation. Early seismic wave propagation applications of the SEM,
-utilizing Legendre basis functions and a perfectly diagonal mass matrix,
-include \citet{CoJoTo93}, \citet{Kom97}, \citet{FaMaPaQu97}, \citet{CaGa97},
-\citet{KoVi98} and \citet{KoTr99}, whereas applications involving
-Chebyshev basis functions and a nondiagonal mass matrix include \citet{SePr94},
-\citet{PrCaSe94} and \citet{SePrPr95}.\\
-
-All SPECFEM3D software is written in Fortran2003 with full portability
-in mind, and conforms strictly to the Fortran2003 standard. It uses
-no obsolete or obsolescent features of Fortran. The package uses parallel
-programming based upon the Message Passing Interface (MPI) \citep{GrLuSk94,Pac97}.\\
-
-SPECFEM3D won the Gordon Bell award for best performance at the SuperComputing~2003
-conference in Phoenix, Arizona (USA) (see \citet{KoTsChTr03} and
-\url{www.sc-conference.org/sc2003/nrfinalaward.html}). It was a finalist
-again in 2008 for a run at 0.16 petaflops (sustained) on 149,784 processors
-of the `Jaguar' Cray XT5 system at Oak Ridge National Laboratories
-(USA) \citep{CaKoLaTiMiLeSnTr08}. It also won the BULL Joseph Fourier
-supercomputing award in 2010.\\
-
-It reached the sustained one petaflop performance level for the first time in February 2013
-on the Blue Waters Cray supercomputer at the National Center for Supercomputing Applications (NCSA), located at the University of Illinois at Urbana-Champaign (USA).\\
-
-This new release of the code includes Convolution or
-Auxiliary Differential Equation Perfectly Matched absorbing Layers
-(C-PML or ADE-PML) \citep{MaKoEz08,MaKoGe08,MaKo09,MaKoGeBr10,KoMa07}.
-It also includes support for GPU graphics card acceleration \citep{Kom11,MiKo10,KoMiEr09,KoErGoMi10}.\\
-
-The next release of the code will use the PT-SCOTCH parallel and threaded
-version of SCOTCH for mesh partitioning instead of the serial version.\\
-
-SPECFEM3D Cartesian includes coupled fluid-solid domains and adjoint
-capabilities, which enables one to address seismological inverse problems,
-but for linear rheologies only so far. To accommodate visco-plastic
-or non-linear rheologies, the reader can refer to the \href{http://geoelse.stru.polimi.it/}{GeoELSE}software
-package \citep{CaGa97,StPaIg09}.
-
-\section{Citation}
-
-If you use SPECFEM3D Cartesian for your own research, please cite
-at least one of the following articles:
-\begin{description}
-\item [{Numerical simulations in general}] ~\\
- Forward and adjoint simulations are described in detail in \citet{TrKoLi08,PeKoLuMaLeCaLeMaLiBlNiBaTr11,VaCaSaKoVi99,KoMiEr09,KoErGoMi10,ChKoViCaVaFe07,MaKoDi09,KoViCh10,CaKoLaTiMiLeSnTr08,TrKoHjLiZhPeBoMcFrTrHu10,KoRiTr02,KoTr02a,KoTr02b,KoTr99}
-or \citet{KoVi98}. Additional aspects of adjoint simulations are
-described in \citet{TrTaLi05,LiTr06,TrKoLi08,LiTr08,TrKoHjLiZhPeBoMcFrTrHu10,PeKoLuMaLeCaLeMaLiBlNiBaTr11}.
-Domain decomposition is explained in detail in \citet{MaKoBlLe08},
-and excellent scaling up to 150,000 processor cores is shown for instance
-in \citet{CaKoLaTiMiLeSnTr08,KoLaMi08a,MaKoBlLe08,KoErGoMi10,Kom11},
-\item [{GPU computing}] ~\\
- Computing on GPU graphics cards for acoustic or seismic wave propagation
-applications is described in detail in \citet{Kom11,MiKo10,KoMiEr09,KoErGoMi10}.
-\end{description}
-\noindent If you use this new version, which has non blocking
-MPI for much better performance for medium or large runs, please cite
-at least one of these six articles, in which results of non blocking
-MPI runs are presented: \citet{PeKoLuMaLeCaLeMaLiBlNiBaTr11,KoErGoMi10,KoViCh10,Kom11,CaKoLaTiMiLeSnTr08,MaKoBlLe08}.\\
-
-\noindent If you work on geophysical applications, you may be interested
-in citing some of these application articles as well, among others:
-\begin{description}
-\item [{Southern California simulations}] ~\\
- \citet{KoLiTrSuStSh04,KrJiKoTr06a,KrJiKoTr06b}.
-
-If you use the 3D southern California model, please cite \citet{SuSh03}
-(Los Angeles model), \citet{lovelyetal06} (Salton Trough), and \citet{hauksson2000}
-(southern California). The Moho map was determined by \citet{zhu&kanamori2000}.
-The 1D SoCal model was developed by \citet{DrHe90}.
-
-\item [{Anisotropy}] ~\\
- \citet{ChTr07,JiTsKoTr05,ChFaKo04,FaChKo04,RiRiKoTrHe02,TrKo00}.
-\item [{Attenuation}] ~\\
- \citet{SaKoTr10,KoTr02a,KoTr99}.
-\item [{Topography}] ~\\
- \citet{LeKoHuTr09,LeChKoHuTr09,LeChLiKoHuTr08,GoAmTaCaSmSaMaKo09,WiKoScTr04}.
-\end{description}
-The corresponding Bib\TeX{} entries may be found in file \texttt{doc/USER\_MANUAL/bibliography.bib}.
-
-\section{Support}
-
-This material is based upon work supported by the USA National Science
-Foundation under Grants No. EAR-0406751 and EAR-0711177, by the French
-CNRS, French INRIA Sud-Ouest MAGIQUE-3D, French ANR NUMASIS under
-Grant No. ANR-05-CIGC-002, and European FP6 Marie Curie International
-Reintegration Grant No. MIRG-CT-2005-017461. Any opinions, findings,
-and conclusions or recommendations expressed in this material are
-those of the authors and do not necessarily reflect the views of the
-USA National Science Foundation, CNRS, INRIA, ANR or the European
-Marie Curie program.
-
-\chapter{\label{cha:Getting-Started}Getting Started}
-
-To download the SPECFEM3D\_Cartesian software package, type this:
-\begin{lyxcode}
-git clone --recursive --branch devel https://github.com/geodynamics/specfem3d.git
-\end{lyxcode}
-{\small \par}
-
-We recommend that you add {\texttt{ulimit -S -s unlimited}} to your
-{\texttt{.bash\_profile}} file and/or {\texttt{limit stacksize
-unlimited }} to your {\texttt{.cshrc}} file to suppress any potential
-limit to the size of the Unix stack.\\
-
-Then, to configure the software for your system, run the \texttt{configure}
-shell script. This script will attempt to guess the appropriate configuration
-values for your system. However, at a minimum, it is recommended that
-you explicitly specify the appropriate command names for your Fortran
-compiler and MPI package (another option is to define FC, CC and MPIF90
-in your .bash\_profile or your .cshrc file):
-\begin{lyxcode}
-./configure~FC=ifort~MPIFC=mpif90
-\end{lyxcode}
-Note that MPI must be installed with MPI-IO enabled because parts of SPECFEM3D perform I/Os through MPI-IO.
-
-Before running the \texttt{configure} script, you should probably
-edit file \texttt{flags.guess} to make sure that it contains the best
-compiler options for your system. Known issues or things to check
-are:
-\begin{description}
-\item [{\texttt{Intel ifort compiler}}] See if you need to add \texttt{-assume byterecl} for your machine.
-\textbf{In the case of that compiler, we have noticed that versions dot zero sometimes have bugs or issues that can lead to wrong results when running the code, thus we \emph{strongly} recommend using versions dot one or above (for instance version 13.1 instead of 13.0, version 14.1 instead of 14.0 and so on)}.
-\item [{\texttt{IBM compiler}}] See if you need to add \texttt{-qsave}
-or \texttt{-qnosave} for your machine.
-\item [{\texttt{Mac OS}}] You will probably need to install \texttt{XCODE}.
-\end{description}
-When compiling on an IBM machine with the \texttt{xlf} and \texttt{xlc}
-compilers, we suggest running the \texttt{configure} script with the
-following options:\\
-
-\noindent \texttt{./configure FC=xlf90\_r MPIFC=mpif90 CC=xlc\_r CFLAGS=\textquotedbl{}-O3
--q64\textquotedbl{} FCFLAGS=\textquotedbl{}-O3 -q64\textquotedbl{}
--with-scotch-dir=...}.\\
-
-On SGI systems, \texttt{flags.guess} automatically informs \texttt{configure}
-to insert `\texttt{`TRAP\_FPE=OFF}'' into the generated \texttt{Makefile}
-in order to turn underflow trapping off.\\
-
-You can add \texttt{--enable-vectorization} to the configuration options to speed up the code in the fluid (acoustic) and elastic parts.
-This works fine if (and only if) your computer always allocates a contiguous memory block for each allocatable array;
-this is the case for most machines and most compilers, but not all. To disable this feature, use option \texttt{--disable-vectorization}.
-For more details see \href{https://github.com/geodynamics/specfem3d/issues/81}{github.com/geodynamics/specfem3d/issues/81} .
-To check if that option works fine on your machine, run the code with and without it for an acoustic/elastic model and make sure the seismograms are identical.\\
-
-Note that we use CUBIT (now called Trelis) to create meshes of hexahedra, but other packages
-can be used as well, for instance GiD from \url{http://gid.cimne.upc.es}
-or Gmsh from \url{http://geuz.org/gmsh} \citep{GeRe09}. Even mesh
-creation packages that generate tetrahedra, for instance TetGen from
-\url{http://tetgen.berlios.de}, can be used because each tetrahedron
-can then easily be decomposed into four hexahedra as shown in the
-picture of the TetGen logo at \url{http://tetgen.berlios.de/figs/Delaunay-Voronoi-3D.gif};
-while this approach does not generate hexahedra of optimal quality,
-it can ease mesh creation in some situations and it has been shown
-that the spectral-element method can very accurately handle distorted
-mesh elements \citep{OlSe11}.
-
-The SPECFEM3D Cartesian software package relies on the SCOTCH library
-to partition meshes created with CUBIT. METIS \citep{KaKu98a,KaKu98c,KaKu98b}
-can also be used instead of SCOTCH if you prefer, by editing file
-src/decompose\_mesh/decompose\_mesh.F90 and uncommenting the flag
-USE\_METIS\_INSTEAD\_OF\_SCOTCH. You will also then need to install
-and compile Metis version 4.0 (do {*}NOT{*} install Metis version
-5.0, which has incompatible function calls) and edit \texttt{src/decompose\_mesh/Makefile.in}
-and uncomment the METIS link flag in that file before running \texttt{configure}.\\
-
-The SCOTCH library \citep{PeRo96} provides efficent static mapping,
-graph and mesh partitioning routines. SCOTCH is a free software package
-developed by Fran\c{c}ois Pellegrini et al. from LaBRI and INRIA in Bordeaux,
-France, downloadable from the web page \url{https://gforge.inria.fr/projects/scotch/}.
-In case no SCOTCH libraries can be found on the system, the configuration
-will bundle the version provided with the source code for compilation.
-The path to an existing SCOTCH installation can to be set explicitly
-with the option \texttt{-{}-with-scotch-dir}. Just as an example:
-\begin{lyxcode}
-./configure~FC=ifort~MPIFC=mpif90~-{}-with-scotch-dir=/opt/scotch
-\end{lyxcode}
-If you use the Intel ifort compiler to compile the code, we recommend
-that you use the Intel icc C compiler to compile Scotch, i.e., use:
-\begin{lyxcode}
-./configure~CC=icc~FC=ifort~MPIFC=mpif90
-\end{lyxcode}
-
-When compiling the SCOTCH source code, if you get a message such as: "ld: cannot find -lz",
-the Zlib compression development library is probably missing on your machine and you will need to install it or ask your system administrator to
-do so. On Linux machines the package is often called "zlib1g-dev" or similar. (thus "sudo apt-get install zlib1g-dev" would install it)
-
-To compile a serial version of the code for small meshes that fits
-on one compute node and can therefore be run serially, run \texttt{configure}
-with the \texttt{-{}-without-mpi} option to suppress all calls to
-MPI.
-
-A summary of the most important configuration variables follows.
-\begin{description}
-\item [{\texttt{F90}}] Path to the Fortran compiler.
-\item [{\texttt{MPIF90}}] Path to MPI Fortran.
-\item [{\texttt{MPI\_FLAGS}}] Some systems require this flag to link to
-MPI libraries.
-\item [{\texttt{FLAGS\_CHECK}}] Compiler flags.
-\end{description}
-The configuration script automatically creates for each executable
-a corresponding \texttt{Makefile} in the \texttt{src/} subdirectoy.
-The \texttt{Makefile} contains a number of suggested entries for various
-compilers, e.g., Portland, Intel, Absoft, NAG, and Lahey. The software
-has run on a wide variety of compute platforms, e.g., various PC clusters
-and machines from Sun, SGI, IBM, Compaq, and NEC. Select the compiler
-you wish to use on your system and choose the related optimization
-flags. Note that the default flags in the \texttt{Makefile} are undoubtedly
-not optimal for your system, so we encourage you to experiment with
-these flags and to solicit advice from your systems administrator.
-Selecting the right compiler and optimization flags can make a tremendous
-difference in terms of performance. We welcome feedback on your experience
-with various compilers and flags.
-
-Now that you have set the compiler information, you need to select
-a number of flags in the \texttt{constants.h} file depending on your
-system:
-\begin{description}
-\item [{\texttt{LOCAL\_PATH\_IS\_ALSO\_GLOBAL}}] Set to \texttt{.false.}
-on most cluster applications. For reasons of speed, the (parallel)
-distributed database generator typically writes a (parallel) database
-for the solver on the local disks of the compute nodes. Some systems
-have no local disks, e.g., BlueGene or the Earth Simulator, and other
-systems have a fast parallel file system, in which case this flag
-should be set to \texttt{.true.}. Note that this flag is not used
-by the database generator or the solver; it is only used for some
-of the post-processing.
-\end{description}
-The package can run either in single or in double precision mode.
-The default is single precision because for almost all calculations
-performed using the spectral-element method using single precision
-is sufficient and gives the same results (i.e. the same seismograms);
-and the single precision code is faster and requires exactly half
-as much memory. Select your preference by selecting the appropriate
-setting in the \texttt{constants.h} file:
-\begin{description}
-\item [{\texttt{CUSTOM\_REAL}}] Set to \texttt{SIZE\_REAL} for single precision
-and \texttt{SIZE\_DOUBLE} for double precision.
-\end{description}
-In the \texttt{precision.h} file:
-\begin{description}
-\item [{\texttt{CUSTOM\_MPI\_TYPE}}] Set to \texttt{MPI\_REAL} for single
-precision and \texttt{MPI\_DOUBLE\_PRECISION} for double precision.
-\end{description}
-On many current processors (e.g., Intel, AMD, IBM Power), single precision
-calculations are significantly faster; the difference can typically
-be 10\% to 25\%. It is therefore better to use single precision. What
-you can do once for the physical problem you want to study is run
-the same calculation in single precision and in double precision on
-your system and compare the seismograms. If they are identical (and
-in most cases they will), you can select single precision for your
-future runs.\\
-
-If your compiler has problems with the \texttt{use mpi} statements that are used in the code, use the script called
-\texttt{replace\_use\_mpi\_with\_include\_mpif\_dot\_h.pl} in the root directory to replace all of them with \texttt{include 'mpif.h'} automatically.
-
-\section{Adding OpenMP support in addition to MPI}
-
-NOTE FROM JULY 2013: OpenMP support is maybe / probably not maintained any more.
-Thus the section below is maybe obsolete.
-
-OpenMP support can be enabled in addition to MPI. However, in many
-cases performance will not improve because our pure MPI implementation
-is already heavily optimized and thus the resulting code will in fact
-be slightly slower. A possible exception could be IBM BlueGene-type
-architectures.
-
-To enable OpenMP, uncomment the OpenMP compiler option in two lines
-in file \texttt{src/specfem3D/Makefile.in} (before running \texttt{configure})
-and also uncomment the \texttt{\#define USE\_OPENMP} statement in
-file\\
- \texttt{src/specfem3D/specfem3D.F90}.
-
-The DO-loop using OpenMP threads has a SCHEDULE property. The OMP\_SCHEDULE
-environment variable can set the scheduling policy of that DO-loop.
-Tests performed by Marcin Zielinski at SARA (The Netherlands) showed
-that often the best scheduling policy is DYNAMIC with the size of
-the chunk equal to the number of OpenMP threads, but most preferably
-being twice as the number of OpenMP threads (thus chunk size = 8 for
-4 OpenMP threads etc). If OMP\_SCHEDULE is not set or is empty, the
-DO-loop will assume generic scheduling policy, which will slow down
-the job quite a bit.
-
-\section{Compiling on an IBM BlueGene}
-
-\underline{More recent installation instruction for IBM BlueGene, from April 2013:}\\
-
-\noindent Edit file \texttt{flags.guess} and put this for \texttt{FLAGS\_CHECK}:\\
-
-\noindent \texttt{-g -qfullpath -O2 -qsave -qstrict -qtune=qp -qarch=qp
--qcache=auto -qhalt=w}\\
- \texttt{-qfree=f90 -qsuffix=f=f90 -qlanglvl=95pure -Q -Q+rank,swap\_all
--Wl,-relax}\\
-
-\noindent The most relevant are the -qarch and -qtune flags, otherwise
-if these flags are set to ``auto'' then they are wrongly assigned
-to the architecture of the frond-end node, which is different from
-that on the compute nodes. You will need to set these flags to the
-right architecture for your BlueGene compute nodes, which is not necessarily
-``qp''; ask your system administrator. On some machines if is necessary
-to use -O2 in these flags instead of -O3 due to a compiler bug of
-the XLF version installed. We thus suggest to first try -O3, and then
-if the code does not compile or does not run fine then switch back
-to -O2. The debug flags (-g, -qfullpath) do not influence performance
-but are useful to get at least some insights in case of problems.\\
-
-\noindent Before running \texttt{configure}, select the XL Fortran
-compiler by typing \texttt{module load bgq-xl/1.0} or \texttt{module
-load bgq-xl} (another, less efficient option is to load the GNU compilers
-using \texttt{module load bgq-gnu/4.4.6} or similar).\\
-
-\noindent Then, to configure the code, type this:\\
-
-\noindent \texttt{./configure FC=bgxlf90\_r MPIFC=mpixlf90\_r CC=bgxlc\_r
-LOCAL\_PATH\_IS\_ALSO\_GLOBAL=true}\\
-
-In order for the SCOTCH domain decomposer to compile, on some (but
-not all) Blue Gene systems you may need to run \texttt{configure}
-with \texttt{CC=gcc} instead of \texttt{CC=bgxlc\_r}.
-
-\noindent \underline{Older installation instruction for IBM BlueGene, from 2011:}\\
-
-\noindent To compile the code on an IBM BlueGene, Laurent L\'eger from
-IDRIS, France, suggests the following: compile the code with
-
-\noindent \texttt{ FLAGS\_CHECK=\textquotedbl{}-O3 -qsave -qstrict
--qtune=auto -qarch=450d -qcache=auto }~\\
- \texttt{ -qfree=f90 -qsuffix=f=f90 -g -qlanglvl=95pure -qhalt=w -Q
--Q+rank,swap\_all -Wl,-relax\textquotedbl{}}
-
-\noindent Option \textquotedbl{}-Wl,-relax\textquotedbl{} must be
-added on many (but not all) BlueGene systems to be able to link the
-binaries \texttt{xmeshfem3D} and \texttt{xspecfem3D} because the final
-link step is done by the GNU \texttt{ld} linker even if one uses \texttt{FC=bgxlf90\_r,
-MPIFC=mpixlf90\_r} and \texttt{CC=bgxlc\_r} to create all the object
-files. On the contrary, on some BlueGene systems that use the native
-AIX linker option \textquotedbl{}-Wl,-relax\textquotedbl{} can lead
-to problems and must be suppressed from \texttt{flags.guess}.
-
-\noindent Also, \texttt{AR=ar, ARFLAGS=cru} and \texttt{RANLIB=ranlib}
-are hardwired in all \texttt{Makefile.in} files by default, but to
-cross-compile on BlueGene/P one needs to change these values to \texttt{AR=bgar,
-ARFLAGS=cru} and \texttt{RANLIB=bgranlib}. Thus the easiest thing
-to do is to modify all \texttt{Makefile.in} files and the \texttt{configure}
-script to set them automatically by \texttt{configure}. One then just
-needs to pass the right commands to the \texttt{configure} script:
-
-\noindent \texttt{ ./configure -{}-prefix=/path/to/SPECFEM3DG\_SP
--{}-host=Babel -{}-build=BGP }~\\
- \texttt{ FC=bgxlf90\_r MPIFC=mpixlf90\_r CC=bgxlc\_r AR=bgar ARFLAGS=cru
-}~\\
- \texttt{ RANLIB=bgranlib LOCAL\_PATH\_IS\_ALSO\_GLOBAL=false}
-
-\noindent This trick can be useful for all hosts on which one needs
-to cross-compile.
-
-\noindent On BlueGene, one also needs to run the \texttt{xcreate\_header\_file}
-binary file manually rather than in the Makefile:
-
-\noindent \texttt{bgrun -np 1 -mode VN -exe ./bin/xcreate\_header\_file}
-
-\section{Visualizing the subroutine calling tree of the source code}
-
-\noindent
-Packages such as \texttt{Doxywizard} can be used to visualize the calling tree of the
-subroutines of the  source code. \texttt{Doxywizard} is
-a GUI front-end for configuring and running \texttt{Doxygen}.
-
-\bigskip
-
-\noindent To do your own call graphs, you can follow these simple steps below.
-
-\begin{enumerate}
-\setcounter{enumi}{-1}
-\item Install \texttt{Doxygen} \underline{and} \texttt{graphviz} (the two are usually in the package manager of classic Linux distribution).
-\item Run in the terminal : \texttt{doxygen -g}, which creates a \texttt{Doxyfile} that tells doxygen what you want it to do.
-\item Edit the Doxyfile. Two Doxyfile-type files have been already committed in the directory \\
-\noindent \texttt{specfem3d/doc/Call\_trees}:
-
-\begin{itemize}
-\item[\textbullet] \texttt{Doxyfile\_truncated\_call\_tree} will generate call graphs with maximum 3 or 4 levels of tree structure,
-\item[\textbullet] \texttt{Doxyfile\_complete\_call\_tree} will generate call graphs with complete tree structure.
-\end{itemize}
-
-The important entries in the Doxyfile are:
-
-\begin{description}
-\item [{\texttt{PROJECT\_NAME}}]
-\item [{\texttt{OPTIMIZE\_FOR\_FORTRAN}}] Set to YES
-\item [{\texttt{EXTRACT\_ALL}}] Set to YES
-\item [{\texttt{EXTRACT\_PRIVATE}}] Set to YES
-\item [{\texttt{EXTRACT\_STATIC}}] Set to YES
-\item [{\texttt{INPUT}}] From the directory \texttt{specfem3d/doc/Call\_trees}, it is \texttt{"../../src/"}
-\item [{\texttt{FILE\_PATTERNS}}] In SPECFEM case, it is \texttt{*.f90* *.F90* *.c* *.cu* *.h*}
-\item [{\texttt{HAVE\_DOT}}] Set to YES
-\item [{\texttt{CALL\_GRAPH}}] Set to YES
-\item [{\texttt{CALLER\_GRAPH}}] Set to YES
-\item [{\texttt{DOT\_PATH}}] The path where is located the dot program graphviz (if it is not in your \$PATH)
-\item [{\texttt{RECURSIVE}}] This tag can be used to turn specify whether or not subdirectories should be searched for input files as well. In the case of SPECFEM, set to YES.
-\item [{\texttt{EXCLUDE}}] Here, you can exclude:  \\
-\texttt{"../../src/specfem3D/older\_not\_maintained\_partial\_OpenMP\_port"}\\
-\texttt{"../../src/decompose\_mesh/scotch"}\\
-\texttt{"../../src/decompose\_mesh/scotch\_5.1.12b"}
-\item [{\texttt{DOT\_GRAPH\_MAX\_NODES}}] to set the maximum number of nodes that will be shown in the graph.
-If the number of nodes in a graph becomes larger than this value, doxygen will truncate the graph, which is visualized by representing a node as a red box.
-Minimum value: 0, maximum value: 10000, default value: 50.
-\item [{\texttt{MAX\_DOT\_GRAPH\_DEPTH}}] to set the maximum depth of the graphs generated by dot.
-A depth value of 3 means that only nodes reachable from the root by following a path via at most 3 edges will be shown.
-Using a depth of 0 means no depth restriction.
-Minimum value: 0, maximum value: 1000, default value: 0.
-
-\end{description}
-
-\item Run : \texttt{doxygen Doxyfile}, HTML and LaTeX files created by default in \texttt{html} and \texttt{latex} subdirectories.
-
-\item To see the call trees, you have to open the file \texttt{html/index.html} in your \underline{browser}.
-You will have many informations about each subroutines of SPECFEM (not only call graphs), you can click on every boxes / subroutines.
-It show you the call, and, the caller graph of each subroutine : the subroutines called by the concerned subroutine, and the previous subroutines who call this subroutine (the previous path), respectively.
-In the case of a truncated calling tree, the boxes with a red border indicates a node that has \underline{more} arrows than are shown (in other words: the graph is truncated with respect to this node).
-
-\end{enumerate}
-
-\medskip
-
-\noindent Finally, some useful links:
-
-\begin{itemize}
-\item[\textbullet] a good and short summary for the basic utilisation of Doxygen:
-
-\url{http://www.softeng.rl.ac.uk/blog/2010/jan/30/callgraph-fortran-doxygen/},
-
-\item[\textbullet] to configure the diagrams :
-
-\url{http://www.stack.nl/~dimitri/doxygen/manual/diagrams.html},
-
-\item[\textbullet] the complete alphabetical index of the tags in Doxyfile:
-
-\url{http://www.stack.nl/~dimitri/doxygen/manual/config.html},
-
-\item[\textbullet] more generally, the Doxygen manual:
-
-\url{http://www.stack.nl/~dimitri/doxygen/manual/index.html}.
-
-\end{itemize}
-
-\medskip
-
-\section{Using the ADIOS library for I/O}
-Regular POSIX I/O can be problematic when dealing with large simulations one large
-clusters (typically more than $10,000$ processes). SPECFEM3D use the ADIOS library~\cite{Liu2013}
-to deal transparently take advantage of advanced paralllel file system features. To enable
-ADIOS, the following steps should be done:
-\begin{enumerate}
-\item Install ADIOS (available from \url{https://www.olcf.ornl.gov/center-projects/adios/}).
-Make sure that your environment variables reference it.
-\item You may want to change ADIOS related values in the \texttt{constants.h.in} file.
-The default values probably suit most cases.
-\item Configure using the \texttt{--with-adios} flag.
-\end{enumerate}
-ADIOS is currently only usable for meshfem3D generated mesh (i.e. not for meshes generated
-with CUBTI). Additional control parameters are discussed in section~\ref{cha:Main-Parameter}.
-
-\section{Becoming a developer of the code, or making small modifications in the source code}
-
-If you want to develop new features in the code, and/or if you want to make small changes, improvements, or bug fixes, you are very welcome to contribute.
-To do so, i.e.
-to access the development branch of the source code with read/write access (in a safe way, no need to worry too much about breaking the package, there is a robot called BuildBot that is in charge of checking and validating all new contributions and changes), please visit this Web page: \url{https://github.com/geodynamics/specfem3d/wiki/Using-Hub}.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\chapter{\label{cha:Mesh-Generation}Mesh Generation}
-
-The first step in running a spectral-element simulation consists of
-constructing a high-quality mesh for the region under consideration.
-We provide two possibilities to do so: (1) relying on the external,
-hexahedral mesher CUBIT, or (2) using the provided, internal mesher
-\texttt{xmeshfem3D}. In the following, we explain these two approaches.
-
-\section{\label{cha:Running-the-Mesher-CUBIT}Meshing with \texttt{CUBIT}}
-
-CUBIT is a meshing tool suite for the creation of finite-element meshes
-for arbitrarily shaped models. It has been developed and maintained
-at Sandia National Laboratories and can be purchased for a small academic
-institutional fee at \url{http://cubit.sandia.gov}. Our experience
-showed that using CUBIT greatly facilitates and speeds up the generation
-and preparation of hexahedral, conforming meshes for a variety of
-geophysical models with increasing complexity.
-
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=0.6\textwidth]{figures/mount-cubit}
-\par\end{centering}
-
-\caption{Example of the graphical user interface of CUBIT. The hexahedral mesh
-shown in the main display consists of a hexahedral discretization
-of a single volume with topography.}
-
-\label{fig:mount.cubit}
-\end{figure}
-
-\noindent
-The basic steps in creating a load-balanced, partitioned mesh with
-CUBIT are:
-\begin{description}
-\item [{1.}] setting up a hexahedral mesh with CUBIT,
-\item [{2.}] exporting the CUBIT mesh into a SPECFEM3D Cartesian file format
-and
-\item [{3.}] partitioning the SPECFEM3D Cartesian mesh files for a chosen
-number of cores.
-\end{description}
-Examples are provided in the SPECFEM3D Cartesian package in the subdirectory
-\texttt{EXAMPLES/}. We strongly encourage you to contribute your own
-example to this package by contacting the CIG Computational Seismology
-Mailing List \urlwithparentheses{cig-seismo@geodynamics.org}.
-
-\subsection{Creating the Mesh with CUBIT}
-
-For the installation and handling of the CUBIT meshing tool suite,
-please refer to the CUBIT user manual and documentation. In order
-to give you a basic understanding of how to use CUBIT for our purposes,
-examples are provided in the SPECFEM3D Cartesian package in the subdirectory
-\texttt{EXAMPLES/}:
-\begin{description}
-\item [{\texttt{homogeneous\_halfspace}}] Creates a single block model
-and assigns elastic material parameters.
-\item [{\texttt{layered\_halfspace}}] Combines two different, elastic material
-volumes and creates a refinement layer between the two. This example
-can be compared for validation against the solutions provided in subdirectory
-~\\
- \texttt{VALIDATION\_3D\_SEM\_SIMPLER\_LAYER\_SOURCE\_DEPTH/}.
-\item [{\texttt{waterlayered\_halfspace}}] Combines an acoustic and elastic
-material volume as in a schematic marine survey example.
-\item [{\texttt{tomographic\_model}}] Creates a single block model whose
-material properties will have to be read in from a tomographic model
-file during the databases creation by \texttt{xgenerate\_databases}.
-\end{description}
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=0.45\textwidth]{figures/example-homogeneous}
-\includegraphics[width=0.45\textwidth]{figures/example-2layers} \\
- \includegraphics[width=0.45\textwidth]{figures/example-water} \includegraphics[width=0.45\textwidth]{figures/example-tomo}
-\par\end{centering}
-
-\caption{Screenshots of the CUBIT examples provided in subdirectory \texttt{EXAMPLES/}:
-homogeneous halfspace (top-left), layered halfspace (top-right), water
-layered halfspace (bottom-left) and tomographic model (bottom-right).}
-
-\label{fig:examples.cubit}
-\end{figure}
-
-In each example subdirectory you will find a \texttt{README} file,
-which explains in a step-by-step tutorial the workflow for the example.
-Please feel free to contribute your own example to this package by
-contacting the CIG Computational Seismology Mailing List \urlwithparentheses{cig-seismo@geodynamics.org}.
-
-In some cases, to re-create the meshes for the examples given, just type "claro ./create\_mesh.py" or similar
-from the command line ("claro" is the command to run CUBIT from the command line).
-
-IMPORTANT note: in order to correctly set up GEOCUBIT and run the examples, please read the file called "EXAMPLES/README";
-in particular, please make sure you correctly set up the Python paths as indicated in that file.
-
-\subsection{Exporting the Mesh with \texttt{run\_cubit2specfem3d.py} }
-
-\label{subsec:Exporting-the-Mesh}
-
-Once the geometric model volumes in CUBIT are meshed, you prepare
-the model for exportation with the definition of material blocks and
-boundary surfaces. Thus, prior to exporting the mesh, you need to
-define blocks specifying the materials and absorbing boundaries in
-CUBIT. This process could be done automatically using the script \texttt{run\_cubit2specfem3d.py}
-if the mesh meets some conditions or manually, following the block
-convention:
-\begin{description}
-\item [{material\_name}] Each material should have a specific block defined
-by a unique name. The name convention of the material is to start
-with either 'elastic' or 'acoustic'. It must be then followed by a
-unique identifier, e.g. 'elastic 1', 'elastic 2', etc. The additional
-attributes to the block define the material description.
-
-For an elastic material:
-\begin{description}
-\item [{material\_id}] An integer value which is unique for this material.
-\item [{Vp}] P-wave speed of the material (given in m/s).
-\item [{Vs}] S-wave speed of the material (given in m/s).
-\item [{rho}] density of the material (given in kg/m$^{3}$).
-\item [{Q}] quality factor to use in case of a simulation with attenuation
-turned on. It should be between 1 and 9000. In case no attenuation
-information is available, it can be set to zero. Please note that
-your Vp- and Vs-speeds are given for a reference frequency. To change
-this reference frequency, you change the value of \texttt{ATTENUATION\_f0\_REFERENCE}
-in the main constants file \texttt{constants.h} found in subdirectory
-\texttt{src/shared/}.
-\item [{anisotropic\_flag}] Flag describing the anisotropic model to use
-in case an anisotropic simulation should be conducted. See the file
-\texttt{model\_aniso.f90} in subdirectory \texttt{src/generate\_databases/}
-for an implementation of the anisotropic models. In case no anisotropy
-is available, it can be set to zero.
-\end{description}
-
-Note that this material block has to be defined using all the volumes
-which belong to this elastic material. For volumes belonging to another,
-different material, you will need to define a new material block.
-
-For an acoustic material:
-\begin{description}
-\item [{material\_id}] An integer value which is unique for this material.
-\item [{Vp}] P-wave speed of the material (given in m/s).
-\item [{0}] S-wave speed of the material is ignored.
-\item [{rho}] density of the material (given in kg/m$^{3}$).
-\end{description}
-\item [{face\_topo}] Block definition for the surface which defines the
-free surface (which can have topography). The name of this block must
-be 'face\_topo', the block has to be defined using all the surfaces
-which constitute the complete free surface of the model.
-\item [{face\_abs\_xmin}] Block definition for the faces on the absorbing
-boundaries, one block for each surface with x=Xmin.
-\item [{face\_abs\_xmax}] Block definition for the faces on the absorbing
-boundaries, one block for each surface with x=Xmax.
-\item [{face\_abs\_ymin}] Block definition for the faces on the absorbing
-boundaries, one block for each surface with y=Ymin.
-\item [{face\_abs\_ymax}] Block definition for the faces on the absorbing
-boundaries, one block for each surface with y=Ymax.
-\item [{face\_abs\_bottom}] Block definition for the faces on the absorbing
-boundaries, one block for each surface with z=bottom.
-\end{description}
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=0.45\textwidth]{figures/mount-surface} \includegraphics[width=0.45\textwidth]{figures/mount-abs}
-\par\end{centering}
-
-\caption{Example of the block definitions for the free surface 'face\_topo'
-(left) and the absorbing boundaries, defined in a single block 'face\_abs\_xmin'
-(right) in CUBIT.}
-
-\label{fig:mount.abs}
-\end{figure}
-
-Optionally, instead of specifying for each surface at the boundaries
-a single block like mentioned above, you can also specify a single
-block for all boundary surfaces and name it as one of the absorbing
-blocks above, e.g. 'face\_abs\_xmin'.
-
-After the block definitions are done, you export the mesh using the
-script \texttt{cubit2specfem3d.py} provided in each of the example
-directories (linked to the common script \texttt{CUBIT\_GEOCUBIT/cubit2specfem3d.py}).
-If the export was successful, you should find the following
-files in a subdirectory \texttt{MESH/}:
-\begin{description}
-\item [{absorbing\_cpml\_file}] \textbf{(only needed in case of C-PML absorbing
-conditions)} Contains on the first line the total number of C-PML
-spectral elements in the mesh, and then on the following line the
-list of all these C-PML elements with two numbers per line: first
-the spectral element number, and then a C-PML flag indicating to which
-C-PML layer(s) that element belongs, according to the following convention:
-
-\begin{itemize}
-\item Flag = 1 : element belongs to a X CPML layer only (either in Xmin
-or in Xmax),
-\item Flag = 2 : element belongs to a Y CPML layer only (either in Ymin
-or in Ymax),
-\item Flag = 3 : element belongs to a Z CPML layer only (either in Zmin
-or in Zmax),
-\item Flag = 4 : element belongs to a X CPML layer and also to a Y CPML
-layer,
-\item Flag = 5 : element belongs to a X CPML layer and also to a Z CPML
-layer,
-\item Flag = 6 : element belongs to a Y CPML layer and also to a Z CPML
-layer,
-\item Flag = 7 : element belongs to a X, to a Y and to a Z CPML layer, i.e.,
-it belongs to a CPML corner.
-\end{itemize}
-
-Note that it does not matter whether an element belongs to a Xmin
-or to a Xmax CPML, the flag is the same in both cases; the same is
-true for Ymin and Ymax, and also Zmin and Zmax.
-
-When you have an existing CUBIT (or similar) mesh stored in SPECFEM3D
-format, i.e., if you have existing \textquotedbl{}nodes\_coords\_file\textquotedbl{}
-and \textquotedbl{}mesh\_file\textquotedbl{} files but do not know
-how to assign CPML flags to them, we have created a small serial Fortran
-program that will do that automatically for you, i.e., which will
-create the \textquotedbl{}absorbing\_cpml\_file\textquotedbl{} for
-you. That program is utils/CPML/convert\_external\_layers\_of\_a\_given\_mesh\_to\_CPML\_layers.f90,
-and a small Makefile is provided in that directory (utils/CPML).
-
-IMPORTANT: it is your responsibility to make sure that in the input
-CUBIT (or similar) mesh that this code will read in SPECFEM3D format
-from files \textquotedbl{}nodes\_coords\_file\textquotedbl{} and \textquotedbl{}mesh\_file\textquotedbl{}
-you have created layers of elements that constitute a layer of constant
-thickness aligned with the coordinate grid axes (X, Y and/or Z), so
-that this code can assign CPML flags to them. This code does NOT check
-that (because it cannot, in any easy way). The mesh inside these CPML
-layers does not need to be structured nor regular, any non-structured
-mesh is fine as long as it has flat PML inner and outer faces, parallel
-to the axes, and thus of a constant thickness. The thickness can be
-different for the X, Y and Z sides. But for X it must not vary, for
-Y it must not vary, and for Z it must not vary. If you do not know
-the exact thickness, you can use a slightly LARGER value in this code
-(say 2\% to 5\% more) and this code will fix that and will adjust
-it; never use a SMALLER value otherwise this code will miss some CPML
-elements.
-
-Note: in a future release we will remove the constraint of having
-CPML layers aligned with the coordinate axes; we will allow for meshes
-that are titled by any constant angle in the horizontal plane. However
-this is not implemented yet.
-
-Note: in the case of fluid-solid layers in contact, the fluid-solid
-interface currently needs to be flat and horizontal inside the CPML
-layer (i.e., bathymetry should become flat and horizontal when entering
-the CPML); this (small) constraint will probably remain in the code
-for a while because it makes fluid-solid matching inside the CPML
-much easier.
-
-\item [{materials\_file}] Contains the material associations for each element.
-The format is:
-
-\begin{lyxcode}
-element\_ID~material\_ID
-\end{lyxcode}
-
-where \texttt{element\_ID} is the element identifier and \texttt{material\_ID}
-is a unique identifier, positive (for materials taken from this list
-of materials, i.e. for which each spectral element has constant material
-properties taken from this list) or negative (for tomographic models,
-i.e. for spectral element whose real velocities and density will be
-assigned later by calling an external function to define model variations,
-for instance in the case of tomographic models; in such a case, material
-properties can vary inside each spectral element, i.e. be different
-at each of its Gauss-Lobatto-Legendre grid points).
-
-\item [{nummaterial\_velocity\_file}] Defines the material properties.
-
-\begin{itemize}
-\item For classical materials (i.e., spectral elements for which the velocity
-and density model will not be assigned by calling an external function
-to define for instance a tomographic model), the format is:
-
-\begin{lyxcode}
-domain\_ID~material\_ID~rho~vp~vs~Qkappa~Qmu~anisotropy\_flag
-\end{lyxcode}
-
-where \texttt{\textbf{domain\_ID}}\textbf{ is 1 for acoustic and 2
-for elastic or viscoelastic materials,} \texttt{material\_ID} a unique
-identifier, \texttt{rho} the density in $kg\, m^{-3}$, \texttt{vp}
-the P-wave speed in $m\, s^{-1}$, \texttt{vs} the S-wave speed in
-$m\, s^{-1}$, \texttt{Q} the quality factor and \texttt{anisotropy\_flag}
-an identifier for anisotropic models. Note that the \texttt{Qkappa}
-value is ignored by the code unless \texttt{FULL\_ATTENUATION\_SOLID}
-is set. Note also that both \texttt{Qkappa} and \texttt{Qmu} are ignored
-by the code unless \texttt{ATTENUATION} is set. If you want a model
-with no \texttt{Qmu} attenuation, both set \texttt{ATTENUATION} to
-\texttt{.false.} in the \texttt{Par\_file} and set \texttt{Qmu} to
-9999 here. If you want a model with no \texttt{Qkappa} attenuation,
-both set \texttt{FULL\_ATTENUATION\_SOLID} to \texttt{.false.} in
-the \texttt{Par\_file} and set \texttt{Qkappa} to 9999 here. %%magnoni
-
-\item For tomographic velocity models, please read Chapter \ref{cha:-Changing-the}
-and Section \ref{sec:Using-tomographic} `Using external tomographic
-Earth models' for further details.
-\end{itemize}
-\item [{nodes\_coords\_file}] Contains the point locations in Cartesian
-coordinates of the mesh element corners.
-\item [{mesh\_file}] Contains the mesh element connectivity. The hexahedral
-elements can have 8 or 27 nodes.\\
- See picture doc/mesh\_numbering\_convention/numbering\_convention\_27\_nodes.jpg
-to see\\
- in which (standard) order the points must be cited. In the case of
-8 nodes, just include the first 8 points.
-\item [{free\_or\_absorbing\_surface\_file\_zmax}] Contains the free surface
-connectivity or \\
- the surface connectivity of the absorbing boundary surface at the
-top (Zmax), \\
- depending on whether the top surface is defined as free or absorbing
-(\texttt{STACEY\_INSTEAD\_OF\_FREE\_SURFACE} in \texttt{DATA/Par\_file}). \\
- You should put both the surface of acoustic regions and of elastic
-regions in that file; that is, list all the element faces that constitute
-the surface of the model in that file.
-\item [{absorbing\_surface\_file\_xmax}] Contains the surface connectivity
-of the absorbing boundary surface at Xmax \\
-(also needed in the case of C-PML absorbing conditions, in order for
-the code to be able to impose Dirichlet conditions on their outer
-edge).
-\item [{absorbing\_surface\_file\_xmin}] Contains the surface connectivity
-of the absorbing boundary surface at Xmin \\
-(also needed in the case of C-PML absorbing conditions, in order for
-the code to be able to impose Dirichlet conditions on their outer
-edge).
-\item [{absorbing\_surface\_file\_ymax}] Contains the surface connectivity
-of the absorbing boundary surface at Ymax \\
-(also needed in the case of C-PML absorbing conditions, in order for
-the code to be able to impose Dirichlet conditions on their outer
-edge).
-\item [{absorbing\_surface\_file\_ymin}] Contains the surface connectivity
-of the absorbing boundary surface at Ymin \\
-(also needed in the case of C-PML absorbing conditions, in order for
-the code to be able to impose Dirichlet conditions on their outer
-edge).
-\item [{absorbing\_surface\_file\_bottom}] Contains the surface connectivity
-of the absorbing boundary surface at the bottom (Zmin) \\
-(also needed in the case of C-PML absorbing conditions, in order for
-the code to be able to impose Dirichlet conditions on their outer
-edge).
-\end{description}
-These mesh files are needed as input files for the partitioner \texttt{xdecompose\_mesh}
-to load-balance the mesh. Please see the next section for further
-details.
-
-In directory "CUBIT\_GEOCUBIT" we provide a script that can help doing the above tasks of exporting a CUBIT mesh
-to SPECFEM3D format automatically for you: "run\_cubit2specfem3d.py".
-Just edit them to indicate the path to your local installation of CUBIT and also the name of the *.cub existing CUBIT mesh file
-that you want to export to SPECFEM3D format. These scripts will do the conversion for you automatically except assigning material properties
-to the different mesh layers. To do so, you will then need to edit the file
-called "nummaterial\_velocity\_file" that will have just been created and change it from the prototype created:
-
-    0 1 vol1 --> syntax: \#material\_domain\_id \#material\_id \#rho \#vp \#vs \#Q\_kappa \#Q\_mu \#anisotropy
-    0 2 vol2 --> syntax: \#material\_domain\_id \#material\_id \#rho \#vp \#vs \#Q\_kappa \#Q\_mu \#anisotropy
-
-(where "vol1" and "vol2" here represent the volume labels that you have set while creating the mesh in CUBIT)
-to for instance
-
-    2 1 1500 2300 1800 9999.0 9999.0 0
-    2 2 1600 2500 20000 9999.0 9999.0 0
-
-\subsubsection*{Checking the mesh quality}
-
-The quality of the mesh may be inspected more precisely based upon
-the serial code in the file \texttt{check\_mesh\_quality\_}~\\
- \texttt{CUBIT\_Abaqus.f90} located in the directory \texttt{src/check\_mesh\_quality\_CUBIT\_Abaqus/}.
-Running this code is optional because no information needed by the
-solver is generated.
-
-Prior to running and compiling this code, you have to export your
-mesh in CUBIT to an ABAQUS (.inp) format. For example, export mesh
-block IDs belonging to volumes in order to check the quality of the
-hexahedral elements. You also have to determine a number of parameters
-of your mesh, such as the number of nodes and number of elements and
-modify the header of the \texttt{check\_mesh\_quality\_CUBIT\_Abaqus.f90}
-source file in directory \texttt{src/check\_mesh\_quality\_CUBIT\_Abaqus/}.
-
-\noindent
-Then, in the main directory, type
-\begin{lyxcode}
-{\small make~xcheck\_mesh\_quality\_CUBIT\_Abaqus}{\small \par}
-\end{lyxcode}
-and use
-\begin{lyxcode}
-{\small ./bin/xcheck\_mesh\_quality\_CUBIT\_Abaqus}{\small \par}
-\end{lyxcode}
-to generate an %not supported: AVS output file (\texttt{\small AVS\_meshquality.inp} in AVS UCD format) or
-OpenDX output file (\texttt{\small DX\_mesh\_quality.dx}{\small )
-that can be used to investigate mesh quality, e.g. skewness of elements,
-and a Gnuplot histogram (}\texttt{\small mesh\_quality\_histogram.txt}{\small )
-that can be plotted with gnuplot (type `}\texttt{\small gnuplot plot\_mesh\_quality\_histogram.gnu}{\small ').
-The histogram is also printed to the screen. Analyze that skewness
-histogram of mesh elements to make sure no element has a skewness
-above approximately 0.75, otherwise the mesh is of poor quality (and
-if even a single element has a skewness value above 0.80, then you
-must definitely improve the mesh). If you want to start designing
-your own meshes, this tool is useful for viewing your creations. You
-are striving for meshes with elements with `cube-like' dimensions,
-e.g., the mesh should contain no very elongated or skewed elements.}{\small \par}
-
-\subsection{Partitioning the Mesh with \texttt{xdecompose\_mesh}}
-
-The SPECFEM3D Cartesian software package performs large scale simulations
-in a parallel 'Single Process Multiple Data' way. The spectral-element
-mesh created with CUBIT needs to be distributed on the processors.
-This partitioning is executed once and for all prior to the execution
-of the solver so it is referred to as a static mapping.
-
-An efficient partitioning is important because it leverages the overall
-running time of the application. It amounts to balance the number
-of elements in each slice while minimizing the communication costs
-resulting from the placement of adjacent elements on different processors.
-\texttt{decompose\_mesh} depends on the SCOTCH library \citep{PeRo96},
-which provides efficent static mapping, graph and mesh partitioning
-routines. SCOTCH is a free software package developed by Fran\c{c}ois
-Pellegrini et al. from LaBRI and INRIA in Bordeaux, France, downloadable
-from the web page \url{https://gforge.inria.fr/projects/scotch/}.
-\\
-
-In most cases, the configuration with \texttt{./configure~FC=ifort}
-should be sufficient. During the configuration process, the script
-tries to find existing SCOTCH installations. In case your system has
-no pre-existing SCOTCH installation, we provide the source code of
-SCOTCH, which is released open source under the French CeCILL-C version
-1 license, in directory \texttt{src/decompose\_mesh/scotch\_5.1.12b}.
-This version gets bundled with the compilation of the SPECFEM3D Cartesian
-package if no libraries could have been found. If this automatic compilation
-of the SCOTCH libraries fails, please refer to file INSTALL.txt in
-that directory to see further details how to compile it on your system.
-In case you want to use a pre-existing installation, make sure you
-have correctly specified the path of the SCOTCH library when using
-the option \texttt{-{}-with-scotch-dir} with the \texttt{./configure}
-script. In the future you should be able to find more recent versions
-at \url{http://www.labri.fr/perso/pelegrin/scotch/scotch_en.html}
-. \\
-
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=0.45\textwidth]{figures/mount-partitions}
-\includegraphics[width=0.45\textwidth]{figures/mount-partitions2}
-\par\end{centering}
-
-\caption{Example of a mesh partitioning onto four cores. Each single core partition
-is colored differently. The executable \texttt{xdecompose\_mesh} can
-equally distribute the mesh on any arbitrary number of cores. Domain
-decomposition is explained in detail in \citet{MaKoBlLe08}, and excellent
-scaling up to 150,000 processor cores in shown for instance in \citet{CaKoLaTiMiLeSnTr08,KoLaMi08a,MaKoBlLe08,KoErGoMi10,Kom11}.}
-
-\label{fig:mount.partitions}
-\end{figure}
-
-When you are ready to compile, in the main directory type
-\begin{lyxcode}
-{\small make~xdecompose\_mesh}{\small \par}
-\end{lyxcode}
-If all paths and flags have been set correctly, the executable \texttt{bin/xdecompose\_mesh}
-should be produced.
-
-The partitioning is done in serial for now (in the next release we
-will provide a parallel version of that code). It needs to be run in the main directory because it expects the \texttt{./DATA/Par\_file}. The synopsis is:
-\begin{lyxcode}
-./bin/xdecompose\_mesh~nparts~input\_directory~output\_directory
-\end{lyxcode}
-where
-\begin{itemize}
-\item \texttt{nparts} is the number of partitions, i.e., the number of cores
-for the parallel simulations,
-\item \texttt{input\_directory} is the directory which holds all the files
-generated by the Python script \texttt{cubit2specfem3d.py} explained
-in the previous Section~\ref{subsec:Exporting-the-Mesh}, e.g. \texttt{./MESH/},
-and
-\item \texttt{output\_directory} is the directory for the output of this
-partitioner which stores ACII-format files named like \texttt{proc{*}{*}{*}{*}{*}{*}\_Database}
-for each partition. These files will be needed for creating the distributed
-databases, and have to reside in the directory \texttt{LOCAL\_PATH}
-specified in the main \texttt{Par\_file}, e.g. in directory \texttt{./OUTPUT\_FILES/DATABASES\_MPI}.
-Please see Chapter~\ref{cha:Creating-Distributed-Databases} for
-further details.
-\end{itemize}
-Note that all the files generated by the Python script \texttt{cubit2specfem3d.py}
-must be placed in the \texttt{input\_directory} folder before running
-the program.
-
-\section{\label{cha:Running-the-Mesher-Meshfem3D}Meshing with \texttt{xmeshfem3D}}
-
-In case you successfully ran the configuration script, you are also
-ready to compile the internal mesher. This is an alternative to CUBIT
-for the mesh generation of relatively simple models.
-
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[scale=0.5]{figures/socal_map_mpi}
-\par\end{centering}
-
-\caption{\label{fig:For-parallel-computing}For parallel computing purposes,
-the model block is subdivided in $\nprocxi\times\nproceta$ slices
-of elements. In this example we use $5^{2}=25$ processors. }
-\end{figure}
-
-\noindent
-In the main directory, type
-\begin{lyxcode}
-{\small make~xmeshfem3D}{\small \par}
-\end{lyxcode}
-If all paths and flags have been set correctly, the mesher should
-now compile and produce the executable \texttt{bin/xmeshfem3D}. Please
-note that \texttt{xmeshfem3D} must be called directly from the main
-directory, as most of the binaries of the package.
-
-Input for the mesh generation program is provided through the parameter
-file \texttt{Mesh\_Par\_file}, which resides in the subdirectory \texttt{DATA/meshfem3D\_files/}.
-(to see how to use it, see the EXAMPLES specific to the internal mesher in directory \texttt{EXAMPLES/meshfem3D\_examples/}.
-Before running the mesher, a number of parameters need to be set in
-the \texttt{Mesh\_Par\_file}. This requires a basic understanding
-of how the SEM is implemented, and we encourage you to read \citet{KoVi98,KoTr99}
-and \citet{KoLiTrSuStSh04}.
-
-The mesher and the solver use UTM coordinates internally, therefore
-you need to define the zone number for the UTM projection (e.g., zone
-11 for Los Angeles). Use decimal values for latitude and longitude
-(no minutes/seconds). These values are approximate; the mesher will
-round them off to define a square mesh in UTM coordinates. When running
-benchmarks on rectangular models, turn the UTM projection off by using
-the flag \texttt{\small SUPPRESS\_UTM\_PROJECTION}{\small , in which
-case all `longitude' parameters simply refer to the $x$~axis, and
-all `latitude' parameters simply refer to the $y$~axis. To run the
-mesher for a global simulation, the following parameters need to be
-set in the }\texttt{\small Mesh\_Par\_file}{\small :}{\small \par}
-\begin{description}
-\item [{\texttt{LATITUDE\_MIN}}] Minimum latitude in the block (negative
-for South).
-\item [{\texttt{LATITUDE\_MAX}}] Maximum latitude in the block.
-\item [{\texttt{LONGITUDE\_MIN}}] Minimum longitude in the block (negative
-for West).
-\item [{\texttt{LONGITUDE\_MAX}}] Maximum longitude in the block.
-\item [{\texttt{DEPTH\_BLOCK\_KM}}] Depth of bottom of mesh in kilometers.
-\item [{\texttt{\noun{UTM\_PROJECTION\_ZONE}}}] UTM projection zone in
-which your model resides, only valid when \texttt{SUPPRESS\_UTM\_}~\\
- \texttt{PROJECTION} is \texttt{.false.}.
-Use a negative zone number for the Southern hemisphere:
-the Northern hemisphere corresponds to zones +1 to +60,
-the Southern hemisphere to zones -1 to -60.
-We use the WGS84 (World Geodetic System 1984) reference ellipsoid for the UTM projection. If you prefer to use the Clarke 1866 ellipsoid,
-edit file \texttt{src/shared/utm\_geo.f90}, uncomment that ellipsoid and recompile the code.
-From http://en.wikipedia.org/wiki/Universal\_Transverse\_Mercator\_coordinate\_system :
-The Universal Transverse Mercator coordinate system was developed by the United States Army Corps of Engineers in the 1940s.
-The system was based on an ellipsoidal model of Earth. For areas within the contiguous United States
-the Clarke Ellipsoid of 1866 was used. For the remaining areas of Earth, including Hawaii, the International Ellipsoid was used.
-The WGS84 ellipsoid is now generally used to model the Earth in the UTM coordinate system,
-which means that current UTM northing at a given point can be 200+ meters different from the old one.
-For different geographic regions, other datum systems (e.g.: ED50, NAD83) can be used.
-\item [{\texttt{SUPPRESS\_UTM\_PROJECTION}}] set to be \texttt{.false.}
-when your model range is specified in geographical coordinates, and
-needs to be \texttt{.true.} when your model is specified in Cartesian
-coordinates. \noun{UTM projection zone in which your simulation region
-resides.}
-\item [{\texttt{INTERFACES\_FILE }}] File which contains the description
-of the topography and of the interfaces between the different layers
-of the model, if any. The number of spectral elements in the vertical
-direction within each layer is also defined in this file.
-\item [{$\nexxi$}] The number of spectral elements along one side of the
-block. This number \textit{must} be 8~$\times$~a multiple of $\nprocxi$
-defined below. Based upon benchmarks against semi-analytical discrete
-wavenumber synthetic seismograms \citep{KoLiTrSuStSh04}, determined
-that a $\nexxi=288$ run is accurate to a shortest period of roughly
-2~s. Therefore, since accuracy is determined by the number of grid
-points per shortest wavelength, for any particular value of $\nexxi$
-the simulation will be accurate to a shortest period determined by
-\begin{equation}
-\mbox{shortest period (s)}=(288/\nexxi)\times2.\label{eq:shortest_period}
-\end{equation}
- The number of grid points in each orthogonal direction of the reference
-element, i.e., the number of Gauss-Lobatto-Legendre points, is determined
-by \texttt{NGLLX} in the \texttt{constants.h} file. We generally use
-$\mbox{\texttt{NGLLX\/}}=5$, for a total of $5^{3}=125$ points per
-elements. We suggest not to change this value.
-\item [{$\nexeta$}] The number of spectral elements along the other side
-of the block. This number \textit{must} be 8~$\times$~a multiple
-of $\nproceta$ defined below.
-\item [{$\nprocxi$}] The number of processors or slices along one side
-of the block (see Figure~\ref{fig:For-parallel-computing}); we must
-have $\nexxi=8\times c\times\nprocxi$, where $c\ge1$ is a positive
-integer.
-\item [{$\nproceta$}] The number of processors or slices along the other
-side of the block; we must have $\nexeta=8\times c\times\nproceta$,
-where $c\ge1$ is a positive integer.
-\item [{\texttt{USE\_REGULAR\_MESH}}] set to be \texttt{.true.} if you
-want a perfectly regular mesh or \texttt{.false.} if you want to add
-doubling horizontal layers to coarsen the mesh. In this case, you
-also need to provide additional information by setting up the next
-three parameters.
-\item [{\texttt{NDOUBLINGS}}] The number of horizontal doubling layers.
-Must be set to \texttt{1} or \texttt{2} if \texttt{USE\_REGULAR\_MESH}
-is set to \texttt{.true.}.
-\item [{\texttt{NZ\_DOUBLING\_1}}] The position of the first doubling layer
-(only interpreted if \texttt{USE\_REGULAR\_MESH} is set to \texttt{.true.}).
-\item [{\texttt{NZ\_DOUBLING\_2}}] The position of the second doubling
-layer (only interpreted if \texttt{USE\_REGULAR\_MESH} is set to \texttt{.true.}
-and if \texttt{NDOUBLINGS} is set to \texttt{2}).
-\item [{\texttt{CREATE\_ABAQUS\_FILES}}] Set this flag to \texttt{.true.}
-to save Abaqus FEA \urlwithparentheses{www.simulia.com} mesh files
-for subsequent viewing. Turning the flag on generates files in the
-\texttt{LOCAL\_PATH} directory. See Section~\ref{sec:Mesh-graphics}
-for a discussion of mesh viewing features.
-\item [{\texttt{CREATE\_DX\_FILES}}] Set this flag to \texttt{.true.} to
-save OpenDX \urlwithparentheses{www.opendx.org} mesh files for subsequent
-viewing.
-\item [{\texttt{LOCAL\_PATH}}] Directory in which the partitions generated
-by the mesher will be written. Generally one uses a directory on the
-local disk of the compute nodes, although on some machines these partitions
-are written on a parallel (global) file system (see also the earlier
-discussion of the \texttt{LOCAL\_PATH\_IS\_ALSO\_GLOBAL} flag in Chapter~\ref{cha:Getting-Started}).
-The mesher generates the necessary partitions in parallel, one set
-for each of the $\nprocxi\times\nproceta$ slices that constitutes
-the mesh (see Figure~\ref{fig:For-parallel-computing}). After the
-mesher finishes, you can log in to one of the compute nodes and view
-the contents of the \texttt{LOCAL\_PATH} directory to see the files
-generated by the mesher. These files will be needed for creating the
-distributed databases, and have to reside in the directory \texttt{LOCAL\_PATH}
-specified in the main \texttt{Par\_file}, e.g. in directory \texttt{OUTPUT\_FILES/DATABASES\_MPI}.
-Please see Chapter~\ref{cha:Creating-Distributed-Databases} for
-further details.
-\item [{\texttt{NMATERIALS}}] The number of different materials in your
-model. In the following lines, each material needs to be defined as
-:
-
-\begin{lyxcode}
-material\_ID~rho~vp~vs~Q~anisotropy\_flag~domain\_ID
-\end{lyxcode}
-
-where
-\begin{itemize}
-\item \texttt{Q} : quality factor (0=no attenuation)
-\item \texttt{anisotropy\_flag} : 0=no anisotropy / 1,2,.. check with implementation
-in \texttt{aniso\_model.f90}
-\item \texttt{domain\_id} : 1=acoustic / 2=elastic
-\end{itemize}
-\item [{\texttt{NREGIONS}}] The number of regions in the mesh. In the following
-lines, because the mesh is regular or 'almost regular', each region
-is defined as :
-
-\begin{lyxcode}
-NEX\_XI\_BEGIN~NEX\_XI\_END~NEX\_ETA\_BEGIN~NEX\_ETA\_END~NZ\_BEGIN~NZ\_END~material\_ID
-\end{lyxcode}
-\end{description}
-The \texttt{INTERFACES\_FILE} parameter of \texttt{Mesh\_Par\_File}
-defines the file which contains the settings of the topography grid
-and of the interfaces grids. Topography is defined as a set of elevation
-values on a regular 2D grid. It is also possible to define interfaces
-between the layers of the model in the same way. The file needs to
-define several parameters:
-\begin{itemize}
-\item The number of interfaces, including the topography. This needs to
-be set at the first line. Then, from the bottom to the top of the
-model, you need to define the grids with:
-\item \texttt{SUPPRESS\_UTM\_PROJECTION} flag as described previously,
-\item number of points along $x$ and $y$ direction (NXI and NETA),
-\item minimal $x$ and $y$ coordinates (LONG\_MIN and LAT\_MIN),
-\item spacing between points along $x$ and $y$ (SPACING\_XI and SPACING\_ETA)
-and
-\item the name of the file which contains the elevation values (in $y$.$x$
-increasing order).
-\end{itemize}
-At the end of this file, you simply need to set the number of spectral
-elements in the vertical direction for each layer. We provide a few
-models in the {\texttt{EXAMPLES/}} directory. \\
-
-Finally, depending on your system, you might need to provide a file
-that tells MPI what compute nodes to use for the simulations. The
-file must have a number of entries (one entry per line) at least equal
-to the number of processors needed for the run. A sample file is provided
-in the file \texttt{mymachines}. This file is not used by the mesher
-or solver, but is required by the \texttt{go\_mesher} and \texttt{go\_solver}
-default job submission scripts. See Chapter~\ref{cha:Scheduler}
-for information about running the code on a system with a scheduler,
-e.g., LSF.
-
-Now that you have set the appropriate parameters in the \texttt{Mesh\_Par\_file}
-and have compiled the mesher, you are ready to launch it! This is
-most easily accomplished based upon the \texttt{go\_mesher} script.
-When you run on a PC cluster, the script assumes that the nodes are
-named n001, n002, etc. If this is not the case, change the \texttt{tr
--d `n'} line in the script. You may also need to edit the last command
-at the end of the script that invokes the \texttt{mpirun} command.
-See Chapter~\ref{cha:Scheduler} for information about running the
-code on a system with a scheduler, e.g., LSF.
-
-Mesher output is provided in the \texttt{OUTPUT\_FILES} directory
-in \texttt{output\_mesher.txt}; this file provides lots of details
-about the mesh that was generated. Please note that the mesher suggests
-a time step \texttt{DT} to run the solver with. The mesher output
-file also contains a table about the quality of the mesh to indicate
-possible problems with the distortions of elements. Alternatively,
-output can be directed to the screen instead by uncommenting a line
-in \texttt{constants.h}:
-\begin{lyxcode}
-!~uncomment~this~to~write~messages~to~the~screen~
-
-!~integer,~parameter~::~IMAIN~=~ISTANDARD\_OUTPUT
-\end{lyxcode}
-To control the quality of the mesh, check the standard output (either
-on the screen or in the \texttt{OUTPUT\_FILES} directory in \texttt{output\_mesher.txt})
-and analyze the skewness histogram of mesh elements to make sure no
-element has a skewness above approximately 0.75, otherwise the mesh
-is of poor quality (and if even a single element has a skewness value
-above 0.80, then you must definitely improve the mesh). To draw the
-skewness histogram on the screen, type \texttt{gnuplot plot\_mesh\_quality\_histogram.gnu}.
-
-\chapter{\label{cha:Creating-Distributed-Databases}Creating the Distributed
-Databases}
-
-After using \texttt{xmeshfem3D} or \texttt{xdecompose\_mesh}, the
-next step in the workflow is to compile \texttt{xgenerate\_}~\\
- \texttt{databases}. This program is going to create all the missing
-information needed by the SEM solver.
-
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=0.6\textwidth]{figures/workflow}
-\par\end{centering}
-
-\caption{Schematic workflow for a SPECFEM3D Cartesian simulation. The executable
-\texttt{xgenerate\_databases} creates the GLL mesh points and assigns
-specific model parameters.}
-
-\label{fig:workflow.databases}
-\end{figure}
-
-\noindent
-In the main directory, type
-\begin{lyxcode}
-{\small make~xgenerate\_databases}{\small \par}
-\end{lyxcode}
-Input for the program is provided through the main parameter file
-\texttt{Par\_file}, which resides in the subdirectory \texttt{DATA}.
-Please note that \texttt{xgenerate\_databases} must be called directly
-from the main directory, as most of the binaries of the package.
-
-\section{\label{cha:Main-Parameter}Main parameter file \texttt{Par\_file}}
-
-Before running \texttt{xgenerate\_databases}, a number of parameters
-need to be set in the main parameter \texttt{Par\_file} located in
-the subdirectory \texttt{DATA}:
-\begin{description}
-\item [{\texttt{SIMULATION\_TYPE}}] is set to 1 for forward simulations,
-2 for adjoint simulations (see Section \ref{sec:Adjoint-simulation-finite})
-and 3 for kernel simulations (see Section \ref{sec:Finite-Frequency-Kernels}).
-\item [{\texttt{SAVE\_FORWARD}}] is only set to \texttt{.true.} for a forward
-simulation with the last frame of the simulation saved, as part of
-the finite-frequency kernel calculations (see Section \ref{sec:Finite-Frequency-Kernels}).
-For a regular forward simulation, leave \texttt{SIMULATION\_TYPE}
-and \texttt{SAVE\_FORWARD} at their default values.
-\item [{\texttt{\noun{UTM\_PROJECTION\_ZONE}}}] UTM projection zone in
-which your model resides, only valid when \texttt{SUPPRESS\_UTM\_}~\\
- \texttt{PROJECTION} is \texttt{.false.}.
-\item [{\texttt{SUPPRESS\_UTM\_PROJECTION}}] set to be \texttt{.false.}
-when your model range is specified in the geographical coordinates,
-and needs to be \texttt{.true.} when your model is specified in a
-cartesian coordinates. \noun{UTM projection zone in which your simulation
-region resides.}
-\item [{\texttt{NPROC}}] The number of MPI processors, each one is assigned
-one slice of the whole mesh.
-\item [{\texttt{NSTEP}}] The number of time steps of the simulation. This
-controls the length of the numerical simulation, i.e., twice the number
-of time steps requires twice as much CPU time. This feature is not
-used at the time of generating the distributed databases but is required
-for the solver, i.e., you may change this parameter after running
-\texttt{xgenerate\_databases}.
-\item [{\texttt{DT}}] The length of each time step in seconds. This feature
-is not used at the time of generating the distributed databases but
-is required for the solver. Please see also Section~\ref{sec:Choosing-the-Time-Step}
-for further details.
-\item [{\texttt{NGNOD}}] The number of nodes for 2D and 3D shape functions
-for hexahedra. We use either 8-node mesh elements (bricks) or 27-node
-elements. If you use the internal mesher, the only option is 8-node
-bricks (27-node elements are not supported). \texttt{CUBIT} does not
-support HEX27 elements either (it can generate them, but they are
-flat, i.e. identical to HEX8). To generate HEX27 elements with curvature
-properly taken into account, you can use \texttt{Gmsh} \url{http://geuz.org/gmsh/}
-\item [{\texttt{MODEL}}] Must be set to one of the following:
-
-\begin{description}
-\item [{\textmd{Models defined by mesh parameters:}}] ~
-
-\begin{description}
-\item [{\texttt{default}}] Uses model parameters as defined by meshing
-procedures described in the previous Chapter~\ref{cha:Mesh-Generation}.
-\end{description}
-
-\item [{\textmd{1D~models~with~real~structure:}}] ~
-
-\begin{description}
-\item [{\texttt{1D\_prem}}] Isotropic version of the spherically symmetric
-Preliminary Reference Earth Model (PREM) \citep{DzAn81}.
-\item [{\texttt{1D\_socal}}] A standard isotropic 1D model for Southern
-California.
-\item [{\texttt{1D\_cascadia}}] Isotropic 1D profile for the Cascadia region.
-\end{description}
-
-\item [{\textmd{Fully~3D~models:}}] ~
-
-\begin{description}
-\item [{\texttt{aniso}}] For a user-specified fully anisotropic model.
-Parameters are set up in routines located in file \texttt{model\_aniso.f90}
-in directory \texttt{src/generate\_databases/}. See Chapter~\ref{cha:-Changing-the}
-for a discussion on how to specify your own 3D model.
-
-\item [{\texttt{external}}] For a user-specified isotropic model which
-uses externally defined model parameters. Uses external model definitions
-set up in routines located in file \texttt{model\_external\_values.f90}
-in directory \texttt{src/generate\_databases/}. Please modify these
-generic template routines to use your own model definitions.
-\item [{\texttt{gll}}] For a user-specified isotropic model which uses
-external binary files for $v_{p}$, $v_{s}$ and $\rho$. Binary files
-are given in the same format as when outputted by the \texttt{xgenerate\_databases}
-executable when using option \texttt{SAVE\_MESH\_FILES}. These binary
-files define the model parameters on all GLL points which can be used
-for iterative inversion procedures.
-\item [{\texttt{salton\_trough}}] A 3D $V_{p}$ model for Southern California.
-Users must provide the corresponding data file \texttt{regrid3\_vel\_p.bin}
-in directory \texttt{DATA/st\_3D\_block\_harvard/}.
-\item [{\texttt{tomo}}] For a user-specified 3D isotropic model which uses
-a tomographic model file \texttt{tomographic\_model.xyz} in directory
-\texttt{DATA}. See Section \ref{sec:Using-tomographic}, for a discussion
-on how to specify your own 3D tomographic model.
-\end{description}
-\end{description}
-\item [{\texttt{APPROXIMATE\_OCEAN\_LOAD}}] Set to \texttt{.true.} if the
-effect of the oceans on seismic wave propagation should be incorporated
-based upon the (rough) approximate treatment discussed in \citet{KoTr02b}.
-This feature is inexpensive from a numerical perspective, both in
-terms of memory requirements and CPU time. This approximation is accurate
-at periods of roughly 20~s and longer. At shorter periods the effect
-of water phases/reverberations is not taken into account, even when
-the flag is on. If you want to model the effect of a fluid-solid model
-at short periods, then set this flag to \texttt{.false.} and mesh
-the fluid layer explicitly in your mesher, so that it is computed
-accurately and without this approximation.
-\item [{\texttt{TOPOGRAPHY}}] This feature is only effective if \texttt{APPROXIMATE\_OCEAN\_LOAD}
-is set to \texttt{.true.}. Set to \texttt{.true.} if topography and
-bathymetry should be read in based upon the topography file specified
-in the main constants file \texttt{constants.h} found in subdirectory
-\texttt{src/shared/} to evaluate elevations. If not set, elevations
-will be read from the numerical mesh.
-\item [{\texttt{ATTENUATION}}] Set to \texttt{.true.} if attenuation should
-be incorporated. Turning this feature on increases the memory requirements
-significantly (roughly by a factor of~1.5), and is numerically fairly
-expensive. See \citet{KoTr99,KoTr02a} for a discussion on the implementation
-of attenuation based upon standard linear solids. Please note that
-the Vp- and Vs-velocities of your model are given for a reference
-frequency. To change this reference frequency, you change the value
-of \texttt{ATTENUATION\_f0\_REFERENCE} in the main constants file
-\texttt{constants.h} found in subdirectory \texttt{src/shared/}.
-\item [{\texttt{FULL\_ATTENUATION\_SOLID}}] Set to \texttt{.true.} to add
-$Q_{\kappa}$ attenuation to the simulations in addition to $Q_{\mu}$
-attenuation when the \texttt{ATTENUATION} flag above is on. Should
-be off in almost all cases for seismic wave propagation, since $Q_{\kappa}$
-is almost always negligible in Earth models. However, this parameter
-can be useful for instance for ocean acoustics simulations, when $Q_{\kappa}$
-is not negligible in the solid layer at the sea bottom.
-\item [{\texttt{ANISOTROPY}}] Set to \texttt{.true.} if you want to use
-an anisotropy model. Please see the file \texttt{model\_aniso.f90}
-in subdirectory \texttt{src/generate\_databases/} for the current
-implementation of anisotropic models.
-\item [{\texttt{TOMOGRAPHY\_PATH}}] Directory in which the tomography files
-are stored for using external tomographic Earth models (please read
-Chapter \ref{cha:-Changing-the} and Section \ref{sec:Using-tomographic}
-`Using external tomographic Earth models' for further details.).
-\item [{\texttt{USE\_OLSEN\_ATTENUATION}}] Set to \texttt{.true.} if you
-want to use the attenuation model that scaled from the S-wave speed
-model using Olsen's empirical relation (see \citet{OlDaBr03}).
-\item [{\texttt{OLSEN\_ATTENUATION\_RATIO}}] Determines the Olsen's constant
-in Olsen's empirical relation (see \citet{OlDaBr03}).
-\item [{\texttt{PML\_CONDITIONS}}] Set to \texttt{.true.} to turn on C-PML
-boundary conditions for a regional simulation. Both fluids and elastic
-solids are supported. Note that xmeshfem3d generated meshes do not support C-PML yet.
-\item [{\texttt{PML\_INSTEAD\_OF\_FREE\_SURFACE}}] Set to \texttt{.true.}
-to turn on C-PML boundary conditions on the top surface instead of
-the usual free surface.
-\item [{\texttt{f0\_FOR\_PML}}] Determines the dominant frequency that
-will be used in the calculation of PML damping profiles; this should
-be set to the same (or similar) dominant frequency as that of the
-source that you will use in your simulation. If you plan to use a
-Dirac source, then use the dominant frequency of the source wavelet
-with which you plan to convolve your seismograms later on in post-processing.
-\item [{\texttt{STACEY\_ABSORBING\_CONDITIONS}}] Set to \texttt{.true.}
-to turn on Clayton-Enquist absorbing boundary conditions (see \citet{KoTr99}).
-In almost all cases it is much better to use CPML absorbing layers
-(see the options above) and leave this flag to \texttt{.false.}.
-\item [{\texttt{STACEY\_INSTEAD\_OF\_FREE\_SURFACE}}] Set to \texttt{.true.}
-to turn on absorbing boundary conditions on the top surface which
-by default constitutes a free surface of the model.
-\item [{\texttt{CREATE\_SHAKEMAP}}] Set this flag to \texttt{.true.} to
-create a ShakeMap\textregistered{}, i.e., a peak ground velocity map
-of the maximum absolute value of the two horizontal components of
-the velocity vector.
-\item [{\texttt{MOVIE\_SURFACE}}] Set to \texttt{.false.}, unless you want
-to create a movie of seismic wave propagation on the Earth's surface.
-Turning this option on generates large output files. See Section~\ref{sec:Movies}
-for a discussion on the generation of movies. This feature is only
-relevant for the solver.
-\item [{\texttt{MOVIE\_TYPE}}] Set this flag to 1 to show the top surface
-(tomography + oceans) only, to 2 to show all external faces of the
-mesh (i.e. topography + vertical edges + bottom) in shakemaps and
-surface movies.
-\item [{\texttt{MOVIE\_VOLUME}}] Set to \texttt{.false.}, unless you want
-to create a movie of seismic wave propagation in the Earth's interior.
-Turning this option on generates huge output files. See Section~\ref{sec:Movies}
-for a discussion on the generation of movies. This feature is only
-relevant for the solver.
-\item [{\texttt{SAVE\_DISPLACEMENT}}] Set this flag to \texttt{.true.}
-if you want to save the displacement instead of velocity for the movie
-frames.
-\item [{\texttt{USE\_HIGHRES\_FOR\_MOVIES}}] Set this flag to \texttt{.true.}
-if you want to save the values at all the NGLL grid points for the
-movie frames.
-\item [{\texttt{NTSTEP\_BETWEEN\_FRAMES}}] Determines the number of timesteps
-between movie frames. Typically you want to save a snapshot every
-100 timesteps. The smaller you make this number the more output will
-be generated! See Section~\ref{sec:Movies} for a discussion on the
-generation of movies. This feature is only relevant for the solver.
-\item [{\texttt{HDUR\_MOVIE}}] Determines the half duration of the source
-time function for the movie simulations. When this parameter is set
-to be 0, a default half duration that corresponds to the accuracy
-of the simulation is provided. Otherwise, it adds this half duration
-to the half duration specified in the source file \texttt{CMTSOLUTION},
-thus simulates longer periods to make the movie images look smoother.
-\item [{\texttt{SAVE\_MESH\_FILES}}] Set this flag to \texttt{.true.} to
-save ParaView \urlwithparentheses{www.paraview.org} mesh files for
-subsequent viewing. Turning the flag on generates large (distributed)
-files in the \texttt{LOCAL\_PATH} directory. See Section~\ref{sec:Mesh-graphics}
-for a discussion of mesh viewing features.
-\item [{\texttt{LOCAL\_PATH}}] Directory in which the distributed databases
-will be written. Generally one uses a directory on the local disk
-of the compute nodes, although on some machines these databases are
-written on a parallel (global) file system (see also the earlier discussion
-of the \texttt{LOCAL\_PATH\_IS\_ALSO\_GLOBAL} flag in Chapter~\ref{cha:Getting-Started}).
-\texttt{xgenerate\_databases} generates the necessary databases in
-parallel, one set for each of the \texttt{NPROC} slices that constitutes
-the mesh (see Figure~\ref{fig:mount.partitions} and Figure~\ref{fig:For-parallel-computing}).
-After the executable finishes, you can log in to one of the compute
-nodes and view the contents of the \texttt{LOCAL\_PATH} directory
-to see the (many) files generated by \texttt{xgenerate\_databases}.
-Please note that the \texttt{LOCAL\_PATH} directory should already
-contain the output files of the partitioner, i.e. from \texttt{xdecompose\_mesh}
-or \texttt{xmeshfem3D}.
-\item [{\texttt{NTSTEP\_BETWEEN\_OUTPUT\_INFO}}] This parameter specifies
-the interval at which basic information about a run is written to
-the file system (\texttt{timestamp{*}} files in the \texttt{OUTPUT\_FILES}
-directory). If you have access to a fast machine, set \texttt{NTSTEP\_BETWEEN\_OUTPUT\_INFO}
-to a relatively high value (e.g., at least 100, or even 1000 or more)
-to avoid writing output text files too often. This feature is not
-used at the time of meshing. One can set this parameter to a larger
-value than the number of time steps to avoid writing output during
-the run.
-\item [{\texttt{NTSTEP\_BETWEEN\_OUTPUT\_SEISMOS}}] This parameter specifies
-the interval at which synthetic seismograms are written in the \texttt{LOCAL\_PATH}
-directory. If a run crashes, you may still find usable (but shorter
-than requested) seismograms in this directory. On a fast machine set
-\texttt{NTSTEP\_BETWEEN\_OUTPUT\_SEISMOS} to a relatively high value
-to avoid writing to the seismograms too often. This feature is only
-relevant for the solver.
-\item [{\texttt{USE\_FORCE\_POINT\_SOURCE}}] Turn this flag on to use a
-(tilted) \texttt{FORCESOLUTION} force point source instead of a \texttt{CMTSOLUTION}
-moment-tensor source. When the force source does not fall exactly
-at a grid point, the solver interpolates the force between grid points
-using Lagrange interpolants. This can be useful e.g. for oil industry
-foothills simulations in which the source is a vertical force, normal
-force, tilted force, or an impact etc. Note that in the \texttt{FORCESOLUTION}
-file, you will need to edit the East, North and vertical components
-of an arbitrary (not necessarily unitary, the code will normalize it automatically) direction vector of the force vector;
-thus refer to Appendix~\ref{cha:Coordinates} for the orientation
-of the reference frame. This vector is made unitary internally in
-the solver and thus only its direction matters here; its norm is ignored
-and the norm of the force used is the factor force source times the
-source time function.
-When using this option, by default the code can locate the force source
-anywhere between mesh points in order to honor its exact location; this
-is more precise than using the closest GLL mesh point, but it is also a bit slower.
-If needed, you can change that default behavior and force the code to use the closest
-GLL mesh point instead by setting flag \texttt{USE\_BEST\_LOCATION} to \texttt{.false.} instead of \texttt{.true.} in file
-\texttt{src/shared/constants.h.in} and running the \texttt{configure} script again and recompiling the code.
-\item [{\texttt{USE\_RICKER\_TIME\_FUNCTION}}] Turn this flag on to use
-a Ricker source time function instead of the source time functions
-set by default to represent a (tilted) \texttt{FORCESOLUTION} force
-point source or a \texttt{CMTSOLUTION} moment-tensor source. Originally,
-if a \texttt{CMTSOLUTION} moment-tensor source is used, a (pseudo)
-Heaviside step function with a very short half duration is defined
-for elastic cases to represent the permanent slip on the fault while
-in the acoustic case a Gaussian source time function with a similarly
-short half duration is defined to physically describe actions within
-the fluid. Otherwise, if a \texttt{FORCESOLUTION} force source is
-used, a (pseudo) Dirac delta source time function is defined by default.
-Any other source-time function may then be obtained by convolution.
-\item [{\texttt{PRINT\_SOURCE\_TIME\_FUNCTION}}] Turn this flag on to print
-information about the source time function in the file \texttt{OUTPUT\_FILES/plot\_source\_time\_function.txt}.
-This feature is only relevant for the solver.
-
-\item [{\texttt{NUMBER\_OF\_SIMULTANEOUS\_RUNS}}] adds the ability to run several calculations (several earthquakes)
-in an embarrassingly-parallel fashion from within the same run;
-this can be useful when using a very large supercomputer to compute
-many earthquakes in a catalog, in which case it can be better from
-a batch job submission point of view to start fewer and much larger jobs,
-each of them computing several earthquakes in parallel.
-To turn that option on, set parameter NUMBER\_OF\_SIMULTANEOUS\_RUNS to a value greater than 1.
-To implement that, we create NUMBER\_OF\_SIMULTANEOUS\_RUNS MPI sub-communicators,
-each of them being labeled "my\_local\_mpi\_comm\_world", and we use them
-in all the routines in "src/shared/parallel.f90", except in MPI\_ABORT() because in that case
-we need to kill the entire run.
-When that option is on, of course the number of processor cores used to start
-the code in the batch system must be a multiple of NUMBER\_OF\_SIMULTANEOUS\_RUNS,
-all the individual runs must use the same number of processor cores,
-which as usual is NPROC in the Par\_file,
-and thus the total number of processor cores to request from the batch system
-should be NUMBER\_OF\_SIMULTANEOUS\_RUNS * NPROC.
-All the runs to perform must be placed in directories called run0001, run0002, run0003 and so on
-(with exactly four digits).
-
-Imagine you have 10 independent calculations to do, each of them on 100 cores; you have three options:
-
-1/ submit 10 jobs to the batch system
-
-2/ submit a single job on 1000 cores to the batch, and in that script create a sub-array of jobs to start 10 jobs,
-each running on 100 cores (see e.g. http://www.schedmd.com/slurmdocs/job\_array.html )
-
-3/ submit a single job on 1000 cores to the batch, start SPECFEM3D on 1000 cores, create 10 sub-communicators,
-cd into one of 10 subdirectories (called e.g. run0001, run0002,... run0010) depending on the sub-communicator
-your MPI rank belongs to, and run normally on 100 cores using that sub-communicator.
-
-The option NUMBER\_OF\_SIMULTANEOUS\_RUNS implements 3/.
-
-\item [{\texttt{BROADCAST\_SAME\_MESH\_AND\_MODEL}}]:
-if we perform simultaneous runs in parallel, if only the source and receivers vary between these runs
-but not the mesh nor the model (velocity and density) then we can also read the mesh and model files
-from a single run in the beginning and broadcast them to all the others; for a large number of simultaneous
-runs for instance when solving inverse problems iteratively this can DRASTICALLY reduce I/Os to disk in the solver
-(by a factor equal to NUMBER\_OF\_SIMULTANEOUS\_RUNS), and reducing I/Os is crucial in the case of huge runs.
-Thus, always set this option to .true. if the mesh and the model are the same for all simultaneous runs.
-In that case there is no need to duplicate the mesh and model file database (the content of the DATABASES\_MPI
-directories) in each of the run0001, run0002,... directories, it is sufficient to have one in run0001
-and the code will broadcast it to the others).
-
-TODO / future work to do: currently the BROADCAST\_SAME\_MESH\_AND\_MODEL option assumes to have the (master) mesh files in run0001/DATABASES\_MPI or run0001/OUTPUT\_FILES/DATABASES\_MPI.
-However, for adjoint runs you still need a DATABASES\_MPI folder in each of the sub-runs directories, e.g.
-run0002/DATABASES\_MPI, etc.
-to store the forward wavefields, kernels etc.
-of each sub-run.
-This would not be needed for forward simulations.
-
-TODO / future work to do: the sensitivity kernel summing and smoothing tools in directory src/tomography are currently not ported to this new option to do many runs simultaneously, only the solver (src/specfem3d) is.
-Thus these tools should work, but in their current version will need to be run for each simulation result independently.
-More precisely, the current kernel summing and smoothing routines work fine, with the exception that you need to move out the mesh files (and also the parameters).
-This works because these routines consider multiple runs by design.
-You simply have to provide them the directories where the kernels are.
-
-\item [{\texttt{GPU\_MODE}}] Turn this flag on to use GPUs.
-
-\item [\texttt{ADIOS\_ENABLED}] Turn this flag on to enable ADIOS. If set to \texttt{.false.}, subsequent ADIOS
-parameters will not be considered.
-\item [\texttt{ADIOS\_FOR\_DATABASES}] Turn this flag on to use ADIOS for xmeshfem3D output and
-xgenerate\_database input.
-\item [\texttt{ADIOS\_FOR\_MESH}]  Turn this flag on to use ADIOS for generated databases.
-\item [\texttt{ADIOS\_FOR\_FORWARD\_ARRAYS}] Turn this flag on to read and write forward arrays using ADIOS.
-\item [\texttt{ADIOS\_FOR\_KERNELS}] Turn this flag on to produce ADIOS kernels that can later be visualized with the ADIOS version of combine\_vol\_data.
-
-\end{description}
-
-The present version of SPECFEM can handle fully saturated porous simulations, Christina Morency implemented Biot equation.
-But the code cannot calculate partially saturated cases in its state.
-Christina Morency is presently working on a dual porosity, dual permeability formulation, type Pride and Berryman, for an other project, but it will not be available for some time.\\
-
-The way we prescribe material property for porous material in SPECFEM3D is as follow:
-We use a file name nummaterial\_poroelastic\_file, which is located in the directory MESH, the format is as follow:\\
- rhos rhof phi c kxx kxy kxz kyy kyz kzz Ks Kf Kfr etaf mufr\\
-where\\
-rho\_s = solid density\\
-rho\_f = fluid density\\
-phi = porosity\\
-tort = tortuosity\\
-kxx = xx component of permeability tensor\\
-kxy = xy,yx components of permeability tensor\\
-kyy = yy component of permeability tensor\\
-kxz = xz,zx components of permeability tensor\\
-kzz = zz component of permeability tensor\\
-kappa\_s = solid bulk modulus\\
-kappa\_f= fluid bulk modulus\\
-kappa\_fr= frame bulk modulus\\
-eta\_f = fluid viscosity\\
-mu\_fr = frame shear modulus.\\
-
-Using an external mesh (for instance coming from CUBIT/TRELIS), poroelastic materials have the ID number 3, while 1 is elastic, 2 is acoustic ?
-see the example in the package: examples/homogeneous\_poroelastic.
-
-If you use PML, the mesh elements that belong to the PML layers can
-be acoustic or elastic, but not viscoelastic nor poroelastic. Then,
-when defining your model, you should define these absorbing elements
-as either acoustic or elastic. In you forget to do that, the code
-will fix the problem by automatically converting the viscoelastic
-or poroelastic PML elements to elastic. This means that strictly speaking
-the PML layer will not be perfectly matched any more, since the physical
-model will change from viscoelastic or poroelastic to elastic at the
-entrance of the PML, but in practice this is sufficient and produces
-only tiny / negligible spurious reflections.
-
-If you use PML and an external mesh (created using an external meshing tool
-such as CUBIT/TRELIS or similar), try to have elements inside the PML as regular as possible,
-i.e. ideally non-deformed cubes obtained by `extrusion' of regular surface mesh elements meshing the
-outer edges of the computational domain without PML; by doing so, the PMLs obtained will be far more stable
-in time (PML being weakly unstable from a mathematical point of view, very deformed mesh elements
-inside the PMLs can trigger instabilities much more quickly).
-
-If you use PML and an external tomographic velocity and density model,
-you should be careful because mathematically a PML cannot handle heterogeneities
-along the normal to the PML edge inside the PML layer. This comes
-from the fact that the damping profile that is defined assumes a constant
-velocity and density model along the normal direction.
-
-Thus, you need to modify your velocity and density model in order
-for it to be 1D inside the PML, as shown in Figure~\ref{fig:modify_external_velocity_model_to_use_PML}.
-
-This applies to the bottom layer as well; there you should make sure
-that your model is 1D and thus constant along the vertical direction.
-
-To summarize, only use a 3D velocity and density model inside the
-physical region, and in all the PML layers extend it by continuity
-from its values along the inner PML edge.
-
-%%
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=3in]{figures/how_to_use_PML_when_a_tomographic_velocity_model_is_used}
-\par\end{centering}
-
-\caption{How to modify your external 3D velocity and density model in order
-to use PML. Such a modification is not needed when using Stacey absorbing
-boundary conditions (but such conditions are significantly less efficient).}
-
-\label{fig:modify_external_velocity_model_to_use_PML}
-\end{figure}
-
-%%
-
-\section{\label{sec:Choosing-the-Time-Step}Choosing the time step \texttt{DT}}
-
-The parameter \texttt{DT} sets the length of each time step in seconds.
-The value of this parameter is crucial for the stability of the spectral-element
-simulation. Your time step \texttt{DT} will depend on the minimum
-ratio between the distance $h$ of neighboring mesh points and the
-wave speeds $v$ defined in your model. The condition for the time
-step $\Delta t$ is:
-\begin{lyxcode}
-$\Delta t<C~\mathrm{min}_{\Omega}(~h/v~)$
-\end{lyxcode}
-where $C$ is the so-called Courant number and $\Omega$ denotes the
-model volume. The distance $h$ depends on the mesh element size and
-the number of GLL points \texttt{NGLL} specified in the main constants
-file \texttt{constants.h} located in the \texttt{src/shared/} subdirectory.
-The wave speed $v$ is determined based on your model's P- (or S-)
-wave speed values.
-
-The database generator \texttt{xgenerate\_databases}, as well as the
-internal mesher \texttt{xmeshfem3D}, are trying to evaluate the value
-of $\Delta t$ for empirically chosen Courant numbers $C\sim0.3$.
-If you used the mesher \texttt{xmeshfem3D} to generate your mesh,
-you should set the value suggested in \texttt{OUTPUT\_FILES/output\_mesher.txt}
-file, which is created after the mesher completed. In case you used
-CUBIT to create the mesh, you might use an arbitrary value when running
-\texttt{xgenerate\_databases} and then use the value suggested in
-the ~\\
- \texttt{OUTPUT\_FILES/output\_mesher.txt} file after the database
-generation completed. Note that the implemented Newmark time scheme
-uses this time step globally, thus your simulations become more expensive
-for very small mesh elements in high wave-speed regions. Please be
-aware of this restriction when constructing your mesh in Chapter~\ref{cha:Mesh-Generation}.
-
-\chapter{\label{cha:Running-the-Solver}Running the Solver \texttt{xspecfem3D}}
-
-Now that you have successfully generated the databases, you are ready
-to compile the solver. In the main directory, type
-\begin{lyxcode}
-{\small make~xspecfem3D}{\small \par}
-\end{lyxcode}
-Please note that \texttt{xspecfem3D} must be called directly from
-the main directory, as most of the binaries of the package.\\
-
-\noindent
-The solver needs three input files in the \texttt{DATA} directory
-to run:
-\begin{description}
-\item [{\texttt{Par\_file}}] the main parameter file which was discussed
-in detail in the previous Chapter~\ref{cha:Creating-Distributed-Databases},
-\item [{\texttt{CMTSOLUTION}} or {\texttt{FORCESOLUTION}}] the earthquake
-source parameter file or the force source parameter file, and
-\item [{\texttt{STATIONS}}] the stations file.
-\end{description}
-Most parameters in the \texttt{Par\_file} should be set prior to running
-the databases generation. Only the following parameters may be changed
-after running \texttt{xgenerate\_databases}:
-\begin{itemize}
-\item the simulation type control parameters: \texttt{SIMULATION\_TYPE}
-and \texttt{SAVE\_FORWARD}
-\item the time step parameters \texttt{NSTEP} and \texttt{DT}
-\item the absorbing boundary control parameter \texttt{PML\_CONDITIONS}
-on condition that the\\
- \texttt{PML\_INSTEAD\_OF\_FREE\_SURFACE} flag remains unmodified
-after running the databases generation.
-\item the movie control parameters \texttt{MOVIE\_SURFACE}, \texttt{MOVIE\_VOLUME},
-and \texttt{NTSTEPS\_BETWEEN\_FRAMES}
-\item the ShakeMap\textregistered{}option \texttt{CREATE\_SHAKEMAP}
-\item the output information parameters \texttt{MOVIE\_TYPE}, \texttt{NTSTEP\_BETWEEN\_OUTPUT\_INFO}
-and \texttt{NTSTEP\_BETWEEN\_OUTPUT\_}~\\
- \texttt{SEISMOS}
-\item the \texttt{PRINT\_SOURCE\_TIME\_FUNCTION} flags
-\end{itemize}
-Any other change to the \texttt{Par\_file} implies rerunning both
-the database generator \texttt{xgenerate\_databases} and the solver
-\texttt{xspecfem3D}.
-
-For any particular earthquake, the \texttt{CMTSOLUTION} file that
-represents the point source may be obtained directly from the Harvard
-Centroid-Moment Tensor (CMT) web page \urlwithparentheses{www.seismology.harvard.edu}.
-It looks like the example shown in Fig. \ref{fig:CMTSOLUTION-file}.
-%\begin{lyxcode}
-\begin{figure}[htp]
-\noindent
-\begin{centering}
-{\small \includegraphics[width=0.8\textwidth]{figures/Hollywood_CMT}
-}
-\par
-\end{centering}
-\caption{\texttt{CMTSOLUTION} file based on the
-format from the Harvard CMT catalog. \textbf{M} is the moment tensor,
-$M_{0}${\small {} }is the seismic moment, and $M_{w}$ is the moment
-magnitude.}
-\label{fig:CMTSOLUTION-file}
-\end{figure}
-%\end{lyxcode}
-The \texttt{CMTSOLUTION} file should be edited in the following way:
-\begin{itemize}
-\item Set the latitude or UTM $x$ coordinate, longitude or UTM $y$ coordinate, depth of the source (in km).
-\item Set the \texttt{time shift} parameter equal to $0.0$ (the solver
-will not run otherwise.) The time shift parameter would simply apply
-an overall time shift to the synthetics, something that can be done
-in the post-processing (see Section \ref{sec:Process-data-and-syn}).
-\item For point-source simulations (see finite sources, page \pageref{To-simulate-a})
-we recommend setting the source half-duration parameter \texttt{half
-duration} equal to zero, which corresponds to simulating a step source-time
-function, i.e., a moment-rate function that is a delta function. If
-\texttt{half duration} is not set to zero, the code will use a Gaussian
-(i.e., a signal with a shape similar to a `smoothed triangle', as
-explained in \citet{KoTr02a} and shown in Fig~\ref{fig:gauss.vs.triangle})
-source-time function with half-width \texttt{half duration}. We prefer
-to run the solver with \texttt{half duration} set to zero and convolve
-the resulting synthetic seismograms in post-processing after the run,
-because this way it is easy to use a variety of source-time functions
-(see Section \ref{sec:Process-data-and-syn}). \citet{KoTr02a} determined
-that the noise generated in the simulation by using a step source
-time function may be safely filtered out afterward based upon a convolution
-with the desired source time function and/or low-pass filtering. Use
-the serial code \texttt{convolve\_source\_timefunction.f90} and the
-script \texttt{convolve\_source\_timefunction.csh} for this purpose,
-or alternatively use signal-processing software packages such as SAC
-\urlwithparentheses{www.llnl.gov/sac}. Type
-
-\begin{lyxcode}
-{\small make~xconvolve\_source\_timefunction}{\small \par}
-\end{lyxcode}
-
-to compile the code and then set the parameter \texttt{hdur} in \texttt{convolve\_source\_timefunction.csh}
-to the desired half-duration.
-
-\item The zero time of the simulation corresponds to the center of the triangle/Gaussian,
-or the centroid time of the earthquake. The start time of the simulation
-is $t=-1.5*\texttt{half duration}$ (the 1.5 is to make sure the moment
-rate function is very close to zero when starting the simulation).
-To convert to absolute time $t_{\mathrm{abs}}$, set
-
-\begin{lyxcode}
-$t_{\mathrm{abs}}=t_{\mathrm{pde}}+\texttt{time shift}+t_{\mathrm{synthetic}}$
-\end{lyxcode}
-
-where $t_{\mathrm{pde}}$ is the time given in the first line of the
-\texttt{CMTSOLUTION}, \texttt{time shift} is the corresponding value
-from the original \texttt{CMTSOLUTION} file and $t_{\mathrm{synthetic}}$
-is the time in the first column of the output seismogram.
-
-\end{itemize}
-\begin{figure}[htp]
-\noindent \begin{centering}
-\includegraphics[width=3.5in]{figures/gauss_vs_triangle_mod}
-\par\end{centering}
-
-\caption{Comparison of the shape of a triangle and the Gaussian function actually
-used.}
-
-\label{fig:gauss.vs.triangle}
-\end{figure}
-
-If you know the earthquake source in strike/dip/rake format rather
-than in \texttt{CMTSOLUTION} format, use the C code \texttt{SPECFEM3D\_GLOBE/utils/strike\_dip\_rake\_to\_CMTSOLUTION.c}
-to convert it. The conversion formulas are given for instance in \citet{AkRi80}.
-Note that the \citet{AkRi80} convention is slightly different from
-the Harvard \texttt{CMTSOLUTION} convention (the sign of some components
-is different). The C code outputs both.
-
-Centroid latitude and longitude should be provided in geographical
-coordinates. The code converts these coordinates to geocentric coordinates~\citep{DaTr98}.
-Of course you may provide your own source representations by designing
-your own \texttt{CMTSOLUTION} file. Just make sure that the resulting
-file adheres to the Harvard CMT conventions (see Appendix~\ref{cha:Coordinates}).
-Note that the first line in the \texttt{CMTSOLUTION} file is the Preliminary
-Determination of Earthquakes (PDE) solution performed by the USGS
-NEIC, which is used as a seed for the Harvard CMT inversion. The PDE
-solution is based upon P waves and often gives the hypocenter of the
-earthquake, i.e., the rupture initiation point, whereas the CMT solution
-gives the `centroid location', which is the location with dominant
-moment release. The PDE solution is not used by our software package
-but must be present anyway in the first line of the file.
-
-\label{To-simulate-a}To simulate a kinematic rupture, i.e., a finite-source
-event, represented in terms of $N_{\mathrm{sources}}$ point sources,
-provide a \texttt{CMTSOLUTION} file that has $N_{\mathrm{sources}}$
-entries, one for each subevent (i.e., concatenate $N_{\mathrm{sources}}$
-\texttt{CMTSOLUTION} files to a single \texttt{CMTSOLUTION} file).
-At least one entry (not necessarily the first) must have a zero \texttt{time
-shift}, and all the other entries must have non-negative \texttt{time
-shift}. Each subevent can have its own half duration, latitude, longitude,
-depth, and moment tensor (effectively, the local moment-density tensor).
-
-Note that the zero in the synthetics does NOT represent the hypocentral
-time or centroid time in general, but the timing of the \textit{center}
-of the source triangle with zero \texttt{time shift} (Fig~\ref{fig:source_timing}).
-
-Although it is convenient to think of each source as a triangle, in
-the simulation they are actually Gaussians (as they have better frequency
-characteristics). The relationship between the triangle and the gaussian
-used is shown in Fig~\ref{fig:gauss.vs.triangle}. For finite fault
-simulations it is usually not advisable to use a zero half duration
-and convolve afterwards, since the half duration is generally fixed
-by the finite fault model.
-
-\vspace{1cm}
-
-\noindent The \texttt{FORCESOLUTION} file should be edited in the
-following way:
-\begin{itemize}
-\item Set the \texttt{time shift} parameter equal to $0.0$ (the solver
-will not run otherwise.) The time shift parameter would simply apply
-an overall time shift to the synthetics, something that can be done
-in the post-processing (see Section \ref{sec:Process-data-and-syn}).
-\item Set the \texttt{f0} parameter (the dominant frequency) of the Ricker source time
-function when {\texttt{USE\_RICKER\_TIME\_FUNCTION}} is turned on
-in the main parameter file {\texttt{Par\_file}}. In case that the
-solver uses a (pseudo) Dirac delta source time function to represent
-a force point source, a very short duration of five time steps is automatically set by default.
-\item Set the latitude or UTM $x$ coordinate, longitude or UTM $y$ coordinate, depth of the source (in km).
-\item Set the magnitude of the force source.
-\item Set the components of a (non-unitary) direction vector for the force
-source in the East/North/Vertical basis (see Appendix A for the orientation
-of the reference frame).
-\end{itemize}
-\noindent Where necessary, set a \texttt{FORCESOLUTION} file in the
-same way you configure a \texttt{CMTSOLUTION} file with $N_{\mathrm{sources}}$
-entries, one for each subevent (i.e., concatenate $N_{\mathrm{sources}}$
-\texttt{FORCESOLUTION} files to a single \texttt{FORCESOLUTION} file).
-At least one entry (not necessarily the first) must have a zero \texttt{time
-shift}, and all the other entries must have non-negative \texttt{time
-shift}. Each subevent can have its own half latitude, longitude, depth,
-\texttt{half duration} and force parameters.
-
-\begin{figure}[H]
-\noindent \begin{centering}
-{\small \includegraphics[width=5in]{figures/source_timing} }
-\par\end{centering}
-
-\caption{Example of timing for three sources. The center of the first source
-triangle is defined to be time zero. Note that this is NOT in general
-the hypocentral time, or the start time of the source (marked as tstart).
-The parameter \texttt{time shift} in the \texttt{CMTSOLUTION} file
-would be t1(=0), t2, t3 in this case, and the parameter \texttt{half
-duration} would be hdur1, hdur2, hdur3 for the sources 1, 2, 3 respectively.}
-
-{\small \label{fig:source_timing} }
-\end{figure}
-
-The solver can calculate seismograms at any number of stations for
-basically the same numerical cost, so the user is encouraged to include
-as many stations as conceivably useful in the \texttt{STATIONS} file,
-which looks like this:
-
-\begin{figure}[H]
-\noindent \begin{centering}
-{\small \includegraphics{figures/STATIONS_basin_explained} }
-\par\end{centering}
-
-\label{fig:Sample-STATIONS-file.}\caption{Sample \texttt{STATIONS} file. Station latitude and longitude should
-be provided in geographical coordinates. The width of the station
-label should be no more than 32 characters (see \texttt{MAX\_LENGTH\_STATION\_NAME}
-in the \texttt{constants.h} file), and the network label should be
-no more than 8 characters (see \texttt{MAX\_LENGTH\_NETWORK\_NAME}
-in the \texttt{constants.h} file).}
-\end{figure}
-
-Each line represents one station in the following format:
-\begin{lyxcode}
-{\small Station~Network~Latitude~(degrees)~Longitude~(degrees)~Elevation~(m)~burial~(m)~}{\small \par}
-\end{lyxcode}
-The solver \texttt{xspecfem3D} filters the list of stations in file
-\texttt{DATA/STATIONS} to exclude stations that are not located within
-the region given in the \texttt{Par\_file} (between \texttt{LATITUDE\_MIN}
-and \texttt{LATITUDE\_MAX} and between \texttt{LONGITUDE\_MIN} and
-\texttt{LONGITUDE\_MAX}). The filtered file is called \texttt{DATA/STATIONS\_FILTERED}.
-
-Solver output is provided in the \texttt{OUTPUT\_FILES} directory
-in the \texttt{output\_solver.txt} file. Output can be directed to
-the screen instead by uncommenting a line in \texttt{constants.h}:
-\begin{lyxcode}
-!~uncomment~this~to~write~messages~to~the~screen~
-
-!~integer,~parameter~::~IMAIN~=~ISTANDARD\_OUTPUT~
-\end{lyxcode}
-On PC clusters the seismogram files are generally written to the local
-disks (the path \texttt{LOCAL\_PATH} in the \texttt{Par\_file}) and
-need to be gathered at the end of the simulation.
-
-While the solver is running, its progress may be tracked by monitoring
-the `\texttt{\small timestamp{*}}{\small ' files in the ~}\\
-{\small{} }\texttt{\small OUTPUT\_FILES}{\small{} directory. These tiny
-files look something like this:}{\small \par}
-\begin{lyxcode}
-Time~step~\#~~~~~~~~~~10000~
-
-Time:~~~~~108.4890~~~~~~seconds~
-
-Elapsed~time~in~seconds~=~~~~1153.28696703911~
-
-Elapsed~time~in~hh:mm:ss~=~~~~~0~h~19~m~13~s~
-
-Mean~elapsed~time~per~time~step~in~seconds~=~~~~~0.115328696703911~
-
-Max~norm~displacement~vector~U~in~all~slices~(m)~=~~~~1.0789589E-02~
-\end{lyxcode}
-The \texttt{\small timestamp{*}}{\small{} files provide the }\texttt{\small Mean
-elapsed time per time step in seconds}{\small , which may be used
-to assess performance on various machines (assuming you are the only
-user on a node), as well as the }\texttt{\small Max norm displacement
-vector U in all slices~(m)}{\small . If something is wrong with the
-model, the mesh, or the source, you will see the code become unstable
-through exponentially growing values of the displacement and fluid
-potential with time, and ultimately the run will be terminated by
-the program. You can control the rate at which the timestamp files
-are written based upon the parameter }\texttt{\small NTSTEP\_BETWEEN\_OUTPUT\_INFO}{\small{}
-in the }\texttt{\small Par\_file}{\small .}{\small \par}
-
-Having set the \texttt{Par\_file} parameters, and having provided
-the \texttt{CMTSOLUTION} (or the \texttt{FORCESOLUTION}) and \texttt{STATIONS}
-files, you are now ready to launch the solver! This is most easily
-accomplished based upon the \texttt{go\_solver} script (See Chapter
-\ref{cha:Scheduler} for information about running through a scheduler,
-e.g., LSF). You may need to edit the last command at the end of the
-script that invokes the \texttt{mpirun} command. The \texttt{runall}
-script compiles and runs both \texttt{xgenerate\_databases} and \texttt{xspecfem3D}
-in sequence. This is a safe approach that ensures using the correct
-combination of distributed database output and solver input.
-
-It is important to realize that the CPU and memory requirements of
-the solver are closely tied to choices about attenuation (\texttt{ATTENUATION})
-and the nature of the model (i.e., isotropic models are cheaper than
-anisotropic models). We encourage you to run a variety of simulations
-with various flags turned on or off to develop a sense for what is
-involved.
-
-For the same model, one can rerun the solver for different events
-by simply changing the \texttt{CMTSOLUTION} or \texttt{FORCESOLUTION}
-file, or for different stations by changing the \texttt{STATIONS}
-file. There is no need to rerun the \texttt{xgenerate\_databases}
-executable. Of course it is best to include as many stations as possible,
-since this does not add to the cost of the simulation.
-
-We have also added the ability to run several calculations (several earthquakes)
-in an embarrassingly-parallel fashion from within the same run;
-this can be useful when using a very large supercomputer to compute
-many earthquakes in a catalog, in which case it can be better from
-a batch job submission point of view to start fewer and much larger jobs,
-each of them computing several earthquakes in parallel.
-To turn that option on, set parameter \texttt{NUMBER\_OF\_SIMULTANEOUS\_RUNS}
-to a value greater than 1 in file setup/constants.h.in before
-configuring and compiling the code.
-When that option is on, of course the number of processor cores used to start
-the code in the batch system must be a multiple of \texttt{NUMBER\_OF\_SIMULTANEOUS\_RUNS},
-all the individual runs must use the same number of processor cores,
-which as usual is \texttt{NPROC} in the input file \texttt{DATA/Par\_file},
-and thus the total number of processor cores to request from the batch system
-should be \texttt{NUMBER\_OF\_SIMULTANEOUS\_RUNS $\times$ NPROC}.
-All the runs to perform must be placed in directories called \texttt{run0001}, \texttt{run0002}, \texttt{run0003} and so on (with exactly four digits)
-and you must create a link from the root directory of the code
-to the first copy of the executable programs by typing \texttt{ln -s run0001/bin bin}.
-
-\chapter{\label{cha:Fault-Kinematics-Dynamics}Kinematic and dynamic fault
-sources}
-
-SPECFEM3D can handle finite fault sources of two kinds:
-\begin{enumerate}
-\item \emph{Kinematic}: the spatio-temporal distribution of slip rate is
-prescribed along the fault surface
-\item \emph{Dynamic}: a friction law and initial stresses are prescribed
-on the fault, a spontaneous rupture process is computed.
-\end{enumerate}
-
-\section{\label{sec:Mesh-Generation-with}Mesh Generation with Split Nodes}
-
-Faults need to be handled in a special way during mesh generation.
-A fault surface must lie at the interface between elements (the mesh
-must honor the fault surfaces). Moreover, a fault is made of two surfaces
-in contact. Each of these two surfaces needs a separate set of nodes.
-This approach is known as \textquotedbl{}split nodes\textquotedbl{}.
-Currently faults can only be run with \texttt{xdecompose\_mesh} and
-\texttt{CUBIT}. \texttt{xmeshfem3D} is not yet ready to handle faults.\\
-To facilitate the mesh generation with split nodes in CUBIT, we need
-to separate the two fault surfaces by a small distance, effectively
-creating a tiny opening of the fault (Figure \ref{fig:examples.splitnodes},
-\ref{fig:examples.splitnodes-surfacetrace}). Note that only the interior
-of the fault must be opened, its edges must remain closed (except
-the edge on the free surface). The fault is automatically closed later
-by SPECFEM3D.\\
-
-Here is an example CUBIT script to generate a mesh with split nodes
-for a buried vertical strike-slip fault:
-\begin{lyxcode}
-~reset~~\\
-~brick~x~10~y~10~z~10~~\\
-~webcut~volume~all~with~plane~xplane~~\\
-~webcut~volume~all~with~plane~yplane~~\\
-~webcut~volume~all~with~plane~xplane~offset~3~~\\
-~webcut~volume~all~with~plane~zplane~offset~3~~\\
-~webcut~volume~all~with~plane~zplane~offset~-3~~\\
-~imprint~all~~\\
-~merge~all~~\\
-~unmerge~surf~160~~\\
-~mesh~vol~all~~\\
-~set~node~constraint~off~~\\
-~node~in~surf~168~move~X~0~Y~0.01~Z~0~~\\
-~node~in~surf~160~move~X~0~Y~-0.01~Z~0~\\
-
-\end{lyxcode}
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[scale=0.3]{figures/faultmesh} \\
- \includegraphics[scale=0.3]{figures/surf168} \includegraphics[scale=0.3]{figures/splitnodes}
-\par\end{centering}
-
-\caption{Screenshots of the CUBIT example for the embedded fault with split
-nodes: The entire mesh is decomposed into several volumes bounding
-the fault (top), the fault surface 168 is shown as orange squares
-(middle) and split nodes of the fault shown as orange dots(bottom).
-Note that only interior nodes of the fault are split while those on
-the edges of the fault surface are touching each other.}
-
-\label{fig:examples.splitnodes}
-\end{figure}
-
-\begin{lyxcode}
-
-\end{lyxcode}
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[scale=0.55]{figures/splitnodes_surfacetrace} \\
-
-\par\end{centering}
-
-\caption{Screenshots of a CUBIT example showing split nodes for a fault reaching
-the surface. Surface trace of the fault is shown in orange. Note that
-the edges of the fault are not split while the interior nodes are
-offset by a small distance on either side of the fault}
-
-\label{fig:examples.splitnodes-surfacetrace}
-\end{figure}
-
-The CUBIT scripts ({*}.jou and {*}.py) in the directory \texttt{EXAMPLES}
-generate more complicated meshes. The {*}.py files are Python scripts
-that execute CUBIT commands and use the CUBIT-python interface for
-SPECFEM3D (see next section). The Python language allows to define
-and manipulate variables to parameterize the mesh. Alternatively,
-the Python script can call a CUBIT journal file ({*}.jou), which looks
-like the example above. Variables can be defined and manipulated there
-using the $\mathtt{APREPRO}$ language built in CUBIT.\\
-
-Note that you should avoid gaps in the list of indices of mesh objects
-with the following CUBIT command:
-\begin{lyxcode}
-$\mathtt{compress\: ids\: hex\: face\: edge\: node}$
-\end{lyxcode}
-(otherwise you will get a segmentation fault during domain decomposition)
-
-\section{CUBIT-Python Scripts for Faults}
-
-The mesh generated in CUBIT needs to be processed and exported in
-a format compatible with SPECFEM3D. This is achieved in the Python
-scripts by calling the Python-CUBIT interface functions defined in
-the CUBIT directory:
-\begin{enumerate}
-\item Function $\mathtt{define\_bc}$ (or $\mathtt{boundary\_definition.py}$)
-must be called to set up the absorbing boundaries database
-\item Function $\mathtt{fault\_input}$ must be called once for each fault
-to set up the fault database
-\item Function $\mathtt{cubit2specfem3d.export2SPECFEM3D}$ must be called
-at the very end of the script to export the mesh in a SPECFEM3D format.
-\end{enumerate}
-The functions in \#1 and \#3 are for the bulk and are documented in
-Section \ref{subsec:Exporting-the-Mesh}. We focus here on \#2:
-\begin{description}
-\item [{Function:}] $\mathtt{fault\_input}$
-\item [{Purpose:}] export a fault mesh database from CUBIT to a SPECFEM3D-compliant
-file
-\item [{Syntax:}] fault\_input(id\_fault, ids\_surf\_1, ids\_surf\_2)
-\item [{Inputs:}]~\end{description}
-\begin{itemize}
-\item id\_fault integer index assigned to the fault. (The user must number
-all the faults, starting at 1, with unit increments)
-\item ids\_surf\_1 list of CUBIT ids of all surfaces that form side 1 of
-the fault.
-\item ids\_surf\_2 list of CUBIT ids of all surfaces that form side 2 of
-the fault. (The user must decide which side of the fault is side 1.
-This choice affects the sign conventions of fault quantities as explained
-in Section \ref{sec:Sign-Convention-for}).\end{itemize}
-\begin{description}
-\item [{Outputs:}] file $\mathtt{fault\_file\_X.dat}$, where X is the
-fault id (id\_fault).
-\item [{Example:}] For the example in Section \ref{sec:Mesh-Generation-with}:\end{description}
-\begin{lyxcode}
-$\hphantom{}$A1~=~{[}168{]}
-
-$\hphantom{}$A2~=~{[}160{]}
-
-$\hphantom{}$fault\_input(1,A1,A2)
-\end{lyxcode}
-
-\section{Examples}
-
-The package includes examples, the SCEC benchmark problems:
-\begin{itemize}
-\item TPV5, a planar vertical strike-slip fault
-\item TPV14 and TPV15, vertical strike-slip fault system with a fault branch
-\item TPV102 and TPV103, rate-and-state benchmarks
-\item TPV22 and TPV23, step-over benchmarks
-\item TPV16, heterogenous initial stress
-\end{itemize}
-and
-\begin{itemize}
-\item Splay fault models from Wendt et al. (2009)\\
-
-\end{itemize}
-Read the documents in the directory \texttt{EXAMPLES/*/description}.
-They contain a description of the example and additional instructions
-to run it. Visualize the results with the Matlab scripts in the directory
-$\mathtt{EXAMPLES/*/post}$
-
-\section{\label{sec:Sign-Convention-for}Sign Convention for Fault Quantities}
-
-During mesh generation, the fault is defined by two surfaces in contact.
-Let's denote as \textquotedbl{}\emph{side 1}\textquotedbl{} the SECOND
-surface declared by the user in the call to the python function \textquotedbl{}\emph{fault\_input}\textquotedbl{},
-and the FIRST surface as \textquotedbl{}\emph{side 2}\textquotedbl{}.
-The local coordinate system on the fault is defined as the right-handed
-coordinate system defined by (strike, dip, normal), where \textquotedbl{}\emph{normal}\textquotedbl{}
-is the normal vector outgoing from side 1, \textquotedbl{}\emph{dip}\textquotedbl{}
-is parallel to the along-dip direction pointing downwards, and \textquotedbl{}\emph{strike}\textquotedbl{}
-is the horizontal along-strike vector such that the system is right-handed.\\
-
-Slip is defined as displacement on side 2 minus displacement on side
-1. In the local coordinate system on the fault, positive along-strike
-slip is right-lateral and positive along-dip slip is thrust if side
-1 is on the hanging wall (normal faulting if side 1 is on the foot
-wall).\\
-
-Traction is defined as the stress induced on side 1 by side 2, which
-is the stress tensor times the normal vector outgoing from side 1.
-In the local coordinate system on the fault, the normal traction is
-negative in compression, positive along-strike traction generates
-right-lateral slip and positive along-dip traction generates thrust
-slip if side 1 is on the hanging wall (normal faulting if side 1 is
-on the foot wall).
-
-\section{Input Files}
-
-Three additional input files are required in the \texttt{DATA/} directory
-for dynamic and kinematic faults. They are $\mathtt{Par\_file\_faults}$,
-$\mathtt{FAULT\_STATIONS}$ and $\mathtt{input\_file.txt}$ (optional).
-If the former does not exist, the code assumes that there are no faults.\\
-
-\begin{lyxlist}{00.00.0000}
-\item [{\textbf{DATA/Par\_file\_faults}}] contains parameters of the fault.
-The first part of this file has a strict format:
-
-\begin{description}
-\item [{Line}] 1: Number of faults (NF)
-\item [{Lines}] 2 to NF+1: Kelvin Voigt damping (in seconds) for each fault.
-(See below how to set this parameter)
-\item [{Line}] NF+2: Type of simulation (1=dynamic , 2 = kinematic)
-\item [{Line}] NF+3: Number of time steps between updates of the time series
-outputs at selected fault points (see DATA/FAULT\_STATIONS), usually
-a large number (100s or 1000s). Note that the sampling rate of the
-time series is usually much higher.
-\item [{Line}] NF+4: Number of time steps between fault snapshot outputs
-(quantities at every fault point exported at regular times), usually
-a large number (100s or 1000s).
-\item [{Line}] NF+5: Slip velocity threshold below which frictional healing
-is set (friction coefficient is reset to its static value). If this
-value is negative healing is disabled.
-\item [{Line}] NF+6: Slip velocity threshold to define the rupture front.
-Only used for outputs.
-\end{description}
-
-The rest of this file is made of namelist input blocks (see \textquotedbl{}namelist\textquotedbl{}
-in a Fortran 9x manual). The input for each fault has the following
-sequence (arguments in {[}brackets{]} are optional):
-
-\&\textbf{BEGIN\_FAULT} /
-
-\&\textbf{INIT\_STRESS} S1, S2, S3 {[},n1, n2, n3{]} /
-
-followed by (n1+n2+n3) \&\textbf{DIST2D} blocks
-
-\&\textbf{SWF} mus, mud, dc {[}, nmus, nmud, ndc{]} /
-
-followed by (nmus+nmud+ndc) \&\textbf{DIST2D} blocks\\
-
-\begin{lyxlist}{00.00.0000}
-\item [{\&\textbf{INIT\_STRESS}}] input block sets the initial fault stresses
-relative to the foot-wall side of the fault. Initial stresses are
-composed of a constant background value possibly overwritten in prescribed
-regions by heterogeneous distributions (see \&\textbf{DIST2D} blocks
-below):
-
-\begin{description}
-\item [{S1}] = constant background value of along-strike shear stress (positive
-in the usual strike direction)
-\item [{S2}] = constant background value of along-dip shear (positive is
-down-dip, normal faulting)
-\item [{S3}] = constant background value of normal stress (negative in
-compression)
-\item [{n1}] = number of heterogeneous items for along-strike shear stress
-{[}default is 0{]}
-\item [{n2}] = number of heterogeneous items for along-dip shear stress
-{[}default is 0{]}
-\item [{n3}] = number of heterogeneous items for normal stress {[}default
-is 0{]}
-\end{description}
-\item [{\&\textbf{SWF}}] input block sets the slip-weakening friction parameters
-of the fault:
-
-\begin{description}
-\item [{mus}] = constant background value of static friction coefficient
-\item [{mud}] = constant background value of dynamic friction coefficient
-\item [{dc}] = constant background value of critical slip-weakening distance
-\item [{nmus}] = number of heterogeneous items for static friction coefficient
-{[}default is 0{]}
-\item [{nmud}] = number of heterogeneous items for dynamic friction coefficient
-{[}default is 0{]}
-\item [{ndc}] = number of heterogeneous items for critical slip-weakening
-distance {[}default is 0{]}\\
-
-\end{description}
-\item [{\&\textbf{DIST2D}}] input blocks modify (overwrite) the value of
-a fault parameter by a heterogeneous spatial distribution:
-
-\&\textbf{DIST2D} shape='$\mathtt{square}$', val, xc, yc, zc, l /
-
-$\;$sets a constant value (val) within a cube with center (xc,yc,zc)
-and edge size l.
-
-\&\textbf{DIST2D} shape='$\mathtt{rectangle}$', val, xc, yc, zc,
-lx, ly, lz /
-
-$\;$sets a constant value (val) within a rectangular prism with center
-(xc,yc,zc) and edge sizes (lx,ly,lz).
-
-\&\textbf{DIST2D} shape='$\mathtt{rectangle-taper}$', val, valh,
-xc, yc, zc, lx, ly, lz /
-
-$\;$sets a vertical linear gradient within a rectangular prism with
-center (xc,yc,zc) and edge sizes (lx,ly,lz). Values vary linearly
-as a function of vertical position z between value val at z = zc-lz/2
-and value valh at z = zc+lz/2 .
-
-\&\textbf{DIST2D} shape='$\mathtt{circular}$', val, xc, yc, zc, r
-/
-
-$\;$sets a constant value (val) within a sphere with center (xc,yc,zc)
-and radius r. \\
-
-\end{lyxlist}
-\end{lyxlist}
-
-\begin{lyxlist}{00.00.0000}
-\item [{\textbf{DATA/FAULT\_STATIONS}}] Stations in the fault plane.
-
-\textbf{Line 1} : number of stations.
-
-\textbf{Line 2 to end}: 5 columns: X, Y, Z (-depth), station name,
-fault-id
-
-The fault-id identifies the fault that contains the station. It is
-the index of appearance in the faults list after line 2 of Par\_file\_faults
-
-\end{lyxlist}
-
-\begin{lyxlist}{00.00.0000}
-\item [{\textbf{DATA/input\_file.txt}}] Heterogeneous stresses and friction
-parameters are documented in page 10 of
-
-\begin{lyxlist}{00.00.0000}
-\item [{$\mathtt{EXAMPLES/tpv16/description/TPV16\_17\_Description\_v03.pdf}$}]~
-\end{lyxlist}
-
-To activate this feature, in fault\_solver\_dynamic.f90 set \textbf{TPV16}
-= .true. then re-compile the code.
-
-\end{lyxlist}
-
-\section{\label{sec:Setting-the-Kelvin-Voigt}Setting the Kelvin-Voigt Damping
-Parameter }
-
-The purpose of the Kelvin-Voigt viscosity in the dynamic fault solver
-is to damp spurious oscillations generated by the fault slip at frequencies
-that are too high to be resolved by the mesh. The viscosity $\mathtt{eta}$
-(in seconds) depends on the size of the elements on the fault. Here
-is how to set it:
-\begin{enumerate}
-\item Determine the average linear size of the elements on the fault plane,
-$\mathtt{h\_fault}$. Usually this value is prescribed by the user
-during mesh generation. Otherwise it can be found by inspection of
-the mesh inside the CUBIT GUI.
-\item Use the Matlab function $\mathtt{utils/critical\_timestep.m}$ to
-compute\\
- $\mathtt{dtc\_fault}=\mathtt{critical\_timestep\left(c_{p},h\_fault,ngll\right)}$.
-\\
-This is the critical time step in an elastic medium for a hypothetical
-element of cubic shape with size equal to $\mathtt{h\_fault}$.
-\item Set $\mathtt{eta}$ in $\mathtt{Par\_file\_faults}$ to (0.1 to 0.3)$\ensuremath{\times}\mathtt{dtc\_fault}$.
-A larger $\mathtt{eta}$ damps high-frequencies more aggressively
-but it might also affect lower frequencies and rupture speed.
-\end{enumerate}
-Viscosity reduces numerical stability: the critical timestep in a
-simulation with Kelvin-Voigt damping needs to be smaller than that
-in a purely elastic simulation. Here is how to set the time step accordingly:
-\begin{enumerate}
-\item Run a test simulation without viscosity ($\mathtt{eta}$=0 and only
-a few time steps)
-\item Look for the \textquotedbl{}maximum suggested time step\textquotedbl{}
-in \texttt{OUTPUT\_FILES/output\_mesher.txt}. This is the critical
-timestep of a purely elastic simulation, $\mathtt{dtc\_bulk}$.
-\item Reset the timestep of the simulation with a Kelvin-Voigt material
-to a value smaller than\\
- $\mathtt{dtc\_kv}=\mathtt{eta}\left(\sqrt{1+\mathtt{dtc\_bulk^{2}}/\mathtt{eta^{2}}}-1\right)$
-\end{enumerate}
-Note that in general $\mathtt{dtc\_bulk}$ is smaller than $\mathtt{dtc\_fault}$,
-because elements off the fault might be smaller or more distorted
-than element faces on the fault.
-
-\section{Output Files }
-
-Several output files are saved in \texttt{OUTPUT\_FILES/}:\\
-
-\begin{enumerate}
-\item Seismograms for each station on the fault plane given in \texttt{DATA/FAUL\_STATIONS}.
-One output file is generated for each station, named after the station.
-The files are ascii and start with a header (22 lines long) followed
-by a data block with the following format, one line per time sample:\\
-
-Column 1 = Time (s)
-
-Column 2 = horizontal right-lateral slip (m)
-
-Column 3 = horizontal right-lateral slip rate (m/s)
-
-Column 4 = horizontal right-lateral shear stress (MPa)
-
-Column 5 = vertical up-dip slip (m)
-
-Column 6 = vertical up-dip slip rate (m/s)
-
-Column 7 = vertical up-dip shear stress (MPa)
-
-Column 8 = normal stress (MPa)\\
-
-The stresses are relative to the footwall side of the fault (this
-convention controls their sign, but not their amplitude). Slip is
-defined as displacement of the hanging wall relative to the footwall.
-
-\item Seismograms at stations in the bulk (out of the fault plane) given
-in \texttt{DATA/STATIONS}.
-\item Rupture time files are named $\mathtt{Rupture\_time\_FAULT}$-$\mathtt{id}$.
-One file is generated for each fault. The files are ascii and start
-with a header (12 lines long) followed by a data block with the following
-format, one line per fault node:
-
-Column 1 = horizontal coordinate, distance along strike (m)
-
-Column 2 = vertical coordinate, distance down-dip (m)
-
-Column 3 = rupture time (s)
-
-\item Fault quantities (slip, slip rate, stresses, etc) at regular times
-are stored in binary data files called $\mathtt{Snapshot\#it\#.bin}$,
-where \#it\# is the timestep number. These can be read in Matlab with
-the function $\mathtt{utils/FSEM3D\_snapshot.m}$
-\end{enumerate}
-
-\section{Post-processing and Visualization }
-
-Some Matlab functions for post-processing and visualization are included
-in directory $\mathtt{utils}$. The functions are internally documented
-(see their matlab help).\\
-
-\texttt{FSEM3D\_snapshot} reads a fault data snapshot\\
-
-The directories \texttt{EXAMPLES/*/post} contain additional Matlab
-scripts to generate figures specific to each example.
-
-\chapter{\label{cha:Adjoint-Simulations}Adjoint Simulations}
-
-Adjoint simulations are generally performed for two distinct applications.
-First, they can be used in point source moment-tensor inversions, or source imaging for
-earthquakes with large ruptures such as the Lander's earthquake \citep{WaHe94}.
-Second, they can be used to generate finite-frequency sensitivity
-kernels that are a critical part of tomographic inversions based upon
-3D reference models \citep{TrTaLi05,LiTr06,TrKoLi08,LiTr08}. In either
-case, source parameter or velocity structure updates are sought to
-minimize a specific misfit function (e.g., waveform or traveltime
-differences), and the adjoint simulation provides a means of computing
-the gradient of the misfit function and further reducing it in successive
-iterations. Applications and procedures pertaining to source studies
-and finite-frequency kernels are discussed in Sections~\ref{sec:Adjoint-simulation-sources}
-and \ref{sec:Adjoint-simulation-finite}, respectively. The two related
-parameters in the \texttt{Par\_file} are \texttt{SIMULATION\_TYPE}
-(1, 2 or 3) and the \texttt{SAVE\_FORWARD} (boolean).
-
-\section{\label{sec:Adjoint-simulation-sources}Adjoint Simulations for Sources
-Only (not for the Model)}
-
-When a specific misfit function between data and synthetics is minimized to invert
-for earthquake source parameters, the gradient of the misfit function
-with respect to these source parameters can be computed by placing
-time-reversed seismograms at the receivers as virtual sources
-in an adjoint simulation. Then the value of the gradient is obtained
-from the adjoint seismograms recorded at the original earthquake location.
-\begin{enumerate}
-\item \textbf{Prepare the adjoint sources} \label{enu:Prepare-the-adjoint}
-
-\begin{enumerate}
-\item First, run a regular forward simlation (\texttt{SIMULATION\_TYPE =
-1} and \texttt{SAVE\_FORWARD = .false.}). You can automatically set
-these two variables using the \texttt{\small utils/change\_simulation\_type.pl}{\small{}
-script:}{\small \par}
-
-\begin{lyxcode}
-utils/change\_simulation\_type.pl~-f~
-\end{lyxcode}
-
-and then collect the recorded seismograms at all the stations given
-in \texttt{DATA/STATIONS}.
-
-\item Then select the stations for which you want to compute the time-reversed
-adjoint sources and run the adjoint simulation, and compile them into
-the \texttt{DATA/STATIONS\_ADJOINT} file, which has the same format
-as the regular \texttt{DATA/STATIONS} file.
-
-\begin{itemize}
-\item Depending on what type of misfit function is used for the source inversion,
-adjoint sources need to be computed from the original recorded seismograms
-for the selected stations and saved in a sub-directory called \texttt{SEM/}
-in the root directory of the code, with the format \texttt{NT.STA.BX?.adj}, where \texttt{NT},
-\texttt{STA} are the network code and station name given in the \texttt{DATA/STATIONS\_ADJOINT}
-file, and \texttt{BX?} represents the component name of a particular
-adjoint seismogram. Please note that the band code can change depending
-on your sampling rate (see Appendix~\ref{cha:channel-codes} for
-further details).
-\item The adjoint seismograms are in the same format as the original seismogram
-(\texttt{NT.STA.BX?.sem?}), with the same start time, time interval
-and record length.
-\end{itemize}
-\item Notice that even if you choose to time reverse only one component
-from one specific station, you still need to supply all three components
-because the code is expecting them (you can set the other two components
-to be zero).
-\item Also note that since time-reversal is done in the code itself, no
-explicit time-reversing is needed for the preparation of the adjoint
-sources, i.e., the adjoint sources are in the same forward time sense
-as the original recorded seismograms.
-\end{enumerate}
-\item \textbf{Set the related parameters and run the adjoint simulation}\\
- In the \texttt{DATA/Par\_file}, set the two related parameters to
-be \texttt{SIMULATION\_TYPE = 2} and \texttt{SAVE\_FORWARD = .false.}.
-More conveniently, use the scripts \texttt{utils/change\_simulation\_type.pl}
-to modify the \texttt{Par\_file} automatically (\texttt{change\_simulation\_type.pl
--a}). Then run the solver to launch the adjoint simulation.
-\item \textbf{Collect the seismograms at the original source location}
-
-After the adjoint simulation has completed successfully, collect the
-seismograms from \texttt{LOCAL\_PATH}.
-\begin{itemize}
-\item These adjoint seismograms are recorded at the locations of the original
-earthquake sources given by the \texttt{DATA/CMTSOLUTION} file, and
-have names of the form \texttt{NT.S?????.S??.sem} for the six-component
-strain tensor (\texttt{SNN,SEE,SZZ,SNE,SNZ,SEZ}) at these locations,
-and ~\\
- \texttt{NT.S?????.BX?.sem} for the three-component displacements
-(\texttt{BXN,BXE,BXZ}) recorded at these locations.
-\item \texttt{S?????} denotes the source number; for example, if the original
-\texttt{CMTSOLUTION} provides only a point source, then the seismograms
-collected will start with \texttt{S00001}.
-\item These adjoint seismograms provide critical information for the computation
-of the gradient of the misfit function.
-\end{itemize}
-\end{enumerate}
-
-\section{\label{sec:Adjoint-simulation-finite}Adjoint Simulations for Finite-Frequency
-Kernels (Kernel Simulation)}
-
-Finite-frequency sensitivity kernels are computed in two successive
-simulations (please refer to \citet{LiTr06} and \citet{TrKoLi08}
-for details).
-\begin{enumerate}
-\item \textbf{Run a forward simulation with the state variables saved at
-the end of the simulation}
-
-Prepare the \texttt{\small CMTSOLUTION}{\small{} and }\texttt{\small STATIONS}{\small{}
-files, set the parameters }\texttt{\small SIMULATION\_TYPE}{\small {}
-}\texttt{\small =}{\small {} }\texttt{\small 1}{\small{} and }\texttt{\small SAVE\_FORWARD
-=}{\small {} }\texttt{\small .true.}{\small{} in the }\texttt{\small Par\_file}{\small{}
-(}\texttt{\small change\_simulation\_type -F}{\small ), and run the
-solver.}{\small \par}
-\begin{itemize}
-\item Notice that attenuation is not implemented yet for the computation
-of finite-frequency kernels; therefore set \texttt{ATTENUATION = .false.}
-in the \texttt{Par\_file}.
-\item We also suggest you modify the half duration of the \texttt{CMTSOLUTION}
-to be similar to the accuracy of the simulation (see Equation \ref{eq:shortest_period})
-to avoid too much high-frequency noise in the forward wavefield, although
-theoretically the high-frequency noise should be eliminated when convolved
-with an adjoint wavefield with the proper frequency content.
-\item This forward simulation differs from the regular simulations (\texttt{\small SIMULATION\_TYPE}{\small {}
-}\texttt{\small =}{\small {} }\texttt{\small 1}{\small{} and }\texttt{\small SAVE\_FORWARD}{\small {}
-}\texttt{\small =}{\small {} }\texttt{\small .false.}{\small ) described
-in the previous chapters in that the state variables for the last
-time step of the simulation, including wavefields of the displacement,
-velocity, acceleration, etc., are saved to the }\texttt{\small LOCAL\_PATH}{\small{}
-to be used for the subsequent simulation. }{\small \par}
-\item For regional simulations, the files recording the absorbing boundary
-contribution are also written to the \texttt{LOCAL\_PATH} when \texttt{SAVE\_FORWARD
-= .true.}.
-\end{itemize}
-\item \textbf{Prepare the adjoint sources}
-
-The adjoint sources need to be prepared the same way as described
-in the Section \ref{enu:Prepare-the-adjoint}.
-\begin{itemize}
-\item In the case of travel-time finite-frequency kernel for one source-receiver
-pair, i.e., point source from the \texttt{CMTSOLUTION}, and one station
-in the \texttt{STATIONS\_ADJOINT} list, we supply a sample program
-in \texttt{utils/adjoint\_sources/traveltime/xcreate\_adjsrc\_traveltime}
-to cut a certain portion of the original displacement seismograms
-and convert it into the proper adjoint source to compute the finite-frequency
-kernel.
-
-\begin{lyxcode}
-xcreate\_adjsrc\_traveltime~t1~t2~ifile{[}0-5{]}~E/N/Z-ascii-files~{[}baz{]}
-\end{lyxcode}
-
-where \texttt{t1} and \texttt{t2} are the start and end time of the
-portion you are interested in, \texttt{ifile} denotes the component
-of the seismograms to be used (0 for all three components, 1 for East,
-2 for North, and 3 for vertical, 4 for transverse, and 5 for radial
-component), \texttt{E/N/Z-ascii-files} indicate the three-component
-displacement seismograms in the right order, and \texttt{baz} is the
-back-azimuth of the station. Note that \texttt{baz} is only supplied
-when \texttt{ifile} = 4 or 5.
-
-\item Similarly, a sample program to compute adjoint sources for amplitude
-finite-frequency kernels may be found in \texttt{utils/adjoint\_sources/amplitude}
-and used in the same way as described for traveltime measurements
-
-\begin{lyxcode}
-xcreate\_adjsrc\_amplitude~t1~t2~ifile{[}0-5{]}~E/N/Z-ascii-files~{[}baz{]}.
-\end{lyxcode}
-\end{itemize}
-\item \textbf{Run the kernel simulation}
-
-With the successful forward simulation and the adjoint source ready
-in the \texttt{SEM/} directory, set \texttt{SIMULATION\_TYPE
-= 3} and \texttt{SAVE\_FORWARD = .false.} in the \texttt{Par\_file}
-(you can use \texttt{change\_simulation\_type.pl -b}), and rerun the
-solver.
-\begin{itemize}
-\item The adjoint simulation is launched together with the back reconstruction
-of the original forward wavefield from the state variables saved from
-the previous forward simulation, and the finite-frequency kernels
-are computed by the interaction of the reconstructed forward wavefield
-and the adjoint wavefield.
-\item The back-reconstructed seismograms at the original station locations
-are saved to the \texttt{LOCAL\_PATH} at the end of the kernel simulations,
-and can be collected to the local disk.
-\item These back-constructed seismograms can be compared with the time-reversed
-original seismograms to assess the accuracy of the backward reconstruction,
-and they should match very well.
-\item The arrays for density, P-wave speed and S-wave speed kernels are
-also saved in the \texttt{LOCAL\_PATH} with the names \texttt{proc??????\_rho(alpha,beta)\_kernel.bin},
-where \texttt{proc??????} represents the processor number, \texttt{rho(alpha,beta)}
-are the different types of kernels.
-\end{itemize}
-\item \textbf{Run the anisotropic kernel simulation}
-
-Instead of the kernels for the isotropic wave speeds, you can also
-compute the kernels for the 21 independent components $C_{IJ},\, I,J=1,...,6$
-(using Voigt's notation) of the elastic tensor in the cartesian coordinate
-system. This is done by setting \texttt{ANISOTROPIC\_KL} \texttt{=}
-\texttt{.true.} in \texttt{constants.h} before compiling the package.
-The definition of the parameters $C_{IJ}$ in terms of the corresponding
-components $c_{ijkl},ijkl,i,j,k,l=1,2,3$ of the elastic tensor in
-cartesian coordinates follows \citet{ChTr07}. The 21 anisotropic
-kernels are saved in the \texttt{LOCAL\_PATH} in one file with the
-name of \texttt{proc??????\_cijkl\_kernel.bin} (with \texttt{proc??????}
-the processor number). The output kernels correspond to absolute perturbations
-$\delta C_{IJ}$ of the elastic parameters and their unit is in $s/GPa/km^{3}$.
-For consistency, the output density kernels with this option turned
-on are for a perturbation $\delta\rho$ (and not $\frac{\delta\rho}{\rho}$)
-and their unit is in s / (kg/m$^{3}$) / km$^{3}$. These `primary'
-anisotropic kernels can then be combined to obtain the kernels for
-different parameterizations of anisotropy. This can be done, for example,
-when combining the kernel files from slices into one mesh file (see
-Section~\ref{sec:Finite-Frequency-Kernels}).
-
-If \texttt{ANISOTROPIC\_KL} \texttt{=} \texttt{.true.} by additionally
-setting \texttt{ANISOTROPIC\_KL} \texttt{=} \texttt{.true.} in \texttt{constants.h}
-the package will save anisotropic kernels paramterized as velocities
-related to transverse isotropy based on the the Chen and Tromp parameters
-\citet{ChTr07}. The kernels are saved as relative perturbations for
-horizontal and vertical P and S velocities, $\alpha_{v},\alpha_{h},\beta_{v},\beta_{h}$.
-Explicit relations can be found in appendix B. of \citet{SiLiTrTr07b}
-
-The anisotropic kernels are only currently available for CPU mode.
-
-\end{enumerate}
-In general, the three steps need to be run sequentially to assure
-proper access to the necessary files. If the simulations are run through
-some cluster scheduling system (e.g., LSF), and the forward simulation
-and the subsequent kernel simulations cannot be assigned to the same
-set of computer nodes, the kernel simulation will not be able to access
-the database files saved by the forward simulation. Solutions for
-this dilemma are provided in Chapter~\ref{cha:Scheduler}. Visualization
-of the finite-frequency kernels is discussed in Section~\ref{sec:Finite-Frequency-Kernels}.
-
-%------------------------------------------------------------------------------------------------%
-
-\chapter{\label{cha:tomo}Doing tomography, i.e., updating the model based on the sensitivity kernels obtained}
-
-%------------------------------------------------------------------------------------------------%
-
-UNDER CONSTRUCTION (March 2015). Will be added around April / May 2015.
-
-\section{Tomographic inversion / imaging using the sensitivity kernels obtained}
-
-OLD VERSION OF THE SECTION, WILL SOON BE IMPROVED AND ENRICHED.\\
-
-One of the fundamental reasons for computing sensitivity kernels (Section~\ref{sec:Adjoint-simulation-finite})
-is to use them within a tomographic inversion. In other words, use
-recorded seismograms, make measurements with synthetic seismograms,
-and use the misfit between the two to iteratively improve the model
-described by (at least) $V_{{\rm p}}$, $V_{{\rm s}}$, and $\rho$
-as a function of space.
-
-Whatever misfit function you use for the tomographic inversion (you can find several
-examples in \citet{TrKoLi08} and \citet{TrTaLi05}), you will weight
-the sensitivity kernels with measurements. Furthermore, you will use
-as many measurements (stations, components, time windows) as possible
-per event; hence, we call these composite kernels ``event kernels,''
-which are volumetric fields representing the gradient of the misfit
-function with respect to one of the variables (\eg $V_{{\rm s}}$).
-The basic features of an adjoint-based tomographic inversion were
-illustrated in \citet{TaLiTr07} and \citet{TrKoLi08} using a conjugate-gradient
-algorithm; there are dozens of versions of gradient-based inversion
-algorithms that could alternatively be used. The tomographic inversion
-of \citet{TaLiMaTr09,TaLiMaTr2010} used SPECFEM3D Cartesian as well
-as several additional components which are also stored on the CIG
-svn server, described next.
-
-The directory containing external utilities for tomographic inversion using
-SPECFEM3D Cartesian (or other packages that evaluate misfit functions
-and gradients) is in directory \texttt{utils/ADJOINT\_TOMOGRAPHY\_TOOLS/}:
-\begin{verbatim}
-flexwin/     -- FLEXWIN algorithm for automated picking of time windows
-measure_adj/ -- reads FLEXWIN output file and makes measurements,
-                with the option for computing adjoint sources
-iterate_adj/ -- various tools for iterative inversion
-                (requires pre-computed "event kernels")
-\end{verbatim}
-This directory also contains
-a brief \verb+README+ file indicating the role of the three subdirectories,
-\verb+flexwin+ \citep{Maggi2009}, \verb+measure_adj+, and \verb+iterate_adj+.
-The components for making the model update are there; however, there
-are no explicit rules for computing the model update, just as with
-any optimization problem. There are options for computing a conjugate
-gradient step, as well as a source subspace projection step.
-
-The best single file to read is probably: \verb+ADJOINT_TOMO/iterate_adj/cluster/README+.
-
-%------------------------------------------------------------------------------------------------%
-
-\chapter{Noise Cross-correlation Simulations}
-
-%------------------------------------------------------------------------------------------------%
-
-Besides earthquake simulations, SPECFEM3D Cartesian includes functionality
-for seismic noise tomography as well. In order to proceed successfully
-in this chapter, it is critical that you have already familiarized
-yourself with procedures for meshing (Chapter \ref{cha:Mesh-Generation}),
-creating distributed databases (Chapter \ref{cha:Creating-Distributed-Databases}),
-running earthquake simulations (Chapters \ref{cha:Running-the-Solver})
-and adjoint simulations (Chapter \ref{cha:Adjoint-Simulations}).
-Also, make sure you read the article `Noise cross-correlation sensitivity
-kernels' \citep{trompetal2010}, in order to understand noise simulations
-from a theoretical perspective.
-
-\section{Input Parameter Files}
-
-As usual, the three main input files are crucial: \texttt{Par\_file},
-\texttt{CMTSOLUTION} and \texttt{STATIONS}. Unless otherwise specified,
-those input files should be located in directory \texttt{DATA/}.\\
-
-\texttt{CMTSOLUTION} is required for all simulations. At a first glance,
-it may seem unexpected to have it here, since the noise simulations
-should have nothing to do with the earthquake -- \texttt{CMTSOLUTION}.
-However, for noise simulations, it is critical to have no earthquakes.
-In other words, the moment tensor specified in \texttt{CMTSOLUTION}
-must be set to zero manually!\\
-
-\texttt{STATIONS} remains the same as in previous earthquake simulations,
-except that the order of receivers listed in \texttt{STATIONS} is
-now important. The order will be used to determine the `master' receiver,
-i.e., the one that simultaneously cross correlates with the others.\\
-
-\texttt{Par\_file} also requires careful attention. A parameter called
-\texttt{NOISE\_TOMOGRAPHY} has been added which specifies the type
-of simulation to be run. \texttt{NOISE\_TOMOGRAPHY} is an integer
-with possible values 0, 1, 2 and 3. For example, when \texttt{NOISE\_TOMOGRAPHY}
-equals 0, a regular earthquake simulation will be run. When it is
-1/2/3, you are about to run step 1/2/3 of the noise simulations respectively.
-Should you be confused by the three steps, refer to \citet{trompetal2010}
-for details.\\
-
-Another change to \texttt{Par\_file} involves the parameter \texttt{NSTEP}.
-While for regular earthquake simulations this parameter specifies
-the length of synthetic seismograms generated, for noise simulations
-it specifies the length of the seismograms used to compute cross correlations.
-The actual cross correlations are thus twice this length, i.e., $2~\mathrm{NSTEP}-1$.
-The code automatically makes the modification accordingly, if \texttt{NOISE\_TOMOGRAPHY}
-is not zero.\\
-
-There are other parameters in \texttt{Par\_file} which should be given
-specific values. For instance, since the first two steps for calculating
-noise sensitivity kernels correspond to forward simulations, \texttt{SIMULATION\_TYPE}
-must be 1 when \texttt{NOISE\_TOMOGRAPHY} equals 1 or 2. Also, we
-have to reconstruct the ensemble forward wavefields in adjoint simulations,
-therefore we need to set \texttt{SAVE\_FORWARD} to \texttt{.true.}
-for the second step, i.e., when \texttt{NOISE\_TOMOGRAPHY} equals
-2. The third step is for kernel constructions. Hence \texttt{SIMULATION\_TYPE}
-should be 3, whereas \texttt{SAVE\_FORWARD} must be \texttt{.false.}.\\
-
-Finally, for most system architectures, please make sure that \texttt{LOCAL\_PATH}
-in \texttt{Par\_file} is in fact local, not globally shared. Because
-we have to save the wavefields at the earth's surface at every time
-step, it is quite problematic to have a globally shared \texttt{LOCAL\_PATH},
-in terms of both disk storage and I/O speed.
-
-\section{Noise Simulations: Step by Step}
-
-Proper parameters in those parameter files are not enough for noise
-simulations to run. We have more parameters to specify: for example,
-the ensemble-averaged noise spectrum, the noise distribution etc.
-However, since there are a few `new' files, it is better to introduce
-them sequentially. In this section, standard procedures for noise
-simulations are described.
-
-\subsection{Pre-simulation}
-\begin{itemize}
-\item As usual, we first configure the software package using:
-
-\texttt{./configure FC=ifort MPIFC=mpif90}
-
-Use the following if SCOTCH is needed:
-
-\texttt{./configure FC=ifort MPIFC=mpif90 -{}-with-scotch-dir=/opt/scotch}\\
-
-\item Next, we need to compile the source code using:
-
-\texttt{make xgenerate\_databases}
-
-\texttt{make xspecfem3D} \\
-
-\item Before we can run noise simulations, we have to specify the noise
-statistics, e.g., the ensemble-averaged noise spectrum. Matlab scripts
-are provided to help you to generate the necessary file.
-
-\texttt{EXAMPLES/noise\_tomography/NOISE\_TOMOGRAPHY.m} (main program)\\
- \texttt{EXAMPLES/noise\_tomography/PetersonNoiseModel.m}
-
-In Matlab, simply run:
-
-\texttt{NOISE\_TOMOGRAPHY(NSTEP, DT, Tmin, Tmax, NOISE\_MODEL)}\\
-
-\texttt{DT} is given in \texttt{Par\_file}, but \texttt{NSTEP} is
-NOT the one specified in \texttt{Par\_file}. Instead, you have to
-feed $2~\mathrm{NSTEP}-1$ to account for the doubled length of cross
-correlations. \texttt{Tmin} and \texttt{Tmax} correspond to the period
-range you are interested in, whereas \texttt{NOISE\_MODEL} denotes
-the noise model you will be using. Details can be found in the Matlab
-script.
-
-After running the Matlab script, you will be given the following information
-(plus a figure in Matlab):
-
-\texttt{{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}}\\
- \texttt{the source time function has been saved in:}\\
- \texttt{/data2/yangl/3D\_NOISE/S\_squared} (note this path must be
-different)\\
- \texttt{S\_squared should be put into directory:}\\
- \texttt{OUTPUT\_FILES/NOISE\_TOMOGRAPHY/ in the SPECFEM3D Cartesian
-package}\\
-
-In other words, the Matlab script creates a file called \texttt{S\_squared},
-which is the first `new' input file we encounter for noise simulations.
-
-One may choose a flat noise spectrum rather than Peterson's noise
-model. This can be done easily by modifying the Matlab script a little.
-
-\item Create a new directory in the SPECFEM3D Cartesian package, name it
-as \texttt{OUTPUT\_FILES/NOISE\_TOMOGRAPHY/}. We will add some parameter
-files later in this folder.
-\item Put the Matlab-generated-file \texttt{S\_squared} in \texttt{OUTPUT\_FILES/NOISE\_TOMOGRAPHY/}.
-
-That's to say, you will have a file \texttt{OUTPUT\_FILES/NOISE\_TOMOGRAPHY/S\_squared}
-in the SPECFEM3D Cartesian package.
-
-\item Create a file called \texttt{OUTPUT\_FILES/NOISE\_TOMOGRAPHY/irec\_master\_noise}.
-Note that this file is located in directory \texttt{OUTPUT\_FILES/NOISE\_TOMOGRAPHY/}
-as well. In general, all noise simulation related parameter files
-go into that directory. \texttt{irec\_master\_noise} contains only
-one integer, which is the ID of the `master' receiver. For example,
-if this file contains 5, it means that the fifth receiver listed in
-\texttt{DATA/STATIONS} becomes the `master'. That's why we mentioned
-previously that the order of receivers in \texttt{DATA/STATIONS} is
-important.
-
-Note that in regional simulations, the \texttt{DATA/STATIONS} might
-contain receivers which are outside of our computational domains.
-Therefore, the integer in \texttt{irec\_master\_noise} is actually
-the ID in \texttt{DATA/STATIONS\_FILTERED} (which is generated by
-\texttt{bin/xgenerate\_databases}).
-
-\item Create a file called \texttt{OUTPUT\_FILES/NOISE\_TOMOGRAPHY/nu\_master}.
-This file holds three numbers, forming a (unit) vector. It describes
-which component we are cross-correlating at the `master' receiver,
-i.e., ${\hat{{\bf \nu}}}^{\alpha}$ in \citet{trompetal2010}. The
-three numbers correspond to E/N/Z components respectively. Most often,
-the vertical component is used, and in those cases the three numbers
-should be 0, 0 and 1.
-\item Describe the noise direction and distributions in \texttt{src/shared/noise\_tomography.f90}.
-Search for a subroutine called \texttt{noise\_distribution\_direction}
-in \texttt{noise\_tomography.f90}. It is actually located at the very
-beginning of \texttt{noise\_tomography.f90}. The default assumes vertical
-noise and a uniform distribution across the whole free surface of
-the model. It should be quite self-explanatory for modifications.
-Should you modify this part, you have to re-compile the source code.
-\end{itemize}
-
-\subsection{Simulations}
-
-With all of the above done, we can finally launch our simulations.
-Again, please make sure that the \texttt{LOCAL\_PATH} in \texttt{DATA/Par\_file}
-is not globally shared. It is quite problematic to have a globally
-shared \texttt{LOCAL\_PATH}, in terms of both disk storage and speed
-of I/O (we have to save the wavefields at the earth's surface at every
-time step).
-
-As discussed in \citet{trompetal2010}, it takes three steps/simulations
-to obtain one contribution of the ensemble sensitivity kernels:
-\begin{itemize}
-\item Step 1: simulation for generating wavefields\\
- \texttt{SIMULATION\_TYPE=1}\\
- \texttt{NOISE\_TOMOGRAPHY=1}\\
- \texttt{SAVE\_FORWARD} not used, can be either .true. or .false.
-\item Step 2: simulation for ensemble forward wavefields\\
- \texttt{SIMULATION\_TYPE=1}\\
- \texttt{NOISE\_TOMOGRAPHY=2}\\
- \texttt{SAVE\_FORWARD=.true.}
-\item Step 3: simulation for ensemble adjoint wavefields and sensitivity
-kernels\\
- \texttt{SIMULATION\_TYPE=3}\\
- \texttt{NOISE\_TOMOGRAPHY=3}\\
- \texttt{SAVE\_FORWARD=.false.}
-
-Note Step 3 is an adjoint simulation, please refer to previous chapters
-on how to prepare adjoint sources and other necessary files, as well
-as how adjoint simulations work.
-
-\end{itemize}
-Note that it is better to run the three steps continuously within
-the same job, otherwise you have to collect the saved surface movies
-from the old nodes/CPUs to the new nodes/CPUs. This process varies
-from cluster to cluster and thus cannot be discussed here. Please
-ask your cluster administrator for information/configuration of the
-cluster you are using.\\
-
-\subsection{Post-simulation}
-
-After those simulations, you have all stuff you need, either in the
-\texttt{OUTPUT\_FILES/} or in the directory specified by \texttt{LOCAL\_PATH}
-in \texttt{DATA/Par\_file} (which are most probably on local nodes).
-Collect whatever you want from the local nodes to your workstation,
-and then visualize them. This process is the same as what you may
-have done for regular earthquake simulations. Refer to other chapters
-if you have problems.\\
-
-Simply speaking, two outputs are the most interesting: the simulated
-ensemble cross correlations and one contribution of the ensemble sensitivity
-kernels.\\
-
-The simulated ensemble cross correlations are obtained after the second
-simulation (Step 2). Seismograms in \texttt{OUTPUT\_FILES/} are actually
-the simulated ensemble cross correlations. Collect them immediately
-after Step 2, or the Step 3 will overwrite them. Note that we have
-a `master' receiver specified by \texttt{irec\_master\_noise}, the
-seismogram at one station corresponds to the cross correlation between
-that station and the `master'. Since the seismograms have three components,
-we may obtain cross correlations for different components as well,
-not necessarily the cross correlations between vertical components.\\
-
-One contribution of the ensemble cross-correlation sensitivity kernels
-are obtained after Step 3, stored in the \texttt{DATA/LOCAL\_PATH}
-on local nodes. The ensemble kernel files are named the same as regular
-earthquake kernels.\\
-
-You need to run another three simulations to get the other contribution
-of the ensemble kernels, using different forward and adjoint sources
-given in \citet{trompetal2010}.
-
-\section{Example}
-
-In order to illustrate noise simulations in an easy way, one example
-is provided in \texttt{EXAMPLES/noise\_tomography/}. See \texttt{EXAMPLES/noise\_tomography/README}
-for explanations. \\
-
-Note, however, that they are created for a specific workstation (CLOVER@PRINCETON),
-which has at least 4 cores with `mpif90' working properly. \\
-
-If your workstation is suitable, you can run the example in \texttt{EXAMPLES/noise\_tomography/}
-using:\\
-
-\texttt{./pre-processing.sh}\\
-
-Even if this script does not work on your workstation, the procedure
-it describes is universal. You may review the whole process described
-in the last section by following the commands in \texttt{pre-processing.sh},
-which should contain enough explanations for all the commands.
-
-%------------------------------------------------------------------------------------------------%
-
-\chapter{Gravity integral calculations for the gravity field of an Earth model}
-
-%------------------------------------------------------------------------------------------------%
-
-SPECFEM can now compute the gravity field as well as its derivatives (i.e., gravity gradiometry)
-generated by any given 3D Earth model at a given elevation, i.e., on a given observation surface.
-That feature is still experimental but should work just fine.\\
-
-To see how it is implemented and to use it, type this in the root directory of the code:\\
-
-\texttt{grep -i GRAVITY\_INTEGRALS src/*/*}
-
-%------------------------------------------------------------------------------------------------%
-
-\chapter{Graphics}
-
-%------------------------------------------------------------------------------------------------%
-
-\section{\label{sec:Mesh-graphics}Meshes}
-
-In case you used the internal mesher \texttt{xmeshfem3D} to create
-and partition your mesh, you can output mesh files in ABAQUS (.INP)
-and DX (.dx) format to visualize them. For this, you must set either
-the flag \texttt{CREATE\_DX\_FILES} or \texttt{CREATE\_ABAQUS\_FILES}
-to \texttt{.true.} in the mesher's parameter file \texttt{Mesh\_Par\_file}
-prior to running the mesher (see Chapter~\ref{cha:Running-the-Mesher-Meshfem3D}
-for details). You can then use AVS \urlwithparentheses{www.avs.com}
-or OpenDX \urlwithparentheses{www.opendx.org} to visualize the mesh
-and MPI partition (slices).
-
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=0.49\textwidth]{figures/vtk_mesh_vp} \includegraphics[width=0.49\textwidth]{figures/vtk_mesh_vs}
-\par\end{centering}
-
-\caption{Visualization using Paraview of VTK files created by \texttt{xgenerate\_databases}
-showing P- and S-wave velocities assigned to the mesh points. The
-mesh was created by \texttt{xmeshfem3D} for 4 processors.}
-
-\label{fig:vtk.mesh}
-\end{figure}
-
-You have also the option to visualize the distributed databases produced
-by \texttt{xgenerate\_databases} using Paraview \urlwithparentheses{www.paraview.org}.
-For this, you must set the flag \texttt{SAVE\_MESH\_FILES} to \texttt{.true.}
-in the main parameter file \texttt{Par\_file} (see Chapter~\ref{cha:Main-Parameter}
-for details). This will create VTK files for each single partition.
-You can then use Paraview \urlwithparentheses{www.paraview.org}
-to visualized these partitions.
-
-\section{\label{sec:Movies}Movies}
-
-To make a surface or volume movie of the simulation, set parameters
-\texttt{MOVIE\_SURFACE}, \texttt{MOVIE\_VOLUME}, \texttt{MOVIE\_TYPE},
-and \texttt{NTSTEP\_BETWEEN\_FRAMES} in the \texttt{Par\_file}. Turning
-on the movie flags, in particular \texttt{MOVIE\_VOLUME}, produces
-large output files. \texttt{MOVIE\_VOLUME} files are saved in the
-\texttt{LOCAL\_PATH} directory, whereas \texttt{MOVIE\_SURFACE} output
-files are saved in the \texttt{OUTPUT\_FILES} directory. We save the
-displacement field if the parameter \texttt{SAVE\_DISPLACEMENT} is
-set, otherwise the velocity field is saved. The look of a movie is
-determined by the half-duration of the source. The half-duration should
-be large enough so that the movie does not contain frequencies that
-are not resolved by the mesh, i.e., it should not contain numerical
-noise. This can be accomplished by selecting a CMT \texttt{HALF\_DURATION}
-> 1.1 $\times$ smallest period (see figure \ref{fig:CMTSOLUTION-file}).
-When \texttt{MOVIE\_SURFACE} = .\texttt{true.}, the half duration
-of each source in the \texttt{CMTSOLUTION} file is replaced by
-\begin{quote}
-\[
-\sqrt{(}\mathrm{\mathtt{HALF\_DURATIO}\mathtt{N}^{2}}+\mathrm{\mathtt{HDUR\_MOVI}\mathtt{E}^{2}})
-\]
- \textbf{NOTE:} If \texttt{HDUR\_MOVIE} is set to 0.0, the code will
-select the appropriate value of 1.1 $\times$ smallest period. As
-usual, for a point source one can set \texttt{half duration} in the
-\texttt{CMTSOLUTION} file to be 0.0 and \texttt{HDUR\_MOVIE} = 0.0
-in the \texttt{Par\_file} to get the highest frequencies resolved
-by the simulation, but for a finite source one would keep all the
-\texttt{half durations} as prescribed by the finite source model and
-set \texttt{HDUR\_MOVIE} = 0.0.
-\end{quote}
-
-\subsection{Movie Surface}
-
-When running \texttt{xspecfem3D} with the \texttt{MOVIE\_SURFACE}
-flag turned on, the code outputs \texttt{moviedata??????} files in
-the \texttt{OUTPUT\_FILES} directory. There are several flags in the
-main parameter file \texttt{Par\_file} that control the output of
-these movie data files (see section \ref{cha:Main-Parameter} for
-details): \texttt{\small NTSTEP\_BETWEEN\_FRAMES}{\small{} to set the
-timesteps between frames, }\texttt{\small SAVE\_DISPLACEMENT}{\small{}
-to save displacement instead of velocity, }\texttt{\small USE\_HIGHRES\_FOR\_MOVIES}{\small{}
-to save values at all GLL point instead of element edges. In order
-to output additionally shakemaps, you would set the parameter }\texttt{\small CREATE\_SHAKEMAP}{\small{}
-to }\texttt{\small .true.}{\small .}{\small \par}
-
-The files are in a fairly complicated binary format, but there is
-a program provided to convert the output into more user friendly formats:
-\begin{description}
-\item [{\texttt{xcreate\_movie\_shakemap\_AVS\_DX\_GMT}}] From \texttt{create\_movie\_shakemap\_AVS\_DX\_GMT.f90},
-it outputs data in ASCII, OpenDX, or AVS format (also readable in
-ParaView). Before compiling the code, make sure you have the file
-\texttt{surface\_from\_mesher.h} in the \texttt{OUTPUT\_FILES/} directory.
-This file will be created by the solver run. Then type
-
-\begin{lyxcode}
-{\footnotesize make~xcreate\_movie\_shakemap\_AVS\_DX\_GMT}{\footnotesize \par}
-\end{lyxcode}
-
-and run the executable \texttt{xcreate\_movie\_shakemap\_AVS\_DX\_GMT}
-in the main directory. It will create visualization files
-in your format of choice. The code will prompt the user for input
-parameters.
-
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=0.32\textwidth]{figures/movie_surf_1} \includegraphics[width=0.32\textwidth]{figures/movie_surf_2}
-\includegraphics[width=0.32\textwidth]{figures/movie_surf_3}
-\par\end{centering}
-
-\caption{Visualization using AVS files created by \texttt{xcreate\_movie\_shakemap\_AVS\_DX\_GMT}
-showing movie snapshots of vertical velocity components at different
-times.}
-
-\label{fig:movie.surf}
-\end{figure}
-
-\end{description}
-The \texttt{SPECFEM3D Cartesian} code is running in near real-time
-to produce animations of southern California earthquakes via the web;
-see Southern California ShakeMovie\textregistered{}\urlwithparentheses{www.shakemovie.caltech.edu}.
-
-\subsection{\label{sub:Movie-Volume}Movie Volume}
-
-When running xspecfem3D with the \texttt{\small MOVIE\_VOLUME}{\small{}
-flag turned on, the code outputs several files in }\texttt{\small LOCAL\_PATH}{\small{}
-specified in the main }\texttt{\small Par\_file}{\small , e.g. in
-directory }\texttt{\small OUTPUT\_FILES/DATABASES\_MPI}{\small .
-The output is saved by each processor at the time interval specified
-by }\texttt{\small NTSTEP\_BETWEEN\_FRAMES}{\small . For all domains,
-the velocity field is output to files: }\texttt{\small proc??????\_velocity\_X\_it??????.bin}{\small ,
-}\texttt{\small proc??????\_velocity\_Y\_it??????.bin}{\small{} and
-}\texttt{\small proc??????\_velocity\_Z\_it??????.bin}{\small . For
-elastic domains, the divergence and curl taken from the velocity field,
-i.e. $\nabla\cdot{\bf {v}}$ and $\nabla\times{\bf {v}}$, get stored
-as well: }\texttt{\small proc??????\_div\_it??????.bin}{\small , }\texttt{\small proc??????\_curl\_X\_t??????.bin}{\small ,
-}\texttt{\small proc??????\_curl\_Y\_it??????.bin}{\small{} and }\texttt{\small proc??????\_curl\_Z\_it??????.bin}{\small .
-The files denoted }\texttt{\small proc??????\_div\_}{\small ~}\\
-{\small{} }\texttt{\small glob\_it??????.bin}{\small{} and }\texttt{\small proc??????\_curl\_glob\_it??????.bin}{\small{}
-are stored on the global points only, all the other arrays are stored
-on all GLL points. Note that the components X/Y/Z can change to E/N/Z
-according to the }\texttt{\small SUPPRESS\_UTM\_PROJECTION}{\small{}
-flag (see also Appendix~\ref{cha:Coordinates} and \ref{cha:channel-codes}).}{\small \par}
-
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=0.32\textwidth]{figures/movie_volume_1} \includegraphics[width=0.32\textwidth]{figures/movie_volume_2}
-\includegraphics[width=0.32\textwidth]{figures/movie_volume_3}
-\par\end{centering}
-
-\caption{Paraview visualization using movie volume files (converted by \texttt{xcombine\_vol\_data}
-and \texttt{mesh2vtu.pl}) and showing snapshots of vertical velocity
-components at different times.}
-
-\label{fig:movie.volume}
-\end{figure}
-
-To visualize these files, we use an auxilliary program \texttt{combine\_vol\_data.f90}
-to combine the data from all slices into one mesh file. To compile
-it in the root directory, type:
-\begin{lyxcode}
-{\footnotesize make~xcombine\_vol\_data~}{\footnotesize \par}
-\end{lyxcode}
-which will create the executable \texttt{bin/xcombine\_vol\_data}.
-To output the usage of this executable, type
-\begin{lyxcode}
-{\footnotesize ./bin/xcombine\_vol\_data~}{\footnotesize \par}
-\end{lyxcode}
-without arguments.
-
-\medskip
-
-\texttt{xcombine\_vol\_data} will combine the data on the different processors (located in \texttt{OUTPUT\_FILES/DATABASES\_MPI}), for a given quantity and a given iteration, to one file.
-For example, if you want to combine \texttt{velocity\_Z}, for the iteration 400 on 4 processors, i.e.
-if you want to combine these files :
-
-\medskip
-
-\texttt{proc000000\_velocity\_Z\_it000400.bin }
-
-\texttt{proc000001\_velocity\_Z\_it000400.bin }
-
-\texttt{proc000002\_velocity\_Z\_it000400.bin }
-
-\texttt{proc000003\_velocity\_Z\_it000400.bin}
-
-\medskip
-
-You have to go in the directory of the concerned example (where are the directories \texttt{DATA}, \texttt{OUTPUT\_FILES}, etc.).
-Then, you can launch \texttt{xcombine\_vol\_data}, specifying the path where it is located.
-As an example, if the \texttt{DATA} and \texttt{OUTPUT\_FILES} directories of the concerned example are in the root specfem3d directory, \texttt{xcombine\_vol\_data} is in \texttt{./bin}, and you have to type:
-
-\begin{lyxcode}
-~{\footnotesize ./bin/xcombine\_vol\_data~0~3~velocity\_Z\_it000400~./OUTPUT\_FILES/DATABASES\_MPI~./OUTPUT\_FILES~0~}{\footnotesize \par}
-\end{lyxcode}
-
-Here, $0$ is the number of the first processor, $3$ the number of the last one, \texttt{velocity\_Z\_it000400} the name of the files we want to combine without the prefix \texttt{proc000000*\_}, \texttt{./OUTPUT\_FILES/DATABASES\_MPI} the directory where the files are located, \texttt{./OUTPUT\_FILES} the directory where the combined file will be stored, and $0$ is the parameter to create a low-resolution mesh file ($1$ for a high-resolution).
-
-\medskip
-
-At the beginning of the source file \texttt{combine\_vol\_data.f90} (located in \texttt{src/auxiliaries/}), there is a logical constant parameter \texttt{USE\_VTK\_OUTPUT} whose default value is \texttt{.true.}, that which will provide the output combined file directly in the VTK format and you can use Paraview to visualize it.
-
-\medskip
-
-If the value of this parameter is \texttt{.false.}, the output mesh file will have the name \texttt{velocity\_Z\_it000400.mesh}. Then we next have to convert the
-\texttt{.mesh} file into the VTU (Unstructured grid file) format which can be viewed in ParaView. For this task, you can use and modify the
-script \texttt{mesh2vtu.pl} located in directory \texttt{utils/Visualization/Paraview/}, for example:
-\begin{lyxcode}
-{\footnotesize mesh2vtu.pl~-i~velocity\_Z\_it000400.mesh~-o~velocity\_Z\_it000400.vtu}{\footnotesize \par}
-\end{lyxcode}
-Notice that this Perl script uses a program \texttt{mesh2vtu} in the \texttt{utils/Visualization/Paraview/mesh2vtu} directory, which further
-uses the VTK \urlwithparentheses{http://www.vtk.org/} run-time library for its execution. Therefore, make sure you have them properly set
-in the script according to your system.
-
-\bigskip
-
-Then, to do a movie with several iterations, you have to repeat this process for each iteration you want to put in your movie.
-
-\section{\label{sec:Finite-Frequency-Kernels}Finite-Frequency Kernels}
-
-The finite-frequency kernels computed as explained in Section \ref{sec:Adjoint-simulation-finite}
-are saved in the \texttt{LOCAL\_PATH} at the end of the simulation.
-Therefore, we first need to collect these files on the front end,
-combine them into one mesh file, and visualize them with some auxilliary
-programs.
-\begin{enumerate}
-\item \textbf{Create slice files}
-
-We will only discuss the case of one source-receiver pair, i.e., the
-so-called banana-doughnut kernels. Although it is possible to collect
-the kernel files from all slices on the front end, it usually takes
-up too much storage space (at least tens of gigabytes). Since the
-sensitivity kernels are the strongest along the source-receiver great
-circle path, it is sufficient to collect only the slices that are
-along or close to the great circle path.
-
-A Perl script \texttt{slice\_number.pl} located in directory \texttt{utils/Visualization/Paraview/}
-can help to figure out the slice numbers that lie along the great
-circle path. It applies to meshes created with the internal mesher
-\texttt{xmeshfem3D}.
-\begin{enumerate}
-\item On machines where you have access to the script, copy the \texttt{Mesh\_Par\_file},
-and \texttt{output\_solver} files, and run:
-
-\begin{lyxcode}
-{\small slice\_number.pl~Mesh\_Par\_file~output\_solver.txt~slice\_file}{\small \par}
-\end{lyxcode}
-
-which will generate a \texttt{slices\_file}.
-
-\item For cases with multiple sources and multiple receivers, you need to
-provide a slice file before proceeding to the next step.
-\end{enumerate}
-\item \textbf{Collect the kernel files}
-
-After obtaining the slice files, you can collect the corresponding
-kernel files from the given slices.
-\begin{enumerate}
-\item You can use or modify the script \texttt{utils/copy\_basin\_database.pl}
-to accomplish this:
-
-\begin{lyxcode}
-{\small utils/copy\_database.pl~slice\_file~lsf\_machine\_file~filename~{[}jobid{]}}{\small \par}
-\end{lyxcode}
-
-where \texttt{\small lsf\_machine\_file}{\small{} is the machine file
-generated by the LSF scheduler, }\texttt{\small filename}{\small{}
-is the kernel name (e.g., }\texttt{\small rho\_kernel}{\small , }\texttt{\small alpha\_kernel}{\small{}
-and }\texttt{\small beta\_kernel}{\small ), and the optional }\texttt{\small jobid}{\small{}
-is the name of the subdirectory under }\texttt{\small LOCAL\_PATH}{\small{}
-where all the kernel files are stored.}{\small \par}
-
-\item After executing this script, all the necessary mesh topology files
-as well as the kernel array files are collected to the local directory
-of the front end.
-\end{enumerate}
-\item \textbf{Combine kernel files into one mesh file}
-
-We use an auxilliary program \texttt{combine\_vol\_data.f90} to combine
-the kernel files from all slices into one mesh file.
-\begin{enumerate}
-\item Compile it in the root directory:
-
-\begin{lyxcode}
-{\footnotesize make~xcombine\_vol\_data~}{\footnotesize \par}
-
-{\footnotesize ./bin/xcombine\_vol\_data~slice\_list~filename~input\_dir~output\_dir~high/low-resolution~}{\footnotesize \par}
-\end{lyxcode}
-
-where \texttt{input\_dir} is the directory where all the individual
-kernel files are stored, and \texttt{output\_dir} is where the mesh
-file will be written.
-
-\item Use 1 for a high-resolution mesh, outputting all the GLL points to
-the mesh file, or use 0 for low resolution, outputting only the corner
-points of the elements to the mesh file.
-\item The output mesh file will have the name \texttt{filename\_rho(alpha,beta).mesh}
-\end{enumerate}
-\item \textbf{Convert mesh files into .vtu files}
-
-\begin{enumerate}
-\item We next convert the \texttt{.mesh} file into the VTU (Unstructured
-grid file) format which can be viewed in ParaView. For this task,
-you can use and modify the script \texttt{mesh2vtu.pl} located in
-directory \texttt{utils/Visualization/Paraview/}, for example:
-
-\begin{lyxcode}
-{\footnotesize mesh2vtu.pl~-i~file.mesh~-o~file.vtu~}{\footnotesize \par}
-\end{lyxcode}
-\item Notice that this Perl script uses a program \texttt{mesh2vtu} in the
-\texttt{utils/Visualization/Paraview/mesh2vtu} directory, which further
-uses the VTK \urlwithparentheses{http://www.vtk.org/} run-time library
-for its execution. Therefore, make sure you have them properly set
-in the script according to your system.
-\end{enumerate}
-\item \textbf{Copy over the source and receiver .vtk file}
-
-In the case of a single source and a single receiver, the simulation
-also generates the file \texttt{sr.vtk} located in the \texttt{OUTPUT\_FILES/}
-directory to describe the source and receiver locations, which can
-also be viewed in Paraview in the next step.
-
-\item \textbf{View the mesh in ParaView}
-
-Finally, we can view the mesh in ParaView \urlwithparentheses{www.paraview.org}.
-\begin{enumerate}
-\item Open ParaView.
-\item From the top menu, \textsf{File} $\rightarrow$\textsf{Open data},
-select \texttt{file.vtu}, and click the \textsf{Accept} button.
-
-\begin{itemize}
-\item If the mesh file is of moderate size, it shows up on the screen; otherwise,
-only the bounding box is shown.
-\end{itemize}
-\item Click \textsf{Display Tab} $\rightarrow$ \textsf{Display Style} $\rightarrow$
-\textsf{Representation} and select \textsf{wireframe of surface} to
-display it.
-\item To create a cross-section of the volumetric mesh, choose \textsf{Filter}
-$\rightarrow$ \textsf{cut}, and under \textsf{Parameters Tab}, choose
-\textsf{Cut Function} $\rightarrow$ \textsf{plane}.
-\item Fill in center and normal information given by the \texttt{global\_slice\_number.pl}
-script (either from the standard output or from \texttt{normal\_plane.txt}
-file).
-\item To change the color scale, go to \textsf{Display Tab} $\rightarrow$
-\textsf{Color} $\rightarrow$ \textsf{Edit Color Map} and reselect
-lower and upper limits, or change the color scheme.
-\item Now load in the source and receiver location file by \textsf{File}
-$\rightarrow$ \textsf{Open data}, select \texttt{sr.vt}k, and click
-the \textsf{Accept} button. Choose \textsf{Filter} $\rightarrow$
-\textsf{Glyph}, and represent the points by `\textsf{spheres}'.
-\item For more information about ParaView, see the ParaView Users Guide
-\urlwithparentheses{www.paraview.org/files/v1.6/ParaViewUsersGuide.PDF}.
-\end{enumerate}
-\end{enumerate}
-\begin{figure}[H]
-\noindent \begin{centering}
-\includegraphics[scale=0.6]{figures/3D-S-Kernel}
-\par\end{centering}
-
-\caption{(a) Top Panel: Vertical source-receiver cross-section of the S-wave
-finite-frequency sensitivity kernel $K_{\beta}$ for station GSC at
-an epicentral distance of 176 km from the September 3, 2002, Yorba
-Linda earthquake. Lower Panel: Vertical source-receiver cross-section
-of the 3D S-wave speed model used for the spectral-element simulations
-\citep{KoLiTrSuStSh04}. (b) The same as (a) but for station HEC at
-an epicentral distance of 165 km \citep{LiTr06}.}
-
-\label{figure:P-wave-speed-finite-frequency}
-\end{figure}
-
-\chapter{\label{cha:Scheduler}Running through a Scheduler}
-
-The code is usually run on large parallel machines, often PC clusters,
-most of which use schedulers, i.e., queuing or batch management systems
-to manage the running of jobs from a large number of users. The following
-considerations need to be taken into account when running on a system
-that uses a scheduler:
-\begin{itemize}
-\item The processors/nodes to be used for each run are assigned dynamically
-by the scheduler, based on availability. Therefore, in order for the
-\texttt{xgenerate\_databases} and the \texttt{xspecfem3D} executables
-(or between successive runs of the solver) to have access to the same
-database files (if they are stored on hard drives local to the nodes
-on which the code is run), they must be launched in sequence as a
-single job.
-\item On some systems, the nodes to which running jobs are assigned are
-not configured for compilation. It may therefore be necessary to pre-compile
-both the \texttt{xgenerate\_databases} and the \texttt{xspecfem3D}
-executables.
-\item One feature of schedulers/queuing systems is that they allow submission
-of multiple jobs in a ``launch and forget'' mode. In order to take
-advantage of this property, care needs to be taken that output and
-intermediate files from separate jobs do not overwrite each other,
-or otherwise interfere with other running jobs.
-\end{itemize}
-Examples of job scripts can be found in the \texttt{\small utils/Cluster/}{\small{}
-directory and can straightforwardly be modified and adapted to meet
-more specific running needs.}{\small \par}
-
-\chapter{{\normalsize \label{cha:-Changing-the}} Changing the Model}
-
-In this section we explain how to change the velocity model used for
-your simulations. These changes involve contributing specific subroutines
-that replace existing subroutines in the \texttt{SPECFEM3D Cartesian}
-package. Note that \texttt{SPECFEM3D Cartesian} can handle Earth models
-with material properties that vary within each spectral element.
-
-\section{{\normalsize \label{sec:Using-tomographic}}Using external tomographic Earth models}
-
-To implement your own external tomographic model(s), you must provide
-your own external tomography file(s), and choose between two possible
-options:\\
- \indent (1) set in the \texttt{Par\_file} the model parameter \texttt{MODEL
-= tomo}, \\
- or for more user control: \\
- \indent (2) set in the \texttt{Par\_file} the model parameter \texttt{MODEL
-= default}, define the negative \texttt{material\_ID} identifier for
-each element in the file \texttt{MESH/materials\_file} and use the
-following format in the file \texttt{MESH/nummaterial\_velocity\_file}
-when using CUBIT to construct your mesh (see also section \ref{subsec:Exporting-the-Mesh}):
-\begin{lyxcode}
-domain\_ID~material\_ID~tomography~elastic~file\_name~1
-\end{lyxcode}
-where: \\
- \indent \texttt{domain\_ID} is 1 for acoustic or 2 for elastic materials,
-\\
- \indent \texttt{material\_ID} a negative, unique identifier (i.e.,
--1,-2,...), \\
- \indent \texttt{tomography} keyword for tomographic material definition,
-\\
- \indent \texttt{elastic} keyword for elastic material definition,
-\\
- \indent \texttt{file\_name} the name of the tomography file and
-1 a positive unique identifier.\\
-
-The external tomographic model is represented by a grid of points
-with assigned material properties and homogeneous resolution along
-each spatial direction x, y and z. The ASCII file \texttt{file\_name}
-that describe the tomography should be located in the \texttt{TOMOGRAPHY\_PATH}
-directory, set in the \texttt{Par\_file}. The format of the file,
-as read from \texttt{model\_tomography.f90} located in the \texttt{src/generate\_databases}
-directory, looks like Figure \ref{fig:tomography_file}, where
-\begin{description}
-\item [{\texttt{ORIG\_X}, \texttt{END\_X}}] are, respectively, the coordinates
-of the initial and final tomographic grid points along the x direction
-(in the mesh units, e.g., $m$);
-\item [{\texttt{ORIG\_Y}, \texttt{END\_Y}}] respectively the coordinates
-of the initial and final tomographic grid points along the y direction
-(in the mesh units, e.g., $m$);
-\item [{\texttt{ORIG\_Z}, \texttt{END\_Z}}] respectively the coordinates
-of the initial and final tomographic grid points along the z direction
-(in the mesh units, e.g., $m$);
-\item [{\texttt{SPACING\_X}, \texttt{SPACING\_Y}, \texttt{SPACING\_Z}}] the
-spacing between the tomographic grid points along the x, y and z directions,
-respectively (in the mesh units, e.g., $m$);
-\item [{\texttt{NX}, \texttt{NY}, \texttt{NZ}}] the number of grid points
-along the spatial directions x, y and z, respectively; \texttt{NX}
-is given by {[}(\texttt{END\_X} - \texttt{ORIG\_X})/\texttt{SPACING\_X}{]}+1;
-\texttt{NY} and \texttt{NZ} are the same as \texttt{NX}, but for the
-y and z directions, respectively;
-\item [{\texttt{VP\_MIN}, \texttt{VP\_MAX}, \texttt{VS\_MIN}, \texttt{VS\_MAX}, \texttt{RHO\_MIN}, \texttt{RHO\_MAX}}] the
-minimum and maximum values of the wave speed \texttt{vp} and \texttt{vs}
-(in $m\, s^{-1}$) and of the density \texttt{rho} (in $kg\, m^{-3}$);
-these values could be the actual limits of the tomographic parameters
-in the grid or the minimum and maximum values to which we force the
-cut of velocity and density in the model.
-\end{description}
-After the first four lines, in the file \texttt{file\_name} the tomographic
-grid points are listed with the corresponding values of \texttt{vp},
-\texttt{vs} and \texttt{rho}, scanning the grid along the x coordinate
-(from \texttt{ORIG\_X} to \texttt{END\_X} with step of \texttt{SPACING\_X})
-for each given y (from \texttt{ORIG\_Y} to \texttt{END\_Y}, with step
-of \texttt{SPACING\_Y}) and each given z (from \texttt{ORIG\_Z} to
-\texttt{END\_Z}, with step of \texttt{SPACING\_Z}). %fig
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=1\textwidth]{figures/tomo_file}
-\par\end{centering}
-
-\caption{Tomography file \texttt{file\_name} that describes an external Earth
-model. The coordinates x, y and z of the grid, their limits ORIG\_X,
-ORIG\_Y, ORIG\_Z, END\_X, END\_Y, END\_Z and the grid spacings SPACING\_X,
-SPACING\_Y, SPACING\_Z should be in the units of the constructed mesh
-(e.g., UTM coordinates in meters).}
-
-\label{fig:tomography_file}
-\end{figure}
-
-The user can implement his own interpolation algorithm for the tomography
-model by changing the routine \texttt{model\_tomography.f90} located
-in the \texttt{src/generate\_databases/} directory. Moreover, for
-models that involve both fully defined materials and a tomography
-description, the \texttt{nummaterial\_velocity\_file} has multiple
-lines each with the corresponding suitable format described above.
-
-%==============
-
-\subsubsection*{Example: External tomography file with variable discretization intervals}
-
-The example in Figure~\ref{fig:tomography_file} is for an external tomography file with uniform spacing in the $x$, $y$, and $z$ directions.
-(Note that $\Delta x$, $\Delta y$, and $\Delta z$ need not be the same as in the example.) In that example, the  \texttt{nummaterial\_velocity\_file} is
-%
-\begin{lyxcode}
-2  -1 tomography elastic tomography\_model.xyz 1
-\end{lyxcode}
-%
-and the file \texttt{tomography\_model.xyz} will need to reside in \texttt{DATA/}.
-All entries in the second column of \texttt{materials\_file} will be \texttt{-1}, which means that each element in the mesh will be interpolated accordind to the values in \texttt{tomography\_model.xyz}.
-
-In some cases it may be desirable to have an external tomography model that is described in more than one file.
-For example, in cases like southern California, the length scale of variation in the structure of the wave speed model is much shorter in the sedimentary basin models within the upper 15~km.
-Therefore one might want to use an external tomography file that is sampled with, say, $\Delta x = 1000$~m, $\Delta y = 1000$~m, and $\Delta z = 250$~m in the uppermost 15~km, and then use $\Delta x = 2000$~m, $\Delta y = 2000$~m, and $\Delta z = 1000$~m below a depth of 15~km.
-If these intervals are chosen appropriately, then it will result in a pair of external tomography file that is much smaller than the alternative of having a single file with the fine discretization.
-In this case  \texttt{nummaterial\_velocity\_file} is
-%
-\begin{lyxcode}
-2  -1 tomography elastic file\_above\_15km.xyz 1 \\
-2  -2 tomography elastic file\_below\_15km.xyz 1
-\end{lyxcode}
-%
-and the files \texttt{file\_above\_15km.xyz} and \texttt{file\_below\_15km.xyz} will need to reside in \texttt{DATA/}.
-All entries in the second column of \texttt{materials\_file} will need to be either \texttt{-1} or \texttt{-2}, depending on whether the element is above or below 15~km depth.
-In this sense, the construction of the mesh and the discretization of the external model will generally be done in tandem.
-
-Other more complicated discretizations can be used following the same procedure. (In particular, there is no limit on the number of different external files that are used.)
-
-%=============
-
-\section{{\normalsize \label{sec:Anelastic-Models}}External (an)elastic Models}
-
-To use your own external model, you can set in the \texttt{Par\_file}
-the model parameter \texttt{MODEL = external}. Three-dimensional acoustic
-and/or (an)elastic (attenuation) models may then be superimposed onto
-the mesh based upon your subroutine in \texttt{model\_external\_values.f90}
-located in directory \texttt{src/generate\_databases}. The call to
-this routine would be as follows
-\begin{lyxcode}
-call~model\_external\_values(xmesh,~ymesh,~zmesh,~~\&~~\\
-~~~~~~~~~~~rho,~vp,~vs,~qmu\_atten,~iflag\_aniso,~idomain\_id)
-\end{lyxcode}
-Input to this routine consists of:
-\begin{description}
-\item [{\texttt{xmesh,ymesh,zmesh}}] location of mesh point
-\end{description}
-Output to this routine consists of:
-\begin{description}
-\item [{\texttt{rho,vp,vs}}] isotropic model parameters for density $\rho$
-($kg/m^{3}$), $v_{p}$ ($m/s$) and $v_{s}$ ($m/s$)
-\item [{\texttt{qmu\_atten}}] Shear wave quality factor: $0<Q_{\mu}<9000$
-\item [{\texttt{iflag\_aniso}}] anisotropic model flag, $0$ indicating
-no anisotropy or $1$ using anisotropic model parameters as defined
-in routine file \texttt{model\_aniso.f90}
-\item [{\texttt{idomain\_id}}] domain identifier, $1$ for acoustic, $2$
-for elastic, $3$ for poroelastic materials.
-\end{description}
-Note that the resolution and maximum value of anelastic models are
-truncated. This speeds the construction of the standard linear solids
-during the meshing stage. To change the resolution, currently at one
-significant figure following the decimal, or the maximum value (9000),
-consult \texttt{shared/constants.h}.
-
-\section{{\normalsize \label{sec:Anisotropic-Models}}Anisotropic Models}
-
-To use your anisotropic models, you can either set in the \texttt{Par\_file}
-the model parameter \texttt{ANISOTROPY} to \texttt{.true.} or set
-the model parameter \texttt{MODEL = aniso}. Three-dimensional anisotropic
-models may be superimposed on the mesh based upon the subroutine in
-file \texttt{model\_aniso.f90} located in directory \texttt{src/generate\_databases}.
-The call to this subroutine is of the form
-\begin{lyxcode}
-call~model\_aniso(iflag\_aniso,~rho,~vp,~vs,~\&~~\\
-~~~~~~~~~~~c11,c12,c13,c14,c15,c16,c22,c23,c24,c25,c26,~\&~~\\
-~~~~~~~~~~~c33,c34,c35,c36,c44,c45,c46,c55,c56,c66)~
-\end{lyxcode}
-Input to this routine consists of:
-\begin{description}
-\item [{\texttt{iflag\_aniso}}] flag indicating the type of the anisotropic
-model, i.e. $0$ for no superimposed anisotropy, $1$ or $2$ for
-generic pre-defined anisotropic models.
-\item [{\texttt{rho,vp,vs}}] reference isotropic model parameters for density
-$\rho$, $v_{p}$ and $v_{s}$.
-\end{description}
-Output from the routine consists of the following non-dimensional
-model parameters:
-\begin{description}
-\item [{\texttt{c11},}] \textbf{$\cdots$,} \texttt{\textbf{c66}} 21~dimensionalized
-anisotropic elastic parameters.
-\end{description}
-You can replace the \texttt{model\_aniso.f90} file by your own version
-\textit{provided you do not change the call structure of the routine},
-i.e., the new routine should take exactly the same input and produce
-the required relative output.
-
-\section{{\normalsize \label{sec:Using-SEP}}Using external SEP models.}
-
-Stanford Exploration Project (SEP) models consists of two files per variables, one header file (\texttt{.H})
-and one binary file (\texttt{.H$@$}).
-Among other information, the header file indicates, for each dimension,
-the offset (\texttt{o1}, \texttt{o2} and \texttt{o3} fields),
-the spatial step size (\texttt{d1}, \texttt{d2} and \texttt{d3} fields)
-and the number of samples (\texttt{n1}, \texttt{n2} and \texttt{n3}) necessary to read
-the linked binary file (\texttt{in} field).
-
-SEP model reading occurs during database generation. It it the user responsibility
-to ensure that proper interfaces between acoustic and elastic domains are set up
-during the meshing phase. For more information, refer to
-section~\ref{cha:Running-the-Mesher-Meshfem3D}.
-
-In \texttt{DATA/Par\_file}, \texttt{\textbf{MODEL}} should be set to \texttt{\textbf{sep}},
-and \texttt{\textbf{SEP\_MODEL\_DIRECTORY}} should point to wherever your sep model files reside.
-
-Note that in order not to overload the parameter file, SEP header files
-should be named \texttt{vp.H}, \texttt{vs.H} and \texttt{rho.H}. The \texttt{in} field
-in these binary files can be any path as long as it is relative to
-the previously set \texttt{SEP\_MODEL\_DIRECTORY}.
-We also assume that any dimensional value inside SEP header files is given in meters and
-that binary files are single precision floating point numbers stored in little endian format.
-There is currently only support for \texttt{vp}, \texttt{vs} and \texttt{rho} variables.
-
-An example demonstrating the use SEP models is given in:
-\texttt{EXAMPLES/meshfem3D\_examples/sep\_bathymetry}.
-
-\chapter{Post-Processing Scripts}
-
-Several post-processing scripts/programs are provided in the \texttt{utils/}
-directory, and most of them need to be adjusted when used on different
-systems, for example, the path of the executable programs. Here we
-only list a few of the available scripts and provide a brief description,
-and you can either refer to the related sections for detailed usage
-or, in a lot of cases, type the script/program name without arguments
-for its usage.
-
-\section{Process Data and Synthetics\label{sec:Process-data-and-syn}}
-
-In many cases, the SEM synthetics are calculated and compared to data
-seismograms recorded at seismic stations. Since the SEM synthetics
-are accurate for a certain frequency range, both the original data
-and the synthetics need to be processed before a comparison can be
-made.
-
-For such comparisons, the following steps are recommended:
-\begin{enumerate}
-\item Make sure that both synthetic and observed seismograms have the correct
-station/event and timing information.
-\item Convolve synthetic seismograms with a source time function with the
-half duration specified in the \texttt{CMTSOLUTION} file, provided,
-as recommended, you used a zero half duration in the SEM simulations.
-\item Resample both observed and synthetic seismograms to a common sampling
-rate.
-\item Cut the records using the same window.
-\item Remove the trend and mean from the records and taper them.
-\item Remove the instrument response from the observed seismograms (recommended)
-or convolve the synthetic seismograms with the instrument response.
-\item Make sure that you apply the same filters to both observed and synthetic
-seismograms. Preferably, avoid filtering your records more than once.
-\item Now, you are ready to compare your synthetic and observed seismograms.
-\end{enumerate}
-\noindent We generally use the following scripts provided in the \texttt{utils/seis\_process/}
-directory:
-
-\subsection{Data processing script \texttt{process\_data.pl}}
-
-This script cuts a given portion of the original data, filters it,
-transfers the data into a displacement record, and picks the first
-P and S arrivals. For more functionality, type `\texttt{process\_data.pl}'
-without any argument. An example of the usage of the script:
-\begin{lyxcode}
-{\footnotesize process\_data.pl~-m~CMTSOLUTION~-s~1.0~-l~0/4000~-i~-f~-t~40/500~-p~-x~bp~DATA/1999.330{*}.BH?.SAC}~
-\end{lyxcode}
-\noindent which has resampled the SAC files to a sampling rate of
-1 seconds, cut them between 0 and 4000 seconds, transfered them into
-displacement records and filtered them between 40 and 500 seconds,
-picked the first P and S arrivals, and added suffix `\texttt{bp}'
-to the file names.
-
-Note that all of the scripts in this section actually use the SAC
-and/or IASP91 to do the core operations; therefore make sure that
-the SAC and IASP91 packages are installed properly on your system,
-and that all the environment variables are set properly before running
-these scripts.
-
-\subsection{Synthetics processing script \texttt{process\_syn.pl}}
-
-This script converts the synthetic output from the SEM code from ASCII
-to SAC format, and performs similar operations as `\texttt{process\_data.pl}'.
-An example of the usage of the script:
-\begin{lyxcode}
-{\footnotesize process\_syn.pl~-m~CMTSOLUTION~-h~-a~STATIONS~-s~1.0~-l~0/4000~-f~-t~40/500~-p~-x~bp~SEM/{*}.BX?.semd}~
-\end{lyxcode}
-which will convolve the synthetics with a triangular source-time function
-from the \texttt{CMTSOLUTION} file, convert the synthetics into SAC
-format, add event and station information into the SAC headers, resample
-the SAC files with a sampling rate of 1 seconds, cut them between
-0 and 4000 seconds, filter them between 40 and 500 seconds with the
-same filter used for the observed data, pick the first P and S arrivals,
-and add the suffix `\texttt{bp}' to the file names.
-
-More options are available for this script, such as adding time shift
-to the origin time of the synthetics, convolving the synthetics with
-a triangular source time function with a given half duration, etc.
-Type \texttt{process\_syn.pl} without any argument for a detailed
-usage.
-
-In order to convert between SAC format and ASCII files, useful scripts
-are provided in the subdirectories ~\\
- \texttt{utils/sac2000\_alpha\_convert/} and \texttt{utils/seis\_process/asc2sac/}.
-
-\subsection{Script \texttt{rotate.pl}}
-
-The original data and synthetics have three components: vertical (BHZ
-resp. BXZ), north (BHN resp. BXN) and east (BHE resp. BXE). However,
-for most seismology applications, transverse and radial components
-are also desirable. Therefore, we need to rotate the horizontal components
-of both the data and the synthetics to the transverse and radial direction,
-and \texttt{\small rotate.pl}{\small{} can be used to accomplish this:}{\small \par}
-\begin{lyxcode}
-rotate.pl~-l~0~-L~180~-d~DATA/{*}.BHE.SAC.bp~
-
-rotate.pl~-l~0~-L~180~SEM/{*}.BXE.semd.sac.bp~
-\end{lyxcode}
-where the first command performs rotation on the SAC data obtained
-through Seismogram Transfer Program (STP) \urlwithparentheses{http://www.data.scec.org/STP/stp.html},
-while the second command rotates the processed SEM synthetics.
-
-\section{Collect Synthetic Seismograms}
-
-The forward and adjoint simulations generate synthetic seismograms
-in the \texttt{OUTPUT\_FILES/} directory by default. For the forward
-simulation, the files are named like \texttt{NT.STA.BX?.semd} for
-two-column time series, or \texttt{NT.STA.BX?.semd.sac} for ASCII
-SAC format, where NT and STA are the network code and station name,
-and \texttt{BX?} stands for the component name. Please see the Appendix~\ref{cha:Coordinates}
-and \ref{cha:channel-codes} for further details.
-
-The adjont simulations generate synthetic seismograms with the name
-\texttt{NT.S?????.S??.sem} (refer to Section \ref{sec:Adjoint-simulation-sources}
-for details). The kernel simulations output the back-reconstructed
-synthetic seismogram in the name \texttt{NT.STA.BX?.semd}, mainly
-for the purpose of checking the accuracy of the reconstruction. Refer
-to Section \ref{sec:Adjoint-simulation-finite} for further details.
-
-You do have further options to change this default output behavior,
-given in the main constants file \texttt{constants.h} located in \texttt{src/shared/}
-directory:
-\begin{description}
-\item [{\texttt{SEISMOGRAMS\_BINARY}}] set to \texttt{.true.} to have seismograms
-written out in binary format.
-\item [{\texttt{WRITE\_SEISMOGRAMS\_BY\_MASTER}}] Set to \texttt{.true.}
-to have only the master process writing out seismograms. This can
-be useful on a cluster, where only the master process node has access
-to the output directory.
-\item [{\texttt{USE\_OUTPUT\_FILES\_PATH}}] Set to \texttt{.false.} to
-have the seismograms output to \texttt{LOCAL\_PATH} directory specified
-in the main parameter file \texttt{DATA/Par\_file}. In this case,
-you could collect the synthetics onto the frontend using the \texttt{collect\_seismo\_lsf\_multi.pl}
-script located in the \texttt{utils/Cluster/lsf/} directory. The usage
-of the script would be e.g.:
-
-\begin{lyxcode}
-collect\_seismo.pl~machines~DATA/Par\_file
-\end{lyxcode}
-\end{description}
-where \texttt{machines} is a file containing the node names and \texttt{DATA/Par\_file}
-the parameter file used to extract the \texttt{LOCAL\_PATH} directory
-used for the simulation.
-
-\section{Clean Local Database}
-
-After all the simulations are done, the seismograms are collected,
-and the useful database files are copied to the frontend, you may
-need to clean the local scratch disk for the next simulation. This
-is especially important in the case of kernel simulation, where very
-large files are generated for the absorbing boundaries to help with
-the reconstruction of the regular forward wavefield. A sample script
-is provided in \texttt{utils/}:
-\begin{lyxcode}
-cleanbase.pl~machines
-\end{lyxcode}
-where \texttt{machines} is a file containing the node names.
-
-\section{Plot Movie Snapshots and Synthetic Shakemaps}
-
-\subsection{Script \texttt{movie2gif.gmt.pl}}
-
-With the movie data saved in \texttt{OUTPUT\_FILES/} at the end of
-a movie simulation (\texttt{\small MOVIE\_SURFACE=.true.}{\small ),
-you can run the }\texttt{\small `create\_movie\_shakemap\_AVS\_DX\_GMT}{\small '
-code to convert these binary movie data into GMT xyz files for futher
-processing. A sample script }\texttt{\small movie2gif.gmt.pl}{\small{}
-is provided to do this conversion, and then plot the movie snapshots
-in GMT, for example:}{\small \par}
-\begin{lyxcode}
-movie2gif.gmt.pl~-m~CMTSOLUTION~-g~-f~1/40~-n~-2~-p~
-\end{lyxcode}
-which for the first through the 40th movie frame, converts the \texttt{moviedata}
-files into GMT xyz files, interpolates them using the 'nearneighbor'
-command in GMT, and plots them on a 2D topography map. Note that `\texttt{-2}'
-and `\texttt{-p}' are both optional.
-
-\subsection{Script \texttt{plot\_shakemap.gmt.pl}}
-
-With the shakemap data saved in \texttt{OUTPUT\_FILES/} at the end
-of a shakemap simulation ~\\
- (\texttt{CREATE\_SHAKEMAP=.true.}), you can also run \texttt{`create\_movie\_shakemap\_AVS\_DX\_GMT}'
-code to convert the binary shakemap data into GMT xyz files. A sample
-script \texttt{plot\_shakemap.gmt.pl} is provided to do this conversion,
-and then plot the shakemaps in GMT, for example:
-\begin{lyxcode}
-plot\_shakemap.gmt.pl~data\_dir~type(1,2,3)~CMTSOLUTION~
-\end{lyxcode}
-where \texttt{type=1} for a displacement shakemap, \texttt{2} for
-velocity, and \texttt{3} for acceleration.
-
-\section{Map Local Database}
-
-A sample program \texttt{remap\_database} is provided to map the local
-database from a set of machines to another set of machines. This is
-especially useful when you want to run mesher and solver, or different
-types of solvers separately through a scheduler (refer to Chapter~\ref{cha:Scheduler}).
-\begin{lyxcode}
-run\_lsf.bash~-{}-gm-no-shmem~-{}-gm-copy-env~remap\_database~old\_machines~150
-\end{lyxcode}
-where \texttt{old\_machines} is the LSF machine file used in the previous
-simulation, and \texttt{150} is the number of processors in total.
-
-%============================================================
-
-\chapter{\label{cha:developers}Information for developers of the code, and for people who want to learn how the technique works}
-
-You can get a very simple 1D version of a demo code (there is one in Fortran and one in Python):\\
-
-git clone --recursive https://github.com/geodynamics/specfem1d.git\\
-
-\noindent We also have simple 3D demo source codes that implement the SEM in a single, small program, in directory\\
-utils/small\_SEM\_solvers\_in\_Fortran\_and\_C\_without\_MPI\_to\_learn of the specfem3d package.
-They are useful to learn how the spectral-element method works, and how to write or modify a code to implement it.
-Also useful to test new ideas by modifying these simple codes to run some tests.\\
-
-\noindent For information on how to contribute to the code, i.e., for how to make your modifications, additions or improvements part of the
-official package, see https://github.com/geodynamics/specfem3d/wiki .
-
-%============================================================
-
-\chapter*{\label{cha:Bug-Reports-and}Bug Reports and Suggestions for Improvements}
-
-To report bugs or suggest improvements to the code, please send an
-e-mail to the CIG Computational Seismology Mailing List \urlwithparentheses{cig-seismo@geodynamics.org}.
-
-\addcontentsline{toc}{chapter}{Notes \& Acknowledgments}
-
-\chapter*{Notes \& Acknowledgments}
-
-In order to keep the software package thread-safe in case a multithreaded
-implementation of MPI is used, developers should not add modules or
-common blocks to the source code but rather use regular subroutine
-arguments (which can be grouped in ``derived types'' if needed for
-clarity).
-
-The Gauss-Lobatto-Legendre subroutines in \texttt{gll\_library.f90}
-are based in part on software libraries from the Massachusetts Institute
-of Technology, Department of Mechanical Engineering (Cambridge, Massachusetts,
-USA). The non-structured global numbering software was provided by
-Paul F. Fischer (Brown University, Providence, Rhode Island, USA,
-now at Argonne National Laboratory, USA).
-
-OpenDX \urlwithparentheses{http://www.opendx.org} is open-source
-based on IBM Data Explorer, AVS \urlwithparentheses{http://www.avs.com}
-is a trademark of Advanced Visualization Systems, and ParaView \urlwithparentheses{http://www.paraview.org}
-is an open-source visualization platform.
-
-Please e-mail your feedback, questions, comments, and suggestions
-to the CIG Computational Seismology Mailing List \urlwithparentheses{cig-seismo@geodynamics.org}.
-
-\addcontentsline{toc}{chapter}{Copyright}
-
-\chapter*{Copyright}
-
-Main historical authors: Dimitri Komatitsch and Jeroen Tromp
-
-Princeton University, USA, and CNRS / University of Marseille, France
-
-$\copyright$ Princeton University and CNRS / University of Marseille, July 2012
-
-This program is free software; you can redistribute it and/or modify
-it under the terms of the GNU General Public License as published
-by the Free Software Foundation (see Appendix \ref{cha:License}).\\
-
-\textbf{\underline{Evolution of the code:}}\\
-
-MPI v. 3.0, December 2014: many developers.
-Convolutional PML, LDDRK time scheme, bulk attenuation support, simultaneous MPI runs,
-ADIOS file I/O support, coupling with external codes, new seismogram names,
-Deville routines for additional GLL degrees, tomography tools, unit/regression test framework,
-improved CUDA GPUs performance, additonal GEOCUBIT support, better make compilation,
-git versioning system. \\
-
-MPI v. 2.1, July 2012: Max Rietmann, Peter Messmer, Daniel Peter, Dimitri
-Komatitsch, Joseph Charles, Zhinan Xie: support for CUDA GPUs, better
-CFL stability for the Stacey absorbing conditions. \\
-
-MPI v. 2.0, November 2010: Dimitri Komatitsch, Nicolas Le Goff, Roland
-Martin and Pieyre Le Loher, University of Pau, France, Daniel Peter,
-Jeroen Tromp and the Princeton group of developers, Princeton University,
-USA, and Emanuele Casarotti, INGV Roma, Italy: support for CUBIT meshes
-decomposed by SCOTCH; much faster solver using Michel Deville's inlined
-matrix products.\\
-
-MPI v. 1.4 Dimitri Komatitsch, University of Pau, Qinya Liu and others,
-Caltech, September 2006: better adjoint and kernel calculations, faster
-and better I/Os on very large systems, many small improvements and
-bug fixes.\\
-
-MPI v. 1.3 Dimitri Komatitsch, University of Pau, and Qinya Liu, Caltech,
-July 2005: serial version, regular mesh, adjoint and kernel calculations,
-ParaView support.\\
-
-MPI v. 1.2 Min Chen and Dimitri Komatitsch, Caltech, July 2004: full
-anisotropy, volume movie.\\
-
-MPI v. 1.1 Dimitri Komatitsch, Caltech, October 2002: Zhu's Moho map,
-scaling of $V_{s}$ with depth, Hauksson's regional model, attenuation,
-oceans, movies.\\
-
-MPI v. 1.0 Dimitri Komatitsch, Caltech, USA, May 2002: first MPI version
-based on global code.\\
-
-Dimitri Komatitsch, IPG Paris, France, December 1996: first 3-D solver
-for the CM-5 Connection Machine, parallelized on 128 processors using
-Connection Machine Fortran.\\
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\include{01_introduction}
+
+\include{02_getting_started}
+
+\include{03_mesh_generation}
+
+\include{04_creating_databases}
+
+\include{05_running_the_solver}
+
+\include{06_fault_sources}
+
+\include{07_adjoint_simulations}
+
+\include{08_doing_tomography}
+
+\include{09_noise_simulations}
+
+\include{10_gravity_calculations}
+
+\include{11_graphics}
+
+\include{12_running_scheduler}
+
+\include{13_changing_the_model}
+
+\include{14_post_processing}
+
+\include{15_informations_for_developers}
+
+\include{bug_reports}
+
+\include{notes_and_acknowledgement}
+
+\include{copyright_and_version}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% REFERENCES
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
 \addcontentsline{toc}{chapter}{References} \bibliography{bibliography}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% APPENDIX
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
-\chapter{\label{cha:Coordinates}Reference Frame Convention}
-
-The code uses the following convention for the Cartesian reference
-frame:
-\begin{itemize}
-\item the $x$ axis points East
-\item the $y$ axis points North
-\item the $z$ axis points up
-\end{itemize}
-Note that this convention is different from both the \citet{AkRi80}
-convention and the Harvard Centroid-Moment Tensor (CMT) convention.
-The Aki \& Richards convention is
-\begin{itemize}
-\item the $x$ axis points North
-\item the $y$ axis points East
-\item the $z$ axis points down
-\end{itemize}
-and the Harvard CMT convention is
-\begin{itemize}
-\item the $x$ axis points South
-\item the $y$ axis points East
-\item the $z$ axis points up
-\end{itemize}
-
-\subsection*{Source and receiver locations}
-
-The SPECFEM3D Cartesian software code internally uses Cartesian coordinates.
-The given locations for sources and receiver locations thus may get
-converted. Sources and receiver locations are read in from the \texttt{CMTSOLUTION}
-(or \texttt{FORCESOLUTION}) and \texttt{STATIONS} files. Note that
-e.g. the \texttt{CMTSOLUTION} and \texttt{FORCESOLUTION} files denotes
-the location by \textquotedbl{}longitude/latitude/depth\textquotedbl{}.
-We read in longitude as $x$ coordinate, latitude as $y$ coordinate.
-
-In case the flag \texttt{SUPPRESS\_UTM\_PROJECTION} is set to \texttt{.false.}
-in the main parameter file (see Chapter~\ref{cha:Creating-Distributed-Databases}),
-the $x$/$y$ coordinates have to be given in degrees and are converted
-to Cartesian coordinates using a UTM conversion with the specified
-UTM zone.
-
-The value for depth (given in $km$ in \texttt{CMTSOLUTION} and \texttt{FORCESOLUTION})
-or burial depth (given in $m$ in \texttt{STATIONS}) is evaluated
-with respect to the surface of the mesh at the specified $x$/$y$
-location to find the corresponding $z$ coordinate. It is possible
-to use this depth value directly as $z$ coordinate by changing the
-flag \texttt{USE\_SOURCES\_RECVS\_Z} to \texttt{.true.} in the file
-\texttt{constants.h} located in the \texttt{src/shared/} subdirectory.
-
-\subsection*{Seismogram outputs}
-
-The seismogram output directions are given in Cartesian $x$/$y$/$z$
-directions and not rotated any further. Changing flags in \texttt{constants.h}
-in the \texttt{src/shared/} subdirectory only rotates the seismogram
-outputs if receivers are forced to be located at the surface (\texttt{RECVS\_CAN\_BE\_BURIED\_EXT\_MESH}
-set to \texttt{.false.}) and the normal to the surface at the receiver
-location should be used (\texttt{EXT\_MESH\_RECV\_NORMAL} set to \texttt{.true.})
-as vertical. In this case, the outputs are rotated to have the vertical
-component normal to the surface of the mesh, $x$ and $y$ directions
-are somewhat arbitrary orthogonal directions along the surface.
-
-For the labeling of the seismogram channels, see Appendix~\ref{cha:channel-codes}.
-Additionally, we add labels to the synthetics using the following
-convention: For a receiver station located in an
-\begin{description}
-\item [{elastic domain:}] ~\\
-
-\begin{itemize}
-\item \texttt{semd} for the displacement vector
-\item \texttt{semv} for the velocity vector
-\item \texttt{sema} for the acceleration vector
-\end{itemize}
-\item [{acoustic domain:}] ~\\
- (please note that receiver stations in acoustic domains must be buried.
-This is due to the free surface condition which enforces a zero pressure
-at the surface.)
-
-\begin{itemize}
-\item \texttt{semd} for the displacement vector
-\item \texttt{semv} for the velocity vector
-\item \texttt{sema} for pressure which will be stored in the vertical component
-\texttt{Z}. Note that pressure in the acoustic media is isotropic
-and thus the pressure record would be the same in the other two component
-directions. We therefore use the other two seismogram components to
-store the acoustic potential in component \texttt{X} (or \texttt{N})
-and the first time derivative of the acoustic potential in component
-\texttt{Y} (or \texttt{E}).
-\end{itemize}
-\end{description}
-The seismograms are by default written out in ASCII-format to the
-\texttt{OUTPUT\_FILES/} subdirectory by each parallel process. You
-can change this behavior by changing the following flags in the \texttt{constants.h}
-file located in the \texttt{src/shared/} subdirectory:
-\begin{description}
-\item [{\texttt{SEISMOGRAMS\_BINARY}}] set to \texttt{.true.} to have seismograms
-written out in binary format.
-\item [{\texttt{WRITE\_SEISMOGRAMS\_BY\_MASTER}}] Set to \texttt{.true.}
-to have only the master process writing out seismograms. This can
-be useful on a cluster, where only the master process node has access
-to the output directory.
-\item [{\texttt{USE\_OUTPUT\_FILES\_PATH}}] Set to \texttt{.false.} to
-have the seismograms output to \texttt{LOCAL\_PATH} directory specified
-in the main parameter file \texttt{DATA/Par\_file}.
-\end{description}
-
-\chapter{\label{cha:channel-codes}Channel Codes of Seismograms}
-
-Seismic networks, such as the Global Seismographic Network (GSN),
-generally involve various types of instruments with different bandwidths,
-sampling properties and component configurations. There are standards
-to name channel codes depending on instrument properties. IRIS \texttt{(www.iris.edu)}
-uses SEED format for channel codes, which are represented by three
-letters, such as \texttt{LHN}, \texttt{BHZ}, etc. In older versions
-of the SPECFEM3D Cartesian package, a common format was used for the
-channel codes of all seismograms, which was \texttt{BHE/BHN/BHZ} for
-three components. To avoid confusion when comparison are made to observed
-data, we are now using the FDSN convention (\url{http://www.fdsn.org/})
-for SEM seismograms. In the following, we give a brief explanation
-of the FDSN convention used by IRIS, and how it is adopted in SEM
-seismograms. Please visit \texttt{www.iris.edu/manuals/SEED\_appA.htm}
-for further information.\\
-
-\noindent \texttt{\textbf{Band code:}} The first letter in the channel
-code denotes the band code of seismograms, which depends on the response
-band and the sampling rate of instruments. The list of band codes
-used by IRIS is shown in Figure \ref{fig:IRIS_band_codes}. The sampling
-rate of SEM synthetics is controlled by the resolution of simulations
-rather than instrument properties. However, for consistency, we follow
-the FDSN convention for SEM seismograms governed by their sampling
-rate. For SEM synthetics, we consider band codes for which $dt\leq1$
-s. IRIS also considers the response band of instruments. For instance,
-short-period and broad-band seismograms with the same sampling rate
-correspond to different band codes, such as S and B, respectively.
-In such cases, we consider SEM seismograms as broad band, ignoring
-the corner period ($\geq10$ s) of the response band of instruments
-(note that at these resolutions, the minimum period in the SEM synthetics
-will be less than $10$ s). Accordingly, when you run a simulation
-the band code will be chosen depending on the resolution of the synthetics,
-and channel codes of SEM seismograms will start with either \texttt{L},
-\texttt{M}, \texttt{B}, \texttt{H}, \texttt{C} or \texttt{F}, shown
-by red color in the figure.\\
-
-\begin{figure}[ht]
-\noindent \begin{centering}
-\includegraphics[scale=0.6]{figures/IRIS_band_codes}
-\par\end{centering}
-
-\caption{The band code convention is based on the sampling rate and the response
-band of instruments. Please visit \texttt{www.iris.edu/manuals/SEED\_appA.htm}
-for further information. Grey rows show the relative band-code range
-in SPECFEM3D Cartesian, and the band codes used to name SEM seismograms
-are denoted in red.}
-
-\label{fig:IRIS_band_codes}
-\end{figure}
-
-\noindent \texttt{\textbf{Instrument code:}} The second letter in
-the channel code corresponds to instrument codes, which specify the
-family to which the sensor belongs. For instance, \texttt{H} and \texttt{L}
-are used for high-gain and low-gain seismometers, respectively. The
-instrument code of SEM seismograms will always be \texttt{X}, as assigned
-by FDSN for synthetic seismograms. \\
-
-\noindent \texttt{\textbf{Orientation code:}} The third letter in
-channel codes is an orientation code, which generally describes the
-physical configuration of the components of instrument packages. SPECFEM3D
-Cartesian uses the traditional orientation code \texttt{E/N/Z} (East-West,
-North-South, Vertical) for three components when a UTM projection
-is used. If the UTM conversion is suppressed, i.e. the flag \texttt{SUPPRESS\_UTM\_PROJECTION}
-is set to \texttt{.true.}, then the three components are labelled
-\texttt{X/Y/Z} according to the Cartesian reference frame. \\
-
-\noindent \texttt{\textbf{EXAMPLE:}} The sampling rate is given by
-\texttt{DT} in the main parameter file \texttt{Par\_file} located
-in the \texttt{DATA/} subdirectory. Depending on the resolution of
-your simulations, if you choose a sampling rate greater than $0.01$
-s and less than $1$ s, a seismogram recording displacements on the
-vertical component of a station \texttt{ASBS} (network \texttt{AZ})
-will be named \texttt{AZ.ASBS.MXZ.semd.sac}, whereas it will be \texttt{AZ.ASBS.BXZ.semd.sac},
-if the sampling rate is greater than 0.0125 and less equal to 0.1
-s.
-
-\chapter{\label{cha:Troubleshooting}Troubleshooting}
-
-\section*{FAQ}
-\begin{description}
-\item [{configuration fails:}]~
-
-Examine the log file 'config.log'. It contains detailed informations.
-In many cases, the paths to these specific compiler commands F90,
-CC and MPIF90 will not be correct if `./configure` fails.
-
-Please make sure that you have a working installation of a Fortran
-compiler, a C compiler and an MPI implementation. You should be able
-to compile this little program code:
-\begin{lyxcode}
-program~main~~~\\
-~include~'mpif.h'~~~\\
-~integer,~parameter~::~CUSTOM\_MPI\_TYPE~=~MPI\_REAL~~~\\
-~integer~ier~~~\\
-~call~MPI\_INIT(ier)~~~\\
-~call~MPI\_BARRIER(MPI\_COMM\_WORLD,ier)~~~\\
-~call~MPI\_FINALIZE(ier)~~~\\
-~end
-\end{lyxcode}
-\item [{compilation fails stating:}] ~\\
-
-\begin{lyxcode}
-...~~~\\
-~obj/program\_generate\_databases.o:~In~function~`MAIN\_\_':~~~\\
-~program\_generate\_databases.f90:(.text+0x14):~undefined~reference~to~`\_gfortran\_set\_std'~~~\\
-~...~~~\\
-
-\end{lyxcode}
-
-Make sure you're pointing to the right 'mpif90' wrapper command.
-
-Normally, this message will appear when you are mixing two different
-Fortran compilers. That is, using e.g. gfortran to compile non-MPI
-files and mpif90, wrapper provided for e.g. ifort, to compile MPI-files.
-
-fix: e.g. specify > ./configure FC=gfortran MPIF90=/usr/local/openmpi-gfortran/bin/mpif90
-
-\item [{after executing \texttt{xmeshfem3D} I've got elements with skewness of 81\% percent, what does this mean:}] Look
-at the skewness table printed in the \texttt{output\_mesher.txt} file
-after executing \texttt{xmeshfem3D} for the example given in \texttt{EXAMPLES/meshfem3D\_examples/simple\_model/}:
-
-\begin{lyxcode}
-...~~\\
-~histogram~of~skewness~(0.~good~-~1.~bad):~~~\\
-~0.0000000E+00~-~5.0000001E-02~27648~81.81818~\%~~~\\
-~5.0000001E-02~-~0.1000000~0~0.0000000E+00~\%~~~\\
-~...~~~\\
-
-\end{lyxcode}
-
-The first line means that you have 27,648 elements with a skewness
-value between 0 and 0.05 (which means the element is basically not
-skewed, just plain regular hexahedral element). The total number of
-elements you have in this mesh is (see in the \texttt{output\_mesher.txt}
-file a bit further down):
-\begin{lyxcode}
-...~~\\
-~total~number~of~elements~in~entire~mesh:~33792~~~\\
-~...~~~\\
-
-\end{lyxcode}
-
-which gives you that: 27,648 / 33,792 $\sim$ 81.8 \% of all elements
-are not skewed, i.e. regular elements. a fantastic value :)
-
-The histogram lists for this mesh also some stronger skewed elements,
-for example the worst ones belong to:
-\begin{lyxcode}
-...~~\\
-~0.6000000~-~0.6500000~2048~6.060606~\%~~~\\
-~...~~~\\
-
-\end{lyxcode}
-
-about 6 \% of all elements have distortions with a skewness value
-between 0.6 and 0.65. The skewness values give you a hint of how good
-your mesh is. In an ideal world, you would want to have no distortions,
-just like the 81\% from above. Those elements give you the best approximate
-values by the GLL quadrature used in the spectral-element method.
-However, having weakly distorted elements is still fine and the solutions
-are still accurate enough. So empirically, values up to around 0.7
-are tolerable, above that you should consider remeshing...
-
-To give you an idea why some of the elements are distorted, see the
-following figure \ref{fig:mesh.vp} of the mesh you obtain in the
-example \texttt{EXAMPLES/meshfem3D\_examples/simple\_model/}.
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=0.9\textwidth]{figures/mesh_vp}
-\par\end{centering}
-
-\caption{Paraview visualization using the mesh vtk-files for the example given
-in \texttt{EXAMPLES/meshfem3D\_examples/simple\_model/}.}
-
-\label{fig:mesh.vp}
-\end{figure}
-
-You will see that the mesh contains a doubling layer, where we stitch
-elements together such that the size of two elements will transition
-to the size of one element (very useful to keep the ratio of wavespeed
-/ element\_size about constant). Those elements in this doubling layer
-have higher skewness values and make up those 6 \% in the histogram.
-
-\item [{the code gives following error message "need at least one receiver":}] This
-means that no stations given in the input file \texttt{DATA/STATIONS}
-could be located within the dimensions of the mesh. This can happen
-for example when the mesh was created with the in-house mesher \texttt{xmeshfem3D}
-while using the Universal Transverse Mercator (UTM) projection but
-the simulation with \texttt{xspecfem3D} was suppressing this projection
-from latitude/longitude to x/y/z coordinates.
-
-In such cases, try to change your \texttt{DATA/Par\_file} and set
-e.g.:
-\begin{lyxcode}
-SUPPRESS\_UTM\_PROJECTION~=~.false.~~~\\
-
-\end{lyxcode}
-
-to be the same in \texttt{Mesh\_Par\_file} and \texttt{Par\_file}.
-This flag should be identical when using the in-house mesher \texttt{xmeshfem3D},
-\texttt{xgenerate\_databases} and \texttt{xspecfem3D} together to
-run simulations.
-
-The flag determines if the coordinates you specify for your source
-and station locations are given as lat/lon degrees and must be converted
-to UTM coordinates. As an example, if you use \texttt{.false.} within
-\texttt{Mesh\_Par\_file} then you create a mesh with \texttt{xmeshfem3D}
-using the UTM projection from lat/lon as input format to UTM projected
-coordinates to store the mesh point positions, which is fine. The
-error then may occur if in the \texttt{Par\_file} you have this set
-to \texttt{.true.} so that the \texttt{xgenerate\_databases} and \texttt{xspecfem3D}
-suppress the UTM projection and assume that all coordinates you use
-now for source and receiver locations are given in meters (that is,
-converted) already. So it won't find the specified locations in the
-used mesh. As a solutions, just change the flag in \texttt{Par\_file}
-to be the same as in \texttt{Mesh\_Par\_file} and rerun \texttt{xgenerate\_databases}
-and \texttt{xspecfem3D} to make sure that your simulation works fine.
-
-\item [{I get the following error message "forward simulation became unstable and blew up":}] In
-most cases this means that your time step size \texttt{DT} is chosen
-too big. Look at your files \texttt{output\_mesher.txt} or \texttt{output\_solver.txt}
-created in the folder \texttt{OUTPUT\_FILES}. In these output files,
-find the section:
-
-\begin{lyxcode}
-...~~~\\
-~{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}~~~\\
-~{*}{*}{*}~Verification~of~simulation~parameters~{*}{*}{*}~~~\\
-~{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}~~~\\
-~...~~~\\
-~{*}{*}{*}~Minimum~period~resolved~=~4.308774~~~\\
-~{*}{*}{*}~Maximum~suggested~time~step~=~6.8863556E-02~~~\\
-~...~~~\\
-~...~~~\\
-
-\end{lyxcode}
-
-then change \texttt{DT} in the \texttt{DATA/Par\_file} to be somewhere
-close to the maximum suggested time step. In the example above: \\
- \texttt{DT = 0.05d0} \\
- would (most probably) work fine. It could be also bigger than the
-0.068 s suggested. This depends a bit on the distortions of your mesh
-elements. The more regular they are, the bigger you can choose \texttt{DT}.
-Just play with this value a bit and see when the simulation becomes
-stable ...
-
-\end{description}
-
-\chapter{\label{cha:License}License }
-
-\textbf{GNU GENERAL PUBLIC LICENSE Version 2, June 1991. Copyright
-(C) 1989, 1991 Free Software Foundation, Inc. 59 Temple Place, Suite
-330, Boston, MA 02111-1307 USA} \\
- Everyone is permitted to copy and distribute verbatim copies of this
-license document, but changing it is not allowed.
-
-\section*{Preamble}
-
-The licenses for most software are designed to take away your freedom
-to share and change it. By contrast, the GNU General Public License
-is intended to guarantee your freedom to share and change free software
--- to make sure the software is free for all its users. This General
-Public License applies to most of the Free Software Foundation's software
-and to any other program whose authors commit to using it. (Some other
-Free Software Foundation software is covered by the GNU Library General
-Public License instead.) You can apply it to your programs, too.
-
-When we speak of free software, we are referring to freedom, not price.
-Our General Public Licenses are designed to make sure that you have
-the freedom to distribute copies of free software (and charge for
-this service if you wish), that you receive source code or can get
-it if you want it, that you can change the software or use pieces
-of it in new free programs; and that you know you can do these things.
-
-To protect your rights, we need to make restrictions that forbid anyone
-to deny you these rights or to ask you to surrender the rights. These
-restrictions translate to certain responsibilities for you if you
-distribute copies of the software, or if you modify it.
-
-For example, if you distribute copies of such a program, whether gratis
-or for a fee, you must give the recipients all the rights that you
-have. You must make sure that they, too, receive or can get the source
-code. And you must show them these terms so they know their rights.
-
-We protect your rights with two steps:
-\begin{enumerate}
-\item Copyright the software, and
-\item Offer you this license which gives you legal permission to copy, distribute
-and/or modify the software.
-\end{enumerate}
-Also, for each author's protection and ours, we want to make certain
-that everyone understands that there is no warranty for this free
-software. If the software is modified by someone else and passed on,
-we want its recipients to know that what they have is not the original,
-so that any problems introduced by others will not reflect on the
-original authors' reputations.
-
-Finally, any free program is threatened constantly by software patents.
-We wish to avoid the danger that redistributors of a free program
-will individually obtain patent licenses, in effect making the program
-proprietary. To prevent this, we have made it clear that any patent
-must be licensed for everyone's free use or not licensed at all.
-
-The precise terms and conditions for copying, distribution and modification
-follow.
-
-\section*{GNU GENERAL PUBLIC LICENSE TERMS AND CONDITIONS FOR COPYING, DISTRIBUTION
-AND MODIFICATION }
-\begin{itemize}
-\item 0.This License applies to any program or other work which contains
-a notice placed by the copyright holder saying it may be distributed
-under the terms of this General Public License. The ``Program''
-below refers to any such program or work, and a ``work based on the
-Program'' means either the Program or any derivative work under copyright
-law: that is to say, a work containing the Program or a portion of
-it, either verbatim or with modifications and/or translated into another
-language. (Hereinafter, translation is included without limitation
-in the term ``modification.'') Each licensee is addressed as ``you.''\\
- \\
- Activities other than copying, distribution and modification are
-not covered by this License; they are outside its scope. The act of
-running the Program is not restricted, and the output from the Program
-is covered only if its contents constitute a work based on the Program
-(independent of having been made by running the Program). Whether
-that is true depends on what the Program does.\end{itemize}
-\begin{enumerate}
-\item You may copy and distribute verbatim copies of the Program's source
-code as you receive it, in any medium, provided that you conspicuously
-and appropriately publish on each copy an appropriate copyright notice
-and disclaimer of warranty; keep intact all the notices that refer
-to this License and to the absence of any warranty; and give any other
-recipients of the Program a copy of this License along with the Program.
-
-You may charge a fee for the physical act of transferring a copy,
-and you may at your option offer warranty protection in exchange for
-a fee.
-
-\item You may modify your copy or copies of the Program or any portion of
-it, thus forming a work based on the Program, and copy and distribute
-such modifications or work under the terms of Section 1 above, provided
-that you also meet all of these conditions:
-
-\begin{enumerate}
-\item You must cause the modified files to carry prominent notices stating
-that you changed the files and the date of any change.
-\item You must cause any work that you distribute or publish, that in whole
-or in part contains or is derived from the Program or any part thereof,
-to be licensed as a whole at no charge to all third parties under
-the terms of this License.
-\item If the modified program normally reads commands interactively when
-run, you must cause it, when started running for such interactive
-use in the most ordinary way, to print or display an announcement
-including an appropriate copyright notice and a notice that there
-is no warranty (or else, saying that you provide a warranty) and that
-users may redistribute the program under these conditions, and telling
-the user how to view a copy of this License. (Exception: if the Program
-itself is interactive but does not normally print such an announcement,
-your work based on the Program is not required to print an announcement.)
-\end{enumerate}
-
-These requirements apply to the modified work as a whole. If identifiable
-sections of that work are not derived from the Program, and can be
-reasonably considered independent and separate works in themselves,
-then this License, and its terms, do not apply to those sections when
-you distribute them as separate works. But when you distribute the
-same sections as part of a whole which is a work based on the Program,
-the distribution of the whole must be on the terms of this License,
-whose permissions for other licensees extend to the entire whole,
-and thus to each and every part regardless of who wrote it.
-
-Thus, it is not the intent of this section to claim rights or contest
-your rights to work written entirely by you; rather, the intent is
-to exercise the right to control the distribution of derivative or
-collective works based on the Program.
-
-In addition, mere aggregation of another work not based on the Program
-with the Program (or with a work based on the Program) on a volume
-of a storage or distribution medium does not bring the other work
-under the scope of this License.
-
-\item You may copy and distribute the Program (or a work based on it, under
-Section 2) in object code or executable form under the terms of Sections
-1 and 2 above provided that you also do one of the following:
-
-\begin{enumerate}
-\item Accompany it with the complete corresponding machine-readable source
-code, which must be distributed under the terms of Sections 1 and
-2 above on a medium customarily used for software interchange; or,
-\item Accompany it with a written offer, valid for at least three years,
-to give any third party, for a charge no more than your cost of physically
-performing source distribution, a complete machine-readable copy of
-the corresponding source code, to be distributed under the terms of
-Sections 1 and 2 above on a medium customarily used for software interchange;
-or,
-\item Accompany it with the information you received as to the offer to
-distribute corresponding source code. (This alternative is allowed
-only for noncommercial distribution and only if you received the program
-in object code or executable form with such an offer, in accord with
-Subsection b above.)
-\end{enumerate}
-
-The source code for a work means the preferred form of the work for
-making modifications to it. For an executable work, complete source
-code means all the source code for all modules it contains, plus any
-associated interface definition files, plus the scripts used to control
-compilation and installation of the executable. However, as a special
-exception, the source code distributed need not include anything that
-is normally distributed (in either source or binary form) with the
-major components (compiler, kernel, and so on) of the operating system
-on which the executable runs, unless that component itself accompanies
-the executable.
-
-If distribution of executable or object code is made by offering access
-to copy from a designated place, then offering equivalent access to
-copy the source code from the same place counts as distribution of
-the source code, even though third parties are not compelled to copy
-the source along with the object code.
-
-\item You may not copy, modify, sublicense, or distribute the Program except
-as expressly provided under this License. Any attempt otherwise to
-copy, modify, sublicense or distribute the Program is void, and will
-automatically terminate your rights under this License. However, parties
-who have received copies, or rights, from you under this License will
-not have their licenses terminated so long as such parties remain
-in full compliance.
-\item You are not required to accept this License, since you have not signed
-it. However, nothing else grants you permission to modify or distribute
-the Program or its derivative works. These actions are prohibited
-by law if you do not accept this License. Therefore, by modifying
-or distributing the Program (or any work based on the Program), you
-indicate your acceptance of this License to do so, and all its terms
-and conditions for copying, distributing or modifying the Program
-or works based on it.
-\item Each time you redistribute the Program (or any work based on the Program),
-the recipient automatically receives a license from the original licensor
-to copy, distribute or modify the Program subject to these terms and
-conditions. You may not impose any further restrictions on the recipients'
-exercise of the rights granted herein. You are not responsible for
-enforcing compliance by third parties to this License.
-\item If, as a consequence of a court judgment or allegation of patent infringement
-or for any other reason (not limited to patent issues), conditions
-are imposed on you (whether by court order, agreement or otherwise)
-that contradict the conditions of this License, they do not excuse
-you from the conditions of this License. If you cannot distribute
-so as to satisfy simultaneously your obligations under this License
-and any other pertinent obligations, then as a consequence you may
-not distribute the Program at all. For example, if a patent license
-would not permit royalty-free redistribution of the Program by all
-those who receive copies directly or indirectly through you, then
-the only way you could satisfy both it and this License would be to
-refrain entirely from distribution of the Program.
-
-If any portion of this section is held invalid or unenforceable under
-any particular circumstance, the balance of the section is intended
-to apply and the section as a whole is intended to apply in other
-circumstances.
-
-It is not the purpose of this section to induce you to infringe any
-patents or other property right claims or to contest validity of any
-such claims; this section has the sole purpose of protecting the integrity
-of the free software distribution system, which is implemented by
-public license practices. Many people have made generous contributions
-to the wide range of software distributed through that system in reliance
-on consistent application of that system; it is up to the author/donor
-to decide if he or she is willing to distribute software through any
-other system and a licensee cannot impose that choice.
-
-This section is intended to make thoroughly clear what is believed
-to be a consequence of the rest of this License.
-
-\item If the distribution and/or use of the Program is restricted in certain
-countries either by patents or by copyrighted interfaces, the original
-copyright holder who places the Program under this License may add
-an explicit geographical distribution limitation excluding those countries,
-so that distribution is permitted only in or among countries not thus
-excluded. In such case, this License incorporates the limitation as
-if written in the body of this License.
-\item The Free Software Foundation may publish revised and/or new versions
-of the General Public License from time to time. Such new versions
-will be similar in spirit to the present version, but may differ in
-detail to address new problems or concerns.
-
-Each version is given a distinguishing version number. If the Program
-specifies a version number of this License which applies to it and
-``any later version,'' you have the option of following the terms
-and conditions either of that version or of any later version published
-by the Free Software Foundation. If the Program does not specify a
-version number of this License, you may choose any version ever published
-by the Free Software Foundation.
-
-\item If you wish to incorporate parts of the Program into other free programs
-whose distribution conditions are different, write to the author to
-ask for permission. For software which is copyrighted by the Free
-Software Foundation, write to the Free Software Foundation; we sometimes
-make exceptions for this. Our decision will be guided by the two goals
-of preserving the free status of all derivatives of our free software
-and of promoting the sharing and reuse of software generally.
-\end{enumerate}
-
-\subsection*{NO WARRANTY }
-\begin{itemize}
-\item 11.BECAUSE THE PROGRAM IS LICENSED FREE OF CHARGE, THERE IS NO WARRANTY
-FOR THE PROGRAM, TO THE EXTENT PERMITTED BY APPLICABLE LAW. EXCEPT
-WHEN OTHERWISE STATED IN WRITING THE COPYRIGHT HOLDERS AND/OR OTHER
-PARTIES PROVIDE THE PROGRAM ``AS IS'' WITHOUT WARRANTY OF ANY KIND,
-EITHER EXPRESSED OR IMPLIED, INCLUDING, BUT NOT LIMITED TO, THE IMPLIED
-WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE.
-THE ENTIRE RISK AS TO THE QUALITY AND PERFORMANCE OF THE PROGRAM IS
-WITH YOU. SHOULD THE PROGRAM PROVE DEFECTIVE, YOU ASSUME THE COST
-OF ALL NECESSARY SERVICING, REPAIR OR CORRECTION.
-\item 12.IN NO EVENT UNLESS REQUIRED BY APPLICABLE LAW OR AGREED TO IN WRITING
-WILL ANY COPYRIGHT HOLDER, OR ANY OTHER PARTY WHO MAY MODIFY AND/OR
-REDISTRIBUTE THE PROGRAM AS PERMITTED ABOVE, BE LIABLE TO YOU FOR
-DAMAGES, INCLUDING ANY GENERAL, SPECIAL, INCIDENTAL OR CONSEQUENTIAL
-DAMAGES ARISING OUT OF THE USE OR INABILITY TO USE THE PROGRAM (INCLUDING
-BUT NOT LIMITED TO LOSS OF DATA OR DATA BEING RENDERED INACCURATE
-OR LOSSES SUSTAINED BY YOU OR THIRD PARTIES OR A FAILURE OF THE PROGRAM
-TO OPERATE WITH ANY OTHER PROGRAMS), EVEN IF SUCH HOLDER OR OTHER
-PARTY HAS BEEN ADVISED OF THE POSSIBILITY OF SUCH DAMAGES.
-\end{itemize}
-
-\section*{END OF TERMS AND CONDITIONS }
-
-\subsection*{How to Apply These Terms to Your New Programs}
-
-If you develop a new program, and you want it to be of the greatest
-possible use to the public, the best way to achieve this is to make
-it free software which everyone can redistribute and change under
-these terms.
-
-To do so, attach the following notices to the program. It is safest
-to attach them to the start of each source file to most effectively
-convey the exclusion of warranty; and each file should have at least
-the ``copyright'' line and a pointer to where the full notice is
-found. For example:
-\begin{quote}
-One line to give the program's name and a brief idea of what it does.
-Copyright {\footnotesize $\copyright$ (}year) (name of author)
-
-This program is free software; you can redistribute it and/or modify
-it under the terms of the GNU General Public License as published
-by the Free Software Foundation; either version 2 of the License,
-or (at your option) any later version.
-
-This program is distributed in the hope that it will be useful, but
-WITHOUT ANY WARRANTY; without even the implied warranty of MERCHANTABILITY
-or FITNESS FOR A PARTICULAR PURPOSE. See the GNU General Public License
-for more details.
-
-You should have received a copy of the GNU General Public License
-along with this program; if not, write to the Free Software Foundation,
-Inc., 59 Temple Place, Suite 330, Boston, MA 02111-1307 USA
-\end{quote}
-Also add information on how to contact you by electronic and paper
-mail.
-
-If the program is interactive, make it output a short notice like
-this when it starts in an interactive mode:
-\begin{quote}
-Gnomovision version 69, Copyright $\copyright$ year name of author
-Gnomovision comes with ABSOLUTELY NO WARRANTY; for details type `show
-w'. This is free software, and you are welcome to redistribute it
-under certain conditions; type `show c' for details.
-\end{quote}
-The hypothetical commands `show w' and `show c' should show the appropriate
-parts of the General Public License. Of course, the commands you use
-may be called something other than `show w' and `show c'; they could
-even be mouse-clicks or menu items -- whatever suits your program.
-
-You should also get your employer (if you work as a programmer) or
-your school, if any, to sign a ``copyright disclaimer'' for the
-program, if necessary. Here is a sample; alter the names:
-\begin{quote}
-Yoyodyne, Inc., hereby disclaims all copyright interest in the program
-`Gnomovision' (which makes passes at compilers) written by James Hacker.
-
-(signature of Ty Coon)\\
- 1 April 1989 \\
- Ty Coon, President of Vice
-\end{quote}
-This General Public License does not permit incorporating your program
-into proprietary programs. If your program is a subroutine library,
-you may consider it more useful to permit linking proprietary applications
-with the library. If this is what you want to do, use the GNU Library
-General Public License instead of this License.
+\include{A_reference_frame}
+
+\include{B_channel_codes}
+
+\include{C_troubleshooting}
+
+\include{D_license}
+
 \end{document}

### 2015-04-24 doc/USER_MANUAL/notes_and_acknowledgement.tex
updates user manual to use facilitate the generation of a markdown version to the github wiki

--- 2012-09-28 doc/USER_MANUAL/manual_SPECFEM3D.tex
@@
-%% LyX 1.5.1 created this file.  For more info, see http://www.lyx.org/.
-%% Do not edit unless you really know what you are doing.
-\documentclass[oneside,english,onecolumn]{book}
-\usepackage[T1]{fontenc}
-\usepackage[latin1]{inputenc}
-\usepackage{geometry}
-\geometry{letterpaper,tmargin=1in,bmargin=1in,lmargin=1in,rmargin=1in}
-\setcounter{secnumdepth}{3}
-\setcounter{tocdepth}{3}
-\usepackage{float}
-\usepackage{textcomp}
-\usepackage{amsmath}
+\addcontentsline{toc}{chapter}{Notes \& Acknowledgments}
 
-% figures
-\usepackage[pdftex]{graphicx}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-% we are running pdflatex, so convert .eps files to .pdf
-\usepackage{epstopdf}
-
-\usepackage{amssymb}
-\IfFileExists{url.sty}{\usepackage{url}}
-                      {\newcommand{\url}{\texttt}}
-\usepackage[authoryear]{natbib}
-
-\makeatletter
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% LyX specific LaTeX commands.
-\newcommand{\noun}[1]{\textsc{#1}}
-%% A simple dot to overcome graphicx limitations
-\newcommand{\lyxdot}{.}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Textclass specific LaTeX commands.
-\newenvironment{lyxcode}
-{\begin{list}{}{
-\setlength{\rightmargin}{\leftmargin}
-\setlength{\listparindent}{0pt}% needed for AMS classes
-\raggedright
-\setlength{\itemsep}{0pt}
-\setlength{\parsep}{0pt}
-\normalfont\ttfamily}%
- \item[]}
-{\end{list}}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% User specified LaTeX commands.
-%\renewcommand{\baselinestretch}{1.5}
-
-% figures
-\usepackage[dvips]{epsfig}
-
-\usepackage{wrapfig}
-
-% fonts
-\usepackage{times}
-
-% hyperlinks to sections and references
-\usepackage[pdftex,bookmarks=true,bookmarksnumbered=true,pdfpagemode=None,pdfstartview=FitH,pdfpagelayout=SinglePage,pdfborder={0 0 0}]{hyperref}
-
-\newcommand{\urlwithparentheses}[1]{(\url{#1})}
-
-% biblio GJI
-\bibliographystyle{abbrvnat}
-
-\newcommand{\toall}[1]{\textbf{*** All: #1 ***}}
-\newcommand{\tojeroen}[1]{\textbf{*** Jeroen: #1 ***}}
-\newcommand{\tobrian}[1]{\textbf{*** Brian: #1 ***}}
-\newcommand{\tovala}[1]{\textbf{*** Vala: #1 ***}}
-\newcommand{\tovalabrian}[1]{\textbf{*** Vala \& Brian: #1 ***}}
-\newcommand{\tovalaqinya}[1]{\textbf{*** Vala \& Qinya: #1 ***}}
-\newcommand{\toqinya}[1]{\textbf{*** Qinya: #1 ***}}
-\newcommand{\tomin}[1]{\textbf{*** Min: #1 ***}}
-\newcommand{\toalessia}[1]{\textbf{*** Alessia: #1 ***}}
-\newcommand{\todimitri}[1]{\textbf{*** Dimitri: #1 ***}}
-
-\newcommand{\nexxi}{\mbox{\texttt{NEX\_XI\/}}}
-\newcommand{\nexeta}{\mbox{\texttt{NEX\_ETA\/}}}
-\newcommand{\nprocxi}{\mbox{\texttt{NPROC\_XI\/}}}
-\newcommand{\nproceta}{\mbox{\texttt{NPROC\_ETA\/}}}
-\newcommand{\nchunks}{\mbox{\texttt{NCHUNKS\/}}}
-\newcommand{\eg}{e.g.,}
-
-% colors for the text
-\usepackage[dvipsnames,usenames]{color}
-\newcommand{\red}[1]{\textcolor{Red}{#1}}
-
-\usepackage{babel}
-\makeatother
-
-\begin{document}
-\begin{center}
-\thispagestyle{empty}%
-\vspace*{-1.8truecm}
-\noindent\makebox[\textwidth]{%
-\noindent\includegraphics[width=0.83\paperwidth]{figures/specfem_3d-cover.pdf}
-}
-\end{center}
-%
-\title{\textbf{SPECFEM3D}\\
-\textbf{User Manual}}
-
-\author{$\copyright$ Princeton University (USA) and\\
-CNRS / INRIA / University of Pau (France)\\
-Version 2.1}
-
-\date{\noindent \today}
-
-\maketitle
-
-\section*{Authors}
-The SPECFEM3D package was first developed by Dimitri Komatitsch and Jeroen Tromp at Harvard University and Caltech, USA, starting in 1998, based on earlier work by Dimitri Komatitsch and Jean-Pierre Vilotte at Institut de Physique du Globe (IPGP) in Paris, France from 1995 to 1997.\\
-
-Since then it has been developed and maintained by a development team: in alphabetical order,
-Piero Basini,
-C\'eline Blitz,
-Ebru Bozda\u{g},
-Emanuele Casarotti,
-Joseph Charles,
-Min Chen,
-Vala Hj\"orleifsd\'ottir,
-Sue Kientz,
-Dimitri Komatitsch,
-Jes\'us Labarta,
-Nicolas Le Goff,
-Pieyre Le Loher,
-Qinya Liu,
-Yang Luo,
-Alessia Maggi,
-Federica Magnoni,
-Roland Martin,
-Dennis McRitchie,
-Matthias Meschede,
-Peter Messmer,
-David Mich\'ea,
-Tarje Nissen-Meyer,
-Daniel Peter,
-Max Rietmann,
-Brian Savage,
-Bernhard Schuberth,
-Anne Sieminski,
-Leif Strand,
-Carl Tape,
-Jeroen Tromp,
-Zhinan Xie,
-Hejun Zhu.\\
-%\toall{add full list of developers here and update the AUTHORS file accordingly}
-
-The cover graphic of the manual was created
-by Santiago Lombeyda from Caltech's Center for Advanced Computing Research (CACR) \urlwithparentheses{http://www.cacr.caltech.edu}. \\
-
-\newpage{}
-
-\tableofcontents{}
-
-\chapter{Introduction}
-
-The software package SPECFEM3D simulates seismic
-wave propagation at the local or regional scale
-based upon the spectral-element method (SEM).
-The SEM is a continuous Galerkin technique \citep{TrKoLi08,PeKoLuMaLeCaLeMaLiBlNiBaTr11},
-which can easily be made discontinous \citep{BeMaPa94,Ch00,KoWoHu02,ChCaVi03,LaWaBe05,Kop06,WiStBuGh10,AcKo11};
-it is then close to a particular case of the discontinuous Galerkin technique \citep{ReHi73,Arn82,FaRi99,HuHuRa99,CoKaSh00,GiHeWa02,RiWh03,MoRi05,GrScSc06,AiMoMu06,BeLaPi06,DuKa06,DeSeWh08,PuAmKa09,WiStBuGh10,DeSe10,EtChViGl10}, with optimized efficiency because of its tensorized basis functions \citep{WiStBuGh10,AcKo11}.
-In particular, it can accurately handle very distorted mesh elements \citep{OlSe11}.\\
-
-It has very good accuracy and convergence properties \citep{MaPa89,SePr94,DeFiMu02,Coh02,DeSe07,SeOl08}.
-The spectral element approach admits spectral rates of convergence and allows exploiting $hp$-convergence schemes.
-It is also very well suited to parallel implementation on very large supercomputers \citep{KoTsChTr03,TsKoChTr03,KoLaMi08a,CaKoLaTiMiLeSnTr08,KoViCh10} as well as on clusters of GPU accelerating graphics cards \citep{KoMiEr09,KoErGoMi10,Kom11}.
-Tensor products inside each element can be optimized to reach very high efficiency \citep{DeFiMu02}, and mesh point and element numbering can be optimized to reduce processor cache misses and improve cache reuse \citep{KoLaMi08a}.
-The SEM can also handle triangular (in 2D) or tetrahedral (in 3D) elements \citep{WinBoyd96,TaWi00,KoMaTrTaWi01,Coh02,MeViSa06} as well as mixed meshes, although with increased cost and reduced accuracy in these elements, as in the discontinuous Galerkin method.\\
-
-Note that in many geological models in the context of seismic wave propagation studies
-(except for instance for fault dynamic rupture studies, in which very high frequencies or supershear rupture need to be modeled near the fault, see e.g.
-\cite{BeGlCrViPi07,BeGlCrVi09,PuAmKa09,TaCrEtViBeSa10})
-a continuous formulation is sufficient because material property contrasts are not drastic and thus
-conforming mesh doubling bricks can efficiently handle mesh size variations \citep{KoTr02a,KoLiTrSuStSh04,LeChLiKoHuTr08,LeChKoHuTr09,LeKoHuTr09}.\\
-
-For a detailed introduction to the SEM as applied to regional
-seismic wave propagation, please consult \citet{PeKoLuMaLeCaLeMaLiBlNiBaTr11,TrKoLi08,KoVi98,KoTr99,ChKoViCaVaFe07} and
-in particular \citet{LeKoHuTr09,LeChKoHuTr09,LeChLiKoHuTr08,GoAmTaCaSmSaMaKo09,WiKoScTr04,KoLiTrSuStSh04}.
-A detailed theoretical analysis of the dispersion
-and stability properties of the SEM is available in \citet{Coh02}, \citet{DeSe07}, \citet{SeOl07} and \citet{SeOl08}.\\
-
-Effects due to lateral variations in compressional-wave speed, shear-wave
-speed, density, a 3D crustal model, topography and bathymetry are
-included. The package can accommodate
-full 21-parameter anisotropy (see~\citet{ChTr07}) as well as lateral
-variations in attenuation \citep{SaKoTr10}. Adjoint capabilities and finite-frequency
-kernel simulations are included \citep{TrKoLi08,PeKoLuMaLeCaLeMaLiBlNiBaTr11,LiTr06,FiIgBuKe09,ViOp09}.\\
-
-The SEM was originally developed in computational fluid dynamics \citep{Pat84,MaPa89}
-and has been successfully adapted to address problems in seismic wave propagation.
-Early seismic wave propagation applications of the SEM, utilizing Legendre basis functions and a
-perfectly diagonal mass matrix, include \cite{CoJoTo93}, \cite{Kom97},
-\cite{FaMaPaQu97}, \cite{CaGa97}, \cite{KoVi98} and \cite{KoTr99},
-whereas applications involving Chebyshev basis functions and a nondiagonal mass matrix
-include \cite{SePr94}, \cite{PrCaSe94} and \cite{SePrPr95}.\\
-
-All SPECFEM3D software is written in Fortran90 with full portability
-in mind, and conforms strictly to the Fortran95 standard. It uses
-no obsolete or obsolescent features of Fortran77. The package uses
-parallel programming based upon the Message Passing Interface (MPI)
-\citep{GrLuSk94,Pac97}.\\
-
-SPECFEM3D won the Gordon Bell award for best performance at the SuperComputing~2003
-conference in Phoenix, Arizona (USA) (see \cite{KoTsChTr03}
-and \url{www.sc-conference.org/sc2003/nr_finalaward.html}).
-It was a finalist again in 2008 for a run at 0.16 petaflops (sustained) on 149,784 processors of the `Jaguar' Cray XT5 system at Oak Ridge National Laboratories (USA) \citep{CaKoLaTiMiLeSnTr08}.
-It also won the BULL Joseph Fourier supercomputing award in 2010.\\
-
-This new release of the code (version 2.1) includes support for GPU graphics card acceleration \citep{KoMiEr09,KoGoErMi10,KoErGoMi10,MiKo10,Kom11}.\\
-
-The next release of the code will include support Convolution or Auxiliary Differential Equation Perfectly Matched absorbing Layers
-(C-PML or ADE-PML) \citep{KoMa07,MaKoEz08,MaKoGe08,MaKo09,MaKoGeBr10}.
-It will also use the PT-SCOTCH parallel library for mesh partitioning.\\
-
-SPECFEM3D includes coupled fluid-solid domains and
-adjoint capabilities, which enables one to address seismological inverse problems, but for linear rheologies only so far.
-To accommodate visco-plastic or non-linear rheologies, the reader can refer to
-the \href{http://geoelse.stru.polimi.it/}{GeoELSE} software package \citep{CaGa97,StPaIg09,ChMoTsBaKrKaStKr10}.
-
-\section{Citation}
-
-If you use SPECFEM3D for your own research, please cite at least one
-of the following articles:
-
-\begin{description}
-
-\item [Numerical simulations in general] ~\\
-Forward and adjoint simulations are described in detail in \cite{TrKoLi08,PeKoLuMaLeCaLeMaLiBlNiBaTr11,VaCaSaKoVi99,KoMiEr09,KoErGoMi10,KoGoErMi10,ChKoViCaVaFe07,MaKoDi09,KoViCh10,CaKoLaTiMiLeSnTr08,TrKoHjLiZhPeBoMcFrTrHu10,KoRiTr02,KoTr02a,KoTr02b,KoTr99} or \cite{KoVi98}.
-Additional aspects dealing with adjoint simulations are described in
-\cite{TrTaLi05,LiTr06,TrKoLi08,LiTr08,TrKoHjLiZhPeBoMcFrTrHu10,PeKoLuMaLeCaLeMaLiBlNiBaTr11}.
-Domain decomposition is explained in detail in \cite{MaKoBlLe08}, and excellent scaling up to 150,000 processor cores is shown for instance in
-\cite{CaKoLaTiMiLeSnTr08,KoLaMi08a,MaKoBlLe08,KoErGoMi10,KoGoErMi10,Kom11},
-
-\item [GPU computing] ~\\
-Computing on GPU graphics cards for acoustic or seismic wave propagation applications
-is described in detail in \cite{KoMiEr09,KoGoErMi10,KoErGoMi10,MiKo10,Kom11}.
-
-\end{description}
-
-\noindent
-If you use this new version 2.1, which has non blocking MPI for much better performance for medium or large runs, please cite at least one of these six articles,
-in which results of non blocking MPI runs are presented: \cite{PeKoLuMaLeCaLeMaLiBlNiBaTr11,KoGoErMi10,KoErGoMi10,KoViCh10,Kom11,CaKoLaTiMiLeSnTr08,MaKoBlLe08}.\\
-
-\noindent
-If you work on geophysical applications, you may be interested in citing some of these application articles as well, among others:
-
-\begin{description}
-
-\item [Southern California simulations] ~\\
-\cite{KoLiTrSuStSh04,KrJiKoTr06a,KrJiKoTr06b}.
-
-If you use the 3D southern California model, please cite
-\citet{SuSh03} (Los Angeles model),
-\citet{lovelyetal06} (Salton Trough), and
-\citet{hauksson2000} (southern California).
-The Moho map was determined by
-\citet{zhu&kanamori2000}.
-The 1D SoCal model was developed by \citet{DrHe90}.
-
-\item [Anisotropy] ~\\
-\cite{ChTr07,JiTsKoTr05,ChFaKo04,FaChKo04,RiRiKoTrHe02,TrKo00}.
-
-\item [Attenuation] ~\\
-\cite{SaKoTr10,KoTr02a,KoTr99}.
-
-\item [Topography] ~\\
-\cite{LeKoHuTr09,LeChKoHuTr09,LeChLiKoHuTr08,GoAmTaCaSmSaMaKo09,WiKoScTr04}.
-
-\end{description}
-
-The corresponding Bib\TeX{} entries may be found
-in file \texttt{doc/USER\_MANUAL/bibliography.bib}.
-
-\section{Support}
-
-This material is based upon work supported by the USA National Science
-Foundation under Grants No. EAR-0406751 and EAR-0711177, by the French
-CNRS, French INRIA Sud-Ouest MAGIQUE-3D, French ANR NUMASIS under
-Grant No. ANR-05-CIGC-002, and European FP6 Marie Curie International
-Reintegration Grant No. MIRG-CT-2005-017461.
-Any opinions, findings, and conclusions or recommendations expressed in this material are
-those of the authors and do not necessarily reflect the views of the
-USA National Science Foundation, CNRS, INRIA, ANR or the European
-Marie Curie program.
-
-\chapter{\label{cha:Getting-Started}Getting Started}
-
-The SPECFEM3D software package comes in a gzipped tar ball. In the
-directory in which you want to install the package, type
-\begin{lyxcode}
-{\small tar~-zxvf~SPECFEM3D\_V2.1.0.tar.gz}{\small \par}
-\end{lyxcode}
-The directory \texttt{\small SPECFEM3D/} will then contain
-the source code.\\
-
-We recommend that you add {\texttt{ulimit -S -s unlimited}} to your {\texttt{.bash\_profile}} file and/or {\texttt{limit stacksize unlimited }} to your {\texttt{.cshrc}} file to suppress any potential limit to the size of the Unix stack.\\
-
-Then, to configure the software for your system, run the
-\texttt{configure} shell script. This script will attempt to guess
-the appropriate configuration values for your system. However, at
-a minimum, it is recommended that you explicitly specify the appropriate
-command names for your Fortran90 compiler and MPI package:
-
-\begin{lyxcode}
-./configure~FC=ifort~MPIFC=mpif90
-\end{lyxcode}
-
-Before running the \texttt{configure} script, you should probably edit file \texttt{flags.guess} to make sure that it contains the best compiler options for your system.
-Known issues or things to check are:
-
-\begin{description}
-\item [{\texttt{Intel ifort compiler}}] See if you need to add \texttt{-assume byterecl} for your machine.
-\item [{\texttt{IBM compiler}}] See if you need to add \texttt{-qsave} or \texttt{-qnosave} for your machine.
-\item [{\texttt{Mac OS}}] You will probably need to install \texttt{XCODE}.
-In addition, the \texttt{clock\_gettime} routine, which is used by the \texttt{SCOTCH} library that we use, does not exist in Mac OS.
-You will need to replace it with \texttt{clock\_get\_time} if you want to use \texttt{SCOTCH}.
-\end{description}
-
-When compiling on an IBM machine with the \texttt{xlf} and \texttt{xlc} compilers, we suggest running the \texttt{configure} script
-with the following options:\\
-
-\noindent\texttt{./configure FC=xlf90\_r MPIFC=mpif90 CC=xlc\_r CFLAGS="-O3 -q64" FCFLAGS="-O3 -q64" -with-scotch-dir=...}.\\
-
-On SGI systems, \texttt{flags.guess} automatically informs \texttt{configure}
-to insert `\texttt{`TRAP\_FPE=OFF}'' into the generated \texttt{Makefile}
-in order to turn underflow trapping off.\\
-
-If you run very large meshes on a relatively small number
-of processors, the memory size needed on each processor might become
-greater than 2 gigabytes, which is the upper limit for 32-bit addressing.
-In this case, on some compilers you may need to add \texttt{``-mcmodel=medium}'' or \texttt{``-mcmodel=medium -shared-intel}''
-to the configure options of CFLAGS, FCFLAGS and LDFLAGS otherwise the compiler will display an error
-message (for instance \texttt{``relocation truncated to fit: R\_X86\_64\_PC32 against .bss''} or something similar);
-on an IBM machine with the \texttt{xlf} and \texttt{xlc} compilers, using \texttt{-q64} is usually sufficient.\\
-
-The SPECFEM3D software package relies on the SCOTCH library to partition meshes created with CUBIT.
-Note that we use CUBIT to create meshes of hexahedra, but other packages can be used as well, for instance GiD from \url{http://gid.cimne.upc.es} or Gmsh from \url{http://geuz.org/gmsh} \citep{GeRe09}.
-Even mesh creation packages that generate tetrahedra, for instance TetGen from \url{http://tetgen.berlios.de}, can be used because each tetrahedron can then easily be decomposed into four hexahedra as shown in the picture of the TetGen logo at \url{http://tetgen.berlios.de/figs/Delaunay-Voronoi-3D.gif}; while this approach does not generate hexahedra of optimal quality, it can ease mesh creation in some situations and it has been shown that the spectral-element method can very accurately handle distorted mesh elements \citep{OlSe11}.
-
-The SCOTCH library \citep{PeRo96}
-provides efficent static mapping, graph and mesh partitioning routines. SCOTCH is a free software package developed by
-Fran\c {c}ois Pellegrini et al.
-from LaBRI and INRIA in Bordeaux, France, downloadable from the web page \url{https://gforge.inria.fr/projects/scotch/}.
-It is more recent than METIS, actively maintained and performs better in many cases.
-A recent version of its source code is provided in directory \texttt{src/decompose\_mesh\_SCOTCH/scotch\_5.1.10b}.
-In case no SCOTCH libraries can be found on the system, the configuration will bundle this version for compilation.
-The path to an existing SCOTCH installation can to be set explicitly with the option \texttt{-{}-with-scotch-dir}. Just as an example:
-\begin{lyxcode}
-./configure~FC=ifort~MPIFC=mpif90~-{}-with-scotch-dir=/opt/scotch
-\end{lyxcode}
-
-If you use the Intel ifort compiler to compile the code, we recommend that you use the Intel icc C compiler to compile Scotch, i.e., use:
-\begin{lyxcode}
-./configure~CC=icc~FC=ifort~MPIFC=mpif90
-\end{lyxcode}
-
-To compile a serial version of the code for small meshes that fits
-on one compute node and can therefore be run serially, run \texttt{configure}
-with the \texttt{-{}-without-mpi} option to suppress all calls to
-MPI.
-
-A summary of the most important configuration variables follows.
-
-\begin{description}
-\item [{\texttt{F90}}] Path to the Fortran90 compiler.
-\item [{\texttt{MPIF90}}] Path to MPI Fortran90.
-\item [{\texttt{MPI\_FLAGS}}] Some systems require this flag to link to
-MPI libraries.
-\item [{\texttt{FLAGS\_CHECK}}] Compiler flag for non-critical subroutines.
-\item [{\texttt{FLAGS\_NO\_CHECK}}] Compiler flag for creating fast, production-run
-code for critical subroutines.
-\end{description}
-The configuration script automatically creates for each executable a corresponding
-\texttt{Makefile} in the \texttt{src/} subdirectoy.
-The \texttt{Makefile} contains a number of suggested entries for various
-compilers, e.g., Portland, Intel, Absoft, NAG, and Lahey. The software
-has run on a wide variety of compute platforms, e.g., various PC clusters
-and machines from Sun, SGI, IBM, Compaq, and NEC. Select the compiler
-you wish to use on your system and choose the related optimization
-flags. Note that the default flags in the \texttt{Makefile} are undoubtedly
-not optimal for your system, so we encourage you to experiment with
-these flags and to solicit advice from your systems administrator.
-Selecting the right compiler and optimization flags can make a tremendous
-difference in terms of performance. We welcome feedback on your experience
-with various compilers and flags.
-
-Now that you have set the compiler information, you need to select
-a number of flags in the \texttt{constants.h} file depending on your
-system:
-
-\begin{description}
-\item [{\texttt{LOCAL\_PATH\_IS\_ALSO\_GLOBAL}}] Set to \texttt{.false.}
-on most cluster applications. For reasons of speed, the (parallel)
-distributed database generator typically writes a (parallel) database for the solver on the
-local disks of the compute nodes. Some systems have no local disks,
-e.g., BlueGene or the Earth Simulator, and other systems have a fast
-parallel file system, in which case this flag should be set to \texttt{.true.}.
-Note that this flag is not used by the database generator or the solver; it is
-only used for some of the post-processing.
-\end{description}
-The package can run either
-in single or in double precision mode. The default is single precision
-because for almost all calculations performed using the spectral-element method
-using single precision is sufficient and gives the same results (i.e. the same seismograms);
-and the single precision code is faster and requires exactly half as much memory.
-Select your preference by selecting the appropriate setting in the \texttt{constants.h} file:
-
-\begin{description}
-\item [{\texttt{CUSTOM\_REAL}}] Set to \texttt{SIZE\_REAL} for single precision
-and \texttt{SIZE\_DOUBLE} for double precision.
-\end{description}
-In the \texttt{precision.h} file:
-
-\begin{description}
-\item [{\texttt{CUSTOM\_MPI\_TYPE}}] Set to \texttt{MPI\_REAL} for single
-precision and \texttt{MPI\_DOUBLE\_PRECISION} for double precision.
-\end{description}
-On many current processors (e.g., Intel, AMD, IBM Power), single precision calculations
-are significantly faster; the difference can typically be 10\%
-to 25\%. It is therefore better to use single precision.
-What you can do once for the physical problem you want to study is run the same calculation in single precision
-and in double precision on your system and compare the seismograms.
-If they are identical (and in most cases they will), you can select single precision for your future runs.\\
-
-\section{Adding OpenMP support in addition to MPI}
-
-OpenMP support can be enabled in addition to MPI. However, in many cases performance will not improve
-because our pure MPI implementation is already heavily optimized and thus the resulting code will in fact
-be slightly slower. A possible exception could be IBM BlueGene-type architectures.
-
-To enable OpenMP, uncomment the OpenMP compiler option in two lines in file \texttt{src/specfem3D/Makefile.in}
-(before running \texttt{configure}) and also uncomment the \texttt{\#define USE\_OPENMP} statement in file\\
-\texttt{src/specfem3D/specfem3D.F90}.
-
-The DO-loop using OpenMP threads has a SCHEDULE property. The OMP\_SCHEDULE
-environment variable can set the scheduling policy of that DO-loop.
-Tests performed by Marcin Zielinski at SARA (The Netherlands) showed that often
-the best scheduling policy is DYNAMIC with the size of the chunk equal to the number of
-OpenMP threads, but most preferably being twice as the number of
-OpenMP threads (thus chunk size = 8 for 4 OpenMP threads etc).
-If OMP\_SCHEDULE is not set or is empty, the DO-loop will assume generic
-scheduling policy, which will slow down the job quite a bit.
-
-\section{Visualizing the subroutine calling tree of the source code}
-
-Packages such as \texttt{doxywizard} can be used to visualize the subroutine calling tree of the source code.
-\texttt{Doxywizard} is a GUI front-end for configuring and running \texttt{doxygen}.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\chapter{\label{cha:Mesh-Generation}Mesh Generation}
-
-The first step in running a spectral-element simulation consists of constructing a high-quality mesh for the region under consideration.
-We provide two possibilities to do so: (1) relying on the external, hexahedral mesher CUBIT, or (2) using the provided, internal mesher \texttt{xmeshfem3D}.
-In the following, we explain these two approaches.
-
-\section{\label{cha:Running-the-Mesher-CUBIT}Meshing with \texttt{CUBIT}}
-
-CUBIT is a meshing tool suite for the creation of finite-element meshes for arbitrarily shaped models.
-It has been developed and maintained at Sandia National Laboratories and can be purchased for a small academic institutional fee at \url{http://cubit.sandia.gov}.
-Our experience showed that using CUBIT greatly facilitates and speeds up the generation and preparation of hexahedral, conforming meshes for a variety of geophysical models with increasing complexity.
-
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=.6\textwidth]{figures/mount-cubit.jpg}
-\par\end{centering}
-
-\caption{Example of the graphical user interface of CUBIT. The hexahedral mesh shown in the main display consists of a hexahedral discretization of a single volume with topography.}
-
-\label{fig:mount.cubit}
-\end{figure}
-
-The basic steps in creating a load-balanced, partitioned mesh with CUBIT are:
-\begin{description}
-\item [1.] setting up a hexahedral mesh with CUBIT,
-\item [2.] exporting the CUBIT mesh into a SPECFEM3D file format and
-\item [3.] partitioning the SPECFEM3D mesh files for a chosen number of cores.
-\end{description}
-
-Examples are provided in the SPECFEM3D package in the subdirectory \texttt{examples/}.
-We strongly encourage you to contribute your own example to this package by contacting the CIG Computational Seismology Mailing List  \urlwithparentheses{cig-seismo@geodynamics.org}.
-
-\subsection{Creating the Mesh with CUBIT}
-
-For the installation and handling of the CUBIT meshing tool suite, please refer to the CUBIT user manual and documentation.
-In order to give you a basic understanding of how to use CUBIT for our purposes, examples are provided in the SPECFEM3D package in the subdirectory \texttt{examples/}:
-\begin{description}
-\item [{\texttt{homogeneous\_halfspace}}] Creates a single block model and assigns elastic material parameters.
-\item [{\texttt{layered\_halfspace}}]  Combines two different, elastic material volumes and creates a refinement layer between the two.
-This example can be compared for validation against the solutions provided in subdirectory ~\\
-\texttt{VALIDATION\_3D\_SEM\_SIMPLER\_LAYER\_SOURCE\_DEPTH/}.
-\item [{\texttt{waterlayered\_halfspace}}] Combines an acoustic and elastic material volume as in a schematic marine survey example.
-\item [{\texttt{tomographic\_model}}] Creates a single block model whose material properties will have to be read in from a tomographic model file during the databases creation by \texttt{xgenerate\_databases}.
-\end{description}
-
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=.45\textwidth]{figures/example-homogeneous.jpg}
-\includegraphics[width=.45\textwidth]{figures/example-2layers.jpg} \\
-\includegraphics[width=.45\textwidth]{figures/example-water.jpg}
-\includegraphics[width=.45\textwidth]{figures/example-tomo.jpg}
-\par\end{centering}
-
-\caption{Screenshots of the CUBIT examples provided in subdirectory  \texttt{examples/}: homogeneous halfspace (top-left), layered halfspace (top-right), water layered halfspace (bottom-left) and tomographic model (bottom-right).}
-
-\label{fig:examples.cubit}
-\end{figure}
-
-In each example subdirectory you will find a \texttt{README} file, which explains in a step-by-step tutorial the workflow for the example.
-Please feel free to contribute your own example to this package by contacting the CIG Computational Seismology Mailing List  \urlwithparentheses{cig-seismo@geodynamics.org}.
-
-\subsection{Exporting the Mesh with \texttt{cubit2specfem3d.py} }
-\label{subsec:Exporting-the-Mesh}
-
-Once the geometric model volumes in CUBIT are meshed, you prepare the model for exportation with the definition of material blocks and boundary surfaces.
-Thus, prior to exporting the mesh, you need to define blocks specifying the materials and absorbing boundaries in CUBIT.
-This process could be done automatically using the script \texttt{boundary\_definition.py} if the mesh meets some conditions or manually, following the block convention:
-
-\begin{description}
-\item [material\_name] Each material should have a specific block defined by a unique name.
-The name convention of the material is to start with either 'elastic' or 'acoustic'.
-It must be then followed by a unique identifier, e.g.
-'elastic 1', 'elastic 2', etc.
-The additional attributes to the block define the material description.
-
-For an elastic material:
-\begin{description}
-\item [material\_id] An integer value which is unique for this material.
-\item [Vp] P-wave speed of the material (given in m/s).
-\item [Vs] S-wave speed of the material (given in m/s).
-\item [rho] density of the material (given in kg/m$^3$).
-\item [Q] quality factor to use in case of a simulation with attenuation turned on.
-It should be between 1 and 9000.
-In case no attenuation information is available, it can be set to zero.
-Please note that your Vp- and Vs-speeds are given for a reference frequency.
-To change this reference frequency, you change the value of
-\texttt{ATTENUATION\_f0\_REFERENCE} in the main constants file \texttt{constants.h} found in subdirectory \texttt{src/shared/}.
-\item [anisotropic\_flag] Flag describing the anisotropic model to use in case an anisotropic simulation should be conducted.
-See the file \texttt{model\_aniso.f90} in subdirectory \texttt{src/generate\_databases/} for an implementation of the anisotropic models.
-In case no anisotropy is available, it can be set to zero.
-\end{description}
-Note that this material block has to be defined using all the volumes which belong to this elastic material.
-For volumes belonging to another, different material, you will need to define a new material block.
-
-For an acoustic material:
-\begin{description}
-\item [material\_id] An integer value which is unique for this material.
-\item [Vp] P-wave speed of the material (given in m/s).
-\item [0] S-wave speed of the material is ignored.
-\item [rho] density of the material (given in kg/m$^3$).
-\end{description}
-
-\item [face\_topo] Block definition for the surface which defines the free surface (which can have topography).
-The name of this block must be 'face\_topo', the block has to be defined using all the surfaces which constitute the complete free surface of the model.
-
-\item [face\_abs\_xmin] Block definition for the faces on the absorbing boundaries, one block for each surface with x=Xmin.
-\item [face\_abs\_xmax] Block definition for the faces on the absorbing boundaries, one block for each surface with x=Xmax.
-\item [face\_abs\_ymin] Block definition for the faces on the absorbing boundaries, one block for each surface with y=Ymin.
-\item [face\_abs\_ymax] Block definition for the faces on the absorbing boundaries, one block for each surface with y=Ymax.
-\item [face\_abs\_bottom] Block definition for the faces on the absorbing boundaries, one block for each surface with z=bottom.
-\end{description}
-
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=.45\textwidth]{figures/mount-surface.jpg}
-\includegraphics[width=.45\textwidth]{figures/mount-abs.jpg}
-\par\end{centering}
-
-\caption{Example of the block definitions for the free surface 'face\_topo' (left) and the absorbing boundaries, defined in a single block 'face\_abs\_xmin' (right) in CUBIT.}
-
-\label{fig:mount.abs}
-\end{figure}
-
-Optionally, instead of specifying for each surface at the boundaries a single block like mentioned above, you can also specify a single block for all boundary surfaces and name it as one of the absorbing blocks above, e.g.
-'face\_abs\_xmin'.
-
-After the block definitions are done, you export the mesh using the script \texttt{cubit2specfem3d.py} provided in each of the example directories.
-If the export was successful, you should find the following files in a subdirectory \texttt{MESH/}:
-
-\begin{description}
-\item [nummaterial\_velocity\_file] Defines the material properties. For fully defined materials, the formats is:
-\begin{lyxcode}
-\texttt{domain\_ID} \texttt{material\_ID}  \texttt{rho}  \texttt{vp}  \texttt{vs}  \texttt{Q}  \texttt{anisotropy\_flag}
-\end{lyxcode}
-where
-\texttt{domain\_ID} is 1 for acoustic or 2 for elastic materials, \texttt{material\_ID} a unique identifier,  \texttt{rho} the density in $kg \, m^{-3}$,  \texttt{vp} the P-wave speed in $m \, s^{-1}$, \texttt{vs}  the S-wave speed in $m \, s^{-1}$, \texttt{Q} the quality factor and \texttt{anisotropy\_flag} an identifier for anisotropic models.
-
-%%magnoni
-For tomographic velocity models, please read Chapter \ref{cha:-Changing-the} and Section \ref{sec:Using-tomographic} `Using external tomographic Earth models' for further details.
-%
-%For a tomographic model, the material definition format is:
-%\begin{lyxcode}
-%\texttt{domain\_ID} \texttt{material\_ID}  \texttt{tomography}  \texttt{name}
-%\end{lyxcode}
-%where
-%\texttt{domain\_ID} is 1 for acoustic or 2 for elastic materials, \texttt{material\_ID} a negative, unique identifier, \texttt{tomography} keyword for tomographic material definition and \texttt{name} the name of the tomography file.
-The name is not used so far, rather change the filename defined in the file \texttt{model\_tomography.f90} located in the
-%\texttt{src/generate\_databases/} directory.
-%
-\item [materials\_file] Contains the material associations for each element.
-\item [nodes\_coords\_file] Contains the point locations in Cartesian coordinates of the mesh element corners.
-\item [mesh\_file] Contains the mesh element connectivity.
-\item [free\_surface\_file] Contains the free surface connectivity.
-\item [absorbing\_surface\_file\_xmax] Contains the surface connectivity of the absorbing boundary surface at the Xmax.
-\item [absorbing\_surface\_file\_xmin] Contains the surface connectivity of the absorbing boundary surface at the Xmin.
-\item [absorbing\_surface\_file\_ymax] Contains the surface connectivity of the absorbing boundary surface at the Ymax.
-\item [absorbing\_surface\_file\_ymin] Contains the surface connectivity of the absorbing boundary surface at the Ymin.
-\item [absorbing\_surface\_file\_bottom] Contains the surface connectivity of the absorbing boundary surface at the bottom.
-\end{description}
-
-These mesh files are needed as input files for the partitioner  \texttt{xdecompose\_mesh\_SCOTCH}
-to load-balance the mesh. Please see the next section for further details.
-
-\subsubsection*{Checking the mesh quality}
-The quality of the mesh may be inspected more precisely based upon the serial code
-in the file  \texttt{check\_mesh\_quality\_}~\\
-\texttt{CUBIT\_Abaqus.f90} located in the directory \texttt{src/check\_mesh\_quality\_CUBIT\_Abaqus/}.
-Running this code is optional because no information needed by the
-solver is generated.
-
-Prior to running and compiling this code, you have to export your mesh in CUBIT to an ABAQUS (.inp) format.
-For example, export mesh block IDs belonging to volumes in order to check the quality of the hexahedral elements.
-You also have to determine a number of parameters of your mesh, such as the number of nodes and number of elements
-and modify the header of the \texttt{check\_mesh\_quality\_CUBIT\_Abaqus.f90} source file. Then,
-in the main directory, type
-\begin{lyxcode}
-{\small make xcheck\_mesh\_quality\_CUBIT\_Abaqus}{\small \par}
-\end{lyxcode}
-and use
-\begin{lyxcode}
-{\small ./bin/xcheck\_mesh\_quality\_CUBIT\_Abaqus}{\small \par}
-\end{lyxcode}
-to generate an
-%not supported: AVS output file (\texttt{\small AVS\_meshquality.inp} in AVS UCD format) or
-OpenDX output file (\texttt{\small DX\_mesh\_quality.dx})
-that can be used to investigate mesh quality, e.g. skewness of elements,
-and a Gnuplot histogram (\texttt{\small mesh\_quality\_histogram.txt})
-that can be plotted with gnuplot (type
-`\texttt{\small gnuplot plot\_mesh\_quality\_histogram.gnu}'). The
-histogram is also printed to the screen.
-Analyze that skewness histogram of mesh elements to make sure no element has a skewness above approximately 0.75, otherwise the mesh is of poor quality (and if even a single element has a skewness value above 0.80, then you must definitely improve the mesh).
-If you want to start designing
-your own meshes, this tool is useful for viewing your creations. You
-are striving for meshes with elements with `cube-like' dimensions,
-e.g., the mesh should contain no very elongated or skewed elements.
-
-\subsection{Partitioning the Mesh with \texttt{xdecompose\_mesh\_SCOTCH}}
-
-The SPECFEM3D software package performs large scale simulations in a parallel 'Single Process Multiple Data' way.
-The spectral-element mesh created with CUBIT needs to be distributed on the processors. This partitioning is executed once and for all prior to
-the execution of the solver so it is referred to as a static mapping.
-
-An efficient partitioning is important because it leverages the overall running time of the application.
-It amounts to balance the number of elements in each slice while minimizing the communication costs
-resulting from the placement of adjacent elements on different processors. \texttt{decompose\_mesh\_SCOTCH} depends on the SCOTCH library \citep{PeRo96},
-which provides efficent static mapping, graph and mesh partitioning routines. SCOTCH is a free software package developed by
-Fran\c {c}ois Pellegrini et al.
-from LaBRI and INRIA in Bordeaux, France, downloadable from the web page \url{https://gforge.inria.fr/projects/scotch/}.
-It is more recent than METIS, actively maintained and performs better in many cases.
-A recent version of its source code is provided in directory \texttt{src/decompose\_mesh\_SCOTCH/scotch\_5.1.10b}.
-\\
-
-In most cases, the configuration with \texttt{./configure~FC=ifort} should be sufficient.
-During the configuration process, the script tries to find existing SCOTCH installations.
-In case your system has no pre-existing SCOTCH installation,
-we provide the source code of SCOTCH, which is released open source under the French CeCILL-C version 1 license,
-in directory
-\texttt{src/decompose\_mesh\_SCOTCH/scotch\_5.1.10b}.
-This version gets bundled with the compilation of the SPECFEM3D package if no libraries could have been found.
-If this automatic compilation of the SCOTCH libraries fails,
-please refer to file INSTALL.txt in that directory to see further details how to compile it on your system.
-In case you want to use a pre-existing installation, make sure you have correctly specified the path
-of the SCOTCH library when using the option \texttt{-{}-with-scotch-dir} with the \texttt{./configure} script.
-In the future you should be able to find more recent versions at  \url{http://www.labri.fr/perso/pelegrin/scotch/scotch\_en.html} .
-\\
-
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=.45\textwidth]{figures/mount-partitions.jpg}
-\includegraphics[width=.45\textwidth]{figures/mount-partitions2.jpg}
-\par\end{centering}
-
-\caption{Example of a mesh partitioning onto four cores.
-Each single core partition is colored differently.
-The executable \texttt{xdecompose\_mesh\_SCOTCH} can equally distribute the mesh on any arbitrary number of cores.
-Domain decomposition is explained in detail in \cite{MaKoBlLe08}, and excellent scaling up to 150,000 processor cores in shown for instance in
-\cite{CaKoLaTiMiLeSnTr08,KoLaMi08a,MaKoBlLe08,KoErGoMi10,KoGoErMi10,Kom11}.}
-
-\label{fig:mount.partitions}
-\end{figure}
-
-When you are ready to compile,
-in the main directory type
-\begin{lyxcode}
-{\small make xdecompose\_mesh\_SCOTCH}{\small \par}
-\end{lyxcode}
-If all paths and flags have been set correctly, the executable \texttt{bin/xdecompose\_mesh\_SCOTCH} should be produced.
-
-The partitioning is done in serial for now (in the next release we will provide a parallel version of that code),
-the synopsis is:
-\begin{lyxcode}
-./bin/xdecompose\_mesh\_SCOTCH  nparts  input\_directory output\_directory
-\end{lyxcode}
-where
-\begin{itemize}
-     \item \texttt{nparts} is the number of partitions, i.e., the number of cores for the parallel simulations,
-     \item \texttt{input\_directory} is the directory which holds all the files generated by the Python script \texttt{cubit2specfem3d.py} explained in the previous Section~\ref{subsec:Exporting-the-Mesh}, e.g.
-\texttt{MESH/}, and
-     \item \texttt{output\_directory} is the directory for the output of this partitioner which stores ACII-format files named like \texttt{proc******\_Database} for each partition.
-These files will be needed for creating the distributed databases, and have to reside in the directory  \texttt{LOCAL\_PATH} specified in the main \texttt{Par\_file}, e.g.
-in directory \texttt{in\_out\_files/DATABASES\_MPI}.
-Please see Chapter~\ref{cha:Creating-Distributed-Databases} for further details.
-\end{itemize}
-Note that all the files generated by the Python script  \texttt{cubit2specfem3d.py} must be placed in the \texttt{input\_directory} folder before running the program.
-
-\section{\label{cha:Running-the-Mesher-Meshfem3D}Meshing with \texttt{xmeshfem3D}}
-
-In case you successfully ran the configuration script, you are also ready to compile the internal mesher.
-This is an alternative to CUBIT
-for the mesh generation of relatively simple geological models. The mesher is no longer
-dedicated to Southern California and more flexiblity is provided in this version of the package.
-
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[scale=0.5]{figures/socal_map_mpi}
-\par\end{centering}
-
-\caption{\label{fig:For-parallel-computing}For parallel computing purposes,
-the model block is subdivided in $\nprocxi\times\nproceta$ slices
-of elements. In this example we use $5^{2}=25$ processors. }
-
-\end{figure}
-
-In the main directory, type
-\begin{lyxcode}
-{\small make xmeshfem3D}{\small \par}
-\end{lyxcode}
-If all paths and flags
-have been set correctly, the mesher should now compile and produce
-the executable \texttt{bin/xmeshfem3D}. Please note that
-\texttt{xmeshfem3D} must be called directly from the \texttt{bin/} directory,
-as most of the binaries of the package.
-
-Input for the mesh generation program is provided through the parameter
-file \texttt{Mesh\_Par\_file}, which resides in the subdirectory \texttt{in\_data\_files/meshfem3D\_files/}.
-Before running the mesher, a number of parameters need to be set in
-the \texttt{Mesh\_Par\_file}. This requires a basic understanding of how
-the SEM is implemented, and we encourage you to read \citet{KoVi98,KoTr99}
-and \citet{KoLiTrSuStSh04}.
-
-The mesher and the solver use UTM coordinates internally, therefore
-you need to define the zone number for the UTM projection (e.g., zone
-11 for Los Angeles). Use decimal values for latitude and longitude
-(no minutes/seconds). These values are approximate; the mesher will
-round them off to define a square mesh in UTM coordinates. When running
-benchmarks on rectangular models, turn the UTM projection off by using
-the flag \texttt{\small SUPPRESS\_UTM\_PROJECTION}, in which case
-all `longitude' parameters simply refer to the $x$~axis, and all
-`latitude' parameters simply refer to the $y$~axis. To run the mesher
-for a global simulation, the following parameters need to be set in
-the \texttt{Mesh\_Par\_file}:
-
-\begin{description}
-\item [{\texttt{LATITUDE\_MIN}}] Minimum latitude in the block (negative
-for South).
-\item [{\texttt{LATITUDE\_MAX}}] Maximum latitude in the block.
-\item [{\texttt{LONGITUDE\_MIN}}] Minimum longitude in the block (negative
-for West).
-\item [{\texttt{LONGITUDE\_MAX}}] Maximum longitude in the block.
-\item [{\texttt{DEPTH\_BLOCK\_KM}}] Depth of bottom of mesh in kilometers.
-\item [{\texttt{\noun{UTM\_PROJECTION\_ZONE}}}] UTM projection zone in
-which your model resides, only valid when \texttt{SUPPRESS\_UTM\_}~\\
-\texttt{PROJECTION} is \texttt{.false.}.
-\item [{\texttt{SUPPRESS\_UTM\_PROJECTION}}] set to be \texttt{.false.}
-when your model range is specified in geographical coordinates,
-and needs to be \texttt{.true.} when your model is specified in
-Cartesian coordinates. \noun{UTM projection zone in which your simulation
-region resides.}
-\item [{\texttt{INTERFACES\_FILE }}] File which contains the description of the topography and
-of the interfaces between the different layers of the model, if any.
-The number of spectral elements in the vertical direction within each layer is also defined in this file.
-
-\item [{$\nexxi$}] The number of spectral elements along one side of the
-block. This number \textit{must} be 8~$\times$~a multiple of $\nprocxi$
-defined below. Based upon benchmarks against semi-analytical discrete
-wavenumber synthetic seismograms \citep{KoLiTrSuStSh04}, determined
-that a $\nexxi=288$ run is accurate to a shortest period of roughly
-2~s. Therefore, since accuracy is determined by the number of grid
-points per shortest wavelength, for any particular value of $\nexxi$
-the simulation will be accurate to a shortest period determined by
-\begin{equation}
-\mbox{shortest period (s)}=(288/\nexxi)\times2.\label{eq:shortest_period}\end{equation}
-The number of grid points in each orthogonal direction of the reference
-element, i.e., the number of Gauss-Lobatto-Legendre points, is determined
-by \texttt{NGLLX} in the \texttt{constants.h} file. We generally use
-$\mbox{\texttt{NGLLX\/}}=5$, for a total of $5^{3}=125$ points per
-elements. We suggest not to change this value.
-\item [{$\nexeta$}] The number of spectral elements along the other side
-of the block. This number \textit{must} be 8~$\times$~a multiple
-of $\nproceta$ defined below.
-\item [{$\nprocxi$}] The number of processors or slices along one side
-of the block (see Figure~\ref{fig:For-parallel-computing}); we must
-have $\nexxi=8\times c\times\nprocxi$, where $c\ge1$ is a positive
-integer.
-\item [{$\nproceta$}] The number of processors or slices along the other
-side of the block; we must have $\nexeta=8\times c\times\nproceta$,
-where $c\ge1$ is a positive integer.
-\item [{\texttt{USE\_REGULAR\_MESH}}] set to be \texttt{.true.} if you want a perfectly regular
-mesh or \texttt{.false.} if you want to add doubling horizontal layers to coarsen the mesh. In this case, you
-also need to provide additional information by setting up the next three parameters.
-\item [{\texttt{NDOUBLINGS}}] The number of horizontal doubling layers. Must be set to \texttt{1} or \texttt{2}
-if \texttt{USE\_REGULAR\_MESH} is set to \texttt{.true.}.
-\item [{\texttt{NZ\_DOUBLING\_1}}] The position of the first doubling layer (only interpreted
-if \texttt{USE\_REGULAR\_MESH} is set to \texttt{.true.}).
-\item [{\texttt{NZ\_DOUBLING\_2}}] The position of the second doubling layer (only interpreted
-if \texttt{USE\_REGULAR\_MESH} is set to \texttt{.true.} and if \texttt{NDOUBLINGS} is set to \texttt{2}).
-\item [{\texttt{CREATE\_ABAQUS\_FILES}}] Set this flag to \texttt{.true.} to save Abaqus FEA \urlwithparentheses{www.simulia.com}
-mesh files for subsequent viewing. Turning the flag on generates files in the \texttt{LOCAL\_PATH} directory.
-See Section~\ref{sec:Mesh-graphics} for a discussion of mesh viewing features.
-\item [{\texttt{CREATE\_DX\_FILES}}] Set this flag to \texttt{.true.} to save OpenDX \urlwithparentheses{www.opendx.org}
-mesh files for subsequent viewing.
-\item [{\texttt{LOCAL\_PATH}}] Directory in which the partitions generated
-by the mesher will be written. Generally one uses a directory on the
-local disk of the compute nodes, although on some machines these partitions
-are written on a parallel (global) file system (see also the earlier
-discussion of the \texttt{LOCAL\_PATH\_IS\_ALSO\_GLOBAL} flag in Chapter~\ref{cha:Getting-Started}).
-The mesher generates the necessary partitions in parallel, one set
-for each of the $\nprocxi\times\nproceta$ slices that constitutes
-the mesh (see Figure~\ref{fig:For-parallel-computing}). After the
-mesher finishes, you can log in to one of the compute nodes and view
-the contents of the \texttt{LOCAL\_PATH} directory to see the
-files generated by the mesher.
-These files will be needed for creating the distributed databases, and have to reside in the directory  \texttt{LOCAL\_PATH} specified in the main \texttt{Par\_file}, e.g.
-in directory \texttt{in\_out\_files/DATABASES\_MPI}.
-Please see Chapter~\ref{cha:Creating-Distributed-Databases} for further details.
-\item [{\texttt{NMATERIALS}}] The number of different materials in your model. In the following lines, each material needs to be defined as :
-\begin{lyxcode}
-\texttt{material\_ID}  \texttt{rho}  \texttt{vp}  \texttt{vs}  \texttt{Q}  \texttt{anisotropy\_flag} \texttt{domain\_ID}
-\end{lyxcode}
-where
-\begin{itemize}
-     \item \texttt{Q}           : quality factor (0=no attenuation)
-     \item \texttt{anisotropy\_flag}  : 0=no anisotropy / 1,2,.. check with implementation in \texttt{aniso\_model.f90}
-     \item \texttt{domain\_id}        : 1=acoustic / 2=elastic
-\end{itemize}
-\item [{\texttt{NREGIONS}}] The number of regions in the mesh.
-In the following lines, because the mesh is regular or 'almost regular', each region is defined as :
-\begin{lyxcode}
-  \texttt{NEX\_XI\_BEGIN}  \texttt{NEX\_XI\_END}  \texttt{NEX\_ETA\_BEGIN}  \texttt{NEX\_ETA\_END}  \texttt{NZ\_BEGIN}  \texttt{NZ\_END}  \texttt{material\_ID}
-\end{lyxcode}
-\end{description}
-
-The \texttt{INTERFACES\_FILE} parameter of \texttt{Mesh\_Par\_File} defines the file which contains the settings of the topography grid and of the interfaces grids.
-Topography is defined as a set of elevation values on a regular 2D grid.
-It is also possible to
-define interfaces between the layers of the model in the same way.
-The file needs to define several parameters:
-\begin{itemize}
-\item The number of interfaces, including the topography. This needs to be set at the first line. Then, from the bottom
-to the top of the model, you need to define the grids with:
-\item \texttt{SUPPRESS\_UTM\_PROJECTION} flag as described previously,
-\item number of points along $x$ and $y$ direction (NXI and NETA),
-\item minimal $x$ and $y$ coordinates (LONG\_MIN and LAT\_MIN),
-\item spacing between points along $x$ and $y$ (SPACING\_XI and SPACING\_ETA) and
-\item the name of the file which contains the elevation values (in $y$.$x$ increasing order).
-\end{itemize}
-At the end of this file, you simply need to set the number of spectral elements in the vertical direction for each layer. We provide a few models in the {\texttt{examples/}} directory.  \\
-
-Finally, depending on your system, you might need to provide a file that tells MPI what compute nodes
-to use for the simulations. The file must have a number of entries
-(one entry per line) at least equal to the number of processors needed
-for the run. A sample file is provided in the file \texttt{mymachines}.
-This file is not used by the mesher or solver, but is required by
-the \texttt{go\_mesher} and \texttt{go\_solver} default job submission
-scripts. See Chapter~\ref{cha:Scheduler} for information about running
-the code on a system with a scheduler, e.g., LSF.
-
-Now that you have set the appropriate parameters in the \texttt{Mesh\_Par\_file}
-and have compiled the mesher, you are ready to launch it! This is
-most easily accomplished based upon the \texttt{go\_mesher} script.
-When you run on a PC cluster, the script assumes that the nodes are
-named n001, n002, etc. If this is not the case, change the \texttt{tr
--d `n'} line in the script. You may also need to edit the last command
-at the end of the script that invokes the \texttt{mpirun} command.
-See Chapter~\ref{cha:Scheduler} for information about running the
-code on a system with a scheduler, e.g., LSF.
-
-Mesher output is provided in the \texttt{in\_out\_files/OUTPUT\_FILES} directory
-in \texttt{output\_mesher.txt}; this file provides lots of details
-about the mesh that was generated. Please note that the mesher suggests a time step \texttt{DT} to run
-the solver with. The mesher output file also contains a table about the quality of the mesh to indicate
-possible problems with the distortions of elements.
-Alternatively, output can be directed
-to the screen instead by uncommenting a line in \texttt{constants.h}:
-\begin{lyxcode}
-!~uncomment~this~to~write~messages~to~the~screen~
-
-!~integer,~parameter~::~IMAIN~=~ISTANDARD\_OUTPUT
-\end{lyxcode}
-
-To control the quality of the mesh, check the standard output (either on the screen or in the \texttt{in\_out\_files/OUTPUT\_FILES} directory in \texttt{output\_mesher.txt}) and analyze the skewness histogram of mesh elements to make sure no element has a skewness above approximately 0.75, otherwise the mesh is of poor quality (and if even a single element has a skewness value above 0.80, then you must definitely improve the mesh).
-To draw the skewness histogram on the screen, type \texttt{gnuplot plot\_mesh\_quality\_histogram.gnu}.
-
-\chapter{\label{cha:Creating-Distributed-Databases}Creating the Distributed Databases}
-
-After using \texttt{xmeshfem3D} or \texttt{xdecompose\_mesh\_SCOTCH}, the next step in the workflow is to compile
-\texttt{xgenerate\_}~\\
-\texttt{databases}. This program is going to create all the missing information needed by the SEM solver.
-
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=.6\textwidth]{figures/workflow.jpg}
-\par\end{centering}
-
-\caption{Schematic workflow for a SPECFEM3D simulation. The executable \texttt{xgenerate\_databases} creates the GLL mesh points and assigns specific model parameters.}
-
-\label{fig:workflow.databases}
-\end{figure}
-
-In the main directory, type
-\begin{lyxcode}
-{\small make xgenerate\_databases}{\small \par}
-\end{lyxcode}
-Input for the program is provided through
-the main parameter file \texttt{Par\_file}, which resides in the subdirectory \texttt{in\_data\_files/}. Please note that
-\texttt{xgenerate\_databases} must be called directly from the \texttt{bin/} directory, as most of the binaries of the package.
-
-\section{\label{cha:Main-Parameter}Main parameter file \texttt{Par\_file}}
-
-Before running \texttt{xgenerate\_databases}, a number of parameters need to be set in the main parameter \texttt{Par\_file}
-located in the subdirectory \texttt{in\_data\_files/}:
-
-\begin{description}
-\item [{\texttt{SIMULATION\_TYPE}}] is set to 1 for forward simulations,
-2 for adjoint simulations (see Section \ref{sec:Adjoint-simulation-finite})
-and 3 for kernel simulations (see Section \ref{sec:Finite-Frequency-Kernels}).
-\item [{\texttt{SAVE\_FORWARD}}] is only set to \texttt{.true.} for a forward
-simulation with the last frame of the simulation saved, as part of
-the finite-frequency kernel calculations (see Section \ref{sec:Finite-Frequency-Kernels}).
-For a regular forward simulation, leave \texttt{SIMULATION\_TYPE}
-and \texttt{SAVE\_FORWARD} at their default values.
-\item [{\texttt{\noun{UTM\_PROJECTION\_ZONE}}}] UTM projection zone in
-which your model resides, only valid when \texttt{SUPPRESS\_UTM\_}~\\
-\texttt{PROJECTION} is \texttt{.false.}.
-\item [{\texttt{SUPPRESS\_UTM\_PROJECTION}}] set to be \texttt{.false.}
-when your model range is specified in the geographical coordinates,
-and needs to be \texttt{.true.} when your model is specified in a
-cartesian coordinates. \noun{UTM projection zone in which your simulation
-region resides.}
-\item [{\texttt{NPROC}}] The number of MPI processors, each one is assigned one slice of the whole mesh.
-\item [{\texttt{NSTEP}}] The number of time steps of the simulation. This controls the
-length of the numerical simulation, i.e., twice the number of time steps
-requires twice as much CPU time. This feature is not used at the time
-of generating the distributed databases but is required for the solver, i.e., you may change this
-parameter after running  \texttt{xgenerate\_databases}.
-\item [{\texttt{DT}}] The length of each time step in seconds.
-This feature is not used at the time of generating the distributed databases but is required for the solver.
-Please see also Section~\ref{sec:Choosing-the-Time-Step} for further details.
-%
-\item [{\texttt{MODEL}}] Must be set to one of the following:
-%
-\begin{description}
-\item [{\textmd{Models defined by mesh parameters:}}]~
-\begin{description}
-\item [{\texttt{default}}] Uses model parameters as defined by meshing procedures described
-in the previous Chapter~\ref{cha:Mesh-Generation}.
-\end{description}
-\end{description}
-%
-\begin{description}
-\item [{\textmd{1D~models~with~real~structure:}}]~
-\begin{description}
-\item [{\texttt{1D\_prem}}] Isotropic version of the spherically
-symmetric Preliminary Reference Earth Model (PREM) \citep{DzAn81}.
-\item [{\texttt{1D\_socal}}] A standard isotropic 1D model for Southern California.
-\item [{\texttt{1D\_cascadia}}] Isotropic 1D profile for the Cascadia region.
-\end{description}
-\end{description}
-%
-\begin{description}
-\item [{\textmd{Fully~3D~models:}}]~
-\begin{description}
-\item [{\texttt{aniso}}] For a user-specified fully anisotropic model.
-Parameters are set up in routines located in file \texttt{model\_aniso.f90} in directory \texttt{src/generate\_databases/}.
-See Chapter~\ref{cha:-Changing-the} for a discussion on how to specify your own 3D model.
-\item [{\texttt{external}}] For a user-specified isotropic model which uses externally defined model parameters.
-Uses external model definitions set up in routines located in file \texttt{model\_external\_values.f90} in directory \texttt{src/generate\_databases/}.
-Please modify these generic template routines to use your own model definitions.
-\item [{\texttt{gll}}] For a user-specified isotropic model which uses external binary files for $v_p$, $v_s$ and $\rho$.
-Binary files are given in the same format as when outputted by the \texttt{xgenerate\_databases} executable when using option \texttt{SAVE\_MESH\_FILES}.
-These binary files define the model parameters on all GLL points which can be used for iterative inversion procedures.
-\item [{\texttt{salton\_trough}}] A 3D $V_p$ model for Southern California.
-Users must provide the corresponding data file \texttt{regrid3\_vel\_p.bin} in directory \texttt{DATA/st\_3D\_block\_harvard/}.
-\item [{\texttt{tomo}}] For a user-specified 3D isotropic model which uses a tomographic model file \texttt{tomographic\_model.xyz} in directory \texttt{in\_data\_files/}.
-See Section \ref{sec:Using-tomographic},
-for a discussion on how to specify your own 3D tomographic model.
-\end{description}
-\end{description}
-%
-%
-\item [{\texttt{OCEANS}}] Set to \texttt{.true.} if the effect of the oceans
-on seismic wave propagation should be incorporated based upon the
-approximate treatment discussed in \citet{KoTr02b}. This feature
-is inexpensive from a numerical perspective, both in terms of memory
-requirements and CPU time. This approximation is accurate at periods
-of roughly 20~s and longer. At shorter periods the effect of water
-phases/reverberations is not taken into account, even when the flag
-is on.
-\item [{\texttt{TOPOGRAPHY}}] This feature is only effective if \texttt{OCEANS} is set to \texttt{.true.}.
-Set to \texttt{.true.} if topography and
-bathymetry should be read in based upon the topography file specified in the main
-constants file \texttt{constants.h} found in subdirectory \texttt{src/shared/} to evaluate
-elevations. If not set, elevations will be read from the numerical mesh.
-\item [{\texttt{ATTENUATION}}] Set to \texttt{.true.} if attenuation should
-be incorporated. Turning this feature on increases the memory requirements
-significantly (roughly by a factor of~1.5), and is numerically fairly
-expensive. See \citet{KoTr99,KoTr02a} for a discussion on the implementation
-of attenuation based upon standard linear solids.
-Please note that the Vp- and Vs-velocities of your model are given for a reference frequency.
-To change this reference frequency, you change the value of
-\texttt{ATTENUATION\_f0\_REFERENCE} in the main constants file \texttt{constants.h} found in subdirectory \texttt{src/shared/}.
-\item [{\texttt{USE\_OLSEN\_ATTENUATION}}] Set to \texttt{.true.} if you
-want to use the attenuation model that scaled from the S-wave speed model
-using Olsen's empirical relation (see \citet{OlDaBr03}).
-\item [{\texttt{ANISOTROPY}}] Set to \texttt{.true.} if you
-want to use an anisotropy model. Please see the file \texttt{model\_aniso.f90} in subdirectory \texttt{src/generate\_databases/} for the current implementation of anisotropic models.
-\item [{\texttt{ABSORBING\_CONDITIONS}}] Set to \texttt{.true.} to turn
-on Clayton-Enquist absorbing boundary conditions (see \citet{KoTr99}).
-\item [{\texttt{ABSORB\_INSTEAD\_OF\_FREE\_SURFACE}}] Set to \texttt{.true.} to turn on 
-absorbing boundary conditions on the top surface which by default constitutes a free surface of the model.
-\item [{\texttt{MOVIE\_SURFACE}}] Set to \texttt{.false.}, unless you want
-to create a movie of seismic wave propagation on the Earth's surface.
-Turning this option on generates large output files. See Section~\ref{sec:Movies}
-for a discussion on the generation of movies. This feature is only relevant for the solver.
-\item [{\texttt{MOVIE\_VOLUME}}] Set to \texttt{.false.}, unless you want
-to create a movie of seismic wave propagation in the Earth's interior.
-Turning this option on generates huge output files. See Section~\ref{sec:Movies}
-for a discussion on the generation of movies. This feature is only relevant for the solver.
-\item [{\texttt{NTSTEP\_BETWEEN\_FRAMES}}] Determines the number of timesteps
-between movie frames. Typically you want to save a snapshot every
-100 timesteps. The smaller you make this number the more output will
-be generated! See Section~\ref{sec:Movies} for a discussion on the
-generation of movies. This feature is only relevant for the solver.
-\item [{\texttt{CREATE\_SHAKEMAP}}] Set this flag to \texttt{.true.} to
-create a ShakeMap\textregistered, i.e., a peak ground velocity map
-of the maximum absolute value of the two horizontal components of the velocity vector.
-\item [{\texttt{SAVE\_DISPLACEMENT}}] Set this flag to \texttt{.true.}
-if you want to save the displacement instead of velocity for the movie
-frames.
-\item [{\texttt{USE\_HIGHRES\_FOR\_MOVIES}}] Set this flag to \texttt{.true.}
-if you want to save the values at all the NGLL grid points for the
-movie frames.
-\item [{\texttt{HDUR\_MOVIE}}] determines the half duration of the source time function for the movie simulations.
-When this parameter is set to be 0, a default half duration that corresponds to the accuracy of the simulation is provided.
-Otherwise, it adds this half duration to the half duration specified in the source file \texttt{CMTSOLUTION},
-thus simulates longer periods to make the movie images look smoother.
-\item [{\texttt{SAVE\_MESH\_FILES}}] Set this flag to \texttt{.true.} to
-save ParaView \urlwithparentheses{www.paraview.org}
-mesh files for subsequent viewing. Turning the flag on generates large
-(distributed) files in the \texttt{LOCAL\_PATH} directory. See Section~\ref{sec:Mesh-graphics}
-for a discussion of mesh viewing features.
-\item [{\texttt{LOCAL\_PATH}}] Directory in which the distributed databases
-will be written. Generally one uses a directory on the
-local disk of the compute nodes, although on some machines these databases
-are written on a parallel (global) file system (see also the earlier
-discussion of the \texttt{LOCAL\_PATH\_IS\_ALSO\_GLOBAL} flag in Chapter~\ref{cha:Getting-Started}).
-\texttt{xgenerate\_databases} generates the necessary databases in parallel, one set
-for each of the \texttt{NPROC} slices that constitutes
-the mesh (see Figure~\ref{fig:mount.partitions} and Figure~\ref{fig:For-parallel-computing}). After
-the executable finishes, you can log in to one of the compute nodes and view
-the contents of the \texttt{LOCAL\_PATH} directory to see the (many)
-files generated by \texttt{xgenerate\_databases}.
-Please note that the \texttt{LOCAL\_PATH} directory should already contain the output files of
-the partitioner, i.e. from \texttt{xdecompose\_mesh\_SCOTCH} or \texttt{xmeshfem3D}.
-\item [{\texttt{NTSTEP\_BETWEEN\_OUTPUT\_INFO}}] This parameter specifies
-the interval at which basic information about a run is written to
-the file system (\texttt{timestamp{*}} files in the \texttt{in\_out\_files/OUTPUT\_FILES}
-directory). If you have access to a fast machine, set \texttt{NTSTEP\_BETWEEN\_OUTPUT\_INFO}
-to a relatively high value (e.g., at least 100, or even 1000 or more)
-to avoid writing output text files too often. This feature is not
-used at the time of meshing. One can set this parameter to a larger
-value than the number of time steps to avoid writing output during
-the run.
-\item [{\texttt{NTSTEP\_BETWEEN\_OUTPUT\_SEISMOS}}] This parameter specifies
-the interval at which synthetic seismograms are written in the \texttt{LOCAL\_PATH}
-directory. If a run crashes, you may still find usable (but shorter
-than requested) seismograms in this directory. On a fast machine set
-\texttt{NTSTEP\_BETWEEN\_OUTPUT\_SEISMOS} to a relatively high value
-to avoid writing to the seismograms too often. This feature is only relevant for the solver.
-\item [{\texttt{USE\_FORCE\_POINT\_SOURCE}}] Turn this flag on to use a \texttt{FORCESOLUTION} force source
-instead of a \texttt{CMTSOLUTION} moment-tensor source. This can be useful e.g. for oil industry 
-foothills simulations in which the source is a vertical force, normal force, 
-tilted force, or an impact etc. You will need to give the East, North and vertical components of an 
-arbitrary (non-unitary) direction vector of the force vector; thus refer to Appendix~\ref{cha:Coordinates} 
-for the orientation of the reference frame.
-This vector is made unitary internally in the code and thus only its direction matters here;
-its norm is ignored and the norm of the force used is the factor force source times the source time function.
-\item [{\texttt{PRINT\_SOURCE\_TIME\_FUNCTION}}] Turn this flag on to print
-information about the source time function in the file \texttt{in\_out\_files/OUTPUT\_FILES/plot\_source\_time\_function.txt}.
-This feature is only relevant for the solver.
-\end{description}
-
-\section{\label{sec:Choosing-the-Time-Step}Choosing the time step \texttt{DT}}
-
-The parameter \texttt{DT} sets the length of each time step in seconds. The value of this parameter is crucial for the stability of
-the spectral-element simulation.
-Your time step \texttt{DT} will depend on the minimum ratio between the distance $h$ of neighboring mesh points and the wave speeds $v$ defined in your model.
-The condition for the time step $\Delta t$ is:
-
-\begin{lyxcode}
-$\Delta t < C ~\mathrm{min}_\Omega (~ h / v ~)$
-\end{lyxcode}
-where $C$ is the so-called Courant number and $\Omega$ denotes the model volume.
-The distance $h$ depends on the mesh element size and the number of
-GLL points \texttt{NGLL} specified in the main constants file \texttt{constants.h} located in the \texttt{src/shared/} subdirectory.
-The wave speed $v$ is determined based on your model's P- (or S-) wave speed values.
-
-The database generator \texttt{xgenerate\_databases}, as well as the internal mesher \texttt{xmeshfem3D}, are trying
-to evaluate the value of $\Delta t$ for empirically chosen Courant numbers $C \sim 0.3$.
-If you used the mesher \texttt{xmeshfem3D} to generate your mesh, you should set the value suggested in \texttt{in\_out\_files/OUTPUT\_FILES/output\_mesher.txt} file, which is created after the mesher completed.
-In case you used CUBIT to create the mesh, you might use an arbitrary value when running \texttt{xgenerate\_databases} and then use the value suggested in the ~\\
-\texttt{in\_out\_files/OUTPUT\_FILES/output\_mesher.txt} file after the database generation completed.
-Note that the implemented Newmark time scheme uses this time step globally, thus your simulations become more expensive for very small mesh elements in high wave-speed regions.
-Please be aware of this restriction when constructing your mesh in Chapter~\ref{cha:Mesh-Generation}.
-
-\chapter{\label{cha:Running-the-Solver}Running the Solver \texttt{xspecfem3D}}
-
-Now that you have successfully generated the databases, you are ready to compile
-the solver. In the main directory, type
-\begin{lyxcode}
-{\small make xspecfem3D}{\small \par}
-\end{lyxcode}
-Please note that \texttt{xspecfem3D} must be called directly from
-the \texttt{bin/} directory, as most of the binaries of the package.
-
-The solver needs three input files in the \texttt{in\_data\_files/} directory
-to run:
-\begin{description}
-\item [{\texttt{Par\_file}}] the main parameter file which was discussed in detail in the previous
-Chapter~\ref{cha:Creating-Distributed-Databases},
-\item [{\texttt{CMTSOLUTION}} or {\texttt{FORCESOLUTION}}] the earthquake source parameter file or the force source parameter file or, and
-\item [{\texttt{STATIONS}}] the stations file.
-\end{description}
-
-Most parameters in the \texttt{Par\_file}
-should be set prior to running the databases generation. Only the following parameters
-may be changed after running \texttt{xgenerate\_databases}:
-
-\begin{itemize}
-\item the simulation type control parameters: \texttt{SIMULATION\_TYPE}
-and \texttt{SAVE\_FORWARD}
-\item the time step parameters \texttt{NSTEP} and \texttt{DT}
-\item the absorbing boundary control parameter \texttt{ABSORBING\_CONDITIONS} on condition that the\\ 
-  \texttt{ABSORB\_INSTEAD\_OF\_FREE\_SURFACE} flag remains unmodified after running the databases generation. 
-\item the movie control parameters \texttt{MOVIE\_SURFACE}, \texttt{MOVIE\_VOLUME},
-and \texttt{NTSTEPS\_BETWEEN\_FRAMES}
-\item the ShakeMap\textregistered option \texttt{CREATE\_SHAKEMAP}
-\item the output information parameters \texttt{NTSTEP\_BETWEEN\_OUTPUT\_INFO}
-and \texttt{NTSTEP\_BETWEEN\_OUTPUT\_}~\\
-\texttt{SEISMOS}
-\item the \texttt{PRINT\_SOURCE\_TIME\_FUNCTION} flags
-\end{itemize}
-Any other change to the \texttt{Par\_file} implies rerunning both
-the database generator \texttt{xgenerate\_databases} and the solver \texttt{xspecfem3D}.
-
-For any particular earthquake, the \texttt{CMTSOLUTION} file that
-represents the point source may be obtained directly from the Harvard Centroid-Moment Tensor (CMT) web page \urlwithparentheses{www.seismology.harvard.edu}.
-It looks like this:
-
-\begin{lyxcode}
-{\small }%
-\begin{figure}[H]
-\noindent \begin{centering}
-{\small \includegraphics[width=1\textwidth]{figures/Hollywood_CMT} }
-\par\end{centering}{\small \par}
-
-\caption{\label{fig:CMTSOLUTION-file}\texttt{CMTSOLUTION} file based on the
-format from the Harvard CMT catalog. \textbf{M} is the moment tensor,
-$M_{0}${\small{} }is the seismic moment, and $M_{w}$ is the moment
-magnitude.}
-
-\end{figure}
-{\small \par}
-\end{lyxcode}
-The \texttt{CMTSOLUTION} should be edited in the following way:
-
-\begin{itemize}
-\item Set the \texttt{time shift} parameter equal to $0.0$ (the solver
-will not run otherwise.) The time shift parameter would simply apply
-an overall time shift to the synthetics, something that can be done
-in the post-processing (see Section \ref{sec:Process-data-and-syn}).
-\item For point-source simulations (see finite sources, page \pageref{To-simulate-a})
-we recommend setting the source half-duration parameter \texttt{half
-duration} equal to zero, which corresponds to simulating a step source-time
-function, i.e., a moment-rate function that is a delta function. If
-\texttt{half duration} is not set to zero, the code will use a Gaussian
-(i.e., a signal with a shape similar to a `smoothed triangle', as
-explained in \citet{KoTr02a} and shown in Fig~\ref{fig:gauss.vs.triangle})
-source-time function with half-width \texttt{half duration}. We prefer
-to run the solver with \texttt{half duration} set to zero and convolve
-the resulting synthetic seismograms in post-processing after the run,
-because this way it is easy to use a variety of source-time functions
-(see Section \ref{sec:Process-data-and-syn}). \citet{KoTr02a} determined
-that the noise generated in the simulation by using a step source
-time function may be safely filtered out afterward based upon a convolution
-with the desired source time function and/or low-pass filtering. Use
-the serial code \texttt{convolve\_source\_timefunction.f90} and the
-script \texttt{convolve\_source\_timefunction.csh} for this purpose,
-or alternatively use signal-processing software packages such as SAC \urlwithparentheses{www.llnl.gov/sac}.
-Type
-\begin{lyxcode}
-{\small make xconvolve\_source\_timefunction}{\small \par}
-\end{lyxcode}
-to compile the code and then set the parameter \texttt{hdur} in \texttt{convolve\_source\_timefunction.csh}
-to the desired half-duration.
-
-\item The zero time of the simulation corresponds to the center of the triangle/Gaussian,
-or the centroid time of the earthquake. The start time of the simulation
-is $t=-1.5*\texttt{half duration}$ (the 1.5 is to make sure the moment
-rate function is very close to zero when starting the simulation).
-To convert to absolute time $t_{\mathrm{abs}}$, set
-
-\begin{lyxcode}
-$t_{\mathrm{abs}}=t_{\mathrm{pde}}+\texttt{time shift}+t_{\mathrm{synthetic}}$
-\end{lyxcode}
-where $t_{\mathrm{pde}}$ is the time given in the first line of the
-\texttt{CMTSOLUTION}, \texttt{time shift} is the corresponding value
-from the original \texttt{CMTSOLUTION} file and $t_{\mathrm{synthetic}}$
-is the time in the first column of the output seismogram.
-
-\end{itemize}
-%
-\begin{figure}
-\noindent \begin{centering}
-%%%\includegraphics[width=3in]{figures/gauss_vs_triangle_mod.eps}
-\includegraphics[width=3in]{figures/gauss_vs_triangle_mod.pdf}
-\par\end{centering}
-
-\caption{Comparison of the shape of a triangle and the Gaussian function actually
-used.}
-
-\label{fig:gauss.vs.triangle}
-\end{figure}
-
-Centroid latitude and longitude should be provided in geographical
-coordinates. The code converts these coordinates to geocentric coordinates~\citep{DaTr98}.
-Of course you may provide your own source representations by designing
-your own \texttt{CMTSOLUTION} file. Just make sure that the resulting
-file adheres to the Harvard CMT conventions (see Appendix~\ref{cha:Coordinates}).
-Note that the first line in the \texttt{CMTSOLUTION} file is the Preliminary Determination of Earthquakes (PDE) solution performed by the USGS NEIC, which is used as a seed for the Harvard CMT inversion.
-The PDE solution is based upon P waves and often gives the hypocenter of the earthquake, i.e., the rupture initiation point, whereas the CMT solution gives the `centroid location', which is the location with dominant moment release.
-The PDE solution is not used by our software package but must be present anyway in the first line of the file.
-
-\label{To-simulate-a}To simulate a kinematic rupture, i.e., a finite-source
-event, represented in terms of $N_{\mathrm{sources}}$ point sources,
-provide a \texttt{CMTSOLUTION} file that has $N_{\mathrm{sources}}$
-entries, one for each subevent (i.e., concatenate $N_{\mathrm{sources}}$
-\texttt{CMTSOLUTION} files to a single \texttt{CMTSOLUTION} file).
-At least one entry (not necessarily the first) must have a zero \texttt{time
-shift}, and all the other entries must have non-negative \texttt{time
-shift}. Each subevent can have its own half duration, latitude, longitude,
-depth, and moment tensor (effectively, the local moment-density tensor).
-
-Note that the zero in the synthetics does NOT represent the hypocentral
-time or centroid time in general, but the timing of the \textit{center}
-of the source triangle with zero \texttt{time shift} (Fig~\ref{fig:source_timing}).
-
-Although it is convenient to think of each source as a triangle, in
-the simulation they are actually Gaussians (as they have better frequency
-characteristics). The relationship between the triangle and the gaussian
-used is shown in Fig~\ref{fig:gauss.vs.triangle}. For finite fault
-simulations it is usually not advisable to use a zero half duration
-and convolve afterwards, since the half duration is generally fixed
-by the finite fault model.
-
-\vspace{1cm}
-
-\noindent The \texttt{FORCESOLUTION} should be edited in the following way:
-
-\begin{itemize}
-\item Set the \texttt{time shift} parameter equal to $0.0$ (the solver
-will not run otherwise.) The time shift parameter would simply apply
-an overall time shift to the synthetics, something that can be done
-in the post-processing (see Section \ref{sec:Process-data-and-syn}).
-\item Set the latitude, longitude, depth of the source in geographical
-  coordinates.
-\item Set the magnitude of the force source.
-\item Set the components of a (non-unitary) direction vector for the
-  force source in the East/North/Vertical basis (see Appendix A for
-  the orientation of the reference frame).
-\end{itemize}
-
-{\small }%
-\begin{figure}[H]
-\noindent \begin{centering}
-{\small \includegraphics[width=5in]{figures/source_timing.pdf} }
-\par\end{centering}{\small \par}
-
-\caption{Example of timing for three sources. The center of the first source
-triangle is defined to be time zero. Note that this is NOT in general
-the hypocentral time, or the start time of the source (marked as tstart).
-The parameter \texttt{time shift} in the \texttt{CMTSOLUTION} file
-would be t1(=0), t2, t3 in this case, and the parameter \texttt{half
-duration} would be hdur1, hdur2, hdur3 for the sources 1, 2, 3 respectively.}
-
-{\small \label{fig:source_timing} }
-\end{figure}
-{\small \par}
-
-The solver can calculate seismograms at any number of stations for
-basically the same numerical cost, so the user is encouraged to include
-as many stations as conceivably useful in the \texttt{STATIONS} file,
-which looks like this:
-
-{\small }%
-\begin{figure}[H]
-\noindent \begin{centering}
-{\small \includegraphics{figures/STATIONS_basin_explained} }
-\par\end{centering}{\small \par}
-
-\caption{Sample \texttt{STATIONS} file. Station latitude and longitude should
-be provided in geographical coordinates. The width of the station
-label should be no more than 32 characters (see \texttt{MAX\_LENGTH\_STATION\_NAME}
-in the \texttt{constants.h} file), and the network label should be
-no more than 8 characters (see \texttt{MAX\_LENGTH\_NETWORK\_NAME}
-in the \texttt{constants.h} file).}
-
-\end{figure}
-{\small \par}
-
-Each line represents one station in the following format:
-
-\begin{lyxcode}
-{\small Station~Network~Latitude~(degrees)~Longitude~(degrees)~Elevation~(m)~burial~(m)~}{\small \par}
-\end{lyxcode}
-The solver \texttt{xspecfem3D} filters the list of stations in file \texttt{in\_data\_files/STATIONS}
-to exclude stations that are not located within the region given in
-the \texttt{Par\_file} (between \texttt{LATITUDE\_MIN} and \texttt{LATITUDE\_MAX}
-and between \texttt{LONGITUDE\_MIN} and \texttt{LONGITUDE\_MAX}).
-The filtered file is called \texttt{in\_data\_files/STATIONS\_FILTERED}.
-
-Solver output is provided in the \texttt{in\_out\_files/OUTPUT\_FILES} directory
-in the \texttt{output\_solver.txt} file. Output can be directed to
-the screen instead by uncommenting a line in \texttt{constants.h}:
-
-\begin{lyxcode}
-!~uncomment~this~to~write~messages~to~the~screen~
-
-!~integer,~parameter~::~IMAIN~=~ISTANDARD\_OUTPUT~
-\end{lyxcode}
-On PC clusters the seismogram files are generally written to the local
-disks (the path \texttt{LOCAL\_PATH} in the \texttt{Par\_file}) and
-need to be gathered at the end of the simulation.
-
-While the solver is running, its progress may be tracked by monitoring
-the `\texttt{\small timestamp{*}}' files in the ~\\
-\texttt{\small in\_out\_files/OUTPUT\_FILES} directory.
-These tiny files look something like this:
-
-\begin{lyxcode}
-Time~step~\#~~~~~~~~~~10000~
-
-Time:~~~~~108.4890~~~~~~seconds~
-
-Elapsed~time~in~seconds~=~~~~1153.28696703911~
-
-Elapsed~time~in~hh:mm:ss~=~~~~~0~h~19~m~13~s~
-
-Mean~elapsed~time~per~time~step~in~seconds~=~~~~~0.115328696703911~
-
-Max~norm~displacement~vector~U~in~all~slices~(m)~=~~~~1.0789589E-02~
-\end{lyxcode}
-The \texttt{\small timestamp{*}} files provide the \texttt{Mean elapsed
-time per time step in seconds}, which may be used to assess performance
-on various machines (assuming you are the only user on a node), as
-well as the \texttt{Max norm displacement vector U in all slices~(m)}.
-If something is wrong with the model, the mesh, or the source, you
-will see the code become unstable through exponentially growing values
-of the displacement and fluid potential with time, and ultimately
-the run will be terminated by the program. You can control the rate
-at which the timestamp files are written based upon the parameter
-\texttt{NTSTEP\_BETWEEN\_OUTPUT\_INFO} in the \texttt{Par\_file}.
-
-Having set the \texttt{Par\_file} parameters, and having provided the
-\texttt{CMTSOLUTION} (or the \texttt{FORCESOLUTION}) and
-\texttt{STATIONS} files, you are now ready to launch the solver! This
-is most easily accomplished based upon the \texttt{go\_solver} script
-(See Chapter \ref{cha:Scheduler} for information about running through
-a scheduler, e.g., LSF). You may need to edit the last command at the
-end of the script that invokes the \texttt{mpirun} command. The
-\texttt{runall} script compiles and runs both
-\texttt{xgenerate\_databases} and \texttt{xspecfem3D} in
-sequence. This is a safe approach that ensures using the correct
-combination of distributed database output and solver input.
-
-It is important to realize that the CPU and memory requirements of
-the solver are closely tied to choices about attenuation (\texttt{ATTENUATION})
-and the nature of the model (i.e., isotropic models are cheaper than
-anisotropic models). We encourage you to run a variety of simulations
-with various flags turned on or off to develop a sense for what is
-involved.
-
-For the same model, one can rerun the solver for different events by
-simply changing the \texttt{CMTSOLUTION} or \texttt{FORCESOLUTION}
-file, or for different stations by changing the \texttt{STATIONS}
-file. There is no need to rerun the \texttt{xgenerate\_databases}
-executable. Of course it is best to include as many stations as
-possible, since this does not add to the cost of the simulation.
-
-\chapter{\label{cha:Adjoint-Simulations}Adjoint Simulations}
-
-Adjoint simulations are generally performed for two distinct applications.
-First, they can be used for earthquake source inversions, especially
-earthquakes with large ruptures such as the Lander's earthquake \citep{WaHe94}.
-Second, they can be used to generate finite-frequency sensitivity
-kernels that are a critical part of tomographic inversions based upon
-3D reference models \citep{TrTaLi05,LiTr06,TrKoLi08,LiTr08}. In either
-case, source parameter or velocity structure updates are sought to
-minimize a specific misfit function (e.g., waveform or traveltime
-differences), and the adjoint simulation provides a means of computing
-the gradient of the misfit function and further reducing it in successive
-iterations. Applications and procedures pertaining to source studies
-and finite-frequency kernels are discussed in Sections~\ref{sec:Adjoint-simulation-sources}
-and \ref{sec:Adjoint-simulation-finite}, respectively. The two related
-parameters in the \texttt{Par\_file} are \texttt{SIMULATION\_TYPE}
-(1 or 2) and the \texttt{SAVE\_FORWARD} (boolean).
-
-\section{\label{sec:Adjoint-simulation-sources}Adjoint Simulations for Sources}
-
-In the case where a specific misfit function is minimized to invert
-for the earthquake source parameters, the gradient of the misfit function
-with respect to these source parameters can be computed by placing
-time-reversed seismograms at the receivers and using them as sources
-in an adjoint simulation, and then the value of the gradient is obtained
-from the adjoint seismograms recorded at the original earthquake location.
-
-\begin{enumerate}
-\item \textbf{Prepare the adjoint sources} \label{enu:Prepare-the-adjoint}
-
-\begin{enumerate}
-\item First, run a regular forward simlation (\texttt{SIMULATION\_TYPE =
-1} and \texttt{SAVE\_FORWARD = .false.}). You can automatically set
-these two variables using the \texttt{\small utils/change\_simulation\_type.pl}
-script:
-
-\begin{lyxcode}
-utils/change\_simulation\_type.pl~-f~
-\end{lyxcode}
-and then collect the recorded seismograms at all the stations given
-in \texttt{in\_data\_files/STATIONS}.
-
-\item Then select the stations for which you want to compute the time-reversed
-adjoint sources and run the adjoint simulation, and compile them into
-the \texttt{in\_data\_files/STATIONS\_ADJOINT} file, which has the same format
-as the regular \texttt{in\_data\_files/STATIONS} file.
-
-\begin{itemize}
-\item Depending on what type of misfit function is used for the source inversion,
-adjoint sources need to be computed from the original recorded seismograms
-for the selected stations and saved in the \texttt{in\_out\_files/SEM/} directory
-with the format \texttt{STA.NT.BX?.adj}, where \texttt{STA}, \texttt{NT}
-are the station name and network code given in the \texttt{in\_data\_files/STATIONS\_ADJOINT}
-file, and \texttt{BX?} represents the component name of a particular
-adjoint seismogram. Please note that the band code can change depending on
-your sampling rate (see Appendix~\ref{cha:channel-codes} for further details).
-\item The adjoint seismograms are in the same format as the original seismogram
-(\texttt{STA.NT.BX?.sem?}), with the same start time, time interval
-and record length.
-\end{itemize}
-\item Notice that even if you choose to time reverse only one component
-from one specific station, you still need to supply all three components
-because the code is expecting them (you can set the other two components
-to be zero).
-\item Also note that since time-reversal is done in the code itself, no
-explicit time-reversing is needed for the preparation of the adjoint
-sources, i.e., the adjoint sources are in the same forward time sense
-as the original recorded seismograms.
-\end{enumerate}
-\item \textbf{Set the related parameters and run the adjoint simulation}\\
-In the \texttt{in\_data\_files/Par\_file}, set the two related parameters to
-be \texttt{SIMULATION\_TYPE = 2} and \texttt{SAVE\_FORWARD = .false.}.
-More conveniently, use the scripts \texttt{utils/change\_simulation\_type.pl}
-to modify the \texttt{Par\_file} automatically (\texttt{change\_simulation\_type.pl
--a}). Then run the solver to launch the adjoint simulation.
-\item \textbf{Collect the seismograms at the original source location}
-
-After the adjoint simulation has completed successfully, collect the
-seismograms from \texttt{LOCAL\_PATH}.
-
-\begin{itemize}
-\item These adjoint seismograms are recorded at the locations of the original
-earthquake sources given by the \texttt{in\_data\_files/CMTSOLUTION} file, and
-have names of the form \texttt{S?????.NT.S??.sem} for the six-component
-strain tensor (\texttt{SNN,SEE,SZZ,SNE,SNZ,SEZ}) at these locations,
-and ~\\
-\texttt{S?????.NT.BX?.sem} for the three-component displacements
-(\texttt{BXN,BXE,BXZ}) recorded at these locations.
-\item \texttt{S?????} denotes the source number; for example, if the original
-\texttt{CMTSOLUTION} provides only a point source, then the seismograms
-collected will start with \texttt{S00001}.
-\item These adjoint seismograms provide critical information for the computation
-of the gradient of the misfit function.
-\end{itemize}
-\end{enumerate}
-
-\section{\label{sec:Adjoint-simulation-finite}Adjoint Simulations for Finite-Frequency
-Kernels (Kernel Simulation)}
-
-Finite-frequency sensitivity kernels are computed in two successive
-simulations (please refer to \citet{LiTr06} and \citet{TrKoLi08} for details).
-
-\begin{enumerate}
-\item \textbf{Run a forward simulation with the state variables saved at
-the end of the simulation}
-
-Prepare the \texttt{\small CMTSOLUTION} and \texttt{\small STATIONS}
-files, set the parameters \texttt{\small SIMULATION\_TYPE}{\small{}
-}\texttt{\small =}{\small{} }\texttt{\small 1} and \texttt{\small SAVE\_FORWARD
-=}{\small{} }\texttt{\small .true.} in the \texttt{Par\_file} (\texttt{change\_simulation\_type
--F}), and run the solver.
-
-\begin{itemize}
-\item Notice that attenuation is not implemented yet for the computation
-of finite-frequency kernels; therefore set \texttt{ATTENUATION = .false.}
-in the \texttt{Par\_file}.
-\item We also suggest you modify the half duration of the \texttt{CMTSOLUTION}
-to be similar to the accuracy of the simulation (see Equation \ref{eq:shortest_period})
-to avoid too much high-frequency noise in the forward wavefield, although
-theoretically the high-frequency noise should be eliminated when convolved
-with an adjoint wavefield with the proper frequency content.
-\item This forward simulation differs from the regular simulations (\texttt{\small SIMULATION\_TYPE}{\small{}
-}\texttt{\small =}{\small{} }\texttt{\small 1} and \texttt{\small SAVE\_FORWARD}{\small{}
-}\texttt{\small =}{\small{} }\texttt{\small .false.}) described in
-the previous chapters in that the state variables for the last time
-step of the simulation, including wavefields of the displacement,
-velocity, acceleration, etc., are saved to the \texttt{LOCAL\_PATH}
-to be used for the subsequent simulation.
-\item For regional simulations, the files recording the absorbing boundary
-contribution are also written to the \texttt{LOCAL\_PATH} when \texttt{SAVE\_FORWARD
-= .true.}.
-\end{itemize}
-\item \textbf{Prepare the adjoint sources}
-
-The adjoint sources need to be prepared the same way as described
-in the Section \ref{enu:Prepare-the-adjoint}.
-
-\begin{itemize}
-\item In the case of travel-time finite-frequency kernel for one source-receiver
-pair, i.e., point source from the \texttt{CMTSOLUTION}, and one station
-in the \texttt{STATIONS\_ADJOINT} list, we supply a sample program
-in \texttt{utils/adjoint\_sources/traveltime/xcreate\_adjsrc\_traveltime} to cut a certain portion of the original
-displacement seismograms and convert it into the proper adjoint source
-to compute the finite-frequency kernel.
-\begin{lyxcode}
-xcreate\_adjsrc\_traveltime~t1~t2~ifile{[}0-5]~E/N/Z-ascii-files~{[}baz]
-\end{lyxcode}
-where \texttt{t1} and \texttt{t2} are the start and end time of the
-portion you are interested in, \texttt{ifile} denotes the component
-of the seismograms to be used (0 for all three components, 1 for East,
-2 for North, and 3 for vertical, 4 for transverse, and 5 for radial
-component), \texttt{E/N/Z-ascii-files} indicate the three-component
-displacement seismograms in the right order, and \texttt{baz} is the
-back-azimuth of the station. Note that \texttt{baz} is only supplied
-when \texttt{ifile} = 4 or 5.
-
-\item Similarly, a sample program to compute adjoint sources for amplitude finite-frequency kernels
-may be found in \texttt{utils/adjoint\_sources/amplitude} and used in the same way as described for
-traveltime measurements
-\begin{lyxcode}
-xcreate\_adjsrc\_amplitude~t1~t2~ifile{[}0-5]~E/N/Z-ascii-files~{[}baz].
-\end{lyxcode}
-
-\end{itemize}
-
-\item \textbf{Run the kernel simulation}
-
-With the successful forward simulation and the adjoint source ready
-in the \texttt{in\_out\_files/SEM/} directory, set \texttt{SIMULATION\_TYPE = 3} and \texttt{SAVE\_FORWARD
-= .false.} in the \texttt{Par\_file} (you can use \texttt{change\_simulation\_type.pl -b}),
-and rerun the solver.
-
-\begin{itemize}
-\item The adjoint simulation is launched together with the back reconstruction
-of the original forward wavefield from the state variables saved from
-the previous forward simulation, and the finite-frequency kernels
-are computed by the interaction of the reconstructed forward wavefield
-and the adjoint wavefield.
-\item The back-reconstructed seismograms at the original station locations
-are saved to the \texttt{LOCAL\_PATH} at the end of the kernel simulations,
-and can be collected to the local disk.
-\item These back-constructed seismograms can be compared with the time-reversed
-original seismograms to assess the accuracy of the backward reconstruction,
-and they should match very well.
-\item The arrays for density, P-wave speed and S-wave speed kernels are
-also saved in the \texttt{LOCAL\_PATH} with the names \texttt{proc??????\_rho(alpha,beta)\_kernel.bin},
-where \texttt{proc??????} represents the processor number, \texttt{rho(alpha,beta)}
-are the different types of kernels.
-\end{itemize}
-\end{enumerate}
-In general, the three steps need to be run sequentially to assure
-proper access to the necessary files. If the simulations are run through
-some cluster scheduling system (e.g., LSF), and the forward simulation
-and the subsequent kernel simulations cannot be assigned to the same
-set of computer nodes, the kernel simulation will not be able to access
-the database files saved by the forward simulation. Solutions for
-this dilemma are provided in Chapter~\ref{cha:Scheduler}. Visualization
-of the finite-frequency kernels is discussed in Section~\ref{sec:Finite-Frequency-Kernels}.
-
-\chapter{Noise Cross-correlation Simulations}
-
-Besides earthquake simulations, SPECFEM3D includes functionality for seismic noise tomography as well.
-In order to proceed successfully in this chapter, it is critical that you have already familiarized yourself
-with procedures for meshing (Chapter \ref{cha:Mesh-Generation}),
-creating distributed databases (Chapter \ref{cha:Creating-Distributed-Databases}),
-running earthquake simulations (Chapters \ref{cha:Running-the-Solver})
-and adjoint simulations (Chapter \ref{cha:Adjoint-Simulations}).
-Also, make sure you read the article `Noise cross-correlation sensitivity kernels' \citep{trompetal2010},
-in order to understand noise simulations from a theoretical perspective.
-
-\section{Input Parameter Files}
-
-As usual, the three main input files are crucial: \texttt{Par\_file}, \texttt{CMTSOLUTION} and \texttt{STATIONS}.
-Unless otherwise specified, those input files should be located in directory \texttt{in\_data\_files/}.\\
-
-\texttt{CMTSOLUTION} is required for all simulations.
-At a first glance, it may seem unexpected to have it here,
-since the noise simulations should have nothing to do with the earthquake -- \texttt{CMTSOLUTION}.
-However, for noise simulations, it is critical to have no earthquakes.
-In other words, the moment tensor specified in \texttt{CMTSOLUTION} must be set to zero manually!\\
-
-\texttt{STATIONS} remains the same as in previous earthquake simulations,
-except that the order of receivers listed in \texttt{STATIONS} is now important.
-The order will be used to determine the `master' receiver,
-i.e., the one that simultaneously cross correlates with the others.\\
-
-\texttt{Par\_file} also requires careful attention.
-A parameter called \texttt{NOISE\_TOMOGRAPHY} has been added which specifies the type of simulation to be run.
-\texttt{NOISE\_TOMOGRAPHY} is an integer with possible values 0, 1, 2 and 3.
-For example, when \texttt{NOISE\_TOMOGRAPHY} equals 0, a regular earthquake simulation will be run.
-When it is 1/2/3, you are about to run
-step 1/2/3 of the noise simulations respectively.
-Should you be confused by the three steps, refer to \citet{trompetal2010} for details.\\
-
-Another change to \texttt{Par\_file} involves the parameter \texttt{NSTEP}.
-While for regular earthquake simulations this parameter specifies the length of synthetic seismograms generated,
-for noise simulations it specifies the length of the seismograms used to compute cross correlations.
-The actual cross correlations are thus twice this length, i.e., $2~\mathrm{NSTEP}-1$.
-The code automatically makes the modification accordingly, if \texttt{NOISE\_TOMOGRAPHY} is not zero.\\
-
-There are other parameters in \texttt{Par\_file} which should be given specific values.
-For instance, since the first two steps for calculating noise sensitivity kernels correspond to
-forward simulations, \texttt{SIMULATION\_TYPE} must be 1 when \texttt{NOISE\_TOMOGRAPHY} equals 1 or 2.
-Also, we have to reconstruct the ensemble forward wavefields in adjoint simulations,
-therefore we need to set \texttt{SAVE\_FORWARD} to \texttt{.true.} for the second step,
-i.e., when \texttt{NOISE\_TOMOGRAPHY} equals 2.
-The third step is for kernel constructions. Hence \texttt{SIMULATION\_TYPE} should be 3, whereas
-\texttt{SAVE\_FORWARD} must be \texttt{.false.}.\\
-
-Finally, for most system architectures,
-please make sure that \texttt{LOCAL\_PATH} in \texttt{Par\_file} is in fact local, not globally shared.
-Because we have to save the wavefields at the earth's surface at every time step,
-it is quite problematic to have a globally shared \texttt{LOCAL\_PATH},
-in terms of both disk storage and I/O speed.
-
-\section{Noise Simulations: Step by Step}
-
-Proper parameters in those parameter files are not enough for noise simulations to run.
-We have more parameters to specify:
-for example, the ensemble-averaged noise spectrum, the noise distribution etc.
-However, since there are a few `new' files, it is better to introduce them sequentially.
-In this section, standard procedures for noise simulations are described.
-
-\subsection{Pre-simulation}
-
-\begin{itemize}
-
-\item
-As usual, we first configure the software package using:
-
-\texttt{./configure FC=ifort MPIFC=mpif90}
-
-Use the following if SCOTCH is needed:
-
-\texttt{./configure FC=ifort MPIFC=mpif90 --with-scotch-dir=/opt/scotch}\\
-
-\item
-Next, we need to compile the source code using:
-
-\texttt{make xgenerate\_databases}
-
-\texttt{make xspecfem3D} \\
-
-\item
-Before we can run noise simulations, we have to specify the noise statistics,
-e.g., the ensemble-averaged noise spectrum.
-Matlab scripts are provided to help you to generate the necessary file.
-
-\texttt{examples/noise\_tomography/NOISE\_TOMOGRAPHY.m}  (main program)\\
-\texttt{examples/noise\_tomography/PetersonNoiseModel.m}
-
-In Matlab, simply run:
-
-\texttt{NOISE\_TOMOGRAPHY(NSTEP, DT, Tmin, Tmax, NOISE\_MODEL)}\\
-
-\texttt{DT} is given in \texttt{Par\_file},
-but \texttt{NSTEP} is NOT the one specified in \texttt{Par\_file}.
-Instead, you have to feed $2~\mathrm{NSTEP}-1$ to account for the doubled length of cross correlations.
-\texttt{Tmin} and \texttt{Tmax} correspond to the period range you are interested in,
-whereas \texttt{NOISE\_MODEL} denotes the noise model you will be using.
-Details can be found in the Matlab script.
-
-After running the Matlab script, you will be given the following information (plus a figure in Matlab):
-
-\texttt{*************************************************************}\\
-\texttt{the source time function has been saved in:}\\
-\texttt{/data2/yangl/3D\_NOISE/S\_squared} (note this path must be different)\\
-\texttt{S\_squared should be put into directory:}\\
-\texttt{in\_out\_files/NOISE\_TOMOGRAPHY/ in the SPECFEM3D package}\\
-
-In other words, the Matlab script creates a file called \texttt{S\_squared},
-which is the first `new' input file we encounter for noise simulations.
-
-One may choose a flat noise spectrum rather than Peterson's noise model.
-This can be done easily by modifying the Matlab script a little.
-
-\item
-Create a new directory in the SPECFEM3D package, name it as \texttt{in\_out\_files/NOISE\_TOMOGRAPHY/}.
-We will add some parameter files later in this folder.
-
-\item
-Put the Matlab-generated-file \texttt{S\_squared} in \texttt{in\_out\_files/NOISE\_TOMOGRAPHY/}.
-
-That's to say, you will have a file \texttt{in\_out\_files/NOISE\_TOMOGRAPHY/S\_squared} in the SPECFEM3D package.
-
-\item
-Create a file called \texttt{in\_out\_files/NOISE\_TOMOGRAPHY/irec\_master\_noise}. Note that this file is located
-in directory \texttt{in\_out\_files/NOISE\_TOMOGRAPHY/} as well. In general, all noise simulation related parameter
-files go into that directory.
-\texttt{irec\_master\_noise} contains only one integer, which is the ID
-of the `master' receiver. For example, if this file contains 5, it means that the fifth receiver
-listed in \texttt{in\_data\_files/STATIONS} becomes the `master'. That's why we mentioned previously that
-the order of receivers in \texttt{in\_data\_files/STATIONS} is important.
-
-Note that in regional simulations, the \texttt{in\_data\_files/STATIONS}
-might contain receivers which are outside of our computational domains.
-Therefore, the integer in \texttt{irec\_master\_noise}
-is actually the ID in \texttt{in\_data\_files/STATIONS\_FILTERED}
-(which is generated by \texttt{bin/xgenerate\_databases}).
-
-\item
-Create a file called \texttt{in\_out\_files/NOISE\_TOMOGRAPHY/nu\_master}. This file holds three numbers,
-forming a (unit) vector. It describes which component we are cross-correlating at the `master'
-receiver, i.e., ${\hat{\bf \nu}}^{\alpha}$ in \citet{trompetal2010}.
-The three numbers correspond to E/N/Z components respectively.
-Most often, the vertical component is used, and in those cases the three numbers should be 0, 0 and 1.
-
-\item
-Describe the noise direction and distributions in \texttt{src/shared/noise\_tomography.f90}.
-Search for a subroutine called \texttt{noise\_distribution\_direction}
-in \texttt{noise\_tomography.f90}. It is actually located at the very beginning
-of \texttt{noise\_tomography.f90}. The default assumes vertical noise and a uniform distribution
-across the whole free surface of the model. It should be quite self-explanatory for modifications.
-Should you modify this part, you have to re-compile the source code.
-
-\end{itemize}
-
-\subsection{Simulations}
-
-With all of the above done, we can finally launch our simulations.
-Again, please make sure that the \texttt{LOCAL\_PATH} in \texttt{in\_data\_files/Par\_file} is not globally shared.
-It is quite problematic to have a globally shared \texttt{LOCAL\_PATH}, in terms of both disk storage and
-speed of I/O (we have to save the wavefields at the earth's surface at every time step).
-
-As discussed in \citet{trompetal2010}, it takes three steps/simulations to obtain one contribution
-of the ensemble sensitivity kernels:
-
-\begin{itemize}
-\item
-Step 1: simulation for generating wavefields\\
-\texttt{SIMULATION\_TYPE=1}\\
-\texttt{NOISE\_TOMOGRAPHY=1}\\
-\texttt{SAVE\_FORWARD} not used, can be either .true. or .false.
-
-\item
-Step 2: simulation for ensemble forward wavefields\\
-\texttt{SIMULATION\_TYPE=1}\\
-\texttt{NOISE\_TOMOGRAPHY=2}\\
-\texttt{SAVE\_FORWARD=.true.}
-
-\item
-Step 3: simulation for ensemble adjoint wavefields and sensitivity kernels\\
-\texttt{SIMULATION\_TYPE=3}\\
-\texttt{NOISE\_TOMOGRAPHY=3}\\
-\texttt{SAVE\_FORWARD=.false.}
-
-Note Step 3 is an adjoint simulation, please refer to previous chapters on how to prepare adjoint sources
-and other necessary files, as well as how adjoint simulations work.
-
-\end{itemize}
-
-Note that it is better to run the three steps continuously within the same job,
-otherwise you have to collect the saved surface movies from the old nodes/CPUs to the new nodes/CPUs.
-This process varies from cluster to cluster and thus cannot be discussed here.
-Please ask your cluster administrator for information/configuration of the cluster you are using.\\
-
-\subsection{Post-simulation}
-
-After those simulations, you have all stuff you need, either in the \texttt{in\_out\_files/OUTPUT\_FILES/} or
-in the directory specified by \texttt{LOCAL\_PATH} in \texttt{in\_data\_files/Par\_file}
-(which are most probably on local nodes).
-Collect whatever you want from the local nodes to your workstation, and then visualize them.
-This process is the same as what you may have done for regular earthquake simulations.
-Refer to other chapters if you have problems.\\
-
-Simply speaking, two outputs are the most interesting: the simulated ensemble cross correlations and
-one contribution of the ensemble sensitivity kernels.\\
-
-The simulated ensemble cross correlations are obtained after the second simulation (Step 2).
-Seismograms in \texttt{in\_out\_files/OUTPUT\_FILES/} are actually the simulated ensemble cross correlations.
-Collect them immediately after Step 2, or the Step 3 will overwrite them.
-Note that we have a `master' receiver specified by \texttt{irec\_master\_noise},
-the seismogram at one station corresponds to the cross correlation between that station and the `master'.
-Since the seismograms have three components, we may obtain cross correlations
-for different components as well, not necessarily the cross correlations between vertical components.\\
-
-One contribution of the ensemble cross-correlation sensitivity kernels are obtained after Step 3,
-stored in the \texttt{in\_data\_files/LOCAL\_PATH} on local nodes.
-The ensemble kernel files are named the same as regular earthquake kernels.\\
-
-You need to run another three simulations to get the other contribution of the ensemble kernels,
-using different forward and adjoint sources given in \citet{trompetal2010}.
-
-\section{Example}
-
-In order to illustrate noise simulations in an easy way, one example is provided in \texttt{examples/noise\_tomography/}.
-See \texttt{examples/noise\_tomography/README} for explanations. \\
-
-Note, however, that they are created for a specific workstation (CLOVER@PRINCETON),
-which has at least 4 cores with `mpif90' working properly. \\
-
-If your workstation is suitable, you can run the example in \texttt{examples/noise\_tomography/} using:\\
-
-\texttt{./pre-processing.sh}\\
-
-Even if this script does not work on your workstation, the procedure it describes is universal.
-You may review the whole process described in the last section by following the commands in \texttt{pre-processing.sh},
-which should contain enough explanations for all the commands.
-
-\chapter{Graphics}
-
-\section{\label{sec:Mesh-graphics}Meshes}
-
-In case you used the internal mesher \texttt{xmeshfem3D} to create and partition your mesh,
-you can output mesh files in ABAQUS (.INP) and DX (.dx) format to visualize them. For this, you must set either the
-flag \texttt{CREATE\_DX\_FILES} or \texttt{CREATE\_ABAQUS\_FILES} to \texttt{.true.}
-in the mesher's parameter file \texttt{Mesh\_Par\_file}
-prior to running the mesher (see Chapter~\ref{cha:Running-the-Mesher-Meshfem3D} for details).
-You can then use AVS \urlwithparentheses{www.avs.com} or OpenDX \urlwithparentheses{www.opendx.org}
-to visualize the mesh and MPI partition (slices).
-
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=.49\textwidth]{figures/vtk_mesh_vp.jpg}
-\includegraphics[width=.49\textwidth]{figures/vtk_mesh_vs.jpg}
-\par\end{centering}
-
-\caption{Visualization using Paraview of VTK files created by \texttt{xgenerate\_databases} showing P- and S-wave velocities assigned to the mesh points.
-The mesh was created by \texttt{xmeshfem3D} for 4 processors.}
-
-\label{fig:vtk.mesh}
-\end{figure}
-
-You have also the option to visualize the distributed databases produced by \texttt{xgenerate\_databases}
-using Paraview \urlwithparentheses{www.paraview.org}. For this, you must set the flag \texttt{SAVE\_MESH\_FILES} to \texttt{.true.}
-in the main parameter file \texttt{Par\_file} (see Chapter~\ref{cha:Main-Parameter} for details).
-This will create VTK files for each single partition. You can then use Paraview \urlwithparentheses{www.paraview.org} to visualized these partitions.
-
-\section{\label{sec:Movies}Movies}
-
-To make a surface or volume movie of the simulation, set parameters
-\texttt{MOVIE\_SURFACE}, \texttt{MOVIE\_VOLUME}, and \texttt{NTSTEP\_BETWEEN\_FRAMES}
-in the \texttt{Par\_file}. Turning on the movie flags, in particular
-\texttt{MOVIE\_VOLUME}, produces large output files. \texttt{MOVIE\_VOLUME}
-files are saved in the \texttt{LOCAL\_PATH} directory, whereas \texttt{MOVIE\_SURFACE}
-output files are saved in the \texttt{in\_out\_files/OUTPUT\_FILES} directory. We
-save the displacement field if the parameter \texttt{SAVE\_DISPLACEMENT} is set, otherwise the velocity field is saved.
-The look of a movie is determined by the
-half-duration of the source. The half-duration should be large enough
-so that the movie does not contain frequencies that are not resolved
-by the mesh, i.e., it should not contain numerical noise. This can
-be accomplished by selecting a CMT \texttt{HALF\_DURATION} > 1.1 $\times$
-smallest period (see figure \ref{fig:CMTSOLUTION-file}). When \texttt{MOVIE\_SURFACE}
-= .\texttt{true.}, the half duration of each source in the \texttt{CMTSOLUTION}
-file is replaced by
-
-\begin{quote}
-\[
-\sqrt{(}\mathrm{\mathtt{HALF\_DURATIO}\mathtt{N}^{2}}+\mathrm{\mathtt{HDUR\_MOVI}\mathtt{E}^{2}})\]
-\textbf{NOTE:} If \texttt{HDUR\_MOVIE} is set to 0.0, the code will
-select the appropriate value of 1.1 $\times$ smallest period. As
-usual, for a point source one can set \texttt{half duration} in the
-\texttt{CMTSOLUTION} file to be 0.0 and \texttt{HDUR\_MOVIE} = 0.0 in the \texttt{Par\_file} to get
-the highest frequencies resolved by the simulation, but for a finite
-source one would keep all the \texttt{half durations} as prescribed
-by the finite source model and set \texttt{HDUR\_MOVIE} = 0.0.
-\end{quote}
-
-\subsection{Movie Surface}
-
-When running \texttt{xspecfem3D} with the \texttt{MOVIE\_SURFACE}
-flag turned on, the code outputs \texttt{moviedata??????} files in
-the \texttt{in\_out\_files/OUTPUT\_FILES} directory. There are several flags
-in the main parameter file \texttt{Par\_file} that control the output of these moviedata files
-(see section \ref{cha:Main-Parameter} for details):
-\texttt{\small NTSTEP\_BETWEEN\_FRAMES} to set the timesteps between frames,
-\texttt{\small SAVE\_DISPLACEMENT} to save displacement instead of velocity,
-\texttt{USE\_HIGHRES\_FOR\_MOVIES} to save values at all GLL point instead of element edges.
-In order to output additionally shakemaps, you would set
-the parameter \texttt{CREATE\_SHAKEMAP} to \texttt{.true.}.
-
-The files are in a fairly complicated
-binary format, but there is a program provided to convert the
-output into more user friendly formats:
-\begin{description}
-\item [\texttt{xcreate\_movie\_shakemap\_AVS\_DX\_GMT}]
-From \texttt{create\_movie\_shakemap\_AVS\_DX\_GMT.f90}, it
-outputs data in ASCII, OpenDX, or AVS format (also readable in ParaView).
-Before compiling the code, make sure you have the file \texttt{surface\_from\_mesher.h}
-in the  \texttt{in\_out\_files/OUTPUT\_FILES/} directory. This file will be created by the solver run.
-Then type
-\begin{lyxcode}
-{\footnotesize make xcreate\_movie\_shakemap\_AVS\_DX\_GMT}{\footnotesize \par}
-\end{lyxcode}
-and run the executable \texttt{xcreate\_movie\_shakemap\_AVS\_DX\_GMT}
-in the \texttt{bin/} subdirectory. It will create visualization files in your format of choice. The code
-will prompt the user for input parameters.
-
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=.32\textwidth]{figures/movie_surf_1.jpg}
-\includegraphics[width=.32\textwidth]{figures/movie_surf_2.jpg}
-\includegraphics[width=.32\textwidth]{figures/movie_surf_3.jpg}
-\par\end{centering}
-
-\caption{Visualization using AVS files created by \texttt{xcreate\_movie\_shakemap\_AVS\_DX\_GMT} showing
-movie snapshots of vertical velocity components at different times.}
-
-\label{fig:movie.surf}
-\end{figure}
-
-\end{description}
-
-The \texttt{SPECFEM3D} code is running in near real-time to produce
-animations of southern California earthquakes via the web; see Southern California ShakeMovie\textregistered \urlwithparentheses{www.shakemovie.caltech.edu}.
-
-\subsection{\label{sub:Movie-Volume}Movie Volume}
-
-When running xspecfem3D with the \texttt{\small MOVIE\_VOLUME} flag
-turned on, the code outputs several files in \texttt{\small LOCAL\_PATH} specified in the main
-\texttt{Par\_file}, e.g. in directory \texttt{in\_out\_files/DATABASES\_MPI}.
-%As the files can be very large, there are several flags in the \texttt{\small Par\_file}
-%that control the region in space and time that is saved. These are:
-%\texttt{\small NTSTEP\_BETWEEN\_FRAMES}
-The output is saved by each processor
-at the time interval specified by \texttt{\small NTSTEP\_BETWEEN\_FRAMES}.
-For all domains, the velocity field is output to files:
-\texttt{\small proc??????\_velocity\_X\_it??????.bin},
-\texttt{\small proc??????\_velocity\_Y\_it??????.bin} and
-\texttt{\small proc??????\_velocity\_Z\_it??????.bin}.
-For elastic domains, the divergence and curl taken from the velocity field, i.e. $\nabla \cdot \bf{v}$
-and $\nabla \times \bf{v}$, get stored as well:
-\texttt{\small proc??????\_div\_it??????.bin},
-\texttt{\small proc??????\_curl\_X\_t??????.bin},
-\texttt{\small proc??????\_curl\_Y\_it??????.bin} and
-\texttt{\small proc??????\_curl\_Z\_it??????.bin}.
-The files denoted
-\texttt{\small proc??????\_div\_}~\\
-\texttt{\small glob\_it??????.bin} and
-\texttt{\small proc??????\_curl\_glob\_it??????.bin}
-are stored on the global points only, all the other arrays
-are stored on all GLL points. Note that the components X/Y/Z can change to E/N/Z according to
-the \texttt{SUPPRESS\_UTM\_PROJECTION} flag (see also Appendix~\ref{cha:Coordinates} and \ref{cha:channel-codes}).
-
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=.32\textwidth]{figures/movie_volume_1.jpg}
-\includegraphics[width=.32\textwidth]{figures/movie_volume_2.jpg}
-\includegraphics[width=.32\textwidth]{figures/movie_volume_3.jpg}
-\par\end{centering}
-
-\caption{Paraview visualization using movie volume files (converted by \texttt{xcombine\_vol\_data} and \texttt{mesh2vtu.pl}) and showing
-snapshots of vertical velocity components at different times.}
-
-\label{fig:movie.volume}
-\end{figure}
-
-To visualize these files, we use an auxilliary program \texttt{combine\_vol\_data.f90}
-to combine the data from all slices into one mesh file. To compile it in the root directory, type:
-\begin{lyxcode}
-{\footnotesize make~xcombine\_vol\_data~}{\footnotesize \par}
-\end{lyxcode}
-which will create the executable \texttt{xcombine\_vol\_data} in the directory \texttt{bin/}. To output the
-usage of this executable, type `./bin/xcombine\_vol\_data` without arguments.
-As an example, to run the executable you would use
-\begin{lyxcode}
-{\footnotesize cd~bin/~}{\footnotesize \par}
-{\footnotesize ./xcombine\_vol\_data~0~3~velocity\_Z\_it000400~../in\_out\_files/DATABASES\_MPI~../in\_out\_files/OUTPUT\_FILES~0~}{\footnotesize \par}
-\end{lyxcode}
-to create a low-resolution mesh file. The output mesh file will have the name \texttt{velocity\_Z\_it000400.mesh}.
-We next convert the \texttt{.mesh} file into the VTU (Unstructured
-grid file) format which can be viewed in ParaView. For this task, you can use and modify the script \texttt{mesh2vtu.pl}
-located in directory \texttt{utils/Visualization/Paraview/}, for example:
-\begin{lyxcode}
-{\footnotesize mesh2vtu.pl~-i~velocity\_Z\_it000400.mesh~-o~velocity\_Z\_it000400.vtu}{\footnotesize \par}
-\end{lyxcode}
-Notice that this Perl script uses a program \texttt{mesh2vtu} in the
-\texttt{utils/Visualization/Paraview/mesh2vtu} directory, which further uses the VTK \urlwithparentheses{http://www.vtk.org/}
-run-time library for its execution. Therefore, make sure you have
-them properly set in the script according to your system.
-
-\section{\label{sec:Finite-Frequency-Kernels}Finite-Frequency Kernels}
-
-The finite-frequency kernels computed as explained in Section \ref{sec:Adjoint-simulation-finite}
-are saved in the \texttt{LOCAL\_PATH} at the end of the simulation.
-Therefore, we first need to collect these files on the front end,
-combine them into one mesh file, and visualize them with some auxilliary
-programs.
-
-\begin{enumerate}
-\item \textbf{Create slice files}
-
-We will only discuss the case of one source-receiver pair, i.e., the
-so-called banana-doughnut kernels. Although it is possible to collect
-the kernel files from all slices on the front end, it usually takes
-up too much storage space (at least tens of gigabytes). Since the
-sensitivity kernels are the strongest along the source-receiver great
-circle path, it is sufficient to collect only the slices that are
-along or close to the great circle path.
-
-A Perl script \texttt{slice\_number.pl} located in directory \texttt{utils/Visualization/Paraview/} can
-help to figure out the slice numbers that lie along the great circle
-path. It applies to meshes created with the internal mesher \texttt{xmeshfem3D}.
-
-\begin{enumerate}
-\item On machines where you have access to the script, copy the \texttt{Mesh\_Par\_file}, and
-\texttt{output\_solver} files, and run:
-
-\begin{lyxcode}
-{\small slice\_number.pl~Mesh\_Par\_file~output\_solver.txt~slice\_file}{\small \par}
-\end{lyxcode}
-which will generate a \texttt{slices\_file}.
-
-\item For cases with multiple sources and multiple receivers, you need to
-provide a slice file before proceeding to the next step.
-\end{enumerate}
-\item \textbf{Collect the kernel files}
-
-After obtaining the slice files, you can collect the corresponding
-kernel files from the given slices.
-
-\begin{enumerate}
-\item You can use or modify the script \texttt{utils/copy\_basin\_database.pl}
-to accomplish this:
-
-\begin{lyxcode}
-{\small utils/copy\_database.pl~slice\_file~lsf\_machine\_file~filename~{[}jobid]}{\small \par}
-\end{lyxcode}
-where \texttt{\small lsf\_machine\_file} is the machine file generated
-by the LSF scheduler, \texttt{\small filename} is the kernel name
-(e.g., \texttt{\small rho\_kernel}, \texttt{\small alpha\_kernel}
-and \texttt{\small beta\_kernel}), and the optional \texttt{\small jobid}
-is the name of the subdirectory under \texttt{\small LOCAL\_PATH}
-where all the kernel files are stored.
-
-\item After executing this script, all the necessary mesh topology files
-as well as the kernel array files are collected to the local directory
-of the front end.
-\end{enumerate}
-\item \textbf{Combine kernel files into one mesh file}
-
-We use an auxilliary program \texttt{combine\_vol\_data.f90}
-to combine the kernel files from all slices into one mesh file.
-
-\begin{enumerate}
-\item Compile it in the root directory:
-
-\begin{lyxcode}
-{\footnotesize make~xcombine\_vol\_data~}{\footnotesize \par}
-
-{\footnotesize ./bin/xcombine\_vol\_data~slice\_list~filename~input\_dir~output\_dir~high/low-resolution~}{\footnotesize \par}
-\end{lyxcode}
-where \texttt{input\_dir} is the directory where all the individual
-kernel files are stored, and \texttt{output\_dir} is where the mesh
-file will be written.
-
-\item Use 1 for a high-resolution mesh, outputting all the GLL points to
-the mesh file, or use 0 for low resolution, outputting only the corner
-points of the elements to the mesh file.
-\item The output mesh file will have the name \texttt{filename\_rho(alpha,beta).mesh}
-\end{enumerate}
-\item \textbf{Convert mesh files into .vtu files}
-
-\begin{enumerate}
-\item We next convert the \texttt{.mesh} file into the VTU (Unstructured
-grid file) format which can be viewed in ParaView. For this task, you can use and modify the script \texttt{mesh2vtu.pl}
-located in directory \texttt{utils/Visualization/Paraview/}, for example:
-
-\begin{lyxcode}
-{\footnotesize mesh2vtu.pl~-i~file.mesh~-o~file.vtu~}{\footnotesize \par}
-\end{lyxcode}
-\item Notice that this Perl script uses a program \texttt{mesh2vtu} in the
-\texttt{utils/Visualization/Paraview/mesh2vtu} directory, which further uses the VTK \urlwithparentheses{http://www.vtk.org/}
-run-time library for its execution. Therefore, make sure you have
-them properly set in the script according to your system.
-\end{enumerate}
-\item \textbf{Copy over the source and receiver .vtk file}
-
-In the case of a single source and a single receiver, the simulation
-also generates the file \texttt{sr.vtk} located in the
-\texttt{in\_out\_files/OUTPUT\_FILES/} directory to describe
-the source and receiver locations, which can also be viewed in Paraview
-in the next step.
-
-\item \textbf{View the mesh in ParaView}
-
-Finally, we can view the mesh in ParaView \urlwithparentheses{www.paraview.org}.
-
-\begin{enumerate}
-\item Open ParaView.
-\item From the top menu, \textsf{File} $\rightarrow$\textsf{Open data},
-select \texttt{file.vtu}, and click the \textsf{Accept} button.
-
-\begin{itemize}
-\item If the mesh file is of moderate size, it shows up on the screen; otherwise,
-only the bounding box is shown.
-\end{itemize}
-\item Click \textsf{Display Tab} $\rightarrow$ \textsf{Display Style} $\rightarrow$
-\textsf{Representation} and select \textsf{wireframe of surface} to
-display it.
-\item To create a cross-section of the volumetric mesh, choose \textsf{Filter}
-$\rightarrow$ \textsf{cut}, and under \textsf{Parameters Tab}, choose
-\textsf{Cut Function} $\rightarrow$ \textsf{plane}.
-\item Fill in center and normal information given by the \texttt{global\_slice\_number.pl}
-script (either from the standard output or from \texttt{normal\_plane.txt}
-file).
-\item To change the color scale, go to \textsf{Display Tab} $\rightarrow$
-\textsf{Color} $\rightarrow$ \textsf{Edit Color Map} and reselect
-lower and upper limits, or change the color scheme.
-\item Now load in the source and receiver location file by \textsf{File}
-$\rightarrow$ \textsf{Open data}, select \texttt{sr.vt}k, and click
-the \textsf{Accept} button. Choose \textsf{Filter} $\rightarrow$
-\textsf{Glyph}, and represent the points by `\textsf{spheres}'.
-\item For more information about ParaView, see the ParaView Users Guide \urlwithparentheses{www.paraview.org/files/v1.6/ParaViewUsersGuide.PDF}.
-\end{enumerate}
-\end{enumerate}
-%
-\begin{figure}[H]
-\noindent \begin{centering}
-\includegraphics[scale=0.6]{figures/3D-S-Kernel}
-\par\end{centering}
-
-\caption{(a) Top Panel: Vertical source-receiver cross-section of the S-wave
-finite-frequency sensitivity kernel $K_{\beta}$ for station GSC at
-an epicentral distance of 176 km from the September 3, 2002, Yorba
-Linda earthquake. Lower Panel: Vertical source-receiver cross-section
-of the 3D S-wave speed model used for the spectral-element simulations
-\citep{KoLiTrSuStSh04}. (b) The same as (a) but for station HEC at
-an epicentral distance of 165 km \citep{LiTr06}.}
-
-\label{figure:P-wave-speed-finite-frequency}
-\end{figure}
-
-\chapter{\label{cha:Scheduler}Running through a Scheduler}
-
-The code is usually run on large parallel machines, often PC clusters,
-most of which use schedulers, i.e., queuing or batch management systems
-to manage the running of jobs from a large number of users. The following
-considerations need to be taken into account when running on a system
-that uses a scheduler:
-
-\begin{itemize}
-\item The processors/nodes to be used for each run are assigned dynamically
-by the scheduler, based on availability. Therefore, in order for the
-\texttt{xgenerate\_databases} and the \texttt{xspecfem3D} executables (or between successive runs of the solver) to
-have access to the same database files (if they are stored on hard
-drives local to the nodes on which the code is run), they must be
-launched in sequence as a single job.
-\item On some systems, the nodes to which running jobs are assigned are
-not configured for compilation. It may therefore be necessary to pre-compile
-both the \texttt{xgenerate\_databases} and the \texttt{xspecfem3D} executables.
-\item One feature of schedulers/queuing systems is that they allow submission
-of multiple jobs in a ``launch and forget'' mode. In order to take
-advantage of this property, care needs to be taken that output and
-intermediate files from separate jobs do not overwrite each other,
-or otherwise interfere with other running jobs.
-\end{itemize}
-Examples of job scripts can be found in the \texttt{\small utils/Cluster/}
-directory and can straightforwardly be modified and adapted to meet
-more specific running needs.
-
-We describe here in some detail a job submission procedure for the
-Caltech 1024-node cluster, CITerra, under the LSF scheduling system.
-We consider the submission of a regular forward simulation using the internal mesher
-to create mesh partitions. The two
-main scripts are \texttt{\small run\_lsf.bash}, which compiles the
-Fortran code and submits the job to the scheduler, and \texttt{\small go\_mesher\_solver\_lsf\_basin.forward},
-which contains the instructions that make up
-the job itself. These scripts can be found in \texttt{\small utils/Cluster/lsf/} directory
-
-\section{Job submission \texttt{run\_lsf.bash}}
-
-This script first sets the job queue to be `normal'. It then compiles
-the mesher, database generator and solver together, figures out the number of processors
-required for this simulation from the \texttt{in\_data\_files/Par\_file}, and
-submits the LSF job.
-
-\begin{lyxcode}
-\#!/bin/bash
-
-\#~use~the~normal~queue~unless~otherwise~directed~queue=\char`\"{}-q~normal\char`\"{}~
-
-if~{[}~\$\#~-eq~1~];~then
-
-~~~~~~~~echo~\char`\"{}Setting~the~queue~to~\$1\char`\"{}
-
-~~~~~~~~queue=\char`\"{}-q~\$1\char`\"{}~
-
-fi~\\
-~\\
-\#~compile~the~mesher~and~the~solver~
-
-d=`date'~echo~\char`\"{}Starting~compilation~\$d\char`\"{}~
-
-make~clean~
-
-make~xmeshfem3D~
-
-make~xgenerate\_databases~
-
-make~xspecfem3D~
-
-d=`date'~
-
-echo~\char`\"{}Finished~compilation~\$d\char`\"{}~\\
-~\\
-\#~get~total~number~of~nodes~needed~for~solver~
-
-NPROC=`grep~NPROC~in\_data\_files/Par\_file~|~cut~-c~34-~'~\\
-~\\
-
-\#~compute~total~number~of~nodes~needed~for~mesher~
-
-NPROC\_XI=`grep~NPROC\_XI~in\_data\_files/meshfem3D\_files/Mesh\_Par\_file~|~cut~-c~34-~'~
-~\\
-
-NPROC\_ETA=`grep~NPROC\_ETA~in\_data\_files/meshfem3D\_files/Mesh\_Par\_file~|~cut~-c~34-~'~
-~\\
-\#~total~number~of~nodes~is~the~product~of~the~values~read~
-
-numnodes=\$((~\$NPROC\_XI~*~\$NPROC\_ETA~))~\\
-~\\
-
-\#~checks~total~number~of~nodes~
-
-if~{[}~\$numnodes~-neq~\$NPROC~];~then
-
-~~~~~~~~echo~\char`\"{}error number of procs mismatch\char`\"{}
-
-~~~~~~~~exit~
-
-fi~\\
-
-~\\
-echo~\char`\"{}Submitting~job\char`\"{}~
-
-bsub~\$queue~-n~\$numnodes~-W~60~-K~<go\_mesher\_solver\_lsf.forward~
-\end{lyxcode}
-
-\section{Job script \texttt{go\_mesher\_solver\_lsf.forward}}
-
-This script describes the job itself, including setup steps that can
-only be done once the scheduler has assigned a job-ID and a set of
-compute nodes to the job, the \texttt{run\_lsf.bash} commands used
-to run the mesher, database generator and the solver, and calls to scripts that collect
-the output seismograms from the compute nodes and perform clean-up
-operations.
-
-\begin{enumerate}
-\item First the script directs the scheduler to save its own output and
-output from \texttt{stdout} into ~\\
-\texttt{\small in\_out\_files/OUTPUT\_FILES/\%J.o},
-where \texttt{\%J} is short-hand for the job-ID; it also tells the
-scheduler what version of \texttt{mpich} to use (\texttt{mpich\_gm})
-and how to name this job (\texttt{go\_mesher\_solver\_lsf}).
-\item The script then creates a list of the nodes allocated to this job
-by echoing the value of a dynamically set environment variable \texttt{LSB\_MCPU\_HOSTS}
-and parsing the output into a one-column list using the Perl script
-\texttt{utils/Cluster/lsf/remap\_lsf\_machines.pl}. It then creates a set of scratch
-directories on these nodes (\texttt{\small /scratch/}~\\
-\texttt{\small \$USER/DATABASES\_MPI}) to be used as the \texttt{LOCAL\_PATH}
-for temporary storage of the database files. The scratch directories
-are created using \texttt{shmux}, a shell multiplexor that can execute
-the same commands on many hosts in parallel. \texttt{shmux} is available
-from Shmux \urlwithparentheses{web.taranis.org/shmux/}. Make sure that the \texttt{LOCAL\_PATH}
-parameter in \texttt{in\_data\_files/Par\_file} is also set properly.
-\item The next portion of the script launches the mesher, database generator and then the solver
-using \texttt{run\_lsf.bash}.
-\item The final portion of the script collects the seismograms and performs
-clean up on the nodes, using the Perl scripts \texttt{collect\_seismo\_lsf\_multi.pl}
-and \texttt{cleanmulti.pl}.
-\end{enumerate}
-\begin{lyxcode}
-\#!/bin/bash~-v~
-
-\#BSUB~-o~in\_out\_files/OUTPUT\_FILES/\%J.o~
-
-\#BSUB~-a~mpich\_gm~
-
-\#BSUB~-J~go\_mesher\_solver\_lsf~\\
-~\\
-\#~set~up~local~scratch~directories
-
-BASEMPIDIR=/scratch/\$USER/DATABASES\_MPI
-
-mkdir~-p~in\_out\_files/OUTPUT\_FILES~
-
-echo~\char`\"{}\$LSB\_MCPU\_HOSTS\char`\"{}~>~in\_out\_files/OUTPUT\_FILES/lsf\_machines~
-
-echo~\char`\"{}\$LSB\_JOBID\char`\"{}~>~in\_out\_files/OUTPUT\_FILES/jobid
-
-remap\_lsf\_machines.pl~in\_out\_files/OUTPUT\_FILES/lsf\_machines~> in\_out\_files/OUTPUT\_FILES/machines
-
-shmux~-M50~-Sall~-c~\char`\"{}rm~-r~-f~/scratch/\$USER;~\textbackslash{}
-
-~~~~~~mkdir~-p~/scratch/\$USER;~mkdir~-p~\$BASEMPIDIR\char`\"{}~\textbackslash{}
-
-~~~~~~-~<~in\_out\_files/OUTPUT\_FILES/machines~>/dev/null~\\
-~\\
-\#~run~the~specfem~program
-
-current\_pwd=\$PWD \\
-
-cd bin/ \\
-
-run\_lsf.bash~-{}-gm-no-shmem~-{}-gm-copy-env~\$current\_pwd/xmeshfem3D~\\
-
-run\_lsf.bash~-{}-gm-no-shmem~-{}-gm-copy-env~\$current\_pwd/xgenerate\_databases~\\
-
-run\_lsf.bash~-{}-gm-no-shmem~-{}-gm-copy-env~\$current\_pwd/xspecfem3D~\\
-~\\
-\#~collect~seismograms~and~clean~up
-
-cd current\_pwd/~ \\
-
-mkdir~-p~in\_out\_files/SEM
-
-cd~in\_out\_files/SEM/~
-
-collect\_seismo.pl~../OUTPUT\_FILES/lsf\_machines
-
-cleanbase.pl~../OUTPUT\_FILES/machines
-\end{lyxcode}
-
-\chapter{{\normalsize \label{cha:-Changing-the}} Changing the Model}
-
-In this section we explain how to change the velocity model used for your simulations.
-These changes involve contributing specific subroutines
-that replace existing subroutines in the \texttt{SPECFEM3D} package.
-Note that \texttt{SPECFEM3D} can handle Earth models with material properties that vary within each spectral element.
-
-%% magnoni 6/12
-\section{{\normalsize \label{sec:Using-tomographic}}Using external tomographic Earth models}
-
-To implement your own external tomographic model, you must provide your own tomography file
-\texttt{tomography\_model.xyz} and choose between two possible options:
-(1) set in the \texttt{Par\_file} the model parameter \texttt{MODEL = tomo}, or
-(2) for more user control, set in the \texttt{Par\_file} the model parameter \texttt{MODEL = default} and use the following format in the file \texttt{MESH/nummaterial\_velocity\_file} when using CUBIT to construct your mesh (see also section \ref{subsec:Exporting-the-Mesh}):
-\begin{lyxcode}
-\texttt{domain\_ID} \texttt{material\_ID} \texttt{tomography} \texttt{elastic} \texttt{file\_name} 1
-\end{lyxcode}
-where
-\texttt{domain\_ID} is 1 for acoustic or 2 for elastic materials,
-\texttt{material\_ID} a negative, unique identifier (i.e., -1),
-\texttt{tomography} keyword for tomographic material definition,
-\texttt{elastic} keyword for elastic material definition,
-\texttt{file\_name} the name of the tomography file and
-1 a positive unique identifier.
-The name of the tomography file is not used so far, rather change the filename (\texttt{TOMO\_FILENAME}) defined in the file \texttt{model\_tomography.f90} located in the \texttt{src/generate\_databases/} directory.
-
-The external tomographic model is represented by a grid of points with assigned material properties and homogeneous resolution along each spatial direction x, y and z.
-The ASCII file \texttt{tomography\_model.xyz} that describes the tomography should be located in the \texttt{in\_data\_files/} directory.
-The format of the file, as read from \texttt{model\_tomography.f90}, looks like Figure \ref{fig:tomography_file},
-where
-%
-\begin{description}
-\item [{\texttt{ORIG\_X}, \texttt{END\_X}}] are, respectively, the coordinates of the initial and final tomographic grid points along the x direction (in the mesh units, e.g., $m$);
-\item [{\texttt{ORIG\_Y}, \texttt{END\_Y}}] respectively the coordinates of the initial and final tomographic grid points along the y direction (in the mesh units, e.g., $m$);
-\item [{\texttt{ORIG\_Z}, \texttt{END\_Z}}] respectively the coordinates of the initial and final tomographic grid points along the z direction (in the mesh units, e.g., $m$);
-\item [\texttt{SPACING\_X}, \texttt{SPACING\_Y}, \texttt{SPACING\_Z}] the spacing between the tomographic grid points along the x, y and z directions, respectively (in the mesh units, e.g., $m$);
-\item [\texttt{NX}, \texttt{NY}, \texttt{NZ}] the number of grid points along the spatial directions x, y and z, respectively; \texttt{NX} is given by [(\texttt{END\_X} - \texttt{ORIG\_X})/\texttt{SPACING\_X}]+1; \texttt{NY} and \texttt{NZ} are the same as \texttt{NX}, but for the y and z directions, respectively;
-\item [\texttt{VP\_MIN}, \texttt{VP\_MAX}, \texttt{VS\_MIN}, \texttt{VS\_MAX}, \texttt{RHO\_MIN}, \texttt{RHO\_MAX}] the minimum and maximum values of the wave speed \texttt{vp} and \texttt{vs} (in $m \, s^{-1}$) and of the density \texttt{rho} (in $kg \, m^{-3}$); these values could be the actual limits of the tomographic parameters in the grid or the minimum and maximum values to which we force the cut of velocity and density in the model.
-\end{description}
-%
-After the first four lines, in the file \texttt{tomography\_model.xyz} the tomographic grid points are listed with the corresponding values of \texttt{vp}, \texttt{vs} and \texttt{rho}, scanning the grid along the x coordinate (from \texttt{ORIG\_X} to \texttt{END\_X} with step of \texttt{SPACING\_X}) for each given y (from \texttt{ORIG\_Y} to \texttt{END\_Y}, with step of \texttt{SPACING\_Y}) and each given z (from \texttt{ORIG\_Z} to \texttt{END\_Z}, with step of \texttt{SPACING\_Z}).
-%fig
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=1.0\textwidth]{figures/tomo_file.jpg}
-\par\end{centering}
-\caption{Tomography file \texttt{tomography\_model.xyz} that describes an external Earth model and that is read by \texttt{model\_tomography.f90}.
-The coordinates x, y and z of the grid, their limits ORIG\_X, ORIG\_Y, ORIG\_Z, END\_X, END\_Y, END\_Z and the grid spacings SPACING\_X, SPACING\_Y, SPACING\_Z should be in the units of the constructed mesh (e.g., UTM coordinates in meters).}
-\label{fig:tomography_file}
-\end{figure}
-
-The user can implement his own interpolation algorithm for the tomography model by changing the routine \texttt{model\_tomography.f90} located in the \texttt{src/generate\_databases/} directory.
-Moreover, for models that involve both fully defined materials and a tomography description, the \texttt{nummaterial\_velocity\_file} has multiple lines each with the corresponding suitable format described above.
-
-\section{{\normalsize \label{sec:Anelastic-Models}}External (an)elastic Models}
-
-To use your own external model, you can set in the \texttt{Par\_file} the model parameter \texttt{MODEL = external}.
-Three-dimensional acoustic and/or (an)elastic (attenuation) models may then be superimposed
-onto the mesh based upon your subroutine in
-\texttt{model\_external\_values.f90} located in directory \texttt{src/generate\_databases}.
-The call to this routine would be as follows
-
-\begin{lyxcode}
-call~model\_external\_values(xmesh,~ymesh,~zmesh, ~\& \\
-~~~~~~~~~~rho,~vp,~vs,~qmu\_atten,~iflag\_aniso,~idomain\_id)
-\end{lyxcode}
-%
-Input to this routine consists of:
-\begin{description}
-\item [{\texttt{xmesh,ymesh,zmesh}}] location of mesh point
-\end{description}
-%
-Output to this routine consists of:
-\begin{description}
-\item [{\texttt{rho,vp,vs}}] isotropic model parameters for density $\rho$ ($kg/m^3$), $v_p$ ($m/s$) and $v_s$ ($m/s$)
-\item [{\texttt{qmu\_atten}}] Shear wave quality factor: $0<Q_{\mu}<9000$
-\item [{\texttt{iflag\_aniso}}] anisotropic model flag, $0$ indicating no anisotropy or $1$ using anisotropic model parameters as defined in routine file \texttt{model\_aniso.f90}
-\item [{\texttt{idomain\_id}}] domain identifier, $1$ for acoustic, $2$ for elastic, $3$ for poroelastic materials.
-\end{description}
-
-Note that the resolution and maximum value of anelastic models are
-truncated. This speeds the construction of the standard linear solids
-during the meshing stage. To change the resolution, currently at one
-significant figure following the decimal, or the maximum value (9000),
-consult \texttt{shared/constants.h}.
-
-\section{{\normalsize \label{sec:Anisotropic-Models}}Anisotropic Models}
-
-To use your anisotropic models, you can either set in the \texttt{Par\_file} the model parameter \texttt{ANISOTROPY} to \texttt{.true.} or set the model parameter \texttt{MODEL = aniso}.
-Three-dimensional anisotropic models may be superimposed on
-the mesh based upon the subroutine in file \texttt{model\_aniso.f90}
-located in directory \texttt{src/generate\_databases}.
-The call to this subroutine is of the form
-
-\begin{lyxcode}
-call~model\_aniso(iflag\_aniso,~rho,~vp,~vs,~\& \\
-~~~~~~~~~~c11,c12,c13,c14,c15,c16,c22,c23,c24,c25,c26,~\& \\
-~~~~~~~~~~c33,c34,c35,c36,c44,c45,c46,c55,c56,c66)~
-\end{lyxcode}
-%
-Input to this routine consists of:
-\begin{description}
-\item [{\texttt{iflag\_aniso}}] flag indicating the type of the anisotropic model, i.e. $0$ for no superimposed anisotropy, $1$ or $2$ for generic pre-defined anisotropic models.
-\item [{\texttt{rho,vp,vs}}] reference isotropic model parameters for density $\rho$, $v_p$ and $v_s$.
-\end{description}
-%
-Output from the routine consists of the following non-dimensional
-model parameters:
-\begin{description}
-\item [{\texttt{c11},}] \textbf{$\cdots$,} \texttt{\textbf{c66}} 21~dimensionalized
-anisotropic elastic parameters.
-\end{description}
-
-You can replace the \texttt{model\_aniso.f90} file by
-your own version \textit{provided you do not change the call structure
-of the routine}, i.e., the new routine should take exactly the same
-input and produce the required relative output.
-
-\chapter{Post-Processing Scripts}
-
-Several post-processing scripts/programs are provided in the \texttt{utils/}
-directory, and most of them need to be adjusted when used on different
-systems, for example, the path of the executable programs. Here we
-only list a few of the available scripts and provide a brief description, and
-you can either refer to the related sections for detailed usage or,
-in a lot of cases, type the script/program name without arguments
-for its usage.
-
-\section{Process Data and Synthetics\label{sec:Process-data-and-syn}}
-
-In many cases, the SEM synthetics are calculated and compared to data
-seismograms recorded at seismic stations. Since the SEM synthetics
-are accurate for a certain frequency range, both the original data
-and the synthetics need to be processed before a comparison can be
-made.
-
- For such comparisons, the following steps are recommended:
-
-\begin{enumerate}
-\item Make sure that both synthetic and observed seismograms have the correct station/event and timing information.
-\item Convolve synthetic seismograms with a source time function with the half duration specified in the \texttt{CMTSOLUTION} file, provided, as recommended, you used a zero half duration in the SEM simulations.
-\item Resample both observed and synthetic seismograms to a common sampling rate.
-\item Cut the records using the same window.
-\item Remove the trend and mean from the records and taper them.
-\item Remove the instrument response from the observed seismograms (recommended) or convolve the synthetic seismograms with the instrument response.
-\item Make sure that you apply the same filters to both observed and synthetic seismograms. Preferably, avoid filtering your records more than once.
-\item Now, you are ready to compare your synthetic and observed seismograms.
-\end{enumerate}
-
-\noindent
-We generally use the following scripts provided in the \texttt{utils/seis\_process/} directory:
-
-\subsection{Data processing script \texttt{process\_data.pl}}
-
-This script cuts a given portion of the original data, filters it,
-transfers the data into a displacement record, and picks the first
-P and S arrivals. For more functionality, type `\texttt{process\_data.pl}'
-without any argument. An example of the usage of the script:
-
-\begin{lyxcode}
-{\footnotesize process\_data.pl~-m~CMTSOLUTION~-s~1.0~-l~0/4000~-i~-f~-t~40/500~-p~-x~bp~DATA/1999.330{*}.BH?.SAC}
-{\footnotesize \par}
-\end{lyxcode}
-
-\noindent which has resampled the SAC files to a sampling rate of 1 seconds, cut
-them between 0 and 4000 seconds, transfered them into displacement
-records and filtered them between 40 and 500 seconds,
-picked the first P and S arrivals, and added suffix `\texttt{bp}' to the file names.
-
-Note that all of the scripts in this section actually use the SAC
-and/or IASP91 to do the core operations; therefore make sure that
-the SAC and IASP91 packages are installed properly on your system,
-and that all the environment variables are set properly before running
-these scripts.
-
-\subsection{Synthetics processing script \texttt{process\_syn.pl}}
-
-This script converts the synthetic output from the SEM code from ASCII
-to SAC format, and performs similar operations as `\texttt{process\_data.pl}'.
-An example of the usage of the script:
-\begin{lyxcode}
-{\footnotesize process\_syn.pl~-m~CMTSOLUTION~-h~-a~STATIONS~-s~1.0~-l~0/4000~-f~-t~40/500~-p~-x~bp~SEM/{*}.BX?.semd}
-{\footnotesize \par}
-\end{lyxcode}
-which will convolve the synthetics with a triangular source-time function
-from the \texttt{CMTSOLUTION} file, convert the synthetics into SAC
-format, add event and station information into the SAC headers, resample the SAC files with a sampling rate of 1 seconds, cut them between 0 and 4000 seconds, filter them between 40 and
-500 seconds with the same filter used for the observed data, pick the first P and S arrivals, and add the suffix `\texttt{bp}'
-to the file names.
-
-More options are available for this script, such as adding time shift
-to the origin time of the synthetics, convolving the synthetics with
-a triangular source time function with a given half duration, etc.
-Type \texttt{process\_syn.pl} without any argument for a detailed
-usage.
-
-In order to convert between SAC format and ASCII files, useful scripts
-are provided in the subdirectories ~\\
-\texttt{utils/sac2000\_alpha\_convert/} and \texttt{utils/seis\_process/asc2sac/}.
-
-\subsection{Script \texttt{rotate.pl}}
-
-The original data and synthetics have three components: vertical (BHZ resp. BXZ),
-north (BHN resp. BXN) and east (BHE resp. BXE). However, for most seismology applications,
-transverse and radial components are also desirable. Therefore, we
-need to rotate the horizontal components of both the data and the
-synthetics to the transverse and radial direction, and \texttt{\small rotate.pl}
-can be used to accomplish this:
-
-\begin{lyxcode}
-rotate.pl~-l~0~-L~180~-d~DATA/{*}.BHE.SAC.bp~
-
-rotate.pl~-l~0~-L~180~SEM/{*}.BXE.semd.sac.bp~
-\end{lyxcode}
-where the first command performs rotation on the SAC data obtained
-through Seismogram Transfer Program (STP) \urlwithparentheses{http://www.data.scec.org/STP/stp.html},
-while the second command rotates the processed SEM synthetics.
-
-\section{Collect Synthetic Seismograms}
-
-The forward and adjoint simulations generate synthetic seismograms
-in the \texttt{in\_out\_files/OUTPUT\_FILES/} directory by default. For the forward simulation, the files
-are named like \texttt{STA.NT.BX?.semd} for two-column time series, or
-\texttt{STA.NT.BX?.semd.sac} for ASCII SAC format, where STA and NT
-are the station name and network code, and \texttt{BX?} stands for
-the component name. Please see the Appendix~\ref{cha:Coordinates} and \ref{cha:channel-codes}
-for further details.
-
-The adjont simulations generate synthetic seismograms
-with the name \texttt{S?????.NT.S??.sem} (refer to Section \ref{sec:Adjoint-simulation-sources}
-for details). The kernel simulations output the back-reconstructed
-synthetic seismogram in the name \texttt{STA.NT.BX?.semd}, mainly
-for the purpose of checking the accuracy of the reconstruction. Refer
-to Section \ref{sec:Adjoint-simulation-finite} for further details.
-
-You do have further options to change this default output behavior, given in the main constants file
-\texttt{constants.h} located in \texttt{src/shared/} directory:
-\begin{description}
-\item [{\texttt{SEISMOGRAMS\_BINARY}}] set to \texttt{.true.} to have seismograms written out in binary format.
-\item [{\texttt{WRITE\_SEISMOGRAMS\_BY\_MASTER}}] Set to \texttt{.true.} to have only the master process writing out seismograms.
-This can be useful on a cluster, where only the master process node has access to the output directory.
-\item [{\texttt{USE\_OUTPUT\_FILES\_PATH}}] Set to \texttt{.false.} to have the seismograms output to \texttt{LOCAL\_PATH} directory specified in the main parameter file \texttt{in\_data\_files/Par\_file}.
-In this case, you could collect the synthetics onto the frontend using the \texttt{collect\_seismo\_lsf\_multi.pl} script
-located in the \texttt{utils/Cluster/lsf/} directory. The usage of the script would be e.g.:
-
-\begin{lyxcode}
-collect\_seismo.pl~machines~in\_data\_files/Par\_file
-\end{lyxcode}
-
-\end{description}
-where  \texttt{machines} is a file containing the node names and  \texttt{in\_data\_files/Par\_file} the parameter file used to extract the \texttt{LOCAL\_PATH} directory used for the simulation.
-
-\section{Clean Local Database}
-
-After all the simulations are done, the seismograms are collected,
-and the useful database files are copied to the frontend, you may
-need to clean the local scratch disk for the next simulation. This
-is especially important in the case of kernel simulation,
-where very large files are generated for the absorbing boundaries
-to help with the reconstruction of the regular forward wavefield.
-A sample script is provided in \texttt{utils/}:
-
-\begin{lyxcode}
-cleanbase.pl~machines
-\end{lyxcode}
-where  \texttt{machines} is a file containing the node names.
-
-\section{Plot Movie Snapshots and Synthetic Shakemaps}
-
-\subsection{Script \texttt{movie2gif.gmt.pl}}
-
-With the movie data saved in \texttt{in\_out\_files/OUTPUT\_FILES/} at the end of
-a movie simulation (\texttt{\small MOVIE\_SURFACE=.true.}), you can
-run the \texttt{`create\_movie\_shakemap\_AVS\_DX\_GMT}' code to convert these binary
-movie data into GMT xyz files for futher processing. A sample script
-\texttt{movie2gif.gmt.pl} is provided to do this conversion, and then
-plot the movie snapshots in GMT, for example:
-
-\begin{lyxcode}
-movie2gif.gmt.pl~-m~CMTSOLUTION~-g~-f~1/40~-n~-2~-p~
-\end{lyxcode}
-which for the first through the 40th movie frame, converts the \texttt{moviedata}
-files into GMT xyz files, interpolates them using the 'nearneighbor'
-command in GMT, and plots them on a 2D topography map. Note that `\texttt{-2}'
-and `\texttt{-p}' are both optional.
-
-\subsection{Script \texttt{plot\_shakemap.gmt.pl}}
-
-With the shakemap data saved in \texttt{in\_out\_files/OUTPUT\_FILES/} at the end
-of a shakemap simulation ~\\
-(\texttt{CREATE\_SHAKEMAP=.true.}), you can
-also run \texttt{`create\_movie\_shakemap\_AVS\_DX\_GMT}' code to convert the binary
-shakemap data into GMT xyz files. A sample script \texttt{plot\_shakemap.gmt.pl}
-is provided to do this conversion, and then plot the shakemaps in
-GMT, for example:
-
-\begin{lyxcode}
-plot\_shakemap.gmt.pl~data\_dir~type(1,2,3)~CMTSOLUTION~
-\end{lyxcode}
-where \texttt{type=1} for a displacement shakemap, \texttt{2} for
-velocity, and \texttt{3} for acceleration.
-
-\section{Map Local Database}
-
-A sample program \texttt{remap\_database} is provided to map the local
-database from a set of machines to another set of machines. This is
-especially useful when you want to run mesher and solver, or different
-types of solvers separately through a scheduler (refer to Chapter~\ref{cha:Scheduler}).
-\begin{lyxcode}
-run\_lsf.bash~-{}-gm-no-shmem~-{}-gm-copy-env~remap\_database~old\_machines~150
-\end{lyxcode}
-where \texttt{old\_machines} is the LSF machine file used in the previous
-simulation, and \texttt{150} is the number of processors in total.
-
-\chapter{\label{cha:tomo}Tomographic inversion using sensitivity kernels}
-
-One of the fundamental reasons for computing sensitivity kernels (Section~\ref{sec:Adjoint-simulation-finite}) is to use them within a tomographic inversion.
-In other words, use recorded seismograms, make measurements with synthetic seismograms, and use the misfit between the two to iteratively improve the model described by (at least) $V_{\rm p}$, $V_{\rm s}$, and $\rho$ as a function of space.
-
-Whatever misfit function you use for the tomographic inversion (several examples in \cite{TrTaLi05} and \cite{TrKoLi08}), you will weight the sensitivity kernels with measurements.
-Furthermore, you will use as many measurements (stations, components, time windows) as possible per event; hence, we call these composite kernels ``event kernels,'' which are volumetric fields representing the gradient of the misfit function with respect to one of the variables (\eg $V_{\rm s}$).
-The basic features of an adjoint-based tomographic inversion were illustrated in \citet{TaLiTr07} and \citet{TrKoLi08} using a conjugate-gradient algorithm; there are dozens of versions of gradient-based inversion algorithms that could alternatively be used.
-The tomographic inversion of \citet{TaLiMaTr09,TaLiMaTr2010} used SPECFEM3D as well as several additional components which are also stored on the CIG svn server, described next.
-
-The directory containing utilities for tomographic inversion using SPECFEM3D (or other packages that evaluate misfit functions and gradients) is here on the CIG svn server:
-%
-\begin{verbatim}
-/cig/seismo/3D/ADJOINT_TOMO/
-   flexwin/     -- FLEXWIN algorithm for automated picking of time windows
-   measure_adj/ -- reads FLEXWIN output file and makes measurements,
-                   with the option for computing adjoint sources
-   iterate_adj/ -- various tools for iterative inversion
-                   (requires pre-computed "event kernels")
-\end{verbatim}
-%
-This directory also contains a brief \verb+README+ file indicating the role of the three subdirectories, \verb+flexwin+ \citep{Maggi2009}, \verb+measure_adj+, and \verb+iterate_adj+.
-The components for making the model update are there; however, there are no explicit rules for computing the model update, just as with any optimization problem.
-There are options for computing a conjugate gradient step, as well as a source subspace projection step.
-The workflow could use substantial ``stitching together,'' and we welcome suggestions; undoubtedly, the tomographic inversion procedure will improve with time.
-{\bf This is a work in progress.}
-
-The best single file to read is probably: \verb+ADJOINT_TOMO/iterate_adj/cluster/README+.
-
-%============================================================
-
-\chapter*{\label{cha:Bug-Reports-and}Bug Reports and Suggestions for Improvements}
-
-To report bugs or suggest improvements to the code, please send an
-e-mail to the CIG Computational Seismology Mailing List \urlwithparentheses{cig-seismo@geodynamics.org}
-or Jeroen Tromp \urlwithparentheses{jtromp-AT-princeton.edu}, and/or use our online
-bug tracking system Roundup \urlwithparentheses{www.geodynamics.org/roundup}.
-
-\addcontentsline{toc}{chapter}{Notes \& Acknowledgments}
 \chapter*{Notes \& Acknowledgments}
@@
 are based in part on software libraries from the Massachusetts Institute
-of Technology, Department of Mechanical Engineering (Cambridge, Massachusetts, USA).
-The non-structured global numbering software was provided by Paul
-F. Fischer (Brown University, Providence, Rhode Island, USA, now at Argonne National Laboratory, USA).
+of Technology, Department of Mechanical Engineering (Cambridge, Massachusetts,
+USA). The non-structured global numbering software was provided by
+Paul F. Fischer (Brown University, Providence, Rhode Island, USA,
+now at Argonne National Laboratory, USA).
 
-OpenDX \urlwithparentheses{http://www.opendx.org} is open-source based on IBM Data
-Explorer, AVS \urlwithparentheses{http://www.avs.com} is a trademark of Advanced
-Visualization Systems, and ParaView \urlwithparentheses{http://www.paraview.org}
+OpenDX \urlwithparentheses{http://www.opendx.org} is open-source
+based on IBM Data Explorer, AVS \urlwithparentheses{http://www.avs.com}
+is a trademark of Advanced Visualization Systems, and ParaView \urlwithparentheses{http://www.paraview.org}
 is an open-source visualization platform.
@@
 Please e-mail your feedback, questions, comments, and suggestions
-to Jeroen Tromp \urlwithparentheses{jtromp-AT-princeton.edu} or to the CIG Computational Seismology Mailing List \urlwithparentheses{cig-seismo@geodynamics.org}.
-
-\addcontentsline{toc}{chapter}{Copyright}
-\chapter*{Copyright}
-
-Main authors: Dimitri Komatitsch and Jeroen Tromp
-
-Princeton University, USA, and CNRS / INRIA / University of Pau, France
-
-$\copyright$ Princeton University and CNRS / INRIA / University of Pau, July 2012
-
-This program is free software; you can redistribute it and/or modify
-it under the terms of the GNU General Public License as published
-by the Free Software Foundation (see Appendix \ref{cha:License}).\\
-
-\textbf{\underline{Evolution of the code:}}\\
-
- MPI v 2.1, July 2012:
- Max Rietmann, Peter Messmer, Daniel Peter, Dimitri Komatitsch, Joseph Charles, Zhinan Xie: support for CUDA GPUs,
-better CFL stability for the Stacey absorbing conditions. \\
-
- MPI v. 2.0, November 2010:
- Dimitri Komatitsch, Nicolas Le Goff, Roland Martin and Pieyre Le Loher, University of Pau, France,
- Jeroen Tromp and the Princeton group of developers, Princeton University, USA,
- and Emanuele Casarotti, INGV Roma, Italy:
-  support for CUBIT meshes decomposed by SCOTCH, METIS or ZOLTAN;
-  much faster solver using Michel Deville's inlined matrix products.\\
-
- MPI v. 1.4 Dimitri Komatitsch, University of Pau, Qinya Liu and others, Caltech, September 2006:
-  better adjoint and kernel calculations, faster and better I/Os
-  on very large systems, many small improvements and bug fixes.\\
-
- MPI v. 1.3 Dimitri Komatitsch, University of Pau, and Qinya Liu, Caltech, July 2005:
-  serial version, regular mesh, adjoint and kernel calculations, ParaView support.\\
-
- MPI v. 1.2 Min Chen and Dimitri Komatitsch, Caltech, July 2004:
-  full anisotropy, volume movie.\\
-
- MPI v. 1.1 Dimitri Komatitsch, Caltech, October 2002: Zhu's Moho map, scaling
-  of $V_s$ with depth, Hauksson's regional model, attenuation, oceans, movies.\\
-
- MPI v. 1.0 Dimitri Komatitsch, Caltech, USA, May 2002: first MPI version based on global code.\\
-
- Dimitri Komatitsch, IPG Paris, France, December 1996: first 3-D solver for the CM-5 Connection Machine,
-    parallelized on 128 processors using Connection Machine Fortran.\\
-
-\addcontentsline{toc}{chapter}{References}
-\bibliography{bibliography}
-
-\appendix
-
-\chapter{\label{cha:Coordinates}Reference Frame Convention}
-
-The code uses the following convention for the Cartesian reference
-frame:
-
-\begin{itemize}
-\item the $x$ axis points East
-\item the $y$ axis points North
-\item the $z$ axis points up
-\end{itemize}
-Note that this convention is different from both the \citet{AkRi80}
-convention and the Harvard Centroid-Moment Tensor (CMT) convention.
-The Aki \& Richards convention is
-
-\begin{itemize}
-\item the $x$ axis points North
-\item the $y$ axis points East
-\item the $z$ axis points down
-\end{itemize}
-and the Harvard CMT convention is
-
-\begin{itemize}
-\item the $x$ axis points South
-\item the $y$ axis points East
-\item the $z$ axis points up
-\end{itemize}
-
-\subsection*{Source and receiver locations}
-
-The SPECFEM3D software code internally uses Cartesian coordinates.
-The given locations for sources and receiver locations thus may get
-converted.  Sources and receiver locations are read in from the
-\texttt{CMTSOLUTION} (or \texttt{FORCESOLUTION}) and \texttt{STATIONS}
-files.  Note that e.g. the \texttt{CMTSOLUTION} and
-\texttt{FORCESOLUTION} files denotes the location by
-"longitude/latitude/depth". We read in longitude as $x$ coordinate,
-latitude as $y$ coordinate.
-
-In case the flag \texttt{SUPPRESS\_UTM\_PROJECTION} is set to \texttt{.false.}
-in the main parameter file (see Chapter~\ref{cha:Creating-Distributed-Databases}),
-the $x$/$y$ coordinates have to be given in degrees and are converted to Cartesian coordinates
-using a UTM conversion with the specified UTM zone.
-
-The value for depth (given in $km$ in \texttt{CMTSOLUTION} and
-\texttt{FORCESOLUTION}) or burial depth (given in $m$ in
-\texttt{STATIONS}) is evaluated with respect to the surface of the
-mesh at the specified $x$/$y$ location to find the corresponding $z$
-coordinate.  It is possible to use this depth value directly as $z$
-coordinate by changing the flag \texttt{USE\_SOURCES\_RECVS\_Z} to
-\texttt{.true.} in the file \texttt{constants.h} located in the
-\texttt{src/shared/} subdirectory.
-
-\subsection*{Seismogram outputs}
-
-The seismogram output directions are given in Cartesian $x$/$y$/$z$ directions and not rotated any further.
-Changing flags in \texttt{constants.h} in the \texttt{src/shared/} subdirectory only rotates the seismogram outputs
-if receivers are forced to be located at the surface (\texttt{RECVS\_CAN\_BE\_BURIED\_EXT\_MESH} set to \texttt{.false.}) and the normal to the surface at the receiver location should be used (\texttt{EXT\_MESH\_RECV\_NORMAL} set to
-\texttt{.true.}) as vertical.
-In this case, the outputs are rotated to have the vertical component normal to the surface of the mesh, $x$ and $y$ directions are somewhat arbitrary orthogonal directions along the surface.
-
-For the labeling of the seismogram channels, see Appendix~\ref{cha:channel-codes}. Additionally, we add labels to the synthetics using the following convention:
-For a receiver station located in an
-\begin{description}
-\item[elastic domain:] ~\\
-\begin{itemize}
-\item \texttt{semd} for the displacement vector
-\item \texttt{semv} for the velocity vector
-\item \texttt{sema} for the acceleration vector
-\end{itemize}
-\item[acoustic domain:] ~\\
-(please note that receiver stations in acoustic domains must be buried. This is due to the free surface condition which enforces a zero pressure at the surface.)
-\begin{itemize}
-\item \texttt{semd} for the displacement vector
-\item \texttt{semv} for the velocity vector
-\item \texttt{sema} for pressure which will be stored in the vertical component \texttt{Z}.
-Note that pressure in the acoustic media is isotropic and thus the pressure record would be the same in the other two component directions.
-We therefore use the other two seismogram components to store the acoustic potential in component \texttt{X} (or \texttt{N}) and the first time derivative of the acoustic potential in component \texttt{Y} (or \texttt{E}).
-\end{itemize}
-\end{description}
-The seismograms are by default written out in ASCII-format to the \texttt{in\_out\_files/OUTPUT\_FILES/} subdirectory by each parallel process.
-You can change this behavior by changing the following flags in the \texttt{constants.h} file located in the \texttt{src/shared/} subdirectory:
-\begin{description}
-\item [{\texttt{SEISMOGRAMS\_BINARY}}] set to \texttt{.true.} to have seismograms written out in binary format.
-\item [{\texttt{WRITE\_SEISMOGRAMS\_BY\_MASTER}}] Set to \texttt{.true.} to have only the master process writing out seismograms.
-This can be useful on a cluster, where only the master process node has access to the output directory.
-\item [{\texttt{USE\_OUTPUT\_FILES\_PATH}}] Set to \texttt{.false.} to have the seismograms output to \texttt{LOCAL\_PATH} directory specified in the main parameter file \texttt{in\_data\_files/Par\_file}.
-\end{description}
-
-\chapter{\label{cha:channel-codes}Channel Codes of Seismograms}
-
-Seismic networks, such as the Global Seismographic Network (GSN), generally involve various types of instruments with different bandwidths, sampling properties and component configurations.
-There are standards to name channel codes depending on instrument properties.
-IRIS  \texttt{(www.iris.edu)} uses SEED format for channel codes, which are represented by three letters, such as \texttt{LHN}, \texttt{BHZ}, etc.
-In older versions of the SPECFEM3D package, a common format was used for the channel codes of all seismograms, which was \texttt{BHE/BHN/BHZ} for three components.
-To avoid confusion when comparison are made to observed data, we are now using the FDSN convention (\url{http://www.fdsn.org/}) for SEM seismograms.
-In the following, we give a brief explanation of the FDSN convention used by IRIS, and how it is adopted in SEM seismograms.
-Please visit \texttt{www.iris.edu/manuals/SEED\_appA.htm} for further information.\\
-
-\noindent \textbf{\texttt{Band code:}} The first letter in the channel code denotes the band code of seismograms, which depends on the response band and the sampling rate of instruments.
-The list of band codes used by IRIS is shown in Figure \ref{fig:IRIS_band_codes}.
-The sampling rate of SEM synthetics is controlled by the resolution of simulations rather than instrument properties.
-However, for consistency, we follow the FDSN convention for SEM seismograms governed by their sampling rate.
-For SEM synthetics, we consider band codes for which $dt \leq 1$ s.
-IRIS also considers the response band of instruments.
-For instance, short-period and broad-band seismograms with the same sampling rate correspond to different band codes, such as S and B, respectively.
-In such cases, we consider SEM seismograms as broad band, ignoring the corner period ($\geq 10$ s) of the response band of instruments (note that at these resolutions, the minimum period in the SEM synthetics will be less than $10$ s).
-Accordingly, when you run a simulation the band code will be chosen depending on the resolution of the synthetics, and channel codes of SEM seismograms will start with either \texttt{L}, \texttt{M}, \texttt{B}, \texttt{H}, \texttt{C} or \texttt{F}, shown by red color in the figure.\\
-
-\begin{figure}[ht]
-\noindent \begin{centering}
-\includegraphics[scale=0.6]{figures/IRIS_band_codes.pdf}
-\par\end{centering}
-
-\caption{The band code convention is based on the sampling rate and the response band of instruments.
-Please visit \texttt{www.iris.edu/manuals/SEED\_appA.htm} for further information.
-Grey rows show the relative band-code range in SPECFEM3D, and the band codes used to name SEM seismograms are denoted in red.}
-
-\label{fig:IRIS_band_codes}
-\end{figure}
-
-\noindent \textbf{\texttt{Instrument code:}} The second letter in the channel code corresponds to instrument codes, which specify the family to which the sensor belongs.
-For instance, \texttt{H} and \texttt{L} are used for high-gain and low-gain seismometers, respectively.
-The instrument code of SEM seismograms will always be \texttt{X}, as assigned by FDSN for synthetic seismograms.
-\\
-
-\noindent \textbf{\texttt{Orientation code:}} The third letter in channel codes is an orientation code, which generally describes the physical configuration of the components of instrument packages.
-SPECFEM3D uses the traditional orientation code  \texttt{E/N/Z} (East-West, North-South, Vertical) for three components when a UTM projection is used.
-If the UTM conversion is suppressed, i.e.
-the flag \texttt{SUPPRESS\_UTM\_PROJECTION} is set to \texttt{.true.}, then the three components are labelled \texttt{X/Y/Z} according to the Cartesian reference frame. \\
-
-\noindent \textbf{\texttt{EXAMPLE:}} The sampling rate is given by \texttt{DT} in the main parameter file \texttt{Par\_file} located in the \texttt{in\_data\_files/} subdirectory.
-Depending on the resolution of your simulations, if you choose a sampling rate greater than $0.01$ s and less than $1$ s, a seismogram recording displacements on the vertical component of a station \texttt{ASBS} (network \texttt{AZ}) will be named \texttt{ASBS.AZ.MXZ.semd.sac}, whereas it will be \texttt{ASBS.AZ.BXZ.semd.sac}, if the sampling rate is greater than 0.0125 and less equal to 0.1 s.
-
-\chapter{\label{cha:Troubleshooting}Troubleshooting}
-
-\section*{FAQ}
-
-\begin{description}
-\item [configuration fails:]
-
-   Examine the log file 'config.log'. It contains detailed informations.
-   In many cases, the path's to these specific compiler commands F90,
-   CC and MPIF90 won't be correct if `./configure` fails.
-
-   Please make sure that you have a working installation of a Fortran compiler,
-   a C compiler and an MPI implementation. You should be able to compile this
-   little program code:
-
-\begin{lyxcode}
-
-      program main ~\\
-        include 'mpif.h' ~\\
-        integer, parameter :: CUSTOM\_MPI\_TYPE = MPI\_REAL ~\\
-        integer ier ~\\
-        call MPI\_INIT(ier) ~\\
-        call MPI\_BARRIER(MPI\_COMM\_WORLD,ier) ~\\
-        call MPI\_FINALIZE(ier) ~\\
-      end
-
-\end{lyxcode}
-
-\item [compilation fails stating:] ~\\
-\begin{lyxcode}
-    ...  ~\\
-    obj/program\_generate\_databases.o: In function `MAIN\_\_':  ~\\
-    program\_generate\_databases.f90:(.text+0x14): undefined reference to `\_gfortran\_set\_std'  ~\\
-    ...  ~\\
-\end{lyxcode}
-
-  Make sure you're pointing to the right 'mpif90' wrapper command.
-
-  Normally, this message will appear when you're mixing two different Fortran
-  compilers. That is, using e.g. gfortran to compile non-MPI files
-  and mpif90, wrapper provided for e.g. ifort, to compile MPI-files.
-
-  fix: e.g. specify > ./configure FC=gfortran MPIF90=/usr/local/openmpi-gfortran/bin/mpif90
-
-\item [after executing \texttt{xmeshfem3D} I've got elements with skewness of 81\% percent, what does this mean:]
-Look at the skewness table printed in the \texttt{output\_mesher.txt} file after executing \texttt{xmeshfem3D} for the
-example given in \texttt{examples/meshfem3D\_examples/simple\_model/}:
-\begin{lyxcode}
-...~\\
-histogram of skewness (0. good - 1. bad): ~\\
- 0.0000000E+00  -   5.0000001E-02       27648     81.81818      \% ~\\
- 5.0000001E-02  -   0.1000000               0    0.0000000E+00  \% ~\\
-... ~\\
-\end{lyxcode}
-The first line means that you have 27,648 elements with a skewness value between 0 and 0.05 (which means the element is basically not skewed, just plain regular hexahedral element).
-The total number of elements you have in this mesh is (see in the \texttt{output\_mesher.txt} file a bit further down):
-\begin{lyxcode}
-...~\\
-total number of elements in entire mesh:        33792 ~\\
-... ~\\
-\end{lyxcode}
-which gives you that:  27,648 / 33,792 $\sim$ 81.8 \% of all elements are not skewed, i.e. regular elements. a fantastic value :)
-
-The histogram lists for this mesh also some stronger skewed elements, for example the worst ones belong to:
-\begin{lyxcode}
-...~\\
-0.6000000      -   0.6500000            2048     6.060606      \% ~\\
-... ~\\
-\end{lyxcode}
-
-about 6 \% of all elements have distortions with a skewness value between 0.6 and 0.65.
-The skewness values give you a hint of how good your mesh is.
-In an ideal world, you would want to have no distortions, just like the 81\% from above.
-Those elements give you the best approximate values by the GLL quadrature used in the spectral-element method.
-However, having weakly distorted elements is still fine and the solutions are still accurate enough.
-So empirically, values up to around 0.7 are tolerable, above that you should consider remeshing...
-
-To give you an idea why some of the elements are distorted, see the following figure \ref{fig:mesh.vp} of the mesh you obtain in the example \texttt{examples/meshfem3D\_examples/simple\_model/}.
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=.9\textwidth]{figures/mesh_vp.jpg}
-\par\end{centering}
-\caption{Paraview visualization using the mesh vtk-files for the example given in \texttt{examples/meshfem3D\_examples/simple\_model/}.}
-\label{fig:mesh.vp}
-\end{figure}
-You will see that the mesh contains a doubling layer, where we stitch elements together such that the size of two elements will transition to the size of one element (very useful to keep the ratio of wavespeed / element\_size about constant).
-Those elements in this doubling layer have higher skewness values and make up those 6 \% in the histogram.
-
-\item [the code gives following error message "need at least one receiver":]
-This means that no stations given in the input file \texttt{in\_data\_files/STATIONS} could be located within the dimensions of the mesh.
-This can happen for example when the mesh was created with the in-house mesher \texttt{xmeshfem3D} while using the Universal Transverse Mercator (UTM) projection but the simulation with \texttt{xspecfem3D} was suppressing this projection from latitude/longitude to x/y/z coordinates.
-
-In such cases, try to change your \texttt{in\_data\_files/Par\_file} and set e.g.:
-\begin{lyxcode}
-SUPPRESS\_UTM\_PROJECTION         = .false. ~\\
-\end{lyxcode}
-to be the same in \texttt{Mesh\_Par\_file} and \texttt{Par\_file}.
-This flag should be identical when using the in-house mesher \texttt{xmeshfem3D}, \texttt{xgenerate\_databases} and \texttt{xspecfem3D} together to run simulations.
-
-The flag determines if the coordinates you specify for your source and station locations are given as lat/lon degrees and
-must be converted to UTM coordinates.
-As an example, if you use \texttt{.false.} within \texttt{Mesh\_Par\_file} then you create a mesh with \texttt{xmeshfem3D} using the UTM projection from lat/lon as input format to UTM projected coordinates to store the mesh point positions, which is fine.
-The error then may occur if in the \texttt{Par\_file} you have this set to \texttt{.true.} so that the \texttt{xgenerate\_databases} and \texttt{xspecfem3D} suppress the UTM projection and assume that all coordinates you use now for source and receiver locations are given in meters (that is, converted) already.
-So it won't find the specified locations in the used mesh.
-As a solutions, just change the flag in \texttt{Par\_file} to be the same as in \texttt{Mesh\_Par\_file} and rerun \texttt{xgenerate\_databases} and \texttt{xspecfem3D} to make sure that your simulation works fine.
-
-\item [I get the following error message "forward simulation became unstable and blew up":]
-In most cases this means that your time step size \texttt{DT} is chosen too big.
-Look at your files \texttt{output\_mesher.txt} or \texttt{output\_solver.txt} created in the folder \texttt{in\_out\_files/OUTPUT\_FILES/}.
-In these output files, find the section:
-\begin{lyxcode}
-... ~\\
-********************************************* ~\\
-*** Verification of simulation parameters *** ~\\
-********************************************* ~\\
-... ~\\
-*** Minimum period resolved =    4.308774 ~\\
-*** Maximum suggested time step =   6.8863556E-02 ~\\
-... ~\\
-... ~\\
-\end{lyxcode}
-
-then change \texttt{DT} in the \texttt{in\_data\_files/Par\_file} to be somewhere close to the maximum suggested time step. In the example above: \\
-\texttt{DT                              = 0.05d0} \\
-would (most probably) work fine. It could be also bigger than the 0.068 s suggested. This depends a bit on the distortions of
-your mesh elements. The more regular they are, the bigger you can choose \texttt{DT}. Just play with this value a bit and see
-when the simulation becomes stable ...
-
-\end{description}
-
-\chapter{\label{cha:License}License }
-
-\textbf{GNU GENERAL PUBLIC LICENSE Version 2, June 1991. Copyright
-(C) 1989, 1991 Free Software Foundation, Inc. 59 Temple Place, Suite
-330, Boston, MA 02111-1307 USA} \\
-Everyone is permitted to copy and distribute verbatim copies of this
-license document, but changing it is not allowed.
-
-\section*{Preamble}
-
-The licenses for most software are designed to take away your freedom
-to share and change it. By contrast, the GNU General Public License
-is intended to guarantee your freedom to share and change free software
--- to make sure the software is free for all its users. This General
-Public License applies to most of the Free Software Foundation's software
-and to any other program whose authors commit to using it. (Some other
-Free Software Foundation software is covered by the GNU Library General
-Public License instead.) You can apply it to your programs, too.
-
-When we speak of free software, we are referring to freedom, not price.
-Our General Public Licenses are designed to make sure that you have
-the freedom to distribute copies of free software (and charge for
-this service if you wish), that you receive source code or can get
-it if you want it, that you can change the software or use pieces
-of it in new free programs; and that you know you can do these things.
-
-To protect your rights, we need to make restrictions that forbid anyone
-to deny you these rights or to ask you to surrender the rights. These
-restrictions translate to certain responsibilities for you if you
-distribute copies of the software, or if you modify it.
-
-For example, if you distribute copies of such a program, whether gratis
-or for a fee, you must give the recipients all the rights that you
-have. You must make sure that they, too, receive or can get the source
-code. And you must show them these terms so they know their rights.
-
-We protect your rights with two steps:
-
-\begin{enumerate}
-\item Copyright the software, and
-\item Offer you this license which gives you legal permission to copy, distribute
-and/or modify the software.
-\end{enumerate}
-Also, for each author's protection and ours, we want to make certain
-that everyone understands that there is no warranty for this free
-software. If the software is modified by someone else and passed on,
-we want its recipients to know that what they have is not the original,
-so that any problems introduced by others will not reflect on the
-original authors' reputations.
-
-Finally, any free program is threatened constantly by software patents.
-We wish to avoid the danger that redistributors of a free program
-will individually obtain patent licenses, in effect making the program
-proprietary. To prevent this, we have made it clear that any patent
-must be licensed for everyone's free use or not licensed at all.
-
-The precise terms and conditions for copying, distribution and modification
-follow.
-
-\section*{GNU GENERAL PUBLIC LICENSE TERMS AND CONDITIONS FOR COPYING, DISTRIBUTION
-AND MODIFICATION }
-
-\begin{itemize}
-
-\item[0.]This License applies to any program or other work which
-contains a notice placed by the copyright holder saying it may be
-distributed under the terms of this General Public License. The ``Program''
-below refers to any such program or work, and a ``work based on the
-Program'' means either the Program or any derivative work under copyright
-law: that is to say, a work containing the Program or a portion of
-it, either verbatim or with modifications and/or translated into another
-language. (Hereinafter, translation is included without limitation
-in the term ``modification.'') Each licensee is addressed as ``you.''\\
-\\
-Activities other than copying, distribution and modification are not
-covered by this License; they are outside its scope. The act of running
-the Program is not restricted, and the output from the Program is
-covered only if its contents constitute a work based on the Program
-(independent of having been made by running the Program). Whether
-that is true depends on what the Program does.
-
-\end{itemize}
-
-\begin{enumerate}
-\item You may copy and distribute verbatim copies of the Program's source
-code as you receive it, in any medium, provided that you conspicuously
-and appropriately publish on each copy an appropriate copyright notice
-and disclaimer of warranty; keep intact all the notices that refer
-to this License and to the absence of any warranty; and give any other
-recipients of the Program a copy of this License along with the Program.
-
-You may charge a fee for the physical act of transferring a copy,
-and you may at your option offer warranty protection in exchange for
-a fee.
-
-\item You may modify your copy or copies of the Program or any portion of
-it, thus forming a work based on the Program, and copy and distribute
-such modifications or work under the terms of Section 1 above, provided
-that you also meet all of these conditions:
-
-\begin{enumerate}
-\item You must cause the modified files to carry prominent notices stating
-that you changed the files and the date of any change.
-\item You must cause any work that you distribute or publish, that in whole
-or in part contains or is derived from the Program or any part thereof,
-to be licensed as a whole at no charge to all third parties under
-the terms of this License.
-\item If the modified program normally reads commands interactively when
-run, you must cause it, when started running for such interactive
-use in the most ordinary way, to print or display an announcement
-including an appropriate copyright notice and a notice that there
-is no warranty (or else, saying that you provide a warranty) and that
-users may redistribute the program under these conditions, and telling
-the user how to view a copy of this License. (Exception: if the Program
-itself is interactive but does not normally print such an announcement,
-your work based on the Program is not required to print an announcement.)
-\end{enumerate}
-These requirements apply to the modified work as a whole. If identifiable
-sections of that work are not derived from the Program, and can be
-reasonably considered independent and separate works in themselves,
-then this License, and its terms, do not apply to those sections when
-you distribute them as separate works. But when you distribute the
-same sections as part of a whole which is a work based on the Program,
-the distribution of the whole must be on the terms of this License,
-whose permissions for other licensees extend to the entire whole,
-and thus to each and every part regardless of who wrote it.
-
-Thus, it is not the intent of this section to claim rights or contest
-your rights to work written entirely by you; rather, the intent is
-to exercise the right to control the distribution of derivative or
-collective works based on the Program.
-
-In addition, mere aggregation of another work not based on the Program
-with the Program (or with a work based on the Program) on a volume
-of a storage or distribution medium does not bring the other work
-under the scope of this License.
-
-\item You may copy and distribute the Program (or a work based on it, under
-Section 2) in object code or executable form under the terms of Sections
-1 and 2 above provided that you also do one of the following:
-
-\begin{enumerate}
-\item Accompany it with the complete corresponding machine-readable source
-code, which must be distributed under the terms of Sections 1 and
-2 above on a medium customarily used for software interchange; or,
-\item Accompany it with a written offer, valid for at least three years,
-to give any third party, for a charge no more than your cost of physically
-performing source distribution, a complete machine-readable copy of
-the corresponding source code, to be distributed under the terms of
-Sections 1 and 2 above on a medium customarily used for software interchange;
-or,
-\item Accompany it with the information you received as to the offer to
-distribute corresponding source code. (This alternative is allowed
-only for noncommercial distribution and only if you received the program
-in object code or executable form with such an offer, in accord with
-Subsection b above.)
-\end{enumerate}
-The source code for a work means the preferred form of the work for
-making modifications to it. For an executable work, complete source
-code means all the source code for all modules it contains, plus any
-associated interface definition files, plus the scripts used to control
-compilation and installation of the executable. However, as a special
-exception, the source code distributed need not include anything that
-is normally distributed (in either source or binary form) with the
-major components (compiler, kernel, and so on) of the operating system
-on which the executable runs, unless that component itself accompanies
-the executable.
-
-If distribution of executable or object code is made by offering access
-to copy from a designated place, then offering equivalent access to
-copy the source code from the same place counts as distribution of
-the source code, even though third parties are not compelled to copy
-the source along with the object code.
-
-\item You may not copy, modify, sublicense, or distribute the Program except
-as expressly provided under this License. Any attempt otherwise to
-copy, modify, sublicense or distribute the Program is void, and will
-automatically terminate your rights under this License. However, parties
-who have received copies, or rights, from you under this License will
-not have their licenses terminated so long as such parties remain
-in full compliance.
-\item You are not required to accept this License, since you have not signed
-it. However, nothing else grants you permission to modify or distribute
-the Program or its derivative works. These actions are prohibited
-by law if you do not accept this License. Therefore, by modifying
-or distributing the Program (or any work based on the Program), you
-indicate your acceptance of this License to do so, and all its terms
-and conditions for copying, distributing or modifying the Program
-or works based on it.
-\item Each time you redistribute the Program (or any work based on the Program),
-the recipient automatically receives a license from the original licensor
-to copy, distribute or modify the Program subject to these terms and
-conditions. You may not impose any further restrictions on the recipients'
-exercise of the rights granted herein. You are not responsible for
-enforcing compliance by third parties to this License.
-\item If, as a consequence of a court judgment or allegation of patent infringement
-or for any other reason (not limited to patent issues), conditions
-are imposed on you (whether by court order, agreement or otherwise)
-that contradict the conditions of this License, they do not excuse
-you from the conditions of this License. If you cannot distribute
-so as to satisfy simultaneously your obligations under this License
-and any other pertinent obligations, then as a consequence you may
-not distribute the Program at all. For example, if a patent license
-would not permit royalty-free redistribution of the Program by all
-those who receive copies directly or indirectly through you, then
-the only way you could satisfy both it and this License would be to
-refrain entirely from distribution of the Program.
-
-If any portion of this section is held invalid or unenforceable under
-any particular circumstance, the balance of the section is intended
-to apply and the section as a whole is intended to apply in other
-circumstances.
-
-It is not the purpose of this section to induce you to infringe any
-patents or other property right claims or to contest validity of any
-such claims; this section has the sole purpose of protecting the integrity
-of the free software distribution system, which is implemented by
-public license practices. Many people have made generous contributions
-to the wide range of software distributed through that system in reliance
-on consistent application of that system; it is up to the author/donor
-to decide if he or she is willing to distribute software through any
-other system and a licensee cannot impose that choice.
-
-This section is intended to make thoroughly clear what is believed
-to be a consequence of the rest of this License.
-
-\item If the distribution and/or use of the Program is restricted in certain
-countries either by patents or by copyrighted interfaces, the original
-copyright holder who places the Program under this License may add
-an explicit geographical distribution limitation excluding those countries,
-so that distribution is permitted only in or among countries not thus
-excluded. In such case, this License incorporates the limitation as
-if written in the body of this License.
-\item The Free Software Foundation may publish revised and/or new versions
-of the General Public License from time to time. Such new versions
-will be similar in spirit to the present version, but may differ in
-detail to address new problems or concerns.
-
-Each version is given a distinguishing version number. If the Program
-specifies a version number of this License which applies to it and
-``any later version,'' you have the option of following the terms
-and conditions either of that version or of any later version published
-by the Free Software Foundation. If the Program does not specify a
-version number of this License, you may choose any version ever published
-by the Free Software Foundation.
-
-\item If you wish to incorporate parts of the Program into other free programs
-whose distribution conditions are different, write to the author to
-ask for permission. For software which is copyrighted by the Free
-Software Foundation, write to the Free Software Foundation; we sometimes
-make exceptions for this. Our decision will be guided by the two goals
-of preserving the free status of all derivatives of our free software
-and of promoting the sharing and reuse of software generally.
-\end{enumerate}
-
-\subsection*{NO WARRANTY }
-
-\begin{itemize}
-
-\item[11.]BECAUSE THE PROGRAM IS LICENSED FREE OF CHARGE, THERE IS
-NO WARRANTY FOR THE PROGRAM, TO THE EXTENT PERMITTED BY APPLICABLE
-LAW. EXCEPT WHEN OTHERWISE STATED IN WRITING THE COPYRIGHT HOLDERS
-AND/OR OTHER PARTIES PROVIDE THE PROGRAM ``AS IS'' WITHOUT WARRANTY
-OF ANY KIND, EITHER EXPRESSED OR IMPLIED, INCLUDING, BUT NOT LIMITED
-TO, THE IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR
-PURPOSE. THE ENTIRE RISK AS TO THE QUALITY AND PERFORMANCE OF THE
-PROGRAM IS WITH YOU. SHOULD THE PROGRAM PROVE DEFECTIVE, YOU ASSUME
-THE COST OF ALL NECESSARY SERVICING, REPAIR OR CORRECTION.
-
-\item[12.]IN NO EVENT UNLESS REQUIRED BY APPLICABLE LAW OR AGREED
-TO IN WRITING WILL ANY COPYRIGHT HOLDER, OR ANY OTHER PARTY WHO MAY
-MODIFY AND/OR REDISTRIBUTE THE PROGRAM AS PERMITTED ABOVE, BE LIABLE
-TO YOU FOR DAMAGES, INCLUDING ANY GENERAL, SPECIAL, INCIDENTAL OR
-CONSEQUENTIAL DAMAGES ARISING OUT OF THE USE OR INABILITY TO USE THE
-PROGRAM (INCLUDING BUT NOT LIMITED TO LOSS OF DATA OR DATA BEING RENDERED
-INACCURATE OR LOSSES SUSTAINED BY YOU OR THIRD PARTIES OR A FAILURE
-OF THE PROGRAM TO OPERATE WITH ANY OTHER PROGRAMS), EVEN IF SUCH HOLDER
-OR OTHER PARTY HAS BEEN ADVISED OF THE POSSIBILITY OF SUCH DAMAGES.
-
-\end{itemize}
-
-\section*{END OF TERMS AND CONDITIONS }
-
-\subsection*{How to Apply These Terms to Your New Programs}
-
-If you develop a new program, and you want it to be of the greatest
-possible use to the public, the best way to achieve this is to make
-it free software which everyone can redistribute and change under
-these terms.
-
-To do so, attach the following notices to the program. It is safest
-to attach them to the start of each source file to most effectively
-convey the exclusion of warranty; and each file should have at least
-the ``copyright'' line and a pointer to where the full notice is found.
-For example:
-
-\begin{quote}
-One line to give the program's name and a brief idea of what it does.
-Copyright {\footnotesize $\copyright$ (}year) (name of author)
-
-This program is free software; you can redistribute it and/or modify
-it under the terms of the GNU General Public License as published
-by the Free Software Foundation; either version 2 of the License,
-or (at your option) any later version.
-
-This program is distributed in the hope that it will be useful, but
-WITHOUT ANY WARRANTY; without even the implied warranty of MERCHANTABILITY
-or FITNESS FOR A PARTICULAR PURPOSE. See the GNU General Public License
-for more details.
-
-You should have received a copy of the GNU General Public License
-along with this program; if not, write to the Free Software Foundation,
-Inc., 59 Temple Place, Suite 330, Boston, MA 02111-1307 USA
-\end{quote}
-Also add information on how to contact you by electronic and paper
-mail.
-
-If the program is interactive, make it output a short notice like
-this when it starts in an interactive mode:
-
-\begin{quote}
-Gnomovision version 69, Copyright $\copyright$ year name of author Gnomovision
-comes with ABSOLUTELY NO WARRANTY; for details type `show w'. This
-is free software, and you are welcome to redistribute it under certain
-conditions; type `show c' for details.
-\end{quote}
-The hypothetical commands `show w' and `show c' should show the appropriate
-parts of the General Public License. Of course, the commands you use
-may be called something other than `show w' and `show c'; they could
-even be mouse-clicks or menu items -- whatever suits your program.
-
-You should also get your employer (if you work as a programmer) or
-your school, if any, to sign a ``copyright disclaimer'' for the program,
-if necessary. Here is a sample; alter the names:
-
-\begin{quote}
-Yoyodyne, Inc., hereby disclaims all copyright interest in the program
-`Gnomovision' (which makes passes at compilers) written by James Hacker.
-
-(signature of Ty Coon)\\
-1 April 1989 \\
-Ty Coon, President of Vice
-\end{quote}
-This General Public License does not permit incorporating your program
-into proprietary programs. If your program is a subroutine library,
-you may consider it more useful to permit linking proprietary applications
-with the library. If this is what you want to do, use the GNU Library
-General Public License instead of this License.
-\end{document}
+to the CIG Computational Seismology Mailing List \urlwithparentheses{cig-seismo@geodynamics.org}.

### 2015-04-24 doc/USER_MANUAL/01_introduction.tex

\chapter{Introduction}

The software package SPECFEM3D Cartesian simulates seismic wave propagation
at the local or regional scale and performs full waveform imaging (FWI) or adjoint tomography based upon the spectral-element method
(SEM). The SEM is a continuous Galerkin technique \citep{TrKoLi08,PeKoLuMaLeCaLeMaLiBlNiBaTr11},
which can easily be made discontinuous \citep{BeMaPa94,Ch00,KoWoHu02,ChCaVi03,LaWaBe05,Kop06,WiStBuGh10,AcKo11};
it is then close to a particular case of the discontinuous Galerkin
technique \citep{ReHi73,LeRa74,Arn82,JoPi86,BoMaHe91,FaRi99,HuHuRa99,CoKaSh00,GiHeWa02,RiWh03,MoRi05,GrScSc06,AiMoMu06,BeLaPi06,DuKa06,DeSeWh08,PuAmKa09,WiStBuGh10,DeSe10,EtChViGl10},
with optimized efficiency because of its tensorized basis functions
\citep{WiStBuGh10,AcKo11}. In particular, it can accurately handle
very distorted mesh elements \citep{OlSe11}.\\

It has very good accuracy and convergence properties \citep{MaPa89,SePr94,DeFiMu02,Coh02,DeSe07,SeOl08,AiWa09,AiWa10,MeStTh12}.
The spectral element approach admits spectral rates of convergence
and allows exploiting $hp$-convergence schemes. It is also very well
suited to parallel implementation on very large supercomputers \citep{KoTsChTr03,TsKoChTr03,KoLaMi08a,CaKoLaTiMiLeSnTr08,KoViCh10}
as well as on clusters of GPU accelerating graphics cards \citep{Kom11,MiKo10,KoMiEr09,KoErGoMi10}.
Tensor products inside each element can be optimized to reach very
high efficiency \citep{DeFiMu02}, and mesh point and element numbering
can be optimized to reduce processor cache misses and improve cache
reuse \citep{KoLaMi08a}. The SEM can also handle triangular (in 2D)
or tetrahedral (in 3D) elements \citep{WinBoyd96,TaWi00,KoMaTrTaWi01,Coh02,MeViSa06}
as well as mixed meshes, although with increased cost and reduced
accuracy in these elements, as in the discontinuous Galerkin method.\\

Note that in many geological models in the context of seismic wave
propagation studies (except for instance for fault dynamic rupture
studies, in which very high frequencies or supershear rupture need
to be modeled near the fault, see e.g. \citet{BeGlCrViPi07,BeGlCrVi09,PuAmKa09,TaCrEtViBeSa10})
a continuous formulation is sufficient because material property contrasts
are not drastic and thus conforming mesh doubling bricks can efficiently
handle mesh size variations \citep{KoTr02a,KoLiTrSuStSh04,LeChLiKoHuTr08,LeChKoHuTr09,LeKoHuTr09}.\\

For a detailed introduction to the SEM as applied to regional seismic
wave propagation, please consult \citet{PeKoLuMaLeCaLeMaLiBlNiBaTr11,TrKoLi08,KoVi98,KoTr99,ChKoViCaVaFe07}
and in particular \citet{LeKoHuTr09,LeChKoHuTr09,LeChLiKoHuTr08,GoAmTaCaSmSaMaKo09,WiKoScTr04,KoLiTrSuStSh04}.
A detailed theoretical analysis of the dispersion
and stability properties of the SEM is available in \citet{Coh02}, \citet{DeSe07}, \citet{SeOl07}, \citet{SeOl08} and \citet{MeStTh12}.\\

Effects due to lateral variations in compressional-wave speed, shear-wave
speed, density, a 3D crustal model, topography and bathymetry are
included. The package can accommodate full 21-parameter anisotropy
(see~\citet{ChTr07}) as well as lateral variations in attenuation
\citep{SaKoTr10}. Adjoint capabilities and finite-frequency kernel
simulations are included \citep{TrKoLi08,PeKoLuMaLeCaLeMaLiBlNiBaTr11,LiTr06,FiIgBuKe09,ViOp09}.\\

The SEM was originally developed in computational fluid dynamics \citep{Pat84,MaPa89}
and has been successfully adapted to address problems in seismic wave
propagation. Early seismic wave propagation applications of the SEM,
utilizing Legendre basis functions and a perfectly diagonal mass matrix,
include \citet{CoJoTo93}, \citet{Kom97}, \citet{FaMaPaQu97}, \citet{CaGa97},
\citet{KoVi98} and \citet{KoTr99}, whereas applications involving
Chebyshev basis functions and a nondiagonal mass matrix include \citet{SePr94},
\citet{PrCaSe94} and \citet{SePrPr95}.\\

All SPECFEM3D software is written in Fortran2003 with full portability
in mind, and conforms strictly to the Fortran2003 standard. It uses
no obsolete or obsolescent features of Fortran. The package uses parallel
programming based upon the Message Passing Interface (MPI) \citep{GrLuSk94,Pac97}.\\

SPECFEM3D won the Gordon Bell award for best performance at the SuperComputing~2003
conference in Phoenix, Arizona (USA) (see \citet{KoTsChTr03} and
\url{www.sc-conference.org/sc2003/nrfinalaward.html}). It was a finalist
again in 2008 for a run at 0.16 petaflops (sustained) on 149,784 processors
of the `Jaguar' Cray XT5 system at Oak Ridge National Laboratories
(USA) \citep{CaKoLaTiMiLeSnTr08}. It also won the BULL Joseph Fourier
supercomputing award in 2010.\\

It reached the sustained one petaflop performance level for the first time in February 2013
on the Blue Waters Cray supercomputer at the National Center for Supercomputing Applications (NCSA), located at the University of Illinois at Urbana-Champaign (USA).\\

This new release of the code includes Convolution or
Auxiliary Differential Equation Perfectly Matched absorbing Layers
(C-PML or ADE-PML) \citep{MaKoEz08,MaKoGe08,MaKo09,MaKoGeBr10,KoMa07}.
It also includes support for GPU graphics card acceleration \citep{Kom11,MiKo10,KoMiEr09,KoErGoMi10}.\\

The next release of the code will use the PT-SCOTCH parallel and threaded
version of SCOTCH for mesh partitioning instead of the serial version.\\

SPECFEM3D Cartesian includes coupled fluid-solid domains and adjoint
capabilities, which enables one to address seismological inverse problems,
but for linear rheologies only so far. To accommodate visco-plastic
or non-linear rheologies, the reader can refer to the \href{http://geoelse.stru.polimi.it/}{GeoELSE}software
package \citep{CaGa97,StPaIg09}.

\section{Citation}

If you use SPECFEM3D Cartesian for your own research, please cite
at least one of the following articles:
\begin{description}
\item [{Numerical simulations in general}] ~\\
 Forward and adjoint simulations are described in detail in \citet{TrKoLi08,PeKoLuMaLeCaLeMaLiBlNiBaTr11,VaCaSaKoVi99,KoMiEr09,KoErGoMi10,ChKoViCaVaFe07,MaKoDi09,KoViCh10,CaKoLaTiMiLeSnTr08,TrKoHjLiZhPeBoMcFrTrHu10,KoRiTr02,KoTr02a,KoTr02b,KoTr99}
or \citet{KoVi98}. Additional aspects of adjoint simulations are
described in \citet{TrTaLi05,LiTr06,TrKoLi08,LiTr08,TrKoHjLiZhPeBoMcFrTrHu10,PeKoLuMaLeCaLeMaLiBlNiBaTr11}.
Domain decomposition is explained in detail in \citet{MaKoBlLe08},
and excellent scaling up to 150,000 processor cores is shown for instance
in \citet{CaKoLaTiMiLeSnTr08,KoLaMi08a,MaKoBlLe08,KoErGoMi10,Kom11},
\item [{GPU computing}] ~\\
 Computing on GPU graphics cards for acoustic or seismic wave propagation
applications is described in detail in \citet{Kom11,MiKo10,KoMiEr09,KoErGoMi10}.
\end{description}
\noindent If you use this new version, which has non blocking
MPI for much better performance for medium or large runs, please cite
at least one of these six articles, in which results of non blocking
MPI runs are presented: \citet{PeKoLuMaLeCaLeMaLiBlNiBaTr11,KoErGoMi10,KoViCh10,Kom11,CaKoLaTiMiLeSnTr08,MaKoBlLe08}.\\

\noindent If you work on geophysical applications, you may be interested
in citing some of these application articles as well, among others:
\begin{description}
\item [{Southern California simulations}] ~\\
 \citet{KoLiTrSuStSh04,KrJiKoTr06a,KrJiKoTr06b}.

If you use the 3D southern California model, please cite \citet{SuSh03}
(Los Angeles model), \citet{lovelyetal06} (Salton Trough), and \citet{hauksson2000}
(southern California). The Moho map was determined by \citet{zhukanamori2000}.
The 1D SoCal model was developed by \citet{DrHe90}.

\item [{Anisotropy}] ~\\
 \citet{ChTr07,JiTsKoTr05,ChFaKo04,FaChKo04,RiRiKoTrHe02,TrKo00}.
\item [{Attenuation}] ~\\
 \citet{SaKoTr10,KoTr02a,KoTr99}.
\item [{Topography}] ~\\
 \citet{LeKoHuTr09,LeChKoHuTr09,LeChLiKoHuTr08,GoAmTaCaSmSaMaKo09,WiKoScTr04}.
\end{description}
The corresponding Bib\TeX{} entries may be found in file \texttt{doc/USER\_MANUAL/bibliography.bib}.

\section{Support}

This material is based upon work supported by the USA National Science
Foundation under Grants No. EAR-0406751 and EAR-0711177, by the French
CNRS, French INRIA Sud-Ouest MAGIQUE-3D, French ANR NUMASIS under
Grant No. ANR-05-CIGC-002, and European FP6 Marie Curie International
Reintegration Grant No. MIRG-CT-2005-017461. Any opinions, findings,
and conclusions or recommendations expressed in this material are
those of the authors and do not necessarily reflect the views of the
USA National Science Foundation, CNRS, INRIA, ANR or the European
Marie Curie program.

### 2015-04-24 doc/USER_MANUAL/03_mesh_generation.tex

\chapter{Mesh Generation}\label{cha:Mesh-Generation}

The first step in running a spectral-element simulation consists of
constructing a high-quality mesh for the region under consideration.
We provide two possibilities to do so: (1) relying on the external,
hexahedral mesher CUBIT, or (2) using the provided, internal mesher
\texttt{xmeshfem3D}. In the following, we explain these two approaches.

\section{Meshing with \texttt{CUBIT}}\label{cha:Running-the-Mesher-CUBIT}

CUBIT is a meshing tool suite for the creation of finite-element meshes
for arbitrarily shaped models. It has been developed and maintained
at Sandia National Laboratories and can be purchased for a small academic
institutional fee at \url{http://cubit.sandia.gov}. Our experience
showed that using CUBIT greatly facilitates and speeds up the generation
and preparation of hexahedral, conforming meshes for a variety of
geophysical models with increasing complexity.

\begin{figure}[htbp]
\noindent \begin{centering}
\includegraphics[width=0.6\textwidth]{figures/mount-cubit.jpg}
\par\end{centering}

\caption{Example of the graphical user interface of CUBIT. The hexahedral mesh
shown in the main display consists of a hexahedral discretization
of a single volume with topography.}

\label{fig:mount.cubit}
\end{figure}

\noindent
The basic steps in creating a load-balanced, partitioned mesh with
CUBIT are:
\begin{description}
\item [{1.}] setting up a hexahedral mesh with CUBIT,
\item [{2.}] exporting the CUBIT mesh into a SPECFEM3D Cartesian file format
and
\item [{3.}] partitioning the SPECFEM3D Cartesian mesh files for a chosen
number of cores.
\end{description}
Examples are provided in the SPECFEM3D Cartesian package in the subdirectory
\texttt{EXAMPLES/}. We strongly encourage you to contribute your own
example to this package by contacting the CIG Computational Seismology
Mailing List \urlwithparentheses{cig-seismo@geodynamics.org}.

\subsection{Creating the Mesh with CUBIT}

For the installation and handling of the CUBIT meshing tool suite,
please refer to the CUBIT user manual and documentation. In order
to give you a basic understanding of how to use CUBIT for our purposes,
examples are provided in the SPECFEM3D Cartesian package in the subdirectory
\texttt{EXAMPLES/}:
\begin{description}
\item [{\texttt{homogeneous\_halfspace}}] Creates a single block model
and assigns elastic material parameters.
\item [{\texttt{layered\_halfspace}}] Combines two different, elastic material
volumes and creates a refinement layer between the two. This example
can be compared for validation against the solutions provided in subdirectory
~\\
 \texttt{VALIDATION\_3D\_SEM\_SIMPLER\_LAYER\_SOURCE\_DEPTH/}.
\item [{\texttt{waterlayered\_halfspace}}] Combines an acoustic and elastic
material volume as in a schematic marine survey example.
\item [{\texttt{tomographic\_model}}] Creates a single block model whose
material properties will have to be read in from a tomographic model
file during the databases creation by \texttt{xgenerate\_databases}.
\end{description}
\begin{figure}[htbp]
\noindent \begin{centering}
\includegraphics[width=0.45\textwidth]{figures/example-homogeneous.jpg}
\includegraphics[width=0.45\textwidth]{figures/example-2layers.jpg} \\
\includegraphics[width=0.45\textwidth]{figures/example-water.jpg} 
\includegraphics[width=0.45\textwidth]{figures/example-tomo.jpg}
\par\end{centering}

\caption{Screenshots of the CUBIT examples provided in subdirectory \texttt{EXAMPLES/}:
homogeneous halfspace (top-left), layered halfspace (top-right), water
layered halfspace (bottom-left) and tomographic model (bottom-right).}

\label{fig:examples.cubit}
\end{figure}

In each example subdirectory you will find a \texttt{README} file,
which explains in a step-by-step tutorial the workflow for the example.
Please feel free to contribute your own example to this package by
contacting the CIG Computational Seismology Mailing List \urlwithparentheses{cig-seismo@geodynamics.org}.

In some cases, to re-create the meshes for the examples given, just type "claro ./create\_mesh.py" or similar
from the command line ("claro" is the command to run CUBIT from the command line).

IMPORTANT note: in order to correctly set up GEOCUBIT and run the examples, please read the file called "EXAMPLES/README";
in particular, please make sure you correctly set up the Python paths as indicated in that file.

\subsection{Exporting the Mesh with \texttt{run\_cubit2specfem3d.py} }\label{subsec:Exporting-the-Mesh}

Once the geometric model volumes in CUBIT are meshed, you prepare
the model for exportation with the definition of material blocks and
boundary surfaces. Thus, prior to exporting the mesh, you need to
define blocks specifying the materials and absorbing boundaries in
CUBIT. This process could be done automatically using the script \texttt{run\_cubit2specfem3d.py}
if the mesh meets some conditions or manually, following the block
convention:
\begin{description}
\item [{material\_name}] Each material should have a specific block defined
by a unique name. The name convention of the material is to start
with either 'elastic' or 'acoustic'. It must be then followed by a
unique identifier, e.g. 'elastic 1', 'elastic 2', etc. The additional
attributes to the block define the material description.

For an elastic material:
\begin{description}
\item [{material\_id}] An integer value which is unique for this material.
\item [{Vp}] P-wave speed of the material (given in m/s).
\item [{Vs}] S-wave speed of the material (given in m/s).
\item [{rho}] density of the material (given in kg/m$^{3}$).
\item [{Q}] quality factor to use in case of a simulation with attenuation
turned on. It should be between 1 and 9000. In case no attenuation
information is available, it can be set to zero. Please note that
your Vp- and Vs-speeds are given for a reference frequency. To change
this reference frequency, you change the value of \texttt{ATTENUATION\_f0\_REFERENCE}
in the main constants file \texttt{constants.h} found in subdirectory
\texttt{src/shared/}.
\item [{anisotropic\_flag}] Flag describing the anisotropic model to use
in case an anisotropic simulation should be conducted. See the file
\texttt{model\_aniso.f90} in subdirectory \texttt{src/generate\_databases/}
for an implementation of the anisotropic models. In case no anisotropy
is available, it can be set to zero.
\end{description}

Note that this material block has to be defined using all the volumes
which belong to this elastic material. For volumes belonging to another,
different material, you will need to define a new material block.

For an acoustic material:
\begin{description}
\item [{material\_id}] An integer value which is unique for this material.
\item [{Vp}] P-wave speed of the material (given in m/s).
\item [{0}] S-wave speed of the material is ignored.
\item [{rho}] density of the material (given in kg/m$^{3}$).
\end{description}
\item [{face\_topo}] Block definition for the surface which defines the
free surface (which can have topography). The name of this block must
be 'face\_topo', the block has to be defined using all the surfaces
which constitute the complete free surface of the model.
\item [{face\_abs\_xmin}] Block definition for the faces on the absorbing
boundaries, one block for each surface with x=Xmin.
\item [{face\_abs\_xmax}] Block definition for the faces on the absorbing
boundaries, one block for each surface with x=Xmax.
\item [{face\_abs\_ymin}] Block definition for the faces on the absorbing
boundaries, one block for each surface with y=Ymin.
\item [{face\_abs\_ymax}] Block definition for the faces on the absorbing
boundaries, one block for each surface with y=Ymax.
\item [{face\_abs\_bottom}] Block definition for the faces on the absorbing
boundaries, one block for each surface with z=bottom.
\end{description}
\begin{figure}[htbp]
\noindent \begin{centering}
\includegraphics[width=0.45\textwidth]{figures/mount-surface.jpg} 
\includegraphics[width=0.45\textwidth]{figures/mount-abs.jpg}
\par\end{centering}

\caption{Example of the block definitions for the free surface 'face\_topo'
(left) and the absorbing boundaries, defined in a single block 'face\_abs\_xmin'
(right) in CUBIT.}

\label{fig:mount.abs}
\end{figure}

Optionally, instead of specifying for each surface at the boundaries
a single block like mentioned above, you can also specify a single
block for all boundary surfaces and name it as one of the absorbing
blocks above, e.g. 'face\_abs\_xmin'.

After the block definitions are done, you export the mesh using the
script \texttt{cubit2specfem3d.py} provided in each of the example
directories (linked to the common script \texttt{CUBIT\_GEOCUBIT/cubit2specfem3d.py}).
If the export was successful, you should find the following
files in a subdirectory \texttt{MESH/}:
\begin{description}
\item [{absorbing\_cpml\_file}] \textbf{(only needed in case of C-PML absorbing
conditions)} Contains on the first line the total number of C-PML
spectral elements in the mesh, and then on the following line the
list of all these C-PML elements with two numbers per line: first
the spectral element number, and then a C-PML flag indicating to which
C-PML layer(s) that element belongs, according to the following convention:

\begin{itemize}
\item Flag = 1 : element belongs to a X CPML layer only (either in Xmin or in Xmax),
\item Flag = 2 : element belongs to a Y CPML layer only (either in Ymin or in Ymax),
\item Flag = 3 : element belongs to a Z CPML layer only (either in Zmin or in Zmax),
\item Flag = 4 : element belongs to a X CPML layer and also to a Y CPML layer,
\item Flag = 5 : element belongs to a X CPML layer and also to a Z CPML layer,
\item Flag = 6 : element belongs to a Y CPML layer and also to a Z CPML layer,
\item Flag = 7 : element belongs to a X, to a Y and to a Z CPML layer, i.e., it belongs to a CPML corner.
\end{itemize}

Note that it does not matter whether an element belongs to a Xmin
or to a Xmax CPML, the flag is the same in both cases; the same is
true for Ymin and Ymax, and also Zmin and Zmax.

When you have an existing CUBIT (or similar) mesh stored in SPECFEM3D
format, i.e., if you have existing \textquotedbl{}nodes\_coords\_file\textquotedbl{}
and \textquotedbl{}mesh\_file\textquotedbl{} files but do not know
how to assign CPML flags to them, we have created a small serial Fortran
program that will do that automatically for you, i.e., which will
create the \textquotedbl{}absorbing\_cpml\_file\textquotedbl{} for
you. That program is utils/CPML/convert\_external\_layers\_of\_a\_given\_mesh\_to\_CPML\_layers.f90,
and a small Makefile is provided in that directory (utils/CPML).

IMPORTANT: it is your responsibility to make sure that in the input
CUBIT (or similar) mesh that this code will read in SPECFEM3D format
from files \textquotedbl{}nodes\_coords\_file\textquotedbl{} and \textquotedbl{}mesh\_file\textquotedbl{}
you have created layers of elements that constitute a layer of constant
thickness aligned with the coordinate grid axes (X, Y and/or Z), so
that this code can assign CPML flags to them. This code does NOT check
that (because it cannot, in any easy way). The mesh inside these CPML
layers does not need to be structured nor regular, any non-structured
mesh is fine as long as it has flat PML inner and outer faces, parallel
to the axes, and thus of a constant thickness. The thickness can be
different for the X, Y and Z sides. But for X it must not vary, for
Y it must not vary, and for Z it must not vary. If you do not know
the exact thickness, you can use a slightly LARGER value in this code
(say 2\% to 5\% more) and this code will fix that and will adjust
it; never use a SMALLER value otherwise this code will miss some CPML
elements.

Note: in a future release we will remove the constraint of having
CPML layers aligned with the coordinate axes; we will allow for meshes
that are titled by any constant angle in the horizontal plane. However
this is not implemented yet.

Note: in the case of fluid-solid layers in contact, the fluid-solid
interface currently needs to be flat and horizontal inside the CPML
layer (i.e., bathymetry should become flat and horizontal when entering
the CPML); this (small) constraint will probably remain in the code
for a while because it makes fluid-solid matching inside the CPML
much easier.

\item [{materials\_file}] Contains the material associations for each element.
The format is:

\begin{lyxcode}
element\_ID~material\_ID
\end{lyxcode}

where \texttt{element\_ID} is the element identifier and \texttt{material\_ID}
is a unique identifier, positive (for materials taken from this list
of materials, i.e. for which each spectral element has constant material
properties taken from this list) or negative (for tomographic models,
i.e. for spectral element whose real velocities and density will be
assigned later by calling an external function to define model variations,
for instance in the case of tomographic models; in such a case, material
properties can vary inside each spectral element, i.e. be different
at each of its Gauss-Lobatto-Legendre grid points).

\item [{nummaterial\_velocity\_file}] Defines the material properties.

\begin{itemize}
\item For classical materials (i.e., spectral elements for which the velocity
and density model will not be assigned by calling an external function
to define for instance a tomographic model), the format is:

\begin{lyxcode}
domain\_ID~material\_ID~rho~vp~vs~Qkappa~Qmu~anisotropy\_flag
\end{lyxcode}

where \texttt{\textbf{domain\_ID}}\textbf{ is 1 for acoustic and 2
for elastic or viscoelastic materials,} \texttt{material\_ID} a unique
identifier, \texttt{rho} the density in $kg\, m^{-3}$, \texttt{vp}
the P-wave speed in $m\, s^{-1}$, \texttt{vs} the S-wave speed in
$m\, s^{-1}$, \texttt{Q} the quality factor and \texttt{anisotropy\_flag}
an identifier for anisotropic models. Note that the \texttt{Qkappa}
value is ignored by the code unless \texttt{FULL\_ATTENUATION\_SOLID}
is set. Note also that both \texttt{Qkappa} and \texttt{Qmu} are ignored
by the code unless \texttt{ATTENUATION} is set. If you want a model
with no \texttt{Qmu} attenuation, both set \texttt{ATTENUATION} to
\texttt{.false.} in the \texttt{Par\_file} and set \texttt{Qmu} to
9999 here. If you want a model with no \texttt{Qkappa} attenuation,
both set \texttt{FULL\_ATTENUATION\_SOLID} to \texttt{.false.} in
the \texttt{Par\_file} and set \texttt{Qkappa} to 9999 here. %%magnoni

\item For tomographic velocity models, please read Chapter \ref{cha:-Changing-the}
and Section \ref{sec:Using-tomographic} `Using external tomographic
Earth models' for further details.
\end{itemize}
\item [{nodes\_coords\_file}] Contains the point locations in Cartesian
coordinates of the mesh element corners.
\item [{mesh\_file}] Contains the mesh element connectivity. The hexahedral
elements can have 8 or 27 nodes.\\
 See picture doc/mesh\_numbering\_convention/numbering\_convention\_27\_nodes.jpg
to see\\
 in which (standard) order the points must be cited. In the case of
8 nodes, just include the first 8 points.
\item [{free\_or\_absorbing\_surface\_file\_zmax}] Contains the free surface
connectivity or \\
 the surface connectivity of the absorbing boundary surface at the
top (Zmax), \\
 depending on whether the top surface is defined as free or absorbing
(\texttt{STACEY\_INSTEAD\_OF\_FREE\_SURFACE} in \texttt{DATA/Par\_file}). \\
 You should put both the surface of acoustic regions and of elastic
regions in that file; that is, list all the element faces that constitute
the surface of the model in that file.
\item [{absorbing\_surface\_file\_xmax}] Contains the surface connectivity
of the absorbing boundary surface at Xmax \\
(also needed in the case of C-PML absorbing conditions, in order for
the code to be able to impose Dirichlet conditions on their outer
edge).
\item [{absorbing\_surface\_file\_xmin}] Contains the surface connectivity
of the absorbing boundary surface at Xmin \\
(also needed in the case of C-PML absorbing conditions, in order for
the code to be able to impose Dirichlet conditions on their outer
edge).
\item [{absorbing\_surface\_file\_ymax}] Contains the surface connectivity
of the absorbing boundary surface at Ymax \\
(also needed in the case of C-PML absorbing conditions, in order for
the code to be able to impose Dirichlet conditions on their outer
edge).
\item [{absorbing\_surface\_file\_ymin}] Contains the surface connectivity
of the absorbing boundary surface at Ymin \\
(also needed in the case of C-PML absorbing conditions, in order for
the code to be able to impose Dirichlet conditions on their outer
edge).
\item [{absorbing\_surface\_file\_bottom}] Contains the surface connectivity
of the absorbing boundary surface at the bottom (Zmin) \\
(also needed in the case of C-PML absorbing conditions, in order for
the code to be able to impose Dirichlet conditions on their outer
edge).
\end{description}
These mesh files are needed as input files for the partitioner \texttt{xdecompose\_mesh}
to load-balance the mesh. Please see the next section for further
details.

In directory "CUBIT\_GEOCUBIT" we provide a script that can help doing the above tasks of exporting a CUBIT mesh
to SPECFEM3D format automatically for you: "run\_cubit2specfem3d.py".
Just edit them to indicate the path to your local installation of CUBIT and also the name of the *.cub existing CUBIT mesh file
that you want to export to SPECFEM3D format. These scripts will do the conversion for you automatically except assigning material properties
to the different mesh layers. To do so, you will then need to edit the file
called "nummaterial\_velocity\_file" that will have just been created and change it from the prototype created:

    0 1 vol1 --> syntax: \#material\_domain\_id \#material\_id \#rho \#vp \#vs \#Q\_kappa \#Q\_mu \#anisotropy
    0 2 vol2 --> syntax: \#material\_domain\_id \#material\_id \#rho \#vp \#vs \#Q\_kappa \#Q\_mu \#anisotropy

(where "vol1" and "vol2" here represent the volume labels that you have set while creating the mesh in CUBIT)
to for instance

    2 1 1500 2300 1800 9999.0 9999.0 0
    2 2 1600 2500 20000 9999.0 9999.0 0

\subsubsection*{Checking the mesh quality}

The quality of the mesh may be inspected more precisely based upon
the serial code in the file \texttt{check\_mesh\_quality\_}~\\
 \texttt{CUBIT\_Abaqus.f90} located in the directory \texttt{src/check\_mesh\_quality\_CUBIT\_Abaqus/}.
Running this code is optional because no information needed by the
solver is generated.

Prior to running and compiling this code, you have to export your
mesh in CUBIT to an ABAQUS (.inp) format. For example, export mesh
block IDs belonging to volumes in order to check the quality of the
hexahedral elements. You also have to determine a number of parameters
of your mesh, such as the number of nodes and number of elements and
modify the header of the \texttt{check\_mesh\_quality\_CUBIT\_Abaqus.f90}
source file in directory \texttt{src/check\_mesh\_quality\_CUBIT\_Abaqus/}.

\noindent
Then, in the main directory, type
\begin{lyxcode}
{\small make~xcheck\_mesh\_quality\_CUBIT\_Abaqus}{\small \par}
\end{lyxcode}
and use
\begin{lyxcode}
{\small ./bin/xcheck\_mesh\_quality\_CUBIT\_Abaqus}{\small \par}
\end{lyxcode}
to generate an %not supported: AVS output file (\texttt{\small AVS\_meshquality.inp} in AVS UCD format) or
OpenDX output file (\texttt{\small DX\_mesh\_quality.dx}{\small )
that can be used to investigate mesh quality, e.g. skewness of elements,
and a Gnuplot histogram (}\texttt{\small mesh\_quality\_histogram.txt}{\small )
that can be plotted with gnuplot (type `}\texttt{\small gnuplot plot\_mesh\_quality\_histogram.gnu}{\small ').
The histogram is also printed to the screen. Analyze that skewness
histogram of mesh elements to make sure no element has a skewness
above approximately 0.75, otherwise the mesh is of poor quality (and
if even a single element has a skewness value above 0.80, then you
must definitely improve the mesh). If you want to start designing
your own meshes, this tool is useful for viewing your creations. You
are striving for meshes with elements with `cube-like' dimensions,
e.g., the mesh should contain no very elongated or skewed elements.}{\small \par}

\subsection{Partitioning the Mesh with \texttt{xdecompose\_mesh}}

The SPECFEM3D Cartesian software package performs large scale simulations
in a parallel 'Single Process Multiple Data' way. The spectral-element
mesh created with CUBIT needs to be distributed on the processors.
This partitioning is executed once and for all prior to the execution
of the solver so it is referred to as a static mapping.

An efficient partitioning is important because it leverages the overall
running time of the application. It amounts to balance the number
of elements in each slice while minimizing the communication costs
resulting from the placement of adjacent elements on different processors.
\texttt{decompose\_mesh} depends on the SCOTCH library \citep{PeRo96},
which provides efficent static mapping, graph and mesh partitioning
routines. SCOTCH is a free software package developed by Fran\c{c}ois
Pellegrini et al. from LaBRI and INRIA in Bordeaux, France, downloadable
from the web page \url{https://gforge.inria.fr/projects/scotch/}.
\\

In most cases, the configuration with \texttt{./configure~FC=ifort}
should be sufficient. During the configuration process, the script
tries to find existing SCOTCH installations. In case your system has
no pre-existing SCOTCH installation, we provide the source code of
SCOTCH, which is released open source under the French CeCILL-C version
1 license, in directory \texttt{src/decompose\_mesh/scotch\_5.1.12b}.
This version gets bundled with the compilation of the SPECFEM3D Cartesian
package if no libraries could have been found. If this automatic compilation
of the SCOTCH libraries fails, please refer to file INSTALL.txt in
that directory to see further details how to compile it on your system.
In case you want to use a pre-existing installation, make sure you
have correctly specified the path of the SCOTCH library when using
the option \texttt{-{}-with-scotch-dir} with the \texttt{./configure}
script. In the future you should be able to find more recent versions
at \url{http://www.labri.fr/perso/pelegrin/scotch/scotch_en.html}
. \\

\begin{figure}[htbp]
\noindent \begin{centering}
\includegraphics[width=0.45\textwidth]{figures/mount-partitions.jpg}
\includegraphics[width=0.45\textwidth]{figures/mount-partitions2.jpg}
\par\end{centering}

\caption{Example of a mesh partitioning onto four cores. Each single core partition
is colored differently. The executable \texttt{xdecompose\_mesh} can
equally distribute the mesh on any arbitrary number of cores. Domain
decomposition is explained in detail in \citet{MaKoBlLe08}, and excellent
scaling up to 150,000 processor cores in shown for instance in \citet{CaKoLaTiMiLeSnTr08,KoLaMi08a,MaKoBlLe08,KoErGoMi10,Kom11}.}

\label{fig:mount.partitions}
\end{figure}

When you are ready to compile, in the main directory type
\begin{lyxcode}
{\small make~xdecompose\_mesh}{\small \par}
\end{lyxcode}
If all paths and flags have been set correctly, the executable \texttt{bin/xdecompose\_mesh}
should be produced.

The partitioning is done in serial for now (in the next release we
will provide a parallel version of that code). It needs to be run in the main directory because it expects the \texttt{./DATA/Par\_file}. The synopsis is:
\begin{lyxcode}
./bin/xdecompose\_mesh~nparts~input\_directory~output\_directory
\end{lyxcode}
where
\begin{itemize}
\item \texttt{nparts} is the number of partitions, i.e., the number of cores
for the parallel simulations,
\item \texttt{input\_directory} is the directory which holds all the files
generated by the Python script \texttt{cubit2specfem3d.py} explained
in the previous Section~\ref{subsec:Exporting-the-Mesh}, e.g. \texttt{./MESH/},
and
\item \texttt{output\_directory} is the directory for the output of this
partitioner which stores ACII-format files named like \texttt{proc{*}{*}{*}{*}{*}{*}\_Database}
for each partition. These files will be needed for creating the distributed
databases, and have to reside in the directory \texttt{LOCAL\_PATH}
specified in the main \texttt{Par\_file}, e.g. in directory \texttt{./OUTPUT\_FILES/DATABASES\_MPI}.
Please see Chapter~\ref{cha:Creating-Distributed-Databases} for
further details.
\end{itemize}
Note that all the files generated by the Python script \texttt{cubit2specfem3d.py}
must be placed in the \texttt{input\_directory} folder before running
the program.

\section{Meshing with \texttt{xmeshfem3D}}\label{cha:Running-the-Mesher-Meshfem3D}

In case you successfully ran the configuration script, you are also
ready to compile the internal mesher. This is an alternative to CUBIT
for the mesh generation of relatively simple models.

\begin{figure}[htbp]
\noindent \begin{centering}
\includegraphics[scale=0.5]{figures/socal_map_mpi.pdf}
\par\end{centering}

\caption{For parallel computing purposes,
the model block is subdivided in $\nprocxi\times\nproceta$ slices
of elements. In this example we use $5^{2}=25$ processors. }
\label{fig:For-parallel-computing}
\end{figure}

\noindent
In the main directory, type
\begin{lyxcode}
{\small make~xmeshfem3D}{\small \par}
\end{lyxcode}
If all paths and flags have been set correctly, the mesher should
now compile and produce the executable \texttt{bin/xmeshfem3D}. Please
note that \texttt{xmeshfem3D} must be called directly from the main
directory, as most of the binaries of the package.

Input for the mesh generation program is provided through the parameter
file \texttt{Mesh\_Par\_file}, which resides in the subdirectory \texttt{DATA/meshfem3D\_files/}.
(to see how to use it, see the EXAMPLES specific to the internal mesher in directory \texttt{EXAMPLES/meshfem3D\_examples/}.
Before running the mesher, a number of parameters need to be set in
the \texttt{Mesh\_Par\_file}. This requires a basic understanding
of how the SEM is implemented, and we encourage you to read \citet{KoVi98,KoTr99}
and \citet{KoLiTrSuStSh04}.

The mesher and the solver use UTM coordinates internally, therefore
you need to define the zone number for the UTM projection (e.g., zone
11 for Los Angeles). Use decimal values for latitude and longitude
(no minutes/seconds). These values are approximate; the mesher will
round them off to define a square mesh in UTM coordinates. When running
benchmarks on rectangular models, turn the UTM projection off by using
the flag \texttt{\small SUPPRESS\_UTM\_PROJECTION}{\small , in which
case all `longitude' parameters simply refer to the $x$~axis, and
all `latitude' parameters simply refer to the $y$~axis. To run the
mesher for a global simulation, the following parameters need to be
set in the }\texttt{\small Mesh\_Par\_file}{\small :}{\small \par}
\begin{description}
\item [{\texttt{LATITUDE\_MIN}}] Minimum latitude in the block (negative
for South).
\item [{\texttt{LATITUDE\_MAX}}] Maximum latitude in the block.
\item [{\texttt{LONGITUDE\_MIN}}] Minimum longitude in the block (negative
for West).
\item [{\texttt{LONGITUDE\_MAX}}] Maximum longitude in the block.
\item [{\texttt{DEPTH\_BLOCK\_KM}}] Depth of bottom of mesh in kilometers.
\item [{\texttt{\noun{UTM\_PROJECTION\_ZONE}}}] UTM projection zone in
which your model resides, only valid when \texttt{SUPPRESS\_UTM\_}~\\
 \texttt{PROJECTION} is \texttt{.false.}.
Use a negative zone number for the Southern hemisphere:
the Northern hemisphere corresponds to zones +1 to +60,
the Southern hemisphere to zones -1 to -60.
We use the WGS84 (World Geodetic System 1984) reference ellipsoid for the UTM projection. If you prefer to use the Clarke 1866 ellipsoid,
edit file \texttt{src/shared/utm\_geo.f90}, uncomment that ellipsoid and recompile the code.
From http://en.wikipedia.org/wiki/Universal\_Transverse\_Mercator\_coordinate\_system :
The Universal Transverse Mercator coordinate system was developed by the United States Army Corps of Engineers in the 1940s.
The system was based on an ellipsoidal model of Earth. For areas within the contiguous United States
the Clarke Ellipsoid of 1866 was used. For the remaining areas of Earth, including Hawaii, the International Ellipsoid was used.
The WGS84 ellipsoid is now generally used to model the Earth in the UTM coordinate system,
which means that current UTM northing at a given point can be 200+ meters different from the old one.
For different geographic regions, other datum systems (e.g.: ED50, NAD83) can be used.
\item [{\texttt{SUPPRESS\_UTM\_PROJECTION}}] set to be \texttt{.false.}
when your model range is specified in geographical coordinates, and
needs to be \texttt{.true.} when your model is specified in Cartesian
coordinates. \noun{UTM projection zone in which your simulation region
resides.}
\item [{\texttt{INTERFACES\_FILE }}] File which contains the description
of the topography and of the interfaces between the different layers
of the model, if any. The number of spectral elements in the vertical
direction within each layer is also defined in this file.
\item [{$\nexxi$}] The number of spectral elements along one side of the
block. This number \textit{must} be 8~$\times$~a multiple of $\nprocxi$
defined below. Based upon benchmarks against semi-analytical discrete
wavenumber synthetic seismograms \citep{KoLiTrSuStSh04}, determined
that a $\nexxi=288$ run is accurate to a shortest period of roughly
2~s. Therefore, since accuracy is determined by the number of grid
points per shortest wavelength, for any particular value of $\nexxi$
the simulation will be accurate to a shortest period determined by
\begin{equation}
\mbox{shortest period (s)}=(288/\nexxi)\times2.\label{eq:shortest_period}
\end{equation}
 The number of grid points in each orthogonal direction of the reference
element, i.e., the number of Gauss-Lobatto-Legendre points, is determined
by \texttt{NGLLX} in the \texttt{constants.h} file. We generally use
$\mbox{\texttt{NGLLX\/}}=5$, for a total of $5^{3}=125$ points per
elements. We suggest not to change this value.
\item [{$\nexeta$}] The number of spectral elements along the other side
of the block. This number \textit{must} be 8~$\times$~a multiple
of $\nproceta$ defined below.
\item [{$\nprocxi$}] The number of processors or slices along one side
of the block (see Figure~\ref{fig:For-parallel-computing}); we must
have $\nexxi=8\times c\times\nprocxi$, where $c\ge1$ is a positive
integer.
\item [{$\nproceta$}] The number of processors or slices along the other
side of the block; we must have $\nexeta=8\times c\times\nproceta$,
where $c\ge1$ is a positive integer.
\item [{\texttt{USE\_REGULAR\_MESH}}] set to be \texttt{.true.} if you
want a perfectly regular mesh or \texttt{.false.} if you want to add
doubling horizontal layers to coarsen the mesh. In this case, you
also need to provide additional information by setting up the next
three parameters.
\item [{\texttt{NDOUBLINGS}}] The number of horizontal doubling layers.
Must be set to \texttt{1} or \texttt{2} if \texttt{USE\_REGULAR\_MESH}
is set to \texttt{.true.}.
\item [{\texttt{NZ\_DOUBLING\_1}}] The position of the first doubling layer
(only interpreted if \texttt{USE\_REGULAR\_MESH} is set to \texttt{.true.}).
\item [{\texttt{NZ\_DOUBLING\_2}}] The position of the second doubling
layer (only interpreted if \texttt{USE\_REGULAR\_MESH} is set to \texttt{.true.}
and if \texttt{NDOUBLINGS} is set to \texttt{2}).
\item [{\texttt{CREATE\_ABAQUS\_FILES}}] Set this flag to \texttt{.true.}
to save Abaqus FEA \urlwithparentheses{www.simulia.com} mesh files
for subsequent viewing. Turning the flag on generates files in the
\texttt{LOCAL\_PATH} directory. See Section~\ref{sec:Mesh-graphics}
for a discussion of mesh viewing features.
\item [{\texttt{CREATE\_DX\_FILES}}] Set this flag to \texttt{.true.} to
save OpenDX \urlwithparentheses{www.opendx.org} mesh files for subsequent
viewing.
\item [{\texttt{LOCAL\_PATH}}] Directory in which the partitions generated
by the mesher will be written. Generally one uses a directory on the
local disk of the compute nodes, although on some machines these partitions
are written on a parallel (global) file system (see also the earlier
discussion of the \texttt{LOCAL\_PATH\_IS\_ALSO\_GLOBAL} flag in Chapter~\ref{cha:Getting-Started}).
The mesher generates the necessary partitions in parallel, one set
for each of the $\nprocxi\times\nproceta$ slices that constitutes
the mesh (see Figure~\ref{fig:For-parallel-computing}). After the
mesher finishes, you can log in to one of the compute nodes and view
the contents of the \texttt{LOCAL\_PATH} directory to see the files
generated by the mesher. These files will be needed for creating the
distributed databases, and have to reside in the directory \texttt{LOCAL\_PATH}
specified in the main \texttt{Par\_file}, e.g. in directory \texttt{OUTPUT\_FILES/DATABASES\_MPI}.
Please see Chapter~\ref{cha:Creating-Distributed-Databases} for
further details.
\item [{\texttt{NMATERIALS}}] The number of different materials in your
model. In the following lines, each material needs to be defined as
:

\begin{lyxcode}
material\_ID~rho~vp~vs~Q~anisotropy\_flag~domain\_ID
\end{lyxcode}

where
\begin{itemize}
\item \texttt{Q} : quality factor (0=no attenuation)
\item \texttt{anisotropy\_flag} : 0=no anisotropy / 1,2,.. check with implementation
in \texttt{aniso\_model.f90}
\item \texttt{domain\_id} : 1=acoustic / 2=elastic
\end{itemize}
\item [{\texttt{NREGIONS}}] The number of regions in the mesh. In the following
lines, because the mesh is regular or 'almost regular', each region
is defined as :

\begin{lyxcode}
NEX\_XI\_BEGIN~NEX\_XI\_END~NEX\_ETA\_BEGIN~NEX\_ETA\_END~NZ\_BEGIN~NZ\_END~material\_ID
\end{lyxcode}
\end{description}
The \texttt{INTERFACES\_FILE} parameter of \texttt{Mesh\_Par\_File}
defines the file which contains the settings of the topography grid
and of the interfaces grids. Topography is defined as a set of elevation
values on a regular 2D grid. It is also possible to define interfaces
between the layers of the model in the same way. The file needs to
define several parameters:
\begin{itemize}
\item The number of interfaces, including the topography. This needs to
be set at the first line. Then, from the bottom to the top of the
model, you need to define the grids with:
\item \texttt{SUPPRESS\_UTM\_PROJECTION} flag as described previously,
\item number of points along $x$ and $y$ direction (NXI and NETA),
\item minimal $x$ and $y$ coordinates (LONG\_MIN and LAT\_MIN),
\item spacing between points along $x$ and $y$ (SPACING\_XI and SPACING\_ETA)
and
\item the name of the file which contains the elevation values (in $y$.$x$
increasing order).
\end{itemize}
At the end of this file, you simply need to set the number of spectral
elements in the vertical direction for each layer. We provide a few
models in the {\texttt{EXAMPLES/}} directory. \\

Finally, depending on your system, you might need to provide a file
that tells MPI what compute nodes to use for the simulations. The
file must have a number of entries (one entry per line) at least equal
to the number of processors needed for the run. A sample file is provided
in the file \texttt{mymachines}. This file is not used by the mesher
or solver, but is required by the \texttt{go\_mesher} and \texttt{go\_solver}
default job submission scripts. See Chapter~\ref{cha:Scheduler}
for information about running the code on a system with a scheduler,
e.g., LSF.

Now that you have set the appropriate parameters in the \texttt{Mesh\_Par\_file}
and have compiled the mesher, you are ready to launch it! This is
most easily accomplished based upon the \texttt{go\_mesher} script.
When you run on a PC cluster, the script assumes that the nodes are
named n001, n002, etc. If this is not the case, change the \texttt{tr
-d `n'} line in the script. You may also need to edit the last command
at the end of the script that invokes the \texttt{mpirun} command.
See Chapter~\ref{cha:Scheduler} for information about running the
code on a system with a scheduler, e.g., LSF.

Mesher output is provided in the \texttt{OUTPUT\_FILES} directory
in \texttt{output\_mesher.txt}; this file provides lots of details
about the mesh that was generated. Please note that the mesher suggests
a time step \texttt{DT} to run the solver with. The mesher output
file also contains a table about the quality of the mesh to indicate
possible problems with the distortions of elements. Alternatively,
output can be directed to the screen instead by uncommenting a line
in \texttt{constants.h}:
\begin{lyxcode}
!~uncomment~this~to~write~messages~to~the~screen~

!~integer,~parameter~::~IMAIN~=~ISTANDARD\_OUTPUT
\end{lyxcode}
To control the quality of the mesh, check the standard output (either
on the screen or in the \texttt{OUTPUT\_FILES} directory in \texttt{output\_mesher.txt})
and analyze the skewness histogram of mesh elements to make sure no
element has a skewness above approximately 0.75, otherwise the mesh
is of poor quality (and if even a single element has a skewness value
above 0.80, then you must definitely improve the mesh). To draw the
skewness histogram on the screen, type \texttt{gnuplot plot\_mesh\_quality\_histogram.gnu}.

### 2015-04-24 doc/USER_MANUAL/04_creating_databases.tex

\chapter{Creating the Distributed Databases}\label{cha:Creating-Distributed-Databases}

After using \texttt{xmeshfem3D} or \texttt{xdecompose\_mesh}, the
next step in the workflow is to compile \texttt{xgenerate\_}~\\
 \texttt{databases}. This program is going to create all the missing
information needed by the SEM solver.

\begin{figure}[htbp]
\noindent \begin{centering}
\includegraphics[width=0.6\textwidth]{figures/workflow.pdf}
\par\end{centering}

\caption{Schematic workflow for a SPECFEM3D Cartesian simulation. The executable
\texttt{xgenerate\_databases} creates the GLL mesh points and assigns
specific model parameters.}

\label{fig:workflow.databases}
\end{figure}

\noindent
In the main directory, type
\begin{lyxcode}
{\small make~xgenerate\_databases}{\small \par}
\end{lyxcode}
Input for the program is provided through the main parameter file
\texttt{Par\_file}, which resides in the subdirectory \texttt{DATA}.
Please note that \texttt{xgenerate\_databases} must be called directly
from the main directory, as most of the binaries of the package.

\section{Main parameter file \texttt{Par\_file}}\label{cha:Main-Parameter}

Before running \texttt{xgenerate\_databases}, a number of parameters
need to be set in the main parameter \texttt{Par\_file} located in
the subdirectory \texttt{DATA}:
\begin{description}
\item [{\texttt{SIMULATION\_TYPE}}] is set to 1 for forward simulations,
2 for adjoint simulations (see Section \ref{sec:Adjoint-simulation-finite})
and 3 for kernel simulations (see Section \ref{sec:Finite-Frequency-Kernels}).
\item [{\texttt{SAVE\_FORWARD}}] is only set to \texttt{.true.} for a forward
simulation with the last frame of the simulation saved, as part of
the finite-frequency kernel calculations (see Section \ref{sec:Finite-Frequency-Kernels}).
For a regular forward simulation, leave \texttt{SIMULATION\_TYPE}
and \texttt{SAVE\_FORWARD} at their default values.
\item [{\texttt{\noun{UTM\_PROJECTION\_ZONE}}}] UTM projection zone in
which your model resides, only valid when \texttt{SUPPRESS\_UTM\_}~\\
 \texttt{PROJECTION} is \texttt{.false.}.
\item [{\texttt{SUPPRESS\_UTM\_PROJECTION}}] set to be \texttt{.false.}
when your model range is specified in the geographical coordinates,
and needs to be \texttt{.true.} when your model is specified in a
cartesian coordinates. \noun{UTM projection zone in which your simulation
region resides.}
\item [{\texttt{NPROC}}] The number of MPI processors, each one is assigned
one slice of the whole mesh.
\item [{\texttt{NSTEP}}] The number of time steps of the simulation. This
controls the length of the numerical simulation, i.e., twice the number
of time steps requires twice as much CPU time. This feature is not
used at the time of generating the distributed databases but is required
for the solver, i.e., you may change this parameter after running
\texttt{xgenerate\_databases}.
\item [{\texttt{DT}}] The length of each time step in seconds. This feature
is not used at the time of generating the distributed databases but
is required for the solver. Please see also Section~\ref{sec:Choosing-the-Time-Step}
for further details.
\item [{\texttt{NGNOD}}] The number of nodes for 2D and 3D shape functions
for hexahedra. We use either 8-node mesh elements (bricks) or 27-node
elements. If you use the internal mesher, the only option is 8-node
bricks (27-node elements are not supported). \texttt{CUBIT} does not
support HEX27 elements either (it can generate them, but they are
flat, i.e. identical to HEX8). To generate HEX27 elements with curvature
properly taken into account, you can use \texttt{Gmsh} \url{http://geuz.org/gmsh/}
\item [{\texttt{MODEL}}] Must be set to one of the following:

\begin{description}
\item [{\textmd{Models defined by mesh parameters:}}] ~

\begin{description}
\item [{\texttt{default}}] Uses model parameters as defined by meshing
procedures described in the previous Chapter~\ref{cha:Mesh-Generation}.
\end{description}

\item [{\textmd{1D~models~with~real~structure:}}] ~

\begin{description}
\item [{\texttt{1D\_prem}}] Isotropic version of the spherically symmetric
Preliminary Reference Earth Model (PREM) \citep{DzAn81}.
\item [{\texttt{1D\_socal}}] A standard isotropic 1D model for Southern
California.
\item [{\texttt{1D\_cascadia}}] Isotropic 1D profile for the Cascadia region.
\end{description}

\item [{\textmd{Fully~3D~models:}}] ~

\begin{description}
\item [{\texttt{aniso}}] For a user-specified fully anisotropic model.
Parameters are set up in routines located in file \texttt{model\_aniso.f90}
in directory \texttt{src/generate\_databases/}. See Chapter~\ref{cha:-Changing-the}
for a discussion on how to specify your own 3D model.

\item [{\texttt{external}}] For a user-specified isotropic model which
uses externally defined model parameters. Uses external model definitions
set up in routines located in file \texttt{model\_external\_values.f90}
in directory \texttt{src/generate\_databases/}. Please modify these
generic template routines to use your own model definitions.
\item [{\texttt{gll}}] For a user-specified isotropic model which uses
external binary files for $v_{p}$, $v_{s}$ and $\rho$. Binary files
are given in the same format as when outputted by the \texttt{xgenerate\_databases}
executable when using option \texttt{SAVE\_MESH\_FILES}. These binary
files define the model parameters on all GLL points which can be used
for iterative inversion procedures.
\item [{\texttt{salton\_trough}}] A 3D $V_{p}$ model for Southern California.
Users must provide the corresponding data file \texttt{regrid3\_vel\_p.bin}
in directory \texttt{DATA/st\_3D\_block\_harvard/}.
\item [{\texttt{tomo}}] For a user-specified 3D isotropic model which uses
a tomographic model file \texttt{tomographic\_model.xyz} in directory
\texttt{DATA}. See Section \ref{sec:Using-tomographic}, for a discussion
on how to specify your own 3D tomographic model.
\end{description}
\end{description}
\item [{\texttt{APPROXIMATE\_OCEAN\_LOAD}}] Set to \texttt{.true.} if the
effect of the oceans on seismic wave propagation should be incorporated
based upon the (rough) approximate treatment discussed in \citet{KoTr02b}.
This feature is inexpensive from a numerical perspective, both in
terms of memory requirements and CPU time. This approximation is accurate
at periods of roughly 20~s and longer. At shorter periods the effect
of water phases/reverberations is not taken into account, even when
the flag is on. If you want to model the effect of a fluid-solid model
at short periods, then set this flag to \texttt{.false.} and mesh
the fluid layer explicitly in your mesher, so that it is computed
accurately and without this approximation.
\item [{\texttt{TOPOGRAPHY}}] This feature is only effective if \texttt{APPROXIMATE\_OCEAN\_LOAD}
is set to \texttt{.true.}. Set to \texttt{.true.} if topography and
bathymetry should be read in based upon the topography file specified
in the main constants file \texttt{constants.h} found in subdirectory
\texttt{src/shared/} to evaluate elevations. If not set, elevations
will be read from the numerical mesh.
\item [{\texttt{ATTENUATION}}] Set to \texttt{.true.} if attenuation should
be incorporated. Turning this feature on increases the memory requirements
significantly (roughly by a factor of~1.5), and is numerically fairly
expensive. See \citet{KoTr99,KoTr02a} for a discussion on the implementation
of attenuation based upon standard linear solids. Please note that
the Vp- and Vs-velocities of your model are given for a reference
frequency. To change this reference frequency, you change the value
of \texttt{ATTENUATION\_f0\_REFERENCE} in the main constants file
\texttt{constants.h} found in subdirectory \texttt{src/shared/}.
\item [{\texttt{FULL\_ATTENUATION\_SOLID}}] Set to \texttt{.true.} to add
$Q_{\kappa}$ attenuation to the simulations in addition to $Q_{\mu}$
attenuation when the \texttt{ATTENUATION} flag above is on. Should
be off in almost all cases for seismic wave propagation, since $Q_{\kappa}$
is almost always negligible in Earth models. However, this parameter
can be useful for instance for ocean acoustics simulations, when $Q_{\kappa}$
is not negligible in the solid layer at the sea bottom.
\item [{\texttt{ANISOTROPY}}] Set to \texttt{.true.} if you want to use
an anisotropy model. Please see the file \texttt{model\_aniso.f90}
in subdirectory \texttt{src/generate\_databases/} for the current
implementation of anisotropic models.
\item [{\texttt{TOMOGRAPHY\_PATH}}] Directory in which the tomography files
are stored for using external tomographic Earth models (please read
Chapter \ref{cha:-Changing-the} and Section \ref{sec:Using-tomographic}
`Using external tomographic Earth models' for further details.).
\item [{\texttt{USE\_OLSEN\_ATTENUATION}}] Set to \texttt{.true.} if you
want to use the attenuation model that scaled from the S-wave speed
model using Olsen's empirical relation (see \citet{OlDaBr03}).
\item [{\texttt{OLSEN\_ATTENUATION\_RATIO}}] Determines the Olsen's constant
in Olsen's empirical relation (see \citet{OlDaBr03}).
\item [{\texttt{PML\_CONDITIONS}}] Set to \texttt{.true.} to turn on C-PML
boundary conditions for a regional simulation. Both fluids and elastic
solids are supported. Note that xmeshfem3d generated meshes do not support C-PML yet.
\item [{\texttt{PML\_INSTEAD\_OF\_FREE\_SURFACE}}] Set to \texttt{.true.}
to turn on C-PML boundary conditions on the top surface instead of
the usual free surface.
\item [{\texttt{f0\_FOR\_PML}}] Determines the dominant frequency that
will be used in the calculation of PML damping profiles; this should
be set to the same (or similar) dominant frequency as that of the
source that you will use in your simulation. If you plan to use a
Dirac source, then use the dominant frequency of the source wavelet
with which you plan to convolve your seismograms later on in post-processing.
\item [{\texttt{STACEY\_ABSORBING\_CONDITIONS}}] Set to \texttt{.true.}
to turn on Clayton-Enquist absorbing boundary conditions (see \citet{KoTr99}).
In almost all cases it is much better to use CPML absorbing layers
(see the options above) and leave this flag to \texttt{.false.}.
\item [{\texttt{STACEY\_INSTEAD\_OF\_FREE\_SURFACE}}] Set to \texttt{.true.}
to turn on absorbing boundary conditions on the top surface which
by default constitutes a free surface of the model.
\item [{\texttt{CREATE\_SHAKEMAP}}] Set this flag to \texttt{.true.} to
create a ShakeMap\textregistered{}, i.e., a peak ground velocity map
of the maximum absolute value of the two horizontal components of
the velocity vector.
\item [{\texttt{MOVIE\_SURFACE}}] Set to \texttt{.false.}, unless you want
to create a movie of seismic wave propagation on the Earth's surface.
Turning this option on generates large output files. See Section~\ref{sec:Movies}
for a discussion on the generation of movies. This feature is only
relevant for the solver.
\item [{\texttt{MOVIE\_TYPE}}] Set this flag to 1 to show the top surface
(tomography + oceans) only, to 2 to show all external faces of the
mesh (i.e. topography + vertical edges + bottom) in shakemaps and
surface movies.
\item [{\texttt{MOVIE\_VOLUME}}] Set to \texttt{.false.}, unless you want
to create a movie of seismic wave propagation in the Earth's interior.
Turning this option on generates huge output files. See Section~\ref{sec:Movies}
for a discussion on the generation of movies. This feature is only
relevant for the solver.
\item [{\texttt{SAVE\_DISPLACEMENT}}] Set this flag to \texttt{.true.}
if you want to save the displacement instead of velocity for the movie
frames.
\item [{\texttt{USE\_HIGHRES\_FOR\_MOVIES}}] Set this flag to \texttt{.true.}
if you want to save the values at all the NGLL grid points for the
movie frames.
\item [{\texttt{NTSTEP\_BETWEEN\_FRAMES}}] Determines the number of timesteps
between movie frames. Typically you want to save a snapshot every
100 timesteps. The smaller you make this number the more output will
be generated! See Section~\ref{sec:Movies} for a discussion on the
generation of movies. This feature is only relevant for the solver.
\item [{\texttt{HDUR\_MOVIE}}] Determines the half duration of the source
time function for the movie simulations. When this parameter is set
to be 0, a default half duration that corresponds to the accuracy
of the simulation is provided. Otherwise, it adds this half duration
to the half duration specified in the source file \texttt{CMTSOLUTION},
thus simulates longer periods to make the movie images look smoother.
\item [{\texttt{SAVE\_MESH\_FILES}}] Set this flag to \texttt{.true.} to
save ParaView \urlwithparentheses{www.paraview.org} mesh files for
subsequent viewing. Turning the flag on generates large (distributed)
files in the \texttt{LOCAL\_PATH} directory. See Section~\ref{sec:Mesh-graphics}
for a discussion of mesh viewing features.
\item [{\texttt{LOCAL\_PATH}}] Directory in which the distributed databases
will be written. Generally one uses a directory on the local disk
of the compute nodes, although on some machines these databases are
written on a parallel (global) file system (see also the earlier discussion
of the \texttt{LOCAL\_PATH\_IS\_ALSO\_GLOBAL} flag in Chapter~\ref{cha:Getting-Started}).
\texttt{xgenerate\_databases} generates the necessary databases in
parallel, one set for each of the \texttt{NPROC} slices that constitutes
the mesh (see Figure~\ref{fig:mount.partitions} and Figure~\ref{fig:For-parallel-computing}).
After the executable finishes, you can log in to one of the compute
nodes and view the contents of the \texttt{LOCAL\_PATH} directory
to see the (many) files generated by \texttt{xgenerate\_databases}.
Please note that the \texttt{LOCAL\_PATH} directory should already
contain the output files of the partitioner, i.e. from \texttt{xdecompose\_mesh}
or \texttt{xmeshfem3D}.
\item [{\texttt{NTSTEP\_BETWEEN\_OUTPUT\_INFO}}] This parameter specifies
the interval at which basic information about a run is written to
the file system (\texttt{timestamp{*}} files in the \texttt{OUTPUT\_FILES}
directory). If you have access to a fast machine, set \texttt{NTSTEP\_BETWEEN\_OUTPUT\_INFO}
to a relatively high value (e.g., at least 100, or even 1000 or more)
to avoid writing output text files too often. This feature is not
used at the time of meshing. One can set this parameter to a larger
value than the number of time steps to avoid writing output during
the run.
\item [{\texttt{NTSTEP\_BETWEEN\_OUTPUT\_SEISMOS}}] This parameter specifies
the interval at which synthetic seismograms are written in the \texttt{LOCAL\_PATH}
directory. If a run crashes, you may still find usable (but shorter
than requested) seismograms in this directory. On a fast machine set
\texttt{NTSTEP\_BETWEEN\_OUTPUT\_SEISMOS} to a relatively high value
to avoid writing to the seismograms too often. This feature is only
relevant for the solver.
\item [{\texttt{USE\_FORCE\_POINT\_SOURCE}}] Turn this flag on to use a
(tilted) \texttt{FORCESOLUTION} force point source instead of a \texttt{CMTSOLUTION}
moment-tensor source. When the force source does not fall exactly
at a grid point, the solver interpolates the force between grid points
using Lagrange interpolants. This can be useful e.g. for oil industry
foothills simulations in which the source is a vertical force, normal
force, tilted force, or an impact etc. Note that in the \texttt{FORCESOLUTION}
file, you will need to edit the East, North and vertical components
of an arbitrary (not necessarily unitary, the code will normalize it automatically) direction vector of the force vector;
thus refer to Appendix~\ref{cha:Coordinates} for the orientation
of the reference frame. This vector is made unitary internally in
the solver and thus only its direction matters here; its norm is ignored
and the norm of the force used is the factor force source times the
source time function.
When using this option, by default the code can locate the force source
anywhere between mesh points in order to honor its exact location; this
is more precise than using the closest GLL mesh point, but it is also a bit slower.
If needed, you can change that default behavior and force the code to use the closest
GLL mesh point instead by setting flag \texttt{USE\_BEST\_LOCATION} to \texttt{.false.} instead of \texttt{.true.} in file
\texttt{src/shared/constants.h.in} and running the \texttt{configure} script again and recompiling the code.
\item [{\texttt{USE\_RICKER\_TIME\_FUNCTION}}] Turn this flag on to use
a Ricker source time function instead of the source time functions
set by default to represent a (tilted) \texttt{FORCESOLUTION} force
point source or a \texttt{CMTSOLUTION} moment-tensor source. Originally,
if a \texttt{CMTSOLUTION} moment-tensor source is used, a (pseudo)
Heaviside step function with a very short half duration is defined
for elastic cases to represent the permanent slip on the fault while
in the acoustic case a Gaussian source time function with a similarly
short half duration is defined to physically describe actions within
the fluid. Otherwise, if a \texttt{FORCESOLUTION} force source is
used, a (pseudo) Dirac delta source time function is defined by default.
Any other source-time function may then be obtained by convolution.
\item [{\texttt{PRINT\_SOURCE\_TIME\_FUNCTION}}] Turn this flag on to print
information about the source time function in the file \texttt{OUTPUT\_FILES/plot\_source\_time\_function.txt}.
This feature is only relevant for the solver.

\item [{\texttt{NUMBER\_OF\_SIMULTANEOUS\_RUNS}}] adds the ability to run several calculations (several earthquakes)
in an embarrassingly-parallel fashion from within the same run;
this can be useful when using a very large supercomputer to compute
many earthquakes in a catalog, in which case it can be better from
a batch job submission point of view to start fewer and much larger jobs,
each of them computing several earthquakes in parallel.
To turn that option on, set parameter NUMBER\_OF\_SIMULTANEOUS\_RUNS to a value greater than 1.
To implement that, we create NUMBER\_OF\_SIMULTANEOUS\_RUNS MPI sub-communicators,
each of them being labeled "my\_local\_mpi\_comm\_world", and we use them
in all the routines in "src/shared/parallel.f90", except in MPI\_ABORT() because in that case
we need to kill the entire run.
When that option is on, of course the number of processor cores used to start
the code in the batch system must be a multiple of NUMBER\_OF\_SIMULTANEOUS\_RUNS,
all the individual runs must use the same number of processor cores,
which as usual is NPROC in the Par\_file,
and thus the total number of processor cores to request from the batch system
should be NUMBER\_OF\_SIMULTANEOUS\_RUNS * NPROC.
All the runs to perform must be placed in directories called run0001, run0002, run0003 and so on
(with exactly four digits).

Imagine you have 10 independent calculations to do, each of them on 100 cores; you have three options:

1/ submit 10 jobs to the batch system

2/ submit a single job on 1000 cores to the batch, and in that script create a sub-array of jobs to start 10 jobs,
each running on 100 cores (see e.g. http://www.schedmd.com/slurmdocs/job\_array.html )

3/ submit a single job on 1000 cores to the batch, start SPECFEM3D on 1000 cores, create 10 sub-communicators,
cd into one of 10 subdirectories (called e.g. run0001, run0002,... run0010) depending on the sub-communicator
your MPI rank belongs to, and run normally on 100 cores using that sub-communicator.

The option NUMBER\_OF\_SIMULTANEOUS\_RUNS implements 3/.

\item [{\texttt{BROADCAST\_SAME\_MESH\_AND\_MODEL}}]:
if we perform simultaneous runs in parallel, if only the source and receivers vary between these runs
but not the mesh nor the model (velocity and density) then we can also read the mesh and model files
from a single run in the beginning and broadcast them to all the others; for a large number of simultaneous
runs for instance when solving inverse problems iteratively this can DRASTICALLY reduce I/Os to disk in the solver
(by a factor equal to NUMBER\_OF\_SIMULTANEOUS\_RUNS), and reducing I/Os is crucial in the case of huge runs.
Thus, always set this option to .true. if the mesh and the model are the same for all simultaneous runs.
In that case there is no need to duplicate the mesh and model file database (the content of the DATABASES\_MPI
directories) in each of the run0001, run0002,... directories, it is sufficient to have one in run0001
and the code will broadcast it to the others).

\item [{\texttt{USE\_FAILSAFE\_MECHANISM}}]:
if one or a few of these simultaneous runs fail, kill all the runs or let the others finish using a fail-safe mechanism
(in most cases, should be set to true).

TODO / future work to do: currently the BROADCAST\_SAME\_MESH\_AND\_MODEL option assumes to have the (master) mesh files in run0001/DATABASES\_MPI or run0001/OUTPUT\_FILES/DATABASES\_MPI. However, for adjoint runs you still need a DATABASES\_MPI folder in each of the sub-runs directories, e.g. run0002/DATABASES\_MPI, etc. to store the forward wavefields, kernels etc. of each sub-run. This would not be needed for forward simulations.

TODO / future work to do: the sensitivity kernel summing and smoothing tools in directory src/tomography are currently not ported to this new option to do many runs simultaneously, only the solver (src/specfem3d) is. Thus these tools should work, but in their current version will need to be run for each simulation result independently.
More precisely, the current kernel summing and smoothing routines work fine, with the exception that you need to move out the mesh files (and also the parameters). This works because these routines consider multiple runs by design. You simply have to provide them the directories where the kernels are.

\item [{\texttt{GPU\_MODE}}] Turn this flag on to use GPUs.

\item [\texttt{ADIOS\_ENABLED}] Turn this flag on to enable ADIOS. If set to \texttt{.false.}, subsequent ADIOS
parameters will not be considered.
\item [\texttt{ADIOS\_FOR\_DATABASES}] Turn this flag on to use ADIOS for xmeshfem3D output and
xgenerate\_database input.
\item [\texttt{ADIOS\_FOR\_MESH}]  Turn this flag on to use ADIOS for generated databases.
\item [\texttt{ADIOS\_FOR\_FORWARD\_ARRAYS}] Turn this flag on to read and write forward arrays using ADIOS.
\item [\texttt{ADIOS\_FOR\_KERNELS}] Turn this flag on to produce ADIOS kernels that can later be visualized with the ADIOS version of combine\_vol\_data.

\end{description}

The present version of SPECFEM can handle fully saturated porous simulations, Christina Morency implemented Biot equation.
But the code cannot calculate partially saturated cases in its state.
Christina Morency is presently working on a dual porosity, dual permeability formulation, type Pride and Berryman, for an other project, but it will not be available for some time.\\

The way we prescribe material property for porous material in SPECFEM3D is as follow:
We use a file name nummaterial\_poroelastic\_file, which is located in the directory MESH, the format is as follow:\\
 rhos rhof phi c kxx kxy kxz kyy kyz kzz Ks Kf Kfr etaf mufr\\
where\\
rho\_s = solid density\\
rho\_f = fluid density\\
phi = porosity\\
tort = tortuosity\\
kxx = xx component of permeability tensor\\
kxy = xy,yx components of permeability tensor\\
kyy = yy component of permeability tensor\\
kxz = xz,zx components of permeability tensor\\
kzz = zz component of permeability tensor\\
kappa\_s = solid bulk modulus\\
kappa\_f= fluid bulk modulus\\
kappa\_fr= frame bulk modulus\\
eta\_f = fluid viscosity\\
mu\_fr = frame shear modulus.\\

Using an external mesh (for instance coming from CUBIT/TRELIS), poroelastic materials have the ID number 3, while 1 is elastic, 2 is acoustic ? see the example in the package: examples/homogeneous\_poroelastic.

If you use PML, the mesh elements that belong to the PML layers can
be acoustic or elastic, but not viscoelastic nor poroelastic. Then,
when defining your model, you should define these absorbing elements
as either acoustic or elastic. In you forget to do that, the code
will fix the problem by automatically converting the viscoelastic
or poroelastic PML elements to elastic. This means that strictly speaking
the PML layer will not be perfectly matched any more, since the physical
model will change from viscoelastic or poroelastic to elastic at the
entrance of the PML, but in practice this is sufficient and produces
only tiny / negligible spurious reflections.

If you use PML and an external mesh (created using an external meshing tool
such as CUBIT/TRELIS or similar), try to have elements inside the PML as regular as possible,
i.e. ideally non-deformed cubes obtained by `extrusion' of regular surface mesh elements meshing the
outer edges of the computational domain without PML; by doing so, the PMLs obtained will be far more stable
in time (PML being weakly unstable from a mathematical point of view, very deformed mesh elements
inside the PMLs can trigger instabilities much more quickly).

If you use PML and an external tomographic velocity and density model,
you should be careful because mathematically a PML cannot handle heterogeneities
along the normal to the PML edge inside the PML layer. This comes
from the fact that the damping profile that is defined assumes a constant
velocity and density model along the normal direction.

Thus, you need to modify your velocity and density model in order
for it to be 1D inside the PML, as shown in Figure~\ref{fig:modify_external_velocity_model_to_use_PML}.

This applies to the bottom layer as well; there you should make sure
that your model is 1D and thus constant along the vertical direction.

To summarize, only use a 3D velocity and density model inside the
physical region, and in all the PML layers extend it by continuity
from its values along the inner PML edge.

%%
\begin{figure}[htbp]
\noindent \begin{centering}
\includegraphics[width=3in]{figures/how_to_use_PML_when_a_tomographic_velocity_model_is_used.png}
\par\end{centering}

\caption{How to modify your external 3D velocity and density model in order
to use PML. Such a modification is not needed when using Stacey absorbing
boundary conditions (but such conditions are significantly less efficient).}

\label{fig:modify_external_velocity_model_to_use_PML}
\end{figure}

%%

\section{Choosing the time step \texttt{DT}}\label{sec:Choosing-the-Time-Step}

The parameter \texttt{DT} sets the length of each time step in seconds.
The value of this parameter is crucial for the stability of the spectral-element
simulation. Your time step \texttt{DT} will depend on the minimum
ratio between the distance $h$ of neighboring mesh points and the
wave speeds $v$ defined in your model. The condition for the time
step $\Delta t$ is:
\begin{lyxcode}
$\Delta t<C~\mathrm{min}_{\Omega}(~h/v~)$
\end{lyxcode}
where $C$ is the so-called Courant number and $\Omega$ denotes the
model volume. The distance $h$ depends on the mesh element size and
the number of GLL points \texttt{NGLL} specified in the main constants
file \texttt{constants.h} located in the \texttt{src/shared/} subdirectory.
The wave speed $v$ is determined based on your model's P- (or S-)
wave speed values.

The database generator \texttt{xgenerate\_databases}, as well as the
internal mesher \texttt{xmeshfem3D}, are trying to evaluate the value
of $\Delta t$ for empirically chosen Courant numbers $C\sim0.3$.
If you used the mesher \texttt{xmeshfem3D} to generate your mesh,
you should set the value suggested in \texttt{OUTPUT\_FILES/output\_mesher.txt}
file, which is created after the mesher completed. In case you used
CUBIT to create the mesh, you might use an arbitrary value when running
\texttt{xgenerate\_databases} and then use the value suggested in
the ~\\
 \texttt{OUTPUT\_FILES/output\_mesher.txt} file after the database
generation completed. Note that the implemented Newmark time scheme
uses this time step globally, thus your simulations become more expensive
for very small mesh elements in high wave-speed regions. Please be
aware of this restriction when constructing your mesh in Chapter~\ref{cha:Mesh-Generation}.

### 2015-04-24 doc/USER_MANUAL/05_running_the_solver.tex

--- 2012-09-28 doc/USER_MANUAL/manual_SPECFEM3D.tex
@@
-%% LyX 1.5.1 created this file.  For more info, see http://www.lyx.org/.
-%% Do not edit unless you really know what you are doing.
-\documentclass[oneside,english,onecolumn]{book}
-\usepackage[T1]{fontenc}
-\usepackage[latin1]{inputenc}
-\usepackage{geometry}
-\geometry{letterpaper,tmargin=1in,bmargin=1in,lmargin=1in,rmargin=1in}
-\setcounter{secnumdepth}{3}
-\setcounter{tocdepth}{3}
-\usepackage{float}
-\usepackage{textcomp}
-\usepackage{amsmath}
-
-% figures
-\usepackage[pdftex]{graphicx}
-
-% we are running pdflatex, so convert .eps files to .pdf
-\usepackage{epstopdf}
-
-\usepackage{amssymb}
-\IfFileExists{url.sty}{\usepackage{url}}
-                      {\newcommand{\url}{\texttt}}
-\usepackage[authoryear]{natbib}
-
-\makeatletter
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% LyX specific LaTeX commands.
-\newcommand{\noun}[1]{\textsc{#1}}
-%% A simple dot to overcome graphicx limitations
-\newcommand{\lyxdot}{.}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Textclass specific LaTeX commands.
-\newenvironment{lyxcode}
-{\begin{list}{}{
-\setlength{\rightmargin}{\leftmargin}
-\setlength{\listparindent}{0pt}% needed for AMS classes
-\raggedright
-\setlength{\itemsep}{0pt}
-\setlength{\parsep}{0pt}
-\normalfont\ttfamily}%
- \item[]}
-{\end{list}}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% User specified LaTeX commands.
-%\renewcommand{\baselinestretch}{1.5}
-
-% figures
-\usepackage[dvips]{epsfig}
-
-\usepackage{wrapfig}
-
-% fonts
-\usepackage{times}
-
-% hyperlinks to sections and references
-\usepackage[pdftex,bookmarks=true,bookmarksnumbered=true,pdfpagemode=None,pdfstartview=FitH,pdfpagelayout=SinglePage,pdfborder={0 0 0}]{hyperref}
-
-\newcommand{\urlwithparentheses}[1]{(\url{#1})}
-
-% biblio GJI
-\bibliographystyle{abbrvnat}
-
-\newcommand{\toall}[1]{\textbf{*** All: #1 ***}}
-\newcommand{\tojeroen}[1]{\textbf{*** Jeroen: #1 ***}}
-\newcommand{\tobrian}[1]{\textbf{*** Brian: #1 ***}}
-\newcommand{\tovala}[1]{\textbf{*** Vala: #1 ***}}
-\newcommand{\tovalabrian}[1]{\textbf{*** Vala \& Brian: #1 ***}}
-\newcommand{\tovalaqinya}[1]{\textbf{*** Vala \& Qinya: #1 ***}}
-\newcommand{\toqinya}[1]{\textbf{*** Qinya: #1 ***}}
-\newcommand{\tomin}[1]{\textbf{*** Min: #1 ***}}
-\newcommand{\toalessia}[1]{\textbf{*** Alessia: #1 ***}}
-\newcommand{\todimitri}[1]{\textbf{*** Dimitri: #1 ***}}
-
-\newcommand{\nexxi}{\mbox{\texttt{NEX\_XI\/}}}
-\newcommand{\nexeta}{\mbox{\texttt{NEX\_ETA\/}}}
-\newcommand{\nprocxi}{\mbox{\texttt{NPROC\_XI\/}}}
-\newcommand{\nproceta}{\mbox{\texttt{NPROC\_ETA\/}}}
-\newcommand{\nchunks}{\mbox{\texttt{NCHUNKS\/}}}
-\newcommand{\eg}{e.g.,}
-
-% colors for the text
-\usepackage[dvipsnames,usenames]{color}
-\newcommand{\red}[1]{\textcolor{Red}{#1}}
-
-\usepackage{babel}
-\makeatother
-
-\begin{document}
-\begin{center}
-\thispagestyle{empty}%
-\vspace*{-1.8truecm}
-\noindent\makebox[\textwidth]{%
-\noindent\includegraphics[width=0.83\paperwidth]{figures/specfem_3d-cover.pdf}
-}
-\end{center}
-%
-\title{\textbf{SPECFEM3D}\\
-\textbf{User Manual}}
-
-\author{$\copyright$ Princeton University (USA) and\\
-CNRS / INRIA / University of Pau (France)\\
-Version 2.1}
-
-\date{\noindent \today}
-
-\maketitle
-
-\section*{Authors}
-The SPECFEM3D package was first developed by Dimitri Komatitsch and Jeroen Tromp at Harvard University and Caltech, USA, starting in 1998, based on earlier work by Dimitri Komatitsch and Jean-Pierre Vilotte at Institut de Physique du Globe (IPGP) in Paris, France from 1995 to 1997.\\
-
-Since then it has been developed and maintained by a development team: in alphabetical order,
-Piero Basini,
-C\'eline Blitz,
-Ebru Bozda\u{g},
-Emanuele Casarotti,
-Joseph Charles,
-Min Chen,
-Vala Hj\"orleifsd\'ottir,
-Sue Kientz,
-Dimitri Komatitsch,
-Jes\'us Labarta,
-Nicolas Le Goff,
-Pieyre Le Loher,
-Qinya Liu,
-Yang Luo,
-Alessia Maggi,
-Federica Magnoni,
-Roland Martin,
-Dennis McRitchie,
-Matthias Meschede,
-Peter Messmer,
-David Mich\'ea,
-Tarje Nissen-Meyer,
-Daniel Peter,
-Max Rietmann,
-Brian Savage,
-Bernhard Schuberth,
-Anne Sieminski,
-Leif Strand,
-Carl Tape,
-Jeroen Tromp,
-Zhinan Xie,
-Hejun Zhu.\\
-%\toall{add full list of developers here and update the AUTHORS file accordingly}
-
-The cover graphic of the manual was created
-by Santiago Lombeyda from Caltech's Center for Advanced Computing Research (CACR) \urlwithparentheses{http://www.cacr.caltech.edu}. \\
-
-\newpage{}
-
-\tableofcontents{}
-
-\chapter{Introduction}
-
-The software package SPECFEM3D simulates seismic
-wave propagation at the local or regional scale
-based upon the spectral-element method (SEM).
-The SEM is a continuous Galerkin technique \citep{TrKoLi08,PeKoLuMaLeCaLeMaLiBlNiBaTr11},
-which can easily be made discontinous \citep{BeMaPa94,Ch00,KoWoHu02,ChCaVi03,LaWaBe05,Kop06,WiStBuGh10,AcKo11};
-it is then close to a particular case of the discontinuous Galerkin technique \citep{ReHi73,Arn82,FaRi99,HuHuRa99,CoKaSh00,GiHeWa02,RiWh03,MoRi05,GrScSc06,AiMoMu06,BeLaPi06,DuKa06,DeSeWh08,PuAmKa09,WiStBuGh10,DeSe10,EtChViGl10}, with optimized efficiency because of its tensorized basis functions \citep{WiStBuGh10,AcKo11}.
-In particular, it can accurately handle very distorted mesh elements \citep{OlSe11}.\\
-
-It has very good accuracy and convergence properties \citep{MaPa89,SePr94,DeFiMu02,Coh02,DeSe07,SeOl08}.
-The spectral element approach admits spectral rates of convergence and allows exploiting $hp$-convergence schemes.
-It is also very well suited to parallel implementation on very large supercomputers \citep{KoTsChTr03,TsKoChTr03,KoLaMi08a,CaKoLaTiMiLeSnTr08,KoViCh10} as well as on clusters of GPU accelerating graphics cards \citep{KoMiEr09,KoErGoMi10,Kom11}.
-Tensor products inside each element can be optimized to reach very high efficiency \citep{DeFiMu02}, and mesh point and element numbering can be optimized to reduce processor cache misses and improve cache reuse \citep{KoLaMi08a}.
-The SEM can also handle triangular (in 2D) or tetrahedral (in 3D) elements \citep{WinBoyd96,TaWi00,KoMaTrTaWi01,Coh02,MeViSa06} as well as mixed meshes, although with increased cost and reduced accuracy in these elements, as in the discontinuous Galerkin method.\\
-
-Note that in many geological models in the context of seismic wave propagation studies
-(except for instance for fault dynamic rupture studies, in which very high frequencies or supershear rupture need to be modeled near the fault, see e.g.
-\cite{BeGlCrViPi07,BeGlCrVi09,PuAmKa09,TaCrEtViBeSa10})
-a continuous formulation is sufficient because material property contrasts are not drastic and thus
-conforming mesh doubling bricks can efficiently handle mesh size variations \citep{KoTr02a,KoLiTrSuStSh04,LeChLiKoHuTr08,LeChKoHuTr09,LeKoHuTr09}.\\
-
-For a detailed introduction to the SEM as applied to regional
-seismic wave propagation, please consult \citet{PeKoLuMaLeCaLeMaLiBlNiBaTr11,TrKoLi08,KoVi98,KoTr99,ChKoViCaVaFe07} and
-in particular \citet{LeKoHuTr09,LeChKoHuTr09,LeChLiKoHuTr08,GoAmTaCaSmSaMaKo09,WiKoScTr04,KoLiTrSuStSh04}.
-A detailed theoretical analysis of the dispersion
-and stability properties of the SEM is available in \citet{Coh02}, \citet{DeSe07}, \citet{SeOl07} and \citet{SeOl08}.\\
-
-Effects due to lateral variations in compressional-wave speed, shear-wave
-speed, density, a 3D crustal model, topography and bathymetry are
-included. The package can accommodate
-full 21-parameter anisotropy (see~\citet{ChTr07}) as well as lateral
-variations in attenuation \citep{SaKoTr10}. Adjoint capabilities and finite-frequency
-kernel simulations are included \citep{TrKoLi08,PeKoLuMaLeCaLeMaLiBlNiBaTr11,LiTr06,FiIgBuKe09,ViOp09}.\\
-
-The SEM was originally developed in computational fluid dynamics \citep{Pat84,MaPa89}
-and has been successfully adapted to address problems in seismic wave propagation.
-Early seismic wave propagation applications of the SEM, utilizing Legendre basis functions and a
-perfectly diagonal mass matrix, include \cite{CoJoTo93}, \cite{Kom97},
-\cite{FaMaPaQu97}, \cite{CaGa97}, \cite{KoVi98} and \cite{KoTr99},
-whereas applications involving Chebyshev basis functions and a nondiagonal mass matrix
-include \cite{SePr94}, \cite{PrCaSe94} and \cite{SePrPr95}.\\
-
-All SPECFEM3D software is written in Fortran90 with full portability
-in mind, and conforms strictly to the Fortran95 standard. It uses
-no obsolete or obsolescent features of Fortran77. The package uses
-parallel programming based upon the Message Passing Interface (MPI)
-\citep{GrLuSk94,Pac97}.\\
-
-SPECFEM3D won the Gordon Bell award for best performance at the SuperComputing~2003
-conference in Phoenix, Arizona (USA) (see \cite{KoTsChTr03}
-and \url{www.sc-conference.org/sc2003/nr_finalaward.html}).
-It was a finalist again in 2008 for a run at 0.16 petaflops (sustained) on 149,784 processors of the `Jaguar' Cray XT5 system at Oak Ridge National Laboratories (USA) \citep{CaKoLaTiMiLeSnTr08}.
-It also won the BULL Joseph Fourier supercomputing award in 2010.\\
-
-This new release of the code (version 2.1) includes support for GPU graphics card acceleration \citep{KoMiEr09,KoGoErMi10,KoErGoMi10,MiKo10,Kom11}.\\
-
-The next release of the code will include support Convolution or Auxiliary Differential Equation Perfectly Matched absorbing Layers
-(C-PML or ADE-PML) \citep{KoMa07,MaKoEz08,MaKoGe08,MaKo09,MaKoGeBr10}.
-It will also use the PT-SCOTCH parallel library for mesh partitioning.\\
-
-SPECFEM3D includes coupled fluid-solid domains and
-adjoint capabilities, which enables one to address seismological inverse problems, but for linear rheologies only so far.
-To accommodate visco-plastic or non-linear rheologies, the reader can refer to
-the \href{http://geoelse.stru.polimi.it/}{GeoELSE} software package \citep{CaGa97,StPaIg09,ChMoTsBaKrKaStKr10}.
-
-\section{Citation}
-
-If you use SPECFEM3D for your own research, please cite at least one
-of the following articles:
-
-\begin{description}
-
-\item [Numerical simulations in general] ~\\
-Forward and adjoint simulations are described in detail in \cite{TrKoLi08,PeKoLuMaLeCaLeMaLiBlNiBaTr11,VaCaSaKoVi99,KoMiEr09,KoErGoMi10,KoGoErMi10,ChKoViCaVaFe07,MaKoDi09,KoViCh10,CaKoLaTiMiLeSnTr08,TrKoHjLiZhPeBoMcFrTrHu10,KoRiTr02,KoTr02a,KoTr02b,KoTr99} or \cite{KoVi98}.
-Additional aspects dealing with adjoint simulations are described in
-\cite{TrTaLi05,LiTr06,TrKoLi08,LiTr08,TrKoHjLiZhPeBoMcFrTrHu10,PeKoLuMaLeCaLeMaLiBlNiBaTr11}.
-Domain decomposition is explained in detail in \cite{MaKoBlLe08}, and excellent scaling up to 150,000 processor cores is shown for instance in
-\cite{CaKoLaTiMiLeSnTr08,KoLaMi08a,MaKoBlLe08,KoErGoMi10,KoGoErMi10,Kom11},
-
-\item [GPU computing] ~\\
-Computing on GPU graphics cards for acoustic or seismic wave propagation applications
-is described in detail in \cite{KoMiEr09,KoGoErMi10,KoErGoMi10,MiKo10,Kom11}.
-
-\end{description}
+\chapter{Running the Solver \texttt{xspecfem3D}}\label{cha:Running-the-Solver}
+
+Now that you have successfully generated the databases, you are ready
+to compile the solver. In the main directory, type
+\begin{lyxcode}
+{\small make~xspecfem3D}{\small \par}
+\end{lyxcode}
+Please note that \texttt{xspecfem3D} must be called directly from
+the main directory, as most of the binaries of the package.\\
 
 \noindent
-If you use this new version 2.1, which has non blocking MPI for much better performance for medium or large runs, please cite at least one of these six articles,
-in which results of non blocking MPI runs are presented: \cite{PeKoLuMaLeCaLeMaLiBlNiBaTr11,KoGoErMi10,KoErGoMi10,KoViCh10,Kom11,CaKoLaTiMiLeSnTr08,MaKoBlLe08}.\\
-
-\noindent
-If you work on geophysical applications, you may be interested in citing some of these application articles as well, among others:
-
-\begin{description}
-
-\item [Southern California simulations] ~\\
-\cite{KoLiTrSuStSh04,KrJiKoTr06a,KrJiKoTr06b}.
-
-If you use the 3D southern California model, please cite
-\citet{SuSh03} (Los Angeles model),
-\citet{lovelyetal06} (Salton Trough), and
-\citet{hauksson2000} (southern California).
-The Moho map was determined by
-\citet{zhu&kanamori2000}.
-The 1D SoCal model was developed by \citet{DrHe90}.
-
-\item [Anisotropy] ~\\
-\cite{ChTr07,JiTsKoTr05,ChFaKo04,FaChKo04,RiRiKoTrHe02,TrKo00}.
-
-\item [Attenuation] ~\\
-\cite{SaKoTr10,KoTr02a,KoTr99}.
-
-\item [Topography] ~\\
-\cite{LeKoHuTr09,LeChKoHuTr09,LeChLiKoHuTr08,GoAmTaCaSmSaMaKo09,WiKoScTr04}.
-
-\end{description}
-
-The corresponding Bib\TeX{} entries may be found
-in file \texttt{doc/USER\_MANUAL/bibliography.bib}.
-
-\section{Support}
-
-This material is based upon work supported by the USA National Science
-Foundation under Grants No. EAR-0406751 and EAR-0711177, by the French
-CNRS, French INRIA Sud-Ouest MAGIQUE-3D, French ANR NUMASIS under
-Grant No. ANR-05-CIGC-002, and European FP6 Marie Curie International
-Reintegration Grant No. MIRG-CT-2005-017461.
-Any opinions, findings, and conclusions or recommendations expressed in this material are
-those of the authors and do not necessarily reflect the views of the
-USA National Science Foundation, CNRS, INRIA, ANR or the European
-Marie Curie program.
-
-\chapter{\label{cha:Getting-Started}Getting Started}
-
-The SPECFEM3D software package comes in a gzipped tar ball. In the
-directory in which you want to install the package, type
-\begin{lyxcode}
-{\small tar~-zxvf~SPECFEM3D\_V2.1.0.tar.gz}{\small \par}
-\end{lyxcode}
-The directory \texttt{\small SPECFEM3D/} will then contain
-the source code.\\
-
-We recommend that you add {\texttt{ulimit -S -s unlimited}} to your {\texttt{.bash\_profile}} file and/or {\texttt{limit stacksize unlimited }} to your {\texttt{.cshrc}} file to suppress any potential limit to the size of the Unix stack.\\
-
-Then, to configure the software for your system, run the
-\texttt{configure} shell script. This script will attempt to guess
-the appropriate configuration values for your system. However, at
-a minimum, it is recommended that you explicitly specify the appropriate
-command names for your Fortran90 compiler and MPI package:
-
-\begin{lyxcode}
-./configure~FC=ifort~MPIFC=mpif90
-\end{lyxcode}
-
-Before running the \texttt{configure} script, you should probably edit file \texttt{flags.guess} to make sure that it contains the best compiler options for your system.
-Known issues or things to check are:
-
-\begin{description}
-\item [{\texttt{Intel ifort compiler}}] See if you need to add \texttt{-assume byterecl} for your machine.
-\item [{\texttt{IBM compiler}}] See if you need to add \texttt{-qsave} or \texttt{-qnosave} for your machine.
-\item [{\texttt{Mac OS}}] You will probably need to install \texttt{XCODE}.
-In addition, the \texttt{clock\_gettime} routine, which is used by the \texttt{SCOTCH} library that we use, does not exist in Mac OS.
-You will need to replace it with \texttt{clock\_get\_time} if you want to use \texttt{SCOTCH}.
-\end{description}
-
-When compiling on an IBM machine with the \texttt{xlf} and \texttt{xlc} compilers, we suggest running the \texttt{configure} script
-with the following options:\\
-
-\noindent\texttt{./configure FC=xlf90\_r MPIFC=mpif90 CC=xlc\_r CFLAGS="-O3 -q64" FCFLAGS="-O3 -q64" -with-scotch-dir=...}.\\
-
-On SGI systems, \texttt{flags.guess} automatically informs \texttt{configure}
-to insert `\texttt{`TRAP\_FPE=OFF}'' into the generated \texttt{Makefile}
-in order to turn underflow trapping off.\\
-
-If you run very large meshes on a relatively small number
-of processors, the memory size needed on each processor might become
-greater than 2 gigabytes, which is the upper limit for 32-bit addressing.
-In this case, on some compilers you may need to add \texttt{``-mcmodel=medium}'' or \texttt{``-mcmodel=medium -shared-intel}''
-to the configure options of CFLAGS, FCFLAGS and LDFLAGS otherwise the compiler will display an error
-message (for instance \texttt{``relocation truncated to fit: R\_X86\_64\_PC32 against .bss''} or something similar);
-on an IBM machine with the \texttt{xlf} and \texttt{xlc} compilers, using \texttt{-q64} is usually sufficient.\\
-
-The SPECFEM3D software package relies on the SCOTCH library to partition meshes created with CUBIT.
-Note that we use CUBIT to create meshes of hexahedra, but other packages can be used as well, for instance GiD from \url{http://gid.cimne.upc.es} or Gmsh from \url{http://geuz.org/gmsh} \citep{GeRe09}.
-Even mesh creation packages that generate tetrahedra, for instance TetGen from \url{http://tetgen.berlios.de}, can be used because each tetrahedron can then easily be decomposed into four hexahedra as shown in the picture of the TetGen logo at \url{http://tetgen.berlios.de/figs/Delaunay-Voronoi-3D.gif}; while this approach does not generate hexahedra of optimal quality, it can ease mesh creation in some situations and it has been shown that the spectral-element method can very accurately handle distorted mesh elements \citep{OlSe11}.
-
-The SCOTCH library \citep{PeRo96}
-provides efficent static mapping, graph and mesh partitioning routines. SCOTCH is a free software package developed by
-Fran\c {c}ois Pellegrini et al.
-from LaBRI and INRIA in Bordeaux, France, downloadable from the web page \url{https://gforge.inria.fr/projects/scotch/}.
-It is more recent than METIS, actively maintained and performs better in many cases.
-A recent version of its source code is provided in directory \texttt{src/decompose\_mesh\_SCOTCH/scotch\_5.1.10b}.
-In case no SCOTCH libraries can be found on the system, the configuration will bundle this version for compilation.
-The path to an existing SCOTCH installation can to be set explicitly with the option \texttt{-{}-with-scotch-dir}. Just as an example:
-\begin{lyxcode}
-./configure~FC=ifort~MPIFC=mpif90~-{}-with-scotch-dir=/opt/scotch
-\end{lyxcode}
-
-If you use the Intel ifort compiler to compile the code, we recommend that you use the Intel icc C compiler to compile Scotch, i.e., use:
-\begin{lyxcode}
-./configure~CC=icc~FC=ifort~MPIFC=mpif90
-\end{lyxcode}
-
-To compile a serial version of the code for small meshes that fits
-on one compute node and can therefore be run serially, run \texttt{configure}
-with the \texttt{-{}-without-mpi} option to suppress all calls to
-MPI.
-
-A summary of the most important configuration variables follows.
-
-\begin{description}
-\item [{\texttt{F90}}] Path to the Fortran90 compiler.
-\item [{\texttt{MPIF90}}] Path to MPI Fortran90.
-\item [{\texttt{MPI\_FLAGS}}] Some systems require this flag to link to
-MPI libraries.
-\item [{\texttt{FLAGS\_CHECK}}] Compiler flag for non-critical subroutines.
-\item [{\texttt{FLAGS\_NO\_CHECK}}] Compiler flag for creating fast, production-run
-code for critical subroutines.
-\end{description}
-The configuration script automatically creates for each executable a corresponding
-\texttt{Makefile} in the \texttt{src/} subdirectoy.
-The \texttt{Makefile} contains a number of suggested entries for various
-compilers, e.g., Portland, Intel, Absoft, NAG, and Lahey. The software
-has run on a wide variety of compute platforms, e.g., various PC clusters
-and machines from Sun, SGI, IBM, Compaq, and NEC. Select the compiler
-you wish to use on your system and choose the related optimization
-flags. Note that the default flags in the \texttt{Makefile} are undoubtedly
-not optimal for your system, so we encourage you to experiment with
-these flags and to solicit advice from your systems administrator.
-Selecting the right compiler and optimization flags can make a tremendous
-difference in terms of performance. We welcome feedback on your experience
-with various compilers and flags.
-
-Now that you have set the compiler information, you need to select
-a number of flags in the \texttt{constants.h} file depending on your
-system:
-
-\begin{description}
-\item [{\texttt{LOCAL\_PATH\_IS\_ALSO\_GLOBAL}}] Set to \texttt{.false.}
-on most cluster applications. For reasons of speed, the (parallel)
-distributed database generator typically writes a (parallel) database for the solver on the
-local disks of the compute nodes. Some systems have no local disks,
-e.g., BlueGene or the Earth Simulator, and other systems have a fast
-parallel file system, in which case this flag should be set to \texttt{.true.}.
-Note that this flag is not used by the database generator or the solver; it is
-only used for some of the post-processing.
-\end{description}
-The package can run either
-in single or in double precision mode. The default is single precision
-because for almost all calculations performed using the spectral-element method
-using single precision is sufficient and gives the same results (i.e. the same seismograms);
-and the single precision code is faster and requires exactly half as much memory.
-Select your preference by selecting the appropriate setting in the \texttt{constants.h} file:
-
-\begin{description}
-\item [{\texttt{CUSTOM\_REAL}}] Set to \texttt{SIZE\_REAL} for single precision
-and \texttt{SIZE\_DOUBLE} for double precision.
-\end{description}
-In the \texttt{precision.h} file:
-
-\begin{description}
-\item [{\texttt{CUSTOM\_MPI\_TYPE}}] Set to \texttt{MPI\_REAL} for single
-precision and \texttt{MPI\_DOUBLE\_PRECISION} for double precision.
-\end{description}
-On many current processors (e.g., Intel, AMD, IBM Power), single precision calculations
-are significantly faster; the difference can typically be 10\%
-to 25\%. It is therefore better to use single precision.
-What you can do once for the physical problem you want to study is run the same calculation in single precision
-and in double precision on your system and compare the seismograms.
-If they are identical (and in most cases they will), you can select single precision for your future runs.\\
-
-\section{Adding OpenMP support in addition to MPI}
-
-OpenMP support can be enabled in addition to MPI. However, in many cases performance will not improve
-because our pure MPI implementation is already heavily optimized and thus the resulting code will in fact
-be slightly slower. A possible exception could be IBM BlueGene-type architectures.
-
-To enable OpenMP, uncomment the OpenMP compiler option in two lines in file \texttt{src/specfem3D/Makefile.in}
-(before running \texttt{configure}) and also uncomment the \texttt{\#define USE\_OPENMP} statement in file\\
-\texttt{src/specfem3D/specfem3D.F90}.
-
-The DO-loop using OpenMP threads has a SCHEDULE property. The OMP\_SCHEDULE
-environment variable can set the scheduling policy of that DO-loop.
-Tests performed by Marcin Zielinski at SARA (The Netherlands) showed that often
-the best scheduling policy is DYNAMIC with the size of the chunk equal to the number of
-OpenMP threads, but most preferably being twice as the number of
-OpenMP threads (thus chunk size = 8 for 4 OpenMP threads etc).
-If OMP\_SCHEDULE is not set or is empty, the DO-loop will assume generic
-scheduling policy, which will slow down the job quite a bit.
-
-\section{Visualizing the subroutine calling tree of the source code}
-
-Packages such as \texttt{doxywizard} can be used to visualize the subroutine calling tree of the source code.
-\texttt{Doxywizard} is a GUI front-end for configuring and running \texttt{doxygen}.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\chapter{\label{cha:Mesh-Generation}Mesh Generation}
-
-The first step in running a spectral-element simulation consists of constructing a high-quality mesh for the region under consideration.
-We provide two possibilities to do so: (1) relying on the external, hexahedral mesher CUBIT, or (2) using the provided, internal mesher \texttt{xmeshfem3D}.
-In the following, we explain these two approaches.
-
-\section{\label{cha:Running-the-Mesher-CUBIT}Meshing with \texttt{CUBIT}}
-
-CUBIT is a meshing tool suite for the creation of finite-element meshes for arbitrarily shaped models.
-It has been developed and maintained at Sandia National Laboratories and can be purchased for a small academic institutional fee at \url{http://cubit.sandia.gov}.
-Our experience showed that using CUBIT greatly facilitates and speeds up the generation and preparation of hexahedral, conforming meshes for a variety of geophysical models with increasing complexity.
-
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=.6\textwidth]{figures/mount-cubit.jpg}
-\par\end{centering}
-
-\caption{Example of the graphical user interface of CUBIT. The hexahedral mesh shown in the main display consists of a hexahedral discretization of a single volume with topography.}
-
-\label{fig:mount.cubit}
-\end{figure}
-
-The basic steps in creating a load-balanced, partitioned mesh with CUBIT are:
-\begin{description}
-\item [1.] setting up a hexahedral mesh with CUBIT,
-\item [2.] exporting the CUBIT mesh into a SPECFEM3D file format and
-\item [3.] partitioning the SPECFEM3D mesh files for a chosen number of cores.
-\end{description}
-
-Examples are provided in the SPECFEM3D package in the subdirectory \texttt{examples/}.
-We strongly encourage you to contribute your own example to this package by contacting the CIG Computational Seismology Mailing List  \urlwithparentheses{cig-seismo@geodynamics.org}.
-
-\subsection{Creating the Mesh with CUBIT}
-
-For the installation and handling of the CUBIT meshing tool suite, please refer to the CUBIT user manual and documentation.
-In order to give you a basic understanding of how to use CUBIT for our purposes, examples are provided in the SPECFEM3D package in the subdirectory \texttt{examples/}:
-\begin{description}
-\item [{\texttt{homogeneous\_halfspace}}] Creates a single block model and assigns elastic material parameters.
-\item [{\texttt{layered\_halfspace}}]  Combines two different, elastic material volumes and creates a refinement layer between the two.
-This example can be compared for validation against the solutions provided in subdirectory ~\\
-\texttt{VALIDATION\_3D\_SEM\_SIMPLER\_LAYER\_SOURCE\_DEPTH/}.
-\item [{\texttt{waterlayered\_halfspace}}] Combines an acoustic and elastic material volume as in a schematic marine survey example.
-\item [{\texttt{tomographic\_model}}] Creates a single block model whose material properties will have to be read in from a tomographic model file during the databases creation by \texttt{xgenerate\_databases}.
-\end{description}
-
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=.45\textwidth]{figures/example-homogeneous.jpg}
-\includegraphics[width=.45\textwidth]{figures/example-2layers.jpg} \\
-\includegraphics[width=.45\textwidth]{figures/example-water.jpg}
-\includegraphics[width=.45\textwidth]{figures/example-tomo.jpg}
-\par\end{centering}
-
-\caption{Screenshots of the CUBIT examples provided in subdirectory  \texttt{examples/}: homogeneous halfspace (top-left), layered halfspace (top-right), water layered halfspace (bottom-left) and tomographic model (bottom-right).}
-
-\label{fig:examples.cubit}
-\end{figure}
-
-In each example subdirectory you will find a \texttt{README} file, which explains in a step-by-step tutorial the workflow for the example.
-Please feel free to contribute your own example to this package by contacting the CIG Computational Seismology Mailing List  \urlwithparentheses{cig-seismo@geodynamics.org}.
-
-\subsection{Exporting the Mesh with \texttt{cubit2specfem3d.py} }
-\label{subsec:Exporting-the-Mesh}
-
-Once the geometric model volumes in CUBIT are meshed, you prepare the model for exportation with the definition of material blocks and boundary surfaces.
-Thus, prior to exporting the mesh, you need to define blocks specifying the materials and absorbing boundaries in CUBIT.
-This process could be done automatically using the script \texttt{boundary\_definition.py} if the mesh meets some conditions or manually, following the block convention:
-
-\begin{description}
-\item [material\_name] Each material should have a specific block defined by a unique name.
-The name convention of the material is to start with either 'elastic' or 'acoustic'.
-It must be then followed by a unique identifier, e.g.
-'elastic 1', 'elastic 2', etc.
-The additional attributes to the block define the material description.
-
-For an elastic material:
-\begin{description}
-\item [material\_id] An integer value which is unique for this material.
-\item [Vp] P-wave speed of the material (given in m/s).
-\item [Vs] S-wave speed of the material (given in m/s).
-\item [rho] density of the material (given in kg/m$^3$).
-\item [Q] quality factor to use in case of a simulation with attenuation turned on.
-It should be between 1 and 9000.
-In case no attenuation information is available, it can be set to zero.
-Please note that your Vp- and Vs-speeds are given for a reference frequency.
-To change this reference frequency, you change the value of
-\texttt{ATTENUATION\_f0\_REFERENCE} in the main constants file \texttt{constants.h} found in subdirectory \texttt{src/shared/}.
-\item [anisotropic\_flag] Flag describing the anisotropic model to use in case an anisotropic simulation should be conducted.
-See the file \texttt{model\_aniso.f90} in subdirectory \texttt{src/generate\_databases/} for an implementation of the anisotropic models.
-In case no anisotropy is available, it can be set to zero.
-\end{description}
-Note that this material block has to be defined using all the volumes which belong to this elastic material.
-For volumes belonging to another, different material, you will need to define a new material block.
-
-For an acoustic material:
-\begin{description}
-\item [material\_id] An integer value which is unique for this material.
-\item [Vp] P-wave speed of the material (given in m/s).
-\item [0] S-wave speed of the material is ignored.
-\item [rho] density of the material (given in kg/m$^3$).
-\end{description}
-
-\item [face\_topo] Block definition for the surface which defines the free surface (which can have topography).
-The name of this block must be 'face\_topo', the block has to be defined using all the surfaces which constitute the complete free surface of the model.
-
-\item [face\_abs\_xmin] Block definition for the faces on the absorbing boundaries, one block for each surface with x=Xmin.
-\item [face\_abs\_xmax] Block definition for the faces on the absorbing boundaries, one block for each surface with x=Xmax.
-\item [face\_abs\_ymin] Block definition for the faces on the absorbing boundaries, one block for each surface with y=Ymin.
-\item [face\_abs\_ymax] Block definition for the faces on the absorbing boundaries, one block for each surface with y=Ymax.
-\item [face\_abs\_bottom] Block definition for the faces on the absorbing boundaries, one block for each surface with z=bottom.
-\end{description}
-
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=.45\textwidth]{figures/mount-surface.jpg}
-\includegraphics[width=.45\textwidth]{figures/mount-abs.jpg}
-\par\end{centering}
-
-\caption{Example of the block definitions for the free surface 'face\_topo' (left) and the absorbing boundaries, defined in a single block 'face\_abs\_xmin' (right) in CUBIT.}
-
-\label{fig:mount.abs}
-\end{figure}
-
-Optionally, instead of specifying for each surface at the boundaries a single block like mentioned above, you can also specify a single block for all boundary surfaces and name it as one of the absorbing blocks above, e.g.
-'face\_abs\_xmin'.
-
-After the block definitions are done, you export the mesh using the script \texttt{cubit2specfem3d.py} provided in each of the example directories.
-If the export was successful, you should find the following files in a subdirectory \texttt{MESH/}:
-
-\begin{description}
-\item [nummaterial\_velocity\_file] Defines the material properties. For fully defined materials, the formats is:
-\begin{lyxcode}
-\texttt{domain\_ID} \texttt{material\_ID}  \texttt{rho}  \texttt{vp}  \texttt{vs}  \texttt{Q}  \texttt{anisotropy\_flag}
-\end{lyxcode}
-where
-\texttt{domain\_ID} is 1 for acoustic or 2 for elastic materials, \texttt{material\_ID} a unique identifier,  \texttt{rho} the density in $kg \, m^{-3}$,  \texttt{vp} the P-wave speed in $m \, s^{-1}$, \texttt{vs}  the S-wave speed in $m \, s^{-1}$, \texttt{Q} the quality factor and \texttt{anisotropy\_flag} an identifier for anisotropic models.
-
-%%magnoni
-For tomographic velocity models, please read Chapter \ref{cha:-Changing-the} and Section \ref{sec:Using-tomographic} `Using external tomographic Earth models' for further details.
-%
-%For a tomographic model, the material definition format is:
-%\begin{lyxcode}
-%\texttt{domain\_ID} \texttt{material\_ID}  \texttt{tomography}  \texttt{name}
-%\end{lyxcode}
-%where
-%\texttt{domain\_ID} is 1 for acoustic or 2 for elastic materials, \texttt{material\_ID} a negative, unique identifier, \texttt{tomography} keyword for tomographic material definition and \texttt{name} the name of the tomography file.
-The name is not used so far, rather change the filename defined in the file \texttt{model\_tomography.f90} located in the
-%\texttt{src/generate\_databases/} directory.
-%
-\item [materials\_file] Contains the material associations for each element.
-\item [nodes\_coords\_file] Contains the point locations in Cartesian coordinates of the mesh element corners.
-\item [mesh\_file] Contains the mesh element connectivity.
-\item [free\_surface\_file] Contains the free surface connectivity.
-\item [absorbing\_surface\_file\_xmax] Contains the surface connectivity of the absorbing boundary surface at the Xmax.
-\item [absorbing\_surface\_file\_xmin] Contains the surface connectivity of the absorbing boundary surface at the Xmin.
-\item [absorbing\_surface\_file\_ymax] Contains the surface connectivity of the absorbing boundary surface at the Ymax.
-\item [absorbing\_surface\_file\_ymin] Contains the surface connectivity of the absorbing boundary surface at the Ymin.
-\item [absorbing\_surface\_file\_bottom] Contains the surface connectivity of the absorbing boundary surface at the bottom.
-\end{description}
-
-These mesh files are needed as input files for the partitioner  \texttt{xdecompose\_mesh\_SCOTCH}
-to load-balance the mesh. Please see the next section for further details.
-
-\subsubsection*{Checking the mesh quality}
-The quality of the mesh may be inspected more precisely based upon the serial code
-in the file  \texttt{check\_mesh\_quality\_}~\\
-\texttt{CUBIT\_Abaqus.f90} located in the directory \texttt{src/check\_mesh\_quality\_CUBIT\_Abaqus/}.
-Running this code is optional because no information needed by the
-solver is generated.
-
-Prior to running and compiling this code, you have to export your mesh in CUBIT to an ABAQUS (.inp) format.
-For example, export mesh block IDs belonging to volumes in order to check the quality of the hexahedral elements.
-You also have to determine a number of parameters of your mesh, such as the number of nodes and number of elements
-and modify the header of the \texttt{check\_mesh\_quality\_CUBIT\_Abaqus.f90} source file. Then,
-in the main directory, type
-\begin{lyxcode}
-{\small make xcheck\_mesh\_quality\_CUBIT\_Abaqus}{\small \par}
-\end{lyxcode}
-and use
-\begin{lyxcode}
-{\small ./bin/xcheck\_mesh\_quality\_CUBIT\_Abaqus}{\small \par}
-\end{lyxcode}
-to generate an
-%not supported: AVS output file (\texttt{\small AVS\_meshquality.inp} in AVS UCD format) or
-OpenDX output file (\texttt{\small DX\_mesh\_quality.dx})
-that can be used to investigate mesh quality, e.g. skewness of elements,
-and a Gnuplot histogram (\texttt{\small mesh\_quality\_histogram.txt})
-that can be plotted with gnuplot (type
-`\texttt{\small gnuplot plot\_mesh\_quality\_histogram.gnu}'). The
-histogram is also printed to the screen.
-Analyze that skewness histogram of mesh elements to make sure no element has a skewness above approximately 0.75, otherwise the mesh is of poor quality (and if even a single element has a skewness value above 0.80, then you must definitely improve the mesh).
-If you want to start designing
-your own meshes, this tool is useful for viewing your creations. You
-are striving for meshes with elements with `cube-like' dimensions,
-e.g., the mesh should contain no very elongated or skewed elements.
-
-\subsection{Partitioning the Mesh with \texttt{xdecompose\_mesh\_SCOTCH}}
-
-The SPECFEM3D software package performs large scale simulations in a parallel 'Single Process Multiple Data' way.
-The spectral-element mesh created with CUBIT needs to be distributed on the processors. This partitioning is executed once and for all prior to
-the execution of the solver so it is referred to as a static mapping.
-
-An efficient partitioning is important because it leverages the overall running time of the application.
-It amounts to balance the number of elements in each slice while minimizing the communication costs
-resulting from the placement of adjacent elements on different processors. \texttt{decompose\_mesh\_SCOTCH} depends on the SCOTCH library \citep{PeRo96},
-which provides efficent static mapping, graph and mesh partitioning routines. SCOTCH is a free software package developed by
-Fran\c {c}ois Pellegrini et al.
-from LaBRI and INRIA in Bordeaux, France, downloadable from the web page \url{https://gforge.inria.fr/projects/scotch/}.
-It is more recent than METIS, actively maintained and performs better in many cases.
-A recent version of its source code is provided in directory \texttt{src/decompose\_mesh\_SCOTCH/scotch\_5.1.10b}.
-\\
-
-In most cases, the configuration with \texttt{./configure~FC=ifort} should be sufficient.
-During the configuration process, the script tries to find existing SCOTCH installations.
-In case your system has no pre-existing SCOTCH installation,
-we provide the source code of SCOTCH, which is released open source under the French CeCILL-C version 1 license,
-in directory
-\texttt{src/decompose\_mesh\_SCOTCH/scotch\_5.1.10b}.
-This version gets bundled with the compilation of the SPECFEM3D package if no libraries could have been found.
-If this automatic compilation of the SCOTCH libraries fails,
-please refer to file INSTALL.txt in that directory to see further details how to compile it on your system.
-In case you want to use a pre-existing installation, make sure you have correctly specified the path
-of the SCOTCH library when using the option \texttt{-{}-with-scotch-dir} with the \texttt{./configure} script.
-In the future you should be able to find more recent versions at  \url{http://www.labri.fr/perso/pelegrin/scotch/scotch\_en.html} .
-\\
-
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=.45\textwidth]{figures/mount-partitions.jpg}
-\includegraphics[width=.45\textwidth]{figures/mount-partitions2.jpg}
-\par\end{centering}
-
-\caption{Example of a mesh partitioning onto four cores.
-Each single core partition is colored differently.
-The executable \texttt{xdecompose\_mesh\_SCOTCH} can equally distribute the mesh on any arbitrary number of cores.
-Domain decomposition is explained in detail in \cite{MaKoBlLe08}, and excellent scaling up to 150,000 processor cores in shown for instance in
-\cite{CaKoLaTiMiLeSnTr08,KoLaMi08a,MaKoBlLe08,KoErGoMi10,KoGoErMi10,Kom11}.}
-
-\label{fig:mount.partitions}
-\end{figure}
-
-When you are ready to compile,
-in the main directory type
-\begin{lyxcode}
-{\small make xdecompose\_mesh\_SCOTCH}{\small \par}
-\end{lyxcode}
-If all paths and flags have been set correctly, the executable \texttt{bin/xdecompose\_mesh\_SCOTCH} should be produced.
-
-The partitioning is done in serial for now (in the next release we will provide a parallel version of that code),
-the synopsis is:
-\begin{lyxcode}
-./bin/xdecompose\_mesh\_SCOTCH  nparts  input\_directory output\_directory
-\end{lyxcode}
-where
-\begin{itemize}
-     \item \texttt{nparts} is the number of partitions, i.e., the number of cores for the parallel simulations,
-     \item \texttt{input\_directory} is the directory which holds all the files generated by the Python script \texttt{cubit2specfem3d.py} explained in the previous Section~\ref{subsec:Exporting-the-Mesh}, e.g.
-\texttt{MESH/}, and
-     \item \texttt{output\_directory} is the directory for the output of this partitioner which stores ACII-format files named like \texttt{proc******\_Database} for each partition.
-These files will be needed for creating the distributed databases, and have to reside in the directory  \texttt{LOCAL\_PATH} specified in the main \texttt{Par\_file}, e.g.
-in directory \texttt{in\_out\_files/DATABASES\_MPI}.
-Please see Chapter~\ref{cha:Creating-Distributed-Databases} for further details.
-\end{itemize}
-Note that all the files generated by the Python script  \texttt{cubit2specfem3d.py} must be placed in the \texttt{input\_directory} folder before running the program.
-
-\section{\label{cha:Running-the-Mesher-Meshfem3D}Meshing with \texttt{xmeshfem3D}}
-
-In case you successfully ran the configuration script, you are also ready to compile the internal mesher.
-This is an alternative to CUBIT
-for the mesh generation of relatively simple geological models. The mesher is no longer
-dedicated to Southern California and more flexiblity is provided in this version of the package.
-
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[scale=0.5]{figures/socal_map_mpi}
-\par\end{centering}
-
-\caption{\label{fig:For-parallel-computing}For parallel computing purposes,
-the model block is subdivided in $\nprocxi\times\nproceta$ slices
-of elements. In this example we use $5^{2}=25$ processors. }
-
-\end{figure}
-
-In the main directory, type
-\begin{lyxcode}
-{\small make xmeshfem3D}{\small \par}
-\end{lyxcode}
-If all paths and flags
-have been set correctly, the mesher should now compile and produce
-the executable \texttt{bin/xmeshfem3D}. Please note that
-\texttt{xmeshfem3D} must be called directly from the \texttt{bin/} directory,
-as most of the binaries of the package.
-
-Input for the mesh generation program is provided through the parameter
-file \texttt{Mesh\_Par\_file}, which resides in the subdirectory \texttt{in\_data\_files/meshfem3D\_files/}.
-Before running the mesher, a number of parameters need to be set in
-the \texttt{Mesh\_Par\_file}. This requires a basic understanding of how
-the SEM is implemented, and we encourage you to read \citet{KoVi98,KoTr99}
-and \citet{KoLiTrSuStSh04}.
-
-The mesher and the solver use UTM coordinates internally, therefore
-you need to define the zone number for the UTM projection (e.g., zone
-11 for Los Angeles). Use decimal values for latitude and longitude
-(no minutes/seconds). These values are approximate; the mesher will
-round them off to define a square mesh in UTM coordinates. When running
-benchmarks on rectangular models, turn the UTM projection off by using
-the flag \texttt{\small SUPPRESS\_UTM\_PROJECTION}, in which case
-all `longitude' parameters simply refer to the $x$~axis, and all
-`latitude' parameters simply refer to the $y$~axis. To run the mesher
-for a global simulation, the following parameters need to be set in
-the \texttt{Mesh\_Par\_file}:
-
-\begin{description}
-\item [{\texttt{LATITUDE\_MIN}}] Minimum latitude in the block (negative
-for South).
-\item [{\texttt{LATITUDE\_MAX}}] Maximum latitude in the block.
-\item [{\texttt{LONGITUDE\_MIN}}] Minimum longitude in the block (negative
-for West).
-\item [{\texttt{LONGITUDE\_MAX}}] Maximum longitude in the block.
-\item [{\texttt{DEPTH\_BLOCK\_KM}}] Depth of bottom of mesh in kilometers.
-\item [{\texttt{\noun{UTM\_PROJECTION\_ZONE}}}] UTM projection zone in
-which your model resides, only valid when \texttt{SUPPRESS\_UTM\_}~\\
-\texttt{PROJECTION} is \texttt{.false.}.
-\item [{\texttt{SUPPRESS\_UTM\_PROJECTION}}] set to be \texttt{.false.}
-when your model range is specified in geographical coordinates,
-and needs to be \texttt{.true.} when your model is specified in
-Cartesian coordinates. \noun{UTM projection zone in which your simulation
-region resides.}
-\item [{\texttt{INTERFACES\_FILE }}] File which contains the description of the topography and
-of the interfaces between the different layers of the model, if any.
-The number of spectral elements in the vertical direction within each layer is also defined in this file.
-
-\item [{$\nexxi$}] The number of spectral elements along one side of the
-block. This number \textit{must} be 8~$\times$~a multiple of $\nprocxi$
-defined below. Based upon benchmarks against semi-analytical discrete
-wavenumber synthetic seismograms \citep{KoLiTrSuStSh04}, determined
-that a $\nexxi=288$ run is accurate to a shortest period of roughly
-2~s. Therefore, since accuracy is determined by the number of grid
-points per shortest wavelength, for any particular value of $\nexxi$
-the simulation will be accurate to a shortest period determined by
-\begin{equation}
-\mbox{shortest period (s)}=(288/\nexxi)\times2.\label{eq:shortest_period}\end{equation}
-The number of grid points in each orthogonal direction of the reference
-element, i.e., the number of Gauss-Lobatto-Legendre points, is determined
-by \texttt{NGLLX} in the \texttt{constants.h} file. We generally use
-$\mbox{\texttt{NGLLX\/}}=5$, for a total of $5^{3}=125$ points per
-elements. We suggest not to change this value.
-\item [{$\nexeta$}] The number of spectral elements along the other side
-of the block. This number \textit{must} be 8~$\times$~a multiple
-of $\nproceta$ defined below.
-\item [{$\nprocxi$}] The number of processors or slices along one side
-of the block (see Figure~\ref{fig:For-parallel-computing}); we must
-have $\nexxi=8\times c\times\nprocxi$, where $c\ge1$ is a positive
-integer.
-\item [{$\nproceta$}] The number of processors or slices along the other
-side of the block; we must have $\nexeta=8\times c\times\nproceta$,
-where $c\ge1$ is a positive integer.
-\item [{\texttt{USE\_REGULAR\_MESH}}] set to be \texttt{.true.} if you want a perfectly regular
-mesh or \texttt{.false.} if you want to add doubling horizontal layers to coarsen the mesh. In this case, you
-also need to provide additional information by setting up the next three parameters.
-\item [{\texttt{NDOUBLINGS}}] The number of horizontal doubling layers. Must be set to \texttt{1} or \texttt{2}
-if \texttt{USE\_REGULAR\_MESH} is set to \texttt{.true.}.
-\item [{\texttt{NZ\_DOUBLING\_1}}] The position of the first doubling layer (only interpreted
-if \texttt{USE\_REGULAR\_MESH} is set to \texttt{.true.}).
-\item [{\texttt{NZ\_DOUBLING\_2}}] The position of the second doubling layer (only interpreted
-if \texttt{USE\_REGULAR\_MESH} is set to \texttt{.true.} and if \texttt{NDOUBLINGS} is set to \texttt{2}).
-\item [{\texttt{CREATE\_ABAQUS\_FILES}}] Set this flag to \texttt{.true.} to save Abaqus FEA \urlwithparentheses{www.simulia.com}
-mesh files for subsequent viewing. Turning the flag on generates files in the \texttt{LOCAL\_PATH} directory.
-See Section~\ref{sec:Mesh-graphics} for a discussion of mesh viewing features.
-\item [{\texttt{CREATE\_DX\_FILES}}] Set this flag to \texttt{.true.} to save OpenDX \urlwithparentheses{www.opendx.org}
-mesh files for subsequent viewing.
-\item [{\texttt{LOCAL\_PATH}}] Directory in which the partitions generated
-by the mesher will be written. Generally one uses a directory on the
-local disk of the compute nodes, although on some machines these partitions
-are written on a parallel (global) file system (see also the earlier
-discussion of the \texttt{LOCAL\_PATH\_IS\_ALSO\_GLOBAL} flag in Chapter~\ref{cha:Getting-Started}).
-The mesher generates the necessary partitions in parallel, one set
-for each of the $\nprocxi\times\nproceta$ slices that constitutes
-the mesh (see Figure~\ref{fig:For-parallel-computing}). After the
-mesher finishes, you can log in to one of the compute nodes and view
-the contents of the \texttt{LOCAL\_PATH} directory to see the
-files generated by the mesher.
-These files will be needed for creating the distributed databases, and have to reside in the directory  \texttt{LOCAL\_PATH} specified in the main \texttt{Par\_file}, e.g.
-in directory \texttt{in\_out\_files/DATABASES\_MPI}.
-Please see Chapter~\ref{cha:Creating-Distributed-Databases} for further details.
-\item [{\texttt{NMATERIALS}}] The number of different materials in your model. In the following lines, each material needs to be defined as :
-\begin{lyxcode}
-\texttt{material\_ID}  \texttt{rho}  \texttt{vp}  \texttt{vs}  \texttt{Q}  \texttt{anisotropy\_flag} \texttt{domain\_ID}
-\end{lyxcode}
-where
-\begin{itemize}
-     \item \texttt{Q}           : quality factor (0=no attenuation)
-     \item \texttt{anisotropy\_flag}  : 0=no anisotropy / 1,2,.. check with implementation in \texttt{aniso\_model.f90}
-     \item \texttt{domain\_id}        : 1=acoustic / 2=elastic
-\end{itemize}
-\item [{\texttt{NREGIONS}}] The number of regions in the mesh.
-In the following lines, because the mesh is regular or 'almost regular', each region is defined as :
-\begin{lyxcode}
-  \texttt{NEX\_XI\_BEGIN}  \texttt{NEX\_XI\_END}  \texttt{NEX\_ETA\_BEGIN}  \texttt{NEX\_ETA\_END}  \texttt{NZ\_BEGIN}  \texttt{NZ\_END}  \texttt{material\_ID}
-\end{lyxcode}
-\end{description}
-
-The \texttt{INTERFACES\_FILE} parameter of \texttt{Mesh\_Par\_File} defines the file which contains the settings of the topography grid and of the interfaces grids.
-Topography is defined as a set of elevation values on a regular 2D grid.
-It is also possible to
-define interfaces between the layers of the model in the same way.
-The file needs to define several parameters:
-\begin{itemize}
-\item The number of interfaces, including the topography. This needs to be set at the first line. Then, from the bottom
-to the top of the model, you need to define the grids with:
-\item \texttt{SUPPRESS\_UTM\_PROJECTION} flag as described previously,
-\item number of points along $x$ and $y$ direction (NXI and NETA),
-\item minimal $x$ and $y$ coordinates (LONG\_MIN and LAT\_MIN),
-\item spacing between points along $x$ and $y$ (SPACING\_XI and SPACING\_ETA) and
-\item the name of the file which contains the elevation values (in $y$.$x$ increasing order).
-\end{itemize}
-At the end of this file, you simply need to set the number of spectral elements in the vertical direction for each layer. We provide a few models in the {\texttt{examples/}} directory.  \\
-
-Finally, depending on your system, you might need to provide a file that tells MPI what compute nodes
-to use for the simulations. The file must have a number of entries
-(one entry per line) at least equal to the number of processors needed
-for the run. A sample file is provided in the file \texttt{mymachines}.
-This file is not used by the mesher or solver, but is required by
-the \texttt{go\_mesher} and \texttt{go\_solver} default job submission
-scripts. See Chapter~\ref{cha:Scheduler} for information about running
-the code on a system with a scheduler, e.g., LSF.
-
-Now that you have set the appropriate parameters in the \texttt{Mesh\_Par\_file}
-and have compiled the mesher, you are ready to launch it! This is
-most easily accomplished based upon the \texttt{go\_mesher} script.
-When you run on a PC cluster, the script assumes that the nodes are
-named n001, n002, etc. If this is not the case, change the \texttt{tr
--d `n'} line in the script. You may also need to edit the last command
-at the end of the script that invokes the \texttt{mpirun} command.
-See Chapter~\ref{cha:Scheduler} for information about running the
-code on a system with a scheduler, e.g., LSF.
-
-Mesher output is provided in the \texttt{in\_out\_files/OUTPUT\_FILES} directory
-in \texttt{output\_mesher.txt}; this file provides lots of details
-about the mesh that was generated. Please note that the mesher suggests a time step \texttt{DT} to run
-the solver with. The mesher output file also contains a table about the quality of the mesh to indicate
-possible problems with the distortions of elements.
-Alternatively, output can be directed
-to the screen instead by uncommenting a line in \texttt{constants.h}:
-\begin{lyxcode}
-!~uncomment~this~to~write~messages~to~the~screen~
-
-!~integer,~parameter~::~IMAIN~=~ISTANDARD\_OUTPUT
-\end{lyxcode}
-
-To control the quality of the mesh, check the standard output (either on the screen or in the \texttt{in\_out\_files/OUTPUT\_FILES} directory in \texttt{output\_mesher.txt}) and analyze the skewness histogram of mesh elements to make sure no element has a skewness above approximately 0.75, otherwise the mesh is of poor quality (and if even a single element has a skewness value above 0.80, then you must definitely improve the mesh).
-To draw the skewness histogram on the screen, type \texttt{gnuplot plot\_mesh\_quality\_histogram.gnu}.
-
-\chapter{\label{cha:Creating-Distributed-Databases}Creating the Distributed Databases}
-
-After using \texttt{xmeshfem3D} or \texttt{xdecompose\_mesh\_SCOTCH}, the next step in the workflow is to compile
-\texttt{xgenerate\_}~\\
-\texttt{databases}. This program is going to create all the missing information needed by the SEM solver.
-
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=.6\textwidth]{figures/workflow.jpg}
-\par\end{centering}
-
-\caption{Schematic workflow for a SPECFEM3D simulation. The executable \texttt{xgenerate\_databases} creates the GLL mesh points and assigns specific model parameters.}
-
-\label{fig:workflow.databases}
-\end{figure}
-
-In the main directory, type
-\begin{lyxcode}
-{\small make xgenerate\_databases}{\small \par}
-\end{lyxcode}
-Input for the program is provided through
-the main parameter file \texttt{Par\_file}, which resides in the subdirectory \texttt{in\_data\_files/}. Please note that
-\texttt{xgenerate\_databases} must be called directly from the \texttt{bin/} directory, as most of the binaries of the package.
-
-\section{\label{cha:Main-Parameter}Main parameter file \texttt{Par\_file}}
-
-Before running \texttt{xgenerate\_databases}, a number of parameters need to be set in the main parameter \texttt{Par\_file}
-located in the subdirectory \texttt{in\_data\_files/}:
-
-\begin{description}
-\item [{\texttt{SIMULATION\_TYPE}}] is set to 1 for forward simulations,
-2 for adjoint simulations (see Section \ref{sec:Adjoint-simulation-finite})
-and 3 for kernel simulations (see Section \ref{sec:Finite-Frequency-Kernels}).
-\item [{\texttt{SAVE\_FORWARD}}] is only set to \texttt{.true.} for a forward
-simulation with the last frame of the simulation saved, as part of
-the finite-frequency kernel calculations (see Section \ref{sec:Finite-Frequency-Kernels}).
-For a regular forward simulation, leave \texttt{SIMULATION\_TYPE}
-and \texttt{SAVE\_FORWARD} at their default values.
-\item [{\texttt{\noun{UTM\_PROJECTION\_ZONE}}}] UTM projection zone in
-which your model resides, only valid when \texttt{SUPPRESS\_UTM\_}~\\
-\texttt{PROJECTION} is \texttt{.false.}.
-\item [{\texttt{SUPPRESS\_UTM\_PROJECTION}}] set to be \texttt{.false.}
-when your model range is specified in the geographical coordinates,
-and needs to be \texttt{.true.} when your model is specified in a
-cartesian coordinates. \noun{UTM projection zone in which your simulation
-region resides.}
-\item [{\texttt{NPROC}}] The number of MPI processors, each one is assigned one slice of the whole mesh.
-\item [{\texttt{NSTEP}}] The number of time steps of the simulation. This controls the
-length of the numerical simulation, i.e., twice the number of time steps
-requires twice as much CPU time. This feature is not used at the time
-of generating the distributed databases but is required for the solver, i.e., you may change this
-parameter after running  \texttt{xgenerate\_databases}.
-\item [{\texttt{DT}}] The length of each time step in seconds.
-This feature is not used at the time of generating the distributed databases but is required for the solver.
-Please see also Section~\ref{sec:Choosing-the-Time-Step} for further details.
-%
-\item [{\texttt{MODEL}}] Must be set to one of the following:
-%
-\begin{description}
-\item [{\textmd{Models defined by mesh parameters:}}]~
-\begin{description}
-\item [{\texttt{default}}] Uses model parameters as defined by meshing procedures described
-in the previous Chapter~\ref{cha:Mesh-Generation}.
-\end{description}
-\end{description}
-%
-\begin{description}
-\item [{\textmd{1D~models~with~real~structure:}}]~
-\begin{description}
-\item [{\texttt{1D\_prem}}] Isotropic version of the spherically
-symmetric Preliminary Reference Earth Model (PREM) \citep{DzAn81}.
-\item [{\texttt{1D\_socal}}] A standard isotropic 1D model for Southern California.
-\item [{\texttt{1D\_cascadia}}] Isotropic 1D profile for the Cascadia region.
-\end{description}
-\end{description}
-%
-\begin{description}
-\item [{\textmd{Fully~3D~models:}}]~
-\begin{description}
-\item [{\texttt{aniso}}] For a user-specified fully anisotropic model.
-Parameters are set up in routines located in file \texttt{model\_aniso.f90} in directory \texttt{src/generate\_databases/}.
-See Chapter~\ref{cha:-Changing-the} for a discussion on how to specify your own 3D model.
-\item [{\texttt{external}}] For a user-specified isotropic model which uses externally defined model parameters.
-Uses external model definitions set up in routines located in file \texttt{model\_external\_values.f90} in directory \texttt{src/generate\_databases/}.
-Please modify these generic template routines to use your own model definitions.
-\item [{\texttt{gll}}] For a user-specified isotropic model which uses external binary files for $v_p$, $v_s$ and $\rho$.
-Binary files are given in the same format as when outputted by the \texttt{xgenerate\_databases} executable when using option \texttt{SAVE\_MESH\_FILES}.
-These binary files define the model parameters on all GLL points which can be used for iterative inversion procedures.
-\item [{\texttt{salton\_trough}}] A 3D $V_p$ model for Southern California.
-Users must provide the corresponding data file \texttt{regrid3\_vel\_p.bin} in directory \texttt{DATA/st\_3D\_block\_harvard/}.
-\item [{\texttt{tomo}}] For a user-specified 3D isotropic model which uses a tomographic model file \texttt{tomographic\_model.xyz} in directory \texttt{in\_data\_files/}.
-See Section \ref{sec:Using-tomographic},
-for a discussion on how to specify your own 3D tomographic model.
-\end{description}
-\end{description}
-%
-%
-\item [{\texttt{OCEANS}}] Set to \texttt{.true.} if the effect of the oceans
-on seismic wave propagation should be incorporated based upon the
-approximate treatment discussed in \citet{KoTr02b}. This feature
-is inexpensive from a numerical perspective, both in terms of memory
-requirements and CPU time. This approximation is accurate at periods
-of roughly 20~s and longer. At shorter periods the effect of water
-phases/reverberations is not taken into account, even when the flag
-is on.
-\item [{\texttt{TOPOGRAPHY}}] This feature is only effective if \texttt{OCEANS} is set to \texttt{.true.}.
-Set to \texttt{.true.} if topography and
-bathymetry should be read in based upon the topography file specified in the main
-constants file \texttt{constants.h} found in subdirectory \texttt{src/shared/} to evaluate
-elevations. If not set, elevations will be read from the numerical mesh.
-\item [{\texttt{ATTENUATION}}] Set to \texttt{.true.} if attenuation should
-be incorporated. Turning this feature on increases the memory requirements
-significantly (roughly by a factor of~1.5), and is numerically fairly
-expensive. See \citet{KoTr99,KoTr02a} for a discussion on the implementation
-of attenuation based upon standard linear solids.
-Please note that the Vp- and Vs-velocities of your model are given for a reference frequency.
-To change this reference frequency, you change the value of
-\texttt{ATTENUATION\_f0\_REFERENCE} in the main constants file \texttt{constants.h} found in subdirectory \texttt{src/shared/}.
-\item [{\texttt{USE\_OLSEN\_ATTENUATION}}] Set to \texttt{.true.} if you
-want to use the attenuation model that scaled from the S-wave speed model
-using Olsen's empirical relation (see \citet{OlDaBr03}).
-\item [{\texttt{ANISOTROPY}}] Set to \texttt{.true.} if you
-want to use an anisotropy model. Please see the file \texttt{model\_aniso.f90} in subdirectory \texttt{src/generate\_databases/} for the current implementation of anisotropic models.
-\item [{\texttt{ABSORBING\_CONDITIONS}}] Set to \texttt{.true.} to turn
-on Clayton-Enquist absorbing boundary conditions (see \citet{KoTr99}).
-\item [{\texttt{ABSORB\_INSTEAD\_OF\_FREE\_SURFACE}}] Set to \texttt{.true.} to turn on 
-absorbing boundary conditions on the top surface which by default constitutes a free surface of the model.
-\item [{\texttt{MOVIE\_SURFACE}}] Set to \texttt{.false.}, unless you want
-to create a movie of seismic wave propagation on the Earth's surface.
-Turning this option on generates large output files. See Section~\ref{sec:Movies}
-for a discussion on the generation of movies. This feature is only relevant for the solver.
-\item [{\texttt{MOVIE\_VOLUME}}] Set to \texttt{.false.}, unless you want
-to create a movie of seismic wave propagation in the Earth's interior.
-Turning this option on generates huge output files. See Section~\ref{sec:Movies}
-for a discussion on the generation of movies. This feature is only relevant for the solver.
-\item [{\texttt{NTSTEP\_BETWEEN\_FRAMES}}] Determines the number of timesteps
-between movie frames. Typically you want to save a snapshot every
-100 timesteps. The smaller you make this number the more output will
-be generated! See Section~\ref{sec:Movies} for a discussion on the
-generation of movies. This feature is only relevant for the solver.
-\item [{\texttt{CREATE\_SHAKEMAP}}] Set this flag to \texttt{.true.} to
-create a ShakeMap\textregistered, i.e., a peak ground velocity map
-of the maximum absolute value of the two horizontal components of the velocity vector.
-\item [{\texttt{SAVE\_DISPLACEMENT}}] Set this flag to \texttt{.true.}
-if you want to save the displacement instead of velocity for the movie
-frames.
-\item [{\texttt{USE\_HIGHRES\_FOR\_MOVIES}}] Set this flag to \texttt{.true.}
-if you want to save the values at all the NGLL grid points for the
-movie frames.
-\item [{\texttt{HDUR\_MOVIE}}] determines the half duration of the source time function for the movie simulations.
-When this parameter is set to be 0, a default half duration that corresponds to the accuracy of the simulation is provided.
-Otherwise, it adds this half duration to the half duration specified in the source file \texttt{CMTSOLUTION},
-thus simulates longer periods to make the movie images look smoother.
-\item [{\texttt{SAVE\_MESH\_FILES}}] Set this flag to \texttt{.true.} to
-save ParaView \urlwithparentheses{www.paraview.org}
-mesh files for subsequent viewing. Turning the flag on generates large
-(distributed) files in the \texttt{LOCAL\_PATH} directory. See Section~\ref{sec:Mesh-graphics}
-for a discussion of mesh viewing features.
-\item [{\texttt{LOCAL\_PATH}}] Directory in which the distributed databases
-will be written. Generally one uses a directory on the
-local disk of the compute nodes, although on some machines these databases
-are written on a parallel (global) file system (see also the earlier
-discussion of the \texttt{LOCAL\_PATH\_IS\_ALSO\_GLOBAL} flag in Chapter~\ref{cha:Getting-Started}).
-\texttt{xgenerate\_databases} generates the necessary databases in parallel, one set
-for each of the \texttt{NPROC} slices that constitutes
-the mesh (see Figure~\ref{fig:mount.partitions} and Figure~\ref{fig:For-parallel-computing}). After
-the executable finishes, you can log in to one of the compute nodes and view
-the contents of the \texttt{LOCAL\_PATH} directory to see the (many)
-files generated by \texttt{xgenerate\_databases}.
-Please note that the \texttt{LOCAL\_PATH} directory should already contain the output files of
-the partitioner, i.e. from \texttt{xdecompose\_mesh\_SCOTCH} or \texttt{xmeshfem3D}.
-\item [{\texttt{NTSTEP\_BETWEEN\_OUTPUT\_INFO}}] This parameter specifies
-the interval at which basic information about a run is written to
-the file system (\texttt{timestamp{*}} files in the \texttt{in\_out\_files/OUTPUT\_FILES}
-directory). If you have access to a fast machine, set \texttt{NTSTEP\_BETWEEN\_OUTPUT\_INFO}
-to a relatively high value (e.g., at least 100, or even 1000 or more)
-to avoid writing output text files too often. This feature is not
-used at the time of meshing. One can set this parameter to a larger
-value than the number of time steps to avoid writing output during
-the run.
-\item [{\texttt{NTSTEP\_BETWEEN\_OUTPUT\_SEISMOS}}] This parameter specifies
-the interval at which synthetic seismograms are written in the \texttt{LOCAL\_PATH}
-directory. If a run crashes, you may still find usable (but shorter
-than requested) seismograms in this directory. On a fast machine set
-\texttt{NTSTEP\_BETWEEN\_OUTPUT\_SEISMOS} to a relatively high value
-to avoid writing to the seismograms too often. This feature is only relevant for the solver.
-\item [{\texttt{USE\_FORCE\_POINT\_SOURCE}}] Turn this flag on to use a \texttt{FORCESOLUTION} force source
-instead of a \texttt{CMTSOLUTION} moment-tensor source. This can be useful e.g. for oil industry 
-foothills simulations in which the source is a vertical force, normal force, 
-tilted force, or an impact etc. You will need to give the East, North and vertical components of an 
-arbitrary (non-unitary) direction vector of the force vector; thus refer to Appendix~\ref{cha:Coordinates} 
-for the orientation of the reference frame.
-This vector is made unitary internally in the code and thus only its direction matters here;
-its norm is ignored and the norm of the force used is the factor force source times the source time function.
-\item [{\texttt{PRINT\_SOURCE\_TIME\_FUNCTION}}] Turn this flag on to print
-information about the source time function in the file \texttt{in\_out\_files/OUTPUT\_FILES/plot\_source\_time\_function.txt}.
-This feature is only relevant for the solver.
-\end{description}
-
-\section{\label{sec:Choosing-the-Time-Step}Choosing the time step \texttt{DT}}
-
-The parameter \texttt{DT} sets the length of each time step in seconds. The value of this parameter is crucial for the stability of
-the spectral-element simulation.
-Your time step \texttt{DT} will depend on the minimum ratio between the distance $h$ of neighboring mesh points and the wave speeds $v$ defined in your model.
-The condition for the time step $\Delta t$ is:
-
-\begin{lyxcode}
-$\Delta t < C ~\mathrm{min}_\Omega (~ h / v ~)$
-\end{lyxcode}
-where $C$ is the so-called Courant number and $\Omega$ denotes the model volume.
-The distance $h$ depends on the mesh element size and the number of
-GLL points \texttt{NGLL} specified in the main constants file \texttt{constants.h} located in the \texttt{src/shared/} subdirectory.
-The wave speed $v$ is determined based on your model's P- (or S-) wave speed values.
-
-The database generator \texttt{xgenerate\_databases}, as well as the internal mesher \texttt{xmeshfem3D}, are trying
-to evaluate the value of $\Delta t$ for empirically chosen Courant numbers $C \sim 0.3$.
-If you used the mesher \texttt{xmeshfem3D} to generate your mesh, you should set the value suggested in \texttt{in\_out\_files/OUTPUT\_FILES/output\_mesher.txt} file, which is created after the mesher completed.
-In case you used CUBIT to create the mesh, you might use an arbitrary value when running \texttt{xgenerate\_databases} and then use the value suggested in the ~\\
-\texttt{in\_out\_files/OUTPUT\_FILES/output\_mesher.txt} file after the database generation completed.
-Note that the implemented Newmark time scheme uses this time step globally, thus your simulations become more expensive for very small mesh elements in high wave-speed regions.
-Please be aware of this restriction when constructing your mesh in Chapter~\ref{cha:Mesh-Generation}.
-
-\chapter{\label{cha:Running-the-Solver}Running the Solver \texttt{xspecfem3D}}
-
-Now that you have successfully generated the databases, you are ready to compile
-the solver. In the main directory, type
-\begin{lyxcode}
-{\small make xspecfem3D}{\small \par}
-\end{lyxcode}
-Please note that \texttt{xspecfem3D} must be called directly from
-the \texttt{bin/} directory, as most of the binaries of the package.
-
-The solver needs three input files in the \texttt{in\_data\_files/} directory
+The solver needs three input files in the \texttt{DATA} directory
 to run:
 \begin{description}
-\item [{\texttt{Par\_file}}] the main parameter file which was discussed in detail in the previous
-Chapter~\ref{cha:Creating-Distributed-Databases},
-\item [{\texttt{CMTSOLUTION}} or {\texttt{FORCESOLUTION}}] the earthquake source parameter file or the force source parameter file or, and
+\item [{\texttt{Par\_file}}] the main parameter file which was discussed
+in detail in the previous Chapter~\ref{cha:Creating-Distributed-Databases},
+\item [{\texttt{CMTSOLUTION}} or {\texttt{FORCESOLUTION}}] the earthquake
+source parameter file or the force source parameter file, and
 \item [{\texttt{STATIONS}}] the stations file.
 \end{description}
-
-Most parameters in the \texttt{Par\_file}
-should be set prior to running the databases generation. Only the following parameters
-may be changed after running \texttt{xgenerate\_databases}:
-
+Most parameters in the \texttt{Par\_file} should be set prior to running
+the databases generation. Only the following parameters may be changed
+after running \texttt{xgenerate\_databases}:
 \begin{itemize}
@@
 \item the time step parameters \texttt{NSTEP} and \texttt{DT}
-\item the absorbing boundary control parameter \texttt{ABSORBING\_CONDITIONS} on condition that the\\ 
-  \texttt{ABSORB\_INSTEAD\_OF\_FREE\_SURFACE} flag remains unmodified after running the databases generation. 
+\item the absorbing boundary control parameter \texttt{PML\_CONDITIONS}
+on condition that the\\
+ \texttt{PML\_INSTEAD\_OF\_FREE\_SURFACE} flag remains unmodified
+after running the databases generation.
 \item the movie control parameters \texttt{MOVIE\_SURFACE}, \texttt{MOVIE\_VOLUME},
 and \texttt{NTSTEPS\_BETWEEN\_FRAMES}
-\item the ShakeMap\textregistered option \texttt{CREATE\_SHAKEMAP}
-\item the output information parameters \texttt{NTSTEP\_BETWEEN\_OUTPUT\_INFO}
+\item the ShakeMap\textregistered{}option \texttt{CREATE\_SHAKEMAP}
+\item the output information parameters \texttt{MOVIE\_TYPE}, \texttt{NTSTEP\_BETWEEN\_OUTPUT\_INFO}
 and \texttt{NTSTEP\_BETWEEN\_OUTPUT\_}~\\
@@
 Any other change to the \texttt{Par\_file} implies rerunning both
-the database generator \texttt{xgenerate\_databases} and the solver \texttt{xspecfem3D}.
+the database generator \texttt{xgenerate\_databases} and the solver
+\texttt{xspecfem3D}.
 
 For any particular earthquake, the \texttt{CMTSOLUTION} file that
-represents the point source may be obtained directly from the Harvard Centroid-Moment Tensor (CMT) web page \urlwithparentheses{www.seismology.harvard.edu}.
-It looks like this:
-
-\begin{lyxcode}
-{\small }%
-\begin{figure}[H]
-\noindent \begin{centering}
-{\small \includegraphics[width=1\textwidth]{figures/Hollywood_CMT} }
-\par\end{centering}{\small \par}
-
-\caption{\label{fig:CMTSOLUTION-file}\texttt{CMTSOLUTION} file based on the
+represents the point source may be obtained directly from the Harvard
+Centroid-Moment Tensor (CMT) web page \urlwithparentheses{www.seismology.harvard.edu}.
+It looks like the example shown in Fig. \ref{fig:CMTSOLUTION-file}.
+%\begin{lyxcode}
+\begin{figure}[htp]
+\noindent
+\begin{centering}
+{\small \includegraphics[width=0.8\textwidth]{figures/Hollywood_CMT.pdf}
+}
+\par
+\end{centering}
+\caption{\texttt{CMTSOLUTION} file based on the
 format from the Harvard CMT catalog. \textbf{M} is the moment tensor,
-$M_{0}${\small{} }is the seismic moment, and $M_{w}$ is the moment
+$M_{0}${\small {} }is the seismic moment, and $M_{w}$ is the moment
 magnitude.}
-
+\label{fig:CMTSOLUTION-file}
 \end{figure}
-{\small \par}
-\end{lyxcode}
-The \texttt{CMTSOLUTION} should be edited in the following way:
-
+%\end{lyxcode}
+The \texttt{CMTSOLUTION} file should be edited in the following way:
 \begin{itemize}
+\item Set the latitude or UTM $x$ coordinate, longitude or UTM $y$ coordinate, depth of the source (in km).
 \item Set the \texttt{time shift} parameter equal to $0.0$ (the solver
@@
 \end{itemize}
-%
-\begin{figure}
+\begin{figure}[htp]
 \noindent \begin{centering}
-%%%\includegraphics[width=3in]{figures/gauss_vs_triangle_mod.eps}
-\includegraphics[width=3in]{figures/gauss_vs_triangle_mod.pdf}
+\includegraphics[width=3.5in]{figures/gauss_vs_triangle_mod.pdf}
 \par\end{centering}
@@
 \end{figure}
+
+If you know the earthquake source in strike/dip/rake format rather
+than in \texttt{CMTSOLUTION} format, use the C code \texttt{SPECFEM3D\_GLOBE/utils/strike\_dip\_rake\_to\_CMTSOLUTION.c}
+to convert it. The conversion formulas are given for instance in \citet{AkRi80}.
+Note that the \citet{AkRi80} convention is slightly different from
+the Harvard \texttt{CMTSOLUTION} convention (the sign of some components
+is different). The C code outputs both.
 
@@
 
-\noindent The \texttt{FORCESOLUTION} should be edited in the following way:
-
+\noindent The \texttt{FORCESOLUTION} file should be edited in the
+following way:
 \begin{itemize}
@@
 in the post-processing (see Section \ref{sec:Process-data-and-syn}).
-\item Set the latitude, longitude, depth of the source in geographical
-  coordinates.
+\item Set the \texttt{f0} parameter (the dominant frequency) of the Ricker source time
+function when {\texttt{USE\_RICKER\_TIME\_FUNCTION}} is turned on
+in the main parameter file {\texttt{Par\_file}}. In case that the
+solver uses a (pseudo) Dirac delta source time function to represent
+a force point source, a very short duration of five time steps is automatically set by default.
+\item Set the latitude or UTM $x$ coordinate, longitude or UTM $y$ coordinate, depth of the source (in km).
 \item Set the magnitude of the force source.
-\item Set the components of a (non-unitary) direction vector for the
-  force source in the East/North/Vertical basis (see Appendix A for
-  the orientation of the reference frame).
+\item Set the components of a (non-unitary) direction vector for the force
+source in the East/North/Vertical basis (see Appendix A for the orientation
+of the reference frame).
 \end{itemize}
-
-{\small }%
+\noindent Where necessary, set a \texttt{FORCESOLUTION} file in the
+same way you configure a \texttt{CMTSOLUTION} file with $N_{\mathrm{sources}}$
+entries, one for each subevent (i.e., concatenate $N_{\mathrm{sources}}$
+\texttt{FORCESOLUTION} files to a single \texttt{FORCESOLUTION} file).
+At least one entry (not necessarily the first) must have a zero \texttt{time
+shift}, and all the other entries must have non-negative \texttt{time
+shift}. Each subevent can have its own half latitude, longitude, depth,
+\texttt{half duration} and force parameters.
+
 \begin{figure}[H]
@@
 {\small \includegraphics[width=5in]{figures/source_timing.pdf} }
-\par\end{centering}{\small \par}
+\par\end{centering}
 
@@
 
-{\small }%
 \begin{figure}[H]
 \noindent \begin{centering}
-{\small \includegraphics{figures/STATIONS_basin_explained} }
-\par\end{centering}{\small \par}
-
+{\small \includegraphics{figures/STATIONS_basin_explained.pdf} }
+\par\end{centering}
+
+\label{fig:Sample-STATIONS-file.}
 \caption{Sample \texttt{STATIONS} file. Station latitude and longitude should
@@
 \end{lyxcode}
-The solver \texttt{xspecfem3D} filters the list of stations in file \texttt{in\_data\_files/STATIONS}
-to exclude stations that are not located within the region given in
-the \texttt{Par\_file} (between \texttt{LATITUDE\_MIN} and \texttt{LATITUDE\_MAX}
-and between \texttt{LONGITUDE\_MIN} and \texttt{LONGITUDE\_MAX}).
-The filtered file is called \texttt{in\_data\_files/STATIONS\_FILTERED}.
-
-Solver output is provided in the \texttt{in\_out\_files/OUTPUT\_FILES} directory
+The solver \texttt{xspecfem3D} filters the list of stations in file
+\texttt{DATA/STATIONS} to exclude stations that are not located within
+the region given in the \texttt{Par\_file} (between \texttt{LATITUDE\_MIN}
+and \texttt{LATITUDE\_MAX} and between \texttt{LONGITUDE\_MIN} and
+\texttt{LONGITUDE\_MAX}). The filtered file is called \texttt{DATA/STATIONS\_FILTERED}.
+
+Solver output is provided in the \texttt{OUTPUT\_FILES} directory
 in the \texttt{output\_solver.txt} file. Output can be directed to
 the screen instead by uncommenting a line in \texttt{constants.h}:
-
 \begin{lyxcode}
@@
 While the solver is running, its progress may be tracked by monitoring
-the `\texttt{\small timestamp{*}}' files in the ~\\
-\texttt{\small in\_out\_files/OUTPUT\_FILES} directory.
-These tiny files look something like this:
-
+the `\texttt{\small timestamp{*}}{\small ' files in the ~}\\
+{\small{} }\texttt{\small OUTPUT\_FILES}{\small{} directory. These tiny
+files look something like this:}{\small \par}
 \begin{lyxcode}
@@
 
-For the same model, one can rerun the solver for different events by
-simply changing the \texttt{CMTSOLUTION} or \texttt{FORCESOLUTION}
+For the same model, one can rerun the solver for different events
+by simply changing the \texttt{CMTSOLUTION} or \texttt{FORCESOLUTION}
 file, or for different stations by changing the \texttt{STATIONS}
 file. There is no need to rerun the \texttt{xgenerate\_databases}
-executable. Of course it is best to include as many stations as
-possible, since this does not add to the cost of the simulation.
-
-\chapter{\label{cha:Adjoint-Simulations}Adjoint Simulations}
-
-Adjoint simulations are generally performed for two distinct applications.
-First, they can be used for earthquake source inversions, especially
-earthquakes with large ruptures such as the Lander's earthquake \citep{WaHe94}.
-Second, they can be used to generate finite-frequency sensitivity
-kernels that are a critical part of tomographic inversions based upon
-3D reference models \citep{TrTaLi05,LiTr06,TrKoLi08,LiTr08}. In either
-case, source parameter or velocity structure updates are sought to
-minimize a specific misfit function (e.g., waveform or traveltime
-differences), and the adjoint simulation provides a means of computing
-the gradient of the misfit function and further reducing it in successive
-iterations. Applications and procedures pertaining to source studies
-and finite-frequency kernels are discussed in Sections~\ref{sec:Adjoint-simulation-sources}
-and \ref{sec:Adjoint-simulation-finite}, respectively. The two related
-parameters in the \texttt{Par\_file} are \texttt{SIMULATION\_TYPE}
-(1 or 2) and the \texttt{SAVE\_FORWARD} (boolean).
-
-\section{\label{sec:Adjoint-simulation-sources}Adjoint Simulations for Sources}
-
-In the case where a specific misfit function is minimized to invert
-for the earthquake source parameters, the gradient of the misfit function
-with respect to these source parameters can be computed by placing
-time-reversed seismograms at the receivers and using them as sources
-in an adjoint simulation, and then the value of the gradient is obtained
-from the adjoint seismograms recorded at the original earthquake location.
-
-\begin{enumerate}
-\item \textbf{Prepare the adjoint sources} \label{enu:Prepare-the-adjoint}
-
-\begin{enumerate}
-\item First, run a regular forward simlation (\texttt{SIMULATION\_TYPE =
-1} and \texttt{SAVE\_FORWARD = .false.}). You can automatically set
-these two variables using the \texttt{\small utils/change\_simulation\_type.pl}
-script:
-
-\begin{lyxcode}
-utils/change\_simulation\_type.pl~-f~
-\end{lyxcode}
-and then collect the recorded seismograms at all the stations given
-in \texttt{in\_data\_files/STATIONS}.
-
-\item Then select the stations for which you want to compute the time-reversed
-adjoint sources and run the adjoint simulation, and compile them into
-the \texttt{in\_data\_files/STATIONS\_ADJOINT} file, which has the same format
-as the regular \texttt{in\_data\_files/STATIONS} file.
-
-\begin{itemize}
-\item Depending on what type of misfit function is used for the source inversion,
-adjoint sources need to be computed from the original recorded seismograms
-for the selected stations and saved in the \texttt{in\_out\_files/SEM/} directory
-with the format \texttt{STA.NT.BX?.adj}, where \texttt{STA}, \texttt{NT}
-are the station name and network code given in the \texttt{in\_data\_files/STATIONS\_ADJOINT}
-file, and \texttt{BX?} represents the component name of a particular
-adjoint seismogram. Please note that the band code can change depending on
-your sampling rate (see Appendix~\ref{cha:channel-codes} for further details).
-\item The adjoint seismograms are in the same format as the original seismogram
-(\texttt{STA.NT.BX?.sem?}), with the same start time, time interval
-and record length.
-\end{itemize}
-\item Notice that even if you choose to time reverse only one component
-from one specific station, you still need to supply all three components
-because the code is expecting them (you can set the other two components
-to be zero).
-\item Also note that since time-reversal is done in the code itself, no
-explicit time-reversing is needed for the preparation of the adjoint
-sources, i.e., the adjoint sources are in the same forward time sense
-as the original recorded seismograms.
-\end{enumerate}
-\item \textbf{Set the related parameters and run the adjoint simulation}\\
-In the \texttt{in\_data\_files/Par\_file}, set the two related parameters to
-be \texttt{SIMULATION\_TYPE = 2} and \texttt{SAVE\_FORWARD = .false.}.
-More conveniently, use the scripts \texttt{utils/change\_simulation\_type.pl}
-to modify the \texttt{Par\_file} automatically (\texttt{change\_simulation\_type.pl
--a}). Then run the solver to launch the adjoint simulation.
-\item \textbf{Collect the seismograms at the original source location}
-
-After the adjoint simulation has completed successfully, collect the
-seismograms from \texttt{LOCAL\_PATH}.
-
-\begin{itemize}
-\item These adjoint seismograms are recorded at the locations of the original
-earthquake sources given by the \texttt{in\_data\_files/CMTSOLUTION} file, and
-have names of the form \texttt{S?????.NT.S??.sem} for the six-component
-strain tensor (\texttt{SNN,SEE,SZZ,SNE,SNZ,SEZ}) at these locations,
-and ~\\
-\texttt{S?????.NT.BX?.sem} for the three-component displacements
-(\texttt{BXN,BXE,BXZ}) recorded at these locations.
-\item \texttt{S?????} denotes the source number; for example, if the original
-\texttt{CMTSOLUTION} provides only a point source, then the seismograms
-collected will start with \texttt{S00001}.
-\item These adjoint seismograms provide critical information for the computation
-of the gradient of the misfit function.
-\end{itemize}
-\end{enumerate}
-
-\section{\label{sec:Adjoint-simulation-finite}Adjoint Simulations for Finite-Frequency
-Kernels (Kernel Simulation)}
-
-Finite-frequency sensitivity kernels are computed in two successive
-simulations (please refer to \citet{LiTr06} and \citet{TrKoLi08} for details).
-
-\begin{enumerate}
-\item \textbf{Run a forward simulation with the state variables saved at
-the end of the simulation}
-
-Prepare the \texttt{\small CMTSOLUTION} and \texttt{\small STATIONS}
-files, set the parameters \texttt{\small SIMULATION\_TYPE}{\small{}
-}\texttt{\small =}{\small{} }\texttt{\small 1} and \texttt{\small SAVE\_FORWARD
-=}{\small{} }\texttt{\small .true.} in the \texttt{Par\_file} (\texttt{change\_simulation\_type
--F}), and run the solver.
-
-\begin{itemize}
-\item Notice that attenuation is not implemented yet for the computation
-of finite-frequency kernels; therefore set \texttt{ATTENUATION = .false.}
-in the \texttt{Par\_file}.
-\item We also suggest you modify the half duration of the \texttt{CMTSOLUTION}
-to be similar to the accuracy of the simulation (see Equation \ref{eq:shortest_period})
-to avoid too much high-frequency noise in the forward wavefield, although
-theoretically the high-frequency noise should be eliminated when convolved
-with an adjoint wavefield with the proper frequency content.
-\item This forward simulation differs from the regular simulations (\texttt{\small SIMULATION\_TYPE}{\small{}
-}\texttt{\small =}{\small{} }\texttt{\small 1} and \texttt{\small SAVE\_FORWARD}{\small{}
-}\texttt{\small =}{\small{} }\texttt{\small .false.}) described in
-the previous chapters in that the state variables for the last time
-step of the simulation, including wavefields of the displacement,
-velocity, acceleration, etc., are saved to the \texttt{LOCAL\_PATH}
-to be used for the subsequent simulation.
-\item For regional simulations, the files recording the absorbing boundary
-contribution are also written to the \texttt{LOCAL\_PATH} when \texttt{SAVE\_FORWARD
-= .true.}.
-\end{itemize}
-\item \textbf{Prepare the adjoint sources}
-
-The adjoint sources need to be prepared the same way as described
-in the Section \ref{enu:Prepare-the-adjoint}.
-
-\begin{itemize}
-\item In the case of travel-time finite-frequency kernel for one source-receiver
-pair, i.e., point source from the \texttt{CMTSOLUTION}, and one station
-in the \texttt{STATIONS\_ADJOINT} list, we supply a sample program
-in \texttt{utils/adjoint\_sources/traveltime/xcreate\_adjsrc\_traveltime} to cut a certain portion of the original
-displacement seismograms and convert it into the proper adjoint source
-to compute the finite-frequency kernel.
-\begin{lyxcode}
-xcreate\_adjsrc\_traveltime~t1~t2~ifile{[}0-5]~E/N/Z-ascii-files~{[}baz]
-\end{lyxcode}
-where \texttt{t1} and \texttt{t2} are the start and end time of the
-portion you are interested in, \texttt{ifile} denotes the component
-of the seismograms to be used (0 for all three components, 1 for East,
-2 for North, and 3 for vertical, 4 for transverse, and 5 for radial
-component), \texttt{E/N/Z-ascii-files} indicate the three-component
-displacement seismograms in the right order, and \texttt{baz} is the
-back-azimuth of the station. Note that \texttt{baz} is only supplied
-when \texttt{ifile} = 4 or 5.
-
-\item Similarly, a sample program to compute adjoint sources for amplitude finite-frequency kernels
-may be found in \texttt{utils/adjoint\_sources/amplitude} and used in the same way as described for
-traveltime measurements
-\begin{lyxcode}
-xcreate\_adjsrc\_amplitude~t1~t2~ifile{[}0-5]~E/N/Z-ascii-files~{[}baz].
-\end{lyxcode}
-
-\end{itemize}
-
-\item \textbf{Run the kernel simulation}
-
-With the successful forward simulation and the adjoint source ready
-in the \texttt{in\_out\_files/SEM/} directory, set \texttt{SIMULATION\_TYPE = 3} and \texttt{SAVE\_FORWARD
-= .false.} in the \texttt{Par\_file} (you can use \texttt{change\_simulation\_type.pl -b}),
-and rerun the solver.
-
-\begin{itemize}
-\item The adjoint simulation is launched together with the back reconstruction
-of the original forward wavefield from the state variables saved from
-the previous forward simulation, and the finite-frequency kernels
-are computed by the interaction of the reconstructed forward wavefield
-and the adjoint wavefield.
-\item The back-reconstructed seismograms at the original station locations
-are saved to the \texttt{LOCAL\_PATH} at the end of the kernel simulations,
-and can be collected to the local disk.
-\item These back-constructed seismograms can be compared with the time-reversed
-original seismograms to assess the accuracy of the backward reconstruction,
-and they should match very well.
-\item The arrays for density, P-wave speed and S-wave speed kernels are
-also saved in the \texttt{LOCAL\_PATH} with the names \texttt{proc??????\_rho(alpha,beta)\_kernel.bin},
-where \texttt{proc??????} represents the processor number, \texttt{rho(alpha,beta)}
-are the different types of kernels.
-\end{itemize}
-\end{enumerate}
-In general, the three steps need to be run sequentially to assure
-proper access to the necessary files. If the simulations are run through
-some cluster scheduling system (e.g., LSF), and the forward simulation
-and the subsequent kernel simulations cannot be assigned to the same
-set of computer nodes, the kernel simulation will not be able to access
-the database files saved by the forward simulation. Solutions for
-this dilemma are provided in Chapter~\ref{cha:Scheduler}. Visualization
-of the finite-frequency kernels is discussed in Section~\ref{sec:Finite-Frequency-Kernels}.
-
-\chapter{Noise Cross-correlation Simulations}
-
-Besides earthquake simulations, SPECFEM3D includes functionality for seismic noise tomography as well.
-In order to proceed successfully in this chapter, it is critical that you have already familiarized yourself
-with procedures for meshing (Chapter \ref{cha:Mesh-Generation}),
-creating distributed databases (Chapter \ref{cha:Creating-Distributed-Databases}),
-running earthquake simulations (Chapters \ref{cha:Running-the-Solver})
-and adjoint simulations (Chapter \ref{cha:Adjoint-Simulations}).
-Also, make sure you read the article `Noise cross-correlation sensitivity kernels' \citep{trompetal2010},
-in order to understand noise simulations from a theoretical perspective.
-
-\section{Input Parameter Files}
-
-As usual, the three main input files are crucial: \texttt{Par\_file}, \texttt{CMTSOLUTION} and \texttt{STATIONS}.
-Unless otherwise specified, those input files should be located in directory \texttt{in\_data\_files/}.\\
-
-\texttt{CMTSOLUTION} is required for all simulations.
-At a first glance, it may seem unexpected to have it here,
-since the noise simulations should have nothing to do with the earthquake -- \texttt{CMTSOLUTION}.
-However, for noise simulations, it is critical to have no earthquakes.
-In other words, the moment tensor specified in \texttt{CMTSOLUTION} must be set to zero manually!\\
-
-\texttt{STATIONS} remains the same as in previous earthquake simulations,
-except that the order of receivers listed in \texttt{STATIONS} is now important.
-The order will be used to determine the `master' receiver,
-i.e., the one that simultaneously cross correlates with the others.\\
-
-\texttt{Par\_file} also requires careful attention.
-A parameter called \texttt{NOISE\_TOMOGRAPHY} has been added which specifies the type of simulation to be run.
-\texttt{NOISE\_TOMOGRAPHY} is an integer with possible values 0, 1, 2 and 3.
-For example, when \texttt{NOISE\_TOMOGRAPHY} equals 0, a regular earthquake simulation will be run.
-When it is 1/2/3, you are about to run
-step 1/2/3 of the noise simulations respectively.
-Should you be confused by the three steps, refer to \citet{trompetal2010} for details.\\
-
-Another change to \texttt{Par\_file} involves the parameter \texttt{NSTEP}.
-While for regular earthquake simulations this parameter specifies the length of synthetic seismograms generated,
-for noise simulations it specifies the length of the seismograms used to compute cross correlations.
-The actual cross correlations are thus twice this length, i.e., $2~\mathrm{NSTEP}-1$.
-The code automatically makes the modification accordingly, if \texttt{NOISE\_TOMOGRAPHY} is not zero.\\
-
-There are other parameters in \texttt{Par\_file} which should be given specific values.
-For instance, since the first two steps for calculating noise sensitivity kernels correspond to
-forward simulations, \texttt{SIMULATION\_TYPE} must be 1 when \texttt{NOISE\_TOMOGRAPHY} equals 1 or 2.
-Also, we have to reconstruct the ensemble forward wavefields in adjoint simulations,
-therefore we need to set \texttt{SAVE\_FORWARD} to \texttt{.true.} for the second step,
-i.e., when \texttt{NOISE\_TOMOGRAPHY} equals 2.
-The third step is for kernel constructions. Hence \texttt{SIMULATION\_TYPE} should be 3, whereas
-\texttt{SAVE\_FORWARD} must be \texttt{.false.}.\\
-
-Finally, for most system architectures,
-please make sure that \texttt{LOCAL\_PATH} in \texttt{Par\_file} is in fact local, not globally shared.
-Because we have to save the wavefields at the earth's surface at every time step,
-it is quite problematic to have a globally shared \texttt{LOCAL\_PATH},
-in terms of both disk storage and I/O speed.
-
-\section{Noise Simulations: Step by Step}
-
-Proper parameters in those parameter files are not enough for noise simulations to run.
-We have more parameters to specify:
-for example, the ensemble-averaged noise spectrum, the noise distribution etc.
-However, since there are a few `new' files, it is better to introduce them sequentially.
-In this section, standard procedures for noise simulations are described.
-
-\subsection{Pre-simulation}
-
-\begin{itemize}
-
-\item
-As usual, we first configure the software package using:
-
-\texttt{./configure FC=ifort MPIFC=mpif90}
-
-Use the following if SCOTCH is needed:
-
-\texttt{./configure FC=ifort MPIFC=mpif90 --with-scotch-dir=/opt/scotch}\\
-
-\item
-Next, we need to compile the source code using:
-
-\texttt{make xgenerate\_databases}
-
-\texttt{make xspecfem3D} \\
-
-\item
-Before we can run noise simulations, we have to specify the noise statistics,
-e.g., the ensemble-averaged noise spectrum.
-Matlab scripts are provided to help you to generate the necessary file.
-
-\texttt{examples/noise\_tomography/NOISE\_TOMOGRAPHY.m}  (main program)\\
-\texttt{examples/noise\_tomography/PetersonNoiseModel.m}
-
-In Matlab, simply run:
-
-\texttt{NOISE\_TOMOGRAPHY(NSTEP, DT, Tmin, Tmax, NOISE\_MODEL)}\\
-
-\texttt{DT} is given in \texttt{Par\_file},
-but \texttt{NSTEP} is NOT the one specified in \texttt{Par\_file}.
-Instead, you have to feed $2~\mathrm{NSTEP}-1$ to account for the doubled length of cross correlations.
-\texttt{Tmin} and \texttt{Tmax} correspond to the period range you are interested in,
-whereas \texttt{NOISE\_MODEL} denotes the noise model you will be using.
-Details can be found in the Matlab script.
-
-After running the Matlab script, you will be given the following information (plus a figure in Matlab):
-
-\texttt{*************************************************************}\\
-\texttt{the source time function has been saved in:}\\
-\texttt{/data2/yangl/3D\_NOISE/S\_squared} (note this path must be different)\\
-\texttt{S\_squared should be put into directory:}\\
-\texttt{in\_out\_files/NOISE\_TOMOGRAPHY/ in the SPECFEM3D package}\\
-
-In other words, the Matlab script creates a file called \texttt{S\_squared},
-which is the first `new' input file we encounter for noise simulations.
-
-One may choose a flat noise spectrum rather than Peterson's noise model.
-This can be done easily by modifying the Matlab script a little.
-
-\item
-Create a new directory in the SPECFEM3D package, name it as \texttt{in\_out\_files/NOISE\_TOMOGRAPHY/}.
-We will add some parameter files later in this folder.
-
-\item
-Put the Matlab-generated-file \texttt{S\_squared} in \texttt{in\_out\_files/NOISE\_TOMOGRAPHY/}.
-
-That's to say, you will have a file \texttt{in\_out\_files/NOISE\_TOMOGRAPHY/S\_squared} in the SPECFEM3D package.
-
-\item
-Create a file called \texttt{in\_out\_files/NOISE\_TOMOGRAPHY/irec\_master\_noise}. Note that this file is located
-in directory \texttt{in\_out\_files/NOISE\_TOMOGRAPHY/} as well. In general, all noise simulation related parameter
-files go into that directory.
-\texttt{irec\_master\_noise} contains only one integer, which is the ID
-of the `master' receiver. For example, if this file contains 5, it means that the fifth receiver
-listed in \texttt{in\_data\_files/STATIONS} becomes the `master'. That's why we mentioned previously that
-the order of receivers in \texttt{in\_data\_files/STATIONS} is important.
-
-Note that in regional simulations, the \texttt{in\_data\_files/STATIONS}
-might contain receivers which are outside of our computational domains.
-Therefore, the integer in \texttt{irec\_master\_noise}
-is actually the ID in \texttt{in\_data\_files/STATIONS\_FILTERED}
-(which is generated by \texttt{bin/xgenerate\_databases}).
-
-\item
-Create a file called \texttt{in\_out\_files/NOISE\_TOMOGRAPHY/nu\_master}. This file holds three numbers,
-forming a (unit) vector. It describes which component we are cross-correlating at the `master'
-receiver, i.e., ${\hat{\bf \nu}}^{\alpha}$ in \citet{trompetal2010}.
-The three numbers correspond to E/N/Z components respectively.
-Most often, the vertical component is used, and in those cases the three numbers should be 0, 0 and 1.
-
-\item
-Describe the noise direction and distributions in \texttt{src/shared/noise\_tomography.f90}.
-Search for a subroutine called \texttt{noise\_distribution\_direction}
-in \texttt{noise\_tomography.f90}. It is actually located at the very beginning
-of \texttt{noise\_tomography.f90}. The default assumes vertical noise and a uniform distribution
-across the whole free surface of the model. It should be quite self-explanatory for modifications.
-Should you modify this part, you have to re-compile the source code.
-
-\end{itemize}
-
-\subsection{Simulations}
-
-With all of the above done, we can finally launch our simulations.
-Again, please make sure that the \texttt{LOCAL\_PATH} in \texttt{in\_data\_files/Par\_file} is not globally shared.
-It is quite problematic to have a globally shared \texttt{LOCAL\_PATH}, in terms of both disk storage and
-speed of I/O (we have to save the wavefields at the earth's surface at every time step).
-
-As discussed in \citet{trompetal2010}, it takes three steps/simulations to obtain one contribution
-of the ensemble sensitivity kernels:
-
-\begin{itemize}
-\item
-Step 1: simulation for generating wavefields\\
-\texttt{SIMULATION\_TYPE=1}\\
-\texttt{NOISE\_TOMOGRAPHY=1}\\
-\texttt{SAVE\_FORWARD} not used, can be either .true. or .false.
-
-\item
-Step 2: simulation for ensemble forward wavefields\\
-\texttt{SIMULATION\_TYPE=1}\\
-\texttt{NOISE\_TOMOGRAPHY=2}\\
-\texttt{SAVE\_FORWARD=.true.}
-
-\item
-Step 3: simulation for ensemble adjoint wavefields and sensitivity kernels\\
-\texttt{SIMULATION\_TYPE=3}\\
-\texttt{NOISE\_TOMOGRAPHY=3}\\
-\texttt{SAVE\_FORWARD=.false.}
-
-Note Step 3 is an adjoint simulation, please refer to previous chapters on how to prepare adjoint sources
-and other necessary files, as well as how adjoint simulations work.
-
-\end{itemize}
-
-Note that it is better to run the three steps continuously within the same job,
-otherwise you have to collect the saved surface movies from the old nodes/CPUs to the new nodes/CPUs.
-This process varies from cluster to cluster and thus cannot be discussed here.
-Please ask your cluster administrator for information/configuration of the cluster you are using.\\
-
-\subsection{Post-simulation}
-
-After those simulations, you have all stuff you need, either in the \texttt{in\_out\_files/OUTPUT\_FILES/} or
-in the directory specified by \texttt{LOCAL\_PATH} in \texttt{in\_data\_files/Par\_file}
-(which are most probably on local nodes).
-Collect whatever you want from the local nodes to your workstation, and then visualize them.
-This process is the same as what you may have done for regular earthquake simulations.
-Refer to other chapters if you have problems.\\
-
-Simply speaking, two outputs are the most interesting: the simulated ensemble cross correlations and
-one contribution of the ensemble sensitivity kernels.\\
-
-The simulated ensemble cross correlations are obtained after the second simulation (Step 2).
-Seismograms in \texttt{in\_out\_files/OUTPUT\_FILES/} are actually the simulated ensemble cross correlations.
-Collect them immediately after Step 2, or the Step 3 will overwrite them.
-Note that we have a `master' receiver specified by \texttt{irec\_master\_noise},
-the seismogram at one station corresponds to the cross correlation between that station and the `master'.
-Since the seismograms have three components, we may obtain cross correlations
-for different components as well, not necessarily the cross correlations between vertical components.\\
-
-One contribution of the ensemble cross-correlation sensitivity kernels are obtained after Step 3,
-stored in the \texttt{in\_data\_files/LOCAL\_PATH} on local nodes.
-The ensemble kernel files are named the same as regular earthquake kernels.\\
-
-You need to run another three simulations to get the other contribution of the ensemble kernels,
-using different forward and adjoint sources given in \citet{trompetal2010}.
-
-\section{Example}
-
-In order to illustrate noise simulations in an easy way, one example is provided in \texttt{examples/noise\_tomography/}.
-See \texttt{examples/noise\_tomography/README} for explanations. \\
-
-Note, however, that they are created for a specific workstation (CLOVER@PRINCETON),
-which has at least 4 cores with `mpif90' working properly. \\
-
-If your workstation is suitable, you can run the example in \texttt{examples/noise\_tomography/} using:\\
-
-\texttt{./pre-processing.sh}\\
-
-Even if this script does not work on your workstation, the procedure it describes is universal.
-You may review the whole process described in the last section by following the commands in \texttt{pre-processing.sh},
-which should contain enough explanations for all the commands.
-
-\chapter{Graphics}
-
-\section{\label{sec:Mesh-graphics}Meshes}
-
-In case you used the internal mesher \texttt{xmeshfem3D} to create and partition your mesh,
-you can output mesh files in ABAQUS (.INP) and DX (.dx) format to visualize them. For this, you must set either the
-flag \texttt{CREATE\_DX\_FILES} or \texttt{CREATE\_ABAQUS\_FILES} to \texttt{.true.}
-in the mesher's parameter file \texttt{Mesh\_Par\_file}
-prior to running the mesher (see Chapter~\ref{cha:Running-the-Mesher-Meshfem3D} for details).
-You can then use AVS \urlwithparentheses{www.avs.com} or OpenDX \urlwithparentheses{www.opendx.org}
-to visualize the mesh and MPI partition (slices).
-
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=.49\textwidth]{figures/vtk_mesh_vp.jpg}
-\includegraphics[width=.49\textwidth]{figures/vtk_mesh_vs.jpg}
-\par\end{centering}
-
-\caption{Visualization using Paraview of VTK files created by \texttt{xgenerate\_databases} showing P- and S-wave velocities assigned to the mesh points.
-The mesh was created by \texttt{xmeshfem3D} for 4 processors.}
-
-\label{fig:vtk.mesh}
-\end{figure}
-
-You have also the option to visualize the distributed databases produced by \texttt{xgenerate\_databases}
-using Paraview \urlwithparentheses{www.paraview.org}. For this, you must set the flag \texttt{SAVE\_MESH\_FILES} to \texttt{.true.}
-in the main parameter file \texttt{Par\_file} (see Chapter~\ref{cha:Main-Parameter} for details).
-This will create VTK files for each single partition. You can then use Paraview \urlwithparentheses{www.paraview.org} to visualized these partitions.
-
-\section{\label{sec:Movies}Movies}
-
-To make a surface or volume movie of the simulation, set parameters
-\texttt{MOVIE\_SURFACE}, \texttt{MOVIE\_VOLUME}, and \texttt{NTSTEP\_BETWEEN\_FRAMES}
-in the \texttt{Par\_file}. Turning on the movie flags, in particular
-\texttt{MOVIE\_VOLUME}, produces large output files. \texttt{MOVIE\_VOLUME}
-files are saved in the \texttt{LOCAL\_PATH} directory, whereas \texttt{MOVIE\_SURFACE}
-output files are saved in the \texttt{in\_out\_files/OUTPUT\_FILES} directory. We
-save the displacement field if the parameter \texttt{SAVE\_DISPLACEMENT} is set, otherwise the velocity field is saved.
-The look of a movie is determined by the
-half-duration of the source. The half-duration should be large enough
-so that the movie does not contain frequencies that are not resolved
-by the mesh, i.e., it should not contain numerical noise. This can
-be accomplished by selecting a CMT \texttt{HALF\_DURATION} > 1.1 $\times$
-smallest period (see figure \ref{fig:CMTSOLUTION-file}). When \texttt{MOVIE\_SURFACE}
-= .\texttt{true.}, the half duration of each source in the \texttt{CMTSOLUTION}
-file is replaced by
-
-\begin{quote}
-\[
-\sqrt{(}\mathrm{\mathtt{HALF\_DURATIO}\mathtt{N}^{2}}+\mathrm{\mathtt{HDUR\_MOVI}\mathtt{E}^{2}})\]
-\textbf{NOTE:} If \texttt{HDUR\_MOVIE} is set to 0.0, the code will
-select the appropriate value of 1.1 $\times$ smallest period. As
-usual, for a point source one can set \texttt{half duration} in the
-\texttt{CMTSOLUTION} file to be 0.0 and \texttt{HDUR\_MOVIE} = 0.0 in the \texttt{Par\_file} to get
-the highest frequencies resolved by the simulation, but for a finite
-source one would keep all the \texttt{half durations} as prescribed
-by the finite source model and set \texttt{HDUR\_MOVIE} = 0.0.
-\end{quote}
-
-\subsection{Movie Surface}
-
-When running \texttt{xspecfem3D} with the \texttt{MOVIE\_SURFACE}
-flag turned on, the code outputs \texttt{moviedata??????} files in
-the \texttt{in\_out\_files/OUTPUT\_FILES} directory. There are several flags
-in the main parameter file \texttt{Par\_file} that control the output of these moviedata files
-(see section \ref{cha:Main-Parameter} for details):
-\texttt{\small NTSTEP\_BETWEEN\_FRAMES} to set the timesteps between frames,
-\texttt{\small SAVE\_DISPLACEMENT} to save displacement instead of velocity,
-\texttt{USE\_HIGHRES\_FOR\_MOVIES} to save values at all GLL point instead of element edges.
-In order to output additionally shakemaps, you would set
-the parameter \texttt{CREATE\_SHAKEMAP} to \texttt{.true.}.
-
-The files are in a fairly complicated
-binary format, but there is a program provided to convert the
-output into more user friendly formats:
-\begin{description}
-\item [\texttt{xcreate\_movie\_shakemap\_AVS\_DX\_GMT}]
-From \texttt{create\_movie\_shakemap\_AVS\_DX\_GMT.f90}, it
-outputs data in ASCII, OpenDX, or AVS format (also readable in ParaView).
-Before compiling the code, make sure you have the file \texttt{surface\_from\_mesher.h}
-in the  \texttt{in\_out\_files/OUTPUT\_FILES/} directory. This file will be created by the solver run.
-Then type
-\begin{lyxcode}
-{\footnotesize make xcreate\_movie\_shakemap\_AVS\_DX\_GMT}{\footnotesize \par}
-\end{lyxcode}
-and run the executable \texttt{xcreate\_movie\_shakemap\_AVS\_DX\_GMT}
-in the \texttt{bin/} subdirectory. It will create visualization files in your format of choice. The code
-will prompt the user for input parameters.
-
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=.32\textwidth]{figures/movie_surf_1.jpg}
-\includegraphics[width=.32\textwidth]{figures/movie_surf_2.jpg}
-\includegraphics[width=.32\textwidth]{figures/movie_surf_3.jpg}
-\par\end{centering}
-
-\caption{Visualization using AVS files created by \texttt{xcreate\_movie\_shakemap\_AVS\_DX\_GMT} showing
-movie snapshots of vertical velocity components at different times.}
-
-\label{fig:movie.surf}
-\end{figure}
-
-\end{description}
-
-The \texttt{SPECFEM3D} code is running in near real-time to produce
-animations of southern California earthquakes via the web; see Southern California ShakeMovie\textregistered \urlwithparentheses{www.shakemovie.caltech.edu}.
-
-\subsection{\label{sub:Movie-Volume}Movie Volume}
-
-When running xspecfem3D with the \texttt{\small MOVIE\_VOLUME} flag
-turned on, the code outputs several files in \texttt{\small LOCAL\_PATH} specified in the main
-\texttt{Par\_file}, e.g. in directory \texttt{in\_out\_files/DATABASES\_MPI}.
-%As the files can be very large, there are several flags in the \texttt{\small Par\_file}
-%that control the region in space and time that is saved. These are:
-%\texttt{\small NTSTEP\_BETWEEN\_FRAMES}
-The output is saved by each processor
-at the time interval specified by \texttt{\small NTSTEP\_BETWEEN\_FRAMES}.
-For all domains, the velocity field is output to files:
-\texttt{\small proc??????\_velocity\_X\_it??????.bin},
-\texttt{\small proc??????\_velocity\_Y\_it??????.bin} and
-\texttt{\small proc??????\_velocity\_Z\_it??????.bin}.
-For elastic domains, the divergence and curl taken from the velocity field, i.e. $\nabla \cdot \bf{v}$
-and $\nabla \times \bf{v}$, get stored as well:
-\texttt{\small proc??????\_div\_it??????.bin},
-\texttt{\small proc??????\_curl\_X\_t??????.bin},
-\texttt{\small proc??????\_curl\_Y\_it??????.bin} and
-\texttt{\small proc??????\_curl\_Z\_it??????.bin}.
-The files denoted
-\texttt{\small proc??????\_div\_}~\\
-\texttt{\small glob\_it??????.bin} and
-\texttt{\small proc??????\_curl\_glob\_it??????.bin}
-are stored on the global points only, all the other arrays
-are stored on all GLL points. Note that the components X/Y/Z can change to E/N/Z according to
-the \texttt{SUPPRESS\_UTM\_PROJECTION} flag (see also Appendix~\ref{cha:Coordinates} and \ref{cha:channel-codes}).
-
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=.32\textwidth]{figures/movie_volume_1.jpg}
-\includegraphics[width=.32\textwidth]{figures/movie_volume_2.jpg}
-\includegraphics[width=.32\textwidth]{figures/movie_volume_3.jpg}
-\par\end{centering}
-
-\caption{Paraview visualization using movie volume files (converted by \texttt{xcombine\_vol\_data} and \texttt{mesh2vtu.pl}) and showing
-snapshots of vertical velocity components at different times.}
-
-\label{fig:movie.volume}
-\end{figure}
-
-To visualize these files, we use an auxilliary program \texttt{combine\_vol\_data.f90}
-to combine the data from all slices into one mesh file. To compile it in the root directory, type:
-\begin{lyxcode}
-{\footnotesize make~xcombine\_vol\_data~}{\footnotesize \par}
-\end{lyxcode}
-which will create the executable \texttt{xcombine\_vol\_data} in the directory \texttt{bin/}. To output the
-usage of this executable, type `./bin/xcombine\_vol\_data` without arguments.
-As an example, to run the executable you would use
-\begin{lyxcode}
-{\footnotesize cd~bin/~}{\footnotesize \par}
-{\footnotesize ./xcombine\_vol\_data~0~3~velocity\_Z\_it000400~../in\_out\_files/DATABASES\_MPI~../in\_out\_files/OUTPUT\_FILES~0~}{\footnotesize \par}
-\end{lyxcode}
-to create a low-resolution mesh file. The output mesh file will have the name \texttt{velocity\_Z\_it000400.mesh}.
-We next convert the \texttt{.mesh} file into the VTU (Unstructured
-grid file) format which can be viewed in ParaView. For this task, you can use and modify the script \texttt{mesh2vtu.pl}
-located in directory \texttt{utils/Visualization/Paraview/}, for example:
-\begin{lyxcode}
-{\footnotesize mesh2vtu.pl~-i~velocity\_Z\_it000400.mesh~-o~velocity\_Z\_it000400.vtu}{\footnotesize \par}
-\end{lyxcode}
-Notice that this Perl script uses a program \texttt{mesh2vtu} in the
-\texttt{utils/Visualization/Paraview/mesh2vtu} directory, which further uses the VTK \urlwithparentheses{http://www.vtk.org/}
-run-time library for its execution. Therefore, make sure you have
-them properly set in the script according to your system.
-
-\section{\label{sec:Finite-Frequency-Kernels}Finite-Frequency Kernels}
-
-The finite-frequency kernels computed as explained in Section \ref{sec:Adjoint-simulation-finite}
-are saved in the \texttt{LOCAL\_PATH} at the end of the simulation.
-Therefore, we first need to collect these files on the front end,
-combine them into one mesh file, and visualize them with some auxilliary
-programs.
-
-\begin{enumerate}
-\item \textbf{Create slice files}
-
-We will only discuss the case of one source-receiver pair, i.e., the
-so-called banana-doughnut kernels. Although it is possible to collect
-the kernel files from all slices on the front end, it usually takes
-up too much storage space (at least tens of gigabytes). Since the
-sensitivity kernels are the strongest along the source-receiver great
-circle path, it is sufficient to collect only the slices that are
-along or close to the great circle path.
-
-A Perl script \texttt{slice\_number.pl} located in directory \texttt{utils/Visualization/Paraview/} can
-help to figure out the slice numbers that lie along the great circle
-path. It applies to meshes created with the internal mesher \texttt{xmeshfem3D}.
-
-\begin{enumerate}
-\item On machines where you have access to the script, copy the \texttt{Mesh\_Par\_file}, and
-\texttt{output\_solver} files, and run:
-
-\begin{lyxcode}
-{\small slice\_number.pl~Mesh\_Par\_file~output\_solver.txt~slice\_file}{\small \par}
-\end{lyxcode}
-which will generate a \texttt{slices\_file}.
-
-\item For cases with multiple sources and multiple receivers, you need to
-provide a slice file before proceeding to the next step.
-\end{enumerate}
-\item \textbf{Collect the kernel files}
-
-After obtaining the slice files, you can collect the corresponding
-kernel files from the given slices.
-
-\begin{enumerate}
-\item You can use or modify the script \texttt{utils/copy\_basin\_database.pl}
-to accomplish this:
-
-\begin{lyxcode}
-{\small utils/copy\_database.pl~slice\_file~lsf\_machine\_file~filename~{[}jobid]}{\small \par}
-\end{lyxcode}
-where \texttt{\small lsf\_machine\_file} is the machine file generated
-by the LSF scheduler, \texttt{\small filename} is the kernel name
-(e.g., \texttt{\small rho\_kernel}, \texttt{\small alpha\_kernel}
-and \texttt{\small beta\_kernel}), and the optional \texttt{\small jobid}
-is the name of the subdirectory under \texttt{\small LOCAL\_PATH}
-where all the kernel files are stored.
-
-\item After executing this script, all the necessary mesh topology files
-as well as the kernel array files are collected to the local directory
-of the front end.
-\end{enumerate}
-\item \textbf{Combine kernel files into one mesh file}
-
-We use an auxilliary program \texttt{combine\_vol\_data.f90}
-to combine the kernel files from all slices into one mesh file.
-
-\begin{enumerate}
-\item Compile it in the root directory:
-
-\begin{lyxcode}
-{\footnotesize make~xcombine\_vol\_data~}{\footnotesize \par}
-
-{\footnotesize ./bin/xcombine\_vol\_data~slice\_list~filename~input\_dir~output\_dir~high/low-resolution~}{\footnotesize \par}
-\end{lyxcode}
-where \texttt{input\_dir} is the directory where all the individual
-kernel files are stored, and \texttt{output\_dir} is where the mesh
-file will be written.
-
-\item Use 1 for a high-resolution mesh, outputting all the GLL points to
-the mesh file, or use 0 for low resolution, outputting only the corner
-points of the elements to the mesh file.
-\item The output mesh file will have the name \texttt{filename\_rho(alpha,beta).mesh}
-\end{enumerate}
-\item \textbf{Convert mesh files into .vtu files}
-
-\begin{enumerate}
-\item We next convert the \texttt{.mesh} file into the VTU (Unstructured
-grid file) format which can be viewed in ParaView. For this task, you can use and modify the script \texttt{mesh2vtu.pl}
-located in directory \texttt{utils/Visualization/Paraview/}, for example:
-
-\begin{lyxcode}
-{\footnotesize mesh2vtu.pl~-i~file.mesh~-o~file.vtu~}{\footnotesize \par}
-\end{lyxcode}
-\item Notice that this Perl script uses a program \texttt{mesh2vtu} in the
-\texttt{utils/Visualization/Paraview/mesh2vtu} directory, which further uses the VTK \urlwithparentheses{http://www.vtk.org/}
-run-time library for its execution. Therefore, make sure you have
-them properly set in the script according to your system.
-\end{enumerate}
-\item \textbf{Copy over the source and receiver .vtk file}
-
-In the case of a single source and a single receiver, the simulation
-also generates the file \texttt{sr.vtk} located in the
-\texttt{in\_out\_files/OUTPUT\_FILES/} directory to describe
-the source and receiver locations, which can also be viewed in Paraview
-in the next step.
-
-\item \textbf{View the mesh in ParaView}
-
-Finally, we can view the mesh in ParaView \urlwithparentheses{www.paraview.org}.
-
-\begin{enumerate}
-\item Open ParaView.
-\item From the top menu, \textsf{File} $\rightarrow$\textsf{Open data},
-select \texttt{file.vtu}, and click the \textsf{Accept} button.
-
-\begin{itemize}
-\item If the mesh file is of moderate size, it shows up on the screen; otherwise,
-only the bounding box is shown.
-\end{itemize}
-\item Click \textsf{Display Tab} $\rightarrow$ \textsf{Display Style} $\rightarrow$
-\textsf{Representation} and select \textsf{wireframe of surface} to
-display it.
-\item To create a cross-section of the volumetric mesh, choose \textsf{Filter}
-$\rightarrow$ \textsf{cut}, and under \textsf{Parameters Tab}, choose
-\textsf{Cut Function} $\rightarrow$ \textsf{plane}.
-\item Fill in center and normal information given by the \texttt{global\_slice\_number.pl}
-script (either from the standard output or from \texttt{normal\_plane.txt}
-file).
-\item To change the color scale, go to \textsf{Display Tab} $\rightarrow$
-\textsf{Color} $\rightarrow$ \textsf{Edit Color Map} and reselect
-lower and upper limits, or change the color scheme.
-\item Now load in the source and receiver location file by \textsf{File}
-$\rightarrow$ \textsf{Open data}, select \texttt{sr.vt}k, and click
-the \textsf{Accept} button. Choose \textsf{Filter} $\rightarrow$
-\textsf{Glyph}, and represent the points by `\textsf{spheres}'.
-\item For more information about ParaView, see the ParaView Users Guide \urlwithparentheses{www.paraview.org/files/v1.6/ParaViewUsersGuide.PDF}.
-\end{enumerate}
-\end{enumerate}
-%
-\begin{figure}[H]
-\noindent \begin{centering}
-\includegraphics[scale=0.6]{figures/3D-S-Kernel}
-\par\end{centering}
-
-\caption{(a) Top Panel: Vertical source-receiver cross-section of the S-wave
-finite-frequency sensitivity kernel $K_{\beta}$ for station GSC at
-an epicentral distance of 176 km from the September 3, 2002, Yorba
-Linda earthquake. Lower Panel: Vertical source-receiver cross-section
-of the 3D S-wave speed model used for the spectral-element simulations
-\citep{KoLiTrSuStSh04}. (b) The same as (a) but for station HEC at
-an epicentral distance of 165 km \citep{LiTr06}.}
-
-\label{figure:P-wave-speed-finite-frequency}
-\end{figure}
-
-\chapter{\label{cha:Scheduler}Running through a Scheduler}
-
-The code is usually run on large parallel machines, often PC clusters,
-most of which use schedulers, i.e., queuing or batch management systems
-to manage the running of jobs from a large number of users. The following
-considerations need to be taken into account when running on a system
-that uses a scheduler:
-
-\begin{itemize}
-\item The processors/nodes to be used for each run are assigned dynamically
-by the scheduler, based on availability. Therefore, in order for the
-\texttt{xgenerate\_databases} and the \texttt{xspecfem3D} executables (or between successive runs of the solver) to
-have access to the same database files (if they are stored on hard
-drives local to the nodes on which the code is run), they must be
-launched in sequence as a single job.
-\item On some systems, the nodes to which running jobs are assigned are
-not configured for compilation. It may therefore be necessary to pre-compile
-both the \texttt{xgenerate\_databases} and the \texttt{xspecfem3D} executables.
-\item One feature of schedulers/queuing systems is that they allow submission
-of multiple jobs in a ``launch and forget'' mode. In order to take
-advantage of this property, care needs to be taken that output and
-intermediate files from separate jobs do not overwrite each other,
-or otherwise interfere with other running jobs.
-\end{itemize}
-Examples of job scripts can be found in the \texttt{\small utils/Cluster/}
-directory and can straightforwardly be modified and adapted to meet
-more specific running needs.
-
-We describe here in some detail a job submission procedure for the
-Caltech 1024-node cluster, CITerra, under the LSF scheduling system.
-We consider the submission of a regular forward simulation using the internal mesher
-to create mesh partitions. The two
-main scripts are \texttt{\small run\_lsf.bash}, which compiles the
-Fortran code and submits the job to the scheduler, and \texttt{\small go\_mesher\_solver\_lsf\_basin.forward},
-which contains the instructions that make up
-the job itself. These scripts can be found in \texttt{\small utils/Cluster/lsf/} directory
-
-\section{Job submission \texttt{run\_lsf.bash}}
-
-This script first sets the job queue to be `normal'. It then compiles
-the mesher, database generator and solver together, figures out the number of processors
-required for this simulation from the \texttt{in\_data\_files/Par\_file}, and
-submits the LSF job.
-
-\begin{lyxcode}
-\#!/bin/bash
-
-\#~use~the~normal~queue~unless~otherwise~directed~queue=\char`\"{}-q~normal\char`\"{}~
-
-if~{[}~\$\#~-eq~1~];~then
-
-~~~~~~~~echo~\char`\"{}Setting~the~queue~to~\$1\char`\"{}
-
-~~~~~~~~queue=\char`\"{}-q~\$1\char`\"{}~
-
-fi~\\
-~\\
-\#~compile~the~mesher~and~the~solver~
-
-d=`date'~echo~\char`\"{}Starting~compilation~\$d\char`\"{}~
-
-make~clean~
-
-make~xmeshfem3D~
-
-make~xgenerate\_databases~
-
-make~xspecfem3D~
-
-d=`date'~
-
-echo~\char`\"{}Finished~compilation~\$d\char`\"{}~\\
-~\\
-\#~get~total~number~of~nodes~needed~for~solver~
-
-NPROC=`grep~NPROC~in\_data\_files/Par\_file~|~cut~-c~34-~'~\\
-~\\
-
-\#~compute~total~number~of~nodes~needed~for~mesher~
-
-NPROC\_XI=`grep~NPROC\_XI~in\_data\_files/meshfem3D\_files/Mesh\_Par\_file~|~cut~-c~34-~'~
-~\\
-
-NPROC\_ETA=`grep~NPROC\_ETA~in\_data\_files/meshfem3D\_files/Mesh\_Par\_file~|~cut~-c~34-~'~
-~\\
-\#~total~number~of~nodes~is~the~product~of~the~values~read~
-
-numnodes=\$((~\$NPROC\_XI~*~\$NPROC\_ETA~))~\\
-~\\
-
-\#~checks~total~number~of~nodes~
-
-if~{[}~\$numnodes~-neq~\$NPROC~];~then
-
-~~~~~~~~echo~\char`\"{}error number of procs mismatch\char`\"{}
-
-~~~~~~~~exit~
-
-fi~\\
-
-~\\
-echo~\char`\"{}Submitting~job\char`\"{}~
-
-bsub~\$queue~-n~\$numnodes~-W~60~-K~<go\_mesher\_solver\_lsf.forward~
-\end{lyxcode}
-
-\section{Job script \texttt{go\_mesher\_solver\_lsf.forward}}
-
-This script describes the job itself, including setup steps that can
-only be done once the scheduler has assigned a job-ID and a set of
-compute nodes to the job, the \texttt{run\_lsf.bash} commands used
-to run the mesher, database generator and the solver, and calls to scripts that collect
-the output seismograms from the compute nodes and perform clean-up
-operations.
-
-\begin{enumerate}
-\item First the script directs the scheduler to save its own output and
-output from \texttt{stdout} into ~\\
-\texttt{\small in\_out\_files/OUTPUT\_FILES/\%J.o},
-where \texttt{\%J} is short-hand for the job-ID; it also tells the
-scheduler what version of \texttt{mpich} to use (\texttt{mpich\_gm})
-and how to name this job (\texttt{go\_mesher\_solver\_lsf}).
-\item The script then creates a list of the nodes allocated to this job
-by echoing the value of a dynamically set environment variable \texttt{LSB\_MCPU\_HOSTS}
-and parsing the output into a one-column list using the Perl script
-\texttt{utils/Cluster/lsf/remap\_lsf\_machines.pl}. It then creates a set of scratch
-directories on these nodes (\texttt{\small /scratch/}~\\
-\texttt{\small \$USER/DATABASES\_MPI}) to be used as the \texttt{LOCAL\_PATH}
-for temporary storage of the database files. The scratch directories
-are created using \texttt{shmux}, a shell multiplexor that can execute
-the same commands on many hosts in parallel. \texttt{shmux} is available
-from Shmux \urlwithparentheses{web.taranis.org/shmux/}. Make sure that the \texttt{LOCAL\_PATH}
-parameter in \texttt{in\_data\_files/Par\_file} is also set properly.
-\item The next portion of the script launches the mesher, database generator and then the solver
-using \texttt{run\_lsf.bash}.
-\item The final portion of the script collects the seismograms and performs
-clean up on the nodes, using the Perl scripts \texttt{collect\_seismo\_lsf\_multi.pl}
-and \texttt{cleanmulti.pl}.
-\end{enumerate}
-\begin{lyxcode}
-\#!/bin/bash~-v~
-
-\#BSUB~-o~in\_out\_files/OUTPUT\_FILES/\%J.o~
-
-\#BSUB~-a~mpich\_gm~
-
-\#BSUB~-J~go\_mesher\_solver\_lsf~\\
-~\\
-\#~set~up~local~scratch~directories
-
-BASEMPIDIR=/scratch/\$USER/DATABASES\_MPI
-
-mkdir~-p~in\_out\_files/OUTPUT\_FILES~
-
-echo~\char`\"{}\$LSB\_MCPU\_HOSTS\char`\"{}~>~in\_out\_files/OUTPUT\_FILES/lsf\_machines~
-
-echo~\char`\"{}\$LSB\_JOBID\char`\"{}~>~in\_out\_files/OUTPUT\_FILES/jobid
-
-remap\_lsf\_machines.pl~in\_out\_files/OUTPUT\_FILES/lsf\_machines~> in\_out\_files/OUTPUT\_FILES/machines
-
-shmux~-M50~-Sall~-c~\char`\"{}rm~-r~-f~/scratch/\$USER;~\textbackslash{}
-
-~~~~~~mkdir~-p~/scratch/\$USER;~mkdir~-p~\$BASEMPIDIR\char`\"{}~\textbackslash{}
-
-~~~~~~-~<~in\_out\_files/OUTPUT\_FILES/machines~>/dev/null~\\
-~\\
-\#~run~the~specfem~program
-
-current\_pwd=\$PWD \\
-
-cd bin/ \\
-
-run\_lsf.bash~-{}-gm-no-shmem~-{}-gm-copy-env~\$current\_pwd/xmeshfem3D~\\
-
-run\_lsf.bash~-{}-gm-no-shmem~-{}-gm-copy-env~\$current\_pwd/xgenerate\_databases~\\
-
-run\_lsf.bash~-{}-gm-no-shmem~-{}-gm-copy-env~\$current\_pwd/xspecfem3D~\\
-~\\
-\#~collect~seismograms~and~clean~up
-
-cd current\_pwd/~ \\
-
-mkdir~-p~in\_out\_files/SEM
-
-cd~in\_out\_files/SEM/~
-
-collect\_seismo.pl~../OUTPUT\_FILES/lsf\_machines
-
-cleanbase.pl~../OUTPUT\_FILES/machines
-\end{lyxcode}
-
-\chapter{{\normalsize \label{cha:-Changing-the}} Changing the Model}
-
-In this section we explain how to change the velocity model used for your simulations.
-These changes involve contributing specific subroutines
-that replace existing subroutines in the \texttt{SPECFEM3D} package.
-Note that \texttt{SPECFEM3D} can handle Earth models with material properties that vary within each spectral element.
-
-%% magnoni 6/12
-\section{{\normalsize \label{sec:Using-tomographic}}Using external tomographic Earth models}
-
-To implement your own external tomographic model, you must provide your own tomography file
-\texttt{tomography\_model.xyz} and choose between two possible options:
-(1) set in the \texttt{Par\_file} the model parameter \texttt{MODEL = tomo}, or
-(2) for more user control, set in the \texttt{Par\_file} the model parameter \texttt{MODEL = default} and use the following format in the file \texttt{MESH/nummaterial\_velocity\_file} when using CUBIT to construct your mesh (see also section \ref{subsec:Exporting-the-Mesh}):
-\begin{lyxcode}
-\texttt{domain\_ID} \texttt{material\_ID} \texttt{tomography} \texttt{elastic} \texttt{file\_name} 1
-\end{lyxcode}
-where
-\texttt{domain\_ID} is 1 for acoustic or 2 for elastic materials,
-\texttt{material\_ID} a negative, unique identifier (i.e., -1),
-\texttt{tomography} keyword for tomographic material definition,
-\texttt{elastic} keyword for elastic material definition,
-\texttt{file\_name} the name of the tomography file and
-1 a positive unique identifier.
-The name of the tomography file is not used so far, rather change the filename (\texttt{TOMO\_FILENAME}) defined in the file \texttt{model\_tomography.f90} located in the \texttt{src/generate\_databases/} directory.
-
-The external tomographic model is represented by a grid of points with assigned material properties and homogeneous resolution along each spatial direction x, y and z.
-The ASCII file \texttt{tomography\_model.xyz} that describes the tomography should be located in the \texttt{in\_data\_files/} directory.
-The format of the file, as read from \texttt{model\_tomography.f90}, looks like Figure \ref{fig:tomography_file},
-where
-%
-\begin{description}
-\item [{\texttt{ORIG\_X}, \texttt{END\_X}}] are, respectively, the coordinates of the initial and final tomographic grid points along the x direction (in the mesh units, e.g., $m$);
-\item [{\texttt{ORIG\_Y}, \texttt{END\_Y}}] respectively the coordinates of the initial and final tomographic grid points along the y direction (in the mesh units, e.g., $m$);
-\item [{\texttt{ORIG\_Z}, \texttt{END\_Z}}] respectively the coordinates of the initial and final tomographic grid points along the z direction (in the mesh units, e.g., $m$);
-\item [\texttt{SPACING\_X}, \texttt{SPACING\_Y}, \texttt{SPACING\_Z}] the spacing between the tomographic grid points along the x, y and z directions, respectively (in the mesh units, e.g., $m$);
-\item [\texttt{NX}, \texttt{NY}, \texttt{NZ}] the number of grid points along the spatial directions x, y and z, respectively; \texttt{NX} is given by [(\texttt{END\_X} - \texttt{ORIG\_X})/\texttt{SPACING\_X}]+1; \texttt{NY} and \texttt{NZ} are the same as \texttt{NX}, but for the y and z directions, respectively;
-\item [\texttt{VP\_MIN}, \texttt{VP\_MAX}, \texttt{VS\_MIN}, \texttt{VS\_MAX}, \texttt{RHO\_MIN}, \texttt{RHO\_MAX}] the minimum and maximum values of the wave speed \texttt{vp} and \texttt{vs} (in $m \, s^{-1}$) and of the density \texttt{rho} (in $kg \, m^{-3}$); these values could be the actual limits of the tomographic parameters in the grid or the minimum and maximum values to which we force the cut of velocity and density in the model.
-\end{description}
-%
-After the first four lines, in the file \texttt{tomography\_model.xyz} the tomographic grid points are listed with the corresponding values of \texttt{vp}, \texttt{vs} and \texttt{rho}, scanning the grid along the x coordinate (from \texttt{ORIG\_X} to \texttt{END\_X} with step of \texttt{SPACING\_X}) for each given y (from \texttt{ORIG\_Y} to \texttt{END\_Y}, with step of \texttt{SPACING\_Y}) and each given z (from \texttt{ORIG\_Z} to \texttt{END\_Z}, with step of \texttt{SPACING\_Z}).
-%fig
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=1.0\textwidth]{figures/tomo_file.jpg}
-\par\end{centering}
-\caption{Tomography file \texttt{tomography\_model.xyz} that describes an external Earth model and that is read by \texttt{model\_tomography.f90}.
-The coordinates x, y and z of the grid, their limits ORIG\_X, ORIG\_Y, ORIG\_Z, END\_X, END\_Y, END\_Z and the grid spacings SPACING\_X, SPACING\_Y, SPACING\_Z should be in the units of the constructed mesh (e.g., UTM coordinates in meters).}
-\label{fig:tomography_file}
-\end{figure}
-
-The user can implement his own interpolation algorithm for the tomography model by changing the routine \texttt{model\_tomography.f90} located in the \texttt{src/generate\_databases/} directory.
-Moreover, for models that involve both fully defined materials and a tomography description, the \texttt{nummaterial\_velocity\_file} has multiple lines each with the corresponding suitable format described above.
-
-\section{{\normalsize \label{sec:Anelastic-Models}}External (an)elastic Models}
-
-To use your own external model, you can set in the \texttt{Par\_file} the model parameter \texttt{MODEL = external}.
-Three-dimensional acoustic and/or (an)elastic (attenuation) models may then be superimposed
-onto the mesh based upon your subroutine in
-\texttt{model\_external\_values.f90} located in directory \texttt{src/generate\_databases}.
-The call to this routine would be as follows
-
-\begin{lyxcode}
-call~model\_external\_values(xmesh,~ymesh,~zmesh, ~\& \\
-~~~~~~~~~~rho,~vp,~vs,~qmu\_atten,~iflag\_aniso,~idomain\_id)
-\end{lyxcode}
-%
-Input to this routine consists of:
-\begin{description}
-\item [{\texttt{xmesh,ymesh,zmesh}}] location of mesh point
-\end{description}
-%
-Output to this routine consists of:
-\begin{description}
-\item [{\texttt{rho,vp,vs}}] isotropic model parameters for density $\rho$ ($kg/m^3$), $v_p$ ($m/s$) and $v_s$ ($m/s$)
-\item [{\texttt{qmu\_atten}}] Shear wave quality factor: $0<Q_{\mu}<9000$
-\item [{\texttt{iflag\_aniso}}] anisotropic model flag, $0$ indicating no anisotropy or $1$ using anisotropic model parameters as defined in routine file \texttt{model\_aniso.f90}
-\item [{\texttt{idomain\_id}}] domain identifier, $1$ for acoustic, $2$ for elastic, $3$ for poroelastic materials.
-\end{description}
-
-Note that the resolution and maximum value of anelastic models are
-truncated. This speeds the construction of the standard linear solids
-during the meshing stage. To change the resolution, currently at one
-significant figure following the decimal, or the maximum value (9000),
-consult \texttt{shared/constants.h}.
-
-\section{{\normalsize \label{sec:Anisotropic-Models}}Anisotropic Models}
-
-To use your anisotropic models, you can either set in the \texttt{Par\_file} the model parameter \texttt{ANISOTROPY} to \texttt{.true.} or set the model parameter \texttt{MODEL = aniso}.
-Three-dimensional anisotropic models may be superimposed on
-the mesh based upon the subroutine in file \texttt{model\_aniso.f90}
-located in directory \texttt{src/generate\_databases}.
-The call to this subroutine is of the form
-
-\begin{lyxcode}
-call~model\_aniso(iflag\_aniso,~rho,~vp,~vs,~\& \\
-~~~~~~~~~~c11,c12,c13,c14,c15,c16,c22,c23,c24,c25,c26,~\& \\
-~~~~~~~~~~c33,c34,c35,c36,c44,c45,c46,c55,c56,c66)~
-\end{lyxcode}
-%
-Input to this routine consists of:
-\begin{description}
-\item [{\texttt{iflag\_aniso}}] flag indicating the type of the anisotropic model, i.e. $0$ for no superimposed anisotropy, $1$ or $2$ for generic pre-defined anisotropic models.
-\item [{\texttt{rho,vp,vs}}] reference isotropic model parameters for density $\rho$, $v_p$ and $v_s$.
-\end{description}
-%
-Output from the routine consists of the following non-dimensional
-model parameters:
-\begin{description}
-\item [{\texttt{c11},}] \textbf{$\cdots$,} \texttt{\textbf{c66}} 21~dimensionalized
-anisotropic elastic parameters.
-\end{description}
-
-You can replace the \texttt{model\_aniso.f90} file by
-your own version \textit{provided you do not change the call structure
-of the routine}, i.e., the new routine should take exactly the same
-input and produce the required relative output.
-
-\chapter{Post-Processing Scripts}
-
-Several post-processing scripts/programs are provided in the \texttt{utils/}
-directory, and most of them need to be adjusted when used on different
-systems, for example, the path of the executable programs. Here we
-only list a few of the available scripts and provide a brief description, and
-you can either refer to the related sections for detailed usage or,
-in a lot of cases, type the script/program name without arguments
-for its usage.
-
-\section{Process Data and Synthetics\label{sec:Process-data-and-syn}}
-
-In many cases, the SEM synthetics are calculated and compared to data
-seismograms recorded at seismic stations. Since the SEM synthetics
-are accurate for a certain frequency range, both the original data
-and the synthetics need to be processed before a comparison can be
-made.
-
- For such comparisons, the following steps are recommended:
-
-\begin{enumerate}
-\item Make sure that both synthetic and observed seismograms have the correct station/event and timing information.
-\item Convolve synthetic seismograms with a source time function with the half duration specified in the \texttt{CMTSOLUTION} file, provided, as recommended, you used a zero half duration in the SEM simulations.
-\item Resample both observed and synthetic seismograms to a common sampling rate.
-\item Cut the records using the same window.
-\item Remove the trend and mean from the records and taper them.
-\item Remove the instrument response from the observed seismograms (recommended) or convolve the synthetic seismograms with the instrument response.
-\item Make sure that you apply the same filters to both observed and synthetic seismograms. Preferably, avoid filtering your records more than once.
-\item Now, you are ready to compare your synthetic and observed seismograms.
-\end{enumerate}
-
-\noindent
-We generally use the following scripts provided in the \texttt{utils/seis\_process/} directory:
-
-\subsection{Data processing script \texttt{process\_data.pl}}
-
-This script cuts a given portion of the original data, filters it,
-transfers the data into a displacement record, and picks the first
-P and S arrivals. For more functionality, type `\texttt{process\_data.pl}'
-without any argument. An example of the usage of the script:
-
-\begin{lyxcode}
-{\footnotesize process\_data.pl~-m~CMTSOLUTION~-s~1.0~-l~0/4000~-i~-f~-t~40/500~-p~-x~bp~DATA/1999.330{*}.BH?.SAC}
-{\footnotesize \par}
-\end{lyxcode}
-
-\noindent which has resampled the SAC files to a sampling rate of 1 seconds, cut
-them between 0 and 4000 seconds, transfered them into displacement
-records and filtered them between 40 and 500 seconds,
-picked the first P and S arrivals, and added suffix `\texttt{bp}' to the file names.
-
-Note that all of the scripts in this section actually use the SAC
-and/or IASP91 to do the core operations; therefore make sure that
-the SAC and IASP91 packages are installed properly on your system,
-and that all the environment variables are set properly before running
-these scripts.
-
-\subsection{Synthetics processing script \texttt{process\_syn.pl}}
-
-This script converts the synthetic output from the SEM code from ASCII
-to SAC format, and performs similar operations as `\texttt{process\_data.pl}'.
-An example of the usage of the script:
-\begin{lyxcode}
-{\footnotesize process\_syn.pl~-m~CMTSOLUTION~-h~-a~STATIONS~-s~1.0~-l~0/4000~-f~-t~40/500~-p~-x~bp~SEM/{*}.BX?.semd}
-{\footnotesize \par}
-\end{lyxcode}
-which will convolve the synthetics with a triangular source-time function
-from the \texttt{CMTSOLUTION} file, convert the synthetics into SAC
-format, add event and station information into the SAC headers, resample the SAC files with a sampling rate of 1 seconds, cut them between 0 and 4000 seconds, filter them between 40 and
-500 seconds with the same filter used for the observed data, pick the first P and S arrivals, and add the suffix `\texttt{bp}'
-to the file names.
-
-More options are available for this script, such as adding time shift
-to the origin time of the synthetics, convolving the synthetics with
-a triangular source time function with a given half duration, etc.
-Type \texttt{process\_syn.pl} without any argument for a detailed
-usage.
-
-In order to convert between SAC format and ASCII files, useful scripts
-are provided in the subdirectories ~\\
-\texttt{utils/sac2000\_alpha\_convert/} and \texttt{utils/seis\_process/asc2sac/}.
-
-\subsection{Script \texttt{rotate.pl}}
-
-The original data and synthetics have three components: vertical (BHZ resp. BXZ),
-north (BHN resp. BXN) and east (BHE resp. BXE). However, for most seismology applications,
-transverse and radial components are also desirable. Therefore, we
-need to rotate the horizontal components of both the data and the
-synthetics to the transverse and radial direction, and \texttt{\small rotate.pl}
-can be used to accomplish this:
-
-\begin{lyxcode}
-rotate.pl~-l~0~-L~180~-d~DATA/{*}.BHE.SAC.bp~
-
-rotate.pl~-l~0~-L~180~SEM/{*}.BXE.semd.sac.bp~
-\end{lyxcode}
-where the first command performs rotation on the SAC data obtained
-through Seismogram Transfer Program (STP) \urlwithparentheses{http://www.data.scec.org/STP/stp.html},
-while the second command rotates the processed SEM synthetics.
-
-\section{Collect Synthetic Seismograms}
-
-The forward and adjoint simulations generate synthetic seismograms
-in the \texttt{in\_out\_files/OUTPUT\_FILES/} directory by default. For the forward simulation, the files
-are named like \texttt{STA.NT.BX?.semd} for two-column time series, or
-\texttt{STA.NT.BX?.semd.sac} for ASCII SAC format, where STA and NT
-are the station name and network code, and \texttt{BX?} stands for
-the component name. Please see the Appendix~\ref{cha:Coordinates} and \ref{cha:channel-codes}
-for further details.
-
-The adjont simulations generate synthetic seismograms
-with the name \texttt{S?????.NT.S??.sem} (refer to Section \ref{sec:Adjoint-simulation-sources}
-for details). The kernel simulations output the back-reconstructed
-synthetic seismogram in the name \texttt{STA.NT.BX?.semd}, mainly
-for the purpose of checking the accuracy of the reconstruction. Refer
-to Section \ref{sec:Adjoint-simulation-finite} for further details.
-
-You do have further options to change this default output behavior, given in the main constants file
-\texttt{constants.h} located in \texttt{src/shared/} directory:
-\begin{description}
-\item [{\texttt{SEISMOGRAMS\_BINARY}}] set to \texttt{.true.} to have seismograms written out in binary format.
-\item [{\texttt{WRITE\_SEISMOGRAMS\_BY\_MASTER}}] Set to \texttt{.true.} to have only the master process writing out seismograms.
-This can be useful on a cluster, where only the master process node has access to the output directory.
-\item [{\texttt{USE\_OUTPUT\_FILES\_PATH}}] Set to \texttt{.false.} to have the seismograms output to \texttt{LOCAL\_PATH} directory specified in the main parameter file \texttt{in\_data\_files/Par\_file}.
-In this case, you could collect the synthetics onto the frontend using the \texttt{collect\_seismo\_lsf\_multi.pl} script
-located in the \texttt{utils/Cluster/lsf/} directory. The usage of the script would be e.g.:
-
-\begin{lyxcode}
-collect\_seismo.pl~machines~in\_data\_files/Par\_file
-\end{lyxcode}
-
-\end{description}
-where  \texttt{machines} is a file containing the node names and  \texttt{in\_data\_files/Par\_file} the parameter file used to extract the \texttt{LOCAL\_PATH} directory used for the simulation.
-
-\section{Clean Local Database}
-
-After all the simulations are done, the seismograms are collected,
-and the useful database files are copied to the frontend, you may
-need to clean the local scratch disk for the next simulation. This
-is especially important in the case of kernel simulation,
-where very large files are generated for the absorbing boundaries
-to help with the reconstruction of the regular forward wavefield.
-A sample script is provided in \texttt{utils/}:
-
-\begin{lyxcode}
-cleanbase.pl~machines
-\end{lyxcode}
-where  \texttt{machines} is a file containing the node names.
-
-\section{Plot Movie Snapshots and Synthetic Shakemaps}
-
-\subsection{Script \texttt{movie2gif.gmt.pl}}
-
-With the movie data saved in \texttt{in\_out\_files/OUTPUT\_FILES/} at the end of
-a movie simulation (\texttt{\small MOVIE\_SURFACE=.true.}), you can
-run the \texttt{`create\_movie\_shakemap\_AVS\_DX\_GMT}' code to convert these binary
-movie data into GMT xyz files for futher processing. A sample script
-\texttt{movie2gif.gmt.pl} is provided to do this conversion, and then
-plot the movie snapshots in GMT, for example:
-
-\begin{lyxcode}
-movie2gif.gmt.pl~-m~CMTSOLUTION~-g~-f~1/40~-n~-2~-p~
-\end{lyxcode}
-which for the first through the 40th movie frame, converts the \texttt{moviedata}
-files into GMT xyz files, interpolates them using the 'nearneighbor'
-command in GMT, and plots them on a 2D topography map. Note that `\texttt{-2}'
-and `\texttt{-p}' are both optional.
-
-\subsection{Script \texttt{plot\_shakemap.gmt.pl}}
-
-With the shakemap data saved in \texttt{in\_out\_files/OUTPUT\_FILES/} at the end
-of a shakemap simulation ~\\
-(\texttt{CREATE\_SHAKEMAP=.true.}), you can
-also run \texttt{`create\_movie\_shakemap\_AVS\_DX\_GMT}' code to convert the binary
-shakemap data into GMT xyz files. A sample script \texttt{plot\_shakemap.gmt.pl}
-is provided to do this conversion, and then plot the shakemaps in
-GMT, for example:
-
-\begin{lyxcode}
-plot\_shakemap.gmt.pl~data\_dir~type(1,2,3)~CMTSOLUTION~
-\end{lyxcode}
-where \texttt{type=1} for a displacement shakemap, \texttt{2} for
-velocity, and \texttt{3} for acceleration.
-
-\section{Map Local Database}
-
-A sample program \texttt{remap\_database} is provided to map the local
-database from a set of machines to another set of machines. This is
-especially useful when you want to run mesher and solver, or different
-types of solvers separately through a scheduler (refer to Chapter~\ref{cha:Scheduler}).
-\begin{lyxcode}
-run\_lsf.bash~-{}-gm-no-shmem~-{}-gm-copy-env~remap\_database~old\_machines~150
-\end{lyxcode}
-where \texttt{old\_machines} is the LSF machine file used in the previous
-simulation, and \texttt{150} is the number of processors in total.
-
-\chapter{\label{cha:tomo}Tomographic inversion using sensitivity kernels}
-
-One of the fundamental reasons for computing sensitivity kernels (Section~\ref{sec:Adjoint-simulation-finite}) is to use them within a tomographic inversion.
-In other words, use recorded seismograms, make measurements with synthetic seismograms, and use the misfit between the two to iteratively improve the model described by (at least) $V_{\rm p}$, $V_{\rm s}$, and $\rho$ as a function of space.
-
-Whatever misfit function you use for the tomographic inversion (several examples in \cite{TrTaLi05} and \cite{TrKoLi08}), you will weight the sensitivity kernels with measurements.
-Furthermore, you will use as many measurements (stations, components, time windows) as possible per event; hence, we call these composite kernels ``event kernels,'' which are volumetric fields representing the gradient of the misfit function with respect to one of the variables (\eg $V_{\rm s}$).
-The basic features of an adjoint-based tomographic inversion were illustrated in \citet{TaLiTr07} and \citet{TrKoLi08} using a conjugate-gradient algorithm; there are dozens of versions of gradient-based inversion algorithms that could alternatively be used.
-The tomographic inversion of \citet{TaLiMaTr09,TaLiMaTr2010} used SPECFEM3D as well as several additional components which are also stored on the CIG svn server, described next.
-
-The directory containing utilities for tomographic inversion using SPECFEM3D (or other packages that evaluate misfit functions and gradients) is here on the CIG svn server:
-%
-\begin{verbatim}
-/cig/seismo/3D/ADJOINT_TOMO/
-   flexwin/     -- FLEXWIN algorithm for automated picking of time windows
-   measure_adj/ -- reads FLEXWIN output file and makes measurements,
-                   with the option for computing adjoint sources
-   iterate_adj/ -- various tools for iterative inversion
-                   (requires pre-computed "event kernels")
-\end{verbatim}
-%
-This directory also contains a brief \verb+README+ file indicating the role of the three subdirectories, \verb+flexwin+ \citep{Maggi2009}, \verb+measure_adj+, and \verb+iterate_adj+.
-The components for making the model update are there; however, there are no explicit rules for computing the model update, just as with any optimization problem.
-There are options for computing a conjugate gradient step, as well as a source subspace projection step.
-The workflow could use substantial ``stitching together,'' and we welcome suggestions; undoubtedly, the tomographic inversion procedure will improve with time.
-{\bf This is a work in progress.}
-
-The best single file to read is probably: \verb+ADJOINT_TOMO/iterate_adj/cluster/README+.
-
-%============================================================
-
-\chapter*{\label{cha:Bug-Reports-and}Bug Reports and Suggestions for Improvements}
-
-To report bugs or suggest improvements to the code, please send an
-e-mail to the CIG Computational Seismology Mailing List \urlwithparentheses{cig-seismo@geodynamics.org}
-or Jeroen Tromp \urlwithparentheses{jtromp-AT-princeton.edu}, and/or use our online
-bug tracking system Roundup \urlwithparentheses{www.geodynamics.org/roundup}.
-
-\addcontentsline{toc}{chapter}{Notes \& Acknowledgments}
-\chapter*{Notes \& Acknowledgments}
-
-In order to keep the software package thread-safe in case a multithreaded
-implementation of MPI is used, developers should not add modules or
-common blocks to the source code but rather use regular subroutine
-arguments (which can be grouped in ``derived types'' if needed for
-clarity).
-
-The Gauss-Lobatto-Legendre subroutines in \texttt{gll\_library.f90}
-are based in part on software libraries from the Massachusetts Institute
-of Technology, Department of Mechanical Engineering (Cambridge, Massachusetts, USA).
-The non-structured global numbering software was provided by Paul
-F. Fischer (Brown University, Providence, Rhode Island, USA, now at Argonne National Laboratory, USA).
-
-OpenDX \urlwithparentheses{http://www.opendx.org} is open-source based on IBM Data
-Explorer, AVS \urlwithparentheses{http://www.avs.com} is a trademark of Advanced
-Visualization Systems, and ParaView \urlwithparentheses{http://www.paraview.org}
-is an open-source visualization platform.
-
-Please e-mail your feedback, questions, comments, and suggestions
-to Jeroen Tromp \urlwithparentheses{jtromp-AT-princeton.edu} or to the CIG Computational Seismology Mailing List \urlwithparentheses{cig-seismo@geodynamics.org}.
-
-\addcontentsline{toc}{chapter}{Copyright}
-\chapter*{Copyright}
-
-Main authors: Dimitri Komatitsch and Jeroen Tromp
-
-Princeton University, USA, and CNRS / INRIA / University of Pau, France
-
-$\copyright$ Princeton University and CNRS / INRIA / University of Pau, July 2012
-
-This program is free software; you can redistribute it and/or modify
-it under the terms of the GNU General Public License as published
-by the Free Software Foundation (see Appendix \ref{cha:License}).\\
-
-\textbf{\underline{Evolution of the code:}}\\
-
- MPI v 2.1, July 2012:
- Max Rietmann, Peter Messmer, Daniel Peter, Dimitri Komatitsch, Joseph Charles, Zhinan Xie: support for CUDA GPUs,
-better CFL stability for the Stacey absorbing conditions. \\
-
- MPI v. 2.0, November 2010:
- Dimitri Komatitsch, Nicolas Le Goff, Roland Martin and Pieyre Le Loher, University of Pau, France,
- Jeroen Tromp and the Princeton group of developers, Princeton University, USA,
- and Emanuele Casarotti, INGV Roma, Italy:
-  support for CUBIT meshes decomposed by SCOTCH, METIS or ZOLTAN;
-  much faster solver using Michel Deville's inlined matrix products.\\
-
- MPI v. 1.4 Dimitri Komatitsch, University of Pau, Qinya Liu and others, Caltech, September 2006:
-  better adjoint and kernel calculations, faster and better I/Os
-  on very large systems, many small improvements and bug fixes.\\
-
- MPI v. 1.3 Dimitri Komatitsch, University of Pau, and Qinya Liu, Caltech, July 2005:
-  serial version, regular mesh, adjoint and kernel calculations, ParaView support.\\
-
- MPI v. 1.2 Min Chen and Dimitri Komatitsch, Caltech, July 2004:
-  full anisotropy, volume movie.\\
-
- MPI v. 1.1 Dimitri Komatitsch, Caltech, October 2002: Zhu's Moho map, scaling
-  of $V_s$ with depth, Hauksson's regional model, attenuation, oceans, movies.\\
-
- MPI v. 1.0 Dimitri Komatitsch, Caltech, USA, May 2002: first MPI version based on global code.\\
-
- Dimitri Komatitsch, IPG Paris, France, December 1996: first 3-D solver for the CM-5 Connection Machine,
-    parallelized on 128 processors using Connection Machine Fortran.\\
-
-\addcontentsline{toc}{chapter}{References}
-\bibliography{bibliography}
-
-\appendix
-
-\chapter{\label{cha:Coordinates}Reference Frame Convention}
-
-The code uses the following convention for the Cartesian reference
-frame:
-
-\begin{itemize}
-\item the $x$ axis points East
-\item the $y$ axis points North
-\item the $z$ axis points up
-\end{itemize}
-Note that this convention is different from both the \citet{AkRi80}
-convention and the Harvard Centroid-Moment Tensor (CMT) convention.
-The Aki \& Richards convention is
-
-\begin{itemize}
-\item the $x$ axis points North
-\item the $y$ axis points East
-\item the $z$ axis points down
-\end{itemize}
-and the Harvard CMT convention is
-
-\begin{itemize}
-\item the $x$ axis points South
-\item the $y$ axis points East
-\item the $z$ axis points up
-\end{itemize}
-
-\subsection*{Source and receiver locations}
-
-The SPECFEM3D software code internally uses Cartesian coordinates.
-The given locations for sources and receiver locations thus may get
-converted.  Sources and receiver locations are read in from the
-\texttt{CMTSOLUTION} (or \texttt{FORCESOLUTION}) and \texttt{STATIONS}
-files.  Note that e.g. the \texttt{CMTSOLUTION} and
-\texttt{FORCESOLUTION} files denotes the location by
-"longitude/latitude/depth". We read in longitude as $x$ coordinate,
-latitude as $y$ coordinate.
-
-In case the flag \texttt{SUPPRESS\_UTM\_PROJECTION} is set to \texttt{.false.}
-in the main parameter file (see Chapter~\ref{cha:Creating-Distributed-Databases}),
-the $x$/$y$ coordinates have to be given in degrees and are converted to Cartesian coordinates
-using a UTM conversion with the specified UTM zone.
-
-The value for depth (given in $km$ in \texttt{CMTSOLUTION} and
-\texttt{FORCESOLUTION}) or burial depth (given in $m$ in
-\texttt{STATIONS}) is evaluated with respect to the surface of the
-mesh at the specified $x$/$y$ location to find the corresponding $z$
-coordinate.  It is possible to use this depth value directly as $z$
-coordinate by changing the flag \texttt{USE\_SOURCES\_RECVS\_Z} to
-\texttt{.true.} in the file \texttt{constants.h} located in the
-\texttt{src/shared/} subdirectory.
-
-\subsection*{Seismogram outputs}
-
-The seismogram output directions are given in Cartesian $x$/$y$/$z$ directions and not rotated any further.
-Changing flags in \texttt{constants.h} in the \texttt{src/shared/} subdirectory only rotates the seismogram outputs
-if receivers are forced to be located at the surface (\texttt{RECVS\_CAN\_BE\_BURIED\_EXT\_MESH} set to \texttt{.false.}) and the normal to the surface at the receiver location should be used (\texttt{EXT\_MESH\_RECV\_NORMAL} set to
-\texttt{.true.}) as vertical.
-In this case, the outputs are rotated to have the vertical component normal to the surface of the mesh, $x$ and $y$ directions are somewhat arbitrary orthogonal directions along the surface.
-
-For the labeling of the seismogram channels, see Appendix~\ref{cha:channel-codes}. Additionally, we add labels to the synthetics using the following convention:
-For a receiver station located in an
-\begin{description}
-\item[elastic domain:] ~\\
-\begin{itemize}
-\item \texttt{semd} for the displacement vector
-\item \texttt{semv} for the velocity vector
-\item \texttt{sema} for the acceleration vector
-\end{itemize}
-\item[acoustic domain:] ~\\
-(please note that receiver stations in acoustic domains must be buried. This is due to the free surface condition which enforces a zero pressure at the surface.)
-\begin{itemize}
-\item \texttt{semd} for the displacement vector
-\item \texttt{semv} for the velocity vector
-\item \texttt{sema} for pressure which will be stored in the vertical component \texttt{Z}.
-Note that pressure in the acoustic media is isotropic and thus the pressure record would be the same in the other two component directions.
-We therefore use the other two seismogram components to store the acoustic potential in component \texttt{X} (or \texttt{N}) and the first time derivative of the acoustic potential in component \texttt{Y} (or \texttt{E}).
-\end{itemize}
-\end{description}
-The seismograms are by default written out in ASCII-format to the \texttt{in\_out\_files/OUTPUT\_FILES/} subdirectory by each parallel process.
-You can change this behavior by changing the following flags in the \texttt{constants.h} file located in the \texttt{src/shared/} subdirectory:
-\begin{description}
-\item [{\texttt{SEISMOGRAMS\_BINARY}}] set to \texttt{.true.} to have seismograms written out in binary format.
-\item [{\texttt{WRITE\_SEISMOGRAMS\_BY\_MASTER}}] Set to \texttt{.true.} to have only the master process writing out seismograms.
-This can be useful on a cluster, where only the master process node has access to the output directory.
-\item [{\texttt{USE\_OUTPUT\_FILES\_PATH}}] Set to \texttt{.false.} to have the seismograms output to \texttt{LOCAL\_PATH} directory specified in the main parameter file \texttt{in\_data\_files/Par\_file}.
-\end{description}
-
-\chapter{\label{cha:channel-codes}Channel Codes of Seismograms}
-
-Seismic networks, such as the Global Seismographic Network (GSN), generally involve various types of instruments with different bandwidths, sampling properties and component configurations.
-There are standards to name channel codes depending on instrument properties.
-IRIS  \texttt{(www.iris.edu)} uses SEED format for channel codes, which are represented by three letters, such as \texttt{LHN}, \texttt{BHZ}, etc.
-In older versions of the SPECFEM3D package, a common format was used for the channel codes of all seismograms, which was \texttt{BHE/BHN/BHZ} for three components.
-To avoid confusion when comparison are made to observed data, we are now using the FDSN convention (\url{http://www.fdsn.org/}) for SEM seismograms.
-In the following, we give a brief explanation of the FDSN convention used by IRIS, and how it is adopted in SEM seismograms.
-Please visit \texttt{www.iris.edu/manuals/SEED\_appA.htm} for further information.\\
-
-\noindent \textbf{\texttt{Band code:}} The first letter in the channel code denotes the band code of seismograms, which depends on the response band and the sampling rate of instruments.
-The list of band codes used by IRIS is shown in Figure \ref{fig:IRIS_band_codes}.
-The sampling rate of SEM synthetics is controlled by the resolution of simulations rather than instrument properties.
-However, for consistency, we follow the FDSN convention for SEM seismograms governed by their sampling rate.
-For SEM synthetics, we consider band codes for which $dt \leq 1$ s.
-IRIS also considers the response band of instruments.
-For instance, short-period and broad-band seismograms with the same sampling rate correspond to different band codes, such as S and B, respectively.
-In such cases, we consider SEM seismograms as broad band, ignoring the corner period ($\geq 10$ s) of the response band of instruments (note that at these resolutions, the minimum period in the SEM synthetics will be less than $10$ s).
-Accordingly, when you run a simulation the band code will be chosen depending on the resolution of the synthetics, and channel codes of SEM seismograms will start with either \texttt{L}, \texttt{M}, \texttt{B}, \texttt{H}, \texttt{C} or \texttt{F}, shown by red color in the figure.\\
-
-\begin{figure}[ht]
-\noindent \begin{centering}
-\includegraphics[scale=0.6]{figures/IRIS_band_codes.pdf}
-\par\end{centering}
-
-\caption{The band code convention is based on the sampling rate and the response band of instruments.
-Please visit \texttt{www.iris.edu/manuals/SEED\_appA.htm} for further information.
-Grey rows show the relative band-code range in SPECFEM3D, and the band codes used to name SEM seismograms are denoted in red.}
-
-\label{fig:IRIS_band_codes}
-\end{figure}
-
-\noindent \textbf{\texttt{Instrument code:}} The second letter in the channel code corresponds to instrument codes, which specify the family to which the sensor belongs.
-For instance, \texttt{H} and \texttt{L} are used for high-gain and low-gain seismometers, respectively.
-The instrument code of SEM seismograms will always be \texttt{X}, as assigned by FDSN for synthetic seismograms.
-\\
-
-\noindent \textbf{\texttt{Orientation code:}} The third letter in channel codes is an orientation code, which generally describes the physical configuration of the components of instrument packages.
-SPECFEM3D uses the traditional orientation code  \texttt{E/N/Z} (East-West, North-South, Vertical) for three components when a UTM projection is used.
-If the UTM conversion is suppressed, i.e.
-the flag \texttt{SUPPRESS\_UTM\_PROJECTION} is set to \texttt{.true.}, then the three components are labelled \texttt{X/Y/Z} according to the Cartesian reference frame. \\
-
-\noindent \textbf{\texttt{EXAMPLE:}} The sampling rate is given by \texttt{DT} in the main parameter file \texttt{Par\_file} located in the \texttt{in\_data\_files/} subdirectory.
-Depending on the resolution of your simulations, if you choose a sampling rate greater than $0.01$ s and less than $1$ s, a seismogram recording displacements on the vertical component of a station \texttt{ASBS} (network \texttt{AZ}) will be named \texttt{ASBS.AZ.MXZ.semd.sac}, whereas it will be \texttt{ASBS.AZ.BXZ.semd.sac}, if the sampling rate is greater than 0.0125 and less equal to 0.1 s.
-
-\chapter{\label{cha:Troubleshooting}Troubleshooting}
-
-\section*{FAQ}
-
-\begin{description}
-\item [configuration fails:]
-
-   Examine the log file 'config.log'. It contains detailed informations.
-   In many cases, the path's to these specific compiler commands F90,
-   CC and MPIF90 won't be correct if `./configure` fails.
-
-   Please make sure that you have a working installation of a Fortran compiler,
-   a C compiler and an MPI implementation. You should be able to compile this
-   little program code:
-
-\begin{lyxcode}
-
-      program main ~\\
-        include 'mpif.h' ~\\
-        integer, parameter :: CUSTOM\_MPI\_TYPE = MPI\_REAL ~\\
-        integer ier ~\\
-        call MPI\_INIT(ier) ~\\
-        call MPI\_BARRIER(MPI\_COMM\_WORLD,ier) ~\\
-        call MPI\_FINALIZE(ier) ~\\
-      end
-
-\end{lyxcode}
-
-\item [compilation fails stating:] ~\\
-\begin{lyxcode}
-    ...  ~\\
-    obj/program\_generate\_databases.o: In function `MAIN\_\_':  ~\\
-    program\_generate\_databases.f90:(.text+0x14): undefined reference to `\_gfortran\_set\_std'  ~\\
-    ...  ~\\
-\end{lyxcode}
-
-  Make sure you're pointing to the right 'mpif90' wrapper command.
-
-  Normally, this message will appear when you're mixing two different Fortran
-  compilers. That is, using e.g. gfortran to compile non-MPI files
-  and mpif90, wrapper provided for e.g. ifort, to compile MPI-files.
-
-  fix: e.g. specify > ./configure FC=gfortran MPIF90=/usr/local/openmpi-gfortran/bin/mpif90
-
-\item [after executing \texttt{xmeshfem3D} I've got elements with skewness of 81\% percent, what does this mean:]
-Look at the skewness table printed in the \texttt{output\_mesher.txt} file after executing \texttt{xmeshfem3D} for the
-example given in \texttt{examples/meshfem3D\_examples/simple\_model/}:
-\begin{lyxcode}
-...~\\
-histogram of skewness (0. good - 1. bad): ~\\
- 0.0000000E+00  -   5.0000001E-02       27648     81.81818      \% ~\\
- 5.0000001E-02  -   0.1000000               0    0.0000000E+00  \% ~\\
-... ~\\
-\end{lyxcode}
-The first line means that you have 27,648 elements with a skewness value between 0 and 0.05 (which means the element is basically not skewed, just plain regular hexahedral element).
-The total number of elements you have in this mesh is (see in the \texttt{output\_mesher.txt} file a bit further down):
-\begin{lyxcode}
-...~\\
-total number of elements in entire mesh:        33792 ~\\
-... ~\\
-\end{lyxcode}
-which gives you that:  27,648 / 33,792 $\sim$ 81.8 \% of all elements are not skewed, i.e. regular elements. a fantastic value :)
-
-The histogram lists for this mesh also some stronger skewed elements, for example the worst ones belong to:
-\begin{lyxcode}
-...~\\
-0.6000000      -   0.6500000            2048     6.060606      \% ~\\
-... ~\\
-\end{lyxcode}
-
-about 6 \% of all elements have distortions with a skewness value between 0.6 and 0.65.
-The skewness values give you a hint of how good your mesh is.
-In an ideal world, you would want to have no distortions, just like the 81\% from above.
-Those elements give you the best approximate values by the GLL quadrature used in the spectral-element method.
-However, having weakly distorted elements is still fine and the solutions are still accurate enough.
-So empirically, values up to around 0.7 are tolerable, above that you should consider remeshing...
-
-To give you an idea why some of the elements are distorted, see the following figure \ref{fig:mesh.vp} of the mesh you obtain in the example \texttt{examples/meshfem3D\_examples/simple\_model/}.
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=.9\textwidth]{figures/mesh_vp.jpg}
-\par\end{centering}
-\caption{Paraview visualization using the mesh vtk-files for the example given in \texttt{examples/meshfem3D\_examples/simple\_model/}.}
-\label{fig:mesh.vp}
-\end{figure}
-You will see that the mesh contains a doubling layer, where we stitch elements together such that the size of two elements will transition to the size of one element (very useful to keep the ratio of wavespeed / element\_size about constant).
-Those elements in this doubling layer have higher skewness values and make up those 6 \% in the histogram.
-
-\item [the code gives following error message "need at least one receiver":]
-This means that no stations given in the input file \texttt{in\_data\_files/STATIONS} could be located within the dimensions of the mesh.
-This can happen for example when the mesh was created with the in-house mesher \texttt{xmeshfem3D} while using the Universal Transverse Mercator (UTM) projection but the simulation with \texttt{xspecfem3D} was suppressing this projection from latitude/longitude to x/y/z coordinates.
-
-In such cases, try to change your \texttt{in\_data\_files/Par\_file} and set e.g.:
-\begin{lyxcode}
-SUPPRESS\_UTM\_PROJECTION         = .false. ~\\
-\end{lyxcode}
-to be the same in \texttt{Mesh\_Par\_file} and \texttt{Par\_file}.
-This flag should be identical when using the in-house mesher \texttt{xmeshfem3D}, \texttt{xgenerate\_databases} and \texttt{xspecfem3D} together to run simulations.
-
-The flag determines if the coordinates you specify for your source and station locations are given as lat/lon degrees and
-must be converted to UTM coordinates.
-As an example, if you use \texttt{.false.} within \texttt{Mesh\_Par\_file} then you create a mesh with \texttt{xmeshfem3D} using the UTM projection from lat/lon as input format to UTM projected coordinates to store the mesh point positions, which is fine.
-The error then may occur if in the \texttt{Par\_file} you have this set to \texttt{.true.} so that the \texttt{xgenerate\_databases} and \texttt{xspecfem3D} suppress the UTM projection and assume that all coordinates you use now for source and receiver locations are given in meters (that is, converted) already.
-So it won't find the specified locations in the used mesh.
-As a solutions, just change the flag in \texttt{Par\_file} to be the same as in \texttt{Mesh\_Par\_file} and rerun \texttt{xgenerate\_databases} and \texttt{xspecfem3D} to make sure that your simulation works fine.
-
-\item [I get the following error message "forward simulation became unstable and blew up":]
-In most cases this means that your time step size \texttt{DT} is chosen too big.
-Look at your files \texttt{output\_mesher.txt} or \texttt{output\_solver.txt} created in the folder \texttt{in\_out\_files/OUTPUT\_FILES/}.
-In these output files, find the section:
-\begin{lyxcode}
-... ~\\
-********************************************* ~\\
-*** Verification of simulation parameters *** ~\\
-********************************************* ~\\
-... ~\\
-*** Minimum period resolved =    4.308774 ~\\
-*** Maximum suggested time step =   6.8863556E-02 ~\\
-... ~\\
-... ~\\
-\end{lyxcode}
-
-then change \texttt{DT} in the \texttt{in\_data\_files/Par\_file} to be somewhere close to the maximum suggested time step. In the example above: \\
-\texttt{DT                              = 0.05d0} \\
-would (most probably) work fine. It could be also bigger than the 0.068 s suggested. This depends a bit on the distortions of
-your mesh elements. The more regular they are, the bigger you can choose \texttt{DT}. Just play with this value a bit and see
-when the simulation becomes stable ...
-
-\end{description}
-
-\chapter{\label{cha:License}License }
-
-\textbf{GNU GENERAL PUBLIC LICENSE Version 2, June 1991. Copyright
-(C) 1989, 1991 Free Software Foundation, Inc. 59 Temple Place, Suite
-330, Boston, MA 02111-1307 USA} \\
-Everyone is permitted to copy and distribute verbatim copies of this
-license document, but changing it is not allowed.
-
-\section*{Preamble}
-
-The licenses for most software are designed to take away your freedom
-to share and change it. By contrast, the GNU General Public License
-is intended to guarantee your freedom to share and change free software
--- to make sure the software is free for all its users. This General
-Public License applies to most of the Free Software Foundation's software
-and to any other program whose authors commit to using it. (Some other
-Free Software Foundation software is covered by the GNU Library General
-Public License instead.) You can apply it to your programs, too.
-
-When we speak of free software, we are referring to freedom, not price.
-Our General Public Licenses are designed to make sure that you have
-the freedom to distribute copies of free software (and charge for
-this service if you wish), that you receive source code or can get
-it if you want it, that you can change the software or use pieces
-of it in new free programs; and that you know you can do these things.
-
-To protect your rights, we need to make restrictions that forbid anyone
-to deny you these rights or to ask you to surrender the rights. These
-restrictions translate to certain responsibilities for you if you
-distribute copies of the software, or if you modify it.
-
-For example, if you distribute copies of such a program, whether gratis
-or for a fee, you must give the recipients all the rights that you
-have. You must make sure that they, too, receive or can get the source
-code. And you must show them these terms so they know their rights.
-
-We protect your rights with two steps:
-
-\begin{enumerate}
-\item Copyright the software, and
-\item Offer you this license which gives you legal permission to copy, distribute
-and/or modify the software.
-\end{enumerate}
-Also, for each author's protection and ours, we want to make certain
-that everyone understands that there is no warranty for this free
-software. If the software is modified by someone else and passed on,
-we want its recipients to know that what they have is not the original,
-so that any problems introduced by others will not reflect on the
-original authors' reputations.
-
-Finally, any free program is threatened constantly by software patents.
-We wish to avoid the danger that redistributors of a free program
-will individually obtain patent licenses, in effect making the program
-proprietary. To prevent this, we have made it clear that any patent
-must be licensed for everyone's free use or not licensed at all.
-
-The precise terms and conditions for copying, distribution and modification
-follow.
-
-\section*{GNU GENERAL PUBLIC LICENSE TERMS AND CONDITIONS FOR COPYING, DISTRIBUTION
-AND MODIFICATION }
-
-\begin{itemize}
-
-\item[0.]This License applies to any program or other work which
-contains a notice placed by the copyright holder saying it may be
-distributed under the terms of this General Public License. The ``Program''
-below refers to any such program or work, and a ``work based on the
-Program'' means either the Program or any derivative work under copyright
-law: that is to say, a work containing the Program or a portion of
-it, either verbatim or with modifications and/or translated into another
-language. (Hereinafter, translation is included without limitation
-in the term ``modification.'') Each licensee is addressed as ``you.''\\
-\\
-Activities other than copying, distribution and modification are not
-covered by this License; they are outside its scope. The act of running
-the Program is not restricted, and the output from the Program is
-covered only if its contents constitute a work based on the Program
-(independent of having been made by running the Program). Whether
-that is true depends on what the Program does.
-
-\end{itemize}
-
-\begin{enumerate}
-\item You may copy and distribute verbatim copies of the Program's source
-code as you receive it, in any medium, provided that you conspicuously
-and appropriately publish on each copy an appropriate copyright notice
-and disclaimer of warranty; keep intact all the notices that refer
-to this License and to the absence of any warranty; and give any other
-recipients of the Program a copy of this License along with the Program.
-
-You may charge a fee for the physical act of transferring a copy,
-and you may at your option offer warranty protection in exchange for
-a fee.
-
-\item You may modify your copy or copies of the Program or any portion of
-it, thus forming a work based on the Program, and copy and distribute
-such modifications or work under the terms of Section 1 above, provided
-that you also meet all of these conditions:
-
-\begin{enumerate}
-\item You must cause the modified files to carry prominent notices stating
-that you changed the files and the date of any change.
-\item You must cause any work that you distribute or publish, that in whole
-or in part contains or is derived from the Program or any part thereof,
-to be licensed as a whole at no charge to all third parties under
-the terms of this License.
-\item If the modified program normally reads commands interactively when
-run, you must cause it, when started running for such interactive
-use in the most ordinary way, to print or display an announcement
-including an appropriate copyright notice and a notice that there
-is no warranty (or else, saying that you provide a warranty) and that
-users may redistribute the program under these conditions, and telling
-the user how to view a copy of this License. (Exception: if the Program
-itself is interactive but does not normally print such an announcement,
-your work based on the Program is not required to print an announcement.)
-\end{enumerate}
-These requirements apply to the modified work as a whole. If identifiable
-sections of that work are not derived from the Program, and can be
-reasonably considered independent and separate works in themselves,
-then this License, and its terms, do not apply to those sections when
-you distribute them as separate works. But when you distribute the
-same sections as part of a whole which is a work based on the Program,
-the distribution of the whole must be on the terms of this License,
-whose permissions for other licensees extend to the entire whole,
-and thus to each and every part regardless of who wrote it.
-
-Thus, it is not the intent of this section to claim rights or contest
-your rights to work written entirely by you; rather, the intent is
-to exercise the right to control the distribution of derivative or
-collective works based on the Program.
-
-In addition, mere aggregation of another work not based on the Program
-with the Program (or with a work based on the Program) on a volume
-of a storage or distribution medium does not bring the other work
-under the scope of this License.
-
-\item You may copy and distribute the Program (or a work based on it, under
-Section 2) in object code or executable form under the terms of Sections
-1 and 2 above provided that you also do one of the following:
-
-\begin{enumerate}
-\item Accompany it with the complete corresponding machine-readable source
-code, which must be distributed under the terms of Sections 1 and
-2 above on a medium customarily used for software interchange; or,
-\item Accompany it with a written offer, valid for at least three years,
-to give any third party, for a charge no more than your cost of physically
-performing source distribution, a complete machine-readable copy of
-the corresponding source code, to be distributed under the terms of
-Sections 1 and 2 above on a medium customarily used for software interchange;
-or,
-\item Accompany it with the information you received as to the offer to
-distribute corresponding source code. (This alternative is allowed
-only for noncommercial distribution and only if you received the program
-in object code or executable form with such an offer, in accord with
-Subsection b above.)
-\end{enumerate}
-The source code for a work means the preferred form of the work for
-making modifications to it. For an executable work, complete source
-code means all the source code for all modules it contains, plus any
-associated interface definition files, plus the scripts used to control
-compilation and installation of the executable. However, as a special
-exception, the source code distributed need not include anything that
-is normally distributed (in either source or binary form) with the
-major components (compiler, kernel, and so on) of the operating system
-on which the executable runs, unless that component itself accompanies
-the executable.
-
-If distribution of executable or object code is made by offering access
-to copy from a designated place, then offering equivalent access to
-copy the source code from the same place counts as distribution of
-the source code, even though third parties are not compelled to copy
-the source along with the object code.
-
-\item You may not copy, modify, sublicense, or distribute the Program except
-as expressly provided under this License. Any attempt otherwise to
-copy, modify, sublicense or distribute the Program is void, and will
-automatically terminate your rights under this License. However, parties
-who have received copies, or rights, from you under this License will
-not have their licenses terminated so long as such parties remain
-in full compliance.
-\item You are not required to accept this License, since you have not signed
-it. However, nothing else grants you permission to modify or distribute
-the Program or its derivative works. These actions are prohibited
-by law if you do not accept this License. Therefore, by modifying
-or distributing the Program (or any work based on the Program), you
-indicate your acceptance of this License to do so, and all its terms
-and conditions for copying, distributing or modifying the Program
-or works based on it.
-\item Each time you redistribute the Program (or any work based on the Program),
-the recipient automatically receives a license from the original licensor
-to copy, distribute or modify the Program subject to these terms and
-conditions. You may not impose any further restrictions on the recipients'
-exercise of the rights granted herein. You are not responsible for
-enforcing compliance by third parties to this License.
-\item If, as a consequence of a court judgment or allegation of patent infringement
-or for any other reason (not limited to patent issues), conditions
-are imposed on you (whether by court order, agreement or otherwise)
-that contradict the conditions of this License, they do not excuse
-you from the conditions of this License. If you cannot distribute
-so as to satisfy simultaneously your obligations under this License
-and any other pertinent obligations, then as a consequence you may
-not distribute the Program at all. For example, if a patent license
-would not permit royalty-free redistribution of the Program by all
-those who receive copies directly or indirectly through you, then
-the only way you could satisfy both it and this License would be to
-refrain entirely from distribution of the Program.
-
-If any portion of this section is held invalid or unenforceable under
-any particular circumstance, the balance of the section is intended
-to apply and the section as a whole is intended to apply in other
-circumstances.
-
-It is not the purpose of this section to induce you to infringe any
-patents or other property right claims or to contest validity of any
-such claims; this section has the sole purpose of protecting the integrity
-of the free software distribution system, which is implemented by
-public license practices. Many people have made generous contributions
-to the wide range of software distributed through that system in reliance
-on consistent application of that system; it is up to the author/donor
-to decide if he or she is willing to distribute software through any
-other system and a licensee cannot impose that choice.
-
-This section is intended to make thoroughly clear what is believed
-to be a consequence of the rest of this License.
-
-\item If the distribution and/or use of the Program is restricted in certain
-countries either by patents or by copyrighted interfaces, the original
-copyright holder who places the Program under this License may add
-an explicit geographical distribution limitation excluding those countries,
-so that distribution is permitted only in or among countries not thus
-excluded. In such case, this License incorporates the limitation as
-if written in the body of this License.
-\item The Free Software Foundation may publish revised and/or new versions
-of the General Public License from time to time. Such new versions
-will be similar in spirit to the present version, but may differ in
-detail to address new problems or concerns.
-
-Each version is given a distinguishing version number. If the Program
-specifies a version number of this License which applies to it and
-``any later version,'' you have the option of following the terms
-and conditions either of that version or of any later version published
-by the Free Software Foundation. If the Program does not specify a
-version number of this License, you may choose any version ever published
-by the Free Software Foundation.
-
-\item If you wish to incorporate parts of the Program into other free programs
-whose distribution conditions are different, write to the author to
-ask for permission. For software which is copyrighted by the Free
-Software Foundation, write to the Free Software Foundation; we sometimes
-make exceptions for this. Our decision will be guided by the two goals
-of preserving the free status of all derivatives of our free software
-and of promoting the sharing and reuse of software generally.
-\end{enumerate}
-
-\subsection*{NO WARRANTY }
-
-\begin{itemize}
-
-\item[11.]BECAUSE THE PROGRAM IS LICENSED FREE OF CHARGE, THERE IS
-NO WARRANTY FOR THE PROGRAM, TO THE EXTENT PERMITTED BY APPLICABLE
-LAW. EXCEPT WHEN OTHERWISE STATED IN WRITING THE COPYRIGHT HOLDERS
-AND/OR OTHER PARTIES PROVIDE THE PROGRAM ``AS IS'' WITHOUT WARRANTY
-OF ANY KIND, EITHER EXPRESSED OR IMPLIED, INCLUDING, BUT NOT LIMITED
-TO, THE IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR
-PURPOSE. THE ENTIRE RISK AS TO THE QUALITY AND PERFORMANCE OF THE
-PROGRAM IS WITH YOU. SHOULD THE PROGRAM PROVE DEFECTIVE, YOU ASSUME
-THE COST OF ALL NECESSARY SERVICING, REPAIR OR CORRECTION.
-
-\item[12.]IN NO EVENT UNLESS REQUIRED BY APPLICABLE LAW OR AGREED
-TO IN WRITING WILL ANY COPYRIGHT HOLDER, OR ANY OTHER PARTY WHO MAY
-MODIFY AND/OR REDISTRIBUTE THE PROGRAM AS PERMITTED ABOVE, BE LIABLE
-TO YOU FOR DAMAGES, INCLUDING ANY GENERAL, SPECIAL, INCIDENTAL OR
-CONSEQUENTIAL DAMAGES ARISING OUT OF THE USE OR INABILITY TO USE THE
-PROGRAM (INCLUDING BUT NOT LIMITED TO LOSS OF DATA OR DATA BEING RENDERED
-INACCURATE OR LOSSES SUSTAINED BY YOU OR THIRD PARTIES OR A FAILURE
-OF THE PROGRAM TO OPERATE WITH ANY OTHER PROGRAMS), EVEN IF SUCH HOLDER
-OR OTHER PARTY HAS BEEN ADVISED OF THE POSSIBILITY OF SUCH DAMAGES.
-
-\end{itemize}
-
-\section*{END OF TERMS AND CONDITIONS }
-
-\subsection*{How to Apply These Terms to Your New Programs}
-
-If you develop a new program, and you want it to be of the greatest
-possible use to the public, the best way to achieve this is to make
-it free software which everyone can redistribute and change under
-these terms.
-
-To do so, attach the following notices to the program. It is safest
-to attach them to the start of each source file to most effectively
-convey the exclusion of warranty; and each file should have at least
-the ``copyright'' line and a pointer to where the full notice is found.
-For example:
-
-\begin{quote}
-One line to give the program's name and a brief idea of what it does.
-Copyright {\footnotesize $\copyright$ (}year) (name of author)
-
-This program is free software; you can redistribute it and/or modify
-it under the terms of the GNU General Public License as published
-by the Free Software Foundation; either version 2 of the License,
-or (at your option) any later version.
-
-This program is distributed in the hope that it will be useful, but
-WITHOUT ANY WARRANTY; without even the implied warranty of MERCHANTABILITY
-or FITNESS FOR A PARTICULAR PURPOSE. See the GNU General Public License
-for more details.
-
-You should have received a copy of the GNU General Public License
-along with this program; if not, write to the Free Software Foundation,
-Inc., 59 Temple Place, Suite 330, Boston, MA 02111-1307 USA
-\end{quote}
-Also add information on how to contact you by electronic and paper
-mail.
-
-If the program is interactive, make it output a short notice like
-this when it starts in an interactive mode:
-
-\begin{quote}
-Gnomovision version 69, Copyright $\copyright$ year name of author Gnomovision
-comes with ABSOLUTELY NO WARRANTY; for details type `show w'. This
-is free software, and you are welcome to redistribute it under certain
-conditions; type `show c' for details.
-\end{quote}
-The hypothetical commands `show w' and `show c' should show the appropriate
-parts of the General Public License. Of course, the commands you use
-may be called something other than `show w' and `show c'; they could
-even be mouse-clicks or menu items -- whatever suits your program.
-
-You should also get your employer (if you work as a programmer) or
-your school, if any, to sign a ``copyright disclaimer'' for the program,
-if necessary. Here is a sample; alter the names:
-
-\begin{quote}
-Yoyodyne, Inc., hereby disclaims all copyright interest in the program
-`Gnomovision' (which makes passes at compilers) written by James Hacker.
-
-(signature of Ty Coon)\\
-1 April 1989 \\
-Ty Coon, President of Vice
-\end{quote}
-This General Public License does not permit incorporating your program
-into proprietary programs. If your program is a subroutine library,
-you may consider it more useful to permit linking proprietary applications
-with the library. If this is what you want to do, use the GNU Library
-General Public License instead of this License.
-\end{document}
+executable. Of course it is best to include as many stations as possible,
+since this does not add to the cost of the simulation.
+
+We have also added the ability to run several calculations (several earthquakes)
+in an embarrassingly-parallel fashion from within the same run;
+this can be useful when using a very large supercomputer to compute
+many earthquakes in a catalog, in which case it can be better from
+a batch job submission point of view to start fewer and much larger jobs,
+each of them computing several earthquakes in parallel.
+To turn that option on, set parameter \texttt{NUMBER\_OF\_SIMULTANEOUS\_RUNS}
+to a value greater than 1 in file setup/constants.h.in before
+configuring and compiling the code.
+When that option is on, of course the number of processor cores used to start
+the code in the batch system must be a multiple of \texttt{NUMBER\_OF\_SIMULTANEOUS\_RUNS},
+all the individual runs must use the same number of processor cores,
+which as usual is \texttt{NPROC} in the input file \texttt{DATA/Par\_file},
+and thus the total number of processor cores to request from the batch system
+should be \texttt{NUMBER\_OF\_SIMULTANEOUS\_RUNS $\times$ NPROC}.
+All the runs to perform must be placed in directories called \texttt{run0001}, \texttt{run0002}, \texttt{run0003} and so on (with exactly four digits)
+and you must create a link from the root directory of the code
+to the first copy of the executable programs by typing \texttt{ln -s run0001/bin bin}.

### 2015-04-24 doc/USER_MANUAL/06_fault_sources.tex

\chapter{Kinematic and dynamic fault sources}\label{cha:Fault-Kinematics-Dynamics}

SPECFEM3D can handle finite fault sources of two kinds:
\begin{enumerate}
\item \emph{Kinematic}: the spatio-temporal distribution of slip rate is
prescribed along the fault surface
\item \emph{Dynamic}: a friction law and initial stresses are prescribed
on the fault, a spontaneous rupture process is computed.
\end{enumerate}

\section{Mesh Generation with Split Nodes}\label{sec:Mesh-Generation-with}

Faults need to be handled in a special way during mesh generation.
A fault surface must lie at the interface between elements (the mesh
must honor the fault surfaces). Moreover, a fault is made of two surfaces
in contact. Each of these two surfaces needs a separate set of nodes.
This approach is known as \textquotedbl{}split nodes\textquotedbl{}.
Currently faults can only be run with \texttt{xdecompose\_mesh} and
\texttt{CUBIT}. \texttt{xmeshfem3D} is not yet ready to handle faults.\\
To facilitate the mesh generation with split nodes in CUBIT, we need
to separate the two fault surfaces by a small distance, effectively
creating a tiny opening of the fault (Figure \ref{fig:examples.splitnodes},
\ref{fig:examples.splitnodes-surfacetrace}). Note that only the interior
of the fault must be opened, its edges must remain closed (except
the edge on the free surface). The fault is automatically closed later
by SPECFEM3D.\\

Here is an example CUBIT script to generate a mesh with split nodes
for a buried vertical strike-slip fault:
\begin{lyxcode}
~reset~~\\
~brick~x~10~y~10~z~10~~\\
~webcut~volume~all~with~plane~xplane~~\\
~webcut~volume~all~with~plane~yplane~~\\
~webcut~volume~all~with~plane~xplane~offset~3~~\\
~webcut~volume~all~with~plane~zplane~offset~3~~\\
~webcut~volume~all~with~plane~zplane~offset~-3~~\\
~imprint~all~~\\
~merge~all~~\\
~unmerge~surf~160~~\\
~mesh~vol~all~~\\
~set~node~constraint~off~~\\
~node~in~surf~168~move~X~0~Y~0.01~Z~0~~\\
~node~in~surf~160~move~X~0~Y~-0.01~Z~0~\\

\end{lyxcode}
\begin{figure}[htbp]
\noindent \begin{centering}
\includegraphics[scale=0.3]{figures/faultmesh.jpg} \\
\includegraphics[scale=0.3]{figures/surf168.jpg} 
\includegraphics[scale=0.3]{figures/splitnodes.jpg}
\par\end{centering}

\caption{Screenshots of the CUBIT example for the embedded fault with split
nodes: The entire mesh is decomposed into several volumes bounding
the fault (top), the fault surface 168 is shown as orange squares
(middle) and split nodes of the fault shown as orange dots(bottom).
Note that only interior nodes of the fault are split while those on
the edges of the fault surface are touching each other.}

\label{fig:examples.splitnodes}
\end{figure}

\begin{lyxcode}

\end{lyxcode}
\begin{figure}[htbp]
\noindent \begin{centering}
\includegraphics[scale=0.55]{figures/splitnodes_surfacetrace.pdf} \\

\par\end{centering}

\caption{Screenshots of a CUBIT example showing split nodes for a fault reaching
the surface. Surface trace of the fault is shown in orange. Note that
the edges of the fault are not split while the interior nodes are
offset by a small distance on either side of the fault}

\label{fig:examples.splitnodes-surfacetrace}
\end{figure}

The CUBIT scripts ({*}.jou and {*}.py) in the directory \texttt{EXAMPLES}
generate more complicated meshes. The {*}.py files are Python scripts
that execute CUBIT commands and use the CUBIT-python interface for
SPECFEM3D (see next section). The Python language allows to define
and manipulate variables to parameterize the mesh. Alternatively,
the Python script can call a CUBIT journal file ({*}.jou), which looks
like the example above. Variables can be defined and manipulated there
using the $\mathtt{APREPRO}$ language built in CUBIT.\\

Note that you should avoid gaps in the list of indices of mesh objects
with the following CUBIT command:
\begin{lyxcode}
$\mathtt{compress\: ids\: hex\: face\: edge\: node}$
\end{lyxcode}
(otherwise you will get a segmentation fault during domain decomposition)

\section{CUBIT-Python Scripts for Faults}

The mesh generated in CUBIT needs to be processed and exported in
a format compatible with SPECFEM3D. This is achieved in the Python
scripts by calling the Python-CUBIT interface functions defined in
the CUBIT directory:
\begin{enumerate}
\item Function $\mathtt{define\_bc}$ (or $\mathtt{boundary\_definition.py}$)
must be called to set up the absorbing boundaries database
\item Function $\mathtt{fault\_input}$ must be called once for each fault
to set up the fault database
\item Function $\mathtt{cubit2specfem3d.export2SPECFEM3D}$ must be called
at the very end of the script to export the mesh in a SPECFEM3D format.
\end{enumerate}
The functions in \#1 and \#3 are for the bulk and are documented in
Section \ref{subsec:Exporting-the-Mesh}. We focus here on \#2:
\begin{description}
\item [{Function:}] $\mathtt{fault\_input}$
\item [{Purpose:}] export a fault mesh database from CUBIT to a SPECFEM3D-compliant
file
\item [{Syntax:}] fault\_input(id\_fault, ids\_surf\_1, ids\_surf\_2)
\item [{Inputs:}]~\end{description}
\begin{itemize}
\item id\_fault integer index assigned to the fault. (The user must number
all the faults, starting at 1, with unit increments)
\item ids\_surf\_1 list of CUBIT ids of all surfaces that form side 1 of
the fault.
\item ids\_surf\_2 list of CUBIT ids of all surfaces that form side 2 of
the fault. (The user must decide which side of the fault is side 1.
This choice affects the sign conventions of fault quantities as explained
in Section \ref{sec:Sign-Convention-for}).\end{itemize}
\begin{description}
\item [{Outputs:}] file $\mathtt{fault\_file\_X.dat}$, where X is the
fault id (id\_fault).
\item [{Example:}] For the example in Section \ref{sec:Mesh-Generation-with}:\end{description}
\begin{lyxcode}
$\hphantom{}$A1~=~{[}168{]}

$\hphantom{}$A2~=~{[}160{]}

$\hphantom{}$fault\_input(1,A1,A2)
\end{lyxcode}

\section{Examples}

The package includes examples, the SCEC benchmark problems:
\begin{itemize}
\item TPV5, a planar vertical strike-slip fault
\item TPV14 and TPV15, vertical strike-slip fault system with a fault branch
\item TPV102 and TPV103, rate-and-state benchmarks
\item TPV22 and TPV23, step-over benchmarks
\item TPV16, heterogenous initial stress
\end{itemize}
and
\begin{itemize}
\item Splay fault models from Wendt et al. (2009)\\

\end{itemize}
Read the documents in the directory \texttt{EXAMPLES/*/description}.
They contain a description of the example and additional instructions
to run it. Visualize the results with the Matlab scripts in the directory
$\mathtt{EXAMPLES/*/post}$

\section{Sign Convention for Fault Quantities}\label{sec:Sign-Convention-for}

During mesh generation, the fault is defined by two surfaces in contact.
Let's denote as \textquotedbl{}\emph{side 1}\textquotedbl{} the SECOND
surface declared by the user in the call to the python function \textquotedbl{}\emph{fault\_input}\textquotedbl{},
and the FIRST surface as \textquotedbl{}\emph{side 2}\textquotedbl{}.
The local coordinate system on the fault is defined as the right-handed
coordinate system defined by (strike, dip, normal), where \textquotedbl{}\emph{normal}\textquotedbl{}
is the normal vector outgoing from side 1, \textquotedbl{}\emph{dip}\textquotedbl{}
is parallel to the along-dip direction pointing downwards, and \textquotedbl{}\emph{strike}\textquotedbl{}
is the horizontal along-strike vector such that the system is right-handed.\\

Slip is defined as displacement on side 2 minus displacement on side
1. In the local coordinate system on the fault, positive along-strike
slip is right-lateral and positive along-dip slip is thrust if side
1 is on the hanging wall (normal faulting if side 1 is on the foot
wall).\\

Traction is defined as the stress induced on side 1 by side 2, which
is the stress tensor times the normal vector outgoing from side 1.
In the local coordinate system on the fault, the normal traction is
negative in compression, positive along-strike traction generates
right-lateral slip and positive along-dip traction generates thrust
slip if side 1 is on the hanging wall (normal faulting if side 1 is
on the foot wall).

\section{Input Files}

Three additional input files are required in the \texttt{DATA/} directory
for dynamic and kinematic faults. They are $\mathtt{Par\_file\_faults}$,
$\mathtt{FAULT\_STATIONS}$ and $\mathtt{input\_file.txt}$ (optional).
If the former does not exist, the code assumes that there are no faults.\\

\begin{lyxlist}{00.00.0000}
\item [{\textbf{DATA/Par\_file\_faults}}] contains parameters of the fault.
The first part of this file has a strict format:

\begin{description}
\item [{Line}] 1: Number of faults (NF)
\item [{Lines}] 2 to NF+1: Kelvin Voigt damping (in seconds) for each fault.
(See below how to set this parameter)
\item [{Line}] NF+2: Type of simulation (1=dynamic , 2 = kinematic)
\item [{Line}] NF+3: Number of time steps between updates of the time series
outputs at selected fault points (see DATA/FAULT\_STATIONS), usually
a large number (100s or 1000s). Note that the sampling rate of the
time series is usually much higher.
\item [{Line}] NF+4: Number of time steps between fault snapshot outputs
(quantities at every fault point exported at regular times), usually
a large number (100s or 1000s).
\item [{Line}] NF+5: Slip velocity threshold below which frictional healing
is set (friction coefficient is reset to its static value). If this
value is negative healing is disabled.
\item [{Line}] NF+6: Slip velocity threshold to define the rupture front.
Only used for outputs.
\end{description}

The rest of this file is made of namelist input blocks (see \textquotedbl{}namelist\textquotedbl{}
in a Fortran 9x manual). The input for each fault has the following
sequence (arguments in {[}brackets{]} are optional):

\&\textbf{BEGIN\_FAULT} /

\&\textbf{INIT\_STRESS} S1, S2, S3 {[},n1, n2, n3{]} /

followed by (n1+n2+n3) \&\textbf{DIST2D} blocks

\&\textbf{SWF} mus, mud, dc {[}, nmus, nmud, ndc{]} /

followed by (nmus+nmud+ndc) \&\textbf{DIST2D} blocks\\

\begin{lyxlist}{00.00.0000}
\item [{\&\textbf{INIT\_STRESS}}] input block sets the initial fault stresses
relative to the foot-wall side of the fault. Initial stresses are
composed of a constant background value possibly overwritten in prescribed
regions by heterogeneous distributions (see \&\textbf{DIST2D} blocks
below):

\begin{description}
\item [{S1}] = constant background value of along-strike shear stress (positive
in the usual strike direction)
\item [{S2}] = constant background value of along-dip shear (positive is
down-dip, normal faulting)
\item [{S3}] = constant background value of normal stress (negative in
compression)
\item [{n1}] = number of heterogeneous items for along-strike shear stress
{[}default is 0{]}
\item [{n2}] = number of heterogeneous items for along-dip shear stress
{[}default is 0{]}
\item [{n3}] = number of heterogeneous items for normal stress {[}default
is 0{]}
\end{description}
\item [{\&\textbf{SWF}}] input block sets the slip-weakening friction parameters
of the fault:

\begin{description}
\item [{mus}] = constant background value of static friction coefficient
\item [{mud}] = constant background value of dynamic friction coefficient
\item [{dc}] = constant background value of critical slip-weakening distance
\item [{nmus}] = number of heterogeneous items for static friction coefficient
{[}default is 0{]}
\item [{nmud}] = number of heterogeneous items for dynamic friction coefficient
{[}default is 0{]}
\item [{ndc}] = number of heterogeneous items for critical slip-weakening
distance {[}default is 0{]}\\

\end{description}
\item [{\&\textbf{DIST2D}}] input blocks modify (overwrite) the value of
a fault parameter by a heterogeneous spatial distribution:

\&\textbf{DIST2D} shape='$\mathtt{square}$', val, xc, yc, zc, l /

$\;$sets a constant value (val) within a cube with center (xc,yc,zc)
and edge size l.

\&\textbf{DIST2D} shape='$\mathtt{rectangle}$', val, xc, yc, zc,
lx, ly, lz /

$\;$sets a constant value (val) within a rectangular prism with center
(xc,yc,zc) and edge sizes (lx,ly,lz).

\&\textbf{DIST2D} shape='$\mathtt{rectangle-taper}$', val, valh,
xc, yc, zc, lx, ly, lz /

$\;$sets a vertical linear gradient within a rectangular prism with
center (xc,yc,zc) and edge sizes (lx,ly,lz). Values vary linearly
as a function of vertical position z between value val at z = zc-lz/2
and value valh at z = zc+lz/2 .

\&\textbf{DIST2D} shape='$\mathtt{circular}$', val, xc, yc, zc, r
/

$\;$sets a constant value (val) within a sphere with center (xc,yc,zc)
and radius r. \\

\end{lyxlist}
\end{lyxlist}

\begin{lyxlist}{00.00.0000}
\item [{\textbf{DATA/FAULT\_STATIONS}}] Stations in the fault plane.

\textbf{Line 1} : number of stations.

\textbf{Line 2 to end}: 5 columns: X, Y, Z (-depth), station name,
fault-id

The fault-id identifies the fault that contains the station. It is
the index of appearance in the faults list after line 2 of Par\_file\_faults

\end{lyxlist}

\begin{lyxlist}{00.00.0000}
\item [{\textbf{DATA/input\_file.txt}}] Heterogeneous stresses and friction
parameters are documented in page 10 of

\begin{lyxlist}{00.00.0000}
\item [{$\mathtt{EXAMPLES/tpv16/description/TPV16\_17\_Description\_v03.pdf}$}]~
\end{lyxlist}

To activate this feature, in fault\_solver\_dynamic.f90 set \textbf{TPV16}
= .true. then re-compile the code.

\end{lyxlist}

\section{Setting the Kelvin-Voigt Damping Parameter}\label{sec:Setting-the-Kelvin-Voigt}

The purpose of the Kelvin-Voigt viscosity in the dynamic fault solver
is to damp spurious oscillations generated by the fault slip at frequencies
that are too high to be resolved by the mesh. The viscosity $\mathtt{eta}$
(in seconds) depends on the size of the elements on the fault. Here
is how to set it:
\begin{enumerate}
\item Determine the average linear size of the elements on the fault plane,
$\mathtt{h\_fault}$. Usually this value is prescribed by the user
during mesh generation. Otherwise it can be found by inspection of
the mesh inside the CUBIT GUI.
\item Use the Matlab function $\mathtt{utils/critical\_timestep.m}$ to
compute\\
 $\mathtt{dtc\_fault}=\mathtt{critical\_timestep\left(c_{p},h\_fault,ngll\right)}$.
\\
This is the critical time step in an elastic medium for a hypothetical
element of cubic shape with size equal to $\mathtt{h\_fault}$.
\item Set $\mathtt{eta}$ in $\mathtt{Par\_file\_faults}$ to (0.1 to 0.3)$\ensuremath{\times}\mathtt{dtc\_fault}$.
A larger $\mathtt{eta}$ damps high-frequencies more aggressively
but it might also affect lower frequencies and rupture speed.
\end{enumerate}
Viscosity reduces numerical stability: the critical timestep in a
simulation with Kelvin-Voigt damping needs to be smaller than that
in a purely elastic simulation. Here is how to set the time step accordingly:
\begin{enumerate}
\item Run a test simulation without viscosity ($\mathtt{eta}$=0 and only
a few time steps)
\item Look for the \textquotedbl{}maximum suggested time step\textquotedbl{}
in \texttt{OUTPUT\_FILES/output\_mesher.txt}. This is the critical
timestep of a purely elastic simulation, $\mathtt{dtc\_bulk}$.
\item Reset the timestep of the simulation with a Kelvin-Voigt material
to a value smaller than\\
 $\mathtt{dtc\_kv}=\mathtt{eta}\left(\sqrt{1+\mathtt{dtc\_bulk^{2}}/\mathtt{eta^{2}}}-1\right)$
\end{enumerate}
Note that in general $\mathtt{dtc\_bulk}$ is smaller than $\mathtt{dtc\_fault}$,
because elements off the fault might be smaller or more distorted
than element faces on the fault.

\section{Output Files }

Several output files are saved in \texttt{OUTPUT\_FILES/}:\\

\begin{enumerate}
\item Seismograms for each station on the fault plane given in \texttt{DATA/FAUL\_STATIONS}.
One output file is generated for each station, named after the station.
The files are ascii and start with a header (22 lines long) followed
by a data block with the following format, one line per time sample:\\

Column 1 = Time (s)

Column 2 = horizontal right-lateral slip (m)

Column 3 = horizontal right-lateral slip rate (m/s)

Column 4 = horizontal right-lateral shear stress (MPa)

Column 5 = vertical up-dip slip (m)

Column 6 = vertical up-dip slip rate (m/s)

Column 7 = vertical up-dip shear stress (MPa)

Column 8 = normal stress (MPa)\\

The stresses are relative to the footwall side of the fault (this
convention controls their sign, but not their amplitude). Slip is
defined as displacement of the hanging wall relative to the footwall.

\item Seismograms at stations in the bulk (out of the fault plane) given
in \texttt{DATA/STATIONS}.
\item Rupture time files are named $\mathtt{Rupture\_time\_FAULT}$-$\mathtt{id}$.
One file is generated for each fault. The files are ascii and start
with a header (12 lines long) followed by a data block with the following
format, one line per fault node:

Column 1 = horizontal coordinate, distance along strike (m)

Column 2 = vertical coordinate, distance down-dip (m)

Column 3 = rupture time (s)

\item Fault quantities (slip, slip rate, stresses, etc) at regular times
are stored in binary data files called $\mathtt{Snapshot\#it\#.bin}$,
where \#it\# is the timestep number. These can be read in Matlab with
the function $\mathtt{utils/FSEM3D\_snapshot.m}$
\end{enumerate}

\section{Post-processing and Visualization}

Some Matlab functions for post-processing and visualization are included
in directory $\mathtt{utils}$. The functions are internally documented
(see their matlab help).\\

\texttt{FSEM3D\_snapshot} reads a fault data snapshot\\

The directories \texttt{EXAMPLES/*/post} contain additional Matlab
scripts to generate figures specific to each example.

### 2015-04-24 doc/USER_MANUAL/08_doing_tomography.tex

\chapter[Doing tomography based on the sensitivity kernels obtained]{Doing tomography, i.e., updating the model based on the sensitivity kernels obtained}\label{cha:tomo}

UNDER CONSTRUCTION (April 2015). Will be added around June 2015.

\section{Tomographic full waveform inversion (FWI) / imaging using the sensitivity kernels obtained}

One of the fundamental reasons for computing sensitivity kernels (Section~\ref{sec:Adjoint-simulation-finite})
is to use them within a tomographic inversion, i.e., perform imaging. In other words, use
recorded seismograms, make measurements with synthetic seismograms,
and use the misfit between the two to iteratively improve the model
described by (at least) $V_{{\rm p}}$, $V_{{\rm s}}$, and $\rho$ as a function of space.

As explained for instance in \cite{MoChKoWa15},
full waveform inversion means that one considers the observed seismograms (possibly filtered) as the
basic observables that one wants to fit. One thus searches for the model that minimizes the mean squared difference between observed
and synthetic seismograms. In other words, the goal is to find a structural model that can explain a larger portion
of seismological records, and not simply the phase of a few seismic arrivals.

Whatever misfit function you use for the tomographic inversion (you can find several
examples in \citet{TrKoLi08} and \citet{TrTaLi05}), you will weight
the sensitivity kernels with measurements. Furthermore, you will use
as many measurements (stations, components, time windows) as possible
per event; hence, we call these composite kernels ``event kernels'',
which are volumetric fields representing the gradient of the misfit
function with respect to one of the variables (\eg $V_{{\rm s}}$).
The basic features of an adjoint-based tomographic inversion were
illustrated in \citet{TrKoLi08} and \citet{TaLiTr07} using a conjugate-gradient algorithm.
Other, more powerful techniques such as the Limited-Memory Broyden-Fletcher-Goldfarb-Shanno (L-BFGS) method can now be used,
as illustrated for instance in \cite{MoChKoWa15}.

\subsection{Principle}

We want to minimize the classical waveform misfit function:
\begin{equation}
\chi \left( \boldm \right) = \sum_{s=1}^N\sum_{r=1}^M \int_0^T \frac{1}{2} \|\boldu(\boldx_r, \boldx_s; t) - \boldd(\boldx_r, \boldx_s; t)\| ^{2} \, dt .
\label{misfit}
\end{equation}
This functional quantifies the $L^2$ difference between the observed waveforms $\boldd(\boldx_r, \boldx_s; t)$ at receivers $\boldx_r$, $r = 1, ..., M$
produced by sources at $\boldx_s$, $s=1,...,N$, and the corresponding synthetic seismograms $\boldu(\boldx_r, \boldx_s; t)$ computed in model $\boldm$.
While this misfit function is indeed classical, it is worth mentioning that in the case of noisy real data other norms could be used,
since in the oil industry for instance it is known that the $L^1$ norm
(\cite{CrPiNoMcTa90,Brossier_2010_WDR}), hybrid $L^1-L^2$ norms (\cite{Bube_1997_HL1}), Hubert norm (\cite{Ha_2009_WIU}),
Student-t distribution (\cite{ArVaHe11,Jeong_2015}) etc... can be more robust that the $L^2$ norm used here in the context of synthetic data with no noise.
In the vicinity of $\boldm$, the misfit function can be expanded into a Taylor series:
\begin{equation}
\chi(\boldm+ \delta\boldm) \approx \chi(\boldm)+\boldg(\boldm)\cdot\delta\boldm
+\delta\boldm\cdot\mathbf{H}(\boldm)\cdot\delta\boldm \, ,
\label{local_misfit}
\end{equation}
where $\boldg(\boldm)$ is the gradient of the waveform misfit function:
\begin{equation}
\boldg(\boldm) = \frac{\partial\chi(\boldm)}{\partial\boldm} \, ,
\end{equation}
and $\mathbf{H}(\boldm)$ the Hessian:
\begin{equation}
\mathbf{H}(\boldm) = \frac{\partial^2\chi(\boldm)}{\partial\boldm^2} \, .
\end{equation}
In the following, for simplicity the dependence of the gradient and Hessian on the model will be implicitly assumed and omitted in the notations.
The nearest minimum of $\chi$ in (\ref{local_misfit}) with respect to the model perturbation $\delta \boldm$ is reached for
\begin{equation}
\delta \boldm = -\mathbf{H}^{-1} \cdot \boldg \, .
\label{Newton}
\end{equation}
The local minimum of (\ref{misfit}) is thus given by perturbing the model in the direction of the gradient preconditioned
by the inverse Hessian.

\subsection{Computation of the gradient based on the adjoint method}

A direct method to compute the gradient is to take the derivative of (\ref{misfit}) with respect to model parameters:
\begin{equation}
\frac{\partial\chi(\boldm)}{\partial\boldm} = -\sum_{s=1}^N\sum_{r=1}^M \int_0^T
\frac{\partial \boldu(\boldx_r, \boldx_s; t)}{\partial\boldm}\cdot\left[\boldu(\boldx_r, \boldx_s; t) - \boldd(\boldx_r, \boldx_s; t)\right] \, dt \, .
\label{calc_gradient}
\end{equation}
This equation can be reformulated as the matrix-vector product:
\begin{equation}
\boldg = -\mathbf{J}^* \cdot \delta \mathbf{d} \, ,
\label{Jacobian}
\end{equation}
where $\mathbf{J}^*$ is the adjoint of the Jacobian matrix of the forward problem that contains the Fr\'echet derivatives of the
data with respect to model parameters, and $\delta \mathbf{d}$ is the vector that contains the data residuals.
The determination of $\mathbf{J}$ would require computing the Fr\'echet derivatives for each time step in the time
window considered and for all the source-station pairs, which is completely prohibitive on current supercomputers
(let us note that this situation may change one day).
However, it is possible to obtain this gradient without computing the Jacobian matrix explicitly.
The approach to determine the gradient without computing the Fr\'echet derivatives was introduced in nonlinear optimization
by \cite{Cha74} working with J. L. Lions, and later applied to seismic exploration problems by
\cite{BaChLa77}, \cite{BaChHeLa82}, \cite{Lai83} and \cite{Tar84}. The idea is to resort to the adjoint state,
which corresponds to the wavefield emitted and back-propagated from the receivers \cite[e.g.,][]{TrTaLi05,TrKoLi08,Plessix_2006_RAS,FiBuIg06}.

Let us give an outline of the theory to compute the gradient based on the adjoint method, and refer the reader to e.g. \cite{TrTaLi05}
and \cite{TrKoLi08} for further details. The perturbation of the misfit function can be expressed as:
\begin{equation}
\delta \chi (\boldm) = \sum_{s=1}^N\sum_{r=1}^M \int_0^T \left[\boldu(\boldx_r, \boldx_s; t) - \boldd(\boldx_r, \boldx_s; t)\right]\cdot \delta
\boldu(\boldx_r, \boldx_s; t) \, dt \, ,
\label{dchi}
\end{equation}
where $\delta\boldu$ is the perturbation of displacement given by the first-order Born approximation \cite[e.g.,][]{Hud77}:
\begin{eqnarray}
\delta \boldu (\boldx_r, \boldx_s; t) & = & -\int_0^t \int_V
\left[\delta\rho(\boldx)\mathbf{G}(\boldx_r,\boldx;t-t')\cdot\partial^2_{t'}\boldu(\boldx,\boldx_s;t') \right. \nonumber \\
& + & \left. \nabla \mathbf{G}(\boldx_r,\boldx;t-t'):\delta \boldc(\boldx): \nabla\boldu(\boldx;t') \right] \, d^3\boldx \, dt' \, .
\label{Born}
\end{eqnarray}
In this expression, $\mathbf{G}$ is the Green's tensor, $\delta\rho$ the perturbation of density, $\delta c$ the perturbation of the fourth-order elasticity
tensor,
and a colon denotes a double tensor contraction operation.
Inserting (\ref{Born}) into (\ref{dchi}) we obtain
\begin{eqnarray}
\delta \chi (\boldm) & = & -\sum_{s=1}^N\sum_{r=1}^M \int_0^T \left[\boldu(\boldx_r, \boldx_s; t) - \boldd(\boldx_r, \boldx_s; t)\right] \int_0^t \int_V
\left[\delta\rho(\boldx)\mathbf{G}(\boldx_r,\boldx;t-t')\cdot\partial^2_{t'}\boldu(\boldx,\boldx_s;t') \right. \nonumber \\
& + & \left. \nabla \mathbf{G}(\boldx_r,\boldx;t-t'):\delta \boldc(\boldx):\nabla\boldu(\boldx,t') \right] \, d^3\boldx \, dt' \, dt \,.
\end{eqnarray}
Defining the waveform adjoint source for each source $\boldx_s$
\begin{equation}
\boldf^\dagger(\boldx,\boldx_s;t) = \sum_{r=1}^M  \left[\boldu(\boldx_r, \boldx_s; T-t) - \boldd(\boldx_r, \boldx_s; T-t)\right] \delta(\boldx-\boldx_r) \, ,
\end{equation}
and the corresponding adjoint wavefield
\begin{equation}
\boldu^\dagger(\boldx,\boldx_s;t) = \int_0^{t'}\int_V\mathbf{G}(\boldx,\boldx';t'-t)\cdot\boldf^\dagger(\boldx',\boldx_s;t)\, d^3\boldx' \, dt \, ,
\end{equation}
the perturbation of the misfit function may be expressed as:
\begin{eqnarray}
\delta\chi(\boldm) & = & - \sum_{s=1}^N\nonumber \int_V \int_0^T \left[\delta\rho \, \boldu^\dagger(\boldx,\boldx_s;T-t)\cdot
\partial^2_{t}\boldu(\boldx,\boldx_s;t) \right. \nonumber \\
& +  & \left. \nabla\boldu^\dagger(\boldx,\boldx_s;T-t):\delta \boldc: \nabla \boldu(\boldx,\boldx_s;t) \right] \, d^3\boldx \, dt \, .
\label{dchi2}
\end{eqnarray}
At this point, we make some assumptions on the nature of the elasticity tensor.
A general fourth-order elasticity tensor is described by 21 elastic parameters, a very large number that makes its complete characterization
way beyond the reach of any tomographic approach. For the time being, let us thus consider isotropic elasticity tensors,
described by the two Lam\'e parameters $\lambda$ and $\mu$:
\begin{equation}
c_{ijkl} = \lambda \delta_{ij}\delta_{kl}+\mu(\delta_{ik}\delta_{jl}+\delta_{il}\delta_{jk}) \, .
\end{equation}
In this case, (\ref{dchi2}) can be written as:
\begin{eqnarray}
\delta\chi(\boldm) & = & - \sum_{s=1}^N\nonumber \int_V \left[ K_\rho(\boldx,\boldx_s)\delta\ln\rho(\boldx)
+K_\lambda(\boldx,\boldx_s)\delta\ln\lambda(\boldx) \right. \nonumber \\
& + & \left. K_\mu(\boldx,\boldx_s)\delta\ln\mu(\boldx)\right] \, d^3\boldx \, ,
\label{dchi2_1}
\end{eqnarray}
where ln() is the natural logarithm and where the Fr\'echet derivatives with respect to the density and Lam\'e parameters are given by:
\begin{eqnarray}
K_\rho(\boldx,\boldx_s) & = & - \int_0^T \rho(\boldx) \boldu^\dagger(\boldx,\boldx_s;T-t)\cdot \partial^2_{t}\boldu(\boldx,\boldx_s;t) \, dt \\
K_\lambda(\boldx,\boldx_s) & = & - \int_0^T \lambda(\boldx) \nabla\cdot\boldu^\dagger(\boldx,\boldx_s;T-t) \nabla\cdot\boldu(\boldx,\boldx_s;t) \, dt \\
K_\mu(\boldx,\boldx_s) & = & -2 \int_0^T \mu(\boldx) (\boldx) \nabla\boldu^\dagger(\boldx,\boldx_s;T-t) :\nabla\boldu(\boldx,\boldx_s;t)\, dt \, .
\end{eqnarray}
Since the propagation of seismic waves mainly depends on compressional wave speed $\alpha$ and shear wave speed $\beta$,
but also because these seismic velocities are easier to interpret, tomographic models are usually described based on these
two parameters. With this new parametrisation, the perturbation of the misfit function may be written as:
\begin{eqnarray}
\delta\chi(\boldm) & = & - \sum_{s=1}^N\nonumber \int_V \left[ K'_\rho(\boldx,\boldx_s)\delta\ln\rho(\boldx)
+K'_\alpha(\boldx,\boldx_s)\delta\ln\alpha(\boldx) \right. \nonumber \\
& + & \left. K'_\beta(\boldx,\boldx_s)\delta\ln\beta(\boldx)\right] \, d^3\boldx \, ,
\end{eqnarray}
where
\begin{eqnarray}
K'_\rho(\boldx,\boldx_s) & = &  K_\rho(\boldx,\boldx_s)+K_\lambda(\boldx,\boldx_s)+K_\mu(\boldx,\boldx_s) \\
K'_\alpha(\boldx,\boldx_s) & = &  2 \left(\frac{\lambda+2\mu}{\lambda}\right) K_\lambda(\boldx,\boldx_s) \\
K'_\beta(\boldx,\boldx_s) & = & 2 K_\mu -\frac{4\mu}{\lambda} K_\lambda (\boldx,\boldx_s) \, .
\end{eqnarray}

As can be seen from these expressions, the principle of the adjoint-state method is to correlate two wavefields:
the direct (i.e. forward) field that propagates from the source to the receivers, and the adjoint field that propagates from all the
receivers backward in time. The same approach can be followed for any type of seismic observable (phase, amplitude,
envelope, time series...), provided the appropriate adjoint source is used \cite[]{TrTaLi05,TrKoLi08}.
For example, for the cross-correlated traveltime of a seismic
phase, the adjoint source is defined as the velocity of that synthetic phase weighted by the travel-time residual.

Computing the gradient based on the adjoint-state method requires performing two simulations per source (forward and adjoint
fields) regardless of the type of observable. However, to define the adjoint field one
must know the adjoint source, and that source is computed from the results of the forward simulation. One must therefore perform
the forward simulation before the adjoint simulation.
A straightforward solution for time-domain methods would be to store the whole forward field to disk at
each time step during the forward run and then read it back during the adjoint simulation to calculate the interaction of these two fields.
In 2-D this is feasible but in the 3-D case for very short seismic periods and without lossy compression, downsampling,
or a large amount of disk or memory checkpointing \cite[e.g.,][]{FiKeIg09,RuHaPuGu14,CyShWi15}
the amount of disk storage required would currently be too large.
However let us note again that this situation will change in the future.
In the mean time, a standard possible solution is to perform three simulations per source \cite[]{TrKoLi08,PeKoLuMaLeCaLeMaLiBlNiBaTr11},
i.e., perform the forward calculation twice, once to compute the adjoint sources and once again
at the same time as the adjoint simulation to correlate the two fields and sum their interaction on the fly over all the time steps.
Doing so for an elastic Earth, one only needs a small amount of disk storage to store the last time
step of the forward run, which is then used as an initial condition to redo the forward run backwards,
as well as the field on the outer edges of the mesh for each time steps in order to be able to undo the absorbing boundary conditions.

\section{OLD VERSION OF THE SECTION, WILL SOON BE IMPROVED AND ENRICHED}

There are dozens of versions of gradient-based inversion
algorithms that could alternatively be used. The tomographic inversion
of \citet{TaLiMaTr09,TaLiMaTr2010} used SPECFEM3D Cartesian as well
as several additional components which are also stored on the CIG
svn server, described next.

The directory containing external utilities for tomographic inversion using
SPECFEM3D Cartesian (or other packages that evaluate misfit functions
and gradients) is in directory \texttt{utils/ADJOINT\_TOMOGRAPHY\_TOOLS/}:
\begin{verbatim}
flexwin/     -- FLEXWIN algorithm for automated picking of time windows
measure_adj/ -- reads FLEXWIN output file and makes measurements,
                with the option for computing adjoint sources
iterate_adj/ -- various tools for iterative inversion
                (requires pre-computed "event kernels")
\end{verbatim}
This directory also contains
a brief \verb+README+ file indicating the role of the three subdirectories,
\verb+flexwin+ \citep{Maggi2009}, \verb+measure_adj+, and \verb+iterate_adj+.
The components for making the model update are there; however, there
are no explicit rules for computing the model update, just as with
any optimization problem. There are options for computing a conjugate
gradient step, as well as a source subspace projection step.

The best single file to read is probably: \verb+ADJOINT_TOMO/iterate_adj/cluster/README+.

### 2015-04-24 doc/USER_MANUAL/11_graphics.tex

--- 2012-09-28 doc/USER_MANUAL/manual_SPECFEM3D.tex
@@
-%% LyX 1.5.1 created this file.  For more info, see http://www.lyx.org/.
-%% Do not edit unless you really know what you are doing.
-\documentclass[oneside,english,onecolumn]{book}
-\usepackage[T1]{fontenc}
-\usepackage[latin1]{inputenc}
-\usepackage{geometry}
-\geometry{letterpaper,tmargin=1in,bmargin=1in,lmargin=1in,rmargin=1in}
-\setcounter{secnumdepth}{3}
-\setcounter{tocdepth}{3}
-\usepackage{float}
-\usepackage{textcomp}
-\usepackage{amsmath}
-
-% figures
-\usepackage[pdftex]{graphicx}
-
-% we are running pdflatex, so convert .eps files to .pdf
-\usepackage{epstopdf}
-
-\usepackage{amssymb}
-\IfFileExists{url.sty}{\usepackage{url}}
-                      {\newcommand{\url}{\texttt}}
-\usepackage[authoryear]{natbib}
-
-\makeatletter
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% LyX specific LaTeX commands.
-\newcommand{\noun}[1]{\textsc{#1}}
-%% A simple dot to overcome graphicx limitations
-\newcommand{\lyxdot}{.}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Textclass specific LaTeX commands.
-\newenvironment{lyxcode}
-{\begin{list}{}{
-\setlength{\rightmargin}{\leftmargin}
-\setlength{\listparindent}{0pt}% needed for AMS classes
-\raggedright
-\setlength{\itemsep}{0pt}
-\setlength{\parsep}{0pt}
-\normalfont\ttfamily}%
- \item[]}
-{\end{list}}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% User specified LaTeX commands.
-%\renewcommand{\baselinestretch}{1.5}
-
-% figures
-\usepackage[dvips]{epsfig}
-
-\usepackage{wrapfig}
-
-% fonts
-\usepackage{times}
-
-% hyperlinks to sections and references
-\usepackage[pdftex,bookmarks=true,bookmarksnumbered=true,pdfpagemode=None,pdfstartview=FitH,pdfpagelayout=SinglePage,pdfborder={0 0 0}]{hyperref}
-
-\newcommand{\urlwithparentheses}[1]{(\url{#1})}
-
-% biblio GJI
-\bibliographystyle{abbrvnat}
-
-\newcommand{\toall}[1]{\textbf{*** All: #1 ***}}
-\newcommand{\tojeroen}[1]{\textbf{*** Jeroen: #1 ***}}
-\newcommand{\tobrian}[1]{\textbf{*** Brian: #1 ***}}
-\newcommand{\tovala}[1]{\textbf{*** Vala: #1 ***}}
-\newcommand{\tovalabrian}[1]{\textbf{*** Vala \& Brian: #1 ***}}
-\newcommand{\tovalaqinya}[1]{\textbf{*** Vala \& Qinya: #1 ***}}
-\newcommand{\toqinya}[1]{\textbf{*** Qinya: #1 ***}}
-\newcommand{\tomin}[1]{\textbf{*** Min: #1 ***}}
-\newcommand{\toalessia}[1]{\textbf{*** Alessia: #1 ***}}
-\newcommand{\todimitri}[1]{\textbf{*** Dimitri: #1 ***}}
-
-\newcommand{\nexxi}{\mbox{\texttt{NEX\_XI\/}}}
-\newcommand{\nexeta}{\mbox{\texttt{NEX\_ETA\/}}}
-\newcommand{\nprocxi}{\mbox{\texttt{NPROC\_XI\/}}}
-\newcommand{\nproceta}{\mbox{\texttt{NPROC\_ETA\/}}}
-\newcommand{\nchunks}{\mbox{\texttt{NCHUNKS\/}}}
-\newcommand{\eg}{e.g.,}
-
-% colors for the text
-\usepackage[dvipsnames,usenames]{color}
-\newcommand{\red}[1]{\textcolor{Red}{#1}}
-
-\usepackage{babel}
-\makeatother
-
-\begin{document}
-\begin{center}
-\thispagestyle{empty}%
-\vspace*{-1.8truecm}
-\noindent\makebox[\textwidth]{%
-\noindent\includegraphics[width=0.83\paperwidth]{figures/specfem_3d-cover.pdf}
-}
-\end{center}
-%
-\title{\textbf{SPECFEM3D}\\
-\textbf{User Manual}}
-
-\author{$\copyright$ Princeton University (USA) and\\
-CNRS / INRIA / University of Pau (France)\\
-Version 2.1}
-
-\date{\noindent \today}
-
-\maketitle
-
-\section*{Authors}
-The SPECFEM3D package was first developed by Dimitri Komatitsch and Jeroen Tromp at Harvard University and Caltech, USA, starting in 1998, based on earlier work by Dimitri Komatitsch and Jean-Pierre Vilotte at Institut de Physique du Globe (IPGP) in Paris, France from 1995 to 1997.\\
-
-Since then it has been developed and maintained by a development team: in alphabetical order,
-Piero Basini,
-C\'eline Blitz,
-Ebru Bozda\u{g},
-Emanuele Casarotti,
-Joseph Charles,
-Min Chen,
-Vala Hj\"orleifsd\'ottir,
-Sue Kientz,
-Dimitri Komatitsch,
-Jes\'us Labarta,
-Nicolas Le Goff,
-Pieyre Le Loher,
-Qinya Liu,
-Yang Luo,
-Alessia Maggi,
-Federica Magnoni,
-Roland Martin,
-Dennis McRitchie,
-Matthias Meschede,
-Peter Messmer,
-David Mich\'ea,
-Tarje Nissen-Meyer,
-Daniel Peter,
-Max Rietmann,
-Brian Savage,
-Bernhard Schuberth,
-Anne Sieminski,
-Leif Strand,
-Carl Tape,
-Jeroen Tromp,
-Zhinan Xie,
-Hejun Zhu.\\
-%\toall{add full list of developers here and update the AUTHORS file accordingly}
-
-The cover graphic of the manual was created
-by Santiago Lombeyda from Caltech's Center for Advanced Computing Research (CACR) \urlwithparentheses{http://www.cacr.caltech.edu}. \\
-
-\newpage{}
-
-\tableofcontents{}
-
-\chapter{Introduction}
-
-The software package SPECFEM3D simulates seismic
-wave propagation at the local or regional scale
-based upon the spectral-element method (SEM).
-The SEM is a continuous Galerkin technique \citep{TrKoLi08,PeKoLuMaLeCaLeMaLiBlNiBaTr11},
-which can easily be made discontinous \citep{BeMaPa94,Ch00,KoWoHu02,ChCaVi03,LaWaBe05,Kop06,WiStBuGh10,AcKo11};
-it is then close to a particular case of the discontinuous Galerkin technique \citep{ReHi73,Arn82,FaRi99,HuHuRa99,CoKaSh00,GiHeWa02,RiWh03,MoRi05,GrScSc06,AiMoMu06,BeLaPi06,DuKa06,DeSeWh08,PuAmKa09,WiStBuGh10,DeSe10,EtChViGl10}, with optimized efficiency because of its tensorized basis functions \citep{WiStBuGh10,AcKo11}.
-In particular, it can accurately handle very distorted mesh elements \citep{OlSe11}.\\
-
-It has very good accuracy and convergence properties \citep{MaPa89,SePr94,DeFiMu02,Coh02,DeSe07,SeOl08}.
-The spectral element approach admits spectral rates of convergence and allows exploiting $hp$-convergence schemes.
-It is also very well suited to parallel implementation on very large supercomputers \citep{KoTsChTr03,TsKoChTr03,KoLaMi08a,CaKoLaTiMiLeSnTr08,KoViCh10} as well as on clusters of GPU accelerating graphics cards \citep{KoMiEr09,KoErGoMi10,Kom11}.
-Tensor products inside each element can be optimized to reach very high efficiency \citep{DeFiMu02}, and mesh point and element numbering can be optimized to reduce processor cache misses and improve cache reuse \citep{KoLaMi08a}.
-The SEM can also handle triangular (in 2D) or tetrahedral (in 3D) elements \citep{WinBoyd96,TaWi00,KoMaTrTaWi01,Coh02,MeViSa06} as well as mixed meshes, although with increased cost and reduced accuracy in these elements, as in the discontinuous Galerkin method.\\
-
-Note that in many geological models in the context of seismic wave propagation studies
-(except for instance for fault dynamic rupture studies, in which very high frequencies or supershear rupture need to be modeled near the fault, see e.g.
-\cite{BeGlCrViPi07,BeGlCrVi09,PuAmKa09,TaCrEtViBeSa10})
-a continuous formulation is sufficient because material property contrasts are not drastic and thus
-conforming mesh doubling bricks can efficiently handle mesh size variations \citep{KoTr02a,KoLiTrSuStSh04,LeChLiKoHuTr08,LeChKoHuTr09,LeKoHuTr09}.\\
-
-For a detailed introduction to the SEM as applied to regional
-seismic wave propagation, please consult \citet{PeKoLuMaLeCaLeMaLiBlNiBaTr11,TrKoLi08,KoVi98,KoTr99,ChKoViCaVaFe07} and
-in particular \citet{LeKoHuTr09,LeChKoHuTr09,LeChLiKoHuTr08,GoAmTaCaSmSaMaKo09,WiKoScTr04,KoLiTrSuStSh04}.
-A detailed theoretical analysis of the dispersion
-and stability properties of the SEM is available in \citet{Coh02}, \citet{DeSe07}, \citet{SeOl07} and \citet{SeOl08}.\\
-
-Effects due to lateral variations in compressional-wave speed, shear-wave
-speed, density, a 3D crustal model, topography and bathymetry are
-included. The package can accommodate
-full 21-parameter anisotropy (see~\citet{ChTr07}) as well as lateral
-variations in attenuation \citep{SaKoTr10}. Adjoint capabilities and finite-frequency
-kernel simulations are included \citep{TrKoLi08,PeKoLuMaLeCaLeMaLiBlNiBaTr11,LiTr06,FiIgBuKe09,ViOp09}.\\
-
-The SEM was originally developed in computational fluid dynamics \citep{Pat84,MaPa89}
-and has been successfully adapted to address problems in seismic wave propagation.
-Early seismic wave propagation applications of the SEM, utilizing Legendre basis functions and a
-perfectly diagonal mass matrix, include \cite{CoJoTo93}, \cite{Kom97},
-\cite{FaMaPaQu97}, \cite{CaGa97}, \cite{KoVi98} and \cite{KoTr99},
-whereas applications involving Chebyshev basis functions and a nondiagonal mass matrix
-include \cite{SePr94}, \cite{PrCaSe94} and \cite{SePrPr95}.\\
-
-All SPECFEM3D software is written in Fortran90 with full portability
-in mind, and conforms strictly to the Fortran95 standard. It uses
-no obsolete or obsolescent features of Fortran77. The package uses
-parallel programming based upon the Message Passing Interface (MPI)
-\citep{GrLuSk94,Pac97}.\\
-
-SPECFEM3D won the Gordon Bell award for best performance at the SuperComputing~2003
-conference in Phoenix, Arizona (USA) (see \cite{KoTsChTr03}
-and \url{www.sc-conference.org/sc2003/nr_finalaward.html}).
-It was a finalist again in 2008 for a run at 0.16 petaflops (sustained) on 149,784 processors of the `Jaguar' Cray XT5 system at Oak Ridge National Laboratories (USA) \citep{CaKoLaTiMiLeSnTr08}.
-It also won the BULL Joseph Fourier supercomputing award in 2010.\\
-
-This new release of the code (version 2.1) includes support for GPU graphics card acceleration \citep{KoMiEr09,KoGoErMi10,KoErGoMi10,MiKo10,Kom11}.\\
-
-The next release of the code will include support Convolution or Auxiliary Differential Equation Perfectly Matched absorbing Layers
-(C-PML or ADE-PML) \citep{KoMa07,MaKoEz08,MaKoGe08,MaKo09,MaKoGeBr10}.
-It will also use the PT-SCOTCH parallel library for mesh partitioning.\\
-
-SPECFEM3D includes coupled fluid-solid domains and
-adjoint capabilities, which enables one to address seismological inverse problems, but for linear rheologies only so far.
-To accommodate visco-plastic or non-linear rheologies, the reader can refer to
-the \href{http://geoelse.stru.polimi.it/}{GeoELSE} software package \citep{CaGa97,StPaIg09,ChMoTsBaKrKaStKr10}.
-
-\section{Citation}
-
-If you use SPECFEM3D for your own research, please cite at least one
-of the following articles:
-
-\begin{description}
-
-\item [Numerical simulations in general] ~\\
-Forward and adjoint simulations are described in detail in \cite{TrKoLi08,PeKoLuMaLeCaLeMaLiBlNiBaTr11,VaCaSaKoVi99,KoMiEr09,KoErGoMi10,KoGoErMi10,ChKoViCaVaFe07,MaKoDi09,KoViCh10,CaKoLaTiMiLeSnTr08,TrKoHjLiZhPeBoMcFrTrHu10,KoRiTr02,KoTr02a,KoTr02b,KoTr99} or \cite{KoVi98}.
-Additional aspects dealing with adjoint simulations are described in
-\cite{TrTaLi05,LiTr06,TrKoLi08,LiTr08,TrKoHjLiZhPeBoMcFrTrHu10,PeKoLuMaLeCaLeMaLiBlNiBaTr11}.
-Domain decomposition is explained in detail in \cite{MaKoBlLe08}, and excellent scaling up to 150,000 processor cores is shown for instance in
-\cite{CaKoLaTiMiLeSnTr08,KoLaMi08a,MaKoBlLe08,KoErGoMi10,KoGoErMi10,Kom11},
-
-\item [GPU computing] ~\\
-Computing on GPU graphics cards for acoustic or seismic wave propagation applications
-is described in detail in \cite{KoMiEr09,KoGoErMi10,KoErGoMi10,MiKo10,Kom11}.
-
-\end{description}
-
-\noindent
-If you use this new version 2.1, which has non blocking MPI for much better performance for medium or large runs, please cite at least one of these six articles,
-in which results of non blocking MPI runs are presented: \cite{PeKoLuMaLeCaLeMaLiBlNiBaTr11,KoGoErMi10,KoErGoMi10,KoViCh10,Kom11,CaKoLaTiMiLeSnTr08,MaKoBlLe08}.\\
-
-\noindent
-If you work on geophysical applications, you may be interested in citing some of these application articles as well, among others:
-
-\begin{description}
-
-\item [Southern California simulations] ~\\
-\cite{KoLiTrSuStSh04,KrJiKoTr06a,KrJiKoTr06b}.
-
-If you use the 3D southern California model, please cite
-\citet{SuSh03} (Los Angeles model),
-\citet{lovelyetal06} (Salton Trough), and
-\citet{hauksson2000} (southern California).
-The Moho map was determined by
-\citet{zhu&kanamori2000}.
-The 1D SoCal model was developed by \citet{DrHe90}.
-
-\item [Anisotropy] ~\\
-\cite{ChTr07,JiTsKoTr05,ChFaKo04,FaChKo04,RiRiKoTrHe02,TrKo00}.
-
-\item [Attenuation] ~\\
-\cite{SaKoTr10,KoTr02a,KoTr99}.
-
-\item [Topography] ~\\
-\cite{LeKoHuTr09,LeChKoHuTr09,LeChLiKoHuTr08,GoAmTaCaSmSaMaKo09,WiKoScTr04}.
-
-\end{description}
-
-The corresponding Bib\TeX{} entries may be found
-in file \texttt{doc/USER\_MANUAL/bibliography.bib}.
-
-\section{Support}
-
-This material is based upon work supported by the USA National Science
-Foundation under Grants No. EAR-0406751 and EAR-0711177, by the French
-CNRS, French INRIA Sud-Ouest MAGIQUE-3D, French ANR NUMASIS under
-Grant No. ANR-05-CIGC-002, and European FP6 Marie Curie International
-Reintegration Grant No. MIRG-CT-2005-017461.
-Any opinions, findings, and conclusions or recommendations expressed in this material are
-those of the authors and do not necessarily reflect the views of the
-USA National Science Foundation, CNRS, INRIA, ANR or the European
-Marie Curie program.
-
-\chapter{\label{cha:Getting-Started}Getting Started}
-
-The SPECFEM3D software package comes in a gzipped tar ball. In the
-directory in which you want to install the package, type
-\begin{lyxcode}
-{\small tar~-zxvf~SPECFEM3D\_V2.1.0.tar.gz}{\small \par}
-\end{lyxcode}
-The directory \texttt{\small SPECFEM3D/} will then contain
-the source code.\\
-
-We recommend that you add {\texttt{ulimit -S -s unlimited}} to your {\texttt{.bash\_profile}} file and/or {\texttt{limit stacksize unlimited }} to your {\texttt{.cshrc}} file to suppress any potential limit to the size of the Unix stack.\\
-
-Then, to configure the software for your system, run the
-\texttt{configure} shell script. This script will attempt to guess
-the appropriate configuration values for your system. However, at
-a minimum, it is recommended that you explicitly specify the appropriate
-command names for your Fortran90 compiler and MPI package:
-
-\begin{lyxcode}
-./configure~FC=ifort~MPIFC=mpif90
-\end{lyxcode}
-
-Before running the \texttt{configure} script, you should probably edit file \texttt{flags.guess} to make sure that it contains the best compiler options for your system.
-Known issues or things to check are:
-
-\begin{description}
-\item [{\texttt{Intel ifort compiler}}] See if you need to add \texttt{-assume byterecl} for your machine.
-\item [{\texttt{IBM compiler}}] See if you need to add \texttt{-qsave} or \texttt{-qnosave} for your machine.
-\item [{\texttt{Mac OS}}] You will probably need to install \texttt{XCODE}.
-In addition, the \texttt{clock\_gettime} routine, which is used by the \texttt{SCOTCH} library that we use, does not exist in Mac OS.
-You will need to replace it with \texttt{clock\_get\_time} if you want to use \texttt{SCOTCH}.
-\end{description}
-
-When compiling on an IBM machine with the \texttt{xlf} and \texttt{xlc} compilers, we suggest running the \texttt{configure} script
-with the following options:\\
-
-\noindent\texttt{./configure FC=xlf90\_r MPIFC=mpif90 CC=xlc\_r CFLAGS="-O3 -q64" FCFLAGS="-O3 -q64" -with-scotch-dir=...}.\\
-
-On SGI systems, \texttt{flags.guess} automatically informs \texttt{configure}
-to insert `\texttt{`TRAP\_FPE=OFF}'' into the generated \texttt{Makefile}
-in order to turn underflow trapping off.\\
-
-If you run very large meshes on a relatively small number
-of processors, the memory size needed on each processor might become
-greater than 2 gigabytes, which is the upper limit for 32-bit addressing.
-In this case, on some compilers you may need to add \texttt{``-mcmodel=medium}'' or \texttt{``-mcmodel=medium -shared-intel}''
-to the configure options of CFLAGS, FCFLAGS and LDFLAGS otherwise the compiler will display an error
-message (for instance \texttt{``relocation truncated to fit: R\_X86\_64\_PC32 against .bss''} or something similar);
-on an IBM machine with the \texttt{xlf} and \texttt{xlc} compilers, using \texttt{-q64} is usually sufficient.\\
-
-The SPECFEM3D software package relies on the SCOTCH library to partition meshes created with CUBIT.
-Note that we use CUBIT to create meshes of hexahedra, but other packages can be used as well, for instance GiD from \url{http://gid.cimne.upc.es} or Gmsh from \url{http://geuz.org/gmsh} \citep{GeRe09}.
-Even mesh creation packages that generate tetrahedra, for instance TetGen from \url{http://tetgen.berlios.de}, can be used because each tetrahedron can then easily be decomposed into four hexahedra as shown in the picture of the TetGen logo at \url{http://tetgen.berlios.de/figs/Delaunay-Voronoi-3D.gif}; while this approach does not generate hexahedra of optimal quality, it can ease mesh creation in some situations and it has been shown that the spectral-element method can very accurately handle distorted mesh elements \citep{OlSe11}.
-
-The SCOTCH library \citep{PeRo96}
-provides efficent static mapping, graph and mesh partitioning routines. SCOTCH is a free software package developed by
-Fran\c {c}ois Pellegrini et al.
-from LaBRI and INRIA in Bordeaux, France, downloadable from the web page \url{https://gforge.inria.fr/projects/scotch/}.
-It is more recent than METIS, actively maintained and performs better in many cases.
-A recent version of its source code is provided in directory \texttt{src/decompose\_mesh\_SCOTCH/scotch\_5.1.10b}.
-In case no SCOTCH libraries can be found on the system, the configuration will bundle this version for compilation.
-The path to an existing SCOTCH installation can to be set explicitly with the option \texttt{-{}-with-scotch-dir}. Just as an example:
-\begin{lyxcode}
-./configure~FC=ifort~MPIFC=mpif90~-{}-with-scotch-dir=/opt/scotch
-\end{lyxcode}
-
-If you use the Intel ifort compiler to compile the code, we recommend that you use the Intel icc C compiler to compile Scotch, i.e., use:
-\begin{lyxcode}
-./configure~CC=icc~FC=ifort~MPIFC=mpif90
-\end{lyxcode}
-
-To compile a serial version of the code for small meshes that fits
-on one compute node and can therefore be run serially, run \texttt{configure}
-with the \texttt{-{}-without-mpi} option to suppress all calls to
-MPI.
-
-A summary of the most important configuration variables follows.
-
-\begin{description}
-\item [{\texttt{F90}}] Path to the Fortran90 compiler.
-\item [{\texttt{MPIF90}}] Path to MPI Fortran90.
-\item [{\texttt{MPI\_FLAGS}}] Some systems require this flag to link to
-MPI libraries.
-\item [{\texttt{FLAGS\_CHECK}}] Compiler flag for non-critical subroutines.
-\item [{\texttt{FLAGS\_NO\_CHECK}}] Compiler flag for creating fast, production-run
-code for critical subroutines.
-\end{description}
-The configuration script automatically creates for each executable a corresponding
-\texttt{Makefile} in the \texttt{src/} subdirectoy.
-The \texttt{Makefile} contains a number of suggested entries for various
-compilers, e.g., Portland, Intel, Absoft, NAG, and Lahey. The software
-has run on a wide variety of compute platforms, e.g., various PC clusters
-and machines from Sun, SGI, IBM, Compaq, and NEC. Select the compiler
-you wish to use on your system and choose the related optimization
-flags. Note that the default flags in the \texttt{Makefile} are undoubtedly
-not optimal for your system, so we encourage you to experiment with
-these flags and to solicit advice from your systems administrator.
-Selecting the right compiler and optimization flags can make a tremendous
-difference in terms of performance. We welcome feedback on your experience
-with various compilers and flags.
-
-Now that you have set the compiler information, you need to select
-a number of flags in the \texttt{constants.h} file depending on your
-system:
-
-\begin{description}
-\item [{\texttt{LOCAL\_PATH\_IS\_ALSO\_GLOBAL}}] Set to \texttt{.false.}
-on most cluster applications. For reasons of speed, the (parallel)
-distributed database generator typically writes a (parallel) database for the solver on the
-local disks of the compute nodes. Some systems have no local disks,
-e.g., BlueGene or the Earth Simulator, and other systems have a fast
-parallel file system, in which case this flag should be set to \texttt{.true.}.
-Note that this flag is not used by the database generator or the solver; it is
-only used for some of the post-processing.
-\end{description}
-The package can run either
-in single or in double precision mode. The default is single precision
-because for almost all calculations performed using the spectral-element method
-using single precision is sufficient and gives the same results (i.e. the same seismograms);
-and the single precision code is faster and requires exactly half as much memory.
-Select your preference by selecting the appropriate setting in the \texttt{constants.h} file:
-
-\begin{description}
-\item [{\texttt{CUSTOM\_REAL}}] Set to \texttt{SIZE\_REAL} for single precision
-and \texttt{SIZE\_DOUBLE} for double precision.
-\end{description}
-In the \texttt{precision.h} file:
-
-\begin{description}
-\item [{\texttt{CUSTOM\_MPI\_TYPE}}] Set to \texttt{MPI\_REAL} for single
-precision and \texttt{MPI\_DOUBLE\_PRECISION} for double precision.
-\end{description}
-On many current processors (e.g., Intel, AMD, IBM Power), single precision calculations
-are significantly faster; the difference can typically be 10\%
-to 25\%. It is therefore better to use single precision.
-What you can do once for the physical problem you want to study is run the same calculation in single precision
-and in double precision on your system and compare the seismograms.
-If they are identical (and in most cases they will), you can select single precision for your future runs.\\
-
-\section{Adding OpenMP support in addition to MPI}
-
-OpenMP support can be enabled in addition to MPI. However, in many cases performance will not improve
-because our pure MPI implementation is already heavily optimized and thus the resulting code will in fact
-be slightly slower. A possible exception could be IBM BlueGene-type architectures.
-
-To enable OpenMP, uncomment the OpenMP compiler option in two lines in file \texttt{src/specfem3D/Makefile.in}
-(before running \texttt{configure}) and also uncomment the \texttt{\#define USE\_OPENMP} statement in file\\
-\texttt{src/specfem3D/specfem3D.F90}.
-
-The DO-loop using OpenMP threads has a SCHEDULE property. The OMP\_SCHEDULE
-environment variable can set the scheduling policy of that DO-loop.
-Tests performed by Marcin Zielinski at SARA (The Netherlands) showed that often
-the best scheduling policy is DYNAMIC with the size of the chunk equal to the number of
-OpenMP threads, but most preferably being twice as the number of
-OpenMP threads (thus chunk size = 8 for 4 OpenMP threads etc).
-If OMP\_SCHEDULE is not set or is empty, the DO-loop will assume generic
-scheduling policy, which will slow down the job quite a bit.
-
-\section{Visualizing the subroutine calling tree of the source code}
-
-Packages such as \texttt{doxywizard} can be used to visualize the subroutine calling tree of the source code.
-\texttt{Doxywizard} is a GUI front-end for configuring and running \texttt{doxygen}.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\chapter{\label{cha:Mesh-Generation}Mesh Generation}
-
-The first step in running a spectral-element simulation consists of constructing a high-quality mesh for the region under consideration.
-We provide two possibilities to do so: (1) relying on the external, hexahedral mesher CUBIT, or (2) using the provided, internal mesher \texttt{xmeshfem3D}.
-In the following, we explain these two approaches.
-
-\section{\label{cha:Running-the-Mesher-CUBIT}Meshing with \texttt{CUBIT}}
-
-CUBIT is a meshing tool suite for the creation of finite-element meshes for arbitrarily shaped models.
-It has been developed and maintained at Sandia National Laboratories and can be purchased for a small academic institutional fee at \url{http://cubit.sandia.gov}.
-Our experience showed that using CUBIT greatly facilitates and speeds up the generation and preparation of hexahedral, conforming meshes for a variety of geophysical models with increasing complexity.
+\chapter{Graphics}
+
+\section{Meshes}\label{sec:Mesh-graphics}
+
+In case you used the internal mesher \texttt{xmeshfem3D} to create
+and partition your mesh, you can output mesh files in ABAQUS (.INP)
+and DX (.dx) format to visualize them. For this, you must set either
+the flag \texttt{CREATE\_DX\_FILES} or \texttt{CREATE\_ABAQUS\_FILES}
+to \texttt{.true.} in the mesher's parameter file \texttt{Mesh\_Par\_file}
+prior to running the mesher (see Chapter~\ref{cha:Running-the-Mesher-Meshfem3D}
+for details). You can then use AVS \urlwithparentheses{www.avs.com}
+or OpenDX \urlwithparentheses{www.opendx.org} to visualize the mesh
+and MPI partition (slices).
 
@@
 \noindent \begin{centering}
-\includegraphics[width=.6\textwidth]{figures/mount-cubit.jpg}
+\includegraphics[width=0.49\textwidth]{figures/vtk_mesh_vp.jpg} 
+\includegraphics[width=0.49\textwidth]{figures/vtk_mesh_vs.jpg}
 \par\end{centering}
 
-\caption{Example of the graphical user interface of CUBIT. The hexahedral mesh shown in the main display consists of a hexahedral discretization of a single volume with topography.}
-
-\label{fig:mount.cubit}
-\end{figure}
-
-The basic steps in creating a load-balanced, partitioned mesh with CUBIT are:
-\begin{description}
-\item [1.] setting up a hexahedral mesh with CUBIT,
-\item [2.] exporting the CUBIT mesh into a SPECFEM3D file format and
-\item [3.] partitioning the SPECFEM3D mesh files for a chosen number of cores.
-\end{description}
-
-Examples are provided in the SPECFEM3D package in the subdirectory \texttt{examples/}.
-We strongly encourage you to contribute your own example to this package by contacting the CIG Computational Seismology Mailing List  \urlwithparentheses{cig-seismo@geodynamics.org}.
-
-\subsection{Creating the Mesh with CUBIT}
-
-For the installation and handling of the CUBIT meshing tool suite, please refer to the CUBIT user manual and documentation.
-In order to give you a basic understanding of how to use CUBIT for our purposes, examples are provided in the SPECFEM3D package in the subdirectory \texttt{examples/}:
-\begin{description}
-\item [{\texttt{homogeneous\_halfspace}}] Creates a single block model and assigns elastic material parameters.
-\item [{\texttt{layered\_halfspace}}]  Combines two different, elastic material volumes and creates a refinement layer between the two.
-This example can be compared for validation against the solutions provided in subdirectory ~\\
-\texttt{VALIDATION\_3D\_SEM\_SIMPLER\_LAYER\_SOURCE\_DEPTH/}.
-\item [{\texttt{waterlayered\_halfspace}}] Combines an acoustic and elastic material volume as in a schematic marine survey example.
-\item [{\texttt{tomographic\_model}}] Creates a single block model whose material properties will have to be read in from a tomographic model file during the databases creation by \texttt{xgenerate\_databases}.
-\end{description}
-
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=.45\textwidth]{figures/example-homogeneous.jpg}
-\includegraphics[width=.45\textwidth]{figures/example-2layers.jpg} \\
-\includegraphics[width=.45\textwidth]{figures/example-water.jpg}
-\includegraphics[width=.45\textwidth]{figures/example-tomo.jpg}
-\par\end{centering}
-
-\caption{Screenshots of the CUBIT examples provided in subdirectory  \texttt{examples/}: homogeneous halfspace (top-left), layered halfspace (top-right), water layered halfspace (bottom-left) and tomographic model (bottom-right).}
-
-\label{fig:examples.cubit}
-\end{figure}
-
-In each example subdirectory you will find a \texttt{README} file, which explains in a step-by-step tutorial the workflow for the example.
-Please feel free to contribute your own example to this package by contacting the CIG Computational Seismology Mailing List  \urlwithparentheses{cig-seismo@geodynamics.org}.
-
-\subsection{Exporting the Mesh with \texttt{cubit2specfem3d.py} }
-\label{subsec:Exporting-the-Mesh}
-
-Once the geometric model volumes in CUBIT are meshed, you prepare the model for exportation with the definition of material blocks and boundary surfaces.
-Thus, prior to exporting the mesh, you need to define blocks specifying the materials and absorbing boundaries in CUBIT.
-This process could be done automatically using the script \texttt{boundary\_definition.py} if the mesh meets some conditions or manually, following the block convention:
-
-\begin{description}
-\item [material\_name] Each material should have a specific block defined by a unique name.
-The name convention of the material is to start with either 'elastic' or 'acoustic'.
-It must be then followed by a unique identifier, e.g.
-'elastic 1', 'elastic 2', etc.
-The additional attributes to the block define the material description.
-
-For an elastic material:
-\begin{description}
-\item [material\_id] An integer value which is unique for this material.
-\item [Vp] P-wave speed of the material (given in m/s).
-\item [Vs] S-wave speed of the material (given in m/s).
-\item [rho] density of the material (given in kg/m$^3$).
-\item [Q] quality factor to use in case of a simulation with attenuation turned on.
-It should be between 1 and 9000.
-In case no attenuation information is available, it can be set to zero.
-Please note that your Vp- and Vs-speeds are given for a reference frequency.
-To change this reference frequency, you change the value of
-\texttt{ATTENUATION\_f0\_REFERENCE} in the main constants file \texttt{constants.h} found in subdirectory \texttt{src/shared/}.
-\item [anisotropic\_flag] Flag describing the anisotropic model to use in case an anisotropic simulation should be conducted.
-See the file \texttt{model\_aniso.f90} in subdirectory \texttt{src/generate\_databases/} for an implementation of the anisotropic models.
-In case no anisotropy is available, it can be set to zero.
-\end{description}
-Note that this material block has to be defined using all the volumes which belong to this elastic material.
-For volumes belonging to another, different material, you will need to define a new material block.
-
-For an acoustic material:
-\begin{description}
-\item [material\_id] An integer value which is unique for this material.
-\item [Vp] P-wave speed of the material (given in m/s).
-\item [0] S-wave speed of the material is ignored.
-\item [rho] density of the material (given in kg/m$^3$).
-\end{description}
-
-\item [face\_topo] Block definition for the surface which defines the free surface (which can have topography).
-The name of this block must be 'face\_topo', the block has to be defined using all the surfaces which constitute the complete free surface of the model.
-
-\item [face\_abs\_xmin] Block definition for the faces on the absorbing boundaries, one block for each surface with x=Xmin.
-\item [face\_abs\_xmax] Block definition for the faces on the absorbing boundaries, one block for each surface with x=Xmax.
-\item [face\_abs\_ymin] Block definition for the faces on the absorbing boundaries, one block for each surface with y=Ymin.
-\item [face\_abs\_ymax] Block definition for the faces on the absorbing boundaries, one block for each surface with y=Ymax.
-\item [face\_abs\_bottom] Block definition for the faces on the absorbing boundaries, one block for each surface with z=bottom.
-\end{description}
-
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=.45\textwidth]{figures/mount-surface.jpg}
-\includegraphics[width=.45\textwidth]{figures/mount-abs.jpg}
-\par\end{centering}
-
-\caption{Example of the block definitions for the free surface 'face\_topo' (left) and the absorbing boundaries, defined in a single block 'face\_abs\_xmin' (right) in CUBIT.}
-
-\label{fig:mount.abs}
-\end{figure}
-
-Optionally, instead of specifying for each surface at the boundaries a single block like mentioned above, you can also specify a single block for all boundary surfaces and name it as one of the absorbing blocks above, e.g.
-'face\_abs\_xmin'.
-
-After the block definitions are done, you export the mesh using the script \texttt{cubit2specfem3d.py} provided in each of the example directories.
-If the export was successful, you should find the following files in a subdirectory \texttt{MESH/}:
-
-\begin{description}
-\item [nummaterial\_velocity\_file] Defines the material properties. For fully defined materials, the formats is:
-\begin{lyxcode}
-\texttt{domain\_ID} \texttt{material\_ID}  \texttt{rho}  \texttt{vp}  \texttt{vs}  \texttt{Q}  \texttt{anisotropy\_flag}
-\end{lyxcode}
-where
-\texttt{domain\_ID} is 1 for acoustic or 2 for elastic materials, \texttt{material\_ID} a unique identifier,  \texttt{rho} the density in $kg \, m^{-3}$,  \texttt{vp} the P-wave speed in $m \, s^{-1}$, \texttt{vs}  the S-wave speed in $m \, s^{-1}$, \texttt{Q} the quality factor and \texttt{anisotropy\_flag} an identifier for anisotropic models.
-
-%%magnoni
-For tomographic velocity models, please read Chapter \ref{cha:-Changing-the} and Section \ref{sec:Using-tomographic} `Using external tomographic Earth models' for further details.
-%
-%For a tomographic model, the material definition format is:
-%\begin{lyxcode}
-%\texttt{domain\_ID} \texttt{material\_ID}  \texttt{tomography}  \texttt{name}
-%\end{lyxcode}
-%where
-%\texttt{domain\_ID} is 1 for acoustic or 2 for elastic materials, \texttt{material\_ID} a negative, unique identifier, \texttt{tomography} keyword for tomographic material definition and \texttt{name} the name of the tomography file.
-The name is not used so far, rather change the filename defined in the file \texttt{model\_tomography.f90} located in the
-%\texttt{src/generate\_databases/} directory.
-%
-\item [materials\_file] Contains the material associations for each element.
-\item [nodes\_coords\_file] Contains the point locations in Cartesian coordinates of the mesh element corners.
-\item [mesh\_file] Contains the mesh element connectivity.
-\item [free\_surface\_file] Contains the free surface connectivity.
-\item [absorbing\_surface\_file\_xmax] Contains the surface connectivity of the absorbing boundary surface at the Xmax.
-\item [absorbing\_surface\_file\_xmin] Contains the surface connectivity of the absorbing boundary surface at the Xmin.
-\item [absorbing\_surface\_file\_ymax] Contains the surface connectivity of the absorbing boundary surface at the Ymax.
-\item [absorbing\_surface\_file\_ymin] Contains the surface connectivity of the absorbing boundary surface at the Ymin.
-\item [absorbing\_surface\_file\_bottom] Contains the surface connectivity of the absorbing boundary surface at the bottom.
-\end{description}
-
-These mesh files are needed as input files for the partitioner  \texttt{xdecompose\_mesh\_SCOTCH}
-to load-balance the mesh. Please see the next section for further details.
-
-\subsubsection*{Checking the mesh quality}
-The quality of the mesh may be inspected more precisely based upon the serial code
-in the file  \texttt{check\_mesh\_quality\_}~\\
-\texttt{CUBIT\_Abaqus.f90} located in the directory \texttt{src/check\_mesh\_quality\_CUBIT\_Abaqus/}.
-Running this code is optional because no information needed by the
-solver is generated.
-
-Prior to running and compiling this code, you have to export your mesh in CUBIT to an ABAQUS (.inp) format.
-For example, export mesh block IDs belonging to volumes in order to check the quality of the hexahedral elements.
-You also have to determine a number of parameters of your mesh, such as the number of nodes and number of elements
-and modify the header of the \texttt{check\_mesh\_quality\_CUBIT\_Abaqus.f90} source file. Then,
-in the main directory, type
-\begin{lyxcode}
-{\small make xcheck\_mesh\_quality\_CUBIT\_Abaqus}{\small \par}
-\end{lyxcode}
-and use
-\begin{lyxcode}
-{\small ./bin/xcheck\_mesh\_quality\_CUBIT\_Abaqus}{\small \par}
-\end{lyxcode}
-to generate an
-%not supported: AVS output file (\texttt{\small AVS\_meshquality.inp} in AVS UCD format) or
-OpenDX output file (\texttt{\small DX\_mesh\_quality.dx})
-that can be used to investigate mesh quality, e.g. skewness of elements,
-and a Gnuplot histogram (\texttt{\small mesh\_quality\_histogram.txt})
-that can be plotted with gnuplot (type
-`\texttt{\small gnuplot plot\_mesh\_quality\_histogram.gnu}'). The
-histogram is also printed to the screen.
-Analyze that skewness histogram of mesh elements to make sure no element has a skewness above approximately 0.75, otherwise the mesh is of poor quality (and if even a single element has a skewness value above 0.80, then you must definitely improve the mesh).
-If you want to start designing
-your own meshes, this tool is useful for viewing your creations. You
-are striving for meshes with elements with `cube-like' dimensions,
-e.g., the mesh should contain no very elongated or skewed elements.
-
-\subsection{Partitioning the Mesh with \texttt{xdecompose\_mesh\_SCOTCH}}
-
-The SPECFEM3D software package performs large scale simulations in a parallel 'Single Process Multiple Data' way.
-The spectral-element mesh created with CUBIT needs to be distributed on the processors. This partitioning is executed once and for all prior to
-the execution of the solver so it is referred to as a static mapping.
-
-An efficient partitioning is important because it leverages the overall running time of the application.
-It amounts to balance the number of elements in each slice while minimizing the communication costs
-resulting from the placement of adjacent elements on different processors. \texttt{decompose\_mesh\_SCOTCH} depends on the SCOTCH library \citep{PeRo96},
-which provides efficent static mapping, graph and mesh partitioning routines. SCOTCH is a free software package developed by
-Fran\c {c}ois Pellegrini et al.
-from LaBRI and INRIA in Bordeaux, France, downloadable from the web page \url{https://gforge.inria.fr/projects/scotch/}.
-It is more recent than METIS, actively maintained and performs better in many cases.
-A recent version of its source code is provided in directory \texttt{src/decompose\_mesh\_SCOTCH/scotch\_5.1.10b}.
-\\
-
-In most cases, the configuration with \texttt{./configure~FC=ifort} should be sufficient.
-During the configuration process, the script tries to find existing SCOTCH installations.
-In case your system has no pre-existing SCOTCH installation,
-we provide the source code of SCOTCH, which is released open source under the French CeCILL-C version 1 license,
-in directory
-\texttt{src/decompose\_mesh\_SCOTCH/scotch\_5.1.10b}.
-This version gets bundled with the compilation of the SPECFEM3D package if no libraries could have been found.
-If this automatic compilation of the SCOTCH libraries fails,
-please refer to file INSTALL.txt in that directory to see further details how to compile it on your system.
-In case you want to use a pre-existing installation, make sure you have correctly specified the path
-of the SCOTCH library when using the option \texttt{-{}-with-scotch-dir} with the \texttt{./configure} script.
-In the future you should be able to find more recent versions at  \url{http://www.labri.fr/perso/pelegrin/scotch/scotch\_en.html} .
-\\
-
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=.45\textwidth]{figures/mount-partitions.jpg}
-\includegraphics[width=.45\textwidth]{figures/mount-partitions2.jpg}
-\par\end{centering}
-
-\caption{Example of a mesh partitioning onto four cores.
-Each single core partition is colored differently.
-The executable \texttt{xdecompose\_mesh\_SCOTCH} can equally distribute the mesh on any arbitrary number of cores.
-Domain decomposition is explained in detail in \cite{MaKoBlLe08}, and excellent scaling up to 150,000 processor cores in shown for instance in
-\cite{CaKoLaTiMiLeSnTr08,KoLaMi08a,MaKoBlLe08,KoErGoMi10,KoGoErMi10,Kom11}.}
-
-\label{fig:mount.partitions}
-\end{figure}
-
-When you are ready to compile,
-in the main directory type
-\begin{lyxcode}
-{\small make xdecompose\_mesh\_SCOTCH}{\small \par}
-\end{lyxcode}
-If all paths and flags have been set correctly, the executable \texttt{bin/xdecompose\_mesh\_SCOTCH} should be produced.
-
-The partitioning is done in serial for now (in the next release we will provide a parallel version of that code),
-the synopsis is:
-\begin{lyxcode}
-./bin/xdecompose\_mesh\_SCOTCH  nparts  input\_directory output\_directory
-\end{lyxcode}
-where
-\begin{itemize}
-     \item \texttt{nparts} is the number of partitions, i.e., the number of cores for the parallel simulations,
-     \item \texttt{input\_directory} is the directory which holds all the files generated by the Python script \texttt{cubit2specfem3d.py} explained in the previous Section~\ref{subsec:Exporting-the-Mesh}, e.g.
-\texttt{MESH/}, and
-     \item \texttt{output\_directory} is the directory for the output of this partitioner which stores ACII-format files named like \texttt{proc******\_Database} for each partition.
-These files will be needed for creating the distributed databases, and have to reside in the directory  \texttt{LOCAL\_PATH} specified in the main \texttt{Par\_file}, e.g.
-in directory \texttt{in\_out\_files/DATABASES\_MPI}.
-Please see Chapter~\ref{cha:Creating-Distributed-Databases} for further details.
-\end{itemize}
-Note that all the files generated by the Python script  \texttt{cubit2specfem3d.py} must be placed in the \texttt{input\_directory} folder before running the program.
-
-\section{\label{cha:Running-the-Mesher-Meshfem3D}Meshing with \texttt{xmeshfem3D}}
-
-In case you successfully ran the configuration script, you are also ready to compile the internal mesher.
-This is an alternative to CUBIT
-for the mesh generation of relatively simple geological models. The mesher is no longer
-dedicated to Southern California and more flexiblity is provided in this version of the package.
-
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[scale=0.5]{figures/socal_map_mpi}
-\par\end{centering}
-
-\caption{\label{fig:For-parallel-computing}For parallel computing purposes,
-the model block is subdivided in $\nprocxi\times\nproceta$ slices
-of elements. In this example we use $5^{2}=25$ processors. }
-
-\end{figure}
-
-In the main directory, type
-\begin{lyxcode}
-{\small make xmeshfem3D}{\small \par}
-\end{lyxcode}
-If all paths and flags
-have been set correctly, the mesher should now compile and produce
-the executable \texttt{bin/xmeshfem3D}. Please note that
-\texttt{xmeshfem3D} must be called directly from the \texttt{bin/} directory,
-as most of the binaries of the package.
-
-Input for the mesh generation program is provided through the parameter
-file \texttt{Mesh\_Par\_file}, which resides in the subdirectory \texttt{in\_data\_files/meshfem3D\_files/}.
-Before running the mesher, a number of parameters need to be set in
-the \texttt{Mesh\_Par\_file}. This requires a basic understanding of how
-the SEM is implemented, and we encourage you to read \citet{KoVi98,KoTr99}
-and \citet{KoLiTrSuStSh04}.
-
-The mesher and the solver use UTM coordinates internally, therefore
-you need to define the zone number for the UTM projection (e.g., zone
-11 for Los Angeles). Use decimal values for latitude and longitude
-(no minutes/seconds). These values are approximate; the mesher will
-round them off to define a square mesh in UTM coordinates. When running
-benchmarks on rectangular models, turn the UTM projection off by using
-the flag \texttt{\small SUPPRESS\_UTM\_PROJECTION}, in which case
-all `longitude' parameters simply refer to the $x$~axis, and all
-`latitude' parameters simply refer to the $y$~axis. To run the mesher
-for a global simulation, the following parameters need to be set in
-the \texttt{Mesh\_Par\_file}:
-
-\begin{description}
-\item [{\texttt{LATITUDE\_MIN}}] Minimum latitude in the block (negative
-for South).
-\item [{\texttt{LATITUDE\_MAX}}] Maximum latitude in the block.
-\item [{\texttt{LONGITUDE\_MIN}}] Minimum longitude in the block (negative
-for West).
-\item [{\texttt{LONGITUDE\_MAX}}] Maximum longitude in the block.
-\item [{\texttt{DEPTH\_BLOCK\_KM}}] Depth of bottom of mesh in kilometers.
-\item [{\texttt{\noun{UTM\_PROJECTION\_ZONE}}}] UTM projection zone in
-which your model resides, only valid when \texttt{SUPPRESS\_UTM\_}~\\
-\texttt{PROJECTION} is \texttt{.false.}.
-\item [{\texttt{SUPPRESS\_UTM\_PROJECTION}}] set to be \texttt{.false.}
-when your model range is specified in geographical coordinates,
-and needs to be \texttt{.true.} when your model is specified in
-Cartesian coordinates. \noun{UTM projection zone in which your simulation
-region resides.}
-\item [{\texttt{INTERFACES\_FILE }}] File which contains the description of the topography and
-of the interfaces between the different layers of the model, if any.
-The number of spectral elements in the vertical direction within each layer is also defined in this file.
-
-\item [{$\nexxi$}] The number of spectral elements along one side of the
-block. This number \textit{must} be 8~$\times$~a multiple of $\nprocxi$
-defined below. Based upon benchmarks against semi-analytical discrete
-wavenumber synthetic seismograms \citep{KoLiTrSuStSh04}, determined
-that a $\nexxi=288$ run is accurate to a shortest period of roughly
-2~s. Therefore, since accuracy is determined by the number of grid
-points per shortest wavelength, for any particular value of $\nexxi$
-the simulation will be accurate to a shortest period determined by
-\begin{equation}
-\mbox{shortest period (s)}=(288/\nexxi)\times2.\label{eq:shortest_period}\end{equation}
-The number of grid points in each orthogonal direction of the reference
-element, i.e., the number of Gauss-Lobatto-Legendre points, is determined
-by \texttt{NGLLX} in the \texttt{constants.h} file. We generally use
-$\mbox{\texttt{NGLLX\/}}=5$, for a total of $5^{3}=125$ points per
-elements. We suggest not to change this value.
-\item [{$\nexeta$}] The number of spectral elements along the other side
-of the block. This number \textit{must} be 8~$\times$~a multiple
-of $\nproceta$ defined below.
-\item [{$\nprocxi$}] The number of processors or slices along one side
-of the block (see Figure~\ref{fig:For-parallel-computing}); we must
-have $\nexxi=8\times c\times\nprocxi$, where $c\ge1$ is a positive
-integer.
-\item [{$\nproceta$}] The number of processors or slices along the other
-side of the block; we must have $\nexeta=8\times c\times\nproceta$,
-where $c\ge1$ is a positive integer.
-\item [{\texttt{USE\_REGULAR\_MESH}}] set to be \texttt{.true.} if you want a perfectly regular
-mesh or \texttt{.false.} if you want to add doubling horizontal layers to coarsen the mesh. In this case, you
-also need to provide additional information by setting up the next three parameters.
-\item [{\texttt{NDOUBLINGS}}] The number of horizontal doubling layers. Must be set to \texttt{1} or \texttt{2}
-if \texttt{USE\_REGULAR\_MESH} is set to \texttt{.true.}.
-\item [{\texttt{NZ\_DOUBLING\_1}}] The position of the first doubling layer (only interpreted
-if \texttt{USE\_REGULAR\_MESH} is set to \texttt{.true.}).
-\item [{\texttt{NZ\_DOUBLING\_2}}] The position of the second doubling layer (only interpreted
-if \texttt{USE\_REGULAR\_MESH} is set to \texttt{.true.} and if \texttt{NDOUBLINGS} is set to \texttt{2}).
-\item [{\texttt{CREATE\_ABAQUS\_FILES}}] Set this flag to \texttt{.true.} to save Abaqus FEA \urlwithparentheses{www.simulia.com}
-mesh files for subsequent viewing. Turning the flag on generates files in the \texttt{LOCAL\_PATH} directory.
-See Section~\ref{sec:Mesh-graphics} for a discussion of mesh viewing features.
-\item [{\texttt{CREATE\_DX\_FILES}}] Set this flag to \texttt{.true.} to save OpenDX \urlwithparentheses{www.opendx.org}
-mesh files for subsequent viewing.
-\item [{\texttt{LOCAL\_PATH}}] Directory in which the partitions generated
-by the mesher will be written. Generally one uses a directory on the
-local disk of the compute nodes, although on some machines these partitions
-are written on a parallel (global) file system (see also the earlier
-discussion of the \texttt{LOCAL\_PATH\_IS\_ALSO\_GLOBAL} flag in Chapter~\ref{cha:Getting-Started}).
-The mesher generates the necessary partitions in parallel, one set
-for each of the $\nprocxi\times\nproceta$ slices that constitutes
-the mesh (see Figure~\ref{fig:For-parallel-computing}). After the
-mesher finishes, you can log in to one of the compute nodes and view
-the contents of the \texttt{LOCAL\_PATH} directory to see the
-files generated by the mesher.
-These files will be needed for creating the distributed databases, and have to reside in the directory  \texttt{LOCAL\_PATH} specified in the main \texttt{Par\_file}, e.g.
-in directory \texttt{in\_out\_files/DATABASES\_MPI}.
-Please see Chapter~\ref{cha:Creating-Distributed-Databases} for further details.
-\item [{\texttt{NMATERIALS}}] The number of different materials in your model. In the following lines, each material needs to be defined as :
-\begin{lyxcode}
-\texttt{material\_ID}  \texttt{rho}  \texttt{vp}  \texttt{vs}  \texttt{Q}  \texttt{anisotropy\_flag} \texttt{domain\_ID}
-\end{lyxcode}
-where
-\begin{itemize}
-     \item \texttt{Q}           : quality factor (0=no attenuation)
-     \item \texttt{anisotropy\_flag}  : 0=no anisotropy / 1,2,.. check with implementation in \texttt{aniso\_model.f90}
-     \item \texttt{domain\_id}        : 1=acoustic / 2=elastic
-\end{itemize}
-\item [{\texttt{NREGIONS}}] The number of regions in the mesh.
-In the following lines, because the mesh is regular or 'almost regular', each region is defined as :
-\begin{lyxcode}
-  \texttt{NEX\_XI\_BEGIN}  \texttt{NEX\_XI\_END}  \texttt{NEX\_ETA\_BEGIN}  \texttt{NEX\_ETA\_END}  \texttt{NZ\_BEGIN}  \texttt{NZ\_END}  \texttt{material\_ID}
-\end{lyxcode}
-\end{description}
-
-The \texttt{INTERFACES\_FILE} parameter of \texttt{Mesh\_Par\_File} defines the file which contains the settings of the topography grid and of the interfaces grids.
-Topography is defined as a set of elevation values on a regular 2D grid.
-It is also possible to
-define interfaces between the layers of the model in the same way.
-The file needs to define several parameters:
-\begin{itemize}
-\item The number of interfaces, including the topography. This needs to be set at the first line. Then, from the bottom
-to the top of the model, you need to define the grids with:
-\item \texttt{SUPPRESS\_UTM\_PROJECTION} flag as described previously,
-\item number of points along $x$ and $y$ direction (NXI and NETA),
-\item minimal $x$ and $y$ coordinates (LONG\_MIN and LAT\_MIN),
-\item spacing between points along $x$ and $y$ (SPACING\_XI and SPACING\_ETA) and
-\item the name of the file which contains the elevation values (in $y$.$x$ increasing order).
-\end{itemize}
-At the end of this file, you simply need to set the number of spectral elements in the vertical direction for each layer. We provide a few models in the {\texttt{examples/}} directory.  \\
-
-Finally, depending on your system, you might need to provide a file that tells MPI what compute nodes
-to use for the simulations. The file must have a number of entries
-(one entry per line) at least equal to the number of processors needed
-for the run. A sample file is provided in the file \texttt{mymachines}.
-This file is not used by the mesher or solver, but is required by
-the \texttt{go\_mesher} and \texttt{go\_solver} default job submission
-scripts. See Chapter~\ref{cha:Scheduler} for information about running
-the code on a system with a scheduler, e.g., LSF.
-
-Now that you have set the appropriate parameters in the \texttt{Mesh\_Par\_file}
-and have compiled the mesher, you are ready to launch it! This is
-most easily accomplished based upon the \texttt{go\_mesher} script.
-When you run on a PC cluster, the script assumes that the nodes are
-named n001, n002, etc. If this is not the case, change the \texttt{tr
--d `n'} line in the script. You may also need to edit the last command
-at the end of the script that invokes the \texttt{mpirun} command.
-See Chapter~\ref{cha:Scheduler} for information about running the
-code on a system with a scheduler, e.g., LSF.
-
-Mesher output is provided in the \texttt{in\_out\_files/OUTPUT\_FILES} directory
-in \texttt{output\_mesher.txt}; this file provides lots of details
-about the mesh that was generated. Please note that the mesher suggests a time step \texttt{DT} to run
-the solver with. The mesher output file also contains a table about the quality of the mesh to indicate
-possible problems with the distortions of elements.
-Alternatively, output can be directed
-to the screen instead by uncommenting a line in \texttt{constants.h}:
-\begin{lyxcode}
-!~uncomment~this~to~write~messages~to~the~screen~
-
-!~integer,~parameter~::~IMAIN~=~ISTANDARD\_OUTPUT
-\end{lyxcode}
-
-To control the quality of the mesh, check the standard output (either on the screen or in the \texttt{in\_out\_files/OUTPUT\_FILES} directory in \texttt{output\_mesher.txt}) and analyze the skewness histogram of mesh elements to make sure no element has a skewness above approximately 0.75, otherwise the mesh is of poor quality (and if even a single element has a skewness value above 0.80, then you must definitely improve the mesh).
-To draw the skewness histogram on the screen, type \texttt{gnuplot plot\_mesh\_quality\_histogram.gnu}.
-
-\chapter{\label{cha:Creating-Distributed-Databases}Creating the Distributed Databases}
-
-After using \texttt{xmeshfem3D} or \texttt{xdecompose\_mesh\_SCOTCH}, the next step in the workflow is to compile
-\texttt{xgenerate\_}~\\
-\texttt{databases}. This program is going to create all the missing information needed by the SEM solver.
-
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=.6\textwidth]{figures/workflow.jpg}
-\par\end{centering}
-
-\caption{Schematic workflow for a SPECFEM3D simulation. The executable \texttt{xgenerate\_databases} creates the GLL mesh points and assigns specific model parameters.}
-
-\label{fig:workflow.databases}
-\end{figure}
-
-In the main directory, type
-\begin{lyxcode}
-{\small make xgenerate\_databases}{\small \par}
-\end{lyxcode}
-Input for the program is provided through
-the main parameter file \texttt{Par\_file}, which resides in the subdirectory \texttt{in\_data\_files/}. Please note that
-\texttt{xgenerate\_databases} must be called directly from the \texttt{bin/} directory, as most of the binaries of the package.
-
-\section{\label{cha:Main-Parameter}Main parameter file \texttt{Par\_file}}
-
-Before running \texttt{xgenerate\_databases}, a number of parameters need to be set in the main parameter \texttt{Par\_file}
-located in the subdirectory \texttt{in\_data\_files/}:
-
-\begin{description}
-\item [{\texttt{SIMULATION\_TYPE}}] is set to 1 for forward simulations,
-2 for adjoint simulations (see Section \ref{sec:Adjoint-simulation-finite})
-and 3 for kernel simulations (see Section \ref{sec:Finite-Frequency-Kernels}).
-\item [{\texttt{SAVE\_FORWARD}}] is only set to \texttt{.true.} for a forward
-simulation with the last frame of the simulation saved, as part of
-the finite-frequency kernel calculations (see Section \ref{sec:Finite-Frequency-Kernels}).
-For a regular forward simulation, leave \texttt{SIMULATION\_TYPE}
-and \texttt{SAVE\_FORWARD} at their default values.
-\item [{\texttt{\noun{UTM\_PROJECTION\_ZONE}}}] UTM projection zone in
-which your model resides, only valid when \texttt{SUPPRESS\_UTM\_}~\\
-\texttt{PROJECTION} is \texttt{.false.}.
-\item [{\texttt{SUPPRESS\_UTM\_PROJECTION}}] set to be \texttt{.false.}
-when your model range is specified in the geographical coordinates,
-and needs to be \texttt{.true.} when your model is specified in a
-cartesian coordinates. \noun{UTM projection zone in which your simulation
-region resides.}
-\item [{\texttt{NPROC}}] The number of MPI processors, each one is assigned one slice of the whole mesh.
-\item [{\texttt{NSTEP}}] The number of time steps of the simulation. This controls the
-length of the numerical simulation, i.e., twice the number of time steps
-requires twice as much CPU time. This feature is not used at the time
-of generating the distributed databases but is required for the solver, i.e., you may change this
-parameter after running  \texttt{xgenerate\_databases}.
-\item [{\texttt{DT}}] The length of each time step in seconds.
-This feature is not used at the time of generating the distributed databases but is required for the solver.
-Please see also Section~\ref{sec:Choosing-the-Time-Step} for further details.
-%
-\item [{\texttt{MODEL}}] Must be set to one of the following:
-%
-\begin{description}
-\item [{\textmd{Models defined by mesh parameters:}}]~
-\begin{description}
-\item [{\texttt{default}}] Uses model parameters as defined by meshing procedures described
-in the previous Chapter~\ref{cha:Mesh-Generation}.
-\end{description}
-\end{description}
-%
-\begin{description}
-\item [{\textmd{1D~models~with~real~structure:}}]~
-\begin{description}
-\item [{\texttt{1D\_prem}}] Isotropic version of the spherically
-symmetric Preliminary Reference Earth Model (PREM) \citep{DzAn81}.
-\item [{\texttt{1D\_socal}}] A standard isotropic 1D model for Southern California.
-\item [{\texttt{1D\_cascadia}}] Isotropic 1D profile for the Cascadia region.
-\end{description}
-\end{description}
-%
-\begin{description}
-\item [{\textmd{Fully~3D~models:}}]~
-\begin{description}
-\item [{\texttt{aniso}}] For a user-specified fully anisotropic model.
-Parameters are set up in routines located in file \texttt{model\_aniso.f90} in directory \texttt{src/generate\_databases/}.
-See Chapter~\ref{cha:-Changing-the} for a discussion on how to specify your own 3D model.
-\item [{\texttt{external}}] For a user-specified isotropic model which uses externally defined model parameters.
-Uses external model definitions set up in routines located in file \texttt{model\_external\_values.f90} in directory \texttt{src/generate\_databases/}.
-Please modify these generic template routines to use your own model definitions.
-\item [{\texttt{gll}}] For a user-specified isotropic model which uses external binary files for $v_p$, $v_s$ and $\rho$.
-Binary files are given in the same format as when outputted by the \texttt{xgenerate\_databases} executable when using option \texttt{SAVE\_MESH\_FILES}.
-These binary files define the model parameters on all GLL points which can be used for iterative inversion procedures.
-\item [{\texttt{salton\_trough}}] A 3D $V_p$ model for Southern California.
-Users must provide the corresponding data file \texttt{regrid3\_vel\_p.bin} in directory \texttt{DATA/st\_3D\_block\_harvard/}.
-\item [{\texttt{tomo}}] For a user-specified 3D isotropic model which uses a tomographic model file \texttt{tomographic\_model.xyz} in directory \texttt{in\_data\_files/}.
-See Section \ref{sec:Using-tomographic},
-for a discussion on how to specify your own 3D tomographic model.
-\end{description}
-\end{description}
-%
-%
-\item [{\texttt{OCEANS}}] Set to \texttt{.true.} if the effect of the oceans
-on seismic wave propagation should be incorporated based upon the
-approximate treatment discussed in \citet{KoTr02b}. This feature
-is inexpensive from a numerical perspective, both in terms of memory
-requirements and CPU time. This approximation is accurate at periods
-of roughly 20~s and longer. At shorter periods the effect of water
-phases/reverberations is not taken into account, even when the flag
-is on.
-\item [{\texttt{TOPOGRAPHY}}] This feature is only effective if \texttt{OCEANS} is set to \texttt{.true.}.
-Set to \texttt{.true.} if topography and
-bathymetry should be read in based upon the topography file specified in the main
-constants file \texttt{constants.h} found in subdirectory \texttt{src/shared/} to evaluate
-elevations. If not set, elevations will be read from the numerical mesh.
-\item [{\texttt{ATTENUATION}}] Set to \texttt{.true.} if attenuation should
-be incorporated. Turning this feature on increases the memory requirements
-significantly (roughly by a factor of~1.5), and is numerically fairly
-expensive. See \citet{KoTr99,KoTr02a} for a discussion on the implementation
-of attenuation based upon standard linear solids.
-Please note that the Vp- and Vs-velocities of your model are given for a reference frequency.
-To change this reference frequency, you change the value of
-\texttt{ATTENUATION\_f0\_REFERENCE} in the main constants file \texttt{constants.h} found in subdirectory \texttt{src/shared/}.
-\item [{\texttt{USE\_OLSEN\_ATTENUATION}}] Set to \texttt{.true.} if you
-want to use the attenuation model that scaled from the S-wave speed model
-using Olsen's empirical relation (see \citet{OlDaBr03}).
-\item [{\texttt{ANISOTROPY}}] Set to \texttt{.true.} if you
-want to use an anisotropy model. Please see the file \texttt{model\_aniso.f90} in subdirectory \texttt{src/generate\_databases/} for the current implementation of anisotropic models.
-\item [{\texttt{ABSORBING\_CONDITIONS}}] Set to \texttt{.true.} to turn
-on Clayton-Enquist absorbing boundary conditions (see \citet{KoTr99}).
-\item [{\texttt{ABSORB\_INSTEAD\_OF\_FREE\_SURFACE}}] Set to \texttt{.true.} to turn on 
-absorbing boundary conditions on the top surface which by default constitutes a free surface of the model.
-\item [{\texttt{MOVIE\_SURFACE}}] Set to \texttt{.false.}, unless you want
-to create a movie of seismic wave propagation on the Earth's surface.
-Turning this option on generates large output files. See Section~\ref{sec:Movies}
-for a discussion on the generation of movies. This feature is only relevant for the solver.
-\item [{\texttt{MOVIE\_VOLUME}}] Set to \texttt{.false.}, unless you want
-to create a movie of seismic wave propagation in the Earth's interior.
-Turning this option on generates huge output files. See Section~\ref{sec:Movies}
-for a discussion on the generation of movies. This feature is only relevant for the solver.
-\item [{\texttt{NTSTEP\_BETWEEN\_FRAMES}}] Determines the number of timesteps
-between movie frames. Typically you want to save a snapshot every
-100 timesteps. The smaller you make this number the more output will
-be generated! See Section~\ref{sec:Movies} for a discussion on the
-generation of movies. This feature is only relevant for the solver.
-\item [{\texttt{CREATE\_SHAKEMAP}}] Set this flag to \texttt{.true.} to
-create a ShakeMap\textregistered, i.e., a peak ground velocity map
-of the maximum absolute value of the two horizontal components of the velocity vector.
-\item [{\texttt{SAVE\_DISPLACEMENT}}] Set this flag to \texttt{.true.}
-if you want to save the displacement instead of velocity for the movie
-frames.
-\item [{\texttt{USE\_HIGHRES\_FOR\_MOVIES}}] Set this flag to \texttt{.true.}
-if you want to save the values at all the NGLL grid points for the
-movie frames.
-\item [{\texttt{HDUR\_MOVIE}}] determines the half duration of the source time function for the movie simulations.
-When this parameter is set to be 0, a default half duration that corresponds to the accuracy of the simulation is provided.
-Otherwise, it adds this half duration to the half duration specified in the source file \texttt{CMTSOLUTION},
-thus simulates longer periods to make the movie images look smoother.
-\item [{\texttt{SAVE\_MESH\_FILES}}] Set this flag to \texttt{.true.} to
-save ParaView \urlwithparentheses{www.paraview.org}
-mesh files for subsequent viewing. Turning the flag on generates large
-(distributed) files in the \texttt{LOCAL\_PATH} directory. See Section~\ref{sec:Mesh-graphics}
-for a discussion of mesh viewing features.
-\item [{\texttt{LOCAL\_PATH}}] Directory in which the distributed databases
-will be written. Generally one uses a directory on the
-local disk of the compute nodes, although on some machines these databases
-are written on a parallel (global) file system (see also the earlier
-discussion of the \texttt{LOCAL\_PATH\_IS\_ALSO\_GLOBAL} flag in Chapter~\ref{cha:Getting-Started}).
-\texttt{xgenerate\_databases} generates the necessary databases in parallel, one set
-for each of the \texttt{NPROC} slices that constitutes
-the mesh (see Figure~\ref{fig:mount.partitions} and Figure~\ref{fig:For-parallel-computing}). After
-the executable finishes, you can log in to one of the compute nodes and view
-the contents of the \texttt{LOCAL\_PATH} directory to see the (many)
-files generated by \texttt{xgenerate\_databases}.
-Please note that the \texttt{LOCAL\_PATH} directory should already contain the output files of
-the partitioner, i.e. from \texttt{xdecompose\_mesh\_SCOTCH} or \texttt{xmeshfem3D}.
-\item [{\texttt{NTSTEP\_BETWEEN\_OUTPUT\_INFO}}] This parameter specifies
-the interval at which basic information about a run is written to
-the file system (\texttt{timestamp{*}} files in the \texttt{in\_out\_files/OUTPUT\_FILES}
-directory). If you have access to a fast machine, set \texttt{NTSTEP\_BETWEEN\_OUTPUT\_INFO}
-to a relatively high value (e.g., at least 100, or even 1000 or more)
-to avoid writing output text files too often. This feature is not
-used at the time of meshing. One can set this parameter to a larger
-value than the number of time steps to avoid writing output during
-the run.
-\item [{\texttt{NTSTEP\_BETWEEN\_OUTPUT\_SEISMOS}}] This parameter specifies
-the interval at which synthetic seismograms are written in the \texttt{LOCAL\_PATH}
-directory. If a run crashes, you may still find usable (but shorter
-than requested) seismograms in this directory. On a fast machine set
-\texttt{NTSTEP\_BETWEEN\_OUTPUT\_SEISMOS} to a relatively high value
-to avoid writing to the seismograms too often. This feature is only relevant for the solver.
-\item [{\texttt{USE\_FORCE\_POINT\_SOURCE}}] Turn this flag on to use a \texttt{FORCESOLUTION} force source
-instead of a \texttt{CMTSOLUTION} moment-tensor source. This can be useful e.g. for oil industry 
-foothills simulations in which the source is a vertical force, normal force, 
-tilted force, or an impact etc. You will need to give the East, North and vertical components of an 
-arbitrary (non-unitary) direction vector of the force vector; thus refer to Appendix~\ref{cha:Coordinates} 
-for the orientation of the reference frame.
-This vector is made unitary internally in the code and thus only its direction matters here;
-its norm is ignored and the norm of the force used is the factor force source times the source time function.
-\item [{\texttt{PRINT\_SOURCE\_TIME\_FUNCTION}}] Turn this flag on to print
-information about the source time function in the file \texttt{in\_out\_files/OUTPUT\_FILES/plot\_source\_time\_function.txt}.
-This feature is only relevant for the solver.
-\end{description}
-
-\section{\label{sec:Choosing-the-Time-Step}Choosing the time step \texttt{DT}}
-
-The parameter \texttt{DT} sets the length of each time step in seconds. The value of this parameter is crucial for the stability of
-the spectral-element simulation.
-Your time step \texttt{DT} will depend on the minimum ratio between the distance $h$ of neighboring mesh points and the wave speeds $v$ defined in your model.
-The condition for the time step $\Delta t$ is:
-
-\begin{lyxcode}
-$\Delta t < C ~\mathrm{min}_\Omega (~ h / v ~)$
-\end{lyxcode}
-where $C$ is the so-called Courant number and $\Omega$ denotes the model volume.
-The distance $h$ depends on the mesh element size and the number of
-GLL points \texttt{NGLL} specified in the main constants file \texttt{constants.h} located in the \texttt{src/shared/} subdirectory.
-The wave speed $v$ is determined based on your model's P- (or S-) wave speed values.
-
-The database generator \texttt{xgenerate\_databases}, as well as the internal mesher \texttt{xmeshfem3D}, are trying
-to evaluate the value of $\Delta t$ for empirically chosen Courant numbers $C \sim 0.3$.
-If you used the mesher \texttt{xmeshfem3D} to generate your mesh, you should set the value suggested in \texttt{in\_out\_files/OUTPUT\_FILES/output\_mesher.txt} file, which is created after the mesher completed.
-In case you used CUBIT to create the mesh, you might use an arbitrary value when running \texttt{xgenerate\_databases} and then use the value suggested in the ~\\
-\texttt{in\_out\_files/OUTPUT\_FILES/output\_mesher.txt} file after the database generation completed.
-Note that the implemented Newmark time scheme uses this time step globally, thus your simulations become more expensive for very small mesh elements in high wave-speed regions.
-Please be aware of this restriction when constructing your mesh in Chapter~\ref{cha:Mesh-Generation}.
-
-\chapter{\label{cha:Running-the-Solver}Running the Solver \texttt{xspecfem3D}}
-
-Now that you have successfully generated the databases, you are ready to compile
-the solver. In the main directory, type
-\begin{lyxcode}
-{\small make xspecfem3D}{\small \par}
-\end{lyxcode}
-Please note that \texttt{xspecfem3D} must be called directly from
-the \texttt{bin/} directory, as most of the binaries of the package.
-
-The solver needs three input files in the \texttt{in\_data\_files/} directory
-to run:
-\begin{description}
-\item [{\texttt{Par\_file}}] the main parameter file which was discussed in detail in the previous
-Chapter~\ref{cha:Creating-Distributed-Databases},
-\item [{\texttt{CMTSOLUTION}} or {\texttt{FORCESOLUTION}}] the earthquake source parameter file or the force source parameter file or, and
-\item [{\texttt{STATIONS}}] the stations file.
-\end{description}
-
-Most parameters in the \texttt{Par\_file}
-should be set prior to running the databases generation. Only the following parameters
-may be changed after running \texttt{xgenerate\_databases}:
-
-\begin{itemize}
-\item the simulation type control parameters: \texttt{SIMULATION\_TYPE}
-and \texttt{SAVE\_FORWARD}
-\item the time step parameters \texttt{NSTEP} and \texttt{DT}
-\item the absorbing boundary control parameter \texttt{ABSORBING\_CONDITIONS} on condition that the\\ 
-  \texttt{ABSORB\_INSTEAD\_OF\_FREE\_SURFACE} flag remains unmodified after running the databases generation. 
-\item the movie control parameters \texttt{MOVIE\_SURFACE}, \texttt{MOVIE\_VOLUME},
-and \texttt{NTSTEPS\_BETWEEN\_FRAMES}
-\item the ShakeMap\textregistered option \texttt{CREATE\_SHAKEMAP}
-\item the output information parameters \texttt{NTSTEP\_BETWEEN\_OUTPUT\_INFO}
-and \texttt{NTSTEP\_BETWEEN\_OUTPUT\_}~\\
-\texttt{SEISMOS}
-\item the \texttt{PRINT\_SOURCE\_TIME\_FUNCTION} flags
-\end{itemize}
-Any other change to the \texttt{Par\_file} implies rerunning both
-the database generator \texttt{xgenerate\_databases} and the solver \texttt{xspecfem3D}.
-
-For any particular earthquake, the \texttt{CMTSOLUTION} file that
-represents the point source may be obtained directly from the Harvard Centroid-Moment Tensor (CMT) web page \urlwithparentheses{www.seismology.harvard.edu}.
-It looks like this:
-
-\begin{lyxcode}
-{\small }%
-\begin{figure}[H]
-\noindent \begin{centering}
-{\small \includegraphics[width=1\textwidth]{figures/Hollywood_CMT} }
-\par\end{centering}{\small \par}
-
-\caption{\label{fig:CMTSOLUTION-file}\texttt{CMTSOLUTION} file based on the
-format from the Harvard CMT catalog. \textbf{M} is the moment tensor,
-$M_{0}${\small{} }is the seismic moment, and $M_{w}$ is the moment
-magnitude.}
-
-\end{figure}
-{\small \par}
-\end{lyxcode}
-The \texttt{CMTSOLUTION} should be edited in the following way:
-
-\begin{itemize}
-\item Set the \texttt{time shift} parameter equal to $0.0$ (the solver
-will not run otherwise.) The time shift parameter would simply apply
-an overall time shift to the synthetics, something that can be done
-in the post-processing (see Section \ref{sec:Process-data-and-syn}).
-\item For point-source simulations (see finite sources, page \pageref{To-simulate-a})
-we recommend setting the source half-duration parameter \texttt{half
-duration} equal to zero, which corresponds to simulating a step source-time
-function, i.e., a moment-rate function that is a delta function. If
-\texttt{half duration} is not set to zero, the code will use a Gaussian
-(i.e., a signal with a shape similar to a `smoothed triangle', as
-explained in \citet{KoTr02a} and shown in Fig~\ref{fig:gauss.vs.triangle})
-source-time function with half-width \texttt{half duration}. We prefer
-to run the solver with \texttt{half duration} set to zero and convolve
-the resulting synthetic seismograms in post-processing after the run,
-because this way it is easy to use a variety of source-time functions
-(see Section \ref{sec:Process-data-and-syn}). \citet{KoTr02a} determined
-that the noise generated in the simulation by using a step source
-time function may be safely filtered out afterward based upon a convolution
-with the desired source time function and/or low-pass filtering. Use
-the serial code \texttt{convolve\_source\_timefunction.f90} and the
-script \texttt{convolve\_source\_timefunction.csh} for this purpose,
-or alternatively use signal-processing software packages such as SAC \urlwithparentheses{www.llnl.gov/sac}.
-Type
-\begin{lyxcode}
-{\small make xconvolve\_source\_timefunction}{\small \par}
-\end{lyxcode}
-to compile the code and then set the parameter \texttt{hdur} in \texttt{convolve\_source\_timefunction.csh}
-to the desired half-duration.
-
-\item The zero time of the simulation corresponds to the center of the triangle/Gaussian,
-or the centroid time of the earthquake. The start time of the simulation
-is $t=-1.5*\texttt{half duration}$ (the 1.5 is to make sure the moment
-rate function is very close to zero when starting the simulation).
-To convert to absolute time $t_{\mathrm{abs}}$, set
-
-\begin{lyxcode}
-$t_{\mathrm{abs}}=t_{\mathrm{pde}}+\texttt{time shift}+t_{\mathrm{synthetic}}$
-\end{lyxcode}
-where $t_{\mathrm{pde}}$ is the time given in the first line of the
-\texttt{CMTSOLUTION}, \texttt{time shift} is the corresponding value
-from the original \texttt{CMTSOLUTION} file and $t_{\mathrm{synthetic}}$
-is the time in the first column of the output seismogram.
-
-\end{itemize}
-%
-\begin{figure}
-\noindent \begin{centering}
-%%%\includegraphics[width=3in]{figures/gauss_vs_triangle_mod.eps}
-\includegraphics[width=3in]{figures/gauss_vs_triangle_mod.pdf}
-\par\end{centering}
-
-\caption{Comparison of the shape of a triangle and the Gaussian function actually
-used.}
-
-\label{fig:gauss.vs.triangle}
-\end{figure}
-
-Centroid latitude and longitude should be provided in geographical
-coordinates. The code converts these coordinates to geocentric coordinates~\citep{DaTr98}.
-Of course you may provide your own source representations by designing
-your own \texttt{CMTSOLUTION} file. Just make sure that the resulting
-file adheres to the Harvard CMT conventions (see Appendix~\ref{cha:Coordinates}).
-Note that the first line in the \texttt{CMTSOLUTION} file is the Preliminary Determination of Earthquakes (PDE) solution performed by the USGS NEIC, which is used as a seed for the Harvard CMT inversion.
-The PDE solution is based upon P waves and often gives the hypocenter of the earthquake, i.e., the rupture initiation point, whereas the CMT solution gives the `centroid location', which is the location with dominant moment release.
-The PDE solution is not used by our software package but must be present anyway in the first line of the file.
-
-\label{To-simulate-a}To simulate a kinematic rupture, i.e., a finite-source
-event, represented in terms of $N_{\mathrm{sources}}$ point sources,
-provide a \texttt{CMTSOLUTION} file that has $N_{\mathrm{sources}}$
-entries, one for each subevent (i.e., concatenate $N_{\mathrm{sources}}$
-\texttt{CMTSOLUTION} files to a single \texttt{CMTSOLUTION} file).
-At least one entry (not necessarily the first) must have a zero \texttt{time
-shift}, and all the other entries must have non-negative \texttt{time
-shift}. Each subevent can have its own half duration, latitude, longitude,
-depth, and moment tensor (effectively, the local moment-density tensor).
-
-Note that the zero in the synthetics does NOT represent the hypocentral
-time or centroid time in general, but the timing of the \textit{center}
-of the source triangle with zero \texttt{time shift} (Fig~\ref{fig:source_timing}).
-
-Although it is convenient to think of each source as a triangle, in
-the simulation they are actually Gaussians (as they have better frequency
-characteristics). The relationship between the triangle and the gaussian
-used is shown in Fig~\ref{fig:gauss.vs.triangle}. For finite fault
-simulations it is usually not advisable to use a zero half duration
-and convolve afterwards, since the half duration is generally fixed
-by the finite fault model.
-
-\vspace{1cm}
-
-\noindent The \texttt{FORCESOLUTION} should be edited in the following way:
-
-\begin{itemize}
-\item Set the \texttt{time shift} parameter equal to $0.0$ (the solver
-will not run otherwise.) The time shift parameter would simply apply
-an overall time shift to the synthetics, something that can be done
-in the post-processing (see Section \ref{sec:Process-data-and-syn}).
-\item Set the latitude, longitude, depth of the source in geographical
-  coordinates.
-\item Set the magnitude of the force source.
-\item Set the components of a (non-unitary) direction vector for the
-  force source in the East/North/Vertical basis (see Appendix A for
-  the orientation of the reference frame).
-\end{itemize}
-
-{\small }%
-\begin{figure}[H]
-\noindent \begin{centering}
-{\small \includegraphics[width=5in]{figures/source_timing.pdf} }
-\par\end{centering}{\small \par}
-
-\caption{Example of timing for three sources. The center of the first source
-triangle is defined to be time zero. Note that this is NOT in general
-the hypocentral time, or the start time of the source (marked as tstart).
-The parameter \texttt{time shift} in the \texttt{CMTSOLUTION} file
-would be t1(=0), t2, t3 in this case, and the parameter \texttt{half
-duration} would be hdur1, hdur2, hdur3 for the sources 1, 2, 3 respectively.}
-
-{\small \label{fig:source_timing} }
-\end{figure}
-{\small \par}
-
-The solver can calculate seismograms at any number of stations for
-basically the same numerical cost, so the user is encouraged to include
-as many stations as conceivably useful in the \texttt{STATIONS} file,
-which looks like this:
-
-{\small }%
-\begin{figure}[H]
-\noindent \begin{centering}
-{\small \includegraphics{figures/STATIONS_basin_explained} }
-\par\end{centering}{\small \par}
-
-\caption{Sample \texttt{STATIONS} file. Station latitude and longitude should
-be provided in geographical coordinates. The width of the station
-label should be no more than 32 characters (see \texttt{MAX\_LENGTH\_STATION\_NAME}
-in the \texttt{constants.h} file), and the network label should be
-no more than 8 characters (see \texttt{MAX\_LENGTH\_NETWORK\_NAME}
-in the \texttt{constants.h} file).}
-
-\end{figure}
-{\small \par}
-
-Each line represents one station in the following format:
-
-\begin{lyxcode}
-{\small Station~Network~Latitude~(degrees)~Longitude~(degrees)~Elevation~(m)~burial~(m)~}{\small \par}
-\end{lyxcode}
-The solver \texttt{xspecfem3D} filters the list of stations in file \texttt{in\_data\_files/STATIONS}
-to exclude stations that are not located within the region given in
-the \texttt{Par\_file} (between \texttt{LATITUDE\_MIN} and \texttt{LATITUDE\_MAX}
-and between \texttt{LONGITUDE\_MIN} and \texttt{LONGITUDE\_MAX}).
-The filtered file is called \texttt{in\_data\_files/STATIONS\_FILTERED}.
-
-Solver output is provided in the \texttt{in\_out\_files/OUTPUT\_FILES} directory
-in the \texttt{output\_solver.txt} file. Output can be directed to
-the screen instead by uncommenting a line in \texttt{constants.h}:
-
-\begin{lyxcode}
-!~uncomment~this~to~write~messages~to~the~screen~
-
-!~integer,~parameter~::~IMAIN~=~ISTANDARD\_OUTPUT~
-\end{lyxcode}
-On PC clusters the seismogram files are generally written to the local
-disks (the path \texttt{LOCAL\_PATH} in the \texttt{Par\_file}) and
-need to be gathered at the end of the simulation.
-
-While the solver is running, its progress may be tracked by monitoring
-the `\texttt{\small timestamp{*}}' files in the ~\\
-\texttt{\small in\_out\_files/OUTPUT\_FILES} directory.
-These tiny files look something like this:
-
-\begin{lyxcode}
-Time~step~\#~~~~~~~~~~10000~
-
-Time:~~~~~108.4890~~~~~~seconds~
-
-Elapsed~time~in~seconds~=~~~~1153.28696703911~
-
-Elapsed~time~in~hh:mm:ss~=~~~~~0~h~19~m~13~s~
-
-Mean~elapsed~time~per~time~step~in~seconds~=~~~~~0.115328696703911~
-
-Max~norm~displacement~vector~U~in~all~slices~(m)~=~~~~1.0789589E-02~
-\end{lyxcode}
-The \texttt{\small timestamp{*}} files provide the \texttt{Mean elapsed
-time per time step in seconds}, which may be used to assess performance
-on various machines (assuming you are the only user on a node), as
-well as the \texttt{Max norm displacement vector U in all slices~(m)}.
-If something is wrong with the model, the mesh, or the source, you
-will see the code become unstable through exponentially growing values
-of the displacement and fluid potential with time, and ultimately
-the run will be terminated by the program. You can control the rate
-at which the timestamp files are written based upon the parameter
-\texttt{NTSTEP\_BETWEEN\_OUTPUT\_INFO} in the \texttt{Par\_file}.
-
-Having set the \texttt{Par\_file} parameters, and having provided the
-\texttt{CMTSOLUTION} (or the \texttt{FORCESOLUTION}) and
-\texttt{STATIONS} files, you are now ready to launch the solver! This
-is most easily accomplished based upon the \texttt{go\_solver} script
-(See Chapter \ref{cha:Scheduler} for information about running through
-a scheduler, e.g., LSF). You may need to edit the last command at the
-end of the script that invokes the \texttt{mpirun} command. The
-\texttt{runall} script compiles and runs both
-\texttt{xgenerate\_databases} and \texttt{xspecfem3D} in
-sequence. This is a safe approach that ensures using the correct
-combination of distributed database output and solver input.
-
-It is important to realize that the CPU and memory requirements of
-the solver are closely tied to choices about attenuation (\texttt{ATTENUATION})
-and the nature of the model (i.e., isotropic models are cheaper than
-anisotropic models). We encourage you to run a variety of simulations
-with various flags turned on or off to develop a sense for what is
-involved.
-
-For the same model, one can rerun the solver for different events by
-simply changing the \texttt{CMTSOLUTION} or \texttt{FORCESOLUTION}
-file, or for different stations by changing the \texttt{STATIONS}
-file. There is no need to rerun the \texttt{xgenerate\_databases}
-executable. Of course it is best to include as many stations as
-possible, since this does not add to the cost of the simulation.
-
-\chapter{\label{cha:Adjoint-Simulations}Adjoint Simulations}
-
-Adjoint simulations are generally performed for two distinct applications.
-First, they can be used for earthquake source inversions, especially
-earthquakes with large ruptures such as the Lander's earthquake \citep{WaHe94}.
-Second, they can be used to generate finite-frequency sensitivity
-kernels that are a critical part of tomographic inversions based upon
-3D reference models \citep{TrTaLi05,LiTr06,TrKoLi08,LiTr08}. In either
-case, source parameter or velocity structure updates are sought to
-minimize a specific misfit function (e.g., waveform or traveltime
-differences), and the adjoint simulation provides a means of computing
-the gradient of the misfit function and further reducing it in successive
-iterations. Applications and procedures pertaining to source studies
-and finite-frequency kernels are discussed in Sections~\ref{sec:Adjoint-simulation-sources}
-and \ref{sec:Adjoint-simulation-finite}, respectively. The two related
-parameters in the \texttt{Par\_file} are \texttt{SIMULATION\_TYPE}
-(1 or 2) and the \texttt{SAVE\_FORWARD} (boolean).
-
-\section{\label{sec:Adjoint-simulation-sources}Adjoint Simulations for Sources}
-
-In the case where a specific misfit function is minimized to invert
-for the earthquake source parameters, the gradient of the misfit function
-with respect to these source parameters can be computed by placing
-time-reversed seismograms at the receivers and using them as sources
-in an adjoint simulation, and then the value of the gradient is obtained
-from the adjoint seismograms recorded at the original earthquake location.
-
-\begin{enumerate}
-\item \textbf{Prepare the adjoint sources} \label{enu:Prepare-the-adjoint}
-
-\begin{enumerate}
-\item First, run a regular forward simlation (\texttt{SIMULATION\_TYPE =
-1} and \texttt{SAVE\_FORWARD = .false.}). You can automatically set
-these two variables using the \texttt{\small utils/change\_simulation\_type.pl}
-script:
-
-\begin{lyxcode}
-utils/change\_simulation\_type.pl~-f~
-\end{lyxcode}
-and then collect the recorded seismograms at all the stations given
-in \texttt{in\_data\_files/STATIONS}.
-
-\item Then select the stations for which you want to compute the time-reversed
-adjoint sources and run the adjoint simulation, and compile them into
-the \texttt{in\_data\_files/STATIONS\_ADJOINT} file, which has the same format
-as the regular \texttt{in\_data\_files/STATIONS} file.
-
-\begin{itemize}
-\item Depending on what type of misfit function is used for the source inversion,
-adjoint sources need to be computed from the original recorded seismograms
-for the selected stations and saved in the \texttt{in\_out\_files/SEM/} directory
-with the format \texttt{STA.NT.BX?.adj}, where \texttt{STA}, \texttt{NT}
-are the station name and network code given in the \texttt{in\_data\_files/STATIONS\_ADJOINT}
-file, and \texttt{BX?} represents the component name of a particular
-adjoint seismogram. Please note that the band code can change depending on
-your sampling rate (see Appendix~\ref{cha:channel-codes} for further details).
-\item The adjoint seismograms are in the same format as the original seismogram
-(\texttt{STA.NT.BX?.sem?}), with the same start time, time interval
-and record length.
-\end{itemize}
-\item Notice that even if you choose to time reverse only one component
-from one specific station, you still need to supply all three components
-because the code is expecting them (you can set the other two components
-to be zero).
-\item Also note that since time-reversal is done in the code itself, no
-explicit time-reversing is needed for the preparation of the adjoint
-sources, i.e., the adjoint sources are in the same forward time sense
-as the original recorded seismograms.
-\end{enumerate}
-\item \textbf{Set the related parameters and run the adjoint simulation}\\
-In the \texttt{in\_data\_files/Par\_file}, set the two related parameters to
-be \texttt{SIMULATION\_TYPE = 2} and \texttt{SAVE\_FORWARD = .false.}.
-More conveniently, use the scripts \texttt{utils/change\_simulation\_type.pl}
-to modify the \texttt{Par\_file} automatically (\texttt{change\_simulation\_type.pl
--a}). Then run the solver to launch the adjoint simulation.
-\item \textbf{Collect the seismograms at the original source location}
-
-After the adjoint simulation has completed successfully, collect the
-seismograms from \texttt{LOCAL\_PATH}.
-
-\begin{itemize}
-\item These adjoint seismograms are recorded at the locations of the original
-earthquake sources given by the \texttt{in\_data\_files/CMTSOLUTION} file, and
-have names of the form \texttt{S?????.NT.S??.sem} for the six-component
-strain tensor (\texttt{SNN,SEE,SZZ,SNE,SNZ,SEZ}) at these locations,
-and ~\\
-\texttt{S?????.NT.BX?.sem} for the three-component displacements
-(\texttt{BXN,BXE,BXZ}) recorded at these locations.
-\item \texttt{S?????} denotes the source number; for example, if the original
-\texttt{CMTSOLUTION} provides only a point source, then the seismograms
-collected will start with \texttt{S00001}.
-\item These adjoint seismograms provide critical information for the computation
-of the gradient of the misfit function.
-\end{itemize}
-\end{enumerate}
-
-\section{\label{sec:Adjoint-simulation-finite}Adjoint Simulations for Finite-Frequency
-Kernels (Kernel Simulation)}
-
-Finite-frequency sensitivity kernels are computed in two successive
-simulations (please refer to \citet{LiTr06} and \citet{TrKoLi08} for details).
-
-\begin{enumerate}
-\item \textbf{Run a forward simulation with the state variables saved at
-the end of the simulation}
-
-Prepare the \texttt{\small CMTSOLUTION} and \texttt{\small STATIONS}
-files, set the parameters \texttt{\small SIMULATION\_TYPE}{\small{}
-}\texttt{\small =}{\small{} }\texttt{\small 1} and \texttt{\small SAVE\_FORWARD
-=}{\small{} }\texttt{\small .true.} in the \texttt{Par\_file} (\texttt{change\_simulation\_type
--F}), and run the solver.
-
-\begin{itemize}
-\item Notice that attenuation is not implemented yet for the computation
-of finite-frequency kernels; therefore set \texttt{ATTENUATION = .false.}
-in the \texttt{Par\_file}.
-\item We also suggest you modify the half duration of the \texttt{CMTSOLUTION}
-to be similar to the accuracy of the simulation (see Equation \ref{eq:shortest_period})
-to avoid too much high-frequency noise in the forward wavefield, although
-theoretically the high-frequency noise should be eliminated when convolved
-with an adjoint wavefield with the proper frequency content.
-\item This forward simulation differs from the regular simulations (\texttt{\small SIMULATION\_TYPE}{\small{}
-}\texttt{\small =}{\small{} }\texttt{\small 1} and \texttt{\small SAVE\_FORWARD}{\small{}
-}\texttt{\small =}{\small{} }\texttt{\small .false.}) described in
-the previous chapters in that the state variables for the last time
-step of the simulation, including wavefields of the displacement,
-velocity, acceleration, etc., are saved to the \texttt{LOCAL\_PATH}
-to be used for the subsequent simulation.
-\item For regional simulations, the files recording the absorbing boundary
-contribution are also written to the \texttt{LOCAL\_PATH} when \texttt{SAVE\_FORWARD
-= .true.}.
-\end{itemize}
-\item \textbf{Prepare the adjoint sources}
-
-The adjoint sources need to be prepared the same way as described
-in the Section \ref{enu:Prepare-the-adjoint}.
-
-\begin{itemize}
-\item In the case of travel-time finite-frequency kernel for one source-receiver
-pair, i.e., point source from the \texttt{CMTSOLUTION}, and one station
-in the \texttt{STATIONS\_ADJOINT} list, we supply a sample program
-in \texttt{utils/adjoint\_sources/traveltime/xcreate\_adjsrc\_traveltime} to cut a certain portion of the original
-displacement seismograms and convert it into the proper adjoint source
-to compute the finite-frequency kernel.
-\begin{lyxcode}
-xcreate\_adjsrc\_traveltime~t1~t2~ifile{[}0-5]~E/N/Z-ascii-files~{[}baz]
-\end{lyxcode}
-where \texttt{t1} and \texttt{t2} are the start and end time of the
-portion you are interested in, \texttt{ifile} denotes the component
-of the seismograms to be used (0 for all three components, 1 for East,
-2 for North, and 3 for vertical, 4 for transverse, and 5 for radial
-component), \texttt{E/N/Z-ascii-files} indicate the three-component
-displacement seismograms in the right order, and \texttt{baz} is the
-back-azimuth of the station. Note that \texttt{baz} is only supplied
-when \texttt{ifile} = 4 or 5.
-
-\item Similarly, a sample program to compute adjoint sources for amplitude finite-frequency kernels
-may be found in \texttt{utils/adjoint\_sources/amplitude} and used in the same way as described for
-traveltime measurements
-\begin{lyxcode}
-xcreate\_adjsrc\_amplitude~t1~t2~ifile{[}0-5]~E/N/Z-ascii-files~{[}baz].
-\end{lyxcode}
-
-\end{itemize}
-
-\item \textbf{Run the kernel simulation}
-
-With the successful forward simulation and the adjoint source ready
-in the \texttt{in\_out\_files/SEM/} directory, set \texttt{SIMULATION\_TYPE = 3} and \texttt{SAVE\_FORWARD
-= .false.} in the \texttt{Par\_file} (you can use \texttt{change\_simulation\_type.pl -b}),
-and rerun the solver.
-
-\begin{itemize}
-\item The adjoint simulation is launched together with the back reconstruction
-of the original forward wavefield from the state variables saved from
-the previous forward simulation, and the finite-frequency kernels
-are computed by the interaction of the reconstructed forward wavefield
-and the adjoint wavefield.
-\item The back-reconstructed seismograms at the original station locations
-are saved to the \texttt{LOCAL\_PATH} at the end of the kernel simulations,
-and can be collected to the local disk.
-\item These back-constructed seismograms can be compared with the time-reversed
-original seismograms to assess the accuracy of the backward reconstruction,
-and they should match very well.
-\item The arrays for density, P-wave speed and S-wave speed kernels are
-also saved in the \texttt{LOCAL\_PATH} with the names \texttt{proc??????\_rho(alpha,beta)\_kernel.bin},
-where \texttt{proc??????} represents the processor number, \texttt{rho(alpha,beta)}
-are the different types of kernels.
-\end{itemize}
-\end{enumerate}
-In general, the three steps need to be run sequentially to assure
-proper access to the necessary files. If the simulations are run through
-some cluster scheduling system (e.g., LSF), and the forward simulation
-and the subsequent kernel simulations cannot be assigned to the same
-set of computer nodes, the kernel simulation will not be able to access
-the database files saved by the forward simulation. Solutions for
-this dilemma are provided in Chapter~\ref{cha:Scheduler}. Visualization
-of the finite-frequency kernels is discussed in Section~\ref{sec:Finite-Frequency-Kernels}.
-
-\chapter{Noise Cross-correlation Simulations}
-
-Besides earthquake simulations, SPECFEM3D includes functionality for seismic noise tomography as well.
-In order to proceed successfully in this chapter, it is critical that you have already familiarized yourself
-with procedures for meshing (Chapter \ref{cha:Mesh-Generation}),
-creating distributed databases (Chapter \ref{cha:Creating-Distributed-Databases}),
-running earthquake simulations (Chapters \ref{cha:Running-the-Solver})
-and adjoint simulations (Chapter \ref{cha:Adjoint-Simulations}).
-Also, make sure you read the article `Noise cross-correlation sensitivity kernels' \citep{trompetal2010},
-in order to understand noise simulations from a theoretical perspective.
-
-\section{Input Parameter Files}
-
-As usual, the three main input files are crucial: \texttt{Par\_file}, \texttt{CMTSOLUTION} and \texttt{STATIONS}.
-Unless otherwise specified, those input files should be located in directory \texttt{in\_data\_files/}.\\
-
-\texttt{CMTSOLUTION} is required for all simulations.
-At a first glance, it may seem unexpected to have it here,
-since the noise simulations should have nothing to do with the earthquake -- \texttt{CMTSOLUTION}.
-However, for noise simulations, it is critical to have no earthquakes.
-In other words, the moment tensor specified in \texttt{CMTSOLUTION} must be set to zero manually!\\
-
-\texttt{STATIONS} remains the same as in previous earthquake simulations,
-except that the order of receivers listed in \texttt{STATIONS} is now important.
-The order will be used to determine the `master' receiver,
-i.e., the one that simultaneously cross correlates with the others.\\
-
-\texttt{Par\_file} also requires careful attention.
-A parameter called \texttt{NOISE\_TOMOGRAPHY} has been added which specifies the type of simulation to be run.
-\texttt{NOISE\_TOMOGRAPHY} is an integer with possible values 0, 1, 2 and 3.
-For example, when \texttt{NOISE\_TOMOGRAPHY} equals 0, a regular earthquake simulation will be run.
-When it is 1/2/3, you are about to run
-step 1/2/3 of the noise simulations respectively.
-Should you be confused by the three steps, refer to \citet{trompetal2010} for details.\\
-
-Another change to \texttt{Par\_file} involves the parameter \texttt{NSTEP}.
-While for regular earthquake simulations this parameter specifies the length of synthetic seismograms generated,
-for noise simulations it specifies the length of the seismograms used to compute cross correlations.
-The actual cross correlations are thus twice this length, i.e., $2~\mathrm{NSTEP}-1$.
-The code automatically makes the modification accordingly, if \texttt{NOISE\_TOMOGRAPHY} is not zero.\\
-
-There are other parameters in \texttt{Par\_file} which should be given specific values.
-For instance, since the first two steps for calculating noise sensitivity kernels correspond to
-forward simulations, \texttt{SIMULATION\_TYPE} must be 1 when \texttt{NOISE\_TOMOGRAPHY} equals 1 or 2.
-Also, we have to reconstruct the ensemble forward wavefields in adjoint simulations,
-therefore we need to set \texttt{SAVE\_FORWARD} to \texttt{.true.} for the second step,
-i.e., when \texttt{NOISE\_TOMOGRAPHY} equals 2.
-The third step is for kernel constructions. Hence \texttt{SIMULATION\_TYPE} should be 3, whereas
-\texttt{SAVE\_FORWARD} must be \texttt{.false.}.\\
-
-Finally, for most system architectures,
-please make sure that \texttt{LOCAL\_PATH} in \texttt{Par\_file} is in fact local, not globally shared.
-Because we have to save the wavefields at the earth's surface at every time step,
-it is quite problematic to have a globally shared \texttt{LOCAL\_PATH},
-in terms of both disk storage and I/O speed.
-
-\section{Noise Simulations: Step by Step}
-
-Proper parameters in those parameter files are not enough for noise simulations to run.
-We have more parameters to specify:
-for example, the ensemble-averaged noise spectrum, the noise distribution etc.
-However, since there are a few `new' files, it is better to introduce them sequentially.
-In this section, standard procedures for noise simulations are described.
-
-\subsection{Pre-simulation}
-
-\begin{itemize}
-
-\item
-As usual, we first configure the software package using:
-
-\texttt{./configure FC=ifort MPIFC=mpif90}
-
-Use the following if SCOTCH is needed:
-
-\texttt{./configure FC=ifort MPIFC=mpif90 --with-scotch-dir=/opt/scotch}\\
-
-\item
-Next, we need to compile the source code using:
-
-\texttt{make xgenerate\_databases}
-
-\texttt{make xspecfem3D} \\
-
-\item
-Before we can run noise simulations, we have to specify the noise statistics,
-e.g., the ensemble-averaged noise spectrum.
-Matlab scripts are provided to help you to generate the necessary file.
-
-\texttt{examples/noise\_tomography/NOISE\_TOMOGRAPHY.m}  (main program)\\
-\texttt{examples/noise\_tomography/PetersonNoiseModel.m}
-
-In Matlab, simply run:
-
-\texttt{NOISE\_TOMOGRAPHY(NSTEP, DT, Tmin, Tmax, NOISE\_MODEL)}\\
-
-\texttt{DT} is given in \texttt{Par\_file},
-but \texttt{NSTEP} is NOT the one specified in \texttt{Par\_file}.
-Instead, you have to feed $2~\mathrm{NSTEP}-1$ to account for the doubled length of cross correlations.
-\texttt{Tmin} and \texttt{Tmax} correspond to the period range you are interested in,
-whereas \texttt{NOISE\_MODEL} denotes the noise model you will be using.
-Details can be found in the Matlab script.
-
-After running the Matlab script, you will be given the following information (plus a figure in Matlab):
-
-\texttt{*************************************************************}\\
-\texttt{the source time function has been saved in:}\\
-\texttt{/data2/yangl/3D\_NOISE/S\_squared} (note this path must be different)\\
-\texttt{S\_squared should be put into directory:}\\
-\texttt{in\_out\_files/NOISE\_TOMOGRAPHY/ in the SPECFEM3D package}\\
-
-In other words, the Matlab script creates a file called \texttt{S\_squared},
-which is the first `new' input file we encounter for noise simulations.
-
-One may choose a flat noise spectrum rather than Peterson's noise model.
-This can be done easily by modifying the Matlab script a little.
-
-\item
-Create a new directory in the SPECFEM3D package, name it as \texttt{in\_out\_files/NOISE\_TOMOGRAPHY/}.
-We will add some parameter files later in this folder.
-
-\item
-Put the Matlab-generated-file \texttt{S\_squared} in \texttt{in\_out\_files/NOISE\_TOMOGRAPHY/}.
-
-That's to say, you will have a file \texttt{in\_out\_files/NOISE\_TOMOGRAPHY/S\_squared} in the SPECFEM3D package.
-
-\item
-Create a file called \texttt{in\_out\_files/NOISE\_TOMOGRAPHY/irec\_master\_noise}. Note that this file is located
-in directory \texttt{in\_out\_files/NOISE\_TOMOGRAPHY/} as well. In general, all noise simulation related parameter
-files go into that directory.
-\texttt{irec\_master\_noise} contains only one integer, which is the ID
-of the `master' receiver. For example, if this file contains 5, it means that the fifth receiver
-listed in \texttt{in\_data\_files/STATIONS} becomes the `master'. That's why we mentioned previously that
-the order of receivers in \texttt{in\_data\_files/STATIONS} is important.
-
-Note that in regional simulations, the \texttt{in\_data\_files/STATIONS}
-might contain receivers which are outside of our computational domains.
-Therefore, the integer in \texttt{irec\_master\_noise}
-is actually the ID in \texttt{in\_data\_files/STATIONS\_FILTERED}
-(which is generated by \texttt{bin/xgenerate\_databases}).
-
-\item
-Create a file called \texttt{in\_out\_files/NOISE\_TOMOGRAPHY/nu\_master}. This file holds three numbers,
-forming a (unit) vector. It describes which component we are cross-correlating at the `master'
-receiver, i.e., ${\hat{\bf \nu}}^{\alpha}$ in \citet{trompetal2010}.
-The three numbers correspond to E/N/Z components respectively.
-Most often, the vertical component is used, and in those cases the three numbers should be 0, 0 and 1.
-
-\item
-Describe the noise direction and distributions in \texttt{src/shared/noise\_tomography.f90}.
-Search for a subroutine called \texttt{noise\_distribution\_direction}
-in \texttt{noise\_tomography.f90}. It is actually located at the very beginning
-of \texttt{noise\_tomography.f90}. The default assumes vertical noise and a uniform distribution
-across the whole free surface of the model. It should be quite self-explanatory for modifications.
-Should you modify this part, you have to re-compile the source code.
-
-\end{itemize}
-
-\subsection{Simulations}
-
-With all of the above done, we can finally launch our simulations.
-Again, please make sure that the \texttt{LOCAL\_PATH} in \texttt{in\_data\_files/Par\_file} is not globally shared.
-It is quite problematic to have a globally shared \texttt{LOCAL\_PATH}, in terms of both disk storage and
-speed of I/O (we have to save the wavefields at the earth's surface at every time step).
-
-As discussed in \citet{trompetal2010}, it takes three steps/simulations to obtain one contribution
-of the ensemble sensitivity kernels:
-
-\begin{itemize}
-\item
-Step 1: simulation for generating wavefields\\
-\texttt{SIMULATION\_TYPE=1}\\
-\texttt{NOISE\_TOMOGRAPHY=1}\\
-\texttt{SAVE\_FORWARD} not used, can be either .true. or .false.
-
-\item
-Step 2: simulation for ensemble forward wavefields\\
-\texttt{SIMULATION\_TYPE=1}\\
-\texttt{NOISE\_TOMOGRAPHY=2}\\
-\texttt{SAVE\_FORWARD=.true.}
-
-\item
-Step 3: simulation for ensemble adjoint wavefields and sensitivity kernels\\
-\texttt{SIMULATION\_TYPE=3}\\
-\texttt{NOISE\_TOMOGRAPHY=3}\\
-\texttt{SAVE\_FORWARD=.false.}
-
-Note Step 3 is an adjoint simulation, please refer to previous chapters on how to prepare adjoint sources
-and other necessary files, as well as how adjoint simulations work.
-
-\end{itemize}
-
-Note that it is better to run the three steps continuously within the same job,
-otherwise you have to collect the saved surface movies from the old nodes/CPUs to the new nodes/CPUs.
-This process varies from cluster to cluster and thus cannot be discussed here.
-Please ask your cluster administrator for information/configuration of the cluster you are using.\\
-
-\subsection{Post-simulation}
-
-After those simulations, you have all stuff you need, either in the \texttt{in\_out\_files/OUTPUT\_FILES/} or
-in the directory specified by \texttt{LOCAL\_PATH} in \texttt{in\_data\_files/Par\_file}
-(which are most probably on local nodes).
-Collect whatever you want from the local nodes to your workstation, and then visualize them.
-This process is the same as what you may have done for regular earthquake simulations.
-Refer to other chapters if you have problems.\\
-
-Simply speaking, two outputs are the most interesting: the simulated ensemble cross correlations and
-one contribution of the ensemble sensitivity kernels.\\
-
-The simulated ensemble cross correlations are obtained after the second simulation (Step 2).
-Seismograms in \texttt{in\_out\_files/OUTPUT\_FILES/} are actually the simulated ensemble cross correlations.
-Collect them immediately after Step 2, or the Step 3 will overwrite them.
-Note that we have a `master' receiver specified by \texttt{irec\_master\_noise},
-the seismogram at one station corresponds to the cross correlation between that station and the `master'.
-Since the seismograms have three components, we may obtain cross correlations
-for different components as well, not necessarily the cross correlations between vertical components.\\
-
-One contribution of the ensemble cross-correlation sensitivity kernels are obtained after Step 3,
-stored in the \texttt{in\_data\_files/LOCAL\_PATH} on local nodes.
-The ensemble kernel files are named the same as regular earthquake kernels.\\
-
-You need to run another three simulations to get the other contribution of the ensemble kernels,
-using different forward and adjoint sources given in \citet{trompetal2010}.
-
-\section{Example}
-
-In order to illustrate noise simulations in an easy way, one example is provided in \texttt{examples/noise\_tomography/}.
-See \texttt{examples/noise\_tomography/README} for explanations. \\
-
-Note, however, that they are created for a specific workstation (CLOVER@PRINCETON),
-which has at least 4 cores with `mpif90' working properly. \\
-
-If your workstation is suitable, you can run the example in \texttt{examples/noise\_tomography/} using:\\
-
-\texttt{./pre-processing.sh}\\
-
-Even if this script does not work on your workstation, the procedure it describes is universal.
-You may review the whole process described in the last section by following the commands in \texttt{pre-processing.sh},
-which should contain enough explanations for all the commands.
-
-\chapter{Graphics}
-
-\section{\label{sec:Mesh-graphics}Meshes}
-
-In case you used the internal mesher \texttt{xmeshfem3D} to create and partition your mesh,
-you can output mesh files in ABAQUS (.INP) and DX (.dx) format to visualize them. For this, you must set either the
-flag \texttt{CREATE\_DX\_FILES} or \texttt{CREATE\_ABAQUS\_FILES} to \texttt{.true.}
-in the mesher's parameter file \texttt{Mesh\_Par\_file}
-prior to running the mesher (see Chapter~\ref{cha:Running-the-Mesher-Meshfem3D} for details).
-You can then use AVS \urlwithparentheses{www.avs.com} or OpenDX \urlwithparentheses{www.opendx.org}
-to visualize the mesh and MPI partition (slices).
-
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=.49\textwidth]{figures/vtk_mesh_vp.jpg}
-\includegraphics[width=.49\textwidth]{figures/vtk_mesh_vs.jpg}
-\par\end{centering}
-
-\caption{Visualization using Paraview of VTK files created by \texttt{xgenerate\_databases} showing P- and S-wave velocities assigned to the mesh points.
-The mesh was created by \texttt{xmeshfem3D} for 4 processors.}
+\caption{Visualization using Paraview of VTK files created by \texttt{xgenerate\_databases}
+showing P- and S-wave velocities assigned to the mesh points. The
+mesh was created by \texttt{xmeshfem3D} for 4 processors.}
 
@@
 
-You have also the option to visualize the distributed databases produced by \texttt{xgenerate\_databases}
-using Paraview \urlwithparentheses{www.paraview.org}. For this, you must set the flag \texttt{SAVE\_MESH\_FILES} to \texttt{.true.}
-in the main parameter file \texttt{Par\_file} (see Chapter~\ref{cha:Main-Parameter} for details).
-This will create VTK files for each single partition. You can then use Paraview \urlwithparentheses{www.paraview.org} to visualized these partitions.
-
-\section{\label{sec:Movies}Movies}
+You have also the option to visualize the distributed databases produced
+by \texttt{xgenerate\_databases} using Paraview \urlwithparentheses{www.paraview.org}.
+For this, you must set the flag \texttt{SAVE\_MESH\_FILES} to \texttt{.true.}
+in the main parameter file \texttt{Par\_file} (see Chapter~\ref{cha:Main-Parameter}
+for details). This will create VTK files for each single partition.
+You can then use Paraview \urlwithparentheses{www.paraview.org}
+to visualized these partitions.
+
+\section{Movies}\label{sec:Movies}
 
 To make a surface or volume movie of the simulation, set parameters
-\texttt{MOVIE\_SURFACE}, \texttt{MOVIE\_VOLUME}, and \texttt{NTSTEP\_BETWEEN\_FRAMES}
-in the \texttt{Par\_file}. Turning on the movie flags, in particular
-\texttt{MOVIE\_VOLUME}, produces large output files. \texttt{MOVIE\_VOLUME}
-files are saved in the \texttt{LOCAL\_PATH} directory, whereas \texttt{MOVIE\_SURFACE}
-output files are saved in the \texttt{in\_out\_files/OUTPUT\_FILES} directory. We
-save the displacement field if the parameter \texttt{SAVE\_DISPLACEMENT} is set, otherwise the velocity field is saved.
-The look of a movie is determined by the
-half-duration of the source. The half-duration should be large enough
-so that the movie does not contain frequencies that are not resolved
-by the mesh, i.e., it should not contain numerical noise. This can
-be accomplished by selecting a CMT \texttt{HALF\_DURATION} > 1.1 $\times$
-smallest period (see figure \ref{fig:CMTSOLUTION-file}). When \texttt{MOVIE\_SURFACE}
-= .\texttt{true.}, the half duration of each source in the \texttt{CMTSOLUTION}
-file is replaced by
-
+\texttt{MOVIE\_SURFACE}, \texttt{MOVIE\_VOLUME}, \texttt{MOVIE\_TYPE},
+and \texttt{NTSTEP\_BETWEEN\_FRAMES} in the \texttt{Par\_file}. Turning
+on the movie flags, in particular \texttt{MOVIE\_VOLUME}, produces
+large output files. \texttt{MOVIE\_VOLUME} files are saved in the
+\texttt{LOCAL\_PATH} directory, whereas \texttt{MOVIE\_SURFACE} output
+files are saved in the \texttt{OUTPUT\_FILES} directory. We save the
+displacement field if the parameter \texttt{SAVE\_DISPLACEMENT} is
+set, otherwise the velocity field is saved. The look of a movie is
+determined by the half-duration of the source. The half-duration should
+be large enough so that the movie does not contain frequencies that
+are not resolved by the mesh, i.e., it should not contain numerical
+noise. This can be accomplished by selecting a CMT \texttt{HALF\_DURATION}
+> 1.1 $\times$ smallest period (see figure \ref{fig:CMTSOLUTION-file}).
+When \texttt{MOVIE\_SURFACE} = .\texttt{true.}, the half duration
+of each source in the \texttt{CMTSOLUTION} file is replaced by
 \begin{quote}
 \[
-\sqrt{(}\mathrm{\mathtt{HALF\_DURATIO}\mathtt{N}^{2}}+\mathrm{\mathtt{HDUR\_MOVI}\mathtt{E}^{2}})\]
+\sqrt{(}\mathrm{\mathtt{HALF\_DURATIO}\mathtt{N}^{2}}+\mathrm{\mathtt{HDUR\_MOVI}\mathtt{E}^{2}})
+\]
  \textbf{NOTE:} If \texttt{HDUR\_MOVIE} is set to 0.0, the code will
@@
 usual, for a point source one can set \texttt{half duration} in the
-\texttt{CMTSOLUTION} file to be 0.0 and \texttt{HDUR\_MOVIE} = 0.0 in the \texttt{Par\_file} to get
-the highest frequencies resolved by the simulation, but for a finite
-source one would keep all the \texttt{half durations} as prescribed
-by the finite source model and set \texttt{HDUR\_MOVIE} = 0.0.
+\texttt{CMTSOLUTION} file to be 0.0 and \texttt{HDUR\_MOVIE} = 0.0
+in the \texttt{Par\_file} to get the highest frequencies resolved
+by the simulation, but for a finite source one would keep all the
+\texttt{half durations} as prescribed by the finite source model and
+set \texttt{HDUR\_MOVIE} = 0.0.
 \end{quote}
@@
 flag turned on, the code outputs \texttt{moviedata??????} files in
-the \texttt{in\_out\_files/OUTPUT\_FILES} directory. There are several flags
-in the main parameter file \texttt{Par\_file} that control the output of these moviedata files
-(see section \ref{cha:Main-Parameter} for details):
-\texttt{\small NTSTEP\_BETWEEN\_FRAMES} to set the timesteps between frames,
-\texttt{\small SAVE\_DISPLACEMENT} to save displacement instead of velocity,
-\texttt{USE\_HIGHRES\_FOR\_MOVIES} to save values at all GLL point instead of element edges.
-In order to output additionally shakemaps, you would set
-the parameter \texttt{CREATE\_SHAKEMAP} to \texttt{.true.}.
-
-The files are in a fairly complicated
-binary format, but there is a program provided to convert the
-output into more user friendly formats:
+the \texttt{OUTPUT\_FILES} directory. There are several flags in the
+main parameter file \texttt{Par\_file} that control the output of
+these movie data files (see section \ref{cha:Main-Parameter} for
+details): \texttt{\small NTSTEP\_BETWEEN\_FRAMES}{\small{} to set the
+timesteps between frames, }\texttt{\small SAVE\_DISPLACEMENT}{\small{}
+to save displacement instead of velocity, }\texttt{\small USE\_HIGHRES\_FOR\_MOVIES}{\small{}
+to save values at all GLL point instead of element edges. In order
+to output additionally shakemaps, you would set the parameter }\texttt{\small CREATE\_SHAKEMAP}{\small{}
+to }\texttt{\small .true.}{\small .}{\small \par}
+
+The files are in a fairly complicated binary format, but there is
+a program provided to convert the output into more user friendly formats:
 \begin{description}
-\item [\texttt{xcreate\_movie\_shakemap\_AVS\_DX\_GMT}]
-From \texttt{create\_movie\_shakemap\_AVS\_DX\_GMT.f90}, it
-outputs data in ASCII, OpenDX, or AVS format (also readable in ParaView).
-Before compiling the code, make sure you have the file \texttt{surface\_from\_mesher.h}
-in the  \texttt{in\_out\_files/OUTPUT\_FILES/} directory. This file will be created by the solver run.
-Then type
-\begin{lyxcode}
-{\footnotesize make xcreate\_movie\_shakemap\_AVS\_DX\_GMT}{\footnotesize \par}
-\end{lyxcode}
+\item [{\texttt{xcreate\_movie\_shakemap\_AVS\_DX\_GMT}}] From \texttt{create\_movie\_shakemap\_AVS\_DX\_GMT.f90},
+it outputs data in ASCII, OpenDX, or AVS format (also readable in
+ParaView). Before compiling the code, make sure you have the file
+\texttt{surface\_from\_mesher.h} in the \texttt{OUTPUT\_FILES/} directory.
+This file will be created by the solver run. Then type
+
+\begin{lyxcode}
+{\footnotesize make~xcreate\_movie\_shakemap\_AVS\_DX\_GMT}{\footnotesize \par}
+\end{lyxcode}
+
 and run the executable \texttt{xcreate\_movie\_shakemap\_AVS\_DX\_GMT}
-in the \texttt{bin/} subdirectory. It will create visualization files in your format of choice. The code
-will prompt the user for input parameters.
+in the main directory. It will create visualization files
+in your format of choice. The code will prompt the user for input
+parameters.
 
@@
 \noindent \begin{centering}
-\includegraphics[width=.32\textwidth]{figures/movie_surf_1.jpg}
-\includegraphics[width=.32\textwidth]{figures/movie_surf_2.jpg}
-\includegraphics[width=.32\textwidth]{figures/movie_surf_3.jpg}
+\includegraphics[width=0.32\textwidth]{figures/movie_surf_1.jpg}
+\includegraphics[width=0.32\textwidth]{figures/movie_surf_2.jpg}
+\includegraphics[width=0.32\textwidth]{figures/movie_surf_3.jpg}
 \par\end{centering}
 
-\caption{Visualization using AVS files created by \texttt{xcreate\_movie\_shakemap\_AVS\_DX\_GMT} showing
-movie snapshots of vertical velocity components at different times.}
+\caption{Visualization using AVS files created by \texttt{xcreate\_movie\_shakemap\_AVS\_DX\_GMT}
+showing movie snapshots of vertical velocity components at different
+times.}
 
@@
 \end{description}
-
-The \texttt{SPECFEM3D} code is running in near real-time to produce
-animations of southern California earthquakes via the web; see Southern California ShakeMovie\textregistered \urlwithparentheses{www.shakemovie.caltech.edu}.
-
-\subsection{\label{sub:Movie-Volume}Movie Volume}
-
-When running xspecfem3D with the \texttt{\small MOVIE\_VOLUME} flag
-turned on, the code outputs several files in \texttt{\small LOCAL\_PATH} specified in the main
-\texttt{Par\_file}, e.g. in directory \texttt{in\_out\_files/DATABASES\_MPI}.
-%As the files can be very large, there are several flags in the \texttt{\small Par\_file}
-%that control the region in space and time that is saved. These are:
-%\texttt{\small NTSTEP\_BETWEEN\_FRAMES}
-The output is saved by each processor
-at the time interval specified by \texttt{\small NTSTEP\_BETWEEN\_FRAMES}.
-For all domains, the velocity field is output to files:
-\texttt{\small proc??????\_velocity\_X\_it??????.bin},
-\texttt{\small proc??????\_velocity\_Y\_it??????.bin} and
-\texttt{\small proc??????\_velocity\_Z\_it??????.bin}.
-For elastic domains, the divergence and curl taken from the velocity field, i.e. $\nabla \cdot \bf{v}$
-and $\nabla \times \bf{v}$, get stored as well:
-\texttt{\small proc??????\_div\_it??????.bin},
-\texttt{\small proc??????\_curl\_X\_t??????.bin},
-\texttt{\small proc??????\_curl\_Y\_it??????.bin} and
-\texttt{\small proc??????\_curl\_Z\_it??????.bin}.
-The files denoted
-\texttt{\small proc??????\_div\_}~\\
-\texttt{\small glob\_it??????.bin} and
-\texttt{\small proc??????\_curl\_glob\_it??????.bin}
-are stored on the global points only, all the other arrays
-are stored on all GLL points. Note that the components X/Y/Z can change to E/N/Z according to
-the \texttt{SUPPRESS\_UTM\_PROJECTION} flag (see also Appendix~\ref{cha:Coordinates} and \ref{cha:channel-codes}).
+The \texttt{SPECFEM3D Cartesian} code is running in near real-time
+to produce animations of southern California earthquakes via the web;
+see Southern California ShakeMovie\textregistered{}\urlwithparentheses{www.shakemovie.caltech.edu}.
+
+\subsection{Movie Volume}\label{sub:Movie-Volume}
+
+When running xspecfem3D with the \texttt{\small MOVIE\_VOLUME}{\small{}
+flag turned on, the code outputs several files in }\texttt{\small LOCAL\_PATH}{\small{}
+specified in the main }\texttt{\small Par\_file}{\small , e.g. in
+directory }\texttt{\small OUTPUT\_FILES/DATABASES\_MPI}{\small .
+The output is saved by each processor at the time interval specified
+by }\texttt{\small NTSTEP\_BETWEEN\_FRAMES}{\small . For all domains,
+the velocity field is output to files: }\texttt{\small proc??????\_velocity\_X\_it??????.bin}{\small ,
+}\texttt{\small proc??????\_velocity\_Y\_it??????.bin}{\small{} and
+}\texttt{\small proc??????\_velocity\_Z\_it??????.bin}{\small . For
+elastic domains, the divergence and curl taken from the velocity field,
+i.e. $\nabla\cdot{\bf {v}}$ and $\nabla\times{\bf {v}}$, get stored
+as well: }\texttt{\small proc??????\_div\_it??????.bin}{\small , }\texttt{\small proc??????\_curl\_X\_t??????.bin}{\small ,
+}\texttt{\small proc??????\_curl\_Y\_it??????.bin}{\small{} and }\texttt{\small proc??????\_curl\_Z\_it??????.bin}{\small .
+The files denoted }\texttt{\small proc??????\_div\_}{\small ~}\\
+{\small{} }\texttt{\small glob\_it??????.bin}{\small{} and }\texttt{\small proc??????\_curl\_glob\_it??????.bin}{\small{}
+are stored on the global points only, all the other arrays are stored
+on all GLL points. Note that the components X/Y/Z can change to E/N/Z
+according to the }\texttt{\small SUPPRESS\_UTM\_PROJECTION}{\small{}
+flag (see also Appendix~\ref{cha:Coordinates} and \ref{cha:channel-codes}).}{\small \par}
 
@@
 \noindent \begin{centering}
-\includegraphics[width=.32\textwidth]{figures/movie_volume_1.jpg}
-\includegraphics[width=.32\textwidth]{figures/movie_volume_2.jpg}
-\includegraphics[width=.32\textwidth]{figures/movie_volume_3.jpg}
+\includegraphics[width=0.32\textwidth]{figures/movie_volume_1.jpg} 
+\includegraphics[width=0.32\textwidth]{figures/movie_volume_2.jpg}
+\includegraphics[width=0.32\textwidth]{figures/movie_volume_3.jpg}
 \par\end{centering}
 
-\caption{Paraview visualization using movie volume files (converted by \texttt{xcombine\_vol\_data} and \texttt{mesh2vtu.pl}) and showing
-snapshots of vertical velocity components at different times.}
+\caption{Paraview visualization using movie volume files (converted by \texttt{xcombine\_vol\_data}
+and \texttt{mesh2vtu.pl}) and showing snapshots of vertical velocity
+components at different times.}
 
@@
 To visualize these files, we use an auxilliary program \texttt{combine\_vol\_data.f90}
-to combine the data from all slices into one mesh file. To compile it in the root directory, type:
+to combine the data from all slices into one mesh file. To compile
+it in the root directory, type:
 \begin{lyxcode}
@@
 \end{lyxcode}
-which will create the executable \texttt{xcombine\_vol\_data} in the directory \texttt{bin/}. To output the
-usage of this executable, type `./bin/xcombine\_vol\_data` without arguments.
-As an example, to run the executable you would use
-\begin{lyxcode}
-{\footnotesize cd~bin/~}{\footnotesize \par}
-{\footnotesize ./xcombine\_vol\_data~0~3~velocity\_Z\_it000400~../in\_out\_files/DATABASES\_MPI~../in\_out\_files/OUTPUT\_FILES~0~}{\footnotesize \par}
-\end{lyxcode}
-to create a low-resolution mesh file. The output mesh file will have the name \texttt{velocity\_Z\_it000400.mesh}.
-We next convert the \texttt{.mesh} file into the VTU (Unstructured
-grid file) format which can be viewed in ParaView. For this task, you can use and modify the script \texttt{mesh2vtu.pl}
-located in directory \texttt{utils/Visualization/Paraview/}, for example:
+which will create the executable \texttt{bin/xcombine\_vol\_data}.
+To output the usage of this executable, type
+\begin{lyxcode}
+{\footnotesize ./bin/xcombine\_vol\_data~}{\footnotesize \par}
+\end{lyxcode}
+without arguments.
+
+\medskip
+
+\texttt{xcombine\_vol\_data} will combine the data on the different processors (located in \texttt{OUTPUT\_FILES/DATABASES\_MPI}), for a given quantity and a given iteration, to one file.
+For example, if you want to combine \texttt{velocity\_Z}, for the iteration 400 on 4 processors, i.e.
+if you want to combine these files :
+
+\medskip
+
+\texttt{proc000000\_velocity\_Z\_it000400.bin }
+
+\texttt{proc000001\_velocity\_Z\_it000400.bin }
+
+\texttt{proc000002\_velocity\_Z\_it000400.bin }
+
+\texttt{proc000003\_velocity\_Z\_it000400.bin}
+
+\medskip
+
+You have to go in the directory of the concerned example (where are the directories \texttt{DATA}, \texttt{OUTPUT\_FILES}, etc.).
+Then, you can launch \texttt{xcombine\_vol\_data}, specifying the path where it is located.
+As an example, if the \texttt{DATA} and \texttt{OUTPUT\_FILES} directories of the concerned example are in the root specfem3d directory, \texttt{xcombine\_vol\_data} is in \texttt{./bin}, and you have to type:
+
+\begin{lyxcode}
+~{\footnotesize ./bin/xcombine\_vol\_data~0~3~velocity\_Z\_it000400~./OUTPUT\_FILES/DATABASES\_MPI~./OUTPUT\_FILES~0~}{\footnotesize \par}
+\end{lyxcode}
+
+Here, $0$ is the number of the first processor, $3$ the number of the last one, \texttt{velocity\_Z\_it000400} the name of the files we want to combine without the prefix \texttt{proc000000*\_}, \texttt{./OUTPUT\_FILES/DATABASES\_MPI} the directory where the files are located, \texttt{./OUTPUT\_FILES} the directory where the combined file will be stored, and $0$ is the parameter to create a low-resolution mesh file ($1$ for a high-resolution).
+
+\medskip
+
+At the beginning of the source file \texttt{combine\_vol\_data.f90} (located in \texttt{src/auxiliaries/}), there is a logical constant parameter \texttt{USE\_VTK\_OUTPUT} whose default value is \texttt{.true.}, that which will provide the output combined file directly in the VTK format and you can use Paraview to visualize it.
+
+\medskip
+
+If the value of this parameter is \texttt{.false.}, the output mesh file will have the name \texttt{velocity\_Z\_it000400.mesh}. Then we next have to convert the
+\texttt{.mesh} file into the VTU (Unstructured grid file) format which can be viewed in ParaView. For this task, you can use and modify the
+script \texttt{mesh2vtu.pl} located in directory \texttt{utils/Visualization/Paraview/}, for example:
 \begin{lyxcode}
@@
 \end{lyxcode}
-Notice that this Perl script uses a program \texttt{mesh2vtu} in the
-\texttt{utils/Visualization/Paraview/mesh2vtu} directory, which further uses the VTK \urlwithparentheses{http://www.vtk.org/}
-run-time library for its execution. Therefore, make sure you have
-them properly set in the script according to your system.
-
-\section{\label{sec:Finite-Frequency-Kernels}Finite-Frequency Kernels}
+Notice that this Perl script uses a program \texttt{mesh2vtu} in the \texttt{utils/Visualization/Paraview/mesh2vtu} directory, which further
+uses the VTK \urlwithparentheses{http://www.vtk.org/} run-time library for its execution. Therefore, make sure you have them properly set
+in the script according to your system.
+
+\bigskip
+
+Then, to do a movie with several iterations, you have to repeat this process for each iteration you want to put in your movie.
+
+\section{Finite-Frequency Kernels}\label{sec:Finite-Frequency-Kernels}
 
@@
 
-A Perl script \texttt{slice\_number.pl} located in directory \texttt{utils/Visualization/Paraview/} can
-help to figure out the slice numbers that lie along the great circle
-path. It applies to meshes created with the internal mesher \texttt{xmeshfem3D}.
-
-\begin{enumerate}
-\item On machines where you have access to the script, copy the \texttt{Mesh\_Par\_file}, and
-\texttt{output\_solver} files, and run:
+A Perl script \texttt{slice\_number.pl} located in directory \texttt{utils/Visualization/Paraview/}
+can help to figure out the slice numbers that lie along the great
+circle path. It applies to meshes created with the internal mesher
+\texttt{xmeshfem3D}.
+\begin{enumerate}
+\item On machines where you have access to the script, copy the \texttt{Mesh\_Par\_file},
+and \texttt{output\_solver} files, and run:
 
@@
 \item Notice that this Perl script uses a program \texttt{mesh2vtu} in the
-\texttt{utils/Visualization/Paraview/mesh2vtu} directory, which further uses the VTK \urlwithparentheses{http://www.vtk.org/}
-run-time library for its execution. Therefore, make sure you have
-them properly set in the script according to your system.
+\texttt{utils/Visualization/Paraview/mesh2vtu} directory, which further
+uses the VTK \urlwithparentheses{http://www.vtk.org/} run-time library
+for its execution. Therefore, make sure you have them properly set
+in the script according to your system.
 \end{enumerate}
@@
 In the case of a single source and a single receiver, the simulation
-also generates the file \texttt{sr.vtk} located in the
-\texttt{in\_out\_files/OUTPUT\_FILES/} directory to describe
-the source and receiver locations, which can also be viewed in Paraview
-in the next step.
+also generates the file \texttt{sr.vtk} located in the \texttt{OUTPUT\_FILES/}
+directory to describe the source and receiver locations, which can
+also be viewed in Paraview in the next step.
 
@@
 \textsf{Glyph}, and represent the points by `\textsf{spheres}'.
-\item For more information about ParaView, see the ParaView Users Guide \urlwithparentheses{www.paraview.org/files/v1.6/ParaViewUsersGuide.PDF}.
-\end{enumerate}
-\end{enumerate}
-%
+\item For more information about ParaView, see the ParaView Users Guide
+\urlwithparentheses{www.paraview.org/files/v1.6/ParaViewUsersGuide.PDF}.
+\end{enumerate}
+\end{enumerate}
 \begin{figure}[H]
 \noindent \begin{centering}
-\includegraphics[scale=0.6]{figures/3D-S-Kernel}
+\includegraphics[scale=0.6]{figures/3D-S-Kernel.pdf}
 \par\end{centering}
@@
 \end{figure}
-
-\chapter{\label{cha:Scheduler}Running through a Scheduler}
-
-The code is usually run on large parallel machines, often PC clusters,
-most of which use schedulers, i.e., queuing or batch management systems
-to manage the running of jobs from a large number of users. The following
-considerations need to be taken into account when running on a system
-that uses a scheduler:
-
-\begin{itemize}
-\item The processors/nodes to be used for each run are assigned dynamically
-by the scheduler, based on availability. Therefore, in order for the
-\texttt{xgenerate\_databases} and the \texttt{xspecfem3D} executables (or between successive runs of the solver) to
-have access to the same database files (if they are stored on hard
-drives local to the nodes on which the code is run), they must be
-launched in sequence as a single job.
-\item On some systems, the nodes to which running jobs are assigned are
-not configured for compilation. It may therefore be necessary to pre-compile
-both the \texttt{xgenerate\_databases} and the \texttt{xspecfem3D} executables.
-\item One feature of schedulers/queuing systems is that they allow submission
-of multiple jobs in a ``launch and forget'' mode. In order to take
-advantage of this property, care needs to be taken that output and
-intermediate files from separate jobs do not overwrite each other,
-or otherwise interfere with other running jobs.
-\end{itemize}
-Examples of job scripts can be found in the \texttt{\small utils/Cluster/}
-directory and can straightforwardly be modified and adapted to meet
-more specific running needs.
-
-We describe here in some detail a job submission procedure for the
-Caltech 1024-node cluster, CITerra, under the LSF scheduling system.
-We consider the submission of a regular forward simulation using the internal mesher
-to create mesh partitions. The two
-main scripts are \texttt{\small run\_lsf.bash}, which compiles the
-Fortran code and submits the job to the scheduler, and \texttt{\small go\_mesher\_solver\_lsf\_basin.forward},
-which contains the instructions that make up
-the job itself. These scripts can be found in \texttt{\small utils/Cluster/lsf/} directory
-
-\section{Job submission \texttt{run\_lsf.bash}}
-
-This script first sets the job queue to be `normal'. It then compiles
-the mesher, database generator and solver together, figures out the number of processors
-required for this simulation from the \texttt{in\_data\_files/Par\_file}, and
-submits the LSF job.
-
-\begin{lyxcode}
-\#!/bin/bash
-
-\#~use~the~normal~queue~unless~otherwise~directed~queue=\char`\"{}-q~normal\char`\"{}~
-
-if~{[}~\$\#~-eq~1~];~then
-
-~~~~~~~~echo~\char`\"{}Setting~the~queue~to~\$1\char`\"{}
-
-~~~~~~~~queue=\char`\"{}-q~\$1\char`\"{}~
-
-fi~\\
-~\\
-\#~compile~the~mesher~and~the~solver~
-
-d=`date'~echo~\char`\"{}Starting~compilation~\$d\char`\"{}~
-
-make~clean~
-
-make~xmeshfem3D~
-
-make~xgenerate\_databases~
-
-make~xspecfem3D~
-
-d=`date'~
-
-echo~\char`\"{}Finished~compilation~\$d\char`\"{}~\\
-~\\
-\#~get~total~number~of~nodes~needed~for~solver~
-
-NPROC=`grep~NPROC~in\_data\_files/Par\_file~|~cut~-c~34-~'~\\
-~\\
-
-\#~compute~total~number~of~nodes~needed~for~mesher~
-
-NPROC\_XI=`grep~NPROC\_XI~in\_data\_files/meshfem3D\_files/Mesh\_Par\_file~|~cut~-c~34-~'~
-~\\
-
-NPROC\_ETA=`grep~NPROC\_ETA~in\_data\_files/meshfem3D\_files/Mesh\_Par\_file~|~cut~-c~34-~'~
-~\\
-\#~total~number~of~nodes~is~the~product~of~the~values~read~
-
-numnodes=\$((~\$NPROC\_XI~*~\$NPROC\_ETA~))~\\
-~\\
-
-\#~checks~total~number~of~nodes~
-
-if~{[}~\$numnodes~-neq~\$NPROC~];~then
-
-~~~~~~~~echo~\char`\"{}error number of procs mismatch\char`\"{}
-
-~~~~~~~~exit~
-
-fi~\\
-
-~\\
-echo~\char`\"{}Submitting~job\char`\"{}~
-
-bsub~\$queue~-n~\$numnodes~-W~60~-K~<go\_mesher\_solver\_lsf.forward~
-\end{lyxcode}
-
-\section{Job script \texttt{go\_mesher\_solver\_lsf.forward}}
-
-This script describes the job itself, including setup steps that can
-only be done once the scheduler has assigned a job-ID and a set of
-compute nodes to the job, the \texttt{run\_lsf.bash} commands used
-to run the mesher, database generator and the solver, and calls to scripts that collect
-the output seismograms from the compute nodes and perform clean-up
-operations.
-
-\begin{enumerate}
-\item First the script directs the scheduler to save its own output and
-output from \texttt{stdout} into ~\\
-\texttt{\small in\_out\_files/OUTPUT\_FILES/\%J.o},
-where \texttt{\%J} is short-hand for the job-ID; it also tells the
-scheduler what version of \texttt{mpich} to use (\texttt{mpich\_gm})
-and how to name this job (\texttt{go\_mesher\_solver\_lsf}).
-\item The script then creates a list of the nodes allocated to this job
-by echoing the value of a dynamically set environment variable \texttt{LSB\_MCPU\_HOSTS}
-and parsing the output into a one-column list using the Perl script
-\texttt{utils/Cluster/lsf/remap\_lsf\_machines.pl}. It then creates a set of scratch
-directories on these nodes (\texttt{\small /scratch/}~\\
-\texttt{\small \$USER/DATABASES\_MPI}) to be used as the \texttt{LOCAL\_PATH}
-for temporary storage of the database files. The scratch directories
-are created using \texttt{shmux}, a shell multiplexor that can execute
-the same commands on many hosts in parallel. \texttt{shmux} is available
-from Shmux \urlwithparentheses{web.taranis.org/shmux/}. Make sure that the \texttt{LOCAL\_PATH}
-parameter in \texttt{in\_data\_files/Par\_file} is also set properly.
-\item The next portion of the script launches the mesher, database generator and then the solver
-using \texttt{run\_lsf.bash}.
-\item The final portion of the script collects the seismograms and performs
-clean up on the nodes, using the Perl scripts \texttt{collect\_seismo\_lsf\_multi.pl}
-and \texttt{cleanmulti.pl}.
-\end{enumerate}
-\begin{lyxcode}
-\#!/bin/bash~-v~
-
-\#BSUB~-o~in\_out\_files/OUTPUT\_FILES/\%J.o~
-
-\#BSUB~-a~mpich\_gm~
-
-\#BSUB~-J~go\_mesher\_solver\_lsf~\\
-~\\
-\#~set~up~local~scratch~directories
-
-BASEMPIDIR=/scratch/\$USER/DATABASES\_MPI
-
-mkdir~-p~in\_out\_files/OUTPUT\_FILES~
-
-echo~\char`\"{}\$LSB\_MCPU\_HOSTS\char`\"{}~>~in\_out\_files/OUTPUT\_FILES/lsf\_machines~
-
-echo~\char`\"{}\$LSB\_JOBID\char`\"{}~>~in\_out\_files/OUTPUT\_FILES/jobid
-
-remap\_lsf\_machines.pl~in\_out\_files/OUTPUT\_FILES/lsf\_machines~> in\_out\_files/OUTPUT\_FILES/machines
-
-shmux~-M50~-Sall~-c~\char`\"{}rm~-r~-f~/scratch/\$USER;~\textbackslash{}
-
-~~~~~~mkdir~-p~/scratch/\$USER;~mkdir~-p~\$BASEMPIDIR\char`\"{}~\textbackslash{}
-
-~~~~~~-~<~in\_out\_files/OUTPUT\_FILES/machines~>/dev/null~\\
-~\\
-\#~run~the~specfem~program
-
-current\_pwd=\$PWD \\
-
-cd bin/ \\
-
-run\_lsf.bash~-{}-gm-no-shmem~-{}-gm-copy-env~\$current\_pwd/xmeshfem3D~\\
-
-run\_lsf.bash~-{}-gm-no-shmem~-{}-gm-copy-env~\$current\_pwd/xgenerate\_databases~\\
-
-run\_lsf.bash~-{}-gm-no-shmem~-{}-gm-copy-env~\$current\_pwd/xspecfem3D~\\
-~\\
-\#~collect~seismograms~and~clean~up
-
-cd current\_pwd/~ \\
-
-mkdir~-p~in\_out\_files/SEM
-
-cd~in\_out\_files/SEM/~
-
-collect\_seismo.pl~../OUTPUT\_FILES/lsf\_machines
-
-cleanbase.pl~../OUTPUT\_FILES/machines
-\end{lyxcode}
-
-\chapter{{\normalsize \label{cha:-Changing-the}} Changing the Model}
-
-In this section we explain how to change the velocity model used for your simulations.
-These changes involve contributing specific subroutines
-that replace existing subroutines in the \texttt{SPECFEM3D} package.
-Note that \texttt{SPECFEM3D} can handle Earth models with material properties that vary within each spectral element.
-
-%% magnoni 6/12
-\section{{\normalsize \label{sec:Using-tomographic}}Using external tomographic Earth models}
-
-To implement your own external tomographic model, you must provide your own tomography file
-\texttt{tomography\_model.xyz} and choose between two possible options:
-(1) set in the \texttt{Par\_file} the model parameter \texttt{MODEL = tomo}, or
-(2) for more user control, set in the \texttt{Par\_file} the model parameter \texttt{MODEL = default} and use the following format in the file \texttt{MESH/nummaterial\_velocity\_file} when using CUBIT to construct your mesh (see also section \ref{subsec:Exporting-the-Mesh}):
-\begin{lyxcode}
-\texttt{domain\_ID} \texttt{material\_ID} \texttt{tomography} \texttt{elastic} \texttt{file\_name} 1
-\end{lyxcode}
-where
-\texttt{domain\_ID} is 1 for acoustic or 2 for elastic materials,
-\texttt{material\_ID} a negative, unique identifier (i.e., -1),
-\texttt{tomography} keyword for tomographic material definition,
-\texttt{elastic} keyword for elastic material definition,
-\texttt{file\_name} the name of the tomography file and
-1 a positive unique identifier.
-The name of the tomography file is not used so far, rather change the filename (\texttt{TOMO\_FILENAME}) defined in the file \texttt{model\_tomography.f90} located in the \texttt{src/generate\_databases/} directory.
-
-The external tomographic model is represented by a grid of points with assigned material properties and homogeneous resolution along each spatial direction x, y and z.
-The ASCII file \texttt{tomography\_model.xyz} that describes the tomography should be located in the \texttt{in\_data\_files/} directory.
-The format of the file, as read from \texttt{model\_tomography.f90}, looks like Figure \ref{fig:tomography_file},
-where
-%
-\begin{description}
-\item [{\texttt{ORIG\_X}, \texttt{END\_X}}] are, respectively, the coordinates of the initial and final tomographic grid points along the x direction (in the mesh units, e.g., $m$);
-\item [{\texttt{ORIG\_Y}, \texttt{END\_Y}}] respectively the coordinates of the initial and final tomographic grid points along the y direction (in the mesh units, e.g., $m$);
-\item [{\texttt{ORIG\_Z}, \texttt{END\_Z}}] respectively the coordinates of the initial and final tomographic grid points along the z direction (in the mesh units, e.g., $m$);
-\item [\texttt{SPACING\_X}, \texttt{SPACING\_Y}, \texttt{SPACING\_Z}] the spacing between the tomographic grid points along the x, y and z directions, respectively (in the mesh units, e.g., $m$);
-\item [\texttt{NX}, \texttt{NY}, \texttt{NZ}] the number of grid points along the spatial directions x, y and z, respectively; \texttt{NX} is given by [(\texttt{END\_X} - \texttt{ORIG\_X})/\texttt{SPACING\_X}]+1; \texttt{NY} and \texttt{NZ} are the same as \texttt{NX}, but for the y and z directions, respectively;
-\item [\texttt{VP\_MIN}, \texttt{VP\_MAX}, \texttt{VS\_MIN}, \texttt{VS\_MAX}, \texttt{RHO\_MIN}, \texttt{RHO\_MAX}] the minimum and maximum values of the wave speed \texttt{vp} and \texttt{vs} (in $m \, s^{-1}$) and of the density \texttt{rho} (in $kg \, m^{-3}$); these values could be the actual limits of the tomographic parameters in the grid or the minimum and maximum values to which we force the cut of velocity and density in the model.
-\end{description}
-%
-After the first four lines, in the file \texttt{tomography\_model.xyz} the tomographic grid points are listed with the corresponding values of \texttt{vp}, \texttt{vs} and \texttt{rho}, scanning the grid along the x coordinate (from \texttt{ORIG\_X} to \texttt{END\_X} with step of \texttt{SPACING\_X}) for each given y (from \texttt{ORIG\_Y} to \texttt{END\_Y}, with step of \texttt{SPACING\_Y}) and each given z (from \texttt{ORIG\_Z} to \texttt{END\_Z}, with step of \texttt{SPACING\_Z}).
-%fig
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=1.0\textwidth]{figures/tomo_file.jpg}
-\par\end{centering}
-\caption{Tomography file \texttt{tomography\_model.xyz} that describes an external Earth model and that is read by \texttt{model\_tomography.f90}.
-The coordinates x, y and z of the grid, their limits ORIG\_X, ORIG\_Y, ORIG\_Z, END\_X, END\_Y, END\_Z and the grid spacings SPACING\_X, SPACING\_Y, SPACING\_Z should be in the units of the constructed mesh (e.g., UTM coordinates in meters).}
-\label{fig:tomography_file}
-\end{figure}
-
-The user can implement his own interpolation algorithm for the tomography model by changing the routine \texttt{model\_tomography.f90} located in the \texttt{src/generate\_databases/} directory.
-Moreover, for models that involve both fully defined materials and a tomography description, the \texttt{nummaterial\_velocity\_file} has multiple lines each with the corresponding suitable format described above.
-
-\section{{\normalsize \label{sec:Anelastic-Models}}External (an)elastic Models}
-
-To use your own external model, you can set in the \texttt{Par\_file} the model parameter \texttt{MODEL = external}.
-Three-dimensional acoustic and/or (an)elastic (attenuation) models may then be superimposed
-onto the mesh based upon your subroutine in
-\texttt{model\_external\_values.f90} located in directory \texttt{src/generate\_databases}.
-The call to this routine would be as follows
-
-\begin{lyxcode}
-call~model\_external\_values(xmesh,~ymesh,~zmesh, ~\& \\
-~~~~~~~~~~rho,~vp,~vs,~qmu\_atten,~iflag\_aniso,~idomain\_id)
-\end{lyxcode}
-%
-Input to this routine consists of:
-\begin{description}
-\item [{\texttt{xmesh,ymesh,zmesh}}] location of mesh point
-\end{description}
-%
-Output to this routine consists of:
-\begin{description}
-\item [{\texttt{rho,vp,vs}}] isotropic model parameters for density $\rho$ ($kg/m^3$), $v_p$ ($m/s$) and $v_s$ ($m/s$)
-\item [{\texttt{qmu\_atten}}] Shear wave quality factor: $0<Q_{\mu}<9000$
-\item [{\texttt{iflag\_aniso}}] anisotropic model flag, $0$ indicating no anisotropy or $1$ using anisotropic model parameters as defined in routine file \texttt{model\_aniso.f90}
-\item [{\texttt{idomain\_id}}] domain identifier, $1$ for acoustic, $2$ for elastic, $3$ for poroelastic materials.
-\end{description}
-
-Note that the resolution and maximum value of anelastic models are
-truncated. This speeds the construction of the standard linear solids
-during the meshing stage. To change the resolution, currently at one
-significant figure following the decimal, or the maximum value (9000),
-consult \texttt{shared/constants.h}.
-
-\section{{\normalsize \label{sec:Anisotropic-Models}}Anisotropic Models}
-
-To use your anisotropic models, you can either set in the \texttt{Par\_file} the model parameter \texttt{ANISOTROPY} to \texttt{.true.} or set the model parameter \texttt{MODEL = aniso}.
-Three-dimensional anisotropic models may be superimposed on
-the mesh based upon the subroutine in file \texttt{model\_aniso.f90}
-located in directory \texttt{src/generate\_databases}.
-The call to this subroutine is of the form
-
-\begin{lyxcode}
-call~model\_aniso(iflag\_aniso,~rho,~vp,~vs,~\& \\
-~~~~~~~~~~c11,c12,c13,c14,c15,c16,c22,c23,c24,c25,c26,~\& \\
-~~~~~~~~~~c33,c34,c35,c36,c44,c45,c46,c55,c56,c66)~
-\end{lyxcode}
-%
-Input to this routine consists of:
-\begin{description}
-\item [{\texttt{iflag\_aniso}}] flag indicating the type of the anisotropic model, i.e. $0$ for no superimposed anisotropy, $1$ or $2$ for generic pre-defined anisotropic models.
-\item [{\texttt{rho,vp,vs}}] reference isotropic model parameters for density $\rho$, $v_p$ and $v_s$.
-\end{description}
-%
-Output from the routine consists of the following non-dimensional
-model parameters:
-\begin{description}
-\item [{\texttt{c11},}] \textbf{$\cdots$,} \texttt{\textbf{c66}} 21~dimensionalized
-anisotropic elastic parameters.
-\end{description}
-
-You can replace the \texttt{model\_aniso.f90} file by
-your own version \textit{provided you do not change the call structure
-of the routine}, i.e., the new routine should take exactly the same
-input and produce the required relative output.
-
-\chapter{Post-Processing Scripts}
-
-Several post-processing scripts/programs are provided in the \texttt{utils/}
-directory, and most of them need to be adjusted when used on different
-systems, for example, the path of the executable programs. Here we
-only list a few of the available scripts and provide a brief description, and
-you can either refer to the related sections for detailed usage or,
-in a lot of cases, type the script/program name without arguments
-for its usage.
-
-\section{Process Data and Synthetics\label{sec:Process-data-and-syn}}
-
-In many cases, the SEM synthetics are calculated and compared to data
-seismograms recorded at seismic stations. Since the SEM synthetics
-are accurate for a certain frequency range, both the original data
-and the synthetics need to be processed before a comparison can be
-made.
-
- For such comparisons, the following steps are recommended:
-
-\begin{enumerate}
-\item Make sure that both synthetic and observed seismograms have the correct station/event and timing information.
-\item Convolve synthetic seismograms with a source time function with the half duration specified in the \texttt{CMTSOLUTION} file, provided, as recommended, you used a zero half duration in the SEM simulations.
-\item Resample both observed and synthetic seismograms to a common sampling rate.
-\item Cut the records using the same window.
-\item Remove the trend and mean from the records and taper them.
-\item Remove the instrument response from the observed seismograms (recommended) or convolve the synthetic seismograms with the instrument response.
-\item Make sure that you apply the same filters to both observed and synthetic seismograms. Preferably, avoid filtering your records more than once.
-\item Now, you are ready to compare your synthetic and observed seismograms.
-\end{enumerate}
-
-\noindent
-We generally use the following scripts provided in the \texttt{utils/seis\_process/} directory:
-
-\subsection{Data processing script \texttt{process\_data.pl}}
-
-This script cuts a given portion of the original data, filters it,
-transfers the data into a displacement record, and picks the first
-P and S arrivals. For more functionality, type `\texttt{process\_data.pl}'
-without any argument. An example of the usage of the script:
-
-\begin{lyxcode}
-{\footnotesize process\_data.pl~-m~CMTSOLUTION~-s~1.0~-l~0/4000~-i~-f~-t~40/500~-p~-x~bp~DATA/1999.330{*}.BH?.SAC}
-{\footnotesize \par}
-\end{lyxcode}
-
-\noindent which has resampled the SAC files to a sampling rate of 1 seconds, cut
-them between 0 and 4000 seconds, transfered them into displacement
-records and filtered them between 40 and 500 seconds,
-picked the first P and S arrivals, and added suffix `\texttt{bp}' to the file names.
-
-Note that all of the scripts in this section actually use the SAC
-and/or IASP91 to do the core operations; therefore make sure that
-the SAC and IASP91 packages are installed properly on your system,
-and that all the environment variables are set properly before running
-these scripts.
-
-\subsection{Synthetics processing script \texttt{process\_syn.pl}}
-
-This script converts the synthetic output from the SEM code from ASCII
-to SAC format, and performs similar operations as `\texttt{process\_data.pl}'.
-An example of the usage of the script:
-\begin{lyxcode}
-{\footnotesize process\_syn.pl~-m~CMTSOLUTION~-h~-a~STATIONS~-s~1.0~-l~0/4000~-f~-t~40/500~-p~-x~bp~SEM/{*}.BX?.semd}
-{\footnotesize \par}
-\end{lyxcode}
-which will convolve the synthetics with a triangular source-time function
-from the \texttt{CMTSOLUTION} file, convert the synthetics into SAC
-format, add event and station information into the SAC headers, resample the SAC files with a sampling rate of 1 seconds, cut them between 0 and 4000 seconds, filter them between 40 and
-500 seconds with the same filter used for the observed data, pick the first P and S arrivals, and add the suffix `\texttt{bp}'
-to the file names.
-
-More options are available for this script, such as adding time shift
-to the origin time of the synthetics, convolving the synthetics with
-a triangular source time function with a given half duration, etc.
-Type \texttt{process\_syn.pl} without any argument for a detailed
-usage.
-
-In order to convert between SAC format and ASCII files, useful scripts
-are provided in the subdirectories ~\\
-\texttt{utils/sac2000\_alpha\_convert/} and \texttt{utils/seis\_process/asc2sac/}.
-
-\subsection{Script \texttt{rotate.pl}}
-
-The original data and synthetics have three components: vertical (BHZ resp. BXZ),
-north (BHN resp. BXN) and east (BHE resp. BXE). However, for most seismology applications,
-transverse and radial components are also desirable. Therefore, we
-need to rotate the horizontal components of both the data and the
-synthetics to the transverse and radial direction, and \texttt{\small rotate.pl}
-can be used to accomplish this:
-
-\begin{lyxcode}
-rotate.pl~-l~0~-L~180~-d~DATA/{*}.BHE.SAC.bp~
-
-rotate.pl~-l~0~-L~180~SEM/{*}.BXE.semd.sac.bp~
-\end{lyxcode}
-where the first command performs rotation on the SAC data obtained
-through Seismogram Transfer Program (STP) \urlwithparentheses{http://www.data.scec.org/STP/stp.html},
-while the second command rotates the processed SEM synthetics.
-
-\section{Collect Synthetic Seismograms}
-
-The forward and adjoint simulations generate synthetic seismograms
-in the \texttt{in\_out\_files/OUTPUT\_FILES/} directory by default. For the forward simulation, the files
-are named like \texttt{STA.NT.BX?.semd} for two-column time series, or
-\texttt{STA.NT.BX?.semd.sac} for ASCII SAC format, where STA and NT
-are the station name and network code, and \texttt{BX?} stands for
-the component name. Please see the Appendix~\ref{cha:Coordinates} and \ref{cha:channel-codes}
-for further details.
-
-The adjont simulations generate synthetic seismograms
-with the name \texttt{S?????.NT.S??.sem} (refer to Section \ref{sec:Adjoint-simulation-sources}
-for details). The kernel simulations output the back-reconstructed
-synthetic seismogram in the name \texttt{STA.NT.BX?.semd}, mainly
-for the purpose of checking the accuracy of the reconstruction. Refer
-to Section \ref{sec:Adjoint-simulation-finite} for further details.
-
-You do have further options to change this default output behavior, given in the main constants file
-\texttt{constants.h} located in \texttt{src/shared/} directory:
-\begin{description}
-\item [{\texttt{SEISMOGRAMS\_BINARY}}] set to \texttt{.true.} to have seismograms written out in binary format.
-\item [{\texttt{WRITE\_SEISMOGRAMS\_BY\_MASTER}}] Set to \texttt{.true.} to have only the master process writing out seismograms.
-This can be useful on a cluster, where only the master process node has access to the output directory.
-\item [{\texttt{USE\_OUTPUT\_FILES\_PATH}}] Set to \texttt{.false.} to have the seismograms output to \texttt{LOCAL\_PATH} directory specified in the main parameter file \texttt{in\_data\_files/Par\_file}.
-In this case, you could collect the synthetics onto the frontend using the \texttt{collect\_seismo\_lsf\_multi.pl} script
-located in the \texttt{utils/Cluster/lsf/} directory. The usage of the script would be e.g.:
-
-\begin{lyxcode}
-collect\_seismo.pl~machines~in\_data\_files/Par\_file
-\end{lyxcode}
-
-\end{description}
-where  \texttt{machines} is a file containing the node names and  \texttt{in\_data\_files/Par\_file} the parameter file used to extract the \texttt{LOCAL\_PATH} directory used for the simulation.
-
-\section{Clean Local Database}
-
-After all the simulations are done, the seismograms are collected,
-and the useful database files are copied to the frontend, you may
-need to clean the local scratch disk for the next simulation. This
-is especially important in the case of kernel simulation,
-where very large files are generated for the absorbing boundaries
-to help with the reconstruction of the regular forward wavefield.
-A sample script is provided in \texttt{utils/}:
-
-\begin{lyxcode}
-cleanbase.pl~machines
-\end{lyxcode}
-where  \texttt{machines} is a file containing the node names.
-
-\section{Plot Movie Snapshots and Synthetic Shakemaps}
-
-\subsection{Script \texttt{movie2gif.gmt.pl}}
-
-With the movie data saved in \texttt{in\_out\_files/OUTPUT\_FILES/} at the end of
-a movie simulation (\texttt{\small MOVIE\_SURFACE=.true.}), you can
-run the \texttt{`create\_movie\_shakemap\_AVS\_DX\_GMT}' code to convert these binary
-movie data into GMT xyz files for futher processing. A sample script
-\texttt{movie2gif.gmt.pl} is provided to do this conversion, and then
-plot the movie snapshots in GMT, for example:
-
-\begin{lyxcode}
-movie2gif.gmt.pl~-m~CMTSOLUTION~-g~-f~1/40~-n~-2~-p~
-\end{lyxcode}
-which for the first through the 40th movie frame, converts the \texttt{moviedata}
-files into GMT xyz files, interpolates them using the 'nearneighbor'
-command in GMT, and plots them on a 2D topography map. Note that `\texttt{-2}'
-and `\texttt{-p}' are both optional.
-
-\subsection{Script \texttt{plot\_shakemap.gmt.pl}}
-
-With the shakemap data saved in \texttt{in\_out\_files/OUTPUT\_FILES/} at the end
-of a shakemap simulation ~\\
-(\texttt{CREATE\_SHAKEMAP=.true.}), you can
-also run \texttt{`create\_movie\_shakemap\_AVS\_DX\_GMT}' code to convert the binary
-shakemap data into GMT xyz files. A sample script \texttt{plot\_shakemap.gmt.pl}
-is provided to do this conversion, and then plot the shakemaps in
-GMT, for example:
-
-\begin{lyxcode}
-plot\_shakemap.gmt.pl~data\_dir~type(1,2,3)~CMTSOLUTION~
-\end{lyxcode}
-where \texttt{type=1} for a displacement shakemap, \texttt{2} for
-velocity, and \texttt{3} for acceleration.
-
-\section{Map Local Database}
-
-A sample program \texttt{remap\_database} is provided to map the local
-database from a set of machines to another set of machines. This is
-especially useful when you want to run mesher and solver, or different
-types of solvers separately through a scheduler (refer to Chapter~\ref{cha:Scheduler}).
-\begin{lyxcode}
-run\_lsf.bash~-{}-gm-no-shmem~-{}-gm-copy-env~remap\_database~old\_machines~150
-\end{lyxcode}
-where \texttt{old\_machines} is the LSF machine file used in the previous
-simulation, and \texttt{150} is the number of processors in total.
-
-\chapter{\label{cha:tomo}Tomographic inversion using sensitivity kernels}
-
-One of the fundamental reasons for computing sensitivity kernels (Section~\ref{sec:Adjoint-simulation-finite}) is to use them within a tomographic inversion.
-In other words, use recorded seismograms, make measurements with synthetic seismograms, and use the misfit between the two to iteratively improve the model described by (at least) $V_{\rm p}$, $V_{\rm s}$, and $\rho$ as a function of space.
-
-Whatever misfit function you use for the tomographic inversion (several examples in \cite{TrTaLi05} and \cite{TrKoLi08}), you will weight the sensitivity kernels with measurements.
-Furthermore, you will use as many measurements (stations, components, time windows) as possible per event; hence, we call these composite kernels ``event kernels,'' which are volumetric fields representing the gradient of the misfit function with respect to one of the variables (\eg $V_{\rm s}$).
-The basic features of an adjoint-based tomographic inversion were illustrated in \citet{TaLiTr07} and \citet{TrKoLi08} using a conjugate-gradient algorithm; there are dozens of versions of gradient-based inversion algorithms that could alternatively be used.
-The tomographic inversion of \citet{TaLiMaTr09,TaLiMaTr2010} used SPECFEM3D as well as several additional components which are also stored on the CIG svn server, described next.
-
-The directory containing utilities for tomographic inversion using SPECFEM3D (or other packages that evaluate misfit functions and gradients) is here on the CIG svn server:
-%
-\begin{verbatim}
-/cig/seismo/3D/ADJOINT_TOMO/
-   flexwin/     -- FLEXWIN algorithm for automated picking of time windows
-   measure_adj/ -- reads FLEXWIN output file and makes measurements,
-                   with the option for computing adjoint sources
-   iterate_adj/ -- various tools for iterative inversion
-                   (requires pre-computed "event kernels")
-\end{verbatim}
-%
-This directory also contains a brief \verb+README+ file indicating the role of the three subdirectories, \verb+flexwin+ \citep{Maggi2009}, \verb+measure_adj+, and \verb+iterate_adj+.
-The components for making the model update are there; however, there are no explicit rules for computing the model update, just as with any optimization problem.
-There are options for computing a conjugate gradient step, as well as a source subspace projection step.
-The workflow could use substantial ``stitching together,'' and we welcome suggestions; undoubtedly, the tomographic inversion procedure will improve with time.
-{\bf This is a work in progress.}
-
-The best single file to read is probably: \verb+ADJOINT_TOMO/iterate_adj/cluster/README+.
-
-%============================================================
-
-\chapter*{\label{cha:Bug-Reports-and}Bug Reports and Suggestions for Improvements}
-
-To report bugs or suggest improvements to the code, please send an
-e-mail to the CIG Computational Seismology Mailing List \urlwithparentheses{cig-seismo@geodynamics.org}
-or Jeroen Tromp \urlwithparentheses{jtromp-AT-princeton.edu}, and/or use our online
-bug tracking system Roundup \urlwithparentheses{www.geodynamics.org/roundup}.
-
-\addcontentsline{toc}{chapter}{Notes \& Acknowledgments}
-\chapter*{Notes \& Acknowledgments}
-
-In order to keep the software package thread-safe in case a multithreaded
-implementation of MPI is used, developers should not add modules or
-common blocks to the source code but rather use regular subroutine
-arguments (which can be grouped in ``derived types'' if needed for
-clarity).
-
-The Gauss-Lobatto-Legendre subroutines in \texttt{gll\_library.f90}
-are based in part on software libraries from the Massachusetts Institute
-of Technology, Department of Mechanical Engineering (Cambridge, Massachusetts, USA).
-The non-structured global numbering software was provided by Paul
-F. Fischer (Brown University, Providence, Rhode Island, USA, now at Argonne National Laboratory, USA).
-
-OpenDX \urlwithparentheses{http://www.opendx.org} is open-source based on IBM Data
-Explorer, AVS \urlwithparentheses{http://www.avs.com} is a trademark of Advanced
-Visualization Systems, and ParaView \urlwithparentheses{http://www.paraview.org}
-is an open-source visualization platform.
-
-Please e-mail your feedback, questions, comments, and suggestions
-to Jeroen Tromp \urlwithparentheses{jtromp-AT-princeton.edu} or to the CIG Computational Seismology Mailing List \urlwithparentheses{cig-seismo@geodynamics.org}.
-
-\addcontentsline{toc}{chapter}{Copyright}
-\chapter*{Copyright}
-
-Main authors: Dimitri Komatitsch and Jeroen Tromp
-
-Princeton University, USA, and CNRS / INRIA / University of Pau, France
-
-$\copyright$ Princeton University and CNRS / INRIA / University of Pau, July 2012
-
-This program is free software; you can redistribute it and/or modify
-it under the terms of the GNU General Public License as published
-by the Free Software Foundation (see Appendix \ref{cha:License}).\\
-
-\textbf{\underline{Evolution of the code:}}\\
-
- MPI v 2.1, July 2012:
- Max Rietmann, Peter Messmer, Daniel Peter, Dimitri Komatitsch, Joseph Charles, Zhinan Xie: support for CUDA GPUs,
-better CFL stability for the Stacey absorbing conditions. \\
-
- MPI v. 2.0, November 2010:
- Dimitri Komatitsch, Nicolas Le Goff, Roland Martin and Pieyre Le Loher, University of Pau, France,
- Jeroen Tromp and the Princeton group of developers, Princeton University, USA,
- and Emanuele Casarotti, INGV Roma, Italy:
-  support for CUBIT meshes decomposed by SCOTCH, METIS or ZOLTAN;
-  much faster solver using Michel Deville's inlined matrix products.\\
-
- MPI v. 1.4 Dimitri Komatitsch, University of Pau, Qinya Liu and others, Caltech, September 2006:
-  better adjoint and kernel calculations, faster and better I/Os
-  on very large systems, many small improvements and bug fixes.\\
-
- MPI v. 1.3 Dimitri Komatitsch, University of Pau, and Qinya Liu, Caltech, July 2005:
-  serial version, regular mesh, adjoint and kernel calculations, ParaView support.\\
-
- MPI v. 1.2 Min Chen and Dimitri Komatitsch, Caltech, July 2004:
-  full anisotropy, volume movie.\\
-
- MPI v. 1.1 Dimitri Komatitsch, Caltech, October 2002: Zhu's Moho map, scaling
-  of $V_s$ with depth, Hauksson's regional model, attenuation, oceans, movies.\\
-
- MPI v. 1.0 Dimitri Komatitsch, Caltech, USA, May 2002: first MPI version based on global code.\\
-
- Dimitri Komatitsch, IPG Paris, France, December 1996: first 3-D solver for the CM-5 Connection Machine,
-    parallelized on 128 processors using Connection Machine Fortran.\\
-
-\addcontentsline{toc}{chapter}{References}
-\bibliography{bibliography}
-
-\appendix
-
-\chapter{\label{cha:Coordinates}Reference Frame Convention}
-
-The code uses the following convention for the Cartesian reference
-frame:
-
-\begin{itemize}
-\item the $x$ axis points East
-\item the $y$ axis points North
-\item the $z$ axis points up
-\end{itemize}
-Note that this convention is different from both the \citet{AkRi80}
-convention and the Harvard Centroid-Moment Tensor (CMT) convention.
-The Aki \& Richards convention is
-
-\begin{itemize}
-\item the $x$ axis points North
-\item the $y$ axis points East
-\item the $z$ axis points down
-\end{itemize}
-and the Harvard CMT convention is
-
-\begin{itemize}
-\item the $x$ axis points South
-\item the $y$ axis points East
-\item the $z$ axis points up
-\end{itemize}
-
-\subsection*{Source and receiver locations}
-
-The SPECFEM3D software code internally uses Cartesian coordinates.
-The given locations for sources and receiver locations thus may get
-converted.  Sources and receiver locations are read in from the
-\texttt{CMTSOLUTION} (or \texttt{FORCESOLUTION}) and \texttt{STATIONS}
-files.  Note that e.g. the \texttt{CMTSOLUTION} and
-\texttt{FORCESOLUTION} files denotes the location by
-"longitude/latitude/depth". We read in longitude as $x$ coordinate,
-latitude as $y$ coordinate.
-
-In case the flag \texttt{SUPPRESS\_UTM\_PROJECTION} is set to \texttt{.false.}
-in the main parameter file (see Chapter~\ref{cha:Creating-Distributed-Databases}),
-the $x$/$y$ coordinates have to be given in degrees and are converted to Cartesian coordinates
-using a UTM conversion with the specified UTM zone.
-
-The value for depth (given in $km$ in \texttt{CMTSOLUTION} and
-\texttt{FORCESOLUTION}) or burial depth (given in $m$ in
-\texttt{STATIONS}) is evaluated with respect to the surface of the
-mesh at the specified $x$/$y$ location to find the corresponding $z$
-coordinate.  It is possible to use this depth value directly as $z$
-coordinate by changing the flag \texttt{USE\_SOURCES\_RECVS\_Z} to
-\texttt{.true.} in the file \texttt{constants.h} located in the
-\texttt{src/shared/} subdirectory.
-
-\subsection*{Seismogram outputs}
-
-The seismogram output directions are given in Cartesian $x$/$y$/$z$ directions and not rotated any further.
-Changing flags in \texttt{constants.h} in the \texttt{src/shared/} subdirectory only rotates the seismogram outputs
-if receivers are forced to be located at the surface (\texttt{RECVS\_CAN\_BE\_BURIED\_EXT\_MESH} set to \texttt{.false.}) and the normal to the surface at the receiver location should be used (\texttt{EXT\_MESH\_RECV\_NORMAL} set to
-\texttt{.true.}) as vertical.
-In this case, the outputs are rotated to have the vertical component normal to the surface of the mesh, $x$ and $y$ directions are somewhat arbitrary orthogonal directions along the surface.
-
-For the labeling of the seismogram channels, see Appendix~\ref{cha:channel-codes}. Additionally, we add labels to the synthetics using the following convention:
-For a receiver station located in an
-\begin{description}
-\item[elastic domain:] ~\\
-\begin{itemize}
-\item \texttt{semd} for the displacement vector
-\item \texttt{semv} for the velocity vector
-\item \texttt{sema} for the acceleration vector
-\end{itemize}
-\item[acoustic domain:] ~\\
-(please note that receiver stations in acoustic domains must be buried. This is due to the free surface condition which enforces a zero pressure at the surface.)
-\begin{itemize}
-\item \texttt{semd} for the displacement vector
-\item \texttt{semv} for the velocity vector
-\item \texttt{sema} for pressure which will be stored in the vertical component \texttt{Z}.
-Note that pressure in the acoustic media is isotropic and thus the pressure record would be the same in the other two component directions.
-We therefore use the other two seismogram components to store the acoustic potential in component \texttt{X} (or \texttt{N}) and the first time derivative of the acoustic potential in component \texttt{Y} (or \texttt{E}).
-\end{itemize}
-\end{description}
-The seismograms are by default written out in ASCII-format to the \texttt{in\_out\_files/OUTPUT\_FILES/} subdirectory by each parallel process.
-You can change this behavior by changing the following flags in the \texttt{constants.h} file located in the \texttt{src/shared/} subdirectory:
-\begin{description}
-\item [{\texttt{SEISMOGRAMS\_BINARY}}] set to \texttt{.true.} to have seismograms written out in binary format.
-\item [{\texttt{WRITE\_SEISMOGRAMS\_BY\_MASTER}}] Set to \texttt{.true.} to have only the master process writing out seismograms.
-This can be useful on a cluster, where only the master process node has access to the output directory.
-\item [{\texttt{USE\_OUTPUT\_FILES\_PATH}}] Set to \texttt{.false.} to have the seismograms output to \texttt{LOCAL\_PATH} directory specified in the main parameter file \texttt{in\_data\_files/Par\_file}.
-\end{description}
-
-\chapter{\label{cha:channel-codes}Channel Codes of Seismograms}
-
-Seismic networks, such as the Global Seismographic Network (GSN), generally involve various types of instruments with different bandwidths, sampling properties and component configurations.
-There are standards to name channel codes depending on instrument properties.
-IRIS  \texttt{(www.iris.edu)} uses SEED format for channel codes, which are represented by three letters, such as \texttt{LHN}, \texttt{BHZ}, etc.
-In older versions of the SPECFEM3D package, a common format was used for the channel codes of all seismograms, which was \texttt{BHE/BHN/BHZ} for three components.
-To avoid confusion when comparison are made to observed data, we are now using the FDSN convention (\url{http://www.fdsn.org/}) for SEM seismograms.
-In the following, we give a brief explanation of the FDSN convention used by IRIS, and how it is adopted in SEM seismograms.
-Please visit \texttt{www.iris.edu/manuals/SEED\_appA.htm} for further information.\\
-
-\noindent \textbf{\texttt{Band code:}} The first letter in the channel code denotes the band code of seismograms, which depends on the response band and the sampling rate of instruments.
-The list of band codes used by IRIS is shown in Figure \ref{fig:IRIS_band_codes}.
-The sampling rate of SEM synthetics is controlled by the resolution of simulations rather than instrument properties.
-However, for consistency, we follow the FDSN convention for SEM seismograms governed by their sampling rate.
-For SEM synthetics, we consider band codes for which $dt \leq 1$ s.
-IRIS also considers the response band of instruments.
-For instance, short-period and broad-band seismograms with the same sampling rate correspond to different band codes, such as S and B, respectively.
-In such cases, we consider SEM seismograms as broad band, ignoring the corner period ($\geq 10$ s) of the response band of instruments (note that at these resolutions, the minimum period in the SEM synthetics will be less than $10$ s).
-Accordingly, when you run a simulation the band code will be chosen depending on the resolution of the synthetics, and channel codes of SEM seismograms will start with either \texttt{L}, \texttt{M}, \texttt{B}, \texttt{H}, \texttt{C} or \texttt{F}, shown by red color in the figure.\\
-
-\begin{figure}[ht]
-\noindent \begin{centering}
-\includegraphics[scale=0.6]{figures/IRIS_band_codes.pdf}
-\par\end{centering}
-
-\caption{The band code convention is based on the sampling rate and the response band of instruments.
-Please visit \texttt{www.iris.edu/manuals/SEED\_appA.htm} for further information.
-Grey rows show the relative band-code range in SPECFEM3D, and the band codes used to name SEM seismograms are denoted in red.}
-
-\label{fig:IRIS_band_codes}
-\end{figure}
-
-\noindent \textbf{\texttt{Instrument code:}} The second letter in the channel code corresponds to instrument codes, which specify the family to which the sensor belongs.
-For instance, \texttt{H} and \texttt{L} are used for high-gain and low-gain seismometers, respectively.
-The instrument code of SEM seismograms will always be \texttt{X}, as assigned by FDSN for synthetic seismograms.
-\\
-
-\noindent \textbf{\texttt{Orientation code:}} The third letter in channel codes is an orientation code, which generally describes the physical configuration of the components of instrument packages.
-SPECFEM3D uses the traditional orientation code  \texttt{E/N/Z} (East-West, North-South, Vertical) for three components when a UTM projection is used.
-If the UTM conversion is suppressed, i.e.
-the flag \texttt{SUPPRESS\_UTM\_PROJECTION} is set to \texttt{.true.}, then the three components are labelled \texttt{X/Y/Z} according to the Cartesian reference frame. \\
-
-\noindent \textbf{\texttt{EXAMPLE:}} The sampling rate is given by \texttt{DT} in the main parameter file \texttt{Par\_file} located in the \texttt{in\_data\_files/} subdirectory.
-Depending on the resolution of your simulations, if you choose a sampling rate greater than $0.01$ s and less than $1$ s, a seismogram recording displacements on the vertical component of a station \texttt{ASBS} (network \texttt{AZ}) will be named \texttt{ASBS.AZ.MXZ.semd.sac}, whereas it will be \texttt{ASBS.AZ.BXZ.semd.sac}, if the sampling rate is greater than 0.0125 and less equal to 0.1 s.
-
-\chapter{\label{cha:Troubleshooting}Troubleshooting}
-
-\section*{FAQ}
-
-\begin{description}
-\item [configuration fails:]
-
-   Examine the log file 'config.log'. It contains detailed informations.
-   In many cases, the path's to these specific compiler commands F90,
-   CC and MPIF90 won't be correct if `./configure` fails.
-
-   Please make sure that you have a working installation of a Fortran compiler,
-   a C compiler and an MPI implementation. You should be able to compile this
-   little program code:
-
-\begin{lyxcode}
-
-      program main ~\\
-        include 'mpif.h' ~\\
-        integer, parameter :: CUSTOM\_MPI\_TYPE = MPI\_REAL ~\\
-        integer ier ~\\
-        call MPI\_INIT(ier) ~\\
-        call MPI\_BARRIER(MPI\_COMM\_WORLD,ier) ~\\
-        call MPI\_FINALIZE(ier) ~\\
-      end
-
-\end{lyxcode}
-
-\item [compilation fails stating:] ~\\
-\begin{lyxcode}
-    ...  ~\\
-    obj/program\_generate\_databases.o: In function `MAIN\_\_':  ~\\
-    program\_generate\_databases.f90:(.text+0x14): undefined reference to `\_gfortran\_set\_std'  ~\\
-    ...  ~\\
-\end{lyxcode}
-
-  Make sure you're pointing to the right 'mpif90' wrapper command.
-
-  Normally, this message will appear when you're mixing two different Fortran
-  compilers. That is, using e.g. gfortran to compile non-MPI files
-  and mpif90, wrapper provided for e.g. ifort, to compile MPI-files.
-
-  fix: e.g. specify > ./configure FC=gfortran MPIF90=/usr/local/openmpi-gfortran/bin/mpif90
-
-\item [after executing \texttt{xmeshfem3D} I've got elements with skewness of 81\% percent, what does this mean:]
-Look at the skewness table printed in the \texttt{output\_mesher.txt} file after executing \texttt{xmeshfem3D} for the
-example given in \texttt{examples/meshfem3D\_examples/simple\_model/}:
-\begin{lyxcode}
-...~\\
-histogram of skewness (0. good - 1. bad): ~\\
- 0.0000000E+00  -   5.0000001E-02       27648     81.81818      \% ~\\
- 5.0000001E-02  -   0.1000000               0    0.0000000E+00  \% ~\\
-... ~\\
-\end{lyxcode}
-The first line means that you have 27,648 elements with a skewness value between 0 and 0.05 (which means the element is basically not skewed, just plain regular hexahedral element).
-The total number of elements you have in this mesh is (see in the \texttt{output\_mesher.txt} file a bit further down):
-\begin{lyxcode}
-...~\\
-total number of elements in entire mesh:        33792 ~\\
-... ~\\
-\end{lyxcode}
-which gives you that:  27,648 / 33,792 $\sim$ 81.8 \% of all elements are not skewed, i.e. regular elements. a fantastic value :)
-
-The histogram lists for this mesh also some stronger skewed elements, for example the worst ones belong to:
-\begin{lyxcode}
-...~\\
-0.6000000      -   0.6500000            2048     6.060606      \% ~\\
-... ~\\
-\end{lyxcode}
-
-about 6 \% of all elements have distortions with a skewness value between 0.6 and 0.65.
-The skewness values give you a hint of how good your mesh is.
-In an ideal world, you would want to have no distortions, just like the 81\% from above.
-Those elements give you the best approximate values by the GLL quadrature used in the spectral-element method.
-However, having weakly distorted elements is still fine and the solutions are still accurate enough.
-So empirically, values up to around 0.7 are tolerable, above that you should consider remeshing...
-
-To give you an idea why some of the elements are distorted, see the following figure \ref{fig:mesh.vp} of the mesh you obtain in the example \texttt{examples/meshfem3D\_examples/simple\_model/}.
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=.9\textwidth]{figures/mesh_vp.jpg}
-\par\end{centering}
-\caption{Paraview visualization using the mesh vtk-files for the example given in \texttt{examples/meshfem3D\_examples/simple\_model/}.}
-\label{fig:mesh.vp}
-\end{figure}
-You will see that the mesh contains a doubling layer, where we stitch elements together such that the size of two elements will transition to the size of one element (very useful to keep the ratio of wavespeed / element\_size about constant).
-Those elements in this doubling layer have higher skewness values and make up those 6 \% in the histogram.
-
-\item [the code gives following error message "need at least one receiver":]
-This means that no stations given in the input file \texttt{in\_data\_files/STATIONS} could be located within the dimensions of the mesh.
-This can happen for example when the mesh was created with the in-house mesher \texttt{xmeshfem3D} while using the Universal Transverse Mercator (UTM) projection but the simulation with \texttt{xspecfem3D} was suppressing this projection from latitude/longitude to x/y/z coordinates.
-
-In such cases, try to change your \texttt{in\_data\_files/Par\_file} and set e.g.:
-\begin{lyxcode}
-SUPPRESS\_UTM\_PROJECTION         = .false. ~\\
-\end{lyxcode}
-to be the same in \texttt{Mesh\_Par\_file} and \texttt{Par\_file}.
-This flag should be identical when using the in-house mesher \texttt{xmeshfem3D}, \texttt{xgenerate\_databases} and \texttt{xspecfem3D} together to run simulations.
-
-The flag determines if the coordinates you specify for your source and station locations are given as lat/lon degrees and
-must be converted to UTM coordinates.
-As an example, if you use \texttt{.false.} within \texttt{Mesh\_Par\_file} then you create a mesh with \texttt{xmeshfem3D} using the UTM projection from lat/lon as input format to UTM projected coordinates to store the mesh point positions, which is fine.
-The error then may occur if in the \texttt{Par\_file} you have this set to \texttt{.true.} so that the \texttt{xgenerate\_databases} and \texttt{xspecfem3D} suppress the UTM projection and assume that all coordinates you use now for source and receiver locations are given in meters (that is, converted) already.
-So it won't find the specified locations in the used mesh.
-As a solutions, just change the flag in \texttt{Par\_file} to be the same as in \texttt{Mesh\_Par\_file} and rerun \texttt{xgenerate\_databases} and \texttt{xspecfem3D} to make sure that your simulation works fine.
-
-\item [I get the following error message "forward simulation became unstable and blew up":]
-In most cases this means that your time step size \texttt{DT} is chosen too big.
-Look at your files \texttt{output\_mesher.txt} or \texttt{output\_solver.txt} created in the folder \texttt{in\_out\_files/OUTPUT\_FILES/}.
-In these output files, find the section:
-\begin{lyxcode}
-... ~\\
-********************************************* ~\\
-*** Verification of simulation parameters *** ~\\
-********************************************* ~\\
-... ~\\
-*** Minimum period resolved =    4.308774 ~\\
-*** Maximum suggested time step =   6.8863556E-02 ~\\
-... ~\\
-... ~\\
-\end{lyxcode}
-
-then change \texttt{DT} in the \texttt{in\_data\_files/Par\_file} to be somewhere close to the maximum suggested time step. In the example above: \\
-\texttt{DT                              = 0.05d0} \\
-would (most probably) work fine. It could be also bigger than the 0.068 s suggested. This depends a bit on the distortions of
-your mesh elements. The more regular they are, the bigger you can choose \texttt{DT}. Just play with this value a bit and see
-when the simulation becomes stable ...
-
-\end{description}
-
-\chapter{\label{cha:License}License }
-
-\textbf{GNU GENERAL PUBLIC LICENSE Version 2, June 1991. Copyright
-(C) 1989, 1991 Free Software Foundation, Inc. 59 Temple Place, Suite
-330, Boston, MA 02111-1307 USA} \\
-Everyone is permitted to copy and distribute verbatim copies of this
-license document, but changing it is not allowed.
-
-\section*{Preamble}
-
-The licenses for most software are designed to take away your freedom
-to share and change it. By contrast, the GNU General Public License
-is intended to guarantee your freedom to share and change free software
--- to make sure the software is free for all its users. This General
-Public License applies to most of the Free Software Foundation's software
-and to any other program whose authors commit to using it. (Some other
-Free Software Foundation software is covered by the GNU Library General
-Public License instead.) You can apply it to your programs, too.
-
-When we speak of free software, we are referring to freedom, not price.
-Our General Public Licenses are designed to make sure that you have
-the freedom to distribute copies of free software (and charge for
-this service if you wish), that you receive source code or can get
-it if you want it, that you can change the software or use pieces
-of it in new free programs; and that you know you can do these things.
-
-To protect your rights, we need to make restrictions that forbid anyone
-to deny you these rights or to ask you to surrender the rights. These
-restrictions translate to certain responsibilities for you if you
-distribute copies of the software, or if you modify it.
-
-For example, if you distribute copies of such a program, whether gratis
-or for a fee, you must give the recipients all the rights that you
-have. You must make sure that they, too, receive or can get the source
-code. And you must show them these terms so they know their rights.
-
-We protect your rights with two steps:
-
-\begin{enumerate}
-\item Copyright the software, and
-\item Offer you this license which gives you legal permission to copy, distribute
-and/or modify the software.
-\end{enumerate}
-Also, for each author's protection and ours, we want to make certain
-that everyone understands that there is no warranty for this free
-software. If the software is modified by someone else and passed on,
-we want its recipients to know that what they have is not the original,
-so that any problems introduced by others will not reflect on the
-original authors' reputations.
-
-Finally, any free program is threatened constantly by software patents.
-We wish to avoid the danger that redistributors of a free program
-will individually obtain patent licenses, in effect making the program
-proprietary. To prevent this, we have made it clear that any patent
-must be licensed for everyone's free use or not licensed at all.
-
-The precise terms and conditions for copying, distribution and modification
-follow.
-
-\section*{GNU GENERAL PUBLIC LICENSE TERMS AND CONDITIONS FOR COPYING, DISTRIBUTION
-AND MODIFICATION }
-
-\begin{itemize}
-
-\item[0.]This License applies to any program or other work which
-contains a notice placed by the copyright holder saying it may be
-distributed under the terms of this General Public License. The ``Program''
-below refers to any such program or work, and a ``work based on the
-Program'' means either the Program or any derivative work under copyright
-law: that is to say, a work containing the Program or a portion of
-it, either verbatim or with modifications and/or translated into another
-language. (Hereinafter, translation is included without limitation
-in the term ``modification.'') Each licensee is addressed as ``you.''\\
-\\
-Activities other than copying, distribution and modification are not
-covered by this License; they are outside its scope. The act of running
-the Program is not restricted, and the output from the Program is
-covered only if its contents constitute a work based on the Program
-(independent of having been made by running the Program). Whether
-that is true depends on what the Program does.
-
-\end{itemize}
-
-\begin{enumerate}
-\item You may copy and distribute verbatim copies of the Program's source
-code as you receive it, in any medium, provided that you conspicuously
-and appropriately publish on each copy an appropriate copyright notice
-and disclaimer of warranty; keep intact all the notices that refer
-to this License and to the absence of any warranty; and give any other
-recipients of the Program a copy of this License along with the Program.
-
-You may charge a fee for the physical act of transferring a copy,
-and you may at your option offer warranty protection in exchange for
-a fee.
-
-\item You may modify your copy or copies of the Program or any portion of
-it, thus forming a work based on the Program, and copy and distribute
-such modifications or work under the terms of Section 1 above, provided
-that you also meet all of these conditions:
-
-\begin{enumerate}
-\item You must cause the modified files to carry prominent notices stating
-that you changed the files and the date of any change.
-\item You must cause any work that you distribute or publish, that in whole
-or in part contains or is derived from the Program or any part thereof,
-to be licensed as a whole at no charge to all third parties under
-the terms of this License.
-\item If the modified program normally reads commands interactively when
-run, you must cause it, when started running for such interactive
-use in the most ordinary way, to print or display an announcement
-including an appropriate copyright notice and a notice that there
-is no warranty (or else, saying that you provide a warranty) and that
-users may redistribute the program under these conditions, and telling
-the user how to view a copy of this License. (Exception: if the Program
-itself is interactive but does not normally print such an announcement,
-your work based on the Program is not required to print an announcement.)
-\end{enumerate}
-These requirements apply to the modified work as a whole. If identifiable
-sections of that work are not derived from the Program, and can be
-reasonably considered independent and separate works in themselves,
-then this License, and its terms, do not apply to those sections when
-you distribute them as separate works. But when you distribute the
-same sections as part of a whole which is a work based on the Program,
-the distribution of the whole must be on the terms of this License,
-whose permissions for other licensees extend to the entire whole,
-and thus to each and every part regardless of who wrote it.
-
-Thus, it is not the intent of this section to claim rights or contest
-your rights to work written entirely by you; rather, the intent is
-to exercise the right to control the distribution of derivative or
-collective works based on the Program.
-
-In addition, mere aggregation of another work not based on the Program
-with the Program (or with a work based on the Program) on a volume
-of a storage or distribution medium does not bring the other work
-under the scope of this License.
-
-\item You may copy and distribute the Program (or a work based on it, under
-Section 2) in object code or executable form under the terms of Sections
-1 and 2 above provided that you also do one of the following:
-
-\begin{enumerate}
-\item Accompany it with the complete corresponding machine-readable source
-code, which must be distributed under the terms of Sections 1 and
-2 above on a medium customarily used for software interchange; or,
-\item Accompany it with a written offer, valid for at least three years,
-to give any third party, for a charge no more than your cost of physically
-performing source distribution, a complete machine-readable copy of
-the corresponding source code, to be distributed under the terms of
-Sections 1 and 2 above on a medium customarily used for software interchange;
-or,
-\item Accompany it with the information you received as to the offer to
-distribute corresponding source code. (This alternative is allowed
-only for noncommercial distribution and only if you received the program
-in object code or executable form with such an offer, in accord with
-Subsection b above.)
-\end{enumerate}
-The source code for a work means the preferred form of the work for
-making modifications to it. For an executable work, complete source
-code means all the source code for all modules it contains, plus any
-associated interface definition files, plus the scripts used to control
-compilation and installation of the executable. However, as a special
-exception, the source code distributed need not include anything that
-is normally distributed (in either source or binary form) with the
-major components (compiler, kernel, and so on) of the operating system
-on which the executable runs, unless that component itself accompanies
-the executable.
-
-If distribution of executable or object code is made by offering access
-to copy from a designated place, then offering equivalent access to
-copy the source code from the same place counts as distribution of
-the source code, even though third parties are not compelled to copy
-the source along with the object code.
-
-\item You may not copy, modify, sublicense, or distribute the Program except
-as expressly provided under this License. Any attempt otherwise to
-copy, modify, sublicense or distribute the Program is void, and will
-automatically terminate your rights under this License. However, parties
-who have received copies, or rights, from you under this License will
-not have their licenses terminated so long as such parties remain
-in full compliance.
-\item You are not required to accept this License, since you have not signed
-it. However, nothing else grants you permission to modify or distribute
-the Program or its derivative works. These actions are prohibited
-by law if you do not accept this License. Therefore, by modifying
-or distributing the Program (or any work based on the Program), you
-indicate your acceptance of this License to do so, and all its terms
-and conditions for copying, distributing or modifying the Program
-or works based on it.
-\item Each time you redistribute the Program (or any work based on the Program),
-the recipient automatically receives a license from the original licensor
-to copy, distribute or modify the Program subject to these terms and
-conditions. You may not impose any further restrictions on the recipients'
-exercise of the rights granted herein. You are not responsible for
-enforcing compliance by third parties to this License.
-\item If, as a consequence of a court judgment or allegation of patent infringement
-or for any other reason (not limited to patent issues), conditions
-are imposed on you (whether by court order, agreement or otherwise)
-that contradict the conditions of this License, they do not excuse
-you from the conditions of this License. If you cannot distribute
-so as to satisfy simultaneously your obligations under this License
-and any other pertinent obligations, then as a consequence you may
-not distribute the Program at all. For example, if a patent license
-would not permit royalty-free redistribution of the Program by all
-those who receive copies directly or indirectly through you, then
-the only way you could satisfy both it and this License would be to
-refrain entirely from distribution of the Program.
-
-If any portion of this section is held invalid or unenforceable under
-any particular circumstance, the balance of the section is intended
-to apply and the section as a whole is intended to apply in other
-circumstances.
-
-It is not the purpose of this section to induce you to infringe any
-patents or other property right claims or to contest validity of any
-such claims; this section has the sole purpose of protecting the integrity
-of the free software distribution system, which is implemented by
-public license practices. Many people have made generous contributions
-to the wide range of software distributed through that system in reliance
-on consistent application of that system; it is up to the author/donor
-to decide if he or she is willing to distribute software through any
-other system and a licensee cannot impose that choice.
-
-This section is intended to make thoroughly clear what is believed
-to be a consequence of the rest of this License.
-
-\item If the distribution and/or use of the Program is restricted in certain
-countries either by patents or by copyrighted interfaces, the original
-copyright holder who places the Program under this License may add
-an explicit geographical distribution limitation excluding those countries,
-so that distribution is permitted only in or among countries not thus
-excluded. In such case, this License incorporates the limitation as
-if written in the body of this License.
-\item The Free Software Foundation may publish revised and/or new versions
-of the General Public License from time to time. Such new versions
-will be similar in spirit to the present version, but may differ in
-detail to address new problems or concerns.
-
-Each version is given a distinguishing version number. If the Program
-specifies a version number of this License which applies to it and
-``any later version,'' you have the option of following the terms
-and conditions either of that version or of any later version published
-by the Free Software Foundation. If the Program does not specify a
-version number of this License, you may choose any version ever published
-by the Free Software Foundation.
-
-\item If you wish to incorporate parts of the Program into other free programs
-whose distribution conditions are different, write to the author to
-ask for permission. For software which is copyrighted by the Free
-Software Foundation, write to the Free Software Foundation; we sometimes
-make exceptions for this. Our decision will be guided by the two goals
-of preserving the free status of all derivatives of our free software
-and of promoting the sharing and reuse of software generally.
-\end{enumerate}
-
-\subsection*{NO WARRANTY }
-
-\begin{itemize}
-
-\item[11.]BECAUSE THE PROGRAM IS LICENSED FREE OF CHARGE, THERE IS
-NO WARRANTY FOR THE PROGRAM, TO THE EXTENT PERMITTED BY APPLICABLE
-LAW. EXCEPT WHEN OTHERWISE STATED IN WRITING THE COPYRIGHT HOLDERS
-AND/OR OTHER PARTIES PROVIDE THE PROGRAM ``AS IS'' WITHOUT WARRANTY
-OF ANY KIND, EITHER EXPRESSED OR IMPLIED, INCLUDING, BUT NOT LIMITED
-TO, THE IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR
-PURPOSE. THE ENTIRE RISK AS TO THE QUALITY AND PERFORMANCE OF THE
-PROGRAM IS WITH YOU. SHOULD THE PROGRAM PROVE DEFECTIVE, YOU ASSUME
-THE COST OF ALL NECESSARY SERVICING, REPAIR OR CORRECTION.
-
-\item[12.]IN NO EVENT UNLESS REQUIRED BY APPLICABLE LAW OR AGREED
-TO IN WRITING WILL ANY COPYRIGHT HOLDER, OR ANY OTHER PARTY WHO MAY
-MODIFY AND/OR REDISTRIBUTE THE PROGRAM AS PERMITTED ABOVE, BE LIABLE
-TO YOU FOR DAMAGES, INCLUDING ANY GENERAL, SPECIAL, INCIDENTAL OR
-CONSEQUENTIAL DAMAGES ARISING OUT OF THE USE OR INABILITY TO USE THE
-PROGRAM (INCLUDING BUT NOT LIMITED TO LOSS OF DATA OR DATA BEING RENDERED
-INACCURATE OR LOSSES SUSTAINED BY YOU OR THIRD PARTIES OR A FAILURE
-OF THE PROGRAM TO OPERATE WITH ANY OTHER PROGRAMS), EVEN IF SUCH HOLDER
-OR OTHER PARTY HAS BEEN ADVISED OF THE POSSIBILITY OF SUCH DAMAGES.
-
-\end{itemize}
-
-\section*{END OF TERMS AND CONDITIONS }
-
-\subsection*{How to Apply These Terms to Your New Programs}
-
-If you develop a new program, and you want it to be of the greatest
-possible use to the public, the best way to achieve this is to make
-it free software which everyone can redistribute and change under
-these terms.
-
-To do so, attach the following notices to the program. It is safest
-to attach them to the start of each source file to most effectively
-convey the exclusion of warranty; and each file should have at least
-the ``copyright'' line and a pointer to where the full notice is found.
-For example:
-
-\begin{quote}
-One line to give the program's name and a brief idea of what it does.
-Copyright {\footnotesize $\copyright$ (}year) (name of author)
-
-This program is free software; you can redistribute it and/or modify
-it under the terms of the GNU General Public License as published
-by the Free Software Foundation; either version 2 of the License,
-or (at your option) any later version.
-
-This program is distributed in the hope that it will be useful, but
-WITHOUT ANY WARRANTY; without even the implied warranty of MERCHANTABILITY
-or FITNESS FOR A PARTICULAR PURPOSE. See the GNU General Public License
-for more details.
-
-You should have received a copy of the GNU General Public License
-along with this program; if not, write to the Free Software Foundation,
-Inc., 59 Temple Place, Suite 330, Boston, MA 02111-1307 USA
-\end{quote}
-Also add information on how to contact you by electronic and paper
-mail.
-
-If the program is interactive, make it output a short notice like
-this when it starts in an interactive mode:
-
-\begin{quote}
-Gnomovision version 69, Copyright $\copyright$ year name of author Gnomovision
-comes with ABSOLUTELY NO WARRANTY; for details type `show w'. This
-is free software, and you are welcome to redistribute it under certain
-conditions; type `show c' for details.
-\end{quote}
-The hypothetical commands `show w' and `show c' should show the appropriate
-parts of the General Public License. Of course, the commands you use
-may be called something other than `show w' and `show c'; they could
-even be mouse-clicks or menu items -- whatever suits your program.
-
-You should also get your employer (if you work as a programmer) or
-your school, if any, to sign a ``copyright disclaimer'' for the program,
-if necessary. Here is a sample; alter the names:
-
-\begin{quote}
-Yoyodyne, Inc., hereby disclaims all copyright interest in the program
-`Gnomovision' (which makes passes at compilers) written by James Hacker.
-
-(signature of Ty Coon)\\
-1 April 1989 \\
-Ty Coon, President of Vice
-\end{quote}
-This General Public License does not permit incorporating your program
-into proprietary programs. If your program is a subroutine library,
-you may consider it more useful to permit linking proprietary applications
-with the library. If this is what you want to do, use the GNU Library
-General Public License instead of this License.
-\end{document}

### 2015-04-24 doc/USER_MANUAL/15_informations_for_developers.tex

\chapter{Information for developers of the code, and for people who want to learn how the technique works}\label{cha:developers}

You can get a very simple 1D version of a demo code (there is one in Fortran and one in Python):\\

git clone --recursive https://github.com/geodynamics/specfem1d.git\\

\noindent We also have simple 3D demo source codes that implement the SEM in a single, small program, in directory\\
utils/small\_SEM\_solvers\_in\_Fortran\_and\_C\_without\_MPI\_to\_learn of the specfem3d package.
They are useful to learn how the spectral-element method works, and how to write or modify a code to implement it.
Also useful to test new ideas by modifying these simple codes to run some tests.\\

\noindent For information on how to contribute to the code, i.e., for how to make your modifications, additions or improvements part of the
official package, see https://github.com/geodynamics/specfem3d/wiki .

### 2015-04-24 doc/USER_MANUAL/C_troubleshooting.tex

--- 2012-09-28 doc/USER_MANUAL/manual_SPECFEM3D.tex
@@
-%% LyX 1.5.1 created this file.  For more info, see http://www.lyx.org/.
-%% Do not edit unless you really know what you are doing.
-\documentclass[oneside,english,onecolumn]{book}
-\usepackage[T1]{fontenc}
-\usepackage[latin1]{inputenc}
-\usepackage{geometry}
-\geometry{letterpaper,tmargin=1in,bmargin=1in,lmargin=1in,rmargin=1in}
-\setcounter{secnumdepth}{3}
-\setcounter{tocdepth}{3}
-\usepackage{float}
-\usepackage{textcomp}
-\usepackage{amsmath}
+\chapter{Troubleshooting}\label{cha:Troubleshooting}
 
-% figures
-\usepackage[pdftex]{graphicx}
+\section*{FAQ}
+\begin{description}
+\item [{configuration fails:}]~
 
-% we are running pdflatex, so convert .eps files to .pdf
-\usepackage{epstopdf}
+Examine the log file 'config.log'. It contains detailed informations.
+In many cases, the paths to these specific compiler commands F90,
+CC and MPIF90 will not be correct if `./configure` fails.
 
-\usepackage{amssymb}
-\IfFileExists{url.sty}{\usepackage{url}}
-                      {\newcommand{\url}{\texttt}}
-\usepackage[authoryear]{natbib}
-
-\makeatletter
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% LyX specific LaTeX commands.
-\newcommand{\noun}[1]{\textsc{#1}}
-%% A simple dot to overcome graphicx limitations
-\newcommand{\lyxdot}{.}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Textclass specific LaTeX commands.
-\newenvironment{lyxcode}
-{\begin{list}{}{
-\setlength{\rightmargin}{\leftmargin}
-\setlength{\listparindent}{0pt}% needed for AMS classes
-\raggedright
-\setlength{\itemsep}{0pt}
-\setlength{\parsep}{0pt}
-\normalfont\ttfamily}%
- \item[]}
-{\end{list}}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% User specified LaTeX commands.
-%\renewcommand{\baselinestretch}{1.5}
-
-% figures
-\usepackage[dvips]{epsfig}
-
-\usepackage{wrapfig}
-
-% fonts
-\usepackage{times}
-
-% hyperlinks to sections and references
-\usepackage[pdftex,bookmarks=true,bookmarksnumbered=true,pdfpagemode=None,pdfstartview=FitH,pdfpagelayout=SinglePage,pdfborder={0 0 0}]{hyperref}
-
-\newcommand{\urlwithparentheses}[1]{(\url{#1})}
-
-% biblio GJI
-\bibliographystyle{abbrvnat}
-
-\newcommand{\toall}[1]{\textbf{*** All: #1 ***}}
-\newcommand{\tojeroen}[1]{\textbf{*** Jeroen: #1 ***}}
-\newcommand{\tobrian}[1]{\textbf{*** Brian: #1 ***}}
-\newcommand{\tovala}[1]{\textbf{*** Vala: #1 ***}}
-\newcommand{\tovalabrian}[1]{\textbf{*** Vala \& Brian: #1 ***}}
-\newcommand{\tovalaqinya}[1]{\textbf{*** Vala \& Qinya: #1 ***}}
-\newcommand{\toqinya}[1]{\textbf{*** Qinya: #1 ***}}
-\newcommand{\tomin}[1]{\textbf{*** Min: #1 ***}}
-\newcommand{\toalessia}[1]{\textbf{*** Alessia: #1 ***}}
-\newcommand{\todimitri}[1]{\textbf{*** Dimitri: #1 ***}}
-
-\newcommand{\nexxi}{\mbox{\texttt{NEX\_XI\/}}}
-\newcommand{\nexeta}{\mbox{\texttt{NEX\_ETA\/}}}
-\newcommand{\nprocxi}{\mbox{\texttt{NPROC\_XI\/}}}
-\newcommand{\nproceta}{\mbox{\texttt{NPROC\_ETA\/}}}
-\newcommand{\nchunks}{\mbox{\texttt{NCHUNKS\/}}}
-\newcommand{\eg}{e.g.,}
-
-% colors for the text
-\usepackage[dvipsnames,usenames]{color}
-\newcommand{\red}[1]{\textcolor{Red}{#1}}
-
-\usepackage{babel}
-\makeatother
-
-\begin{document}
-\begin{center}
-\thispagestyle{empty}%
-\vspace*{-1.8truecm}
-\noindent\makebox[\textwidth]{%
-\noindent\includegraphics[width=0.83\paperwidth]{figures/specfem_3d-cover.pdf}
-}
-\end{center}
-%
-\title{\textbf{SPECFEM3D}\\
-\textbf{User Manual}}
-
-\author{$\copyright$ Princeton University (USA) and\\
-CNRS / INRIA / University of Pau (France)\\
-Version 2.1}
-
-\date{\noindent \today}
-
-\maketitle
-
-\section*{Authors}
-The SPECFEM3D package was first developed by Dimitri Komatitsch and Jeroen Tromp at Harvard University and Caltech, USA, starting in 1998, based on earlier work by Dimitri Komatitsch and Jean-Pierre Vilotte at Institut de Physique du Globe (IPGP) in Paris, France from 1995 to 1997.\\
-
-Since then it has been developed and maintained by a development team: in alphabetical order,
-Piero Basini,
-C\'eline Blitz,
-Ebru Bozda\u{g},
-Emanuele Casarotti,
-Joseph Charles,
-Min Chen,
-Vala Hj\"orleifsd\'ottir,
-Sue Kientz,
-Dimitri Komatitsch,
-Jes\'us Labarta,
-Nicolas Le Goff,
-Pieyre Le Loher,
-Qinya Liu,
-Yang Luo,
-Alessia Maggi,
-Federica Magnoni,
-Roland Martin,
-Dennis McRitchie,
-Matthias Meschede,
-Peter Messmer,
-David Mich\'ea,
-Tarje Nissen-Meyer,
-Daniel Peter,
-Max Rietmann,
-Brian Savage,
-Bernhard Schuberth,
-Anne Sieminski,
-Leif Strand,
-Carl Tape,
-Jeroen Tromp,
-Zhinan Xie,
-Hejun Zhu.\\
-%\toall{add full list of developers here and update the AUTHORS file accordingly}
-
-The cover graphic of the manual was created
-by Santiago Lombeyda from Caltech's Center for Advanced Computing Research (CACR) \urlwithparentheses{http://www.cacr.caltech.edu}. \\
-
-\newpage{}
-
-\tableofcontents{}
-
-\chapter{Introduction}
-
-The software package SPECFEM3D simulates seismic
-wave propagation at the local or regional scale
-based upon the spectral-element method (SEM).
-The SEM is a continuous Galerkin technique \citep{TrKoLi08,PeKoLuMaLeCaLeMaLiBlNiBaTr11},
-which can easily be made discontinous \citep{BeMaPa94,Ch00,KoWoHu02,ChCaVi03,LaWaBe05,Kop06,WiStBuGh10,AcKo11};
-it is then close to a particular case of the discontinuous Galerkin technique \citep{ReHi73,Arn82,FaRi99,HuHuRa99,CoKaSh00,GiHeWa02,RiWh03,MoRi05,GrScSc06,AiMoMu06,BeLaPi06,DuKa06,DeSeWh08,PuAmKa09,WiStBuGh10,DeSe10,EtChViGl10}, with optimized efficiency because of its tensorized basis functions \citep{WiStBuGh10,AcKo11}.
-In particular, it can accurately handle very distorted mesh elements \citep{OlSe11}.\\
-
-It has very good accuracy and convergence properties \citep{MaPa89,SePr94,DeFiMu02,Coh02,DeSe07,SeOl08}.
-The spectral element approach admits spectral rates of convergence and allows exploiting $hp$-convergence schemes.
-It is also very well suited to parallel implementation on very large supercomputers \citep{KoTsChTr03,TsKoChTr03,KoLaMi08a,CaKoLaTiMiLeSnTr08,KoViCh10} as well as on clusters of GPU accelerating graphics cards \citep{KoMiEr09,KoErGoMi10,Kom11}.
-Tensor products inside each element can be optimized to reach very high efficiency \citep{DeFiMu02}, and mesh point and element numbering can be optimized to reduce processor cache misses and improve cache reuse \citep{KoLaMi08a}.
-The SEM can also handle triangular (in 2D) or tetrahedral (in 3D) elements \citep{WinBoyd96,TaWi00,KoMaTrTaWi01,Coh02,MeViSa06} as well as mixed meshes, although with increased cost and reduced accuracy in these elements, as in the discontinuous Galerkin method.\\
-
-Note that in many geological models in the context of seismic wave propagation studies
-(except for instance for fault dynamic rupture studies, in which very high frequencies or supershear rupture need to be modeled near the fault, see e.g.
-\cite{BeGlCrViPi07,BeGlCrVi09,PuAmKa09,TaCrEtViBeSa10})
-a continuous formulation is sufficient because material property contrasts are not drastic and thus
-conforming mesh doubling bricks can efficiently handle mesh size variations \citep{KoTr02a,KoLiTrSuStSh04,LeChLiKoHuTr08,LeChKoHuTr09,LeKoHuTr09}.\\
-
-For a detailed introduction to the SEM as applied to regional
-seismic wave propagation, please consult \citet{PeKoLuMaLeCaLeMaLiBlNiBaTr11,TrKoLi08,KoVi98,KoTr99,ChKoViCaVaFe07} and
-in particular \citet{LeKoHuTr09,LeChKoHuTr09,LeChLiKoHuTr08,GoAmTaCaSmSaMaKo09,WiKoScTr04,KoLiTrSuStSh04}.
-A detailed theoretical analysis of the dispersion
-and stability properties of the SEM is available in \citet{Coh02}, \citet{DeSe07}, \citet{SeOl07} and \citet{SeOl08}.\\
-
-Effects due to lateral variations in compressional-wave speed, shear-wave
-speed, density, a 3D crustal model, topography and bathymetry are
-included. The package can accommodate
-full 21-parameter anisotropy (see~\citet{ChTr07}) as well as lateral
-variations in attenuation \citep{SaKoTr10}. Adjoint capabilities and finite-frequency
-kernel simulations are included \citep{TrKoLi08,PeKoLuMaLeCaLeMaLiBlNiBaTr11,LiTr06,FiIgBuKe09,ViOp09}.\\
-
-The SEM was originally developed in computational fluid dynamics \citep{Pat84,MaPa89}
-and has been successfully adapted to address problems in seismic wave propagation.
-Early seismic wave propagation applications of the SEM, utilizing Legendre basis functions and a
-perfectly diagonal mass matrix, include \cite{CoJoTo93}, \cite{Kom97},
-\cite{FaMaPaQu97}, \cite{CaGa97}, \cite{KoVi98} and \cite{KoTr99},
-whereas applications involving Chebyshev basis functions and a nondiagonal mass matrix
-include \cite{SePr94}, \cite{PrCaSe94} and \cite{SePrPr95}.\\
-
-All SPECFEM3D software is written in Fortran90 with full portability
-in mind, and conforms strictly to the Fortran95 standard. It uses
-no obsolete or obsolescent features of Fortran77. The package uses
-parallel programming based upon the Message Passing Interface (MPI)
-\citep{GrLuSk94,Pac97}.\\
-
-SPECFEM3D won the Gordon Bell award for best performance at the SuperComputing~2003
-conference in Phoenix, Arizona (USA) (see \cite{KoTsChTr03}
-and \url{www.sc-conference.org/sc2003/nr_finalaward.html}).
-It was a finalist again in 2008 for a run at 0.16 petaflops (sustained) on 149,784 processors of the `Jaguar' Cray XT5 system at Oak Ridge National Laboratories (USA) \citep{CaKoLaTiMiLeSnTr08}.
-It also won the BULL Joseph Fourier supercomputing award in 2010.\\
-
-This new release of the code (version 2.1) includes support for GPU graphics card acceleration \citep{KoMiEr09,KoGoErMi10,KoErGoMi10,MiKo10,Kom11}.\\
-
-The next release of the code will include support Convolution or Auxiliary Differential Equation Perfectly Matched absorbing Layers
-(C-PML or ADE-PML) \citep{KoMa07,MaKoEz08,MaKoGe08,MaKo09,MaKoGeBr10}.
-It will also use the PT-SCOTCH parallel library for mesh partitioning.\\
-
-SPECFEM3D includes coupled fluid-solid domains and
-adjoint capabilities, which enables one to address seismological inverse problems, but for linear rheologies only so far.
-To accommodate visco-plastic or non-linear rheologies, the reader can refer to
-the \href{http://geoelse.stru.polimi.it/}{GeoELSE} software package \citep{CaGa97,StPaIg09,ChMoTsBaKrKaStKr10}.
-
-\section{Citation}
-
-If you use SPECFEM3D for your own research, please cite at least one
-of the following articles:
-
-\begin{description}
-
-\item [Numerical simulations in general] ~\\
-Forward and adjoint simulations are described in detail in \cite{TrKoLi08,PeKoLuMaLeCaLeMaLiBlNiBaTr11,VaCaSaKoVi99,KoMiEr09,KoErGoMi10,KoGoErMi10,ChKoViCaVaFe07,MaKoDi09,KoViCh10,CaKoLaTiMiLeSnTr08,TrKoHjLiZhPeBoMcFrTrHu10,KoRiTr02,KoTr02a,KoTr02b,KoTr99} or \cite{KoVi98}.
-Additional aspects dealing with adjoint simulations are described in
-\cite{TrTaLi05,LiTr06,TrKoLi08,LiTr08,TrKoHjLiZhPeBoMcFrTrHu10,PeKoLuMaLeCaLeMaLiBlNiBaTr11}.
-Domain decomposition is explained in detail in \cite{MaKoBlLe08}, and excellent scaling up to 150,000 processor cores is shown for instance in
-\cite{CaKoLaTiMiLeSnTr08,KoLaMi08a,MaKoBlLe08,KoErGoMi10,KoGoErMi10,Kom11},
-
-\item [GPU computing] ~\\
-Computing on GPU graphics cards for acoustic or seismic wave propagation applications
-is described in detail in \cite{KoMiEr09,KoGoErMi10,KoErGoMi10,MiKo10,Kom11}.
-
-\end{description}
-
-\noindent
-If you use this new version 2.1, which has non blocking MPI for much better performance for medium or large runs, please cite at least one of these six articles,
-in which results of non blocking MPI runs are presented: \cite{PeKoLuMaLeCaLeMaLiBlNiBaTr11,KoGoErMi10,KoErGoMi10,KoViCh10,Kom11,CaKoLaTiMiLeSnTr08,MaKoBlLe08}.\\
-
-\noindent
-If you work on geophysical applications, you may be interested in citing some of these application articles as well, among others:
-
-\begin{description}
-
-\item [Southern California simulations] ~\\
-\cite{KoLiTrSuStSh04,KrJiKoTr06a,KrJiKoTr06b}.
-
-If you use the 3D southern California model, please cite
-\citet{SuSh03} (Los Angeles model),
-\citet{lovelyetal06} (Salton Trough), and
-\citet{hauksson2000} (southern California).
-The Moho map was determined by
-\citet{zhu&kanamori2000}.
-The 1D SoCal model was developed by \citet{DrHe90}.
-
-\item [Anisotropy] ~\\
-\cite{ChTr07,JiTsKoTr05,ChFaKo04,FaChKo04,RiRiKoTrHe02,TrKo00}.
-
-\item [Attenuation] ~\\
-\cite{SaKoTr10,KoTr02a,KoTr99}.
-
-\item [Topography] ~\\
-\cite{LeKoHuTr09,LeChKoHuTr09,LeChLiKoHuTr08,GoAmTaCaSmSaMaKo09,WiKoScTr04}.
-
-\end{description}
-
-The corresponding Bib\TeX{} entries may be found
-in file \texttt{doc/USER\_MANUAL/bibliography.bib}.
-
-\section{Support}
-
-This material is based upon work supported by the USA National Science
-Foundation under Grants No. EAR-0406751 and EAR-0711177, by the French
-CNRS, French INRIA Sud-Ouest MAGIQUE-3D, French ANR NUMASIS under
-Grant No. ANR-05-CIGC-002, and European FP6 Marie Curie International
-Reintegration Grant No. MIRG-CT-2005-017461.
-Any opinions, findings, and conclusions or recommendations expressed in this material are
-those of the authors and do not necessarily reflect the views of the
-USA National Science Foundation, CNRS, INRIA, ANR or the European
-Marie Curie program.
-
-\chapter{\label{cha:Getting-Started}Getting Started}
-
-The SPECFEM3D software package comes in a gzipped tar ball. In the
-directory in which you want to install the package, type
+Please make sure that you have a working installation of a Fortran
+compiler, a C compiler and an MPI implementation. You should be able
+to compile this little program code:
 \begin{lyxcode}
-{\small tar~-zxvf~SPECFEM3D\_V2.1.0.tar.gz}{\small \par}
+program~main~~~\\
+~include~'mpif.h'~~~\\
+~integer,~parameter~::~CUSTOM\_MPI\_TYPE~=~MPI\_REAL~~~\\
+~integer~ier~~~\\
+~call~MPI\_INIT(ier)~~~\\
+~call~MPI\_BARRIER(MPI\_COMM\_WORLD,ier)~~~\\
+~call~MPI\_FINALIZE(ier)~~~\\
+~end
 \end{lyxcode}
-The directory \texttt{\small SPECFEM3D/} will then contain
-the source code.\\
-
-We recommend that you add {\texttt{ulimit -S -s unlimited}} to your {\texttt{.bash\_profile}} file and/or {\texttt{limit stacksize unlimited }} to your {\texttt{.cshrc}} file to suppress any potential limit to the size of the Unix stack.\\
-
-Then, to configure the software for your system, run the
-\texttt{configure} shell script. This script will attempt to guess
-the appropriate configuration values for your system. However, at
-a minimum, it is recommended that you explicitly specify the appropriate
-command names for your Fortran90 compiler and MPI package:
+\item [{compilation fails stating:}] ~\\
 
 \begin{lyxcode}
-./configure~FC=ifort~MPIFC=mpif90
-\end{lyxcode}
+...~~~\\
+~obj/program\_generate\_databases.o:~In~function~`MAIN\_\_':~~~\\
+~program\_generate\_databases.f90:(.text+0x14):~undefined~\\~~~reference~to~`\_gfortran\_set\_std'~~~\\
+~...~~~\\
 
-Before running the \texttt{configure} script, you should probably edit file \texttt{flags.guess} to make sure that it contains the best compiler options for your system.
-Known issues or things to check are:
-
-\begin{description}
-\item [{\texttt{Intel ifort compiler}}] See if you need to add \texttt{-assume byterecl} for your machine.
-\item [{\texttt{IBM compiler}}] See if you need to add \texttt{-qsave} or \texttt{-qnosave} for your machine.
-\item [{\texttt{Mac OS}}] You will probably need to install \texttt{XCODE}.
-In addition, the \texttt{clock\_gettime} routine, which is used by the \texttt{SCOTCH} library that we use, does not exist in Mac OS.
-You will need to replace it with \texttt{clock\_get\_time} if you want to use \texttt{SCOTCH}.
-\end{description}
-
-When compiling on an IBM machine with the \texttt{xlf} and \texttt{xlc} compilers, we suggest running the \texttt{configure} script
-with the following options:\\
-
-\noindent\texttt{./configure FC=xlf90\_r MPIFC=mpif90 CC=xlc\_r CFLAGS="-O3 -q64" FCFLAGS="-O3 -q64" -with-scotch-dir=...}.\\
-
-On SGI systems, \texttt{flags.guess} automatically informs \texttt{configure}
-to insert `\texttt{`TRAP\_FPE=OFF}'' into the generated \texttt{Makefile}
-in order to turn underflow trapping off.\\
-
-If you run very large meshes on a relatively small number
-of processors, the memory size needed on each processor might become
-greater than 2 gigabytes, which is the upper limit for 32-bit addressing.
-In this case, on some compilers you may need to add \texttt{``-mcmodel=medium}'' or \texttt{``-mcmodel=medium -shared-intel}''
-to the configure options of CFLAGS, FCFLAGS and LDFLAGS otherwise the compiler will display an error
-message (for instance \texttt{``relocation truncated to fit: R\_X86\_64\_PC32 against .bss''} or something similar);
-on an IBM machine with the \texttt{xlf} and \texttt{xlc} compilers, using \texttt{-q64} is usually sufficient.\\
-
-The SPECFEM3D software package relies on the SCOTCH library to partition meshes created with CUBIT.
-Note that we use CUBIT to create meshes of hexahedra, but other packages can be used as well, for instance GiD from \url{http://gid.cimne.upc.es} or Gmsh from \url{http://geuz.org/gmsh} \citep{GeRe09}.
-Even mesh creation packages that generate tetrahedra, for instance TetGen from \url{http://tetgen.berlios.de}, can be used because each tetrahedron can then easily be decomposed into four hexahedra as shown in the picture of the TetGen logo at \url{http://tetgen.berlios.de/figs/Delaunay-Voronoi-3D.gif}; while this approach does not generate hexahedra of optimal quality, it can ease mesh creation in some situations and it has been shown that the spectral-element method can very accurately handle distorted mesh elements \citep{OlSe11}.
-
-The SCOTCH library \citep{PeRo96}
-provides efficent static mapping, graph and mesh partitioning routines. SCOTCH is a free software package developed by
-Fran\c {c}ois Pellegrini et al.
-from LaBRI and INRIA in Bordeaux, France, downloadable from the web page \url{https://gforge.inria.fr/projects/scotch/}.
-It is more recent than METIS, actively maintained and performs better in many cases.
-A recent version of its source code is provided in directory \texttt{src/decompose\_mesh\_SCOTCH/scotch\_5.1.10b}.
-In case no SCOTCH libraries can be found on the system, the configuration will bundle this version for compilation.
-The path to an existing SCOTCH installation can to be set explicitly with the option \texttt{-{}-with-scotch-dir}. Just as an example:
-\begin{lyxcode}
-./configure~FC=ifort~MPIFC=mpif90~-{}-with-scotch-dir=/opt/scotch
-\end{lyxcode}
-
-If you use the Intel ifort compiler to compile the code, we recommend that you use the Intel icc C compiler to compile Scotch, i.e., use:
-\begin{lyxcode}
-./configure~CC=icc~FC=ifort~MPIFC=mpif90
-\end{lyxcode}
-
-To compile a serial version of the code for small meshes that fits
-on one compute node and can therefore be run serially, run \texttt{configure}
-with the \texttt{-{}-without-mpi} option to suppress all calls to
-MPI.
-
-A summary of the most important configuration variables follows.
-
-\begin{description}
-\item [{\texttt{F90}}] Path to the Fortran90 compiler.
-\item [{\texttt{MPIF90}}] Path to MPI Fortran90.
-\item [{\texttt{MPI\_FLAGS}}] Some systems require this flag to link to
-MPI libraries.
-\item [{\texttt{FLAGS\_CHECK}}] Compiler flag for non-critical subroutines.
-\item [{\texttt{FLAGS\_NO\_CHECK}}] Compiler flag for creating fast, production-run
-code for critical subroutines.
-\end{description}
-The configuration script automatically creates for each executable a corresponding
-\texttt{Makefile} in the \texttt{src/} subdirectoy.
-The \texttt{Makefile} contains a number of suggested entries for various
-compilers, e.g., Portland, Intel, Absoft, NAG, and Lahey. The software
-has run on a wide variety of compute platforms, e.g., various PC clusters
-and machines from Sun, SGI, IBM, Compaq, and NEC. Select the compiler
-you wish to use on your system and choose the related optimization
-flags. Note that the default flags in the \texttt{Makefile} are undoubtedly
-not optimal for your system, so we encourage you to experiment with
-these flags and to solicit advice from your systems administrator.
-Selecting the right compiler and optimization flags can make a tremendous
-difference in terms of performance. We welcome feedback on your experience
-with various compilers and flags.
-
-Now that you have set the compiler information, you need to select
-a number of flags in the \texttt{constants.h} file depending on your
-system:
-
-\begin{description}
-\item [{\texttt{LOCAL\_PATH\_IS\_ALSO\_GLOBAL}}] Set to \texttt{.false.}
-on most cluster applications. For reasons of speed, the (parallel)
-distributed database generator typically writes a (parallel) database for the solver on the
-local disks of the compute nodes. Some systems have no local disks,
-e.g., BlueGene or the Earth Simulator, and other systems have a fast
-parallel file system, in which case this flag should be set to \texttt{.true.}.
-Note that this flag is not used by the database generator or the solver; it is
-only used for some of the post-processing.
-\end{description}
-The package can run either
-in single or in double precision mode. The default is single precision
-because for almost all calculations performed using the spectral-element method
-using single precision is sufficient and gives the same results (i.e. the same seismograms);
-and the single precision code is faster and requires exactly half as much memory.
-Select your preference by selecting the appropriate setting in the \texttt{constants.h} file:
-
-\begin{description}
-\item [{\texttt{CUSTOM\_REAL}}] Set to \texttt{SIZE\_REAL} for single precision
-and \texttt{SIZE\_DOUBLE} for double precision.
-\end{description}
-In the \texttt{precision.h} file:
-
-\begin{description}
-\item [{\texttt{CUSTOM\_MPI\_TYPE}}] Set to \texttt{MPI\_REAL} for single
-precision and \texttt{MPI\_DOUBLE\_PRECISION} for double precision.
-\end{description}
-On many current processors (e.g., Intel, AMD, IBM Power), single precision calculations
-are significantly faster; the difference can typically be 10\%
-to 25\%. It is therefore better to use single precision.
-What you can do once for the physical problem you want to study is run the same calculation in single precision
-and in double precision on your system and compare the seismograms.
-If they are identical (and in most cases they will), you can select single precision for your future runs.\\
-
-\section{Adding OpenMP support in addition to MPI}
-
-OpenMP support can be enabled in addition to MPI. However, in many cases performance will not improve
-because our pure MPI implementation is already heavily optimized and thus the resulting code will in fact
-be slightly slower. A possible exception could be IBM BlueGene-type architectures.
-
-To enable OpenMP, uncomment the OpenMP compiler option in two lines in file \texttt{src/specfem3D/Makefile.in}
-(before running \texttt{configure}) and also uncomment the \texttt{\#define USE\_OPENMP} statement in file\\
-\texttt{src/specfem3D/specfem3D.F90}.
-
-The DO-loop using OpenMP threads has a SCHEDULE property. The OMP\_SCHEDULE
-environment variable can set the scheduling policy of that DO-loop.
-Tests performed by Marcin Zielinski at SARA (The Netherlands) showed that often
-the best scheduling policy is DYNAMIC with the size of the chunk equal to the number of
-OpenMP threads, but most preferably being twice as the number of
-OpenMP threads (thus chunk size = 8 for 4 OpenMP threads etc).
-If OMP\_SCHEDULE is not set or is empty, the DO-loop will assume generic
-scheduling policy, which will slow down the job quite a bit.
-
-\section{Visualizing the subroutine calling tree of the source code}
-
-Packages such as \texttt{doxywizard} can be used to visualize the subroutine calling tree of the source code.
-\texttt{Doxywizard} is a GUI front-end for configuring and running \texttt{doxygen}.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\chapter{\label{cha:Mesh-Generation}Mesh Generation}
-
-The first step in running a spectral-element simulation consists of constructing a high-quality mesh for the region under consideration.
-We provide two possibilities to do so: (1) relying on the external, hexahedral mesher CUBIT, or (2) using the provided, internal mesher \texttt{xmeshfem3D}.
-In the following, we explain these two approaches.
-
-\section{\label{cha:Running-the-Mesher-CUBIT}Meshing with \texttt{CUBIT}}
-
-CUBIT is a meshing tool suite for the creation of finite-element meshes for arbitrarily shaped models.
-It has been developed and maintained at Sandia National Laboratories and can be purchased for a small academic institutional fee at \url{http://cubit.sandia.gov}.
-Our experience showed that using CUBIT greatly facilitates and speeds up the generation and preparation of hexahedral, conforming meshes for a variety of geophysical models with increasing complexity.
-
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=.6\textwidth]{figures/mount-cubit.jpg}
-\par\end{centering}
-
-\caption{Example of the graphical user interface of CUBIT. The hexahedral mesh shown in the main display consists of a hexahedral discretization of a single volume with topography.}
-
-\label{fig:mount.cubit}
-\end{figure}
-
-The basic steps in creating a load-balanced, partitioned mesh with CUBIT are:
-\begin{description}
-\item [1.] setting up a hexahedral mesh with CUBIT,
-\item [2.] exporting the CUBIT mesh into a SPECFEM3D file format and
-\item [3.] partitioning the SPECFEM3D mesh files for a chosen number of cores.
-\end{description}
-
-Examples are provided in the SPECFEM3D package in the subdirectory \texttt{examples/}.
-We strongly encourage you to contribute your own example to this package by contacting the CIG Computational Seismology Mailing List  \urlwithparentheses{cig-seismo@geodynamics.org}.
-
-\subsection{Creating the Mesh with CUBIT}
-
-For the installation and handling of the CUBIT meshing tool suite, please refer to the CUBIT user manual and documentation.
-In order to give you a basic understanding of how to use CUBIT for our purposes, examples are provided in the SPECFEM3D package in the subdirectory \texttt{examples/}:
-\begin{description}
-\item [{\texttt{homogeneous\_halfspace}}] Creates a single block model and assigns elastic material parameters.
-\item [{\texttt{layered\_halfspace}}]  Combines two different, elastic material volumes and creates a refinement layer between the two.
-This example can be compared for validation against the solutions provided in subdirectory ~\\
-\texttt{VALIDATION\_3D\_SEM\_SIMPLER\_LAYER\_SOURCE\_DEPTH/}.
-\item [{\texttt{waterlayered\_halfspace}}] Combines an acoustic and elastic material volume as in a schematic marine survey example.
-\item [{\texttt{tomographic\_model}}] Creates a single block model whose material properties will have to be read in from a tomographic model file during the databases creation by \texttt{xgenerate\_databases}.
-\end{description}
-
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=.45\textwidth]{figures/example-homogeneous.jpg}
-\includegraphics[width=.45\textwidth]{figures/example-2layers.jpg} \\
-\includegraphics[width=.45\textwidth]{figures/example-water.jpg}
-\includegraphics[width=.45\textwidth]{figures/example-tomo.jpg}
-\par\end{centering}
-
-\caption{Screenshots of the CUBIT examples provided in subdirectory  \texttt{examples/}: homogeneous halfspace (top-left), layered halfspace (top-right), water layered halfspace (bottom-left) and tomographic model (bottom-right).}
-
-\label{fig:examples.cubit}
-\end{figure}
-
-In each example subdirectory you will find a \texttt{README} file, which explains in a step-by-step tutorial the workflow for the example.
-Please feel free to contribute your own example to this package by contacting the CIG Computational Seismology Mailing List  \urlwithparentheses{cig-seismo@geodynamics.org}.
-
-\subsection{Exporting the Mesh with \texttt{cubit2specfem3d.py} }
-\label{subsec:Exporting-the-Mesh}
-
-Once the geometric model volumes in CUBIT are meshed, you prepare the model for exportation with the definition of material blocks and boundary surfaces.
-Thus, prior to exporting the mesh, you need to define blocks specifying the materials and absorbing boundaries in CUBIT.
-This process could be done automatically using the script \texttt{boundary\_definition.py} if the mesh meets some conditions or manually, following the block convention:
-
-\begin{description}
-\item [material\_name] Each material should have a specific block defined by a unique name.
-The name convention of the material is to start with either 'elastic' or 'acoustic'.
-It must be then followed by a unique identifier, e.g.
-'elastic 1', 'elastic 2', etc.
-The additional attributes to the block define the material description.
-
-For an elastic material:
-\begin{description}
-\item [material\_id] An integer value which is unique for this material.
-\item [Vp] P-wave speed of the material (given in m/s).
-\item [Vs] S-wave speed of the material (given in m/s).
-\item [rho] density of the material (given in kg/m$^3$).
-\item [Q] quality factor to use in case of a simulation with attenuation turned on.
-It should be between 1 and 9000.
-In case no attenuation information is available, it can be set to zero.
-Please note that your Vp- and Vs-speeds are given for a reference frequency.
-To change this reference frequency, you change the value of
-\texttt{ATTENUATION\_f0\_REFERENCE} in the main constants file \texttt{constants.h} found in subdirectory \texttt{src/shared/}.
-\item [anisotropic\_flag] Flag describing the anisotropic model to use in case an anisotropic simulation should be conducted.
-See the file \texttt{model\_aniso.f90} in subdirectory \texttt{src/generate\_databases/} for an implementation of the anisotropic models.
-In case no anisotropy is available, it can be set to zero.
-\end{description}
-Note that this material block has to be defined using all the volumes which belong to this elastic material.
-For volumes belonging to another, different material, you will need to define a new material block.
-
-For an acoustic material:
-\begin{description}
-\item [material\_id] An integer value which is unique for this material.
-\item [Vp] P-wave speed of the material (given in m/s).
-\item [0] S-wave speed of the material is ignored.
-\item [rho] density of the material (given in kg/m$^3$).
-\end{description}
-
-\item [face\_topo] Block definition for the surface which defines the free surface (which can have topography).
-The name of this block must be 'face\_topo', the block has to be defined using all the surfaces which constitute the complete free surface of the model.
-
-\item [face\_abs\_xmin] Block definition for the faces on the absorbing boundaries, one block for each surface with x=Xmin.
-\item [face\_abs\_xmax] Block definition for the faces on the absorbing boundaries, one block for each surface with x=Xmax.
-\item [face\_abs\_ymin] Block definition for the faces on the absorbing boundaries, one block for each surface with y=Ymin.
-\item [face\_abs\_ymax] Block definition for the faces on the absorbing boundaries, one block for each surface with y=Ymax.
-\item [face\_abs\_bottom] Block definition for the faces on the absorbing boundaries, one block for each surface with z=bottom.
-\end{description}
-
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=.45\textwidth]{figures/mount-surface.jpg}
-\includegraphics[width=.45\textwidth]{figures/mount-abs.jpg}
-\par\end{centering}
-
-\caption{Example of the block definitions for the free surface 'face\_topo' (left) and the absorbing boundaries, defined in a single block 'face\_abs\_xmin' (right) in CUBIT.}
-
-\label{fig:mount.abs}
-\end{figure}
-
-Optionally, instead of specifying for each surface at the boundaries a single block like mentioned above, you can also specify a single block for all boundary surfaces and name it as one of the absorbing blocks above, e.g.
-'face\_abs\_xmin'.
-
-After the block definitions are done, you export the mesh using the script \texttt{cubit2specfem3d.py} provided in each of the example directories.
-If the export was successful, you should find the following files in a subdirectory \texttt{MESH/}:
-
-\begin{description}
-\item [nummaterial\_velocity\_file] Defines the material properties. For fully defined materials, the formats is:
-\begin{lyxcode}
-\texttt{domain\_ID} \texttt{material\_ID}  \texttt{rho}  \texttt{vp}  \texttt{vs}  \texttt{Q}  \texttt{anisotropy\_flag}
-\end{lyxcode}
-where
-\texttt{domain\_ID} is 1 for acoustic or 2 for elastic materials, \texttt{material\_ID} a unique identifier,  \texttt{rho} the density in $kg \, m^{-3}$,  \texttt{vp} the P-wave speed in $m \, s^{-1}$, \texttt{vs}  the S-wave speed in $m \, s^{-1}$, \texttt{Q} the quality factor and \texttt{anisotropy\_flag} an identifier for anisotropic models.
-
-%%magnoni
-For tomographic velocity models, please read Chapter \ref{cha:-Changing-the} and Section \ref{sec:Using-tomographic} `Using external tomographic Earth models' for further details.
-%
-%For a tomographic model, the material definition format is:
-%\begin{lyxcode}
-%\texttt{domain\_ID} \texttt{material\_ID}  \texttt{tomography}  \texttt{name}
-%\end{lyxcode}
-%where
-%\texttt{domain\_ID} is 1 for acoustic or 2 for elastic materials, \texttt{material\_ID} a negative, unique identifier, \texttt{tomography} keyword for tomographic material definition and \texttt{name} the name of the tomography file.
-The name is not used so far, rather change the filename defined in the file \texttt{model\_tomography.f90} located in the
-%\texttt{src/generate\_databases/} directory.
-%
-\item [materials\_file] Contains the material associations for each element.
-\item [nodes\_coords\_file] Contains the point locations in Cartesian coordinates of the mesh element corners.
-\item [mesh\_file] Contains the mesh element connectivity.
-\item [free\_surface\_file] Contains the free surface connectivity.
-\item [absorbing\_surface\_file\_xmax] Contains the surface connectivity of the absorbing boundary surface at the Xmax.
-\item [absorbing\_surface\_file\_xmin] Contains the surface connectivity of the absorbing boundary surface at the Xmin.
-\item [absorbing\_surface\_file\_ymax] Contains the surface connectivity of the absorbing boundary surface at the Ymax.
-\item [absorbing\_surface\_file\_ymin] Contains the surface connectivity of the absorbing boundary surface at the Ymin.
-\item [absorbing\_surface\_file\_bottom] Contains the surface connectivity of the absorbing boundary surface at the bottom.
-\end{description}
-
-These mesh files are needed as input files for the partitioner  \texttt{xdecompose\_mesh\_SCOTCH}
-to load-balance the mesh. Please see the next section for further details.
-
-\subsubsection*{Checking the mesh quality}
-The quality of the mesh may be inspected more precisely based upon the serial code
-in the file  \texttt{check\_mesh\_quality\_}~\\
-\texttt{CUBIT\_Abaqus.f90} located in the directory \texttt{src/check\_mesh\_quality\_CUBIT\_Abaqus/}.
-Running this code is optional because no information needed by the
-solver is generated.
-
-Prior to running and compiling this code, you have to export your mesh in CUBIT to an ABAQUS (.inp) format.
-For example, export mesh block IDs belonging to volumes in order to check the quality of the hexahedral elements.
-You also have to determine a number of parameters of your mesh, such as the number of nodes and number of elements
-and modify the header of the \texttt{check\_mesh\_quality\_CUBIT\_Abaqus.f90} source file. Then,
-in the main directory, type
-\begin{lyxcode}
-{\small make xcheck\_mesh\_quality\_CUBIT\_Abaqus}{\small \par}
-\end{lyxcode}
-and use
-\begin{lyxcode}
-{\small ./bin/xcheck\_mesh\_quality\_CUBIT\_Abaqus}{\small \par}
-\end{lyxcode}
-to generate an
-%not supported: AVS output file (\texttt{\small AVS\_meshquality.inp} in AVS UCD format) or
-OpenDX output file (\texttt{\small DX\_mesh\_quality.dx})
-that can be used to investigate mesh quality, e.g. skewness of elements,
-and a Gnuplot histogram (\texttt{\small mesh\_quality\_histogram.txt})
-that can be plotted with gnuplot (type
-`\texttt{\small gnuplot plot\_mesh\_quality\_histogram.gnu}'). The
-histogram is also printed to the screen.
-Analyze that skewness histogram of mesh elements to make sure no element has a skewness above approximately 0.75, otherwise the mesh is of poor quality (and if even a single element has a skewness value above 0.80, then you must definitely improve the mesh).
-If you want to start designing
-your own meshes, this tool is useful for viewing your creations. You
-are striving for meshes with elements with `cube-like' dimensions,
-e.g., the mesh should contain no very elongated or skewed elements.
-
-\subsection{Partitioning the Mesh with \texttt{xdecompose\_mesh\_SCOTCH}}
-
-The SPECFEM3D software package performs large scale simulations in a parallel 'Single Process Multiple Data' way.
-The spectral-element mesh created with CUBIT needs to be distributed on the processors. This partitioning is executed once and for all prior to
-the execution of the solver so it is referred to as a static mapping.
-
-An efficient partitioning is important because it leverages the overall running time of the application.
-It amounts to balance the number of elements in each slice while minimizing the communication costs
-resulting from the placement of adjacent elements on different processors. \texttt{decompose\_mesh\_SCOTCH} depends on the SCOTCH library \citep{PeRo96},
-which provides efficent static mapping, graph and mesh partitioning routines. SCOTCH is a free software package developed by
-Fran\c {c}ois Pellegrini et al.
-from LaBRI and INRIA in Bordeaux, France, downloadable from the web page \url{https://gforge.inria.fr/projects/scotch/}.
-It is more recent than METIS, actively maintained and performs better in many cases.
-A recent version of its source code is provided in directory \texttt{src/decompose\_mesh\_SCOTCH/scotch\_5.1.10b}.
-\\
-
-In most cases, the configuration with \texttt{./configure~FC=ifort} should be sufficient.
-During the configuration process, the script tries to find existing SCOTCH installations.
-In case your system has no pre-existing SCOTCH installation,
-we provide the source code of SCOTCH, which is released open source under the French CeCILL-C version 1 license,
-in directory
-\texttt{src/decompose\_mesh\_SCOTCH/scotch\_5.1.10b}.
-This version gets bundled with the compilation of the SPECFEM3D package if no libraries could have been found.
-If this automatic compilation of the SCOTCH libraries fails,
-please refer to file INSTALL.txt in that directory to see further details how to compile it on your system.
-In case you want to use a pre-existing installation, make sure you have correctly specified the path
-of the SCOTCH library when using the option \texttt{-{}-with-scotch-dir} with the \texttt{./configure} script.
-In the future you should be able to find more recent versions at  \url{http://www.labri.fr/perso/pelegrin/scotch/scotch\_en.html} .
-\\
-
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=.45\textwidth]{figures/mount-partitions.jpg}
-\includegraphics[width=.45\textwidth]{figures/mount-partitions2.jpg}
-\par\end{centering}
-
-\caption{Example of a mesh partitioning onto four cores.
-Each single core partition is colored differently.
-The executable \texttt{xdecompose\_mesh\_SCOTCH} can equally distribute the mesh on any arbitrary number of cores.
-Domain decomposition is explained in detail in \cite{MaKoBlLe08}, and excellent scaling up to 150,000 processor cores in shown for instance in
-\cite{CaKoLaTiMiLeSnTr08,KoLaMi08a,MaKoBlLe08,KoErGoMi10,KoGoErMi10,Kom11}.}
-
-\label{fig:mount.partitions}
-\end{figure}
-
-When you are ready to compile,
-in the main directory type
-\begin{lyxcode}
-{\small make xdecompose\_mesh\_SCOTCH}{\small \par}
-\end{lyxcode}
-If all paths and flags have been set correctly, the executable \texttt{bin/xdecompose\_mesh\_SCOTCH} should be produced.
-
-The partitioning is done in serial for now (in the next release we will provide a parallel version of that code),
-the synopsis is:
-\begin{lyxcode}
-./bin/xdecompose\_mesh\_SCOTCH  nparts  input\_directory output\_directory
-\end{lyxcode}
-where
-\begin{itemize}
-     \item \texttt{nparts} is the number of partitions, i.e., the number of cores for the parallel simulations,
-     \item \texttt{input\_directory} is the directory which holds all the files generated by the Python script \texttt{cubit2specfem3d.py} explained in the previous Section~\ref{subsec:Exporting-the-Mesh}, e.g.
-\texttt{MESH/}, and
-     \item \texttt{output\_directory} is the directory for the output of this partitioner which stores ACII-format files named like \texttt{proc******\_Database} for each partition.
-These files will be needed for creating the distributed databases, and have to reside in the directory  \texttt{LOCAL\_PATH} specified in the main \texttt{Par\_file}, e.g.
-in directory \texttt{in\_out\_files/DATABASES\_MPI}.
-Please see Chapter~\ref{cha:Creating-Distributed-Databases} for further details.
-\end{itemize}
-Note that all the files generated by the Python script  \texttt{cubit2specfem3d.py} must be placed in the \texttt{input\_directory} folder before running the program.
-
-\section{\label{cha:Running-the-Mesher-Meshfem3D}Meshing with \texttt{xmeshfem3D}}
-
-In case you successfully ran the configuration script, you are also ready to compile the internal mesher.
-This is an alternative to CUBIT
-for the mesh generation of relatively simple geological models. The mesher is no longer
-dedicated to Southern California and more flexiblity is provided in this version of the package.
-
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[scale=0.5]{figures/socal_map_mpi}
-\par\end{centering}
-
-\caption{\label{fig:For-parallel-computing}For parallel computing purposes,
-the model block is subdivided in $\nprocxi\times\nproceta$ slices
-of elements. In this example we use $5^{2}=25$ processors. }
-
-\end{figure}
-
-In the main directory, type
-\begin{lyxcode}
-{\small make xmeshfem3D}{\small \par}
-\end{lyxcode}
-If all paths and flags
-have been set correctly, the mesher should now compile and produce
-the executable \texttt{bin/xmeshfem3D}. Please note that
-\texttt{xmeshfem3D} must be called directly from the \texttt{bin/} directory,
-as most of the binaries of the package.
-
-Input for the mesh generation program is provided through the parameter
-file \texttt{Mesh\_Par\_file}, which resides in the subdirectory \texttt{in\_data\_files/meshfem3D\_files/}.
-Before running the mesher, a number of parameters need to be set in
-the \texttt{Mesh\_Par\_file}. This requires a basic understanding of how
-the SEM is implemented, and we encourage you to read \citet{KoVi98,KoTr99}
-and \citet{KoLiTrSuStSh04}.
-
-The mesher and the solver use UTM coordinates internally, therefore
-you need to define the zone number for the UTM projection (e.g., zone
-11 for Los Angeles). Use decimal values for latitude and longitude
-(no minutes/seconds). These values are approximate; the mesher will
-round them off to define a square mesh in UTM coordinates. When running
-benchmarks on rectangular models, turn the UTM projection off by using
-the flag \texttt{\small SUPPRESS\_UTM\_PROJECTION}, in which case
-all `longitude' parameters simply refer to the $x$~axis, and all
-`latitude' parameters simply refer to the $y$~axis. To run the mesher
-for a global simulation, the following parameters need to be set in
-the \texttt{Mesh\_Par\_file}:
-
-\begin{description}
-\item [{\texttt{LATITUDE\_MIN}}] Minimum latitude in the block (negative
-for South).
-\item [{\texttt{LATITUDE\_MAX}}] Maximum latitude in the block.
-\item [{\texttt{LONGITUDE\_MIN}}] Minimum longitude in the block (negative
-for West).
-\item [{\texttt{LONGITUDE\_MAX}}] Maximum longitude in the block.
-\item [{\texttt{DEPTH\_BLOCK\_KM}}] Depth of bottom of mesh in kilometers.
-\item [{\texttt{\noun{UTM\_PROJECTION\_ZONE}}}] UTM projection zone in
-which your model resides, only valid when \texttt{SUPPRESS\_UTM\_}~\\
-\texttt{PROJECTION} is \texttt{.false.}.
-\item [{\texttt{SUPPRESS\_UTM\_PROJECTION}}] set to be \texttt{.false.}
-when your model range is specified in geographical coordinates,
-and needs to be \texttt{.true.} when your model is specified in
-Cartesian coordinates. \noun{UTM projection zone in which your simulation
-region resides.}
-\item [{\texttt{INTERFACES\_FILE }}] File which contains the description of the topography and
-of the interfaces between the different layers of the model, if any.
-The number of spectral elements in the vertical direction within each layer is also defined in this file.
-
-\item [{$\nexxi$}] The number of spectral elements along one side of the
-block. This number \textit{must} be 8~$\times$~a multiple of $\nprocxi$
-defined below. Based upon benchmarks against semi-analytical discrete
-wavenumber synthetic seismograms \citep{KoLiTrSuStSh04}, determined
-that a $\nexxi=288$ run is accurate to a shortest period of roughly
-2~s. Therefore, since accuracy is determined by the number of grid
-points per shortest wavelength, for any particular value of $\nexxi$
-the simulation will be accurate to a shortest period determined by
-\begin{equation}
-\mbox{shortest period (s)}=(288/\nexxi)\times2.\label{eq:shortest_period}\end{equation}
-The number of grid points in each orthogonal direction of the reference
-element, i.e., the number of Gauss-Lobatto-Legendre points, is determined
-by \texttt{NGLLX} in the \texttt{constants.h} file. We generally use
-$\mbox{\texttt{NGLLX\/}}=5$, for a total of $5^{3}=125$ points per
-elements. We suggest not to change this value.
-\item [{$\nexeta$}] The number of spectral elements along the other side
-of the block. This number \textit{must} be 8~$\times$~a multiple
-of $\nproceta$ defined below.
-\item [{$\nprocxi$}] The number of processors or slices along one side
-of the block (see Figure~\ref{fig:For-parallel-computing}); we must
-have $\nexxi=8\times c\times\nprocxi$, where $c\ge1$ is a positive
-integer.
-\item [{$\nproceta$}] The number of processors or slices along the other
-side of the block; we must have $\nexeta=8\times c\times\nproceta$,
-where $c\ge1$ is a positive integer.
-\item [{\texttt{USE\_REGULAR\_MESH}}] set to be \texttt{.true.} if you want a perfectly regular
-mesh or \texttt{.false.} if you want to add doubling horizontal layers to coarsen the mesh. In this case, you
-also need to provide additional information by setting up the next three parameters.
-\item [{\texttt{NDOUBLINGS}}] The number of horizontal doubling layers. Must be set to \texttt{1} or \texttt{2}
-if \texttt{USE\_REGULAR\_MESH} is set to \texttt{.true.}.
-\item [{\texttt{NZ\_DOUBLING\_1}}] The position of the first doubling layer (only interpreted
-if \texttt{USE\_REGULAR\_MESH} is set to \texttt{.true.}).
-\item [{\texttt{NZ\_DOUBLING\_2}}] The position of the second doubling layer (only interpreted
-if \texttt{USE\_REGULAR\_MESH} is set to \texttt{.true.} and if \texttt{NDOUBLINGS} is set to \texttt{2}).
-\item [{\texttt{CREATE\_ABAQUS\_FILES}}] Set this flag to \texttt{.true.} to save Abaqus FEA \urlwithparentheses{www.simulia.com}
-mesh files for subsequent viewing. Turning the flag on generates files in the \texttt{LOCAL\_PATH} directory.
-See Section~\ref{sec:Mesh-graphics} for a discussion of mesh viewing features.
-\item [{\texttt{CREATE\_DX\_FILES}}] Set this flag to \texttt{.true.} to save OpenDX \urlwithparentheses{www.opendx.org}
-mesh files for subsequent viewing.
-\item [{\texttt{LOCAL\_PATH}}] Directory in which the partitions generated
-by the mesher will be written. Generally one uses a directory on the
-local disk of the compute nodes, although on some machines these partitions
-are written on a parallel (global) file system (see also the earlier
-discussion of the \texttt{LOCAL\_PATH\_IS\_ALSO\_GLOBAL} flag in Chapter~\ref{cha:Getting-Started}).
-The mesher generates the necessary partitions in parallel, one set
-for each of the $\nprocxi\times\nproceta$ slices that constitutes
-the mesh (see Figure~\ref{fig:For-parallel-computing}). After the
-mesher finishes, you can log in to one of the compute nodes and view
-the contents of the \texttt{LOCAL\_PATH} directory to see the
-files generated by the mesher.
-These files will be needed for creating the distributed databases, and have to reside in the directory  \texttt{LOCAL\_PATH} specified in the main \texttt{Par\_file}, e.g.
-in directory \texttt{in\_out\_files/DATABASES\_MPI}.
-Please see Chapter~\ref{cha:Creating-Distributed-Databases} for further details.
-\item [{\texttt{NMATERIALS}}] The number of different materials in your model. In the following lines, each material needs to be defined as :
-\begin{lyxcode}
-\texttt{material\_ID}  \texttt{rho}  \texttt{vp}  \texttt{vs}  \texttt{Q}  \texttt{anisotropy\_flag} \texttt{domain\_ID}
-\end{lyxcode}
-where
-\begin{itemize}
-     \item \texttt{Q}           : quality factor (0=no attenuation)
-     \item \texttt{anisotropy\_flag}  : 0=no anisotropy / 1,2,.. check with implementation in \texttt{aniso\_model.f90}
-     \item \texttt{domain\_id}        : 1=acoustic / 2=elastic
-\end{itemize}
-\item [{\texttt{NREGIONS}}] The number of regions in the mesh.
-In the following lines, because the mesh is regular or 'almost regular', each region is defined as :
-\begin{lyxcode}
-  \texttt{NEX\_XI\_BEGIN}  \texttt{NEX\_XI\_END}  \texttt{NEX\_ETA\_BEGIN}  \texttt{NEX\_ETA\_END}  \texttt{NZ\_BEGIN}  \texttt{NZ\_END}  \texttt{material\_ID}
-\end{lyxcode}
-\end{description}
-
-The \texttt{INTERFACES\_FILE} parameter of \texttt{Mesh\_Par\_File} defines the file which contains the settings of the topography grid and of the interfaces grids.
-Topography is defined as a set of elevation values on a regular 2D grid.
-It is also possible to
-define interfaces between the layers of the model in the same way.
-The file needs to define several parameters:
-\begin{itemize}
-\item The number of interfaces, including the topography. This needs to be set at the first line. Then, from the bottom
-to the top of the model, you need to define the grids with:
-\item \texttt{SUPPRESS\_UTM\_PROJECTION} flag as described previously,
-\item number of points along $x$ and $y$ direction (NXI and NETA),
-\item minimal $x$ and $y$ coordinates (LONG\_MIN and LAT\_MIN),
-\item spacing between points along $x$ and $y$ (SPACING\_XI and SPACING\_ETA) and
-\item the name of the file which contains the elevation values (in $y$.$x$ increasing order).
-\end{itemize}
-At the end of this file, you simply need to set the number of spectral elements in the vertical direction for each layer. We provide a few models in the {\texttt{examples/}} directory.  \\
-
-Finally, depending on your system, you might need to provide a file that tells MPI what compute nodes
-to use for the simulations. The file must have a number of entries
-(one entry per line) at least equal to the number of processors needed
-for the run. A sample file is provided in the file \texttt{mymachines}.
-This file is not used by the mesher or solver, but is required by
-the \texttt{go\_mesher} and \texttt{go\_solver} default job submission
-scripts. See Chapter~\ref{cha:Scheduler} for information about running
-the code on a system with a scheduler, e.g., LSF.
-
-Now that you have set the appropriate parameters in the \texttt{Mesh\_Par\_file}
-and have compiled the mesher, you are ready to launch it! This is
-most easily accomplished based upon the \texttt{go\_mesher} script.
-When you run on a PC cluster, the script assumes that the nodes are
-named n001, n002, etc. If this is not the case, change the \texttt{tr
--d `n'} line in the script. You may also need to edit the last command
-at the end of the script that invokes the \texttt{mpirun} command.
-See Chapter~\ref{cha:Scheduler} for information about running the
-code on a system with a scheduler, e.g., LSF.
-
-Mesher output is provided in the \texttt{in\_out\_files/OUTPUT\_FILES} directory
-in \texttt{output\_mesher.txt}; this file provides lots of details
-about the mesh that was generated. Please note that the mesher suggests a time step \texttt{DT} to run
-the solver with. The mesher output file also contains a table about the quality of the mesh to indicate
-possible problems with the distortions of elements.
-Alternatively, output can be directed
-to the screen instead by uncommenting a line in \texttt{constants.h}:
-\begin{lyxcode}
-!~uncomment~this~to~write~messages~to~the~screen~
-
-!~integer,~parameter~::~IMAIN~=~ISTANDARD\_OUTPUT
-\end{lyxcode}
-
-To control the quality of the mesh, check the standard output (either on the screen or in the \texttt{in\_out\_files/OUTPUT\_FILES} directory in \texttt{output\_mesher.txt}) and analyze the skewness histogram of mesh elements to make sure no element has a skewness above approximately 0.75, otherwise the mesh is of poor quality (and if even a single element has a skewness value above 0.80, then you must definitely improve the mesh).
-To draw the skewness histogram on the screen, type \texttt{gnuplot plot\_mesh\_quality\_histogram.gnu}.
-
-\chapter{\label{cha:Creating-Distributed-Databases}Creating the Distributed Databases}
-
-After using \texttt{xmeshfem3D} or \texttt{xdecompose\_mesh\_SCOTCH}, the next step in the workflow is to compile
-\texttt{xgenerate\_}~\\
-\texttt{databases}. This program is going to create all the missing information needed by the SEM solver.
-
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=.6\textwidth]{figures/workflow.jpg}
-\par\end{centering}
-
-\caption{Schematic workflow for a SPECFEM3D simulation. The executable \texttt{xgenerate\_databases} creates the GLL mesh points and assigns specific model parameters.}
-
-\label{fig:workflow.databases}
-\end{figure}
-
-In the main directory, type
-\begin{lyxcode}
-{\small make xgenerate\_databases}{\small \par}
-\end{lyxcode}
-Input for the program is provided through
-the main parameter file \texttt{Par\_file}, which resides in the subdirectory \texttt{in\_data\_files/}. Please note that
-\texttt{xgenerate\_databases} must be called directly from the \texttt{bin/} directory, as most of the binaries of the package.
-
-\section{\label{cha:Main-Parameter}Main parameter file \texttt{Par\_file}}
-
-Before running \texttt{xgenerate\_databases}, a number of parameters need to be set in the main parameter \texttt{Par\_file}
-located in the subdirectory \texttt{in\_data\_files/}:
-
-\begin{description}
-\item [{\texttt{SIMULATION\_TYPE}}] is set to 1 for forward simulations,
-2 for adjoint simulations (see Section \ref{sec:Adjoint-simulation-finite})
-and 3 for kernel simulations (see Section \ref{sec:Finite-Frequency-Kernels}).
-\item [{\texttt{SAVE\_FORWARD}}] is only set to \texttt{.true.} for a forward
-simulation with the last frame of the simulation saved, as part of
-the finite-frequency kernel calculations (see Section \ref{sec:Finite-Frequency-Kernels}).
-For a regular forward simulation, leave \texttt{SIMULATION\_TYPE}
-and \texttt{SAVE\_FORWARD} at their default values.
-\item [{\texttt{\noun{UTM\_PROJECTION\_ZONE}}}] UTM projection zone in
-which your model resides, only valid when \texttt{SUPPRESS\_UTM\_}~\\
-\texttt{PROJECTION} is \texttt{.false.}.
-\item [{\texttt{SUPPRESS\_UTM\_PROJECTION}}] set to be \texttt{.false.}
-when your model range is specified in the geographical coordinates,
-and needs to be \texttt{.true.} when your model is specified in a
-cartesian coordinates. \noun{UTM projection zone in which your simulation
-region resides.}
-\item [{\texttt{NPROC}}] The number of MPI processors, each one is assigned one slice of the whole mesh.
-\item [{\texttt{NSTEP}}] The number of time steps of the simulation. This controls the
-length of the numerical simulation, i.e., twice the number of time steps
-requires twice as much CPU time. This feature is not used at the time
-of generating the distributed databases but is required for the solver, i.e., you may change this
-parameter after running  \texttt{xgenerate\_databases}.
-\item [{\texttt{DT}}] The length of each time step in seconds.
-This feature is not used at the time of generating the distributed databases but is required for the solver.
-Please see also Section~\ref{sec:Choosing-the-Time-Step} for further details.
-%
-\item [{\texttt{MODEL}}] Must be set to one of the following:
-%
-\begin{description}
-\item [{\textmd{Models defined by mesh parameters:}}]~
-\begin{description}
-\item [{\texttt{default}}] Uses model parameters as defined by meshing procedures described
-in the previous Chapter~\ref{cha:Mesh-Generation}.
-\end{description}
-\end{description}
-%
-\begin{description}
-\item [{\textmd{1D~models~with~real~structure:}}]~
-\begin{description}
-\item [{\texttt{1D\_prem}}] Isotropic version of the spherically
-symmetric Preliminary Reference Earth Model (PREM) \citep{DzAn81}.
-\item [{\texttt{1D\_socal}}] A standard isotropic 1D model for Southern California.
-\item [{\texttt{1D\_cascadia}}] Isotropic 1D profile for the Cascadia region.
-\end{description}
-\end{description}
-%
-\begin{description}
-\item [{\textmd{Fully~3D~models:}}]~
-\begin{description}
-\item [{\texttt{aniso}}] For a user-specified fully anisotropic model.
-Parameters are set up in routines located in file \texttt{model\_aniso.f90} in directory \texttt{src/generate\_databases/}.
-See Chapter~\ref{cha:-Changing-the} for a discussion on how to specify your own 3D model.
-\item [{\texttt{external}}] For a user-specified isotropic model which uses externally defined model parameters.
-Uses external model definitions set up in routines located in file \texttt{model\_external\_values.f90} in directory \texttt{src/generate\_databases/}.
-Please modify these generic template routines to use your own model definitions.
-\item [{\texttt{gll}}] For a user-specified isotropic model which uses external binary files for $v_p$, $v_s$ and $\rho$.
-Binary files are given in the same format as when outputted by the \texttt{xgenerate\_databases} executable when using option \texttt{SAVE\_MESH\_FILES}.
-These binary files define the model parameters on all GLL points which can be used for iterative inversion procedures.
-\item [{\texttt{salton\_trough}}] A 3D $V_p$ model for Southern California.
-Users must provide the corresponding data file \texttt{regrid3\_vel\_p.bin} in directory \texttt{DATA/st\_3D\_block\_harvard/}.
-\item [{\texttt{tomo}}] For a user-specified 3D isotropic model which uses a tomographic model file \texttt{tomographic\_model.xyz} in directory \texttt{in\_data\_files/}.
-See Section \ref{sec:Using-tomographic},
-for a discussion on how to specify your own 3D tomographic model.
-\end{description}
-\end{description}
-%
-%
-\item [{\texttt{OCEANS}}] Set to \texttt{.true.} if the effect of the oceans
-on seismic wave propagation should be incorporated based upon the
-approximate treatment discussed in \citet{KoTr02b}. This feature
-is inexpensive from a numerical perspective, both in terms of memory
-requirements and CPU time. This approximation is accurate at periods
-of roughly 20~s and longer. At shorter periods the effect of water
-phases/reverberations is not taken into account, even when the flag
-is on.
-\item [{\texttt{TOPOGRAPHY}}] This feature is only effective if \texttt{OCEANS} is set to \texttt{.true.}.
-Set to \texttt{.true.} if topography and
-bathymetry should be read in based upon the topography file specified in the main
-constants file \texttt{constants.h} found in subdirectory \texttt{src/shared/} to evaluate
-elevations. If not set, elevations will be read from the numerical mesh.
-\item [{\texttt{ATTENUATION}}] Set to \texttt{.true.} if attenuation should
-be incorporated. Turning this feature on increases the memory requirements
-significantly (roughly by a factor of~1.5), and is numerically fairly
-expensive. See \citet{KoTr99,KoTr02a} for a discussion on the implementation
-of attenuation based upon standard linear solids.
-Please note that the Vp- and Vs-velocities of your model are given for a reference frequency.
-To change this reference frequency, you change the value of
-\texttt{ATTENUATION\_f0\_REFERENCE} in the main constants file \texttt{constants.h} found in subdirectory \texttt{src/shared/}.
-\item [{\texttt{USE\_OLSEN\_ATTENUATION}}] Set to \texttt{.true.} if you
-want to use the attenuation model that scaled from the S-wave speed model
-using Olsen's empirical relation (see \citet{OlDaBr03}).
-\item [{\texttt{ANISOTROPY}}] Set to \texttt{.true.} if you
-want to use an anisotropy model. Please see the file \texttt{model\_aniso.f90} in subdirectory \texttt{src/generate\_databases/} for the current implementation of anisotropic models.
-\item [{\texttt{ABSORBING\_CONDITIONS}}] Set to \texttt{.true.} to turn
-on Clayton-Enquist absorbing boundary conditions (see \citet{KoTr99}).
-\item [{\texttt{ABSORB\_INSTEAD\_OF\_FREE\_SURFACE}}] Set to \texttt{.true.} to turn on 
-absorbing boundary conditions on the top surface which by default constitutes a free surface of the model.
-\item [{\texttt{MOVIE\_SURFACE}}] Set to \texttt{.false.}, unless you want
-to create a movie of seismic wave propagation on the Earth's surface.
-Turning this option on generates large output files. See Section~\ref{sec:Movies}
-for a discussion on the generation of movies. This feature is only relevant for the solver.
-\item [{\texttt{MOVIE\_VOLUME}}] Set to \texttt{.false.}, unless you want
-to create a movie of seismic wave propagation in the Earth's interior.
-Turning this option on generates huge output files. See Section~\ref{sec:Movies}
-for a discussion on the generation of movies. This feature is only relevant for the solver.
-\item [{\texttt{NTSTEP\_BETWEEN\_FRAMES}}] Determines the number of timesteps
-between movie frames. Typically you want to save a snapshot every
-100 timesteps. The smaller you make this number the more output will
-be generated! See Section~\ref{sec:Movies} for a discussion on the
-generation of movies. This feature is only relevant for the solver.
-\item [{\texttt{CREATE\_SHAKEMAP}}] Set this flag to \texttt{.true.} to
-create a ShakeMap\textregistered, i.e., a peak ground velocity map
-of the maximum absolute value of the two horizontal components of the velocity vector.
-\item [{\texttt{SAVE\_DISPLACEMENT}}] Set this flag to \texttt{.true.}
-if you want to save the displacement instead of velocity for the movie
-frames.
-\item [{\texttt{USE\_HIGHRES\_FOR\_MOVIES}}] Set this flag to \texttt{.true.}
-if you want to save the values at all the NGLL grid points for the
-movie frames.
-\item [{\texttt{HDUR\_MOVIE}}] determines the half duration of the source time function for the movie simulations.
-When this parameter is set to be 0, a default half duration that corresponds to the accuracy of the simulation is provided.
-Otherwise, it adds this half duration to the half duration specified in the source file \texttt{CMTSOLUTION},
-thus simulates longer periods to make the movie images look smoother.
-\item [{\texttt{SAVE\_MESH\_FILES}}] Set this flag to \texttt{.true.} to
-save ParaView \urlwithparentheses{www.paraview.org}
-mesh files for subsequent viewing. Turning the flag on generates large
-(distributed) files in the \texttt{LOCAL\_PATH} directory. See Section~\ref{sec:Mesh-graphics}
-for a discussion of mesh viewing features.
-\item [{\texttt{LOCAL\_PATH}}] Directory in which the distributed databases
-will be written. Generally one uses a directory on the
-local disk of the compute nodes, although on some machines these databases
-are written on a parallel (global) file system (see also the earlier
-discussion of the \texttt{LOCAL\_PATH\_IS\_ALSO\_GLOBAL} flag in Chapter~\ref{cha:Getting-Started}).
-\texttt{xgenerate\_databases} generates the necessary databases in parallel, one set
-for each of the \texttt{NPROC} slices that constitutes
-the mesh (see Figure~\ref{fig:mount.partitions} and Figure~\ref{fig:For-parallel-computing}). After
-the executable finishes, you can log in to one of the compute nodes and view
-the contents of the \texttt{LOCAL\_PATH} directory to see the (many)
-files generated by \texttt{xgenerate\_databases}.
-Please note that the \texttt{LOCAL\_PATH} directory should already contain the output files of
-the partitioner, i.e. from \texttt{xdecompose\_mesh\_SCOTCH} or \texttt{xmeshfem3D}.
-\item [{\texttt{NTSTEP\_BETWEEN\_OUTPUT\_INFO}}] This parameter specifies
-the interval at which basic information about a run is written to
-the file system (\texttt{timestamp{*}} files in the \texttt{in\_out\_files/OUTPUT\_FILES}
-directory). If you have access to a fast machine, set \texttt{NTSTEP\_BETWEEN\_OUTPUT\_INFO}
-to a relatively high value (e.g., at least 100, or even 1000 or more)
-to avoid writing output text files too often. This feature is not
-used at the time of meshing. One can set this parameter to a larger
-value than the number of time steps to avoid writing output during
-the run.
-\item [{\texttt{NTSTEP\_BETWEEN\_OUTPUT\_SEISMOS}}] This parameter specifies
-the interval at which synthetic seismograms are written in the \texttt{LOCAL\_PATH}
-directory. If a run crashes, you may still find usable (but shorter
-than requested) seismograms in this directory. On a fast machine set
-\texttt{NTSTEP\_BETWEEN\_OUTPUT\_SEISMOS} to a relatively high value
-to avoid writing to the seismograms too often. This feature is only relevant for the solver.
-\item [{\texttt{USE\_FORCE\_POINT\_SOURCE}}] Turn this flag on to use a \texttt{FORCESOLUTION} force source
-instead of a \texttt{CMTSOLUTION} moment-tensor source. This can be useful e.g. for oil industry 
-foothills simulations in which the source is a vertical force, normal force, 
-tilted force, or an impact etc. You will need to give the East, North and vertical components of an 
-arbitrary (non-unitary) direction vector of the force vector; thus refer to Appendix~\ref{cha:Coordinates} 
-for the orientation of the reference frame.
-This vector is made unitary internally in the code and thus only its direction matters here;
-its norm is ignored and the norm of the force used is the factor force source times the source time function.
-\item [{\texttt{PRINT\_SOURCE\_TIME\_FUNCTION}}] Turn this flag on to print
-information about the source time function in the file \texttt{in\_out\_files/OUTPUT\_FILES/plot\_source\_time\_function.txt}.
-This feature is only relevant for the solver.
-\end{description}
-
-\section{\label{sec:Choosing-the-Time-Step}Choosing the time step \texttt{DT}}
-
-The parameter \texttt{DT} sets the length of each time step in seconds. The value of this parameter is crucial for the stability of
-the spectral-element simulation.
-Your time step \texttt{DT} will depend on the minimum ratio between the distance $h$ of neighboring mesh points and the wave speeds $v$ defined in your model.
-The condition for the time step $\Delta t$ is:
-
-\begin{lyxcode}
-$\Delta t < C ~\mathrm{min}_\Omega (~ h / v ~)$
-\end{lyxcode}
-where $C$ is the so-called Courant number and $\Omega$ denotes the model volume.
-The distance $h$ depends on the mesh element size and the number of
-GLL points \texttt{NGLL} specified in the main constants file \texttt{constants.h} located in the \texttt{src/shared/} subdirectory.
-The wave speed $v$ is determined based on your model's P- (or S-) wave speed values.
-
-The database generator \texttt{xgenerate\_databases}, as well as the internal mesher \texttt{xmeshfem3D}, are trying
-to evaluate the value of $\Delta t$ for empirically chosen Courant numbers $C \sim 0.3$.
-If you used the mesher \texttt{xmeshfem3D} to generate your mesh, you should set the value suggested in \texttt{in\_out\_files/OUTPUT\_FILES/output\_mesher.txt} file, which is created after the mesher completed.
-In case you used CUBIT to create the mesh, you might use an arbitrary value when running \texttt{xgenerate\_databases} and then use the value suggested in the ~\\
-\texttt{in\_out\_files/OUTPUT\_FILES/output\_mesher.txt} file after the database generation completed.
-Note that the implemented Newmark time scheme uses this time step globally, thus your simulations become more expensive for very small mesh elements in high wave-speed regions.
-Please be aware of this restriction when constructing your mesh in Chapter~\ref{cha:Mesh-Generation}.
-
-\chapter{\label{cha:Running-the-Solver}Running the Solver \texttt{xspecfem3D}}
-
-Now that you have successfully generated the databases, you are ready to compile
-the solver. In the main directory, type
-\begin{lyxcode}
-{\small make xspecfem3D}{\small \par}
-\end{lyxcode}
-Please note that \texttt{xspecfem3D} must be called directly from
-the \texttt{bin/} directory, as most of the binaries of the package.
-
-The solver needs three input files in the \texttt{in\_data\_files/} directory
-to run:
-\begin{description}
-\item [{\texttt{Par\_file}}] the main parameter file which was discussed in detail in the previous
-Chapter~\ref{cha:Creating-Distributed-Databases},
-\item [{\texttt{CMTSOLUTION}} or {\texttt{FORCESOLUTION}}] the earthquake source parameter file or the force source parameter file or, and
-\item [{\texttt{STATIONS}}] the stations file.
-\end{description}
-
-Most parameters in the \texttt{Par\_file}
-should be set prior to running the databases generation. Only the following parameters
-may be changed after running \texttt{xgenerate\_databases}:
-
-\begin{itemize}
-\item the simulation type control parameters: \texttt{SIMULATION\_TYPE}
-and \texttt{SAVE\_FORWARD}
-\item the time step parameters \texttt{NSTEP} and \texttt{DT}
-\item the absorbing boundary control parameter \texttt{ABSORBING\_CONDITIONS} on condition that the\\ 
-  \texttt{ABSORB\_INSTEAD\_OF\_FREE\_SURFACE} flag remains unmodified after running the databases generation. 
-\item the movie control parameters \texttt{MOVIE\_SURFACE}, \texttt{MOVIE\_VOLUME},
-and \texttt{NTSTEPS\_BETWEEN\_FRAMES}
-\item the ShakeMap\textregistered option \texttt{CREATE\_SHAKEMAP}
-\item the output information parameters \texttt{NTSTEP\_BETWEEN\_OUTPUT\_INFO}
-and \texttt{NTSTEP\_BETWEEN\_OUTPUT\_}~\\
-\texttt{SEISMOS}
-\item the \texttt{PRINT\_SOURCE\_TIME\_FUNCTION} flags
-\end{itemize}
-Any other change to the \texttt{Par\_file} implies rerunning both
-the database generator \texttt{xgenerate\_databases} and the solver \texttt{xspecfem3D}.
-
-For any particular earthquake, the \texttt{CMTSOLUTION} file that
-represents the point source may be obtained directly from the Harvard Centroid-Moment Tensor (CMT) web page \urlwithparentheses{www.seismology.harvard.edu}.
-It looks like this:
-
-\begin{lyxcode}
-{\small }%
-\begin{figure}[H]
-\noindent \begin{centering}
-{\small \includegraphics[width=1\textwidth]{figures/Hollywood_CMT} }
-\par\end{centering}{\small \par}
-
-\caption{\label{fig:CMTSOLUTION-file}\texttt{CMTSOLUTION} file based on the
-format from the Harvard CMT catalog. \textbf{M} is the moment tensor,
-$M_{0}${\small{} }is the seismic moment, and $M_{w}$ is the moment
-magnitude.}
-
-\end{figure}
-{\small \par}
-\end{lyxcode}
-The \texttt{CMTSOLUTION} should be edited in the following way:
-
-\begin{itemize}
-\item Set the \texttt{time shift} parameter equal to $0.0$ (the solver
-will not run otherwise.) The time shift parameter would simply apply
-an overall time shift to the synthetics, something that can be done
-in the post-processing (see Section \ref{sec:Process-data-and-syn}).
-\item For point-source simulations (see finite sources, page \pageref{To-simulate-a})
-we recommend setting the source half-duration parameter \texttt{half
-duration} equal to zero, which corresponds to simulating a step source-time
-function, i.e., a moment-rate function that is a delta function. If
-\texttt{half duration} is not set to zero, the code will use a Gaussian
-(i.e., a signal with a shape similar to a `smoothed triangle', as
-explained in \citet{KoTr02a} and shown in Fig~\ref{fig:gauss.vs.triangle})
-source-time function with half-width \texttt{half duration}. We prefer
-to run the solver with \texttt{half duration} set to zero and convolve
-the resulting synthetic seismograms in post-processing after the run,
-because this way it is easy to use a variety of source-time functions
-(see Section \ref{sec:Process-data-and-syn}). \citet{KoTr02a} determined
-that the noise generated in the simulation by using a step source
-time function may be safely filtered out afterward based upon a convolution
-with the desired source time function and/or low-pass filtering. Use
-the serial code \texttt{convolve\_source\_timefunction.f90} and the
-script \texttt{convolve\_source\_timefunction.csh} for this purpose,
-or alternatively use signal-processing software packages such as SAC \urlwithparentheses{www.llnl.gov/sac}.
-Type
-\begin{lyxcode}
-{\small make xconvolve\_source\_timefunction}{\small \par}
-\end{lyxcode}
-to compile the code and then set the parameter \texttt{hdur} in \texttt{convolve\_source\_timefunction.csh}
-to the desired half-duration.
-
-\item The zero time of the simulation corresponds to the center of the triangle/Gaussian,
-or the centroid time of the earthquake. The start time of the simulation
-is $t=-1.5*\texttt{half duration}$ (the 1.5 is to make sure the moment
-rate function is very close to zero when starting the simulation).
-To convert to absolute time $t_{\mathrm{abs}}$, set
-
-\begin{lyxcode}
-$t_{\mathrm{abs}}=t_{\mathrm{pde}}+\texttt{time shift}+t_{\mathrm{synthetic}}$
-\end{lyxcode}
-where $t_{\mathrm{pde}}$ is the time given in the first line of the
-\texttt{CMTSOLUTION}, \texttt{time shift} is the corresponding value
-from the original \texttt{CMTSOLUTION} file and $t_{\mathrm{synthetic}}$
-is the time in the first column of the output seismogram.
-
-\end{itemize}
-%
-\begin{figure}
-\noindent \begin{centering}
-%%%\includegraphics[width=3in]{figures/gauss_vs_triangle_mod.eps}
-\includegraphics[width=3in]{figures/gauss_vs_triangle_mod.pdf}
-\par\end{centering}
-
-\caption{Comparison of the shape of a triangle and the Gaussian function actually
-used.}
-
-\label{fig:gauss.vs.triangle}
-\end{figure}
-
-Centroid latitude and longitude should be provided in geographical
-coordinates. The code converts these coordinates to geocentric coordinates~\citep{DaTr98}.
-Of course you may provide your own source representations by designing
-your own \texttt{CMTSOLUTION} file. Just make sure that the resulting
-file adheres to the Harvard CMT conventions (see Appendix~\ref{cha:Coordinates}).
-Note that the first line in the \texttt{CMTSOLUTION} file is the Preliminary Determination of Earthquakes (PDE) solution performed by the USGS NEIC, which is used as a seed for the Harvard CMT inversion.
-The PDE solution is based upon P waves and often gives the hypocenter of the earthquake, i.e., the rupture initiation point, whereas the CMT solution gives the `centroid location', which is the location with dominant moment release.
-The PDE solution is not used by our software package but must be present anyway in the first line of the file.
-
-\label{To-simulate-a}To simulate a kinematic rupture, i.e., a finite-source
-event, represented in terms of $N_{\mathrm{sources}}$ point sources,
-provide a \texttt{CMTSOLUTION} file that has $N_{\mathrm{sources}}$
-entries, one for each subevent (i.e., concatenate $N_{\mathrm{sources}}$
-\texttt{CMTSOLUTION} files to a single \texttt{CMTSOLUTION} file).
-At least one entry (not necessarily the first) must have a zero \texttt{time
-shift}, and all the other entries must have non-negative \texttt{time
-shift}. Each subevent can have its own half duration, latitude, longitude,
-depth, and moment tensor (effectively, the local moment-density tensor).
-
-Note that the zero in the synthetics does NOT represent the hypocentral
-time or centroid time in general, but the timing of the \textit{center}
-of the source triangle with zero \texttt{time shift} (Fig~\ref{fig:source_timing}).
-
-Although it is convenient to think of each source as a triangle, in
-the simulation they are actually Gaussians (as they have better frequency
-characteristics). The relationship between the triangle and the gaussian
-used is shown in Fig~\ref{fig:gauss.vs.triangle}. For finite fault
-simulations it is usually not advisable to use a zero half duration
-and convolve afterwards, since the half duration is generally fixed
-by the finite fault model.
-
-\vspace{1cm}
-
-\noindent The \texttt{FORCESOLUTION} should be edited in the following way:
-
-\begin{itemize}
-\item Set the \texttt{time shift} parameter equal to $0.0$ (the solver
-will not run otherwise.) The time shift parameter would simply apply
-an overall time shift to the synthetics, something that can be done
-in the post-processing (see Section \ref{sec:Process-data-and-syn}).
-\item Set the latitude, longitude, depth of the source in geographical
-  coordinates.
-\item Set the magnitude of the force source.
-\item Set the components of a (non-unitary) direction vector for the
-  force source in the East/North/Vertical basis (see Appendix A for
-  the orientation of the reference frame).
-\end{itemize}
-
-{\small }%
-\begin{figure}[H]
-\noindent \begin{centering}
-{\small \includegraphics[width=5in]{figures/source_timing.pdf} }
-\par\end{centering}{\small \par}
-
-\caption{Example of timing for three sources. The center of the first source
-triangle is defined to be time zero. Note that this is NOT in general
-the hypocentral time, or the start time of the source (marked as tstart).
-The parameter \texttt{time shift} in the \texttt{CMTSOLUTION} file
-would be t1(=0), t2, t3 in this case, and the parameter \texttt{half
-duration} would be hdur1, hdur2, hdur3 for the sources 1, 2, 3 respectively.}
-
-{\small \label{fig:source_timing} }
-\end{figure}
-{\small \par}
-
-The solver can calculate seismograms at any number of stations for
-basically the same numerical cost, so the user is encouraged to include
-as many stations as conceivably useful in the \texttt{STATIONS} file,
-which looks like this:
-
-{\small }%
-\begin{figure}[H]
-\noindent \begin{centering}
-{\small \includegraphics{figures/STATIONS_basin_explained} }
-\par\end{centering}{\small \par}
-
-\caption{Sample \texttt{STATIONS} file. Station latitude and longitude should
-be provided in geographical coordinates. The width of the station
-label should be no more than 32 characters (see \texttt{MAX\_LENGTH\_STATION\_NAME}
-in the \texttt{constants.h} file), and the network label should be
-no more than 8 characters (see \texttt{MAX\_LENGTH\_NETWORK\_NAME}
-in the \texttt{constants.h} file).}
-
-\end{figure}
-{\small \par}
-
-Each line represents one station in the following format:
-
-\begin{lyxcode}
-{\small Station~Network~Latitude~(degrees)~Longitude~(degrees)~Elevation~(m)~burial~(m)~}{\small \par}
-\end{lyxcode}
-The solver \texttt{xspecfem3D} filters the list of stations in file \texttt{in\_data\_files/STATIONS}
-to exclude stations that are not located within the region given in
-the \texttt{Par\_file} (between \texttt{LATITUDE\_MIN} and \texttt{LATITUDE\_MAX}
-and between \texttt{LONGITUDE\_MIN} and \texttt{LONGITUDE\_MAX}).
-The filtered file is called \texttt{in\_data\_files/STATIONS\_FILTERED}.
-
-Solver output is provided in the \texttt{in\_out\_files/OUTPUT\_FILES} directory
-in the \texttt{output\_solver.txt} file. Output can be directed to
-the screen instead by uncommenting a line in \texttt{constants.h}:
-
-\begin{lyxcode}
-!~uncomment~this~to~write~messages~to~the~screen~
-
-!~integer,~parameter~::~IMAIN~=~ISTANDARD\_OUTPUT~
-\end{lyxcode}
-On PC clusters the seismogram files are generally written to the local
-disks (the path \texttt{LOCAL\_PATH} in the \texttt{Par\_file}) and
-need to be gathered at the end of the simulation.
-
-While the solver is running, its progress may be tracked by monitoring
-the `\texttt{\small timestamp{*}}' files in the ~\\
-\texttt{\small in\_out\_files/OUTPUT\_FILES} directory.
-These tiny files look something like this:
-
-\begin{lyxcode}
-Time~step~\#~~~~~~~~~~10000~
-
-Time:~~~~~108.4890~~~~~~seconds~
-
-Elapsed~time~in~seconds~=~~~~1153.28696703911~
-
-Elapsed~time~in~hh:mm:ss~=~~~~~0~h~19~m~13~s~
-
-Mean~elapsed~time~per~time~step~in~seconds~=~~~~~0.115328696703911~
-
-Max~norm~displacement~vector~U~in~all~slices~(m)~=~~~~1.0789589E-02~
-\end{lyxcode}
-The \texttt{\small timestamp{*}} files provide the \texttt{Mean elapsed
-time per time step in seconds}, which may be used to assess performance
-on various machines (assuming you are the only user on a node), as
-well as the \texttt{Max norm displacement vector U in all slices~(m)}.
-If something is wrong with the model, the mesh, or the source, you
-will see the code become unstable through exponentially growing values
-of the displacement and fluid potential with time, and ultimately
-the run will be terminated by the program. You can control the rate
-at which the timestamp files are written based upon the parameter
-\texttt{NTSTEP\_BETWEEN\_OUTPUT\_INFO} in the \texttt{Par\_file}.
-
-Having set the \texttt{Par\_file} parameters, and having provided the
-\texttt{CMTSOLUTION} (or the \texttt{FORCESOLUTION}) and
-\texttt{STATIONS} files, you are now ready to launch the solver! This
-is most easily accomplished based upon the \texttt{go\_solver} script
-(See Chapter \ref{cha:Scheduler} for information about running through
-a scheduler, e.g., LSF). You may need to edit the last command at the
-end of the script that invokes the \texttt{mpirun} command. The
-\texttt{runall} script compiles and runs both
-\texttt{xgenerate\_databases} and \texttt{xspecfem3D} in
-sequence. This is a safe approach that ensures using the correct
-combination of distributed database output and solver input.
-
-It is important to realize that the CPU and memory requirements of
-the solver are closely tied to choices about attenuation (\texttt{ATTENUATION})
-and the nature of the model (i.e., isotropic models are cheaper than
-anisotropic models). We encourage you to run a variety of simulations
-with various flags turned on or off to develop a sense for what is
-involved.
-
-For the same model, one can rerun the solver for different events by
-simply changing the \texttt{CMTSOLUTION} or \texttt{FORCESOLUTION}
-file, or for different stations by changing the \texttt{STATIONS}
-file. There is no need to rerun the \texttt{xgenerate\_databases}
-executable. Of course it is best to include as many stations as
-possible, since this does not add to the cost of the simulation.
-
-\chapter{\label{cha:Adjoint-Simulations}Adjoint Simulations}
-
-Adjoint simulations are generally performed for two distinct applications.
-First, they can be used for earthquake source inversions, especially
-earthquakes with large ruptures such as the Lander's earthquake \citep{WaHe94}.
-Second, they can be used to generate finite-frequency sensitivity
-kernels that are a critical part of tomographic inversions based upon
-3D reference models \citep{TrTaLi05,LiTr06,TrKoLi08,LiTr08}. In either
-case, source parameter or velocity structure updates are sought to
-minimize a specific misfit function (e.g., waveform or traveltime
-differences), and the adjoint simulation provides a means of computing
-the gradient of the misfit function and further reducing it in successive
-iterations. Applications and procedures pertaining to source studies
-and finite-frequency kernels are discussed in Sections~\ref{sec:Adjoint-simulation-sources}
-and \ref{sec:Adjoint-simulation-finite}, respectively. The two related
-parameters in the \texttt{Par\_file} are \texttt{SIMULATION\_TYPE}
-(1 or 2) and the \texttt{SAVE\_FORWARD} (boolean).
-
-\section{\label{sec:Adjoint-simulation-sources}Adjoint Simulations for Sources}
-
-In the case where a specific misfit function is minimized to invert
-for the earthquake source parameters, the gradient of the misfit function
-with respect to these source parameters can be computed by placing
-time-reversed seismograms at the receivers and using them as sources
-in an adjoint simulation, and then the value of the gradient is obtained
-from the adjoint seismograms recorded at the original earthquake location.
-
-\begin{enumerate}
-\item \textbf{Prepare the adjoint sources} \label{enu:Prepare-the-adjoint}
-
-\begin{enumerate}
-\item First, run a regular forward simlation (\texttt{SIMULATION\_TYPE =
-1} and \texttt{SAVE\_FORWARD = .false.}). You can automatically set
-these two variables using the \texttt{\small utils/change\_simulation\_type.pl}
-script:
-
-\begin{lyxcode}
-utils/change\_simulation\_type.pl~-f~
-\end{lyxcode}
-and then collect the recorded seismograms at all the stations given
-in \texttt{in\_data\_files/STATIONS}.
-
-\item Then select the stations for which you want to compute the time-reversed
-adjoint sources and run the adjoint simulation, and compile them into
-the \texttt{in\_data\_files/STATIONS\_ADJOINT} file, which has the same format
-as the regular \texttt{in\_data\_files/STATIONS} file.
-
-\begin{itemize}
-\item Depending on what type of misfit function is used for the source inversion,
-adjoint sources need to be computed from the original recorded seismograms
-for the selected stations and saved in the \texttt{in\_out\_files/SEM/} directory
-with the format \texttt{STA.NT.BX?.adj}, where \texttt{STA}, \texttt{NT}
-are the station name and network code given in the \texttt{in\_data\_files/STATIONS\_ADJOINT}
-file, and \texttt{BX?} represents the component name of a particular
-adjoint seismogram. Please note that the band code can change depending on
-your sampling rate (see Appendix~\ref{cha:channel-codes} for further details).
-\item The adjoint seismograms are in the same format as the original seismogram
-(\texttt{STA.NT.BX?.sem?}), with the same start time, time interval
-and record length.
-\end{itemize}
-\item Notice that even if you choose to time reverse only one component
-from one specific station, you still need to supply all three components
-because the code is expecting them (you can set the other two components
-to be zero).
-\item Also note that since time-reversal is done in the code itself, no
-explicit time-reversing is needed for the preparation of the adjoint
-sources, i.e., the adjoint sources are in the same forward time sense
-as the original recorded seismograms.
-\end{enumerate}
-\item \textbf{Set the related parameters and run the adjoint simulation}\\
-In the \texttt{in\_data\_files/Par\_file}, set the two related parameters to
-be \texttt{SIMULATION\_TYPE = 2} and \texttt{SAVE\_FORWARD = .false.}.
-More conveniently, use the scripts \texttt{utils/change\_simulation\_type.pl}
-to modify the \texttt{Par\_file} automatically (\texttt{change\_simulation\_type.pl
--a}). Then run the solver to launch the adjoint simulation.
-\item \textbf{Collect the seismograms at the original source location}
-
-After the adjoint simulation has completed successfully, collect the
-seismograms from \texttt{LOCAL\_PATH}.
-
-\begin{itemize}
-\item These adjoint seismograms are recorded at the locations of the original
-earthquake sources given by the \texttt{in\_data\_files/CMTSOLUTION} file, and
-have names of the form \texttt{S?????.NT.S??.sem} for the six-component
-strain tensor (\texttt{SNN,SEE,SZZ,SNE,SNZ,SEZ}) at these locations,
-and ~\\
-\texttt{S?????.NT.BX?.sem} for the three-component displacements
-(\texttt{BXN,BXE,BXZ}) recorded at these locations.
-\item \texttt{S?????} denotes the source number; for example, if the original
-\texttt{CMTSOLUTION} provides only a point source, then the seismograms
-collected will start with \texttt{S00001}.
-\item These adjoint seismograms provide critical information for the computation
-of the gradient of the misfit function.
-\end{itemize}
-\end{enumerate}
-
-\section{\label{sec:Adjoint-simulation-finite}Adjoint Simulations for Finite-Frequency
-Kernels (Kernel Simulation)}
-
-Finite-frequency sensitivity kernels are computed in two successive
-simulations (please refer to \citet{LiTr06} and \citet{TrKoLi08} for details).
-
-\begin{enumerate}
-\item \textbf{Run a forward simulation with the state variables saved at
-the end of the simulation}
-
-Prepare the \texttt{\small CMTSOLUTION} and \texttt{\small STATIONS}
-files, set the parameters \texttt{\small SIMULATION\_TYPE}{\small{}
-}\texttt{\small =}{\small{} }\texttt{\small 1} and \texttt{\small SAVE\_FORWARD
-=}{\small{} }\texttt{\small .true.} in the \texttt{Par\_file} (\texttt{change\_simulation\_type
--F}), and run the solver.
-
-\begin{itemize}
-\item Notice that attenuation is not implemented yet for the computation
-of finite-frequency kernels; therefore set \texttt{ATTENUATION = .false.}
-in the \texttt{Par\_file}.
-\item We also suggest you modify the half duration of the \texttt{CMTSOLUTION}
-to be similar to the accuracy of the simulation (see Equation \ref{eq:shortest_period})
-to avoid too much high-frequency noise in the forward wavefield, although
-theoretically the high-frequency noise should be eliminated when convolved
-with an adjoint wavefield with the proper frequency content.
-\item This forward simulation differs from the regular simulations (\texttt{\small SIMULATION\_TYPE}{\small{}
-}\texttt{\small =}{\small{} }\texttt{\small 1} and \texttt{\small SAVE\_FORWARD}{\small{}
-}\texttt{\small =}{\small{} }\texttt{\small .false.}) described in
-the previous chapters in that the state variables for the last time
-step of the simulation, including wavefields of the displacement,
-velocity, acceleration, etc., are saved to the \texttt{LOCAL\_PATH}
-to be used for the subsequent simulation.
-\item For regional simulations, the files recording the absorbing boundary
-contribution are also written to the \texttt{LOCAL\_PATH} when \texttt{SAVE\_FORWARD
-= .true.}.
-\end{itemize}
-\item \textbf{Prepare the adjoint sources}
-
-The adjoint sources need to be prepared the same way as described
-in the Section \ref{enu:Prepare-the-adjoint}.
-
-\begin{itemize}
-\item In the case of travel-time finite-frequency kernel for one source-receiver
-pair, i.e., point source from the \texttt{CMTSOLUTION}, and one station
-in the \texttt{STATIONS\_ADJOINT} list, we supply a sample program
-in \texttt{utils/adjoint\_sources/traveltime/xcreate\_adjsrc\_traveltime} to cut a certain portion of the original
-displacement seismograms and convert it into the proper adjoint source
-to compute the finite-frequency kernel.
-\begin{lyxcode}
-xcreate\_adjsrc\_traveltime~t1~t2~ifile{[}0-5]~E/N/Z-ascii-files~{[}baz]
-\end{lyxcode}
-where \texttt{t1} and \texttt{t2} are the start and end time of the
-portion you are interested in, \texttt{ifile} denotes the component
-of the seismograms to be used (0 for all three components, 1 for East,
-2 for North, and 3 for vertical, 4 for transverse, and 5 for radial
-component), \texttt{E/N/Z-ascii-files} indicate the three-component
-displacement seismograms in the right order, and \texttt{baz} is the
-back-azimuth of the station. Note that \texttt{baz} is only supplied
-when \texttt{ifile} = 4 or 5.
-
-\item Similarly, a sample program to compute adjoint sources for amplitude finite-frequency kernels
-may be found in \texttt{utils/adjoint\_sources/amplitude} and used in the same way as described for
-traveltime measurements
-\begin{lyxcode}
-xcreate\_adjsrc\_amplitude~t1~t2~ifile{[}0-5]~E/N/Z-ascii-files~{[}baz].
-\end{lyxcode}
-
-\end{itemize}
-
-\item \textbf{Run the kernel simulation}
-
-With the successful forward simulation and the adjoint source ready
-in the \texttt{in\_out\_files/SEM/} directory, set \texttt{SIMULATION\_TYPE = 3} and \texttt{SAVE\_FORWARD
-= .false.} in the \texttt{Par\_file} (you can use \texttt{change\_simulation\_type.pl -b}),
-and rerun the solver.
-
-\begin{itemize}
-\item The adjoint simulation is launched together with the back reconstruction
-of the original forward wavefield from the state variables saved from
-the previous forward simulation, and the finite-frequency kernels
-are computed by the interaction of the reconstructed forward wavefield
-and the adjoint wavefield.
-\item The back-reconstructed seismograms at the original station locations
-are saved to the \texttt{LOCAL\_PATH} at the end of the kernel simulations,
-and can be collected to the local disk.
-\item These back-constructed seismograms can be compared with the time-reversed
-original seismograms to assess the accuracy of the backward reconstruction,
-and they should match very well.
-\item The arrays for density, P-wave speed and S-wave speed kernels are
-also saved in the \texttt{LOCAL\_PATH} with the names \texttt{proc??????\_rho(alpha,beta)\_kernel.bin},
-where \texttt{proc??????} represents the processor number, \texttt{rho(alpha,beta)}
-are the different types of kernels.
-\end{itemize}
-\end{enumerate}
-In general, the three steps need to be run sequentially to assure
-proper access to the necessary files. If the simulations are run through
-some cluster scheduling system (e.g., LSF), and the forward simulation
-and the subsequent kernel simulations cannot be assigned to the same
-set of computer nodes, the kernel simulation will not be able to access
-the database files saved by the forward simulation. Solutions for
-this dilemma are provided in Chapter~\ref{cha:Scheduler}. Visualization
-of the finite-frequency kernels is discussed in Section~\ref{sec:Finite-Frequency-Kernels}.
-
-\chapter{Noise Cross-correlation Simulations}
-
-Besides earthquake simulations, SPECFEM3D includes functionality for seismic noise tomography as well.
-In order to proceed successfully in this chapter, it is critical that you have already familiarized yourself
-with procedures for meshing (Chapter \ref{cha:Mesh-Generation}),
-creating distributed databases (Chapter \ref{cha:Creating-Distributed-Databases}),
-running earthquake simulations (Chapters \ref{cha:Running-the-Solver})
-and adjoint simulations (Chapter \ref{cha:Adjoint-Simulations}).
-Also, make sure you read the article `Noise cross-correlation sensitivity kernels' \citep{trompetal2010},
-in order to understand noise simulations from a theoretical perspective.
-
-\section{Input Parameter Files}
-
-As usual, the three main input files are crucial: \texttt{Par\_file}, \texttt{CMTSOLUTION} and \texttt{STATIONS}.
-Unless otherwise specified, those input files should be located in directory \texttt{in\_data\_files/}.\\
-
-\texttt{CMTSOLUTION} is required for all simulations.
-At a first glance, it may seem unexpected to have it here,
-since the noise simulations should have nothing to do with the earthquake -- \texttt{CMTSOLUTION}.
-However, for noise simulations, it is critical to have no earthquakes.
-In other words, the moment tensor specified in \texttt{CMTSOLUTION} must be set to zero manually!\\
-
-\texttt{STATIONS} remains the same as in previous earthquake simulations,
-except that the order of receivers listed in \texttt{STATIONS} is now important.
-The order will be used to determine the `master' receiver,
-i.e., the one that simultaneously cross correlates with the others.\\
-
-\texttt{Par\_file} also requires careful attention.
-A parameter called \texttt{NOISE\_TOMOGRAPHY} has been added which specifies the type of simulation to be run.
-\texttt{NOISE\_TOMOGRAPHY} is an integer with possible values 0, 1, 2 and 3.
-For example, when \texttt{NOISE\_TOMOGRAPHY} equals 0, a regular earthquake simulation will be run.
-When it is 1/2/3, you are about to run
-step 1/2/3 of the noise simulations respectively.
-Should you be confused by the three steps, refer to \citet{trompetal2010} for details.\\
-
-Another change to \texttt{Par\_file} involves the parameter \texttt{NSTEP}.
-While for regular earthquake simulations this parameter specifies the length of synthetic seismograms generated,
-for noise simulations it specifies the length of the seismograms used to compute cross correlations.
-The actual cross correlations are thus twice this length, i.e., $2~\mathrm{NSTEP}-1$.
-The code automatically makes the modification accordingly, if \texttt{NOISE\_TOMOGRAPHY} is not zero.\\
-
-There are other parameters in \texttt{Par\_file} which should be given specific values.
-For instance, since the first two steps for calculating noise sensitivity kernels correspond to
-forward simulations, \texttt{SIMULATION\_TYPE} must be 1 when \texttt{NOISE\_TOMOGRAPHY} equals 1 or 2.
-Also, we have to reconstruct the ensemble forward wavefields in adjoint simulations,
-therefore we need to set \texttt{SAVE\_FORWARD} to \texttt{.true.} for the second step,
-i.e., when \texttt{NOISE\_TOMOGRAPHY} equals 2.
-The third step is for kernel constructions. Hence \texttt{SIMULATION\_TYPE} should be 3, whereas
-\texttt{SAVE\_FORWARD} must be \texttt{.false.}.\\
-
-Finally, for most system architectures,
-please make sure that \texttt{LOCAL\_PATH} in \texttt{Par\_file} is in fact local, not globally shared.
-Because we have to save the wavefields at the earth's surface at every time step,
-it is quite problematic to have a globally shared \texttt{LOCAL\_PATH},
-in terms of both disk storage and I/O speed.
-
-\section{Noise Simulations: Step by Step}
-
-Proper parameters in those parameter files are not enough for noise simulations to run.
-We have more parameters to specify:
-for example, the ensemble-averaged noise spectrum, the noise distribution etc.
-However, since there are a few `new' files, it is better to introduce them sequentially.
-In this section, standard procedures for noise simulations are described.
-
-\subsection{Pre-simulation}
-
-\begin{itemize}
-
-\item
-As usual, we first configure the software package using:
-
-\texttt{./configure FC=ifort MPIFC=mpif90}
-
-Use the following if SCOTCH is needed:
-
-\texttt{./configure FC=ifort MPIFC=mpif90 --with-scotch-dir=/opt/scotch}\\
-
-\item
-Next, we need to compile the source code using:
-
-\texttt{make xgenerate\_databases}
-
-\texttt{make xspecfem3D} \\
-
-\item
-Before we can run noise simulations, we have to specify the noise statistics,
-e.g., the ensemble-averaged noise spectrum.
-Matlab scripts are provided to help you to generate the necessary file.
-
-\texttt{examples/noise\_tomography/NOISE\_TOMOGRAPHY.m}  (main program)\\
-\texttt{examples/noise\_tomography/PetersonNoiseModel.m}
-
-In Matlab, simply run:
-
-\texttt{NOISE\_TOMOGRAPHY(NSTEP, DT, Tmin, Tmax, NOISE\_MODEL)}\\
-
-\texttt{DT} is given in \texttt{Par\_file},
-but \texttt{NSTEP} is NOT the one specified in \texttt{Par\_file}.
-Instead, you have to feed $2~\mathrm{NSTEP}-1$ to account for the doubled length of cross correlations.
-\texttt{Tmin} and \texttt{Tmax} correspond to the period range you are interested in,
-whereas \texttt{NOISE\_MODEL} denotes the noise model you will be using.
-Details can be found in the Matlab script.
-
-After running the Matlab script, you will be given the following information (plus a figure in Matlab):
-
-\texttt{*************************************************************}\\
-\texttt{the source time function has been saved in:}\\
-\texttt{/data2/yangl/3D\_NOISE/S\_squared} (note this path must be different)\\
-\texttt{S\_squared should be put into directory:}\\
-\texttt{in\_out\_files/NOISE\_TOMOGRAPHY/ in the SPECFEM3D package}\\
-
-In other words, the Matlab script creates a file called \texttt{S\_squared},
-which is the first `new' input file we encounter for noise simulations.
-
-One may choose a flat noise spectrum rather than Peterson's noise model.
-This can be done easily by modifying the Matlab script a little.
-
-\item
-Create a new directory in the SPECFEM3D package, name it as \texttt{in\_out\_files/NOISE\_TOMOGRAPHY/}.
-We will add some parameter files later in this folder.
-
-\item
-Put the Matlab-generated-file \texttt{S\_squared} in \texttt{in\_out\_files/NOISE\_TOMOGRAPHY/}.
-
-That's to say, you will have a file \texttt{in\_out\_files/NOISE\_TOMOGRAPHY/S\_squared} in the SPECFEM3D package.
-
-\item
-Create a file called \texttt{in\_out\_files/NOISE\_TOMOGRAPHY/irec\_master\_noise}. Note that this file is located
-in directory \texttt{in\_out\_files/NOISE\_TOMOGRAPHY/} as well. In general, all noise simulation related parameter
-files go into that directory.
-\texttt{irec\_master\_noise} contains only one integer, which is the ID
-of the `master' receiver. For example, if this file contains 5, it means that the fifth receiver
-listed in \texttt{in\_data\_files/STATIONS} becomes the `master'. That's why we mentioned previously that
-the order of receivers in \texttt{in\_data\_files/STATIONS} is important.
-
-Note that in regional simulations, the \texttt{in\_data\_files/STATIONS}
-might contain receivers which are outside of our computational domains.
-Therefore, the integer in \texttt{irec\_master\_noise}
-is actually the ID in \texttt{in\_data\_files/STATIONS\_FILTERED}
-(which is generated by \texttt{bin/xgenerate\_databases}).
-
-\item
-Create a file called \texttt{in\_out\_files/NOISE\_TOMOGRAPHY/nu\_master}. This file holds three numbers,
-forming a (unit) vector. It describes which component we are cross-correlating at the `master'
-receiver, i.e., ${\hat{\bf \nu}}^{\alpha}$ in \citet{trompetal2010}.
-The three numbers correspond to E/N/Z components respectively.
-Most often, the vertical component is used, and in those cases the three numbers should be 0, 0 and 1.
-
-\item
-Describe the noise direction and distributions in \texttt{src/shared/noise\_tomography.f90}.
-Search for a subroutine called \texttt{noise\_distribution\_direction}
-in \texttt{noise\_tomography.f90}. It is actually located at the very beginning
-of \texttt{noise\_tomography.f90}. The default assumes vertical noise and a uniform distribution
-across the whole free surface of the model. It should be quite self-explanatory for modifications.
-Should you modify this part, you have to re-compile the source code.
-
-\end{itemize}
-
-\subsection{Simulations}
-
-With all of the above done, we can finally launch our simulations.
-Again, please make sure that the \texttt{LOCAL\_PATH} in \texttt{in\_data\_files/Par\_file} is not globally shared.
-It is quite problematic to have a globally shared \texttt{LOCAL\_PATH}, in terms of both disk storage and
-speed of I/O (we have to save the wavefields at the earth's surface at every time step).
-
-As discussed in \citet{trompetal2010}, it takes three steps/simulations to obtain one contribution
-of the ensemble sensitivity kernels:
-
-\begin{itemize}
-\item
-Step 1: simulation for generating wavefields\\
-\texttt{SIMULATION\_TYPE=1}\\
-\texttt{NOISE\_TOMOGRAPHY=1}\\
-\texttt{SAVE\_FORWARD} not used, can be either .true. or .false.
-
-\item
-Step 2: simulation for ensemble forward wavefields\\
-\texttt{SIMULATION\_TYPE=1}\\
-\texttt{NOISE\_TOMOGRAPHY=2}\\
-\texttt{SAVE\_FORWARD=.true.}
-
-\item
-Step 3: simulation for ensemble adjoint wavefields and sensitivity kernels\\
-\texttt{SIMULATION\_TYPE=3}\\
-\texttt{NOISE\_TOMOGRAPHY=3}\\
-\texttt{SAVE\_FORWARD=.false.}
-
-Note Step 3 is an adjoint simulation, please refer to previous chapters on how to prepare adjoint sources
-and other necessary files, as well as how adjoint simulations work.
-
-\end{itemize}
-
-Note that it is better to run the three steps continuously within the same job,
-otherwise you have to collect the saved surface movies from the old nodes/CPUs to the new nodes/CPUs.
-This process varies from cluster to cluster and thus cannot be discussed here.
-Please ask your cluster administrator for information/configuration of the cluster you are using.\\
-
-\subsection{Post-simulation}
-
-After those simulations, you have all stuff you need, either in the \texttt{in\_out\_files/OUTPUT\_FILES/} or
-in the directory specified by \texttt{LOCAL\_PATH} in \texttt{in\_data\_files/Par\_file}
-(which are most probably on local nodes).
-Collect whatever you want from the local nodes to your workstation, and then visualize them.
-This process is the same as what you may have done for regular earthquake simulations.
-Refer to other chapters if you have problems.\\
-
-Simply speaking, two outputs are the most interesting: the simulated ensemble cross correlations and
-one contribution of the ensemble sensitivity kernels.\\
-
-The simulated ensemble cross correlations are obtained after the second simulation (Step 2).
-Seismograms in \texttt{in\_out\_files/OUTPUT\_FILES/} are actually the simulated ensemble cross correlations.
-Collect them immediately after Step 2, or the Step 3 will overwrite them.
-Note that we have a `master' receiver specified by \texttt{irec\_master\_noise},
-the seismogram at one station corresponds to the cross correlation between that station and the `master'.
-Since the seismograms have three components, we may obtain cross correlations
-for different components as well, not necessarily the cross correlations between vertical components.\\
-
-One contribution of the ensemble cross-correlation sensitivity kernels are obtained after Step 3,
-stored in the \texttt{in\_data\_files/LOCAL\_PATH} on local nodes.
-The ensemble kernel files are named the same as regular earthquake kernels.\\
-
-You need to run another three simulations to get the other contribution of the ensemble kernels,
-using different forward and adjoint sources given in \citet{trompetal2010}.
-
-\section{Example}
-
-In order to illustrate noise simulations in an easy way, one example is provided in \texttt{examples/noise\_tomography/}.
-See \texttt{examples/noise\_tomography/README} for explanations. \\
-
-Note, however, that they are created for a specific workstation (CLOVER@PRINCETON),
-which has at least 4 cores with `mpif90' working properly. \\
-
-If your workstation is suitable, you can run the example in \texttt{examples/noise\_tomography/} using:\\
-
-\texttt{./pre-processing.sh}\\
-
-Even if this script does not work on your workstation, the procedure it describes is universal.
-You may review the whole process described in the last section by following the commands in \texttt{pre-processing.sh},
-which should contain enough explanations for all the commands.
-
-\chapter{Graphics}
-
-\section{\label{sec:Mesh-graphics}Meshes}
-
-In case you used the internal mesher \texttt{xmeshfem3D} to create and partition your mesh,
-you can output mesh files in ABAQUS (.INP) and DX (.dx) format to visualize them. For this, you must set either the
-flag \texttt{CREATE\_DX\_FILES} or \texttt{CREATE\_ABAQUS\_FILES} to \texttt{.true.}
-in the mesher's parameter file \texttt{Mesh\_Par\_file}
-prior to running the mesher (see Chapter~\ref{cha:Running-the-Mesher-Meshfem3D} for details).
-You can then use AVS \urlwithparentheses{www.avs.com} or OpenDX \urlwithparentheses{www.opendx.org}
-to visualize the mesh and MPI partition (slices).
-
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=.49\textwidth]{figures/vtk_mesh_vp.jpg}
-\includegraphics[width=.49\textwidth]{figures/vtk_mesh_vs.jpg}
-\par\end{centering}
-
-\caption{Visualization using Paraview of VTK files created by \texttt{xgenerate\_databases} showing P- and S-wave velocities assigned to the mesh points.
-The mesh was created by \texttt{xmeshfem3D} for 4 processors.}
-
-\label{fig:vtk.mesh}
-\end{figure}
-
-You have also the option to visualize the distributed databases produced by \texttt{xgenerate\_databases}
-using Paraview \urlwithparentheses{www.paraview.org}. For this, you must set the flag \texttt{SAVE\_MESH\_FILES} to \texttt{.true.}
-in the main parameter file \texttt{Par\_file} (see Chapter~\ref{cha:Main-Parameter} for details).
-This will create VTK files for each single partition. You can then use Paraview \urlwithparentheses{www.paraview.org} to visualized these partitions.
-
-\section{\label{sec:Movies}Movies}
-
-To make a surface or volume movie of the simulation, set parameters
-\texttt{MOVIE\_SURFACE}, \texttt{MOVIE\_VOLUME}, and \texttt{NTSTEP\_BETWEEN\_FRAMES}
-in the \texttt{Par\_file}. Turning on the movie flags, in particular
-\texttt{MOVIE\_VOLUME}, produces large output files. \texttt{MOVIE\_VOLUME}
-files are saved in the \texttt{LOCAL\_PATH} directory, whereas \texttt{MOVIE\_SURFACE}
-output files are saved in the \texttt{in\_out\_files/OUTPUT\_FILES} directory. We
-save the displacement field if the parameter \texttt{SAVE\_DISPLACEMENT} is set, otherwise the velocity field is saved.
-The look of a movie is determined by the
-half-duration of the source. The half-duration should be large enough
-so that the movie does not contain frequencies that are not resolved
-by the mesh, i.e., it should not contain numerical noise. This can
-be accomplished by selecting a CMT \texttt{HALF\_DURATION} > 1.1 $\times$
-smallest period (see figure \ref{fig:CMTSOLUTION-file}). When \texttt{MOVIE\_SURFACE}
-= .\texttt{true.}, the half duration of each source in the \texttt{CMTSOLUTION}
-file is replaced by
-
-\begin{quote}
-\[
-\sqrt{(}\mathrm{\mathtt{HALF\_DURATIO}\mathtt{N}^{2}}+\mathrm{\mathtt{HDUR\_MOVI}\mathtt{E}^{2}})\]
-\textbf{NOTE:} If \texttt{HDUR\_MOVIE} is set to 0.0, the code will
-select the appropriate value of 1.1 $\times$ smallest period. As
-usual, for a point source one can set \texttt{half duration} in the
-\texttt{CMTSOLUTION} file to be 0.0 and \texttt{HDUR\_MOVIE} = 0.0 in the \texttt{Par\_file} to get
-the highest frequencies resolved by the simulation, but for a finite
-source one would keep all the \texttt{half durations} as prescribed
-by the finite source model and set \texttt{HDUR\_MOVIE} = 0.0.
-\end{quote}
-
-\subsection{Movie Surface}
-
-When running \texttt{xspecfem3D} with the \texttt{MOVIE\_SURFACE}
-flag turned on, the code outputs \texttt{moviedata??????} files in
-the \texttt{in\_out\_files/OUTPUT\_FILES} directory. There are several flags
-in the main parameter file \texttt{Par\_file} that control the output of these moviedata files
-(see section \ref{cha:Main-Parameter} for details):
-\texttt{\small NTSTEP\_BETWEEN\_FRAMES} to set the timesteps between frames,
-\texttt{\small SAVE\_DISPLACEMENT} to save displacement instead of velocity,
-\texttt{USE\_HIGHRES\_FOR\_MOVIES} to save values at all GLL point instead of element edges.
-In order to output additionally shakemaps, you would set
-the parameter \texttt{CREATE\_SHAKEMAP} to \texttt{.true.}.
-
-The files are in a fairly complicated
-binary format, but there is a program provided to convert the
-output into more user friendly formats:
-\begin{description}
-\item [\texttt{xcreate\_movie\_shakemap\_AVS\_DX\_GMT}]
-From \texttt{create\_movie\_shakemap\_AVS\_DX\_GMT.f90}, it
-outputs data in ASCII, OpenDX, or AVS format (also readable in ParaView).
-Before compiling the code, make sure you have the file \texttt{surface\_from\_mesher.h}
-in the  \texttt{in\_out\_files/OUTPUT\_FILES/} directory. This file will be created by the solver run.
-Then type
-\begin{lyxcode}
-{\footnotesize make xcreate\_movie\_shakemap\_AVS\_DX\_GMT}{\footnotesize \par}
-\end{lyxcode}
-and run the executable \texttt{xcreate\_movie\_shakemap\_AVS\_DX\_GMT}
-in the \texttt{bin/} subdirectory. It will create visualization files in your format of choice. The code
-will prompt the user for input parameters.
-
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=.32\textwidth]{figures/movie_surf_1.jpg}
-\includegraphics[width=.32\textwidth]{figures/movie_surf_2.jpg}
-\includegraphics[width=.32\textwidth]{figures/movie_surf_3.jpg}
-\par\end{centering}
-
-\caption{Visualization using AVS files created by \texttt{xcreate\_movie\_shakemap\_AVS\_DX\_GMT} showing
-movie snapshots of vertical velocity components at different times.}
-
-\label{fig:movie.surf}
-\end{figure}
-
-\end{description}
-
-The \texttt{SPECFEM3D} code is running in near real-time to produce
-animations of southern California earthquakes via the web; see Southern California ShakeMovie\textregistered \urlwithparentheses{www.shakemovie.caltech.edu}.
-
-\subsection{\label{sub:Movie-Volume}Movie Volume}
-
-When running xspecfem3D with the \texttt{\small MOVIE\_VOLUME} flag
-turned on, the code outputs several files in \texttt{\small LOCAL\_PATH} specified in the main
-\texttt{Par\_file}, e.g. in directory \texttt{in\_out\_files/DATABASES\_MPI}.
-%As the files can be very large, there are several flags in the \texttt{\small Par\_file}
-%that control the region in space and time that is saved. These are:
-%\texttt{\small NTSTEP\_BETWEEN\_FRAMES}
-The output is saved by each processor
-at the time interval specified by \texttt{\small NTSTEP\_BETWEEN\_FRAMES}.
-For all domains, the velocity field is output to files:
-\texttt{\small proc??????\_velocity\_X\_it??????.bin},
-\texttt{\small proc??????\_velocity\_Y\_it??????.bin} and
-\texttt{\small proc??????\_velocity\_Z\_it??????.bin}.
-For elastic domains, the divergence and curl taken from the velocity field, i.e. $\nabla \cdot \bf{v}$
-and $\nabla \times \bf{v}$, get stored as well:
-\texttt{\small proc??????\_div\_it??????.bin},
-\texttt{\small proc??????\_curl\_X\_t??????.bin},
-\texttt{\small proc??????\_curl\_Y\_it??????.bin} and
-\texttt{\small proc??????\_curl\_Z\_it??????.bin}.
-The files denoted
-\texttt{\small proc??????\_div\_}~\\
-\texttt{\small glob\_it??????.bin} and
-\texttt{\small proc??????\_curl\_glob\_it??????.bin}
-are stored on the global points only, all the other arrays
-are stored on all GLL points. Note that the components X/Y/Z can change to E/N/Z according to
-the \texttt{SUPPRESS\_UTM\_PROJECTION} flag (see also Appendix~\ref{cha:Coordinates} and \ref{cha:channel-codes}).
-
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=.32\textwidth]{figures/movie_volume_1.jpg}
-\includegraphics[width=.32\textwidth]{figures/movie_volume_2.jpg}
-\includegraphics[width=.32\textwidth]{figures/movie_volume_3.jpg}
-\par\end{centering}
-
-\caption{Paraview visualization using movie volume files (converted by \texttt{xcombine\_vol\_data} and \texttt{mesh2vtu.pl}) and showing
-snapshots of vertical velocity components at different times.}
-
-\label{fig:movie.volume}
-\end{figure}
-
-To visualize these files, we use an auxilliary program \texttt{combine\_vol\_data.f90}
-to combine the data from all slices into one mesh file. To compile it in the root directory, type:
-\begin{lyxcode}
-{\footnotesize make~xcombine\_vol\_data~}{\footnotesize \par}
-\end{lyxcode}
-which will create the executable \texttt{xcombine\_vol\_data} in the directory \texttt{bin/}. To output the
-usage of this executable, type `./bin/xcombine\_vol\_data` without arguments.
-As an example, to run the executable you would use
-\begin{lyxcode}
-{\footnotesize cd~bin/~}{\footnotesize \par}
-{\footnotesize ./xcombine\_vol\_data~0~3~velocity\_Z\_it000400~../in\_out\_files/DATABASES\_MPI~../in\_out\_files/OUTPUT\_FILES~0~}{\footnotesize \par}
-\end{lyxcode}
-to create a low-resolution mesh file. The output mesh file will have the name \texttt{velocity\_Z\_it000400.mesh}.
-We next convert the \texttt{.mesh} file into the VTU (Unstructured
-grid file) format which can be viewed in ParaView. For this task, you can use and modify the script \texttt{mesh2vtu.pl}
-located in directory \texttt{utils/Visualization/Paraview/}, for example:
-\begin{lyxcode}
-{\footnotesize mesh2vtu.pl~-i~velocity\_Z\_it000400.mesh~-o~velocity\_Z\_it000400.vtu}{\footnotesize \par}
-\end{lyxcode}
-Notice that this Perl script uses a program \texttt{mesh2vtu} in the
-\texttt{utils/Visualization/Paraview/mesh2vtu} directory, which further uses the VTK \urlwithparentheses{http://www.vtk.org/}
-run-time library for its execution. Therefore, make sure you have
-them properly set in the script according to your system.
-
-\section{\label{sec:Finite-Frequency-Kernels}Finite-Frequency Kernels}
-
-The finite-frequency kernels computed as explained in Section \ref{sec:Adjoint-simulation-finite}
-are saved in the \texttt{LOCAL\_PATH} at the end of the simulation.
-Therefore, we first need to collect these files on the front end,
-combine them into one mesh file, and visualize them with some auxilliary
-programs.
-
-\begin{enumerate}
-\item \textbf{Create slice files}
-
-We will only discuss the case of one source-receiver pair, i.e., the
-so-called banana-doughnut kernels. Although it is possible to collect
-the kernel files from all slices on the front end, it usually takes
-up too much storage space (at least tens of gigabytes). Since the
-sensitivity kernels are the strongest along the source-receiver great
-circle path, it is sufficient to collect only the slices that are
-along or close to the great circle path.
-
-A Perl script \texttt{slice\_number.pl} located in directory \texttt{utils/Visualization/Paraview/} can
-help to figure out the slice numbers that lie along the great circle
-path. It applies to meshes created with the internal mesher \texttt{xmeshfem3D}.
-
-\begin{enumerate}
-\item On machines where you have access to the script, copy the \texttt{Mesh\_Par\_file}, and
-\texttt{output\_solver} files, and run:
-
-\begin{lyxcode}
-{\small slice\_number.pl~Mesh\_Par\_file~output\_solver.txt~slice\_file}{\small \par}
-\end{lyxcode}
-which will generate a \texttt{slices\_file}.
-
-\item For cases with multiple sources and multiple receivers, you need to
-provide a slice file before proceeding to the next step.
-\end{enumerate}
-\item \textbf{Collect the kernel files}
-
-After obtaining the slice files, you can collect the corresponding
-kernel files from the given slices.
-
-\begin{enumerate}
-\item You can use or modify the script \texttt{utils/copy\_basin\_database.pl}
-to accomplish this:
-
-\begin{lyxcode}
-{\small utils/copy\_database.pl~slice\_file~lsf\_machine\_file~filename~{[}jobid]}{\small \par}
-\end{lyxcode}
-where \texttt{\small lsf\_machine\_file} is the machine file generated
-by the LSF scheduler, \texttt{\small filename} is the kernel name
-(e.g., \texttt{\small rho\_kernel}, \texttt{\small alpha\_kernel}
-and \texttt{\small beta\_kernel}), and the optional \texttt{\small jobid}
-is the name of the subdirectory under \texttt{\small LOCAL\_PATH}
-where all the kernel files are stored.
-
-\item After executing this script, all the necessary mesh topology files
-as well as the kernel array files are collected to the local directory
-of the front end.
-\end{enumerate}
-\item \textbf{Combine kernel files into one mesh file}
-
-We use an auxilliary program \texttt{combine\_vol\_data.f90}
-to combine the kernel files from all slices into one mesh file.
-
-\begin{enumerate}
-\item Compile it in the root directory:
-
-\begin{lyxcode}
-{\footnotesize make~xcombine\_vol\_data~}{\footnotesize \par}
-
-{\footnotesize ./bin/xcombine\_vol\_data~slice\_list~filename~input\_dir~output\_dir~high/low-resolution~}{\footnotesize \par}
-\end{lyxcode}
-where \texttt{input\_dir} is the directory where all the individual
-kernel files are stored, and \texttt{output\_dir} is where the mesh
-file will be written.
-
-\item Use 1 for a high-resolution mesh, outputting all the GLL points to
-the mesh file, or use 0 for low resolution, outputting only the corner
-points of the elements to the mesh file.
-\item The output mesh file will have the name \texttt{filename\_rho(alpha,beta).mesh}
-\end{enumerate}
-\item \textbf{Convert mesh files into .vtu files}
-
-\begin{enumerate}
-\item We next convert the \texttt{.mesh} file into the VTU (Unstructured
-grid file) format which can be viewed in ParaView. For this task, you can use and modify the script \texttt{mesh2vtu.pl}
-located in directory \texttt{utils/Visualization/Paraview/}, for example:
-
-\begin{lyxcode}
-{\footnotesize mesh2vtu.pl~-i~file.mesh~-o~file.vtu~}{\footnotesize \par}
-\end{lyxcode}
-\item Notice that this Perl script uses a program \texttt{mesh2vtu} in the
-\texttt{utils/Visualization/Paraview/mesh2vtu} directory, which further uses the VTK \urlwithparentheses{http://www.vtk.org/}
-run-time library for its execution. Therefore, make sure you have
-them properly set in the script according to your system.
-\end{enumerate}
-\item \textbf{Copy over the source and receiver .vtk file}
-
-In the case of a single source and a single receiver, the simulation
-also generates the file \texttt{sr.vtk} located in the
-\texttt{in\_out\_files/OUTPUT\_FILES/} directory to describe
-the source and receiver locations, which can also be viewed in Paraview
-in the next step.
-
-\item \textbf{View the mesh in ParaView}
-
-Finally, we can view the mesh in ParaView \urlwithparentheses{www.paraview.org}.
-
-\begin{enumerate}
-\item Open ParaView.
-\item From the top menu, \textsf{File} $\rightarrow$\textsf{Open data},
-select \texttt{file.vtu}, and click the \textsf{Accept} button.
-
-\begin{itemize}
-\item If the mesh file is of moderate size, it shows up on the screen; otherwise,
-only the bounding box is shown.
-\end{itemize}
-\item Click \textsf{Display Tab} $\rightarrow$ \textsf{Display Style} $\rightarrow$
-\textsf{Representation} and select \textsf{wireframe of surface} to
-display it.
-\item To create a cross-section of the volumetric mesh, choose \textsf{Filter}
-$\rightarrow$ \textsf{cut}, and under \textsf{Parameters Tab}, choose
-\textsf{Cut Function} $\rightarrow$ \textsf{plane}.
-\item Fill in center and normal information given by the \texttt{global\_slice\_number.pl}
-script (either from the standard output or from \texttt{normal\_plane.txt}
-file).
-\item To change the color scale, go to \textsf{Display Tab} $\rightarrow$
-\textsf{Color} $\rightarrow$ \textsf{Edit Color Map} and reselect
-lower and upper limits, or change the color scheme.
-\item Now load in the source and receiver location file by \textsf{File}
-$\rightarrow$ \textsf{Open data}, select \texttt{sr.vt}k, and click
-the \textsf{Accept} button. Choose \textsf{Filter} $\rightarrow$
-\textsf{Glyph}, and represent the points by `\textsf{spheres}'.
-\item For more information about ParaView, see the ParaView Users Guide \urlwithparentheses{www.paraview.org/files/v1.6/ParaViewUsersGuide.PDF}.
-\end{enumerate}
-\end{enumerate}
-%
-\begin{figure}[H]
-\noindent \begin{centering}
-\includegraphics[scale=0.6]{figures/3D-S-Kernel}
-\par\end{centering}
-
-\caption{(a) Top Panel: Vertical source-receiver cross-section of the S-wave
-finite-frequency sensitivity kernel $K_{\beta}$ for station GSC at
-an epicentral distance of 176 km from the September 3, 2002, Yorba
-Linda earthquake. Lower Panel: Vertical source-receiver cross-section
-of the 3D S-wave speed model used for the spectral-element simulations
-\citep{KoLiTrSuStSh04}. (b) The same as (a) but for station HEC at
-an epicentral distance of 165 km \citep{LiTr06}.}
-
-\label{figure:P-wave-speed-finite-frequency}
-\end{figure}
-
-\chapter{\label{cha:Scheduler}Running through a Scheduler}
-
-The code is usually run on large parallel machines, often PC clusters,
-most of which use schedulers, i.e., queuing or batch management systems
-to manage the running of jobs from a large number of users. The following
-considerations need to be taken into account when running on a system
-that uses a scheduler:
-
-\begin{itemize}
-\item The processors/nodes to be used for each run are assigned dynamically
-by the scheduler, based on availability. Therefore, in order for the
-\texttt{xgenerate\_databases} and the \texttt{xspecfem3D} executables (or between successive runs of the solver) to
-have access to the same database files (if they are stored on hard
-drives local to the nodes on which the code is run), they must be
-launched in sequence as a single job.
-\item On some systems, the nodes to which running jobs are assigned are
-not configured for compilation. It may therefore be necessary to pre-compile
-both the \texttt{xgenerate\_databases} and the \texttt{xspecfem3D} executables.
-\item One feature of schedulers/queuing systems is that they allow submission
-of multiple jobs in a ``launch and forget'' mode. In order to take
-advantage of this property, care needs to be taken that output and
-intermediate files from separate jobs do not overwrite each other,
-or otherwise interfere with other running jobs.
-\end{itemize}
-Examples of job scripts can be found in the \texttt{\small utils/Cluster/}
-directory and can straightforwardly be modified and adapted to meet
-more specific running needs.
-
-We describe here in some detail a job submission procedure for the
-Caltech 1024-node cluster, CITerra, under the LSF scheduling system.
-We consider the submission of a regular forward simulation using the internal mesher
-to create mesh partitions. The two
-main scripts are \texttt{\small run\_lsf.bash}, which compiles the
-Fortran code and submits the job to the scheduler, and \texttt{\small go\_mesher\_solver\_lsf\_basin.forward},
-which contains the instructions that make up
-the job itself. These scripts can be found in \texttt{\small utils/Cluster/lsf/} directory
-
-\section{Job submission \texttt{run\_lsf.bash}}
-
-This script first sets the job queue to be `normal'. It then compiles
-the mesher, database generator and solver together, figures out the number of processors
-required for this simulation from the \texttt{in\_data\_files/Par\_file}, and
-submits the LSF job.
-
-\begin{lyxcode}
-\#!/bin/bash
-
-\#~use~the~normal~queue~unless~otherwise~directed~queue=\char`\"{}-q~normal\char`\"{}~
-
-if~{[}~\$\#~-eq~1~];~then
-
-~~~~~~~~echo~\char`\"{}Setting~the~queue~to~\$1\char`\"{}
-
-~~~~~~~~queue=\char`\"{}-q~\$1\char`\"{}~
-
-fi~\\
-~\\
-\#~compile~the~mesher~and~the~solver~
-
-d=`date'~echo~\char`\"{}Starting~compilation~\$d\char`\"{}~
-
-make~clean~
-
-make~xmeshfem3D~
-
-make~xgenerate\_databases~
-
-make~xspecfem3D~
-
-d=`date'~
-
-echo~\char`\"{}Finished~compilation~\$d\char`\"{}~\\
-~\\
-\#~get~total~number~of~nodes~needed~for~solver~
-
-NPROC=`grep~NPROC~in\_data\_files/Par\_file~|~cut~-c~34-~'~\\
-~\\
-
-\#~compute~total~number~of~nodes~needed~for~mesher~
-
-NPROC\_XI=`grep~NPROC\_XI~in\_data\_files/meshfem3D\_files/Mesh\_Par\_file~|~cut~-c~34-~'~
-~\\
-
-NPROC\_ETA=`grep~NPROC\_ETA~in\_data\_files/meshfem3D\_files/Mesh\_Par\_file~|~cut~-c~34-~'~
-~\\
-\#~total~number~of~nodes~is~the~product~of~the~values~read~
-
-numnodes=\$((~\$NPROC\_XI~*~\$NPROC\_ETA~))~\\
-~\\
-
-\#~checks~total~number~of~nodes~
-
-if~{[}~\$numnodes~-neq~\$NPROC~];~then
-
-~~~~~~~~echo~\char`\"{}error number of procs mismatch\char`\"{}
-
-~~~~~~~~exit~
-
-fi~\\
-
-~\\
-echo~\char`\"{}Submitting~job\char`\"{}~
-
-bsub~\$queue~-n~\$numnodes~-W~60~-K~<go\_mesher\_solver\_lsf.forward~
-\end{lyxcode}
-
-\section{Job script \texttt{go\_mesher\_solver\_lsf.forward}}
-
-This script describes the job itself, including setup steps that can
-only be done once the scheduler has assigned a job-ID and a set of
-compute nodes to the job, the \texttt{run\_lsf.bash} commands used
-to run the mesher, database generator and the solver, and calls to scripts that collect
-the output seismograms from the compute nodes and perform clean-up
-operations.
-
-\begin{enumerate}
-\item First the script directs the scheduler to save its own output and
-output from \texttt{stdout} into ~\\
-\texttt{\small in\_out\_files/OUTPUT\_FILES/\%J.o},
-where \texttt{\%J} is short-hand for the job-ID; it also tells the
-scheduler what version of \texttt{mpich} to use (\texttt{mpich\_gm})
-and how to name this job (\texttt{go\_mesher\_solver\_lsf}).
-\item The script then creates a list of the nodes allocated to this job
-by echoing the value of a dynamically set environment variable \texttt{LSB\_MCPU\_HOSTS}
-and parsing the output into a one-column list using the Perl script
-\texttt{utils/Cluster/lsf/remap\_lsf\_machines.pl}. It then creates a set of scratch
-directories on these nodes (\texttt{\small /scratch/}~\\
-\texttt{\small \$USER/DATABASES\_MPI}) to be used as the \texttt{LOCAL\_PATH}
-for temporary storage of the database files. The scratch directories
-are created using \texttt{shmux}, a shell multiplexor that can execute
-the same commands on many hosts in parallel. \texttt{shmux} is available
-from Shmux \urlwithparentheses{web.taranis.org/shmux/}. Make sure that the \texttt{LOCAL\_PATH}
-parameter in \texttt{in\_data\_files/Par\_file} is also set properly.
-\item The next portion of the script launches the mesher, database generator and then the solver
-using \texttt{run\_lsf.bash}.
-\item The final portion of the script collects the seismograms and performs
-clean up on the nodes, using the Perl scripts \texttt{collect\_seismo\_lsf\_multi.pl}
-and \texttt{cleanmulti.pl}.
-\end{enumerate}
-\begin{lyxcode}
-\#!/bin/bash~-v~
-
-\#BSUB~-o~in\_out\_files/OUTPUT\_FILES/\%J.o~
-
-\#BSUB~-a~mpich\_gm~
-
-\#BSUB~-J~go\_mesher\_solver\_lsf~\\
-~\\
-\#~set~up~local~scratch~directories
-
-BASEMPIDIR=/scratch/\$USER/DATABASES\_MPI
-
-mkdir~-p~in\_out\_files/OUTPUT\_FILES~
-
-echo~\char`\"{}\$LSB\_MCPU\_HOSTS\char`\"{}~>~in\_out\_files/OUTPUT\_FILES/lsf\_machines~
-
-echo~\char`\"{}\$LSB\_JOBID\char`\"{}~>~in\_out\_files/OUTPUT\_FILES/jobid
-
-remap\_lsf\_machines.pl~in\_out\_files/OUTPUT\_FILES/lsf\_machines~> in\_out\_files/OUTPUT\_FILES/machines
-
-shmux~-M50~-Sall~-c~\char`\"{}rm~-r~-f~/scratch/\$USER;~\textbackslash{}
-
-~~~~~~mkdir~-p~/scratch/\$USER;~mkdir~-p~\$BASEMPIDIR\char`\"{}~\textbackslash{}
-
-~~~~~~-~<~in\_out\_files/OUTPUT\_FILES/machines~>/dev/null~\\
-~\\
-\#~run~the~specfem~program
-
-current\_pwd=\$PWD \\
-
-cd bin/ \\
-
-run\_lsf.bash~-{}-gm-no-shmem~-{}-gm-copy-env~\$current\_pwd/xmeshfem3D~\\
-
-run\_lsf.bash~-{}-gm-no-shmem~-{}-gm-copy-env~\$current\_pwd/xgenerate\_databases~\\
-
-run\_lsf.bash~-{}-gm-no-shmem~-{}-gm-copy-env~\$current\_pwd/xspecfem3D~\\
-~\\
-\#~collect~seismograms~and~clean~up
-
-cd current\_pwd/~ \\
-
-mkdir~-p~in\_out\_files/SEM
-
-cd~in\_out\_files/SEM/~
-
-collect\_seismo.pl~../OUTPUT\_FILES/lsf\_machines
-
-cleanbase.pl~../OUTPUT\_FILES/machines
-\end{lyxcode}
-
-\chapter{{\normalsize \label{cha:-Changing-the}} Changing the Model}
-
-In this section we explain how to change the velocity model used for your simulations.
-These changes involve contributing specific subroutines
-that replace existing subroutines in the \texttt{SPECFEM3D} package.
-Note that \texttt{SPECFEM3D} can handle Earth models with material properties that vary within each spectral element.
-
-%% magnoni 6/12
-\section{{\normalsize \label{sec:Using-tomographic}}Using external tomographic Earth models}
-
-To implement your own external tomographic model, you must provide your own tomography file
-\texttt{tomography\_model.xyz} and choose between two possible options:
-(1) set in the \texttt{Par\_file} the model parameter \texttt{MODEL = tomo}, or
-(2) for more user control, set in the \texttt{Par\_file} the model parameter \texttt{MODEL = default} and use the following format in the file \texttt{MESH/nummaterial\_velocity\_file} when using CUBIT to construct your mesh (see also section \ref{subsec:Exporting-the-Mesh}):
-\begin{lyxcode}
-\texttt{domain\_ID} \texttt{material\_ID} \texttt{tomography} \texttt{elastic} \texttt{file\_name} 1
-\end{lyxcode}
-where
-\texttt{domain\_ID} is 1 for acoustic or 2 for elastic materials,
-\texttt{material\_ID} a negative, unique identifier (i.e., -1),
-\texttt{tomography} keyword for tomographic material definition,
-\texttt{elastic} keyword for elastic material definition,
-\texttt{file\_name} the name of the tomography file and
-1 a positive unique identifier.
-The name of the tomography file is not used so far, rather change the filename (\texttt{TOMO\_FILENAME}) defined in the file \texttt{model\_tomography.f90} located in the \texttt{src/generate\_databases/} directory.
-
-The external tomographic model is represented by a grid of points with assigned material properties and homogeneous resolution along each spatial direction x, y and z.
-The ASCII file \texttt{tomography\_model.xyz} that describes the tomography should be located in the \texttt{in\_data\_files/} directory.
-The format of the file, as read from \texttt{model\_tomography.f90}, looks like Figure \ref{fig:tomography_file},
-where
-%
-\begin{description}
-\item [{\texttt{ORIG\_X}, \texttt{END\_X}}] are, respectively, the coordinates of the initial and final tomographic grid points along the x direction (in the mesh units, e.g., $m$);
-\item [{\texttt{ORIG\_Y}, \texttt{END\_Y}}] respectively the coordinates of the initial and final tomographic grid points along the y direction (in the mesh units, e.g., $m$);
-\item [{\texttt{ORIG\_Z}, \texttt{END\_Z}}] respectively the coordinates of the initial and final tomographic grid points along the z direction (in the mesh units, e.g., $m$);
-\item [\texttt{SPACING\_X}, \texttt{SPACING\_Y}, \texttt{SPACING\_Z}] the spacing between the tomographic grid points along the x, y and z directions, respectively (in the mesh units, e.g., $m$);
-\item [\texttt{NX}, \texttt{NY}, \texttt{NZ}] the number of grid points along the spatial directions x, y and z, respectively; \texttt{NX} is given by [(\texttt{END\_X} - \texttt{ORIG\_X})/\texttt{SPACING\_X}]+1; \texttt{NY} and \texttt{NZ} are the same as \texttt{NX}, but for the y and z directions, respectively;
-\item [\texttt{VP\_MIN}, \texttt{VP\_MAX}, \texttt{VS\_MIN}, \texttt{VS\_MAX}, \texttt{RHO\_MIN}, \texttt{RHO\_MAX}] the minimum and maximum values of the wave speed \texttt{vp} and \texttt{vs} (in $m \, s^{-1}$) and of the density \texttt{rho} (in $kg \, m^{-3}$); these values could be the actual limits of the tomographic parameters in the grid or the minimum and maximum values to which we force the cut of velocity and density in the model.
-\end{description}
-%
-After the first four lines, in the file \texttt{tomography\_model.xyz} the tomographic grid points are listed with the corresponding values of \texttt{vp}, \texttt{vs} and \texttt{rho}, scanning the grid along the x coordinate (from \texttt{ORIG\_X} to \texttt{END\_X} with step of \texttt{SPACING\_X}) for each given y (from \texttt{ORIG\_Y} to \texttt{END\_Y}, with step of \texttt{SPACING\_Y}) and each given z (from \texttt{ORIG\_Z} to \texttt{END\_Z}, with step of \texttt{SPACING\_Z}).
-%fig
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=1.0\textwidth]{figures/tomo_file.jpg}
-\par\end{centering}
-\caption{Tomography file \texttt{tomography\_model.xyz} that describes an external Earth model and that is read by \texttt{model\_tomography.f90}.
-The coordinates x, y and z of the grid, their limits ORIG\_X, ORIG\_Y, ORIG\_Z, END\_X, END\_Y, END\_Z and the grid spacings SPACING\_X, SPACING\_Y, SPACING\_Z should be in the units of the constructed mesh (e.g., UTM coordinates in meters).}
-\label{fig:tomography_file}
-\end{figure}
-
-The user can implement his own interpolation algorithm for the tomography model by changing the routine \texttt{model\_tomography.f90} located in the \texttt{src/generate\_databases/} directory.
-Moreover, for models that involve both fully defined materials and a tomography description, the \texttt{nummaterial\_velocity\_file} has multiple lines each with the corresponding suitable format described above.
-
-\section{{\normalsize \label{sec:Anelastic-Models}}External (an)elastic Models}
-
-To use your own external model, you can set in the \texttt{Par\_file} the model parameter \texttt{MODEL = external}.
-Three-dimensional acoustic and/or (an)elastic (attenuation) models may then be superimposed
-onto the mesh based upon your subroutine in
-\texttt{model\_external\_values.f90} located in directory \texttt{src/generate\_databases}.
-The call to this routine would be as follows
-
-\begin{lyxcode}
-call~model\_external\_values(xmesh,~ymesh,~zmesh, ~\& \\
-~~~~~~~~~~rho,~vp,~vs,~qmu\_atten,~iflag\_aniso,~idomain\_id)
-\end{lyxcode}
-%
-Input to this routine consists of:
-\begin{description}
-\item [{\texttt{xmesh,ymesh,zmesh}}] location of mesh point
-\end{description}
-%
-Output to this routine consists of:
-\begin{description}
-\item [{\texttt{rho,vp,vs}}] isotropic model parameters for density $\rho$ ($kg/m^3$), $v_p$ ($m/s$) and $v_s$ ($m/s$)
-\item [{\texttt{qmu\_atten}}] Shear wave quality factor: $0<Q_{\mu}<9000$
-\item [{\texttt{iflag\_aniso}}] anisotropic model flag, $0$ indicating no anisotropy or $1$ using anisotropic model parameters as defined in routine file \texttt{model\_aniso.f90}
-\item [{\texttt{idomain\_id}}] domain identifier, $1$ for acoustic, $2$ for elastic, $3$ for poroelastic materials.
-\end{description}
-
-Note that the resolution and maximum value of anelastic models are
-truncated. This speeds the construction of the standard linear solids
-during the meshing stage. To change the resolution, currently at one
-significant figure following the decimal, or the maximum value (9000),
-consult \texttt{shared/constants.h}.
-
-\section{{\normalsize \label{sec:Anisotropic-Models}}Anisotropic Models}
-
-To use your anisotropic models, you can either set in the \texttt{Par\_file} the model parameter \texttt{ANISOTROPY} to \texttt{.true.} or set the model parameter \texttt{MODEL = aniso}.
-Three-dimensional anisotropic models may be superimposed on
-the mesh based upon the subroutine in file \texttt{model\_aniso.f90}
-located in directory \texttt{src/generate\_databases}.
-The call to this subroutine is of the form
-
-\begin{lyxcode}
-call~model\_aniso(iflag\_aniso,~rho,~vp,~vs,~\& \\
-~~~~~~~~~~c11,c12,c13,c14,c15,c16,c22,c23,c24,c25,c26,~\& \\
-~~~~~~~~~~c33,c34,c35,c36,c44,c45,c46,c55,c56,c66)~
-\end{lyxcode}
-%
-Input to this routine consists of:
-\begin{description}
-\item [{\texttt{iflag\_aniso}}] flag indicating the type of the anisotropic model, i.e. $0$ for no superimposed anisotropy, $1$ or $2$ for generic pre-defined anisotropic models.
-\item [{\texttt{rho,vp,vs}}] reference isotropic model parameters for density $\rho$, $v_p$ and $v_s$.
-\end{description}
-%
-Output from the routine consists of the following non-dimensional
-model parameters:
-\begin{description}
-\item [{\texttt{c11},}] \textbf{$\cdots$,} \texttt{\textbf{c66}} 21~dimensionalized
-anisotropic elastic parameters.
-\end{description}
-
-You can replace the \texttt{model\_aniso.f90} file by
-your own version \textit{provided you do not change the call structure
-of the routine}, i.e., the new routine should take exactly the same
-input and produce the required relative output.
-
-\chapter{Post-Processing Scripts}
-
-Several post-processing scripts/programs are provided in the \texttt{utils/}
-directory, and most of them need to be adjusted when used on different
-systems, for example, the path of the executable programs. Here we
-only list a few of the available scripts and provide a brief description, and
-you can either refer to the related sections for detailed usage or,
-in a lot of cases, type the script/program name without arguments
-for its usage.
-
-\section{Process Data and Synthetics\label{sec:Process-data-and-syn}}
-
-In many cases, the SEM synthetics are calculated and compared to data
-seismograms recorded at seismic stations. Since the SEM synthetics
-are accurate for a certain frequency range, both the original data
-and the synthetics need to be processed before a comparison can be
-made.
-
- For such comparisons, the following steps are recommended:
-
-\begin{enumerate}
-\item Make sure that both synthetic and observed seismograms have the correct station/event and timing information.
-\item Convolve synthetic seismograms with a source time function with the half duration specified in the \texttt{CMTSOLUTION} file, provided, as recommended, you used a zero half duration in the SEM simulations.
-\item Resample both observed and synthetic seismograms to a common sampling rate.
-\item Cut the records using the same window.
-\item Remove the trend and mean from the records and taper them.
-\item Remove the instrument response from the observed seismograms (recommended) or convolve the synthetic seismograms with the instrument response.
-\item Make sure that you apply the same filters to both observed and synthetic seismograms. Preferably, avoid filtering your records more than once.
-\item Now, you are ready to compare your synthetic and observed seismograms.
-\end{enumerate}
-
-\noindent
-We generally use the following scripts provided in the \texttt{utils/seis\_process/} directory:
-
-\subsection{Data processing script \texttt{process\_data.pl}}
-
-This script cuts a given portion of the original data, filters it,
-transfers the data into a displacement record, and picks the first
-P and S arrivals. For more functionality, type `\texttt{process\_data.pl}'
-without any argument. An example of the usage of the script:
-
-\begin{lyxcode}
-{\footnotesize process\_data.pl~-m~CMTSOLUTION~-s~1.0~-l~0/4000~-i~-f~-t~40/500~-p~-x~bp~DATA/1999.330{*}.BH?.SAC}
-{\footnotesize \par}
-\end{lyxcode}
-
-\noindent which has resampled the SAC files to a sampling rate of 1 seconds, cut
-them between 0 and 4000 seconds, transfered them into displacement
-records and filtered them between 40 and 500 seconds,
-picked the first P and S arrivals, and added suffix `\texttt{bp}' to the file names.
-
-Note that all of the scripts in this section actually use the SAC
-and/or IASP91 to do the core operations; therefore make sure that
-the SAC and IASP91 packages are installed properly on your system,
-and that all the environment variables are set properly before running
-these scripts.
-
-\subsection{Synthetics processing script \texttt{process\_syn.pl}}
-
-This script converts the synthetic output from the SEM code from ASCII
-to SAC format, and performs similar operations as `\texttt{process\_data.pl}'.
-An example of the usage of the script:
-\begin{lyxcode}
-{\footnotesize process\_syn.pl~-m~CMTSOLUTION~-h~-a~STATIONS~-s~1.0~-l~0/4000~-f~-t~40/500~-p~-x~bp~SEM/{*}.BX?.semd}
-{\footnotesize \par}
-\end{lyxcode}
-which will convolve the synthetics with a triangular source-time function
-from the \texttt{CMTSOLUTION} file, convert the synthetics into SAC
-format, add event and station information into the SAC headers, resample the SAC files with a sampling rate of 1 seconds, cut them between 0 and 4000 seconds, filter them between 40 and
-500 seconds with the same filter used for the observed data, pick the first P and S arrivals, and add the suffix `\texttt{bp}'
-to the file names.
-
-More options are available for this script, such as adding time shift
-to the origin time of the synthetics, convolving the synthetics with
-a triangular source time function with a given half duration, etc.
-Type \texttt{process\_syn.pl} without any argument for a detailed
-usage.
-
-In order to convert between SAC format and ASCII files, useful scripts
-are provided in the subdirectories ~\\
-\texttt{utils/sac2000\_alpha\_convert/} and \texttt{utils/seis\_process/asc2sac/}.
-
-\subsection{Script \texttt{rotate.pl}}
-
-The original data and synthetics have three components: vertical (BHZ resp. BXZ),
-north (BHN resp. BXN) and east (BHE resp. BXE). However, for most seismology applications,
-transverse and radial components are also desirable. Therefore, we
-need to rotate the horizontal components of both the data and the
-synthetics to the transverse and radial direction, and \texttt{\small rotate.pl}
-can be used to accomplish this:
-
-\begin{lyxcode}
-rotate.pl~-l~0~-L~180~-d~DATA/{*}.BHE.SAC.bp~
-
-rotate.pl~-l~0~-L~180~SEM/{*}.BXE.semd.sac.bp~
-\end{lyxcode}
-where the first command performs rotation on the SAC data obtained
-through Seismogram Transfer Program (STP) \urlwithparentheses{http://www.data.scec.org/STP/stp.html},
-while the second command rotates the processed SEM synthetics.
-
-\section{Collect Synthetic Seismograms}
-
-The forward and adjoint simulations generate synthetic seismograms
-in the \texttt{in\_out\_files/OUTPUT\_FILES/} directory by default. For the forward simulation, the files
-are named like \texttt{STA.NT.BX?.semd} for two-column time series, or
-\texttt{STA.NT.BX?.semd.sac} for ASCII SAC format, where STA and NT
-are the station name and network code, and \texttt{BX?} stands for
-the component name. Please see the Appendix~\ref{cha:Coordinates} and \ref{cha:channel-codes}
-for further details.
-
-The adjont simulations generate synthetic seismograms
-with the name \texttt{S?????.NT.S??.sem} (refer to Section \ref{sec:Adjoint-simulation-sources}
-for details). The kernel simulations output the back-reconstructed
-synthetic seismogram in the name \texttt{STA.NT.BX?.semd}, mainly
-for the purpose of checking the accuracy of the reconstruction. Refer
-to Section \ref{sec:Adjoint-simulation-finite} for further details.
-
-You do have further options to change this default output behavior, given in the main constants file
-\texttt{constants.h} located in \texttt{src/shared/} directory:
-\begin{description}
-\item [{\texttt{SEISMOGRAMS\_BINARY}}] set to \texttt{.true.} to have seismograms written out in binary format.
-\item [{\texttt{WRITE\_SEISMOGRAMS\_BY\_MASTER}}] Set to \texttt{.true.} to have only the master process writing out seismograms.
-This can be useful on a cluster, where only the master process node has access to the output directory.
-\item [{\texttt{USE\_OUTPUT\_FILES\_PATH}}] Set to \texttt{.false.} to have the seismograms output to \texttt{LOCAL\_PATH} directory specified in the main parameter file \texttt{in\_data\_files/Par\_file}.
-In this case, you could collect the synthetics onto the frontend using the \texttt{collect\_seismo\_lsf\_multi.pl} script
-located in the \texttt{utils/Cluster/lsf/} directory. The usage of the script would be e.g.:
-
-\begin{lyxcode}
-collect\_seismo.pl~machines~in\_data\_files/Par\_file
-\end{lyxcode}
-
-\end{description}
-where  \texttt{machines} is a file containing the node names and  \texttt{in\_data\_files/Par\_file} the parameter file used to extract the \texttt{LOCAL\_PATH} directory used for the simulation.
-
-\section{Clean Local Database}
-
-After all the simulations are done, the seismograms are collected,
-and the useful database files are copied to the frontend, you may
-need to clean the local scratch disk for the next simulation. This
-is especially important in the case of kernel simulation,
-where very large files are generated for the absorbing boundaries
-to help with the reconstruction of the regular forward wavefield.
-A sample script is provided in \texttt{utils/}:
-
-\begin{lyxcode}
-cleanbase.pl~machines
-\end{lyxcode}
-where  \texttt{machines} is a file containing the node names.
-
-\section{Plot Movie Snapshots and Synthetic Shakemaps}
-
-\subsection{Script \texttt{movie2gif.gmt.pl}}
-
-With the movie data saved in \texttt{in\_out\_files/OUTPUT\_FILES/} at the end of
-a movie simulation (\texttt{\small MOVIE\_SURFACE=.true.}), you can
-run the \texttt{`create\_movie\_shakemap\_AVS\_DX\_GMT}' code to convert these binary
-movie data into GMT xyz files for futher processing. A sample script
-\texttt{movie2gif.gmt.pl} is provided to do this conversion, and then
-plot the movie snapshots in GMT, for example:
-
-\begin{lyxcode}
-movie2gif.gmt.pl~-m~CMTSOLUTION~-g~-f~1/40~-n~-2~-p~
-\end{lyxcode}
-which for the first through the 40th movie frame, converts the \texttt{moviedata}
-files into GMT xyz files, interpolates them using the 'nearneighbor'
-command in GMT, and plots them on a 2D topography map. Note that `\texttt{-2}'
-and `\texttt{-p}' are both optional.
-
-\subsection{Script \texttt{plot\_shakemap.gmt.pl}}
-
-With the shakemap data saved in \texttt{in\_out\_files/OUTPUT\_FILES/} at the end
-of a shakemap simulation ~\\
-(\texttt{CREATE\_SHAKEMAP=.true.}), you can
-also run \texttt{`create\_movie\_shakemap\_AVS\_DX\_GMT}' code to convert the binary
-shakemap data into GMT xyz files. A sample script \texttt{plot\_shakemap.gmt.pl}
-is provided to do this conversion, and then plot the shakemaps in
-GMT, for example:
-
-\begin{lyxcode}
-plot\_shakemap.gmt.pl~data\_dir~type(1,2,3)~CMTSOLUTION~
-\end{lyxcode}
-where \texttt{type=1} for a displacement shakemap, \texttt{2} for
-velocity, and \texttt{3} for acceleration.
-
-\section{Map Local Database}
-
-A sample program \texttt{remap\_database} is provided to map the local
-database from a set of machines to another set of machines. This is
-especially useful when you want to run mesher and solver, or different
-types of solvers separately through a scheduler (refer to Chapter~\ref{cha:Scheduler}).
-\begin{lyxcode}
-run\_lsf.bash~-{}-gm-no-shmem~-{}-gm-copy-env~remap\_database~old\_machines~150
-\end{lyxcode}
-where \texttt{old\_machines} is the LSF machine file used in the previous
-simulation, and \texttt{150} is the number of processors in total.
-
-\chapter{\label{cha:tomo}Tomographic inversion using sensitivity kernels}
-
-One of the fundamental reasons for computing sensitivity kernels (Section~\ref{sec:Adjoint-simulation-finite}) is to use them within a tomographic inversion.
-In other words, use recorded seismograms, make measurements with synthetic seismograms, and use the misfit between the two to iteratively improve the model described by (at least) $V_{\rm p}$, $V_{\rm s}$, and $\rho$ as a function of space.
-
-Whatever misfit function you use for the tomographic inversion (several examples in \cite{TrTaLi05} and \cite{TrKoLi08}), you will weight the sensitivity kernels with measurements.
-Furthermore, you will use as many measurements (stations, components, time windows) as possible per event; hence, we call these composite kernels ``event kernels,'' which are volumetric fields representing the gradient of the misfit function with respect to one of the variables (\eg $V_{\rm s}$).
-The basic features of an adjoint-based tomographic inversion were illustrated in \citet{TaLiTr07} and \citet{TrKoLi08} using a conjugate-gradient algorithm; there are dozens of versions of gradient-based inversion algorithms that could alternatively be used.
-The tomographic inversion of \citet{TaLiMaTr09,TaLiMaTr2010} used SPECFEM3D as well as several additional components which are also stored on the CIG svn server, described next.
-
-The directory containing utilities for tomographic inversion using SPECFEM3D (or other packages that evaluate misfit functions and gradients) is here on the CIG svn server:
-%
-\begin{verbatim}
-/cig/seismo/3D/ADJOINT_TOMO/
-   flexwin/     -- FLEXWIN algorithm for automated picking of time windows
-   measure_adj/ -- reads FLEXWIN output file and makes measurements,
-                   with the option for computing adjoint sources
-   iterate_adj/ -- various tools for iterative inversion
-                   (requires pre-computed "event kernels")
-\end{verbatim}
-%
-This directory also contains a brief \verb+README+ file indicating the role of the three subdirectories, \verb+flexwin+ \citep{Maggi2009}, \verb+measure_adj+, and \verb+iterate_adj+.
-The components for making the model update are there; however, there are no explicit rules for computing the model update, just as with any optimization problem.
-There are options for computing a conjugate gradient step, as well as a source subspace projection step.
-The workflow could use substantial ``stitching together,'' and we welcome suggestions; undoubtedly, the tomographic inversion procedure will improve with time.
-{\bf This is a work in progress.}
-
-The best single file to read is probably: \verb+ADJOINT_TOMO/iterate_adj/cluster/README+.
-
-%============================================================
-
-\chapter*{\label{cha:Bug-Reports-and}Bug Reports and Suggestions for Improvements}
-
-To report bugs or suggest improvements to the code, please send an
-e-mail to the CIG Computational Seismology Mailing List \urlwithparentheses{cig-seismo@geodynamics.org}
-or Jeroen Tromp \urlwithparentheses{jtromp-AT-princeton.edu}, and/or use our online
-bug tracking system Roundup \urlwithparentheses{www.geodynamics.org/roundup}.
-
-\addcontentsline{toc}{chapter}{Notes \& Acknowledgments}
-\chapter*{Notes \& Acknowledgments}
-
-In order to keep the software package thread-safe in case a multithreaded
-implementation of MPI is used, developers should not add modules or
-common blocks to the source code but rather use regular subroutine
-arguments (which can be grouped in ``derived types'' if needed for
-clarity).
-
-The Gauss-Lobatto-Legendre subroutines in \texttt{gll\_library.f90}
-are based in part on software libraries from the Massachusetts Institute
-of Technology, Department of Mechanical Engineering (Cambridge, Massachusetts, USA).
-The non-structured global numbering software was provided by Paul
-F. Fischer (Brown University, Providence, Rhode Island, USA, now at Argonne National Laboratory, USA).
-
-OpenDX \urlwithparentheses{http://www.opendx.org} is open-source based on IBM Data
-Explorer, AVS \urlwithparentheses{http://www.avs.com} is a trademark of Advanced
-Visualization Systems, and ParaView \urlwithparentheses{http://www.paraview.org}
-is an open-source visualization platform.
-
-Please e-mail your feedback, questions, comments, and suggestions
-to Jeroen Tromp \urlwithparentheses{jtromp-AT-princeton.edu} or to the CIG Computational Seismology Mailing List \urlwithparentheses{cig-seismo@geodynamics.org}.
-
-\addcontentsline{toc}{chapter}{Copyright}
-\chapter*{Copyright}
-
-Main authors: Dimitri Komatitsch and Jeroen Tromp
-
-Princeton University, USA, and CNRS / INRIA / University of Pau, France
-
-$\copyright$ Princeton University and CNRS / INRIA / University of Pau, July 2012
-
-This program is free software; you can redistribute it and/or modify
-it under the terms of the GNU General Public License as published
-by the Free Software Foundation (see Appendix \ref{cha:License}).\\
-
-\textbf{\underline{Evolution of the code:}}\\
-
- MPI v 2.1, July 2012:
- Max Rietmann, Peter Messmer, Daniel Peter, Dimitri Komatitsch, Joseph Charles, Zhinan Xie: support for CUDA GPUs,
-better CFL stability for the Stacey absorbing conditions. \\
-
- MPI v. 2.0, November 2010:
- Dimitri Komatitsch, Nicolas Le Goff, Roland Martin and Pieyre Le Loher, University of Pau, France,
- Jeroen Tromp and the Princeton group of developers, Princeton University, USA,
- and Emanuele Casarotti, INGV Roma, Italy:
-  support for CUBIT meshes decomposed by SCOTCH, METIS or ZOLTAN;
-  much faster solver using Michel Deville's inlined matrix products.\\
-
- MPI v. 1.4 Dimitri Komatitsch, University of Pau, Qinya Liu and others, Caltech, September 2006:
-  better adjoint and kernel calculations, faster and better I/Os
-  on very large systems, many small improvements and bug fixes.\\
-
- MPI v. 1.3 Dimitri Komatitsch, University of Pau, and Qinya Liu, Caltech, July 2005:
-  serial version, regular mesh, adjoint and kernel calculations, ParaView support.\\
-
- MPI v. 1.2 Min Chen and Dimitri Komatitsch, Caltech, July 2004:
-  full anisotropy, volume movie.\\
-
- MPI v. 1.1 Dimitri Komatitsch, Caltech, October 2002: Zhu's Moho map, scaling
-  of $V_s$ with depth, Hauksson's regional model, attenuation, oceans, movies.\\
-
- MPI v. 1.0 Dimitri Komatitsch, Caltech, USA, May 2002: first MPI version based on global code.\\
-
- Dimitri Komatitsch, IPG Paris, France, December 1996: first 3-D solver for the CM-5 Connection Machine,
-    parallelized on 128 processors using Connection Machine Fortran.\\
-
-\addcontentsline{toc}{chapter}{References}
-\bibliography{bibliography}
-
-\appendix
-
-\chapter{\label{cha:Coordinates}Reference Frame Convention}
-
-The code uses the following convention for the Cartesian reference
-frame:
-
-\begin{itemize}
-\item the $x$ axis points East
-\item the $y$ axis points North
-\item the $z$ axis points up
-\end{itemize}
-Note that this convention is different from both the \citet{AkRi80}
-convention and the Harvard Centroid-Moment Tensor (CMT) convention.
-The Aki \& Richards convention is
-
-\begin{itemize}
-\item the $x$ axis points North
-\item the $y$ axis points East
-\item the $z$ axis points down
-\end{itemize}
-and the Harvard CMT convention is
-
-\begin{itemize}
-\item the $x$ axis points South
-\item the $y$ axis points East
-\item the $z$ axis points up
-\end{itemize}
-
-\subsection*{Source and receiver locations}
-
-The SPECFEM3D software code internally uses Cartesian coordinates.
-The given locations for sources and receiver locations thus may get
-converted.  Sources and receiver locations are read in from the
-\texttt{CMTSOLUTION} (or \texttt{FORCESOLUTION}) and \texttt{STATIONS}
-files.  Note that e.g. the \texttt{CMTSOLUTION} and
-\texttt{FORCESOLUTION} files denotes the location by
-"longitude/latitude/depth". We read in longitude as $x$ coordinate,
-latitude as $y$ coordinate.
-
-In case the flag \texttt{SUPPRESS\_UTM\_PROJECTION} is set to \texttt{.false.}
-in the main parameter file (see Chapter~\ref{cha:Creating-Distributed-Databases}),
-the $x$/$y$ coordinates have to be given in degrees and are converted to Cartesian coordinates
-using a UTM conversion with the specified UTM zone.
-
-The value for depth (given in $km$ in \texttt{CMTSOLUTION} and
-\texttt{FORCESOLUTION}) or burial depth (given in $m$ in
-\texttt{STATIONS}) is evaluated with respect to the surface of the
-mesh at the specified $x$/$y$ location to find the corresponding $z$
-coordinate.  It is possible to use this depth value directly as $z$
-coordinate by changing the flag \texttt{USE\_SOURCES\_RECVS\_Z} to
-\texttt{.true.} in the file \texttt{constants.h} located in the
-\texttt{src/shared/} subdirectory.
-
-\subsection*{Seismogram outputs}
-
-The seismogram output directions are given in Cartesian $x$/$y$/$z$ directions and not rotated any further.
-Changing flags in \texttt{constants.h} in the \texttt{src/shared/} subdirectory only rotates the seismogram outputs
-if receivers are forced to be located at the surface (\texttt{RECVS\_CAN\_BE\_BURIED\_EXT\_MESH} set to \texttt{.false.}) and the normal to the surface at the receiver location should be used (\texttt{EXT\_MESH\_RECV\_NORMAL} set to
-\texttt{.true.}) as vertical.
-In this case, the outputs are rotated to have the vertical component normal to the surface of the mesh, $x$ and $y$ directions are somewhat arbitrary orthogonal directions along the surface.
-
-For the labeling of the seismogram channels, see Appendix~\ref{cha:channel-codes}. Additionally, we add labels to the synthetics using the following convention:
-For a receiver station located in an
-\begin{description}
-\item[elastic domain:] ~\\
-\begin{itemize}
-\item \texttt{semd} for the displacement vector
-\item \texttt{semv} for the velocity vector
-\item \texttt{sema} for the acceleration vector
-\end{itemize}
-\item[acoustic domain:] ~\\
-(please note that receiver stations in acoustic domains must be buried. This is due to the free surface condition which enforces a zero pressure at the surface.)
-\begin{itemize}
-\item \texttt{semd} for the displacement vector
-\item \texttt{semv} for the velocity vector
-\item \texttt{sema} for pressure which will be stored in the vertical component \texttt{Z}.
-Note that pressure in the acoustic media is isotropic and thus the pressure record would be the same in the other two component directions.
-We therefore use the other two seismogram components to store the acoustic potential in component \texttt{X} (or \texttt{N}) and the first time derivative of the acoustic potential in component \texttt{Y} (or \texttt{E}).
-\end{itemize}
-\end{description}
-The seismograms are by default written out in ASCII-format to the \texttt{in\_out\_files/OUTPUT\_FILES/} subdirectory by each parallel process.
-You can change this behavior by changing the following flags in the \texttt{constants.h} file located in the \texttt{src/shared/} subdirectory:
-\begin{description}
-\item [{\texttt{SEISMOGRAMS\_BINARY}}] set to \texttt{.true.} to have seismograms written out in binary format.
-\item [{\texttt{WRITE\_SEISMOGRAMS\_BY\_MASTER}}] Set to \texttt{.true.} to have only the master process writing out seismograms.
-This can be useful on a cluster, where only the master process node has access to the output directory.
-\item [{\texttt{USE\_OUTPUT\_FILES\_PATH}}] Set to \texttt{.false.} to have the seismograms output to \texttt{LOCAL\_PATH} directory specified in the main parameter file \texttt{in\_data\_files/Par\_file}.
-\end{description}
-
-\chapter{\label{cha:channel-codes}Channel Codes of Seismograms}
-
-Seismic networks, such as the Global Seismographic Network (GSN), generally involve various types of instruments with different bandwidths, sampling properties and component configurations.
-There are standards to name channel codes depending on instrument properties.
-IRIS  \texttt{(www.iris.edu)} uses SEED format for channel codes, which are represented by three letters, such as \texttt{LHN}, \texttt{BHZ}, etc.
-In older versions of the SPECFEM3D package, a common format was used for the channel codes of all seismograms, which was \texttt{BHE/BHN/BHZ} for three components.
-To avoid confusion when comparison are made to observed data, we are now using the FDSN convention (\url{http://www.fdsn.org/}) for SEM seismograms.
-In the following, we give a brief explanation of the FDSN convention used by IRIS, and how it is adopted in SEM seismograms.
-Please visit \texttt{www.iris.edu/manuals/SEED\_appA.htm} for further information.\\
-
-\noindent \textbf{\texttt{Band code:}} The first letter in the channel code denotes the band code of seismograms, which depends on the response band and the sampling rate of instruments.
-The list of band codes used by IRIS is shown in Figure \ref{fig:IRIS_band_codes}.
-The sampling rate of SEM synthetics is controlled by the resolution of simulations rather than instrument properties.
-However, for consistency, we follow the FDSN convention for SEM seismograms governed by their sampling rate.
-For SEM synthetics, we consider band codes for which $dt \leq 1$ s.
-IRIS also considers the response band of instruments.
-For instance, short-period and broad-band seismograms with the same sampling rate correspond to different band codes, such as S and B, respectively.
-In such cases, we consider SEM seismograms as broad band, ignoring the corner period ($\geq 10$ s) of the response band of instruments (note that at these resolutions, the minimum period in the SEM synthetics will be less than $10$ s).
-Accordingly, when you run a simulation the band code will be chosen depending on the resolution of the synthetics, and channel codes of SEM seismograms will start with either \texttt{L}, \texttt{M}, \texttt{B}, \texttt{H}, \texttt{C} or \texttt{F}, shown by red color in the figure.\\
-
-\begin{figure}[ht]
-\noindent \begin{centering}
-\includegraphics[scale=0.6]{figures/IRIS_band_codes.pdf}
-\par\end{centering}
-
-\caption{The band code convention is based on the sampling rate and the response band of instruments.
-Please visit \texttt{www.iris.edu/manuals/SEED\_appA.htm} for further information.
-Grey rows show the relative band-code range in SPECFEM3D, and the band codes used to name SEM seismograms are denoted in red.}
-
-\label{fig:IRIS_band_codes}
-\end{figure}
-
-\noindent \textbf{\texttt{Instrument code:}} The second letter in the channel code corresponds to instrument codes, which specify the family to which the sensor belongs.
-For instance, \texttt{H} and \texttt{L} are used for high-gain and low-gain seismometers, respectively.
-The instrument code of SEM seismograms will always be \texttt{X}, as assigned by FDSN for synthetic seismograms.
-\\
-
-\noindent \textbf{\texttt{Orientation code:}} The third letter in channel codes is an orientation code, which generally describes the physical configuration of the components of instrument packages.
-SPECFEM3D uses the traditional orientation code  \texttt{E/N/Z} (East-West, North-South, Vertical) for three components when a UTM projection is used.
-If the UTM conversion is suppressed, i.e.
-the flag \texttt{SUPPRESS\_UTM\_PROJECTION} is set to \texttt{.true.}, then the three components are labelled \texttt{X/Y/Z} according to the Cartesian reference frame. \\
-
-\noindent \textbf{\texttt{EXAMPLE:}} The sampling rate is given by \texttt{DT} in the main parameter file \texttt{Par\_file} located in the \texttt{in\_data\_files/} subdirectory.
-Depending on the resolution of your simulations, if you choose a sampling rate greater than $0.01$ s and less than $1$ s, a seismogram recording displacements on the vertical component of a station \texttt{ASBS} (network \texttt{AZ}) will be named \texttt{ASBS.AZ.MXZ.semd.sac}, whereas it will be \texttt{ASBS.AZ.BXZ.semd.sac}, if the sampling rate is greater than 0.0125 and less equal to 0.1 s.
-
-\chapter{\label{cha:Troubleshooting}Troubleshooting}
-
-\section*{FAQ}
-
-\begin{description}
-\item [configuration fails:]
-
-   Examine the log file 'config.log'. It contains detailed informations.
-   In many cases, the path's to these specific compiler commands F90,
-   CC and MPIF90 won't be correct if `./configure` fails.
-
-   Please make sure that you have a working installation of a Fortran compiler,
-   a C compiler and an MPI implementation. You should be able to compile this
-   little program code:
-
-\begin{lyxcode}
-
-      program main ~\\
-        include 'mpif.h' ~\\
-        integer, parameter :: CUSTOM\_MPI\_TYPE = MPI\_REAL ~\\
-        integer ier ~\\
-        call MPI\_INIT(ier) ~\\
-        call MPI\_BARRIER(MPI\_COMM\_WORLD,ier) ~\\
-        call MPI\_FINALIZE(ier) ~\\
-      end
-
-\end{lyxcode}
-
-\item [compilation fails stating:] ~\\
-\begin{lyxcode}
-    ...  ~\\
-    obj/program\_generate\_databases.o: In function `MAIN\_\_':  ~\\
-    program\_generate\_databases.f90:(.text+0x14): undefined reference to `\_gfortran\_set\_std'  ~\\
-    ...  ~\\
 \end{lyxcode}
@@
 
-  Normally, this message will appear when you're mixing two different Fortran
-  compilers. That is, using e.g. gfortran to compile non-MPI files
-  and mpif90, wrapper provided for e.g. ifort, to compile MPI-files.
+Normally, this message will appear when you are mixing two different
+Fortran compilers. That is, using e.g. gfortran to compile non-MPI
+files and mpif90, wrapper provided for e.g. ifort, to compile MPI-files.
 
@@
 
-\item [after executing \texttt{xmeshfem3D} I've got elements with skewness of 81\% percent, what does this mean:]
-Look at the skewness table printed in the \texttt{output\_mesher.txt} file after executing \texttt{xmeshfem3D} for the
-example given in \texttt{examples/meshfem3D\_examples/simple\_model/}:
+\item [{after executing \texttt{xmeshfem3D} I've got elements with skewness of 81\% percent, what does this mean:}] Look
+at the skewness table printed in the \texttt{output\_mesher.txt} file
+after executing \texttt{xmeshfem3D} for the example given in \texttt{EXAMPLES/meshfem3D\_examples/simple\_model/}:
+
 \begin{lyxcode}
-...~\\
-histogram of skewness (0. good - 1. bad): ~\\
- 0.0000000E+00  -   5.0000001E-02       27648     81.81818      \% ~\\
- 5.0000001E-02  -   0.1000000               0    0.0000000E+00  \% ~\\
-... ~\\
-\end{lyxcode}
-The first line means that you have 27,648 elements with a skewness value between 0 and 0.05 (which means the element is basically not skewed, just plain regular hexahedral element).
-The total number of elements you have in this mesh is (see in the \texttt{output\_mesher.txt} file a bit further down):
-\begin{lyxcode}
-...~\\
-total number of elements in entire mesh:        33792 ~\\
-... ~\\
-\end{lyxcode}
-which gives you that:  27,648 / 33,792 $\sim$ 81.8 \% of all elements are not skewed, i.e. regular elements. a fantastic value :)
+...~~\\
+~histogram~of~skewness~(0.~good~-~1.~bad):~~~\\
+~0.0000000E+00~-~5.0000001E-02~27648~81.81818~\%~~~\\
+~5.0000001E-02~-~0.1000000~0~0.0000000E+00~\%~~~\\
+~...~~~\\
 
-The histogram lists for this mesh also some stronger skewed elements, for example the worst ones belong to:
-\begin{lyxcode}
-...~\\
-0.6000000      -   0.6500000            2048     6.060606      \% ~\\
-... ~\\
 \end{lyxcode}
 
-about 6 \% of all elements have distortions with a skewness value between 0.6 and 0.65.
-The skewness values give you a hint of how good your mesh is.
-In an ideal world, you would want to have no distortions, just like the 81\% from above.
-Those elements give you the best approximate values by the GLL quadrature used in the spectral-element method.
-However, having weakly distorted elements is still fine and the solutions are still accurate enough.
-So empirically, values up to around 0.7 are tolerable, above that you should consider remeshing...
+The first line means that you have 27,648 elements with a skewness
+value between 0 and 0.05 (which means the element is basically not
+skewed, just plain regular hexahedral element). The total number of
+elements you have in this mesh is (see in the \texttt{output\_mesher.txt}
+file a bit further down):
+\begin{lyxcode}
+...~~\\
+~total~number~of~elements~in~entire~mesh:~33792~~~\\
+~...~~~\\
 
-To give you an idea why some of the elements are distorted, see the following figure \ref{fig:mesh.vp} of the mesh you obtain in the example \texttt{examples/meshfem3D\_examples/simple\_model/}.
+\end{lyxcode}
+
+which gives you that: 27,648 / 33,792 $\sim$ 81.8 \% of all elements
+are not skewed, i.e. regular elements. a fantastic value :)
+
+The histogram lists for this mesh also some stronger skewed elements,
+for example the worst ones belong to:
+\begin{lyxcode}
+...~~\\
+~0.6000000~-~0.6500000~2048~6.060606~\%~~~\\
+~...~~~\\
+
+\end{lyxcode}
+
+about 6 \% of all elements have distortions with a skewness value
+between 0.6 and 0.65. The skewness values give you a hint of how good
+your mesh is. In an ideal world, you would want to have no distortions,
+just like the 81\% from above. Those elements give you the best approximate
+values by the GLL quadrature used in the spectral-element method.
+However, having weakly distorted elements is still fine and the solutions
+are still accurate enough. So empirically, values up to around 0.7
+are tolerable, above that you should consider remeshing...
+
+To give you an idea why some of the elements are distorted, see the
+following figure \ref{fig:mesh.vp} of the mesh you obtain in the
+example \texttt{EXAMPLES/meshfem3D\_examples/simple\_model/}.
 \begin{figure}[htbp]
 \noindent \begin{centering}
-\includegraphics[width=.9\textwidth]{figures/mesh_vp.jpg}
+\includegraphics[width=0.9\textwidth]{figures/mesh_vp.jpg}
 \par\end{centering}
-\caption{Paraview visualization using the mesh vtk-files for the example given in \texttt{examples/meshfem3D\_examples/simple\_model/}.}
+
+\caption{Paraview visualization using the mesh vtk-files for the example given
+in \texttt{EXAMPLES/meshfem3D\_examples/simple\_model/}.}
+
 \label{fig:mesh.vp}
 \end{figure}
-You will see that the mesh contains a doubling layer, where we stitch elements together such that the size of two elements will transition to the size of one element (very useful to keep the ratio of wavespeed / element\_size about constant).
-Those elements in this doubling layer have higher skewness values and make up those 6 \% in the histogram.
 
-\item [the code gives following error message "need at least one receiver":]
-This means that no stations given in the input file \texttt{in\_data\_files/STATIONS} could be located within the dimensions of the mesh.
-This can happen for example when the mesh was created with the in-house mesher \texttt{xmeshfem3D} while using the Universal Transverse Mercator (UTM) projection but the simulation with \texttt{xspecfem3D} was suppressing this projection from latitude/longitude to x/y/z coordinates.
+You will see that the mesh contains a doubling layer, where we stitch
+elements together such that the size of two elements will transition
+to the size of one element (very useful to keep the ratio of wavespeed
+/ element\_size about constant). Those elements in this doubling layer
+have higher skewness values and make up those 6 \% in the histogram.
 
-In such cases, try to change your \texttt{in\_data\_files/Par\_file} and set e.g.:
+\item [{the code gives following error message "need at least one receiver":}] This
+means that no stations given in the input file \texttt{DATA/STATIONS}
+could be located within the dimensions of the mesh. This can happen
+for example when the mesh was created with the in-house mesher \texttt{xmeshfem3D}
+while using the Universal Transverse Mercator (UTM) projection but
+the simulation with \texttt{xspecfem3D} was suppressing this projection
+from latitude/longitude to x/y/z coordinates.
+
+In such cases, try to change your \texttt{DATA/Par\_file} and set
+e.g.:
 \begin{lyxcode}
-SUPPRESS\_UTM\_PROJECTION         = .false. ~\\
-\end{lyxcode}
-to be the same in \texttt{Mesh\_Par\_file} and \texttt{Par\_file}.
-This flag should be identical when using the in-house mesher \texttt{xmeshfem3D}, \texttt{xgenerate\_databases} and \texttt{xspecfem3D} together to run simulations.
+SUPPRESS\_UTM\_PROJECTION~=~.false.~~~\\
 
-The flag determines if the coordinates you specify for your source and station locations are given as lat/lon degrees and
-must be converted to UTM coordinates.
-As an example, if you use \texttt{.false.} within \texttt{Mesh\_Par\_file} then you create a mesh with \texttt{xmeshfem3D} using the UTM projection from lat/lon as input format to UTM projected coordinates to store the mesh point positions, which is fine.
-The error then may occur if in the \texttt{Par\_file} you have this set to \texttt{.true.} so that the \texttt{xgenerate\_databases} and \texttt{xspecfem3D} suppress the UTM projection and assume that all coordinates you use now for source and receiver locations are given in meters (that is, converted) already.
-So it won't find the specified locations in the used mesh.
-As a solutions, just change the flag in \texttt{Par\_file} to be the same as in \texttt{Mesh\_Par\_file} and rerun \texttt{xgenerate\_databases} and \texttt{xspecfem3D} to make sure that your simulation works fine.
-
-\item [I get the following error message "forward simulation became unstable and blew up":]
-In most cases this means that your time step size \texttt{DT} is chosen too big.
-Look at your files \texttt{output\_mesher.txt} or \texttt{output\_solver.txt} created in the folder \texttt{in\_out\_files/OUTPUT\_FILES/}.
-In these output files, find the section:
-\begin{lyxcode}
-... ~\\
-********************************************* ~\\
-*** Verification of simulation parameters *** ~\\
-********************************************* ~\\
-... ~\\
-*** Minimum period resolved =    4.308774 ~\\
-*** Maximum suggested time step =   6.8863556E-02 ~\\
-... ~\\
-... ~\\
 \end{lyxcode}
 
-then change \texttt{DT} in the \texttt{in\_data\_files/Par\_file} to be somewhere close to the maximum suggested time step. In the example above: \\
+to be the same in \texttt{Mesh\_Par\_file} and \texttt{Par\_file}.
+This flag should be identical when using the in-house mesher \texttt{xmeshfem3D},
+\texttt{xgenerate\_databases} and \texttt{xspecfem3D} together to
+run simulations.
+
+The flag determines if the coordinates you specify for your source
+and station locations are given as lat/lon degrees and must be converted
+to UTM coordinates. As an example, if you use \texttt{.false.} within
+\texttt{Mesh\_Par\_file} then you create a mesh with \texttt{xmeshfem3D}
+using the UTM projection from lat/lon as input format to UTM projected
+coordinates to store the mesh point positions, which is fine. The
+error then may occur if in the \texttt{Par\_file} you have this set
+to \texttt{.true.} so that the \texttt{xgenerate\_databases} and \texttt{xspecfem3D}
+suppress the UTM projection and assume that all coordinates you use
+now for source and receiver locations are given in meters (that is,
+converted) already. So it won't find the specified locations in the
+used mesh. As a solutions, just change the flag in \texttt{Par\_file}
+to be the same as in \texttt{Mesh\_Par\_file} and rerun \texttt{xgenerate\_databases}
+and \texttt{xspecfem3D} to make sure that your simulation works fine.
+
+\item [{I get the following error message "forward simulation became unstable and blew up":}] In
+most cases this means that your time step size \texttt{DT} is chosen
+too big. Look at your files \texttt{output\_mesher.txt} or \texttt{output\_solver.txt}
+created in the folder \texttt{OUTPUT\_FILES}. In these output files,
+find the section:
+
+\begin{lyxcode}
+...~~~\\
+~{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}~~~\\
+~{*}{*}{*}~Verification~of~simulation~parameters~{*}{*}{*}~~~\\
+~{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}~~~\\
+~...~~~\\
+~{*}{*}{*}~Minimum~period~resolved~=~4.308774~~~\\
+~{*}{*}{*}~Maximum~suggested~time~step~=~6.8863556E-02~~~\\
+~...~~~\\
+~...~~~\\
+
+\end{lyxcode}
+
+then change \texttt{DT} in the \texttt{DATA/Par\_file} to be somewhere
+close to the maximum suggested time step. In the example above: \\
  \texttt{DT = 0.05d0} \\
-would (most probably) work fine. It could be also bigger than the 0.068 s suggested. This depends a bit on the distortions of
-your mesh elements. The more regular they are, the bigger you can choose \texttt{DT}. Just play with this value a bit and see
-when the simulation becomes stable ...
+ would (most probably) work fine. It could be also bigger than the
+0.068 s suggested. This depends a bit on the distortions of your mesh
+elements. The more regular they are, the bigger you can choose \texttt{DT}.
+Just play with this value a bit and see when the simulation becomes
+stable ...
 
 \end{description}
-
-\chapter{\label{cha:License}License }
-
-\textbf{GNU GENERAL PUBLIC LICENSE Version 2, June 1991. Copyright
-(C) 1989, 1991 Free Software Foundation, Inc. 59 Temple Place, Suite
-330, Boston, MA 02111-1307 USA} \\
-Everyone is permitted to copy and distribute verbatim copies of this
-license document, but changing it is not allowed.
-
-\section*{Preamble}
-
-The licenses for most software are designed to take away your freedom
-to share and change it. By contrast, the GNU General Public License
-is intended to guarantee your freedom to share and change free software
--- to make sure the software is free for all its users. This General
-Public License applies to most of the Free Software Foundation's software
-and to any other program whose authors commit to using it. (Some other
-Free Software Foundation software is covered by the GNU Library General
-Public License instead.) You can apply it to your programs, too.
-
-When we speak of free software, we are referring to freedom, not price.
-Our General Public Licenses are designed to make sure that you have
-the freedom to distribute copies of free software (and charge for
-this service if you wish), that you receive source code or can get
-it if you want it, that you can change the software or use pieces
-of it in new free programs; and that you know you can do these things.
-
-To protect your rights, we need to make restrictions that forbid anyone
-to deny you these rights or to ask you to surrender the rights. These
-restrictions translate to certain responsibilities for you if you
-distribute copies of the software, or if you modify it.
-
-For example, if you distribute copies of such a program, whether gratis
-or for a fee, you must give the recipients all the rights that you
-have. You must make sure that they, too, receive or can get the source
-code. And you must show them these terms so they know their rights.
-
-We protect your rights with two steps:
-
-\begin{enumerate}
-\item Copyright the software, and
-\item Offer you this license which gives you legal permission to copy, distribute
-and/or modify the software.
-\end{enumerate}
-Also, for each author's protection and ours, we want to make certain
-that everyone understands that there is no warranty for this free
-software. If the software is modified by someone else and passed on,
-we want its recipients to know that what they have is not the original,
-so that any problems introduced by others will not reflect on the
-original authors' reputations.
-
-Finally, any free program is threatened constantly by software patents.
-We wish to avoid the danger that redistributors of a free program
-will individually obtain patent licenses, in effect making the program
-proprietary. To prevent this, we have made it clear that any patent
-must be licensed for everyone's free use or not licensed at all.
-
-The precise terms and conditions for copying, distribution and modification
-follow.
-
-\section*{GNU GENERAL PUBLIC LICENSE TERMS AND CONDITIONS FOR COPYING, DISTRIBUTION
-AND MODIFICATION }
-
-\begin{itemize}
-
-\item[0.]This License applies to any program or other work which
-contains a notice placed by the copyright holder saying it may be
-distributed under the terms of this General Public License. The ``Program''
-below refers to any such program or work, and a ``work based on the
-Program'' means either the Program or any derivative work under copyright
-law: that is to say, a work containing the Program or a portion of
-it, either verbatim or with modifications and/or translated into another
-language. (Hereinafter, translation is included without limitation
-in the term ``modification.'') Each licensee is addressed as ``you.''\\
-\\
-Activities other than copying, distribution and modification are not
-covered by this License; they are outside its scope. The act of running
-the Program is not restricted, and the output from the Program is
-covered only if its contents constitute a work based on the Program
-(independent of having been made by running the Program). Whether
-that is true depends on what the Program does.
-
-\end{itemize}
-
-\begin{enumerate}
-\item You may copy and distribute verbatim copies of the Program's source
-code as you receive it, in any medium, provided that you conspicuously
-and appropriately publish on each copy an appropriate copyright notice
-and disclaimer of warranty; keep intact all the notices that refer
-to this License and to the absence of any warranty; and give any other
-recipients of the Program a copy of this License along with the Program.
-
-You may charge a fee for the physical act of transferring a copy,
-and you may at your option offer warranty protection in exchange for
-a fee.
-
-\item You may modify your copy or copies of the Program or any portion of
-it, thus forming a work based on the Program, and copy and distribute
-such modifications or work under the terms of Section 1 above, provided
-that you also meet all of these conditions:
-
-\begin{enumerate}
-\item You must cause the modified files to carry prominent notices stating
-that you changed the files and the date of any change.
-\item You must cause any work that you distribute or publish, that in whole
-or in part contains or is derived from the Program or any part thereof,
-to be licensed as a whole at no charge to all third parties under
-the terms of this License.
-\item If the modified program normally reads commands interactively when
-run, you must cause it, when started running for such interactive
-use in the most ordinary way, to print or display an announcement
-including an appropriate copyright notice and a notice that there
-is no warranty (or else, saying that you provide a warranty) and that
-users may redistribute the program under these conditions, and telling
-the user how to view a copy of this License. (Exception: if the Program
-itself is interactive but does not normally print such an announcement,
-your work based on the Program is not required to print an announcement.)
-\end{enumerate}
-These requirements apply to the modified work as a whole. If identifiable
-sections of that work are not derived from the Program, and can be
-reasonably considered independent and separate works in themselves,
-then this License, and its terms, do not apply to those sections when
-you distribute them as separate works. But when you distribute the
-same sections as part of a whole which is a work based on the Program,
-the distribution of the whole must be on the terms of this License,
-whose permissions for other licensees extend to the entire whole,
-and thus to each and every part regardless of who wrote it.
-
-Thus, it is not the intent of this section to claim rights or contest
-your rights to work written entirely by you; rather, the intent is
-to exercise the right to control the distribution of derivative or
-collective works based on the Program.
-
-In addition, mere aggregation of another work not based on the Program
-with the Program (or with a work based on the Program) on a volume
-of a storage or distribution medium does not bring the other work
-under the scope of this License.
-
-\item You may copy and distribute the Program (or a work based on it, under
-Section 2) in object code or executable form under the terms of Sections
-1 and 2 above provided that you also do one of the following:
-
-\begin{enumerate}
-\item Accompany it with the complete corresponding machine-readable source
-code, which must be distributed under the terms of Sections 1 and
-2 above on a medium customarily used for software interchange; or,
-\item Accompany it with a written offer, valid for at least three years,
-to give any third party, for a charge no more than your cost of physically
-performing source distribution, a complete machine-readable copy of
-the corresponding source code, to be distributed under the terms of
-Sections 1 and 2 above on a medium customarily used for software interchange;
-or,
-\item Accompany it with the information you received as to the offer to
-distribute corresponding source code. (This alternative is allowed
-only for noncommercial distribution and only if you received the program
-in object code or executable form with such an offer, in accord with
-Subsection b above.)
-\end{enumerate}
-The source code for a work means the preferred form of the work for
-making modifications to it. For an executable work, complete source
-code means all the source code for all modules it contains, plus any
-associated interface definition files, plus the scripts used to control
-compilation and installation of the executable. However, as a special
-exception, the source code distributed need not include anything that
-is normally distributed (in either source or binary form) with the
-major components (compiler, kernel, and so on) of the operating system
-on which the executable runs, unless that component itself accompanies
-the executable.
-
-If distribution of executable or object code is made by offering access
-to copy from a designated place, then offering equivalent access to
-copy the source code from the same place counts as distribution of
-the source code, even though third parties are not compelled to copy
-the source along with the object code.
-
-\item You may not copy, modify, sublicense, or distribute the Program except
-as expressly provided under this License. Any attempt otherwise to
-copy, modify, sublicense or distribute the Program is void, and will
-automatically terminate your rights under this License. However, parties
-who have received copies, or rights, from you under this License will
-not have their licenses terminated so long as such parties remain
-in full compliance.
-\item You are not required to accept this License, since you have not signed
-it. However, nothing else grants you permission to modify or distribute
-the Program or its derivative works. These actions are prohibited
-by law if you do not accept this License. Therefore, by modifying
-or distributing the Program (or any work based on the Program), you
-indicate your acceptance of this License to do so, and all its terms
-and conditions for copying, distributing or modifying the Program
-or works based on it.
-\item Each time you redistribute the Program (or any work based on the Program),
-the recipient automatically receives a license from the original licensor
-to copy, distribute or modify the Program subject to these terms and
-conditions. You may not impose any further restrictions on the recipients'
-exercise of the rights granted herein. You are not responsible for
-enforcing compliance by third parties to this License.
-\item If, as a consequence of a court judgment or allegation of patent infringement
-or for any other reason (not limited to patent issues), conditions
-are imposed on you (whether by court order, agreement or otherwise)
-that contradict the conditions of this License, they do not excuse
-you from the conditions of this License. If you cannot distribute
-so as to satisfy simultaneously your obligations under this License
-and any other pertinent obligations, then as a consequence you may
-not distribute the Program at all. For example, if a patent license
-would not permit royalty-free redistribution of the Program by all
-those who receive copies directly or indirectly through you, then
-the only way you could satisfy both it and this License would be to
-refrain entirely from distribution of the Program.
-
-If any portion of this section is held invalid or unenforceable under
-any particular circumstance, the balance of the section is intended
-to apply and the section as a whole is intended to apply in other
-circumstances.
-
-It is not the purpose of this section to induce you to infringe any
-patents or other property right claims or to contest validity of any
-such claims; this section has the sole purpose of protecting the integrity
-of the free software distribution system, which is implemented by
-public license practices. Many people have made generous contributions
-to the wide range of software distributed through that system in reliance
-on consistent application of that system; it is up to the author/donor
-to decide if he or she is willing to distribute software through any
-other system and a licensee cannot impose that choice.
-
-This section is intended to make thoroughly clear what is believed
-to be a consequence of the rest of this License.
-
-\item If the distribution and/or use of the Program is restricted in certain
-countries either by patents or by copyrighted interfaces, the original
-copyright holder who places the Program under this License may add
-an explicit geographical distribution limitation excluding those countries,
-so that distribution is permitted only in or among countries not thus
-excluded. In such case, this License incorporates the limitation as
-if written in the body of this License.
-\item The Free Software Foundation may publish revised and/or new versions
-of the General Public License from time to time. Such new versions
-will be similar in spirit to the present version, but may differ in
-detail to address new problems or concerns.
-
-Each version is given a distinguishing version number. If the Program
-specifies a version number of this License which applies to it and
-``any later version,'' you have the option of following the terms
-and conditions either of that version or of any later version published
-by the Free Software Foundation. If the Program does not specify a
-version number of this License, you may choose any version ever published
-by the Free Software Foundation.
-
-\item If you wish to incorporate parts of the Program into other free programs
-whose distribution conditions are different, write to the author to
-ask for permission. For software which is copyrighted by the Free
-Software Foundation, write to the Free Software Foundation; we sometimes
-make exceptions for this. Our decision will be guided by the two goals
-of preserving the free status of all derivatives of our free software
-and of promoting the sharing and reuse of software generally.
-\end{enumerate}
-
-\subsection*{NO WARRANTY }
-
-\begin{itemize}
-
-\item[11.]BECAUSE THE PROGRAM IS LICENSED FREE OF CHARGE, THERE IS
-NO WARRANTY FOR THE PROGRAM, TO THE EXTENT PERMITTED BY APPLICABLE
-LAW. EXCEPT WHEN OTHERWISE STATED IN WRITING THE COPYRIGHT HOLDERS
-AND/OR OTHER PARTIES PROVIDE THE PROGRAM ``AS IS'' WITHOUT WARRANTY
-OF ANY KIND, EITHER EXPRESSED OR IMPLIED, INCLUDING, BUT NOT LIMITED
-TO, THE IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR
-PURPOSE. THE ENTIRE RISK AS TO THE QUALITY AND PERFORMANCE OF THE
-PROGRAM IS WITH YOU. SHOULD THE PROGRAM PROVE DEFECTIVE, YOU ASSUME
-THE COST OF ALL NECESSARY SERVICING, REPAIR OR CORRECTION.
-
-\item[12.]IN NO EVENT UNLESS REQUIRED BY APPLICABLE LAW OR AGREED
-TO IN WRITING WILL ANY COPYRIGHT HOLDER, OR ANY OTHER PARTY WHO MAY
-MODIFY AND/OR REDISTRIBUTE THE PROGRAM AS PERMITTED ABOVE, BE LIABLE
-TO YOU FOR DAMAGES, INCLUDING ANY GENERAL, SPECIAL, INCIDENTAL OR
-CONSEQUENTIAL DAMAGES ARISING OUT OF THE USE OR INABILITY TO USE THE
-PROGRAM (INCLUDING BUT NOT LIMITED TO LOSS OF DATA OR DATA BEING RENDERED
-INACCURATE OR LOSSES SUSTAINED BY YOU OR THIRD PARTIES OR A FAILURE
-OF THE PROGRAM TO OPERATE WITH ANY OTHER PROGRAMS), EVEN IF SUCH HOLDER
-OR OTHER PARTY HAS BEEN ADVISED OF THE POSSIBILITY OF SUCH DAMAGES.
-
-\end{itemize}
-
-\section*{END OF TERMS AND CONDITIONS }
-
-\subsection*{How to Apply These Terms to Your New Programs}
-
-If you develop a new program, and you want it to be of the greatest
-possible use to the public, the best way to achieve this is to make
-it free software which everyone can redistribute and change under
-these terms.
-
-To do so, attach the following notices to the program. It is safest
-to attach them to the start of each source file to most effectively
-convey the exclusion of warranty; and each file should have at least
-the ``copyright'' line and a pointer to where the full notice is found.
-For example:
-
-\begin{quote}
-One line to give the program's name and a brief idea of what it does.
-Copyright {\footnotesize $\copyright$ (}year) (name of author)
-
-This program is free software; you can redistribute it and/or modify
-it under the terms of the GNU General Public License as published
-by the Free Software Foundation; either version 2 of the License,
-or (at your option) any later version.
-
-This program is distributed in the hope that it will be useful, but
-WITHOUT ANY WARRANTY; without even the implied warranty of MERCHANTABILITY
-or FITNESS FOR A PARTICULAR PURPOSE. See the GNU General Public License
-for more details.
-
-You should have received a copy of the GNU General Public License
-along with this program; if not, write to the Free Software Foundation,
-Inc., 59 Temple Place, Suite 330, Boston, MA 02111-1307 USA
-\end{quote}
-Also add information on how to contact you by electronic and paper
-mail.
-
-If the program is interactive, make it output a short notice like
-this when it starts in an interactive mode:
-
-\begin{quote}
-Gnomovision version 69, Copyright $\copyright$ year name of author Gnomovision
-comes with ABSOLUTELY NO WARRANTY; for details type `show w'. This
-is free software, and you are welcome to redistribute it under certain
-conditions; type `show c' for details.
-\end{quote}
-The hypothetical commands `show w' and `show c' should show the appropriate
-parts of the General Public License. Of course, the commands you use
-may be called something other than `show w' and `show c'; they could
-even be mouse-clicks or menu items -- whatever suits your program.
-
-You should also get your employer (if you work as a programmer) or
-your school, if any, to sign a ``copyright disclaimer'' for the program,
-if necessary. Here is a sample; alter the names:
-
-\begin{quote}
-Yoyodyne, Inc., hereby disclaims all copyright interest in the program
-`Gnomovision' (which makes passes at compilers) written by James Hacker.
-
-(signature of Ty Coon)\\
-1 April 1989 \\
-Ty Coon, President of Vice
-\end{quote}
-This General Public License does not permit incorporating your program
-into proprietary programs. If your program is a subroutine library,
-you may consider it more useful to permit linking proprietary applications
-with the library. If this is what you want to do, use the GNU Library
-General Public License instead of this License.
-\end{document}

### 2015-04-24 doc/USER_MANUAL/D_license.tex

--- 2012-09-28 doc/USER_MANUAL/manual_SPECFEM3D.tex
@@
-%% LyX 1.5.1 created this file.  For more info, see http://www.lyx.org/.
-%% Do not edit unless you really know what you are doing.
-\documentclass[oneside,english,onecolumn]{book}
-\usepackage[T1]{fontenc}
-\usepackage[latin1]{inputenc}
-\usepackage{geometry}
-\geometry{letterpaper,tmargin=1in,bmargin=1in,lmargin=1in,rmargin=1in}
-\setcounter{secnumdepth}{3}
-\setcounter{tocdepth}{3}
-\usepackage{float}
-\usepackage{textcomp}
-\usepackage{amsmath}
-
-% figures
-\usepackage[pdftex]{graphicx}
-
-% we are running pdflatex, so convert .eps files to .pdf
-\usepackage{epstopdf}
-
-\usepackage{amssymb}
-\IfFileExists{url.sty}{\usepackage{url}}
-                      {\newcommand{\url}{\texttt}}
-\usepackage[authoryear]{natbib}
-
-\makeatletter
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% LyX specific LaTeX commands.
-\newcommand{\noun}[1]{\textsc{#1}}
-%% A simple dot to overcome graphicx limitations
-\newcommand{\lyxdot}{.}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Textclass specific LaTeX commands.
-\newenvironment{lyxcode}
-{\begin{list}{}{
-\setlength{\rightmargin}{\leftmargin}
-\setlength{\listparindent}{0pt}% needed for AMS classes
-\raggedright
-\setlength{\itemsep}{0pt}
-\setlength{\parsep}{0pt}
-\normalfont\ttfamily}%
- \item[]}
-{\end{list}}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% User specified LaTeX commands.
-%\renewcommand{\baselinestretch}{1.5}
-
-% figures
-\usepackage[dvips]{epsfig}
-
-\usepackage{wrapfig}
-
-% fonts
-\usepackage{times}
-
-% hyperlinks to sections and references
-\usepackage[pdftex,bookmarks=true,bookmarksnumbered=true,pdfpagemode=None,pdfstartview=FitH,pdfpagelayout=SinglePage,pdfborder={0 0 0}]{hyperref}
-
-\newcommand{\urlwithparentheses}[1]{(\url{#1})}
-
-% biblio GJI
-\bibliographystyle{abbrvnat}
-
-\newcommand{\toall}[1]{\textbf{*** All: #1 ***}}
-\newcommand{\tojeroen}[1]{\textbf{*** Jeroen: #1 ***}}
-\newcommand{\tobrian}[1]{\textbf{*** Brian: #1 ***}}
-\newcommand{\tovala}[1]{\textbf{*** Vala: #1 ***}}
-\newcommand{\tovalabrian}[1]{\textbf{*** Vala \& Brian: #1 ***}}
-\newcommand{\tovalaqinya}[1]{\textbf{*** Vala \& Qinya: #1 ***}}
-\newcommand{\toqinya}[1]{\textbf{*** Qinya: #1 ***}}
-\newcommand{\tomin}[1]{\textbf{*** Min: #1 ***}}
-\newcommand{\toalessia}[1]{\textbf{*** Alessia: #1 ***}}
-\newcommand{\todimitri}[1]{\textbf{*** Dimitri: #1 ***}}
-
-\newcommand{\nexxi}{\mbox{\texttt{NEX\_XI\/}}}
-\newcommand{\nexeta}{\mbox{\texttt{NEX\_ETA\/}}}
-\newcommand{\nprocxi}{\mbox{\texttt{NPROC\_XI\/}}}
-\newcommand{\nproceta}{\mbox{\texttt{NPROC\_ETA\/}}}
-\newcommand{\nchunks}{\mbox{\texttt{NCHUNKS\/}}}
-\newcommand{\eg}{e.g.,}
-
-% colors for the text
-\usepackage[dvipsnames,usenames]{color}
-\newcommand{\red}[1]{\textcolor{Red}{#1}}
-
-\usepackage{babel}
-\makeatother
-
-\begin{document}
-\begin{center}
-\thispagestyle{empty}%
-\vspace*{-1.8truecm}
-\noindent\makebox[\textwidth]{%
-\noindent\includegraphics[width=0.83\paperwidth]{figures/specfem_3d-cover.pdf}
-}
-\end{center}
-%
-\title{\textbf{SPECFEM3D}\\
-\textbf{User Manual}}
-
-\author{$\copyright$ Princeton University (USA) and\\
-CNRS / INRIA / University of Pau (France)\\
-Version 2.1}
-
-\date{\noindent \today}
-
-\maketitle
-
-\section*{Authors}
-The SPECFEM3D package was first developed by Dimitri Komatitsch and Jeroen Tromp at Harvard University and Caltech, USA, starting in 1998, based on earlier work by Dimitri Komatitsch and Jean-Pierre Vilotte at Institut de Physique du Globe (IPGP) in Paris, France from 1995 to 1997.\\
-
-Since then it has been developed and maintained by a development team: in alphabetical order,
-Piero Basini,
-C\'eline Blitz,
-Ebru Bozda\u{g},
-Emanuele Casarotti,
-Joseph Charles,
-Min Chen,
-Vala Hj\"orleifsd\'ottir,
-Sue Kientz,
-Dimitri Komatitsch,
-Jes\'us Labarta,
-Nicolas Le Goff,
-Pieyre Le Loher,
-Qinya Liu,
-Yang Luo,
-Alessia Maggi,
-Federica Magnoni,
-Roland Martin,
-Dennis McRitchie,
-Matthias Meschede,
-Peter Messmer,
-David Mich\'ea,
-Tarje Nissen-Meyer,
-Daniel Peter,
-Max Rietmann,
-Brian Savage,
-Bernhard Schuberth,
-Anne Sieminski,
-Leif Strand,
-Carl Tape,
-Jeroen Tromp,
-Zhinan Xie,
-Hejun Zhu.\\
-%\toall{add full list of developers here and update the AUTHORS file accordingly}
-
-The cover graphic of the manual was created
-by Santiago Lombeyda from Caltech's Center for Advanced Computing Research (CACR) \urlwithparentheses{http://www.cacr.caltech.edu}. \\
-
-\newpage{}
-
-\tableofcontents{}
-
-\chapter{Introduction}
-
-The software package SPECFEM3D simulates seismic
-wave propagation at the local or regional scale
-based upon the spectral-element method (SEM).
-The SEM is a continuous Galerkin technique \citep{TrKoLi08,PeKoLuMaLeCaLeMaLiBlNiBaTr11},
-which can easily be made discontinous \citep{BeMaPa94,Ch00,KoWoHu02,ChCaVi03,LaWaBe05,Kop06,WiStBuGh10,AcKo11};
-it is then close to a particular case of the discontinuous Galerkin technique \citep{ReHi73,Arn82,FaRi99,HuHuRa99,CoKaSh00,GiHeWa02,RiWh03,MoRi05,GrScSc06,AiMoMu06,BeLaPi06,DuKa06,DeSeWh08,PuAmKa09,WiStBuGh10,DeSe10,EtChViGl10}, with optimized efficiency because of its tensorized basis functions \citep{WiStBuGh10,AcKo11}.
-In particular, it can accurately handle very distorted mesh elements \citep{OlSe11}.\\
-
-It has very good accuracy and convergence properties \citep{MaPa89,SePr94,DeFiMu02,Coh02,DeSe07,SeOl08}.
-The spectral element approach admits spectral rates of convergence and allows exploiting $hp$-convergence schemes.
-It is also very well suited to parallel implementation on very large supercomputers \citep{KoTsChTr03,TsKoChTr03,KoLaMi08a,CaKoLaTiMiLeSnTr08,KoViCh10} as well as on clusters of GPU accelerating graphics cards \citep{KoMiEr09,KoErGoMi10,Kom11}.
-Tensor products inside each element can be optimized to reach very high efficiency \citep{DeFiMu02}, and mesh point and element numbering can be optimized to reduce processor cache misses and improve cache reuse \citep{KoLaMi08a}.
-The SEM can also handle triangular (in 2D) or tetrahedral (in 3D) elements \citep{WinBoyd96,TaWi00,KoMaTrTaWi01,Coh02,MeViSa06} as well as mixed meshes, although with increased cost and reduced accuracy in these elements, as in the discontinuous Galerkin method.\\
-
-Note that in many geological models in the context of seismic wave propagation studies
-(except for instance for fault dynamic rupture studies, in which very high frequencies or supershear rupture need to be modeled near the fault, see e.g.
-\cite{BeGlCrViPi07,BeGlCrVi09,PuAmKa09,TaCrEtViBeSa10})
-a continuous formulation is sufficient because material property contrasts are not drastic and thus
-conforming mesh doubling bricks can efficiently handle mesh size variations \citep{KoTr02a,KoLiTrSuStSh04,LeChLiKoHuTr08,LeChKoHuTr09,LeKoHuTr09}.\\
-
-For a detailed introduction to the SEM as applied to regional
-seismic wave propagation, please consult \citet{PeKoLuMaLeCaLeMaLiBlNiBaTr11,TrKoLi08,KoVi98,KoTr99,ChKoViCaVaFe07} and
-in particular \citet{LeKoHuTr09,LeChKoHuTr09,LeChLiKoHuTr08,GoAmTaCaSmSaMaKo09,WiKoScTr04,KoLiTrSuStSh04}.
-A detailed theoretical analysis of the dispersion
-and stability properties of the SEM is available in \citet{Coh02}, \citet{DeSe07}, \citet{SeOl07} and \citet{SeOl08}.\\
-
-Effects due to lateral variations in compressional-wave speed, shear-wave
-speed, density, a 3D crustal model, topography and bathymetry are
-included. The package can accommodate
-full 21-parameter anisotropy (see~\citet{ChTr07}) as well as lateral
-variations in attenuation \citep{SaKoTr10}. Adjoint capabilities and finite-frequency
-kernel simulations are included \citep{TrKoLi08,PeKoLuMaLeCaLeMaLiBlNiBaTr11,LiTr06,FiIgBuKe09,ViOp09}.\\
-
-The SEM was originally developed in computational fluid dynamics \citep{Pat84,MaPa89}
-and has been successfully adapted to address problems in seismic wave propagation.
-Early seismic wave propagation applications of the SEM, utilizing Legendre basis functions and a
-perfectly diagonal mass matrix, include \cite{CoJoTo93}, \cite{Kom97},
-\cite{FaMaPaQu97}, \cite{CaGa97}, \cite{KoVi98} and \cite{KoTr99},
-whereas applications involving Chebyshev basis functions and a nondiagonal mass matrix
-include \cite{SePr94}, \cite{PrCaSe94} and \cite{SePrPr95}.\\
-
-All SPECFEM3D software is written in Fortran90 with full portability
-in mind, and conforms strictly to the Fortran95 standard. It uses
-no obsolete or obsolescent features of Fortran77. The package uses
-parallel programming based upon the Message Passing Interface (MPI)
-\citep{GrLuSk94,Pac97}.\\
-
-SPECFEM3D won the Gordon Bell award for best performance at the SuperComputing~2003
-conference in Phoenix, Arizona (USA) (see \cite{KoTsChTr03}
-and \url{www.sc-conference.org/sc2003/nr_finalaward.html}).
-It was a finalist again in 2008 for a run at 0.16 petaflops (sustained) on 149,784 processors of the `Jaguar' Cray XT5 system at Oak Ridge National Laboratories (USA) \citep{CaKoLaTiMiLeSnTr08}.
-It also won the BULL Joseph Fourier supercomputing award in 2010.\\
-
-This new release of the code (version 2.1) includes support for GPU graphics card acceleration \citep{KoMiEr09,KoGoErMi10,KoErGoMi10,MiKo10,Kom11}.\\
-
-The next release of the code will include support Convolution or Auxiliary Differential Equation Perfectly Matched absorbing Layers
-(C-PML or ADE-PML) \citep{KoMa07,MaKoEz08,MaKoGe08,MaKo09,MaKoGeBr10}.
-It will also use the PT-SCOTCH parallel library for mesh partitioning.\\
-
-SPECFEM3D includes coupled fluid-solid domains and
-adjoint capabilities, which enables one to address seismological inverse problems, but for linear rheologies only so far.
-To accommodate visco-plastic or non-linear rheologies, the reader can refer to
-the \href{http://geoelse.stru.polimi.it/}{GeoELSE} software package \citep{CaGa97,StPaIg09,ChMoTsBaKrKaStKr10}.
-
-\section{Citation}
-
-If you use SPECFEM3D for your own research, please cite at least one
-of the following articles:
-
-\begin{description}
-
-\item [Numerical simulations in general] ~\\
-Forward and adjoint simulations are described in detail in \cite{TrKoLi08,PeKoLuMaLeCaLeMaLiBlNiBaTr11,VaCaSaKoVi99,KoMiEr09,KoErGoMi10,KoGoErMi10,ChKoViCaVaFe07,MaKoDi09,KoViCh10,CaKoLaTiMiLeSnTr08,TrKoHjLiZhPeBoMcFrTrHu10,KoRiTr02,KoTr02a,KoTr02b,KoTr99} or \cite{KoVi98}.
-Additional aspects dealing with adjoint simulations are described in
-\cite{TrTaLi05,LiTr06,TrKoLi08,LiTr08,TrKoHjLiZhPeBoMcFrTrHu10,PeKoLuMaLeCaLeMaLiBlNiBaTr11}.
-Domain decomposition is explained in detail in \cite{MaKoBlLe08}, and excellent scaling up to 150,000 processor cores is shown for instance in
-\cite{CaKoLaTiMiLeSnTr08,KoLaMi08a,MaKoBlLe08,KoErGoMi10,KoGoErMi10,Kom11},
-
-\item [GPU computing] ~\\
-Computing on GPU graphics cards for acoustic or seismic wave propagation applications
-is described in detail in \cite{KoMiEr09,KoGoErMi10,KoErGoMi10,MiKo10,Kom11}.
-
-\end{description}
-
-\noindent
-If you use this new version 2.1, which has non blocking MPI for much better performance for medium or large runs, please cite at least one of these six articles,
-in which results of non blocking MPI runs are presented: \cite{PeKoLuMaLeCaLeMaLiBlNiBaTr11,KoGoErMi10,KoErGoMi10,KoViCh10,Kom11,CaKoLaTiMiLeSnTr08,MaKoBlLe08}.\\
-
-\noindent
-If you work on geophysical applications, you may be interested in citing some of these application articles as well, among others:
-
-\begin{description}
-
-\item [Southern California simulations] ~\\
-\cite{KoLiTrSuStSh04,KrJiKoTr06a,KrJiKoTr06b}.
-
-If you use the 3D southern California model, please cite
-\citet{SuSh03} (Los Angeles model),
-\citet{lovelyetal06} (Salton Trough), and
-\citet{hauksson2000} (southern California).
-The Moho map was determined by
-\citet{zhu&kanamori2000}.
-The 1D SoCal model was developed by \citet{DrHe90}.
-
-\item [Anisotropy] ~\\
-\cite{ChTr07,JiTsKoTr05,ChFaKo04,FaChKo04,RiRiKoTrHe02,TrKo00}.
-
-\item [Attenuation] ~\\
-\cite{SaKoTr10,KoTr02a,KoTr99}.
-
-\item [Topography] ~\\
-\cite{LeKoHuTr09,LeChKoHuTr09,LeChLiKoHuTr08,GoAmTaCaSmSaMaKo09,WiKoScTr04}.
-
-\end{description}
-
-The corresponding Bib\TeX{} entries may be found
-in file \texttt{doc/USER\_MANUAL/bibliography.bib}.
-
-\section{Support}
-
-This material is based upon work supported by the USA National Science
-Foundation under Grants No. EAR-0406751 and EAR-0711177, by the French
-CNRS, French INRIA Sud-Ouest MAGIQUE-3D, French ANR NUMASIS under
-Grant No. ANR-05-CIGC-002, and European FP6 Marie Curie International
-Reintegration Grant No. MIRG-CT-2005-017461.
-Any opinions, findings, and conclusions or recommendations expressed in this material are
-those of the authors and do not necessarily reflect the views of the
-USA National Science Foundation, CNRS, INRIA, ANR or the European
-Marie Curie program.
-
-\chapter{\label{cha:Getting-Started}Getting Started}
-
-The SPECFEM3D software package comes in a gzipped tar ball. In the
-directory in which you want to install the package, type
-\begin{lyxcode}
-{\small tar~-zxvf~SPECFEM3D\_V2.1.0.tar.gz}{\small \par}
-\end{lyxcode}
-The directory \texttt{\small SPECFEM3D/} will then contain
-the source code.\\
-
-We recommend that you add {\texttt{ulimit -S -s unlimited}} to your {\texttt{.bash\_profile}} file and/or {\texttt{limit stacksize unlimited }} to your {\texttt{.cshrc}} file to suppress any potential limit to the size of the Unix stack.\\
-
-Then, to configure the software for your system, run the
-\texttt{configure} shell script. This script will attempt to guess
-the appropriate configuration values for your system. However, at
-a minimum, it is recommended that you explicitly specify the appropriate
-command names for your Fortran90 compiler and MPI package:
-
-\begin{lyxcode}
-./configure~FC=ifort~MPIFC=mpif90
-\end{lyxcode}
-
-Before running the \texttt{configure} script, you should probably edit file \texttt{flags.guess} to make sure that it contains the best compiler options for your system.
-Known issues or things to check are:
-
-\begin{description}
-\item [{\texttt{Intel ifort compiler}}] See if you need to add \texttt{-assume byterecl} for your machine.
-\item [{\texttt{IBM compiler}}] See if you need to add \texttt{-qsave} or \texttt{-qnosave} for your machine.
-\item [{\texttt{Mac OS}}] You will probably need to install \texttt{XCODE}.
-In addition, the \texttt{clock\_gettime} routine, which is used by the \texttt{SCOTCH} library that we use, does not exist in Mac OS.
-You will need to replace it with \texttt{clock\_get\_time} if you want to use \texttt{SCOTCH}.
-\end{description}
-
-When compiling on an IBM machine with the \texttt{xlf} and \texttt{xlc} compilers, we suggest running the \texttt{configure} script
-with the following options:\\
-
-\noindent\texttt{./configure FC=xlf90\_r MPIFC=mpif90 CC=xlc\_r CFLAGS="-O3 -q64" FCFLAGS="-O3 -q64" -with-scotch-dir=...}.\\
-
-On SGI systems, \texttt{flags.guess} automatically informs \texttt{configure}
-to insert `\texttt{`TRAP\_FPE=OFF}'' into the generated \texttt{Makefile}
-in order to turn underflow trapping off.\\
-
-If you run very large meshes on a relatively small number
-of processors, the memory size needed on each processor might become
-greater than 2 gigabytes, which is the upper limit for 32-bit addressing.
-In this case, on some compilers you may need to add \texttt{``-mcmodel=medium}'' or \texttt{``-mcmodel=medium -shared-intel}''
-to the configure options of CFLAGS, FCFLAGS and LDFLAGS otherwise the compiler will display an error
-message (for instance \texttt{``relocation truncated to fit: R\_X86\_64\_PC32 against .bss''} or something similar);
-on an IBM machine with the \texttt{xlf} and \texttt{xlc} compilers, using \texttt{-q64} is usually sufficient.\\
-
-The SPECFEM3D software package relies on the SCOTCH library to partition meshes created with CUBIT.
-Note that we use CUBIT to create meshes of hexahedra, but other packages can be used as well, for instance GiD from \url{http://gid.cimne.upc.es} or Gmsh from \url{http://geuz.org/gmsh} \citep{GeRe09}.
-Even mesh creation packages that generate tetrahedra, for instance TetGen from \url{http://tetgen.berlios.de}, can be used because each tetrahedron can then easily be decomposed into four hexahedra as shown in the picture of the TetGen logo at \url{http://tetgen.berlios.de/figs/Delaunay-Voronoi-3D.gif}; while this approach does not generate hexahedra of optimal quality, it can ease mesh creation in some situations and it has been shown that the spectral-element method can very accurately handle distorted mesh elements \citep{OlSe11}.
-
-The SCOTCH library \citep{PeRo96}
-provides efficent static mapping, graph and mesh partitioning routines. SCOTCH is a free software package developed by
-Fran\c {c}ois Pellegrini et al.
-from LaBRI and INRIA in Bordeaux, France, downloadable from the web page \url{https://gforge.inria.fr/projects/scotch/}.
-It is more recent than METIS, actively maintained and performs better in many cases.
-A recent version of its source code is provided in directory \texttt{src/decompose\_mesh\_SCOTCH/scotch\_5.1.10b}.
-In case no SCOTCH libraries can be found on the system, the configuration will bundle this version for compilation.
-The path to an existing SCOTCH installation can to be set explicitly with the option \texttt{-{}-with-scotch-dir}. Just as an example:
-\begin{lyxcode}
-./configure~FC=ifort~MPIFC=mpif90~-{}-with-scotch-dir=/opt/scotch
-\end{lyxcode}
-
-If you use the Intel ifort compiler to compile the code, we recommend that you use the Intel icc C compiler to compile Scotch, i.e., use:
-\begin{lyxcode}
-./configure~CC=icc~FC=ifort~MPIFC=mpif90
-\end{lyxcode}
-
-To compile a serial version of the code for small meshes that fits
-on one compute node and can therefore be run serially, run \texttt{configure}
-with the \texttt{-{}-without-mpi} option to suppress all calls to
-MPI.
-
-A summary of the most important configuration variables follows.
-
-\begin{description}
-\item [{\texttt{F90}}] Path to the Fortran90 compiler.
-\item [{\texttt{MPIF90}}] Path to MPI Fortran90.
-\item [{\texttt{MPI\_FLAGS}}] Some systems require this flag to link to
-MPI libraries.
-\item [{\texttt{FLAGS\_CHECK}}] Compiler flag for non-critical subroutines.
-\item [{\texttt{FLAGS\_NO\_CHECK}}] Compiler flag for creating fast, production-run
-code for critical subroutines.
-\end{description}
-The configuration script automatically creates for each executable a corresponding
-\texttt{Makefile} in the \texttt{src/} subdirectoy.
-The \texttt{Makefile} contains a number of suggested entries for various
-compilers, e.g., Portland, Intel, Absoft, NAG, and Lahey. The software
-has run on a wide variety of compute platforms, e.g., various PC clusters
-and machines from Sun, SGI, IBM, Compaq, and NEC. Select the compiler
-you wish to use on your system and choose the related optimization
-flags. Note that the default flags in the \texttt{Makefile} are undoubtedly
-not optimal for your system, so we encourage you to experiment with
-these flags and to solicit advice from your systems administrator.
-Selecting the right compiler and optimization flags can make a tremendous
-difference in terms of performance. We welcome feedback on your experience
-with various compilers and flags.
-
-Now that you have set the compiler information, you need to select
-a number of flags in the \texttt{constants.h} file depending on your
-system:
-
-\begin{description}
-\item [{\texttt{LOCAL\_PATH\_IS\_ALSO\_GLOBAL}}] Set to \texttt{.false.}
-on most cluster applications. For reasons of speed, the (parallel)
-distributed database generator typically writes a (parallel) database for the solver on the
-local disks of the compute nodes. Some systems have no local disks,
-e.g., BlueGene or the Earth Simulator, and other systems have a fast
-parallel file system, in which case this flag should be set to \texttt{.true.}.
-Note that this flag is not used by the database generator or the solver; it is
-only used for some of the post-processing.
-\end{description}
-The package can run either
-in single or in double precision mode. The default is single precision
-because for almost all calculations performed using the spectral-element method
-using single precision is sufficient and gives the same results (i.e. the same seismograms);
-and the single precision code is faster and requires exactly half as much memory.
-Select your preference by selecting the appropriate setting in the \texttt{constants.h} file:
-
-\begin{description}
-\item [{\texttt{CUSTOM\_REAL}}] Set to \texttt{SIZE\_REAL} for single precision
-and \texttt{SIZE\_DOUBLE} for double precision.
-\end{description}
-In the \texttt{precision.h} file:
-
-\begin{description}
-\item [{\texttt{CUSTOM\_MPI\_TYPE}}] Set to \texttt{MPI\_REAL} for single
-precision and \texttt{MPI\_DOUBLE\_PRECISION} for double precision.
-\end{description}
-On many current processors (e.g., Intel, AMD, IBM Power), single precision calculations
-are significantly faster; the difference can typically be 10\%
-to 25\%. It is therefore better to use single precision.
-What you can do once for the physical problem you want to study is run the same calculation in single precision
-and in double precision on your system and compare the seismograms.
-If they are identical (and in most cases they will), you can select single precision for your future runs.\\
-
-\section{Adding OpenMP support in addition to MPI}
-
-OpenMP support can be enabled in addition to MPI. However, in many cases performance will not improve
-because our pure MPI implementation is already heavily optimized and thus the resulting code will in fact
-be slightly slower. A possible exception could be IBM BlueGene-type architectures.
-
-To enable OpenMP, uncomment the OpenMP compiler option in two lines in file \texttt{src/specfem3D/Makefile.in}
-(before running \texttt{configure}) and also uncomment the \texttt{\#define USE\_OPENMP} statement in file\\
-\texttt{src/specfem3D/specfem3D.F90}.
-
-The DO-loop using OpenMP threads has a SCHEDULE property. The OMP\_SCHEDULE
-environment variable can set the scheduling policy of that DO-loop.
-Tests performed by Marcin Zielinski at SARA (The Netherlands) showed that often
-the best scheduling policy is DYNAMIC with the size of the chunk equal to the number of
-OpenMP threads, but most preferably being twice as the number of
-OpenMP threads (thus chunk size = 8 for 4 OpenMP threads etc).
-If OMP\_SCHEDULE is not set or is empty, the DO-loop will assume generic
-scheduling policy, which will slow down the job quite a bit.
-
-\section{Visualizing the subroutine calling tree of the source code}
-
-Packages such as \texttt{doxywizard} can be used to visualize the subroutine calling tree of the source code.
-\texttt{Doxywizard} is a GUI front-end for configuring and running \texttt{doxygen}.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\chapter{\label{cha:Mesh-Generation}Mesh Generation}
-
-The first step in running a spectral-element simulation consists of constructing a high-quality mesh for the region under consideration.
-We provide two possibilities to do so: (1) relying on the external, hexahedral mesher CUBIT, or (2) using the provided, internal mesher \texttt{xmeshfem3D}.
-In the following, we explain these two approaches.
-
-\section{\label{cha:Running-the-Mesher-CUBIT}Meshing with \texttt{CUBIT}}
-
-CUBIT is a meshing tool suite for the creation of finite-element meshes for arbitrarily shaped models.
-It has been developed and maintained at Sandia National Laboratories and can be purchased for a small academic institutional fee at \url{http://cubit.sandia.gov}.
-Our experience showed that using CUBIT greatly facilitates and speeds up the generation and preparation of hexahedral, conforming meshes for a variety of geophysical models with increasing complexity.
-
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=.6\textwidth]{figures/mount-cubit.jpg}
-\par\end{centering}
-
-\caption{Example of the graphical user interface of CUBIT. The hexahedral mesh shown in the main display consists of a hexahedral discretization of a single volume with topography.}
-
-\label{fig:mount.cubit}
-\end{figure}
-
-The basic steps in creating a load-balanced, partitioned mesh with CUBIT are:
-\begin{description}
-\item [1.] setting up a hexahedral mesh with CUBIT,
-\item [2.] exporting the CUBIT mesh into a SPECFEM3D file format and
-\item [3.] partitioning the SPECFEM3D mesh files for a chosen number of cores.
-\end{description}
-
-Examples are provided in the SPECFEM3D package in the subdirectory \texttt{examples/}.
-We strongly encourage you to contribute your own example to this package by contacting the CIG Computational Seismology Mailing List  \urlwithparentheses{cig-seismo@geodynamics.org}.
-
-\subsection{Creating the Mesh with CUBIT}
-
-For the installation and handling of the CUBIT meshing tool suite, please refer to the CUBIT user manual and documentation.
-In order to give you a basic understanding of how to use CUBIT for our purposes, examples are provided in the SPECFEM3D package in the subdirectory \texttt{examples/}:
-\begin{description}
-\item [{\texttt{homogeneous\_halfspace}}] Creates a single block model and assigns elastic material parameters.
-\item [{\texttt{layered\_halfspace}}]  Combines two different, elastic material volumes and creates a refinement layer between the two.
-This example can be compared for validation against the solutions provided in subdirectory ~\\
-\texttt{VALIDATION\_3D\_SEM\_SIMPLER\_LAYER\_SOURCE\_DEPTH/}.
-\item [{\texttt{waterlayered\_halfspace}}] Combines an acoustic and elastic material volume as in a schematic marine survey example.
-\item [{\texttt{tomographic\_model}}] Creates a single block model whose material properties will have to be read in from a tomographic model file during the databases creation by \texttt{xgenerate\_databases}.
-\end{description}
-
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=.45\textwidth]{figures/example-homogeneous.jpg}
-\includegraphics[width=.45\textwidth]{figures/example-2layers.jpg} \\
-\includegraphics[width=.45\textwidth]{figures/example-water.jpg}
-\includegraphics[width=.45\textwidth]{figures/example-tomo.jpg}
-\par\end{centering}
-
-\caption{Screenshots of the CUBIT examples provided in subdirectory  \texttt{examples/}: homogeneous halfspace (top-left), layered halfspace (top-right), water layered halfspace (bottom-left) and tomographic model (bottom-right).}
-
-\label{fig:examples.cubit}
-\end{figure}
-
-In each example subdirectory you will find a \texttt{README} file, which explains in a step-by-step tutorial the workflow for the example.
-Please feel free to contribute your own example to this package by contacting the CIG Computational Seismology Mailing List  \urlwithparentheses{cig-seismo@geodynamics.org}.
-
-\subsection{Exporting the Mesh with \texttt{cubit2specfem3d.py} }
-\label{subsec:Exporting-the-Mesh}
-
-Once the geometric model volumes in CUBIT are meshed, you prepare the model for exportation with the definition of material blocks and boundary surfaces.
-Thus, prior to exporting the mesh, you need to define blocks specifying the materials and absorbing boundaries in CUBIT.
-This process could be done automatically using the script \texttt{boundary\_definition.py} if the mesh meets some conditions or manually, following the block convention:
-
-\begin{description}
-\item [material\_name] Each material should have a specific block defined by a unique name.
-The name convention of the material is to start with either 'elastic' or 'acoustic'.
-It must be then followed by a unique identifier, e.g.
-'elastic 1', 'elastic 2', etc.
-The additional attributes to the block define the material description.
-
-For an elastic material:
-\begin{description}
-\item [material\_id] An integer value which is unique for this material.
-\item [Vp] P-wave speed of the material (given in m/s).
-\item [Vs] S-wave speed of the material (given in m/s).
-\item [rho] density of the material (given in kg/m$^3$).
-\item [Q] quality factor to use in case of a simulation with attenuation turned on.
-It should be between 1 and 9000.
-In case no attenuation information is available, it can be set to zero.
-Please note that your Vp- and Vs-speeds are given for a reference frequency.
-To change this reference frequency, you change the value of
-\texttt{ATTENUATION\_f0\_REFERENCE} in the main constants file \texttt{constants.h} found in subdirectory \texttt{src/shared/}.
-\item [anisotropic\_flag] Flag describing the anisotropic model to use in case an anisotropic simulation should be conducted.
-See the file \texttt{model\_aniso.f90} in subdirectory \texttt{src/generate\_databases/} for an implementation of the anisotropic models.
-In case no anisotropy is available, it can be set to zero.
-\end{description}
-Note that this material block has to be defined using all the volumes which belong to this elastic material.
-For volumes belonging to another, different material, you will need to define a new material block.
-
-For an acoustic material:
-\begin{description}
-\item [material\_id] An integer value which is unique for this material.
-\item [Vp] P-wave speed of the material (given in m/s).
-\item [0] S-wave speed of the material is ignored.
-\item [rho] density of the material (given in kg/m$^3$).
-\end{description}
-
-\item [face\_topo] Block definition for the surface which defines the free surface (which can have topography).
-The name of this block must be 'face\_topo', the block has to be defined using all the surfaces which constitute the complete free surface of the model.
-
-\item [face\_abs\_xmin] Block definition for the faces on the absorbing boundaries, one block for each surface with x=Xmin.
-\item [face\_abs\_xmax] Block definition for the faces on the absorbing boundaries, one block for each surface with x=Xmax.
-\item [face\_abs\_ymin] Block definition for the faces on the absorbing boundaries, one block for each surface with y=Ymin.
-\item [face\_abs\_ymax] Block definition for the faces on the absorbing boundaries, one block for each surface with y=Ymax.
-\item [face\_abs\_bottom] Block definition for the faces on the absorbing boundaries, one block for each surface with z=bottom.
-\end{description}
-
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=.45\textwidth]{figures/mount-surface.jpg}
-\includegraphics[width=.45\textwidth]{figures/mount-abs.jpg}
-\par\end{centering}
-
-\caption{Example of the block definitions for the free surface 'face\_topo' (left) and the absorbing boundaries, defined in a single block 'face\_abs\_xmin' (right) in CUBIT.}
-
-\label{fig:mount.abs}
-\end{figure}
-
-Optionally, instead of specifying for each surface at the boundaries a single block like mentioned above, you can also specify a single block for all boundary surfaces and name it as one of the absorbing blocks above, e.g.
-'face\_abs\_xmin'.
-
-After the block definitions are done, you export the mesh using the script \texttt{cubit2specfem3d.py} provided in each of the example directories.
-If the export was successful, you should find the following files in a subdirectory \texttt{MESH/}:
-
-\begin{description}
-\item [nummaterial\_velocity\_file] Defines the material properties. For fully defined materials, the formats is:
-\begin{lyxcode}
-\texttt{domain\_ID} \texttt{material\_ID}  \texttt{rho}  \texttt{vp}  \texttt{vs}  \texttt{Q}  \texttt{anisotropy\_flag}
-\end{lyxcode}
-where
-\texttt{domain\_ID} is 1 for acoustic or 2 for elastic materials, \texttt{material\_ID} a unique identifier,  \texttt{rho} the density in $kg \, m^{-3}$,  \texttt{vp} the P-wave speed in $m \, s^{-1}$, \texttt{vs}  the S-wave speed in $m \, s^{-1}$, \texttt{Q} the quality factor and \texttt{anisotropy\_flag} an identifier for anisotropic models.
-
-%%magnoni
-For tomographic velocity models, please read Chapter \ref{cha:-Changing-the} and Section \ref{sec:Using-tomographic} `Using external tomographic Earth models' for further details.
-%
-%For a tomographic model, the material definition format is:
-%\begin{lyxcode}
-%\texttt{domain\_ID} \texttt{material\_ID}  \texttt{tomography}  \texttt{name}
-%\end{lyxcode}
-%where
-%\texttt{domain\_ID} is 1 for acoustic or 2 for elastic materials, \texttt{material\_ID} a negative, unique identifier, \texttt{tomography} keyword for tomographic material definition and \texttt{name} the name of the tomography file.
-The name is not used so far, rather change the filename defined in the file \texttt{model\_tomography.f90} located in the
-%\texttt{src/generate\_databases/} directory.
-%
-\item [materials\_file] Contains the material associations for each element.
-\item [nodes\_coords\_file] Contains the point locations in Cartesian coordinates of the mesh element corners.
-\item [mesh\_file] Contains the mesh element connectivity.
-\item [free\_surface\_file] Contains the free surface connectivity.
-\item [absorbing\_surface\_file\_xmax] Contains the surface connectivity of the absorbing boundary surface at the Xmax.
-\item [absorbing\_surface\_file\_xmin] Contains the surface connectivity of the absorbing boundary surface at the Xmin.
-\item [absorbing\_surface\_file\_ymax] Contains the surface connectivity of the absorbing boundary surface at the Ymax.
-\item [absorbing\_surface\_file\_ymin] Contains the surface connectivity of the absorbing boundary surface at the Ymin.
-\item [absorbing\_surface\_file\_bottom] Contains the surface connectivity of the absorbing boundary surface at the bottom.
-\end{description}
-
-These mesh files are needed as input files for the partitioner  \texttt{xdecompose\_mesh\_SCOTCH}
-to load-balance the mesh. Please see the next section for further details.
-
-\subsubsection*{Checking the mesh quality}
-The quality of the mesh may be inspected more precisely based upon the serial code
-in the file  \texttt{check\_mesh\_quality\_}~\\
-\texttt{CUBIT\_Abaqus.f90} located in the directory \texttt{src/check\_mesh\_quality\_CUBIT\_Abaqus/}.
-Running this code is optional because no information needed by the
-solver is generated.
-
-Prior to running and compiling this code, you have to export your mesh in CUBIT to an ABAQUS (.inp) format.
-For example, export mesh block IDs belonging to volumes in order to check the quality of the hexahedral elements.
-You also have to determine a number of parameters of your mesh, such as the number of nodes and number of elements
-and modify the header of the \texttt{check\_mesh\_quality\_CUBIT\_Abaqus.f90} source file. Then,
-in the main directory, type
-\begin{lyxcode}
-{\small make xcheck\_mesh\_quality\_CUBIT\_Abaqus}{\small \par}
-\end{lyxcode}
-and use
-\begin{lyxcode}
-{\small ./bin/xcheck\_mesh\_quality\_CUBIT\_Abaqus}{\small \par}
-\end{lyxcode}
-to generate an
-%not supported: AVS output file (\texttt{\small AVS\_meshquality.inp} in AVS UCD format) or
-OpenDX output file (\texttt{\small DX\_mesh\_quality.dx})
-that can be used to investigate mesh quality, e.g. skewness of elements,
-and a Gnuplot histogram (\texttt{\small mesh\_quality\_histogram.txt})
-that can be plotted with gnuplot (type
-`\texttt{\small gnuplot plot\_mesh\_quality\_histogram.gnu}'). The
-histogram is also printed to the screen.
-Analyze that skewness histogram of mesh elements to make sure no element has a skewness above approximately 0.75, otherwise the mesh is of poor quality (and if even a single element has a skewness value above 0.80, then you must definitely improve the mesh).
-If you want to start designing
-your own meshes, this tool is useful for viewing your creations. You
-are striving for meshes with elements with `cube-like' dimensions,
-e.g., the mesh should contain no very elongated or skewed elements.
-
-\subsection{Partitioning the Mesh with \texttt{xdecompose\_mesh\_SCOTCH}}
-
-The SPECFEM3D software package performs large scale simulations in a parallel 'Single Process Multiple Data' way.
-The spectral-element mesh created with CUBIT needs to be distributed on the processors. This partitioning is executed once and for all prior to
-the execution of the solver so it is referred to as a static mapping.
-
-An efficient partitioning is important because it leverages the overall running time of the application.
-It amounts to balance the number of elements in each slice while minimizing the communication costs
-resulting from the placement of adjacent elements on different processors. \texttt{decompose\_mesh\_SCOTCH} depends on the SCOTCH library \citep{PeRo96},
-which provides efficent static mapping, graph and mesh partitioning routines. SCOTCH is a free software package developed by
-Fran\c {c}ois Pellegrini et al.
-from LaBRI and INRIA in Bordeaux, France, downloadable from the web page \url{https://gforge.inria.fr/projects/scotch/}.
-It is more recent than METIS, actively maintained and performs better in many cases.
-A recent version of its source code is provided in directory \texttt{src/decompose\_mesh\_SCOTCH/scotch\_5.1.10b}.
-\\
-
-In most cases, the configuration with \texttt{./configure~FC=ifort} should be sufficient.
-During the configuration process, the script tries to find existing SCOTCH installations.
-In case your system has no pre-existing SCOTCH installation,
-we provide the source code of SCOTCH, which is released open source under the French CeCILL-C version 1 license,
-in directory
-\texttt{src/decompose\_mesh\_SCOTCH/scotch\_5.1.10b}.
-This version gets bundled with the compilation of the SPECFEM3D package if no libraries could have been found.
-If this automatic compilation of the SCOTCH libraries fails,
-please refer to file INSTALL.txt in that directory to see further details how to compile it on your system.
-In case you want to use a pre-existing installation, make sure you have correctly specified the path
-of the SCOTCH library when using the option \texttt{-{}-with-scotch-dir} with the \texttt{./configure} script.
-In the future you should be able to find more recent versions at  \url{http://www.labri.fr/perso/pelegrin/scotch/scotch\_en.html} .
-\\
-
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=.45\textwidth]{figures/mount-partitions.jpg}
-\includegraphics[width=.45\textwidth]{figures/mount-partitions2.jpg}
-\par\end{centering}
-
-\caption{Example of a mesh partitioning onto four cores.
-Each single core partition is colored differently.
-The executable \texttt{xdecompose\_mesh\_SCOTCH} can equally distribute the mesh on any arbitrary number of cores.
-Domain decomposition is explained in detail in \cite{MaKoBlLe08}, and excellent scaling up to 150,000 processor cores in shown for instance in
-\cite{CaKoLaTiMiLeSnTr08,KoLaMi08a,MaKoBlLe08,KoErGoMi10,KoGoErMi10,Kom11}.}
-
-\label{fig:mount.partitions}
-\end{figure}
-
-When you are ready to compile,
-in the main directory type
-\begin{lyxcode}
-{\small make xdecompose\_mesh\_SCOTCH}{\small \par}
-\end{lyxcode}
-If all paths and flags have been set correctly, the executable \texttt{bin/xdecompose\_mesh\_SCOTCH} should be produced.
-
-The partitioning is done in serial for now (in the next release we will provide a parallel version of that code),
-the synopsis is:
-\begin{lyxcode}
-./bin/xdecompose\_mesh\_SCOTCH  nparts  input\_directory output\_directory
-\end{lyxcode}
-where
-\begin{itemize}
-     \item \texttt{nparts} is the number of partitions, i.e., the number of cores for the parallel simulations,
-     \item \texttt{input\_directory} is the directory which holds all the files generated by the Python script \texttt{cubit2specfem3d.py} explained in the previous Section~\ref{subsec:Exporting-the-Mesh}, e.g.
-\texttt{MESH/}, and
-     \item \texttt{output\_directory} is the directory for the output of this partitioner which stores ACII-format files named like \texttt{proc******\_Database} for each partition.
-These files will be needed for creating the distributed databases, and have to reside in the directory  \texttt{LOCAL\_PATH} specified in the main \texttt{Par\_file}, e.g.
-in directory \texttt{in\_out\_files/DATABASES\_MPI}.
-Please see Chapter~\ref{cha:Creating-Distributed-Databases} for further details.
-\end{itemize}
-Note that all the files generated by the Python script  \texttt{cubit2specfem3d.py} must be placed in the \texttt{input\_directory} folder before running the program.
-
-\section{\label{cha:Running-the-Mesher-Meshfem3D}Meshing with \texttt{xmeshfem3D}}
-
-In case you successfully ran the configuration script, you are also ready to compile the internal mesher.
-This is an alternative to CUBIT
-for the mesh generation of relatively simple geological models. The mesher is no longer
-dedicated to Southern California and more flexiblity is provided in this version of the package.
-
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[scale=0.5]{figures/socal_map_mpi}
-\par\end{centering}
-
-\caption{\label{fig:For-parallel-computing}For parallel computing purposes,
-the model block is subdivided in $\nprocxi\times\nproceta$ slices
-of elements. In this example we use $5^{2}=25$ processors. }
-
-\end{figure}
-
-In the main directory, type
-\begin{lyxcode}
-{\small make xmeshfem3D}{\small \par}
-\end{lyxcode}
-If all paths and flags
-have been set correctly, the mesher should now compile and produce
-the executable \texttt{bin/xmeshfem3D}. Please note that
-\texttt{xmeshfem3D} must be called directly from the \texttt{bin/} directory,
-as most of the binaries of the package.
-
-Input for the mesh generation program is provided through the parameter
-file \texttt{Mesh\_Par\_file}, which resides in the subdirectory \texttt{in\_data\_files/meshfem3D\_files/}.
-Before running the mesher, a number of parameters need to be set in
-the \texttt{Mesh\_Par\_file}. This requires a basic understanding of how
-the SEM is implemented, and we encourage you to read \citet{KoVi98,KoTr99}
-and \citet{KoLiTrSuStSh04}.
-
-The mesher and the solver use UTM coordinates internally, therefore
-you need to define the zone number for the UTM projection (e.g., zone
-11 for Los Angeles). Use decimal values for latitude and longitude
-(no minutes/seconds). These values are approximate; the mesher will
-round them off to define a square mesh in UTM coordinates. When running
-benchmarks on rectangular models, turn the UTM projection off by using
-the flag \texttt{\small SUPPRESS\_UTM\_PROJECTION}, in which case
-all `longitude' parameters simply refer to the $x$~axis, and all
-`latitude' parameters simply refer to the $y$~axis. To run the mesher
-for a global simulation, the following parameters need to be set in
-the \texttt{Mesh\_Par\_file}:
-
-\begin{description}
-\item [{\texttt{LATITUDE\_MIN}}] Minimum latitude in the block (negative
-for South).
-\item [{\texttt{LATITUDE\_MAX}}] Maximum latitude in the block.
-\item [{\texttt{LONGITUDE\_MIN}}] Minimum longitude in the block (negative
-for West).
-\item [{\texttt{LONGITUDE\_MAX}}] Maximum longitude in the block.
-\item [{\texttt{DEPTH\_BLOCK\_KM}}] Depth of bottom of mesh in kilometers.
-\item [{\texttt{\noun{UTM\_PROJECTION\_ZONE}}}] UTM projection zone in
-which your model resides, only valid when \texttt{SUPPRESS\_UTM\_}~\\
-\texttt{PROJECTION} is \texttt{.false.}.
-\item [{\texttt{SUPPRESS\_UTM\_PROJECTION}}] set to be \texttt{.false.}
-when your model range is specified in geographical coordinates,
-and needs to be \texttt{.true.} when your model is specified in
-Cartesian coordinates. \noun{UTM projection zone in which your simulation
-region resides.}
-\item [{\texttt{INTERFACES\_FILE }}] File which contains the description of the topography and
-of the interfaces between the different layers of the model, if any.
-The number of spectral elements in the vertical direction within each layer is also defined in this file.
-
-\item [{$\nexxi$}] The number of spectral elements along one side of the
-block. This number \textit{must} be 8~$\times$~a multiple of $\nprocxi$
-defined below. Based upon benchmarks against semi-analytical discrete
-wavenumber synthetic seismograms \citep{KoLiTrSuStSh04}, determined
-that a $\nexxi=288$ run is accurate to a shortest period of roughly
-2~s. Therefore, since accuracy is determined by the number of grid
-points per shortest wavelength, for any particular value of $\nexxi$
-the simulation will be accurate to a shortest period determined by
-\begin{equation}
-\mbox{shortest period (s)}=(288/\nexxi)\times2.\label{eq:shortest_period}\end{equation}
-The number of grid points in each orthogonal direction of the reference
-element, i.e., the number of Gauss-Lobatto-Legendre points, is determined
-by \texttt{NGLLX} in the \texttt{constants.h} file. We generally use
-$\mbox{\texttt{NGLLX\/}}=5$, for a total of $5^{3}=125$ points per
-elements. We suggest not to change this value.
-\item [{$\nexeta$}] The number of spectral elements along the other side
-of the block. This number \textit{must} be 8~$\times$~a multiple
-of $\nproceta$ defined below.
-\item [{$\nprocxi$}] The number of processors or slices along one side
-of the block (see Figure~\ref{fig:For-parallel-computing}); we must
-have $\nexxi=8\times c\times\nprocxi$, where $c\ge1$ is a positive
-integer.
-\item [{$\nproceta$}] The number of processors or slices along the other
-side of the block; we must have $\nexeta=8\times c\times\nproceta$,
-where $c\ge1$ is a positive integer.
-\item [{\texttt{USE\_REGULAR\_MESH}}] set to be \texttt{.true.} if you want a perfectly regular
-mesh or \texttt{.false.} if you want to add doubling horizontal layers to coarsen the mesh. In this case, you
-also need to provide additional information by setting up the next three parameters.
-\item [{\texttt{NDOUBLINGS}}] The number of horizontal doubling layers. Must be set to \texttt{1} or \texttt{2}
-if \texttt{USE\_REGULAR\_MESH} is set to \texttt{.true.}.
-\item [{\texttt{NZ\_DOUBLING\_1}}] The position of the first doubling layer (only interpreted
-if \texttt{USE\_REGULAR\_MESH} is set to \texttt{.true.}).
-\item [{\texttt{NZ\_DOUBLING\_2}}] The position of the second doubling layer (only interpreted
-if \texttt{USE\_REGULAR\_MESH} is set to \texttt{.true.} and if \texttt{NDOUBLINGS} is set to \texttt{2}).
-\item [{\texttt{CREATE\_ABAQUS\_FILES}}] Set this flag to \texttt{.true.} to save Abaqus FEA \urlwithparentheses{www.simulia.com}
-mesh files for subsequent viewing. Turning the flag on generates files in the \texttt{LOCAL\_PATH} directory.
-See Section~\ref{sec:Mesh-graphics} for a discussion of mesh viewing features.
-\item [{\texttt{CREATE\_DX\_FILES}}] Set this flag to \texttt{.true.} to save OpenDX \urlwithparentheses{www.opendx.org}
-mesh files for subsequent viewing.
-\item [{\texttt{LOCAL\_PATH}}] Directory in which the partitions generated
-by the mesher will be written. Generally one uses a directory on the
-local disk of the compute nodes, although on some machines these partitions
-are written on a parallel (global) file system (see also the earlier
-discussion of the \texttt{LOCAL\_PATH\_IS\_ALSO\_GLOBAL} flag in Chapter~\ref{cha:Getting-Started}).
-The mesher generates the necessary partitions in parallel, one set
-for each of the $\nprocxi\times\nproceta$ slices that constitutes
-the mesh (see Figure~\ref{fig:For-parallel-computing}). After the
-mesher finishes, you can log in to one of the compute nodes and view
-the contents of the \texttt{LOCAL\_PATH} directory to see the
-files generated by the mesher.
-These files will be needed for creating the distributed databases, and have to reside in the directory  \texttt{LOCAL\_PATH} specified in the main \texttt{Par\_file}, e.g.
-in directory \texttt{in\_out\_files/DATABASES\_MPI}.
-Please see Chapter~\ref{cha:Creating-Distributed-Databases} for further details.
-\item [{\texttt{NMATERIALS}}] The number of different materials in your model. In the following lines, each material needs to be defined as :
-\begin{lyxcode}
-\texttt{material\_ID}  \texttt{rho}  \texttt{vp}  \texttt{vs}  \texttt{Q}  \texttt{anisotropy\_flag} \texttt{domain\_ID}
-\end{lyxcode}
-where
-\begin{itemize}
-     \item \texttt{Q}           : quality factor (0=no attenuation)
-     \item \texttt{anisotropy\_flag}  : 0=no anisotropy / 1,2,.. check with implementation in \texttt{aniso\_model.f90}
-     \item \texttt{domain\_id}        : 1=acoustic / 2=elastic
-\end{itemize}
-\item [{\texttt{NREGIONS}}] The number of regions in the mesh.
-In the following lines, because the mesh is regular or 'almost regular', each region is defined as :
-\begin{lyxcode}
-  \texttt{NEX\_XI\_BEGIN}  \texttt{NEX\_XI\_END}  \texttt{NEX\_ETA\_BEGIN}  \texttt{NEX\_ETA\_END}  \texttt{NZ\_BEGIN}  \texttt{NZ\_END}  \texttt{material\_ID}
-\end{lyxcode}
-\end{description}
-
-The \texttt{INTERFACES\_FILE} parameter of \texttt{Mesh\_Par\_File} defines the file which contains the settings of the topography grid and of the interfaces grids.
-Topography is defined as a set of elevation values on a regular 2D grid.
-It is also possible to
-define interfaces between the layers of the model in the same way.
-The file needs to define several parameters:
-\begin{itemize}
-\item The number of interfaces, including the topography. This needs to be set at the first line. Then, from the bottom
-to the top of the model, you need to define the grids with:
-\item \texttt{SUPPRESS\_UTM\_PROJECTION} flag as described previously,
-\item number of points along $x$ and $y$ direction (NXI and NETA),
-\item minimal $x$ and $y$ coordinates (LONG\_MIN and LAT\_MIN),
-\item spacing between points along $x$ and $y$ (SPACING\_XI and SPACING\_ETA) and
-\item the name of the file which contains the elevation values (in $y$.$x$ increasing order).
-\end{itemize}
-At the end of this file, you simply need to set the number of spectral elements in the vertical direction for each layer. We provide a few models in the {\texttt{examples/}} directory.  \\
-
-Finally, depending on your system, you might need to provide a file that tells MPI what compute nodes
-to use for the simulations. The file must have a number of entries
-(one entry per line) at least equal to the number of processors needed
-for the run. A sample file is provided in the file \texttt{mymachines}.
-This file is not used by the mesher or solver, but is required by
-the \texttt{go\_mesher} and \texttt{go\_solver} default job submission
-scripts. See Chapter~\ref{cha:Scheduler} for information about running
-the code on a system with a scheduler, e.g., LSF.
-
-Now that you have set the appropriate parameters in the \texttt{Mesh\_Par\_file}
-and have compiled the mesher, you are ready to launch it! This is
-most easily accomplished based upon the \texttt{go\_mesher} script.
-When you run on a PC cluster, the script assumes that the nodes are
-named n001, n002, etc. If this is not the case, change the \texttt{tr
--d `n'} line in the script. You may also need to edit the last command
-at the end of the script that invokes the \texttt{mpirun} command.
-See Chapter~\ref{cha:Scheduler} for information about running the
-code on a system with a scheduler, e.g., LSF.
-
-Mesher output is provided in the \texttt{in\_out\_files/OUTPUT\_FILES} directory
-in \texttt{output\_mesher.txt}; this file provides lots of details
-about the mesh that was generated. Please note that the mesher suggests a time step \texttt{DT} to run
-the solver with. The mesher output file also contains a table about the quality of the mesh to indicate
-possible problems with the distortions of elements.
-Alternatively, output can be directed
-to the screen instead by uncommenting a line in \texttt{constants.h}:
-\begin{lyxcode}
-!~uncomment~this~to~write~messages~to~the~screen~
-
-!~integer,~parameter~::~IMAIN~=~ISTANDARD\_OUTPUT
-\end{lyxcode}
-
-To control the quality of the mesh, check the standard output (either on the screen or in the \texttt{in\_out\_files/OUTPUT\_FILES} directory in \texttt{output\_mesher.txt}) and analyze the skewness histogram of mesh elements to make sure no element has a skewness above approximately 0.75, otherwise the mesh is of poor quality (and if even a single element has a skewness value above 0.80, then you must definitely improve the mesh).
-To draw the skewness histogram on the screen, type \texttt{gnuplot plot\_mesh\_quality\_histogram.gnu}.
-
-\chapter{\label{cha:Creating-Distributed-Databases}Creating the Distributed Databases}
-
-After using \texttt{xmeshfem3D} or \texttt{xdecompose\_mesh\_SCOTCH}, the next step in the workflow is to compile
-\texttt{xgenerate\_}~\\
-\texttt{databases}. This program is going to create all the missing information needed by the SEM solver.
-
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=.6\textwidth]{figures/workflow.jpg}
-\par\end{centering}
-
-\caption{Schematic workflow for a SPECFEM3D simulation. The executable \texttt{xgenerate\_databases} creates the GLL mesh points and assigns specific model parameters.}
-
-\label{fig:workflow.databases}
-\end{figure}
-
-In the main directory, type
-\begin{lyxcode}
-{\small make xgenerate\_databases}{\small \par}
-\end{lyxcode}
-Input for the program is provided through
-the main parameter file \texttt{Par\_file}, which resides in the subdirectory \texttt{in\_data\_files/}. Please note that
-\texttt{xgenerate\_databases} must be called directly from the \texttt{bin/} directory, as most of the binaries of the package.
-
-\section{\label{cha:Main-Parameter}Main parameter file \texttt{Par\_file}}
-
-Before running \texttt{xgenerate\_databases}, a number of parameters need to be set in the main parameter \texttt{Par\_file}
-located in the subdirectory \texttt{in\_data\_files/}:
-
-\begin{description}
-\item [{\texttt{SIMULATION\_TYPE}}] is set to 1 for forward simulations,
-2 for adjoint simulations (see Section \ref{sec:Adjoint-simulation-finite})
-and 3 for kernel simulations (see Section \ref{sec:Finite-Frequency-Kernels}).
-\item [{\texttt{SAVE\_FORWARD}}] is only set to \texttt{.true.} for a forward
-simulation with the last frame of the simulation saved, as part of
-the finite-frequency kernel calculations (see Section \ref{sec:Finite-Frequency-Kernels}).
-For a regular forward simulation, leave \texttt{SIMULATION\_TYPE}
-and \texttt{SAVE\_FORWARD} at their default values.
-\item [{\texttt{\noun{UTM\_PROJECTION\_ZONE}}}] UTM projection zone in
-which your model resides, only valid when \texttt{SUPPRESS\_UTM\_}~\\
-\texttt{PROJECTION} is \texttt{.false.}.
-\item [{\texttt{SUPPRESS\_UTM\_PROJECTION}}] set to be \texttt{.false.}
-when your model range is specified in the geographical coordinates,
-and needs to be \texttt{.true.} when your model is specified in a
-cartesian coordinates. \noun{UTM projection zone in which your simulation
-region resides.}
-\item [{\texttt{NPROC}}] The number of MPI processors, each one is assigned one slice of the whole mesh.
-\item [{\texttt{NSTEP}}] The number of time steps of the simulation. This controls the
-length of the numerical simulation, i.e., twice the number of time steps
-requires twice as much CPU time. This feature is not used at the time
-of generating the distributed databases but is required for the solver, i.e., you may change this
-parameter after running  \texttt{xgenerate\_databases}.
-\item [{\texttt{DT}}] The length of each time step in seconds.
-This feature is not used at the time of generating the distributed databases but is required for the solver.
-Please see also Section~\ref{sec:Choosing-the-Time-Step} for further details.
-%
-\item [{\texttt{MODEL}}] Must be set to one of the following:
-%
-\begin{description}
-\item [{\textmd{Models defined by mesh parameters:}}]~
-\begin{description}
-\item [{\texttt{default}}] Uses model parameters as defined by meshing procedures described
-in the previous Chapter~\ref{cha:Mesh-Generation}.
-\end{description}
-\end{description}
-%
-\begin{description}
-\item [{\textmd{1D~models~with~real~structure:}}]~
-\begin{description}
-\item [{\texttt{1D\_prem}}] Isotropic version of the spherically
-symmetric Preliminary Reference Earth Model (PREM) \citep{DzAn81}.
-\item [{\texttt{1D\_socal}}] A standard isotropic 1D model for Southern California.
-\item [{\texttt{1D\_cascadia}}] Isotropic 1D profile for the Cascadia region.
-\end{description}
-\end{description}
-%
-\begin{description}
-\item [{\textmd{Fully~3D~models:}}]~
-\begin{description}
-\item [{\texttt{aniso}}] For a user-specified fully anisotropic model.
-Parameters are set up in routines located in file \texttt{model\_aniso.f90} in directory \texttt{src/generate\_databases/}.
-See Chapter~\ref{cha:-Changing-the} for a discussion on how to specify your own 3D model.
-\item [{\texttt{external}}] For a user-specified isotropic model which uses externally defined model parameters.
-Uses external model definitions set up in routines located in file \texttt{model\_external\_values.f90} in directory \texttt{src/generate\_databases/}.
-Please modify these generic template routines to use your own model definitions.
-\item [{\texttt{gll}}] For a user-specified isotropic model which uses external binary files for $v_p$, $v_s$ and $\rho$.
-Binary files are given in the same format as when outputted by the \texttt{xgenerate\_databases} executable when using option \texttt{SAVE\_MESH\_FILES}.
-These binary files define the model parameters on all GLL points which can be used for iterative inversion procedures.
-\item [{\texttt{salton\_trough}}] A 3D $V_p$ model for Southern California.
-Users must provide the corresponding data file \texttt{regrid3\_vel\_p.bin} in directory \texttt{DATA/st\_3D\_block\_harvard/}.
-\item [{\texttt{tomo}}] For a user-specified 3D isotropic model which uses a tomographic model file \texttt{tomographic\_model.xyz} in directory \texttt{in\_data\_files/}.
-See Section \ref{sec:Using-tomographic},
-for a discussion on how to specify your own 3D tomographic model.
-\end{description}
-\end{description}
-%
-%
-\item [{\texttt{OCEANS}}] Set to \texttt{.true.} if the effect of the oceans
-on seismic wave propagation should be incorporated based upon the
-approximate treatment discussed in \citet{KoTr02b}. This feature
-is inexpensive from a numerical perspective, both in terms of memory
-requirements and CPU time. This approximation is accurate at periods
-of roughly 20~s and longer. At shorter periods the effect of water
-phases/reverberations is not taken into account, even when the flag
-is on.
-\item [{\texttt{TOPOGRAPHY}}] This feature is only effective if \texttt{OCEANS} is set to \texttt{.true.}.
-Set to \texttt{.true.} if topography and
-bathymetry should be read in based upon the topography file specified in the main
-constants file \texttt{constants.h} found in subdirectory \texttt{src/shared/} to evaluate
-elevations. If not set, elevations will be read from the numerical mesh.
-\item [{\texttt{ATTENUATION}}] Set to \texttt{.true.} if attenuation should
-be incorporated. Turning this feature on increases the memory requirements
-significantly (roughly by a factor of~1.5), and is numerically fairly
-expensive. See \citet{KoTr99,KoTr02a} for a discussion on the implementation
-of attenuation based upon standard linear solids.
-Please note that the Vp- and Vs-velocities of your model are given for a reference frequency.
-To change this reference frequency, you change the value of
-\texttt{ATTENUATION\_f0\_REFERENCE} in the main constants file \texttt{constants.h} found in subdirectory \texttt{src/shared/}.
-\item [{\texttt{USE\_OLSEN\_ATTENUATION}}] Set to \texttt{.true.} if you
-want to use the attenuation model that scaled from the S-wave speed model
-using Olsen's empirical relation (see \citet{OlDaBr03}).
-\item [{\texttt{ANISOTROPY}}] Set to \texttt{.true.} if you
-want to use an anisotropy model. Please see the file \texttt{model\_aniso.f90} in subdirectory \texttt{src/generate\_databases/} for the current implementation of anisotropic models.
-\item [{\texttt{ABSORBING\_CONDITIONS}}] Set to \texttt{.true.} to turn
-on Clayton-Enquist absorbing boundary conditions (see \citet{KoTr99}).
-\item [{\texttt{ABSORB\_INSTEAD\_OF\_FREE\_SURFACE}}] Set to \texttt{.true.} to turn on 
-absorbing boundary conditions on the top surface which by default constitutes a free surface of the model.
-\item [{\texttt{MOVIE\_SURFACE}}] Set to \texttt{.false.}, unless you want
-to create a movie of seismic wave propagation on the Earth's surface.
-Turning this option on generates large output files. See Section~\ref{sec:Movies}
-for a discussion on the generation of movies. This feature is only relevant for the solver.
-\item [{\texttt{MOVIE\_VOLUME}}] Set to \texttt{.false.}, unless you want
-to create a movie of seismic wave propagation in the Earth's interior.
-Turning this option on generates huge output files. See Section~\ref{sec:Movies}
-for a discussion on the generation of movies. This feature is only relevant for the solver.
-\item [{\texttt{NTSTEP\_BETWEEN\_FRAMES}}] Determines the number of timesteps
-between movie frames. Typically you want to save a snapshot every
-100 timesteps. The smaller you make this number the more output will
-be generated! See Section~\ref{sec:Movies} for a discussion on the
-generation of movies. This feature is only relevant for the solver.
-\item [{\texttt{CREATE\_SHAKEMAP}}] Set this flag to \texttt{.true.} to
-create a ShakeMap\textregistered, i.e., a peak ground velocity map
-of the maximum absolute value of the two horizontal components of the velocity vector.
-\item [{\texttt{SAVE\_DISPLACEMENT}}] Set this flag to \texttt{.true.}
-if you want to save the displacement instead of velocity for the movie
-frames.
-\item [{\texttt{USE\_HIGHRES\_FOR\_MOVIES}}] Set this flag to \texttt{.true.}
-if you want to save the values at all the NGLL grid points for the
-movie frames.
-\item [{\texttt{HDUR\_MOVIE}}] determines the half duration of the source time function for the movie simulations.
-When this parameter is set to be 0, a default half duration that corresponds to the accuracy of the simulation is provided.
-Otherwise, it adds this half duration to the half duration specified in the source file \texttt{CMTSOLUTION},
-thus simulates longer periods to make the movie images look smoother.
-\item [{\texttt{SAVE\_MESH\_FILES}}] Set this flag to \texttt{.true.} to
-save ParaView \urlwithparentheses{www.paraview.org}
-mesh files for subsequent viewing. Turning the flag on generates large
-(distributed) files in the \texttt{LOCAL\_PATH} directory. See Section~\ref{sec:Mesh-graphics}
-for a discussion of mesh viewing features.
-\item [{\texttt{LOCAL\_PATH}}] Directory in which the distributed databases
-will be written. Generally one uses a directory on the
-local disk of the compute nodes, although on some machines these databases
-are written on a parallel (global) file system (see also the earlier
-discussion of the \texttt{LOCAL\_PATH\_IS\_ALSO\_GLOBAL} flag in Chapter~\ref{cha:Getting-Started}).
-\texttt{xgenerate\_databases} generates the necessary databases in parallel, one set
-for each of the \texttt{NPROC} slices that constitutes
-the mesh (see Figure~\ref{fig:mount.partitions} and Figure~\ref{fig:For-parallel-computing}). After
-the executable finishes, you can log in to one of the compute nodes and view
-the contents of the \texttt{LOCAL\_PATH} directory to see the (many)
-files generated by \texttt{xgenerate\_databases}.
-Please note that the \texttt{LOCAL\_PATH} directory should already contain the output files of
-the partitioner, i.e. from \texttt{xdecompose\_mesh\_SCOTCH} or \texttt{xmeshfem3D}.
-\item [{\texttt{NTSTEP\_BETWEEN\_OUTPUT\_INFO}}] This parameter specifies
-the interval at which basic information about a run is written to
-the file system (\texttt{timestamp{*}} files in the \texttt{in\_out\_files/OUTPUT\_FILES}
-directory). If you have access to a fast machine, set \texttt{NTSTEP\_BETWEEN\_OUTPUT\_INFO}
-to a relatively high value (e.g., at least 100, or even 1000 or more)
-to avoid writing output text files too often. This feature is not
-used at the time of meshing. One can set this parameter to a larger
-value than the number of time steps to avoid writing output during
-the run.
-\item [{\texttt{NTSTEP\_BETWEEN\_OUTPUT\_SEISMOS}}] This parameter specifies
-the interval at which synthetic seismograms are written in the \texttt{LOCAL\_PATH}
-directory. If a run crashes, you may still find usable (but shorter
-than requested) seismograms in this directory. On a fast machine set
-\texttt{NTSTEP\_BETWEEN\_OUTPUT\_SEISMOS} to a relatively high value
-to avoid writing to the seismograms too often. This feature is only relevant for the solver.
-\item [{\texttt{USE\_FORCE\_POINT\_SOURCE}}] Turn this flag on to use a \texttt{FORCESOLUTION} force source
-instead of a \texttt{CMTSOLUTION} moment-tensor source. This can be useful e.g. for oil industry 
-foothills simulations in which the source is a vertical force, normal force, 
-tilted force, or an impact etc. You will need to give the East, North and vertical components of an 
-arbitrary (non-unitary) direction vector of the force vector; thus refer to Appendix~\ref{cha:Coordinates} 
-for the orientation of the reference frame.
-This vector is made unitary internally in the code and thus only its direction matters here;
-its norm is ignored and the norm of the force used is the factor force source times the source time function.
-\item [{\texttt{PRINT\_SOURCE\_TIME\_FUNCTION}}] Turn this flag on to print
-information about the source time function in the file \texttt{in\_out\_files/OUTPUT\_FILES/plot\_source\_time\_function.txt}.
-This feature is only relevant for the solver.
-\end{description}
-
-\section{\label{sec:Choosing-the-Time-Step}Choosing the time step \texttt{DT}}
-
-The parameter \texttt{DT} sets the length of each time step in seconds. The value of this parameter is crucial for the stability of
-the spectral-element simulation.
-Your time step \texttt{DT} will depend on the minimum ratio between the distance $h$ of neighboring mesh points and the wave speeds $v$ defined in your model.
-The condition for the time step $\Delta t$ is:
-
-\begin{lyxcode}
-$\Delta t < C ~\mathrm{min}_\Omega (~ h / v ~)$
-\end{lyxcode}
-where $C$ is the so-called Courant number and $\Omega$ denotes the model volume.
-The distance $h$ depends on the mesh element size and the number of
-GLL points \texttt{NGLL} specified in the main constants file \texttt{constants.h} located in the \texttt{src/shared/} subdirectory.
-The wave speed $v$ is determined based on your model's P- (or S-) wave speed values.
-
-The database generator \texttt{xgenerate\_databases}, as well as the internal mesher \texttt{xmeshfem3D}, are trying
-to evaluate the value of $\Delta t$ for empirically chosen Courant numbers $C \sim 0.3$.
-If you used the mesher \texttt{xmeshfem3D} to generate your mesh, you should set the value suggested in \texttt{in\_out\_files/OUTPUT\_FILES/output\_mesher.txt} file, which is created after the mesher completed.
-In case you used CUBIT to create the mesh, you might use an arbitrary value when running \texttt{xgenerate\_databases} and then use the value suggested in the ~\\
-\texttt{in\_out\_files/OUTPUT\_FILES/output\_mesher.txt} file after the database generation completed.
-Note that the implemented Newmark time scheme uses this time step globally, thus your simulations become more expensive for very small mesh elements in high wave-speed regions.
-Please be aware of this restriction when constructing your mesh in Chapter~\ref{cha:Mesh-Generation}.
-
-\chapter{\label{cha:Running-the-Solver}Running the Solver \texttt{xspecfem3D}}
-
-Now that you have successfully generated the databases, you are ready to compile
-the solver. In the main directory, type
-\begin{lyxcode}
-{\small make xspecfem3D}{\small \par}
-\end{lyxcode}
-Please note that \texttt{xspecfem3D} must be called directly from
-the \texttt{bin/} directory, as most of the binaries of the package.
-
-The solver needs three input files in the \texttt{in\_data\_files/} directory
-to run:
-\begin{description}
-\item [{\texttt{Par\_file}}] the main parameter file which was discussed in detail in the previous
-Chapter~\ref{cha:Creating-Distributed-Databases},
-\item [{\texttt{CMTSOLUTION}} or {\texttt{FORCESOLUTION}}] the earthquake source parameter file or the force source parameter file or, and
-\item [{\texttt{STATIONS}}] the stations file.
-\end{description}
-
-Most parameters in the \texttt{Par\_file}
-should be set prior to running the databases generation. Only the following parameters
-may be changed after running \texttt{xgenerate\_databases}:
-
-\begin{itemize}
-\item the simulation type control parameters: \texttt{SIMULATION\_TYPE}
-and \texttt{SAVE\_FORWARD}
-\item the time step parameters \texttt{NSTEP} and \texttt{DT}
-\item the absorbing boundary control parameter \texttt{ABSORBING\_CONDITIONS} on condition that the\\ 
-  \texttt{ABSORB\_INSTEAD\_OF\_FREE\_SURFACE} flag remains unmodified after running the databases generation. 
-\item the movie control parameters \texttt{MOVIE\_SURFACE}, \texttt{MOVIE\_VOLUME},
-and \texttt{NTSTEPS\_BETWEEN\_FRAMES}
-\item the ShakeMap\textregistered option \texttt{CREATE\_SHAKEMAP}
-\item the output information parameters \texttt{NTSTEP\_BETWEEN\_OUTPUT\_INFO}
-and \texttt{NTSTEP\_BETWEEN\_OUTPUT\_}~\\
-\texttt{SEISMOS}
-\item the \texttt{PRINT\_SOURCE\_TIME\_FUNCTION} flags
-\end{itemize}
-Any other change to the \texttt{Par\_file} implies rerunning both
-the database generator \texttt{xgenerate\_databases} and the solver \texttt{xspecfem3D}.
-
-For any particular earthquake, the \texttt{CMTSOLUTION} file that
-represents the point source may be obtained directly from the Harvard Centroid-Moment Tensor (CMT) web page \urlwithparentheses{www.seismology.harvard.edu}.
-It looks like this:
-
-\begin{lyxcode}
-{\small }%
-\begin{figure}[H]
-\noindent \begin{centering}
-{\small \includegraphics[width=1\textwidth]{figures/Hollywood_CMT} }
-\par\end{centering}{\small \par}
-
-\caption{\label{fig:CMTSOLUTION-file}\texttt{CMTSOLUTION} file based on the
-format from the Harvard CMT catalog. \textbf{M} is the moment tensor,
-$M_{0}${\small{} }is the seismic moment, and $M_{w}$ is the moment
-magnitude.}
-
-\end{figure}
-{\small \par}
-\end{lyxcode}
-The \texttt{CMTSOLUTION} should be edited in the following way:
-
-\begin{itemize}
-\item Set the \texttt{time shift} parameter equal to $0.0$ (the solver
-will not run otherwise.) The time shift parameter would simply apply
-an overall time shift to the synthetics, something that can be done
-in the post-processing (see Section \ref{sec:Process-data-and-syn}).
-\item For point-source simulations (see finite sources, page \pageref{To-simulate-a})
-we recommend setting the source half-duration parameter \texttt{half
-duration} equal to zero, which corresponds to simulating a step source-time
-function, i.e., a moment-rate function that is a delta function. If
-\texttt{half duration} is not set to zero, the code will use a Gaussian
-(i.e., a signal with a shape similar to a `smoothed triangle', as
-explained in \citet{KoTr02a} and shown in Fig~\ref{fig:gauss.vs.triangle})
-source-time function with half-width \texttt{half duration}. We prefer
-to run the solver with \texttt{half duration} set to zero and convolve
-the resulting synthetic seismograms in post-processing after the run,
-because this way it is easy to use a variety of source-time functions
-(see Section \ref{sec:Process-data-and-syn}). \citet{KoTr02a} determined
-that the noise generated in the simulation by using a step source
-time function may be safely filtered out afterward based upon a convolution
-with the desired source time function and/or low-pass filtering. Use
-the serial code \texttt{convolve\_source\_timefunction.f90} and the
-script \texttt{convolve\_source\_timefunction.csh} for this purpose,
-or alternatively use signal-processing software packages such as SAC \urlwithparentheses{www.llnl.gov/sac}.
-Type
-\begin{lyxcode}
-{\small make xconvolve\_source\_timefunction}{\small \par}
-\end{lyxcode}
-to compile the code and then set the parameter \texttt{hdur} in \texttt{convolve\_source\_timefunction.csh}
-to the desired half-duration.
-
-\item The zero time of the simulation corresponds to the center of the triangle/Gaussian,
-or the centroid time of the earthquake. The start time of the simulation
-is $t=-1.5*\texttt{half duration}$ (the 1.5 is to make sure the moment
-rate function is very close to zero when starting the simulation).
-To convert to absolute time $t_{\mathrm{abs}}$, set
-
-\begin{lyxcode}
-$t_{\mathrm{abs}}=t_{\mathrm{pde}}+\texttt{time shift}+t_{\mathrm{synthetic}}$
-\end{lyxcode}
-where $t_{\mathrm{pde}}$ is the time given in the first line of the
-\texttt{CMTSOLUTION}, \texttt{time shift} is the corresponding value
-from the original \texttt{CMTSOLUTION} file and $t_{\mathrm{synthetic}}$
-is the time in the first column of the output seismogram.
-
-\end{itemize}
-%
-\begin{figure}
-\noindent \begin{centering}
-%%%\includegraphics[width=3in]{figures/gauss_vs_triangle_mod.eps}
-\includegraphics[width=3in]{figures/gauss_vs_triangle_mod.pdf}
-\par\end{centering}
-
-\caption{Comparison of the shape of a triangle and the Gaussian function actually
-used.}
-
-\label{fig:gauss.vs.triangle}
-\end{figure}
-
-Centroid latitude and longitude should be provided in geographical
-coordinates. The code converts these coordinates to geocentric coordinates~\citep{DaTr98}.
-Of course you may provide your own source representations by designing
-your own \texttt{CMTSOLUTION} file. Just make sure that the resulting
-file adheres to the Harvard CMT conventions (see Appendix~\ref{cha:Coordinates}).
-Note that the first line in the \texttt{CMTSOLUTION} file is the Preliminary Determination of Earthquakes (PDE) solution performed by the USGS NEIC, which is used as a seed for the Harvard CMT inversion.
-The PDE solution is based upon P waves and often gives the hypocenter of the earthquake, i.e., the rupture initiation point, whereas the CMT solution gives the `centroid location', which is the location with dominant moment release.
-The PDE solution is not used by our software package but must be present anyway in the first line of the file.
-
-\label{To-simulate-a}To simulate a kinematic rupture, i.e., a finite-source
-event, represented in terms of $N_{\mathrm{sources}}$ point sources,
-provide a \texttt{CMTSOLUTION} file that has $N_{\mathrm{sources}}$
-entries, one for each subevent (i.e., concatenate $N_{\mathrm{sources}}$
-\texttt{CMTSOLUTION} files to a single \texttt{CMTSOLUTION} file).
-At least one entry (not necessarily the first) must have a zero \texttt{time
-shift}, and all the other entries must have non-negative \texttt{time
-shift}. Each subevent can have its own half duration, latitude, longitude,
-depth, and moment tensor (effectively, the local moment-density tensor).
-
-Note that the zero in the synthetics does NOT represent the hypocentral
-time or centroid time in general, but the timing of the \textit{center}
-of the source triangle with zero \texttt{time shift} (Fig~\ref{fig:source_timing}).
-
-Although it is convenient to think of each source as a triangle, in
-the simulation they are actually Gaussians (as they have better frequency
-characteristics). The relationship between the triangle and the gaussian
-used is shown in Fig~\ref{fig:gauss.vs.triangle}. For finite fault
-simulations it is usually not advisable to use a zero half duration
-and convolve afterwards, since the half duration is generally fixed
-by the finite fault model.
-
-\vspace{1cm}
-
-\noindent The \texttt{FORCESOLUTION} should be edited in the following way:
-
-\begin{itemize}
-\item Set the \texttt{time shift} parameter equal to $0.0$ (the solver
-will not run otherwise.) The time shift parameter would simply apply
-an overall time shift to the synthetics, something that can be done
-in the post-processing (see Section \ref{sec:Process-data-and-syn}).
-\item Set the latitude, longitude, depth of the source in geographical
-  coordinates.
-\item Set the magnitude of the force source.
-\item Set the components of a (non-unitary) direction vector for the
-  force source in the East/North/Vertical basis (see Appendix A for
-  the orientation of the reference frame).
-\end{itemize}
-
-{\small }%
-\begin{figure}[H]
-\noindent \begin{centering}
-{\small \includegraphics[width=5in]{figures/source_timing.pdf} }
-\par\end{centering}{\small \par}
-
-\caption{Example of timing for three sources. The center of the first source
-triangle is defined to be time zero. Note that this is NOT in general
-the hypocentral time, or the start time of the source (marked as tstart).
-The parameter \texttt{time shift} in the \texttt{CMTSOLUTION} file
-would be t1(=0), t2, t3 in this case, and the parameter \texttt{half
-duration} would be hdur1, hdur2, hdur3 for the sources 1, 2, 3 respectively.}
-
-{\small \label{fig:source_timing} }
-\end{figure}
-{\small \par}
-
-The solver can calculate seismograms at any number of stations for
-basically the same numerical cost, so the user is encouraged to include
-as many stations as conceivably useful in the \texttt{STATIONS} file,
-which looks like this:
-
-{\small }%
-\begin{figure}[H]
-\noindent \begin{centering}
-{\small \includegraphics{figures/STATIONS_basin_explained} }
-\par\end{centering}{\small \par}
-
-\caption{Sample \texttt{STATIONS} file. Station latitude and longitude should
-be provided in geographical coordinates. The width of the station
-label should be no more than 32 characters (see \texttt{MAX\_LENGTH\_STATION\_NAME}
-in the \texttt{constants.h} file), and the network label should be
-no more than 8 characters (see \texttt{MAX\_LENGTH\_NETWORK\_NAME}
-in the \texttt{constants.h} file).}
-
-\end{figure}
-{\small \par}
-
-Each line represents one station in the following format:
-
-\begin{lyxcode}
-{\small Station~Network~Latitude~(degrees)~Longitude~(degrees)~Elevation~(m)~burial~(m)~}{\small \par}
-\end{lyxcode}
-The solver \texttt{xspecfem3D} filters the list of stations in file \texttt{in\_data\_files/STATIONS}
-to exclude stations that are not located within the region given in
-the \texttt{Par\_file} (between \texttt{LATITUDE\_MIN} and \texttt{LATITUDE\_MAX}
-and between \texttt{LONGITUDE\_MIN} and \texttt{LONGITUDE\_MAX}).
-The filtered file is called \texttt{in\_data\_files/STATIONS\_FILTERED}.
-
-Solver output is provided in the \texttt{in\_out\_files/OUTPUT\_FILES} directory
-in the \texttt{output\_solver.txt} file. Output can be directed to
-the screen instead by uncommenting a line in \texttt{constants.h}:
-
-\begin{lyxcode}
-!~uncomment~this~to~write~messages~to~the~screen~
-
-!~integer,~parameter~::~IMAIN~=~ISTANDARD\_OUTPUT~
-\end{lyxcode}
-On PC clusters the seismogram files are generally written to the local
-disks (the path \texttt{LOCAL\_PATH} in the \texttt{Par\_file}) and
-need to be gathered at the end of the simulation.
-
-While the solver is running, its progress may be tracked by monitoring
-the `\texttt{\small timestamp{*}}' files in the ~\\
-\texttt{\small in\_out\_files/OUTPUT\_FILES} directory.
-These tiny files look something like this:
-
-\begin{lyxcode}
-Time~step~\#~~~~~~~~~~10000~
-
-Time:~~~~~108.4890~~~~~~seconds~
-
-Elapsed~time~in~seconds~=~~~~1153.28696703911~
-
-Elapsed~time~in~hh:mm:ss~=~~~~~0~h~19~m~13~s~
-
-Mean~elapsed~time~per~time~step~in~seconds~=~~~~~0.115328696703911~
-
-Max~norm~displacement~vector~U~in~all~slices~(m)~=~~~~1.0789589E-02~
-\end{lyxcode}
-The \texttt{\small timestamp{*}} files provide the \texttt{Mean elapsed
-time per time step in seconds}, which may be used to assess performance
-on various machines (assuming you are the only user on a node), as
-well as the \texttt{Max norm displacement vector U in all slices~(m)}.
-If something is wrong with the model, the mesh, or the source, you
-will see the code become unstable through exponentially growing values
-of the displacement and fluid potential with time, and ultimately
-the run will be terminated by the program. You can control the rate
-at which the timestamp files are written based upon the parameter
-\texttt{NTSTEP\_BETWEEN\_OUTPUT\_INFO} in the \texttt{Par\_file}.
-
-Having set the \texttt{Par\_file} parameters, and having provided the
-\texttt{CMTSOLUTION} (or the \texttt{FORCESOLUTION}) and
-\texttt{STATIONS} files, you are now ready to launch the solver! This
-is most easily accomplished based upon the \texttt{go\_solver} script
-(See Chapter \ref{cha:Scheduler} for information about running through
-a scheduler, e.g., LSF). You may need to edit the last command at the
-end of the script that invokes the \texttt{mpirun} command. The
-\texttt{runall} script compiles and runs both
-\texttt{xgenerate\_databases} and \texttt{xspecfem3D} in
-sequence. This is a safe approach that ensures using the correct
-combination of distributed database output and solver input.
-
-It is important to realize that the CPU and memory requirements of
-the solver are closely tied to choices about attenuation (\texttt{ATTENUATION})
-and the nature of the model (i.e., isotropic models are cheaper than
-anisotropic models). We encourage you to run a variety of simulations
-with various flags turned on or off to develop a sense for what is
-involved.
-
-For the same model, one can rerun the solver for different events by
-simply changing the \texttt{CMTSOLUTION} or \texttt{FORCESOLUTION}
-file, or for different stations by changing the \texttt{STATIONS}
-file. There is no need to rerun the \texttt{xgenerate\_databases}
-executable. Of course it is best to include as many stations as
-possible, since this does not add to the cost of the simulation.
-
-\chapter{\label{cha:Adjoint-Simulations}Adjoint Simulations}
-
-Adjoint simulations are generally performed for two distinct applications.
-First, they can be used for earthquake source inversions, especially
-earthquakes with large ruptures such as the Lander's earthquake \citep{WaHe94}.
-Second, they can be used to generate finite-frequency sensitivity
-kernels that are a critical part of tomographic inversions based upon
-3D reference models \citep{TrTaLi05,LiTr06,TrKoLi08,LiTr08}. In either
-case, source parameter or velocity structure updates are sought to
-minimize a specific misfit function (e.g., waveform or traveltime
-differences), and the adjoint simulation provides a means of computing
-the gradient of the misfit function and further reducing it in successive
-iterations. Applications and procedures pertaining to source studies
-and finite-frequency kernels are discussed in Sections~\ref{sec:Adjoint-simulation-sources}
-and \ref{sec:Adjoint-simulation-finite}, respectively. The two related
-parameters in the \texttt{Par\_file} are \texttt{SIMULATION\_TYPE}
-(1 or 2) and the \texttt{SAVE\_FORWARD} (boolean).
-
-\section{\label{sec:Adjoint-simulation-sources}Adjoint Simulations for Sources}
-
-In the case where a specific misfit function is minimized to invert
-for the earthquake source parameters, the gradient of the misfit function
-with respect to these source parameters can be computed by placing
-time-reversed seismograms at the receivers and using them as sources
-in an adjoint simulation, and then the value of the gradient is obtained
-from the adjoint seismograms recorded at the original earthquake location.
-
-\begin{enumerate}
-\item \textbf{Prepare the adjoint sources} \label{enu:Prepare-the-adjoint}
-
-\begin{enumerate}
-\item First, run a regular forward simlation (\texttt{SIMULATION\_TYPE =
-1} and \texttt{SAVE\_FORWARD = .false.}). You can automatically set
-these two variables using the \texttt{\small utils/change\_simulation\_type.pl}
-script:
-
-\begin{lyxcode}
-utils/change\_simulation\_type.pl~-f~
-\end{lyxcode}
-and then collect the recorded seismograms at all the stations given
-in \texttt{in\_data\_files/STATIONS}.
-
-\item Then select the stations for which you want to compute the time-reversed
-adjoint sources and run the adjoint simulation, and compile them into
-the \texttt{in\_data\_files/STATIONS\_ADJOINT} file, which has the same format
-as the regular \texttt{in\_data\_files/STATIONS} file.
-
-\begin{itemize}
-\item Depending on what type of misfit function is used for the source inversion,
-adjoint sources need to be computed from the original recorded seismograms
-for the selected stations and saved in the \texttt{in\_out\_files/SEM/} directory
-with the format \texttt{STA.NT.BX?.adj}, where \texttt{STA}, \texttt{NT}
-are the station name and network code given in the \texttt{in\_data\_files/STATIONS\_ADJOINT}
-file, and \texttt{BX?} represents the component name of a particular
-adjoint seismogram. Please note that the band code can change depending on
-your sampling rate (see Appendix~\ref{cha:channel-codes} for further details).
-\item The adjoint seismograms are in the same format as the original seismogram
-(\texttt{STA.NT.BX?.sem?}), with the same start time, time interval
-and record length.
-\end{itemize}
-\item Notice that even if you choose to time reverse only one component
-from one specific station, you still need to supply all three components
-because the code is expecting them (you can set the other two components
-to be zero).
-\item Also note that since time-reversal is done in the code itself, no
-explicit time-reversing is needed for the preparation of the adjoint
-sources, i.e., the adjoint sources are in the same forward time sense
-as the original recorded seismograms.
-\end{enumerate}
-\item \textbf{Set the related parameters and run the adjoint simulation}\\
-In the \texttt{in\_data\_files/Par\_file}, set the two related parameters to
-be \texttt{SIMULATION\_TYPE = 2} and \texttt{SAVE\_FORWARD = .false.}.
-More conveniently, use the scripts \texttt{utils/change\_simulation\_type.pl}
-to modify the \texttt{Par\_file} automatically (\texttt{change\_simulation\_type.pl
--a}). Then run the solver to launch the adjoint simulation.
-\item \textbf{Collect the seismograms at the original source location}
-
-After the adjoint simulation has completed successfully, collect the
-seismograms from \texttt{LOCAL\_PATH}.
-
-\begin{itemize}
-\item These adjoint seismograms are recorded at the locations of the original
-earthquake sources given by the \texttt{in\_data\_files/CMTSOLUTION} file, and
-have names of the form \texttt{S?????.NT.S??.sem} for the six-component
-strain tensor (\texttt{SNN,SEE,SZZ,SNE,SNZ,SEZ}) at these locations,
-and ~\\
-\texttt{S?????.NT.BX?.sem} for the three-component displacements
-(\texttt{BXN,BXE,BXZ}) recorded at these locations.
-\item \texttt{S?????} denotes the source number; for example, if the original
-\texttt{CMTSOLUTION} provides only a point source, then the seismograms
-collected will start with \texttt{S00001}.
-\item These adjoint seismograms provide critical information for the computation
-of the gradient of the misfit function.
-\end{itemize}
-\end{enumerate}
-
-\section{\label{sec:Adjoint-simulation-finite}Adjoint Simulations for Finite-Frequency
-Kernels (Kernel Simulation)}
-
-Finite-frequency sensitivity kernels are computed in two successive
-simulations (please refer to \citet{LiTr06} and \citet{TrKoLi08} for details).
-
-\begin{enumerate}
-\item \textbf{Run a forward simulation with the state variables saved at
-the end of the simulation}
-
-Prepare the \texttt{\small CMTSOLUTION} and \texttt{\small STATIONS}
-files, set the parameters \texttt{\small SIMULATION\_TYPE}{\small{}
-}\texttt{\small =}{\small{} }\texttt{\small 1} and \texttt{\small SAVE\_FORWARD
-=}{\small{} }\texttt{\small .true.} in the \texttt{Par\_file} (\texttt{change\_simulation\_type
--F}), and run the solver.
-
-\begin{itemize}
-\item Notice that attenuation is not implemented yet for the computation
-of finite-frequency kernels; therefore set \texttt{ATTENUATION = .false.}
-in the \texttt{Par\_file}.
-\item We also suggest you modify the half duration of the \texttt{CMTSOLUTION}
-to be similar to the accuracy of the simulation (see Equation \ref{eq:shortest_period})
-to avoid too much high-frequency noise in the forward wavefield, although
-theoretically the high-frequency noise should be eliminated when convolved
-with an adjoint wavefield with the proper frequency content.
-\item This forward simulation differs from the regular simulations (\texttt{\small SIMULATION\_TYPE}{\small{}
-}\texttt{\small =}{\small{} }\texttt{\small 1} and \texttt{\small SAVE\_FORWARD}{\small{}
-}\texttt{\small =}{\small{} }\texttt{\small .false.}) described in
-the previous chapters in that the state variables for the last time
-step of the simulation, including wavefields of the displacement,
-velocity, acceleration, etc., are saved to the \texttt{LOCAL\_PATH}
-to be used for the subsequent simulation.
-\item For regional simulations, the files recording the absorbing boundary
-contribution are also written to the \texttt{LOCAL\_PATH} when \texttt{SAVE\_FORWARD
-= .true.}.
-\end{itemize}
-\item \textbf{Prepare the adjoint sources}
-
-The adjoint sources need to be prepared the same way as described
-in the Section \ref{enu:Prepare-the-adjoint}.
-
-\begin{itemize}
-\item In the case of travel-time finite-frequency kernel for one source-receiver
-pair, i.e., point source from the \texttt{CMTSOLUTION}, and one station
-in the \texttt{STATIONS\_ADJOINT} list, we supply a sample program
-in \texttt{utils/adjoint\_sources/traveltime/xcreate\_adjsrc\_traveltime} to cut a certain portion of the original
-displacement seismograms and convert it into the proper adjoint source
-to compute the finite-frequency kernel.
-\begin{lyxcode}
-xcreate\_adjsrc\_traveltime~t1~t2~ifile{[}0-5]~E/N/Z-ascii-files~{[}baz]
-\end{lyxcode}
-where \texttt{t1} and \texttt{t2} are the start and end time of the
-portion you are interested in, \texttt{ifile} denotes the component
-of the seismograms to be used (0 for all three components, 1 for East,
-2 for North, and 3 for vertical, 4 for transverse, and 5 for radial
-component), \texttt{E/N/Z-ascii-files} indicate the three-component
-displacement seismograms in the right order, and \texttt{baz} is the
-back-azimuth of the station. Note that \texttt{baz} is only supplied
-when \texttt{ifile} = 4 or 5.
-
-\item Similarly, a sample program to compute adjoint sources for amplitude finite-frequency kernels
-may be found in \texttt{utils/adjoint\_sources/amplitude} and used in the same way as described for
-traveltime measurements
-\begin{lyxcode}
-xcreate\_adjsrc\_amplitude~t1~t2~ifile{[}0-5]~E/N/Z-ascii-files~{[}baz].
-\end{lyxcode}
-
-\end{itemize}
-
-\item \textbf{Run the kernel simulation}
-
-With the successful forward simulation and the adjoint source ready
-in the \texttt{in\_out\_files/SEM/} directory, set \texttt{SIMULATION\_TYPE = 3} and \texttt{SAVE\_FORWARD
-= .false.} in the \texttt{Par\_file} (you can use \texttt{change\_simulation\_type.pl -b}),
-and rerun the solver.
-
-\begin{itemize}
-\item The adjoint simulation is launched together with the back reconstruction
-of the original forward wavefield from the state variables saved from
-the previous forward simulation, and the finite-frequency kernels
-are computed by the interaction of the reconstructed forward wavefield
-and the adjoint wavefield.
-\item The back-reconstructed seismograms at the original station locations
-are saved to the \texttt{LOCAL\_PATH} at the end of the kernel simulations,
-and can be collected to the local disk.
-\item These back-constructed seismograms can be compared with the time-reversed
-original seismograms to assess the accuracy of the backward reconstruction,
-and they should match very well.
-\item The arrays for density, P-wave speed and S-wave speed kernels are
-also saved in the \texttt{LOCAL\_PATH} with the names \texttt{proc??????\_rho(alpha,beta)\_kernel.bin},
-where \texttt{proc??????} represents the processor number, \texttt{rho(alpha,beta)}
-are the different types of kernels.
-\end{itemize}
-\end{enumerate}
-In general, the three steps need to be run sequentially to assure
-proper access to the necessary files. If the simulations are run through
-some cluster scheduling system (e.g., LSF), and the forward simulation
-and the subsequent kernel simulations cannot be assigned to the same
-set of computer nodes, the kernel simulation will not be able to access
-the database files saved by the forward simulation. Solutions for
-this dilemma are provided in Chapter~\ref{cha:Scheduler}. Visualization
-of the finite-frequency kernels is discussed in Section~\ref{sec:Finite-Frequency-Kernels}.
-
-\chapter{Noise Cross-correlation Simulations}
-
-Besides earthquake simulations, SPECFEM3D includes functionality for seismic noise tomography as well.
-In order to proceed successfully in this chapter, it is critical that you have already familiarized yourself
-with procedures for meshing (Chapter \ref{cha:Mesh-Generation}),
-creating distributed databases (Chapter \ref{cha:Creating-Distributed-Databases}),
-running earthquake simulations (Chapters \ref{cha:Running-the-Solver})
-and adjoint simulations (Chapter \ref{cha:Adjoint-Simulations}).
-Also, make sure you read the article `Noise cross-correlation sensitivity kernels' \citep{trompetal2010},
-in order to understand noise simulations from a theoretical perspective.
-
-\section{Input Parameter Files}
-
-As usual, the three main input files are crucial: \texttt{Par\_file}, \texttt{CMTSOLUTION} and \texttt{STATIONS}.
-Unless otherwise specified, those input files should be located in directory \texttt{in\_data\_files/}.\\
-
-\texttt{CMTSOLUTION} is required for all simulations.
-At a first glance, it may seem unexpected to have it here,
-since the noise simulations should have nothing to do with the earthquake -- \texttt{CMTSOLUTION}.
-However, for noise simulations, it is critical to have no earthquakes.
-In other words, the moment tensor specified in \texttt{CMTSOLUTION} must be set to zero manually!\\
-
-\texttt{STATIONS} remains the same as in previous earthquake simulations,
-except that the order of receivers listed in \texttt{STATIONS} is now important.
-The order will be used to determine the `master' receiver,
-i.e., the one that simultaneously cross correlates with the others.\\
-
-\texttt{Par\_file} also requires careful attention.
-A parameter called \texttt{NOISE\_TOMOGRAPHY} has been added which specifies the type of simulation to be run.
-\texttt{NOISE\_TOMOGRAPHY} is an integer with possible values 0, 1, 2 and 3.
-For example, when \texttt{NOISE\_TOMOGRAPHY} equals 0, a regular earthquake simulation will be run.
-When it is 1/2/3, you are about to run
-step 1/2/3 of the noise simulations respectively.
-Should you be confused by the three steps, refer to \citet{trompetal2010} for details.\\
-
-Another change to \texttt{Par\_file} involves the parameter \texttt{NSTEP}.
-While for regular earthquake simulations this parameter specifies the length of synthetic seismograms generated,
-for noise simulations it specifies the length of the seismograms used to compute cross correlations.
-The actual cross correlations are thus twice this length, i.e., $2~\mathrm{NSTEP}-1$.
-The code automatically makes the modification accordingly, if \texttt{NOISE\_TOMOGRAPHY} is not zero.\\
-
-There are other parameters in \texttt{Par\_file} which should be given specific values.
-For instance, since the first two steps for calculating noise sensitivity kernels correspond to
-forward simulations, \texttt{SIMULATION\_TYPE} must be 1 when \texttt{NOISE\_TOMOGRAPHY} equals 1 or 2.
-Also, we have to reconstruct the ensemble forward wavefields in adjoint simulations,
-therefore we need to set \texttt{SAVE\_FORWARD} to \texttt{.true.} for the second step,
-i.e., when \texttt{NOISE\_TOMOGRAPHY} equals 2.
-The third step is for kernel constructions. Hence \texttt{SIMULATION\_TYPE} should be 3, whereas
-\texttt{SAVE\_FORWARD} must be \texttt{.false.}.\\
-
-Finally, for most system architectures,
-please make sure that \texttt{LOCAL\_PATH} in \texttt{Par\_file} is in fact local, not globally shared.
-Because we have to save the wavefields at the earth's surface at every time step,
-it is quite problematic to have a globally shared \texttt{LOCAL\_PATH},
-in terms of both disk storage and I/O speed.
-
-\section{Noise Simulations: Step by Step}
-
-Proper parameters in those parameter files are not enough for noise simulations to run.
-We have more parameters to specify:
-for example, the ensemble-averaged noise spectrum, the noise distribution etc.
-However, since there are a few `new' files, it is better to introduce them sequentially.
-In this section, standard procedures for noise simulations are described.
-
-\subsection{Pre-simulation}
-
-\begin{itemize}
-
-\item
-As usual, we first configure the software package using:
-
-\texttt{./configure FC=ifort MPIFC=mpif90}
-
-Use the following if SCOTCH is needed:
-
-\texttt{./configure FC=ifort MPIFC=mpif90 --with-scotch-dir=/opt/scotch}\\
-
-\item
-Next, we need to compile the source code using:
-
-\texttt{make xgenerate\_databases}
-
-\texttt{make xspecfem3D} \\
-
-\item
-Before we can run noise simulations, we have to specify the noise statistics,
-e.g., the ensemble-averaged noise spectrum.
-Matlab scripts are provided to help you to generate the necessary file.
-
-\texttt{examples/noise\_tomography/NOISE\_TOMOGRAPHY.m}  (main program)\\
-\texttt{examples/noise\_tomography/PetersonNoiseModel.m}
-
-In Matlab, simply run:
-
-\texttt{NOISE\_TOMOGRAPHY(NSTEP, DT, Tmin, Tmax, NOISE\_MODEL)}\\
-
-\texttt{DT} is given in \texttt{Par\_file},
-but \texttt{NSTEP} is NOT the one specified in \texttt{Par\_file}.
-Instead, you have to feed $2~\mathrm{NSTEP}-1$ to account for the doubled length of cross correlations.
-\texttt{Tmin} and \texttt{Tmax} correspond to the period range you are interested in,
-whereas \texttt{NOISE\_MODEL} denotes the noise model you will be using.
-Details can be found in the Matlab script.
-
-After running the Matlab script, you will be given the following information (plus a figure in Matlab):
-
-\texttt{*************************************************************}\\
-\texttt{the source time function has been saved in:}\\
-\texttt{/data2/yangl/3D\_NOISE/S\_squared} (note this path must be different)\\
-\texttt{S\_squared should be put into directory:}\\
-\texttt{in\_out\_files/NOISE\_TOMOGRAPHY/ in the SPECFEM3D package}\\
-
-In other words, the Matlab script creates a file called \texttt{S\_squared},
-which is the first `new' input file we encounter for noise simulations.
-
-One may choose a flat noise spectrum rather than Peterson's noise model.
-This can be done easily by modifying the Matlab script a little.
-
-\item
-Create a new directory in the SPECFEM3D package, name it as \texttt{in\_out\_files/NOISE\_TOMOGRAPHY/}.
-We will add some parameter files later in this folder.
-
-\item
-Put the Matlab-generated-file \texttt{S\_squared} in \texttt{in\_out\_files/NOISE\_TOMOGRAPHY/}.
-
-That's to say, you will have a file \texttt{in\_out\_files/NOISE\_TOMOGRAPHY/S\_squared} in the SPECFEM3D package.
-
-\item
-Create a file called \texttt{in\_out\_files/NOISE\_TOMOGRAPHY/irec\_master\_noise}. Note that this file is located
-in directory \texttt{in\_out\_files/NOISE\_TOMOGRAPHY/} as well. In general, all noise simulation related parameter
-files go into that directory.
-\texttt{irec\_master\_noise} contains only one integer, which is the ID
-of the `master' receiver. For example, if this file contains 5, it means that the fifth receiver
-listed in \texttt{in\_data\_files/STATIONS} becomes the `master'. That's why we mentioned previously that
-the order of receivers in \texttt{in\_data\_files/STATIONS} is important.
-
-Note that in regional simulations, the \texttt{in\_data\_files/STATIONS}
-might contain receivers which are outside of our computational domains.
-Therefore, the integer in \texttt{irec\_master\_noise}
-is actually the ID in \texttt{in\_data\_files/STATIONS\_FILTERED}
-(which is generated by \texttt{bin/xgenerate\_databases}).
-
-\item
-Create a file called \texttt{in\_out\_files/NOISE\_TOMOGRAPHY/nu\_master}. This file holds three numbers,
-forming a (unit) vector. It describes which component we are cross-correlating at the `master'
-receiver, i.e., ${\hat{\bf \nu}}^{\alpha}$ in \citet{trompetal2010}.
-The three numbers correspond to E/N/Z components respectively.
-Most often, the vertical component is used, and in those cases the three numbers should be 0, 0 and 1.
-
-\item
-Describe the noise direction and distributions in \texttt{src/shared/noise\_tomography.f90}.
-Search for a subroutine called \texttt{noise\_distribution\_direction}
-in \texttt{noise\_tomography.f90}. It is actually located at the very beginning
-of \texttt{noise\_tomography.f90}. The default assumes vertical noise and a uniform distribution
-across the whole free surface of the model. It should be quite self-explanatory for modifications.
-Should you modify this part, you have to re-compile the source code.
-
-\end{itemize}
-
-\subsection{Simulations}
-
-With all of the above done, we can finally launch our simulations.
-Again, please make sure that the \texttt{LOCAL\_PATH} in \texttt{in\_data\_files/Par\_file} is not globally shared.
-It is quite problematic to have a globally shared \texttt{LOCAL\_PATH}, in terms of both disk storage and
-speed of I/O (we have to save the wavefields at the earth's surface at every time step).
-
-As discussed in \citet{trompetal2010}, it takes three steps/simulations to obtain one contribution
-of the ensemble sensitivity kernels:
-
-\begin{itemize}
-\item
-Step 1: simulation for generating wavefields\\
-\texttt{SIMULATION\_TYPE=1}\\
-\texttt{NOISE\_TOMOGRAPHY=1}\\
-\texttt{SAVE\_FORWARD} not used, can be either .true. or .false.
-
-\item
-Step 2: simulation for ensemble forward wavefields\\
-\texttt{SIMULATION\_TYPE=1}\\
-\texttt{NOISE\_TOMOGRAPHY=2}\\
-\texttt{SAVE\_FORWARD=.true.}
-
-\item
-Step 3: simulation for ensemble adjoint wavefields and sensitivity kernels\\
-\texttt{SIMULATION\_TYPE=3}\\
-\texttt{NOISE\_TOMOGRAPHY=3}\\
-\texttt{SAVE\_FORWARD=.false.}
-
-Note Step 3 is an adjoint simulation, please refer to previous chapters on how to prepare adjoint sources
-and other necessary files, as well as how adjoint simulations work.
-
-\end{itemize}
-
-Note that it is better to run the three steps continuously within the same job,
-otherwise you have to collect the saved surface movies from the old nodes/CPUs to the new nodes/CPUs.
-This process varies from cluster to cluster and thus cannot be discussed here.
-Please ask your cluster administrator for information/configuration of the cluster you are using.\\
-
-\subsection{Post-simulation}
-
-After those simulations, you have all stuff you need, either in the \texttt{in\_out\_files/OUTPUT\_FILES/} or
-in the directory specified by \texttt{LOCAL\_PATH} in \texttt{in\_data\_files/Par\_file}
-(which are most probably on local nodes).
-Collect whatever you want from the local nodes to your workstation, and then visualize them.
-This process is the same as what you may have done for regular earthquake simulations.
-Refer to other chapters if you have problems.\\
-
-Simply speaking, two outputs are the most interesting: the simulated ensemble cross correlations and
-one contribution of the ensemble sensitivity kernels.\\
-
-The simulated ensemble cross correlations are obtained after the second simulation (Step 2).
-Seismograms in \texttt{in\_out\_files/OUTPUT\_FILES/} are actually the simulated ensemble cross correlations.
-Collect them immediately after Step 2, or the Step 3 will overwrite them.
-Note that we have a `master' receiver specified by \texttt{irec\_master\_noise},
-the seismogram at one station corresponds to the cross correlation between that station and the `master'.
-Since the seismograms have three components, we may obtain cross correlations
-for different components as well, not necessarily the cross correlations between vertical components.\\
-
-One contribution of the ensemble cross-correlation sensitivity kernels are obtained after Step 3,
-stored in the \texttt{in\_data\_files/LOCAL\_PATH} on local nodes.
-The ensemble kernel files are named the same as regular earthquake kernels.\\
-
-You need to run another three simulations to get the other contribution of the ensemble kernels,
-using different forward and adjoint sources given in \citet{trompetal2010}.
-
-\section{Example}
-
-In order to illustrate noise simulations in an easy way, one example is provided in \texttt{examples/noise\_tomography/}.
-See \texttt{examples/noise\_tomography/README} for explanations. \\
-
-Note, however, that they are created for a specific workstation (CLOVER@PRINCETON),
-which has at least 4 cores with `mpif90' working properly. \\
-
-If your workstation is suitable, you can run the example in \texttt{examples/noise\_tomography/} using:\\
-
-\texttt{./pre-processing.sh}\\
-
-Even if this script does not work on your workstation, the procedure it describes is universal.
-You may review the whole process described in the last section by following the commands in \texttt{pre-processing.sh},
-which should contain enough explanations for all the commands.
-
-\chapter{Graphics}
-
-\section{\label{sec:Mesh-graphics}Meshes}
-
-In case you used the internal mesher \texttt{xmeshfem3D} to create and partition your mesh,
-you can output mesh files in ABAQUS (.INP) and DX (.dx) format to visualize them. For this, you must set either the
-flag \texttt{CREATE\_DX\_FILES} or \texttt{CREATE\_ABAQUS\_FILES} to \texttt{.true.}
-in the mesher's parameter file \texttt{Mesh\_Par\_file}
-prior to running the mesher (see Chapter~\ref{cha:Running-the-Mesher-Meshfem3D} for details).
-You can then use AVS \urlwithparentheses{www.avs.com} or OpenDX \urlwithparentheses{www.opendx.org}
-to visualize the mesh and MPI partition (slices).
-
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=.49\textwidth]{figures/vtk_mesh_vp.jpg}
-\includegraphics[width=.49\textwidth]{figures/vtk_mesh_vs.jpg}
-\par\end{centering}
-
-\caption{Visualization using Paraview of VTK files created by \texttt{xgenerate\_databases} showing P- and S-wave velocities assigned to the mesh points.
-The mesh was created by \texttt{xmeshfem3D} for 4 processors.}
-
-\label{fig:vtk.mesh}
-\end{figure}
-
-You have also the option to visualize the distributed databases produced by \texttt{xgenerate\_databases}
-using Paraview \urlwithparentheses{www.paraview.org}. For this, you must set the flag \texttt{SAVE\_MESH\_FILES} to \texttt{.true.}
-in the main parameter file \texttt{Par\_file} (see Chapter~\ref{cha:Main-Parameter} for details).
-This will create VTK files for each single partition. You can then use Paraview \urlwithparentheses{www.paraview.org} to visualized these partitions.
-
-\section{\label{sec:Movies}Movies}
-
-To make a surface or volume movie of the simulation, set parameters
-\texttt{MOVIE\_SURFACE}, \texttt{MOVIE\_VOLUME}, and \texttt{NTSTEP\_BETWEEN\_FRAMES}
-in the \texttt{Par\_file}. Turning on the movie flags, in particular
-\texttt{MOVIE\_VOLUME}, produces large output files. \texttt{MOVIE\_VOLUME}
-files are saved in the \texttt{LOCAL\_PATH} directory, whereas \texttt{MOVIE\_SURFACE}
-output files are saved in the \texttt{in\_out\_files/OUTPUT\_FILES} directory. We
-save the displacement field if the parameter \texttt{SAVE\_DISPLACEMENT} is set, otherwise the velocity field is saved.
-The look of a movie is determined by the
-half-duration of the source. The half-duration should be large enough
-so that the movie does not contain frequencies that are not resolved
-by the mesh, i.e., it should not contain numerical noise. This can
-be accomplished by selecting a CMT \texttt{HALF\_DURATION} > 1.1 $\times$
-smallest period (see figure \ref{fig:CMTSOLUTION-file}). When \texttt{MOVIE\_SURFACE}
-= .\texttt{true.}, the half duration of each source in the \texttt{CMTSOLUTION}
-file is replaced by
-
-\begin{quote}
-\[
-\sqrt{(}\mathrm{\mathtt{HALF\_DURATIO}\mathtt{N}^{2}}+\mathrm{\mathtt{HDUR\_MOVI}\mathtt{E}^{2}})\]
-\textbf{NOTE:} If \texttt{HDUR\_MOVIE} is set to 0.0, the code will
-select the appropriate value of 1.1 $\times$ smallest period. As
-usual, for a point source one can set \texttt{half duration} in the
-\texttt{CMTSOLUTION} file to be 0.0 and \texttt{HDUR\_MOVIE} = 0.0 in the \texttt{Par\_file} to get
-the highest frequencies resolved by the simulation, but for a finite
-source one would keep all the \texttt{half durations} as prescribed
-by the finite source model and set \texttt{HDUR\_MOVIE} = 0.0.
-\end{quote}
-
-\subsection{Movie Surface}
-
-When running \texttt{xspecfem3D} with the \texttt{MOVIE\_SURFACE}
-flag turned on, the code outputs \texttt{moviedata??????} files in
-the \texttt{in\_out\_files/OUTPUT\_FILES} directory. There are several flags
-in the main parameter file \texttt{Par\_file} that control the output of these moviedata files
-(see section \ref{cha:Main-Parameter} for details):
-\texttt{\small NTSTEP\_BETWEEN\_FRAMES} to set the timesteps between frames,
-\texttt{\small SAVE\_DISPLACEMENT} to save displacement instead of velocity,
-\texttt{USE\_HIGHRES\_FOR\_MOVIES} to save values at all GLL point instead of element edges.
-In order to output additionally shakemaps, you would set
-the parameter \texttt{CREATE\_SHAKEMAP} to \texttt{.true.}.
-
-The files are in a fairly complicated
-binary format, but there is a program provided to convert the
-output into more user friendly formats:
-\begin{description}
-\item [\texttt{xcreate\_movie\_shakemap\_AVS\_DX\_GMT}]
-From \texttt{create\_movie\_shakemap\_AVS\_DX\_GMT.f90}, it
-outputs data in ASCII, OpenDX, or AVS format (also readable in ParaView).
-Before compiling the code, make sure you have the file \texttt{surface\_from\_mesher.h}
-in the  \texttt{in\_out\_files/OUTPUT\_FILES/} directory. This file will be created by the solver run.
-Then type
-\begin{lyxcode}
-{\footnotesize make xcreate\_movie\_shakemap\_AVS\_DX\_GMT}{\footnotesize \par}
-\end{lyxcode}
-and run the executable \texttt{xcreate\_movie\_shakemap\_AVS\_DX\_GMT}
-in the \texttt{bin/} subdirectory. It will create visualization files in your format of choice. The code
-will prompt the user for input parameters.
-
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=.32\textwidth]{figures/movie_surf_1.jpg}
-\includegraphics[width=.32\textwidth]{figures/movie_surf_2.jpg}
-\includegraphics[width=.32\textwidth]{figures/movie_surf_3.jpg}
-\par\end{centering}
-
-\caption{Visualization using AVS files created by \texttt{xcreate\_movie\_shakemap\_AVS\_DX\_GMT} showing
-movie snapshots of vertical velocity components at different times.}
-
-\label{fig:movie.surf}
-\end{figure}
-
-\end{description}
-
-The \texttt{SPECFEM3D} code is running in near real-time to produce
-animations of southern California earthquakes via the web; see Southern California ShakeMovie\textregistered \urlwithparentheses{www.shakemovie.caltech.edu}.
-
-\subsection{\label{sub:Movie-Volume}Movie Volume}
-
-When running xspecfem3D with the \texttt{\small MOVIE\_VOLUME} flag
-turned on, the code outputs several files in \texttt{\small LOCAL\_PATH} specified in the main
-\texttt{Par\_file}, e.g. in directory \texttt{in\_out\_files/DATABASES\_MPI}.
-%As the files can be very large, there are several flags in the \texttt{\small Par\_file}
-%that control the region in space and time that is saved. These are:
-%\texttt{\small NTSTEP\_BETWEEN\_FRAMES}
-The output is saved by each processor
-at the time interval specified by \texttt{\small NTSTEP\_BETWEEN\_FRAMES}.
-For all domains, the velocity field is output to files:
-\texttt{\small proc??????\_velocity\_X\_it??????.bin},
-\texttt{\small proc??????\_velocity\_Y\_it??????.bin} and
-\texttt{\small proc??????\_velocity\_Z\_it??????.bin}.
-For elastic domains, the divergence and curl taken from the velocity field, i.e. $\nabla \cdot \bf{v}$
-and $\nabla \times \bf{v}$, get stored as well:
-\texttt{\small proc??????\_div\_it??????.bin},
-\texttt{\small proc??????\_curl\_X\_t??????.bin},
-\texttt{\small proc??????\_curl\_Y\_it??????.bin} and
-\texttt{\small proc??????\_curl\_Z\_it??????.bin}.
-The files denoted
-\texttt{\small proc??????\_div\_}~\\
-\texttt{\small glob\_it??????.bin} and
-\texttt{\small proc??????\_curl\_glob\_it??????.bin}
-are stored on the global points only, all the other arrays
-are stored on all GLL points. Note that the components X/Y/Z can change to E/N/Z according to
-the \texttt{SUPPRESS\_UTM\_PROJECTION} flag (see also Appendix~\ref{cha:Coordinates} and \ref{cha:channel-codes}).
-
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=.32\textwidth]{figures/movie_volume_1.jpg}
-\includegraphics[width=.32\textwidth]{figures/movie_volume_2.jpg}
-\includegraphics[width=.32\textwidth]{figures/movie_volume_3.jpg}
-\par\end{centering}
-
-\caption{Paraview visualization using movie volume files (converted by \texttt{xcombine\_vol\_data} and \texttt{mesh2vtu.pl}) and showing
-snapshots of vertical velocity components at different times.}
-
-\label{fig:movie.volume}
-\end{figure}
-
-To visualize these files, we use an auxilliary program \texttt{combine\_vol\_data.f90}
-to combine the data from all slices into one mesh file. To compile it in the root directory, type:
-\begin{lyxcode}
-{\footnotesize make~xcombine\_vol\_data~}{\footnotesize \par}
-\end{lyxcode}
-which will create the executable \texttt{xcombine\_vol\_data} in the directory \texttt{bin/}. To output the
-usage of this executable, type `./bin/xcombine\_vol\_data` without arguments.
-As an example, to run the executable you would use
-\begin{lyxcode}
-{\footnotesize cd~bin/~}{\footnotesize \par}
-{\footnotesize ./xcombine\_vol\_data~0~3~velocity\_Z\_it000400~../in\_out\_files/DATABASES\_MPI~../in\_out\_files/OUTPUT\_FILES~0~}{\footnotesize \par}
-\end{lyxcode}
-to create a low-resolution mesh file. The output mesh file will have the name \texttt{velocity\_Z\_it000400.mesh}.
-We next convert the \texttt{.mesh} file into the VTU (Unstructured
-grid file) format which can be viewed in ParaView. For this task, you can use and modify the script \texttt{mesh2vtu.pl}
-located in directory \texttt{utils/Visualization/Paraview/}, for example:
-\begin{lyxcode}
-{\footnotesize mesh2vtu.pl~-i~velocity\_Z\_it000400.mesh~-o~velocity\_Z\_it000400.vtu}{\footnotesize \par}
-\end{lyxcode}
-Notice that this Perl script uses a program \texttt{mesh2vtu} in the
-\texttt{utils/Visualization/Paraview/mesh2vtu} directory, which further uses the VTK \urlwithparentheses{http://www.vtk.org/}
-run-time library for its execution. Therefore, make sure you have
-them properly set in the script according to your system.
-
-\section{\label{sec:Finite-Frequency-Kernels}Finite-Frequency Kernels}
-
-The finite-frequency kernels computed as explained in Section \ref{sec:Adjoint-simulation-finite}
-are saved in the \texttt{LOCAL\_PATH} at the end of the simulation.
-Therefore, we first need to collect these files on the front end,
-combine them into one mesh file, and visualize them with some auxilliary
-programs.
-
-\begin{enumerate}
-\item \textbf{Create slice files}
-
-We will only discuss the case of one source-receiver pair, i.e., the
-so-called banana-doughnut kernels. Although it is possible to collect
-the kernel files from all slices on the front end, it usually takes
-up too much storage space (at least tens of gigabytes). Since the
-sensitivity kernels are the strongest along the source-receiver great
-circle path, it is sufficient to collect only the slices that are
-along or close to the great circle path.
-
-A Perl script \texttt{slice\_number.pl} located in directory \texttt{utils/Visualization/Paraview/} can
-help to figure out the slice numbers that lie along the great circle
-path. It applies to meshes created with the internal mesher \texttt{xmeshfem3D}.
-
-\begin{enumerate}
-\item On machines where you have access to the script, copy the \texttt{Mesh\_Par\_file}, and
-\texttt{output\_solver} files, and run:
-
-\begin{lyxcode}
-{\small slice\_number.pl~Mesh\_Par\_file~output\_solver.txt~slice\_file}{\small \par}
-\end{lyxcode}
-which will generate a \texttt{slices\_file}.
-
-\item For cases with multiple sources and multiple receivers, you need to
-provide a slice file before proceeding to the next step.
-\end{enumerate}
-\item \textbf{Collect the kernel files}
-
-After obtaining the slice files, you can collect the corresponding
-kernel files from the given slices.
-
-\begin{enumerate}
-\item You can use or modify the script \texttt{utils/copy\_basin\_database.pl}
-to accomplish this:
-
-\begin{lyxcode}
-{\small utils/copy\_database.pl~slice\_file~lsf\_machine\_file~filename~{[}jobid]}{\small \par}
-\end{lyxcode}
-where \texttt{\small lsf\_machine\_file} is the machine file generated
-by the LSF scheduler, \texttt{\small filename} is the kernel name
-(e.g., \texttt{\small rho\_kernel}, \texttt{\small alpha\_kernel}
-and \texttt{\small beta\_kernel}), and the optional \texttt{\small jobid}
-is the name of the subdirectory under \texttt{\small LOCAL\_PATH}
-where all the kernel files are stored.
-
-\item After executing this script, all the necessary mesh topology files
-as well as the kernel array files are collected to the local directory
-of the front end.
-\end{enumerate}
-\item \textbf{Combine kernel files into one mesh file}
-
-We use an auxilliary program \texttt{combine\_vol\_data.f90}
-to combine the kernel files from all slices into one mesh file.
-
-\begin{enumerate}
-\item Compile it in the root directory:
-
-\begin{lyxcode}
-{\footnotesize make~xcombine\_vol\_data~}{\footnotesize \par}
-
-{\footnotesize ./bin/xcombine\_vol\_data~slice\_list~filename~input\_dir~output\_dir~high/low-resolution~}{\footnotesize \par}
-\end{lyxcode}
-where \texttt{input\_dir} is the directory where all the individual
-kernel files are stored, and \texttt{output\_dir} is where the mesh
-file will be written.
-
-\item Use 1 for a high-resolution mesh, outputting all the GLL points to
-the mesh file, or use 0 for low resolution, outputting only the corner
-points of the elements to the mesh file.
-\item The output mesh file will have the name \texttt{filename\_rho(alpha,beta).mesh}
-\end{enumerate}
-\item \textbf{Convert mesh files into .vtu files}
-
-\begin{enumerate}
-\item We next convert the \texttt{.mesh} file into the VTU (Unstructured
-grid file) format which can be viewed in ParaView. For this task, you can use and modify the script \texttt{mesh2vtu.pl}
-located in directory \texttt{utils/Visualization/Paraview/}, for example:
-
-\begin{lyxcode}
-{\footnotesize mesh2vtu.pl~-i~file.mesh~-o~file.vtu~}{\footnotesize \par}
-\end{lyxcode}
-\item Notice that this Perl script uses a program \texttt{mesh2vtu} in the
-\texttt{utils/Visualization/Paraview/mesh2vtu} directory, which further uses the VTK \urlwithparentheses{http://www.vtk.org/}
-run-time library for its execution. Therefore, make sure you have
-them properly set in the script according to your system.
-\end{enumerate}
-\item \textbf{Copy over the source and receiver .vtk file}
-
-In the case of a single source and a single receiver, the simulation
-also generates the file \texttt{sr.vtk} located in the
-\texttt{in\_out\_files/OUTPUT\_FILES/} directory to describe
-the source and receiver locations, which can also be viewed in Paraview
-in the next step.
-
-\item \textbf{View the mesh in ParaView}
-
-Finally, we can view the mesh in ParaView \urlwithparentheses{www.paraview.org}.
-
-\begin{enumerate}
-\item Open ParaView.
-\item From the top menu, \textsf{File} $\rightarrow$\textsf{Open data},
-select \texttt{file.vtu}, and click the \textsf{Accept} button.
-
-\begin{itemize}
-\item If the mesh file is of moderate size, it shows up on the screen; otherwise,
-only the bounding box is shown.
-\end{itemize}
-\item Click \textsf{Display Tab} $\rightarrow$ \textsf{Display Style} $\rightarrow$
-\textsf{Representation} and select \textsf{wireframe of surface} to
-display it.
-\item To create a cross-section of the volumetric mesh, choose \textsf{Filter}
-$\rightarrow$ \textsf{cut}, and under \textsf{Parameters Tab}, choose
-\textsf{Cut Function} $\rightarrow$ \textsf{plane}.
-\item Fill in center and normal information given by the \texttt{global\_slice\_number.pl}
-script (either from the standard output or from \texttt{normal\_plane.txt}
-file).
-\item To change the color scale, go to \textsf{Display Tab} $\rightarrow$
-\textsf{Color} $\rightarrow$ \textsf{Edit Color Map} and reselect
-lower and upper limits, or change the color scheme.
-\item Now load in the source and receiver location file by \textsf{File}
-$\rightarrow$ \textsf{Open data}, select \texttt{sr.vt}k, and click
-the \textsf{Accept} button. Choose \textsf{Filter} $\rightarrow$
-\textsf{Glyph}, and represent the points by `\textsf{spheres}'.
-\item For more information about ParaView, see the ParaView Users Guide \urlwithparentheses{www.paraview.org/files/v1.6/ParaViewUsersGuide.PDF}.
-\end{enumerate}
-\end{enumerate}
-%
-\begin{figure}[H]
-\noindent \begin{centering}
-\includegraphics[scale=0.6]{figures/3D-S-Kernel}
-\par\end{centering}
-
-\caption{(a) Top Panel: Vertical source-receiver cross-section of the S-wave
-finite-frequency sensitivity kernel $K_{\beta}$ for station GSC at
-an epicentral distance of 176 km from the September 3, 2002, Yorba
-Linda earthquake. Lower Panel: Vertical source-receiver cross-section
-of the 3D S-wave speed model used for the spectral-element simulations
-\citep{KoLiTrSuStSh04}. (b) The same as (a) but for station HEC at
-an epicentral distance of 165 km \citep{LiTr06}.}
-
-\label{figure:P-wave-speed-finite-frequency}
-\end{figure}
-
-\chapter{\label{cha:Scheduler}Running through a Scheduler}
-
-The code is usually run on large parallel machines, often PC clusters,
-most of which use schedulers, i.e., queuing or batch management systems
-to manage the running of jobs from a large number of users. The following
-considerations need to be taken into account when running on a system
-that uses a scheduler:
-
-\begin{itemize}
-\item The processors/nodes to be used for each run are assigned dynamically
-by the scheduler, based on availability. Therefore, in order for the
-\texttt{xgenerate\_databases} and the \texttt{xspecfem3D} executables (or between successive runs of the solver) to
-have access to the same database files (if they are stored on hard
-drives local to the nodes on which the code is run), they must be
-launched in sequence as a single job.
-\item On some systems, the nodes to which running jobs are assigned are
-not configured for compilation. It may therefore be necessary to pre-compile
-both the \texttt{xgenerate\_databases} and the \texttt{xspecfem3D} executables.
-\item One feature of schedulers/queuing systems is that they allow submission
-of multiple jobs in a ``launch and forget'' mode. In order to take
-advantage of this property, care needs to be taken that output and
-intermediate files from separate jobs do not overwrite each other,
-or otherwise interfere with other running jobs.
-\end{itemize}
-Examples of job scripts can be found in the \texttt{\small utils/Cluster/}
-directory and can straightforwardly be modified and adapted to meet
-more specific running needs.
-
-We describe here in some detail a job submission procedure for the
-Caltech 1024-node cluster, CITerra, under the LSF scheduling system.
-We consider the submission of a regular forward simulation using the internal mesher
-to create mesh partitions. The two
-main scripts are \texttt{\small run\_lsf.bash}, which compiles the
-Fortran code and submits the job to the scheduler, and \texttt{\small go\_mesher\_solver\_lsf\_basin.forward},
-which contains the instructions that make up
-the job itself. These scripts can be found in \texttt{\small utils/Cluster/lsf/} directory
-
-\section{Job submission \texttt{run\_lsf.bash}}
-
-This script first sets the job queue to be `normal'. It then compiles
-the mesher, database generator and solver together, figures out the number of processors
-required for this simulation from the \texttt{in\_data\_files/Par\_file}, and
-submits the LSF job.
-
-\begin{lyxcode}
-\#!/bin/bash
-
-\#~use~the~normal~queue~unless~otherwise~directed~queue=\char`\"{}-q~normal\char`\"{}~
-
-if~{[}~\$\#~-eq~1~];~then
-
-~~~~~~~~echo~\char`\"{}Setting~the~queue~to~\$1\char`\"{}
-
-~~~~~~~~queue=\char`\"{}-q~\$1\char`\"{}~
-
-fi~\\
-~\\
-\#~compile~the~mesher~and~the~solver~
-
-d=`date'~echo~\char`\"{}Starting~compilation~\$d\char`\"{}~
-
-make~clean~
-
-make~xmeshfem3D~
-
-make~xgenerate\_databases~
-
-make~xspecfem3D~
-
-d=`date'~
-
-echo~\char`\"{}Finished~compilation~\$d\char`\"{}~\\
-~\\
-\#~get~total~number~of~nodes~needed~for~solver~
-
-NPROC=`grep~NPROC~in\_data\_files/Par\_file~|~cut~-c~34-~'~\\
-~\\
-
-\#~compute~total~number~of~nodes~needed~for~mesher~
-
-NPROC\_XI=`grep~NPROC\_XI~in\_data\_files/meshfem3D\_files/Mesh\_Par\_file~|~cut~-c~34-~'~
-~\\
-
-NPROC\_ETA=`grep~NPROC\_ETA~in\_data\_files/meshfem3D\_files/Mesh\_Par\_file~|~cut~-c~34-~'~
-~\\
-\#~total~number~of~nodes~is~the~product~of~the~values~read~
-
-numnodes=\$((~\$NPROC\_XI~*~\$NPROC\_ETA~))~\\
-~\\
-
-\#~checks~total~number~of~nodes~
-
-if~{[}~\$numnodes~-neq~\$NPROC~];~then
-
-~~~~~~~~echo~\char`\"{}error number of procs mismatch\char`\"{}
-
-~~~~~~~~exit~
-
-fi~\\
-
-~\\
-echo~\char`\"{}Submitting~job\char`\"{}~
-
-bsub~\$queue~-n~\$numnodes~-W~60~-K~<go\_mesher\_solver\_lsf.forward~
-\end{lyxcode}
-
-\section{Job script \texttt{go\_mesher\_solver\_lsf.forward}}
-
-This script describes the job itself, including setup steps that can
-only be done once the scheduler has assigned a job-ID and a set of
-compute nodes to the job, the \texttt{run\_lsf.bash} commands used
-to run the mesher, database generator and the solver, and calls to scripts that collect
-the output seismograms from the compute nodes and perform clean-up
-operations.
-
-\begin{enumerate}
-\item First the script directs the scheduler to save its own output and
-output from \texttt{stdout} into ~\\
-\texttt{\small in\_out\_files/OUTPUT\_FILES/\%J.o},
-where \texttt{\%J} is short-hand for the job-ID; it also tells the
-scheduler what version of \texttt{mpich} to use (\texttt{mpich\_gm})
-and how to name this job (\texttt{go\_mesher\_solver\_lsf}).
-\item The script then creates a list of the nodes allocated to this job
-by echoing the value of a dynamically set environment variable \texttt{LSB\_MCPU\_HOSTS}
-and parsing the output into a one-column list using the Perl script
-\texttt{utils/Cluster/lsf/remap\_lsf\_machines.pl}. It then creates a set of scratch
-directories on these nodes (\texttt{\small /scratch/}~\\
-\texttt{\small \$USER/DATABASES\_MPI}) to be used as the \texttt{LOCAL\_PATH}
-for temporary storage of the database files. The scratch directories
-are created using \texttt{shmux}, a shell multiplexor that can execute
-the same commands on many hosts in parallel. \texttt{shmux} is available
-from Shmux \urlwithparentheses{web.taranis.org/shmux/}. Make sure that the \texttt{LOCAL\_PATH}
-parameter in \texttt{in\_data\_files/Par\_file} is also set properly.
-\item The next portion of the script launches the mesher, database generator and then the solver
-using \texttt{run\_lsf.bash}.
-\item The final portion of the script collects the seismograms and performs
-clean up on the nodes, using the Perl scripts \texttt{collect\_seismo\_lsf\_multi.pl}
-and \texttt{cleanmulti.pl}.
-\end{enumerate}
-\begin{lyxcode}
-\#!/bin/bash~-v~
-
-\#BSUB~-o~in\_out\_files/OUTPUT\_FILES/\%J.o~
-
-\#BSUB~-a~mpich\_gm~
-
-\#BSUB~-J~go\_mesher\_solver\_lsf~\\
-~\\
-\#~set~up~local~scratch~directories
-
-BASEMPIDIR=/scratch/\$USER/DATABASES\_MPI
-
-mkdir~-p~in\_out\_files/OUTPUT\_FILES~
-
-echo~\char`\"{}\$LSB\_MCPU\_HOSTS\char`\"{}~>~in\_out\_files/OUTPUT\_FILES/lsf\_machines~
-
-echo~\char`\"{}\$LSB\_JOBID\char`\"{}~>~in\_out\_files/OUTPUT\_FILES/jobid
-
-remap\_lsf\_machines.pl~in\_out\_files/OUTPUT\_FILES/lsf\_machines~> in\_out\_files/OUTPUT\_FILES/machines
-
-shmux~-M50~-Sall~-c~\char`\"{}rm~-r~-f~/scratch/\$USER;~\textbackslash{}
-
-~~~~~~mkdir~-p~/scratch/\$USER;~mkdir~-p~\$BASEMPIDIR\char`\"{}~\textbackslash{}
-
-~~~~~~-~<~in\_out\_files/OUTPUT\_FILES/machines~>/dev/null~\\
-~\\
-\#~run~the~specfem~program
-
-current\_pwd=\$PWD \\
-
-cd bin/ \\
-
-run\_lsf.bash~-{}-gm-no-shmem~-{}-gm-copy-env~\$current\_pwd/xmeshfem3D~\\
-
-run\_lsf.bash~-{}-gm-no-shmem~-{}-gm-copy-env~\$current\_pwd/xgenerate\_databases~\\
-
-run\_lsf.bash~-{}-gm-no-shmem~-{}-gm-copy-env~\$current\_pwd/xspecfem3D~\\
-~\\
-\#~collect~seismograms~and~clean~up
-
-cd current\_pwd/~ \\
-
-mkdir~-p~in\_out\_files/SEM
-
-cd~in\_out\_files/SEM/~
-
-collect\_seismo.pl~../OUTPUT\_FILES/lsf\_machines
-
-cleanbase.pl~../OUTPUT\_FILES/machines
-\end{lyxcode}
-
-\chapter{{\normalsize \label{cha:-Changing-the}} Changing the Model}
-
-In this section we explain how to change the velocity model used for your simulations.
-These changes involve contributing specific subroutines
-that replace existing subroutines in the \texttt{SPECFEM3D} package.
-Note that \texttt{SPECFEM3D} can handle Earth models with material properties that vary within each spectral element.
-
-%% magnoni 6/12
-\section{{\normalsize \label{sec:Using-tomographic}}Using external tomographic Earth models}
-
-To implement your own external tomographic model, you must provide your own tomography file
-\texttt{tomography\_model.xyz} and choose between two possible options:
-(1) set in the \texttt{Par\_file} the model parameter \texttt{MODEL = tomo}, or
-(2) for more user control, set in the \texttt{Par\_file} the model parameter \texttt{MODEL = default} and use the following format in the file \texttt{MESH/nummaterial\_velocity\_file} when using CUBIT to construct your mesh (see also section \ref{subsec:Exporting-the-Mesh}):
-\begin{lyxcode}
-\texttt{domain\_ID} \texttt{material\_ID} \texttt{tomography} \texttt{elastic} \texttt{file\_name} 1
-\end{lyxcode}
-where
-\texttt{domain\_ID} is 1 for acoustic or 2 for elastic materials,
-\texttt{material\_ID} a negative, unique identifier (i.e., -1),
-\texttt{tomography} keyword for tomographic material definition,
-\texttt{elastic} keyword for elastic material definition,
-\texttt{file\_name} the name of the tomography file and
-1 a positive unique identifier.
-The name of the tomography file is not used so far, rather change the filename (\texttt{TOMO\_FILENAME}) defined in the file \texttt{model\_tomography.f90} located in the \texttt{src/generate\_databases/} directory.
-
-The external tomographic model is represented by a grid of points with assigned material properties and homogeneous resolution along each spatial direction x, y and z.
-The ASCII file \texttt{tomography\_model.xyz} that describes the tomography should be located in the \texttt{in\_data\_files/} directory.
-The format of the file, as read from \texttt{model\_tomography.f90}, looks like Figure \ref{fig:tomography_file},
-where
-%
-\begin{description}
-\item [{\texttt{ORIG\_X}, \texttt{END\_X}}] are, respectively, the coordinates of the initial and final tomographic grid points along the x direction (in the mesh units, e.g., $m$);
-\item [{\texttt{ORIG\_Y}, \texttt{END\_Y}}] respectively the coordinates of the initial and final tomographic grid points along the y direction (in the mesh units, e.g., $m$);
-\item [{\texttt{ORIG\_Z}, \texttt{END\_Z}}] respectively the coordinates of the initial and final tomographic grid points along the z direction (in the mesh units, e.g., $m$);
-\item [\texttt{SPACING\_X}, \texttt{SPACING\_Y}, \texttt{SPACING\_Z}] the spacing between the tomographic grid points along the x, y and z directions, respectively (in the mesh units, e.g., $m$);
-\item [\texttt{NX}, \texttt{NY}, \texttt{NZ}] the number of grid points along the spatial directions x, y and z, respectively; \texttt{NX} is given by [(\texttt{END\_X} - \texttt{ORIG\_X})/\texttt{SPACING\_X}]+1; \texttt{NY} and \texttt{NZ} are the same as \texttt{NX}, but for the y and z directions, respectively;
-\item [\texttt{VP\_MIN}, \texttt{VP\_MAX}, \texttt{VS\_MIN}, \texttt{VS\_MAX}, \texttt{RHO\_MIN}, \texttt{RHO\_MAX}] the minimum and maximum values of the wave speed \texttt{vp} and \texttt{vs} (in $m \, s^{-1}$) and of the density \texttt{rho} (in $kg \, m^{-3}$); these values could be the actual limits of the tomographic parameters in the grid or the minimum and maximum values to which we force the cut of velocity and density in the model.
-\end{description}
-%
-After the first four lines, in the file \texttt{tomography\_model.xyz} the tomographic grid points are listed with the corresponding values of \texttt{vp}, \texttt{vs} and \texttt{rho}, scanning the grid along the x coordinate (from \texttt{ORIG\_X} to \texttt{END\_X} with step of \texttt{SPACING\_X}) for each given y (from \texttt{ORIG\_Y} to \texttt{END\_Y}, with step of \texttt{SPACING\_Y}) and each given z (from \texttt{ORIG\_Z} to \texttt{END\_Z}, with step of \texttt{SPACING\_Z}).
-%fig
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=1.0\textwidth]{figures/tomo_file.jpg}
-\par\end{centering}
-\caption{Tomography file \texttt{tomography\_model.xyz} that describes an external Earth model and that is read by \texttt{model\_tomography.f90}.
-The coordinates x, y and z of the grid, their limits ORIG\_X, ORIG\_Y, ORIG\_Z, END\_X, END\_Y, END\_Z and the grid spacings SPACING\_X, SPACING\_Y, SPACING\_Z should be in the units of the constructed mesh (e.g., UTM coordinates in meters).}
-\label{fig:tomography_file}
-\end{figure}
-
-The user can implement his own interpolation algorithm for the tomography model by changing the routine \texttt{model\_tomography.f90} located in the \texttt{src/generate\_databases/} directory.
-Moreover, for models that involve both fully defined materials and a tomography description, the \texttt{nummaterial\_velocity\_file} has multiple lines each with the corresponding suitable format described above.
-
-\section{{\normalsize \label{sec:Anelastic-Models}}External (an)elastic Models}
-
-To use your own external model, you can set in the \texttt{Par\_file} the model parameter \texttt{MODEL = external}.
-Three-dimensional acoustic and/or (an)elastic (attenuation) models may then be superimposed
-onto the mesh based upon your subroutine in
-\texttt{model\_external\_values.f90} located in directory \texttt{src/generate\_databases}.
-The call to this routine would be as follows
-
-\begin{lyxcode}
-call~model\_external\_values(xmesh,~ymesh,~zmesh, ~\& \\
-~~~~~~~~~~rho,~vp,~vs,~qmu\_atten,~iflag\_aniso,~idomain\_id)
-\end{lyxcode}
-%
-Input to this routine consists of:
-\begin{description}
-\item [{\texttt{xmesh,ymesh,zmesh}}] location of mesh point
-\end{description}
-%
-Output to this routine consists of:
-\begin{description}
-\item [{\texttt{rho,vp,vs}}] isotropic model parameters for density $\rho$ ($kg/m^3$), $v_p$ ($m/s$) and $v_s$ ($m/s$)
-\item [{\texttt{qmu\_atten}}] Shear wave quality factor: $0<Q_{\mu}<9000$
-\item [{\texttt{iflag\_aniso}}] anisotropic model flag, $0$ indicating no anisotropy or $1$ using anisotropic model parameters as defined in routine file \texttt{model\_aniso.f90}
-\item [{\texttt{idomain\_id}}] domain identifier, $1$ for acoustic, $2$ for elastic, $3$ for poroelastic materials.
-\end{description}
-
-Note that the resolution and maximum value of anelastic models are
-truncated. This speeds the construction of the standard linear solids
-during the meshing stage. To change the resolution, currently at one
-significant figure following the decimal, or the maximum value (9000),
-consult \texttt{shared/constants.h}.
-
-\section{{\normalsize \label{sec:Anisotropic-Models}}Anisotropic Models}
-
-To use your anisotropic models, you can either set in the \texttt{Par\_file} the model parameter \texttt{ANISOTROPY} to \texttt{.true.} or set the model parameter \texttt{MODEL = aniso}.
-Three-dimensional anisotropic models may be superimposed on
-the mesh based upon the subroutine in file \texttt{model\_aniso.f90}
-located in directory \texttt{src/generate\_databases}.
-The call to this subroutine is of the form
-
-\begin{lyxcode}
-call~model\_aniso(iflag\_aniso,~rho,~vp,~vs,~\& \\
-~~~~~~~~~~c11,c12,c13,c14,c15,c16,c22,c23,c24,c25,c26,~\& \\
-~~~~~~~~~~c33,c34,c35,c36,c44,c45,c46,c55,c56,c66)~
-\end{lyxcode}
-%
-Input to this routine consists of:
-\begin{description}
-\item [{\texttt{iflag\_aniso}}] flag indicating the type of the anisotropic model, i.e. $0$ for no superimposed anisotropy, $1$ or $2$ for generic pre-defined anisotropic models.
-\item [{\texttt{rho,vp,vs}}] reference isotropic model parameters for density $\rho$, $v_p$ and $v_s$.
-\end{description}
-%
-Output from the routine consists of the following non-dimensional
-model parameters:
-\begin{description}
-\item [{\texttt{c11},}] \textbf{$\cdots$,} \texttt{\textbf{c66}} 21~dimensionalized
-anisotropic elastic parameters.
-\end{description}
-
-You can replace the \texttt{model\_aniso.f90} file by
-your own version \textit{provided you do not change the call structure
-of the routine}, i.e., the new routine should take exactly the same
-input and produce the required relative output.
-
-\chapter{Post-Processing Scripts}
-
-Several post-processing scripts/programs are provided in the \texttt{utils/}
-directory, and most of them need to be adjusted when used on different
-systems, for example, the path of the executable programs. Here we
-only list a few of the available scripts and provide a brief description, and
-you can either refer to the related sections for detailed usage or,
-in a lot of cases, type the script/program name without arguments
-for its usage.
-
-\section{Process Data and Synthetics\label{sec:Process-data-and-syn}}
-
-In many cases, the SEM synthetics are calculated and compared to data
-seismograms recorded at seismic stations. Since the SEM synthetics
-are accurate for a certain frequency range, both the original data
-and the synthetics need to be processed before a comparison can be
-made.
-
- For such comparisons, the following steps are recommended:
-
-\begin{enumerate}
-\item Make sure that both synthetic and observed seismograms have the correct station/event and timing information.
-\item Convolve synthetic seismograms with a source time function with the half duration specified in the \texttt{CMTSOLUTION} file, provided, as recommended, you used a zero half duration in the SEM simulations.
-\item Resample both observed and synthetic seismograms to a common sampling rate.
-\item Cut the records using the same window.
-\item Remove the trend and mean from the records and taper them.
-\item Remove the instrument response from the observed seismograms (recommended) or convolve the synthetic seismograms with the instrument response.
-\item Make sure that you apply the same filters to both observed and synthetic seismograms. Preferably, avoid filtering your records more than once.
-\item Now, you are ready to compare your synthetic and observed seismograms.
-\end{enumerate}
-
-\noindent
-We generally use the following scripts provided in the \texttt{utils/seis\_process/} directory:
-
-\subsection{Data processing script \texttt{process\_data.pl}}
-
-This script cuts a given portion of the original data, filters it,
-transfers the data into a displacement record, and picks the first
-P and S arrivals. For more functionality, type `\texttt{process\_data.pl}'
-without any argument. An example of the usage of the script:
-
-\begin{lyxcode}
-{\footnotesize process\_data.pl~-m~CMTSOLUTION~-s~1.0~-l~0/4000~-i~-f~-t~40/500~-p~-x~bp~DATA/1999.330{*}.BH?.SAC}
-{\footnotesize \par}
-\end{lyxcode}
-
-\noindent which has resampled the SAC files to a sampling rate of 1 seconds, cut
-them between 0 and 4000 seconds, transfered them into displacement
-records and filtered them between 40 and 500 seconds,
-picked the first P and S arrivals, and added suffix `\texttt{bp}' to the file names.
-
-Note that all of the scripts in this section actually use the SAC
-and/or IASP91 to do the core operations; therefore make sure that
-the SAC and IASP91 packages are installed properly on your system,
-and that all the environment variables are set properly before running
-these scripts.
-
-\subsection{Synthetics processing script \texttt{process\_syn.pl}}
-
-This script converts the synthetic output from the SEM code from ASCII
-to SAC format, and performs similar operations as `\texttt{process\_data.pl}'.
-An example of the usage of the script:
-\begin{lyxcode}
-{\footnotesize process\_syn.pl~-m~CMTSOLUTION~-h~-a~STATIONS~-s~1.0~-l~0/4000~-f~-t~40/500~-p~-x~bp~SEM/{*}.BX?.semd}
-{\footnotesize \par}
-\end{lyxcode}
-which will convolve the synthetics with a triangular source-time function
-from the \texttt{CMTSOLUTION} file, convert the synthetics into SAC
-format, add event and station information into the SAC headers, resample the SAC files with a sampling rate of 1 seconds, cut them between 0 and 4000 seconds, filter them between 40 and
-500 seconds with the same filter used for the observed data, pick the first P and S arrivals, and add the suffix `\texttt{bp}'
-to the file names.
-
-More options are available for this script, such as adding time shift
-to the origin time of the synthetics, convolving the synthetics with
-a triangular source time function with a given half duration, etc.
-Type \texttt{process\_syn.pl} without any argument for a detailed
-usage.
-
-In order to convert between SAC format and ASCII files, useful scripts
-are provided in the subdirectories ~\\
-\texttt{utils/sac2000\_alpha\_convert/} and \texttt{utils/seis\_process/asc2sac/}.
-
-\subsection{Script \texttt{rotate.pl}}
-
-The original data and synthetics have three components: vertical (BHZ resp. BXZ),
-north (BHN resp. BXN) and east (BHE resp. BXE). However, for most seismology applications,
-transverse and radial components are also desirable. Therefore, we
-need to rotate the horizontal components of both the data and the
-synthetics to the transverse and radial direction, and \texttt{\small rotate.pl}
-can be used to accomplish this:
-
-\begin{lyxcode}
-rotate.pl~-l~0~-L~180~-d~DATA/{*}.BHE.SAC.bp~
-
-rotate.pl~-l~0~-L~180~SEM/{*}.BXE.semd.sac.bp~
-\end{lyxcode}
-where the first command performs rotation on the SAC data obtained
-through Seismogram Transfer Program (STP) \urlwithparentheses{http://www.data.scec.org/STP/stp.html},
-while the second command rotates the processed SEM synthetics.
-
-\section{Collect Synthetic Seismograms}
-
-The forward and adjoint simulations generate synthetic seismograms
-in the \texttt{in\_out\_files/OUTPUT\_FILES/} directory by default. For the forward simulation, the files
-are named like \texttt{STA.NT.BX?.semd} for two-column time series, or
-\texttt{STA.NT.BX?.semd.sac} for ASCII SAC format, where STA and NT
-are the station name and network code, and \texttt{BX?} stands for
-the component name. Please see the Appendix~\ref{cha:Coordinates} and \ref{cha:channel-codes}
-for further details.
-
-The adjont simulations generate synthetic seismograms
-with the name \texttt{S?????.NT.S??.sem} (refer to Section \ref{sec:Adjoint-simulation-sources}
-for details). The kernel simulations output the back-reconstructed
-synthetic seismogram in the name \texttt{STA.NT.BX?.semd}, mainly
-for the purpose of checking the accuracy of the reconstruction. Refer
-to Section \ref{sec:Adjoint-simulation-finite} for further details.
-
-You do have further options to change this default output behavior, given in the main constants file
-\texttt{constants.h} located in \texttt{src/shared/} directory:
-\begin{description}
-\item [{\texttt{SEISMOGRAMS\_BINARY}}] set to \texttt{.true.} to have seismograms written out in binary format.
-\item [{\texttt{WRITE\_SEISMOGRAMS\_BY\_MASTER}}] Set to \texttt{.true.} to have only the master process writing out seismograms.
-This can be useful on a cluster, where only the master process node has access to the output directory.
-\item [{\texttt{USE\_OUTPUT\_FILES\_PATH}}] Set to \texttt{.false.} to have the seismograms output to \texttt{LOCAL\_PATH} directory specified in the main parameter file \texttt{in\_data\_files/Par\_file}.
-In this case, you could collect the synthetics onto the frontend using the \texttt{collect\_seismo\_lsf\_multi.pl} script
-located in the \texttt{utils/Cluster/lsf/} directory. The usage of the script would be e.g.:
-
-\begin{lyxcode}
-collect\_seismo.pl~machines~in\_data\_files/Par\_file
-\end{lyxcode}
-
-\end{description}
-where  \texttt{machines} is a file containing the node names and  \texttt{in\_data\_files/Par\_file} the parameter file used to extract the \texttt{LOCAL\_PATH} directory used for the simulation.
-
-\section{Clean Local Database}
-
-After all the simulations are done, the seismograms are collected,
-and the useful database files are copied to the frontend, you may
-need to clean the local scratch disk for the next simulation. This
-is especially important in the case of kernel simulation,
-where very large files are generated for the absorbing boundaries
-to help with the reconstruction of the regular forward wavefield.
-A sample script is provided in \texttt{utils/}:
-
-\begin{lyxcode}
-cleanbase.pl~machines
-\end{lyxcode}
-where  \texttt{machines} is a file containing the node names.
-
-\section{Plot Movie Snapshots and Synthetic Shakemaps}
-
-\subsection{Script \texttt{movie2gif.gmt.pl}}
-
-With the movie data saved in \texttt{in\_out\_files/OUTPUT\_FILES/} at the end of
-a movie simulation (\texttt{\small MOVIE\_SURFACE=.true.}), you can
-run the \texttt{`create\_movie\_shakemap\_AVS\_DX\_GMT}' code to convert these binary
-movie data into GMT xyz files for futher processing. A sample script
-\texttt{movie2gif.gmt.pl} is provided to do this conversion, and then
-plot the movie snapshots in GMT, for example:
-
-\begin{lyxcode}
-movie2gif.gmt.pl~-m~CMTSOLUTION~-g~-f~1/40~-n~-2~-p~
-\end{lyxcode}
-which for the first through the 40th movie frame, converts the \texttt{moviedata}
-files into GMT xyz files, interpolates them using the 'nearneighbor'
-command in GMT, and plots them on a 2D topography map. Note that `\texttt{-2}'
-and `\texttt{-p}' are both optional.
-
-\subsection{Script \texttt{plot\_shakemap.gmt.pl}}
-
-With the shakemap data saved in \texttt{in\_out\_files/OUTPUT\_FILES/} at the end
-of a shakemap simulation ~\\
-(\texttt{CREATE\_SHAKEMAP=.true.}), you can
-also run \texttt{`create\_movie\_shakemap\_AVS\_DX\_GMT}' code to convert the binary
-shakemap data into GMT xyz files. A sample script \texttt{plot\_shakemap.gmt.pl}
-is provided to do this conversion, and then plot the shakemaps in
-GMT, for example:
-
-\begin{lyxcode}
-plot\_shakemap.gmt.pl~data\_dir~type(1,2,3)~CMTSOLUTION~
-\end{lyxcode}
-where \texttt{type=1} for a displacement shakemap, \texttt{2} for
-velocity, and \texttt{3} for acceleration.
-
-\section{Map Local Database}
-
-A sample program \texttt{remap\_database} is provided to map the local
-database from a set of machines to another set of machines. This is
-especially useful when you want to run mesher and solver, or different
-types of solvers separately through a scheduler (refer to Chapter~\ref{cha:Scheduler}).
-\begin{lyxcode}
-run\_lsf.bash~-{}-gm-no-shmem~-{}-gm-copy-env~remap\_database~old\_machines~150
-\end{lyxcode}
-where \texttt{old\_machines} is the LSF machine file used in the previous
-simulation, and \texttt{150} is the number of processors in total.
-
-\chapter{\label{cha:tomo}Tomographic inversion using sensitivity kernels}
-
-One of the fundamental reasons for computing sensitivity kernels (Section~\ref{sec:Adjoint-simulation-finite}) is to use them within a tomographic inversion.
-In other words, use recorded seismograms, make measurements with synthetic seismograms, and use the misfit between the two to iteratively improve the model described by (at least) $V_{\rm p}$, $V_{\rm s}$, and $\rho$ as a function of space.
-
-Whatever misfit function you use for the tomographic inversion (several examples in \cite{TrTaLi05} and \cite{TrKoLi08}), you will weight the sensitivity kernels with measurements.
-Furthermore, you will use as many measurements (stations, components, time windows) as possible per event; hence, we call these composite kernels ``event kernels,'' which are volumetric fields representing the gradient of the misfit function with respect to one of the variables (\eg $V_{\rm s}$).
-The basic features of an adjoint-based tomographic inversion were illustrated in \citet{TaLiTr07} and \citet{TrKoLi08} using a conjugate-gradient algorithm; there are dozens of versions of gradient-based inversion algorithms that could alternatively be used.
-The tomographic inversion of \citet{TaLiMaTr09,TaLiMaTr2010} used SPECFEM3D as well as several additional components which are also stored on the CIG svn server, described next.
-
-The directory containing utilities for tomographic inversion using SPECFEM3D (or other packages that evaluate misfit functions and gradients) is here on the CIG svn server:
-%
-\begin{verbatim}
-/cig/seismo/3D/ADJOINT_TOMO/
-   flexwin/     -- FLEXWIN algorithm for automated picking of time windows
-   measure_adj/ -- reads FLEXWIN output file and makes measurements,
-                   with the option for computing adjoint sources
-   iterate_adj/ -- various tools for iterative inversion
-                   (requires pre-computed "event kernels")
-\end{verbatim}
-%
-This directory also contains a brief \verb+README+ file indicating the role of the three subdirectories, \verb+flexwin+ \citep{Maggi2009}, \verb+measure_adj+, and \verb+iterate_adj+.
-The components for making the model update are there; however, there are no explicit rules for computing the model update, just as with any optimization problem.
-There are options for computing a conjugate gradient step, as well as a source subspace projection step.
-The workflow could use substantial ``stitching together,'' and we welcome suggestions; undoubtedly, the tomographic inversion procedure will improve with time.
-{\bf This is a work in progress.}
-
-The best single file to read is probably: \verb+ADJOINT_TOMO/iterate_adj/cluster/README+.
-
-%============================================================
-
-\chapter*{\label{cha:Bug-Reports-and}Bug Reports and Suggestions for Improvements}
-
-To report bugs or suggest improvements to the code, please send an
-e-mail to the CIG Computational Seismology Mailing List \urlwithparentheses{cig-seismo@geodynamics.org}
-or Jeroen Tromp \urlwithparentheses{jtromp-AT-princeton.edu}, and/or use our online
-bug tracking system Roundup \urlwithparentheses{www.geodynamics.org/roundup}.
-
-\addcontentsline{toc}{chapter}{Notes \& Acknowledgments}
-\chapter*{Notes \& Acknowledgments}
-
-In order to keep the software package thread-safe in case a multithreaded
-implementation of MPI is used, developers should not add modules or
-common blocks to the source code but rather use regular subroutine
-arguments (which can be grouped in ``derived types'' if needed for
-clarity).
-
-The Gauss-Lobatto-Legendre subroutines in \texttt{gll\_library.f90}
-are based in part on software libraries from the Massachusetts Institute
-of Technology, Department of Mechanical Engineering (Cambridge, Massachusetts, USA).
-The non-structured global numbering software was provided by Paul
-F. Fischer (Brown University, Providence, Rhode Island, USA, now at Argonne National Laboratory, USA).
-
-OpenDX \urlwithparentheses{http://www.opendx.org} is open-source based on IBM Data
-Explorer, AVS \urlwithparentheses{http://www.avs.com} is a trademark of Advanced
-Visualization Systems, and ParaView \urlwithparentheses{http://www.paraview.org}
-is an open-source visualization platform.
-
-Please e-mail your feedback, questions, comments, and suggestions
-to Jeroen Tromp \urlwithparentheses{jtromp-AT-princeton.edu} or to the CIG Computational Seismology Mailing List \urlwithparentheses{cig-seismo@geodynamics.org}.
-
-\addcontentsline{toc}{chapter}{Copyright}
-\chapter*{Copyright}
-
-Main authors: Dimitri Komatitsch and Jeroen Tromp
-
-Princeton University, USA, and CNRS / INRIA / University of Pau, France
-
-$\copyright$ Princeton University and CNRS / INRIA / University of Pau, July 2012
-
-This program is free software; you can redistribute it and/or modify
-it under the terms of the GNU General Public License as published
-by the Free Software Foundation (see Appendix \ref{cha:License}).\\
-
-\textbf{\underline{Evolution of the code:}}\\
-
- MPI v 2.1, July 2012:
- Max Rietmann, Peter Messmer, Daniel Peter, Dimitri Komatitsch, Joseph Charles, Zhinan Xie: support for CUDA GPUs,
-better CFL stability for the Stacey absorbing conditions. \\
-
- MPI v. 2.0, November 2010:
- Dimitri Komatitsch, Nicolas Le Goff, Roland Martin and Pieyre Le Loher, University of Pau, France,
- Jeroen Tromp and the Princeton group of developers, Princeton University, USA,
- and Emanuele Casarotti, INGV Roma, Italy:
-  support for CUBIT meshes decomposed by SCOTCH, METIS or ZOLTAN;
-  much faster solver using Michel Deville's inlined matrix products.\\
-
- MPI v. 1.4 Dimitri Komatitsch, University of Pau, Qinya Liu and others, Caltech, September 2006:
-  better adjoint and kernel calculations, faster and better I/Os
-  on very large systems, many small improvements and bug fixes.\\
-
- MPI v. 1.3 Dimitri Komatitsch, University of Pau, and Qinya Liu, Caltech, July 2005:
-  serial version, regular mesh, adjoint and kernel calculations, ParaView support.\\
-
- MPI v. 1.2 Min Chen and Dimitri Komatitsch, Caltech, July 2004:
-  full anisotropy, volume movie.\\
-
- MPI v. 1.1 Dimitri Komatitsch, Caltech, October 2002: Zhu's Moho map, scaling
-  of $V_s$ with depth, Hauksson's regional model, attenuation, oceans, movies.\\
-
- MPI v. 1.0 Dimitri Komatitsch, Caltech, USA, May 2002: first MPI version based on global code.\\
-
- Dimitri Komatitsch, IPG Paris, France, December 1996: first 3-D solver for the CM-5 Connection Machine,
-    parallelized on 128 processors using Connection Machine Fortran.\\
-
-\addcontentsline{toc}{chapter}{References}
-\bibliography{bibliography}
-
-\appendix
-
-\chapter{\label{cha:Coordinates}Reference Frame Convention}
-
-The code uses the following convention for the Cartesian reference
-frame:
-
-\begin{itemize}
-\item the $x$ axis points East
-\item the $y$ axis points North
-\item the $z$ axis points up
-\end{itemize}
-Note that this convention is different from both the \citet{AkRi80}
-convention and the Harvard Centroid-Moment Tensor (CMT) convention.
-The Aki \& Richards convention is
-
-\begin{itemize}
-\item the $x$ axis points North
-\item the $y$ axis points East
-\item the $z$ axis points down
-\end{itemize}
-and the Harvard CMT convention is
-
-\begin{itemize}
-\item the $x$ axis points South
-\item the $y$ axis points East
-\item the $z$ axis points up
-\end{itemize}
-
-\subsection*{Source and receiver locations}
-
-The SPECFEM3D software code internally uses Cartesian coordinates.
-The given locations for sources and receiver locations thus may get
-converted.  Sources and receiver locations are read in from the
-\texttt{CMTSOLUTION} (or \texttt{FORCESOLUTION}) and \texttt{STATIONS}
-files.  Note that e.g. the \texttt{CMTSOLUTION} and
-\texttt{FORCESOLUTION} files denotes the location by
-"longitude/latitude/depth". We read in longitude as $x$ coordinate,
-latitude as $y$ coordinate.
-
-In case the flag \texttt{SUPPRESS\_UTM\_PROJECTION} is set to \texttt{.false.}
-in the main parameter file (see Chapter~\ref{cha:Creating-Distributed-Databases}),
-the $x$/$y$ coordinates have to be given in degrees and are converted to Cartesian coordinates
-using a UTM conversion with the specified UTM zone.
-
-The value for depth (given in $km$ in \texttt{CMTSOLUTION} and
-\texttt{FORCESOLUTION}) or burial depth (given in $m$ in
-\texttt{STATIONS}) is evaluated with respect to the surface of the
-mesh at the specified $x$/$y$ location to find the corresponding $z$
-coordinate.  It is possible to use this depth value directly as $z$
-coordinate by changing the flag \texttt{USE\_SOURCES\_RECVS\_Z} to
-\texttt{.true.} in the file \texttt{constants.h} located in the
-\texttt{src/shared/} subdirectory.
-
-\subsection*{Seismogram outputs}
-
-The seismogram output directions are given in Cartesian $x$/$y$/$z$ directions and not rotated any further.
-Changing flags in \texttt{constants.h} in the \texttt{src/shared/} subdirectory only rotates the seismogram outputs
-if receivers are forced to be located at the surface (\texttt{RECVS\_CAN\_BE\_BURIED\_EXT\_MESH} set to \texttt{.false.}) and the normal to the surface at the receiver location should be used (\texttt{EXT\_MESH\_RECV\_NORMAL} set to
-\texttt{.true.}) as vertical.
-In this case, the outputs are rotated to have the vertical component normal to the surface of the mesh, $x$ and $y$ directions are somewhat arbitrary orthogonal directions along the surface.
-
-For the labeling of the seismogram channels, see Appendix~\ref{cha:channel-codes}. Additionally, we add labels to the synthetics using the following convention:
-For a receiver station located in an
-\begin{description}
-\item[elastic domain:] ~\\
-\begin{itemize}
-\item \texttt{semd} for the displacement vector
-\item \texttt{semv} for the velocity vector
-\item \texttt{sema} for the acceleration vector
-\end{itemize}
-\item[acoustic domain:] ~\\
-(please note that receiver stations in acoustic domains must be buried. This is due to the free surface condition which enforces a zero pressure at the surface.)
-\begin{itemize}
-\item \texttt{semd} for the displacement vector
-\item \texttt{semv} for the velocity vector
-\item \texttt{sema} for pressure which will be stored in the vertical component \texttt{Z}.
-Note that pressure in the acoustic media is isotropic and thus the pressure record would be the same in the other two component directions.
-We therefore use the other two seismogram components to store the acoustic potential in component \texttt{X} (or \texttt{N}) and the first time derivative of the acoustic potential in component \texttt{Y} (or \texttt{E}).
-\end{itemize}
-\end{description}
-The seismograms are by default written out in ASCII-format to the \texttt{in\_out\_files/OUTPUT\_FILES/} subdirectory by each parallel process.
-You can change this behavior by changing the following flags in the \texttt{constants.h} file located in the \texttt{src/shared/} subdirectory:
-\begin{description}
-\item [{\texttt{SEISMOGRAMS\_BINARY}}] set to \texttt{.true.} to have seismograms written out in binary format.
-\item [{\texttt{WRITE\_SEISMOGRAMS\_BY\_MASTER}}] Set to \texttt{.true.} to have only the master process writing out seismograms.
-This can be useful on a cluster, where only the master process node has access to the output directory.
-\item [{\texttt{USE\_OUTPUT\_FILES\_PATH}}] Set to \texttt{.false.} to have the seismograms output to \texttt{LOCAL\_PATH} directory specified in the main parameter file \texttt{in\_data\_files/Par\_file}.
-\end{description}
-
-\chapter{\label{cha:channel-codes}Channel Codes of Seismograms}
-
-Seismic networks, such as the Global Seismographic Network (GSN), generally involve various types of instruments with different bandwidths, sampling properties and component configurations.
-There are standards to name channel codes depending on instrument properties.
-IRIS  \texttt{(www.iris.edu)} uses SEED format for channel codes, which are represented by three letters, such as \texttt{LHN}, \texttt{BHZ}, etc.
-In older versions of the SPECFEM3D package, a common format was used for the channel codes of all seismograms, which was \texttt{BHE/BHN/BHZ} for three components.
-To avoid confusion when comparison are made to observed data, we are now using the FDSN convention (\url{http://www.fdsn.org/}) for SEM seismograms.
-In the following, we give a brief explanation of the FDSN convention used by IRIS, and how it is adopted in SEM seismograms.
-Please visit \texttt{www.iris.edu/manuals/SEED\_appA.htm} for further information.\\
-
-\noindent \textbf{\texttt{Band code:}} The first letter in the channel code denotes the band code of seismograms, which depends on the response band and the sampling rate of instruments.
-The list of band codes used by IRIS is shown in Figure \ref{fig:IRIS_band_codes}.
-The sampling rate of SEM synthetics is controlled by the resolution of simulations rather than instrument properties.
-However, for consistency, we follow the FDSN convention for SEM seismograms governed by their sampling rate.
-For SEM synthetics, we consider band codes for which $dt \leq 1$ s.
-IRIS also considers the response band of instruments.
-For instance, short-period and broad-band seismograms with the same sampling rate correspond to different band codes, such as S and B, respectively.
-In such cases, we consider SEM seismograms as broad band, ignoring the corner period ($\geq 10$ s) of the response band of instruments (note that at these resolutions, the minimum period in the SEM synthetics will be less than $10$ s).
-Accordingly, when you run a simulation the band code will be chosen depending on the resolution of the synthetics, and channel codes of SEM seismograms will start with either \texttt{L}, \texttt{M}, \texttt{B}, \texttt{H}, \texttt{C} or \texttt{F}, shown by red color in the figure.\\
-
-\begin{figure}[ht]
-\noindent \begin{centering}
-\includegraphics[scale=0.6]{figures/IRIS_band_codes.pdf}
-\par\end{centering}
-
-\caption{The band code convention is based on the sampling rate and the response band of instruments.
-Please visit \texttt{www.iris.edu/manuals/SEED\_appA.htm} for further information.
-Grey rows show the relative band-code range in SPECFEM3D, and the band codes used to name SEM seismograms are denoted in red.}
-
-\label{fig:IRIS_band_codes}
-\end{figure}
-
-\noindent \textbf{\texttt{Instrument code:}} The second letter in the channel code corresponds to instrument codes, which specify the family to which the sensor belongs.
-For instance, \texttt{H} and \texttt{L} are used for high-gain and low-gain seismometers, respectively.
-The instrument code of SEM seismograms will always be \texttt{X}, as assigned by FDSN for synthetic seismograms.
-\\
-
-\noindent \textbf{\texttt{Orientation code:}} The third letter in channel codes is an orientation code, which generally describes the physical configuration of the components of instrument packages.
-SPECFEM3D uses the traditional orientation code  \texttt{E/N/Z} (East-West, North-South, Vertical) for three components when a UTM projection is used.
-If the UTM conversion is suppressed, i.e.
-the flag \texttt{SUPPRESS\_UTM\_PROJECTION} is set to \texttt{.true.}, then the three components are labelled \texttt{X/Y/Z} according to the Cartesian reference frame. \\
-
-\noindent \textbf{\texttt{EXAMPLE:}} The sampling rate is given by \texttt{DT} in the main parameter file \texttt{Par\_file} located in the \texttt{in\_data\_files/} subdirectory.
-Depending on the resolution of your simulations, if you choose a sampling rate greater than $0.01$ s and less than $1$ s, a seismogram recording displacements on the vertical component of a station \texttt{ASBS} (network \texttt{AZ}) will be named \texttt{ASBS.AZ.MXZ.semd.sac}, whereas it will be \texttt{ASBS.AZ.BXZ.semd.sac}, if the sampling rate is greater than 0.0125 and less equal to 0.1 s.
-
-\chapter{\label{cha:Troubleshooting}Troubleshooting}
-
-\section*{FAQ}
-
-\begin{description}
-\item [configuration fails:]
-
-   Examine the log file 'config.log'. It contains detailed informations.
-   In many cases, the path's to these specific compiler commands F90,
-   CC and MPIF90 won't be correct if `./configure` fails.
-
-   Please make sure that you have a working installation of a Fortran compiler,
-   a C compiler and an MPI implementation. You should be able to compile this
-   little program code:
-
-\begin{lyxcode}
-
-      program main ~\\
-        include 'mpif.h' ~\\
-        integer, parameter :: CUSTOM\_MPI\_TYPE = MPI\_REAL ~\\
-        integer ier ~\\
-        call MPI\_INIT(ier) ~\\
-        call MPI\_BARRIER(MPI\_COMM\_WORLD,ier) ~\\
-        call MPI\_FINALIZE(ier) ~\\
-      end
-
-\end{lyxcode}
-
-\item [compilation fails stating:] ~\\
-\begin{lyxcode}
-    ...  ~\\
-    obj/program\_generate\_databases.o: In function `MAIN\_\_':  ~\\
-    program\_generate\_databases.f90:(.text+0x14): undefined reference to `\_gfortran\_set\_std'  ~\\
-    ...  ~\\
-\end{lyxcode}
-
-  Make sure you're pointing to the right 'mpif90' wrapper command.
-
-  Normally, this message will appear when you're mixing two different Fortran
-  compilers. That is, using e.g. gfortran to compile non-MPI files
-  and mpif90, wrapper provided for e.g. ifort, to compile MPI-files.
-
-  fix: e.g. specify > ./configure FC=gfortran MPIF90=/usr/local/openmpi-gfortran/bin/mpif90
-
-\item [after executing \texttt{xmeshfem3D} I've got elements with skewness of 81\% percent, what does this mean:]
-Look at the skewness table printed in the \texttt{output\_mesher.txt} file after executing \texttt{xmeshfem3D} for the
-example given in \texttt{examples/meshfem3D\_examples/simple\_model/}:
-\begin{lyxcode}
-...~\\
-histogram of skewness (0. good - 1. bad): ~\\
- 0.0000000E+00  -   5.0000001E-02       27648     81.81818      \% ~\\
- 5.0000001E-02  -   0.1000000               0    0.0000000E+00  \% ~\\
-... ~\\
-\end{lyxcode}
-The first line means that you have 27,648 elements with a skewness value between 0 and 0.05 (which means the element is basically not skewed, just plain regular hexahedral element).
-The total number of elements you have in this mesh is (see in the \texttt{output\_mesher.txt} file a bit further down):
-\begin{lyxcode}
-...~\\
-total number of elements in entire mesh:        33792 ~\\
-... ~\\
-\end{lyxcode}
-which gives you that:  27,648 / 33,792 $\sim$ 81.8 \% of all elements are not skewed, i.e. regular elements. a fantastic value :)
-
-The histogram lists for this mesh also some stronger skewed elements, for example the worst ones belong to:
-\begin{lyxcode}
-...~\\
-0.6000000      -   0.6500000            2048     6.060606      \% ~\\
-... ~\\
-\end{lyxcode}
-
-about 6 \% of all elements have distortions with a skewness value between 0.6 and 0.65.
-The skewness values give you a hint of how good your mesh is.
-In an ideal world, you would want to have no distortions, just like the 81\% from above.
-Those elements give you the best approximate values by the GLL quadrature used in the spectral-element method.
-However, having weakly distorted elements is still fine and the solutions are still accurate enough.
-So empirically, values up to around 0.7 are tolerable, above that you should consider remeshing...
-
-To give you an idea why some of the elements are distorted, see the following figure \ref{fig:mesh.vp} of the mesh you obtain in the example \texttt{examples/meshfem3D\_examples/simple\_model/}.
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=.9\textwidth]{figures/mesh_vp.jpg}
-\par\end{centering}
-\caption{Paraview visualization using the mesh vtk-files for the example given in \texttt{examples/meshfem3D\_examples/simple\_model/}.}
-\label{fig:mesh.vp}
-\end{figure}
-You will see that the mesh contains a doubling layer, where we stitch elements together such that the size of two elements will transition to the size of one element (very useful to keep the ratio of wavespeed / element\_size about constant).
-Those elements in this doubling layer have higher skewness values and make up those 6 \% in the histogram.
-
-\item [the code gives following error message "need at least one receiver":]
-This means that no stations given in the input file \texttt{in\_data\_files/STATIONS} could be located within the dimensions of the mesh.
-This can happen for example when the mesh was created with the in-house mesher \texttt{xmeshfem3D} while using the Universal Transverse Mercator (UTM) projection but the simulation with \texttt{xspecfem3D} was suppressing this projection from latitude/longitude to x/y/z coordinates.
-
-In such cases, try to change your \texttt{in\_data\_files/Par\_file} and set e.g.:
-\begin{lyxcode}
-SUPPRESS\_UTM\_PROJECTION         = .false. ~\\
-\end{lyxcode}
-to be the same in \texttt{Mesh\_Par\_file} and \texttt{Par\_file}.
-This flag should be identical when using the in-house mesher \texttt{xmeshfem3D}, \texttt{xgenerate\_databases} and \texttt{xspecfem3D} together to run simulations.
-
-The flag determines if the coordinates you specify for your source and station locations are given as lat/lon degrees and
-must be converted to UTM coordinates.
-As an example, if you use \texttt{.false.} within \texttt{Mesh\_Par\_file} then you create a mesh with \texttt{xmeshfem3D} using the UTM projection from lat/lon as input format to UTM projected coordinates to store the mesh point positions, which is fine.
-The error then may occur if in the \texttt{Par\_file} you have this set to \texttt{.true.} so that the \texttt{xgenerate\_databases} and \texttt{xspecfem3D} suppress the UTM projection and assume that all coordinates you use now for source and receiver locations are given in meters (that is, converted) already.
-So it won't find the specified locations in the used mesh.
-As a solutions, just change the flag in \texttt{Par\_file} to be the same as in \texttt{Mesh\_Par\_file} and rerun \texttt{xgenerate\_databases} and \texttt{xspecfem3D} to make sure that your simulation works fine.
-
-\item [I get the following error message "forward simulation became unstable and blew up":]
-In most cases this means that your time step size \texttt{DT} is chosen too big.
-Look at your files \texttt{output\_mesher.txt} or \texttt{output\_solver.txt} created in the folder \texttt{in\_out\_files/OUTPUT\_FILES/}.
-In these output files, find the section:
-\begin{lyxcode}
-... ~\\
-********************************************* ~\\
-*** Verification of simulation parameters *** ~\\
-********************************************* ~\\
-... ~\\
-*** Minimum period resolved =    4.308774 ~\\
-*** Maximum suggested time step =   6.8863556E-02 ~\\
-... ~\\
-... ~\\
-\end{lyxcode}
-
-then change \texttt{DT} in the \texttt{in\_data\_files/Par\_file} to be somewhere close to the maximum suggested time step. In the example above: \\
-\texttt{DT                              = 0.05d0} \\
-would (most probably) work fine. It could be also bigger than the 0.068 s suggested. This depends a bit on the distortions of
-your mesh elements. The more regular they are, the bigger you can choose \texttt{DT}. Just play with this value a bit and see
-when the simulation becomes stable ...
-
-\end{description}
-
-\chapter{\label{cha:License}License }
+\chapter{License }\label{cha:License}
 
@@
 this when it starts in an interactive mode:
-
 \begin{quote}
-Gnomovision version 69, Copyright $\copyright$ year name of author Gnomovision
-comes with ABSOLUTELY NO WARRANTY; for details type `show w'. This
-is free software, and you are welcome to redistribute it under certain
-conditions; type `show c' for details.
+Gnomovision version 69, Copyright $\copyright$ year name of author
+Gnomovision comes with ABSOLUTELY NO WARRANTY; for details type `show
+w'. This is free software, and you are welcome to redistribute it
+under certain conditions; type `show c' for details.
 \end{quote}
@@
 General Public License instead of this License.
-\end{document}

### 2015-04-24 doc/USER_MANUAL/copyright_and_version.tex

--- 2012-09-28 doc/USER_MANUAL/manual_SPECFEM3D.tex
@@
-%% LyX 1.5.1 created this file.  For more info, see http://www.lyx.org/.
-%% Do not edit unless you really know what you are doing.
-\documentclass[oneside,english,onecolumn]{book}
-\usepackage[T1]{fontenc}
-\usepackage[latin1]{inputenc}
-\usepackage{geometry}
-\geometry{letterpaper,tmargin=1in,bmargin=1in,lmargin=1in,rmargin=1in}
-\setcounter{secnumdepth}{3}
-\setcounter{tocdepth}{3}
-\usepackage{float}
-\usepackage{textcomp}
-\usepackage{amsmath}
+\addcontentsline{toc}{chapter}{Copyright}
 
-% figures
-\usepackage[pdftex]{graphicx}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-% we are running pdflatex, so convert .eps files to .pdf
-\usepackage{epstopdf}
-
-\usepackage{amssymb}
-\IfFileExists{url.sty}{\usepackage{url}}
-                      {\newcommand{\url}{\texttt}}
-\usepackage[authoryear]{natbib}
-
-\makeatletter
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% LyX specific LaTeX commands.
-\newcommand{\noun}[1]{\textsc{#1}}
-%% A simple dot to overcome graphicx limitations
-\newcommand{\lyxdot}{.}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Textclass specific LaTeX commands.
-\newenvironment{lyxcode}
-{\begin{list}{}{
-\setlength{\rightmargin}{\leftmargin}
-\setlength{\listparindent}{0pt}% needed for AMS classes
-\raggedright
-\setlength{\itemsep}{0pt}
-\setlength{\parsep}{0pt}
-\normalfont\ttfamily}%
- \item[]}
-{\end{list}}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% User specified LaTeX commands.
-%\renewcommand{\baselinestretch}{1.5}
-
-% figures
-\usepackage[dvips]{epsfig}
-
-\usepackage{wrapfig}
-
-% fonts
-\usepackage{times}
-
-% hyperlinks to sections and references
-\usepackage[pdftex,bookmarks=true,bookmarksnumbered=true,pdfpagemode=None,pdfstartview=FitH,pdfpagelayout=SinglePage,pdfborder={0 0 0}]{hyperref}
-
-\newcommand{\urlwithparentheses}[1]{(\url{#1})}
-
-% biblio GJI
-\bibliographystyle{abbrvnat}
-
-\newcommand{\toall}[1]{\textbf{*** All: #1 ***}}
-\newcommand{\tojeroen}[1]{\textbf{*** Jeroen: #1 ***}}
-\newcommand{\tobrian}[1]{\textbf{*** Brian: #1 ***}}
-\newcommand{\tovala}[1]{\textbf{*** Vala: #1 ***}}
-\newcommand{\tovalabrian}[1]{\textbf{*** Vala \& Brian: #1 ***}}
-\newcommand{\tovalaqinya}[1]{\textbf{*** Vala \& Qinya: #1 ***}}
-\newcommand{\toqinya}[1]{\textbf{*** Qinya: #1 ***}}
-\newcommand{\tomin}[1]{\textbf{*** Min: #1 ***}}
-\newcommand{\toalessia}[1]{\textbf{*** Alessia: #1 ***}}
-\newcommand{\todimitri}[1]{\textbf{*** Dimitri: #1 ***}}
-
-\newcommand{\nexxi}{\mbox{\texttt{NEX\_XI\/}}}
-\newcommand{\nexeta}{\mbox{\texttt{NEX\_ETA\/}}}
-\newcommand{\nprocxi}{\mbox{\texttt{NPROC\_XI\/}}}
-\newcommand{\nproceta}{\mbox{\texttt{NPROC\_ETA\/}}}
-\newcommand{\nchunks}{\mbox{\texttt{NCHUNKS\/}}}
-\newcommand{\eg}{e.g.,}
-
-% colors for the text
-\usepackage[dvipsnames,usenames]{color}
-\newcommand{\red}[1]{\textcolor{Red}{#1}}
-
-\usepackage{babel}
-\makeatother
-
-\begin{document}
-\begin{center}
-\thispagestyle{empty}%
-\vspace*{-1.8truecm}
-\noindent\makebox[\textwidth]{%
-\noindent\includegraphics[width=0.83\paperwidth]{figures/specfem_3d-cover.pdf}
-}
-\end{center}
-%
-\title{\textbf{SPECFEM3D}\\
-\textbf{User Manual}}
-
-\author{$\copyright$ Princeton University (USA) and\\
-CNRS / INRIA / University of Pau (France)\\
-Version 2.1}
-
-\date{\noindent \today}
-
-\maketitle
-
-\section*{Authors}
-The SPECFEM3D package was first developed by Dimitri Komatitsch and Jeroen Tromp at Harvard University and Caltech, USA, starting in 1998, based on earlier work by Dimitri Komatitsch and Jean-Pierre Vilotte at Institut de Physique du Globe (IPGP) in Paris, France from 1995 to 1997.\\
-
-Since then it has been developed and maintained by a development team: in alphabetical order,
-Piero Basini,
-C\'eline Blitz,
-Ebru Bozda\u{g},
-Emanuele Casarotti,
-Joseph Charles,
-Min Chen,
-Vala Hj\"orleifsd\'ottir,
-Sue Kientz,
-Dimitri Komatitsch,
-Jes\'us Labarta,
-Nicolas Le Goff,
-Pieyre Le Loher,
-Qinya Liu,
-Yang Luo,
-Alessia Maggi,
-Federica Magnoni,
-Roland Martin,
-Dennis McRitchie,
-Matthias Meschede,
-Peter Messmer,
-David Mich\'ea,
-Tarje Nissen-Meyer,
-Daniel Peter,
-Max Rietmann,
-Brian Savage,
-Bernhard Schuberth,
-Anne Sieminski,
-Leif Strand,
-Carl Tape,
-Jeroen Tromp,
-Zhinan Xie,
-Hejun Zhu.\\
-%\toall{add full list of developers here and update the AUTHORS file accordingly}
-
-The cover graphic of the manual was created
-by Santiago Lombeyda from Caltech's Center for Advanced Computing Research (CACR) \urlwithparentheses{http://www.cacr.caltech.edu}. \\
-
-\newpage{}
-
-\tableofcontents{}
-
-\chapter{Introduction}
-
-The software package SPECFEM3D simulates seismic
-wave propagation at the local or regional scale
-based upon the spectral-element method (SEM).
-The SEM is a continuous Galerkin technique \citep{TrKoLi08,PeKoLuMaLeCaLeMaLiBlNiBaTr11},
-which can easily be made discontinous \citep{BeMaPa94,Ch00,KoWoHu02,ChCaVi03,LaWaBe05,Kop06,WiStBuGh10,AcKo11};
-it is then close to a particular case of the discontinuous Galerkin technique \citep{ReHi73,Arn82,FaRi99,HuHuRa99,CoKaSh00,GiHeWa02,RiWh03,MoRi05,GrScSc06,AiMoMu06,BeLaPi06,DuKa06,DeSeWh08,PuAmKa09,WiStBuGh10,DeSe10,EtChViGl10}, with optimized efficiency because of its tensorized basis functions \citep{WiStBuGh10,AcKo11}.
-In particular, it can accurately handle very distorted mesh elements \citep{OlSe11}.\\
-
-It has very good accuracy and convergence properties \citep{MaPa89,SePr94,DeFiMu02,Coh02,DeSe07,SeOl08}.
-The spectral element approach admits spectral rates of convergence and allows exploiting $hp$-convergence schemes.
-It is also very well suited to parallel implementation on very large supercomputers \citep{KoTsChTr03,TsKoChTr03,KoLaMi08a,CaKoLaTiMiLeSnTr08,KoViCh10} as well as on clusters of GPU accelerating graphics cards \citep{KoMiEr09,KoErGoMi10,Kom11}.
-Tensor products inside each element can be optimized to reach very high efficiency \citep{DeFiMu02}, and mesh point and element numbering can be optimized to reduce processor cache misses and improve cache reuse \citep{KoLaMi08a}.
-The SEM can also handle triangular (in 2D) or tetrahedral (in 3D) elements \citep{WinBoyd96,TaWi00,KoMaTrTaWi01,Coh02,MeViSa06} as well as mixed meshes, although with increased cost and reduced accuracy in these elements, as in the discontinuous Galerkin method.\\
-
-Note that in many geological models in the context of seismic wave propagation studies
-(except for instance for fault dynamic rupture studies, in which very high frequencies or supershear rupture need to be modeled near the fault, see e.g.
-\cite{BeGlCrViPi07,BeGlCrVi09,PuAmKa09,TaCrEtViBeSa10})
-a continuous formulation is sufficient because material property contrasts are not drastic and thus
-conforming mesh doubling bricks can efficiently handle mesh size variations \citep{KoTr02a,KoLiTrSuStSh04,LeChLiKoHuTr08,LeChKoHuTr09,LeKoHuTr09}.\\
-
-For a detailed introduction to the SEM as applied to regional
-seismic wave propagation, please consult \citet{PeKoLuMaLeCaLeMaLiBlNiBaTr11,TrKoLi08,KoVi98,KoTr99,ChKoViCaVaFe07} and
-in particular \citet{LeKoHuTr09,LeChKoHuTr09,LeChLiKoHuTr08,GoAmTaCaSmSaMaKo09,WiKoScTr04,KoLiTrSuStSh04}.
-A detailed theoretical analysis of the dispersion
-and stability properties of the SEM is available in \citet{Coh02}, \citet{DeSe07}, \citet{SeOl07} and \citet{SeOl08}.\\
-
-Effects due to lateral variations in compressional-wave speed, shear-wave
-speed, density, a 3D crustal model, topography and bathymetry are
-included. The package can accommodate
-full 21-parameter anisotropy (see~\citet{ChTr07}) as well as lateral
-variations in attenuation \citep{SaKoTr10}. Adjoint capabilities and finite-frequency
-kernel simulations are included \citep{TrKoLi08,PeKoLuMaLeCaLeMaLiBlNiBaTr11,LiTr06,FiIgBuKe09,ViOp09}.\\
-
-The SEM was originally developed in computational fluid dynamics \citep{Pat84,MaPa89}
-and has been successfully adapted to address problems in seismic wave propagation.
-Early seismic wave propagation applications of the SEM, utilizing Legendre basis functions and a
-perfectly diagonal mass matrix, include \cite{CoJoTo93}, \cite{Kom97},
-\cite{FaMaPaQu97}, \cite{CaGa97}, \cite{KoVi98} and \cite{KoTr99},
-whereas applications involving Chebyshev basis functions and a nondiagonal mass matrix
-include \cite{SePr94}, \cite{PrCaSe94} and \cite{SePrPr95}.\\
-
-All SPECFEM3D software is written in Fortran90 with full portability
-in mind, and conforms strictly to the Fortran95 standard. It uses
-no obsolete or obsolescent features of Fortran77. The package uses
-parallel programming based upon the Message Passing Interface (MPI)
-\citep{GrLuSk94,Pac97}.\\
-
-SPECFEM3D won the Gordon Bell award for best performance at the SuperComputing~2003
-conference in Phoenix, Arizona (USA) (see \cite{KoTsChTr03}
-and \url{www.sc-conference.org/sc2003/nr_finalaward.html}).
-It was a finalist again in 2008 for a run at 0.16 petaflops (sustained) on 149,784 processors of the `Jaguar' Cray XT5 system at Oak Ridge National Laboratories (USA) \citep{CaKoLaTiMiLeSnTr08}.
-It also won the BULL Joseph Fourier supercomputing award in 2010.\\
-
-This new release of the code (version 2.1) includes support for GPU graphics card acceleration \citep{KoMiEr09,KoGoErMi10,KoErGoMi10,MiKo10,Kom11}.\\
-
-The next release of the code will include support Convolution or Auxiliary Differential Equation Perfectly Matched absorbing Layers
-(C-PML or ADE-PML) \citep{KoMa07,MaKoEz08,MaKoGe08,MaKo09,MaKoGeBr10}.
-It will also use the PT-SCOTCH parallel library for mesh partitioning.\\
-
-SPECFEM3D includes coupled fluid-solid domains and
-adjoint capabilities, which enables one to address seismological inverse problems, but for linear rheologies only so far.
-To accommodate visco-plastic or non-linear rheologies, the reader can refer to
-the \href{http://geoelse.stru.polimi.it/}{GeoELSE} software package \citep{CaGa97,StPaIg09,ChMoTsBaKrKaStKr10}.
-
-\section{Citation}
-
-If you use SPECFEM3D for your own research, please cite at least one
-of the following articles:
-
-\begin{description}
-
-\item [Numerical simulations in general] ~\\
-Forward and adjoint simulations are described in detail in \cite{TrKoLi08,PeKoLuMaLeCaLeMaLiBlNiBaTr11,VaCaSaKoVi99,KoMiEr09,KoErGoMi10,KoGoErMi10,ChKoViCaVaFe07,MaKoDi09,KoViCh10,CaKoLaTiMiLeSnTr08,TrKoHjLiZhPeBoMcFrTrHu10,KoRiTr02,KoTr02a,KoTr02b,KoTr99} or \cite{KoVi98}.
-Additional aspects dealing with adjoint simulations are described in
-\cite{TrTaLi05,LiTr06,TrKoLi08,LiTr08,TrKoHjLiZhPeBoMcFrTrHu10,PeKoLuMaLeCaLeMaLiBlNiBaTr11}.
-Domain decomposition is explained in detail in \cite{MaKoBlLe08}, and excellent scaling up to 150,000 processor cores is shown for instance in
-\cite{CaKoLaTiMiLeSnTr08,KoLaMi08a,MaKoBlLe08,KoErGoMi10,KoGoErMi10,Kom11},
-
-\item [GPU computing] ~\\
-Computing on GPU graphics cards for acoustic or seismic wave propagation applications
-is described in detail in \cite{KoMiEr09,KoGoErMi10,KoErGoMi10,MiKo10,Kom11}.
-
-\end{description}
-
-\noindent
-If you use this new version 2.1, which has non blocking MPI for much better performance for medium or large runs, please cite at least one of these six articles,
-in which results of non blocking MPI runs are presented: \cite{PeKoLuMaLeCaLeMaLiBlNiBaTr11,KoGoErMi10,KoErGoMi10,KoViCh10,Kom11,CaKoLaTiMiLeSnTr08,MaKoBlLe08}.\\
-
-\noindent
-If you work on geophysical applications, you may be interested in citing some of these application articles as well, among others:
-
-\begin{description}
-
-\item [Southern California simulations] ~\\
-\cite{KoLiTrSuStSh04,KrJiKoTr06a,KrJiKoTr06b}.
-
-If you use the 3D southern California model, please cite
-\citet{SuSh03} (Los Angeles model),
-\citet{lovelyetal06} (Salton Trough), and
-\citet{hauksson2000} (southern California).
-The Moho map was determined by
-\citet{zhu&kanamori2000}.
-The 1D SoCal model was developed by \citet{DrHe90}.
-
-\item [Anisotropy] ~\\
-\cite{ChTr07,JiTsKoTr05,ChFaKo04,FaChKo04,RiRiKoTrHe02,TrKo00}.
-
-\item [Attenuation] ~\\
-\cite{SaKoTr10,KoTr02a,KoTr99}.
-
-\item [Topography] ~\\
-\cite{LeKoHuTr09,LeChKoHuTr09,LeChLiKoHuTr08,GoAmTaCaSmSaMaKo09,WiKoScTr04}.
-
-\end{description}
-
-The corresponding Bib\TeX{} entries may be found
-in file \texttt{doc/USER\_MANUAL/bibliography.bib}.
-
-\section{Support}
-
-This material is based upon work supported by the USA National Science
-Foundation under Grants No. EAR-0406751 and EAR-0711177, by the French
-CNRS, French INRIA Sud-Ouest MAGIQUE-3D, French ANR NUMASIS under
-Grant No. ANR-05-CIGC-002, and European FP6 Marie Curie International
-Reintegration Grant No. MIRG-CT-2005-017461.
-Any opinions, findings, and conclusions or recommendations expressed in this material are
-those of the authors and do not necessarily reflect the views of the
-USA National Science Foundation, CNRS, INRIA, ANR or the European
-Marie Curie program.
-
-\chapter{\label{cha:Getting-Started}Getting Started}
-
-The SPECFEM3D software package comes in a gzipped tar ball. In the
-directory in which you want to install the package, type
-\begin{lyxcode}
-{\small tar~-zxvf~SPECFEM3D\_V2.1.0.tar.gz}{\small \par}
-\end{lyxcode}
-The directory \texttt{\small SPECFEM3D/} will then contain
-the source code.\\
-
-We recommend that you add {\texttt{ulimit -S -s unlimited}} to your {\texttt{.bash\_profile}} file and/or {\texttt{limit stacksize unlimited }} to your {\texttt{.cshrc}} file to suppress any potential limit to the size of the Unix stack.\\
-
-Then, to configure the software for your system, run the
-\texttt{configure} shell script. This script will attempt to guess
-the appropriate configuration values for your system. However, at
-a minimum, it is recommended that you explicitly specify the appropriate
-command names for your Fortran90 compiler and MPI package:
-
-\begin{lyxcode}
-./configure~FC=ifort~MPIFC=mpif90
-\end{lyxcode}
-
-Before running the \texttt{configure} script, you should probably edit file \texttt{flags.guess} to make sure that it contains the best compiler options for your system.
-Known issues or things to check are:
-
-\begin{description}
-\item [{\texttt{Intel ifort compiler}}] See if you need to add \texttt{-assume byterecl} for your machine.
-\item [{\texttt{IBM compiler}}] See if you need to add \texttt{-qsave} or \texttt{-qnosave} for your machine.
-\item [{\texttt{Mac OS}}] You will probably need to install \texttt{XCODE}.
-In addition, the \texttt{clock\_gettime} routine, which is used by the \texttt{SCOTCH} library that we use, does not exist in Mac OS.
-You will need to replace it with \texttt{clock\_get\_time} if you want to use \texttt{SCOTCH}.
-\end{description}
-
-When compiling on an IBM machine with the \texttt{xlf} and \texttt{xlc} compilers, we suggest running the \texttt{configure} script
-with the following options:\\
-
-\noindent\texttt{./configure FC=xlf90\_r MPIFC=mpif90 CC=xlc\_r CFLAGS="-O3 -q64" FCFLAGS="-O3 -q64" -with-scotch-dir=...}.\\
-
-On SGI systems, \texttt{flags.guess} automatically informs \texttt{configure}
-to insert `\texttt{`TRAP\_FPE=OFF}'' into the generated \texttt{Makefile}
-in order to turn underflow trapping off.\\
-
-If you run very large meshes on a relatively small number
-of processors, the memory size needed on each processor might become
-greater than 2 gigabytes, which is the upper limit for 32-bit addressing.
-In this case, on some compilers you may need to add \texttt{``-mcmodel=medium}'' or \texttt{``-mcmodel=medium -shared-intel}''
-to the configure options of CFLAGS, FCFLAGS and LDFLAGS otherwise the compiler will display an error
-message (for instance \texttt{``relocation truncated to fit: R\_X86\_64\_PC32 against .bss''} or something similar);
-on an IBM machine with the \texttt{xlf} and \texttt{xlc} compilers, using \texttt{-q64} is usually sufficient.\\
-
-The SPECFEM3D software package relies on the SCOTCH library to partition meshes created with CUBIT.
-Note that we use CUBIT to create meshes of hexahedra, but other packages can be used as well, for instance GiD from \url{http://gid.cimne.upc.es} or Gmsh from \url{http://geuz.org/gmsh} \citep{GeRe09}.
-Even mesh creation packages that generate tetrahedra, for instance TetGen from \url{http://tetgen.berlios.de}, can be used because each tetrahedron can then easily be decomposed into four hexahedra as shown in the picture of the TetGen logo at \url{http://tetgen.berlios.de/figs/Delaunay-Voronoi-3D.gif}; while this approach does not generate hexahedra of optimal quality, it can ease mesh creation in some situations and it has been shown that the spectral-element method can very accurately handle distorted mesh elements \citep{OlSe11}.
-
-The SCOTCH library \citep{PeRo96}
-provides efficent static mapping, graph and mesh partitioning routines. SCOTCH is a free software package developed by
-Fran\c {c}ois Pellegrini et al.
-from LaBRI and INRIA in Bordeaux, France, downloadable from the web page \url{https://gforge.inria.fr/projects/scotch/}.
-It is more recent than METIS, actively maintained and performs better in many cases.
-A recent version of its source code is provided in directory \texttt{src/decompose\_mesh\_SCOTCH/scotch\_5.1.10b}.
-In case no SCOTCH libraries can be found on the system, the configuration will bundle this version for compilation.
-The path to an existing SCOTCH installation can to be set explicitly with the option \texttt{-{}-with-scotch-dir}. Just as an example:
-\begin{lyxcode}
-./configure~FC=ifort~MPIFC=mpif90~-{}-with-scotch-dir=/opt/scotch
-\end{lyxcode}
-
-If you use the Intel ifort compiler to compile the code, we recommend that you use the Intel icc C compiler to compile Scotch, i.e., use:
-\begin{lyxcode}
-./configure~CC=icc~FC=ifort~MPIFC=mpif90
-\end{lyxcode}
-
-To compile a serial version of the code for small meshes that fits
-on one compute node and can therefore be run serially, run \texttt{configure}
-with the \texttt{-{}-without-mpi} option to suppress all calls to
-MPI.
-
-A summary of the most important configuration variables follows.
-
-\begin{description}
-\item [{\texttt{F90}}] Path to the Fortran90 compiler.
-\item [{\texttt{MPIF90}}] Path to MPI Fortran90.
-\item [{\texttt{MPI\_FLAGS}}] Some systems require this flag to link to
-MPI libraries.
-\item [{\texttt{FLAGS\_CHECK}}] Compiler flag for non-critical subroutines.
-\item [{\texttt{FLAGS\_NO\_CHECK}}] Compiler flag for creating fast, production-run
-code for critical subroutines.
-\end{description}
-The configuration script automatically creates for each executable a corresponding
-\texttt{Makefile} in the \texttt{src/} subdirectoy.
-The \texttt{Makefile} contains a number of suggested entries for various
-compilers, e.g., Portland, Intel, Absoft, NAG, and Lahey. The software
-has run on a wide variety of compute platforms, e.g., various PC clusters
-and machines from Sun, SGI, IBM, Compaq, and NEC. Select the compiler
-you wish to use on your system and choose the related optimization
-flags. Note that the default flags in the \texttt{Makefile} are undoubtedly
-not optimal for your system, so we encourage you to experiment with
-these flags and to solicit advice from your systems administrator.
-Selecting the right compiler and optimization flags can make a tremendous
-difference in terms of performance. We welcome feedback on your experience
-with various compilers and flags.
-
-Now that you have set the compiler information, you need to select
-a number of flags in the \texttt{constants.h} file depending on your
-system:
-
-\begin{description}
-\item [{\texttt{LOCAL\_PATH\_IS\_ALSO\_GLOBAL}}] Set to \texttt{.false.}
-on most cluster applications. For reasons of speed, the (parallel)
-distributed database generator typically writes a (parallel) database for the solver on the
-local disks of the compute nodes. Some systems have no local disks,
-e.g., BlueGene or the Earth Simulator, and other systems have a fast
-parallel file system, in which case this flag should be set to \texttt{.true.}.
-Note that this flag is not used by the database generator or the solver; it is
-only used for some of the post-processing.
-\end{description}
-The package can run either
-in single or in double precision mode. The default is single precision
-because for almost all calculations performed using the spectral-element method
-using single precision is sufficient and gives the same results (i.e. the same seismograms);
-and the single precision code is faster and requires exactly half as much memory.
-Select your preference by selecting the appropriate setting in the \texttt{constants.h} file:
-
-\begin{description}
-\item [{\texttt{CUSTOM\_REAL}}] Set to \texttt{SIZE\_REAL} for single precision
-and \texttt{SIZE\_DOUBLE} for double precision.
-\end{description}
-In the \texttt{precision.h} file:
-
-\begin{description}
-\item [{\texttt{CUSTOM\_MPI\_TYPE}}] Set to \texttt{MPI\_REAL} for single
-precision and \texttt{MPI\_DOUBLE\_PRECISION} for double precision.
-\end{description}
-On many current processors (e.g., Intel, AMD, IBM Power), single precision calculations
-are significantly faster; the difference can typically be 10\%
-to 25\%. It is therefore better to use single precision.
-What you can do once for the physical problem you want to study is run the same calculation in single precision
-and in double precision on your system and compare the seismograms.
-If they are identical (and in most cases they will), you can select single precision for your future runs.\\
-
-\section{Adding OpenMP support in addition to MPI}
-
-OpenMP support can be enabled in addition to MPI. However, in many cases performance will not improve
-because our pure MPI implementation is already heavily optimized and thus the resulting code will in fact
-be slightly slower. A possible exception could be IBM BlueGene-type architectures.
-
-To enable OpenMP, uncomment the OpenMP compiler option in two lines in file \texttt{src/specfem3D/Makefile.in}
-(before running \texttt{configure}) and also uncomment the \texttt{\#define USE\_OPENMP} statement in file\\
-\texttt{src/specfem3D/specfem3D.F90}.
-
-The DO-loop using OpenMP threads has a SCHEDULE property. The OMP\_SCHEDULE
-environment variable can set the scheduling policy of that DO-loop.
-Tests performed by Marcin Zielinski at SARA (The Netherlands) showed that often
-the best scheduling policy is DYNAMIC with the size of the chunk equal to the number of
-OpenMP threads, but most preferably being twice as the number of
-OpenMP threads (thus chunk size = 8 for 4 OpenMP threads etc).
-If OMP\_SCHEDULE is not set or is empty, the DO-loop will assume generic
-scheduling policy, which will slow down the job quite a bit.
-
-\section{Visualizing the subroutine calling tree of the source code}
-
-Packages such as \texttt{doxywizard} can be used to visualize the subroutine calling tree of the source code.
-\texttt{Doxywizard} is a GUI front-end for configuring and running \texttt{doxygen}.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\chapter{\label{cha:Mesh-Generation}Mesh Generation}
-
-The first step in running a spectral-element simulation consists of constructing a high-quality mesh for the region under consideration.
-We provide two possibilities to do so: (1) relying on the external, hexahedral mesher CUBIT, or (2) using the provided, internal mesher \texttt{xmeshfem3D}.
-In the following, we explain these two approaches.
-
-\section{\label{cha:Running-the-Mesher-CUBIT}Meshing with \texttt{CUBIT}}
-
-CUBIT is a meshing tool suite for the creation of finite-element meshes for arbitrarily shaped models.
-It has been developed and maintained at Sandia National Laboratories and can be purchased for a small academic institutional fee at \url{http://cubit.sandia.gov}.
-Our experience showed that using CUBIT greatly facilitates and speeds up the generation and preparation of hexahedral, conforming meshes for a variety of geophysical models with increasing complexity.
-
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=.6\textwidth]{figures/mount-cubit.jpg}
-\par\end{centering}
-
-\caption{Example of the graphical user interface of CUBIT. The hexahedral mesh shown in the main display consists of a hexahedral discretization of a single volume with topography.}
-
-\label{fig:mount.cubit}
-\end{figure}
-
-The basic steps in creating a load-balanced, partitioned mesh with CUBIT are:
-\begin{description}
-\item [1.] setting up a hexahedral mesh with CUBIT,
-\item [2.] exporting the CUBIT mesh into a SPECFEM3D file format and
-\item [3.] partitioning the SPECFEM3D mesh files for a chosen number of cores.
-\end{description}
-
-Examples are provided in the SPECFEM3D package in the subdirectory \texttt{examples/}.
-We strongly encourage you to contribute your own example to this package by contacting the CIG Computational Seismology Mailing List  \urlwithparentheses{cig-seismo@geodynamics.org}.
-
-\subsection{Creating the Mesh with CUBIT}
-
-For the installation and handling of the CUBIT meshing tool suite, please refer to the CUBIT user manual and documentation.
-In order to give you a basic understanding of how to use CUBIT for our purposes, examples are provided in the SPECFEM3D package in the subdirectory \texttt{examples/}:
-\begin{description}
-\item [{\texttt{homogeneous\_halfspace}}] Creates a single block model and assigns elastic material parameters.
-\item [{\texttt{layered\_halfspace}}]  Combines two different, elastic material volumes and creates a refinement layer between the two.
-This example can be compared for validation against the solutions provided in subdirectory ~\\
-\texttt{VALIDATION\_3D\_SEM\_SIMPLER\_LAYER\_SOURCE\_DEPTH/}.
-\item [{\texttt{waterlayered\_halfspace}}] Combines an acoustic and elastic material volume as in a schematic marine survey example.
-\item [{\texttt{tomographic\_model}}] Creates a single block model whose material properties will have to be read in from a tomographic model file during the databases creation by \texttt{xgenerate\_databases}.
-\end{description}
-
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=.45\textwidth]{figures/example-homogeneous.jpg}
-\includegraphics[width=.45\textwidth]{figures/example-2layers.jpg} \\
-\includegraphics[width=.45\textwidth]{figures/example-water.jpg}
-\includegraphics[width=.45\textwidth]{figures/example-tomo.jpg}
-\par\end{centering}
-
-\caption{Screenshots of the CUBIT examples provided in subdirectory  \texttt{examples/}: homogeneous halfspace (top-left), layered halfspace (top-right), water layered halfspace (bottom-left) and tomographic model (bottom-right).}
-
-\label{fig:examples.cubit}
-\end{figure}
-
-In each example subdirectory you will find a \texttt{README} file, which explains in a step-by-step tutorial the workflow for the example.
-Please feel free to contribute your own example to this package by contacting the CIG Computational Seismology Mailing List  \urlwithparentheses{cig-seismo@geodynamics.org}.
-
-\subsection{Exporting the Mesh with \texttt{cubit2specfem3d.py} }
-\label{subsec:Exporting-the-Mesh}
-
-Once the geometric model volumes in CUBIT are meshed, you prepare the model for exportation with the definition of material blocks and boundary surfaces.
-Thus, prior to exporting the mesh, you need to define blocks specifying the materials and absorbing boundaries in CUBIT.
-This process could be done automatically using the script \texttt{boundary\_definition.py} if the mesh meets some conditions or manually, following the block convention:
-
-\begin{description}
-\item [material\_name] Each material should have a specific block defined by a unique name.
-The name convention of the material is to start with either 'elastic' or 'acoustic'.
-It must be then followed by a unique identifier, e.g.
-'elastic 1', 'elastic 2', etc.
-The additional attributes to the block define the material description.
-
-For an elastic material:
-\begin{description}
-\item [material\_id] An integer value which is unique for this material.
-\item [Vp] P-wave speed of the material (given in m/s).
-\item [Vs] S-wave speed of the material (given in m/s).
-\item [rho] density of the material (given in kg/m$^3$).
-\item [Q] quality factor to use in case of a simulation with attenuation turned on.
-It should be between 1 and 9000.
-In case no attenuation information is available, it can be set to zero.
-Please note that your Vp- and Vs-speeds are given for a reference frequency.
-To change this reference frequency, you change the value of
-\texttt{ATTENUATION\_f0\_REFERENCE} in the main constants file \texttt{constants.h} found in subdirectory \texttt{src/shared/}.
-\item [anisotropic\_flag] Flag describing the anisotropic model to use in case an anisotropic simulation should be conducted.
-See the file \texttt{model\_aniso.f90} in subdirectory \texttt{src/generate\_databases/} for an implementation of the anisotropic models.
-In case no anisotropy is available, it can be set to zero.
-\end{description}
-Note that this material block has to be defined using all the volumes which belong to this elastic material.
-For volumes belonging to another, different material, you will need to define a new material block.
-
-For an acoustic material:
-\begin{description}
-\item [material\_id] An integer value which is unique for this material.
-\item [Vp] P-wave speed of the material (given in m/s).
-\item [0] S-wave speed of the material is ignored.
-\item [rho] density of the material (given in kg/m$^3$).
-\end{description}
-
-\item [face\_topo] Block definition for the surface which defines the free surface (which can have topography).
-The name of this block must be 'face\_topo', the block has to be defined using all the surfaces which constitute the complete free surface of the model.
-
-\item [face\_abs\_xmin] Block definition for the faces on the absorbing boundaries, one block for each surface with x=Xmin.
-\item [face\_abs\_xmax] Block definition for the faces on the absorbing boundaries, one block for each surface with x=Xmax.
-\item [face\_abs\_ymin] Block definition for the faces on the absorbing boundaries, one block for each surface with y=Ymin.
-\item [face\_abs\_ymax] Block definition for the faces on the absorbing boundaries, one block for each surface with y=Ymax.
-\item [face\_abs\_bottom] Block definition for the faces on the absorbing boundaries, one block for each surface with z=bottom.
-\end{description}
-
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=.45\textwidth]{figures/mount-surface.jpg}
-\includegraphics[width=.45\textwidth]{figures/mount-abs.jpg}
-\par\end{centering}
-
-\caption{Example of the block definitions for the free surface 'face\_topo' (left) and the absorbing boundaries, defined in a single block 'face\_abs\_xmin' (right) in CUBIT.}
-
-\label{fig:mount.abs}
-\end{figure}
-
-Optionally, instead of specifying for each surface at the boundaries a single block like mentioned above, you can also specify a single block for all boundary surfaces and name it as one of the absorbing blocks above, e.g.
-'face\_abs\_xmin'.
-
-After the block definitions are done, you export the mesh using the script \texttt{cubit2specfem3d.py} provided in each of the example directories.
-If the export was successful, you should find the following files in a subdirectory \texttt{MESH/}:
-
-\begin{description}
-\item [nummaterial\_velocity\_file] Defines the material properties. For fully defined materials, the formats is:
-\begin{lyxcode}
-\texttt{domain\_ID} \texttt{material\_ID}  \texttt{rho}  \texttt{vp}  \texttt{vs}  \texttt{Q}  \texttt{anisotropy\_flag}
-\end{lyxcode}
-where
-\texttt{domain\_ID} is 1 for acoustic or 2 for elastic materials, \texttt{material\_ID} a unique identifier,  \texttt{rho} the density in $kg \, m^{-3}$,  \texttt{vp} the P-wave speed in $m \, s^{-1}$, \texttt{vs}  the S-wave speed in $m \, s^{-1}$, \texttt{Q} the quality factor and \texttt{anisotropy\_flag} an identifier for anisotropic models.
-
-%%magnoni
-For tomographic velocity models, please read Chapter \ref{cha:-Changing-the} and Section \ref{sec:Using-tomographic} `Using external tomographic Earth models' for further details.
-%
-%For a tomographic model, the material definition format is:
-%\begin{lyxcode}
-%\texttt{domain\_ID} \texttt{material\_ID}  \texttt{tomography}  \texttt{name}
-%\end{lyxcode}
-%where
-%\texttt{domain\_ID} is 1 for acoustic or 2 for elastic materials, \texttt{material\_ID} a negative, unique identifier, \texttt{tomography} keyword for tomographic material definition and \texttt{name} the name of the tomography file.
-The name is not used so far, rather change the filename defined in the file \texttt{model\_tomography.f90} located in the
-%\texttt{src/generate\_databases/} directory.
-%
-\item [materials\_file] Contains the material associations for each element.
-\item [nodes\_coords\_file] Contains the point locations in Cartesian coordinates of the mesh element corners.
-\item [mesh\_file] Contains the mesh element connectivity.
-\item [free\_surface\_file] Contains the free surface connectivity.
-\item [absorbing\_surface\_file\_xmax] Contains the surface connectivity of the absorbing boundary surface at the Xmax.
-\item [absorbing\_surface\_file\_xmin] Contains the surface connectivity of the absorbing boundary surface at the Xmin.
-\item [absorbing\_surface\_file\_ymax] Contains the surface connectivity of the absorbing boundary surface at the Ymax.
-\item [absorbing\_surface\_file\_ymin] Contains the surface connectivity of the absorbing boundary surface at the Ymin.
-\item [absorbing\_surface\_file\_bottom] Contains the surface connectivity of the absorbing boundary surface at the bottom.
-\end{description}
-
-These mesh files are needed as input files for the partitioner  \texttt{xdecompose\_mesh\_SCOTCH}
-to load-balance the mesh. Please see the next section for further details.
-
-\subsubsection*{Checking the mesh quality}
-The quality of the mesh may be inspected more precisely based upon the serial code
-in the file  \texttt{check\_mesh\_quality\_}~\\
-\texttt{CUBIT\_Abaqus.f90} located in the directory \texttt{src/check\_mesh\_quality\_CUBIT\_Abaqus/}.
-Running this code is optional because no information needed by the
-solver is generated.
-
-Prior to running and compiling this code, you have to export your mesh in CUBIT to an ABAQUS (.inp) format.
-For example, export mesh block IDs belonging to volumes in order to check the quality of the hexahedral elements.
-You also have to determine a number of parameters of your mesh, such as the number of nodes and number of elements
-and modify the header of the \texttt{check\_mesh\_quality\_CUBIT\_Abaqus.f90} source file. Then,
-in the main directory, type
-\begin{lyxcode}
-{\small make xcheck\_mesh\_quality\_CUBIT\_Abaqus}{\small \par}
-\end{lyxcode}
-and use
-\begin{lyxcode}
-{\small ./bin/xcheck\_mesh\_quality\_CUBIT\_Abaqus}{\small \par}
-\end{lyxcode}
-to generate an
-%not supported: AVS output file (\texttt{\small AVS\_meshquality.inp} in AVS UCD format) or
-OpenDX output file (\texttt{\small DX\_mesh\_quality.dx})
-that can be used to investigate mesh quality, e.g. skewness of elements,
-and a Gnuplot histogram (\texttt{\small mesh\_quality\_histogram.txt})
-that can be plotted with gnuplot (type
-`\texttt{\small gnuplot plot\_mesh\_quality\_histogram.gnu}'). The
-histogram is also printed to the screen.
-Analyze that skewness histogram of mesh elements to make sure no element has a skewness above approximately 0.75, otherwise the mesh is of poor quality (and if even a single element has a skewness value above 0.80, then you must definitely improve the mesh).
-If you want to start designing
-your own meshes, this tool is useful for viewing your creations. You
-are striving for meshes with elements with `cube-like' dimensions,
-e.g., the mesh should contain no very elongated or skewed elements.
-
-\subsection{Partitioning the Mesh with \texttt{xdecompose\_mesh\_SCOTCH}}
-
-The SPECFEM3D software package performs large scale simulations in a parallel 'Single Process Multiple Data' way.
-The spectral-element mesh created with CUBIT needs to be distributed on the processors. This partitioning is executed once and for all prior to
-the execution of the solver so it is referred to as a static mapping.
-
-An efficient partitioning is important because it leverages the overall running time of the application.
-It amounts to balance the number of elements in each slice while minimizing the communication costs
-resulting from the placement of adjacent elements on different processors. \texttt{decompose\_mesh\_SCOTCH} depends on the SCOTCH library \citep{PeRo96},
-which provides efficent static mapping, graph and mesh partitioning routines. SCOTCH is a free software package developed by
-Fran\c {c}ois Pellegrini et al.
-from LaBRI and INRIA in Bordeaux, France, downloadable from the web page \url{https://gforge.inria.fr/projects/scotch/}.
-It is more recent than METIS, actively maintained and performs better in many cases.
-A recent version of its source code is provided in directory \texttt{src/decompose\_mesh\_SCOTCH/scotch\_5.1.10b}.
-\\
-
-In most cases, the configuration with \texttt{./configure~FC=ifort} should be sufficient.
-During the configuration process, the script tries to find existing SCOTCH installations.
-In case your system has no pre-existing SCOTCH installation,
-we provide the source code of SCOTCH, which is released open source under the French CeCILL-C version 1 license,
-in directory
-\texttt{src/decompose\_mesh\_SCOTCH/scotch\_5.1.10b}.
-This version gets bundled with the compilation of the SPECFEM3D package if no libraries could have been found.
-If this automatic compilation of the SCOTCH libraries fails,
-please refer to file INSTALL.txt in that directory to see further details how to compile it on your system.
-In case you want to use a pre-existing installation, make sure you have correctly specified the path
-of the SCOTCH library when using the option \texttt{-{}-with-scotch-dir} with the \texttt{./configure} script.
-In the future you should be able to find more recent versions at  \url{http://www.labri.fr/perso/pelegrin/scotch/scotch\_en.html} .
-\\
-
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=.45\textwidth]{figures/mount-partitions.jpg}
-\includegraphics[width=.45\textwidth]{figures/mount-partitions2.jpg}
-\par\end{centering}
-
-\caption{Example of a mesh partitioning onto four cores.
-Each single core partition is colored differently.
-The executable \texttt{xdecompose\_mesh\_SCOTCH} can equally distribute the mesh on any arbitrary number of cores.
-Domain decomposition is explained in detail in \cite{MaKoBlLe08}, and excellent scaling up to 150,000 processor cores in shown for instance in
-\cite{CaKoLaTiMiLeSnTr08,KoLaMi08a,MaKoBlLe08,KoErGoMi10,KoGoErMi10,Kom11}.}
-
-\label{fig:mount.partitions}
-\end{figure}
-
-When you are ready to compile,
-in the main directory type
-\begin{lyxcode}
-{\small make xdecompose\_mesh\_SCOTCH}{\small \par}
-\end{lyxcode}
-If all paths and flags have been set correctly, the executable \texttt{bin/xdecompose\_mesh\_SCOTCH} should be produced.
-
-The partitioning is done in serial for now (in the next release we will provide a parallel version of that code),
-the synopsis is:
-\begin{lyxcode}
-./bin/xdecompose\_mesh\_SCOTCH  nparts  input\_directory output\_directory
-\end{lyxcode}
-where
-\begin{itemize}
-     \item \texttt{nparts} is the number of partitions, i.e., the number of cores for the parallel simulations,
-     \item \texttt{input\_directory} is the directory which holds all the files generated by the Python script \texttt{cubit2specfem3d.py} explained in the previous Section~\ref{subsec:Exporting-the-Mesh}, e.g.
-\texttt{MESH/}, and
-     \item \texttt{output\_directory} is the directory for the output of this partitioner which stores ACII-format files named like \texttt{proc******\_Database} for each partition.
-These files will be needed for creating the distributed databases, and have to reside in the directory  \texttt{LOCAL\_PATH} specified in the main \texttt{Par\_file}, e.g.
-in directory \texttt{in\_out\_files/DATABASES\_MPI}.
-Please see Chapter~\ref{cha:Creating-Distributed-Databases} for further details.
-\end{itemize}
-Note that all the files generated by the Python script  \texttt{cubit2specfem3d.py} must be placed in the \texttt{input\_directory} folder before running the program.
-
-\section{\label{cha:Running-the-Mesher-Meshfem3D}Meshing with \texttt{xmeshfem3D}}
-
-In case you successfully ran the configuration script, you are also ready to compile the internal mesher.
-This is an alternative to CUBIT
-for the mesh generation of relatively simple geological models. The mesher is no longer
-dedicated to Southern California and more flexiblity is provided in this version of the package.
-
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[scale=0.5]{figures/socal_map_mpi}
-\par\end{centering}
-
-\caption{\label{fig:For-parallel-computing}For parallel computing purposes,
-the model block is subdivided in $\nprocxi\times\nproceta$ slices
-of elements. In this example we use $5^{2}=25$ processors. }
-
-\end{figure}
-
-In the main directory, type
-\begin{lyxcode}
-{\small make xmeshfem3D}{\small \par}
-\end{lyxcode}
-If all paths and flags
-have been set correctly, the mesher should now compile and produce
-the executable \texttt{bin/xmeshfem3D}. Please note that
-\texttt{xmeshfem3D} must be called directly from the \texttt{bin/} directory,
-as most of the binaries of the package.
-
-Input for the mesh generation program is provided through the parameter
-file \texttt{Mesh\_Par\_file}, which resides in the subdirectory \texttt{in\_data\_files/meshfem3D\_files/}.
-Before running the mesher, a number of parameters need to be set in
-the \texttt{Mesh\_Par\_file}. This requires a basic understanding of how
-the SEM is implemented, and we encourage you to read \citet{KoVi98,KoTr99}
-and \citet{KoLiTrSuStSh04}.
-
-The mesher and the solver use UTM coordinates internally, therefore
-you need to define the zone number for the UTM projection (e.g., zone
-11 for Los Angeles). Use decimal values for latitude and longitude
-(no minutes/seconds). These values are approximate; the mesher will
-round them off to define a square mesh in UTM coordinates. When running
-benchmarks on rectangular models, turn the UTM projection off by using
-the flag \texttt{\small SUPPRESS\_UTM\_PROJECTION}, in which case
-all `longitude' parameters simply refer to the $x$~axis, and all
-`latitude' parameters simply refer to the $y$~axis. To run the mesher
-for a global simulation, the following parameters need to be set in
-the \texttt{Mesh\_Par\_file}:
-
-\begin{description}
-\item [{\texttt{LATITUDE\_MIN}}] Minimum latitude in the block (negative
-for South).
-\item [{\texttt{LATITUDE\_MAX}}] Maximum latitude in the block.
-\item [{\texttt{LONGITUDE\_MIN}}] Minimum longitude in the block (negative
-for West).
-\item [{\texttt{LONGITUDE\_MAX}}] Maximum longitude in the block.
-\item [{\texttt{DEPTH\_BLOCK\_KM}}] Depth of bottom of mesh in kilometers.
-\item [{\texttt{\noun{UTM\_PROJECTION\_ZONE}}}] UTM projection zone in
-which your model resides, only valid when \texttt{SUPPRESS\_UTM\_}~\\
-\texttt{PROJECTION} is \texttt{.false.}.
-\item [{\texttt{SUPPRESS\_UTM\_PROJECTION}}] set to be \texttt{.false.}
-when your model range is specified in geographical coordinates,
-and needs to be \texttt{.true.} when your model is specified in
-Cartesian coordinates. \noun{UTM projection zone in which your simulation
-region resides.}
-\item [{\texttt{INTERFACES\_FILE }}] File which contains the description of the topography and
-of the interfaces between the different layers of the model, if any.
-The number of spectral elements in the vertical direction within each layer is also defined in this file.
-
-\item [{$\nexxi$}] The number of spectral elements along one side of the
-block. This number \textit{must} be 8~$\times$~a multiple of $\nprocxi$
-defined below. Based upon benchmarks against semi-analytical discrete
-wavenumber synthetic seismograms \citep{KoLiTrSuStSh04}, determined
-that a $\nexxi=288$ run is accurate to a shortest period of roughly
-2~s. Therefore, since accuracy is determined by the number of grid
-points per shortest wavelength, for any particular value of $\nexxi$
-the simulation will be accurate to a shortest period determined by
-\begin{equation}
-\mbox{shortest period (s)}=(288/\nexxi)\times2.\label{eq:shortest_period}\end{equation}
-The number of grid points in each orthogonal direction of the reference
-element, i.e., the number of Gauss-Lobatto-Legendre points, is determined
-by \texttt{NGLLX} in the \texttt{constants.h} file. We generally use
-$\mbox{\texttt{NGLLX\/}}=5$, for a total of $5^{3}=125$ points per
-elements. We suggest not to change this value.
-\item [{$\nexeta$}] The number of spectral elements along the other side
-of the block. This number \textit{must} be 8~$\times$~a multiple
-of $\nproceta$ defined below.
-\item [{$\nprocxi$}] The number of processors or slices along one side
-of the block (see Figure~\ref{fig:For-parallel-computing}); we must
-have $\nexxi=8\times c\times\nprocxi$, where $c\ge1$ is a positive
-integer.
-\item [{$\nproceta$}] The number of processors or slices along the other
-side of the block; we must have $\nexeta=8\times c\times\nproceta$,
-where $c\ge1$ is a positive integer.
-\item [{\texttt{USE\_REGULAR\_MESH}}] set to be \texttt{.true.} if you want a perfectly regular
-mesh or \texttt{.false.} if you want to add doubling horizontal layers to coarsen the mesh. In this case, you
-also need to provide additional information by setting up the next three parameters.
-\item [{\texttt{NDOUBLINGS}}] The number of horizontal doubling layers. Must be set to \texttt{1} or \texttt{2}
-if \texttt{USE\_REGULAR\_MESH} is set to \texttt{.true.}.
-\item [{\texttt{NZ\_DOUBLING\_1}}] The position of the first doubling layer (only interpreted
-if \texttt{USE\_REGULAR\_MESH} is set to \texttt{.true.}).
-\item [{\texttt{NZ\_DOUBLING\_2}}] The position of the second doubling layer (only interpreted
-if \texttt{USE\_REGULAR\_MESH} is set to \texttt{.true.} and if \texttt{NDOUBLINGS} is set to \texttt{2}).
-\item [{\texttt{CREATE\_ABAQUS\_FILES}}] Set this flag to \texttt{.true.} to save Abaqus FEA \urlwithparentheses{www.simulia.com}
-mesh files for subsequent viewing. Turning the flag on generates files in the \texttt{LOCAL\_PATH} directory.
-See Section~\ref{sec:Mesh-graphics} for a discussion of mesh viewing features.
-\item [{\texttt{CREATE\_DX\_FILES}}] Set this flag to \texttt{.true.} to save OpenDX \urlwithparentheses{www.opendx.org}
-mesh files for subsequent viewing.
-\item [{\texttt{LOCAL\_PATH}}] Directory in which the partitions generated
-by the mesher will be written. Generally one uses a directory on the
-local disk of the compute nodes, although on some machines these partitions
-are written on a parallel (global) file system (see also the earlier
-discussion of the \texttt{LOCAL\_PATH\_IS\_ALSO\_GLOBAL} flag in Chapter~\ref{cha:Getting-Started}).
-The mesher generates the necessary partitions in parallel, one set
-for each of the $\nprocxi\times\nproceta$ slices that constitutes
-the mesh (see Figure~\ref{fig:For-parallel-computing}). After the
-mesher finishes, you can log in to one of the compute nodes and view
-the contents of the \texttt{LOCAL\_PATH} directory to see the
-files generated by the mesher.
-These files will be needed for creating the distributed databases, and have to reside in the directory  \texttt{LOCAL\_PATH} specified in the main \texttt{Par\_file}, e.g.
-in directory \texttt{in\_out\_files/DATABASES\_MPI}.
-Please see Chapter~\ref{cha:Creating-Distributed-Databases} for further details.
-\item [{\texttt{NMATERIALS}}] The number of different materials in your model. In the following lines, each material needs to be defined as :
-\begin{lyxcode}
-\texttt{material\_ID}  \texttt{rho}  \texttt{vp}  \texttt{vs}  \texttt{Q}  \texttt{anisotropy\_flag} \texttt{domain\_ID}
-\end{lyxcode}
-where
-\begin{itemize}
-     \item \texttt{Q}           : quality factor (0=no attenuation)
-     \item \texttt{anisotropy\_flag}  : 0=no anisotropy / 1,2,.. check with implementation in \texttt{aniso\_model.f90}
-     \item \texttt{domain\_id}        : 1=acoustic / 2=elastic
-\end{itemize}
-\item [{\texttt{NREGIONS}}] The number of regions in the mesh.
-In the following lines, because the mesh is regular or 'almost regular', each region is defined as :
-\begin{lyxcode}
-  \texttt{NEX\_XI\_BEGIN}  \texttt{NEX\_XI\_END}  \texttt{NEX\_ETA\_BEGIN}  \texttt{NEX\_ETA\_END}  \texttt{NZ\_BEGIN}  \texttt{NZ\_END}  \texttt{material\_ID}
-\end{lyxcode}
-\end{description}
-
-The \texttt{INTERFACES\_FILE} parameter of \texttt{Mesh\_Par\_File} defines the file which contains the settings of the topography grid and of the interfaces grids.
-Topography is defined as a set of elevation values on a regular 2D grid.
-It is also possible to
-define interfaces between the layers of the model in the same way.
-The file needs to define several parameters:
-\begin{itemize}
-\item The number of interfaces, including the topography. This needs to be set at the first line. Then, from the bottom
-to the top of the model, you need to define the grids with:
-\item \texttt{SUPPRESS\_UTM\_PROJECTION} flag as described previously,
-\item number of points along $x$ and $y$ direction (NXI and NETA),
-\item minimal $x$ and $y$ coordinates (LONG\_MIN and LAT\_MIN),
-\item spacing between points along $x$ and $y$ (SPACING\_XI and SPACING\_ETA) and
-\item the name of the file which contains the elevation values (in $y$.$x$ increasing order).
-\end{itemize}
-At the end of this file, you simply need to set the number of spectral elements in the vertical direction for each layer. We provide a few models in the {\texttt{examples/}} directory.  \\
-
-Finally, depending on your system, you might need to provide a file that tells MPI what compute nodes
-to use for the simulations. The file must have a number of entries
-(one entry per line) at least equal to the number of processors needed
-for the run. A sample file is provided in the file \texttt{mymachines}.
-This file is not used by the mesher or solver, but is required by
-the \texttt{go\_mesher} and \texttt{go\_solver} default job submission
-scripts. See Chapter~\ref{cha:Scheduler} for information about running
-the code on a system with a scheduler, e.g., LSF.
-
-Now that you have set the appropriate parameters in the \texttt{Mesh\_Par\_file}
-and have compiled the mesher, you are ready to launch it! This is
-most easily accomplished based upon the \texttt{go\_mesher} script.
-When you run on a PC cluster, the script assumes that the nodes are
-named n001, n002, etc. If this is not the case, change the \texttt{tr
--d `n'} line in the script. You may also need to edit the last command
-at the end of the script that invokes the \texttt{mpirun} command.
-See Chapter~\ref{cha:Scheduler} for information about running the
-code on a system with a scheduler, e.g., LSF.
-
-Mesher output is provided in the \texttt{in\_out\_files/OUTPUT\_FILES} directory
-in \texttt{output\_mesher.txt}; this file provides lots of details
-about the mesh that was generated. Please note that the mesher suggests a time step \texttt{DT} to run
-the solver with. The mesher output file also contains a table about the quality of the mesh to indicate
-possible problems with the distortions of elements.
-Alternatively, output can be directed
-to the screen instead by uncommenting a line in \texttt{constants.h}:
-\begin{lyxcode}
-!~uncomment~this~to~write~messages~to~the~screen~
-
-!~integer,~parameter~::~IMAIN~=~ISTANDARD\_OUTPUT
-\end{lyxcode}
-
-To control the quality of the mesh, check the standard output (either on the screen or in the \texttt{in\_out\_files/OUTPUT\_FILES} directory in \texttt{output\_mesher.txt}) and analyze the skewness histogram of mesh elements to make sure no element has a skewness above approximately 0.75, otherwise the mesh is of poor quality (and if even a single element has a skewness value above 0.80, then you must definitely improve the mesh).
-To draw the skewness histogram on the screen, type \texttt{gnuplot plot\_mesh\_quality\_histogram.gnu}.
-
-\chapter{\label{cha:Creating-Distributed-Databases}Creating the Distributed Databases}
-
-After using \texttt{xmeshfem3D} or \texttt{xdecompose\_mesh\_SCOTCH}, the next step in the workflow is to compile
-\texttt{xgenerate\_}~\\
-\texttt{databases}. This program is going to create all the missing information needed by the SEM solver.
-
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=.6\textwidth]{figures/workflow.jpg}
-\par\end{centering}
-
-\caption{Schematic workflow for a SPECFEM3D simulation. The executable \texttt{xgenerate\_databases} creates the GLL mesh points and assigns specific model parameters.}
-
-\label{fig:workflow.databases}
-\end{figure}
-
-In the main directory, type
-\begin{lyxcode}
-{\small make xgenerate\_databases}{\small \par}
-\end{lyxcode}
-Input for the program is provided through
-the main parameter file \texttt{Par\_file}, which resides in the subdirectory \texttt{in\_data\_files/}. Please note that
-\texttt{xgenerate\_databases} must be called directly from the \texttt{bin/} directory, as most of the binaries of the package.
-
-\section{\label{cha:Main-Parameter}Main parameter file \texttt{Par\_file}}
-
-Before running \texttt{xgenerate\_databases}, a number of parameters need to be set in the main parameter \texttt{Par\_file}
-located in the subdirectory \texttt{in\_data\_files/}:
-
-\begin{description}
-\item [{\texttt{SIMULATION\_TYPE}}] is set to 1 for forward simulations,
-2 for adjoint simulations (see Section \ref{sec:Adjoint-simulation-finite})
-and 3 for kernel simulations (see Section \ref{sec:Finite-Frequency-Kernels}).
-\item [{\texttt{SAVE\_FORWARD}}] is only set to \texttt{.true.} for a forward
-simulation with the last frame of the simulation saved, as part of
-the finite-frequency kernel calculations (see Section \ref{sec:Finite-Frequency-Kernels}).
-For a regular forward simulation, leave \texttt{SIMULATION\_TYPE}
-and \texttt{SAVE\_FORWARD} at their default values.
-\item [{\texttt{\noun{UTM\_PROJECTION\_ZONE}}}] UTM projection zone in
-which your model resides, only valid when \texttt{SUPPRESS\_UTM\_}~\\
-\texttt{PROJECTION} is \texttt{.false.}.
-\item [{\texttt{SUPPRESS\_UTM\_PROJECTION}}] set to be \texttt{.false.}
-when your model range is specified in the geographical coordinates,
-and needs to be \texttt{.true.} when your model is specified in a
-cartesian coordinates. \noun{UTM projection zone in which your simulation
-region resides.}
-\item [{\texttt{NPROC}}] The number of MPI processors, each one is assigned one slice of the whole mesh.
-\item [{\texttt{NSTEP}}] The number of time steps of the simulation. This controls the
-length of the numerical simulation, i.e., twice the number of time steps
-requires twice as much CPU time. This feature is not used at the time
-of generating the distributed databases but is required for the solver, i.e., you may change this
-parameter after running  \texttt{xgenerate\_databases}.
-\item [{\texttt{DT}}] The length of each time step in seconds.
-This feature is not used at the time of generating the distributed databases but is required for the solver.
-Please see also Section~\ref{sec:Choosing-the-Time-Step} for further details.
-%
-\item [{\texttt{MODEL}}] Must be set to one of the following:
-%
-\begin{description}
-\item [{\textmd{Models defined by mesh parameters:}}]~
-\begin{description}
-\item [{\texttt{default}}] Uses model parameters as defined by meshing procedures described
-in the previous Chapter~\ref{cha:Mesh-Generation}.
-\end{description}
-\end{description}
-%
-\begin{description}
-\item [{\textmd{1D~models~with~real~structure:}}]~
-\begin{description}
-\item [{\texttt{1D\_prem}}] Isotropic version of the spherically
-symmetric Preliminary Reference Earth Model (PREM) \citep{DzAn81}.
-\item [{\texttt{1D\_socal}}] A standard isotropic 1D model for Southern California.
-\item [{\texttt{1D\_cascadia}}] Isotropic 1D profile for the Cascadia region.
-\end{description}
-\end{description}
-%
-\begin{description}
-\item [{\textmd{Fully~3D~models:}}]~
-\begin{description}
-\item [{\texttt{aniso}}] For a user-specified fully anisotropic model.
-Parameters are set up in routines located in file \texttt{model\_aniso.f90} in directory \texttt{src/generate\_databases/}.
-See Chapter~\ref{cha:-Changing-the} for a discussion on how to specify your own 3D model.
-\item [{\texttt{external}}] For a user-specified isotropic model which uses externally defined model parameters.
-Uses external model definitions set up in routines located in file \texttt{model\_external\_values.f90} in directory \texttt{src/generate\_databases/}.
-Please modify these generic template routines to use your own model definitions.
-\item [{\texttt{gll}}] For a user-specified isotropic model which uses external binary files for $v_p$, $v_s$ and $\rho$.
-Binary files are given in the same format as when outputted by the \texttt{xgenerate\_databases} executable when using option \texttt{SAVE\_MESH\_FILES}.
-These binary files define the model parameters on all GLL points which can be used for iterative inversion procedures.
-\item [{\texttt{salton\_trough}}] A 3D $V_p$ model for Southern California.
-Users must provide the corresponding data file \texttt{regrid3\_vel\_p.bin} in directory \texttt{DATA/st\_3D\_block\_harvard/}.
-\item [{\texttt{tomo}}] For a user-specified 3D isotropic model which uses a tomographic model file \texttt{tomographic\_model.xyz} in directory \texttt{in\_data\_files/}.
-See Section \ref{sec:Using-tomographic},
-for a discussion on how to specify your own 3D tomographic model.
-\end{description}
-\end{description}
-%
-%
-\item [{\texttt{OCEANS}}] Set to \texttt{.true.} if the effect of the oceans
-on seismic wave propagation should be incorporated based upon the
-approximate treatment discussed in \citet{KoTr02b}. This feature
-is inexpensive from a numerical perspective, both in terms of memory
-requirements and CPU time. This approximation is accurate at periods
-of roughly 20~s and longer. At shorter periods the effect of water
-phases/reverberations is not taken into account, even when the flag
-is on.
-\item [{\texttt{TOPOGRAPHY}}] This feature is only effective if \texttt{OCEANS} is set to \texttt{.true.}.
-Set to \texttt{.true.} if topography and
-bathymetry should be read in based upon the topography file specified in the main
-constants file \texttt{constants.h} found in subdirectory \texttt{src/shared/} to evaluate
-elevations. If not set, elevations will be read from the numerical mesh.
-\item [{\texttt{ATTENUATION}}] Set to \texttt{.true.} if attenuation should
-be incorporated. Turning this feature on increases the memory requirements
-significantly (roughly by a factor of~1.5), and is numerically fairly
-expensive. See \citet{KoTr99,KoTr02a} for a discussion on the implementation
-of attenuation based upon standard linear solids.
-Please note that the Vp- and Vs-velocities of your model are given for a reference frequency.
-To change this reference frequency, you change the value of
-\texttt{ATTENUATION\_f0\_REFERENCE} in the main constants file \texttt{constants.h} found in subdirectory \texttt{src/shared/}.
-\item [{\texttt{USE\_OLSEN\_ATTENUATION}}] Set to \texttt{.true.} if you
-want to use the attenuation model that scaled from the S-wave speed model
-using Olsen's empirical relation (see \citet{OlDaBr03}).
-\item [{\texttt{ANISOTROPY}}] Set to \texttt{.true.} if you
-want to use an anisotropy model. Please see the file \texttt{model\_aniso.f90} in subdirectory \texttt{src/generate\_databases/} for the current implementation of anisotropic models.
-\item [{\texttt{ABSORBING\_CONDITIONS}}] Set to \texttt{.true.} to turn
-on Clayton-Enquist absorbing boundary conditions (see \citet{KoTr99}).
-\item [{\texttt{ABSORB\_INSTEAD\_OF\_FREE\_SURFACE}}] Set to \texttt{.true.} to turn on 
-absorbing boundary conditions on the top surface which by default constitutes a free surface of the model.
-\item [{\texttt{MOVIE\_SURFACE}}] Set to \texttt{.false.}, unless you want
-to create a movie of seismic wave propagation on the Earth's surface.
-Turning this option on generates large output files. See Section~\ref{sec:Movies}
-for a discussion on the generation of movies. This feature is only relevant for the solver.
-\item [{\texttt{MOVIE\_VOLUME}}] Set to \texttt{.false.}, unless you want
-to create a movie of seismic wave propagation in the Earth's interior.
-Turning this option on generates huge output files. See Section~\ref{sec:Movies}
-for a discussion on the generation of movies. This feature is only relevant for the solver.
-\item [{\texttt{NTSTEP\_BETWEEN\_FRAMES}}] Determines the number of timesteps
-between movie frames. Typically you want to save a snapshot every
-100 timesteps. The smaller you make this number the more output will
-be generated! See Section~\ref{sec:Movies} for a discussion on the
-generation of movies. This feature is only relevant for the solver.
-\item [{\texttt{CREATE\_SHAKEMAP}}] Set this flag to \texttt{.true.} to
-create a ShakeMap\textregistered, i.e., a peak ground velocity map
-of the maximum absolute value of the two horizontal components of the velocity vector.
-\item [{\texttt{SAVE\_DISPLACEMENT}}] Set this flag to \texttt{.true.}
-if you want to save the displacement instead of velocity for the movie
-frames.
-\item [{\texttt{USE\_HIGHRES\_FOR\_MOVIES}}] Set this flag to \texttt{.true.}
-if you want to save the values at all the NGLL grid points for the
-movie frames.
-\item [{\texttt{HDUR\_MOVIE}}] determines the half duration of the source time function for the movie simulations.
-When this parameter is set to be 0, a default half duration that corresponds to the accuracy of the simulation is provided.
-Otherwise, it adds this half duration to the half duration specified in the source file \texttt{CMTSOLUTION},
-thus simulates longer periods to make the movie images look smoother.
-\item [{\texttt{SAVE\_MESH\_FILES}}] Set this flag to \texttt{.true.} to
-save ParaView \urlwithparentheses{www.paraview.org}
-mesh files for subsequent viewing. Turning the flag on generates large
-(distributed) files in the \texttt{LOCAL\_PATH} directory. See Section~\ref{sec:Mesh-graphics}
-for a discussion of mesh viewing features.
-\item [{\texttt{LOCAL\_PATH}}] Directory in which the distributed databases
-will be written. Generally one uses a directory on the
-local disk of the compute nodes, although on some machines these databases
-are written on a parallel (global) file system (see also the earlier
-discussion of the \texttt{LOCAL\_PATH\_IS\_ALSO\_GLOBAL} flag in Chapter~\ref{cha:Getting-Started}).
-\texttt{xgenerate\_databases} generates the necessary databases in parallel, one set
-for each of the \texttt{NPROC} slices that constitutes
-the mesh (see Figure~\ref{fig:mount.partitions} and Figure~\ref{fig:For-parallel-computing}). After
-the executable finishes, you can log in to one of the compute nodes and view
-the contents of the \texttt{LOCAL\_PATH} directory to see the (many)
-files generated by \texttt{xgenerate\_databases}.
-Please note that the \texttt{LOCAL\_PATH} directory should already contain the output files of
-the partitioner, i.e. from \texttt{xdecompose\_mesh\_SCOTCH} or \texttt{xmeshfem3D}.
-\item [{\texttt{NTSTEP\_BETWEEN\_OUTPUT\_INFO}}] This parameter specifies
-the interval at which basic information about a run is written to
-the file system (\texttt{timestamp{*}} files in the \texttt{in\_out\_files/OUTPUT\_FILES}
-directory). If you have access to a fast machine, set \texttt{NTSTEP\_BETWEEN\_OUTPUT\_INFO}
-to a relatively high value (e.g., at least 100, or even 1000 or more)
-to avoid writing output text files too often. This feature is not
-used at the time of meshing. One can set this parameter to a larger
-value than the number of time steps to avoid writing output during
-the run.
-\item [{\texttt{NTSTEP\_BETWEEN\_OUTPUT\_SEISMOS}}] This parameter specifies
-the interval at which synthetic seismograms are written in the \texttt{LOCAL\_PATH}
-directory. If a run crashes, you may still find usable (but shorter
-than requested) seismograms in this directory. On a fast machine set
-\texttt{NTSTEP\_BETWEEN\_OUTPUT\_SEISMOS} to a relatively high value
-to avoid writing to the seismograms too often. This feature is only relevant for the solver.
-\item [{\texttt{USE\_FORCE\_POINT\_SOURCE}}] Turn this flag on to use a \texttt{FORCESOLUTION} force source
-instead of a \texttt{CMTSOLUTION} moment-tensor source. This can be useful e.g. for oil industry 
-foothills simulations in which the source is a vertical force, normal force, 
-tilted force, or an impact etc. You will need to give the East, North and vertical components of an 
-arbitrary (non-unitary) direction vector of the force vector; thus refer to Appendix~\ref{cha:Coordinates} 
-for the orientation of the reference frame.
-This vector is made unitary internally in the code and thus only its direction matters here;
-its norm is ignored and the norm of the force used is the factor force source times the source time function.
-\item [{\texttt{PRINT\_SOURCE\_TIME\_FUNCTION}}] Turn this flag on to print
-information about the source time function in the file \texttt{in\_out\_files/OUTPUT\_FILES/plot\_source\_time\_function.txt}.
-This feature is only relevant for the solver.
-\end{description}
-
-\section{\label{sec:Choosing-the-Time-Step}Choosing the time step \texttt{DT}}
-
-The parameter \texttt{DT} sets the length of each time step in seconds. The value of this parameter is crucial for the stability of
-the spectral-element simulation.
-Your time step \texttt{DT} will depend on the minimum ratio between the distance $h$ of neighboring mesh points and the wave speeds $v$ defined in your model.
-The condition for the time step $\Delta t$ is:
-
-\begin{lyxcode}
-$\Delta t < C ~\mathrm{min}_\Omega (~ h / v ~)$
-\end{lyxcode}
-where $C$ is the so-called Courant number and $\Omega$ denotes the model volume.
-The distance $h$ depends on the mesh element size and the number of
-GLL points \texttt{NGLL} specified in the main constants file \texttt{constants.h} located in the \texttt{src/shared/} subdirectory.
-The wave speed $v$ is determined based on your model's P- (or S-) wave speed values.
-
-The database generator \texttt{xgenerate\_databases}, as well as the internal mesher \texttt{xmeshfem3D}, are trying
-to evaluate the value of $\Delta t$ for empirically chosen Courant numbers $C \sim 0.3$.
-If you used the mesher \texttt{xmeshfem3D} to generate your mesh, you should set the value suggested in \texttt{in\_out\_files/OUTPUT\_FILES/output\_mesher.txt} file, which is created after the mesher completed.
-In case you used CUBIT to create the mesh, you might use an arbitrary value when running \texttt{xgenerate\_databases} and then use the value suggested in the ~\\
-\texttt{in\_out\_files/OUTPUT\_FILES/output\_mesher.txt} file after the database generation completed.
-Note that the implemented Newmark time scheme uses this time step globally, thus your simulations become more expensive for very small mesh elements in high wave-speed regions.
-Please be aware of this restriction when constructing your mesh in Chapter~\ref{cha:Mesh-Generation}.
-
-\chapter{\label{cha:Running-the-Solver}Running the Solver \texttt{xspecfem3D}}
-
-Now that you have successfully generated the databases, you are ready to compile
-the solver. In the main directory, type
-\begin{lyxcode}
-{\small make xspecfem3D}{\small \par}
-\end{lyxcode}
-Please note that \texttt{xspecfem3D} must be called directly from
-the \texttt{bin/} directory, as most of the binaries of the package.
-
-The solver needs three input files in the \texttt{in\_data\_files/} directory
-to run:
-\begin{description}
-\item [{\texttt{Par\_file}}] the main parameter file which was discussed in detail in the previous
-Chapter~\ref{cha:Creating-Distributed-Databases},
-\item [{\texttt{CMTSOLUTION}} or {\texttt{FORCESOLUTION}}] the earthquake source parameter file or the force source parameter file or, and
-\item [{\texttt{STATIONS}}] the stations file.
-\end{description}
-
-Most parameters in the \texttt{Par\_file}
-should be set prior to running the databases generation. Only the following parameters
-may be changed after running \texttt{xgenerate\_databases}:
-
-\begin{itemize}
-\item the simulation type control parameters: \texttt{SIMULATION\_TYPE}
-and \texttt{SAVE\_FORWARD}
-\item the time step parameters \texttt{NSTEP} and \texttt{DT}
-\item the absorbing boundary control parameter \texttt{ABSORBING\_CONDITIONS} on condition that the\\ 
-  \texttt{ABSORB\_INSTEAD\_OF\_FREE\_SURFACE} flag remains unmodified after running the databases generation. 
-\item the movie control parameters \texttt{MOVIE\_SURFACE}, \texttt{MOVIE\_VOLUME},
-and \texttt{NTSTEPS\_BETWEEN\_FRAMES}
-\item the ShakeMap\textregistered option \texttt{CREATE\_SHAKEMAP}
-\item the output information parameters \texttt{NTSTEP\_BETWEEN\_OUTPUT\_INFO}
-and \texttt{NTSTEP\_BETWEEN\_OUTPUT\_}~\\
-\texttt{SEISMOS}
-\item the \texttt{PRINT\_SOURCE\_TIME\_FUNCTION} flags
-\end{itemize}
-Any other change to the \texttt{Par\_file} implies rerunning both
-the database generator \texttt{xgenerate\_databases} and the solver \texttt{xspecfem3D}.
-
-For any particular earthquake, the \texttt{CMTSOLUTION} file that
-represents the point source may be obtained directly from the Harvard Centroid-Moment Tensor (CMT) web page \urlwithparentheses{www.seismology.harvard.edu}.
-It looks like this:
-
-\begin{lyxcode}
-{\small }%
-\begin{figure}[H]
-\noindent \begin{centering}
-{\small \includegraphics[width=1\textwidth]{figures/Hollywood_CMT} }
-\par\end{centering}{\small \par}
-
-\caption{\label{fig:CMTSOLUTION-file}\texttt{CMTSOLUTION} file based on the
-format from the Harvard CMT catalog. \textbf{M} is the moment tensor,
-$M_{0}${\small{} }is the seismic moment, and $M_{w}$ is the moment
-magnitude.}
-
-\end{figure}
-{\small \par}
-\end{lyxcode}
-The \texttt{CMTSOLUTION} should be edited in the following way:
-
-\begin{itemize}
-\item Set the \texttt{time shift} parameter equal to $0.0$ (the solver
-will not run otherwise.) The time shift parameter would simply apply
-an overall time shift to the synthetics, something that can be done
-in the post-processing (see Section \ref{sec:Process-data-and-syn}).
-\item For point-source simulations (see finite sources, page \pageref{To-simulate-a})
-we recommend setting the source half-duration parameter \texttt{half
-duration} equal to zero, which corresponds to simulating a step source-time
-function, i.e., a moment-rate function that is a delta function. If
-\texttt{half duration} is not set to zero, the code will use a Gaussian
-(i.e., a signal with a shape similar to a `smoothed triangle', as
-explained in \citet{KoTr02a} and shown in Fig~\ref{fig:gauss.vs.triangle})
-source-time function with half-width \texttt{half duration}. We prefer
-to run the solver with \texttt{half duration} set to zero and convolve
-the resulting synthetic seismograms in post-processing after the run,
-because this way it is easy to use a variety of source-time functions
-(see Section \ref{sec:Process-data-and-syn}). \citet{KoTr02a} determined
-that the noise generated in the simulation by using a step source
-time function may be safely filtered out afterward based upon a convolution
-with the desired source time function and/or low-pass filtering. Use
-the serial code \texttt{convolve\_source\_timefunction.f90} and the
-script \texttt{convolve\_source\_timefunction.csh} for this purpose,
-or alternatively use signal-processing software packages such as SAC \urlwithparentheses{www.llnl.gov/sac}.
-Type
-\begin{lyxcode}
-{\small make xconvolve\_source\_timefunction}{\small \par}
-\end{lyxcode}
-to compile the code and then set the parameter \texttt{hdur} in \texttt{convolve\_source\_timefunction.csh}
-to the desired half-duration.
-
-\item The zero time of the simulation corresponds to the center of the triangle/Gaussian,
-or the centroid time of the earthquake. The start time of the simulation
-is $t=-1.5*\texttt{half duration}$ (the 1.5 is to make sure the moment
-rate function is very close to zero when starting the simulation).
-To convert to absolute time $t_{\mathrm{abs}}$, set
-
-\begin{lyxcode}
-$t_{\mathrm{abs}}=t_{\mathrm{pde}}+\texttt{time shift}+t_{\mathrm{synthetic}}$
-\end{lyxcode}
-where $t_{\mathrm{pde}}$ is the time given in the first line of the
-\texttt{CMTSOLUTION}, \texttt{time shift} is the corresponding value
-from the original \texttt{CMTSOLUTION} file and $t_{\mathrm{synthetic}}$
-is the time in the first column of the output seismogram.
-
-\end{itemize}
-%
-\begin{figure}
-\noindent \begin{centering}
-%%%\includegraphics[width=3in]{figures/gauss_vs_triangle_mod.eps}
-\includegraphics[width=3in]{figures/gauss_vs_triangle_mod.pdf}
-\par\end{centering}
-
-\caption{Comparison of the shape of a triangle and the Gaussian function actually
-used.}
-
-\label{fig:gauss.vs.triangle}
-\end{figure}
-
-Centroid latitude and longitude should be provided in geographical
-coordinates. The code converts these coordinates to geocentric coordinates~\citep{DaTr98}.
-Of course you may provide your own source representations by designing
-your own \texttt{CMTSOLUTION} file. Just make sure that the resulting
-file adheres to the Harvard CMT conventions (see Appendix~\ref{cha:Coordinates}).
-Note that the first line in the \texttt{CMTSOLUTION} file is the Preliminary Determination of Earthquakes (PDE) solution performed by the USGS NEIC, which is used as a seed for the Harvard CMT inversion.
-The PDE solution is based upon P waves and often gives the hypocenter of the earthquake, i.e., the rupture initiation point, whereas the CMT solution gives the `centroid location', which is the location with dominant moment release.
-The PDE solution is not used by our software package but must be present anyway in the first line of the file.
-
-\label{To-simulate-a}To simulate a kinematic rupture, i.e., a finite-source
-event, represented in terms of $N_{\mathrm{sources}}$ point sources,
-provide a \texttt{CMTSOLUTION} file that has $N_{\mathrm{sources}}$
-entries, one for each subevent (i.e., concatenate $N_{\mathrm{sources}}$
-\texttt{CMTSOLUTION} files to a single \texttt{CMTSOLUTION} file).
-At least one entry (not necessarily the first) must have a zero \texttt{time
-shift}, and all the other entries must have non-negative \texttt{time
-shift}. Each subevent can have its own half duration, latitude, longitude,
-depth, and moment tensor (effectively, the local moment-density tensor).
-
-Note that the zero in the synthetics does NOT represent the hypocentral
-time or centroid time in general, but the timing of the \textit{center}
-of the source triangle with zero \texttt{time shift} (Fig~\ref{fig:source_timing}).
-
-Although it is convenient to think of each source as a triangle, in
-the simulation they are actually Gaussians (as they have better frequency
-characteristics). The relationship between the triangle and the gaussian
-used is shown in Fig~\ref{fig:gauss.vs.triangle}. For finite fault
-simulations it is usually not advisable to use a zero half duration
-and convolve afterwards, since the half duration is generally fixed
-by the finite fault model.
-
-\vspace{1cm}
-
-\noindent The \texttt{FORCESOLUTION} should be edited in the following way:
-
-\begin{itemize}
-\item Set the \texttt{time shift} parameter equal to $0.0$ (the solver
-will not run otherwise.) The time shift parameter would simply apply
-an overall time shift to the synthetics, something that can be done
-in the post-processing (see Section \ref{sec:Process-data-and-syn}).
-\item Set the latitude, longitude, depth of the source in geographical
-  coordinates.
-\item Set the magnitude of the force source.
-\item Set the components of a (non-unitary) direction vector for the
-  force source in the East/North/Vertical basis (see Appendix A for
-  the orientation of the reference frame).
-\end{itemize}
-
-{\small }%
-\begin{figure}[H]
-\noindent \begin{centering}
-{\small \includegraphics[width=5in]{figures/source_timing.pdf} }
-\par\end{centering}{\small \par}
-
-\caption{Example of timing for three sources. The center of the first source
-triangle is defined to be time zero. Note that this is NOT in general
-the hypocentral time, or the start time of the source (marked as tstart).
-The parameter \texttt{time shift} in the \texttt{CMTSOLUTION} file
-would be t1(=0), t2, t3 in this case, and the parameter \texttt{half
-duration} would be hdur1, hdur2, hdur3 for the sources 1, 2, 3 respectively.}
-
-{\small \label{fig:source_timing} }
-\end{figure}
-{\small \par}
-
-The solver can calculate seismograms at any number of stations for
-basically the same numerical cost, so the user is encouraged to include
-as many stations as conceivably useful in the \texttt{STATIONS} file,
-which looks like this:
-
-{\small }%
-\begin{figure}[H]
-\noindent \begin{centering}
-{\small \includegraphics{figures/STATIONS_basin_explained} }
-\par\end{centering}{\small \par}
-
-\caption{Sample \texttt{STATIONS} file. Station latitude and longitude should
-be provided in geographical coordinates. The width of the station
-label should be no more than 32 characters (see \texttt{MAX\_LENGTH\_STATION\_NAME}
-in the \texttt{constants.h} file), and the network label should be
-no more than 8 characters (see \texttt{MAX\_LENGTH\_NETWORK\_NAME}
-in the \texttt{constants.h} file).}
-
-\end{figure}
-{\small \par}
-
-Each line represents one station in the following format:
-
-\begin{lyxcode}
-{\small Station~Network~Latitude~(degrees)~Longitude~(degrees)~Elevation~(m)~burial~(m)~}{\small \par}
-\end{lyxcode}
-The solver \texttt{xspecfem3D} filters the list of stations in file \texttt{in\_data\_files/STATIONS}
-to exclude stations that are not located within the region given in
-the \texttt{Par\_file} (between \texttt{LATITUDE\_MIN} and \texttt{LATITUDE\_MAX}
-and between \texttt{LONGITUDE\_MIN} and \texttt{LONGITUDE\_MAX}).
-The filtered file is called \texttt{in\_data\_files/STATIONS\_FILTERED}.
-
-Solver output is provided in the \texttt{in\_out\_files/OUTPUT\_FILES} directory
-in the \texttt{output\_solver.txt} file. Output can be directed to
-the screen instead by uncommenting a line in \texttt{constants.h}:
-
-\begin{lyxcode}
-!~uncomment~this~to~write~messages~to~the~screen~
-
-!~integer,~parameter~::~IMAIN~=~ISTANDARD\_OUTPUT~
-\end{lyxcode}
-On PC clusters the seismogram files are generally written to the local
-disks (the path \texttt{LOCAL\_PATH} in the \texttt{Par\_file}) and
-need to be gathered at the end of the simulation.
-
-While the solver is running, its progress may be tracked by monitoring
-the `\texttt{\small timestamp{*}}' files in the ~\\
-\texttt{\small in\_out\_files/OUTPUT\_FILES} directory.
-These tiny files look something like this:
-
-\begin{lyxcode}
-Time~step~\#~~~~~~~~~~10000~
-
-Time:~~~~~108.4890~~~~~~seconds~
-
-Elapsed~time~in~seconds~=~~~~1153.28696703911~
-
-Elapsed~time~in~hh:mm:ss~=~~~~~0~h~19~m~13~s~
-
-Mean~elapsed~time~per~time~step~in~seconds~=~~~~~0.115328696703911~
-
-Max~norm~displacement~vector~U~in~all~slices~(m)~=~~~~1.0789589E-02~
-\end{lyxcode}
-The \texttt{\small timestamp{*}} files provide the \texttt{Mean elapsed
-time per time step in seconds}, which may be used to assess performance
-on various machines (assuming you are the only user on a node), as
-well as the \texttt{Max norm displacement vector U in all slices~(m)}.
-If something is wrong with the model, the mesh, or the source, you
-will see the code become unstable through exponentially growing values
-of the displacement and fluid potential with time, and ultimately
-the run will be terminated by the program. You can control the rate
-at which the timestamp files are written based upon the parameter
-\texttt{NTSTEP\_BETWEEN\_OUTPUT\_INFO} in the \texttt{Par\_file}.
-
-Having set the \texttt{Par\_file} parameters, and having provided the
-\texttt{CMTSOLUTION} (or the \texttt{FORCESOLUTION}) and
-\texttt{STATIONS} files, you are now ready to launch the solver! This
-is most easily accomplished based upon the \texttt{go\_solver} script
-(See Chapter \ref{cha:Scheduler} for information about running through
-a scheduler, e.g., LSF). You may need to edit the last command at the
-end of the script that invokes the \texttt{mpirun} command. The
-\texttt{runall} script compiles and runs both
-\texttt{xgenerate\_databases} and \texttt{xspecfem3D} in
-sequence. This is a safe approach that ensures using the correct
-combination of distributed database output and solver input.
-
-It is important to realize that the CPU and memory requirements of
-the solver are closely tied to choices about attenuation (\texttt{ATTENUATION})
-and the nature of the model (i.e., isotropic models are cheaper than
-anisotropic models). We encourage you to run a variety of simulations
-with various flags turned on or off to develop a sense for what is
-involved.
-
-For the same model, one can rerun the solver for different events by
-simply changing the \texttt{CMTSOLUTION} or \texttt{FORCESOLUTION}
-file, or for different stations by changing the \texttt{STATIONS}
-file. There is no need to rerun the \texttt{xgenerate\_databases}
-executable. Of course it is best to include as many stations as
-possible, since this does not add to the cost of the simulation.
-
-\chapter{\label{cha:Adjoint-Simulations}Adjoint Simulations}
-
-Adjoint simulations are generally performed for two distinct applications.
-First, they can be used for earthquake source inversions, especially
-earthquakes with large ruptures such as the Lander's earthquake \citep{WaHe94}.
-Second, they can be used to generate finite-frequency sensitivity
-kernels that are a critical part of tomographic inversions based upon
-3D reference models \citep{TrTaLi05,LiTr06,TrKoLi08,LiTr08}. In either
-case, source parameter or velocity structure updates are sought to
-minimize a specific misfit function (e.g., waveform or traveltime
-differences), and the adjoint simulation provides a means of computing
-the gradient of the misfit function and further reducing it in successive
-iterations. Applications and procedures pertaining to source studies
-and finite-frequency kernels are discussed in Sections~\ref{sec:Adjoint-simulation-sources}
-and \ref{sec:Adjoint-simulation-finite}, respectively. The two related
-parameters in the \texttt{Par\_file} are \texttt{SIMULATION\_TYPE}
-(1 or 2) and the \texttt{SAVE\_FORWARD} (boolean).
-
-\section{\label{sec:Adjoint-simulation-sources}Adjoint Simulations for Sources}
-
-In the case where a specific misfit function is minimized to invert
-for the earthquake source parameters, the gradient of the misfit function
-with respect to these source parameters can be computed by placing
-time-reversed seismograms at the receivers and using them as sources
-in an adjoint simulation, and then the value of the gradient is obtained
-from the adjoint seismograms recorded at the original earthquake location.
-
-\begin{enumerate}
-\item \textbf{Prepare the adjoint sources} \label{enu:Prepare-the-adjoint}
-
-\begin{enumerate}
-\item First, run a regular forward simlation (\texttt{SIMULATION\_TYPE =
-1} and \texttt{SAVE\_FORWARD = .false.}). You can automatically set
-these two variables using the \texttt{\small utils/change\_simulation\_type.pl}
-script:
-
-\begin{lyxcode}
-utils/change\_simulation\_type.pl~-f~
-\end{lyxcode}
-and then collect the recorded seismograms at all the stations given
-in \texttt{in\_data\_files/STATIONS}.
-
-\item Then select the stations for which you want to compute the time-reversed
-adjoint sources and run the adjoint simulation, and compile them into
-the \texttt{in\_data\_files/STATIONS\_ADJOINT} file, which has the same format
-as the regular \texttt{in\_data\_files/STATIONS} file.
-
-\begin{itemize}
-\item Depending on what type of misfit function is used for the source inversion,
-adjoint sources need to be computed from the original recorded seismograms
-for the selected stations and saved in the \texttt{in\_out\_files/SEM/} directory
-with the format \texttt{STA.NT.BX?.adj}, where \texttt{STA}, \texttt{NT}
-are the station name and network code given in the \texttt{in\_data\_files/STATIONS\_ADJOINT}
-file, and \texttt{BX?} represents the component name of a particular
-adjoint seismogram. Please note that the band code can change depending on
-your sampling rate (see Appendix~\ref{cha:channel-codes} for further details).
-\item The adjoint seismograms are in the same format as the original seismogram
-(\texttt{STA.NT.BX?.sem?}), with the same start time, time interval
-and record length.
-\end{itemize}
-\item Notice that even if you choose to time reverse only one component
-from one specific station, you still need to supply all three components
-because the code is expecting them (you can set the other two components
-to be zero).
-\item Also note that since time-reversal is done in the code itself, no
-explicit time-reversing is needed for the preparation of the adjoint
-sources, i.e., the adjoint sources are in the same forward time sense
-as the original recorded seismograms.
-\end{enumerate}
-\item \textbf{Set the related parameters and run the adjoint simulation}\\
-In the \texttt{in\_data\_files/Par\_file}, set the two related parameters to
-be \texttt{SIMULATION\_TYPE = 2} and \texttt{SAVE\_FORWARD = .false.}.
-More conveniently, use the scripts \texttt{utils/change\_simulation\_type.pl}
-to modify the \texttt{Par\_file} automatically (\texttt{change\_simulation\_type.pl
--a}). Then run the solver to launch the adjoint simulation.
-\item \textbf{Collect the seismograms at the original source location}
-
-After the adjoint simulation has completed successfully, collect the
-seismograms from \texttt{LOCAL\_PATH}.
-
-\begin{itemize}
-\item These adjoint seismograms are recorded at the locations of the original
-earthquake sources given by the \texttt{in\_data\_files/CMTSOLUTION} file, and
-have names of the form \texttt{S?????.NT.S??.sem} for the six-component
-strain tensor (\texttt{SNN,SEE,SZZ,SNE,SNZ,SEZ}) at these locations,
-and ~\\
-\texttt{S?????.NT.BX?.sem} for the three-component displacements
-(\texttt{BXN,BXE,BXZ}) recorded at these locations.
-\item \texttt{S?????} denotes the source number; for example, if the original
-\texttt{CMTSOLUTION} provides only a point source, then the seismograms
-collected will start with \texttt{S00001}.
-\item These adjoint seismograms provide critical information for the computation
-of the gradient of the misfit function.
-\end{itemize}
-\end{enumerate}
-
-\section{\label{sec:Adjoint-simulation-finite}Adjoint Simulations for Finite-Frequency
-Kernels (Kernel Simulation)}
-
-Finite-frequency sensitivity kernels are computed in two successive
-simulations (please refer to \citet{LiTr06} and \citet{TrKoLi08} for details).
-
-\begin{enumerate}
-\item \textbf{Run a forward simulation with the state variables saved at
-the end of the simulation}
-
-Prepare the \texttt{\small CMTSOLUTION} and \texttt{\small STATIONS}
-files, set the parameters \texttt{\small SIMULATION\_TYPE}{\small{}
-}\texttt{\small =}{\small{} }\texttt{\small 1} and \texttt{\small SAVE\_FORWARD
-=}{\small{} }\texttt{\small .true.} in the \texttt{Par\_file} (\texttt{change\_simulation\_type
--F}), and run the solver.
-
-\begin{itemize}
-\item Notice that attenuation is not implemented yet for the computation
-of finite-frequency kernels; therefore set \texttt{ATTENUATION = .false.}
-in the \texttt{Par\_file}.
-\item We also suggest you modify the half duration of the \texttt{CMTSOLUTION}
-to be similar to the accuracy of the simulation (see Equation \ref{eq:shortest_period})
-to avoid too much high-frequency noise in the forward wavefield, although
-theoretically the high-frequency noise should be eliminated when convolved
-with an adjoint wavefield with the proper frequency content.
-\item This forward simulation differs from the regular simulations (\texttt{\small SIMULATION\_TYPE}{\small{}
-}\texttt{\small =}{\small{} }\texttt{\small 1} and \texttt{\small SAVE\_FORWARD}{\small{}
-}\texttt{\small =}{\small{} }\texttt{\small .false.}) described in
-the previous chapters in that the state variables for the last time
-step of the simulation, including wavefields of the displacement,
-velocity, acceleration, etc., are saved to the \texttt{LOCAL\_PATH}
-to be used for the subsequent simulation.
-\item For regional simulations, the files recording the absorbing boundary
-contribution are also written to the \texttt{LOCAL\_PATH} when \texttt{SAVE\_FORWARD
-= .true.}.
-\end{itemize}
-\item \textbf{Prepare the adjoint sources}
-
-The adjoint sources need to be prepared the same way as described
-in the Section \ref{enu:Prepare-the-adjoint}.
-
-\begin{itemize}
-\item In the case of travel-time finite-frequency kernel for one source-receiver
-pair, i.e., point source from the \texttt{CMTSOLUTION}, and one station
-in the \texttt{STATIONS\_ADJOINT} list, we supply a sample program
-in \texttt{utils/adjoint\_sources/traveltime/xcreate\_adjsrc\_traveltime} to cut a certain portion of the original
-displacement seismograms and convert it into the proper adjoint source
-to compute the finite-frequency kernel.
-\begin{lyxcode}
-xcreate\_adjsrc\_traveltime~t1~t2~ifile{[}0-5]~E/N/Z-ascii-files~{[}baz]
-\end{lyxcode}
-where \texttt{t1} and \texttt{t2} are the start and end time of the
-portion you are interested in, \texttt{ifile} denotes the component
-of the seismograms to be used (0 for all three components, 1 for East,
-2 for North, and 3 for vertical, 4 for transverse, and 5 for radial
-component), \texttt{E/N/Z-ascii-files} indicate the three-component
-displacement seismograms in the right order, and \texttt{baz} is the
-back-azimuth of the station. Note that \texttt{baz} is only supplied
-when \texttt{ifile} = 4 or 5.
-
-\item Similarly, a sample program to compute adjoint sources for amplitude finite-frequency kernels
-may be found in \texttt{utils/adjoint\_sources/amplitude} and used in the same way as described for
-traveltime measurements
-\begin{lyxcode}
-xcreate\_adjsrc\_amplitude~t1~t2~ifile{[}0-5]~E/N/Z-ascii-files~{[}baz].
-\end{lyxcode}
-
-\end{itemize}
-
-\item \textbf{Run the kernel simulation}
-
-With the successful forward simulation and the adjoint source ready
-in the \texttt{in\_out\_files/SEM/} directory, set \texttt{SIMULATION\_TYPE = 3} and \texttt{SAVE\_FORWARD
-= .false.} in the \texttt{Par\_file} (you can use \texttt{change\_simulation\_type.pl -b}),
-and rerun the solver.
-
-\begin{itemize}
-\item The adjoint simulation is launched together with the back reconstruction
-of the original forward wavefield from the state variables saved from
-the previous forward simulation, and the finite-frequency kernels
-are computed by the interaction of the reconstructed forward wavefield
-and the adjoint wavefield.
-\item The back-reconstructed seismograms at the original station locations
-are saved to the \texttt{LOCAL\_PATH} at the end of the kernel simulations,
-and can be collected to the local disk.
-\item These back-constructed seismograms can be compared with the time-reversed
-original seismograms to assess the accuracy of the backward reconstruction,
-and they should match very well.
-\item The arrays for density, P-wave speed and S-wave speed kernels are
-also saved in the \texttt{LOCAL\_PATH} with the names \texttt{proc??????\_rho(alpha,beta)\_kernel.bin},
-where \texttt{proc??????} represents the processor number, \texttt{rho(alpha,beta)}
-are the different types of kernels.
-\end{itemize}
-\end{enumerate}
-In general, the three steps need to be run sequentially to assure
-proper access to the necessary files. If the simulations are run through
-some cluster scheduling system (e.g., LSF), and the forward simulation
-and the subsequent kernel simulations cannot be assigned to the same
-set of computer nodes, the kernel simulation will not be able to access
-the database files saved by the forward simulation. Solutions for
-this dilemma are provided in Chapter~\ref{cha:Scheduler}. Visualization
-of the finite-frequency kernels is discussed in Section~\ref{sec:Finite-Frequency-Kernels}.
-
-\chapter{Noise Cross-correlation Simulations}
-
-Besides earthquake simulations, SPECFEM3D includes functionality for seismic noise tomography as well.
-In order to proceed successfully in this chapter, it is critical that you have already familiarized yourself
-with procedures for meshing (Chapter \ref{cha:Mesh-Generation}),
-creating distributed databases (Chapter \ref{cha:Creating-Distributed-Databases}),
-running earthquake simulations (Chapters \ref{cha:Running-the-Solver})
-and adjoint simulations (Chapter \ref{cha:Adjoint-Simulations}).
-Also, make sure you read the article `Noise cross-correlation sensitivity kernels' \citep{trompetal2010},
-in order to understand noise simulations from a theoretical perspective.
-
-\section{Input Parameter Files}
-
-As usual, the three main input files are crucial: \texttt{Par\_file}, \texttt{CMTSOLUTION} and \texttt{STATIONS}.
-Unless otherwise specified, those input files should be located in directory \texttt{in\_data\_files/}.\\
-
-\texttt{CMTSOLUTION} is required for all simulations.
-At a first glance, it may seem unexpected to have it here,
-since the noise simulations should have nothing to do with the earthquake -- \texttt{CMTSOLUTION}.
-However, for noise simulations, it is critical to have no earthquakes.
-In other words, the moment tensor specified in \texttt{CMTSOLUTION} must be set to zero manually!\\
-
-\texttt{STATIONS} remains the same as in previous earthquake simulations,
-except that the order of receivers listed in \texttt{STATIONS} is now important.
-The order will be used to determine the `master' receiver,
-i.e., the one that simultaneously cross correlates with the others.\\
-
-\texttt{Par\_file} also requires careful attention.
-A parameter called \texttt{NOISE\_TOMOGRAPHY} has been added which specifies the type of simulation to be run.
-\texttt{NOISE\_TOMOGRAPHY} is an integer with possible values 0, 1, 2 and 3.
-For example, when \texttt{NOISE\_TOMOGRAPHY} equals 0, a regular earthquake simulation will be run.
-When it is 1/2/3, you are about to run
-step 1/2/3 of the noise simulations respectively.
-Should you be confused by the three steps, refer to \citet{trompetal2010} for details.\\
-
-Another change to \texttt{Par\_file} involves the parameter \texttt{NSTEP}.
-While for regular earthquake simulations this parameter specifies the length of synthetic seismograms generated,
-for noise simulations it specifies the length of the seismograms used to compute cross correlations.
-The actual cross correlations are thus twice this length, i.e., $2~\mathrm{NSTEP}-1$.
-The code automatically makes the modification accordingly, if \texttt{NOISE\_TOMOGRAPHY} is not zero.\\
-
-There are other parameters in \texttt{Par\_file} which should be given specific values.
-For instance, since the first two steps for calculating noise sensitivity kernels correspond to
-forward simulations, \texttt{SIMULATION\_TYPE} must be 1 when \texttt{NOISE\_TOMOGRAPHY} equals 1 or 2.
-Also, we have to reconstruct the ensemble forward wavefields in adjoint simulations,
-therefore we need to set \texttt{SAVE\_FORWARD} to \texttt{.true.} for the second step,
-i.e., when \texttt{NOISE\_TOMOGRAPHY} equals 2.
-The third step is for kernel constructions. Hence \texttt{SIMULATION\_TYPE} should be 3, whereas
-\texttt{SAVE\_FORWARD} must be \texttt{.false.}.\\
-
-Finally, for most system architectures,
-please make sure that \texttt{LOCAL\_PATH} in \texttt{Par\_file} is in fact local, not globally shared.
-Because we have to save the wavefields at the earth's surface at every time step,
-it is quite problematic to have a globally shared \texttt{LOCAL\_PATH},
-in terms of both disk storage and I/O speed.
-
-\section{Noise Simulations: Step by Step}
-
-Proper parameters in those parameter files are not enough for noise simulations to run.
-We have more parameters to specify:
-for example, the ensemble-averaged noise spectrum, the noise distribution etc.
-However, since there are a few `new' files, it is better to introduce them sequentially.
-In this section, standard procedures for noise simulations are described.
-
-\subsection{Pre-simulation}
-
-\begin{itemize}
-
-\item
-As usual, we first configure the software package using:
-
-\texttt{./configure FC=ifort MPIFC=mpif90}
-
-Use the following if SCOTCH is needed:
-
-\texttt{./configure FC=ifort MPIFC=mpif90 --with-scotch-dir=/opt/scotch}\\
-
-\item
-Next, we need to compile the source code using:
-
-\texttt{make xgenerate\_databases}
-
-\texttt{make xspecfem3D} \\
-
-\item
-Before we can run noise simulations, we have to specify the noise statistics,
-e.g., the ensemble-averaged noise spectrum.
-Matlab scripts are provided to help you to generate the necessary file.
-
-\texttt{examples/noise\_tomography/NOISE\_TOMOGRAPHY.m}  (main program)\\
-\texttt{examples/noise\_tomography/PetersonNoiseModel.m}
-
-In Matlab, simply run:
-
-\texttt{NOISE\_TOMOGRAPHY(NSTEP, DT, Tmin, Tmax, NOISE\_MODEL)}\\
-
-\texttt{DT} is given in \texttt{Par\_file},
-but \texttt{NSTEP} is NOT the one specified in \texttt{Par\_file}.
-Instead, you have to feed $2~\mathrm{NSTEP}-1$ to account for the doubled length of cross correlations.
-\texttt{Tmin} and \texttt{Tmax} correspond to the period range you are interested in,
-whereas \texttt{NOISE\_MODEL} denotes the noise model you will be using.
-Details can be found in the Matlab script.
-
-After running the Matlab script, you will be given the following information (plus a figure in Matlab):
-
-\texttt{*************************************************************}\\
-\texttt{the source time function has been saved in:}\\
-\texttt{/data2/yangl/3D\_NOISE/S\_squared} (note this path must be different)\\
-\texttt{S\_squared should be put into directory:}\\
-\texttt{in\_out\_files/NOISE\_TOMOGRAPHY/ in the SPECFEM3D package}\\
-
-In other words, the Matlab script creates a file called \texttt{S\_squared},
-which is the first `new' input file we encounter for noise simulations.
-
-One may choose a flat noise spectrum rather than Peterson's noise model.
-This can be done easily by modifying the Matlab script a little.
-
-\item
-Create a new directory in the SPECFEM3D package, name it as \texttt{in\_out\_files/NOISE\_TOMOGRAPHY/}.
-We will add some parameter files later in this folder.
-
-\item
-Put the Matlab-generated-file \texttt{S\_squared} in \texttt{in\_out\_files/NOISE\_TOMOGRAPHY/}.
-
-That's to say, you will have a file \texttt{in\_out\_files/NOISE\_TOMOGRAPHY/S\_squared} in the SPECFEM3D package.
-
-\item
-Create a file called \texttt{in\_out\_files/NOISE\_TOMOGRAPHY/irec\_master\_noise}. Note that this file is located
-in directory \texttt{in\_out\_files/NOISE\_TOMOGRAPHY/} as well. In general, all noise simulation related parameter
-files go into that directory.
-\texttt{irec\_master\_noise} contains only one integer, which is the ID
-of the `master' receiver. For example, if this file contains 5, it means that the fifth receiver
-listed in \texttt{in\_data\_files/STATIONS} becomes the `master'. That's why we mentioned previously that
-the order of receivers in \texttt{in\_data\_files/STATIONS} is important.
-
-Note that in regional simulations, the \texttt{in\_data\_files/STATIONS}
-might contain receivers which are outside of our computational domains.
-Therefore, the integer in \texttt{irec\_master\_noise}
-is actually the ID in \texttt{in\_data\_files/STATIONS\_FILTERED}
-(which is generated by \texttt{bin/xgenerate\_databases}).
-
-\item
-Create a file called \texttt{in\_out\_files/NOISE\_TOMOGRAPHY/nu\_master}. This file holds three numbers,
-forming a (unit) vector. It describes which component we are cross-correlating at the `master'
-receiver, i.e., ${\hat{\bf \nu}}^{\alpha}$ in \citet{trompetal2010}.
-The three numbers correspond to E/N/Z components respectively.
-Most often, the vertical component is used, and in those cases the three numbers should be 0, 0 and 1.
-
-\item
-Describe the noise direction and distributions in \texttt{src/shared/noise\_tomography.f90}.
-Search for a subroutine called \texttt{noise\_distribution\_direction}
-in \texttt{noise\_tomography.f90}. It is actually located at the very beginning
-of \texttt{noise\_tomography.f90}. The default assumes vertical noise and a uniform distribution
-across the whole free surface of the model. It should be quite self-explanatory for modifications.
-Should you modify this part, you have to re-compile the source code.
-
-\end{itemize}
-
-\subsection{Simulations}
-
-With all of the above done, we can finally launch our simulations.
-Again, please make sure that the \texttt{LOCAL\_PATH} in \texttt{in\_data\_files/Par\_file} is not globally shared.
-It is quite problematic to have a globally shared \texttt{LOCAL\_PATH}, in terms of both disk storage and
-speed of I/O (we have to save the wavefields at the earth's surface at every time step).
-
-As discussed in \citet{trompetal2010}, it takes three steps/simulations to obtain one contribution
-of the ensemble sensitivity kernels:
-
-\begin{itemize}
-\item
-Step 1: simulation for generating wavefields\\
-\texttt{SIMULATION\_TYPE=1}\\
-\texttt{NOISE\_TOMOGRAPHY=1}\\
-\texttt{SAVE\_FORWARD} not used, can be either .true. or .false.
-
-\item
-Step 2: simulation for ensemble forward wavefields\\
-\texttt{SIMULATION\_TYPE=1}\\
-\texttt{NOISE\_TOMOGRAPHY=2}\\
-\texttt{SAVE\_FORWARD=.true.}
-
-\item
-Step 3: simulation for ensemble adjoint wavefields and sensitivity kernels\\
-\texttt{SIMULATION\_TYPE=3}\\
-\texttt{NOISE\_TOMOGRAPHY=3}\\
-\texttt{SAVE\_FORWARD=.false.}
-
-Note Step 3 is an adjoint simulation, please refer to previous chapters on how to prepare adjoint sources
-and other necessary files, as well as how adjoint simulations work.
-
-\end{itemize}
-
-Note that it is better to run the three steps continuously within the same job,
-otherwise you have to collect the saved surface movies from the old nodes/CPUs to the new nodes/CPUs.
-This process varies from cluster to cluster and thus cannot be discussed here.
-Please ask your cluster administrator for information/configuration of the cluster you are using.\\
-
-\subsection{Post-simulation}
-
-After those simulations, you have all stuff you need, either in the \texttt{in\_out\_files/OUTPUT\_FILES/} or
-in the directory specified by \texttt{LOCAL\_PATH} in \texttt{in\_data\_files/Par\_file}
-(which are most probably on local nodes).
-Collect whatever you want from the local nodes to your workstation, and then visualize them.
-This process is the same as what you may have done for regular earthquake simulations.
-Refer to other chapters if you have problems.\\
-
-Simply speaking, two outputs are the most interesting: the simulated ensemble cross correlations and
-one contribution of the ensemble sensitivity kernels.\\
-
-The simulated ensemble cross correlations are obtained after the second simulation (Step 2).
-Seismograms in \texttt{in\_out\_files/OUTPUT\_FILES/} are actually the simulated ensemble cross correlations.
-Collect them immediately after Step 2, or the Step 3 will overwrite them.
-Note that we have a `master' receiver specified by \texttt{irec\_master\_noise},
-the seismogram at one station corresponds to the cross correlation between that station and the `master'.
-Since the seismograms have three components, we may obtain cross correlations
-for different components as well, not necessarily the cross correlations between vertical components.\\
-
-One contribution of the ensemble cross-correlation sensitivity kernels are obtained after Step 3,
-stored in the \texttt{in\_data\_files/LOCAL\_PATH} on local nodes.
-The ensemble kernel files are named the same as regular earthquake kernels.\\
-
-You need to run another three simulations to get the other contribution of the ensemble kernels,
-using different forward and adjoint sources given in \citet{trompetal2010}.
-
-\section{Example}
-
-In order to illustrate noise simulations in an easy way, one example is provided in \texttt{examples/noise\_tomography/}.
-See \texttt{examples/noise\_tomography/README} for explanations. \\
-
-Note, however, that they are created for a specific workstation (CLOVER@PRINCETON),
-which has at least 4 cores with `mpif90' working properly. \\
-
-If your workstation is suitable, you can run the example in \texttt{examples/noise\_tomography/} using:\\
-
-\texttt{./pre-processing.sh}\\
-
-Even if this script does not work on your workstation, the procedure it describes is universal.
-You may review the whole process described in the last section by following the commands in \texttt{pre-processing.sh},
-which should contain enough explanations for all the commands.
-
-\chapter{Graphics}
-
-\section{\label{sec:Mesh-graphics}Meshes}
-
-In case you used the internal mesher \texttt{xmeshfem3D} to create and partition your mesh,
-you can output mesh files in ABAQUS (.INP) and DX (.dx) format to visualize them. For this, you must set either the
-flag \texttt{CREATE\_DX\_FILES} or \texttt{CREATE\_ABAQUS\_FILES} to \texttt{.true.}
-in the mesher's parameter file \texttt{Mesh\_Par\_file}
-prior to running the mesher (see Chapter~\ref{cha:Running-the-Mesher-Meshfem3D} for details).
-You can then use AVS \urlwithparentheses{www.avs.com} or OpenDX \urlwithparentheses{www.opendx.org}
-to visualize the mesh and MPI partition (slices).
-
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=.49\textwidth]{figures/vtk_mesh_vp.jpg}
-\includegraphics[width=.49\textwidth]{figures/vtk_mesh_vs.jpg}
-\par\end{centering}
-
-\caption{Visualization using Paraview of VTK files created by \texttt{xgenerate\_databases} showing P- and S-wave velocities assigned to the mesh points.
-The mesh was created by \texttt{xmeshfem3D} for 4 processors.}
-
-\label{fig:vtk.mesh}
-\end{figure}
-
-You have also the option to visualize the distributed databases produced by \texttt{xgenerate\_databases}
-using Paraview \urlwithparentheses{www.paraview.org}. For this, you must set the flag \texttt{SAVE\_MESH\_FILES} to \texttt{.true.}
-in the main parameter file \texttt{Par\_file} (see Chapter~\ref{cha:Main-Parameter} for details).
-This will create VTK files for each single partition. You can then use Paraview \urlwithparentheses{www.paraview.org} to visualized these partitions.
-
-\section{\label{sec:Movies}Movies}
-
-To make a surface or volume movie of the simulation, set parameters
-\texttt{MOVIE\_SURFACE}, \texttt{MOVIE\_VOLUME}, and \texttt{NTSTEP\_BETWEEN\_FRAMES}
-in the \texttt{Par\_file}. Turning on the movie flags, in particular
-\texttt{MOVIE\_VOLUME}, produces large output files. \texttt{MOVIE\_VOLUME}
-files are saved in the \texttt{LOCAL\_PATH} directory, whereas \texttt{MOVIE\_SURFACE}
-output files are saved in the \texttt{in\_out\_files/OUTPUT\_FILES} directory. We
-save the displacement field if the parameter \texttt{SAVE\_DISPLACEMENT} is set, otherwise the velocity field is saved.
-The look of a movie is determined by the
-half-duration of the source. The half-duration should be large enough
-so that the movie does not contain frequencies that are not resolved
-by the mesh, i.e., it should not contain numerical noise. This can
-be accomplished by selecting a CMT \texttt{HALF\_DURATION} > 1.1 $\times$
-smallest period (see figure \ref{fig:CMTSOLUTION-file}). When \texttt{MOVIE\_SURFACE}
-= .\texttt{true.}, the half duration of each source in the \texttt{CMTSOLUTION}
-file is replaced by
-
-\begin{quote}
-\[
-\sqrt{(}\mathrm{\mathtt{HALF\_DURATIO}\mathtt{N}^{2}}+\mathrm{\mathtt{HDUR\_MOVI}\mathtt{E}^{2}})\]
-\textbf{NOTE:} If \texttt{HDUR\_MOVIE} is set to 0.0, the code will
-select the appropriate value of 1.1 $\times$ smallest period. As
-usual, for a point source one can set \texttt{half duration} in the
-\texttt{CMTSOLUTION} file to be 0.0 and \texttt{HDUR\_MOVIE} = 0.0 in the \texttt{Par\_file} to get
-the highest frequencies resolved by the simulation, but for a finite
-source one would keep all the \texttt{half durations} as prescribed
-by the finite source model and set \texttt{HDUR\_MOVIE} = 0.0.
-\end{quote}
-
-\subsection{Movie Surface}
-
-When running \texttt{xspecfem3D} with the \texttt{MOVIE\_SURFACE}
-flag turned on, the code outputs \texttt{moviedata??????} files in
-the \texttt{in\_out\_files/OUTPUT\_FILES} directory. There are several flags
-in the main parameter file \texttt{Par\_file} that control the output of these moviedata files
-(see section \ref{cha:Main-Parameter} for details):
-\texttt{\small NTSTEP\_BETWEEN\_FRAMES} to set the timesteps between frames,
-\texttt{\small SAVE\_DISPLACEMENT} to save displacement instead of velocity,
-\texttt{USE\_HIGHRES\_FOR\_MOVIES} to save values at all GLL point instead of element edges.
-In order to output additionally shakemaps, you would set
-the parameter \texttt{CREATE\_SHAKEMAP} to \texttt{.true.}.
-
-The files are in a fairly complicated
-binary format, but there is a program provided to convert the
-output into more user friendly formats:
-\begin{description}
-\item [\texttt{xcreate\_movie\_shakemap\_AVS\_DX\_GMT}]
-From \texttt{create\_movie\_shakemap\_AVS\_DX\_GMT.f90}, it
-outputs data in ASCII, OpenDX, or AVS format (also readable in ParaView).
-Before compiling the code, make sure you have the file \texttt{surface\_from\_mesher.h}
-in the  \texttt{in\_out\_files/OUTPUT\_FILES/} directory. This file will be created by the solver run.
-Then type
-\begin{lyxcode}
-{\footnotesize make xcreate\_movie\_shakemap\_AVS\_DX\_GMT}{\footnotesize \par}
-\end{lyxcode}
-and run the executable \texttt{xcreate\_movie\_shakemap\_AVS\_DX\_GMT}
-in the \texttt{bin/} subdirectory. It will create visualization files in your format of choice. The code
-will prompt the user for input parameters.
-
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=.32\textwidth]{figures/movie_surf_1.jpg}
-\includegraphics[width=.32\textwidth]{figures/movie_surf_2.jpg}
-\includegraphics[width=.32\textwidth]{figures/movie_surf_3.jpg}
-\par\end{centering}
-
-\caption{Visualization using AVS files created by \texttt{xcreate\_movie\_shakemap\_AVS\_DX\_GMT} showing
-movie snapshots of vertical velocity components at different times.}
-
-\label{fig:movie.surf}
-\end{figure}
-
-\end{description}
-
-The \texttt{SPECFEM3D} code is running in near real-time to produce
-animations of southern California earthquakes via the web; see Southern California ShakeMovie\textregistered \urlwithparentheses{www.shakemovie.caltech.edu}.
-
-\subsection{\label{sub:Movie-Volume}Movie Volume}
-
-When running xspecfem3D with the \texttt{\small MOVIE\_VOLUME} flag
-turned on, the code outputs several files in \texttt{\small LOCAL\_PATH} specified in the main
-\texttt{Par\_file}, e.g. in directory \texttt{in\_out\_files/DATABASES\_MPI}.
-%As the files can be very large, there are several flags in the \texttt{\small Par\_file}
-%that control the region in space and time that is saved. These are:
-%\texttt{\small NTSTEP\_BETWEEN\_FRAMES}
-The output is saved by each processor
-at the time interval specified by \texttt{\small NTSTEP\_BETWEEN\_FRAMES}.
-For all domains, the velocity field is output to files:
-\texttt{\small proc??????\_velocity\_X\_it??????.bin},
-\texttt{\small proc??????\_velocity\_Y\_it??????.bin} and
-\texttt{\small proc??????\_velocity\_Z\_it??????.bin}.
-For elastic domains, the divergence and curl taken from the velocity field, i.e. $\nabla \cdot \bf{v}$
-and $\nabla \times \bf{v}$, get stored as well:
-\texttt{\small proc??????\_div\_it??????.bin},
-\texttt{\small proc??????\_curl\_X\_t??????.bin},
-\texttt{\small proc??????\_curl\_Y\_it??????.bin} and
-\texttt{\small proc??????\_curl\_Z\_it??????.bin}.
-The files denoted
-\texttt{\small proc??????\_div\_}~\\
-\texttt{\small glob\_it??????.bin} and
-\texttt{\small proc??????\_curl\_glob\_it??????.bin}
-are stored on the global points only, all the other arrays
-are stored on all GLL points. Note that the components X/Y/Z can change to E/N/Z according to
-the \texttt{SUPPRESS\_UTM\_PROJECTION} flag (see also Appendix~\ref{cha:Coordinates} and \ref{cha:channel-codes}).
-
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=.32\textwidth]{figures/movie_volume_1.jpg}
-\includegraphics[width=.32\textwidth]{figures/movie_volume_2.jpg}
-\includegraphics[width=.32\textwidth]{figures/movie_volume_3.jpg}
-\par\end{centering}
-
-\caption{Paraview visualization using movie volume files (converted by \texttt{xcombine\_vol\_data} and \texttt{mesh2vtu.pl}) and showing
-snapshots of vertical velocity components at different times.}
-
-\label{fig:movie.volume}
-\end{figure}
-
-To visualize these files, we use an auxilliary program \texttt{combine\_vol\_data.f90}
-to combine the data from all slices into one mesh file. To compile it in the root directory, type:
-\begin{lyxcode}
-{\footnotesize make~xcombine\_vol\_data~}{\footnotesize \par}
-\end{lyxcode}
-which will create the executable \texttt{xcombine\_vol\_data} in the directory \texttt{bin/}. To output the
-usage of this executable, type `./bin/xcombine\_vol\_data` without arguments.
-As an example, to run the executable you would use
-\begin{lyxcode}
-{\footnotesize cd~bin/~}{\footnotesize \par}
-{\footnotesize ./xcombine\_vol\_data~0~3~velocity\_Z\_it000400~../in\_out\_files/DATABASES\_MPI~../in\_out\_files/OUTPUT\_FILES~0~}{\footnotesize \par}
-\end{lyxcode}
-to create a low-resolution mesh file. The output mesh file will have the name \texttt{velocity\_Z\_it000400.mesh}.
-We next convert the \texttt{.mesh} file into the VTU (Unstructured
-grid file) format which can be viewed in ParaView. For this task, you can use and modify the script \texttt{mesh2vtu.pl}
-located in directory \texttt{utils/Visualization/Paraview/}, for example:
-\begin{lyxcode}
-{\footnotesize mesh2vtu.pl~-i~velocity\_Z\_it000400.mesh~-o~velocity\_Z\_it000400.vtu}{\footnotesize \par}
-\end{lyxcode}
-Notice that this Perl script uses a program \texttt{mesh2vtu} in the
-\texttt{utils/Visualization/Paraview/mesh2vtu} directory, which further uses the VTK \urlwithparentheses{http://www.vtk.org/}
-run-time library for its execution. Therefore, make sure you have
-them properly set in the script according to your system.
-
-\section{\label{sec:Finite-Frequency-Kernels}Finite-Frequency Kernels}
-
-The finite-frequency kernels computed as explained in Section \ref{sec:Adjoint-simulation-finite}
-are saved in the \texttt{LOCAL\_PATH} at the end of the simulation.
-Therefore, we first need to collect these files on the front end,
-combine them into one mesh file, and visualize them with some auxilliary
-programs.
-
-\begin{enumerate}
-\item \textbf{Create slice files}
-
-We will only discuss the case of one source-receiver pair, i.e., the
-so-called banana-doughnut kernels. Although it is possible to collect
-the kernel files from all slices on the front end, it usually takes
-up too much storage space (at least tens of gigabytes). Since the
-sensitivity kernels are the strongest along the source-receiver great
-circle path, it is sufficient to collect only the slices that are
-along or close to the great circle path.
-
-A Perl script \texttt{slice\_number.pl} located in directory \texttt{utils/Visualization/Paraview/} can
-help to figure out the slice numbers that lie along the great circle
-path. It applies to meshes created with the internal mesher \texttt{xmeshfem3D}.
-
-\begin{enumerate}
-\item On machines where you have access to the script, copy the \texttt{Mesh\_Par\_file}, and
-\texttt{output\_solver} files, and run:
-
-\begin{lyxcode}
-{\small slice\_number.pl~Mesh\_Par\_file~output\_solver.txt~slice\_file}{\small \par}
-\end{lyxcode}
-which will generate a \texttt{slices\_file}.
-
-\item For cases with multiple sources and multiple receivers, you need to
-provide a slice file before proceeding to the next step.
-\end{enumerate}
-\item \textbf{Collect the kernel files}
-
-After obtaining the slice files, you can collect the corresponding
-kernel files from the given slices.
-
-\begin{enumerate}
-\item You can use or modify the script \texttt{utils/copy\_basin\_database.pl}
-to accomplish this:
-
-\begin{lyxcode}
-{\small utils/copy\_database.pl~slice\_file~lsf\_machine\_file~filename~{[}jobid]}{\small \par}
-\end{lyxcode}
-where \texttt{\small lsf\_machine\_file} is the machine file generated
-by the LSF scheduler, \texttt{\small filename} is the kernel name
-(e.g., \texttt{\small rho\_kernel}, \texttt{\small alpha\_kernel}
-and \texttt{\small beta\_kernel}), and the optional \texttt{\small jobid}
-is the name of the subdirectory under \texttt{\small LOCAL\_PATH}
-where all the kernel files are stored.
-
-\item After executing this script, all the necessary mesh topology files
-as well as the kernel array files are collected to the local directory
-of the front end.
-\end{enumerate}
-\item \textbf{Combine kernel files into one mesh file}
-
-We use an auxilliary program \texttt{combine\_vol\_data.f90}
-to combine the kernel files from all slices into one mesh file.
-
-\begin{enumerate}
-\item Compile it in the root directory:
-
-\begin{lyxcode}
-{\footnotesize make~xcombine\_vol\_data~}{\footnotesize \par}
-
-{\footnotesize ./bin/xcombine\_vol\_data~slice\_list~filename~input\_dir~output\_dir~high/low-resolution~}{\footnotesize \par}
-\end{lyxcode}
-where \texttt{input\_dir} is the directory where all the individual
-kernel files are stored, and \texttt{output\_dir} is where the mesh
-file will be written.
-
-\item Use 1 for a high-resolution mesh, outputting all the GLL points to
-the mesh file, or use 0 for low resolution, outputting only the corner
-points of the elements to the mesh file.
-\item The output mesh file will have the name \texttt{filename\_rho(alpha,beta).mesh}
-\end{enumerate}
-\item \textbf{Convert mesh files into .vtu files}
-
-\begin{enumerate}
-\item We next convert the \texttt{.mesh} file into the VTU (Unstructured
-grid file) format which can be viewed in ParaView. For this task, you can use and modify the script \texttt{mesh2vtu.pl}
-located in directory \texttt{utils/Visualization/Paraview/}, for example:
-
-\begin{lyxcode}
-{\footnotesize mesh2vtu.pl~-i~file.mesh~-o~file.vtu~}{\footnotesize \par}
-\end{lyxcode}
-\item Notice that this Perl script uses a program \texttt{mesh2vtu} in the
-\texttt{utils/Visualization/Paraview/mesh2vtu} directory, which further uses the VTK \urlwithparentheses{http://www.vtk.org/}
-run-time library for its execution. Therefore, make sure you have
-them properly set in the script according to your system.
-\end{enumerate}
-\item \textbf{Copy over the source and receiver .vtk file}
-
-In the case of a single source and a single receiver, the simulation
-also generates the file \texttt{sr.vtk} located in the
-\texttt{in\_out\_files/OUTPUT\_FILES/} directory to describe
-the source and receiver locations, which can also be viewed in Paraview
-in the next step.
-
-\item \textbf{View the mesh in ParaView}
-
-Finally, we can view the mesh in ParaView \urlwithparentheses{www.paraview.org}.
-
-\begin{enumerate}
-\item Open ParaView.
-\item From the top menu, \textsf{File} $\rightarrow$\textsf{Open data},
-select \texttt{file.vtu}, and click the \textsf{Accept} button.
-
-\begin{itemize}
-\item If the mesh file is of moderate size, it shows up on the screen; otherwise,
-only the bounding box is shown.
-\end{itemize}
-\item Click \textsf{Display Tab} $\rightarrow$ \textsf{Display Style} $\rightarrow$
-\textsf{Representation} and select \textsf{wireframe of surface} to
-display it.
-\item To create a cross-section of the volumetric mesh, choose \textsf{Filter}
-$\rightarrow$ \textsf{cut}, and under \textsf{Parameters Tab}, choose
-\textsf{Cut Function} $\rightarrow$ \textsf{plane}.
-\item Fill in center and normal information given by the \texttt{global\_slice\_number.pl}
-script (either from the standard output or from \texttt{normal\_plane.txt}
-file).
-\item To change the color scale, go to \textsf{Display Tab} $\rightarrow$
-\textsf{Color} $\rightarrow$ \textsf{Edit Color Map} and reselect
-lower and upper limits, or change the color scheme.
-\item Now load in the source and receiver location file by \textsf{File}
-$\rightarrow$ \textsf{Open data}, select \texttt{sr.vt}k, and click
-the \textsf{Accept} button. Choose \textsf{Filter} $\rightarrow$
-\textsf{Glyph}, and represent the points by `\textsf{spheres}'.
-\item For more information about ParaView, see the ParaView Users Guide \urlwithparentheses{www.paraview.org/files/v1.6/ParaViewUsersGuide.PDF}.
-\end{enumerate}
-\end{enumerate}
-%
-\begin{figure}[H]
-\noindent \begin{centering}
-\includegraphics[scale=0.6]{figures/3D-S-Kernel}
-\par\end{centering}
-
-\caption{(a) Top Panel: Vertical source-receiver cross-section of the S-wave
-finite-frequency sensitivity kernel $K_{\beta}$ for station GSC at
-an epicentral distance of 176 km from the September 3, 2002, Yorba
-Linda earthquake. Lower Panel: Vertical source-receiver cross-section
-of the 3D S-wave speed model used for the spectral-element simulations
-\citep{KoLiTrSuStSh04}. (b) The same as (a) but for station HEC at
-an epicentral distance of 165 km \citep{LiTr06}.}
-
-\label{figure:P-wave-speed-finite-frequency}
-\end{figure}
-
-\chapter{\label{cha:Scheduler}Running through a Scheduler}
-
-The code is usually run on large parallel machines, often PC clusters,
-most of which use schedulers, i.e., queuing or batch management systems
-to manage the running of jobs from a large number of users. The following
-considerations need to be taken into account when running on a system
-that uses a scheduler:
-
-\begin{itemize}
-\item The processors/nodes to be used for each run are assigned dynamically
-by the scheduler, based on availability. Therefore, in order for the
-\texttt{xgenerate\_databases} and the \texttt{xspecfem3D} executables (or between successive runs of the solver) to
-have access to the same database files (if they are stored on hard
-drives local to the nodes on which the code is run), they must be
-launched in sequence as a single job.
-\item On some systems, the nodes to which running jobs are assigned are
-not configured for compilation. It may therefore be necessary to pre-compile
-both the \texttt{xgenerate\_databases} and the \texttt{xspecfem3D} executables.
-\item One feature of schedulers/queuing systems is that they allow submission
-of multiple jobs in a ``launch and forget'' mode. In order to take
-advantage of this property, care needs to be taken that output and
-intermediate files from separate jobs do not overwrite each other,
-or otherwise interfere with other running jobs.
-\end{itemize}
-Examples of job scripts can be found in the \texttt{\small utils/Cluster/}
-directory and can straightforwardly be modified and adapted to meet
-more specific running needs.
-
-We describe here in some detail a job submission procedure for the
-Caltech 1024-node cluster, CITerra, under the LSF scheduling system.
-We consider the submission of a regular forward simulation using the internal mesher
-to create mesh partitions. The two
-main scripts are \texttt{\small run\_lsf.bash}, which compiles the
-Fortran code and submits the job to the scheduler, and \texttt{\small go\_mesher\_solver\_lsf\_basin.forward},
-which contains the instructions that make up
-the job itself. These scripts can be found in \texttt{\small utils/Cluster/lsf/} directory
-
-\section{Job submission \texttt{run\_lsf.bash}}
-
-This script first sets the job queue to be `normal'. It then compiles
-the mesher, database generator and solver together, figures out the number of processors
-required for this simulation from the \texttt{in\_data\_files/Par\_file}, and
-submits the LSF job.
-
-\begin{lyxcode}
-\#!/bin/bash
-
-\#~use~the~normal~queue~unless~otherwise~directed~queue=\char`\"{}-q~normal\char`\"{}~
-
-if~{[}~\$\#~-eq~1~];~then
-
-~~~~~~~~echo~\char`\"{}Setting~the~queue~to~\$1\char`\"{}
-
-~~~~~~~~queue=\char`\"{}-q~\$1\char`\"{}~
-
-fi~\\
-~\\
-\#~compile~the~mesher~and~the~solver~
-
-d=`date'~echo~\char`\"{}Starting~compilation~\$d\char`\"{}~
-
-make~clean~
-
-make~xmeshfem3D~
-
-make~xgenerate\_databases~
-
-make~xspecfem3D~
-
-d=`date'~
-
-echo~\char`\"{}Finished~compilation~\$d\char`\"{}~\\
-~\\
-\#~get~total~number~of~nodes~needed~for~solver~
-
-NPROC=`grep~NPROC~in\_data\_files/Par\_file~|~cut~-c~34-~'~\\
-~\\
-
-\#~compute~total~number~of~nodes~needed~for~mesher~
-
-NPROC\_XI=`grep~NPROC\_XI~in\_data\_files/meshfem3D\_files/Mesh\_Par\_file~|~cut~-c~34-~'~
-~\\
-
-NPROC\_ETA=`grep~NPROC\_ETA~in\_data\_files/meshfem3D\_files/Mesh\_Par\_file~|~cut~-c~34-~'~
-~\\
-\#~total~number~of~nodes~is~the~product~of~the~values~read~
-
-numnodes=\$((~\$NPROC\_XI~*~\$NPROC\_ETA~))~\\
-~\\
-
-\#~checks~total~number~of~nodes~
-
-if~{[}~\$numnodes~-neq~\$NPROC~];~then
-
-~~~~~~~~echo~\char`\"{}error number of procs mismatch\char`\"{}
-
-~~~~~~~~exit~
-
-fi~\\
-
-~\\
-echo~\char`\"{}Submitting~job\char`\"{}~
-
-bsub~\$queue~-n~\$numnodes~-W~60~-K~<go\_mesher\_solver\_lsf.forward~
-\end{lyxcode}
-
-\section{Job script \texttt{go\_mesher\_solver\_lsf.forward}}
-
-This script describes the job itself, including setup steps that can
-only be done once the scheduler has assigned a job-ID and a set of
-compute nodes to the job, the \texttt{run\_lsf.bash} commands used
-to run the mesher, database generator and the solver, and calls to scripts that collect
-the output seismograms from the compute nodes and perform clean-up
-operations.
-
-\begin{enumerate}
-\item First the script directs the scheduler to save its own output and
-output from \texttt{stdout} into ~\\
-\texttt{\small in\_out\_files/OUTPUT\_FILES/\%J.o},
-where \texttt{\%J} is short-hand for the job-ID; it also tells the
-scheduler what version of \texttt{mpich} to use (\texttt{mpich\_gm})
-and how to name this job (\texttt{go\_mesher\_solver\_lsf}).
-\item The script then creates a list of the nodes allocated to this job
-by echoing the value of a dynamically set environment variable \texttt{LSB\_MCPU\_HOSTS}
-and parsing the output into a one-column list using the Perl script
-\texttt{utils/Cluster/lsf/remap\_lsf\_machines.pl}. It then creates a set of scratch
-directories on these nodes (\texttt{\small /scratch/}~\\
-\texttt{\small \$USER/DATABASES\_MPI}) to be used as the \texttt{LOCAL\_PATH}
-for temporary storage of the database files. The scratch directories
-are created using \texttt{shmux}, a shell multiplexor that can execute
-the same commands on many hosts in parallel. \texttt{shmux} is available
-from Shmux \urlwithparentheses{web.taranis.org/shmux/}. Make sure that the \texttt{LOCAL\_PATH}
-parameter in \texttt{in\_data\_files/Par\_file} is also set properly.
-\item The next portion of the script launches the mesher, database generator and then the solver
-using \texttt{run\_lsf.bash}.
-\item The final portion of the script collects the seismograms and performs
-clean up on the nodes, using the Perl scripts \texttt{collect\_seismo\_lsf\_multi.pl}
-and \texttt{cleanmulti.pl}.
-\end{enumerate}
-\begin{lyxcode}
-\#!/bin/bash~-v~
-
-\#BSUB~-o~in\_out\_files/OUTPUT\_FILES/\%J.o~
-
-\#BSUB~-a~mpich\_gm~
-
-\#BSUB~-J~go\_mesher\_solver\_lsf~\\
-~\\
-\#~set~up~local~scratch~directories
-
-BASEMPIDIR=/scratch/\$USER/DATABASES\_MPI
-
-mkdir~-p~in\_out\_files/OUTPUT\_FILES~
-
-echo~\char`\"{}\$LSB\_MCPU\_HOSTS\char`\"{}~>~in\_out\_files/OUTPUT\_FILES/lsf\_machines~
-
-echo~\char`\"{}\$LSB\_JOBID\char`\"{}~>~in\_out\_files/OUTPUT\_FILES/jobid
-
-remap\_lsf\_machines.pl~in\_out\_files/OUTPUT\_FILES/lsf\_machines~> in\_out\_files/OUTPUT\_FILES/machines
-
-shmux~-M50~-Sall~-c~\char`\"{}rm~-r~-f~/scratch/\$USER;~\textbackslash{}
-
-~~~~~~mkdir~-p~/scratch/\$USER;~mkdir~-p~\$BASEMPIDIR\char`\"{}~\textbackslash{}
-
-~~~~~~-~<~in\_out\_files/OUTPUT\_FILES/machines~>/dev/null~\\
-~\\
-\#~run~the~specfem~program
-
-current\_pwd=\$PWD \\
-
-cd bin/ \\
-
-run\_lsf.bash~-{}-gm-no-shmem~-{}-gm-copy-env~\$current\_pwd/xmeshfem3D~\\
-
-run\_lsf.bash~-{}-gm-no-shmem~-{}-gm-copy-env~\$current\_pwd/xgenerate\_databases~\\
-
-run\_lsf.bash~-{}-gm-no-shmem~-{}-gm-copy-env~\$current\_pwd/xspecfem3D~\\
-~\\
-\#~collect~seismograms~and~clean~up
-
-cd current\_pwd/~ \\
-
-mkdir~-p~in\_out\_files/SEM
-
-cd~in\_out\_files/SEM/~
-
-collect\_seismo.pl~../OUTPUT\_FILES/lsf\_machines
-
-cleanbase.pl~../OUTPUT\_FILES/machines
-\end{lyxcode}
-
-\chapter{{\normalsize \label{cha:-Changing-the}} Changing the Model}
-
-In this section we explain how to change the velocity model used for your simulations.
-These changes involve contributing specific subroutines
-that replace existing subroutines in the \texttt{SPECFEM3D} package.
-Note that \texttt{SPECFEM3D} can handle Earth models with material properties that vary within each spectral element.
-
-%% magnoni 6/12
-\section{{\normalsize \label{sec:Using-tomographic}}Using external tomographic Earth models}
-
-To implement your own external tomographic model, you must provide your own tomography file
-\texttt{tomography\_model.xyz} and choose between two possible options:
-(1) set in the \texttt{Par\_file} the model parameter \texttt{MODEL = tomo}, or
-(2) for more user control, set in the \texttt{Par\_file} the model parameter \texttt{MODEL = default} and use the following format in the file \texttt{MESH/nummaterial\_velocity\_file} when using CUBIT to construct your mesh (see also section \ref{subsec:Exporting-the-Mesh}):
-\begin{lyxcode}
-\texttt{domain\_ID} \texttt{material\_ID} \texttt{tomography} \texttt{elastic} \texttt{file\_name} 1
-\end{lyxcode}
-where
-\texttt{domain\_ID} is 1 for acoustic or 2 for elastic materials,
-\texttt{material\_ID} a negative, unique identifier (i.e., -1),
-\texttt{tomography} keyword for tomographic material definition,
-\texttt{elastic} keyword for elastic material definition,
-\texttt{file\_name} the name of the tomography file and
-1 a positive unique identifier.
-The name of the tomography file is not used so far, rather change the filename (\texttt{TOMO\_FILENAME}) defined in the file \texttt{model\_tomography.f90} located in the \texttt{src/generate\_databases/} directory.
-
-The external tomographic model is represented by a grid of points with assigned material properties and homogeneous resolution along each spatial direction x, y and z.
-The ASCII file \texttt{tomography\_model.xyz} that describes the tomography should be located in the \texttt{in\_data\_files/} directory.
-The format of the file, as read from \texttt{model\_tomography.f90}, looks like Figure \ref{fig:tomography_file},
-where
-%
-\begin{description}
-\item [{\texttt{ORIG\_X}, \texttt{END\_X}}] are, respectively, the coordinates of the initial and final tomographic grid points along the x direction (in the mesh units, e.g., $m$);
-\item [{\texttt{ORIG\_Y}, \texttt{END\_Y}}] respectively the coordinates of the initial and final tomographic grid points along the y direction (in the mesh units, e.g., $m$);
-\item [{\texttt{ORIG\_Z}, \texttt{END\_Z}}] respectively the coordinates of the initial and final tomographic grid points along the z direction (in the mesh units, e.g., $m$);
-\item [\texttt{SPACING\_X}, \texttt{SPACING\_Y}, \texttt{SPACING\_Z}] the spacing between the tomographic grid points along the x, y and z directions, respectively (in the mesh units, e.g., $m$);
-\item [\texttt{NX}, \texttt{NY}, \texttt{NZ}] the number of grid points along the spatial directions x, y and z, respectively; \texttt{NX} is given by [(\texttt{END\_X} - \texttt{ORIG\_X})/\texttt{SPACING\_X}]+1; \texttt{NY} and \texttt{NZ} are the same as \texttt{NX}, but for the y and z directions, respectively;
-\item [\texttt{VP\_MIN}, \texttt{VP\_MAX}, \texttt{VS\_MIN}, \texttt{VS\_MAX}, \texttt{RHO\_MIN}, \texttt{RHO\_MAX}] the minimum and maximum values of the wave speed \texttt{vp} and \texttt{vs} (in $m \, s^{-1}$) and of the density \texttt{rho} (in $kg \, m^{-3}$); these values could be the actual limits of the tomographic parameters in the grid or the minimum and maximum values to which we force the cut of velocity and density in the model.
-\end{description}
-%
-After the first four lines, in the file \texttt{tomography\_model.xyz} the tomographic grid points are listed with the corresponding values of \texttt{vp}, \texttt{vs} and \texttt{rho}, scanning the grid along the x coordinate (from \texttt{ORIG\_X} to \texttt{END\_X} with step of \texttt{SPACING\_X}) for each given y (from \texttt{ORIG\_Y} to \texttt{END\_Y}, with step of \texttt{SPACING\_Y}) and each given z (from \texttt{ORIG\_Z} to \texttt{END\_Z}, with step of \texttt{SPACING\_Z}).
-%fig
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=1.0\textwidth]{figures/tomo_file.jpg}
-\par\end{centering}
-\caption{Tomography file \texttt{tomography\_model.xyz} that describes an external Earth model and that is read by \texttt{model\_tomography.f90}.
-The coordinates x, y and z of the grid, their limits ORIG\_X, ORIG\_Y, ORIG\_Z, END\_X, END\_Y, END\_Z and the grid spacings SPACING\_X, SPACING\_Y, SPACING\_Z should be in the units of the constructed mesh (e.g., UTM coordinates in meters).}
-\label{fig:tomography_file}
-\end{figure}
-
-The user can implement his own interpolation algorithm for the tomography model by changing the routine \texttt{model\_tomography.f90} located in the \texttt{src/generate\_databases/} directory.
-Moreover, for models that involve both fully defined materials and a tomography description, the \texttt{nummaterial\_velocity\_file} has multiple lines each with the corresponding suitable format described above.
-
-\section{{\normalsize \label{sec:Anelastic-Models}}External (an)elastic Models}
-
-To use your own external model, you can set in the \texttt{Par\_file} the model parameter \texttt{MODEL = external}.
-Three-dimensional acoustic and/or (an)elastic (attenuation) models may then be superimposed
-onto the mesh based upon your subroutine in
-\texttt{model\_external\_values.f90} located in directory \texttt{src/generate\_databases}.
-The call to this routine would be as follows
-
-\begin{lyxcode}
-call~model\_external\_values(xmesh,~ymesh,~zmesh, ~\& \\
-~~~~~~~~~~rho,~vp,~vs,~qmu\_atten,~iflag\_aniso,~idomain\_id)
-\end{lyxcode}
-%
-Input to this routine consists of:
-\begin{description}
-\item [{\texttt{xmesh,ymesh,zmesh}}] location of mesh point
-\end{description}
-%
-Output to this routine consists of:
-\begin{description}
-\item [{\texttt{rho,vp,vs}}] isotropic model parameters for density $\rho$ ($kg/m^3$), $v_p$ ($m/s$) and $v_s$ ($m/s$)
-\item [{\texttt{qmu\_atten}}] Shear wave quality factor: $0<Q_{\mu}<9000$
-\item [{\texttt{iflag\_aniso}}] anisotropic model flag, $0$ indicating no anisotropy or $1$ using anisotropic model parameters as defined in routine file \texttt{model\_aniso.f90}
-\item [{\texttt{idomain\_id}}] domain identifier, $1$ for acoustic, $2$ for elastic, $3$ for poroelastic materials.
-\end{description}
-
-Note that the resolution and maximum value of anelastic models are
-truncated. This speeds the construction of the standard linear solids
-during the meshing stage. To change the resolution, currently at one
-significant figure following the decimal, or the maximum value (9000),
-consult \texttt{shared/constants.h}.
-
-\section{{\normalsize \label{sec:Anisotropic-Models}}Anisotropic Models}
-
-To use your anisotropic models, you can either set in the \texttt{Par\_file} the model parameter \texttt{ANISOTROPY} to \texttt{.true.} or set the model parameter \texttt{MODEL = aniso}.
-Three-dimensional anisotropic models may be superimposed on
-the mesh based upon the subroutine in file \texttt{model\_aniso.f90}
-located in directory \texttt{src/generate\_databases}.
-The call to this subroutine is of the form
-
-\begin{lyxcode}
-call~model\_aniso(iflag\_aniso,~rho,~vp,~vs,~\& \\
-~~~~~~~~~~c11,c12,c13,c14,c15,c16,c22,c23,c24,c25,c26,~\& \\
-~~~~~~~~~~c33,c34,c35,c36,c44,c45,c46,c55,c56,c66)~
-\end{lyxcode}
-%
-Input to this routine consists of:
-\begin{description}
-\item [{\texttt{iflag\_aniso}}] flag indicating the type of the anisotropic model, i.e. $0$ for no superimposed anisotropy, $1$ or $2$ for generic pre-defined anisotropic models.
-\item [{\texttt{rho,vp,vs}}] reference isotropic model parameters for density $\rho$, $v_p$ and $v_s$.
-\end{description}
-%
-Output from the routine consists of the following non-dimensional
-model parameters:
-\begin{description}
-\item [{\texttt{c11},}] \textbf{$\cdots$,} \texttt{\textbf{c66}} 21~dimensionalized
-anisotropic elastic parameters.
-\end{description}
-
-You can replace the \texttt{model\_aniso.f90} file by
-your own version \textit{provided you do not change the call structure
-of the routine}, i.e., the new routine should take exactly the same
-input and produce the required relative output.
-
-\chapter{Post-Processing Scripts}
-
-Several post-processing scripts/programs are provided in the \texttt{utils/}
-directory, and most of them need to be adjusted when used on different
-systems, for example, the path of the executable programs. Here we
-only list a few of the available scripts and provide a brief description, and
-you can either refer to the related sections for detailed usage or,
-in a lot of cases, type the script/program name without arguments
-for its usage.
-
-\section{Process Data and Synthetics\label{sec:Process-data-and-syn}}
-
-In many cases, the SEM synthetics are calculated and compared to data
-seismograms recorded at seismic stations. Since the SEM synthetics
-are accurate for a certain frequency range, both the original data
-and the synthetics need to be processed before a comparison can be
-made.
-
- For such comparisons, the following steps are recommended:
-
-\begin{enumerate}
-\item Make sure that both synthetic and observed seismograms have the correct station/event and timing information.
-\item Convolve synthetic seismograms with a source time function with the half duration specified in the \texttt{CMTSOLUTION} file, provided, as recommended, you used a zero half duration in the SEM simulations.
-\item Resample both observed and synthetic seismograms to a common sampling rate.
-\item Cut the records using the same window.
-\item Remove the trend and mean from the records and taper them.
-\item Remove the instrument response from the observed seismograms (recommended) or convolve the synthetic seismograms with the instrument response.
-\item Make sure that you apply the same filters to both observed and synthetic seismograms. Preferably, avoid filtering your records more than once.
-\item Now, you are ready to compare your synthetic and observed seismograms.
-\end{enumerate}
-
-\noindent
-We generally use the following scripts provided in the \texttt{utils/seis\_process/} directory:
-
-\subsection{Data processing script \texttt{process\_data.pl}}
-
-This script cuts a given portion of the original data, filters it,
-transfers the data into a displacement record, and picks the first
-P and S arrivals. For more functionality, type `\texttt{process\_data.pl}'
-without any argument. An example of the usage of the script:
-
-\begin{lyxcode}
-{\footnotesize process\_data.pl~-m~CMTSOLUTION~-s~1.0~-l~0/4000~-i~-f~-t~40/500~-p~-x~bp~DATA/1999.330{*}.BH?.SAC}
-{\footnotesize \par}
-\end{lyxcode}
-
-\noindent which has resampled the SAC files to a sampling rate of 1 seconds, cut
-them between 0 and 4000 seconds, transfered them into displacement
-records and filtered them between 40 and 500 seconds,
-picked the first P and S arrivals, and added suffix `\texttt{bp}' to the file names.
-
-Note that all of the scripts in this section actually use the SAC
-and/or IASP91 to do the core operations; therefore make sure that
-the SAC and IASP91 packages are installed properly on your system,
-and that all the environment variables are set properly before running
-these scripts.
-
-\subsection{Synthetics processing script \texttt{process\_syn.pl}}
-
-This script converts the synthetic output from the SEM code from ASCII
-to SAC format, and performs similar operations as `\texttt{process\_data.pl}'.
-An example of the usage of the script:
-\begin{lyxcode}
-{\footnotesize process\_syn.pl~-m~CMTSOLUTION~-h~-a~STATIONS~-s~1.0~-l~0/4000~-f~-t~40/500~-p~-x~bp~SEM/{*}.BX?.semd}
-{\footnotesize \par}
-\end{lyxcode}
-which will convolve the synthetics with a triangular source-time function
-from the \texttt{CMTSOLUTION} file, convert the synthetics into SAC
-format, add event and station information into the SAC headers, resample the SAC files with a sampling rate of 1 seconds, cut them between 0 and 4000 seconds, filter them between 40 and
-500 seconds with the same filter used for the observed data, pick the first P and S arrivals, and add the suffix `\texttt{bp}'
-to the file names.
-
-More options are available for this script, such as adding time shift
-to the origin time of the synthetics, convolving the synthetics with
-a triangular source time function with a given half duration, etc.
-Type \texttt{process\_syn.pl} without any argument for a detailed
-usage.
-
-In order to convert between SAC format and ASCII files, useful scripts
-are provided in the subdirectories ~\\
-\texttt{utils/sac2000\_alpha\_convert/} and \texttt{utils/seis\_process/asc2sac/}.
-
-\subsection{Script \texttt{rotate.pl}}
-
-The original data and synthetics have three components: vertical (BHZ resp. BXZ),
-north (BHN resp. BXN) and east (BHE resp. BXE). However, for most seismology applications,
-transverse and radial components are also desirable. Therefore, we
-need to rotate the horizontal components of both the data and the
-synthetics to the transverse and radial direction, and \texttt{\small rotate.pl}
-can be used to accomplish this:
-
-\begin{lyxcode}
-rotate.pl~-l~0~-L~180~-d~DATA/{*}.BHE.SAC.bp~
-
-rotate.pl~-l~0~-L~180~SEM/{*}.BXE.semd.sac.bp~
-\end{lyxcode}
-where the first command performs rotation on the SAC data obtained
-through Seismogram Transfer Program (STP) \urlwithparentheses{http://www.data.scec.org/STP/stp.html},
-while the second command rotates the processed SEM synthetics.
-
-\section{Collect Synthetic Seismograms}
-
-The forward and adjoint simulations generate synthetic seismograms
-in the \texttt{in\_out\_files/OUTPUT\_FILES/} directory by default. For the forward simulation, the files
-are named like \texttt{STA.NT.BX?.semd} for two-column time series, or
-\texttt{STA.NT.BX?.semd.sac} for ASCII SAC format, where STA and NT
-are the station name and network code, and \texttt{BX?} stands for
-the component name. Please see the Appendix~\ref{cha:Coordinates} and \ref{cha:channel-codes}
-for further details.
-
-The adjont simulations generate synthetic seismograms
-with the name \texttt{S?????.NT.S??.sem} (refer to Section \ref{sec:Adjoint-simulation-sources}
-for details). The kernel simulations output the back-reconstructed
-synthetic seismogram in the name \texttt{STA.NT.BX?.semd}, mainly
-for the purpose of checking the accuracy of the reconstruction. Refer
-to Section \ref{sec:Adjoint-simulation-finite} for further details.
-
-You do have further options to change this default output behavior, given in the main constants file
-\texttt{constants.h} located in \texttt{src/shared/} directory:
-\begin{description}
-\item [{\texttt{SEISMOGRAMS\_BINARY}}] set to \texttt{.true.} to have seismograms written out in binary format.
-\item [{\texttt{WRITE\_SEISMOGRAMS\_BY\_MASTER}}] Set to \texttt{.true.} to have only the master process writing out seismograms.
-This can be useful on a cluster, where only the master process node has access to the output directory.
-\item [{\texttt{USE\_OUTPUT\_FILES\_PATH}}] Set to \texttt{.false.} to have the seismograms output to \texttt{LOCAL\_PATH} directory specified in the main parameter file \texttt{in\_data\_files/Par\_file}.
-In this case, you could collect the synthetics onto the frontend using the \texttt{collect\_seismo\_lsf\_multi.pl} script
-located in the \texttt{utils/Cluster/lsf/} directory. The usage of the script would be e.g.:
-
-\begin{lyxcode}
-collect\_seismo.pl~machines~in\_data\_files/Par\_file
-\end{lyxcode}
-
-\end{description}
-where  \texttt{machines} is a file containing the node names and  \texttt{in\_data\_files/Par\_file} the parameter file used to extract the \texttt{LOCAL\_PATH} directory used for the simulation.
-
-\section{Clean Local Database}
-
-After all the simulations are done, the seismograms are collected,
-and the useful database files are copied to the frontend, you may
-need to clean the local scratch disk for the next simulation. This
-is especially important in the case of kernel simulation,
-where very large files are generated for the absorbing boundaries
-to help with the reconstruction of the regular forward wavefield.
-A sample script is provided in \texttt{utils/}:
-
-\begin{lyxcode}
-cleanbase.pl~machines
-\end{lyxcode}
-where  \texttt{machines} is a file containing the node names.
-
-\section{Plot Movie Snapshots and Synthetic Shakemaps}
-
-\subsection{Script \texttt{movie2gif.gmt.pl}}
-
-With the movie data saved in \texttt{in\_out\_files/OUTPUT\_FILES/} at the end of
-a movie simulation (\texttt{\small MOVIE\_SURFACE=.true.}), you can
-run the \texttt{`create\_movie\_shakemap\_AVS\_DX\_GMT}' code to convert these binary
-movie data into GMT xyz files for futher processing. A sample script
-\texttt{movie2gif.gmt.pl} is provided to do this conversion, and then
-plot the movie snapshots in GMT, for example:
-
-\begin{lyxcode}
-movie2gif.gmt.pl~-m~CMTSOLUTION~-g~-f~1/40~-n~-2~-p~
-\end{lyxcode}
-which for the first through the 40th movie frame, converts the \texttt{moviedata}
-files into GMT xyz files, interpolates them using the 'nearneighbor'
-command in GMT, and plots them on a 2D topography map. Note that `\texttt{-2}'
-and `\texttt{-p}' are both optional.
-
-\subsection{Script \texttt{plot\_shakemap.gmt.pl}}
-
-With the shakemap data saved in \texttt{in\_out\_files/OUTPUT\_FILES/} at the end
-of a shakemap simulation ~\\
-(\texttt{CREATE\_SHAKEMAP=.true.}), you can
-also run \texttt{`create\_movie\_shakemap\_AVS\_DX\_GMT}' code to convert the binary
-shakemap data into GMT xyz files. A sample script \texttt{plot\_shakemap.gmt.pl}
-is provided to do this conversion, and then plot the shakemaps in
-GMT, for example:
-
-\begin{lyxcode}
-plot\_shakemap.gmt.pl~data\_dir~type(1,2,3)~CMTSOLUTION~
-\end{lyxcode}
-where \texttt{type=1} for a displacement shakemap, \texttt{2} for
-velocity, and \texttt{3} for acceleration.
-
-\section{Map Local Database}
-
-A sample program \texttt{remap\_database} is provided to map the local
-database from a set of machines to another set of machines. This is
-especially useful when you want to run mesher and solver, or different
-types of solvers separately through a scheduler (refer to Chapter~\ref{cha:Scheduler}).
-\begin{lyxcode}
-run\_lsf.bash~-{}-gm-no-shmem~-{}-gm-copy-env~remap\_database~old\_machines~150
-\end{lyxcode}
-where \texttt{old\_machines} is the LSF machine file used in the previous
-simulation, and \texttt{150} is the number of processors in total.
-
-\chapter{\label{cha:tomo}Tomographic inversion using sensitivity kernels}
-
-One of the fundamental reasons for computing sensitivity kernels (Section~\ref{sec:Adjoint-simulation-finite}) is to use them within a tomographic inversion.
-In other words, use recorded seismograms, make measurements with synthetic seismograms, and use the misfit between the two to iteratively improve the model described by (at least) $V_{\rm p}$, $V_{\rm s}$, and $\rho$ as a function of space.
-
-Whatever misfit function you use for the tomographic inversion (several examples in \cite{TrTaLi05} and \cite{TrKoLi08}), you will weight the sensitivity kernels with measurements.
-Furthermore, you will use as many measurements (stations, components, time windows) as possible per event; hence, we call these composite kernels ``event kernels,'' which are volumetric fields representing the gradient of the misfit function with respect to one of the variables (\eg $V_{\rm s}$).
-The basic features of an adjoint-based tomographic inversion were illustrated in \citet{TaLiTr07} and \citet{TrKoLi08} using a conjugate-gradient algorithm; there are dozens of versions of gradient-based inversion algorithms that could alternatively be used.
-The tomographic inversion of \citet{TaLiMaTr09,TaLiMaTr2010} used SPECFEM3D as well as several additional components which are also stored on the CIG svn server, described next.
-
-The directory containing utilities for tomographic inversion using SPECFEM3D (or other packages that evaluate misfit functions and gradients) is here on the CIG svn server:
-%
-\begin{verbatim}
-/cig/seismo/3D/ADJOINT_TOMO/
-   flexwin/     -- FLEXWIN algorithm for automated picking of time windows
-   measure_adj/ -- reads FLEXWIN output file and makes measurements,
-                   with the option for computing adjoint sources
-   iterate_adj/ -- various tools for iterative inversion
-                   (requires pre-computed "event kernels")
-\end{verbatim}
-%
-This directory also contains a brief \verb+README+ file indicating the role of the three subdirectories, \verb+flexwin+ \citep{Maggi2009}, \verb+measure_adj+, and \verb+iterate_adj+.
-The components for making the model update are there; however, there are no explicit rules for computing the model update, just as with any optimization problem.
-There are options for computing a conjugate gradient step, as well as a source subspace projection step.
-The workflow could use substantial ``stitching together,'' and we welcome suggestions; undoubtedly, the tomographic inversion procedure will improve with time.
-{\bf This is a work in progress.}
-
-The best single file to read is probably: \verb+ADJOINT_TOMO/iterate_adj/cluster/README+.
-
-%============================================================
-
-\chapter*{\label{cha:Bug-Reports-and}Bug Reports and Suggestions for Improvements}
-
-To report bugs or suggest improvements to the code, please send an
-e-mail to the CIG Computational Seismology Mailing List \urlwithparentheses{cig-seismo@geodynamics.org}
-or Jeroen Tromp \urlwithparentheses{jtromp-AT-princeton.edu}, and/or use our online
-bug tracking system Roundup \urlwithparentheses{www.geodynamics.org/roundup}.
-
-\addcontentsline{toc}{chapter}{Notes \& Acknowledgments}
-\chapter*{Notes \& Acknowledgments}
-
-In order to keep the software package thread-safe in case a multithreaded
-implementation of MPI is used, developers should not add modules or
-common blocks to the source code but rather use regular subroutine
-arguments (which can be grouped in ``derived types'' if needed for
-clarity).
-
-The Gauss-Lobatto-Legendre subroutines in \texttt{gll\_library.f90}
-are based in part on software libraries from the Massachusetts Institute
-of Technology, Department of Mechanical Engineering (Cambridge, Massachusetts, USA).
-The non-structured global numbering software was provided by Paul
-F. Fischer (Brown University, Providence, Rhode Island, USA, now at Argonne National Laboratory, USA).
-
-OpenDX \urlwithparentheses{http://www.opendx.org} is open-source based on IBM Data
-Explorer, AVS \urlwithparentheses{http://www.avs.com} is a trademark of Advanced
-Visualization Systems, and ParaView \urlwithparentheses{http://www.paraview.org}
-is an open-source visualization platform.
-
-Please e-mail your feedback, questions, comments, and suggestions
-to Jeroen Tromp \urlwithparentheses{jtromp-AT-princeton.edu} or to the CIG Computational Seismology Mailing List \urlwithparentheses{cig-seismo@geodynamics.org}.
-
-\addcontentsline{toc}{chapter}{Copyright}
 \chapter*{Copyright}
 
-Main authors: Dimitri Komatitsch and Jeroen Tromp
+Main historical authors: Dimitri Komatitsch and Jeroen Tromp
 
-Princeton University, USA, and CNRS / INRIA / University of Pau, France
+Princeton University, USA, and CNRS / University of Marseille, France
 
-$\copyright$ Princeton University and CNRS / INRIA / University of Pau, July 2012
+$\copyright$ Princeton University and CNRS / University of Marseille, July 2012
 
@@
 
- MPI v 2.1, July 2012:
- Max Rietmann, Peter Messmer, Daniel Peter, Dimitri Komatitsch, Joseph Charles, Zhinan Xie: support for CUDA GPUs,
-better CFL stability for the Stacey absorbing conditions. \\
+MPI v. 3.0, December 2014: many developers.
+Convolutional PML, LDDRK time scheme, bulk attenuation support, simultaneous MPI runs,
+ADIOS file I/O support, coupling with external codes, new seismogram names,
+Deville routines for additional GLL degrees, tomography tools, unit/regression test framework,
+improved CUDA GPUs performance, additonal GEOCUBIT support, better make compilation,
+git versioning system. \\
 
- MPI v. 2.0, November 2010:
- Dimitri Komatitsch, Nicolas Le Goff, Roland Martin and Pieyre Le Loher, University of Pau, France,
- Jeroen Tromp and the Princeton group of developers, Princeton University, USA,
- and Emanuele Casarotti, INGV Roma, Italy:
-  support for CUBIT meshes decomposed by SCOTCH, METIS or ZOLTAN;
-  much faster solver using Michel Deville's inlined matrix products.\\
+MPI v. 2.1, July 2012: Max Rietmann, Peter Messmer, Daniel Peter, Dimitri
+Komatitsch, Joseph Charles, Zhinan Xie: support for CUDA GPUs, better
+CFL stability for the Stacey absorbing conditions. \\
 
- MPI v. 1.4 Dimitri Komatitsch, University of Pau, Qinya Liu and others, Caltech, September 2006:
-  better adjoint and kernel calculations, faster and better I/Os
-  on very large systems, many small improvements and bug fixes.\\
+MPI v. 2.0, November 2010: Dimitri Komatitsch, Nicolas Le Goff, Roland
+Martin and Pieyre Le Loher, University of Pau, France, Daniel Peter,
+Jeroen Tromp and the Princeton group of developers, Princeton University,
+USA, and Emanuele Casarotti, INGV Roma, Italy: support for CUBIT meshes
+decomposed by SCOTCH; much faster solver using Michel Deville's inlined
+matrix products.\\
 
- MPI v. 1.3 Dimitri Komatitsch, University of Pau, and Qinya Liu, Caltech, July 2005:
-  serial version, regular mesh, adjoint and kernel calculations, ParaView support.\\
+MPI v. 1.4 Dimitri Komatitsch, University of Pau, Qinya Liu and others,
+Caltech, September 2006: better adjoint and kernel calculations, faster
+and better I/Os on very large systems, many small improvements and
+bug fixes.\\
 
- MPI v. 1.2 Min Chen and Dimitri Komatitsch, Caltech, July 2004:
-  full anisotropy, volume movie.\\
+MPI v. 1.3 Dimitri Komatitsch, University of Pau, and Qinya Liu, Caltech,
+July 2005: serial version, regular mesh, adjoint and kernel calculations,
+ParaView support.\\
 
- MPI v. 1.1 Dimitri Komatitsch, Caltech, October 2002: Zhu's Moho map, scaling
-  of $V_s$ with depth, Hauksson's regional model, attenuation, oceans, movies.\\
+MPI v. 1.2 Min Chen and Dimitri Komatitsch, Caltech, July 2004: full
+anisotropy, volume movie.\\
 
- MPI v. 1.0 Dimitri Komatitsch, Caltech, USA, May 2002: first MPI version based on global code.\\
+MPI v. 1.1 Dimitri Komatitsch, Caltech, October 2002: Zhu's Moho map,
+scaling of $V_{s}$ with depth, Hauksson's regional model, attenuation,
+oceans, movies.\\
 
- Dimitri Komatitsch, IPG Paris, France, December 1996: first 3-D solver for the CM-5 Connection Machine,
-    parallelized on 128 processors using Connection Machine Fortran.\\
+MPI v. 1.0 Dimitri Komatitsch, Caltech, USA, May 2002: first MPI version
+based on global code.\\
 
-\addcontentsline{toc}{chapter}{References}
-\bibliography{bibliography}
-
-\appendix
-
-\chapter{\label{cha:Coordinates}Reference Frame Convention}
-
-The code uses the following convention for the Cartesian reference
-frame:
-
-\begin{itemize}
-\item the $x$ axis points East
-\item the $y$ axis points North
-\item the $z$ axis points up
-\end{itemize}
-Note that this convention is different from both the \citet{AkRi80}
-convention and the Harvard Centroid-Moment Tensor (CMT) convention.
-The Aki \& Richards convention is
-
-\begin{itemize}
-\item the $x$ axis points North
-\item the $y$ axis points East
-\item the $z$ axis points down
-\end{itemize}
-and the Harvard CMT convention is
-
-\begin{itemize}
-\item the $x$ axis points South
-\item the $y$ axis points East
-\item the $z$ axis points up
-\end{itemize}
-
-\subsection*{Source and receiver locations}
-
-The SPECFEM3D software code internally uses Cartesian coordinates.
-The given locations for sources and receiver locations thus may get
-converted.  Sources and receiver locations are read in from the
-\texttt{CMTSOLUTION} (or \texttt{FORCESOLUTION}) and \texttt{STATIONS}
-files.  Note that e.g. the \texttt{CMTSOLUTION} and
-\texttt{FORCESOLUTION} files denotes the location by
-"longitude/latitude/depth". We read in longitude as $x$ coordinate,
-latitude as $y$ coordinate.
-
-In case the flag \texttt{SUPPRESS\_UTM\_PROJECTION} is set to \texttt{.false.}
-in the main parameter file (see Chapter~\ref{cha:Creating-Distributed-Databases}),
-the $x$/$y$ coordinates have to be given in degrees and are converted to Cartesian coordinates
-using a UTM conversion with the specified UTM zone.
-
-The value for depth (given in $km$ in \texttt{CMTSOLUTION} and
-\texttt{FORCESOLUTION}) or burial depth (given in $m$ in
-\texttt{STATIONS}) is evaluated with respect to the surface of the
-mesh at the specified $x$/$y$ location to find the corresponding $z$
-coordinate.  It is possible to use this depth value directly as $z$
-coordinate by changing the flag \texttt{USE\_SOURCES\_RECVS\_Z} to
-\texttt{.true.} in the file \texttt{constants.h} located in the
-\texttt{src/shared/} subdirectory.
-
-\subsection*{Seismogram outputs}
-
-The seismogram output directions are given in Cartesian $x$/$y$/$z$ directions and not rotated any further.
-Changing flags in \texttt{constants.h} in the \texttt{src/shared/} subdirectory only rotates the seismogram outputs
-if receivers are forced to be located at the surface (\texttt{RECVS\_CAN\_BE\_BURIED\_EXT\_MESH} set to \texttt{.false.}) and the normal to the surface at the receiver location should be used (\texttt{EXT\_MESH\_RECV\_NORMAL} set to
-\texttt{.true.}) as vertical.
-In this case, the outputs are rotated to have the vertical component normal to the surface of the mesh, $x$ and $y$ directions are somewhat arbitrary orthogonal directions along the surface.
-
-For the labeling of the seismogram channels, see Appendix~\ref{cha:channel-codes}. Additionally, we add labels to the synthetics using the following convention:
-For a receiver station located in an
-\begin{description}
-\item[elastic domain:] ~\\
-\begin{itemize}
-\item \texttt{semd} for the displacement vector
-\item \texttt{semv} for the velocity vector
-\item \texttt{sema} for the acceleration vector
-\end{itemize}
-\item[acoustic domain:] ~\\
-(please note that receiver stations in acoustic domains must be buried. This is due to the free surface condition which enforces a zero pressure at the surface.)
-\begin{itemize}
-\item \texttt{semd} for the displacement vector
-\item \texttt{semv} for the velocity vector
-\item \texttt{sema} for pressure which will be stored in the vertical component \texttt{Z}.
-Note that pressure in the acoustic media is isotropic and thus the pressure record would be the same in the other two component directions.
-We therefore use the other two seismogram components to store the acoustic potential in component \texttt{X} (or \texttt{N}) and the first time derivative of the acoustic potential in component \texttt{Y} (or \texttt{E}).
-\end{itemize}
-\end{description}
-The seismograms are by default written out in ASCII-format to the \texttt{in\_out\_files/OUTPUT\_FILES/} subdirectory by each parallel process.
-You can change this behavior by changing the following flags in the \texttt{constants.h} file located in the \texttt{src/shared/} subdirectory:
-\begin{description}
-\item [{\texttt{SEISMOGRAMS\_BINARY}}] set to \texttt{.true.} to have seismograms written out in binary format.
-\item [{\texttt{WRITE\_SEISMOGRAMS\_BY\_MASTER}}] Set to \texttt{.true.} to have only the master process writing out seismograms.
-This can be useful on a cluster, where only the master process node has access to the output directory.
-\item [{\texttt{USE\_OUTPUT\_FILES\_PATH}}] Set to \texttt{.false.} to have the seismograms output to \texttt{LOCAL\_PATH} directory specified in the main parameter file \texttt{in\_data\_files/Par\_file}.
-\end{description}
-
-\chapter{\label{cha:channel-codes}Channel Codes of Seismograms}
-
-Seismic networks, such as the Global Seismographic Network (GSN), generally involve various types of instruments with different bandwidths, sampling properties and component configurations.
-There are standards to name channel codes depending on instrument properties.
-IRIS  \texttt{(www.iris.edu)} uses SEED format for channel codes, which are represented by three letters, such as \texttt{LHN}, \texttt{BHZ}, etc.
-In older versions of the SPECFEM3D package, a common format was used for the channel codes of all seismograms, which was \texttt{BHE/BHN/BHZ} for three components.
-To avoid confusion when comparison are made to observed data, we are now using the FDSN convention (\url{http://www.fdsn.org/}) for SEM seismograms.
-In the following, we give a brief explanation of the FDSN convention used by IRIS, and how it is adopted in SEM seismograms.
-Please visit \texttt{www.iris.edu/manuals/SEED\_appA.htm} for further information.\\
-
-\noindent \textbf{\texttt{Band code:}} The first letter in the channel code denotes the band code of seismograms, which depends on the response band and the sampling rate of instruments.
-The list of band codes used by IRIS is shown in Figure \ref{fig:IRIS_band_codes}.
-The sampling rate of SEM synthetics is controlled by the resolution of simulations rather than instrument properties.
-However, for consistency, we follow the FDSN convention for SEM seismograms governed by their sampling rate.
-For SEM synthetics, we consider band codes for which $dt \leq 1$ s.
-IRIS also considers the response band of instruments.
-For instance, short-period and broad-band seismograms with the same sampling rate correspond to different band codes, such as S and B, respectively.
-In such cases, we consider SEM seismograms as broad band, ignoring the corner period ($\geq 10$ s) of the response band of instruments (note that at these resolutions, the minimum period in the SEM synthetics will be less than $10$ s).
-Accordingly, when you run a simulation the band code will be chosen depending on the resolution of the synthetics, and channel codes of SEM seismograms will start with either \texttt{L}, \texttt{M}, \texttt{B}, \texttt{H}, \texttt{C} or \texttt{F}, shown by red color in the figure.\\
-
-\begin{figure}[ht]
-\noindent \begin{centering}
-\includegraphics[scale=0.6]{figures/IRIS_band_codes.pdf}
-\par\end{centering}
-
-\caption{The band code convention is based on the sampling rate and the response band of instruments.
-Please visit \texttt{www.iris.edu/manuals/SEED\_appA.htm} for further information.
-Grey rows show the relative band-code range in SPECFEM3D, and the band codes used to name SEM seismograms are denoted in red.}
-
-\label{fig:IRIS_band_codes}
-\end{figure}
-
-\noindent \textbf{\texttt{Instrument code:}} The second letter in the channel code corresponds to instrument codes, which specify the family to which the sensor belongs.
-For instance, \texttt{H} and \texttt{L} are used for high-gain and low-gain seismometers, respectively.
-The instrument code of SEM seismograms will always be \texttt{X}, as assigned by FDSN for synthetic seismograms.
-\\
-
-\noindent \textbf{\texttt{Orientation code:}} The third letter in channel codes is an orientation code, which generally describes the physical configuration of the components of instrument packages.
-SPECFEM3D uses the traditional orientation code  \texttt{E/N/Z} (East-West, North-South, Vertical) for three components when a UTM projection is used.
-If the UTM conversion is suppressed, i.e.
-the flag \texttt{SUPPRESS\_UTM\_PROJECTION} is set to \texttt{.true.}, then the three components are labelled \texttt{X/Y/Z} according to the Cartesian reference frame. \\
-
-\noindent \textbf{\texttt{EXAMPLE:}} The sampling rate is given by \texttt{DT} in the main parameter file \texttt{Par\_file} located in the \texttt{in\_data\_files/} subdirectory.
-Depending on the resolution of your simulations, if you choose a sampling rate greater than $0.01$ s and less than $1$ s, a seismogram recording displacements on the vertical component of a station \texttt{ASBS} (network \texttt{AZ}) will be named \texttt{ASBS.AZ.MXZ.semd.sac}, whereas it will be \texttt{ASBS.AZ.BXZ.semd.sac}, if the sampling rate is greater than 0.0125 and less equal to 0.1 s.
-
-\chapter{\label{cha:Troubleshooting}Troubleshooting}
-
-\section*{FAQ}
-
-\begin{description}
-\item [configuration fails:]
-
-   Examine the log file 'config.log'. It contains detailed informations.
-   In many cases, the path's to these specific compiler commands F90,
-   CC and MPIF90 won't be correct if `./configure` fails.
-
-   Please make sure that you have a working installation of a Fortran compiler,
-   a C compiler and an MPI implementation. You should be able to compile this
-   little program code:
-
-\begin{lyxcode}
-
-      program main ~\\
-        include 'mpif.h' ~\\
-        integer, parameter :: CUSTOM\_MPI\_TYPE = MPI\_REAL ~\\
-        integer ier ~\\
-        call MPI\_INIT(ier) ~\\
-        call MPI\_BARRIER(MPI\_COMM\_WORLD,ier) ~\\
-        call MPI\_FINALIZE(ier) ~\\
-      end
-
-\end{lyxcode}
-
-\item [compilation fails stating:] ~\\
-\begin{lyxcode}
-    ...  ~\\
-    obj/program\_generate\_databases.o: In function `MAIN\_\_':  ~\\
-    program\_generate\_databases.f90:(.text+0x14): undefined reference to `\_gfortran\_set\_std'  ~\\
-    ...  ~\\
-\end{lyxcode}
-
-  Make sure you're pointing to the right 'mpif90' wrapper command.
-
-  Normally, this message will appear when you're mixing two different Fortran
-  compilers. That is, using e.g. gfortran to compile non-MPI files
-  and mpif90, wrapper provided for e.g. ifort, to compile MPI-files.
-
-  fix: e.g. specify > ./configure FC=gfortran MPIF90=/usr/local/openmpi-gfortran/bin/mpif90
-
-\item [after executing \texttt{xmeshfem3D} I've got elements with skewness of 81\% percent, what does this mean:]
-Look at the skewness table printed in the \texttt{output\_mesher.txt} file after executing \texttt{xmeshfem3D} for the
-example given in \texttt{examples/meshfem3D\_examples/simple\_model/}:
-\begin{lyxcode}
-...~\\
-histogram of skewness (0. good - 1. bad): ~\\
- 0.0000000E+00  -   5.0000001E-02       27648     81.81818      \% ~\\
- 5.0000001E-02  -   0.1000000               0    0.0000000E+00  \% ~\\
-... ~\\
-\end{lyxcode}
-The first line means that you have 27,648 elements with a skewness value between 0 and 0.05 (which means the element is basically not skewed, just plain regular hexahedral element).
-The total number of elements you have in this mesh is (see in the \texttt{output\_mesher.txt} file a bit further down):
-\begin{lyxcode}
-...~\\
-total number of elements in entire mesh:        33792 ~\\
-... ~\\
-\end{lyxcode}
-which gives you that:  27,648 / 33,792 $\sim$ 81.8 \% of all elements are not skewed, i.e. regular elements. a fantastic value :)
-
-The histogram lists for this mesh also some stronger skewed elements, for example the worst ones belong to:
-\begin{lyxcode}
-...~\\
-0.6000000      -   0.6500000            2048     6.060606      \% ~\\
-... ~\\
-\end{lyxcode}
-
-about 6 \% of all elements have distortions with a skewness value between 0.6 and 0.65.
-The skewness values give you a hint of how good your mesh is.
-In an ideal world, you would want to have no distortions, just like the 81\% from above.
-Those elements give you the best approximate values by the GLL quadrature used in the spectral-element method.
-However, having weakly distorted elements is still fine and the solutions are still accurate enough.
-So empirically, values up to around 0.7 are tolerable, above that you should consider remeshing...
-
-To give you an idea why some of the elements are distorted, see the following figure \ref{fig:mesh.vp} of the mesh you obtain in the example \texttt{examples/meshfem3D\_examples/simple\_model/}.
-\begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=.9\textwidth]{figures/mesh_vp.jpg}
-\par\end{centering}
-\caption{Paraview visualization using the mesh vtk-files for the example given in \texttt{examples/meshfem3D\_examples/simple\_model/}.}
-\label{fig:mesh.vp}
-\end{figure}
-You will see that the mesh contains a doubling layer, where we stitch elements together such that the size of two elements will transition to the size of one element (very useful to keep the ratio of wavespeed / element\_size about constant).
-Those elements in this doubling layer have higher skewness values and make up those 6 \% in the histogram.
-
-\item [the code gives following error message "need at least one receiver":]
-This means that no stations given in the input file \texttt{in\_data\_files/STATIONS} could be located within the dimensions of the mesh.
-This can happen for example when the mesh was created with the in-house mesher \texttt{xmeshfem3D} while using the Universal Transverse Mercator (UTM) projection but the simulation with \texttt{xspecfem3D} was suppressing this projection from latitude/longitude to x/y/z coordinates.
-
-In such cases, try to change your \texttt{in\_data\_files/Par\_file} and set e.g.:
-\begin{lyxcode}
-SUPPRESS\_UTM\_PROJECTION         = .false. ~\\
-\end{lyxcode}
-to be the same in \texttt{Mesh\_Par\_file} and \texttt{Par\_file}.
-This flag should be identical when using the in-house mesher \texttt{xmeshfem3D}, \texttt{xgenerate\_databases} and \texttt{xspecfem3D} together to run simulations.
-
-The flag determines if the coordinates you specify for your source and station locations are given as lat/lon degrees and
-must be converted to UTM coordinates.
-As an example, if you use \texttt{.false.} within \texttt{Mesh\_Par\_file} then you create a mesh with \texttt{xmeshfem3D} using the UTM projection from lat/lon as input format to UTM projected coordinates to store the mesh point positions, which is fine.
-The error then may occur if in the \texttt{Par\_file} you have this set to \texttt{.true.} so that the \texttt{xgenerate\_databases} and \texttt{xspecfem3D} suppress the UTM projection and assume that all coordinates you use now for source and receiver locations are given in meters (that is, converted) already.
-So it won't find the specified locations in the used mesh.
-As a solutions, just change the flag in \texttt{Par\_file} to be the same as in \texttt{Mesh\_Par\_file} and rerun \texttt{xgenerate\_databases} and \texttt{xspecfem3D} to make sure that your simulation works fine.
-
-\item [I get the following error message "forward simulation became unstable and blew up":]
-In most cases this means that your time step size \texttt{DT} is chosen too big.
-Look at your files \texttt{output\_mesher.txt} or \texttt{output\_solver.txt} created in the folder \texttt{in\_out\_files/OUTPUT\_FILES/}.
-In these output files, find the section:
-\begin{lyxcode}
-... ~\\
-********************************************* ~\\
-*** Verification of simulation parameters *** ~\\
-********************************************* ~\\
-... ~\\
-*** Minimum period resolved =    4.308774 ~\\
-*** Maximum suggested time step =   6.8863556E-02 ~\\
-... ~\\
-... ~\\
-\end{lyxcode}
-
-then change \texttt{DT} in the \texttt{in\_data\_files/Par\_file} to be somewhere close to the maximum suggested time step. In the example above: \\
-\texttt{DT                              = 0.05d0} \\
-would (most probably) work fine. It could be also bigger than the 0.068 s suggested. This depends a bit on the distortions of
-your mesh elements. The more regular they are, the bigger you can choose \texttt{DT}. Just play with this value a bit and see
-when the simulation becomes stable ...
-
-\end{description}
-
-\chapter{\label{cha:License}License }
-
-\textbf{GNU GENERAL PUBLIC LICENSE Version 2, June 1991. Copyright
-(C) 1989, 1991 Free Software Foundation, Inc. 59 Temple Place, Suite
-330, Boston, MA 02111-1307 USA} \\
-Everyone is permitted to copy and distribute verbatim copies of this
-license document, but changing it is not allowed.
-
-\section*{Preamble}
-
-The licenses for most software are designed to take away your freedom
-to share and change it. By contrast, the GNU General Public License
-is intended to guarantee your freedom to share and change free software
--- to make sure the software is free for all its users. This General
-Public License applies to most of the Free Software Foundation's software
-and to any other program whose authors commit to using it. (Some other
-Free Software Foundation software is covered by the GNU Library General
-Public License instead.) You can apply it to your programs, too.
-
-When we speak of free software, we are referring to freedom, not price.
-Our General Public Licenses are designed to make sure that you have
-the freedom to distribute copies of free software (and charge for
-this service if you wish), that you receive source code or can get
-it if you want it, that you can change the software or use pieces
-of it in new free programs; and that you know you can do these things.
-
-To protect your rights, we need to make restrictions that forbid anyone
-to deny you these rights or to ask you to surrender the rights. These
-restrictions translate to certain responsibilities for you if you
-distribute copies of the software, or if you modify it.
-
-For example, if you distribute copies of such a program, whether gratis
-or for a fee, you must give the recipients all the rights that you
-have. You must make sure that they, too, receive or can get the source
-code. And you must show them these terms so they know their rights.
-
-We protect your rights with two steps:
-
-\begin{enumerate}
-\item Copyright the software, and
-\item Offer you this license which gives you legal permission to copy, distribute
-and/or modify the software.
-\end{enumerate}
-Also, for each author's protection and ours, we want to make certain
-that everyone understands that there is no warranty for this free
-software. If the software is modified by someone else and passed on,
-we want its recipients to know that what they have is not the original,
-so that any problems introduced by others will not reflect on the
-original authors' reputations.
-
-Finally, any free program is threatened constantly by software patents.
-We wish to avoid the danger that redistributors of a free program
-will individually obtain patent licenses, in effect making the program
-proprietary. To prevent this, we have made it clear that any patent
-must be licensed for everyone's free use or not licensed at all.
-
-The precise terms and conditions for copying, distribution and modification
-follow.
-
-\section*{GNU GENERAL PUBLIC LICENSE TERMS AND CONDITIONS FOR COPYING, DISTRIBUTION
-AND MODIFICATION }
-
-\begin{itemize}
-
-\item[0.]This License applies to any program or other work which
-contains a notice placed by the copyright holder saying it may be
-distributed under the terms of this General Public License. The ``Program''
-below refers to any such program or work, and a ``work based on the
-Program'' means either the Program or any derivative work under copyright
-law: that is to say, a work containing the Program or a portion of
-it, either verbatim or with modifications and/or translated into another
-language. (Hereinafter, translation is included without limitation
-in the term ``modification.'') Each licensee is addressed as ``you.''\\
-\\
-Activities other than copying, distribution and modification are not
-covered by this License; they are outside its scope. The act of running
-the Program is not restricted, and the output from the Program is
-covered only if its contents constitute a work based on the Program
-(independent of having been made by running the Program). Whether
-that is true depends on what the Program does.
-
-\end{itemize}
-
-\begin{enumerate}
-\item You may copy and distribute verbatim copies of the Program's source
-code as you receive it, in any medium, provided that you conspicuously
-and appropriately publish on each copy an appropriate copyright notice
-and disclaimer of warranty; keep intact all the notices that refer
-to this License and to the absence of any warranty; and give any other
-recipients of the Program a copy of this License along with the Program.
-
-You may charge a fee for the physical act of transferring a copy,
-and you may at your option offer warranty protection in exchange for
-a fee.
-
-\item You may modify your copy or copies of the Program or any portion of
-it, thus forming a work based on the Program, and copy and distribute
-such modifications or work under the terms of Section 1 above, provided
-that you also meet all of these conditions:
-
-\begin{enumerate}
-\item You must cause the modified files to carry prominent notices stating
-that you changed the files and the date of any change.
-\item You must cause any work that you distribute or publish, that in whole
-or in part contains or is derived from the Program or any part thereof,
-to be licensed as a whole at no charge to all third parties under
-the terms of this License.
-\item If the modified program normally reads commands interactively when
-run, you must cause it, when started running for such interactive
-use in the most ordinary way, to print or display an announcement
-including an appropriate copyright notice and a notice that there
-is no warranty (or else, saying that you provide a warranty) and that
-users may redistribute the program under these conditions, and telling
-the user how to view a copy of this License. (Exception: if the Program
-itself is interactive but does not normally print such an announcement,
-your work based on the Program is not required to print an announcement.)
-\end{enumerate}
-These requirements apply to the modified work as a whole. If identifiable
-sections of that work are not derived from the Program, and can be
-reasonably considered independent and separate works in themselves,
-then this License, and its terms, do not apply to those sections when
-you distribute them as separate works. But when you distribute the
-same sections as part of a whole which is a work based on the Program,
-the distribution of the whole must be on the terms of this License,
-whose permissions for other licensees extend to the entire whole,
-and thus to each and every part regardless of who wrote it.
-
-Thus, it is not the intent of this section to claim rights or contest
-your rights to work written entirely by you; rather, the intent is
-to exercise the right to control the distribution of derivative or
-collective works based on the Program.
-
-In addition, mere aggregation of another work not based on the Program
-with the Program (or with a work based on the Program) on a volume
-of a storage or distribution medium does not bring the other work
-under the scope of this License.
-
-\item You may copy and distribute the Program (or a work based on it, under
-Section 2) in object code or executable form under the terms of Sections
-1 and 2 above provided that you also do one of the following:
-
-\begin{enumerate}
-\item Accompany it with the complete corresponding machine-readable source
-code, which must be distributed under the terms of Sections 1 and
-2 above on a medium customarily used for software interchange; or,
-\item Accompany it with a written offer, valid for at least three years,
-to give any third party, for a charge no more than your cost of physically
-performing source distribution, a complete machine-readable copy of
-the corresponding source code, to be distributed under the terms of
-Sections 1 and 2 above on a medium customarily used for software interchange;
-or,
-\item Accompany it with the information you received as to the offer to
-distribute corresponding source code. (This alternative is allowed
-only for noncommercial distribution and only if you received the program
-in object code or executable form with such an offer, in accord with
-Subsection b above.)
-\end{enumerate}
-The source code for a work means the preferred form of the work for
-making modifications to it. For an executable work, complete source
-code means all the source code for all modules it contains, plus any
-associated interface definition files, plus the scripts used to control
-compilation and installation of the executable. However, as a special
-exception, the source code distributed need not include anything that
-is normally distributed (in either source or binary form) with the
-major components (compiler, kernel, and so on) of the operating system
-on which the executable runs, unless that component itself accompanies
-the executable.
-
-If distribution of executable or object code is made by offering access
-to copy from a designated place, then offering equivalent access to
-copy the source code from the same place counts as distribution of
-the source code, even though third parties are not compelled to copy
-the source along with the object code.
-
-\item You may not copy, modify, sublicense, or distribute the Program except
-as expressly provided under this License. Any attempt otherwise to
-copy, modify, sublicense or distribute the Program is void, and will
-automatically terminate your rights under this License. However, parties
-who have received copies, or rights, from you under this License will
-not have their licenses terminated so long as such parties remain
-in full compliance.
-\item You are not required to accept this License, since you have not signed
-it. However, nothing else grants you permission to modify or distribute
-the Program or its derivative works. These actions are prohibited
-by law if you do not accept this License. Therefore, by modifying
-or distributing the Program (or any work based on the Program), you
-indicate your acceptance of this License to do so, and all its terms
-and conditions for copying, distributing or modifying the Program
-or works based on it.
-\item Each time you redistribute the Program (or any work based on the Program),
-the recipient automatically receives a license from the original licensor
-to copy, distribute or modify the Program subject to these terms and
-conditions. You may not impose any further restrictions on the recipients'
-exercise of the rights granted herein. You are not responsible for
-enforcing compliance by third parties to this License.
-\item If, as a consequence of a court judgment or allegation of patent infringement
-or for any other reason (not limited to patent issues), conditions
-are imposed on you (whether by court order, agreement or otherwise)
-that contradict the conditions of this License, they do not excuse
-you from the conditions of this License. If you cannot distribute
-so as to satisfy simultaneously your obligations under this License
-and any other pertinent obligations, then as a consequence you may
-not distribute the Program at all. For example, if a patent license
-would not permit royalty-free redistribution of the Program by all
-those who receive copies directly or indirectly through you, then
-the only way you could satisfy both it and this License would be to
-refrain entirely from distribution of the Program.
-
-If any portion of this section is held invalid or unenforceable under
-any particular circumstance, the balance of the section is intended
-to apply and the section as a whole is intended to apply in other
-circumstances.
-
-It is not the purpose of this section to induce you to infringe any
-patents or other property right claims or to contest validity of any
-such claims; this section has the sole purpose of protecting the integrity
-of the free software distribution system, which is implemented by
-public license practices. Many people have made generous contributions
-to the wide range of software distributed through that system in reliance
-on consistent application of that system; it is up to the author/donor
-to decide if he or she is willing to distribute software through any
-other system and a licensee cannot impose that choice.
-
-This section is intended to make thoroughly clear what is believed
-to be a consequence of the rest of this License.
-
-\item If the distribution and/or use of the Program is restricted in certain
-countries either by patents or by copyrighted interfaces, the original
-copyright holder who places the Program under this License may add
-an explicit geographical distribution limitation excluding those countries,
-so that distribution is permitted only in or among countries not thus
-excluded. In such case, this License incorporates the limitation as
-if written in the body of this License.
-\item The Free Software Foundation may publish revised and/or new versions
-of the General Public License from time to time. Such new versions
-will be similar in spirit to the present version, but may differ in
-detail to address new problems or concerns.
-
-Each version is given a distinguishing version number. If the Program
-specifies a version number of this License which applies to it and
-``any later version,'' you have the option of following the terms
-and conditions either of that version or of any later version published
-by the Free Software Foundation. If the Program does not specify a
-version number of this License, you may choose any version ever published
-by the Free Software Foundation.
-
-\item If you wish to incorporate parts of the Program into other free programs
-whose distribution conditions are different, write to the author to
-ask for permission. For software which is copyrighted by the Free
-Software Foundation, write to the Free Software Foundation; we sometimes
-make exceptions for this. Our decision will be guided by the two goals
-of preserving the free status of all derivatives of our free software
-and of promoting the sharing and reuse of software generally.
-\end{enumerate}
-
-\subsection*{NO WARRANTY }
-
-\begin{itemize}
-
-\item[11.]BECAUSE THE PROGRAM IS LICENSED FREE OF CHARGE, THERE IS
-NO WARRANTY FOR THE PROGRAM, TO THE EXTENT PERMITTED BY APPLICABLE
-LAW. EXCEPT WHEN OTHERWISE STATED IN WRITING THE COPYRIGHT HOLDERS
-AND/OR OTHER PARTIES PROVIDE THE PROGRAM ``AS IS'' WITHOUT WARRANTY
-OF ANY KIND, EITHER EXPRESSED OR IMPLIED, INCLUDING, BUT NOT LIMITED
-TO, THE IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR
-PURPOSE. THE ENTIRE RISK AS TO THE QUALITY AND PERFORMANCE OF THE
-PROGRAM IS WITH YOU. SHOULD THE PROGRAM PROVE DEFECTIVE, YOU ASSUME
-THE COST OF ALL NECESSARY SERVICING, REPAIR OR CORRECTION.
-
-\item[12.]IN NO EVENT UNLESS REQUIRED BY APPLICABLE LAW OR AGREED
-TO IN WRITING WILL ANY COPYRIGHT HOLDER, OR ANY OTHER PARTY WHO MAY
-MODIFY AND/OR REDISTRIBUTE THE PROGRAM AS PERMITTED ABOVE, BE LIABLE
-TO YOU FOR DAMAGES, INCLUDING ANY GENERAL, SPECIAL, INCIDENTAL OR
-CONSEQUENTIAL DAMAGES ARISING OUT OF THE USE OR INABILITY TO USE THE
-PROGRAM (INCLUDING BUT NOT LIMITED TO LOSS OF DATA OR DATA BEING RENDERED
-INACCURATE OR LOSSES SUSTAINED BY YOU OR THIRD PARTIES OR A FAILURE
-OF THE PROGRAM TO OPERATE WITH ANY OTHER PROGRAMS), EVEN IF SUCH HOLDER
-OR OTHER PARTY HAS BEEN ADVISED OF THE POSSIBILITY OF SUCH DAMAGES.
-
-\end{itemize}
-
-\section*{END OF TERMS AND CONDITIONS }
-
-\subsection*{How to Apply These Terms to Your New Programs}
-
-If you develop a new program, and you want it to be of the greatest
-possible use to the public, the best way to achieve this is to make
-it free software which everyone can redistribute and change under
-these terms.
-
-To do so, attach the following notices to the program. It is safest
-to attach them to the start of each source file to most effectively
-convey the exclusion of warranty; and each file should have at least
-the ``copyright'' line and a pointer to where the full notice is found.
-For example:
-
-\begin{quote}
-One line to give the program's name and a brief idea of what it does.
-Copyright {\footnotesize $\copyright$ (}year) (name of author)
-
-This program is free software; you can redistribute it and/or modify
-it under the terms of the GNU General Public License as published
-by the Free Software Foundation; either version 2 of the License,
-or (at your option) any later version.
-
-This program is distributed in the hope that it will be useful, but
-WITHOUT ANY WARRANTY; without even the implied warranty of MERCHANTABILITY
-or FITNESS FOR A PARTICULAR PURPOSE. See the GNU General Public License
-for more details.
-
-You should have received a copy of the GNU General Public License
-along with this program; if not, write to the Free Software Foundation,
-Inc., 59 Temple Place, Suite 330, Boston, MA 02111-1307 USA
-\end{quote}
-Also add information on how to contact you by electronic and paper
-mail.
-
-If the program is interactive, make it output a short notice like
-this when it starts in an interactive mode:
-
-\begin{quote}
-Gnomovision version 69, Copyright $\copyright$ year name of author Gnomovision
-comes with ABSOLUTELY NO WARRANTY; for details type `show w'. This
-is free software, and you are welcome to redistribute it under certain
-conditions; type `show c' for details.
-\end{quote}
-The hypothetical commands `show w' and `show c' should show the appropriate
-parts of the General Public License. Of course, the commands you use
-may be called something other than `show w' and `show c'; they could
-even be mouse-clicks or menu items -- whatever suits your program.
-
-You should also get your employer (if you work as a programmer) or
-your school, if any, to sign a ``copyright disclaimer'' for the program,
-if necessary. Here is a sample; alter the names:
-
-\begin{quote}
-Yoyodyne, Inc., hereby disclaims all copyright interest in the program
-`Gnomovision' (which makes passes at compilers) written by James Hacker.
-
-(signature of Ty Coon)\\
-1 April 1989 \\
-Ty Coon, President of Vice
-\end{quote}
-This General Public License does not permit incorporating your program
-into proprietary programs. If your program is a subroutine library,
-you may consider it more useful to permit linking proprietary applications
-with the library. If this is what you want to do, use the GNU Library
-General Public License instead of this License.
-\end{document}
+Dimitri Komatitsch, IPG Paris, France, December 1996: first 3-D solver
+for the CM-5 Connection Machine, parallelized on 128 processors using
+Connection Machine Fortran.\\

### 2015-05-20 doc/USER_MANUAL/02_getting_started.tex
updates manual using verbatim for better display in markdown

@@
 To download the SPECFEM3D\_Cartesian software package, type this:
-\begin{lyxcode}
+\begin{verbatim}
 git clone --recursive --branch devel https://github.com/geodynamics/specfem3d.git
-\end{lyxcode}
-{\small \par}
-
+\end{verbatim}
+%
 We recommend that you add {\texttt{ulimit -S -s unlimited}} to your
@@
 in your .bash\_profile or your .cshrc file):
-\begin{lyxcode}
-./configure~FC=ifort~MPIFC=mpif90
-\end{lyxcode}
+\begin{verbatim}
+./configure FC=ifort MPIFC=mpif90
+\end{verbatim}
 Note that MPI must be installed with MPI-IO enabled because parts of SPECFEM3D perform I/Os through MPI-IO.
@@
 with the option \texttt{-{}-with-scotch-dir}. Just as an example:
-\begin{lyxcode}
-./configure~FC=ifort~MPIFC=mpif90~-{}-with-scotch-dir=/opt/scotch
-\end{lyxcode}
+\begin{verbatim}
+./configure FC=ifort MPIFC=mpif90 --with-scotch-dir=/opt/scotch
+\end{verbatim}
 If you use the Intel ifort compiler to compile the code, we recommend
 that you use the Intel icc C compiler to compile Scotch, i.e., use:
-\begin{lyxcode}
-./configure~CC=icc~FC=ifort~MPIFC=mpif90
-\end{lyxcode}
+\begin{verbatim}
+./configure CC=icc FC=ifort MPIFC=mpif90
+\end{verbatim}
 

### 2015-05-20 doc/USER_MANUAL/03_mesh_generation.tex
updates manual using verbatim for better display in markdown

@@
 The format is:
-
-\begin{lyxcode}
-element\_ID~material\_ID
-\end{lyxcode}
+\begin{verbatim}
+element_ID material_ID
+\end{verbatim}
 
@@
 
-\begin{lyxcode}
-domain\_ID~material\_ID~rho~vp~vs~Qkappa~Qmu~anisotropy\_flag
-\end{lyxcode}
+\begin{verbatim}
+domain_ID material_ID rho vp vs Qkappa Qmu anisotropy_flag
+\end{verbatim}
 
@@
 called "nummaterial\_velocity\_file" that will have just been created and change it from the prototype created:
-
-    0 1 vol1 --> syntax: \#material\_domain\_id \#material\_id \#rho \#vp \#vs \#Q\_kappa \#Q\_mu \#anisotropy
-    0 2 vol2 --> syntax: \#material\_domain\_id \#material\_id \#rho \#vp \#vs \#Q\_kappa \#Q\_mu \#anisotropy
-
+{\small
+\begin{verbatim}
+    0 1 vol1 --> syntax: #material_domain_id #material_id #rho #vp #vs #Q_kappa #Q_mu #anisotropy
+    0 2 vol2 --> syntax: #material_domain_id #material_id #rho #vp #vs #Q_kappa #Q_mu #anisotropy
+\end{verbatim}
+}
 (where "vol1" and "vol2" here represent the volume labels that you have set while creating the mesh in CUBIT)
 to for instance
-
+{\small
+\begin{verbatim}
     2 1 1500 2300 1800 9999.0 9999.0 0
     2 2 1600 2500 20000 9999.0 9999.0 0
+\end{verbatim}
+}
 
@@
 Then, in the main directory, type
-\begin{lyxcode}
-{\small make~xcheck\_mesh\_quality\_CUBIT\_Abaqus}{\small \par}
-\end{lyxcode}
+{\small
+\begin{verbatim}
+make xcheck_mesh_quality_CUBIT_Abaqus
+\end{verbatim}
+}
 and use
-\begin{lyxcode}
-{\small ./bin/xcheck\_mesh\_quality\_CUBIT\_Abaqus}{\small \par}
-\end{lyxcode}
+{\small
+\begin{verbatim}
+./bin/xcheck_mesh_quality_CUBIT_Abaqus
+\end{verbatim}
+}
 to generate an %not supported: AVS output file (\texttt{\small AVS\_meshquality.inp} in AVS UCD format) or
@@
 When you are ready to compile, in the main directory type
-\begin{lyxcode}
-{\small make~xdecompose\_mesh}{\small \par}
-\end{lyxcode}
+{\small
+\begin{verbatim}
+make xdecompose_mesh
+\end{verbatim}
+}
 If all paths and flags have been set correctly, the executable \texttt{bin/xdecompose\_mesh}
@@
 will provide a parallel version of that code). It needs to be run in the main directory because it expects the \texttt{./DATA/Par\_file}. The synopsis is:
-\begin{lyxcode}
-./bin/xdecompose\_mesh~nparts~input\_directory~output\_directory
-\end{lyxcode}
+\begin{verbatim}
+./bin/xdecompose_mesh nparts input_directory output_directory
+\end{verbatim}
 where
@@
 In the main directory, type
-\begin{lyxcode}
-{\small make~xmeshfem3D}{\small \par}
-\end{lyxcode}
+{\small
+\begin{verbatim}
+make xmeshfem3D
+\end{verbatim}
+}
 If all paths and flags have been set correctly, the mesher should
@@
 edit file \texttt{src/shared/utm\_geo.f90}, uncomment that ellipsoid and recompile the code.
-From http://en.wikipedia.org/wiki/Universal\_Transverse\_Mercator\_coordinate\_system :
+From \url{http://en.wikipedia.org/wiki/Universal_Transverse_Mercator_coordinate_system}:\\
 The Universal Transverse Mercator coordinate system was developed by the United States Army Corps of Engineers in the 1940s.
@@
 For different geographic regions, other datum systems (e.g.: ED50, NAD83) can be used.
+%
 \item [{\texttt{SUPPRESS\_UTM\_PROJECTION}}] set to be \texttt{.false.}
@@
 
-\begin{lyxcode}
-material\_ID~rho~vp~vs~Q~anisotropy\_flag~domain\_ID
-\end{lyxcode}
+\begin{verbatim}
+material_ID rho vp vs Q anisotropy_flag domain_ID
+\end{verbatim}
 
@@
 
-\begin{lyxcode}
-NEX\_XI\_BEGIN~NEX\_XI\_END~NEX\_ETA\_BEGIN~NEX\_ETA\_END~NZ\_BEGIN~NZ\_END~material\_ID
-\end{lyxcode}
+\begin{verbatim}
+NEX_XI_BEGIN NEX_XI_END NEX_ETA_BEGIN NEX_ETA_END NZ_BEGIN NZ_END material_ID
+\end{verbatim}
 \end{description}
@@
 in \texttt{constants.h}:
-\begin{lyxcode}
-!~uncomment~this~to~write~messages~to~the~screen~
-
-!~integer,~parameter~::~IMAIN~=~ISTANDARD\_OUTPUT
-\end{lyxcode}
+\begin{verbatim}
+! uncomment this to write messages to the screen
+! integer, parameter :: IMAIN = ISTANDARD_OUTPUT
+\end{verbatim}
 To control the quality of the mesh, check the standard output (either

### 2015-05-20 doc/USER_MANUAL/04_creating_databases.tex
updates manual using verbatim for better display in markdown

@@
 In the main directory, type
-\begin{lyxcode}
-{\small make~xgenerate\_databases}{\small \par}
-\end{lyxcode}
+{\small
+\begin{verbatim}
+make xgenerate_databases
+\end{verbatim}
+}
 Input for the program is provided through the main parameter file
@@
 2/ submit a single job on 1000 cores to the batch, and in that script create a sub-array of jobs to start 10 jobs,
-each running on 100 cores (see e.g. http://www.schedmd.com/slurmdocs/job\_array.html )
+each running on 100 cores (see e.g. \url{http://www.schedmd.com/slurmdocs/job_array.html})
 
@@
 The way we prescribe material property for porous material in SPECFEM3D is as follow:
-We use a file name nummaterial\_poroelastic\_file, which is located in the directory MESH, the format is as follow:\\
- rhos rhof phi c kxx kxy kxz kyy kyz kzz Ks Kf Kfr etaf mufr\\
+We use a file name nummaterial\_poroelastic\_file, which is located in the directory MESH, the format is as follow:
+\begin{verbatim}
+ rhos rhof phi c kxx kxy kxz kyy kyz kzz Ks Kf Kfr etaf mufr
+\end{verbatim} 
 where\\

### 2015-05-20 doc/USER_MANUAL/05_running_the_solver.tex
updates manual using verbatim for better display in markdown

@@
 to compile the solver. In the main directory, type
-\begin{lyxcode}
-{\small make~xspecfem3D}{\small \par}
-\end{lyxcode}
+{\small
+\begin{verbatim}
+make xspecfem3D
+\end{verbatim}
+}
 Please note that \texttt{xspecfem3D} must be called directly from
@@
 It looks like the example shown in Fig. \ref{fig:CMTSOLUTION-file}.
-%\begin{lyxcode}
 \begin{figure}[htp]
@@
 \end{figure}
-%\end{lyxcode}
+%
 The \texttt{CMTSOLUTION} file should be edited in the following way:
@@
 \urlwithparentheses{www.llnl.gov/sac}. Type
-
-\begin{lyxcode}
-{\small make~xconvolve\_source\_timefunction}{\small \par}
-\end{lyxcode}
-
+{\small
+\begin{verbatim}
+make xconvolve_source_timefunction
+\end{verbatim}
+}
 to compile the code and then set the parameter \texttt{hdur} in \texttt{convolve\_source\_timefunction.csh}
@@
 To convert to absolute time $t_{\mathrm{abs}}$, set
-
+%
 \begin{lyxcode}
@@
 Each line represents one station in the following format:
-\begin{lyxcode}
-{\small Station~Network~Latitude~(degrees)~Longitude~(degrees)~Elevation~(m)~burial~(m)~}{\small \par}
-\end{lyxcode}
+{\small
+\begin{verbatim}
+Station Network Latitude(degrees) Longitude(degrees) Elevation(m) burial(m)
+\end{verbatim}
+}
 The solver \texttt{xspecfem3D} filters the list of stations in file
@@
 the screen instead by uncommenting a line in \texttt{constants.h}:
-\begin{lyxcode}
-!~uncomment~this~to~write~messages~to~the~screen~
-
-!~integer,~parameter~::~IMAIN~=~ISTANDARD\_OUTPUT~
-\end{lyxcode}
+\begin{verbatim}
+! uncomment this to write messages to the screen
+! integer, parameter :: IMAIN = ISTANDARD_OUTPUT
+\end{verbatim}
 On PC clusters the seismogram files are generally written to the local
@@
 While the solver is running, its progress may be tracked by monitoring
-the `\texttt{\small timestamp{*}}{\small ' files in the ~}\\
-{\small{} }\texttt{\small OUTPUT\_FILES}{\small{} directory. These tiny
-files look something like this:}{\small \par}
-\begin{lyxcode}
-Time~step~\#~~~~~~~~~~10000~
-
-Time:~~~~~108.4890~~~~~~seconds~
-
-Elapsed~time~in~seconds~=~~~~1153.28696703911~
-
-Elapsed~time~in~hh:mm:ss~=~~~~~0~h~19~m~13~s~
-
-Mean~elapsed~time~per~time~step~in~seconds~=~~~~~0.115328696703911~
-
-Max~norm~displacement~vector~U~in~all~slices~(m)~=~~~~1.0789589E-02~
-\end{lyxcode}
-The \texttt{\small timestamp{*}}{\small{} files provide the }\texttt{\small Mean
-elapsed time per time step in seconds}{\small , which may be used
+the `\texttt{\small timestamp{*}}' files in the \texttt{\small OUTPUT\_FILES/} directory. These tiny
+files look something like this:
+{\small
+\begin{verbatim}
+Time step #          10000
+Time:     108.4890      seconds
+Elapsed time in seconds =    1153.28696703911
+Elapsed time in hh:mm:ss =     0 h 19 m 13 s
+Mean elapsed time per time step in seconds =     0.115328696703911
+Max norm displacement vector U in all slices (m) =     1.0789589E-02
+\end{verbatim}
+}
+The \texttt{\small timestamp{*}} files provide the 
+\texttt{\small Mean elapsed time per time step in seconds}, which may be used
 to assess performance on various machines (assuming you are the only
-user on a node), as well as the }\texttt{\small Max norm displacement
-vector U in all slices~(m)}{\small . If something is wrong with the
+user on a node), as well as the 
+\texttt{\small Max norm displacement vector U in all slices~(m)}. 
+If something is wrong with the
 model, the mesh, or the source, you will see the code become unstable
@@
 the program. You can control the rate at which the timestamp files
-are written based upon the parameter }\texttt{\small NTSTEP\_BETWEEN\_OUTPUT\_INFO}{\small{}
-in the }\texttt{\small Par\_file}{\small .}{\small \par}
+are written based upon the parameter 
+\texttt{\small NTSTEP\_BETWEEN\_OUTPUT\_INFO}
+in the \texttt{\small Par\_file}.
 

### 2015-05-20 doc/USER_MANUAL/11_graphics.tex
Buildbot ran the code beautifier.
updates manual using verbatim for better display in markdown

@@
 these movie data files (see section \ref{cha:Main-Parameter} for
-details): \texttt{\small NTSTEP\_BETWEEN\_FRAMES}{\small{} to set the
-timesteps between frames, }\texttt{\small SAVE\_DISPLACEMENT}{\small{}
-to save displacement instead of velocity, }\texttt{\small USE\_HIGHRES\_FOR\_MOVIES}{\small{}
+details): 
+\texttt{\small NTSTEP\_BETWEEN\_FRAMES} to set the
+timesteps between frames, 
+\texttt{\small SAVE\_DISPLACEMENT}
+to save displacement instead of velocity,
+\texttt{\small USE\_HIGHRES\_FOR\_MOVIES}
 to save values at all GLL point instead of element edges. In order
-to output additionally shakemaps, you would set the parameter }\texttt{\small CREATE\_SHAKEMAP}{\small{}
-to }\texttt{\small .true.}{\small .}{\small \par}
+to output additionally shakemaps, you would set the parameter
+\texttt{\small CREATE\_SHAKEMAP}
+to \texttt{\small .true.}.
 
@@
 \begin{description}
-\item [{\texttt{xcreate\_movie\_shakemap\_AVS\_DX\_GMT}}] From \texttt{create\_movie\_shakemap\_AVS\_DX\_GMT.f90},
+\item [{\texttt{xcreate\_movie\_shakemap\_AVS\_DX\_GMT}}] 
+From \texttt{create\_movie\_shakemap\_AVS\_DX\_GMT.f90},
 it outputs data in ASCII, OpenDX, or AVS format (also readable in
@@
 
-\begin{lyxcode}
-{\footnotesize make~xcreate\_movie\_shakemap\_AVS\_DX\_GMT}{\footnotesize \par}
-\end{lyxcode}
-
+{\footnotesize
+\begin{verbatim}
+make xcreate_movie_shakemap_AVS_DX_GMT
+\end{verbatim}
+}
 and run the executable \texttt{xcreate\_movie\_shakemap\_AVS\_DX\_GMT}
@@
 
-When running xspecfem3D with the \texttt{\small MOVIE\_VOLUME}{\small{}
-flag turned on, the code outputs several files in }\texttt{\small LOCAL\_PATH}{\small{}
-specified in the main }\texttt{\small Par\_file}{\small , e.g. in
-directory }\texttt{\small OUTPUT\_FILES/DATABASES\_MPI}{\small .
+When running xspecfem3D with the 
+\texttt{\small MOVIE\_VOLUME}
+flag turned on, the code outputs several files in \texttt{\small LOCAL\_PATH}
+specified in the main \texttt{\small Par\_file}, 
+e.g. in directory \texttt{\small OUTPUT\_FILES/DATABASES\_MPI}.
 The output is saved by each processor at the time interval specified
-by }\texttt{\small NTSTEP\_BETWEEN\_FRAMES}{\small . For all domains,
-the velocity field is output to files: }\texttt{\small proc??????\_velocity\_X\_it??????.bin}{\small ,
-}\texttt{\small proc??????\_velocity\_Y\_it??????.bin}{\small{} and
-}\texttt{\small proc??????\_velocity\_Z\_it??????.bin}{\small . For
-elastic domains, the divergence and curl taken from the velocity field,
+by 
+\texttt{\small NTSTEP\_BETWEEN\_FRAMES}. 
+For all domains,
+the velocity field is output to files: 
+{\small
+\begin{verbatim}
+proc??????_velocity_X_it??????.bin
+proc??????_velocity_Y_it??????.bin
+proc??????_velocity_Z_it??????.bin 
+\end{verbatim}
+}
+For elastic domains, the divergence and curl taken from the velocity field,
 i.e. $\nabla\cdot{\bf {v}}$ and $\nabla\times{\bf {v}}$, get stored
-as well: }\texttt{\small proc??????\_div\_it??????.bin}{\small , }\texttt{\small proc??????\_curl\_X\_t??????.bin}{\small ,
-}\texttt{\small proc??????\_curl\_Y\_it??????.bin}{\small{} and }\texttt{\small proc??????\_curl\_Z\_it??????.bin}{\small .
-The files denoted }\texttt{\small proc??????\_div\_}{\small ~}\\
-{\small{} }\texttt{\small glob\_it??????.bin}{\small{} and }\texttt{\small proc??????\_curl\_glob\_it??????.bin}{\small{}
+as well: 
+{\small
+\begin{verbatim}
+proc??????_div_it??????.bin
+proc??????_curl_X_t??????.bin
+proc??????_curl_Y_it??????.bin
+proc??????_curl_Z_it??????.bin
+\end{verbatim}
+}
+The files denoted \texttt{\small proc??????\_div\_glob\_it??????.bin} and 
+\texttt{\small proc??????\_curl\_glob\_it??????.bin}
 are stored on the global points only, all the other arrays are stored
 on all GLL points. Note that the components X/Y/Z can change to E/N/Z
-according to the }\texttt{\small SUPPRESS\_UTM\_PROJECTION}{\small{}
-flag (see also Appendix~\ref{cha:Coordinates} and \ref{cha:channel-codes}).}{\small \par}
+according to the 
+\texttt{\small SUPPRESS\_UTM\_PROJECTION}
+flag (see also Appendix~\ref{cha:Coordinates} and \ref{cha:channel-codes}).
 
@@
 it in the root directory, type:
-\begin{lyxcode}
-{\footnotesize make~xcombine\_vol\_data~}{\footnotesize \par}
-\end{lyxcode}
+
+{\footnotesize
+\begin{verbatim}
+make xcombine_vol_data
+\end{verbatim}
+}
 which will create the executable \texttt{bin/xcombine\_vol\_data}.
 To output the usage of this executable, type
-\begin{lyxcode}
-{\footnotesize ./bin/xcombine\_vol\_data~}{\footnotesize \par}
-\end{lyxcode}
+{\footnotesize
+\begin{verbatim}
+./bin/xcombine_vol_data
+\end{verbatim}
+}
 without arguments.
@@
 if you want to combine these files :
-
-\medskip
-
-\texttt{proc000000\_velocity\_Z\_it000400.bin }
-
-\texttt{proc000001\_velocity\_Z\_it000400.bin }
-
-\texttt{proc000002\_velocity\_Z\_it000400.bin }
-
-\texttt{proc000003\_velocity\_Z\_it000400.bin}
-
-\medskip
+{\small
+\begin{verbatim}
+proc000000_velocity_Z_it000400.bin
+proc000001_velocity_Z_it000400.bin
+proc000002_velocity_Z_it000400.bin
+proc000003_velocity_Z_it000400.bin
+\end{verbatim}
+}
 
@@
 
-\begin{lyxcode}
-~{\footnotesize ./bin/xcombine\_vol\_data~0~3~velocity\_Z\_it000400~./OUTPUT\_FILES/DATABASES\_MPI~./OUTPUT\_FILES~0~}{\footnotesize \par}
-\end{lyxcode}
-
+{\footnotesize
+\begin{verbatim}
+./bin/xcombine_vol_data 0 3 velocity_Z_it000400 ./OUTPUT_FILES/DATABASES_MPI ./OUTPUT_FILES 0
+\end{verbatim}
+}
 Here, $0$ is the number of the first processor, $3$ the number of the last one, \texttt{velocity\_Z\_it000400} the name of the files we want to combine without the prefix \texttt{proc000000*\_}, \texttt{./OUTPUT\_FILES/DATABASES\_MPI} the directory where the files are located, \texttt{./OUTPUT\_FILES} the directory where the combined file will be stored, and $0$ is the parameter to create a low-resolution mesh file ($1$ for a high-resolution).
@@
 If the value of this parameter is \texttt{.false.}, the output mesh file will have the name \texttt{velocity\_Z\_it000400.mesh}. Then we next have to convert the
-\texttt{.mesh} file into the VTU (Unstructured grid file) format which can be viewed in ParaView. For this task, you can use and modify the
-script \texttt{mesh2vtu.pl} located in directory \texttt{utils/Visualization/Paraview/}, for example:
-\begin{lyxcode}
-{\footnotesize mesh2vtu.pl~-i~velocity\_Z\_it000400.mesh~-o~velocity\_Z\_it000400.vtu}{\footnotesize \par}
-\end{lyxcode}
+\texttt{.mesh} file into the VTU (Unstructured grid file) format which can be viewed in ParaView.
+
+\noindent
+For this task, you can use and modify the
+script \texttt{mesh2vtu.pl} located in directory \texttt{\small utils/Visualization/Paraview/}, for example:
+
+{\footnotesize
+\begin{verbatim}
+mesh2vtu.pl -i velocity_Z_it000400.mesh -o velocity_Z_it000400.vtu
+\end{verbatim}
+}
+
 Notice that this Perl script uses a program \texttt{mesh2vtu} in the \texttt{utils/Visualization/Paraview/mesh2vtu} directory, which further
@@
 
-\begin{lyxcode}
-{\small slice\_number.pl~Mesh\_Par\_file~output\_solver.txt~slice\_file}{\small \par}
-\end{lyxcode}
-
+{\small
+\begin{verbatim}
+slice_number.pl Mesh_Par_file output_solver.txt slice_file
+\end{verbatim}
+}
 which will generate a \texttt{slices\_file}.
@@
 
-\begin{lyxcode}
-{\small utils/copy\_database.pl~slice\_file~lsf\_machine\_file~filename~{[}jobid{]}}{\small \par}
-\end{lyxcode}
-
+{\small
+\begin{verbatim}
+utils/copy_database.pl slice_file lsf_machine_file filename [jobid]
+\end{verbatim}
+}
 where \texttt{\small lsf\_machine\_file}{\small{} is the machine file
@@
 \item Compile it in the root directory:
-
-\begin{lyxcode}
-{\footnotesize make~xcombine\_vol\_data~}{\footnotesize \par}
-
-{\footnotesize ./bin/xcombine\_vol\_data~slice\_list~filename~input\_dir~output\_dir~high/low-resolution~}{\footnotesize \par}
-\end{lyxcode}
-
+{\footnotesize
+\begin{verbatim}
+make xcombine_vol_data
+./bin/xcombine_vol_data slice_list filename input_dir output_dir high/low-resolution
+\end{verbatim}
+}
 where \texttt{input\_dir} is the directory where all the individual
@@
 \end{enumerate}
+
 \item \textbf{Convert mesh files into .vtu files}
@@
 
-\begin{lyxcode}
-{\footnotesize mesh2vtu.pl~-i~file.mesh~-o~file.vtu~}{\footnotesize \par}
-\end{lyxcode}
+{\footnotesize
+\begin{verbatim}
+mesh2vtu.pl -i file.mesh -o file.vtu
+\end{verbatim}
+}
+
 \item Notice that this Perl script uses a program \texttt{mesh2vtu} in the
-\texttt{utils/Visualization/Paraview/mesh2vtu} directory, which further
+\texttt{\small utils/Visualization/Paraview/mesh2vtu} directory, which further
 uses the VTK \urlwithparentheses{http://www.vtk.org/} run-time library
@@
 \end{enumerate}
+
 \item \textbf{Copy over the source and receiver .vtk file}

### 2015-05-20 doc/USER_MANUAL/13_changing_the_model.tex
updates manual using verbatim for better display in markdown

@@
 when using CUBIT to construct your mesh (see also section \ref{subsec:Exporting-the-Mesh}):
-\begin{lyxcode}
-domain\_ID~material\_ID~tomography~elastic~file\_name~1
-\end{lyxcode}
+\begin{verbatim}
+domain_ID material_ID tomography elastic file_name 1
+\end{verbatim}
 where: \\
@@
 %
-\begin{lyxcode}
-2  -1 tomography elastic tomography\_model.xyz 1
-\end{lyxcode}
+\begin{verbatim}
+2  -1 tomography elastic tomography_model.xyz 1
+\end{verbatim}
 %
@@
 %
-\begin{lyxcode}
-2  -1 tomography elastic file\_above\_15km.xyz 1 \\
-2  -2 tomography elastic file\_below\_15km.xyz 1
-\end{lyxcode}
+\begin{verbatim}
+2  -1 tomography elastic file_above_15km.xyz 1
+2  -2 tomography elastic file_below_15km.xyz 1
+\end{verbatim}
 %
@@
 this routine would be as follows
-\begin{lyxcode}
-call~model\_external\_values(xmesh,~ymesh,~zmesh,~~\&~~\\
-~~~~~~~~~~~rho,~vp,~vs,~qmu\_atten,~iflag\_aniso,~idomain\_id)
-\end{lyxcode}
+\begin{verbatim}
+call model_external_values(xmesh, ymesh, zmesh, &
+                           rho, vp, vs, qmu_atten, iflag_aniso, idomain_id)
+\end{verbatim}
 Input to this routine consists of:
@@
 The call to this subroutine is of the form
-\begin{lyxcode}
-call~model\_aniso(iflag\_aniso,~rho,~vp,~vs,~\&~~\\
-~~~~~~~~~~~c11,c12,c13,c14,c15,c16,c22,c23,c24,c25,c26,~\&~~\\
-~~~~~~~~~~~c33,c34,c35,c36,c44,c45,c46,c55,c56,c66)~
-\end{lyxcode}
+\begin{verbatim}
+call model_aniso(iflag_aniso, rho, vp, vs, &
+                 c11,c12,c13,c14,c15,c16,c22,c23,c24,c25,c26, &
+                 c33,c34,c35,c36,c44,c45,c46,c55,c56,c66)
+\end{verbatim}
 Input to this routine consists of:

### 2015-05-20 doc/USER_MANUAL/14_post_processing.tex
updates manual using verbatim for better display in markdown

@@
 without any argument. An example of the usage of the script:
-\begin{lyxcode}
-{\footnotesize process\_data.pl~-m~CMTSOLUTION~-s~1.0~-l~0/4000~-i~-f~-t~40/500~-p~-x~bp~DATA/1999.330{*}.BH?.SAC}~
-\end{lyxcode}
-\noindent which has resampled the SAC files to a sampling rate of
+
+{\footnotesize
+\begin{verbatim}
+process_data.pl -m CMTSOLUTION -s 1.0 -l 0/4000 -i -f -t 40/500 -p -x bp DATA/1999.330*.BH?.SAC
+\end{verbatim}
+}
+
+\noindent 
+which has resampled the SAC files to a sampling rate of
 1 seconds, cut them between 0 and 4000 seconds, transfered them into
@@
 An example of the usage of the script:
-\begin{lyxcode}
-{\footnotesize process\_syn.pl~-m~CMTSOLUTION~-h~-a~STATIONS~-s~1.0~-l~0/4000~-f~-t~40/500~-p~-x~bp~SEM/{*}.BX?.semd}~
-\end{lyxcode}
+
+{\footnotesize
+\begin{verbatim}
+process_syn.pl -m CMTSOLUTION -h -a STATIONS -s 1.0 -l 0/4000 -f -t 40/500 -p -x bp SEM/*.BX?.semd
+\end{verbatim}
+}
+\noindent
 which will convolve the synthetics with a triangular source-time function
@@
 and \texttt{\small rotate.pl}{\small{} can be used to accomplish this:}{\small \par}
-\begin{lyxcode}
-rotate.pl~-l~0~-L~180~-d~DATA/{*}.BHE.SAC.bp~
-
-rotate.pl~-l~0~-L~180~SEM/{*}.BXE.semd.sac.bp~
-\end{lyxcode}
+\begin{verbatim}
+rotate.pl -l 0 -L 180 -d DATA/*.BHE.SAC.bp
+rotate.pl -l 0 -L 180 SEM/*.BXE.semd.sac.bp
+\end{verbatim}
 where the first command performs rotation on the SAC data obtained
@@
 
-\begin{lyxcode}
-collect\_seismo.pl~machines~DATA/Par\_file
-\end{lyxcode}
+\begin{verbatim}
+collect_seismo.pl machines DATA/Par_file
+\end{verbatim}
+
 \end{description}
+
 where \texttt{machines} is a file containing the node names and \texttt{DATA/Par\_file}
@@
 is provided in \texttt{utils/}:
-\begin{lyxcode}
-cleanbase.pl~machines
-\end{lyxcode}
+\begin{verbatim}
+cleanbase.pl machines
+\end{verbatim}
 where \texttt{machines} is a file containing the node names.
@@
 in GMT, for example:}{\small \par}
-\begin{lyxcode}
-movie2gif.gmt.pl~-m~CMTSOLUTION~-g~-f~1/40~-n~-2~-p~
-\end{lyxcode}
+\begin{verbatim}
+movie2gif.gmt.pl -m CMTSOLUTION -g -f 1/40 -n -2 -p
+\end{verbatim}
 which for the first through the 40th movie frame, converts the \texttt{moviedata}
@@
 and then plot the shakemaps in GMT, for example:
-\begin{lyxcode}
-plot\_shakemap.gmt.pl~data\_dir~type(1,2,3)~CMTSOLUTION~
-\end{lyxcode}
+\begin{verbatim}
+plot_shakemap.gmt.pl data _dir type(1,2,3) CMTSOLUTION
+\end{verbatim}
 where \texttt{type=1} for a displacement shakemap, \texttt{2} for
@@
 types of solvers separately through a scheduler (refer to Chapter~\ref{cha:Scheduler}).
-\begin{lyxcode}
-run\_lsf.bash~-{}-gm-no-shmem~-{}-gm-copy-env~remap\_database~old\_machines~150
-\end{lyxcode}
+\begin{verbatim}
+run_lsf.bash --gm-no-shmem --gm-copy-env remap_database old_machines 150
+\end{verbatim}
 where \texttt{old\_machines} is the LSF machine file used in the previous

### 2015-05-21 doc/USER_MANUAL/06_fault_sources.tex
Buildbot ran the code beautifier.

@@
 for a buried vertical strike-slip fault:
-\begin{lyxcode}
-~reset~~\\
-~brick~x~10~y~10~z~10~~\\
-~webcut~volume~all~with~plane~xplane~~\\
-~webcut~volume~all~with~plane~yplane~~\\
-~webcut~volume~all~with~plane~xplane~offset~3~~\\
-~webcut~volume~all~with~plane~zplane~offset~3~~\\
-~webcut~volume~all~with~plane~zplane~offset~-3~~\\
-~imprint~all~~\\
-~merge~all~~\\
-~unmerge~surf~160~~\\
-~mesh~vol~all~~\\
-~set~node~constraint~off~~\\
-~node~in~surf~168~move~X~0~Y~0.01~Z~0~~\\
-~node~in~surf~160~move~X~0~Y~-0.01~Z~0~\\
-
-\end{lyxcode}
+{\footnotesize
+\begin{verbatim}
+ reset
+ brick x 10 y 10 z 10
+ webcut volume all with plane xplane
+ webcut volume all with plane yplane
+ webcut volume all with plane xplane offset 3
+ webcut volume all with plane zplane offset 3
+ webcut volume all with plane zplane offset -3
+ imprint all
+ merge all
+ unmerge surf 160
+ mesh vol all
+ set node constraint off
+ node in surf 168 move X 0 Y 0.01 Z 0
+ node in surf 160 move X 0 Y -0.01 Z 0
+\end{verbatim}
+}
+%
 \begin{figure}[htbp]
@@
 the edges of the fault surface are touching each other.}
-
 \label{fig:examples.splitnodes}
@@
 
-\begin{lyxcode}
-
-\end{lyxcode}
 \begin{figure}[htbp]
@@
 with the following CUBIT command:
-\begin{lyxcode}
-$\mathtt{compress\: ids\: hex\: face\: edge\: node}$
-\end{lyxcode}
+{\footnotesize
+\begin{verbatim}
+compress ids hex face edge node
+\end{verbatim}
+}
 (otherwise you will get a segmentation fault during domain decomposition)
@@
 fault id (id\_fault).
-\item [{Example:}] For the example in Section \ref{sec:Mesh-Generation-with}:\end{description}
-\begin{lyxcode}
-$\hphantom{}$A1~=~{[}168{]}
-
-$\hphantom{}$A2~=~{[}160{]}
-
-$\hphantom{}$fault\_input(1,A1,A2)
-\end{lyxcode}
+\item [{Example:}] For the example in Section \ref{sec:Mesh-Generation-with}:
+\begin{verbatim}
+  A1 = [168]
+  A2 = [160]
+  fault_input(1,A1,A2)
+\end{verbatim}
+\end{description}
 

### 2015-05-21 doc/USER_MANUAL/02_getting_started.tex
updates manual for better wiki display

@@
 %
+\noindent
 We recommend that you add {\texttt{ulimit -S -s unlimited}} to your
@@
 \end{verbatim}
+\noindent
 Note that MPI must be installed with MPI-IO enabled because parts of SPECFEM3D perform I/Os through MPI-IO.
@@
 compilers, we suggest running the \texttt{configure} script with the
-following options:\\
+following options:
 
-\noindent \texttt{./configure FC=xlf90\_r MPIFC=mpif90 CC=xlc\_r CFLAGS=\textquotedbl{}-O3
--q64\textquotedbl{} FCFLAGS=\textquotedbl{}-O3 -q64\textquotedbl{}
--with-scotch-dir=...}.\\
+{\footnotesize
+\begin{verbatim}
+./configure FC=xlf90_r MPIFC=mpif90 CC=xlc_r CFLAGS="-O3 -q64" FCFLAGS="-O3 -q64" -with-scotch-dir=...
+\end{verbatim}
+}
 
@@
 \end{verbatim}
+\noindent
 If you use the Intel ifort compiler to compile the code, we recommend
@@
 in file \texttt{src/specfem3D/Makefile.in} (before running \texttt{configure})
-and also uncomment the \texttt{\#define USE\_OPENMP} statement in
-file\\
+and also uncomment the 
+\texttt{\#define USE\_OPENMP} statement in file\\ 
 \texttt{src/specfem3D/specfem3D.F90}.
 
-The DO-loop using OpenMP threads has a SCHEDULE property. The OMP\_SCHEDULE
+The DO-loop using OpenMP threads has a SCHEDULE property. The \texttt{OMP\_SCHEDULE}
 environment variable can set the scheduling policy of that DO-loop.
@@
 being twice as the number of OpenMP threads (thus chunk size = 8 for
-4 OpenMP threads etc). If OMP\_SCHEDULE is not set or is empty, the
+4 OpenMP threads etc). If \texttt{OMP\_SCHEDULE} is not set or is empty, the
 DO-loop will assume generic scheduling policy, which will slow down
@@
 
-\noindent Edit file \texttt{flags.guess} and put this for \texttt{FLAGS\_CHECK}:\\
-
-\noindent \texttt{-g -qfullpath -O2 -qsave -qstrict -qtune=qp -qarch=qp
--qcache=auto -qhalt=w}\\
- \texttt{-qfree=f90 -qsuffix=f=f90 -qlanglvl=95pure -Q -Q+rank,swap\_all
--Wl,-relax}\\
+\noindent Edit file \texttt{flags.guess} and put this for \texttt{FLAGS\_CHECK}:
+\begin{verbatim}
+-g -qfullpath -O2 -qsave -qstrict -qtune=qp -qarch=qp -qcache=auto -qhalt=w
+-qfree=f90 -qsuffix=f=f90 -qlanglvl=95pure -Q -Q+rank,swap_all -Wl,-relax
+\end{verbatim}
 
@@
 
-\noindent Then, to configure the code, type this:\\
-
-\noindent \texttt{./configure FC=bgxlf90\_r MPIFC=mpixlf90\_r CC=bgxlc\_r
-LOCAL\_PATH\_IS\_ALSO\_GLOBAL=true}\\
+\noindent Then, to configure the code, type this:
+\begin{verbatim}
+./configure FC=bgxlf90_r MPIFC=mpixlf90_r CC=bgxlc_r LOCAL_PATH_IS_ALSO_GLOBAL=true
+\end{verbatim}
 
@@
 IDRIS, France, suggests the following: compile the code with
+\begin{verbatim}
+FLAGS_CHECK="-O3 -qsave -qstrict -qtune=auto -qarch=450d -qcache=auto \
+  -qfree=f90 -qsuffix=f=f90 -g -qlanglvl=95pure -qhalt=w -Q \
+  -Q+rank,swap_all -Wl,-relax"
+\end{verbatim}
 
-\noindent \texttt{ FLAGS\_CHECK=\textquotedbl{}-O3 -qsave -qstrict
--qtune=auto -qarch=450d -qcache=auto }~\\
- \texttt{ -qfree=f90 -qsuffix=f=f90 -g -qlanglvl=95pure -qhalt=w -Q
--Q+rank,swap\_all -Wl,-relax\textquotedbl{}}
-
-\noindent Option \textquotedbl{}-Wl,-relax\textquotedbl{} must be
+\noindent 
+Option \textquotedbl{}-Wl,-relax\textquotedbl{} must be
 added on many (but not all) BlueGene systems to be able to link the
@@
 needs to pass the right commands to the \texttt{configure} script:
-
-\noindent \texttt{ ./configure -{}-prefix=/path/to/SPECFEM3DG\_SP
--{}-host=Babel -{}-build=BGP }~\\
- \texttt{ FC=bgxlf90\_r MPIFC=mpixlf90\_r CC=bgxlc\_r AR=bgar ARFLAGS=cru
-}~\\
- \texttt{ RANLIB=bgranlib LOCAL\_PATH\_IS\_ALSO\_GLOBAL=false}
+\begin{verbatim}
+./configure --prefix=/path/to/SPECFEM3DG_SP --host=Babel --build=BGP \
+  FC=bgxlf90_r MPIFC=mpixlf90_r CC=bgxlc_r AR=bgar ARFLAGS=cru \
+  RANLIB=bgranlib LOCAL_PATH_IS_ALSO_GLOBAL=false
+\end{verbatim}
 
@@
 
-\noindent \texttt{bgrun -np 1 -mode VN -exe ./bin/xcreate\_header\_file}
+\noindent 
+\begin{verbatim}
+bgrun -np 1 -mode VN -exe ./bin/xcreate_header_file
+\end{verbatim}
 
@@
 \item [{\texttt{RECURSIVE}}] This tag can be used to turn specify whether or not subdirectories should be searched for input files as well. In the case of SPECFEM, set to YES.
-\item [{\texttt{EXCLUDE}}] Here, you can exclude:  \\
-\texttt{"../../src/specfem3D/older\_not\_maintained\_partial\_OpenMP\_port"}\\
-\texttt{"../../src/decompose\_mesh/scotch"}\\
-\texttt{"../../src/decompose\_mesh/scotch\_5.1.12b"}
+\item [{\texttt{EXCLUDE}}] Here, you can exclude:
+\begin{verbatim}
+../../src/specfem3D/older_not_maintained_partial_OpenMP_port
+../../src/decompose_mesh/scotch
+../../src/decompose_mesh/scotch_5.1.12b
+\end{verbatim}
 \item [{\texttt{DOT\_GRAPH\_MAX\_NODES}}] to set the maximum number of nodes that will be shown in the graph.

### 2015-05-21 doc/USER_MANUAL/03_mesh_generation.tex
updates manual for better wiki display

@@
 \begin{figure}[htbp]
-\noindent \begin{centering}
+\begin{centering}
 \includegraphics[width=0.6\textwidth]{figures/mount-cubit.jpg}
-\par\end{centering}
-
+\par
+\end{centering}
 \caption{Example of the graphical user interface of CUBIT. The hexahedral mesh
@@
 of a single volume with topography.}
-
 \label{fig:mount.cubit}
@@
 \end{description}
+%
 \begin{figure}[htbp]
-\noindent \begin{centering}
+\begin{centering}
 \includegraphics[width=0.45\textwidth]{figures/example-homogeneous.jpg}
@@
 \includegraphics[width=0.45\textwidth]{figures/example-tomo.jpg}
-\par\end{centering}
-
+\par
+\end{centering}
 \caption{Screenshots of the CUBIT examples provided in subdirectory \texttt{EXAMPLES/}:
@@
 layered halfspace (bottom-left) and tomographic model (bottom-right).}
-
 \label{fig:examples.cubit}
@@
 \end{description}
+%
 \begin{figure}[htbp]
-\noindent \begin{centering}
+\begin{centering}
 \includegraphics[width=0.45\textwidth]{figures/mount-surface.jpg}
 \includegraphics[width=0.45\textwidth]{figures/mount-abs.jpg}
-\par\end{centering}
-
+\par
+\end{centering}
 \caption{Example of the block definitions for the free surface 'face\_topo'
@@
 (right) in CUBIT.}
-
 \label{fig:mount.abs}
@@
 called "nummaterial\_velocity\_file" that will have just been created and change it from the prototype created:
-{\small
+
+{\footnotesize
 \begin{verbatim}
@@
 }
+
 (where "vol1" and "vol2" here represent the volume labels that you have set while creating the mesh in CUBIT)
 to for instance
-{\small
+
+{\footnotesize
 \begin{verbatim}
@@
 Pellegrini et al. from LaBRI and INRIA in Bordeaux, France, downloadable
-from the web page \url{https://gforge.inria.fr/projects/scotch/}.
-\\
+from the web page \url{https://gforge.inria.fr/projects/scotch/}.\\
 
@@
 script. In the future you should be able to find more recent versions
-at \url{http://www.labri.fr/perso/pelegrin/scotch/scotch_en.html}
-. \\
+at \url{http://www.labri.fr/perso/pelegrin/scotch/scotch_en.html}.\\
 
 \begin{figure}[htbp]
-\noindent \begin{centering}
+\begin{centering}
 \includegraphics[width=0.45\textwidth]{figures/mount-partitions.jpg}
 \includegraphics[width=0.45\textwidth]{figures/mount-partitions2.jpg}
-\par\end{centering}
-
+\par
+\end{centering}
 \caption{Example of a mesh partitioning onto four cores. Each single core partition
@@
 decomposition is explained in detail in \citet{MaKoBlLe08}, and excellent
-scaling up to 150,000 processor cores in shown for instance in \citet{CaKoLaTiMiLeSnTr08,KoLaMi08a,MaKoBlLe08,KoErGoMi10,Kom11}.}
-
+scaling up to 150,000 processor cores in shown for instance in 
+\citet{CaKoLaTiMiLeSnTr08,KoLaMi08a,MaKoBlLe08,KoErGoMi10,Kom11}.}
 \label{fig:mount.partitions}
@@
 }
+\noindent
 If all paths and flags have been set correctly, the executable \texttt{bin/xdecompose\_mesh}
@@
 The partitioning is done in serial for now (in the next release we
-will provide a parallel version of that code). It needs to be run in the main directory because it expects the \texttt{./DATA/Par\_file}. The synopsis is:
+will provide a parallel version of that code). It needs to be run in the main directory because it expects the \texttt{./DATA/Par\_file}. 
+The synopsis is:
+%
 \begin{verbatim}
@@
 \end{verbatim}
+%
+\noindent
 where
@@
 for the mesh generation of relatively simple models.
-
+%
 \begin{figure}[htbp]
-\noindent \begin{centering}
+\begin{centering}
 \includegraphics[scale=0.5]{figures/socal_map_mpi.pdf}
-\par\end{centering}
-
+\par
+\end{centering}
 \caption{For parallel computing purposes,
@@
 }
+\noindent
 If all paths and flags have been set correctly, the mesher should
@@
 benchmarks on rectangular models, turn the UTM projection off by using
-the flag \texttt{\small SUPPRESS\_UTM\_PROJECTION}{\small , in which
-case all `longitude' parameters simply refer to the $x$~axis, and
+the flag \texttt{\small SUPPRESS\_UTM\_PROJECTION}, 
+in which case all `longitude' parameters simply refer to the $x$~axis, and
 all `latitude' parameters simply refer to the $y$~axis. To run the
 mesher for a global simulation, the following parameters need to be
-set in the }\texttt{\small Mesh\_Par\_file}{\small :}{\small \par}
+set in the \texttt{\small Mesh\_Par\_file}:
+%
 \begin{description}
@@
 \item [{\texttt{DEPTH\_BLOCK\_KM}}] Depth of bottom of mesh in kilometers.
-\item [{\texttt{\noun{UTM\_PROJECTION\_ZONE}}}] UTM projection zone in
-which your model resides, only valid when \texttt{SUPPRESS\_UTM\_}~\\
- \texttt{PROJECTION} is \texttt{.false.}.
+\item [{\texttt{UTM\_PROJECTION\_ZONE}}] UTM projection zone in
+which your model resides, only valid when 
+\texttt{SUPPRESS\_UTM\_PROJECTION} is \texttt{.false.}.
 Use a negative zone number for the Southern hemisphere:
@@
 the Southern hemisphere to zones -1 to -60.
+
 We use the WGS84 (World Geodetic System 1984) reference ellipsoid for the UTM projection. If you prefer to use the Clarke 1866 ellipsoid,
 edit file \texttt{src/shared/utm\_geo.f90}, uncomment that ellipsoid and recompile the code.
+
 From \url{http://en.wikipedia.org/wiki/Universal_Transverse_Mercator_coordinate_system}:\\
@@
 For different geographic regions, other datum systems (e.g.: ED50, NAD83) can be used.
-%
+
 \item [{\texttt{SUPPRESS\_UTM\_PROJECTION}}] set to be \texttt{.false.}
@@
 resides.}
+%
 \item [{\texttt{INTERFACES\_FILE }}] File which contains the description
@@
 direction within each layer is also defined in this file.
-\item [{$\nexxi$}] The number of spectral elements along one side of the
+%
+\item [{\texttt{NEX\_XI}}] The number of spectral elements along one side of the
 block. This number \textit{must} be 8~$\times$~a multiple of $\nprocxi$
@@
 elements. We suggest not to change this value.
-\item [{$\nexeta$}] The number of spectral elements along the other side
+%
+\item [{\texttt{NEX\_ETA}}] The number of spectral elements along the other side
 of the block. This number \textit{must} be 8~$\times$~a multiple
 of $\nproceta$ defined below.
-\item [{$\nprocxi$}] The number of processors or slices along one side
+%
+\item [{\texttt{NPROC\_XI}}] The number of processors or slices along one side
 of the block (see Figure~\ref{fig:For-parallel-computing}); we must
@@
 integer.
-\item [{$\nproceta$}] The number of processors or slices along the other
+%
+\item [{\texttt{NPROC\_ETA}}] The number of processors or slices along the other
 side of the block; we must have $\nexeta=8\times c\times\nproceta$,
 where $c\ge1$ is a positive integer.
+%
 \item [{\texttt{USE\_REGULAR\_MESH}}] set to be \texttt{.true.} if you
@@
 three parameters.
+%
 \item [{\texttt{NDOUBLINGS}}] The number of horizontal doubling layers.
@@
 discussion of the \texttt{LOCAL\_PATH\_IS\_ALSO\_GLOBAL} flag in Chapter~\ref{cha:Getting-Started}).
+
 The mesher generates the necessary partitions in parallel, one set
@@
 further details.
+%
 \item [{\texttt{NMATERIALS}}] The number of different materials in your
-model. In the following lines, each material needs to be defined as
-:
+model. In the following lines, each material needs to be defined as:
 
@@
 
+\noindent
 where
@@
 lines, because the mesh is regular or 'almost regular', each region
-is defined as :
-
+is defined as:
 \begin{verbatim}
@@
 \end{description}
+%
 The \texttt{INTERFACES\_FILE} parameter of \texttt{Mesh\_Par\_File}
@@
 \end{itemize}
+%
 At the end of this file, you simply need to set the number of spectral
 elements in the vertical direction for each layer. We provide a few
-models in the {\texttt{EXAMPLES/}} directory. \\
+models in the {\texttt{EXAMPLES/}} directory.\\
 

### 2015-05-21 doc/USER_MANUAL/04_creating_databases.tex
Buildbot ran the code beautifier.
updates manual for better wiki display

@@
 \begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[width=0.6\textwidth]{figures/workflow.pdf}
-\par\end{centering}
-
+\begin{centering}
+\includegraphics[width=0.6\textwidth]{figures/workflow.jpg}
+\par
+\end{centering}
 \caption{Schematic workflow for a SPECFEM3D Cartesian simulation. The executable
@@
 specific model parameters.}
-
 \label{fig:workflow.databases}
@@
 }
+\noindent
 Input for the program is provided through the main parameter file
@@
 and \texttt{SAVE\_FORWARD} at their default values.
-\item [{\texttt{\noun{UTM\_PROJECTION\_ZONE}}}] UTM projection zone in
-which your model resides, only valid when \texttt{SUPPRESS\_UTM\_}~\\
- \texttt{PROJECTION} is \texttt{.false.}.
+\item [{\texttt{UTM\_PROJECTION\_ZONE}}] UTM projection zone in
+which your model resides, only valid when \texttt{SUPPRESS\_UTM\_PROJECTION} is \texttt{.false.}.
 \item [{\texttt{SUPPRESS\_UTM\_PROJECTION}}] set to be \texttt{.false.}
@@
 flat, i.e. identical to HEX8). To generate HEX27 elements with curvature
-properly taken into account, you can use \texttt{Gmsh} \url{http://geuz.org/gmsh/}
+properly taken into account, you can use Gmsh \url{http://geuz.org/gmsh/}
 \item [{\texttt{MODEL}}] Must be set to one of the following:
@@
 source time function.
+
 When using this option, by default the code can locate the force source
@@
 This feature is only relevant for the solver.
-
+%
 \item [{\texttt{NUMBER\_OF\_SIMULTANEOUS\_RUNS}}] adds the ability to run several calculations (several earthquakes)
@@
 each of them computing several earthquakes in parallel.
-To turn that option on, set parameter NUMBER\_OF\_SIMULTANEOUS\_RUNS to a value greater than 1.
-To implement that, we create NUMBER\_OF\_SIMULTANEOUS\_RUNS MPI sub-communicators,
-each of them being labeled "my\_local\_mpi\_comm\_world", and we use them
+
+To turn that option on, set parameter \texttt{NUMBER\_OF\_SIMULTANEOUS\_RUNS} to a value greater than 1.
+To implement that, we create \texttt{NUMBER\_OF\_SIMULTANEOUS\_RUNS} MPI sub-communicators,
+each of them being labeled \texttt{my\_local\_mpi\_comm\_world}, and we use them
 in all the routines in "src/shared/parallel.f90", except in MPI\_ABORT() because in that case
 we need to kill the entire run.
+
 When that option is on, of course the number of processor cores used to start
-the code in the batch system must be a multiple of NUMBER\_OF\_SIMULTANEOUS\_RUNS,
+the code in the batch system must be a multiple of \texttt{NUMBER\_OF\_SIMULTANEOUS\_RUNS},
 all the individual runs must use the same number of processor cores,
@@
 and thus the total number of processor cores to request from the batch system
-should be NUMBER\_OF\_SIMULTANEOUS\_RUNS * NPROC.
+should be \texttt{NUMBER\_OF\_SIMULTANEOUS\_RUNS}~*~\texttt{NPROC}.
 All the runs to perform must be placed in directories called run0001, run0002, run0003 and so on
@@
 
-The option NUMBER\_OF\_SIMULTANEOUS\_RUNS implements 3/.
+The option \texttt{NUMBER\_OF\_SIMULTANEOUS\_RUNS} implements 3/.
 
-\item [{\texttt{BROADCAST\_SAME\_MESH\_AND\_MODEL}}]:
-if we perform simultaneous runs in parallel, if only the source and receivers vary between these runs
+\item [{\texttt{BROADCAST\_SAME\_MESH\_AND\_MODEL}}] if we perform simultaneous runs in parallel, 
+if only the source and receivers vary between these runs
 but not the mesh nor the model (velocity and density) then we can also read the mesh and model files
@@
 runs for instance when solving inverse problems iteratively this can DRASTICALLY reduce I/Os to disk in the solver
-(by a factor equal to NUMBER\_OF\_SIMULTANEOUS\_RUNS), and reducing I/Os is crucial in the case of huge runs.
+(by a factor equal to \texttt{NUMBER\_OF\_SIMULTANEOUS\_RUNS}), and reducing I/Os is crucial in the case of huge runs.
 Thus, always set this option to .true. if the mesh and the model are the same for all simultaneous runs.
@@
 
-\item [{\texttt{USE\_FAILSAFE\_MECHANISM}}]:
-if one or a few of these simultaneous runs fail, kill all the runs or let the others finish using a fail-safe mechanism
+\item [{\texttt{USE\_FAILSAFE\_MECHANISM}}] if one or a few of these simultaneous runs fail, 
+kill all the runs or let the others finish using a fail-safe mechanism
 (in most cases, should be set to true).
 
-TODO / future work to do: currently the BROADCAST\_SAME\_MESH\_AND\_MODEL option assumes to have the (master) mesh files in run0001/DATABASES\_MPI or run0001/OUTPUT\_FILES/DATABASES\_MPI.
-However, for adjoint runs you still need a DATABASES\_MPI folder in each of the sub-runs directories, e.g.
-run0002/DATABASES\_MPI, etc.
-to store the forward wavefields, kernels etc.
-of each sub-run.
+TODO / future work to do: currently the \texttt{BROADCAST\_SAME\_MESH\_AND\_MODEL} 
+option assumes to have the (master) mesh files in run0001/DATABASES\_MPI or run0001/OUTPUT\_FILES/DATABASES\_MPI. 
+However, for adjoint runs you still need a DATABASES\_MPI folder in each of the sub-runs directories,
+e.g. run0002/DATABASES\_MPI, etc. to store the forward wavefields, kernels etc. of each sub-run.
 This would not be needed for forward simulations.
@@
 Thus these tools should work, but in their current version will need to be run for each simulation result independently.
+
 More precisely, the current kernel summing and smoothing routines work fine, with the exception that you need to move out the mesh files (and also the parameters).
@@
 where\\
-rho\_s = solid density\\
-rho\_f = fluid density\\
-phi = porosity\\
-tort = tortuosity\\
-kxx = xx component of permeability tensor\\
-kxy = xy,yx components of permeability tensor\\
-kyy = yy component of permeability tensor\\
-kxz = xz,zx components of permeability tensor\\
-kzz = zz component of permeability tensor\\
-kappa\_s = solid bulk modulus\\
-kappa\_f= fluid bulk modulus\\
-kappa\_fr= frame bulk modulus\\
-eta\_f = fluid viscosity\\
-mu\_fr = frame shear modulus.\\
+\texttt{rho\_s} = solid density,\\
+\texttt{rho\_f} = fluid density,\\
+\texttt{phi} = porosity,\\
+\texttt{tort} = tortuosity,\\
+\texttt{kxx} = xx component of permeability tensor,\\
+\texttt{kxy} = xy,yx components of permeability tensor,\\
+\texttt{kyy} = yy component of permeability tensor,\\
+\texttt{kxz} = xz,zx components of permeability tensor,\\
+\texttt{kzz} = zz component of permeability tensor,\\
+\texttt{kappa\_s} = solid bulk modulus,\\
+\texttt{kappa\_f} = fluid bulk modulus,\\
+\texttt{kappa\_fr} = frame bulk modulus,\\
+\texttt{eta\_f} = fluid viscosity,\\
+\texttt{mu\_fr} = frame shear modulus.\\
 

### 2015-05-21 doc/USER_MANUAL/05_running_the_solver.tex
Buildbot ran the code beautifier.
updates manual for better wiki display

@@
 }
+\noindent
 Please note that \texttt{xspecfem3D} must be called directly from
@@
 and \texttt{NTSTEPS\_BETWEEN\_FRAMES}
-\item the ShakeMap\textregistered{}option \texttt{CREATE\_SHAKEMAP}
-\item the output information parameters \texttt{MOVIE\_TYPE}, \texttt{NTSTEP\_BETWEEN\_OUTPUT\_INFO}
-and \texttt{NTSTEP\_BETWEEN\_OUTPUT\_}~\\
- \texttt{SEISMOS}
+\item the ShakeMap\textregistered{} option \texttt{CREATE\_SHAKEMAP}
+\item the output information parameters \texttt{MOVIE\_TYPE}, \texttt{NTSTEP\_BETWEEN\_OUTPUT\_INFO} and\\
+\texttt{NTSTEP\_BETWEEN\_OUTPUT\_SEISMOS}
 \item the \texttt{PRINT\_SOURCE\_TIME\_FUNCTION} flags
@@
 \begin{figure}[htp]
-\noindent
 \begin{centering}
-{\small \includegraphics[width=0.8\textwidth]{figures/Hollywood_CMT.pdf}
-}
+\includegraphics[width=0.8\textwidth]{figures/Hollywood_CMT.jpg}
 \par
@@
 is the time in the first column of the output seismogram.
-
 \end{itemize}
+%
 \begin{figure}[htp]
-\noindent \begin{centering}
-\includegraphics[width=3.5in]{figures/gauss_vs_triangle_mod.pdf}
-\par\end{centering}
-
-\caption{Comparison of the shape of a triangle and the Gaussian function actually
-used.}
-
+\begin{centering}
+\includegraphics[width=3.5in]{figures/gauss_vs_triangle_mod.jpg}
+\par
+\end{centering}
+\caption{Comparison of the shape of a triangle and the Gaussian function actually used.}
 \label{fig:gauss.vs.triangle}
@@
 \begin{figure}[H]
-\noindent \begin{centering}
-{\small \includegraphics[width=5in]{figures/source_timing.pdf} }
-\par\end{centering}
-
+\begin{centering}
+\includegraphics[width=4in]{figures/source_timing.jpg}
+\par
+\end{centering}
 \caption{Example of timing for three sources. The center of the first source
@@
 duration} would be hdur1, hdur2, hdur3 for the sources 1, 2, 3 respectively.}
-
 {\small \label{fig:source_timing} }
@@
 which looks like this:
-
+%
 \begin{figure}[H]
-\noindent \begin{centering}
-{\small \includegraphics{figures/STATIONS_basin_explained.pdf} }
-\par\end{centering}
-
-\label{fig:Sample-STATIONS-file.}
+\begin{centering}
+\includegraphics[width=4in]{figures/STATIONS_basin_explained.jpg}
+\par
+\end{centering}
 \caption{Sample \texttt{STATIONS} file. Station latitude and longitude should
@@
 in the \texttt{constants.h} file).}
+\label{fig:Sample-STATIONS-file.}
 \end{figure}

### 2015-05-21 doc/USER_MANUAL/11_graphics.tex
Buildbot ran the code beautifier.
updates manual for better wiki display

@@
 \]
+\end{quote}
+
 \textbf{NOTE:} If \texttt{HDUR\_MOVIE} is set to 0.0, the code will
@@
 set \texttt{HDUR\_MOVIE} = 0.0.
-\end{quote}
 
@@
 \begin{figure}[htbp]
-\noindent \begin{centering}
+\begin{centering}
 \includegraphics[width=0.32\textwidth]{figures/movie_surf_1.jpg}
@@
 \includegraphics[width=0.32\textwidth]{figures/movie_surf_3.jpg}
-\par\end{centering}
-
+\par
+\end{centering}
 \caption{Visualization using AVS files created by \texttt{xcreate\_movie\_shakemap\_AVS\_DX\_GMT}
-showing movie snapshots of vertical velocity components at different
-times.}
-
+showing movie snapshots of vertical velocity components at different times.}
 \label{fig:movie.surf}
@@
 }
+\noindent
 For elastic domains, the divergence and curl taken from the velocity field,
@@
 }
+\noindent
 The files denoted \texttt{\small proc??????\_div\_glob\_it??????.bin} and
@@
 \begin{figure}[htbp]
-\noindent \begin{centering}
+\begin{centering}
 \includegraphics[width=0.32\textwidth]{figures/movie_volume_1.jpg}
@@
 \includegraphics[width=0.32\textwidth]{figures/movie_volume_3.jpg}
-\par\end{centering}
-
+\par
+\end{centering}
 \caption{Paraview visualization using movie volume files (converted by \texttt{xcombine\_vol\_data}
@@
 components at different times.}
-
 \label{fig:movie.volume}
@@
 }
+\noindent
 which will create the executable \texttt{bin/xcombine\_vol\_data}.
 To output the usage of this executable, type
+
 {\footnotesize
@@
 }
+\noindent
 without arguments.
@@
 \end{enumerate}
+%
 \begin{figure}[H]
-\noindent \begin{centering}
-\includegraphics[scale=0.6]{figures/3D-S-Kernel.pdf}
-\par\end{centering}
-
+\begin{centering}
+\includegraphics[width=6in]{figures/3D-S-Kernel.jpg}
+\par
+\end{centering}
 \caption{(a) Top Panel: Vertical source-receiver cross-section of the S-wave
@@
 an epicentral distance of 165 km \citep{LiTr06}.}
-
 \label{figure:P-wave-speed-finite-frequency}

### 2015-05-21 doc/USER_MANUAL/06_fault_sources.tex

@@
 \texttt{CUBIT}. \texttt{xmeshfem3D} is not yet ready to handle faults.\\
+
 To facilitate the mesh generation with split nodes in CUBIT, we need
@@
 
+\noindent
 Here is an example CUBIT script to generate a mesh with split nodes
 for a buried vertical strike-slip fault:
+
 {\footnotesize
@@
 }
-%
+
 \begin{figure}[htbp]
-\noindent \begin{centering}
+\begin{centering}
 \includegraphics[scale=0.3]{figures/faultmesh.jpg} \\
@@
 \includegraphics[scale=0.3]{figures/splitnodes.jpg}
-\par\end{centering}
-
+\par
+\end{centering}
 \caption{Screenshots of the CUBIT example for the embedded fault with split
@@
 \begin{figure}[htbp]
-\noindent \begin{centering}
-\includegraphics[scale=0.55]{figures/splitnodes_surfacetrace.pdf} \\
-
-\par\end{centering}
-
+\begin{centering}
+\includegraphics[scale=0.55]{figures/splitnodes_surfacetrace.pdf}
+\par
+\end{centering}
 \caption{Screenshots of a CUBIT example showing split nodes for a fault reaching
@@
 offset by a small distance on either side of the fault}
-
 \label{fig:examples.splitnodes-surfacetrace}
@@
 
+\noindent
 The CUBIT scripts ({*}.jou and {*}.py) in the directory \texttt{EXAMPLES}
@@
 
+\newpage
+\noindent
 Note that you should avoid gaps in the list of indices of mesh objects
 with the following CUBIT command:
+
 {\footnotesize
@@
 }
+\noindent
 (otherwise you will get a segmentation fault during domain decomposition)
@@
 
+\noindent
 The directories \texttt{EXAMPLES/*/post} contain additional Matlab

### 2015-05-21 doc/USER_MANUAL/07_adjoint_simulations.tex

@@
 script:}{\small \par}
-
-\begin{lyxcode}
-utils/change\_simulation\_type.pl~-f~
-\end{lyxcode}
+\begin{verbatim}
+utils/change_simulation_type.pl -f
+\end{verbatim}
 
@@
 
-\begin{lyxcode}
-xcreate\_adjsrc\_traveltime~t1~t2~ifile{[}0-5{]}~E/N/Z-ascii-files~{[}baz{]}
-\end{lyxcode}
-
+\begin{verbatim}
+xcreate_adjsrc_traveltime t1 t2 ifile[0-5] E/N/Z-ascii-files [baz]
+\end{verbatim}
 where \texttt{t1} and \texttt{t2} are the start and end time of the
@@
 and used in the same way as described for traveltime measurements
-
-\begin{lyxcode}
-xcreate\_adjsrc\_amplitude~t1~t2~ifile{[}0-5{]}~E/N/Z-ascii-files~{[}baz{]}.
-\end{lyxcode}
+\begin{verbatim}
+xcreate_adjsrc_amplitude t1 t2 ifile[0-5] E/N/Z-ascii-files [baz]
+\end{verbatim}
 \end{itemize}

### 2015-05-21 doc/USER_MANUAL/C_troubleshooting.tex

@@
 to compile this little program code:
-\begin{lyxcode}
-program~main~~~\\
-~include~'mpif.h'~~~\\
-~integer,~parameter~::~CUSTOM\_MPI\_TYPE~=~MPI\_REAL~~~\\
-~integer~ier~~~\\
-~call~MPI\_INIT(ier)~~~\\
-~call~MPI\_BARRIER(MPI\_COMM\_WORLD,ier)~~~\\
-~call~MPI\_FINALIZE(ier)~~~\\
-~end
-\end{lyxcode}
-\item [{compilation fails stating:}] ~\\
 
-\begin{lyxcode}
-...~~~\\
-~obj/program\_generate\_databases.o:~In~function~`MAIN\_\_':~~~\\
-~program\_generate\_databases.f90:(.text+0x14):~undefined~\\~~~reference~to~`\_gfortran\_set\_std'~~~\\
-~...~~~\\
+{\footnotesize
+\begin{verbatim}
+ program main
+ include 'mpif.h'
+ integer, parameter :: CUSTOM_MPI_TYPE = MPI_REAL
+ integer ier
+ call MPI_INIT(ier)
+ call MPI_BARRIER(MPI_COMM_WORLD,ier)
+ call MPI_FINALIZE(ier)
+ end
+\end{verbatim}
+}
 
-\end{lyxcode}
+\item [{compilation fails stating:}] ~
+{\footnotesize
+\begin{verbatim}
+...
+ obj/program_generate_databases.o: In function `MAIN__':
+ program_generate_databases.f90:(.text+0x14): undefined reference to `_gfortran_set_std'
+...
+\end{verbatim}
+}
 
@@
 
-\begin{lyxcode}
-...~~\\
-~histogram~of~skewness~(0.~good~-~1.~bad):~~~\\
-~0.0000000E+00~-~5.0000001E-02~27648~81.81818~\%~~~\\
-~5.0000001E-02~-~0.1000000~0~0.0000000E+00~\%~~~\\
-~...~~~\\
-
-\end{lyxcode}
+{\footnotesize
+\begin{verbatim}
+...
+ histogram of skewness (0. good - 1. bad):
+ 0.0000000E+00 - 5.0000001E-02 27648 81.81818 %
+ 5.0000001E-02 - 0.1000000 0 0.0000000E+00 %
+...
+\end{verbatim}
+}
 
@@
 file a bit further down):
-\begin{lyxcode}
-...~~\\
-~total~number~of~elements~in~entire~mesh:~33792~~~\\
-~...~~~\\
 
-\end{lyxcode}
+{\footnotesize
+\begin{verbatim}
+...
+ total number of elements in entire mesh: 33792
+...
+\end{verbatim}
+}
 
@@
 for example the worst ones belong to:
-\begin{lyxcode}
-...~~\\
-~0.6000000~-~0.6500000~2048~6.060606~\%~~~\\
-~...~~~\\
 
-\end{lyxcode}
+{\footnotesize
+\begin{verbatim}
+...
+ 0.6000000 - 0.6500000 2048 6.060606 %
+...
+
+\end{verbatim}
+}
 
@@
 e.g.:
-\begin{lyxcode}
-SUPPRESS\_UTM\_PROJECTION~=~.false.~~~\\
-
-\end{lyxcode}
-
+\begin{verbatim}
+SUPPRESS_UTM_PROJECTION = .false.
+\end{verbatim}
 to be the same in \texttt{Mesh\_Par\_file} and \texttt{Par\_file}.
@@
 
-\begin{lyxcode}
-...~~~\\
-~{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}~~~\\
-~{*}{*}{*}~Verification~of~simulation~parameters~{*}{*}{*}~~~\\
-~{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}~~~\\
-~...~~~\\
-~{*}{*}{*}~Minimum~period~resolved~=~4.308774~~~\\
-~{*}{*}{*}~Maximum~suggested~time~step~=~6.8863556E-02~~~\\
-~...~~~\\
-~...~~~\\
-
-\end{lyxcode}
+{\footnotesize
+\begin{verbatim}
+...
+ *********************************************
+ *** Verification of simulation parameters ***
+ *********************************************
+...
+ *** Minimum period resolved = 4.308774
+ *** Maximum suggested time step = 6.8863556E-02
+...
+\end{verbatim}
+}
 

### 2015-06-02 utils/EXTERNAL_CODES_coupled_with_SPECFEM3D/AxiSEM_for_SPECFEM3D/user_manual_coupling_AxiSEM_Specfem3D.tex

% Type de document
\documentclass[a4paper,11pt]{article}

% Chargement des extensions
\usepackage[latin1]{inputenc}
 \usepackage{graphicx} 
\begin{document}
{\centerline{\bf{Coupling AxiSEM and Specfem3D}}}
{\bf{Authors:}} V. Monteiller

\section{Meshfem3D}

running Meshfem3D with parameters : \newline

In the directory MESH/ \newline

ParFileMeshChunk  : parameters for the chunk meshing \newline

iasp91\_dsm or prem\_dsm : background model used in both AxiSEM ans Specfem3D \newline

In the Directory DATA/  \newline

CMTSOLUTION  : with 0 in order to not use source inside \newline

coeff\_poly\_deg12 : to generate smooth intitial solution  \newline

STATIONS  : stations files  \newline

Par\_file : with \newline

SIMULATION\_TYPE=1 \newline

SAVE\_FORWARD = .false. \newline

COUPLE\_WITH\_EXTERNAL\_CODE = .true \newline

EXTERNAL\_CODE\_TYPE    = 2 \newline

\section{AxiSEM mesher}

running AxiSEM mesher

\section{AxiSEM solver}
copy 2 files produced by meshfem3D in the running AxisSEM directory: \newline

input\_box.txt and input\_box\_sem\_cart.txt  \newline

add one first line to indicate the number of line to be read \newline

running AxiSEM solver with parameters (differences with *.TEMPLATES provided
bny AxiSEM) : \newline

inparam\_basic : \newline

ATTENUATION         false \newline
SAVE\_SNAPSHOTS     true \newline

inparam\_advanced : \newline

KERNEL\_WAVEFIELDS   true \newline
KERNEL\_IBEG         0 \newline
KERNEL\_IEND         4 \newline

\section{Specfem Partitionning}

running scotch partitionning

\section{Specfem Generate database}

running generate database

\section{Interface : expand 2D to 3D}

mpi run with arbitrary number of processes  \newline

parameter file : expand\_2D\_3D.par (in SOLVER directory)\newline

input\_box.txt \newline
input\_box\_sem\_cart.txt \newline
8                          \# number of AxiSEM mpi processes used in solver  \newline
60. 0.                     \# source position (lat lon)  \newline
0.  60.                    \# chnuk center (lat lon)    \newline
1                          \# number of axisem simus depends on moment tensor
used  \newline
8                          \# number of Specfem3D MPI processes   \newline
../../run\_synth\_alps/create\_mesh/MESH \# Specfem MESH directory  \newline
../../run\_synth\_alps/create\_mesh/OUTPUT\_FILES/DATABASES\_MPI \# Specfem

TRACTION DIRECTORY \newline

\section{Interface : reformat}

mpi run with the **SAME** number of processes that Specfem3D will use. One
file is created by one process for one Spefem3D partion of domain. \newline

copy input\_box.txt and input\_box\_sem\_cart.txt inside the directory where
AxiSEM did the run and add a new parameter file: \newline

reformat.par  \newline
25.         \# output sampling in Hz (time step that will use in Specfem3D simu) \newline
650. 700.   \# begin time and end time (s.) \newline

\section{Running specfem3D}

\section{Set up scripts}

\end{document}

### 2015-06-07 doc/USER_MANUAL/02_getting_started.tex
clarified the "configure" command

@@
 
-Then, to configure the software for your system, run the \texttt{configure}
-shell script. This script will attempt to guess the appropriate configuration
-values for your system. However, at a minimum, it is recommended that
-you explicitly specify the appropriate command names for your Fortran
-compiler and MPI package (another option is to define FC, CC and MPIF90
-in your .bash\_profile or your .cshrc file):
+Then, to configure the software for your system, run the
+\texttt{configure} shell script. This script will attempt to guess
+the appropriate configuration values for your system. However, at
+a minimum, it is recommended that you explicitly specify the appropriate
+command names for your Fortran compiler (another option is to define FC, CC and MPIF90 in your .bash\_profile
+or your .cshrc file):
+%
 \begin{verbatim}
-./configure FC=ifort MPIFC=mpif90
+    ./configure FC=gfortran CC=gcc
 \end{verbatim}
-\noindent
+%
+If you want to run in parallel, i.e., using more than one processor core, then you would type
+%
+\begin{verbatim}
+    ./configure FC=gfortran CC=gcc MPIFC=mpif90 --with-mpi
+\end{verbatim}
+
+You can replace the GNU compilers above (gfortran and gcc) with other compilers if you want to; for instance for Intel ifort and icc use FC=ifort CC=icc instead.
+
 Note that MPI must be installed with MPI-IO enabled because parts of SPECFEM3D perform I/Os through MPI-IO.
@@
 to access the development branch of the source code with read/write access (in a safe way, no need to worry too much about breaking the package, there is a robot called BuildBot that is in charge of checking and validating all new contributions and changes), please visit this Web page: \url{https://github.com/geodynamics/specfem3d/wiki/Using-Hub}.
+
+To visualize the call tree (calling tree) of the source code, you can see the Doxygen tool available in directory \texttt{doc/call\_trees\_of\_the\_source\_code}.

### 2015-06-18 doc/USER_MANUAL/02_getting_started.tex
updates user manual

@@
 
-The SCOTCH library \citep{PeRo96} provides efficent static mapping,
+The SCOTCH library \citep{PeRo96} provides efficient static mapping,
 graph and mesh partitioning routines. SCOTCH is a free software package
@@
 The configuration script automatically creates for each executable
-a corresponding \texttt{Makefile} in the \texttt{src/} subdirectoy.
+a corresponding \texttt{Makefile} in the \texttt{src/} subdirectory.
 The \texttt{Makefile} contains a number of suggested entries for various
@@
 clusters (typically more than $10,000$ processes). SPECFEM3D use the ADIOS library~\cite{Liu2013}
-to deal transparently take advantage of advanced paralllel file system features. To enable
+to deal transparently take advantage of advanced parallel file system features. To enable
 ADIOS, the following steps should be done:

### 2015-06-18 doc/USER_MANUAL/04_creating_databases.tex
updates user manual

@@
 generic template routines to use your own model definitions.
+
 \item [{\texttt{gll}}] For a user-specified isotropic model which uses
@@
 files define the model parameters on all GLL points which can be used
-for iterative inversion procedures.
+for iterative inversion procedures. Note that for simulation setups with attenuation, 
+it will also read in the external binary mesh files for $Q_{\kappa}$ and $Q_{\mu}$.
+
 \item [{\texttt{salton\_trough}}] A 3D $V_{p}$ model for Southern California.
@@
 in directory \texttt{DATA/st\_3D\_block\_harvard/}.
+
 \item [{\texttt{tomo}}] For a user-specified 3D isotropic model which uses
@@
 \end{description}
+
 \item [{\texttt{APPROXIMATE\_OCEAN\_LOAD}}] Set to \texttt{.true.} if the
@@
 TODO / future work to do: currently the \texttt{BROADCAST\_SAME\_MESH\_AND\_MODEL}
-option assumes to have the (master) mesh files in run0001/DATABASES\_MPI or run0001/OUTPUT\_FILES/DATABASES\_MPI. 
-However, for adjoint runs you still need a DATABASES\_MPI folder in each of the sub-runs directories,
-e.g. run0002/DATABASES\_MPI, etc. to store the forward wavefields, kernels etc. of each sub-run.
+option assumes to have the (master) mesh files in \texttt{run0001/DATABASES\_MPI} or 
+\texttt{run0001/OUTPUT\_FILES/DATABASES\_MPI}.
+However, for adjoint runs you still need a \texttt{DATABASES\_MPI/} folder in each of the sub-runs directories,
+e.g. \texttt{run0002/DATABASES\_MPI}, etc. to store the forward wavefields, kernels etc. of each sub-run.
 This would not be needed for forward simulations.
@@
 The way we prescribe material property for porous material in SPECFEM3D is as follow:
-We use a file name nummaterial\_poroelastic\_file, which is located in the directory MESH, the format is as follow:
+We use a file name "nummaterial\_poroelastic\_file", which is located in the directory \texttt{MESH/}, the format is as follow:
 \begin{verbatim}
@@
 
-Using an external mesh (for instance coming from CUBIT/TRELIS), poroelastic materials have the ID number 3, while 1 is elastic, 2 is acoustic ?
-see the example in the package: examples/homogeneous\_poroelastic.
+Using an external mesh (for instance coming from CUBIT/TRELIS), poroelastic materials have the ID number 3, while 1 is acoustic and 2 is elastic (see the example in the package: EXAMPLES/homogeneous\_poroelastic).
 

### 2015-06-18 doc/USER_MANUAL/11_graphics.tex
updates user manual

@@
 
-To visualize these files, we use an auxilliary program \texttt{combine\_vol\_data.f90}
+To visualize these files, we use an auxiliary program \texttt{combine\_vol\_data.f90}
 to combine the data from all slices into one mesh file. To compile
@@
 Therefore, we first need to collect these files on the front end,
-combine them into one mesh file, and visualize them with some auxilliary
+combine them into one mesh file, and visualize them with some auxiliary
 programs.
@@
 
-We use an auxilliary program \texttt{combine\_vol\_data.f90} to combine
+We use an auxiliary program \texttt{combine\_vol\_data.f90} to combine
 the kernel files from all slices into one mesh file.

### 2015-06-18 doc/USER_MANUAL/13_changing_the_model.tex
updates user manual

@@
 as read from \texttt{model\_tomography.f90} located in the \texttt{src/generate\_databases}
-directory, looks like Figure \ref{fig:tomography_file}, where
+directory, looks like Figure \ref{fig:tomography_file}, starting with a header information where
 \begin{description}
@@
 \end{description}
-After the first four lines, in the file \texttt{file\_name} the tomographic
+After the first four lines, the tomography file \texttt{file\_name} lists the data record where all tomographic
 grid points are listed with the corresponding values of \texttt{vp},
-\texttt{vs} and \texttt{rho}, scanning the grid along the x coordinate
+\texttt{vs} and \texttt{rho} (and optionally $Q_{p}$ and $Q_{s}$), scanning the grid along the x coordinate
 (from \texttt{ORIG\_X} to \texttt{END\_X} with step of \texttt{SPACING\_X})
@@
 of \texttt{SPACING\_Y}) and each given z (from \texttt{ORIG\_Z} to
-\texttt{END\_Z}, with step of \texttt{SPACING\_Z}). %fig
+\texttt{END\_Z}, with step of \texttt{SPACING\_Z}).
+
+\noindent
+Each data record line provides the velocity model values in a format like:  
+{\small
+\begin{verbatim}
+# data record format: purely isotropic model
+x	y	z		vp		vs		rho
+..
+\end{verbatim}
+}
+\noindent
+or
+{\small
+\begin{verbatim}
+# data record format: anelastic isotropic model
+x	y	z		vp		vs		rho	Qp	Qs
+..
+\end{verbatim}
+}
+\noindent
+where  \texttt{x}, \texttt{y}, \texttt{z} are the grid point position, \texttt{vp}, \texttt{vs} and \texttt{rho} the P- and S-wave speeds and density, and \texttt{Qp} and \texttt{Qs} the quality factors for P- and S-wave speeds.
+The quality factors are optional and can be omitted for purely elastic models (the tomography routine will recognize both formats).
+Internally, the quality factors will be converted to bulk and shear attenuation values,
+$Q_{\kappa}$ and $Q_{\mu}$ respectively \citep{AnHa78}. 
+For simulations with attenuation, please note that the Vp- and Vs-velocities of your model are given for a reference
+frequency. To change this reference frequency, you change the value
+of \texttt{ATTENUATION\_f0\_REFERENCE} in the main constants file
+\texttt{constants.h} found in subdirectory \texttt{src/shared/}.
+
+%fig%
 \begin{figure}[htbp]
@@
 
-\caption{Tomography file \texttt{file\_name} that describes an external Earth
+\caption{Tomography file \texttt{file\_name} that describes an external, purely elastic Earth
 model. The coordinates x, y and z of the grid, their limits ORIG\_X,
@@
 and the file \texttt{tomography\_model.xyz} will need to reside in \texttt{DATA/}.
-All entries in the second column of \texttt{materials\_file} will be \texttt{-1}, which means that each element in the mesh will be interpolated accordind to the values in \texttt{tomography\_model.xyz}.
+All entries in the second column of \texttt{materials\_file} will be \texttt{-1}, which means that each element in the mesh will be interpolated according to the values in \texttt{tomography\_model.xyz}.
 
@@
 call model_external_values(xmesh, ymesh, zmesh, &
-                           rho, vp, vs, qmu_atten, iflag_aniso, idomain_id)
+            rho, vp, vs, qkappa_atten, qmu_atten, iflag_aniso, idomain_id)
 \end{verbatim}
@@
 ($kg/m^{3}$), $v_{p}$ ($m/s$) and $v_{s}$ ($m/s$)
-\item [{\texttt{qmu\_atten}}] Shear wave quality factor: $0<Q_{\mu}<9000$
+\item [{\texttt{qkappa\_atten,qmu\_atten}}] Bulk and shear wave quality factor: $0<Q_{\kappa,\mu}<9000$
 \item [{\texttt{iflag\_aniso}}] anisotropic model flag, $0$ indicating
@@
 \begin{description}
-\item [{\texttt{c11},}] \textbf{$\cdots$,} \texttt{\textbf{c66}} 21~dimensionalized
+\item [{\texttt{c11,..,c66}}] 21~dimensionalized
 anisotropic elastic parameters.

### 2015-06-19 utils/EXTERNAL_CODES_coupled_with_SPECFEM3D/AxiSEM_for_SPECFEM3D/Manual_coupling_AxiSEM_Specfem3D.tex
Step 7 coupling AxiSEM validated + edit coupling Manual

\documentclass[11pt]{article}
%
%\usepackage{kmath,kerkis} 
\usepackage{mathpazo}
%\usepackage{palatino}
\usepackage[T1]{fontenc} 
\usepackage{amsfonts}
\usepackage{mathrsfs}
\usepackage{amsthm}
\usepackage{dsfont}
%\usepackage{marvosym}
\usepackage{eurosym}
\usepackage{stmaryrd}
%\usepackage{gensymb}
\usepackage{hhline}
\usepackage{pifont}
\usepackage{fancybox}
\usepackage{enumerate}
%\usepackage{enumitem}
\usepackage{multirow}
\usepackage[table]{xcolor}
\usepackage[normalem]{ulem}
\usepackage{amsmath,amssymb}             % AMS Math
\usepackage[french]{babel}
\frenchbsetup{StandardLists=true}
%\usepackage{esint}
%\usepackage[latin1]{inputenc}

%\usepackage[hmargin=2cm,vmargin=4cm]{geometry}
%\usepackage[left=2.7cm,right=2cm,top=2.3cm,bottom=2.3cm,includefoot,includehead,headheight=13.6pt]{geometry}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\definecolor{goldp}{rgb}{0.99 0.92 0.8}
\definecolor{oxblue}{RGB}{2 33 71}
\definecolor{cobalt}{RGB}{0 71 171}
\definecolor{impb}{RGB}{20 35 149}
\definecolor{graylight}{gray}{0.85}
\definecolor{denim}{RGB}{21 96 189}
\definecolor{bgreen}{RGB}{13 152 186}
\definecolor{bdeaux}{RGB}{153 12 20}
\definecolor{linkcol}{rgb}{0,0,0.4} 
\definecolor{citecol}{rgb}{0.5,0.01,0.02}
\definecolor{lsb}{RGB}{120, 200, 200}

\DeclareSymbolFont{calletters}{OMS}{cmsy}{m}{n}
\DeclareSymbolFontAlphabet{\mathcal}{calletters}

\newcommand{\bw}{\mathbf{w}} 
\newcommand{\be}{\mathbf{e}}  
\newcommand{\bh}{\mathbf{h}}  
\newcommand{\bT}{\mathbf{T}}  
\newcommand{\bU}{\mathbf{U}} 
\newcommand{\bV}{\mathbf{V}} 
\newcommand{\bB}{\mathbf{B}}     
\newcommand{\bD}{\mathbf{D}}     
\newcommand{\bE}{\mathbf{E}}     
\newcommand{\bH}{\mathbf{H}}     
\newcommand{\bJ}{\mathbf{J}}     
\newcommand{\bS}{\mathbf{S}}  
\newcommand{\bW}{\mathbf{W}}  
\newcommand{\bX}{\mathbf{X}}     
\newcommand{\bY}{\mathbf{Y}}    
\newcommand{\bn}{\mathbf{n}}     
\newcommand{\bx}{\mathbf{x}}     
\newcommand{\rmT}{\mathrm{T}} 
\newcommand{\rmS}{\mathrm{S}} 
\newcommand{\ci}{\mathbf{i}}     
\newcommand{\ds}{\displaystyle} 
\newcommand{\tr}{^{\mathrm{T}}}
\newcommand{\ie}{i.e.~}

\usepackage{hyperref}
\usepackage{amsfonts}
\usepackage{mathrsfs}
\usepackage{amssymb}
\usepackage{dsfont}
\usepackage{color}
%\usepackage[left=2.7cm,right=2cm,top=2.3cm,bottom=2.3cm,includefoot,includehead,headheight=13.6pt]{geometry}
\usepackage[hmargin=2.44cm,vmargin=2.2cm]{geometry}
\usepackage{graphics}
\usepackage{graphicx}

%
\def\R{\hbox{I\kern-0.2em\hbox{R}}}
\def\expg#1#2{{\vphantom{#1}}^{#2}\!#1}
\def\t#1{\expg{\!#1}{t}}

\def\restriction#1#2{\mathchoice
              {\setbox1\hbox{${\displaystyle #1}_{\scriptstyle #2}$}
              \restrictionaux{#1}{#2}}
              {\setbox1\hbox{${\textstyle #1}_{\scriptstyle #2}$}
              \restrictionaux{#1}{#2}}
              {\setbox1\hbox{${\scriptstyle #1}_{\scriptscriptstyle #2}$}
              \restrictionaux{#1}{#2}}
              {\setbox1\hbox{${\scriptscriptstyle #1}_{\scriptscriptstyle #2}$}
              \restrictionaux{#1}{#2}}}
\def\restrictionaux#1#2{{#1\,\smash{\vrule height .8\ht1 depth .85\dp1}}_{\,#2}}

\date{}

\begin{document}
{\huge {\sc {\centerline{Coupling AxiSEM and Specfem3D}} }}

\bigskip
\bigskip
\bigskip

\noindent {\bf{Authors:}} V. Monteiller, C. Durochat

\bigskip
\bigskip

\noindent NOTE : We have to finally choose which AxiSEM version we add and definitely use for coupling with SPECFEM3D.

\medskip

\noindent The paths present here are form specfem3d home directory.

\medskip

\noindent {\LARGE Here, \texttt{'run'}  is an alias for Curie defined as : \texttt{run='ccc\_msub -q standard'}.}

\bigskip

\section{STEP 1 : meshfem3D}

\begin{itemize}

\item[\textbullet] Go in 'EXAMPLES/coupling\_with\_EXTERNAL\_CODES/with\_AxiSEM/example\_1st\_for\_validation'

\item[\textbullet] Type : \\

\noindent ./clean\_this\_example\_dir.sh

\smallskip

\noindent run ./batch\_coupling\_step1\_create\_dir\_and\_paths\_+\_meshfem3d\_for\_AxiSEM.sh

\end{itemize}

\noindent It generate mesh files (on 1 proc on Curie), and copy files 'list\_gll\_boundary\_spherical.txt' and 'list\_gll\_boundary\_cartesian.txt' in AxiSEM solver dir

\medskip

\noindent Also, check Meshfem3D with parameters :

\begin{itemize}

\item[\textbullet] In the directory MESH/ :

\begin{itemize}

\item[\textbullet] ParFileMeshChunk  : parameters for the chunk meshing

\item[\textbullet] iasp91\_dsm or prem\_dsm : background model used in both AxiSEM ans Specfem3D

\end{itemize}

\item[\textbullet] In the Directory DATA/

\item[\textbullet] CMTSOLUTION  : with 0 in order to not use source inside

\item[\textbullet] coeff\_poly\_deg12 : to generate smooth intitial solution

\item[\textbullet] STATIONS  : stations files

\item[\textbullet] Par\_file : with

\item[\textbullet] SIMULATION\_TYPE=1

\item[\textbullet] SAVE\_FORWARD = .false.

\item[\textbullet] COUPLE\_WITH\_EXTERNAL\_CODE = .true

\item[\textbullet] EXTERNAL\_CODE\_TYPE    = 2

\end{itemize}

\section{STEP 2 : AxiSEM mesher}

\noindent Go in  : {\scriptsize  'utils/EXTERNAL\_CODES\_coupled\_with\_SPECFEM3D/AxiSEM\_for\_SPECFEM3D/AxiSEM\_modif\_for\_coupling\_with\_specfem/MESHER'}.

\medskip

\noindent Check the file 'inparam\_mesh'

\medskip

\noindent Type : ./submit.csh

\medskip

\noindent It run the mesher in serial : check file OUTPUT (wait for "....DONE WITH MESHER" to appear in OUTPUT)

\medskip

\noindent Type : ./movemesh.csh NAME\_OF\_DESTINATION\_MESH\_DIR

\medskip

\noindent It moves mesh files to ../SOLVER/MESHES/NAME\_OF\_DESTINATION\_MESH\_DIR

\medskip

\noindent For more details, see Axisem manual.

\section{STEP 3 : AxiSEM solver}

\noindent 2 files produced by meshfem3D (input\_box.txt and input\_box\_sem\_cart.txt) were copied (and the first line indicating the number of line to be read were add at the beginning of the files too) during the STEP 1 in the AxiSEM SOLVER dir.

\medskip

\noindent Go in {\scriptsize 'utils/EXTERNAL\_CODES\_coupled\_with\_SPECFEM3D/AxiSEM\_for\_SPECFEM3D/AxiSEM\_modif\_for\_coupling\_with\_specfem/SOLVER'}

\medskip

\noindent Check 'inparam\_basic' (set the value for MESHNAME to the meshname from above, among others), and also (differences with *.TEMPLATES provided
by AxiSEM) :

\begin{itemize}

\item[\textbullet] ATTENUATION         false 
\item[\textbullet] SAVE\_SNAPSHOTS     true

\end{itemize}

\medskip

\noindent Check 'inparam\_advanced' :

\begin{itemize}

\item[\textbullet] KERNEL\_WAVEFIELDS   true
\item[\textbullet] KERNEL\_IBEG         0 
\item[\textbullet] KERNEL\_IEND         4

\end{itemize}

\medskip

\noindent Type : ./sub\_launch\_script\_for\_run\_AxiSEM\_Curie\_CD.csh NAME\_OF\_RESULTS\_DIR

\smallskip

\noindent !! CAUTION : don't forget "NAME\_OF\_RESULTS\_DIR" (it will be your results folder, and a subfolder of SOLVER/) !!

\smallskip

\noindent It run, on N procs in batch on Curie, the compilation and the execution of  AxiSEM solver.
Among others, this script calls : \\

\noindent ./add\_line\_number\_in\_points\_lists.sh (rename files 'list\_gll\_boundary\_spherical.txt' \& \\ 'list\_gll\_boundary\_cartesian.txt', and add number of lines) \\

\noindent run ./sub\_called\_batch\_for\_AxiSEM\_Curie\_CD.sh (submit the job on Curie)

\section{STEP 4 : Specfem Partitionning ~ \& ~ STEP 5 : Specfem Generate database}

\noindent Go back in 'EXAMPLES/coupling\_with\_EXTERNAL\_CODES/with\_AxiSEM/example\_1st\_for\_validation'

\medskip

\noindent Type : run ./batch\_coupling\_step4\_and\_step5\_scotch\_part\_and\_generdata\_for\_AxiSEM.sh

\medskip

\noindent !! CAUTION : the number of procs in DATA/Par\_file, have to be the same as in the 'inparam\_mesh' (cf STEP 2) of AxiSEM !!

\section{Interface $\Longrightarrow$ STEP 6 : expand 2D to 3D ~ \& ~ STEP 7 : reformat}

\noindent Go in {\scriptsize 'utils/EXTERNAL\_CODES\_coupled\_with\_SPECFEM3D/AxiSEM\_for\_SPECFEM3D/AxiSEM\_modif\_for\_coupling\_with\_specfem/SOLVER'}

\medskip

\begin{itemize}

\item[\textbullet] Check the file 'expand\_2D\_3D.par', cf the paths in the 3 last lines !! ==> Caution : for the tractions path (Tractions/1/), the directory need to be created !! ==> it is created in the script 'sub\_launch\_script\_for\_make\_+\_run\_expand2D3D\_and\_reformat\_Curie\_CD.sh', but if you don't use this script, you have to create the directory yourself

\smallskip

\noindent Example : 
\begin{verbatim}

input_box.txt 
input_box_sem_cart.txt 
32                          # number of AxiSEM mpi processes used in solver  
90. 0.                      # source position (lat lon)  
0.  60.                     # chunk center (lat lon)    
1                           # number of axisem simus depends on moment tensor used  
32                          # number of Specfem3D MPI processes  
/[...]/example_1st_for_validation/MESH                           # Specfem MESH dir
/[...]/example_1st_for_validation/OUTPUT_FILES/DATABASES_MPI     # Specfem databases dir
/[...]/AxiSEM_modif_for_coupling_with_specfem/SOLVER/Tractions/1 # AxiSEM tractions dir

\end{verbatim}

\item[\textbullet] Check the file 'reformat.par'

\smallskip

\noindent Example : 
\begin{verbatim}

10.         # output sampling in Hz (time step that will use in Specfem3D simu) \newline
710. 1800.  # begin time and end time (s.) \newline

\end{verbatim}

\end{itemize}

\medskip

\noindent You can launch the compilation (of expand\_2D\_3D and reformat) AND the execution of this two programs with the single script :

\smallskip

\noindent ./sub\_launch\_script\_for\_make\_+\_run\_expand2D3D\_and\_reformat\_Curie\_CD.sh NAME\_OF\_RESULTS\_DIR

\medskip

\noindent !! CAUTION : don't forget "NAME\_OF\_RESULTS\_DIR" !!

\medskip

\noindent It will do the compilation, launch ./sub\_called\_batch\_for\_expand2D3D\_Curie\_CD.sh, and at the end, launch ./sub\_called\_batch\_for\_reformat\_Curie\_CD.sh $\Longleftarrow$ IMPORTANT : for the expand\_2D\_3D part, the number processes is arbitrary, BUT for the reformat part you have to configure the batch with the **SAME** number of processes that Specfem3D will use. Onefile is created by one process for one Spefem3D partion of domain.

\smallskip

\noindent All the results will be in 'NAME\_OF\_RESULTS\_DIR'.

\bigskip

\noindent If you prefer you do this step by step :

\begin{itemize}

\item[\textbullet] Go in {\scriptsize 'utils/EXTERNAL\_CODES\_coupled\_with\_SPECFEM3D/AxiSEM\_for\_SPECFEM3D/UTILS\_COUPLING\_SpecFEM'} to make the compilation.
\item[\textbullet] Go in NAME\_OF\_RESULTS\_DIR and type : run ../sub\_called\_batch\_for\_expand2D3D\_Curie\_CD.sh to submit expand\_2D\_3D
\item[\textbullet] Go in NAME\_OF\_RESULTS\_DIR and type : run ../sub\_called\_batch\_for\_reformat\_Curie\_CD.sh to submit reformat

\end{itemize}

\section{STEP 8 : Running specfem3D}

\section{STEP ? : Set up scripts}

\end{document}

### 2015-06-22 doc/USER_MANUAL/03_mesh_generation.tex
Buildbot ran the code beautifier.
renamed check_mesh_quality_CUBIT_Abaqus.f90 to check_mesh_quality.f90

@@
 turned on. It should be between 1 and 9000. In case no attenuation
-information is available, it can be set to zero. Please note that
-your Vp- and Vs-speeds are given for a reference frequency. To change
+information is available, it can be set to zero. You can either specify a single Q value,
+in which case it will be assumed to be pure shear attenuation $Q_{\mu}$, or two
+separate values for bulk and shear attenuation, $Q_{\kappa}$ and $Q_{\mu}$ respectively.
+
+Please note that your Vp- and Vs-speeds are given for a reference frequency. To change
 this reference frequency, you change the value of \texttt{ATTENUATION\_f0\_REFERENCE}
@@
 
-In directory "CUBIT\_GEOCUBIT" we provide a script that can help doing the above tasks of exporting a CUBIT mesh
+In directory "CUBIT\_GEOCUBIT/" we provide a script that can help doing the above tasks of exporting a CUBIT mesh
 to SPECFEM3D format automatically for you: "run\_cubit2specfem3d.py".
@@
 }
-
+\noindent
 (where "vol1" and "vol2" here represent the volume labels that you have set while creating the mesh in CUBIT)
@@
 \begin{verbatim}
-make xcheck_mesh_quality_CUBIT_Abaqus
+make xcheck_mesh_quality
 \end{verbatim}
@@
 \begin{verbatim}
-./bin/xcheck_mesh_quality_CUBIT_Abaqus
+./bin/xcheck_mesh_quality
 \end{verbatim}
@@
 \texttt{decompose\_mesh} depends on the SCOTCH library \citep{PeRo96},
-which provides efficent static mapping, graph and mesh partitioning
+which provides efficient static mapping, graph and mesh partitioning
 routines. SCOTCH is a free software package developed by Fran\c{c}ois
@@
 \begin{itemize}
-\item \texttt{Q} : quality factor (0=no attenuation)
+\item \texttt{Q} : quality factor (0=no attenuation) for shear attenuation $Q_{\mu}$
 \item \texttt{anisotropy\_flag} : 0=no anisotropy / 1,2,.. check with implementation

### 2015-06-22 doc/USER_MANUAL/07_adjoint_simulations.tex

@@
 \begin{enumerate}
-\item First, run a regular forward simlation (\texttt{SIMULATION\_TYPE =
+\item First, run a regular forward simulation (\texttt{SIMULATION\_TYPE =
 1} and \texttt{SAVE\_FORWARD = .false.}). You can automatically set
@@
 setting \texttt{ANISOTROPIC\_KL} \texttt{=} \texttt{.true.} in \texttt{constants.h}
-the package will save anisotropic kernels paramterized as velocities
+the package will save anisotropic kernels parameterized as velocities
 related to transverse isotropy based on the the Chen and Tromp parameters

### 2015-06-22 doc/USER_MANUAL/09_noise_simulations.tex

@@
 \item As usual, we first configure the software package using:
-
-\texttt{./configure FC=ifort MPIFC=mpif90}
+\begin{verbatim}
+./configure FC=ifort MPIFC=mpif90
+\end{verbatim}
 
 Use the following if SCOTCH is needed:
-
-\texttt{./configure FC=ifort MPIFC=mpif90 -{}-with-scotch-dir=/opt/scotch}\\
+\begin{verbatim}
+./configure FC=ifort MPIFC=mpif90 --with-scotch-dir=/opt/scotch
+\end{verbatim}
 
 \item Next, we need to compile the source code using:
-
-\texttt{make xgenerate\_databases}
-
-\texttt{make xspecfem3D} \\
+\begin{verbatim}
+make xgenerate_databases
+make xspecfem3D
+\end{verbatim}
 
@@
 statistics, e.g., the ensemble-averaged noise spectrum. Matlab scripts
-are provided to help you to generate the necessary file.
-
+are provided to help you to generate the necessary file:
 \texttt{EXAMPLES/noise\_tomography/NOISE\_TOMOGRAPHY.m} (main program)\\
@@
 In Matlab, simply run:
-
-\texttt{NOISE\_TOMOGRAPHY(NSTEP, DT, Tmin, Tmax, NOISE\_MODEL)}\\
+\begin{verbatim}
+NOISE_TOMOGRAPHY(NSTEP, DT, Tmin, Tmax, NOISE_MODEL)
+\end{verbatim}
 
@@
 (plus a figure in Matlab):
-
-\texttt{{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}}\\
- \texttt{the source time function has been saved in:}\\
- \texttt{/data2/yangl/3D\_NOISE/S\_squared} (note this path must be
-different)\\
- \texttt{S\_squared should be put into directory:}\\
- \texttt{OUTPUT\_FILES/NOISE\_TOMOGRAPHY/ in the SPECFEM3D Cartesian
-package}\\
+\begin{verbatim}
+*************************************************************
+the source time function has been saved in:
+..../S_squared (note this path must be different)
+S_squared should be put into directory:
+OUTPUT_FILES/NOISE_TOMOGRAPHY/ in the SPECFEM3D Cartesian package
+\end{verbatim}
 
@@
 \begin{itemize}
-\item Step 1: simulation for generating wavefields\\
- \texttt{SIMULATION\_TYPE=1}\\
- \texttt{NOISE\_TOMOGRAPHY=1}\\
- \texttt{SAVE\_FORWARD} not used, can be either .true. or .false.
-\item Step 2: simulation for ensemble forward wavefields\\
- \texttt{SIMULATION\_TYPE=1}\\
- \texttt{NOISE\_TOMOGRAPHY=2}\\
- \texttt{SAVE\_FORWARD=.true.}
+\item Step 1: simulation for generating wavefields
+\begin{verbatim}
+SIMULATION_TYPE = 1
+NOISE_TOMOGRAPHY = 1
+SAVE_FORWARD (not used, can be either .true. or .false.)
+\end{verbatim}
+
+\item Step 2: simulation for ensemble forward wavefields
+\begin{verbatim}
+SIMULATION_TYPE = 1
+NOISE_TOMOGRAPHY = 2
+SAVE_FORWARD = .true.
+\end{verbatim}
+
 \item Step 3: simulation for ensemble adjoint wavefields and sensitivity
-kernels\\
- \texttt{SIMULATION\_TYPE=3}\\
- \texttt{NOISE\_TOMOGRAPHY=3}\\
- \texttt{SAVE\_FORWARD=.false.}
+kernels
+\begin{verbatim}
+SIMULATION_TYPE = 3
+NOISE_TOMOGRAPHY = 3
+SAVE_FORWARD = .false.
+\end{verbatim}
 

### 2015-06-22 doc/USER_MANUAL/15_informations_for_developers.tex

@@
 
-You can get a very simple 1D version of a demo code (there is one in Fortran and one in Python):\\
-
-git clone --recursive https://github.com/geodynamics/specfem1d.git\\
+You can get a very simple 1D version of a demo code (there is one in Fortran and one in Python):
+\begin{verbatim}
+git clone --recursive https://github.com/geodynamics/specfem1d.git
+\end{verbatim}
 
@@
 \noindent For information on how to contribute to the code, i.e., for how to make your modifications, additions or improvements part of the
-official package, see https://github.com/geodynamics/specfem3d/wiki .
+official package, see \url{https://github.com/geodynamics/specfem3d/wiki} .

### 2015-06-22 doc/USER_MANUAL/B_channel_codes.tex

@@
 sampling properties and component configurations. There are standards
-to name channel codes depending on instrument properties. IRIS \texttt{(www.iris.edu)}
+to name channel codes depending on instrument properties. IRIS (\url{http://www.iris.edu})
 uses SEED format for channel codes, which are represented by three
@@
 of the FDSN convention used by IRIS, and how it is adopted in SEM
-seismograms. Please visit \texttt{www.iris.edu/manuals/SEED\_appA.htm}
+seismograms. Please visit \url{http://www.iris.edu/manuals/SEED_appA.htm}
 for further information.\\
@@
 \caption{The band code convention is based on the sampling rate and the response
-band of instruments. Please visit \texttt{www.iris.edu/manuals/SEED\_appA.htm}
+band of instruments. Please visit \url{http://www.iris.edu/manuals/SEED_appA.htm}
 for further information. Grey rows show the relative band-code range

### 2015-06-22 doc/USER_MANUAL/authors.tex

\section*{Authors}

\noindent The SPECFEM3D Cartesian package was first developed by Dimitri
Komatitsch and Jean-Pierre Vilotte at Institut de Physique du Globe
(IPGP) in Paris, France from 1995 to 1997 and then by Dimitri Komatitsch
and Jeroen Tromp at Harvard University and Caltech, USA, starting in 1998.
The story started on April 4, 1995, when Prof. Yvon Maday from CNRS and University of Paris, France, gave a lecture to
Dimitri Komatitsch and Jean-Pierre Vilotte at IPG about the nice properties of the Legendre spectral-element method with diagonal mass matrix that he had used for
other equations. We are deeply indebted and thankful to him for that.
That followed a visit by Dimitri Komatitsch to OGS (Istituto Nazionale di Oceanografia e di Geofisica Sperimentale) in Trieste, Italy, in February 1995
to meet with G\'eza Seriani and Enrico Priolo, who introduced him to their 2D Chebyshev version of the spectral-element method with a non-diagonal mass matrix.
We are deeply indebted and thankful to them for that.

%%%%%%%% please do NOT merge the lines of the paragraph below, it makes it difficult to insert new names in alphabetical order; leave a single name per line
%%%%%%%% please do NOT merge the lines of the paragraph below, it makes it difficult to insert new names in alphabetical order; leave a single name per line
%%%%%%%% please do NOT merge the lines of the paragraph below, it makes it difficult to insert new names in alphabetical order; leave a single name per line
Since then it has been developed and maintained by a development team:
in alphabetical order,
Jean-Paul (Pablo) Ampuero,
Kangchen Bai,
Piero Basini,
C\'eline Blitz,
Alexis Bottero,
Ebru Bozda\u{g},
Emanuele Casarotti,
Joseph Charles,
Min Chen,
Paul Cristini,
Percy Galvez,
Dominik G\"oddeke,
Vala Hj\"orleifsd\'ottir,
Sue Kientz,
Dimitri Komatitsch,
Jes\'us Labarta,
Nicolas Le Goff,
Pieyre Le Loher,
Matthieu Lefebvre,
Qinya Liu,
Yang Luo,
Alessia Maggi,
Federica Magnoni,
Roland Martin,
Ren\'e Matzen,
Dennis McRitchie,
Matthias Meschede,
Peter Messmer,
David Mich\'ea,
Surendra Nadh Somala,
Tarje Nissen-Meyer,
Daniel Peter,
Max Rietmann,
Elliott Sales de Andrade,
Brian Savage,
Bernhard Schuberth,
Anne Sieminski,
Leif Strand,
Carl Tape,
Jeroen Tromp,
Jean-Pierre Vilotte,
Zhinan Xie,
Chang-Hua Zhang,
Hejun Zhu.\\

The cover graphic of the manual was created by Santiago Lombeyda from
Caltech's Center for Advanced Computing Research (CACR) \urlwithparentheses{http://www.cacr.caltech.edu}.\\

The code is released open-source under the GNU version 2+ license, see the license at the end of this manual.\\

### 2015-06-22 doc/USER_MANUAL/sponsors.tex

\section*{Current and past main participants or main sponsors of the SPECFEM project (in no particular order)}
%
\begin{figure}[htbp]
\noindent \begin{centering}
\includegraphics[width=0.125\textwidth]{figures/logo_cnrs.png}\vspace*{2truemm}
\includegraphics[width=0.125\textwidth]{figures/logo_princeton.jpg}\vspace*{2truemm}
\includegraphics[width=0.15\textwidth]{figures/logo_aix_marseille_universite.png}\vspace*{0.02truemm}
\includegraphics[width=0.15\textwidth]{figures/logo_ETH.jpg}\vspace*{2truemm}
\includegraphics[width=0.12\textwidth]{figures/logo_CSC_China.jpg}\vspace*{0.02truemm}
\includegraphics[width=0.15\textwidth]{figures/logo_inria.jpg}\vspace*{2truemm}
\includegraphics[width=0.14\textwidth]{figures/logo_UPPA.png}
\par\end{centering}
%
\vspace*{-2truemm}
%
\noindent \begin{centering}
\includegraphics[width=0.15\textwidth]{figures/logo_NVIDIA.jpg}\vspace*{2truemm}
\includegraphics[width=0.125\textwidth]{figures/logo_IUF.jpg}\vspace*{2truemm}
\includegraphics[width=0.135\textwidth]{figures/logo_Caltech.png}\vspace*{2truemm}
\includegraphics[width=0.135\textwidth]{figures/logo_Harvard.jpg}\vspace*{2truemm}
\includegraphics[width=0.135\textwidth]{figures/logo_IPGP.jpg}\vspace*{2truemm}
\includegraphics[width=0.112\textwidth]{figures/logo_ANR.png}\vspace*{2truemm}
\includegraphics[width=0.112\textwidth]{figures/logo_NSF.png}\vspace*{2truemm}
\par\end{centering}
%
\vspace*{-2truemm}
%
\noindent \begin{centering}
\includegraphics[width=0.190\textwidth]{figures/logo_European_Union.png}\vspace*{2truemm}
\includegraphics[width=0.112\textwidth]{figures/logo_GENCI.jpg}\vspace*{2truemm}
\includegraphics[width=0.112\textwidth]{figures/logo_PRACE.jpg}\vspace*{2truemm}
\includegraphics[width=0.112\textwidth]{figures/logo_CINES.png}\vspace*{2truemm}
\includegraphics[width=0.130\textwidth]{figures/logo_Oak_Ridge.png}\vspace*{2truemm}
\hspace*{3mm}\includegraphics[width=0.130\textwidth]{figures/logo_fondation_Del_Duca.png}
\includegraphics[width=0.130\textwidth]{figures/logo_CIG.png}\vspace*{2truemm}
\par\end{centering}
%
\vspace*{-2truemm}
%
\noindent \begin{centering}
\includegraphics[width=0.112\textwidth]{figures/logo_University_of_Toronto.jpg}\vspace*{2truemm}
\includegraphics[width=0.109\textwidth]{figures/logo_INGV.jpg}\vspace*{2truemm}
\includegraphics[width=0.102\textwidth]{figures/logo_Univ_Toulouse.jpg}\vspace*{2truemm}
\includegraphics[width=0.112\textwidth]{figures/logo_TOTAL.jpg}\vspace*{2truemm}
\includegraphics[width=0.130\textwidth]{figures/logo_Fairbanks.jpg}\vspace*{2truemm}
\includegraphics[width=0.120\textwidth]{figures/logo_CINECA.jpg}\vspace*{2truemm}
\includegraphics[width=0.140\textwidth]{figures/logo_Intel_Exascale_Labs.png}\vspace*{2truemm}
\includegraphics[width=0.130\textwidth]{figures/logo_Maison_Simulation.jpg}
\par\end{centering}
\end{figure}

### 2015-06-29 utils/EXTERNAL_CODES_coupled_with_SPECFEM3D/AxiSEM_for_SPECFEM3D/Manual_coupling_AxiSEM_Specfem3D.tex
Final step of coupling w AxiSEM OK (left to check conindence of sismos) + edit coupling Manual

@@
 /[...]/example_1st_for_validation/OUTPUT_FILES/DATABASES_MPI    # Specfem databases dir
-/[...]/AxiSEM_modif_for_coupling_with_specfem/SOLVER/Tractions/1 # AxiSEM tractions dir
+/[...]/example_1st_for_validation/DATA/AxiSEM_tractions/1       # AxiSEM tractions dir
 
@@
 
-10.         # output sampling in Hz (time step that will use in Specfem3D simu) \newline
-710. 1800.  # begin time and end time (s.) \newline
+10.         # = 1/DT, i.e. output sampling in Hz (time step that will use in Specfem3D simu) 
+710. 1800.  # begin time and end time (s.)
 
@@
 
+\noindent Go in 'EXAMPLES/coupling\_with\_EXTERNAL\_CODES/with\_AxiSEM/example\_1st\_for\_validation'
+
+\medskip
+
+\noindent Check in 'DATA/Par\_file', the lines after the comment "\# time step parameters", these parameters must coincide with the parameters in  'reformat.par' (cf STEP 7) :
+
+\begin{itemize}
+
+\item[\textbullet] NSTEP = 10900 ~~ \# $\Longrightarrow$ cf the two values in the second line of 'reformat.par'. This value equals to : 
+$$\ds \dfrac{\hbox{second value} - \hbox{the first}}{\text{DT}} ~~ \left(= \dfrac{\hbox{duration of physical time}}{\text{DT}} \right)$$
+
+\noindent So here, it is : $\ds \dfrac{1800 - 710}{0.1} = 10900$
+
+\medskip
+
+\item[\textbullet] DT = 0.1 ~~ \# $\Longrightarrow$ the inverse of the first line in 'reformat.par' (which is 1/DT)
+
+\end{itemize}
+
+\medskip
+
+\noindent Also check in 'DATA/Par\_file', the lines after the comment "\# to couple with an external code (such as DSM, AxiSEM, or FK)" :
+
+\begin{itemize}
+
+\item[\textbullet] COUPLE\_WITH\_EXTERNAL\_CODE = .true. ~~ \# $\Longrightarrow$ always set it to '.true.' in this case
+
+\item[\textbullet] EXTERNAL\_CODE\_TYPE = 2  ~ \# 1 = DSM, 2 = AxiSEM, 3 = FK ~~ $\Longrightarrow$ always set it to '2' for AxiSEM
+
+\item[\textbullet] TRACTION\_PATH  = ./DATA/AxiSEM\_tractions/1/ ~~   \# $\Longrightarrow$ This is where there are the tractions from AxiSEM (from steps 6 and 7) !!
+CAUTION : verify that they are in this dir, in the form of files 'proc0000**\_sol\_axisem', etc.
+!!
+
+\item[\textbullet] MESH\_A\_CHUNK\_OF\_THE\_EARTH  = .true.  ~~  \# $\Longrightarrow$ always set it to '.true.' in this case
+
+\end{itemize}
+
+\medskip
+
+\noindent And finally, type : run ./batch\_coupling\_step8\_xspecfem3d.sh
+
 \section{STEP ? : Set up scripts}

### 2015-07-01 utils/EXTERNAL_CODES_coupled_with_SPECFEM3D/AxiSEM_for_SPECFEM3D/Manual_coupling_AxiSEM_Specfem3D.tex
Buildbot ran the code beautifier.
Modif in AxiSEM/specfem coupling manual

@@
 
-\item[\textbullet] Go in 'EXAMPLES/coupling\_with\_EXTERNAL\_CODES/with\_AxiSEM/example\_1st\_for\_validation'
+\item[\textbullet] Go in 'EXAMPLES/coupling\_with\_EXTERNAL\_CODES/with\_AxiSEM/example\_1st\_for\_validation',
+
+\smallskip
+
+(OR, create another subdirectory in '[...]/with\_AxiSEM', copy files from '[...]/example\_1st\_for\_validation' and modify them if you need)
+
+\item[\textbullet] {\color{red} CAUTION !!
+: Check the file 'paths\_for\_coupling\_SPECFEM3D\_AxiSEM.sh' and modify the variable 'HOME\_SPECFEM3D'} $\Longrightarrow$ this is the only path you have to modify for specfem3d
 
@@
 
-\item[\textbullet] Check the file 'expand\_2D\_3D.par', cf the paths in the 3 last lines !!
-==> Caution : for the tractions path (Tractions/1/), the directory need to be created !!
-==> it is created in the script 'sub\_launch\_script\_for\_make\_+\_run\_expand2D3D\_and\_reformat\_Curie\_CD.sh', but if you don't use this script, you have to create the directory yourself
+\item[\textbullet] {\color{red} Check the file 'expand\_2D\_3D.par', cf THE PATHS IN THE 3 LAST LINES (normally, these are the only paths you to modify in the AxiSEM dir) !!} ==> Caution : for the tractions path (Tractions/1/), the directory need to be created !!
+==>
+
+it is created in the script 'sub\_launch\_script\_for\_make\_+\_run\_expand2D3D\_and\_reformat\_Curie\_CD.sh', but if you don't use this script, you have to create the directory yourself
 

### 2015-07-13 doc/USER_MANUAL/authors.tex
merged the SPECFEM3D and SPECFEM3D_GLOBE list of authors (too difficult to distinguish them, and does not make much sense)

@@
 
-\noindent The SPECFEM3D Cartesian package was first developed by Dimitri
+\noindent The SPECFEM3D package was first developed by Dimitri
 Komatitsch and Jean-Pierre Vilotte at Institut de Physique du Globe
@@
 %%%%%%%% please do NOT merge the lines of the paragraph below, it makes it difficult to insert new names in alphabetical order; leave a single name per line
-Since then it has been developed and maintained by a development team:
-in alphabetical order,
+Since then it has been developed and maintained by a development team: in alphabetical order,
+Michael Afanasiev,
 Jean-Paul (Pablo) Ampuero,
@@
 Paul Cristini,
+Cl\'ement Durochat,
 Percy Galvez,
@@
 Qinya Liu,
+David Luet,
 Yang Luo,
@@
 David Mich\'ea,
+Vadim Monteiller,
 Surendra Nadh Somala,
@@
 Daniel Peter,
+Kevin Pouget,
 Max Rietmann,
@@
 Anne Sieminski,
+James Smith,
 Leif Strand,
@@
 Jeroen Tromp,
+Brice Videau,
 Jean-Pierre Vilotte,
@@
 The cover graphic of the manual was created by Santiago Lombeyda from
-Caltech's Center for Advanced Computing Research (CACR) \urlwithparentheses{http://www.cacr.caltech.edu}.\\
+Caltech's Center for Advanced Computing Research (CACR), USA, with free satellite clipart pictures
+from \url{http://www.4vector.com} and \url{http://www.clker.com} added to it.\\
 

### 2015-07-15 utils/EXTERNAL_CODES_coupled_with_SPECFEM3D/AxiSEM_for_SPECFEM3D/Manual_coupling_AxiSEM_Specfem3D.tex
Buildbot ran the code beautifier.
Change Axisem coupling manual

@@
 
-\noindent It generate mesh files (on 1 proc on Curie), and copy files 'list\_gll\_boundary\_spherical.txt' and 'list\_gll\_boundary\_cartesian.txt' in AxiSEM solver dir
+\noindent It generate mesh files (on 1 proc on Curie), and copy files 'list\_gll\_boundary\_spherical.txt' and 'list\_gll\_boundary\_cartesian.txt' in AxiSEM solver dir.
 
+\bigskip
 \medskip
 
-\noindent Also, check Meshfem3D with parameters :
+\noindent Also, check Meshfem3D with parameters $\Longrightarrow$ In the directory MESH/ :
 
@@
 
-\item[\textbullet] In the directory MESH/ :
+\item[\textbullet] 'ParFileMeshChunk'  : parameters for the chunk meshing
+
+\item[\textbullet] 'iasp91\_dsm' or 'prem\_dsm' : background model used in both AxiSEM ans Specfem3D
+
+\end{itemize}
+
+\bigskip
+
+\noindent In the Directory DATA/ :
 
@@
 
-\item[\textbullet] ParFileMeshChunk  : parameters for the chunk meshing
+\item[\textbullet] 'CMTSOLUTION'  : with 0 in order to not use source inside
 
-\item[\textbullet] iasp91\_dsm or prem\_dsm : background model used in both AxiSEM ans Specfem3D
+\item[\textbullet] 'coeff\_poly\_deg12' : to generate smooth intitial solution
 
+\item[\textbullet] 'STATIONS'  : stations files
+
+\item[\textbullet] Set the file 'Par\_file' with these parmeters :
+
+\begin{itemize}
+\item[\textbullet] SIMULATION\_TYPE=1
+\item[\textbullet] SAVE\_FORWARD = .false.
+\item[\textbullet] COUPLE\_WITH\_EXTERNAL\_CODE = .true
+\item[\textbullet] EXTERNAL\_CODE\_TYPE    = 2
 \end{itemize}
-
-\item[\textbullet] In the Directory DATA/
-
-\item[\textbullet] CMTSOLUTION  : with 0 in order to not use source inside
-
-\item[\textbullet] coeff\_poly\_deg12 : to generate smooth intitial solution
-
-\item[\textbullet] STATIONS  : stations files
-
-\item[\textbullet] Par\_file : with
-
-\item[\textbullet] SIMULATION\_TYPE=1
-
-\item[\textbullet] SAVE\_FORWARD = .false.
-
-\item[\textbullet] COUPLE\_WITH\_EXTERNAL\_CODE = .true
-
-\item[\textbullet] EXTERNAL\_CODE\_TYPE    = 2
 

### 2015-07-16 utils/EXTERNAL_CODES_coupled_with_SPECFEM3D/AxiSEM_for_SPECFEM3D/Manual_coupling_AxiSEM_Specfem3D.tex
Fix big in get_model for coupling w Axisem + manual correction

@@
 
-\noindent It generate mesh files (on 1 proc on Curie), and copy files 'list\_gll\_boundary\_spherical.txt' and 'list\_gll\_boundary\_cartesian.txt' in AxiSEM solver dir.
+\noindent It generates mesh files (on 1 proc on Curie), and copies files 'list\_gll\_boundary\_spherical.txt' and 'list\_gll\_boundary\_cartesian.txt' in AxiSEM solver dir.
 
@@
 
-\item[\textbullet] 'iasp91\_dsm' or 'prem\_dsm' : background model used in both AxiSEM ans Specfem3D
+\item[\textbullet] 'iasp91\_dsm' or 'prem\_dsm' : background model used in both AxiSEM and Specfem3D
 
@@
 
-\item[\textbullet] Set the file 'Par\_file' with these parmeters :
+\item[\textbullet] Set the file 'Par\_file' with these parameters :
 

### 2015-07-16 utils/EXTERNAL_CODES_coupled_with_SPECFEM3D/AxiSEM_for_SPECFEM3D/Manual_coupling_AxiSEM_Specfem3D.tex
Modif in coupling w AxiSEM manual

@@
 
-\item[\textbullet] {\color{red} Check the file 'expand\_2D\_3D.par', cf THE PATHS IN THE 3 LAST LINES (normally, these are the only paths you to modify in the AxiSEM dir) !!} ==> Caution : for the tractions path (Tractions/1/), the directory need to be created !!
-==>
+\item[\textbullet] {\color{red} Check the file 'expand\_2D\_3D.par', cf THE PATHS IN THE 3 LAST LINES (normally, these are the only paths you to modify in the AxiSEM dir) !!} ==> Caution : for the tractions path (for example: /[...]/AxiSEM\_tractions/1/), the directory need to be created !!!!
+{\color{red} ~ ==>}
 
-it is created in the script 'sub\_launch\_script\_for\_make\_+\_run\_expand2D3D\_and\_reformat\_Curie\_CD.sh', but if you don't use this script, you have to create the directory yourself
+the script 'sub\_launch\_script\_for\_make\_+\_run\_expand2D3D\_and\_reformat\_Curie\_CD.sh' create a traction directory here:
+
+\smallskip
+
+{\footnotesize 'EXAMPLES/coupling\_with\_EXTERNAL\_CODES/with\_AxiSEM/example\_1st\_for\_validation/DATA/AxiSEM\_tractions/1/', }
+
+BUT, if you don't use this script or if you want to create another tractions dir, you have to create the directory yourself.
 

### 2015-07-23 utils/EXTERNAL_CODES_coupled_with_SPECFEM3D/AxiSEM_for_SPECFEM3D/Manual_coupling_AxiSEM_Specfem3D.tex
Buildbot ran the code beautifier.
Final modifs in Axisem dir: 1

@@
 
-\noindent 2 files produced by meshfem3D (input\_box.txt and input\_box\_sem\_cart.txt) were copied (and the first line indicating the number of line to be read were add at the beginning of the files too) during the STEP 1 in the AxiSEM SOLVER dir.
+\noindent 2 files produced by meshfem3D ('list\_gll\_boundary\_spherical.txt' and 'list\_gll\_boundary\_cartesian.txt') were copied during the STEP 1 in the AxiSEM SOLVER dir.
 
@@
 
-\noindent ./add\_line\_number\_in\_points\_lists.sh (rename files 'list\_gll\_boundary\_spherical.txt' \& \\ 'list\_gll\_boundary\_cartesian.txt', and add number of lines) \\
+\noindent ./add\_line\_number\_in\_points\_lists.sh : rename files 'list\_gll\_boundary\_spherical.txt' and \\ 'list\_gll\_boundary\_cartesian.txt' to 'input\_box.txt' and 'input\_box\_sem\_cart.txt', resp.
+and add number of lines at the beginning of the two files too.
+\\
 
-\noindent run ./sub\_called\_batch\_for\_AxiSEM\_Curie\_CD.sh (submit the job on Curie)
+\noindent run ./sub\_called\_batch\_for\_AxiSEM\_Curie\_CD.sh : submit the job on Curie.
 

### 2015-08-06 utils/EXTERNAL_CODES_coupled_with_SPECFEM3D/AxiSEM_for_SPECFEM3D/Manual_coupling_AxiSEM_Specfem3D.tex
Final modifs in Axisem dir: 1 (big and last commit about this)
Buildbot ran the code beautifier.
Modif manual of coupling w Axisem

@@
 
-\noindent {\bf{Authors:}} V. Monteiller, C. Durochat
+\noindent {\bf{Authors:}} C. Durochat, V. Monteiller
 
@@
 \noindent {\LARGE Here, \texttt{'run'}  is an alias for Curie defined as : \texttt{run='ccc\_msub -q standard'}.}
+
+\medskip
+
+\noindent We show the process of the coupling, for the example:
+
+\noindent 'EXAMPLES/coupling\_with\_EXTERNAL\_CODES/with\_AxiSEM/example\_1st\_for\_validation', see the file 'README\_example\_details' for some particulars of this example.
+
+\medskip
+
+\noindent There are other examples in 'EXAMPLES/coupling\_with\_EXTERNAL\_CODES/with\_AxiSEM/' with other set of parameters.
 
@@
 \item[\textbullet] KERNEL\_IEND         4
+\item[\textbullet] KERNEL\_DUMPTYPE     coupling\_box
 
@@
 \smallskip
+
+\noindent (or 'sub\_launch\_script\_for\_run\_AxiSEM\_Curie\_CD\_VM.csh', it depends on examples)
+
+\medskip
 
@@
 
-\noindent run ./sub\_called\_batch\_for\_AxiSEM\_Curie\_CD.sh : submit the job on Curie.
+\noindent run ./sub\_called\_batch\_for\_AxiSEM\_Curie\_CD\textit{(\_VM)}.sh : submit the job on Curie.
 
@@
 
-\noindent Go in {\scriptsize 'utils/EXTERNAL\_CODES\_coupled\_with\_SPECFEM3D/AxiSEM\_for\_SPECFEM3D/AxiSEM\_modif\_for\_coupling\_with\_specfem/SOLVER'}
+\noindent Go in {\scriptsize 'utils/EXTERNAL\_CODES\_coupled\_with\_SPECFEM3D/AxiSEM\_for\_SPECFEM3D/UTILS\_COUPLING\_SpecFEM'} and make the compilation (make clean ; make all).
+
+\bigskip
+
+\noindent Go in {\scriptsize 'utils/EXTERNAL\_CODES\_coupled\_with\_SPECFEM3D/AxiSEM\_for\_SPECFEM3D/AxiSEM\_modif\_for\_coupling\_with\_specfem/SOLVER'} :
 
@@
 
-\item[\textbullet] {\color{red} Check the file 'expand\_2D\_3D.par', cf THE PATHS IN THE 3 LAST LINES (normally, these are the only paths you to modify in the AxiSEM dir) !!} ==> Caution : for the tractions path (for example: /[...]/AxiSEM\_tractions/1/), the directory need to be created !!!!
-{\color{red} ~ ==>}
-
-the script 'sub\_launch\_script\_for\_make\_+\_run\_expand2D3D\_and\_reformat\_Curie\_CD.sh' create a traction directory here:
-
-\smallskip
-
-{\footnotesize 'EXAMPLES/coupling\_with\_EXTERNAL\_CODES/with\_AxiSEM/example\_1st\_for\_validation/DATA/AxiSEM\_tractions/1/', }
-
-BUT, if you don't use this script or if you want to create another tractions dir, you have to create the directory yourself.
+\item[\textbullet] {\color{red} Check the file 'expand\_2D\_3D.par', cf THE PATHS IN THE 3 LAST LINES (normally, these are the only paths you to modify in the AxiSEM dir) !!} ==> Caution : for the tractions path (for example: /[...]/AxiSEM\_tractions/1/), the directory need to be created !!!
 
@@
 \begin{verbatim}
-
 input_box.txt
@@
 /[...]/example_1st_for_validation/DATA/AxiSEM_tractions/1       # AxiSEM tractions dir
-
 \end{verbatim}
 
-\item[\textbullet] Check the file 'reformat.par'
+\item[\textbullet] Go in NAME\_OF\_RESULTS\_DIR and type : run ../sub\_called\_batch\_for\_expand2D3D\_Curie\_CD.sh to submit expand\_2D\_3D (the number processes is here arbitrary)
+
+\item[\textbullet] At the end of expand\_2D\_3D, check the file 'StaLta.txt' to see at what time, the seismos begin to be non-zero.
+With this information, you can {\color{red} configure} begin time (corresponding to a little before the time when the values in 'StaLta.txt' begin to be non-zero) {\color{red} in the file 'reformat.par'}:
 
@@
 \begin{verbatim}
+10.         # = 1/DT, i.e. output sampling in Hz (time step that will use in specfem simu)
+710. 1800.  # begin time and end time (s.) (corresponding to the SpecFEM simulation time)
+\end{verbatim}
 
-10.         # = 1/DT, i.e. output sampling in Hz (time step that will use in Specfem3D simu)
-710. 1800.  # begin time and end time (s.)
-
-\end{verbatim}
+\item[\textbullet] Stay in NAME\_OF\_RESULTS\_DIR and type : run ../sub\_called\_batch\_for\_reformat\_Curie\_CD.sh to submit reformat.
+IMPORTANT : you have to configure the batch with the **SAME** number of processes that Specfem3D will use.
+One file is created by one process for one Spefem3D partition of domain.
 
@@
 
+\bigskip
+
+\noindent {\color{red} NOTE:} If you already know the value of the begin time and if it is already configured in 'reformat.par', you can launch the compilation of expand\_2D\_3D and reformat, AND the execution of this two programs with the single script :
+
 \medskip
-
-\noindent You can launch the compilation (of expand\_2D\_3D and reformat) AND the execution of this two programs with the single script :
-
-\smallskip
 
@@
 
-\medskip
+\bigskip
 
-\noindent It will do the compilation, launch ./sub\_called\_batch\_for\_expand2D3D\_Curie\_CD.sh, and at the end, launch ./sub\_called\_batch\_for\_reformat\_Curie\_CD.sh $\Longleftarrow$ IMPORTANT : for the expand\_2D\_3D part, the number processes is arbitrary, BUT for the reformat part you have to configure the batch with the **SAME** number of processes that Specfem3D will use.
-Onefile is created by one process for one Spefem3D partion of domain.
+\noindent It will do the compilation, create a traction directory here:
+
+\noindent {\footnotesize 'EXAMPLES/coupling\_with\_EXTERNAL\_CODES/with\_AxiSEM/example\_1st\_for\_validation/DATA/AxiSEM\_tractions/1/'},
 
@@
 
-\noindent All the results will be in 'NAME\_OF\_RESULTS\_DIR'.
+\noindent AND launch ./sub\_called\_batch\_for\_expand2D3D\_Curie\_CD.sh, AND at the end, launch \\
+\noindent ./sub\_called\_batch\_for\_reformat\_Curie\_CD.sh
+
+\medskip
+
+\noindent {\color{red} BUT}, if you don't use this script or if you want to create another tractions dir, you have to create the directory yourself and launch the complication and the two run separately.
 
@@
 
-\noindent If you prefer you do this step by step :
-
-\begin{itemize}
-
-\item[\textbullet] Go in {\scriptsize 'utils/EXTERNAL\_CODES\_coupled\_with\_SPECFEM3D/AxiSEM\_for\_SPECFEM3D/UTILS\_COUPLING\_SpecFEM'} to make the compilation.
-\item[\textbullet] Go in NAME\_OF\_RESULTS\_DIR and type : run ../sub\_called\_batch\_for\_expand2D3D\_Curie\_CD.sh to submit expand\_2D\_3D
-\item[\textbullet] Go in NAME\_OF\_RESULTS\_DIR and type : run ../sub\_called\_batch\_for\_reformat\_Curie\_CD.sh to submit reformat
-
-\end{itemize}
+\noindent All the results will be in 'NAME\_OF\_RESULTS\_DIR'.
 

### 2015-08-10 utils/EXTERNAL_CODES_coupled_with_SPECFEM3D/AxiSEM_for_SPECFEM3D/AxiSEM_modif_for_coupling_with_specfem/MANUAL/theoretical_foundations.tex
Add NEW VERSION of AxiSEM

\section{Theoretical foundations}
The following sections are excerpts from papers (1)-(4), and not necessarily coherent.

\subsection{Elastodynamic equations of motion: weak form}
For the remainder of this paper, we will treat a non-rotating, 
non-gravitating, isotropic (Lam\'{e} parameters $\lambda$, $\mu$)
earth model. The 3-D elastodynamic weak form comprises the following
three integral equalities for the displacement-potential vector 
$(\bu,\chi)$ in an appropriate space with square-integrable derivatives
for all admissible test or trial functions $(\bw,w)$ \citep{nissen+:07a}:
%
\eqa \label{eq:weak1}
\lefteqn{0=\int_{\rm SIC}\rho\hspace{0.1em}\bw\cdot\ddot{\bu}\,d^3\bx
-\int_{\rm ICB}(\brh\cdot\bw) \ddot{\chi}\,d^2\bx +
}
\nonumber\\
\lefteqn{\hspace{1.5em}
\int_{\rm SIC}\left\{\lambda(\bdel\cdot\bw)
(\bdel\cdot\bu)
+\mu\bdel\bw\!:\!\left[\bdel\bu+\left(\bdel\bu\right)^T\right]\right\}
\,d^3\bx,
}
\ena
%
\eqa \label{eq:weak2}
\lefteqn{0=\int_{\rm FOC}\kappa^{-1}w\hspace{0.1em} \ddot{\chi}\,d^3\bx
+\int_{\rm ICB}w(\brh\cdot\bu)\,d^2\bx -
%}\nonumber\\
%\lefteqn{\hspace{0.5em}
\int_{\rm CMB}w(\brh\cdot\bu)\,d^2\bx
+\int_{\rm FOC}\rho^{-1}\bdel w\cdot\bdel\chi\,d^3\bx,}
\ena
%
\eqa \label{eq:weak3}
\lefteqn{\left\{
\begin{array}{c}
\bph\cdot\bw(r_{\rm r}\hspace{0.1em}\bzh)\delta(t)
\\ \bM:\bdel\bw(r_{\rm s}\hspace{0.1em}\bzh)H(t)
\end{array}\right\}=
%}\nonumber\\ \lefteqn{\hspace{0.5em}
\int_{\rm SMC}\rho\hspace{0.1em}\bw\cdot
\ddot{\bu}\,d^3\bx+\int_{\rm CMB}(\brh\cdot\bw)
\ddot{\chi}\, d^2\bx +}\nonumber\\
\lefteqn{\hspace{11.5em}\int_{\rm SMC}
\left\{\lambda(\bdel\cdot\bw)(\bdel\cdot\bu)
+\mu\bdel\bw\!:\!\left[\bdel\bu+\left(\bdel\bu\right)^T\right]\right\}d^3\bx}
\ena
%
where $\rho$ is the mass density, $\kappa$ the incompressibility, 
$\bph\delta(t)$ the impulsive unity single-force vector acting at the 
receiver at axial radius $r_{\rm r}$ and 
$\bM H(t)$ the moment tensor at the source at axial radius $r_{\rm s}$ as a 
step function in time, and second time derivatives are abbreviated as 
$\ddot{\chi}=\partial^2_t{\chi}$.
The regions solid inner core SIC, fluid outer core FOC, and solid mantle and 
crust SMC in eqs~(\ref{eq:weak1})--(\ref{eq:weak3}) are coupled by virtue of 
the boundary surface integrals over the inner-core (ICB) and core-mantle 
boundary (CMB). The fluid region eq.~(\ref{eq:weak2}) is treated via a 
displacement potential formulation \citep{nissen+:07a}
%
\eq\label{eq:potential}
\bu=\rho^{-1}\bdel\chi,
\en
%
such that the fluid-domain wave equation becomes
\eq\label{eq:fluid_waveequation}
\kappa^{-1}\ddot{\chi}=\bdel\cdot(\rho^{-1}\bdel\chi).
\en
%
Note that the inclusion
of the inverse density in this definition of the potential $\chi$ results in 
boundary integrals in eqs~(\ref{eq:weak1})--(\ref{eq:weak3}) that are 
independent of any intrinsic material properties.
This is slightly different from other spectral-element methods
\citep{KoTr02a,manu04} in which the density-free potential leads
to solid-fluid boundary terms that depend on the discontinuity jumps.
This is necessitated by their inclusion of gravitation, but in our 
non-gravitating case it is indeed feasible to pursue this convenient definition 
in eq.~(\ref{eq:potential}).
For spherically symmetric background models, the 3-D weak system 
eqs~(\ref{eq:weak1})--(\ref{eq:weak3}) may be reduced to its 2-D equivalent 
by factorizing $(\bu,\chi)$ and $(\bw,w)$ into their $(s,z)$ and $\phi$ 
dependencies and analytically evaluating the longitudinal integrals 
such as $\int_0^{2\pi}\sin^2{\phi}\,d\phi=\pi$ as shown by 
\citet[][Section~4.4]{nissen+:07a}.
%
While retaining the $(s,\phi,z)$ cylindrical coordinate
system for the monopole and quadrupole cases, we develop the dipole system 
in the $(+,-,z)$ system, such that
%
\eqa
\lefteqn{u_{\pm}(s,z)=\textstyle{\frac{1}{2}}[u_s(s,z)\pm u_{\phi}(s,z)],}\\
\lefteqn{w_{\pm}(s,z)=\textstyle{\frac{1}{2}}[w_s(s,z)\pm w_{\phi}(s,z)], }
\ena
%
inasmuch as this enables easier implementation of 
the axial boundary conditions given by
\eqa 
\lefteqn{\begin{array}{ll}
\textrm{Monopole:} &
\left(u_s=w_s\right)_{s=0}=0, \label{eq:ax_bc_mo}\\
\textrm{Dipole:}&
\left(u_-=w_-=u_z=w_z\right)_{s=0}=0, \label{eq:ax_bc_di}\\
\textrm{Quadrupole:}&
\left(u_s=w_s=u_\phi=w_\phi=u_z=w_z\right)_{s=0}=0 
\label{eq:ax_bc_qu}.
\end{array}}
\ena
%

\subsection{Meshing}
\begin{figure*}[htb!]
\begin{center}
\includegraphics[scale=0.68]{prem_mesh.eps}
\caption{\textbf{Left:} The semicircular, solid-fluid 
domain $\Omega=\Omega^{\rm s} + \Omega^{\rm f}$
discretized for the PREM background model using
quadrilateral elements $\Omega_e$ for dominant source period $T_0=20\, \textrm{s}$.
Note that all discontinuities are honored and several 
conforming coarsening levels are included to maintain a relatively 
constant resolution throughout the domain. 
\textbf{Top right:} Enlargement of the crust for one (right) and two (left) crustal
layers, and the upper mantle, including one mesh coarsening region. 
Note the variable vertical spacing due to discontinuities.
\textbf{Bottom right:} The central region for two resolutions. 
To circumvent the singularity at the center, 
we apply the following analytical expressions to reshape rectangular elements: 
$\left|{x}\right|^p+\left|{y}\right|^p=\left|{r}\right|^p$, 
$x=s+z,\,y=s-z,\,1\le p \le 2$.
This guarantees an easy handle on grid spacing which varies maximally 
at the outermost, deformed elements of this central region and hence 
controls stability and resolution.}
\label{img:prem_mesh}
\end{center}
\end{figure*}

Let us first define some geometric notation. As shown in 
Fig.~\ref{img:prem_mesh} on the left, we work in a semi- disk 
$\Omega=\Omega^{\rm s} + \Omega^{\rm f}$ of outer 
radius $r_0$ spanned by $\bxh^{\rm 2D}=\bsh + \bzh=\brh+\bthetah$,
where $s\in [0,r_0]$, $z\in [-r_0,r_0]$ refer to cylindrical coordinates, and
radius $r\in [0,r_0]$ and colatitude $\theta\in [0,\pi]$ denote the spherical system. 
The longitude $\phi\in [0,2\pi[$ is identical for both coordinate systems, 
$\bxh^{\rm 3D}=\bxh^{\rm 2D}+\bphih$.  Material properties are invariant in 
$\phi$, and the seismic wavefields either invariant (monopole sources) or analytically 
continued from the $(s,z)$ plane \citep{nissen+:07a}; as a result we do not need 
to discretize this third dimension. The grid structure represents the tiling into
non-overlapping solid ($\Omega_e^{\rm s}$) and fluid ($\Omega_e^{\rm f}$) 
elements within which functions are analytically mapped to a reference square
$[-1,+1]^2$ to be expanded upon a polynomial basis 
(see Section~\ref{section:spatial_discretization}).
Fig.~\ref{img:prem_mesh} also depicts magnified regions, showing 
the crust and upper mantle (top right panel) for a two-layered PREM \citep{prem}
crust (left) and a one-layered ``crust'' (right). Note the variable vertical spacing due to the 
discretization of adjacent discontinuities such as $600\,\textrm{km}$ and 
$660\, \textrm{km}$.

The generation of a tangible mesh is no trivial matter. 
In global seismology, we face the additional complication that the seismic
velocity $v_{\rm p,s}(r)$ generally increases, while horizontal grid spacing
(for spheroidal topologies) decreases with depth which works 
as a doubly detrimental effect for meshing purposes.
Thus, we need to employ mesh coarsening to remediate these issues.
Additionally, a spheroidal grid cannot account for the 
center of the sphere such that one needs to introduce a different (e. g. linear) topology.
The following factors also need to be considered as specific constraints
\citep[see also][Sections~3.1 and 4.1]{nissen+:07b}:
%
\begin{enumerate}
\item Quadrilateral element shapes,
\item Non-overlapping element boundaries,
\item Discontinuities to coincide with element boundaries.
\end{enumerate}
%
Point (i) is a minor issue inasmuch as the nature of 
sharp global boundaries is spherical, i.e. mildly deformed and we do not 
need to accommodate sharp wedges such as in regional velocity models which may
considerably influence the spacing variability and accuracy. Unless the 
expensive mortar element method \citep{manu03} 
or a discontinuous Galerkin approach based on numerical fluxes 
\citep{KaeserDumbser:05} are used, 
constraint (ii) only permits conforming coarsening architectures 
\citep{KoTr02a}, which is what we chose \citep{nissen+:07b}.
Due to the last point (iii), however, we are faced with problems 
such as discretizing the thin crustal layers even for long periods, thus 
introducing local oversampling, or the existence of the fluid outer core, and 
drastic velocity changes across the inner-core boundary 
(see Section~\ref{section:central_region}).
For elastic wave phenomena, one seeks a discretization in which the local
$P$- or $S$-velocity-dependent wavelength $\Lambda(v_{\rm p,s}(r))$ is 
sampled by a minimal variation in the number of grid points per wavelength $n_\Lambda$ 
anywhere in the spatial domain, and, simultaneously, minimal variation 
in the number of time samples $T_0/\Delta{t}$ for the source-induced, 
constant dominant (i.e., peak spectrum) period $T_0=\Lambda(v(r))/v(r)$. 
The limiting values for the number of grid points per 
wavelength, $n_\Lambda^0=\min{\left[n_\Lambda (r,\theta)\right]}$, controlling the 
resolution, and the Courant number, 
${\mathcal C}^0=\max{\left[{\mathcal C}(r,\theta))\right]}$, controlling the stability,
are inherent properties of the numerical scheme.
They are related to $T_0$ and $\Delta t$ via the 
characteristic lead time $\tau_{\rm p,s}(r,\theta)=\Delta x(r,\theta)/v_{\rm p,s}(r)$
\footnote{Subscripts p,s are mnemonic indicators for adhering to the respectively 
largest and smallest velocities for a given location; $\tau_{\rm s}$ in the fluid for example
is obviously defined by the overarching $P$-wave velocity, and near the surface by surface 
wave velocities.}:
%
\eqa 
\lefteqn{T_0 =n_\Lambda^0 \max{\left[\tau_{\rm s}(r,\theta)\right]},} \label{eq:period}\\
\lefteqn{\Delta{t}= {\mathcal C}^0 \min{\left[\tau_{\rm p}(r,\theta)\right]},}\label{eq:timestep}
\ena
%
where for each element, grid spacing $\Delta x(r,\theta)$ also depends on $\theta$ due to the 
irregular Gauss-Lobatto-Legendre point distribution across elements \citep{nissen+:07b}, 
the occurrence of coarsening levels and the linear discretization of the central region.
Meshing in light of these constraints and ideas 
principally aims at simultaneously 
accommodating two factors: 
(i) obtaining a smallest possible total number of elements, i.e. minimizing 
the run-time memory cost with 
(ii) the least variation in $\tau_{\rm p,s}(r,\theta)$. 
This non-unique meshing procedure can be addressed from many angles.
%
\subsubsection{The background-model based spherical mesh}
%
Given the above constraints and the two-dimensional geometry, 
we chose a straightforward approach to define our mesh: 
Given an anticipated dominant period $T_0$ 
to be resolved, the minimal number of grid points per wavelength $n^0_\Lambda$, 
and the location and velocity jump across discontinuities,
we calculate the maximally allowed, smallest global grid spacing  
$\Delta x(\bx_{\rm minx})=T_0/n^0_\Lambda v_{\rm s}(r_{\rm minx})$  for that location 
$\bx_{\rm minx}$ (typically just below the inner-core boundary, or at the surface). 
We determine the exact lowermost radius $r_{\rm sl}$ of the spheroidal domain 
(which characterizes the transition to rectangular element shapes)
by assuming that velocities within the inner core vary very little, starting with 
the grid-spacing constraint $r_{\rm sl}\le r_{\rm minx}$ and moving to 
the next possible lower radius
$r_{\rm sl}=\min\left[r_{\rm icb},{\rm int}\left(r_{\rm minx}/\Delta x(\bx_{\rm minx})\right)\right]$.
Additionally given the number of coarsening levels based 
on the overall change in velocity, and the number of processors $N_{\rm proc}$
(see Section~\ref{section:parallel}), we obtain a permissible number of lateral 
elements at the surface with grid spacing $\Delta x_{\rm lat}(r_0)$ 
upon adjusting $\Delta x(\bx_{\rm minx})$ such that 
$\pi r_0$ is an even multiple of $N_{\rm proc} \Delta x_{\rm lat}(r_0)$.
Progressively moving downwards from the surface while honoring all 
discontinuities and keeping vertical spacing between them similar to 
the ideal value $T_0/n^0_\Lambda \min[v_{\rm s}(r)]$, a coarsening depth 
$\textstyle{r_{\rm c}^i}$ is 
found once $\textstyle{2\tau_{\rm s}(r_{\rm c}^i,\theta)\le T_0/n^0_\Lambda}$, 
i.e. the spacing-velocity combination small enough to double and still resolve $T_0$. 
Once this process is completed, the global time step is found given the limiting 
Courant number ${\mathcal C}^0$ and $\min[\tau_{\rm p}]$ from the newly 
created mesh.
For the SEM, the minimal possible number of points per wavelength is 
$n_\Lambda^0\approx4$ \citep[e.g.][]{Ampuero+:07}, whereas ${\mathcal C}^0$ is 
usually determined empirically and depends on the type and 
order of the time extrapolation scheme (see Section~\ref{section:time_scheme})
and model-space dimension.
We obtain stable, accurate simulations up to ${\mathcal C}^0=0.6$, but one should 
treat this choice carefully for each different mesh and time scheme 
(see Section~\ref{section:time_scheme}). 
Clearly, for any mesh containing model property or grid spacing variations,
these values for $\Delta{t}$ and $T_0$ are the global worst-case combinations, 
and the actual, local Courant number and number of grid points per wavelength are 
variable across the mesh. It is therefore our maxime to construct $\Delta x(r,\theta)$ 
in a way to minimize such variations. All applications in this paper are undertaken 
with a polynomial order $N_{\rm pol}=4$, i.e. having $5$ grid points within an element 
along a given dimension (including its edges); for a more thorough quantification of 
meshing one may however also vary the polynomial order. Note that small polynomial 
orders ease the meshing process in two ways: the smallest grid spacing within elements 
upon a Gauss-Lobatto-Legendre basis varies as $N^{-2}_{\rm pol}$, and, secondly, 
smaller element sizes allow for easier adaptation to material interfaces, especially 
regarding closely spaced discontinuities like in the crust.
Fig.~\ref{img:prem_mesh} shows an example of a PREM model discretization using 
this method for $T_0=20 \, \textrm{s}$ (left), and two magnified regions of interest, i.e. 
the crust and upper mantle (top right) and the central region 
(bottom right, see next section). Clearly, lower resolutions (being confronted with 
the same background model as higher resolution realizations), are 
less effective and expose larger spacing variations, as seen in the crust. 
It is noteworthy to highlight that meshing is done only upon the lowest velocities
(surface waves near the surface, $S$ waves in deeper solid, $P$ waves in fluid), but one
subsequently determines the time step (i.e. the other component of the cost of the 
scheme besides the number of elements ) based on largest ($P$ wave) velocities.

\begin{figure*}[t!]
\begin{center}
\includegraphics[scale=0.7]
{char_times_fig.eps}
\caption{Elementally minimal and maximal characteristic lead times scaled by 
the time step and Courant number, 
$\tau_{\rm p,s} {\mathcal C}^0/\Delta{t}$ in the spherical part of the model space as  
a function of radius.
We depict PREM meshes for source periods $T_0=10\,\textrm{s}$ (left) and 
$T_0=20\, \textrm{s}$ (right). The vertical line to the left 
denotes unity, i.e. minimal possible $\min[\tau_{\rm p}]$ due to the definition of $\Delta{t}$ in 
eq.~(\ref{eq:timestep}), and the vertical line to the right the corresponding maximal 
value given by the relationship for the source period, eq.~(\ref{eq:period}), i.e. 
$T_0 {\mathcal C}^0/(n^0_{\Lambda}\Delta{t})$.}
\label{img:char_time}
\end{center}
\end{figure*}
%
It is thus desirable to at least assess this entirely non-unique, potentially iterative process 
a posteriori in terms of efficiency and overall cost. 
In the interest of generality, we present examples of different mesh 
realizations in terms of non-dimensionalized parameters, namely the 
characteristic lead time 
%
\eq
\frac{\tau_{\rm p,s}}{\min[\tau_{\rm p}]}=\tau_{\rm p,s}\frac{{\mathcal C^0}}{\Delta{t}},
\en
%
and the local oversampling ratio
%
\eq
\frac{T_0}{\Delta{t}^{\rm eff}(\bx)}=\frac{T_0}{n_\Lambda^0\tau_{\rm p}(\bx)}.
\en
%
\begin{table*}[htb!]
\begin{center}
\caption{Characteristic lead times and time steps for various mesh resolutions and Courant 
number ${\mathcal C}^0=0.6$ and $n_\Lambda^0=6$.}
\label{table:char_time} 
\begin{tabular}{@{}cccccccc}
\hline\hline
$T_0\, [s]$ & $\min{\left(\tau_{\rm p}^{\rm lat} \right)}\,[{\rm s}]$ &  
$\min{\left(\tau_{\rm p}^{\rm rad} \right)}\,[{\rm s}]$ &
$\max{\left(\tau_{\rm s}^{\rm lat} \right)}\,[{\rm s}]$ & 
%$\max{\left(\tau_{\rm s}^{\rm rad} \right)}\,[{\rm s}]$ & 
$\Delta t\,[{\rm s}]$ &  $\max{\left(\tau_{\rm s} \right)} {\mathcal C}^0/\Delta t$ & 
$T_0/\min\left(\Delta t^{\rm eff} \right)$
\\
\hline\\
$5$   & 0.0917& 0.117 & 0.832 & 0.0555 & 9.08 & 8.99 \\[10pt]
$10$ & 0.182 & 0.234 & 1.66 & 0.111 & 8.97 & 9.16 \\[10pt]
$20$ & 0.356 & 0.239 & 3.32 & 0.121 & 16.46 & 13.9 \\[10pt]
$40$ & 0.687 & 0.239 & 6.45 & 0.121 & 32.0 & 27.9 \\[10pt]
%
\hline
\end{tabular}
\end{center}
\end{table*}
%
In both cases, the definition contains all relevant constant parameters such that 
this analysis can be viewed as independent both of the actual mesh resolution and the 
choice of the spatio-temporal discretization scheme (spectral elements, 
finite differences etc.) and may therefore be
useful in estimating the mesh quality for varying resolutions and methods.
In Fig.~\ref{img:char_time}, we show the variation of the non-dimensional 
characteristic lead time for the spherical part 
of the domain as a function of radius. The panels are for different 
PREM model realizations with $T_0=10\, \textrm{s}$ on the left and
 $T_0=20\, \textrm{s}$ on the right. 
For each mesh, we plot radial/lateral and maximal/minimal values of 
$\tau_{\rm p,s}$ for each element, respectively.
Clearly, the mesh coarsening is reflected by the jumps in lateral spacing, whereas 
the radial structure is smoothly kept within the bounds set by the lateral values. 
Another apparent feature is the fact that the lateral characteristic lead times 
$\tau_{\rm p}$ seem to follow two profiles, i.e. for a given radius, there are 
two minimal spacings. This is merely the result of the fact that we employ 
a different polynomial basis within axial elements 
(Gauss-Lobatto-Jacobi (GLJ) points) than for non-axial elements 
(Gauss-Lobatto-Legendre (GLL) points): The profile with 
smaller values is due to axial elements, and all others follow the profile for 
larger values \citep[see][Section~3]{nissen+:07b}. Note also that the distinction 
only exists for lateral $\tau$ as we utilize the same Gauss-Lobatto-Legendre 
points in the radial direction for all axial and non-axial elements. The difference 
in largest elemental spacing within $\tau_{\rm s}$ is much smaller between GLJ and 
GLL, but still visible in Fig.~\ref{img:char_time}. 
Otherwise, the distribution of lead times is similar in both cases, suggesting that 
the meshing process is consistent and independent of the actual resolution.
Adhering to the non-dimensional nature of the function, it is also curious to note that 
we are faced with a variation of less than an order of magnitude
for the characteristic lead time in the $10\,{\rm s}$ case, but a  
doubled variation for $T_0=20\, \textrm{s}$ due to the fact that the 
crustal layer thicknesses at such low frequencies 
determine the smallest spacing rather than the resolution. Table~\ref{table:char_time} illuminates
this issue in a quantitative way by listing the characteristic lead times for meshes 
between $T_0=5\,{\rm s}$ and $T_0=40\,{\rm s}$. Evidently, the smallest radial spacing 
is constant for meshes above $10\,{\rm s}$ such that the time step does not increase with the
period, and the above-mentioned ratio of minimal versus maximal $\tau$ increases with the 
period. Apart from these cases of $T_0>10\,{\rm s}$, 
the meshing process is consistent and independent of the resolution in that
source period $T_0$ and $\tau$ are linearly related and their ratio constant.

\begin{figure}[htb!]
\begin{center}
\includegraphics[scale=0.55]
{prem_10sec_coars3_oversampling_rad_el.eps}
\caption{Local temporal oversampling in terms of period and time step for the mesh down to 
$10\, \textrm{s}$. Unity would be equivalent to the ideal case 
of no variation for 
the characteristic lead time $\tau$ whatsoever, impossible for any elastic medium.
A more realistic aim, the empty circles to the left depict the ratio $\tau_{\rm p}/\tau_{\rm s}$ which 
represents the limiting case of ideal meshing for the given background model \textit{and} the 
inevitable grid spacing variation due to the polynomial basis of Gauss-Lobatto-Legendre points
\citep{nissen+:07b}, i.e. with no oversampling due to suboptimal meshing.
The vertical line to the right 
is the constant ratio $(T_0/\Delta{t})({\mathcal C}^0/n^0_\Lambda)$, i.e. the global worst case scenario 
after which these parameters have been chosen. }
\label{img:spacing_ratios}
\end{center}
\end{figure}
%
Fig.~\ref{img:spacing_ratios} is an additional characteristic for the mesh 
efficiency, painting its local oversampling ratio as a function of
radius. This function essentially denotes the local numerical gap between 
defining grid spacing upon the period and largest possible spacing, 
but simultaneously noting the smallest spacing to abide stability for the temporal 
extrapolation. The lower bound is the mere $\tau_{\rm p}/\tau_{\rm s}$ ratio, 
i.e. the ideal case with local oversampling due to meshing, and the higher bound is 
the scaled ratio of period and time step $T_0/\Delta{t}$, i.e. the global worst case.
As before, the radial spacing follows a relatively smooth profile along the idealized 
case of the characteristic lead time ratio, whereas the lateral spacing is subject to
coarsening layers and the main contributor to spacing variability or deviation 
from the smooth profile. Table~\ref{table:char_time} lists this oversampling 
ratio in the last column for different resolutions. Similar to the characteristic 
lead time, the non-dimensionalized oversampling is constant and independent 
of the actual resolution once discontinuity locations do not determine the 
grid spacing as for $T_0>10\,{\rm s}$.
Again, it  may be interesting to conduct a thorough 
comparison of these functions for a variety of background models and numerical 
methods. In our case, it is verified that the meshing process is consistent for 
varying resolutions and optimally discretized where possible.
Both these latter plots give an idea about the mesh quality, inasmuch as they 
generally delineate excursions from the idealized respective minimal values 
\citep[compare][]{KoTr02a}. We shall take this to the next level of estimating 
a non-dimensionalized, global mesh cost after the next section on the central region.
%
\subsubsection{The central region}\label{section:central_region}
%
The singularity at $r=0$ is circumvented by rectangular element discretization
\citep{manuthesis} which introduces an additional problem, namely the vast 
spacing variations between the rectangular $r<r_{\rm sl}$ 
and circular regions $r>r_{\rm sl}$. We accommodate this by defining 
(see Fig.~\ref{img:prem_mesh})
%
\eq \label{eq:central}
\left|{x}\right|^p+\left|{y}\right|^p=\left|{r}\right|^p,\; 
x=s+z,\,y=s-z,\,1\le p \le 2.
\en
%
where the exponent $p$ can vary either linearly, quadratically, or cubically 
with the radius $r$. We thereby maintain an acceptable
$\min{[\tau_{\rm p}]}\le\tau_{\rm p,s}(r<r_{\rm sl},\theta)\le\max{[\tau_{\rm s}]}$ 
where we take the extremal values of $\tau_{\rm p,s}$ from the spherical 
domain $r\ge r_{\rm sl}$, thereby avoiding the case that these linear elements
determine the overall cost. 
Applying the mapping eq.~(\ref{eq:central}), elements along the 
diagonal directions $x=0$ and $y=0$ are extremely deformed, and therefore 
subsequently stretched outwards by half an element size, respectively 
(see Fig.~\ref{img:prem_mesh} c).
The high mesh density in the inner core stems from the fact that we effectively follow
a drastic velocity drop across the inner-core boundary since we only need to resolve
outer-core $P$ waves but inner-core $S$ waves.
Fig.~\ref{img:central} illustrates the central region in terms of relative points 
per wavelength (left) $n_{\Lambda}(\bx)/n^0_\Lambda$, and local oversampling 
(right) $T_0/(\tau_{\rm p}(\bx) n^0_\Lambda)$ as (interpolated) 2-D functions across 
the central region. The colorbars respectively range from the minimal to 
the maximal values of the spherical domain, which is taken as the 
goal to stay within. In both critical cases, our approach to minimize element 
distortion via eq.~(\ref{eq:central}) seems to be verified in that all elements in 
this region maintain characteristic lead times within the bounds of the 
spherical domain and thus do not further deteriorate the overall trade-off 
between $T_0$, total number of elements, and $\Delta t$.
%
\begin{figure}[tb!]
\begin{center}
\includegraphics[scale=0.6]
{prem_10sec_coars3_centralregion_period_dt_per_over_dt.eps}
\caption{Numerical parameters in the central region for the PREM mesh down to 
$10\, {\rm s}$. \textbf{Left:} Number of points per 
$S$ wavelength scaled by the minimal value $n_\Lambda^0=4$. The fact that we need to resolve 
$S$ waves in the inner core as opposed to outer core results in a drastic minimal velocity drop across 
the ICB and accordingly grid densification. \textbf{Right:}  Local oversampling ratio, 
i.e. the period $T_0$ versus characteristic lead spacing time $\tau_{\rm p}$. 
The colorbar bounds represent the extremal characteristic lead times in the spherical domain above.}
\label{img:central}
\end{center}
\end{figure}
%
\subsubsection{Mesh scaling and computational cost}
%
Based on the two previous sections, we now revert to estimating the overall 
cost of the scheme that is launched upon a mesh as defined here. 
In a recent study, \citet{Ampuero+:07} suggest for the cost
${\$\$}$ on a topologically and structurally homogeneous mesh the 
compact relationship
%
\eq\label{eq:abscost_homo}
\lefteqn{{\$\$} =  E(N_{\rm pol}) \left(\frac{L}{h}\right)^D \frac{L/v}{\Delta t},}
\en
%
where $E(N_{\rm pol})$ denotes the number of floating-point multiplications per element 
as a function of polynomial order $N_{\rm pol}$, $L$ is the total propagation length, 
and $D$ the dimension.
Assessing our highly non-regular mesh, we modify this expression into
the generalized, non-dimensional philosophy such that we can 
determine the quality and scaling properties as a function of the resolution, 
also bypassing $E(N_{\rm pol})$ since we keep $N_{\rm pol}$ fixed in this 
analysis for different meshes.
Let the cost $\$\$={\mathcal I}{\mathcal E}$ be factorized into
instantaneous cost (or an indicator for run-time memory occupation)
${\mathcal I}$ formed by the total number of elements $N_{\rm el}$ in the mesh, 
and secondly the evolving/extrapolation cost ${\mathcal E}$, 
i.e. the number of time steps:
%
\eqa
\lefteqn{{\mathcal I}=\frac{N_{\rm el}}{N_{\rm el}^{\rm ref}}=
\frac{2}{\pi r_0^2} (N_{\rm pol}\Delta x_{\rm max})^2 
N_{\rm el} ,} \label{eq:instcost}\\
\lefteqn{{\mathcal E}=\frac{\Delta{t}^{\rm ref}}{\Delta{t}}=\frac{T_0}{2\Delta{t} } . }
\label{eq:extrapolcost}%\\
\ena
%
In eq.~(\ref{eq:instcost}), we scale the total number of elements $N_{\rm el}$ with an approximate 
reference number of fewest possible elements, 
i.e. assuming global coverage via the largest grid spacing, such that $\min[{\mathcal I}]=1$ 
represents the idealized scenario were there no velocity and spacing variation for a given period $T_0$. 
In eq.~(\ref{eq:extrapolcost}), we take the idealized sampling rate from the Nyquist sampling 
theorem as $\Delta{t}^{\rm ref}=T_0/2$, such that again $\min[{\mathcal E}]=1$ represents the 
cheapest possible setting. In this scale-invariant formulation, we forego the inclusion of the total 
simulation length $L/v$ as in eq.~(\ref{eq:abscost_homo}) since this would simply result in the 
same multiplicative factor for all cases.
%
\begin{figure}[tb!]
\begin{center}
\includegraphics[scale=0.66]{cost_fig.eps}
\caption{Computational cost of the mesh. \textbf{Left:} Instantaneous cost, i.e. the 
size of the mesh. \textbf{Right:} Extrapolation cost, i.e. the number of time steps. 
Both functions are constructed in a non-dimensionalized fashion
such that they cancel out resolution dependencies when 
plotted as a function of the source period $T_0$.}
\label{img:cost}
\end{center}
\end{figure}
%
In contrast to \cite{Ampuero+:07}, we also refrain from computing a multiplicative total cost 
since the actual non-dimensionalized values of the two respective definitions 
eqs~(\ref{eq:instcost})--(\ref{eq:extrapolcost}) 
bear no interconnection, especially with regard to large values.
It is nevertheless noteworthy that the overall decisive factor for the cost of a given mesh 
for $T_0$ turns out as $N_{\rm pol}^2 \Delta x_{\rm max}^2 N_{\rm el}/\Delta{t}$.
Also note that the respective dependence on the period  (via $\Delta x_{\rm max}$ in eq.~(\ref{eq:instcost})) 
counteracts  the 
expected dependence of the 2-D spatial and 1-D temporal discretization such that the actual 
cost as defined in eqs~(\ref{eq:instcost})--(\ref{eq:extrapolcost}) 
is independent of the resolution itself by virtue of the chosen normalization. 
Fig.~\ref{img:cost} depicts the instantaneous and 
extrapolation cost, respectively, as a function of period for PREM and PREM-1-crust 
realizations of the mesh strategy described in the previous sections. 
Evidently, low resolution is relatively ineffective inasmuch as the time step is controlled 
by the smallest grid spacing which in turn is the mere radial separation of the two closest 
discontinuities, mostly in the crust. For both cost functions, we enter another regime below 
$10\,\textrm{s}$ , where the meshing process itself controls the computational cost and the 
two models are equally expensive. 
The slight increase of instantaneous cost with higher resolution below $7\,\textrm{s}$ 
represents the cutoff resolution below which the instantaneous cost does not decrease 
any further due to the background model heterogeneities, overall velocity 
contrasts and the associated necessity to honor smaller velocities with smaller 
element sizes, independent from the actual resolution.
The seemingly asymptotic convergence  to $50$ for the extrapolation cost reflects 
the ratio $T_0/\Delta t$ as shown in Fig.~\ref{img:spacing_ratios} which is then 
independent of the resolution but merely the difference in maximal and minimal 
characteristic lead times. In both instantaneous and extrapolating cost functions, 
the mesh is more cost-effective with increasing resolution, which is 
expected as we asymptotically mimic a continuous medium 
when increasing resolution at fixed background structure: We are progressively 
less limited by grid constraints such as discontinuity locations, permissible lateral 
number of elements due to the number of processors, coarsening levels, and 
hemispheric mirror symmetry as will be further elucidated in the next section.
%
\begin{figure*}[tb!]
\begin{center}
\includegraphics[scale=0.54]{dd_fig.eps}
\caption{Domain decomposition for $T_0=10\,{\rm s}$ 
and four (left), eight (middle), and sixteen (right two panels) processors.
Note that each processor has the same number of elements 
(``efficiency''), and maximally two neighbors (minimal latency)
with small shared boundaries (i.e., small bandwidth). 
Each processor touches the axis, albeit to varying amounts. Axial terms 
are however 1-D, and as such do not add any countable CPU time to the 
overall scheme.The central region is subdivided such that each processor 
has exactly the same amount of elements.}
\label{img:dd}
\end{center}
\end{figure*}
%
\subsubsection{Parallelization and domain decomposition} \label{section:parallel}
%
The meshes described in the previous section, even though only 2-D, are dense enough 
to reach typical memory limits when going below $T_0\sim 5 \, \textrm{s}$: 
The mesh at $5\,\textrm{s}$ contains about $10^7$ grid points, 
each carrying up to $40$ (single-precision) global, scalar floating-point 
numbers at run-time memory, summing up to $1.5\, \textrm{GByte}$ shared memory.
This shall not remain a centennial problem in the wake of hardware improvements, 
but as we do strive to compute kernels down to $1\, \textrm{s}$, it is indispensible 
to parallelize the method, especially in light of the extensive wavefield output.
With the potentially limited life-time of the parallelized version in mind and the fact 
that one may compute synthetics and kernels for lower resolutions, 
the code is written in a parallel modular way such that all message passing, 
communication and domain decomposition issues are confined to 
a single module and moderate amount of call statements hence it being trivial 
to denounce the parallel world and adopt the method for entirely serial simulations.
When spreading the computational workload across several separate CPUs and 
memory units, one seeks to minimize \textit{inefficiency} (work imbalance), 
\textit{latency} (number of neighboring processors) and \textit{bandwidth} 
(vector length to be communicated). We follow an approach based on global 
numbering \citep{TufoFischer:01,dfm}, predefining one index vector of 
arbitrarily located grid points for each processor-processor pair to be 
exchanged such that processors may harbor separate subdomains or share 
multiple edges but always exchange one message with any neighbor. 
This flexibility enables us to e.g. let each CPU belabor fluid {\it and} solid regions, 
hence reducing imbalance.
%
Fig.~\ref{img:dd} shows examples of our domain decomposition strategy 
for the mesh down to $T_0=10\,{\rm s}$ for four, eight, and sixteen 
processors from left to right: 
We chose to split the domain laterally such that each processor has an 
equal amount of fluid and solid elements at the generally inevitable expense of 
focusing the axial workload (additional terms and source) onto two 
processors that occupy most of the axis (e.g. black and green domains in the 
left panel). 
%
As will be seen, this shall not hamper the favorable scaling of 
our parallelization. At this point, it is important to recall our application of 
saving spatio-temporal wavefields for the computation of sensitivity kernels, 
i.e. such extensive I/O shall overshadow any slight processor imbalance due to
additional axial terms or the source by and large. 
The actual number of processors needs to be an even 
multiplicative of $2$ due to the equatorial symmetry and the connection
of the spheroidal with the central region. While the former is trivial to 
decompose simply based upon colatitude, the latter central part poses 
a problem in keeping with the load-balance philosophy.
%
\begin{figure}[tb!]
\begin{center}
\includegraphics[scale=0.4]{parallelscaling_both_2.eps}
\caption{Relative simulation times of the time loop for different numbers of processors 
$N_{\rm proc}$, either keeping the total memory (i.e., constant mesh and resolution) 
or the per-processor memory constant (i.e., constant job load per processor). $t_0$ 
is the CPU time for the one-processor case;
the diagonal constitutes the case for which CPU times are unaffected by the parallelization. 
In both our cases, the parallelization 
plays such a minor role that simulation times are actually \textit{shorter} for more processors, 
adhering to smaller memory occupation and negligible parallel communication patterns.} 
\label{img:scaling}
\end{center}
\end{figure}
%
We tackled this issue as seen in the magnified panel on the right of 
Fig.~\ref{img:dd} by defining discrete polynomials $z(s)$ up to 
second order for each processor's central region under the premises of 
covering the same discrete area, maintaining a respective 
vertical and diagonal ``thickness'' of at least one element,
choosing outer end locations given by the 
spheroidal processor latitudes, and touching the axis with at least one 
element. During this elemental fitting to the polynomial, 
we ``fix'' non-smooth processor boundaries including those that 
develop elemental plateaus and corners.
%
Nevertheless, this procedure carries with its iterative, element-by-element 
search some inevitable, unaesthetic remnants such as the spikes in the processor 
boundary shapes, but such negligible deviations from the same-surface ideal 
seem inevitable when working with such domains.  
Most importantly, this approach guarantees exact load balancing in terms of 
number of elements, and minimal communication such that each processor only 
touches maximally two neighbors, taking an inherent topological 
advantage of our 2-D semi-disk bounded by the axis. 
%
%
Again, we strive to quantify this parallelization approach in a 
relative, non-dimensional sense. Two end-members are considered, 
namely the fixation of total and processor-specific memory. 
Keeping the total memory constant means that the code ideally speeds up 
by the factor $N_{\rm proc}$, whereas constant memory per processor means 
that for constant run-time, the resolution should accordingly 
increase with $N_{\rm proc}$. Fig.~\ref{img:scaling} shows some results from 
simulations up to $16$ processors: In both end-member cases, the curves lie 
below the diagonal which is the communication-blind scaling where CPU times 
are inversely proportional to $N_{\rm proc}$, such that our parallelization is 
actually favorable: The total CPU time of a parallel job is \textit{smaller} than 
the according CPU time of the single-processor job. This result is mainly attributed 
to the memory occupation, but more importantly, validates our domain decomposition 
and parallelization scheme as completely transparent to the CPU times. Generally though, 
minute hardware and network fluctuations between different nodes may have well affected
these CPU times such that the actual values shall be viewed with caution. Clearly, though, 
this is an extremely favorable result, and shall be repeated in a thorough benchmark 
on different architectures in the future.  
%
Finally, Fig.~\ref{img:message_surf_vol} depicts another quality control on the
domain decomposition for varying resolutions: We plot the sum of message
interfaces (counted in unique grid points) versus the number of unique grid
points of the entire domain as a function of $N_{\rm proc}$, in other
words describing the fractional spatial size of messages related to the total mesh. 
For all feasible cases of up to $32$ processors (extrapolating this function), 
we retain moderate, sub-percentile values such that CPU times shall not be 
significantly affected by the mere size of messages.
If one however extrapolated these to very large numbers of processors (beyond $32$), 
it may be worthwhile to consider other domain decomposition strategies 
which keep these ratios smaller (e.g. 2-D decomposition).
See Appendix~\ref{appsection:runtime_breakdown}
for more on fractional runtimes and parallel scaling.
%
\begin{figure}[t!]
\begin{center}
\includegraphics[scale=0.4]{sum_messages_surf_vol.eps}
\caption{The sum over all message sizes versus total volume size 
(both counted in unique grid points) for each mesh and $N_{\rm proc}$, 
i.e. the grid fraction that communicates. 
If this function reaches a significantly large value, then the 
parallelization and message passing will take up larger fractions 
of the CPU time, but as our case indicates, it shall asymptotically 
approach value below $1\,\%$ for all projected, feasible cases of 
up to $32$ processors. }
\label{img:message_surf_vol}
\end{center}
\end{figure}
%
%
\subsection{Spectral-element discretization}
In this section, we give a brief overview of the aspects of the 
spectral-element method (SEM) that are relevant for the discretization, 
which we take up in the following sections. 
The reader is referred to the detailed literature
for a full description of the approach, e.g. \citet{dfm} and 
\citet{karniadakis} 
on the mathematical foundations of the SEM with a focus on fluid dynamics; 
\citet{KoVi98}, \citet{KoTr99}, \citet{KoTr02a}, \citet{manu03}, and 
\citet{Chaljub+:06} on the 
elastodynamic SEM; and \citet{bernardi} and \citet{fournier04} on 
discretizing an axisymmetric system.
%
%
\begin{figure}[htb!]
\begin{center}
\includegraphics[scale=0.8]{fig1.eps}
\caption{The mesh architecture. The geometric mapping mentioned in 
Section~\ref{section:geometrical} transforms the unit square element (A), 
within which all computations are undertaken, into any quadrilateral element 
$\Omega_e$ or $\bar{\Omega}_e$ (B). (C) shows the actual mesh used 
in the simulations described in Section~\ref{section:validation}. 
Three different element geometries, as defined in 
Appendix~\ref{appsection:geom_mapping} and 
Fig.~\ref{figa1}, are used to construct this mesh such that spacing 
variations are kept small. The radius of the earth is $r_0$, and 
the left straight edge is the non-physical axis of 
symmetry for which we employ different discretization rules due to the 
appearance of singularities $s^{-1}$.}
\label{fig1}
\end{center}
\end{figure}
%
\subsubsection{Geometrical mapping and discretization}\label{section:geometrical}
%
As in any element-based method, we discretize and decompose the continuous, 
open domain $\Omega$ with boundary $\partial\Omega$ 
into a union of non-overlapping elements. 
For clarity, we shall distinguish the so-called axial elements, which share an 
edge with the axis of symmetry $s=0$, from the remaining ones. Any axial 
element is then denoted by $\bar{\Omega}_{e}$, $e=1,...,\bar{n}$, 
while a non-axial element is represented by $\Omega_e$, $e=1,...,n$, such that 
the total number of elements forming the skeleton of the spectral-element 
mesh is $n+\bar{n}$.
Throughout the paper, we will consistently utilize an overbar to 
identify axial quantities. This decomposition permits us to break any global 
integration over $\Omega$ up into $n+\bar{n}$ local integrations, each 
spanning their respective elements, such that we define all integrands on 
the elemental level and only connect these to the global system when 
marching forward in time, as detailed in 
Section~\ref{section:discr_eq_motion}. 
In the 2-D spectral-element discretization used here, we confine $\Omega_e$ or 
$\bar{\Omega}_{e}$ to be a quadrilateral image of a reference square 
spanning $-1\le\xi\le 1$, $-1\le\eta\le 1$, defined by the invertible 
mapping $s=s(\xi,\eta)$, $z=z(\xi,\eta)$ (Fig.~\ref{fig1}~(A),~(B)). 
The exclusion of other geometries 
such as triangular elements will not hamper the feasibility of our approach 
given the spherical nature of spherical-earth discontinuities, i.e., the 
mildly deformed elemental shape. 
The D-shaped, planar domain with radius $r_0$ is shown in Fig.~\ref{fig1}~(C),
along with the mesh discretization as used throughout this paper.
Note that we included several conforming coarsening levels to 
reduce grid spacing variations \citep{KoTr02a} which generally trade 
off the time step (depending on the minimal spacing) with the frequency 
resolution (depending on the maximal spacing) in any numerical scheme. 
The center is discretized using rectangular elements \citep{manuthesis}.
The expression of the mapping depends on the elemental shape and is
either analytical or subparametric \citep{fournier04}. Explicit formulae and 
illustrations are given in Appendix~\ref{appsection:geom_mapping}.
With a given element shape, we can then compute the elemental Jacobian
\eq \label{eq:jacob}
{\mathcal J}(\xi,\eta)=
\frac{\partial(s,z)}{\partial(\xi,\eta)}=
\det\left(\begin{array}{cc}
\partial_\xi s & \partial_\eta s \\
\partial_\xi z & \partial_\eta z
\end{array}\right).
\en
%
Derivatives $\partial_s$ and $\partial_z$ of a function 
$u(s(\xi,\eta),z(\xi,\eta))$ map into the reference square as
\eq \label{eq:ds_u}
\partial_{s}u(s,z)=
\left[\partial_{\eta}z(\xi,\eta)\partial_{\xi}u(s(\xi,\eta),z(\xi,\eta))-
\partial_{\xi}z(\xi,\eta)\partial_{\eta}u(s(\xi,\eta),z(\xi,\eta))\right]
{\mathcal J}^{-1}(\xi,\eta),
\en
%
\eq \label{eq:dz_u}
\partial_{z}u(s,z)=
\left[-\partial_{\eta}s(\xi,\eta)\partial_{\xi}u(s(\xi,\eta),z(\xi,\eta))+
\partial_{\xi}s(\xi,\eta)\partial_{\eta}u(s(\xi,\eta),z(\xi,\eta)) \right]
{\mathcal J}^{-1}(\xi,\eta).
\en
%
Elements $\bar{\Omega}_{e}$ adjacent to the non-physical boundary $s=0$ 
(i.e., the axis of symmetry) are treated separately to accommodate 
singularities arising from terms 
involving $s^{-1}$ in the equations of motion (\ref{monosmc})--(\ref{quadsmc}).
We describe the treatment of the non-axial elements in 
Section~\ref{section:nonax} and briefly discuss the additional complexities 
for the axial elements in Section~\ref{section:ax}.
%
%
\subsubsection{General non-axial functional discretization}\label{section:nonax}
%
The spectral-element approach presented here relies on quadrature of the 
Gauss-Lobatto type, allowing us to enforce continuity across element 
boundaries conveniently by including the nodes $\xi=-1,1$. 
$N$th order Legendre polynomials $P_N$ 
(see Appendix~\ref{appsection:sem_lib}, eq.~(\ref{appeq:legendrepol})) 
are then invoked to define the most accurate and natural quadrature rule 
\citep{dfm}. Quadrature nodes are hereupon defined as the zeroes of 
$(1-\xi^2)\partial_\xi P_N(\xi)$, 
denoted as Gauss-Lobatto-Legendre (GLL) points $\xi_p,\;0\le p \le N$.
To interpolate functions upon these GLL nodes, we introduce Lagrange 
interpolating functions $l^N_i(\xi_p)$ of polynomial order $N$, 
$0\le i,p\le N$, such that: $l^N_i(\xi_p)=\delta_{ip}$ (see 
Appendix~\ref{appsection:gll}). For clarity, we shall denote GLL nodes 
in the $\xi$ direction as $\xi_i$ and in the $\eta$ direction as $\eta_j$.
The choice of Lagrangian interpolation upon GLL nodes yields the desirable 
property of a diagonal mass matrix, thereby greatly reducing the effort on 
time marching (see Section~\ref{section:discr_eq_motion}).
Furthermore, the orthogonality of coordinates $\xi$ and $\eta$ in a 
quadrilateral element geometry allows for separation of variables and 
leads to efficiently implemented tensor products (see 
Appendix~\ref{appsection:stiffness_mxm}).
Elemental integrals over $(s,z)$ may now be mapped into $(\xi,\eta)$ and 
discretized as
%
\eqa \label{eq:gll_quad_nonax}
\lefteqn{\int_{\Omega_e}u\left(s,z,t\right)s\,ds\,dz
=\int_{-1}^{1}\int_{-1}^{1}u\left(s(\xi,\eta),z(\xi,\eta),t\right)s(\xi,\eta)
{\mathcal J}(\xi,\eta)\;d\xi \,d\eta }\nonumber \\
&&\mbox{}\hspace{2.3cm} %extra
\approx \sum\limits_{p,q=0}^{N}
\sigma^N_p \sigma^N_{q}s(\xi_p,\eta_q){\mathcal J}(\xi_p,\eta_q) 
{u\left(\xi_p,\eta_q,t\right)},
\ena
where $\sigma_{p,q}=\int_{-1}^{1}l^N_{p,q}(\xi)\;d\xi$ 
are the integration weights for non-axial elements.

Time-dependent field variables such as the displacement are approximated as
%
\eq \label{eq:u_pol}
u_\alpha(\xi,\eta,t)\approx\sum\limits_{i,j=0}^{N}
u_\alpha^{ij}(t)l^N_i(\xi)l^N_j(\eta),
\en
%
where $u_\alpha$ stands for any of the components $u_s$, $u_z$, $u_\phi$, or 
$u_\pm$. 
The coefficients $u_\alpha^{ij}(t)$ carry the actual function values at the 
nodes $(\xi_i,\eta_j)$. Similarly, we define test functions $w_\beta$ as
\eq \label{eq:w_pol}
w_\beta(\xi,\eta)\approx\sum\limits_{I,J=0}^{N}w_\beta^{IJ}
l^N_I(\xi)l^N_J(\eta), 
\en
%
where $\beta$ is the analog of $\alpha$ for test functions. 
We set the coefficients $w^{IJ}_\beta$ in the discretized version of 
eqs~(\ref{monosmc})--(\ref{quadsmc}) equal to zero for all points except 
$(\xi_I,\eta_J)$, effectively canceling the summation and replacing 
elemental integrations with quantities defined at each point $(I,J)$.
For the remainder of the paper, we will adhere to the index convention 
$(I,J,\beta)$ for test functions, $(i,j,\alpha)$ for displacements, 
and $(p,q)$ arising from the quadrature.
%
Using eqs~(\ref{eq:ds_u}) and (\ref{eq:dz_u}), 
derivatives are expanded, e.g., as 
\eqa \label{eq:discrete_deriv}
\lefteqn{\partial_s u(s(\xi,\eta),z(\xi,\eta))\approx }\nonumber\\
&&\mbox{}\hspace{1.cm} %extra
\left[\partial_{\eta}z(\xi,\eta)\sum^N_{i,j=0}\partial_\xi l^N_i(\xi) 
l^N_j(\eta)u^{ij}
-\partial_{\xi}z(\xi,\eta)\sum^N_{i,j=0}l^N_i(\xi) 
\partial_\eta l^N_j(\eta)u^{ij}\right]{\mathcal J}^{-1}(\xi,\eta).
\ena
%
In what follows, we will drop dependencies on $t$, $(s,z)$, $(\xi,\eta)$, 
and $N$ for conciseness unless indispensable. We will also adopt the 
abbreviations 
$l^{\,\prime}_i(\xi)=\partial_\xi l_i(\xi)$, 
$l^{\,\prime}_j(\eta)=\partial_\eta l_j(\eta)$, and, e.g.,  
$s^{ij}_\xi=\partial_\xi s\left(\xi_i,\eta_j\right)$. 
Summation over $I,J$, $i,j$, and $p,q$ always goes from $0$ to $N$, 
unless otherwise noted.
Each elemental integral term is computed by discretizing 
$u_\alpha$ and $w_\beta$, evaluating the integral with GLL quadrature,
and recasting the system such that we obtain expressions over $I,J$ 
as the free indices (see Section~\ref{section:discr_eq_motion}). 
For example, a typical term within the elemental 
monopole stiffness integral may be discretized as 
%
\eq
\int_{\Omega_e}\lambda\,\partial_z{w_z}
s^{-1}u_s s\,ds\,dz  \approx
\sum_{IJ}w_z^{IJ}
\left[\sum_j\lambda^{Ij}\sigma_I
\sigma_j s_\xi^{Ij}l^{\,\prime}_{J}(\eta_j)u_s^{Ij}
-\sum_i\lambda^{iJ}\sigma_i
\sigma_J s^{iJ}_{\eta} l^{\,\prime}_{I}(\xi_i)u_s^{iJ}
\right].
\en
%
We will elaborate upon the whole system in Section~\ref{section:stiffness} 
and Appendix~\ref{appsection:stiffness}.
%
\subsubsection{Gauss-Lobatto-Jacobi (0,1) quadrature in axial elements}
\label{section:ax}
%
Particular attention must be focused on the elements adjacent to the axis
for which the case $s=0$ needs to be addressed. A convenient remedy is to 
introduce an additional factor in the quadrature formulae, and apply 
L'Hospital's rule if needed. Consequently, while retaining the
GLL quadrature in the $\eta$ direction, we resort to a Gauss-Lobatto-Jacobi
(GLJ) (0,1) quadrature rule in the $\xi$ direction and introduce Lagrange 
interpolating functions $\bar{l}_i(\bar{\xi}_p)$ 
over the set of associated quadrature nodes $\bar{\xi}_p,\;0\le p\le N$ 
\citep[e.g.][see Fig.~\ref{fig2} for the relative distribution of GLJ versus 
GLL points, and 
Appendix~\ref{appsection:glj_ax} for definitions]{fournier05,bernardi}. 
Elemental integrals within axial elements are then expressed as
%
\eq \label{eq:glj_quad_ax}
\int_{\bar{\Omega}_{e}}u\left(s,z\right)
s\,ds\,dz\approx \sum\limits_{pq}\bar{\sigma}_p(1+\bar{\xi}_p)^{-1} 
\sigma_q s(\bar{\xi}_p,\eta_q){\mathcal J}(\bar{\xi}_p,\eta_q) 
{u\left(\bar{\xi}_p,\eta_q\right)}.
\en
%
This definition introduces the singular term $(1+\bar{\xi}_p)^{-1}$ on 
the axis $\bar{\xi}_0=-1$, which can be removed by applying L'Hospital's 
rule $s(\bar{\xi}_0,\eta_j)(1+\bar{\xi}_0)^{-1}=s_\xi^{0j}$. 
When evaluating stiffness integral expressions, we also make use of the 
axial identities
$\bar{l}_i(\bar{\xi}_0)(1+\bar{\xi}_0)^{-1}=\partial_{\xi}
\bar{l}_i(\bar{\xi}_0)$, $z_\eta|_{s=0}=({\mathcal J}s_\xi^{-1})|_{s=0}$
and $\partial_s u|_{s=0}= s^{-1}u|_{s=0}$ 
if $u|_{s=0}=0$. 
For each integrand in eqs~(\ref{monosmc})--(\ref{quadsmc}), we need to honor 
the two cases $p>0$ and $p=0$. For the latter location on the axis, 
L'Hospital's rule is invoked to evaluate terms involving $s^{-1}$.
Similarly to the non-axial case, we obtain for example
%
\eqa
\lefteqn{\int_{\bar{\Omega}_{e}}\lambda\,\partial_z{w_z}
s^{-1}u_s s\,ds\,dz }  \nonumber \\
&&\mbox{}\hspace{0.0em}
\approx\sum_{IJ}w_z^{IJ}\biggl[ \sum_i-\lambda^{iJ}\bar{\sigma_i}\sigma_J
(1+\bar{\xi}_i)^{-1}s_\eta^{iJ}\bar{l}^{\,\prime}_{I}(\bar{\xi}_i)
u_s^{iJ}+\sum_j\lambda^{Ij}\bar{\sigma_I}\sigma_j
(1+\bar{\xi}_I)^{-1} s_\xi^{Ij}l^{\,\prime}_{J}(\eta_j)u_s^{Ij}\nonumber\\
&&\mbox{}\hspace{1.em}
+\sum_j\bar{\sigma}_0\sigma_j\lambda^{0j}s_\xi^{0j}
\bar{l}^{\,\prime}_I(\bar{\xi}_0)l^{\,\prime}_J(\eta_j)u_s^{0j}\biggr]+ 
\sum_J w_z^{0J}\biggl[
\bar{\sigma}_0\sigma_J\lambda^{0J}s_\xi^{0J}
\sum_i \bar{l}^{\,\prime}_i(\bar{\xi}_0)\sum_jl^{\,\prime}_J(\eta_j)
u_s^{ij}\biggr].
\ena
%
The two terms in the second row appear due to the singularity removal via 
l'Hospital's rule; the first depends upon the 
displacement along the axis and contributes to the whole element $w^{IJ}$, 
whereas the second depends on the displacement in the whole element 
and appears only on the axis $w^{0J}$. 
Depending on the essential boundary condition, maximally one of these two 
terms is non-zero since in all cases either $w^{0J}=0$ or $u^{0j}=0$ for this 
type of integral. 
Section~\ref{section:stiffness} contains details on the different terms and 
the full stiffness system is derived in Appendix~\ref{appsection:stiffness}.
%
%
\subsubsection{Geometrical mapping} \label{appsection:geom_mapping}
%
\begin{figure}[htb!]
\begin{center}
\includegraphics[scale=0.7]{figa1.eps}
\caption{Elemental shapes. Sketch illustrating the three fundamental 
element-shape geometries used to construct the mesh. 
The mesh on the left is a low resolution example to simply highlight 
the skeleton. (A) is the analytically expressed circular
element shape, (B) is the purely rectangular 
element shape necessary to mesh the central part of the domain and computed in 
a subparametric fashion (``serendipity quadrilateral''), and (C) is the 
connecting type between (A) and (B), mapped using an analytical formula 
similar to type (A). 
Both (A) and (C) take on two different shapes, one of which is located 
within the coarsening region, respectively.
Numbers denote the collocation point 
indices as used in Appendix~\ref{appsection:geom_mapping}.}
\label{figa1}
\end{center}
\end{figure}
%
%
Depending on the geometrical complexity of elements, we employ either 
analytical or subparametric mapping between the reference system 
$(\xi,\eta)$ and the physical coordinates $\bx=(s,z)$ \citep{fournier04}. 
We distinguish three fundamental elemental geometries as depicted in 
Fig.~\ref{figa1}. Given the set of vertex coordinates 
$\{\bx_1,\bx_3,\bx_5,\bx_7\}$ for the most common, circular type (A) and 
the circular-linear transition type (C), any $(s,z)$ corresponding to 
reference coordinates $(\xi,\eta)$  is computed by analytical transformation 
formulae for the $z\ge 0$ hemisphere
%
\eq \label{appeq:generic_sz_a_c}
\lefteqn{s(\xi,\eta)=
\textstyle{\frac{1}{2}}\left[\left(1-\eta\right)\tilde{s}_{\rm bot}(\xi)+ 
\left(1+\eta\right)\tilde{s}_{\rm top} (\xi)\right],\,
z(\xi,\eta)=
\textstyle{\frac{1}{2}}\left[\left(1-\eta\right)\tilde{z}_{\rm bot}(\xi)+ 
\left(1+\eta\right)\tilde{z}_{\rm top}(\xi)\right],}
\en
%
where $\{\tilde{s},\tilde{z}\}_{\rm top}$ are the same for both types:
%
\eq \label{appeq:sz_top_ac}
\lefteqn{\tilde{s}_{\rm top}(\xi)=r_7
\sin\left\{\textstyle{\frac{1}{2}}\left[\left(1-\xi\right)\theta_7+
\left(1+\xi\right)\theta_5\right]\right\},\,
\tilde{z}_{\rm top}(\xi)=r_7
\cos\left\{\textstyle{\frac{1}{2}}\left[\left(1-\xi\right)\theta_7+
\left(1+\xi\right)\theta_5\right]\right\}.}
\en
%
$\{\tilde{s},\tilde{z}\}_{\rm bot}$ for type (A) read
%
\eq \label{appeq:sz_bot_a}
\lefteqn{\tilde{s}_{\rm bot}(\xi)=r_1
\sin\left\{\textstyle{\frac{1}{2}}\left[\left(1-\xi\right)\theta_1+
\left(1+\xi\right)\theta_3\right]\right\},\,
\tilde{z}_{\rm bot}(\xi)=r_1
\cos\left\{\textstyle{\frac{1}{2}}\left[\left(1-\xi\right)\theta_1+
\left(1+\xi\right)\theta_3\right]\right\},}
\en
%
and for type (C)
\eq \label{appeq:sz_bot_c}
\lefteqn{\tilde{s}_{\rm bot}(\xi)=\textstyle{\frac{1}{2}}r_1
\left[\left(1-\xi\right)\sin\theta_1+\left(1+\xi\right)\sin\theta_3\right],\,
\tilde{z}_{\rm bot}=\textstyle{\frac{1}{2}}r_1
\left[\left(1-\xi\right)\cos\theta_1+\left(1+\xi\right)\cos\theta_3\right].}
\en
%
For $z<0$, the mesh is mirrored by swapping the respective vertices 
$\bx_1 \leftrightarrow \bx_7$ and $\bx_3 \leftrightarrow \bx_5$ 
in the above formulae.
The linear element type (B) is parameterized via eight control nodes as shown 
in Fig.~\ref{figa1}.

We approximate the mapping by \cite[e.g.][]{KoTr99}
%
\eq
\bx(\xi,\eta)=\sum^{n_a}_{a=1}N_a(\xi,\eta)\bx_a,\; \textrm{and derivatives as}
\;\;\;\frac{d\bx(\xi,\eta)}{d(\xi,\eta)}=
\sum^{n_a}_{a=1}\frac{dN_a(\xi,\eta)}{d(\xi,\eta)}\bx_a,
\en
where $N_a$ are shape functions defining the geometry anchored at the $n_a=8$ 
control points $\bx_a$, which define the so-called serendipity quadrilateral
\citep{hughes}. The shape functions are given by
%
\eqa 
\lefteqn{N_1(\xi,\eta)= (1-\xi)(1-\eta)(-1-\xi-\eta)/4,}
\label{appeq:serend_shp8_1}\\
\lefteqn{N_2(\xi,\eta)=(1-\xi^2)(1-\eta)/2,}\\
\lefteqn{N_3(\xi,\eta)= (1+\xi)(1-\eta)(-1+\xi-\eta)/4,}\\
\lefteqn{N_4(\xi,\eta)=(1+\xi)(1-\eta^2)/2,}\\
\lefteqn{N_5(\xi,\eta)= (1+\xi)(1+\eta)(-1+\xi+\eta)/4,}\\
\lefteqn{N_6(\xi,\eta)=(1-\xi^2)(1+\eta)/2,}\\
\lefteqn{N_7(\xi,\eta)= (1-\xi)(1+\eta)(-1-\xi+\eta)/4,}\\
\lefteqn{N_8(\xi,\eta)=(1-\xi)(1-\eta^2)/2.}\label{appeq:serend_shp8_8}
\ena
%
Given the coordinates $\bx_a$ at the eight control nodes, one may then readily 
compute coordinates at any point $\bx(\xi,\eta)$, 
particularly the grid points used in the quadrature, 
as defined in the following section. 
%
Partial derivatives $\partial\bx/\partial(\xi,\eta)$ and the Jacobian 
${\mathcal J}=\partial_\xi{s}\partial_\eta{z}-\partial_\xi{z}\partial_\eta{s}$
are obtained by straightforward differentiation of 
eqs~(\ref{appeq:generic_sz_a_c})--(\ref{appeq:serend_shp8_8}).
%
%
\subsubsection{Gauss-Lobatto-Legendre quadrature} \label{appsection:gll}
%
The solution to the differential equation 
\eq \label{appeq:legendre_ode}
\partial_\xi \left[\left(1-\xi^2\right)\partial_\xi P_N\right] 
= -N\left(N+1\right)P_N
\en
%
with $P_N(1)=1$ and $P_N(-1)=(-1)^N$ is a Legendre polynomial $P_N$ 
of order $N$
\eq \label{appeq:legendrepol}
P_N(\xi)=\frac{1}{2^N N!}\frac{d^N}{d\xi^N}(\xi^2-1)^N
\en
which may be computed using the induction formula
\eq
P_N(\xi)=\frac{1}{N}\left[(2N-1)\xi P_{N-1}(\xi)-(N-1)P_{N-2}(\xi)\right]
\textrm{, $P_0(\xi)=1$, $P_1(\xi)=\xi$}.
\en
%
Legendre Polynomials are orthogonal in $\mathbb{L}^2$:
\eq
\int_{-1}^{1}P_{N_1}(\xi)P_{N_2}(\xi)\;d\xi=\left\{
\begin{array}{cr}
0,  & (N_1\ne N_2),\\ 
1/(N_1+1/2), & (N_1=N_2).
\end{array}\right.
\en
%
We utilize Gauss-Lobatto-Legendre nodes as quadrature points
$\xi^N_i,\; 0\le i\le N$ as the zeroes of $(1-\xi^2)\partial_\xi P_N(\xi)$. 
Gauss-Lobatto-Legendre quadrature weights $\sigma^N_i$ for non-axial elements 
are given by
%
\eq
\sigma^N_i=\frac{2}{N(N+1)[P_N(\xi^N_i)]^2} 
\en
%
and the corresponding basis functions, the Lagrange interpolating functions 
$l_i^N(\xi)$, may be calculated via 
%
\eq
l_i^N(\xi)=\left\{
\begin{array}{ll}
(-1)^N \frac{(1-\xi)P^{\,\prime}_N(\xi)}{N(N+1)}, & i=0, \\ [8pt]
\frac{1}{N(N+1)P_N(\xi_i^N)}\frac{(1-\xi^2)P^{\,\prime}_N(\xi)}{\xi_i^N-\xi},&
0<i<N,\\[8pt]
\frac{(1+\xi)P^{\,\prime}_N(\xi)}{N(N+1)}, & i=N.
\end{array}\right.
\en
%
Note also that $l^{N}_i(\xi_j)=\delta_{ij}$, which is an important 
property inasmuch as it gives rise to the diagonality of the mass matrix.
Derivatives $\partial_\xi l_i^N(\xi)$ are found using 
eq.~(\ref{appeq:legendre_ode}), as \citep{fournierthesis}
%
\eq
\partial_\xi l_i(\xi_I)=
\left\{
\begin{array}{ll}
\frac{P_N(\xi_I)}{P_N(\xi_i)} \frac{1}{\xi_I-\xi_i} & i\ne I, \\ [8pt]
\frac{-N(N+1)}{4} & i=I=0, \\ [8pt]
\frac{N(N+1)}{4} & i=I=N, \\ [8pt]
0 & \textrm{otherwise}. 
\end{array}\right.
\en
%
All of the above holds true for polynomial 
representation in the $\eta$ direction for all elements, and the $\xi$ 
direction for non-axial elements. 
%
\subsubsection{Gauss-Lobatto-Jacobi (0,1) quadrature} \label{appsection:glj_ax}
%
A detailed description of different quadrature rules can be found in 
\cite{bernardi} and \citet[][Appendix~B]{karniadakis}, 
where our specific case of 
(0,1) is given by setting the integrand powers equal to $\alpha=0$ and 
$\beta=1$.
The polynomial representation of the $\xi$ direction within axial elements is 
constructed using a different set of polynomials, defined by
%
\eq
\bar{P}_N(\xi)=\frac{P_N(\xi)+P_{N+1}(\xi)}{1+\xi},
\en
%
satisfying the differential equation
%
\eq \label{appeq:jacobi_ode}
\partial_\xi\left((1+\xi^2)(1-\xi)\partial_\xi\bar{P}_N \right)=
-N(N+2)(1+\xi)\bar{P}_N.
\en
%
These polynomials are orthogonal in $\mathbb{L}^2_1$:
%
\eq
\int_{-1}^{1} \bar{P}_{N_1}(\xi)\bar{P}_{N_2}(\xi)(1+\xi)\;d\xi=\left\{
\begin{array}{cr}
0 & (N_1\ne N_2), \\
2/(N_1+1) & (N_1= N_2).
\end{array}\right.
\en
%
With starting values $\bar{P}_0(\xi)=1$ and 
$\bar{P}_1(\xi)=\frac{1}{2}(3\xi-1)$, the induction formula for $N\ge 1$ reads
%
\eq
\bar{P}_{N+1}(\xi)=
\left[\frac{2N+3}{N+2}\xi-\frac{1}{(2N+1)(N+2)}\right]\bar{P}_N(\xi)-
\frac{N(N+2)}{(2N+1)(2N+3)}\bar{P}_{N-1}(\xi).
\en
%
Here, we define Gauss-Lobatto-Jabobi points 
$\bar{\xi}^N_i,\; 0 \le i \le N$ as the zeroes of
$(1-\xi^2)\partial_\xi \bar{P}_N(\xi)$. 
Gauss-Lobatto-Jacobi (0,1) quadrature weights $\bar{\sigma}^N_i$ are
%
\eq
\bar{\sigma}^N_i=\frac{4}{N(N+2)\bar{P}_N^2(\bar{\xi}_i^N)}
\textrm{ for } 1 \le i \le N \textrm{ and } 
\bar{\sigma}_0^N=\frac{8}{N(N+2)(N+1)^2}.
\en
%
The corresponding basis functions may then be computed as
%
\eq
\bar{l}_i^N(\xi)=\left\{
\begin{array}{ll}
\frac{2(-1)^N(\xi-1)\partial_\xi \bar{P}_N(\xi) }{N(N+1)(N+2)} & i=0, \\ [8pt]
\frac{1 }{N(N+2)\bar{P}_N(\bar{\xi}_i^N )}
\frac{(1-\xi^2)\partial_\xi\bar{P}_N(\xi)}{\bar{\xi}_i^N-\xi} & 0<i<N,\\[8pt]
\frac{(1+\xi)\partial_\xi \bar{P}_N(\xi) }{N(N+2)} & i=N.
\end{array}\right.
\en
%
Again, derivatives $\partial_\xi\bar{l}_i^N(\xi)$ may be found using 
eq.~(\ref{appeq:jacobi_ode}), leading to \citep{fournierthesis}
%
\eq
\partial_\xi \bar{l}_i(\bar{\xi}_I)=\left\{
\begin{array}{ll}
\frac{-N(N+2)}{6} & i=I=0, \\ [8pt]
\frac{2(-1)^N \bar{P}_N(\bar{\xi}_I)}{(1+\bar{\xi}_I)(N+1)} & 
i=0,\; 1\le I \le N-1, \\ [8pt]
\frac{(-1)^N}{N+1} & i=0,\;I=N, \\ [8pt]
\frac{(-1)^{N+1}(N+1)}{2\bar{P}_N(\bar{\xi}_i)(1+\bar{\xi}_i)}
& 1 \le i \le N-1,\;I=0, \\ [8pt]
\frac{1}{\bar{\xi}_I-\bar{\xi}_i}
\frac{\bar{P}_N(\bar{\xi}_I)}{\bar{P}_N(\bar{\xi}_i)} & 
1 \le i \le N-1,\; 1\le I \le N-1, i\ne I, \\ [8pt]
\frac{-1}{2(1+\bar{\xi}_i)} & 1 \le i \le N -1 , \; I=i, \\ [8pt]
\frac{1}{\bar{P}_N(\bar{\xi}_I)(1-\bar{\xi}_i)} & 1 \le i \le N-1,\;I=N,\\[8pt]
\frac{(-1)^{N+1}(N+1)}{4} & i=N,\; I=0, \\ [8pt]
\frac{-\bar{P}_N(\bar{\xi}_I)}{(1-\bar{\xi}_I)} & i=N,\;1\le I \le N-1, \\[8pt]
\frac{N(N+2)-1}{4} & i=N,\; I=N.
\end{array}\right.
\en
%
%

%
\subsubsection{Discretized equations of motion}\label{section:discr_eq_motion}
%
In this section, we will present the equations of motion for the 
multipole source system in discretized form, elaborate the scheme for 
time marching, and present the expanded expressions for all terms involved. 
Upon inserting eqs~(\ref{eq:u_pol}) and (\ref{eq:w_pol}) into the 
equations of motion (\ref{monosmc})--(\ref{quadsmc}), we obtain a system 
of ordinary differential equations in time for each element
\eq \label{eq:ode}
\sum_{IJ}\sum_{\beta}w_\beta^{IJ}\Big[\sum_{ij}\sum_{\alpha}
M^{IJij}_{\beta\alpha}\ddot{u}_\alpha^{ij}(t)+
\sum_{ij}\sum_{\alpha}K^{IJij}_{\beta\alpha}
u_\alpha^{ij}(t)\Big]=\sum_{IJ}\sum_{\beta}w_\beta^{IJ}f^{IJ}_\beta(t),
\en
where a double dot denotes the second partial derivative with respect to time, 
$f_\beta^{IJ}$ are the expansion coefficients of the source term, and 
we abbreviated the elemental mass matrix by $M^{IJij}_{\beta\alpha}$
and the elemental stiffness matrix by $K^{IJij}_{\beta\alpha}$.
In matrix notation, let $\bK_e$ now be such a stiffness matrix of element $e$.
We may then merge these elemental (local) contributions to a 
block-diagonal matrix $\bK_{\rm{L}}$ (unassembled, global stiffness matrix) 
defined by
%
\eqa
\bK_{\rm L}=\left(
\begin{array}{cccccc}
\bK_1 & & & & &   \\
& \ddots & & & &  \\
& & \bK_{n}  & &   & \\
& & & \bar{\bGamma}\bar{\bK}_1\bar{\bGamma}  & & \\
& & & & \ddots & \\
& & & & & \bar{\bGamma}\bar{\bK}_{\bar{n}}\bar{\bGamma}\\
\end{array}
\right),
\ena
%
where $\bar{\bGamma}$ is the diagonal matrix that acts as a ``mask'' to set 
all components to zero which vanish on the axis in accordance with the 
essential axial boundary conditions eqs~(\ref{eq:ax_bc_qu})
\citep{nissen+:07a,fournier04,bernardi}. We associate with each distinct node 
(i.e., counting nodes at $\xi,\eta=-1,1$ only once) in $\Omega$ a unique 
global number \citep{dfm} upon which we define the global displacement $\bsfu$.
We shall always use sans serif font to refer 
to global matrices of this assembled form. A Boolean connectivity 
matrix $\bsfQ$ mapping $\bsfu$ to $\bu_{\rm{L}}$, i.e., 
$\bu_{\rm{L}}=\bsfQ\bsfu$, copies values from the global vector into the 
collection of local vectors (``scatter''). The transpose operation 
$\bsfQ^{\rm{T}}$ (``gather'') sums up element edge/corner
contributions. The joint application of $\bsfQ\bsfQ^{\rm{T}}$ is called the 
direct stiffness summation and constitutes the stiffness assembly stage; we
denote the global, assembled stiffness matrix as 
$\bsfK=\bsfQ^{\rm{T}}\bK_{\rm{L}}\bsfQ$. 
Summing over all elemental stiffness terms 
$w_\beta^{ij} K^{IJij}_{\beta\alpha}u_\alpha^{ij}$ as in eq.~(\ref{eq:ode}), 
we write 
%
\eq \label{eq:stiff_el_glob}
\sum^{n+\bar{n}}_{e=1}\bw_e^{\rm{T}}\bK_e \bu_e=
\bw_{\rm{L}}^{\rm{T}} \bK_{\rm{L}} \bu_{\rm L}=
\bsfw^{\rm{T}}\bsfQ^{\rm{T}} \bK_{\rm L} \bsfQ \bsfu = 
\bsfw^{\rm{T}} \bsfK \bsfu.
\en
%
The same concept applies to the mass and source terms $\bM_{\rm{L}}$ and 
$\mathbf{f}_{\rm{L}}$. Construction of a global coordinate mesh and its 
relation to elemental nodes are based upon a global numbering 
technique \citep{dfm}. To obtain the semi-discretized version of 
eqs~(\ref{monosmc})--(\ref{quadsmc}), we span the discrete space of test 
functions by setting each $\bsfw$ to $1$ at a specific grid point, and 
zero everywhere else. We can then drop summations over $(I,J)$ and 
$\beta$ to form an assembled, linear algebraic system of ordinary differential 
equations in time 
%
\eq \label{eq:global_ode}
\bsfM\ddot{\bsfu}(t)+\bsfK\bsfu(t)=\bsff(t).
\en
%
\subsubsection{Source terms}\label{section:source}
%
Due to the inherent cylindrical symmetry, point sources are always located 
on the axis such that $\rsubs=r_{\rm s} \bzh$ for moment tensors and 
$\rsubr=r_{\rm r} \bzh$ for point forces. 
We additionally assume here that sources coincide with mesh nodes which 
is not a compulsory limitation of the method, but will suffice for immediate 
purposes. Discretized moment-tensor source terms effectively get 
``spread out'' over the bearing element due to the polynomial representation of
the $\bdel \bw$ operation, as can be verified from the sum appearing in terms 
such as eq.~(\ref{eq:discrete_deriv}).
Point forces $p_{x,y,z}$ on the other hand remain non-zero only at 
$\rsubr=(0,r_{\rm r})$. 
%
We compute the non-zero components $f_\beta$ of the respective source types 
appearing in eqs~(\ref{monosmc})--(\ref{quadsmc}) for the monopole as
%
\eqa \label{eq:src_mo}
\lefteqn{f_z^{IJ}(p_z,t)=(2\pi)^{-1}
\delta_{I0}\delta_{JJ_{\rm r}} \delta(t),} \vspace*{0.2cm}\\
\lefteqn{f_z^{IJ}(M_{zz},t)= (2\pi)^{-1}M_0
\delta_{I0}l^{\,\prime}_J(\eta_{\rm s})z_\eta^{0J_{\rm s}}H(t),}
\vspace*{0.2cm}\\
\lefteqn{f_s^{IJ}\left(\left(M_{xx}+M_{yy}\right)/2,t\right)=
(2\pi)^{-1}M_0\delta_{JJ_{\rm s}}(1-\delta_{I0})
\bar{l}^{\,\prime}_I(\bar{\xi}_0)s_\xi^{0J_{\rm s}} H(t),}
\ena
%
for the dipole as
%
\eqa \label{eq:src_di}
\lefteqn{f_+^{IJ}(p_x,t)=f_+^{IJ}(p_y,t)=\pi^{-1}
\delta_{I0}\delta_{JJ_{\rm r}}\delta(t),} \vspace*{0.2cm}\\
\lefteqn{f_+^{IJ}(M_{xz},t)=f_+^{IJ}(M_{yz},t)=\pi^{-1}\delta_{I0}
l^{\,\prime}_J(\eta_{\rm s})z_\eta^{0J_{\rm s}}H(t),} 
\vspace*{0.2cm}\\
\lefteqn{f_z^{IJ}(M_{xz},t)=f_z^{IJ}(M_{yz},t)=
\pi^{-1}M_0\delta_{JJ_{\rm s}}(1-\delta_{I0})
\bar{l}^{\,\prime}_I(\bar{\xi}_0)s_\xi^{0J_{\rm s}} H(t),}
\ena
%
and for the quadrupole as
%
\eqa \label{eq:src_qu}
\lefteqn{f_{s,\phi}^{IJ}\left(M_{xy},t\right)=
f_{s,\phi}^{IJ}\left(\left(M_{xx}-M_{yy}\right)/2,t\right)=
\pi^{-1}M_0\delta_{JJ_{\rm s}}(1-\delta_{I0})\bar{l}^{\,\prime}_I(\bar{\xi}_0)
s_\xi^{0J_{\rm s}} H(t),}
\ena
%
where $M_0$ is the scalar moment.
\begin{figure}[htb!]
\begin{center}
\includegraphics[scale=0.6]{fig2.eps}
\caption{Moment-tensor source gallery. Due to the polynomial representation of 
functions in the spectral-element method, a point-like moment-tensor source,
containing $\bdel\mathbf{w}$, spreads out over the bearing 
element in its discretized form. The figure shows all non-zero source vector 
components $\mathbf{f}_\beta$ as defined in 
eqs~(\ref{eq:src_mo})--(\ref{eq:src_qu}) for 
the element containing the source at polynomial order $N=8$. 
Arrow lengths represent the relative values of the magnitude of 
$\mathbf{f}_\beta$. The physical source location on the axis $s=0$ is denoted 
by a large black dot. Neighboring elements (not shown) share 
the same values as the edge of the shown element, thus a maximum of four 
elements may have non-zero source terms. Note that the location and relative 
spacing of grid points within the element is different for directions 
parallel and perpendicular to the axis, reflecting the axial discretization in
terms of GLJ points for the $\xi$ (and $s$) directions, and GLL points for the 
$\eta$ (and $z$) directions
(see Sections~\ref{section:nonax},~\ref{section:ax} and 
Appendices~\ref{appsection:gll},~\ref{appsection:glj_ax}).}
\label{fig2}
\end{center}
\end{figure}
%
Fig.~\ref{fig2} illustrates the spatial distribution of functions 
$f^{IJ}_\beta$ for the moment-tensor sources to emphasize the spreading of 
the source representation across the whole element. 
The actual physical source location (large black dot) has zero
source-vector amplitude. Note also that up to three elements neighboring the 
source element carry non-zero values as assembled from the element edges 
shared with the source-bearing element. 
%
\subsubsection{Mass terms}\label{section:mass}
%
The mass matrix is diagonal by construction 
and therefore readily invertible. One can deduce from 
eq.~(\ref{eq:gll_quad_nonax}) that the mass terms of non-axial elements
(as in eq.~(\ref{eq:ode})) take the form for monopole sources, 
%
\eqa
\lefteqn{\sum_{ij}\sum_\alpha 
(M_{s\alpha}^{IJij}+M_{z\alpha}^{IJij})u_\alpha^{ij}=
\rho^{IJ}\sigma_I\sigma_Js^{IJ}{\mathcal J}^{IJ}(\ddot{u}^{IJ}_s+
\ddot{u}^{IJ}_z),}
\ena
for dipole sources,
\eqa
\lefteqn{\sum_{ij}\sum_\alpha (M_{+\alpha}^{IJij}+M_{-\alpha}^{IJij}+
M_{z\alpha}^{IJij})u_\alpha^{ij}=
\rho^{IJ}\sigma_I\sigma_Js^{IJ}{\mathcal J}^{IJ}(2\ddot{u}^{IJ}_+ + 
2\ddot{u}^{IJ}_- +
\ddot{u}^{IJ}_z),}
\ena
%
and for quadrupole sources,
\eqa
\lefteqn{\sum_{ij}\sum_\alpha (M_{s\alpha}^{IJij}+M_{\phi\alpha}^{IJij}+
M_{z\alpha}^{IJij})u_\alpha^{ij}=
\rho^{IJ}\sigma_I\sigma_Js^{IJ}{\mathcal J}^{IJ}(\ddot{u}^{IJ}_s+
\ddot{u}^{IJ}_\phi+
\ddot{u}^{IJ}_z).}
\ena
%
The factors preceding the acceleration components are statically precomputed 
and inverted, hence the operation $\bM^{-1}$ in step 2 of 
eq.~(\ref{eq:time_marching}) is merely a point-by-point multiplication. 
The axial case is equivalent after the substitution
$\sigma_I\rightarrow\bar{\sigma}_I(1+\bar{\xi}_I)^{-1}$ and application of 
L'Hospital's rule on the axis: 
$s^{0J}(1+\bar{\xi}_0)^{-1}=\left(\partial_\xi s\right)^{0J}$.
%
%
\subsubsection{Stiffness terms}\label{section:stiffness}
%
\begin{table*} 
\begin{minipage}{156mm}
\begin{center}
\caption{Definitions for precomputable matrices of the stiffness term
($\epsilon=\lambda,~\mu$ or any combination thereof).}
\label{table:precomp} 
\begin{tabular}{@{}llllll}
\hline\hline
Matrix & Non-axial elements & Axial elements $(i>0)$ & Axial elements $(i=0)$\\
\hline\\
${}_{\epsilon}{A}^{ij}$  &  
$\epsilon^{ij}\sigma_i\sigma_j (s^{ij})^{-1}{\mathcal J}^{ij}$ &
$\epsilon^{ij}
\bar{\sigma_i}(1+\bar{\xi}_i)^{-1}\sigma_j (s^{ij})^{-1}{\mathcal J}^{ij}$ &
${}_\epsilon{A}^{0j}=\epsilon^{0j}
\bar{\sigma_0} \sigma_j {\mathcal J}^{0j}(s_\xi^{0j})^{-1}$ \\[10pt]
${}_\epsilon{B}_{z_\eta}^{ij}$ & 
$\epsilon^{ij}\sigma_i\sigma_j z_\eta^{ij}$ &
$\epsilon^{ij}\bar{\sigma_i}(1+\bar{\xi}_i)^{-1}\sigma_j z_\eta^{ij}$ &
${}_\epsilon{B}^{0j}_{z_\eta}=\epsilon^{0j}
\bar{\sigma_0}\sigma_j z_\eta^{0j}={}_\epsilon{A}^{0j}   $ \\[10pt]
$C^{ij}$ & 
$\sigma_i\sigma_j s^{ij} ({\mathcal J}^{ij})^{-1}$ &
$\bar{\sigma_i}\sigma_js^{ij}(1+\bar{\xi}_i)^{-1} ({\mathcal J}^{ij})^{-1}$ &
$C^{0j}=\bar{\sigma_0}\sigma_js_\xi^{0j} ({\mathcal J}^{0j})^{-1}$ \\[10pt]
$D_{\xi}^{Ii}$ & 
$\partial_\xi l_I(\xi_i)$ &
$\partial_\xi \bar{l}_I(\bar{\xi}_i)$ &
$\partial_\xi \bar{l}_I(\bar{\xi}_0)$ \\[10pt]
$D_{\eta}^{Jj}$ & 
$\partial_\eta l_J(\eta_j)=\partial_\xi l_J(\xi_j)$ &
$\partial_\eta l_J(\eta_j)$ & \\[5pt]
%
\hline\hline
\end{tabular}
%
\vspace{0.2cm}\\ 
\begin{tabular}{@{}llllll}
$(G_k^{xy})^{ij}$ &$\hspace{1em} k=1$ &$\hspace{1em} k=2$ &$\hspace{1em} k=3 $
&$\hspace{1em} k=4$ \\
\hline\vspace{0.1cm} 
$x=s,\; y=s$ &$\hspace{1em} z_\xi^{ij} z_\eta^{ij} $
&$\hspace{1em} z_\eta^{ij} z_\eta^{ij}$ &$\hspace{1em} z_\xi^{ij} 
z_\eta^{ij}$ &$\hspace{1em} z_\xi^{ij} z_\xi^{ij}$ \\[5pt]
$x=s,\; y=z$ &$\hspace{1em} z_\eta^{ij} s_\xi^{ij} $
&$\hspace{1em} z_\eta^{ij} s_\eta^{ij}$ &$\hspace{1em} z_\xi^{ij} 
s_\eta^{ij}$ &$\hspace{1em} z_\xi^{ij} s_\xi^{ij}$ \\[5pt]
$x=z,\; y=s$ &$\hspace{1em} s_\eta^{ij} z_\xi^{ij}$
&$\hspace{1em} s_\eta^{ij} z_\eta^{ij}$ &$\hspace{1em} s_\xi^{ij} 
z_\eta^{ij}$ &$\hspace{1em} s_\xi^{ij} z_\xi^{ij}$ \\[5pt]
$x=z,\; y=z$ &$\hspace{1em} s_\xi^{ij} s_\eta^{ij}$ &$\hspace{1em}
s_\eta^{ij} s_\eta^{ij} $&$\hspace{1em} s_\xi^{ij} 
s_\eta^{ij}$ &$\hspace{1em} s_\xi^{ij} s_\xi^{ij}$ \\[5pt] 
\hline 
\end{tabular}
\end{center}
\end{minipage}
\end{table*}
%
We relegate the description of the composite stiffness term to 
Appendix~\ref{appsection:stiffness} and only sketch the solution to the 
different types of integrals here.
Discretization yields a number of precomputable quantities 
$A$, $B$, $C$, $D$, and $G$ which are collectively defined 
in Table~\ref{table:precomp} for axial and non-axial elements using the 
definitions in Section~\ref{section:nonax} and the axial expressions from 
Section~\ref{section:ax}.
The terms involved in the stiffness matrix are computationally 
most demanding and need to be optimized. The two contributing factors of (i) 
recasting terms to avoid repeated operations and 
(ii) usage of cache-access optimized tensor products are 
explained in Appendix~\ref{appsection:stiffness}.

Here, we simply depict the different types of integrals, 
e.g., for non-axial elements
%
\eqa \label{eq:scheme_wu}
\int_{\Omega_e} \epsilon \frac{w_\beta}{s} \frac{u_\alpha}{s}s\,ds\,dz \approx 
\sum_{IJ}w_\beta^{IJ} \left({_\epsilon}A^{IJ}u^{IJ}_\alpha\right)=
\sum_{IJ}w_\beta^{IJ} (R^\alpha_1)^{IJ},
\ena
%
\eqa \label{eq:scheme_wdu}
\int_{\Omega_e} \epsilon \frac{w_\beta}{s} \partial_x u_\alpha
s\,ds\,dz \approx 
\sum_{IJ}w_\beta^{IJ}
\left({}_\epsilon{B}_{\chi_\eta}^{IJ}\sum_iD_\xi^{iI}u^{iJ}_\alpha
+{}_\epsilon{B}_{\chi_\xi}^{IJ}\sum_jD_\eta^{jJ}u^{Ij}_\alpha\right)=
\sum_{IJ}w_\beta^{IJ} (R^\alpha_2)^{IJ},
\ena
%
\eqa \label{eq:scheme_dwu}
\int_{\Omega_e}\epsilon \partial_x w_\beta \frac{u_\alpha}{s}s\,ds\,dz \approx 
\sum_{IJ}w_\beta^{IJ}
\left(\sum_i {}_\epsilon{B}_{\chi_\eta}^{iJ}D_\xi^{Ii}u^{iJ}_\alpha+
\sum_j {}_\epsilon{B}_{\chi_\xi}^{Ij}D_\eta^{Jj}u^{Ij}_\alpha\right)=
\sum_{IJ}w_\beta^{IJ} (R^\alpha_3)^{IJ},
\ena
%
\eqa \label{eq:scheme_dwdu}
\lefteqn{\int_{\Omega_e}
\epsilon \partial_x w_\beta \partial_y u_\alpha s\,ds\,dz 
}  \nonumber \\
&&\mbox{}\hspace{-0.9em}
\approx\sum_{IJ}w_\beta^{IJ}\Big[
\sum_i D_\xi^{Ii} C^{iJ}(G_{1}^{xy})^{iJ} \sum_j D_\eta^{jJ}u^{ij}_\alpha+
\sum_p D_\xi^{Ip} C^{pJ}(G_{2}^{xy})^{pJ} \sum_i D_\xi^{ip}u^{iJ}_\alpha 
\nonumber\\
&&\mbox{}\hspace{-0.9em}
+\sum_j D_\eta^{Jj} C^{Ij}(G_{3}^{xy})^{Ij}\sum_i D_\xi^{iI}u^{ij}_\alpha+
\sum_q D_\eta^{Jq} C^{Iq}(G_{4}^{xy})^{Iq}\sum_j D_\eta^{jq}u^{Ij}_\alpha\Big]=
\sum_{IJ}w_\beta^{IJ} (R^\alpha_4)^{IJ},
\ena
%
where the precomputed matrices $A^{ij}$, $B_{\chi_\xi}^{ij}$, $C^{ij}$ as 
given in Table~\ref{table:precomp} may generally contain integration weights, 
the Jacobian, mapping derivatives, the coordinate $s$, and elastic parameters. 
If $\partial_x=\partial_s$ then $\chi=z$, and if $\partial_x=\partial_z$ then 
$\chi=s$. We abbreviated the quantities in brackets as 
$(R^\alpha_k)^{IJ}$, which contain the actual operations to be carried out 
to obtain $\bK\bu$, the stiffness matrix acting on the displacement.
In addition, we collected twice-appearing mapping derivatives into 
$(G_k^{xy})^{ij}$, defined in Table~\ref{table:precomp}.
%
Inspecting the equations of motion (\ref{monosmc})--(\ref{quadsmc}), 
we note that whenever factors $s^{-1} w_\beta$ appear, 
$w_\beta$ vanishes at the axis; and equivalently for $u_\alpha$. Thus, 
a number of sums are developed for the two cases $I>0$ and $I=0$, 
resulting in summation for $I>0$ and additional terms for $I=0$. An 
equivalent scenario prevails for $u_\alpha$ and the corresponding summation 
over $i$. The respective integrals for axial elements are then approximated as
%
\eq \label{eq:scheme_wu_ax}
\int_{\bar{\Omega}^e} \epsilon \frac{w_\beta}{s} \frac{u_\alpha}
{s}s\,ds\,dz \approx 
\sum_{I>0\,J}w_\beta^{IJ} \left({_\epsilon}A^{IJ}u^{IJ}_\beta
+D_\xi^{I0}A^{0J}\sum_{i>0} D_\xi^{i0}u_\beta^{iJ}\right)=
\sum_{I>0\,J}w_\beta^{IJ} (\bar{R}^\alpha_1)^{IJ},
\en
%
\eqa \label{eq:scheme_wdu_ax}
\lefteqn{\int_{\bar{\Omega}^e} \epsilon \frac{w_\beta}{s} \partial_x u_\alpha
s\,ds\,dz  }\nonumber \\
&&\mbox{}
\approx\sum_{I>0\,J}w_\beta^{IJ}
\left[{}_\epsilon{B}_{\chi_\eta}^{IJ}\sum_iD_\xi^{iI}u^{iJ}_\alpha
+{}_\epsilon{B}_{\chi_\xi}^{IJ}\sum_jD_\eta^{jJ}u^{Ij}_\alpha 
+\bar{L}_1(u_\alpha)^{IJ}\right]=
\sum_{I>0\,J}w_\beta^{IJ} (\bar{R}^\alpha_2)^{IJ},
\ena
%
\eqa \label{eq:scheme_dwu_ax}
\lefteqn{\int_{\bar{\Omega}^e}\epsilon \partial_x w_\beta \frac{u_\alpha}
{s}s\,ds\,dz }\nonumber \\
&&\mbox{}
 \approx\sum_{IJ}w_\beta^{IJ}
\left[\sum_{i>0} {}_{_\epsilon}B_{\chi_\eta}^{iJ}D_\xi^{Ii}u^{iJ}_\alpha+
\sum_j {}_{_\epsilon}B_{\chi_\xi}^{Ij}D_\eta^{Jj}u^{Ij}_\alpha
+\bar{L}_2(u_\alpha)^{IJ}\right]=
\sum_{IJ}w_\beta^{IJ} (\bar{R}^\alpha_3)^{IJ},
\ena
%
\eqa \label{eq:scheme_dwdu_ax}
\lefteqn{\int_{\bar{\Omega}^e}
\epsilon \partial_x w_\beta \partial_y u_\alpha s\,ds\,dz 
}\nonumber \\
&&\mbox{}\hspace{-0.9em} 
\approx \sum_{IJ}w_\beta^{IJ}\Big[
\sum_i D_\xi^{Ii} C^{iJ}(G_{1}^{xy})^{iJ} \sum_j D_\eta^{jJ}u^{ij}_\alpha+
\sum_p D_\xi^{Ip} C^{pJ}(G_{2}^{xy})^{pJ} \sum_i D_\xi^{ip}u^{iJ}_\alpha 
\nonumber\\
&&\mbox{}\hspace{-0.9em}
+\sum_j D_\eta^{Jj} C^{Ij}(G_{3}^{xy})^{Ij}\sum_i D_\xi^{iI}u^{ij}_\alpha+
\sum_q D_\eta^{Jq} C^{Iq}(G_{4}^{xy})^{Iq}\sum_j D_\eta^{jq}u^{Ij}_\alpha\Big]=
\sum_{IJ}w_\beta^{IJ} (\bar{R}^\alpha_4)^{IJ}.
\ena
%
We do not  explicitly differentiate between axial and non-axial 
precomputed matrices of Table~\ref{table:precomp} 
but rather acknowledge that the usage automatically 
conforms with the respective definitions.
The additional terms $\bar{L}_{1,2}^{IJ}$ depend on 
the direction of the derivative:
%
\eqa \label{eq:stiff_add_ax_s}
\lefteqn{\textrm{Case }\partial_x=\partial_s:\;\;
\bar{L}_1(u_\alpha)^{IJ}=\bar{L}_2(u_\alpha)^{IJ}=
D_\xi^{I0}{}_\epsilon{A}^{0J}\sum_i D_\xi^{i0}u_\alpha^{iJ},}\\
\lefteqn{\textrm{Case }\partial_x=\partial_z:\;\;
\bar{L}_1(u_\alpha)^{IJ}=
D_\xi^{I0}{}_\epsilon{B}_{s_\xi}^{0J}\sum_j D_\eta^{jJ}u_\alpha^{0j}+
{}_\epsilon{B}_{s_\xi}^{0J}\sum_j D_\eta^{jJ}\sum_i D_\xi^{i0}u_\alpha^{ij},}
\label{eq:stiff_add_ax_z1}\\
\lefteqn{\mbox{}\hspace{7.2em}\bar{L}_2(u_\alpha)^{IJ}=
D_\xi^{I0}\sum_j D_\eta^{Jj}{}_\epsilon{B}_{s_\xi}^{0j}u_\alpha^{0j}+
\sum_j D_\eta^{Jj}{}_\epsilon{B}_{s_\xi}^{0j}\sum_{i>0} D_\xi^{i0}u_\alpha^{ij}.}
\label{eq:stiff_add_ax_z2}
\ena
%
Since $z_\xi|_{s=0}=s_\eta|_{s=0}=0$, there are no terms involving 
$B^{0j}_{z_\xi}$ and $B^{0j}_{s_\eta}$.
Also note that in the case of $\partial_z$ 
(eq.~(\ref{eq:stiff_add_ax_z1})--(\ref{eq:stiff_add_ax_z2})), the respective 
former term depends upon the displacement along the axis and contributes to 
the whole axial element in $(\bar{R}^\alpha_{(2,3)})^{IJ}$, whereas the latter 
term is synthesized from the displacement in the whole element and only 
inhabits the axis within $(\bar{R}^\alpha_{(2,3)})^{0J}$. Hence, if e.g., 
$u_\alpha|_{s=0}=w_\beta|_{s=0}=0$, then there is no additional axial term 
for $\partial_x=\partial_z$, i.e. $\bar{L}_1=\bar{L}_2=0$.
%
The expressions $(R_k^\alpha)^{IJ}$ and 
$(\bar{R}_k^\alpha)^{IJ}$ are the building blocks for the discretized 
stiffness term, subject to rearrangement and tensorization 
for optimization purposes. All details are discussed in 
Appendix~\ref{appsection:stiffness}. In summary, the stiffness term is 
approximated as follows:
%
\begin{enumerate}
\item Axial masking of displacement $\bu$,
\item Computation of elemental stiffness terms $\bK\bu$,
\item Axial masking of $\bK\bu$,
\item Assembly to obtain $\bsfK\bsfu$.
\end{enumerate}
%
\subsubsection{Matrix notation, tensor products and recast stiffness system}
\label{appsection:stiffness_mxm}
%
\begin{table*} 
\begin{minipage}{150mm}
\caption{Definitions of tensor notations and product operations, 
based on definitions in Table~\ref{table:precomp}. }
\label{apptable:matrix_op}
\begin{tabular}{@{}lllll}
 &&&&\\
Type & Matrix notation & Index notation & Axial vector 
& Axial index \\
\hline\hline\\
Precomputed tensor & ${}_\epsilon \bB_{s_\xi}$ & 
${}_\epsilon B_{s_\xi}^{ij}$ & ${}_\epsilon \bB^0_{s_\xi}$ & 
${}_\epsilon B_{s_\xi}^{0j}$ \\[10pt]
Derivative tensor $\xi$ & $\bD_\xi$ & $D_\xi^{Ii}$ & 
$\bD_{\xi}^0$ & $D^{I0}_{xi}$\\[10pt]
Transpose of $\bD_\xi$ & $\bD_\xi^{\rm{T}}$ & $D_\xi^{iI}$ & 
$(\bD_{\xi}^0)^{\rm{T}}$ & $D^{i0}_{\xi}$ \\[10pt]
Derivative tensor $\eta$ & $\bD_\eta$ & $D_\eta^{jJ}$ & 
&\\[10pt]
Matrix product & $\bX=\bu\otimes\bD_{\eta}$ & 
$
%X^{IJ}=
\sum_j u^{Ij}D^{jJ}_{\eta}$ & $\bX^0=\bu^0\otimes \bD_\eta$ & 
$X^{0J}=
\sum_j u^{0j}D^{jJ}_\eta$ \\[10pt]
Hadamard product & $\bX=\bA\odot \bu$ & 
$
%X^{IJ}=
A^{IJ}u^{IJ}$ &
$\bX^0={}_\epsilon\bB^0_{s_\xi}\odot\bu^0$ & 
$X^{0j}={}_\epsilon B_{z_\eta}^{0j}u^{0j}$\\[10pt]
Dyadic product & & & $\bX=\bD^0_\xi\bu^0$ & $X^{IJ}=D_\xi^{I0}u^{0J}$ \\
\hline
\end{tabular}
\end{minipage}
\end{table*}
%
In the interest of a succinct description, we revert to matrix/vector 
notation as defined in Table~\ref{apptable:matrix_op}.
Let us start by rewriting the non-axial expressions 
eqs~(\ref{eq:scheme_wu})--(\ref{eq:scheme_dwdu}) in terms of these matrix 
operations as
%
\eq \label{appeq:R1}
\bR^\alpha_1={}_{\epsilon}\bA\odot\bu_\alpha,
\en
%
\eq \label{appeq:R2}
\bR^\alpha_2={}_\epsilon\bB_{\chi_\eta}\odot
\left(\bD_\xi^{\rm{T}}\otimes\bu_\alpha\right)+
{}_\epsilon\bB_{\chi_\xi}\odot\left(\bu_\alpha\otimes\bD_\eta\right),
\en
%
\eq \label{appeq:R3}
\bR^\alpha_3=\bD_\xi\otimes\left({}_\epsilon\bB_{\chi_\eta}\odot
\bu_\alpha\right)+
\left({}_\epsilon\bB_{\chi_\xi}\odot\bu_\alpha\right)\otimes\bD_\eta^{\rm{T}},
\en
%
\eqa \label{appeq:R4}
\lefteqn{
\bR^\alpha_4=\bD_\xi\otimes\left[\bC\odot\bG_1^{xy}\odot
\left(\bu_\alpha\otimes\bD_\eta\right)\right]+
\bD_\xi\otimes\left[\bC\odot\bG_2^{xy}\odot
\left(\bD_\xi^{\rm{T}}\otimes\bu_\alpha\right)\right]
}\nonumber\\&&\mbox{}\hspace{0.2em}
+\left[\bC\odot\bG_2^{xy}\odot
\left(\bD_\xi^{\rm{T}}\otimes\bu_\alpha\right)\right]\otimes\bD_\eta^{\rm{T}}+
\left[\bC\odot\bG_3^{xy}\odot
\left(\bu_\alpha\otimes\bD_\eta\right)\right]\otimes\bD_\eta^{\rm{T}}.
\ena
%
The axial discretization
eqs~(\ref{eq:scheme_wu_ax})--(\ref{eq:scheme_dwdu_ax}) takes the form 
%
\eq \label{appeq:R1ax}
\bar{\bR}^\alpha_1=\bR^\alpha_1+
\bD_\xi^0 \left[{}_{\epsilon}\bA_0\odot
\left((\bD_\xi^0)^{\rm{T}}\otimes\bu_\alpha\right)\right],
\en
%
\eq \label{appeq:R2ax}
\bar{\bR}^\alpha_2=
\bR^\alpha_2+\left\{
\begin{array}{lr}
\bD^0_\xi\left[
{}_\epsilon\bA_0\odot
\left((\bD_\xi^0)^{\rm{T}}\otimes\bu_\alpha\right)\right]
& (\partial_x=\partial_s) \vspace{0.5em}\\ 
\bD^0_\xi\left[{}_\epsilon\bB^0_{s_\xi}\odot
\left(\bu^0_{\alpha}\otimes\bD_{\eta}\right)\right]+
{}_\epsilon\bB^0_{s_\xi}\odot\left[\left((\bD_\xi^0)^{\rm{T}}\otimes
\bu_\alpha\right)
\otimes\bD_{\eta}\right] & (\partial_x=\partial_z) 
\end{array}\right\},
\en
%
\eq \label{appeq:R3ax}
\bar{\bR}^\alpha_3=
\bR^\alpha_3+\left\{
\begin{array}{lr}
\bD^0_\xi\left[
_\epsilon\bA_0\odot\left((\bD_\xi^0)^{\rm{T}}\otimes
\bu_\alpha\right)\right]
& (\partial_x=\partial_s) \vspace{0.5em}\\ 
\bD^0_\xi\left[
\left(_\epsilon\bB^0_{s_\xi}\odot\bu^0_\alpha\right)\otimes
\bD_{\eta}^{\rm{T}}\right]+
\left[_\epsilon\bB^0_{s_\xi}\odot
\left((\bD_\xi^0)^{\rm{T}}\otimes\bu_\alpha\right)\right]
\otimes\bD_{\eta}^{\rm{T}} & (\partial_x=\partial_z) 
\end{array}\right\},
\en
%
\eqa \label{appeq:R4ax}
\bar{\bR}^\alpha_4=\bR^\alpha_4.
\ena
%
We assumed here that whenever summation starts at $I>0$ 
as in eqs~(\ref{eq:scheme_wu_ax})--(\ref{eq:scheme_dwu_ax}), 
then precomputed matrices vanish at $I=0$.
Both elemental operations $\otimes$ and $\odot$ are optimized using 
unrolled loops and unit-stride cache access \citep{dfm}. 
Schematically, we start out by computing all necessary $\otimes$ and $\odot$ 
operations, and then successively add different terms before 
computing the next round of operations $\otimes$ or $\odot$. 
%
For notational clarity, let $S=\sum_\beta S_\beta$, $\beta=s,\phi,z$ 
or $\beta=+,-,z$ be the original, continuous elemental stiffness integral 
such that $S\approx\bw^{\rm{T}} \bK \bu$. We are then interested in evaluating 
the ``components'' $(\bK\bu)_\beta$. Additionally, we rearrange the integrand 
into the two sets
$S_\beta=S_{\beta}^{\partial\partial}+S_{\beta}^{\partial}$ and accordingly, 
$(\bK\bu)_\beta=(\bK\bu)_\beta^{\partial\partial}+(\bK\bu)_\beta^{\partial}$,
where $S_{\beta}^{\partial\partial}$ is the collection of integrands 
containing derivatives of both $w_\beta$ and $u_\alpha$ (``leading order''), 
and $S_{\beta}^{\partial}$ is the remaining part with maximally one derivative
 (``lower order'').
%
%
\subsubsection{Leading order terms}
%
For completeness, the analytical, initial integral form of the leading order 
terms is, for the monopole
%
\eq
S_s^{\partial\partial}=\int_{\Omega_e}\big[\left(\lambda+2\mu\right)
\partial_s{w_s}\partial_s{u_s} +
\lambda\partial_s{w_s}\partial_z{u_z} +
\mu\partial_z{w_s}\left(\partial_s{u_z}+
\partial_z{u_s}\right)\big]s\,ds\,dz,
\en
%
\eq
S_z^{\partial\partial}=\int_{\Omega_e}
\big[\left(\lambda+2\mu\right)\partial_z{w_z}\partial_z{u_z}+
\lambda\partial_z{w_z}\partial_s{u_s} +
\mu\partial_s{w_z}\left(\partial_s{u_z}+\partial_z{u_s}\right)\big]s\,ds\,dz,
\en
%
and for the dipole,
%
\eqa
\lefteqn{S^{\partial\partial}_+=\int_{\Omega_e}
\big[\left(\lambda+3\mu\right)\partial_s{w_+}\partial_s u_+ +
\left(\lambda+\mu\right)\partial_s{w_+}\partial_s u_- +
2\mu\partial_z{w_+}\partial_z{u_+}}
\nonumber\\&&\mbox{} %extra 
+\lambda\partial_s w_+ \partial_z u_z +
\mu\partial_z w_+ \partial_s u_z
\big]s\,ds\,dz,
\ena
%
%
\eqa
\lefteqn{S^{\partial\partial}_-=\int_{\Omega_e}
\big[\left(\lambda+3\mu\right)\partial_s{w_-}\partial_s u_- +
\left(\lambda+\mu\right)\partial_s{w_-}\partial_s u_+ 
+2\mu\partial_z{w_-}\partial_z{u_-} }
\nonumber\\&&\mbox{} %extra
+\lambda\partial_s w_- \partial_z u_z +
\mu\partial_z w_- \partial_s u_z 
\big]s\,ds\,dz,
\ena
%
\eqa
\lefteqn{S^{\partial\partial}_z=\int_{\Omega_e}
\big[\partial_z{w_z}\left[\left(\lambda+2\mu\right)\partial_z u_z +
\lambda\left(\partial_s u_+ +\partial_s u_- \right)\right]}
\nonumber\\&&\mbox{} %extra
+\mu\partial_s{w_z}\left(\partial_s{u_z}+
\partial_z{u_+}+\partial_z{u_-}\right)\big]s\,ds\,dz.
\ena
%
%
In the quadrupole case, $S^{\partial\partial}_s$ and $S^{\partial\partial}_z$ 
are identical to the monopole case, and we get additionally
\eq
S^{\partial\partial}_\phi=\int_{\Omega_e}
\big[\partial_s{w_\phi}\partial_s{u_\phi}+
\partial_z{w_\phi}\partial_z{u_\phi}\big]s\,ds\,dz.
\en
%
Rearranging these terms into groups of identical or partly similar operations, 
we end up with a system of precomputed matrices detailed in 
Table~A2
for the terms of the $\bR^\alpha_4$-type and 
each source type. For any given element, this entire leading-order part of the 
stiffness matrix is then compacted into 
%
\eqa \label{eq:stiff_4th_sums}
\lefteqn{\left(\bK\bU\right)_\beta^{\partial\partial}=\sum_{\alpha}
\biggl\{\bD_\xi\otimes\left[\bC\odot\bE_{\beta\alpha}^{(1)}
\odot\left(\bu_\alpha\otimes\bD_\eta\right)\right]
+\bD_\xi\otimes\left[\bC\odot\bE_{\beta\alpha}^{(2)}
\odot\left(\bD^{\rm{T}}_\eta\otimes\bu_\alpha\right)\right]} \nonumber\\
&&\mbox{}\hspace{4.5em}+\left[\bC\odot\bE_{\beta\alpha}^{(3)}
\odot\left(\bD^{\rm{T}}_\eta\otimes\bu_\alpha\right)\right]\otimes
\bD^{\rm{T}}_\xi
+\left[\bC\odot\bE_{\beta\alpha}^{(4)}
\odot\left(\bu_\alpha\otimes\bD_\eta\right)\right]\otimes
\bD^{\rm{T}}_\xi\biggr\},
\ena
%
where the difference between source types is concentrated in 
$\bE_{\beta\alpha}^{(k)}$. 
The quantities $\bE_{\beta\alpha}^{(k)}$ are listed in 
Table~A2 for monopole, dipole, and quadrupole 
source types, respectively. The difference in axial versus non-axial elements 
is abundant in quantities $\bC$ and $\bD$ as defined in 
Table~\ref{table:precomp}. The explicit expansion of all lower order 
terms for axial or non-axial elements is given in the following sections
for each source type separately.
%
\begin{sidewaystable}
{\bf Table A2.} Precomputed matrices $(E_{\beta\alpha}^{(k)})^{ij}$ of the 
leading order terms.\\
\begin{tabular}{ | l | l  l  l  l }
$\alpha$ & $k=1$ & $k=2$  & $k=3$ & $k=4$ \\ \hline\hline
\multicolumn{5}{|c|} {Monopole \& Quadrupole $\beta=s$}  \\[3pt]
$s$ &  $(\lambda^{ij}+
2\mu^{ij})z_\xi^{ij}z_\eta^{ij} +\mu^{ij}s_\xi^{ij}s_\eta^{ij}$ & 
$(\lambda^{ij}+2\mu^{ij})(z_\eta^{ij})^2+\mu^{ij}(s_\eta^{ij})^2$ & 
$(\lambda^{ij}+2\mu^{ij})z_\xi^{ij}z_\eta^{ij} +
\mu^{ij}s_\xi^{ij}s_\eta^{ij}$ & 
$(\lambda^{ij}+2\mu^{ij})(z_\xi^{ij})^2+\mu^{ij}(s_\xi^{ij})^2$\\
$z$ & 
$\lambda^{ij}s_\xi^{ij}z_\eta^{ij}+\mu^{ij}z_\xi^{ij}s_\eta^{ij}$ & 
$(\lambda^{ij}+\mu^{ij})s_\eta^{ij}z_\eta^{ij}$ & 
$\lambda^{ij}z_\xi^{ij}s_\eta^{ij}+\mu^{ij}s_\xi^{ij}z_\eta^{ij}$ & 
$(\lambda^{ij}+\mu^{ij})s_\xi^{ij}z_\xi^{ij}$ \\
\hline
\multicolumn{5}{|c|} {Monopole \& Quadrupole $\beta=z$}  \\[3pt]
$s$  & $\mu^{ij}z_\xi^{ij}z_\eta^{ij} +(\lambda^{ij}+
2\mu^{ij})s_\xi^{ij}s_\eta^{ij}$ & 
$\mu^{ij}(z_\eta^{ij})^2+(\lambda^{ij}+2\mu^{ij})(s_\eta^{ij})^2$ & 
$\mu^{ij}z_\xi^{ij}z_\eta^{ij} +(\lambda^{ij}+
2\mu^{ij})s_\xi^{ij}s_\eta^{ij}$ & 
$\mu^{ij}(z_\xi^{ij})^2+(\lambda^{ij}+2\mu^{ij})(s_\xi^{ij})^2$\\
$z$ & 
$\lambda^{ij}z_\xi^{ij}s_\eta^{ij}+\mu^{ij}s_\xi^{ij}z_\eta^{ij}$ & 
$(\lambda^{ij}+\mu^{ij})s_\eta^{ij}z_\eta^{ij}$ & 
$\lambda^{ij}s_\xi^{ij}z_\eta^{ij}+\mu^{ij}z_\xi^{ij}s_\eta^{ij}$ & 
$(\lambda^{ij}+\mu^{ij})s_\xi^{ij}z_\xi^{ij}$ \\
\hline\hline 
\multicolumn{5}{|c|} {Quadrupole $\beta=\phi$}  \\[3pt]
$\phi$ & 
$\mu^{ij}(z_\xi^{ij}z_\eta^{ij} +s_\xi^{ij}s_\eta^{ij})$ & 
$\mu^{ij}((z_\eta^{ij})^2+(s_\eta^{ij})^2)$ & 
$\mu^{ij}(z_\xi^{ij}z_\eta^{ij} +s_\xi^{ij}s_\eta^{ij})$ & 
$\mu^{ij}((z_\xi^{ij})^2+(s_\xi^{ij})^2)$ \\                      
\hline\hline
\multicolumn{5}{|c|} {Dipole $\beta=+$}  \\[3pt]
$+$ & $(\lambda^{ij}+
3\mu^{ij})z_\xi^{ij}z_\eta^{ij} +
2\mu^{ij}s_\xi^{ij}s_\eta^{ij}$ & 
$(\lambda^{ij}+3\mu^{ij})(z_\eta^{ij})^2 +
2\mu^{ij}(s_\eta^{ij})^2$ & 
$(\lambda^{ij}+3\mu^{ij})z_\xi^{ij}z_\eta^{ij} +
2\mu^{ij}s_\xi^{ij}s_\eta^{ij}$ &
$(\lambda^{ij}+3\mu^{ij})(z_\xi^{ij})^2 +
2\mu^{ij}(s_\xi^{ij})^2$ \\
$-$ & $(\lambda^{ij}+
\mu^{ij})z_\xi^{ij}z_\eta^{ij}$ & 
$(\lambda^{ij}+\mu^{ij})(z_\eta^{ij})^2$ & 
$(\lambda^{ij}+\mu^{ij})z_\xi^{ij}z_\eta^{ij}$ &
$(\lambda^{ij}+\mu^{ij})(z_\xi^{ij})^2$ \\ 
$z$ & 
$\lambda^{ij}s_\xi^{ij}z_\eta^{ij}+\mu^{ij}z_\xi^{ij}s_\eta^{ij}$ & 
$(\lambda^{ij}+\mu^{ij})s_\eta^{ij}z_\eta^{ij}$ & 
$\lambda^{ij}z_\xi^{ij}s_\eta^{ij}+\mu^{ij}s_\xi^{ij}z_\eta^{ij}$ & 
$(\lambda^{ij}+\mu^{ij})s_\xi^{ij}z_\xi^{ij}$ \\
\hline
\multicolumn{5}{|c|} {Dipole $\beta=-$}  \\[3pt]
$+$ &  $(\lambda^{ij}+
\mu^{ij})z_\xi^{ij}z_\eta^{ij}$ & 
$(\lambda^{ij}+\mu^{ij})(z_\eta^{ij})^2$ & 
$(\lambda^{ij}+\mu^{ij})z_\xi^{ij}z_\eta^{ij}$ &
$(\lambda^{ij}+\mu^{ij})(z_\xi^{ij})^2$ \\
$-$ & $(\lambda^{ij}+
3\mu^{ij})z_\xi^{ij}z_\eta^{ij} +
2\mu^{ij}s_\xi^{ij}s_\eta^{ij}$ & 
$(\lambda^{ij}+3\mu^{ij})(z_\eta^{ij})^2 +
2\mu^{ij}(s_\eta^{ij})^2$ & 
$(\lambda^{ij}+3\mu^{ij})z_\xi^{ij}z_\eta^{ij} +
2\mu^{ij}s_\xi^{ij}s_\eta^{ij}$ &
$(\lambda^{ij}+3\mu^{ij})(z_\xi^{ij})^2 +
2\mu^{ij}(s_\xi^{ij})^2$ \\ 
$z$ & 
$\lambda^{ij}s_\xi^{ij}z_\eta^{ij}+\mu^{ij}z_\xi^{ij}s_\eta^{ij}$ & 
$(\lambda^{ij}+\mu^{ij})s_\eta^{ij}z_\eta^{ij}$ & 
$\lambda^{ij}z_\xi^{ij}s_\eta^{ij}+\mu^{ij}s_\xi^{ij}z_\eta^{ij}$ & 
$(\lambda^{ij}+\mu^{ij})s_\xi^{ij}z_\xi^{ij}$ \\
\hline
\multicolumn{5}{|c|} {Dipole $\beta=z$}  \\[3pt]
$\pm$ & 
$\lambda^{ij}z_\xi^{ij}s_\eta^{ij}+\mu^{ij}s_\xi^{ij}z_\eta^{ij}$ & 
$(\lambda^{ij}+\mu^{ij})s_\eta^{ij}z_\eta^{ij}$ & 
$\lambda^{ij}s_\xi^{ij}z_\eta^{ij}+\mu^{ij}z_\xi^{ij}s_\eta^{ij}$ & 
$(\lambda^{ij}+\mu^{ij})s_\xi^{ij}z_\xi^{ij}$ \\
$z$ & $\mu^{ij}z_\xi^{ij}z_\eta^{ij} +
(\lambda^{ij}+2\mu^{ij})s_\xi^{ij}s_\eta^{ij}$ & 
$\mu^{ij}(z_\eta^{ij})^2 +
(\lambda^{ij}+2\mu^{ij})(s_\eta^{ij})^2$ &
 $\mu^{ij}z_\xi^{ij}z_\eta^{ij} +
(\lambda^{ij}+2\mu^{ij})s_\xi^{ij}s_\eta^{ij}$ &
$\mu^{ij}(z_\xi^{ij})^2 +(\lambda^{ij}+ 2\mu^{ij})(s_\xi^{ij})^2$ \\ 
\hline\hline              
\end{tabular}
\label{apptable:precomp_E}
\end{sidewaystable}
%
%
\subsubsection{Lower order terms: monopole}
%
The lower order terms for the monopole case are
%
\eq
S_s^{\partial}=\int_{\Omega_e}\Big\{\left(\lambda+2\mu\right)
\frac{w_s{u_s}}{s^2} +\lambda\left[\partial_s{w_s}\frac{u_s}{s}
+\frac{w_s}{s}\left(\partial_s{u_s}+\partial_z{u_z}\right)\right]\Big\}
s\,ds\,dz,
\en
%
\eq
S_z^{\partial}=\int_{\Omega_e}\lambda\partial_z{w_z}\frac{u_s}{s}s\,ds\,dz.
\en
%
Honoring the additional terms exclusive to axial elements 
(see eqs~(\ref{appeq:R1ax})--(\ref{appeq:R3ax})), we will develop the 
discretization with these additional terms in curly braces being 
preceded by $\delta_{e\bar{e}}$, such that the description applies to 
all elements. The discretized form for any element unfolds as
%
\eqa \label{appeq:loword_monos}
\lefteqn{
(\bK\bU)_s^{\partial}=
{}_\lambda{\bB}_{s_\xi} \odot \left(\bu_z\otimes\bD_{\eta}\right)+
{}_\lambda{\bB}_{s_\eta} \odot \left(\bD_\xi^{\rm{T}}\otimes{\bu_z}\right)+
{}_\lambda{\bB}_{z_\xi} \odot 
\left(\bu_s{}\otimes\bD_{\eta}\right) }\nonumber\\
&&\mbox{}\hspace{1.3em} 
+{_\lambda\bB}_{z_\eta} \odot 
\left(\bD_\xi^{\rm{T}}\otimes{\bu_s}\right)+
\bD_{\xi}\otimes\left({}_\lambda{\bB}_{z_\eta} \odot {\bu_s}\right)
+\left({}_\lambda{\bB}_{z_\xi} \odot {\bu_s}\right)\otimes\bD_{\eta}^{\rm{T}}
\nonumber\\
&&\mbox{}\hspace{1.3em}
+{}_{(\lambda+2\mu)}{\bA}\odot\bu_s +
\delta_{e\bar{e}}\Big\{\bD^0_{\xi}
\left[{}_\lambda{\bB}^0_{s_\xi} \odot \left(\bu_z^0\otimes\bD_{\eta}\right)+
{}_{3\lambda+2\mu}{\bA_0} \odot 
\left(\left(\bD^0_{\xi}\right)^{T}\otimes\bu_s\right)
\right]\Big\},
\ena
%
\eq \label{appeq:loword_monoz}
(\bK\bU)_z^{\partial}=\bD_{\xi}\otimes
\left({{}_\lambda\bB}_{s_\eta} \odot {\bu_s}\right)+
\left({{}_\lambda\bB}_{s_\xi} \odot 
{\bu_s}\right)\otimes{\bD_{\eta}^{\rm{T}}}+
\delta_{e\bar{e}}\Big\{\Bigl[{}_\lambda\bB^0_{s_\xi} \odot 
\left(\left(\bD^0_\xi\right)^{\rm{T}}
\otimes\bu_s\right)\Bigr]\otimes\bD_{\eta}^{\rm{T}}\Big\}.
\en
%
%
\subsubsection{Lower order terms: dipole}
%
We again write the the dipole system in the $(+,-,z)$ coordinate system to 
properly implement the axial boundary conditions in eq.~(\ref{eq:ax_bc_di}). 
After rearrangement, the stiffness system of lower order terms 
then takes the form
%
\eq
S^{\partial}_+=\int_{\Omega_e}
\Big\{2\left(\lambda+\mu\right)\partial_s{w_+}\frac{u_-}{s}
+\mu\partial_z{w_+}\frac{u_z}{s}\Big\}s\,ds\,dz,
\en
%
\eqa
\lefteqn{S^{\partial}_-=\int_{\Omega_e}\Big\{2\left(\lambda-\mu\right)
\partial_s{w_-}\frac{u_-}{s}-\mu\partial_z{w_-}\frac{u_z}{s} }\nonumber\\
&&\mbox{}\hspace{2.5em} 
+2\frac{w_-}{s}\left[2\left(\lambda+3\mu\right)\frac{u_-}{s}+
\left(\lambda+\mu\right)\partial_s u_+ + 
\left(\lambda-\mu\right)\partial_s u_- + \lambda \partial_z u_z\right]\Big\}
s\,ds\,dz,
\ena
%
\eq
S^{\partial}_z=\int_{\Omega_e}\Big\{2\lambda\partial_z{w_z}
\frac{u_-}{s}+\mu\frac{w_z}{s}\left(\partial_z u_+ 
-\partial_z u_- +\frac{u_z}{s}\right)\Big\}
s\,ds\,dz.
\en
%
For any element, the discretized version of the lower order terms 
of the dipole stiffness term is
%
\eqa
\lefteqn{(\bK\bU)_+^{\partial} = \bD_\xi\otimes
\left({}_{2\left(\lambda+\mu\right)}\bB_{z_\eta}\odot \bu_-\right) + 
\left({}_{2\left(\lambda+\mu\right)}\bB_{z_\xi}\odot \bu_-\right)
\otimes \bD^{\rm{T}}_\eta 
}\nonumber\\
&&\mbox{}\hspace{2.0em} +\bD_\xi\otimes\left({}_\mu\bB_{s_\eta}\odot \bu_z\right) + 
\left({}_\mu\bB_{s_\xi}\odot \bu_z\right)\otimes \bD^{\rm{T}}_\eta  
\nonumber\\
&&\mbox{}\hspace{2.0em} +\delta_{e\bar{e}}\Big\{\bD^0_{\xi}
\left[{}_{2\left(\lambda+\mu\right)}{\bA_0} \odot 
\left(\left(\bD^0_\xi\right)^{\rm{T}} \otimes \bu_-\right)\right] + 
\left[{}_{\mu}\bB^0_{s_\xi} \odot 
\left(\left(\bD^0_\xi\right)^{\rm{T}} \otimes \bu_z\right)\right]
 \otimes \bD_\eta^{\rm{T}}\Big\},
\ena
%
\eqa
\lefteqn{(\bK\bU)_-^{\partial} =\bD_\xi\otimes
\left({}_{2\left(\lambda-\mu\right)}\bB_{z_\eta}\odot \bu_-\right) + 
\left({}_{2\left(\lambda-\mu\right)}\bB_{z_\xi}\odot \bu_-\right)
\otimes \bD^{\rm{T}}_\eta -
\bD_\xi\otimes\left({}_\mu\bB_{s_\eta}\odot \bu_z\right) }\nonumber\\
&&\mbox{}\hspace{2.2em} 
-\left({}_\mu\bB_{s_\xi}\odot \bu_z\right)\otimes \bD^{\rm{T}}_\eta 
 + {}_{4\left(\lambda+3\mu\right)}\bA\odot \bu_- +
2{}_{\lambda}\bB_{s_\xi}\odot\left(\bu_z\otimes\bD_\eta\right)
\nonumber\\&&\mbox{}\hspace{2.2em}
+2{}_{\lambda}\bB_{s_\eta}\odot\left(\bD^{\rm{T}}_\xi\otimes \bu_z\right)
+ {}_{2\left(\lambda+\mu\right)}\bB_{z_\eta}\odot
\left(\bD^{\rm{T}}_\xi\otimes \bu_+\right) 
+ {}_{2\left(\lambda-\mu\right)}\bB_{z_\eta}\odot
\left(\bD^{\rm{T}}_\xi\otimes \bu_-\right)
\nonumber\\&&\mbox{}\hspace{2.2em}
+ {}_{2\left(\lambda+\mu\right)}\bB_{z_\xi}\odot
\left(\bu_+\otimes \bD_\eta\right) 
+{}_{2\left(\lambda-\mu\right)}\bB_{z_\xi}\odot
\left(\bu_-\otimes \bD_\eta\right)  
\nonumber\\&&\mbox{}\hspace{2.2em}
+ \delta_{e\bar{e}}\Big\{\bD^0_\xi 
\Bigl[{}_{2\left(\lambda+\mu\right)}{\bA_0} \odot 
\left(4\left(\bD^0_\xi\right)^{\rm{T}} \otimes \bu_-+
\left(\bD^0_\xi\right)^{\rm{T}} \otimes \bu_+\right)\Bigr]
\Big\},
\ena
%
\eqa
\lefteqn{(\bK\bU)_z^{\partial} =
\bD_\xi\otimes
\left({}_{2\lambda}\bB_{s_\eta}\odot \bu_-\right) + 
\left({}_{2\lambda}\bB_{s_\xi}\odot \bu_-\right)
\otimes \bD^{\rm{T}}_\eta + {}_\mu\bA\odot \bu_z 
}\nonumber\\
&&\mbox{}\hspace{2.2em}
+{}_\mu\bB_{s_\xi}\odot
\left[\left(\bu_+ - \bu_-\right)\otimes \bD_\eta\right]
+ {}_\mu\bB_{s_\eta}\odot
\left[\bD^{\rm{T}}_\xi\otimes \left(\bu_+-\bu_-\right)\right]
\nonumber\\
&&\mbox{}\hspace{2.2em} 
+\delta_{e\bar{e}}\Big\{
\bD^0_\xi\left[{}_{\mu}{\bA_0} \odot 
\left(\left(\bD^0_\xi\right)^{\rm{T}} \otimes \bu_z\right)+
{}_{\mu}\bB^0_{s_\xi} \odot \left(\bu^0_+ \otimes \bD_\eta\right)\right]\Big\}.
\ena
%
This system is computationally slightly more expensive than the monopole case.
%
%
\subsubsection{Lower order terms: quadrupole}
%
In the quadrupole case, lower order terms of the integral form 
eq.~(\ref{quadsmc}) are
%
\begin{eqnarray}\label{appeq:loword_quads}
\lefteqn{S^{\partial}_s=\int_{\Omega_e}\Big\{\left(\lambda+2\mu\right)
\frac{w_s}{s^2}\left(u_s-2u_\phi\right)
+ \lambda\left[\partial_s{w_s}\frac{u_s-2u_\phi}{s}
+\frac{w_s}{s}\left(\partial_s{u_s}+\partial_z{u_z}\right)\right] }
\nonumber\\
&&\mbox{}\hspace{2.5em} %extra
+2\mu\frac{w_s}{s}\left(\partial_s{u_\phi}+
\frac{2u_s-u_\phi}{s}\right)\Big\}s\,ds\,dz,
\end{eqnarray}
%
\begin{eqnarray}\label{appeq:loword_quadphi}
\lefteqn{S^{\partial}_\phi=\int_{\Omega_e}\Big\{\frac{w_\phi}{s}\left[
-\left(2\lambda+6\mu\right)\frac{u_s}{s} 
+\left(4\lambda+9\mu\right)\frac{u_\phi}{s} -
\lambda\partial_s{u_s}+\lambda\partial_z{u_z} -\mu \partial_s u_\phi\right] }
\nonumber\\
&&\mbox{}\hspace{2.5em} %extra
+\mu\partial_s{w_\phi}\frac{2u_s-u_\phi}{s} +
2\mu\partial_z{w_\phi}\frac{u_z}{s}\Big\}s\,ds\,dz,
\end{eqnarray}
%
\begin{eqnarray}\label{appeq:loword_quadz}
\lefteqn{S^{\partial}_z=\int_{\Omega_e}\Big\{\lambda\partial_z{w_z}
\frac{u_s-2u_\phi}{s}+2\mu\frac{w_z}{s}\left(\partial_z{u_\phi}+
\frac{2u_z}{s}\right)\Big\}s\,ds\,dz.}
\end{eqnarray}
%
The discretized version for the lower order terms in the quadrupole case reads
%
\begin{eqnarray}
\lefteqn{
(\bK\bU)_s^{\partial}=
{{}_\lambda\bB}_{s_\xi} \odot \left(\bu_z\otimes\bD_{\eta}\right)+
{{}_\lambda\bB}_{s_\eta} \odot \left(\bD_\xi^{\rm{T}}\otimes{\bu_z}\right)+
{{}_\lambda\bB}_{z_\xi} \odot 
\left(\bu_s{}\otimes\bD_{\eta}\right)}
\nonumber\\&&\mbox{}\hspace{2.2em}
+{{}_\lambda\bB}_{z_\eta} \odot \left(\bD_\xi^{\rm{T}}\otimes{\bu_s}\right)
+ \bD_{\xi}\otimes\left({{}_\lambda\bB}_{z_\eta} \odot {\bu_s}\right)
+\left({{}_\lambda\bB}_{z_\xi} \odot {\bu_s}\right)\otimes\bD_{\eta}^{\rm{T}}
\nonumber\\&&\mbox{}\hspace{2.2em}
+ _{(\lambda+6\mu)}\bA\odot\bu_s 
-\bD_\xi\otimes\left({}_{2\lambda}\bB_{z_\eta}\odot{\bu_\phi}\right)
- \left({}_{2\lambda}\bB_{z_\xi}\odot{\bu_\phi}\right)\otimes\bD^{\rm{T}}_\eta 
 \nonumber\\&&\mbox{}\hspace{2.2em} 
+{}_{2\mu}\bB_{z_\xi}\odot\left(\bu_\phi\otimes\bD_\eta\right)
+{}_{2\mu}\bB_{z_\eta}\odot\left(\bD^{\rm{T}}_\xi\otimes \bu_\phi\right)
 - {}_{2\lambda+6\mu}\bA\odot \bu_\phi
 \nonumber\\&&\mbox{}\hspace{2.2em}  
+ \delta_{e\bar{e}}\Bigl\{
\bD^0_\xi\Bigl[{}_{3(\lambda+2\mu)}{\bA_0}\odot
\left(\left(\bD^0_\xi\right)^{\rm{T}}\otimes \bu_s\right)
-{}_{4(\lambda+\mu)}{\bA_0}
\odot\left(\left(\bD^0_\xi\right)^{\rm{T}} \otimes \bu_\phi\right)\Bigr]
\Bigr\},
\end{eqnarray}
%
\begin{eqnarray}
\lefteqn{(\bK\bU)_\phi^{\partial} =
\bD_\xi\otimes\left({}_{2\mu}\bB_{z_\eta}\odot{\bu_s}\right)+
\left({}_{2\mu}\bB_{z_\xi}\odot{\bu_s}\right)\otimes\bD^{\rm{T}}_\eta-
\bD_\xi\otimes\left({}_{\mu}\bB_{z_\eta}\odot{\bu_\phi}\right)-
\left({}_{\mu}\bB_{z_\xi}\odot{\bu_\phi}\right)\otimes\bD^{\rm{T}}_\eta 
}\nonumber\\&&\mbox{}\hspace{2.2em}
+\bD_\xi\otimes\left(_{2\mu}\bB_{s_\eta}\odot{\bu_z}\right)
+\left({}_{2\mu}\bB_{s_\xi}\odot{\bu_z}\right)\otimes\bD^{\rm{T}}_\eta-
2{}_\lambda\bB_{z_\xi}\odot\left(\bu_s\otimes\bD_\eta\right)
\nonumber\\&&\mbox{}\hspace{2.2em} 
-2{}_\lambda\bB_{z_\eta}\odot\left(\bD^{\rm{T}}_\xi\otimes \bu_s\right) 
-2{}_\lambda\bB_{s_\xi}\odot\left(\bu_z\otimes\bD_\eta\right)
- 2{}_\lambda\bB_{s_\eta}\odot\left(\bD^{\rm{T}}_\xi\otimes \bu_z\right)
\nonumber\\&&\mbox{}\hspace{2.2em} 
-{}_{\mu}\bB_{z_\xi}\odot\left(\bu_\phi\otimes\bD_\eta\right)
+{}_{\mu}\bB_{z_\eta}\odot\left(\bD^{\rm{T}}_\xi\otimes \bu_\phi\right) 
-{}_{2\lambda+6\mu}\bA\odot \bu_s 
+ {}_{4\lambda+9\mu}\bA\odot \bu_\phi 
\nonumber\\&&\mbox{}\hspace{2.2em}
+\delta_{e\bar{e}}\Bigl\{\bD^0_\xi\Bigl[
{}_{-4(\lambda+\mu)}{\bA_0}
\odot\left(\left(\bD^0_\xi\right)^{\rm{T}} \otimes \bu_s\right)
+{}_{4\lambda+7\mu}{\bA_0}
\odot\left(\left(\bD^0_\xi\right)^{\rm{T}} \otimes \bu_\phi\right)\Bigr]
\Bigr\},
\end{eqnarray}
%
\begin{eqnarray}
\lefteqn{(\bK\bU)_z^{\partial} =
\bD_\xi\otimes\left({}_{\lambda}\bB_{s_\eta}\odot{\bu_s}\right)+
\left({}_{\lambda}\bB_{s_\xi}\odot{\bu_s}\right)\otimes\bD^{\rm{T}}_\eta-
\bD_\xi\otimes\left({}_{2\lambda}\bB_{s_\eta}\odot{\bu_\phi}\right)}
\nonumber\\&&\mbox{}\hspace{2.2em}
-\left({}_{2\lambda}\bB_{s_\xi}\odot{\bu_\phi}\right)\otimes\bD^{\rm{T}}_\eta
+{}_{2\mu}\bB_{s_\xi}\odot\left(\bu_\phi\otimes\bD_\eta\right)+
{}_{2\mu}\bB_{s_\eta}\odot\left(\bD^{\rm{T}}_\xi\otimes \bu_\phi\right)+
{}_{4\mu}\bA\odot \bu_z 
\nonumber\\&&\mbox{}\hspace{2.2em}
+\delta_{e\bar{e}}\Bigl\{
\bD^0_\xi\Bigl[{}_{4\mu}{\bA_0}\odot\left(\left(\bD^0_\xi\right)^{\rm{T}}\otimes 
\bu_z\right)\Bigr]\Bigr\}.
\end{eqnarray}
%
Since all components vanish at the axis, all non-zero additional terms are of 
the same kind and for $\partial_x=\partial_s$. 
Note that the non-axial quadrupole case includes the non-axial monopole case. 
%
\subsubsection{Fluid mass and stiffness terms}
%
Using the notation of \citet[][Section~4]{nissen+:07b}, the mass term inside 
the fluid region for a non-axial element becomes
%
\eqa
\lefteqn{\int_{\Omega_e^{\rm f}} \kappa^{-1} w \ddot{\chi}  s\,ds\, dz\approx
\sum_{IJ}w^{IJ}\sum_{ij}M_{\rm f}^{IJij}\ddot{\chi}^{ij}=
%}\nonumber\\
%\lefteqn{\hspace{1.5em}
\sum_{IJ}w^{IJ}
(\kappa^{IJ})^{-1}\sigma_I\sigma_Js^{IJ}{\mathcal J}^{IJ}\ddot{\chi}^{IJ}.}
\ena
%
Axial elements are treated equivalently after the substitution
$\sigma_I\rightarrow\bar{\sigma}_I(1+\bar{\xi}_I)^{-1}$ and application of 
L'Hospital's rule on the axis
$s^{0J}(1+\bar{\xi}_0)^{-1}=\left(\partial_\xi s\right)^{0J}$ 
\citep[][Section~3.3]{nissen+:07b}.
Note that the mass terms are the same for monopole, dipole and quadrupole 
source types, respectively, since the average over $\sin^2{\phi}$ and 
$\cos^2{\phi}$ is the same as over $\sin^2{2\phi}$ and $\cos^2{2\phi}$.
Compared to the lengthy solid stiffness terms in 
\citet[][Appendix~B]{nissen+:07b}, the fluid stiffness terms take on a 
straightforward, computationally light form. All source types contain 
the basic scalar, monopole Laplacian integral
%
\eqa \label{eq:scheme_dwdu}
\lefteqn{\int_{\Omega_e^{\rm f}}  \rho^{-1}\left(\partial_s{w}\partial_s{\chi}+
\partial_z{w}\partial_z{\chi}\right)s\,ds\,dz
\approx } \nonumber\\
\lefteqn{\hspace{3.5em}
\sum_{IJ}w^{IJ}\Big[
\sum_i D_\xi^{Ii} C^{iJ}G_{1}^{iJ} \sum_j D_\eta^{jJ}\chi^{ij}+
\sum_p D_\xi^{Ip} C^{pJ}G_{2}^{pJ} \sum_i D_\xi^{ip}\chi^{iJ}+ }
\nonumber\\
\lefteqn{\hspace{7.5em}
\sum_j D_\eta^{Jj} C^{Ij}G_{3}^{Ij}\sum_i D_\xi^{iI}\chi^{ij}+
\sum_q D_\eta^{Jq} C^{Iq}G_{4}^{Iq}\sum_j D_\eta^{jq}\chi^{Ij}\Big].}
\ena
%
The dipole fluid stiffness term is completed by adding to 
eq.~(\ref{eq:scheme_dwdu}) once, and with the quadrupole fluid stiffness
term four times, the expression
%
\eq \label{eq:scheme_wu_nonax}
\int_{\Omega_e^{\rm f}} \rho^{-1}s^{-2}w\chi s\,ds\,dz \approx
\sum_{IJ}w^{IJ} (\rho^{IJ})^{-1}\sigma_I\sigma_J (s^{IJ})^{-1}
{\mathcal J}^{IJ}\chi^{IJ}
\en
%
for non-axial elements and 
%\eq \label{eq:scheme_wu_ax}
%\int_{\rm FOC}\rho^{-1}s^{-2}w\chi s\,ds\,dz \approx
%\sum_{I>0\,J}w^{IJ} \sigma_J\left[
%(\rho^{IJ})^{-1}\sigma_I (s^{IJ})^{-1}
%{\mathcal J}^{IJ}\chi^{IJ}+
%D_{\xi}^{I0}(\rho^{0J})^{-1}\bar{\sigma}_0 {\mathcal J}^{0J} 
%(\partial_\xi s^{0J})^{-1}\sum_{i>0}D_{\xi}^{i0}\chi^{iJ}
%\right]
%\en
%
\eqa \label{eq:scheme_wu_ax}
\lefteqn{\int_{\Omega_e^{\rm f}} \rho^{-1}s^{-2}w\chi s\,ds\,dz \approx
} \nonumber\\
\lefteqn{\hspace{2.5em}
\sum_{I>0\,J}w^{IJ} \sigma_J\Bigl[
(\rho^{IJ})^{-1}\sigma_I (s^{IJ})^{-1}
{\mathcal J}^{IJ}\chi^{IJ}
+D_{\xi}^{I0}(\rho^{0J})^{-1}\bar{\sigma}_0 {\mathcal J}^{0J} 
(\partial_\xi s^{0J})^{-1}\sum_{i>0}D_{\xi}^{i0}\chi^{iJ}\Bigr]}
\ena
%
for axial elements. We abbreviated several quantities as 
in \citet[][Table~1]{nissen+:07b}:
%
\eqa
\lefteqn{C^{ij}=\sigma_i\sigma_j s^{ij} ({\mathcal J}^{ij})^{-1}\;
\textrm{(non-axial)},}\nonumber\\
\lefteqn{C^{ij}=\bar{\sigma_i}\sigma_j s^{ij}(1+\bar{\xi}_i)^{-1}
({\mathcal J}^{ij})^{-1}\;\textrm{(axial, $i>0$)} ,}\nonumber\\
\lefteqn{C^{0j}=\bar{\sigma_0}\sigma_j s_\xi^{0j} ({\mathcal J}^{0j})^{-1} 
\;\textrm{(axial, $i=0$)}, }\nonumber\\
\lefteqn{D_{\xi}^{Ii}=\partial_\xi l_I(\xi_i),\;
D_{\eta}^{Jj}=\partial_\eta l_J(\eta_j)=\partial_\xi l_J(\xi_j),}\nonumber\\
\lefteqn{G_{1}^{iJ}=z_\xi^{iJ}z_\eta^{iJ}+s_\xi^{iJ}s_\eta^{iJ},\;
G_{2}^{pJ}=z_\eta^{pJ} z_\eta^{pJ}+ s_\eta^{pJ} s_\eta^{pJ},}\nonumber\\
\lefteqn{G_{3}^{Ij}=z_\xi^{Ij}z_\eta^{Ij}+s_\xi^{Ij}s_\eta^{Ij},\;
G_{4}^{Iq}=z_\xi^{Iq} z_\xi^{Iq}+ s_\xi^{Iq} s_\xi^{Iq}.}\nonumber
\ena
%
where e.g. $s_\xi=\partial_\xi s$.
% 
\begin{figure}[b!]
\begin{center}
\includegraphics[scale=0.5]{boundaryplot.eps}
\caption{Sketch to illustrate our nomenclature around a solid-fluid boundary 
(in this case CMB). $r_{\rm sf}$ is the radius, and $\Sigma_{\rm f}^{\rm s}$ 
the surface of the boundary. $(.,.)$ denote the respective elemental indices 
$(I,J)$ of Gauss-Lobatto-Legendre points of the boundary-hugging corner 
points of elements above and below.}
\label{img:bdry}
\end{center}
\end{figure}
%
\subsubsection{Solid-fluid coupling terms}
%
Let $\sum=\sum^{\rm s}_{\rm f}\bigcup\sum^{\rm f}_{\rm s}$ denote the union 
of solid-fluid boundaries such that we can e.g. specify those non-axial 
element edges in the solid that touch the boundary at radius $r_{\rm sf}$ 
from above by $\sum^{\rm s}_e$ and those axial element edges in the fluid that 
touch the boundary from below by $\overline{\sum}_{\rm f}^e$, as shown in 
Fig.~\ref{img:bdry}. Note that only purely circular element types 
\citep[see][Appendix~A]{nissen+:07b} may be located at 
discontinuities.
We also define the colatitudinal element boundaries such that within an element
$\theta_1\leq\theta\leq\theta_2$ for the northern hemisphere and vice versa 
for the southern hemisphere.

We can then represent the two types of 
boundary terms as
%
\eqa \label{eq:boundary_solid}
\lefteqn{\int_{\sum_e^{\rm s}}\left(w_s\sin{\theta}+
w_z\cos{\theta}\right)\ddot{\chi}\;r^2\,\sin{\theta}\;d\theta\approx} 
\nonumber\\
\lefteqn{\hspace{2.5em}\frac{r^2_{\rm sf}}{2}\left|\theta_2-\theta_1\right|\sum_{I}
\sigma_I\left(\sin{\theta}\right)_{I0} 
\left[w_s^{I0}\left(\sin{\theta}\right)_{I0}+
w_z^{I0}\left(\cos{\theta}\right)_{I0}\right]\ddot{\chi}_{I0}}
\ena
%
for non-axial elements inside the solid region, and
%
\eqa \label{eq:boundary_fluid}
\lefteqn{\int_{\sum^e_{\rm f}}w\left(u_s\sin{\theta}+
u_z\cos{\theta}\right)\;r^2\,\sin{\theta}\;d\theta\approx } 
\nonumber\\
\lefteqn{\hspace{2.5em} \frac{r^2_{\rm sf}}{2}\left|\theta_2-\theta_1\right| \sum_I w^{I0}
\sigma_I\left(\sin{\theta}\right)_{I0}
\left[u_s^{I0}\left(\sin{\theta}\right)_{I0}+
u_z^{I0}\left(\cos{\theta}\right)_{I0}\right]}
\ena
%
for non-axial elements inside the fluid. Note that the statically 
precomputable factors such as
$\frac{r^2_{\rm sf}}{2}\left|\theta_2-\theta_1\right|\sigma_I
\left(\sin{\theta}\right)_{I0}^2$ 
are identical in both cases and on both sides of the boundary since our
potential formulation eq.~(\ref{eq:potential}) results in such 
material-independent boundary terms. Furthermore, the coupling terms are 
identical for all excitation types for the same reason as the mass terms.
For axial elements, we again replace 
$\sigma_I$ by $\bar{\sigma}_I\left(1+\xi_0\right)^{-1}$
in eqs~(\ref{eq:boundary_solid}) and (\ref{eq:boundary_fluid}), respectively.
For the axis $I=0$ itself, we apply L'Hospital's rule to remove 
the singularity: 
$\textstyle{\left(\sin{\theta}\right)_{00} (1+\xi_0)^{-1}=
(\partial_\xi{s})_{00}=\frac{1}{2}\theta_2r_{\rm sf}}$. 
%
In practice, the solid-fluid earth model leads to two distinct domains 
with respective global numbers such that the corresponding indices from 
both sides in the boundary terms need to be found based on a coordinate search.
Note that being a 1-D operator, the boundary term is computationally invisible 
to the overall cost of the scheme, similar to purely axial stiffness terms.

\subsection{Time marching}
Upon inserting the polynomial function representation of our spectral-element
approach \citep[][eqs~(12)--(13)]{nissen+:07b}, the 2-D set of weak equations 
translates to its discretized counterpart, a coupled, global system of 
ordinary differential equations in time, which, following the same procedure 
as by \citet[][Section~4]{nissen+:07b}, reads 
%
\eqa \label{eq:global_ode}
\lefteqn{\bsfM_{\rm s}\ddot{\bsfu}(t)+\bsfK_{\rm s}\bsfu(t)+
\bsfB_{\rm s}^{\rm f}\ddot{\bsfchi}(t)=\bsfzero,}\label{eq:global_ode1}\\
\lefteqn{\bsfM_{\rm f}\ddot{\bsfchi}(t)+\bsfK_{\rm f}\bsfchi(t)-
\bsfB_{\rm f}^{\rm s}\bsfu(t)+\bsfB_{\rm s}^{\rm f}\bsfu(t)=\bsfzero ,}
\label{eq:global_ode2}\\
\lefteqn{\bsfM_{\rm s}\ddot{\bsfu}(t)+\bsfK_{\rm s}\bsfu(t)+
\bsfB^{\rm s}_{\rm f}\ddot{\bsfchi}(t)=\bsff(t).}\label{eq:global_ode3}
\ena
%
Temporal evolution problems such as 
eqs~(\ref{eq:global_ode1})--(\ref{eq:global_ode3}) are ubiquitous in a 
variety of physical phenomena, hence a zoo of possible numerical 
solutions exists. 
The common principle is to find a numerical expression for the ODE type
$\ddot{u}=f(u)$ at time $t$ discretized by steps $\Delta t$
after applying the inverse mass matrix ($\bsfM^{-1}$), and calculating
internal (e.g., $\bsfK\bsfu$) and external ($\bsff$) forces.
In seismology, a second-order Newmark scheme has primarily been applied to 
solve such systems \citep{KoVi98}. 
Here, we employ this Newmark scheme, but will also argue for symplectic 
integration based on superior accuracy at small additional cost and 
straightforward implementation. We explicitly present the respective schemes 
in this section and provide a tutorial-style introduction and summary of 
symplectic integration in Appendix~\ref{appsection:symplectic}. 
For brevity, let us concatenate internal and external forces for the 
fluid as 
$\bsfXi_{\rm f}\left(\bsfu,\bsfchi\right)=
-\bsfK_{\rm f}\bsfchi+\bsfB_{\rm f}\bsfu$ and solid domains as
$\bsfXi_{\rm s}\left(\bsfu,\ddot{\bsfchi}\right)=
-\bsfK_{\rm s}\bsfu-\bsfB_{\rm s}
\ddot{\bsfchi}
+\bsff$.
%
%
\subsubsection{Second-order Newmark scheme}
%
The second-order, acceleration-driven explicit Newmark scheme we employed in 
\citet{nissen+:07b} is extended to the solid-fluid case and reads 
\citep{manu03}
\eqa 
\lefteqn{\bsfchi(t+\Delta{t})=\bsfchi(t)+
\Delta{t}\dot{\bsfchi}(t)+
\textstyle{\frac{1}{2}}\Delta{t}^2\ddot{\bsfchi}(t),}\label{eq:newmark1}\\
\lefteqn{\bsfu(t+\Delta{t})=\bsfu(t)+\Delta{t}\dot{\bsfu}(t)+
\textstyle{\frac{1}{2}}\Delta{t}^2\ddot{\bsfu}(t),}\label{eq:newmark2}\\
\lefteqn{\ddot{\bsfchi}(t+\Delta{t})=\bsfM_{\rm f}^{-1}
\bsfXi_{\rm f}\left[\bsfu(t+\Delta t),\bsfchi(t+\Delta t)\right],}
\label{eq:newmark3}\\
\lefteqn{\ddot{\bsfu}(t+\Delta{t})=\bsfM_{\rm s}^{-1}
\bsfXi_{\rm s}\left[\bsfu(t+\Delta t),\ddot{\bsfchi}(t+\Delta t)\right],}
\label{eq:newmark4}\\
\lefteqn{\dot{\bsfchi}(t+\Delta{t})=\dot{\bsfchi}(t)+
\textstyle{\frac{1}{2}}\Delta{t}\left[\ddot{\bsfchi}(t)+
\ddot{\bsfchi}(t+\Delta{t})\right],}\label{eq:newmark5}\\
\lefteqn{\dot{\bsfu}(t+\Delta{t})=\dot{\bsfu}(t)+
\textstyle{\frac{1}{2}}\Delta{t}
\left[\ddot{\bsfu}(t)+\ddot{\bsfu}(t+\Delta{t})\right].}\label{eq:newmark6}
\ena
%
This scheme is fully explicit and requires one evaluation of the global 
forces per time step which is the dominant factor for CPU time. Note that the 
order of evaluation is important as we need $\ddot{\bsfchi}$ on 
solid-fluid boundaries from eq.~(\ref{eq:newmark3})
for the internal forces on the solid side (eq.~(\ref{eq:newmark4})).
%
\subsubsection{Fourth-order symplectic scheme} \label{section:symplectic}
%
Alternatively, we implemented a fourth-order symplectic 
scheme which has been suggested for the elastodynamic 
spectral-element method by \citet{Ampuero+:07}, following
\citet[][eq.~(22)]{omelyan+:02} and \citet{omelyan+:03}.
This essentially means that we evaluate all terms four 
times in between $t$ and $t+\Delta{t}$.
Setting $\bsfu_0=\bsfu(t),\,\dot{\bsfu}_0=\dot{\bsfu}(t)$ and 
$\bsfu_5=\bsfu(t+\Delta t),\,\dot{\bsfu}_4=\dot{\bsfu}(t+ \Delta t)$
(ditto for $\bsfchi,\,\dot{\bsfchi}$) and rewriting
$\bsfXi_{\rm s}\left(\bsfu,\bsfXi_{\rm f}\left(\bsfu,\bsfchi\right)\right)=
-\bsfK_{\rm s}\bsfu-\bsfB_{\rm s}
\bsfXi_{\rm f}\left(\bsfu,\bsfchi\right)
+\bsff$ to eliminate $\ddot{\bsfchi}$, the scheme reads
%
\eqa \label{eq:pefrl}
\lefteqn{\bsfchi_i= \bsfchi_{i-1}+\kappa_i \Delta t \dot{\bsfchi}_{i-1},}\\
\lefteqn{\bsfu_i= \bsfu_{i-1}+\kappa_i \Delta t \dot{\bsfu}_{i-1},}\\
\lefteqn{\dot{\bsfchi}_i= \dot{\bsfchi}_{i-1}+\pi_i \Delta t \bsfM^{-1}_{\rm f}
\bsfXi_{\rm f}\left(\bsfu_i,\bsfchi_i\right),}\\
\lefteqn{\dot{\bsfu}_i= \dot{\bsfu}_{i-1}+\pi_i \Delta t \bsfM^{-1}_{\rm s}
\bsfXi_{\rm s}\left(\bsfu_i,\bsfXi_{\rm f}(\bsfu_i,\bsfchi_i)\right),}\\
\lefteqn{\bsfchi_5= \bsfchi_{4}+\kappa_5 \Delta t \dot{\bsfchi}_{4},}\\
\lefteqn{\bsfu_5= \bsfu_{4}+\kappa_5 \Delta t \dot{\bsfu}_{4},}
\ena
%
where $i=1,...,4$ denotes an inner loop of updates that requires a four-fold
evaluation of all forces per time step. Note that this scheme is explicit, 
does not require storage of past time steps and avoids the usage of the 
acceleration resulting in less run-time memory occupation than the Newmark 
scheme eqs~(\ref{eq:newmark1})--(\ref{eq:newmark6}). 
Factorization coefficients $\kappa_i,\,\pi_i$ are usually determined 
by symbolic algebra; the position extended Forest-Ruth like 
(PEFRL) scheme we adopted takes the values
%
\eqa
\lefteqn{\kappa_1=\kappa_5=\xi,\,\kappa_2=\kappa_4=\chi,\,
\kappa_3=1-2(\chi+\xi),}\\
\lefteqn{\pi_1=\pi_4=1/2-\lambda,\,\pi_2=\pi_4=\lambda,}
\ena
%
where
\eqa
\lefteqn{\xi=+0.1786178958448091,}\\ 
\lefteqn{\lambda=-0.2123418310626054,\,\textrm{ and} }\\
\lefteqn{\chi=-0.06626458266981849.}
\ena
See Appendix~\ref{appsection:symplectic} for details on symplectic time 
integration and an outline to obtain factorization coefficients. 
%
%
In general, this scheme satisfies more stringent conditions on preserving 
geometrical properties such as total energy, (angular) momentum or 
time-reversal symmetry which are not generally respected by other schemes 
such as Newmark.
While short-time phenomena are expected to be sufficiently discretized 
using standard, non-symplectic techniques, any long-time issues including 
propagation over many wavelengths shall be much better represented by 
minimizing numerical dissipation of these invariants. 
More importantly, as shown in \citet{Ampuero+:07}, it is conceivable to consider 
that seismic-wave propagation is at, or at least will soon enter, the stage of deducing 
information from high-frequency arrivals at large offsets for which minimal 
dispersion and accurate techniques are crucial.

\subsection{On-the-fly accuracy tests}
%
In this appendix, we summarize some of our default, on-the-fly tests on various 
aspects of the method, in some instances critical cases for the 
demeanor of the settings and therefore subject to simulation homicide.
%
\subsubsection{Surfaces, volumes, and masses}
%
A trivial and fast test on the meshing and functional discretization 
is the computation of spherical surfaces and volumes compared to the analytical 
values. Following surface discretization expression eq.~(\ref{eq:boundary_solid}),
each spherical surface error $\Delta {\mathcal S}$ at radius 
$r_{\rm sph}$ is defined by the numerical difference to $4\pi r_{\rm sph}^2$ as
%
\eq
\Delta {\mathcal S}=\left| \left[\sum_e
\left| \theta_2^e - \theta_1^e \right| \sum_i  \sigma_i \left(\sin{\theta}\right)^e_i \right]-
2\right|,
\en
%
where we sum over all elements $e$ that touch boundary $r_{\rm sph}$. 
This effectively tests the quadrature rule and location of neighboring elements, 
and specifically the solid-fluid boundary term 
eqs~(\ref{eq:boundary_solid})--(\ref{eq:boundary_fluid}).
%
Equivalently, solid and fluid subvolume errors are given by
%
\eq
\Delta {\mathcal V}=\left| \left[\sum_e
\sum_{ij} M_{ij}^e  \right]- 
\frac{2}{3}\left(r^3_{\rm max} -r^3_{\rm min} \right)\right|,
\en
%
where $M_{ij}^e=\sigma_i\sigma_j s_{ij}^e {\mathcal J}_{ij}^e $ is the elemental, 
diagonal mass term and we sum over either solid or fluid elements $e$ and use their 
corresponding, bounding radii $r_{\rm max,min}$. This tests the actual mass 
matrices $M_{ij}^e$ as used in the time extrapolation schemes 
for both solid and fluid subdomains, the overall tiling into non-overlapping 
elements, and the location of solid and fluid boundaries. 
Finally, it is conceivable to repeat this volumetric exercise but honor the mass 
matrix' namesake by multiplying it with the local density and hence computing 
the mass of each spherical layer between discontinuities in the earth. 
This of course additionally takes the density structure, specifically the jumps 
across discontinuities and the honoring thereof into account. 
The reference mass is computed using a very fine radial resolution for each 
layer. Table~\ref{table:surfvol}
depicts some surface, volume, and mass errors for different mesh resolutions.
The actual values are somewhat insignificant since the errors are at machine 
precision, verifying the analytical mapping \citep[][Appendix~A]{nissen+:07b}. 
This test was done in single precision. Our code however defines all 
mesh coordinates in double precision by default, hence the surfaces which are 
purely geometric are at double precision.
%
\begin{table}[htb!]
\begin{center}
\caption{Accuracy of spherical surfaces ${\mathcal S}$, volumes ${\mathcal V}$, and masses 
${\mathcal M}$ for different mesh resolutions.}
\label{table:surfvol} 
\begin{tabular}{@{}ccccc}
&&&&\\
\hline\hline
$T_0\, [s]$ & $\max{\left[\Delta {\mathcal S}\right]}$ & $\Delta {\mathcal V_{\rm s}}$ & 
$\Delta {\mathcal V_{\rm f}}$ & $\max{\left[\Delta {\mathcal M}\right]}$  \\
\hline\\
$5$ & $3.4\cdot 10^{-15}$  & $4.97\cdot 10^{-8}$ & $1.02\cdot 10^{-8}$ & $2.96\cdot 10^{-8}$ \\[10pt]
$10$ & $1.6\cdot 10^{-15}$  & $4.97\cdot 10^{-8}$ & $1.02\cdot 10^{-8}$ & $2.96\cdot 10^{-8}$\\[10pt]
$20$ &  $1.4\cdot 10^{-15}$ & $2.72\cdot 10^{-8}$ & $1.02\cdot 10^{-8}$  & $2.96\cdot 10^{-8}$\\[10pt]
$40$ & $2.7\cdot 10^{-15}$  & $2.72\cdot 10^{-8}$ & $1.02\cdot 10^{-8}$ & $2.96\cdot 10^{-8}$ \\[10pt]
%
\hline
\end{tabular}
\end{center}
\end{table}
%
%
\subsubsection{Discrete material property variations}
%
Aside from computing radial profiles at the grid points to compare to the input 
polynomials that specify the background model, a specific check on the validity 
of our mesh is the variation of material 
properties within each element. Conventionally, it is permissible to discretize smoothly 
varying properties, especially with our polynomial orders above $4$. A generic check 
on the variation however certifies the lack of jumps in any element and gives an idea 
about the coarseness of the mesh with respect to background model variations. 
Table~\ref{table:material_across_elements} lists the variations in $\rho$, $\lambda$, 
and $\mu$ for different mesh resolutions. Of course, the variation for the same
 model directly scales with the resolution, i.e. element sizes, itself. Note that at 
$40\, \textrm{s}$, variations reach $10\,\%$.
%
\begin{table}[htb!]
\begin{center}
\caption{Material property variations across elements.}
\label{table:material_across_elements} 
\begin{tabular}{@{}ccccc}
&&&\\
\hline\hline
$T_0\, [s]$ & $\max{\left[\Delta \rho \right]}$ &  $\max{\left[\Delta \lambda \right]}$ & 
$\max{\left[\Delta \mu \right]}$ \\
\hline\\
$5$ & $4.2\cdot 10^{-3}$ &  $ 1.7\cdot 10^{-2}$  & $ 1.9\cdot 10^{-2}$ \\[10pt]
$10$ & $8.4\cdot 10^{-3}$  & $ 3.5\cdot 10^{-2}$  & $ 3.7\cdot 10^{-2}$ \\[10pt]
$20$ &  $1.7\cdot 10^{-2}$  & $ 6.6\cdot 10^{-2}$  & $ 7.2\cdot 10^{-2}$ \\[10pt]
$40$ &  $2.3\cdot 10^{-2}$  & $ 8.6\cdot 10^{-2}$  & $ 1.1\cdot 10^{-1}$ \\[10pt]
%
\hline
\end{tabular}
\end{center}
\end{table}
%
\subsubsection{Resolution trigonometry}
%
Moving beyond typical meshing concerns, we also verify the mesh upon its 
expected resolution capacity by defining radial sine functions whose frequency 
content depends upon background-model velocity and source period, and integrate
these global sine functions numerically and analytically on an element-by-element 
basis:
%
\eq
\Delta {\mathcal R} =\sum_e \left| \left[ \sum_{ij} M_{ij}^e
\sin{\left( a_e r_e\right)}\right]-\left[ {\mathcal I} (r_e^{\rm max}) 
- {\mathcal I} (r_e^{\rm min}) \right]\right| 
\en
%
where $a_e=2\pi \left(T_0 {v_{\rm p,s}^e}\right)^{-1}$ is a source-period and background-model 
dependent wavenumber as a monochromatic approximation to how seismic waves are 
sampled by the given settings. The analytical integral within an element is then given by
%
\eq
{\mathcal I}(r_e) = 2r_e/a_e^2 \sin{(a_e r_e)} - (r_e^2/a_e - 2/a_e^3) \cos{(a_e r_e)},
\en
%
provided that $a_e\neq f(r)$ such that we take extremal values for $v_{\rm p,s}$ for 
each element as shown in Table~\ref{table:resol_sine}.
%
\begin{table}[htb!]
\label{table:resol_sine}
\begin{center}
\caption{Radial sine function accuracy for various resolutions.}
\begin{tabular}{@{}ccccc}
&&&\\
\hline\hline
$T_0\, [s]$ & $\min{\left[\Delta v_{\rm s} \right]}$ &  $\min{\left[\Delta v_{\rm p} \right]}$ 
& $\max{\left[\Delta v_{\rm s} \right]}$ &  $\max{\left[\Delta v_{\rm p} \right]}$  \\
\hline\\
$5$ &  $4.2\cdot 10^{-6}$ & $3.3\cdot 10^{-6}$  & $1.4\cdot 10^{-5}$ & $1.7\cdot 10^{-6}$ \\[10pt]
$10$ & $6.1\cdot 10^{-6}$ & $4.3\cdot 10^{-7}$ & $7.2\cdot 10^{-6}$ & $1.0\cdot 10^{-5}$   \\[10pt]
$20$ & $5.3\cdot 10^{-6}$ & $4.4\cdot 10^{-6}$ & $7.0\cdot 10^{-6}$ & $1.6\cdot 10^{-8}$    \\[10pt]
$40$ & $6.2\cdot 10^{-6}$ & $2.8\cdot 10^{-6}$ & $7.7\cdot 10^{-7}$ & $7.7\cdot 10^{-7}$    \\[10pt]
%
\hline
\end{tabular}
%\end{minipage}
\end{center}
\end{table}
%
%
\subsubsection{Run-time breakdown}\label{appsection:runtime_breakdown}
%
Finally, we present the fractional CPU times occupied by several major 
parts of the time loop to give an idea about where the main computational 
burden lies. Fig.~\ref{img:frac_runtimes} depicts the fractional subprocess
CPU times as a function of number of processors.
Clearly, the message-passing CPU time increases with the number of 
processors, but to an insignificant amount compared to the overall 
computational burden. This is a further indicator that our method 
scales extremely well and is largely blind to the number of processors 
utilized. 
%
\begin{figure}[htb!]
\begin{center}
\includegraphics[scale=0.45]{timing_vs_timeloop.eps}
\caption{Fractional simulation times for the various subprocesses such as stiffness 
summation, message passing, assembly, and output of seismograms 
as a function of number of processors. $t_0$ denotes the simulation time 
using only one processor.}
\label{img:frac_runtimes}
\end{center}
\end{figure}
%
Table~\ref{table:runtimes} furthermore lists the dependencies 
of the subprocesses on different settings regarding the amount of output. 
Evidently, strain tensor calculation and output of wavefields for kernel
calculations are highly influential on the fractional CPU time in that 
they rise the output routine to take up to half of the computation time. 
The simulation including the computation of the energy occupies most 
of the added time within the routine that dumps stuff since it invokes 
the stiffness routine once more per time step. ``Seis 4'' and ``seis 8'' 
refer to the default setting of dumping a series of seismograms only, 
but for either $N_{\rm proc}=4$ or  $N_{\rm proc}=8$. Clearly, the 
increase in number of processors does not add any significant 
fractional CPU time compared to the simulations with intensified output.
The actual values in Table~\ref{table:runtimes} should not be taken 
too strictly as they are somewhat variable from one system to another, 
but merely as a an approximate hint as to what the relative CPU times of 
various settings may be.
%
\begin{table}[b!]
\begin{center}
\caption{Fractional run-times for several simulation types at $T_0=10\, \textrm{s}$
for $1900$ time steps.}
\label{table:runtimes}
\begin{tabular}{@{}cccccc}
&&&\\
\hline\hline
Activity & seis 4 & seis 8 & snaps & strain  & energy \\
\hline\\
assembly & 0.448 & 0.341& 0.320 & 0.319 & 0.178  \\[10pt]
MPI & 0.0051 &  0.018 & 0.0162 & 0.160 & 0.002 \\[10pt]
stiffness &0.301 & 0.358 & 0.306 &0.172 & 0.192  \\[10pt]
output & 0.127 & 0.153  & 0.250  &0.438 &  0.551 \\[10pt]
Total/$\Delta t$ & 0.636 & 0.273 &  0.323&  0.571 &  0.508\\ [10pt]
%
\hline
\end{tabular}
\end{center}
\end{table}
%

### 2015-08-10 utils/EXTERNAL_CODES_coupled_with_SPECFEM3D/AxiSEM_for_SPECFEM3D/AxiSEM_modif_for_coupling_with_specfem/MANUAL/virtual_box_tutorial.tex
Add NEW VERSION of AxiSEM

\subsection{Virtual box: 3-hour tutorial}

COMMENTS: (with permission, I (Kasra) highlight the important parts for next references)

* Tarje:
Hi Kasra, on a quick look, this seems excellent, but from my experience it’s better to have a very easy and quick first run of the tasks, e.g. on ONE page with all the important calls and major points we wish to achieve... and delegate images and details to an appendix. So, let’s keep all this in but reshuffle it to have a concise early section. Tarje

I would structure it as I said in the email... pasting it here again:
Tutorial:
1) 10min intro
2) 10min big run (maybe drop this, depending on progress next week)
3) 60min virtual box: data and AxiSEM
4) 20min big run post processing
5) installation of AxiSEM, obspy locally

(TAS÷
1) Load data from one of the 5 events (Kasra) with Obspy, plot cross sections
2) change AxiSEM input parameters to run this scenario, submit job
3) check mesh, background model, source-receiver geometry (google earth)
4) run post processing, filter, sum, movie snapshots
5) Movie snapshots
6) plot data vs. synthetics
7) plot AxiSEM synthetics vs SPECFEM
8) change input parameters (source CMT), load run with different background model

The virtual box should therefore contain:
(DONE)        - all 5 events (metadata and data), each in one directory 
(PENDING) - AxiSEM source code structure (mesher, solver, manual, testing)
(DONE*)       - all axisem runs replicating the data for 2-3 background models and source choices
(PENDING)  - plots of seismograms, movie 
*: for each event there are: prem\_aniso and iasp91 for 5 (PENDING),10,50,100 seconds

AXISEM vs DATA:

Start from within the SCRIPTS directory:
1. python sta\_event\_plot.py ../EVENTS/EVENT-1: plot the event and all the stations in the selected directory (EVENT-1 in this example).
2. run AXISEM for one of these events. The required information is located in /EVENT-1/INFO.
3.

% 1. Introduction
% In this part of the tutorial, we want to compare AXISEM waveforms with real data. For this reason, three events are selected (Figure 1). Detailed information for each event can be found in APPENDIX-1.

% Figure 1: beach ball diagrams (based on GCMT catalog) of event-1 to event-3

% Figure-2 shows how these events and their meta-data are organized in the Virtual-box.

% ?????????????????SHOUDL BE ADDED?????????????????????
% Figure 2: folder structure

% In SCRIPTS directory, two python scripts are provided that we will use here:
% 1. epi_plot.py: plotting tool for comparison purposes.
% 2. tt_plot.py: project the time shift derived by cross correlating the AXISEM waveforms and real data.

% 2. Run AXISEM
% As the first step, we want to simulate event-2 (Figure-3) with AXISEM. All the required AXISEM input files are provided in ????folder???.  APPENDIX-2 gives a quick overview on how to run AXISEM using PyAxi (Python interface for AXISEM).

% Figure 3: event and station configuration for event-2

% Once your simulation is done, all the AXISEM synthetic waveforms should be moved to ????folder???:
% cp <synthetic_waveforms> <destination????>

% 3. Compare the results
% To have an idea on how the synthetic and real data look like:
% python epi_plot.py ????address???

% Figure 4: AXISEM and real data waveforms arranged by epicentral distance

% For one specific seismic phase (Pdiff in this example), the following command plots the results of AXISEM against the real data: (to change the frequency?????)
% python epi_plot.py ../EVENTS/EVENT-2/AXISEM_PRE_SIMULATED/PREM\_ANISO_5sec/ Pdiff

% Figure 5: AXISEM and real data waveforms arranged by epicentral distance (Pdiff seismic phase)

% SPECFEM3D waveforms can be also added to the comparison by: (how to retrieve SPECFEM?!?!?!?!?)
% python epi_plot.py ../EVENTS/EVENT-2/AXISEM_PRE_SIMULATED/PREM_ANISO\_5sec/ Pdiff specfem3D
% Figure 6: AXISEM, SPECFEM3D and real data waveforms arranged by epicentral distance (Pdiff seismic phase)

% 4. Travel time measurements
% The time shift between AXISEM and real data (Figure 5) can be roughly measured by cross-correlating the time series. epi_plot.py can perform this and shift the synthetics in order to match them up:
% python epi_plot.py ../EVENTS/EVENT-2/AXISEM_PRE_SIMULATED/PREM_ANISO_5sec/ Pdiff shift_synthetics

% Figure 7: AXISEM and real data waveforms arranged by epicentral distance (Pdiff seismic phase), synthetic waveforms are shifted based on cross-correlation analysis.

% The calculated time shift can be mapped on the relevant station for the selected phase:
% python tt_plot.py ../EVENTS/EVENT-2/

% APPENDIX-1: Events
% ????????

% APPENDIX-2: A Quick Guide to PyAxi
% PyAxi (Python interface for AXISEM) is a Python script to run AXISEM automatically.

% python PyAxi <inpython.cfg> <STATIONS>

% and the rest should be done automatically. Therefore, to run the AXISEM for the provided examples (I mean! AXISEM_pre_simulated), it is enough to run:

% APPENDIX-3: Retrieving SPECFEM3D seismograms

% Event-1:
% ./obspyDMT.py --datapath EVENT-1 --min_date 2009-07-15 --max_date 2009-07-16 --min_depth 20 --list_stas /home/hosseini/Desktop/ALASKA_VBOX_LATEST/EVENTS/EVENT-1/INFO/STATIONS --specfem3D --offset 3600 --min_mag 7.0 --req_parallel --arc N

% Event-2:
% ./obspyDMT.py --datapath EVENT-2 --min_date 2009-09-30 --max_date 2009-10-01 --min_depth 70 --list_stas /home/hosseini/Desktop/ALASKA_VBOX_LATEST/EVENTS/EVENT-2/INFO/STATIONS --specfem3D --offset 3600 --min_mag 7.0 --req_parallel --arc N

% Event-3:
% ./obspyDMT.py --datapath EVENT-3 --min_date 2006-10-15 --max_date 2006-10-16 --min_depth 20 --list_stas /home/hosseini/Desktop/ALASKA_VBOX_LATEST/EVENTS/EVENT-3/INFO/STATIONS --specfem3D --offset 3600 --min_mag 6.0 --req_parallel --arc N

### 2015-08-10 utils/EXTERNAL_CODES_coupled_with_SPECFEM3D/AxiSEM_for_SPECFEM3D/AxiSEM_modif_for_coupling_with_specfem/MESHER/discont_meshing.f90

@@ -1,6 +1,6 @@
 !
 !    Copyright 2013, Tarje Nissen-Meyer, Alexandre Fournier, Martin van Driel
-!                    Simon Stahler, Kasra Hosseini, Stefanie Hempel
+!                    Simon Stähler, Kasra Hosseini, Stefanie Hempel
 !
 !    This file is part of AxiSEM.
 !    It is distributed from the webpage <http://www.axisem.info>
@@ -19,13 +19,14 @@
 !    along with AxiSEM.  If not, see <http://www.gnu.org/licenses/>.
 !
 
+!=========================================================================================
 module discont_meshing
 
   use global_parameters
   use data_bkgrdmodel
   use data_spec
   use model_discontinuities, only : define_discont
-  use background_models
+  use background_models 
   use data_diag
 
   implicit none
@@ -38,12 +39,12 @@ module discont_meshing
 !-----------------------------------------------------------------------------------------
 subroutine create_subregions
 
-  real(kind=dp)   , allocatable                 :: rdisc_top(:), rdisc_bot(:)
-  real(kind=dp)   , dimension(:), allocatable   :: ds_glob, radius_arr
-  real(kind=dp)   , dimension(:), allocatable   :: vp_arr, vs_arr
+  real(kind=dp), allocatable                 :: rdisc_top(:), rdisc_bot(:)
+  real(kind=dp), dimension(:), allocatable   :: ds_glob, radius_arr
+  real(kind=dp), dimension(:), allocatable   :: vp_arr, vs_arr
   real(kind=dp)     :: ds, minh, maxh, aveh, dz, current_radius
   integer           :: idom, ic, icount_glob, iz_glob
-  integer           :: ns_ref,  ndisc_stop,  ns_ref_icb1, ns_ref_icb, ns_ref_surf
+  integer           :: ns_ref,  ns_ref_icb1, ns_ref_icb, ns_ref_surf
   logical           :: memorydz, current
   integer, dimension(1)             :: iloc1, iloc2
   real(kind=realkind), dimension(1) :: rad1, rad2
@@ -64,259 +65,230 @@ subroutine create_subregions
   ! 7) coarsening levels need to be between discontinuities
   ! ==> total number of subdomains: ndisc+ncoars
[…]

### 2015-08-10 utils/EXTERNAL_CODES_coupled_with_SPECFEM3D/AxiSEM_for_SPECFEM3D/AxiSEM_modif_for_coupling_with_specfem/SOLVER/UTILS/post_processing.F90

@@ -1,6 +1,6 @@
 !
 !    Copyright 2013, Tarje Nissen-Meyer, Alexandre Fournier, Martin van Driel
-!                    Simon Stahler, Kasra Hosseini, Stefanie Hempel
+!                    Simon Stähler, Kasra Hosseini, Stefanie Hempel
 !
 !    This file is part of AxiSEM.
 !    It is distributed from the webpage <http://www.axisem.info>
@@ -19,22 +19,22 @@
 !    along with AxiSEM.  If not, see <http://www.gnu.org/licenses/>.
 !
 
-!-----------------------------------------------------------------------------
+!=========================================================================================
 module data_all
 
   implicit none
   public
 
-  integer, parameter, private         :: sp = selected_real_kind(6, 37)
-  integer, parameter, private         :: dp = selected_real_kind(15, 307)
-  integer, parameter         :: qp = selected_real_kind(33, 4931)
+  integer, parameter, private       :: sp = selected_real_kind(6, 37)
+  integer, parameter, private       :: dp = selected_real_kind(15, 307)
+  integer, parameter                :: qp = selected_real_kind(33, 4931)
   logical                           :: rot_src_rec_true
   character(len=100)                :: stf_file, rot_rec_file, filename, fname
 
-  character(len=7),allocatable, dimension(:)        :: stf_type
-  character(len=040),allocatable, dimension(:)      :: recname
-  character(len=100),allocatable, dimension(:,:)    :: outname, outname2, rec_full_name
-  character(len=10),allocatable, dimension(:,:)     :: src_type
+  character(len=7), allocatable, dimension(:)       :: stf_type
+  character(len=040), allocatable, dimension(:)     :: recname
+  character(len=100), allocatable, dimension(:,:)   :: outname, outname2, rec_full_name
+  character(len=10), allocatable, dimension(:,:)    :: src_type
   integer                                           :: nrec, nt_seis
 
   character(len=1),dimension(3)       :: reccomp
@@ -54,8 +54,9 @@ module data_all
   logical                             :: load_snaps
   logical                             :: negative_time
   logical                             :: detailed_output
[…]

### 2015-08-10 utils/EXTERNAL_CODES_coupled_with_SPECFEM3D/AxiSEM_for_SPECFEM3D/AxiSEM_modif_for_coupling_with_specfem/SOLVER/meshes_io.F90

@@ -1,6 +1,6 @@
 !
 !    Copyright 2013, Tarje Nissen-Meyer, Alexandre Fournier, Martin van Driel
-!                    Simon Stahler, Kasra Hosseini, Stefanie Hempel
+!                    Simon Stähler, Kasra Hosseini, Stefanie Hempel
 !
 !    This file is part of AxiSEM.
 !    It is distributed from the webpage <http://www.axisem.info>
@@ -19,6 +19,7 @@
 !    along with AxiSEM.  If not, see <http://www.gnu.org/licenses/>.
 !
 
+!=========================================================================================
 !> This module contains routines that compute and dump the respective meshes
 !! underlying the actual wavefields to be dumped in the time loop
 !! which is done in wavefields_io. This module needs pre-loop variables such
@@ -29,42 +30,44 @@ module meshes_io
   use data_mesh
   use data_proc
   use data_io
-
-  use utlity, ONLY : scoord, zcoord, rcoord, thetacoord
-
+  
+  use utlity, only : scoord, zcoord, rcoord, thetacoord
+  
   implicit none
-
+  
   private
-  public :: fldout_cyl2
   public :: finish_xdmf_xml
   public :: dump_wavefields_mesh_1d
   public :: dump_glob_grid_midpoint
   public :: dump_xdmf_grid
-  public :: dump_solid_grid
-  public :: dump_fluid_grid
   public :: prepare_mesh_memoryvar_vtk
+  public :: build_kwf_grid
+  public :: dump_kwf_midpoint_xdmf
+  public :: dump_kwf_fem_xdmf
+  public :: dump_kwf_sem_xdmf
+  public :: dump_kwf_gll_xdmf
 
 contains
 
-!-----------------------------------------------------------------------------
-!> Dumps the mesh (s,z) [m] in ASCII format as needed to visualize global
+!-----------------------------------------------------------------------------------------
+!> Dumps the mesh (s,z) [m] in ASCII format as needed to visualize global 
 !! snapshots, and additionally the constant factors preceding the displacement
 !! in the fluid, namely rho^{-1} and (rho s)^{-1}.
-!! When reading the fluid wavefield, one therefore needs to multiply all
[…]

### 2015-08-10 utils/EXTERNAL_CODES_coupled_with_SPECFEM3D/AxiSEM_for_SPECFEM3D/AxiSEM_modif_for_coupling_with_specfem/MANUAL/AXISEM_TUTORIAL/AXISEMTutorial.tex

%&pdfLaTeX
% !TEX encoding = UTF-8 Unicode
\documentclass{article}
\usepackage{ifxetex}
\ifxetex
\usepackage{fontspec}
\setmainfont[Mapping=tex-text]{STIXGeneral}
\else
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\fi
\usepackage{textcomp}

\usepackage{tabularx}
\usepackage{graphicx}
\usepackage{ulem}
\usepackage{amssymb}
\usepackage{fancyhdr}
\usepackage{float}

\newcommand{\ttilde}[0]{\raise.17ex\hbox{$\scriptstyle\sim$}}
\renewcommand{\vec}[1]{\ensuremath{\mathbf{#1}}}
\newcommand{\evec}[1]{\ensuremath{\mathbf{\hat #1}}}

\renewcommand{\headrulewidth}{0pt}
\renewcommand{\footrulewidth}{0pt}
\usepackage{color}
\usepackage[a4paper,vmargin={20mm,20mm},hmargin={20mm,10mm}]{geometry}
% \usepackage{listings}
% \lstset{inputencoding=ansinew}

\definecolor{color18}{rgb}{0.07,0.33,0.80}

\setlength\parindent{0pt}

\begin{document}

\begin{center}
    \begin{tabularx}{\textwidth}{lX}
        \begin{minipage}{0.3\textwidth}
            \begin{center}
                \includegraphics[width=\textwidth]{Axisem2png.png}
            \end{center}
        \end{minipage}
        &
        \begin{minipage}{0.6\textwidth}
            \begin{center}
                \title{}
                \LARGE{ \textbf{\sc AxiSEM Tutorial}}
                \vspace*{0.6cm}\\
                {\large 
                Kasra Hosseini\textsuperscript{1)}, 
                Martin van Driel\textsuperscript{2)},
                Simon St\"{a}hler\textsuperscript{1)}, \\
                \vspace*{-0.3cm}
                Lion Krischer\textsuperscript{1)},
                Tarje Nissen-Meyer\textsuperscript{3)}}
                \vspace*{0.3cm}\\
                {\small \textsuperscript{1)} LMU Munich, \textsuperscript{2)} ETH Zurich, 
                \textsuperscript{3)} Oxford University \\
                {\large Fairbanks, Alaska, July 15th 2013}}
            \end{center}
        \end{minipage}
    \end{tabularx}
\end{center}

%
\section{Tutorial Overview}
%
\subsection*{ObsPy \small{(http://www.obspy.org)}}
\vspace{1ex}
\begin{center}
    \begin{tabularx}{\textwidth}{lX}
        \begin{minipage}{0.3\textwidth}
            \begin{center}
                \includegraphics[width=0.9\columnwidth]{obspy.png}
            \end{center}
        \end{minipage}
        &
        \begin{minipage}{0.65\textwidth}
            ObsPy is a community-driven, open-source project dedicated to
            provide a Python framework for processing seismological data. It
            provides parsers for common file formats, clients to access data
            centers and seismological signal processing routines which allow
            the manipulation of seismological time series. The goal of the
            ObsPy project is to facilitate rapid application and workflow
            development for seismology.
        \end{minipage}
    \end{tabularx}
\end{center}

Some of the tools employed in this tutorial use ObsPy but the tutorial
does not have a formal introduction to ObsPy due to temporal
constraints. The VirtualBox image contains an extensive amount of
training material for Python, ObsPy and some third party libraries
intended to get you started. Furthermore one of the core developers of
ObsPy is present and available for questions. You can find all material related
to ObsPy at \ttilde\verb|/Desktop/ObsPy|.
%
\subsection*{AxiSEM - Hands On}
The goal of this initial task is to familiarize users with the basic
principles behind AxiSEM, its input/output structures and
peculiarities such as post-processing. An end-to-end approach (meshing to
wavefield movie and seismograms) will be conducted for long-period
settings within the virtual box.
%
\subsection*{Data and Synthetics: ObsPy, AxiSEM and SPECFEM}
The main goal of this part of the tutorial is to use AxiSEM for realistic
scenarios, compare the results with real seismograms and explore the effects of
source parameters and background models on the waveforms. In particular, we will
learn how to:

\begin{itemize}
    \item Load data with ObsPy and plot seismograms.
    \item compare AxiSEM and SPECFEM synthetics with data including
        different frequency ranges and background model.
    \item analyze the influence of different source mechanisms on waveforms.
\end{itemize}

\subsection*{Virtual Box content}
On the desktop you find four folders:
\begin{itemize}
    \item \verb|ObsPy|: ObsPy training material.
    \item \verb|axisem|: AxiSEM source code and input files ready for the tutorial.
    \item \verb|EVENTS|: Data and precomputed synthetics for three events and various background models.
    \item \verb|VIDEOS|: Illustrative 3D wave propagation video and high-resolution snapshots thereof.
\end{itemize}

\newpage

\section{AxiSEM - Hands On}

\subsection*{The AxiSEM Concept}

\begin{center}
    \begin{minipage}[t]{0.4\paperwidth}
        Source Decomposition:\\
        \begin{minipage}[c]{0.1\paperwidth}
            \includegraphics[width=\textwidth]{radpat_mono.pdf}%
        \end{minipage}%
        \begin{minipage}[c]{0.4\paperwidth}
            $\scriptsize{\vec u = \vec u (s,z)}$ \\
        \end{minipage}

        \begin{minipage}[c]{0.1\paperwidth}
            \includegraphics[width=\textwidth]{radpat_di.pdf}
        \end{minipage}%
        \begin{minipage}[c]{0.4\paperwidth}
            $\scriptsize{\vec u = \vec u (s,z)} \cdot f(\sin\phi, \cos\phi)$ \\
        \end{minipage}

        \begin{minipage}[c]{0.1\paperwidth}
            \includegraphics[width=\textwidth]{radpat_quad.pdf}
        \end{minipage}%
        \begin{minipage}[c]{0.4\paperwidth}
            $\scriptsize{\vec u = \vec u (s,z)} \cdot f(\sin(2\phi), \cos(2\phi))$ \\
        \end{minipage}%
    \end{minipage}%
    \begin{minipage}[t]{0.25\paperwidth}
        2D numerical problems:

        \includegraphics[width=.7\textwidth]{mesh.png}
    \end{minipage}%
\end{center}

The basic idea behind AxiSEM is to take advantage of axial symmetry with respect
to an axis going through the center of the earth and the source. In such axisymmetric 
models, the response to a moment tensor or single force point source can be expanded in a
series of multipoles (mono-, di- and quadrupole). The dependence of the wavefield on
azimuth $\phi$ can be solved analytically and the remaining 2D problems (four of them for
a full moment tensor source) are solved numerically using a spectral element approach.

\subsection*{MESHER - generate a Mesh}

\begin{enumerate}
    \item Open a terminal, go to the \ttilde\verb|/Desktop/axisem/MESHER| folder and open
    the \verb|inparam_mesh| file with your favourite editor:
    %
    \begin{verbatim}
    $ cd Desktop/axisem/MESHER
    $ vi inparam_mesh
    \end{verbatim}
    %
    The parameters should be readily set, but you might want to double check and verify:
    %
    \begin{verbatim}
    BACKGROUND_MODEL    'prem_ani_light'
    DOMINANT_PERIOD     100.0
    NCPU                1
    WRITE_VTK           true
    COARSENING_LAYERS   2
    \end{verbatim}
    %
    The file should be self-explanatory. NB: Models without crust ('light') allow for a
    larger time step and hence run a lot faster on the box. The virtual box only has a
    single processor, so parallelization does not speed up the simulation.

    \item Run the mesher, and watch the progress:
    %
    \begin{verbatim}
    $ ./submit.csh
    $ tailf OUTPUT
    \end{verbatim}
    %
    The meshing should be really fast for the chosen parameters. Wait for 
    \verb|....DONE WITH MESHER !| to appear.

    \item Take a look at the mesh with paraview
    \begin{verbatim}
    $ paraview
    \end{verbatim}
    %
    Open one of the vtk files in the subfolder \verb|Diags|, e.g.\
    \verb|mesh_vp.vtk| and click apply in the properties panel on the left (you
    might get an OpenGL Error on the virtual box, which you can ignore). To see
    the mesh, change the representation from 'surface' to 'surface with edges'
    (On some host systems, the dropdown menu seems to be messed up, in that
    case go to the 'Display' context in the 'Properties' panel on the left. If
    the plot appears all yellow, click on play).  You can open other vtk files
    to look at other properties of the model and the mesh. You might need to
    rescale the color range by clicking on the left-right arrow symbol in the
    top left.

    \begin{center}
    \includegraphics[height=30mm]{paraview.png}
    \hspace{5mm}
    \includegraphics[height=30mm]{paraview3.png}
    \end{center}

    \item Move the mesh to the solver directory and give it a meaningful name:
    \begin{verbatim}
    $ ./movemesh.csh prem_ani_light_100s
    \end{verbatim}

\end{enumerate}

\subsection*{SOLVER - solve the elastic wave equation}

\begin{enumerate}
    \item Go to the \ttilde\verb|/Desktop/axisem/SOLVER| folder and open the
    \verb|inparam_basic| file with your favourite editor:
    %
    \begin{verbatim}
    $ cd ../SOLVER
    $ vi inparam_basic
    \end{verbatim}
    %
    The parameters should be readily set, but you might want to double check and verify:
    %
    \begin{verbatim}
    SIMULATION_TYPE     single
    SEISMOGRAM_LENGTH   1800.
    RECFILE_TYPE        stations 
    MESHNAME            prem_ani_light_100s
    ATTENUATION         true 
    SAVE_SNAPSHOTS      true  
    \end{verbatim}
    %
    \item First, we are taking a look at a basic sourcetype: a vertical dipole, which has
    a monopole radiation pattern. This is set by \verb|SIMULATION_TYPE single| and defined
    in the \verb|sourceparams.dat| file. Run the solver, giving the run a meaningful name:
    %
    \begin{verbatim}
    $ ./submit.csh prem_ani_light_100s_mzz
    \end{verbatim}
    %
    This command compiles the code if needed and starts the simulation. You can observe
    the progress in the outputfile:
    %
    \begin{verbatim}
    $ cd prem_ani_light_100s_mzz
    $ tailf OUTPUT_prem_ani_light_100s_mzz
    \end{verbatim}
    %
    Once the run is finished, take a look at the wavefield with \verb|paraview|: open
    the \verb|prem_ani_light_100s_mzz/| \verb|Data/xdmf_xml_0000.xdmf|
    file and click apply. Go to the last snapshot and rescale the color range,
    then click on play to see the wave propagate.  You can also choose
    different components of the wavefield or the absolute value.  For paraview
    experienced users: choose absolute value and a logarithmic colorscale to
    see all wave types at once (e.g.\ 'black body radiation' looks nice).  You
    can find a 5s period movie of these snapshots in the \verb|VIDEOS| folder on
    the desktop (use \verb|gpicview| to open the \verb|.png| files and
    \verb|gnome-mplayer| to open the movie \verb|SeismicWavePropagation.mov|).

    \begin{center}
    \includegraphics[width=110mm]{paraview2.png} \hspace{5mm}
    \includegraphics[width=40mm]{0600.png}
    \end{center}

    \item Now simulate seismograms for a full moment tensor source: the source is defined
    in the \verb|CMTSOLUTION| file and the one referred to as 'event-1' in the later tasks.
    Stations are defined in the \verb|STATIONS| file. Go back to the \verb|SOLVER|
    directory and change the \verb|inparam_basic| file such that:
    %
    \begin{verbatim}
    SIMULATION_TYPE     moment
    SAVE_SNAPSHOTS      false
    \end{verbatim}
    %
    Run the solver, giving the run a meaningful name:
    %
    \begin{verbatim}
    $ ./submit.csh prem_ani_light_100s_event1
    \end{verbatim}
    %
    This command compiles the code if needed and starts four simulations at once, each
    simulating a basic source type (two monopoles, a dipole and a quadrupole, for details
    see Nissen-Meyer et al 2007). You can observe the progress in the outputfiles in
    each job's subdirectory
    %
    \begin{verbatim}
    $ cd prem_ani_light_100s_event1
    $ tailf MZZ/OUTPUT_MZZ
    \end{verbatim}
    %
    Once all the jobs are done (check with \verb|htop|), you can proceed with
    postprocessing.

\end{enumerate}

\subsection*{POSTPROCESSING - rotate and sum seismograms and wavefields}

Postprocessing is a key feature of AxiSEM: the source mechanism and source time function
can be modified without redoing the more expensive simulation.

\begin{enumerate}
    \item For the previous simulation, the contribution of the elemental sources needs
    to be summed up to get seismograms for a full moment tensor source. In the
    main rundirectory (\verb|prem_ani_light_100s_event1|) open the file
    \verb|param_postprocessing|. It should contain these settings (auto generated by the
    solver):
    %
    \begin{verbatim}
    REC_COMP_SYS    enz
    CONV_PERIOD     100.0000
    CONV_STF        gauss_0
    \end{verbatim}
    %
    The source mechanism (depth and location cannot be changed in postprocessing) is read
    from the \verb|CMTSOLUTION| file in the same directory. Start the postprocessing:
    %
    \begin{verbatim}
    $ ./postprocessing.csh
    \end{verbatim}
    %
    The resulting seismograms and plots can be found in the directory 
    \verb|Data_Postprocessing|. Seismograms can be viewed with
    %
    \begin{verbatim}
    $ cd Data_Postprocessing/GRAPHICS
    $ gpicview <filename.gif>
    \end{verbatim}
    %
    For a nice overview, you can use \verb|google-earth| (might not run on all computers
    and depends on internet connection). Open the \verb|googleearth_src_rec_seis.kml| file
    in the \verb|Data_Postprocessing/| directory (double check the exact path,
    google-earth might have something older from history which is quite confusing).
    You should now see the earthquake and the receivers in the places menu on the left.

    \begin{center}
    \includegraphics[height=70mm]{google-earth.png}
    \hspace{5mm}
    \includegraphics[height=70mm]{google-earth2.png}
    \end{center}

    Click on the stations or source to see more...

\end{enumerate}

%\emph{Low-frequency simulation at period 50/100 s (20min)}
%\begin{enumerate}
%    \item change AxiSEM input parameters to run this scenario, submit job
%    \item check mesh, background model, source-receiver geometry (google earth)
%    \item post processing on 50s run: filter, sum, rotate, movie snapshots
%\end{enumerate}

\newpage

\section{Data and Synthetics: ObsPy, AxiSEM and SPECFEM}
The main goal of this part of the tutorial is to use AxiSEM for realistic scenarios, 
compare the results with real seismograms and 3D synthetics and 
explore the effects of source parameters and background models on the waveforms. 
% In case that you want more information, 
% refer to Appendix-5 in which a complete example with all the commands and results is presented.
For a brief walk-through, follow this: \\

Start from within the \ttilde\verb|/Desktop/EVENTS/| directory:

\begin{enumerate}

    \item Plot one of the events (listed in EVENTS directory, we choose EVENT-1
    in this example):

    \begin{verbatim}
    $ plot_station_event_distribution.py EVENT-1
    \end{verbatim}
    * For more information on the folder structure of your VirtualBox image,
    refer to Appendix-2.

    \item To get an overview of both real data and AxiSEM synthetics 
    (e.g.\ \textit{PREM\_ANISO} for 5 seconds dominant period) [Figure-1]:
    \begin{verbatim}
    $ plot_seismograms.py EVENT-1/AXISEM/PREM_ANISO_5sec
    \end{verbatim}

    \item Window the waveforms arounf Pdiff and PKiKP seismic phases and plot
    the seismograms (compared with real data):
    \begin{verbatim}
    $ plot_seismograms.py EVENT-1/AXISEM/PREM_ANISO_5sec Pdiff
    $ plot_seismograms.py EVENT-1/AXISEM/PREM_ANISO_5sec PKiKP
    \end{verbatim}

    \item Change the filter in the
    \ttilde\verb|/Desktop/EVENTS/SCRIPTS/plot_seismograms.py| script and
    compare the results with SPECFEM3D [Figure-2]:

    Note: for changing the filter, open \textit{plot\_seismograms.py} and
    change values at top of the file. \\
    Note: resolution in SPECFEM3D seismograms is 17 - 500 sec.
    \begin{verbatim}
    $ plot_seismograms.py EVENT-1/AXISEM/PREM_ANISO_5sec Pdiff specfem3d
    $ plot_seismograms.py EVENT-1/AXISEM/PREM_ANISO_5sec PKiKP specfem3d
    \end{verbatim}

    \item Compare the results for two different background models
    (\textit{PREM\_ANISO\_5sec} with \textit{IASP91\_5sec)} e.g.\ Pdiff phase:

    \begin{verbatim}
    $ plot_seismograms.py EVENT-1/AXISEM/PREM_ANISO_5sec Pdiff EVENT-1/AXISEM/IASP91_5sec
    \end{verbatim}

    \item Change the filter, as explained in step 4, and repeat step 5.

    \item Compare the seismograms calculated for two different source
    mechanisms (\textit{PREM\_ANISO\_5sec} and \\
    \textit{PREM\_ANSIO\_5sec\_GCMT}) for Pdiff phase [Figure-3]:

    Note: for more information about the source parameters, refer to Appendix-1.\\
    Note: depth and location of the source can not be changed.
    \begin{verbatim}
    $ plot_seismograms.py EVENT-1/AXISEM/PREM_ANISO_5sec Pdiff 
    EVENT-1/AXISEM/PREM_ANISO_5sec_GCMT
    \end{verbatim}

    \item Try different filter settings and compare the results.
    (Note: dominant period in synthetic seismograms is 5 sec.)

    \item Find the time shift between the synthetics and real data (maximum of
    cross-correlation function), shift the synthetics accordingly and plot the results:

    \begin{verbatim}
    $ plot_seismograms.py EVENT-1/AXISEM/PREM_ANISO_5sec Pdiff shift_synthetics
    \end{verbatim}

    \item Map the calculated time shifts in step 9 on the station locations:

    \begin{verbatim}
    $ plot_travel_time_map.py EVENT-1
    \end{verbatim}

\end{enumerate}

\newpage
\begin{figure}[H]
    \centering
    \includegraphics[width=1.\linewidth]{AXISEMTutorial-fig007.png}
    \begin{center}
    {\small{}Figure 1: Real and AXISEM waveforms for EVENT-1.}
    \end{center}
\end{figure}

\begin{figure}[H]
    \centering
    \begin{minipage}{.5\textwidth}
        \centering
        \includegraphics[width=0.9\linewidth]{AXISEMTutorial-fig010.pdf}
        {\small{}Figure 2: Comparison between real, AXISEM and SPECFEM3D waveforms for 
        Pdiff phase. \\ Filter: 20 - 100 sec.}
    \end{minipage}%
    \begin{minipage}{.5\textwidth}
        \centering
        \includegraphics[width=0.9\linewidth]{AXISEMTutorial-fig014.png}
        {\small{}Figure 3: Comparison between real and AXISEM waveforms for two different 
        source parameters (Pdiff). \\ Filter: 5 - 20 sec.}
    \end{minipage}
\end{figure}

\newpage
\appendix
\section{APPENDIX-1: Events}

Three events are selected for this tutorial (Figure-A1) with the following source 
characteristics:

\begin{center}
%%\begin{figure}[htbp]
\includegraphics[width=372pt, height=178pt, keepaspectratio=true]{AXISEMTutorial-fig001.png}
%%\caption{This should be the caption for \texttt{AXISEMTutorial-fig001.png}.}
%%\end{figure}

{\small{}Figure A1: beach ball diagrams of event-1 to event-3 (based on GCMT catalog)}

\vspace{1cm}
%%\begin{figure}[htbp]
\includegraphics[width=312pt, height=110pt, keepaspectratio=true]{AXISEMTutorial-fig002.png}
%%\caption{This should be the caption for \texttt{AXISEMTutorial-fig002.png}.}
%%\end{figure}

\vspace{1cm}
%%\begin{figure}[htbp]
\includegraphics[width=312pt, height=110pt, keepaspectratio=true]{AXISEMTutorial-fig003.png}
%%\caption{This should be the caption for \texttt{AXISEMTutorial-fig003.png}.}
%%\end{figure}

\vspace{1cm}
%%\begin{figure}[htbp]
\includegraphics[width=312pt, height=110pt, keepaspectratio=true]{AXISEMTutorial-fig004.png}
%%\caption{This should be the caption for \texttt{AXISEMTutorial-fig004.png}.}
%%\end{figure}

\end{center}

%\baselineskip=13pt
%\leftskip=0pt

\newpage

\section{APPENDIX-2: Folder structure}

Figure A2 shows how the events (and their meta-data), waveforms and scripts are 
organized in the Virtual-box:

\begin{center}
%%\begin{figure}[htbp]
\includegraphics[width=334pt, height=362pt, keepaspectratio=true]{Folder_structure.jpeg}

Figure A2: Folder structure
%%\end{figure}
\end{center}

In \textit{SCRIPTS} directory, there are three python scripts that we use here:

\begin{enumerate}
    \item \textbf{plot\_station\_event\_distribution.py}: maps event and stations of an
    event directory.
    \item \textbf{plot\_seismograms.py}: plotting/filtering seismograms for comparison purposes.
    \item \textbf{plot\_travel\_time\_map.py}: project the time shift measured by cross
    correlating the AXISEM waveforms and real data.
\end{enumerate}

In \textit{EVENTS} directory, there are three events, each with the following
sub-directories:

\begin{enumerate}
    \item \textbf{AXISEM}: contains seismograms simulated by AXISEM with
    the required PyAxi (Appendix-3) input files (INPUT\_FILES) to re-produce them.
    \item \textbf{REAL\_DATA}: seismograms retrieved from \textit{IRIS. }(refer to
    APPENDIX-4)
    \item \textbf{SPECFEM3D}: waveforms simulated by \textit{SPECFEM3D} for comparison purposes. (downloaded from\textit{ IRIS}, refer to APPENDIX-4)
    \item \textbf{INFO}: information about the event and stations: \textit{event\_1.xml} and
    \textit{STATIONS}.
\end{enumerate}

\newpage
\section{APPENDIX-3: A Quick Guide to PyAxi}

PyAxi is a Python script developed as an interface for AXISEM. All the options 
available in AXISEM are included in only one input file (\textit{inpython.cfg}). By running 
the script, all the necessary steps (MESHER, SOLVER and Post-Processing) will be 
done automatically. \\

All you should do to run PyAxi for an input file (\textit{inpython.cfg}) and station list 
(STATIONS) is:
\begin{verbatim}
    $ python PyAxi <inpython.cfg> <STATIONS>
\end{verbatim}

and the rest should be done automatically. \\
\textit{inpython.cfg} is a conguration file that 
contains all the AXISEM options. To change the input file, open \textit{inpython.cfg} with 
an editor. However, you could find some already prepared \textit{inpython.cfg} files for 
the events in the Virtual-box. (refer to APPENDIX-2 for more information; \textit{INPUT\_FILES} in Figure A2) 
Therefore, to run AXISEM for the provided events (EVENT-1 
and IASP91-5sec as an example), it is enough to replace:

\begin{verbatim}
    <inpython.cfg>: 
    ~/Desktop/EVENTS/EVENT-1/AXISEM/IASP91_5sec/INPUT_FILES/inpython.cfg

    <STATIONS>:
    ~/Desktop/EVENTS/EVENT-1/AXISEM/IASP91_5sec/INPUT_FILES/STATIONS
\end{verbatim}

WARNING: this will take ages on the Virtual-box because of the dominant period (5sec) and available computing resources.

\newpage
\section{APPENDIX-4: Retrieving real data and SPECFEM3D seismograms automatically}

{\color{color18} \emph{obspyDMT}} (ObsPy Data Management Tool) is a command line 
tool for retrieving, processing and management of massive seismological data in 
a fully automatic way which could be run in serial or in parallel. \\

This tool is developed to mainly address the following tasks automatically:

1. Retrieval of waveforms (MSEED or SAC), response files and metadata from {\color{color18} \emph{IRIS}} 
and {\color{color18} \emph{ORFEUS}} (via {\color{color18} \emph{ArcLink}}) archives. 
This could be done in \textit{serial} or in \textit{parallel} for single or large 
requests.

2. Supports event-based and continuous requests.

3. Extracting the information of all the events via user-defined options (time 
span, magnitude, depth and event location) from {\color{color18} \emph{IRIS}} and 
{\color{color18} \emph{EMSC}} (European Mediterranean Seismological Centre).

4. Updating the existing archives (waveforms, response files and metadata).

5. Processing the data in \textit{serial} or in \textit{parallel} (e.g. \textit{Tapering, 
removing the trend of the time series, filtering and Instrument correction}).

6. Management of large seismic datasets.

7. Plotting tools (events and/or station locations, Ray coverage (event-station 
pair), epicentral-distance plots for all archived waveforms and seismicity maps).

\vspace{13pt}
Here, we use obspyDMT to retrieve both real data and SPECFEM3D seismograms. For 
more information about this tool please refer to the following webpage:

{\color{color18} \emph{https://github.com/kasra-hosseini/obspyDMT}}  \\

obspyDMT is installed on your virtual machine. By running the following commands, 
the real data used in this tutorial can be retrieved automatically:

\textbf{Event-1:}

\begin{verbatim}
    ./obspyDMT.py --datapath EVENT-1_real --min_date 2009-07-15 --max_date 2009-07-16 
    --min_mag 7.0 --min_depth 20 --list_stas ~/Desktop/EVENT-1/INFO/STATIONS 
    --offset 3600 --req_parallel --arc N
\end{verbatim}

\textbf{Event-2:}

\begin{verbatim}
    ./obspyDMT.py --datapath EVENT-2_real --min_date 2009-09-30 --max_date 2009-10-01 
    --min_mag 7.0 --min_depth 70 --list_stas ~/Desktop/EVENT-1/INFO/STATIONS 
    --offset 3600 --req_parallel --arc N
\end{verbatim}

\textbf{Event-3:}

\begin{verbatim}
    ./obspyDMT.py --datapath EVENT-3_real --min_date 2006-10-15 --max_date 2006-10-16 
    --min_mag 6.0 --min_depth 20 --list_stas ~/Desktop/EVENT-1/INFO/STATIONS 
    --offset 3600 --req_parallel --arc N
\end{verbatim}

Moreover, the SPECFEM3D seismograms can be also retrieved in the same manner:

\textbf{Event-1:}

\begin{verbatim}
    ./obspyDMT.py --datapath EVENT-1 --min_date 2009-07-15 --max_date 2009-07-16 
    --min_mag 7.0 --min_depth 20 --list_stas ~/Desktop/EVENT-1/INFO/STATIONS 
    --specfem3D --offset 3600 --req_parallel --arc N
\end{verbatim}

\textbf{Event-2:}

\begin{verbatim}
    ./obspyDMT.py --datapath EVENT-2 --min_date 2009-09-30 --max_date 2009-10-01 
    --min_mag 7.0 --min_depth 70 --list_stas ~/Desktop/EVENT-1/INFO/STATIONS 
    --specfem3D --offset 3600 --req_parallel --arc N
\end{verbatim}

\textbf{Event-3:}

\begin{verbatim}
    ./obspyDMT.py --datapath EVENT-3 --min_date 2006-10-15 --max_date 2006-10-16 
    --min_mag 6.0 --min_depth 20 --list_stas ~/Desktop/EVENT-1/INFO/STATIONS 
    --specfem3D --offset 3600 --req_parallel --arc N
\end{verbatim}

\newpage

% \section{APPENDIX-5: Data versus modeling (EVENT-1)}
% 
% In this appendix, we follow the steps in the main tutorial and show the commands 
% and outcomes for each step. For this reason, we focus on one of the events, EVENT-1. \\
% 
% Start from within the \ttilde\verb|/Desktop/EVENTS| directory:
% 
% 1. Plot one of the events (listed in EVENTS directory, we choose EVENT-1 in this example):
% \begin{verbatim}
%     $ plot_station_event_distribution.py EVENT-1
% \end{verbatim}
% * For more information on the folder structure of your Virtual-Box, refer to Appendix-2.
% 
% \begin{center}
% %%\begin{figure}[htbp]
% \includegraphics[width=444pt, height=240pt, keepaspectratio=true]{AXISEMTutorial-fig006.png}
% %%\caption{This should be the caption for \texttt{AXISEMTutorial-fig006.png}.}
% %%\end{figure}
% 
% {\small{}Figure A2: Event-station configuration for EVENT-1}
% \end{center}
% 
% \baselineskip=13pt
% \leftskip=0pt
% 2. To get an overview on both real data and pre-simulated seismograms: (e.g. PREM\_ANISO 
% for 5 seconds dominant period) [Figure A3]
% \begin{verbatim}
%     $ plot_seismograms.py EVENT-1/AXISEM/PREM_ANISO_5sec
% \end{verbatim}
% 
% 3.Cut the waveforms for Pdiff and PKiKP seismic phases and plot the seismograms: 
% (compared with real data)
% \begin{verbatim}
%     $ plot_seismograms.py EVENT-1/AXISEM/PREM_ANISO_5sec Pdiff
%     $ plot_seismograms.py EVENT-1/AXISEM/PREM_ANISO_5sec PKiKP
% \end{verbatim}
% 
% \begin{figure}
% %%\begin{figure}[htbp]
% \includegraphics[width=497pt, height=287pt, keepaspectratio=true]{AXISEMTutorial-fig007.png}
% %%\caption{This should be the caption for \texttt{AXISEMTutorial-fig007.png}.}
% %%\end{figure}
% \begin{center}
% {\small{}Figure A3:Real and AXISEM waveforms for EVENT-1}
% \end{center}
% \end{figure}
% 
% \begin{figure}
% \centering
% \begin{minipage}{.5\textwidth}
%   \centering
%   \includegraphics[width=1.\linewidth]{AXISEMTutorial-fig008.pdf}
%   %\includegraphics[width=.4\linewidth]{image1}
% \end{minipage}%
% \begin{minipage}{.5\textwidth}
%   \centering
%   \includegraphics[width=1.\linewidth]{AXISEMTutorial-fig009.pdf}
% %   \includegraphics[width=.4\linewidth]{image1}
% \end{minipage}
% \begin{center}
% {\small{}Figure A4: Comparison between real and synthetic waveforms for Pdiff and 
% PKiKP phases. (without alignment)}
% \end{center}
% \end{figure}
% 
% 
% 4. Change the filter in \textit{plot\_seismograms.py} script and compare the results with 
% \textit{SPECFEM3D}: (for changing the filter, open \textit{plot\_seismograms.py} and change 
% values at top of the file) \\
% For this example, ????? we change hfreq (high frequency) in line 34 to 0.05Hz (20sec) and \textit{lfreq }to 0.01Hz: (Figure 
% A5)
% 
% \begin{verbatim}
%     $ plot_seismograms.py EVENT-1/AXISEM/PREM_ANISO_5sec Pdiff specfem3d
%     $ plot_seismograms.py EVENT-1/AXISEM/PREM_ANISO_5sec PKiKP specfem3d
% \end{verbatim}
% 
% 
% \begin{figure}
% \centering
% \begin{minipage}{.5\textwidth}
%   \centering
%   \includegraphics[width=1.\linewidth]{AXISEMTutorial-fig010.pdf}
%   %\includegraphics[width=.4\linewidth]{image1}
% \end{minipage}%
% \begin{minipage}{.5\textwidth}
%   \centering
%   \includegraphics[width=1.\linewidth]{AXISEMTutorial-fig011.pdf}
% %   \includegraphics[width=.4\linewidth]{image1}
% \end{minipage}
% \begin{center}
% {\small{}Figure A5: Comparison between real, AXISEM and SPECFEM3D waveforms for 
% Pdiff and PKiKP phases.}
% \end{center}
% \end{figure}
% 
% 5. Compare the results of two different background models (\textit{PREM\_ANISO\_5sec} with 
% \textit{IASP91\_5sec)} for Pdiff phase: (Figure A6).
% 
% \begin{verbatim}
%     $ plot_seismograms.py EVENT-1/AXISEM/PREM_ANISO_5sec Pdiff 
%     EVENT-1/AXISEM/IASP91_5sec
% \end{verbatim}
% 
% 6. Change the filter, as explained in step 4, and repeat step 5. Here, we increase 
% the \textit{hfreq} to 0.2Hz and \textit{lfreq} to 0.05Hz (Figure A7).
% 
% \begin{figure}
% \centering
% %%\begin{figure}[htbp]
% \includegraphics[width=234pt, height=310pt, keepaspectratio=true]{AXISEMTutorial-fig012.png}
% \begin{center}
% {\small{}Figure A6: Comparison between real and AXISEM waveforms for two different 
% background models (Pdiff).}
% \end{center}
% \end{figure}
% 
% \begin{figure}
% \centering
% %%\begin{figure}[htbp]
% \includegraphics[width=241pt, height=320pt, keepaspectratio=true]{AXISEMTutorial-fig013.png}
% \begin{center}
% {\small{}Figure A7: Comparison between real and AXISEM waveforms for two different 
% background models (Pdiff).} 
% \end{center}
% \end{figure}
% 
% 7. Compare the seismograms calculated for two different source parameters (PREM\_ANISO\_5sec and \\
% PREM\_ANSIO\_5sec\_GCMT) for Pdiff phase: (For more information about the source 
% parameters, refer to Appendix-1) [Figure A8]
% 
% \begin{verbatim}
%     $ plot_seismograms.py EVENT-1/AXISEM/PREM_ANISO_5sec Pdiff 
%     EVENT-1/AXISEM/PREM_ANISO_5sec_GCMT
% \end{verbatim}
% 
% \begin{figure}
% \centering
% \includegraphics[width=234pt, height=310pt, keepaspectratio=true]{AXISEMTutorial-fig014.png}
% \begin{center}
% {\small{}Figure A8: Comparison between real and AXISEM waveforms for two different 
% source parameters (Pdiff).}
% \end{center}
% \end{figure}
% 
% 8. Change the filter, as explained in step 4, and repeat step 7. ????? Here, we decrease 
% the \textit{hfreq} to 0.05Hz and \textit{lfreq} to 0.01Hz. (Figure A9)
% 
% 9. Find the time shift between the synthetics and real data, shift the synthetics 
% accordingly and plot the results: (Figure A10)
% 
% \begin{verbatim}
%     $ plot_seismograms.py EVENT-1/AXISEM/PREM_ANISO_5sec Pdiff shift_synthetics
% \end{verbatim}
% 
% \vspace{13pt}
% 10. Map the calculated time shifts in step 9 on the station locations: (note that 
% it always plots the results of the latest comparison, i.e. Pdiff here) 
% 
% \begin{verbatim}
%     $ plot_travel_time_map.py EVENT-1
% \end{verbatim}
% 
% \begin{center}
% %%\begin{figure}[htbp]
% \includegraphics[width=242pt, height=322pt, keepaspectratio=true]{AXISEMTutorial-fig015.png}
% %%\caption{This should be the caption for \texttt{AXISEMTutorial-fig015.png}.}
% %%\end{figure}
% 
% {\small{}Figure A9: Comparison between real and AXISEM waveforms for two different 
% source parameters (Pdiff).}
% 
% %%\begin{figure}[htbp]
% \includegraphics[width=249pt, height=331pt, keepaspectratio=true]{AXISEMTutorial-fig016.png}
% %%\caption{This should be the caption for \texttt{AXISEMTutorial-fig016.png}.}
% %%\end{figure}
% 
% {\small{}Figure A10: Comparison between real and AXISEM waveforms for Pdiff. AXISEM 
% waveforms are shifted in order to gain the maximum cross correlation coefficient.}
% 
% \vspace{1cm}
% 
% %%\begin{figure}[htbp]
% \includegraphics[width=446pt, height=217pt, keepaspectratio=true]{AXISEMTutorial-fig017.png}
% %%\caption{This should be the caption for \texttt{AXISEMTutorial-fig017.png}.}
% %%\end{figure}
% 
% {\small{}Figure A11: Time shift calculated by cross correlation between real and 
% AXISEM waveforns. Blue color indicates that Pdiff in real data arrived sooner than 
% that in the synthetic one.}
% \end{center}
% 
% \baselineskip=13pt
% \leftskip=0pt

\end{document}

### 2015-08-10 utils/EXTERNAL_CODES_coupled_with_SPECFEM3D/AxiSEM_for_SPECFEM3D/AxiSEM_modif_for_coupling_with_specfem/MANUAL/automated_testing.tex

\subsection{Automated testing}

This section is mainly useful for the develpers or those want to change/add/remove some parts of the code and compare the new changes with the reference solutions.
For the reference solutions, 5 different tests have been designed and 
the results from \textit{yspec} [Al-Attar \& Woodhouse (2008)] and AXISEM have been included in the \textit{TESTING/automatic} directory.
The whole procedure (running the code, compare the results and plot) is automatic with the least user intervention: \\

Start from within the {\tt AXISEM} directory:
 \begin{enumerate}
 \item {\tt cd TESTING}
 \item {\tt python test\_axisem.py}
 \end{enumerate}

Enter test number(s) and this is all you should do! \\

\noindent As an example, we want to run the script \textit{test\_axisem.py} for test number 5. (Figure~\ref{test_axisem})

\begin{figure*}[htb]
  \begin{verbatim}
   ==========================================
  Please enter the test number(s): 
  1. test01: explosion
  2. test02: dipole (mxz)
  3. test03: quadpole (mxy)
  4. test04: CMT (location: north pole)
  5. test05: CMT (location: 70°N 50°E)

  (format = 01,02 OR 1,2,3)
  \end{verbatim}
  \caption{\textit{Screenshot while running test\_axisem.py}}
\label{test_axisem}
\end{figure*}

\noindent At the end, it plots three figures (one for each channel) 
in which the new AXISEM waveforms are compared with both the original ones and \textit{yspec} results for the same simulation.
%In Figure~\ref{channel_Z}, only the Z channel has been plotted.
It should be noted that these tests are designed for rough comparison purposes (in terms of the pulse shape and sign) and they should not be considered as a detailed benchmark with respect to YSPEC. \\

\begin{figure}
  \includegraphics[width=0.48\textwidth]{./PYAXI/record_section_Z.png}
  \includegraphics[width=0.48\textwidth]{./PYAXI/l2_misfit.png}
  \caption{\textit{Comparing new AXISEM results with the reference solution and \textit{yspec} waveforms. (Z channel)}}
  \label{channel_Z}
\end{figure}

\noindent Another output is the RMS misfit (Figure~\ref{l2_misfit}) between the new and reference data 
in which the traces can be compared in a quantitative sense.

% \begin{figure}
% 
%   \caption{\textit{RMS misfit between the new and reference data.}}
%   \label{l2_misfit}
% \end{figure}

### 2015-08-10 utils/EXTERNAL_CODES_coupled_with_SPECFEM3D/AxiSEM_for_SPECFEM3D/AxiSEM_modif_for_coupling_with_specfem/MANUAL/manual_axisem1.1.tex

%&pdfLaTeX
% !TEX encoding = UTF-8 Unicode
\documentclass{article}
\usepackage{ifxetex}
\ifxetex
\usepackage{fontspec}
\setmainfont[Mapping=tex-text]{STIXGeneral}
\else
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\fi
\usepackage{textcomp}

\usepackage{tabularx}
\usepackage{graphicx}
\usepackage{ulem}
\usepackage{amssymb}
\usepackage{fancyhdr}
\usepackage{float}
\usepackage{url}
\usepackage[colorlinks=true,
            linkcolor=red,
            urlcolor=blue,
            citecolor=gray]{hyperref}
\newcommand{\ttilde}[0]{\raise.17ex\hbox{$\scriptstyle\sim$}}
\renewcommand{\vec}[1]{\ensuremath{\mathbf{#1}}}
\newcommand{\evec}[1]{\ensuremath{\mathbf{\hat #1}}}

\renewcommand{\headrulewidth}{0pt}
\renewcommand{\footrulewidth}{0pt}
\usepackage{color}
\usepackage[a4paper,vmargin={20mm,20mm},hmargin={20mm,10mm}]{geometry}
% \usepackage{listings}
% \lstset{inputencoding=ansinew}

\definecolor{color18}{rgb}{0.07,0.33,0.80}

\setlength\parindent{0pt}

\usepackage{times}
\usepackage{textpos}

\begin{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% START OF CIG MANUAL COVER TEMPLATE %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% This should be pasted at the start of manuals and appropriate strings entered at locations indicated with FILL.
% Be sure the TeX file includes the following packages.
% \usepackage{graphicx}
% \usepackage{times}
% \usepackage{textpos}

\definecolor{dark_grey}{gray}{0.3}

%LINE 1%
{
\renewcommand{\familydefault}{\sfdefault}

\pagenumbering{gobble}
\begin{center}
\resizebox{\textwidth}{!}{\textcolor{dark_grey}{\fontfamily{\sfdefault}\selectfont 
COMPUTATIONAL INFRASTRUCTURE FOR GEODYNAMICS (CIG)
}}

\hrule

%LINE 2%
\color{dark_grey}
\rule{\textwidth}{2pt}

%LINE 3%
\color{dark_grey}
% FILL: additional organizations
% e.g.: {\Large Organization 1\\Organization 2}
{\Large }
\end{center}

%COLOR AND CODENAME BLOCK%
\begin{center}
\resizebox{\textwidth}{!}{\colorbox
% FILL: color of code name text box
% e.g. blue
{blue}{\fontfamily{\rmdefault}\selectfont \textcolor{white} {
% FILL: name of the code
% e.g. PyLith
% You may want to add \hspace to both sides of the codename to better center it, such as:
% \newcommand{\codename}{\hspace{0.1in}CodeName\hspace{0.1in}}
\hspace{0.1in}AxiSEM\hspace{0.1in}
}}}
\end{center}

%MAIN PICTURE%
\begin{textblock*}{0in}(0.7in,0.2in)
% FILL: image height
% e.g. height=6.5in
\begin{center}
\vspace{.5in}
\includegraphics[height=5.5in]
% FILL: image file name
% e.g. cover_image.png
{Axisem2png.png}
\end{center}
\end{textblock*}

%USER MANUAL%
\color{dark_grey}
\hfill{\Huge \fontfamily{\sfdefault}\selectfont User Manual \\
% FILL: manual version
% e.g. 1.0
\raggedleft \huge \fontfamily{\sfdefault}\selectfont Version {1.1}\\}

%AUTHOR(S) & WEBSITE%
\null
\vfill
\color{dark_grey}
\Large \hfill {\raggedleft \fontfamily{\sfdefault}\selectfont 
% FILL: author list
% e.g. Author One\\Author Two\\Author Three\\
% be sure to have a newline (\\) after the final author
Tarje Nissen-Meyer\\Martin van Driel\\Simon St\"{a}hler\\Kasra Hosseini\\Stefanie Hempel\\Alexandre Fournier\\
}
{\fontfamily{\sfdefault}\selectfont www.geodynamics.org}

\hrule

%LINE% 
\color{dark_grey}
\rule{\textwidth}{2pt}

}

\pagebreak
\pagenumbering{arabic}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%   END OF CIG MANUAL COVER TEMPLATE    %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{center}
    \begin{tabularx}{\textwidth}{lX}
        \begin{minipage}{0.3\textwidth}
            \begin{center}
                \includegraphics[width=\textwidth]{Axisem2png.png}
            \end{center}
        \end{minipage}
        &
        \begin{minipage}{0.6\textwidth}
            \begin{center}
                \title{}
                \LARGE{ \textbf{\sc AxiSEMv1.1 Manual}}
                \vspace*{0.6cm}\\
                {\large 
                Tarje Nissen-Meyer\textsuperscript{1},
                Martin van Driel\textsuperscript{2},
                Simon St\"{a}hler\textsuperscript{3},
                Kasra Hosseini\textsuperscript{3},
                Stefanie Hempel\textsuperscript{4}, 
                Alexandre Fournier\textsuperscript{5}}
                \vspace*{0.3cm}\\
                {\small  \textsuperscript{1} Oxford University (UK),
                  \textsuperscript{2} ETH Zurich (Switzerland),  \textsuperscript{3}
                  LMU M\"{u}nchen (Germany), \textsuperscript{4} Universit\"{a}t
                  M\"{u}nster (Germany), \textsuperscript{5} IPG Paris
                  (France)}\\
                {\large Oxford, UK -- July 7, 2014}
            \end{center}
        \end{minipage}
    \end{tabularx}
\end{center}

%
%
\noindent {AxiSEM} is a \textbf{parallel spectral-element method} to solve
3D wave propagation in a sphere with axisymmetric or spherically symmetric
visco-elastic, acoustic, anisotropic structures. Such media allow the
computational domain to be collapsed to a 2D disk, where the third, azimuthal
dimension is solved analytically on-the-fly posteriori. This leads to extreme
speedup by many orders of magnitude with respect to methods that discretize the
3D domain, and enables a full coverage of the seismic body- and surface wave
frequency spectrum between 0.001-1Hz.  The time-domain code delivers full
spatio-temporal wavefields that can be stored on disk and transformed to
frequency domain. Due to the dimensional reduction, global wave propagation at
typical seismic of \textbf{periods down to 5 seconds can be tackled on
laptops}, and at 1Hz on moderate clusters.

The Fortran 90 code is divided into a \textbf{Mesher}, a \textbf{Solver}
utilizing the message-passing interface (MPI) for communication between
separate domains, and comprehensive \textbf{post processing} for ease of
visualization.
The essential raison-d'\^{e}tre of this method is the efficient
calculation of seismograms, wavefield movies, and those wavefields that underly
sensitivity kernels to allow for tomographic inversions of any portion of a
seismogram at any relevant frequency.

\begin{center}
\textbf{Portal for this code:} \href{http://www.axisem.info}{www.axisem.info}\\
\textbf{Contact:} \href{mailto:info@axisem.info}{info@axisem.info}
\end{center}

%
\noindent \textbf{Principal authors:} \\
Tarje Nissen-Meyer, Alexandre
Fournier, Martin van Driel, Simon St\"{a}hler, Kasra Hosseini, Stefanie Hempel.\vspace*{0.3cm}\\
\noindent \textbf{Contributions:}\\
 J.-P. Ampuero, E. Chaljub, A. Colombi, F. A. Dahlen, D. Komatitsch,
 G. Nolet, J. Tromp.\vspace*{0.3cm}\\
\noindent \textbf{Research funding:} \\Princeton University, NSF, HP2C Petaquake, 
QUEST ITN (Marie Curie), ETH Zurich, Oxford University.\vspace*{0.3cm}\\
%
\noindent \textbf{Copyright}\\
\copyright  \hspace*{0.1cm} 
2013, Tarje Nissen-Meyer,
Alexandre Fournier, Martin van Driel, Simon St\"{a}hler, Kasra Hosseini, Stefanie Hempel.\vspace*{0.3cm}\\

AxiSEM is free software: you may redistribute it and/or modify it
under the terms of the GNU General Public License as published by the
Free Software Foundation, either version 3 of the License, or any
later version.

AxiSEM is distributed in the hope that it will be useful, but WITHOUT
ANY WARRANTY; without even the implied warranty of MERCHANTABILITY or
FITNESS FOR A PARTICULAR PURPOSE. Commercial use must be discussed
with the authors prior to usage. See the GNU General Public License
for more details: \verb|LICENSE_GPL.txt|
\newpage
\tableofcontents
%
\vspace*{1cm}

\section{Preliminaries}

\subsection{The AxiSEM Concept}

The basic idea behind AxiSEM is to take advantage of axial symmetry with respect
to an axis going through the center of the earth and the source. In such axisymmetric 
models, the response to a moment tensor or single force point source can be expanded in a
series of multipoles (mono-, di- and quadrupole). The dependence of the 3D wavefield on
azimuth $\phi$ can be solved analytically and the remaining 2D problems (four of them for
a full moment tensor source) are solved numerically using a spectral
element approach.\\

\begin{center}
    \begin{minipage}[t]{0.4\paperwidth}
        Source Decomposition:\\
        \begin{minipage}[c]{0.1\paperwidth}
            \includegraphics[width=\textwidth]{radpat_mono.pdf}%
        \end{minipage}%
        \begin{minipage}[c]{0.4\paperwidth}
            $\scriptsize{\vec u = \vec u (s,z)}$ \\
        \end{minipage}

        \begin{minipage}[c]{0.1\paperwidth}
            \includegraphics[width=\textwidth]{radpat_di.pdf}
        \end{minipage}%
        \begin{minipage}[c]{0.4\paperwidth}
            $\scriptsize{\vec u = \vec u (s,z)} \cdot f(\sin\phi, \cos\phi)$ \\
        \end{minipage}

        \begin{minipage}[c]{0.1\paperwidth}
            \includegraphics[width=\textwidth]{radpat_quad.pdf}
        \end{minipage}%
        \begin{minipage}[c]{0.4\paperwidth}
            $\scriptsize{\vec u = \vec u (s,z)} \cdot f(\sin(2\phi), \cos(2\phi))$ \\
        \end{minipage}%
    \end{minipage}%
    \begin{minipage}[t]{0.25\paperwidth}
        2D numerical problems:

        \includegraphics[width=.7\textwidth]{mesh.png}
    \end{minipage}%
\end{center}

\newpage

\subsection{Software and hardware requirements}

\subsubsection{Essential requirements}

\begin{description}
    \item[Operating system] The software should run on any UNIX-like operating system and has
          been tested on Linux (Debian, Ubuntu, SuSe, RedHat, Mageia, Cray et al) and MacOS X. 
    \item[Compilers:] Fortran 90 compiler (tested on ifort, gfortran-4.6, portland, Cray)
    \item[Libraries:] MPI, NetCDF (optional), fftw (optional)
    \item[Systems:] Unix-based OS on desktop scale (tested on Linux and MacOS)\\ and HPC scale (tested on Cray XT4, XE6 and XK7, HECToR (UK National Supercomputing Service), SuperMUC (Leibniz-Rechenzentrum, Garching), Pacman (Arctic Region Supercomputing Center).
    \item[Tools:] tcshell, perl
\end{description}\vspace*{0.3cm}

\textbf{Useful tools}

\begin{description}
    \item[ObsPy:] Needed for the automated tests. Follow the instructions for your system
          on \url{http://www.obspy.org}.
    \item[Paraview:] Can be used to check meshes and watch wavefield movies.
    \item[Google Earth:] Can be used for a quick overview on seismograms at the locations
          of the receivers.
    \item[gnuplot:] Used to make quick overview plots
    \item[matlab:] Used for plotting record section (script included)
\end{description}

\subsubsection{NetCDF}

AxiSEM allows to output larger datasets, especially wavefields in the NetCDF format. The
current version also has full support for binary dumps, but the development will move
towards NetCDF containers.\\
Unfortunately, the installation of the NetCDF libraries may still be cumbersome; below we list some limited support for installing NetCDF..

\begin{description}
    \item[HPC systems:] The libraries should be provided by the system. Use the
          recommended settings.
    \item[Ubuntu 12.10 and newer, Debian Wheezy and newer]: The code is working with the NetCDF libraries delivered
          with the system (for gfortran). They can be installed by \\
          \verb|sudo apt-get install libnetcdff5|
    \item[Ubuntu 12.04 and older; MacOS]: The libraries delivered with Ubuntu 12.04 and
          earlier do not seem to work reliably. We therefore generally recommend to
          compile the NetCDF libraries from source. This can be done with the script
          \textit{make\_netcdf.sh} in the \verb|SOLVER/UTILS| directory. It downloads
          current versions of the zlib, hdf5 and netcdf4 libraries from
          \url{http://www.unidata.ucar.edu}, compiles them and runs the included tests. By
          default, the new libraries are installed in \verb|$(HOME)/local|. In the first
          lines of the script, specify your compiler (has to be the same as the one you
          are using for AxiSEM). The script should be run from a scratch directory like
          \verb|/tmp|:\\
          \verb|cd /tmp|\\
          \verb|$AXISEM_DIRECTORY/SOLVER/UTILS/make_netcdf.sh|\\
          Especially the HDF5 compilation can half an hour and will produce tons of warnings. They can be
          ignored, as long as the tests pass. If one of the tests should fail, the reason
          is most likely a wrong compiler configuration. We can offer only very limited
          support for the compilation of the libraries.
    \item[Windows]: While we never tested it, installation of NetCDF on Windows should be
          possible: \url{http://www.unidata.ucar.edu/software/netcdf/docs/faq.html#windows_netcdf4_2}
\end{description}

\subsection{Preparation of a Debian/Ubuntu Linux system}
To prepare a fresh Debian-based Linux system, the absolutely necessary packages can be installed with:\\ \\
 \verb|sudo apt-get install gfortran build-essential tcsh openmpi-bin libopenmpi-dev|\\ \\
The processing and visualization tools can be installed with:\\ \\
 \verb|sudo apt-get install paraview gnuplot|

\newpage

\section{Running the code}

\subsection{Quick start}

This is the step-by-step, blackbox procedure, i.e. running a workflow from raw source code
to analyzing seismograms and wavefield movies upon pre-set parameters. It assumes your
system fulfils all requirements mentioned above.\\

\begin{center}
    \fbox{\parbox{0.65\textwidth}{
    The default simulation parameters are: 
    \begin{description}
        \item[PREM] velocity model (isotropic, anelastic, continental crust)
        \item[50 s] dominant period of the mesh
        \item[2 CPUs] used for the SOLVER
        \item[1800s] seismogram length
        \item[Vertical dipole] source
        \item[100 km] source depth
    \end{description}
    } }
\end{center}

\noindent Start from within the {\tt AXISEM} directory:

\begin{enumerate}
    \item \verb|./copytemplates.csh| $\Rightarrow$ creates various input files from templates
    \item Check file \verb|make_axisem.macros| whether the compiler settings fit your
          system.
    \item \verb|cd MESHER| 
    \item Check file {\tt inparam\_mesh} for background model, period of simulation and
          number of CPUs \\
          Default is \textit{PREM}, \textit{50 s} and \textit{2 CPUs} to be running within a
          few minutes on a modern PC.
    \item \verb|./submit.csh| $\Rightarrow$ Check file {\tt OUTPUT}.
    \item Wait for ``{\tt ....DONE WITH MESHER}'' to appear in {\tt OUTPUT}.
    \item move mesh files to \verb|../SOLVER/MESHES/| directory and give it a name
          \verb|./movemesh.csh PREM_50s| 
    \item \verb|cd ../SOLVER|
    \item In \verb|inparam_basic| set the value for \verb|MESHNAME| to the meshname from
          above (here: \verb|PREM_50s|)
    \item \verb|./submit.csh PREM_mrr_50s_gauss_1800s|  $\Rightarrow$ compiles and runs
          the code
    \item \verb|cd PREM_mrr_50s_gauss_1800s| $\Rightarrow$ go to the run directory.
    \item Wait for ``\verb|PROGRAM axisem FINISHED|'' to appear in
          \verb|OUTPUT_PREM_mrr_50s_gauss_1800s| \\
          (use \verb|tail -f OUTPUT_PREM_mrr_50s_gauss_1800s|).
    \item \verb|./post_processing.csh|
    \item \verb|cd Data_Postprocessing| 
    \item \verb|googleearth|, open {\tt googleearth\_src\_rec\_processed.kml}, click
            earthquake (info), receivers (seismograms).
    \item {\tt matlab}, run {\tt plot\_record\_section.m}, plotting all components
            of displacement seismograms.

\end{enumerate}
If the Solver is re-run with different parameters but the same mesh, you may start at step 9. 
To change model, frequency or number of CPUs, repeat steps 3. to 7. and select the new
mesh in \verb|SOLVER/inparam_basic|. \\
The solver input can be changed in \verb|inparam_basic| between 8. and 9.,
changing post-processing input between 11. and 12. Using a new mesh requires
recompilation of the solver (done automatically in step 9.). If post
processing parameters are changed, also change the post processing directory or
delete the old one.

\subsection{MESHER - generate a Mesh}

\begin{enumerate}
    \item Open a terminal, go to the \ttilde\verb|MESHER| folder and open
    the \verb|inparam_mesh| file with your favourite editor:
    %
    \begin{verbatim}
    $ cd MESHER
    $ vi inparam_mesh
    \end{verbatim}
    %
    The parameters should be readily set, but you might want to double check and verify:
    %
    \begin{verbatim}
    BACKGROUND_MODEL    'prem_iso'
    DOMINANT_PERIOD     50.0
    NTHETA_SLICES       2
    NRADIAL_SLICES      1 
    WRITE_VTK           true
    COARSENING_LAYERS   3
    \end{verbatim}
    %
    The file should be self-explanatory. NB: Models without crust ('light') allow for a
    larger time step and hence run a lot faster. \\
    WARNING: Only write vtk files if the dominant period is rather large, i.e. above 10 or
    20s, as these files become exceedingly large.

    \item Run the mesher, and watch the progress:
    %
    \begin{verbatim}
    $ ./submit.csh
    $ tail -f OUTPUT
    \end{verbatim}
    %
    The meshing should be really fast (i.e. on the order of seconds) for the chosen parameters. Wait for 
    \verb|....DONE WITH MESHER !| to appear.

    \item Take a look at the mesh with paraview
    \begin{verbatim}
    $ paraview
    \end{verbatim}
    %
    Open one of the vtk files in the subfolder \verb|Diags|, e.g.\
    \verb|mesh_vp.vtk| and click apply in the properties panel on the left (you
    might get an OpenGL Error on the virtual box, which you can ignore). To see
    the mesh, change the representation from 'surface' to 'surface with edges'
    (On some host systems, the dropdown menu seems to be messed up, in that
    case go to the 'Display' context in the 'Properties' panel on the left. If
    the plot appears all yellow, click on play).  You can open other vtk files
    to look at other properties of the model and the mesh. You might need to
    rescale the color range by clicking on the left-right arrow symbol in the
    top left.

    \begin{center}
    \includegraphics[height=30mm]{paraview.png}
    \hspace{5mm}
    \includegraphics[height=30mm]{paraview3.png}
    \end{center}

    \item Move the mesh to the solver directory and give it a meaningful name:
    \begin{verbatim}
    $ ./movemesh.csh PREM_50s
    \end{verbatim}

\end{enumerate}

\newpage
\subsection{SOLVER - solve the elastic wave equation}

\begin{enumerate}
    \item Go to the \ttilde\verb|SOLVER| folder and open the
    \verb|inparam_basic| file with your favourite editor:
    %
    \begin{verbatim}
    $ cd ../SOLVER
    $ vi inparam_basic
    \end{verbatim}
    %
    Set these parameters:
    %
    \begin{verbatim}
    SIMULATION_TYPE     single
    SEISMOGRAM_LENGTH   1800.
    RECFILE_TYPE        stations 
    MESHNAME            PREM_50s
    ATTENUATION         true 
    SAVE_SNAPSHOTS      true  
    \end{verbatim}
    %
    WARNING: Only save snapshots if the mesh is rather low resolution,
    e.g. above 20s as these files become exceedingly large. You may
    alternatively opt to only plot a fraction of the 2D domain which
    can be set in \verb|inparam_advanced|.

    \item First, we are taking a look at a basic sourcetype: a vertical dipole, which has
    a monopole radiation pattern. This is set by \verb|SIMULATION_TYPE single| and defined
    in the \verb|inparam_source| file. Run the solver, giving the run a meaningful name:
    %
    \begin{verbatim}
    $ ./submit.csh PREM_mrr_50s_gauss_1800s
    \end{verbatim}
    %
    This command compiles the code if needed and starts the simulation. You can observe
    the progress in the outputfile:
    %
    \begin{verbatim}
    $ cd PREM_mrr_50s_gauss_1800s
    $ tail -f OUTPUT_PREM_mrr_50s_gauss_1800s
    \end{verbatim}
    %
    Once the run is finished, take a look at the wavefield with \verb|paraview|: open
    the \verb|PREM_mrr_50s_gauss_1800s/| \verb|Data/xdmf_xml_0000.xdmf|
    file and click apply. Go to the last snapshot and rescale the color range,
    then click on play to see the wave propagate.  You can also choose
    different components of the wavefield or the absolute value.  For paraview
    experienced users: choose absolute value and a logarithmic colorscale to
    see all wave types at once (e.g.\ 'black body radiation' looks nice).

    \begin{center}
    \includegraphics[width=110mm]{paraview2.png} \hspace{5mm}
%    \includegraphics[width=40mm]{0600.png}
    \end{center}

    \item Now simulate seismograms for a full moment tensor source: the source is defined
    in the \verb|CMTSOLUTION| file and the one referred to as 'event-1' in the later tasks.
    Stations are defined in the \verb|STATIONS| file. Go back to the \verb|SOLVER|
    directory and change the \verb|inparam_basic| file such that:
    %
    \begin{verbatim}
    SIMULATION_TYPE     moment
    SAVE_SNAPSHOTS      false
    \end{verbatim}
    %
    Run the solver, giving the run a meaningful name:
    %
    \begin{verbatim}
    $ ./submit.csh prem_50s_event1
    \end{verbatim}
    %
    This command compiles the code if needed and starts four simulations at once, each
    simulating a basic source type (two monopoles, a dipole and a quadrupole, for details
    see \textit{Nissen-Meyer et al., 2007}). You can observe the progress in the outputfiles in
    each job's subdirectory
    %
    \begin{verbatim}
    $ cd prem_50s_event1
    $ tail -f MZZ/OUTPUT_MZZ
    \end{verbatim}
    %
    Once all the jobs are done (check with \verb|htop|), you can proceed with
    postprocessing.

\end{enumerate}

\subsection{POSTPROCESSING - rotate and sum seismograms and wavefields}

Postprocessing is a key feature of AxiSEM: the source mechanism and source time function
can be modified without redoing the more expensive simulation.

\begin{enumerate}
    \item For the previous simulation, the contribution of the elemental sources needs
    to be summed up to get seismograms for a full moment tensor source. In the
    main rundirectory (\verb|prem_50s_event1|) open the file
    \verb|param_postprocessing|. It should contain these settings (auto generated by the
    solver):
    %
    \begin{verbatim}
    REC_COMP_SYS    enz
    CONV_PERIOD     0.
    CONV_STF        gauss_0
    \end{verbatim}
    %
    The source mechanism (depth and location cannot be changed in postprocessing) is read
    from the \verb|CMTSOLUTION| file in the same directory. Start the postprocessing:
    %
    \begin{verbatim}
    $ ./postprocessing.csh
    \end{verbatim}
    %
    The resulting seismograms and plots can be found in the directory 
    \verb|Data_Postprocessing|. Seismograms can be viewed with your favorite image viewer,
    e.g.\ \verb|eog|:
    %
    \begin{verbatim}
    $ cd Data_Postprocessing/GRAPHICS
    $ eog <filename.gif>
    \end{verbatim}
    %
    For a nice overview, you can use \verb|google-earth| (might not run on all computers
    and depends on internet connection). Open the \verb|googleearth_src_rec_seis.kml| file
    in the \verb|Data_Postprocessing/| directory (double check the exact path,
    google-earth might have something older from history which is quite confusing).
    You should now see the earthquake and the receivers in the places menu on the left.

    \begin{center}
    \includegraphics[height=70mm]{google-earth.png}
    \hspace{5mm}
    \includegraphics[height=70mm]{google-earth2.jpg}
    \end{center}

    Click on the stations or source to see more...

\end{enumerate}

\subsection{Computational aspects}

Running in a 2D computational domain, the code is (obviously) significantly faster
than comparable 3D methods.

\begin{center}
    %\includegraphics[height=50mm]{mesher_comp_cost.pdf}
    \includegraphics[height=50mm]{memory_usage_new.pdf}
    \hspace{5mm}
    \includegraphics[height=50mm]{solver_totalcpu.pdf}
\end{center}

On the left, you may deduct the mesher's RAM occupation as a function of frequency. Going
towards very high resolution (around and above 1Hz), you will need a rather fat node (>
16GB RAM) for the (shared memory) meshing.
On the right, we depict the computational cost
associated with the solver to compute seismograms of 30 min length. The relation between
seismic period and CPU-hours is an approximate proxy to estimate how many cores and
wall-clock time is optimal for your infrastructure. Scaling of CPU hours is approximately 
to the third power of the maximum frequency.

\begin{center}
    \includegraphics[height=50mm]{COMPUTATIONAL_COST/strong_scaling_new.pdf}
    \hspace{5mm}
    \includegraphics[height=50mm]{COMPUTATIONAL_COST/weak_scaling_new.pdf}
\end{center}

Scalability plots from test runs on a Cray machine at CSCS, Switzerland. The left shows
strong scaling (fixed global problem size), the right plot weak scaling (fixed problem
size per CPU).

\section{Typical use cases}

\subsection{Change source type}

\verb|SOLVER/inparam_source|, \verb|SOLVER/CMTSOLUTION| and \verb|SOLVER/inparam_basic|.
AxiSEM has two principal modes, which are selected by the value \verb|SIMULATION_TYPE| in
\verb|SOLVER/inparam_basic|.

\begin{enumerate}
\item \verb|SIMULATION_TYPE single|:\\ The Solver simulates one basic source, which can be
      one of the following:\\
      \begin{center}
        \begin{tabular}{|l|cccc|} \hline
         monopoles:   & $M_{rr}$ & explosion  & $M_{\theta\theta}+M_{\phi\phi}$  & $P_r$ \\
         dipoles:     & $M_{\theta r}$ & $M_{\phi r}$        & $P_{\theta}$ & $P_{\phi}$\\
         quadrupoles: & $M_{\theta \phi}$ & $M_{\theta\theta}-M_{\phi\phi}$ & & \\\hline
        \end{tabular}
      \end{center}
      where $M_{ii}$ are moment tensor sources with the mentioned components of $M$ set to
      one and the others to zero, $P_i$ is the same for single forces.\\ Choose the source
      type and set the source depth and amplitude in \verb|SOLVER/inparam_source|.

\item \verb|SIMULATION_TYPE moment|:\\
      The submit.csh script starts four separate simulations for the basic types $M_{rr}$,
      $M_{\theta\theta}+M_{\phi\phi}$, $M_{\theta r}$, $M_{\theta \phi}$. You have to run
      the postprocessing script after the simulation to sum them up correctly. \\ Before
      the simulation, set the source depth and the moment tensor in
      \verb|SOLVER/CMTSOLUTION|. \\Run \verb|postprocessing.csh| in the simulation
      directory afterwards. You can run postprocessing for different moment tensors on the
      same simulation, but not for different depths (since the forward simulation depends
      on the depth).  The \verb|CMTSOLUTION| file is a standard format and can be
      downloaded from many sites in the web, including
      \url{http://www.globalcmt.org/CMTsearch.html}
\end{enumerate}

\subsection{Change station locations}

\verb|SOLVER/STATIONS| or \verb|SOLVER/receivers.dat|, depending on
\verb|SOLVER/inparam_basic|, parameter \verb|RECFILE_TYPE|\\

\verb|SOLVER/STATIONS|: Similar to \textit{SPECFEM3D Globe}, an ASCII file with six
columns, which are: station name, network name, latitude, longitude, elevation, depth
(n.b: AxiSEM puts all receivers to the surface, the last two rows are ignored).

\begin{verbatim}
 RAYN   GD  23.52   45.50  0.0  0.0
 PALK   GD   7.27   80.70  0.0  0.0
 MAJO   GD  36.54  138.21  0.0  0.0
 ERM    GD  42.02  143.16  0.0  0.0
\end{verbatim}

The station names are used by post\_processing to assign names to the seismogram files.\\
\verb|SOLVER/receivers.dat|: Plain ASCII file with number of receivers in the first line
and then nrec lines with colatitude and longitude.

\begin{verbatim}
 7
0.0 0.0
30.0 0.0
60.0 0.0
90.0 0.0
120.0 0.0
150.0 0.0
180.0 0.0
\end{verbatim}

\subsection{Change background model}

\verb|MESHER/inparam_mesh|, parameter \verb|BACKGROUND_MODEL|:\\
afterwards the steps from 5 have to be rerun. Supported models are

\begin{verbatim}
# prem_iso:               Isotropic continental PREM model
# prem_iso_solid:         like 'prem_iso', replace fluid outer core with vs=vp/sqrt(3)
# prem_iso_onecrust:      like 'prem_iso' but extend lower crust to surface
# prem_iso_light:         like 'prem_iso' but with mantle material extended to surface
# prem_iso_solid_light:   like 'prem_light', but in fluid outer core vs=vp/sqrt(3)
#
# prem_ani:               Anisotropic continental PREM model (actual PREM)
# prem_ani_onecrust:      like 'prem_ani' but extend lower crust to surface
# prem_ani_light:         like 'prem_ani' but with mantle material extended to surface
# 
# ak135                   AK135 (Isotropic, PREM attenuation)
# ak135f                  AK135 (Isotropic, own attenuation)
# iasp91:                 Isotropic IASP91 model with PREM density and attenuation
# external:               Layered external model, give file name in EXT_MODEL, the 
#                         inner core needs to be big enough, check VTK output.
\end{verbatim}

\subsection{Use external 1D velocity model}

\verb|MESHER/inparam_mesh|, change parameter \verb|BACKGROUND_MODEL| to \verb|external|
and \verb|EXT_MODEL| to the filename of your model. The model should be stored in a file
of the following form:

\begin{verbatim}
ANELASTIC       T
ANISOTROPIC     T
UNITS           m
COLUMNS     radius    rho    vpv    vsv    qka    qmu  vph   vsh   eta
            6371000.  2600.  5800.  3200.  57827. 600. 5800. 3200. 1.00000
            6356000.  2600.  5800.  3200.  57827. 600. 5800. 3200. 1.00000
#          Discontinuity   1, depth:      15.00 km
            6356000.  2900.  6800.  3900.  57827. 600. 6800. 3900. 1.00000
            6346600.  2900.  6800.  3900.  57827. 600. 6800. 3900. 1.00000
#          Discontinuity   2, depth:      24.40 km
            6346600.  3380.  8190.  4396.  57827. 600. 8190. 4611. 0.90039
            6341600.  3380.  8186.  4397.  57827. 600. 8186. 4607. 0.90233
            6336600.  3379.  8183.  4398.  57827. 600. 8183. 4602. 0.90428
            6331600.  3379.  8179.  4399.  57827. 600. 8179. 4598. 0.90622

\end{verbatim}

In the header, four keywords are mandatory to describe the following model:
\begin{itemize}
 \item \verb|ANELASTIC|: Is the model anelastic (viscoelastic) or not?
 \item \verb|ANISOTROPIC|: Is the model anisotropic or not?
 \item \verb|UNITS|: Are the units in this file given in \textit{SI}-units? Allowed values:
       \begin{itemize}
        \item \verb|m| \textit{SI}-units: (m, m/s, kg/m$^3$)
        \item \verb|km| km, km/s, g/cm$^3$. These units are often used because the values are between 1 and 10.
       \end{itemize}
 \item \verb|COLUMNS|: Describe the column order in the file. Necessary values: 
	\begin{itemize}
	\item For an elastic, isotropic model: 
		\begin{itemize}
		\item \verb|radius|/\verb|depth| (radius or depth of the layer. Allows to define the position of the layers in radius (beginning from the center) or depth (beginning from the surface)
		\item \verb|rho| ($\rho$, density)
		\item \verb|vpv| ($\alpha$, vertical P-velocity)
		\item \verb|vsv| ($\beta$, vertical S-velocity)
		\end{itemize}
	\item For an anelastic model additionally: 
		\begin{itemize}
		\item \verb|qka| ($Q_{\kappa}$)
		\item \verb|qmu| ($Q_{\mu}$)
		\end{itemize}
	\item For an anisotropic model additionally: 
		\begin{itemize}
		\item \verb|vph| (horizontal P-velocity)
		\item \verb|vsh| (horizontal S-velocity)
		\item \verb|eta| (anisotropic parameter $\eta$)
		\end{itemize}
	\end{itemize}
\end{itemize}

This header is followed by an arbritrary number of lines with velocity layers, where \verb|#| marks comment lines. The header variables \verb|UNITS| and \verb|COLUMNS| allow to customize the structure of these lines, which should facilitate the import of external models from other programs.

First order discontinuities are enforced by double layers with the same radius (see layers 2/3 and 4/5
in the example) and are honoured by the MESHER.

For an example file, select one of the internal models in \verb|MESHER/inparam_mesh|, enable the option \verb|WRITE_1DMODEL| and run the MESHER. It will write a valid input file of the selected model into \verb|MESHER/Diags/1dmodel_axisem.bm|. Modify this file according to your needs.

The overall radius of the body is given by the radius of the outermost layer and can take
any reasonable value (Read: Planets, moons and tennis balls can be simulated, single electrons probably not).

\subsection{Change number of CPUs}

\verb|MESHER/inparam_mesh|, parameter \verb|NTHETA_SLICES| and \verb|NRADIAL_SLICES|: \\
The number of CPUs used is the product of the two parameters. \\
\verb|NTHETA_SLICES| needs to be 1, 2, 4 or a multiple of 4. To get a suggestion for
optimal decomposition, run the Mesher with \verb|ONLY_SUGGEST_NTHETA true| and check the
OUTPUT file.\\
\verb|NRADIAL_SLICES| should be on the order of 8, larger numbers work for very high
frequencies. It can be left at 1 for \verb|NTHETA_SLICES|<64 CPUs, but should be increased
then to reduce MPI communication. To ensure scaling, each processor should have at least
about 500 elements.\\
N.B: This value is for ONE simulation. To calculate the wavefield of a full moment tensor,
4 parallel simulations have to be run and the number of necessary CPUs is
\verb|NTHETA_SLICES|* \verb|NRADIAL_SLICES|*4.

\subsection{Change the maximum frequency of the simulation}
\verb|MESHER/inparam_mesh|, parameter \verb|DOMINANT_PERIOD|: \\
As a rule of thumb: Simulations with \verb|DOMINANT_PERIOD|>10s can be run with 2 or 4
CPUs on a modern workstation and cost around 1 CPUh.

\subsection{Calculate the response to a full moment tensor solution}
\verb|SOLVER/inparam_basic|, change parameter \verb|SIMULATION_TYPE| to \verb|moment|:\\
The moment tensor, depth and location of the source must be set in the file
\verb|CMTSOLUTION|. The submit.csh script starts four separate runs in parallel, the
postprocessing script sums the results to get correct seismograms.

\subsection{Change source time function}
\verb|SOLVER/inparam_advanced|, change parameter \verb|SOURCE_FUNCTION||:\\
AxiSEM generally calculates displacement seismograms. The source time function selected here is equivalent to the moment function $m(t)$. Note that to calculate seismograms similar to those of an earthquake (i.e. with a persistent displacement at the source), the setting \verb|errorf| has to be used. Note that that is different to the literature that usually defines the source time function as the moment rate function $\dot{m}(t)$. However, it is consistent with SpecFEM.\\
To create seismograms with a flat, zero-phased source spectrum, use the setting \verb|dirac_0|. Seismograms calculated with this setting can be convolved with other source functions in AxiSEM postprocessing or manually with a program of your choice.\\
However, note that seismograms or wavefields calculated with a \verb|dirac_0| STF will contain numerical noise. Therefore, visualization and wavefield plotting should be done using \verb|errorf| or one of the Gaussian STFs.

\subsection{Change seismogram length or sampling rate}
\verb|SOLVER/inparam_basic|, change parameter \verb|SEISMOGRAM_LENGTH|:\\
Default value is 1800~s, although the exact length is rounded to the next multiple of the
simulation time step. There is no maximum limit, AxiSEM has been run for 400000s (5 days)
to compare amplitude spectra with a normal modes summation (Nissen-Meyer et al, 2014).\\
\verb|SOLVER/inparam_advanced|, change parameter \verb|SAMPLING_RATE|:\\
By default, the sampling rate is set to the time step length of the simulation. We
strongly recommend to leave it as such to avoid aliasing. The resampling can better be
done with \textit{ObsPy} or another tool that supports filtering.

\subsection{Include lateral heterogeneities (2.5D simulation)}
\verb|SOLVER/inparam_basic|, change parameter \verb|LAT_HETEROGENEITY| to true:\\
The actual heterogeneity model is set in \verb|SOLVER/inparam_hetero|. 
This functionality is currently under development and not documented. See the example files \verb|SOLVER/inparam_hetero.TEMPLATE| for an idea of what to do.

\newpage
\section{References}

\noindent \textbf{Directly dealing with this code:}\vspace*{0.2cm}\\
When using this code, please cite one or more of these publications. \vspace*{0.2cm}\\

(1) Tarje Nissen-Meyer, M. van Driel, S. C. Stähler, K. Hosseini, S. Hempel, L. Auer, 
A. Colombi, and A. Fournier (2014), \textit{AxiSEM: broadband 3-D seismic wavefields
in axisymmetric media},  Solid Earth, 5, 425-445. \\
doi:10.5194/se-5-425-2014\\

(2) Tarje Nissen-Meyer, F. A. Dahlen, A. Fournier (2007),
\textit{Spherical-earth Fr\'{e}chet sensitivity kernels},        
Geophysical Journal International 168(3),1051-1066. \\
doi:10.1111/j.1365-246X.2006.03123.x                \\

(3) Tarje Nissen-Meyer, A. Fournier, F. A. Dahlen (2007), 
\textit{A two-dimensional spectral-element method for
spherical-earth seismograms-I. Moment-tensor source}, 
Geophysical Journal International 168(3), 1067-1092. \\
doi:10.1111/j.1365-246X.2006.03121.x                 \\

(4) Tarje Nissen-Meyer, A. Fournier, F. A. Dahlen (2008),  
\textit{A two-dimensional spectral-element method for   
spherical-earth seismograms - II. Waves in solid-fluid media},
Geophysical Journal International, 174(3), 873-888.\\
doi:10.1111/j.1365-246X.2008.03813.x\\

(5) Tarje Nissen-Meyer (2007),
\textit{Full-wave seismic sensitivity in a spherical Earth},
Ph.D. thesis, Princeton University
(This includes refs (2)-(4) and more details.)\\
% 
% (5) Jean-Paul Ampuero, Tarje Nissen-Meyer (2011),
% \textit{High-order conservative time schemes in spectral-element methods 
% for seismic wave propagation.}, To be submitted to Geophys. J. Int.\\

%\noindent \textbf{NOTE:} Most important sections of these papers are in the Theory section below.\\

\noindent \textbf{Other references:}\vspace*{0.2cm}

(6) Deville, M. O., Fischer, P. F., Mund, E. H. (2002), 
\textit{High-Order Methods for Incompressible Fluid Flow}, 
Vol. 2, Cambridge monographs on Sppl. \& Comp. Math., Cambridge University Press.\\

(7) Tufo, H. M., Fischer, P. F. (2001), \textit{Fast Parallel Direct Solvers For Coarse
Grid Problems}, 61, 151-177, J. Par. and Dist. Comput.\\

(8) Bernardi, C., Dauge, M., Maday, Y. (1999), \textit{Spectral Methods for Axisymmetric
Domains}, Vol. 3, Series in Appl. Math., Gauthier-Villars, Paris.\\

(9) Chaljub, E. (2000), \textit{Mod{\'{e}}lisation num{\'{e}}rique de la 
propagation d'ondes sismiques en g{\'{e}}om{\'{e}}trie sph{\'{e}}rique:
Application {\`{a}} la sismologie globale}, 
Ph.D. thesis, Universit{\'{e}} de Paris 7.\\

(10) Komatitsch D., Tromp, J. (2002), \textit{Spectral-element simulations of
global seismic wave propagation---I. Validation},
149, 390-412, Geophys. J. Int.

\end{document}

### 2015-08-10 utils/EXTERNAL_CODES_coupled_with_SPECFEM3D/AxiSEM_for_SPECFEM3D/AxiSEM_modif_for_coupling_with_specfem/MANUAL/manual_axisem1.2.tex

--- 2015-08-10 utils/EXTERNAL_CODES_coupled_with_SPECFEM3D/AxiSEM_for_SPECFEM3D/AxiSEM_modif_for_coupling_with_specfem/MANUAL/manual_axisem1.1.tex
@@
 % e.g. 1.0
-\raggedleft \huge \fontfamily{\sfdefault}\selectfont Version {1.1}\\}
+\raggedleft \huge \fontfamily{\sfdefault}\selectfont Version {1.2}\\}
 
@@
                 \title{}
-                \LARGE{ \textbf{\sc AxiSEMv1.1 Manual}}
+                \LARGE{ \textbf{\sc AxiSEMv1.2 Manual}}
                 \vspace*{0.6cm}\\
@@
 \subsubsection{NetCDF}
+\label{sssec:netcdf}
 
-AxiSEM allows to output larger datasets, especially wavefields in the NetCDF format. The
-current version also has full support for binary dumps, but the development will move
-towards NetCDF containers.\\
-Unfortunately, the installation of the NetCDF libraries may still be cumbersome; below we list some limited support for installing NetCDF..
+AxiSEM allows to output larger datasets, especially wavefields in the NetCDF format. \\
+Unfortunately, the installation of the NetCDF libraries may still be cumbersome; below we
+list some limited support for installing NetCDF..
 
@@
     \hspace{5mm}
-    \includegraphics[height=50mm]{solver_totalcpu.pdf}
+    %\includegraphics[height=50mm]{solver_totalcpu.pdf}
+    \includegraphics[height=50mm]{db_generation_cost.pdf}
 \end{center}
@@
 On the right, we depict the computational cost
-associated with the solver to compute seismograms of 30 min length. The relation between
-seismic period and CPU-hours is an approximate proxy to estimate how many cores and
-wall-clock time is optimal for your infrastructure. Scaling of CPU hours is approximately 
-to the third power of the maximum frequency.
+associated with the solver to compute seismograms of one hour length for Earth and Mars
+and for the second order Newmark as well as the fourth order symplectic time scheme. The
+relation between seismic period and CPU-hours is an approximate proxy to estimate how many
+cores and wall-clock time is optimal for your infrastructure. Scaling of CPU hours is
+approximately to the third power of the maximum frequency. As a rule of thumb, using the
+symplectic scheme is advisable when propagating waves for more than 100 wavelengths.
 
@@
 \verb|SOLVER/inparam_source|, \verb|SOLVER/CMTSOLUTION| and \verb|SOLVER/inparam_basic|.
-AxiSEM has two principal modes, which are selected by the value \verb|SIMULATION_TYPE| in
+AxiSEM has three principal modes, which are selected by the value \verb|SIMULATION_TYPE| in
 \verb|SOLVER/inparam_basic|.
@@
       \url{http://www.globalcmt.org/CMTsearch.html}
+
+\item \verb|SIMULATION_TYPE force|:\\
+      The submit.csh script starts two separate simulations for the basic types $P_r$ and
+      $P_\theta$. This mode is used to compute the databases for instaseis. Note, that the
+      postprocessing does not handle these runs yet appropriately.
 \end{enumerate}
@@
 \verb|SOLVER/inparam_advanced|, change parameter \verb|SOURCE_FUNCTION||:\\
-AxiSEM generally calculates displacement seismograms.
-The source time function selected here is equivalent to the moment function $m(t)$.
-Note that to calculate seismograms similar to those of an earthquake (i.e.
-with a persistent displacement at the source), the setting \verb|errorf| has to be used.
-Note that that is different to the literature that usually defines the source time function as the moment rate function $\dot{m}(t)$.
+AxiSEM generally calculates displacement seismograms. The source time function selected
+here is equivalent to the moment function $m(t)$. Note that to calculate seismograms
+similar to those of an earthquake (i.e. with a persistent displacement at the source), the
+setting \verb|errorf| has to be used. Note that that is different to the literature that
+usually defines the source time function as the moment rate function $\dot{m}(t)$.
 However, it is consistent with SpecFEM.\\
-To create seismograms with a flat, zero-phased source spectrum, use the setting \verb|dirac_0|.
-Seismograms calculated with this setting can be convolved with other source functions in AxiSEM postprocessing or manually with a program of your choice.\\
-However, note that seismograms or wavefields calculated with a \verb|dirac_0| STF will contain numerical noise.
-Therefore, visualization and wavefield plotting should be done using \verb|errorf| or one of the Gaussian STFs.
+To create seismograms with a flat, zero-phased source spectrum, use the setting
+\verb|dirac_0|. Seismograms calculated with this setting can be convolved with other
+source functions in AxiSEM postprocessing or manually with a program of your choice.\\
+However, note that seismograms or wavefields calculated with a \verb|dirac_0| STF will
+contain numerical noise. Therefore, visualization and wavefield plotting should be done
+using \verb|errorf| or one of the Gaussian STFs.
 
@@
 
+\subsection{Output wavefields in netcdf format needed for \textit{Instaseis}}
+
+\begin{itemize}
+    \item Compile with netcdf (see section~\ref{sssec:netcdf} for hints on the
+    installation): in \verb|make_axisem.macros| set \verb|USE_NETCDF| to \verb|true|
+    and if netcdf is not in your \verb|PATH|, also set \verb|NETCDF_PATH| to the location
+    of your netcdf installation. If you have netcdf compiled with MPI and run on a
+    supercomputer, using \verb|USE_PARALLEL_NETCDF true| might be beneficial for short
+    period runs.
+    \item In \verb|SOLVER/inparam_advanced| set \verb|USE_NETCDF| to \verb|true| to
+    use netcdf output in the solver
+    \item In \verb|SOLVER/inparam_advanced| set \verb|DATA_DIR| to some location where you
+    have enough free space and in case you use parallel IO this should be on a parallel
+    file system
+    \item further settings in \verb|SOLVER/inparam_advanced| to generate a full database,
+    that is all epicentral distances and source down to 700km and using the fourth order
+    spatial scheme:
+
+\begin{verbatim}
+SOURCE_FUNCTION     gauss_0
+
+KERNEL_WAVEFIELDS   true
+
+KERNEL_DUMPTYPE     displ_only
+KERNEL_SPP          4
+KERNEL_SOURCE       igno
+
+KERNEL_COLAT_MIN    0.
+KERNEL_COLAT_MAX    180.
+
+KERNEL_RMIN         5671.
+KERNEL_RMAX         7000.
+\end{verbatim}
+
+    \item in \verb|SOLVER/inparam_basic| set \verb|SIMULATION_TYPE| to \verb|force| to
+    simulate both the vertical and the horizontal database. To generate just one of the
+    two, set it to \verb|single| and select either \verb|thetaforce| or \verb|vertforce|
+    in \\\verb|SOLVER/inparam_source|
+
+    \item once the AxiSEM runs are finished, go to the rundirectory and run
+    \verb|./fieltransform.sh| (in case of single runs just run \verb|./fieltransform|
+    directly). This rechunks the wavefields in the netcdf file to optimize them for
+    reading as time series instead of snapshots.
+
+\end{itemize}
+
 \subsection{Include lateral heterogeneities (2.5D simulation)}
@@
 The actual heterogeneity model is set in \verb|SOLVER/inparam_hetero|. 
-This functionality is currently under development and not documented. See the example files \verb|SOLVER/inparam_hetero.TEMPLATE| for an idea of what to do.
+This functionality is currently under development and not documented. See the example
+files \verb|SOLVER/inparam_hetero.TEMPLATE| for an idea of what to do.
 
@@
 (This includes refs (2)-(4) and more details.)\\
+
+(6) Martin van Driel and Tarje Nissen-Meyer (2014),
+\textit{Seismic Wave Propagation in Fully Anisotropic Axisymmetric Media},
+Geophysical Journal International 199 (2), 880–893.\\
+doi:10.1093/gji/ggu269.\\
+
+(7) Martin van Driel and Tarje Nissen-Meyer (2014),
+\textit{Optimized Viscoelastic Wave Propagation for Weakly Dissipative Media},
+Geophysical Journal International 199 (2) 1078–1093.\\
+doi:10.1093/gji/ggu314.\\
+
 %
@@
 
-(6) Deville, M. O., Fischer, P. F., Mund, E. H. (2002), 
+(8) Martin van Driel, Lion Krischer, Simon Stähler, Kasra Hosseini, and Tarje Nissen-Meyer (2015),
+\textit{Instaseis: Instant Global Broadband Seismograms Based on a Waveform Database},
+To be submitted to Solid Earth\\
+
+(9) Deville, M. O., Fischer, P. F., Mund, E. H. (2002), 
 \textit{High-Order Methods for Incompressible Fluid Flow}, 
@@
 
-(7) Tufo, H. M., Fischer, P. F. (2001), \textit{Fast Parallel Direct Solvers For Coarse
+(10) Tufo, H. M., Fischer, P. F. (2001), \textit{Fast Parallel Direct Solvers For Coarse
 Grid Problems}, 61, 151-177, J. Par. and Dist. Comput.\\
 
-(8) Bernardi, C., Dauge, M., Maday, Y. (1999), \textit{Spectral Methods for Axisymmetric
+(11) Bernardi, C., Dauge, M., Maday, Y. (1999), \textit{Spectral Methods for Axisymmetric
 Domains}, Vol. 3, Series in Appl. Math., Gauthier-Villars, Paris.\\
 
-(9) Chaljub, E. (2000), \textit{Mod{\'{e}}lisation num{\'{e}}rique de la 
+(12) Chaljub, E. (2000), \textit{Mod{\'{e}}lisation num{\'{e}}rique de la 
 propagation d'ondes sismiques en g{\'{e}}om{\'{e}}trie sph{\'{e}}rique:
@@
 
-(10) Komatitsch D., Tromp, J. (2002), \textit{Spectral-element simulations of
+(13) Komatitsch D., Tromp, J. (2002), \textit{Spectral-element simulations of
 global seismic wave propagation---I. Validation},

### 2015-08-10 utils/EXTERNAL_CODES_coupled_with_SPECFEM3D/AxiSEM_for_SPECFEM3D/AxiSEM_modif_for_coupling_with_specfem/MESHER/gllmeshgen.f90
Clean NEW Axisem DIR

@@ -23,9 +23,9 @@
 module gllmeshgen
 
   use data_mesh
-  use analytic_mapping 
+  use analytic_mapping
   use data_spec
-  use data_gllmesh 
+  use data_gllmesh
   use splib
 
   implicit none
@@ -77,7 +77,7 @@ subroutine create_gllmesh
   do ishp = 0, npol
      call hn_jprime(eta, ishp, npol, df)
      G2(ishp,:) = df
-  end do
+  enddo
   G2T = transpose(G2)
 
   ! Derivative in s-direction: \partial_\xi (\bar{l}_i(\xi_p))
@@ -85,7 +85,7 @@ subroutine create_gllmesh
   do ishp = 0, npol
      call lag_interp_deriv_wgl(df,xi_k,ishp,npol)
      G1(ishp,:) = df
-  end do
+  enddo
   G1T = transpose(G1)
 
   ! Axial vector
@@ -93,7 +93,7 @@ subroutine create_gllmesh
 
   !$omp parallel do shared(sgll, zgll) private(crd_nodes, jpol, ipol, stest)
   do iel = 1, neltot
-  
+
      ! define dummy coordinate arrays
      crd_nodes(:,:) = 0.d0
      crd_nodes(1,1) = sg(lnodesg(1,iel))
@@ -104,29 +104,29 @@ subroutine create_gllmesh
      crd_nodes(5,2) = zg(lnodesg(3,iel))
      crd_nodes(7,1) = sg(lnodesg(4,iel))
      crd_nodes(7,2) = zg(lnodesg(4,iel))
-     
+
      crd_nodes(2,:) = .5d0 * ( crd_nodes(1,:) + crd_nodes(3,:) )  ! midpoints are necessary
      crd_nodes(4,:) = .5d0 * ( crd_nodes(3,:) + crd_nodes(5,:) )  ! for subparametric mapping
      crd_nodes(6,:) = .5d0 * ( crd_nodes(5,:) + crd_nodes(7,:) )  ! (Serendipity elements).
-     crd_nodes(8,:) = .5d0 * ( crd_nodes(7,:) + crd_nodes(1,:) )  !  
+     crd_nodes(8,:) = .5d0 * ( crd_nodes(7,:) + crd_nodes(1,:) )  !
 
      stest = minval(sg(lnodesg(1:4,iel)))
-     if ( stest < smallval_dble ) then 
+     if ( stest < smallval_dble ) then
         do jpol = 0, npol
            do ipol = 0, npol
[…]

### 2015-08-10 utils/EXTERNAL_CODES_coupled_with_SPECFEM3D/AxiSEM_for_SPECFEM3D/AxiSEM_modif_for_coupling_with_specfem/MESHER/main.f90

@@ -51,11 +51,11 @@ program gllmesh
   ! 3) bkgrdmodel & discontinuity radii
   call read_params ! input
   !- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
-  
+
   write(6,*)'MAIN: creating subregions/discontinuity model..........'; call flush(6)
   call create_subregions ! discont_meshing
 
-  southern = .true. 
+  southern = .true.
 
   write(6,*)'MAIN: generating skeleton..............................'; call flush(6)
   iclock02 = tick()
@@ -86,7 +86,7 @@ program gllmesh
   call def_fluid_regions ! mesh_info
   call def_solid_regions ! mesh_info
   call extract_fluid_solid_submeshes ! gllmeshgen
-  
+
   write(6,*)'MAIN: glob-slob/flob numbering.........................'; call flush(6)
   iclock08 = tick()
   if (have_fluid) &
@@ -102,7 +102,7 @@ program gllmesh
   ! Parallelization
   write(6,*)'MAIN: domain decomposition.............................'; call flush(6)
   call create_domain_decomposition !parallelization
- 
+
   write(6,*)'MAIN: creating parallel database.......................'; call flush(6)
   iclock11 = tick()
   call create_pdb ! pdb
@@ -110,9 +110,9 @@ program gllmesh
 
   ! clean up
   call empty_data_mesh
-  
+
   call end_clock ! clocks
-  
+
   write(6,*)''
   write(6,*)'....DONE WITH MESHER !'
 

### 2015-08-10 utils/EXTERNAL_CODES_coupled_with_SPECFEM3D/AxiSEM_for_SPECFEM3D/AxiSEM_modif_for_coupling_with_specfem/MANUAL/manual_axisem1.2.tex
Clean NEW Axisem DIR ==> 2

@@
 
-(1) Tarje Nissen-Meyer, M. van Driel, S. C. Stähler, K. Hosseini, S. Hempel, L. Auer, 
+(1) Tarje Nissen-Meyer, M. van Driel, S. C. Stahler, K. Hosseini, S. Hempel, L. Auer,
 A. Colombi, and A. Fournier (2014), \textit{AxiSEM: broadband 3-D seismic wavefields
@@
 \textit{Seismic Wave Propagation in Fully Anisotropic Axisymmetric Media},
-Geophysical Journal International 199 (2), 880–893.\\
+Geophysical Journal International 199 (2), 880-893.\\
 doi:10.1093/gji/ggu269.\\
@@
 \textit{Optimized Viscoelastic Wave Propagation for Weakly Dissipative Media},
-Geophysical Journal International 199 (2) 1078–1093.\\
+Geophysical Journal International 199 (2) 1078-1093.\\
 doi:10.1093/gji/ggu314.\\
@@
 
-(8) Martin van Driel, Lion Krischer, Simon Stähler, Kasra Hosseini, and Tarje Nissen-Meyer (2015),
+(8) Martin van Driel, Lion Krischer, Simon Stahler, Kasra Hosseini, and Tarje Nissen-Meyer (2015),
 \textit{Instaseis: Instant Global Broadband Seismograms Based on a Waveform Database},

### 2015-08-10 utils/EXTERNAL_CODES_coupled_with_SPECFEM3D/AxiSEM_for_SPECFEM3D/AxiSEM_modif_for_coupling_with_specfem/MESHER/makemake.pl

@@ -1,7 +1,7 @@
 #! /usr/bin/perl
 #
 #    Copyright 2013, Tarje Nissen-Meyer, Alexandre Fournier, Martin van Driel
-#                    Simon Stähler, Kasra Hosseini, Stefanie Hempel
+#                    Simon Stahler, Kasra Hosseini, Stefanie Hempel
 #
 #    This file is part of AxiSEM.
 #    It is distributed from the webpage <http://www.axisem.info>
@@ -21,13 +21,13 @@
 #
 #    Generate a Makefile from the sources in the current directory.  The source
 #    files may be in either C, FORTRAN 77, Fortran 90 or some combination of
-#    these languages.  
+#    these languages.
 #
 #    Original version written by Michael Wester <wester@math.unm.edu> February 16, 1995
 #    Cotopaxi (Consulting), Albuquerque, New Mexico
 #
-#    Modified by Martin van Driel, ETH Zürich and Simon Stähler, 
-#    LMU München to fit the needs of Axisem. The compiler version is
+#    Modified by Martin van Driel, ETH Zurich and Simon Stahler,
+#    LMU Munchen to fit the needs of Axisem. The compiler version is
 #    now set in the file ../make_axisem.macros
 
 open(MAKEFILE, "> Makefile");
@@ -257,7 +257,7 @@ sub MakeDependsf90 {
          #
          }
       }
-   
+
 
 print "\nCheck Makefile to make sure you're happy with it.\n\n";
 system("perl -ni -e 'print unless /^kdtree2.o/' Makefile ");

### 2015-08-10 utils/EXTERNAL_CODES_coupled_with_SPECFEM3D/AxiSEM_for_SPECFEM3D/AxiSEM_modif_for_coupling_with_specfem/SOLVER/UTILS/fieldtransform_force.sh

@@ -1,4 +1,4 @@
-#!/bin/bash 
+#!/bin/bash
 set -e
 cd PX
 echo "Transforming PX"

### 2015-08-10 utils/EXTERNAL_CODES_coupled_with_SPECFEM3D/AxiSEM_for_SPECFEM3D/AxiSEM_modif_for_coupling_with_specfem/MANUAL/virtual_box_tutorial.tex
Clean NEW Axisem DIR

@@
 * Tarje:
-Hi Kasra, on a quick look, this seems excellent, but from my experience it’s better to have a very easy and quick first run of the tasks, e.g.
+Hi Kasra, on a quick look, this seems excellent, but from my experience it's better to have a very easy and quick first run of the tasks, e.g.
 on ONE page with all the important calls and major points we wish to achieve...
 and delegate images and details to an appendix.
-So, let’s keep all this in but reshuffle it to have a concise early section.
+So, let's keep all this in but reshuffle it to have a concise early section.
 Tarje
@@
 
-(TAS÷
+(TAS?
 1) Load data from one of the 5 events (Kasra) with Obspy, plot cross sections

### 2015-08-11 utils/EXTERNAL_CODES_coupled_with_SPECFEM3D/AxiSEM_for_SPECFEM3D/Manual_coupling_AxiSEM_Specfem3D.tex
Update README files and Manual AxiSEM + clean coupling UTILS

@@
 
+\medskip
+
+\noindent In Axisem directory, there are all the parameter files, scripts, stations, etc., to put in Axisem mesher and solver directories, for each example, here:
+
+\noindent {\small 'utils/EXTERNAL\_CODES\_coupled\_with\_SPECFEM3D/AxiSEM\_for\_SPECFEM3D/Parfiles\_and\_scripts\_for\_examples'}
+
+\noindent Cf 'README' in this directory for more details.
+
 \bigskip
@@
 
-\item[\textbullet] ATTENUATION         false
-\item[\textbullet] SAVE\_SNAPSHOTS     true
+\item[\textbullet] MESHNAME            ~~ NAME\_OF\_DESTINATION\_MESH\_DIR
+\item[\textbullet] ATTENUATION         ~~ false
+\item[\textbullet] SAVE\_SNAPSHOTS     ~~ true
 
@@
 
-\noindent {\color{red} NOTE:} If you already know the value of the begin time and if it is already configured in 'reformat.par', you can launch the compilation of expand\_2D\_3D and reformat, AND the execution of this two programs with the single script :
+\noindent {\color{red} NOTE:} If you already know the value of the begin time and if it is already configured in 'reformat.par', you can launch the compilation of expand\_2D\_3D and reformat, AND the execution of this two programs with the single script (AND uncomment the LAST LINE of the script 'sub\_called\_batch\_for\_expand2D3D\_Curie\_CD.sh' BEFORE that) :
 
@@
 \noindent !! CAUTION : don't forget "NAME\_OF\_RESULTS\_DIR" !!
+
+\medskip
+
+\noindent !! CAUTION : don't forget to UNCOMMENT the last line
+
+\noindent (\#\#\#\#ccc\_msub -q standard ../sub\_called\_batch\_for\_reformat\_Curie\_CD.sh)
+
+\noindent of 'sub\_called\_batch\_for\_expand2D3D\_Curie\_CD.sh', which will launch the reformat script at the end !!
 
@@
 
-\section{STEP ? : Set up scripts}
+%%\section{STEP ? : Set up scripts}
 

### 2015-09-09 doc/USER_MANUAL/06_fault_sources.tex
Modify the fault solver part of the user manual.

@@
 for dynamic and kinematic faults. They are $\mathtt{Par\_file\_faults}$,
-$\mathtt{FAULT\_STATIONS}$ and $\mathtt{input\_file.txt}$ (optional).
+$\mathtt{FAULT\_STATIONS}$ and $\mathtt{input\_file.txt}$ or $\mathtt{rsf\_hete\_input\_file.txt}$(optional).
 If the former does not exist, the code assumes that there are no faults.\\
@@
 
+\&\textbf{RUPTURE\_SWITCHES}  /
+RATE\_AND\_STATE ,TPV16 ,HETE\_RSF ,TPV10X
+
 \&\textbf{BEGIN\_FAULT} /
+
+\&\textbf{STRESS\_TENSOR} Sigma = $\sigma_{xx}$,$\sigma_{yy}$,$\sigma_{zz}$,$\sigma_{xy}$,$\sigma_{xz}$,$\sigma_{yz}$ /
 
@@
 
-followed by (nmus+nmud+ndc) \&\textbf{DIST2D} blocks\\
+\&\textbf{RSF} V0,f0,a,b,L,V\_init,theta\_init,C,StateLaw {[} nV0,nf0,na,nb,nL,nV\_init,ntheta\_init,nC {]} /
+
+followed by (nV0+nf0+na+nb+nL+nV\_init+ntheta\_init+nC) \&\textbf{DIST2D} blocks\\
 
 \begin{lyxlist}{00.00.0000}
+\item [{\&\textbf{RUPTURE\_SWITCHES}}] input block sets some switches in the simulation process.
+  \begin{description}
+    \item [{RATE\_AND\_STATE}] 
+         = .TRUE. use rate and state friction,=.FALSE. use slip weakening friction
+    \item [{TPV16}]
+         =.TRUE turn on heterogeneity fault property input for slip weakening friction from DATA\/input\_file.txt,
+         =.FALSE. turn off such feature 
+    \item [{HETE\_RSF}]
+         =.TRUE. turn on heterogeneity fault property input for rate and state friction 
+              using input from DATA\/rsf\_hete\_input\_filr.txt, 
+	 =.FALSE. turn off such feature
+    \item[{TPV10X}]
+         =.TRUE. turn on some ad hoc features for TPV101-104 simulations, a rate strengthening 
+	 layer surrounding the rate weakening region. 
+	 =.FALSE. turn off such feature
+   \end{description}
+
+\item [{\&\textbf{STRESS\_TENSOR}}] input block sets the initial fault stresses 
+  by projecting a uniform regional stress field onto the fault plane.So that the 
+  tractions $\tau = \sigma : n$ while $n$ is the local norm of the fault plane.
+  Sigma =  $\sigma_{xx}$,$\sigma_{yy}$,$\sigma_{zz}$,$\sigma_{xy}$,$\sigma_{xz}$,$\sigma_{yz}$ 
+  (in Pa)
+
 \item [{\&\textbf{INIT\_STRESS}}] input block sets the initial fault stresses
@@
 \end{description}
+
+\item [{\&\textbf{RSF}}] input block sets the rate and state friction parameters
+of the fault:
+
+\begin{description}
+  \item []
+    We refer to the well known rate and state friction formula:
+    $$\mu = f_0 + a* log(\frac{v}{v_0}) + b* log(\frac{v_0}{\theta L}))  $$
+\item [{V0}] = constant background value of $V_0$
+\item [{f0}] = constant background value of $f_0$
+\item [{a}] = constant background value of $a$
+\item [{b}] = constant background value of $b$
+\item [{L}] = constant background value of $L$
+\item [{V\_init}] = constant background value of along strike right lateral slip rate at time 0
+\item [{theta\_init}] = constant background value of state variable $\theta$ at time 0
+\item [{C}] = constant background value of cohesion (in Pa)
+\item [{StateLaw}] = 1 or 2. 1 for aging law , 2 for slip law with strong dynamic weakening.
+\item [{n**}] = number of heterogeneous items for quantity **
+\end{description}
+
 \item [{\&\textbf{DIST2D}}] input blocks modify (overwrite) the value of
@@
 \begin{lyxlist}{00.00.0000}
-\item [{\textbf{DATA/input\_file.txt}}] Heterogeneous stresses and friction
+\item [{\textbf{DATA/input\_file.txt}}] Heterogeneous stresses and friction for 
+linear slip weakening friction
 parameters are documented in page 10 of
@@
 
-To activate this feature, in fault\_solver\_dynamic.f90 set \textbf{TPV16}
-= .true. then re-compile the code.
+To activate this feature, in Par\_file\_faults name list \&\textbf{RUPTURE\_SWITCHES}, set TPV16=.TRUE..
 
+\end{lyxlist}
+\begin{lyxlist}{00.00.0000}
+\item [{\textbf{DATA/rsf\_hete\_input\_file.txt}}]  Heterogeneous stresses and friction input for 
+  rate and state friction. 
+To activate this feature, in Par\_file\_faults name list \&\textbf{RUPTURE\_SWITCHES}, set RSF\_HETE=.TRUE..
+
+  The format of  {\textbf{DATA/rsf\_hete\_input\_file.txt}} is as such, in the first line there are four integers
+
+  that are sequentially documenting the number of divisions along strike, number of divisions along
+
+  dip, cell size along strike , cell size along dip. Then the following N(N=NumberOfDivisionsAlongStrike
+
+  * NumberOfDivisionsAlongDip) lines will document the stress and friction properties on the grid.
+
+  There are a total of 13 columns.
+  \begin{enumerate}
+  \item{Column1} = Along strike distance(m)
+
+  \item{Column2} = Along dip distances(m) The distance should be negative for points at depth
+
+  \item{Column3} = Normal stress(Pa)
+
+  \item{Column4} = Horizontal right-lateral shear stress(Pa)
+
+  \item{Column5} = Vertical up-dip shear stress(Pa)
+
+  \item{Column6} = $V_0$ value in RSF(m/s)
+
+  \item{Column7} = $f_0$ value in RSF
+
+  \item{Column8} = a value in RSF
+
+  \item{Column9} = b value in RSF
+
+  \item{Column10} = L value in RSF
+
+  \item{Column11} = Initial slip velocity(m/s)
+
+  \item{Column12} = State variable $\theta$
+
+  \item{Column13} = Cohesion(Pa)  
+  \end{enumerate}
+  Note that the input grid file do not have to coincide with the mesh. For each fault node on the mesh, they will use the value from the nearest node on the input grid.
 \end{lyxlist}

### 2015-09-10 doc/USER_MANUAL/06_fault_sources.tex

@@
   \begin{enumerate}
-  \item{Column1} = Along strike distance(m)
+  \item{Column1} = Along strike distance
 
-  \item{Column2} = Along dip distances(m) The distance should be negative for points at depth
+  \item{Column2} = Along dip distances
 
-  \item{Column3} = Normal stress(Pa)
+  \item{Column3} = Normal stress(MPa)
 
-  \item{Column4} = Horizontal right-lateral shear stress(Pa)
+  \item{Column4} = Horizontal right-lateral shear stress(MPa)
 
-  \item{Column5} = Vertical up-dip shear stress(Pa)
+  \item{Column5} = Vertical up-dip shear stress(MPa)
 
-  \item{Column6} = $V_0$ value in RSF(m/s)
+  \item{Column6} = $V_0$ value in RSF
 
@@
 
-  \item{Column11} = Initial slip velocity(m/s)
+  \item{Column11} = Initial slip velocity
 

### 2015-09-10 utils/EXTERNAL_CODES_coupled_with_SPECFEM3D/AxiSEM_for_SPECFEM3D/Manual_coupling_AxiSEM_Specfem3D.tex
Buildbot ran the code beautifier.
Change Axisem coupling Manual

@@
 90. 0.                      # source position (lat lon)
-0.  60.                     # chunk center (lat lon)
+0.  60.  0.                 # chunk center (lat lon azi_rot)
 1                           # number of axisem simus depends on moment tensor used

### 2015-10-12 doc/USER_MANUAL/03_mesh_generation.tex

@@
 \item [{\texttt{NDOUBLINGS}}] The number of horizontal doubling layers.
-Must be set to \texttt{1} or \texttt{2} if \texttt{USE\_REGULAR\_MESH}
-is set to \texttt{.true.}.
+Must be set at least to \texttt{1} if \texttt{USE\_REGULAR\_MESH} is set to \texttt{.true.}.
+Multiple mesh doublings can be chosen, for which each an \texttt{NZ\_DOUBLING\_**} entry must be given.
+By default, we only provide two possible entries in the \texttt{Mesh\_Par\_file}.
+For higher numbers of doubling layers, additional entries must be added.
 \item [{\texttt{NZ\_DOUBLING\_1}}] The position of the first doubling layer
@@
 layer (only interpreted if \texttt{USE\_REGULAR\_MESH} is set to \texttt{.true.}
-and if \texttt{NDOUBLINGS} is set to \texttt{2}).
+and if \texttt{NDOUBLINGS} is set to \texttt{2}). Doubling layers must be at least \texttt{2} layers apart.
+The layer count starts from the bottom layer.
+More entries must be listed by the user if \texttt{NDOUBLINGS} is larger than \texttt{2}.
 \item [{\texttt{CREATE\_ABAQUS\_FILES}}] Set this flag to \texttt{.true.}

### 2015-10-12 doc/USER_MANUAL/08_doing_tomography.tex
updates user manual

@@
 
-UNDER CONSTRUCTION (April 2015). Will be added around June 2015.
+UNDER CONSTRUCTION (July 2015). New content will be added soon.
 
@@
 but also because these seismic velocities are easier to interpret, tomographic models are usually described based on these
-two parameters. With this new parametrisation, the perturbation of the misfit function may be written as:
+two parameters. With this new parametrization, the perturbation of the misfit function may be written as:
 \begin{eqnarray}
@@
 the amount of disk storage required would currently be too large.
+
 However let us note again that this situation will change in the future.
@@
 
+\section{Tomographic tools}
+
+Besides the ability to compute adjoint sensitivity kernels, the SPECFEM3D package provides a number of tools useful for post-processing kernels, gradients and conducting tomographic model updates.
+The tomographic and post-processing tools can be compiled by typing in the main directory:
+\begin{verbatim}
+make tomography
+make postprocess
+\end{verbatim}
+You can modify specific settings affecting, e.g., the (isotropic) parameterization or directory setup structure in file \texttt{setup/constants\_tomography.h}.
+The compiled binaries can be found in directory \texttt{bin/}.\\
+
+\noindent
+The iterative inversion workflow may comprise three main steps for which tools are provided:
+\begin{enumerate}
+\item Summing event kernels, e.g. $K_{\alpha,\beta,\rho}$, to build the misfit gradient $\boldg(\boldm)$
+\item Smoothing and post-processing of the gradient $\boldg(\boldm)$
+\item Updating the model $\boldm_{i+1} = \boldm_i + \delta\boldm$.
+\end{enumerate}
+
+Please note that the tomographic routines are very similar for both SPECFEM3D\_Cartesian and SPECFEM3D\_GLOBE versions.
+Some differences occur, e.g., when reading in mesh files and when the maximum of the gradient is taken for the update step length.
+Also note that currently only the binary database format is supported, no ADIOS file format support is implemented yet.
+In future, we plan to extend and merge these into the same set of tools within a common SPECFEM function library.\\
+
+\subsection{Summing kernels}
+
+For summing different event kernels, the executables \texttt{xsum\_kernels} and \texttt{xsum\_preconditioned\_kernels} can be used.
+The binaries sum up either a set of isotropic or transversely isotropic kernel files, depending on the setting in file \texttt{setup/constants\_tomography.h}.
+The following parameterizations can be chosen:
+\begin{enumerate}
+\item [-] isotropic kernels $K_{\alpha,\beta,\rho}$ or
+\item [-] isotropic bulk kernels $K_{c,\beta,\rho}$, or
+\item [-] transversely isotropic bulk kernels $K_{c,\beta_v,\beta_h,\eta}$.
+\end{enumerate}
+
+\noindent
+The tools for summing event kernels can be used in parallel, using the same number of processes (\texttt{NPROC}) as the forward/kernel simulations.
+The summation tools use the following input/output format (see Fig.~\ref{fig:tomo-dir-struct} for the default directory structure):
+\begin{itemize}
+\item [-]{\bf Input:}
+Input is provided by a list of event directories, given in file \texttt{kernels\_list.txt} (default setting for \texttt{KERNEL\_FILE\_LIST} in \texttt{constants\_tomography.h}).
+The file lists all event kernel directories which should be summed together.
+All event directories have to be located in directory \texttt{INPUT\_KERNELS/}, which can be setup as links to all individual event kernel directories.
+
+\item [-]{\bf Output:}
+The summed kernels will be stored in corresponding kernel files in output directory \texttt{OUTPUT\_SUM/}.\\
+\end{itemize}
+
+\begin{figure}[htp]
+\begin{centering}
+\includegraphics[width=4in]{figures/tomo_dir_struct.pdf}
+\par\end{centering}
+%
+\caption{
+Example directory structure when using tomographic tools for summation. Event directory names (red) can be chosen freely with corresponding name entries in file \texttt{kernels\_list.txt}}
+\label{fig:tomo-dir-struct}
+\end{figure}
+
+\begin{description}
+\item [Kernel summation:] As example with a total of 4 MPI processes
+\begin{verbatim}
+mpirun -np 4 ./bin/xsum_kernels
+\end{verbatim}
+adds sensitivity kernels $K$ from different events together, i.e. outputs $\boldg$ where
+\begin{equation}
+\boldg = \Sum^{N}_i K^{i} \nonumber
+\end{equation}
+for events $i = 1,..,N$ as specified in file \texttt{kernel\_list.txt}.
+
+\item [Kernel Summation with Preconditioning:] As example with a total of 4 MPI processes
+\begin{verbatim}
+mpirun -np 4 ./bin/xsum_preconditioned_kernels
+\end{verbatim}
+adds sensitivity kernels $K$ together from different events and "preconditions" the sum by dividing with the sum of the corresponding approximate Hessians $\tilde{H}$, i.e. outputs $\boldg$ where
+\begin{equation}
+\boldg = \frac{1}{\Sum_i \tilde{H}^{i}} \; \Sum_i  K^{i} \nonumber
+\end{equation}
+In this case, the factor $\frac{1}{\Sum_i \tilde{H}^{i}}$ acts as preconditioner, approximating the inverse of the Hessian $\mathbf{H}^{-1}$.
+\end{description}
+
+\noindent
+{\bf Kernel names}: The summation tool will look for kernels with following names:
+\begin{itemize}
+\item [-] for an {\it isotropic} model parameterization, kernel names are:\\
+\begin{tabular}{ c c }
+\begin{minipage}{3in}
+\begin{verbatim}
+## Cartesian version
+proc***_alpha_kernel.bin
+proc***_beta_kernel.bin
+proc***_rho_kernel.bin
+\end{verbatim}
+\end{minipage}
+&
+\begin{minipage}{3in}
+\begin{verbatim}
+## GLOBE version
+proc***_reg1_alpha_kernel.bin
+proc***_reg1_beta_kernel.bin
+proc***_reg1_rho_kernel.bin
+\end{verbatim}
+\end{minipage}
+\end{tabular}\\
+
+\item [-] for an {\it isotropic bulk} model parameterization, kernel names are:\\
+\begin{tabular}{ c c }
+\begin{minipage}{3in}
+\begin{verbatim}
+## Cartesian version
+proc***_bulk_kernel.bin
+proc***_bulk_beta_kernel.bin
+proc***_rho_kernel.bin
+\end{verbatim}
+\end{minipage}
+&
+\begin{minipage}{3in}
+\begin{verbatim}
+## GLOBE version
+proc***_reg1_bulk_kernel.bin
+proc***_reg1_bulk_beta_kernel.bin
+proc***_reg1_rho_kernel.bin
+\end{verbatim}
+\end{minipage}
+\end{tabular}\\
+
+\item [-] for a {\it transversely isotropic} model parameterization, kernel names are:\\
+\begin{tabular}{ c c }
+\begin{minipage}{3in}
+\begin{verbatim}
+## Cartesian version
+proc***_bulk_c_kernel.bin
+proc***_bulk_betav_kernel.bin
+proc***_bulk_betah_kernel.bin
+proc***_eta_kernel.bin
+\end{verbatim}
+\end{minipage}
+&
+\begin{minipage}{3in}
+\begin{verbatim}
+## GLOBE version
+proc***_reg1_bulk_c_kernel.bin
+proc***_reg1_bulk_betav_kernel.bin
+proc***_reg1_bulk_betah_kernel.bin
+proc***_reg1_eta_kernel.bin
+\end{verbatim}
+\end{minipage}
+\end{tabular}\\
+
+Note that these event kernels are stored after a kernel simulation (\texttt{SIMULATION\_TYPE = 3}) in the \texttt{LOCAL\_PATH} directory, which by default points to directory \texttt{DATABASES\_MPI/}.
+Isotropic kernels will be created by default.
+To create transversely isotropic kernels, you set \texttt{ANISOTROPIC\_KL} and \texttt{SAVE\_TRANSVERSE\_KL} to \texttt{.true.} in the parameter file \texttt{DATA/Par\_file}.
+\\
+
+\item [- ] for preconditioning, the approximate Hessian kernels taken for summation are:\\
+\begin{tabular}{ c c }
+\begin{minipage}{3in}
+\begin{verbatim}
+## Cartesian version
+proc***_hess_kernel.bin
+\end{verbatim}
+\end{minipage}
+&
+\begin{minipage}{3in}
+\begin{verbatim}
+## GLOBE version
+proc***_reg1_hess_kernel.bin
+\end{verbatim}
+\end{minipage}
+\end{tabular}\\
+
+These Hessian kernels approximate the diagonal elements of the Hessian matrix and can be used as preconditioner in a gradient optimization scheme.
+To create these kernels, you have to set \texttt{APPROXIMATE\_HESS\_KL} to \texttt{.true.} in the parameter file \texttt{DATA/Par\_file}.\\
+
+\end{itemize}
+
+For SPECFEM3D\_GLOBE, by default only kernels in the crust/mantle region ("reg1") will be considered for summation.
+You can change the default by setting parameter \texttt{REG} in file \texttt{constants\_tomography.h} to the region of interest.
+\\
+
+Note that although we provide the preconditioned summation here, we recommend to smooth first both, the summed kernels and summed approximate Hessians, before inverting the summed Hessian and applying it as preconditioner to the gradient.
+From experience, we prefer doing
+\begin{equation}
+\boldg =  \frac{1}{\mathcal{F}_{smooth} \left( \Sum^{N}_i \tilde{H}^{i} \right) }  \; \mathcal{F}_{smooth} \left( \Sum_i K^{i}  \right) \nonumber
+\end{equation}
+rather than
+\begin{equation}
+\boldg = \mathcal{F}_{smooth} \left( \frac{1}{\Sum^{N}_i \tilde{H}^{i}}  \;  \Sum_i K^{i}  \right) \: . \nonumber
+\end{equation}
+Due to the large sensitivity kernel values close to source and receiver locations, the inverse of the thresholded summed Hessian becomes better balanced in the former.
+
+\subsection{Smoothing and post-processing}
+For additional smoothing and post-processing of kernels, gradients, or models, we provide tools useful for gradient and model preparation.
+These additional tools for post-processing sensitivity kernels are provided together with the main package. For compilation of these tools, you type in the main directory:
+\begin{verbatim}
+make postprocess
+\end{verbatim}
+
+\noindent
+The following tools are provided:
+\begin{enumerate}
+\item [-] \texttt{xsmooth\_sem} is used for smoothing with a gaussian function.
+\item [-] \texttt{xclip\_sem} is used to threshold a kernel, gradient or model, clipping to a minimum and maximum value.
+\item [-] \texttt{xcombine\_sem} is used to combine different event kernels, gradient or model files, summing all individually specified files together (similar to \texttt{xsum\_kernels} but more flexible).
+\end{enumerate}
+The post-processing tools are primarily intented to be used to process kernel files. They can be used though on
+any scalar field of dimension \texttt{(NGLLX,NGLLY,NGLLZ,NSPEC)}.
+The tools are parallel programs -- they must be invoked with mpirun or other
+appropriate utility.  Operations are performed in embarrassingly-parallel fashion.
+
+\begin{description}
+\item [Smoothing:] To smooth a kernel, gradient or model file, use
+\begin{verbatim}
+mpirun -np NPROC bin/xsmooth_sem SIGMA_H SIGMA_V KERNEL_NAME INPUT_DIR OUPUT_DIR
+\end{verbatim}
+with command line arguments:
+\texttt{SIGMA\_H} horizontal smoothing radius,
+\texttt{SIGMA\_V} vertical smoothing radius,
+\texttt{KERNEL\_NAME} kernel name, e.g. \texttt{alpha\_kernel},
+\texttt{INPUT\_DIR} directory from which kernels are read,
+\texttt{OUTPUT\_DIR} directory to which smoothed kernels are written.
+
+Smooths kernels by convolution with a Gaussian. Writes the resulting
+smoothed kernels to \texttt{OUTPUT\_DIR}.
+
+Files written to \texttt{OUTPUT\_DIR} have the suffix 'smooth' appended,
+e.g. \texttt{proc***alpha\_kernel.bin} becomes \texttt{proc***alpha\_kernel\_smooth.bin}.
+
+\item [Clipping:] Values in a kernel, gradient or model binary file can be clipped with a minimum/maximum threshold value using
+\begin{verbatim}
+mpirun -np NPROC bin/xclip_sem MIN_VAL MAX_VAL KERNEL_NAMES INPUT_FILE OUTPUT_DIR
+\end{verbatim}
+with command line arguments:
+\texttt{MIN\_VAL} threshold below which array values are clipped,
+\texttt{MAX\_VAL} threshold above which array values are clipped,
+\texttt{KERNEL\_NAMES} one or more kernel names separated by commas,
+\texttt{INPUT\_DIR} directory from which arrays are read,
+\texttt{OUTPUT\_DIR} directory to which clipped array are written.
+
+For each name in \texttt{KERNEL\_NAMES}, reads kernels from \texttt{INPUT\_DIR}, applies
+thresholds, and writes the resulting clipped kernels to \texttt{OUTPUT\_DIR}.
+
+\texttt{KERNEL\_NAMES} is a comma-delimited list of kernel names,
+e.g. \texttt{alphav\_kernel,alphah\_kernel}.
+\begin{itemize}
+\item [] For example, in this case type
+{\small
+\begin{verbatim}
+mpirun -np 4 ./bin/xclip_sem -0.1 0.1 alphav_kernel,alphah_kernel DIR1/ DIR1/
+\end{verbatim}
+}
+to clip the corresponding kernels in directory \texttt{DIR1} to be in range $[-0.1,0.1]$.
+\end{itemize}
+
+Files written to \texttt{OUTPUT\_DIR} have the suffix 'clip' appended,
+e.g. \texttt{proc***alpha\_kernel.bin} becomes \texttt{proc***alpha\_kernel\_clip.bin}
+
+\item [Combining:] To combine kernel values from different kernel directories, you can use
+\begin{verbatim}
+mpirun -np NPROC bin/xcombine_sem KERNEL_NAMES INPUT_FILE OUTPUT_DIR
+\end{verbatim}
+with command line arguments:
+\texttt{KERNEL\_NAMES} one or more kernel names separated by commas,
+\texttt{INPUT\_FILE} text file containing list of kernel directories,
+\texttt{OUTPUT\_PATH} directory to which summed kernels are written.
+
+For each name in \texttt{KERNEL\_NAMES}, sums kernels from directories specified in
+\texttt{INPUT\_FILE}. Writes the resulting sums to \texttt{OUTPUT\_DIR}.
+
+\texttt{INPUT\_FILE} is a text file containing a list of absolute or relative paths to
+kernel directories, one directory per line.
+
+\texttt{KERNEL\_NAMES} is a comma-delimited list of kernel names,
+e.g. \texttt{alpha\_kernel,beta\_kernel,rho\_kernel}.
+
+\end{description}
+
+\subsection{Model updating}
+We provide (simple) tools for updating the current model $\boldm_i$ using a gradient $\boldg$ and steplength $\alpha$:, i.e.
+\begin{equation}
+\boldm_{i+i} = \boldm_i + \delta \boldm  \nonumber
+\end{equation}
+with
+\begin{equation}
+\delta \boldm = - \alpha \, \boldg \nonumber \: .
+\end{equation}
+\noindent
+The following tomographic tools are provided:
+\begin{enumerate}
+\item [-] \texttt{xadd\_model\_iso} can be used to update isotropic model files with a (summed \& smoothed) gradient.
+The gradient files are given for isotropic parameters or isotropic bulk parameters.
+
+\end{enumerate}
+
+Note that instead of using the density kernels $K_{\rho}$ for model updates, setting \texttt{USE\_RHO\_SCALING} to \texttt{.true.} will
+ignore the density kernel and use density perturbations scaled from isotropic $V_s$ pertubations.\\
+
+\begin{description}
+\item [Isotropic model update:] The program \texttt{xadd\_model\_iso} can be used to update {\it isotropic} model files with
+(smoothed \& summed) event kernels:
+\begin{verbatim}
+mpirun -np 4 bin/xadd_model_iso step_factor [INPUT-KERNELS-DIR/] [OUTPUT-MODEL-DIR/]
+\end{verbatim}
+with command line arguments:
+\texttt{step\_factor} the step length to scale the gradient, e.g. $0.03$ for a $\pm 3$\% update,
+\texttt{INPUT-KERNELS-DIR/} (optional) directory which holds summed kernels (e.g. \texttt{proc***alpha\_kernel.bin},.., by default
+directory \texttt{INPUT\_GRADIENT/} is used),
+\texttt{OUTPUT-MODEL-DIR/} (optional) directory which will hold new model files (e.g. \texttt{proc***vp\_new.bin},.., by default directory
+\texttt{OUTPUT\_MODEL/} is used).
+
+The gradients are given for isotropic parameters \texttt{(alpha,beta,rho)} or \texttt{(bulk\_c,beta,rho)}.
+
+The algorithm uses a steepest descent method with a step length
+determined by the given maximum update percentage.
+
+By default, the directory and file setup used is:
+\begin{itemize}
+\item [-] directory \texttt{INPUT\_MODEL/} contains:
+\begin{verbatim}
+proc***_vs.bin
+proc***_vp.bin
+proc***_rho.bin
+\end{verbatim}
+
+\item [-] directory \texttt{INPUT\_GRADIENT/} contains:
+\begin{verbatim}
+proc***_bulk_c_kernel_smooth.bin
+proc***_bulk_beta_kernel_smooth.bin
+proc***_rho_kernel_smooth.bin
+\end{verbatim}
+or
+\begin{verbatim}
+proc***_alpha_kernel_smooth.bin
+proc***_beta_kernel_smooth.bin
+proc***_rho_kernel_smooth.bin
+\end{verbatim}
+depending on the model parameterization.
+
+\item [-] directory \texttt{topo/} contains:
+\begin{verbatim}
+proc***_external_mesh.bin
+\end{verbatim}
+for the Cartesian version, and
+\begin{verbatim}
+proc***_solver_data.bin
+\end{verbatim}
+for the GLOBE version.
+\end{itemize}
+
+\noindent
+The new model files are stored in directory \texttt{OUTPUT\_MODEL/} as
+\begin{verbatim}
+proc***_vp_new.bin
+proc***_vs_new.bin
+proc***_rho_new.bin
+\end{verbatim}
+
+\end{description}
+
+Note that additional post-processing and model update tools are provided for the SPECFEM3D\_GLOBE version and will be provided for the Cartesian version as well in future.
+
 \section{OLD VERSION OF THE SECTION, WILL SOON BE IMPROVED AND ENRICHED}
@@
 algorithms that could alternatively be used. The tomographic inversion
-of \citet{TaLiMaTr09,TaLiMaTr2010} used SPECFEM3D Cartesian as well
+of \citet{TaLiMaTr09,TaLiMaTr2010} used SPECFEM3D\_Cartesian as well
 as several additional components which are also stored on the CIG

### 2015-10-12 doc/USER_MANUAL/figures/tomo_dir_struct.tex

\documentclass{standalone}

\usepackage{amsmath}
\usepackage{pgfplots}
\pgfplotsset{compat=1.3}

% Latex generated graphics
\usepackage{tikz}

% tikz libraries
\usetikzlibrary{trees}

%\usetikzlibrary{positioning}
%\usetikzlibrary{shapes}
%\usetikzlibrary{calc}
%\usetikzlibrary{backgrounds}
%\usetikzlibrary{fit}
%\usetikzlibrary{patterns}

%\tikzset{timed/.style=dashed}

% colors
\usepackage{color}
\definecolor{myred}{rgb}{0.5,0.2,0.2}

\begin{document}

\tikzstyle{every node}=[draw=black,thick,anchor=west,inner sep=2pt,minimum size=1pt]

\tikzstyle{links}=[draw=red,fill=myred!20]
%\tikzstyle{file}=[dashed,fill=gray!50]
\tikzstyle{file}=[draw=none]
\tikzstyle{missing}=[draw=white]

\begin{tikzpicture}[%
  grow via three points={one child at (0.5,-0.7) and
  two children at (0.5,-0.7) and (0.5,-1.4)},
  edge from parent path={(\tikzparentnode.south) |- (\tikzchildnode.west)}
  ]
  \node {TOMO\_ROOT\_DIR}
    child { node {bin}}
    child { node [file] {kernels\_list.txt}}
    child { node {INPUT\_KERNELS}
        child { node [links] {event1\_DATABASES\_MPI}
          child { node [file] {proc***\_alpha\_kernel.bin}}
          child { node [file] {proc***\_beta\_kernel.bin}}
          child { node [file] {proc***\_rho\_kernel.bin}}
          child { node [file] {proc***\_hess\_kernel.bin}}
      }
     child [missing] {}
     child [missing] {}
     child [missing] {}
     child [missing] {}
     child { node [links] {event2\_DATABASES\_MPI}
       child { node [file] {....}}
      }
      child [missing] {}
      child { node [links] {....}}
    }
    child [missing] {}
    child [missing] {}
    child [missing] {}
    child [missing] {}
    child [missing] {}
    child [missing] {}
    child [missing] {}
    child [missing] {}
    child { node {OUTPUT\_SUM}};
\end{tikzpicture}

\end{document}

### 2015-10-20 doc/USER_MANUAL/05_running_the_solver.tex
added Quentin Brissaud's note on the acoustic formulation to the users manual

@@
 to the first copy of the executable programs by typing \texttt{ln -s run0001/bin bin}.
+
+%------------------------------------------------------------------------------------------------%
+
+%------------------------------------------------------------------------------------------------%
+
+\chapter{Note on the SPECFEM acoustic formulation}
+
+{\Large Note from Quentin Brissaud, Dimitri Komatitsch and Rapha\"el Garcia, October 2015}
+
+\section{About the potential formulation used in SPECFEM for pure acoustic waves (i.e. in the absence of gravity terms)}
+
+In SPECFEM, when considering purely acoustic waves (neglecting gravity acceleration and viscosity)
+in fluid regions one uses a potential formulation to represent the "density times displacement" perturbation
+\cite{ChKoViCaVaFe07}:
+\begin{equation}
+\rho \mathbf{u} = \nabla \chi
+\label{potential}
+\end{equation}
+where $\mathbf{u}$ is the displacement perturbation, $\rho$ the background density and $\chi$ is the scalar potential.\\
+This formulation has two great advantages: it reduces the number of unknowns from three to one, and it also automatically removes artefacts
+in finite-element methods (FEM) caused by the displacement formulation \cite{HaOuVe78}.
+This is crucial since it is the main reason for which a displacement formulation cannot be used in fluid regions in a FEM.
+\newline
+
+However, this formulation assumes that the density is spatially constant, or locally slowly varying (e.g. inside a given spectral element),
+because the use of a potential for acoustic waves is justified by the fact that acoustic displacements are irrotational.
+In the cases for which the background density has strong local variations, that formulation is not valid any more.
+Indeed, taking the rotational of \eqref{potential}
+\begin{eqnarray*}
+\nabla \times (\rho \mathbf{u}) &=& \nabla\times\nabla \chi
+\end{eqnarray*}
+Since for any scalar $\chi$, $ \nabla\times\nabla \chi = 0$, this yields
+\begin{eqnarray*}
+\nabla\rho\times\mathbf{u} + \rho\nabla\times\mathbf{u} &=& \mathbf{0}\\
+\nabla\times\mathbf{u}&=&-(1/\rho)\nabla\rho\times\mathbf{u}
+\end{eqnarray*}
+Then, one notices that in the case of strong density gradients the rotational of $\mathbf{u}$ is no longer zero.
+Thus, the potential formulation \eqref{potential} is no more valid. It is valid only when $(1/\rho)\nabla\rho\times\mathbf{u}$ can be considered negligible,
+i.e. when both the gradient of density and the value of displacement are small, or if the displacement is parallel or quasi-parallel to the gradient of density,
+in which case their rotational is also close to zero.
+\newline
+
+Another approach exists for acoustic waves, based on a displacement potential instead of on a potential of $\rho \mathbf{u}$:
+\begin{equation}
+\mathbf{u} = \nabla \chi \, ,
+\end{equation}
+which gives $\nabla\times\mathbf{u} = \mathbf{0}$.
+But in this case the gradient of density is neglected in the Eulerian momentum equation in order to identify gradient terms.
+Indeed, from the Eulerian momentum equation for inviscid acoustic waves only
+\begin{eqnarray*}
+\rho\partial_t^2\mathbf{u} &=& -\nabla p\\
+\nabla (\partial_t^2\chi)&=& -(1/\rho)\nabla p \, .
+\end{eqnarray*}
+When $\nabla\rho \gg \mathbf{0}$ strictly speaking one cannot identify $\partial_t^2\chi$ with $p$.
+
+The solution to this is to work with pressure as the scalar unknown (\cite{God11}, equation 23),
+or even better with the time integral (the primitive) of pressure (see e.g. \cite{Eve81});
+this was the approach that was originally implemented in SPECFEM until 10 years ago or so, but \cite{ChCaVi03} showed that switching to a displacement potential
+or to a potential of $\rho \mathbf{u}$ enables one to couple the fluid and solid regions without iterating on the coupling condition in the time scheme,
+thus resulting in much easier and cheaper implementation; since at that time SPECFEM users were not studying media with strong local variations of density (in solid Earth
+geophysics such large local variations in fluid regions never happen), the SPECFEM developers permanently switched to such a displacement-based potential formulation.
+However, for media with strong local variations of density in a fluid region we should reconsider the issue and consider trying the pressure-based formulation again.
+
+Cases in which this can be problematic may include e.g. upper atmosphere studies: the density there is very small but displacements can become very large;
+in 1D atmosphere models $\frac{\nabla \rho}{\rho}$ is approximately constant and equal to -1/H, where H is the scaling height,
+which leads to $\frac{\nabla \rho}{\rho} \simeq$ 3.42*1e-5, i.e. very small; however displacements there can be large
+and thus it is not clear if $(1/\rho)\nabla\rho\times\mathbf{u}$ can be considered negligible there.

### 2015-10-20 doc/USER_MANUAL/05_running_the_solver.tex
improved the note on the acoustic formulation

@@
 
-However, this formulation assumes that the density is spatially constant, or locally slowly varying (e.g. inside a given spectral element),
-because the use of a potential for acoustic waves is justified by the fact that acoustic displacements are irrotational.
-In the cases for which the background density has strong local variations, that formulation is not valid any more.
+However, this formulation indirectly assumes that the density is spatially constant, or locally slowly varying (e.g. inside a given spectral element),
+because acoustic displacements must be irrotational. Since we know that $\nabla \times \mathbf{u}$ must be equal to zero
+we can see that when density has strong local variations, equation (\ref{potential}) is not valid any more.
 Indeed, taking the rotational of \eqref{potential}
 \begin{eqnarray*}
-\nabla \times (\rho \mathbf{u}) &=& \nabla\times\nabla \chi
+\nabla \times (\rho \mathbf{u}) &=& \nabla\times\nabla \chi \, ,
 \end{eqnarray*}
-Since for any scalar $\chi$, $ \nabla\times\nabla \chi = 0$, this yields
+since the curl of a gradient is always zero i.e. for any scalar $\chi$, $ \nabla\times\nabla \chi = 0$, we get
 \begin{eqnarray*}
@@
 \end{equation}
-which gives $\nabla\times\mathbf{u} = \mathbf{0}$.
+which correctly gives $\nabla\times\mathbf{u} = \mathbf{0}$.
 But in this case the gradient of density is neglected in the Eulerian momentum equation in order to identify gradient terms.

### 2015-10-21 doc/USER_MANUAL/05_running_the_solver.tex
better version of the notes on the acoustic formulation

@@
 in fluid regions one uses a potential formulation to represent the "density times displacement" perturbation
-\cite{ChKoViCaVaFe07}:
+\citep{ChKoViCaVaFe07}:
 \begin{equation}
@@
 where $\mathbf{u}$ is the displacement perturbation, $\rho$ the background density and $\chi$ is the scalar potential.\\
-This formulation has two great advantages: it reduces the number of unknowns from three to one, and it also automatically removes artefacts
-in finite-element methods (FEM) caused by the displacement formulation \cite{HaOuVe78}.
+This formulation has several great advantages: it reduces the number of unknowns from three to one, and it also automatically removes artefacts
+in finite-element methods (FEM) caused by the displacement formulation \citep{HaOuVe78}.
 This is crucial since it is the main reason for which a displacement formulation cannot be used in fluid regions in a FEM.
-\newline
+It also permits acoustic-elastic coupling based on a non-iterative time scheme \citep{ChVa04}.
+Displacement is then: u = grad(Chi) / rho,
+velocity is then: v = grad(Chi\_dot) / rho (Chi\_dot being the time derivative of Chi)
+and pressure is: p = - Chi\_dot\_dot  (Chi\_dot\_dot being the time second derivative of Chi).
+First-order acoustic-acoustic discontinuities are also handled automatically
+because pressure is continuous at such an interface, therefore Chi\_dot\_dot
+is continuous, therefore Chi is also continuous, which is consistent with
+the spectral-element basis functions and with the assembling process.
+This is the reason why a simple displacement potential u = grad(Chi) would
+not work because it would be discontinuous at such an interface and would
+therefore not be consistent with the basis functions. And first-order acoustic-acoustic discontinuities are used
+routinely in several domains, e.g. in the oil industry when dealing with geological models as acoustic only (no S waves),
+or in ocean acoustics when considering the ocean bottom as fluid (again, no S waves).
 
@@
 
-The solution to this is to work with pressure as the scalar unknown (\cite{God11}, equation 23),
+A solution to this could be to work with pressure as the scalar unknown (\cite{God11}, equation 23),
 or even better with the time integral (the primitive) of pressure (see e.g. \cite{Eve81});
-this was the approach that was originally implemented in SPECFEM until 10 years ago or so, but \cite{ChCaVi03} showed that switching to a displacement potential
+this was the approach that was originally implemented in SPECFEM until 10 years ago or so, but \cite{ChVa04} showed that switching to a displacement potential
 or to a potential of $\rho \mathbf{u}$ enables one to couple the fluid and solid regions without iterating on the coupling condition in the time scheme,
@@
 geophysics such large local variations in fluid regions never happen), the SPECFEM developers permanently switched to such a displacement-based potential formulation.
-However, for media with strong local variations of density in a fluid region we should reconsider the issue and consider trying the pressure-based formulation again.
+However, for media with strong local variations of density in a fluid region we should reconsider the issue and consider trying the pressure-based formulation again
+to see if there is a way of using it without having to numerically iterate on the fluid-solid matching condition.
+
+Another possibility, suggested by Emmanuel Chaljub, would be to use two potentials instead of a single one (see Section~2.4 of \cite{ChKoViCaVaFe07}).
 

### 2015-10-21 doc/USER_MANUAL/05_running_the_solver.tex
final version of the section on the acoustic formulation

@@
 
-{\Large Note from Quentin Brissaud, Dimitri Komatitsch and Rapha\"el Garcia, October 2015}
+{\Large Note from Quentin Brissaud, Dimitri Komatitsch and Rapha\"el Garcia, October 2015,with feedback from Jeroen Tromp and from Emmanuel Chaljub}
 
-\section{About the potential formulation used in SPECFEM for pure acoustic waves (i.e. in the absence of gravity terms)}
+\section{About the potential formulation used in SPECFEM for purely acoustic waves (i.e. in the absence of gravity terms)}
 
@@
 or in ocean acoustics when considering the ocean bottom as fluid (again, no S waves).
+\newline
 
@@
 When $\nabla\rho \gg \mathbf{0}$ strictly speaking one cannot identify $\partial_t^2\chi$ with $p$.
+\newline
 
@@
 to see if there is a way of using it without having to numerically iterate on the fluid-solid matching condition.
+\newline
 
 Another possibility, suggested by Emmanuel Chaljub, would be to use two potentials instead of a single one (see Section~2.4 of \cite{ChKoViCaVaFe07}).
+\newline
 
@@
 and thus it is not clear if $(1/\rho)\nabla\rho\times\mathbf{u}$ can be considered negligible there.
+\newline
+
+Note that it is only when gravity is off that there is potentially a problem.
+More details on the case with gravity can be found in \cite{ChVa04} (in particular in their Figure 2 and Section 2.1)
+and in \cite{KoTsTr05}. As summarized by Jeroen Tromp, note that in the presence of gravity and rotation we do not assume that density is constant in the fluid
+(see Section~3.3 in \cite{KoTsTr05}). Instead, we assume that the fluid is stably stratified, i.e., that the Brunt-V\"ais\"al\"a frequency vanishes.
+Only in the step leading to eqns. (35) and (36) in \cite{KoTsTr05} do we make the Cowling approximation (\cite{Cow41}) and set the Brunt-V\"ais\"al\"a frequency equal to zero.
+If we make the Cowling approximation, assume the core is stably stratified, and ignore rotation, then we obtain eqn. (41).
+In that case the displacement is the gradient of a potential (not density times the gradient of a potential).
+More generally, the displacement is given by eqn. (28). The results with and without rotation (always assuming stable stratification), are implemented in SPECFEM3D\_GLOBE.
+\newline
+
+The question for regional simulations, e.g., in an ocean layer, is whether or not we should incorporate gravity in the analysis, and whether or not we can assume that, on the time scales of seismic waves, the ocean is stably stratified.
+\newline
+
+Note that one cannot have density gradients coming out of nowhere, i.e., in principle one cannot ignore gravity.
+In practice, density gradients are controlled by the Brunt-V\"ais\"al\"a frequency, be it in the oceans or the outer core.
+If we assume stable stratification, then we are naturally led to (35) and (36) in \cite{KoTsTr05}, or (41) in the absence of rotation.
+Jeroen Tromp thinks we either need to assume $N^2=0$ or $N^2>0$, since when $N^2<0$ the fluid will overturn.
+SPECFEM3D/SPECFEM3D\_GLOBE currently implement the $N^2=0$ case, so the question is whether we need to worry about the case $N^2>0$.
+The latter corresponds to eqns.
+(33) and (34) in Komatitsch et al.
+(2007) (with phi=0 in the Cowling approximation).
+\newline
+
+Regarding always including the effect of gravity in order to avoid this issue,
+one (important) exception could be the oil industry case when they use P waves only, in which case they do not include gravity and their spatially-varying density is just rho taken from the elastic model...
+These days they mostly run fully elastic simulations instead?
+(although some groups still often uses the acoustic approximation for large inverse problems); that's also true in ocean acoustics because the approximation often made is to consider the solid ocean bottom as fluid (no S waves, again just taking rho and cp from the elastic model).
+\newline
+
+So, as summarized by Jeroen, the question is whether we can safely assume $N^2=0$ as we currently do, or whether we also need to consider the case $N^2>0$.

### 2015-11-12 doc/USER_MANUAL/13_changing_the_model.tex
Buildbot ran the code beautifier.

@@
 $Q_{\kappa}$ and $Q_{\mu}$ respectively \citep{AnHa78}.
+Note that Qmu is always equal to Qs, but Qkappa is in general not equal to Qp.
+To convert one to the other see doc/note\_on\_Qkappa\_versus\_Qp.pdf and utils/attenuation/conversion\_from\_Qkappa\_Qmu\_to\_Qp\_Qs\_from\_Dahlen\_Tromp\_959\_960.f90.
 For simulations with attenuation, please note that the Vp- and Vs-velocities of your model are given for a reference
@@
 ($kg/m^{3}$), $v_{p}$ ($m/s$) and $v_{s}$ ($m/s$)
-\item [{\texttt{qkappa\_atten,qmu\_atten}}] Bulk and shear wave quality factor: $0<Q_{\kappa,\mu}<9000$
+\item [{\texttt{qkappa\_atten,qmu\_atten}}] Bulk and shear wave quality factor: $0<Q_{\kappa,\mu}<9000$.
+Note that Qmu is always equal to Qs, but Qkappa is in general not equal to Qp.
+To convert one to the other see doc/note\_on\_Qkappa\_versus\_Qp.pdf and utils/attenuation/conversion\_from\_Qkappa\_Qmu\_to\_Qp\_Qs\_from\_Dahlen\_Tromp\_959\_960.f90.
 \item [{\texttt{iflag\_aniso}}] anisotropic model flag, $0$ indicating

### 2015-11-12 doc/USER_MANUAL/15_informations_for_developers.tex

@@
 They are useful to learn how the spectral-element method works, and how to write or modify a code to implement it.
-Also useful to test new ideas by modifying these simple codes to run some tests.\\
+Also useful to test new ideas by modifying these simple codes to run some tests.
+We also have a similar, even simpler, demo source code for the 2D case in directory\\
+utils/small\_SEM\_solver\_in\_Fortran\_without\_MPI\_to\_learn of the specfem2d package.
 

### 2015-11-12 doc/USER_MANUAL/01_introduction.tex
added a sentence about varying density in fluids to the users manual

@@
 very distorted mesh elements \citep{OlSe11}.\\
+
+\red{In fluids, when gravity is turned off, SPECFEM2D uses the classical
+linearized Euler equation; thus if you have sharp local variations of
+density in the fluid (highly heterogeneous fluids
+in terms of density) or if density becomes extremely small in some
+regions of your model (e.g. for upper-atmosphere studies), before using
+the code please make sure the linearized Euler equation is a valid
+approximation in the case you want to study, and/or if you should turn gravity on.
+For more details on that see e.g. \cite{COA2011}.}\\
 
@@
 \citet{KoVi98} and \citet{KoTr99}, whereas applications involving
-Chebyshev basis functions and a nondiagonal mass matrix include \citet{SePr94},
+Chebyshev basis functions and a non-diagonal mass matrix include \citet{SePr94},
 \citet{PrCaSe94} and \citet{SePrPr95}.\\

### 2015-11-12 doc/USER_MANUAL/01_introduction.tex
fixed two typos

--- 2015-04-24 doc/USER_MANUAL/01_introduction.tex
@@
 very distorted mesh elements \citep{OlSe11}.\\
+
+\red{In fluids, when gravity is turned off, SPECFEM3D uses the classical
+linearized Euler equation; thus if you have sharp local variations of
+density in the fluid (highly heterogeneous fluids
+in terms of density) or if density becomes extremely small in some
+regions of your model (e.g. for upper-atmosphere studies), before using
+the code please make sure the linearized Euler equation is a valid
+approximation in the case you want to study, and/or see if you should turn gravity on.
+For more details on that see e.g. \cite{COA2011}.}\\
 

### 2015-11-24 doc/USER_MANUAL/copyright_and_version.tex
updates manual, makes latex header more similar in all specfem versions; add a sentence about contributions and copyrights

@@
-\addcontentsline{toc}{chapter}{Copyright}
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \chapter*{Copyright}
+\addcontentsline{toc}{chapter}{Copyright}
 
@@
 
-Princeton University, USA, and CNRS / University of Marseille, France
+Princeton University, USA, and CNRS / University of Marseille, France\\
+$\copyright$ Princeton University and CNRS / University of Marseille, July 2012\\
 
-$\copyright$ Princeton University and CNRS / University of Marseille, July 2012
-
+\noindent
 This program is free software; you can redistribute it and/or modify
@@
 
+\noindent
+Please note that by contributing to this code, the developer understands and agrees that this project and contribution
+are public and fall under the open source license mentioned above.\\
+
+\noindent
 \textbf{\underline{Evolution of the code:}}\\

### 2015-11-24 doc/USER_MANUAL/manual_SPECFEM3D_Cartesian.tex
added a sentence about varying density in fluids to the users manual
updates manual, makes latex header more similar in all specfem versions; add a sentence about contributions and copyrights

@@
 \geometry{verbose,tmargin=1in,bmargin=1in,lmargin=1in,rmargin=1in}
+
+% table of contents and numbering
+\usepackage{tocbibind}
 \setcounter{secnumdepth}{3}
 \setcounter{tocdepth}{3}
+
 \usepackage{babel}
@@
 \usepackage{textcomp}
-\usepackage{url}
 \usepackage{amsmath}
 \usepackage{amssymb}
+
+\IfFileExists{url.sty}{\usepackage{url}}
+                      {\newcommand{\url}{\texttt}}
+
+% biblio
+\usepackage[authoryear]{natbib}
+% biblio GJI
+\bibliographystyle{abbrvnat}
+
+% fonts
+\usepackage{times}
+
+% figures
 \usepackage{graphicx}
-\usepackage[authoryear]{natbib}
-\usepackage[unicode=true,pdfusetitle,
- bookmarks=true,bookmarksnumbered=true,bookmarksopen=false,
- breaklinks=false,pdfborder={0 0 0},backref=section,colorlinks=false,pdfpagemode=None]
- {hyperref}
+
+% we are running pdflatex, so convert .eps files to .pdf
+%\usepackage[dvips]{epsfig}
+
+% colors to show the corrections
+\usepackage[dvipsnames,usenames]{xcolor}
+\definecolor{light-gray}{rgb}{0.7,0.7,0.7}
+\definecolor{dark-gray}{rgb}{0.1,0.1,0.3}
+\definecolor{light-blue}{rgb}{0.7,0.7,0.9}
+\definecolor{dark-blue}{rgb}{0.2,0.2,0.5}
+
+% hyperlinks to sections and references
+\usepackage[pdftex]{hyperref}
+% more options
 \hypersetup{
- pdftex,pdfstartview=FitH,pdfpagelayout=SinglePage}
+  bookmarks=true,
+  bookmarksnumbered=true,
+  pdfpagemode=UseNone,
+  pdfstartview=FitH,
+  pdfpagelayout=SinglePage,
+  pdfborder={0 0 0},
+  colorlinks=true,      % color links
+  citecolor=dark-gray,  % citation links
+  filecolor=dark-gray,  % file links
+  linkcolor=dark-gray,  % internal link (rgb values)
+  urlcolor=dark-blue    % external weblinks
+}
+
+% create thumbnails
+%\usepackage[pdftex]{thumbpdf}
+
 \usepackage{breakurl}
@@
 
-% figures
-%\usepackage[dvips]{epsfig}
-
-% fonts
-\usepackage{times}
-
 % hyperlinks to sections and references
-
 \newcommand{\urlwithparentheses}[1]{(\url{#1})}
-
-% biblio GJI
-\bibliographystyle{abbrvnat}
 
@@
 % colors for the text
-\@ifundefined{definecolor}
- {\usepackage{color}}{}
-\newcommand{\red}[1]{\textcolor{Red}{#1}}
+\newcommand{\red}[1]{\textbf{\textcolor{Red}{#1}}}
 
@@
 
-\usepackage{babel}
-
 \makeatother
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Start of document.
 
@@
 
+% date of last edit
 \date{\today}
@@
 
-\addcontentsline{toc}{chapter}{References} \bibliography{bibliography}
+\bibliography{bibliography}
 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \end{document}

### 2016-01-03 doc/USER_MANUAL/08_doing_tomography.tex
added a reference to the users manual

@@
 since in the oil industry for instance it is known that the $L^1$ norm
-(\cite{CrPiNoMcTa90,Brossier_2010_WDR}), hybrid $L^1-L^2$ norms (\cite{Bube_1997_HL1}), Hubert norm (\cite{Ha_2009_WIU}),
+(\cite{CrPiNoMcTa90,BrOpVi2010}), hybrid $L^1-L^2$ norms (\cite{Bube_1997_HL1}), Hubert norm (\cite{Ha_2009_WIU}),
 Student-t distribution (\cite{ArVaHe11,Jeong_2015}) etc... can be more robust that the $L^2$ norm used here in the context of synthetic data with no noise.

### 2016-02-15 doc/USER_MANUAL/04_creating_databases.tex
added notes about attenuation
larger figure in the manual (for an important one)

@@
 it will also read in the external binary mesh files for $Q_{\kappa}$ and $Q_{\mu}$.
+Note that Qmu is always equal to Qs, but Qkappa is in general not equal to Qp.
+To convert one to the other see doc/note\_on\_Qkappa\_versus\_Qp.pdf and utils/attenuation/conversion\_from\_Qkappa\_Qmu\_to\_Qp\_Qs\_from\_Dahlen\_Tromp\_959\_960.f90.
 
@@
 is not negligible in the solid layer at the sea bottom.
+Note that Qmu is always equal to Qs, but Qkappa is in general not equal to Qp.
+To convert one to the other see doc/note\_on\_Qkappa\_versus\_Qp.pdf and utils/attenuation/conversion\_from\_Qkappa\_Qmu\_to\_Qp\_Qs\_from\_Dahlen\_Tromp\_959\_960.f90.
 \item [{\texttt{ANISOTROPY}}] Set to \texttt{.true.} if you want to use
@@
 \noindent \begin{centering}
-\includegraphics[width=3in]{figures/how_to_use_PML_when_a_tomographic_velocity_model_is_used.png}
+\includegraphics[width=6in]{figures/how_to_use_PML_when_a_tomographic_velocity_model_is_used.png}
 \par\end{centering}

### 2016-05-01 doc/USER_MANUAL/copyright_and_version.tex
removed old copy of AxiSEM, as requested by several developers (there was no reason to have one, and it was making the code more complex for no reason)

@@
 Convolutional PML, LDDRK time scheme, bulk attenuation support, simultaneous MPI runs,
-ADIOS file I/O support, coupling with external codes, new seismogram names,
+ADIOS file I/O support, new seismogram names,
 Deville routines for additional GLL degrees, tomography tools, unit/regression test framework,

### 2016-05-01 EXAMPLES/coupling_with_EXTERNAL_CODES/with_AxiSEM/Compare_and_rotate_sismos_Specfem_Axisem/compare_gnuplot_and_makepdf.sh

@@ -1,31 +1,0 @@
-#!/bin/bash
-
-### Launch the script from the directory containing subdirs SISMOS_RUN_1 and SISMOS_RUN_2
-
-mkdir SISMOS_PLOT_COMPARED_Axisem_and_Specfem_converted
-
-gnuplot allgn.plt
-
-mv *.eps SISMOS_PLOT_COMPARED_Axisem_and_Specfem_converted
-cd SISMOS_PLOT_COMPARED_Axisem_and_Specfem_converted
-
-epstopdf SY.S0001.E.semd_COMPARED.eps
-epstopdf SY.S0001.N.semd_COMPARED.eps
-epstopdf SY.S0001.Z.semd_COMPARED.eps
-epstopdf SY.S0127.E.semd_COMPARED.eps
-epstopdf SY.S0127.N.semd_COMPARED.eps
-epstopdf SY.S0127.Z.semd_COMPARED.eps
-epstopdf SY.S0230.E.semd_COMPARED.eps
-epstopdf SY.S0230.N.semd_COMPARED.eps
-epstopdf SY.S0230.Z.semd_COMPARED.eps
-epstopdf SY.S0253.E.semd_COMPARED.eps
-epstopdf SY.S0253.N.semd_COMPARED.eps
-epstopdf SY.S0253.Z.semd_COMPARED.eps
-epstopdf SY.S0076.E.semd_COMPARED.eps
-epstopdf SY.S0076.N.semd_COMPARED.eps
-epstopdf SY.S0076.Z.semd_COMPARED.eps
-
-rm *.eps
-
-gs -dBATCH -dNOPAUSE -q -sDEVICE=pdfwrite -sOutputFile=concatall.pdf *.pdf
-

### 2016-05-01 EXAMPLES/coupling_with_EXTERNAL_CODES/with_AxiSEM/example_IASP91_model/batch_coupling_step8_xspecfem3d.sh

@@ -1,48 +1,0 @@
-#!/bin/bash
-
-
-#             ------------ BACTH AND SPECIFIC CLUSTER DIRECTIVES  ------
-
-#MSUB -r Coupling_step8_xspecfem3d       # Nom du job
-#MSUB -n 32
-#MSUB -N 2
-#MSUB -T 5400
-#MSUB -q standard
-#MSUB -e Xspecfem3D_32p.e
-#MSUB -o Xspecfem3D_32p.o
-#MSUB -A gen7165
-
-set -x
-cd ${BRIDGE_MSUB_PWD}
-
-######################################################################################################################
-
-# NBPROC is declared as integer (important do not change)
-declare -i NPROC NPROC_MINUS_ONE CPUS CHOICE MIDDLE
-
-# NUMBER OF MPI PROCESSES
-NPROC=32
-CPUS=32
-
-# MPIRUN COMMAND
-MPIRUN=ccc_mprun
-
-# ENTER OPTION FOR MPIRUN
-OPTION=
-
-# do not change
-NPROC_MINUS_ONE="$NPROC-1"
-
-# Define different paths and folders
-source paths_for_coupling_SPECFEM3D_AxiSEM.sh
-
-### Check AxiSEM directory for previous steps
-
-# ----------------------------------------------------
-# 8 / ------- Run simulation with specfem3D
-# ----------------------------------------------------
-
-source ${HOME_AxiSEM}/scripts_SPECFEM3D_for_AxiSEM.sh
-
-run_specfem3d_simulation
-

### 2016-05-01 src/generate_databases/get_model.f90

@@ -1,614 +1,0 @@
-!=====================================================================
-!
-!               S p e c f e m 3 D  V e r s i o n  3 . 0
-!               ---------------------------------------
-!
-!     Main historical authors: Dimitri Komatitsch and Jeroen Tromp
-!                        Princeton University, USA
-!                and CNRS / University of Marseille, France
-!                 (there are currently many more authors!)
-! (c) Princeton University and CNRS / University of Marseille, July 2012
-!
-! This program is free software; you can redistribute it and/or modify
-! it under the terms of the GNU General Public License as published by
-! the Free Software Foundation; either version 2 of the License, or
-! (at your option) any later version.
-!
-! This program is distributed in the hope that it will be useful,
-! but WITHOUT ANY WARRANTY; without even the implied warranty of
-! MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
-! GNU General Public License for more details.
-!
-! You should have received a copy of the GNU General Public License along
-! with this program; if not, write to the Free Software Foundation, Inc.,
-! 51 Franklin Street, Fifth Floor, Boston, MA 02110-1301 USA.
-!
-!=====================================================================
-
-
-  subroutine get_model(myrank)
-
-  use generate_databases_par, only: IMODEL, &
-    IMODEL_DEFAULT,IMODEL_GLL,IMODEL_1D_PREM,IMODEL_1D_CASCADIA,IMODEL_1D_SOCAL, &
-    IMODEL_SALTON_TROUGH,IMODEL_TOMO,IMODEL_USER_EXTERNAL,IMODEL_IPATI,IMODEL_IPATI_WATER, &
-    IDOMAIN_ACOUSTIC,IDOMAIN_ELASTIC,IDOMAIN_POROELASTIC, &
-    nspec => NSPEC_AB,ibool,mat_ext_mesh, &
-    materials_ext_mesh,nmat_ext_mesh,undef_mat_prop,nundefMat_ext_mesh, &
-    ANISOTROPY,COUPLE_WITH_EXTERNAL_CODE,MESH_A_CHUNK_OF_THE_EARTH, &
-    NGLLX,NGLLY,NGLLZ, &
[…]

### 2016-05-01 src/shared/read_parameter_file.f90

@@ -1,624 +1,0 @@
-!=====================================================================
-!
-!               S p e c f e m 3 D  V e r s i o n  3 . 0
-!               ---------------------------------------
-!
-!     Main historical authors: Dimitri Komatitsch and Jeroen Tromp
-!                        Princeton University, USA
-!                and CNRS / University of Marseille, France
-!                 (there are currently many more authors!)
-! (c) Princeton University and CNRS / University of Marseille, July 2012
-!
-! This program is free software; you can redistribute it and/or modify
-! it under the terms of the GNU General Public License as published by
-! the Free Software Foundation; either version 2 of the License, or
-! (at your option) any later version.
-!
-! This program is distributed in the hope that it will be useful,
-! but WITHOUT ANY WARRANTY; without even the implied warranty of
-! MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
-! GNU General Public License for more details.
-!
-! You should have received a copy of the GNU General Public License along
-! with this program; if not, write to the Free Software Foundation, Inc.,
-! 51 Franklin Street, Fifth Floor, Boston, MA 02110-1301 USA.
-!
-!=====================================================================
-
-  subroutine read_parameter_file(myrank,BROADCAST_AFTER_READ)
-
-  use constants
-
-  use shared_parameters
-
-  implicit none
-
-  integer, intent(in) :: myrank
-  logical, intent(in) :: BROADCAST_AFTER_READ
-
-! local variables
-  integer :: icounter,isource,idummy,ier
-  integer :: nproc_eta_old,nproc_xi_old
-  double precision :: hdur,minval_hdur
-  character(len=MAX_STRING_LEN) :: dummystring
-
[…]

### 2016-05-01 doc/USER_MANUAL/01_introduction.tex
added a reference to the users manual

@@
 Chebyshev basis functions and a non-diagonal mass matrix include \citet{SePr94},
-\citet{PrCaSe94} and \citet{SePrPr95}.\\
+\citet{PrCaSe94} and \citet{SePrPr95}.
+In the Legendre version that we use in SPECFEM the mass matrix is purposely slightly inexact but diagonal (but can be made exact if needed, see \cite{Teu15}),
+while in the Chebyshev version it is exact but non diagonal.\\
 
@@
 
+If you use the C-PML absorbing layer capabilities of the code, please cite at least one article
+written by the developers of the package, for instance:
+%
+\begin{itemize}
+\item \cite{XiKoMaMa14},
+\item \cite{XiMaCrKoMa16}.
+\end{itemize}
+%
+If you use the attenuation (anelastic/viscoelastic) capabilities of the code, please cite at least one article
+written by the developers of the package, for instance:
+%
+\begin{itemize}
+\item \cite{BlKoChLoXi16}.
+\end{itemize}
+%
+If you use the kernel capabilities of the code, please cite at least one article
+written by the developers of the package, for instance:
+%
+\begin{itemize}
+\item \cite{TrKoLi08},
+\item \cite{PeKoLuMaLeCaLeMaLiBlNiBaTr11},
+\item \cite{LiTr06},
+\item \cite{MoLuTr09}.
+\end{itemize}
+
 \noindent If you work on geophysical applications, you may be interested

### 2016-05-01 doc/USER_MANUAL/05_running_the_solver.tex
fixed a typo

--- 2015-10-20 doc/USER_MANUAL/05_running_the_solver.tex
@@
 
-%------------------------------------------------------------------------------------------------%
+\section{Note on the viscoelastic model used}
 
-%------------------------------------------------------------------------------------------------%
+\noindent
+The model used is a constant $Q$, thus with no dependence on frequency ($Q(f)$ = constant).
+See e.g. \cite{BlKoChLoXi16}. \\
 
-\chapter{Note on the SPECFEM acoustic formulation}
-
-{\Large Note from Quentin Brissaud, Dimitri Komatitsch and Rapha\"el Garcia, October 2015}
-
-\section{About the potential formulation used in SPECFEM for pure acoustic waves (i.e. in the absence of gravity terms)}
-
-In SPECFEM, when considering purely acoustic waves (neglecting gravity acceleration and viscosity)
-in fluid regions one uses a potential formulation to represent the "density times displacement" perturbation
-\cite{ChKoViCaVaFe07}:
-\begin{equation}
-\rho \mathbf{u} = \nabla \chi
-\label{potential}
-\end{equation}
-where $\mathbf{u}$ is the displacement perturbation, $\rho$ the background density and $\chi$ is the scalar potential.\\
-This formulation has two great advantages: it reduces the number of unknowns from three to one, and it also automatically removes artefacts
-in finite-element methods (FEM) caused by the displacement formulation \cite{HaOuVe78}.
-This is crucial since it is the main reason for which a displacement formulation cannot be used in fluid regions in a FEM.
-\newline
-
-However, this formulation indirectly assumes that the density is spatially constant, or locally slowly varying (e.g. inside a given spectral element),
-because acoustic displacements must be irrotational. Since we know that $\nabla \times \mathbf{u}$ must be equal to zero
-we can see that when density has strong local variations, equation (\ref{potential}) is not valid any more.
-Indeed, taking the rotational of \eqref{potential}
-\begin{eqnarray*}
-\nabla \times (\rho \mathbf{u}) &=& \nabla\times\nabla \chi \, ,
-\end{eqnarray*}
-since the curl of a gradient is always zero i.e. for any scalar $\chi$, $ \nabla\times\nabla \chi = 0$, we get
-\begin{eqnarray*}
-\nabla\rho\times\mathbf{u} + \rho\nabla\times\mathbf{u} &=& \mathbf{0}\\
-\nabla\times\mathbf{u}&=&-(1/\rho)\nabla\rho\times\mathbf{u}
-\end{eqnarray*}
-Then, one notices that in the case of strong density gradients the rotational of $\mathbf{u}$ is no longer zero.
-Thus, the potential formulation \eqref{potential} is no more valid. It is valid only when $(1/\rho)\nabla\rho\times\mathbf{u}$ can be considered negligible,
-i.e. when both the gradient of density and the value of displacement are small, or if the displacement is parallel or quasi-parallel to the gradient of density,
-in which case their rotational is also close to zero.
-\newline
-
-Another approach exists for acoustic waves, based on a displacement potential instead of on a potential of $\rho \mathbf{u}$:
-\begin{equation}
-\mathbf{u} = \nabla \chi \, ,
-\end{equation}
-which correctly gives $\nabla\times\mathbf{u} = \mathbf{0}$.
-But in this case the gradient of density is neglected in the Eulerian momentum equation in order to identify gradient terms.
-Indeed, from the Eulerian momentum equation for inviscid acoustic waves only
-\begin{eqnarray*}
-\rho\partial_t^2\mathbf{u} &=& -\nabla p\\
-\nabla (\partial_t^2\chi)&=& -(1/\rho)\nabla p \, .
-\end{eqnarray*}
-When $\nabla\rho \gg \mathbf{0}$ strictly speaking one cannot identify $\partial_t^2\chi$ with $p$.
-
-The solution to this is to work with pressure as the scalar unknown (\cite{God11}, equation 23),
-or even better with the time integral (the primitive) of pressure (see e.g. \cite{Eve81});
-this was the approach that was originally implemented in SPECFEM until 10 years ago or so, but \cite{ChCaVi03} showed that switching to a displacement potential
-or to a potential of $\rho \mathbf{u}$ enables one to couple the fluid and solid regions without iterating on the coupling condition in the time scheme,
-thus resulting in much easier and cheaper implementation; since at that time SPECFEM users were not studying media with strong local variations of density (in solid Earth
-geophysics such large local variations in fluid regions never happen), the SPECFEM developers permanently switched to such a displacement-based potential formulation.
-However, for media with strong local variations of density in a fluid region we should reconsider the issue and consider trying the pressure-based formulation again.
-
-Cases in which this can be problematic may include e.g. upper atmosphere studies: the density there is very small but displacements can become very large;
-in 1D atmosphere models $\frac{\nabla \rho}{\rho}$ is approximately constant and equal to -1/H, where H is the scaling height,
-which leads to $\frac{\nabla \rho}{\rho} \simeq$ 3.42*1e-5, i.e. very small; however displacements there can be large
-and thus it is not clear if $(1/\rho)\nabla\rho\times\mathbf{u}$ can be considered negligible there.
+\noindent
+However in practice for technical reasons it is approximated based on the sum of different Generalized Zener body mechanisms
+and thus the code outputs the band in which the approximation is very good, outside of that range it can be less accurate.
+The logarithmic center of that frequency band is the \texttt{ATTENUATION\_f0} parameter defined (in Hz) in input file \texttt{DATA/Par\_file}.

### 2016-05-01 doc/USER_MANUAL/08_doing_tomography.tex
made a sentence more precise in the users manual

@@
 Whatever misfit function you use for the tomographic inversion (you can find several
-examples in \citet{TrKoLi08} and \citet{TrTaLi05}), you will weight
+examples in \citet{TrKoLi08} and \citet{TrTaLi05}), you will weigh
 the sensitivity kernels with measurements. Furthermore, you will use
@@
 
-There are dozens of versions of gradient-based inversion
-algorithms that could alternatively be used. The tomographic inversion
-of \citet{TaLiMaTr09,TaLiMaTr2010} used SPECFEM3D\_Cartesian as well
+There are many other versions of gradient-based inversion
+algorithms that could alternatively be used (see e.g. \cite{ViOp09,MoChKoWa15} for a list).
+The tomographic inversion of \citet{TaLiMaTr09,TaLiMaTr2010} used SPECFEM3D\_Cartesian as well
 as several additional components which are also stored on the CIG

### 2016-05-01 doc/USER_MANUAL/15_informations_for_developers.tex

--- 2015-06-22 doc/USER_MANUAL/15_informations_for_developers.tex
@@
 They are useful to learn how the spectral-element method works, and how to write or modify a code to implement it.
-Also useful to test new ideas by modifying these simple codes to run some tests.\\
+Also useful to test new ideas by modifying these simple codes to run some tests.
+We also have a similar, even simpler, demo source code for the 2D case in directory\\
+utils/small\_SEM\_solver\_in\_Fortran\_without\_MPI\_to\_learn of the specfem2d package.\\
 

### 2016-05-01 doc/USER_MANUAL/notes_and_acknowledgement.tex

@@
-\addcontentsline{toc}{chapter}{Notes \& Acknowledgments}
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \chapter*{Notes \& Acknowledgments}
+\addcontentsline{toc}{chapter}{Notes \& Acknowledgments}
 
@@
 is a trademark of Advanced Visualization Systems, and ParaView \urlwithparentheses{http://www.paraview.org}
-is an open-source visualization platform.
+is an open-source visualization platform.\\
 

### 2016-05-19 doc/USER_MANUAL/05_running_the_solver.tex
added Hom Nath's paragraph to doc/USER_MANUAL/05_running_the_solver.tex

@@
 
+Elevation and burial are generally applicable to geographical regions.
+Burial is measured down from the top surface. For other problems in other
+fields (ultrasonic testing, medical imaging etc...),
+it may be confusing. We generally follow either of the following
+procedures for those kind of problems:
+
+Procedure 1:\\
+============\\
+
+- Always put the origin on the top of the model.\\
+
+- Let's say you want to place two receivers at (x1,y1,z1) and
+(x2,y2,z2). Your STATIONS file should look like:\\
+
+BONE  GR  y1  x1  0.00  -z1\\
+BONE  GR  y2  x2  0.00  -z2\\
+
+Depending on the geometry of the models it may not work in some cases.\\
+
+Procedure 2: (less recommended)\\
+============\\
+
+- Open setup/constants.h.in file and change the line
+logical, parameter :: USE\_SOURCES\_RECEIVERS\_Z = .false.\\
+\\
+to\\
+\\
+logical, parameter :: USE\_SOURCES\_RECEIVERS\_Z = .true.\\
+\\
+configure the package and compile.\\
+
+- Let's say you want to place two receivers at (x1,y1,z1) and
+(x2,y2,z2). Your STATIONS file should look like:\\
+\\
+BONE  GR  y1  x1  0.00  z1\\
+BONE  GR  y2  x2  0.00  z2\\
+\\
+The option set to .true. in 1 will discard the elevation and set
+burial as the $z$ coordinate.\\
+\\
+Third column is Y and Fourth is X due to the latitude/longitude convention.\\
+\\
+You can replace the station name "BONE" with any word of length less
+than 32, and the network name "GR" with any word of length less than
+8.\\
+\\
+You can always plot OUTPUT\_FILES/sr.vtk file in ParaView to check the
+source/receiver locations after your simulation.\\
+
 Solver output is provided in the \texttt{OUTPUT\_FILES} directory

### 2016-06-01 doc/USER_MANUAL/05_running_the_solver.tex
Add new option for external source time function: modification of Par_files and add some lines in manual

@@
 
+\vspace{1cm}
+
+ In addition to inbuild source time function, the solver can also use an external source time function defined by the user. This option can be  
+activated by setting   \texttt{EXTERNAL\_SOURCE\_FILE} to true in \texttt{Par\_File} and by adding the name of the file
+containing the source time function at the end of \texttt{FORCESOLUTION} or \texttt{CMTSOLUTION} files. The source time function file must 
+contain 2 columns with the current time and the amplitude of the soure function. The time step must be the same as used for the 
+simulation. Note  when the  \texttt{EXTERNAL\_SOURCE\_FILE} is set to false then the line with the external source time function file must not appear in 
+the files  \texttt{FORCESOLUTION} and \texttt{CMTSOLUTION} otherwise the solver will exit with an error.
+
+\vspace{1cm}
+
 The solver can calculate seismograms at any number of stations for

### 2016-06-01 src/shared/read_parameter_file.F90

@@ -243,26 +243,9 @@
 
   !! VM VM add possibility of external source file
   call read_value_logical(EXTERNAL_STF, 'EXTERNAL_SOURCE_FILE', ier)
-  if (ier /=0) then
-     write(IMAIN,*)
-     write(IMAIN,*)  ' USING INBUILT SOURCE '
-     write(IMAIN,*) 
-     EXTERNAL_STF=.false.
-  else
-     if (EXTERNAL_STF) then
-        write(IMAIN,*)
-        write(IMAIN,*)
-        write(IMAIN,*) ' USE EXTERNAL SOURCE FILE THEN  '
-        write(IMAIN,*) ' SET USE_RICKER_TIME_FUNCTION to .false.' 
-        write(IMAIN,*) 
-        write(IMAIN,*)
-        USE_RICKER_TIME_FUNCTION=.false.
-     else
-        write(IMAIN,*)
-        write(IMAIN,*) 'external source file will not use '
-        write(IMAIN,*)
-     end if
-  end if
+  !! VM VM to be consistent with external source file option turned to on
+  if (EXTERNAL_STF)  USE_RICKER_TIME_FUNCTION=.false.
+     
 
 #ifdef DEBUG_COUPLED
     include "../../../add_to_read_parameter_file_1.F90"
@@ -375,10 +358,14 @@
       if (ier == 0) icounter = icounter + 1
     enddo
     close(21)
-
-    if (mod(icounter,NLINES_PER_CMTSOLUTION_SOURCE) /= 0) &
-      stop 'Error total number of lines in CMTSOLUTION file should be a multiple of NLINES_PER_CMTSOLUTION_SOURCE'
-
+    if (.not. EXTERNAL_STF) then
+       if (mod(icounter,NLINES_PER_CMTSOLUTION_SOURCE) /= 0) &
+            stop 'Error total number of lines in CMTSOLUTION file should be a multiple of NLINES_PER_CMTSOLUTION_SOURCE'
+    else
+       !! VM VM in case of EXTERNAL_STF we have to read one additional line per source (the name of external source file) 
+       NSOURCES = icounter / (NLINES_PER_FORCESOLUTION_SOURCE+1)
+    end if
+       
     NSOURCES = icounter / NLINES_PER_CMTSOLUTION_SOURCE
     if (NSOURCES < 1) stop 'Error need at least one source in CMTSOLUTION file'
 

### 2016-06-01 src/specfem3D/compute_add_sources_acoustic.f90

@@ -42,7 +42,7 @@
                         station_name,network_name,adj_source_file,nrec_local,number_receiver_global, &
                         pm1_source_encoding,nsources_local,USE_FORCE_POINT_SOURCE, &
                         USE_RICKER_TIME_FUNCTION,SU_FORMAT,USE_TRICK_FOR_BETTER_PRESSURE,USE_SOURCE_ENCODING, &
-                        USE_LDDRK,istage,EXTERNAL_STF
+                        USE_LDDRK,istage,EXTERNAL_STF,user_source_time_function
 
   implicit none
 
@@ -103,6 +103,10 @@
   !equivalence (i2head,i4head,r4head)    ! share the same 240-byte-memory
   double precision :: hxir(NGLLX), hpxir(NGLLX), hetar(NGLLY), hpetar(NGLLY),hgammar(NGLLZ), hpgammar(NGLLZ)
 
+! VM VM to know if we used the source in this domain 
+  integer :: source_is_in_this_domain,source_is_in_this_domain_all
+
+  source_is_in_this_domain=0
 ! plotting source time function
   if (PRINT_SOURCE_TIME_FUNCTION .and. .not. phase_is_inner) then
     ! initializes total
@@ -123,7 +127,7 @@
         if (ispec_is_inner(ispec) .eqv. phase_is_inner) then
 
           if (ispec_is_acoustic(ispec)) then
-
+             source_is_in_this_domain=1
             if (USE_LDDRK) then
               time_source_dble = dble(it-1)*DT+dble(C_LDDRK(istage))*DT-t0-tshift_src(isource)
             else
@@ -408,8 +412,10 @@
     endif ! nadj_rec_local > 0
   endif
 
+  call sum_all_i( source_is_in_this_domain, source_is_in_this_domain_all)
+  call bcast_all_singlei(source_is_in_this_domain_all)  
   ! master prints out source time function to file
-  if (PRINT_SOURCE_TIME_FUNCTION .and. phase_is_inner) then
+  if (PRINT_SOURCE_TIME_FUNCTION .and. source_is_in_this_domain_all > 0) then
     time_source = (it-1)*DT - t0
     call sum_all_cr(stf_used_total,stf_used_total_all)
     if (myrank == 0) write(IOSTF,*) time_source,stf_used_total_all
@@ -620,7 +626,8 @@
                         xigll,yigll,zigll,xi_receiver,eta_receiver,gamma_receiver,&
[…]

### 2016-06-15 doc/USER_MANUAL/04_creating_databases.tex
added doc/USER_MANUAL/figures/how_to_use_a_transition_mesh_layer_to_stabilize_PML.png and some explanations

@@
 in time (PML being weakly unstable from a mathematical point of view, very deformed mesh elements
-inside the PMLs can trigger instabilities much more quickly).
+inside the PMLs can trigger instabilities much more quickly). \red{We have two utilities in directory utils/CPML that do that automatically
+and that are very fast}. To stabilize PMLs it also helps to add a transition layer of geometrically-regular non-PML elements, in which attenuation is also 
+turned off (i.e. $Q_\kappa = Q_\mu = 9999$ in that layer), as in the red layer of Figure~\ref{fig:mesh_extrusion}.
+Our tools in directory utils/CPML implement that transition layer automatically.
 
@@
 \noindent \begin{centering}
+\includegraphics[width=4.5in]{figures/how_to_use_a_transition_mesh_layer_to_stabilize_PML.png}
+\par\end{centering}
+\caption{Mesh extrusion for PML (green elements) and a non-PML stabilization layer (red elements).}
+\label{fig:mesh_extrusion}
+\end{figure}
+%%
+
+%%
+\begin{figure}[htbp]
+\noindent \begin{centering}
 \includegraphics[width=6in]{figures/how_to_use_PML_when_a_tomographic_velocity_model_is_used.png}
 \par\end{centering}
-
 \caption{How to modify your external 3D velocity and density model in order
@@
 boundary conditions (but such conditions are significantly less efficient).}
-
 \label{fig:modify_external_velocity_model_to_use_PML}
 \end{figure}
-
 %%

### 2016-06-16 doc/USER_MANUAL/03_mesh_generation.tex
added notes about attenuation
updates doc

@@
 Please feel free to contribute your own example to this package by
-contacting the CIG Computational Seismology Mailing List \urlwithparentheses{cig-seismo@geodynamics.org}.
+contacting the CIG Computational Seismology Mailing List \urlwithparentheses{cig-seismo@geodynamics.org}.\\
 
-In some cases, to re-create the meshes for the examples given, just type "claro ./create\_mesh.py" or similar
-from the command line ("claro" is the command to run CUBIT from the command line).
+In some cases, to re-create the meshes for the examples given, just type
+\begin{verbatim}
+claro ./create_mesh.py
+\end{verbatim}
+or similar from the command line (\texttt{claro} is the command to run CUBIT from the command line).
 
-IMPORTANT note: in order to correctly set up GEOCUBIT and run the examples, please read the file called "EXAMPLES/README";
+{\it IMPORTANT:} In order to correctly set up GEOCUBIT and run the examples, please read the file called \texttt{EXAMPLES/README};
 in particular, please make sure you correctly set up the Python paths as indicated in that file.
+
+Concerning the script \texttt{create\_mesh.py}, you may find the option \texttt{use\_explicit} which chooses how to assign the material properties for the volume (or domain):
+\begin{itemize}
+\item[(1)] one way is to explicitly assign block attributes to the block of the corresponding volume, by commands like:
+\begin{verbatim}
+cubit.cmd('block '+str(id_block)+' attribute index 2 2800') # vp
+\end{verbatim}
+This is done when using \texttt{use\_explicit = 1} in the script. The final command:
+\begin{verbatim}
+cubit2specfem3d.export2SPECFEM3D('MESH/')
+\end{verbatim}
+will then create the corresponding \texttt{MESH/nummaterial\_velocity\_file} for all such defined block volumes/domains.
+
+\item[(2)] the other option with \texttt{use\_explicit = 0} is to let GEOCUBIT deal with a dummy entry and then overwrite the \texttt{nummaterial\_velocity\_file} at the end with a corresponding command section:
+\begin{verbatim}
+f = open(nummaterial_velocity_file,'w')
+\end{verbatim}
+This second way is chosen by default because GEOCUBIT can handle partitions for several processes and glues everything together automatically.
+Thus, it adds some more sophistication when the model gets more complicated than just this single volume example.
+\end{itemize}
+You will find out by experimenting what is easier for your case.
 
@@
 by a unique name. The name convention of the material is to start
-with either 'elastic' or 'acoustic'. It must be then followed by a
-unique identifier, e.g. 'elastic 1', 'elastic 2', etc. The additional
+with either \textbf{'elastic'} or \textbf{'acoustic'}. It must be then followed by a
+unique identifier, e.g. \textbf{'elastic 1'}, \textbf{'elastic 2'}, etc. The additional
 attributes to the block define the material description.
@@
 separate values for bulk and shear attenuation, $Q_{\kappa}$ and $Q_{\mu}$ respectively.
+Note that Qmu is always equal to Qs, but Qkappa is in general not equal to Qp.
+To convert one to the other see doc/note\_on\_Qkappa\_versus\_Qp.pdf and utils/attenuation/conversion\_from\_Qkappa\_Qmu\_to\_Qp\_Qs\_from\_Dahlen\_Tromp\_959\_960.f90.
 
@@
 When you have an existing CUBIT (or similar) mesh stored in SPECFEM3D
-format, i.e., if you have existing \textquotedbl{}nodes\_coords\_file\textquotedbl{}
-and \textquotedbl{}mesh\_file\textquotedbl{} files but do not know
+format, i.e., if you have existing \texttt{nodes\_coords\_file}
+and \texttt{mesh\_file} files but do not know
 how to assign CPML flags to them, we have created a small serial Fortran
 program that will do that automatically for you, i.e., which will
-create the \textquotedbl{}absorbing\_cpml\_file\textquotedbl{} for
-you. That program is utils/CPML/convert\_external\_layers\_of\_a\_given\_mesh\_to\_CPML\_layers.f90,
+create the \texttt{absorbing\_cpml\_file} for
+you. That program is: \\
+\texttt{utils/CPML/convert\_external\_layers\_of\_a\_given\_mesh\_to\_CPML\_layers.f90},\\
 and a small Makefile is provided in that directory (utils/CPML).
@@
 CUBIT (or similar) mesh that this code will read in SPECFEM3D format
-from files \textquotedbl{}nodes\_coords\_file\textquotedbl{} and \textquotedbl{}mesh\_file\textquotedbl{}
+from files \texttt{nodes\_coords\_file} and \texttt{mesh\_file}
 you have created layers of elements that constitute a layer of constant
@@
 both set \texttt{FULL\_ATTENUATION\_SOLID} to \texttt{.false.} in
-the \texttt{Par\_file} and set \texttt{Qkappa} to 9999 here. %%magnoni
+the \texttt{Par\_file} and set \texttt{Qkappa} to 9999 here.
+Note that Qmu is always equal to Qs, but Qkappa is in general not equal to Qp.
+To convert one to the other see doc/note\_on\_Qkappa\_versus\_Qp.pdf and utils/attenuation/conversion\_from\_Qkappa\_Qmu\_to\_Qp\_Qs\_from\_Dahlen\_Tromp\_959\_960.f90.
 

### 2016-07-11 doc/USER_MANUAL/09_noise_simulations.tex
updates doc

@@
 are provided to help you to generate the necessary file:
-\texttt{EXAMPLES/noise\_tomography/NOISE\_TOMOGRAPHY.m} (main program)\\
-\texttt{EXAMPLES/noise\_tomography/PetersonNoiseModel.m}
+\begin{verbatim}
+EXAMPLES/noise_tomography/NOISE_TOMOGRAPHY.m  (main program)
+EXAMPLES/noise_tomography/PetersonNoiseModel.m
+\end{verbatim}
 
@@
 
-\texttt{DT} is given in \texttt{Par\_file}, but \texttt{NSTEP} is
-NOT the one specified in \texttt{Par\_file}. Instead, you have to
+\texttt{DT} is given in \texttt{Par\_file}, but {\color{red} \texttt{NSTEP} is
+NOT the one specified in \texttt{Par\_file}}. {\bf Instead, you have to
 feed $2~\mathrm{NSTEP}-1$ to account for the doubled length of cross
-correlations. \texttt{Tmin} and \texttt{Tmax} correspond to the period
+correlations}. \texttt{Tmin} and \texttt{Tmax} correspond to the period
 range you are interested in, whereas \texttt{NOISE\_MODEL} denotes
-the noise model you will be using. Details can be found in the Matlab
-script.
+the noise model you will be using 
+(\texttt{'NLNM'} for New Low Noise Model or \texttt{'NHNM'} for New High Noise Model). 
+Details can be found in the Matlab script.
 
@@
 S_squared should be put into directory:
-OUTPUT_FILES/NOISE_TOMOGRAPHY/ in the SPECFEM3D Cartesian package
+./NOISE_TOMOGRAPHY/ in the SPECFEM3D Cartesian package
 \end{verbatim}
@@
 \item Create a new directory in the SPECFEM3D Cartesian package, name it
-as \texttt{OUTPUT\_FILES/NOISE\_TOMOGRAPHY/}. We will add some parameter
+as \texttt{./NOISE\_TOMOGRAPHY/}. We will add some parameter
 files later in this folder.
-\item Put the Matlab-generated-file \texttt{S\_squared} in \texttt{OUTPUT\_FILES/NOISE\_TOMOGRAPHY/}.
 
-That's to say, you will have a file \texttt{OUTPUT\_FILES/NOISE\_TOMOGRAPHY/S\_squared}
-in the SPECFEM3D Cartesian package.
+\item Put the Matlab-generated-file \texttt{S\_squared} in \texttt{./NOISE\_TOMOGRAPHY/}.
 
-\item Create a file called \texttt{OUTPUT\_FILES/NOISE\_TOMOGRAPHY/irec\_master\_noise}.
-Note that this file is located in directory \texttt{OUTPUT\_FILES/NOISE\_TOMOGRAPHY/}
+That's to say, you will have a file \texttt{./NOISE\_TOMOGRAPHY/S\_squared}
+in the example provided in the SPECFEM3D Cartesian package.
+
+\item Create a file called \texttt{./NOISE\_TOMOGRAPHY/irec\_master\_noise}.
+Note that this file is located in directory \texttt{./NOISE\_TOMOGRAPHY/}
 as well. In general, all noise simulation related parameter files
@@
 
-Note that in regional simulations, the \texttt{DATA/STATIONS} might
+Note that in some simulations, the \texttt{DATA/STATIONS} might
 contain receivers which are outside of our computational domains.
@@
 
-\item Create a file called \texttt{OUTPUT\_FILES/NOISE\_TOMOGRAPHY/nu\_master}.
+\item Create a file called \texttt{./NOISE\_TOMOGRAPHY/nu\_master}.
 This file holds three numbers, forming a (unit) vector. It describes
@@
 should be 0, 0 and 1.
-\item Describe the noise direction and distributions in \texttt{src/shared/noise\_tomography.f90}.
+
+\item Describe the noise direction and distributions in \texttt{src/specfem3d/noise\_tomography.f90}.
 Search for a subroutine called \texttt{noise\_distribution\_direction}
@@
 \end{itemize}
-Note that it is better to run the three steps continuously within
-the same job, otherwise you have to collect the saved surface movies
-from the old nodes/CPUs to the new nodes/CPUs. This process varies
+
+It is better to run the three steps continuously within
+the same job on a cluster, otherwise you have to collect the saved surface movies
+from the old nodes to the new nodes. This process varies
 from cluster to cluster and thus cannot be discussed here. Please

### 2016-07-12 doc/USER_MANUAL/11_graphics.tex
updates surface movie and shakemap routines

@@
 
-\subsection{Movie Surface}
+\subsection{Movie Surface and Shakemaps}
 
@@
 main parameter file \texttt{Par\_file} that control the output of
-these movie data files (see section \ref{cha:Main-Parameter} for
-details):
-\texttt{\small NTSTEP\_BETWEEN\_FRAMES} to set the
-timesteps between frames,
-\texttt{\small SAVE\_DISPLACEMENT}
-to save displacement instead of velocity,
-\texttt{\small USE\_HIGHRES\_FOR\_MOVIES}
-to save values at all GLL point instead of element edges. In order
-to output additionally shakemaps, you would set the parameter
-\texttt{\small CREATE\_SHAKEMAP}
-to \texttt{\small .true.}.
+these movie data files, please see section \ref{cha:Main-Parameter} for
+more details. Especially, the following parameters affect the output:
+\begin{itemize}
+\item \texttt{\small SAVE\_DISPLACEMENT}: to save displacement instead of velocity,
+\item \texttt{\small NTSTEP\_BETWEEN\_FRAMES}: to set the timesteps between frames,
+\item \texttt{\small USE\_HIGHRES\_FOR\_MOVIES}: to save values at all GLL point instead of element edges.
+\end{itemize}
 
-The files are in a fairly complicated binary format, but there is
+You can also output shakemaps independant of the \texttt{MOVIE\_SURFACE} flag.
+Shakemaps show the peak-ground values of the simulation.
+For creating shakemaps (at the end of the simulation), you would set the parameter
+\texttt{\small CREATE\_SHAKEMAP} to \texttt{\small .true.}.
+For shakemaps, both the parameters \texttt{\small MOVIE\_TYPE} and \texttt{\small USE\_HIGHRES\_FOR\_MOVIES} affect the output. The following setting has an additional affect:
+\begin{itemize}
+\item \texttt{\small MOVIE\_TYPE}: if set to $1$, the horizontal peak-ground values of displacement/velocity/acceleration are output.
+if set to $2$, the maximum length of the particle displacement/velocity/acceleration vector is output.
+Please be aware that these peak-ground values can differ from each other.
+\end{itemize}
+
+The movie and shakemap output files are in a custom binary format, but there is
 a program provided to convert the output into more user friendly formats:

### 2016-07-12 src/specfem3D/detect_mesh_surfaces.f90

@@ -41,16 +41,15 @@
           iglob_is_surface_external_mesh(NGLOB_AB),stat=ier)
   if (ier /= 0) stop 'error allocating array for mesh surface'
 
-! determines model surface
-  if (.not. RECEIVERS_CAN_BE_BURIED .or. MOVIE_TYPE == 2) then
-
+  ! determines model surface
+  if (.not. RECEIVERS_CAN_BE_BURIED .or. MOVIE_SURFACE .or. CREATE_SHAKEMAP) then
     ! returns surface points/elements
     ! in ispec_is_surface_external_mesh / iglob_is_surface_external_mesh and
-    ! number of faces in nfaces_surface_ext_mesh
+    ! number of faces in nfaces_surface
     call detect_surface(NPROC,NGLOB_AB,NSPEC_AB,ibool,&
                       ispec_is_surface_external_mesh, &
                       iglob_is_surface_external_mesh, &
-                      nfaces_surface_ext_mesh, &
+                      nfaces_surface, &
                       num_interfaces_ext_mesh, &
                       max_nibool_interfaces_ext_mesh, &
                       nibool_interfaces_ext_mesh, &
@@ -58,55 +57,48 @@
                       ibool_interfaces_ext_mesh)
   endif
 
-! takes cross-section surfaces instead
-  if (MOVIE_TYPE == 2 .and. PLOT_CROSS_SECTIONS) then
-    call detect_surface_cross_section(NPROC,NGLOB_AB,NSPEC_AB,ibool,&
-                            ispec_is_surface_external_mesh, &
-                            iglob_is_surface_external_mesh, &
-                            nfaces_surface_ext_mesh, &
-                            num_interfaces_ext_mesh, &
-                            max_nibool_interfaces_ext_mesh, &
-                            nibool_interfaces_ext_mesh, &
-                            my_neighbours_ext_mesh, &
-                            ibool_interfaces_ext_mesh,&
-                            CROSS_SECTION_X,CROSS_SECTION_Y,CROSS_SECTION_Z, &
-                            xstore,ystore,zstore,myrank)
+  ! takes cross-section surfaces instead
+  if (MOVIE_SURFACE .or. CREATE_SHAKEMAP) then
[…]

### 2016-07-17 doc/USER_MANUAL/04_creating_databases.tex
Buildbot ran the code beautifier.
added a sentence about free surfaces versus rigid surfaces
clarified a couple of print statements about attenuation, and updated the users manual accordingly

@@
 \texttt{constants.h} found in subdirectory \texttt{src/shared/}.
+The code uses a constant $Q$ quality factor, write(IMAIN,*) "but approximated based on a series of Zener standard linear solids (SLS).
+The approximation is thus performed in a given frequency band determined based on that \texttt{ATTENUATION\_f0\_REFERENCE} reference frequency.
 \item [{\texttt{FULL\_ATTENUATION\_SOLID}}] Set to \texttt{.true.} to add
@@
 by default constitutes a free surface of the model.
+\item [{\texttt{BOTTOM\_FREE\_SURFACE}}]
+When STACEY\_ABSORBING\_CONDITIONS is set to .true. :
+absorbing conditions are defined in xmin, xmax, ymin, ymax and zmin
+this option BOTTOM\_FREE\_SURFACE can be set to .true. to
+make zmin free surface instead of absorbing condition.
 \item [{\texttt{CREATE\_SHAKEMAP}}] Set this flag to \texttt{.true.} to

### 2016-09-17 doc/USER_MANUAL/04_creating_databases.tex
suppressed FULL_ATTENUATION_SOLID from the users manual

@@
 The approximation is thus performed in a given frequency band determined based on that \texttt{ATTENUATION\_f0\_REFERENCE} reference frequency.
-\item [{\texttt{FULL\_ATTENUATION\_SOLID}}] Set to \texttt{.true.} to add
-$Q_{\kappa}$ attenuation to the simulations in addition to $Q_{\mu}$
-attenuation when the \texttt{ATTENUATION} flag above is on. Should
-be off in almost all cases for seismic wave propagation, since $Q_{\kappa}$
-is almost always negligible in Earth models. However, this parameter
-can be useful for instance for ocean acoustics simulations, when $Q_{\kappa}$
-is not negligible in the solid layer at the sea bottom.
 Note that Qmu is always equal to Qs, but Qkappa is in general not equal to Qp.

### 2016-10-25 doc/USER_MANUAL/01_introduction.tex
better BibTeX

@@
 
+You can find all the references below in \BibTeX format in file \texttt{doc/USER\_MANUAL/bibliography.bib}.\\
+
 If you use SPECFEM3D Cartesian for your own research, please cite
@@
 %
-If you use the attenuation (anelastic/viscoelastic) capabilities of the code, please cite at least one article
+If you use the UNDO\_ATTENUATION option of the code in order to produce full anelastic/viscoelastic sensitivity kernels, please cite at least one article
+written by the developers of the package, for instance (and in particular):
+%
+\begin{itemize}
+\item \cite{KoXiBoPeSaLiTr16}.
+\end{itemize}
+%
+More generally, if you use the attenuation (anelastic/viscoelastic) capabilities of the code, please cite at least one article
 written by the developers of the package, for instance:
@@
 \begin{itemize}
+\item \cite{KoXiBoPeSaLiTr16},
 \item \cite{BlKoChLoXi16}.

### 2016-10-25 doc/USER_MANUAL/03_mesh_generation.tex

@@
 \texttt{src/shared/}.
+The code uses a constant $Q$ quality factor, write(IMAIN,*) "but approximated based on a series of Zener standard linear solids (SLS).
+The approximation is thus performed in a given frequency band determined based on that \texttt{ATTENUATION\_f0\_REFERENCE} reference frequency.
 \item [{anisotropic\_flag}] Flag describing the anisotropic model to use
@@
 $m\, s^{-1}$, \texttt{Q} the quality factor and \texttt{anisotropy\_flag}
-an identifier for anisotropic models. Note that the \texttt{Qkappa}
-value is ignored by the code unless \texttt{FULL\_ATTENUATION\_SOLID}
-is set. Note also that both \texttt{Qkappa} and \texttt{Qmu} are ignored
+an identifier for anisotropic models.
+Note that both \texttt{Qkappa} and \texttt{Qmu} are ignored
 by the code unless \texttt{ATTENUATION} is set. If you want a model
@@
 \texttt{.false.} in the \texttt{Par\_file} and set \texttt{Qmu} to
-9999 here. If you want a model with no \texttt{Qkappa} attenuation,
-both set \texttt{FULL\_ATTENUATION\_SOLID} to \texttt{.false.} in
-the \texttt{Par\_file} and set \texttt{Qkappa} to 9999 here.
+9999 here. If you want a model with no \texttt{Qkappa} attenuation, set \texttt{Qkappa} to 9999 here.
 Note that Qmu is always equal to Qs, but Qkappa is in general not equal to Qp.

### 2016-10-25 doc/USER_MANUAL/06_fault_sources.tex
Modified contents in user manual chapter 6.
Buildbot ran the code beautifier.

--- 2015-09-09 doc/USER_MANUAL/06_fault_sources.tex
@@
 creating a tiny opening of the fault (Figure \ref{fig:examples.splitnodes},
-\ref{fig:examples.splitnodes-surfacetrace}). Note that only the interior
+\ref{fig:examples.splitnodes-surfacetrace}). Note that the opening distance
+should not be too small especially in single precision simulations
+(For example, if your fault is at $y=y_0$ plane, then the opening distance
+should be at least $10^{-6}y_0$ since single precision can ensure precision
+within only 7 digits). Note that only the interior
 of the fault must be opened, its edges must remain closed (except
@@
 
-\&\textbf{DIST2D} shape='$\mathtt{square}$', val, xc, yc, zc, l /
+\&\textbf{DIST2D} shapeval='$\mathtt{square}$', val, xc, yc, zc, l /
 
@@
 
-\&\textbf{DIST2D} shape='$\mathtt{rectangle}$', val, xc, yc, zc,
+\&\textbf{DIST2D} shapeval='$\mathtt{rectangle}$', val, xc, yc, zc,
 lx, ly, lz /
@@
 
-\&\textbf{DIST2D} shape='$\mathtt{rectangle-taper}$', val, valh,
+\&\textbf{DIST2D} shapeval='$\mathtt{rectangle-taper}$', val, valh,
 xc, yc, zc, lx, ly, lz /
@@
 
-\&\textbf{DIST2D} shape='$\mathtt{circular}$', val, xc, yc, zc, r
+\&\textbf{DIST2D} shapeval='$\mathtt{circular}$', val, xc, yc, zc, r
 /

### 2017-01-20 doc/USER_MANUAL/authors.tex

@@
 Qinya Liu,
+Youshan Liu,
 David Luet,

### 2017-01-23 doc/USER_MANUAL/02_getting_started.tex
added Daniel Peter's instructions for CUDA8

@@
 \end{verbatim}
+
+When compiling for GPU cards i.e. using CUDA, you can use:
+
+\texttt{./configure --with-cuda}
+
+for CUDA version 4, or
+
+\texttt{./configure --with-cuda=cuda5}
+
+for CUDA version 5, or
+
+\texttt{./configure --with-cuda=cuda6}
+
+for CUDA version 6, or
+
+\texttt{./configure --with-cuda=cuda8}
+
+for CUDA version 8.
+
+Before CUDA version 5, one version supported basically one new architecture and needed a different kind of compilation.
+Since version 5, the compilation has stayed the same, but newer versions supported newer architectures.
+However at the moment, we still have one version linked to one specific architecture:
+
+- CUDA 4 for Fermi,
+- CUDA 5 for Tesla K20
+- CUDA 6 for Tesla K80
+- CUDA 8 for Pascal P100.
+
+So even if you have the new CUDA toolkit version 8, but you want to run on say a K20 GPU, then you would still configure with:
+
+\texttt{./configure --with-cuda=cuda5}
+
+The compilation should work and the cuda5 setting chooses the right architecture (-gencode=arch=compute\_35,code=sm\_35 for K20 cards).
+
+So, type
+
+\texttt{./configure --with-cuda=cuda8}
+
+for Pascal P100 boards.
 

### 2017-02-01 doc/USER_MANUAL/06_fault_sources.tex
Modify fault solver part of the manual.

@@
 is parallel to the along-dip direction pointing downwards, and \textquotedbl{}\emph{strike}\textquotedbl{}
-is the horizontal along-strike vector such that the system is right-handed.\\
+is the horizontal along-strike vector such that the system is right-handed. In places where the fault plane is horizontal,
+we define the alone strike direction to be (1,0,0).\\
 

### 2017-03-18 doc/USER_MANUAL/01_introduction.tex
added a sentence about how to configure the code on CRAY machines, and added an announcement for the fact that SPECFEM3D can now perform gravity modeling in addition to seismic modeling

@@
 \chapter{Introduction}
+
+\red{\Large ANNOUNCEMENT: SPECFEM3D can now perform gravity field calculations in addition (or instead of) seismic wave propagation only.
+See flag \texttt{GRAVITY\_INTEGRALS} in file \texttt{setup/constants.h.in}. Please also refer to \url{http://komatitsch.free.fr/preprints/GJI_Martin_gravimetry_2017.pdf}.
+And yes, that is the reason why I added a falling apple on the cover of the manual :-)}\\
 

### 2017-03-18 doc/USER_MANUAL/02_getting_started.tex
added a sentence about how to configure the code on CRAY machines, and added an announcement for the fact that SPECFEM3D can now perform gravity modeling in addition to seismic modeling

@@
 }
+
+If you have problems configuring the code on a Cray machine, i.e. for instance if you get an error message from the \texttt{configure} script, try exporting these two variables:
+\texttt{MPI\_INC=\${CRAY\_MPICH2\_DIR}/include and FCLIBS=" "}, and for more details if needed you can refer to the \texttt{utils/Cray\_compiler\_information} directory.
 

### 2017-04-10 doc/USER_MANUAL/02_getting_started.tex
better sentence for Intel ifort

@@
 \item [{\texttt{Intel ifort compiler}}] See if you need to add \texttt{-assume byterecl} for your machine.
-\textbf{In the case of that compiler, we have noticed that versions dot zero sometimes have bugs or issues that can lead to wrong results when running the code, thus we \emph{strongly} recommend using versions dot one or above (for instance version 13.1 instead of 13.0, version 14.1 instead of 14.0 and so on)}.
+\textbf{In the case of that compiler, we have noticed that initial release versions sometimes have bugs or issues that can lead to wrong results when running the code, thus we \emph{strongly} recommend using a version for which at least one service pack or update has been installed.}
 \item [{\texttt{IBM compiler}}] See if you need to add \texttt{-qsave}

### 2017-05-24 doc/USER_MANUAL/authors.tex
- Enhanced the cavity option. Now the multiple cavities can be added.

--- 2015-07-13 doc/USER_MANUAL/authors.tex
@@
 Percy Galvez,
+Hom Nath Gharti,
 Dominik G\"oddeke,

### 2017-06-06 doc/USER_MANUAL/authors.tex
added Etienne Bachmann to the list of authors of the 3D code

--- 2015-07-13 doc/USER_MANUAL/authors.tex
@@
 Jean-Paul (Pablo) Ampuero,
+\'Etienne Bachmann,
 Kangchen Bai,

### 2017-06-22 doc/USER_MANUAL/01_introduction.tex
switched the license to GNU version 3 (now that we use pyCMT3D, which is LGPL v3, the combination automatically becomes GNU version 3). Also added FWI inversion information to the users manual.

@@
 
-\red{\Large ANNOUNCEMENT: SPECFEM3D can now perform gravity field calculations in addition (or instead of) seismic wave propagation only.
+\red{\Large ANNOUNCEMENT: SPECFEM3D can now perform full waveform inversion (FWI), i.e. invert for models in an iterative fashion, and it can also perform source inversions in a constant model;
+please refer to the two new directories, \texttt{inverse\_problem\_for\_model} and \texttt{inverse\_problem\_for\_source}, and the README files they contain.
+For FWI inversions for the model, also refer to the new examples provided in
+directory \texttt{EXAMPLES}.}\\
+
+\red{OTHER ANNOUNCEMENT: SPECFEM3D can now perform gravity field calculations in addition (or instead of) seismic wave propagation only.
 See flag \texttt{GRAVITY\_INTEGRALS} in file \texttt{setup/constants.h.in}. Please also refer to \url{http://komatitsch.free.fr/preprints/GJI_Martin_gravimetry_2017.pdf}.

### 2017-06-22 doc/USER_MANUAL/D_license.tex
switched the license to GNU version 3 (now that we use pyCMT3D, which is LGPL v3, the combination automatically becomes GNU version 3). Also added FWI inversion information to the users manual.

@@
-\chapter{License }\label{cha:License}
+\chapter{License}\label{cha:License}
 
-\textbf{GNU GENERAL PUBLIC LICENSE Version 2, June 1991. Copyright
-(C) 1989, 1991 Free Software Foundation, Inc. 59 Temple Place, Suite
-330, Boston, MA 02111-1307 USA} \\
+\begin{center}
+
+\textbf{GNU GENERAL PUBLIC LICENSE}
+
+\textbf{Version 3, 29 June 2007}
+
+{\parindent 0in
+
+Copyright \copyright\  2007 Free Software Foundation, Inc. \texttt{http://fsf.org/}
+
+\bigskip
 Everyone is permitted to copy and distribute verbatim copies of this
-license document, but changing it is not allowed.
 
-\section*{Preamble}
+license document, but changing it is not allowed.}
 
-The licenses for most software are designed to take away your freedom
-to share and change it. By contrast, the GNU General Public License
-is intended to guarantee your freedom to share and change free software
--- to make sure the software is free for all its users. This General
-Public License applies to most of the Free Software Foundation's software
-and to any other program whose authors commit to using it. (Some other
-Free Software Foundation software is covered by the GNU Library General
-Public License instead.) You can apply it to your programs, too.
+\end{center}
 
-When we speak of free software, we are referring to freedom, not price.
-Our General Public Licenses are designed to make sure that you have
-the freedom to distribute copies of free software (and charge for
-this service if you wish), that you receive source code or can get
-it if you want it, that you can change the software or use pieces
-of it in new free programs; and that you know you can do these things.
+\subsection*{Preamble}
+The GNU General Public License is a free, copyleft license for
+software and other kinds of works.
 
-To protect your rights, we need to make restrictions that forbid anyone
-to deny you these rights or to ask you to surrender the rights. These
-restrictions translate to certain responsibilities for you if you
-distribute copies of the software, or if you modify it.
+The licenses for most software and other practical works are designed
+to take away your freedom to share and change the works.  By contrast,
+the GNU General Public License is intended to guarantee your freedom to
+share and change all versions of a program--to make sure it remains free
+software for all its users.  We, the Free Software Foundation, use the
+GNU General Public License for most of our software; it applies also to
+any other work released this way by its authors.  You can apply it to
+your programs, too.
 
-For example, if you distribute copies of such a program, whether gratis
-or for a fee, you must give the recipients all the rights that you
-have. You must make sure that they, too, receive or can get the source
-code. And you must show them these terms so they know their rights.
+When we speak of free software, we are referring to freedom, not
+price.  Our General Public Licenses are designed to make sure that you
+have the freedom to distribute copies of free software (and charge for
+them if you wish), that you receive source code or can get it if you
+want it, that you can change the software or use pieces of it in new
+free programs, and that you know you can do these things.
 
-We protect your rights with two steps:
-\begin{enumerate}
-\item Copyright the software, and
-\item Offer you this license which gives you legal permission to copy, distribute
-and/or modify the software.
-\end{enumerate}
-Also, for each author's protection and ours, we want to make certain
-that everyone understands that there is no warranty for this free
-software. If the software is modified by someone else and passed on,
-we want its recipients to know that what they have is not the original,
-so that any problems introduced by others will not reflect on the
-original authors' reputations.
+To protect your rights, we need to prevent others from denying you
+these rights or asking you to surrender the rights.  Therefore, you have
+certain responsibilities if you distribute copies of the software, or if
+you modify it: responsibilities to respect the freedom of others.
 
-Finally, any free program is threatened constantly by software patents.
-We wish to avoid the danger that redistributors of a free program
-will individually obtain patent licenses, in effect making the program
-proprietary. To prevent this, we have made it clear that any patent
-must be licensed for everyone's free use or not licensed at all.
+For example, if you distribute copies of such a program, whether
+gratis or for a fee, you must pass on to the recipients the same
+freedoms that you received.  You must make sure that they, too, receive
+or can get the source code.  And you must show them these terms so they
+know their rights.
 
-The precise terms and conditions for copying, distribution and modification
-follow.
+Developers that use the GNU GPL protect your rights with two steps:
+(1) assert copyright on the software, and (2) offer you this License
+giving you legal permission to copy, distribute and/or modify it.
 
-\section*{GNU GENERAL PUBLIC LICENSE TERMS AND CONDITIONS FOR COPYING, DISTRIBUTION
-AND MODIFICATION }
-\begin{itemize}
-\item 0.This License applies to any program or other work which contains
-a notice placed by the copyright holder saying it may be distributed
-under the terms of this General Public License. The ``Program''
-below refers to any such program or work, and a ``work based on the
-Program'' means either the Program or any derivative work under copyright
-law: that is to say, a work containing the Program or a portion of
-it, either verbatim or with modifications and/or translated into another
-language. (Hereinafter, translation is included without limitation
-in the term ``modification.'') Each licensee is addressed as ``you.''\\
- \\
- Activities other than copying, distribution and modification are
-not covered by this License; they are outside its scope. The act of
-running the Program is not restricted, and the output from the Program
-is covered only if its contents constitute a work based on the Program
-(independent of having been made by running the Program). Whether
-that is true depends on what the Program does.\end{itemize}
-\begin{enumerate}
-\item You may copy and distribute verbatim copies of the Program's source
-code as you receive it, in any medium, provided that you conspicuously
-and appropriately publish on each copy an appropriate copyright notice
-and disclaimer of warranty; keep intact all the notices that refer
-to this License and to the absence of any warranty; and give any other
-recipients of the Program a copy of this License along with the Program.
+For the developers' and authors' protection, the GPL clearly explains
+that there is no warranty for this free software.  For both users' and
+authors' sake, the GPL requires that modified versions be marked as
+changed, so that their problems will not be attributed erroneously to
+authors of previous versions.
 
-You may charge a fee for the physical act of transferring a copy,
-and you may at your option offer warranty protection in exchange for
-a fee.
+Some devices are designed to deny users access to install or run
+modified versions of the software inside them, although the manufacturer
+can do so.  This is fundamentally incompatible with the aim of
+protecting users' freedom to change the software.  The systematic
+pattern of such abuse occurs in the area of products for individuals to
+use, which is precisely where it is most unacceptable.  Therefore, we
+have designed this version of the GPL to prohibit the practice for those
+products.  If such problems arise substantially in other domains, we
+stand ready to extend this provision to those domains in future versions
+of the GPL, as needed to protect the freedom of users.
 
-\item You may modify your copy or copies of the Program or any portion of
-it, thus forming a work based on the Program, and copy and distribute
-such modifications or work under the terms of Section 1 above, provided
-that you also meet all of these conditions:
+Finally, every program is threatened constantly by software patents.
+States should not allow patents to restrict development and use of
+software on general-purpose computers, but in those that do, we wish to
+avoid the special danger that patents applied to a free program could
+make it effectively proprietary.  To prevent this, the GPL assures that
+patents cannot be used to render the program non-free.
+
+The precise terms and conditions for copying, distribution and
+modification follow.
+
+\begin{center}
+{\Large \sc Terms and Conditions}
+\end{center}
 
 \begin{enumerate}
-\item You must cause the modified files to carry prominent notices stating
-that you changed the files and the date of any change.
-\item You must cause any work that you distribute or publish, that in whole
-or in part contains or is derived from the Program or any part thereof,
-to be licensed as a whole at no charge to all third parties under
-the terms of this License.
-\item If the modified program normally reads commands interactively when
-run, you must cause it, when started running for such interactive
-use in the most ordinary way, to print or display an announcement
-including an appropriate copyright notice and a notice that there
-is no warranty (or else, saying that you provide a warranty) and that
-users may redistribute the program under these conditions, and telling
-the user how to view a copy of this License. (Exception: if the Program
-itself is interactive but does not normally print such an announcement,
-your work based on the Program is not required to print an announcement.)
+
+\addtocounter{enumi}{-1}
+
+\item Definitions.
+
+``This License'' refers to version 3 of the GNU General Public License.
+
+``Copyright'' also means copyright-like laws that apply to other kinds of
+works, such as semiconductor masks.
+
+``The Program'' refers to any copyrightable work licensed under this
+License.  Each licensee is addressed as ``you''.  ``Licensees'' and
+``recipients'' may be individuals or organizations.
+
+To ``modify'' a work means to copy from or adapt all or part of the work
+in a fashion requiring copyright permission, other than the making of an
+exact copy.  The resulting work is called a ``modified version'' of the
+earlier work or a work ``based on'' the earlier work.
+
+A ``covered work'' means either the unmodified Program or a work based
+on the Program.
+
+To ``propagate'' a work means to do anything with it that, without
+permission, would make you directly or secondarily liable for
+infringement under applicable copyright law, except executing it on a
+computer or modifying a private copy.  Propagation includes copying,
+distribution (with or without modification), making available to the
+public, and in some countries other activities as well.
+
+To ``convey'' a work means any kind of propagation that enables other
+parties to make or receive copies.  Mere interaction with a user through
+a computer network, with no transfer of a copy, is not conveying.
+
+An interactive user interface displays ``Appropriate Legal Notices''
+to the extent that it includes a convenient and prominently visible
+feature that (1) displays an appropriate copyright notice, and (2)
+tells the user that there is no warranty for the work (except to the
+extent that warranties are provided), that licensees may convey the
+work under this License, and how to view a copy of this License.  If
+the interface presents a list of user commands or options, such as a
+menu, a prominent item in the list meets this criterion.
+
+\item Source Code.
+
+The ``source code'' for a work means the preferred form of the work
+for making modifications to it.  ``Object code'' means any non-source
+form of a work.
+
+A ``Standard Interface'' means an interface that either is an official
+standard defined by a recognized standards body, or, in the case of
+interfaces specified for a particular programming language, one that
+is widely used among developers working in that language.
+
+The ``System Libraries'' of an executable work include anything, other
+than the work as a whole, that (a) is included in the normal form of
+packaging a Major Component, but which is not part of that Major
+Component, and (b) serves only to enable use of the work with that
+Major Component, or to implement a Standard Interface for which an
+implementation is available to the public in source code form.  A
+``Major Component'', in this context, means a major essential component
+(kernel, window system, and so on) of the specific operating system
+(if any) on which the executable work runs, or a compiler used to
+produce the work, or an object code interpreter used to run it.
+
+The ``Corresponding Source'' for a work in object code form means all
+the source code needed to generate, install, and (for an executable
+work) run the object code and to modify the work, including scripts to
+control those activities.  However, it does not include the work's
+System Libraries, or general-purpose tools or generally available free
+programs which are used unmodified in performing those activities but
+which are not part of the work.  For example, Corresponding Source
+includes interface definition files associated with source files for
+the work, and the source code for shared libraries and dynamically
+linked subprograms that the work is specifically designed to require,
+such as by intimate data communication or control flow between those
+subprograms and other parts of the work.
+
+The Corresponding Source need not include anything that users
+can regenerate automatically from other parts of the Corresponding
+Source.
+
+The Corresponding Source for a work in source code form is that
+same work.
+
+\item Basic Permissions.
+
+All rights granted under this License are granted for the term of
+copyright on the Program, and are irrevocable provided the stated
+conditions are met.  This License explicitly affirms your unlimited
+permission to run the unmodified Program.  The output from running a
+covered work is covered by this License only if the output, given its
+content, constitutes a covered work.  This License acknowledges your
+rights of fair use or other equivalent, as provided by copyright law.
+
+You may make, run and propagate covered works that you do not
+convey, without conditions so long as your license otherwise remains
+in force.  You may convey covered works to others for the sole purpose
+of having them make modifications exclusively for you, or provide you
+with facilities for running those works, provided that you comply with
+the terms of this License in conveying all material for which you do
+not control copyright.  Those thus making or running the covered works
+for you must do so exclusively on your behalf, under your direction
+and control, on terms that prohibit them from making any copies of
+your copyrighted material outside their relationship with you.
+
+Conveying under any other circumstances is permitted solely under
+the conditions stated below.  Sublicensing is not allowed; section 10
+makes it unnecessary.
+
+\item Protecting Users' Legal Rights From Anti-Circumvention Law.
+
+No covered work shall be deemed part of an effective technological
+measure under any applicable law fulfilling obligations under article
+11 of the WIPO copyright treaty adopted on 20 December 1996, or
+similar laws prohibiting or restricting circumvention of such
+measures.
+
+When you convey a covered work, you waive any legal power to forbid
+circumvention of technological measures to the extent such circumvention
+is effected by exercising rights under this License with respect to
+the covered work, and you disclaim any intention to limit operation or
+modification of the work as a means of enforcing, against the work's
+users, your or third parties' legal rights to forbid circumvention of
+technological measures.
+
+\item Conveying Verbatim Copies.
+
+You may convey verbatim copies of the Program's source code as you
+receive it, in any medium, provided that you conspicuously and
+appropriately publish on each copy an appropriate copyright notice;
+keep intact all notices stating that this License and any
+non-permissive terms added in accord with section 7 apply to the code;
+keep intact all notices of the absence of any warranty; and give all
+recipients a copy of this License along with the Program.
+
+You may charge any price or no price for each copy that you convey,
+and you may offer support or warranty protection for a fee.
+
+\item Conveying Modified Source Versions.
+
+You may convey a work based on the Program, or the modifications to
+produce it from the Program, in the form of source code under the
+terms of section 4, provided that you also meet all of these conditions:
+  \begin{enumerate}
+  \item The work must carry prominent notices stating that you modified
+  it, and giving a relevant date.
+
+  \item The work must carry prominent notices stating that it is
+  released under this License and any conditions added under section
+  7.  This requirement modifies the requirement in section 4 to
+  ``keep intact all notices''.
+
+  \item You must license the entire work, as a whole, under this
+  License to anyone who comes into possession of a copy.  This
+  License will therefore apply, along with any applicable section 7
+  additional terms, to the whole of the work, and all its parts,
+  regardless of how they are packaged.  This License gives no
+  permission to license the work in any other way, but it does not
+  invalidate such permission if you have separately received it.
+
+  \item If the work has interactive user interfaces, each must display
+  Appropriate Legal Notices; however, if the Program has interactive
+  interfaces that do not display Appropriate Legal Notices, your
+  work need not make them do so.
+\end{enumerate}
+A compilation of a covered work with other separate and independent
+works, which are not by their nature extensions of the covered work,
+and which are not combined with it such as to form a larger program,
+in or on a volume of a storage or distribution medium, is called an
+``aggregate'' if the compilation and its resulting copyright are not
+used to limit the access or legal rights of the compilation's users
+beyond what the individual works permit.  Inclusion of a covered work
+in an aggregate does not cause this License to apply to the other
+parts of the aggregate.
+
+\item Conveying Non-Source Forms.
+
+You may convey a covered work in object code form under the terms
+of sections 4 and 5, provided that you also convey the
+machine-readable Corresponding Source under the terms of this License,
+in one of these ways:
+  \begin{enumerate}
+  \item Convey the object code in, or embodied in, a physical product
+  (including a physical distribution medium), accompanied by the
+  Corresponding Source fixed on a durable physical medium
+  customarily used for software interchange.
+
+  \item Convey the object code in, or embodied in, a physical product
+  (including a physical distribution medium), accompanied by a
+  written offer, valid for at least three years and valid for as
+  long as you offer spare parts or customer support for that product
+  model, to give anyone who possesses the object code either (1) a
+  copy of the Corresponding Source for all the software in the
+  product that is covered by this License, on a durable physical
+  medium customarily used for software interchange, for a price no
+  more than your reasonable cost of physically performing this
+  conveying of source, or (2) access to copy the
+  Corresponding Source from a network server at no charge.
+
+  \item Convey individual copies of the object code with a copy of the
+  written offer to provide the Corresponding Source.  This
+  alternative is allowed only occasionally and noncommercially, and
+  only if you received the object code with such an offer, in accord
+  with subsection 6b.
+
+  \item Convey the object code by offering access from a designated
+  place (gratis or for a charge), and offer equivalent access to the
+  Corresponding Source in the same way through the same place at no
+  further charge.  You need not require recipients to copy the
+  Corresponding Source along with the object code.  If the place to
+  copy the object code is a network server, the Corresponding Source
+  may be on a different server (operated by you or a third party)
+  that supports equivalent copying facilities, provided you maintain
+  clear directions next to the object code saying where to find the
+  Corresponding Source.  Regardless of what server hosts the
+  Corresponding Source, you remain obligated to ensure that it is
+  available for as long as needed to satisfy these requirements.
+
+  \item Convey the object code using peer-to-peer transmission, provided
+  you inform other peers where the object code and Corresponding
+  Source of the work are being offered to the general public at no
+  charge under subsection 6d.
   \end{enumerate}
 
-These requirements apply to the modified work as a whole. If identifiable
-sections of that work are not derived from the Program, and can be
-reasonably considered independent and separate works in themselves,
-then this License, and its terms, do not apply to those sections when
-you distribute them as separate works. But when you distribute the
-same sections as part of a whole which is a work based on the Program,
-the distribution of the whole must be on the terms of this License,
-whose permissions for other licensees extend to the entire whole,
-and thus to each and every part regardless of who wrote it.
+A separable portion of the object code, whose source code is excluded
+from the Corresponding Source as a System Library, need not be
+included in conveying the object code work.
 
-Thus, it is not the intent of this section to claim rights or contest
-your rights to work written entirely by you; rather, the intent is
-to exercise the right to control the distribution of derivative or
-collective works based on the Program.
+A ``User Product'' is either (1) a ``consumer product'', which means any
+tangible personal property which is normally used for personal, family,
+or household purposes, or (2) anything designed or sold for incorporation
+into a dwelling.  In determining whether a product is a consumer product,
+doubtful cases shall be resolved in favor of coverage.  For a particular
+product received by a particular user, ``normally used'' refers to a
+typical or common use of that class of product, regardless of the status
+of the particular user or of the way in which the particular user
+actually uses, or expects or is expected to use, the product.  A product
+is a consumer product regardless of whether the product has substantial
+commercial, industrial or non-consumer uses, unless such uses represent
+the only significant mode of use of the product.
 
-In addition, mere aggregation of another work not based on the Program
-with the Program (or with a work based on the Program) on a volume
-of a storage or distribution medium does not bring the other work
-under the scope of this License.
+``Installation Information'' for a User Product means any methods,
+procedures, authorization keys, or other information required to install
+and execute modified versions of a covered work in that User Product from
+a modified version of its Corresponding Source.  The information must
+suffice to ensure that the continued functioning of the modified object
+code is in no case prevented or interfered with solely because
+modification has been made.
 
-\item You may copy and distribute the Program (or a work based on it, under
-Section 2) in object code or executable form under the terms of Sections
-1 and 2 above provided that you also do one of the following:
+If you convey an object code work under this section in, or with, or
+specifically for use in, a User Product, and the conveying occurs as
+part of a transaction in which the right of possession and use of the
+User Product is transferred to the recipient in perpetuity or for a
+fixed term (regardless of how the transaction is characterized), the
+Corresponding Source conveyed under this section must be accompanied
+by the Installation Information.  But this requirement does not apply
+if neither you nor any third party retains the ability to install
+modified object code on the User Product (for example, the work has
+been installed in ROM).
 
+The requirement to provide Installation Information does not include a
+requirement to continue to provide support service, warranty, or updates
+for a work that has been modified or installed by the recipient, or for
+the User Product in which it has been modified or installed.  Access to a
+network may be denied when the modification itself materially and
+adversely affects the operation of the network or violates the rules and
+protocols for communication across the network.
+
+Corresponding Source conveyed, and Installation Information provided,
+in accord with this section must be in a format that is publicly
+documented (and with an implementation available to the public in
+source code form), and must require no special password or key for
+unpacking, reading or copying.
+
+\item Additional Terms.
+
+``Additional permissions'' are terms that supplement the terms of this
+License by making exceptions from one or more of its conditions.
+Additional permissions that are applicable to the entire Program shall
+be treated as though they were included in this License, to the extent
+that they are valid under applicable law.  If additional permissions
+apply only to part of the Program, that part may be used separately
+under those permissions, but the entire Program remains governed by
+this License without regard to the additional permissions.
+
+When you convey a copy of a covered work, you may at your option
+remove any additional permissions from that copy, or from any part of
+it.  (Additional permissions may be written to require their own
+removal in certain cases when you modify the work.)  You may place
+additional permissions on material, added by you to a covered work,
+for which you have or can give appropriate copyright permission.
+
+Notwithstanding any other provision of this License, for material you
+add to a covered work, you may (if authorized by the copyright holders of
+that material) supplement the terms of this License with terms:
   \begin{enumerate}
-\item Accompany it with the complete corresponding machine-readable source
-code, which must be distributed under the terms of Sections 1 and
-2 above on a medium customarily used for software interchange; or,
-\item Accompany it with a written offer, valid for at least three years,
-to give any third party, for a charge no more than your cost of physically
-performing source distribution, a complete machine-readable copy of
-the corresponding source code, to be distributed under the terms of
-Sections 1 and 2 above on a medium customarily used for software interchange;
-or,
-\item Accompany it with the information you received as to the offer to
-distribute corresponding source code. (This alternative is allowed
-only for noncommercial distribution and only if you received the program
-in object code or executable form with such an offer, in accord with
-Subsection b above.)
+  \item Disclaiming warranty or limiting liability differently from the
+  terms of sections 15 and 16 of this License; or
+
+  \item Requiring preservation of specified reasonable legal notices or
+  author attributions in that material or in the Appropriate Legal
+  Notices displayed by works containing it; or
+
+  \item Prohibiting misrepresentation of the origin of that material, or
+  requiring that modified versions of such material be marked in
+  reasonable ways as different from the original version; or
+
+  \item Limiting the use for publicity purposes of names of licensors or
+  authors of the material; or
+
+  \item Declining to grant rights under trademark law for use of some
+  trade names, trademarks, or service marks; or
+
+  \item Requiring indemnification of licensors and authors of that
+  material by anyone who conveys the material (or modified versions of
+  it) with contractual assumptions of liability to the recipient, for
+  any liability that these contractual assumptions directly impose on
+  those licensors and authors.
   \end{enumerate}
 
-The source code for a work means the preferred form of the work for
-making modifications to it. For an executable work, complete source
-code means all the source code for all modules it contains, plus any
-associated interface definition files, plus the scripts used to control
-compilation and installation of the executable. However, as a special
-exception, the source code distributed need not include anything that
-is normally distributed (in either source or binary form) with the
-major components (compiler, kernel, and so on) of the operating system
-on which the executable runs, unless that component itself accompanies
-the executable.
+All other non-permissive additional terms are considered ``further
+restrictions'' within the meaning of section 10.  If the Program as you
+received it, or any part of it, contains a notice stating that it is
+governed by this License along with a term that is a further
+restriction, you may remove that term.  If a license document contains
+a further restriction but permits relicensing or conveying under this
+License, you may add to a covered work material governed by the terms
+of that license document, provided that the further restriction does
+not survive such relicensing or conveying.
 
-If distribution of executable or object code is made by offering access
-to copy from a designated place, then offering equivalent access to
-copy the source code from the same place counts as distribution of
-the source code, even though third parties are not compelled to copy
-the source along with the object code.
+If you add terms to a covered work in accord with this section, you
+must place, in the relevant source files, a statement of the
+additional terms that apply to those files, or a notice indicating
+where to find the applicable terms.
 
-\item You may not copy, modify, sublicense, or distribute the Program except
-as expressly provided under this License. Any attempt otherwise to
-copy, modify, sublicense or distribute the Program is void, and will
-automatically terminate your rights under this License. However, parties
-who have received copies, or rights, from you under this License will
-not have their licenses terminated so long as such parties remain
-in full compliance.
-\item You are not required to accept this License, since you have not signed
-it. However, nothing else grants you permission to modify or distribute
-the Program or its derivative works. These actions are prohibited
-by law if you do not accept this License. Therefore, by modifying
-or distributing the Program (or any work based on the Program), you
-indicate your acceptance of this License to do so, and all its terms
-and conditions for copying, distributing or modifying the Program
-or works based on it.
-\item Each time you redistribute the Program (or any work based on the Program),
-the recipient automatically receives a license from the original licensor
-to copy, distribute or modify the Program subject to these terms and
-conditions. You may not impose any further restrictions on the recipients'
-exercise of the rights granted herein. You are not responsible for
-enforcing compliance by third parties to this License.
-\item If, as a consequence of a court judgment or allegation of patent infringement
-or for any other reason (not limited to patent issues), conditions
-are imposed on you (whether by court order, agreement or otherwise)
-that contradict the conditions of this License, they do not excuse
-you from the conditions of this License. If you cannot distribute
-so as to satisfy simultaneously your obligations under this License
-and any other pertinent obligations, then as a consequence you may
-not distribute the Program at all. For example, if a patent license
-would not permit royalty-free redistribution of the Program by all
-those who receive copies directly or indirectly through you, then
-the only way you could satisfy both it and this License would be to
-refrain entirely from distribution of the Program.
+Additional terms, permissive or non-permissive, may be stated in the
+form of a separately written license, or stated as exceptions;
+the above requirements apply either way.
 
-If any portion of this section is held invalid or unenforceable under
-any particular circumstance, the balance of the section is intended
-to apply and the section as a whole is intended to apply in other
-circumstances.
+\item Termination.
 
-It is not the purpose of this section to induce you to infringe any
-patents or other property right claims or to contest validity of any
-such claims; this section has the sole purpose of protecting the integrity
-of the free software distribution system, which is implemented by
-public license practices. Many people have made generous contributions
-to the wide range of software distributed through that system in reliance
-on consistent application of that system; it is up to the author/donor
-to decide if he or she is willing to distribute software through any
-other system and a licensee cannot impose that choice.
+You may not propagate or modify a covered work except as expressly
+provided under this License.  Any attempt otherwise to propagate or
+modify it is void, and will automatically terminate your rights under
+this License (including any patent licenses granted under the third
+paragraph of section 11).
 
-This section is intended to make thoroughly clear what is believed
-to be a consequence of the rest of this License.
+However, if you cease all violation of this License, then your
+license from a particular copyright holder is reinstated (a)
+provisionally, unless and until the copyright holder explicitly and
+finally terminates your license, and (b) permanently, if the copyright
+holder fails to notify you of the violation by some reasonable means
+prior to 60 days after the cessation.
 
-\item If the distribution and/or use of the Program is restricted in certain
-countries either by patents or by copyrighted interfaces, the original
-copyright holder who places the Program under this License may add
-an explicit geographical distribution limitation excluding those countries,
-so that distribution is permitted only in or among countries not thus
-excluded. In such case, this License incorporates the limitation as
-if written in the body of this License.
-\item The Free Software Foundation may publish revised and/or new versions
-of the General Public License from time to time. Such new versions
-will be similar in spirit to the present version, but may differ in
-detail to address new problems or concerns.
+Moreover, your license from a particular copyright holder is
+reinstated permanently if the copyright holder notifies you of the
+violation by some reasonable means, this is the first time you have
+received notice of violation of this License (for any work) from that
+copyright holder, and you cure the violation prior to 30 days after
+your receipt of the notice.
 
-Each version is given a distinguishing version number. If the Program
-specifies a version number of this License which applies to it and
-``any later version,'' you have the option of following the terms
-and conditions either of that version or of any later version published
-by the Free Software Foundation. If the Program does not specify a
-version number of this License, you may choose any version ever published
+Termination of your rights under this section does not terminate the
+licenses of parties who have received copies or rights from you under
+this License.  If your rights have been terminated and not permanently
+reinstated, you do not qualify to receive new licenses for the same
+material under section 10.
+
+\item Acceptance Not Required for Having Copies.
+
+You are not required to accept this License in order to receive or
+run a copy of the Program.  Ancillary propagation of a covered work
+occurring solely as a consequence of using peer-to-peer transmission
+to receive a copy likewise does not require acceptance.  However,
+nothing other than this License grants you permission to propagate or
+modify any covered work.  These actions infringe copyright if you do
+not accept this License.  Therefore, by modifying or propagating a
+covered work, you indicate your acceptance of this License to do so.
+
+\item Automatic Licensing of Downstream Recipients.
+
+Each time you convey a covered work, the recipient automatically
+receives a license from the original licensors, to run, modify and
+propagate that work, subject to this License.  You are not responsible
+for enforcing compliance by third parties with this License.
+
+An ``entity transaction'' is a transaction transferring control of an
+organization, or substantially all assets of one, or subdividing an
+organization, or merging organizations.  If propagation of a covered
+work results from an entity transaction, each party to that
+transaction who receives a copy of the work also receives whatever
+licenses to the work the party's predecessor in interest had or could
+give under the previous paragraph, plus a right to possession of the
+Corresponding Source of the work from the predecessor in interest, if
+the predecessor has it or can get it with reasonable efforts.
+
+You may not impose any further restrictions on the exercise of the
+rights granted or affirmed under this License.  For example, you may
+not impose a license fee, royalty, or other charge for exercise of
+rights granted under this License, and you may not initiate litigation
+(including a cross-claim or counterclaim in a lawsuit) alleging that
+any patent claim is infringed by making, using, selling, offering for
+sale, or importing the Program or any portion of it.
+
+\item Patents.
+
+A ``contributor'' is a copyright holder who authorizes use under this
+License of the Program or a work on which the Program is based.  The
+work thus licensed is called the contributor's ``contributor version''.
+
+A contributor's ``essential patent claims'' are all patent claims
+owned or controlled by the contributor, whether already acquired or
+hereafter acquired, that would be infringed by some manner, permitted
+by this License, of making, using, or selling its contributor version,
+but do not include claims that would be infringed only as a
+consequence of further modification of the contributor version.  For
+purposes of this definition, ``control'' includes the right to grant
+patent sublicenses in a manner consistent with the requirements of
+this License.
+
+Each contributor grants you a non-exclusive, worldwide, royalty-free
+patent license under the contributor's essential patent claims, to
+make, use, sell, offer for sale, import and otherwise run, modify and
+propagate the contents of its contributor version.
+
+In the following three paragraphs, a ``patent license'' is any express
+agreement or commitment, however denominated, not to enforce a patent
+(such as an express permission to practice a patent or covenant not to
+sue for patent infringement).  To ``grant'' such a patent license to a
+party means to make such an agreement or commitment not to enforce a
+patent against the party.
+
+If you convey a covered work, knowingly relying on a patent license,
+and the Corresponding Source of the work is not available for anyone
+to copy, free of charge and under the terms of this License, through a
+publicly available network server or other readily accessible means,
+then you must either (1) cause the Corresponding Source to be so
+available, or (2) arrange to deprive yourself of the benefit of the
+patent license for this particular work, or (3) arrange, in a manner
+consistent with the requirements of this License, to extend the patent
+license to downstream recipients.  ``Knowingly relying'' means you have
+actual knowledge that, but for the patent license, your conveying the
+covered work in a country, or your recipient's use of the covered work
+in a country, would infringe one or more identifiable patents in that
+country that you have reason to believe are valid.
+
+If, pursuant to or in connection with a single transaction or
+arrangement, you convey, or propagate by procuring conveyance of, a
+covered work, and grant a patent license to some of the parties
+receiving the covered work authorizing them to use, propagate, modify
+or convey a specific copy of the covered work, then the patent license
+you grant is automatically extended to all recipients of the covered
+work and works based on it.
+
+A patent license is ``discriminatory'' if it does not include within
+the scope of its coverage, prohibits the exercise of, or is
+conditioned on the non-exercise of one or more of the rights that are
+specifically granted under this License.  You may not convey a covered
+work if you are a party to an arrangement with a third party that is
+in the business of distributing software, under which you make payment
+to the third party based on the extent of your activity of conveying
+the work, and under which the third party grants, to any of the
+parties who would receive the covered work from you, a discriminatory
+patent license (a) in connection with copies of the covered work
+conveyed by you (or copies made from those copies), or (b) primarily
+for and in connection with specific products or compilations that
+contain the covered work, unless you entered into that arrangement,
+or that patent license was granted, prior to 28 March 2007.
+
+Nothing in this License shall be construed as excluding or limiting
+any implied license or other defenses to infringement that may
+otherwise be available to you under applicable patent law.
+
+\item No Surrender of Others' Freedom.
+
+If conditions are imposed on you (whether by court order, agreement or
+otherwise) that contradict the conditions of this License, they do not
+excuse you from the conditions of this License.  If you cannot convey a
+covered work so as to satisfy simultaneously your obligations under this
+License and any other pertinent obligations, then as a consequence you may
+not convey it at all.  For example, if you agree to terms that obligate you
+to collect a royalty for further conveying from those to whom you convey
+the Program, the only way you could satisfy both those terms and this
+License would be to refrain entirely from conveying the Program.
+
+\item Use with the GNU Affero General Public License.
+
+Notwithstanding any other provision of this License, you have
+permission to link or combine any covered work with a work licensed
+under version 3 of the GNU Affero General Public License into a single
+combined work, and to convey the resulting work.  The terms of this
+License will continue to apply to the part which is the covered work,
+but the special requirements of the GNU Affero General Public License,
+section 13, concerning interaction through a network will apply to the
+combination as such.
+
+\item Revised Versions of this License.
+
+The Free Software Foundation may publish revised and/or new versions of
+the GNU General Public License from time to time.  Such new versions will
+be similar in spirit to the present version, but may differ in detail to
+address new problems or concerns.
+
+Each version is given a distinguishing version number.  If the
+Program specifies that a certain numbered version of the GNU General
+Public License ``or any later version'' applies to it, you have the
+option of following the terms and conditions either of that numbered
+version or of any later version published by the Free Software
+Foundation.  If the Program does not specify a version number of the
+GNU General Public License, you may choose any version ever published
 by the Free Software Foundation.
 
-\item If you wish to incorporate parts of the Program into other free programs
-whose distribution conditions are different, write to the author to
-ask for permission. For software which is copyrighted by the Free
-Software Foundation, write to the Free Software Foundation; we sometimes
-make exceptions for this. Our decision will be guided by the two goals
-of preserving the free status of all derivatives of our free software
-and of promoting the sharing and reuse of software generally.
-\end{enumerate}
+If the Program specifies that a proxy can decide which future
+versions of the GNU General Public License can be used, that proxy's
+public statement of acceptance of a version permanently authorizes you
+to choose that version for the Program.
 
-\subsection*{NO WARRANTY }
-\begin{itemize}
-\item 11.BECAUSE THE PROGRAM IS LICENSED FREE OF CHARGE, THERE IS NO WARRANTY
-FOR THE PROGRAM, TO THE EXTENT PERMITTED BY APPLICABLE LAW. EXCEPT
-WHEN OTHERWISE STATED IN WRITING THE COPYRIGHT HOLDERS AND/OR OTHER
-PARTIES PROVIDE THE PROGRAM ``AS IS'' WITHOUT WARRANTY OF ANY KIND,
-EITHER EXPRESSED OR IMPLIED, INCLUDING, BUT NOT LIMITED TO, THE IMPLIED
-WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE.
-THE ENTIRE RISK AS TO THE QUALITY AND PERFORMANCE OF THE PROGRAM IS
-WITH YOU. SHOULD THE PROGRAM PROVE DEFECTIVE, YOU ASSUME THE COST
-OF ALL NECESSARY SERVICING, REPAIR OR CORRECTION.
-\item 12.IN NO EVENT UNLESS REQUIRED BY APPLICABLE LAW OR AGREED TO IN WRITING
-WILL ANY COPYRIGHT HOLDER, OR ANY OTHER PARTY WHO MAY MODIFY AND/OR
-REDISTRIBUTE THE PROGRAM AS PERMITTED ABOVE, BE LIABLE TO YOU FOR
+Later license versions may give you additional or different
+permissions.  However, no additional obligations are imposed on any
+author or copyright holder as a result of your choosing to follow a
+later version.
+
+\item Disclaimer of Warranty.
+
+\begin{sloppypar}
+ THERE IS NO WARRANTY FOR THE PROGRAM, TO THE EXTENT PERMITTED BY
+ APPLICABLE LAW.  EXCEPT WHEN OTHERWISE STATED IN WRITING THE
+ COPYRIGHT HOLDERS AND/OR OTHER PARTIES PROVIDE THE PROGRAM ``AS IS''
+ WITHOUT WARRANTY OF ANY KIND, EITHER EXPRESSED OR IMPLIED,
+ INCLUDING, BUT NOT LIMITED TO, THE IMPLIED WARRANTIES OF
+ MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE.  THE ENTIRE
+ RISK AS TO THE QUALITY AND PERFORMANCE OF THE PROGRAM IS WITH YOU.
+ SHOULD THE PROGRAM PROVE DEFECTIVE, YOU ASSUME THE COST OF ALL
+ NECESSARY SERVICING, REPAIR OR CORRECTION.
+\end{sloppypar}
+
+\item Limitation of Liability.
+
+ IN NO EVENT UNLESS REQUIRED BY APPLICABLE LAW OR AGREED TO IN
+ WRITING WILL ANY COPYRIGHT HOLDER, OR ANY OTHER PARTY WHO MODIFIES
+ AND/OR CONVEYS THE PROGRAM AS PERMITTED ABOVE, BE LIABLE TO YOU FOR
  DAMAGES, INCLUDING ANY GENERAL, SPECIAL, INCIDENTAL OR CONSEQUENTIAL
-DAMAGES ARISING OUT OF THE USE OR INABILITY TO USE THE PROGRAM (INCLUDING
-BUT NOT LIMITED TO LOSS OF DATA OR DATA BEING RENDERED INACCURATE
-OR LOSSES SUSTAINED BY YOU OR THIRD PARTIES OR A FAILURE OF THE PROGRAM
-TO OPERATE WITH ANY OTHER PROGRAMS), EVEN IF SUCH HOLDER OR OTHER
-PARTY HAS BEEN ADVISED OF THE POSSIBILITY OF SUCH DAMAGES.
-\end{itemize}
+ DAMAGES ARISING OUT OF THE USE OR INABILITY TO USE THE PROGRAM
+ (INCLUDING BUT NOT LIMITED TO LOSS OF DATA OR DATA BEING RENDERED
+ INACCURATE OR LOSSES SUSTAINED BY YOU OR THIRD PARTIES OR A FAILURE
+ OF THE PROGRAM TO OPERATE WITH ANY OTHER PROGRAMS), EVEN IF SUCH
+ HOLDER OR OTHER PARTY HAS BEEN ADVISED OF THE POSSIBILITY OF SUCH
+ DAMAGES.
 
-\section*{END OF TERMS AND CONDITIONS }
+\item Interpretation of Sections 15 and 16.
 
-\subsection*{How to Apply These Terms to Your New Programs}
+If the disclaimer of warranty and limitation of liability provided
+above cannot be given local legal effect according to their terms,
+reviewing courts shall apply local law that most closely approximates
+an absolute waiver of all civil liability in connection with the
+Program, unless a warranty or assumption of liability accompanies a
+copy of the Program in return for a fee.
+
+\begin{center}
+{\Large\sc End of Terms and Conditions}
+
+\bigskip
+How to Apply These Terms to Your New Programs
+\end{center}
 
 If you develop a new program, and you want it to be of the greatest
-possible use to the public, the best way to achieve this is to make
-it free software which everyone can redistribute and change under
-these terms.
+possible use to the public, the best way to achieve this is to make it
+free software which everyone can redistribute and change under these terms.
 
@@
 to attach them to the start of each source file to most effectively
-convey the exclusion of warranty; and each file should have at least
-the ``copyright'' line and a pointer to where the full notice is
-found. For example:
-\begin{quote}
-One line to give the program's name and a brief idea of what it does.
-Copyright {\footnotesize $\copyright$ (}year) (name of author)
+state the exclusion of warranty; and each file should have at least
+the ``copyright'' line and a pointer to where the full notice is found.
 
-This program is free software; you can redistribute it and/or modify
-it under the terms of the GNU General Public License as published
-by the Free Software Foundation; either version 2 of the License,
-or (at your option) any later version.
+{\footnotesize
+\begin{verbatim}
+<one line to give the program's name and a brief idea of what it does.>
 
-This program is distributed in the hope that it will be useful, but
-WITHOUT ANY WARRANTY; without even the implied warranty of MERCHANTABILITY
-or FITNESS FOR A PARTICULAR PURPOSE. See the GNU General Public License
-for more details.
+Copyright (C) <textyear>  <name of author>
+
+This program is free software: you can redistribute it and/or modify
+it under the terms of the GNU General Public License as published by
+the Free Software Foundation, either version 3 of the License, or
+(at your option) any later version.
+
+This program is distributed in the hope that it will be useful,
+but WITHOUT ANY WARRANTY; without even the implied warranty of
+MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+GNU General Public License for more details.
 
 You should have received a copy of the GNU General Public License
-along with this program; if not, write to the Free Software Foundation,
-Inc., 59 Temple Place, Suite 330, Boston, MA 02111-1307 USA
-\end{quote}
-Also add information on how to contact you by electronic and paper
-mail.
+along with this program.  If not, see <http://www.gnu.org/licenses/>.
+\end{verbatim}
+}
 
-If the program is interactive, make it output a short notice like
-this when it starts in an interactive mode:
-\begin{quote}
-Gnomovision version 69, Copyright $\copyright$ year name of author
-Gnomovision comes with ABSOLUTELY NO WARRANTY; for details type `show
-w'. This is free software, and you are welcome to redistribute it
+Also add information on how to contact you by electronic and paper mail.
+
+If the program does terminal interaction, make it output a short
+notice like this when it starts in an interactive mode:
+
+{\footnotesize
+\begin{verbatim}
+<program>  Copyright (C) <year>  <name of author>
+
+This program comes with ABSOLUTELY NO WARRANTY; for details type `show w'.
+This is free software, and you are welcome to redistribute it
 under certain conditions; type `show c' for details.
-\end{quote}
-The hypothetical commands `show w' and `show c' should show the appropriate
-parts of the General Public License. Of course, the commands you use
-may be called something other than `show w' and `show c'; they could
-even be mouse-clicks or menu items -- whatever suits your program.
+\end{verbatim}
+}
+
+The hypothetical commands {\tt show w} and {\tt show c} should show
+the appropriate
+parts of the General Public License.  Of course, your program's commands
+might be different; for a GUI interface, you would use an ``about box''.
 
 You should also get your employer (if you work as a programmer) or
-your school, if any, to sign a ``copyright disclaimer'' for the
-program, if necessary. Here is a sample; alter the names:
-\begin{quote}
-Yoyodyne, Inc., hereby disclaims all copyright interest in the program
-`Gnomovision' (which makes passes at compilers) written by James Hacker.
+school, if any, to sign a ``copyright disclaimer'' for the program, if
+necessary.  For more information on this, and how to apply and follow
+the GNU GPL, see \texttt{http://www.gnu.org/licenses/}.
 
-(signature of Ty Coon)\\
- 1 April 1989 \\
- Ty Coon, President of Vice
-\end{quote}
-This General Public License does not permit incorporating your program
-into proprietary programs. If your program is a subroutine library,
-you may consider it more useful to permit linking proprietary applications
-with the library. If this is what you want to do, use the GNU Library
-General Public License instead of this License.
+The GNU General Public License does not permit incorporating your
+program into proprietary programs.  If your program is a subroutine
+library, you may consider it more useful to permit linking proprietary
+applications with the library.  If this is what you want to do, use
+the GNU Lesser General Public License instead of this License.  But
+first, please read \texttt{http://www.gnu.org/philosophy/why-not-lgpl.html}.
+
+\end{enumerate}
+
+\end{document}

### 2017-06-22 doc/USER_MANUAL/08b_performing_full_waveform_inversion_FWI_or_source_inversions.tex
fixed a typo in the users manual

--- 2016-10-25 doc/USER_MANUAL/01_introduction.tex
@@
-\chapter{Introduction}
+\chapter{Performing full waveform inversion (FWI) or source inversions}\label{cha:FWI}
 
-The software package SPECFEM3D Cartesian simulates seismic wave propagation
-at the local or regional scale and performs full waveform imaging (FWI) or adjoint tomography based upon the spectral-element method
-(SEM). The SEM is a continuous Galerkin technique \citep{TrKoLi08,PeKoLuMaLeCaLeMaLiBlNiBaTr11},
-which can easily be made discontinuous \citep{BeMaPa94,Ch00,KoWoHu02,ChCaVi03,LaWaBe05,Kop06,WiStBuGh10,AcKo11};
-it is then close to a particular case of the discontinuous Galerkin
-technique \citep{ReHi73,LeRa74,Arn82,JoPi86,BoMaHe91,FaRi99,HuHuRa99,CoKaSh00,GiHeWa02,RiWh03,MoRi05,GrScSc06,AiMoMu06,BeLaPi06,DuKa06,DeSeWh08,PuAmKa09,WiStBuGh10,DeSe10,EtChViGl10},
-with optimized efficiency because of its tensorized basis functions
-\citep{WiStBuGh10,AcKo11}. In particular, it can accurately handle
-very distorted mesh elements \citep{OlSe11}.\\
-
-\red{In fluids, when gravity is turned off, SPECFEM3D uses the classical
-linearized Euler equation; thus if you have sharp local variations of
-density in the fluid (highly heterogeneous fluids
-in terms of density) or if density becomes extremely small in some
-regions of your model (e.g. for upper-atmosphere studies), before using
-the code please make sure the linearized Euler equation is a valid
-approximation in the case you want to study, and/or see if you should turn gravity on.
-For more details on that see e.g. \cite{COA2011}.}\\
-
-It has very good accuracy and convergence properties \citep{MaPa89,SePr94,DeFiMu02,Coh02,DeSe07,SeOl08,AiWa09,AiWa10,MeStTh12}.
-The spectral element approach admits spectral rates of convergence
-and allows exploiting $hp$-convergence schemes. It is also very well
-suited to parallel implementation on very large supercomputers \citep{KoTsChTr03,TsKoChTr03,KoLaMi08a,CaKoLaTiMiLeSnTr08,KoViCh10}
-as well as on clusters of GPU accelerating graphics cards \citep{Kom11,MiKo10,KoMiEr09,KoErGoMi10}.
-Tensor products inside each element can be optimized to reach very
-high efficiency \citep{DeFiMu02}, and mesh point and element numbering
-can be optimized to reduce processor cache misses and improve cache
-reuse \citep{KoLaMi08a}. The SEM can also handle triangular (in 2D)
-or tetrahedral (in 3D) elements \citep{WinBoyd96,TaWi00,KoMaTrTaWi01,Coh02,MeViSa06}
-as well as mixed meshes, although with increased cost and reduced
-accuracy in these elements, as in the discontinuous Galerkin method.\\
-
-Note that in many geological models in the context of seismic wave
-propagation studies (except for instance for fault dynamic rupture
-studies, in which very high frequencies or supershear rupture need
-to be modeled near the fault, see e.g. \citet{BeGlCrViPi07,BeGlCrVi09,PuAmKa09,TaCrEtViBeSa10})
-a continuous formulation is sufficient because material property contrasts
-are not drastic and thus conforming mesh doubling bricks can efficiently
-handle mesh size variations \citep{KoTr02a,KoLiTrSuStSh04,LeChLiKoHuTr08,LeChKoHuTr09,LeKoHuTr09}.\\
-
-For a detailed introduction to the SEM as applied to regional seismic
-wave propagation, please consult \citet{PeKoLuMaLeCaLeMaLiBlNiBaTr11,TrKoLi08,KoVi98,KoTr99,ChKoViCaVaFe07}
-and in particular \citet{LeKoHuTr09,LeChKoHuTr09,LeChLiKoHuTr08,GoAmTaCaSmSaMaKo09,WiKoScTr04,KoLiTrSuStSh04}.
-A detailed theoretical analysis of the dispersion
-and stability properties of the SEM is available in \citet{Coh02}, \citet{DeSe07}, \citet{SeOl07}, \citet{SeOl08} and \citet{MeStTh12}.\\
-
-Effects due to lateral variations in compressional-wave speed, shear-wave
-speed, density, a 3D crustal model, topography and bathymetry are
-included. The package can accommodate full 21-parameter anisotropy
-(see~\citet{ChTr07}) as well as lateral variations in attenuation
-\citep{SaKoTr10}. Adjoint capabilities and finite-frequency kernel
-simulations are included \citep{TrKoLi08,PeKoLuMaLeCaLeMaLiBlNiBaTr11,LiTr06,FiIgBuKe09,ViOp09}.\\
-
-The SEM was originally developed in computational fluid dynamics \citep{Pat84,MaPa89}
-and has been successfully adapted to address problems in seismic wave
-propagation. Early seismic wave propagation applications of the SEM,
-utilizing Legendre basis functions and a perfectly diagonal mass matrix,
-include \citet{CoJoTo93}, \citet{Kom97}, \citet{FaMaPaQu97}, \citet{CaGa97},
-\citet{KoVi98} and \citet{KoTr99}, whereas applications involving
-Chebyshev basis functions and a non-diagonal mass matrix include \citet{SePr94},
-\citet{PrCaSe94} and \citet{SePrPr95}.
-In the Legendre version that we use in SPECFEM the mass matrix is purposely slightly inexact but diagonal (but can be made exact if needed, see \cite{Teu15}),
-while in the Chebyshev version it is exact but non diagonal.\\
-
-All SPECFEM3D software is written in Fortran2003 with full portability
-in mind, and conforms strictly to the Fortran2003 standard. It uses
-no obsolete or obsolescent features of Fortran. The package uses parallel
-programming based upon the Message Passing Interface (MPI) \citep{GrLuSk94,Pac97}.\\
-
-SPECFEM3D won the Gordon Bell award for best performance at the SuperComputing~2003
-conference in Phoenix, Arizona (USA) (see \citet{KoTsChTr03} and
-\url{www.sc-conference.org/sc2003/nrfinalaward.html}). It was a finalist
-again in 2008 for a run at 0.16 petaflops (sustained) on 149,784 processors
-of the `Jaguar' Cray XT5 system at Oak Ridge National Laboratories
-(USA) \citep{CaKoLaTiMiLeSnTr08}. It also won the BULL Joseph Fourier
-supercomputing award in 2010.\\
-
-It reached the sustained one petaflop performance level for the first time in February 2013
-on the Blue Waters Cray supercomputer at the National Center for Supercomputing Applications (NCSA), located at the University of Illinois at Urbana-Champaign (USA).\\
-
-This new release of the code includes Convolution or
-Auxiliary Differential Equation Perfectly Matched absorbing Layers
-(C-PML or ADE-PML) \citep{MaKoEz08,MaKoGe08,MaKo09,MaKoGeBr10,KoMa07}.
-It also includes support for GPU graphics card acceleration \citep{Kom11,MiKo10,KoMiEr09,KoErGoMi10}.\\
-
-The next release of the code will use the PT-SCOTCH parallel and threaded
-version of SCOTCH for mesh partitioning instead of the serial version.\\
-
-SPECFEM3D Cartesian includes coupled fluid-solid domains and adjoint
-capabilities, which enables one to address seismological inverse problems,
-but for linear rheologies only so far. To accommodate visco-plastic
-or non-linear rheologies, the reader can refer to the \href{http://geoelse.stru.polimi.it/}{GeoELSE}software
-package \citep{CaGa97,StPaIg09}.
-
-\section{Citation}
-
-You can find all the references below in \BibTeX format in file \texttt{doc/USER\_MANUAL/bibliography.bib}.\\
-
-If you use SPECFEM3D Cartesian for your own research, please cite
-at least one of the following articles:
-\begin{description}
-\item [{Numerical simulations in general}] ~\\
- Forward and adjoint simulations are described in detail in \citet{TrKoLi08,PeKoLuMaLeCaLeMaLiBlNiBaTr11,VaCaSaKoVi99,KoMiEr09,KoErGoMi10,ChKoViCaVaFe07,MaKoDi09,KoViCh10,CaKoLaTiMiLeSnTr08,TrKoHjLiZhPeBoMcFrTrHu10,KoRiTr02,KoTr02a,KoTr02b,KoTr99}
-or \citet{KoVi98}. Additional aspects of adjoint simulations are
-described in \citet{TrTaLi05,LiTr06,TrKoLi08,LiTr08,TrKoHjLiZhPeBoMcFrTrHu10,PeKoLuMaLeCaLeMaLiBlNiBaTr11}.
-Domain decomposition is explained in detail in \citet{MaKoBlLe08},
-and excellent scaling up to 150,000 processor cores is shown for instance
-in \citet{CaKoLaTiMiLeSnTr08,KoLaMi08a,MaKoBlLe08,KoErGoMi10,Kom11},
-\item [{GPU computing}] ~\\
- Computing on GPU graphics cards for acoustic or seismic wave propagation
-applications is described in detail in \citet{Kom11,MiKo10,KoMiEr09,KoErGoMi10}.
-\end{description}
-\noindent If you use this new version, which has non blocking
-MPI for much better performance for medium or large runs, please cite
-at least one of these six articles, in which results of non blocking
-MPI runs are presented: \citet{PeKoLuMaLeCaLeMaLiBlNiBaTr11,KoErGoMi10,KoViCh10,Kom11,CaKoLaTiMiLeSnTr08,MaKoBlLe08}.\\
-
-If you use the C-PML absorbing layer capabilities of the code, please cite at least one article
-written by the developers of the package, for instance:
-%
-\begin{itemize}
-\item \cite{XiKoMaMa14},
-\item \cite{XiMaCrKoMa16}.
-\end{itemize}
-%
-If you use the UNDO\_ATTENUATION option of the code in order to produce full anelastic/viscoelastic sensitivity kernels, please cite at least one article
-written by the developers of the package, for instance (and in particular):
-%
-\begin{itemize}
-\item \cite{KoXiBoPeSaLiTr16}.
-\end{itemize}
-%
-More generally, if you use the attenuation (anelastic/viscoelastic) capabilities of the code, please cite at least one article
-written by the developers of the package, for instance:
-%
-\begin{itemize}
-\item \cite{KoXiBoPeSaLiTr16},
-\item \cite{BlKoChLoXi16}.
-\end{itemize}
-%
-If you use the kernel capabilities of the code, please cite at least one article
-written by the developers of the package, for instance:
-%
-\begin{itemize}
-\item \cite{TrKoLi08},
-\item \cite{PeKoLuMaLeCaLeMaLiBlNiBaTr11},
-\item \cite{LiTr06},
-\item \cite{MoLuTr09}.
-\end{itemize}
-
-\noindent If you work on geophysical applications, you may be interested
-in citing some of these application articles as well, among others:
-\begin{description}
-\item [{Southern California simulations}] ~\\
- \citet{KoLiTrSuStSh04,KrJiKoTr06a,KrJiKoTr06b}.
-
-If you use the 3D southern California model, please cite \citet{SuSh03}
-(Los Angeles model), \citet{lovelyetal06} (Salton Trough), and \citet{hauksson2000}
-(southern California). The Moho map was determined by \citet{zhukanamori2000}.
-The 1D SoCal model was developed by \citet{DrHe90}.
-
-\item [{Anisotropy}] ~\\
- \citet{ChTr07,JiTsKoTr05,ChFaKo04,FaChKo04,RiRiKoTrHe02,TrKo00}.
-\item [{Attenuation}] ~\\
- \citet{SaKoTr10,KoTr02a,KoTr99}.
-\item [{Topography}] ~\\
- \citet{LeKoHuTr09,LeChKoHuTr09,LeChLiKoHuTr08,GoAmTaCaSmSaMaKo09,WiKoScTr04}.
-\end{description}
-The corresponding Bib\TeX{} entries may be found in file \texttt{doc/USER\_MANUAL/bibliography.bib}.
-
-\section{Support}
-
-This material is based upon work supported by the USA National Science
-Foundation under Grants No. EAR-0406751 and EAR-0711177, by the French
-CNRS, French INRIA Sud-Ouest MAGIQUE-3D, French ANR NUMASIS under
-Grant No. ANR-05-CIGC-002, and European FP6 Marie Curie International
-Reintegration Grant No. MIRG-CT-2005-017461. Any opinions, findings,
-and conclusions or recommendations expressed in this material are
-those of the authors and do not necessarily reflect the views of the
-USA National Science Foundation, CNRS, INRIA, ANR or the European
-Marie Curie program.
+\red{\Large ANNOUNCEMENT: SPECFEM3D can now perform full waveform inversion (FWI), i.e. invert for models in an iterative fashion, and it can also perform source inversions in a constant model;
+please refer to the two new directories, \texttt{inverse\_problem\_for\_model} and \texttt{inverse\_problem\_for\_source}, and the README files they contain.
+For FWI inversions for the model, also refer to the new examples provided in
+directory \texttt{EXAMPLES}.}

### 2017-06-24 src/inverse_problem_for_source/CMT3D/cmt3d/readme/readme.tex
re-added src/inverse_problem_for_source/CMT3D as a local copy instead of as a submodule

%
%=======================================
%  template.tex
%=======================================
%
\documentclass[12pt,titlepage,fleqn]{article}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{latexsym}
\usepackage[round]{natbib}
\usepackage{xspace}
%\usepackage{epsfig}
\usepackage{graphicx}
\usepackage{bm}
\usepackage[usenames]{color}
\usepackage{enumerate}

% do not use when showing Table of Contents
%\usepackage{fancybox}
\usepackage{fancyhdr}
\pagestyle{fancy}
\fancyheadoffset[R]{12pt}
\lhead{}
\chead{}
\rhead{Q Liu}

% SPACING COMMANDS (Latex Companion, p. 52)
\usepackage{setspace}
\renewcommand{\baselinestretch}{1.0}
\textwidth 470pt
\textheight 630pt
\oddsidemargin 0pt
\evensidemargin 0pt

% see Latex Companion, p. 85
%\voffset     -50pt
\topmargin     0pt
\headsep      20pt
\headheight   15pt
\headheight    0pt
\footskip     30pt
\hoffset       0pt

% nicer fonts
\usepackage{times}

% self-defined commands
\include{newcommands}

% bibliography
\bibliographystyle{gji}

%======================================================
\begin{document}
%======================================================
\begin{spacing}{1.0} % empty lines are required!
\begin{center}

\Large {\vspace{0.1in} \bf \verb+CMT3D_FLEXWIN+ Manual}

\small
Qinya Liu \\
Department of Physics \\
%Faculty of Arts and Science \\
University of Toronto

\today

\end{center}

\end{spacing}

%======================================================
\verb=CMT3D_FLEXWIN= is re-written from my old \verb=CMT3D= program to reflect the gist of the program, which is to use Gaussian elimination to solve the source inverse problem, In particular,
\begin{itemize}
\item directly use the output windows file from Alessia's \verb=FLEXWIN= package, which selects time windows with decent match between data and synthetics.
\item eliminate bandpass filter as it should have been applied before entering \verb+FLEXWIN+.
\item  by default, inversions are done in local coordinates \{\verb+dep,lon,lat+\}. However, with \verb+global_coord=.true.+, it also allows inversions to be done in global coordinates \{\verb+X,Y,Z+\}, which may come in handy for future multiple source inversions. The  bottom line is that the code is written such that it should be able to take the old parameter files and run exactly the same as before (in local coordinates) without even specifying this logical.
\item incorporate centroid time shift and half duration as inversion parameters (npar=10 and 11), although derivatives for hdur and tshift have to be precomputed by python scripts \red{(which one?)}
\end{itemize}
What still remains to be added is the possiblity of \verb+CMT3D+ inversion for multiple subevents, which may be useful if $N_{event}$ is relatively small, and may be handy for efficiency comparisons to \verb+CMT3D_CG+.

%===========================================================
\section{Setup}

\subsection{Setup derivative cmt solutions}
According to the theory in the Appendix, we need to compute the derivative synthetics by brute force difference formula
\eq \papa{s}{m} \sim \frac{s(m+\Delta m)-s(m)}{\Delta m} \en
for all source parameters involved in the inversion. Depending on how many parameters are chosen (\verb+npar+), we need to run at least $6-9$ SEM simulations to gather all the derivatives. Each SEM simulation requires a corresponding \verb+CMTSOLUTION+ for $m+\Delta m$, and the $\Delta m$ should be chosen carefully to avoid either numerical inaccuracy or instability. In the local coordinates  \{\verb+dep,lon,lat+\}, for the global code (\verb+SPECFEM3D_GLOBE+), a choice may be
\eq
\Delta M = 1\times 10^{24}\text{ dyne.cm},\quad \Delta X = 2\text{ km (depth)}, \quad \Delta\delta=0.02 \text{ deg  (lat/lon)}
\en
and for the basin code (\verb+SPECFEM3D+), a choice may be
\eq
\Delta M = 1\times 10^{22}\text{ dyne.cm},\quad \Delta X = 0.5\text{ km}, \quad \Delta\delta=0.01
\en
where $\Delta X$ refers to depth perturbations and $\Delta \delta$ refers to perturbation in latitude and longitude.

Depending on what coordinate system you choose to calculate derivatives and run \verb+CMT3D+ inversions, different extensions are expected for derivative \verb+CMTSOLUTION+s as well as derivative synthetics (generated by \verb+gen_der_cmt()+ python script):
\begin{itemize}
\item local (depth, lon, lat) coordinates (\verb+global_coord=.false.+):\\
   \verb+ext=['Mrr','Mtt','Mpp','Mrt','Mrp','Mtp','dep','lon','lat']+\\
This system should used as default for most applications.

\item global code, global (X,Y,Z) coordinates (\verb+global_coord=.true.+):\\
   \verb+ext=['Mxx','Myy','Mzz','Mxy','Mxz','Myz','xxx','yyy','zzz']+\\
This coordinate system may only be used to accommodate multiple event inversions in the future. For derivations of rotation matrices, please refer to the Appendix.

\end{itemize}

Note:
\begin{enumerate}
\item
Python script and modules are provided in \verb+scripts+ directory to facilitate this process, and allow full compatibility with \verb+CMT3D+ code:
\begin{verbatim}
 gen_cmt_der('CMTSOLUTION',npar=9,dmoment=1.e24,
              ddepth=2,dlocation=0.02,utm=0)
\end{verbatim}
where $dmoment$, $ddepth$ and $dlocation$ are given in dyne$\cdot$cm, km and degrees ($\Delta X=111.12\Delta \delta $), and $utm=0$ refers to local coordinates  (depth, lon, lat).

If $utm=-1$, it refers to global coordinates (X, Y, Z), and \verb+ddepth=dX+, \verb+dlocation=dY=dZ+ (may need to modified in future).
Note that \verb+gen_cmt_der+ does allow utm to be 1-60, and map  small regions into a cartesian geometry (depth, E, N). But that is NOT yet used for the \verb+cmt3d+ package.
\item Also as discussed in the theory section, derivatives for centroid time and half duration simply depend on the original synthetics, and  no extra simulation is necessary. \red{how to compute?}
\item After simulations for these derivative \verb+CMTSOLUTION+s are finished, collect synthetic seismograms, convert them to sac files and add proper extension to the sac files according to the choice of coordinate system.

\item Note synthetics for the original \verb+CMTSOLUTION+ do not need to be obtained through an extra simulation, a simple linear combination of moment tensor derivatives can be realized by:
\begin{verbatim}
 xadd_frechet_derivatives s syn_list cmt moment_scale(1e24)
\end{verbatim}

\end{enumerate}

\subsection{Pre-processing data and synthetics}

After gather raw data from IRIS (for global earthquakes) or SCEDC (for socal earthquakes), and collect synthetics as well derivative synthetics from SEM simulations, you need to go through the pre-processing step to present properly cut and filtered data and synthetics to the \verb=FLEXWIN= package. I have come a long way in writing such scripts, and right now I am settled on a recent one in python:
\begin{verbatim}
 process_ds.py  \
  -d data,sac -s syn,  -p CMTSOLUTION,STATIONS\
  -t 6,30 -w -D [-r]
\end{verbatim}
%\verb=-d= and \verb=-s= give directory names and extension names of the raw data and synthetics. \verb=-0=,\verb=-1=\verb=-2=,\verb=-3= represents different pre-processing stages which you can choose to run separately. During \verb=-0=, information is gathered from \verb=CMTSOLUTION= and \verb=station_file= and written into sac headers. New directories \verb=data_PP= and \verb=syn_PP= are created to host these pre-processed data and syn files. During \verb=-1= stage, data and synthetics are padded before the beginning time to a value specified by \verb=-b=, and proper cutting between data and synthetics are performed to ensure futher operations. \verb=-2= stage mainly deals with band-passing data and synthetics between the frequence bands given by \verb=-t=, and moving the new processed files to directories \verb=data_T006_T030= and \verb=syn_T006_T030=. \verb=-3= stage writes out the input file \verb=flexwin_T006_T030.input= for \verb=FLEXWIN= package after all the above steps have been executed. The script structure and many implementation details are inspired by a similar script written by Carl Tape.
For detailed usage, simple type \verb+process_ds.py+. The outcomes of this script are
\begin{itemize}
\item filtered data and (derivative) synthetics in \verb+data_6_30+ and \verb+syn_6_30+.
\item an input file to \verb+FLEXWIN+ named \verb+INPUT_flexwin_6_30+ will be generated.
\end{itemize}

\subsection{Executing FLEXWIN}
After choosing the right parameters in \verb=PAR_FILE= and user functions for flexwin package (please refer to \verb+FLEXWIN+ manual for details), simply run
\begin{verbatim}
./flexwin < INPUT_flexwin_6_30 > OUTPUT_flexwin_6_30
\end{verbatim}
After running \verb+FLEXWIN+ through data and synthetics filtered at various period bands (e.g., \verb+3-30+ and \verb+6-30+ secs),  all individual windows files in \verb+MEASURE_t1_t2+ directories can be cat into an input for \verb=cmt3d_flexwin= package.
\begin{verbatim}
write_flexwin_out.py MEASURE_6_30 MEASURE_20_50  \
                     INPUT_cmt3d_flexwin_noweight
\end{verbatim}
Now it is possible to pad time shift and weight information to the composite output windows file (after the time range of the window), either to facilitate a direct evaluation of misfit function, or to eliminate the weight calculations in \verb+CMT3D+ package, which can be rather cumbersome and inflexible.
\begin{verbatim}
write_flexwin_shift_weight.py -c 2/2/1 -d 1.15/0.55/0.78 -a 0.2 -r 2 -t \
      -D data -s MEASURE  INPUT_cmt3d_flexwin_noweight \
      INPUT_cmt3d_flexwin_withweight
\end{verbatim}
Obviously you can not just blindly use this package without checking. Here are two auxilliary scripts:
\begin{verbatim}
plot_flexwin.py MEASURE
plot_data_and_syn.pl -d data_dir,data_ext -s syn_dir,syn_ext -m cmt \
   -M OUTPUT_flexwin -A (Z/R/T) 0.05/0.05/0.08
\end{verbatim}
The first script gives \verb+flexwin_seismo.pdf+ and the second one gives \verb+ds_all.pdf+.
%============================================
\section{CMT3D package}
\subsection{Parameter file}
For structures of the main program and subroutine calls, refer to \verb=call-graph=. The name of the parameter file is hard-wired in the main program as \verb=INVERSION.PAR=.

\begin{verbatim}
CMTSOLUTION
CMTSOLUTION_NEW
6  .false.                      -- npar[, global_coord]
0.01 1.0  1.0e22                -- dlocation,ddepth,dmoment
flexwin.out
.true. .false.                 -- weigh_data_files[, read_weight]
2 2 1  0  1.15 0.55 0.78       -- weights of data comp,az,dist
.true.                         -- station_correction
.true.  .true. 0.0             -- zero_trace,double_couple,damping
.false.                        -- write_new_syn (psuedo-synthetics)
\end{verbatim}
The first line gives the old cmt solution file name. This is also the cmt solution you have used to generate derivative cmts and compute the synthetics from linear combinations. The second line gives name of the new cmt solution file output after the inversion is done.\\
File \verb=flexwin.out= holds the start and end time of the windows outputed from FLEXWIN package, in the following format:
\begin{verbatim}
nfiles
data-file-1
syn-file-1
 nwins-for-this-data-and-syn
 tstart(1) tend(1) [tshift(1) weight(1)]
 ...
 tstart(nwin) tend(nwin) [tshift(nwin) weight(nwin)]
data-file-2
syn-file-2
...
\end{verbatim}
Only data points inside the windows from \verb+flexwin.out+ file will be used to perform time integration and assemble the inversion matrices. Note the values in [] are recent additions where we allow weights to be specified in the input file together with an indicative logical \verb+read_weight+. For now, \verb+tshift+ is still a dummy variable not used for aligning data and synthetics (a separate xcorr subroutine calculates it in the code, but future merge may be possible).

All other parameters are pretty self-explanatory:
\begin{itemize}
 \item Number of parameters to be inverted can only be 6(moment-only), 7(moment+depth), 9(moment+locations). \verb+global_coord+ indicates if the input derivative synthetics are in terms of $(X,Y,Z)$ or $(depth,lon,lat)$.
 \item For local coordinates, the derivative synthetics should have been calculated based on the \verb=dlocation= (in degrees), \verb=ddepth= (in km), \verb=dmoment= (in dyne$\cdot$cm) values. However, notice that strictly speaking, only the ones for moment are actual derivative synthetics; the ones for depth, longitude, and latitude (or \red{R, S, E} depending on coordinate system or utm) are actual synthetics, from which the code subtracts the old synthetics and divides it by ddepth or ddelta (refer to Appendix for more details).
 \item Weighting subroutine \verb=compute_data_weights()=  needs to be tweaked if you choose to \verb=weigh_data_files=, and not importing weights directly from the input file by \verb=read_weight=. The idea is that you utilize the \verb=dist_km=, \verb=dist_deg=, \verb=azimuth=, and other file information from the sac headers to come up with a sensible weighting scheme for your particular scale/region.
 \item Station correction is used to align the data with the synthetics to accommondate 3D path effects that may bias your source inversion results.
 \item Different constraints on the moment-tensor are allowed, including zero-trace and double-couple. To improve the stability of the matrix inverse (when only very few windows are used), damping is also allowed before Gaussian elimination.
 \item Finally synthetics for the new cmt solution can be written, however, notice that these are obtained from `linearized' approximation, not from an actual SEM simulation.
 \end{itemize}

Some more words on weighting:
\begin{itemize}
\item If you want to invert either only surface waves or body waves, make sure you set the optimal parameters in \verb+FLEXWIN+ to get the correct sets of windows.

\item How do the weights work in \verb=compute_data_weights()=:
\begin{verbatim}
        data_weights(nwint) =  cmp_weight(nwint) &
             * (dist/REF_DIST) ** dist_exp_weight(nwint) &
             * corr(nwint) ** corr_exp_weight(nwint) &
             / ( naz(k) ** az_exp_weight)
\end{verbatim}
 where component weights are applied linearly, distance and azimuth weights are applied exponentially (such that $0$ means no weight and $1$ means linear weight). \verb=naz(k)= is the number of traces available in a particular azimuth bin (\verb=+1=). One can set \verb=az_exp_weight= to be 0 to mute out the azimuthal effect, the same for distance weights.

\end{itemize}

\subsection{Synthetic data tests}
This package has been fully tested for $6,7,9,10$ parameter inversions based on synthetic data.

%======================================================

\appendix
\section{CMT3D Theory}
\subsection{Moment-tensor Inversions}
We seek to minimize the difference between data $\{d_i(t),i=1,\ldots,N\}$ and corresponding synthetics $\{s_i(\bbm,t),i=1,\ldots,N\}$  computed for a given model parameters $\bbm=\{m_j,j=1,\ldots,M\}$ after applying some bandpass filter:
\eq
\phi(\bbm)=\oneovertwo \sum_{rq}  \int\,W_{rq}(t)\left[ \bs^b(\bbm,\bx_r,t)-\bd^b_i(\bx_r,t)\right]^2\,dt
\en
where $W_{rq}(t)$ combines the taper function with weighting factor for the q'th window at r'th receiver. $\bs^b$ and $\bd^b$ denote filtered data and synthetics.

If we take the first and second order derivatives of this misfit function at some given reference model $\bbm^0$, we obtain
\eqa
\papa{\phi(\bbm^0)}{m_j} &=& \sum_{i=rq} \int\,W_{rq}(t)\left[ \bs^b(\bbm^0,\bx_r,t)-\bd^b(\bx_r,t)\right]\papa{s_i(\bbm^0,\bx_r,t)}{m_j}\,dt \\
\papapan{\phi(\bbm^0)}{m_j}{m_k} &\sim& \sum_{i=rq}  \int\,W_{rq}(t)\papa{\bs^b(\bbm^0,\bx_r,t)}{m_j}\papa{\bs^b(\bbm^0,\bx_r,t)}{m_k}\,dt
\label{eq:first-second-der}
\ena
where for the second-order derivatives, we have omitted the term related to second-order synthetics derivatives, i.e. approximate Hessian

By assuming that at the optimal model parameter $\bbm$, we have $\papa{\phi}{\bbm}=0$, which translates to the quasi-Newton's method:
\eq
\papa{\phi(\bbm^0)}{m_j}+\papapan{\phi(\bbm^0)}{m_j}{m_k}(m_k-m_k^0) = 0
\en
Therefore, we just need to compute the first and second-order derivatives according to equations \ref{eq:first-second-der}, which further points to the computation of derivative synthetic seismograms $\papa{s_i(\bbm^0)}{m_j}$.
%======================================================

\subsection{Derivative Synthetics}
We know in general, the synthetics for model $\bbm$ can be expanded with respect to a reference model $\bbm^0$:
\eq
s(\bbm,t)=s(\bbm^0,t) + \sum_{j=1}^M \papa{s_i(\bbm^0,t)}{m_j} (m_j-m_j^0)
\en
If source parameters only include moment-tensor elements, then the linear dependences can be written as:
\eq
s(\bbm,t,t) = \sum_{j=1}^{M}\papa{s_i(\bbm^0,t)}{m_j} m_j
\en
Therefore derivatives for moment-tensor elements can be computed by
\eq
\papa{s_i(\bbm^0,t)}{M_j} = s(M_j^0,t)/M_j^0
\en
where $s(M_j^0,t)$ is the synthetics calculated by setting all the moment tensor elements other than the $j$th component to $0$.

For other source parameters, including event location and depth, a finite-difference formula has to be used:
\eq
\papa{s_i(\bbm^0,t)}{m_j} = \frac{s(m_j^0+\De m_j^0,t)-s(m_j^0,t)}{\De m_j}
\en
where $s(m_j^0+\De m_j^0,t)$ refers to synthetics computed from only perturbing the $j$'th component of the source parameters. $\Delta m_j$ needs to be selected with care to avoid numerical noise. Typically $2-5$ km for global earthquakes and $1$ km for local earthquakes.

Derivatives for time shift and half duration can be derived from the original synthetics itself. Given a Gaussian source time function
\eq
g(t) = \frac{1}{\sqrt{\pi}t_h}e^{-\left(\frac{t-t_s}{t_h}\right)^2} \quad
\text{or}\quad g(\om) = e^{-\frac{\om^2 t_h^2}{4}} e^{-i\om t_s}
\en
under the Fourier transformation pairs:
\eq
F(\om) = \int_{-\infty}^{\infty} f(t) e^{-i\om t}dt, \quad f(t+a) \Leftrightarrow F(\om) e^{i\om a},
\quad \dot{f}(t) \Leftrightarrow (i\om) F(\om)
\en
Given Green's functions for a point source with delta moment function, we can express the synthetics for an earthquake as
\eq
\bs(\bx_r,\om) = \frac{M}{i\om} : \nab G(\bx_r,\bx_s,\om) g(\om)
\en
i.e., synthetics for error function as moment function.  Therefore
\eq
\papa{\bs(\bx_r,\om)}{t_{s,h}} = \frac{M}{i\om} : \nab G(\bx_r,\bx_s,\om)\papa{g(\om)}{t_{s,h}}.
\en
Since
\eqa
\papa{g(\om)}{t_s} &=& -(i\om) g(\om)  \nn \\
\papa{g(\om)}{t_h} &=& -\frac{1}{2} t_h \om^2 g(\om) = -\oneovertwo t_h (i\om) \papa{g(\om)}{t_s}
\ena
we have
\eqa
\papa{s(\bbm_0,t)}{t_s} &=& -\dot{\bs}(\bbm_0,t) \nn \\
\papa{s(\bbm_0,t)}{t_h} &=& -\frac{1}{2}t_h \pa_{t} [\papa{s(\bbm_0,t)}{t_s}]
\ena
%======================================================
\subsection{Rotation between Global and Local coordinates}
A series rotation is needed to tranform local $(r,\tht,\phi)$ coordinates
(which is also what CMT moment tensor uses) into the global coordinates $(X,Y,Z)$, where $X$ is pointing to $\Deg{0}$ longitude, and vice versa.

For a given point with $(\tht=90-Lat,\phi=Long)$ on a \textbf{unit} sphere, the coordinates in the cartesian global coordinate system is given by
\eq
X=\sin\tht\cos\phi,\quad Y=\sin\tht\sin\phi,\quad Z=\cos\tht
\en
Conversely, given the global coordates $(X,Y,Z)$,
\eq
\tht = acos(Z/r),\quad  \phi=atan2(Y/\sin\tht, X/\sin\tht)
\en
Note here if $\tht=0$ (North pole), then $\phi$ is undetermined, and can be basically any number.

For rotation matrices, as an example, in 2-D, for the same point, the coordinates under new coordinate system is related to the old coordinates by
\eq
\begin{bmatrix} X_{new} \\ Y_{new}\end{bmatrix}
= \begin{bmatrix} \cos\tht & \sin\tht \\ -\sin\tht & \cos\tht \end{bmatrix}
\begin{bmatrix} X_{old} \\ Y_{old}\end{bmatrix}
\en
\begin{figure}[h]
\putfigure{5in}{rot.png}
\caption{Rotation matrices}
\end{figure}
After the first two simple rotations ($Z (\phi)$ and $Y(\tht)$) demonstrated by Figure 1, we have the relation between $(X,Y,Z)$ and $(\tht,\phi,r)$ systems:
\eq
\begin{bmatrix} x_{\tht} \\ x_{\phi} \\ x_{r} \end{bmatrix}
= \begin{bmatrix} \cos\tht & 0 & - \sin\tht \\
0 & 1 & 0 \\ \sin\tht & 0 & \cos\tht \end{bmatrix}
\begin{bmatrix} \cos\phi & \sin\phi & 0 \\
-\sin\phi & \cos\phi & 0 \\ 0 & 0 & 1 \end{bmatrix}
\begin{bmatrix} X \\ Y \\ Z\end{bmatrix}
\en
This is not the end of story, as the usual order of local coordinate system is $(r,\tht,\phi)$ instead of $(\tht,\phi,r)$. Therefore two extra rotations ($Y(-\Deg{90})$ and $X (-\Deg{90})$) are needed, which give an extra (left-applied) rotation matrix
\eq
\begin{bmatrix} 1 & 0  &  0 \\
0 & \cos(-\Deg{90}) & \sin(-\Deg{90}) \\
0 & -\sin(-\Deg{90}) & \cos(-\Deg{90}) \end{bmatrix}
\begin{bmatrix} \cos(-\Deg{90}) & 0 & -\sin(-\Deg{90}) \\
0 & 1 & 0 \\
\sin(-\Deg{90}) & 0  & \cos(-\Deg{90}) \end{bmatrix}
\en
Multiply these two matrices together, we obtain the final rotation matrix from $(X,Y,Z)$ to $(r,\tht,\phi)$:
\eq
R^T=
\begin{bmatrix}
\sin\tht\cos\phi & \sin\tht\sin\phi  &  \cos\tht \\
\cos\tht\cos\phi & \cos\tht\sin\phi & -\sin\tht \\
-\sin\phi & \cos\phi & 0 \end{bmatrix}
\en
which can be verified by
\eq
\begin{bmatrix} 1 \\ 0 \\ 0\end{bmatrix}=
R^T
\begin{bmatrix} \sin\tht\cos\phi \\ \sin\tht\sin\phi \\ \cos\tht\end{bmatrix}
\en
The rotation matrix from local $(r,\tht,\phi)$ to global $(X,Y,Z)$ is simply:
\eq
R=\begin{bmatrix}
\sin\tht\cos\phi & \cos\tht\cos\phi  &  -\sin\phi \\
\sin\tht\sin\phi & \cos\tht\sin\phi & \cos\phi\\
\cos\tht & -\sin\tht & 0 \end{bmatrix}
\en
%======================================================

%============================================

\section{Testing (further update awaits)}

Two types of datasets are provided to accurately test this inversion package, one with station correction, and one without. Synthetics and derivative synthetics for corresponding \verb=CMTSOLUTION= are always in \verb=syn_T006_T030=. No pre-processing are needed for these dataset as they have gone through all the procedures. Although to achieve inversion result that is accurate to several digits after the decimal point, several tweaks with the usual pre-processing procedures have to be done.

\begin{itemize}
\item  \blue{(obsolete, needs fix)} we should not use \verb=-h= in the synthetics processing to ensure that the synthetics data and the derivative synthetics go through the exact same set of processing procedures. we also need to take out the \verb=rtrend= and \verb=rmean= line in the filtering steps of \verb=process_syn_new.pl= and \verb=process_cal_data.pl= (\red{?}). Since the data and synthetics presented in \verb=test_dataset= have gone through the script:
\begin{verbatim}
  process_data_and_syn.pl  -d data,sac -s syn,  \
         -0 -m CMTSOLUTION -a station_file   \
         -1 -b -26 -2 -t 6/30
\end{verbatim}
which generated \verb=process_ds.csh= script for all the commands, we can just tweak individual lines in \verb=process_ds.csh= to achieve our test dataset. Also
note that the \verb=-i none= option for data processing and \verb=-h= option for synthetics processing have to be taken out from the script to make sure that synthetics and data are processed in exactly the same way.

\item The synthetic data for the 7 paramter case are generated by the following command:
\begin{verbatim}
  xadd_frechet_full s syn_file_list old_cmt true_cmt \
                    doment depth dx[dy]
\end{verbatim}
which is a fancier version of the original
\begin{verbatim}
  xadd_frechet_derivatives s syn_file_list \
                           cmt_file moment_scale
\end{verbatim}
Of course we can't really use the true synthetics for the true cmt solution in a 7-par case, since we know that the synthetics is a non-linear function of depth, and the inversion won't accurately recover depth. Linearized `pseudo-synthetics' will ensure accurate results, which is the purpose of testing.

\item It is impossible to design test dataset for 7 parameter inversion with time shift, since time shifts interfer with the effect of waveform change caused by depth variation, therefore, may not give exact solution, and sometimes may even produce the incorrect solution. This is part of the reason why I also wrote the grid search package \verb=GRID3D_FLEXWIN=.
\end{itemize}
%=======================================================

\section{Auxilliary Scripts and Programs}
A list of the scripts that are used before and after the package:
\begin{itemize}
\item \verb=process_data_and_syn.pl=, which includes \\
 \verb=process_data_new.pl=(global/regional) or \verb=process_cal_data.pl= (socal), \\
 \verb=process_syn_new.pl= (global/regional) or \verb=process_trinet_syn_new.pl=(socal), \\
 \verb=pad_zeros.pl= (include \verb=pad_zero_to_syn=), \verb=rotate.pl=, \verb=saclst=. Of course \verb=sac= has to be present in the system.
\item \verb=prepare_meas_all.pl=
\item \verb=xadd_frechet_derivatives= and \verb=xadd_frechet_full=
\end{itemize}

\end{document}

### 2017-06-24 src/inverse_problem_for_source/CMT3D/grid3d/readme/readme.tex
re-added src/inverse_problem_for_source/CMT3D as a local copy instead of as a submodule

%
%=======================================
%  template.tex
%=======================================
%
\documentclass[12pt,titlepage,fleqn]{article}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{latexsym}
\usepackage[round]{natbib}
\usepackage{xspace}
%\usepackage{epsfig}
\usepackage{graphicx}
\usepackage{bm}

% SPACING COMMANDS (Latex Companion, p. 52)
\usepackage{setspace}
\renewcommand{\baselinestretch}{1.0}
\textwidth 470pt
\textheight 690pt
\oddsidemargin 0pt
\evensidemargin 0pt

% see Latex Companion, p. 85
%\voffset     -50pt
\topmargin     0pt
\headsep      20pt
\headheight   15pt
\headheight    0pt
\footskip     30pt
\hoffset       0pt

% nicer fonts
\usepackage{times}

% self-defined commands
\include{newcommands}

% bibliography
\bibliographystyle{gji}

%======================================================
\begin{document}
%======================================================
\begin{spacing}{1.0} % empty lines are required!
\begin{center}

\Large {\vspace{0.1in} \bf \verb=GRID3D_FLEXWIN= Readme}
\end{center}

\end{spacing}

\verb=GRID3D_FLEXWIN= is re-written from my old \verb=GRID3D= program to adapt to the output of Alessia's \verb=FLEXWIN= package, and provide an unbiased initial solution for the \verb=cmt3d_FLEXWIN= program, although this best solution is only for a fixed depth. Notice the pre-processing step to pick the usable data and synthetic windows is exactly the same as stated in section 1 of the \verb=cmt3d_FLEXWIN= Readme file.
%=====%======================================================

\section{Parameter file}

For the structure of the program and subroutine calls, refer to \verb=call-graph=. The name of the parameter file is hard-wired in the main program: \verb=GRID3D.PAR=.

\begin{verbatim}
CMTSOLUTION
1.0e22                         -- dmoment
flexwin_T006_T030.output
.false.                        -- weigh_data_files
2 2 1  0  1.15 0.55 0.78       -- weights of data comp,az,dist
.false.                        -- station_correction
.true. 3                       -- global_search, ncalc
10 16 0.5                      -- strike
20 26 0.5                      -- dip
30 36 0.5                      -- rake
4.2 4.4 0.1                    -- Mw, only valid with local search
.true.                         -- write cmt-file-new
\end{verbatim}

The first line gives a cmt solution file name. Unlike \verb=cmt3d_flexwin=, this is only used to obtain a reasonable starting $M_w$ for global grid search, therefore may be elliminated in the future. However, since we are not short of initial solutions, I decided to keep it there. The data weighting procedures are exactly the same as outlined in the readme file for \verb=cmt3d_flexwin=. Station correction must be allowed for actual data (as oppose to synthetic data), which otherwise defeats the purpose of this program, although significantly longer search time should be expected with this extra operation.

Global search goes through \verb-strike=0:10:180-, \verb-dip=0:10:90-, \verb?rake=-180:30:180?, \verb-Mw=0.9:0.05:1.1*mw- in search for an optimal misfit value. Therefore the values listed in the next 4 lines are only used when \verb-global_search=.false.-, and ignored when \\
\verb-global_search=.true.-. Also \verb-ncalc- is only significant in the case of global search, where iteratively refined searches are done to pinpoint the global minimum.

\begin{itemize}
\item The output \verb=CMTSOLUTION_NEW= file can be copied directly to \verb=cmt3d_flexwin= as an initial cmt solution, and the corresponding synthetics should be calculated by the program \verb=xadd_frechet_derivatives=.
\item No synthetics is needed (derivative synthetics of course have to be present), since all the synthetic header information is taken from \verb=syn.Mrr= files.
\item New cmt solution will bear the name \verb-CMTSOLUTION_NEW-, which is hard-wired.
\item A typical search array size can be 20x20x40x5 for strike,dip,rake and Mw. although it is their product (80000) that really restricts the number of double-couple moment tensor sources that can be searched.
\end{itemize}

%============================================

\section{Plotting}

Although the program will output the best strike,dip,rake,Mw that ive rises to the global or local minimum of the misfit function, it is much more straightforward to read that information off some graphs. These graphs can also display visually the trade-off between different parameters, therefore, are more useful in terms of understanding the significance of our minimum solution.

The output files from the \verb-grid3d_flexwin- package have the names
\begin{verbatim}
 grid3d_misfit[.icalc].n_mw
\end{verbatim}
Each file documents the strike,dip,rake and corresponding minimum value (normalized by minimums for all Mw) for a particular choice of $M_w$, as well as a particular iteration of global search. I have written an auxilliary program to help produce plots to visualize this 3D function:
\begin{verbatim}
 plot_grid.pl grid-output-files
\end{verbatim}
For each output file, 6 strike, dip and rake are selected close to the global minimum, and corresponding slices are taken from the grid file. Each slice is plotted and contoured to show how the misfit varies as a function of two variables with other variables fixed. This script is by no means general, since it is very hard to write general GMT programs that will fit all cases. But the users can play around either with this perl script or the ouput bash file for fine tuning. The perl script used some perl libary files that are available for Caltech/GPS users. A sample output is attached to this readme.

\begin{figure}
\begin{center}
     \includegraphics[width=6in]{grid_0.eps}%
  \end{center}
\end{figure}

%=======================================================

\end{document}

### 2017-06-24 src/inverse_problem_for_source/CMT3D/cmt3d/cmt3d_utils.f

@@ -1,0 +1,397 @@
+c This subroutine takes the value of a moment tensor
+c and outputs the strike-dip-rake of the best-fitting 
+c double couple, and their moment. 
+
+	 subroutine mij(m,epsilon,s1,d1,r1,s2,d2,r2,m0,m00,iflag)
+	 implicit  REAL*8(A-H,O-Z)
+
+	 dimension a(3,3),d(3),v(3,3),ae(3),an(3)
+	 dimension ae1(3),an1(3),an2(3),ae2(3)
+	 real*8  m(6), m0, m00
+	 common/mpa/ft,fd,fl,ft1,fd1,fl1,azt,aiht,azp,aihp,az0,aih0,
+     *d,clvd
+
+         if (iflag == 0) then ! Hiroo's convention
+            a(1,1)=m(1)
+            a(2,2)=m(2)
+            a(3,3)=m(3)
+            a(1,2)=m(4)
+            a(1,3)=m(5)
+            a(2,3)=m(6)
+            a(2,1)=a(1,2)
+            a(3,1)=a(1,3)
+            a(3,2)=a(2,3)
+         else                 ! Harvard's convention
+            a(1,1)=m(2)
+            a(2,2)=m(3)
+            a(3,3)=m(1)
+            a(1,2)=m(6)
+            a(1,3)=-m(4)
+            a(2,3)=-m(5)
+            a(2,1)=a(1,2)
+            a(3,1)=a(1,3)
+            a(3,2)=a(2,3)
+         endif
+
+c        calculate the full moment
+         m00 =0.
+         do i = 1, 3
+            do j = 1, 3
+               m00 = m00 + a(i,j) ** 2
+            enddo
+         enddo
+         m00 = sqrt(m00 /2 )
+c       take eigen values and eigen vectors
+	 call jacobi(a,3,3,d,v,nrot)
+11	 format(1x,i3,3(f5.2,2x))
+
+c       subtract isotropic part
+         sum = 0.
+         do i = 1, 3
+            sum = sum + d(i)
+         enddo
+         do i = 1, 3
+            d(i) = d(i) - sum/3
[…]

### 2017-06-24 src/inverse_problem_for_source/CMT3D/scripts/mech_table2.pl

@@ -1,0 +1,133 @@
+#!/usr/bin/perl
+use lib '/opt/seismo/lib/perl';
+use GMT_PLOT;
+use GMT_PLOT_SC;
+use CMT_TOOLS;
+use POSIX;
+use List::Util qw[max];
+
+## ok, this version includes both 2, 3 and 6 seconds measurements
+# parameter files
+$inv_file="INVERSION.PAR.SAVE";
+$outdir="mech"; $psfile="mech.ps";
+
+$weigh_data="2 2 1 0.5 1.15 0.55 0.78";
+$weigh_data_nocomp="1 1 1 0.5 1.15 0.55 0.78";
+$weigh_data_noaz="2 2 1 0 1.15 0.55 0.78";
+
+if (not -f "CMTSOLUTION") {die("Check cmt solution file\n");}
+($evid)=split(" ",`cat CMTSOLUTION | awk 'NR == 2 {print \$3}'`);
+$mwin="MEASUREMENT_WINDOWS";
+
+@names=("Initial","Grid Search","7 Par+ZT",
+        "7 Par+no-SC","6 par+DC","7 Par+DC",
+        "7 Par(2-30s)","7 Par(3-30s)","7 Par(6-30s)");
+@npars=(6,6,7, 7,6,7, 7,7,7);
+@wdatas=($weigh_data, $weigh_data, $weigh_data, $weigh_data,$weigh_data,$weigh_data, $weigh_data,$weigh_data,$weigh_data);
+@scorrs=("true","true",".true.",  ".false.", ".true.", ".true.",  ".true.",".true.",".true.");
+@cons=(".true. true. 0.0", ".true. .true. 0.0", ".true. .false. 0.0",
+       ".true. .false. 0.0", "true. .true. 0.0", "true. .true. 0.0",
+       ".true. .true. 0.0",".true. .true. 0.0",".true. .true. 0.0");
+@meas=("${mwin}_${evid}_ALL","${mwin}_${evid}_ALL","${mwin}_${evid}_ALL",
+       "${mwin}_${evid}_ALL","${mwin}_${evid}_ALL","${mwin}_${evid}_ALL",
+       "${mwin}_${evid}_T002_T030_m12","${mwin}_${evid}_T003_T030_m12","${mwin}_${evid}_T006_T030_m12");
+
+$nbeach=@npars;
[…]

### 2017-06-30 doc/USER_MANUAL/05_running_the_solver.tex
added Zhinan's table to the manual

@@
 and thus the code outputs the band in which the approximation is very good, outside of that range it can be less accurate.
-The logarithmic center of that frequency band is the \texttt{ATTENUATION\_f0} parameter defined (in Hz) in input file \texttt{DATA/Par\_file}.
+The logarithmic center of that frequency band is the \texttt{ATTENUATION\_f0} parameter defined (in Hz) in input file \texttt{DATA/Par\_file}. \\
+
+\noindent
+Regarding attenuation (viscoelasticity), in the Par\_file you need to select the number of standard linear solids (N\_SLS) to use to mimic a constant $Q$ quality factor.
+Using N\_SLS = 3 is always safe. If (and only if) you know what you are doing, you can try to reduce that in order to reduce the cost of the simulations.
+Figure~\ref{fig:selectNSLS} shows values that you can consider using (again, if and only if you know what you are doing). That table has been created by Zhinan Xie using
+a comparison between results obtained with a truly-constant $Q$ and results obtained with its approximation based on N\_SLS standard linear solids.
+The comparison is performed using the time-frequency misfit and goodness-of-fit criteria proposed by \cite{Kristekova_2009}.
+The table is drawn for a dimensionless parameter representing the distance of propagation.
+%%
+\begin{figure}[htbp]
+\centering
+\includegraphics[width=5in]{figures/minimum_number_of_SLS_that_can_be_used_in_viscoelastic_simulation.png}
+\caption{Table showing how you can select a value of N\_SLS smaller than 3, if and only if you know what you are doing.}
+\label{fig:selectNSLS}
+\end{figure}
+%%

### 2017-07-04 doc/USER_MANUAL/05_running_the_solver.tex
clarified the depth definition in doc/USER_MANUAL/manual_SPECFEM3D_Cartesian.pdf

@@
 \begin{itemize}
-\item Set the latitude or UTM $x$ coordinate, longitude or UTM $y$ coordinate, depth of the source (in km).
+\item Set the latitude or UTM $x$ coordinate, longitude or UTM $y$ coordinate, depth of the source (in km). Remark: In principle in the international CMTSOLUTION format in geophysics
+the depth is given in kilometers; however for users in other fields (non-destructive testing, medical imaging, near-surface studies...) who may prefer to give the position of the source (rather than its depth from the surface) we provide an option called USE\_SOURCES\_RECEIVERS\_Z in the Par\_file, and if so that position is read from CMTSOLUTION in meters rather than kilometers (and again, it is then the true position in the mesh, not the depth).
+When option USE\_SOURCES\_RECEIVERS\_Z in the Par\_file is on, this remark applies to the position of the receivers as well.
 \item Set the \texttt{time shift} parameter equal to $0.0$ (the solver

### 2017-07-17 doc/USER_MANUAL/05_running_the_solver.tex

--- 2017-06-30 doc/USER_MANUAL/05_running_the_solver.tex
@@
 \begin{itemize}
-\item Set the latitude or UTM $x$ coordinate, longitude or UTM $y$ coordinate, depth of the source (in km).
+\item Set the latitude or UTM $x$ coordinate, longitude or UTM $y$ coordinate, depth of the source (in km). Remark: In principle in the international CMTSOLUTION format in geophysics
+the depth is given in kilometers; however for users in other fields (non-destructive testing, medical imaging, near-surface studies...) who may prefer to give the position of the source (rather than its depth from the surface), or for people who use FORCESOLUTION to describe the source rather than CMTSOLUTION, we provide an option called USE\_SOURCES\_RECEIVERS\_Z in the Par\_file, and if so that position is read from CMTSOLUTION in meters rather than kilometers (and again, it is then the true position in the mesh, not the depth).
+When option USE\_SOURCES\_RECEIVERS\_Z in the Par\_file is on, this remark applies to the position of the receivers as well.
 \item Set the \texttt{time shift} parameter equal to $0.0$ (the solver
@@
 
-Procedure 1:\\
-============\\
+Procedure 1 (mostly for geophysics, when the top surface is a free surface (topography) and the five other edges of the mesh are absorbing surfaces):\\
+====================================================================================================================================================\\
 
-- Always put the origin on the top of the model.\\
+- Put the origin on the top of the model.\\
 
@@
 
-Depending on the geometry of the models it may not work in some cases.\\
+Procedure 2 (useful for other application domains, in which using the absolute $Z$ position of the sources and receivers is more standard than using their depth from the surface):\\
+==================================================================================================================================================================================
 
-Procedure 2: (less recommended)\\
-============\\
-
-- Open setup/constants.h.in file and change the line
-logical, parameter :: USE\_SOURCES\_RECEIVERS\_Z = .false.\\
-\\
-to\\
-\\
-logical, parameter :: USE\_SOURCES\_RECEIVERS\_Z = .true.\\
-\\
-configure the package and compile.\\
+- In principle in the international CMTSOLUTION format in geophysics
+the depth is given in kilometers; however for users in other fields (non-destructive testing, medical imaging, near-surface studies...) who may prefer to give the position of the source (rather than its depth from the surface), or for people who use FORCESOLUTION to describe the source rather than CMTSOLUTION, we provide an option called USE\_SOURCES\_RECEIVERS\_Z in the Par\_file, and if so that position is read from CMTSOLUTION in meters rather than kilometers (and again, it is then the true position in the mesh, not the depth).
+When option USE\_SOURCES\_RECEIVERS\_Z in the Par\_file is on, this remark applies to the position of the receivers as well.\\
 
 - Let's say you want to place two receivers at (x1,y1,z1) and
-(x2,y2,z2). Your STATIONS file should look like:\\
+(x2,y2,z2). Your STATIONS file should then look like:\\
 \\
@@
 \\
-The option set to .true. in 1 will discard the elevation and set
+The option USE\_SOURCES\_RECEIVERS\_Z set to .true. will then discard the elevation and set
 burial as the $z$ coordinate.\\

### 2017-07-21 doc/USER_MANUAL/04_creating_databases.tex
better sentence for f0_for_PML in the users manual

@@
 \item [{\texttt{f0\_FOR\_PML}}] Determines the dominant frequency that
-will be used in the calculation of PML damping profiles; this should
-be set to the same (or similar) dominant frequency as that of the
-source that you will use in your simulation. If you plan to use a
-Dirac source, then use the dominant frequency of the source wavelet
+will be used in the calculation of PML damping profiles;
+\red{This should be set to the same (or similar) dominant frequency as that of the
+source that you will use in your simulation. It is \textbf{VERY IMPORTANT} to do that, otherwise the PML 
+absorbing conditions can become unstable.}
+If you plan to use a Dirac source, then use the dominant frequency of the source wavelet
 with which you plan to convolve your seismograms later on in post-processing.

### 2017-10-02 doc/USER_MANUAL/05_running_the_solver.tex

@@
 the files  \texttt{FORCESOLUTION} and \texttt{CMTSOLUTION} otherwise the solver will exit with an error.
+When using an external source file, you can still set up the source location and directivity as in the default case.
+In the FORCESOLUTION file: you set "latorUTM", "longorUTM" and "depth" to define the position of your point source.
+Then if you want to define a directivity, then change the following lines: "component dir vect source E", "component dir vect source N" and "component dir vect source Z\_UP".
+What you are doing is simply that you define the source position and directivity the same way as in the default case, but in addition you are specifying the path to read in a non-default source time function from an external file.
 

### 2017-10-31 src/cuda/compute_forces_acoustic_cuda.cu
updated headers

@@ -5,10 +5,10 @@
  !               ---------------------------------------
  !
  !     Main historical authors: Dimitri Komatitsch and Jeroen Tromp
- !                        Princeton University, USA
- !                and CNRS / University of Marseille, France
+ !                              CNRS, France
+ !                       and Princeton University, USA
  !                 (there are currently many more authors!)
- ! (c) Princeton University and CNRS / University of Marseille, July 2012
+ !                           (c) October 2017
  !
  ! This program is free software; you can redistribute it and/or modify
  ! it under the terms of the GNU General Public License as published by

### 2017-11-13 doc/USER_MANUAL/manual_SPECFEM3D_Cartesian.tex
switched the license to GNU version 3 (now that we use pyCMT3D, which is LGPL v3, the combination automatically becomes GNU version 3). Also added FWI inversion information to the users manual.

@@
 \usepackage{breakurl}
+
+% BibTeX logo
+\newcommand{\BibTeX}{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em{T}\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
 
@@
 
-\title{\textbf{SPECFEM3D Cartesian}\\
+\title{\textbf{SPECFEM3D}\\
+\textbf{Cartesian}\\
  \textbf{User Manual}}
 
-\author{\copyright Princeton University (USA), CNRS and University of Marseille (France),\\
+\author{\copyright \, CNRS (France), Princeton University (USA),\\
 and ETH Z\"urich (Switzerland)\\
@@
 \include{08_doing_tomography}
+
+\include{08b_performing_full_waveform_inversion_FWI_or_source_inversions}
 

### 2018-02-10 doc/USER_MANUAL/02_getting_started.tex
added comment about compiler bug in some releases of Intel ifort version 17 when option -assume buffered_io is used

@@
 \textbf{In the case of that compiler, we have noticed that initial release versions sometimes have bugs or issues that can lead to wrong results when running the code, thus we \emph{strongly} recommend using a version for which at least one service pack or update has been installed.}
+\red{In particular, for version 17 of that compiler, users have reported problems (making the code crash at run time) with the \texttt{-assume buffered\_io} option; if you notice problems,
+remove that option from file \texttt{flags.guess} or change it to \texttt{-assume nobuffered\_io} and try again.}
 \item [{\texttt{IBM compiler}}] See if you need to add \texttt{-qsave}

### 2018-02-11 doc/USER_MANUAL/01_introduction.tex
added description of Hourglass modes that can appear in some cases in the source element to the users manual

@@
 while in the Chebyshev version it is exact but non diagonal.\\
+
+\red{Beware that, in a spectral-element method, some spurious modes (that have some similarities with classical so-called "Hourglass modes" in finite-element techniques,
+although in the SEM they are not zero-energy modes) can appear in some (but not all) cases in the spectral element in which the source is located.
+Fortunately, they do not propagate away from the source element.
+However, this means that if you put a receiver in the same spectral element as a source, the recorded signals may in some cases be wrong, typically exhibiting some spurious
+oscillations, which are often even non causal.
+If that is the case, an easy option is to slightly change the mesh in the source region in order to get rid of these Hourglass-like spurious modes,
+as explained in \cite{DuLiScGa14}, in which this phenomenon is described in details, and in which practical solutions to avoid it are suggested.}\\
 

### 2018-04-09 doc/USER_MANUAL/01_introduction.tex
added paragraph from Surendra Nadh Somala to the users manual

@@
 See flag \texttt{GRAVITY\_INTEGRALS} in file \texttt{setup/constants.h.in}. Please also refer to \url{http://komatitsch.free.fr/preprints/GJI_Martin_gravimetry_2017.pdf}.
-And yes, that is the reason why I added a falling apple on the cover of the manual :-)}\\
+And yes, that is the reason why I added a falling apple on the cover of the manual :-).
+Note that SPECFEM3D can also model transient gravity perturbations induced by earthquake rupture, as developed and explained in \cite{HaAmBaChMoSoWh2015}.
+These are two different things, and both are implemented and avaible in SPECFEM3D.
+To use the second feature, please refer to doc/how\_to/README\_gravity\_perturbation.txt.}\\
 

### 2018-04-20 doc/USER_MANUAL/01_introduction.tex
added description of DSM and FK coupling to the users manual

@@
 
-\red{\Large ANNOUNCEMENT: SPECFEM3D can now perform full waveform inversion (FWI), i.e. invert for models in an iterative fashion, and it can also perform source inversions in a constant model;
+\red{FIRST ANNOUNCEMENT: SPECFEM3D can now perform full waveform inversion (FWI), i.e. invert for models in an iterative fashion, and it can also perform source inversions in a constant model;
 please refer to the two new directories, \texttt{inverse\_problem\_for\_model} and \texttt{inverse\_problem\_for\_source}, and the README files they contain.
@@
 
-\red{OTHER ANNOUNCEMENT: SPECFEM3D can now perform gravity field calculations in addition (or instead of) seismic wave propagation only.
+\red{SECOND ANNOUNCEMENT: SPECFEM3D can now perform coupling with an external code (DSM, AxiSEM or FK) based on a database of displacement vectors and traction vectors
+on the outer edges of the mesh created once and for all (see \cite{MoChKoFu13,MoChKoWa15,WaChMoKoMo16,ToChKoBaLi14,ToKoTsHuChBaLi14}, and if you use that feature of the code please cite at least one of these papers).} \\
+To use coupling with FK, just use the set of parameters that is in the DATA/Par\_file input file of the code: \\
+\#----------------------------------------------------------- \\
+\# \\
+\# Coupling with an injection technique (DSM, AxiSEM, or FK) \\
+\# \\
+\#----------------------------------------------------------- \\
+COUPLE\_WITH\_INJECTION\_TECHNIQUE = .false. \\
+INJECTION\_TECHNIQUE\_TYPE        = 3   \# 1 = DSM, 2 = AxiSEM, 3 = FK \\
+MESH\_A\_CHUNK\_OF\_THE\_EARTH       = .false. \\
+TRACTION\_PATH                   = ./DATA/AxiSEM\_tractions/3/ \\
+FKMODEL\_FILE                    = FKmodel \\
+RECIPROCITY\_AND\_KH\_INTEGRAL     = .false.   \# does not work yet \\
+That part (coupling with FK) is actively maintained and works fine. See e.g. \url{http://komatitsch.free.fr/preprints/GRL_Ping_Tong_2014.pdf} for some examples.
+There is also an example that is provided with the code: specfem3d/EXAMPLES/small\_example\_coupling\_FK\_specfem. \\
+Regarding coupling with DSM, that part is not actively maintained any more, but it is still included in the code, you may have to test it again and make minor adjustments if needed.
+The necessary tools are in directory specfem3d/EXTERNAL\_PACKAGES\_coupled\_with\_SPECFEM3D/DSM\_for\_SPECFEM3D , and there is a README file in specfem3d/EXTERNAL\_PACKAGES\_coupled\_with\_SPECFEM3D that should be more or less up-to-date (there are about four steps to follow in total, the first one being creating the database of DSM tractions and displacements on the edges of the coupling box).
+\\
+See e.g. \url{http://komatitsch.free.fr/preprints/GJI1_Vadim_2013.pdf} and \url{http://komatitsch.free.fr/preprints/GJI2_Vadim_2015.pdf} for some examples.\\
+
+\red{THIRD ANNOUNCEMENT: SPECFEM3D can now perform gravity field calculations in addition (or instead of) seismic wave propagation only.
 See flag \texttt{GRAVITY\_INTEGRALS} in file \texttt{setup/constants.h.in}. Please also refer to \url{http://komatitsch.free.fr/preprints/GJI_Martin_gravimetry_2017.pdf}.

### 2018-05-11 doc/USER_MANUAL/04_creating_databases.tex
Buildbot ran the code beautifier.
added definition of the Ricker to the users manual

@@
 \item [{\texttt{USE\_RICKER\_TIME\_FUNCTION}}] Turn this flag on to use
-a Ricker source time function instead of the source time functions
+a Ricker source time function, i.e., the second derivative of a Gaussian, instead of the source time functions
 set by default to represent a (tilted) \texttt{FORCESOLUTION} force
-point source or a \texttt{CMTSOLUTION} moment-tensor source. Originally,
-if a \texttt{CMTSOLUTION} moment-tensor source is used, a (pseudo)
+point source or a \texttt{CMTSOLUTION} moment-tensor source.
+Note that we use the standard definition of a Ricker, for a dominant frequency $f_0$:
+$\mathrm{Ricker}(t) = (1 - 2 a t^2) e^{-a t^2}$, with $a = \pi^2 f_0^2$,
+whose Fourier transform is thus:
+$\frac{1}{2} \frac{\sqrt{\pi}\omega^2}{a^{3/2}}e^{-\frac{\omega^2}{4 a}}$
+This gives the wavelet of Figure~\ref{fig:RickerWavelet}.
+%%
+\begin{figure}[htbp]
+\centering
+\includegraphics[width=3in]{figures/Ricker_wavelet.png}
+\caption{We use the standard definition of a Ricker (i.e., second derivative of a Gaussian). Image taken from \url{http://subsurfwiki.org}.}
+\label{fig:RickerWavelet}
+\end{figure}
+%%
+Originally, if a \texttt{CMTSOLUTION} moment-tensor source is used, a (pseudo)
 Heaviside step function with a very short half duration is defined

### 2018-05-11 doc/USER_MANUAL/05_running_the_solver.tex
added definition of the Ricker to the users manual

@@
 \item Set the \texttt{f0} parameter (the dominant frequency) of the Ricker source time
-function when {\texttt{USE\_RICKER\_TIME\_FUNCTION}} is turned on
+function (i.e., the second derivative of a Gaussian) when {\texttt{USE\_RICKER\_TIME\_FUNCTION}} is turned on
 in the main parameter file {\texttt{Par\_file}}. In case that the
@@
 a force point source, a very short duration of five time steps is automatically set by default.
+Note that we use the standard definition of a Ricker, for a dominant frequency $f_0$:
+$\mathrm{Ricker}(t) = (1 - 2 a t^2) e^{-a t^2}$, with $a = \pi^2 f_0^2$,
+whose Fourier transform is thus:
+$\frac{1}{2} \frac{\sqrt{\pi}\omega^2}{a^{3/2}}e^{-\frac{\omega^2}{4 a}}$
+This gives the wavelet of Figure~\ref{fig:RickerWavelet}.
 \item Set the latitude or UTM $x$ coordinate, longitude or UTM $y$ coordinate, depth of the source (in km).

### 2018-05-30 doc/USER_MANUAL/authors.tex
fixed a typo in the users manual
added Masaru Nagaso as author, and added Windows installation information

@@
 Surendra Nadh Somala,
+Masaru Nagaso,
 Tarje Nissen-Meyer,
@@
 
-The code is released open-source under the GNU version 2+ license, see the license at the end of this manual.\\
+The code is released open-source under the GNU version 3 license, see the license at the end of this manual.\\

### 2018-06-04 doc/USER_MANUAL/06_fault_sources.tex
documented cohesion in slip-weakening friction

@@
 below):
-
 \begin{description}
@@
 \end{description}
+
 \item [{\&\textbf{SWF}}] input block sets the slip-weakening friction parameters
 of the fault:
-
 \begin{description}
@@
 \item [{dc}] = constant background value of critical slip-weakening distance
+\item [{C}] = constant background value of cohesion (in Pa)
 \item [{nmus}] = number of heterogeneous items for static friction coefficient
@@
 \item [{ndc}] = number of heterogeneous items for critical slip-weakening
-distance {[}default is 0{]}\\
-
+distance {[}default is 0{]}
+\item [{nC}] = number of heterogeneous items for cohesion
 \end{description}
@@
 of the fault:
-
 \begin{description}
@@
 To activate this feature, in Par\_file\_faults name list \&\textbf{RUPTURE\_SWITCHES}, set RSF\_HETE=.TRUE..
-
   The format of  {\textbf{DATA/rsf\_hete\_input\_file.txt}} is as such, in the first line there are four integers
-
   that are sequentially documenting the number of divisions along strike, number of divisions along
-
-  dip, cell size along strike , cell size along dip. Then the following N(N=NumberOfDivisionsAlongStrike
-
+  dip, cell size along strike , cell size along dip. Then the following N (N=NumberOfDivisionsAlongStrike
   * NumberOfDivisionsAlongDip) lines will document the stress and friction properties on the grid.
-
   There are a total of 13 columns.

### 2018-06-08 doc/USER_MANUAL/11_graphics.tex
added a sentence about ParaView to the users manual
fixed the USE_VTK_INSTEAD_OF_MESH issue reported by Huihui

@@
 to visualized these partitions.
+
+{\red To visualize seismograms with Paraview, you should turn off the flag SU\_FORMAT, and turn on the use binary seismograms flag.
+This will generate .bin files, that you can open with Paraview.
+You have to select the way you open it ("raw"), and specify the dimensions of the file (NSTEP and NREC).}
 
@@
 
-At the beginning of the source file \texttt{combine\_vol\_data.f90} (located in \texttt{src/auxiliaries/}), there is a logical constant parameter \texttt{USE\_VTK\_OUTPUT} whose default value is \texttt{.true.}, that which will provide the output combined file directly in the VTK format and you can use Paraview to visualize it.
+At the beginning of the source file \texttt{src/auxiliaries/combine\_vol\_data\_impl.f90} there is a logical parameter \texttt{USE\_VTK\_INSTEAD\_OF\_MESH},
+whose default value is \texttt{.false.}, which can be used to generate the combined output file directly in VTK format, which can then be visualized using Paraview.
 
@@
 
-If the value of this parameter is \texttt{.false.}, the output mesh file will have the name \texttt{velocity\_Z\_it000400.mesh}. Then we next have to convert the
-\texttt{.mesh} file into the VTU (Unstructured grid file) format which can be viewed in ParaView.
+If the value of this parameter is \texttt{.false.}, the output mesh file will have a name of the form \texttt{velocity\_Z\_it000400.mesh}.
+One then has to convert the \texttt{.mesh} file into the VTU (Unstructured grid file) format which can be viewed in ParaView.
 
@@
 For this task, you can use and modify the
-script \texttt{mesh2vtu.pl} located in directory \texttt{\small utils/Visualization/Paraview/}, for example:
+script \texttt{mesh2vtu.pl} located in directory \texttt{\small utils/Visualization/Paraview}, for example:
 
@@
 Notice that this Perl script uses a program \texttt{mesh2vtu} in the \texttt{utils/Visualization/Paraview/mesh2vtu} directory, which further
-uses the VTK \urlwithparentheses{http://www.vtk.org/} run-time library for its execution. Therefore, make sure you have them properly set
+uses the VTK \urlwithparentheses{http://www.vtk.org} run-time library for its execution. Therefore, make sure you have them properly set
 in the script according to your system.

### 2018-06-08 doc/USER_MANUAL/11_graphics.tex
reverted the change made yesterday for the VTK executable. It turns out there was no bug, it was just the users manual that was saying something wrong (i.e. obsolete).

@@
 
-\medskip
-
-At the beginning of the source file \texttt{src/auxiliaries/combine\_vol\_data\_impl.f90} there is a logical parameter \texttt{USE\_VTK\_INSTEAD\_OF\_MESH},
-whose default value is \texttt{.false.}, which can be used to generate the combined output file directly in VTK format, which can then be visualized using Paraview.
-
-\medskip
-
-If the value of this parameter is \texttt{.false.}, the output mesh file will have a name of the form \texttt{velocity\_Z\_it000400.mesh}.
-One then has to convert the \texttt{.mesh} file into the VTU (Unstructured grid file) format which can be viewed in ParaView.
+When compiling the code, you will get two executables related to this: xcombine\_vol\_data, and xcombine\_vol\_data\_vtk.
+Use the executable that has the \_vtk extension if you want to directly create VTK files.
+If you use the executable that does not have the \_vtk extension, then the output mesh file will have a name of the form \texttt{velocity\_Z\_it000400.mesh}.
+You will then have to convert the \texttt{.mesh} file into the VTU (Unstructured grid file) format which can be viewed in ParaView.
 

### 2018-06-15 doc/USER_MANUAL/05_running_the_solver.tex
nice bug fix by Lei Zhang in src/shared/read_parameter_file.F90

@@
 activated by setting   \texttt{EXTERNAL\_SOURCE\_FILE} to true in \texttt{Par\_File} and by adding the name of the file
-containing the source time function at the end of \texttt{FORCESOLUTION} or \texttt{CMTSOLUTION} files. The source time function file must
-contain 2 columns with the current time and the amplitude of the soure function. The time step must be the same as used for the
-simulation. Note  when the  \texttt{EXTERNAL\_SOURCE\_FILE} is set to false then the line with the external source time function file must not appear in
+containing the source time function at the end of \texttt{FORCESOLUTION} or \texttt{CMTSOLUTION} files.
+The source time function file must contain the value of the time step on its first line, and then a single column with the amplitude of the source time function for all the time steps.
+The time step must be exactly the same as that used for the simulation.
+Note  when the  \texttt{EXTERNAL\_SOURCE\_FILE} is set to false then the line with the external source time function file must not appear in
 the files  \texttt{FORCESOLUTION} and \texttt{CMTSOLUTION} otherwise the solver will exit with an error.

### 2018-07-26 doc/USER_MANUAL/01_introduction.tex
Buildbot ran the code beautifier.
updates doc

@@
 To use coupling with FK, just use the set of parameters that is in the DATA/Par\_file input file of the code: \\
-\#----------------------------------------------------------- \\
-\# \\
-\# Coupling with an injection technique (DSM, AxiSEM, or FK) \\
-\# \\
-\#----------------------------------------------------------- \\
-COUPLE\_WITH\_INJECTION\_TECHNIQUE = .false. \\
-INJECTION\_TECHNIQUE\_TYPE        = 3   \# 1 = DSM, 2 = AxiSEM, 3 = FK \\
-MESH\_A\_CHUNK\_OF\_THE\_EARTH       = .false. \\
-TRACTION\_PATH                   = ./DATA/AxiSEM\_tractions/3/ \\
-FKMODEL\_FILE                    = FKmodel \\
-RECIPROCITY\_AND\_KH\_INTEGRAL     = .false.   \# does not work yet \\
+%
+\begin{verbatim}
+#-----------------------------------------------------------
+#
+# Coupling with an injection technique (DSM, AxiSEM, or FK)
+#
+#-----------------------------------------------------------
+COUPLE_WITH_INJECTION_TECHNIQUE = .false.
+INJECTION_TECHNIQUE_TYPE        = 3       # 1 = DSM, 2 = AxiSEM, 3 = FK
+MESH_A_CHUNK_OF_THE_EARTH       = .false.
+TRACTION_PATH                   = ./DATA/AxiSEM_tractions/3/
+FKMODEL_FILE                    = FKmodel
+RECIPROCITY_AND_KH_INTEGRAL     = .false. # does not work yet
+\end{verbatim}
 That part (coupling with FK) is actively maintained and works fine. See e.g. \url{http://komatitsch.free.fr/preprints/GRL_Ping_Tong_2014.pdf} for some examples.
@@
 Regarding coupling with DSM, that part is not actively maintained any more, but it is still included in the code, you may have to test it again and make minor adjustments if needed.
-The necessary tools are in directory specfem3d/EXTERNAL\_PACKAGES\_coupled\_with\_SPECFEM3D/DSM\_for\_SPECFEM3D , and there is a README file in specfem3d/EXTERNAL\_PACKAGES\_coupled\_with\_SPECFEM3D that should be more or less up-to-date (there are about four steps to follow in total, the first one being creating the database of DSM tractions and displacements on the edges of the coupling box).
+The necessary tools are in directory \texttt{specfem3d/EXTERNAL\_PACKAGES\_coupled\_with\_SPECFEM3D/DSM\_for\_SPECFEM3D}, and there is a \texttt{README} file in \texttt{specfem3d/EXTERNAL\_PACKAGES\_coupled\_with\_SPECFEM3D} that should be more or less up-to-date (there are about four steps to follow in total, the first one being creating the database of DSM tractions and displacements on the edges of the coupling box).
 \\
@@
 These are two different things, and both are implemented and avaible in SPECFEM3D.
-To use the second feature, please refer to doc/how\_to/README\_gravity\_perturbation.txt.}\\
+To use the second feature, please refer to \texttt{doc/how\_to/README\_gravity\_perturbation.txt}.}\\
 

### 2018-08-30 doc/USER_MANUAL/02_getting_started.tex
added Masaru Nagaso as author, and added Windows installation information

@@
 %
+\noindent
+\red{Note: for people who would like to run the package on Windows rather than on Unix machines, you can install Docker or VirtualBox (installing a Linux in VirtualBox in that latter case) and run it easily from inside that.}\\
+
 \noindent
@@
 \texttt{MPI\_INC=\${CRAY\_MPICH2\_DIR}/include and FCLIBS=" "}, and for more details if needed you can refer to the \texttt{utils/Cray\_compiler\_information} directory.
+You can also have a look at the configure script called \texttt{utils/Cray\_compiler\_information/configure\_SPECFEM\_for\_Piz\_Daint.bash}.
 

### 2018-11-07 doc/USER_MANUAL/06_fault_sources.tex
Revised the USER_MANUAL and added an mesh example (scripts for CUBIT meshing)

@@
 \&\textbf{RUPTURE\_SWITCHES}  /
-RATE\_AND\_STATE ,TPV16 ,HETE\_RSF ,TPV10X
+RATE\_AND\_STATE ,TPV16 ,HETE\_RSF ,TPV10X ,TWF
 
@@
 \&\textbf{SWF} mus, mud, dc {[}, nmus, nmud, ndc{]} /
+
+\&\textbf{TWF} nuc\_x, nuc\_y, nuc\_z, nuc\_r, nuc\_t0, nuc\_v /
 
@@
    =.FALSE. turn off such feature
+    \item[{TWF}]
+         =.TRUE. turn on time-weakening nucleation. This function is activated only when                    
+         RATE\_AND\_STATE=.FALSE. and TWF=.TRUE.
    \end{description}
@@
 \item [{nC}] = number of heterogeneous items for cohesion
+\end{description}
+
+\item [{\&\textbf{TWF}}] input block sets the time-weakening friction parameters
+of the fault:
+\begin{description}
+\item [{nuc\_x}]  = x coordinate of the hypocenter (in m)
+\item [{nuc\_y}]  = y coordinate of the hypocenter (in m)
+\item [{nuc\_z}]  = z coordinate of the hypocenter (in m)
+\item [{nuc\_r}]  = the radius of time-weakening nucleation (in m) 
+\item [{nuc\_t0}] = the cohesive time. The friction linearly decreases from static friction to dynamic friction during this time (in s)
+\item [{nuc\_v}]  = the time-weakening nucleation speed (in m/s)
 \end{description}

### 2020-07-06 doc/USER_MANUAL/04_creating_databases.tex
fixed a bug in decompose_mesh_MPI introduced when I added support for QKappa; removed an old sentence in the users manual that said that the internal mesher did not support PMLs (it now does)
corrects typo, default name is tomograpy_model.xyz for tomo models

@@
 \item [{\texttt{tomo}}] For a user-specified 3D isotropic model which uses
-a tomographic model file \texttt{tomographic\_model.xyz} in directory
+a tomographic model file \texttt{tomography\_model.xyz} in directory
 \texttt{DATA}. See Section \ref{sec:Using-tomographic}, for a discussion
@@
 \item [{\texttt{PML\_CONDITIONS}}] Set to \texttt{.true.} to turn on C-PML
-boundary conditions for a regional simulation. Both fluids and elastic
-solids are supported. Note that xmeshfem3d generated meshes do not support C-PML yet.
+boundary conditions for a regional simulation. Both fluids and elastic solids are supported.
 \item [{\texttt{PML\_INSTEAD\_OF\_FREE\_SURFACE}}] Set to \texttt{.true.}

### 2020-07-06 doc/USER_MANUAL/04_creating_databases.tex
updates doc

@@
 Note that Qmu is always equal to Qs, but Qkappa is in general not equal to Qp.
-To convert one to the other see doc/note\_on\_Qkappa\_versus\_Qp.pdf and utils/attenuation/conversion\_from\_Qkappa\_Qmu\_to\_Qp\_Qs\_from\_Dahlen\_Tromp\_959\_960.f90.
+To convert one to the other see \texttt{doc/note\_on\_Qkappa\_versus\_Qp.pdf}
+and in folder \texttt{utils/attenuation}, the tool
+\texttt{conversion\_from\_Qkappa\_Qmu\_to\_Qp\_Qs\_from\_Dahlen\_Tromp\_959\_960.f90}.
 
@@
 Note that Qmu is always equal to Qs, but Qkappa is in general not equal to Qp.
-To convert one to the other see doc/note\_on\_Qkappa\_versus\_Qp.pdf and utils/attenuation/conversion\_from\_Qkappa\_Qmu\_to\_Qp\_Qs\_from\_Dahlen\_Tromp\_959\_960.f90.
+To convert one to the other see \texttt{doc/note\_on\_Qkappa\_versus\_Qp.pdf}
+and in folder \texttt{utils/attenuation}, the tool
+\texttt{conversion\_from\_Qkappa\_Qmu\_to\_Qp\_Qs\_from\_Dahlen\_Tromp\_959\_960.f90}.
+
 \item [{\texttt{ANISOTROPY}}] Set to \texttt{.true.} if you want to use

### 2020-07-12 doc/USER_MANUAL/14_post_processing.tex
Buildbot ran the code beautifier.
master/slave -> main/secondary after new linux language

@@
 written out in binary format.
-\item [{\texttt{WRITE\_SEISMOGRAMS\_BY\_MASTER}}] Set to \texttt{.true.}
-to have only the master process writing out seismograms. This can
-be useful on a cluster, where only the master process node has access
+\item [{\texttt{WRITE\_SEISMOGRAMS\_BY\_MAIN}}] Set to \texttt{.true.}
+to have only the main process writing out seismograms. This can
+be useful on a cluster, where only the main process node has access
 to the output directory.

### 2020-07-12 src/specfem3D/fault_solver_dynamic.f90

@@ -1499,7 +1499,7 @@ contains
 
     bc%dataXZ_all%npoin = npoin_all
 
-    ! only master ranks needs to allocate gathering arrays **_all
+    ! only main ranks needs to allocate gathering arrays **_all
     if (myrank == 0) then
       allocate(bc%dataXZ_all%xcoord(npoin_all),stat=ier)
       if (ier /= 0) call exit_MPI_without_rank('error allocating array 1404')

### 2020-07-12 src/specfem3D/write_seismograms.f90

@@ -92,7 +92,7 @@
       write_time_begin = wtime()
 
       ! checks if anything to do
-      if (nrec_local > 0 .or. (WRITE_SEISMOGRAMS_BY_MASTER .and. myrank == 0) .or. ASDF_FORMAT) then
+      if (nrec_local > 0 .or. (WRITE_SEISMOGRAMS_BY_MAIN .and. myrank == 0) .or. ASDF_FORMAT) then
 
         ! writes out seismogram files
         select case(SIMULATION_TYPE)
@@ -162,8 +162,8 @@
         ! output
         write(IMAIN,*)
         write(IMAIN,*) 'Total number of time steps written: ', it-it_begin+1
-        if (WRITE_SEISMOGRAMS_BY_MASTER) then
-          write(IMAIN,*) 'Writing the seismograms by master proc alone took ',sngl(write_time),' seconds'
+        if (WRITE_SEISMOGRAMS_BY_MAIN) then
+          write(IMAIN,*) 'Writing the seismograms by main proc alone took ',sngl(write_time),' seconds'
         else
           write(IMAIN,*) 'Writing the seismograms in parallel took ',sngl(write_time),' seconds'
         endif
@@ -194,7 +194,7 @@
           nrec,nrec_local,islice_selected_rec, &
           seismo_offset,seismo_current, &
           NSTEP,NTSTEP_BETWEEN_OUTPUT_SEISMOS,ASDF_FORMAT, &
-          WRITE_SEISMOGRAMS_BY_MASTER,SAVE_ALL_SEISMOS_IN_ONE_FILE,USE_BINARY_FOR_SEISMOGRAMS
+          WRITE_SEISMOGRAMS_BY_MAIN,SAVE_ALL_SEISMOS_IN_ONE_FILE,USE_BINARY_FOR_SEISMOGRAMS
 
   implicit none
 
@@ -206,7 +206,7 @@
 
   character(len=1) component
 
-  ! parameters for master collects seismograms
+  ! parameters for main collects seismograms
   real(kind=CUSTOM_REAL), dimension(:,:), allocatable :: one_seismogram
   integer :: nrec_local_received,total_seismos,receiver,sender
   integer :: iproc,ier,i
@@ -233,7 +233,7 @@
   if (ier /= 0) stop 'error while allocating one temporary seismogram'
 
   ! write out seismograms: all processes write their local seismograms themselves
-  if (.not. WRITE_SEISMOGRAMS_BY_MASTER) then
[…]

### 2020-10-23 doc/USER_MANUAL/A_reference_frame.tex

@@
 \begin{description}
-\item [{elastic domain:}] ~\\
-
+\item [{elastic domain:}] ~
 \begin{itemize}
@@
 at the surface.)
-
 \begin{itemize}

### 2020-10-24 doc/USER_MANUAL/04_creating_databases.tex

@@
 TODO / future work to do: currently the \texttt{BROADCAST\_SAME\_MESH\_AND\_MODEL}
-option assumes to have the (master) mesh files in \texttt{run0001/DATABASES\_MPI} or
+option assumes to have the (main) mesh files in \texttt{run0001/DATABASES\_MPI} or
 \texttt{run0001/OUTPUT\_FILES/DATABASES\_MPI}.

### 2020-10-24 doc/USER_MANUAL/09_noise_simulations.tex
adds version info; removes white space

@@
 except that the order of receivers listed in \texttt{STATIONS} is
-now important. The order will be used to determine the `master' receiver,
+now important. The order will be used to determine the `main' receiver,
 i.e., the one that simultaneously cross correlates with the others.\\
@@
 
-\item Create a file called \texttt{./NOISE\_TOMOGRAPHY/irec\_master\_noise}.
+\item Create a file called \texttt{./NOISE\_TOMOGRAPHY/irec\_main\_noise}.
 Note that this file is located in directory \texttt{./NOISE\_TOMOGRAPHY/}
 as well. In general, all noise simulation related parameter files
-go into that directory. \texttt{irec\_master\_noise} contains only
-one integer, which is the ID of the `master' receiver. For example,
+go into that directory. \texttt{irec\_main\_noise} contains only
+one integer, which is the ID of the `main' receiver. For example,
 if this file contains 5, it means that the fifth receiver listed in
-\texttt{DATA/STATIONS} becomes the `master'. That's why we mentioned
+\texttt{DATA/STATIONS} becomes the `main'. That's why we mentioned
 previously that the order of receivers in \texttt{DATA/STATIONS} is
@@
 contain receivers which are outside of our computational domains.
-Therefore, the integer in \texttt{irec\_master\_noise} is actually
+Therefore, the integer in \texttt{irec\_main\_noise} is actually
 the ID in \texttt{DATA/STATIONS\_FILTERED} (which is generated by
@@
 
-\item Create a file called \texttt{./NOISE\_TOMOGRAPHY/nu\_master}.
+\item Create a file called \texttt{./NOISE\_TOMOGRAPHY/nu\_main}.
 This file holds three numbers, forming a (unit) vector. It describes
-which component we are cross-correlating at the `master' receiver,
+which component we are cross-correlating at the `main' receiver,
 i.e., ${\hat{{\bf \nu}}}^{\alpha}$ in \citet{trompetal2010}. The
@@
 after Step 2, or the Step 3 will overwrite them. Note that we have
-a `master' receiver specified by \texttt{irec\_master\_noise}, the
+a `main' receiver specified by \texttt{irec\_main\_noise}, the
 seismogram at one station corresponds to the cross correlation between
-that station and the `master'. Since the seismograms have three components,
+that station and the `main'. Since the seismograms have three components,
 we may obtain cross correlations for different components as well,

### 2021-04-14 doc/USER_MANUAL/02_getting_started.tex
updates manual

@@
 
-When compiling for GPU cards i.e. using CUDA, you can use:
-
-\texttt{./configure --with-cuda}
-
-for CUDA version 4, or
-
-\texttt{./configure --with-cuda=cuda5}
-
-for CUDA version 5, or
-
-\texttt{./configure --with-cuda=cuda6}
-
-for CUDA version 6, or
-
-\texttt{./configure --with-cuda=cuda8}
-
-for CUDA version 8.
-
+\noindent
+When compiling for GPU cards, you can use CUDA with:
+\begin{verbatim}
+  ./configure --with-cuda ..
+\end{verbatim}
+or
+\begin{verbatim}
+  ./configure --with-cuda=cuda4 ..
+\end{verbatim}
+where for example \texttt{cuda4,cuda5,cuda6,cuda7,..} specifies the target GPU architecture of your card,
+(e.g., with CUDA 4 this refers to Tesla K10 cards), rather than the installed version of the CUDA toolkit.
 Before CUDA version 5, one version supported basically one new architecture and needed a different kind of compilation.
@@
 However at the moment, we still have one version linked to one specific architecture:
+\begin{verbatim}
+- CUDA 4 for Tesla,   cards like K10, Geforce GTX 650, ..
+- CUDA 5 for Kepler,  like K20
+- CUDA 6 for Kepler,  like K80
+- CUDA 7 for Maxwell, like Quadro K2200
+- CUDA 8 for Pascal,  like P100
+- CUDA 9 for Volta,   like V100
+- CUDA 10 for Turing, like GeForce RTX 2080
+- CUDA 11 for Ampere, like A100
+\end{verbatim}
+So even if you have the new CUDA toolkit version 11, but you want to run on say a K20 GPU, then you would still configure with:
+\begin{verbatim}
+  ./configure --with-cuda=cuda5
+\end{verbatim}
+The compilation with the cuda5 setting chooses then the right architecture (\texttt{-gencode=arch=compute\_35,code=sm\_35} for K20 cards).
 
-- CUDA 4 for Fermi,
-- CUDA 5 for Tesla K20
-- CUDA 6 for Tesla K80
-- CUDA 8 for Pascal P100.
-
-So even if you have the new CUDA toolkit version 8, but you want to run on say a K20 GPU, then you would still configure with:
-
-\texttt{./configure --with-cuda=cuda5}
-
-The compilation should work and the cuda5 setting chooses the right architecture (-gencode=arch=compute\_35,code=sm\_35 for K20 cards).
-
-So, type
-
-\texttt{./configure --with-cuda=cuda8}
-
-for Pascal P100 boards.
+The same applies to compilation for AMD cards with HIP:
+\begin{verbatim}
+  ./configure --with-hip ..
+\end{verbatim}
+or
+\begin{verbatim}
+  ./configure --with-hip=MI8 ..
+\end{verbatim}
+where for example \texttt{MI8,MI25,MI50,MI100,..} specifies the target GPU architecture of your card.\\
 
@@
 the Zlib compression development library is probably missing on your machine and you will need to install it or ask your system administrator to
-do so. On Linux machines the package is often called "zlib1g-dev" or similar. (thus "sudo apt-get install zlib1g-dev" would install it)
+do so. On Linux machines the package is often called "zlib1g-dev" or similar. (thus "sudo apt-get install zlib1g-dev" would install it)\\
 
@@
 with the \texttt{-{}-without-mpi} option to suppress all calls to
-MPI.
+MPI.\\
 

### 2021-07-05 doc/USER_MANUAL/02_getting_started.tex

@@
 %
-\noindent
-\red{Note: for people who would like to run the package on Windows rather than on Unix machines, you can install Docker or VirtualBox (installing a Linux in VirtualBox in that latter case) and run it easily from inside that.}\\
-
-\noindent
-We recommend that you add {\texttt{ulimit -S -s unlimited}} to your
-{\texttt{.bash\_profile}} file and/or {\texttt{limit stacksize
-unlimited }} to your {\texttt{.cshrc}} file to suppress any potential
-limit to the size of the Unix stack.\\
 
@@
 You can replace the GNU compilers above (gfortran and gcc) with other compilers if you want to; for instance for Intel ifort and icc use FC=ifort CC=icc instead.
-
 Note that MPI must be installed with MPI-IO enabled because parts of SPECFEM3D perform I/Os through MPI-IO.
@@
 \item [{\texttt{Intel ifort compiler}}] See if you need to add \texttt{-assume byterecl} for your machine.
-\textbf{In the case of that compiler, we have noticed that initial release versions sometimes have bugs or issues that can lead to wrong results when running the code, thus we \emph{strongly} recommend using a version for which at least one service pack or update has been installed.}
-\red{In particular, for version 17 of that compiler, users have reported problems (making the code crash at run time) with the \texttt{-assume buffered\_io} option; if you notice problems,
-remove that option from file \texttt{flags.guess} or change it to \texttt{-assume nobuffered\_io} and try again.}
+In the case of that compiler, we have noticed that initial release versions sometimes have bugs or issues that can lead to wrong results when running the code, thus we \emph{strongly} recommend using a version for which at least one service pack or update has been installed.
+In particular, for version 17 of that compiler, users have reported problems (making the code crash at run time) with the \texttt{-assume buffered\_io} option; if you notice problems,
+remove that option from file \texttt{flags.guess} or change it to \texttt{-assume nobuffered\_io} and try again.
+
 \item [{\texttt{IBM compiler}}] See if you need to add \texttt{-qsave}
 or \texttt{-qnosave} for your machine.
+
 \item [{\texttt{Mac OS}}] You will probably need to install \texttt{XCODE}.
 \end{description}
+
 When compiling on an IBM machine with the \texttt{xlf} and \texttt{xlc}
@@
 \texttt{MPI\_INC=\${CRAY\_MPICH2\_DIR}/include and FCLIBS=" "}, and for more details if needed you can refer to the \texttt{utils/Cray\_compiler\_information} directory.
-You can also have a look at the configure script called \texttt{utils/Cray\_compiler\_information/configure\_SPECFEM\_for\_Piz\_Daint.bash}.
+You can also have a look at the configure script called:\\
+\texttt{utils/Cray\_compiler\_information/configure\_SPECFEM\_for\_Piz\_Daint.bash}.
 
@@
 
-You can add \texttt{--enable-vectorization} to the configuration options to speed up the code in the fluid (acoustic) and elastic parts.
+You can add \texttt{-{}-enable-vectorization} to the configuration options to speed up the code in the fluid (acoustic) and elastic parts.
 This works fine if (and only if) your computer always allocates a contiguous memory block for each allocatable array;
-this is the case for most machines and most compilers, but not all. To disable this feature, use option \texttt{--disable-vectorization}.
+this is the case for most machines and most compilers, but not all. To disable this feature, use option \texttt{-{}-disable-vectorization}.
 For more details see \href{https://github.com/geodynamics/specfem3d/issues/81}{github.com/geodynamics/specfem3d/issues/81} .
@@
 that the spectral-element method can very accurately handle distorted
-mesh elements \citep{OlSe11}.
+mesh elements \citep{OlSe11}.\\
 
@@
 
+When compiling the SCOTCH source code, if you get a message such as: "ld: cannot find -lz",
+the Zlib compression development library is probably missing on your machine and you will need to install it or ask your system administrator to
+do so. On Linux machines the package is often called "zlib1g-dev" or similar. (thus "sudo apt-get install zlib1g-dev" would install it)\\
+
+To compile a serial version of the code for small meshes that fits
+on one compute node and can therefore be run serially, run \texttt{configure}
+with the \texttt{-{}-without-mpi} option to suppress all calls to
+MPI.\\
+
+For people who would like to run the package on Windows rather than on Unix machines, you can install Docker or VirtualBox (installing a Linux in VirtualBox in that latter case) and run it easily from inside that.\\
+
+We recommend that you add {\texttt{ulimit -S -s unlimited}} to your
+{\texttt{.bash\_profile}} file and/or {\texttt{limit stacksize
+unlimited }} to your {\texttt{.cshrc}} file to suppress any potential
+limit to the size of the Unix stack.\\
+
+%-----------------------------------------------------------------------------------------------------------------------------------%
+\section{Using the GPU version of the code}
+%-----------------------------------------------------------------------------------------------------------------------------------%
+
 \noindent
-When compiling for GPU cards, you can use CUDA with:
+SPECFEM3D now supports CUDA and HIP GPU acceleration.
+When compiling for GPU cards, you can enable the CUDA version with:
 \begin{verbatim}
@@
 \begin{verbatim}
-  ./configure --with-cuda=cuda4 ..
+  ./configure --with-cuda=cuda9 ..
 \end{verbatim}
 where for example \texttt{cuda4,cuda5,cuda6,cuda7,..} specifies the target GPU architecture of your card,
-(e.g., with CUDA 4 this refers to Tesla K10 cards), rather than the installed version of the CUDA toolkit.
+(e.g., with CUDA 9 this refers to Volta V100 cards), rather than the installed version of the CUDA toolkit.
 Before CUDA version 5, one version supported basically one new architecture and needed a different kind of compilation.
@@
 \end{verbatim}
-The compilation with the cuda5 setting chooses then the right architecture (\texttt{-gencode=arch=compute\_35,code=sm\_35} for K20 cards).
+The compilation with the cuda5 setting chooses then the right architecture (\texttt{-gencode=arch=compute\_35,code=sm\_35} for K20 cards).\\
 
@@
 
-When compiling the SCOTCH source code, if you get a message such as: "ld: cannot find -lz",
-the Zlib compression development library is probably missing on your machine and you will need to install it or ask your system administrator to
-do so. On Linux machines the package is often called "zlib1g-dev" or similar. (thus "sudo apt-get install zlib1g-dev" would install it)\\
+%-----------------------------------------------------------------------------------------------------------------------------------%
+\section{Using the ADIOS library for I/O}
+%-----------------------------------------------------------------------------------------------------------------------------------%
 
-To compile a serial version of the code for small meshes that fits
-on one compute node and can therefore be run serially, run \texttt{configure}
-with the \texttt{-{}-without-mpi} option to suppress all calls to
-MPI.\\
+Regular POSIX I/O can be problematic when dealing with large simulations on large
+clusters (typically more than $10,000$ MPI processes). SPECFEM3D can use the ADIOS library~\cite{Liu2013}
+to take advantage of advanced parallel file system features. To enable
+ADIOS, the following steps should be done:
+\begin{enumerate}
+\item Install ADIOS (available from \url{https://www.olcf.ornl.gov/center-projects/adios/}).
+Make sure that your environment variables reference it.
+\item You may want to change ADIOS related values in the \texttt{constants.h.in} file.
+The default values probably suit most cases.
+\item Configure using the \texttt{-{}-with-adios} flag.
+\end{enumerate}
+ADIOS is currently only usable for meshfem3D generated mesh (i.e. not for meshes generated
+with CUBIT). Additional control parameters are discussed in section~\ref{cha:Main-Parameter}.
 
+%-----------------------------------------------------------------------------------------------------------------------------------%
+\section{Adding OpenMP support in addition to MPI}
+%-----------------------------------------------------------------------------------------------------------------------------------%
+
+OpenMP support can be enabled in addition to MPI. However, in many
+cases performance will not improve because our pure MPI implementation
+is already heavily optimized and thus the resulting code will in fact
+be slightly slower. A possible exception could be IBM BlueGene-type
+architectures.\\
+
+\noindent
+To enable OpenMP,  add the flag \texttt{-{}-enable-openmp} to the configuration:
+\begin{verbatim}
+./configure --enable-openmp ..
+\end{verbatim}
+This will add the corresponding OpenMP flag for the chosen Fortran compiler.\\
+
+The DO-loop using OpenMP threads has a SCHEDULE property. The \texttt{OMP\_SCHEDULE}
+environment variable can set the scheduling policy of that DO-loop.
+Tests performed by Marcin Zielinski at SARA (The Netherlands) showed
+that often the best scheduling policy is DYNAMIC with the size of
+the chunk equal to the number of OpenMP threads, but most preferably
+being twice as the number of OpenMP threads (thus chunk size = 8 for
+4 OpenMP threads etc). If \texttt{OMP\_SCHEDULE} is not set or is empty, the
+DO-loop will assume generic scheduling policy, which will slow down
+the job quite a bit.
+
+%-----------------------------------------------------------------------------------------------------------------------------------%
+\section{Configuration summary}
+%-----------------------------------------------------------------------------------------------------------------------------------%
+
+\noindent
 A summary of the most important configuration variables follows.
@@
 \item [{\texttt{F90}}] Path to the Fortran compiler.
+
 \item [{\texttt{MPIF90}}] Path to MPI Fortran.
+
 \item [{\texttt{MPI\_FLAGS}}] Some systems require this flag to link to
 MPI libraries.
+
 \item [{\texttt{FLAGS\_CHECK}}] Compiler flags.
 \end{description}
+%
 The configuration script automatically creates for each executable
@@
 difference in terms of performance. We welcome feedback on your experience
-with various compilers and flags.
+with various compilers and flags.\\
 
@@
 \end{description}
+%
 The package can run either in single or in double precision mode.
@@
 \end{description}
+%
 In the \texttt{precision.h} file:
@@
 \end{description}
+%
 On many current processors (e.g., Intel, AMD, IBM Power), single precision
@@
 
-\section{Adding OpenMP support in addition to MPI}
+%-----------------------------------------------------------------------------------------------------------------------------------%
+\section{Compiling on an IBM BlueGene}
+%-----------------------------------------------------------------------------------------------------------------------------------%
 
-NOTE FROM JULY 2013: OpenMP support is maybe / probably not maintained any more.
-Thus the section below is maybe obsolete.
-
-OpenMP support can be enabled in addition to MPI. However, in many
-cases performance will not improve because our pure MPI implementation
-is already heavily optimized and thus the resulting code will in fact
-be slightly slower. A possible exception could be IBM BlueGene-type
-architectures.
-
-To enable OpenMP, uncomment the OpenMP compiler option in two lines
-in file \texttt{src/specfem3D/Makefile.in} (before running \texttt{configure})
-and also uncomment the
-\texttt{\#define USE\_OPENMP} statement in file\\
-\texttt{src/specfem3D/specfem3D.F90}.
-
-The DO-loop using OpenMP threads has a SCHEDULE property. The \texttt{OMP\_SCHEDULE}
-environment variable can set the scheduling policy of that DO-loop.
-Tests performed by Marcin Zielinski at SARA (The Netherlands) showed
-that often the best scheduling policy is DYNAMIC with the size of
-the chunk equal to the number of OpenMP threads, but most preferably
-being twice as the number of OpenMP threads (thus chunk size = 8 for
-4 OpenMP threads etc). If \texttt{OMP\_SCHEDULE} is not set or is empty, the
-DO-loop will assume generic scheduling policy, which will slow down
-the job quite a bit.
-
-\section{Compiling on an IBM BlueGene}
-
-\underline{More recent installation instruction for IBM BlueGene, from April 2013:}\\
+\noindent
+Installation instructions for IBM BlueGene (from April 2013):\\
 
@@
 
+%-----------------------------------------------------------------------------------------------------------------------------------%
 \section{Visualizing the subroutine calling tree of the source code}
+%-----------------------------------------------------------------------------------------------------------------------------------%
 
@@
 a GUI front-end for configuring and running \texttt{Doxygen}.
+
+To visualize the call tree (calling tree) of the source code, you can see the Doxygen tool available in directory \texttt{doc/call\_trees\_of\_the\_source\_code}.
 
@@
 
-\section{Using the ADIOS library for I/O}
-
-Regular POSIX I/O can be problematic when dealing with large simulations one large
-clusters (typically more than $10,000$ processes). SPECFEM3D use the ADIOS library~\cite{Liu2013}
-to deal transparently take advantage of advanced parallel file system features. To enable
-ADIOS, the following steps should be done:
-\begin{enumerate}
-\item Install ADIOS (available from \url{https://www.olcf.ornl.gov/center-projects/adios/}).
-Make sure that your environment variables reference it.
-\item You may want to change ADIOS related values in the \texttt{constants.h.in} file.
-The default values probably suit most cases.
-\item Configure using the \texttt{--with-adios} flag.
-\end{enumerate}
-ADIOS is currently only usable for meshfem3D generated mesh (i.e. not for meshes generated
-with CUBTI). Additional control parameters are discussed in section~\ref{cha:Main-Parameter}.
-
+%-----------------------------------------------------------------------------------------------------------------------------------%
 \section{Becoming a developer of the code, or making small modifications in the source code}
+%-----------------------------------------------------------------------------------------------------------------------------------%
 
@@
 To do so, i.e.
-to access the development branch of the source code with read/write access (in a safe way, no need to worry too much about breaking the package, there is a robot called BuildBot that is in charge of checking and validating all new contributions and changes), please visit this Web page: \url{https://github.com/geodynamics/specfem3d/wiki/Using-Hub}.
-
-To visualize the call tree (calling tree) of the source code, you can see the Doxygen tool available in directory \texttt{doc/call\_trees\_of\_the\_source\_code}.
+to access the development branch of the source code with read/write access (in a safe way, no need to worry too much about breaking the package, there are CI tests based on BuildBot, Travis-CI and Jenkins in place that are checking and validating all new contributions and changes), please visit this Web page:\\
+\url{https://github.com/geodynamics/specfem3d/wiki}

### 2021-07-05 doc/USER_MANUAL/08b_performing_full_waveform_inversion_FWI_or_source_inversions.tex

@@
 
-\red{\Large ANNOUNCEMENT: SPECFEM3D can now perform full waveform inversion (FWI), i.e. invert for models in an iterative fashion, and it can also perform source inversions in a constant model;
+SPECFEM3D can now perform full waveform inversion (FWI), i.e. invert for models in an iterative fashion, and it can also perform source inversions in a constant model;
 please refer to the two new directories, \texttt{inverse\_problem\_for\_model} and \texttt{inverse\_problem\_for\_source}, and the README files they contain.
 For FWI inversions for the model, also refer to the new examples provided in
-directory \texttt{EXAMPLES}.}
+directory \texttt{EXAMPLES}.

### 2021-10-13 doc/USER_MANUAL/05_running_the_solver.tex
updates doc about STF

@@
 or the centroid time of the earthquake. The start time of the simulation
-is $t=-1.5*\texttt{half duration}$ (the 1.5 is to make sure the moment
-rate function is very close to zero when starting the simulation).
+is $t=-2.0*\texttt{half duration}$ (the 2.0 factor is to make sure the moment
+rate function is very close to zero when starting the simulation. This avoids spurious high-frequency oscillations.
+Also, for acoustic simulations using a Gaussian STF this is set to a factor 3.0, while Ricker source time functions use a factor 1.2).
 To convert to absolute time $t_{\mathrm{abs}}$, set

### 2021-10-20 doc/USER_MANUAL/02_getting_started.tex
Update manual about cluster version issue
Update manual about dynamic fault config parameters

@@
 limit to the size of the Unix stack.\\
+
+Beware that some cluster systems that run a recent version many not run and\slash compile an older version of the code. \\
+
+When using dynamic fault in parallel with the developer version, we suggest you to set the configuration parameters 
+of {\texttt{FAULT_DISPL_VELOC}} and {\texttt{FAULT_SYNCHRONIZE_ACCEL}} as {\texttt{.true.}}.\\
 

### 2021-11-16 doc/USER_MANUAL/06_fault_sources.tex
Buildbot ran the code beautifier.
Edits of User_Manual for adding exponential slip-weakening friction law

@@
 should be at least $10^{-6}y_0$ since single precision can ensure precision
-within only 7 digits). Note that only the interior
+within only 7 digits; furthermore, to properly recognize split nodes, the dimension
+of the mesh also affect numerical roundoffs, and therefore the split nodes
+should also be at least \texttt{10.d-10 * (Xmax-Xmin)} apart). Note that only the interior
 of the fault must be opened, its edges must remain closed (except
@@
 
-\&\textbf{SWF} mus, mud, dc {[}, nmus, nmud, ndc{]} /
+\&\textbf{SWF} mus, mud, dc {[}, nmus, nmud, ndc{]} (weakening_kind=1 for linear (default); weakening_kind=2 for exponential)/
 
@@
 \item [{nC}] = number of heterogeneous items for cohesion
+\item [{weakening_kind}] = 1 for linear slip-weakening law (default); 2 for exponential slip-weakening law: $\mu_d + (\mu_s-\mu_d) \exp(-\frac{u}{d_c})$
 \end{description}

### 2021-11-16 doc/USER_MANUAL/06_fault_sources.tex
Edits of User_Manual for adding exponential slip-weakening friction law

@@
 \item [{nC}] = number of heterogeneous items for cohesion
-\item [{weakening_kind}] = 1 for linear slip-weakening law (default); 2 for exponential slip-weakening law: $\mu_d + (\mu_s-\mu_d) \exp(-\frac{u}{d_c})$
+\item [{weakening_kind}] = 1 for linear slip-weakening law (default); 2 for exponential slip-weakening law: $$\mu_d + (\mu_s-\mu_d) \exp(-\frac{u}{d_c})$$
 \end{description}

### 2021-11-16 doc/USER_MANUAL/06_fault_sources.tex

--- 2021-11-16 doc/USER_MANUAL/06_fault_sources.tex
@@
 \item [{nC}] = number of heterogeneous items for cohesion
-\item [{weakening_kind}] = 1 for linear slip-weakening law (default); 2 for exponential slip-weakening law: $\mu_d + (\mu_s-\mu_d) \exp(-\frac{u}{d_c})$
+\item [{weakening_kind}] = 1 for linear slip-weakening law (default); 2 for exponential slip-weakening law:
+                            $$\mu = \mu_d + (\mu_s-\mu_d) exp(-\frac{u}{d_c}) $$
 \end{description}

### 2021-11-17 doc/USER_MANUAL/03_mesh_generation.tex
renames input parameter subamp_seismos to NTSTEP_BETWEEN_OUTPUT_SAMPLE (making naming more uniform across different SPECFEM codes)

@@
 
+\noindent
 When you are ready to compile, in the main directory type
@@
 ready to compile the internal mesher. This is an alternative to CUBIT
-for the mesh generation of relatively simple models.
-%
+for the mesh generation of relatively simple models.\\
+
 \begin{figure}[htbp]
@@
 note that \texttt{xmeshfem3D} must be called directly from the main
-directory, as most of the binaries of the package.
+directory, as most of the binaries of the package.\\
 
@@
 file \texttt{Mesh\_Par\_file}, which resides in the subdirectory \texttt{DATA/meshfem3D\_files/}.
-(to see how to use it, see the EXAMPLES specific to the internal mesher in directory \texttt{EXAMPLES/meshfem3D\_examples/}.
+(To see how to use it, see the EXAMPLES specific to the internal mesher in directory \texttt{EXAMPLES/meshfem3D\_examples/}.)
 Before running the mesher, a number of parameters need to be set in
@@
 of how the SEM is implemented, and we encourage you to read \citet{KoVi98,KoTr99}
-and \citet{KoLiTrSuStSh04}.
+and \citet{KoLiTrSuStSh04}.\\
 
@@
 in which case all `longitude' parameters simply refer to the $x$~axis, and
-all `latitude' parameters simply refer to the $y$~axis. To run the
-mesher for a global simulation, the following parameters need to be
+all `latitude' parameters simply refer to the $y$~axis.\\
+
+\noindent
+To run the
+internal mesher for a regional simulation, the following parameters need to be
 set in the \texttt{\small Mesh\_Par\_file}:
@@
 
-Finally, depending on your system, you might need to provide a file
-that tells MPI what compute nodes to use for the simulations. The
+Now that you have set the appropriate parameters in the \texttt{Mesh\_Par\_file}
+and have compiled the mesher, you are ready to run it!
+\footnote{
+Depending on your system, please check how to run executables with MPI.
+To run the mesher on your cluster, we provide some example scripts for you to modify accordingly in \texttt{utils/Cluster/}.
+For some installations, you might need to provide a file
+that tells MPI what compute nodes to use for the simulations. In this case, the
 file must have a number of entries (one entry per line) at least equal
-to the number of processors needed for the run. A sample file is provided
-in the file \texttt{mymachines}. This file is not used by the mesher
-or solver, but is required by the \texttt{go\_mesher} and \texttt{go\_solver}
-default job submission scripts. See Chapter~\ref{cha:Scheduler}
-for information about running the code on a system with a scheduler,
-e.g., LSF.
-
-Now that you have set the appropriate parameters in the \texttt{Mesh\_Par\_file}
-and have compiled the mesher, you are ready to launch it! This is
-most easily accomplished based upon the \texttt{go\_mesher} script.
+to the number of processors needed for the run.
+A sample file name is \texttt{mymachines}. This file is not used by the mesher
+or solver, but is required by the \texttt{go\_mesher***} and \texttt{go\_solver**}
+default job submission scripts provided in directory \texttt{utils/Cluster/}.
+To run the mesher, it is most easily accomplished based upon the \texttt{go\_mesher***} script.
 When you run on a PC cluster, the script assumes that the nodes are
@@
 at the end of the script that invokes the \texttt{mpirun} command.
-See Chapter~\ref{cha:Scheduler} for information about running the
-code on a system with a scheduler, e.g., LSF.
+See Chapter~\ref{cha:Scheduler}
+for information about running the code on a system with a scheduler,
+e.g., LSF.
+}
+Mesher output is provided in the \texttt{OUTPUT\_FILES/} directory
+in \texttt{output\_mesher.txt}; this file provides lots of details
+about the mesh that was generated.\\
 
-Mesher output is provided in the \texttt{OUTPUT\_FILES} directory
-in \texttt{output\_mesher.txt}; this file provides lots of details
-about the mesh that was generated. Please note that the mesher suggests
+Please note that the mesher suggests
 a time step \texttt{DT} to run the solver with. The mesher output
@@
 in \texttt{constants.h}:
+{\small
 \begin{verbatim}
@@
 ! integer, parameter :: IMAIN = ISTANDARD_OUTPUT
-\end{verbatim}
+\end{verbatim}}
+
 To control the quality of the mesh, check the standard output (either
-on the screen or in the \texttt{OUTPUT\_FILES} directory in \texttt{output\_mesher.txt})
+on the screen or in the \texttt{OUTPUT\_FILES/} directory in \texttt{output\_mesher.txt})
 and analyze the skewness histogram of mesh elements to make sure no
@@
 above 0.80, then you must definitely improve the mesh). To draw the
-skewness histogram on the screen, type \texttt{gnuplot plot\_mesh\_quality\_histogram.gnu}.
+skewness histogram on the screen, you could type
+{\small \begin{verbatim}
+  gnuplot plot_mesh_quality_histogram.gnu
+\end{verbatim}}

### 2021-11-17 doc/USER_MANUAL/04_creating_databases.tex
renames input parameter subamp_seismos to NTSTEP_BETWEEN_OUTPUT_SAMPLE (making naming more uniform across different SPECFEM codes)

@@
 
+\newpage
 \section{Main parameter file \texttt{Par\_file}}\label{cha:Main-Parameter}
@@
 need to be set in the main parameter \texttt{Par\_file} located in
-the subdirectory \texttt{DATA}:
+the subdirectory \texttt{DATA}:\\
+
+\noindent {\bf $\bullet$ Simulation setup}
 \begin{description}
@@
 for further details.
+\end{description}
+
+\vspace{1cm}
+\noindent {\bf $\bullet$ Mesh setup}
+\begin{description}
 \item [{\texttt{NGNOD}}] The number of nodes for 2D and 3D shape functions
@@
 in Olsen's empirical relation (see \citet{OlDaBr03}).
+\end{description}
+
+\vspace{1cm}
+\noindent {\bf $\bullet$ Absorbing boundary conditions setup}
+\begin{description}
 \item [{\texttt{PML\_CONDITIONS}}] Set to \texttt{.true.} to turn on C-PML
@@
 make zmin free surface instead of absorbing condition.
+\end{description}
+
+\vspace{1cm}
+\noindent {\bf $\bullet$ Visualization setup}
+\begin{description}
 \item [{\texttt{CREATE\_SHAKEMAP}}] Set this flag to \texttt{.true.} to
@@
 the run.
-\item [{\texttt{NTSTEP\_BETWEEN\_OUTPUT\_SEISMOS}}] This parameter specifies
-the interval at which synthetic seismograms are written in the \texttt{LOCAL\_PATH}
-directory. If a run crashes, you may still find usable (but shorter
-than requested) seismograms in this directory. On a fast machine set
-\texttt{NTSTEP\_BETWEEN\_OUTPUT\_SEISMOS} to a relatively high value
-to avoid writing to the seismograms too often. This feature is only
-relevant for the solver.
+\end{description}
+
+\vspace{1cm}
+\noindent {\bf $\bullet$ Sources setup}
+\begin{description}
 \item [{\texttt{USE\_FORCE\_POINT\_SOURCE}}] Turn this flag on to use a
@@
 %
+\end{description}
+
+\vspace{1cm}
+\noindent {\bf $\bullet$ Additionals}
+\begin{description}
+\item [{\texttt{NTSTEP\_BETWEEN\_OUTPUT\_SEISMOS}}] This parameter specifies
+the interval at which synthetic seismograms are written in the \texttt{LOCAL\_PATH}
+directory. If a run crashes, you may still find usable (but shorter
+than requested) seismograms in this directory. On a fast machine set
+\texttt{NTSTEP\_BETWEEN\_OUTPUT\_SEISMOS} to a relatively high value
+to avoid writing to the seismograms too often. This feature is only
+relevant for the solver.
+
 \item [{\texttt{NUMBER\_OF\_SIMULTANEOUS\_RUNS}}] adds the ability to run several calculations (several earthquakes)
@@
 
-\item [{\texttt{USE\_FAILSAFE\_MECHANISM}}] if one or a few of these simultaneous runs fail,
-kill all the runs or let the others finish using a fail-safe mechanism
-(in most cases, should be set to true).
+%\item [{\texttt{USE\_FAILSAFE\_MECHANISM}}] if one or a few of these simultaneous runs fail,
+%kill all the runs or let the others finish using a fail-safe mechanism
+%(in most cases, should be set to true).
 
@@
 \item [\texttt{ADIOS\_FOR\_KERNELS}] Turn this flag on to produce ADIOS kernels that can later be visualized with the ADIOS version of combine\_vol\_data.
+\end{description}
+There are quite a few more parameters to change the default setup of your runs. Please check the comments in the \texttt{Par\_file} directly for further explanations.\\
 
-\end{description}
-
-The present version of SPECFEM can handle fully saturated porous simulations, Christina Morency implemented Biot equation.
-But the code cannot calculate partially saturated cases in its state.
-Christina Morency is presently working on a dual porosity, dual permeability formulation, type Pride and Berryman, for an other project, but it will not be available for some time.\\
-
-The way we prescribe material property for porous material in SPECFEM3D is as follow:
-We use a file name "nummaterial\_poroelastic\_file", which is located in the directory \texttt{MESH/}, the format is as follow:
-\begin{verbatim}
+The present version of SPECFEM can handle fully saturated porous simulations.
+(Christina Morency implemented Biot's equation.
+However, the code cannot calculate partially saturated cases in its state.
+Christina Morency is presently working on a dual porosity, dual permeability formulation, type Pride and Berryman, for an other project, but it will not be available for some time.)
+The way we prescribe material property for porous material in SPECFEM3D is as follow.
+We use a file \texttt{nummaterial\_poroelastic\_file}, which is located in the directory \texttt{MESH/}. The line format is as follow:
+{\small \begin{verbatim}
   rhos rhof phi c kxx kxy kxz kyy kyz kzz Ks Kf Kfr etaf mufr
-\end{verbatim}
-where\\
+\end{verbatim}}
+\begin{description}
+\item where\\
 \texttt{rho\_s} = solid density,\\
@@
 \texttt{mu\_fr} = frame shear modulus.\\
-
-Using an external mesh (for instance coming from CUBIT/TRELIS), poroelastic materials have the ID number 3, while 1 is acoustic and 2 is elastic (see the example in the package: EXAMPLES/homogeneous\_poroelastic).
+\end{description}
+Using an external mesh (for instance coming from CUBIT/TRELIS), poroelastic materials have the ID number 3, while 1 is acoustic and 2 is elastic (see the example in the package: EXAMPLES/homogeneous\_poroelastic).\\
 
@@
 velocity and density model along the normal direction.
-
 Thus, you need to modify your velocity and density model in order
 for it to be 1D inside the PML, as shown in Figure~\ref{fig:modify_external_velocity_model_to_use_PML}.
-
 This applies to the bottom layer as well; there you should make sure

### 2021-11-17 doc/USER_MANUAL/05_running_the_solver.tex
renames input parameter subamp_seismos to NTSTEP_BETWEEN_OUTPUT_SAMPLE (making naming more uniform across different SPECFEM codes)

@@
 to run:
-\begin{description}
-\item [{\texttt{Par\_file}}] the main parameter file which was discussed
+\begin{itemize}
+\item {\bf \texttt{Par\_file}} the main parameter file which was discussed
 in detail in the previous Chapter~\ref{cha:Creating-Distributed-Databases},
-\item [{\texttt{CMTSOLUTION}} or {\texttt{FORCESOLUTION}}] the earthquake
+\item {\bf \texttt{CMTSOLUTION}} or {\bf \texttt{FORCESOLUTION}} the earthquake
 source parameter file or the force source parameter file, and
-\item [{\texttt{STATIONS}}] the stations file.
-\end{description}
+\item {\bf \texttt{STATIONS}} the stations file.
+\end{itemize}
+
+\noindent
 Most parameters in the \texttt{Par\_file} should be set prior to running
@@
 \begin{itemize}
-\item the simulation type control parameters: \texttt{SIMULATION\_TYPE}
+\item the simulation type control parameters \texttt{SIMULATION\_TYPE}
 and \texttt{SAVE\_FORWARD}
@@
 the database generator \texttt{xgenerate\_databases} and the solver
-\texttt{xspecfem3D}.
+\texttt{xspecfem3D}.\\
 
@@
 
+\noindent
 Each line represents one station in the following format:
@@
 \texttt{LONGITUDE\_MAX}). The filtered file is called \texttt{DATA/STATIONS\_FILTERED}.
+Elevation and burial are generally applicable to geographical regions.
+Burial is measured down from the top surface.\\
 
-Elevation and burial are generally applicable to geographical regions.
-Burial is measured down from the top surface. For other problems in other
+\noindent
+For other problems in other
 fields (ultrasonic testing, medical imaging etc...),
-it may be confusing. We generally follow either of the following
+it may be confusing. We generally follow either one of the following
 procedures for those kind of problems:
+\begin{itemize}
+\item {\it Procedure 1:} mostly for geophysics, when the top surface is a free surface (topography) and the five other edges of the mesh are absorbing surfaces
+\begin{enumerate}
+\item [-] Put the origin on the top of the model.\\
+\item [-] Let's say you want to place two receivers at (x1,y1,z1) and
+(x2,y2,z2). Your STATIONS file should look like:
+{\small \begin{verbatim}
+BONE  GR  y1  x1  0.00  -z1
+BONE  GR  y2  x2  0.00  -z2
+\end{verbatim}}
+\end{enumerate}
 
-Procedure 1 (mostly for geophysics, when the top surface is a free surface (topography) and the five other edges of the mesh are absorbing surfaces):\\
-====================================================================================================================================================\\
-
-- Put the origin on the top of the model.\\
-
-- Let's say you want to place two receivers at (x1,y1,z1) and
-(x2,y2,z2). Your STATIONS file should look like:\\
-
-BONE  GR  y1  x1  0.00  -z1\\
-BONE  GR  y2  x2  0.00  -z2\\
-
-Procedure 2 (useful for other application domains, in which using the absolute $Z$ position of the sources and receivers is more standard than using their depth from the surface):\\
-==================================================================================================================================================================================
-
-- In principle in the international CMTSOLUTION format in geophysics
+\item {\it Procedure 2:} useful for other application domains, in which using the absolute $Z$ position of the sources and receivers is more standard than using their depth from the surface
+\begin{enumerate}
+\item [-] In principle in the international CMTSOLUTION format in geophysics
 the depth is given in kilometers; however for users in other fields (non-destructive testing, medical imaging, near-surface studies...) who may prefer to give the position of the source (rather than its depth from the surface), or for people who use FORCESOLUTION to describe the source rather than CMTSOLUTION, we provide an option called USE\_SOURCES\_RECEIVERS\_Z in the Par\_file, and if so that position is read from CMTSOLUTION in meters rather than kilometers (and again, it is then the true position in the mesh, not the depth).
 When option USE\_SOURCES\_RECEIVERS\_Z in the Par\_file is on, this remark applies to the position of the receivers as well.\\
+\item [-] Let's say you want to place two receivers at (x1,y1,z1) and
+(x2,y2,z2). Your STATIONS file should then look like:
+{\small \begin{verbatim}
+BONE  GR  y1  x1  0.00  z1
+BONE  GR  y2  x2  0.00  z2
+\end{verbatim}}
+The option USE\_SOURCES\_RECEIVERS\_Z set to .true. will then discard the elevation and set
+burial as the $z$ coordinate.
+Third column is Y and Fourth is X due to the latitude/longitude convention.\\
+\end{enumerate}
 
-- Let's say you want to place two receivers at (x1,y1,z1) and
-(x2,y2,z2). Your STATIONS file should then look like:\\
-\\
-BONE  GR  y1  x1  0.00  z1\\
-BONE  GR  y2  x2  0.00  z2\\
-\\
-The option USE\_SOURCES\_RECEIVERS\_Z set to .true. will then discard the elevation and set
-burial as the $z$ coordinate.\\
-\\
-Third column is Y and Fourth is X due to the latitude/longitude convention.\\
-\\
 You can replace the station name "BONE" with any word of length less
 than 32, and the network name "GR" with any word of length less than
-8.\\
-\\
-You can always plot OUTPUT\_FILES/sr.vtk file in ParaView to check the
+8.
+And you can always plot OUTPUT\_FILES/sr.vtk file in ParaView to check the
 source/receiver locations after your simulation.\\
+\end{itemize}
 

### 2021-11-19 doc/USER_MANUAL/02_getting_started.tex
minor edit
Buildbot ran the code beautifier.
updates doc

--- 2021-07-05 doc/USER_MANUAL/02_getting_started.tex
@@
 limit to the size of the Unix stack.\\
+
+Beware that some cluster systems that run a recent version may not run and\slash compile an older version of the code. \\
+
+When using dynamic fault in parallel with the developer version, we suggest you to set the configuration parameters
+of \texttt{FAULT\_DISPL\_VELOC} and \texttt{FAULT\_SYNCHRONIZE\_ACCEL} as \texttt{.true.}.\\
 

### 2021-11-19 doc/USER_MANUAL/06_fault_sources.tex
updates doc

--- 2021-11-16 doc/USER_MANUAL/06_fault_sources.tex
@@
 \item [{Example:}] For the example in Section \ref{sec:Mesh-Generation-with}:
+{\footnotesize
 \begin{verbatim}
@@
   fault_input(1,A1,A2)
-\end{verbatim}
+\end{verbatim}}
 \end{description}
@@
 in a Fortran 9x manual). The input for each fault has the following
-sequence (arguments in {[}brackets{]} are optional):
+sequence (arguments in {[}brackets{]} are optional):\\
 
-\&\textbf{RUPTURE\_SWITCHES}  /
-RATE\_AND\_STATE ,TPV16 ,HETE\_RSF ,TPV10X ,TWF
+\&\textbf{RUPTURE\_SWITCHES}  / RATE\_AND\_STATE ,TPV16 ,HETE\_RSF ,TPV10X ,TWF\\
+\&\textbf{BEGIN\_FAULT} /\\
+\&\textbf{STRESS\_TENSOR} Sigma = $\sigma_{xx}$,$\sigma_{yy}$,$\sigma_{zz}$,$\sigma_{xy}$,$\sigma_{xz}$,$\sigma_{yz}$ /\\
+\&\textbf{INIT\_STRESS} S1, S2, S3 {[},n1, n2, n3{]} /\\
 
-\&\textbf{BEGIN\_FAULT} /
+followed by (n1+n2+n3) \&\textbf{DIST2D} blocks\\
 
-\&\textbf{STRESS\_TENSOR} Sigma = $\sigma_{xx}$,$\sigma_{yy}$,$\sigma_{zz}$,$\sigma_{xy}$,$\sigma_{xz}$,$\sigma_{yz}$ /
-
-\&\textbf{INIT\_STRESS} S1, S2, S3 {[},n1, n2, n3{]} /
-
-followed by (n1+n2+n3) \&\textbf{DIST2D} blocks
-
-\&\textbf{SWF} mus, mud, dc {[}, nmus, nmud, ndc{]} (weakening_kind=1 for linear (default); weakening_kind=2 for exponential)/
-
-\&\textbf{TWF} nuc\_x, nuc\_y, nuc\_z, nuc\_r, nuc\_t0, nuc\_v /
-
-\&\textbf{RSF} V0,f0,a,b,L,V\_init,theta\_init,C,StateLaw {[} nV0,nf0,na,nb,nL,nV\_init,ntheta\_init,nC {]} /
+\&\textbf{SWF} mus, mud, dc {[}, nmus, nmud, ndc{]} (weakening\_kind=1 for linear (default); weakening\_kind=2 for exponential) /\\
+\&\textbf{TWF} nuc\_x, nuc\_y, nuc\_z, nuc\_r, nuc\_t0, nuc\_v /\\
+\&\textbf{RSF} V0,f0,a,b,L,V\_init,theta\_init,C,StateLaw {[} nV0,nf0,na,nb,nL,nV\_init,ntheta\_init,nC {]} /\\
 
@@
 \item [{nC}] = number of heterogeneous items for cohesion
-\item [{weakening_kind}] = 1 for linear slip-weakening law (default); 2 for exponential slip-weakening law: $\mu_d + (\mu_s-\mu_d) \exp(-\frac{u}{d_c})$
+\item [{weakening\_kind}] = 1 for linear slip-weakening law (default); 2 for exponential slip-weakening law:
+                            $$\mu = \mu_d + (\mu_s-\mu_d) exp(-\frac{u}{d_c}) $$
 \end{description}

### 2021-11-25 doc/USER_MANUAL/06_fault_sources.tex
Add read-from-file option for heterogeneous fault parameters (#2)
* Add read-from-file option for heterogeneous fault parameters

* Update user manual for read-from-file

* Minor edits of user manual

* Minor edits of user manual

* Minor edits in User_manual

* Minor edits in User_manual

* Minor edits in User_manual

* Minor edits in User_manual

* Minor edits in User_manual

* Minor edits in User_manual

--- 2021-11-16 doc/USER_MANUAL/06_fault_sources.tex
@@
 
-\&\textbf{SWF} mus, mud, dc {[}, nmus, nmud, ndc{]} (weakening_kind=1 for linear (default); weakening_kind=2 for exponential)/
+\&\textbf{SWF} mus, mud, dc {[}, nmus, nmud, ndc{]} (weakening\_kind=1 for linear (default); weakening\_kind=2 for exponential)/
 
@@
 \item [{nC}] = number of heterogeneous items for cohesion
-\item [{weakening_kind}] = 1 for linear slip-weakening law (default); 2 for exponential slip-weakening law: $$\mu_d + (\mu_s-\mu_d) \exp(-\frac{u}{d_c})$$
+\item [{weakening\_kind}] = 1 for linear slip-weakening law (default); 2 for exponential slip-weakening law:
+                            $$\mu = \mu_d + (\mu_s-\mu_d) exp(-\frac{u}{d_c}) $$
 \end{description}
@@
 $\;$sets a constant value (val) within a sphere with center (xc,yc,zc)
-and radius r. \\
+and radius r.
+
+\&\textbf{DIST2D} shapeval='$\mathtt{read-from-file}$', filename
+/
+
+$\;$sets the name of the filename saved in the DATA directory. 
+The first line in the file shall be the number of data points and the 
+following lines in the file has the format: 
+x, y, z, val. Method of interpolation is nearest. 
+\\
 

### 2022-03-04 doc/USER_MANUAL/04_creating_databases.tex
corrects typos

@@
 \item [{\texttt{STACEY\_ABSORBING\_CONDITIONS}}] Set to \texttt{.true.}
-to turn on Clayton-Enquist absorbing boundary conditions (see \citet{KoTr99}).
+to turn on Clayton-Engquist absorbing boundary conditions (see \citet{KoTr99}).
 In almost all cases it is much better to use CPML absorbing layers

### 2022-03-05 doc/USER_MANUAL/05_running_the_solver.tex
updates reading Par_file location for simultaneous runs (single file, should be in ./DATA/, not in run**/DATA/)

@@
 
+\section{Note on the simultaneous simulation of several earthquakes}
+
 We have also added the ability to run several calculations (several earthquakes)
@@
 To turn that option on, set parameter \texttt{NUMBER\_OF\_SIMULTANEOUS\_RUNS}
-to a value greater than 1 in file setup/constants.h.in before
-configuring and compiling the code.
+to a value greater than 1 in file \texttt{DATA/Par\_file}.
+
 When that option is on, of course the number of processor cores used to start
@@
 should be \texttt{NUMBER\_OF\_SIMULTANEOUS\_RUNS $\times$ NPROC}.
-All the runs to perform must be placed in directories called \texttt{run0001}, \texttt{run0002}, \texttt{run0003} and so on (with exactly four digits)
-and you must create a link from the root directory of the code
-to the first copy of the executable programs by typing \texttt{ln -s run0001/bin bin}.
+
+\begin{figure}[H]
+\begin{centering}
+\includegraphics[width=2.2in]{figures/simultaneous_dir_struct.pdf}
+\par\end{centering}
+%
+\caption{
+Directory structure when simulating several earthquakes at once.
+To improve readability, only directories have been drawn.}
+\label{fig:simultaneous_dir_struct}
+\end{figure}
+
+Figure~\ref{fig:simultaneous_dir_struct} shows what the directory structure
+should looks like when simulating multiple earthquakes at ones.
+All the runs to perform must be placed in directories called \texttt{run0001}, \texttt{run0002}, \texttt{run0003} and so on (with exactly four digits).
+
+\begin{itemize}
+\item The simulation is launched within the root directory \texttt{EXAMPLE\_ROOT\_DIR}\\
+(usually \texttt{mpirun -np N ./bin/xspecfem3D}).
+\item \texttt{DATA} should contain the \texttt{Par\_file} parameter file with
+\texttt{NUMBER\_OF\_SIMULTANEOUS\_RUNS} as explained in Chapter~\ref{cha:Creating-Distributed-Databases}.
+\item \texttt{DATABASES\_MPI} and {OUTPUT\_FILES} directory may contain the mesher output but they
+are not required as they are superseded by the ones in the \texttt{runXXX} directories.
+\item \texttt{runXXXX} directories must be created beforehand. There should be be as many as
+\texttt{NUMBER\_OF\_SIMULTANEOUS\_RUNS} and the numbering should be contiguous, starting from \texttt{0001}.
+They all should have \texttt{DATA},
+\texttt{DATABASES\_MPI} and \texttt{OUTPUT\_FILES} directories. Additionally a \texttt{SEM} directory
+containing adjoint sources have to be created to perform adjoint simulations.
+\item \texttt{runXXXX/DATA} directories must all contain a \texttt{CMTSOLUTION} file,
+a \texttt{STATIONS} file along with an eventual \texttt{STATIONS\_ADJOINT} file.
+\item If \texttt{BROADCAST\_SAME\_MESH\_AND\_MODEL} is set to \texttt{.true.} in \texttt{DATA/Par\_file},
+only \texttt{run0001/OUTPUT\_FILES} and \texttt{run0001/DATABASES\_MPI} directories need to contain the files
+outputted by the mesher.
+\item If \texttt{BROADCAST\_SAME\_MESH\_AND\_MODEL} is set to \texttt{.false.} in \texttt{DATA/Par\_file},
+every \texttt{runXXXX/OUTPUT\_FILES} and \texttt{runXXXX/DATABASES\_MPI} directories need to contain the files
+outputted by the mesher. Note that while the meshes might have been created from different models and
+parameter sets, they should have been created using the same number of MPI processes.
+\end{itemize}
 

### 2022-03-25 doc/USER_MANUAL/03_mesh_generation.tex
updates doc

@@
 Note that Qmu is always equal to Qs, but Qkappa is in general not equal to Qp.
-To convert one to the other see doc/note\_on\_Qkappa\_versus\_Qp.pdf and utils/attenuation/conversion\_from\_Qkappa\_Qmu\_to\_Qp\_Qs\_from\_Dahlen\_Tromp\_959\_960.f90.
+
+To convert one to the other see \texttt{doc/note\_on\_Qkappa\_versus\_Qp.pdf} and the helper code\\
+\texttt{conversion\_from\_Qkappa\_Qmu\_to\_Qp\_Qs\_from\_Dahlen\_Tromp\_959\_960.f90} in folder
+\texttt{utils/attenuation/}.
 
@@
 elements can have 8 or 27 nodes.\\
- See picture doc/mesh\_numbering\_convention/numbering\_convention\_27\_nodes.jpg
+ See picture \texttt{doc/mesh\_numbering\_convention/numbering\_convention\_27\_nodes.jpg}
 to see\\
@@
 Pellegrini et al. from LaBRI and INRIA in Bordeaux, France, downloadable
-from the web page \url{https://gforge.inria.fr/projects/scotch/}.\\
+from the web page \url{https://gitlab.inria.fr/scotch/scotch}.\\
 
@@
 model. In the following lines, each material needs to be defined as:
-
+%
+\begin{itemize}[leftmargin=-8ex,itemindent=8ex]
+\item[] acoustic/elastic material
+{\small
 \begin{verbatim}
-material_ID rho vp vs Q anisotropy_flag domain_ID
+material_ID rho vp vs Qkappa Qmu anisotropy_flag domain_ID
 \end{verbatim}
+}
+\item[] poroelastic material
+{\small
+\begin{verbatim}
+material_id rho_s rho_f phi tort kxx kxy kxz kyy kyz kzz kappa_s kappa_f kappa_fr eta mu_fr domain_id
+\end{verbatim}
+}
+\item[] tomographic
+{\small
+\begin{verbatim}
+material_id type-keyword domain-name tomo-filename tomo_id domain_id
+\end{verbatim}
+}
+\end{itemize}
+For acoustic and elastic material, you would specify
+\begin{itemize}
+\item \texttt{material\_ID} : a material ID number (1,2,..) to identify the material for the following mesh region selection.
+\item \texttt{rho}, \texttt{vp}, \texttt{vs} : density (in $kg/m^3$), $V_p$ and $V_s$ (in $m/s$) wave speeds
+\item \texttt{Qkappa, Qmu} : quality factor (0=no attenuation) for bulk $Q_{\kappa}$ and shear attenuation $Q_{\mu}$
+\item \texttt{anisotropy\_flag} : 0=no anisotropy / 1,2,.. check with implementation
+in \texttt{aniso\_model.f90}
+\item \texttt{domain\_id} : 1=acoustic / 2=elastic / 3=poroelastic
+\end{itemize}
+\noindent
+For poroelastic materials, you would specify the corresponding poroelastic properties.
+Please read the section below for more on poroelastic parameters.\\
 
 \noindent
-where
-\begin{itemize}
-\item \texttt{Q} : quality factor (0=no attenuation) for shear attenuation $Q_{\mu}$
-\item \texttt{anisotropy\_flag} : 0=no anisotropy / 1,2,.. check with implementation
-in \texttt{aniso\_model.f90}
-\item \texttt{domain\_id} : 1=acoustic / 2=elastic
-\end{itemize}
+For tomographic models, you would need to specify a negative \texttt{material\_id} (-1,-2,..)
+and a corresponding \texttt{domain\_id} for either acoustic (1) or elastic (2) domains. For example a line like:
+{\small
+\begin{verbatim}
+-1 tomography elastic tomography_model.xyz 0 2
+\end{verbatim}
+}
+where the specifications \texttt{tomography}, \texttt{tomography\_model.xyz}, \texttt{0} are always used by default and
+the keyword \texttt{elastic} or \texttt{acoustic} will be defined internally according to your specified \texttt{domain\_id}.
+%
 \item [{\texttt{NREGIONS}}] The number of regions in the mesh. In the following
@@
 is defined as:
+{\small
 \begin{verbatim}
@@
 \end{verbatim}
+}
 \end{description}
@@
 Now that you have set the appropriate parameters in the \texttt{Mesh\_Par\_file}
-and have compiled the mesher, you are ready to run it!
-\footnote{
+and have compiled the mesher, you are ready to run it!\\
+
 Depending on your system, please check how to run executables with MPI.
@@
 When you run on a PC cluster, the script assumes that the nodes are
-named n001, n002, etc. If this is not the case, change the \texttt{tr
--d `n'} line in the script. You may also need to edit the last command
+named \texttt{n001}, \texttt{n002}, etc.
+If this is not the case, change the \texttt{tr -d `n'} line in the script.
+You may also need to edit the last command
 at the end of the script that invokes the \texttt{mpirun} command.
@@
 for information about running the code on a system with a scheduler,
-e.g., LSF.
-}
+e.g., LSF.\\
+
 Mesher output is provided in the \texttt{OUTPUT\_FILES/} directory
@@
 \end{verbatim}}
+
+\section{Poroelastic materials}\label{cha:Poroelastic-materials}
+
+The present version of SPECFEM can handle fully saturated porous simulations.
+At the moment, the code implements Biot's equation.
+However, the code cannot calculate partially saturated cases in its current state.\\
+
+The way we prescribe material property for porous material in SPECFEM3D depends on wheter the internal mesher \texttt{xmeshfem3D}
+or an external mesher like CUBIT is taken.
+For external meshers which provide the resulting meshes in a folder like \texttt{MESH/},
+we use a file \texttt{nummaterial\_poroelastic\_file}, which needs to be added to this directory \texttt{MESH/}.
+The line format for poroelastic materials is as follow:
+{\small \begin{verbatim}
+  rhos rhof phi c kxx kxy kxz kyy kyz kzz Ks Kf Kfr etaf mufr
+\end{verbatim}}
+\begin{description}
+\item where\\
+\texttt{rho\_s} = solid density,\\
+\texttt{rho\_f} = fluid density,\\
+\texttt{phi} = porosity,\\
+\texttt{tort} = tortuosity,\\
+\texttt{kxx} = xx component of permeability tensor,\\
+\texttt{kxy} = xy,yx components of permeability tensor,\\
+\texttt{kyy} = yy component of permeability tensor,\\
+\texttt{kxz} = xz,zx components of permeability tensor,\\
+\texttt{kzz} = zz component of permeability tensor,\\
+\texttt{kappa\_s} = solid bulk modulus,\\
+\texttt{kappa\_f} = fluid bulk modulus,\\
+\texttt{kappa\_fr} = frame bulk modulus,\\
+\texttt{eta\_f} = fluid viscosity,\\
+\texttt{mu\_fr} = frame shear modulus.\\
+\end{description}
+Using an external mesh (for instance coming from CUBIT/TRELIS), poroelastic materials have the ID number 3, while 1 is acoustic and 2 is elastic (see the example in the package: EXAMPLES/homogeneous\_poroelastic).\\
+
+\noindent
+In case the internal mesher \texttt{xmeshfem3D} is used, the same poroelastic material properties will be defined in the \texttt{Mesh\_Par\_file}.
+See the above section~\ref{cha:Running-the-Mesher-Meshfem3D} on how the corresponding line format would look like.\\
+
+\noindent
+Note that attenuation is not implemented yet for poroelastic domains, thus there is no need for specifiying Q values.

### 2022-03-26 doc/USER_MANUAL/02_getting_started.tex

@@
 to partition meshes created with CUBIT. METIS \citep{KaKu98a,KaKu98c,KaKu98b}
-can also be used instead of SCOTCH if you prefer, by editing file
-src/decompose\_mesh/decompose\_mesh.F90 and uncommenting the flag
-USE\_METIS\_INSTEAD\_OF\_SCOTCH. You will also then need to install
+can also be used instead of SCOTCH if you prefer, by changing the parameter
+\texttt{PARTITIONING\_TYPE} in the \texttt{Par\_file}. You will also then need to install
 and compile Metis version 4.0 (do {*}NOT{*} install Metis version
-5.0, which has incompatible function calls) and edit \texttt{src/decompose\_mesh/Makefile.in}
+5.0, which has incompatible function calls) and edit \texttt{Makefile.in}
 and uncomment the METIS link flag in that file before running \texttt{configure}.\\
@@
 developed by Fran\c{c}ois Pellegrini et al. from LaBRI and INRIA in Bordeaux,
-France, downloadable from the web page \url{https://gforge.inria.fr/projects/scotch/}.
+France, downloadable from the web page \url{https://gitlab.inria.fr/scotch/scotch}.
 In case no SCOTCH libraries can be found on the system, the configuration

### 2022-03-26 doc/USER_MANUAL/04_creating_databases.tex

@@
 
-The present version of SPECFEM can handle fully saturated porous simulations.
-(Christina Morency implemented Biot's equation.
-However, the code cannot calculate partially saturated cases in its state.
-Christina Morency is presently working on a dual porosity, dual permeability formulation, type Pride and Berryman, for an other project, but it will not be available for some time.)
-The way we prescribe material property for porous material in SPECFEM3D is as follow.
-We use a file \texttt{nummaterial\_poroelastic\_file}, which is located in the directory \texttt{MESH/}. The line format is as follow:
-{\small \begin{verbatim}
-  rhos rhof phi c kxx kxy kxz kyy kyz kzz Ks Kf Kfr etaf mufr
-\end{verbatim}}
-\begin{description}
-\item where\\
-\texttt{rho\_s} = solid density,\\
-\texttt{rho\_f} = fluid density,\\
-\texttt{phi} = porosity,\\
-\texttt{tort} = tortuosity,\\
-\texttt{kxx} = xx component of permeability tensor,\\
-\texttt{kxy} = xy,yx components of permeability tensor,\\
-\texttt{kyy} = yy component of permeability tensor,\\
-\texttt{kxz} = xz,zx components of permeability tensor,\\
-\texttt{kzz} = zz component of permeability tensor,\\
-\texttt{kappa\_s} = solid bulk modulus,\\
-\texttt{kappa\_f} = fluid bulk modulus,\\
-\texttt{kappa\_fr} = frame bulk modulus,\\
-\texttt{eta\_f} = fluid viscosity,\\
-\texttt{mu\_fr} = frame shear modulus.\\
-\end{description}
-Using an external mesh (for instance coming from CUBIT/TRELIS), poroelastic materials have the ID number 3, while 1 is acoustic and 2 is elastic (see the example in the package: EXAMPLES/homogeneous\_poroelastic).\\
+\subsection{PML absorbing boundary layers}
 

### 2022-03-30 doc/USER_MANUAL/06_fault_sources.tex
Remove a conflict

--- 2021-11-16 doc/USER_MANUAL/06_fault_sources.tex
@@
 
-\&\textbf{SWF} mus, mud, dc {[}, nmus, nmud, ndc{]} (weakening_kind=1 for linear (default); weakening_kind=2 for exponential)/
+\&\textbf{SWF} mus, mud, dc {[}, nmus, nmud, ndc{]} /
 
@@
 \item [{nC}] = number of heterogeneous items for cohesion
-\item [{weakening_kind}] = 1 for linear slip-weakening law (default); 2 for exponential slip-weakening law:
-                            $$\mu = \mu_d + (\mu_s-\mu_d) exp(-\frac{u}{d_c}) $$
 \end{description}

### 2022-03-31 doc/USER_MANUAL/11_graphics.tex

@@
 For all domains,
-the velocity field is output to files:
+the either the velocity or displacement field is output to files (depending on setting \texttt{SAVE\_DISPLACEMENT}):
 {\small
@@
 \noindent
-For elastic domains, the divergence and curl taken from the velocity field,
-i.e. $\nabla\cdot{\bf {v}}$ and $\nabla\times{\bf {v}}$, get stored
+or
+{\small
+\begin{verbatim}
+proc??????_displ_X_it??????.bin
+proc??????_displ_Y_it??????.bin
+proc??????_displ_Z_it??????.bin
+\end{verbatim}
+}
+\noindent
+For elastic domains, the divergence and curl taken from the velocity (or displacement) field,
+i.e. $\nabla\cdot{\bf {v}}$ and $\nabla\times{\bf {v}}$ (or $\nabla\cdot{\bf {d}}$ and $\nabla\times{\bf {d}}$), get stored
 as well:

### 2022-05-17 doc/USER_MANUAL/11_graphics.tex
adds option (MOVIE_TYPE == 3) for shakemap to output geometric mean of horizontal PG-values

@@
 \begin{itemize}
-\item \texttt{\small MOVIE\_TYPE}: if set to $1$, the horizontal peak-ground values of displacement/velocity/acceleration are output.
-if set to $2$, the maximum length of the particle displacement/velocity/acceleration vector is output.
+\item \texttt{\small MOVIE\_TYPE}: if set to $1$, the horizontal peak-ground values of displacement/velocity/acceleration are output; if set to $2$, the maximum length of the particle displacement/velocity/acceleration vector is output; if set to $3$, the geometric mean of horizontal peak-ground values of displacement/velocity/acceleration are output.
 Please be aware that these peak-ground values can differ from each other.

### 2022-05-17 src/specfem3D/initialize_simulation.F90

@@ -378,8 +378,8 @@
   endif
 
   ! checks the MOVIE_TYPE parameter
-  if (MOVIE_TYPE /= 1 .and. MOVIE_TYPE /= 2) then
-    stop 'error: MOVIE_TYPE must be either 1 or 2! Please modify Par_file and recompile solver'
+  if (MOVIE_TYPE < 1 .or. MOVIE_TYPE > 3) then
+    stop 'error: MOVIE_TYPE must be either 1, 2 or 3! Please modify Par_file and recompile solver'
   endif
 
   ! check that the code has been compiled with the right values

### 2022-08-17 doc/USER_MANUAL/01_introduction.tex

--- 2018-04-09 doc/USER_MANUAL/01_introduction.tex
@@
 \chapter{Introduction}
-
-\red{\Large ANNOUNCEMENT: SPECFEM3D can now perform full waveform inversion (FWI), i.e. invert for models in an iterative fashion, and it can also perform source inversions in a constant model;
-please refer to the two new directories, \texttt{inverse\_problem\_for\_model} and \texttt{inverse\_problem\_for\_source}, and the README files they contain.
-For FWI inversions for the model, also refer to the new examples provided in
-directory \texttt{EXAMPLES}.}\\
-
-\red{OTHER ANNOUNCEMENT: SPECFEM3D can now perform gravity field calculations in addition (or instead of) seismic wave propagation only.
-See flag \texttt{GRAVITY\_INTEGRALS} in file \texttt{setup/constants.h.in}. Please also refer to \url{http://komatitsch.free.fr/preprints/GJI_Martin_gravimetry_2017.pdf}.
-And yes, that is the reason why I added a falling apple on the cover of the manual :-).
-Note that SPECFEM3D can also model transient gravity perturbations induced by earthquake rupture, as developed and explained in \cite{HaAmBaChMoSoWh2015}.
-These are two different things, and both are implemented and avaible in SPECFEM3D.
-To use the second feature, please refer to doc/how\_to/README\_gravity\_perturbation.txt.}\\
 
@@
 
-\red{In fluids, when gravity is turned off, SPECFEM3D uses the classical
+In fluids, when gravity is turned off, SPECFEM3D uses the classical
 linearized Euler equation; thus if you have sharp local variations of
@@
 approximation in the case you want to study, and/or see if you should turn gravity on.
-For more details on that see e.g. \cite{COA2011}.}\\
+For more details on that see e.g. \cite{COA2011}.\\
 
@@
 
-\red{Beware that, in a spectral-element method, some spurious modes (that have some similarities with classical so-called "Hourglass modes" in finite-element techniques,
+Beware that, in a spectral-element method, some spurious modes (that have some similarities with classical so-called "Hourglass modes" in finite-element techniques,
 although in the SEM they are not zero-energy modes) can appear in some (but not all) cases in the spectral element in which the source is located.
@@
 If that is the case, an easy option is to slightly change the mesh in the source region in order to get rid of these Hourglass-like spurious modes,
-as explained in \cite{DuLiScGa14}, in which this phenomenon is described in details, and in which practical solutions to avoid it are suggested.}\\
+as explained in \cite{DuLiScGa14}, in which this phenomenon is described in details, and in which practical solutions to avoid it are suggested.\\
 
@@
 but for linear rheologies only so far. To accommodate visco-plastic
-or non-linear rheologies, the reader can refer to the \href{http://geoelse.stru.polimi.it/}{GeoELSE}software
+or non-linear rheologies, the reader can refer to the \href{http://geoelse.stru.polimi.it/}{GeoELSE} software
 package \citep{CaGa97,StPaIg09}.
+
+\subsection*{Announcements}
+\begin{itemize}
+\item {\bf FWI}: SPECFEM3D can now perform full waveform inversion (FWI), i.e. invert for models in an iterative fashion, and it can also perform source inversions in a constant model;
+please refer to the two new directories, \texttt{inverse\_problem\_for\_model} and \texttt{inverse\_problem\_for\_source}, and the README files they contain.
+For FWI inversions for the model, also refer to the new examples provided in
+directory \texttt{EXAMPLES}.\\
+
+\item {\bf External coupling}: SPECFEM3D can now perform coupling with an external code (DSM, AxiSEM or FK) based on a database of displacement vectors and traction vectors
+on the outer edges of the mesh created once and for all (see \cite{MoChKoFu13,MoChKoWa15,WaChMoKoMo16,ToChKoBaLi14,ToKoTsHuChBaLi14}, and if you use that feature of the code please cite at least one of these papers).
+\\
+
+To use coupling with FK, just use the set of parameters that is in the DATA/Par\_file input file of the code: \\
+%
+\begin{verbatim}
+#-----------------------------------------------------------
+#
+# Coupling with an injection technique (DSM, AxiSEM, or FK)
+#
+#-----------------------------------------------------------
+COUPLE_WITH_INJECTION_TECHNIQUE = .false.
+INJECTION_TECHNIQUE_TYPE        = 3       # 1 = DSM, 2 = AxiSEM, 3 = FK
+MESH_A_CHUNK_OF_THE_EARTH       = .false.
+TRACTION_PATH                   = ./DATA/AxiSEM_tractions/3/
+FKMODEL_FILE                    = FKmodel
+RECIPROCITY_AND_KH_INTEGRAL     = .false. # does not work yet
+\end{verbatim}
+That part (coupling with FK) is actively maintained and works fine. See e.g. \href{https://doi.org/10.1002/2014GL061644}{GRL\_Ping\_Tong\_2014} for some examples.
+There is also an example that is provided with the code:\\
+\texttt{specfem3d/EXAMPLES/small\_example\_coupling\_FK\_specfem}. \\
+
+Regarding coupling with DSM, that part is not actively maintained any more, but it is still included in the code, you may have to test it again and make minor adjustments if needed.
+The necessary tools are in directory \texttt{specfem3d/EXTERNAL\_PACKAGES\_coupled\_with\_SPECFEM3D/DSM\_for\_SPECFEM3D}, and there is a \texttt{README} file in \texttt{specfem3d/EXTERNAL\_PACKAGES\_coupled\_with\_SPECFEM3D} that should be more or less up-to-date (there are about four steps to follow in total, the first one being creating the database of DSM tractions and displacements on the edges of the coupling box).
+See e.g. \href{https://doi.org/10.1093/gji/ggs006}{GJI\_Vadim\_2013} and \href{https://doi.org/10.1093/gji/ggv189}{GJI\_Vadim\_2015} for some examples.\\
+
+\item {\bf Gravity}: SPECFEM3D can now perform gravity field calculations in addition (or instead of) seismic wave propagation only.
+See flag \texttt{GRAVITY\_INTEGRALS} in file \texttt{setup/constants.h.in}. Please also refer to \href{https://doi.org/10.1093/gji/ggx010}{GJI\_Martin\_gravimetry\_2017}.
+And yes, that is the reason why Dimitri added a falling apple on the cover of the manual :-).
+Note that SPECFEM3D can also model transient gravity perturbations induced by earthquake rupture, as developed and explained in \cite{HaAmBaChMoSoWh2015}.
+These are two different things, and both are implemented and avaible in SPECFEM3D.
+To use the second feature, please refer to \texttt{doc/how\_to/README\_gravity\_perturbation.txt}.\\
+
+\end{itemize}
 

### 2022-09-16 doc/USER_MANUAL/02_getting_started.tex
adds AMD MI250 configuration support (for Frontier)

@@
 \end{verbatim}
-where for example \texttt{MI8,MI25,MI50,MI100,..} specifies the target GPU architecture of your card.\\
+where for example \texttt{MI8,MI25,MI50,MI100,MI250,..} specifies the target GPU architecture of your card.
+Additional compilation flags can be added by specifying \texttt{HIP\_FLAGS}, as for example:
+{\small
+\begin{verbatim}
+  ./configure --with-hip=MI250 \
+      HIP_FLAGS="-fPIC -ftemplate-depth-2048 -fno-gpu-rdc -std=c++17 \
+                 -O2 -fdenormal-fp-math=ieee -fcuda-flush-denormals-to-zero -munsafe-fp-atomics" \
+      ..
+\end{verbatim}
+}
 

### 2022-10-28 doc/USER_MANUAL/authors.tex
Update small issues in documentation [#1531]

@@
-\section*{Authors}
+\section*{https://specfem3d.readthedocs.io/en/latest/authors/}
 
@@
 %%%%%%%% please do NOT merge the lines of the paragraph below, it makes it difficult to insert new names in alphabetical order; leave a single name per line
-Since then it has been developed and maintained by a development team: in alphabetical order,
-Michael Afanasiev,
-Jean-Paul (Pablo) Ampuero,
-\'Etienne Bachmann,
-Kangchen Bai,
-Piero Basini,
-C\'eline Blitz,
-Alexis Bottero,
-Ebru Bozda\u{g},
-Emanuele Casarotti,
-Joseph Charles,
-Min Chen,
-Paul Cristini,
-Cl\'ement Durochat,
-Percy Galvez,
-Hom Nath Gharti,
-Dominik G\"oddeke,
-Vala Hj\"orleifsd\'ottir,
-Sue Kientz,
-Dimitri Komatitsch,
-Jes\'us Labarta,
-Nicolas Le Goff,
-Pieyre Le Loher,
-Matthieu Lefebvre,
-Qinya Liu,
-Youshan Liu,
-David Luet,
-Yang Luo,
-Alessia Maggi,
-Federica Magnoni,
-Roland Martin,
-Ren\'e Matzen,
-Dennis McRitchie,
-Matthias Meschede,
-Peter Messmer,
-David Mich\'ea,
-Vadim Monteiller,
-Surendra Nadh Somala,
-Masaru Nagaso,
-Tarje Nissen-Meyer,
-Daniel Peter,
-Kevin Pouget,
-Max Rietmann,
-Elliott Sales de Andrade,
-Brian Savage,
-Bernhard Schuberth,
-Anne Sieminski,
-James Smith,
-Leif Strand,
-Carl Tape,
-Jeroen Tromp,
-Brice Videau,
-Jean-Pierre Vilotte,
-Zhinan Xie,
-Chang-Hua Zhang,
-Hejun Zhu.\\
+Nikhil Vaidyar
 

### 2022-11-09 doc/USER_MANUAL/03_mesh_generation.tex
updates headers and usage of geodynamics

@@
 \texttt{EXAMPLES/}. We strongly encourage you to contribute your own
-example to this package by contacting the CIG Computational Seismology
-Mailing List \urlwithparentheses{cig-seismo@geodynamics.org}.
+example to this package through the github repository!
 
@@
 which explains in a step-by-step tutorial the workflow for the example.
-Please feel free to contribute your own example to this package by
-contacting the CIG Computational Seismology Mailing List \urlwithparentheses{cig-seismo@geodynamics.org}.\\
+Please feel free to contribute your own example to this package through the github repository!
 

### 2022-11-09 utils/ADJOINT_TOMOGRAPHY_TOOLS/measure_adj/USER_MANUAL/measure_adj_manual.tex
updates headers and usage of geodynamics

@@
 
-To report bugs or suggest improvements to the code, please send an email to the CIG Computational Seismology Mailing List (cig-seismo@geodynamics.org) or Carl Tape (carltape@fas.harvard.edu), and/or use our online bug tracking system Roundup (www.geodynamics.org/roundup).
+To report bugs or suggest improvements to the code, please use the Issues feature on the github repository.
 

### 2022-11-09 doc/USER_MANUAL/01_introduction.tex
Path updates for parameters moved from constants.h to Par_file and vice versa, and other minor corrections.

@@
 
-To use coupling with FK, just use the set of parameters that is in the DATA/Par\_file input file of the code: \\
+To use coupling with FK, just use the set of parameters that is in the \texttt{DATA/Par\_file} input file of the code: \\
 %
@@
 \item {\bf Gravity}: SPECFEM3D can now perform gravity field calculations in addition (or instead of) seismic wave propagation only.
-See flag \texttt{GRAVITY\_INTEGRALS} in file \texttt{setup/constants.h.in}. Please also refer to \href{https://doi.org/10.1093/gji/ggx010}{GJI\_Martin\_gravimetry\_2017}.
+See flag \texttt{GRAVITY\_INTEGRALS} in file \texttt{setup/constants.h}. Please also refer to \href{https://doi.org/10.1093/gji/ggx010}{GJI\_Martin\_gravimetry\_2017}.
 And yes, that is the reason why Dimitri added a falling apple on the cover of the manual :-).

### 2022-11-09 doc/USER_MANUAL/02_getting_started.tex
Path updates for parameters moved from constants.h to Par_file and vice versa, and other minor corrections.

@@
 can also be used instead of SCOTCH if you prefer, by changing the parameter
-\texttt{PARTITIONING\_TYPE} in the \texttt{Par\_file}. You will also then need to install
+\texttt{PARTITIONING\_TYPE} in the \texttt{DATA/Par\_file}. You will also then need to install
 and compile Metis version 4.0 (do {*}NOT{*} install Metis version
@@
 Make sure that your environment variables reference it.
-\item You may want to change ADIOS related values in the \texttt{constants.h.in} file.
+\item You may want to change ADIOS related values in the \texttt{setup/constants.h} file.
 The default values probably suit most cases.
@@
 Now that you have set the compiler information, you need to select
-a number of flags in the \texttt{constants.h} file depending on your
+a number of flags in the \texttt{setup/constants.h} file depending on your
 system:
@@
 as much memory. Select your preference by selecting the appropriate
-setting in the \texttt{constants.h} file:
+setting in the \texttt{setup/constants.h} file:
 \begin{description}

### 2022-11-09 doc/USER_MANUAL/03_mesh_generation.tex
Path updates for parameters moved from constants.h to Par_file and vice versa, and other minor corrections.

--- 2022-03-25 doc/USER_MANUAL/03_mesh_generation.tex
@@
 this reference frequency, you change the value of \texttt{ATTENUATION\_f0\_REFERENCE}
-in the main constants file \texttt{constants.h} found in subdirectory
-\texttt{src/shared/}.
+defined (in Hz) in input file \texttt{DATA/Par\_file}.
 The code uses a constant $Q$ quality factor, write(IMAIN,*) "but approximated based on a series of Zener standard linear solids (SLS).
@@
 with no \texttt{Qmu} attenuation, both set \texttt{ATTENUATION} to
-\texttt{.false.} in the \texttt{Par\_file} and set \texttt{Qmu} to
+\texttt{.false.} in the \texttt{DATA/Par\_file} and set \texttt{Qmu} to
 9999 here. If you want a model with no \texttt{Qkappa} attenuation, set \texttt{Qkappa} to 9999 here.
@@
 The partitioning is done in serial for now (in the next release we
-will provide a parallel version of that code). It needs to be run in the main directory because it expects the \texttt{./DATA/Par\_file}.
+will provide a parallel version of that code). It needs to be run in the main directory because it expects the \texttt{DATA/Par\_file}.
 The synopsis is:
@@
 databases, and have to reside in the directory \texttt{LOCAL\_PATH}
-specified in the main \texttt{Par\_file}, e.g. in directory \texttt{./OUTPUT\_FILES/DATABASES\_MPI}.
+specified in the main \texttt{DATA/Par\_file}, e.g. in directory \texttt{./OUTPUT\_FILES/DATABASES\_MPI}.
 Please see Chapter~\ref{cha:Creating-Distributed-Databases} for
@@
 element, i.e., the number of Gauss-Lobatto-Legendre points, is determined
-by \texttt{NGLLX} in the \texttt{constants.h} file. We generally use
+by \texttt{NGLLX} in the \texttt{setup/constants.h} file. We generally use
 $\mbox{\texttt{NGLLX\/}}=5$, for a total of $5^{3}=125$ points per
@@
 \item [{\texttt{NDOUBLINGS}}] The number of horizontal doubling layers.
-Must be set at least to \texttt{1} if \texttt{USE\_REGULAR\_MESH} is set to \texttt{.true.}.
+Must be set at least to \texttt{1} if \texttt{USE\_REGULAR\_MESH} is set to \texttt{.false.}.
 Multiple mesh doublings can be chosen, for which each an \texttt{NZ\_DOUBLING\_**} entry must be given.
@@
 \item [{\texttt{NZ\_DOUBLING\_1}}] The position of the first doubling layer
-(only interpreted if \texttt{USE\_REGULAR\_MESH} is set to \texttt{.true.}).
+(only interpreted if \texttt{USE\_REGULAR\_MESH} is set to \texttt{.false.}).
 \item [{\texttt{NZ\_DOUBLING\_2}}] The position of the second doubling
-layer (only interpreted if \texttt{USE\_REGULAR\_MESH} is set to \texttt{.true.}
+layer (only interpreted if \texttt{USE\_REGULAR\_MESH} is set to \texttt{.false.}
 and if \texttt{NDOUBLINGS} is set to \texttt{2}). Doubling layers must be at least \texttt{2} layers apart.
@@
 distributed databases, and have to reside in the directory \texttt{LOCAL\_PATH}
-specified in the main \texttt{Par\_file}, e.g. in directory \texttt{OUTPUT\_FILES/DATABASES\_MPI}.
+specified in the main \texttt{DATA/Par\_file}, e.g. in directory \texttt{OUTPUT\_FILES/DATABASES\_MPI}.
 Please see Chapter~\ref{cha:Creating-Distributed-Databases} for
@@
 output can be directed to the screen instead by uncommenting a line
-in \texttt{constants.h}:
+in \texttt{setup/constants.h}:
 {\small

### 2022-11-09 doc/USER_MANUAL/04_creating_databases.tex
adds option (MOVIE_TYPE == 3) for shakemap to output geometric mean of horizontal PG-values
Path updates for parameters moved from constants.h to Par_file and vice versa, and other minor corrections.

@@
 bathymetry should be read in based upon the topography file specified
-in the main constants file \texttt{constants.h} found in subdirectory
-\texttt{src/shared/} to evaluate elevations. If not set, elevations
+in the main constants file \texttt{setup/constants.h} to evaluate elevations. If not set, elevations
 will be read from the numerical mesh.
@@
 frequency. To change this reference frequency, you change the value
-of \texttt{ATTENUATION\_f0\_REFERENCE} in the main constants file
-\texttt{constants.h} found in subdirectory \texttt{src/shared/}.
+of \texttt{ATTENUATION\_f0\_REFERENCE} defined (in Hz) in input file \texttt{DATA/Par\_file}.
 The code uses a constant $Q$ quality factor, write(IMAIN,*) "but approximated based on a series of Zener standard linear solids (SLS).
@@
 \item [{\texttt{CREATE\_SHAKEMAP}}] Set this flag to \texttt{.true.} to
-create a ShakeMap\textregistered{}, i.e., a peak ground velocity map
+create a ShakeMap\textregistered{}, e.g., a peak ground velocity map
 of the maximum absolute value of the two horizontal components of
@@
 GLL mesh point instead by setting flag \texttt{USE\_BEST\_LOCATION} to \texttt{.false.} instead of \texttt{.true.} in file
-\texttt{src/shared/constants.h.in} and running the \texttt{configure} script again and recompiling the code.
+\texttt{setup/constants.h} and recompiling the code.
 \item [{\texttt{USE\_RICKER\_TIME\_FUNCTION}}] Turn this flag on to use
@@
 all the individual runs must use the same number of processor cores,
-which as usual is NPROC in the Par\_file,
+which as usual is NPROC in the \texttt{DATA/Par\_file},
 and thus the total number of processor cores to request from the batch system
@@
 \end{description}
-There are quite a few more parameters to change the default setup of your runs. Please check the comments in the \texttt{Par\_file} directly for further explanations.\\
+There are quite a few more parameters to change the default setup of your runs. Please check the comments in the \texttt{DATA/Par\_file} directly for further explanations.\\
 
@@
 the number of GLL points \texttt{NGLL} specified in the main constants
-file \texttt{constants.h} located in the \texttt{src/shared/} subdirectory.
-The wave speed $v$ is determined based on your model's P- (or S-)
+file \texttt{setup/constants.h}. The wave speed $v$ is determined based on your model's P- (or S-)
 wave speed values.

### 2022-11-09 doc/USER_MANUAL/05_running_the_solver.tex
Path updates for parameters moved from constants.h to Par_file and vice versa, and other minor corrections.

@@
 \noindent
-Most parameters in the \texttt{Par\_file} should be set prior to running
+Most parameters in the \texttt{DATA/Par\_file} should be set prior to running
 the databases generation. Only the following parameters may be changed
@@
 \end{itemize}
-Any other change to the \texttt{Par\_file} implies rerunning both
+Any other change to the \texttt{DATA/Par\_file} implies rerunning both
 the database generator \texttt{xgenerate\_databases} and the solver
@@
 \item Set the latitude or UTM $x$ coordinate, longitude or UTM $y$ coordinate, depth of the source (in km). Remark: In principle in the international CMTSOLUTION format in geophysics
-the depth is given in kilometers; however for users in other fields (non-destructive testing, medical imaging, near-surface studies...) who may prefer to give the position of the source (rather than its depth from the surface), or for people who use FORCESOLUTION to describe the source rather than CMTSOLUTION, we provide an option called USE\_SOURCES\_RECEIVERS\_Z in the Par\_file, and if so that position is read from CMTSOLUTION in meters rather than kilometers (and again, it is then the true position in the mesh, not the depth).
-When option USE\_SOURCES\_RECEIVERS\_Z in the Par\_file is on, this remark applies to the position of the receivers as well.
+the depth is given in kilometers; however for users in other fields (non-destructive testing, medical imaging, near-surface studies...) who may prefer to give the position of the source (rather than its depth from the surface), or for people who use FORCESOLUTION to describe the source rather than CMTSOLUTION, we provide an option called USE\_SOURCES\_RECEIVERS\_Z in the \texttt{DATA/Par\_file}, and if so that position is read from CMTSOLUTION in meters rather than kilometers (and again, it is then the true position in the mesh, not the depth).
+When option USE\_SOURCES\_RECEIVERS\_Z in the \texttt{DATA/Par\_file} is on, this remark applies to the position of the receivers as well.
 \item Set the \texttt{time shift} parameter equal to $0.0$ (the solver
@@
 function (i.e., the second derivative of a Gaussian) when {\texttt{USE\_RICKER\_TIME\_FUNCTION}} is turned on
-in the main parameter file {\texttt{Par\_file}}. In case that the
+in the main parameter file \texttt{DATA/Par\_file}. In case that the
 solver uses a (pseudo) Dirac delta source time function to represent
@@
 shift}, and all the other entries must have non-negative \texttt{time
-shift}. Each subevent can have its own half latitude, longitude, depth,
+shift}. Each subevent can have its own latitude, longitude, depth,
 \texttt{half duration} and force parameters.
@@
 
-\vspace{1cm}
-
-In addition to inbuild source time function, the solver can also use an external source time function defined by the user. This option can be
-activated by setting   \texttt{EXTERNAL\_SOURCE\_FILE} to true in \texttt{Par\_File} and by adding the name of the file
-containing the source time function at the end of \texttt{FORCESOLUTION} or \texttt{CMTSOLUTION} files.
+In addition to inbuild source time function, the solver can also use an external source time function defined by the user.
+This option can be activated by setting \texttt{USE\_EXTERNAL\_SOURCE\_FILE} to true in \texttt{DATA/Par\_file} and by adding the name of the file containing the source time function at the end of \texttt{FORCESOLUTION} or \texttt{CMTSOLUTION} files.
 The source time function file must contain the value of the time step on its first line, and then a single column with the amplitude of the source time function for all the time steps.
 The time step must be exactly the same as that used for the simulation.
-Note  when the  \texttt{EXTERNAL\_SOURCE\_FILE} is set to false then the line with the external source time function file must not appear in
-the files  \texttt{FORCESOLUTION} and \texttt{CMTSOLUTION} otherwise the solver will exit with an error.
+Note  when the  \texttt{EXTERNAL\_SOURCE\_FILE} is set to false then the line with the external source time function file must not appear in the files  \texttt{FORCESOLUTION} and \texttt{CMTSOLUTION} otherwise the solver will exit with an error.
 When using an external source file, you can still set up the source location and directivity as in the default case.
@@
 label should be no more than 32 characters (see \texttt{MAX\_LENGTH\_STATION\_NAME}
-in the \texttt{constants.h} file), and the network label should be
+in the \texttt{setup/constants.h} file), and the network label should be
 no more than 8 characters (see \texttt{MAX\_LENGTH\_NETWORK\_NAME}
-in the \texttt{constants.h} file).}
+in the \texttt{setup/constants.h} file).}
 \label{fig:Sample-STATIONS-file.}
@@
 \texttt{DATA/STATIONS} to exclude stations that are not located within
-the region given in the \texttt{Par\_file} (between \texttt{LATITUDE\_MIN}
+the region given in the \texttt{Mesh\_Par\_file} (between \texttt{LATITUDE\_MIN}
 and \texttt{LATITUDE\_MAX} and between \texttt{LONGITUDE\_MIN} and
@@
 \item [-] In principle in the international CMTSOLUTION format in geophysics
-the depth is given in kilometers; however for users in other fields (non-destructive testing, medical imaging, near-surface studies...) who may prefer to give the position of the source (rather than its depth from the surface), or for people who use FORCESOLUTION to describe the source rather than CMTSOLUTION, we provide an option called USE\_SOURCES\_RECEIVERS\_Z in the Par\_file, and if so that position is read from CMTSOLUTION in meters rather than kilometers (and again, it is then the true position in the mesh, not the depth).
-When option USE\_SOURCES\_RECEIVERS\_Z in the Par\_file is on, this remark applies to the position of the receivers as well.\\
+the depth is given in kilometers; however for users in other fields (non-destructive testing, medical imaging, near-surface studies...) who may prefer to give the position of the source (rather than its depth from the surface), or for people who use FORCESOLUTION to describe the source rather than CMTSOLUTION, we provide an option called USE\_SOURCES\_RECEIVERS\_Z in the \texttt{DATA/Par\_file}, and if so that position is read from CMTSOLUTION in meters rather than kilometers (and again, it is then the true position in the mesh, not the depth).
+When option USE\_SOURCES\_RECEIVERS\_Z in the \texttt{DATA/Par\_file} is on, this remark applies to the position of the receivers as well.\\
 \item [-] Let's say you want to place two receivers at (x1,y1,z1) and
@@
 in the \texttt{output\_solver.txt} file. Output can be directed to
-the screen instead by uncommenting a line in \texttt{constants.h}:
+the screen instead by uncommenting a line in \texttt{setup/constants.h}:
 \begin{verbatim}
@@
 On PC clusters the seismogram files are generally written to the local
-disks (the path \texttt{LOCAL\_PATH} in the \texttt{Par\_file}) and
+disks (the path \texttt{LOCAL\_PATH} in the \texttt{DATA/Par\_file} and
 need to be gathered at the end of the simulation.
@@
 \texttt{\small NTSTEP\_BETWEEN\_OUTPUT\_INFO}
-in the \texttt{\small Par\_file}.
+in the \texttt{DATA/Par\_file}.
 
-Having set the \texttt{Par\_file} parameters, and having provided
+Having set the \texttt{DATA/Par\_file} parameters, and having provided
 the \texttt{CMTSOLUTION} (or the \texttt{FORCESOLUTION}) and \texttt{STATIONS}
@@
 \noindent
-Regarding attenuation (viscoelasticity), in the Par\_file you need to select the number of standard linear solids (N\_SLS) to use to mimic a constant $Q$ quality factor.
+Regarding attenuation (viscoelasticity), in the \texttt{setup/constants.h} you need to select the number of standard linear solids (N\_SLS) to use to mimic a constant $Q$ quality factor.
 Using N\_SLS = 3 is always safe. If (and only if) you know what you are doing, you can try to reduce that in order to reduce the cost of the simulations.

### 2022-11-09 doc/USER_MANUAL/06_fault_sources.tex
Buildbot ran the code beautifier.
Remove a conflict
Path updates for parameters moved from constants.h to Par_file and vice versa, and other minor corrections.

--- 2021-11-19 doc/USER_MANUAL/06_fault_sources.tex
@@
 $\;$sets a constant value (val) within a sphere with center (xc,yc,zc)
-and radius r. \\
+and radius r.
+
+\&\textbf{DIST2D} shapeval='$\mathtt{read-from-file}$', filename
+/
+
+$\;$sets the name of the filename saved in the DATA directory.
+The first line in the file shall be the number of data points and the
+following lines in the file has the format:
+x, y, z, val. Method of interpolation is nearest.
+\\
 
@@
 The fault-id identifies the fault that contains the station. It is
-the index of appearance in the faults list after line 2 of Par\_file\_faults
+the index of appearance in the faults list after line 2 of \texttt{DATA/Par\_file\_faults}
 
@@
 
-To activate this feature, in Par\_file\_faults name list \&\textbf{RUPTURE\_SWITCHES}, set TPV16=.TRUE..
+To activate this feature, in \texttt{DATA/Par\_file\_faults} name list \&\textbf{RUPTURE\_SWITCHES}, set TPV16=.TRUE..
 
@@
   rate and state friction.
-To activate this feature, in Par\_file\_faults name list \&\textbf{RUPTURE\_SWITCHES}, set RSF\_HETE=.TRUE..
+To activate this feature, in \texttt{DATA/Par\_file\_faults} name list \&\textbf{RUPTURE\_SWITCHES}, set RSF\_HETE=.TRUE..
   The format of  {\textbf{DATA/rsf\_hete\_input\_file.txt}} is as such, in the first line there are four integers
@@
 element of cubic shape with size equal to $\mathtt{h\_fault}$.
-\item Set $\mathtt{eta}$ in $\mathtt{Par\_file\_faults}$ to (0.1 to 0.3)$\ensuremath{\times}\mathtt{dtc\_fault}$.
+\item Set $\mathtt{eta}$ in $\mathtt{DATA/Par\_file\_faults}$ to (0.1 to 0.3)$\ensuremath{\times}\mathtt{dtc\_fault}$.
 A larger $\mathtt{eta}$ damps high-frequencies more aggressively

### 2022-11-09 doc/USER_MANUAL/11_graphics.tex
Path updates for parameters moved from constants.h to Par_file and vice versa, and other minor corrections.

@@
 For this, you must set the flag \texttt{SAVE\_MESH\_FILES} to \texttt{.true.}
-in the main parameter file \texttt{Par\_file} (see Chapter~\ref{cha:Main-Parameter}
+in the main parameter file \texttt{DATA/Par\_file} (see Chapter~\ref{cha:Main-Parameter}
 for details). This will create VTK files for each single partition.
@@
 \texttt{MOVIE\_SURFACE}, \texttt{MOVIE\_VOLUME}, \texttt{MOVIE\_TYPE},
-and \texttt{NTSTEP\_BETWEEN\_FRAMES} in the \texttt{Par\_file}. Turning
+and \texttt{NTSTEP\_BETWEEN\_FRAMES} in the \texttt{DATA/Par\_file}. Turning
 on the movie flags, in particular \texttt{MOVIE\_VOLUME}, produces
@@
 \texttt{CMTSOLUTION} file to be 0.0 and \texttt{HDUR\_MOVIE} = 0.0
-in the \texttt{Par\_file} to get the highest frequencies resolved
+in the \texttt{DATA/Par\_file} to get the highest frequencies resolved
 by the simulation, but for a finite source one would keep all the
@@
 the \texttt{OUTPUT\_FILES} directory. There are several flags in the
-main parameter file \texttt{Par\_file} that control the output of
+main parameter file \texttt{DATA/Par\_file} that control the output of
 these movie data files, please see section \ref{cha:Main-Parameter} for
@@
 flag turned on, the code outputs several files in \texttt{\small LOCAL\_PATH}
-specified in the main \texttt{\small Par\_file},
+specified in the main \texttt{DATA/Par\_file},
 e.g. in directory \texttt{\small OUTPUT\_FILES/DATABASES\_MPI}.

### 2022-11-09 doc/USER_MANUAL/13_changing_the_model.tex
Path updates for parameters moved from constants.h to Par_file and vice versa, and other minor corrections.

@@
 options:\\
- \indent (1) set in the \texttt{Par\_file} the model parameter \texttt{MODEL
+ \indent (1) set in the \texttt{DATA/Par\_file} the model parameter \texttt{MODEL
 = tomo}, \\
  or for more user control: \\
- \indent (2) set in the \texttt{Par\_file} the model parameter \texttt{MODEL
+ \indent (2) set in the \texttt{DATA/Par\_file} the model parameter \texttt{MODEL
 = default}, define the negative \texttt{material\_ID} identifier for
@@
 that describe the tomography should be located in the \texttt{TOMOGRAPHY\_PATH}
-directory, set in the \texttt{Par\_file}. The format of the file,
+directory, set in the \texttt{DATA/Par\_file}. The format of the file,
 as read from \texttt{model\_tomography.f90} located in the \texttt{src/generate\_databases}
@@
 frequency. To change this reference frequency, you change the value
-of \texttt{ATTENUATION\_f0\_REFERENCE} in the main constants file
-\texttt{constants.h} found in subdirectory \texttt{src/shared/}.
+of \texttt{ATTENUATION\_f0\_REFERENCE} defined (in Hz) in input file \texttt{DATA/Par\_file}.
+
+The code uses a constant $Q$ quality factor, write(IMAIN,*) "but approximated based on a series of Zener standard linear solids (SLS).
+The approximation is thus performed in a given frequency band determined based on that \texttt{ATTENUATION\_f0\_REFERENCE} reference frequency.
 
@@
 
-To use your own external model, you can set in the \texttt{Par\_file}
+To use your own external model, you can set in the \texttt{DATA/Par\_file}
 the model parameter \texttt{MODEL = external}. Three-dimensional acoustic
@@
 significant figure following the decimal, or the maximum value (9000),
-consult \texttt{shared/constants.h}.
+consult \texttt{setup/constants.h}.
 
@@
 
-To use your anisotropic models, you can either set in the \texttt{Par\_file}
+To use your anisotropic models, you can either set in the \texttt{DATA/Par\_file}
 the model parameter \texttt{ANISOTROPY} to \texttt{.true.} or set

### 2022-11-09 doc/USER_MANUAL/14_post_processing.tex
Path updates for parameters moved from constants.h to Par_file and vice versa, and other minor corrections.

@@
 You do have further options to change this default output behavior,
-given in the main constants file \texttt{constants.h} located in \texttt{src/shared/}
-directory:
+given in the \texttt{DATA/Par\_file}:
 \begin{description}
-\item [{\texttt{SEISMOGRAMS\_BINARY}}] set to \texttt{.true.} to have seismograms
+\item [{\texttt{USE\_BINARY\_FOR\_SEISMOGRAMS}}] Set to \texttt{.true.} to have seismograms
 written out in binary format.

### 2022-11-09 doc/USER_MANUAL/02_getting_started.tex

--- 2022-09-16 doc/USER_MANUAL/02_getting_started.tex
@@
 \begin{verbatim}
-git clone --recursive --branch devel https://github.com/geodynamics/specfem3d.git
+git clone --recursive --branch devel https://github.com/SPECFEM/specfem3d.git
 \end{verbatim}
@@
 this is the case for most machines and most compilers, but not all. To disable this feature, use option \texttt{-{}-disable-vectorization}.
-For more details see \href{https://github.com/geodynamics/specfem3d/issues/81}{github.com/geodynamics/specfem3d/issues/81} .
+For more details see \href{https://github.com/SPECFEM/specfem3d/issues/81}{github.com/SPECFEM/specfem3d/issues/81} .
 To check if that option works fine on your machine, run the code with and without it for an acoustic/elastic model and make sure the seismograms are identical.\\
@@
 to access the development branch of the source code with read/write access (in a safe way, no need to worry too much about breaking the package, there are CI tests based on BuildBot, Travis-CI and Jenkins in place that are checking and validating all new contributions and changes), please visit this Web page:\\
-\url{https://github.com/geodynamics/specfem3d/wiki}
+\url{https://github.com/SPECFEM/specfem3d/wiki}

### 2022-11-09 doc/USER_MANUAL/15_informations_for_developers.tex

@@
 \begin{verbatim}
-git clone --recursive https://github.com/geodynamics/specfem1d.git
+git clone --recursive https://github.com/SPECFEM/specfem1d.git
 \end{verbatim}
@@
 \noindent For information on how to contribute to the code, i.e., for how to make your modifications, additions or improvements part of the
-official package, see \url{https://github.com/geodynamics/specfem3d/wiki} .
+official package, see \url{https://github.com/SPECFEM/specfem3d/wiki} .

### 2022-11-11 doc/USER_MANUAL/07_adjoint_simulations.tex

@@
 and \ref{sec:Adjoint-simulation-finite}, respectively. The two related
-parameters in the \texttt{Par\_file} are \texttt{SIMULATION\_TYPE}
+parameters in the \texttt{DATA/Par\_file} are \texttt{SIMULATION\_TYPE}
 (1, 2 or 3) and the \texttt{SAVE\_FORWARD} (boolean).
@@
 More conveniently, use the scripts \texttt{utils/change\_simulation\_type.pl}
-to modify the \texttt{Par\_file} automatically (\texttt{change\_simulation\_type.pl
+to modify the \texttt{DATA/Par\_file} automatically (\texttt{change\_simulation\_type.pl
 -a}). Then run the solver to launch the adjoint simulation.
@@
 }\texttt{\small =}{\small {} }\texttt{\small 1}{\small{} and }\texttt{\small SAVE\_FORWARD
-=}{\small {} }\texttt{\small .true.}{\small{} in the }\texttt{\small Par\_file}{\small{}
+=}{\small {} }\texttt{\small .true.}{\small{} in the }\texttt{DATA/Par\_file}{\small{}
 (}\texttt{\small change\_simulation\_type -F}{\small ), and run the
@@
 of finite-frequency kernels; therefore set \texttt{ATTENUATION = .false.}
-in the \texttt{Par\_file}.
+in the \texttt{DATA/Par\_file}.
 \item We also suggest you modify the half duration of the \texttt{CMTSOLUTION}
@@
 in the \texttt{SEM/} directory, set \texttt{SIMULATION\_TYPE
-= 3} and \texttt{SAVE\_FORWARD = .false.} in the \texttt{Par\_file}
+= 3} and \texttt{SAVE\_FORWARD = .false.} in the \texttt{DATA/Par\_file}
 (you can use \texttt{change\_simulation\_type.pl -b}), and rerun the
@@
 system. This is done by setting \texttt{ANISOTROPIC\_KL} \texttt{=}
-\texttt{.true.} in \texttt{constants.h} before compiling the package.
+\texttt{.true.} in \texttt{DATA/Par\_file} before compiling the package.
 The definition of the parameters $C_{IJ}$ in terms of the corresponding
@@
 If \texttt{ANISOTROPIC\_KL} \texttt{=} \texttt{.true.} by additionally
-setting \texttt{ANISOTROPIC\_KL} \texttt{=} \texttt{.true.} in \texttt{constants.h}
+setting \texttt{SAVE\_TRANSVERSE\_KL} \texttt{=} \texttt{.true.} in \texttt{DATA/Par\_file}
 the package will save anisotropic kernels parameterized as velocities

### 2022-11-11 doc/USER_MANUAL/09_noise_simulations.tex

@@
 With all of the above done, we can finally launch our simulations.
-Again, please make sure that the \texttt{LOCAL\_PATH} in \texttt{DATA/Par\_file}
+Again, please make sure that the \texttt{LOCAL\_PATH} in \texttt{Par\_file}
 is not globally shared. It is quite problematic to have a globally
@@
 \texttt{OUTPUT\_FILES/} or in the directory specified by \texttt{LOCAL\_PATH}
-in \texttt{DATA/Par\_file} (which are most probably on local nodes).
+in \texttt{Par\_file} (which are most probably on local nodes).
 Collect whatever you want from the local nodes to your workstation,

### 2022-11-11 doc/USER_MANUAL/A_reference_frame.tex

@@
 to use this depth value directly as $z$ coordinate by changing the
-flag \texttt{USE\_SOURCES\_RECVS\_Z} to \texttt{.true.} in the file
-\texttt{constants.h} located in the \texttt{src/shared/} subdirectory.
+flag \texttt{USE\_SOURCES\_RECEIVERS\_Z} to \texttt{.true.} in
+\texttt{DATA/Par\_file}.
 
@@
 The seismogram output directions are given in Cartesian $x$/$y$/$z$
-directions and not rotated any further. Changing flags in \texttt{constants.h}
-in the \texttt{src/shared/} subdirectory only rotates the seismogram
-outputs if receivers are forced to be located at the surface (\texttt{RECVS\_CAN\_BE\_BURIED\_EXT\_MESH}
+directions and not rotated any further. Changing flags in \texttt{setup/constants.h} only rotates the seismogram
+outputs if receivers are forced to be located at the surface (\texttt{RECEIVERS\_CAN\_BE\_BURIED}
 set to \texttt{.false.}) and the normal to the surface at the receiver
-location should be used (\texttt{EXT\_MESH\_RECV\_NORMAL} set to \texttt{.true.})
+location should be used (\texttt{EXTERNAL\_MESH\_RECEIVERS\_NORMAL} set to \texttt{.true.})
 as vertical. In this case, the outputs are rotated to have the vertical
@@
 \texttt{OUTPUT\_FILES/} subdirectory by each parallel process. You
-can change this behavior by changing the following flags in the \texttt{constants.h}
-file located in the \texttt{src/shared/} subdirectory:
+can change this behavior by changing the following flags in the \texttt{DATA/Par\_file}:
 \begin{description}
-\item [{\texttt{SEISMOGRAMS\_BINARY}}] set to \texttt{.true.} to have seismograms
+\item [{\texttt{USE\_BINARY\_FOR\_SEISMOGRAMS}}] set to \texttt{.true.} to have seismograms
 written out in binary format.

### 2022-11-12 doc/USER_MANUAL/03_mesh_generation.tex
updates manual (minimum period estimate)

@@
 defined below. Based upon benchmarks against semi-analytical discrete
-wavenumber synthetic seismograms \citep{KoLiTrSuStSh04}, determined
-that a $\nexxi=288$ run is accurate to a shortest period of roughly
+wavenumber synthetic seismograms, \cite{KoLiTrSuStSh04} determined
+that a $\nexxi=288$ run for a 516 km x 507 km mesh
+is accurate to a shortest period of roughly
 2~s. Therefore, since accuracy is determined by the number of grid
 points per shortest wavelength, for any particular value of $\nexxi$
-the simulation will be accurate to a shortest period determined by
+the simulation with a mesh extent $\Delta$ (in km)
+will be accurate approximately to a shortest period determined by
 \begin{equation}
-\mbox{shortest period (s)}=(288/\nexxi)\times2.\label{eq:shortest_period}
+\mbox{shortest period (s)}=2. \times \frac{288}{\nexxi} \times \frac{\Delta}{500km}\label{eq:shortest_period}
 \end{equation}
+This estimate also depends on the velocity model. Therefore, the binary \texttt{xgenerate\_databases}
+calculates a minimum period resolved by the chosen mesh discretization and velocity model,
+output into the file \texttt{OUTPUT\_FILES/output\_generate\_databases.txt}.
+
  The number of grid points in each orthogonal direction of the reference
@@
 $\mbox{\texttt{NGLLX\/}}=5$, for a total of $5^{3}=125$ points per
-elements. We suggest not to change this value.
+elements. We suggest this value as it represents a fair trade-off between element deformations to adapt to
+topographic variations and numerical accuracy.
 %

### 2022-11-12 doc/USER_MANUAL/01_introduction.tex
updates tex files (\\ to \newline) for better pandoc conversions to wiki pages

@@
 \citep{WiStBuGh10,AcKo11}. In particular, it can accurately handle
-very distorted mesh elements \citep{OlSe11}.\\
+very distorted mesh elements \citep{OlSe11}.\newline
 
@@
 approximation in the case you want to study, and/or see if you should turn gravity on.
-For more details on that see e.g. \cite{COA2011}.\\
+For more details on that see e.g. \cite{COA2011}.\newline
 
@@
 as well as mixed meshes, although with increased cost and reduced
-accuracy in these elements, as in the discontinuous Galerkin method.\\
+accuracy in these elements, as in the discontinuous Galerkin method.\newline
 
@@
 are not drastic and thus conforming mesh doubling bricks can efficiently
-handle mesh size variations \citep{KoTr02a,KoLiTrSuStSh04,LeChLiKoHuTr08,LeChKoHuTr09,LeKoHuTr09}.\\
+handle mesh size variations \citep{KoTr02a,KoLiTrSuStSh04,LeChLiKoHuTr08,LeChKoHuTr09,LeKoHuTr09}.\newline
 
@@
 A detailed theoretical analysis of the dispersion
-and stability properties of the SEM is available in \citet{Coh02}, \citet{DeSe07}, \citet{SeOl07}, \citet{SeOl08} and \citet{MeStTh12}.\\
+and stability properties of the SEM is available in \citet{Coh02}, \citet{DeSe07}, \citet{SeOl07}, \citet{SeOl08} and \citet{MeStTh12}.\newline
 
@@
 \citep{SaKoTr10}. Adjoint capabilities and finite-frequency kernel
-simulations are included \citep{TrKoLi08,PeKoLuMaLeCaLeMaLiBlNiBaTr11,LiTr06,FiIgBuKe09,ViOp09}.\\
+simulations are included \citep{TrKoLi08,PeKoLuMaLeCaLeMaLiBlNiBaTr11,LiTr06,FiIgBuKe09,ViOp09}.\newline
 
@@
 In the Legendre version that we use in SPECFEM the mass matrix is purposely slightly inexact but diagonal (but can be made exact if needed, see \cite{Teu15}),
-while in the Chebyshev version it is exact but non diagonal.\\
+while in the Chebyshev version it is exact but non diagonal.\newline
 
@@
 If that is the case, an easy option is to slightly change the mesh in the source region in order to get rid of these Hourglass-like spurious modes,
-as explained in \cite{DuLiScGa14}, in which this phenomenon is described in details, and in which practical solutions to avoid it are suggested.\\
+as explained in \cite{DuLiScGa14}, in which this phenomenon is described in details, and in which practical solutions to avoid it are suggested.\newline
 
@@
 no obsolete or obsolescent features of Fortran. The package uses parallel
-programming based upon the Message Passing Interface (MPI) \citep{GrLuSk94,Pac97}.\\
+programming based upon the Message Passing Interface (MPI) \citep{GrLuSk94,Pac97}.\newline
 
@@
 (USA) \citep{CaKoLaTiMiLeSnTr08}. It also won the BULL Joseph Fourier
-supercomputing award in 2010.\\
+supercomputing award in 2010.\newline
 
 It reached the sustained one petaflop performance level for the first time in February 2013
-on the Blue Waters Cray supercomputer at the National Center for Supercomputing Applications (NCSA), located at the University of Illinois at Urbana-Champaign (USA).\\
+on the Blue Waters Cray supercomputer at the National Center for Supercomputing Applications (NCSA), located at the University of Illinois at Urbana-Champaign (USA).\newline
 
@@
 (C-PML or ADE-PML) \citep{MaKoEz08,MaKoGe08,MaKo09,MaKoGeBr10,KoMa07}.
-It also includes support for GPU graphics card acceleration \citep{Kom11,MiKo10,KoMiEr09,KoErGoMi10}.\\
+It also includes support for GPU graphics card acceleration \citep{Kom11,MiKo10,KoMiEr09,KoErGoMi10}.\newline
 
 The next release of the code will use the PT-SCOTCH parallel and threaded
-version of SCOTCH for mesh partitioning instead of the serial version.\\
+version of SCOTCH for mesh partitioning instead of the serial version.\newline
 
@@
 For FWI inversions for the model, also refer to the new examples provided in
-directory \texttt{EXAMPLES}.\\
+directory \texttt{EXAMPLES}.\newline
 
@@
 on the outer edges of the mesh created once and for all (see \cite{MoChKoFu13,MoChKoWa15,WaChMoKoMo16,ToChKoBaLi14,ToKoTsHuChBaLi14}, and if you use that feature of the code please cite at least one of these papers).
-\\
+\newline
 
-To use coupling with FK, just use the set of parameters that is in the \texttt{DATA/Par\_file} input file of the code: \\
+To use coupling with FK, just use the set of parameters that is in the \texttt{DATA/Par\_file} input file of the code: \newline
 %
@@
 That part (coupling with FK) is actively maintained and works fine. See e.g. \href{https://doi.org/10.1002/2014GL061644}{GRL\_Ping\_Tong\_2014} for some examples.
-There is also an example that is provided with the code:\\
-\texttt{specfem3d/EXAMPLES/small\_example\_coupling\_FK\_specfem}. \\
+There is also an example that is provided with the code:\newline
+\texttt{specfem3d/EXAMPLES/small\_example\_coupling\_FK\_specfem}. \newline
 
@@
 The necessary tools are in directory \texttt{specfem3d/EXTERNAL\_PACKAGES\_coupled\_with\_SPECFEM3D/DSM\_for\_SPECFEM3D}, and there is a \texttt{README} file in \texttt{specfem3d/EXTERNAL\_PACKAGES\_coupled\_with\_SPECFEM3D} that should be more or less up-to-date (there are about four steps to follow in total, the first one being creating the database of DSM tractions and displacements on the edges of the coupling box).
-See e.g. \href{https://doi.org/10.1093/gji/ggs006}{GJI\_Vadim\_2013} and \href{https://doi.org/10.1093/gji/ggv189}{GJI\_Vadim\_2015} for some examples.\\
+See e.g. \href{https://doi.org/10.1093/gji/ggs006}{GJI\_Vadim\_2013} and \href{https://doi.org/10.1093/gji/ggv189}{GJI\_Vadim\_2015} for some examples.\newline
 
@@
 These are two different things, and both are implemented and avaible in SPECFEM3D.
-To use the second feature, please refer to \texttt{doc/how\_to/README\_gravity\_perturbation.txt}.\\
+To use the second feature, please refer to\newline
+\texttt{doc/how\_to/README\_gravity\_perturbation.txt}.\newline
 
@@
 
-You can find all the references below in \BibTeX format in file \texttt{doc/USER\_MANUAL/bibliography.bib}.\\
+You can find all the references below in \BibTeX format in file \texttt{doc/USER\_MANUAL/bibliography.bib}.\newline
 
@@
 \begin{description}
-\item [{Numerical simulations in general}] ~\\
+\item [{Numerical simulations in general}] ~\newline
  Forward and adjoint simulations are described in detail in \citet{TrKoLi08,PeKoLuMaLeCaLeMaLiBlNiBaTr11,VaCaSaKoVi99,KoMiEr09,KoErGoMi10,ChKoViCaVaFe07,MaKoDi09,KoViCh10,CaKoLaTiMiLeSnTr08,TrKoHjLiZhPeBoMcFrTrHu10,KoRiTr02,KoTr02a,KoTr02b,KoTr99}
@@
 in \citet{CaKoLaTiMiLeSnTr08,KoLaMi08a,MaKoBlLe08,KoErGoMi10,Kom11},
-\item [{GPU computing}] ~\\
+\item [{GPU computing}] ~\newline
  Computing on GPU graphics cards for acoustic or seismic wave propagation
@@
 at least one of these six articles, in which results of non blocking
-MPI runs are presented: \citet{PeKoLuMaLeCaLeMaLiBlNiBaTr11,KoErGoMi10,KoViCh10,Kom11,CaKoLaTiMiLeSnTr08,MaKoBlLe08}.\\
+MPI runs are presented: \citet{PeKoLuMaLeCaLeMaLiBlNiBaTr11,KoErGoMi10,KoViCh10,Kom11,CaKoLaTiMiLeSnTr08,MaKoBlLe08}.\newline
 
@@
 \begin{description}
-\item [{Southern California simulations}] ~\\
+\item [{Southern California simulations}] ~\newline
  \citet{KoLiTrSuStSh04,KrJiKoTr06a,KrJiKoTr06b}.
@@
 
-\item [{Anisotropy}] ~\\
+\item [{Anisotropy}] ~\newline
  \citet{ChTr07,JiTsKoTr05,ChFaKo04,FaChKo04,RiRiKoTrHe02,TrKo00}.
-\item [{Attenuation}] ~\\
+\item [{Attenuation}] ~\newline
  \citet{SaKoTr10,KoTr02a,KoTr99}.
-\item [{Topography}] ~\\
+\item [{Topography}] ~\newline
  \citet{LeKoHuTr09,LeChKoHuTr09,LeChLiKoHuTr08,GoAmTaCaSmSaMaKo09,WiKoScTr04}.

### 2022-11-12 doc/USER_MANUAL/02_getting_started.tex
updates tex files (\\ to \newline) for better pandoc conversions to wiki pages

@@
 \texttt{MPI\_INC=\${CRAY\_MPICH2\_DIR}/include and FCLIBS=" "}, and for more details if needed you can refer to the \texttt{utils/Cray\_compiler\_information} directory.
-You can also have a look at the configure script called:\\
+You can also have a look at the configure script called:\newline
 \texttt{utils/Cray\_compiler\_information/configure\_SPECFEM\_for\_Piz\_Daint.bash}.
@@
 to insert `\texttt{`TRAP\_FPE=OFF}'' into the generated \texttt{Makefile}
-in order to turn underflow trapping off.\\
+in order to turn underflow trapping off.\newline
 
@@
 For more details see \href{https://github.com/SPECFEM/specfem3d/issues/81}{github.com/SPECFEM/specfem3d/issues/81} .
-To check if that option works fine on your machine, run the code with and without it for an acoustic/elastic model and make sure the seismograms are identical.\\
+To check if that option works fine on your machine, run the code with and without it for an acoustic/elastic model and make sure the seismograms are identical.\newline
 
@@
 that the spectral-element method can very accurately handle distorted
-mesh elements \citep{OlSe11}.\\
+mesh elements \citep{OlSe11}.\newline
 
@@
 5.0, which has incompatible function calls) and edit \texttt{Makefile.in}
-and uncomment the METIS link flag in that file before running \texttt{configure}.\\
+and uncomment the METIS link flag in that file before running \texttt{configure}.\newline
 
@@
 the Zlib compression development library is probably missing on your machine and you will need to install it or ask your system administrator to
-do so. On Linux machines the package is often called "zlib1g-dev" or similar. (thus "sudo apt-get install zlib1g-dev" would install it)\\
+do so. On Linux machines the package is often called "zlib1g-dev" or similar. (thus "sudo apt-get install zlib1g-dev" would install it)\newline
 
@@
 with the \texttt{-{}-without-mpi} option to suppress all calls to
-MPI.\\
+MPI.\newline
 
-For people who would like to run the package on Windows rather than on Unix machines, you can install Docker or VirtualBox (installing a Linux in VirtualBox in that latter case) and run it easily from inside that.\\
+For people who would like to run the package on Windows rather than on Unix machines, you can install Docker or VirtualBox (installing a Linux in VirtualBox in that latter case) and run it easily from inside that.\newline
 
@@
 unlimited }} to your {\texttt{.cshrc}} file to suppress any potential
-limit to the size of the Unix stack.\\
+limit to the size of the Unix stack.\newline
 
-Beware that some cluster systems that run a recent version may not run and\slash compile an older version of the code. \\
+Beware that some cluster systems that run a recent version may not run and\slash compile an older version of the code. \newline
 
 When using dynamic fault in parallel with the developer version, we suggest you to set the configuration parameters
-of \texttt{FAULT\_DISPL\_VELOC} and \texttt{FAULT\_SYNCHRONIZE\_ACCEL} as \texttt{.true.}.\\
+of \texttt{FAULT\_DISPL\_VELOC} and \texttt{FAULT\_SYNCHRONIZE\_ACCEL} as \texttt{.true.}.\newline
 
@@
 \end{verbatim}
-The compilation with the cuda5 setting chooses then the right architecture (\texttt{-gencode=arch=compute\_35,code=sm\_35} for K20 cards).\\
+The compilation with the cuda5 setting chooses then the right architecture (\texttt{-gencode=arch=compute\_35,code=sm\_35} for K20 cards).\newline
 
@@
 be slightly slower. A possible exception could be IBM BlueGene-type
-architectures.\\
+architectures.\newline
 
@@
 \end{verbatim}
-This will add the corresponding OpenMP flag for the chosen Fortran compiler.\\
+This will add the corresponding OpenMP flag for the chosen Fortran compiler.\newline
 
@@
 difference in terms of performance. We welcome feedback on your experience
-with various compilers and flags.\\
+with various compilers and flags.\newline
 
@@
 in most cases they will), you can select single precision for your
-future runs.\\
+future runs.\newline
 
@@
 \noindent
-Installation instructions for IBM BlueGene (from April 2013):\\
+Installation instructions for IBM BlueGene (from April 2013):\newline
 
@@
 to -O2. The debug flags (-g, -qfullpath) do not influence performance
-but are useful to get at least some insights in case of problems.\\
+but are useful to get at least some insights in case of problems.\newline
 
@@
 load bgq-xl} (another, less efficient option is to load the GNU compilers
-using \texttt{module load bgq-gnu/4.4.6} or similar).\\
+using \texttt{module load bgq-gnu/4.4.6} or similar).\newline
 
@@
 
-\noindent \underline{Older installation instruction for IBM BlueGene, from 2011:}\\
+\noindent \underline{Older installation instruction for IBM BlueGene, from 2011:}\newline
 
@@
 \item Run in the terminal : \texttt{doxygen -g}, which creates a \texttt{Doxyfile} that tells doxygen what you want it to do.
-\item Edit the Doxyfile. Two Doxyfile-type files have been already committed in the directory \\
+\item Edit the Doxyfile. Two Doxyfile-type files have been already committed in the directory \newline
 \noindent \texttt{specfem3d/doc/Call\_trees}:
@@
 To do so, i.e.
-to access the development branch of the source code with read/write access (in a safe way, no need to worry too much about breaking the package, there are CI tests based on BuildBot, Travis-CI and Jenkins in place that are checking and validating all new contributions and changes), please visit this Web page:\\
+to access the development branch of the source code with read/write access (in a safe way, no need to worry too much about breaking the package, there are CI tests based on BuildBot, Travis-CI and Jenkins in place that are checking and validating all new contributions and changes), please visit this Web page:\newline
 \url{https://github.com/SPECFEM/specfem3d/wiki}

### 2022-11-12 doc/USER_MANUAL/04_creating_databases.tex
updates tex files (\\ to \newline) for better pandoc conversions to wiki pages

@@
 After using \texttt{xmeshfem3D} or \texttt{xdecompose\_mesh}, the
-next step in the workflow is to compile \texttt{xgenerate\_}~\\
+next step in the workflow is to compile \texttt{xgenerate\_}~\newline
  \texttt{databases}. This program is going to create all the missing
@@
 need to be set in the main parameter \texttt{Par\_file} located in
-the subdirectory \texttt{DATA}:\\
+the subdirectory \texttt{DATA}:\newline
 
@@
 \end{description}
-There are quite a few more parameters to change the default setup of your runs. Please check the comments in the \texttt{DATA/Par\_file} directly for further explanations.\\
+There are quite a few more parameters to change the default setup of your runs. Please check the comments in the \texttt{DATA/Par\_file} directly for further explanations.\newline
 
@@
 \texttt{xgenerate\_databases} and then use the value suggested in
-the ~\\
+the ~\newline
  \texttt{OUTPUT\_FILES/output\_mesher.txt} file after the database

### 2022-11-12 doc/USER_MANUAL/05_running_the_solver.tex
updates tex files (\\ to \newline) for better pandoc conversions to wiki pages

@@
 Please note that \texttt{xspecfem3D} must be called directly from
-the main directory, as most of the binaries of the package.\\
+the main directory, as most of the binaries of the package.\newline
 
@@
 \item the absorbing boundary control parameter \texttt{PML\_CONDITIONS}
-on condition that the\\
+on condition that the\newline
  \texttt{PML\_INSTEAD\_OF\_FREE\_SURFACE} flag remains unmodified
@@
 \item the ShakeMap\textregistered{} option \texttt{CREATE\_SHAKEMAP}
-\item the output information parameters \texttt{MOVIE\_TYPE}, \texttt{NTSTEP\_BETWEEN\_OUTPUT\_INFO} and\\
+\item the output information parameters \texttt{MOVIE\_TYPE}, \texttt{NTSTEP\_BETWEEN\_OUTPUT\_INFO} and\newline
 \texttt{NTSTEP\_BETWEEN\_OUTPUT\_SEISMOS}
@@
 the database generator \texttt{xgenerate\_databases} and the solver
-\texttt{xspecfem3D}.\\
+\texttt{xspecfem3D}.\newline
 
@@
 Elevation and burial are generally applicable to geographical regions.
-Burial is measured down from the top surface.\\
+Burial is measured down from the top surface.\newline
 
@@
 \begin{enumerate}
-\item [-] Put the origin on the top of the model.\\
+\item [-] Put the origin on the top of the model.\newline
 \item [-] Let's say you want to place two receivers at (x1,y1,z1) and
@@
 the depth is given in kilometers; however for users in other fields (non-destructive testing, medical imaging, near-surface studies...) who may prefer to give the position of the source (rather than its depth from the surface), or for people who use FORCESOLUTION to describe the source rather than CMTSOLUTION, we provide an option called USE\_SOURCES\_RECEIVERS\_Z in the \texttt{DATA/Par\_file}, and if so that position is read from CMTSOLUTION in meters rather than kilometers (and again, it is then the true position in the mesh, not the depth).
-When option USE\_SOURCES\_RECEIVERS\_Z in the \texttt{DATA/Par\_file} is on, this remark applies to the position of the receivers as well.\\
+When option USE\_SOURCES\_RECEIVERS\_Z in the \texttt{DATA/Par\_file} is on, this remark applies to the position of the receivers as well.\newline
 \item [-] Let's say you want to place two receivers at (x1,y1,z1) and
@@
 burial as the $z$ coordinate.
-Third column is Y and Fourth is X due to the latitude/longitude convention.\\
+Third column is Y and Fourth is X due to the latitude/longitude convention.\newline
 \end{enumerate}
@@
 And you can always plot OUTPUT\_FILES/sr.vtk file in ParaView to check the
-source/receiver locations after your simulation.\\
+source/receiver locations after your simulation.\newline
 \end{itemize}
@@
 \begin{itemize}
-\item The simulation is launched within the root directory \texttt{EXAMPLE\_ROOT\_DIR}\\
+\item The simulation is launched within the root directory \texttt{EXAMPLE\_ROOT\_DIR}\newline
 (usually \texttt{mpirun -np N ./bin/xspecfem3D}).
@@
 The model used is a constant $Q$, thus with no dependence on frequency ($Q(f)$ = constant).
-See e.g. \cite{BlKoChLoXi16}. \\
+See e.g. \cite{BlKoChLoXi16}. \newline
 
@@
 and thus the code outputs the band in which the approximation is very good, outside of that range it can be less accurate.
-The logarithmic center of that frequency band is the \texttt{ATTENUATION\_f0} parameter defined (in Hz) in input file \texttt{DATA/Par\_file}. \\
+The logarithmic center of that frequency band is the \texttt{ATTENUATION\_f0} parameter defined (in Hz) in input file \texttt{DATA/Par\_file}. \newline
 

### 2022-11-12 doc/USER_MANUAL/06_fault_sources.tex
updates tex files (\\ to \newline) for better pandoc conversions to wiki pages

@@
 Currently faults can only be run with \texttt{xdecompose\_mesh} and
-\texttt{CUBIT}. \texttt{xmeshfem3D} is not yet ready to handle faults.\\
+\texttt{CUBIT}. \texttt{xmeshfem3D} is not yet ready to handle faults.\newline
 
@@
 the edge on the free surface). The fault is automatically closed later
-by SPECFEM3D.\\
+by SPECFEM3D.\newline
 
@@
 \begin{centering}
-\includegraphics[scale=0.3]{figures/faultmesh.jpg} \\
+\includegraphics[scale=0.3]{figures/faultmesh.jpg} \newline
 \includegraphics[scale=0.3]{figures/surf168.jpg}
@@
 like the example above. Variables can be defined and manipulated there
-using the $\mathtt{APREPRO}$ language built in CUBIT.\\
+using the $\mathtt{APREPRO}$ language built in CUBIT.\newline
 
@@
 \begin{itemize}
-\item Splay fault models from Wendt et al. (2009)\\
+\item Splay fault models from Wendt et al. (2009)\newline
 
@@
 is the horizontal along-strike vector such that the system is right-handed. In places where the fault plane is horizontal,
-we define the alone strike direction to be (1,0,0).\\
+we define the alone strike direction to be (1,0,0).\newline
 
@@
 1 is on the hanging wall (normal faulting if side 1 is on the foot
-wall).\\
+wall).\newline
 
@@
 $\mathtt{FAULT\_STATIONS}$ and $\mathtt{input\_file.txt}$ or $\mathtt{rsf\_hete\_input\_file.txt}$(optional).
-If the former does not exist, the code assumes that there are no faults.\\
+If the former does not exist, the code assumes that there are no faults.\newline
 
@@
 in a Fortran 9x manual). The input for each fault has the following
-sequence (arguments in {[}brackets{]} are optional):\\
+sequence (arguments in {[}brackets{]} are optional):\newline
 
-\&\textbf{RUPTURE\_SWITCHES}  / RATE\_AND\_STATE ,TPV16 ,HETE\_RSF ,TPV10X ,TWF\\
-\&\textbf{BEGIN\_FAULT} /\\
-\&\textbf{STRESS\_TENSOR} Sigma = $\sigma_{xx}$,$\sigma_{yy}$,$\sigma_{zz}$,$\sigma_{xy}$,$\sigma_{xz}$,$\sigma_{yz}$ /\\
-\&\textbf{INIT\_STRESS} S1, S2, S3 {[},n1, n2, n3{]} /\\
+\&\textbf{RUPTURE\_SWITCHES}  / RATE\_AND\_STATE ,TPV16 ,HETE\_RSF ,TPV10X ,TWF\newline
+\&\textbf{BEGIN\_FAULT} /\newline
+\&\textbf{STRESS\_TENSOR} Sigma = $\sigma_{xx}$,$\sigma_{yy}$,$\sigma_{zz}$,$\sigma_{xy}$,$\sigma_{xz}$,$\sigma_{yz}$ /\newline
+\&\textbf{INIT\_STRESS} S1, S2, S3 {[},n1, n2, n3{]} /\newline
 
-followed by (n1+n2+n3) \&\textbf{DIST2D} blocks\\
+followed by (n1+n2+n3) \&\textbf{DIST2D} blocks\newline
 
-\&\textbf{SWF} mus, mud, dc {[}, nmus, nmud, ndc{]} (weakening\_kind=1 for linear (default); weakening\_kind=2 for exponential) /\\
-\&\textbf{TWF} nuc\_x, nuc\_y, nuc\_z, nuc\_r, nuc\_t0, nuc\_v /\\
-\&\textbf{RSF} V0,f0,a,b,L,V\_init,theta\_init,C,StateLaw {[} nV0,nf0,na,nb,nL,nV\_init,ntheta\_init,nC {]} /\\
+\&\textbf{SWF} mus, mud, dc {[}, nmus, nmud, ndc{]} (weakening\_kind=1 for linear (default); weakening\_kind=2 for exponential) /\newline
+\&\textbf{TWF} nuc\_x, nuc\_y, nuc\_z, nuc\_r, nuc\_t0, nuc\_v /\newline
+\&\textbf{RSF} V0,f0,a,b,L,V\_init,theta\_init,C,StateLaw {[} nV0,nf0,na,nb,nL,nV\_init,ntheta\_init,nC {]} /\newline
 
-followed by (nV0+nf0+na+nb+nL+nV\_init+ntheta\_init+nC) \&\textbf{DIST2D} blocks\\
+followed by (nV0+nf0+na+nb+nL+nV\_init+ntheta\_init+nC) \&\textbf{DIST2D} blocks\newline
 
@@
 x, y, z, val. Method of interpolation is nearest.
-\\
+\newline
 
@@
 \item Use the Matlab function $\mathtt{utils/critical\_timestep.m}$ to
-compute\\
+compute\newline
  $\mathtt{dtc\_fault}=\mathtt{critical\_timestep\left(c_{p},h\_fault,ngll\right)}$.
-\\
+\newline
 This is the critical time step in an elastic medium for a hypothetical
@@
 \item Reset the timestep of the simulation with a Kelvin-Voigt material
-to a value smaller than\\
+to a value smaller than\newline
  $\mathtt{dtc\_kv}=\mathtt{eta}\left(\sqrt{1+\mathtt{dtc\_bulk^{2}}/\mathtt{eta^{2}}}-1\right)$
@@
 
-Several output files are saved in \texttt{OUTPUT\_FILES/}:\\
+Several output files are saved in \texttt{OUTPUT\_FILES/}:\newline
 
@@
 The files are ascii and start with a header (22 lines long) followed
-by a data block with the following format, one line per time sample:\\
+by a data block with the following format, one line per time sample:\newline
 
@@
 
-Column 8 = normal stress (MPa)\\
+Column 8 = normal stress (MPa)\newline
 
@@
 in directory $\mathtt{utils}$. The functions are internally documented
-(see their matlab help).\\
+(see their matlab help).\newline
 
-\texttt{FSEM3D\_snapshot} reads a fault data snapshot\\
+\texttt{FSEM3D\_snapshot} reads a fault data snapshot\newline
 

### 2022-11-12 doc/USER_MANUAL/07_adjoint_simulations.tex
updates tex files (\\ to \newline) for better pandoc conversions to wiki pages

@@
 \end{enumerate}
-\item \textbf{Set the related parameters and run the adjoint simulation}\\
+\item \textbf{Set the related parameters and run the adjoint simulation}\newline
 In the \texttt{DATA/Par\_file}, set the two related parameters to
@@
 strain tensor (\texttt{SNN,SEE,SZZ,SNE,SNZ,SEZ}) at these locations,
-and ~\\
+and ~\newline
  \texttt{NT.S?????.BX?.sem} for the three-component displacements

### 2022-11-12 doc/USER_MANUAL/08_doing_tomography.tex
switched the license to GNU version 3 (now that we use pyCMT3D, which is LGPL v3, the combination automatically becomes GNU version 3). Also added FWI inversion information to the users manual.
updates tex files (\\ to \newline) for better pandoc conversions to wiki pages

@@
 \chapter[Doing tomography based on the sensitivity kernels obtained]{Doing tomography, i.e., updating the model based on the sensitivity kernels obtained}\label{cha:tomo}
+
+\red{\Large ANNOUNCEMENT (June 2017): See also the next chapter (Chapter~\ref{cha:FWI}) about how to perform full waveform inversion (FWI) or source inversions.}\newline
 
@@
 You can modify specific settings affecting, e.g., the (isotropic) parameterization or directory setup structure in file \texttt{setup/constants\_tomography.h}.
-The compiled binaries can be found in directory \texttt{bin/}.\\
+The compiled binaries can be found in directory \texttt{bin/}.\newline
 
@@
 Also note that currently only the binary database format is supported, no ADIOS file format support is implemented yet.
-In future, we plan to extend and merge these into the same set of tools within a common SPECFEM function library.\\
+In future, we plan to extend and merge these into the same set of tools within a common SPECFEM function library.\newline
 
@@
 \item [-]{\bf Output:}
-The summed kernels will be stored in corresponding kernel files in output directory \texttt{OUTPUT\_SUM/}.\\
+The summed kernels will be stored in corresponding kernel files in output directory \texttt{OUTPUT\_SUM/}.\newline
 \end{itemize}
@@
 \begin{itemize}
-\item [-] for an {\it isotropic} model parameterization, kernel names are:\\
+\item [-] for an {\it isotropic} model parameterization, kernel names are:\newline
 \begin{tabular}{ c c }
@@
 \end{minipage}
-\end{tabular}\\
+\end{tabular}\newline
 
-\item [-] for an {\it isotropic bulk} model parameterization, kernel names are:\\
+\item [-] for an {\it isotropic bulk} model parameterization, kernel names are:\newline
 \begin{tabular}{ c c }
@@
 \end{minipage}
-\end{tabular}\\
+\end{tabular}\newline
 
-\item [-] for a {\it transversely isotropic} model parameterization, kernel names are:\\
+\item [-] for a {\it transversely isotropic} model parameterization, kernel names are:\newline
 \begin{tabular}{ c c }
@@
 \end{minipage}
-\end{tabular}\\
+\end{tabular}\newline
 
@@
 To create transversely isotropic kernels, you set \texttt{ANISOTROPIC\_KL} and \texttt{SAVE\_TRANSVERSE\_KL} to \texttt{.true.} in the parameter file \texttt{DATA/Par\_file}.
-\\
+\newline
 
-\item [- ] for preconditioning, the approximate Hessian kernels taken for summation are:\\
+\item [- ] for preconditioning, the approximate Hessian kernels taken for summation are:\newline
 \begin{tabular}{ c c }
@@
 \end{minipage}
-\end{tabular}\\
+\end{tabular}\newline
 
 These Hessian kernels approximate the diagonal elements of the Hessian matrix and can be used as preconditioner in a gradient optimization scheme.
-To create these kernels, you have to set \texttt{APPROXIMATE\_HESS\_KL} to \texttt{.true.} in the parameter file \texttt{DATA/Par\_file}.\\
+To create these kernels, you have to set \texttt{APPROXIMATE\_HESS\_KL} to \texttt{.true.} in the parameter file \texttt{DATA/Par\_file}.\newline
 
@@
 You can change the default by setting parameter \texttt{REG} in file \texttt{constants\_tomography.h} to the region of interest.
-\\
+\newline
 
@@
 Note that instead of using the density kernels $K_{\rho}$ for model updates, setting \texttt{USE\_RHO\_SCALING} to \texttt{.true.} will
-ignore the density kernel and use density perturbations scaled from isotropic $V_s$ pertubations.\\
+ignore the density kernel and use density perturbations scaled from isotropic $V_s$ pertubations.\newline
 

### 2022-11-12 doc/USER_MANUAL/09_noise_simulations.tex
updates tex files (\\ to \newline) for better pandoc conversions to wiki pages

@@
 \texttt{CMTSOLUTION} and \texttt{STATIONS}. Unless otherwise specified,
-those input files should be located in directory \texttt{DATA/}.\\
+those input files should be located in directory \texttt{DATA/}.\newline
 
@@
 In other words, the moment tensor specified in \texttt{CMTSOLUTION}
-must be set to zero manually!\\
+must be set to zero manually!\newline
 
@@
 now important. The order will be used to determine the `main' receiver,
-i.e., the one that simultaneously cross correlates with the others.\\
+i.e., the one that simultaneously cross correlates with the others.\newline
 
@@
 Should you be confused by the three steps, refer to \citet{trompetal2010}
-for details.\\
+for details.\newline
 
@@
 The code automatically makes the modification accordingly, if \texttt{NOISE\_TOMOGRAPHY}
-is not zero.\\
+is not zero.\newline
 
@@
 2. The third step is for kernel constructions. Hence \texttt{SIMULATION\_TYPE}
-should be 3, whereas \texttt{SAVE\_FORWARD} must be \texttt{.false.}.\\
+should be 3, whereas \texttt{SAVE\_FORWARD} must be \texttt{.false.}.\newline
 
@@
 ask your cluster administrator for information/configuration of the
-cluster you are using.\\
+cluster you are using.\newline
 
@@
 have done for regular earthquake simulations. Refer to other chapters
-if you have problems.\\
+if you have problems.\newline
 
@@
 ensemble cross correlations and one contribution of the ensemble sensitivity
-kernels.\\
+kernels.\newline
 
@@
 we may obtain cross correlations for different components as well,
-not necessarily the cross correlations between vertical components.\\
+not necessarily the cross correlations between vertical components.\newline
 
@@
 on local nodes. The ensemble kernel files are named the same as regular
-earthquake kernels.\\
+earthquake kernels.\newline
 
@@
 is provided in \texttt{EXAMPLES/noise\_tomography/}. See \texttt{EXAMPLES/noise\_tomography/README}
-for explanations. \\
+for explanations. \newline
 
 Note, however, that they are created for a specific workstation (CLOVER@PRINCETON),
-which has at least 4 cores with `mpif90' working properly. \\
+which has at least 4 cores with `mpif90' working properly. \newline
 
 If your workstation is suitable, you can run the example in \texttt{EXAMPLES/noise\_tomography/}
-using:\\
+using:\newline
 
-\texttt{./pre-processing.sh}\\
+\texttt{./pre-processing.sh}\newline
 

### 2022-11-12 doc/USER_MANUAL/10_gravity_calculations.tex
updates tex files (\\ to \newline) for better pandoc conversions to wiki pages

@@
 generated by any given 3D Earth model at a given elevation, i.e., on a given observation surface.
-That feature is still experimental but should work just fine.\\
+That feature is still experimental but should work just fine.\newline
 
-To see how it is implemented and to use it, type this in the root directory of the code:\\
-
-\texttt{grep -i GRAVITY\_INTEGRALS src/*/*}
+To see how it is implemented and to use it, type this in the root directory of the code:
+\begin{verbatim}
+grep -i GRAVITY_INTEGRALS src/*/*
+\end{verbatim}

### 2022-11-12 doc/USER_MANUAL/13_changing_the_model.tex
updates tex files (\\ to \newline) for better pandoc conversions to wiki pages

@@
 your own external tomography file(s), and choose between two possible
-options:\\
+options:\newline
  \indent (1) set in the \texttt{DATA/Par\_file} the model parameter \texttt{MODEL
-= tomo}, \\
- or for more user control: \\
+= tomo}, \newline
+ or for more user control: \newline
  \indent (2) set in the \texttt{DATA/Par\_file} the model parameter \texttt{MODEL
@@
 \end{verbatim}
-where: \\
+where: \newline
  \indent \texttt{domain\_ID} is 1 for acoustic or 2 for elastic materials,
-\\
+\newline
  \indent \texttt{material\_ID} a negative, unique identifier (i.e.,
--1,-2,...), \\
+-1,-2,...), \newline
  \indent \texttt{tomography} keyword for tomographic material definition,
-\\
+\newline
  \indent \texttt{elastic} keyword for elastic material definition,
-\\
+\newline
  \indent \texttt{file\_name} the name of the tomography file and
-1 a positive unique identifier.\\
+1 a positive unique identifier.\newline
 

### 2022-11-12 doc/USER_MANUAL/14_post_processing.tex
updates tex files (\\ to \newline) for better pandoc conversions to wiki pages

@@
 In order to convert between SAC format and ASCII files, useful scripts
-are provided in the subdirectories ~\\
+are provided in the subdirectories ~\newline
  \texttt{utils/sac2000\_alpha\_convert/} and \texttt{utils/seis\_process/asc2sac/}.
@@
 With the shakemap data saved in \texttt{OUTPUT\_FILES/} at the end
-of a shakemap simulation ~\\
+of a shakemap simulation ~\newline
  (\texttt{CREATE\_SHAKEMAP=.true.}), you can also run \texttt{`create\_movie\_shakemap\_AVS\_DX\_GMT}'

### 2022-11-12 doc/USER_MANUAL/authors.tex
updates tex files (\\ to \newline) for better pandoc conversions to wiki pages

@@
-\section*{https://specfem3d.readthedocs.io/en/latest/authors/}
+\section*{Authors}
 
@@
 %%%%%%%% please do NOT merge the lines of the paragraph below, it makes it difficult to insert new names in alphabetical order; leave a single name per line
-Nikhil Vaidyar
+Since then it has been developed and maintained by a development team: in alphabetical order,
+Michael Afanasiev,
+Jean-Paul (Pablo) Ampuero,
+\'Etienne Bachmann,
+Kangchen Bai,
+Piero Basini,
+C\'eline Blitz,
+Alexis Bottero,
+Ebru Bozda\u{g},
+Emanuele Casarotti,
+Joseph Charles,
+Min Chen,
+Paul Cristini,
+Cl\'ement Durochat,
+Percy Galvez,
+Hom Nath Gharti,
+Dominik G\"oddeke,
+Vala Hj\"orleifsd\'ottir,
+Sue Kientz,
+Dimitri Komatitsch,
+Jes\'us Labarta,
+Nicolas Le Goff,
+Pieyre Le Loher,
+Matthieu Lefebvre,
+Qinya Liu,
+Youshan Liu,
+David Luet,
+Yang Luo,
+Alessia Maggi,
+Federica Magnoni,
+Roland Martin,
+Ren\'e Matzen,
+Dennis McRitchie,
+Matthias Meschede,
+Peter Messmer,
+David Mich\'ea,
+Vadim Monteiller,
+Surendra Nadh Somala,
+Masaru Nagaso,
+Tarje Nissen-Meyer,
+Daniel Peter,
+Kevin Pouget,
+Max Rietmann,
+Elliott Sales de Andrade,
+Brian Savage,
+Bernhard Schuberth,
+Anne Sieminski,
+James Smith,
+Leif Strand,
+Carl Tape,
+Jeroen Tromp,
+Brice Videau,
+Jean-Pierre Vilotte,
+Zhinan Xie,
+Chang-Hua Zhang,
+Hejun Zhu.\newline
 
@@
 Caltech's Center for Advanced Computing Research (CACR), USA, with free satellite clipart pictures
-from \url{http://www.4vector.com} and \url{http://www.clker.com} added to it.\\
+from \url{http://www.4vector.com} and \url{http://www.clker.com} added to it.\newline
 
-The code is released open-source under the GNU version 3 license, see the license at the end of this manual.\\
+The code is released open-source under the GNU version 3 license, see the license at the end of this manual.\newline

### 2022-11-12 doc/USER_MANUAL/notes_and_acknowledgement.tex
updates tex files (\\ to \newline) for better pandoc conversions to wiki pages

--- 2015-04-24 doc/USER_MANUAL/notes_and_acknowledgement.tex
@@
 is a trademark of Advanced Visualization Systems, and ParaView \urlwithparentheses{http://www.paraview.org}
-is an open-source visualization platform.
+is an open-source visualization platform.\newline
 
-Please e-mail your feedback, questions, comments, and suggestions
-to the CIG Computational Seismology Mailing List \urlwithparentheses{cig-seismo@geodynamics.org}.
+Please provide your feedback, questions, comments, and suggestions through the github repository.

### 2022-11-12 doc/USER_MANUAL/03_mesh_generation.tex

@@
 can be compared for validation against the solutions provided in subdirectory
-~\\
+~\newline
  \texttt{VALIDATION\_3D\_SEM\_SIMPLER\_LAYER\_SOURCE\_DEPTH/}.
@@
 \includegraphics[width=0.45\textwidth]{figures/example-homogeneous.jpg}
-\includegraphics[width=0.45\textwidth]{figures/example-2layers.jpg} \\
+\includegraphics[width=0.45\textwidth]{figures/example-2layers.jpg} \newline
 \includegraphics[width=0.45\textwidth]{figures/example-water.jpg}
@@
 create the \texttt{absorbing\_cpml\_file} for
-you. That program is: \\
-\texttt{utils/CPML/convert\_external\_layers\_of\_a\_given\_mesh\_to\_CPML\_layers.f90},\\
+you. That program is: \newline
+\texttt{utils/CPML/convert\_external\_layers\_of\_a\_given\_mesh\_to\_CPML\_layers.f90},\newline
 and a small Makefile is provided in that directory (utils/CPML).
@@
 
-To convert one to the other see \texttt{doc/note\_on\_Qkappa\_versus\_Qp.pdf} and the helper code\\
+To convert one to the other see \texttt{doc/note\_on\_Qkappa\_versus\_Qp.pdf} and the helper code\newline
 \texttt{conversion\_from\_Qkappa\_Qmu\_to\_Qp\_Qs\_from\_Dahlen\_Tromp\_959\_960.f90} in folder
@@
 \item [{mesh\_file}] Contains the mesh element connectivity. The hexahedral
-elements can have 8 or 27 nodes.\\
+elements can have 8 or 27 nodes.\newline
  See picture \texttt{doc/mesh\_numbering\_convention/numbering\_convention\_27\_nodes.jpg}
-to see\\
+to see\newline
  in which (standard) order the points must be cited. In the case of
@@
 \item [{free\_or\_absorbing\_surface\_file\_zmax}] Contains the free surface
-connectivity or \\
+connectivity or \newline
  the surface connectivity of the absorbing boundary surface at the
-top (Zmax), \\
+top (Zmax), \newline
  depending on whether the top surface is defined as free or absorbing
-(\texttt{STACEY\_INSTEAD\_OF\_FREE\_SURFACE} in \texttt{DATA/Par\_file}). \\
+(\texttt{STACEY\_INSTEAD\_OF\_FREE\_SURFACE} in \texttt{DATA/Par\_file}). \newline
  You should put both the surface of acoustic regions and of elastic
@@
 \item [{absorbing\_surface\_file\_xmax}] Contains the surface connectivity
-of the absorbing boundary surface at Xmax \\
+of the absorbing boundary surface at Xmax \newline
 (also needed in the case of C-PML absorbing conditions, in order for
@@
 \item [{absorbing\_surface\_file\_xmin}] Contains the surface connectivity
-of the absorbing boundary surface at Xmin \\
+of the absorbing boundary surface at Xmin \newline
 (also needed in the case of C-PML absorbing conditions, in order for
@@
 \item [{absorbing\_surface\_file\_ymax}] Contains the surface connectivity
-of the absorbing boundary surface at Ymax \\
+of the absorbing boundary surface at Ymax \newline
 (also needed in the case of C-PML absorbing conditions, in order for
@@
 \item [{absorbing\_surface\_file\_ymin}] Contains the surface connectivity
-of the absorbing boundary surface at Ymin \\
+of the absorbing boundary surface at Ymin \newline
 (also needed in the case of C-PML absorbing conditions, in order for
@@
 \item [{absorbing\_surface\_file\_bottom}] Contains the surface connectivity
-of the absorbing boundary surface at the bottom (Zmin) \\
+of the absorbing boundary surface at the bottom (Zmin) \newline
 (also needed in the case of C-PML absorbing conditions, in order for
@@
 The quality of the mesh may be inspected more precisely based upon
-the serial code in the file \texttt{check\_mesh\_quality\_}~\\
+the serial code in the file \texttt{check\_mesh\_quality\_}~\newline
  \texttt{CUBIT\_Abaqus.f90} located in the directory \texttt{src/check\_mesh\_quality\_CUBIT\_Abaqus/}.
@@
 Pellegrini et al. from LaBRI and INRIA in Bordeaux, France, downloadable
-from the web page \url{https://gitlab.inria.fr/scotch/scotch}.\\
+from the web page \url{https://gitlab.inria.fr/scotch/scotch}.\newline
 
@@
 script. In the future you should be able to find more recent versions
-at \url{http://www.labri.fr/perso/pelegrin/scotch/scotch_en.html}.\\
+at \url{http://www.labri.fr/perso/pelegrin/scotch/scotch_en.html}.\newline
 
@@
 ready to compile the internal mesher. This is an alternative to CUBIT
-for the mesh generation of relatively simple models.\\
+for the mesh generation of relatively simple models.\newline
 
@@
 note that \texttt{xmeshfem3D} must be called directly from the main
-directory, as most of the binaries of the package.\\
+directory, as most of the binaries of the package.\newline
 
@@
 of how the SEM is implemented, and we encourage you to read \citet{KoVi98,KoTr99}
-and \citet{KoLiTrSuStSh04}.\\
+and \citet{KoLiTrSuStSh04}.\newline
 
@@
 in which case all `longitude' parameters simply refer to the $x$~axis, and
-all `latitude' parameters simply refer to the $y$~axis.\\
+all `latitude' parameters simply refer to the $y$~axis.\newline
 
@@
 
-From \url{http://en.wikipedia.org/wiki/Universal_Transverse_Mercator_coordinate_system}:\\
+From \url{http://en.wikipedia.org/wiki/Universal_Transverse_Mercator_coordinate_system}:\newline
 The Universal Transverse Mercator coordinate system was developed by the United States Army Corps of Engineers in the 1940s.
@@
 For poroelastic materials, you would specify the corresponding poroelastic properties.
-Please read the section below for more on poroelastic parameters.\\
+Please read the section below for more on poroelastic parameters.\newline
 
@@
 elements in the vertical direction for each layer. We provide a few
-models in the {\texttt{EXAMPLES/}} directory.\\
+models in the {\texttt{EXAMPLES/}} directory.\newline
 
 Now that you have set the appropriate parameters in the \texttt{Mesh\_Par\_file}
-and have compiled the mesher, you are ready to run it!\\
+and have compiled the mesher, you are ready to run it!\newline
 
@@
 for information about running the code on a system with a scheduler,
-e.g., LSF.\\
+e.g., LSF.\newline
 
@@
 in \texttt{output\_mesher.txt}; this file provides lots of details
-about the mesh that was generated.\\
+about the mesh that was generated.\newline
 
@@
 At the moment, the code implements Biot's equation.
-However, the code cannot calculate partially saturated cases in its current state.\\
+However, the code cannot calculate partially saturated cases in its current state.\newline
 
@@
 \begin{description}
-\item where\\
-\texttt{rho\_s} = solid density,\\
-\texttt{rho\_f} = fluid density,\\
-\texttt{phi} = porosity,\\
-\texttt{tort} = tortuosity,\\
-\texttt{kxx} = xx component of permeability tensor,\\
-\texttt{kxy} = xy,yx components of permeability tensor,\\
-\texttt{kyy} = yy component of permeability tensor,\\
-\texttt{kxz} = xz,zx components of permeability tensor,\\
-\texttt{kzz} = zz component of permeability tensor,\\
-\texttt{kappa\_s} = solid bulk modulus,\\
-\texttt{kappa\_f} = fluid bulk modulus,\\
-\texttt{kappa\_fr} = frame bulk modulus,\\
-\texttt{eta\_f} = fluid viscosity,\\
-\texttt{mu\_fr} = frame shear modulus.\\
+\item where\newline
+\texttt{rho\_s} = solid density,\newline
+\texttt{rho\_f} = fluid density,\newline
+\texttt{phi} = porosity,\newline
+\texttt{tort} = tortuosity,\newline
+\texttt{kxx} = xx component of permeability tensor,\newline
+\texttt{kxy} = xy,yx components of permeability tensor,\newline
+\texttt{kyy} = yy component of permeability tensor,\newline
+\texttt{kxz} = xz,zx components of permeability tensor,\newline
+\texttt{kzz} = zz component of permeability tensor,\newline
+\texttt{kappa\_s} = solid bulk modulus,\newline
+\texttt{kappa\_f} = fluid bulk modulus,\newline
+\texttt{kappa\_fr} = frame bulk modulus,\newline
+\texttt{eta\_f} = fluid viscosity,\newline
+\texttt{mu\_fr} = frame shear modulus.\newline
 \end{description}
-Using an external mesh (for instance coming from CUBIT/TRELIS), poroelastic materials have the ID number 3, while 1 is acoustic and 2 is elastic (see the example in the package: EXAMPLES/homogeneous\_poroelastic).\\
+Using an external mesh (for instance coming from CUBIT/TRELIS), poroelastic materials have the ID number 3, while 1 is acoustic and 2 is elastic (see the example in the package: EXAMPLES/homogeneous\_poroelastic).\newline
 
@@
 In case the internal mesher \texttt{xmeshfem3D} is used, the same poroelastic material properties will be defined in the \texttt{Mesh\_Par\_file}.
-See the above section~\ref{cha:Running-the-Mesher-Meshfem3D} on how the corresponding line format would look like.\\
+See the above section~\ref{cha:Running-the-Mesher-Meshfem3D} on how the corresponding line format would look like.\newline
 

### 2022-11-12 doc/USER_MANUAL/15_informations_for_developers.tex

@@
 
-\noindent We also have simple 3D demo source codes that implement the SEM in a single, small program, in directory\\
+\noindent We also have simple 3D demo source codes that implement the SEM in a single, small program, in directory\newline
 utils/small\_SEM\_solvers\_in\_Fortran\_and\_C\_without\_MPI\_to\_learn of the specfem3d package.
@@
 Also useful to test new ideas by modifying these simple codes to run some tests.
-We also have a similar, even simpler, demo source code for the 2D case in directory\\
-utils/small\_SEM\_solver\_in\_Fortran\_without\_MPI\_to\_learn of the specfem2d package.\\
+We also have a similar, even simpler, demo source code for the 2D case in directory\newline
+utils/small\_SEM\_solver\_in\_Fortran\_without\_MPI\_to\_learn of the specfem2d package.\newline
 

### 2022-11-12 doc/USER_MANUAL/copyright_and_version.tex
updated headers

@@
 
-Princeton University, USA, and CNRS / University of Marseille, France\\
-$\copyright$ Princeton University and CNRS / University of Marseille, July 2012\\
+CNRS, France and Princeton University, USA\newline
+$\copyright$ October 2017\newline
 
@@
 it under the terms of the GNU General Public License as published
-by the Free Software Foundation (see Appendix \ref{cha:License}).\\
+by the Free Software Foundation (see Appendix \ref{cha:License}).\newline
 
@@
 Please note that by contributing to this code, the developer understands and agrees that this project and contribution
-are public and fall under the open source license mentioned above.\\
+are public and fall under the open source license mentioned above.\newline
 
 \noindent
-\textbf{\underline{Evolution of the code:}}\\
+\textbf{\underline{Evolution of the code:}}\newline
 
@@
 improved CUDA GPUs performance, additonal GEOCUBIT support, better make compilation,
-git versioning system. \\
+git versioning system. \newline
 
@@
 Komatitsch, Joseph Charles, Zhinan Xie: support for CUDA GPUs, better
-CFL stability for the Stacey absorbing conditions. \\
+CFL stability for the Stacey absorbing conditions. \newline
 
@@
 decomposed by SCOTCH; much faster solver using Michel Deville's inlined
-matrix products.\\
+matrix products.\newline
 
@@
 and better I/Os on very large systems, many small improvements and
-bug fixes.\\
+bug fixes.\newline
 
@@
 July 2005: serial version, regular mesh, adjoint and kernel calculations,
-ParaView support.\\
+ParaView support.\newline
 
 MPI v. 1.2 Min Chen and Dimitri Komatitsch, Caltech, July 2004: full
-anisotropy, volume movie.\\
+anisotropy, volume movie.\newline
 
@@
 scaling of $V_{s}$ with depth, Hauksson's regional model, attenuation,
-oceans, movies.\\
+oceans, movies.\newline
 
 MPI v. 1.0 Dimitri Komatitsch, Caltech, USA, May 2002: first MPI version
-based on global code.\\
+based on global code.\newline
 
@@
 for the CM-5 Connection Machine, parallelized on 128 processors using
-Connection Machine Fortran.\\
+Connection Machine Fortran.\newline

### 2022-11-21 doc/USER_MANUAL/06_fault_sources.tex
updates doc (newlines for better conversion with pandoc)

@@
 with a header (12 lines long) followed by a data block with the following
-format, one line per fault node:
+format, one line per fault node:\newline
 
-Column 1 = horizontal coordinate, distance along strike (m)
+Column 1 = horizontal coordinate, distance along strike (m)\newline
 
-Column 2 = vertical coordinate, distance down-dip (m)
+Column 2 = vertical coordinate, distance down-dip (m)\newline
 
-Column 3 = rupture time (s)
+Column 3 = rupture time (s)\newline
 

### 2022-11-21 doc/USER_MANUAL/07_adjoint_simulations.tex
updates doc (newlines for better conversion with pandoc)

@@
 when combining the kernel files from slices into one mesh file (see
-Section~\ref{sec:Finite-Frequency-Kernels}).
+Section~\ref{sec:Finite-Frequency-Kernels}).\newline
 
@@
 horizontal and vertical P and S velocities, $\alpha_{v},\alpha_{h},\beta_{v},\beta_{h}$.
-Explicit relations can be found in appendix B. of \citet{SiLiTrTr07b}
-
-The anisotropic kernels are only currently available for CPU mode.
+Explicit relations can be found in appendix B. of \citet{SiLiTrTr07b}\newline
 
 \end{enumerate}
+
 In general, the three steps need to be run sequentially to assure

### 2022-11-21 doc/USER_MANUAL/14_post_processing.tex
updates doc (newlines for better conversion with pandoc)

@@
 and the synthetics need to be processed before a comparison can be
-made.
+made.\newline
 
@@
 picked the first P and S arrivals, and added suffix `\texttt{bp}'
-to the file names.
+to the file names.\newline
 
@@
 same filter used for the observed data, pick the first P and S arrivals,
-and add the suffix `\texttt{bp}' to the file names.
+and add the suffix `\texttt{bp}' to the file names.\newline
 
@@
 Type \texttt{process\_syn.pl} without any argument for a detailed
-usage.
+usage.\newline
 
@@
 and \texttt{BX?} stands for the component name. Please see the Appendix~\ref{cha:Coordinates}
-and \ref{cha:channel-codes} for further details.
+and \ref{cha:channel-codes} for further details.\newline
 
@@
 for the purpose of checking the accuracy of the reconstruction. Refer
-to Section \ref{sec:Adjoint-simulation-finite} for further details.
+to Section \ref{sec:Adjoint-simulation-finite} for further details.\newline
 

### 2022-11-30 doc/USER_MANUAL/03_mesh_generation.tex

@@
 which explains in a step-by-step tutorial the workflow for the example.
-Please feel free to contribute your own example to this package through the github repository!
+Please feel free to contribute your own example to this package through the github repository!\newline
 
@@
 {\it IMPORTANT:} In order to correctly set up GEOCUBIT and run the examples, please read the file called \texttt{EXAMPLES/README};
-in particular, please make sure you correctly set up the Python paths as indicated in that file.
+in particular, please make sure you correctly set up the Python paths as indicated in that file.\newline
 
@@
 unique identifier, e.g. \textbf{'elastic 1'}, \textbf{'elastic 2'}, etc. The additional
-attributes to the block define the material description.
+attributes to the block define the material description.\newline
 
@@
 separate values for bulk and shear attenuation, $Q_{\kappa}$ and $Q_{\mu}$ respectively.
-Note that Qmu is always equal to Qs, but Qkappa is in general not equal to Qp.
-To convert one to the other see doc/note\_on\_Qkappa\_versus\_Qp.pdf and utils/attenuation/conversion\_from\_Qkappa\_Qmu\_to\_Qp\_Qs\_from\_Dahlen\_Tromp\_959\_960.f90.
+Note that Qmu is always equal to Qs, but Qkappa is in general not equal to Qp. To convert one to the other
+see \texttt{doc/note\_on\_Qkappa\_versus\_Qp.pdf} and\newline
+\texttt{utils/attenuation/conversion\_from\_Qkappa\_Qmu\_to\_Qp\_Qs\_from\_Dahlen\_Tromp\_959\_960.f90}.\newline
 
@@
 which belong to this elastic material. For volumes belonging to another,
-different material, you will need to define a new material block.
+different material, you will need to define a new material block.\newline
 
@@
 block for all boundary surfaces and name it as one of the absorbing
-blocks above, e.g. 'face\_abs\_xmin'.
+blocks above, e.g. 'face\_abs\_xmin'.\newline
 
@@
 or to a Xmax CPML, the flag is the same in both cases; the same is
-true for Ymin and Ymax, and also Zmin and Zmax.
+true for Ymin and Ymax, and also Zmin and Zmax.\newline
 
@@
 \texttt{utils/CPML/convert\_external\_layers\_of\_a\_given\_mesh\_to\_CPML\_layers.f90},\newline
-and a small Makefile is provided in that directory (utils/CPML).
+and a small Makefile is provided in that directory (utils/CPML).\newline
 
@@
 it; never use a SMALLER value otherwise this code will miss some CPML
-elements.
+elements.\newline
 
@@
 that are titled by any constant angle in the horizontal plane. However
-this is not implemented yet.
+this is not implemented yet.\newline
 
@@
 for a while because it makes fluid-solid matching inside the CPML
-much easier.
+much easier.\newline
 
@@
 9999 here. If you want a model with no \texttt{Qkappa} attenuation, set \texttt{Qkappa} to 9999 here.
-Note that Qmu is always equal to Qs, but Qkappa is in general not equal to Qp.
+Note that Qmu is always equal to Qs, but Qkappa is in general not equal to Qp.\newline
 
@@
 to load-balance the mesh. Please see the next section for further
-details.
+details.\newline
 
@@
 Running this code is optional because no information needed by the
-solver is generated.
+solver is generated.\newline
 
@@
 modify the header of the \texttt{check\_mesh\_quality\_CUBIT\_Abaqus.f90}
-source file in directory \texttt{src/check\_mesh\_quality\_CUBIT\_Abaqus/}.
+source file in directory \texttt{src/check\_mesh\_quality\_CUBIT\_Abaqus/}.\newline
 
@@
 This partitioning is executed once and for all prior to the execution
-of the solver so it is referred to as a static mapping.
+of the solver so it is referred to as a static mapping.\newline
 
@@
 If all paths and flags have been set correctly, the executable \texttt{bin/xdecompose\_mesh}
-should be produced.
+should be produced.\newline
 
@@
 the Northern hemisphere corresponds to zones +1 to +60,
-the Southern hemisphere to zones -1 to -60.
+the Southern hemisphere to zones -1 to -60.\newline
 
@@
 which means that current UTM northing at a given point can be 200+ meters different from the old one.
-For different geographic regions, other datum systems (e.g.: ED50, NAD83) can be used.
+For different geographic regions, other datum systems (e.g.: ED50, NAD83) can be used.\newline
 
@@
 calculates a minimum period resolved by the chosen mesh discretization and velocity model,
-output into the file \texttt{OUTPUT\_FILES/output\_generate\_databases.txt}.
+output into the file \texttt{OUTPUT\_FILES/output\_generate\_databases.txt}.\newline
 
@@
 are written on a parallel (global) file system (see also the earlier
-discussion of the \texttt{LOCAL\_PATH\_IS\_ALSO\_GLOBAL} flag in Chapter~\ref{cha:Getting-Started}).
+discussion of the \texttt{LOCAL\_PATH\_IS\_ALSO\_GLOBAL} flag in Chapter~\ref{cha:Getting-Started}).\newline
 

### 2022-11-30 doc/USER_MANUAL/04_creating_databases.tex

@@
 All the runs to perform must be placed in directories called run0001, run0002, run0003 and so on
-(with exactly four digits).
+(with exactly four digits).\newline
 
-Imagine you have 10 independent calculations to do, each of them on 100 cores; you have three options:
+Imagine you have 10 independent calculations to do, each of them on 100 cores; you have three options:\newline
 
-1/ submit 10 jobs to the batch system
+1/ submit 10 jobs to the batch system\newline
 
 2/ submit a single job on 1000 cores to the batch, and in that script create a sub-array of jobs to start 10 jobs,
-each running on 100 cores (see e.g. \url{http://www.schedmd.com/slurmdocs/job_array.html})
+each running on 100 cores (see e.g. \url{http://www.schedmd.com/slurmdocs/job_array.html})\newline
 
@@
 cd into one of 10 subdirectories (called e.g. run0001, run0002,... run0010) depending on the sub-communicator
-your MPI rank belongs to, and run normally on 100 cores using that sub-communicator.
+your MPI rank belongs to, and run normally on 100 cores using that sub-communicator.\newline
 
-The option \texttt{NUMBER\_OF\_SIMULTANEOUS\_RUNS} implements 3/.
+The option \texttt{NUMBER\_OF\_SIMULTANEOUS\_RUNS} implements 3/.\newline
 
@@
 directories) in each of the run0001, run0002,... directories, it is sufficient to have one in run0001
-and the code will broadcast it to the others).
+and the code will broadcast it to the others).\newline
 
@@
 e.g. \texttt{run0002/DATABASES\_MPI}, etc. to store the forward wavefields, kernels etc. of each sub-run.
-This would not be needed for forward simulations.
+This would not be needed for forward simulations.\newline
 
 TODO / future work to do: the sensitivity kernel summing and smoothing tools in directory src/tomography are currently not ported to this new option to do many runs simultaneously, only the solver (src/specfem3d) is.
-Thus these tools should work, but in their current version will need to be run for each simulation result independently.
+Thus these tools should work, but in their current version will need to be run for each simulation result independently.\newline
 
@@
 This works because these routines consider multiple runs by design.
-You simply have to provide them the directories where the kernels are.
+You simply have to provide them the directories where the kernels are.\newline
 
@@
 entrance of the PML, but in practice this is sufficient and produces
-only tiny / negligible spurious reflections.
+only tiny / negligible spurious reflections.\newline
 
@@
 turned off (i.e. $Q_\kappa = Q_\mu = 9999$ in that layer), as in the red layer of Figure~\ref{fig:mesh_extrusion}.
-Our tools in directory utils/CPML implement that transition layer automatically.
+Our tools in directory utils/CPML implement that transition layer automatically.\newline
 
@@
 This applies to the bottom layer as well; there you should make sure
-that your model is 1D and thus constant along the vertical direction.
+that your model is 1D and thus constant along the vertical direction.\newline
 
@@
 physical region, and in all the PML layers extend it by continuity
-from its values along the inner PML edge.
+from its values along the inner PML edge.\newline
 
@@
 file \texttt{setup/constants.h}. The wave speed $v$ is determined based on your model's P- (or S-)
-wave speed values.
+wave speed values.\newline
 

### 2022-11-30 doc/USER_MANUAL/05_running_the_solver.tex

@@
 the Harvard \texttt{CMTSOLUTION} convention (the sign of some components
-is different). The C code outputs both.
+is different). The C code outputs both.\newline
 
@@
 moment release. The PDE solution is not used by our software package
-but must be present anyway in the first line of the file.
+but must be present anyway in the first line of the file.\newline
 
@@
 shift}. Each subevent can have its own half duration, latitude, longitude,
-depth, and moment tensor (effectively, the local moment-density tensor).
+depth, and moment tensor (effectively, the local moment-density tensor).\newline
 
@@
 time or centroid time in general, but the timing of the \textit{center}
-of the source triangle with zero \texttt{time shift} (Fig~\ref{fig:source_timing}).
+of the source triangle with zero \texttt{time shift} (Fig~\ref{fig:source_timing}).\newline
 
@@
 and convolve afterwards, since the half duration is generally fixed
-by the finite fault model.
+by the finite fault model.\newline
 
@@
 disks (the path \texttt{LOCAL\_PATH} in the \texttt{DATA/Par\_file} and
-need to be gathered at the end of the simulation.
+need to be gathered at the end of the simulation.\newline
 
@@
 \texttt{\small NTSTEP\_BETWEEN\_OUTPUT\_INFO}
-in the \texttt{DATA/Par\_file}.
+in the \texttt{DATA/Par\_file}.\newline
 
@@
 in sequence. This is a safe approach that ensures using the correct
-combination of distributed database output and solver input.
+combination of distributed database output and solver input.\newline
 
@@
 with various flags turned on or off to develop a sense for what is
-involved.
+involved.\newline
 
@@
 To turn that option on, set parameter \texttt{NUMBER\_OF\_SIMULTANEOUS\_RUNS}
-to a value greater than 1 in file \texttt{DATA/Par\_file}.
+to a value greater than 1 in file \texttt{DATA/Par\_file}.\newline
 
@@
 and thus the total number of processor cores to request from the batch system
-should be \texttt{NUMBER\_OF\_SIMULTANEOUS\_RUNS $\times$ NPROC}.
+should be \texttt{NUMBER\_OF\_SIMULTANEOUS\_RUNS $\times$ NPROC}.\newline
 

### 2022-11-30 doc/USER_MANUAL/11_graphics.tex

@@
 You can then use Paraview \urlwithparentheses{www.paraview.org}
-to visualized these partitions.
+to visualized these partitions.\newline
 
@@
 }
-Here, $0$ is the number of the first processor, $3$ the number of the last one, \texttt{velocity\_Z\_it000400} the name of the files we want to combine without the prefix \texttt{proc000000*\_}, \texttt{./OUTPUT\_FILES/DATABASES\_MPI} the directory where the files are located, \texttt{./OUTPUT\_FILES} the directory where the combined file will be stored, and $0$ is the parameter to create a low-resolution mesh file ($1$ for a high-resolution).
+Here, $0$ is the number of the first processor, $3$ the number of the last one, \texttt{velocity\_Z\_it000400} the name of the files we want to combine without the prefix \texttt{proc000000*\_}, \texttt{./OUTPUT\_FILES/DATABASES\_MPI} the directory where the files are located, \texttt{./OUTPUT\_FILES} the directory where the combined file will be stored, and $0$ is the parameter to create a low-resolution mesh file ($1$ for a high-resolution).\newline
 
@@
 If you use the executable that does not have the \_vtk extension, then the output mesh file will have a name of the form \texttt{velocity\_Z\_it000400.mesh}.
-You will then have to convert the \texttt{.mesh} file into the VTU (Unstructured grid file) format which can be viewed in ParaView.
+You will then have to convert the \texttt{.mesh} file into the VTU (Unstructured grid file) format which can be viewed in ParaView.\newline
 
@@
 uses the VTK \urlwithparentheses{http://www.vtk.org} run-time library for its execution. Therefore, make sure you have them properly set
-in the script according to your system.
+in the script according to your system.\newline
 
@@
 circle path, it is sufficient to collect only the slices that are
-along or close to the great circle path.
+along or close to the great circle path.\newline
 

### 2022-11-30 doc/USER_MANUAL/A_reference_frame.tex
updates tex files (\\ to \newline) for better pandoc conversions to wiki pages

@@
 the location by \textquotedbl{}longitude/latitude/depth\textquotedbl{}.
-We read in longitude as $x$ coordinate, latitude as $y$ coordinate.
+We read in longitude as $x$ coordinate, latitude as $y$ coordinate.\newline
 
@@
 to Cartesian coordinates using a UTM conversion with the specified
-UTM zone.
+UTM zone.\newline
 
@@
 flag \texttt{USE\_SOURCES\_RECEIVERS\_Z} to \texttt{.true.} in
-\texttt{DATA/Par\_file}.
+\texttt{DATA/Par\_file}.\newline
 
@@
 component normal to the surface of the mesh, $x$ and $y$ directions
-are somewhat arbitrary orthogonal directions along the surface.
+are somewhat arbitrary orthogonal directions along the surface.\newline
 
@@
 \end{itemize}
-\item [{acoustic domain:}] ~\\
+\item [{acoustic domain:}] ~\newline
  (please note that receiver stations in acoustic domains must be buried.

### 2022-11-30 doc/USER_MANUAL/C_troubleshooting.tex

@@
 In many cases, the paths to these specific compiler commands F90,
-CC and MPIF90 will not be correct if `./configure` fails.
+CC and MPIF90 will not be correct if `./configure` fails.\newline
 
@@
 
-Make sure you're pointing to the right 'mpif90' wrapper command.
+Make sure you're pointing to the right 'mpif90' wrapper command.\newline
 
@@
 Fortran compilers. That is, using e.g. gfortran to compile non-MPI
-files and mpif90, wrapper provided for e.g. ifort, to compile MPI-files.
+files and mpif90, wrapper provided for e.g. ifort, to compile MPI-files.\newline
 
@@
 which gives you that: 27,648 / 33,792 $\sim$ 81.8 \% of all elements
-are not skewed, i.e. regular elements. a fantastic value :)
+are not skewed, i.e. regular elements. a fantastic value :)\newline
 
@@
 are still accurate enough. So empirically, values up to around 0.7
-are tolerable, above that you should consider remeshing...
+are tolerable, above that you should consider remeshing...\newline
 
@@
 / element\_size about constant). Those elements in this doubling layer
-have higher skewness values and make up those 6 \% in the histogram.
+have higher skewness values and make up those 6 \% in the histogram.\newline
 
@@
 the simulation with \texttt{xspecfem3D} was suppressing this projection
-from latitude/longitude to x/y/z coordinates.
+from latitude/longitude to x/y/z coordinates.\newline
 
@@
 \texttt{xgenerate\_databases} and \texttt{xspecfem3D} together to
-run simulations.
+run simulations.\newline
 
@@
 to be the same as in \texttt{Mesh\_Par\_file} and rerun \texttt{xgenerate\_databases}
-and \texttt{xspecfem3D} to make sure that your simulation works fine.
+and \texttt{xspecfem3D} to make sure that your simulation works fine.\newline
 
@@
 then change \texttt{DT} in the \texttt{DATA/Par\_file} to be somewhere
-close to the maximum suggested time step. In the example above: \\
- \texttt{DT = 0.05d0} \\
+close to the maximum suggested time step. In the example above: \newline
+ \texttt{DT = 0.05d0} \newline
  would (most probably) work fine. It could be also bigger than the

### 2022-11-30 doc/USER_MANUAL/notes_and_acknowledgement.tex

@@
 arguments (which can be grouped in ``derived types'' if needed for
-clarity).
+clarity).\newline
 
@@
 Paul F. Fischer (Brown University, Providence, Rhode Island, USA,
-now at Argonne National Laboratory, USA).
+now at Argonne National Laboratory, USA).\newline
 

### 2022-12-01 doc/USER_MANUAL/07_adjoint_simulations.tex
user manual updated - minor updates

@@
 cartesian coordinates follows \citet{ChTr07}. The 21 anisotropic
-kernels are saved in the \texttt{LOCAL\_PATH} in one file with the
-name of \texttt{proc??????\_cijkl\_kernel.bin} (with \texttt{proc??????}
+kernels are saved in the \texttt{LOCAL\_PATH} as 21 files with names 
+\texttt{proc??????\_c??\_kernel.bin} (with \texttt{proc??????}
 the processor number). The output kernels correspond to absolute perturbations

### 2022-12-01 doc/USER_MANUAL/14_post_processing.tex
user manual updated - minor updates

@@
 to the output directory.
-\item [{\texttt{USE\_OUTPUT\_FILES\_PATH}}] Set to \texttt{.false.} to
-have the seismograms output to \texttt{LOCAL\_PATH} directory specified
-in the main parameter file \texttt{DATA/Par\_file}. In this case,
-you could collect the synthetics onto the frontend using the \texttt{collect\_seismo\_lsf\_multi.pl}
-script located in the \texttt{utils/Cluster/lsf/} directory. The usage
-of the script would be e.g.:
-
-\begin{verbatim}
-collect_seismo.pl machines DATA/Par_file
-\end{verbatim}
-
 \end{description}
-
-where \texttt{machines} is a file containing the node names and \texttt{DATA/Par\_file}
-the parameter file used to extract the \texttt{LOCAL\_PATH} directory
-used for the simulation.
 

### 2022-12-08 doc/USER_MANUAL/04_creating_databases.tex
user manual updated - minor updates
updates doc

@@
 and the norm of the force used is the factor force source times the
-source time function.
+source-time function.
 
@@
 is more precise than using the closest GLL mesh point, but it is also a bit slower.
-If needed, you can change that default behavior and force the code to use the closest
-GLL mesh point instead by setting flag \texttt{USE\_BEST\_LOCATION} to \texttt{.false.} instead of \texttt{.true.} in file
-\texttt{setup/constants.h} and recompiling the code.
 \item [{\texttt{USE\_RICKER\_TIME\_FUNCTION}}] Turn this flag on to use
-a Ricker source time function, i.e., the second derivative of a Gaussian, instead of the source time functions
+a Ricker source-time function, i.e., the second derivative of a Gaussian, instead of the source-time functions
 set by default to represent a (tilted) \texttt{FORCESOLUTION} force
@@
 for elastic cases to represent the permanent slip on the fault while
-in the acoustic case a Gaussian source time function with a similarly
+in the acoustic case a Gaussian source-time function with a similarly
 short half duration is defined to physically describe actions within
 the fluid. Otherwise, if a \texttt{FORCESOLUTION} force source is
-used, a (pseudo) Dirac delta source time function is defined by default.
+used, a (pseudo) Dirac delta source-time function is defined by default.
 Any other source-time function may then be obtained by convolution.
 \item [{\texttt{PRINT\_SOURCE\_TIME\_FUNCTION}}] Turn this flag on to print
-information about the source time function in the file \texttt{OUTPUT\_FILES/plot\_source\_time\_function.txt}.
+information about the source-time function in the file \texttt{OUTPUT\_FILES/plot\_source\_time\_function.txt}.
 This feature is only relevant for the solver.

### 2022-12-08 doc/USER_MANUAL/05_running_the_solver.tex
updates doc

@@
 For any particular earthquake, the \texttt{CMTSOLUTION} file that
-represents the point source may be obtained directly from the Harvard
-Centroid-Moment Tensor (CMT) web page \urlwithparentheses{www.seismology.harvard.edu}.
+represents the point source may be obtained directly from the Global
+Centroid-Moment Tensor (CMT) web page \urlwithparentheses{www.globalcmt.org}.
 It looks like the example shown in Fig. \ref{fig:CMTSOLUTION-file}.
@@
 \caption{\texttt{CMTSOLUTION} file based on the
-format from the Harvard CMT catalog. \textbf{M} is the moment tensor,
+format from the Global CMT catalog. \textbf{M} is the moment tensor,
 $M_{0}${\small {} }is the seismic moment, and $M_{w}$ is the moment
@@
 \item Set the latitude or UTM $x$ coordinate, longitude or UTM $y$ coordinate, depth of the source (in km). Remark: In principle in the international CMTSOLUTION format in geophysics
-the depth is given in kilometers; however for users in other fields (non-destructive testing, medical imaging, near-surface studies...) who may prefer to give the position of the source (rather than its depth from the surface), or for people who use FORCESOLUTION to describe the source rather than CMTSOLUTION, we provide an option called USE\_SOURCES\_RECEIVERS\_Z in the \texttt{DATA/Par\_file}, and if so that position is read from CMTSOLUTION in meters rather than kilometers (and again, it is then the true position in the mesh, not the depth).
-When option USE\_SOURCES\_RECEIVERS\_Z in the \texttt{DATA/Par\_file} is on, this remark applies to the position of the receivers as well.
+the depth is given in kilometers; however for users in other fields (non-destructive testing, medical imaging, near-surface studies...) who may prefer to give the position of the source (rather than its depth from the surface), or for people who use \texttt{FORCESOLUTION} to describe the source rather than \texttt{CMTSOLUTION}, we provide an option called \texttt{USE\_SOURCES\_RECEIVERS\_Z} in the \texttt{DATA/Par\_file}, and if so that position is read from \texttt{CMTSOLUTION} in meters rather than kilometers (and again, it is then the true position in the mesh, not the depth).
+When option \texttt{USE\_SOURCES\_RECEIVERS\_Z} in the \texttt{DATA/Par\_file} is on, this remark applies to the position of the receivers as well.
 \item Set the \texttt{time shift} parameter equal to $0.0$ (the solver
@@
 in the post-processing (see Section \ref{sec:Process-data-and-syn}).
-\item For point-source simulations (see finite sources, page \pageref{To-simulate-a})
-we recommend setting the source half-duration parameter \texttt{half
-duration} equal to zero, which corresponds to simulating a step source-time
-function, i.e., a moment-rate function that is a delta function. If
-\texttt{half duration} is not set to zero, the code will use a Gaussian
+\item For point-source simulations (see finite sources, page \pageref{To-simulate-a}),
+setting the source half-duration parameter \texttt{half duration} equal to zero corresponds to simulating a step source-time
+function (Heaviside), i.e., to using a moment-rate function that is a delta function. If
+\texttt{half duration} is not set to zero, the code will use a smooth (pseudo) Heaviside source-time function
+with a corresponding Gaussian moment-rate function
 (i.e., a signal with a shape similar to a `smoothed triangle', as
 explained in \citet{KoTr02a} and shown in Fig~\ref{fig:gauss.vs.triangle})
-source-time function with half-width \texttt{half duration}. We prefer
-to run the solver with \texttt{half duration} set to zero and convolve
+with half-width \texttt{half duration}.
+
+Often, it is preferable to run the solver with \texttt{half duration} set to zero and convolve
 the resulting synthetic seismograms in post-processing after the run,
@@
 time function may be safely filtered out afterward based upon a convolution
-with the desired source time function and/or low-pass filtering. Use
+with the desired source-time function and/or low-pass filtering. Use
 the serial code \texttt{convolve\_source\_timefunction.f90} and the
@@
 or the centroid time of the earthquake. The start time of the simulation
-is $t=-2.0*\texttt{half duration}$ (the 2.0 factor is to make sure the moment
-rate function is very close to zero when starting the simulation. This avoids spurious high-frequency oscillations.
-Also, for acoustic simulations using a Gaussian STF this is set to a factor 3.0, while Ricker source time functions use a factor 1.2).
+is $t=-2.0*\texttt{half duration}$ (the 2.0 factor is to make sure the moment-rate function is very close to zero
+when starting the simulation. This avoids spurious high-frequency oscillations.
+Also, for acoustic simulations using a Gaussian STF this is set to a factor 3.0, while Ricker source-time functions use a factor 1.2).
 To convert to absolute time $t_{\mathrm{abs}}$, set
@@
 Note that the \citet{AkRi80} convention is slightly different from
-the Harvard \texttt{CMTSOLUTION} convention (the sign of some components
+the Global/Harvard \texttt{CMTSOLUTION} convention (the sign of some components
 is different). The C code outputs both.\newline
@@
 your own \texttt{CMTSOLUTION} file. Just make sure that the resulting
-file adheres to the Harvard CMT conventions (see Appendix~\ref{cha:Coordinates}).
+file adheres to the Global/Harvard CMT conventions (see Appendix~\ref{cha:Coordinates}).
 Note that the first line in the \texttt{CMTSOLUTION} file is the Preliminary
 Determination of Earthquakes (PDE) solution performed by the USGS
-NEIC, which is used as a seed for the Harvard CMT inversion. The PDE
+NEIC, which is used as a seed for the Global/Harvard CMT inversion. The PDE
 solution is based upon P waves and often gives the hypocenter of the
@@
 \begin{itemize}
-\item Set the \texttt{time shift} parameter equal to $0.0$ (the solver
-will not run otherwise.) The time shift parameter would simply apply
-an overall time shift to the synthetics, something that can be done
-in the post-processing (see Section \ref{sec:Process-data-and-syn}).
-\item Set the \texttt{f0} parameter (the dominant frequency) of the Ricker source time
-function (i.e., the second derivative of a Gaussian) when {\texttt{USE\_RICKER\_TIME\_FUNCTION}} is turned on
-in the main parameter file \texttt{DATA/Par\_file}. In case that the
-solver uses a (pseudo) Dirac delta source time function to represent
-a force point source, a very short duration of five time steps is automatically set by default.
+\item The first line is only the header for the force solution, which can
+be used as the identifier for the force source.
+\item \texttt{time shift:} For a single force source, set
+this  parameter equal to $0.0$ (The solver will not run otherwise!); the time
+shift parameter would simply apply an overall time shift to the synthetics,
+something that can be done in the post-processing (see Section \ref{sec:Process-data-and-syn}).
+\item \texttt{hdurorf0:} Set the half duration value (s) for step source-time
+function, or the dominant frequency (Hz) for Ricker source-time functions.
+In case that the source uses a (pseudo) Dirac delta source-time function (i.e., a Gaussian with a zero half duration) to represent
+a force point source, a very short half duration of five time steps is automatically set by default.
+For a Ricker source-time function, set the dominant frequency value (f0) in Hz.
+See the parameter \texttt{source time function:} below for source-time functions.
+\item \texttt{latorUTM:} Set the latitude or UTM $x$ coordinate.
+\item \texttt{longorUTM:} Set the longitude or UTM $y$ coordinate.
+\item \texttt{depth:} Set the depth of the source (in km).
+\item \texttt{source time function:} Set the type of source-time function: 0 = Gaussian function, 1 = Ricker function, 2 = Heaviside (step) function, 3 = monochromatic function, 4 = Gaussian function as defined in \citet{Meschede2011}.
+When {\texttt{USE\_RICKER\_TIME\_FUNCTION}} is turned on in the main parameter file \texttt{DATA/Par\_file},
+it will override this source time function type selection and always use a Ricker wavelet.
 Note that we use the standard definition of a Ricker, for a dominant frequency $f_0$:
@@
 This gives the wavelet of Figure~\ref{fig:RickerWavelet}.
-\item Set the latitude or UTM $x$ coordinate, longitude or UTM $y$ coordinate, depth of the source (in km).
-\item Set the magnitude of the force source.
-\item Set the components of a (non-unitary) direction vector for the force
-source in the East/North/Vertical basis (see Appendix A for the orientation
+\item \texttt{factor force source:} Set the magnitude of the force source (units in Newton N).
+\item \texttt{component dir vect source E:} Set the East component of a direction vector for the force
+source. Direction vector is not necessarily a unit vector.
+\item \texttt{component dir vect source N:} Set the North component of a direction vector for the force
+source.
+\item \texttt{component dir vect source Z\_up:} Set the vertical component of a direction vector for the force
+source. Sign convention follows the positive upward direction (see Appendix A for the orientation
 of the reference frame).
@@
 At least one entry (not necessarily the first) must have a zero \texttt{time
-shift}, and all the other entries must have non-negative \texttt{time
-shift}. Each subevent can have its own latitude, longitude, depth,
-\texttt{half duration} and force parameters.
+shift}, and all the other entries must have non-negative \texttt{time shift}.
+Each subevent can have its own set of parameters \texttt{latitude}, \texttt{longitude}, \texttt{depth},
+\texttt{half duration} etc.\newline
 
@@
 
-In addition to inbuild source time function, the solver can also use an external source time function defined by the user.
-This option can be activated by setting \texttt{USE\_EXTERNAL\_SOURCE\_FILE} to true in \texttt{DATA/Par\_file} and by adding the name of the file containing the source time function at the end of \texttt{FORCESOLUTION} or \texttt{CMTSOLUTION} files.
-The source time function file must contain the value of the time step on its first line, and then a single column with the amplitude of the source time function for all the time steps.
+In addition to inbuild source-time functions, the solver can also use an external source-time function defined by the user.
+This option can be activated by setting \texttt{USE\_EXTERNAL\_SOURCE\_FILE} to \texttt{.true.} in \texttt{DATA/Par\_file} and by adding the name of the file containing the source-time function at the end of \texttt{FORCESOLUTION} or \texttt{CMTSOLUTION} files.
+The source-time function file must contain a single column with the amplitudes of the source-time function for all the time steps.
 The time step must be exactly the same as that used for the simulation.
-Note  when the  \texttt{EXTERNAL\_SOURCE\_FILE} is set to false then the line with the external source time function file must not appear in the files  \texttt{FORCESOLUTION} and \texttt{CMTSOLUTION} otherwise the solver will exit with an error.
+When the flag is set to \texttt{.false.}, then the line with the external source-time function file must not appear in the files \texttt{FORCESOLUTION} and \texttt{CMTSOLUTION}, otherwise the solver will exit with an error.
 When using an external source file, you can still set up the source location and directivity as in the default case.
-In the FORCESOLUTION file: you set "latorUTM", "longorUTM" and "depth" to define the position of your point source.
-Then if you want to define a directivity, then change the following lines: "component dir vect source E", "component dir vect source N" and "component dir vect source Z\_UP".
-What you are doing is simply that you define the source position and directivity the same way as in the default case, but in addition you are specifying the path to read in a non-default source time function from an external file.
+In the \texttt{FORCESOLUTION} file: you set "latorUTM", "longorUTM" and "depth" to define the position of your point source.
+Then if you want to define a directivity, change the following lines: "component dir vect source E", "component dir vect source N" and "component dir vect source Z\_UP".
+What you are doing is simply that you define the source position and directivity the same way as in the default case, but in addition you are specifying the path to read in a non-default source-time function from an external file.
 
@@
 \item [-] In principle in the international CMTSOLUTION format in geophysics
-the depth is given in kilometers; however for users in other fields (non-destructive testing, medical imaging, near-surface studies...) who may prefer to give the position of the source (rather than its depth from the surface), or for people who use FORCESOLUTION to describe the source rather than CMTSOLUTION, we provide an option called USE\_SOURCES\_RECEIVERS\_Z in the \texttt{DATA/Par\_file}, and if so that position is read from CMTSOLUTION in meters rather than kilometers (and again, it is then the true position in the mesh, not the depth).
-When option USE\_SOURCES\_RECEIVERS\_Z in the \texttt{DATA/Par\_file} is on, this remark applies to the position of the receivers as well.\newline
+the depth is given in kilometers; however for users in other fields (non-destructive testing, medical imaging, near-surface studies...) who may prefer to give the position of the source (rather than its depth from the surface), or for people who use \texttt{FORCESOLUTION} to describe the source rather than \texttt{CMTSOLUTION}, we provide an option called \texttt{USE\_SOURCES\_RECEIVERS\_Z} in the \texttt{DATA/Par\_file}, and if so that position is read from \texttt{CMTSOLUTION} in meters rather than kilometers (and again, it is then the true position in the mesh, not the depth).
+When option \texttt{USE\_SOURCES\_RECEIVERS\_Z} in the \texttt{DATA/Par\_file} is on, this remark applies to the position of the receivers as well.\newline
 \item [-] Let's say you want to place two receivers at (x1,y1,z1) and

### 2022-12-08 doc/USER_MANUAL/14_post_processing.tex
updates doc

@@
 station/event and timing information.
-\item Convolve synthetic seismograms with a source time function with the
+\item Convolve synthetic seismograms with a source-time function with the
 half duration specified in the \texttt{CMTSOLUTION} file, provided,
@@
 to the origin time of the synthetics, convolving the synthetics with
-a triangular source time function with a given half duration, etc.
+a triangular source-time function with a given half duration, etc.
 Type \texttt{process\_syn.pl} without any argument for a detailed

### 2023-03-09 doc/USER_MANUAL/02_getting_started.tex
updates doc

@@
 Note that we use CUBIT (now called Trelis) to create meshes of hexahedra, but other packages
-can be used as well, for instance GiD from \url{http://gid.cimne.upc.es}
-or Gmsh from \url{http://geuz.org/gmsh} \citep{GeRe09}. Even mesh
+can be used as well, for instance GiD from \url{https://www.gidsimulation.com}
+or Gmsh from \url{http://gmsh.info} \citep{GeRe09}. Even mesh
 creation packages that generate tetrahedra, for instance TetGen from
@@
 \begin{itemize}
-\item[\textbullet] a good and short summary for the basic utilisation of Doxygen:
+\item[\textbullet] short summary for the basic utilisation of Doxygen:
 
-\url{http://www.softeng.rl.ac.uk/blog/2010/jan/30/callgraph-fortran-doxygen/},
+\url{https://www.doxygen.nl/manual/starting.html},
 
@@
 
-\url{http://www.stack.nl/~dimitri/doxygen/manual/diagrams.html},
+\url{https://www.doxygen.nl/manual/diagrams.html},
 
@@
 
-\url{http://www.stack.nl/~dimitri/doxygen/manual/config.html},
+\url{https://www.doxygen.nl/manual/config.html},
 
@@
 
-\url{http://www.stack.nl/~dimitri/doxygen/manual/index.html}.
+\url{https://www.doxygen.nl/manual/index.html}.
 

### 2023-03-09 doc/USER_MANUAL/05_running_the_solver.tex
updates doc

@@
 represents the point source may be obtained directly from the Global
-Centroid-Moment Tensor (CMT) web page \urlwithparentheses{www.globalcmt.org}.
+Centroid-Moment Tensor (CMT) \href{www.globalcmt.org}{web page}.
 It looks like the example shown in Fig. \ref{fig:CMTSOLUTION-file}.
@@
 the serial code \texttt{convolve\_source\_timefunction.f90} and the
-script \texttt{convolve\_source\_timefunction.csh} for this purpose,
-or alternatively use signal-processing software packages such as SAC
-\urlwithparentheses{www.llnl.gov/sac}. Type
+script \texttt{convolve\_source\_timefunction.csh} for this purpose
+(or alternatively use signal-processing software packages such as \href{https://seiscode.iris.washington.edu/projects/sac}{SAC}).
+Type
 {\small

### 2023-03-09 doc/USER_MANUAL/11_graphics.tex
updates doc

@@
 prior to running the mesher (see Chapter~\ref{cha:Running-the-Mesher-Meshfem3D}
-for details). You can then use AVS \urlwithparentheses{www.avs.com}
-or OpenDX \urlwithparentheses{www.opendx.org} to visualize the mesh
+for details). You can then use \href{www.avs.com}{AVS}
+or \href{www.opendx.org}{OpenDX} to visualize the mesh
 and MPI partition (slices).
@@
 You have also the option to visualize the distributed databases produced
-by \texttt{xgenerate\_databases} using Paraview \urlwithparentheses{www.paraview.org}.
+by \texttt{xgenerate\_databases} using \href{www.paraview.org}{Paraview}.
 For this, you must set the flag \texttt{SAVE\_MESH\_FILES} to \texttt{.true.}
@@
 for details). This will create VTK files for each single partition.
-You can then use Paraview \urlwithparentheses{www.paraview.org}
+You can then use \href{www.paraview.org}{Paraview}
 to visualized these partitions.\newline
@@
 to produce animations of southern California earthquakes via the web;
-see Southern California ShakeMovie\textregistered{}\urlwithparentheses{www.shakemovie.caltech.edu}.
+see Southern California \href{www.shakemovie.caltech.edu}{ShakeMovie\textregistered{}}.
 
@@
 Notice that this Perl script uses a program \texttt{mesh2vtu} in the \texttt{utils/Visualization/Paraview/mesh2vtu} directory, which further
-uses the VTK \urlwithparentheses{http://www.vtk.org} run-time library for its execution. Therefore, make sure you have them properly set
+uses the \href{http://www.vtk.org}{VTK} run-time library for its execution. Therefore, make sure you have them properly set
 in the script according to your system.\newline
@@
 \texttt{\small utils/Visualization/Paraview/mesh2vtu} directory, which further
-uses the VTK \urlwithparentheses{http://www.vtk.org/} run-time library
+uses the \href{http://www.vtk.org/}{VTK} run-time library
 for its execution. Therefore, make sure you have them properly set
@@
 
-Finally, we can view the mesh in ParaView \urlwithparentheses{www.paraview.org}.
+Finally, we can view the mesh in \href{www.paraview.org}{ParaView}.
 \begin{enumerate}
@@
 \textsf{Glyph}, and represent the points by `\textsf{spheres}'.
-\item For more information about ParaView, see the ParaView Users Guide
-\urlwithparentheses{www.paraview.org/files/v1.6/ParaViewUsersGuide.PDF}.
+\item For more information about ParaView, see the
+\href{www.paraview.org/files/v1.6/ParaViewUsersGuide.PDF}{ParaView Users Guide}.
 \end{enumerate}

### 2023-03-09 doc/USER_MANUAL/14_post_processing.tex
updates doc

@@
 where the first command performs rotation on the SAC data obtained
-through Seismogram Transfer Program (STP) \urlwithparentheses{http://www.data.scec.org/STP/stp.html},
+through Seismogram Transfer Program (\href{http://www.data.scec.org/STP/stp.html}{STP}),
 while the second command rotates the processed SEM synthetics.

### 2023-03-09 doc/USER_MANUAL/copyright_and_version.tex
updates doc

@@
 
-MPI v. 3.0, December 2014: many developers.
+version 4.0, December 2022:
+Rafael Almada, Jean-Paul Ampuero, Etienne Bachmann, Kangchen Bai, Stephen Beller, Jordan Bishop, Alexis Bottero,
+Emanuele Casarotti, Clement Durochat, Rene Gassmoeller, Hom Nath Gharti, Leopold Grinberg, Aakash Gupta,
+Foivos Karakostas, Dimitri Komatitsch, Qinya Liu, Geordie McBain, Ryan Modrak, Vadim Monteiller, Masaru Nagaso, Elif Oral,
+Daniel Peter, Elliott Sales de Andrade,  James Smith, Carl Tape, Eduardo Valero Cano, Huihui Weng, Zhinan Xie:
+various code improvements (fault solver, FK/DSM/AxiSEM coupling, gravity perturbations, energy integrals);
+inverse problem module; parallel heuristic mesh decomposer; ADIOS2 file I/O support;
+ASDF seismogram output; HIP GPU support.\newline
+
+version 3.0, December 2014: many developers.
 Convolutional PML, LDDRK time scheme, bulk attenuation support, simultaneous MPI runs,
@@
 
-MPI v. 2.1, July 2012: Max Rietmann, Peter Messmer, Daniel Peter, Dimitri
+version 2.1, July 2012: Max Rietmann, Peter Messmer, Daniel Peter, Dimitri
 Komatitsch, Joseph Charles, Zhinan Xie: support for CUDA GPUs, better
@@
 
-MPI v. 2.0, November 2010: Dimitri Komatitsch, Nicolas Le Goff, Roland
+version 2.0, November 2010: Dimitri Komatitsch, Nicolas Le Goff, Roland
 Martin and Pieyre Le Loher, University of Pau, France, Daniel Peter,
@@
 
-MPI v. 1.4 Dimitri Komatitsch, University of Pau, Qinya Liu and others,
+version 1.4 Dimitri Komatitsch, University of Pau, Qinya Liu and others,
 Caltech, September 2006: better adjoint and kernel calculations, faster
@@
 
-MPI v. 1.3 Dimitri Komatitsch, University of Pau, and Qinya Liu, Caltech,
+version 1.3 Dimitri Komatitsch, University of Pau, and Qinya Liu, Caltech,
 July 2005: serial version, regular mesh, adjoint and kernel calculations,
@@
 
-MPI v. 1.2 Min Chen and Dimitri Komatitsch, Caltech, July 2004: full
+version 1.2 Min Chen and Dimitri Komatitsch, Caltech, July 2004: full
 anisotropy, volume movie.\newline
 
-MPI v. 1.1 Dimitri Komatitsch, Caltech, October 2002: Zhu's Moho map,
+version 1.1 Dimitri Komatitsch, Caltech, October 2002: Zhu's Moho map,
 scaling of $V_{s}$ with depth, Hauksson's regional model, attenuation,
@@
 
-MPI v. 1.0 Dimitri Komatitsch, Caltech, USA, May 2002: first MPI version
+version 1.0 (MPI) Dimitri Komatitsch, Caltech, USA, May 2002: first MPI version
 based on global code.\newline

### 2023-04-23 doc/USER_MANUAL/03_mesh_generation.tex

@@
 \item [{\texttt{CREATE\_ABAQUS\_FILES}}] Set this flag to \texttt{.true.}
-to save Abaqus FEA \urlwithparentheses{www.simulia.com} mesh files
+to save \href{www.simulia.com}{Abaqus FEA} mesh files
 for subsequent viewing. Turning the flag on generates files in the
@@
 \item [{\texttt{CREATE\_DX\_FILES}}] Set this flag to \texttt{.true.} to
-save OpenDX \urlwithparentheses{www.opendx.org} mesh files for subsequent
+save \href{www.opendx.org}{OpenDX} mesh files for subsequent
 viewing.

### 2023-04-23 doc/USER_MANUAL/04_creating_databases.tex

@@
 flat, i.e. identical to HEX8). To generate HEX27 elements with curvature
-properly taken into account, you can use Gmsh \url{http://geuz.org/gmsh/}
+properly taken into account, you can use Gmsh \url{http://gmsh.info}
 \item [{\texttt{MODEL}}] Must be set to one of the following:
@@
 \item [{\texttt{SAVE\_MESH\_FILES}}] Set this flag to \texttt{.true.} to
-save ParaView \urlwithparentheses{www.paraview.org} mesh files for
+save \href{www.paraview.org}{ParaView} mesh files for
 subsequent viewing. Turning the flag on generates large (distributed)

### 2023-04-23 doc/USER_MANUAL/08_doing_tomography.tex

@@
 
-\red{\Large ANNOUNCEMENT (June 2017): See also the next chapter (Chapter~\ref{cha:FWI}) about how to perform full waveform inversion (FWI) or source inversions.}\newline
+\red{See also the next chapter (Chapter~\ref{cha:FWI}) about how to perform full waveform inversion (FWI) or source inversions.}\newline
 
-UNDER CONSTRUCTION (July 2015). New content will be added soon.
+STILL UNDER CONSTRUCTION ... New content will be added soon.
 
@@
 and use the misfit between the two to iteratively improve the model
-described by (at least) $V_{{\rm p}}$, $V_{{\rm s}}$, and $\rho$ as a function of space.
+described by (at least) $V_{{\rm p}}$, $V_{{\rm s}}$, and $\rho$ as a function of space.\newline
 
@@
 and synthetic seismograms. In other words, the goal is to find a structural model that can explain a larger portion
-of seismological records, and not simply the phase of a few seismic arrivals.
+of seismological records, and not simply the phase of a few seismic arrivals.\newline
 
@@
 \cite{BaChLa77}, \cite{BaChHeLa82}, \cite{Lai83} and \cite{Tar84}. The idea is to resort to the adjoint state,
-which corresponds to the wavefield emitted and back-propagated from the receivers \cite[e.g.,][]{TrTaLi05,TrKoLi08,Plessix_2006_RAS,FiBuIg06}.
+which corresponds to the wavefield emitted and back-propagated from the receivers \cite[e.g.,][]{TrTaLi05,TrKoLi08,Plessix_2006_RAS,FiBuIg06}.\newline
 
@@
 For example, for the cross-correlated traveltime of a seismic
-phase, the adjoint source is defined as the velocity of that synthetic phase weighted by the travel-time residual.
+phase, the adjoint source is defined as the velocity of that synthetic phase weighted by the travel-time residual.\newline
 
@@
 or a large amount of disk or memory checkpointing \cite[e.g.,][]{FiKeIg09,RuHaPuGu14,CyShWi15}
-the amount of disk storage required would currently be too large.
+the amount of disk storage required would currently be too large.\newline
 
@@
 step of the forward run, which is then used as an initial condition to redo the forward run backwards,
-as well as the field on the outer edges of the mesh for each time steps in order to be able to undo the absorbing boundary conditions.
+as well as the field on the outer edges of the mesh for each time steps in order to be able to undo the absorbing boundary conditions.\newline
 
@@
 \texttt{INPUT\_DIR} directory from which kernels are read,
-\texttt{OUTPUT\_DIR} directory to which smoothed kernels are written.
+\texttt{OUTPUT\_DIR} directory to which smoothed kernels are written.\newline
 
-Smooths kernels by convolution with a Gaussian. Writes the resulting
-smoothed kernels to \texttt{OUTPUT\_DIR}.
+This smooths kernels by convolution with a Gaussian, and writes the resulting
+smoothed kernels to \texttt{OUTPUT\_DIR}.\newline
 
@@
 \texttt{INPUT\_DIR} directory from which arrays are read,
-\texttt{OUTPUT\_DIR} directory to which clipped array are written.
+\texttt{OUTPUT\_DIR} directory to which clipped array are written.\newline
 
 For each name in \texttt{KERNEL\_NAMES}, reads kernels from \texttt{INPUT\_DIR}, applies
-thresholds, and writes the resulting clipped kernels to \texttt{OUTPUT\_DIR}.
+thresholds, and writes the resulting clipped kernels to \texttt{OUTPUT\_DIR}.\newline
 
@@
 \texttt{INPUT\_FILE} text file containing list of kernel directories,
-\texttt{OUTPUT\_PATH} directory to which summed kernels are written.
+\texttt{OUTPUT\_PATH} directory to which summed kernels are written.\newline
 
 For each name in \texttt{KERNEL\_NAMES}, sums kernels from directories specified in
-\texttt{INPUT\_FILE}. Writes the resulting sums to \texttt{OUTPUT\_DIR}.
+\texttt{INPUT\_FILE}. Writes the resulting sums to \texttt{OUTPUT\_DIR}.\newline
 
 \texttt{INPUT\_FILE} is a text file containing a list of absolute or relative paths to
-kernel directories, one directory per line.
+kernel directories, one directory per line.\newline
 
@@
 \texttt{OUTPUT-MODEL-DIR/} (optional) directory which will hold new model files (e.g. \texttt{proc***vp\_new.bin},.., by default directory
-\texttt{OUTPUT\_MODEL/} is used).
+\texttt{OUTPUT\_MODEL/} is used).\newline
 
-The gradients are given for isotropic parameters \texttt{(alpha,beta,rho)} or \texttt{(bulk\_c,beta,rho)}.
+The gradients are given for isotropic parameters \texttt{(alpha,beta,rho)} or \texttt{(bulk\_c,beta,rho)}.\newline
 
 The algorithm uses a steepest descent method with a step length
-determined by the given maximum update percentage.
+determined by the given maximum update percentage.\newline
 
@@
 as several additional components which are also stored on the CIG
-svn server, described next.
+svn server, described next.\newline
 
@@
 any optimization problem. There are options for computing a conjugate
-gradient step, as well as a source subspace projection step.
+gradient step, as well as a source subspace projection step.\newline
 

### 2023-04-23 doc/USER_MANUAL/13_changing_the_model.tex

@@
 frequency. To change this reference frequency, you change the value
-of \texttt{ATTENUATION\_f0\_REFERENCE} defined (in Hz) in input file \texttt{DATA/Par\_file}.
+of \texttt{ATTENUATION\_f0\_REFERENCE} defined (in Hz) in input file \texttt{DATA/Par\_file}.\newline
 
@@
 and the file \texttt{tomography\_model.xyz} will need to reside in \texttt{DATA/}.
-All entries in the second column of \texttt{materials\_file} will be \texttt{-1}, which means that each element in the mesh will be interpolated according to the values in \texttt{tomography\_model.xyz}.
+All entries in the second column of \texttt{materials\_file} will be \texttt{-1}, which means that each element in the mesh will be interpolated according to the values in \texttt{tomography\_model.xyz}.\newline
 
@@
 All entries in the second column of \texttt{materials\_file} will need to be either \texttt{-1} or \texttt{-2}, depending on whether the element is above or below 15~km depth.
-In this sense, the construction of the mesh and the discretization of the external model will generally be done in tandem.
+In this sense, the construction of the mesh and the discretization of the external model will generally be done in tandem.\newline
 
@@
 and the number of samples (\texttt{n1}, \texttt{n2} and \texttt{n3}) necessary to read
-the linked binary file (\texttt{in} field).
+the linked binary file (\texttt{in} field).\newline
 
@@
 during the meshing phase. For more information, refer to
-section~\ref{cha:Running-the-Mesher-Meshfem3D}.
+section~\ref{cha:Running-the-Mesher-Meshfem3D}.\newline
 
 In \texttt{DATA/Par\_file}, \texttt{\textbf{MODEL}} should be set to \texttt{\textbf{sep}},
-and \texttt{\textbf{SEP\_MODEL\_DIRECTORY}} should point to wherever your sep model files reside.
+and \texttt{\textbf{SEP\_MODEL\_DIRECTORY}} should point to wherever your sep model files reside.\newline
 
@@
 that binary files are single precision floating point numbers stored in little endian format.
-There is currently only support for \texttt{vp}, \texttt{vs} and \texttt{rho} variables.
+There is currently only support for \texttt{vp}, \texttt{vs} and \texttt{rho} variables.\newline
 

### 2023-04-23 doc/USER_MANUAL/15_informations_for_developers.tex

--- 2022-11-09 doc/USER_MANUAL/15_informations_for_developers.tex
@@
 
-\noindent We also have simple 3D demo source codes that implement the SEM in a single, small program, in directory\\
-utils/small\_SEM\_solvers\_in\_Fortran\_and\_C\_without\_MPI\_to\_learn of the specfem3d package.
+\noindent
+We also have simple 3D demo source codes that implement the SEM in a single, small program, in directory\newline
+\texttt{utils/small\_SEM\_solvers\_in\_Fortran\_and\_C\_without\_MPI\_to\_learn} of the specfem3d package.
 They are useful to learn how the spectral-element method works, and how to write or modify a code to implement it.
 Also useful to test new ideas by modifying these simple codes to run some tests.
-We also have a similar, even simpler, demo source code for the 2D case in directory\\
-utils/small\_SEM\_solver\_in\_Fortran\_without\_MPI\_to\_learn of the specfem2d package.\\
+We also have a similar, even simpler, demo source code for the 2D case in directory\newline
+\texttt{utils/small\_SEM\_solver\_in\_Fortran\_without\_MPI\_to\_learn} of the specfem2d package.\newline
 
-\noindent For information on how to contribute to the code, i.e., for how to make your modifications, additions or improvements part of the
+\noindent
+For information on how to contribute to the code, i.e., for how to make your modifications, additions or improvements part of the
 official package, see \url{https://github.com/SPECFEM/specfem3d/wiki} .

### 2023-04-23 doc/USER_MANUAL/A_reference_frame.tex

@@
 Note that this convention is different from both the \citet{AkRi80}
-convention and the Harvard Centroid-Moment Tensor (CMT) convention.
+convention and the Global/Harvard Centroid-Moment Tensor (CMT) convention.
 The Aki \& Richards convention is
@@
 \end{itemize}
-and the Harvard CMT convention is
+and the Global/Harvard CMT convention is
 \begin{itemize}

### 2023-04-23 doc/USER_MANUAL/B_channel_codes.tex
Path updates for parameters moved from constants.h to Par_file and vice versa, and other minor corrections.

@@
 seismograms. Please visit \url{http://www.iris.edu/manuals/SEED_appA.htm}
-for further information.\\
+for further information.\newline
 
@@
 \texttt{M}, \texttt{B}, \texttt{H}, \texttt{C} or \texttt{F}, shown
-by red color in the figure.\\
+by red color in the figure.\newline
 
@@
 instrument code of SEM seismograms will always be \texttt{X}, as assigned
-by FDSN for synthetic seismograms. \\
+by FDSN for synthetic seismograms. \newline
 
@@
 is set to \texttt{.true.}, then the three components are labelled
-\texttt{X/Y/Z} according to the Cartesian reference frame. \\
+\texttt{X/Y/Z} according to the Cartesian reference frame. \newline
 
 \noindent \texttt{\textbf{EXAMPLE:}} The sampling rate is given by
-\texttt{DT} in the main parameter file \texttt{Par\_file} located
-in the \texttt{DATA/} subdirectory. Depending on the resolution of
+\texttt{DT} in the main parameter file \texttt{DATA/Par\_file}. Depending on the resolution of
 your simulations, if you choose a sampling rate greater than $0.01$

### 2023-04-23 doc/USER_MANUAL/authors.tex

--- 2022-10-28 doc/USER_MANUAL/authors.tex
@@
 to meet with G\'eza Seriani and Enrico Priolo, who introduced him to their 2D Chebyshev version of the spectral-element method with a non-diagonal mass matrix.
-We are deeply indebted and thankful to them for that.
+We are deeply indebted and thankful to them for that.\newline
 
@@
 %%%%%%%% please do NOT merge the lines of the paragraph below, it makes it difficult to insert new names in alphabetical order; leave a single name per line
-Nikhil Vaidyar
+Since then it has been developed and maintained by a development team: in alphabetical order,
+Michael Afanasiev,
+Rafael Almada,
+Jean-Paul (Pablo) Ampuero,
+\'Etienne Bachmann,
+Kangchen Bai,
+Piero Basini,
+Stephen Beller,
+Jordan Bishop,
+C\'eline Blitz,
+Alexis Bottero,
+Ebru Bozda\u{g},
+Emanuele Casarotti,
+Joseph Charles,
+Min Chen,
+Paul Cristini,
+Cl\'ement Durochat,
+Percy Galvez,
+Rene Gassmoeller,
+Hom Nath Gharti,
+Dominik G\"oddeke,
+Leopold Grinberg,
+Aakash Gupta,
+Vala Hj\"orleifsd\'ottir,
+Foivos Karakostas,
+Sue Kientz,
+Dimitri Komatitsch,
+Jes\'us Labarta,
+Nicolas Le Goff,
+Pieyre Le Loher,
+Matthieu Lefebvre,
+Qinya Liu,
+Youshan Liu,
+David Luet,
+Yang Luo,
+Alessia Maggi,
+Federica Magnoni,
+Roland Martin,
+Ren\'e Matzen,
+Geordie McBain,
+Dennis McRitchie,
+Matthias Meschede,
+Peter Messmer,
+David Mich\'ea,
+Ryan Modrak,
+Vadim Monteiller,
+Surendra Nadh Somala,
+Masaru Nagaso,
+Tarje Nissen-Meyer,
+Elif Oral,
+Daniel Peter,
+Kevin Pouget,
+Max Rietmann,
+Elliott Sales de Andrade,
+Brian Savage,
+Bernhard Schuberth,
+Anne Sieminski,
+James Smith,
+Leif Strand,
+Carl Tape,
+Jeroen Tromp,
+Eduardo Valero Cano,
+Brice Videau,
+Jean-Pierre Vilotte,
+Huihui Weng,
+Zhinan Xie,
+Chang-Hua Zhang,
+Hejun Zhu.\newline
 

### 2023-05-04 doc/USER_MANUAL/02_getting_started.tex

@@
 - CUDA 11 for Ampere, like A100
+- CUDA 12 for Hopper, like H100
 \end{verbatim}

### 2023-06-02 doc/USER_MANUAL/04_creating_databases.tex
- added a new parameter "MOVIE_VOLUME_STRESS" in the Par_file to save the stress tensor components in the entire volume for the viscoelastic simulatons.

--- 2022-11-30 doc/USER_MANUAL/04_creating_databases.tex
@@
 frames.
+\item [{\texttt{MOVIE\_VOLUME\_STRESS}}] Set this flag to \texttt{.true.}
+if you want to save the stress tensor for the volume movie.
+To activate this option, \texttt{MOVIE\_VOLUME} must be \texttt{.true.}.
+Please note that this option will generate huge output because it will save six independent stress tensor components in the entire volume.
+This option is currently implemented only for the viscoelastic simulations.
 \item [{\texttt{USE\_HIGHRES\_FOR\_MOVIES}}] Set this flag to \texttt{.true.}
@@
 is more precise than using the closest GLL mesh point, but it is also a bit slower.
-If needed, you can change that default behavior and force the code to use the closest
-GLL mesh point instead by setting flag \texttt{USE\_BEST\_LOCATION} to \texttt{.false.} instead of \texttt{.true.} in file
-\texttt{setup/constants.h} and recompiling the code.
 \item [{\texttt{USE\_RICKER\_TIME\_FUNCTION}}] Turn this flag on to use

### 2023-06-02 doc/USER_MANUAL/copyright_and_version.tex

@@
 
-version 4.0, December 2022:
+version 4.0, March 2023:
 Rafael Almada, Jean-Paul Ampuero, Etienne Bachmann, Kangchen Bai, Stephen Beller, Jordan Bishop, Alexis Bottero,

### 2023-06-02 doc/USER_MANUAL/notes_and_acknowledgement.tex

@@
 
-OpenDX \urlwithparentheses{http://www.opendx.org} is open-source
-based on IBM Data Explorer, AVS \urlwithparentheses{http://www.avs.com}
-is a trademark of Advanced Visualization Systems, and ParaView \urlwithparentheses{http://www.paraview.org}
+\href{http://www.opendx.org}{OpenDX} is open-source
+based on IBM Data Explorer, \href{http://www.avs.com}{AVS}
+is a trademark of Advanced Visualization Systems, and \href{http://www.paraview.org}{ParaView}
 is an open-source visualization platform.\newline

### 2023-11-07 doc/USER_MANUAL/05_running_the_solver.tex
- added description for "reuse" feature for external source time function file

@@
 The time step must be exactly the same as that used for the simulation.
-When the flag is set to \texttt{.false.}, then the line with the external source-time function file must not appear in the files \texttt{FORCESOLUTION} and \texttt{CMTSOLUTION}, otherwise the solver will exit with an error.
+If the external source time functions are same for all the sources, differing only in their respective time shift, you can define the source time function file only for the first source and set the external source function file to 'reuse' and define the appropriate value of \texttt{time shift} for all the other sources.When the flag is set to \texttt{.false.}, then the line with the external source-time function file must not appear in the files \texttt{FORCESOLUTION} and \texttt{CMTSOLUTION}, otherwise the solver will exit with an error.
 When using an external source file, you can still set up the source location and directivity as in the default case.

### 2023-11-07 doc/USER_MANUAL/05_running_the_solver.tex
- minor edit of the 'reuse' feature description.

--- 2023-03-09 doc/USER_MANUAL/05_running_the_solver.tex
@@
 The time step must be exactly the same as that used for the simulation.
+If the external source-time functions are same for all the sources, differing only in their respective time shift, you can define the source-time function file only for the first source and set the external source-time function file to 'reuse' and define the appropriate value of \texttt{time shift} for all the other sources.
 When the flag is set to \texttt{.false.}, then the line with the external source-time function file must not appear in the files \texttt{FORCESOLUTION} and \texttt{CMTSOLUTION}, otherwise the solver will exit with an error.

### 2023-11-14 doc/USER_MANUAL/02_getting_started.tex
updates version number to 4.1.0

@@
 with CUBIT). Additional control parameters are discussed in section~\ref{cha:Main-Parameter}.
+
+%-----------------------------------------------------------------------------------------------------------------------------------%
+\section{Using HDF5 for file I/O}
+%-----------------------------------------------------------------------------------------------------------------------------------%
+
+As file I/O can be a bottleneck in large-scale simulations, SPECFEM3D supports file I/O using the HDF5 format
+for movie snapshots and database files. To support this feature, you will need to compile the code with corresponding HDF5 flags.
+The configuration of the package could look for example like:
+{\small
+\begin{verbatim}
+  ./configure --with-hdf5 HDF5_INC="/opt/homebrew/include" HDF5_LIBS="-L/opt/homebrew/lib" \
+      ..
+\end{verbatim}
+}
+In the main \texttt{Par\_file}, you will then have to turn on the HDF5 flag \texttt{HDF5\_ENABLED}.
+Note that additional MPI processes can be launched specifically to handle the file I/O in an asynchronous way.
+The number of these additional MPI processes is specified by the parameter \texttt{HDF5\_IO\_NODES}, such that
+the total number of MPI processes to launch the executables becomes \texttt{NPROC + HDF5\_IO\_NODES}.
 

### 2023-11-14 doc/USER_MANUAL/authors.tex
updates version number to 4.1.0

@@
 Percy Galvez,
+Rahul Garg,
 Rene Gassmoeller,

### 2023-11-14 doc/USER_MANUAL/04_creating_databases.tex
adds support for MOVIE_VOLUME_STRESS flag and output

@@
 for further details.
+\item [{\texttt{LTS\_MODE}}] Turning this flag on will use a local-time stepping scheme for the Newmark time scheme.
+It will re-distribute the mesh partitioning to balance out elements with similar local-time step sizes.
+This feature can help improve the time-to-solution when the mesh contains elements of various sizes,
+in particular when the mesh contains only a few very small elements that would limit the global time step size.
+
 \end{description}
@@
 To activate this option, \texttt{MOVIE\_VOLUME} must be \texttt{.true.}.
-Please note that this option will generate huge output because it will save six independent stress tensor components in the entire volume.
+Please note that this option will generate huge outputs, because it will save six independent stress tensor components in the entire volume.
 This option is currently implemented only for the viscoelastic simulations.
@@
 
+%% ADIOS
 \item [\texttt{ADIOS\_ENABLED}] Turn this flag on to enable ADIOS. If set to \texttt{.false.}, subsequent ADIOS
@@
 \item [\texttt{ADIOS\_FOR\_KERNELS}] Turn this flag on to produce ADIOS kernels that can later be visualized with the ADIOS version of combine\_vol\_data.
+
+%% HDF5
+\item [\texttt{HDF5\_ENABLED}] Turn this flag on to enable HDF5 file I/O. If set to \texttt{.false.}, subsequent HDF5
+parameters will not be considered.
+\item [\texttt{HDF5\_FOR\_MOVIES}] Turn this flag on to use HDF5 format for movie outputs.
+\item [\texttt{HDF5\_IO\_NODES}] Setting this number >0 will allocate additional MPI processes specifically to handle
+the HDF5 file I/O in an asynchronous way.
+
 \end{description}

### 2023-11-15 doc/USER_MANUAL/02_getting_started.tex
updates doc

@@
 If you have problems configuring the code on a Cray machine, i.e. for instance if you get an error message from the \texttt{configure} script, try exporting these two variables:
-\texttt{MPI\_INC=\${CRAY\_MPICH2\_DIR}/include and FCLIBS=" "}, and for more details if needed you can refer to the \texttt{utils/Cray\_compiler\_information} directory.
+\texttt{MPI\_INC=\${CRAY\_MPICH2\_DIR}/include and FCLIBS=" "}, and for more details if needed you can refer to the \texttt{utils/infos/Cray\_compiler\_information} directory.
 You can also have a look at the configure script called:\newline
-\texttt{utils/Cray\_compiler\_information/configure\_SPECFEM\_for\_Piz\_Daint.bash}.
+\texttt{utils/infos/Cray\_compiler\_information/configure\_SPECFEM\_for\_Piz\_Daint.bash}.
 

### 2023-11-15 doc/USER_MANUAL/04_creating_databases.tex
updates doc

@@
 for further details.
-\item [{\texttt{LTS\_MODE}}] Turning this flag on will use a local-time stepping scheme for the Newmark time scheme.
+\item [{\texttt{LTS\_MODE}}] Turning this flag on will use a local-time stepping scheme for the Newmark time scheme \citep{Rietmann2015,Rietmann2016}.
 It will re-distribute the mesh partitioning to balance out elements with similar local-time step sizes.
@@
 To convert one to the other see \texttt{doc/note\_on\_Qkappa\_versus\_Qp.pdf}
-and in folder \texttt{utils/attenuation}, the tool
+and in folder \texttt{utils/small\_utilities/attenuation}, the tool
 \texttt{conversion\_from\_Qkappa\_Qmu\_to\_Qp\_Qs\_from\_Dahlen\_Tromp\_959\_960.f90}.
@@
 in time (PML being weakly unstable from a mathematical point of view, very deformed mesh elements
-inside the PMLs can trigger instabilities much more quickly). \red{We have two utilities in directory utils/CPML that do that automatically
+inside the PMLs can trigger instabilities much more quickly). \red{We have two utilities in directory \texttt{utils/CPML} that do that automatically
 and that are very fast}. To stabilize PMLs it also helps to add a transition layer of geometrically-regular non-PML elements, in which attenuation is also
 turned off (i.e. $Q_\kappa = Q_\mu = 9999$ in that layer), as in the red layer of Figure~\ref{fig:mesh_extrusion}.
-Our tools in directory utils/CPML implement that transition layer automatically.\newline
+Our tools in directory \texttt{utils/CPML} implement that transition layer automatically.\newline
 

### 2023-11-15 doc/USER_MANUAL/06_fault_sources.tex
updates doc

@@
 \begin{lyxlist}{00.00.0000}
-\item [{$\mathtt{EXAMPLES/tpv16/description/TPV16\_17\_Description\_v03.pdf}$}]~
+\item [{$\mathtt{EXAMPLES/applications/fault\_examples/tpv16/description/TPV16\_17\_Description\_v03.pdf}$}]~
 \end{lyxlist}
@@
 the mesh inside the CUBIT GUI.
-\item Use the Matlab function $\mathtt{utils/critical\_timestep.m}$ to
+\item Use the Matlab function $\mathtt{utils/small\_utilities/critical\_timestep.m}$ to
 compute\newline
@@
 where \#it\# is the timestep number. These can be read in Matlab with
-the function $\mathtt{utils/FSEM3D\_snapshot.m}$
+the function $\mathtt{utils/dynamic\_rupture/FSEM3D\_snapshot.m}$
 \end{enumerate}

### 2023-11-15 doc/USER_MANUAL/09_noise_simulations.tex
updates doc (newlines for better conversion with pandoc)
updates doc

@@
 \begin{verbatim}
-EXAMPLES/noise_tomography/NOISE_TOMOGRAPHY.m  (main program)
-EXAMPLES/noise_tomography/PetersonNoiseModel.m
+EXAMPLES/applications/noise_tomography/NOISE_TOMOGRAPHY.m  (main program)
+EXAMPLES/applications/noise_tomography/PetersonNoiseModel.m
 \end{verbatim}
@@
 previously that the order of receivers in \texttt{DATA/STATIONS} is
-important.
+important.\newline
 
@@
 In order to illustrate noise simulations in an easy way, one example
-is provided in \texttt{EXAMPLES/noise\_tomography/}. See \texttt{EXAMPLES/noise\_tomography/README}
+is provided in \texttt{EXAMPLES/applications/noise\_tomography/}. See \texttt{EXAMPLES/applications/noise\_tomography/README}
 for explanations. \newline
@@
 
-If your workstation is suitable, you can run the example in \texttt{EXAMPLES/noise\_tomography/}
+If your workstation is suitable, you can run the example in \texttt{EXAMPLES/applications/noise\_tomography/}
 using:\newline

### 2023-11-15 doc/USER_MANUAL/14_post_processing.tex
updates doc

@@
 \end{enumerate}
-\noindent We generally use the following scripts provided in the \texttt{utils/seis\_process/}
+\noindent We generally use the following scripts provided in the \texttt{utils/scripts/seis\_process/}
 directory:
@@
 are provided in the subdirectories ~\newline
- \texttt{utils/sac2000\_alpha\_convert/} and \texttt{utils/seis\_process/asc2sac/}.
+ \texttt{utils/scripts/sac2000\_alpha\_convert/} and \texttt{utils/scripts/seis\_process/asc2sac/}.
 
@@
 the reconstruction of the regular forward wavefield. A sample script
-is provided in \texttt{utils/}:
+is provided in \texttt{utils/scripts/Cluster}:
 \begin{verbatim}

### 2023-11-15 doc/USER_MANUAL/15_informations_for_developers.tex
updates doc

@@
-\chapter{Information for developers of the code, and for people who want to learn how the technique works}\label{cha:developers}
+\chapter{Information for developers of the code}\label{cha:developers}
 
@@
 \end{verbatim}
+This is helpful for everyone who wants to learn how the spectral-element method technique works.\newline
 

### 2023-11-15 doc/USER_MANUAL/03_mesh_generation.tex

@@
 see \texttt{doc/note\_on\_Qkappa\_versus\_Qp.pdf} and\newline
-\texttt{utils/attenuation/conversion\_from\_Qkappa\_Qmu\_to\_Qp\_Qs\_from\_Dahlen\_Tromp\_959\_960.f90}.\newline
+\texttt{utils/small\_utilities/attenuation/conversion\_from\_Qkappa\_Qmu\_to\_Qp\_Qs\_from\_Dahlen\_Tromp\_959\_960.f90}.\newline
 
@@
 \texttt{conversion\_from\_Qkappa\_Qmu\_to\_Qp\_Qs\_from\_Dahlen\_Tromp\_959\_960.f90} in folder
-\texttt{utils/attenuation/}.
+\texttt{utils/small\_utilities/attenuation/}.
 
@@
 file \texttt{Mesh\_Par\_file}, which resides in the subdirectory \texttt{DATA/meshfem3D\_files/}.
-(To see how to use it, see the EXAMPLES specific to the internal mesher in directory \texttt{EXAMPLES/meshfem3D\_examples/}.)
+(To see how to use it, see the EXAMPLES specific to the internal mesher in directory \texttt{EXAMPLES/applications/meshfem3D\_examples/}.)
 Before running the mesher, a number of parameters need to be set in
@@
 Depending on your system, please check how to run executables with MPI.
-To run the mesher on your cluster, we provide some example scripts for you to modify accordingly in \texttt{utils/Cluster/}.
+To run the mesher on your cluster, we provide some example scripts for you to modify accordingly in \texttt{utils/scripts/Cluster/}.
 For some installations, you might need to provide a file
@@
 or solver, but is required by the \texttt{go\_mesher***} and \texttt{go\_solver**}
-default job submission scripts provided in directory \texttt{utils/Cluster/}.
+default job submission scripts provided in directory \texttt{utils/scripts/Cluster/}.
 To run the mesher, it is most easily accomplished based upon the \texttt{go\_mesher***} script.
@@
 \end{description}
-Using an external mesh (for instance coming from CUBIT/TRELIS), poroelastic materials have the ID number 3, while 1 is acoustic and 2 is elastic (see the example in the package: EXAMPLES/homogeneous\_poroelastic).\newline
+Using an external mesh (for instance coming from CUBIT/TRELIS), poroelastic materials have the ID number 3, while 1 is acoustic and 2 is elastic (see the example in the package: \texttt{EXAMPLES/applications/homogeneous\_poroelastic}).\newline
 

### 2023-11-15 doc/USER_MANUAL/07_adjoint_simulations.tex

@@
 1} and \texttt{SAVE\_FORWARD = .false.}). You can automatically set
-these two variables using the \texttt{\small utils/change\_simulation\_type.pl}{\small{}
+these two variables using the \texttt{\small utils/scripts/change\_simulation\_type.pl}{\small{}
 script:}{\small \par}
 \begin{verbatim}
-utils/change_simulation_type.pl -f
+utils/scripts/change_simulation_type.pl -f
 \end{verbatim}
@@
 be \texttt{SIMULATION\_TYPE = 2} and \texttt{SAVE\_FORWARD = .false.}.
-More conveniently, use the scripts \texttt{utils/change\_simulation\_type.pl}
+More conveniently, use the scripts \texttt{utils/scripts/change\_simulation\_type.pl}
 to modify the \texttt{DATA/Par\_file} automatically (\texttt{change\_simulation\_type.pl

### 2023-11-15 doc/USER_MANUAL/11_graphics.tex
- added a new parameter "MOVIE_VOLUME_STRESS" in the Par_file to save the stress tensor components in the entire volume for the viscoelastic simulatons.
adds support for MOVIE_VOLUME_STRESS flag and output

@@
 displacement field if the parameter \texttt{SAVE\_DISPLACEMENT} is
-set, otherwise the velocity field is saved. The look of a movie is
-determined by the half-duration of the source. The half-duration should
-be large enough so that the movie does not contain frequencies that
+set, otherwise the velocity field is saved.
+Additionally, if the \texttt{MOVIE\_VOLUME\_STRESS} is set to \texttt{.true.},
+stress tensor components are saved in the entire volume.\newline
+
+The look of a movie is determined by the half-duration of the source.
+The half-duration should be large enough so that the movie does not contain frequencies that
 are not resolved by the mesh, i.e., it should not contain numerical
@@
 As an example, if the \texttt{DATA} and \texttt{OUTPUT\_FILES} directories of the concerned example are in the root specfem3d directory, \texttt{xcombine\_vol\_data} is in \texttt{./bin}, and you have to type:
-
 {\footnotesize
@@
 script \texttt{mesh2vtu.pl} located in directory \texttt{\small utils/Visualization/Paraview}, for example:
-
 {\footnotesize
@@
 and \texttt{output\_solver} files, and run:
-
 {\small
@@
 \begin{enumerate}
-\item You can use or modify the script \texttt{utils/copy\_basin\_database.pl}
+\item You can use or modify the script \texttt{utils/scripts/Cluster/copy\_basin\_database.pl}
 to accomplish this:
-
 {\small
 \begin{verbatim}
-utils/copy_database.pl slice_file lsf_machine_file filename [jobid]
+utils/scripts/Cluster/copy_basin_database.pl slice_file lsf_machine_file filename [jobid]
 \end{verbatim}
@@
 directory \texttt{utils/Visualization/Paraview/}, for example:
-
 {\footnotesize

### 2023-11-15 doc/USER_MANUAL/12_running_scheduler.tex

@@
 \end{itemize}
-Examples of job scripts can be found in the \texttt{\small utils/Cluster/}{\small{}
+Examples of job scripts can be found in the \texttt{\small utils/scripts/Cluster/}{\small{}
 directory and can straightforwardly be modified and adapted to meet

### 2023-11-15 doc/USER_MANUAL/13_changing_the_model.tex

@@
 Note that Qmu is always equal to Qs, but Qkappa is in general not equal to Qp.
-To convert one to the other see doc/note\_on\_Qkappa\_versus\_Qp.pdf and utils/attenuation/conversion\_from\_Qkappa\_Qmu\_to\_Qp\_Qs\_from\_Dahlen\_Tromp\_959\_960.f90.
+To convert one to the other see \texttt{doc/note\_on\_Qkappa\_versus\_Qp.pdf} and \texttt{utils/small\_utilities/attenuation/conversion\_from\_Qkappa\_Qmu\_to\_Qp\_Qs\_from\_Dahlen\_Tromp\_959\_960.f90}.
 For simulations with attenuation, please note that the Vp- and Vs-velocities of your model are given for a reference
@@
 Note that Qmu is always equal to Qs, but Qkappa is in general not equal to Qp.
-To convert one to the other see doc/note\_on\_Qkappa\_versus\_Qp.pdf and utils/attenuation/conversion\_from\_Qkappa\_Qmu\_to\_Qp\_Qs\_from\_Dahlen\_Tromp\_959\_960.f90.
+To convert one to the other see \texttt{doc/note\_on\_Qkappa\_versus\_Qp.pdf} and \texttt{utils/small\_utilities/attenuation/conversion\_from\_Qkappa\_Qmu\_to\_Qp\_Qs\_from\_Dahlen\_Tromp\_959\_960.f90}.
 \item [{\texttt{iflag\_aniso}}] anisotropic model flag, $0$ indicating
@@
 An example demonstrating the use SEP models is given in:
-\texttt{EXAMPLES/meshfem3D\_examples/sep\_bathymetry}.
+\texttt{EXAMPLES/applications/meshfem3D\_examples/sep\_bathymetry}.

### 2023-11-15 doc/USER_MANUAL/C_troubleshooting.tex

@@
 at the skewness table printed in the \texttt{output\_mesher.txt} file
-after executing \texttt{xmeshfem3D} for the example given in \texttt{EXAMPLES/meshfem3D\_examples/simple\_model/}:
+after executing \texttt{xmeshfem3D} for the example given in \texttt{EXAMPLES/applications/meshfem3D\_examples/simple\_model/}:
 
@@
 following figure \ref{fig:mesh.vp} of the mesh you obtain in the
-example \texttt{EXAMPLES/meshfem3D\_examples/simple\_model/}.
+example \texttt{EXAMPLES/applications/meshfem3D\_examples/simple\_model/}.
 \begin{figure}[htbp]
@@
 \caption{Paraview visualization using the mesh vtk-files for the example given
-in \texttt{EXAMPLES/meshfem3D\_examples/simple\_model/}.}
+in \texttt{EXAMPLES/applications/meshfem3D\_examples/simple\_model/}.}
 

### 2023-11-16 doc/USER_MANUAL/05_running_the_solver.tex
- added Brune and Smoothed Brune functions

@@
 \item \texttt{depth:} Set the depth of the source (in km).
-\item \texttt{source time function:} Set the type of source-time function: 0 = Gaussian function, 1 = Ricker function, 2 = Heaviside (step) function, 3 = monochromatic function, 4 = Gaussian function as defined in \citet{Meschede2011}.
+\item \texttt{source time function:} Set the type of source-time function: 0 = Gaussian function, 1 = Ricker function, 2 = Heaviside (step) function, 3 = monochromatic function, 4 = Gaussian function as defined in \citet{Meschede2011}, 5 = Brune function, and 6 = Smoothed Brune function.
+For the Brune and Smoothed Brune source time functions \texttt{hdurorf0} is the rise time.
+The Brune and Smoothed Brune functions are currently implemented only for the viscoelastic simulations.
 When {\texttt{USE\_RICKER\_TIME\_FUNCTION}} is turned on in the main parameter file \texttt{DATA/Par\_file},

### 2023-11-17 doc/USER_MANUAL/05_running_the_solver.tex
- added Brune and Smoothed Brune source time functions

@@
 \item \texttt{source time function:} Set the type of source-time function: 0 = Gaussian function, 1 = Ricker function, 2 = Heaviside (step) function, 3 = monochromatic function, 4 = Gaussian function as defined in \citet{Meschede2011}, 5 = Brune function, and 6 = Smoothed Brune function.
-For the Brune and Smoothed Brune source time functions \texttt{hdurorf0} is the rise time.
+Please note that we have implemented time derivatives of the Brune and smoothed Brune functions as the moment rate functions.
+For these source time functions, \texttt{hdurorf0} is the source duration or the rise time.
 The Brune and Smoothed Brune functions are currently implemented only for the viscoelastic simulations.

### 2023-12-09 doc/USER_MANUAL/04_creating_databases.tex
- added USE_OTHER_TIME_FUNCTION in Par_file to implement the Brune and Smooth Brune functions for CMTSOLUTION

@@
 Any other source-time function may then be obtained by convolution.
+\item [{\texttt{USE\_OTHER\_TIME\_FUNCTION}}] Turn this flag on to use other source time functions for the CMTSOLUTION (Brune and Smooth Brune).
+If this flag is on, one line with the other source time function name (\textt{brune} or \texttt{smooth\_brune}) must be added at the end of each \texttt{CMTSOLUTION}.
 \item [{\texttt{PRINT\_SOURCE\_TIME\_FUNCTION}}] Turn this flag on to print

### 2023-12-09 src/specfem3D/get_cmt.f90

@@ -27,10 +27,10 @@
 
   subroutine get_cmt(CMTSOLUTION,yr,jda,mo,da,ho,mi,sec, &
                      tshift_src,hdur,lat,long,depth,moment_tensor, &
-                     DT,NSOURCES,min_tshift_src_original,user_source_time_function)
+                     DT,NSOURCES,min_tshift_src_original,cmt_stf,user_source_time_function)
 
   use constants, only: IIN,MAX_STRING_LEN,CUSTOM_REAL
-  use shared_parameters, only: USE_EXTERNAL_SOURCE_FILE,NSTEP_STF,NSOURCES_STF,NOISE_TOMOGRAPHY
+  use shared_parameters, only: USE_OTHER_TIME_FUNCTION,USE_EXTERNAL_SOURCE_FILE,NSTEP_STF,NSOURCES_STF,NOISE_TOMOGRAPHY
 
   implicit none
 
@@ -38,6 +38,7 @@
 
   character(len=MAX_STRING_LEN), intent(in) :: CMTSOLUTION
   integer, intent(in) :: NSOURCES
+  integer, dimension(NSOURCES), intent(out) :: cmt_stf
   double precision, intent(in) :: DT
 
   ! PDE time info (from first header line)
@@ -367,6 +368,22 @@
     ! null half-duration indicates a Heaviside
     ! replace with very short error function
     if (hdur(isource) < 5.d0 * DT) hdur(isource) = 5.d0 * DT
+
+    ! reads OTHER SOURCE TIME FUNCTION
+    if (USE_OTHER_TIME_FUNCTION) then
+      ! gets other STF name
+      read(IIN,"(a)") string
+      if (trim(string) == 'BRUNE' .or. &
+          trim(string) == 'Brune' .or. &
+          trim(string) == 'brune') then
+          cmt_stf(isource) = 5  
+      endif
+      if (trim(string) == 'SMOOTH_BRUNE' .or. &
+          trim(string) == 'Smooth_Brune' .or. &
+          trim(string) == 'smooth_brune') then
+          cmt_stf(isource) = 6  
+      endif
+    endif
 
     ! reads USER EXTERNAL SOURCE if needed
     if (USE_EXTERNAL_SOURCE_FILE) then

### 2023-12-09 src/specfem3D/locate_source.F90

@@ -34,7 +34,7 @@
 
   use specfem_par, only: USE_FORCE_POINT_SOURCE,USE_RICKER_TIME_FUNCTION, &
       UTM_PROJECTION_ZONE,SUPPRESS_UTM_PROJECTION,USE_SOURCES_RECEIVERS_Z, &
-      USE_EXTERNAL_SOURCE_FILE, &
+      USE_OTHER_TIME_FUNCTION,USE_EXTERNAL_SOURCE_FILE, &
       USE_TRICK_FOR_BETTER_PRESSURE,COUPLE_WITH_INJECTION_TECHNIQUE, &
       myrank,NSPEC_AB,NGLOB_AB,ibool,xstore,ystore,zstore,DT, &
       NSOURCES,HAS_FINITE_FAULT_SOURCE,LTS_MODE
@@ -42,7 +42,7 @@
   ! sources arrays
   use specfem_par, only: Mxx,Myy,Mzz,Mxy,Mxz,Myz,hdur, &
     tshift_src,utm_x_source,utm_y_source, &
-    force_stf,factor_force_source,comp_dir_vect_source_E,comp_dir_vect_source_N,comp_dir_vect_source_Z_UP, &
+    cmt_stf,force_stf,factor_force_source,comp_dir_vect_source_E,comp_dir_vect_source_N,comp_dir_vect_source_Z_UP, &
     xi_source,eta_source,gamma_source,nu_source, &
     ispec_selected_source,islice_selected_source
 
@@ -520,6 +520,60 @@
             case default
               stop 'unsupported force_stf value!'
             end select
+          elseif (USE_OTHER_TIME_FUNCTION) then
+            ! CMT sources with other time functions
+            select case(cmt_stf(isource))
+            case (0)
+              ! Gaussian
+              write(IMAIN,*) '    using Gaussian source time function'
+              write(IMAIN,*) '             half duration: ',hdur(isource),' seconds'
+              write(IMAIN,*) '    Gaussian half duration: ',hdur(isource)/SOURCE_DECAY_MIMIC_TRIANGLE,' seconds'
+            case (1)
+              ! Ricker
+              write(IMAIN,*) '    using Ricker source time function'
+              ! prints frequency content for point forces
+              f0 = hdur(isource)
+              t0_ricker = 1.2d0/f0
+              write(IMAIN,*)
+              write(IMAIN,*) '    using a source of dominant frequency ',f0
[…]

### 2023-12-20 doc/USER_MANUAL/authors.tex
updates authors

--- 2023-04-23 doc/USER_MANUAL/authors.tex
@@
 Vadim Monteiller,
+Christina Morency,
 Surendra Nadh Somala,

### 2024-01-13 doc/USER_MANUAL/13_changing_the_model.tex

--- 2023-04-23 doc/USER_MANUAL/13_changing_the_model.tex
@@
 each element in the file \texttt{MESH/materials\_file} and use the
-following format in the file \texttt{MESH/nummaterial\_velocity\_file}
+following format in the file \newline
+\texttt{MESH/nummaterial\_velocity\_file}
 when using CUBIT to construct your mesh (see also section \ref{subsec:Exporting-the-Mesh}):
@@
 where  \texttt{x}, \texttt{y}, \texttt{z} are the grid point position, \texttt{vp}, \texttt{vs} and \texttt{rho} the P- and S-wave speeds and density, and \texttt{Qp} and \texttt{Qs} the quality factors for P- and S-wave speeds.
-The quality factors are optional and can be omitted for purely elastic models (the tomography routine will recognize both formats).
+The quality factors are optional and can be omitted for purely elastic models (the tomography routine will recognize both formats).\newline
+
+\noindent
 Internally, the quality factors will be converted to bulk and shear attenuation values,
@@
 Note that Qmu is always equal to Qs, but Qkappa is in general not equal to Qp.
-To convert one to the other see doc/note\_on\_Qkappa\_versus\_Qp.pdf and utils/attenuation/conversion\_from\_Qkappa\_Qmu\_to\_Qp\_Qs\_from\_Dahlen\_Tromp\_959\_960.f90.
+To convert one to the other see \texttt{doc/note\_on\_Qkappa\_versus\_Qp.pdf}.
+There is also a source code file\newline \texttt{conversion\_from\_Qkappa\_Qmu\_to\_Qp\_Qs\_from\_Dahlen\_Tromp\_959\_960.f90}
+provided in folder \texttt{utils/small\_utilities/attenuation/}.
 For simulations with attenuation, please note that the Vp- and Vs-velocities of your model are given for a reference
@@
 
+\noindent
 The code uses a constant $Q$ quality factor, write(IMAIN,*) "but approximated based on a series of Zener standard linear solids (SLS).
-The approximation is thus performed in a given frequency band determined based on that \texttt{ATTENUATION\_f0\_REFERENCE} reference frequency.
+The approximation is thus performed in a given frequency band determined based on that \texttt{ATTENUATION\_f0\_REFERENCE} reference frequency.\newline
 
@@
 
+\noindent
 The user can implement his own interpolation algorithm for the tomography
-model by changing the routine \texttt{model\_tomography.f90} located
+model by changing the routines in file \texttt{model\_tomography.f90} located
 in the \texttt{src/generate\_databases/} directory. Moreover, for
@@
 description, the \texttt{nummaterial\_velocity\_file} has multiple
-lines each with the corresponding suitable format described above.
+lines each with the corresponding suitable format described above.\newline
+
+\noindent
+Finally, the tomography file can have as many comment lines starting with a hash sign "\#" as you like.
+In case you want to provide a tomography file with regular gridded lon/lat/depth coordinates instead of x/y/z,
+you can add a comment line like this into the header section of the file:
+{\small
+\begin{verbatim}
+# coordinate format   : lon / lat / depth
+\end{verbatim}
+}
+\noindent
+This can be helpful when working with a simulation setup for a specific local region using a corresponding UTM-projection (and topography).
+In such cases, it is more convenient to use lon/lat/depth coordinates instead of UTM coordinates directly.\newline
+As an example for how to create such a tomography file, a python script \texttt{run\_convert\_IRIS\_EMC\_netCDF\_2\_tomo.py} is provided in
+folder \texttt{utils/scripts/} to extract a tomography file from an \href{https://ds.iris.edu/ds/products/emc-earthmodels/}{IRIS EMC model}
+that gets provided in netCDF format.
 
@@
 In some cases it may be desirable to have an external tomography model that is described in more than one file.
-For example, in cases like southern California, the length scale of variation in the structure of the wave speed model is much shorter in the sedimentary basin models within the upper 15~km.
+For example, in cases like Southern California, the length scale of variation in the structure of the wave speed model is much shorter in the sedimentary basin models within the upper 15~km.
 Therefore one might want to use an external tomography file that is sampled with, say, $\Delta x = 1000$~m, $\Delta y = 1000$~m, and $\Delta z = 250$~m in the uppermost 15~km, and then use $\Delta x = 2000$~m, $\Delta y = 2000$~m, and $\Delta z = 1000$~m below a depth of 15~km.
@@
 \item [{\texttt{qkappa\_atten,qmu\_atten}}] Bulk and shear wave quality factor: $0<Q_{\kappa,\mu}<9000$.
-Note that Qmu is always equal to Qs, but Qkappa is in general not equal to Qp.
-To convert one to the other see doc/note\_on\_Qkappa\_versus\_Qp.pdf and utils/attenuation/conversion\_from\_Qkappa\_Qmu\_to\_Qp\_Qs\_from\_Dahlen\_Tromp\_959\_960.f90.
+Note that Qmu is always equal to Qs, but Qkappa is in general not equal to Qp. To convert one to the other see\newline
+\texttt{doc/note\_on\_Qkappa\_versus\_Qp.pdf} and source code file\newline
+\texttt{conversion\_from\_Qkappa\_Qmu\_to\_Qp\_Qs\_from\_Dahlen\_Tromp\_959\_960.f90}\newline
+in folder \texttt{utils/small\_utilities/attenuation}.
 \item [{\texttt{iflag\_aniso}}] anisotropic model flag, $0$ indicating
@@
 
-An example demonstrating the use SEP models is given in:
-\texttt{EXAMPLES/meshfem3D\_examples/sep\_bathymetry}.
+An example demonstrating the use of SEP models is given in:\newline
+\texttt{EXAMPLES/applications/meshfem3D\_examples/sep\_bathymetry}.\newline
+
+\noindent
+Please consider adding improvements and/or new model implementations to this package - feel free to contribute!
